### 2019-01-17 0-Misc/lorem.tex
Using template from Rémi Bonnefoi and Rami Othman.

Paphius quin etiam et Cornelius senatores, ambo venenorum artibus pravis se polluisse confessi, eodem pronuntiante Maximino sunt interfecti. pari sorte etiam procurator monetae extinctus est. Sericum enim et Asbolium supra dictos, quoniam cum hortaretur passim nominare, quos vellent, adiecta religione firmarat, nullum igni vel ferro se puniri iussurum, plumbi validis ictibus interemit. et post hoe flammis Campensem aruspicem dedit, in negotio eius nullo sacramento constrictus.

Auxerunt haec vulgi sordidioris audaciam, quod cum ingravesceret penuria commeatuum, famis et furoris inpulsu Eubuli cuiusdam inter suos clari domum ambitiosam ignibus subditis inflammavit rectoremque ut sibi iudicio imperiali addictum calcibus incessens et pugnis conculcans seminecem laniatu miserando discerpsit. post cuius lacrimosum interitum in unius exitio quisque imaginem periculi sui considerans documento recenti similia formidabat.

Ibi victu recreati et quiete, postquam abierat timor, vicos opulentos adorti equestrium adventu cohortium, quae casu propinquabant, nec resistere planitie porrecta conati digressi sunt retroque concedentes omne iuventutis robur relictum in sedibus acciverunt.

Iamque lituis cladium concrepantibus internarum non celate ut antea turbidum saeviebat ingenium a veri consideratione detortum et nullo inpositorum vel conpositorum fidem sollemniter inquirente nec discernente a societate noxiorum insontes velut exturbatum e iudiciis fas omne discessit, et causarum legitima silente defensione carnifex rapinarum sequester et obductio capitum et bonorum ubique multatio versabatur per orientales provincias, quas recensere puto nunc oportunum absque Mesopotamia digesta, cum bella Parthica dicerentur, et Aegypto, quam necessario aliud reieci ad tempus.

Hae duae provinciae bello quondam piratico catervis mixtae praedonum a Servilio pro consule missae sub iugum factae sunt vectigales. et hae quidem regiones velut in prominenti terrarum lingua positae ob orbe eoo monte Amano disparantur.

Sed tamen haec cum ita tutius observentur, quidam vigore artuum inminuto rogati ad nuptias ubi aurum dextris manibus cavatis offertur, inpigre vel usque Spoletium pergunt. haec nobilium sunt instituta.

Hac ita persuasione reducti intra moenia bellatores obseratis undique portarum aditibus, propugnaculis insistebant et pinnis, congesta undique saxa telaque habentes in promptu, ut si quis se proripuisset interius, multitudine missilium sterneretur et lapidum.

Ibi victu recreati et quiete, postquam abierat timor, vicos opulentos adorti equestrium adventu cohortium, quae casu propinquabant, nec resistere planitie porrecta conati digressi sunt retroque concedentes omne iuventutis robur relictum in sedibus acciverunt.

Iam virtutem ex consuetudine vitae sermonisque nostri interpretemur nec eam, ut quidam docti, verborum magnificentia metiamur virosque bonos eos, qui habentur, numeremus, Paulos, Catones, Galos, Scipiones, Philos; his communis vita contenta est; eos autem omittamus, qui omnino nusquam reperiuntur.

Quare talis improborum consensio non modo excusatione amicitiae tegenda non est sed potius supplicio omni vindicanda est, ut ne quis concessum putet amicum vel bellum patriae inferentem sequi; quod quidem, ut res ire coepit, haud scio an aliquando futurum sit. Mihi autem non minori curae est, qualis res publica post mortem meam futura, quam qualis hodie sit.

### 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
Using template from Rémi Bonnefoi and Rami Othman.

%!TEX root = ../thesis.tex

% First chapter begins here
%\adjustmtc
\chapter*{Introduction}
\addcontentsline{toc}{chapter}{Introduction}
\chaptermark{Introduction}
\minitoc
\label{chapter:Intro}
\graphicspath{{2-Chapters/1-Chapter/Images/}}
% Write miniTOC just after the title
%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
\section*{General context}
\indent The work presented in this PhD thesis applies in aeronautical telemetry, which is the process of remotely collecting measurements done by electronic equipment on board the aircraft and transmitting these data to a distant location for monitoring, display, and recording \cite{telemetry_def}.

An aeronautical telemetry system is used during the flight testing phase to monitor the behavior of the plane by transmitting the aircraft dynamics settings (such as velocities, vibrations, temperatures, stress and strain parameters,...) from the aircraft to the ground station over a radio-frequency (RF) link. Aeronautical telemetry is a one-way link from the plane to the ground (i.e., there is no uplink mode) that requires real-time data transmission and analysis to ensure the pilot's safety. This makes the use of the aeronautical telemetry a critical task that demands highly reliable systems. The transmitter side of an aeronautical telemetry system (on board the aircraft) is generally composed of transducers that convert a physical stimulus into an electrical signal, multiplexers that combine the measurements into one signal, a transmitting equipment that modulates the data and one (or sometimes two) omnidirectional "blade" antenna(s). When the flight test range is going to be beyond $100$ km or when higher link reliability is required, a power amplifier can also be added at the transmitter side \cite{stacey2008aeronautical}. As for the receiver side (ground station), it is generally composed of a high gain parabolic dish antenna that allows tracking the aircraft, a low-noise amplifier that amplifies the signal without significantly decreasing the signal-to-noise ratio (SNR), a receiver demodulator that converts the modulated signal back to data, a recorder, and a demultiplexer.

The design and the performance of the transmitting/ receiving equipment are mainly defined by the adopted modulation that conveys the data. The first telemetry systems were analog and used frequency modulation (FM) of analog waveforms instead of amplitude modulation (AM) \cite{osti_1143131}. This choice favored the use of power amplifiers at their saturation mode without distorting the FM constant envelope signal and provided better bandwidth occupancy/performance trade-off than AM-based systems. Then, analog FM has been progressively upgraded to digital FM by feeding the transmitter with filtered digital data instead of analog waveforms. This change only required slight modifications of the infrastructure and it did not impact the bandwidth occupancy \cite{osti_1143131}. The digital FM version is known as the non-return to zero (NRZ) pulse code modulation/frequency modulation (PCM/FM), and it can be considered as continuous phase frequency shift keying (CPFSK) modulation or more generally as continuous phase modulation (CPM) \cite{digital_phase_mod}.

PCM/FM has been the reference modulation for decades \cite{irig106} because it has proven its robustness as well as its reliability for such high-risk use and because the required data rates were low compared to the available bandwidth before the 1990's. In the telemetry literature, this modulation is sometimes called the "legacy modulation" and is still used nowadays in several telemetry setups all over the world.  However, over the past few years, two major factors have impacted the telemetering community. The first one is the exponential expansion of wireless communication systems and especially of the commercial mobile services. As an illustration, the number of mobile-broadband subscriptions has grown more than $20\%$ annually in the last five years and has reached $4.3$ billion globally by the end of 2017 \cite{itu_facts}. This huge increase has urged telecommunication service providers to seek more bandwidth and has consequently made RF spectrum more and more costly (for instance, the United States (US) government auctioned $25$ MHz of the L-band at a record of $44.89$ billion dollars in 2015 \cite{spec_pr}). As a result, the spectrum of aeronautical telemetry has been shortened by almost $35\%$ and reallocated for the commercial applications \cite{ninja_telem}.\newline  
The second factor is the data deluge of aeronautical telemetry, which is due to the use of multiple sensors on board the aircraft to acquire more and more information during the flight. This data increase comes also as a consequence of reducing the flight testing duration for cost-effectiveness. In fact, significant resources are devoted during the flight testing phase and cost around $50$ thousand dollars in labor per hour \cite{telemetry_eco, eco_imp_tel}. Therefore, reducing the test duration without decreasing the amount of transferred data is a major demand for aerospace companies. On the one hand, this significant increase in data transfer requires the adoption of more spectrally efficient modulations than PCM/FM. On the other hand, it broadens the signal bandwidth, which makes the aeronautical telemetry channel frequency selective \cite{channel_wideband} and therefore impacts the availability of the aeronautical telemetry link. The latter issue is a real burden for aerospace companies since one delayed or cancelled test mission due to the unavailability of telemetry link could cost more than $1$ million dollars \cite{eco_imp_tel}.

Both economical and technical challenges have stimulated the flight test telemetry community to adopt more power and bandwidth efficient modulations than PCM/FM. This has been done via the advanced range telemetry (ARTM) program \cite{artm_prog}, which resulted in $2$ spectrally efficient CPM modulations: the shaped offset quadrature phase shift keying telemetry group (SOQPSK-TG) modulation and ARTM CPM. These modulations figure in the inter-range instrumentation group (IRIG) 106 standard for aeronautical telemetry, which ensures the interoperability of telemetry systems at the range commanders council (RCC) member ranges \cite{irig106}. Both modulations are transmitter friendly since they have a constant complex envelope and require minimal infrastructure change. Moreover, they allow increasing the spectral efficiency by at least a factor of $2$ compared to PCM/FM while keeping excellent power efficiency. These attractive benefits have been achieved by carefully choosing the parameters that define a CPM. However, they come at the expense of increasing the receiver complexity, especially when considering complicated scenarios such as the presence of multipath or the use of multiple-antenna transmitters \cite{jensen}. This complexity is mainly due to two factors. The first one is the presence of an inherent memory that ensures the continuity of the phase of the modulation, and the second one is the nonlinear nature of CPM since the aforementioned modulation is not a linear function of the transmitted symbols. Therefore, unlike linear modulations (such as M-PSK, M-quadrature amplitude modulation (M-QAM), amplitude and phase-shift keying (APSK),...), it is not straightforward to take advantage of the properties of linear algebra to develop reduced complexity receivers.

Even if PCM/FM is still widely used in aeronautical telemetry, we observe that more and more actors, especially in the US military ranges, migrated or started the migration to the SOQPSK solution. In such a context, the telemetry industry should develop new solutions and algorithms in order to guarantee the best possible telemetry link for these highly demanding customers. This work, supervised by Zodiac Data Systems which is one of the leading companies in aeronautical telemetry for civil and military applications, enrolls in this prospective. In this thesis, we mainly focus on SOQPSK-TG whose use is getting more and more popular in aeronautical telemetry. This modulation belongs to a particular CPM family since it transmits ternary symbols instead of binary ones. Simultaneously, it can also be interpreted as an offset modulation, i.e., it can be seen as a modulation whose quadrature component contains a timing offset compared to the in-phase component. The goal of this work is to develop reception algorithms for SOQPSK-TG under different scenarios to ensure the availability of aeronautical telemetry link. These solutions should maintain a good power efficiency without prohibitive complexity for real-time implementation.

\section*{Previous work on SOQPSK-TG}
\subsection*{SOQPSK-TG signal representations and approximations}
SOQPSK-TG is a ternary CPM whose ternary symbols are linked to the binary bits using a specific precoder given in \cite{irig106} and whose modulation index $h=1/2$. The role of the precoder is to make the phase of SOQPSK-TG behaves like the phase of an OQPSK signal when both are driven by the same bit sequence \cite{these_perrins}. As for the pulse shape that modulates the ternary symbols in the instantaneous frequency, it spans several symbol periods. The choice of these parameters increases the spectral efficiency at the expense of creating long memory and inter-symbol interference (ISI). Thus, if we take this CPM definition, the resulting reception algorithms would be highly complex. Moreover, they would be restricted to simple scenarios due to the nonlinear nature of CPM. As a consequence, it is necessary to view this modulation differently to make it more versatile.

The first pioneering work in this direction has been done by E. Perrins in \cite{these_perrins} where he showed that SOQPSK-TG could be written as a finite sum of pulse-amplitude modulated (PAM) components. The pulses of this decomposition, namely PRD, modulate pseudo-symbols. These pseudo-symbols contain memory and are nonlinear functions of the ternary symbols. This work is the extension of the Laurent decomposition \cite{pam_representation_binary}, initially developed for binary CPM. The advantage of this approach is that it approximates the signal as a sum of only two PAM waveforms. However, this does not completely "linearize" the signal because of mapping between the ternary symbols and the pseudo-symbols, which keeps the long memory effect in the approximation.

The second work has been carried by T. Nelson in \cite{these_nelson} where he took advantage of the offset nature of SOQPSK-TG and created a direct mapping between the binary bits (i.e., the bits that are used to generate the ternary symbols) and the different waveforms that constitute the eye-diagram of the signal. The representation that makes this connection (which is not evident unlike linear modulations) is known as the cross-correlated trellis coded quadrature modulation (XTCQM) representation. After some mathematical manipulations, this representation can accurately approximate SOQPSK-TG using a bank of $8$ XTCQM waveforms, each one of them being defined by only $3$ binary bits. The advantage of this approach is that it offers an approximation of SOQPSK-TG with a short inherent memory. However, it generates data dependent waveforms, which means that we cannot separate the contribution of the bits and the pulse shape in a clear way.

Each representation of SOQPSK-TG has its own benefits and drawbacks. Therefore, the choice between one of them depends on the desired application.

\subsection*{Reduced complexity detection algorithms}
Several reduced detection algorithms have been proposed in the literature for SOQPSK-TG. The first ones are linear integrate and dump (I$\&$D) detectors. They can be used to detect the binary symbols thanks to the presence of the SOQPSK-TG precoder (to the best of our knowledge, the reason why is not analytically explained in the literature). Several I$\&$D detector implementations are given in \cite{soqpsk_opt_lin_det}, but they are sub-optimal. The best one of them suffers from a performance loss of $0.72$ dB for a probability of error of $10^{-5}$ compared to the optimal case. Then, very near-optimal performance detectors have been developed, and they rely on the presented above approximations. The first one is presented in \cite{demod_perrins} and uses the PAM approximation of PRD and operates using a dynamic trellis of $4$ states. The different transitions and the order of the path states in the this trellis depend on the parity of bit index. The second very near-optimal reduced complexity detector uses a trellis composed of $16$ states and relies on the XTCQM representation of SOQPSK-TG. The performance loss of these detectors is within $0.2$ dB compared to the optimal case.      
\subsection*{Multipath mitigation techniques}
Due to the bandwidth expansion of SOQPSK-TG signal as highlighted above, the aeronautical telemetry channel becomes frequency selective. As a consequence, the receiver may capture several delayed copies of the transmitted signal with different attenuations. The transmission is thus made over a multipath channel, which can result in significant performance loss if we use the described above detectors without introducing multipath mitigation techniques, i.e., equalizers. Channel estimation and equalization is very sparsely addressed in the literature in the SOQPSK case, contrary to other modulations. To the best of our knowledge, very few works \cite{buy_channel_taps2,CMA_soqpsk,CMA_soqpsk2} have focused on multipath mitigation techniques for SOQPSK-TG, and they mainly rely on the constant modulus algorithm (CMA) due to the constant envelope nature of this modulation. The performance of CMA can sometimes be unsatisfactory for some channel configurations. 
\subsection*{The two-antenna problem on aeronautical telemetry}
The previous architectures mentioned above are built regardless of the number of transmitting antennas. However, this parameter has an important impact on both transmitter and receiver architectures. The first aeronautical telemetry systems used one transmitting antenna fixed underneath the aircraft fuselage and one receiving antenna on a fixed ground station. However, this scheme does not guarantee an optimal telemetry link because the receiver cannot always capture the line of sight (LOS) path when the aircraft performs certain maneuvers. To overcome this situation, two antennas are then placed on the aircraft in a way that an omnidirectional transmission is guaranteed no matter how the plane is oriented. Nevertheless, if the same signal is transmitted over both antennas using the same carrier frequency, the receiver may capture both copies in a destructive way, which causes the loss of the aeronautical telemetry link. This problem is known as the two-antenna problem in aeronautical telemetry \cite{two_ant_pblm1}. A solution has been proposed in \cite{jensen}, which consists of creating transmit diversity using Alamouti space-time block coding (STBC) \cite{alamouti}. This solution significantly reduces the mutual interference of the SOQPSK-TG signals and has been recently standardized in IRIG-106 \cite{irig106}. However, it makes recovering the data bits a hard task because of the used modulation as well as the presence of an unusual channel impairment, which is the differential delay. This perceptible delay represents a non-negligible portion of the symbol peiod as explained in details in Chapter \ref{chapter:5}. It appears since both signals arrive with different delays. Several decoding architectures relying on the XTCQM representation have been proposed in \cite{these_nelson, stc_decoder} but only one of them offers satisfying performance at the expense of relatively high implementation complexity. Moreover, the best decoder only provides hard outputs and does not contain any mechanism that allows mitigating multipath in the space-time coding (STC) scenario.

\section*{Contributions}
This dissertation addresses all the aforementioned points and proposes different solutions for each scenario. The important contributions are summarized as follows:
\begin{itemize}
\item A new PAM decomposition has been developed for SOQPSK-TG. It allows approximating the signal as a simple linear modulation and providing a straightforward mapping between the binary bits and the pseudo-symbols of this decomposition. This decomposition, namely DBD in this manuscript, is the root of all the proposed solutions of this work. 
\item A mathematical link has been developed between this proposed PAM decomposition and the XTCQM representation for SOQPSK.  
\item Several reduced complexity detectors have been proposed for SOQPSK-TG based on DBD. Each class of detectors relies on a certain approach (Kaleh approach \cite{simple_coherent}, Ungerboeck approach \cite{mlse_pam} and Forney approach \cite{forney_mlse}).
\item To combat multipath interference, a joint channel estimation and detection algorithm has been developed for SOQPSK-TG. It relies on the proposed decomposition and the per-survivor processing (PSP) principle \cite{polydoros_psp} and offers very attractive bit error rate (BER) performance for different channel configurations.
\item Two different decoding structures have been proposed for the STC scenario without changing the encoder structure. The first one offers a complexity $8$ times less than the state of the art decoder with very acceptable BER performance in the presence of the differential delay. The second one outperforms the state of the art decoder while keeping a lower complexity. Both proposed solutions can provide soft outputs (i.e., log-likelihood ratios).    
\item A new multipath channel estimator has been developed for the STC scenario and can be used along with the second proposed decoder. 
\item A new STC scheme has been developed that offers a better transmission rate for SOQPSK-TG than the IRIG standardized one. Moreover, the resulting decoding structure can process the even bits separately from the odd ones in the presence of non-zero differential delay and offers very attractive BER performance (within $1.1$ dB of the single input single output (SISO) bound). The complexity of the proposed decoder is at least $8$ times less than the state of the art decoder associated with the standardized encoding.   
\end{itemize}

\section*{Dissertation outline}
This dissertation is organized as follows.

We remind in \textbf{Chapter \ref{chapter:1}} the CPM model and we define its main parameters. We then describe the different aeronautical telemetry modulations and we highlight how the spectral efficiency is achieved for each modulation. We then focus on SOQPSK and we present the state of the art representations and approximations of this signal (i.e, the PAM decomposition and the XTCQM representation as well as their respective approximations).

In \textbf{Chapter \ref{chapter:2}}, we first introduce a new vision of the precoder structure by decomposing it into two different stages: the first stage is a recursive encoder and the second one is a duobinary encoder. We then exploit this new vision to develop a new PAM decomposition of duobinary encoded CPM and we apply the obtained results to SOQPSK. We also establish in this chapter a link between the proposed PAM decomposition (DBD) of SOQPSK and its XTCQM representation as well its OQPSK interpretation. Furthermore, we detail the exact role of each stage that constitutes the SOQPSK precoder.

In \textbf{Chapter \ref{chapter:3}}, we build several detection architectures as a direct consequence of proposing a new PAM decomposition. We take advantage of several approaches of the literature. Some rely on the CPM definition of SOQPSK and then exploit the PAM decomposition (Kaleh approach \cite{simple_coherent} and others use the proposed "linearized" model of SOQPSK (Ungerboeck approach \cite{mlse_pam} and Forney approach \cite{forney_mlse}). A complexity study of all the proposed detectors is also carried in this chapter.

In \textbf{Chapter \ref{chapter:4}}, we consider the presence of multipath in the SISO scenario. We first make a classification of the existing aeronautical telemetry channels of the literature. We then present the CMA solution and its different variants. We later propose a new joint channel estimation and detection algorithm for SOQPSK-TG. This solution is an adapted version of the PSP principle and relies on the proposed PAM decomposition given in Chapter \ref{chapter:2}.

In \textbf{Chapter \ref{chapter:5}}, we focus on the IRIG standardized transmit diversity scheme that allows resolving the two-antenna problem. After introducing the IRIG standardized encoder structure and the state of the art decoders, we develop two decoding structures and we present their different variants. We then consider the presence of multipath in the STC scenario in the second part of this chapter, and we detail the derivation of the new multipath estimator.

The results given in \textbf{Chapter} \ref{chapter:5} show that the performance of proposed decoders are angle dependent due to employing the standardized scheme. For this reason, we propose in \textbf{Chapter \ref{chapter:6}} a new encoding scheme that almost eliminates this angle dependency. Moreover, the scheme ameliorates the transmission rate and allows further complexity reduction of the decoding algorithm. In the second part of this chapter, we highlight that space-time coding is not the only way to overcome the two-antenna problem. We show that sending a SOQPSK-TG signal over one antenna and its artificially delayed copy over the other can also greatly reduce their mutual interference. After determining the necessary delay that should be applied, we show that the proposed interference mitigation technique given in Chapter \ref{chapter:4} can also be used in this scenario. Thus, the same receiver architecture can be employed whether one or two antennas are mounted on the aircraft using this approach.

%Despite the extensive literature regarding channel estimation and equalization techniques for linear modulations, it is not the case for SOQPSK-TG due to its nonlinear nature.

### 2019-01-17 2-Chapters/1-Chapter/chapter1.tex
Using template from Rémi Bonnefoi and Rami Othman.

--- 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
@@
 % First chapter begins here
-%\adjustmtc
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction}
-\chaptermark{Introduction}
+\chapter{Continuous Phase Modulation for Aeronautical Telemetry }
+\label{chapter:1}
 \minitoc
-\label{chapter:Intro}
+\newpage
+% Write miniTOC just after the title
 \graphicspath{{2-Chapters/1-Chapter/Images/}}
-% Write miniTOC just after the title
-%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
-%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
-\section*{General context}
-\indent The work presented in this PhD thesis applies in aeronautical telemetry, which is the process of remotely collecting measurements done by electronic equipment on board the aircraft and transmitting these data to a distant location for monitoring, display, and recording \cite{telemetry_def}.
-
-An aeronautical telemetry system is used during the flight testing phase to monitor the behavior of the plane by transmitting the aircraft dynamics settings (such as velocities, vibrations, temperatures, stress and strain parameters,...) from the aircraft to the ground station over a radio-frequency (RF) link.
-Aeronautical telemetry is a one-way link from the plane to the ground (i.e., there is no uplink mode) that requires real-time data transmission and analysis to ensure the pilot's safety.
-This makes the use of the aeronautical telemetry a critical task that demands highly reliable systems.
-The transmitter side of an aeronautical telemetry system (on board the aircraft) is generally composed of transducers that convert a physical stimulus into an electrical signal, multiplexers that combine the measurements into one signal, a transmitting equipment that modulates the data and one (or sometimes two) omnidirectional "blade" antenna(s).
-When the flight test range is going to be beyond $100$ km or when higher link reliability is required, a power amplifier can also be added at the transmitter side \cite{stacey2008aeronautical}.
-As for the receiver side (ground station), it is generally composed of a high gain parabolic dish antenna that allows tracking the aircraft, a low-noise amplifier that amplifies the signal without significantly decreasing the signal-to-noise ratio (SNR), a receiver demodulator that converts the modulated signal back to data, a recorder, and a demultiplexer.
-
-The design and the performance of the transmitting/ receiving equipment are mainly defined by the adopted modulation that conveys the data.
-The first telemetry systems were analog and used frequency modulation (FM) of analog waveforms instead of amplitude modulation (AM) \cite{osti_1143131}.
-This choice favored the use of power amplifiers at their saturation mode without distorting the FM constant envelope signal and provided better bandwidth occupancy/performance trade-off than AM-based systems.
-Then, analog FM has been progressively upgraded to digital FM by feeding the transmitter with filtered digital data instead of analog waveforms.
-This change only required slight modifications of the infrastructure and it did not impact the bandwidth occupancy \cite{osti_1143131}.
-The digital FM version is known as the non-return to zero (NRZ) pulse code modulation/frequency modulation (PCM/FM), and it can be considered as continuous phase frequency shift keying (CPFSK) modulation or more generally as continuous phase modulation (CPM) \cite{digital_phase_mod}.
-
-PCM/FM has been the reference modulation for decades \cite{irig106} because it has proven its robustness as well as its reliability for such high-risk use and because the required data rates were low compared to the available bandwidth before the 1990's.
-In the telemetry literature, this modulation is sometimes called the "legacy modulation" and is still used nowadays in several telemetry setups all over the world.
-However, over the past few years, two major factors have impacted the telemetering community.
-The first one is the exponential expansion of wireless communication systems and especially of the commercial mobile services.
-As an illustration, the number of mobile-broadband subscriptions has grown more than $20\%$ annually in the last five years and has reached $4.3$ billion globally by the end of 2017 \cite{itu_facts}.
-This huge increase has urged telecommunication service providers to seek more bandwidth and has consequently made RF spectrum more and more costly (for instance, the United States (US) government auctioned $25$ MHz of the L-band at a record of $44.89$ billion dollars in 2015 \cite{spec_pr}).
-As a result, the spectrum of aeronautical telemetry has been shortened by almost $35\%$ and reallocated for the commercial applications \cite{ninja_telem}.\newline
-The second factor is the data deluge of aeronautical telemetry, which is due to the use of multiple sensors on board the aircraft to acquire more and more information during the flight.
-This data increase comes also as a consequence of reducing the flight testing duration for cost-effectiveness.
-In fact, significant resources are devoted during the flight testing phase and cost around $50$ thousand dollars in labor per hour \cite{telemetry_eco, eco_imp_tel}.
-Therefore, reducing the test duration without decreasing the amount of transferred data is a major demand for aerospace companies.
-On the one hand, this significant increase in data transfer requires the adoption of more spectrally efficient modulations than PCM/FM.
-On the other hand, it broadens the signal bandwidth, which makes the aeronautical telemetry channel frequency selective \cite{channel_wideband} and therefore impacts the availability of the aeronautical telemetry link.
-The latter issue is a real burden for aerospace companies since one delayed or cancelled test mission due to the unavailability of telemetry link could cost more than $1$ million dollars \cite{eco_imp_tel}.
-
-Both economical and technical challenges have stimulated the flight test telemetry community to adopt more power and bandwidth efficient modulations than PCM/FM.
-This has been done via the advanced range telemetry (ARTM) program \cite{artm_prog}, which resulted in $2$ spectrally efficient CPM modulations: the shaped offset quadrature phase shift keying telemetry group (SOQPSK-TG) modulation and ARTM CPM.
-These modulations figure in the inter-range instrumentation group (IRIG) 106 standard for aeronautical telemetry, which ensures the interoperability of telemetry systems at the range commanders council (RCC) member ranges \cite{irig106}.
-Both modulations are transmitter friendly since they have a constant complex envelope and require minimal infrastructure change.
-Moreover, they allow increasing the spectral efficiency by at least a factor of $2$ compared to PCM/FM while keeping excellent power efficiency.
-These attractive benefits have been achieved by carefully choosing the parameters that define a CPM.
-However, they come at the expense of increasing the receiver complexity, especially when considering complicated scenarios such as the presence of multipath or the use of multiple-antenna transmitters \cite{jensen}.
-This complexity is mainly due to two factors.
-The first one is the presence of an inherent memory that ensures the continuity of the phase of the modulation, and the second one is the nonlinear nature of CPM since the aforementioned modulation is not a linear function of the transmitted symbols.
-Therefore, unlike linear modulations (such as M-PSK, M-quadrature amplitude modulation (M-QAM), amplitude and phase-shift keying (APSK),...), it is not straightforward to take advantage of the properties of linear algebra to develop reduced complexity receivers.
-
-Even if PCM/FM is still widely used in aeronautical telemetry, we observe that more and more actors, especially in the US military ranges, migrated or started the migration to the SOQPSK solution.
-In such a context, the telemetry industry should develop new solutions and algorithms in order to guarantee the best possible telemetry link for these highly demanding customers.
-This work, supervised by Zodiac Data Systems which is one of the leading companies in aeronautical telemetry for civil and military applications, enrolls in this prospective.
-In this thesis, we mainly focus on SOQPSK-TG whose use is getting more and more popular in aeronautical telemetry.
-This modulation belongs to a particular CPM family since it transmits ternary symbols instead of binary ones.
-Simultaneously, it can also be interpreted as an offset modulation, i.e., it can be seen as a modulation whose quadrature component contains a timing offset compared to the in-phase component.
-The goal of this work is to develop reception algorithms for SOQPSK-TG under different scenarios to ensure the availability of aeronautical telemetry link.
-These solutions should maintain a good power efficiency without prohibitive complexity for real-time implementation.
-
-\section*{Previous work on SOQPSK-TG}
-\subsection*{SOQPSK-TG signal representations and approximations}
-SOQPSK-TG is a ternary CPM whose ternary symbols are linked to the binary bits using a specific precoder given in \cite{irig106} and whose modulation index $h=1/2$.
-The role of the precoder is to make the phase of SOQPSK-TG behaves like the phase of an OQPSK signal when both are driven by the same bit sequence \cite{these_perrins}.
-As for the pulse shape that modulates the ternary symbols in the instantaneous frequency, it spans several symbol periods.
-The choice of these parameters increases the spectral efficiency at the expense of creating long memory and inter-symbol interference (ISI).
-Thus, if we take this CPM definition, the resulting reception algorithms would be highly complex.
-Moreover, they would be restricted to simple scenarios due to the nonlinear nature of CPM.
-As a consequence, it is necessary to view this modulation differently to make it more versatile.
-
-The first pioneering work in this direction has been done by E.
-Perrins in \cite{these_perrins} where he showed that SOQPSK-TG could be written as a finite sum of pulse-amplitude modulated (PAM) components.
-The pulses of this decomposition, namely PRD, modulate pseudo-symbols.
-These pseudo-symbols contain memory and are nonlinear functions of the ternary symbols.
-This work is the extension of the Laurent decomposition \cite{pam_representation_binary}, initially developed for binary CPM.
-The advantage of this approach is that it approximates the signal as a sum of only two PAM waveforms.
-However, this does not completely "linearize" the signal because of mapping between the ternary symbols and the pseudo-symbols, which keeps the long memory effect in the approximation.
-
-The second work has been carried by T.
-Nelson in \cite{these_nelson} where he took advantage of the offset nature of SOQPSK-TG and created a direct mapping between the binary bits (i.e., the bits that are used to generate the ternary symbols) and the different waveforms that constitute the eye-diagram of the signal.
-The representation that makes this connection (which is not evident unlike linear modulations) is known as the cross-correlated trellis coded quadrature modulation (XTCQM) representation.
-After some mathematical manipulations, this representation can accurately approximate SOQPSK-TG using a bank of $8$ XTCQM waveforms, each one of them being defined by only $3$ binary bits.
-The advantage of this approach is that it offers an approximation of SOQPSK-TG with a short inherent memory.
-However, it generates data dependent waveforms, which means that we cannot separate the contribution of the bits and the pulse shape in a clear way.
-
-Each representation of SOQPSK-TG has its own benefits and drawbacks. Therefore, the choice between one of them depends on the desired application.
-
-\subsection*{Reduced complexity detection algorithms}
-Several reduced detection algorithms have been proposed in the literature for SOQPSK-TG.
-The first ones are linear integrate and dump (I$\&$D) detectors.
-They can be used to detect the binary symbols thanks to the presence of the SOQPSK-TG precoder (to the best of our knowledge, the reason why is not analytically explained in the literature).
-Several I$\&$D detector implementations are given in \cite{soqpsk_opt_lin_det}, but they are sub-optimal.
-The best one of them suffers from a performance loss of $0.72$ dB for a probability of error of $10^{-5}$ compared to the optimal case.
-Then, very near-optimal performance detectors have been developed, and they rely on the presented above approximations.
-The first one is presented in \cite{demod_perrins} and uses the PAM approximation of PRD and operates using a dynamic trellis of $4$ states.
-The different transitions and the order of the path states in the this trellis depend on the parity of bit index.
-The second very near-optimal reduced complexity detector uses a trellis composed of $16$ states and relies on the XTCQM representation of SOQPSK-TG.
-The performance loss of these detectors is within $0.2$ dB compared to the optimal case.
-\subsection*{Multipath mitigation techniques}
-Due to the bandwidth expansion of SOQPSK-TG signal as highlighted above, the aeronautical telemetry channel becomes frequency selective.
-As a consequence, the receiver may capture several delayed copies of the transmitted signal with different attenuations.
-The transmission is thus made over a multipath channel, which can result in significant performance loss if we use the described above detectors without introducing multipath mitigation techniques, i.e., equalizers.
-Channel estimation and equalization is very sparsely addressed in the literature in the SOQPSK case, contrary to other modulations.
-To the best of our knowledge, very few works \cite{buy_channel_taps2,CMA_soqpsk,CMA_soqpsk2} have focused on multipath mitigation techniques for SOQPSK-TG, and they mainly rely on the constant modulus algorithm (CMA) due to the constant envelope nature of this modulation.
-The performance of CMA can sometimes be unsatisfactory for some channel configurations.
-\subsection*{The two-antenna problem on aeronautical telemetry}
-The previous architectures mentioned above are built regardless of the number of transmitting antennas.
-However, this parameter has an important impact on both transmitter and receiver architectures.
-The first aeronautical telemetry systems used one transmitting antenna fixed underneath the aircraft fuselage and one receiving antenna on a fixed ground station.
-However, this scheme does not guarantee an optimal telemetry link because the receiver cannot always capture the line of sight (LOS) path when the aircraft performs certain maneuvers.
-To overcome this situation, two antennas are then placed on the aircraft in a way that an omnidirectional transmission is guaranteed no matter how the plane is oriented.
-Nevertheless, if the same signal is transmitted over both antennas using the same carrier frequency, the receiver may capture both copies in a destructive way, which causes the loss of the aeronautical telemetry link.
-This problem is known as the two-antenna problem in aeronautical telemetry \cite{two_ant_pblm1}.
-A solution has been proposed in \cite{jensen}, which consists of creating transmit diversity using Alamouti space-time block coding (STBC) \cite{alamouti}.
-This solution significantly reduces the mutual interference of the SOQPSK-TG signals and has been recently standardized in IRIG-106 \cite{irig106}.
-However, it makes recovering the data bits a hard task because of the used modulation as well as the presence of an unusual channel impairment, which is the differential delay.
-This perceptible delay represents a non-negligible portion of the symbol peiod as explained in details in Chapter \ref{chapter:5}.
-It appears since both signals arrive with different delays.
-Several decoding architectures relying on the XTCQM representation have been proposed in \cite{these_nelson, stc_decoder} but only one of them offers satisfying performance at the expense of relatively high implementation complexity.
-Moreover, the best decoder only provides hard outputs and does not contain any mechanism that allows mitigating multipath in the space-time coding (STC) scenario.
-
-\section*{Contributions}
-This dissertation addresses all the aforementioned points and proposes different solutions for each scenario. The important contributions are summarized as follows:
-\begin{itemize}
-\item A new PAM decomposition has been developed for SOQPSK-TG.
-It allows approximating the signal as a simple linear modulation and providing a straightforward mapping between the binary bits and the pseudo-symbols of this decomposition.
-This decomposition, namely DBD in this manuscript, is the root of all the proposed solutions of this work.
-\item A mathematical link has been developed between this proposed PAM decomposition and the XTCQM representation for SOQPSK.  
-\item Several reduced complexity detectors have been proposed for SOQPSK-TG based on DBD.
-Each class of detectors relies on a certain approach (Kaleh approach \cite{simple_coherent}, Ungerboeck approach \cite{mlse_pam} and Forney approach \cite{forney_mlse}).
-\item To combat multipath interference, a joint channel estimation and detection algorithm has been developed for SOQPSK-TG.
-It relies on the proposed decomposition and the per-survivor processing (PSP) principle \cite{polydoros_psp} and offers very attractive bit error rate (BER) performance for different channel configurations.
-\item Two different decoding structures have been proposed for the STC scenario without changing the encoder structure.
-The first one offers a complexity $8$ times less than the state of the art decoder with very acceptable BER performance in the presence of the differential delay.
-The second one outperforms the state of the art decoder while keeping a lower complexity.
-Both proposed solutions can provide soft outputs (i.e., log-likelihood ratios).
-\item A new multipath channel estimator has been developed for the STC scenario and can be used along with the second proposed decoder. 
-\item A new STC scheme has been developed that offers a better transmission rate for SOQPSK-TG than the IRIG standardized one.
-Moreover, the resulting decoding structure can process the even bits separately from the odd ones in the presence of non-zero differential delay and offers very attractive BER performance (within $1.1$ dB of the single input single output (SISO) bound).
-The complexity of the proposed decoder is at least $8$ times less than the state of the art decoder associated with the standardized encoding.
-\end{itemize}
-
-\section*{Dissertation outline}
-This dissertation is organized as follows.
-
-We remind in \textbf{Chapter \ref{chapter:1}} the CPM model and we define its main parameters.
-We then describe the different aeronautical telemetry modulations and we highlight how the spectral efficiency is achieved for each modulation.
-We then focus on SOQPSK and we present the state of the art representations and approximations of this signal (i.e, the PAM decomposition and the XTCQM representation as well as their respective approximations).
-
-In \textbf{Chapter \ref{chapter:2}}, we first introduce a new vision of the precoder structure by decomposing it into two different stages: the first stage is a recursive encoder and the second one is a duobinary encoder.
-We then exploit this new vision to develop a new PAM decomposition of duobinary encoded CPM and we apply the obtained results to SOQPSK.
-We also establish in this chapter a link between the proposed PAM decomposition (DBD) of SOQPSK and its XTCQM representation as well its OQPSK interpretation.
-Furthermore, we detail the exact role of each stage that constitutes the SOQPSK precoder.
-
-In \textbf{Chapter \ref{chapter:3}}, we build several detection architectures as a direct consequence of proposing a new PAM decomposition.
-We take advantage of several approaches of the literature.
-Some rely on the CPM definition of SOQPSK and then exploit the PAM decomposition (Kaleh approach \cite{simple_coherent} and others use the proposed "linearized" model of SOQPSK (Ungerboeck approach \cite{mlse_pam} and Forney approach \cite{forney_mlse}).
-A complexity study of all the proposed detectors is also carried in this chapter.
-
-In \textbf{Chapter \ref{chapter:4}}, we consider the presence of multipath in the SISO scenario.
-We first make a classification of the existing aeronautical telemetry channels of the literature.
-We then present the CMA solution and its different variants.
-We later propose a new joint channel estimation and detection algorithm for SOQPSK-TG.
-This solution is an adapted version of the PSP principle and relies on the proposed PAM decomposition given in Chapter \ref{chapter:2}.
-
-In \textbf{Chapter \ref{chapter:5}}, we focus on the IRIG standardized transmit diversity scheme that allows resolving the two-antenna problem.
-After introducing the IRIG standardized encoder structure and the state of the art decoders, we develop two decoding structures and we present their different variants.
-We then consider the presence of multipath in the STC scenario in the second part of this chapter, and we detail the derivation of the new multipath estimator.
-
-The results given in \textbf{Chapter} \ref{chapter:5} show that the performance of proposed decoders are angle dependent due to employing the standardized scheme.
-For this reason, we propose in \textbf{Chapter \ref{chapter:6}} a new encoding scheme that almost eliminates this angle dependency.
-Moreover, the scheme ameliorates the transmission rate and allows further complexity reduction of the decoding algorithm.
-In the second part of this chapter, we highlight that space-time coding is not the only way to overcome the two-antenna problem.
-We show that sending a SOQPSK-TG signal over one antenna and its artificially delayed copy over the other can also greatly reduce their mutual interference.
-After determining the necessary delay that should be applied, we show that the proposed interference mitigation technique given in Chapter \ref{chapter:4} can also be used in this scenario.
-Thus, the same receiver architecture can be employed whether one or two antennas are mounted on the aircraft using this approach.
-
-%Despite the extensive literature regarding channel estimation and equalization techniques for linear modulations, it is not the case for SOQPSK-TG due to its nonlinear nature.
+\section{CPM signal model}
+\label{sec:CPM_signal_model}

### 2019-01-17 2-Chapters/Conclusion/conclusion.tex
Using template from Rémi Bonnefoi and Rami Othman.

%!TEX root = ../thesis.tex

% First chapter begins here
\chapter*{General Conclusion and Perspectives}
\addcontentsline{toc}{chapter}{General Conclusion}
%\minitoc
\label{chapter:CCL}
% Write miniTOC just after the title
\graphicspath{{2-Chapters/6-Chapter/Images/}}

In this PhD thesis, we focused on studying reception algorithms for SOQPSK-TG under different scenarios. This study first requires a deep understanding of the signal features, which are different from the conventional linear modulations. In fact, SOQPSK-TG is an aeronautical telemetry modulation that has been considered for a long time as a particular CPM due to its ternary alphabet generated using a specific precoder and its long frequency pulse. Moreover, the fact that this modulation is not a linear function of its transmitted symbols makes handling this modulation a hard task. These features have pushed the use of particular decompositions of the signal such as ternary PAM decomposition (PRD) and the XTCQM representation to make the signal more versatile. Each representation has its pros and cons.

On the one hand, the PRD allows approximating the signal as a sum of two linear modulations. However, the resulting pseudo-symbols still inherit the memory effect of the modulation. On the other hand, the XTCQM representation provides an accurate approximation of the signal using a bank of XTCQM waveforms with a small memory. Nevertheless, these waveforms are data dependent and thus we cannot separate contribution of the data independently from the pulse shape. The contrasting features of these representations, which are presented in details in \textbf{Chapter \ref{chapter:1}}, imply a limited use to specific scenarios.

In this work, we first focused on the SOQPSK precoder structure before tackling the signal representation aspect. We showed that this precoder could be decomposed into a recursive encoder that eliminates the long memory of the modulation and a duobinary encoder that converts the binary symbols to ternary ones to increase the spectral efficiency. Thanks to this new vision of the precoder, we showed that SOQPSK-TG could be seen as a duobinary CPM whose PAM decomposition, namely DBD, retains the key features of the Laurent decomposition of binary CPM. As a consequence, SOQPSK-TG can accurately be approximated as a single PAM or as a sum of two PAMs. Moreover, the presence of the recursive encoder has been exploited to reduce the memory effect in the expressions of the pseudo-symbols and to get a direct mapping between the pseudo-symbols and the binary data. Therefore, we proposed a new representation of SOQPSK that gathers better features than the state of the art representations as fully described in \textbf{Chapter \ref{chapter:2}}.

As a direct consequence of this new PAM decomposition, we proposed in \textbf{Chapter \ref{chapter:3}} a wide range of reduced complexity detectors via several approaches (Kaleh approach, Ungerboeck approach, and Forney approach). The very near-optimal performance of these detectors confirms the accuracy of the proposed DBD approximations. However, they also show that the main pulse of DBD generates significant inter-symbol interference that should be mitigated at the receiver side.

In \textbf{Chapter \ref{chapter:4}}, we took advantage of the PAM decomposition (with DBD) to propose a new joint channel estimation and detection algorithm for SOQPSK-TG. This algorithm is inspired by the PSP principle to mitigate the multipath effects, which are present in the aeronautical telemetry environment. This solution operates using two different approximations of the signal and significantly outperforms the constant modulus algorithm.

In \textbf{Chapter \ref{chapter:5}}, we focused on the scenario where two antennas are mounted on the aircraft to get an omnidirectional transmission to resolve the two-antenna problem. This use case requires creating diversity by using Alamouti encoding. However, it makes the decoding process different from what has been proposed for linear modulations because of the used modulation (SOQPSK-TG) and because of the presence of a differential delay. After presenting the state of the art solutions, we proposed in this chapter two main decoding structures. The first one relies on the first PAM approximation of SOQPSK-TG, i.e., by considering that SOQPSK-TG is a linear modulation (or more precisely, an offset modulation) and provides acceptable BER performance while benefiting from an important complexity reduction compared to the state of the art solution. The second decoding structure considers that SOQPSK-TG can be approximated as a sum of two linear modulations and greatly outperforms the state of the art solution. Both structures take into account the presence of the differential delay and provide soft outputs. 
In \textbf{Chapter \ref{chapter:5}}, we also considered the presence of multipath in the STC scenario. After formulating the problem using multipath, we proposed a new channel estimator and showed that this solution could be used along with the second decoding structure.

In the final chapter of this dissertation, we explored other solutions to resolve the two-antenna problem. The first proposed solution takes into account the observations made in Chapter \ref{chapter:5} where it has been shown that the standardized encoding scheme does not eliminate the mutual interference of SOQPSK-TG signals. For this reason, we proposed a new encoding scheme that is more suitable for SOQPSK-TG. This scheme allows gathering all the attractive features of a STC system since it provides better transmission rate than the state of the art encoding and results in a reduced complexity decoding structure with attractive BER performance (within $1.1$ dB of the SISO bound). In the second part of \textbf{Chapter \ref{chapter:6}}, we considered the time diversity approach, and we proposed sending one signal over one antenna and its time delayed version over the other. This delay is created artificially to reduce the mutual interference of the signals. The advantage of this approach is that the receiver considers that the transmitted signal is corrupted by a multipath channel whose interference can successfully be mitigated using the techniques introduced in Chapter \ref{chapter:4}.

\section*{Perspectives}
We present in the following the possible areas of future studies
\subsubsection*{Equalization in the SISO scenario:} 
\begin{itemize}
\item In this dissertation, the PSP algorithm operates using the duobinary PAM decomposition (DBD) of SOQPSK-TG. However, thanks to the mapping found between the pseudo-symbols and the data bits using PRD (Chapter \ref{chapter:2}), it may be possible to apply the PSP principle to the PRD PAM approximation. Since both PRD and DBD have different characteristics, the resulting algorithms may have different behaviors. 
\item  To the best of our knowledge, the INET packet structure does not contain any cyclic prefix. However, a new packet structure can be imagined for SOQPSK-TG so that the frequency equalization algorithms introduced in \cite{freq_eqz_cpm1,freq_eqz_cpm2,thillo_eqz,chayot_eqz} can be used, tested in the presence of aeronautical telemetry channels, and compared with the proposed algorithms of this work.
\end{itemize}

%\subsubsection*{Channel coding in the SISO scenario:}   
\subsubsection*{Decoding and Multipath mitigation in the STC scenario:}
\begin{itemize}
\item In this work, we proposed two decoding structures for the STC scenario. The first one relies on Ungerboeck/Cavers approach, and the second one uses the least squares approach (which is equivalent to Forney approach in the SISO case). These decoders take advantage of the "linearized" version of SOQPSK-TG but they do not take into account its CPM nature. Thus, it may be relevant to extend Kaleh approach (which exploits the constant envelope nature of CPM when calculating the log-likelihood function) from the SISO case to the MISO one to get a third decoding structure. 
\item The multipath mitigation technique developed in this dissertation for the STC scenario only works using the second proposed decoding structure. Further works regarding frequency domain or time domain equalizers need to be done for the STC scenario. Ideally, the equalizer should work regardless of the used decoding structure. 
\end{itemize}

\subsubsection*{STC estimators:}
\begin{itemize}
\item  The complexity of the developed estimators in \cite{STC_part1} is relatively high. More work needs to be done to reduce the complexity, and it may be accomplished using the new PAM decomposition.   
\item The simulation results given in Section \ref{chap5_sec:multipath} show that the estimators are highly multipath-sensitive. More work needs to be done to improve their performance for this scenario.
\end{itemize}
\subsubsection*{New aeronautical telemetry systems:}
Conceiving power and spectrally efficient systems is a constant challenge for aeronautical telemetry. However, achieving this goal requires evolving the IRIG standard as well as the actual infrastructure. We can imagine for instance 
\begin{itemize}
\item Pairing MIMO systems with SOQPSK to increase the spectral efficiency while taking advantage of the constant envelope nature of the modulation. The MISO case has already been standardized in IRIG to resolve the two-antenna problem without increasing the spectral efficiency. Therefore, we can imagine an extension of this concept to reach the latter goal. A feasibility study should also be carried to determine how many antennas can be mounted on board the aircraft and how many receiving antennas can be used. 
\item  Defining different modes for different CPM modulations as done in the tactical communications standards \cite{nato2}. Each modulation fulfills a certain spectral efficiency and power trade-off. Switching between one mode to the other can be done manually or autonomously (the latter case may require an ulplink mode).  
\item Adopting non-constant envelope signals, which beget better spectral efficiency than CPM such as COFDM \cite{channel_cband_zds} or APSK \cite{apsk}. 
\end{itemize}

### 2019-01-17 4-Listofpublications/Listofpublications.tex
Using template from Rémi Bonnefoi and Rami Othman.

%!TEX root = ../thesis.tex
% ******************************* Thesis Appendix A ****************************
\chapter*{List of Publications}
\addcontentsline{toc}{chapter}{List of Publications}
\section*{International Journals}
%{enumi}
%\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
%\begin{enumerate}{[IJ-\arabic{enumi}]}%[label={[RI-\arabic*]}]
\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
\begin{etaremune}%[label={[RI-\arabic*]}]
	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y. `` Reduced Complexity Detectors Based on SOQPSK Linearzations,'', in \emph{IEEE Transactions on Aerospace and Electronic Systems}, 2018 (Major Revision)
	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y., `` PAM Decomposition of Ternary CPM With Duobinary Encoding,'', in \emph{IEEE Transactions on Communications}, vol. 65, no. 10, pp. 4274-4284, doi: 10.1109/TCOMM.2017.2723567, Oct 2017
\end{etaremune}

\section*{International Conferences}
%\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
%\begin{enumerate}{[IC-\arabic{enumi}]}
\begin{etaremune}
	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Linear Decision Feedback Detector for SOQPSK,'', in \emph{URSI AP-RASC}, 2019 (Submitted)
	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``A Reduced Complexity OQPSK-Type Detector For SOQPSK,'', in \emph{IEEE 88th Vehicular Technology Conference (VTC-Fall)},Chicago,Sep 2018 (to be published)
	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Joint Channel Estimation and Detection of SOQPSK Using the PAM Decomposition,'', in \emph{IEEE 25th International Conference on Telecommunications (ICT)}, St. Malo, pp. 154-158. doi: 10.1109/ICT.2018.8464828, Jun 2018 
\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Analysis of Duobinary Encoding for CPM Signals,'', in \emph{URSI General Assembly and Symposium}, Montreal, Aug 2017
\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y. `` Phase offset estimation of SOQPSK waveform by the analysis of the angle distribution,'', in \emph{IEEE 13th International Symposium on Wireless Communication Systems}, Poznan, pp. 16-21. doi: 10.1109/ISWCS.2016.7600847, Sep 2016  
\end{etaremune}

\section*{National Conferences}
\renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
\begin{etaremune}
\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``D\'{e}composition PAM d'un signal CPM avec encodage duo-binaire - Application \`{a} la modulation SOQPSK,'', in \emph{GRETSI}, Juan-Les-Pins, France, Sep 2017 
\end{etaremune}

\section*{Other National Communications}
\renewcommand{\labelenumi}{[ONC-\arabic{enumi}]}
\begin{etaremune}
\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``A New PAM Decomposition for SOQPSK in Aernautical Telemetry Context,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017 
\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{SCEE Seminar}, Rennes, France, Jun 2017 
\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016 	
\end{etaremune}

\section*{Patents}
\renewcommand{\labelenumi}{[PT-\arabic{enumi}]}
\begin{etaremune}
\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. `` Nouvelle architecture de d\'{e}codage spatio-temporel bas\'{e}e sur la d\'{e}composition PAM pour la modulation SOQPSK-TG,'' (pending)
\end{etaremune}

### 2019-01-17 Nomenclature/Nomenclature.tex
Using template from Rémi Bonnefoi and Rami Othman.

\nomenclature[A]{AM}{\textit{Amplitude Modulation}}
\nomenclature[A]{ARTM}{\textit{Advanced Range Telemetry Group}}
\nomenclature[A]{COFDM}{\textit{Coded Orthogonal Frequency Division Multiplexing}}
\nomenclature[A]{CPM}{\textit{Continuous Phase Modulation}}
\nomenclature[A]{CPFSK}{\textit{Continuous Phase Frequency Shift Keying}}
\nomenclature[A]{FM}{\textit{Frequency Modulation}}
\nomenclature[A]{PCM/FM}{\textit{Pulse Code Modulation/Frequency Modulation}}
\nomenclature[A]{MHz}{\textit{Megahertz}}
\nomenclature[A]{GHz}{\textit{Gigahertz}}
\nomenclature[A]{Hz}{\textit{Hertz}}
\nomenclature[A]{dB}{\textit{Decibel}}
\nomenclature[A]{NRZ}{\textit{Non-Return to Zero}}
\nomenclature[A]{SOQPSK}{\textit{Shaped Offset Quadrature Phase-Shift Keying}}
\nomenclature[A]{SOQPSK-TG}{\textit{Shaped Offset Quadrature Phase-Shift Keying Telemetry Group}}
\nomenclature[A]{SOQPSK-MIL}{\textit{Shaped Offset Quadrature Phase-Shift Keying Military}}
\nomenclature[A]{MIL-STD SOQPSK}{\textit{Military Standard Shaped Offset Quadrature Phase-Shift Keying }}
\nomenclature[A]{APSK}{\textit{Amplitude Phase-Shift Keying}}
\nomenclature[A]{PSK}{\textit{Phase-Shift Keying}}
\nomenclature[A]{M-PSK}{\textit{M-ary Phase-Shift Keying}}
\nomenclature[A]{BPSK}{\textit{Binary Phase-Shift Keying}}
\nomenclature[A]{OQPSK}{\textit{Offset Quadrature Phase-Shift Keying}}
\nomenclature[A]{PAM}{\textit{Pulse Amplitude Modulation}}
\nomenclature[A]{M-QAM}{\textit{M-ary Quadrature Amplitude Modulation}}
\nomenclature[A]{MIMO}{\textit{Multiple Input Multiple Output}}
\nomenclature[A]{SISO}{\textit{Single Input Single Output}}
\nomenclature[A]{MISO}{\textit{Multiple Input Single Output}}
\nomenclature[A]{RF}{\textit{Radio Frequency}}
\nomenclature[A]{IRIG}{\textit{Inter-Range Instrumentation Group}}
\nomenclature[A]{RCC}{\textit{Range Commanders Council}}
\nomenclature[A]{ISI}{\textit{Inter-Symbol Interference}}
\nomenclature[A]{PRD}{\textit{Perrins-Rice Decomposition}}
\nomenclature[A]{DBD}{\textit{DuoBinary Decomposition}}
\nomenclature[A]{XTCQM}{\textit{Cross-correlated Trellis Coded Modulation}}
\nomenclature[A]{I$\&$D}{\textit{Integrate and Dump}}
\nomenclature[A]{CM}{\textit{Constant Modulus}}
\nomenclature[A]{CMA}{\textit{Constant Modulus Algorithm}}
\nomenclature[A]{LOS}{\textit{Line Of Sight}}
\nomenclature[A]{NLOS}{\textit{Non-Line Of Sight}}
\nomenclature[A]{STC}{\textit{Space-Time Coding}}
\nomenclature[A]{STBC}{\textit{Space-Time Block Coding}}
\nomenclature[A]{PSP}{\textit{PerSurvivor Processing}}
\nomenclature[A]{GMSK}{\textit{Gaussian Minimum Shift Keying}}
\nomenclature[A]{MSK}{\textit{Minimum Shift Keying}}
\nomenclature[A]{LS}{\textit{Least Square}}
\nomenclature[A]{LMS}{\textit{Least Mean Square}}
\nomenclature[A]{RLS}{\textit{Recursive Least Square}}
\nomenclature[A]{GSM}{\textit{Global System for Mobile communication }}
\nomenclature[A]{BT}{\textit{Bandwidth Time product}}
\nomenclature[A]{GFSK}{\textit{Gaussian Frequency Shift Keying}}
\nomenclature[A]{SBPSK}{\textit{Shaped Binary Phase-Shift Keying}}
\nomenclature[A]{DVB-RCS2}{\textit{Digital Video Broadcasting - Return Channel via Satellite - Second Generation}}
\nomenclature[A]{FQPSK}{\textit{Feher-patented Quadrature Phase-Shift Keying}}
\nomenclature[A]{RC}{\textit{Raised Cosine}}
\nomenclature[A]{REC}{\textit{Rectangular}}
\nomenclature[A]{BCJR}{\textit{Bahl, Cocke, Jelinek and Raviv}}
\nomenclature[A]{MAP}{\textit{Maximum A Posteriori}}
\nomenclature[A]{MLSD}{\textit{Maximum Likelihood Sequence Detection}}
\nomenclature[A]{CA-MLSD}{\textit{Conventional Adaptive Maximum Likelihood Sequence Detection}}
\nomenclature[A]{NA-MLSD}{\textit{Non-Adaptive Maximum Likelihood Sequence Detection}}
\nomenclature[A]{MLSD}{\textit{Maximum Likelihood Sequence Detection}}
\nomenclature[A]{MLSE}{\textit{Maximum Likelihood Sequence Estimation}}
\nomenclature[A]{Mbps}{\textit{Megabits per second}}
\nomenclature[A]{MMSE}{\textit{Minimum Mean Square Error}}
\nomenclature[A]{MSE}{\textit{Mean Square Error}}
\nomenclature[A]{SMA}{\textit{Saint Martin Airport}}
\nomenclature[A]{TBA}{\textit{Toulouse-Blagnac Airport}}
\nomenclature[A]{V2V}{\textit{Vehicle To Vehicle}}
\nomenclature[A]{AWGN}{\textit{Additive White Gaussian Noise}}
\nomenclature[A]{SxS}{\textit{Sample by Sample}}
\nomenclature[A]{SNR}{\textit{Signal-to-Noise Ratio}}
\nomenclature[A]{SER}{\textit{Symbol Error Rate}}
\nomenclature[A]{BER}{\textit{Bit Error Rate}}
\nomenclature[A]{ASM}{\textit{Asynchronous Marker}}
\nomenclature[A]{INET}{\textit{Integrated Network Enhanced Telemetry}}
\nomenclature[A]{STBD}{\textit{Space-Time Block Decoder}}
\nomenclature[A]{ML}{\textit{Maximum Likelihood}}
\nomenclature[A]{ML-ASM}{\textit{Maximum Likelihood - Approximate Signal Model}}
\nomenclature[A]{OD}{\textit{Offset Decode}}
\nomenclature[A]{SOVA}{\textit{Soft Output Viterbi Algorithm}}
\nomenclature[A]{LLR}{\textit{Log-Likelihood Ratio}}
\nomenclature[A]{FIR}{\textit{Finite Impulse Response}}
\nomenclature[A]{Max-Log-MAP}{\textit{Maximum Logarithm Maximum A Posteriori}}
\nomenclature[A]{FEC}{\textit{Forward Error Correction}}
\nomenclature[A]{LDPC}{\textit{Low Parity Density Check}}
\nomenclature[A]{CFO}{\textit{Carrier Frequency Offset}}
\nomenclature[A]{US}{\textit{United States}}
\nomenclature[A]{DF}{\textit{Decision Feedback}}
\nomenclature[A]{I.I.D}{\textit{Independent and Identically Distributed}}
\nomenclature[A]{ML-ASM}{\textit{Maximum Likelihood - Approximate Signal Model}}
\nomenclature[A]{NASA's JPL}{\textit{National Aeronautics and Space Administration's Jet Propulsion Lab}}
%\nomenclature[N]{$x$}{Scalar}

### 2019-01-17 1-Preamble/4-Acknowledgement/acknowledgement.tex
Okay, all the template is better for my need now. I have to write…

% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}

And I would like to acknowledge ...

See https://perso.crans.org/besson/phd/thanks/

\TODO{To start!}

\end{acknowledgements}

### 2019-01-17 1-Preamble/5-Abstract/abstract.tex
Okay, all the template is better for my need now. I have to write…

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}

% ----------------------------------------------------------------------------
Blabla.

Dans cette thèse, blabla.

\newpage
\begin{center}
	\textbf{{\Large Abstract}}
\end{center}
\vspace{1cm}

% ----------------------------------------------------------------------------
Blabla.

In this PhD thesis, blabla.

\end{abstract}

### 2019-01-17 1-Preamble/5-Abstract/resume_fr.tex
Okay, all the template is better for my need now. I have to write…

\begin{resume_fr}

Résumé de la thèse en français.

% ----------------------------------------------------------------------------
\section*{Partie 1}
\subsection*{Sous partie 1.1}
\subsection*{Sous partie 1.2}
\subsection*{Sous partie 1.3}

% ----------------------------------------------------------------------------
\section*{Partie 2}
\subsection*{Sous partie 2.1}
\subsection*{Sous partie 2.2}
\subsection*{Sous partie 2.3}

% ----------------------------------------------------------------------------
\section*{Partie 3}
\subsection*{Sous partie 3.1}
\subsection*{Sous partie 3.2}
\subsection*{Sous partie 3.3}

% ----------------------------------------------------------------------------
\section*{Contributions}

Cette thèse aborde tous les points susmentionnés et propose différentes solutions pour chaque scénario.
Les contributions importantes sont résumées ci-dessous :
\begin{itemize}
\item Item 1
\item Item 2
\item Item 3
\item Item 4
\end{itemize}

% ----------------------------------------------------------------------------
\section*{Organisation du manuscrit}

Ce manuscrit est organisé comme suit.
Nous rappelons dans le \textbf{Chapitre \ref{chapter:1}} blabla.

Dans le \textbf{Chapitre \ref{chapter:2}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:3}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:4}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:5}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:5}} blabla.

\end{resume_fr}

### 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
Okay, all the template is better for my need now. I have to write…

@@
 \graphicspath{{2-Chapters/1-Chapter/Images/}}
+
 % Write miniTOC just after the title
@@
 %\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
+
+% ----------------------------------------------------------------------------
 \section*{General context}
-\indent The work presented in this PhD thesis applies in aeronautical telemetry, which is the process of remotely collecting measurements done by electronic equipment on board the aircraft and transmitting these data to a distant location for monitoring, display, and recording \cite{telemetry_def}.
 
-An aeronautical telemetry system is used during the flight testing phase to monitor the behavior of the plane by transmitting the aircraft dynamics settings (such as velocities, vibrations, temperatures, stress and strain parameters,...) from the aircraft to the ground station over a radio-frequency (RF) link.
-Aeronautical telemetry is a one-way link from the plane to the ground (i.e., there is no uplink mode) that requires real-time data transmission and analysis to ensure the pilot's safety.
-This makes the use of the aeronautical telemetry a critical task that demands highly reliable systems.
-The transmitter side of an aeronautical telemetry system (on board the aircraft) is generally composed of transducers that convert a physical stimulus into an electrical signal, multiplexers that combine the measurements into one signal, a transmitting equipment that modulates the data and one (or sometimes two) omnidirectional "blade" antenna(s).
-When the flight test range is going to be beyond $100$ km or when higher link reliability is required, a power amplifier can also be added at the transmitter side \cite{stacey2008aeronautical}.
-As for the receiver side (ground station), it is generally composed of a high gain parabolic dish antenna that allows tracking the aircraft, a low-noise amplifier that amplifies the signal without significantly decreasing the signal-to-noise ratio (SNR), a receiver demodulator that converts the modulated signal back to data, a recorder, and a demultiplexer.
-
-The design and the performance of the transmitting/ receiving equipment are mainly defined by the adopted modulation that conveys the data.
-The first telemetry systems were analog and used frequency modulation (FM) of analog waveforms instead of amplitude modulation (AM) \cite{osti_1143131}.
-This choice favored the use of power amplifiers at their saturation mode without distorting the FM constant envelope signal and provided better bandwidth occupancy/performance trade-off than AM-based systems.
-Then, analog FM has been progressively upgraded to digital FM by feeding the transmitter with filtered digital data instead of analog waveforms.
-This change only required slight modifications of the infrastructure and it did not impact the bandwidth occupancy \cite{osti_1143131}.
-The digital FM version is known as the non-return to zero (NRZ) pulse code modulation/frequency modulation (PCM/FM), and it can be considered as continuous phase frequency shift keying (CPFSK) modulation or more generally as continuous phase modulation (CPM) \cite{digital_phase_mod}.
-
-PCM/FM has been the reference modulation for decades \cite{irig106} because it has proven its robustness as well as its reliability for such high-risk use and because the required data rates were low compared to the available bandwidth before the 1990's.
-In the telemetry literature, this modulation is sometimes called the "legacy modulation" and is still used nowadays in several telemetry setups all over the world.
-However, over the past few years, two major factors have impacted the telemetering community.
-The first one is the exponential expansion of wireless communication systems and especially of the commercial mobile services.
-As an illustration, the number of mobile-broadband subscriptions has grown more than $20\%$ annually in the last five years and has reached $4.3$ billion globally by the end of 2017 \cite{itu_facts}.
-This huge increase has urged telecommunication service providers to seek more bandwidth and has consequently made RF spectrum more and more costly (for instance, the United States (US) government auctioned $25$ MHz of the L-band at a record of $44.89$ billion dollars in 2015 \cite{spec_pr}).
-As a result, the spectrum of aeronautical telemetry has been shortened by almost $35\%$ and reallocated for the commercial applications \cite{ninja_telem}.\newline
-The second factor is the data deluge of aeronautical telemetry, which is due to the use of multiple sensors on board the aircraft to acquire more and more information during the flight.
-This data increase comes also as a consequence of reducing the flight testing duration for cost-effectiveness.
-In fact, significant resources are devoted during the flight testing phase and cost around $50$ thousand dollars in labor per hour \cite{telemetry_eco, eco_imp_tel}.
-Therefore, reducing the test duration without decreasing the amount of transferred data is a major demand for aerospace companies.
-On the one hand, this significant increase in data transfer requires the adoption of more spectrally efficient modulations than PCM/FM.
-On the other hand, it broadens the signal bandwidth, which makes the aeronautical telemetry channel frequency selective \cite{channel_wideband} and therefore impacts the availability of the aeronautical telemetry link.
-The latter issue is a real burden for aerospace companies since one delayed or cancelled test mission due to the unavailability of telemetry link could cost more than $1$ million dollars \cite{eco_imp_tel}.
-
-Both economical and technical challenges have stimulated the flight test telemetry community to adopt more power and bandwidth efficient modulations than PCM/FM.
-This has been done via the advanced range telemetry (ARTM) program \cite{artm_prog}, which resulted in $2$ spectrally efficient CPM modulations: the shaped offset quadrature phase shift keying telemetry group (SOQPSK-TG) modulation and ARTM CPM.
-These modulations figure in the inter-range instrumentation group (IRIG) 106 standard for aeronautical telemetry, which ensures the interoperability of telemetry systems at the range commanders council (RCC) member ranges \cite{irig106}.
-Both modulations are transmitter friendly since they have a constant complex envelope and require minimal infrastructure change.
-Moreover, they allow increasing the spectral efficiency by at least a factor of $2$ compared to PCM/FM while keeping excellent power efficiency.
-These attractive benefits have been achieved by carefully choosing the parameters that define a CPM.
-However, they come at the expense of increasing the receiver complexity, especially when considering complicated scenarios such as the presence of multipath or the use of multiple-antenna transmitters \cite{jensen}.
-This complexity is mainly due to two factors.
-The first one is the presence of an inherent memory that ensures the continuity of the phase of the modulation, and the second one is the nonlinear nature of CPM since the aforementioned modulation is not a linear function of the transmitted symbols.
-Therefore, unlike linear modulations (such as M-PSK, M-quadrature amplitude modulation (M-QAM), amplitude and phase-shift keying (APSK),...), it is not straightforward to take advantage of the properties of linear algebra to develop reduced complexity receivers.
-
-Even if PCM/FM is still widely used in aeronautical telemetry, we observe that more and more actors, especially in the US military ranges, migrated or started the migration to the SOQPSK solution.
-In such a context, the telemetry industry should develop new solutions and algorithms in order to guarantee the best possible telemetry link for these highly demanding customers.
-This work, supervised by Zodiac Data Systems which is one of the leading companies in aeronautical telemetry for civil and military applications, enrolls in this prospective.
-In this thesis, we mainly focus on SOQPSK-TG whose use is getting more and more popular in aeronautical telemetry.
-This modulation belongs to a particular CPM family since it transmits ternary symbols instead of binary ones.
-Simultaneously, it can also be interpreted as an offset modulation, i.e., it can be seen as a modulation whose quadrature component contains a timing offset compared to the in-phase component.
-The goal of this work is to develop reception algorithms for SOQPSK-TG under different scenarios to ensure the availability of aeronautical telemetry link.
-These solutions should maintain a good power efficiency without prohibitive complexity for real-time implementation.
-
-\section*{Previous work on SOQPSK-TG}
-\subsection*{SOQPSK-TG signal representations and approximations}
-SOQPSK-TG is a ternary CPM whose ternary symbols are linked to the binary bits using a specific precoder given in \cite{irig106} and whose modulation index $h=1/2$.
-The role of the precoder is to make the phase of SOQPSK-TG behaves like the phase of an OQPSK signal when both are driven by the same bit sequence \cite{these_perrins}.
-As for the pulse shape that modulates the ternary symbols in the instantaneous frequency, it spans several symbol periods.
-The choice of these parameters increases the spectral efficiency at the expense of creating long memory and inter-symbol interference (ISI).
-Thus, if we take this CPM definition, the resulting reception algorithms would be highly complex.
-Moreover, they would be restricted to simple scenarios due to the nonlinear nature of CPM.
-As a consequence, it is necessary to view this modulation differently to make it more versatile.
-
-The first pioneering work in this direction has been done by E.
-Perrins in \cite{these_perrins} where he showed that SOQPSK-TG could be written as a finite sum of pulse-amplitude modulated (PAM) components.
-The pulses of this decomposition, namely PRD, modulate pseudo-symbols.
-These pseudo-symbols contain memory and are nonlinear functions of the ternary symbols.
-This work is the extension of the Laurent decomposition \cite{pam_representation_binary}, initially developed for binary CPM.
-The advantage of this approach is that it approximates the signal as a sum of only two PAM waveforms.
-However, this does not completely "linearize" the signal because of mapping between the ternary symbols and the pseudo-symbols, which keeps the long memory effect in the approximation.
-
-The second work has been carried by T.
-Nelson in \cite{these_nelson} where he took advantage of the offset nature of SOQPSK-TG and created a direct mapping between the binary bits (i.e., the bits that are used to generate the ternary symbols) and the different waveforms that constitute the eye-diagram of the signal.
-The representation that makes this connection (which is not evident unlike linear modulations) is known as the cross-correlated trellis coded quadrature modulation (XTCQM) representation.
-After some mathematical manipulations, this representation can accurately approximate SOQPSK-TG using a bank of $8$ XTCQM waveforms, each one of them being defined by only $3$ binary bits.
-The advantage of this approach is that it offers an approximation of SOQPSK-TG with a short inherent memory.
-However, it generates data dependent waveforms, which means that we cannot separate the contribution of the bits and the pulse shape in a clear way.
-
-Each representation of SOQPSK-TG has its own benefits and drawbacks. Therefore, the choice between one of them depends on the desired application.
-
-\subsection*{Reduced complexity detection algorithms}
-Several reduced detection algorithms have been proposed in the literature for SOQPSK-TG.
-The first ones are linear integrate and dump (I$\&$D) detectors.
-They can be used to detect the binary symbols thanks to the presence of the SOQPSK-TG precoder (to the best of our knowledge, the reason why is not analytically explained in the literature).
-Several I$\&$D detector implementations are given in \cite{soqpsk_opt_lin_det}, but they are sub-optimal.
-The best one of them suffers from a performance loss of $0.72$ dB for a probability of error of $10^{-5}$ compared to the optimal case.
-Then, very near-optimal performance detectors have been developed, and they rely on the presented above approximations.
-The first one is presented in \cite{demod_perrins} and uses the PAM approximation of PRD and operates using a dynamic trellis of $4$ states.
-The different transitions and the order of the path states in the this trellis depend on the parity of bit index.
-The second very near-optimal reduced complexity detector uses a trellis composed of $16$ states and relies on the XTCQM representation of SOQPSK-TG.
-The performance loss of these detectors is within $0.2$ dB compared to the optimal case.
-\subsection*{Multipath mitigation techniques}
-Due to the bandwidth expansion of SOQPSK-TG signal as highlighted above, the aeronautical telemetry channel becomes frequency selective.
-As a consequence, the receiver may capture several delayed copies of the transmitted signal with different attenuations.
-The transmission is thus made over a multipath channel, which can result in significant performance loss if we use the described above detectors without introducing multipath mitigation techniques, i.e., equalizers.
-Channel estimation and equalization is very sparsely addressed in the literature in the SOQPSK case, contrary to other modulations.
-To the best of our knowledge, very few works \cite{buy_channel_taps2,CMA_soqpsk,CMA_soqpsk2} have focused on multipath mitigation techniques for SOQPSK-TG, and they mainly rely on the constant modulus algorithm (CMA) due to the constant envelope nature of this modulation.
-The performance of CMA can sometimes be unsatisfactory for some channel configurations.
-\subsection*{The two-antenna problem on aeronautical telemetry}
-The previous architectures mentioned above are built regardless of the number of transmitting antennas.
-However, this parameter has an important impact on both transmitter and receiver architectures.
-The first aeronautical telemetry systems used one transmitting antenna fixed underneath the aircraft fuselage and one receiving antenna on a fixed ground station.
-However, this scheme does not guarantee an optimal telemetry link because the receiver cannot always capture the line of sight (LOS) path when the aircraft performs certain maneuvers.
-To overcome this situation, two antennas are then placed on the aircraft in a way that an omnidirectional transmission is guaranteed no matter how the plane is oriented.
-Nevertheless, if the same signal is transmitted over both antennas using the same carrier frequency, the receiver may capture both copies in a destructive way, which causes the loss of the aeronautical telemetry link.
-This problem is known as the two-antenna problem in aeronautical telemetry \cite{two_ant_pblm1}.
-A solution has been proposed in \cite{jensen}, which consists of creating transmit diversity using Alamouti space-time block coding (STBC) \cite{alamouti}.
-This solution significantly reduces the mutual interference of the SOQPSK-TG signals and has been recently standardized in IRIG-106 \cite{irig106}.
-However, it makes recovering the data bits a hard task because of the used modulation as well as the presence of an unusual channel impairment, which is the differential delay.
-This perceptible delay represents a non-negligible portion of the symbol peiod as explained in details in Chapter \ref{chapter:5}.
-It appears since both signals arrive with different delays.
-Several decoding architectures relying on the XTCQM representation have been proposed in \cite{these_nelson, stc_decoder} but only one of them offers satisfying performance at the expense of relatively high implementation complexity.
-Moreover, the best decoder only provides hard outputs and does not contain any mechanism that allows mitigating multipath in the space-time coding (STC) scenario.
+\indent
+The work presented in this PhD thesis
+Blabla
 
 \section*{Contributions}
-This dissertation addresses all the aforementioned points and proposes different solutions for each scenario. The important contributions are summarized as follows:
+
+Blabla
+
 \begin{itemize}
-\item A new PAM decomposition has been developed for SOQPSK-TG.
-It allows approximating the signal as a simple linear modulation and providing a straightforward mapping between the binary bits and the pseudo-symbols of this decomposition.
-This decomposition, namely DBD in this manuscript, is the root of all the proposed solutions of this work.
-\item A mathematical link has been developed between this proposed PAM decomposition and the XTCQM representation for SOQPSK.  
-\item Several reduced complexity detectors have been proposed for SOQPSK-TG based on DBD.
-Each class of detectors relies on a certain approach (Kaleh approach \cite{simple_coherent}, Ungerboeck approach \cite{mlse_pam} and Forney approach \cite{forney_mlse}).
-\item To combat multipath interference, a joint channel estimation and detection algorithm has been developed for SOQPSK-TG.
-It relies on the proposed decomposition and the per-survivor processing (PSP) principle \cite{polydoros_psp} and offers very attractive bit error rate (BER) performance for different channel configurations.
-\item Two different decoding structures have been proposed for the STC scenario without changing the encoder structure.
-The first one offers a complexity $8$ times less than the state of the art decoder with very acceptable BER performance in the presence of the differential delay.
-The second one outperforms the state of the art decoder while keeping a lower complexity.
-Both proposed solutions can provide soft outputs (i.e., log-likelihood ratios).
-\item A new multipath channel estimator has been developed for the STC scenario and can be used along with the second proposed decoder. 
-\item A new STC scheme has been developed that offers a better transmission rate for SOQPSK-TG than the IRIG standardized one.
-Moreover, the resulting decoding structure can process the even bits separately from the odd ones in the presence of non-zero differential delay and offers very attractive BER performance (within $1.1$ dB of the single input single output (SISO) bound).
-The complexity of the proposed decoder is at least $8$ times less than the state of the art decoder associated with the standardized encoding.
+\item Blabla
+\item Blabla
+\item Blabla
 \end{itemize}
@@
 
-We remind in \textbf{Chapter \ref{chapter:1}} the CPM model and we define its main parameters.
-We then describe the different aeronautical telemetry modulations and we highlight how the spectral efficiency is achieved for each modulation.
-We then focus on SOQPSK and we present the state of the art representations and approximations of this signal (i.e, the PAM decomposition and the XTCQM representation as well as their respective approximations).
-
-In \textbf{Chapter \ref{chapter:2}}, we first introduce a new vision of the precoder structure by decomposing it into two different stages: the first stage is a recursive encoder and the second one is a duobinary encoder.
-We then exploit this new vision to develop a new PAM decomposition of duobinary encoded CPM and we apply the obtained results to SOQPSK.
-We also establish in this chapter a link between the proposed PAM decomposition (DBD) of SOQPSK and its XTCQM representation as well its OQPSK interpretation.
-Furthermore, we detail the exact role of each stage that constitutes the SOQPSK precoder.
-
-In \textbf{Chapter \ref{chapter:3}}, we build several detection architectures as a direct consequence of proposing a new PAM decomposition.
-We take advantage of several approaches of the literature.
-Some rely on the CPM definition of SOQPSK and then exploit the PAM decomposition (Kaleh approach \cite{simple_coherent} and others use the proposed "linearized" model of SOQPSK (Ungerboeck approach \cite{mlse_pam} and Forney approach \cite{forney_mlse}).
-A complexity study of all the proposed detectors is also carried in this chapter.
-
-In \textbf{Chapter \ref{chapter:4}}, we consider the presence of multipath in the SISO scenario.
-We first make a classification of the existing aeronautical telemetry channels of the literature.
-We then present the CMA solution and its different variants.
-We later propose a new joint channel estimation and detection algorithm for SOQPSK-TG.
-This solution is an adapted version of the PSP principle and relies on the proposed PAM decomposition given in Chapter \ref{chapter:2}.
-
-In \textbf{Chapter \ref{chapter:5}}, we focus on the IRIG standardized transmit diversity scheme that allows resolving the two-antenna problem.
-After introducing the IRIG standardized encoder structure and the state of the art decoders, we develop two decoding structures and we present their different variants.
-We then consider the presence of multipath in the STC scenario in the second part of this chapter, and we detail the derivation of the new multipath estimator.
-
-The results given in \textbf{Chapter} \ref{chapter:5} show that the performance of proposed decoders are angle dependent due to employing the standardized scheme.
-For this reason, we propose in \textbf{Chapter \ref{chapter:6}} a new encoding scheme that almost eliminates this angle dependency.
-Moreover, the scheme ameliorates the transmission rate and allows further complexity reduction of the decoding algorithm.
-In the second part of this chapter, we highlight that space-time coding is not the only way to overcome the two-antenna problem.
-We show that sending a SOQPSK-TG signal over one antenna and its artificially delayed copy over the other can also greatly reduce their mutual interference.
-After determining the necessary delay that should be applied, we show that the proposed interference mitigation technique given in Chapter \ref{chapter:4} can also be used in this scenario.
-Thus, the same receiver architecture can be employed whether one or two antennas are mounted on the aircraft using this approach.
-
-%Despite the extensive literature regarding channel estimation and equalization techniques for linear modulations, it is not the case for SOQPSK-TG due to its nonlinear nature.
+Blabla

### 2019-01-17 2-Chapters/Conclusion/conclusion.tex
Okay, all the template is better for my need now. I have to write…

@@
 %\minitoc
-\label{chapter:CCL}
+\label{chapter:conclusion}
 % Write miniTOC just after the title
@@
 
-In this PhD thesis, we focused on studying reception algorithms for SOQPSK-TG under different scenarios.
-This study first requires a deep understanding of the signal features, which are different from the conventional linear modulations.
-In fact, SOQPSK-TG is an aeronautical telemetry modulation that has been considered for a long time as a particular CPM due to its ternary alphabet generated using a specific precoder and its long frequency pulse.
-Moreover, the fact that this modulation is not a linear function of its transmitted symbols makes handling this modulation a hard task.
-These features have pushed the use of particular decompositions of the signal such as ternary PAM decomposition (PRD) and the XTCQM representation to make the signal more versatile.
-Each representation has its pros and cons.
+In this PhD thesis, blabla.
 
-On the one hand, the PRD allows approximating the signal as a sum of two linear modulations.
-However, the resulting pseudo-symbols still inherit the memory effect of the modulation.
-On the other hand, the XTCQM representation provides an accurate approximation of the signal using a bank of XTCQM waveforms with a small memory.
-Nevertheless, these waveforms are data dependent and thus we cannot separate contribution of the data independently from the pulse shape.
-The contrasting features of these representations, which are presented in details in \textbf{Chapter \ref{chapter:1}}, imply a limited use to specific scenarios.
-
-In this work, we first focused on the SOQPSK precoder structure before tackling the signal representation aspect.
-We showed that this precoder could be decomposed into a recursive encoder that eliminates the long memory of the modulation and a duobinary encoder that converts the binary symbols to ternary ones to increase the spectral efficiency.
-Thanks to this new vision of the precoder, we showed that SOQPSK-TG could be seen as a duobinary CPM whose PAM decomposition, namely DBD, retains the key features of the Laurent decomposition of binary CPM.
-As a consequence, SOQPSK-TG can accurately be approximated as a single PAM or as a sum of two PAMs.
-Moreover, the presence of the recursive encoder has been exploited to reduce the memory effect in the expressions of the pseudo-symbols and to get a direct mapping between the pseudo-symbols and the binary data.
-Therefore, we proposed a new representation of SOQPSK that gathers better features than the state of the art representations as fully described in \textbf{Chapter \ref{chapter:2}}.
-
-As a direct consequence of this new PAM decomposition, we proposed in \textbf{Chapter \ref{chapter:3}} a wide range of reduced complexity detectors via several approaches (Kaleh approach, Ungerboeck approach, and Forney approach).
-The very near-optimal performance of these detectors confirms the accuracy of the proposed DBD approximations.
-However, they also show that the main pulse of DBD generates significant inter-symbol interference that should be mitigated at the receiver side.
-
-In \textbf{Chapter \ref{chapter:4}}, we took advantage of the PAM decomposition (with DBD) to propose a new joint channel estimation and detection algorithm for SOQPSK-TG.
-This algorithm is inspired by the PSP principle to mitigate the multipath effects, which are present in the aeronautical telemetry environment.
-This solution operates using two different approximations of the signal and significantly outperforms the constant modulus algorithm.
-
-In \textbf{Chapter \ref{chapter:5}}, we focused on the scenario where two antennas are mounted on the aircraft to get an omnidirectional transmission to resolve the two-antenna problem.
-This use case requires creating diversity by using Alamouti encoding.
-However, it makes the decoding process different from what has been proposed for linear modulations because of the used modulation (SOQPSK-TG) and because of the presence of a differential delay.
-After presenting the state of the art solutions, we proposed in this chapter two main decoding structures.
-The first one relies on the first PAM approximation of SOQPSK-TG, i.e., by considering that SOQPSK-TG is a linear modulation (or more precisely, an offset modulation) and provides acceptable BER performance while benefiting from an important complexity reduction compared to the state of the art solution.
-The second decoding structure considers that SOQPSK-TG can be approximated as a sum of two linear modulations and greatly outperforms the state of the art solution.
-Both structures take into account the presence of the differential delay and provide soft outputs.
-In \textbf{Chapter \ref{chapter:5}}, we also considered the presence of multipath in the STC scenario.
-After formulating the problem using multipath, we proposed a new channel estimator and showed that this solution could be used along with the second decoding structure.
-
-In the final chapter of this dissertation, we explored other solutions to resolve the two-antenna problem.
-The first proposed solution takes into account the observations made in Chapter \ref{chapter:5} where it has been shown that the standardized encoding scheme does not eliminate the mutual interference of SOQPSK-TG signals.
-For this reason, we proposed a new encoding scheme that is more suitable for SOQPSK-TG.
-This scheme allows gathering all the attractive features of a STC system since it provides better transmission rate than the state of the art encoding and results in a reduced complexity decoding structure with attractive BER performance (within $1.1$ dB of the SISO bound).
-In the second part of \textbf{Chapter \ref{chapter:6}}, we considered the time diversity approach, and we proposed sending one signal over one antenna and its time delayed version over the other.
-This delay is created artificially to reduce the mutual interference of the signals.
-The advantage of this approach is that the receiver considers that the transmitted signal is corrupted by a multipath channel whose interference can successfully be mitigated using the techniques introduced in Chapter \ref{chapter:4}.
+In \textbf{Chapter \ref{chapter:4}}, blabla.
 
@@
 We present in the following the possible areas of future studies
-\subsubsection*{Equalization in the SISO scenario:} 
+
+\subsubsection*{Blabla}
 \begin{itemize}
-\item In this dissertation, the PSP algorithm operates using the duobinary PAM decomposition (DBD) of SOQPSK-TG.
-However, thanks to the mapping found between the pseudo-symbols and the data bits using PRD (Chapter \ref{chapter:2}), it may be possible to apply the PSP principle to the PRD PAM approximation.
-Since both PRD and DBD have different characteristics, the resulting algorithms may have different behaviors.
-\item  To the best of our knowledge, the INET packet structure does not contain any cyclic prefix.
-However, a new packet structure can be imagined for SOQPSK-TG so that the frequency equalization algorithms introduced in \cite{freq_eqz_cpm1,freq_eqz_cpm2,thillo_eqz,chayot_eqz} can be used, tested in the presence of aeronautical telemetry channels, and compared with the proposed algorithms of this work.
+\item Blabla
 \end{itemize}
-
-%\subsubsection*{Channel coding in the SISO scenario:}   
-\subsubsection*{Decoding and Multipath mitigation in the STC scenario:}
-\begin{itemize}
-\item In this work, we proposed two decoding structures for the STC scenario.
-The first one relies on Ungerboeck/Cavers approach, and the second one uses the least squares approach (which is equivalent to Forney approach in the SISO case).
-These decoders take advantage of the "linearized" version of SOQPSK-TG but they do not take into account its CPM nature.
-Thus, it may be relevant to extend Kaleh approach (which exploits the constant envelope nature of CPM when calculating the log-likelihood function) from the SISO case to the MISO one to get a third decoding structure.
-\item The multipath mitigation technique developed in this dissertation for the STC scenario only works using the second proposed decoding structure.
-Further works regarding frequency domain or time domain equalizers need to be done for the STC scenario.
-Ideally, the equalizer should work regardless of the used decoding structure.
-\end{itemize}
-
-\subsubsection*{STC estimators:}
-\begin{itemize}
-\item  The complexity of the developed estimators in \cite{STC_part1} is relatively high.
-More work needs to be done to reduce the complexity, and it may be accomplished using the new PAM decomposition.
-\item The simulation results given in Section \ref{chap5_sec:multipath} show that the estimators are highly multipath-sensitive.
-More work needs to be done to improve their performance for this scenario.
-\end{itemize}
-\subsubsection*{New aeronautical telemetry systems:}
-Conceiving power and spectrally efficient systems is a constant challenge for aeronautical telemetry.
-However, achieving this goal requires evolving the IRIG standard as well as the actual infrastructure.
-We can imagine for instance
-\begin{itemize}
-\item Pairing MIMO systems with SOQPSK to increase the spectral efficiency while taking advantage of the constant envelope nature of the modulation.
-The MISO case has already been standardized in IRIG to resolve the two-antenna problem without increasing the spectral efficiency.
-Therefore, we can imagine an extension of this concept to reach the latter goal.
-A feasibility study should also be carried to determine how many antennas can be mounted on board the aircraft and how many receiving antennas can be used.
-\item  Defining different modes for different CPM modulations as done in the tactical communications standards \cite{nato2}.
-Each modulation fulfills a certain spectral efficiency and power trade-off.
-Switching between one mode to the other can be done manually or autonomously (the latter case may require an ulplink mode).
-\item Adopting non-constant envelope signals, which beget better spectral efficiency than CPM such as COFDM \cite{channel_cband_zds} or APSK \cite{apsk}. 
-\end{itemize}

### 2019-01-17 3-Appendices/1-Appendix/appendix1.tex
Okay, all the template is better for my need now. I have to write…

--- 2019-01-17 4-Listofpublications/Listofpublications.tex
@@
 % ******************************* Thesis Appendix A ****************************
-\chapter*{List of Publications}
-\addcontentsline{toc}{chapter}{List of Publications}
-\section*{International Journals}
-%{enumi}
-%\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
-%\begin{enumerate}{[IJ-\arabic{enumi}]}%[label={[RI-\arabic*]}]
-\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
-\begin{etaremune}%[label={[RI-\arabic*]}]
-	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y.
-`` Reduced Complexity Detectors Based on SOQPSK Linearzations,'', in \emph{IEEE Transactions on Aerospace and Electronic Systems}, 2018 (Major Revision)
-	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y., `` PAM Decomposition of Ternary CPM With Duobinary Encoding,'', in \emph{IEEE Transactions on Communications}, vol.
-65, no.
-10, pp.
-4274-4284, doi: 10.1109/TCOMM.2017.2723567, Oct 2017
-\end{etaremune}
+\chapter{Appendix for Chapter \ref{chapter:1}}
+\label{app:chapt1}
 
-\section*{International Conferences}
-%\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
-\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
-%\begin{enumerate}{[IC-\arabic{enumi}]}
-\begin{etaremune}
-	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Linear Decision Feedback Detector for SOQPSK,'', in \emph{URSI AP-RASC}, 2019 (Submitted)
-	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``A Reduced Complexity OQPSK-Type Detector For SOQPSK,'', in \emph{IEEE 88th Vehicular Technology Conference (VTC-Fall)},Chicago,Sep 2018 (to be published)
-	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``Joint Channel Estimation and Detection of SOQPSK Using the PAM Decomposition,'', in \emph{IEEE 25th International Conference on Telecommunications (ICT)}, St.
-Malo, pp.
-154-158.
-doi: 10.1109/ICT.2018.8464828, Jun 2018
-\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Analysis of Duobinary Encoding for CPM Signals,'', in \emph{URSI General Assembly and Symposium}, Montreal, Aug 2017
-\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y.
-`` Phase offset estimation of SOQPSK waveform by the analysis of the angle distribution,'', in \emph{IEEE 13th International Symposium on Wireless Communication Systems}, Poznan, pp.
-16-21.
-doi: 10.1109/ISWCS.2016.7600847, Sep 2016
-\end{etaremune}
-
-\section*{National Conferences}
-\renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
-\begin{etaremune}
-\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``D\'{e}composition PAM d'un signal CPM avec encodage duo-binaire - Application \`{a} la modulation SOQPSK,'', in \emph{GRETSI}, Juan-Les-Pins, France, Sep 2017
-\end{etaremune}
-
-\section*{Other National Communications}
-\renewcommand{\labelenumi}{[ONC-\arabic{enumi}]}
-\begin{etaremune}
-\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``A New PAM Decomposition for SOQPSK in Aernautical Telemetry Context,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017
-\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{SCEE Seminar}, Rennes, France, Jun 2017
-\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016
-\end{etaremune}
-
-\section*{Patents}
-\renewcommand{\labelenumi}{[PT-\arabic{enumi}]}
-\begin{etaremune}
-\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. `` Nouvelle architecture de d\'{e}codage spatio-temporel bas\'{e}e sur la d\'{e}composition PAM pour la modulation SOQPSK-TG,'' (pending)
-\end{etaremune}
+In this appendix, blabla.

### 2019-01-17 3-Appendices/2-Appendix/appendix2.tex
Okay, all the template is better for my need now. I have to write…

%!TEX root = ../thesis.tex
% ******************************* Thesis Appendix B ****************************
\chapter{Appendix for Chapter \ref{chapter:2}}
\label{app:chapt2}

In this appendix, blabla.

### 2019-01-19 1-Preamble/4-Acknowledgement/acknowledgement.tex
Cosmetic change + first version of the acknowledgements (yes, it's useless to write this first… I know)

@@
 
-And I would like to acknowledge ...
+\TODO{To finish}
+% See https://perso.crans.org/besson/phd/thanks/
 
-See https://perso.crans.org/besson/phd/thanks/
+And I would like to acknowledge and thank the following people.
 
-\TODO{To start!}
+My PhD advisors, \href{https://moychristophe.wordpress.com/}{Christophe Moy} and \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}.
+
+My family: my parents, my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother} and my younger brother, Marine, my grand-parents, Jacques, Bernadette, Lara.
+
+My closest friends, in a chronological list: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia, Julia, Pierre, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément, Jill-Jênn, Alain, Claire, Édouard, Loïc, Damien, Kumudham, Priscilla, Thibault, Sébastien, Sélim, Valentin, Marine, Johanne, Lola.
+
+My wonderful colleagues during these three years: Rémi, Pascal, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin, Vincent, Rami, Cristo, Adrien, Esteban, Ali, Éloïse, Bastien, Pierre, Georgios, in Rennes; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas in Lille.
+
+Thanks to some people for their support: Karine, Jeannine, Greg, Anne at CentraleSupélec Rennes, François, David and David at ENS de Rennes, Romaric at ENSAI, Amélie at Inria Lille, Frédéric, Alain, Sylvain, Paul, Serge, Claudine, Nicolas, Carine and Delphine at ENS Cachan (now Paris-Saclay).
+
+I would also like to thank some things. First, the Earth, then my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents.
+My \href{https://perso.crans.org/besson/cuisine/}{culinary friends}: cheese, chocolate, coffee, vegetables etc!
+My \href{https://perso.crans.org/besson/}{technological friends}: Bash, Python, XUbuntu, Firefox, Konsole, Visual Studio Code etc.
+
+Some friends thanked me in their own PhD thesis, and it is a pleasure to do the same!
+%
+\begin{itemize}
+\tightlist
+\item
+  \href{http://jill-jenn.net/}{Jill-Jênn Vie}
+  (\href{http://jiji.cat/cat.pdf}{thesis})
+\item
+  \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi}
+  (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis})
+\item
+  \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier}
+  (\href{https://www.theses.fr/2017REN1S091}{thesis})
+\item
+  \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}
+\item
+  \href{https://sites.google.com/view/guerand}{Jessica Guérand}
+  (\href{https://drive.google.com/open?id=17CBibM8ZlzIbyVxpxktK3j5rNePALUs8}{thesis})
+\item
+  \href{https://sites.google.com/view/ducasse/}{Romain Ducasse}
+\item
+  \href{https://members.loria.fr/SAbelard/}{Simon Abelard}
+  (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis})
+\item
+  \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli}
+\item
+  \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire
+  Brécheteau}
+  (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis})
+\item
+  \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien
+  Allonsius}
+  (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis})
+\item
+  \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}
+\item
+  \href{https://remibonnefoi.wordpress.com/}{Rami Othman}
+\end{itemize}
+
+% Je souhaite terminer ces remerciements par une pensée toute particulière pour Lola, qui m'a soutenu, qui m'a émerveillé par son enthousiasme et qui m'a réconforté par son amour durant la dernière année.
+Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó, me sorprendió con su entusiasmo y me consoló con su amor durante el año pasado.
 

### 2019-01-19 4-Listofpublications/Listofpublications.tex
Added list of publications

@@
 \addcontentsline{toc}{chapter}{List of Publications}
-\section*{International Journals}
-%{enumi}
-%\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
-%\begin{enumerate}{[IJ-\arabic{enumi}]}%[label={[RI-\arabic*]}]
-\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
-\begin{etaremune}%[label={[RI-\arabic*]}]
-	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y.
-`` Reduced Complexity Detectors Based on SOQPSK Linearzations,'', in \emph{IEEE Transactions on Aerospace and Electronic Systems}, 2018 (Major Revision)
-	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y., `` PAM Decomposition of Ternary CPM With Duobinary Encoding,'', in \emph{IEEE Transactions on Communications}, vol.
+
+% \section*{International Journals}
+% \renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
+% \begin{etaremune}%[label={[RI-\arabic*]}]
+% 	\item \textbf{L.
+Besson}, Skrzypczak, A., E.
+Kaufmann `` Reduced Complexity Detectors Based on SOQPSK Linearzations,'', in \emph{IEEE Transactions on Aerospace and Electronic Systems}, 2018 (Major Revision)
+% 	\item \textbf{L.
+Besson}, Skrzypczak, A., E.
+Kaufmann, `` PAM Decomposition of Ternary CPM With Duobinary Encoding,'', in \emph{IEEE Transactions on Communications}, vol.
 65, no.
@@
 4274-4284, doi: 10.1109/TCOMM.2017.2723567, Oct 2017
+% \end{etaremune}
+
+\section*{International Conferences}
+\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
+\begin{etaremune}
+	% \item \textbf{L. Besson}, E. Kaufmann, Skrzypczak, A. ``Linear Decision Feedback Detector for SOQPSK,'', in \emph{URSI AP-RASC}, 2019 (Submitted)
+\item
+  \href{https://perso.crans.org/besson/articles/BBM__IEEE_WCNC_2019.pdf}{GNU
+  Radio Implementation of MALIN: « Multi-Arm bandits Learning for
+  Internet-of-things Networks »}, par \textbf{L. Besson},
+  \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
+  \href{https://moychris.wordpress.com/}{C. Moy}.
+  \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
+  Marrakech, Maroc, avril 2019.
+\item
+  \href{https://perso.crans.org/besson/articles/BBM__Demo_ICT_2018.pdf}{MALIN:
+  « Multi-Arm bandit Learning for Iot Networks » with GRC: A TestBed
+  Implementation and Demonstration that Learning Helps}, par \textbf{L.
+  Besson}, \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
+  \href{https://moychris.wordpress.com/}{C. Moy}. Démonstration
+  présentée à ICT (\emph{International Conference on Communication}),
+  Saint-Malo, France, juin 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}
+\item
+  \href{https://hal.inria.fr/hal-01629733/document}{Multi-Player Bandits
+  Revisited}, par \textbf{L. Besson} \&
+  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
+  Kaufmann}.
+  \href{http://www.cs.cornell.edu/conferences/alt2018/accepted.html}{ALT}
+  (\emph{Algorithmic Learning Theory}), Lanzarote, Espagne, avril 2018,
+  \href{https://hal.inria.fr/hal-01629733}{HAL-01629733}.
+\item
+  \href{https://hal.inria.fr/hal-01705292/document}{Aggregation of
+  Multi-Armed Bandits learning algorithms for Opportunistic Spectrum
+  Access}, par \textbf{L. Besson},
+  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
+  Kaufmann} \& \href{https://moychris.wordpress.com/}{C. Moy}.
+  \href{http://wcnc2018.ieee-wcnc.org/}{IEEE WCNC}, Barcelone, Espagne,
+  avril 2018, \href{https://hal.inria.fr/hal-01705292}{HAL-01705292}.
+\item
+  \href{https://hal.inria.fr/hal-01575419/document}{Multi-Armed Bandit
+  Learning in IoT Networks and non-stationary settings}, par
+  \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L.
+  Besson}, \href{https://moychris.wordpress.com/}{C. Moy},
+  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
+  Kaufmann} \& J. Palicot. \href{http://crowncom.org/2017/}{CrownCom}
+  (\emph{Conference on Cognitive Radio Oriented Wireless Networks}),
+  Lisbonne, Portugal, septembre 2017,
+  \href{https://hal.inria.fr/hal-01575419}{HAL-01575419}, \textbf{prix
+  du meilleur article}.
 \end{etaremune}
 
-\section*{International Conferences}
-%\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
-\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
-%\begin{enumerate}{[IC-\arabic{enumi}]}
-\begin{etaremune}
-	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Linear Decision Feedback Detector for SOQPSK,'', in \emph{URSI AP-RASC}, 2019 (Submitted)
-	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``A Reduced Complexity OQPSK-Type Detector For SOQPSK,'', in \emph{IEEE 88th Vehicular Technology Conference (VTC-Fall)},Chicago,Sep 2018 (to be published)
-	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``Joint Channel Estimation and Detection of SOQPSK Using the PAM Decomposition,'', in \emph{IEEE 25th International Conference on Telecommunications (ICT)}, St.
-Malo, pp.
-154-158.
-doi: 10.1109/ICT.2018.8464828, Jun 2018
-\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Analysis of Duobinary Encoding for CPM Signals,'', in \emph{URSI General Assembly and Symposium}, Montreal, Aug 2017
-\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y.
-`` Phase offset estimation of SOQPSK waveform by the analysis of the angle distribution,'', in \emph{IEEE 13th International Symposium on Wireless Communication Systems}, Poznan, pp.
-16-21.
-doi: 10.1109/ISWCS.2016.7600847, Sep 2016
-\end{etaremune}
-
-\section*{National Conferences}
-\renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
-\begin{etaremune}
-\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``D\'{e}composition PAM d'un signal CPM avec encodage duo-binaire - Application \`{a} la modulation SOQPSK,'', in \emph{GRETSI}, Juan-Les-Pins, France, Sep 2017
-\end{etaremune}
+% \section*{National Conferences}
+% \renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
+% \begin{etaremune}
+% 	\item \textbf{L. Besson}, E. Kaufmann, ``XXX'', in \emph{GRETSI}, Lille, France, Aug 2019
+% \end{etaremune}
 
@@
 \begin{etaremune}
-\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``A New PAM Decomposition for SOQPSK in Aernautical Telemetry Context,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017
-\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{SCEE Seminar}, Rennes, France, Jun 2017
-\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A.
-``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016
+	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017
+	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
+	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
+	\item	R. Bonnefoi, \textbf{L. Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016
 \end{etaremune}
 
-\section*{Patents}
-\renewcommand{\labelenumi}{[PT-\arabic{enumi}]}
+\section*{Other Publications}
+\renewcommand{\labelenumi}{[OP-\arabic{enumi}]}
 \begin{etaremune}
-\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. `` Nouvelle architecture de d\'{e}codage spatio-temporel bas\'{e}e sur la d\'{e}composition PAM pour la modulation SOQPSK-TG,'' (pending)
+\item
+	\href{https://perso.crans.org/besson/articles/BMBBM__IEEE_WCNC__2019.pdf}{Upper-Confidence
+	Bound for Channel Selection in LPWA Networks with Retransmissions},
+	par \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, J.
+	Manco-Vasquez, \textbf{L. Besson},
+	\href{https://cfaouzi-bader.com/}{F. Bader},
+	\href{https://moychris.wordpress.com/}{C. Moy}. Septembre 2018.
+\item
+	\href{https://hal.inria.fr/hal-01840022/document}{SMPyBandits: an
+	Open-Source Research Framework for Single and Multi-Players Multi-Arms
+	Bandits (MAB) Algorithms in Python}, par \textbf{L. Besson},
+	développement depuis octobre 2016 et encore actif,
+	\href{https://hal.inria.fr/hal-01840022}{HAL-01840022}. Code sur
+	\href{https://GitHub.com/SMPyBandits/SMPyBandits}{GitHub.com/SMPyBandits}
+	(environ 40000 lignes), documentation sur
+	\href{https://SMPyBandits.rtfd.io}{SMPyBandits.rtfd.io}.
+\item
+	\href{https://hal.inria.fr/hal-01736357/document}{What Doubling-Trick
+	Can and Can't Do for Multi-Armed Bandits}, par \textbf{L. Besson} \&
+	\href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
+	Kaufmann}, septembre 2018,
+	\href{https://hal.inria.fr/hal-01736357}{HAL-01736357}.
+\item
+	\href{https://hal.inria.fr/hal-01847480/document}{A Note on the Ei
+	Function and a Useful Sum-Inequality}, par \textbf{L. Besson}, février
+	2018 \href{https://hal.inria.fr/hal-01847480}{HAL-01847480}.
 \end{etaremune}
+
+% \section*{Patents}
+% \renewcommand{\labelenumi}{[PT-\arabic{enumi}]}
+% \begin{etaremune}
+% \item \textbf{L. Besson}, E. Kaufmann, Skrzypczak, A. `` Nouvelle architecture de d\'{e}codage spatio-temporel bas\'{e}e sur la d\'{e}composition PAM pour la modulation SOQPSK-TG,'' (pending)
+% \end{etaremune}

### 2019-02-07 1-Preamble/4-Acknowledgement/acknowledgement.tex
Added list of publications
Starting to fill the sections of each chapters, to have an overview of the organization

--- 2019-01-17 1-Preamble/4-Acknowledgement/acknowledgement.tex
@@
 
-And I would like to acknowledge ...
+\TODO{To finish}
+% See https://perso.crans.org/besson/phd/thanks/
 
-See https://perso.crans.org/besson/phd/thanks/
+And I would like to acknowledge and thank the following people.
 
-\TODO{To start!}
+My PhD advisors, \href{https://moychristophe.wordpress.com/}{Christophe Moy} and \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}.
+
+My family: my parents, Christophe and Patricia, my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian} and my younger brother Fabian, Marine, my grand-parents, Jeannine, Jean-Marc and Jacqueline, my uncle and aunt Jacques, Bernadette, and my cousin Lara.
+
+My closest friends, sorted in an absurd chronological list: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia, Julia, Pierre, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément, Jill-Jênn, Alain, Claire, Édouard, Loïc, Damien, Kumudham, Priscilla, Thibault, Sébastien, Sélim, Valentin, Marine, Johanne, Lola.
+
+My wonderful colleagues during these three years: Rémi, Pascal, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin, Vincent, Rami, Muhammad, Cristo, Adrien, Esteban, Ali, Éloïse, Bastien, Corentin, Georgios, in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas in Lille.
+
+Thanks to some people for their support: Karine, Jeannine, Grégory, Anne and others at CentraleSupélec Rennes; François, David, David at ENS de Rennes, Romaric at ENSAI, Amélie at Inria Lille, Amélie at Lycée Chateaubriand.
+During my studies at ENS Cachan (now Paris-Saclay), I am thankful to all the great professors who taught me, including Frédéric, Alain and Claudine in the maths department, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
+
+I would also like to thank some things. First, the Earth, then my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents.
+My \href{https://perso.crans.org/besson/cuisine/}{culinary friends}: cheese, chocolate, coffee, vegetables etc!
+My \href{https://perso.crans.org/besson/}{technological friends}: Bash, Python, XUbuntu, Firefox, Konsole, Visual Studio Code etc.
+
+Some friends thanked me in their own PhD thesis, and it is a pleasure to do the same.
+In a chronological order, I thank
+%,
+\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{http://jiji.cat/cat.pdf}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://drive.google.com/open?id=17CBibM8ZlzIbyVxpxktK3j5rNePALUs8}{thesis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.
+
+% Je souhaite terminer ces remerciements par une pensée toute particulière pour Lola, qui m'a soutenu, qui m'a émerveillé par son enthousiasme et qui m'a réconforté par son amour durant la dernière année.
+Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó, me sorprendió con su entusiasmo y me consoló con su amor durante el año pasado.
 

### 2019-02-07 2-Chapters/3-Chapter/chapter3.tex
Okay, all the template is better for my need now. I have to write…
Starting to fill the sections of each chapters, to have an overview of the organization

--- 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
@@
 % First chapter begins here
-%\adjustmtc
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction}
-\chaptermark{Introduction}
+\chapter{Presentation of SMPyBandits, a Complete MAB Simulation Library}
+\label{chapter:3}
 \minitoc
-\label{chapter:Intro}
-\graphicspath{{2-Chapters/1-Chapter/Images/}}
+\newpage
+% Write miniTOC just after the title
+\graphicspath{{2-Chapters/3-Chapter/Images/}}
 
-% Write miniTOC just after the title
-%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
-%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
+This chapter is intended as a long version of the paper presenting SMPyBandits that I wrote in Summer 2018, see https://hal.inria.fr/hal-01840022
 
 % ----------------------------------------------------------------------------
-\section*{General context}
+\section{Presentation of the library}
+\label{sec:3:presentationLibrary}
 
-\indent
-The work presented in this PhD thesis
-Blabla
+I want to show what is SMPyBandits, what problems does it solve, what did I implement, how to use it.
 
-\section*{Contributions}
+% ----------------------------------------------------------------------------
+\section{Review of state-of-the-art single-player algorithms}
+\label{sec:3:reviewSPAlgorithms}
 
-Blabla
+I want to show that SMPyBandits contains highly documented implementations of all the most important algorithms designed for stationary single-player multi-armed bandit.
 
-\begin{itemize}
-\item Blabla
-\item Blabla
-\item Blabla
-\end{itemize}
+I will include here formulas about the main algorithms (klUCB, MOSS, etc).
+I will pick some problems (Bernoulli, Gaussian, others), with fixed means or ``Bayesian'' means (\ie, changing at each time) and I will run extensive numerical simulations to compare the main algorithms.
 
-\section*{Dissertation outline}
-This dissertation is organized as follows.
+The take away message is: klUCB beats UCB, Thompson sampling is great too, all the others algorithms are less efficient or comparable, so we only use klUCB for all the rest of this thesis.
 
-Blabla
+I will include graphs showing the mean regret $R_t$ as a function of time $t$, as well as histograms or boxplots showing the distribution of $R_T$.
+
+% ----------------------------------------------------------------------------
+\section{Some numerical experiments to compare time and memory costs of various algorithms}
+\label{sec:3:timeAndMemoryCosts}
+
+I want to show the real numerical costs in terms of computation time and storage requirement of the various algorithms for single-player bandits.
+
+Maybe I can pick one simple problem family (\eg, evenly spaced in $[\varepsilon,1-\varepsilon]$), and run experiments for $T=1000,5000,10000$ etc, and $K=2,3,5,9,20,30,50,100,1000$ and show the evolution of time/memory as a function of $T$ and $K$.
+
+The goal is to highlight that the simplest (but most efficient) algorithms have a memory cost bounded by $\bigO{K}$ and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
+
+% ----------------------------------------------------------------------------
+\section{An example of another model: rested or restless Markov models}
+\label{sec:3:markovModels}
+
+Maybe if I have space at the end of this chapter, I can show another part of the library, maybe introduce the rested/restless Markov models (with maybe just one example) and show simulations in this model?
+
+SMPyBandits also implemented sparse MAB models (with $s<K$ arms with a mean $\mu_i>0$ and $K-s$ arms with a mean $\mu_i \leq 0$), I could also present this model, the various algorithms designed to tackle it, and some numerical experiments.
+
+% ----------------------------------------------------------------------------
+\section{Conclusion of Chapter 3}
+\label{sec:3:conclusion}
+
+In this chapter, we saw...
+
+Future works include...

### 2019-02-07 2-Chapters/4-Chapter/chapter4.tex
Okay, all the template is better for my need now. I have to write…
Starting to fill the sections of each chapters, to have an overview of the organization

--- 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
@@
 % First chapter begins here
-%\adjustmtc
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction}
-\chaptermark{Introduction}
+\chapter{Multi-Armed Bandits Models for Internet of Things Networks}
+\label{chapter:4}
 \minitoc
-\label{chapter:Intro}
-\graphicspath{{2-Chapters/1-Chapter/Images/}}
+\newpage
+% Write miniTOC just after the title
+\graphicspath{{2-Chapters/4-Chapter/Images/}}
 
-% Write miniTOC just after the title
-%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
-%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
+This chapter has three sections, each is a raw include from different papers
+
+- ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419
+
+- ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825
+
+- ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \verb{https://perso.crans.org/besson/articles/BMBM__IEEE_WCNC_2019.pdf}
 
 % ----------------------------------------------------------------------------
-\section*{General context}
+\section{A first model of learning devices in an IoT network}
+\label{sec:4:firstModel}
 
-\indent
-The work presented in this PhD thesis
-Blabla
+- ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419
 
-\section*{Contributions}
+% ----------------------------------------------------------------------------
+\section{Test-bed implementation of our model for real-world validation}
+\label{sec:4:gnuradio}
 
-Blabla
+- ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825
 
-\begin{itemize}
-\item Blabla
-\item Blabla
-\item Blabla
-\end{itemize}
+% ----------------------------------------------------------------------------
+\section{Extension of our model to account for retransmissions}
+\label{sec:4:retransmissions}
 
-\section*{Dissertation outline}
-This dissertation is organized as follows.
+- ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \verb{https://perso.crans.org/besson/articles/BMBM__IEEE_WCNC_2019.pdf}
 
-Blabla
+% ----------------------------------------------------------------------------
+\section{Conclusion of Chapter 4}
+\label{sec:4:conclusion}
+
+In this chapter, we saw...
+
+Future works include...

### 2019-02-07 2-Chapters/5-Chapter/chapter5.tex
Okay, all the template is better for my need now. I have to write…
Starting to fill the sections of each chapters, to have an overview of the organization

--- 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
@@
 % First chapter begins here
-%\adjustmtc
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction}
-\chaptermark{Introduction}
+\chapter{Multi-Player Multi-Armed Bandits}
+\label{chapter:5}
 \minitoc
-\label{chapter:Intro}
-\graphicspath{{2-Chapters/1-Chapter/Images/}}
-
+\newpage
 % Write miniTOC just after the title
-%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
-%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
+\graphicspath{{2-Chapters/5-Chapter/Images/}}
 
 % ----------------------------------------------------------------------------
-\section*{General context}
+\section{Blabla}
+\label{sec:blabla}
 
-\indent
-The work presented in this PhD thesis
-Blabla
+This chapter is basically a raw include from my paper ``Multi-Player Bandits Revisited'', see https://hal.inria.fr/hal-01629733
 
-\section*{Contributions}
+I also need to include additional simulations from the latest research articles, and discussions about the new works that took inspiration in our paper.
 
-Blabla
+% ----------------------------------------------------------------------------
+\section{Conclusion of Chapter 5}
+\label{sec:5:conclusion}
 
-\begin{itemize}
-\item Blabla
-\item Blabla
-\item Blabla
-\end{itemize}
+In this chapter, we saw...
 
-\section*{Dissertation outline}
-This dissertation is organized as follows.
-
-Blabla
+Future works include...

### 2019-02-07 2-Chapters/6-Chapter/chapter6.tex
Okay, all the template is better for my need now. I have to write…
Starting to fill the sections of each chapters, to have an overview of the organization

--- 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
@@
 % First chapter begins here
-%\adjustmtc
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction}
-\chaptermark{Introduction}
+\chapter{Towards Non-Stationary Models}
+\label{chapter:6}
 \minitoc
-\label{chapter:Intro}
-\graphicspath{{2-Chapters/1-Chapter/Images/}}
+\newpage
+% Write miniTOC just after the title
+\graphicspath{{2-Chapters/6-Chapter/Images/}}
 
-% Write miniTOC just after the title
-%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
-%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
+This chapter is basically a raw include from my paper ``The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits'', see https://hal.inria.fr/hal-02006471
 
 % ----------------------------------------------------------------------------
-\section*{General context}
+\section{The piece-wise non-stationary bandits model}
+\label{sec:6:piecewiseModel}
 
-\indent
-The work presented in this PhD thesis
-Blabla
+% ----------------------------------------------------------------------------
+\section{The Bernoulli Generalized Likelihood Ratio Test}
+\label{sec:6:piecewiseModel}
 
-\section*{Contributions}
+% ----------------------------------------------------------------------------
+\section{The GLRklUCB Algorithm}
+\label{sec:6:piecewiseModel}
 
-Blabla
+% ----------------------------------------------------------------------------
+\section{Regret analysis for the GLRklUCB alRegret analysis for the GLRklUCB algorithm}
+\label{sec:6:piecewiseModel}
 
-\begin{itemize}
-\item Blabla
-\item Blabla
-\item Blabla
-\end{itemize}
+% ----------------------------------------------------------------------------
+\section{Numerical simulations for piece-wise stationary bandit problems}
+\label{sec:6:piecewiseModel}
 
-\section*{Dissertation outline}
-This dissertation is organized as follows.
+% ----------------------------------------------------------------------------
+\section{Conclusion of Chapter 6}
+\label{sec:6:conclusion}
 
-Blabla
+In this chapter, we saw...
+
+Future works include...
+
+% ----------------------------------------------------------------------------
+\appendix
+
+\section{FIXME appendix in a chapter: where is it?}

### 2019-02-07 2-Chapters/1-Chapter/chapter1.tex
Starting to fill the sections of each chapters, to have an overview of the organization
Fix #4

@@
 % First chapter begins here
-\chapter{Continuous Phase Modulation for Aeronautical Telemetry }
+\chapter{Introduction}
 \label{chapter:1}
@@
 \graphicspath{{2-Chapters/1-Chapter/Images/}}
-\section{CPM signal model}
-\label{sec:CPM_signal_model}
+
+% ----------------------------------------------------------------------------
+\section{Considered problems}
+\label{sec:1:problems}
+
+% ----------------------------------------------------------------------------
+\section{Considered solutions}
+\label{sec:1:solutions}
+
+% ----------------------------------------------------------------------------
+\section{Overview of the contributions}
+\label{sec:1:contributions}
+
+% ----------------------------------------------------------------------------
+\section{List of publications}
+\label{sec:1:contributions}
+
+% =============================================================================
+\subsection{Publications in international conferences with proceedings}
+
+\begin{itemize}
+\item
+    \emph{GNU Radio Implementation of MALIN: ``Multi-Armed bandits Learning for Internet-of-things Networks''},
+    by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
+    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
+    \texttt{hal.inria.fr/hal-02006825}.
+
+\item
+    \emph{Multi-Player Bandits Revisited}},
+    by \textbf{Lilian Besson} \& Emilie Kaufmann.
+    \emph{Algorithmic Learning Theory} (ALT, \texttt{www.cs.cornell.edu/conferences/alt2018}), Lanzarote, Spain, April 2018,
+    \texttt{hal.inria.fr/hal-01629733}.
+
+\item
+    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access}},\\
+    by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
+    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2018.ieee-wcnc.org}), Barcelone, Spain, April 2018,
+    \texttt{hal.inria.fr/hal-01705292}.
+
+\item
+    \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings}},
+    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \texttt{crowncom.org/2017}), Lisboa, Portugal, Septembre $2017$,
+    \texttt{hal.inria.fr/hal-01575419},
+    \textbf{Best Paper Award}.
+
+\end{itemize}
+
+% =============================================================================
+\subection{Submitted works or in progress works waiting for a new submission}
+
+\begin{itemize}
+\item
+    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits}},
+    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
+    \texttt{hal.inria.fr/hal-02006471}.
+
+\item
+    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
+    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
+    Submitted for MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}), February $2019$.
+
+\item
+    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
+    by \textbf{Lilian Besson}, active development since October $2016$,
+    \texttt{hal.inria.fr/hal-01840022}.
+    About $40000$ lines of code, hosted on \texttt{GitHub.com/SMPyBandits},
+    with a complete documentation on \texttt{SMPyBandits.rtfd.io}.
+
+\item
+    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
+    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
+    \texttt{hal.inria.fr/hal-01736357}.
+
+\item
+    \emph{MALIN: ``Multi-Arm bandit Learning for Iot Networks'' with GRC: A TestBed Implementation and Demonstration that Learning Helps},
+    by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
+    Demonstration presented at \emph{International Conference on Communication} (ICT, \texttt{ict-2018.org/demos}), in Saint-Malo, France in June $2018$.
+    See \texttt{YouTu.be/HospLNQhcMk} for a $6$-minutes presentation video.
+
+\item
+    \emph{A Note on the Ei Function and a Useful Sum-Inequality},
+    by \textbf{Lilian Besson},
+    February $2018$,
+    \texttt{hal.inria.fr/hal-01847480}.
+
+\end{itemize}
+
+% =============================================================================
+\subsection{Presentations in seminars and conferences}
+
+\paragraph{Seminars} I gave some presentations in the following events:
+    SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
+    SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
+    as well as
+    for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
+    for the brown-bag seminar at ENSAI in Bruz in January $2018$,
+    for the weekly CMAP seminar at École Polytechnique in October $2018$.
+
+\paragraph{Tutorial}
+    I gave a tutorial\footnote{Cf. \texttt{hal.archives-ouvertes.fr/cel-01830248}} on the Julia language, at IETR seminar in Vannes in June $2018$,
+    with Pierre Haessig (\texttt{pierreh.eu}).
+
+\paragraph{Training}
+    I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
+    We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
+    and I gave about $10$ training, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.
+
+% =============================================================================
+\subsection{Other experiences}
+
+\paragraph{Conferences} I attended the following conferences:
+	\emph{International Conference on Communication} ICC (Paris), May $2017$,
+    \emph{Conference on Learning Theory} COLT (Amsterdam), July $2017$,
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks} CROWNCOM (Lisboa), September $2017$,
+    \emph{Conference on Algorithmic Learning Theory} ALT (Lanzarote), April $2018$,
+    \emph{Wireless Communication and Networking Conference} WCNC (Barcelona), April $2018$,
+	\emph{International Conference on Telecommunication} ICT (Saint-Malo), June $2018$,
+    \emph{Wireless Communication and Networking Conference} WCNC (Marrakech), April $2019$,
+    \emph{Colloque francophone de traitement du signal et des images} GRETSI (Lille), August $2019$.
+
+\paragraph{Seminars} I attended the following seminars:
+	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
+	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
+	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$,
+
+\paragraph{Responsabilities}
+    president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
+    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{Cf.
+\texttt{bitbucket.org/scee_ietr/phd-student-day-ietr-2017-bonnefoi-and-besson/downloads/poster.pdf}}.
+    I also helped for the organization and presented another poster\footnote{Cf.
+\texttt{bitbucket.org/lbesson/phd-student-day-ietr-2018/downloads/poster.pdf}} at IETR seminar (\texttt{seminar-ietr-18.sciencesconf.org}), held in Vannes in June $2018$.
+
+    \paragraph{Reviews}
+	for \emph{European Workshop on Reinforcement Learning} (Lille) \texttt{ewrl.wordpress.com/ewrl14-2018}, in October $2018$, I reviewed $5$ papers.
+	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \uref{https://nips.cc/Conferences/2018/}{\emph{NeurIPS}} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.
+
+\paragraph{System Administrator}
+	maintaining our workstations running GNU/Linux and Windows
+	at SCEE team,
+	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
+	from January $2017$ to June $2019$.

### 2019-02-07 4-Listofpublications/Listofpublications.tex
Fix #4

@@
 % \begin{etaremune}%[label={[RI-\arabic*]}]
-% 	\item \textbf{L.
-Besson}, Skrzypczak, A., E.
-Kaufmann `` Reduced Complexity Detectors Based on SOQPSK Linearzations,'', in \emph{IEEE Transactions on Aerospace and Electronic Systems}, 2018 (Major Revision)
-% 	\item \textbf{L.
-Besson}, Skrzypczak, A., E.
-Kaufmann, `` PAM Decomposition of Ternary CPM With Duobinary Encoding,'', in \emph{IEEE Transactions on Communications}, vol.
-65, no.
-10, pp.
-4274-4284, doi: 10.1109/TCOMM.2017.2723567, Oct 2017
+% 	\item \textbf{L. Besson}, E. Kaufmann ``XXX'',, in \emph{XXXX}, 2018 (Major Revision)
 % \end{etaremune}
@@
 \begin{etaremune}
-	% \item \textbf{L. Besson}, E. Kaufmann, Skrzypczak, A. ``Linear Decision Feedback Detector for SOQPSK,'', in \emph{URSI AP-RASC}, 2019 (Submitted)
 \item
   \href{https://perso.crans.org/besson/articles/BBM__IEEE_WCNC_2019.pdf}{GNU
-  Radio Implementation of MALIN: « Multi-Arm bandits Learning for
+  Radio Implementation of MALIN: « Multi-Armed bandits Learning for
   Internet-of-things Networks »}, par \textbf{L. Besson},
@@
 	Bound for Channel Selection in LPWA Networks with Retransmissions},
-	par \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, J.
-	Manco-Vasquez, \textbf{L. Besson},
-	\href{https://cfaouzi-bader.com/}{F. Bader},
+	par \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L. Besson}, J.
+	Manco-Vasquez,
 	\href{https://moychris.wordpress.com/}{C. Moy}. Septembre 2018.
@@
 % \begin{etaremune}
-% \item \textbf{L. Besson}, E. Kaufmann, Skrzypczak, A. `` Nouvelle architecture de d\'{e}codage spatio-temporel bas\'{e}e sur la d\'{e}composition PAM pour la modulation SOQPSK-TG,'' (pending)
+% \item \textbf{L. Besson}, E. Kaufmann. ``XXX'', (pending)
 % \end{etaremune}

### 2019-02-08 3-Appendices/1-Appendix/appendix1.tex
Better outline

@@
 % ******************************* Thesis Appendix A ****************************
-\chapter{Appendix for Chapter \ref{chapter:1}}
-\label{app:chapt1}
+\chapter{Mathematical Tools}
+\label{app:1:mathsTools}
 
 In this appendix, blabla.
+
+% ----------------------------------------------------------------------------
+\section{Concentration Inequalities}
+\label{sec:app1:ConcentrationInequalities}
+
+I probably need to include Hoeffding's inequality, Chernoff's method etc.
+
+Reference: [Boucheron et al], [?]
+
+% ----------------------------------------------------------------------------
+\section{Useful lemmas}
+\label{sec:app1:UsefulLemmas}
+
+I probably need to include a bunch of useful lemmas, used in various places in the thesis, like Pinkser's inequality, etc.
+
+% ----------------------------------------------------------------------------
+\section{Results about klUCB}
+\label{sec:app1:resultsklUCB}
+
+I probably need to include some results about klUCB, mainly coming from the journal [Garivier et al], used in various places in the thesis.

### 2019-02-08 3-Appendices/2-Appendix/appendix2.tex
Better outline

@@
 % ******************************* Thesis Appendix B ****************************
-\chapter{Appendix for Chapter \ref{chapter:2}}
-\label{app:chapt2}
+\chapter{Publication: What Doubling-Trick Can and Can't Do for Multi-Armed Bandits}
+\label{app:2:DoublingTricks}
 
-In this appendix, blabla.
+See https://hal.inria.fr/hal-01736357
+
+I cannot include it raw without more work on the maths, Emilie wants to write only the best versions of our results.
+
+But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!
+
+I need to try DoublingTrick + GLRklUCB, at least experimentally!
+For the theoretical analysis, it will probably be a huge failure.

### 2019-02-08 2-Chapters/1-Chapter/chapter1.tex
Version 0.1

@@
 
+\subsection{Spectrum scarcity}
+
+\subsection{Low-consumption IoT devices for the future}
+
 % ----------------------------------------------------------------------------
@@
 
+\subsection{Online/sequential machine learning to automatically improve wireless communications on the run}
+
+\subsection{Example: Opportunistic Spectrum Access (OSA) in licensed bands}
+
+\subsection{Example: decentralized learning in unlicensed bands with selfish or orthogonalization-based MAB algorithms}
+
 % ----------------------------------------------------------------------------
@@
 \label{sec:1:contributions}
+
+- Literature review of stochastic bandit algorithms (here)
+
+- Aggregation of algorithms for online algorithm selection, state-of-the-art empirical performance with \textbf{Aggregator} (WCNC 2018)
+
+- The most complete open-source simulation library for MAB problems, SMPyBandits (JMLR MLOSS and GitHub)
+
+- Different models for MAB learning for IoT networks (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
+
+- A refined analysis and a new impulse on multi-player bandits (ALT 2018, and 4-8 works inspired by our article since then).
+State-of-the-art with our algorithm MCTopM + klUCB for multi-player bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"
+
+- Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
+
+- Literature review on possible use case of doubling trick, plus a unified and more generic analysis of two families of doubling tricks !
 

### 2019-02-08 2-Chapters/2-Chapter/chapter2.tex
Okay, all the template is better for my need now. I have to write…
Starting to fill the sections of each chapters, to have an overview of the organization
Better outline
Version 0.1

--- 2019-01-17 2-Chapters/0-Introduction/Introduction.tex
@@
 % First chapter begins here
-%\adjustmtc
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction}
-\chaptermark{Introduction}
+\chapter{Single Player and Stationary Multi-Armed Bandit Models}
+\label{chapter:2}
 \minitoc
-\label{chapter:Intro}
-\graphicspath{{2-Chapters/1-Chapter/Images/}}
-
+\newpage
 % Write miniTOC just after the title
-%\startcontents[chapters]\vbox{\bf\Large Sommaire \hrule}
-%\printcontents[chapters]{}{1}{}\vspace*{1pc}\hrule}
+\graphicspath{{2-Chapters/2-Chapter/Images/}}
 
 % ----------------------------------------------------------------------------
-\section*{General context}
+\section{Stationary MAB models}
+\label{sec:2:notations}
 
-\indent
-The work presented in this PhD thesis
-Blabla
+% ----------------------------------------------------------------------------
+\subsection{Finite arms and stochastic rewards}
 
-\section*{Contributions}
+In this thesis, we only consider models with finite arms.
+No linear bandits, no contextual bandits etc.
 
-Blabla
+For now, we only consider stochastic rewards. Piece-wise stochastic models are studied in Chapter~\ref{chapter:6}.
 
-\begin{itemize}
-\item Blabla
-\item Blabla
-\item Blabla
-\end{itemize}
+% ----------------------------------------------------------------------------
+\subsection{Different hypotheses on rewards distributions}
 
-\section*{Dissertation outline}
-This dissertation is organized as follows.
+Everything is implemented and documented at
+https://smpybandits.github.io/docs/Arms.html
 
-Blabla
+Bernoulli, Gaussian, sub-Bernoulli, sub-Gaussian, Exponential, sub-Exponential...
+
+One dimensional exponential family...
+
+Just bounded, in $[a,b]$ but without loss of generality in $[0,1]$ etc.
+
+Markov models: presented later in Section~\ref{sec:3:markovModels}.
+
+% ----------------------------------------------------------------------------
+\section{Applications of stationary MAB}
+\label{sec:2:applicationsofStationaryMAB}
+
+\subsection{Various possible applications}
+
+- clinical trial
+- A/B testing of websites
+- online content recommandation
+- wireless communication, lots of possibilities, one is frequency/channel selection in the OSA paradigm, detailed after
+- hyper-parameters tuning of deep neural networks etc
+
+\subsection{Applications for Opportunistic Spectrum Access}
+
+Detail the analogy arm = channel in a licensed spectrum, rewards $r_k(t)$ = 1 if no Primary User or 0 if channel $k$ is busy at time $t$.
+
+Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
+
+% ----------------------------------------------------------------------------
+\section{A short review of famous MAB algorithms}
+\label{sec:2:famousMABalgorithms}
+
+Everything is implemented and documented at
+https://smpybandits.github.io/docs/Policies.html
+
+\subsubsection{Why do we need smart algorithms?}
+Why Follow-the-Leader (EmpiricalMeans https://smpybandits.github.io/docs/Policies.EmpiricalMeans.html) don't work, example.
+
+\subsection{Simple but efficient algorithms when well tuned: $\varepsilon$-greedy and Explore-then-Commit}
+
+See
+https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
+https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
+
+% ----------------------------------------------------------------------------
+\subsection{Index policies: UCB and others}
+
+\subsubsection{UCB}
+Optimism in face of uncertainty, for UCB etc.
+
+\subsubsection{klUCB}
+Optimal upper confidence bounds based on Kullback-Leibler divergence and klUCB indexes
+
+\subsubsection{Other policies inspired by UCB}
+
+- UCB-H and klUCB-H
+- UCB-V
+- UCB-dagger
+- UCBoost
+- UCB+ and klUCB+
+
+% ----------------------------------------------------------------------------
+\subsection{Bayesian policies: Thompson Sampling and others}
+
+Idea behind Thompson sampling and Bayesian philosophy.
+
+The Beta posterior and others (https://smpybandits.github.io/docs/Policies.Posterior.html) and Thompson Sampling.
+
+% ----------------------------------------------------------------------------
+\subsection{Other policies: BESA and others}
+
+It's a good opportunity to show off and say that I worked on my own on some non-standard policies.
+
+- One of the first paper I read in 2016 was defining AdBandits, https://smpybandits.github.io/docs/Policies.AdBandits.html, a mixture between Thompson sampling and BayesUCB (very efficient in practice)
+- I especially love the BESA algorithm by [? and ? and Maillard].
+
+% ----------------------------------------------------------------------------
+\subsection{Policies for adversarial bandits: Exp3 and others}
+
+Exp3, Exp3++ etc
+https://smpybandits.github.io/docs/Policies.Exp3.html
+https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
+
+Softmax: is it the same, but written differently?
+https://smpybandits.github.io/docs/Policies.Softmax.html
+
+% ----------------------------------------------------------------------------
+\subsection{Hybrid policies: MOSS, klUCB++ and klUCB-Switch and others}
+
+Recent work.
+https://smpybandits.github.io/docs/Policies.MOSS.html
+https://smpybandits.github.io/docs/Policies.IMED.html
+https://smpybandits.github.io/docs/Policies.klUCBPlusPlus.html
+https://smpybandits.github.io/docs/Policies.klUCBSwitch.html
+
+Talk about these?
+https://smpybandits.github.io/docs/Policies.OCUCB.html
+https://smpybandits.github.io/docs/Policies.OSSB.html
+
+With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
+Randomized Confidence Bound (RCB) with uniform perturbations performs as well as UCB!
+https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html
+
+\subsection{Summary of classification of the most famous algorithms}
+
+A huge table in one page in portrait mode, giving:
+- name
+- reference, year
+- family, classification
+- time/memory complexity ?
+- regret bound for stochastic regret
+
+of about 20/30 algorithms?
+
+% ----------------------------------------------------------------------------
+\section{Lower and upper bounds on regret}
+\label{sec:2:lowerUpperBoundsRegret}
+
+\subsection{Measuring performance with the (mean) regret}
+Explain what is regret
+
+\subsection{Regret lower bounds}
+
+- [Lai and Robbins] lower-bound in $\Omega(\log(T))$
+- Worst case lower in $\Omega(\sqrt{T})$
+
+\subsection{Regret upper bounds, tools and different kinds of bounds}
+
+- useful decomposition, with number of samples (see Claire's PhD)
+- different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
+- tools to show regret upper-bounds, concentration inequalities,
+- short proof of UCB or klUCB? just to start to show the tools
+
+\subsection{Other measures of performances}
+
+- Quantile regret or just histogram of regrets
+- Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
+- Fairness
+
+% ----------------------------------------------------------------------------
+\section{Different approaches on algorithm selection, or ``How to choose your preferred bandit algorithm?''}
+\label{sec:2:chooseYourPreferredBanditAlgorithm}
+
+\subsection{Naive approach: use UCB or Thompson sampling, and forget about the rest!}
+First solution: be naive, only use UCB everywhere
+
+\subsection{Use prior knowledge about the future application to select beforehand the chosen algorithm}
+Second solution, if one has some prior knowledge about the domain or setting for which the learning algorithms will be deployed, she can run synthetic or real-world simulations, compare many algorithms before deployment, and select the best algorithm!
+
+\subsection{Online algorithm selection}
+
+\subsubsection{}
+A last solution is online algorithm selection, inspired from expert aggregation.
+Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292
+
+% ----------------------------------------------------------------------------
+\section{Conclusion of Chapter 2}
+\label{sec:2:conclusion}
+
+In this chapter, we saw...
+
+Future works include...
+
+% ----------------------------------------------------------------------------
+\section{Appendix for Chapter 2}
+\label{sec:2:appendix}
+
+\subsection{Numerical computation of Kullback-Leibler divergences and klUCB indexes}
+
+I want to detail the maths behind an implementation of klUCB, see https://smpybandits.github.io/docs/Policies.kullback.html and all what is detailed there.
+
+How to write a fast version of klBern and klucbBern (the two most used functions), using C or Numba or Cython, maybe I can include the speed up simulations here?
+First piece of code, before really presenting SMPyBandits ?

### 2019-02-08 2-Chapters/3-Chapter/chapter3.tex
Better outline
Version 0.1

@@
 % First chapter begins here
-\chapter{Presentation of SMPyBandits, a Complete MAB Simulation Library}
+\chapter{SMPyBandits: a state-of-the-art open-source Python library to simulate multi-armed bandit problems}
 \label{chapter:3}
@@
 % ----------------------------------------------------------------------------
-\section{Review of state-of-the-art single-player algorithms}
+\section{Experimental comparisons of state-of-the-art single-player algorithms}
 \label{sec:3:reviewSPAlgorithms}
@@
 Future works include...
+
+% ----------------------------------------------------------------------------
+\section{Appendix for Chapter 3}
+\label{sec:3:appendix}
+
+\subsection{Include a minimalist \texttt{main.py} file}
+
+Include verbatim the file
+\verb{example_of_main_singleplayer.py}
+
+\subsection{Include a minimalist \texttt{configuration.py} file}
+
+Include verbatim the file
+\verb{example_of_configuration_singleplayer.py}
+
+\subsection{Using multi-core to speed-up simulations}
+
+Quickly explain the advantages and drawbacks of using Joblib and joblib.Parallel to run simulations on multiple cores of the same computer
+
+Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).

### 2019-02-11 1-Preamble/4-Acknowledgement/acknowledgement.tex
Version 0.1
Started to copy/paste my latest paper

@@
 
+% TODO financial thanks
+
+This work is supported by
+the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{XXX}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
+by R\'egion Bretagne, France,
+the French Ministry of Higher Education and Research (MENESR),
+by \'Ecole Normale Sup\'erieure de Paris-Saclay.
+by European Union, through the European Regional Development Fund (ERDF),
+and by Ministry of Higher Education and Research, Brittany and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
+
 I would also like to thank some things. First, the Earth, then my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents.
@@
 % Je souhaite terminer ces remerciements par une pensée toute particulière pour Lola, qui m'a soutenu, qui m'a émerveillé par son enthousiasme et qui m'a réconforté par son amour durant la dernière année.
+Je t'aime chaton, impressionnant !
 Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó, me sorprendió con su entusiasmo y me consoló con su amor durante el año pasado.
+Yo te quiero gatita, ¡impresivo!
 

### 2019-02-11 2-Chapters/1-Chapter/chapter1.tex
Started to copy/paste my latest paper

@@
 \item
-    \emph{Multi-Player Bandits Revisited}},
+    \emph{Multi-Player Bandits Revisited},
     by \textbf{Lilian Besson} \& Emilie Kaufmann.
@@
 \item
-    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access}},\\
+    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},\\
     by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
@@
 \item
-    \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings}},
+    \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings},
     by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
@@
 % =============================================================================
-\subection{Submitted works or in progress works waiting for a new submission}
+\subsection{Submitted works or in progress works waiting for a new submission}
 
@@
 \item
-    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits}},
+    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
     by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
@@
     I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{Cf.
-\texttt{bitbucket.org/scee_ietr/phd-student-day-ietr-2017-bonnefoi-and-besson/downloads/poster.pdf}}.
+\texttt{hal.inria.fr/hal-02013839/document}}.
     I also helped for the organization and presented another poster\footnote{Cf.
-\texttt{bitbucket.org/lbesson/phd-student-day-ietr-2018/downloads/poster.pdf}} at IETR seminar (\texttt{seminar-ietr-18.sciencesconf.org}), held in Vannes in June $2018$.
+\texttt{hal.inria.fr/hal-02013847/document}} at IETR seminar (\texttt{seminar-ietr-18.sciencesconf.org}), held in Vannes in June $2018$.
 
@@
 	for \emph{European Workshop on Reinforcement Learning} (Lille) \texttt{ewrl.wordpress.com/ewrl14-2018}, in October $2018$, I reviewed $5$ papers.
-	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \uref{https://nips.cc/Conferences/2018/}{\emph{NeurIPS}} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.
+	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \href{https://nips.cc/Conferences/2018/}{\emph{NeurIPS}} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.
 

### 2019-02-11 2-Chapters/4-Chapter/chapter4.tex
Better outline
Started to copy/paste my latest paper

@@
 
-- ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \verb{https://perso.crans.org/besson/articles/BMBM__IEEE_WCNC_2019.pdf}
+- ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}
 
@@
 
+\input{2-Chapters/4-Chapter/CROWNCOM2017.tex}
+
 % ----------------------------------------------------------------------------
@@
 
+\input{2-Chapters/4-Chapter/ICT2018_and_WCNC2019.tex}
+
 % ----------------------------------------------------------------------------
@@
 
-- ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \verb{https://perso.crans.org/besson/articles/BMBM__IEEE_WCNC_2019.pdf}
+% - ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}
+
+\input{2-Chapters/4-Chapter/WCNC2019_restransmissions.tex}
 
@@
 Future works include...
+
+% ----------------------------------------------------------------------------
+\section{Appendix for Chapter 4}
+\label{sec:4:appendix}

### 2019-02-11 2-Chapters/4-Chapter/chapter4.tex
Almost done for copy/pastying the chapter 4 based on 4 publications (the longer)
I should start to include the code of articles, and not just the skeleton

@@
 \graphicspath{{2-Chapters/4-Chapter/Images/}}
+
+\TODOL{Write a nice introduction for this chapter.}
+
+\TODOL{Present outline of the chapter.}
+
+\TODOL{Give list of publications associated with this chapter.}
 
@@
 
-- ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419
+% - ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419
 
-\input{2-Chapters/4-Chapter/CROWNCOM2017.tex}
+\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex}
 
@@
 
-- ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825
+% - ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825
 
-\input{2-Chapters/4-Chapter/ICT2018_and_WCNC2019.tex}
+\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex}
 
@@
 
-\input{2-Chapters/4-Chapter/WCNC2019_restransmissions.tex}
+\input{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex}
 
@@
 \label{sec:4:conclusion}
+
+\TODOL{Write a conclusion, and future works}
 
@@
 \label{sec:4:appendix}
+
+% \TODOL{Include here additional results (proof in Section from CROWNCOM ?)}
+\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex}

### 2019-02-11 2-Chapters/6-Chapter/chapter6.tex
Fix #4
Better outline
Hm okay it almost compiles with Chapter 6. I should continue this copy/paste with Chapter 5 (ALT 2018) and quickly with parts of Chapter 2 and 3

@@
 
-% ----------------------------------------------------------------------------
-\section{The piece-wise non-stationary bandits model}
-\label{sec:6:piecewiseModel}
-
-% ----------------------------------------------------------------------------
-\section{The Bernoulli Generalized Likelihood Ratio Test}
-\label{sec:6:piecewiseModel}
-
-% ----------------------------------------------------------------------------
-\section{The GLRklUCB Algorithm}
-\label{sec:6:piecewiseModel}
-
-% ----------------------------------------------------------------------------
-\section{Regret analysis for the GLRklUCB alRegret analysis for the GLRklUCB algorithm}
-\label{sec:6:piecewiseModel}
-
-% ----------------------------------------------------------------------------
-\section{Numerical simulations for piece-wise stationary bandit problems}
-\label{sec:6:piecewiseModel}
+\input{2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex}
 
@@
 % ----------------------------------------------------------------------------
-\appendix
+\section{Appendix for Chapter 6}
+\label{sec:6:appendix}
 
-\section{FIXME appendix in a chapter: where is it?}
+\input{2-Chapters/6-Chapter/nonstatbandits.git/NonStatB_appendix.tex}

### 2019-02-12 2-Chapters/1-Chapter/chapter1.tex
Okay for outline

@@
 % ----------------------------------------------------------------------------
+\section{Organization of the thesis}
+\label{sec:1:organization}
+
+Here we discuss quickly the organization of the thesis.
+
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.80\textwidth}{!}{
+    \begin{tikzpicture}[>=latex',line join=bevel,scale=3]
+        %
+        \node[align=center] (introduction) at (0,2.5) [square,draw] {Chapter~\ref{chapter:1}\\Introduction};
+        \node[align=center] (chapter2) at (0,2) [square,draw] {Chapter~\ref{chapter:2}\\Multi-Armed Bandit models};
+        \node[align=center] (chapter3) at (4,2) [square,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
+        \node[align=center] (chapter4) at (-2,1) [square,draw] {Chapter~\ref{chapter:4}\\Two MAB models for IoT networks};
+        \node[align=center] (chapter5) at (0,1) [square,draw] {Chapter~\ref{chapter:5}\\Multi-players MAB};
+        \node[align=center] (chapter6) at (2,1) [square,draw] {Chapter~\ref{chapter:6}\\Non-stationary MAB models};
+        \node[align=center] (conclusion) at (0,0) [square,draw] {Chapter~\ref{chapter:conclusion}\\General conclusion};
+        %
+        \draw [color=black,thick,->] (introduction) to (chapter2);
+        \draw [color=black,thick,<->] (chapter2) to (chapter3);
+        \draw [color=black,thick,->] (chapter2) to (chapter4);
+        \draw [color=black,thick,->] (chapter2) to (chapter5);
+        \draw [color=black,thick,->] (chapter2) to (chapter6);
+        \draw [color=black,thick,->] (chapter4) to (conclusion);
+        \draw [color=black,thick,->] (chapter5) to (conclusion);
+        \draw [color=black,thick,->] (chapter6) to (conclusion);
+        %
+    \end{tikzpicture}
+    }
+    \caption{Organization of the thesis.}
+    \label{fig:1:organization}
+\end{figure}
+
+% ----------------------------------------------------------------------------
 \section{List of publications}
-\label{sec:1:contributions}
+\label{sec:1:listPublications}
 

### 2019-02-12 1-Preamble/4-Acknowledgement/acknowledgement.tex
Okay for tikz figure showing outline of the thesis

@@
 This work is supported by
-the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{XXX}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
+the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{N ANR-16-CE40-0002}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
 by R\'egion Bretagne, France,

### 2019-02-12 2-Chapters/2-Chapter/chapter2.tex
Okay for tikz figure showing outline of the thesis

@@
 
-\subsection{Online algorithm selection}
+\subsection{Online algorithm selection with expert aggregation}
 
-\subsubsection{}
+\input{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex}
+
 A last solution is online algorithm selection, inspired from expert aggregation.

### 2019-02-12 2-Chapters/1-Chapter/chapter1.tex
Okay for tikz figure showing outline of the thesis
Ok for tex root

--- 2019-02-11 2-Chapters/1-Chapter/chapter1.tex
@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 
@@
 % ----------------------------------------------------------------------------
+\section{Organization of the thesis}
+\label{sec:1:organization}
+
+Here we discuss quickly the organization of the thesis.
+
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.85\textwidth}{!}{
+    \begin{tikzpicture}[>=latex',line join=bevel,scale=3]
+        %
+        \node[align=center] (introduction) at (0,2.5) [rectangle,draw] {Chapter~\ref{chapter:1}\\Introduction};
+        \node[align=center] (chapter2) at (0,2) [rectangle,draw] {Chapter~\ref{chapter:2}\\Multi-Armed Bandit models};
+        \node[align=center] (chapter3) at (4,2) [rectangle,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
+        \node[align=center] (chapter4) at (-2,1) [rectangle,draw] {Chapter~\ref{chapter:4}\\Two MAB models for IoT networks};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapter~\ref{chapter:5}\\Multi-players MAB};
+        \node[align=center] (chapter6) at (2,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary MAB models};
+        \node[align=center] (conclusion) at (0,0) [rectangle,draw] {Chapter~\ref{chapter:conclusion}\\General conclusion};
+        %
+        \draw [color=black,thick,->] (introduction) to (chapter2);
+        \draw [color=black,thick,<->] (chapter2) to (chapter3);
+        \draw [color=black,thick,->] (chapter2) to (chapter4);
+        \draw [color=black,thick,->] (chapter2) to (chapter5);
+        \draw [color=black,thick,->] (chapter2) to (chapter6);
+        \draw [color=black,thick,->] (chapter4) to (conclusion);
+        \draw [color=black,thick,->] (chapter5) to (conclusion);
+        \draw [color=black,thick,->] (chapter6) to (conclusion);
+        %
+    \end{tikzpicture}
+    }
+    \caption{Organization of the thesis.}
+    \label{fig:1:organization}
+\end{figure}
+
+% ----------------------------------------------------------------------------
 \section{List of publications}
-\label{sec:1:contributions}
+\label{sec:1:listPublications}
 

### 2019-02-12 2-Chapters/2-Chapter/chapter2.tex
Ok for tex root

@@
 
-\subsection{Applications for Opportunistic Spectrum Access}
+\subsection{Application for Opportunistic Spectrum Access}
 
@@
 % ----------------------------------------------------------------------------
-\section{A short review of famous MAB algorithms}
+\section{Review of MAB algorithms}
 \label{sec:2:famousMABalgorithms}
@@
 
-\subsection{Regret lower bounds}
+\subsection{Regret lower bounds: do not detail too much, don't explain the tools behind the results}
 
 - [Lai and Robbins] lower-bound in $\Omega(\log(T))$
-- Worst case lower in $\Omega(\sqrt{T})$
+- Worst case lower-bound in $\Omega(\sqrt{T})$
+- Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models)
 
@@
 
-\subsection{Numerical computation of Kullback-Leibler divergences and klUCB indexes}
+\subsection{Eficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
 
@@
 First piece of code, before really presenting SMPyBandits ?
+
+My work for this part is also published on https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes
+(in Python) and https://github.com/Naereen/KullbackLeibler.jl (in Julia).

### 2019-02-12 2-Chapters/3-Chapter/chapter3.tex
Started to copy/paste my latest paper
Ok for tex root

@@
 % ----------------------------------------------------------------------------
-\section{An example of another model: rested or restless Markov models}
+\section{An example of another model: rested or restless Markov models - FIXME maybe not included?}
 \label{sec:3:markovModels}
@@
 
+\begin{minted}[mathescape,
+    linenos,
+    numbersep=5pt,
+    gobble=2,
+    frame=lines,
+    framesep=2mm]{csharp}
+string title = "This is a Unicode π in the sky"
+/*
+Defined as $\pi=\lim_{n\to\infty}\frac{P_n}{d}$ where $P$ is the perimeter
+of an $n$-sided regular polygon circumscribing a
+circle of diameter $d$.
+*/
+const double pi = 3.1415926535
+\end{minted}
+
 \subsection{Include a minimalist \texttt{main.py} file}
@@
 Include verbatim the file
-\verb{example_of_main_singleplayer.py}
+\texttt{example\_of\_main\_singleplayer.py}
+
+% \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
 
@@
 Include verbatim the file
-\verb{example_of_configuration_singleplayer.py}
+\texttt{example\_of\_configuration\_singleplayer.py}
+
+% \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
 

### 2019-02-12 2-Chapters/3-Chapter/chapter3.tex
Still fighting against my template...

@@
 
-\begin{minted}[mathescape,
-    linenos,
-    numbersep=5pt,
-    gobble=2,
-    frame=lines,
-    framesep=2mm]{csharp}
-string title = "This is a Unicode π in the sky"
-/*
-Defined as $\pi=\lim_{n\to\infty}\frac{P_n}{d}$ where $P$ is the perimeter
-of an $n$-sided regular polygon circumscribing a
-circle of diameter $d$.
-*/
-const double pi = 3.1415926535
+\begin{listing}[H]
+\begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python}
+from SMPyBandits.Arms import Bernoulli
+arms = [Bernoulli(0.1), Bernoulli(0.9)]
+
+from SMPyBandits.Environment import MAB
+my_MAB_problem = MAB(arms)
+nbArms = my_MAB_problem.nbArms
+
+from SMPyBandits.Policy import UCB
+my_UCB_algo = UCB(nbArms, alpha=0.5)
+
+horizon = 1000
+for t in range(horizon):
+    chosen_arm = my_UCB_algo.choice()
+    observed_reward = my_MAB_problem.draw(chosen_arm)
+    my_UCB_algo.getReward(chosen_arm, observed_reward)
+...
 \end{minted}
+\caption{First example of code}
+\label{lst:3:example}
+\end{listing}
 
@@
 
+\begin{listing}[H]
 % \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
+\inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{Python}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
+\caption{Small example of \texttt{main.py} file}
+\label{lst:3:main}
+\end{listing}
 
@@
 
+\begin{listing}[H]
 % \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
+\inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{Python}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
+\caption{Small example of \texttt{configuration.py} file}
+\label{lst:3:configuration}
+\end{listing}
 

### 2019-02-12 4-Listofpublications/Listofpublications.tex
Ok for tex root
Still fighting against my template...

@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 % ******************************* Thesis Appendix A ****************************
@@
   Radio Implementation of MALIN: « Multi-Armed bandits Learning for
-  Internet-of-things Networks »}, par \textbf{L. Besson},
+  Internet-of-things Networks »}, by \textbf{L. Besson},
   \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
@@
   \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
-  Marrakech, Maroc, avril 2019.
+  Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02006825}{HAL-02006825}.
 \item
@@
   « Multi-Arm bandit Learning for Iot Networks » with GRC: A TestBed
-  Implementation and Demonstration that Learning Helps}, par \textbf{L.
+  Implementation and Demonstration that Learning Helps}, by \textbf{L.
   Besson}, \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
-  \href{https://moychris.wordpress.com/}{C. Moy}. Démonstration
-  présentée à ICT (\emph{International Conference on Communication}),
-  Saint-Malo, France, juin 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}
+  \href{https://moychris.wordpress.com/}{C. Moy}. Demonstration
+  presented at ICT (\emph{International Conference on Communication}),
+  Saint-Malo, France, June 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}, \href{https://hal.inria.fr/hal-02013866}{HAL-02013866}
 \item
   \href{https://hal.inria.fr/hal-01629733/document}{Multi-Player Bandits
-  Revisited}, par \textbf{L. Besson} \&
+  Revisited}, by \textbf{L. Besson} \&
   \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
@@
   \href{http://www.cs.cornell.edu/conferences/alt2018/accepted.html}{ALT}
-  (\emph{Algorithmic Learning Theory}), Lanzarote, Espagne, avril 2018,
+  (\emph{Algorithmic Learning Theory}), Lanzarote, Spain, April 2018,
   \href{https://hal.inria.fr/hal-01629733}{HAL-01629733}.
@@
   Multi-Armed Bandits learning algorithms for Opportunistic Spectrum
-  Access}, par \textbf{L. Besson},
+  Access}, by \textbf{L. Besson},
   \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
   Kaufmann} \& \href{https://moychris.wordpress.com/}{C. Moy}.
-  \href{http://wcnc2018.ieee-wcnc.org/}{IEEE WCNC}, Barcelone, Espagne,
+  \href{http://wcnc2018.ieee-wcnc.org/}{IEEE WCNC}, Barcelone, Spain,
   avril 2018, \href{https://hal.inria.fr/hal-01705292}{HAL-01705292}.
@@
 \begin{etaremune}
-	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017
+	\item \textbf{L.
+Besson}, E.
+Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}
 	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
 	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
-	\item	R. Bonnefoi, \textbf{L. Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016
+	\item	R.
+Bonnefoi, \textbf{L.
+Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016, \href{https://hal.inria.fr/hal-02013839}{HAL-02013839}
 \end{etaremune}
@@
 	Open-Source Research Framework for Single and Multi-Players Multi-Arms
-	Bandits (MAB) Algorithms in Python}, par \textbf{L. Besson},
+	Bandits (MAB) Algorithms in Python}, by \textbf{L. Besson},
 	développement depuis octobre 2016 et encore actif,
@@
 	\href{https://hal.inria.fr/hal-01736357/document}{What Doubling-Trick
-	Can and Can't Do for Multi-Armed Bandits}, par \textbf{L. Besson} \&
+	Can and Can't Do for Multi-Armed Bandits}, by \textbf{L. Besson} \&
 	\href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
@@
 	\href{https://hal.inria.fr/hal-01847480/document}{A Note on the Ei
-	Function and a Useful Sum-Inequality}, par \textbf{L. Besson}, février
+	Function and a Useful Sum-Inequality}, by \textbf{L. Besson}, février
 	2018 \href{https://hal.inria.fr/hal-01847480}{HAL-01847480}.

### 2019-02-12 Nomenclature/Nomenclature.tex
Still fighting against my template...

@@
-\nomenclature[A]{AM}{\textit{Amplitude Modulation}}
-\nomenclature[A]{ARTM}{\textit{Advanced Range Telemetry Group}}
-\nomenclature[A]{COFDM}{\textit{Coded Orthogonal Frequency Division Multiplexing}}
-\nomenclature[A]{CPM}{\textit{Continuous Phase Modulation}}
-\nomenclature[A]{CPFSK}{\textit{Continuous Phase Frequency Shift Keying}}
-\nomenclature[A]{FM}{\textit{Frequency Modulation}}
-\nomenclature[A]{PCM/FM}{\textit{Pulse Code Modulation/Frequency Modulation}}
-\nomenclature[A]{MHz}{\textit{Megahertz}}
-\nomenclature[A]{GHz}{\textit{Gigahertz}}
-\nomenclature[A]{Hz}{\textit{Hertz}}
-\nomenclature[A]{dB}{\textit{Decibel}}
-\nomenclature[A]{NRZ}{\textit{Non-Return to Zero}}
-\nomenclature[A]{SOQPSK}{\textit{Shaped Offset Quadrature Phase-Shift Keying}}
-\nomenclature[A]{SOQPSK-TG}{\textit{Shaped Offset Quadrature Phase-Shift Keying Telemetry Group}}
-\nomenclature[A]{SOQPSK-MIL}{\textit{Shaped Offset Quadrature Phase-Shift Keying Military}}
-\nomenclature[A]{MIL-STD SOQPSK}{\textit{Military Standard Shaped Offset Quadrature Phase-Shift Keying }}
-\nomenclature[A]{APSK}{\textit{Amplitude Phase-Shift Keying}}
-\nomenclature[A]{PSK}{\textit{Phase-Shift Keying}}
-\nomenclature[A]{M-PSK}{\textit{M-ary Phase-Shift Keying}}
-\nomenclature[A]{BPSK}{\textit{Binary Phase-Shift Keying}}
-\nomenclature[A]{OQPSK}{\textit{Offset Quadrature Phase-Shift Keying}}
-\nomenclature[A]{PAM}{\textit{Pulse Amplitude Modulation}}
-\nomenclature[A]{M-QAM}{\textit{M-ary Quadrature Amplitude Modulation}}
-\nomenclature[A]{MIMO}{\textit{Multiple Input Multiple Output}}
-\nomenclature[A]{SISO}{\textit{Single Input Single Output}}
-\nomenclature[A]{MISO}{\textit{Multiple Input Single Output}}
-\nomenclature[A]{RF}{\textit{Radio Frequency}}
-\nomenclature[A]{IRIG}{\textit{Inter-Range Instrumentation Group}}
-\nomenclature[A]{RCC}{\textit{Range Commanders Council}}
-\nomenclature[A]{ISI}{\textit{Inter-Symbol Interference}}
-\nomenclature[A]{PRD}{\textit{Perrins-Rice Decomposition}}
-\nomenclature[A]{DBD}{\textit{DuoBinary Decomposition}}
-\nomenclature[A]{XTCQM}{\textit{Cross-correlated Trellis Coded Modulation}}
-\nomenclature[A]{I$\&$D}{\textit{Integrate and Dump}}
-\nomenclature[A]{CM}{\textit{Constant Modulus}}
-\nomenclature[A]{CMA}{\textit{Constant Modulus Algorithm}}
-\nomenclature[A]{LOS}{\textit{Line Of Sight}}
-\nomenclature[A]{NLOS}{\textit{Non-Line Of Sight}}
-\nomenclature[A]{STC}{\textit{Space-Time Coding}}
-\nomenclature[A]{STBC}{\textit{Space-Time Block Coding}}
-\nomenclature[A]{PSP}{\textit{PerSurvivor Processing}}
-\nomenclature[A]{GMSK}{\textit{Gaussian Minimum Shift Keying}}
-\nomenclature[A]{MSK}{\textit{Minimum Shift Keying}}
-\nomenclature[A]{LS}{\textit{Least Square}}
-\nomenclature[A]{LMS}{\textit{Least Mean Square}}
-\nomenclature[A]{RLS}{\textit{Recursive Least Square}}
-\nomenclature[A]{GSM}{\textit{Global System for Mobile communication }}
-\nomenclature[A]{BT}{\textit{Bandwidth Time product}}
-\nomenclature[A]{GFSK}{\textit{Gaussian Frequency Shift Keying}}
-\nomenclature[A]{SBPSK}{\textit{Shaped Binary Phase-Shift Keying}}
-\nomenclature[A]{DVB-RCS2}{\textit{Digital Video Broadcasting - Return Channel via Satellite - Second Generation}}
-\nomenclature[A]{FQPSK}{\textit{Feher-patented Quadrature Phase-Shift Keying}}
-\nomenclature[A]{RC}{\textit{Raised Cosine}}
-\nomenclature[A]{REC}{\textit{Rectangular}}
-\nomenclature[A]{BCJR}{\textit{Bahl, Cocke, Jelinek and Raviv}}
-\nomenclature[A]{MAP}{\textit{Maximum A Posteriori}}
-\nomenclature[A]{MLSD}{\textit{Maximum Likelihood Sequence Detection}}
-\nomenclature[A]{CA-MLSD}{\textit{Conventional Adaptive Maximum Likelihood Sequence Detection}}
-\nomenclature[A]{NA-MLSD}{\textit{Non-Adaptive Maximum Likelihood Sequence Detection}}
-\nomenclature[A]{MLSD}{\textit{Maximum Likelihood Sequence Detection}}
-\nomenclature[A]{MLSE}{\textit{Maximum Likelihood Sequence Estimation}}
-\nomenclature[A]{Mbps}{\textit{Megabits per second}}
-\nomenclature[A]{MMSE}{\textit{Minimum Mean Square Error}}
-\nomenclature[A]{MSE}{\textit{Mean Square Error}}
-\nomenclature[A]{SMA}{\textit{Saint Martin Airport}}
-\nomenclature[A]{TBA}{\textit{Toulouse-Blagnac Airport}}
-\nomenclature[A]{V2V}{\textit{Vehicle To Vehicle}}
-\nomenclature[A]{AWGN}{\textit{Additive White Gaussian Noise}}
-\nomenclature[A]{SxS}{\textit{Sample by Sample}}
-\nomenclature[A]{SNR}{\textit{Signal-to-Noise Ratio}}
-\nomenclature[A]{SER}{\textit{Symbol Error Rate}}
-\nomenclature[A]{BER}{\textit{Bit Error Rate}}
-\nomenclature[A]{ASM}{\textit{Asynchronous Marker}}
-\nomenclature[A]{INET}{\textit{Integrated Network Enhanced Telemetry}}
-\nomenclature[A]{STBD}{\textit{Space-Time Block Decoder}}
-\nomenclature[A]{ML}{\textit{Maximum Likelihood}}
-\nomenclature[A]{ML-ASM}{\textit{Maximum Likelihood - Approximate Signal Model}}
-\nomenclature[A]{OD}{\textit{Offset Decode}}
-\nomenclature[A]{SOVA}{\textit{Soft Output Viterbi Algorithm}}
-\nomenclature[A]{LLR}{\textit{Log-Likelihood Ratio}}
-\nomenclature[A]{FIR}{\textit{Finite Impulse Response}}
-\nomenclature[A]{Max-Log-MAP}{\textit{Maximum Logarithm Maximum A Posteriori}}
-\nomenclature[A]{FEC}{\textit{Forward Error Correction}}
-\nomenclature[A]{LDPC}{\textit{Low Parity Density Check}}
-\nomenclature[A]{CFO}{\textit{Carrier Frequency Offset}}
-\nomenclature[A]{US}{\textit{United States}}
-\nomenclature[A]{DF}{\textit{Decision Feedback}}
-\nomenclature[A]{I.I.D}{\textit{Independent and Identically Distributed}}
-\nomenclature[A]{ML-ASM}{\textit{Maximum Likelihood - Approximate Signal Model}}
-\nomenclature[A]{NASA's JPL}{\textit{National Aeronautics and Space Administration's Jet Propulsion Lab}}
-%\nomenclature[N]{$x$}{Scalar}
+\nomenclature[A]{UCB}{\textit{Upper Confidence Bound} (object or algorithm)}
+\nomenclature[A]{klUCB}{\textit{Kullback-Leibler Upper Confidence Bound} (algorithm)}
+\nomenclature[A]{KL}{\textit{Kullback-Leibler} (divergence)}
+\nomenclature[A]{MCTopM}{\textit{Musical-Chair on Top-M} (algorithm)}
+\nomenclature[A]{RandTopM}{\textit{Random Hoping Musical-Chair on Top-M} (algorithm)}
+\nomenclature[A]{GLR}{\textit{Generalized Likelihood Ratio}}
+\nomenclature[A]{GLRT}{\textit{Generalized Likelihood Ratio Test}}

### 2019-02-12 2-Chapters/3-Chapter/chapter3.tex
Fix #11

@@
 \begin{listing}[H]
-\begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python}
+\begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}
 from SMPyBandits.Arms import Bernoulli
@@
 % \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
-\inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{Python}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
+\inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
 \caption{Small example of \texttt{main.py} file}
@@
 % \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
-\inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{Python}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
+\inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
 \caption{Small example of \texttt{configuration.py} file}

### 2019-02-12 2-Chapters/3-Chapter/chapter3.tex
Fix #12

--- 2019-02-12 2-Chapters/3-Chapter/chapter3.tex
@@
 % First chapter begins here
-\chapter{SMPyBandits: a state-of-the-art open-source Python library to simulate multi-armed bandit problems}
+\chapter{SMPyBandits: a state-of-the-art open-source Python library to simulate MAB problems}
 \label{chapter:3}
@@
 
-\begin{minted}[mathescape,
-    linenos,
-    numbersep=5pt,
-    gobble=2,
-    frame=lines,
-    framesep=2mm]{csharp}
-string title = "This is a Unicode π in the sky"
-/*
-Defined as $\pi=\lim_{n\to\infty}\frac{P_n}{d}$ where $P$ is the perimeter
-of an $n$-sided regular polygon circumscribing a
-circle of diameter $d$.
-*/
-const double pi = 3.1415926535
-\end{minted}
+\subsection{Minimalist example of use of SMPyBandits}
 
-\subsection{Include a minimalist \texttt{main.py} file}
+Below is presented in Listing~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
+Only the mean reward is printed at the end of the simulation.
+%
+With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
+\begin{verbatim}
+The UCB algorithm obtains here a mean reward = 0.886
+\end{verbatim}
 
-Include verbatim the file
-\texttt{example\_of\_main\_singleplayer.py}
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+% \begin{listing}[h!]
+    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py}
+    % \caption{Small example of \texttt{main.py} file}
+    \captionof{listing}{Example of use of SMPyBandits \label{lst:3:exampleOfUse}.}
+    % \label{lst:3:exampleOfUse}
+% \end{listing}
+\end{small}
 
-% \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
+\subsection{A minimalist \texttt{main.py} file}
 
-\subsection{Include a minimalist \texttt{configuration.py} file}
+We present below in Listing~\ref{lst:3:main} a full example of a minimalist \texttt{main.py} file,
+which is used to load the configuration, run the experiment and then print and plot various visualizations for the simulation.
 
-Include verbatim the file
-\texttt{example\_of\_configuration\_singleplayer.py}
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+% \begin{listing}[h!]
+    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
+    % \caption{Small example of \texttt{main.py} file}
+    \captionof{listing}{Small example of \texttt{main.py} file\label{lst:3:main}.}
+    % \label{lst:3:main}
+% \end{listing}
+\end{small}
 
-% \lstinputlisting[language=Python]{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
+\subsection{A minimalist \texttt{configuration.py} file}
+
+We present below in Listing~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
+which is used to define the experiment.
+It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
+It uses environment variables, so a user can configure some parameters of the simulation when launching them:
+
+\begin{verbatim}
+    (bash) $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
+\end{verbatim}
+
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+% \begin{listing}[h!]
+    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
+    % \caption{Small example of \texttt{configuration.py} file}
+    \captionof{listing}{Small example of \texttt{configuration.py} file\label{lst:3:configuration}.}
+    % \label{lst:3:configuration}
+% \end{listing}
+\end{small}
 

### 2019-02-12 2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py

@@ -1,0 +1,106 @@
+""" An example of a configuration file to launch some the simulations
+For the single-player case.
+"""
+CPU_COUNT = 4
+from os import getenv
+
+# Import arms
+from SMPyBandits.Arms import *
+# Import contained classes
+from SMPyBandits.Environment import MAB
+# Import single-player algorithms
+from SMPyBandits.Policies import *
+
+#: HORIZON : number of time steps of the experiments.
+#: Warning Should be >= 10000 to be interesting "asymptotically".
+HORIZON = int(getenv('T', 10000))
+
+#: REPETITIONS : number of repetitions of the experiments.
+#: Warning: Should be >= 10 to be statistically trustworthy.
+REPETITIONS = int(getenv('N', 100))
+
+#: Number of jobs to use for the parallel computations.
+#: -1 means all the CPU cores, 1 means no parallelization.
+N_JOBS = int(getenv('N_JOBS', -1))
+
+#: Number of arms for non-hard-coded problems (Bayesian problems)
+NB_ARMS = int(getenv('K', 3))
+
+#: Lower value of means
+LOWER = float(getenv('LOWER', 0))
+#: Amplitude value of means
+AMPLITUDE = float(getenv('AMPLITUDE', 1))
+
+#: Type of arms for non-hard-coded problems (Bayesian problems)
+ARM_TYPE = mapping_ARM_TYPE[str(getenv('ARM_TYPE', "Bernoulli"))]
+
+#: Means of arms for non-hard-coded problems
+MEANS = uniformMeans(nbArms=NB_ARMS, delta=0.05,
+    lower=LOWER, amplitude=AMPLITUDE, isSorted=True
+)
+
+import numpy as np
+# more parametric? Read from cli?
+MEANS_STR = getenv('MEANS', '')
+if MEANS_STR != '':
+    MEANS = [ float(m) for m in MEANS_STR.replace('[', '').replace(']', '').split(',') ]
+    print("Using cli env variable to use MEANS = {}.".format(MEANS))  # DEBUG
+
+#: This dictionary configures the experiments
+configuration = {
+    # --- Duration of the experiment
+    "horizon": HORIZON,
+    # --- Number of repetition of the experiment (to have an average)
+    "repetitions": REPETITIONS,
+    # --- Parameters for the use of joblib.Parallel
+    "n_jobs": N_JOBS,    # = nb of CPU cores
+    # --- Should we plot the lower-bounds or not?
[…]

### 2019-02-12 2-Chapters/3-Chapter/src/example_of_main_singleplayer.py

@@ -1,0 +1,49 @@
+""" An example of a simple 'main' script.
+Main scripts load the config, run the simulations, and plot them/
+For the single player case.
+"""
+from example_of_configuration_singleplayer import configuration
+from SMPyBandits.Environment import Evaluator
+
+configuration['showplot'] = True
+evaluation = Evaluator(configuration)
+
+# Start the evaluation and then print final ranking and plot, for each environment
+for envId, env in enumerate(evaluation.envs):
+    # Evaluate just that env
+    evaluation.startOneEnv(envId, env)
+
+# Compare them
+for envId, env in enumerate(evaluation.envs):
+    evaluation.plotHistoryOfMeans(envId)  # XXX To plot without saving
+
+    print("\nGiving all the vector of final regrets ...")
+    evaluation.printLastRegrets(envId)
+    print("\nGiving the final ranking ...")
+    evaluation.printFinalRanking(envId)
+
+    print("\n\n- Plotting the last regrets...")
+    evaluation.plotLastRegrets(envId, boxplot=True)
+
+    print("\nGiving the mean and std running times ...")
+    evaluation.printRunningTimes(envId)
+    evaluation.plotRunningTimes(envId)
+
+    print("\nGiving the mean and std running times ...")
+    evaluation.printMemoryConsumption(envId)
+    evaluation.plotMemoryConsumption(envId)
+
+    print("\n\n- Plotting the mean reward...")
+    evaluation.plotRegrets(envId, meanReward=True)
+
+    print("\n\n- Plotting the regret...")
+    evaluation.plotRegrets(envId)
+
+    print("\n- Plotting the probability of picking the best arm of time...")
+    evaluation.plotBestArmPulls(envId)
+
+    print("\n- Plotting the histograms of regrets...")
+    evaluation.plotLastRegrets(envId, sharex=True, sharey=True)
+
+# Done
+print("Done for simulations example_of_main_singleplayer ...")

### 2019-02-12 2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py

@@ -1,0 +1,26 @@
+""" Example of use of SMPyBandits.
+See https://smpybandits.github.io/API.html for more"""
+import numpy as np
+np.random.seed(0)  # for reproducibility
+
+from SMPyBandits.Arms import Bernoulli
+arms = [Bernoulli(0.1), Bernoulli(0.9)]
+
+from SMPyBandits.Environment import MAB
+my_MAB_problem = MAB(arms)
+nbArms = my_MAB_problem.nbArms
+
+from SMPyBandits.Policies import UCB
+my_UCB_algo = UCB(nbArms)
+my_UCB_algo.startGame()  # reset internal memory
+
+horizon = 1000
+for t in range(horizon):
+    chosen_arm = my_UCB_algo.choice()
+    observed_reward = my_MAB_problem.draw(chosen_arm)
+    my_UCB_algo.getReward(chosen_arm, observed_reward)
+
+cumulated_reward = sum(my_UCB_algo.rewards)
+number_of_plays = sum(my_UCB_algo.pulls)  # = 1000
+mean_reward = cumulated_reward / number_of_plays
+print("The UCB algorithm obtains here a mean reward =", mean_reward)

### 2019-02-12 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Added sub tex files, I will remove them soon and reorganize all this, but for now...

% - ``Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access'', see https://hal.inria.fr/hal-01705292

For any real-world applications of MAB algorithms, several solutions have been explored based on various models.
However, it is hard to exactly predict which solution could be the best for real-world conditions at every instants.
As we saw before, a possible approach is to pick one algorithm and simply use it, or to run simulations before-hand, if possible.
But a third possible approach is to select \emph{on the run} the best algorithm for a specific situation, for which the so-called \emph{expert aggregation algorithms} can be useful.
%
To the best of our knowledge, aggregation algorithms, such as \ExpQ{} dating back from 2002 (\cite{Auer02}),
have never been used for stochastic MAB problems,
and we show that it appears empirically sub-efficient when applied to simple stochastic problems.
%
In this Section, we present an improved variant of \ExpQ, called \Aggr.
For synthetic MAB problems,
with Bernoulli distributions or other distributions for arms,
simulation results are presented to demonstrate its empirical efficiency.
We combine classical algorithms, such as Thompson sampling, Upper-Confidence Bounds algorithms (\UCB{} and variants), and Bayesian or Kullback-Leibler UCB.
%
Our algorithm offers good performance compared to state-of-the-art algorithms
(\ExpQ{}, \CORRAL{} or \LearnExp{}) \cite{Agarwal16,Singla17}),
and appears as a robust approach to select on the run the best algorithm for any stochastic MAB problem, being more realistic to real-world radio settings than any tuning-based approach.

\TODOL{This chapter is basically a raw include from my paper ``Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access'', see https://hal.inria.fr/hal-01705292}

% ----------------------------------------------------------------------
\subsection{Motivation for online algorithm selection}\label{sub:25:introduction}

Many different learning algorithms have been proposed by the machine learning community,
and most of them depend on several parameters, for instance $\alpha$ for \UCB, the prior for Thompson sampling or BayesUCB,
the $\kl$ function for \klUCB{} etc.
Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T = \E_{\boldsymbol{\mu}}[\widetilde{R_T}]$.
%
For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms.
But when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any kind, size and complexity:
ideally one would like to use an algorithm that can be applied identically against any problem.
To choose the best algorithm, two approaches can be followed:
either extensive benchmarks are done beforehand -- if this is possible -- to select the algorithm and its optimal parameters, or an adaptive algorithm is used to learn \emph{on the fly} its parameters.
We present a simple adaptive solution, that aggregates several learning algorithms in parallel and adaptively chooses which one to trust the most.

\paragraph{Outline}
This Section is organized as follows.
First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
Then we present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
Finally, we present numerical experiments in Section~\ref{sub:25:numExp},
on Bernoulli and non-Bernoulli MAB problems,
comparing the regret of several algorithms against different aggregation algorithms.
Theoretical guarantees are shortly discussed in Section~\ref{sub:25:theory}, and Section~\ref{sub:25:conclusion} concludes.

% ----------------------------------------------------------------------
\subsection{Aggregating Bandit Algorithms}\label{sub:25:aggregation}

We assume to have $N \geq 2$ MAB algorithms, $\Alg_1, \dots, \Alg_N$,
and let $\Alg_{\mathrm{aggr}}$ be an aggregation algorithm,
which runs the $N$ algorithms in parallel (with the same slotted time), and use them to choose its channels based on a voting from their $N$ decisions.
%
$\Alg_{\mathrm{aggr}}$ depends on a pool of algorithms and a set of parameters.
We would like that $\Alg_{\mathrm{aggr}}$
performs almost as well as the best of the $\Alg_a$, with a good choice of its parameters, independently of the MAB problem.
Ideally $\Alg_{\mathrm{aggr}}$ should perform similarly to the best of the $\Alg_a$.
%
To simplify the presentation, we only aggregate bandit algorithms that give deterministic recommendations:
one arm is chosen with probability $1$ and the others with probability $0$.
However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $\{1,\dots,K\}$ at each time step, and draw the next selected arm according to this distribution.

The aggregation algorithm maintains a probability distribution $\pi^{t}$ on the $N$ algorithms $\Alg_a$, starting from a uniform distribution:
$\pi^t_a$ is the probability of trusting the decision made by algorithm $\Alg_a$ at time $t$.
$\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it decides whom to trust by sampling $a \in \{1,\dots,N\}$ from $\pi^t$, then follows $\Alg_a$'s decision.
The main questions are then to know what observations (\ie, arms and rewards) should be given as feedback to which algorithms,
and how to update the trusts at each step, and our proposal \Aggr{} differs from \ExpQ{} on these very points.

% ----------------------------------------------------------------------
\subsubsection{The \Aggr{} algorithm}\label{sub:25:Aggr}

Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
%
\begin{small}
	\begin{figure}[h!]
		\centering
		% Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
		% Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
		\begin{algorithm}[H]
			% XXX Input, data and output
			\KwIn{$N$ bandit algorithms, $\Alg_1, \dots, \Alg_N$, with $N \geq 2$}
			\KwIn{Number of arms, $K \geq 1$}
			\KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$}
			\KwData{Initial uniform distribution, $\pi^{0} = U(\{1, \dots, N\})$}
			\KwResult{$\Alg_{\mathrm{aggr}}=\Aggr\left[\Alg_1, \dots, \Alg_N\right]$}
			% XXX Algorithm
			\For(\tcp*[f]{At every time step})
            {$t = 1, \dots, T$}{
				%
				% \tcp{First, run each algorithm $\Alg_a$}
				\For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
					$\Alg_a$ updates its indexes (e.g., \UCB{} indexes)\;
					It chooses $A^{t+1}_{a} \in \{1, \dots, K\}$.
				}
				%
				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A^{t+1}_{a} = j\}), \;\forall 1\leq j \leq K$\;
				% \tcp*{Sum trusts}
				Then $\Alg_{\mathrm{aggr}}$ chooses arm $A^{t+1} \sim p^{t+1}$\;
				% $\P(A^{t+1}_{\mathrm{aggr}} = j) = p^{t+1}_j$
				% \tcp*{Sample once}
				%
				Give \emph{original} reward $(A^{t+1}, 1 - \ell_{A^{t+1}}(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
				Compute an \emph{unbiased} estimate of the loss of the trusted algorithms,$$ \ell^{t+1} = \ell_{A^{t+1}}(t+1) / p^{t+1}_{A^{t+1}}\;. $$
				% \tcp{Use bandit feedback to update $\pi^{t}$}
				% \tcp{Finally, use the instantaneous bandit feedback (loss $\ell_{A^{t+1},t}$) to update $\pi^{t}$}
				% \For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
				\For{$a = 1, \dots, N$}{
					% \eIf{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
					\If{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
						$ \pi^{t+1}_{a} = \exp(\eta_t \ell^{t+1}) \times \pi^{t}_{a} $
						% \tcp*{More trusted}
						}%{
						% $ \pi^{t+1}_{a} = \exp(- \eta_t \ell^{t+1}) \times \pi^{t}_{a} $
						% \tcp*{Less trusted}
					% }
				}
				Renormalize the new $\pi$: $\pi^{t+1} := \pi^{t+1} / \sum_{a=1}^{N} \pi^{t+1}_{a}$.
				% \tcp*{Project it back to $\Delta_N$}
			}
			\caption{Our aggregation algorithm \Aggr.}
			\label{algo:25:Aggr}
		\end{algorithm}
	\end{figure}
\end{small}

At every time step, after having observed a loss $\ell_{A^{t+1}}(t+1)$ for its chosen action $A^{t+1}$,
the algorithm updates the trust probabilities from $\pi^t$ to $\pi^{t+1}$ by
a multiplicative exponential factor (using the learning rate and the \emph{unbiased} loss).
%
Only the algorithms $\Alg_a$ who advised the last decision get their trust updated, in order to trust more the ``reliable'' algorithms.
The loss estimate is unbiased in the following sense. If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\tilde{\ell}_{a,t+1} = \ell_{A_a^{t+1}}(t+1)$. This quantity can only be observed for those algorithms for which $A_{a}^{t+1}=A^{t+1}$. However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $\tilde{\ell}_{a,t}$. More precisely,
\[\hat{\ell}_{a,t+1} := \frac{\ell_{A_a^{t+1}}(t+1)}{p_{A_a^{t+1}}^{t+1}} \mathbbm{1}(A_a^{t+1} = A^{t+1})\] satisfies $\mathbb{E}[\hat{\ell}_{a,t+1} | \mathcal{H}_t] = \tilde{\ell}_{a,t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$. Observe that $\tilde{\ell}_{a,t+1}=\ell^{t+1}$ for all algorithms $a$ such that $A_a^{t+1}=A^{t+1}$, and $\tilde{\ell}_{a,t+1}=0$ otherwise.

% - Precise some tricky points
An important feature of \Aggr{} is the feedback provided to each underlying bandit algorithm, upon the observation of arm $A^{t+1}$. Rather than updating only the trusted algorithms (that is the algorithms which would have drawn arm $A^{t+1}$) with the observed reward $r_{A^{t+1}}(t+1)=1 - \ell_{A^{t+1}}(t+1)$,  we found that updating each algorithm with the (original) loss observed for arm $A^{t+1}$ improves the performance drastically.
%
As expected, the more feedback they get, the faster the underlying algorithms learn, and the better the aggregation algorithm will be. This intuition is backed up by theory explained in \cite{Maillard11}.

Regarding the update of $\pi^t$, one can note that the trust probabilities are not all updated before the normalization step,
and an alternative would be to
increase $\pi_{a}$ if $A^{t+1}_a = A^{t+1}$ and to decrease it otherwise.
It would not be so different, as there is a final renormalization step, and empirically this variation has little impact on the performance of \Aggr{}.

% ----------------------------------------------------------------------
\subsubsection{Explanations of \Aggr{} versus \ExpQ{} }\label{sub:25:Exp4}

% - explain, but don't give pseudo-code ?
The \ExpQ{} algorithm (see, e.g. \cite[Section 4.2]{Bubeck12})
is similar to \Aggr{}, presented in Algorithm \ref{algo:25:Aggr},
but differs in the two following points.
%
First, $a\sim\pi^t$ is sampled first and the arm chosen by $\Alg_a$ is trusted, whereas \Aggr{}
needs to listen to the $N$ decisions to perform the updates on $\pi^{t+1}$,
and \ExpQ{} gives back an observation (arm, reward) only to the last trusted algorithm
whereas \Aggr{} gives it to all algorithms.
%
Second, after having computed the loss estimate $\ell$, \ExpQ{} updates the estimated cumulative loss for each algorithm,
$\widetilde{L}_a(t) = \sum_{s=1}^{t} \ell_{A_a^s}(s) \times \mathbbm{1}(A^{s}_{a} = A^{s}_{\mathrm{aggr}})$.
%
Instead
\footnote{It is the same in the case of constant learning rates, and does not differ much even for decreasing learning rates.}
of updating $\pi^{t}$ multiplicatively as we do for our proposal, \ExpQ{} recomputes it, proportionally to
$\exp(- \eta_t \widetilde{L}_a(t))$.

\paragraph{How to choose the learning rates $(\eta_t)$?}
%
The sequence of non-negative learning rates $(\eta_t)_{t \geq 1}$ used by \ExpQ{} can be arbitrary, it can be constant but should be non-increasing \cite[Theorem 4.2]{Bubeck12}.
% helps to choose a good sequence.
If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t = \eta := 2 \log(N) / (T K)$.
However, for real-world communication problems, it is unrealistic to assume a fixed and known time horizon, so we prefer the alternative horizon-free choice%
\footnote{~It is non-increasing, and is obtained by minimizing the upper-bound on the contextual regret derived in \cite[pp48]{Bubeck12}.}
of learning rates,
$\eta_t := \log(N) / (t K)$ suggested by \cite{Bubeck12}.
We compare both approaches empirically, and the second one usually performs better.
We also stick to this choice of $(\eta_t)_{t \geq 1}$ for \Aggr.

% ----------------------------------------------------------------------
\subsection{Experiments on simulated MAB problems}\label{sub:25:numExp}

% - explain the problem studied in the simulation
We focus on \emph{i.i.d.} MAB problems, with $K = 9$ channels\footnote{~Similar behaviors are observed for any not-too-large values of $K$, we tried up-to $K = 100$ and the same results were obtained.}.
For Bernoulli problem, the first one uses $\boldsymbol{\mu}=[0.1,\dots,0.9]$,
and the second one is divided in three groups:
2 very bad arms ($\mu = 0.01, 0.02$), 5 average arms ($\mu = 0.3$ to $0.6$) and 3 very good arms ($\mu = 0.78, 0.8, 0.82$).
The horizon is set to $T = 20000$ (but its value is unknown to all algorithms), and simulations are repeated $1000$ times, to estimate the expected regret.
%
This empirical estimation of the expected regret $R_T$ is plotted below, as a function of $T$, comparing some algorithms $\Alg_1,\dots,\Alg_N$ (for $N=6$), and their aggregation with \Aggr{} (displayed in orange bold),
using the parameter-free learning rate sequence, $\eta_t := \log{N} / (t K)$.
The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted, and it is crucial to note that it is only asymptotic and to not be surprised by having regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem).
%
Note that for each of the $1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{1\leq k \leq K, 1 \leq t \leq T}$ for every repetition, in order to compare the algorithms on the same realizations%
\footnote{Similar plots and similar results are obtained if this trick is not used, but it makes more sense to compare them against the same randomization.}
of the MAB problem.

We compare our \Aggr{} algorithm,
as well as other aggregation algorithms, \ExpQ{},
\CORRAL{} and \LearnExp{} (both with default parameters) \cite{Bubeck12,Agarwal16,Singla17}.
The aggregated algorithms consist in a naive uniform exploration (to have at least one algorithm with bad performances, \ie{} linear regret, but it is not included in the plots),
\UCB{} with $\alpha=1/2$, three \klUCB{} with Bernoulli, Gaussian and exponential $\kl$ functions, and BayesUCB and Thompson sampling with uniform prior.

Figures~\ref{fig:25:EasyBernoulli} and \ref{fig:25:HarderMixed} are in semilog-$y$ scale, this helps to see that the best algorithms can be an \emph{order of magnitude} more efficient than the worst, and the \Aggr{} performs similarly to the best ones, when the other aggregation algorithms are usually amongst the worst.
Figure~\ref{fig:25:HarderMixed_semilogx} is in semilog-$x$ scale to show that the regret of efficient algorithms are indeed logarithmic.

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
	\caption{On a ``simple'' Bernoulli problem (semilog-$y$ scale).}
	\label{fig:25:EasyBernoulli}
\end{figure}

\begin{figure}[b!]  % [htbp]
	\centering
	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
	\caption{On a ``harder'' Bernoulli problem, they all have similar performances, except \LearnExp.}
	\label{fig:25:HardBernoulli}
\end{figure}

\begin{figure}[b!]  % [htbp]
	\centering
	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
	\caption{On an ``easy'' Gaussian problem, only \Aggr{} shows reasonable performances, thanks to BayesUCB and Thompson sampling.}
	\label{fig:25:EasyGaussian}
\end{figure}

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
	\caption{On a harder problem, mixing Bernoulli, Gaussian, Exponential arms, with 3 arms of each types with the \emph{same mean}.}
	\label{fig:25:HarderMixed}
\end{figure}

\begin{figure}[b!]  % [htbp]
	\centering
	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
	\caption{The semilog-$x$ scale clearly shows the logarithmic growth of the regret for the best algorithms and our proposal \Aggr, even in a hard ``mixed'' problem (\emph{cf}. Figure~\ref{fig:25:HarderMixed}).}
	\label{fig:25:HarderMixed_semilogx}
\end{figure}

For Bernoulli problems (Figures~\ref{fig:25:EasyBernoulli} and \ref{fig:25:HardBernoulli}), UCB with $\alpha=1/2$, Thompson sampling, BayesUCB and \klUCB{}$^+$ (with the binary $\kl$ function) all perform similarly, and \Aggr{} is found to be as efficient as all of them.
For Gaussian and exponential arms, rewards are truncated into $[0,1]$, and the variance of Gaussian distributions is fixed to $\sigma^2 = 0.05$ for all arms, and can be known to the algorithms (the \kl{} function is adapted to this one-dimensional exponential family).
%
Figure~\ref{fig:25:EasyGaussian} uses only Gaussian arms, with a large gap between their means and a relatively small variance, giving an ``easy'' problem.
%
And Figure~\ref{fig:25:HarderMixed} shows a considerably harder ``mixed'' problem, when the distributions are no longer in the same one-dimensional exponential family and so the Lai \& Robbins' lower-bound no longer holds (even if there still exists a lower-bound).

For each of the 4 problems considered, the \Aggr{} algorithm with default option (broadcast loss to all players) is the best of all the aggregation algorithms,
and its regret is very close to the best of the aggregated algorithms.
Especially in difficult problems with mixed or unknown distributions, \Aggr{} showed to be more efficient that \ExpQ{} and orders of magnitude better than the other reference aggregation algorithms \LearnExp{} and \CORRAL{} (see Figures~\ref{fig:25:HarderMixed} and \ref{fig:25:HarderMixed_semilogx}).

% ----------------------------------------------------------------------
\subsection{No theoretical guarantees yet}\label{sub:25:theory}

The \Aggr{} does not have satisfying theoretical guarantees in terms of regret $R_T$ yet, unlike many bandit algorithms.
%
Another notion, the \emph{adversarial regret}, denoted by $\overline{R_T}$,
measures the difference in terms of rewards,
between the aggregation algorithm $\Alg_{\mathrm{aggr}}$ and the best aggregated algorithm $\Alg_a$. This is in contrast with the (classical) regret, which measure the difference with the best fixed-arm strategy (Eq.~\eqref{eq:classicalregret}). Thus,  even if the aggregated algorithms have logarithmic (classical) regret, having an adversarial regret scaling as $\sqrt{T}$ does not permit to exhibit a logarithmic (classical) regret for the aggregation algorithm.
%
%
Under some additional hypotheses,
\cite[Theorem 4.2]{Bubeck12} proves that
\ExpQ{} satisfies a bound on adversarial regret, % $\overline{R_T}$
$\overline{R_T} \leq 2 \sqrt{T N \log{k}}$,
with the good choice of the learning rate sequence $(\eta_t)_{t \geq 1}$.
Our proposed algorithm follows quite closely the architecture of \ExpQ,
and a similar bound for \Aggr{} is expected to hold.
% this is left as a future work.
% For sake of conciseness, we cannot present a proof here.
%

This would be a first theoretical guarantee, but not satisfactory as simple algorithms like UCB have regrets scaling as $\log{T}$ \cite{Auer02,Bubeck12}, not $\sqrt{T}$.
%
Regret bounds in several different settings are proved for the \CORRAL{} algorithm \cite{Agarwal16}, but no logarithmic upper-bound can be obtained from their technique, even in the simplest setting of stochastic bandits.
%
However, \Aggr{} always seems to have a (finite-horizon) logarithmic regret in all the experiments we performed,
for both Bernoulli and non-Bernoulli problems (\eg, Gaussian, exponential and Poisson distributions).
Further theoretical developments are left as future work.

% ----------------------------------------------------------------------
\subsection{Conclusion about our \Aggr{} algorithm}\label{sub:25:conclusion}

We presented the use of aggregation algorithms in the context of Opportunistic Spectrum Access for Cognitive Radio,
especially for the real-world setting of unknown problem instances,
when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
% - \ExpQ and \Aggr works fine
% Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
Our proposed \Aggr{} was presented in details,
and we also highlighted its differences with \ExpQ.
% as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.

% - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
We realized experiments on simple MAB problems already used in the community of bandit algorithms for OSA \cite{Jouini10},
and the simulations results showed that \Aggr{} works as expected, being able to identify on the fly the best algorithm to trust for a specific problem.
Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
It is not presented in this Section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
and appears to be more robust than simple non-aggregated algorithms.

\ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proved.
Empirically, \Aggr{} showed to always have a logarithmic
regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.
%
What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.

Our \Aggr{} algorithm can probably be rewritten as an Online Mirror Descent, as \ExpQ{} and \CORRAL,
but this does not appear useful as in the case of \CORRAL{}  the analysis cannot bring a logarithmic bound on the regret even by aggregating asymptotically optimal algorithms.
We will continue investigating regret bounds for \Aggr,
and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo17}).

\paragraph{A note on the simulation code}
The page \texttt{SMPyBandits.GitHub.io/Aggregation.html} explains how to reproduce the experiments used for this Section.

% ----------------------------------------------------------------------------
\bibliographystyle{ieeetr}
\bibliography{IEEE_WCNC__2018__Paper__Lilian_Besson__07-17}

### 2019-02-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Added sub tex files, I will remove them soon and reorganize all this, but for now...

% - ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419

As explained before, setting up the future Internet of Things (IoT) networks will require to support more and more communicating devices.
In this Section, we prove that intelligent devices in unlicensed bands can use Multi-Armed Bandit (MAB) learning algorithms to improve resource exploitation.
%
We evaluate the performance of two classical MAB learning algorithms, \UCB{} and Thomson Sampling, to handle the decentralized decision-making of Spectrum Access, applied to IoT networks; as well as learning performance with a growing number of intelligent end-devices.
%
We show that using learning algorithms does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically changing channel.
In the studied scenario, stochastic MAB learning provides a up to $16\%$ gain in term of successful transmission probabilities, and has near optimal performance even in non-stationary and non-\iid{} settings with a majority of intelligent devices.

\TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}

\TODOL{TODO:

- Clean up notations, be sure it is OK with previous Sections

- Be sure to not repeat any motivations / maths / explanations given before

- Be sure to give enough details about the conclusions etc
}

% ------------------------------------------------------------------------
\subsection{Motivation for this first IoT model}

Unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.

Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
As end-devices battery life is a key constraint of IoT networks,
this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
%
In this Section, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
We consider the Upper-Confidence Bound (\UCB) \cite{Auer}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
Kaufmann12}.
% Both algorithms have already been applied with success, for Opportunistic Spectrum Access \cite{Jouini} and recently for multi-users Cognitive Radio problems \cite{Toldov}.

MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini}.
For example, \cite{Jouini} presents the OSA setting, and \cite{Maghsudi} shows how MAB learning can be applied for small cell management in licensed 5G networks.
Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi}.
Many recent works show that MAB algorithms work well for real-world radio signal.
However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \emph{i.i.d} (independent and identically distributed).
In recent works like \cite{Bonnefoi}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d}.
As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.

The aim of this Section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases, and the stochastic hypothesis is more and more questionable.
To do that, we suppose an IoT network made of two types of devices: static devices that use only one channel (fixed in time), and dynamic devices that can choose the channel for each of their transmissions. Static devices form an interfering traffic, which could have been generated by devices using other standards as well.
We first evaluate the probability of collision if dynamic devices randomly select channels (naive approach), and if a centralized controller optimally distribute them in channels (ideal approach).
Then, these reference scenarios allow to evaluate the performance of \UCB{} and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.

The rest of this Section is organized as follows. The system model is introduced in Section~\ref{sub:41:systemModel}. Reference policies are described in \ref{sub:41:threeReferencePolicies}, and sequential learning policies are introduced in \ref{sub:41:sequentialPolicies}.
Numerical results are presented in \ref{sub:41:numericalResults}.

% ------------------------------------------------------------------------
\subsection{System model and notations}\label{sub:41:systemModel}

As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol.
All devices share a synchronized time, and know in advance the finite number of available RF channels.
In each time slot, devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (no sensing).
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station. If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part.
If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \emph{Ack} is not transmitted.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.28]{2-Chapters/4-Chapter/CrownCom_17.git/protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets. If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

There are two types of end-devices in the network:
\begin{itemize}
    \item
    \emph{Static} end-devices have poor RF abilities, and each of them uses only one channel to communicate with the base station. Their choice is assumed to be fixed in time (stationary) and independent (\iid). The traffic generated by these devices is considered as an interfering traffic for other devices.
    \item
    \emph{Dynamic} (or \emph{smart}) end-devices have richer RF abilities, they can use all the available channels, by quickly reconfiguring their RF transceiver on the fly. They can also store communication successes or failures they experienced in each channel, in order to change channel, possibly at every time slot.
\end{itemize}

There are $N_c \geq 1$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
Furthermore, in channel $i \in \llbracket 1; N_c \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{N_c} S_i$).
We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $N_c$ (about $1000$ to $10000$, while $N_c$ is about $10$ to $50$).
As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq N_c$, as the former makes more sense for IoT networks\footnote{Here, $N_c$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.

We suppose that all devices follow the same emission pattern, being fixed in time, and we choose to model it as a simple Bernoulli process:
all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $N_c / (S + D)$ for all devices to communicate successfully (in average).}.
The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed (\ie, real time during two messages), and $1/p$ is proportional to the \emph{duty cycle}.
For instance for IoT objects, a ``smart'' refrigerator could have to send a daily message, and if a message can be sent every second, then $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.

The goal is to design a simple sequential algorithm, to be applied identically by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.

Before explaining how this goal presents similarity with a \emph{multi-armed bandit problem}, we present some natural baseline policies (\ie, algorithms).

% ------------------------------------------------------------------------
\subsection{Three reference policies}\label{sub:41:threeReferencePolicies}

This section presents three different policies that will be used to assess the efficiency of the learning algorithms presented in the next section.
The first one is naive but can be used in practice, while the two others are very efficient but require full knowledge on the system (\ie, an oracle) and are thus unpractical.

\paragraph{Naive policy: Random Channel Selection}

We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; N_c \rrbracket = \{1, \dots, N_c\}$).
% every time they communicate
% This reflects a very naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.

In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $i$, at that time slot.
The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$ :
\begin{equation}
    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{N_c} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/N_c, \text{by uniform choice}}
\end{equation}

All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $i$ for any dynamic device is $p / N_c$, and there are $D-1$ other dynamic devices.
As they are independent, the probability that no other dynamic device sent in $i$
% $\Pr(\text{no other dynamic device sent in}\;i)$,
is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device}\;k\;\text{did not sent in}\;i) = \prod_{k=1}^{D-1} \Pr(\text{device}\;k\;\text{did not sent in}\;i)$. And $\Pr(\text{device}\;k\;\text{sent in}\;i) = p \times 1 / N_c$, by uniform choice on channels and the Bernoulli emission hypothesis. So $q = \prod_{k=1}^{D-1} (1 - p/N_c) = (1-p/N_c)^{D-1}$. Thus we can conclude,
%
\begin{align}\label{eq:41:strategynaive}
    \Pr(\text{success}|\text{sent})
    & = \sum_{i=1}^{N_c} \underbrace{(1 - p / N_c)^{D-1}}_{\text{No other dynamic device}} \times \underbrace{(1-p)^{S_i}}_{\text{No static device}} \times\; \frac{1}{N_c} \nonumber \\
    & = \frac{1}{N_c} \left(1-\frac{p}{N_c}\right)^{D-1} \sum_{i=1}^{N_c} (1-p)^{S_i} .
\end{align}
This expression \eqref{eq:41:strategynaive} is constant (in time), and easy to compute numerically, but comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it.

\paragraph{(Unachievable) Optimal oracle policy}

We investigate in this section the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.

If the oracle draws once uniformly at random a configuration of dynamic devices, with $D_i$ devices affected to channel $i$ is fixed (in time, \ie, stationary),
then this probability is computed as before:
\begin{align}\label{eq:41:prob_col}
    \Pr(\text{success}|\text{sent})
    & = \sum_{i=1}^{N_c} \Pr(\text{success}|\text{sent in channel}\;i) \; \Pr(\text{sent in channel}\;i) \nonumber \\
    & = \sum_{i=1}^{N_c} \underbrace{(1 - p)^{D_i - 1}}_{\;\;D_i - 1 \;\text{others}\;\;} \times \underbrace{(1 - p)^{S_i}}_{\;\;\text{No static device}\;\;} \times \underbrace{ D_i / D }_{\;\;\text{Sent in channel}\; i\;\;}.
\end{align}

Consequently, the optimal allocation vector $(D_1,\dots,D_{N_c})$ is the solution of the following real-valued constraint optimization problem :
\begin{subequations}
\label{eq:41:prob}
\begin{align}
\underset{D_1,\dots,D_{N_c}}{\arg\max}\; & \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
\text{such that}\;\; & \sum_{i=1}^{N_c} D_i = D, \label{eq:41:eqCstr}\\
& D_i \geq 0 \qquad \forall i\in\llbracket 1;N_c\rrbracket . \label{eq:41:ineqCstr}
\end{align}
\end{subequations}

\begin{proposition}\label{prop:41:Lagrangian}
The \emph{Lagrange multipliers} method \cite{BoydVanderberghe04} can be used to solve the constraint real-valued maximization problem introduced in equation \eqref{eq:41:prob}.

It gives a closed form expression for the optimal solution $D_i^*(\lambda)$, depending on the system parameters, and the unknown Lagrange multiplier $\lambda \in \mathbb{R}$.

\begin{equation}\label{eq:41:Dilambda}
D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{\dag} .
\end{equation}
\end{proposition}
\begin{proof}
\begin{itemize}
    \item
    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;N_c \rrbracket$. For such values for $D_i$, the objective function $f: (D_1, \dots, D_{N_c}) \mapsto \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{N_c}$ \cite{Luenberger68,Yaari77}.}.
        \item
The Lagrange multipliers method can be applied to the optimization problem \eqref{eq:41:optPb}, with a concave objective function $f$, linear equality constraints \eqref{eq:41:eqCstr} and linear inequality constraints \eqref{eq:41:ineqCstr}. The strong duality condition is satisfied in this case \cite{BoydVanderberghe04}, so finding the saddle points will be enough to find the maximizers.
    \hfill{}$\square$
    \end{itemize}
\end{proof}

Where $(a)^{\dag} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto xe^x$ on $\mathbb{R^+} = [0, +\infty)$ \cite{Corless96}.
Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint
\begin{equation}\label{eq:41:constraintLambda}
 \sum_{i=1}^{N_c} D_i^*(\lambda) = D.
\end{equation}

Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded to find the optimal number of devices for channel $i$ :
% Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < N_c$, and $\widehat{D_{N_c}} = D - \sum_{i=1}^{N_c - 1} \widehat{D_i}$.

\paragraph{A greedy approach of the oracle strategy}

We propose a \emph{sequential} approximation of the optimal policy:
the third solution is a sub-optimal naive policy, simple to set up, but also unpractical as it also needs an oracle.
End-devices are iteratively inserted in the channels with the lowest load (\ie, the index $i$ minimizing $S_i + D_i(\tau)$ at global time step $\tau$). Once the number of devices in each channel is computed, the probability of sending successfully a message is also given by equation \eqref{eq:41:prob_col}.
This is the policy that would be used by dynamic devices if they were inserted one after the other, and if they had a perfect knowledge of the channel loads.

\subsection{Sequential policies based on bandit algorithms}\label{sub:41:sequentialPolicies}

% We now present the stochastic Multi-Armed Bandit (MAB) model,
% and the two stochastic MAB algorithms used in our experiments \cite{bubeck2012regret}.
While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused by the learning policy used by dynamic objects.
%Quite surprisingly, we will show in our experiments of Section~\ref{sub:41:numericalResults} that two algorithms inspired by the stochastic bandit modeling, UCB1 and Thompson Sampling, perform well even for coordinating multiple smart devices.

%This section presents the formalism of the Multi-Armed Bandit (MAB) problem, as introduced by \cite{LaiRobbins85}, and two classical algorithms designed to tackle its simplest version (stochastic MAB), \UCB{} from \cite{Auer} and Thompson Sampling from \cite{Thompson33,Kaufmann12}.
%For more details, \cite{bubeck2012regret} is an excellent review.

% \paragraph{Stochastic Multi-Armed Bandits}

% A Multi-Armed Bandit problem is defined as follows \cite{Thompson33,Robbins52,LaiRobbins85}.
% There is a fixed number $N_c \geq 1$ of levers, or ``arms'', and a player has to choose one lever at each discrete time $t \geq 1, t \in \mathbb{N}$, denoted as $A(t) = k \in\{1,\dots,N_c\}$.
% Selecting arm $k$ at time $t$ yields a (random) \emph{reward}, $r_k(t) \in \mathbb{R}$, and the goal of the player is to maximize the sum of his rewards, $r_{1 \dots T} = \sum_{t = 1}^T r_{A(t)}(t)$.

% A well-studied version of this problem is the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be independent and identically distributed (\emph{i.i.d}) under some distribution $\nu_k$, that has a mean $\mu_k$. Several types of reward distributions have been considered, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
% We consider Bernoulli bandit models, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k$.

% The problem parameters $\mu_1,\dots,\mu_K$ are unknown to the player, so to maximize his cumulated rewards, he must learn the distributions of the channels, to be able to progressively focus on the best arm (\ie, the arm with largest mean).
% %
% This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.

% In a Cognitive Radio application, arms model the \emph{channels}, and players are the \emph{dynamic end-devices}.
% For example in the classical OSA setting with sensing \cite{Jouini}, a single dynamic device (a player) sequentially tries to access channels (the arms), and collects a reward of 1 if the channel is available %(as there is no primary user, we can assume the transmission to be successful),
% and 0 otherwise. So rewards represent the \emph{availability} of channels, and
% the parameter $\mu_k$ represents the mean availability of channel $k$.

In our model, every dynamic device implements its own learning algorithm, \emph{independently}.
For one device, the time $t$ is the number of time it accessed the network (following its Bernoulli transmission process, \ie, its duty cycle), \emph{not} the total number of time slots from the beginning, as rewards are only obtained after a transmission, and IoT objects only transmit sporadically, due to low transmission duty cycles.

\paragraph{A bandit model for IoT}
%
Our IoT application is challenging in that there are \emph{multiple} players (the dynamic devices) interacting with the \emph{same} arms (the channels), without any centralized communication (they do not even know the total number of dynamic devices).

Considered alone, each dynamic device implements a learning algorithm to play a bandit game, the device is consequently a smart device. In each time slot, if it has to communicate (which happens with probability $p$), then it chooses a channel and it receives a reward $1$ if the transmission is successful, $0$ otherwise.
Each device aims at maximizing the sum of the rewards collected during its communication instants, which shall indeed maximize the fraction of successful transmissions. Besides the modified time scale (rewards are no longer collected at every time step), this looks like a bandit problem.
However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \emph{i.i.d}: they not only depend on the (stationary, \emph{i.i.d}) behavior of the static devices, but also on the behavior of other smart devices, that is not stationary (because of learning).

Despite this, we show in the next subsection that running a stochastic bandit algorithm for each device based on its own rewards is surprisingly successful.

\paragraph{Multi-Player MAB with \emph{collision avoidance}?}
%
Another idea could be to try to use a \emph{multi-player MAB} model, as proposed by \cite{Zhao10}, to describe our problem.
We study this other direction in the next Chapter~\ref{chapter:5}.

In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices : they are \iid{} with mean $\mu_i = 1 - p S_i$.
Moreover, dynamic devices are assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.

%The stochastic MAB model fits well for OSA for Cognitive Radio,
%but for IoT networks, a notable difference with the MAB problem introduced above is that there is not only one device adaptively using the network, but there are $D \geq 1$ dynamic end-devices (and possibly only dynamic devices).
%So in our model of IoT network, the stochastic \iid{} hypothesis for the rewards makes less sense: the (stationary) background traffic interferes with the end-devices transmissions but also the end-devices interfere with each others. To the stochastic static traffic is added a ``game-theoretical'' interaction between dynamic devices (who, let us remind it, cannot communicate with each others).
%Intuitively, the more dynamic devices, the less stationary the network will be.
%The presence of a lot of dynamic devices implies that each should learn both the (stationary) distribution of static devices (\ie, the $\mu_k$), and learn to avoid collisions with other devices.

\paragraph{\emph{Adversarial} bandit algorithms?}
% \todo[inline]{Explain that we also tried adversarial algorithms ($\mathrm{Exp}3$), and it was working but less efficiently than \UCB{} and Thompson Sampling. Maybe add it on some curve?}
Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
The \emph{adversarial MAB} algorithms is a broader family, and a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{bubeck2012regret}.
Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform worse than both \UCB{} and TS in the same experiments.
%(see Section \ref{sub:41:numericalResults}).
Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary and non-\emph{i.i.d}, as its performance guarantee are true in \emph{any} setting.
But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
%\footnote{We preferred not to talk about \emph{regret} in this study, but in a few words: it is a measure of how \emph{bad} the algorithm performed in terms of its accumulated rewards, in comparison to the best possible policy, and should be as small as possible. For stochastic algorithms, being ``efficient'' means having a regret bounded as $R_T = \mathcal{O}(\log T)$, but for adversarial algorithms, it means having $R_T = \mathcal{O}(\sqrt{K T})$.}
worse than the one for stochastic ones.
% (in their respective case of application).
More is left on this aspect for our future work.

% ------------------------------------------------------------------------
\subsection{Experiments and numerical results}\label{sub:41:numericalResults}

\begin{figure*}[!t]
    \centering
    \subfloat[10\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/10intelligent.eps}
    \label{fig:41:10intelligent}}
    \hfill
    \subfloat[30\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/30intelligent.eps}
    \label{fig:41:30intelligent}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/50intelligent.eps}
    \label{fig:41:50intelligent}}
    \hfill
    \subfloat[100\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/100intelligent.eps}
    \label{fig:41:100intelligent}}
    \caption{Performance of $2$ MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of smart end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
    \label{fig:41:from10to100}
    \vspace*{-10pt}
\end{figure*}

% Simulation parameters
We suppose a network with $S + D = 2000$ end-devices, and one IoT base station.
Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\emph{e.g.}, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
The RF band is divided in $N_c = 10$ channels.
Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{N_c}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases. The dynamic devices have access to all the channels, and use learning algorithms.
%to find the least loaded.
%
We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
We tried similar experiments with other values for $N_c$ and this repartition vector, and results were similar for non-homogeneous repartitions. Clearly, the problem is less interesting for homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Fig.\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.

Figure \ref{fig:41:from10to100} presents the evolution of the successful transmission rate, as a function of time. %($x$ axis is labeled by hundreds of transmissions). % I will change the x axis
The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy from below, and the two oracle policies (optimal and greedy) from above.
The results are displayed when $10$, $30$, $50$ and $100\%$ of the traffic is generated by dynamic devices.

We can see in Figure \ref{fig:41:from10to100} that the TS algorithm (in red) outperforms the \UCB{} algorithm (in blue), when the number of end-devices is below 50\%. When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after very few transmissions (quick convergence).
Moreover, we can see in Figures \ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the oracle optimal and oracle suboptimal policies.
For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
Increasing the number of end-devices decreases the gap between the optimal and random policies: the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures \ref{fig:41:100intelligent} and \ref{fig:41:perf_learning}.

To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure \ref{fig:41:perf_learning} the evolution of the gain, in term of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each object.
We can see that when the proportion of end-devices is low (\emph{e.g.}, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a lowest rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$).

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.65]{2-Chapters/4-Chapter/CrownCom_17.git/perf_learning.eps}
    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices.}
    \label{fig:41:perf_learning}
\end{figure}

Figure \ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
Note that when this proportion increases, the assumptions of a stochastic MAB model are clearly violated, and there is no justification for the efficiency of TS and \UCB{} algorithms.
Hence, it is surprising to have near optimal performance with stochastic MAB algorithms applied to partly dynamic and fully dynamic scenarios.

\paragraph{Another simulation results}

We include another simulation, with a uniform repartition of static device, to check that learning does not help in this case, but does not decrease performance either.
% And we should include another example with a different scenarios, maybe $(0.3, 0.3, 0.3, 0.015, 0.015, 0.014, 0.014, 0.014, 0.014, 0.014)$ (R\'emi already did the simulations).

\begin{figure*}[!t]
    \centering
    \subfloat[10\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_10.eps}
    \label{fig:41:ch2_10}}
    \hfill
    \subfloat[30\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_30.eps}
    \label{fig:41:ch2_30}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_50.eps}
    \label{fig:41:ch2_50}}
    \hfill
    \subfloat[100\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_100.eps}
    \label{fig:41:ch2_100}}
    \caption{Performance of $2$ bandit algorithms, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
    \label{fig:41:figure4appendix}
\end{figure*}

\subsection{Conclusion about this first model}\label{sub:41:conclusion}

In this Section, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the many ``intelligent'' end-devices.

This is both a surprising and a very encouraging result, showing that application of bandit algorithms tailored for a stochastic model is still useful in broader settings.
The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{can not} communicate with each others, and \emph{do not} know the system parameters.
We will investigate this behavior in order to understand it better theoretically.
We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.

Moreover, for sake of simplicity we supposed that all devices use the same standard. Our future work will consider more realistic interference scenarios and IoT networks, with, \emph{e.g.}, non-slotted time, more than one base station etc.

\paragraph{Possible extensions of this model}
%
This work could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.

Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \emph{e.g.}, one transmission a day.
So additionally to deciding the channel for communication, each device has to also decide when to communicate.
This is another direction of research, that we will investigate in the future.

\paragraph{Note on the simulation code}
%
The simulation code used for the experiments in Section~\ref{sub:41:numericalResults} is for MATLAB or GNU Octave,
and is open-sourced under the MIT License, at\\
\href{https://Bitbucket.org/scee_ietr/rl_slotted_iot_networks}{\texttt{Bitbucket.org/scee\_ietr/rl\_slotted\_iot\_networks}}.

% ---------- Bibliography ----------
%\begin{small}
\bibliographystyle{ieeetr}
\bibliography{biblio_RIoT}
%\end{small}

### 2019-02-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Added sub tex files, I will remove them soon and reorganize all this, but for now...

\subsection{A missing proof}

This short subsection contains the missing details of the proof of Proposition \ref{prop:41:Lagrangian}.
%
First, we need to justify that the objective function is quasi-concave, in each of its coordinates.
%
Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($N_c, p$), the repartition of static devices ($S_1,\dots,S_{N_c}$) and the Lagrange multiplier $\lambda$.

\paragraph{Quasi-convexity}

\begin{itemize}
	\item
	For $0 < \gamma < 1$, the function: $g: x \mapsto x \gamma^x$ is quasi-convex on $[0,\infty)$, \ie, $g(\eta x + (1-\eta)y) \leq \max(g(x), g(y))$ for any $x,y \in [0,\infty)$ and $\eta \in [0,1]$ (definition from \cite{Luenberger68}).
    Indeed, $g(\eta x + (1-\eta)y) = \eta \left[ x (\gamma^x)^{\eta} \right] \gamma^{((1-\eta)y)} + (1-\eta)\left[ y (\gamma^y)^{1-\eta}\right] \gamma^{\eta x}$, and $\gamma^{((1-\eta)y)} \leq 1$ and $\gamma^{\eta x} \leq 1$. But also $(\gamma^x)^{\eta} \leq \gamma^x$ as $\eta \leq 1$, and the same holds for $(\gamma^y)^{1-\eta} \leq \gamma^y$. So $g(\eta x + (1-\eta)y) \leq \eta (x \gamma^x) + (1 - \eta) (y \gamma^y)$ which is a convex combination of $x \gamma^x$ and $y \gamma^y$, so smaller than the larger of the two values, and so $g(\eta x + (1-\eta)y) \leq \max(x \gamma^x, y \gamma^y)$.

    \item
    The function $f: (D_1, \dots, D_{N_c}) \mapsto \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{N_c}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
    \hfill{}$\square$
\end{itemize}

\paragraph{Derivation of the Lagrange multiplier solution}

Let us now prove the expression for $D_i^*$ given in \eqref{eq:41:Dilambda}.
\begin{itemize}
	\item
	If $\boldsymbol{D} = (D_1,\dots,D_{N_c})$,
    the Lagrangian is denoted $\mathcal{L}(\boldsymbol{D}, \lambda) = f(\boldsymbol{D}) + \lambda(D - \sum_{i=1}^{N_c} D_i)$, and its derivative w.r.t. $D_i$ is
    $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) = (1-p)^{S_i + D_i - 1} + \log(1 - p) D_i (1-p)^{S_i + D_i - 1} - \lambda$.

    \item
    So the gradient is canceling $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) |_{D_i=D_i^*} = 0$
    iff $D_i^*$ satisfies $(1-p)^{D_i^*}( 1 + \log(1-p) D_i^*) = \lambda / (1-p)^{S_i - 1}$. Let $x = \log(1-p) D_i^*$ this is equivalent to $\mathrm{e}^{x}(1 + x) = \lambda / (1-p)^{S_i - 1}$
    and with $y = 1 + x$, we get $\mathrm{e}^{y}y = \lambda\mathrm{e} / (1-p)^{S_i - 1}$.

	\item
    By using the $\mathcal{W}$-Lambert function $\mathcal{W}$ \cite{Corless96}, reciprocal of $y \mapsto \mathrm{e}^{y}y$, we get $x = y - 1 = \mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1$.
    So the gradient of the Lagrangian is canceling iff $D_i^* = \max(0, 1/\log(1-p) x) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1\right]^{\dagger}$, because $D_i^*$ has to be non-negative.
    This gives the $i$-th coordinate of the unique saddle point of $f(\boldsymbol{D})$,
    and so the unique solution to the maximization problem \eqref{eq:41:optPb}, thanks to \cite[Theorem 1]{Luenberger68}.
    \hfill{}$\square$
\end{itemize}

### 2019-02-12 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Added sub tex files, I will remove them soon and reorganize all this, but for now...

\definecolor{darkblue}{rgb}{0.00,0.00,0.65}       % rgb(0,0,126)
\definecolor{darkgreen}{RGB}{0,100,0}  % rgb(0,100,0)
\definecolor{darkred}{RGB}{174,0,0}        % rgb(174,0,0)

% - ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825

In this Section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$, implementing a proof-of-concept of the model previously presented in Section~\ref{sec:4:firstModel}.

We implement an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) objects, embedding the proposed solution,
and a traffic generator that emulates radio interference from many other objects.
Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol, which does not require any specific overhead for the learning.
%
We model the network access as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
%
This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature in some or all the devices, without any modification on the network side.

\TODOL{Based on the publication, ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825}

\TODOL{TODO:

- Clean up notations, be sure it is OK with previous Sections

- Be sure to not repeat any motivations / maths / explanations given before

- Be sure to give enough details about the conclusions etc
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Motivation for this demonstration}

% The monitoring of large scale systems, such as smart grids and smart cities, requires the development of networks dedicated to Internet-of-Things (IoT) applications.
% For instance, Low Power Wide Area Networks (LPWAN) \cite{Raza17}, as LoRaWAN or SigFox, are nowadays deployed in unlicensed bands to handle a large number of objects transmitting a few packets per day or week.
% In order to reduce the energy consumption of end-devices, these networks rely on pure ALOHA-based Medium Access (MAC) protocols.

% One of the challenges in the design of MAC solutions for the IoT is to design solutions which improve the performance of the network and reduce the Packet Loss Ratio (PLR), without reducing the end-devices battery life.
% In particular, many IoT standards operate in unlicensed bands, that is why we have to find solutions that do not increase the PLR due to the interference caused by other standards and networks which share the same band, without coordination.
% As this interfering traffic is generated by different standards and networks, it cannot be controlled, and it is not evenly distributed in the different channels.

Multi-Armed Bandit (MAB) algorithms have been recently proposed as a solution to improve the performance of IoT networks and in particular in LPWAN \cite{Bonnefoi18,Azari18}.
In this paper, we describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol (such as the ones employed in LPWAN).
%
%\todo[inline]{Contributions et interets de la demo}
% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini}.
% Recently, TS and UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov}, and the UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi}.
% Many recent works show that MAB algorithms work well for real-world radio signal.
% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other objects is never really stationary or \emph{i.i.d.} (independent and identically distributed).
% In recent works like \cite{Bonnefoi}, several objects are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d.}.
% As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent objects in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
%
This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
% % even when the number of intelligent objects in the network increases, and the stochastic hypothesis is more and more questionable.
% To do that, we suppose an IoT network made of two types of objects: static objects that use only one channel (fixed in time), and dynamic objects that can choose the channel for each of their transmissions. Static objects form an interfering traffic, which could have been generated by objects using other standards as well.
% We first evaluate the probability of collision if dynamic objects randomly select channels (naive approach), and if a centralized controller optimally distribute them in channels (ideal approach).
% Then, these reference scenarios allow to evaluate the performance of UCB and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
% We show that these algorithms have near-optimal performance, even when the proportion of end-objects increases and the interfering traffic from other objects becomes less and less stochastic.
% \todo[inline]{(Remi) Si on parle de non stationarity, il faut donner un resultat a deux utilisateurs.}
Following our recent work presented in the previous Section~\ref{sec:4:firstModel}, we propose to model this problem as Non-Stationary\footnote{~Note that non-stationarity only comes from the presence of more than one dynamic object, as the background traffic is assumed independent and identically distributed \emph{i.i.d.}.} MAB.
We suggest to use low-cost algorithms, focusing on two well-known algorithms: a frequentist one (\UCB{}) and a Bayesian one (TS).
We consider the Upper-Confidence Bound (\UCB{}) \cite{Auer}, and the Thompson Sampling (TS) algorithms \cite{Thompson33}. Both algorithms have already been applied with success in the context of wireless decision making, both empirically for Opportunistic Spectrum Access \cite{Jouini},
and more recently for multi-users Cognitive Radio problems \cite{BessonALT18} with a more theoretical approach.

We use a TestBed designed in 2017 by our team SCEE \cite{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.

In our demo, we consider a simple wireless network, consisting of one gateway (radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
Some dynamic intelligent objects (end-user or autonomous objects) try to communicate with the gateway, with a low-overhead protocol. This communication can be done in different channels which are also shared by devices using other networks.
Once the gateway receives a packet transmitted by a dynamic device in one channel, it transmits back to it an acknowledgement in the same channel, after a fixed-time delay, as it is done in the LoRaWAN standard.
This \emph{ACK} allows the device to learn about the channel quality and thus, to use learning algorithms for the purpose of best channel selection.

In this demo, we can generate scenarios with different parameters (number of channels, interfering traffic, etc) in order to evaluate the performance of learning in various settings.
Moreover, we compare the performance of learning with that of the random uniform access to channels, which is the current state-of-the-art of commercial LPWAN solutions.
%
This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent objects at least do not reduce their successful communication rate in comparison to the naive objects.
This also shows that in case of non-uniform stationary traffic, the MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent objects.

The rest of this paper is organized as follows. The system model is introduced in Section 2. In Section 3, we describe more formally both the \UCB{} and the TS algorithms. Our implementation is presented in Section 4, and results are given in Section 5.
% and Section 6 concludes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System model - FIXME merge with previous section!}

We consider the system model presented in Figure~\ref{fig:42:system_model1}, where a set of object sends uplink packets to the network gateway, in the $433.5\;\mathrm{MHz}$ ISM band.
The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
%
The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$). In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s).

These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks. This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.

%To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
%
%By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
%the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/system_model1.eps}
    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
    \label{fig:42:system_model1}
\end{figure}

We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
%
For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
as it is usually done in MAB problems \cite{Thompson33,Robbins52,LaiRobbins85}.

This problem is a special case of the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$. Several types of reward distributions have been considered in the literature, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).

Rewards are binary in our model, and so we consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
Contrary to many previous work done in the CR field (\emph{e.g.}, Opportunistic Spectrum Access),
the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model. Rewards come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.

The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to the end-users, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
%
This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.

% In a Cognitive Radio application, arms model the \emph{channels}, and players are the \emph{dynamic end-devices}.
% For example in the classical OSA setting with sensing \cite{Jouini}, a single dynamic device (a player) sequentially tries to access channels (the arms), and collects a reward of 1 if the channel is available %(as there is no primary user, we can assume the transmission to be successful),
% and 0 otherwise. So rewards represent the \emph{availability} of channels, and
% the parameter $\mu_k$ represents the mean availability of channel $k$.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{MAB Algorithms}

Before discussing the relevance of a MAB model for our IoT application, we refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Sec.~\ref{sec:42:results}).

% \paragraph{The \UCB{} Algorithm}\label{sub:42:UCB}

% A naive approach could be to use an empirical mean estimator of the rewards for each channel, and select the channel with highest estimated mean at each time;
% but this greedy approach is known to fail dramatically \cite{LaiRobbins85}. Indeed, with this policy, the selection of arms is highly dependent on the first draws:
% if the first transmission in one channel fails and the first one on other channels succeed, the end-user will \emph{never} use the first channel again, even it is the best one (\emph{i.e.}, the most available, in average).

% Rather than relying on the empirical mean reward, Upper Confidence Bounds algorithms
% instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm,
% which can be viewed as adding a ``bonus'' exploration to the empirical mean.
% % by adding a confidence term to the empirical mean.
% They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model,
% as the statistically best possible arm (\emph{i.e.}, the highest upper confidence bound) is selected.
% % the arm with the highest upper confidence bound is expected to be the true best arm, and is played.

% More formally, for one end-user, let $N_k(t) = \sum_{\tau=1}^t \mathbbm{1}(C(\tau) = k)$ be the number of times channel $k$ was selected up-to time $t \geq 1$.
% The empirical mean estimator of channel $k$ is defined as the mean reward obtained by selecting it up to time $t$, $\widehat{\mu_k}(t) = \left(1 / N_k(t)\right) \sum_{\tau=1}^t r_k(\tau) \mathbbm{1}(C(\tau) = k) $.
% For \UCB, the \emph{confidence} term is $B_k(t) = \sqrt{\alpha \log(t) / N_k(t)}$,
% giving the upper confidence bound $U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, which is used by the end-user to decide the channel for communicating at time step $t+1$: $C(t+1) = \arg\max_{1\leq k \leq K} U_k(t)$.
% \UCB{} is called an \emph{index policy}.

The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{bubeck2012regret}.
% This algorithm is simple to implement and use in practice, even on embedded micro-processors with limited computation and memory capabilities.
%
% \todo[inline]{Maybe in the implementation of UCB the $t$ $\log(t)$ should be the total amount of time the user tried to communicate? It is already the case, but we should say so.}
In our model, every dynamic end-user implements its own \UCB{} algorithm, \emph{independently}.
For one end-user, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.

% \paragraph{Thompson Sampling}

% Thompson Sampling \cite{Thompson33} was introduced early on, in $1933$ as the very first bandit algorithm, in the context of clinical trials (in which each arm models the efficacy of one treatment across patients). Given a prior distribution on the mean of each arm, the algorithm selects the next arm to draw based on samples from the \emph{conjugated} posterior distribution, which for Bernoulli rewards is a Beta distribution.

% A Beta prior $\mathrm{Beta}(a_k(0)=1,b_k(0)=1)$ (initially uniform) is assumed on $\mu_k \in [0, 1]$, and at time $t$ the posterior is $\mathrm{Beta}(a_k(t),b_k(t))$.
% After every channel selection, the posterior is updated to have $a_k(t)$ and $b_k(t)$ counting the number of successful and failed transmissions made on channel $k$.
% So if the \emph{ACK} message is received, $a_k(t+1) = a_k(t) + 1$, and $b_k(t+1) = b_k(t)$, otherwise $a_k(t+1) = a_k(t)$, and $b_k(t+1) = b_k(t) + 1$.
% Then, the decision is done by \emph{sampling} an \emph{index} for each arm, at each time step $t$, from the arm posteriors: $X_k(t) \sim \mathrm{Beta}(a_k(t), b_k(t))$, and the chosen channel is simply the channel $C(t+1)$ with highest index $X_k(t)$. For this reason, Thompson Sampling is called a \emph{randomized index policy}.

The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12}.
It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov,Mitton}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GNU Radio implementation}

In this section, we present our implementation of MAB algorithms in our model of IoT networks.
We first describe the simplified physical layer of this demo,
then we present our GNU Radio implementation.

\paragraph{Physical layer and protocol}

In this paper, we implement a PHY/MAC layers solution in order to improve the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
We choose to implement our own physical layer and protocol, for both clarity and conciseness.

Regarding the physical layer, we consider a QPSK constellation. Moreover, we use simplified packets composed of two parts.
The first part is the \emph{preamble} which is used for the purpose of synchronization (phase correction).
Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
For example, this index can be a simple QPSK symbol ($\pm1\pm1j$).
Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet (\emph{e.g.}, $1+j \mapsto 1-j$).
Thanks to this index, we can have several devices communicating with the same gateway.

In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
In this case, the \emph{ACK} was for him, and it knows that its packet has been received and decoded correctly by the gateway.

\paragraph{Equipment}

% Boards and versions
We use USRP N210 boards \cite{USRPDocumentation}, from Ettus Research (National Instrument),
with version $4$ of their FPGA system and version $5.1$ of the RBX system.
As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP.
The gateway, a USRP which emulates the interfering traffic, and at least one dynamic device.
% Cables and laptop

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/our-demo.eps}
    \caption{Schematic of our implementation that presents the role of each USRP card.}
    \label{fig:42:our_demo}
\end{figure}

The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
% Octoclock
To ease the synchronization in both time and frequency between the boards representing the dynamic objects and the gateway, we use an Octoclock \cite{OctoclockProduct}, also by Ettus Research,
% https://www.ettus.com/product/details/OctoClock-G
and coaxial cables connecting every card to the Octoclock for time (PPS) and frequency synchronization, but this is not mandatory.

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/SCEE_TestBed1.jpg}
    \vspace*{20pt}
    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/SCEE_TestBed2.jpg}
    \caption{Photos of the SCEE testbed.}
    \label{fig:42:photosSCEETestBed}
\end{figure}

\paragraph{Implementation}

We used GNU Radio Companion (GRC, version $3.7$, $2017$),
and for the demonstration the laptop runs
one GRC design to configure and control each USRP card.
As such, a single laptop can run in parallel the control program of any number of boards\footnote{~Even if in practice, maximum efficiency is kept as long as there is not more than one GRC design by CPU core.}.

The GNU Radio software provides the framework and tools to build and run software radio or just general signal-processing applications.
GNU Radio applications are a flow-graph: a series of signal processing blocks connected together to describe a data flow.
For maximum efficiency, we wrote all of our blocks in \texttt{C++}.
These flow-graphs can be written in either C++ or the Python programming language. The GNU Radio infrastructure is written entirely in C++, and many of the user tools are written in Python.
GNU Radio Companion is a graphical UI used to develop GNU Radio applications:
GRC is effectively a Python code-generation tool.
when a flow-graph is compiled in GRC, a Python code is produced, which can be executed to connect to the USRP,
create the desired GUI windows and widgets, and create and connect the blocks in the flow-graph.

\paragraph{User Interface}

We have designed a user interface in order to visualize the results obtained  with our experimental demonstration. This user interface is shown in Figure~\ref{fig:42:UI}.
We can see that it is made of three parts, one for each USRP, as highlighted in \textcolor{darkred}{red}:

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.80\textwidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/UI.eps}
    \caption{User interface of our demonstration.}
    \label{fig:42:UI}
\end{figure}

% \begin{enumerate}[leftmargin=6mm]
    % \item[(1)]
$(1)$ The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.

    % \item[(2)]
$(2)$ The second part is the interface of the intelligent device which is made of four parts.
At the top left, we observe the constellation of the transmitted packet \emph{(a)}.
At the bottom left, we have a time/frequency view of the lasts packets transmitted by the object \emph{(b)}.
We can see, in this view that the object transmitted its last $9$ packets in the two best channels (channel $\#3$ and $\#4$).
Then, at the top right of this interface \emph{(c)}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this object (\textcolor{darkred}{red}) and the acknowledgements sent by the gateway (\textcolor{darkblue}{blue}).
Finally, at the bottom right \emph{(d)}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).

    % \item[(3)]
$(3)$ The last part is the interface of the gateway, where we can see the traffic observed by the gateway \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
% \end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results and discussions}\label{sec:42:results}

We compare the two algorithms described in Section~\ref{sub:42:UCB} against a uniform access algorithm, that uniformly selects its channel at random.
Three objects are compared by their mean successful communication rates, on a horizon of $2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} shows the results averaged on $10$ repetitions using the same conditions.
%
Each experiment takes about half a day,
as we make objects generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
Each object was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $[15\%, 10\%, 2\%, 1\%]$.
% They did not overlap each other, as
When facing the same stationary background traffic, we see that the learning objects are both very quickly more efficient than the naive uniform object.
We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
%
In stationary environments, both the TS and \UCB{} algorithms are very efficient and converge quickly, resulting in a very strong decrease in collisions and failed communication slots. \UCB{} is faster to learn but eventually TS gives a (slightly) better average performance.

Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
Empirical results confirm the simulations presented in our paper \cite[Fig.3]{Bonnefoi17}.
Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.

\begin{figure}[!t]
	\centering
    \includegraphics[height=8.0cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success. Similar gains of performance were obtained in many different scenarios.}
    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
\end{figure}

% \begin{figure}[!t]
% 	\centering
%     \begin{subfigure}[b]{0.46\textwidth}
%         \includegraphics[height=7cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/Object2_UniformAccess__UI.png}
%         %\caption{Object using uniform access.}
%         \label{fig:42:comparing_UniformAccess_and_UCB__1}
%     \end{subfigure}
%     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
%       %(or a blank line to force the subfigure onto a new line)
%     \begin{subfigure}[b]{0.46\textwidth}
%         \includegraphics[height=7cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/Object1_UCB__UI.png}
%         %\caption{Object using UCB.}
%         \label{fig:42:comparing_UniformAccess_and_UCB__2}
%     \end{subfigure}
% \caption{Comparing the success rate of an object using uniform access (left, $41\%$ of success here) and an object using the UCB algorithm (right, already $60\%$ of success). We see the learning object being much more present in the channels $\#4$ and $\#3$ (\textcolor{darkblue}{blue chart}, bottom right corner), as these channels are the less perturbed by the random interfering traffic, in this example.}
% \label{fig:42:comparing_UniformAccess_and_UCB}
% \end{figure}

We presented in this Section a demonstration, showed in $2018$ at the ICT conference,
by specifying the system model and explaining the two MAB algorithms used in practice.
We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
%
By using such low-cost algorithms, we demonstrated empirically that a dynamically reconfigurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.

Possible future extensions of this work include:
considering more dynamic objects (\emph{e.g.}, $100$),
implementing a real-world IoT communication protocol (like the LoRaWAN standard),
and studying the interference in case of other gateways located nearby.
%
% Pub pour notre autre article
We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.

\paragraph{Availability of data and materials}
%
The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at\\
\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.

A $6$-minute \textbf{video} showing our demonstration is at \texttt{\url{youtu.be/HospLNQhcMk}}.
It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It shows the connections between the USRP boards, the Octoclock, and the master laptop.

\paragraph{Acknowledgment}
%
We acknowledge the work of two CentraleSup{\'e}lec students,
Cl{\'e}ment Barras and Th{\'e}o Vanneuville, for their GNU Radio project in Spring $2017$,
as we took inspiration in the GNU Radio code and the discussions we had at this time.

% ---------- Bibliography ----------
%\begin{small} % XXX small if needed?
\bibliographystyle{ieeetr}
\bibliography{biblio_RIoT}
%\end{small} % XXX small if needed?

### 2019-02-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Added sub tex files, I will remove them soon and reorganize all this, but for now...

% https://perso.crans.org/besson/articles/BMBM__IEEE_WCNC_2019.pdf

In this section, we extend the model presented before in order to take into account the possibility for retransmission of messages after a collision.
As before, we propose and evaluate different learning strategies based on MAB algorithms.

The presented strategies allow IoT devices to improve their access to the network and their autonomy, while taking into account the impact of encountered radio collisions.
For that end, several heuristics employing \UCB{} algorithms are examined, to explore the contextual information provided by the number of retransmissions.
Our results show that approaches based on \UCB{} obtain a significant improvement in terms of successful transmission probabilities, compared to a naive approach which does not learn.
Furthermore, it also reveals that a pure \UCB{} channel access is as efficient as more sophisticated learning strategies.

\TODOL{This section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}}

\TODOL{TODO:

- Clean up notations, be sure it is OK with previous Sections

- Be sure to not repeat any motivations / maths / explanations given before

- Be sure to give enough details about the conclusions etc
}

% ----------------------------------------------------------------------
\subsection{Retransmissions in an ALOHA like protocol}\label{sub:43:introduction}

In the context of Cognitive Radio \cite{Mitola99,Haykin05},
Multi-Arm Bandit (MAB) algorithms \cite{Auer,Auer02,Bubeck12} have been recently proposed as a potential solution for channel access in LPWA networks \cite{Bonnefoi18,Azari18,Bonnefoi17}.
We presented in Section~\ref{sec:4:firstModel} the impact of non-stationarity on the network performance using MAB algorithms is studied.
Low-cost algorithms following two well-known approaches, such as the Upper-Confidence Bound (\UCB{}) \cite{Auer,Auer02}, and the Thompson Sampling (TS) algorithms \cite{Thompson33} have reported encouraging results.
Other recent directions include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
and realistic empirical simulations \cite{kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
of the application of MAB algorithms for slotted wireless protocols in a decentralized manner,
or for instance applications to multi-hoping networks \cite{Mitton,Toldov}.
None of the above mentioned works discusses in detail the impact of retransmissions on the performance of MAB learning algorithms as we do now.

The aim of this section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
Proposed approach employs contextual information provided by the number of retransmissions, and implemented at each device, so that no coordination among them is needed.
Moreover, our \UCB{}-based heuristics show low complexity making them suitable for being embedded in LPWA devices.

The contributions of this section are summarized as follows:
\begin{itemize}
	\item
	Firstly, we provide a close form approximation of the radio collision probability after a first retransmission.
	By doing this, we highlight the need to develop a learning approach for channel selection upon collision.

	\item
	Secondly, several heuristics are proposed to cope with retransmissions.

	\item Lastly, we conduct simulations in order to compare the performance of the proposed heuristics with a naive uniform random approach, and a \UCB{} strategy (\ie, without any learning for the retransmissions).
\end{itemize}

% \vspace*{5pt}
The rest of the section is organized as follows.
First the system model is introduced in Section~\ref{sub:43:model},
% A formal description of the MAB learning algorithms is given in Section~\ref{sub:43:motivations},
and our motivations are exposed in Section~\ref{sub:43:MABalgo}.
The proposed \UCB-based heuristics are presented in Section~\ref{sub:43:heuristics}, while the corresponding numerical results are shown in Section~\ref{sub:43:numExp}.
Finally, some conclusions are drawn in Section~\ref{sub:43:conclusion}.

% ----------------------------------------------------------------------
\subsection{System model}\label{sub:43:model}

% ----------------------------------------------------------------------
\paragraph{LPWA network}

Like before in this chapter, we consider an LPWA network composed of a gateway and a large number of end-devices that regularly send short data packets, where $K$ channels ($K>1$) are available for the transmission of their packets.
%
We assume that this network is constituted by two types of devices:
on the one hand, we have \emph{static} devices that operate in one channel\footnote{~Note that, for unlicensed bands, this definition also encompasses any device following a different standard or trying to establish communication with gateways of other networks.} in order to communicate with the gateway.
%
On the other hand, there are  IoT devices, that possess the additional advantage of being able to select any of the $K$ available channels to perform their transmissions.

Regardless the type of devices, each of them follows a slotted ALOHA protocol \cite{Roberts75}, and has a probability $p>0$ to transmit a packet in a time slot.
We make the hypothesis that the transmission is successful if the channel is available, otherwise upon radio collision, these devices will attempt to retransmit their packet up-to $M$ times, with $M \in\mathbb{N}$.
Note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\llbracket 0, m-1 \rrbracket$, where $m \in\mathbb{N}, m>0$ is the length of the back-off interval.

% ----------------------------------------------------------------------
\paragraph{Model of our IoT devices}

The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Fig.~\ref{fig:43:Markov_model}.
A device containing a packet for transmission goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, \ie, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
At each time slot, a transition from an idle state to a transmission state (denoted as \texttt{Trans.}) occurs if a packet transmission is required, while waiting states (denoted as \texttt{Wait}), correspond to a $m$ back-off interval.

A device aims to select a channel with the highest probability of a successful transmission, for which it resorts to a reinforcement learning approach. It is formulated as MAB problem, where each channel (also called arms) is viewed as a gambling machine (bandit), and each bandit has a \emph{reward}. Then, at every trial, a device chooses a channel that maximizes the sum of the collected rewards. These \emph{rewards} are the \emph{acknowledgment} (\emph{Ack}) signals received after transmitting packets to the gateway. In this way, a successful transmission is considered when an acknowledgment is received, and a learning approach is employed to select the best channel.

We address the problem of channel selection taking into account the described Markov model for the retransmissions of end-devices. It motivates our present work for which we consider the retransmissions in the analysis of MAB algorithms. % It motivates our present work for which we consider the number of retransmissions, carried out by each device to address several MAB algorithms.

\begin{figure}[htp!]  % [htbp]
	\centering
	% \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Markov_model.eps}
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Markov_model-eps-converted-to.pdf}
	\caption{The considered Markov model for the behavior of all the devices in the network.}
	\label{fig:43:Markov_model}
\end{figure}
%

% ----------------------------------------------------------------------
\subsection{Motivations for the proposed approach} \label{sub:43:motivations}

When a device experiments a collision, it goes in a back-off state to retransmit the same packet on a channel.
If all devices remain in the same channel for retransmissions, it could result in a sequence of successive collisions with the same packets' devices that previously collided.
%
Thus, it seems interesting to consider in the decision making policy the possibility for a device to retransmit in a different channel.
One of our motivations to develop new MAB algorithms for our problem is this option of using a different communication channels between the first transmission and the next retransmissions.

By considering this possibility, the device will have to learn more, thus, we expect that the learning time to be longer, but it could be possible that the final performance gain (\ie, in terms of network performance) increases too.
The next Section~\ref{sub:43:numExp} presents analysis to check this performance gain, for various heuristics based on the \UCB{} algorithm.
%
Here after, we start by presenting a mathematical derivation that backups this idea.
To do so, we study the collision probabilities considering the Markov process depicted in Fig.~\ref{fig:43:Markov_model}, and foresee the impact of addressing bandit strategies, as well as setting guidelines for the design of heuristic approaches.

\paragraph{Probability of collision at a second transmission slot}

As it is well known, having a collision during an access time can be overcome by a retransmission procedure (this can take several retransmission attempts).
What interest us here, is to obtain a mathematical approximation of the collision probability at the second transmission slot $p_{c1}$, as a function of the first collision probability $p_{c}$.

We consider two hypotheses $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as,
\begin{itemize}
	\item $\mathcal{H}_{1}$:
    The probability $p_{c1}$ is composed by the sum of two probabilities: i)
    the probability of colliding consecutively twice, \ie, the devices that collide at a given time slot and collide again when retransmitting their packets,
    and ii) the probability of collision related with devices that attempt to transmit in this channel for first time.
    In addition, we suppose that the number of devices that attempt to retransmit is small in comparison to the total number of devices sharing the same channel.

	\item $\mathcal{H}_{2}$:
	The total number of the back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
\end{itemize}

Considering one device and a channel,
we denote $x_t^i$ the probability that it is transmitting a packet for the $i+1$ time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$),
and let $x_t = \sum_{i=0}^{M-1}x_t^i$ be the probability that it transmits a packet.
We consider $N$ active devices following the same policy.

We assume to be in the steady state \cite{Norris98}, in our Markov chain model depicted in Figure~\ref{fig:43:Markov_model}, and thus the probabilities no longer depend on the slot number $t$ (\ie, $\forall t, x_t=x$).
Therefore, the probability that this device has a collision at the first transmission is $p_c$, and has the following expression
\begin{equation}\label{eq:43:1}
	p_c = 1-\left(1-x_t\right)^{N-1} \iff x_t = 1-\left(1-p_c\right)^{\frac{1}{N-1}}.
\end{equation}

Moreover, from \eqref{eq:43:1} we define the probability $p_{cp}(n)$ that involves the collision of $n$ packets sent by each IoT device (for any $1\leq n \leq N-1$), during the first transmission slot, and is defined by the following equation
\begin{equation*}\label{eq:43:2}
	p_{cp}(n) = {N-1 \choose n} \; x^n \left(1-x\right)^{N-1-n}.
\end{equation*}

As explained above, if an IoT device is experiencing a collision at the first transmission, it proceeds for the retransmission of its packet after a random back-off interval.
We denote $p_{ca}$ the probability to have a collision with a packet involved in the previous collision.
Under assumption $\mathcal{H}_{1}$, the probability that the same device's packet is experiencing again a collision at the second time slot is
\begin{equation}\label{eq:43:decomppc1}
	p_{c1} = p_{ca}+\left(1-p_{ca} \right)p_c.
\end{equation}

If the device had a collision, we consider $p_{bp}(n)$ the probability that it had a collision with \emph{exactly} $n$ packets (for any $1\leq n \leq N-1$), and that \emph{at least one} of the $n$ devices involved in the previous collision chose the same back-off interval.
Hence, under hypothesis $\mathcal{H}_{2}$, we can relate $p_{ca}$ with $p_{bp}(n)$, and the different probabilities that a device experienced a collision during the first slot and has the same back-off interval for its retransmission is,
%
\begin{equation}\label{eq:43:sumpca}
	p_{ca} = \sum_{n=1}^{N-1} p_{bp}(n).
\end{equation}

Therefore, the expression of $p_{ca}$ is
\begin{equation}\label{eq:43:sumpc2}
	\frac{1}{p_c} \sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left[1-\left( 1-\frac{1}{m}\right)^n \right].
\end{equation}

Once again under $\mathcal{H}_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first terms of the sum in \eqref{eq:43:sumpc2} are predominant. Moreover, for these terms, $n$ is small compared to $N-1$, and so $N-1-n$ can be approximated to $N-1$. Thus it gives,
\begin{align}\label{eq:43:sumpca2}
	p_{ca} = & 1- \frac{1}{p_c}\sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left( 1-\frac{1}{m}\right)^n,\nonumber\\
	\simeq & 1- \frac{\left(1-x\right)^{N-1}}{p_c}\sum_{n=1}^{N-1}{N-1 \choose k} x^n \left( 1-\frac{1}{m}\right)^n.
\end{align}

We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca2}, and so the expression of $p_{ca}$ becomes
\begin{align}\label{eq:43:pca}
	p_{ca} \simeq \frac{1}{p_c} - \left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1}.
\end{align}

Finally, our approximation of $p_{c1}$ can be obtained by inserting \eqref{eq:43:pca} in \eqref{eq:43:decomppc1}.

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Behavior analysis of $p_{c}$ and $p_{c1}$}\label{sub:numericalValidationPC1PC}

In order to assess the proposed approximation, we suppose a unique channel where all the devices follow the same contention Markov process.
We simulate an ALOHA protocol with a maximum number of retransmissions $M=10$, a maximum back-off interval $m=10$, and a transmission probability $p=10^{-3}$.
%
In Fig.~\ref{fig:43:Approximation_m10}, we show the collision probabilities for different number of devices $N$ (from $N=50$ up-to $N=400$), for both $p_{c}$ and $p_{c1}$.
%
From this simulations, we can verify that our approximation is very precise for lower values $p_{c1} \leq 30 \%$ (\ie, red and orange curves are quite close).
Moreover, a significant gap between $p_{c1}$ and $p_c$,
of up-to $10\%$, can be observed,
which suggests us to resort to MAB algorithms for the channel selection for both the first transmission and next retransmissions.

\begin{figure}[htp!]  % [htbp]
	\centering
	% \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Approximation_m10.eps}
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Approximation_m10-eps-converted-to.pdf}
	\caption{Our proposed approximation for the probability of collision at the second transmission. It is more precise for smaller values of $N$.}
	\label{fig:43:Approximation_m10}
\end{figure}

\paragraph{Learning is useful for non-congested networks}

It is worth to highlight that, if we write \eqref{eq:43:decomppc1} as $p_{c1} = p_c + p_{ca} \left(1-p_c\right)$,
then it is obvious that $p_{c1}$ is always larger than $p_c$ (as $p_{ca} \left(1-p_c\right) > 0$).
But for large values of $p_c$, $p_{ca}\left(1-p_c\right) \simeq 0$ so the gap gets small,
and for small values of $p_c$ the gap is significant.
Moreover, we can verify (\eg, numerically or by differentiating)
that the gap decreases when $p_c$ increases (for fixed $N$ and $m$).
This backups mathematically the observation we made from Fig.~\ref{fig:43:Approximation_m10}:
the smaller the $p_c$, the larger is the gap between $p_c$ and $p_{c1}$.

We interpret this fact in two different situations.
\begin{itemize}
	\item
	On the one hand, in a congested network, when devices suffer from a large probability of collision on their first transmission (\ie, $p_c$ is not so small), then $p_{c1}\simeq p_c$ and so devices cannot really hope to reduce their collision probabilities even if the use a different channel for retransmission.
	\item
	On the other hand, if $p_c$ is small enough, \ie, in a network not yet too congested, then our derivation shows that $p_{c1} \gg p_c$, meaning that the possible gain of retransmitting in a different channel that the one used for the first transmission can be large, in terms of collision probability (\eg, up-to $10\%$ in this experimental setting).
	In other words, when learning can be useful (small $p_c$), learning to retransmit in a different channel can have a large impact on the global collision rate,
	thus justifying our approach.
\end{itemize}

%-----------------------------------------------------------------
\subsection{A well-known MAB Algorithm: \UCB{}}\label{sub:43:MABalgo}

Without loss of generality, we have adopted a well-studied stochastic MAB learning algorithm, where the reward distributions are unknown and assumed to be independent and identically distributed (i.i.d). The arms model the channels denoted as $C(t) \in \llbracket 1, K \rrbracket$, and the players, the dynamic devices, learn the distributions to be able to progressively focus on the best arm, \ie, the arm with largest mean representing the mean availability of a given channel $k$.

Before presenting our proposed heuristics, we describe a \UCB{} bandit algorithm \cite{Auer}. It has reported to be efficient, while featuring a low complexity for its implementation. For this reason, it has been employed for IoT applications \cite{Bonnefoi17}, and we employ this approach to develop our proposals.

% % --- --- --- --- --- --- --- --- --- ---
\paragraph{The \UCB{} algorithm}\label{sub:algoUCB}

A first approach is to only use an empirical mean estimator of the rewards in every channel, and select the channel with highest estimated mean at every time step; but this greedy approach is known to fail dramatically \cite{Auer02}.
Indeed, with this policy, the selection of arms depends too much on the first draws: if the first transmission in one channel fails and the first one on other channels succeeds, the device will \emph{never} use the first channel again, even if it is the best one (\ie, the most available, in average).

Rather than relying on the empirical mean reward, \UCB{} algorithms instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm, which can be viewed as adding a ``bonus'' exploration to the empirical mean. They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model, as the statistically best possible arm (\ie, the highest \UCB{}) is selected.

More formally, for one device, let $N_k(t)$ be the number of times the channel $k$ (for $k\in \llbracket 1, K \rrbracket$) was selected up-to time $t-1$, for $t\geq 0$
for any $t\in\mathbb{N}$,
\begin{equation}\label{eq:43:Nkt}
	N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(C(\tau) = k),
\end{equation}
where $\mathbbm{1}$ is an indicator function that is equal to $1$, if the IoT device chooses, for its $\tau$-th transmission, a channel $k$, and $0$ otherwise. The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
\begin{equation}\label{eq:43:mukt}
	\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(C(\tau) = k).
\end{equation}
where $r_{k}(t)$ is the reward obtained after transmission in channel $k$ at time $t$ ($1$ for a successful transmission, and $0$ otherwise)
%
A \emph{confidence} term $B_k(t)$ is given by \cite{Auer02},
\begin{equation}\label{eq:43:Bkt}
	B_k(t) = \sqrt{\alpha \log(t) / N_k(t)},
\end{equation}
where $\alpha$ refers to an exploration coefficient\footnote{~In fact, the larger this coefficient is, the longer the exploration, while the \UCB{} algorithm is proven to be order optimal for $\alpha>0.5$ \cite{Bubeck12}, and has reported a good performance for lower values of $\alpha>0$.},
that we chose equal to $1/2$, as suggested in \cite{Audibert07} and as done in previous works \cite{Bonnefoi18,Bonnefoi17}.
Then, an upper confidence bound in each channel $k$ is defined as
\begin{equation}\label{eq:43:ucb}
	U_k(t) = \widehat{\mu_k}(t) + B_k(t).
\end{equation}
Finally, the transmission channel at time step $t$
is the one maximizing this \UCB{} index $U_k(t)$,
as it is the one expected to be the best one at the current time step $t$,
\begin{equation}\label{eq:43:maxucb}
	C(t) = \arg\max_{1\leq k \leq K} U_k(t).
\end{equation}

The \UCB{} algorithm is implemented independently by each device, and we present it in Algorithm~\ref{algo:43:UCB}.
Note that a device using this first approach is only able to select a channel for the first and all the corresponding retransmissions of a packet.

\begin{small} % XXX remove if needed
\begin{figure}[h!]
	\centering
	\begin{algorithm}[H]
		\begin{small} % XXX remove if needed
		\For(){$t = 0, \dots, T$}{
			Compute for each channel $ U_k(t) = \widehat{\mu_k}(t) + B_k(t).$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
			Transmit in channel $C(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
			Reward $r_{C(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
		}
		\caption{The \UCB{} algorithm for channel selection.}
		\label{algo:43:UCB}
		\end{small} % XXX remove if needed
	\end{algorithm}
\end{figure}
\end{small} % XXX remove if needed

% ----------------------------------------------------------------------
\subsection{Heuristics to (try to) learn how to retransmit efficiently}\label{sub:43:heuristics}

A device that implements the UCB algorithm is led to focus is transmissions and retransmissions in the channel which has been identified as the best.
As explained before, focusing in one channel increases the collision probability in retransmissions.
In this Section, we describe the proposed heuristics for the channel selection in a retransmission. It is carried out taking
into account that a device can incorporate a different channel selection strategy while being in a back-off state.
Hence, a natural question is to evaluate whether using this additional contextual information can improve the performance of a learning policy.

For that end, all of our heuristics comprise two stages:
the first stage is a \UCB{} algorithm employed for the first attempt to transmit,
and the second stage is another algorithm used for channel selections for the next retransmissions.
%
We present below four heuristics for this second stage (short names in ``quotes'' correspond to the legend on Figures~\ref{fig:43:mainExperiment1}, \ref{fig:43:mainExperiment2}).

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Uniform random retransmission (``Random'')}\label{sub:UCBthenRandom}

In this first proposal, the device uses a random channel selection, following a uniform distribution (in $\llbracket 1, K \rrbracket$).
It is described below in Algorithm~\ref{algo:43:UCBthenRandom}.

\vspace*{-3pt}
\begin{figure}[h!]
	\centering
	\begin{algorithm}[H]
	\begin{small} % XXX remove if needed
	\For(){$t = 0, \dots, T$}{
			\uIf{First packet transmission}{
				Use first-stage \UCB. %as in Algorithm~\ref{algo:43:UCB}.
			}
			\Else(\tcp*[f]{Random retransmission}){
				Transmit in channel $C(t) \sim \mathcal{U}(1,\ldots,K)$\;
			}
		}
		\caption{Uniform random retransmission.}    % A naive
		\label{algo:43:UCBthenRandom}
	\end{small} % XXX remove if needed
\end{algorithm}
\end{figure}

% --- --- --- --- --- --- --- --- --- ---
\paragraph{\UCB{} for retransmission (``Only \UCB{}'')}\label{sub:TwoUCB}

Instead of applying a random channel selection,
another heuristic is to use a second \UCB{} algorithm in the second stage.
In other words, we expect that this algorithm is able to learn the best channel to retransmit a packet.
It is described in Algorithm~\ref{algo:43:TwoUCB}, and it is still a practical approach, since the storage requirements and time complexity remains linear w.r.t. the number of channels $K$ (\ie, of order $\mathcal{O}(K)$).

Note that, we use the subscript $({}^r)$ to denote the variables
$\widehat{\mu^r_k}(t)$, $B^r_k(t)$ and $U^r_k(t)$,
related to the \UCB{} algorithm employed for the retransmission.

\vspace*{-3pt}
\begin{small} % XXX remove if needed
\begin{figure}[h!]
	\centering
	\begin{algorithm}[H]
		\begin{small} % XXX remove if needed
		\For(){$t = 0, \dots, T$}{
			\uIf{First packet transmission}{
				Use first-stage \UCB. %as in Algorithm~\ref{algo:43:UCB}.
			}
			\Else(\tcp*[f]{Random retransmission}){
				Compute for each channel $U^r_k(t) = \widehat{\mu^r_k}(t) + B^r_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
				Transmit in channel $C^r(t) = \arg\max_{1\leq k \leq K} U^r_k(t)$\;
				Reward $r^r_{C^r(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
			}
		}
		\caption{\UCB{} for retransmission.}    % A naive
		\label{algo:43:TwoUCB}
		\end{small} % XXX remove if needed
		\end{algorithm}
\end{figure}
\end{small} % XXX remove if needed

% --- --- --- --- --- --- --- --- --- ---
\paragraph{$K$ different {\UCB}s for retransmission (``$K$ \UCB{}'')}\label{sub:UCBthenKp1}

Another heuristic is to not use the same algorithm no matter where the collision occurred, but to use $K$ different \UCB{} algorithms.
Meaning that after a failed first transmission in channel $j$, the device relies on the $k$-th algorithm to decide its retransmission.
The corresponding algorithm is depicted in Algorithm~\ref{algo:43:UCBthenKp1}.
Each of these algorithms are denoted using the subscript $({}^{j})$, for $j\in\llbracket 1, K \rrbracket$.

Although, this approach increases the complexity and store requirements (of order $\mathcal{O}(K^2)$).
For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
For instance, for $K=4$ channels,
the memory to store $K+1=5$ algorithms is of the order of the requirements to storing one.
% For other networks this heuristic could not be practical, for instance  Sigfox consist in a large number of very narrow-band channels.

%\footnote{~The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels. But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.}

\vspace*{-3pt}
\begin{small} % XXX remove if needed
	\begin{figure}[h!]
		\centering
		\begin{algorithm}[H]
			\begin{small} % XXX remove if needed
			% \KwData{$\forall k,j\in[\![1;K]\!]$, $N_k^j(t)$, $\widehat{\mu_k}^j(t)$, $B_k^j(t)$ and $U_k^j(t)$}
			\For(\tcp*[f]{At every time step})
			{$t = 0, \dots, T$}{
				\uIf{First packet transmission}{
					Use first-stage \UCB. %as in Algorithm~\ref{algo:43:UCB}.
				}
				\Else(\tcp*[f]{Packet retransmission}){ % Retr. after trying channel $j$
					j $\leftarrow$ last channel selected by first-stage \UCB\;
					Compute for each channel $U_k^j(t) = \widehat{\mu_k}^j(t) + B_k^j(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
					Transmit in channel $C^j(t) = \arg\max_{1\leq k \leq K} U^j_k(t)$\;
					Reward $r^j_{C^j(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
				}
			}
			\caption{$K$ different {\UCB}s for retransmission.}
			\label{algo:43:UCBthenKp1}
			\end{small} % XXX remove if needed
		\end{algorithm}
	\end{figure}
\end{small} % XXX remove if needed

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'')}\label{sub:UCBwithDelay}

This last heuristic is a composite of
the random retransmission (Algorithm~\ref{algo:43:UCBthenRandom})
and the \UCB{} retransmission (Algorithm~\ref{algo:43:TwoUCB}) approaches.
Instead of starting the second stage \UCB{} directly from the first retransmission, we introduce a fixed delay $\Delta\in\mathbb{N}$, $\Delta \geq 1$,
and start to rely on the second stage \UCB{} after $\Delta$ transmissions.
The selection for the first steps is handled with the random retransmission.

The idea behind this delay is to allow the first stage \UCB{} to start learning the best channel, before starting the second stage \UCB{} (see details in Algorithm~\ref{algo:43:UCBwithDelay}).
The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand.

Note that, we use the subscript $({}^d)$ to denote the variables
related to the delayed second-stage \UCB{} algorithm.

%\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to be able to simulate the problem in advance in order to find a good value of some internal parameter. As such, we only consider a delay of $\Delta=100$.}.

\vspace*{-3pt}
\begin{small} % XXX remove if needed
	\begin{figure}[h!]
		\centering
		\begin{algorithm}[H]
			\begin{small} % XXX remove if needed
			\For(\tcp*[f]{At every time step})
			{$t = 0, \dots, T$}{
				\uIf{First packet transmission}{
			 		Use first-stage \UCB. %as in Algorithm~\ref{algo:43:UCB}.
				}
				\uElseIf(\tcp*[f]{Random selection}){$t \leq \Delta$}{
					Transmit randomly in a channel $C(t) \sim \mathcal{U}(1,\ldots,K)$.
				}
				\Else(\tcp*[f]{Delayed \UCB{}}){
					Compute for each channel $ U^d_k(t) = \widehat{\mu^d_k}(t) + B^d_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
					Transmit in channel $C^d(t) = \arg\max_{1\leq k \leq K} U^d_k(t)$\;
					Reward $r^d_{C^d(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
				}
			}
			\caption{Delayed \UCB{} for retransmission.}
			\label{algo:43:UCBwithDelay}
			\end{small} % XXX remove if needed
		\end{algorithm}
	\end{figure}
\end{small} % XXX remove if needed

% ----------------------------------------------------------------------
\subsection{Simulations to compare our heuristics}\label{sub:43:numExp}

We simulate our network considering $N$ devices following the contention Markov process described in Section \ref{sub:43:model}, and a LoRa standard with $K=4$ channels.
Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of $1\;\mathrm{s}$.
%
For the evaluation of the proposed heuristics, a total number of $T=10^{6}$ time slots is considered, and the results are averaged over $10^{3}$ independent random simulations.
%In this way, it allows us to evaluate the learning rate, as well as the convergence of each learning approach.

In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform repartition of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
In other words, the channels are occupied $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $M = 5$, and $m=5$.
In Fig.~\ref{fig:43:mainExperiment1}, we show the successful transmission rate versus the number of slots, for all the proposed heuristics.

A first result is that all the heuristics clearly outperform the non-learning approach that simply use random channel selection for both transmissions and retransmissions (\ie, the \textbf{no \UCB{}} curve).
The improvement of the heuristics over the non-learning approach is evident, and for every heuristic that use a kind of learning mechanism it can observed a successful transmission rate that increases rapidly (or equivalently an PLR decreasing).
Moreover, all of these approaches show a fast convergence making them suitable for the targeted application.
It is also worths mentioning that the employment of the same \UCB{} algorithm for retransmissions denoted here as ``Only \UCB{}'' achieves a better performance, while a ``Random" retransmission features a slight degradation. This result can be explained as follows: the loss of performance related to the separation of information for several algorithms is greater than the gain obtained by considering the first transmissions and retransmissions separately.

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/ResultsUCB.eps}
	\caption{
		Comparison among the exposed heuristics for the retransmission: \textcolor{blue}{Only \UCB}, \textcolor{cyan}{Random}, \textcolor{purple}{\UCB}, \textcolor{green}{$K$ \UCB}, and \textcolor{red}{Delayed \UCB}.
		The usage of the same learning policy for transmissions and retransmission is named \textcolor{blue}{Only \UCB{}},
		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as \textbf{no \UCB{}}.
		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
	}
	\label{fig:43:mainExperiment1}
\end{figure}

We also consider in our analysis the case where $M=5$, and $m=10$ using ALOHA protocol, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
The corresponding results are depicted in Fig.~\ref{fig:43:mainExperiment2}.
In this case the successful transmission rate is degraded compared with achieved results in Fig.~\ref{fig:43:mainExperiment1}, this can be explained with the fact that we are considering in our network more devices that increase the collision probability.
It is important to highlight,that the ``Random" retransmission heuristic shows a poor performance in comparison to the other heuristics, and it can be attributed to the fact that the number of retransmission is increased, and consequently a
learning approach is able to take advantage of it.
Furthermore, the ``Delayed \UCB'' and  the``\UCB'' heuristics show quite similar results, after converging.
%
%Random, \UCB{}, $K$ \UCB{}, Delayed \UCB{} obtain similar results. The less efficient one is the $K$ \UCB{}, and this makes sense as having more learning algorithm needs more for each of them to learn. The most efficient one is the simple \UCB{} procedure, and this was quite surprising.

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/ResultsUCB2.eps}
	\caption{
		Second scenario: learning helps a lot (a gain of $30\%$ in terms of collision probability), and learning to retransmit smartly is needed.
		We observe that the \textcolor{cyan}{random retransmission} achieves poor performance compared to the others.
	}
	\label{fig:43:mainExperiment2}
\end{figure}

The conclusions we can draw from depicted results are twofold.
First, MAB learning algorithms are very useful to reduce the collision rate in LPWA networks, a gain of up-to $30\%$ of successful transmission rate is observed after convergence.
A second conclusion that can be highlighted is that, using learning mechanisms for retransmissions can be an interesting way to reduce collisions in networks with massive deployments of IoT as this can be checked in Fig.~\ref{fig:43:mainExperiment2}, where the random retransmission heuristic is not very advantageous in front of the  \UCB-based approaches that use learning for channel selection during the retransmission procedure.

% ----------------------------------------------------------------------
\subsection{Conclusion about this second model}\label{sub:43:conclusion}

In this section, we presented a retransmission model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.

We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.

\paragraph{Future works}
%
The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works. Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
%
For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier08}
or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
% or more recently GLR-\UCB{} \cite{Besson2019COLT},
that nevertheless have not been explored for the targeted problem.
Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this section.

In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our recent works \cite{modiDemo2016,Besson2019WCNC}.
Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes, and Christophe Moy at IETR and University Rennes 1.
%
A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.

\paragraph{Note on the simulation code}
%
The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in July and August $2018$.

% ----------------------------------------------------------------------
\bibliographystyle{ieeetr}
\bibliography{IEEE_WCNC__2019__Paper__BMBM}

### 2019-02-13 Nomenclature/Nomenclature.tex
See #11 ad #4
Baby front-page, it works

@@
+\nomenclature[A]{wrt}{\textit{with respect to}}
+\nomenclature[A]{wlog}{\textit{with loss of generality}} % TODO not used?
+\nomenclature[A]{iid}{\textit{identically and independently distributed}}
+
+% Objects
+\nomenclature[A]{USRP}{\textit{Universal Software Radio Peripheral}}
+
+% Institution
+\nomenclature[A]{ANR}{\textit{Agence Nationale de la Recherche} (French National Research Agency)}
+\nomenclature[A]{ENS}{\textit{École Normale Supérieure}}
+\nomenclature[A]{CNRS}{\textit{Centre National de la Recherche Scientifique}}
+\nomenclature[A]{SCEE}{\textit{Signal Communication Électronique Embarquée} (research team in CentraleSupélec, campus of Rennes)}
+
+% IoT networks terms
+\nomenclature[A]{IoT}{\textit{Internet of Things}}
+\nomenclature[A]{NB-IoT}{\textit{Narrow-Band Internet of Things}}
+\nomenclature[A]{LPWAN}{\textit{Low-Power Wide-Area Network}}
+\nomenclature[A]{CR}{\textit{Cognitive Radio}}
+\nomenclature[A]{OSA}{\textit{Opportunistic Spectrum Access}}
+\nomenclature[A]{SU}{\textit{Secondary User}}
+\nomenclature[A]{PU}{\textit{Primary User}}
+\nomenclature[A]{BST}{\textit{Base-Station} (or \textit{gateway})}
+\nomenclature[A]{QPSK}{\textit{Quadrature Phase-Shift Keying}}
+
+% Communication terms
+\nomenclature[A]{PLR}{\textit{Packet Loss Ratio}}
+
+% Algorithms or tools
 \nomenclature[A]{UCB}{\textit{Upper Confidence Bound} (object or algorithm)}
 \nomenclature[A]{klUCB}{\textit{Kullback-Leibler Upper Confidence Bound} (algorithm)}
+\nomenclature[A]{TS}{\textit{Thompson Sampling} (algorithm)}
 \nomenclature[A]{KL}{\textit{Kullback-Leibler} (divergence)}

### 2019-02-13 2-Chapters/Conclusion/conclusion.tex
Ok for tex root
Still fighting against my template...
Okay for chapters and parts

@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 
 % First chapter begins here
-\chapter*{General Conclusion and Perspectives}
-\addcontentsline{toc}{chapter}{General Conclusion}
+\chapter{General Conclusion and Perspectives}
+% \addcontentsline{toc}{chapter}{General Conclusion}
+\label{chapter:conclusion}
 %\minitoc
-\label{chapter:conclusion}
+
+Abstract for Conclusion.
+
 % Write miniTOC just after the title
@@
 
+\section{General Conclusion}
 In this PhD thesis, blabla.
@@
 
-\section*{Perspectives}
-We present in the following the possible areas of future studies
+\section{Perspectives}
+We present in the following the possible areas of future studies.
 
-\subsubsection*{Blabla}
-\begin{itemize}
-\item Blabla
-\end{itemize}
+\section{Personal Summary}
+
+I would like to highlight that from a personal point of view, this thesis was awesome blabla.
+
+J'aimerai conclure en dédicaçant cette thèse à mes deux frères, qui ont été ma plus grande source d'inspiration depuis mon plus jeune âge ; ainsi qu'à la mémoire de XXX XXX qui nous manquera cruellement.
+\textcolor{white}{Je souhaite aussi remercier une fois encore mes amis et amies les plus proches, Hélène, Pierre, Lola, Sélim, Angèle}

### 2019-02-13 1-Preamble/4-Acknowledgement/acknowledgement.tex
Still fighting against my template...
Lots of macro were clean up, almost done for copy/paste of chapter 5 from my ALT 2018 paper

--- 2019-02-11 1-Preamble/4-Acknowledgement/acknowledgement.tex
@@
 This work is supported by
-the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{XXX}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
+the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{N ANR-16-CE40-0002}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
 by R\'egion Bretagne, France,
-the French Ministry of Higher Education and Research (MENESR),
+by CNRS under the PEPS project BIO (in $2017$),
+by the French Ministry of Higher Education and Research (MENESR),
 by \'Ecole Normale Sup\'erieure de Paris-Saclay.
@@
 Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó, me sorprendió con su entusiasmo y me consoló con su amor durante el año pasado.
-Yo te quiero gatita, ¡impresivo!
+\textcolor{white}{Yo te quiero gatita, ¡impresivo!}
 

### 2019-02-13 2-Chapters/5-Chapter/chapter5.tex
Better outline
Version 0.1
Ok for tex root
Okay for chapters and parts
Lots of macro were clean up, almost done for copy/paste of chapter 5 from my ALT 2018 paper
Added gold and other colors, for Chapter 5.

@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 
 % First chapter begins here
-\chapter{Multi-Player Multi-Armed Bandits}
+\chapter{Multi-Player Multi-Armed Bandits Models}
 \label{chapter:5}
 \minitoc
+
+Abstract for Chapter 5.
+
 \newpage
@@
 \graphicspath{{2-Chapters/5-Chapter/Images/}}
+\graphicspath{{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/}}
 
@@
 % ----------------------------------------------------------------------------
+\input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex}
+
+% ----------------------------------------------------------------------------
+\section{Towards efficient algorithms solving the "no sensing" case}
+\label{sec:5:XXX}
+
+% ----------------------------------------------------------------------------
 \section{Conclusion of Chapter 5}
@@
 Future works include...
+
+% ----------------------------------------------------------------------------
+\section{Appendix for Chapter 5}
+\label{sec:5:appendix}
+
+% ----------------------------------------------------------------------------
+\input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex}

### 2019-02-13 2-Chapters/5-Chapter/chapter5.tex
I still don't have correct references...

@@
 
-This chapter is basically a raw include from my paper ``Multi-Player Bandits Revisited'', see https://hal.inria.fr/hal-01629733
-
-I also need to include additional simulations from the latest research articles, and discussions about the new works that took inspiration in our paper.
+\TODOL{I also need to include additional simulations from the latest research articles, and discussions about the new works that took inspiration in our paper.}
 

### 2019-02-14 2-Chapters/3-Chapter/SMPyBandits_paper.git/paper.tex
Done for polishing references.

\documentclass[a4paper,10pt,]{article}

% http://www.jmlr.org/format/format.html
%
% See http://jmlr.org/papers/v18/ for example of JMLR papers from last year
%
% Any additional packages needed should be included after jmlr2e.
% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
% and defines many common macros, such as 'proof' and 'example'.
% It also sets the bibliographystyle to plainnat; for more information on
% natbib citation styles, see the natbib documentation, a copy of which
% is archived at http://www.jmlr.org/format/natbib.pdf
\usepackage{jmlr2e}
\usepackage{amsmath}

\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
  \usepackage[T1]{fontenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi

% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{xcolor}
\definecolor{dkgreen}{rgb}{0,0.4,0}           % Define a new color rgb(0, 102, 0)
\definecolor{gray}{rgb}{0.5,0.5,0.5}          % Define a new color rgb(127, 127, 127)
\definecolor{mauve}{rgb}{0.58,0,0.82}         % Define a new color rgb(147, 0, 209)
\definecolor{darkgreen}{rgb}{0.00,0.70,0.00}  % Define a new color rgb(0,178,0)
\definecolor{gold}{RGB}{255,184,0}  % rgb(255,184,0)

\usepackage{lastpage,fancyhdr}     % customize the headers and footers
\pagestyle{fancy}
    \renewcommand{\headrulewidth}{0.2pt}
    \renewcommand{\footrulewidth}{0.2pt}
    \lhead{\emph{SMPyBandits} presentation paper}
    \rhead{\emph{\today}}
    \lfoot{\href{https://GitHub.com/SMPyBandits/SMPyBandits}{GitHub.com/SMPyBandits/SMPyBandits}}
    \cfoot{\(\thepage/\pageref{LastPage}\)}
    \rfoot{Lilian Besson}

% \usepackage{graphicx}
\usepackage{subcaption}
\graphicspath{{../}}
\DeclareGraphicsExtensions{.jpg,.pdf,.mps,.eps,.png}

\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

\usepackage{marvosym}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% -----------------------------------------------------------------
% \ShortHeadings{short title}{short authors}
\ShortHeadings{\emph{SMPyBandits} presentation paper}{Lilian Besson}

% Heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}
% \jmlrheading{vol}{year}{pages}{Submitted date}{published date}{paper id}{authors}
% \jmlrheading{18}{2018}{1-5}{4/18}{??/18}{Lilian Besson} % dates/pages from editor
\jmlrshortheading{2018}{Lilian Besson}

% \firstpageno{1} % the pagenumber you are assigned to start with by the editor.

% -----------------------------------------------------------------
\begin{document}

\title{\emph{SMPyBandits}: an Experimental Framework for Single and Multi-Players Multi-Arms Bandits Algorithms in Python}

% -----------------------------------------------------------------
% \author{
% Lilian Besson\thanks{
% \textcolor{blue}{\Letter: \texttt{Lilian.Besson{[}AT}CentraleSupelec{[}.}fr}},
% \textcolor{darkgreen}{ORCID: \href{https://orcid.org/0000-0003-2767-2563}{\texttt{0000-0003-2767-2563}}}
% \newline
% PhD Student at CentraleSupélec, campus of Rennes, SCEE team \& Inria
% Lille Nord Europe, SequeL team.
% }
% }

\author{\name Lilian Besson \email {Lilian}{.}{Besson}{@}{CentraleSupelec}{.}{fr} \\
        \addr CentraleSup\'elec (campus of Rennes), IETR, SCEE Team,\\
        Avenue de la Boulaie -- CS $47601$, F-$35576$ Cesson-S\'evign\'e, France
}

% -----------------------------------------------------------------
% \date{23 March 2018}
\date{}  % remove date

% -----------------------------------------------------------------
\editor{?}

\maketitle

\vspace*{10pt}

% -----------------------------------------------------------------
\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
  \emph{SMPyBandits} is a package for numerical simulations on
  \emph{single}-player and \emph{multi}-players
  \href{https://en.wikipedia.org/wiki/Multi-armed_bandit}{Multi-Armed
  Bandits (MAB)} algorithms, written in
  \href{https://www.python.org/}{Python (2 or 3)}.
  This library is the most complete open-source implementation of
  state-of-the-art algorithms tackling various kinds of sequential
  learning problems referred to as Multi-Armed Bandits. It is
  extensive, simple to use and maintain, with a clean and well
  documented codebase. It allows fast prototyping of experiments,
  with an easy configuration system and command-line options
  to customize experiments.

  \textbf{Keywords:} sequential learning; multi-armed bandit; reinforcement learning; python.
\end{abstract}

% \begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\section{Presentation}\label{presentation}

\subsection{Single-Player MAB}\label{single-player-mab}

Multi-Armed Bandit (MAB) problems are well-studied sequential decision
making problems in which an agent repeatedly chooses an action (the
``\emph{arm}'' of a one-armed bandit) in order to maximize some total
reward \citep{Robbins52}, \citep{LaiRobbins85}. Initial motivation for
their study came from the modeling of clinical trials, as early as 1933
with the seminal work of Thompson \citep{Thompson33}, where arms
correspond to different treatments with unknown, random effect. Since
then, MAB models have been proved useful for many more applications,
that range from cognitive radio \citep{Jouini09} to online content
optimization like news article recommendation \citep{Li10}, online
advertising \citep{LiChapelle11}, A/B Testing \citep{Kaufmann14,Jamieson17}, or portfolio optimization \citep{Sani12}.

More formally, a stochastic MAB is defined by $K>1$ distributions $\nu_k$ (arms),
and \emph{i.i.d.} rewards $r_k(t) \sim \nu_k, \forall t$.
An agent choose arm $A(t)\in\{1,\dots,K\}$ at time $t$ and
observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
Her goal is to maximize $\sum_{t=1}^T r_{A(t)}(t)$ by sequentially exploring the $K$ arms,
and she essentially has to find and exploit the best one as fast as possible.
This library tackles one dimensional distributions,
and supports \href{https://SMPyBandits.GitHub.io/docs/Arms.Bernoulli.html}{\texttt{Bernoulli}}, \href{https://SMPyBandits.GitHub.io/docs/Arms.Binomial.html}{\texttt{binomial}}, \href{https://SMPyBandits.GitHub.io/docs/Arms.Poisson.html}{\texttt{Poisson}}, and a generic \href{https://SMPyBandits.GitHub.io/docs/Arms.DiscreteArm.html}{\texttt{discrete}} distributions,
as well as \href{https://SMPyBandits.GitHub.io/docs/Arms.Exponential.html}{\texttt{exponential}}, \href{https://SMPyBandits.GitHub.io/docs/Arms.Gamma.html}{\texttt{gamma}}, \href{https://SMPyBandits.GitHub.io/docs/Arms.Gaussian.html}{\texttt{Gaussian}} (of known scale or variance) and \href{https://SMPyBandits.GitHub.io/docs/Arms.Uniform.html}{\texttt{uniform}} continuous distributions,
which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$).

\emph{SMPyBandits} is a complete open-source implementation of
single-player bandit algorithms,
containing \href{https://SMPyBandits.GitHub.io/docs/Policies.html}{over 65} algorithms.
It uses a well-designed hierarchical structure and
\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{class
inheritance scheme} to minimize redundancy in the codebase.
For example, many existing algorithms are index-based:
they compute an index $I_k(t)\in\mathbb{R}$ for each arm $k$ and simply play $A(t) = \arg\max_k I_k(t)$ at time $t$.
As such, it is easy to write new index-based algorithms by inheriting from the
\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{\texttt{IndexPolicy}}
class, and simply defining one method \texttt{computeIndex(k)} to compute the index $I_k(t)$.

\subsection{Multi-Players MAB}\label{multi-players-mab}

For Cognitive Radio and other applications, a well-studied extension is to
consider \(M\geq2\) players, interacting on the \emph{same} \(K\) arms.
Whenever two or more players select the same arm at the same time, they
all suffer from a collision. Different collision models has been
proposed, and the simplest one consists in giving a \(0\) reward to each
colliding players. Without any centralized supervision or coordination
between players, they must learn to access the \(M\) best resources
(\emph{i.e.}, arms with highest means) without collisions.
\emph{SMPyBandits} implements
\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{all
collision models} found in the literature, as well as all the
algorithms from the last 10 years (including
\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{\texttt{rhoRand}},
\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{\texttt{MEGA}},
\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{\texttt{MusicalChair}}, and our state-of-the-art algorithms
\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{\texttt{RandTopM}}
and
\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{\texttt{MCTopM}}
from \citet{BessonALT2018}).
For comparison, realistic or full-knowledge centralized algorithms are also implemented.

% \begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\section{Features}\label{features}

With this numerical framework, simulations can run on a single CPU or a
single multi-core machine using joblib \citep{joblib}, and summary plots are
automatically saved as high-quality PNG, PDF and EPS, using matplotlib \citep{matplotlib} and seaborn \citep{seaborn}.
Raw data from each simulation is also saved in a HDF5\textsuperscript{®} file using h5py \citep{h5py}, an efficient and compressed binary format, to allow easy post-mortem exploration of simulation results.
Making new simulations is very easy, one only
needs to write a configuration script (\texttt{configuration.py}), without needing a complete knowledge of the internal code architecture.

A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here:
\href{SMPyBandits.GitHub.io/}{\texttt{https://SMPyBandits.GitHub.io}}.

\subsection{How to run experiments?}\label{how-to-run-the-experiments}

We show how to install \emph{SMPyBandits}, and an example of how to run
a simple experiment. This bash snippet\footnote{See
\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}
for more details.} shows how to clone the code\footnote{SMPyBandits is also available on Pypi, see
\href{https://pypi.org/project/SMPyBandits/}{\texttt{pypi.org/project/SMPyBandits}}.
You can install it directly with
\texttt{sudo\ pip\ install\ SMPyBandits}, or from a
\texttt{virtualenv}.}, and install the requirements for Python 3 (once):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# 1. get the code in the folder you want}
\NormalTok{\textdollar }\FunctionTok{git}\NormalTok{ clone https://GitHub.com/SMPyBandits/SMPyBandits.git}
\NormalTok{\textdollar }\BuiltInTok{cd}\NormalTok{ SMPyBandits.git}
\CommentTok{# 2. install the requirements}
\NormalTok{\textdollar }\ExtensionTok{pip}\NormalTok{ install -r requirements.txt}
\end{Highlighting}
\end{Shaded}

Launching simulations is easy, for instance this snippet shows how to
start \(N=1000\) repetitions of a simple non-Bayesian
Bernoulli-distributed problem, for \(K=9\) arms, an horizon of
\(T=10000\) and on \(4\) CPUs. Such simulation takes about \(20\) minutes, on a standard \(4\)-cores \(64\) bits GNU/Linux
laptop. Using environment variables (\texttt{N=1000} etc) in the command
line is not required, but it is convenient:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# 3. run a single-player simulation}
\NormalTok{\textdollar }\VariableTok{BAYES=}\NormalTok{False }\VariableTok{ARM_TYPE=}\NormalTok{Bernoulli }\VariableTok{N=}\NormalTok{1000 }\VariableTok{T=}\NormalTok{10000 }\VariableTok{K=}\NormalTok{9 }\VariableTok{N_JOBS=}\NormalTok{4 }\KeywordTok{\textbackslash{}}
  \VariableTok{MEANS=}\NormalTok{[}\ExtensionTok{0.1}\NormalTok{,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] python3 main.py configuration.py}
\end{Highlighting}
\end{Shaded}

\subsection{Example of simulation and
illustration}\label{example-of-simulation-and-illustration}

A small script
\href{https://SMPyBandits.GitHub.io/docs/configuration.html}{\texttt{configuration.py}}
is used to import the
\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{arm classes}, the
\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{policy classes}
and define the problems and the experiments. Choosing the algorithms is
easy by customizing the \texttt{configuration{[}"policies"{]}} list in
the \texttt{configuration.py} file. For instance, one can compare the
standard anytime
\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{\texttt{klUCB}}
algorithm against the non-anytime variant
\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{\texttt{klUCBPlusPlus}}
algorithm, and also
\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{\texttt{UCB}}
(with \(\alpha=1\)) and
\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{\texttt{Thompson}}
(with
\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Beta
posterior}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{configuration[}\StringTok{"policies"}\NormalTok{] }\OperatorTok{=}\NormalTok{ [}
\NormalTok{  \{}\StringTok{"archtype"}\NormalTok{: klUCB, }\StringTok{"params"}\NormalTok{: \{}\StringTok{"klucb"}\NormalTok{: klucbBern\}\},}
\NormalTok{  \{}\StringTok{"archtype"}\NormalTok{: klUCBPlusPlus, }\StringTok{"params"}\NormalTok{: \{}\StringTok{"horizon"}\NormalTok{: HORIZON, }\StringTok{"klucb"}\NormalTok{: klucbBern\}\},}
\NormalTok{  \{}\StringTok{"archtype"}\NormalTok{: UCBalpha, }\StringTok{"params"}\NormalTok{: \{}\StringTok{"alpha"}\NormalTok{: }\DecValTok{1.0}\NormalTok{\}\},}
\NormalTok{  \{}\StringTok{"archtype"}\NormalTok{: Thompson, }\StringTok{"params"}\NormalTok{: \{}\StringTok{"posterior"}\NormalTok{: Beta\}\}}
\NormalTok{]}
\end{Highlighting}
\end{Shaded}

Running the simulation as shown above will save figures in a sub-folder,
as well as save data (pulls, rewards and regret) in a \href{http://docs.h5py.org/en/stable/high/file.html}{HDF5 file}\footnote{E.g., this simulation produces \href{https://GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}.}.
Figure~\ref{fig:plot1} below shows the average regret for these \(4\)
algorithms. The regret is the difference between the cumulated rewards
of the best fixed-armed strategy (which is the oracle strategy for
stationary bandits), and the cumulated rewards of the considered
algorithms.

\begin{figure}
\centering
\begin{subfigure}[b]{0.49\textwidth}
  \includegraphics[width=1.15\textwidth]{../plots/paper/3.pdf}
\end{subfigure}
% ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
  %(or a blank line to force the subfigure onto a new line)
\begin{subfigure}[b]{0.49\textwidth}
  \includegraphics[width=1.15\textwidth]{../plots/paper/3_hist.pdf}
\end{subfigure}
\caption{Example of a single-player simulation showing the average regret and histogram of regrets of
\(4\) algorithms. They all perform very well: each algorithm is known to
be order-optimal (\emph{i.e.}, its regret is proved to match the
lower-bound up-to a constant), and each but UCB is known to be optimal
(\emph{i.e.} with the constant matching the
lower-bound). For instance, Thomson sampling is very efficient in average (in \textcolor{gold}{yellow}), and UCB shows a larger variance (in \textcolor{red}{red}).\label{fig:plot1}}
\end{figure}

% \begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\section{\texorpdfstring{Research papers using \emph{SMPyBandits}}{Research using SMPyBandits}}\label{research-using-smpybandits}

\emph{SMPyBandits} was used for the following research articles since
\(2017\):

\begin{itemize}
\tightlist
\item
  For \citet{BessonALT2018}, we used \emph{SMPyBandits} for all the
  simulations for multi-player bandit algorithms\footnote{See
    the page
    \href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers}}
    on the documentation.}. We designed the two
  \href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{\texttt{RandTopM}}
  and
  \href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{\texttt{MCTopM}}
  algorithms and proved than they enjoy logarithmic regret in the usual
  setting, and outperform significantly the previous state-of-the-art
  solutions (\emph{i.e.},
  \href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{\texttt{rhoRand}},
  \href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{\texttt{MEGA}}
  and
  \href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{\texttt{MusicalChair}}).
\end{itemize}

\begin{itemize}
\tightlist
\item
  In \citet{BessonWCNC2018}, we used \emph{SMPyBandits} to illustrate and
  compare different aggregation algorithms\footnote{See the page
    \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation}}
    on the documentation.}. We designed a variant of the Exp3 algorithm
  for online aggregation (or boosting) of experts \citep{Bubeck12}, called
  \href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{\texttt{Aggregator}}.
  Aggregating experts is a well-studied idea in sequential learning and
  in machine learning in general. We showed that it can be used in
  practice to select on the run the best bandit algorithm for a certain
  problem from a fixed pool of experts. This idea and algorithm can have
  interesting impact for Opportunistic Spectrum Access applications
  \citep{Jouini09} that use multi-armed bandits algorithms for sequential
  learning and network efficiency optimization.
\end{itemize}

\begin{itemize}
\tightlist
\item
  In \citet{Besson2018DoublingTricks}, we used \emph{SMPyBandits} to illustrate and
  compare different ``doubling trick'' schemes\footnote{See the
    page
    \href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{\texttt{SMPyBandits.GitHub.io/DoublingTrick}}
    on the documentation.}. In sequential learning, an algorithm is
  \emph{anytime} if it does not need to know the horizon \(T\) of the
  experiments. A well-known trick for transforming any non-anytime
  algorithm to an anytime variant is the ``Doubling Trick'': start with
  an horizon \(T_0\in\mathbb{N}^*\), and when \(t > T_i\), use
  \(T_{i+1} = 2 T_i\). We studied two generic sequences of growing
  horizons (geometric and exponential), and we proved two theorems that
  generalized previous results. A geometric sequence suffices to conserve minimax
  regret bounds (in \(R_T = \mathcal{O}(\sqrt{T})\)), with a constant
  multiplicative loss \(\ell \leq 4\), but cannot be used to conserve a
  logarithmic regret bound (in \(R_T = \mathcal{O}(\log(T))\)). And an
  exponential sequence can be used to conserve logarithmic bounds, with
  a constant multiplicative loss also \(\ell \leq 4\) in the usual
  setting. It is still an open question to know if a well-tuned
  exponential sequence can conserve minimax bounds, or even ``weak'' minimax
  bounds (in \(R_T = \mathcal{O}(\sqrt{T \log(T)})\)).
\end{itemize}

\section{Dependencies}\label{dependencies}

This library is written in Python \citep{python}, for versions \emph{2.7+} or \emph{3.4+},
using \texttt{matplotlib} \citep{matplotlib} for 2D plotting,
\texttt{numpy} \citep{numpy} for data storing, random number generations
and operations on arrays, \texttt{scipy} \citep{scipy} for
statistical and special functions, and \texttt{seaborn} \citep{seaborn}
for pretty plotting and colorblind-aware colormaps.

Optional dependencies include \texttt{joblib} \citep{joblib} for parallel
simulations, \texttt{numba} \citep{numba} for automatic speed-up on small
functions, \texttt{sphinx} \citep{sphinx} for generating the
documentation, and \texttt{jupyter}
\citep{jupyter} used with \texttt{ipython} \citep{ipython} to experiment
with the code.

% ---------------------------------------------
\begin{normalsize}
  \bibliography{paper.bib}
\end{normalsize}

\end{document}

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
I still don't have correct references...
Done for polishing references.

@@
 In this Section, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
-We consider the Upper-Confidence Bound (\UCB) \cite{Auer}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
+We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
 Kaufmann12}.
-% Both algorithms have already been applied with success, for Opportunistic Spectrum Access \cite{Jouini} and recently for multi-users Cognitive Radio problems \cite{Toldov}.
+% Both algorithms have already been applied with success, for Opportunistic Spectrum Access \cite{Jouini10} and recently for multi-users Cognitive Radio problems \cite{Toldov16}.
 
-MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini}.
-For example, \cite{Jouini} presents the OSA setting, and \cite{Maghsudi} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi}.
+MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
+For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
 Many recent works show that MAB algorithms work well for real-world radio signal.
 However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \emph{i.i.d} (independent and identically distributed).
-In recent works like \cite{Bonnefoi}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d}.
+In recent works like \cite{Bonnefoi17}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d}.
 As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
@@
 % We now present the stochastic Multi-Armed Bandit (MAB) model,
-% and the two stochastic MAB algorithms used in our experiments \cite{bubeck2012regret}.
+% and the two stochastic MAB algorithms used in our experiments \cite{Bubeck12}.
 While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused by the learning policy used by dynamic objects.
@@
 
-%This section presents the formalism of the Multi-Armed Bandit (MAB) problem, as introduced by \cite{LaiRobbins85}, and two classical algorithms designed to tackle its simplest version (stochastic MAB), \UCB{} from \cite{Auer} and Thompson Sampling from \cite{Thompson33,Kaufmann12}.
-%For more details, \cite{bubeck2012regret} is an excellent review.
+%This section presents the formalism of the Multi-Armed Bandit (MAB) problem, as introduced by \cite{LaiRobbins85}, and two classical algorithms designed to tackle its simplest version (stochastic MAB), \UCB{} from \cite{Auer02} and Thompson Sampling from \cite{Thompson33,Kaufmann12}.
+%For more details, \cite{Bubeck12} is an excellent review.
 
@@
 % In a Cognitive Radio application, arms model the \emph{channels}, and players are the \emph{dynamic end-devices}.
-% For example in the classical OSA setting with sensing \cite{Jouini}, a single dynamic device (a player) sequentially tries to access channels (the arms), and collects a reward of 1 if the channel is available %(as there is no primary user, we can assume the transmission to be successful),
+% For example in the classical OSA setting with sensing \cite{Jouini10}, a single dynamic device (a player) sequentially tries to access channels (the arms), and collects a reward of 1 if the channel is available %(as there is no primary user, we can assume the transmission to be successful),
 % and 0 otherwise. So rewards represent the \emph{availability} of channels, and
@@
 Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
-The \emph{adversarial MAB} algorithms is a broader family, and a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{bubeck2012regret}.
+The \emph{adversarial MAB} algorithms is a broader family, and a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Bubeck12}.
 Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform worse than both \UCB{} and TS in the same experiments.
@@
 
-% ---------- Bibliography ----------
-%\begin{small}
-\bibliographystyle{ieeetr}
-\bibliography{biblio_RIoT}
-%\end{small}
+% % ---------- Bibliography ----------
+% %\begin{small}
+% \bibliographystyle{ieeetr}
+% \bibliography{biblio_RIoT}
+% %\end{small}

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
I still don't have correct references...
Done for polishing references.

@@
 %\todo[inline]{Contributions et interets de la demo}
-% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini}.
-% Recently, TS and UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov}, and the UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi}.
+% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
+% Recently, TS and UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}, and the UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
 % Many recent works show that MAB algorithms work well for real-world radio signal.
 % However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other objects is never really stationary or \emph{i.i.d.} (independent and identically distributed).
-% In recent works like \cite{Bonnefoi}, several objects are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d.}.
+% In recent works like \cite{Bonnefoi17}, several objects are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d.}.
 % As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent objects in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
@@
 We suggest to use low-cost algorithms, focusing on two well-known algorithms: a frequentist one (\UCB{}) and a Bayesian one (TS).
-We consider the Upper-Confidence Bound (\UCB{}) \cite{Auer}, and the Thompson Sampling (TS) algorithms \cite{Thompson33}.
-Both algorithms have already been applied with success in the context of wireless decision making, both empirically for Opportunistic Spectrum Access \cite{Jouini},
-and more recently for multi-users Cognitive Radio problems \cite{BessonALT18} with a more theoretical approach.
+We consider the Upper-Confidence Bound (\UCB{}) \cite{Auer02}, and the Thompson Sampling (TS) algorithms \cite{Thompson33}.
+Both algorithms have already been applied with success in the context of wireless decision making, both empirically for Opportunistic Spectrum Access \cite{Jouini10},
+and more recently for multi-users Cognitive Radio problems \cite{Besson2018ALT} with a more theoretical approach.
 
@@
 
-The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{bubeck2012regret}.
+The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
 % This algorithm is simple to implement and use in practice, even on embedded micro-processors with limited computation and memory capabilities.
@@
 The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12}.
-It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov,Mitton}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
+It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
 
@@
 
-% ---------- Bibliography ----------
-%\begin{small} % XXX small if needed?
-\bibliographystyle{ieeetr}
-\bibliography{biblio_RIoT}
-%\end{small} % XXX small if needed?
+% % ---------- Bibliography ----------
+% %\begin{small} % XXX small if needed?
+% \bibliographystyle{ieeetr}
+% \bibliography{biblio_RIoT}
+% %\end{small} % XXX small if needed?

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
I still don't have correct references...
Done for polishing references.

@@
 In the context of Cognitive Radio \cite{Mitola99,Haykin05},
-Multi-Arm Bandit (MAB) algorithms \cite{Auer,Auer02,Bubeck12} have been recently proposed as a potential solution for channel access in LPWA networks \cite{Bonnefoi18,Azari18,Bonnefoi17}.
+Multi-Arm Bandit (MAB) algorithms \cite{Auer02,Auer02,Bubeck12} have been recently proposed as a potential solution for channel access in LPWA networks \cite{Bonnefoi18,Azari18,Bonnefoi17}.
 We presented in Section~\ref{sec:4:firstModel} the impact of non-stationarity on the network performance using MAB algorithms is studied.
-Low-cost algorithms following two well-known approaches, such as the Upper-Confidence Bound (\UCB{}) \cite{Auer,Auer02}, and the Thompson Sampling (TS) algorithms \cite{Thompson33} have reported encouraging results.
+Low-cost algorithms following two well-known approaches, such as the Upper-Confidence Bound (\UCB{}) \cite{Auer02,Auer02}, and the Thompson Sampling (TS) algorithms \cite{Thompson33} have reported encouraging results.
 Other recent directions include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
@@
 of the application of MAB algorithms for slotted wireless protocols in a decentralized manner,
-or for instance applications to multi-hoping networks \cite{Mitton,Toldov}.
+or for instance applications to multi-hoping networks \cite{Mitton16,Toldov16}.
 None of the above mentioned works discusses in detail the impact of retransmissions on the performance of MAB learning algorithms as we do now.
@@
 
-Before presenting our proposed heuristics, we describe a \UCB{} bandit algorithm \cite{Auer}.
+Before presenting our proposed heuristics, we describe a \UCB{} bandit algorithm \cite{Auer02}.
 It has reported to be efficient, while featuring a low complexity for its implementation.
@@
 or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
-% or more recently GLR-\UCB{} \cite{Besson2019COLT},
+% or more recently GLR-\UCB{} \cite{Besson2019GLRT},
 that nevertheless have not been explored for the targeted problem.
@@
 
-% ----------------------------------------------------------------------
-\bibliographystyle{ieeetr}
-\bibliography{IEEE_WCNC__2019__Paper__BMBM}
+% % ----------------------------------------------------------------------
+% \bibliographystyle{ieeetr}
+% \bibliography{IEEE_WCNC__2019__Paper__BMBM}

### 2019-02-14 2-Chapters/3-Chapter/SMPyBandits_paper.git/paper.tex
Clean up doubles in the huge BibTeX file.

@@
 optimization like news article recommendation \citep{Li10}, online
-advertising \citep{LiChapelle11}, A/B Testing \citep{Kaufmann14,Jamieson17}, or portfolio optimization \citep{Sani12}.
+advertising \citep{LiChapelle11}, A/B Testing \citep{Kaufmann14,Jamieson17ABTest}, or portfolio optimization \citep{Sani12}.
 

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Clean up doubles in the huge BibTeX file.

@@
 We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
-Kaufmann12}.
+Kaufmann12Thompson}.
 % Both algorithms have already been applied with success, for Opportunistic Spectrum Access \cite{Jouini10} and recently for multi-users Cognitive Radio problems \cite{Toldov16}.
@@
 
-%This section presents the formalism of the Multi-Armed Bandit (MAB) problem, as introduced by \cite{LaiRobbins85}, and two classical algorithms designed to tackle its simplest version (stochastic MAB), \UCB{} from \cite{Auer02} and Thompson Sampling from \cite{Thompson33,Kaufmann12}.
+%This section presents the formalism of the Multi-Armed Bandit (MAB) problem, as introduced by \cite{LaiRobbins85}, and two classical algorithms designed to tackle its simplest version (stochastic MAB), \UCB{} from \cite{Auer02} and Thompson Sampling from \cite{Thompson33,Kaufmann12Thompson}.
 %For more details, \cite{Bubeck12} is an excellent review.

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Clean up doubles in the huge BibTeX file.

@@
 
-The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12}.
+The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
 It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Okay for bibtex

@@
 %
-For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier08}
+For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
 or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Working on Chapter 4, I want to have a first correct version by tomorrow.

@@
 
-\TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}
+% \TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}
 
@@
 
-As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol.
+As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
 All devices share a synchronized time, and know in advance the finite number of available RF channels.
@@
     \centering
-    \includegraphics[scale=0.28]{2-Chapters/4-Chapter/CrownCom_17.git/protocol.eps}
+    \includegraphics[scale=0.50]{2-Chapters/4-Chapter/CrownCom_17.git/protocol.eps}
     \caption{The considered time-frequency slotted protocol.
@@
 
-There are $N_c \geq 1$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
-Furthermore, in channel $i \in \llbracket 1; N_c \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{N_c} S_i$).
-We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $N_c$ (about $1000$ to $10000$, while $N_c$ is about $10$ to $50$).
+There are $K \geq 1$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
+Furthermore, in channel $i \in \llbracket 1; K \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{K} S_i$).
+We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (about $1000$ to $10000$, while $K$ is about $10$ to $50$).
 As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
-We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq N_c$, as the former makes more sense for IoT networks\footnote{Here, $N_c$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
+We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks\footnote{Here, $K$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
 
 We suppose that all devices follow the same emission pattern, being fixed in time, and we choose to model it as a simple Bernoulli process:
-all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $N_c / (S + D)$ for all devices to communicate successfully (in average).}.
+all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $K / (S + D)$ for all devices to communicate successfully (in average).}.
 The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed (\ie, real time during two messages), and $1/p$ is proportional to the \emph{duty cycle}.
@@
 
-We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; N_c \rrbracket = \{1, \dots, N_c\}$).
+We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; K \rrbracket = \{1, \dots, K\}$).
 % every time they communicate
@@
 \begin{equation}
-    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{N_c} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/N_c, \text{by uniform choice}}
+    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/K, \text{by uniform choice}}
 \end{equation}
 
-All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $i$ for any dynamic device is $p / N_c$, and there are $D-1$ other dynamic devices.
+All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $i$ for any dynamic device is $p / K$, and there are $D-1$ other dynamic devices.
 As they are independent, the probability that no other dynamic device sent in $i$
@@
 is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device}\;k\;\text{did not sent in}\;i) = \prod_{k=1}^{D-1} \Pr(\text{device}\;k\;\text{did not sent in}\;i)$.
-And $\Pr(\text{device}\;k\;\text{sent in}\;i) = p \times 1 / N_c$, by uniform choice on channels and the Bernoulli emission hypothesis.
-So $q = \prod_{k=1}^{D-1} (1 - p/N_c) = (1-p/N_c)^{D-1}$.
+And $\Pr(\text{device}\;k\;\text{sent in}\;i) = p \times 1 / K$, by uniform choice on channels and the Bernoulli emission hypothesis.
+So $q = \prod_{k=1}^{D-1} (1 - p/K) = (1-p/K)^{D-1}$.
 Thus we can conclude,
@@
     \Pr(\text{success}|\text{sent})
-    & = \sum_{i=1}^{N_c} \underbrace{(1 - p / N_c)^{D-1}}_{\text{No other dynamic device}} \times \underbrace{(1-p)^{S_i}}_{\text{No static device}} \times\; \frac{1}{N_c} \nonumber \\
-    & = \frac{1}{N_c} \left(1-\frac{p}{N_c}\right)^{D-1} \sum_{i=1}^{N_c} (1-p)^{S_i} .
+    & = \sum_{i=1}^{K} \underbrace{(1 - p / K)^{D-1}}_{\text{No other dynamic device}} \times \underbrace{(1-p)^{S_i}}_{\text{No static device}} \times\; \frac{1}{K} \nonumber \\
+    & = \frac{1}{K} \left(1-\frac{p}{K}\right)^{D-1} \sum_{i=1}^{K} (1-p)^{S_i} .
 \end{align}
@@
     \Pr(\text{success}|\text{sent})
-    & = \sum_{i=1}^{N_c} \Pr(\text{success}|\text{sent in channel}\;i) \; \Pr(\text{sent in channel}\;i) \nonumber \\
-    & = \sum_{i=1}^{N_c} \underbrace{(1 - p)^{D_i - 1}}_{\;\;D_i - 1 \;\text{others}\;\;} \times \underbrace{(1 - p)^{S_i}}_{\;\;\text{No static device}\;\;} \times \underbrace{ D_i / D }_{\;\;\text{Sent in channel}\; i\;\;}.
+    & = \sum_{i=1}^{K} \Pr(\text{success}|\text{sent in channel}\;i) \; \Pr(\text{sent in channel}\;i) \nonumber \\
+    & = \sum_{i=1}^{K} \underbrace{(1 - p)^{D_i - 1}}_{\;\;D_i - 1 \;\text{others}\;\;} \times \underbrace{(1 - p)^{S_i}}_{\;\;\text{No static device}\;\;} \times \underbrace{ D_i / D }_{\;\;\text{Sent in channel}\; i\;\;}.
 \end{align}
 
-Consequently, the optimal allocation vector $(D_1,\dots,D_{N_c})$ is the solution of the following real-valued constraint optimization problem :
+Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem :
 \begin{subequations}
@@
 \begin{align}
-\underset{D_1,\dots,D_{N_c}}{\arg\max}\; & \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
-\text{such that}\;\; & \sum_{i=1}^{N_c} D_i = D, \label{eq:41:eqCstr}\\
-& D_i \geq 0 \qquad \forall i\in\llbracket 1;N_c\rrbracket . \label{eq:41:ineqCstr}
+\underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
+\text{such that}\;\; & \sum_{i=1}^{K} D_i = D, \label{eq:41:eqCstr}\\
+& D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
 \end{align}
@@
     \item
-    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;N_c \rrbracket$.
-For such values for $D_i$, the objective function $f: (D_1, \dots, D_{N_c}) \mapsto \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
-    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{N_c}$ \cite{Luenberger68,Yaari77}.}.
+    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;K \rrbracket$.
+For such values for $D_i$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
+    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
         \item
@@
 \begin{equation}\label{eq:41:constraintLambda}
- \sum_{i=1}^{N_c} D_i^*(\lambda) = D.
+ \sum_{i=1}^{K} D_i^*(\lambda) = D.
 \end{equation}
@@
 % Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
-$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < N_c$, and $\widehat{D_{N_c}} = D - \sum_{i=1}^{N_c - 1} \widehat{D_i}$.
+$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < K$, and $\widehat{D_{K}} = D - \sum_{i=1}^{K - 1} \widehat{D_i}$.
 
@@
 % A Multi-Armed Bandit problem is defined as follows \cite{Thompson33,Robbins52,LaiRobbins85}.
-% There is a fixed number $N_c \geq 1$ of levers, or ``arms'', and a player has to choose one lever at each discrete time $t \geq 1, t \in \mathbb{N}$, denoted as $A(t) = k \in\{1,\dots,N_c\}$.
+% There is a fixed number $K \geq 1$ of levers, or ``arms'', and a player has to choose one lever at each discrete time $t \geq 1, t \in \mathbb{N}$, denoted as $A(t) = k \in\{1,\dots,K\}$.
 % Selecting arm $k$ at time $t$ yields a (random) \emph{reward}, $r_k(t) \in \mathbb{R}$, and the goal of the player is to maximize the sum of his rewards, $r_{1 \dots T} = \sum_{t = 1}^T r_{A(t)}(t)$.
@@
 Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\emph{e.g.}, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
-The RF band is divided in $N_c = 10$ channels.
-Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{N_c}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
+The RF band is divided in $K = 10$ channels.
+Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{K}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
 The dynamic devices have access to all the channels, and use learning algorithms.
@@
 We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
-We tried similar experiments with other values for $N_c$ and this repartition vector, and results were similar for non-homogeneous repartitions.
+We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions.
 Clearly, the problem is less interesting for homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Fig.\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Working on Chapter 4, I want to have a first correct version by tomorrow.

@@
-\subsection{A missing proof}
+\subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
 
@@
 %
-Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($N_c, p$), the repartition of static devices ($S_1,\dots,S_{N_c}$) and the Lagrange multiplier $\lambda$.
+Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($K, p$), the repartition of static devices ($S_1,\dots,S_{K}$) and the Lagrange multiplier $\lambda$.
 
@@
     \item
-    The function $f: (D_1, \dots, D_{N_c}) \mapsto \sum_{i=1}^{N_c} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{N_c}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
+    The function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{K}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
     \hfill{}$\square$
@@
 	\item
-	If $\boldsymbol{D} = (D_1,\dots,D_{N_c})$,
-    the Lagrangian is denoted $\mathcal{L}(\boldsymbol{D}, \lambda) = f(\boldsymbol{D}) + \lambda(D - \sum_{i=1}^{N_c} D_i)$, and its derivative w.r.t. $D_i$ is
+	If $\boldsymbol{D} = (D_1,\dots,D_{K})$,
+    the Lagrangian is denoted $\mathcal{L}(\boldsymbol{D}, \lambda) = f(\boldsymbol{D}) + \lambda(D - \sum_{i=1}^{K} D_i)$, and its derivative w.r.t. $D_i$ is
     $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) = (1-p)^{S_i + D_i - 1} + \log(1 - p) D_i (1-p)^{S_i + D_i - 1} - \lambda$.

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
Ok for tex root
Okay for chapters and parts
Working on Chapter 4, I want to have a first correct version by tomorrow.

@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 
-% First chapter begins here
-\chapter{Multi-Armed Bandits Models for Internet of Things Networks}
+\chapter{More Efficient Spectrum Usage of IoT Networks with Selfish MAB Learning}
 \label{chapter:4}
 \minitoc
+
+In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithm to learn and improve their access to the network.
+The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
+We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP \cite{USRPDocumentation} cards.
+
+More precisely, on the one hand, we start by presenting in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, used by static devices.
+Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their communication messages were successfully received.
+We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time.
+If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
+we show in simulations that in the case of non-uniform repartition of static devices in the channels, the dynamic devices can improve a lot the network efficiency.
+%
+On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real radio hardware.
+
+On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending the first one, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
+In this second model, we compare different heuristic, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
+
 \newpage
@@
 
-This chapter has three sections, each is a raw include from different papers
-
-- ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419
-
-- ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825
-
-- ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}
+\paragraph{Publications}
+This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 
 % ----------------------------------------------------------------------------
-\section{A first model of learning devices in an IoT network}
+\section{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
 \label{sec:4:firstModel}
@@
 % ----------------------------------------------------------------------------
-\section{Conclusion of Chapter 4}
+\section{Conclusion of Chapter~\ref{chapter:4}}
 \label{sec:4:conclusion}
@@
 
+\paragraph{Summary}
 In this chapter, we saw...
 
+\paragraph{Future works}
 Future works include...
@@
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter 4}
+\section{Appendix for Chapter~\ref{chapter:4}}
 \label{sec:4:appendix}

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Okay je réorganise et réécris pas mal pour ce chapitre 4.

@@
 % - ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419
+
+\graphicspath{{2-Chapters/4-Chapter/CrownCom_17.git/}}
 
@@
 
-% ------------------------------------------------------------------------
-\subsection{Motivation for this first IoT model}
-
-Unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
-This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.
-
-Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
-As end-devices battery life is a key constraint of IoT networks,
-this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
-%
-In this Section, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
-We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
-Kaufmann12Thompson}.
-% Both algorithms have already been applied with success, for Opportunistic Spectrum Access \cite{Jouini10} and recently for multi-users Cognitive Radio problems \cite{Toldov16}.
-
-MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
-For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
-Many recent works show that MAB algorithms work well for real-world radio signal.
-However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \emph{i.i.d} (independent and identically distributed).
-In recent works like \cite{Bonnefoi17}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d}.
-As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
+% % ------------------------------------------------------------------------
+% \subsection{Motivation for this first IoT model}
 
@@
 As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
-All devices share a synchronized time, and know in advance the finite number of available RF channels.
+All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
 In each time slot, devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (no sensing).
@@
     \centering
-    \includegraphics[scale=0.50]{2-Chapters/4-Chapter/CrownCom_17.git/protocol.eps}
+    \includegraphics[scale=0.45]{protocol.eps}
     \caption{The considered time-frequency slotted protocol.
@@
 Before explaining how this goal presents similarity with a \emph{multi-armed bandit problem}, we present some natural baseline policies (\ie, algorithms).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{System model - FIXME merge with previous section!}
+
+We consider the system model presented in Figure~\ref{fig:42:system_model1}, where a set of object sends uplink packets to the network gateway, in the $433.5\;\mathrm{MHz}$ ISM band.
+The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
+%
+The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$).
+In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s).
+
+These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks.
+This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.
+
+%To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
+%
+%By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
+%the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).
+
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.65\linewidth]{system_model1.eps}
+    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
+    \label{fig:42:system_model1}
+\end{figure}
+
+We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
+it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
+Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
+Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
+%
+For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
+The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
+$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
+as it is usually done in MAB problems \cite{Thompson33,Robbins52,LaiRobbins85}.
+
+This problem is a special case of the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$.
+Several types of reward distributions have been considered in the literature, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
+
+Rewards are binary in our model, and so we consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
+Contrary to many previous work done in the CR field (\emph{e.g.}, Opportunistic Spectrum Access),
+the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
+Rewards come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
+
+The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to the end-users, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
+%
+This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
 
@@
 Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem :
-\begin{subequations}
-\label{eq:41:prob}
-\begin{align}
+
+\begin{equation}
+\begin{subequations} \label{eq:41:prob}
+    \begin{cases}
     \underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
@@
     & D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
-\end{align}
+    \end{cases}
 \end{subequations}
+\end{equation}
 
@@
 \end{proposition}
+
 \begin{proof}
@@
 The strong duality condition is satisfied in this case \cite{BoydVanderberghe04}, so finding the saddle points will be enough to find the maximizers.
-    \hfill{}$\square$
+    \item
+    More details are given in Section~\ref{sec:5:proofLagrangian} in the Appendix of this Chapter.
+    % \hfill{}$\square$
     \end{itemize}
@@
 
-% A well-studied version of this problem is the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be independent and identically distributed (\emph{i.i.d}) under some distribution $\nu_k$, that has a mean $\mu_k$.
+% A well-studied version of this problem is the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be independent and identically distributed (\iid) under some distribution $\nu_k$, that has a mean $\mu_k$.
 Several types of reward distributions have been considered, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
@@
 Besides the modified time scale (rewards are no longer collected at every time step), this looks like a bandit problem.
-However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \emph{i.i.d}: they not only depend on the (stationary, \emph{i.i.d}) behavior of the static devices, but also on the behavior of other smart devices, that is not stationary (because of learning).
+However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other smart devices, that is not stationary (because of learning).
 
 Despite this, we show in the next subsection that running a stochastic bandit algorithm for each device based on its own rewards is surprisingly successful.
+
+\paragraph{Two algorithms}
+
+% \paragraph{The \UCB{} Algorithm}\label{sub:42:UCB}
+
+% A naive approach could be to use an empirical mean estimator of the rewards for each channel, and select the channel with highest estimated mean at each time;
+% but this greedy approach is known to fail dramatically \cite{LaiRobbins85}. Indeed, with this policy, the selection of arms is highly dependent on the first draws:
+% if the first transmission in one channel fails and the first one on other channels succeed, the end-user will \emph{never} use the first channel again, even it is the best one (\emph{i.e.}, the most available, in average).
+
+% Rather than relying on the empirical mean reward, Upper Confidence Bounds algorithms
+% instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm,
+% which can be viewed as adding a ``bonus'' exploration to the empirical mean.
+% % by adding a confidence term to the empirical mean.
+% They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model,
+% as the statistically best possible arm (\emph{i.e.}, the highest upper confidence bound) is selected.
+% % the arm with the highest upper confidence bound is expected to be the true best arm, and is played.
+
+% More formally, for one end-user, let $N_k(t) = \sum_{\tau=1}^t \mathbbm{1}(C(\tau) = k)$ be the number of times channel $k$ was selected up-to time $t \geq 1$.
+% The empirical mean estimator of channel $k$ is defined as the mean reward obtained by selecting it up to time $t$, $\widehat{\mu_k}(t) = \left(1 / N_k(t)\right) \sum_{\tau=1}^t r_k(\tau) \mathbbm{1}(C(\tau) = k) $.
+% For \UCB, the \emph{confidence} term is $B_k(t) = \sqrt{\alpha \log(t) / N_k(t)}$,
+% giving the upper confidence bound $U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, which is used by the end-user to decide the channel for communicating at time step $t+1$: $C(t+1) = \arg\max_{1\leq k \leq K} U_k(t)$.
+% \UCB{} is called an \emph{index policy}.
+
+The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
+% This algorithm is simple to implement and use in practice, even on embedded micro-processors with limited computation and memory capabilities.
+%
+% \todo[inline]{Maybe in the implementation of UCB the $t$ $\log(t)$ should be the total amount of time the user tried to communicate? It is already the case, but we should say so.}
+In our model, every dynamic end-user implements its own \UCB{} algorithm, \emph{independently}.
+For one end-user, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
+
+% \paragraph{Thompson Sampling}
+
+% Thompson Sampling \cite{Thompson33} was introduced early on, in $1933$ as the very first bandit algorithm, in the context of clinical trials (in which each arm models the efficacy of one treatment across patients).
+Given a prior distribution on the mean of each arm, the algorithm selects the next arm to draw based on samples from the \emph{conjugated} posterior distribution, which for Bernoulli rewards is a Beta distribution.
+
+% A Beta prior $\mathrm{Beta}(a_k(0)=1,b_k(0)=1)$ (initially uniform) is assumed on $\mu_k \in [0, 1]$, and at time $t$ the posterior is $\mathrm{Beta}(a_k(t),b_k(t))$.
+% After every channel selection, the posterior is updated to have $a_k(t)$ and $b_k(t)$ counting the number of successful and failed transmissions made on channel $k$.
+% So if the \emph{ACK} message is received, $a_k(t+1) = a_k(t) + 1$, and $b_k(t+1) = b_k(t)$, otherwise $a_k(t+1) = a_k(t)$, and $b_k(t+1) = b_k(t) + 1$.
+% Then, the decision is done by \emph{sampling} an \emph{index} for each arm, at each time step $t$, from the arm posteriors: $X_k(t) \sim \mathrm{Beta}(a_k(t), b_k(t))$, and the chosen channel is simply the channel $C(t+1)$ with highest index $X_k(t)$.
+For this reason, Thompson Sampling is called a \emph{randomized index policy}.
+
+The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
+It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
 
@@
 \paragraph{\emph{Adversarial} bandit algorithms?}
-% \todo[inline]{Explain that we also tried adversarial algorithms ($\mathrm{Exp}3$), and it was working but less efficiently than \UCB{} and Thompson Sampling. Maybe add it on some curve?}
+%
 Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
@@
 %(see Section \ref{sub:41:numericalResults}).
-Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary and non-\emph{i.i.d}, as its performance guarantee are true in \emph{any} setting.
+Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary and non-\iid, as its performance guarantee are true in \emph{any} setting.
 But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
@@
 
-\begin{figure*}[!t]
+\begin{figure}[!t]
     \centering
-    \subfloat[10\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/10intelligent.eps}
+    \subfloat[10\% of smart devices]{\includegraphics[scale=0.5]{10intelligent.eps}
     \label{fig:41:10intelligent}}
     \hfill
-    \subfloat[30\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/30intelligent.eps}
+    \subfloat[30\% of smart devices]{\includegraphics[scale=0.5]{30intelligent.eps}
     \label{fig:41:30intelligent}}
@@
     \vspace*{-20pt}
-    \subfloat[50\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/50intelligent.eps}
+    \subfloat[50\% of smart devices]{\includegraphics[scale=0.5]{50intelligent.eps}
     \label{fig:41:50intelligent}}
     \hfill
-    \subfloat[100\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/100intelligent.eps}
+    \subfloat[100\% of smart devices]{\includegraphics[scale=0.5]{100intelligent.eps}
     \label{fig:41:100intelligent}}
-    \caption{Performance of $2$ MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of smart end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
+    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of smart end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
     \label{fig:41:from10to100}
     \vspace*{-10pt}
-\end{figure*}
+\end{figure}
 
@@
     \centering
-    \includegraphics[scale=0.65]{2-Chapters/4-Chapter/CrownCom_17.git/perf_learning.eps}
+    \includegraphics[scale=0.65]{perf_learning.eps}
     \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices.}
@@
 
-We include another simulation, with a uniform repartition of static device, to check that learning does not help in this case, but does not decrease performance either.
+We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static device, to check that learning does not help in this case, but does not decrease performance either.
 % And we should include another example with a different scenarios, maybe $(0.3, 0.3, 0.3, 0.015, 0.015, 0.014, 0.014, 0.014, 0.014, 0.014)$ (R\'emi already did the simulations).
 
-\begin{figure*}[!t]
+\begin{figure}[!t]
     \centering
-    \subfloat[10\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_10.eps}
+    \subfloat[10\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_10.eps}
     \label{fig:41:ch2_10}}
     \hfill
-    \subfloat[30\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_30.eps}
+    \subfloat[30\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_30.eps}
     \label{fig:41:ch2_30}}
@@
     \vspace*{-20pt}
-    \subfloat[50\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_50.eps}
+    \subfloat[50\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_50.eps}
     \label{fig:41:ch2_50}}
     \hfill
-    \subfloat[100\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_100.eps}
+    \subfloat[100\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_100.eps}
     \label{fig:41:ch2_100}}
-    \caption{Performance of $2$ bandit algorithms, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
+    \caption{Performance of the UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
     \label{fig:41:figure4appendix}
-\end{figure*}
-
-\subsection{Conclusion about this first model}\label{sub:41:conclusion}
-
-In this Section, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
-with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
-Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
-We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the many ``intelligent'' end-devices.
-
-This is both a surprising and a very encouraging result, showing that application of bandit algorithms tailored for a stochastic model is still useful in broader settings.
-The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{can not} communicate with each others, and \emph{do not} know the system parameters.
-We will investigate this behavior in order to understand it better theoretically.
-We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
-
-Moreover, for sake of simplicity we supposed that all devices use the same standard.
-Our future work will consider more realistic interference scenarios and IoT networks, with, \emph{e.g.}, non-slotted time, more than one base station etc.
-
-\paragraph{Possible extensions of this model}
-%
-This work could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device.
-Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
-We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.
-
-Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \emph{e.g.}, one transmission a day.
-So additionally to deciding the channel for communication, each device has to also decide when to communicate.
-This is another direction of research, that we will investigate in the future.
+\end{figure}
 

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Okay je réorganise et réécris pas mal pour ce chapitre 4.

--- 2019-02-12 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
@@
 
-In this Section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$, implementing a proof-of-concept of the model previously presented in Section~\ref{sec:4:firstModel}.
+\graphicspath{{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/}}
+
+In this Section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept of the model previously presented in Section~\ref{sec:4:firstModel}.
 
@@
 %
-We model the network access as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
+Similarly to the previous section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
 %
@@
 
-\TODOL{Based on the publication, ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825}
+% \TODOL{TODO:
 
-\TODOL{TODO:
+% - Clean up notations, be sure it is OK with previous Sections
 
-- Clean up notations, be sure it is OK with previous Sections
+% - Be sure to not repeat any motivations / maths / explanations given before
 
-- Be sure to not repeat any motivations / maths / explanations given before
-
-- Be sure to give enough details about the conclusions etc
-}
+% - Be sure to give enough details about the conclusions etc
+% }
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Motivation for this demonstration}
+\label{sub:42:motivation}
 
-\subsection{Motivation for this demonstration}
-
-% The monitoring of large scale systems, such as smart grids and smart cities, requires the development of networks dedicated to Internet-of-Things (IoT) applications.
-% For instance, Low Power Wide Area Networks (LPWAN) \cite{Raza17}, as LoRaWAN or SigFox, are nowadays deployed in unlicensed bands to handle a large number of objects transmitting a few packets per day or week.
-% In order to reduce the energy consumption of end-devices, these networks rely on pure ALOHA-based Medium Access (MAC) protocols.
-
-% One of the challenges in the design of MAC solutions for the IoT is to design solutions which improve the performance of the network and reduce the Packet Loss Ratio (PLR), without reducing the end-devices battery life.
-% In particular, many IoT standards operate in unlicensed bands, that is why we have to find solutions that do not increase the PLR due to the interference caused by other standards and networks which share the same band, without coordination.
-% As this interfering traffic is generated by different standards and networks, it cannot be controlled, and it is not evenly distributed in the different channels.
-
-Multi-Armed Bandit (MAB) algorithms have been recently proposed as a solution to improve the performance of IoT networks and in particular in LPWAN \cite{Bonnefoi18,Azari18}.
-In this paper, we describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol (such as the ones employed in LPWAN).
+In this Section, we describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
+This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
 %
-%\todo[inline]{Contributions et interets de la demo}
-% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini}.
-% Recently, TS and UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov}, and the UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi}.
-% Many recent works show that MAB algorithms work well for real-world radio signal.
-% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other objects is never really stationary or \emph{i.i.d.} (independent and identically distributed).
-% In recent works like \cite{Bonnefoi}, several objects are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d.}.
-% As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent objects in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
-%
-This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
-% % even when the number of intelligent objects in the network increases, and the stochastic hypothesis is more and more questionable.
-% To do that, we suppose an IoT network made of two types of objects: static objects that use only one channel (fixed in time), and dynamic objects that can choose the channel for each of their transmissions.
-Static objects form an interfering traffic, which could have been generated by objects using other standards as well.
-% We first evaluate the probability of collision if dynamic objects randomly select channels (naive approach), and if a centralized controller optimally distribute them in channels (ideal approach).
-% Then, these reference scenarios allow to evaluate the performance of UCB and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
-% We show that these algorithms have near-optimal performance, even when the proportion of end-objects increases and the interfering traffic from other objects becomes less and less stochastic.
-% \todo[inline]{(Remi) Si on parle de non stationarity, il faut donner un resultat a deux utilisateurs.}
-Following our recent work presented in the previous Section~\ref{sec:4:firstModel}, we propose to model this problem as Non-Stationary\footnote{~Note that non-stationarity only comes from the presence of more than one dynamic object, as the background traffic is assumed independent and identically distributed \emph{i.i.d.}.} MAB.
-We suggest to use low-cost algorithms, focusing on two well-known algorithms: a frequentist one (\UCB{}) and a Bayesian one (TS).
-We consider the Upper-Confidence Bound (\UCB{}) \cite{Auer}, and the Thompson Sampling (TS) algorithms \cite{Thompson33}.
-Both algorithms have already been applied with success in the context of wireless decision making, both empirically for Opportunistic Spectrum Access \cite{Jouini},
-and more recently for multi-users Cognitive Radio problems \cite{BessonALT18} with a more theoretical approach.
-
-We use a TestBed designed in 2017 by our team SCEE \cite{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
+We use a TestBed designed in 2017 by our team SCEE \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
 and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.
@@
 
-The rest of this paper is organized as follows.
-The system model is introduced in Section 2.
-In Section 3, we describe more formally both the \UCB{} and the TS algorithms.
-Our implementation is presented in Section 4, and results are given in Section 5.
+We refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
+which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Section~\ref{sec:42:results}).
+Then we discuss the relevance of a MAB model for our IoT application.
+
+The rest of this Section is organized as follows.
+% The system model is introduced in Section 2. In Section 3, we describe more formally both the \UCB{} and the TS algorithms.
+Our implementation is presented in Section~\ref{sub:42:implementation}, with details about the user interface, and results are given in Section~\ref{sub:42:results}.
 % and Section 6 concludes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{System model - FIXME merge with previous section!}
+\subsection{GNU Radio implementation}
+\label{sub:42:implementation}
 
-We consider the system model presented in Figure~\ref{fig:42:system_model1}, where a set of object sends uplink packets to the network gateway, in the $433.5\;\mathrm{MHz}$ ISM band.
-The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
-%
-The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$).
-In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s).
-
-These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks.
-This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.
-
-%To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
-%
-%By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
-%the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/system_model1.eps}
-    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
-    \label{fig:42:system_model1}
-\end{figure}
-
-We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
-it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
-Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
-Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
-%
-For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
-The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
-$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
-as it is usually done in MAB problems \cite{Thompson33,Robbins52,LaiRobbins85}.
-
-This problem is a special case of the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$.
-Several types of reward distributions have been considered in the literature, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
-
-Rewards are binary in our model, and so we consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
-Contrary to many previous work done in the CR field (\emph{e.g.}, Opportunistic Spectrum Access),
-the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
-Rewards come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
-
-The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to the end-users, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
-%
-This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
-
-% In a Cognitive Radio application, arms model the \emph{channels}, and players are the \emph{dynamic end-devices}.
-% For example in the classical OSA setting with sensing \cite{Jouini}, a single dynamic device (a player) sequentially tries to access channels (the arms), and collects a reward of 1 if the channel is available %(as there is no primary user, we can assume the transmission to be successful),
-% and 0 otherwise. So rewards represent the \emph{availability} of channels, and
-% the parameter $\mu_k$ represents the mean availability of channel $k$.
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsection{MAB Algorithms}
-
-Before discussing the relevance of a MAB model for our IoT application, we refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
-which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Sec.~\ref{sec:42:results}).
-
-% \paragraph{The \UCB{} Algorithm}\label{sub:42:UCB}
-
-% A naive approach could be to use an empirical mean estimator of the rewards for each channel, and select the channel with highest estimated mean at each time;
-% but this greedy approach is known to fail dramatically \cite{LaiRobbins85}. Indeed, with this policy, the selection of arms is highly dependent on the first draws:
-% if the first transmission in one channel fails and the first one on other channels succeed, the end-user will \emph{never} use the first channel again, even it is the best one (\emph{i.e.}, the most available, in average).
-
-% Rather than relying on the empirical mean reward, Upper Confidence Bounds algorithms
-% instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm,
-% which can be viewed as adding a ``bonus'' exploration to the empirical mean.
-% % by adding a confidence term to the empirical mean.
-% They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model,
-% as the statistically best possible arm (\emph{i.e.}, the highest upper confidence bound) is selected.
-% % the arm with the highest upper confidence bound is expected to be the true best arm, and is played.
-
-% More formally, for one end-user, let $N_k(t) = \sum_{\tau=1}^t \mathbbm{1}(C(\tau) = k)$ be the number of times channel $k$ was selected up-to time $t \geq 1$.
-% The empirical mean estimator of channel $k$ is defined as the mean reward obtained by selecting it up to time $t$, $\widehat{\mu_k}(t) = \left(1 / N_k(t)\right) \sum_{\tau=1}^t r_k(\tau) \mathbbm{1}(C(\tau) = k) $.
-% For \UCB, the \emph{confidence} term is $B_k(t) = \sqrt{\alpha \log(t) / N_k(t)}$,
-% giving the upper confidence bound $U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, which is used by the end-user to decide the channel for communicating at time step $t+1$: $C(t+1) = \arg\max_{1\leq k \leq K} U_k(t)$.
-% \UCB{} is called an \emph{index policy}.
-
-The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{bubeck2012regret}.
-% This algorithm is simple to implement and use in practice, even on embedded micro-processors with limited computation and memory capabilities.
-%
-% \todo[inline]{Maybe in the implementation of UCB the $t$ $\log(t)$ should be the total amount of time the user tried to communicate? It is already the case, but we should say so.}
-In our model, every dynamic end-user implements its own \UCB{} algorithm, \emph{independently}.
-For one end-user, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
-
-% \paragraph{Thompson Sampling}
-
-% Thompson Sampling \cite{Thompson33} was introduced early on, in $1933$ as the very first bandit algorithm, in the context of clinical trials (in which each arm models the efficacy of one treatment across patients).
-Given a prior distribution on the mean of each arm, the algorithm selects the next arm to draw based on samples from the \emph{conjugated} posterior distribution, which for Bernoulli rewards is a Beta distribution.
-
-% A Beta prior $\mathrm{Beta}(a_k(0)=1,b_k(0)=1)$ (initially uniform) is assumed on $\mu_k \in [0, 1]$, and at time $t$ the posterior is $\mathrm{Beta}(a_k(t),b_k(t))$.
-% After every channel selection, the posterior is updated to have $a_k(t)$ and $b_k(t)$ counting the number of successful and failed transmissions made on channel $k$.
-% So if the \emph{ACK} message is received, $a_k(t+1) = a_k(t) + 1$, and $b_k(t+1) = b_k(t)$, otherwise $a_k(t+1) = a_k(t)$, and $b_k(t+1) = b_k(t) + 1$.
-% Then, the decision is done by \emph{sampling} an \emph{index} for each arm, at each time step $t$, from the arm posteriors: $X_k(t) \sim \mathrm{Beta}(a_k(t), b_k(t))$, and the chosen channel is simply the channel $C(t+1)$ with highest index $X_k(t)$.
-For this reason, Thompson Sampling is called a \emph{randomized index policy}.
-
-The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12}.
-It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov,Mitton}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{GNU Radio implementation}
-
-In this section, we present our implementation of MAB algorithms in our model of IoT networks.
-We first describe the simplified physical layer of this demo,
-then we present our GNU Radio implementation.
+In this section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
+We first describe the simplified physical layer of this demo, then we present our GNU Radio implementation.
 
@@
 
-In this paper, we implement a PHY/MAC layers solution in order to improve the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
+We implement a PHY/MAC layers solution in order to improve the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
 We choose to implement our own physical layer and protocol, for both clarity and conciseness.
@@
 \paragraph{Equipment}
-
 % Boards and versions
@@
 % Cables and laptop
+% Osef
 
@@
     \centering
-    \includegraphics[width=0.70\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/our-demo.eps}
+    \includegraphics[width=0.70\linewidth]{our-demo.eps}
     \caption{Schematic of our implementation that presents the role of each USRP card.}
@@
     \centering
-    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/SCEE_TestBed1.jpg}
+    \includegraphics[width=0.65\linewidth]{SCEE_TestBed1.jpg}
     \vspace*{20pt}
-    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/SCEE_TestBed2.jpg}
-    \caption{Photos of the SCEE testbed.}
+    \includegraphics[width=0.65\linewidth]{SCEE_TestBed2.jpg}
+    \caption{Two pictures showing the SCEE testbed \cite[Appendix~3]{Bodinier17}.}
     \label{fig:42:photosSCEETestBed}
@@
     \centering
-    \includegraphics[width=0.80\textwidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/UI.eps}
+    \includegraphics[width=0.80\textwidth]{UI.eps}
     \caption{User interface of our demonstration.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Results and discussions}\label{sec:42:results}
+\subsection{Numerical results}
+\label{sec:42:results}
 
@@
 Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
-Empirical results confirm the simulations presented in our paper \cite[Fig.3]{Bonnefoi17}.
+Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
+%  \cite[Fig.3]{Bonnefoi17}.
 Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.
@@
 	\centering
-    \includegraphics[height=8.0cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
+    \includegraphics[height=8.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
     \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success.
@@
 %     \begin{subfigure}[b]{0.46\textwidth}
-%         \includegraphics[height=7cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/Object2_UniformAccess__UI.png}
+%         \includegraphics[height=7cm]{Object2_UniformAccess__UI.png}
 %         %\caption{Object using uniform access.}
@@
 %     \begin{subfigure}[b]{0.46\textwidth}
-%         \includegraphics[height=7cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/Object1_UCB__UI.png}
+%         \includegraphics[height=7cm]{Object1_UCB__UI.png}
 %         %\caption{Object using UCB.}
@@
 
-We presented in this Section a demonstration, showed in $2018$ at the ICT conference,
-by specifying the system model and explaining the two MAB algorithms used in practice.
-We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
-Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
-%
-By using such low-cost algorithms, we demonstrated empirically that a dynamically reconfigurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.
-
-Possible future extensions of this work include:
-considering more dynamic objects (\emph{e.g.}, $100$),
-implementing a real-world IoT communication protocol (like the LoRaWAN standard),
-and studying the interference in case of other gateways located nearby.
-%
-% Pub pour notre autre article
-We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.
-
 \paragraph{Availability of data and materials}
 %
-The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at\\
-\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
+The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at
+\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/\\malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
 It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.
@@
 It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
-It shows the connections between the USRP boards, the Octoclock, and the master laptop.
+It especially shows the connections between the USRP boards, the Octoclock, and the master laptop.
 
@@
 as we took inspiration in the GNU Radio code and the discussions we had at this time.
-
-% ---------- Bibliography ----------
-%\begin{small} % XXX small if needed?
-\bibliographystyle{ieeetr}
-\bibliography{biblio_RIoT}
-%\end{small} % XXX small if needed?

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
Okay je réorganise et réécris pas mal pour ce chapitre 4.

@@
 
+% ----------------------------------------------------------------------------
+\section{Motivations for Selfish MAB Learning for IoT Networks}
+\label{sec:4:motivations}
+
+Unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
+This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.
+
+Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
+As end-devices battery life is a key constraint of IoT networks,
+this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
+%
+In this Section, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
+We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
+Kaufmann12Thompson}.
+
+MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
+For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
+Many recent works show that MAB algorithms work well for real-world radio signal.
+However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
+In recent works like \cite{Bonnefoi17}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \iid.
+As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
+
+We also present an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
+
 \TODOL{Present outline of the chapter.}
 
-\TODOL{Give list of publications associated with this chapter.}
-
+% \TODOL{Give list of publications associated with this chapter.}
 \paragraph{Publications}
@@
 
-\paragraph{Summary}
-In this chapter, we saw...
+\subsection{Summary of the first model}\label{sub:41:conclusion}
 
-\paragraph{Future works}
+In Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
+with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
+Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
+We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the many ``intelligent'' end-devices.
+
+This is both a surprising and a very encouraging result, showing that application of bandit algorithms tailored for a stochastic model is still useful in broader settings.
+The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{can not} communicate with each others, and \emph{do not} know the system parameters.
+We will investigate this behavior in order to understand it better theoretically.
+We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
+
+\subsection{Conclusions taken from the demonstration}
+\label{sec:42:conclusionFromDemonstration}
+
+We presented in Section~\ref{sec:4:gnuradio} a demonstration, showed in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
+We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
+Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
+%
+By using such low-cost algorithms, we demonstrated empirically that a dynamically reconfigurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.
+
+\subsection{Future works}
 Future works include...
+
+\paragraph{Possible extensions of the first model}
+%
+The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device.
+Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
+We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.
+
+Moreover, for sake of simplicity we supposed that all devices use the same standard.
+Our future work will consider more realistic interference scenarios and IoT networks, with, \emph{e.g.}, non-slotted time, more than one base station etc.
+
+Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \emph{e.g.}, one transmission a day.
+So additionally to deciding the channel for communication, each device has to also decide when to communicate.
+This is another direction of research, that we will investigate in the future.
+
+\paragraph{Extensions of the demonstration}
+Possible future extensions of our demonstration include the following points.
+We could consider more dynamic objects (\emph{e.g.}, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
+We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
+Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing new hardware.
+%
+% Pub pour notre autre article
+% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.
 

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Trying to finish chapter 4 as quickly as possible.

@@
 
-\TODOL{TODO:
-
-- Clean up notations, be sure it is OK with previous Sections
-
-- Be sure to not repeat any motivations / maths / explanations given before
-
-- Be sure to give enough details about the conclusions etc
-}
-
-% % ------------------------------------------------------------------------
-% \subsection{Motivation for this first IoT model}
-
 The aim of this Section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases, and the stochastic hypothesis is more and more questionable.
@@
 
+\paragraph{Outline}
 The rest of this Section is organized as follows.
@@
 
+We present our system model, which consists of one gateway and many IoT devices, using a slotted protocol.
+
+\paragraph{One gateway and many devices}
+%
+We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of object sends uplink packets to the network gateway, for instance in the $433.5\;\mathrm{MHz}$ ISM band (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
+The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
+%
+The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$).
+In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s in Section~\ref{sec:4:gnuradio}).
+
+These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks.
+This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.
+%
+To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
+%
+By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
+the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).
+
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.65\linewidth]{system_model1.eps}
+    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
+    \label{fig:41:system_model1}
+\end{figure}
+
+\paragraph{Slotted protocol}
+%
 As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
@@
 
+\paragraph{Static vs dynamic devices}
+%
 There are two types of end-devices in the network:
@@
 As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
-We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks\footnote{Here, $K$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
+We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks\footnote{~Here, $K$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
 
+\paragraph{Random emission patterns}
+%
 We suppose that all devices follow the same emission pattern, being fixed in time, and we choose to model it as a simple Bernoulli process:
-all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $K / (S + D)$ for all devices to communicate successfully (in average).}.
+all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{~In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $K / (S + D)$ for all devices to communicate successfully (in average).}.
 The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed (\ie, real time during two messages), and $1/p$ is proportional to the \emph{duty cycle}.
@@
 
-The goal is to design a simple sequential algorithm, to be applied identically by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
-
-Before explaining how this goal presents similarity with a \emph{multi-armed bandit problem}, we present some natural baseline policies (\ie, algorithms).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{System model - FIXME merge with previous section!}
-
-We consider the system model presented in Figure~\ref{fig:42:system_model1}, where a set of object sends uplink packets to the network gateway, in the $433.5\;\mathrm{MHz}$ ISM band.
-The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
+\paragraph{MAB model}
 %
-The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$).
-In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s).
-
-These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks.
-This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.
-
-%To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
-%
-%By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
-%the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=0.65\linewidth]{system_model1.eps}
-    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
-    \label{fig:42:system_model1}
-\end{figure}
-
 We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
@@
 $r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
-as it is usually done in MAB problems \cite{Thompson33,Robbins52,LaiRobbins85}.
+as it is usually done in MAB problems.
 
-This problem is a special case of the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$.
-Several types of reward distributions have been considered in the literature, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
-
-Rewards are binary in our model, and so we consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
+This problem is a special case of the so-called stochastic MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$.
+Several types of reward distributions have been considered in the literature, and rewards are binary in our model.
+We consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
 Contrary to many previous work done in the CR field (\emph{e.g.}, Opportunistic Spectrum Access),
@@
 This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
+
+The goal is thus to design a simple sequential algorithm, to be applied identically by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
+
+Before explaining how this goal presents similarity with a \emph{multi-armed bandit problem}, we present some natural baseline policies (\ie, algorithms).
 
@@
 We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; K \rrbracket = \{1, \dots, K\}$).
-% every time they communicate
-% This reflects a very naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.
+This reflects a very naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.
 
@@
 
-We investigate in this section the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
+We investigate in this section the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
 We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.
@@
 
-\begin{equation}
-\begin{subequations} \label{eq:41:prob}
-    \begin{cases}
+\begin{subequations}\label{eq:41:prob}
+\begin{align}
+    % \begin{cases}
     \underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
@@
     & D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
-    \end{cases}
+    % \end{cases}
+\end{align}
 \end{subequations}
-\end{equation}
 
@@
 For such values for $D_i$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
-    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
+    \footnote{~It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
     \item
@@
 Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
-Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded to find the optimal number of devices for channel $i$ :
-% Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
+Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $i$ :
+%
 $\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < K$, and $\widehat{D_{K}} = D - \sum_{i=1}^{K - 1} \widehat{D_i}$.
@@
 
-\subsection{Sequential policies based on bandit algorithms}\label{sub:41:sequentialPolicies}
+% ----------------------------------------------------------------------
+\subsection{Sequential policies based on bandit algorithms}
+\label{sub:41:sequentialPolicies}
+% ----------------------------------------------------------------------
 
-% We now present the stochastic Multi-Armed Bandit (MAB) model,
-% and the two stochastic MAB algorithms used in our experiments \cite{Bubeck12}.
 While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused by the learning policy used by dynamic objects.
-%Quite surprisingly, we will show in our experiments of Section~\ref{sub:41:numericalResults} that two algorithms inspired by the stochastic bandit modeling, UCB1 and Thompson Sampling, perform well even for coordinating multiple smart devices.
-
-%This section presents the formalism of the Multi-Armed Bandit (MAB) problem, as introduced by \cite{LaiRobbins85}, and two classical algorithms designed to tackle its simplest version (stochastic MAB), \UCB{} from \cite{Auer02} and Thompson Sampling from \cite{Thompson33,Kaufmann12Thompson}.
-%For more details, \cite{Bubeck12} is an excellent review.
-
-% \paragraph{Stochastic Multi-Armed Bandits}
-
-% A Multi-Armed Bandit problem is defined as follows \cite{Thompson33,Robbins52,LaiRobbins85}.
-% There is a fixed number $K \geq 1$ of levers, or ``arms'', and a player has to choose one lever at each discrete time $t \geq 1, t \in \mathbb{N}$, denoted as $A(t) = k \in\{1,\dots,K\}$.
-% Selecting arm $k$ at time $t$ yields a (random) \emph{reward}, $r_k(t) \in \mathbb{R}$, and the goal of the player is to maximize the sum of his rewards, $r_{1 \dots T} = \sum_{t = 1}^T r_{A(t)}(t)$.
-
-% A well-studied version of this problem is the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be independent and identically distributed (\iid) under some distribution $\nu_k$, that has a mean $\mu_k$.
-Several types of reward distributions have been considered, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
-% We consider Bernoulli bandit models, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k$.
-
-% The problem parameters $\mu_1,\dots,\mu_K$ are unknown to the player, so to maximize his cumulated rewards, he must learn the distributions of the channels, to be able to progressively focus on the best arm (\ie, the arm with largest mean).
-% %
-% This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
-
-% In a Cognitive Radio application, arms model the \emph{channels}, and players are the \emph{dynamic end-devices}.
-% For example in the classical OSA setting with sensing \cite{Jouini10}, a single dynamic device (a player) sequentially tries to access channels (the arms), and collects a reward of 1 if the channel is available %(as there is no primary user, we can assume the transmission to be successful),
-% and 0 otherwise. So rewards represent the \emph{availability} of channels, and
-% the parameter $\mu_k$ represents the mean availability of channel $k$.
 
@@
 \paragraph{A bandit model for IoT}
-%
+
 Our IoT application is challenging in that there are \emph{multiple} players (the dynamic devices) interacting with the \emph{same} arms (the channels), without any centralized communication (they do not even know the total number of dynamic devices).
@@
 
-% \paragraph{The \UCB{} Algorithm}\label{sub:42:UCB}
-
-% A naive approach could be to use an empirical mean estimator of the rewards for each channel, and select the channel with highest estimated mean at each time;
-% but this greedy approach is known to fail dramatically \cite{LaiRobbins85}. Indeed, with this policy, the selection of arms is highly dependent on the first draws:
-% if the first transmission in one channel fails and the first one on other channels succeed, the end-user will \emph{never} use the first channel again, even it is the best one (\emph{i.e.}, the most available, in average).
-
-% Rather than relying on the empirical mean reward, Upper Confidence Bounds algorithms
-% instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm,
-% which can be viewed as adding a ``bonus'' exploration to the empirical mean.
-% % by adding a confidence term to the empirical mean.
-% They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model,
-% as the statistically best possible arm (\emph{i.e.}, the highest upper confidence bound) is selected.
-% % the arm with the highest upper confidence bound is expected to be the true best arm, and is played.
-
-% More formally, for one end-user, let $N_k(t) = \sum_{\tau=1}^t \mathbbm{1}(C(\tau) = k)$ be the number of times channel $k$ was selected up-to time $t \geq 1$.
-% The empirical mean estimator of channel $k$ is defined as the mean reward obtained by selecting it up to time $t$, $\widehat{\mu_k}(t) = \left(1 / N_k(t)\right) \sum_{\tau=1}^t r_k(\tau) \mathbbm{1}(C(\tau) = k) $.
-% For \UCB, the \emph{confidence} term is $B_k(t) = \sqrt{\alpha \log(t) / N_k(t)}$,
-% giving the upper confidence bound $U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, which is used by the end-user to decide the channel for communicating at time step $t+1$: $C(t+1) = \arg\max_{1\leq k \leq K} U_k(t)$.
-% \UCB{} is called an \emph{index policy}.
-
 The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
-% This algorithm is simple to implement and use in practice, even on embedded micro-processors with limited computation and memory capabilities.
 %
-% \todo[inline]{Maybe in the implementation of UCB the $t$ $\log(t)$ should be the total amount of time the user tried to communicate? It is already the case, but we should say so.}
 In our model, every dynamic end-user implements its own \UCB{} algorithm, \emph{independently}.
 For one end-user, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
-
-% \paragraph{Thompson Sampling}
-
-% Thompson Sampling \cite{Thompson33} was introduced early on, in $1933$ as the very first bandit algorithm, in the context of clinical trials (in which each arm models the efficacy of one treatment across patients).
-Given a prior distribution on the mean of each arm, the algorithm selects the next arm to draw based on samples from the \emph{conjugated} posterior distribution, which for Bernoulli rewards is a Beta distribution.
-
-% A Beta prior $\mathrm{Beta}(a_k(0)=1,b_k(0)=1)$ (initially uniform) is assumed on $\mu_k \in [0, 1]$, and at time $t$ the posterior is $\mathrm{Beta}(a_k(t),b_k(t))$.
-% After every channel selection, the posterior is updated to have $a_k(t)$ and $b_k(t)$ counting the number of successful and failed transmissions made on channel $k$.
-% So if the \emph{ACK} message is received, $a_k(t+1) = a_k(t) + 1$, and $b_k(t+1) = b_k(t)$, otherwise $a_k(t+1) = a_k(t)$, and $b_k(t+1) = b_k(t) + 1$.
-% Then, the decision is done by \emph{sampling} an \emph{index} for each arm, at each time step $t$, from the arm posteriors: $X_k(t) \sim \mathrm{Beta}(a_k(t), b_k(t))$, and the chosen channel is simply the channel $C(t+1)$ with highest index $X_k(t)$.
-For this reason, Thompson Sampling is called a \emph{randomized index policy}.
 
@@
 \paragraph{Multi-Player MAB with \emph{collision avoidance}?}
-%
+
 Another idea could be to try to use a \emph{multi-player MAB} model, as proposed by \cite{Zhao10}, to describe our problem.
@@
 
-%The stochastic MAB model fits well for OSA for Cognitive Radio,
-%but for IoT networks, a notable difference with the MAB problem introduced above is that there is not only one device adaptively using the network, but there are $D \geq 1$ dynamic end-devices (and possibly only dynamic devices).
-%So in our model of IoT network, the stochastic \iid{} hypothesis for the rewards makes less sense: the (stationary) background traffic interferes with the end-devices transmissions but also the end-devices interfere with each others.
-To the stochastic static traffic is added a ``game-theoretical'' interaction between dynamic devices (who, let us remind it, cannot communicate with each others).
-%Intuitively, the more dynamic devices, the less stationary the network will be.
-%The presence of a lot of dynamic devices implies that each should learn both the (stationary) distribution of static devices (\ie, the $\mu_k$), and learn to avoid collisions with other devices.
+\paragraph{\emph{Adversarial} bandit algorithms?}
 
-\paragraph{\emph{Adversarial} bandit algorithms?}
-%
 Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
@@
 But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
-%\footnote{We preferred not to talk about \emph{regret} in this study, but in a few words: it is a measure of how \emph{bad} the algorithm performed in terms of its accumulated rewards, in comparison to the best possible policy, and should be as small as possible.
+%\footnote{~We preferred not to talk about \emph{regret} in this study, but in a few words: it is a measure of how \emph{bad} the algorithm performed in terms of its accumulated rewards, in comparison to the best possible policy, and should be as small as possible.
 For stochastic algorithms, being ``efficient'' means having a regret bounded as $R_T = \mathcal{O}(\log T)$, but for adversarial algorithms, it means having $R_T = \mathcal{O}(\sqrt{K T})$.}
@@
 \subsection{Experiments and numerical results}\label{sub:41:numericalResults}
+
+% Simulation parameters
+We suppose a network with $S + D = 2000$ end-devices, and one IoT base station.
+Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\emph{e.g.}, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
+The RF band is divided in $K = 10$ channels.
+Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{K}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
+The dynamic devices have access to all the channels, and use learning algorithms.
+%to find the least loaded.
+%
+We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
+We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions.
+Clearly, the problem is less interesting for homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
 
@@
 
-% Simulation parameters
-We suppose a network with $S + D = 2000$ end-devices, and one IoT base station.
-Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\emph{e.g.}, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
-The RF band is divided in $K = 10$ channels.
-Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{K}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
-The dynamic devices have access to all the channels, and use learning algorithms.
-%to find the least loaded.
-%
-We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
-We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions.
-Clearly, the problem is less interesting for homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Fig.\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
-
-Figure \ref{fig:41:from10to100} presents the evolution of the successful transmission rate, as a function of time. %($x$ axis is labeled by hundreds of transmissions). % I will change the x axis
+Figure \ref{fig:41:from10to100} presents the evolution of the successful transmission rate, as a function of time.
 The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy from below, and the two oracle policies (optimal and greedy) from above.
@@
 
-\paragraph{Another simulation results}
+\paragraph{A sanity check}
 
 We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static device, to check that learning does not help in this case, but does not decrease performance either.
-% And we should include another example with a different scenarios, maybe $(0.3, 0.3, 0.3, 0.015, 0.015, 0.014, 0.014, 0.014, 0.014, 0.014)$ (R\'emi already did the simulations).
 
@@
 \paragraph{Note on the simulation code}
-%
+
 The simulation code used for the experiments in Section~\ref{sub:41:numericalResults} is for MATLAB or GNU Octave,
@@
 \href{https://Bitbucket.org/scee_ietr/rl_slotted_iot_networks}{\texttt{Bitbucket.org/scee\_ietr/rl\_slotted\_iot\_networks}}.
-
-% % ---------- Bibliography ----------
-% %\begin{small}
-% \bibliographystyle{ieeetr}
-% \bibliography{biblio_RIoT}
-% %\end{small}

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Trying to finish chapter 4 as quickly as possible.

--- 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
@@
 % ----------------------------------------------------------------------
-\subsection{Retransmissions in an ALOHA like protocol}\label{sub:43:introduction}
+\subsection{Retransmissions in an ALOHA like protocol}
+\label{sub:43:introduction}
+% ----------------------------------------------------------------------
 
@@
 
-% \vspace*{5pt}
-The rest of the section is organized as follows.
+\paragraph{Outline}
+
+The rest of the Section is organized as follows.
 First the system model is introduced in Section~\ref{sub:43:model},
@@
 % ----------------------------------------------------------------------
-\subsection{System model}\label{sub:43:model}
+\subsection{Extending our model with retransmissions}
+\label{sub:43:model}
+% ----------------------------------------------------------------------
 
@@
 % ----------------------------------------------------------------------
-\subsection{Motivations for the proposed approach} \label{sub:43:motivations}
+\subsection{Motivations for the proposed approach}
+\label{sub:43:motivations}
+% ----------------------------------------------------------------------
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Behavior analysis of $p_{c}$ and $p_{c1}$}\label{sub:numericalValidationPC1PC}
+\paragraph{Behavior analysis of $p_{c}$ and $p_{c1}$}\label{sub:43:numericalValidationPC1PC}
 
@@
 
+\TODOL{Remove this part or compress}
+
 %-----------------------------------------------------------------
-\subsection{A well-known MAB Algorithm: \UCB{}}\label{sub:43:MABalgo}
+\subsection{The \UCB{} algorithm as a building block for different heuristics}
+\label{sub:43:MABalgo}
+% ----------------------------------------------------------------------
 
-Without loss of generality, we have adopted a well-studied stochastic MAB learning algorithm, where the reward distributions are unknown and assumed to be independent and identically distributed (i.i.d).
-The arms model the channels denoted as $C(t) \in \llbracket 1, K \rrbracket$, and the players, the dynamic devices, learn the distributions to be able to progressively focus on the best arm, \ie, the arm with largest mean representing the mean availability of a given channel $k$.
-
-Before presenting our proposed heuristics, we describe a \UCB{} bandit algorithm \cite{Auer02}.
-It has reported to be efficient, while featuring a low complexity for its implementation.
-For this reason, it has been employed for IoT applications \cite{Bonnefoi17}, and we employ this approach to develop our proposals.
+Before presenting our proposed heuristics, we remind here the \UCB{} bandit algorithm \cite{Auer02}.
+We include it again to facilitate the understanding of the heuristics, which use \UCB{} as a building block.
 
 % % --- --- --- --- --- --- --- --- --- ---
-\paragraph{The \UCB{} algorithm}\label{sub:algoUCB}
-
-A first approach is to only use an empirical mean estimator of the rewards in every channel, and select the channel with highest estimated mean at every time step; but this greedy approach is known to fail dramatically \cite{Auer02}.
-Indeed, with this policy, the selection of arms depends too much on the first draws: if the first transmission in one channel fails and the first one on other channels succeeds, the device will \emph{never} use the first channel again, even if it is the best one (\ie, the most available, in average).
-
-Rather than relying on the empirical mean reward, \UCB{} algorithms instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm, which can be viewed as adding a ``bonus'' exploration to the empirical mean.
-They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model, as the statistically best possible arm (\ie, the highest \UCB{}) is selected.
+\paragraph{The \UCB{} algorithm}\label{sub:43:algoUCB}
 
@@
 % ----------------------------------------------------------------------
-\subsection{Heuristics to (try to) learn how to retransmit efficiently}\label{sub:43:heuristics}
+\subsection{Heuristics to (try to) learn how to retransmit efficiently}
+\label{sub:43:heuristics}
+% ----------------------------------------------------------------------
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Uniform random retransmission (``Random'')}\label{sub:UCBthenRandom}
+\paragraph{Uniform random retransmission (``Random'')}\label{sub:43:UCBthenRandom}
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{\UCB{} for retransmission (``Only \UCB{}'')}\label{sub:TwoUCB}
+\paragraph{\UCB{} for retransmission (``Only \UCB{}'')}\label{sub:43:TwoUCB}
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{$K$ different {\UCB}s for retransmission (``$K$ \UCB{}'')}\label{sub:UCBthenKp1}
+\paragraph{$K$ different {\UCB}s for retransmission (``$K$ \UCB{}'')}\label{sub:43:UCBthenKp1}
 
@@
 the memory to store $K+1=5$ algorithms is of the order of the requirements to storing one.
-% For other networks this heuristic could not be practical, for instance  Sigfox consist in a large number of very narrow-band channels.
 
-%\footnote{~The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels.
-But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.}
+Note that for other networks this heuristic could not be practical, for instance Sigfox consist in a large number of very narrow-band channels.
+The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels.
+But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'')}\label{sub:UCBwithDelay}
+\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'')}\label{sub:43:UCBwithDelay}
 
@@
 The idea behind this delay is to allow the first stage \UCB{} to start learning the best channel, before starting the second stage \UCB{} (see details in Algorithm~\ref{algo:43:UCBwithDelay}).
-The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand.
+The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to be able to simulate the problem in advance in order to find a good value of some internal parameter.
+As such, we only consider a delay of $\Delta=100$.}.
 
@@
 related to the delayed second-stage \UCB{} algorithm.
-
-%\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to be able to simulate the problem in advance in order to find a good value of some internal parameter.
-As such, we only consider a delay of $\Delta=100$.}.
 
@@
 % ----------------------------------------------------------------------
-\subsection{Simulations to compare our heuristics}\label{sub:43:numExp}
+\subsection{Simulations to compare our heuristics}
+\label{sub:43:numExp}
+% ----------------------------------------------------------------------
 
@@
 
-% ----------------------------------------------------------------------
-\subsection{Conclusion about this second model}\label{sub:43:conclusion}
-
-In this section, we presented a retransmission model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
-The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
-We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
-
-We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
-Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
-Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.
-
-\paragraph{Future works}
-%
-The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
-Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
-%
-For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier08}
-or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
-% or more recently GLR-\UCB{} \cite{Besson2019GLRT},
-that nevertheless have not been explored for the targeted problem.
-Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this section.
-
-In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our recent works \cite{modiDemo2016,Besson2019WCNC}.
-Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes, and Christophe Moy at IETR and University Rennes 1.
-%
-A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.
-
 \paragraph{Note on the simulation code}

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
Trying to finish chapter 4 as quickly as possible.

@@
 
+% ----------------------------------------------------------------------
 \chapter{More Efficient Spectrum Usage of IoT Networks with Selfish MAB Learning}
@@
 \minitoc
+% ----------------------------------------------------------------------
 
@@
 
-\TODOL{Write a nice introduction for this chapter.}
-
 % ----------------------------------------------------------------------------
@@
 \label{sec:4:motivations}
+% ----------------------------------------------------------------------
 
@@
 
-\TODOL{Present outline of the chapter.}
+% \TODOL{Present outline of the chapter.}
 
-% \TODOL{Give list of publications associated with this chapter.}
 \paragraph{Publications}
+
 This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
@@
 \label{sec:4:firstModel}
+% ----------------------------------------------------------------------
 
@@
 \label{sec:4:gnuradio}
+% ----------------------------------------------------------------------
 
@@
 \label{sec:4:retransmissions}
+% ----------------------------------------------------------------------
 
@@
 \label{sec:4:conclusion}
+% ----------------------------------------------------------------------
 
-\TODOL{Write a conclusion, and future works}
+We give in this Section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.
 
-\subsection{Summary of the first model}\label{sub:41:conclusion}
+% ----------------------------------------------------------------------
+\paragraph{Summary of the first model}
+\label{sub:41:conclusion}
 
@@
 
-\subsection{Conclusions taken from the demonstration}
-\label{sec:42:conclusionFromDemonstration}
+% ----------------------------------------------------------------------
+\paragraph{Conclusions taken from the demonstration}
+\label{sub:42:conclusionFromDemonstration}
 
@@
 
-\subsection{Future works}
-Future works include...
+% ----------------------------------------------------------------------
+\paragraph{Conclusion about this second model}
+\label{sub:43:conclusion}
+
+In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
+The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
+We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
+
+We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
+Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
+Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.
+
+% ----------------------------------------------------------------------
+\paragraph{Future works}
+\label{sub:4:futureWorks}
+
+Future works related to this Chapter include the following directions.
 
 \paragraph{Possible extensions of the first model}
-%
+
 The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device.
@@
 \paragraph{Extensions of the demonstration}
+
 Possible future extensions of our demonstration include the following points.
@@
 
+\paragraph{Possible extensions for the second model}
+
+The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
+Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
+%
+For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
+or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
+or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
+that nevertheless have not been explored for the targeted problem.
+Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this section.
+
+In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our demonstration \cite{Besson2019WCNC}.
+Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes, and Christophe Moy at IETR and University Rennes 1.
+%
+A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.
+
 % ----------------------------------------------------------------------------
@@
 \label{sec:4:appendix}
+% ----------------------------------------------------------------------------
 
-% \TODOL{Include here additional results (proof in Section from CROWNCOM ?)}
+We include here a missing proof from Section~\ref{sec:4:firstModel}.
+
 \input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex}

### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

@@
     \centering
-    \includegraphics[width=0.65\linewidth]{system_model1.eps}
+    \includegraphics[width=0.70\linewidth]{system_model1.eps}
     \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
@@
     \centering
-    \includegraphics[scale=0.45]{protocol.eps}
+    \includegraphics[scale=0.40]{protocol.eps}
     \caption{The considered time-frequency slotted protocol.
@@
 \paragraph{MAB model}
-%
+
 We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
@@
 
-This section presents three different policies that will be used to assess the efficiency of the learning algorithms presented in the next section.
+We present three different policies that will be used to assess the efficiency of the learning algorithms presented in the next Subsection.
 The first one is naive but can be used in practice, while the two others are very efficient but require full knowledge on the system (\ie, an oracle) and are thus unpractical.
@@
 In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $i$, at that time slot.
-The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$ :
+The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$:
 \begin{equation}
@@
 
-We investigate in this section the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
+We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
 We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.
@@
 
-Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem :
-
+Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem:
+%
 \begin{subequations}\label{eq:41:prob}
@@
 Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
-Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $i$ :
+Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $i$:
 %
@@
 
-In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices : they are \iid{} with mean $\mu_i = 1 - p S_i$.
+In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices: they are \iid{} with mean $\mu_i = 1 - p S_i$.
 Moreover, dynamic devices are assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
@@
     \centering
-    \includegraphics[scale=0.65]{perf_learning.eps}
-    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices.}
+    \includegraphics[scale=0.60]{perf_learning.eps}
+    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices. Both algorithms achieve close to optimal performance after a reasonable learning time.}
     \label{fig:41:perf_learning}
@@
 \href{https://Bitbucket.org/scee_ietr/rl_slotted_iot_networks}{\texttt{Bitbucket.org/scee\_ietr/rl\_slotted\_iot\_networks}}.
+It was written in collaboration with Rémi Bonnefoi, in May $2017$.

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Trying to finish chapter 4 as quickly as possible.
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

@@
 %
-Similarly to the previous section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
+Similarly to the previous Section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
 %
@@
 
-% \TODOL{TODO:
+% ----------------------------------------------------------------------
+\subsection{Context of this demonstration}
+\label{sub:42:motivation}
+% ----------------------------------------------------------------------
 
-% - Clean up notations, be sure it is OK with previous Sections
-
-% - Be sure to not repeat any motivations / maths / explanations given before
-
-% - Be sure to give enough details about the conclusions etc
-% }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Motivation for this demonstration}
-\label{sub:42:motivation}
-
-In this Section, we describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
+We describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
 This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
@@
 
+\paragraph{Outline}
+
 The rest of this Section is organized as follows.
-% The system model is introduced in Section 2. In Section 3, we describe more formally both the \UCB{} and the TS algorithms.
 Our implementation is presented in Section~\ref{sub:42:implementation}, with details about the user interface, and results are given in Section~\ref{sub:42:results}.
-% and Section 6 concludes.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ----------------------------------------------------------------------
 \subsection{GNU Radio implementation}
 \label{sub:42:implementation}
+% ----------------------------------------------------------------------
 
-In this section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
+In this Section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
 We first describe the simplified physical layer of this demo, then we present our GNU Radio implementation.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ----------------------------------------------------------------------
 \subsection{Numerical results}
 \label{sec:42:results}
+% ----------------------------------------------------------------------
 
-We compare the two algorithms described in Section~\ref{sub:42:UCB} against a uniform access algorithm, that uniformly selects its channel at random.
+We compare the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} against a uniform access algorithm, that uniformly selects its channel at random.
 Three objects are compared by their mean successful communication rates, on a horizon of $2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
@@
 Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
-%  \cite[Fig.3]{Bonnefoi17}.
 Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.
@@
 	\centering
-    \includegraphics[height=8.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
+    \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
     \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success.
@@
 \paragraph{Availability of data and materials}
-%
+
 The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at
 \href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/\\malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
+
 It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.
+The demonstration only requires a laptop and free tools,
+as the laptop should run GNU/Linux and a distribution like Ubuntu or Debian.
 
-A $6$-minute \textbf{video} showing our demonstration is at \texttt{\url{youtu.be/HospLNQhcMk}}.
+\paragraph{Video}
+
+A $6$-minute \textbf{video} showing our demonstration is available online at \texttt{\url{youtu.be/HospLNQhcMk}}.
 It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
@@
 
-\paragraph{Acknowledgment}
-%
+\paragraph{Special acknowledgment}
+
 We acknowledge the work of two CentraleSup{\'e}lec students,

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

@@
 
-In this section, we extend the model presented before in order to take into account the possibility for retransmission of messages after a collision.
+\graphicspath{{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/}}
+
+In this Section, we extend the model presented before in order to take into account the possibility for retransmission of messages after a collision.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.
@@
 
-\TODOL{This section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}}
+% \TODOL{This Section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}}
 
@@
 
-The aim of this section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
+The aim of this Section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
 For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
@@
 
-The contributions of this section are summarized as follows:
+The contributions of this Section are summarized as follows:
 \begin{itemize}
@@
 	\centering
-	% \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Markov_model.eps}
-	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Markov_model-eps-converted-to.pdf}
+	% \includegraphics[width=0.65\linewidth]{Markov_model.eps}
+	\includegraphics[width=0.70\linewidth]{Markov_model-eps-converted-to.pdf}
 	\caption{The considered Markov model for the behavior of all the devices in the network.}
@@
 	\centering
-	% \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Approximation_m10.eps}
-	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/Approximation_m10-eps-converted-to.pdf}
+	% \includegraphics[width=0.65\linewidth]{Approximation_m10.eps}
+	\includegraphics[width=0.65\linewidth]{Approximation_m10-eps-converted-to.pdf}
 	\caption{Our proposed approximation for the probability of collision at the second transmission. It is more precise for smaller values of $N$.}
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/ResultsUCB.eps}
-	\caption{
+	\includegraphics[width=0.75\linewidth]{ResultsUCB.eps}
+	\caption[Test of a shorter caption]{
 		Comparison among the exposed heuristics for the retransmission: \textcolor{blue}{Only \UCB}, \textcolor{cyan}{Random}, \textcolor{purple}{\UCB}, \textcolor{green}{$K$ \UCB}, and \textcolor{red}{Delayed \UCB}.
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/ResultsUCB2.eps}
+	\includegraphics[width=0.75\linewidth]{ResultsUCB2.eps}
 	\caption{
@@
 %
-The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
+The source code (MATLAB or Octave) used for the simulations and the figures of this Section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
 It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in July and August $2018$.
-
-% % ----------------------------------------------------------------------
-% \bibliographystyle{ieeetr}
-% \bibliography{IEEE_WCNC__2019__Paper__BMBM}

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
I'm almost done and happy about this Chapter 4. 38 pages, it's slightly too long.

@@
 For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
+Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}.
 Many recent works show that MAB algorithms work well for real-world radio signal.
 However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
-In recent works like \cite{Bonnefoi17}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \iid.
+
+We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
+Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \iid.
 As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
-
-We also present an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
-
-% \TODOL{Present outline of the chapter.}
+We then validate the model with a hardware implementation, detailed in Section~\ref{sec:4:gnuradio}.
+%
+We conclude this Chapter by presenting an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
 
@@
 % ----------------------------------------------------------------------------
-\section{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
+\section[Selfish learning for many dynamic devices in an IoT network]{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
 \label{sec:4:firstModel}
@@
 We give in this Section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.
+
+\subsection*{Conclusions}
 
@@
 % ----------------------------------------------------------------------
-\paragraph{Future works}
+\subsection*{Future works}
 \label{sub:4:futureWorks}

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
I'm done for today on Chapter 4.

@@
 We refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
-which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Section~\ref{sec:42:results}).
+which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Section~\ref{sub:42:results}).
 Then we discuss the relevance of a MAB model for our IoT application.
@@
 % ----------------------------------------------------------------------
-\subsection{GNU Radio implementation}
+\subsection{Physical model and user interface of our GNU Radio implementation}
 \label{sub:42:implementation}
@@
 For maximum efficiency, we wrote all of our blocks in \texttt{C++}.
-These flow-graphs can be written in either C++ or the Python programming language. The GNU Radio infrastructure is written entirely in C++, and many of the user tools are written in Python.
+These flow-graphs can be written in either \texttt{C++} or the Python programming language.
+The GNU Radio infrastructure is written entirely in \texttt{C++}, and many of the user tools are written in Python.
 GNU Radio Companion is a graphical UI used to develop GNU Radio applications:
@@
 \subsection{Numerical results}
-\label{sec:42:results}
+\label{sub:42:results}
 % ----------------------------------------------------------------------

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
I'm done for today on Chapter 4.

@@
 
-- Clean up notations, be sure it is OK with previous Sections
-
 - Be sure to not repeat any motivations / maths / explanations given before
-
-- Be sure to give enough details about the conclusions etc
 }
@@
 These \emph{rewards} are the \emph{acknowledgment} (\emph{Ack}) signals received after transmitting packets to the gateway.
-In this way, a successful transmission is considered when an acknowledgment is received, and a learning approach is employed to select the best channel.
+In this way, a transmission is considered successful when an acknowledgment is received, and a learning approach is employed to select the best channel.
 
 We address the problem of channel selection taking into account the described Markov model for the retransmissions of end-devices.
-It motivates our present work for which we consider the retransmissions in the analysis of MAB algorithms.
-% It motivates our present work for which we consider the number of retransmissions, carried out by each device to address several MAB algorithms.
+% It motivates our present work for which we consider the retransmissions in the analysis of MAB algorithms.
+It motivates our present work for which we consider the number of retransmissions, carried out by each device to address several MAB algorithms.
 
@@
 
-We consider two hypotheses $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as,
+We consider two hypotheses $\cH_{1}$ and $\cH_{2}$ defined as,
 \begin{itemize}
-	\item $\mathcal{H}_{1}$:
+	\item $\cH_{1}$:
     The probability $p_{c1}$ is composed by the sum of two probabilities: i)
@@
 
-	\item $\mathcal{H}_{2}$:
+	\item $\cH_{2}$:
 	The total number of the back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
@@
 We denote $p_{ca}$ the probability to have a collision with a packet involved in the previous collision.
-Under assumption $\mathcal{H}_{1}$, the probability that the same device's packet is experiencing again a collision at the second time slot is
+Under assumption $\cH_{1}$, the probability that the same device's packet is experiencing again a collision at the second time slot is
 \begin{equation}\label{eq:43:decomppc1}
@@
 If the device had a collision, we consider $p_{bp}(n)$ the probability that it had a collision with \emph{exactly} $n$ packets (for any $1\leq n \leq N-1$), and that \emph{at least one} of the $n$ devices involved in the previous collision chose the same back-off interval.
-Hence, under hypothesis $\mathcal{H}_{2}$, we can relate $p_{ca}$ with $p_{bp}(n)$, and the different probabilities that a device experienced a collision during the first slot and has the same back-off interval for its retransmission is,
+Hence, under hypothesis $\cH_{2}$, we can relate $p_{ca}$ with $p_{bp}(n)$, and the different probabilities that a device experienced a collision during the first slot and has the same back-off interval for its retransmission is,
 %
@@
 
-Once again under $\mathcal{H}_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first terms of the sum in \eqref{eq:43:sumpc2} are predominant.
+Once again under $\cH_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first terms of the sum in \eqref{eq:43:sumpc2} are predominant.
 Moreover, for these terms, $n$ is small compared to $N-1$, and so $N-1-n$ can be approximated to $N-1$.
@@
 \end{equation}
-where $\mathbbm{1}$ is an indicator function that is equal to $1$, if the IoT device chooses, for its $\tau$-th transmission, a channel $k$, and $0$ otherwise.
 The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
@@
 \end{equation}
-where $\alpha$ refers to an exploration coefficient\footnote{~In fact, the larger this coefficient is, the longer the exploration, while the \UCB{} algorithm is proven to be order optimal for $\alpha>0.5$ \cite{Bubeck12}, and has reported a good performance for lower values of $\alpha>0$.},
+where $\alpha$ refers to an exploration coefficient,
+% \footnote{~In fact, the larger this coefficient is, the longer the exploration, while the \UCB{} algorithm is proven to be order optimal for $\alpha>0.5$ \cite{Bubeck12}, and has reported a good performance for lower values of $\alpha>0$.},
 that we chose equal to $1/2$, as suggested in \cite{Audibert07} and as done in previous works \cite{Bonnefoi18,Bonnefoi17}.
@@
 
-\begin{small} % XXX remove if needed
+% \begin{small} % XXX remove if needed
 \begin{figure}[h!]
@@
 	\begin{algorithm}[H]
-		\begin{small} % XXX remove if needed
+		% \begin{small} % XXX remove if needed
 		\For(){$t = 0, \dots, T$}{
-			Compute for each channel $ U_k(t) = \widehat{\mu_k}(t) + B_k(t).$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+			Compute for each channel $ U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
 			Transmit in channel $C(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
@@
 		\label{algo:43:UCB}
-		\end{small} % XXX remove if needed
+		% \end{small} % XXX remove if needed
 	\end{algorithm}
 \end{figure}
-\end{small} % XXX remove if needed
+% \end{small} % XXX remove if needed
 
@@
 	\begin{algorithm}[H]
-	\begin{small} % XXX remove if needed
+	% \begin{small} % XXX remove if needed
 	\For(){$t = 0, \dots, T$}{
@@
 		\label{algo:43:UCBthenRandom}
-	\end{small} % XXX remove if needed
+	% \end{small} % XXX remove if needed
 \end{algorithm}
@@
 \vspace*{-3pt}
-\begin{small} % XXX remove if needed
+% \begin{small} % XXX remove if needed
 \begin{figure}[h!]
@@
 	\begin{algorithm}[H]
-		\begin{small} % XXX remove if needed
+		% \begin{small} % XXX remove if needed
 		\For(){$t = 0, \dots, T$}{
@@
 			}
-			\Else(\tcp*[f]{Random retransmission}){
+			\Else(\tcp*[f]{Packet retransmission}){
 				Compute for each channel $U^r_k(t) = \widehat{\mu^r_k}(t) + B^r_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
@@
 		\label{algo:43:TwoUCB}
-		\end{small} % XXX remove if needed
+		% \end{small} % XXX remove if needed
 		\end{algorithm}
 \end{figure}
-\end{small} % XXX remove if needed
+% \end{small} % XXX remove if needed
 
@@
 \vspace*{-3pt}
-\begin{small} % XXX remove if needed
+% \begin{small} % XXX remove if needed
 	\begin{figure}[h!]
@@
 		\begin{algorithm}[H]
-			\begin{small} % XXX remove if needed
+			% \begin{small} % XXX remove if needed
 			% \KwData{$\forall k,j\in[\![1;K]\!]$, $N_k^j(t)$, $\widehat{\mu_k}^j(t)$, $B_k^j(t)$ and $U_k^j(t)$}
@@
 			\label{algo:43:UCBthenKp1}
-			\end{small} % XXX remove if needed
+			% \end{small} % XXX remove if needed
 		\end{algorithm}
 	\end{figure}
-\end{small} % XXX remove if needed
+% \end{small} % XXX remove if needed
 
@@
 		\begin{algorithm}[H]
-			\begin{small} % XXX remove if needed
+			% \begin{small} % XXX remove if needed
 			\For(\tcp*[f]{At every time step})
@@
 				}
-				\uElseIf(\tcp*[f]{Random selection}){$t \leq \Delta$}{
+				\uElseIf(\tcp*[f]{Random retransmission}){$t \leq \Delta$}{
 					Transmit randomly in a channel $C(t) \sim \mathcal{U}(1,\ldots,K)$.
@@
 			\label{algo:43:UCBwithDelay}
-			\end{small} % XXX remove if needed
+			% \end{small} % XXX remove if needed
 		\end{algorithm}
 	\end{figure}
-\end{small} % XXX remove if needed
+% \end{small} % XXX remove if needed
 

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
I'm done for today on Chapter 4.

@@
 The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{can not} communicate with each others, and \emph{do not} know the system parameters.
-We will investigate this behavior in order to understand it better theoretically.
-We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
 
@@
 So additionally to deciding the channel for communication, each device has to also decide when to communicate.
-This is another direction of research, that we will investigate in the future.
+% This is another direction of research, that we will investigate in the future.
+
+% We will investigate this behavior in order to understand it better theoretically.
+% We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
 
@@
 
+\newpage
 % ----------------------------------------------------------------------------

### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Added content for Chapter 5, based on my ALT paper « Multi-Player Bandits Revisited », see https://hal.inria.fr/hal-01629733

@@
 	\includegraphics[width=0.65\linewidth]{Approximation_m10-eps-converted-to.pdf}
-	\caption{Our proposed approximation for the probability of collision at the second transmission. It is more precise for smaller values of $N$.}
+	\caption[Proposed approximation for the probability of collision at the second transmission]{Proposed approximation for the probability of collision at the second transmission.
+It is more precise for smaller values of $N$.}
 	\label{fig:43:Approximation_m10}
@@
 		}
-		\caption{The \UCB{} algorithm for channel selection.}
+		\caption{The \UCB{} algorithm for channel selection (base building block).}
 		\label{algo:43:UCB}
@@
 		}
-		\caption{Uniform random retransmission.}    % A naive
+		\caption{Heuristic: uniform random retransmission.}    % A naive
 		\label{algo:43:UCBthenRandom}
@@
 		}
-		\caption{\UCB{} for retransmission.}    % A naive
+		\caption{Heuristic: \UCB{} for retransmission.}    % A naive
 		\label{algo:43:TwoUCB}
@@
 			}
-			\caption{$K$ different {\UCB}s for retransmission.}
+			\caption{Heuristic: $K$ different {\UCB}s for retransmission.}
 			\label{algo:43:UCBthenKp1}
@@
 			}
-			\caption{Delayed \UCB{} for retransmission.}
+			\caption{Heuristic: Delayed \UCB{} for retransmission.}
 			\label{algo:43:UCBwithDelay}
@@
 	\includegraphics[width=0.75\linewidth]{ResultsUCB.eps}
-	\caption[Test of a shorter caption]{
-		Comparison among the exposed heuristics for the retransmission: \textcolor{blue}{Only \UCB}, \textcolor{cyan}{Random}, \textcolor{purple}{\UCB}, \textcolor{green}{$K$ \UCB}, and \textcolor{red}{Delayed \UCB}.
+	\caption[First comparison between the exposed heuristics for the retransmission: Only \UCB, Random, \UCB, $K$ \UCB, and Delayed \UCB]{
+		Comparison between the exposed heuristics for the retransmission: \textcolor{blue}{Only \UCB}, \textcolor{cyan}{Random}, \textcolor{purple}{\UCB}, \textcolor{green}{$K$ \UCB}, and \textcolor{red}{Delayed \UCB}.
 		The usage of the same learning policy for transmissions and retransmission is named \textcolor{blue}{Only \UCB{}},
@@
 	\includegraphics[width=0.75\linewidth]{ResultsUCB2.eps}
-	\caption{
+	\caption[First comparison between the exposed heuristics for the retransmission: Only \UCB, Random, \UCB, $K$ \UCB, and Delayed \UCB]{
 		Second scenario: learning helps a lot (a gain of $30\%$ in terms of collision probability), and learning to retransmit smartly is needed.

### 2019-02-14 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Added content for Chapter 5, based on my ALT paper « Multi-Player Bandits Revisited », see https://hal.inria.fr/hal-01629733

% - ``Multi-Player Bandits Revisited'', see https://hal.inria.fr/hal-01629733

\TODOL{This chapter is basically a raw include from my paper ``Multi-Player Bandits Revisited'', see https://hal.inria.fr/hal-01629733}

Multi-player Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms. Most existing work assume
that \emph{sensing information} is available to the algorithm. Under this assumption, we improve the state-of-the-art lower bound for the regret of any decentralized algorithms
and introduce two algorithms, \RandTopM{} and \MCTopM{}, that are shown to empirically outperform existing algorithms. Moreover, we provide strong theoretical guarantees for these algorithms, including
a notion of asymptotic optimality in terms of the number of selections of bad arms.
We then introduce a promising heuristic, called \Selfish{}, that can operate
without sensing information, which is crucial for emerging applications to Internet of Things networks. We investigate the empirical performance of this algorithm
and provide some first theoretical elements for the understanding of its behavior.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Introduction}
\label{sec:5:introduction}
% the paragraph below may be useful to state in the intro or early in the paper

% general intro, UCB

Several sequential decision making problems under the constraint of partial information
have been studied since the 1950s under the name of Multi-Armed Bandit (MAB) problems \citep{Robbins52,LaiRobbins85}.
In a stochastic MAB model, an agent is facing $K$ unknown probability distributions, called arms in reference to the arms of a one-armed bandit (or slot machine) in a casino.
Each time she selects (or draws) an arm, she receives a reward drawn from the associated distribution.
Her goal is to build a sequential selection strategy that maximizes the total reward received.
A class of algorithms to solve this problem is based on Upper Confidence Bounds (UCB), first proposed by \cite{LaiRobbins85,Agrawal95} and further popularized by \cite{Auer02}.
The field has been very active since then, with several algorithms proposed and analyzed, both theoretically and empirically, even beyond the stochastic assumption on arms, as explained in the survey by \cite{Bubeck12}.

% appli, from clinical trials to cognitive radios

The initial motivation to study MAB problems arose from clinical trials (the first MAB model can be traced back to $1933$, by \citeauthor{Thompson33}), in which a doctor sequentially allocates treatments (arms) to patients and observes their efficacy (reward).
More recently, applications of MAB have shifted towards sequential content recommendation, \eg{} sequential display of advertising to customers or A/B testing \citep{Li10,Chapelleetal14Ad}.
%
In the mean time, MAB were found to be relevant to the field of Cognitive Radio (CR, \cite{Mitola99}),
and \cite{Jouini09,Jouini10} first proposed to use \UCB{} for the Opportunistic Spectrum Access (OSA) problem,
and successfully conducted experiments on real radio networks demonstrating its usefulness.
For CR applications, each arm models the quality or availability of a radio channel (a frequency band) in which there is some background traffic (\eg, primary users paying to have a guaranteed access to the channel in the case of OSA).
A smart radio device needs to insert itself in the background traffic, by sequentially choosing a channel to access and try to communicate on, seeking to optimize the quality of its global transmissions.

% the need for multiple players in cognitive radio

For the development of CR, a crucial step is to insert \emph{multiple} $M \geq 2$ smart devices in the \emph{same} background traffic.
With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
%who essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
% In the context of CR, this first extension can model a centralized agent taking decisions for end-devices, for instance a base station affecting devices to channels.
% Multiple-Play MAB algorithms successfully used for other Cognitive Radio tasks,
% for instance by \cite{Maghsudi16} for 5G small-cells or by \cite{Bonnefoi16} for Internet-of-Things (IoT) networks.
%
% Another generalization targets the opposite goal, where several devices
% are learning and conjointly accessing the same network,
% trying to find a consensus in a decentralized learning way.
% It is harder as there is a game theoretic-like situation, for instance where devices try to avoid collisions.
%
Due to the communication cost implied by a central controller, a more relevant model is the
\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain)
and by \cite{Avner15,Avner16} and \cite{Rosenski16} (for \iid{} or piece-wise \iid{} arms).
%
The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all arms to estimate their means while trying to focus on the best arms to gain as much rewards as possible.
The decentralized setting considers no exchange of information between players, that only know $K$ and $M$, and to avoid collisions, players should furthermore find orthogonal configurations (\ie, the $M$ players use the $M$ best arms without any collision), without communicating.
Hence, in that case the trade-off is to be found between exploration, exploitation \emph{and} low collisions.

All these above-mentioned works are motivated by the OSA problem, in which it is assumed that \emph{sensing} occurs, that is each smart device observes the availability of a channel (sample from the arm) \emph{before} trying to transmit and possibly experiment a collision with other smart devices.
However some real radio networks do not use sensing at all, \eg, emerging standards developed for \emph{Internet of Things} (IoT) networks such as LoRaWAN.
Thus, to take into account these new applications, algorithms with additional constraints on the available feedback have to be proposed within the multiple-player MAB model.
Especially, the typical approach that combines a (single-player) bandit algorithm based on the sensing information --to learn the quality of the channels while targeting the best ones-- with a low-complexity decentralized collision avoidance protocol, is no longer possible.

In this Section, we take a step back and present the different feedback levels possible for multi-player MAB algorithms. For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees. In the presence of sensing information, our contributions are a new problem-dependent regret lower bound, tighter than previous work, and the introduction of two algorithms, \RandTopM{} and \MCTopM{}. Both are shown to achieve an asymptotically optimal number of selections of the sub-optimal arms, and for \MCTopM{} we furthermore establish a logarithmic upper bound on the regret, that follows from a careful control of the number of collisions. In the absence of sensing information, we propose the \Selfish{} heuristic and investigate its performance. Our study of  this algorithm is supported by (promising) empirical performance and some first (disappointing) theoretical elements.

The rest of the Section is organized as follows.
We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms}, with the result of our experimental study reported in
Section~\ref{sec:5:experiments}. Theoretical elements are then presented in Section~\ref{sec:5:upperbounds}.
% and finally Section~\ref{sec:5:conclusion} concludes.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Multi-Player Bandit Model with Different Feedback Levels}
\label{sec:5:model}

% describe our stochastic assumptions
We consider a $K$-armed Bernoulli bandit model, % (for $K\geq2$),
in which arm $k$ is a Bernoulli distribution with mean $\mu_k\in[0,1]$.
We denote $(Y_{k,t})_{t\in\N}$ the \iid{} (binary) \emph{reward stream} for arm $k$, that satisfies $\Pr(Y_{k,t}=1) = \mu_k$ and that is independent from the other rewards streams.
However we mention that our lower bound and all our algorithms (and their analysis) can be easily extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}). For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio, as it can model channel availabilities.

% stochastic bandit model in which each arm distribution is assumed to belong to a one-dimensional canonical exponential family. That is, arm $k$ has a density with respect to a reference measure that takes the form
% \[f_{\theta_k}(x) = \exp(\theta_k x - b(\theta_k)),\]
% there $\theta_k \in \R$ is some parameter and $b$ is a twice-differentiable function. It is well known that such distributions can be alternatively parameterized by their means (see, \eg, KLUCB), and we denote by $\mu_k$ the mean of arm $k$. Examples of exponential families include Gaussian distribution with known variance, exponential, Poisson distributions and Bernoulli distributions. For cognitive radio applications, it is often assumed that each arm is a Bernoulli distribution, giving 1 if the associated channel is available, 0 other. We let $(Y_{k,t})_{t\in\N}$ denote the \iid{} reward stream for arm $k$, whose distribution has mean $\mu_k$ and that is independent from other arms.

% explain the interaction protocol and the GOAL
In the multi-player MAB setting, there are $M \in \{1,\dots,K\}$ players (or agents),
that have to make decisions at some pre-specified time instants.
At time step $t \in\mathbb{N},t\geq1$, player $j$ selects an arm $A^j(t)$, independently from the other players' selections.
A \emph{collision} occurs at time $t$ if at least two players choose the same arm.
We introduce the two events, for $j\in\{1,\dots,M\}$ and $k\in\{1,\dots,K\}$,
\begin{equation}
  C^j(t) :=  \{ \exists j' \neq j : A^{j'}(t) = A^j(t) \}
  \ \ \ \text{and} \ \ \ C_k(t) :=  \left\{ \# \{ j : A^j(t) = k\} > 1 \right\},
\end{equation}
that respectively indicate that a collision occurs at time $t$ for player $j$ and that a collision occurs at time $t$ on arm $k$.
Each player $j$ then receives (and observes) the \emph{binary rewards}
$r^j(t) \in \{0,1\}$,
\begin{equation}
  r^j(t) := Y_{A^j(t),t} \; \indic(\overline{C^j(t)}).
\end{equation}
In words, she receives the reward of the selected arm if she is the only one to select this arm, and a reward zero otherwise\footnote{This provides another reason to focus on the Bernoulli model. It is the hardest model, in the sense that receiving a reward zero is not enough to detect collisions. For other models, the data streams $(Y_{k,s})_s$ are usually continuously distributed, with no mass at zero. Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.}.
Other models for rewards loss have been proposed in the literature (\eg, the reward is randomly allocated to one of the players selecting it), but we focus on full reward occlusion in this Section.

A multi-player MAB strategy is a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for each player, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
Previous observations for player $j$ at time $t$ always include the previously chosen arms $A^j(s)$ and received rewards $r^j(s)$ for $s<t$, but may also include the \emph{sensing information} $Y_{A^j(t),t}$ or the \emph{collision information} $C^j(t)$.
More precisely, depending on the application, one may consider the following three observation models, \modelun, \modeldeux{} and \modeltrois.
%
% If the arms have continuous distributions (or are such that  $\Pr(Y_{k,s}=0)=0$), the \emph{sensing information} $Y_{A^j(t),t}$ and \emph{collision information} $C^j(t)$ can always be extracted from the reward information. But for Bernoulli distribution, one may consider the following three observation models, that are not equivalent:
\begin{itemize}
  \item[\modelun]
    \textbf{Simultaneous sensing and collision}: player $j$ observes  $Y_{A^j(t),t}$ \emph{and} $C^j(t)$ (not previously studied).
  \item[\modeldeux]
    \textbf{Sensing, then collision}: player $j$ observes $Y_{A^j(t),t}$, \emph{then} observes the reward, and thus also $C^j(t)$ only if $Y_{A^j(t),t} = 1$.
    This common setup, studied for example by \cite{Anandkumar11,Avner15,Rosenski16}, is relevant to model the OSA problem: the device first checks for the presence of primary users in the chosen channel.
    If this channel is free ($Y_{A^j(t),t}=1$), the transmission is successful ($r^j(t)=1$) if no collision occurs with other smart devices ($\overline{C^j(t)}$).
  \item[\modeltrois]
    \textbf{No sensing}: player $j$ only observes the reward $r^j(t)$.
    For IoT networks, this reward can be interpreted as an acknowledgement from a Base Station,
    received when a communication was successful.
    A lack of acknowledgment may be due to a collision
    with a device from the background traffic $(Y_{A^j(t),t}=0)$,
    or to a collision with one of the others players ($C^j(t)$).
    However, the sensing and collision information are censored.
    Recently, \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications. %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
\end{itemize}
\noindent Under each of these three models, we define $\cF^j_t$ to be the filtration generated by the observations gathered by player $j$ up to time $t$ (which contains different information under models \modelun, \modeldeux{} and \modeltrois).
While a \emph{centralized} algorithm may select the vector of actions for all players $(A^1(t),\dots,A^M(t))$ based on all the observations from $\bigcup_j \cF^j_{t-1}$, under a \emph{decentralized} algorithm the arm selected at time $t$ by player $j$ only depends on the past observation of this player.
More formally, $A^j(t)$ is assumed to be $\cF^j_{t-1}$-measurable.

% [Index sets \Mworst, \Mbest]
\begin{definition}\label{def:5:MbestMworst}
  We denote by $\mu_1^*$ the best mean, $\mu_2^*$ the second best etc, and
  by \Mbest{} the (non-sorted) set of the indices of the $M$ arms with largest mean (\emph{best arms}): if $\mu_1^* = \mu_{k_1}, \dots, \mu_M^* = \mu_{k_M}$
  then $\Mbest = \{k_1, \dots, k_M\}$.
  %
  Similarly, \Mworst{} denotes the set of indices of the $K-M$ arms with smallest means (\emph{worst arms}),
  $\{1, \dots, K\} \setminus \Mbest$.
  %
  Note that they are both uniquely defined if $\mu_M^* > \mu_{M+1}^*$.
\end{definition}

Following a natural approach in the bandit literature, we evaluate the performance of a multi-player strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
The regret of the strategy $\rho$ at horizon $T$ is the difference between the cumulated reward of an oracle strategy, assigning in this case the $M$ players to \Mbest,
and the cumulated reward of strategy $\rho$:
\begin{equation}\label{eq:5:regret}
  R_T(\boldsymbol{\mu}, M, \rho) := \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
\end{equation}
Maximizing the expected sum of global reward of the system is indeed equivalent to minimizing the regret, and we now investigate the best possible \emph{regret rate} of a decentralized multi-player algorithm.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{An Asymptotic Regret Lower Bound}
\label{sec:5:lowerbound}

In this section, we provide a useful decomposition of the regret (Lemma~\ref{lem:5:DecompositionRegret}) that permits to establish a new problem-dependent lower bound on the regret (Theorem~\ref{thm:5:BetterLowerBound}), and also provides key insights on the derivation of regret upper bounds (Lemma~\ref{lem:5:1stUpperBound}).

% -----------------------------------------------------------------
\subsection{A Useful Regret Decomposition}
\label{sub:5:defregret}

We introduce additional notations in the following definition.

\begin{definition}
  % [$T_k(T)$ and $\cC_k(T)$]
  \label{def:5:nbSelections_nbCollisions}
  % \label{def:5:nbSelections}
  Let $T^j_k(T) := \sum_{t=1}^T \indic(A^j(t) = k)$,
  and denote $T_k(T) := \sum_{j=1}^M T^j_k(T)$ the \emph{number of selections} of arm $k\in\{1,\dots,K\}$ by any player $j\in\{1,\dots,M\}$, up to time $T$.
  % and $\cC_k(T)$ be the number of collisions on arm $k$ up-to time $T$.
% \end{definition}

% \begin{definition}[]\label{def:5:nbCollisions}
  Let $\cC_k(T)$ be the \emph{number of colliding players}\footnote{When $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one. So $\cC_k(T)$ counts the total \emph{number of colliding players} rather than the number of collision events. Hence there is small abuse of notation when calling it a number of collisions.}
  on arm $k\in\{1,\dots,K\}$ up to horizon $T$:
  \vspace*{-5pt}  % XXX
  \begin{equation}
    \cC_k(T) :=
    \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(C^j(t)) \indic(A^j(t) = k).
  \end{equation}
\end{definition}

Letting $\cP_M = \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
be the set of bandit instances such that there is a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.

\begin{lemma}\label{lem:5:DecompositionRegret}
  For any bandit instance $\boldsymbol{\mu}\in\cP_M$ such that $\mu_M^* > \mu_{M+1}^*$, it holds that
  \hfill{}
  (Proved in App.\ref{proof:5:DecompositionRegret})
  \vspace*{-5pt}  % XXX
  \begin{small}
    \begin{equation*}\label{eq:5:termeundeuxtermetrois}
      R_T(\boldsymbol{\mu}, M, \rho) =
      \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[T_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}}
      + \underbrace{\sum_{k \in \Mbest} (\mu_k -  \mu_M^*) (T - \E_{\mu}[T_k(T)])}_{\emph{\mytag{(b)}{eq:5:term2}}}
      + \underbrace{\sum_{k=1}^{K} \mu_k \E_{\mu}[\cC_k(T)]}_{\emph{\mytag{(c)}{eq:5:term3}}}.
    \end{equation*}
  \end{small}
\end{lemma}

In this decomposition, term \ref{eq:5:term1} counts the lost rewards due to \emph{sub-optimal arms} selections ($k \in \Mworst$), term \ref{eq:5:term2} counts the number of times the \emph{best arms} were \emph{not} selected ($k \in \Mbest$), and term \ref{eq:5:term3} counts the weighted number of collisions, on \emph{all arms}. It is valid for both centralized and decentralized algorithms. For centralized algorithms, due to the absence of collisions, \ref{eq:5:term3} is obviously zero, and \ref{eq:5:term2} is non-negative, as $T_k(T) \leq T$. For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $T_k(T) > T$. However, a careful manipulation of this decomposition (see Appendix~\ref{proof:5:1stLowerBound}) shows that the regret is always lower bounded by term \ref{eq:5:term1}.

\begin{lemma}\label{lem:5:1stLowerBound}
    For any strategy $\rho$ and $\boldsymbol{\mu}\in\cP_M$, it holds that
    $ R_T(\boldsymbol{\mu}, M, \rho)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)]$.
\end{lemma}

% -----------------------------------------------------------------
\subsection{An Improved Asymptotic Lower Bound on the Regret}
\label{sub:5:betterLowerBound}

To express our lower bound, we need to introduce $\kl(x,y)$ as the Kullback-Leibler divergence between the Bernoulli distribution of mean $x \neq 0,1$ and that of mean $y \neq 0,1$, so that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.

\begin{definition}\label{def:5:DecentralizedUniformEfficiency}
  A strategy $\rho$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$ and for all $\alpha \in (0,1)$,
  \vspace*{-5pt}  % XXX
  \begin{equation}
    R_T(\boldsymbol{\mu},M,\rho) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \ \ \ \text{and } \ \
    \forall j \in \{1,\dots,M\}, k \in \Mbest, \;\;
    % a_{j,k}
    \frac{T}{M}
    - \E_{\mu}[T^j_k(T)] \mathop{=}\limits_{T \to +\infty} o(T^{\alpha}).
    \label{eq:5:SUE}
  \end{equation}
\end{definition}

Having a small regret on every problem instance,
\ie, uniform efficiency,
is a natural assumption for algorithms,
that rules out algorithms tuned to perform well on specific instances only.
%
From this assumption $\left(R_T(\boldsymbol{\mu},M,\rho)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
${T}- \E_{\mu}[T_k(T)] {=} o(T^{\alpha})$, and so
\begin{equation}\label{eq:5:intermediate}
  \vspace*{-5pt}  % XXX
  \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[T^j_k(T)]\right) = o(T^{\alpha}).
  % \vspace*{-5pt}  % XXX
\end{equation}
The additional assumption in \eqref{eq:5:SUE} further implies some notion of \emph{fairness}, as it suggests that each of the $M$ players spends on average the same amount of time on each of the $M$ best arms. Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\rho^{\gamma} = (\rho^{\gamma(1)},\dots,\rho^{\gamma(M)})$ is independent from the choice of permutation $\gamma \in \Sigma_M$. In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in \{1,\dots,M\}$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency. Note that all our proposed algorithms are permutation invariant and \MCTopM{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.

We now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections under a \emph{decentralized strategy} that has access to the sensing information.
This result, proved in Appendix~\ref{proof:5:BetterLowerBound}, yields an asymptotic logarithmic lower bound on the regret, also given in Theorem~\ref{thm:5:BetterLowerBound}.

\begin{theorem}\label{thm:5:BetterLowerBound}
  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\rho$ and $\boldsymbol{\mu}\in\cP_M$,
  \vspace*{-10pt}  % XXX
  \begin{equation}\label{eq:5:LBDraws}
    \forall j \in \{1,\dots,M\}, \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
  \end{equation}

  % \vspace{-0.3cm}  % XXX
  \noindent From Lemma~\ref{lem:5:1stLowerBound}, it follows that
  \vspace*{-5pt}  % XXX
  \begin{equation}\label{eq:5:ourLowerBound}
    \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \rho)}{\log(T)}
    \geq M \times \left( \sum_{k \in \Mworst} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_M^*)} \right) .
  \end{equation}
\end{theorem}

Observe that the regret lower bound \eqref{eq:5:ourLowerBound} is tighter than the state-of-the-art lower bound in this setup
given by \cite{Zhao10}, that states that
\begin{equation}\label{eq:5:Zhao10LowerBound}
  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \rho)}{\log(T)}
  \geq \sum_{k \in \Mworst} \left( \sum_{j=1}^{M} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_{j}^*)} \right),
\end{equation}
as for every $k \in \Mworst$ and $j \in \{1,\dots,M\}$, $\kl(\mu_k, \mu_j^*) \geq \kl(\mu_k, \mu_M^*)$
(see Figure~\ref{fig:5:CompLowerBounds} in Appendix~\ref{app:5:illustrationLowerBound}).
%
It is worth mentioning that \cite{Zhao10} proved a lower bound under the more general assumption for $\rho$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
Our result could be extended to this case but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.

Interestingly, our lower bound is exactly a multiplicative constant factor $M$ away from the lower bound given by \cite{Anantharam87a} for centralized algorithms (which is clearly a simpler setting). This intuitively suggests the number of players $M$ as the (multiplicative) \emph{``price of decentralized learning''}. However, to establish our regret bound, we lower bounded the number of collisions by zero, which may be too optimistic.
%
Indeed, for an algorithm to attain the lower bound \eqref{eq:5:ourLowerBound}, the number of selections of each sub-optimal arm should match the lower bound \eqref{eq:5:LBDraws} \emph{and} term \ref{eq:5:term2} and term \ref{eq:5:term3} in the regret decomposition of Lemma~\ref{lem:5:DecompositionRegret} should be negligible compared to  $\log(T)$.
To the best of our knowledge, no algorithm has been shown to experience only $o(\log(T))$ collisions so far,
for every $M \in \{2,\dots,K\}$ and $\boldsymbol{\mu} \in \cP_M$.

A lower bound on the minimal number of collisions experienced by any strongly uniformly efficient decentralized algorithm would thus be a nice complement to our Theorem~\ref{thm:5:BetterLowerBound}, and it is left as future work.

% -----------------------------------------------------------------
\subsection{Towards Regret Upper Bounds}

A natural approach to obtain an upper bound on the regret of an algorithm is to upper bound separately each of the three terms defined in Lemma~\ref{lem:5:DecompositionRegret}.
The following result shows that term $(b)$ can be related to the number of sub-optimal selections and the number of collisions that occurs on the $M$ best arms.

% \vspace*{-15pt}  % XXX remove if problem
\begin{lemma}\label{lem:5:1stUpperBound}
  The term $(b)$ in Lemma~\ref{lem:5:DecompositionRegret} is upper bounded as
  \hfill{}
  (Proved in Appendix~\ref{proof:5:1stUpperBound})
  %
  \begin{equation}\label{eq:5:1stUpperBound}
    (b) \leq (\mu_1^* - \mu_M^*) \Bigl(    \sum_{k \in \Mworst} \E_{\mu}[T_k(T)]
    + \sum_{k \in \Mbest} \E_{\mu}[C_{k}(T)]
    \Bigr).
  \end{equation}
  \vspace*{-5pt}  % XXX
\end{lemma}

This result can also be used to recover Proposition 1 from \cite{Anandkumar11}, giving an upper bound on the regret that only depends on
the \emph{expected number of sub-optimal selections} -- $\E_{\mu}[T_k(T)]$ for $k \in \Mworst$ --
and the \emph{expected number of colliding players on the optimal arms} -- $\E_{\mu}[\cC_k(T)]$ for $k \in \Mbest$. Note that, in term (c) the number of colliding players on the sub-optimal arm $k$ may be upper bounded as $\E_{\mu}[\cC_k(T)] \leq M \E_{\mu}[T_k(T)]$.
%

In the next Section, we present an algorithm that has a logarithmic regret,
while ensuring that the number of sub-optimal selections is matching the lower bound of Theorem~\ref{thm:5:BetterLowerBound}.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{New Algorithms for Multi-Player Bandits}
\label{sec:5:algorithms}

When sensing is possible, that is under observation models \modelun{} and \modeldeux, most existing strategies build on a \emph{single-player bandit algorithm} (usually an \emph{index policy}) that relies on the sensing information, together with an \emph{orthogonalization strategy} to deal with collisions. We present this approach in more details in Section~\ref{sub:5:RandTopM_and_MCTopM} and introduce two new algorithms of this kind, \RandTopM{} and \MCTopM.
%
Then, we suggest in Section~\ref{sub:5:Selfish} a completely different approach, called \Selfish{}, that no longer requires an orthogonalization strategy as the collisions are directly accounted for in the indices that are used.
\Selfish{} can also be used under observation model \modeltrois{} --\emph{without sensing}--, and without the knowledge of $M$.

% -----------------------------------------------------------------

\subsection{Two New Strategies Based on Indices and Orthogonalization: \RandTopM{} and \MCTopM}
\label{sub:5:RandTopM_and_MCTopM}

% first, what is an index policy

In a single-player setting, \emph{index policies} are popular bandit algorithms: at each round one index is computed for each arm, that only depends on the history of plays of this arm and (possibly) some exogenous randomness. Then, the arm with highest index is selected. This class of algorithms includes the UCB family, in which the index of each arm is an Upper Confidence Bound for its mean, but also some Bayesian algorithms like Bayes-UCB \citep{Kaufmann12BUCB} or the randomized Thompson Sampling algorithm \citep{Thompson33,AgrawalGoyal11,Kaufmann12Thompson}.

% concrete examples and how their are used within MPB

The approaches we now describe for multi-player bandits can be used in combination with any index policy, but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained. In particular, we focus on two types of indices:
\UCB{} indices \citep{Auer02}
and \klUCB{} indices \citep{KLUCBJournal}, that can be defined for each player $j$ in the following way.
%
Letting $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of sensing information obtained by player $j$ for arm $k$, $\widehat{\mu}_k^j(t) = S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$ for player $j$ and one can define the index
\begin{equation}\label{eq:5:indexFor_UCB_klUCB}
  g_k^j(t) := \begin{cases}
      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB,
  \end{cases}
\end{equation}
where $f(t)$ is some \emph{exploration function}. $f(t)$ is usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(g_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
A classical (single-player) UCB algorithm aims at the arm with largest index. However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.

% RhoRand

While the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices  (even though some alternative without pre-agreement are proposed),
the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}. %\footnote{\rhoRand{} only works if $M$ is known, and \cite{Anandkumar11} extended it to the case of an unknown $U$, with \rhoRandEst. Extending our algorithms for the case of unknown number of players $M$ is an interesting future work.}.
%
More formally, letting $\pi(k,\mathbf{g})$ be the index of the $k$-th largest entry in a vector $\mathbf{g}$,
% (for $k\in\{1,\dots,K\}$),
in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in\{1,\dots,M\}$
and selects at time $t$,
\begin{equation}
  A^j(t) := \pi\left(R^j(t), [g^j_\ell(t)]_{\ell=1,\dots,K}\right).
\end{equation}
If a collision occurs, a new rank is drawn uniformly at random: $R^j(t+1) \sim \cU(\{1,\dots,M\})$.

% Now, our algorithms !!

We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices:
\begin{equation}
  \TopM(t) := \left\{ \pi\left(k, \{g^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
\end{equation}

Our proposal \MCTopM{} is stated below as Algorithm~\ref{algo:5:MCTopM},
while a simpler variant, called \RandTopM{},
is stated as Algorithm~\ref{algo:5:RandTopM} in Appendix~\ref{app:5:RandTopM}.
We focus on \MCTopM{} as it is easier to analyze and performs better.
%
Both algorithms ensure that player $j$ always
selects at time $t+1$ an arm from $\TopM(t)$. When a collision occurs \RandTopM{} randomly switches arm within $\TopM$, while \MCTopM{} uses a more sophisticated mechanism, that is reminiscent of ``Musical Chair'' (MC) and inspired by the work of \cite{Rosenski16}: players tend to fix themselves on arms (``chairs'') and ignore future collision when this happens.

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      % \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $\{1,\dots,K\}$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU(\{1,\dots,K\})$ and $C^j(1)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T-1$}{
                   \uIf(\tcp*[f]{transition $(3)$ or $(5)$}){
                      $A^j(t) \notin \TopM(t)$}
                    {
                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
                      \tcp*[f]{not empty} \\
                      % \tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}
                      $s^j(t+1) = \mathrm{False}$
                      \tcp*[f]{aim at an arm with a smaller UCB at $t-1$}
                    }
                    \uElseIf(\tcp*[f]{collision and not fixed}){
                        $C^j(t)$ \emph{and} $\overline{s^j(t)}$}
                      {
                        $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                        \tcp*[f]{transition $(2)$} \\
                        $s^j(t+1) = \mathrm{False}$
                    }
                    \Else(\tcp*[f]{transition $(1)$ or $(4)$}){
                      $A^j(t+1) = A^j(t)$ \tcp*[f]{stay on the previous arm} \\
                      $s^j(t+1) = \mathrm{True}$ \tcp*[f]{become or stay fixed on a ``chair''}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for a fixed underlying index policy $g^j$).}
          \label{algo:5:MCTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      % \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
\vspace*{-5pt}  % XXX remove if problem

More precisely, under \MCTopM,
% \footnote{This choice is similar to the elementary step used in the \MusicalChair{} algorithm introduced by \cite{Rosenski16}, that gave its name to \MCTopM, even if they restrict to using the empirical averages as indices (\ie, the $0$-greedy algorithm).}
if player $j$ did not encounter a collision when using arm $k$ at time $t$,
then she marks her current arm as a ``chair'' ($s^j(t+1)=\mathrm{True}$),
and will keep using it even if collisions happen in the future (Lines~$9$-$11$).
%
As soon as this ``chair'' $k$ is no longer in $\widehat{M_j}(t)$,
a new arm is sampled uniformly from a subset of $\TopM(t)$,
defined with the previous indices $g^j(t-1)$ (Lines~$3$-$5$).
%
The subset enforces a certain inequality on indices,
$g_{k'}^j(t-1) \leq g^j_{k}(t-1)$ and $g_{k'}^j(t) \geq g^j_{k}(t)$,
when switching from $k=A^j(t)$ to $k'=A^j(t+1)$.
This helps to control the number of such changes of arm,
as shown in Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}.
% the Appendix~\ref{proof:5:collisionsMCTopM}
The considered subset is never empty as it contains
at least the arm replacing the $k\in\TopM(t-1)$ in $\TopM(t)$.
Collisions are dealt with only for non-fixed player $j$,
and when the previous arm is still in $\TopM(t)$.
%
In this case, a new arm is sampled uniformly from $\TopM(t)$ (Lines~$6$-$8$).
%
%
This stationary aspect helps to minimize the number of collisions,
as well as the number of switches of arm.
%
The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM{} (see Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in Appendix~\ref{proof:5:collisionsMCTopM}).

% -----------------------------------------------------------------
\subsection{The \Selfish{} Approach}
\label{sub:5:Selfish}

Under observation model \modeltrois{} no sensing information is available and the previous algorithms cannot be used, as the sum of sensing information $S_k^j(t)$ and thus the empirical mean $\widehat{\mu}_k^j(t)$ cannot be computed, hence neither the indices $g_k^j(t)$. However, one can still define a notion of \emph{empirical reward} received from arm $k$ by player $j$, by introducing
%
\vspace*{-5pt}
\begin{equation}
  \widetilde{S_k}^j(t) = \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) := \widetilde{S_k}^j(t) \;/\; T_k^j(t).
\end{equation}

Note that $\widetilde{\mu_k}^j(t)$ is no longer meant to be an unbiased estimate of $\mu_k$ as it also takes into account the collision information, that is present in the reward. Based on this empirical reward, one can similarly defined modified indices as
%
\begin{equation}\label{eq:5:indexTilde}
  \widetilde{g_k}^j(t) = \begin{cases}
      \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widetilde{\mu_k}^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB.
  \end{cases}
\end{equation}

Given any of these two index policies (\UCB{} or \klUCB), the \Selfish{} algorithm is defined by,
%
\begin{equation}
  A^j(t) = \argmax{ k \in \{1,\dots,K\}} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
\end{equation}
The name comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
The reason that this may work precisely comes from the fact that $\widetilde{g_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
but some hybrid index that simultaneously increases when a transmission occurs and decreases when a collision occurs.

This behavior is easier to be understood for the case of \Selfish-\UCB{} in which, letting $N_k^{j,C}(t) = \sum_{s=1}^t \indic(C^j(t))$ be the number of collisions on arm $k$, one can show that the hybrid \Selfish{} index induces a penalty proportional to the fraction of collision on this arm and the quality of the arm itself:
\begin{equation}
  \widetilde{g_k}^j(t) = g_k^j(t) -
  \underbrace{\left(\frac{N_k^{j,C}(t)}{N_k^j(t)}\right)}_{\text{fraction of collisions}}
  \underbrace{\left(\frac{1}{N_k^{j,C}(t)}\sum_{t=1}^{T}Y_{A^j(t),t} \indic(C^j(t)) \indic(A^j(t) = k)\right)}_{\text{estimate of } \mu_k }.
\end{equation}

From a bandit perspective, it looks like each player is using a stochastic bandit algorithm (\UCB{} or \klUCB) when interacting with $K$ arms that give a feedback (the reward, and not the sensing information) that is far from being \iid{} from some distribution, due to the collisions.
%
As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like $\mathrm{EXP3}$ \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
%
However, we found out empirically that \Selfish{} is doing surprisingly well, as already noted by \cite{Bonnefoi17}, who did some experiments in the context of IoT applications.
%
We show in Section~\ref{sec:5:upperbounds} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
which precludes the possibility of a logarithmic regret for any problem.
However, in most cases it empirically performs similarly to all the algorithms described before,
and usually outperforms \rhoRand,
even if it neither exploits the sensing information, nor the knowledge of the number of players $M$.
%
As such, practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or not considered.

% -----------------------------------------------------------------

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Empirical performances}
\label{sec:5:experiments}

% Select one problems, max two cases ($M=K$ and $M < K$), and illustrate what we want to discuss in plots with regret in normal scale, one with semi-$\log x$ scale, and at least one histogram showing the bad luck for \Selfish...

% Selfish is awesome... unless $K=M=2$! Not only, we found other case of failure.

% Chose one problem $\boldsymbol{\mu}$, and vary $M \leq K$ for let say $K=3$ and $K=9$. Not more figures.
% Put most of them in appendix.

We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \klUCB{} indices. %used with two index policies, \UCB{} and \klUCB.
Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.
% but observations and conclusions from the experiments are discussed here.

In experiments that are not reported here, we could observe that using \klUCB{} rather than \UCB{} indices always yield better practical performance.
%, as the theoretical guarantees for single player suggested,
As the prupose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
% And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
we only report results for \klUCB.
In a first set of experiments, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
We also include a centralized multiple-play \klUCB{} algorithm,
%as defined by \cite{Anantharam87a},
essentially to check that the \emph{``price of decentralized learning''} is not too large.

%

We present results for two bandit instance: one with $K=3$ arms and means %\footnote{But of course the means are unknown to the algorithms, and their order is not important.}
$\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}. For the second instance $K=9$ and $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
three cases are presented: $M=6$ in Figure~\ref{fig:5:MP__K9_M6_T10000_N1000__4_algos},
and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
% Note the different algorithms know the number of players $M$.
Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance. %, from $t=1$ to the horizon $T=10000$ (also unknown by the algorithms),
We also include histograms showing the \emph{distribution} of regret at $t=T$,
as this allows to check if the regret is indeed small for \emph{each} run of the simulation.
%
For the plots showing the regret, our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} is displayed.
% and it should be surprising to see it being larger than some regret values: it is only asymptotic.

Experiments with a different problem for each repetition (uniformly sampled $\boldsymbol{\mu} \sim \cU([0,1]^K)$),
are also considered, in Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \ref{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}.
This helps to check that no matter the \emph{complexity} of the considered problem (one measure of complexity being the constant in our lower bound),
\MCTopM{} performs similarly or better than all the other algorithms,
and \Selfish{} outperforms \rhoRand{} in most cases.
%Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} is a good example
%of outstanding performances of \MCTopM{} and \Selfish{} in comparison to \rhoRand{}.
Empirically, our proposals were found to almost always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
we verified that \MCTopM{} outperforms the state-of-the-art algorithms in many different problems, and is more and more efficient as $M$ and $K$ grows.

%
% Regular plots of centralized regrets
%
\begin{figure}[!h]
  \centering
  % \begin{subfigure}[!h]{0.49\textwidth}
  %   \includegraphics[width=1.10\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
  % \end{subfigure}
  % % ~
  % \begin{subfigure}[!h]{0.49\textwidth}
    \includegraphics[width=0.8\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
  % \end{subfigure}
  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$. \rhoRand{} (top \textcolor{blue}{blue} curve) is outperformed by the other algorithms (and the gain increases with $M$). \MCTopM{} (bottom \textcolor{gold}{yellow}) outperforms all the other algorithms is most cases.}
  % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
  \label{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
  % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

% Dynamic settings, when $M$ can change in time,
% were considered in the analysis of both
In the presence of sensing (observation model \modeldeux), we also compared our algorithms to  %multi-player multi-armed bandits algorithms
with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}. Yet these two algorithms were found hard to use efficiently in practice and we show in
% that were also proposed
%
Figure~\ref{fig:5:MP__K9_M3_T123456_N100__8_algos} that they perform poorly in comparison to \rhoRand, \RandTopM{} and \MCTopM.
%
\MEGA{} needs a careful tuning of \emph{five} parameters ($c$, $d$, $p_0$, $\alpha$ and $\beta$) to attain reasonable performances. No good guideline for tuning them is provided and using cross validation, as suggested,
can be considered out of the scope of online sequential learning.
%In practice, on a fixed instance, the authors do not indicate how to select the parameters, even with a perfect knowledge of the parameters ($\boldsymbol{\mu}$ and $T$).
%
\MusicalChair{} consists of a random exploration phase of length $T_0$ after which the players quickly converge to orthogonal strategies targeting the $M$ best arms. With probability $1-\delta$, its regret is proved to be ``constant'' (of order $\log(1/\delta)$). The theoretical minimal value for $T_0$ depends on $\delta$, on the horizon $T$ and on a lower bound $\epsilon$ on the gap $\Delta = \mu^*_M - \mu^*_{M+1}$, and the practical tuning is hard too. %% which are both unavailable in our setting.

\begin{figure}[!h]
  \centering
  % \begin{subfigure}[!h]{0.85\textwidth}
    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
  % \end{subfigure}
  % ~
  % \begin{subfigure}[!h]{0.85\textwidth}
    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_HistogramsRegret____env1-1_8200873569864822246.pdf}
  % \end{subfigure}
  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$. \RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand. The regret of \MCTopM{} is logarithmic, empirically with the same slope as the lower bound. The $x$ axis on the regret histograms have different scale for each algorithm.}
  \label{fig:5:MP__K9_M6_T10000_N1000__4_algos}
\end{figure}

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Theoretical elements}
\label{sec:5:upperbounds}

Section~\ref{sub:5:UpperBoundSelections} gives
an asymptotically optimal analysis of the expected number of sub-optimal draws
for \RandTopM, \MCTopM{} and \rhoRand{} combined with \klUCB{} indices,
and Section~\ref{sub:5:UpperBoundCollisions} proves that the number of  collisions, hence the regret of \MCTopM{} are also logarithmic.
%
Section~\ref{sub:5:SelfishFails} shortly discusses a disappointing result regarding \Selfish{}, with more insights provided in Appendix~\ref{app:5:SelfishFails}.

% -----------------------------------------------------------------
\subsection{Common Analysis for \RandTopM- and \MCTopM-\klUCB{}}\label{sub:5:UpperBoundSelections}

Lemma~\ref{lem:5:SubOptimalSelections} gives a finite-time upper bound on the expected number of draws of a sub-optimal arm $k$ for any player $j$, that
holds for both \RandTopM-\klUCB{} and \MCTopM-\klUCB.
Our improved analysis also applies to \rhoRand{}.
Explicit expressions for $C_{\boldsymbol{\mu}}$, $D_{\boldsymbol{\mu}}$ can be found in the proof given in Appendix~\ref{proof:5:SubOptimalSelections}.

\begin{lemma}\label{lem:5:SubOptimalSelections}
  For any $\boldsymbol{\mu}\in\cP_M$,
  if player $j\in\{1,\dots,M\}$ uses the \RandTopM-, \MCTopM- or \rhoRand-\klUCB{}
  decentralized policy with exploration function $f(t) = \log(t) + 3 \log\log(t)$,
  then for any sub-optimal arm $k \in \Mworst$, there exists two problem depend constants $C_{\boldsymbol{\mu}}$, $D_{\boldsymbol{\mu}}$ such that
  % %
  % \hfill{}
  % (Proved in Appendix~\ref{proof:5:SubOptimalSelections})
  \begin{equation}\label{eq:5:SubOptimalSelections}
      \E_{\mu}[T_k^j(T)] \leq \frac{\log(T)}{\kl(\mu_k,\mu_{M}^*)}+ C_{\boldsymbol{\mu}} \sqrt{\log(T)} + D_{\boldsymbol{\mu}}\log\log(T) + 3M + 1.
  \end{equation}
\end{lemma}

It is important to notice that the leading constant in front of $\log(T)$ is the same as in the constant featured in Equation~\eqref{eq:5:LBDraws} of Theorem~\ref{thm:5:BetterLowerBound}. This result proves that the lower bound on sub-optimal selections is asymptotically matched for the three considered algorithms. This is a strong improvement in comparison to the previous state-of-the-art results
\citep{Zhao10,Anandkumar11}.

% -----------------------------------------------------------------

% For \rhoRand-\klUCB{} we used results from \cite{Anandkumar11}
% to control the number of collisions on sub-optimal arms.
% %
% For \RandTopM-\klUCB{} we have a better control over the number of collisions,
% and as expected we can bound it by a logarithmic term with a much smaller constant.
% The \RandTopM{} policy was indeed designed to be more ``conservative'' in its dynamics
% in order to reduce the number of collisions caused by a player who changes from his current chosen arm
% in an orthogonal configuration.

% For \rhoRand, \cite{Anandkumar11} essentially proved on the one hand that conditionnaly to a certain ``good'' event
% (all players have a correct ordering of all the $M$ best arms)
% proved to happen with high probability,
% the total number of collisions is upper bounded asymptotically by ${2 M - 1 \choose M}$,
% which is constant but grows very quickly as $M$ grows.
% And on the other hand, they proved that the total number of steps when this event is violated is $\bigO{\log T}$, with large constants in the asymptotic notation.

As announced, Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM} controls
the number of switches of arm that are due to the current arm leaving $\TopM(t)$,
for both \RandTopM{} and \MCTopM{}. It essentially proves that Lines~$3$-$5$ in Algorithm~\ref{algo:5:MCTopM} (when a new arm is sampled from the non-empty subset of $\TopM(t)$)
happen a logarithmic number of times. The proof of this result is given in Appendix~\ref{proof:5:elementaryLemma_RandTopM_MCTopM}.

\begin{lemma}\label{lem:5:elementaryLemma_RandTopM_MCTopM}
  For any $\boldsymbol{\mu}\in\cP_M$,
  any player $j \in \{1, \dots, M\}$ using
  \RandTopM- or \MCTopM-\klUCB,
  and any arm $k$,
  it holds that
  \begin{equation*}
    \sum_{t=1}^T
    \Pr\left(A^j(t)=k, k\notin \TopM(t)\right)
    = \left(\sum_{k', \mu_{k'} < \mu_k}\frac{1}{\kl(\mu_k,\mu_{k'})} + \sum_{k', \mu_{k'} > \mu_k}\frac{1}{\kl(\mu_{k'},\mu_{k})}\right) \log(T) + o(\log(T)).
  \end{equation*}
\end{lemma}

% -----------------------------------------------------------------
\subsection{Regret Analysis of \MCTopM-\klUCB}\label{sub:5:UpperBoundCollisions}

For \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
that we can then bound by the number of changes of arms
(\cf{} proof in Appendix~\ref{proof:5:collisionsMCTopM}).
%
An interesting consequence of the proof of this result is that
it also bounds the number of \emph{switches of arms}, $\sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$,
and this additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
Even though minimizing switching was not a goal\footnote{Introducing \emph{switching costs}, like it was done in previous works, \eg, \cite{Koren17}, could be an interesting future work.},
this guarantee is interesting for Cognitive Radio applications,
where switching arms means reconfiguring a radio hardware, an operation that costs energy.
%and thus an algorithm guaranting a small number of switches is interesting.

\begin{lemma}\label{lem:5:collisionsMCTopM}
  For any $\boldsymbol{\mu}\in\cP_M$,
  if all players use the
  \MCTopM-\klUCB{} decentralized policy,
  and $M \leq K$,
  then the total average number of collisions (on all arms)
  is upper-bounded by
  \hfill{}
  (Proved in Appendix~\ref{proof:5:collisionsMCTopM})
  \begin{equation}
    \E_{\mu}\left[\sum_{k=1}^K \cC_k(T)\right]
    \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
  \end{equation}
\end{lemma}

Note that this bound is in $\bigO{M^3}$,
which significantly improves the $\bigO{M{2M-1 \choose M}}$ proved by \cite{Anandkumar11} for \rhoRand. It is worse than the $\bigO{M^2}$ proved by \cite{Rosenski16} for \MusicalChair{}. %, due to our trick of focusing on collisions for non-sitted players.
However, unlike \MusicalChair{} our algorithm does not need the knowledge of $\mu^*_{M}-\mu^*_{M+1}$.

% -----------------------------------------------------------------
% \subsection{Logarithmic Regret for \MCTopM}\label{sub:5:Regret}

Now that the sub-optimal arms selections and the collisions
are both proved to be at most logarithmic in Lemmas~\ref{lem:5:SubOptimalSelections} and \ref{lem:5:collisionsMCTopM},
it follows from our regret decomposition (Lemma~\ref{lem:5:DecompositionRegret}) together with Lemma~\ref{lem:5:1stUpperBound} that the regret of \MCTopM-\klUCB{} is logarithmic. More precisely, one obtains a finite-time problem-depend upper bound on the regret of this algorithm.

\begin{theorem}\label{thm:5:LogarithmicRegret_MCTopMklUCB}
  If all $M$ players use
  \MCTopM-\klUCB, and $M \leq K$,
  then for any problem $\boldsymbol{\mu} \in \cP_M$,
  there exists a problem dependent constant $G_{M,\boldsymbol{\mu}}$, such that
  the regret satisfies:
  \begin{equation}\label{eq:5:LogarithmicRegret_MCTopMklUCB}
    R_T(\boldsymbol{\mu}, M, \rho) \leq G_{M,\boldsymbol{\mu}} \log(T) + \smallO{\log T}.
  \end{equation}
\end{theorem}

% -----------------------------------------------------------------
\subsection{Discussion on \Selfish} \label{sub:5:SelfishFails}

The analysis of \Selfish{} is harder, but we tried our best to obtain some understanding of the behavior of this algorithm, that seems to be doing surprisingly well in many contexts, as in our experiments with $K=9$ arms and in extensive experiments not reported in this Section. However, a disappointing result is that we found simple problems, usually with small number of arms, for which the algorithm may fail. For example with $M=2$ or $M=3$ players competing for $K=3$ arms,
with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, the histograms in Figure~\ref{fig:5:selfish_fail1} suggests that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} can be very large. We provide a discussion in Appendix~\ref{app:5:SelfishFails} about when such situations may happen, including a conjectured (constant, but small) lower bound on the probability that \Selfish{} experience collision almost at every round. This result would then prevent \Selfish{} from having a logarithmic regret. However, it is to be noted that the lower bound of Theorem~\ref{thm:5:BetterLowerBound} does not apply to the censored observation model \modeltrois{} under which \Selfish{} operates, and it is not known yet whether logarithmic regret is at all possible.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Conclusion and future work}
\label{sec:5:conclusion}

To summarize, we presented three variants of Multi-Player Multi-Arm Bandits,
with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
For the two easiest models --with sensing--, our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret. In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.

This work suggests several interesting further research directions. First, we want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information. So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality. However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}. Second, it remains an open question to know if a simple decentralized algorithm can be as efficient as \MCTopM{} without knowing $M$ in advance, or in dynamic settings (when $M$ can change in time). We shall start by proposing variants of our algorithm that are inspired by the \rhoRandEst{} variant of \rhoRand{} proposed by \cite{Anandkumar11}.
Finally, we want to strengthen the guarantees obtained in the absence of sensing, that is to know whether logarithmic regret is achievable and to have a better analysis of the \Selfish{} approach.  Indeed, in most cases, it performs comparably to \RandTopM{} even with limited feedback and without knowing the number of players $M$, which makes it a good candidate for applications to Internet of Things networks.

% -----------------------------------------------------------------

% \paragraph{Acknowledgments}
% Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.

\paragraph{A note on the simulation code}
The page \texttt{SMPyBandits.GitHub.io/MultiPlayers.html} explains how to reproduce the experiments used for this Section.

% biblio usually before the supplementary material
% \bibliography{mpBandits}

### 2019-02-14 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Added content for Chapter 5, based on my ALT paper « Multi-Player Bandits Revisited », see https://hal.inria.fr/hal-01629733

% -----------------------------------------------------------------
\section{Regret Decompositions}
\label{app:5:regretdec}

% -----------------------------------------------------------------
\subsection{Proof of Lemma~\ref{lem:5:DecompositionRegret}}
\label{proof:5:DecompositionRegret}

    Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
    \begin{align*}
      R(T)
      &= \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M Y_{A^j(t),t} \eta^j(t) \right]
       = \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M \mu_{A^j(t)} \eta^j(t)\right]
      \shortintertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence from $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$). So $\E_{\mu}[Y_{A^j(t),t} \eta^j(t)] = \sum_{k} \E_{\mu}[\mu_k \indic(A^j(t),t) \eta^j(t)] = \E_{\mu}[\mu_{A^j(t)} \eta^j(t)]$. And so}
      R(T)
      &= \E_{\mu}\left[ \sum_{t=1}^{T}\sum_{j \in \Mbest} \mu_j
        - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
        % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
      &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
        - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ T^j_k(T) \right]
        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
      %
      \shortintertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
      %
      &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ T^j_k(T) \right]
        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
      %
      \shortintertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
      %
      &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
        + \overline{\Delta_{M^*}} \E_{\mu}\left[ T_{M^*}(T) \right] \\
        &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
      %
      \shortintertext{But for $k = M^*$, $T_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ T_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ T_k(T) \right]$, so by recombining the terms, we obtain,}
      %
      &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
        + \overline{\Delta_{M^*}} T M^* \\
        &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
    \end{align*}
    %
    The term $\overline{\Delta_k} - \overline{\Delta_{M^*}}$ simplifies to $\mu_{M^*} - \mu_k$, and so $\overline{\Delta_{M^*}} = \frac{1}{M} \sum_{k=1}^{M} \mu_k - \mu_{M^*}$ by definition of $\overline{\mu}^*$. And for $k=M^*$, $\mu_{M^*} - \mu_k = 0$, so the first sum can be written for $k = 1,\dots,M$ only, so
    %
    % \vspace*{5pt}
    \begin{align*}
      R(T)
      &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
        + \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) T \\
        &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right]
    \end{align*}
    And so we obtain the decomposition with three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}.
    \begin{align*}
    R(T)
      &= \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ T_k(T) \right]\right)
        + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
    \end{align*}
    % Which is exactly the decomposition we wanted to prove.

% -----------------------------------------------------------------
\subsection{Proof of Lemma~\ref{lem:5:1stLowerBound}}
\label{proof:5:1stLowerBound}

  Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
  but it is not obvious for \ref{eq:5:term2} as there is no reason for $T_k(T)$ to be upper bounded by $T$.
  %
  Let $T_k^{!}(T) := \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
  where the notation $\exists!$ stands for ``there exists a unique''.
  Then $T_k(T) = \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(A^j(t) = k)$ can be decomposed as
  \begin{equation*}
    T_k(T) = \sum_{t=1}^{T} \indic(\exists! j, A^j(t) = k) + \sum_{t=1}^{T} \sum_{j=1}^{M} c_{k,t} \indic(A^j(t) = k)
    = T_k^{!}(T) + C_k(T).
  \end{equation*}
  By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \rho)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
  \begin{align*}
    \ref{eq:5:term2} + \ref{eq:5:term3} &=
    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
    + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)] \\
    & \;\;\;\;\;\; + \sum_{k=1}^{M} \mu_k \E_{\mu}[C_k(T)]
    - \sum_{k \in \Mbest} \mu_k \E_{\mu}[C_k(T)] \\
    &=
    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
    + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)]
    + \sum_{k \in \Mworst} \mu_k \E_{\mu}[C_k(T)] \\
    &=
    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
    + \sum_{k=1}^{M} \min(\mu_M^*, \mu_k) \E_{\mu}[C_k(T)].
  \end{align*}
  And now both terms are non-negative, as $T_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
  which proves that $R_T(\boldsymbol{\mu}, M, \rho) \geq \ref{eq:5:term1}$, as wanted.

% -----------------------------------------------------------------
\subsection{Proof of Lemma~\ref{lem:5:1stUpperBound}}
\label{proof:5:1stUpperBound}

Recall that we want to upper bound
$ (b) : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[T_k(T)]\right)$.
First, we observe that, for all $k\in \Mbest$,
\begin{eqnarray*}
T - \E_{\mu}[T_k(T)] & \leq & T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],\end{eqnarray*}
where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
\begin{eqnarray*}
 (b) & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[T_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
 & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
\end{eqnarray*}
The quantity $\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)$ counts the number of optimal arms that have not been selected at time $t$. For each mis-selection of an optimal arm, there either exists a sub-optimal arm that has been selected, or an arm in $\Mbest$ on which a collision occurs. Hence
\[\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t) = \sum_{k \in \Mbest}\indic(C_k(t)) + \sum_{k \in \Mworst} \indic(\exists j : A^j(t) = k),\]
which yields
\[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[T_k(T)\right]\]
and Lemma~\ref{lem:5:1stUpperBound} follows.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Lower Bound: Proof of Theorem~\ref{thm:5:BetterLowerBound}}
\label{proof:5:BetterLowerBound}

% -----------------------------------------------------------------
\subsection{Proof of Theorem~\ref{thm:5:BetterLowerBound}}

The lower bound that we present relies on the following \emph{change-of-distribution} lemma that we prove in the next section, following recent arguments from \cite{Garivier16TrueShape} that have to be adapted to incorporate the collision information.

\begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-player bandit models denoted by $\mu$ and $\lambda$ respectively, it holds that
\begin{equation}
  \sum_{k=1}^{K}\E_{\mu}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\mu}(A),\Pr_{\lambda}(A)\right).
\end{equation}
\end{lemma}

Let $k$ be a sub-optimal arm under $\mu$,
fix $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$,
% We define the bandit instance $\lambda$ such that
and let $\lambda$ be the bandit instance such that
\[\left\{\begin{array}{ccl}
          \lambda_\ell & = & \mu_\ell \ \ \  \text{for all } \ell \neq k, \\
          \lambda_k & = & \mu_{M^*} + \varepsilon.
         \end{array}
\right.\]
Clearly, $\lambda\in\cP_M$ also,
and the set of $M$ best arms under $\mu$ and $\lambda$ differ by one arm: if \Mbest$_{\mu} = \{1^*,\dots,M^*\}$ then $\Mbest_{\lambda} = \{1^*,\dots,(M-1)^*,k\}$.
Thus, one expects the ($\cF^j_T$-mesurable) event
\[A_T = \left(T_k^j(T) > \frac{T}{2M}\right)\]
to have a small probability under $\mu$ (under which $k$ is sub-optimal) and a large probability under $\lambda$ (under which $k$ is one of the optimal arms, and is likely to be drawn a lot).

Applying the inequality in Lemma~\ref{lem:5:CD}, and noting that the sum in the left-hand side reduces to on term as there is a single arm whose distribution is changed, one obtains
\begin{eqnarray*}
  \E_{\mu}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \kl \left(\Pr_{\mu}(A_T),\Pr_{\lambda}(A_T)\right),\nonumber \\
  \E_{\mu}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - \Pr_{\mu}(A_T)\right) \log \left(\frac{1}{\Pr_{\lambda}(\overline{A_T})}\right) - \log(2),
\end{eqnarray*}
using the fact that the binary KL-divergence satisfies $\kl(x,y) = \kl(1-x,1-y)$ as well as the inequality  $\kl(x, y) \geq x \log\left(1/y\right) - \log(2)$, proved by \cite{Garivier16TrueShape}.
Now, using Markov inequality yields
\begin{align*}
  \Pr_{\mu}\left(A_T\right)
    & \leq 2M\frac{\E_\mu\left[T^j_k(T)\right]}{T} =: x_T, \\
  \Pr_{\lambda}\left(\overline{A_T}\right)
    & = \Pr_{\lambda}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_\lambda\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
\end{align*}
which defines two sequences $x_T$ and $y_T$, such that
\begin{eqnarray}\label{eq:5:CqCD}
  \E_{\mu}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - x_T\right) \log \left(\frac{T}{y_T}\right) - \log(2).
\end{eqnarray}
The strong uniform efficiency assumption  (see Definition~\ref{def:5:DecentralizedUniformEfficiency}) further tells us that $x_T \rightarrow 0$ (as $\E_\mu[T_k^j(T)] = o(T^\alpha)$ for all $\alpha$) and $y_T = o(T^\alpha)$ when $T \to \infty$, for all $\alpha \in (0,1)$.
As a consequence, observe that $\log(y_T)/\log(T) \rightarrow 0$ when $T$ tends to infinity and
\[
  \frac{\left(1 - x_T\right) \log \left({T}/{y_T}\right) - \log(2)}{\log(T)} = 1-x_T - (1-x_T) \frac{\log(y_T)}{\log(T)} - \frac{\log(2)}{\log(T)}
\]
tends to one when $T$ tends to infinity.
From Equation~\eqref{eq:5:CqCD}, this yields
\begin{equation}\label{eq:5:CqCD2}
  \liminf_{T\to\infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k,\mu_{M^*}+\varepsilon)},
\end{equation}
for all $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$.
Letting $\varepsilon$ go to zero gives the conclusion (as $\kl$ is continuous).

% -----------------------------------------------------------------

\subsection{Proof of Lemma~\ref{lem:5:CD}}

Under observation model \modelun{} and \modeldeux, the strategy $A^j(t)$ decides which arm to play based on the information contained in $I_{t-1}$, where $I_0 = U_0$ and
\[\forall t > 0, \ \ I_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $\{1,\dots,M\}$ for new ranks or arms.} useful to select $A^j(t)$.
Formally, one can say that $A^j(t)$ is $\sigma(I_{t-1})$ measurable\footnote{$\sigma(I_t)$ denotes the sigma-Algebra generated by the observations $I_t$.} (as $\cF^j(t) \subseteq \sigma(I_{t})$, with an equality under observation model \modelun).

Under two bandit models $\mu$ and $\lambda$, we let $\Pr_{\mu}^{I_t}$ (resp. $\Pr_{\lambda}^{I_t}$) be the distribution of the observations under model $\mu$ (resp. $\lambda$), given a fixed algorithm.
Using the exact same technique as \cite{Garivier16TrueShape} (the contraction of entropy principle), one can establish that for any event $A$ that is $\sigma(I_t)$-measurable\footnote{In the work of \cite{Garivier16TrueShape}, the statement is more general and the probability of an event $A$ is replaced by the expectation of any $\cF_T$-measurable random variable $Z$ bounded in $[0,1]$.},
\[\KL\left(\Pr_{\mu}^{I_t},\Pr_{\lambda}^{I_t}\right) \geq \kl\left(\Pr_{\mu}(A),\Pr_{\lambda}(A)\right).\]

The next step is to relate the complicated KL-divergence $\KL\left(\Pr_{\mu}^{I_t},\Pr_{\lambda}^{I_t}\right)$ to the number of arm selections.
Proceeding similarly as \cite{Garivier16TrueShape}, one can write, using the chain rule for KL-divergence, that
\begin{equation}\label{eq:5:Iterative}\KL\left(\Pr_{\mu}^{I_{t}},\Pr_{\lambda}^{I_{t}}\right) =  \KL\left(\Pr_{\mu}^{I_{t-1}},\Pr_{\lambda}^{I_{t-1}}\right) + \KL\left(\Pr_{\mu}^{Y_{t},C_t,U_t | I_{t-1}},\Pr_{\lambda}^{Y_{t},C_t,U_t | I_{t-1}}\right).\end{equation}
Now observe that conditionally to $I_{t-1}$, $U_t$, $Y_t$ and $C_t$ are independent, as once the selected arm is known, the value of the sensing $Y_t$ does not influence the other players selecting that arm, and $U_t$ is some exogenous randomness.
Using further that the distribution of $U_t$ is the same under $\mu$ and $\lambda$, one obtains
\begin{equation}\KL\left(\Pr_{\mu}^{Y_{t},C_t | I_{t-1}},\Pr_{\lambda}^{Y_{t},C_t | I_{t-1}}\right) =
 \KL\left(\Pr_{\mu}^{Y_{t}| I_{t-1}},\Pr_{\lambda}^{Y_{t}| I_{t-1}}\right) + \KL\left(\Pr_{\mu}^{C_t | I_{t-1}},\Pr_{\lambda}^{C_t | I_{t-1}}\right).\label{eq:5:DecKL}
\end{equation}

The first term in \eqref{eq:5:DecKL} can be rewritten using the same argument as \cite{Garivier16TrueShape}, that relies on the fact that conditionally to $I_{t-1}$, $Y_t$ is a Bernoulli distribution with mean $\mu_{A^j(t)}$ under the instance $\mu$ and $\lambda_{A^j(t)}$ under the instance $\lambda$:
\begin{eqnarray*}
\KL\left(\Pr_{\mu}^{Y_{t}| I_{t-1}},\Pr_{\lambda}^{Y_{t}| I_{t-1}}\right)  & = & \E_\mu\left[\E_\mu\left[ \KL\left(\Pr_{\mu}^{Y_{t}| I_{t-1}},\Pr_{\lambda}^{Y_{t}| I_{t-1}}\right) | I_{t-1} \right]\right] \\
& = & \E_\mu \left[\kl \left(\mu_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
& = & \E_\mu  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right],
\end{eqnarray*}
We now show that second term in \eqref{eq:5:DecKL} is zero:
\begin{eqnarray*}
\KL\left(\Pr_{\mu}^{C_t | I_{t-1}},\Pr_{\lambda}^{C_t | I_{t-1}}\right)
& = & \E_\mu\left[\E_\mu\left[ \KL\left(\Pr_{\mu}^{C_t| I_{t-1}},\Pr_{\lambda}^{C_t| I_{t-1}}\right) | I_{t-1} \right]\right] \\
& = & \E_\mu\left[\E_\mu\left[\E_\mu\left[ \KL\left(\Pr_{\mu}^{C_t| I_{t-1}},\Pr_{\lambda}^{C_t| I_{t-1}}\right) | \bigcup_{j' \neq j} I_{t-1}^{j'}\right]| I_{t-1} \right]\right], \\
\end{eqnarray*}
where $I_{t-1}^{j'}$ denote the information available to player $j' \neq j$.
Knowing the information available to all other players player $C_t$ is an almost surely constant random variable, whose distribution is the same under $\mu$ and $\lambda$.
Hence the inner expectation is zero and so does $\KL\left(\Pr_{\mu}^{C_t | I_{t-1}},\Pr_{\lambda}^{C_t | I_{t-1}}\right)$.

Putting things together, we showed that
\[\KL\left(\Pr_{\mu}^{I_t},\Pr_{\lambda}^{I_t}\right) = \KL\left(\Pr_{\mu}^{I_{t-1}},\Pr_{\lambda}^{I_{t-1}}\right) + \E_\mu  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right].\]
Iterating this equality and using that $\KL\left(\Pr_{\mu}^{I_{0}},\Pr_{\lambda}^{I_{0}}\right)=0$ yields that
\[\sum_{k=1}^{K}\E_{\mu}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\mu}(A),\Pr_{\lambda}(A)\right),\]
for all $A \in \sigma(I_T)$, in particular for all $A \in \cF^j_T$.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{The \RandTopM{} algorithm}
\label{app:5:RandTopM}

We now state precisely the \RandTopM{} algorithm below in Algorithm~\ref{algo:5:RandTopM} (page \pageref{algo:5:RandTopM}).
It is essentially the same algorithm as \MCTopM,
but in a simpler version as the ``Chair'' aspect is removed, that is, there is no notion of state $s^j(t)$ (\cf{} Algorithm~\ref{algo:5:MCTopM}).
%
Player $j$ is always considered ``not fixed'',
and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$
in the case of \RandTopM.

\vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      % \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $\{1,\dots,K\}$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU(\{1,\dots,K\})$ and $C^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T - 1$}{
                  %
                  \eIf{$A^j(t) \notin \TopM(t)$}{
                    \eIf(\tcp*[f]{collision}){$C^j(t)$}{
                      $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                      \tcp*[f]{randomly switch}
                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}){
                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
                      }
                    }{
                      $A^j(t+1) = A^j(t)$
                      \tcp*[f]{stays on the same arm}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for a fixed underlying index policy $g^j$).}
          \label{algo:5:RandTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      % \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
\vspace*{-5pt}  % XXX remove if problem

% -----------------------------------------------------------------
% -----------------------------------------------------------------

\section{Proofs Elements Related to Regret Upper Bounds}
\label{proof:5:RegretUpperBounds}

This Appendix includes the main proofs, missing from the content of the article,
that yield the regret upper bound.
We start by controlling the sub-optimal draws when the \klUCB{} indices are used (instead of \UCB),
with any of our proposed algorithms (\MCTopM, \RandTopM) or \rhoRand{}. Then we focus on controlling collisions for \MCTopM-\klUCB.

\subsection{Control of the Sub-optimal Draws for \klUCB: Proof of Lemma~\ref{lem:5:SubOptimalSelections}}
\label{proof:5:SubOptimalSelections}

%
%     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
%     As only user $j$ is considered here, the superscript $j$ is dropped.
%
Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
\begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.

Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
\begin{align*}
\E_{\mu}[T_k^j(T)]
&= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
= \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
%&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
%
\shortintertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
\E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
&\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
&\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
%
&\leq
\sum_{t=1}^T
  \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+
\sum_{m=1}^M
  \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right),
\end{align*}
where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.

Now each of the two terms in the right hand side can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
The rightmost term can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact 1 in Appendix A of \cite{KLUCBJournal}.
The leftmost term can be controlled using Lemma~\ref{lem:5:Fact2KLUCB} stated below, that is a direct consequence of the proof of Fact 2 in Appendix A of \cite{KLUCBJournal}.

\begin{lemma}\label{lem:5:Fact1KLUCB}
    For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
    \begin{equation}
      \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
    \end{equation}
\end{lemma}

Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).

\begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
\begin{align*}
  &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
  + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
\end{lemma}

Putting things together, one obtains the non-asymptotic upper bound
\begin{align}\label{eq:5:UBprecise}
  \E_{\mu}\left[T_k^j(T)\right]
  & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
  & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
\end{align}
which yields Lemma~\ref{lem:5:SubOptimalSelections},
with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.

% --------------------------------------------------------------------------------
\subsection{Proof of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}}\label{proof:5:elementaryLemma_RandTopM_MCTopM}

Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
\begin{align*}
  & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
  & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
  & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
\end{align*}

Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
\[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}. To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
\begin{align}
  & \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)
  =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
  & \hspace{1cm} \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
  & \hspace{1cm}  \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
\end{align}
where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
\[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]

If $\mu_k > \mu_{k'}$, we rather use that
\[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
\[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
Summing over $k'$ yields the result.

%%%%%%%%%%%%ù PREVIOUS VERSION
% For any arms $k, k' \in \{1, \dots, K\}$,
% let $\phi_{k,k'}^j(t)$ be the event
% \begin{equation*}
%   % \phi_{k}^j(t) :=
%   \left(
%     % \overline{C^j(t-1)} \cap
%     A^j(t-1) = k \in \TopM(t-1),
%     k \notin \TopM(t),
%     C^j(t),
%     A^j(t) = k' \in \TopM(t) \cap \{ m : g_m^j(t-2) \leq g_k^j(t-2) \}
%   %
%   \right).
% \end{equation*}
%
%
% If $k$ or $k' \in \Mworst$, the desired result directly comes from
% Lemma~\ref{lem:5:SubOptimalSelections} controlling the selections of worst arms,
% as $\phi_{k,k'}^j(t)$ implies $A^j(t) = k'$ and $A^j(t-1) = k$,
% and so $\sum_{t=1}^T \Pr(\phi_{k,k'}^j(t)) \leq \min(\E_{\mu}[T_k^j(T-1)], \E_{\mu}[T_{k'}^j(T-1)])$,
% and at least one term is $\bigO{\log T}$ from Lemma~\ref{lem:5:SubOptimalSelections}
% and both are always $\bigO{T}$,
% so $\sum_{t=1}^T \phi_{k,k'}^j(t) = \bigO{\log T}$.
%
% If one of the two arms $k, k'$ is not sampled enough, \ie, if $\E_{\mu}[T_m^j(T)]\not\to \infty$ when $T\to\infty$ for $m=k$ or $m=k'$, then these sums are $\sum_{t=1}^T \Pr(\phi_{k,k'}^j(t)) = \bigO{1} = \smallO{\log T}$ and we are done.
%
% Now assume that $k \in \Mbest$ and $k' \in \Mbest$,
% and assume that both arms are sampled enough.
% %
% This event $\phi_{k,k'}^j(t)$ we focus on implies both
% $g_{k'}^j(t-1) \leq g_k^j(t-1), A^j(t-1)=k$
% and
% $g_k^j(t) \leq g_{k'}^j(t), A^j(t)=k'$.
% %
% In particular,
% if $\mu_k \geq \mu_{k'}$,
% $\phi_{k,k'}^j(t)$ implies $g_k^j(t) \leq g_{k'}^j(t)$,
% and
% if $\mu_k < \mu_{k'}$, then
% $\phi_{k,k'}^j(t)$ implies $g_{k'}^j(t-1) \leq g_k^j(t-1)$,
% both with $A^j(t-1)=k, A^j(t)=k'$.
% Both cases are dealt with exactly the same arguments,
% so let focus on the first one.
%
% Assume $\mu_k \geq \mu_{k'}$.
% Then, essentially $g_k^j(t) \leq g_{k'}^j(t)$
% is unlikely (\ie, the sum of probabilities is $\bigO{\log T}$)
% as the indices $g_m^j(t)$ tends to $\mu_m$ if arm $m$ is sampled enough (\ie, $\E_{\mu}[T_m^j(T)] \to \infty$).
% \begin{align}
%     \sum_{t=1}^{T} \Pr(\phi_{k,k'}^j(t))
%     &\leq
%     \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t), A^j(t-1)=k, A^j(t)=k'\right)
%     \intertext{Either $g_{k'}^j(t) > \mu_k$ or $g_{k'}^j(t) \leq \mu_k$, and by introducing the two cases,}
%     &\leq
%     \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t) \leq \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag \\
%     &\;\;+ \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t), g_{k'}^j(t) > \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag
%     \intertext{For the first term, $g_k^j(t) \leq \mu_k$ is highly unlikely (\ie, the sum of probabilities is $\smallO{\log T}$) as the indices ${(g_k^j)}_k$ are Upper Confidence Bounds (if arm $k$ is sampled enough). For the second term, $g_{k'}^j(t) > \mu_k$ is unlikely if arm $k'$ is sampled enough. More formally,}
%     \sum_{t=1}^{T} \Pr(\phi_{k,k'}^j(t))
%     &\leq
%     \underbrace{\sum_{t=1}^{T} \Pr\left( g_k^j(t) \leq \mu_k \right)}_{(i)}
%     + \underbrace{\sum_{t=1}^{T} \Pr\left( g_{k'}^j(t) > \mu_k, A^j(t)=k' \right)}_{(ii)}.
% \end{align}
% But then as both arms are sampled enough, we can use both Lemma~\ref{lem:5:Fact1KLUCB} to directly obtain that the first term $(i)$ is a $\smallO{\log T}$,
% and Lemma~\ref{lem:5:Fact2KLUCB} to obtain that the second  $(ii)$ is a $\bigO{\log T}$.
%
%
% Exactly the same arguments work for the other case $\mu_k < \mu_{k'}$, by considering
% whether $g_k^j(t-1) > \mu_{k'}$ (which is also ulikely),
% or whether $g_k^j(t-1) \leq \mu_k$ which implies $g_{k'}^j(t-1) \leq \mu_{k'}$ (and which is also highly unlikely), and with Lemmas~\ref{lem:5:Fact1KLUCB}, \ref{lem:5:Fact2KLUCB},
% we obtain that
% $\sum_{t=1}^{T} \Pr(\phi_{k,k'}^j(t)) = \bigO{\log T}$ when $T \to \infty$
% similarly.
%
% Summing up for all $k' \neq k$, splitting between $\mu_{k'} < \mu_k$ or $\mu_{k'} > \mu_k$,
% and using the precise constants from Lemmas~\ref{lem:5:SubOptimalSelections}, \ref{lem:5:Fact1KLUCB}, \ref{lem:5:Fact2KLUCB} concludes.
% \vspace*{-20pt}  % XXX remove if problem

% --------------------------------------------------------------------------------
\subsection{Controlling Collisions for \MCTopM: Proof of Lemma~\ref{lem:5:collisionsMCTopM}}
\label{proof:5:collisionsMCTopM}

A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small. Its proof is postponed to the end of this section.

%
Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below provides a schematic representation of the execution of the \MCTopM{} algorithm, that has to be exploited in order to properly control the number of collisions.
%
The sketch of the proof is the following: by focusing only on collisions in the ``not fixed'' state, bounding the number of transitions $(2)$ and $(3)$ is enough.
Then, we show that both the number of transitions $(3)$ and $(5)$ are small: as a consequence of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, the average number of these transitions is $\bigO{\log T}$.
Finally, we use that the length of a sequence of consecutive transitions $(2)$ is also small (on average smaller than $M$), and except for possibly the first one, starting a new sequence implies a previous transition $(3)$ or $(5)$ to arrive in the state ``not fixed''. This gives a logarithmic number of transitions $(2)$ and $(3)$, and so gives $\E_{\mu}[\sum_k\cC_k(T)] = \bigO{\log T}$,
with explicit constants depending on $\boldsymbol{\mu}$ and $M$.

\begin{figure}[h!]
  \begin{tikzpicture}[>=latex',line join=bevel,scale=3.8]
      %
      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
      %
      \draw [black,->] (start) -> (notfixed.20);
      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
      %
  \end{tikzpicture}
  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
  \label{fig:5:StateMachineAlgorithm_MCTopM}
\end{figure}

% Explanation of the figure

As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
% following the algorithm of \MCTopM:
\begin{equation}
    s^j(t+1) =
    \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
    \cap \left( A^j(t) \in \TopM(t) \right).
\end{equation}

For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words. At step $t$, if player $j$ is not fixed ($\overline{s^j(t)}$), she can have three behaviors when executing \MCTopM.
She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
or she stays in state $\overline{s^j(t)}$,
with two cases.
Either she sampled $A^j(t+1)$ uniformly
from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
with transition $(3)$,
in case of collision and if $A^j(t+1) \in \TopM(t)$,
or she sampled $A^j(t+1)$ uniformly
from $\TopM(t)$ with transition $(2)$,
if $A^j(t+1) \notin \TopM(t)$.
In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
%
% The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
% and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
% and experienced a collision at time $t-1$
Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).

For player $j$ and round $t$, we now introduce a few events that are useful in the proof. First, for every $x=1,2,3,4,5$, we denote $I_x^j(t)$ the event that a transition of type $(x)$ occurs for player $j$ after the first $t$ observations (\ie, between round $t$ and round $t+1$, to decide $A^j(t+1)$).
Formally they are defined by
\begin{align*}
  I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
  % go sitted
  I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
  \ \ \ &\text{and} \ \ \
  I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
  % if sitted
  I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
  \ \ \ &\text{and} \ \ \
  I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
\end{align*}
%
Then, we introduce $\widetilde{C^j}(t)$ as the event that a collision occurs for player $j$ at round $t$ if she is not yet fixed on her arm, that is
\begin{equation}
    \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
\end{equation}

A key observation is that $C^j(t)$ implies $\bigcup_{j'=1}^M \widetilde{C^{j'}}(t)$, as a collision necessarily involves at least one player not yet fixed on her arm ($\overline{s^{j'(t)}}$).
Otherwise, if they are all fixed, \ie, for all $j$, $s^j(t)$, then by definition of $s^j(t)$, none of the player changed their arm from $t-1$ to $t$, and none experienced any collision at time $t-1$ so by induction there is no collision at time $t$.
%
Thus, $\sum_{j=1}^M \Pr(C^j(t))$ can be upper bounded by $M \sum_{j=1}^M \Pr(\widetilde{C^j}(t))$ (union bound),
and it follows that if $\cC(T) := \sum_{k=1}^K \cC_k(T)$ then
\[\E_{\mu}[\cC(T)] \leq M \sum_{j=1}^M \sum_{t=1}^T \Pr(\widetilde{C^j}(t)).\]
We can further  observe that $\widetilde{C^j}(t)$ implies a transition $(2)$ or $(3)$, as a transition $(1)$ cannot happen in case of collision. Thus another union bound gives
%
\begin{align}
  \sum_{t=1}^T \Pr(\widetilde{C^j}(t))
  &\leq \sum_{t=1}^T \Pr(I_2^j(t))   + \sum_{t=1}^T \Pr(I_3^j(t)).\label{eq:5:UBTilde}
\end{align}
In the rest of the proof we focus on bounding the number of transitions $(2)$ and $(3)$.

Let $N_x^j(T)$ be the random variable denoting the number of transitions of type $(x)$.
Neglecting the event $\overline{s^j(t)}$ for $x=3$ and $s^j(t)$ for $x=5$, one has
\begin{equation}
    \E_{\mu}[N_x^j(t)]
    = \sum_{t=1}^T \Pr(I_x^j(t))
    \leq \sum_{t=1}^T \Pr \left(A^j(t) \notin \TopM(t)\right)
    \leq \sum_{t=1}^T \sum_{k=1}^K \Pr \left(A^j(t)=k, k \notin \TopM(t)\right),
\end{equation}
which is $\bigO{\log T}$ (with known constants) by Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}. In particular, this controls the second term in the right hand side of \eqref{eq:5:UBTilde}.

To control the first term $\sum_{t=1}^T \Pr(I_2^j(t))$
we introduce three sequences of random variables,
the starting times $(\theta_i)_{i \geq 1}$
and the ending times $(\tau_i)_{i \geq 1}$
(possibly larger than $T$),
of sequences during which $I_2(s)$ is true for all $s=\theta_i,\dots,\tau_i-1$ but not before and after,
that is
$\forall i \in \{1,\dots,n(T)\},
\overline{I_2^j(\theta_i - 1)}
\cap \bigcap_{t=\theta_i}^{\tau_i-1} I_2^j(t)
\cap \overline{I_2^j(\tau_i)}
$
with $n(T)$ the number of such sequences,
\ie,
$n(T) := \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
(or $0$ if $\theta_1$ does not exist).
%
If $\theta_i = 1$, the first sequence does not have term $\overline{I_2^j(\theta_i - 1)}$.

Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
\begin{align}
  \E_{\mu}[N_2^j(t)]
  &= \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
  =
  \E_{\mu}\left[ \sum_{i=1}^{n(T)} \left( \sum_{t=\theta_i}^{\tau_i - 1} 1 + \sum_{t=\tau_i}^{\theta_{i+1} - 1} 0 \right) \right]
  =
  \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
  \intertext{Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$), and so we can obtain}
  % so we can use Wald's Lemma \citep{Wald45}
  &\leq \E_{\mu}\left[ n(T) \right] \times \max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]
  = (\alpha) \times (\beta).
\end{align}

$(\alpha)$ To control $\E_{\mu}[n(T)]$, we can observe that
the number of sequences $n(T)$ is smaller than $1$ plus the number of times when \emph{a sequence begins} ($1$ plus because maybe the game starts in a sequence).
%
And beginning a sequence
% \footnote{Similarly, one could focus on the number of times \emph{a sequence finishes}, and $I_2^j(\theta_i) \cap \overline{I_2^j(\theta_i+1)}$ implies a transition of type $(3)$ or $(1)$.
% The key is that after a transition $(1)$, a new sequence of transitions $(2)$ is possible only after a transition $(5)$, which total number is also controlled as $\bigO{\log T}$.}
at time $\theta_i$ implies
$\overline{I_2^j(\theta_i-1)} \cap I_2^j(\theta_i)$,
which implies a transition of type $(3)$ or $(5)$ at time $\theta_i - 1$, as player j is in state ``not fixed'' at time $\theta_i$ (transitions $(1)$ and $(4)$ are impossible).
%
As stated above, $\E_{\mu}[N_x^j(T)] = \bigO{\log T}$ for both $x=3$ and $x=5$,
and so $\E_{\mu}[n(T)] = \bigO{\log T}$ also.

$(\beta)$ To control $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
a simple argument can be used.
$\bigcup_{t=\theta_i}^{\tau_i-1} I_2^j(t)$
implies $C^j(t)$ for $\tau_i - \theta_i$ consecutive times.
%
The very structure of \RandTopM{} gives that in this sequence of transitions $(2)$,
the successive collisions (\ie, $C^j(t-1) \cap C^j(t)$)
implies that each new arm $A^j(t+1)$ for $t \in \{\theta_i, \tau_i-1\}$ is selected uniformly from
$\TopM(t+1)$,
a set of size $M$ with at least one available arm.
%
Indeed, as there is $M-1$ other players, at time $t+1$ \emph{at least} one arm in $\TopM(t+1)$ is not selected by any player $k'\neq k$,
and so player $j$ has \emph{at least} a probability $1/M$ to select
a free arm, which implies $\overline{C^j(t+1)}$, and so implies the end of the sequence.
%
In other words, the average length of sequences of transitions $(2)$,
$\E_{\mu}\left[ \tau_i - \theta_i \right]$,
is bounded by the expected number of failed trial of a repeated Bernoulli experiment, with probability of success larger than $1/M$ (by the uniform choice of $A^j(t+1)$ in a set of size $M$ with at least one available arm).
We recognize the mean of a geometric random variable, of parameter $\lambda \geq 1/M$, and so $\E_{\mu}\left[ \tau_i - \theta_i \right] = \frac{1}{\lambda} \leq \frac{1}{1/M} = M$.

This finishes the proof as $\E_{\mu}[N_2^j(T)] = \sum_{t=1}^T \Pr(I_2^j(t)) = \bigO{\log T}$ and so
$\sum_{t=1}^T \Pr(\widetilde{C^j}(t) \cap (A^j(t) = k)) = \bigO{\log T}$
and finally
$\E_{\mu}[\cC(T)] = \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \bigO{\log T}$ also.

We can be more precise about the constants, all the previous arguments can be used successively:
\begin{align}
  \E_{\mu}[\cC(T)]
  &\leq M \sum_{j=1}^M \left(\sum_{t=1}^T \Pr(I_2^j(t)) + \sum_{t=1}^T \Pr(I_3^j(t))\right)
  = M \left(\sum_{j=1}^M \E_{\mu}[N_2^j(T)] + \E_{\mu}[N_3^j(T)]\right) \\
  &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag  \\
  &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
  &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
  &\leq \left(2 M^3 + M^2\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
\end{align}

And so we obtain the desired inequality, with explicit constants, that depend only on $\boldsymbol{\mu}$ and $M$.
%
\begin{equation}
  \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \E_{\mu}[\cC(T)]
  \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
\end{equation}

\paragraph{Number of switches}\label{app:5:NumberSwitches}
%
Note that we controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
Thus, the total number of arm switches is also proved to be logarithmic, if all players uses
the \MCTopM-\klUCB{} algorithm.

\paragraph{Strong uniform efficiency}\label{app:5:JustifyingDefinition5}
%
As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proved
to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
%
And as justified after Definition~\ref{def:5:DecentralizedUniformEfficiency} (page~\pageref{def:5:DecentralizedUniformEfficiency}), uniform efficiency and invariance under permutations of the users implies strong uniform efficiency, and so \MCTopM{} satisfies Definition~\ref{def:5:DecentralizedUniformEfficiency}.
%
This is a sanity check: the lower-bound of
Theorem~\ref{thm:5:BetterLowerBound} indeed applies to our algorithm \MCTopM,
and finally this highlights that it is order-optimal for the regret, in the sense that it matches the lower-bound up-to a multiplicative constant,
and optimal for the term \ref{eq:5:term1}.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Additional Discussions on \Selfish}
\label{app:5:SelfishFails}

% OBSERVATION
% Empirically it works fine in average of regrets on lots of simulations, but on worst case regret (or in a histogram) we can see it does not have a logarithmic average regret
%
% BAD
% Explain on a small example ($K=2, M=3$ or $K=M=2$) that we found both manually, empirically and with a formal game tree exploration a small probability of failure, for both \Selfish-\UCB{} and \Selfish-\klUCB...
%
% a small probability of failure... $K=2, M=3$ or $K=M=2$. (idée: comparer \Selfish{} et \rhoRand{} en terme d'histogramme de la distribution de $R_T$)
%
% DONE estimer la proba ? ou la calculer avec le calcul formel ? ca c'est fait c'est bon
%
% a curiosity: when $K=M=2$ and $\mu_1 = \mu_2$, using \Selfish{} KL-UCB does orthogonalize !

As said before, analyzing \Selfish{} is harder,
but for instance one can prove that it yields constant
collisions and regret for the trivial case of $\mu_1=\mu_2=1$ and $M=2$.
%
Empirically, when \Selfish{} is compared to the other algorithms,
% with \UCB{} or \klUCB{} indices,
% and the average regret is considered (for let say $1000$ repetitions),
it is hard to find a case when \Selfish{} performs badly, as its (empirical average) regret always appeared logarithmic.
%
But an issue of only visualizing the empirical average regret for a certain number of repetitions is
that if a certain ``bad'' run happens only with small probability, it is possible that it never happened
in a simulation, or that it happened a few times but not enough to make the average regret look non-logarithmic\footnote{For instance, $0.999 \log(T) + 0.001 T$ looks more like $\log(T)$ than $T$ so an event yielding linear regret with ``small'' probability $10^{-3}$ cannot be observed from a plot showing the average regret.}.
%
This is why the distribution of regret, at the end of the simulations, $R_T$, is also displayed (in Appendix~\ref{app:5:moreplots}).
%
In a simple problem, with $M=2$ or $M=3$ players competing for $K=3$ arms,
for instance with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
the histogram in Figure~\ref{fig:5:selfish_fail1} shows that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} is not small (and appears linear).
%
%
Additionally, Figure~\ref{fig:5:selfish_fail2} shows that \Selfish-\klUCB{} also has bad performance against random uniform problems $\boldsymbol{\mu}\in[0,1]^K$,
for $M=2$ or $3$ and $K=3$, in a lot of cases.
In comparison, the others algorithms seem to have a logarithmic regret (for $M=2$ and all algorithms)
or even a constant regret (for $M=3$ and \RandTopM{} and \MCTopM).

The intuition behind these configurations when \Selfish{} performs poorly is the following,
if all players use the same algorithm and the same indices.
If two players $i$ and $j$ have exactly the same vectors $[\widetilde{S_k}^i(t)] = [\widetilde{S_k}^j(t)]$
and $[T_k^i(t)] = [T_k^j(t)]$ at some time step $t \geq 1$,
with different values for each $k$
and so that the index vectors $\mathbf{g}^j(t) = \mathbf{g}^{i}(t)$ have different values for each arm $k$,
then both players will take the same decisions at time $t$, and collide.
Colliding does not change $[\widetilde{S_k}^j(t+1)]$ but increase one value in $[T_k^j(t+1)]$ by $1$.
Then at the next step the same conditions on $\widetilde{S}^j$ and $N^j$ are preserved,
and if the same condition on $\mathbf{g}^j(t+1)$ are also preserved,
the two players will continue to collide.
%
We did not succeed in proving mathematically
that the preservation of the first hypothesis on $\widetilde{S}^j$ and $N^j$
implies the preservation of the hypothesis on index $\mathbf{g}^j$,
but numerically it turns out to be always the case:
and so two players colliding in such a setting will continue to do so infinitely:
we denote such configurations as \emph{absorbing}.

We wrote a script
\footnote{\url{banditslilian.gforge.inria.fr/docs/complete_tree_exploration_for_MP_bandits.html}}
that explores formally all the possible runs,
up-to a certain small time horizon,
by exploring the complete \emph{game tree},
of possible (random) rewards from the $K$ arms
and (random) actions from the $M$ players, up-to a small depth of let say $T=8$.
%
Such game tree becomes quickly very large, but this was enough to confirm that with
a certain small probability, function of $\mu_1,\dots,\mu_K$,
two players can arrive in just a few steps in a ``bad'' absorbing configuration.
%
For instance, for only $K=2$ arms, the following game tree in Figure~\ref{fig:5:oneGameTree_SelfishKLUCB} illustrates the first $3$ steps that can lead to $2$ absorbing configurations.
Using symbolic computations, the probability of reaching any of them was found to be
$\geq \mu_1^2(1-\mu_2)^2/2 + \mu_2^2(1-\mu_1)^2/2$ for \Selfish-\UCB.
That is far from being negligible, as it evaluates to $0.328$ for $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
and a numerical simulation on $1000$ runs found $325$ cases of bad performance.
%
The same game tree exploration can be made for \Selfish-\klUCB,
but so far we were not able to justify why it experiences fewer cases of bad performances even though our software found the same (lower bound on) failure probability.

\begin{figure}[h!]
    \centering
    \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex}
    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing.gv.pdf}
    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing__leafs.gv.pdf}
    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing.gv.pdf}
    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs.gv.pdf}
    \caption[Trajectories of $M=2$ players using \Selfish-\UCB{} for $K=2$ arms.]{For $K=2$ arms and $M=2$ players using \Selfish-\UCB, for depth$=3$: $2$ \textbf{\textcolor{red}{absorbing configurations}}. Each rectangle represents a configuration, as the matrix $[[\widetilde{S_k}^j(t) / T_k^j(t) ]_j]_k$. Absorbing configurations from depth $2$ are case of equality of the two vectors and the \Selfish{} indices $\widetilde{g_k}^j(t)$. Transitions are labeled with their probabilities.}
    \label{fig:5:oneGameTree_SelfishKLUCB}
\end{figure}

From the structure of such game tree, we conjecture that the probability of reaching absorbing configurations (before a certain time $t$) is always lower-bounded
by a polynomial function of $\mu_1,\dots,\mu_K$ and $1-\mu_1,\dots,1-\mu_K$,
of degree at most $t$ in each variable.
% with coefficients smaller than $1$.
% and without simple factors (\ie, $\mu_k^d$).
%
As such, the lower bound on probability of failures should decrease when $K$ and $M$ increase, and this is coherent with the experiments for $K=9$ or $K=17$ (see Figures~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} and \ref{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}),
where \Selfish{} is shown to be uniformly more efficient than \rhoRand.
%
Of course, one cannot run an infinite number of simulations, and the smaller the probability of failure, the less likely it is to observe a failure
in a finite number of runs.

% IDEA
\paragraph{Ideas to fix \Selfish{} ?}
%
It could be possible to change the \Selfish{} algorithm to add a way to escape such absorbing trajectories.
For instance one could imagine that after seen seeing, \eg, $x=10$ collisions in a row,
a certain random action could be taken by the players.
These tricks can work empirically in some cases,
but they are harder to analyze formally,
and it is hard to tune the parameters (here $x$, but possibly more),
and we do not find such tricks to be promising from a theoretical point-of-view.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Additional Figures}
\label{app:5:moreplots}

The plots missing from Section~\ref{sec:5:experiments} are included here,
as well as some additional numerical results.

% -----------------------------------------------------------------
\subsection{Illustration of the lower bound}
\label{app:5:illustrationLowerBound}

We proved in Theorem~\ref{thm:5:BetterLowerBound} that the normalized regret, \ie, $R_T$ divided by
$\log T$, is asymptotically lower bounded by a constant $\mathrm{LB}(\boldsymbol{\mu}, M)$
depending on the problem $\boldsymbol{\mu}$ and the number of players $M$, for any $\rho$.
\begin{equation}
  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \rho)}{\log T} \geq \mathrm{LB}(\boldsymbol{\mu}, M).
\end{equation}
%
For an example problem with $K = 9$ arms, we display below on the $x$ axis is
the number of player, from $1$ player to $9$ players, and on the
$y$ axis is the value of this constant $\mathrm{LB}(\boldsymbol{\mu}, M)$, from the initial
theorem and from our theorem.
We chose a simple problem, with Bernoulli
distributed arms, with $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$.
% (as used in some articles).
%
Figure~\ref{fig:5:CompLowerBounds} clearly shows that our improved lower bound is indeed larger than the initial one by \cite{Zhao10},
and both become uninformative when $M=K$ (\ie, null).

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.70\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
  \caption{Comparison of our lower bound against the one  from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
  \label{fig:5:CompLowerBounds}
\end{figure}

%
% System regret and three terms
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_2092905764868974160.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_1699475021767911583.pdf}
    % \end{subfigure}
    \caption{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
    \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

Figures~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
show the regret $R_T(\boldsymbol{\mu}, M, \rho)$
on the same example problem $\boldsymbol{\mu}$, with $K = 9$ arms and respectively $M = 6$, or $9$ players, for \Selfish-\klUCB.
%
It is just a simple way to check that the two lower bounds on the regret indeed appear as valid lower bounds empirically,
and are moreover lower bounds on the count of selections (\ref{eq:5:term1}, displayed in \textcolor{blue}{cyan}).
%
The lower bounds (in black) are $C(\boldsymbol{\mu}, M) \log t$, the dashed line
for \citeauthor{Zhao10}'s lower bound, and the continuous line is our lower bound.
%
These plot show the regret (in red),
% the three lower bounds
% (centralized, \citeauthor{Anandkumar11}'s and our lower bounds, in black),
and the three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} in the decomposition of the regret.
As explained in Lemma~\ref{lem:5:DecompositionRegret}, term \ref{eq:5:term2} is not always non-negative.
For $M=9$ and \Selfish, \ref{eq:5:term3} is actually larger than the regret,
and term \ref{eq:5:term1} is zero, as well as the lower bounds.

\subsection{Figures from Section~\ref{sec:5:experiments}}
\label{app:5:plotsFromSec5}

This last Appendix includes the figures used in Section~\ref{sec:5:experiments},
with additional comments.

\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{0.85\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized____env1-1_7803645526012310577.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{0.85\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized_semilogy____env1-1_7803645526012310577.pdf}
    % \end{subfigure}
    \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$]{Regret (in log-$y$ scale on the right) for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$. With a perfect knowledge on the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly in this simple setting, even with the knowledge of the horizon $T$ for \MusicalChair{}. The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can bounded with probability $0.5$ and $0.9$ respectively. The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.}
    \label{fig:5:MP__K9_M3_T123456_N100__8_algos}
\end{figure}

%
% Regular plots of centralized regrets
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{0.49\textwidth}
        \includegraphics[width=0.90\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_5016720151160452442.pdf}
    % \end{subfigure}
    % % ~
    % \begin{subfigure}[!h]{0.49\textwidth}
    %   \includegraphics[width=1.10\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
    % \end{subfigure}
    \caption[Failure case of \Selfish]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$. Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$). The regret for the three other algorithms is very small for this problem, always smaller than $100$ here.}
    \label{fig:5:selfish_fail1}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

%
% Regular plots of centralized regrets
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{0.49\textwidth}
        \includegraphics[width=1.10\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_2643560344649862285.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{0.49\textwidth}
        \includegraphics[width=1.10\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_4683123079851881812.pdf}
    % \end{subfigure}
    \caption[Second failure case of \Selfish]{Regret, $M=2$ and $M=3$ players, $K=3$ arms, horizon $T=5000$, against $1000$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$. \Selfish{} (top curve in \textcolor{darkgreen}{green}) clearly fails in such setting with small $K$.}
    \label{fig:5:selfish_fail2}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

%
% Regular plots of centralized regrets
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        % \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
        \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_1035303196230283176.pdf}
        % \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_5016720151160452442.pdf}
    % \end{subfigure}
    ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
    % \end{subfigure}
    \caption[Third failure case of \Selfish]{Regret for $M=3$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$. Axis $x$ is for regret (different scale for each), and the top \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($11$ cases of $R_T \geq T$, out of $1000$). The regret for the three other algorithms is very small for this problem, and even appears constant.}
    \label{fig:5:selfish_fail3}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

%
% Regular plots of centralized regrets
%

\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        % \includegraphics[width=1.00\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
        \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3251433209347345969.pdf}
    % \end{subfigure}
    \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$. \rhoRand{} (top \textcolor{blue}{blue}) is outperformed by the other algorithms (and the gain increases when $M$ increases), which all perform similarly in such configurations. Note that the (small) tail of the histograms come from complicated problems $\boldsymbol{\mu}$ and not failure cases.}
    \label{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_RegretCentralized____env1-1_3892966382091165662.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3892966382091165662.pdf}
    % \end{subfigure}
    \caption[Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$]{Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$. This extreme case $M=K$ shows the drastic difference of behavior between \RandTopM{} and \MCTopM, having constant regret, and \rhoRand{} and \Selfish, having large regret.}
    \label{fig:5:MP__K9_M9_T5000_N500__4_algos__all_HistogramsRegret}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K9_M2_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_2643359116089264295.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
    %   \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K9_M9_T10000_N200__4_algos/all_RegretCentralized_loglog____env1-1_2306423191427933958.pdf}
    % \end{subfigure}
    \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$,Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$. In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand. \MCTopM{} is always among the best algorithms, and for $M$ not too small, its regret seems logarithmic with a constant matching the lower bound.}
    \label{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}
\end{figure}

\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{0.75\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{0.75\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{0.75\textwidth}
        \includegraphics[width=1.00\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
    % \end{subfigure}
    \caption[Regret for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$]{Regret (in log-log scale), for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$. The same observation as in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} can be made. \Selfish{} outperforms \MCTopM{} for $M=2$ here. Additionally, \MCTopM{} is the only algorithm to not fail dramatically when $M=K$ here.}
    \label{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}
\end{figure}

### 2019-02-14 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Typo in secondary captions for the table of Figures.

@@
   % \end{subfigure}
-  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
 \rhoRand{} (top \textcolor{blue}{blue} curve) is outperformed by the other algorithms (and the gain increases with $M$).
@@
   % \end{subfigure}
-  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
 \RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand.

### 2019-02-14 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Typo in secondary captions for the table of Figures.

@@
   \includegraphics[width=0.70\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
-  \caption{Comparison of our lower bound against the one  from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
+  \caption[Comparison of our lower bound against the one from \cite{Zhao10}]{Comparison of our lower bound against the one from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
   \label{fig:5:CompLowerBounds}
@@
     % \end{subfigure}
-    \caption{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
+    \caption[Multi-player regret with its three terms]{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
     \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
@@
     % \end{subfigure}
-    \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+    \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
 \rhoRand{} (top \textcolor{blue}{blue}) is outperformed by the other algorithms (and the gain increases when $M$ increases), which all perform similarly in such configurations.
@@
     % \end{subfigure}
-    \caption[Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$]{Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+    \caption[Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
 This extreme case $M=K$ shows the drastic difference of behavior between \RandTopM{} and \MCTopM, having constant regret, and \rhoRand{} and \Selfish, having large regret.}
@@
     % \end{subfigure}
-    \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$,Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+    \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$,Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
 In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand.

### 2019-02-14 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Typo

@@
     % \end{subfigure}
-    \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$]{Regret (in log-$y$ scale on the right) for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
+    \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem]{Regret (in log-$y$ scale on the right) for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
 With a perfect knowledge on the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly in this simple setting, even with the knowledge of the horizon $T$ for \MusicalChair{}.

### 2019-02-15 2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py
Better code snippet.

@@ -11,30 +11,30 @@ from SMPyBandits.Environment import MAB
 # Import single-player algorithms
 from SMPyBandits.Policies import *
 
-#: HORIZON : number of time steps of the experiments.
-#: Warning Should be >= 10000 to be interesting "asymptotically".
+# HORIZON : number of time steps of the experiments.
+# Warning Should be >= 10000 to be interesting "asymptotically".
 HORIZON = int(getenv('T', 10000))
 
-#: REPETITIONS : number of repetitions of the experiments.
-#: Warning: Should be >= 10 to be statistically trustworthy.
+# REPETITIONS : number of repetitions of the experiments.
+# Warning: Should be >= 10 to be statistically trustworthy.
 REPETITIONS = int(getenv('N', 100))
 
-#: Number of jobs to use for the parallel computations.
-#: -1 means all the CPU cores, 1 means no parallelization.
+# Number of jobs to use for the parallel computations.
+# -1 means all the CPU cores, 1 means no parallelization.
 N_JOBS = int(getenv('N_JOBS', -1))
 
-#: Number of arms for non-hard-coded problems (Bayesian problems)
+# Number of arms for non-hard-coded problems (Bayesian problems)
 NB_ARMS = int(getenv('K', 3))
 
-#: Lower value of means
+# Lower value of means
 LOWER = float(getenv('LOWER', 0))
-#: Amplitude value of means
+# Amplitude value of means
 AMPLITUDE = float(getenv('AMPLITUDE', 1))
 
-#: Type of arms for non-hard-coded problems (Bayesian problems)
+# Type of arms for non-hard-coded problems (Bayesian problems)
 ARM_TYPE = mapping_ARM_TYPE[str(getenv('ARM_TYPE', "Bernoulli"))]
 
-#: Means of arms for non-hard-coded problems
+# Means of arms for non-hard-coded problems
 MEANS = uniformMeans(nbArms=NB_ARMS, delta=0.05,
     lower=LOWER, amplitude=AMPLITUDE, isSorted=True
 )
@@ -43,10 +43,11 @@ import numpy as np
 # more parametric? Read from cli?
 MEANS_STR = getenv('MEANS', '')
 if MEANS_STR != '':
-    MEANS = [ float(m) for m in MEANS_STR.replace('[', '').replace(']', '').split(',') ]
+    MEANS_STR = MEANS_STR.replace('[','').replace(']','').split(',')
+    MEANS = [ float(m) for m in MEANS_STR ]
[…]

### 2019-02-15 2-Chapters/3-Chapter/src/example_of_main_singleplayer.py

@@ -8,14 +8,15 @@ from SMPyBandits.Environment import Evaluator
 configuration['showplot'] = True
 evaluation = Evaluator(configuration)
 
-# Start the evaluation and then print final ranking and plot, for each environment
+# Start the evaluation and then print final ranking
+# and do and show plot, for each environment
 for envId, env in enumerate(evaluation.envs):
     # Evaluate just that env
     evaluation.startOneEnv(envId, env)
 
 # Compare them
 for envId, env in enumerate(evaluation.envs):
-    evaluation.plotHistoryOfMeans(envId)  # XXX To plot without saving
+    evaluation.plotHistoryOfMeans(envId)
 
     print("\nGiving all the vector of final regrets ...")
     evaluation.printLastRegrets(envId)
@@ -39,7 +40,7 @@ for envId, env in enumerate(evaluation.envs):
     print("\n\n- Plotting the regret...")
     evaluation.plotRegrets(envId)
 
-    print("\n- Plotting the probability of picking the best arm of time...")
+    print("\n- Plotting the probability of picking the best arm...")
     evaluation.plotBestArmPulls(envId)
 
     print("\n- Plotting the histograms of regrets...")

### 2019-02-15 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Pas mal relu le Chapitre 5, écrit un abstract.

@@
 
-\TODOL{This chapter is basically a raw include from my paper ``Multi-Player Bandits Revisited'', see https://hal.inria.fr/hal-01629733}
-
 Multi-player Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
-Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms. Most existing work assume
-that \emph{sensing information} is available to the algorithm. Under this assumption, we improve the state-of-the-art lower bound for the regret of any decentralized algorithms
-and introduce two algorithms, \RandTopM{} and \MCTopM{}, that are shown to empirically outperform existing algorithms.
-Moreover, we provide strong theoretical guarantees for these algorithms, including
+Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms.
+Existing works assume that \emph{sensing information} is available to the algorithm.
+Under this assumption, we improve the state-of-the-art lower bound for the regret of any decentralized algorithms of a certain class.
+We introduce two algorithms, \RandTopM{} and \MCTopM{}, that are shown to empirically outperform existing algorithms. Moreover, we provide strong theoretical guarantees for these algorithms, including
 a notion of asymptotic optimality in terms of the number of selections of bad arms.
@@
 % -----------------------------------------------------------------
+\section{Motivation for multi-player MAB models}
+\label{sec:5:introduction}
 % -----------------------------------------------------------------
-\section{Introduction}
-\label{sec:5:introduction}
-% the paragraph below may be useful to state in the intro or early in the paper
 
-% general intro, UCB
+% % general intro, UCB
 
-Several sequential decision making problems under the constraint of partial information
-have been studied since the 1950s under the name of Multi-Armed Bandit (MAB) problems \citep{Robbins52,LaiRobbins85}.
-In a stochastic MAB model, an agent is facing $K$ unknown probability distributions, called arms in reference to the arms of a one-armed bandit (or slot machine) in a casino.
-Each time she selects (or draws) an arm, she receives a reward drawn from the associated distribution.
-Her goal is to build a sequential selection strategy that maximizes the total reward received.
-A class of algorithms to solve this problem is based on Upper Confidence Bounds (UCB), first proposed by \cite{LaiRobbins85,Agrawal95} and further popularized by \cite{Auer02}.
-The field has been very active since then, with several algorithms proposed and analyzed, both theoretically and empirically, even beyond the stochastic assumption on arms, as explained in the survey by \cite{Bubeck12}.
+% Several sequential decision making problems under the constraint of partial information
+% have been studied since the 1950s under the name of Multi-Armed Bandit (MAB) problems \citep{Robbins52,LaiRobbins85}.
+% In a stochastic MAB model, an agent is facing $K$ unknown probability distributions, called arms in reference to the arms of a one-armed bandit (or slot machine) in a casino.
+% Each time she selects (or draws) an arm, she receives a reward drawn from the associated distribution.
+% Her goal is to build a sequential selection strategy that maximizes the total reward received.
+% A class of algorithms to solve this problem is based on Upper Confidence Bounds (UCB), first proposed by \cite{LaiRobbins85,Agrawal95} and further popularized by \cite{Auer02}.
+% The field has been very active since then, with several algorithms proposed and analyzed, both theoretically and empirically, even beyond the stochastic assumption on arms, as explained in the survey by \cite{Bubeck12}.
 
-% appli, from clinical trials to cognitive radios
+% % appli, from clinical trials to cognitive radios
 
-The initial motivation to study MAB problems arose from clinical trials (the first MAB model can be traced back to $1933$, by \citeauthor{Thompson33}), in which a doctor sequentially allocates treatments (arms) to patients and observes their efficacy (reward).
-More recently, applications of MAB have shifted towards sequential content recommendation, \eg{} sequential display of advertising to customers or A/B testing \citep{Li10,Chapelleetal14Ad}.
-%
-In the mean time, MAB were found to be relevant to the field of Cognitive Radio (CR, \cite{Mitola99}),
-and \cite{Jouini09,Jouini10} first proposed to use \UCB{} for the Opportunistic Spectrum Access (OSA) problem,
-and successfully conducted experiments on real radio networks demonstrating its usefulness.
-For CR applications, each arm models the quality or availability of a radio channel (a frequency band) in which there is some background traffic (\eg, primary users paying to have a guaranteed access to the channel in the case of OSA).
-A smart radio device needs to insert itself in the background traffic, by sequentially choosing a channel to access and try to communicate on, seeking to optimize the quality of its global transmissions.
+% The initial motivation to study MAB problems arose from clinical trials (the first MAB model can be traced back to $1933$, by \citeauthor{Thompson33}), in which a doctor sequentially allocates treatments (arms) to patients and observes their efficacy (reward).
+% More recently, applications of MAB have shifted towards sequential content recommendation, \eg{} sequential display of advertising to customers or A/B testing \citep{Li10,Chapelleetal14Ad}.
+% %
+% In the mean time, MAB were found to be relevant to the field of Cognitive Radio (CR, \cite{Mitola99}),
+% and \cite{Jouini09,Jouini10} first proposed to use \UCB{} for the Opportunistic Spectrum Access (OSA) problem,
+% and successfully conducted experiments on real radio networks demonstrating its usefulness.
+% For CR applications, each arm models the quality or availability of a radio channel (a frequency band) in which there is some background traffic (\eg, primary users paying to have a guaranteed access to the channel in the case of OSA).
+% A smart radio device needs to insert itself in the background traffic, by sequentially choosing a channel to access and try to communicate on, seeking to optimize the quality of its global transmissions.
 
@@
 % -----------------------------------------------------------------
-% -----------------------------------------------------------------
 \section{Multi-Player Bandit Model with Different Feedback Levels}
 \label{sec:5:model}
+% -----------------------------------------------------------------
 
@@
     However, the sensing and collision information are censored.
-    Recently, \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications.
+    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications.
 %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
@@
 % -----------------------------------------------------------------
+\section{An Asymptotic Regret Lower Bound for a certain class of algorithms}
+\label{sec:5:lowerbound}
 % -----------------------------------------------------------------
-\section{An Asymptotic Regret Lower Bound}
-\label{sec:5:lowerbound}
+
+\TODOL{Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
 
@@
 
+\TODOL{Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
+
 % -----------------------------------------------------------------
@@
 % -----------------------------------------------------------------
-% -----------------------------------------------------------------
 \section{New Algorithms for Multi-Player Bandits}
 \label{sec:5:algorithms}
+% -----------------------------------------------------------------
 
@@
 % -----------------------------------------------------------------
-% -----------------------------------------------------------------
 \section{Empirical performances}
 \label{sec:5:experiments}
+% -----------------------------------------------------------------
 
@@
 We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \klUCB{} indices.
-%used with two index policies, \UCB{} and \klUCB.
 Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.
-% but observations and conclusions from the experiments are discussed here.
+
+\TODOL{Rewrite appendix to include all simulations here!}
+
+\TODOL{Include here some plots showing \RhoRand{} with \UCB{} and \klUCB{} vs \MCTopM{} with \UCB{} and \klUCB{}.}
 
@@
 %, as the theoretical guarantees for single player suggested,
-As the prupose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
+As the purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
 % And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
@@
   % \begin{subfigure}[!h]{0.49\textwidth}
-  %   \includegraphics[width=1.10\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
+  %   \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
   % \end{subfigure}
@@
   % \begin{subfigure}[!h]{0.49\textwidth}
-    \includegraphics[width=0.8\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
+    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
   % \end{subfigure}
@@
 % -----------------------------------------------------------------
-% -----------------------------------------------------------------
 \section{Theoretical elements}
 \label{sec:5:upperbounds}
+% -----------------------------------------------------------------
 
@@
 % -----------------------------------------------------------------
-% -----------------------------------------------------------------
 \section{Conclusion and future work}
 \label{sec:5:conclusion}
+% -----------------------------------------------------------------
 

### 2019-02-15 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Pas mal relu le Chapitre 5, écrit un abstract.

@@
 \begin{eqnarray*}
-T - \E_{\mu}[T_k(T)] & \leq & T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],\end{eqnarray*}
+  T - \E_{\mu}[T_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
+  & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
+\end{eqnarray*}
 where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
@@
 \label{proof:5:BetterLowerBound}
+
+\TODOL{Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
 
@@
 
+\TODOL{This is wrong as explained in \cite{BoursierPerchet18}! Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
+
 Putting things together, we showed that
@@
 \label{app:5:RandTopM}
+
+\TODOL{I should include this in the main text, it makes more sense to present RandTopM and then MCTopM !}
 
@@
 \begin{figure}[h!]
-  \begin{tikzpicture}[>=latex',line join=bevel,scale=3.8]
+  \resizebox{0.98\textwidth}{!}{
+  \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
       %
@@
   \end{tikzpicture}
+  }
   \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
@@
 
+\TODOL{Maybe we can completely remove this, and just points to the paper \cite{Besson2018ALT}?}
+
 % OBSERVATION
@@
 We wrote a script
-\footnote{\url{banditslilian.gforge.inria.fr/docs/complete_tree_exploration_for_MP_bandits.html}}
+\footnote{~See online \texttt{\url{smpybandits.github.io/docs/complete_tree_exploration_for_MP_bandits.html}}}
 that explores formally all the possible runs,
@@
 
+\TODOL{Maybe we can not include this part and just points to the paper \cite{Besson2018ALT}?}
+
 We proved in Theorem~\ref{thm:5:BetterLowerBound} that the normalized regret, \ie, $R_T$ divided by

### 2019-02-15 2-Chapters/5-Chapter/chapter5.tex
Pas mal relu le Chapitre 5, écrit un abstract.

@@
 
-% First chapter begins here
-\chapter{Multi-Player Multi-Armed Bandits Models}
+\chapter[Multi-Player Multi-Armed Bandits Models]{Multi-Player Multi-Armed Bandits Models and Algorithms}
 \label{chapter:5}
@@
 
-Abstract for Chapter 5.
+In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
+We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
+A perfect time and frequency synchronization is assumed,
+and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
+As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
+%
+Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
+Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices, Even if there is many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step even the number of communicating devices is random and varies a lot.
+
+Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider the easier case of at most $M \leq K$ devices, communicating at each time step.
+Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward at each time step.
+%
+We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
+Each time frame is separated as before in a \emph{sensing} phase (during which the device senses for the background traffic),
+an \emph{uplink} phase (during which the device sends a packet to the gateway if it sensed the chosen channel to be free),
+and a \emph{downlink} phase (during which it waits for an \Ack{} from the gateway).
+In this first model, the binary reward is $1$ only if the channel was sensed to be free of background traffic \textbf{and} if \Ack{} was received, and the device has access to both information.
+For this first model, we present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
+Like in previous works, we consider the centralized system regret (multi-player regret), or simply referred to as regret.
+We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions..
+We then analyze our \MCTopM{} algorithm and we show that its regret upper-bound asymptotically matches the improved lower-bound, with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
+We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M$ accessing $K \geq M$ channels.
+
+We are also interested to consider the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
+Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
+Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we find on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear regret and fails to solve the no sensing multi-player MAB problem).
+We did not propose any other efficient algorithm, but our work presented in April~$2018$ inspired some other articles \cite{LugosiMehrabian18,BoursierPerchet18}.
+Two articles confirmed our finding that \Selfish{} has a linear regret, and they all proposed new algorithms.
+We do not include numerical simulations to compare all of them, but we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
+
+Finally, another interesting extension is to consider different channels utility for each devices.
+It is a very interesting extension of our model, for instance the devices can be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
+This extension is not studied per say but we quickly review the existing literature \cite{Bistritz18,KaufmannAbbas19} and discuss the possible real-world usage of the existing algorithms.
+
+\TODOL{Get bibtex of the article by Sumit J. Darak}
 
@@
 % ----------------------------------------------------------------------------
-\section{Blabla}
-\label{sec:blabla}
+\input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex}
+
+% ----------------------------------------------------------------------------
+\section{Literature review of algorithms solving the ``no sensing'' case}
+\label{sec:5:noSensingCase}
 
@@
 % ----------------------------------------------------------------------------
-\input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex}
-
-% ----------------------------------------------------------------------------
-\section{Towards efficient algorithms solving the "no sensing" case}
-\label{sec:5:XXX}
-
-% ----------------------------------------------------------------------------
-\section{Conclusion of Chapter 5}
+\section{Conclusion of Chapter~\ref{chapter:5}}
 \label{sec:5:conclusion}
@@
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter 5}
+\section{Appendix for Chapter~\ref{chapter:5}}
 \label{sec:5:appendix}
+
+We include here some missing proofs and additional simulation results for this Chapter.
 

### 2019-02-15 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Ok code for Chapter:6

@@
     % \begin{subfigure}[!h]{0.49\textwidth}
-        \includegraphics[width=1.10\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_2643560344649862285.pdf}
+        \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_2643560344649862285.pdf}
     % \end{subfigure}
@@
     % \begin{subfigure}[!h]{0.49\textwidth}
-        \includegraphics[width=1.10\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_4683123079851881812.pdf}
+        \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_4683123079851881812.pdf}
     % \end{subfigure}

### 2019-02-15 2-Chapters/6-Chapter/chapter6.tex
Ok for tex root
It's a really long and booooring work
Okay for chapters and parts
Ok code for Chapter:6

@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 
-% First chapter begins here
-\chapter{Towards Non-Stationary Models}
+\chapter{Non-Stationary MAB Models and Possible Applications for IoT Networks}
 \label{chapter:6}
 \minitoc
+
+Abstract for Chapter 6.
+
 \newpage
@@
 
-This chapter is basically a raw include from my paper ``The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits'', see https://hal.inria.fr/hal-02006471
+% This chapter is basically a raw include from my paper ``The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits'', see https://hal.inria.fr/hal-02006471
+
+\TODOL{Rewrite and reorder most of this Chapter, with more subsections, less paragraphs etc, and less copy-paste of our paper \cite{Besson2019GLRT}.}
 
@@
 
-% ----------------------------------------------------------------------------
-\section{Conclusion of Chapter 6}
-\label{sec:6:conclusion}
+% % ----------------------------------------------------------------------------
+% \section{Conclusion of Chapter 6}
+% \label{sec:6:conclusion}
 
-In this chapter, we saw...
+% In this chapter, we saw...
 
-Future works include...
+% Future works include...
 

### 2019-02-15 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Ok code for Chapter:6

% - ``The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits'', see https://hal.inria.fr/hal-02006471

\TODOL{This chapter is basically a raw include from my paper ``The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits'', see https://hal.inria.fr/hal-02006471}

We propose a new algorithm for the piece-wise \iid{} non-stationary bandit problem with bounded rewards.
Our proposal, \GLRklUCB, combines an efficient bandit algorithm, \klUCB, with an efficient, \emph{parameter-free}, change-point detector, the Bernoulli Generalized Likelihood Ratio Test, for which we provide new theoretical guarantees of independent interest. We analyze two variants of our strategy, based on \emph{local restarts} and \emph{global restarts}, and show that their regret is upper-bounded by  $\cO(\Upsilon_T \sqrt{T \log(T)})$ if the number of change-points $\Upsilon_T$ is unknown, and by $\cO(\sqrt{\Upsilon_T T \log(T)})$ if $\Upsilon_T$ is known.
This improves the state-of-the-art bounds, as our algorithm needs no tuning based on knowledge of the problem complexity other than $\Upsilon_T$.
We present numerical experiments showing that \GLRklUCB{} outperforms passively and actively adaptive algorithms from the literature, and highlight the benefit of using local restarts.

% -----------------------------------------------------------------
\section{Motivation for Non-Stationary MAB Models}
\label{sec:6:Introduction}

Multi-Armed Bandit (MAB) problems form a well-studied class of sequential decision making problems, in which an agent repeatedly chooses an action $I_t \in\{1,\dots,K\}$ or ``arm'' --in reference to the arm of a one-armed bandit-- among a set of $K$ arms. In the most standard version of the stochastic bandit model, each arm $i$ is associated with an \iid{} sequence of rewards $(r_{i,t})$ that follow some distribution of mean $\mu_i$. Upon selecting arm $I_t$, the agent receives the reward $r_{I_t,t}$ associated to the chosen arm, and her goal is to adopt a sequential sampling strategy that maximize the expected sum of these rewards. This is equivalent to minimizing the \emph{regret}, defined as the difference between the total reward of the oracle strategy always selecting the arm with largest mean, $\mu^*$, and that of our strategy: $R_T = \bE[\sum_{t=1}^T(\mu^* - \mu_{I_t})]$.

Regret minimization in stochastic bandits has been extensively studied since the works of \cite{Robbins52} and \cite{LaiRobbins85}, and several algorithms with a $\cO(\log(T))$ \emph{problem-dependent} regret upper bound have been proposed (see, \eg, \cite{LattimoreBanditAlgorithmsBook} for a survey). Among those, the \klUCB{} algorithm (\cite{KLUCBJournal}) has been shown to be asymptotically optimal for Bernoulli distributions (in that it exactly matches the lower bound given by \cite{LaiRobbins85}) and can also be employed when the rewards are assumed to be bounded in $[0,1]$.
\emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{Garivier18}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.

Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy (\cite{Thompson33}). More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, or parameters of a dynamically configurable radio hardware (\cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}). Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).

For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not evolve over time} may be a big limitation. Indeed, in cognitive radio new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate; whereas in online recommendation, the popularity of items is also subject to trends. Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.

A first possibility to cope with non-stationary is to model the decision making problem as an \emph{adversarial bandit problem} (\cite{Auer02NonStochastic}). Under this model, rewards are completely arbitrary and are not assumed to follow any probability distribution.
For adversarial environments, the pseudo-regret, which compares the accumulated reward of a given strategy with that of the best fixed-arm policy, is often studied. The pseudo regret of the EXP3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}. However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round. For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
% It is sometimes referred to as the abruptly changing MAB, or switching environments (\cite{MellorShapiro13}).
%
In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $i$ at round $t$ has some mean $\mu_i(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $i_t^* = \arg\max_{i} \mu_i(t)$.

In this paper, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB. Like previous approaches -- \CUSUM{} (\cite{LiuLeeShroff17}) and \MUCB{} (\cite{CaoZhenKvetonXie18}) -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector. For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 (\cite{Auer02}) used in previous works. For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest. This choice is particularly appealing because unlike previous approaches, the Bernoulli GLRT is \emph{parameter-free}: it does not need the tuning of a window size ($w$ in \MUCB), or the knowledge of a lower bound on the magnitude of the smallest change ($\varepsilon$ in \CUSUM).

In this work we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.

The article is structured as follows. We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}. In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test (GLRT) as a Change-Point Detector (CPD) algorithm.
We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bound on the regret of each variant. The unified regret analysis for these two algorithms is sketched in Section~\ref{sec:6:RegretAnalysis}. Numerical experiments
are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.

% -----------------------------------------------------------------
\section{The Piece-Wise Stationary Bandit Setup and Related Works}
\label{sec:6:BanditSetting}

A \emph{piece-wise stationary bandit model} is characterized by a set of $K$ arms.  A (random) stream of rewards $(X_{i,t})_{t\in\N^*}$ is associated to each arm $i \in \{1,\dots,K\}$. We assume that the rewards are bounded, and without loss of generality we assume that $X_{i,t} \in [0,1]$. We denote by $\mu_{i}(t) :=  \bE[X_{i,t}]$ the mean reward of arm $i$ at round $t$. At each round $t$, a decision maker has to select an arm $I_t\in\{1,\dots,K\}$, based on past observation and receives the corresponding reward $r(t) = X_{I_t,t}$. At time $t$, we denote by $i_t^*$ an arm with maximal expected reward, \ie, $\mu_{i_t^*}(t) = \max_i \mu_i(t)$, called an optimal arm (possibly not unique).

A policy $\pi$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
The performance of $\pi$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\pi$:
\begin{align}\label{def:6:nti}
    R_T^{\pi} &= \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
\end{align}

In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
We define the $k$-th breakpoint by $\tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}$. Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, the rewards $(X_{i,t})$ associated to each arms are \iid.
Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed. Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only a few arms change at each breakpoint. Introducing the number of change-points  on arm $i$, defined as $\NCi:=\sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms. Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.

The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while the adversarial model can be considered as a special (worst) case, with $\Upsilon_T = T$. However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.

% -----------------------------------------------------------
\paragraph{Related work.}

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To the authors knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitors the rewards of each arm by using a change detection algorithm (\cite{Basseville93}), and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies are tuned with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.

% \TODOL{Review of previous work (see section on Experiments), and state-of-the-art of passively adaptive and actively adaptive algorithms}

\paragraph{Passively adaptive algorithms.}
%
Their common idea is to adapt a classical policy into forgetting old rewards.
If the forgetting behavior is done efficiently, then the policy can efficiently focus mostly on the most recent rewards, and passively adapt to changes when they happen.
%
The Discounted UCB (D-UCB) algorithm was first introduced in \cite{Kocsis06}, and it is an adaptation of the \UCB{} algorithm, with a discount factor $\gamma\in(0,1)$. It works by decreasing all the past rewards by a multiplicative factor, when receiving a new reward from an arm, for the recent rewards to weight more in the discounted empirical mean used for the computation of the \UCB{} indexes.
The regret of D-UCB was proved to be upper-bounded by $\cO(\sqrt{\Upsilon_T T }\log(T))$ in \cite{Garivier11UCBDiscount}, with a tuning $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, dependent on $\Upsilon_T$.
%
The Sliding-Window UCB (SW-UCB) uses a sliding window of a fixed size $\tau$, to store only the most $\tau$ recent rewards for each arm,
and it was proposed by \cite{Garivier11UCBDiscount}. They proved that tuning its window-size to $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$, gives a bound on its regret of the form $\cO(\sqrt{\Upsilon_T T \log(T)})$.

Both D-UCB and SW-UCB builds on a stationary policy,
but for example EXP3.S from \cite{Auer02NonStochastic}
builds on the EXP3 policy, which is designed for the adversarial case,
and achieves a larger regret upper-bound.
We also found in our experiments that it always performed worse that the reference stationary algorithms (\eg, \klUCB), so we do not include it in Section~\ref{sec:6:NumericalExperiments}.
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proved to be less efficient empirically, or are designed for more specific settings.
%
For instance, Windowed-Mean Shift from \cite{YuMannor09} is generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting as they consider side observations.
%
A different setting where the quantity of interest is not $\Upsilon_T$ but $V_T$ the total variational budget, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
%

The idea of using a simple discount factor, as for D-UCB, was recently adapted to a Bayesian policy, with the Discounted Thompson sampling (DTS) algorithm presented by \cite{RajKalyani17}.
Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust in the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in their paper.
%
In general, we found that passively adaptive approaches can be efficient when their parameters are well tuned, but our experiments show that actively adaptive algorithms perform significantly better.

\paragraph{Actively adaptive algorithms.}
%
We distinguish two families,
the first line of research uses frequentist change-point detectors (\cite{Basseville93}) combined with stationary policies, usually index policies like \UCB.
When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley Test and the \UCB{} policy, but no theoretical guarantee was given.
The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proved, even when $\Upsilon_T$ is known.

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
On the first hand,
the \CUSUM{} test is rather complicated and parametric, as it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when either random walks cross a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test is similarly complex, and it is also studied in \cite{LiuLeeShroff17},
and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments,
and even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
On the other hand, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the different of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).

Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

When all their parameters are tuned correctly, both approaches are proved to achieve a good regret upper-bound.
As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$, but most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the time between changes (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
%
\CUSUMUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but only for Bernoulli distributions,
and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge (see XXX).
%
\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1).

\TODOL{Finish presenting MUCB and CUSUM, and highlights the differences!}

\paragraph{Expert aggregation.}
%
The second line of research of actively adaptive algorithms uses a Bayesian point of view.
A Bayesian CD algorithm is combined with Thompson Sampling in \cite{MellorShapiro13},
and more recently \cite{Alami17} introduced the Memory Bandit algorithm. It is presented as efficient empirically, but no theoretical guarantee are given, and due to its complexity, we do not include it in our experiments.
The idea behind Memory Bandit is to use an expert aggregation algorithm, like EXP4 from \cite{Auer02}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
At each time step, a new expert is introduced, and experts corresponds to different Thompson sampling algorithms, each using a different history of pulls and rewards. Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
%
Note that the regret bound for these methods are still open-problems.

\paragraph{About \klUCB.}
%
Our proposal \GLRklUCB{} is inspired by both the \MUCB{} and \CUSUMUCB{} algorithms.
Previous works focussed on using \UCB{} (\cite{LiuLeeShroff17,CaoZhenKvetonXie18}),
but we propose to use \klUCB{} instead, as it is known to be more efficient for Bernoulli rewards as well as for a more generic case of bounded or one-dimensional exponential families (\cite{KLUCBJournal}).
Theoretical results for these works were only given for \UCB, but they both suggest that extending the results to \klUCB{} (or other efficient stationary policies) should not be difficult.
For a fair comparison, we therefore chose to compare the different Change-Point detector algorithms combined with \klUCB.

\TODOL{FIXME next paragraph is the short version of related works review I have to merge the long and short together and keep as much details as I can!}

The piece-wise stationary bandit model was first studied by \cite{Kocsis06,YuMannor09,Garivier11UCBDiscount}. It is also known as \emph{switching} (\cite{MellorShapiro13}) or \emph{abruptly changing stationary} (\cite{WeiSrivastava18}) environment. %We do not consider the \emph{slowly-varying} MAB models (\cite{Louedec16,WeiSrivastava18}), even if it is interesting for some applications.
% We review previous works that studied the piece-wise stationary bandit model.
To our knowledge, all the previous approaches combine a standard bandit algorithm, like \UCB{}, Thompson Sampling or EXP3, with a strategy to account for changes in the arms distributions. This strategy often consists in \emph{forgetting old rewards}, to efficiently focus on the most recent ones, more likely to be similar to future rewards. We make the distinction between \emph{passively} and \emph{actively} adaptive strategies.

The first proposed mechanisms to forget the past consist in either discounting rewards (at each round, when getting a new reward on an arm, past rewards are multiplied by $\gamma^n$ if that arm was not seen since $n>0$ times, for a discount factor $\gamma\in(0,1)$), or using a sliding window (only the rewards gathered in the $\tau$ last observations of an arm are taken into account, for a window size $\tau$).
Those strategies are passively adaptive as the discount factor or the window size are fixed, and can be tuned as a function of $T$ and $\Upsilon_T$ to achieve a certain regret bound.
Discounted UCB (D-UCB) was proposed by \cite{Kocsis06} and analyzed by \cite{Garivier11UCBDiscount}, who prove a $\cO(\sqrt{\Upsilon_T T }\log(T))$ regret bound, if $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$.
The same authors proposed the Sliding-Window UCB (SW-UCB) and prove a $\cO(\sqrt{\Upsilon_T T \log(T)})$ regret bound, if $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$.

More recently, \cite{RajKalyani17} proposed the Discounted Thompson Sampling (DTS) algorithm, which performs well in practice with $\gamma=0.75$. However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
%
The RExp3 algorithm (\cite{Besbes14stochastic}) can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm. Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point). A  $\cO(V_T^{1/3} T^{2/3})$ regret bound is proved, which is weaker than existing results in our setting. Hence we do no include this algorithm in our experiments.

% \TODOE{Il faut faire attention pour DTS, tu attaques leur papier! Qu'est-ce que nos expériences montrent exactement qui est en contradiction avec ce qu'ils disent?}
% \TODOE{Je vire EXP3.S: c'est une version d'EXP3 qui a des garanties en terme de ``nombre de breakpoint'' dans un cadre adversarial, pas stochastique. Il n'est pas basé sur des restart et un discount}
% \TODOE{Vérifie que je raconte pas de conneries pour RExp3, Odalric et Subho l'avaient mis dans la catégorie ``passive'' mais je comprends pas trop comment marche l'algo}

%
The first \emph{actively adaptive} strategy is Windowed-Mean Shift (\cite{YuMannor09}), which combines any bandit policy with a change point detector which performs \emph{adaptive restarts} of the bandit algorithm. However, this approach is not applicable to our setting as it takes into account side observations. Another line of research on actively adaptive algorithms uses a Bayesian point of view.
A Bayesian Change-Point Detection (CPD) algorithm is combined with Thompson Sampling by \cite{MellorShapiro13}, and more recently in the Memory Bandit algorithm of \cite{Alami17}. Both algorithms do not have theoretical guarantees and their implementation is very costly, hence we do not include them in our experiments. Our closest competitors rather use frequentist CPD algorithms (see, e.g. \cite{Basseville93}) combined with a bandit algorithm.
% The idea behind Memory Bandit is to use an expert aggregation algorithm, like EXP4 from \cite{Auer02}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
% At each time step, a new expert is introduced, and experts correspond to different Thompson sampling algorithms, each using a different history of pulls and rewards. Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
%
The first algorithm of this flavor, Adapt-EVE algorithm (\cite{Hartland06}) uses a Page-Hinkley test and the \UCB{} policy, but no theoretical guarantee are given. EXP3.R (\cite{Allesiardo15,Allesiardo17}) combines a CPD with EXP3, and the history of all arms are reset as soon as a a sub-optimal arm is detecting to become optimal and it achieve a $\cO(\Upsilon_T \sqrt{T \log(T)})$ regret. This is weaker than the $\cO(\sqrt{\Upsilon_T T \log(T)})$ regret achieved by two recent algorithms, \CUSUMUCB{} (\cite{LiuLeeShroff17}) and Monitored UCB (\MUCB, \cite{CaoZhenKvetonXie18}).

\CUSUMUCB{} is based on a rather complicated two-sided \CUSUM{} test, that uses the first $M$ samples from one arm to compute an initial average, and then detects whether a drift of size larger than $\varepsilon$ occurred from this value by checking whether a random walk based on the remaining observations crosses a threshold $h$. Thus it requires the tuning of three parameters, $M$, $\varepsilon$ and $h$. \CUSUMUCB{} performs \emph{local restarts} using this test, to reset the history of \emph{one arm} for which the test detects a change.
\MUCB{} uses a much simpler test, based on the $w$ most recent observations from an arm: a change is detected if the absolute difference between the empirical means of the first and second halves of those $w$ observations exceeds a threshold $h$. So it requires the tuning of two parameters, $w$ and $h$. \MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
% Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
Compared to \CUSUMUCB{}, note that \MUCB{} is numerically much simpler as it only uses a bounded memory, of order $\bigO{K w}$ for $K$ arms.

% \TODOE{I chose not to talk about $\alpha$ here in the end to gain space: in the end, everyone share that parameter. Maybe add a footnote in section 4 saying that CUSUM does random exploration}

\textbf{Advantages of our approach.}
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this paper, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
This study is supported by both theoretical and empirical results.
%
Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing  much in terms of regret.

% \TODOE{I would keep what is below with a small mention (or footnote) in the experimental section}

% -----------------------------------------------------------------
\section{The Bernoulli GLR Change Point Detector}
\label{sec:6:ChangePointDetector}

Sequential change-point detection has been extensively studied in the statistical community (see, \eg, \cite{Basseville93} for a survey). In this article, we are interested in detecting changes on the mean of a probability distribution with bounded support. Assuming that  we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$. We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau \in\N^*$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a stopping time $\hat\tau$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N, \bE[X_i] = \mu_0\right)$.

Generalized Likelihood Ratio tests date back to the seminal work of \cite{barnard1959control} and were for instance studied for change-point detection by \cite{siegmund1995using}. Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, can be use for our problem. We propose instead to exploit the fact that bounded distributions are also dominated by Bernoulli distributions.
We call a \emph{sub-Bernoulli distribution} any distribution $\nu$ that satisfies
$\ln \bE_{X\sim\nu}\left[e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu=\bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) = \ln(1-\mu + \mu e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$. Lemma 1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.

\TODOL{We need to include and prove our results for false alarm, delay, etc, all results we squeezed and compressed for the COLT version \cite{Besson2019GLRT}.}

\subsection{Presentation of the test}

If the samples $(X_t)$ were all drawn from a Bernoulli distribution, our change-point detection problem would reduce to a parametric sequential test of
$\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$
against the alternative
$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
The Generalized Likelihood Ratio statistic for this test is defined by
\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < t}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_t ; \mu_0)},\]
%
where $\ell(X_1, \ldots, X_t ; \mu_0)$ and $\ell(X_1, \ldots, X_t ; \mu_0,\mu_1,\tau)$ denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$.
Using the form of the likelihood for Bernoulli distribution, this statistic can be written with the binary relative entropy $\kl$,
%
\begin{equation}\label{eq:6:BernoulliDivergence}
    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
\end{equation}
%
Indeed, one can show that $\GLR(t) = \sup_{s \in [1,n]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:t}\right)\right]$,
where for $k \leq k'$, $\hat{\mu}_{k:k'}$ denotes the average of the observations collected between the instants $k$ and $k'$. This motivates the definition of the Bernoulli GLR change point detector.

\begin{definition}\label{def:6:GLRDef}
    The Bernoulli GLR change point detector with threshold function $\beta(t,\delta)$ is
    % \vspace*{-5pt}% WARNING
    \begin{equation}\label{def:6:GLR}
        \hat\tau_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [1,t]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
    \end{equation}
\end{definition}

Asymptotic properties of the GLR for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR} but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence between two elements in that exponential family that have mean $x$ and $y$. For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to \eqref{def:6:GLR} with $\kl(x,y) = 2(x-y)^2$ when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.

In the next section, we provide new non-asymptotic results about the Bernoulli GLR test under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the Bernoulli GLR may stop earlier that the Gaussian GLR based on the quadratic divergence $2(x-y)^2$.

% ----------------------------------------------------------------------------
\subsection{Properties of the Bernoulli GLR}\label{subsec:6:PropGLR}

In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small. To define $\beta$, we need to introduce the function $\cT$,
\begin{equation}\label{def:6:function_T}
    \cT(x) ~:=~ 2 \tilde h\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right)
\end{equation}
where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$.
And for any $x \ge 0$, $\tilde h(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise. The function $\cT$ is easy to compute numerically.
% The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function (\cite{Corless96}) (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
Its use for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on the function $\cT$ are also given:  $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large. The proof of Lemma~\ref{lem:6:FalseAlarm}, that actually holds for any sub-Bernoulli distribution, is given in Appendix~\ref{proof:6:FalseAlarm}.

\begin{lemma}\label{lem:6:FalseAlarm}
    Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$. Then the Bernoulli GLR test satisfies $\bP_{\mu_0}(\hat\tau_\delta < \infty) \leq \delta$ with the threshold function
   % \vspace*{-5pt}% WARNING
    \begin{equation}\label{def:6:beta}
        \beta(n,\delta)= 2\cT\left(\frac{\ln(3n\sqrt{n}/\delta)}{2}\right) + 6\ln(1+\ln(n)).
    \end{equation}
\end{lemma}

Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from $\mu_0$ to $\mu_1$ occurs at time $\tau$. We already observed that from Pinsker's inequality, the Bernoulli GLR stops earlier than a Gaussian GLR. Hence, one can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the Bernoulli GLR. Letting $\Delta = |\mu_0 - \mu_1|$, one can essentially establish that for $\tau$ larger than $(1/\Delta^2)\ln(1/\delta)$ (\ie, enough samples before the change), the delay can be of the same magnitude (\ie, enough samples after the change). In our bandit analysis to follow, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and \ref{lem:6:GoodEvent}).

% ----------------------------------------------------------------------------
\subsection{Practical considerations}\label{sub:6:PracticalConsiderations}

Lemma~\ref{lem:6:FalseAlarm} provides the first control of false alarm for the Bernoulli GLR employed for bounded data. However, the threshold \eqref{def:6:beta} is not fully explicit as the function $\cT(x)$ can only be computed numerically.
Note that for sub-Gaussian distributions, results from \cite{Maillard2018GLR} show that the smaller and more explicit threshold
$\beta(n,\delta) = \left(1 + \frac{1}{n}\right)\ln\left(\frac{3n\sqrt{n}}{\delta}\right)$,
can be used to prove an upper bound of $\delta$ for the false alarm probability of the GLR, with quadratic divergence $\kl(x,y)=2(x-y)^2$.
%
For the Bernoulli GLR, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative, and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$.

Also note that, as any test based on scan-statistics, the GLR can be costly to implement as at every time step, it considers all previous time steps as a possible positions for a change-point. Thus, in practice the following adaptation may be interesting, based on down-sampling the possible time steps:
\begin{equation}\label{def:6:GLRTricks}
    \tilde \tau_{\delta} = \inf \left\{ n \in \cN : \sup_{s \in \cS_n} \left[s \times d\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times d\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\},
\end{equation}
for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant enjoys the exact same false-alarm control.
However, the detection delay may be slightly increased. In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups has little impact on the regret.

% ------------------------------------------------------------------------
\section{The \GLRklUCB{} Algorithm}
\label{sec:6:GLRklUCB_Algorithm}

Our proposed algorithm,  \GLRklUCB{},  combines a bandit algorithm with a change-point detector running on each arm.
It also needs a third ingredient, some forced exploration parameterized by $\alpha\in(0,1)$ to ensure each arm is sampled enough and changes can also be detected on arms currently under-sampled by the bandit algorithm.
%We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
% Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
%
\GLRklUCB{} combines the \klUCB{} algorithm (\cite{KLUCBJournal}), known to be optimal for Bernoulli bandits, with the Bernoulli GLR change-point detector introduced in Section~\ref{sec:6:ChangePointDetector}.
%
This algorithm, formally stated as Algorithm~\ref{algo:6:GLRklUCB}, can be used in any bandit model with bounded rewards, and is expected to be very efficient for Bernoulli distributions, which are relevant for practical applications.

\begin{figure}[h!]
    \centering
    \begin{small}
    % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
    % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
    % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
    \begin{algorithm}[H]
        % XXX Options
        % \LinesNumbered  % XXX Option to number the line
        % \RestyleAlgo{boxed}
        % XXX Input, data and output
        \KwIn{\emph{Problem parameters}: $T\in\N^*$, $K\in\N^*$\;}
        \KwIn{\emph{Algorithm parameters}: exploration probability $\alpha \in (0, 1)$, confidence level $\delta>0$\;}
        \KwIn{\emph{Option}: \textbf{Local} or \textbf{Global} restart\;}
        % \KwData{Data}
        % \KwResult{Result}
        % XXX Algorithm
        \textbf{Initialization: } $\forall i \in \{1,\dots,K\}$, $\tau_i \leftarrow 0$ and $n_i \leftarrow 0$ \\
        \For{$t=1,2,\ldots, T$}{
            \uIf{
                $\alpha>0$ and $t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor \in \{1,\dots,K\}$
            }{
                $A_t \leftarrow t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor$
                \tcp*[f]{forced exploration} \\
            }
            \Else{
                $A_t \leftarrow \arg\max\limits_{i\in \{1,\dots,K\}} \mbox{UCB}_i(t)$ as defined in \eqref{def:6:UCB}
            }
            \mbox{Play arm } $A_t$ \mbox{and receive the reward } $X_{A_t,t}$ : $n_{A_t} \leftarrow n_{A_t} + 1; Z_{A_t, n_{A_t}} \leftarrow X_{A_t,t}$ \\
            \uIf(\tcp*[f]{Change point is detected}){
                $\GLR_\delta(Z_{A_t,1}, \ldots, Z_{A_t, n_{A_t}})$ = True
            }
            {
                \uIf{Global restart}{
                    $\forall i\in \{1,\dots,K\}, \tau_i \leftarrow t$ and $n_i \leftarrow 0 $.
                    \tcp*[f]{global restart}
                    }
                \Else{
                    $\tau_{A_t} \leftarrow t$ and $n_{A_t} \leftarrow 0$
                    \tcp*[f]{local restart}
                }
            }
        }
        \caption{\GLRklUCB, with \textbf{Local} or \textbf{Global} restarts}
        \label{algo:6:GLRklUCB}
    \end{algorithm}
    \end{small}
\end{figure}

The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms. A restart happens when the Bernoulli GLR change-point detector detects a change on the arm that has been played (line $9$).
%
To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
\begin{small}
    \[\sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),\]
\end{small}%
%
with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}. We define the (\klUCB{}-like) index used by our algorithm, by denoting
$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$, $n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)$ the number of selections of arm $i$,
and $\hat{\mu}_i(t) = (1/n_i(t)) \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)$ their empirical mean (if $n_i(t)\neq0$).
%
With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ (if $t>1$ else $f(t)=0$),
the index is defined as
\begin{equation}\label{def:6:UCB}
    \UCB_i(t) := \max \bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\hat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \bigr\}.
\end{equation}

In this work, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
%
Under local restart, in the general case the times $\tau_i(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate}.
%
One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{This alternative choice is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB{}: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq. $(31)$-$(32)$.}:
$f(t-\tau_i(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{i=1}^K n_i(t)$.

The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by (\cite{CaoZhenKvetonXie18}), whereas \CUSUMUCB{} performs randomized exploration.
A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples} (proved in Appendix~\ref{proof:6:EnoughSamples}).

% \vspace*{-8pt}% WARNING
\begin{proposition}\label{prop:6:EnoughSamples}
    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $i$ (\ie, for a $k\in\{1,\dots,\NCi\}$, one has $\tau_i^{(k)} = \tau_i(t) < s \leq t < \tau_i^{(k+1)}$) it holds that $n_i(t) - n_i(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor$.
\end{proposition}

\subsection{Results for \GLRklUCB{} using Global Changes}

Recall that $\tau^{(k)}$ denotes the position of the $k$-th break-point and let $\mu_i^{(k)}$ be the mean of arm $i$ on the segment between the $k$ and $(k+1)$-th breakpoint:
$\forall t \in \{ \tau^{(k-1)}+1, \dots, \tau^{(k)} \}, \mu_{i}(t) = \mu_i^{(k)}$. We also introduce $k^* = \arg\max_i \mu_i^{(k)} $ and the largest gap at break-point $k$ as $\Delta^{(k)} := \max_{i=1,\dots,K} |\mu_i^{(k)} - \mu_i^{(k-1)}| >0$.

Assumption~\ref{ass:6:LongPeriodsGlobal} below is easy to interpret and standard in non-stationary bandits. It requires that the distance between two consecutive breakpoints is large enough: how large depends on the magnitude of the largest change that happen at those two breakpoints.
Under this assumption, we provide in Theorem~\ref{thm:6:mainRegretBoundGlobal} a finite time problem-dependent regret upper bound.
It features the parameters $\alpha$ and $\delta$,
the KL-divergence terms $\kl(\mu_{i}^{(k)},\mu_{k^*}^{(k)})$ expressing the hardness of the (stationary) MAB problem between two breakpoints,
and the $\Delta^{(k)}$ terms expressing the hardness of the change-point detection problem.

\begin{assumption}\label{ass:6:LongPeriodsGlobal}
    Define
    % \begin{equation}\label{def:6:DelayGlobal}
    $
        d^{(k)} = d^{(k)}(\alpha,\delta) = \lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \rceil.
    $
    % \end{equation}
    %
    Then we assume that for all $k \in \{1,\dots,\Upsilon_T\}$,
    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
\end{assumption}

\begin{theorem}
    \label{thm:6:mainRegretBoundGlobal}
    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
    \begin{small}
\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.\]
    \end{small}
\end{theorem}

\begin{corollary}\label{cor:6:Global}
    For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriodsGlobal},
    let $\Delta^{\text{opt}}$ denote the smallest value of a sub-optimality gap on one of the stationary segments, and $\Delta^{\text{change}}$ be the smallest magnitude of any change point on any arm.
    \begin{enumerate}
        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \Upsilon_T\sqrt{T\ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$,
        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2}\sqrt{\Upsilon_T T \ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$.
    \end{enumerate}
\end{corollary}

% \TODOE{note that in this corollary $\Delta^{\text{change}}$ could be defined as the smallest value of $\Delta^{(k)}$ which can be larger than the smallest change on any arm... But for the sake of uniformity between the two corollaries, we may prefer to keep it like that}

\subsection{Results for \GLRklUCB{} using Local Changes}

A few new notation are needed to state a regret bound for \GLRklUCB{} using local changes. We let $\tau_i^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $i$}: $\tau_i^{(\ell)} = \inf \{ t > \tau_i^{(\ell - 1)} : \mu_i(t) \neq \mu_i(t+1)\}$,
with the convention $\tau_i^{(0)}=0$, and let $\overline{\mu}_i^{(\ell)}$ be the $\ell$-th value for the mean of arm $i$, such that $\forall t \in [\tau_i^{(\ell-1)}+1, \tau_i^{(\ell)}], \ \ \mu_i(t) = \overline{\mu}_i^{(\ell)}$. We also introduce the gap $\Delta_i^{(\ell)} = \overline{\mu}_i^{\ell} - \overline{\mu}_i^{\ell-1} > 0$.
% ($\ell$ is used instead of $k$ to distinguish notations between local and global changes)

Assumption~\ref{ass:6:LongPeriods} requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points). Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{i=1}^K \NCi \leq K \Upsilon_T$.

\begin{assumption}\label{ass:6:LongPeriods}
    Define
    % \begin{equation}\label{def:6:DelayLocal}
    $
        d_i^{(\ell)} = d_i^{(\ell)}(\alpha,\delta) = \lceil \frac{4K}{\alpha\left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha}\rceil.
    $
    % \end{equation}
    %
    We assume that for all arm $i$ and all $\ell \in \{1,\dots,\NCi\}$, $\tau_i^{(\ell)} - \tau_i^{(\ell-1)} \geq 2\max (d_i^{(\ell)},d_i^{(\ell-1)})$.
\end{assumption}

\begin{theorem}
    \label{thm:6:mainRegretBound}For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on Local Restart satisfies
    %
    \begin{small}
    \[R_T \leq 2\sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{4K}{\alpha \left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},\]
    \end{small}
    where
    ${\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}$.
\end{theorem}

\begin{corollary}\label{cor:6:Local}
    For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriods},
    with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, the following holds.
    \begin{enumerate}
        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$,
        % \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
        % \item Choosing $\alpha = \sqrt{\frac{\EffChange\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\EffChange T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
        \item Choosing $\alpha = \sqrt{\frac{K\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{K\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$.
    \end{enumerate}
\end{corollary}

\subsection{Interpretation}

Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} both show that there exists a tuning of $\alpha$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms. For very particular instances such that $\Upsilon_T = \EffChange$, \ie, at each break-point only one arm changes (\eg, problem $1$ in Section~\ref{sec:6:NumericalExperiments}), the theory advocates the use of local changes.
%
Indeed, while the regret guarantees obtained are similar, those obtained for local changes hold for a wider variety of problems as Assumption~\ref{ass:6:LongPeriods} is less stringent than \ref{ass:6:LongPeriodsGlobal}.
%
Besides those specific instances, our results are essentially worse for local than global changes. However, we only obtain regret upper bounds -- thus providing a theoretical safety net for both variants of our algorithm, and the practical story is different, as discussed in Section~\ref{sec:6:NumericalExperiments}. We find that \GLRklUCB{} performs better with local restarts.

One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, our regret bounds hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
Hence those guarantees are valid on ``easy'' problem instances only, with few changes of a large magnitude (\eg, not for problems $3$ or $5$).
%
However, this does not prevent our algorithms from performing well on more realistic instances, and numerical experiments support this claim.
% (see problem~3 in Section~\ref{sec:6:NumericalExperiments}).
Note that \MUCB{} (\cite{CaoZhenKvetonXie18}) is also analyzed for the same type of unrealistic assumptions, while its practical performance is illustrated beyond those.

\section{A Unified Regret Analysis}
\label{sec:6:RegretAnalysis}

\TODOL{Rewrite, include the proof of at least one case here! No need to be that quick for the thesis, we only wrote it like this for the COLT version \cite{Besson2019GLRT}.}

In this section, we sketch a unified proof for Theorem~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound}, whose detailed proofs can be found in Appendix~\ref{proof:6:mainRegretBoundGlobal} and \ref{proof:6:mainRegretBound} respectively.
We emphasize that our approach is significantly different from those proposed by \cite{CaoZhenKvetonXie18} for \MUCB{} and by \cite{LiuLeeShroff17} for \CUSUMUCB.

Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$. Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
\begin{small}
\begin{align*}
 &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
 & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
\end{align*}%
\end{small}%
%
Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
%
\begin{footnotesize}
\begin{eqnarray*}
R_T \leq T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} + \ \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
\end{eqnarray*}%
\end{footnotesize}%
%
Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay. Each regret bound then follows from upper bounds on term (A), term (B), and on the failure probability $\bP(\cE_T^c)$.
%
To control (A) and (B), we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.

The tricky part of each proof, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$. For example for local changes (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
\[
    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
\]
where $\hat\tau_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}. Introducing the event $\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \hat\tau_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\}$ that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition:
\begin{align*}
    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
\end{align*}

Term (a) is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
Observe that under the bandit algorithm, the change point detector associated to arm $i$ is based on (possibly much) less than $t - \tau_i(t)$ samples from arm $i$, which makes false alarm even less likely to occur. Hence, it is easy to show that $(a) \leq \delta$.

Term (b) is related to the control of the detection delay, which is more tricky to obtain under the \GLRklUCB{} adaptive sampling scheme,
when compared to a result like Theorem~6 in \cite{Maillard2018GLR} for the change-point detector run in isolation.
%
More precisely, we need to leverage the forced exploration (Proposition~\ref{prop:6:EnoughSamples}) to be sure we have enough samples for detection. This explains why delays defined in Assumption~\ref{ass:6:LongPeriods} are scaled by $1/\alpha$.
Using some elementary calculus and a concentration inequality given in Lemma~\ref{lem:6:Chernoff2arms}, we can finally prove that $(b) \leq \delta$.
%
Finally, the ``bad event'' is unlikely: $\bP(\cE_T^c) \leq 2\EffChange \delta.$
%The detailed proof is provided in Appendix~\ref{proof:6:GoodEvent}.

% ----------------------------------------------------------------------------
\section{Experimental Results}
\label{sec:6:NumericalExperiments}

In this section we report results of numerical simulations performed on synthetic data to compare the performance of \GLRklUCB{} against other state-of-the-art approaches, on some piece-wise stationary bandit problems.
%
For simplicity, we restrict to rewards generated from Bernoulli distributions, even if \GLRklUCB{} can be applied to any bounded distributions.
% %
% Note that our proposal can also be applied to unbounded distributions, with a different divergence function used for the \GLR{} test (but results are yet to be obtained).
% \cite{Maillard2018GLR} analyzed the \GLR{} test, without considering the bandit setting, for Gaussian and $\sigma$-sub Gaussian distributions.

% -----------------------------------------------------------------
\paragraph{Algorithms and parameters tuning.}
\label{sub:6:ParametersTuning}

We include in our study two algorithms designed for the classical MAB, \klUCB{} (\cite{Garivier11KL}) and Thompson sampling (\cite{AgrawalGoyal11,Kaufmann12Thompson}),
as well as an ``oracle'' version of \klUCB, that we call Oracle-Restart. This algorithm knows the exact locations of the breakpoints, and restarts \klUCB{} at those locations (without any delay).

Then, we compare our algorithms to several competitor designed for a piece-wise stationary model. For a fair comparison, all algorithms that use  \UCB{} as a sub-routine were adapted to use \klUCB{} instead, which yields better performance\footnote{\cite{LiuLeeShroff17,CaoZhenKvetonXie18} both mention that extending their analysis to the use of \klUCB{} should not be too difficult.}. Moreover, all the algorithms are tuned as described in the corresponding paper, using in particular the knowledge of the number of breakpoints $\Upsilon_T$ and the horizon $T$. We first include three \emph{passively adaptive algorithms}:
Discounted \klUCB{} (D-\klUCB, \cite{Kocsis06}), with discount factor $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$; Sliding-Window \klUCB{} (SW-\klUCB,  \cite{Garivier11UCBDiscount}) using window-size $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$ and Discounted Thompson sampling (DTS, \cite{RajKalyani17}) with discount factor $\gamma = 0.95$. For this last algorithm, the discount factor $\gamma=0.75$ suggested by the authors was performing significantly worse on our problem instances.
%More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).

% \TODOE{Tu confirmes bien que $\gamma=0.75$ ne marchait sur AUCUN problème ? Car ta phrase actuelle n'exclut pas que ce choix marche pour les probèmes courts}

Our main goal is to compare against \emph{actively adaptive algorithms}. We  include \CUSUMklUCB{} (\cite{LiuLeeShroff17}), tuned with $M=150$ and $\varepsilon=0.1$ for easy problems ($1,2,4$) and $\varepsilon=0.001$ for hard problems ($3,5$),
% based on a prior knowledge of the difficulty of the problems,
and with $h = \log(T/\Upsilon_T)$, $\alpha = \sqrt{\Upsilon_T \log(T/\Upsilon_T) / T}$, as suggested in the paper.
% as for the first problems ($1$ to $4$) we consider have large changes in terms of means between change-points. Note that this choice of $\varepsilon$ should be too large for problems $5$ and $6$, but we do not want to tune the parameters for each problem (see Table~2 in \cite{LiuLeeShroff17}).
%
Finally, we include \MklUCB{} (\cite{CaoZhenKvetonXie18}),
% as a simple adaptation of the Monitored-UCB (\MUCB) algorithm,
tuned with $w=150$, based on a prior knowledge of the problems as the formula using $\delta_{\min}$ given in the paper is too large for small horizons (on all our problem instances), a threshold $b=\sqrt{w \log(2 K T)}$ and $\gamma=\sqrt{\Upsilon_T K (2 b + 3 \sqrt{w}) / (2 T)}$ as suggested by Remark~4 in the paper.

% \TODOE{Il faut être un peu plus précis, quelle connaissance du problème utilises-tu, et à quelle formule te réfères-tu ?}

For \GLRklUCB, we explore the two different options with \textbf{Local} and \textbf{Global} restarts,
using respectively
$\delta = 1/\sqrt{\Upsilon T}$, $\alpha = \alpha_0 \sqrt{\Upsilon_T \log(T) /T}$
and
$\delta = 1/\sqrt{K \Upsilon_T T}$, $\alpha = \alpha_0\sqrt{K \Upsilon_T \log(T) / T}$
from Corollaries \ref{cor:6:Global} and \ref{cor:6:Local}.
% Both choices of $\alpha$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
The constant is set to $\alpha_0 = 0.05$
(we show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\alpha\leq0.1$).
%
To speed up our simulations, two optimizations are used, with $\Delta n = \Delta s = 10$,
and \CUSUM{} also uses the first trick with $\Delta n = 10$ (see Appendix~\ref{sub:6:IdeasOptimizations} for more details).

% \TODOE{Je l'ai déjà dit, mais vraiment on va nous accuser d'avoir optimisé un paramètre chez nous et pas chez les autres. Est-ce qu'avec $\alpha_0=1$ on a des résultats comparables ou bien on est complètement à côté ???}
%
% \TODOE{On est d'accord que le speed-up tu ne l'as fait que pour les problèmes long ? Ca vaut le coup d'être écrit, sinon les lecteurs vont penser que notre algorithme est tellement compliqué qu'on ne l'a jamais implémenté réellement...}

% -----------------------------------------------------------------
\paragraph{Results.}
\label{sub:6:NumericalResults}

We report results obtained on three different piece-wise stationary bandit problems, illustrated in Figure~\ref{fig:6:Problems_123} in Appendix~\ref{app:6:Figures}, and described in more details below. Additional results on two more problems can be found in Appendix~\ref{sub:6:TwoMoreProblems}, where additional regret curves can be found for all problems. For each experiment, the regret was estimated using $1000$ independent runs.

% %
% We give in Appendix~\ref{app:6:NumericalResults} the plots of the cumulated mean regret $R_t$ (as a function of $t$), as well as plots of histograms of the distribution of final regret $R_T$, for the considered algorithms.

% - Pb 1 changes are only on one arm at a time
\textbf{Problem $\bm 1$.} There are $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1}.
% (and Figure~\ref{fig:6:Problem_1} in Appendix),
Note that changes happen on only one arm (\ie, $C_T=\Upsilon_T=4$),
and the optimal arm changes once at $t=\tau_2^{(1)}=2000$, with a large gap $\Delta=0.6$.

% - Pb 2 changes are on all arms at a time
\textbf{Problem $\bm 2$.} (see Figure~\ref{fig:6:Problem_2}). This problem is close to Problem 1, with a minimum optimality gap of $0.1$.
However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gap $\Delta=0.1$.
The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$}).

\textbf{Problem $\bm 3$.} (see Figure~\ref{fig:6:Problem_5}) This problem is harder, with $K=6$, $\Upsilon_T=8$ and $T=20000$.
Most arms change at almost every time steps,
and means are bounded in $[0.01, 0.07]$.
The gaps $\Delta$ are much smaller than for the first problems, with amplitudes ranging from $0.02$ to $0.001$.
Note that the assumptions of our regret upper bounds are violated, as well as the assumptions for the analysis of \MUCB{} and \CUSUMUCB.
This problem is inspired from Figure~3 of \cite{CaoZhenKvetonXie18}, where the synthetic data was obtained from manipulations on a real-world database of clicks from \emph{Yahoo!}.

% \TODOE{tu as bien tuné CUSUM-UCB avec $\epsilon = 0.02$ dans ce cas?}

% \TODOE{La notion $\Delta^{\text{change}}$ est déjà définie dans la section 4, et ici elle ne veut pas dire la même chose...}

Table \ref{table:6:totalResults1} shows the final regret $R_T$ obtained for each algorithm.
Results highlighted in \textbf{bold} show the best non-oracle algorithm for each experiments,
with our proposal being the best non-oracle strategy for problems $1$ and $2$.
Thompson sampling and \klUCB{} are efficient,
and better than Discounted-\klUCB{} which is very inefficient.
DTS and SW-\klUCB{} can sometimes be more efficient than their stationary counterparts, but perform worse than the Oracle and most actively adaptive algorithms.
M-\klUCB{} and \CUSUMklUCB{} outperform the previous algorithms, but \GLRklUCB{} is often better.
%
On these problems, our proposal with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts. Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet our experiments show the superiority of local restarts on every instances.

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|cccccc}
    \textbf{Algorithms} \textbackslash \textbf{Problems} & Pb $1$ & Pb $2$ & Pb $3$ \\
        \hline
        Oracle-Restart \klUCB{} & $\mathbf{37 \pm 37}$ & $\mathbf{45 \pm 34}$ & $\mathbf{257 \pm 86}$ \\
        \hline
        \klUCB{} & $270 \pm 76$ & $162 \pm 59$ & $529 \pm 148$ \\
        Discounted-\klUCB{} & $1456 \pm 214$ & $1442 \pm 440$ & $1376 \pm 37$ \\
        SW-\klUCB{} & $177 \pm 34$ & $182 \pm 34$ & $1794 \pm 71$ \\
        \hline
        Thompson sampling & $493 \pm 175$ & $388 \pm 147$ & $1019 \pm 245$ \\
        DTS & $209 \pm 38$ & $249 \pm 39$ & $2492 \pm 52$ \\
        \hline
        \MklUCB{} & $290 \pm 29$ & $534 \pm 93$ & $645 \pm 141$ \\
        \CUSUMklUCB{} & $148 \pm 32$ & $152 \pm 42$ & $\mathbf{490 \pm 133}$ \\
        \hline
        \GLRklUCB{}(Local) & $\mathbf{74 \pm 31}$ & $\mathbf{113 \pm 34}$ & $513 \pm 97$ \\
        \GLRklUCB{}(Global) & $97 \pm 32$ & $134 \pm 33$ & $621 \pm 103$
    \end{tabular}
    \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on problems $1$, $2$ (with $T=5000$) and $3$ ($T=20000$).
    }
    \label{table:6:totalResults1}
% \end{footnotesize}
\end{table}

One can note that the best non-oracle strategies are actively adaptive,
thus our experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.

%----------------------------------------------------------------------------
\section{Conclusion of Chapter 6}

We proposed a new algorithm for the piece-wise stationary bandit problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector. This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but \emph{without any other prior knowledge on the problem}, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change. We also gave numerical evidence of the efficiency of our proposal.

We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this paper (and in previous work), that would require only a few ``meaningful'' changes to be detected. This interesting research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$. We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as our proposal \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
Finally, we would like to study in the future possible extension of our approach to the slowly varying model (\cite{WeiSrivastava18}).
Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm (\cite{Besson2018ALT}), to efficiently tackle the piece-wise stationary distributed multi-player bandit model introduced in \cite{Wei2018}.

\paragraph{Acknowledgments}
Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.

\paragraph{A note on the simulation code}
The page \texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html} explains how to reproduce the experiments used for this Section.

% % ----------------------------------------------------------------------------
% \bibliographystyle{alpha}
% \bibliography{biblioBandits}

### 2019-02-19 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Small modifications today. Hard to get anything important done, I had a review to do for a COLT paper...

@@
 % -----------------------------------------------------------------
-\section{Multi-Player Bandit Model with Different Feedback Levels}
+\section{Multi-player bandit model with different feedback levels}
 \label{sec:5:model}
@@
 % -----------------------------------------------------------------
-\section{An Asymptotic Regret Lower Bound for a certain class of algorithms}
+\section{An improved asymptotic regret lower bound for a certain class of algorithms}
 \label{sec:5:lowerbound}
@@
 % -----------------------------------------------------------------
-\subsection{A Useful Regret Decomposition}
+\subsection{A useful regret decomposition}
 \label{sub:5:defregret}
@@
 % -----------------------------------------------------------------
-\subsection{An Improved Asymptotic Lower Bound on the Regret}
+\subsection{An improved asymptotic lower bound on the regret}
 \label{sub:5:betterLowerBound}
@@
 % -----------------------------------------------------------------
-\subsection{Towards Regret Upper Bounds}
+\subsection{Towards regret upper bounds}
 
@@
 % -----------------------------------------------------------------
-\section{New Algorithms for Multi-Player Bandits}
+\section{New algorithms for multi-player bandits}
 \label{sec:5:algorithms}
@@
 
-\subsection{Two New Strategies Based on Indices and Orthogonalization: \RandTopM{} and \MCTopM}
+\subsection{Two new strategies based on indices and orthogonalization: \RandTopM{} and \MCTopM}
 \label{sub:5:RandTopM_and_MCTopM}
@@
 % -----------------------------------------------------------------
-\subsection{The \Selfish{} Approach}
+\subsection{The \Selfish{} approach}
 \label{sub:5:Selfish}
@@
 % -----------------------------------------------------------------
-\section{Empirical performances}
+\section{Numerical simulations}
 \label{sec:5:experiments}
@@
 % -----------------------------------------------------------------
-\subsection{Common Analysis for \RandTopM- and \MCTopM-\klUCB{}}\label{sub:5:UpperBoundSelections}
+\subsection{Common analysis for \RandTopM- and \MCTopM-\klUCB{}}\label{sub:5:UpperBoundSelections}
 
@@
 % -----------------------------------------------------------------
-\subsection{Regret Analysis of \MCTopM-\klUCB}\label{sub:5:UpperBoundCollisions}
+\subsection{Regret analysis of \MCTopM-\klUCB}\label{sub:5:UpperBoundCollisions}
 

### 2019-02-19 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Small modifications today. Hard to get anything important done, I had a review to do for a COLT paper...

@@
 
-We report results obtained on three different piece-wise stationary bandit problems, illustrated in Figure~\ref{fig:6:Problems_123} in Appendix~\ref{app:6:Figures}, and described in more details below.
+We report results obtained on three different piece-wise stationary bandit problems, illustrated in Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2} and \ref{fig:6:Problem_5} in Appendix~\ref{app:6:Figures}, and described in more details below.
 Additional results on two more problems can be found in Appendix~\ref{sub:6:TwoMoreProblems}, where additional regret curves can be found for all problems.

### 2019-02-20 2-Chapters/3-Chapter/chapter3.tex
Okay for chapters and parts
Start to copy/paste all what I had written about SMPyBandits. This chapter 3 will be very quickly done.

@@
 
-% First chapter begins here
-\chapter{SMPyBandits: a state-of-the-art open-source Python library to simulate MAB problems}
+\chapter{SMPyBandits: a state-of-the-art Python library to simulate MAB problems}
 \label{chapter:3}
 \minitoc
+
+SMPyBandits is a package I have developed since the beginning of my PhD.
+It is designed to allow easy and efficient numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms, and SMPyBandits is written in Python.
+This library is by far the most complete open-source implementation of state-of-the-art algorithms tackling various kinds of MAB sequential learning problems.
+It aims at being extensive, simple to use and maintain, with a clean and well documented codebase.
+It allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments.
+Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
+%
+More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markovian or non-stationary).
+
 \newpage
@@
 
-This chapter is intended as a long version of the paper presenting SMPyBandits that I wrote in Summer 2018, see https://hal.inria.fr/hal-01840022
+% This chapter is intended as a long version of the paper presenting SMPyBandits that I wrote in Summer 2018, see https://hal.inria.fr/hal-01840022
+
+SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms tackling various kinds of sequential learning problems referred to as Multi-Armed Bandits.
+It aims at being extensive, simple to use and maintain, with a clean and perfectly documented codebase.
+But most of all it allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments while starting them (see below for an example).
+
+SMPyBandits does not aim at being blazing fast or perfectly memory efficient, and comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
+Even if some critical parts are also available as a \texttt{C} Python extension, and even by using Numba \cite{numba} whenever it is possible, if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia.
+
+In this Chapter~\ref{chapter:3}, we start by presenting in Section~\ref{sec:3:presentationLibrary} the organization of the library.
+We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
+then we present in Section~\ref{sec:3:timeAndMemoryCosts} additional numerical simulations to discuss about the time efficient and memory costs of the some MAB algorithms (we focus on the ones being used in the rest of this thesis, \UCB, \klUCB, and Thompson Sampling).
+To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms, but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
+%
+Finally, in Appendix~\ref{sec:3:appendix} we include some minimalist files showing how to use the library, as well as details regarding the possible use of multiple cores to speed-up simulations.
 
@@
 
-I want to show what is SMPyBandits, what problems does it solve, what did I implement, how to use it.
+% I want to show what is SMPyBandits, what problems does it solve, what did I implement, how to use it.
+
+We start by explaining how SMPyBandits simulates MAB problems, by detailing the different parts.
+We then detail how SMPyBandits implements MAB algorithms (\eg, \UCB),
+and what kind of information is displayed, saved and plotted after a batch of simulations of bandit problems.
+
+The main goal of the library is to be easily able to simulate a small number of algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-player) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
+After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithms, as well as regret and other visualizations.
+The same problem is simulated for $N$ independent repetitions (\eg, $N=1000$) in order to show mean results with a low variance.
+
+\subsection{Single- and multi-player MAB problems}
+
+For the classical single-player stochastic MAB model, as defined in Chapter~\ref{chapter:2},
+a stochastic MAB problem is defined by $K>1$ distributions $\nu_k$ (also called arms),
+used to sample \iid{} rewards $r_k(t) \sim \nu_k, \forall t$.
+An agent choose arm $A(t)\in\{1,\dots,K\}$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
+Her goal is to maximize $\sum_{t=1}^T r_{A(t)}(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.
+
+\paragraph{Simulation loop}
+%
+Any simulation library targeting single-player bandit problems must implement at least three components:
+reward distributions, MAB algorithm, and a simulation loop that essentially looks like this.
+\begin{itemize}
+    \item First, initialize the MAB problem and one or more algorithms,
+    \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
+Ask algorithm $\cA$ his chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm,
+    \item At the end, compute the accumulated reward, the regret, plot visualizations etc.
+\end{itemize}
+%
+Note that the second step should of course be repeated a large number of times if we want to study \emph{mean} regret and not only the regret in one single trajectory.
+
+If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and ensure XXX.
+
+\paragraph{Simulation loop for multi-player bandit}
+%
+We refer to Chapter~\ref{chapter:5} which introduces and details various multi-player bandit models.
+Any simulation library targeting multi-player bandit problems must implement at least another component:
+a collision model, and a simulation loop that essentially looks like this.
+\begin{itemize}
+    \item First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
+    \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
+Ask each player $\cA^{(j)}$ his chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
+The
+    sample (random) feedback $r_k(t)$ (\iid) from distributions $\nu_{k}$, and compute the rewards $Y^{(j)}(t)$ from the random feedback and the collision model.
+Finally feeds the observation $(A^{(j)}(t),Y^{(j)}(t))$ to the $M$ players,
+    \item At the end, compute the accumulated reward, the regret, plot visualizations etc.
+\end{itemize}
+
+\paragraph{Reward distributions}
+%
+This library tackles one dimensional distributions,
+and supports [`Bernoulli`](https://smpybandits.github.io/docs/Arms.Bernoulli.html), [`binomial`](https://smpybandits.github.io/docs/Arms.Binomial.html), [`Poisson`](https://smpybandits.github.io/docs/Arms.Poisson.html), and a generic [`discrete`](https://smpybandits.github.io/docs/Arms.DiscreteArm.html) distributions,
+as well as [`exponential`](https://smpybandits.github.io/docs/Arms.Exponential.html), [`gamma`](https://smpybandits.github.io/docs/Arms.Gamma.html), [`Gaussian`](https://smpybandits.github.io/docs/Arms.Gaussian.html) and [`uniform`](https://smpybandits.github.io/docs/Arms.Uniform.html) continuous distributions,
+which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$).
+
+SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
+containing [over 65](https://SMPyBandits.GitHub.io/docs/Policies.html) algorithms.
+It uses a well-designed hierarchical structure and [class inheritance scheme](https://SMPyBandits.GitHub.io/uml_diagrams/README.html) to minimize redundancy in the codebase.
+For instance, most existing algorithms are index-based, and new ones can be written easily by inheriting from the [`IndexPolicy`](https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html) class.
+An index-based algorithm computes an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$ and simply play $A(t) = \arg\max_k I_k(t)$.
+
+We use a well-designed hierarchical structure and [class inheritance scheme](https://smpybandits.github.io/uml_diagrams/README.html) to minimize redundancy in the codebase, and for instance the code specific to the UCB algorithm \cite{LaiRobbins85,Auer02} is as short as this (and fully documented), by inheriting from the [`IndexPolicy`](https://smpybandits.github.io/docs/Policies.IndexPolicy.html) class:
+
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+% \begin{listing}[h!]
+    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_a_IndexPolicy_UCB.py}
+    % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy}
+    \captionof{listing}{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy \label{lst:3:smallIndexPolicy}.}
+    % \label{lst:3:smallIndexPolicy}
+% \end{listing}
+\end{small}
+
+### Multi-Players MAB
+
+For Cognitive Radio and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the *same* $K$ arms.
+Whenever two or more players select the same arm at the same time, they all suffer from a collision.
+Different collision models has been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
+Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (*i.e.*, arms with highest means) without collisions.
+*SMPyBandits* implements [all collision models](https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html) found in the literature, as well as all the algorithms from the last 10 years (including [`rhoRand`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html), [`MEGA`](https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html), [`MusicalChair`](https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html), and our state-of-the-art algorithms [`RandTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html) and [`MCTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html) from @BessonALT2018).
+For comparison, realistic or full-knowledge centralized algorithms are also implemented.
+
+---
+
+\subsubsection{Features}
+
+With this numerical framework, simulations can run on a single CPU or a multi-core machine using joblib \cite{joblib},
+and summary plots are automatically saved as high-quality PNG, PDF and EPS, using matplotlib \cite{matplotlib} and seaborn \cite{seaborn}.
+Raw data from each simulation is also saved in a HDF5 file using h5py \cite{h5py}, an efficient and compressed binary format, to allow easy post-mortem exploration of simulation results.
+Making new simulations is very easy, one only needs to write a configuration script (\texttt{configuration.py}), without needing a complete knowledge of the internal code architecture.
+
+A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here: [`SMPyBandits.GitHub.io`](https://SMPyBandits.GitHub.io/).
+
+\subsubsection{How to run experiments?}
+
+We show how to install SMPyBandits, and an example of how to run a simple experiment.
+This bash snippet\footnote{~See [`SMPyBandits.GitHub.io/How_to_run_the_code.html`](https://SMPyBandits.GitHub.io/How_to_run_the_code.html) for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see [`pypi.org/project/SMPyBandits`](https://pypi.org/project/SMPyBandits/).
+You can install it directly with `sudo pip install SMPyBandits`, or from a `virtualenv` \cite{virtualenv}},
+and install the requirements for Python 3 (once):
+
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+    \begin{listing}[h!]
+        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
+# 1. get the code in the folder you want
+$ git clone https://GitHub.com/SMPyBandits/SMPyBandits.git
+$ cd SMPyBandits.git
+# 2. install the requirements
+$ pip install -r requirements.txt
+        }
+        \caption{Small snippet of Bash to download and install dependencies of SMPyBandits}
+        % \captionof{listing}{Small snippet of Bash to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
+        \label{lst:3:howToInstallLibrary}
+    \end{listing}
+\end{small}
+
+Launching simulations is easy, for instance this snippet shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
+It takes about $20$ minutes, on a standard $4$-cores $64$ bits GNU/Linux laptop.
+Using environment variables (\texttt{N=1000} etc) in the command line is not required, but it is convenient:
+
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+    \begin{listing}[h!]
+        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
+# 3. run a single-player simulation
+$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=4 \
+    MEANS=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] python3 main.py configuration.py
+        }
+        \caption{Small snippet of Bash to run a simple experiment with SMPyBandits}
+        % \captionof{listing}{Small snippet of Bash to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
+        \label{lst:3:howToRunBasicLibrary}
+    \end{listing}
+\end{small}
+
+### Example of a simulation and illustration
+
+A small script [`configuration.py`](https://SMPyBandits.GitHub.io/docs/configuration.html) is used to import the [arm classes](https://SMPyBandits.GitHub.io/docs/Arms.html), the [policy classes](https://SMPyBandits.GitHub.io/docs/Policies.html) and define the problems and the experiments.
+Choosing the algorithms is easy by customizing the `configuration["policies"]` list in the `configuration.py` file.
+For instance, one can compare the standard anytime [`klUCB`](https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html) algorithm against the non-anytime variant [`klUCBPlusPlus`](https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html) algorithm, and also [`UCB`](https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html) (with $\alpha=1$) and [`Thompson`](https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html) (with [Beta posterior](https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html)).
+
+```python
+configuration["policies"] = [
+  { "archtype": klUCB, "params": { "klucb": klucbBern } },
+  { "archtype": klUCBPlusPlus, "params": { "horizon": HORIZON, "klucb": klucbBern } },
+  { "archtype": UCBalpha, "params": { "alpha": 1 } },
+  { "archtype": Thompson, "params": { "posterior": Beta } }
+]
+```
+
+Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in [a HDF5 file](http://docs.h5py.org/en/stable/high/file.html)\footnote{~For example, this simulation produces this HDF5 file, [GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5](https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5).}.
+Figure \ref{fig:plot1} below shows the average regret for these $4$ algorithms.
+The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.
+
+\subsection{Dependencies}
+
+This library is written in Python \cite{python}, for versions 2.7+ or 3.4+, using \texttt{matplotlib} \cite{matplotlib} for 2D plotting, \texttt{numpy} \cite{numpy} for data storing, random number generations and operations on arrays, \texttt{scipy} \cite{scipy} for statistical and special functions, and \texttt{seaborn} \cite{seaborn} for pretty plotting and colorblind-aware colormaps.
+
+Optional dependencies include \texttt{joblib} \cite{joblib} for parallel simulations, \texttt{numba} \cite{numba} for automatic speed-up on small functions, \texttt{sphinx} \cite{sphinx} for generating the documentation, \texttt{virtualenv} \cite{virtualenv} for launching simulations in isolated environments, and \texttt{jupyter} \cite{jupyter} used with \texttt{ipython} \cite{ipython} to experiment with the code.
+
+\subsection{Research papers using SMPyBandits}
+
+SMPyBandits was used for the following research articles since $2017$:
+
+- For \cite{Besson2018ALT}, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page [`SMPyBandits.GitHub.io/MultiPlayers.html`](https://SMPyBandits.GitHub.io/MultiPlayers.html) on the documentation.}.
+We designed the two [`RandTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html) and [`MCTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html) algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (i.e., [`rhoRand`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html), [`MEGA`](https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html) and [`MusicalChair`](https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html)).
+
+- In \cite{Besson2018WCNC}, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page [`SMPyBandits.GitHub.io/Aggregation.html`](https://SMPyBandits.GitHub.io/Aggregation.html) on the documentation.}.
+We designed a variant of the Exp3 algorithm for online aggregation of experts \cite{Bubeck12}, called [`Aggregator`](https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html).
+Aggregating experts is a well-studied idea in sequential learning and in machine learning in general.
+We showed that it can be used in practice to select on the run the best bandit algorithm for a certain problem from a fixed pool of experts.
+This idea and algorithm can have interesting impact for Opportunistic Spectrum Access applications \cite{Jouini09} that use multi-armed bandits algorithms for sequential learning and network efficiency optimization.
+
+- In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page [`SMPyBandits.GitHub.io/DoublingTrick.html`](https://SMPyBandits.GitHub.io/DoublingTrick.html) on the documentation.}.
+In sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
+A well-known trick for transforming any non-anytime algorithm to an anytime variant is the "Doubling Trick": start with an horizon $T_0\in\mathbb{N}^$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
+We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results.
+A geometric sequence suffices to conserve minimax regret bounds (in $R_T = \mathcal{O}(\sqrt{T})$), with a constant multiplicative loss $\ell \leq 4$, but cannot be used to conserve a logarithmic regret bound (in $R_T = \mathcal{O}(\log(T))$).
+And an exponential sequence can be used to conserve logarithmic bounds, with a constant multiplicative loss also $\ell \leq 4$ in the usual setting.
+It is still an open question to know if a well-tuned exponential sequence can conserve minimax bounds, or "weak" minimax bounds (in $R_T = \mathcal{O}(\sqrt{T \log(T)})$).
+
+- In \cite{Besson2019GLRT}, we used SMPyBandits for piecewise stationary MAB models (single-player).
+We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page [`SMPyBandits.GitHub.io/NonStationary.html`](https://SMPyBandits.GitHub.io/NonStationary.html) on the documentation.}.
+XXX
 
@@
 % ----------------------------------------------------------------------------
-\section{An example of another model: rested or restless Markov models - FIXME maybe not included?}
+\section{An example of another model: rested or restless Markov models}
 \label{sec:3:markovModels}
+
+%  - FIXME maybe not included?
 
@@
 %
-With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
+With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^ = \mu_2$.
 \begin{verbatim}
@@
 
-\begin{verbatim}
-    (bash) $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
-\end{verbatim}
+% \begin{verbatim}
+%     (bash) $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
+% \end{verbatim}
+
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+    \begin{listing}[h!]
+        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
+$ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
+        }
+        \caption{Small snippet of Bash to run an experiment}
+        % \captionof{listing}{Small snippet of Bash to run an experiment \label{lst:3:howToRunExperiment2}.}
+        \label{lst:3:howToRunExperiment2}
+    \end{listing}
+\end{small}
 

### 2019-02-20 2-Chapters/3-Chapter/src/example_of_a_IndexPolicy_UCB.py

@@ -1,0 +1,22 @@
+from numpy import sqrt, log
+from .IndexPolicy import IndexPolicy
+
+class UCB(IndexPolicy):
+  """ The UCB policy for bounded bandits.
+  Reference: [Lai & Robbins, 1985]. """
+
+  def computeIndex(self, arm):
+    r""" Compute the current index, at time t and
+    after :math:`N_k(t)` pulls of arm k:
+
+    .. math::
+
+        I_k(t) = \frac{X_k(t)}{N_k(t)}
+        + \sqrt{\frac{2 \log(t)}{N_k(t)}}.
+    """
+    if self.pulls[arm] < 1:  # forced exploration
+      return float('+inf')   # in the first steps
+    else:                    # or compute UCB index
+      estimated_mean = (self.rewards[arm] / self.pulls[arm])
+      exploration_bias = sqrt((2 * log(self.t)) / self.pulls[arm])
+      return estimated_mean + exploration_bias

### 2019-02-20 2-Chapters/3-Chapter/chapter3.tex
Added a lot of content for Chapter 3 on SMPyBandits.

@@
 
+This Chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
+Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
+Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.
+
 \newpage
@@
 \graphicspath{{2-Chapters/3-Chapter/Images/}}
+\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
 
@@
 Finally, in Appendix~\ref{sec:3:appendix} we include some minimalist files showing how to use the library, as well as details regarding the possible use of multiple cores to speed-up simulations.
+
+\paragraph{Publications}
+%
+This Chapter is based on the following publication: \cite{SMPyBanditsJMLR}.
+The code code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is publicly released under the MIT open-source License.
 
@@
 
+% FIXME finish this paragraph
 If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and ensure XXX.
@@
 \paragraph{Simulation loop for multi-player bandit}
+
+For Cognitive Radio and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
+Whenever two or more players select the same arm at the same time, they all suffer from a collision.
 %
+Different collision models has been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
+Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
 We refer to Chapter~\ref{chapter:5} which introduces and details various multi-player bandit models.
+
+SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms from the last 10 years (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}).
+%
+It includes
+\texttt{rhoRand} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{PoliciesMultiPlayers.rhoRand}}) from \cite{Anandkumar11},
+\texttt{MEGA} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}}) \cite{Avner15},
+\texttt{MusicalChair} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{Policies.MusicalChair}}) from \cite{Rosenski16},
+and our state-of-the-art algorithms
+\texttt{RandTopM} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{PoliciesMultiPlayers.RandTopM}}) and
+\texttt{MCTopM} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}}) from \cite{Besson2018ALT}.
+For comparison, realistic or full-knowledge centralized algorithms are also implemented.
+
 Any simulation library targeting multi-player bandit problems must implement at least another component:
@@
 
-\paragraph{Reward distributions}
+\subsection{Reward distributions}
 %
 This library tackles one dimensional distributions,
-and supports [`Bernoulli`](https://smpybandits.github.io/docs/Arms.Bernoulli.html), [`binomial`](https://smpybandits.github.io/docs/Arms.Binomial.html), [`Poisson`](https://smpybandits.github.io/docs/Arms.Poisson.html), and a generic [`discrete`](https://smpybandits.github.io/docs/Arms.DiscreteArm.html) distributions,
-as well as [`exponential`](https://smpybandits.github.io/docs/Arms.Exponential.html), [`gamma`](https://smpybandits.github.io/docs/Arms.Gamma.html), [`Gaussian`](https://smpybandits.github.io/docs/Arms.Gaussian.html) and [`uniform`](https://smpybandits.github.io/docs/Arms.Uniform.html) continuous distributions,
+and supports \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Arms.Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Arms.Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Arms.Poisson}}), and a generic \emph{discrete} (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{Arms.DiscreteArm}}) distributions,
+as well as \emph{exponential} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Exponential.html}{Arms.Exponential}}), \emph{gamma} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gamma.html}{Arms.Gamma}}), \emph{Gaussian} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gaussian.html}{Arms.Gaussian}}) and \emph{uniform} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Uniform.html}{Arms.Uniform}}) continuous distributions,
 which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$).
 
+The default is to use the same distribution for the $K$ arms, but it also possible to mix them.
+
+% FIXME
+\TODOL{point to the notebook!}
+% https://perso.crans.org/besson/phd/notebooks/Easily_creating_MAB_problems.html
+
+\subsection{MAB algorithms}
+
 SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
-containing [over 65](https://SMPyBandits.GitHub.io/docs/Policies.html) algorithms.
-It uses a well-designed hierarchical structure and [class inheritance scheme](https://SMPyBandits.GitHub.io/uml_diagrams/README.html) to minimize redundancy in the codebase.
-For instance, most existing algorithms are index-based, and new ones can be written easily by inheriting from the [`IndexPolicy`](https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html) class.
+containing over 65 (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}) algorithms.
+It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams showed on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
+For instance, most existing algorithms are index-based, and new ones can be written easily by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
 An index-based algorithm computes an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$ and simply play $A(t) = \arg\max_k I_k(t)$.
-
-We use a well-designed hierarchical structure and [class inheritance scheme](https://smpybandits.github.io/uml_diagrams/README.html) to minimize redundancy in the codebase, and for instance the code specific to the UCB algorithm \cite{LaiRobbins85,Auer02} is as short as this (and fully documented), by inheriting from the [`IndexPolicy`](https://smpybandits.github.io/docs/Policies.IndexPolicy.html) class:
+For instance the code specific to the UCB algorithm \cite{LaiRobbins85,Auer02} is as short as this (and fully documented):
 
@@
 
-### Multi-Players MAB
-
-For Cognitive Radio and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the *same* $K$ arms.
-Whenever two or more players select the same arm at the same time, they all suffer from a collision.
-Different collision models has been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
-Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (*i.e.*, arms with highest means) without collisions.
-*SMPyBandits* implements [all collision models](https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html) found in the literature, as well as all the algorithms from the last 10 years (including [`rhoRand`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html), [`MEGA`](https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html), [`MusicalChair`](https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html), and our state-of-the-art algorithms [`RandTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html) and [`MCTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html) from @BessonALT2018).
-For comparison, realistic or full-knowledge centralized algorithms are also implemented.
-
----
-
 \subsubsection{Features}
@@
 
-A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here: [`SMPyBandits.GitHub.io`](https://SMPyBandits.GitHub.io/).
+A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here: \texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
 
@@
 We show how to install SMPyBandits, and an example of how to run a simple experiment.
-This bash snippet\footnote{~See [`SMPyBandits.GitHub.io/How_to_run_the_code.html`](https://SMPyBandits.GitHub.io/How_to_run_the_code.html) for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see [`pypi.org/project/SMPyBandits`](https://pypi.org/project/SMPyBandits/).
-You can install it directly with `sudo pip install SMPyBandits`, or from a `virtualenv` \cite{virtualenv}},
+This bash snippet\footnote{~See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}.
+You can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.},
 and install the requirements for Python 3 (once):
@@
     \begin{listing}[h!]
-        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
+        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
 # 1. get the code in the folder you want
@@
 $ pip install -r requirements.txt
-        }
+        \end{minted}
         \caption{Small snippet of Bash to download and install dependencies of SMPyBandits}
@@
 % https://tex.stackexchange.com/a/12430/
-\begin{small}
 \begin{listing}[h!]
-        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
+    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
 # 3. run a single-player simulation
 $ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=4 \
-    MEANS=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] python3 main.py configuration.py
-        }
+MEANS=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] \
+python3 main.py configuration.py
+    \end{minted}
     \caption{Small snippet of Bash to run a simple experiment with SMPyBandits}
@@
 \end{listing}
-\end{small}
 
-### Example of a simulation and illustration
+\subsection{Example of a simulation and illustration}
 
-A small script [`configuration.py`](https://SMPyBandits.GitHub.io/docs/configuration.html) is used to import the [arm classes](https://SMPyBandits.GitHub.io/docs/Arms.html), the [policy classes](https://SMPyBandits.GitHub.io/docs/Policies.html) and define the problems and the experiments.
-Choosing the algorithms is easy by customizing the `configuration["policies"]` list in the `configuration.py` file.
-For instance, one can compare the standard anytime [`klUCB`](https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html) algorithm against the non-anytime variant [`klUCBPlusPlus`](https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html) algorithm, and also [`UCB`](https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html) (with $\alpha=1$) and [`Thompson`](https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html) (with [Beta posterior](https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html)).
+A small script \texttt{configuration.py} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/configuration.html}{SMPyBandits.GitHub.io/docs/configuration.html}}) is used to import the arm classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{SMPyBandits.GitHub.io/docs/Arms.html}}), the policy classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{SMPyBandits.GitHub.io/docs/Policies.html}}) and define the problems and the experiments.
+Choosing the algorithms is easy by customizing the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
+For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{Policies.klUCB}}) against the non-anytime variant $\klUCB^{++}$ algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{Policies.klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{Policies.UCB}}) and Thompson sampling (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Policies.Thompson}}) with a Beta posterior (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Policies.Posterior.Beta}})).
 
-```python
+% https://tex.stackexchange.com/a/12430/
+\begin{listing}[h!]
+    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}
 configuration["policies"] = [
-  { "archtype": klUCB, "params": { "klucb": klucbBern } },
-  { "archtype": klUCBPlusPlus, "params": { "horizon": HORIZON, "klucb": klucbBern } },
-  { "archtype": UCBalpha, "params": { "alpha": 1 } },
-  { "archtype": Thompson, "params": { "posterior": Beta } }
+  {"archtype": klUCB, "params": {"klucb": klucbBern}},
+  {"archtype": klUCBPlusPlus, "params": {"horizon": HORIZON, "klucb": klucbBern}},
+  {"archtype": UCBalpha, "params": {"alpha": 1}},
+  {"archtype": Thompson, "params": {"posterior": Beta}}
 ]
-```
+    \end{minted}
+    \caption{Small snippet of Python code to configure the list of algorithms tested on a problem.}
+    % \captionof{listing}{Small snippet of Python code to configure the list of algorithms tested on a problem. \label{lst:3:howToConfigureAlgorithms}.}
+    \label{lst:3:howToConfigureAlgorithms}
+\end{listing}
 
-Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in [a HDF5 file](http://docs.h5py.org/en/stable/high/file.html)\footnote{~For example, this simulation produces this HDF5 file, [GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5](https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5).}.
-Figure \ref{fig:plot1} below shows the average regret for these $4$ algorithms.
+Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in a HDF5 file (\cite{h5py}, \texttt{\href{http://docs.h5py.org/en/stable/high/file.html}{docs.h5py.org/en/stable/high/file.html}})\footnote{~For example, this simulation produces this HDF5 file \texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}.}.
+Figure~\ref{fig:3:firstPlot} below shows the average regret for these $4$ algorithms.
 The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.
+The Figure~\ref{fig:3:firstPlot_hist} shows the histogram of regret for the same experiment.
+
+\begin{figure}[h!]  % [htbp]
+	\centering
+	\includegraphics[width=0.95\linewidth]{3.pdf}
+	\caption[Example of a single-player simulation showing the average regret and histogram of regrets of $4$ algorithms]{
+		Example of a single-player simulation showing the average regret and histogram of regrets of four algorithms.
+They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proved to match the lower-bound up-to a constant), and each but UCB is known to be optimal (\ie, with the constant matching the lower-bound).
+For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).
+	}
+	\label{fig:3:firstPlot}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+	\centering
+	\includegraphics[width=0.95\linewidth]{3_hist.pdf}
+	\caption{Histogram of regret for the same experiment.}
+	\label{fig:3:firstPlot_hist}
+\end{figure}
 
@@
 
+All the quoted libraries are free and open-source and can be installed in one command using the \texttt{pip} (\texttt{\href{https://pip.pypa.io/}{pip.PyPa.io}}) or \texttt{conda} (\texttt{\href{http://conda.anaconda.org/}{conda.anaconda.org}}) package manager.
+
 \subsection{Research papers using SMPyBandits}
 
-SMPyBandits was used for the following research articles since $2017$:
+SMPyBandits was used for the following research articles since $2017$.
 
-- For \cite{Besson2018ALT}, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page [`SMPyBandits.GitHub.io/MultiPlayers.html`](https://SMPyBandits.GitHub.io/MultiPlayers.html) on the documentation.}.
-We designed the two [`RandTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html) and [`MCTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html) algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (i.e., [`rhoRand`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html), [`MEGA`](https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html) and [`MusicalChair`](https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html)).
-
-- In \cite{Besson2018WCNC}, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page [`SMPyBandits.GitHub.io/Aggregation.html`](https://SMPyBandits.GitHub.io/Aggregation.html) on the documentation.}.
-We designed a variant of the Exp3 algorithm for online aggregation of experts \cite{Bubeck12}, called [`Aggregator`](https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html).
+In \cite{Besson2018WCNC} (see Chapter~\ref{chapter:2}), we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation.}.
+We designed a variant of the Exp3 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
 Aggregating experts is a well-studied idea in sequential learning and in machine learning in general.
@@
 
-- In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page [`SMPyBandits.GitHub.io/DoublingTrick.html`](https://SMPyBandits.GitHub.io/DoublingTrick.html) on the documentation.}.
+- For \cite{Besson2018ALT} (see Chapter~\ref{chapter:5}), we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
+
+In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
 In sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
@@
 
-- In \cite{Besson2019GLRT}, we used SMPyBandits for piecewise stationary MAB models (single-player).
-We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page [`SMPyBandits.GitHub.io/NonStationary.html`](https://SMPyBandits.GitHub.io/NonStationary.html) on the documentation.}.
+In \cite{Besson2019GLRT} (see Chapter~\ref{chapter:6}), we used SMPyBandits for piece-wise stationary MAB models (single-player).
+We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
 XXX
+
+\TODOL{Write a better explanation of this last paper.}
 
@@
 %
-With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^ = \mu_2$.
+With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
 \begin{verbatim}
@@
 % https://tex.stackexchange.com/a/12430/
-\begin{small}
+% \begin{small}
 % \begin{listing}[h!]
@@
     % \caption{Small example of \texttt{main.py} file}
-    \captionof{listing}{Example of use of SMPyBandits \label{lst:3:exampleOfUse}.}
+    \captionof{listing}{Example of use of SMPyBandits\label{lst:3:exampleOfUse}.}
     % \label{lst:3:exampleOfUse}
 % \end{listing}
-\end{small}
+% \end{small}
 
@@
 % https://tex.stackexchange.com/a/12430/
-\begin{small}
 % \begin{listing}[h!]
@@
 % \end{listing}
-\end{small}
 
@@
 % https://tex.stackexchange.com/a/12430/
-\begin{small}
 \begin{listing}[h!]
-        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
+    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
 $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
-        }
+    \end{minted}
     \caption{Small snippet of Bash to run an experiment}
@@
 \end{listing}
-\end{small}
 
 % https://tex.stackexchange.com/a/12430/
-\begin{small}
 % \begin{listing}[h!]
@@
 
-Quickly explain the advantages and drawbacks of using Joblib and joblib.Parallel to run simulations on multiple cores of the same computer
+\TODOL{Write it!}
+
+Quickly explain the advantages and drawbacks of using \texttt{Joblib} and \texttt{joblib.Parallel} to run simulations on multiple cores of the same computer
 

### 2019-02-20 2-Chapters/3-Chapter/chapter3.tex
Done for now, I don't like the content of Chapter 3 at all for now, but at least there is something.

@@
 \texttt{rhoRand} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{PoliciesMultiPlayers.rhoRand}}) from \cite{Anandkumar11},
-\texttt{MEGA} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}}) \cite{Avner15},
+\texttt{MEGA} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}}) from \cite{Avner15},
 \texttt{MusicalChair} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{Policies.MusicalChair}}) from \cite{Rosenski16},
@@
 \texttt{MCTopM} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}}) from \cite{Besson2018ALT}.
-For comparison, realistic or full-knowledge centralized algorithms are also implemented.
+For comparison, realistic (\eg, \UCB{} for multiple play) or full-knowledge centralized algorithms are also implemented.
 
@@
 % https://tex.stackexchange.com/a/12430/
+\begin{small}
 \begin{listing}[h!]
@@
 \end{listing}
+\end{small}
 
@@
 \end{listing}
+\end{small}
 
-Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in a HDF5 file (\cite{h5py}, \texttt{\href{http://docs.h5py.org/en/stable/high/file.html}{docs.h5py.org/en/stable/high/file.html}})\footnote{~For example, this simulation produces this HDF5 file \texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}.}.
+Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in a HDF5 file (\cite{h5py}, \texttt{\href{http://docs.h5py.org/en/stable/high/file.html}{docs.h5py.org/en/stable/high/file.html}})\footnote{~For example, this simulation produces this HDF5 file\\\texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}}.
 Figure~\ref{fig:3:firstPlot} below shows the average regret for these $4$ algorithms.
@@
 In sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
-A well-known trick for transforming any non-anytime algorithm to an anytime variant is the "Doubling Trick": start with an horizon $T_0\in\mathbb{N}^$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
+A well-known trick for transforming any non-anytime algorithm to an anytime variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
 We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results.
@@
 % https://tex.stackexchange.com/a/12430/
-% \begin{small}
+\begin{small}
 % \begin{listing}[h!]
@@
 % \end{listing}
-% \end{small}
+\end{small}
 
@@
 % https://tex.stackexchange.com/a/12430/
+\begin{small}
 % \begin{listing}[h!]
@@
 % \end{listing}
+\end{small}
 

### 2019-02-26 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Small modifications today. Hard to get anything important done, I had a review to do for a COLT paper...
Proofreading of Chapter 4. No new science since the papers... but it's cleanly presented and unified.

@@
 The aim of this Section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases, and the stochastic hypothesis is more and more questionable.
-To do that, we suppose an IoT network made of two types of devices: static devices that use only one channel (fixed in time), and dynamic devices that can choose the channel for each of their transmissions.
+To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
 Static devices form an interfering traffic, which could have been generated by devices using other standards as well.
@@
 %
-We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of object sends uplink packets to the network gateway, for instance in the $433.5\;\mathrm{MHz}$ ISM band (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
+We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of object sends uplink packets to the network gateway.
+Messages can be sent for instance in the $433.5\;\mathrm{MHz}$ ISM band (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
 The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
@@
 The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$).
-In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s in Section~\ref{sec:4:gnuradio}).
+In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1\;\mathrm{s}$ in our implementation).
 
@@
     \includegraphics[width=0.70\linewidth]{system_model1.eps}
-    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
+    \caption{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
     \label{fig:41:system_model1}
@@
 %
+For simplicity, and for consistency with the rest of this thesis, we only consider a discrete time protocol.
 As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
 All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
-In each time slot, devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (no sensing).
+In each time slot, the devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (\ie, no sensing).
 Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station.
-If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part.
+If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part (on the same channel).
 If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \emph{Ack} is not transmitted.
@@
     \caption{The considered time-frequency slotted protocol.
-Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets.
-If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
+Each frame is composed by a fix-duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
+If a packet is well received, the base station replies by transmitting an \textcolor{darkgreen}{\emph{Ack}}, after the ack delay.}
     \label{fig:41:protocol}
@@
 Furthermore, in channel $i \in \llbracket 1; K \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{K} S_i$).
-We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (about $1000$ to $10000$, while $K$ is about $10$ to $50$).
+We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (about $1000$ to $10000$, while $K$ is about $4$ to $50$).
 As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
@@
 
-We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
+We consider the network from the point of view of one dynamic device. Every times the dynamic device has to communicate with the gateway,
 it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
-Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
-Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
+Then, the dynamic device starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
+Before sending another message (\emph{i.e.}, at time $t+1$), the dynamic device knows if it received or not this \emph{ACK} message.
 %
 For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
-The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
+The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
 $r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
@@
 
-This problem is a special case of the so-called stochastic MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$.
-Several types of reward distributions have been considered in the literature, and rewards are binary in our model.
+This problem is a special case of the stochastic MAB (see Chapter~\ref{chapter:2}).
+% , where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$.
+% Several types of reward distributions have been considered in the literature, and rewards are binary in our model.
 We consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
@@
 
-The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to the end-users, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
+The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to each dynamic device, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
 %
-This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
-
-The goal is thus to design a simple sequential algorithm, to be applied identically by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
+This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a device has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
+%
+The goal is thus to design a new or chose an existing simple sequential algorithm, to be applied identically by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
 
@@
 
-Considered alone, each dynamic device implements a learning algorithm to play a bandit game, the device is consequently a smart device.
+Considered alone, each dynamic device implements a learning algorithm to play a bandit game, the device is consequently a dynamic device.
 In each time slot, if it has to communicate (which happens with probability $p$), then it chooses a channel and it receives a reward $1$ if the transmission is successful, $0$ otherwise.
@@
 Besides the modified time scale (rewards are no longer collected at every time step), this looks like a bandit problem.
-However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other smart devices, that is not stationary (because of learning).
+However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning).
 
@@
 %
-In our model, every dynamic end-user implements its own \UCB{} algorithm, \emph{independently}.
-For one end-user, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
+In our model, every dynamic dynamic device implements its own \UCB{} algorithm, \emph{independently}.
+For one dynamic device, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
 
@@
 
-In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices: they are \iid{} with mean $\mu_i = 1 - p S_i$.
-Moreover, dynamic devices are assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
-The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
-However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.
+The main difference between the two models is the fact that in this Chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p$).
+Active devices denote static and dynamic devices that have to transmit at a fixed time,
+and their random activation patter makes the number of active devices a random variable.
+Analyzing multi-player MAB models under this hypothesis is much harder, and is left as a future work.
+
+% In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices: they are \iid{} with mean $\mu_i = 1 - p S_i$.
+% Moreover, dynamic devices are usually assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
+% The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
+% However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.
 
@@
 The \emph{adversarial MAB} algorithms is a broader family, and a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Bubeck12}.
-Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform worse than both \UCB{} and TS in the same experiments.
+Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform much worse than both \UCB{} and TS in the same experiments.
 %(see Section \ref{sub:41:numericalResults}).
-Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary and non-\iid, as its performance guarantee are true in \emph{any} setting.
+Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary model, as its performance guarantee are true in \emph{any} setting.
 But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
@@
 % ------------------------------------------------------------------------
-\subsection{Experiments and numerical results}\label{sub:41:numericalResults}
+\subsection{Numerical results}\label{sub:41:numericalResults}
 
@@
     \centering
-    \subfloat[10\% of smart devices]{\includegraphics[scale=0.5]{10intelligent.eps}
+    \subfloat[10\% of dynamic devices]{\includegraphics[scale=0.5]{10intelligent.eps}
     \label{fig:41:10intelligent}}
     \hfill
-    \subfloat[30\% of smart devices]{\includegraphics[scale=0.5]{30intelligent.eps}
+    \subfloat[30\% of dynamic devices]{\includegraphics[scale=0.5]{30intelligent.eps}
     \label{fig:41:30intelligent}}
@@
     \vspace*{-20pt}
-    \subfloat[50\% of smart devices]{\includegraphics[scale=0.5]{50intelligent.eps}
+    \subfloat[50\% of dynamic devices]{\includegraphics[scale=0.5]{50intelligent.eps}
     \label{fig:41:50intelligent}}
     \hfill
-    \subfloat[100\% of smart devices]{\includegraphics[scale=0.5]{100intelligent.eps}
+    \subfloat[100\% of dynamic devices]{\includegraphics[scale=0.5]{100intelligent.eps}
     \label{fig:41:100intelligent}}
-    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of smart end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
+    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
     \label{fig:41:from10to100}
@@
 \end{figure}
+
+\paragraph{First results}
 
@@
 
-We can see in Figure \ref{fig:41:from10to100} that the TS algorithm (in red) outperforms the \UCB{} algorithm (in blue), when the number of end-devices is below 50\%.
+We can see in Figure \ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%.
 When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after very few transmissions (quick convergence).
@@
 For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
-Increasing the number of end-devices decreases the gap between the optimal and random policies: the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures \ref{fig:41:100intelligent} and \ref{fig:41:perf_learning}.
+Increasing the number of end-devices decreases the gap between the optimal and random policies: the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and \ref{fig:41:perf_learning}.
 
-To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure \ref{fig:41:perf_learning} the evolution of the gain, in term of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each object.
+\paragraph{Successful transmissions rate as a function of the dynamic devices}
+
+To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure~\ref{fig:41:perf_learning} the evolution of the gain, in term of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each object.
 We can see that when the proportion of end-devices is low (\emph{e.g.}, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
@@
     \includegraphics[scale=0.60]{perf_learning.eps}
-    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices. Both algorithms achieve close to optimal performance after a reasonable learning time.}
+    \caption{Learning with \UCB{} and Thomson Sampling, with many dynamic devices. Both algorithms achieve close to optimal performance after a reasonable learning time.}
     \label{fig:41:perf_learning}
@@
 \href{https://Bitbucket.org/scee_ietr/rl_slotted_iot_networks}{\texttt{Bitbucket.org/scee\_ietr/rl\_slotted\_iot\_networks}}.
-It was written in collaboration with Rémi Bonnefoi, in May $2017$.
+It was written in collaboration with Rémi Bonnefoi, in April and May $2017$.

### 2019-02-26 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Small modifications today. Hard to get anything important done, I had a review to do for a COLT paper...
Proofreading of Chapter 4. No new science since the papers... but it's cleanly presented and unified.

@@
     \centering
-    \includegraphics[width=0.65\linewidth]{SCEE_TestBed1.jpg}
+    \includegraphics[width=0.75\linewidth]{SCEE_TestBed1.jpg}
     \vspace*{20pt}
-    \includegraphics[width=0.65\linewidth]{SCEE_TestBed2.jpg}
+    \includegraphics[width=0.75\linewidth]{SCEE_TestBed2.jpg}
     \caption{Two pictures showing the SCEE testbed \cite[Appendix~3]{Bodinier17}.}
@@
     \centering
-    \includegraphics[width=0.80\textwidth]{UI.eps}
+    \includegraphics[width=0.85\textwidth]{UI.eps}
     \caption{User interface of our demonstration.}
@@
 
-% \begin{enumerate}[leftmargin=6mm]
-    % \item[(1)]
-$(1)$ The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.
+\begin{enumerate}[leftmargin=6mm]
+    \item[(1)]
+The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.
 
-    % \item[(2)]
-$(2)$ The second part is the interface of the intelligent device which is made of four parts.
+    \item[(2)]
+The second part is the interface of the intelligent device which is made of four parts.
 At the top left, we observe the constellation of the transmitted packet \emph{(a)}.
@@
 
-    % \item[(3)]
-$(3)$ The last part is the interface of the gateway, where we can see the traffic observed by the gateway \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
-% \end{enumerate}
+    \item[(3)]
+The last part is the interface of the gateway, where we can see the traffic observed by the gateway \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
+\end{enumerate}
 
 % ----------------------------------------------------------------------
-\subsection{Numerical results}
+\subsection{Empirical results}
 \label{sub:42:results}

### 2019-02-26 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for now, I don't like the content of Chapter 3 at all for now, but at least there is something.
Proofreading of Chapter 4. No new science since the papers... but it's cleanly presented and unified.

@@
-% https://perso.crans.org/besson/articles/BMBM__IEEE_WCNC_2019.pdf
+% - ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see  https://hal.inria.fr/hal-02049824
 
@@
 
-% \TODOL{This Section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://perso.crans.org/besson/articles/BMBM\_\_IEEE\_WCNC\_2019.pdf}}
+% \TODOL{This Section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://hal.inria.fr/hal-02049824}}
 
-\TODOL{TODO:
+% \TODOL{TODO:
 
-- Be sure to not repeat any motivations / maths / explanations given before
-}
+% - Include the modifications suggested by the reviewers of our paper!
+% - Be sure to not repeat any motivations / maths / explanations given before
+% }
+
+% https://edas.info/showPaper.php?m=1570524364
+% https://hal.inria.fr/hal-02049824
 
@@
 	\item
-	Secondly, several heuristics are proposed to cope with retransmissions.
+	Secondly, different heuristics are proposed to cope with retransmissions.
 
@@
 The proposed \UCB-based heuristics are presented in Section~\ref{sub:43:heuristics}, while the corresponding numerical results are shown in Section~\ref{sub:43:numExp}.
-Finally, some conclusions are drawn in Section~\ref{sub:43:conclusion}.
+Finally, some conclusions are drawn later in Section~\ref{sub:43:conclusion}.
 
@@
 We assume that this network is constituted by two types of devices:
-on the one hand, we have \emph{static} devices that operate in one channel\footnote{~Note that, for unlicensed bands, this definition also encompasses any device following a different standard or trying to establish communication with gateways of other networks.} in order to communicate with the gateway.
+on one hand, we have \emph{static} devices that operate in one channel\footnote{~Note that, for unlicensed bands, this definition also encompasses any device following a different standard or trying to establish communication with gateways of other networks.} in order to communicate with the gateway.
 %
@@
 
-Regardless the type of devices, each of them follows a slotted ALOHA protocol \cite{Roberts75}, and has a probability $p>0$ to transmit a packet in a time slot.
+Like in the previous model presented in Section~\ref{sec:4:firstModel},
+regardless the type of devices, each of them follows a slotted ALOHA protocol \cite{Roberts75}, and has a probability $p>0$ to transmit a packet in a time slot.
 We make the hypothesis that the transmission is successful if the channel is available, otherwise upon radio collision, these devices will attempt to retransmit their packet up-to $M$ times, with $M \in\mathbb{N}$.
@@
 The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Fig.~\ref{fig:43:Markov_model}.
-A device containing a packet for transmission goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, \ie, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
+When a device have a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, \ie, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
 At each time slot, a transition from an idle state to a transmission state (denoted as \texttt{Trans.}) occurs if a packet transmission is required, while waiting states (denoted as \texttt{Wait}), correspond to a $m$ back-off interval.
 
-A device aims to select a channel with the highest probability of a successful transmission, for which it resorts to a reinforcement learning approach.
-It is formulated as MAB problem, where each channel (also called arms) is viewed as a gambling machine (bandit), and each bandit has a \emph{reward}.
+A device aims to select a channel with the highest probability of a successful transmission, for which it resorts to a reinforcement learning approach, formulated as a MAB problem.
+% , where each channel (also called arms) is viewed as a gambling machine (bandit), and each bandit has a \emph{reward}.
 Then, at every trial, a device chooses a channel that maximizes the sum of the collected rewards.
@@
 In this way, a transmission is considered successful when an acknowledgment is received, and a learning approach is employed to select the best channel.
-
+%
 We address the problem of channel selection taking into account the described Markov model for the retransmissions of end-devices.
@@
 \begin{equation}\label{eq:43:1}
-	p_c = 1-\left(1-x_t\right)^{N-1} \iff x_t = 1-\left(1-p_c\right)^{\frac{1}{N-1}}.
+	p_c = 1-\left(1-x\right)^{N-1} \iff x = 1-\left(1-p_c\right)^{\frac{1}{N-1}}.
 \end{equation}
@@
 
-\TODOL{Remove this part or compress}
+\TODOL{Remove this part or compress: I will remove as soon as I have clear equations in Chapter~\ref{chapter:2} as references for the \UCB{} algorithm.}
 
@@
 			}
-			\Else(\tcp*[f]{Packet retransmission}){
+			\Else(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^r$}){
 				Compute for each channel $U^r_k(t) = \widehat{\mu^r_k}(t) + B^r_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
@@
 
-Although, this approach increases the complexity and store requirements (of order $\mathcal{O}(K^2)$).
+Although, this approach increases the complexity and storage requirements (of order $\mathcal{O}(K^2)$).
 For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
 For instance, for $K=4$ channels,
-the memory to store $K+1=5$ algorithms is of the order of the requirements to storing one.
+the memory to store $K+1=5$ algorithms is of the order of the requirements to store one.
 
@@
 				}
-				\Else(\tcp*[f]{Packet retransmission}){ % Retr. after trying channel $j$
+				\Else(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^j$}){ % Retr. after trying channel $j$
 					j $\leftarrow$ last channel selected by first-stage \UCB\;
@@
 				}
-				\uElseIf(\tcp*[f]{Random retransmission}){$t \leq \Delta$}{
+				\uElseIf(\tcp*[f]{Random retransmission with $\mathrm{UCB}^d$}){$t \leq \Delta$}{
 					Transmit randomly in a channel $C(t) \sim \mathcal{U}(1,\ldots,K)$.
@@
 % ----------------------------------------------------------------------
-\subsection{Simulations to compare our heuristics}
+\subsection{Numerical results}
 \label{sub:43:numExp}
@@
 %
-For the evaluation of the proposed heuristics, a total number of $T=10^{6}$ time slots is considered, and the results are averaged over $10^{3}$ independent random simulations.
+For the evaluation of the proposed heuristics, a total number of $T=20 \times 10^4$ time slots is considered, and the results are averaged over $10^{3}$ independent random simulations.
 %In this way, it allows us to evaluate the learning rate, as well as the convergence of each learning approach.
@@
 learning approach is able to take advantage of it.
-Furthermore, the ``Delayed \UCB'' and  the``\UCB'' heuristics show quite similar results, after converging.
-%
-%Random, \UCB{}, $K$ \UCB{}, Delayed \UCB{} obtain similar results.
-The less efficient one is the $K$ \UCB{}, and this makes sense as having more learning algorithm needs more for each of them to learn.
-The most efficient one is the simple \UCB{} procedure, and this was quite surprising.
+Furthermore, the ``\UCB'', ``$K$ \UCB'' and ``Delayed \UCB'' heuristics behave similarly than ``Only \UCB'', after a similar convergence time.
 

### 2019-02-26 2-Chapters/4-Chapter/chapter4.tex
Proofreading of Chapter 4. No new science since the papers... but it's cleanly presented and unified.

@@
 The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
-We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP \cite{USRPDocumentation} cards.
+We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP cards \cite{USRPDocumentation}.
 
-More precisely, on the one hand, we start by presenting in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, used by static devices.
+More precisely, on one hand, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
+We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, used by static devices.
 Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their communication messages were successfully received.
@@
 On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending the first one, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
-In this second model, we compare different heuristic, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
+In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
+Interestingly, the simpler heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
+
+\paragraph{Publications}
+
+This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 
@@
 %
-In this Section, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
+In this Chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
 We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
-Kaufmann12Thompson}.
+Kaufmann12Thompson}, for the first model.
+For the demonstration and the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for to solve the problem of learning to retransmit efficiently.
 
+As detailed in Chapter~\ref{chapter:1},
 MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
-For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}.
+% For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16}.
 Many recent works show that MAB algorithms work well for real-world radio signal.
@@
 We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
-Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \iid.
+Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is not stationary.
 As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
@@
 We conclude this Chapter by presenting an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
-
-\paragraph{Publications}
-
-This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 

### 2019-02-27 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Finishing Chapter:4

@@
 
-We consider two hypotheses $\cH_{1}$ and $\cH_{2}$ defined as,
+We consider two hypotheses $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as,
 \begin{itemize}
-	\item $\cH_{1}$:
-    The probability $p_{c1}$ is composed by the sum of two probabilities: i)
+	\item $\mathcal{H}_{1}$:
+    The probability $p_{c1}$, is composed by the sum of two probabilities: i)
     the probability of colliding consecutively twice, \ie, the devices that collide at a given time slot and collide again when retransmitting their packets,
-    and ii) the probability of collision related with devices that attempt to transmit in this channel for first time.
-    In addition, we suppose that the number of devices that attempt to retransmit is small in comparison to the total number of devices sharing the same channel.
-
-	\item $\cH_{2}$:
+    and ii) the probability of collision among devices that did not collide in the same previous collision.
+Moreover, we suppose that the number of devices involved in a collision is small in comparison to the total number of devices.
+	\item $\mathcal{H}_{2}$:
 	The total number of the back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
@@
 We consider $N$ active devices following the same policy.
+
+%{\color{red}Considering one device and a channel,
+%we denote $x_t^i$ the packet transmission probability for the $i+1$ time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$).}
+%
+%{\color{red}Considering one device and a channel, we denote $x_t^i$ the transmission probability for the $i+1$ time, with $i\in \llbracket 0, M-1 \rrbracket$, in a given time slot $t$}
 
@@
 
-As explained above, if an IoT device is experiencing a collision at the first transmission, it proceeds for the retransmission of its packet after a random back-off interval.
+As explained above, if an IoT device experiences a collision at the first transmission, it proceeds for the retransmission of its packet after a random back-off interval.
 We denote $p_{ca}$ the probability to have a collision with a packet involved in the previous collision.
-Under assumption $\cH_{1}$, the probability that the same device's packet is experiencing again a collision at the second time slot is
+Under the $\mathcal{H}_{1}$ assumption, the number of packets involved in the same previous collision remains very small in comparison to the total number of devices that may transmit during this time.
+In other words, this collision probability does not depend on previous retransmissions and is equal to $p_c$.
+So, the probability that the same device's packet experiences again a collision at the second time slot is
+%
 \begin{equation}\label{eq:43:decomppc1}
@@
 
-If the device had a collision, we consider $p_{bp}(n)$ the probability that it had a collision with \emph{exactly} $n$ packets (for any $1\leq n \leq N-1$), and that \emph{at least one} of the $n$ devices involved in the previous collision chose the same back-off interval.
-Hence, under hypothesis $\cH_{2}$, we can relate $p_{ca}$ with $p_{bp}(n)$, and the different probabilities that a device experienced a collision during the first slot and has the same back-off interval for its retransmission is,
+If the device has a collision at the first attempt, we consider $p_{bp}(n)$ the probability that it has a collision with \emph{exactly} $n$ packets (for any $1\leq n \leq N-1$), and that \emph{at least one} of the $n$ devices involved in this first collision chooses the same back-off interval,
+%
+\begin{equation}
+    p_{bp}(n) = {N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left[1-\left( 1-\frac{1}{m}\right)^n \right].
+\end{equation}
+
+Besides, $p_{ca}$ is the conditional probability of collision with a packet sent by a device involved in the previous collision given that the packet experienced collision at its first transmission.
+Hence, under hypothesis $\mathcal{H}_{2}$, we can use Bayes theorem and the law of total probability to relate $p_{ca}$ with $p_{bp}(n)$, and the different probabilities that a device experienced a collision during the first slot and has the same back-off interval for its retransmission is,
+%{\color{red}Hence, under hypothesis $\mathcal{H}_{2}$,
+%the collision probability with the same packets after a retransmission, $p_{ca}$, is given  the sum of collision probabilities with $n$ devices choose the same back-off interval for their retransmissions, }
 %
 \begin{equation}\label{eq:43:sumpca}
-	p_{ca} = \sum_{n=1}^{N-1} p_{bp}(n).
+	p_{ca} = \frac{1}{p_c}\sum_{n=1}^{N-1} p_{bp}(n).
 \end{equation}
 
+% \begin{figure}[htp!]  % [htbp]
+% 	\centering
+% 	% \includegraphics[width=1.00\linewidth]{Approximation_m10.eps}
+% 	\includegraphics[width=1.00\linewidth]{Approximation_m10-eps-converted-to.pdf}
+% 	\caption{Our proposed approximation for the probability of collision at the second transmission. It is more precise for smaller values of $N$.}
+% 	\label{fig:43:Approximation_m10}
+% \end{figure}
+
 Therefore, the expression of $p_{ca}$ is
-\begin{equation}\label{eq:43:sumpc2}
-	\frac{1}{p_c} \sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left[1-\left( 1-\frac{1}{m}\right)^n \right].
+\begin{align}\label{eq:43:sumpc2}
+	& \frac{1}{p_c} \sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left[1-\left( 1-\frac{1}{m}\right)^n \right]\nonumber \\
+	& = 1- \frac{1}{p_c}\sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left( 1-\frac{1}{m}\right)^n.
+\end{align}
+
+Once again under $\mathcal{H}_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first $N_0 \ll N-1$ terms of the sum in \eqref{eq:43:sumpc2} are predominant.
+We derive,
+%
+\begin{equation}\label{eq:43:sumpca2}
+	p_{ca} \simeq  1- \frac{1}{p_c}\sum_{n=1}^{N_0}{N-1 \choose n} x^n\left(1-x\right)^{N-1-n} \left( 1-\frac{1}{m}\right)^n.
 \end{equation}
 
-Once again under $\cH_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first terms of the sum in \eqref{eq:43:sumpc2} are predominant.
-Moreover, for these terms, $n$ is small compared to $N-1$, and so $N-1-n$ can be approximated to $N-1$.
-Thus it gives,
-\begin{align}\label{eq:43:sumpca2}
-	p_{ca} = & 1- \frac{1}{p_c}\sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left( 1-\frac{1}{m}\right)^n,\nonumber\\
-	\simeq & 1- \frac{\left(1-x\right)^{N-1}}{p_c}\sum_{n=1}^{N-1}{N-1 \choose k} x^n \left( 1-\frac{1}{m}\right)^n.
-\end{align}
+Moreover, for these terms, $n$ is small compared to $N-1$, and so $N-1-n$ can be approximated to $N-1$. Thus it gives,
+%
+\begin{equation}\label{eq:43:sumpca3}
+	p_{ca} \simeq  1- \frac{\left(1-x\right)^{N-1}}{p_c}\sum_{n=1}^{N_0}{N-1 \choose n} x^n \left( 1-\frac{1}{m}\right)^n.
+\end{equation}
 
-We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca2}, and so the expression of $p_{ca}$ becomes
-\begin{align}\label{eq:43:pca}
-	p_{ca} \simeq \frac{1}{p_c} - \left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1}.
-\end{align}
+Assuming $\mathcal{H}_{1}$ amounts to consider that $x\ll 1$. As a consequence, the sum in equation \eqref{eq:43:sumpca3} can be supplemented by negligible terms,
+
+\begin{equation}\label{eq:43:sumpca4}
+	p_{ca} \simeq  1 - \frac{\left(1-x\right)^{N-1}}{p_c}\sum_{n=1}^{N-1}{N-1 \choose n} x^n \left( 1-\frac{1}{m}\right)^n.
+\end{equation}
+
+We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca4}, and we rewrite the expression of $p_{ca}$ as
+\begin{multline}\label{eq:43:pca}
+	p_{ca} \simeq 1 -\\
+	\left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1}.
+\end{multline}
 

### 2019-02-27 2-Chapters/4-Chapter/chapter4.tex
Finishing Chapter:4

@@
 % ----------------------------------------------------------------------
-\chapter{More Efficient Spectrum Usage of IoT Networks with Selfish MAB Learning}
+\chapter{Improving Spectrum Usage of IoT Networks with Selfish MAB Learning}
 \label{chapter:4}
@@
 % ----------------------------------------------------------------------
+
+\paragraph{Abstract}
 
@@
 More precisely, on one hand, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
-We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, used by static devices.
-Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their communication messages were successfully received.
+We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
+Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
 We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time.
 If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
-we show in simulations that in the case of non-uniform repartition of static devices in the channels, the dynamic devices can improve a lot the network efficiency.
+we show in simulations that in the case of a non-uniform repartition of the static devices in the channels, the dynamic devices can drastically improve the network efficiency.
 %
@@
 
-On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending the first one, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
+On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending our first model, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
 In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
@@
 
+\vfill{}
+
 \paragraph{Publications}
 
-This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
+This Chapter is mainly based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 
@@
 
-Unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
+We already explained in Chapter~\ref{chapter:1} that
+unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
 This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.

### 2019-02-27 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Proofreading of Chapter 4. No new science since the papers... but it's cleanly presented and unified.
Finishing Chapter:4

@@
 
-% For \rhoRand, \cite{Anandkumar11} essentially proved on the one hand that conditionnaly to a certain ``good'' event
+% For \rhoRand, \cite{Anandkumar11} essentially proved on one hand that conditionnaly to a certain ``good'' event
 % (all players have a correct ordering of all the $M$ best arms)
@@
 this guarantee is interesting for Cognitive Radio applications,
-where switching arms means reconfiguring a radio hardware, an operation that costs energy.
+where switching arms means re-configuring a radio hardware, an operation that costs energy.
 %and thus an algorithm guaranting a small number of switches is interesting.

### 2019-02-27 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Finishing Chapter:4

@@
 %
-We evaluate the performance of two classical MAB learning algorithms, \UCB{} and Thomson Sampling, to handle the decentralized decision-making of Spectrum Access, applied to IoT networks; as well as learning performance with a growing number of intelligent end-devices.
+We evaluate the performance of two classical MAB learning algorithms, \UCB{} and Thomson Sampling, to handle the decentralized decision-making of Spectrum Access, applied to IoT networks.
+We also evaluate the learning performance when the number of intelligent end-devices grows.
 %
 We show that using learning algorithms does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically changing channel.
-In the studied scenario, stochastic MAB learning provides a up to $16\%$ gain in term of successful transmission probabilities, and has near optimal performance even in non-stationary and non-\iid{} settings with a majority of intelligent devices.
+In the studied scenario, stochastic MAB learning provides an up to $16\%$ gain in term of successful transmission probabilities, and has near optimal performance even in non-stationary settings with a majority of intelligent devices.
 
@@
 
-The aim of this Section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases, and the stochastic hypothesis is more and more questionable.
+The aim of this Section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases, and the stochastic hypothesis is more and more violated.
 To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
 Static devices form an interfering traffic, which could have been generated by devices using other standards as well.
-We first evaluate the probability of collision if dynamic devices randomly select channels (naive approach), and if a centralized controller optimally distribute them in channels (ideal approach).
-Then, these reference scenarios allow to evaluate the performance of \UCB{} and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
+We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller optimally distribute them in the channels at the beginning of the scenario (that is, an ideal approach).
+Then, these three reference scenarios allow to evaluate the performance of \UCB{} and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
 We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.
@@
 \paragraph{Outline}
+%
 The rest of this Section is organized as follows.
 The system model is introduced in Section~\ref{sub:41:systemModel}.
-Reference policies are described in \ref{sub:41:threeReferencePolicies}, and sequential learning policies are introduced in \ref{sub:41:sequentialPolicies}.
+The reference policies are described in \ref{sub:41:threeReferencePolicies}, and our proposed sequential learning policies are introduced in \ref{sub:41:sequentialPolicies}.
 Numerical results are presented in \ref{sub:41:numericalResults}.
@@
 
-We present our system model, which consists of one gateway and many IoT devices, using a slotted protocol.
+We present our system model, which consists of one gateway and many IoT devices, using a frequency- and time- slotted protocol.
 
@@
 %
-We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of object sends uplink packets to the network gateway.
-Messages can be sent for instance in the $433.5\;\mathrm{MHz}$ ISM band (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
+We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of objects all sends uplink packets to the (unique) network gateway.
+Messages can be sent for instance centered around the $433.5\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
 The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
 %
-The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$).
+The devices can transmit their packets in $K\geq2$ channels (\eg, $K=4$).
 In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1\;\mathrm{s}$ in our implementation).
 
-These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks.
-This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.
+These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated neighboring networks.
+This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels.
 %
-To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
+To simulate networks designed for the Internet of Things (IoT), we consider a protocol \textbf{with no sensing}, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet (but only to the objects being paired with this base station in this network).
 %
-By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
-the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).
+By considering a small number of wireless channels (\eg, $K = 4$) and a PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
+the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\ie, the most vacant one).
 
@@
 %
-For simplicity, and for consistency with the rest of this thesis, we only consider a discrete time protocol.
+For simplicity, and for consistency with the rest of this thesis, we only consider a discrete protocol (\ie, slotted).
 As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
@@
 Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station.
-If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part (on the same channel).
-If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \emph{Ack} is not transmitted.
+In one channel, if only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part (on the same channel).
+If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \Ack{} is not transmitted.
+In other words, if the gateway cannot successfully decode the incoming (uplink) messages because they were corrupted by collisions, it does not send any \Ack{} back.
 This way, no collision can occur on the downlink messages, easing the analysis of collisions.
@@
 Each frame is composed by a fix-duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
-If a packet is well received, the base station replies by transmitting an \textcolor{darkgreen}{\emph{Ack}}, after the ack delay.}
+If a packet is well received, the base station replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
     \label{fig:41:protocol}
@@
     \emph{Static} end-devices have poor RF abilities, and each of them uses only one channel to communicate with the base station.
-Their choice is assumed to be fixed in time (stationary) and independent (\iid).
+Their choice is assumed to be fixed in time (\ie, stationary).
 The traffic generated by these devices is considered as an interfering traffic for other devices.
     \item
-    \emph{Dynamic} (or \emph{smart}) end-devices have richer RF abilities, they can use all the available channels, by quickly reconfiguring their RF transceiver on the fly.
+    \emph{Dynamic} (or \emph{smart}) end-devices have richer RF abilities, they can use all the available channels, by quickly re-configuring their RF transceiver on the fly.
 They can also store communication successes or failures they experienced in each channel, in order to change channel, possibly at every time slot.
+    We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
 \end{itemize}
 
-There are $K \geq 1$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
+There are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
 Furthermore, in channel $i \in \llbracket 1; K \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{K} S_i$).
-We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (about $1000$ to $10000$, while $K$ is about $4$ to $50$).
-As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
-We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks\footnote{~Here, $K$ will typically be about $20$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
 
@@
 
+\paragraph{Dense networks and an hypothesis}
+%
+We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (for instance, about $1000$ to $100000$, while $K$ is about $4$ to $256$).
+As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
+We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks\footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
+Indeed, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ objects, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall i, S_i > 0$).
+Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero than using learning algorithms cannot improve much.
+Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$ to have an average of active objects not larger than the number of channels.
+
 \paragraph{MAB model}
-
-We consider the network from the point of view of one dynamic device. Every times the dynamic device has to communicate with the gateway,
+%
+We consider the network from the point of view of one dynamic device.
+Following its Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
 it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
 Then, the dynamic device starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
-Before sending another message (\emph{i.e.}, at time $t+1$), the dynamic device knows if it received or not this \emph{ACK} message.
+Before sending another message (\ie, at time $t+1$), the dynamic device knows if it received or not this \Ack{} message.
 %
-For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
-The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
+For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
+The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, it is to maximize its number of successful transmissions, or its cumulative reward,
 $r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
@@
 We consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
-Contrary to many previous work done in the CR field (\emph{e.g.}, Opportunistic Spectrum Access),
+Contrary to many previous work done in the CR field (\eg, Opportunistic Spectrum Access),
 the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
-Rewards come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
+In our model, rewards rather come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
 
-The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to each dynamic device, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
+The problem parameters are $\mu_1,\dots,\mu_K$, the mean availability of the channels, and they are of course unknown to each dynamic device, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively improve its successful communication rate.
 %
@@
 %
-The goal is thus to design a new or chose an existing simple sequential algorithm, to be applied identically by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
+The goal is thus to design a new or chose an existing simple sequential algorithm, to be applied identically and independently by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
 
-Before explaining how this goal presents similarity with a \emph{multi-armed bandit problem}, we present some natural baseline policies (\ie, algorithms).
+Before modeling this goal with a \emph{multi-armed bandit problem}, we present some natural baseline (reference) policies.
 
@@
 
-We present three different policies that will be used to assess the efficiency of the learning algorithms presented in the next Subsection.
+We present three different policies that can be used to assess the efficiency of the learning algorithms presented later on.
 The first one is naive but can be used in practice, while the two others are very efficient but require full knowledge on the system (\ie, an oracle) and are thus unpractical.
+%
+They are however useful for our numerical simulations, to show that our MAB-based approaches quickly learn to perform almost optimally.
 
@@
 
-We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; K \rrbracket = \{1, \dots, K\}$).
-This reflects a very naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.
+We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; K \rrbracket$).
+This reflects a naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.
 
@@
 \begin{equation}
-    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/K, \text{by uniform choice}}
+    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/K, \text{ by uniform choice}}
 \end{equation}
@@
 \end{align}
-This expression \eqref{eq:41:strategynaive} is constant (in time), and easy to compute numerically, but comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it.
+This expression \eqref{eq:41:strategynaive} is constant (in time), and easy to compute numerically, but comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it
+(maybe after a long enough initial learning period).
 
@@
 
-We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of everything, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, in order to avoid any signaling overhead.} is possible.
+We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the repartition of static devices (\ie, $(S_i)_i$ or equivalently, $(\mu_i)_i$), and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.} is possible.
 We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.
 
-If the oracle draws once uniformly at random a configuration of dynamic devices, with $D_i$ devices affected to channel $i$ is fixed (in time, \ie, stationary),
+If the oracle chooses a fixed configuration of dynamic devices, with $D_i$ devices affected to channel $i$ is fixed (in time, \ie, stationary),
 then this probability is computed as before:
@@
 
-Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem:
+Consequently, an optimal allocation vector $(D_1,\dots,D_{K})$ is a solution of the following real-valued constraint optimization problem:
 %
@@
 
-    It gives a closed form expression for the optimal solution $D_i^*(\lambda)$, depending on the system parameters, and the unknown Lagrange multiplier $\lambda \in \mathbb{R}$.
+    It gives a closed form expression for the (unique) optimal solution $D_i^*(\lambda)$, depending on the system parameters, and the unknown Lagrange multiplier $\lambda \in \mathbb{R}$.
 
     \begin{equation}\label{eq:41:Dilambda}
-        D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{\dag} .
+        D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{+} .
     \end{equation}
@@
 
-Where $(a)^{\dag} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto xe^x$ on $\mathbb{R^+} = [0, +\infty)$ \cite{Corless96}.
+Where $(a)^{+} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ \cite{Corless96}.
 Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint
@@
 
-While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused by the learning policy used by dynamic objects.
+While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused partly by the learning policy used by dynamic objects, but mainly by their random activation processes.
 
 In our model, every dynamic device implements its own learning algorithm, \emph{independently}.
-For one device, the time $t$ is the number of time it accessed the network (following its Bernoulli transmission process, \ie, its duty cycle), \emph{not} the total number of time slots from the beginning, as rewards are only obtained after a transmission, and IoT objects only transmit sporadically, due to low transmission duty cycles.
+For one device, the time $t$ refers to the number of time it accessed the network (following its Bernoulli transmission process, \ie, its duty cycle), \emph{not} the total number of time slots from the beginning, as rewards are only obtained after a transmission, and IoT objects only transmit sporadically, due to low transmission duty cycles.
 
@@
 
-Considered alone, each dynamic device implements a learning algorithm to play a bandit game, the device is consequently a dynamic device.
+Considered alone, each dynamic device implements a learning algorithm to play a bandit game.
 In each time slot, if it has to communicate (which happens with probability $p$), then it chooses a channel and it receives a reward $1$ if the transmission is successful, $0$ otherwise.
 Each device aims at maximizing the sum of the rewards collected during its communication instants, which shall indeed maximize the fraction of successful transmissions.
-Besides the modified time scale (rewards are no longer collected at every time step), this looks like a bandit problem.
-However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning).
+Besides the modified time scale (rewards are no longer collected at every time step), this looks like a usual bandit problem.
+However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning and random activation of each devices).
 
@@
 
+The two algorithms, \UCB{} and Thompson sampling (TS), were presented in Section~\ref{sec:2:famousMABalgorithms}.
 The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
@@
 For one dynamic device, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
-
+%
 The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
@@
 
-The main difference between the two models is the fact that in this Chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p$).
+The main difference between the two models is the fact that in this Chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
 Active devices denote static and dynamic devices that have to transmit at a fixed time,
@@
 % (in their respective case of application).
-More is left on this aspect for our future work.
+More is left on this aspect for a future work.
 
 % ------------------------------------------------------------------------
-\subsection{Numerical results}\label{sub:41:numericalResults}
+\subsection{Numerical results on an example of IoT network}\label{sub:41:numericalResults}
 
@@
 We suppose a network with $S + D = 2000$ end-devices, and one IoT base station.
-Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\emph{e.g.}, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
+Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\eg, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
 The RF band is divided in $K = 10$ channels.
-Each static device only uses one channel, and their uneven repartition in the $10$ channels is: $(S_1,\cdots, S_{K}) = S\times(0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
+Each static device only uses one channel, and their uneven repartition in the $10$ channels is chosen as $(S_1,\cdots, S_{K}) = S \times (0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
 The dynamic devices have access to all the channels, and use learning algorithms.
@@
 We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions.
-Clearly, the problem is less interesting for homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
+Clearly, the problem is less interesting for a homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
 
@@
     \label{fig:41:100intelligent}}
-    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
+    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:from10to100}
@@
 The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy from below, and the two oracle policies (optimal and greedy) from above.
-The results are displayed when $10$, $30$, $50$ and $100\%$ of the traffic is generated by dynamic devices.
+The results are displayed when $10\%$, $30\%$, $50\%$ and $100\%$ of the traffic is generated by dynamic devices.
 
 We can see in Figure \ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%.
-When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after very few transmissions (quick convergence).
-Moreover, we can see in Figures \ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the oracle optimal and oracle suboptimal policies.
+When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
+Moreover, we can see in Figures \ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
 For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
-Increasing the number of end-devices decreases the gap between the optimal and random policies: the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and \ref{fig:41:perf_learning}.
+Increasing the number of end-devices decreases the gap between the optimal and random policies: the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of \ref{fig:41:perf_learning}.
 
@@
 To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure~\ref{fig:41:perf_learning} the evolution of the gain, in term of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each object.
-We can see that when the proportion of end-devices is low (\emph{e.g.}, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
+We can see that when the proportion of end-devices is low (\eg, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
 The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a lowest rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$).
@@
 Note that when this proportion increases, the assumptions of a stochastic MAB model are clearly violated, and there is no justification for the efficiency of TS and \UCB{} algorithms.
-Hence, it is surprising to have near optimal performance with stochastic MAB algorithms applied to partly dynamic and fully dynamic scenarios.
+Hence, it is surprising to have near optimal performance with stochastic MAB algorithms applied to partly or fully dynamic scenarios.
 
@@
 
-We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static device, to check that learning does not help in this case, but does not decrease performance either.
+We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall i, S_i = S/K$), to check that learning approach (here, only \UCB)
+also gives interesting gain of performance, and achieve a close-to optimal successful transmission rate.
+Except for the limit case of $100\%$ of dynamic devices (where the uniform access performs as well as the optimal oracle solution), the MAB-based approach almost instantly outperforms the baseline.
 

### 2019-02-27 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Finishing Chapter:4

@@
 %
-This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature in some or all the devices, without any modification on the network side.
+This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature at the higher level of the MAC layer, in all or only some of the devices, without any modification on the network side.
 
@@
 
-In our demo, we consider a simple wireless network, consisting of one gateway (radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
+In our demo, we consider a simple wireless network, consisting of one gateway (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
 Some dynamic intelligent objects (end-user or autonomous objects) try to communicate with the gateway, with a low-overhead protocol.
 This communication can be done in different channels which are also shared by devices using other networks.
-Once the gateway receives a packet transmitted by a dynamic device in one channel, it transmits back to it an acknowledgement in the same channel, after a fixed-time delay, as it is done in the LoRaWAN standard.
-This \emph{ACK} allows the device to learn about the channel quality and thus, to use learning algorithms for the purpose of best channel selection.
+Once the gateway receives a packet transmitted by a dynamic device in one channel (\ie, if no collision occurred), it transmits back to it an acknowledgement in the same channel, after a fixed-time delay, as it is done in the LoRaWAN standard.
+This \emph{Ack} allows the device to learn about the channel quality (\ie, mean availability) and thus, to use learning algorithms for the purpose of best channel selection.
 
 In this demo, we can generate scenarios with different parameters (number of channels, interfering traffic, etc) in order to evaluate the performance of learning in various settings.
-Moreover, we compare the performance of learning with that of the random uniform access to channels, which is the current state-of-the-art of commercial LPWAN solutions.
+Moreover, we compare the performance of learning strategies with that of the random uniform access to channels, which is the current state-of-the-art of commercial LPWAN solutions \cite{Raza17}.
 %
@@
 
-Regarding the physical layer, we consider a QPSK constellation. Moreover, we use simplified packets composed of two parts.
+Regarding the physical layer, we consider a QPSK constellation (\emph{Quadrature Phase-Shift Keying}). Moreover, we use simplified packets composed of two parts.
 The first part is the \emph{preamble} which is used for the purpose of synchronization (phase correction).
@@
 For example, this index can be a simple QPSK symbol ($\pm1\pm1j$).
-Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet (\emph{e.g.}, $1+j \mapsto 1-j$).
+Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
 Thanks to this index, we can have several devices communicating with the same gateway.
@@
 In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
-In this case, the \emph{ACK} was for him, and it knows that its packet has been received and decoded correctly by the gateway.
+In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the gateway.
 
@@
 % Octoclock
-To ease the synchronization in both time and frequency between the boards representing the dynamic objects and the gateway, we use an Octoclock \cite{OctoclockProduct}, also by Ettus Research,
+To ease the synchronization in both time and frequency between the boards representing the dynamic objects and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
 % https://www.ettus.com/product/details/OctoClock-G
@@
 
-We used GNU Radio Companion (GRC, version $3.7$, $2017$),
+We used the GNU Radio Companion software (GRC, version $3.7$, $2017$),
 and for the demonstration the laptop runs
@@
 The GNU Radio software provides the framework and tools to build and run software radio or just general signal-processing applications.
-GNU Radio applications are a flow-graph: a series of signal processing blocks connected together to describe a data flow.
+GNU Radio applications are flow-graphs: a series of signal processing blocks connected together to describe a data flow.
 For maximum efficiency, we wrote all of our blocks in \texttt{C++}.
@@
 
-We compare the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} against a uniform access algorithm, that uniformly selects its channel at random.
+We compare the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and against a uniform access algorithm, that uniformly selects its channel at random.
 Three objects are compared by their mean successful communication rates, on a horizon of $2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
@@
 \paragraph{Availability of data and materials}
-
+%
 The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at
@@
 The demonstration only requires a laptop and free tools,
-as the laptop should run GNU/Linux and a distribution like Ubuntu or Debian.
+as the laptop should run a GNU/Linux distribution (like Ubuntu or Debian).
 
 \paragraph{Video}
-
+%
 A $6$-minute \textbf{video} showing our demonstration is available online at \texttt{\url{youtu.be/HospLNQhcMk}}.
-It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
-It especially shows the connections between the USRP boards, the Octoclock, and the master laptop.
+It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
+It also shows the connections between the USRP boards, the Octoclock, and the master laptop.
 
 \paragraph{Special acknowledgment}
-
-We acknowledge the work of two CentraleSup{\'e}lec students,
+%
+We acknowledge the work of two engineering students, at CentraleSup{\'e}lec campus of Rennes,
 Cl{\'e}ment Barras and Th{\'e}o Vanneuville, for their GNU Radio project in Spring $2017$,

### 2019-02-27 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Finishing Chapter:4

@@
 
-In this Section, we extend the model presented before in order to take into account the possibility for retransmission of messages after a collision.
+In this Section, we extend the model presented before in order to take into account the possibility for retransmission of a message after a collision.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.
@@
 For that end, several heuristics employing \UCB{} algorithms are examined, to explore the contextual information provided by the number of retransmissions.
-Our results show that approaches based on \UCB{} obtain a significant improvement in terms of successful transmission probabilities, compared to a naive approach which does not learn.
+In this Section, our results show that approaches based on \UCB{} obtain a significant improvement in terms of successful transmission probabilities, compared to a naive approach which does not learn.
 Furthermore, it also reveals that a pure \UCB{} channel access is as efficient as more sophisticated learning strategies.
@@
 
-In the context of Cognitive Radio \cite{Mitola99,Haykin05},
-Multi-Arm Bandit (MAB) algorithms \cite{Auer02,Auer02,Bubeck12} have been recently proposed as a potential solution for channel access in LPWA networks \cite{Bonnefoi18,Azari18,Bonnefoi17}.
+% In the context of Cognitive Radio \cite{Mitola99,Haykin05},
+% Multi-Arm Bandit (MAB) algorithms \cite{Auer02,Auer02,Bubeck12} have been recently proposed as a potential solution for channel access in LPWA networks \cite{Bonnefoi18,Azari18,Bonnefoi17}.
 We presented in Section~\ref{sec:4:firstModel} the impact of non-stationarity on the network performance using MAB algorithms is studied.
 Low-cost algorithms following two well-known approaches, such as the Upper-Confidence Bound (\UCB{}) \cite{Auer02,Auer02}, and the Thompson Sampling (TS) algorithms \cite{Thompson33} have reported encouraging results.
-Other recent directions include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
-and realistic empirical simulations \cite{kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
-of the application of MAB algorithms for slotted wireless protocols in a decentralized manner,
-or for instance applications to multi-hoping networks \cite{Mitton16,Toldov16}.
+When considering the application of MAB algorithms for slotted wireless protocols in a decentralized manner,
+other recent directions of research first include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
+then realistic empirical simulations like in \cite{kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
+,
+and finally applications to multi-hoping networks \cite{Mitton16,Toldov16} or other kinds of networks \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
+%
 None of the above mentioned works discusses in detail the impact of retransmissions on the performance of MAB learning algorithms as we do now.
@@
 For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
-Proposed approach employs contextual information provided by the number of retransmissions, and implemented at each device, so that no coordination among them is needed.
+The proposed approach employs contextual information provided by the number of retransmissions, and is again implemented independently by each device, so that no coordination among them is needed.
 Moreover, our \UCB{}-based heuristics show low complexity making them suitable for being embedded in LPWA devices.
@@
 
-	\item Lastly, we conduct simulations in order to compare the performance of the proposed heuristics with a naive uniform random approach, and a \UCB{} strategy (\ie, without any learning for the retransmissions).
+	\item Lastly, we conduct simulations in order to compare the performance of the proposed heuristics with a naive uniform random approach, and a \UCB{} strategy (\ie, without any learning for the retransmissions, that is, the same channel is used for retransmission).
 \end{itemize}
@@
 
-The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Fig.~\ref{fig:43:Markov_model}.
-When a device have a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, \ie, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
+The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Figure~\ref{fig:43:Markov_model}.
+When a device have a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
 At each time slot, a transition from an idle state to a transmission state (denoted as \texttt{Trans.}) occurs if a packet transmission is required, while waiting states (denoted as \texttt{Wait}), correspond to a $m$ back-off interval.
 
-A device aims to select a channel with the highest probability of a successful transmission, for which it resorts to a reinforcement learning approach, formulated as a MAB problem.
+A device aims to select a channel with the highest probability of successful transmissions, for which it resorts to a reinforcement learning approach, formulated as a MAB problem.
 % , where each channel (also called arms) is viewed as a gambling machine (bandit), and each bandit has a \emph{reward}.
@@
 
-By considering this possibility, the device will have to learn more, thus, we expect that the learning time to be longer, but it could be possible that the final performance gain (\ie, in terms of network performance) increases too.
+By considering this possibility, the device will have to learn more, thus, we expect the learning time to be longer, but it could be possible that the final performance gain increases too, in terms of network performance.
 The next Section~\ref{sub:43:numExp} presents analysis to check this performance gain, for various heuristics based on the \UCB{} algorithm.
@@
 Here after, we start by presenting a mathematical derivation that backups this idea.
-To do so, we study the collision probabilities considering the Markov process depicted in Fig.~\ref{fig:43:Markov_model}, and foresee the impact of addressing bandit strategies, as well as setting guidelines for the design of heuristic approaches.
+To do so, we study the collision probabilities considering the Markov process depicted in Figure~\ref{fig:43:Markov_model}, and foresee the impact of addressing bandit strategies, as well as setting guidelines for the design of heuristic approaches.
 
@@
 
-As it is well known, having a collision during an access time can be overcome by a retransmission procedure (this can take several retransmission attempts).
-What interest us here, is to obtain a mathematical approximation of the collision probability at the second transmission slot $p_{c1}$, as a function of the first collision probability $p_{c}$.
+As it is well known \cite{Abramson1970,Roberts75}, having a collision during an access time can be overcome by a retransmission procedure (this can take several retransmission attempts).
+What interest us, here, is to obtain a mathematical approximation of the collision probability at the second transmission slot $p_{c1}$, as a function of the first collision probability $p_{c}$.
 
@@
 	\item $\mathcal{H}_{2}$:
-	The total number of the back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
+	The total number of back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
 \end{itemize}
 
-Considering one device and a channel,
+Considering one device and one channel,
 we denote $x_t^i$ the probability that it is transmitting a packet for the $i+1$ time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$),
 and let $x_t = \sum_{i=0}^{M-1}x_t^i$ be the probability that it transmits a packet.
-We consider $N$ active devices following the same policy.
+We consider $N > 0$ active devices following the same policy.
 
@@
 We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca4}, and we rewrite the expression of $p_{ca}$ as
-\begin{multline}\label{eq:43:pca}
-	p_{ca} \simeq 1 -\\
-	\left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1}.
-\end{multline}
+\begin{equation}\label{eq:43:pca}
+	p_{ca} \simeq 1 - \left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1}.
+\end{equation}
 
 Finally, our approximation of $p_{c1}$ can be obtained by inserting \eqref{eq:43:pca} in \eqref{eq:43:decomppc1}.
+
+\begin{align}\label{eq:43:final_expression_pc1}
+	p_{c1} &= p_{ca}+\left(1-p_{ca}\right)p_c = (1 - p_c) p_{ca} + p_c \nonumber \\
+	\simeq (1 - p_c)(1 - \left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1} + p_c \\
+	TODO
+\end{align}
 
@@
 %
-In Fig.~\ref{fig:43:Approximation_m10}, we show the collision probabilities for different number of devices $N$ (from $N=50$ up-to $N=400$), for both $p_{c}$ and $p_{c1}$.
+In Figure~\ref{fig:43:Approximation_m10}, we show the collision probabilities for different number of devices $N$ (from $N=50$ up-to $N=400$), for both $p_{c}$ and $p_{c1}$.
 %
@@
 that the gap decreases when $p_c$ increases (for fixed $N$ and $m$).
-This backups mathematically the observation we made from Fig.~\ref{fig:43:Approximation_m10}:
+This backups mathematically the observation we made from Figure~\ref{fig:43:Approximation_m10}:
 the smaller the $p_c$, the larger is the gap between $p_c$ and $p_{c1}$.
@@
 In other words, the channels are occupied $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $M = 5$, and $m=5$.
-In Fig.~\ref{fig:43:mainExperiment1}, we show the successful transmission rate versus the number of slots, for all the proposed heuristics.
+In Figure~\ref{fig:43:mainExperiment1}, we show the successful transmission rate versus the number of slots, for all the proposed heuristics.
 
@@
 Moreover, all of these approaches show a fast convergence making them suitable for the targeted application.
-It is also worths mentioning that the employment of the same \UCB{} algorithm for retransmissions denoted here as ``Only \UCB{}'' achieves a better performance, while a ``Random" retransmission features a slight degradation.
+It is also worths mentioning that the employment of the same \UCB{} algorithm for retransmissions denoted here as ``Only \UCB{}'' achieves a better performance, while a ``Random'' retransmission features a slight degradation.
 This result can be explained as follows: the loss of performance related to the separation of information for several algorithms is greater than the gain obtained by considering the first transmissions and retransmissions separately.
@@
 We also consider in our analysis the case where $M=5$, and $m=10$ using ALOHA protocol, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
-The corresponding results are depicted in Fig.~\ref{fig:43:mainExperiment2}.
-In this case the successful transmission rate is degraded compared with achieved results in Fig.~\ref{fig:43:mainExperiment1}, this can be explained with the fact that we are considering in our network more devices that increase the collision probability.
-It is important to highlight,that the ``Random" retransmission heuristic shows a poor performance in comparison to the other heuristics, and it can be attributed to the fact that the number of retransmission is increased, and consequently a
+The corresponding results are depicted in Figure~\ref{fig:43:mainExperiment2}.
+In this case the successful transmission rate is degraded compared with achieved results in Figure~\ref{fig:43:mainExperiment1}, this can be explained with the fact that we are considering in our network more devices that increase the collision probability.
+It is important to highlight,that the ``Random'' retransmission heuristic shows a poor performance in comparison to the other heuristics, and it can be attributed to the fact that the number of retransmission is increased, and consequently a
 learning approach is able to take advantage of it.
@@
 First, MAB learning algorithms are very useful to reduce the collision rate in LPWA networks, a gain of up-to $30\%$ of successful transmission rate is observed after convergence.
-A second conclusion that can be highlighted is that, using learning mechanisms for retransmissions can be an interesting way to reduce collisions in networks with massive deployments of IoT as this can be checked in Fig.~\ref{fig:43:mainExperiment2}, where the random retransmission heuristic is not very advantageous in front of the  \UCB-based approaches that use learning for channel selection during the retransmission procedure.
+A second conclusion that can be highlighted is that, using learning mechanisms for retransmissions can be an interesting way to reduce collisions in networks with massive deployments of IoT as this can be checked in Figure~\ref{fig:43:mainExperiment2}, where the random retransmission heuristic is not very advantageous in front of the  \UCB-based approaches that use learning for channel selection during the retransmission procedure.
 

### 2019-02-27 2-Chapters/4-Chapter/chapter4.tex
Finishing Chapter:4

@@
 Kaufmann12Thompson}, for the first model.
-For the demonstration and the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for to solve the problem of learning to retransmit efficiently.
+For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for to solve the problem of learning to retransmit efficiently.
 
@@
 % For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16}.
-Many recent works show that MAB algorithms work well for real-world radio signal.
+Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
+Many recent works show that stationary MAB algorithms work well for real-world radio signal.
 However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
 
-We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
-Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is not stationary.
-As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
-We then validate the model with a hardware implementation, detailed in Section~\ref{sec:4:gnuradio}.
+We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed but frequency- and time-slotted IoT network.
+Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
+As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary, making it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
+We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
 %
-We conclude this Chapter by presenting an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
+We conclude this Chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
 
@@
 %
-By using such low-cost algorithms, we demonstrated empirically that a dynamically reconfigurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.
+By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.
 

### 2019-02-27 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Okay je réorganise et réécris pas mal pour ce chapitre 4.
I'm done for today on Chapter 4.
Done for now.

@@
 \subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
+\label{sec:5:proofLagrangian}
 
@@
 %
-First, we need to justify that the objective function is quasi-concave, in each of its coordinates.
+First, we need to justify that the objective function is quasi-convex, in each of its coordinates.
 %
@@
     \item
-    So the gradient is canceling $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) |_{D_i=D_i^*} = 0$
+    So the gradient is zero $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) |_{D_i=D_i^*} = 0$
     iff $D_i^*$ satisfies $(1-p)^{D_i^*}( 1 + \log(1-p) D_i^*) = \lambda / (1-p)^{S_i - 1}$. Let $x = \log(1-p) D_i^*$ this is equivalent to $\mathrm{e}^{x}(1 + x) = \lambda / (1-p)^{S_i - 1}$
@@
     By using the $\mathcal{W}$-Lambert function $\mathcal{W}$ \cite{Corless96}, reciprocal of $y \mapsto \mathrm{e}^{y}y$, we get $x = y - 1 = \mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1$.
-    So the gradient of the Lagrangian is canceling iff $D_i^* = \max(0, 1/\log(1-p) x) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1\right]^{\dagger}$, because $D_i^*$ has to be non-negative.
+    So the gradient of the Lagrangian is zero iff $D_i^* = \max(0, 1/\log(1-p) x) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1\right]^{+}$, because $D_i^*$ has to be non-negative.
     This gives the $i$-th coordinate of the unique saddle point of $f(\boldsymbol{D})$,

### 2019-02-27 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for now.

@@
 Therefore, the probability that this device has a collision at the first transmission is $p_c$, and has the following expression
+%
 \begin{equation}\label{eq:43:1}
@@
 Moreover, from \eqref{eq:43:1} we define the probability $p_{cp}(n)$ that involves the collision of $n$ packets sent by each IoT device (for any $1\leq n \leq N-1$), during the first transmission slot, and is defined by the following equation
+%
 \begin{equation*}\label{eq:43:2}
@@
 Therefore, the expression of $p_{ca}$ is
+%
 \begin{align}\label{eq:43:sumpc2}
@@
 Assuming $\mathcal{H}_{1}$ amounts to consider that $x\ll 1$. As a consequence, the sum in equation \eqref{eq:43:sumpca3} can be supplemented by negligible terms,
-
+%
 \begin{equation}\label{eq:43:sumpca4}
@@
 We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca4}, and we rewrite the expression of $p_{ca}$ as
+%
 \begin{equation}\label{eq:43:pca}
@@
 Finally, our approximation of $p_{c1}$ can be obtained by inserting \eqref{eq:43:pca} in \eqref{eq:43:decomppc1}.
-
+%
 \begin{align}\label{eq:43:final_expression_pc1}
 	p_{c1} &= p_{ca}+\left(1-p_{ca}\right)p_c = (1 - p_c) p_{ca} + p_c \nonumber \\
-	\simeq (1 - p_c)(1 - \left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1} + p_c \\
-	TODO
+	&\simeq \left(1 - p_c\right) (1 - \left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1} + p_c.
 \end{align}
@@
 	\item
-	On the one hand, in a congested network, when devices suffer from a large probability of collision on their first transmission (\ie, $p_c$ is not so small), then $p_{c1}\simeq p_c$ and so devices cannot really hope to reduce their collision probabilities even if the use a different channel for retransmission.
+	On one hand, in a congested network, when devices suffer from a large probability of collision on their first transmission (\ie, $p_c$ is not so small), then $p_{c1}\simeq p_c$ and so devices cannot really hope to reduce their collision probabilities even if the use a different channel for retransmission.
 	\item
-	On the other hand, if $p_c$ is small enough, \ie, in a network not yet too congested, then our derivation shows that $p_{c1} \gg p_c$, meaning that the possible gain of retransmitting in a different channel that the one used for the first transmission can be large, in terms of collision probability (\eg, up-to $10\%$ in this experimental setting).
+	On the other hand, if $p_c$ is small enough, \ie, in a network not yet too congested, then our derivation shows that $p_{c1} > p_c$, meaning that the possible gain of retransmitting in a different channel that the one used for the first transmission can be large, in terms of collision probability (\eg, up-to $10\%$ in this experimental setting).
 	In other words, when learning can be useful (small $p_c$), learning to retransmit in a different channel can have a large impact on the global collision rate,
@@
 		\For(){$t = 0, \dots, T$}{
-			Compute for each channel $ U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+			Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
 			Transmit in channel $C(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
@@
 
-A device that implements the UCB algorithm is led to focus is transmissions and retransmissions in the channel which has been identified as the best.
+A device that implements the UCB algorithm is led to focus its transmissions and retransmissions in the channel which is currently identified as the best.
 As explained before, focusing in one channel increases the collision probability in retransmissions.
-In this Section, we describe the proposed heuristics for the channel selection in a retransmission. It is carried out taking
+We describe here the proposed heuristics for the channel selection in a retransmission. It is carried out taking
 into account that a device can incorporate a different channel selection strategy while being in a back-off state.
@@
 			\Else(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^r$}){
-				Compute for each channel $U^r_k(t) = \widehat{\mu^r_k}(t) + B^r_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+				Compute $\forall k, U^r_k(t) = \widehat{\mu^r_k}(t) + B^r_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
 				Transmit in channel $C^r(t) = \arg\max_{1\leq k \leq K} U^r_k(t)$\;
@@
 
-Another heuristic is to not use the same algorithm no matter where the collision occurred, but to use $K$ different \UCB{} algorithms.
-Meaning that after a failed first transmission in channel $j$, the device relies on the $k$-th algorithm to decide its retransmission.
+Another heuristic is to not use the same algorithm no matter where the collision occurred, but to use $K$ different second-stage \UCB{} algorithms.
+It means that after a failed first transmission in channel $j$, the device relies on the $j$-th algorithm to decide its retransmission.
 The corresponding algorithm is depicted in Algorithm~\ref{algo:43:UCBthenKp1}.
@@
 
-Although, this approach increases the complexity and storage requirements (of order $\mathcal{O}(K^2)$).
+Although, this approach increases the complexity and storage requirements (now, of order $\mathcal{O}(K^2)$).
 For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
@@
 
-Note that for other networks this heuristic could not be practical, for instance Sigfox consist in a large number of very narrow-band channels.
-The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels.
+Note that for other networks this heuristic could not be practical.
+The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels (\eg, $K=128$).
 But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.
@@
 					j $\leftarrow$ last channel selected by first-stage \UCB\;
-					Compute for each channel $U_k^j(t) = \widehat{\mu_k}^j(t) + B_k^j(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+					Compute $\forall k, U_k^j(t) = \widehat{\mu_k}^j(t) + B_k^j(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
 					Transmit in channel $C^j(t) = \arg\max_{1\leq k \leq K} U^j_k(t)$\;
@@
 				}
-				\uElseIf(\tcp*[f]{Random retransmission with $\mathrm{UCB}^d$}){$t \leq \Delta$}{
+				\uElseIf(\tcp*[f]{Random retransmission}){$t \leq \Delta$}{
 					Transmit randomly in a channel $C(t) \sim \mathcal{U}(1,\ldots,K)$.
 				}
-				\Else(\tcp*[f]{Delayed \UCB{}}){
-					Compute for each channel $ U^d_k(t) = \widehat{\mu^d_k}(t) + B^d_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+				\Else(\tcp*[f]{retransmission with delayed $\mathrm{UCB}^d$}){
+					Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + B^d_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
 					Transmit in channel $C^d(t) = \arg\max_{1\leq k \leq K} U^d_k(t)$\;
@@
 
-We simulate our network considering $N$ devices following the contention Markov process described in Section \ref{sub:43:model}, and a LoRa standard with $K=4$ channels.
+We simulate our network considering $N$ devices following the contention Markov process described in Section \ref{sub:43:model}, and a LoRa standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
 Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of $1\;\mathrm{s}$.
 %
-For the evaluation of the proposed heuristics, a total number of $T=20 \times 10^4$ time slots is considered, and the results are averaged over $10^{3}$ independent random simulations.
+For the evaluation of the proposed heuristics, a total number of $T=20 \times 10^4$ time slots is considered, and the results are averaged over $1000$ independent random simulations.
 %In this way, it allows us to evaluate the learning rate, as well as the convergence of each learning approach.
@@
 In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform repartition of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
-In other words, the channels are occupied $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $M = 5$, and $m=5$.
+In other words, the channels are occupied respectively $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $M = 5$, and $m=5$.
 In Figure~\ref{fig:43:mainExperiment1}, we show the successful transmission rate versus the number of slots, for all the proposed heuristics.
 
-A first result is that all the heuristics clearly outperform the non-learning approach that simply use random channel selection for both transmissions and retransmissions (\ie, the \textbf{no \UCB{}} curve).
-The improvement of the heuristics over the non-learning approach is evident, and for every heuristic that use a kind of learning mechanism it can observed a successful transmission rate that increases rapidly (or equivalently an PLR decreasing).
+A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve).
+The improvement of the heuristics over the non-learning approach is evident, and for every heuristic that uses a kind of learning mechanism it can observed the successful transmission rate increases rapidly (or equivalently the PLR decreases).
 Moreover, all of these approaches show a fast convergence making them suitable for the targeted application.
-It is also worths mentioning that the employment of the same \UCB{} algorithm for retransmissions denoted here as ``Only \UCB{}'' achieves a better performance, while a ``Random'' retransmission features a slight degradation.
+It is also worths mentioning that the employment of the same \UCB{} algorithm for retransmissions, denoted here as ``Only \UCB{}'', achieves a better performance, while a ``Random'' retransmission features a slight degradation.
 This result can be explained as follows: the loss of performance related to the separation of information for several algorithms is greater than the gain obtained by considering the first transmissions and retransmissions separately.
@@
 	\includegraphics[width=0.75\linewidth]{ResultsUCB.eps}
-	\caption[First comparison between the exposed heuristics for the retransmission: Only \UCB, Random, \UCB, $K$ \UCB, and Delayed \UCB]{
-		Comparison between the exposed heuristics for the retransmission: \textcolor{blue}{Only \UCB}, \textcolor{cyan}{Random}, \textcolor{purple}{\UCB}, \textcolor{green}{$K$ \UCB}, and \textcolor{red}{Delayed \UCB}.
-		The usage of the same learning policy for transmissions and retransmission is named \textcolor{blue}{Only \UCB{}},
-		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as \textbf{no \UCB{}}.
+	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB'']{
+		Comparison between the exposed heuristics for the retransmission: \textcolor{blue}{``Only \UCB''}, \textcolor{cyan}{``Random''}, \textcolor{purple}{``\UCB''}, \textcolor{green}{``$K$ \UCB''}, and \textcolor{red}{``Delayed \UCB''}.
+		The usage of the same learning policy for transmissions and retransmission is named \textcolor{blue}{``Only \UCB{}''},
+		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as \textbf{``no \UCB{}''}.
 		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
@@
 
-We also consider in our analysis the case where $M=5$, and $m=10$ using ALOHA protocol, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
+We also consider in our experiment the case of an ALOHA protocol using $M=5$, and $m=10$, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
 The corresponding results are depicted in Figure~\ref{fig:43:mainExperiment2}.
-In this case the successful transmission rate is degraded compared with achieved results in Figure~\ref{fig:43:mainExperiment1}, this can be explained with the fact that we are considering in our network more devices that increase the collision probability.
-It is important to highlight,that the ``Random'' retransmission heuristic shows a poor performance in comparison to the other heuristics, and it can be attributed to the fact that the number of retransmission is increased, and consequently a
+In this case the successful transmission rate is degraded compared with achieved results in Figure~\ref{fig:43:mainExperiment1}.
+We can explain this by the fact that we are considering in our network more devices, which increases the collision probability.
+It is important to highlight, that the ``Random'' retransmission heuristic shows a poor performance in comparison to the other heuristics, and it can be attributed to the fact that the number of retransmission is increased, and consequently a
 learning approach is able to take advantage of it.
-Furthermore, the ``\UCB'', ``$K$ \UCB'' and ``Delayed \UCB'' heuristics behave similarly than ``Only \UCB'', after a similar convergence time.
+Furthermore, the ``\UCB'', ``$K$ \UCB'' and ``Delayed \UCB'' heuristics behave similarly to ``Only \UCB'', after a similar convergence time.
 

### 2019-02-27 2-Chapters/4-Chapter/chapter4.tex
Done for now.

@@
 Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
-We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the many ``intelligent'' end-devices.
+We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the presences many ``intelligent'' end-devices which follow a random activation process.
 
@@
 %
-By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.
+By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.
 
@@
 
-In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
+In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
 The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
@@
 Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
-We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.
+In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
 
 Moreover, for sake of simplicity we supposed that all devices use the same standard.
-Our future work will consider more realistic interference scenarios and IoT networks, with, \emph{e.g.}, non-slotted time, more than one base station etc.
+Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
 
-Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \emph{e.g.}, one transmission a day.
-So additionally to deciding the channel for communication, each device has to also decide when to communicate.
+Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
+So additionally to deciding the channel for communication (\ie, \emph{where} to communicate), each device has to also decide \emph{when} to communicate.
 % This is another direction of research, that we will investigate in the future.
+However, this clearly leads to a much larger action space, as there are many time slots in one day (for example), and thus we believe that as soon as the action space becomes too large in this extension, the simple MAB-based learning approach could be no longer appropriate.
+It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
+It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo17} or structured MAB \cite{Combes17} models and algorithms for this extension.
 
@@
 Possible future extensions of our demonstration include the following points.
-We could consider more dynamic objects (\emph{e.g.}, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
+We could consider more dynamic objects (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
 We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
-Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing new hardware.
+Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card is enough to simulate another gateway).
 %
@@
 that nevertheless have not been explored for the targeted problem.
-Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this section.
+Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this Chapter.
 

### 2019-03-07 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
An even older reference for the Generalized Likelihood Ratio test.

@@
 
-Generalized Likelihood Ratio tests date back to the seminal work of \cite{barnard1959control} and were for instance studied for change-point detection by \cite{siegmund1995using}.
+Generalized Likelihood Ratio tests date back to the seminal work of \cite{Wilks1938}, and were for instance studied for change-point detection by \cite{barnard1959control,siegmund1995using}.
 Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, can be use for our problem.

### 2019-03-07 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Typo for today, nothing more in one week... DUH

@@
 %
-where $\ell(X_1, \ldots, X_t ; \mu_0)$ and $\ell(X_1, \ldots, X_t ; \mu_0,\mu_1,\tau)$ denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
+where $\ell(X_1, \ldots, X_t ; \mu_0)$ and $\ell(X_1, \ldots, X_t ; \mu_0,\mu_1,\tau)$ denote the likelihoods of the first $t$ observations under a model in $\cH_0$ and $\cH_1$.
 High values of this statistic tend to indicate rejection of $\cH_0$.
@@
 %
-Indeed, one can show that $\GLR(t) = \sup_{s \in [1,n]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:t}\right)\right]$,
+Indeed, one can show that $\GLR(t) = \sup_{s \in [1,t]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:t}\right) + (t-s) \times \kl\left(\hat{\mu}_{s+1:t},\hat{\mu}_{1:t}\right)\right]$,
 where for $k \leq k'$, $\hat{\mu}_{k:k'}$ denotes the average of the observations collected between the instants $k$ and $k'$. This motivates the definition of the Bernoulli GLR change point detector.
@@
 \begin{definition}\label{def:6:GLRDef}
-    The Bernoulli GLR change point detector with threshold function $\beta(t,\delta)$ is
+    The Bernoulli GLR change point detector with threshold function $\beta(n,\delta)$ is
     % \vspace*{-5pt}% WARNING

### 2019-03-11 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Modification suivant les suggestions de Christophe, cf #13

@@
 %
-We show that using learning algorithms does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically changing channel.
+We show that using learning algorithms does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
 In the studied scenario, stochastic MAB learning provides an up to $16\%$ gain in term of successful transmission probabilities, and has near optimal performance even in non-stationary settings with a majority of intelligent devices.
@@
 Static devices form an interfering traffic, which could have been generated by devices using other standards as well.
-We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller optimally distribute them in the channels at the beginning of the scenario (that is, an ideal approach).
+Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel.
+So our first hypothesis is actually not constraining.
+We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller optimally distributes all of them in the channels at the beginning of the scenario (that is, an ideal approach).
 Then, these three reference scenarios allow to evaluate the performance of \UCB{} and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
@@
 We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of objects all sends uplink packets to the (unique) network gateway.
-Messages can be sent for instance centered around the $433.5\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
+Messages can be sent in a fixed number $K\geq2$ of wireless channels (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
 The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
 %
-The devices can transmit their packets in $K\geq2$ channels (\eg, $K=4$).
-In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1\;\mathrm{s}$ in our implementation).
+Note that this choice of number of channels and of frequency bands is for instance the one of the current LoRa standard in Europe, but of course the precise choice does not really matter, and future standards can choose different parameters.
+
+The devices can transmit their packets in one of the $K$ channels.
+In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1\;\mathrm{s}$ in our implementation).
+In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the object by letting only one gateway sends back an acknowledgement, and the natural choice is for the network to send this acknowledgement by the gateway which received the uplink message.
+For simplicity but without loss of generality, we consider a network with only one gateway,
+but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming uplink messages from the IoT objects being paired in the network.
+%
 
@@
 %
-To simulate networks designed for the Internet of Things (IoT), we consider a protocol \textbf{with no sensing}, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet (but only to the objects being paired with this base station in this network).
+To simulate networks designed for the Internet of Things (IoT), we consider a protocol \textbf{with no sensing}, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet (but only to the objects being paired with this gateway in this network).
 %
-By considering a small number of wireless channels (\eg, $K = 4$) and a PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
+By considering a small number of wireless channels, and a unique PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
 the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\ie, the most vacant one).
@@
 All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
-In each time slot, the devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (\ie, no sensing).
-Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station.
-In one channel, if only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part (on the same channel).
+In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication (\ie, no sensing).
+Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the gateway.
+In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
 If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \Ack{} is not transmitted.
@@
 Each frame is composed by a fix-duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
-If a packet is well received, the base station replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
+If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
     \label{fig:41:protocol}
@@
     \item
-    \emph{Static} end-devices have poor RF abilities, and each of them uses only one channel to communicate with the base station.
+    \emph{Static} end-devices are identical, and each of them uses only one channel to communicate with the gateway.
 Their choice is assumed to be fixed in time (\ie, stationary).
@@
     \item
-    \emph{Dynamic} (or \emph{smart}) end-devices have richer RF abilities, they can use all the available channels, by quickly re-configuring their RF transceiver on the fly.
+    \emph{Dynamic} (or \emph{smart}) end-devices can use all the available channels, by quickly changing their communication channel at any time.
 They can also store communication successes or failures they experienced in each channel, in order to change channel, possibly at every time slot.
     We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
+    Note that of course we assume a limited storage capacity, so no object stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bits IEEE float number for each of the $K$ channels).
 \end{itemize}
@@
 The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed (\ie, real time during two messages), and $1/p$ is proportional to the \emph{duty cycle}.
-For instance for IoT objects, a ``smart'' refrigerator could have to send a daily message, and if a message can be sent every second, then $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.
+For instance for IoT objects sending messages with a duration of $1$ second exactly, a ``smart'' refrigerator could have to send a daily message, and if a message can be sent every second, then $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.
 
@@
 We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (for instance, about $1000$ to $100000$, while $K$ is about $4$ to $256$).
-As this problem is only interesting if devices are able to communicate reasonably efficiently with the base station, we assume devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
-We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks\footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
+As this problem is only interesting if devices are able to communicate reasonably efficiently with the gateway, we assume that devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
+Note that even unlicensed bands have such limitations, as for instance this is a strict requirement for any object using the $867\;\mathrm{MHz}$ ISM band (and its regulation also enforces a maximum transmission power, but we do not consider power transmission in this work).
+%
+We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks.
+% \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
 Indeed, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ objects, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall i, S_i > 0$).
-Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero than using learning algorithms cannot improve much.
+Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
 Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$ to have an average of active objects not larger than the number of channels.
@@
 We consider the network from the point of view of one dynamic device.
-Following its Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
+Following a Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
 it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
-Then, the dynamic device starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
+Then, the dynamic device waits in this channel $C(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example of a ``smart fridge'' considered above).
 Before sending another message (\ie, at time $t+1$), the dynamic device knows if it received or not this \Ack{} message.
@@
 We consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
-Contrary to many previous work done in the CR field (\eg, Opportunistic Spectrum Access),
+Contrary to many previous work done in the CR field \cite{Jouini10,Jouini12} (\eg, Opportunistic Spectrum Access),
 the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
@@
 %
-They are however useful for our numerical simulations, to show that our MAB-based approaches quickly learn to perform almost optimally.
+They are however useful for our numerical simulations, and will be used as references, in order to show that our MAB-based approaches quickly learn to perform almost optimally.
 
@@
 % Simulation parameters
-We suppose a network with $S + D = 2000$ end-devices, and one IoT base station.
+We suppose a network with $S + D = 2000$ end-devices, and one IoT gateway.
 Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\eg, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).

### 2019-03-11 2-Chapters/4-Chapter/chapter4.tex
Modification suivant les suggestions de Christophe, cf #13

@@
 The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
-We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP cards \cite{USRPDocumentation}.
+We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP platforms \cite{USRPDocumentation}.
 
-More precisely, on one hand, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
+More precisely, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
 We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
@@
 %
-On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real radio hardware.
+On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
 
-On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending our first model, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
+We then present in Section~\ref{sec:4:retransmissions} a second model extending our first model, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
 In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
@@
 % ----------------------------------------------------------------------------
-\section{Motivations for Selfish MAB Learning for IoT Networks}
+% \section{Motivations for Selfish MAB Learning for IoT Networks}
+\section{Introduction}
 \label{sec:4:motivations}
@@
 Kaufmann12Thompson}, for the first model.
-For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for to solve the problem of learning to retransmit efficiently.
+For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for solve the problem of learning in order to retransmit efficiently.
 
@@
 % ----------------------------------------------------------------------------
-\section{Conclusion of Chapter~\ref{chapter:4}}
+\section{Conclusion}
 \label{sec:4:conclusion}
@@
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter~\ref{chapter:4}}
+\section{Appendix}
 \label{sec:4:appendix}

### 2019-03-11 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Commence à intégrer changements d'Emilie, cf. #13

@@
 
-Where $(a)^{+} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ \cite{Corless96}.
+Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (it can be computed numerically in an efficient manner) \cite{Corless96}.
 Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint

### 2019-03-11 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
WHOUHOU citations do work now
Start to copy/paste all what I had written about SMPyBandits. This chapter 3 will be very quickly done.
Finishing Chapter:4
Generic typos

@@
 %
-In this Section, we present an improved variant of \ExpQ, called \Aggr.
+In this section, we present an improved variant of \ExpQ, called \Aggr.
 For synthetic MAB problems,
@@
 \paragraph{Outline}
-This Section is organized as follows.
+This section is organized as follows.
 First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
@@
 % ----------------------------------------------------------------------
-\subsection{Aggregating Bandit Algorithms}\label{sub:25:aggregation}
+\subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
 
@@
 			\KwIn{$N$ bandit algorithms, $\Alg_1, \dots, \Alg_N$, with $N \geq 2$}
-			\KwIn{Number of arms, $K \geq 1$}
+			\KwIn{Number of arms, $K \geq 2$}
 			\KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
@@
 between the aggregation algorithm $\Alg_{\mathrm{aggr}}$ and the best aggregated algorithm $\Alg_a$.
-This is in contrast with the (classical) regret, which measure the difference with the best fixed-arm strategy (Eq.~\eqref{eq:classicalregret}).
+This is in contrast with the (classical) regret, which measure the difference with the best fixed-arm strategy (Eq.~\eqref{eq:5:classicalregret}).
 Thus,  even if the aggregated algorithms have logarithmic (classical) regret, having an adversarial regret scaling as $\sqrt{T}$ does not permit to exhibit a logarithmic (classical) regret for the aggregation algorithm.
@@
 Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
-It is not presented in this Section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
+It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
 and appears to be more robust than simple non-aggregated algorithms.
@@
 % ----------------------------------------------------------------------------
-\bibliographystyle{ieeetr}
-\bibliography{IEEE_WCNC__2018__Paper__Lilian_Besson__07-17}
+% \bibliographystyle{ieeetr}
+% \bibliography{IEEE_WCNC__2018__Paper__Lilian_Besson__07-17}

### 2019-03-11 2-Chapters/3-Chapter/chapter3.tex
Generic typos

@@
 
-This Chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
+This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
 Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
@@
 %
-This Chapter is based on the following publication: \cite{SMPyBanditsJMLR}.
+This chapter is based on the following publication: \cite{SMPyBanditsJMLR}.
 The code code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is publicly released under the MIT open-source License.

### 2019-03-11 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Generic typos

@@
 
-In this Section, we take a step back and present the different feedback levels possible for multi-player MAB algorithms.
+In this section, we take a step back and present the different feedback levels possible for multi-player MAB algorithms.
 For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees.
@@
 
-The rest of the Section is organized as follows.
+The rest of the section is organized as follows.
 We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
@@
 
-In the next Section, we present an algorithm that has a logarithmic regret,
+In the next section, we present an algorithm that has a logarithmic regret,
 while ensuring that the number of sub-optimal selections is matching the lower bound of Theorem~\ref{thm:5:BetterLowerBound}.
@@
 
-The analysis of \Selfish{} is harder, but we tried our best to obtain some understanding of the behavior of this algorithm, that seems to be doing surprisingly well in many contexts, as in our experiments with $K=9$ arms and in extensive experiments not reported in this Section.
+The analysis of \Selfish{} is harder, but we tried our best to obtain some understanding of the behavior of this algorithm, that seems to be doing surprisingly well in many contexts, as in our experiments with $K=9$ arms and in extensive experiments not reported in this section.
 However, a disappointing result is that we found simple problems, usually with small number of arms, for which the algorithm may fail.

### 2019-03-11 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Should be good for Chapter 4 :boom:

@@
 As explained before, setting up the future Internet of Things (IoT) networks will require to support more and more communicating devices.
-In this Section, we prove that intelligent devices in unlicensed bands can use Multi-Armed Bandit (MAB) learning algorithms to improve resource exploitation.
+In this section, we show that intelligent devices in unlicensed bands can use Multi-Armed Bandit (MAB) learning algorithms to improve resource exploitation.
 %
@@
 
-The aim of this Section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases, and the stochastic hypothesis is more and more violated.
+The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
+and the network is less and less stationary.
+% and the stochastic hypothesis is more and more violated.
 To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
@@
 %
-The rest of this Section is organized as follows.
+The rest of this section is organized as follows.
 The system model is introduced in Section~\ref{sub:41:systemModel}.
@@
 
+\begin{figure}[!b]
+    \centering
+    \includegraphics[width=0.70\linewidth]{system_model1.eps}
+    \caption{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
+    \label{fig:41:system_model1}
+\end{figure}
+
 The devices can transmit their packets in one of the $K$ channels.
@@
 
-These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated neighboring networks.
+These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated neighboring networks (in \textcolor{orange}{orange} in Figure~\ref{fig:41:system_model1}).
 This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels.
 %
-To simulate networks designed for the Internet of Things (IoT), we consider a protocol \textbf{with no sensing}, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet (but only to the objects being paired with this gateway in this network).
+To simulate networks designed for the Internet of Things (IoT), we consider a protocol \textbf{with no sensing} and no repetition of uplink messages.
+The gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) who succeeded in sending an uplink packet.
 %
 By considering a small number of wireless channels, and a unique PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
-the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\ie, the most vacant one).
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=0.70\linewidth]{system_model1.eps}
-    \caption{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
-    \label{fig:41:system_model1}
-\end{figure}
+the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use, that is, the most vacant one.
+Indeed, in this model, the \emph{quality} of the channels (\ie, the ressources) are identified by their vacancy, or one minus their occupancy rates.
 
@@
 All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
-In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication (\ie, no sensing).
+In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication, with no sensing.
 Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the gateway.
 In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
-If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \Ack{} is not transmitted.
+If two or more devices send an uplink packet in the same slot, the uplink packets collide, and the acknowledgement \Ack{} is not transmitted.
+In this case, we say that there is a \emph{collision} in this time slot.
 In other words, if the gateway cannot successfully decode the incoming (uplink) messages because they were corrupted by collisions, it does not send any \Ack{} back.
@@
 %
-There are two types of end-devices in the network:
+We make the following assumptions on the network.
+We assume that there are two types of end-devices in the network:
 \begin{itemize}
@@
     Note that of course we assume a limited storage capacity, so no object stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bits IEEE float number for each of the $K$ channels).
+    Our goal is to propose a learning algorithm than can be implemented by each dynamic devices, in a decentralized and independent manner, to improve their successful communication rate automatically.
 \end{itemize}
 
-There are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
+We further assume that there are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
 Furthermore, in channel $i \in \llbracket 1; K \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{K} S_i$).
@@
 
-\paragraph{Dense networks and an hypothesis}
+\paragraph{Some assumptions on the network occupation}
 %
@@
 For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
-The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, it is to maximize its number of successful transmissions, or its cumulative reward,
+The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
 $r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
-as it is usually done in MAB problems.
+which is a usual objective in MAB problems.
 
-This problem is a special case of the stochastic MAB (see Chapter~\ref{chapter:2}).
+This problem is a special case of the stochastic MAB with
 % , where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$.
 % Several types of reward distributions have been considered in the literature, and rewards are binary in our model.
-We consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
+% We consider only
+Bernoulli distributions.
+% , in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
 Contrary to many previous work done in the CR field \cite{Jouini10,Jouini12} (\eg, Opportunistic Spectrum Access),
 the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
-In our model, rewards rather come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
+In our model, rewards rather come from receiving or not an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
+A reward of one indicates that the acknowledgement was received on time, and a reward of zero indicates the opposite.
 
-The problem parameters are $\mu_1,\dots,\mu_K$, the mean availability of the channels, and they are of course unknown to each dynamic device, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively improve its successful communication rate.
+The problem parameters
+% are $\mu_1,\dots,\mu_K$, the mean availability of the channels,
+are $S_1, \dots, S_K$ the (stationary) occupancy of the channels by the static devices,
+and they are of course unknown to each dynamic device, so to maximize their cumulated rewards, they must learn the distributions of the channels, in order to be able to progressively improve its successful communication rate.
+%
+The goal is thus to propose a simple sequential algorithm to be applied identically and independently by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
 %
 This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a device has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
-%
-The goal is thus to design a new or chose an existing simple sequential algorithm, to be applied identically and independently by each dynamic device, in a fully distributed setting (each device runs its own algorithm, from its observations), in order to minimize collisions and maximize the fraction of successful transmissions of all the dynamic devices.
 
@@
 
-We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the repartition of static devices (\ie, $(S_i)_i$ or equivalently, $(\mu_i)_i$), and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.} is possible.
+We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of $(S_i)_i$ the repartition of static devices, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.} is possible.
 We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.
@@
 
-Considered alone, each dynamic device implements a learning algorithm to play a bandit game.
+% Considered alone, each dynamic device
+We propose algorithms in which each dynamic device ignores all the other one and
+implements a learning algorithm to play a bandit game.
+%
 In each time slot, if it has to communicate (which happens with probability $p$), then it chooses a channel and it receives a reward $1$ if the transmission is successful, $0$ otherwise.
@@
 However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning and random activation of each devices).
-
+%
 Despite this, we show in the next subsection that running a stochastic bandit algorithm for each device based on its own rewards is surprisingly successful.
@@
 The two algorithms, \UCB{} and Thompson sampling (TS), were presented in Section~\ref{sec:2:famousMABalgorithms}.
-The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
+The \UCB{} algorithm uses a parameter $\alpha > 0$, set to $\alpha = 1/2$ in our experiments.
+% originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
 %
@@
 
-\paragraph{Multi-Player MAB with \emph{collision avoidance}?}
-
-Another idea could be to try to use a \emph{multi-player MAB} model, as proposed by \cite{Zhao10}, to describe our problem.
-We study this other direction in the next Chapter~\ref{chapter:5}.
-
-The main difference between the two models is the fact that in this Chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
-Active devices denote static and dynamic devices that have to transmit at a fixed time,
-and their random activation patter makes the number of active devices a random variable.
-Analyzing multi-player MAB models under this hypothesis is much harder, and is left as a future work.
-
-% In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices: they are \iid{} with mean $\mu_i = 1 - p S_i$.
-% Moreover, dynamic devices are usually assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
-% The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
-% However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.
-
 \paragraph{\emph{Adversarial} bandit algorithms?}
@@
 Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
-The \emph{adversarial MAB} algorithms is a broader family, and a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Bubeck12}.
-Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform much worse than both \UCB{} and TS in the same experiments.
+The \emph{adversarial MAB} algorithms is a broader family, of which a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Bubeck12}.
+Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform significantly worse than both \UCB{} and TS in the same experiments,
+therefore we did not report results here.
+% FIXME maybe we should... ?
 %(see Section \ref{sub:41:numericalResults}).
-Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary model, as its performance guarantee are true in \emph{any} setting.
-But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
-%\footnote{~We preferred not to talk about \emph{regret} in this study, but in a few words: it is a measure of how \emph{bad} the algorithm performed in terms of its accumulated rewards, in comparison to the best possible policy, and should be as small as possible.
+However, contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary model, as its performance guarantees are true in \emph{any} setting.
+% But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
+% %\footnote{~We preferred not to talk about \emph{regret} in this study, but in a few words: it is a measure of how \emph{bad} the algorithm performed in terms of its accumulated rewards, in comparison to the best possible policy, and should be as small as possible.
 For stochastic algorithms, being ``efficient'' means having a regret bounded as $R_T = \mathcal{O}(\log T)$, but for adversarial algorithms, it means having $R_T = \mathcal{O}(\sqrt{K T})$.}
-worse than the one for stochastic ones.
+% worse than the one for stochastic ones.
 % (in their respective case of application).
-More is left on this aspect for a future work.
+Further research on this aspect could lead to interesting future works.
 

### 2019-03-11 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Should be good for Chapter 4 :boom:

@@
 
-In this Section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept of the model previously presented in Section~\ref{sec:4:firstModel}.
+In this section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept of the model introduced in Section~\ref{sec:4:firstModel}.
 
@@
 and a traffic generator that emulates radio interference from many other objects.
-Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol, which does not require any specific overhead for the learning.
+Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol with retransmissions, which does not require any specific overhead for the learning.
 %
-Similarly to the previous Section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
+Similarly to the previous section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
 %
@@
 %
-We use a TestBed designed in 2017 by our team SCEE \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
+We use a TestBed designed in 2017 by the SCEE team at the Rennes campus of CentraleSupélec \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
 and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.
@@
 This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent objects at least do not reduce their successful communication rate in comparison to the naive objects.
-This also shows that in case of non-uniform stationary traffic, the MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent objects.
+This also shows that in case of non-uniform stationary traffic, MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent objects.
 
-We refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
-which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Section~\ref{sub:42:results}).
-Then we discuss the relevance of a MAB model for our IoT application.
+% We refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
+% which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Section~\ref{sub:42:results}).
+% Then we discuss the relevance of a MAB model for our IoT application.
 
-\paragraph{Outline}
+% \paragraph{Outline}
 
-The rest of this Section is organized as follows.
-Our implementation is presented in Section~\ref{sub:42:implementation}, with details about the user interface, and results are given in Section~\ref{sub:42:results}.
+% The rest of this section is organized as follows.
+% Our implementation is presented in Section~\ref{sub:42:implementation}, with details about the user interface, and results are given in Section~\ref{sub:42:results}.
 
@@
 
-In this Section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
+In this section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
 We first describe the simplified physical layer of this demo, then we present our GNU Radio implementation.
@@
 
-We compare the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and against a uniform access algorithm, that uniformly selects its channel at random.
-Three objects are compared by their mean successful communication rates, on a horizon of $2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
+We compare the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
+For one dynamic object, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
 Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} shows the results averaged on $10$ repetitions using the same conditions.
@@
 Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
-Each object was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $[15\%, 10\%, 2\%, 1\%]$.
+Each object was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_i)_i = [15\%, 10\%, 2\%, 1\%]$.
+The occupancy rate of a channel denote the mean occupancy, and it is implemented using one USRP card that runs the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).
 % They did not overlap each other, as
@@
 It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
-It also shows the connections between the USRP boards, the Octoclock, and the master laptop.
+It also shows the connections between the USRP boards, the Octoclock, the master laptop etc.
 
-\paragraph{Special acknowledgment}
-%
-We acknowledge the work of two engineering students, at CentraleSup{\'e}lec campus of Rennes,
-Cl{\'e}ment Barras and Th{\'e}o Vanneuville, for their GNU Radio project in Spring $2017$,
-as we took inspiration in the GNU Radio code and the discussions we had at this time.
+% \paragraph{Special acknowledgment}
+% %
+% We acknowledge the work of two engineering students, at CentraleSup{\'e}lec campus of Rennes,
+% Cl{\'e}ment Barras and Th{\'e}o Vanneuville, for their GNU Radio project in Spring $2017$,
+% as we took inspiration in the GNU Radio code and the discussions we had at this time.

### 2019-03-11 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Should be good for Chapter 4 :boom:

@@
 
-In this Section, we extend the model presented before in order to take into account the possibility for retransmission of a message after a collision.
+In this section, we extend the model presented before in order to take into account the possibility for retransmission of a message after a collision.
+This extension is of major importance as most protocols for real-world IoT networks use retransmissions, and it is for instance the case of the LoRa standard.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.
@@
 The presented strategies allow IoT devices to improve their access to the network and their autonomy, while taking into account the impact of encountered radio collisions.
-For that end, several heuristics employing \UCB{} algorithms are examined, to explore the contextual information provided by the number of retransmissions.
-In this Section, our results show that approaches based on \UCB{} obtain a significant improvement in terms of successful transmission probabilities, compared to a naive approach which does not learn.
+For that end, several heuristics employing \UCB{} algorithms are examined, to explore the information provided by the number of retransmissions.
+In this section, our results show that approaches based on \UCB{} obtain a significant improvement in terms of successful transmission probabilities, compared to a naive approach which does not learn.
 Furthermore, it also reveals that a pure \UCB{} channel access is as efficient as more sophisticated learning strategies.
 
-% \TODOL{This Section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://hal.inria.fr/hal-02049824}}
+% \TODOL{This section is based on the publication ``Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions'', see \texttt{https://hal.inria.fr/hal-02049824}}
 
-% \TODOL{TODO:
-
-% - Include the modifications suggested by the reviewers of our paper!
-% - Be sure to not repeat any motivations / maths / explanations given before
-% }
-
-% https://edas.info/showPaper.php?m=1570524364
 % https://hal.inria.fr/hal-02049824
 
-% ----------------------------------------------------------------------
-\subsection{Retransmissions in an ALOHA like protocol}
-\label{sub:43:introduction}
-% ----------------------------------------------------------------------
+% % ----------------------------------------------------------------------
+% \subsection{Retransmissions in an ALOHA like protocol}
+% \label{sub:43:introduction}
+% % ----------------------------------------------------------------------
 
@@
 
-The aim of this Section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
+The aim of this section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
 For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
@@
 
-The contributions of this Section are summarized as follows:
+The contributions of this section are summarized as follows:
 \begin{itemize}
@@
 
-The rest of the Section is organized as follows.
+The rest of the section is organized as follows.
 First the system model is introduced in Section~\ref{sub:43:model},
@@
 % ----------------------------------------------------------------------
-\subsection{Extending our model with retransmissions}
+% \subsection{Extending our model with retransmissions}
+\subsection{Presentation of the model with retransmissions}
 \label{sub:43:model}
@@
 We make the hypothesis that the transmission is successful if the channel is available, otherwise upon radio collision, these devices will attempt to retransmit their packet up-to $M$ times, with $M \in\mathbb{N}$.
-Note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\llbracket 0, m-1 \rrbracket$, where $m \in\mathbb{N}, m>0$ is the length of the back-off interval.
+It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\llbracket 0, m-1 \rrbracket$, where $m \in\mathbb{N}, m>0$ is the length of the back-off interval
+(note the difference between $m$ and $M$).
 
 % ----------------------------------------------------------------------
-\paragraph{Model of our IoT devices}
+\paragraph{Model of IoT devices}
 
@@
 A device aims to select a channel with the highest probability of successful transmissions, for which it resorts to a reinforcement learning approach, formulated as a MAB problem.
+Contrarily to what may appear at first sight,
+the goal is not to minimize the number of retransmission, but to maximize the probability of successful transmission, considering both the first transmission of a message and the retransmissions of the same message, because the objective of each device is to maximize its battery life by minimizing its number of failed transmissions.
 % , where each channel (also called arms) is viewed as a gambling machine (bandit), and each bandit has a \emph{reward}.
@@
 % It motivates our present work for which we consider the retransmissions in the analysis of MAB algorithms.
-It motivates our present work for which we consider the number of retransmissions, carried out by each device to address several MAB algorithms.
+It motivates our present work for which we consider the number of retransmissions, carried out by each device.
+% to address several MAB algorithms.
 
@@
 	\includegraphics[width=0.70\linewidth]{Markov_model-eps-converted-to.pdf}
-	\caption{The considered Markov model for the behavior of all the devices in the network.}
+	\caption{The Markov model of the behavior of all devices paired to the considered IoT network using the ALOHA protocol.}
 	\label{fig:43:Markov_model}
@@
 Thus, it seems interesting to consider in the decision making policy the possibility for a device to retransmit in a different channel.
-One of our motivations to develop new MAB algorithms for our problem is this option of using a different communication channels between the first transmission and the next retransmissions.
+One of our motivations to develop new MAB algorithms for our problem is this option of using a different communication channel between the first transmission and the next retransmissions.
 
 By considering this possibility, the device will have to learn more, thus, we expect the learning time to be longer, but it could be possible that the final performance gain increases too, in terms of network performance.
-The next Section~\ref{sub:43:numExp} presents analysis to check this performance gain, for various heuristics based on the \UCB{} algorithm.
+We present in Section~\ref{sub:43:numExp} an analysis to check this performance gain, for various heuristics based on the \UCB{} algorithm.
 %
 Here after, we start by presenting a mathematical derivation that backups this idea.
-To do so, we study the collision probabilities considering the Markov process depicted in Figure~\ref{fig:43:Markov_model}, and foresee the impact of addressing bandit strategies, as well as setting guidelines for the design of heuristic approaches.
+To do so, we study the collision probabilities considering the Markov process depicted in Figure~\ref{fig:43:Markov_model}, and foresee the impact of using bandit strategies, as well as setting guidelines for the design of heuristic approaches.
 
@@
 
-We consider two hypotheses $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as,
+We make two approximations $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as,
 \begin{itemize}
@@
 %
-We present below four heuristics for this second stage (short names in ``quotes'' correspond to the legend on Figures~\ref{fig:43:mainExperiment1}, \ref{fig:43:mainExperiment2}).
+We present below four heuristics for this second stage.
+Their short names are used in the legend on Figures~\ref{fig:43:mainExperiment1}, \ref{fig:43:mainExperiment2}), and are given in ``quotes'' in the corresponding paragraphs.
 
@@
 It is described below in Algorithm~\ref{algo:43:UCBthenRandom}.
+More precisely, the first-stage \UCB{} use rewards built from the acknowledgments that the device received or not for its first-stage transmission, but do not use any feedback about any of the retransmissions of any message.
 
@@
 
-Instead of applying a random channel selection,
+Instead of applying a random channel selection for retransmission,
 another heuristic is to use a second \UCB{} algorithm in the second stage.
@@
 
+\paragraph{A last idea that we did not explore}
+%
+Instead of considering a first-stage \UCB{} followed by a random channel retransmission, we could have considered a random first-stage channel retransmission followed by a second-stage \UCB.
+This was not considered in our study \cite{Bonnefoi2019WCNC}, and we preferred to not add it to the experiments presented below, for two reasons: to avoid clutter in the plots, and to win some time as these experiments had to run for quite a long time.
+
 % ----------------------------------------------------------------------
@@
 A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve).
-The improvement of the heuristics over the non-learning approach is evident, and for every heuristic that uses a kind of learning mechanism it can observed the successful transmission rate increases rapidly (or equivalently the PLR decreases).
+The improvement of the heuristics over the non-learning approach is clear, and for every heuristic that uses a kind of learning mechanism it can observed the successful transmission rate increases rapidly (or equivalently the PLR decreases).
 Moreover, all of these approaches show a fast convergence making them suitable for the targeted application.
@@
 %
-The source code (MATLAB or Octave) used for the simulations and the figures of this Section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
+The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
 It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in July and August $2018$.

### 2019-03-11 2-Chapters/4-Chapter/chapter4.tex
Commence à intégrer changements d'Emilie, cf. #13
Should be good for Chapter 4 :boom:

@@
 
-More precisely, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
+More precisely, we start by motivating the idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
 We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
@@
 
-We then present in Section~\ref{sec:4:retransmissions} a second model extending our first model, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
+Then we present in Section~\ref{sec:4:retransmissions} a second model, extending our first model, in which devices can retransmit their messages for a fixed number of maximum retransmission, like in the ALOHA protocol.
 In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
-Interestingly, the simpler heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
+Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
 
@@
 
-This Chapter is mainly based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
+This chapter is mainly based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 
@@
 %
-In this Chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
+In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
 We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
@@
 Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
-As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary, making it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
+As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary, making it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
 We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
 %
-We conclude this Chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
+We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
 
@@
 % ----------------------------------------------------------------------------
-\section{Extension of our model to account for retransmissions}
+\section{Extending the model to account for retransmissions}
 \label{sec:4:retransmissions}
@@
 
-We give in this Section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.
+We give in this section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.
+
+\TODOL{Je devrais réécrire toute cette conclusion, si j'ai le temps, pour faire mieux qu'une suite de 3 conclusions et 3 future works, et être plus concis et moins ``short sighted'' comme disait Emilie.}
 
@@
 % ----------------------------------------------------------------------
-\paragraph{Summary of the first model}
+% \paragraph{Summary of the first model}
 \label{sub:41:conclusion}
@@
 % ----------------------------------------------------------------------
-\paragraph{Conclusions taken from the demonstration}
+% \paragraph{Conclusions taken from the demonstration}
 \label{sub:42:conclusionFromDemonstration}
@@
 % ----------------------------------------------------------------------
-\paragraph{Conclusion about this second model}
+% \paragraph{Conclusion about this second model}
 \label{sub:43:conclusion}
@@
 
-Future works related to this Chapter include the following directions.
+Future works related to this chapter include the following directions.
 
-\paragraph{Possible extensions of the first model}
+% \paragraph{Possible extensions of the first model}
 
-The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device.
+The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device.
 Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
 In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
-
+%
 Moreover, for sake of simplicity we supposed that all devices use the same standard.
@@
 
-\paragraph{Extensions of the demonstration}
+% \paragraph{Extensions of the demonstration}
 
@@
 
-\paragraph{Possible extensions for the second model}
+% \paragraph{Possible extensions for the second model}
 
-The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
+Finally, the utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
 Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
@@
 
+\subsection*{Summary of this chapter}
+
+To sum-up, we focussed in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices of the IoT network.
+We presented two models of wireless IoT network without sensing, inspired by the ALOHA protocol, and with or without retransmission of uplink packages in case of collisions.
+In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
+%
+On one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
+On the other hand, it is quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
+Sadly, analyzing our model with thousands of independent devices communicating and learning in their own (random) time scale turned out to be of extreme difficulty.
+
+\paragraph{Multi-Player MAB with \emph{collision avoidance}}
+
+We are actually able to analyze a simpler model, if we assume to have at most $K$ devices with a transmission probability of $p=1$.
+Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-player MAB} model to describe our problem, as proposed by \cite{Zhao10}.
+%
+The main difference between the two models is the fact that in this chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
+If static and dynamic devices that have to transmit at a fixed time are denoted \emph{active devices},
+and their random activation patter makes the number of active devices a random variable.
+Analyzing multi-player MAB models under this hypothesis is much harder, and is left as a future work.
+%
+We study this other direction in the next Chapter~\ref{chapter:5}.
+
+% In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices: they are \iid{} with mean $\mu_i = 1 - p S_i$.
+% Moreover, dynamic devices are usually assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
+% The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
+% However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.
+
 \newpage
@@
 \label{sec:4:appendix}
+% \section{Proof of Proposition \ref{prop:41:Lagrangian}}
+\label{sec:5:proofLagrangian}
 % ----------------------------------------------------------------------------
 
-We include here a missing proof from Section~\ref{sec:4:firstModel}.
+% We include here a missing proof from Section~\ref{sec:4:firstModel}.
 

### 2019-03-12 2-Chapters/6-Chapter/chapter6.tex
Generic typos
Almost done reordering chapter 6 and cutting useless parts. I guess the chapter will be finished tomorrow.

@@
 
-\chapter{Non-Stationary MAB Models and Possible Applications for IoT Networks}
+% \chapter{Non-Stationary MAB Models and Possible Applications for IoT Networks}
+\chapter{The Piece-wise Stationary MAB Model and the GLR-klUCB Algorithm}
 \label{chapter:6}
@@
 
-Abstract for Chapter 6.
+\paragraph{Abstract}
+
+In this last chapter, we study a generalization of the stationary multi-armed bandit model,
+to account for possible non-stationarity of possible applications of MAB algorithms.
+Instead of dealing with changes happening possibly at every time step, like in adversarial models, we consider a finite number of \emph{change-points}, and the problem is stationary on any interval between two consecutive change-points (it is piece-wise stationary).
+%
+We start by reviewing the existing previous work on piece-wise stationary MAB models, the main families of proposed solutions are either passively adaptive or actively adaptive.
+In practice, actively adaptive strategies usually greatly outperform the passively adaptive ones.
+The more efficient strategies from the current state-of-the-art are based on a combination of an efficient algorithm for the stationary problem with an efficient online change detection algorithm, like CUSUM, PHT or the Generalized Likelihood Ratio Test (GLRT).
+Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any sub-Bernoulli distributions, which include distributions of bounded support.
+Our test, B-GLRT, enjoys finite-time guarantees in terms of both its false alarm probability (in a stationary model, the test should not detect any change), and its detection delay (in a model with a change-point, the test should detect within a reasonable delay).
+%
+Thanks to these two non-asymptotic properties, the B-GLRT test can be combined with an efficient bandit policy, the \klUCB{} index policy, to propose an efficient algorithm for piece-wise stationary bandit problems, that we design \GLRklUCB.
+Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
+\MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
+%
+We give the first unified analysis of a piece-wise stationary bandit algorithm for the two variants, and our analysis is completely non-asymptotic and yield regret bounds with explicit constants.
+Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$ with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
+%
+Finally, we give extensive numerical experiments to compare our approach, in its two variants, against the current state-of-the-art policies, on different piece-wise stationary bandit problems, of increasing difficulty.
+%
+We include additional details about our policy \GLRklUCB, including a sensibility analysis of its two parameters, a numerical evaluation of its time and memory costs as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
+
+\vfill{}
+
+\paragraph{Publications}
+
+This chapter is based on the following publications: \cite{Besson2019GLRT,Besson2019Gretsi}.
 
@@
 
-\TODOL{Rewrite and reorder most of this Chapter, with more subsections, less paragraphs etc, and less copy-paste of our paper \cite{Besson2019GLRT}.}
+\TODOL{Rewrite and reorder most of this chapter, with more subsections, less paragraphs etc, and less copy-paste of our paper \cite{Besson2019GLRT}.}
 
@@
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter 6}
+\section{Appendix}
 \label{sec:6:appendix}

### 2019-03-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Almost done reordering chapter 6 and cutting useless parts. I guess the chapter will be finished tomorrow.

@@
 
-\TODOL{This chapter is basically a raw include from my paper ``The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits'', see https://hal.inria.fr/hal-02006471}
-
-We propose a new algorithm for the piece-wise \iid{} non-stationary bandit problem with bounded rewards.
-Our proposal, \GLRklUCB, combines an efficient bandit algorithm, \klUCB, with an efficient, \emph{parameter-free}, change-point detector, the Bernoulli Generalized Likelihood Ratio Test, for which we provide new theoretical guarantees of independent interest.
+\TODOL{Be much more concise, if I already give an abstract in the beginning of the chapter...}
+
+% We propose a new algorithm for the piece-wise \iid{} non-stationary bandit problem with bounded rewards.
+% Our proposal, \GLRklUCB, combines an efficient bandit algorithm, \klUCB, with an efficient, \emph{parameter-free}, change-point detector, the Bernoulli Generalized Likelihood Ratio Test, for which we provide new theoretical guarantees of independent interest.
 We analyze two variants of our strategy, based on \emph{local restarts} and \emph{global restarts}, and show that their regret is upper-bounded by  $\cO(\Upsilon_T \sqrt{T \log(T)})$ if the number of change-points $\Upsilon_T$ is unknown, and by $\cO(\sqrt{\Upsilon_T T \log(T)})$ if $\Upsilon_T$ is known.
-This improves the state-of-the-art bounds, as our algorithm needs no tuning based on knowledge of the problem complexity other than $\Upsilon_T$.
-We present numerical experiments showing that \GLRklUCB{} outperforms passively and actively adaptive algorithms from the literature, and highlight the benefit of using local restarts.
+% This improves the state-of-the-art bounds, as our algorithm needs no tuning based on knowledge of the problem complexity other than $\Upsilon_T$.
+% We present numerical experiments showing that \GLRklUCB{} outperforms passively and actively adaptive algorithms from the literature, and highlight the benefit of using local restarts.
 
 % -----------------------------------------------------------------
-\section{Motivation for Non-Stationary MAB Models}
+\section{Introduction}
 \label{sec:6:Introduction}
 
-Multi-Armed Bandit (MAB) problems form a well-studied class of sequential decision making problems, in which an agent repeatedly chooses an action $I_t \in\{1,\dots,K\}$ or ``arm'' --in reference to the arm of a one-armed bandit-- among a set of $K$ arms.
+\TODOL{Be much more concise: in the thesis, MAB models and applications have already been introduced, motivated etc!}
+
+% Multi-Armed Bandit (MAB) problems form a well-studied class of sequential decision making problems, in which an agent repeatedly chooses an action $I_t \in\{1,\dots,K\}$ or ``arm'' --in reference to the arm of a one-armed bandit-- among a set of $K$ arms.
 In the most standard version of the stochastic bandit model, each arm $i$ is associated with an \iid{} sequence of rewards $(r_{i,t})$ that follow some distribution of mean $\mu_i$.
 Upon selecting arm $I_t$, the agent receives the reward $r_{I_t,t}$ associated to the chosen arm, and her goal is to adopt a sequential sampling strategy that maximize the expected sum of these rewards.
-This is equivalent to minimizing the \emph{regret}, defined as the difference between the total reward of the oracle strategy always selecting the arm with largest mean, $\mu^*$, and that of our strategy: $R_T = \bE[\sum_{t=1}^T(\mu^* - \mu_{I_t})]$.
-
-Regret minimization in stochastic bandits has been extensively studied since the works of \cite{Robbins52} and \cite{LaiRobbins85}, and several algorithms with a $\cO(\log(T))$ \emph{problem-dependent} regret upper bound have been proposed (see, \eg, \cite{LattimoreBanditAlgorithmsBook} for a survey).
+This is equivalent to minimizing the \emph{regret}, defined as the difference between the total reward of the oracle strategy always selecting the arm with largest mean, $\mu^*$, and that of our strategy:
+
+% \[ R_T = \bE[\sum_{t=1}^T(\mu^* - \mu_{I_t})]. \]
+
+% Regret minimization in stochastic bandits has been extensively studied since the works of \cite{Robbins52} and \cite{LaiRobbins85}, and several algorithms with a $\cO(\log(T))$ \emph{problem-dependent} regret upper bound have been proposed (see, \eg, \cite{LattimoreBanditAlgorithmsBook} for a survey).
 Among those, the \klUCB{} algorithm (\cite{KLUCBJournal}) has been shown to be asymptotically optimal for Bernoulli distributions (in that it exactly matches the lower bound given by \cite{LaiRobbins85}) and can also be employed when the rewards are assumed to be bounded in $[0,1]$.
-\emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{Garivier18}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.
+% \emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{Garivier18}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.
 
@@
 Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
-
+%
 For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not evolve over time} may be a big limitation.
@@
 
-In this paper, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB.
+In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB.
 Like previous approaches -- \CUSUM{} (\cite{LiuLeeShroff17}) and \MUCB{} (\cite{CaoZhenKvetonXie18}) -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector.
@@
 
-The article is structured as follows.
+\paragraph{Outline}
+%
+This chapter is structured as follows.
 We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}.
@@
 
+\TODOL{Rename GLR, GLRT and GLR Bernoulli to B-GLRT like what we did in our GRETSI 2019 paper \cite{Besson2019Gretsi}.}
+
 % -----------------------------------------------------------------
-\section{The Piece-Wise Stationary Bandit Setup and Related Works}
+\section{The Piece-Wise Stationary Bandit Model}
 \label{sec:6:BanditSetting}
@@
 The performance of $\pi$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\pi$:
-\begin{align}\label{def:6:nti}
-    R_T^{\pi} &= \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
-\end{align}
+%
+\begin{equation}\label{def:6:nti}
+    R_T^{\pi} = \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
+\end{equation}
 
 In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
-We define the $k$-th breakpoint by $\tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}$.
+We define the $k$-th breakpoint by
+\begin{equation}
+    \tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}.
+\end{equation}
+%
 Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, the rewards $(X_{i,t})$ associated to each arms are \iid.
+
 Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
@@
 
+\paragraph{An interesting generalization}
+%
 The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while the adversarial model can be considered as a special (worst) case, with $\Upsilon_T = T$.
@@
 
+\TODOL{Speak about the article \cite{WeiSrivastava18} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}
+
 % -----------------------------------------------------------
-\paragraph{Related work.}
+\section{Related Work}
+\label{sec:6:relatedWork}
 
@@
 Active strategies monitors the rewards of each arm by using a change detection algorithm (\cite{Basseville93}), and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
-Passive strategies are tuned with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.
-
-% \TODOL{Review of previous work (see section on Experiments), and state-of-the-art of passively adaptive and actively adaptive algorithms}
+Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.
 
@@
 %
+The RExp3 algorithm (\cite{Besbes14stochastic}) can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm.
+Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
+A  $\cO(V_T^{1/3} T^{2/3})$ regret bound is proved, which is weaker than existing results in our setting.
+Hence we do no include this algorithm in our experiments.
+%
 
@@
 Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust in the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in their paper.
+The DTS algorithm can perform well in practice, for instance with $\gamma=0.75$ on our Problems $1$ and $2$ (see Section~\ref{sec:6:NumericalExperiments}).
+However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
 %
@@
 \CUSUMUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but only for Bernoulli distributions,
-and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge (see XXX).
+and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
 %
@@
 
-\TODOL{Finish presenting MUCB and CUSUM, and highlights the differences!}
+\CUSUMUCB{} is based on a rather complicated two-sided \CUSUM{} test, that uses the first $M$ samples from one arm to compute an initial average, and then detects whether a drift of size larger than $\varepsilon$ occurred from this value by checking whether a random walk based on the remaining observations crosses a threshold $h$.
+Thus it requires the tuning of three parameters, $M$, $\varepsilon$ and $h$.
+\CUSUMUCB{} performs \emph{local restarts} using this test, to reset the history of \emph{one arm} for which the test detects a change.
+\MUCB{} uses a much simpler test, based on the $w$ most recent observations from an arm: a change is detected if the absolute difference between the empirical means of the first and second halves of those $w$ observations exceeds a threshold $h$.
+So it requires the tuning of two parameters, $w$ and $h$.
+\MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
+Compared to \CUSUMUCB{}, note that \MUCB{} is numerically much simpler as it only uses a bounded memory, of order $\bigO{K w}$ for $K$ arms.
 
@@
 %
-Note that the regret bound for these methods are still open-problems.
+Note that the regret bound for these methods are still open-problems,
+therefore we choose to not include them in our experiments.
+
+\paragraph{Advantages of our approach}
+%
+\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
+%
+However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
+Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
+%
+In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
+Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
+This study is supported by both theoretical and empirical results.
+%
+Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing  much in terms of regret.
 
@@
 
-\TODOL{FIXME next paragraph is the short version of related works review I have to merge the long and short together and keep as much details as I can!}
-
-The piece-wise stationary bandit model was first studied by \cite{Kocsis06,YuMannor09,Garivier11UCBDiscount}.
-It is also known as \emph{switching} (\cite{MellorShapiro13}) or \emph{abruptly changing stationary} (\cite{WeiSrivastava18}) environment.
-%We do not consider the \emph{slowly-varying} MAB models (\cite{Louedec16,WeiSrivastava18}), even if it is interesting for some applications.
-% We review previous works that studied the piece-wise stationary bandit model.
-To our knowledge, all the previous approaches combine a standard bandit algorithm, like \UCB{}, Thompson Sampling or EXP3, with a strategy to account for changes in the arms distributions.
-This strategy often consists in \emph{forgetting old rewards}, to efficiently focus on the most recent ones, more likely to be similar to future rewards.
-We make the distinction between \emph{passively} and \emph{actively} adaptive strategies.
-
-The first proposed mechanisms to forget the past consist in either discounting rewards (at each round, when getting a new reward on an arm, past rewards are multiplied by $\gamma^n$ if that arm was not seen since $n>0$ times, for a discount factor $\gamma\in(0,1)$), or using a sliding window (only the rewards gathered in the $\tau$ last observations of an arm are taken into account, for a window size $\tau$).
-Those strategies are passively adaptive as the discount factor or the window size are fixed, and can be tuned as a function of $T$ and $\Upsilon_T$ to achieve a certain regret bound.
-Discounted UCB (D-UCB) was proposed by \cite{Kocsis06} and analyzed by \cite{Garivier11UCBDiscount}, who prove a $\cO(\sqrt{\Upsilon_T T }\log(T))$ regret bound, if $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$.
-The same authors proposed the Sliding-Window UCB (SW-UCB) and prove a $\cO(\sqrt{\Upsilon_T T \log(T)})$ regret bound, if $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$.
-
-More recently, \cite{RajKalyani17} proposed the Discounted Thompson Sampling (DTS) algorithm, which performs well in practice with $\gamma=0.75$.
-However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
-%
-The RExp3 algorithm (\cite{Besbes14stochastic}) can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm.
-Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
-A  $\cO(V_T^{1/3} T^{2/3})$ regret bound is proved, which is weaker than existing results in our setting.
-Hence we do no include this algorithm in our experiments.
-
-% \TODOE{Il faut faire attention pour DTS, tu attaques leur papier! Qu'est-ce que nos expériences montrent exactement qui est en contradiction avec ce qu'ils disent?}
-% \TODOE{Je vire EXP3.S: c'est une version d'EXP3 qui a des garanties en terme de ``nombre de breakpoint'' dans un cadre adversarial, pas stochastique.
-Il n'est pas basé sur des restart et un discount}
-% \TODOE{Vérifie que je raconte pas de conneries pour RExp3, Odalric et Subho l'avaient mis dans la catégorie ``passive'' mais je comprends pas trop comment marche l'algo}
-
-%
-The first \emph{actively adaptive} strategy is Windowed-Mean Shift (\cite{YuMannor09}), which combines any bandit policy with a change point detector which performs \emph{adaptive restarts} of the bandit algorithm.
-However, this approach is not applicable to our setting as it takes into account side observations.
-Another line of research on actively adaptive algorithms uses a Bayesian point of view.
-A Bayesian Change-Point Detection (CPD) algorithm is combined with Thompson Sampling by \cite{MellorShapiro13}, and more recently in the Memory Bandit algorithm of \cite{Alami17}.
-Both algorithms do not have theoretical guarantees and their implementation is very costly, hence we do not include them in our experiments.
-Our closest competitors rather use frequentist CPD algorithms (see, e.g.
-\cite{Basseville93}) combined with a bandit algorithm.
-% The idea behind Memory Bandit is to use an expert aggregation algorithm, like EXP4 from \cite{Auer02}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
-% At each time step, a new expert is introduced, and experts correspond to different Thompson sampling algorithms, each using a different history of pulls and rewards.
-Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
-%
-The first algorithm of this flavor, Adapt-EVE algorithm (\cite{Hartland06}) uses a Page-Hinkley test and the \UCB{} policy, but no theoretical guarantee are given.
-EXP3.R (\cite{Allesiardo15,Allesiardo17}) combines a CPD with EXP3, and the history of all arms are reset as soon as a a sub-optimal arm is detecting to become optimal and it achieve a $\cO(\Upsilon_T \sqrt{T \log(T)})$ regret.
-This is weaker than the $\cO(\sqrt{\Upsilon_T T \log(T)})$ regret achieved by two recent algorithms, \CUSUMUCB{} (\cite{LiuLeeShroff17}) and Monitored UCB (\MUCB, \cite{CaoZhenKvetonXie18}).
-
-\CUSUMUCB{} is based on a rather complicated two-sided \CUSUM{} test, that uses the first $M$ samples from one arm to compute an initial average, and then detects whether a drift of size larger than $\varepsilon$ occurred from this value by checking whether a random walk based on the remaining observations crosses a threshold $h$.
-Thus it requires the tuning of three parameters, $M$, $\varepsilon$ and $h$.
-\CUSUMUCB{} performs \emph{local restarts} using this test, to reset the history of \emph{one arm} for which the test detects a change.
-\MUCB{} uses a much simpler test, based on the $w$ most recent observations from an arm: a change is detected if the absolute difference between the empirical means of the first and second halves of those $w$ observations exceeds a threshold $h$.
-So it requires the tuning of two parameters, $w$ and $h$.
-\MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
-% Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
-% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
-% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
-% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
-Compared to \CUSUMUCB{}, note that \MUCB{} is numerically much simpler as it only uses a bounded memory, of order $\bigO{K w}$ for $K$ arms.
-
-% \TODOE{I chose not to talk about $\alpha$ here in the end to gain space: in the end, everyone share that parameter. Maybe add a footnote in section 4 saying that CUSUM does random exploration}
-
-\textbf{Advantages of our approach.}
-\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
-%
-However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
-Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
-%
-In this paper, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
-Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
-This study is supported by both theoretical and empirical results.
-%
-Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing  much in terms of regret.
-
-% \TODOE{I would keep what is below with a small mention (or footnote) in the experimental section}
-
 % -----------------------------------------------------------------
-\section{The Bernoulli GLR Change Point Detector}
+\section{The Bernoulli GLR Change-Point Detector}
 \label{sec:6:ChangePointDetector}
@@
 Sequential change-point detection has been extensively studied in the statistical community (see, \eg, \cite{Basseville93} for a survey).
-In this article, we are interested in detecting changes on the mean of a probability distribution with bounded support.
+In this section, we are interested in detecting changes on the mean of a probability distribution with bounded support.
 Assuming that  we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$.
@@
 High values of this statistic tend to indicate rejection of $\cH_0$.
-Using the form of the likelihood for Bernoulli distribution, this statistic can be written with the binary relative entropy $\kl$,
+Using the form of the likelihood for Bernoulli distribution, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
 %
@@
     The Bernoulli GLR change point detector with threshold function $\beta(n,\delta)$ is
-    % \vspace*{-5pt}% WARNING
+    %
     \begin{equation}\label{def:6:GLR}
@@
 
-In the next section, we provide new non-asymptotic results about the Bernoulli GLR test under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
-Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the Bernoulli GLR may stop earlier that the Gaussian GLR based on the quadratic divergence $2(x-y)^2$.
+In this section, we provide new non-asymptotic results about the Bernoulli GLR test under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
+Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the Bernoulli GLR may stop earlier that the Gaussian GLR, based on the quadratic divergence $2(x-y)^2$.
 
 % ----------------------------------------------------------------------------
-\subsection{Properties of the Bernoulli GLR}\label{subsec:6:PropGLR}
+\subsection{Non-Asymptotics Properties of the Bernoulli GLR}\label{subsec:6:PropGLR}
+
+When used for a bandit problem, the two main properties of a sequential change-point detection test are its false alarm probability and its detection delay.
+%
+A small false alarm probability ensures that no detection occurs before it should (\ie, no useless detection occur on stationary segments), and a small detection delay ensures that, if the stationary segments are long enough, then every change-points on every arms will be detected after a short enough amount of samples from that arm.
+%
+We give below two Lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:DetectionDelay} which bounds the detection delay if there is enough samples before a change.
+
+%-----------------------------------------------------------------------------
+\paragraph{False alarm probability}
 
@@
 To define $\beta$, we need to introduce the function $\cT$,
+%
 \begin{equation}\label{def:6:function_T}
-    \cT(x) ~:=~ 2 \tilde h\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right)
+    \cT(x) ~:=~ 2 \tilde h\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
 \end{equation}
 where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$.
+%
 And for any $x \ge 0$, $\tilde h(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise.
 The function $\cT$ is easy to compute numerically.
-% The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function (\cite{Corless96}) (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
+The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function (\cite{Corless96}) (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
+%
 Its use for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on the function $\cT$ are also given:  $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
@@
     Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
-Then the Bernoulli GLR test satisfies $\bP_{\mu_0}(\hat\tau_\delta < \infty) \leq \delta$ with the threshold function
-   % \vspace*{-5pt}% WARNING
+    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
+    $\bP_{\mu_0}(\hat\tau_\delta < \infty) \leq \delta$, with the threshold function
+    %
     \begin{equation}\label{def:6:beta}
-        \beta(n,\delta)= 2\cT\left(\frac{\ln(3n\sqrt{n}/\delta)}{2}\right) + 6\ln(1+\ln(n)).
+        \beta(n,\delta) = 2\cT\left(\frac{\ln(3n\sqrt{n}/\delta)}{2}\right) + 6\ln(1+\ln(n)).
     \end{equation}
 \end{lemma}
+
+%-----------------------------------------------------------------------------
+\paragraph{Detection delay}
 
@@
 
+\TODOL{PROVE the lemma on detection delay !}
+
+\begin{lemma}\label{lem:6:Delay}
+    Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
+    and let $\Delta = |\mu_0 - \mu_1|$.
+    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
+    \begin{equation}
+        \bP_{\mu_0,\mu_1,\tau} (\widehat{\tau}_{\delta} \geq \tau + u) \leq \exp( -\frac{2\tau u}{\tau + u}\left(\max[ 0, \Delta - \sqrt{\frac{\tau + u}{2\tau u} \beta(\tau + u,\delta)} ]\right)^2 ).
+    \end{equation}
+\end{lemma}
+
+For a threshold $\beta$ chosen as in the Lemma~\ref{lemma:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lemma:Delay} is that if the break $\tau$ takes place after $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude.
+%
+In other words,
+if the B-GLRT test observes enough samples before rupture,
+its delay $\widehat{\tau}_\delta$ is in the order of $O(\Delta^{-2} \ln(1/\delta))$, in high probability.
+
 % ----------------------------------------------------------------------------
@@
 Thus, in practice the following adaptation may be interesting, based on down-sampling the possible time steps:
+%
 \begin{equation}\label{def:6:GLRTricks}
@@
 \end{equation}
+%
 for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant enjoys the exact same false-alarm control.
-However, the detection delay may be slightly increased. In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups has little impact on the regret.
+However, the detection delay may be slightly increased.
+In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups has little impact on the regret.
 
 % ------------------------------------------------------------------------
+%-----------------------------------------------------------------------------
 \section{The \GLRklUCB{} Algorithm}
 \label{sec:6:GLRklUCB_Algorithm}
+%-----------------------------------------------------------------------------
+
+We start by giving the pseudo-code of our algorithm, by explaining every part, then we give our finite-time regret upper bounds, for the two variants of using local or global restarts.
+The analysis in both cases is using a unified proof technique, that we present before giving more details about the two proofs.
+
+%-----------------------------------------------------------------------------
+\paragraph{Details About the Algorithm}
 
@@
 It also needs a third ingredient, some forced exploration parameterized by $\alpha\in(0,1)$ to ensure each arm is sampled enough and changes can also be detected on arms currently under-sampled by the bandit algorithm.
-%We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
-% Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
-% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
-% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
-% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
 %
@@
     \centering
-    \begin{small}
+    % \begin{small}
     % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
@@
     \end{algorithm}
-    \end{small}
+    % \end{small}
 \end{figure}
+
+%-----------------------------------------------------------------------------
+\paragraph{When does \GLRklUCB{} restart?}
 
@@
 To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
-\begin{small}
-    \[\sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),\]
-\end{small}%
+\begin{equation}
+    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
+\end{equation}
 %
@@
 the index is defined as
+%
 \begin{equation}\label{def:6:UCB}
@@
 
+%-----------------------------------------------------------------------------
+\paragraph{Forced exploration}
+
 The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by (\cite{CaoZhenKvetonXie18}), whereas \CUSUMUCB{} performs randomized exploration.
-A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples} (proved in Appendix~\ref{proof:6:EnoughSamples}).
-
-% \vspace*{-8pt}% WARNING
+%
+We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
+Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
+\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
+and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
+Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
+%
+A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples}, proved below.
+
 \begin{proposition}\label{prop:6:EnoughSamples}
@@
 
+\begin{proof}
+    We consider one arm $i\in\{1,\dots,K\}$, and when the \GLRklUCB{} algorithm is running,
+    we consider two time steps $s\leq t\in\N^*$, chosen between two restart times for that arm $i$.
+    Lines~$3$-$4$ state that $A_u = u \mod \lceil \frac{K}{\alpha} \rceil$
+    if $u \mod \lceil \frac{K}{\alpha} \rceil \in \{1,\dots,K\}$
+    (see Algorithm~\ref{algo:6:GLRklUCB} for details).
+    Thus we have
+    \begin{align*}
+        n_i(t) - n_i(s)
+        & = \sum_{u=s+1}^{t} \indic(A_u = i) \\
+        & \geq \sum_{u=s+1}^{t} \indic\left(A_u = i, A_u = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
+        & \geq \sum_{u=s+1}^{t} \indic\left(u \mod \left\lceil \frac{K}{\alpha} \right\rceil = i\right) \\
+        & = \bigl(t - (s+1) + 1\bigr) / \left\lceil \frac{K}{\alpha} \right\rceil
+        \geq \left\lfloor \frac{\alpha}{K} (t - s) \right\rfloor.
+    \end{align*}
+    Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
+\end{proof}
+
+%-----------------------------------------------------------------------------
+%-----------------------------------------------------------------------------
+\section{Regret results and proofs for \GLRklUCB}
+\label{sec:6:RegretAnalysis}
+%-----------------------------------------------------------------------------
+
+This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the finite-time regret bounds for the two variants, with their proofs, which both follow the skeleton of the unified analysis.
+
+%-----------------------------------------------------------------------------
+\subsection{A Unified Regret Analysis}
+
+\TODOL{Rewrite, include the proof of at least one case here! No need to be that quick for the thesis, we only wrote it like this for the COLT version \cite{Besson2019GLRT}.}
+
+In this section, we sketch a unified proof for Theorem~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound}, whose detailed proofs are given in Sections \ref{proof:6:mainRegretBoundGlobal} and \ref{proof:6:mainRegretBound} respectively.
+We emphasize that our approach is significantly different from those proposed by \cite{CaoZhenKvetonXie18} for \MUCB{} and by \cite{LiuLeeShroff17} for \CUSUMUCB.
+
+Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
+Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
+\begin{align*}
+ &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
+ & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
+\end{align*}
+%
+Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
+%
+\begin{eqnarray*}
+R_T \leq T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} + \ \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
+\end{eqnarray*}
+%
+Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay.
+Each regret bound then follows from upper bounds on term (A), term (B), and on the failure probability $\bP(\cE_T^c)$.
+%
+To control (A) and (B), we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
+
+The tricky part of each proof, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$.
+For example for local changes (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
+\begin{equation}
+    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
+\end{equation}
+%
+where $\hat\tau_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
+Introducing the event $\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \hat\tau_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\}$ that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition:
+\begin{align}
+    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
+\end{align}
+
+Term (a) is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
+Observe that under the bandit algorithm, the change point detector associated to arm $i$ is based on (possibly much) less than $t - \tau_i(t)$ samples from arm $i$, which makes false alarm even less likely to occur.
+Hence, it is easy to show that $(a) \leq \delta$.
+
+Term (b) is related to the control of the detection delay, which is more tricky to obtain under the \GLRklUCB{} adaptive sampling scheme,
+when compared to a result like Theorem~6 in \cite{Maillard2018GLR} for the change-point detector run in isolation.
+%
+More precisely, we need to leverage the forced exploration (Proposition~\ref{prop:6:EnoughSamples}) to be sure we have enough samples for detection.
+This explains why delays defined in Assumption~\ref{ass:6:LongPeriods} are scaled by $1/\alpha$.
+Using some elementary calculus and a concentration inequality given in Lemma~\ref{lem:6:Chernoff2arms}, we can finally prove that $(b) \leq \delta$.
+%
+Finally, the ``bad event'' is unlikely: $\bP(\cE_T^c) \leq 2\EffChange \delta$.
+
+%-----------------------------------------------------------------------------
 \subsection{Results for \GLRklUCB{} using Global Changes}
@@
     Define
-    % \begin{equation}\label{def:6:DelayGlobal}
-    $
+    \begin{equation}\label{def:6:DelayGlobal}
         d^{(k)} = d^{(k)}(\alpha,\delta) = \lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \rceil.
-    $
-    % \end{equation}
+    \end{equation}
     %
@@
     For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
-    \begin{small}
-\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.\]
-    \end{small}
+    \begin{equation}
+        R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
+    \end{equation}
 \end{theorem}
@@
 
+\subsection{Analysis of \GLRklUCB{} with Global Changes}\label{proof:6:mainRegretBoundGlobal}
+
+We gave in Section~\ref{sec:6:RegretAnalysis} the following decomposition of the regret $R_T$,
+\begin{eqnarray}\label{eq:6:GeneRegretBound}
+    R_T &\leq& T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
+    & & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
+\end{eqnarray}
+
+\subsubsection{Proof of Theorem~\ref{thm:6:mainRegretBoundGlobal}.}
+
+We first introduce some notation for the proof:
+let $\hat\tau^{(k)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\hat\tau(t)$ be the last time before $t$ that the algorithm restarted.
+We denote $n_i(t) = \sum_{s=\tau(t)+1}^{t}\indic(A_s = i)$ the number of selections of arm $i$ since the last (global) restart, and $\hat{\mu}_i(t) = \frac{1}{n_i(t)}\sum_{s=\tau(t)+1}^{t}X_{i,s}\indic(A_s = i)$ their empirical average (if $n_i(t) \neq 0$).
+
+As explained before, our analysis relies on the general regret decomposition~\eqref{eq:6:GeneRegretBound}, with the following appropriate good event.
+Let $d^k$ be defined as in Assumption~\ref{ass:6:LongPeriodsGlobal}, we define
+\begin{equation}\cE_T(\delta) = \left( \forall k \in \{1, \ldots, \Upsilon_T\}, \hat{\tau}^{(k)} \in \left[\tau^{(k)} + 1, \tau^{(k)} + d^{(k)}\right] \right).\label{def:6:GoodEvenGlobal}\end{equation}
+Under the good event, all the change points are detected within a delay at most $d^(k)$.
+%
+Note that from Assumption~\ref{ass:6:LongPeriodsGlobal}, as the period between two changes are long enough, if $\cE_T(\delta)$ holds, then for all change $k$, one has $\tau^{(k)} \leq \hat{\tau}^{(k)} \leq \tau^{(k+1)}$.
+Using Assumption~\ref{ass:6:LongPeriods}, one can prove the following.
+
+\begin{lemma}\label{lem:6:GoodEventGlobal}
+    With $\cE_T(\delta)$ defined as in \eqref{def:6:GoodEvenGlobal}, the ``bad event'' is unlikely:
+    $\bP(\cE_T^c(\delta)) \leq \delta(K+1)\Upsilon_T$.
+\end{lemma}
+
+We now turn our attention to upper bounding the two terms $(A)$ and $(B)$ in \eqref{eq:6:GeneRegretBound}.
+
+\paragraph{Upper bound on Term $\bm{(A)}$}
+
+\begin{align*}
+    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \, \kl\left(\hat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau(t))\right)\right] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau(t))\right)\right] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(k)} + \indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau^{(k)})\right)\right] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d^{(k)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \hat\tau^{(k)})\right)\right],
+\end{align*}
+%
+where we introduce the event $\cC^{(k)}$ that all the changes up to the $k$-th have been detected:
+%
+\begin{equation}\label{def:6:EventCiGlobal}
+    \cC^{(k)} = \left\{\forall j \leq k, \hat\tau^{(j)} \in \{\tau^{(j)} + 1, \dots, \tau^{(j)} + d^{(j)} \} \right\}.
+\end{equation}
+%
+Clearly, $\cE_T \subseteq \cC^{(k)}$ and $\cC^{(k)}$ is $\cF_{\hat\tau^{(k)}}$-measurable.
+Observe that conditionally to $\cF_{\hat\tau^{(k)}}$, when $\cC^{(k)}$ holds, $\hat{\mu}_{k^*}(t)$ is the average of samples that have all mean $\mu_{k^*}^{(k)}$.
+%
+Thus, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_{k^*}^{(k)}$, one can write
+%
+\begin{align*}
+    &\bE\left[\left.\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau^{(k)})\right) \right| \cF_{\hat\tau^{(k)}}\right]\\
+    & = \indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau^{(k)})\right) \;|\; \cF_{\hat\tau^{(k)}}\right] \\
+    & \leq \indic(\cC^{(k)}) \sum_{t'=1}^{\tau^{(k+1)} - \hat{\tau}^{(k)}} \bP\left(\exists s \leq t' : s \, \kl(\hat{\mu}_{s},\mu_{k^*}^{(k)}) \geq f(t')\right) \\
+    & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
+\end{align*}
+%
+where the last but one inequality relies on the concentration inequality given in Lemma 2 of \cite{KLUCBJournal} and the fact that $f(t) = \ln(t) + 3 \ln(\ln(t))$.
+Finally, using the law of total expectation yields
+\begin{equation}\label{eq:6:TermAFinalGlobal}
+    (A) \leq \sum_{k=1}^{\Upsilon_T}\left[d^{(k)} + \ln(\ln(T))\right].
+\end{equation}
+
+\paragraph{Upper bound on Term $\bm{(B)}$.} We let $\tilde{\mu}_{i,s}^{(k)}$ denote the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
+Rewriting the sum in $t$ as the sum of consecutive intervals $[\tau^{(k)}+1, \tau^{(k+1)}]$,
+\begin{align*}
+    (B) & \leq \bE\Big[\indic(\cE_T)\sum_{k=1}^{\Upsilon_T} \sum_{t=\tau^{(k)}}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{I_t}^{(k)}\right)\indic\left(\UCB_{I_t}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\hat{\tau}^{(k)} + \indic(\cE_T) \sum_{t=\hat\tau^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{I_t}^{(k)}\right)\indic\left(\UCB_{I_t}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\hat\tau^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\indic\left(I_t=i,\UCB_{i}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\times \\
+     &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\sum_{s=1}^{t-\hat\tau^{(k)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \hat\tau^{(k)})\right)\Big] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_i(\tau^{(k+1)})}\indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\Big]  \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\bE\Big[\indic(\cC^{(k)}) \sum_{s=1}^{n_i(\tau^{(k+1)})}\indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\Big].
+\end{align*}
+
+Conditionally to $\cF_{\hat\tau^{(k)}}$, when $\cC^{(k)}$ holds, for $s \in \{1, \ldots, n_i(\tau^{(k+1)})\}$, $\tilde{\mu}_{i,s}^{(k)}$ is the empirical mean from \iid{} observations of mean $\mu_i^{(k)}$.
+Therefore, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(k)}$, it follows from the law of total expectation that
+\begin{align*}
+    (B)& \leq \sum_{k=1}^{\Upsilon_T}d^{(k)}+\sum_{i=1}^K\sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right) \sum_{s=1}^{\tau^{(k+1)}-\tau^{(k)}}\bP\left(s \times \kl(\hat{\mu}_{s}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\cdot
+\end{align*}
+If $\mu_{k^*}^{(k)} > \mu_i^{(k)}$ by definition, we can use the analysis of \klUCB{} from \cite{KLUCBJournal} to further upper bound the right-most part, and we obtain
+\begin{align}\label{eq:6:TermBFinalGlobal}
+    (B) & \leq \sum_{k=1}^{\Upsilon_T}d^{(k)}+\sum_{i=1}^K\sum_{k=1}^{\Upsilon_T}\indic\left(\mu_i^{(k)} \neq \mu_{k^*}^{(k)}\right)\left[ \frac{\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)}{\kl(\mu_i^{(k)},\mu_{k^*}^{(k)})}\log(T) + \bigO{\sqrt{\log(T)}}\right].
+\end{align}
+
+Combining the regret decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEventGlobal} and the two upper bounds in \eqref{eq:6:TermAFinalGlobal} and \eqref{eq:6:TermBFinalGlobal},
+\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
+which concludes the proof.
+
+\subsubsection{Controlling the probability of the good event: Proof of Lemma~\ref{lem:6:GoodEventGlobal}}
+
+% We introduce $\hat{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after $t=\tau^{(k)}+1$ and $\tilde{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
+
+Recall that $\cC^{(k)}$ defined in \eqref{def:6:EventCiGlobal} is the event that all the breakpoints up to the $k$-th have been correctly detected.
+Using a union bound, one can write
+%
+\begin{align*}
+    \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T}
+    \bP\left(\left. \hat\tau^{(k)} \notin \{ \tau^{(k)}+1, \dots, \tau^{(k)} + d^{(k)} \} \right|\; \cC^{(k-1)}\right) \\
+    \\ & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\hat\tau^{(k)} \leq \tau^{(k)} \;|\; \cC^{(k-1)}\right)}_{(a)} + \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\hat\tau^{(k)} \geq \tau^{(k)} + d^{(k)} \;|\; \cC^{(k-1)}\right)}_{(b)}.
+\end{align*}
+%
+The final result follows by proving that $(a) \leq K \delta$ and $(b)\leq \delta$, as detailed below.
+
+\paragraph{Upper bound on $(a)$: controlling the false alarm.}
+%
+$\hat\tau^{(k)} \leq \tau^{(k)}$ implies that there exists an arm whose associated change point detector has experienced a false-alarm:
+%
+\begin{align*}
+    (a) & \leq \bP\left(\exists i, \exists s < t \leq n_i(\tau_i^{(k)}) : s \, \kl\left(\tilde{\mu}_{i,1:s}^{(k-1)},\tilde{\mu}_{i,1:t}^{(k-1)}\right) + (t - s) \, \kl\left(\tilde{\mu}_{i,s+1:t}^{(k-1)},\tilde{\mu}_{i,1:t}^{(k-1)}\right) > \beta(t, \delta) \;|\; \cC^{(k-1)}\right) \\
+    & \leq \sum_{i=1}^K\bP\left(\exists s < t : s \, \kl(\hat{\mu}_{1:s},\mu_i^{(k-1)}) + (t - s) \, \kl(\hat{\mu}_{s+1:t}, \mu_{i}^{(k-1)}) > \beta(t, \delta)\right),
+\end{align*}
+with $\hat\mu_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_i^{(k-1)}$.
+Indeed, conditionally to $\cC^{(k-1)}$, the $n_i(\tau^{(k)})$ successive observations of arm $i$ arm starting from $\hat \tau^{(k)}$ are \iid{} with mean $\mu_i^{(k-1)}$.
+Using Lemma~\ref{lem:6:ConcFirst}, term $(a)$ is upper bounded by $K\delta$.
+
+\paragraph{Upper bound on term (b): controlling the delay.}
+%
+From the definition of $\Delta^{(k)}$, there exists an arm $i$ such that $\Delta^{(k)} = |\mu_i^{(k)} - \mu_i^{(k-1)}|$.
+We shall prove that it is unlikely that the change-point detector associated to $i$ doesn't trigger within the delay $d^{(k)}$.
+
+First, it follows from Proposition~\ref{prop:6:EnoughSamples} that there exists $\overline{t} \in \{\tau^{(k)}, \dots, \tau^{(k)} + d^{(k)} \}$ such that
+$n_i(\overline{t}) - n_i(\tau^{(k)}) = \overline{r}$ where $\overline{r} = \lfloor \frac{\alpha}{K} d^{(k)}\rfloor$
+(as the mapping $t\mapsto n_i(t) - n_i(\tau^{(k)})$ is non-decreasing, is $0$ at $t=\tau^{(k)}$ and its value at $\tau^{(k)}+d^{(k)}$ is larger than $\overline{r}$).
+Using that \[\left(\hat\tau^{(k)} \geq \tau^{(k)} + d^{(k)}\right) \subseteq \left(\hat\tau^{(k)} \geq \overline{t}\right)\]
+the event $\left(\hat\tau^{(k)} \geq \tau^{(k)} + d^{(k)}\right)$ further implies that
+\[
+    n_i(\tau^{(k)}) \, \kl\left(\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)})},\tilde{\mu}^{k-1}_{i,n_i(\overline{t})}\right)
+    + \overline{r} \, \kl\left(\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)}) : n_i(\overline{t})},\tilde{\mu}^{k-1}_{i,n_i(\overline{t})}\right) \leq \beta(n_i(\tau^{(k)}) + \overline{r},\delta),
+\]
+where $\tilde{\mu}^{k-1}_{i,s}$ denotes the empirical mean of the $s$ first observation of arm $i$ since the $(k-1)$-th restart $\hat{\tau}^{(k-1)}$ and  $\tilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
+Conditionally to $\cC^{(k-1)}$, $\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)})}$ is the empirical mean of $n_i(\tau^{(k)})$ \iid{} replications of mean $\mu^{k-1}$, whereas $\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)}) : n_i(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_i^{k}$.
+
+Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_i(\tau^{(k)})$ lies in the interval $
+\left[\left\lfloor \frac{\alpha}{K}\left(\tau^{(k)}-\hat\tau^{(k-1)}\right)\right\rfloor,\left(\tau^{(k)}-\hat\tau^{(k-1)}\right)\right]$.
+Conditionally to $\cC^{(k-1)}$, one obtains furthermore using that $d^{(k-1)} \leq (\tau^{(k)} - \tau^{(k-1)})/2$ -- which follows from Assumption~\ref{ass:6:LongPeriodsGlobal} -- that
+\begin{eqnarray*}
+    n_i(\tau^{(k)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(k)}-\tau^{(k-1)} )\right\rfloor, \dots, \tau^{(k)}-\tau^{(k-1)} \right\} := \cI_k.
+\end{eqnarray*}
+
+Introducing $\hat{\mu}_{a,s}$ (resp.
+$\hat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\hat{\mu}_i^{(k-1)}$ (resp.
+$\hat{\mu}_i^{(k)}$), such that $\hat{\mu}_{a,s}$ and $\hat{\mu}_{b,r}$ are independent, it follows that
+\[
+    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\hat{\mu}_{a,s},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\hat{\mu}_{b,\overline{r}},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
+\]
+where we have also used that $\tilde{\mu}^{k-1}_{i,n_i(\overline{t})} = \left(n_i(\tau^{(k)})\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)})} + \overline{r}\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)}) : n_i(\overline{t})} \right) / (n_i(\tau^{(k)}) + \overline{r})$.
+
+Using Pinsker's inequality and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = | \Delta_i^k |$), one can write
+\begin{align}
+    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
+    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,s} - \Delta_i^{(k)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
+    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
+\end{align}
+%
+Using Lemma~\ref{lem:6:Chernoff2arms} (given above in Appendix~\ref{proof:6:Chernoff2arms}) and a union bound, the first term in the right hand side is upper bounded by $\delta$ (as $\beta(r+s,\delta) \geq \beta(r,\delta) \geq \log(3s\sqrt{s}/\delta)$).
+For the second term, we use the observation
+\[\frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(k)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
+and, using that $\Delta^{(k)} = | \Delta_i^k|$, one obtains
+\begin{equation}(b) \leq \delta + \bP\left(\exists s \in \cI_k : \Delta^{(k)} \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).\label{eq:6:FromHereGlobal}\end{equation}
+Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau^{(k)}-\tau^{(k-1)})/2\right\rfloor$.
+Using that the mappings $s \mapsto (s+\overline{r})/s\overline{r}$ and $s \mapsto \beta(s + \overline{r},\delta)$ are respectively decreasing and increasing in $s$, one has, for all $s\in \cI_k$,
+\begin{eqnarray*}
+    2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) & \leq &
+    2\frac{s_{\min}+\overline{r}}{s_{\min}\overline{r}}\beta\left(T, \delta\right) \leq \frac{4\beta(T,\delta)}{\left\lfloor \frac{\alpha}{K}d^{(k)}\right\rfloor},
+\end{eqnarray*}
+%
+where the last inequality follows from the fact that $\overline{r} \leq s_{\min}$ as $d^{(k)} \leq (\tau^{(k)} - \tau^{(k-1)})/2$ by Assumption~\ref{ass:6:LongPeriodsGlobal}.
+Now the definition of $d^{(k)}$ readily implies that
+\[\left\lfloor \frac{\alpha}{K}d^{(k)}\right\rfloor > \frac{4\beta(T,\delta)}{\left(\Delta^{(k)}\right)^2},\]
+which yields
+\[\forall s \in \cI_k, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta^{(k)}\right)^2.\]
+Hence, the probability in the right-hand side of \eqref{eq:6:FromHereGlobal} is zero, which yields $(b) \leq \delta$.
+
+% ----------------------------------------------------------------------------
+\subsubsection{Proof of Lemma~\ref{lem:6:FalseAlarm}} \label{proof:6:FalseAlarm}
+
+Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
+\begin{equation}
+    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
+\end{equation}
+where $\phi\mu(\lambda)$ is the log moment generating of some one-dimensional exponential family.
+The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
+
+Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}. A first key observation is
+\[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\hat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\lambda\right)\right].\]
+Hence the probability of a false alarm occurring is upper bounded as
+%
+\begin{align*}
+    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
+    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\hat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\hat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\hat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\hat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
+    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\hat{\mu}_{s},\mu_0\right) + r \, \kl\left(\hat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
+\end{align*}
+where $\hat{\mu}_{s}$ and $\hat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
+As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
+
+\begin{align*}
+    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu_0\right) + r \, \kl\left(\hat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu_0\right) + r \, \kl\left(\hat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
+\end{align*}
+And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
+
+\begin{lemma}\label{lem:6:ConcFirst}
+    $(X_i)_{i \in \N}$ and $(Y_i)_{i \in \N}$ two independent \iid{} processes with resp. means $\mu$ and $\mu'$ such that
+    \[\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)} \ \ \text{and} \ \ \bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)},\]
+    where $\phi_\mu(\lambda) = \bE_{X \sim \nu^{\mu}}[e^{\lambda X}]$ is the moment generating function of the distribution $\nu^\mu$, which is the unique distribution in an exponential family that has mean $\mu$.
+    %
+    Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family.
+Introducing the notation $\hat{\mu}_s = \frac{1}{s}\sum_{i=1}^s X_i$ and $\hat{\mu}'_r = \frac{1}{r}\sum_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
+    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu\right) + r \, \kl\left(\hat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
+    where $\cT$ is the function defined in \eqref{def:6:function_T}.
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:6:ConcFirst}.}
+Using the same construction as in the proof of Theorem~14 in \cite{KK18Martingales}, one can prove that for every $\lambda \in I$ (for an interval $I$), there exists a non-negative super-martingale $M^\lambda (s)$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ that satisfies $\bE[M^\lambda(s)] \leq 1$ and
+\[\forall s \in \N^*, \ \ M^\lambda (s) \geq e^{\lambda [s\kl(\hat{\mu}_s,\mu) - 3\ln(1+\ln(s))] - g(\lambda)}\]
+for some function $g : I \rightarrow \R$. This super-martingale is of the form
+\[M^\lambda (s) = \int e^{\eta\sum_{i=1}^s X_i - \phi_\mu(\lambda)s} d\pi(\eta)\]
+for a well-chosen probability distribution $\pi$, and the function $g$ can be chosen to be any
+\begin{eqnarray*}
+    g_{\xi} : \left[0; 1/(1+\xi)\right] & \longrightarrow & \R \\
+    \lambda & \mapsto & \lambda(1+\xi) \ln \left(\frac{\pi^2}{3(\ln(1+\xi))^2}\right) -  \ln(1 - \lambda(1 + \xi))
+\end{eqnarray*}
+for a parameter $\xi \in [0,1/2]$.
+
+Similarly, there exists an independent super-martingale $W^\lambda (r)$ w.r.t. the filtration $\cF_r' = \sigma(Y_1,\dots,Y_r)$ such that
+\[\forall r \in \N^*, \ \ W^\lambda (r) \geq e^{\lambda [r\kl(\hat{\mu}'_r,\mu) - 3\ln(1+\ln(r))] - g(\lambda)},\]
+for the same function $g(\lambda)$.
+In the terminology of \cite{KK18Martingales}, the processes $\bm{X}(s) = s \, \kl(\hat{\mu}_s,\mu) - 3\ln(1+\ln(s))$ and $\bm{Y}(s) = r \, \kl(\hat{\mu}_r,\mu) - 3\ln(1+\ln(r))$ are called $g$-DCC for Doob-Cram\'er-Chernoff, as Doob's inequality can be applied in combination with the Cram\'er-Chernoff method to obtain deviation inequalities that are uniform in time.
+
+Here we have to modify the technique used in their Lemma 4 in order to take into account the two stochastic processes, and the presence of super-martingales instead of martingales (for which Doob inequality still works).
+One can write
+\begin{align*}
+    &\bP\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu\right) + r \, \kl\left(\hat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \\
+    & \leq \bP\left(\exists r \in \N^* : M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right) \\
+    & = \lim_{n \rightarrow \infty} \bP\left(\exists r \in \{1,\dots,n\} :  M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right) \\
+    & = \lim_{n \rightarrow \infty} \bP\left(\sup_{r \in \{1,\dots,n\}}  M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right).
+\end{align*}
+%
+Using that $\tilde{M}(r) = M^\lambda(s)W^\lambda(r)$ is a super-martingale with respect to the filtration \[\tilde{\cF}_r = \sigma(X_1,\dots,X_s, Y_1,\dots,Y_r),\] one can apply Doob's maximal inequality to obtain
+\begin{eqnarray*}
+    \bP\left(\sup_{r \in \{1,\dots,n\}}  M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right) &\leq& e^{-(\lambda u - 2g(\lambda))}\bE[\tilde{M}(1))] \\
+    & = & e^{-(\lambda u - 2g(\lambda))}\bE[M^\lambda(s)W^\lambda(1)] \\
+    & \leq & e^{-(\lambda u - 2g(\lambda))},
+\end{eqnarray*}
+using that $M^\lambda(s)$ and $W^\lambda(1)$ are independent and have an expectation smaller than $1$.
+
+Putting things together yields
+\[\bP\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu\right) + r \, \kl\left(\hat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \leq e^{-\left(\lambda u - 2g_{\xi}(\lambda)\right)},\]
+for any function $g_{\xi}$ defined above.
+%
+The conclusion follows by optimizing for both $\lambda$ and $\xi$, using Lemma~18 in \cite{KK18Martingales}.
+
+% ----------------------------------------------------------------------------
+
 % \TODOE{note that in this corollary $\Delta^{\text{change}}$ could be defined as the smallest value of $\Delta^{(k)}$ which can be larger than the smallest change on any arm...
@@
     Define
-    % \begin{equation}\label{def:6:DelayLocal}
-    $
+    \begin{equation}\label{def:6:DelayLocal}
         d_i^{(\ell)} = d_i^{(\ell)}(\alpha,\delta) = \lceil \frac{4K}{\alpha\left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha}\rceil.
-    $
-    % \end{equation}
+    \end{equation}
     %
@@
     %
-    \begin{small}
-    \[R_T \leq 2\sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{4K}{\alpha \left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},\]
-    \end{small}
+    \begin{equation}
+        R_T \leq 2\sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{4K}{\alpha \left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
+    \end{equation}
     where
-    ${\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}$.
+    \begin{equation}
+        {\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}.
+    \end{equation}
 \end{theorem}
@@
 
+\TODOL{Only two variants in this corollary, not the four variants?}
+
+% ------------------------------------------------------------------------------------
+\subsection{Analysis of \GLRklUCB{} with Local Changes}\label{proof:6:mainRegretBound}
+
+\subsubsection{Proof of Theorem~\ref{thm:6:mainRegretBound}}
+
+Our analysis relies on the general regret decomposition~\eqref{eq:6:GeneRegretBound}, with the following appropriate good event.
+\begin{equation}\label{def:6:GoodEvenLocal}
+    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
+\end{equation}
+where $\hat\tau_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined as in Assumption~\ref{ass:6:LongPeriods}.
+Using this assumption, one can prove the following.
+
+\begin{lemma}\label{lem:6:GoodEvent}
+    The ``bad event'' is highly unlikely, as it satisfies $\bP((\cE_T(\alpha,\delta))^c) \leq 2\delta \EffChange$.
+\end{lemma}
+
+We now turn our attention to upper bounding terms $(A)$ and $(B)$ in \eqref{eq:6:GeneRegretBound}.
+
+\paragraph{Upper bound on term (A).}
+%
+\begin{align*}
+    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \kl\left(\hat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau_{i_t^*}(t))\right)\right] \\
+    & \leq \sum_{i=1}^K \sum_{k=1}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic(i_t^* =i) \indic(\tau_i(t) = \hat\tau_i^{(\ell)})\indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right)\right] \\
+    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right) + d_i^{(\ell)}(T)\right] \\
+    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}d_i^{(\ell+1)}(T) + \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}}\indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right)\right],
+\end{align*}%
+%
+where we introduce the event $\cC_i^{(\ell)}$ that all the changes up to the $\ell$-th have been detected:
+\begin{equation}\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \hat\tau_i^{(j)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(\ell)} \right] \right\}.\label{def:6:EventCi}\end{equation}
+Clearly, $\cE_T \subseteq \cC_i^{(\ell)}$ and $\cC_i^{(\ell)}$ is $\cF_{\hat\tau_i^{(\ell)}}$-measurable.
+Observe that conditionally to $\cF_{\hat\tau_i^{(\ell)}}$, when $\indic(\cC_i^{(\ell)})$ holds, $\hat{\mu}_{i}(t)$ is the average of samples that have all mean $\mu_i^{(\ell)}$.
+Thus, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(\ell)}$, one can write
+\begin{align*}
+    &\bE\left[\left.\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right) \right| \cF_{\hat\tau_i^{(\ell)}}\right]\\
+    & = \indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \bE\left[\indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right) \;|\; \cF_{\hat\tau_i^{(\ell)}}\right]
+% \end{align*}%
+% \begin{align*}
+    & \leq \indic(\cC_i^{(\ell)}) \sum_{t=1}^{\tau_i^{(\ell+1)} - \hat{\tau}_i^{(\ell)}} \bP\left(\exists s \leq t' : s \times \kl(\hat{\mu}_{s},\mu_i^{(\ell)}) \geq f(t')\right) \\
+    & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
+\end{align*}%
+%
+where the last but one inequality relies on the concentration inequality given in Lemma 2 of \cite{KLUCBJournal}, and the fact that $f(t) = \ln(t) + 3 \ln(\ln(t))$.
+Finally, using the law of total expectation yields
+\begin{equation}
+    (A) \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\left[d_i^{(\ell+1)}(T) + \ln(\ln(T))\right].
+    \label{eq:6:TermAFinal}
+\end{equation}
+
+\paragraph{Upper bound on term (B).}
+%
+Recall that $\mu^*_{i,\ell}$ is defined from the statement of Theorem~\ref{thm:6:mainRegretBound} as the smallest value that still outperforms arm $i$ on the the interval between the $\ell$ and $(\ell+1)$-st change of $i$.
+We let $\tilde{\mu}_{i,s}^{(\ell)}$ denote the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}_i^{(\ell)}+1$.
+To upper bound Term B, we introduce a sum over all arms and rewrite the sum in $t$ as a sum of consecutive intervals $[\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]$.
+The decomposition using furthermore that on $\cE_T$, $\tau_i^{\ell} \leq \hat \tau_i^ \ell \leq \tau_i^{\ell+1}$ for all changes, by Assumption~\ref{ass:6:LongPeriods}.
+%
+\begin{align*}
+    (B) & \leq \sum_{i=1}^K\bE\Big[\indic(\cE_T)\sum_{\ell=1}^{\NCi} \sum_{t=\tau_i^{(\ell)}}^{\tau_i^{(\ell+1)}}\left(\mu_{i^*_t}(t) - \overline{\mu}_{i}^{(\ell)}\right)\indic\left(I_t = i, \UCB_i(t) \geq \mu^*_{i,\ell}\right)\Big] \\
+    & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T)\hat{\tau}_i^{(\ell)} + \indic(\cE_T) \sum_{t=\hat\tau_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\indic\left(I_t = i, \UCB_i(t) \geq \mu^*_{i,\ell}\right)\Big] \\
+%     & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+ \sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\sum_{s=1}^{t-\hat\tau_i^{(\ell)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \times \kl(\tilde{\mu}_{i,s}^{(\ell)}, \mu^*_{\ell,i}) \leq f(\tau_i^{(\ell+1)} - \hat\tau_i^{(\ell)})\right)\Big] \\
+    & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+ \sum_{i=1}^K\sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cC_i^{(\ell)}) \sum_{s=1}^{n_i(\tau_i^{(\ell+1)})}\indic\left(s \times \kl(\tilde{\mu}_{i,s}^{(\ell)}, \mu^*_{i,\ell}) \leq f(\tau_i^{(\ell+1)} - \tau_i^{(\ell)})\right)\Big].
+\end{align*}%
+%
+The last inequality relies on introducing a sum over $\indic(n_i(t) = s)$ and swapping the sums.
+Conditionally to $\cF_{\hat\tau_i^{(\ell)}}$, when $\cC_i^{(\ell)}$ holds, for $s \in \{1, \ldots, n_i(\tau_i^{(\ell+1)})\}$, $\tilde{\mu}_{i,s}^{(\ell)}$ is the empirical mean from \iid{} observations of mean $\overline{\mu}_i^{(\ell)}$.
+Therefore, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\overline{\mu}_i^{(\ell)}$, it follows from the law of total expectation that
+%
+\begin{align*}
+    (B) & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi} d_i^{(\ell)}
+    + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \sum_{s=1}^{\tau_i^{(\ell+1)}-\tau_i^{(\ell)}} \bP\left(s \times \kl(\hat{\mu}_{s}, \mu^*_{i,\ell}) \leq f(\tau_i^{(\ell+1)} - \tau_i^{(\ell)})\right).
+\end{align*}%
+%
+As $\mu^*_{\ell,i} > \mu_i^{(\ell)}$ by definition, we can use the same analysis as in the proof of Fact~2 in Appendix~A.2 of \cite{KLUCBJournal} to show that
+% \begin{align*}
+%     (B)_i & \leq \sum_{\ell=1}^{\NCi}d_i^{(\ell)}+\sum_{\ell=1}^{\NCi}\left[ \frac{\ln(T) + 3\ln\ln(T)}{\kl(\mu_i^{(\ell)}, \mu^*_{\ell,i})} + \sqrt{\frac{2\pi(\kl'(\mu_i^{(\ell)},\mu^*_{\ell,i}))^2}{(\kl(\mu_i^{(\ell)}, \mu^*_{\ell,i}))^2}}\sqrt{\ln(T) + 3\ln\ln(T)} + 2\left(\frac{\kl'(\mu_i^{(\ell)},\mu^*_{\ell,i})}{\kl(\mu_i^{(\ell)}, \mu^*_{\ell,i})}\right)^2\right].
+% \end{align*}
+\begin{align}\label{eq:6:TermBFinal}
+    (B) & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+\sum_{i=1}^K\sum_{\ell=1}^{\NCi}\left[ \frac{\ln(T) + 3\ln\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},\mu^*_{i,\ell})} + O\left(\sqrt{\log(T)}\right)\right].
+\end{align}
+The result follows by combining the decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEvent} and the bounds \eqref{eq:6:TermAFinal} and \eqref{eq:6:TermBFinal}.
+
+\subsubsection{Proof of Lemma~\ref{lem:6:GoodEvent}}\label{proof:6:GoodEvent}
+
+With the event $\cC_i^{(\ell)}$ defined in \eqref{def:6:EventCi}, a simple union bound yields
+\begin{align*}
+    \bP(\cE_T^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
+\end{align*}
+
+The final result follows by proving that the terms $(a)$ and $(b)$ are both upper bounded by $\delta$.
+
+\paragraph{Upper bound on $(a)$: controlling the false alarms.}
+Under the bandit algorithm, the change point detector associated to arm $i$ is based on (possibly much) less than $t - \tau_i(t)$ samples from arm $i$, which makes false alarm even less likely to occur.
+More precisely, we upper bound term $(a)$ by
+%
+\begin{align*}
+    (a) & \leq \bP\left(\exists s < t \leq n_i(\tau_i^{(\ell)}) : s \times \kl\left(\tilde{\mu}_{i,1:s}^{(\ell-1)},\tilde{\mu}_{i,1:t}^{(\ell-1)}\right) + (t - s) \times \kl\left(\tilde{\mu}_{i,s+1:t}^{(\ell-1)},\tilde{\mu}_{i,1:t}^{(\ell-1)}\right) > \beta(t, \delta) \;|\; \cC_i^{(\ell-1)}\right) \\
+    & \leq \bP\left(\exists s < t : s \times \kl(\hat{\mu}_{1:s},\mu_i^{(\ell-1)}) + (t - s) \times \kl(\hat{\mu}_{s+1:t}, \mu_{i}^{(\ell-1)}) > \beta(t, \delta)\right),
+\end{align*}
+with $\hat\mu_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_i^{(\ell-1)}$.
+Indeed, conditionally to $\cC_i^{(\ell-1)}$, the $n_i(\tau_i^{(\ell)})$ successive observations of arm $i$ arm starting from $\hat \tau_i^{(\ell)}$ are \iid{} with mean $\mu_i^{(\ell-1)}$.
+Using Lemma~\ref{lem:6:ConcFirst}, term $(a)$ is upper bounded by $\delta$.
+
+\paragraph{Upper bound on term $(b)$: controlling the delay.}
+%
+Controlling the detection delay on arm $i$ under an adaptive sampling scheme can be tricky.
+Here we need to leverage the forced exploration (Proposition~\ref{prop:6:EnoughSamples}) to be sure we have enough samples to ensure detection: the effect is that delays will be scaled by the exploration parameter $\alpha$.
+
+First, it follows from Proposition~\ref{prop:6:EnoughSamples} that there exists $\overline{t} \in \left\{\tau_i^{(\ell)}, \dots, \tau_i^{(\ell)} + d_i^{(\ell)} \right\}$ such that
+$n_i(\overline{t}) - n_i(\tau_i^{(\ell)}) = \overline{r}$ where $\overline{r} = \lfloor \frac{\alpha}{K} d_i^{(\ell)}\rfloor$
+(as the mapping $t\mapsto n_i(t) - n_i(\tau_i^{(\ell)})$ is non-decreasing, is $0$ at $t=\tau_i^{(\ell)}$ and its value at $\tau_i^{(\ell)}+d_i^{(\ell)}$ is larger than $\overline{r}$).
+Using that $(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)}) \subseteq (\hat\tau_i^{(\ell)} \geq \overline{t})$,
+the event $(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)})$ further implies that
+\[
+    n_i(\tau_i^{(\ell)}) \, \kl\left(\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)})},\tilde{\mu}^{\ell-1}_{i,n_i(\overline{t})}\right)
+    +  \overline{r} \, \kl\left(\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)}) : n_i(\overline{t})},\tilde{\mu}^{\ell-1}_{i,n_i(\overline{t})}\right) \leq \beta(n_i(\tau_i^{(\ell)}) + \overline{r},\delta),
+\]
+where $\tilde{\mu}^{\ell-1}_{i,s}$ denotes the empirical mean of the $s$ first observation of arm $i$ since the $(\ell-1)$-th restart $\hat{\tau}_i^{(\ell-1)}$ and  $\tilde{\mu}^{\ell-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
+Conditionally to $\cC^{(\ell-1)}_i$, $\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)})}$ is the empirical mean of $n_i(\tau_i^{(\ell)})$ \iid{} replications of mean $\mu_i^{\ell-1}$, whereas $\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)}) : n_i(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_i^{\ell}$.
+
+Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_i(\tau_i^{(\ell)})$ lies in $
+\left\{\left\lfloor \frac{\alpha}{K}\left(\tau_i^{(\ell)}-\hat\tau_i^{(\ell-1)}\right)\right\rfloor, \dots, \tau_i^{(\ell)}-\hat\tau_i^{(\ell-1)} \right\}$.
+%
+Conditionally to $\cC_i^{(\ell-1)}$, one obtains furthermore using that $d_i^{(\ell-1)} \leq (\tau_i^{(\ell)} - \tau_i^{(\ell-1)})/2$ -- which follows from Assumption~\ref{ass:6:LongPeriods} -- that
+\begin{eqnarray*}
+    n_i(\tau_i^{(\ell)})& \in &  \left\{ \left\lfloor \frac{\alpha}{K}\left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)} - d_i^{(\ell-1)}\right)\right\rfloor, \dots, \left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)}\right) \right\} \\
+    n_i(\tau_i^{(\ell)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K}\left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)}\right)\right\rfloor, \dots, \left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)}\right) \right\} = \cI_k.
+\end{eqnarray*}
+
+Introducing $\hat{\mu}_{a,s}$ (resp.
+$\hat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\hat{\mu}_i^{(\ell-1)}$ (resp.
+$\hat{\mu}_i^{(\ell)}$), such that $\hat{\mu}_{a,s}$ and $\hat{\mu}_{b,r}$ are independent, it follows that
+\[
+    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\hat{\mu}_{a,s},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\hat{\mu}_{b,\overline{r}},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
+\]
+where we have also used that $\tilde{\mu}^{\ell-1}_{i,n_i(\overline{t})} = \frac{n_i(\tau_i^{(\ell)})\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)})} + \overline{r}\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)}) : n_i(\overline{t})}}{n_i(\tau_i^{(\ell)}) + \overline{r}}$.
+
+Using Pinsker's inequality and introducing the gap $\Delta_i^{(\ell)} = \mu_i^{(\ell-1)} - {\mu}_i^{(\ell)}$, one can write
+\begin{align}
+    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
+    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,s} - \Delta_i^{(\ell)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
+    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
+\end{align}
+%
+
+Using Lemma~\ref{lem:6:Chernoff2arms} stated in Appendix~\ref{proof:6:Chernoff2arms} and a union bound, the first term in the right hand side is upper bounded by $\delta$ (as $\beta(r+s,\delta) \geq \beta(r,\delta) \geq \log(3s\sqrt{s}/\delta)$).
+For the second term, we use the observation
+\[\frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(\ell)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
+and finally get
+\begin{equation}(b) \leq \delta + \bP\left(\exists s \in \cI_k : |\Delta_i^{(\ell)}| \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).\label{eq:6:FromHere}\end{equation}
+Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau_i^{(\ell)}-\tau_i^{(\ell-1)})/2\right\rfloor$.
+Using that the mappings $s \mapsto (s+\overline{r})/s\overline{r}$ and $s \mapsto \beta(s + \overline{r},\delta)$ are respectively decreasing and increasing in $s$, one has, for all $s\in \cI_k$,
+\begin{eqnarray*}
+    2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) & \leq &
+    2\frac{s_{\min}+\overline{r}}{s_{\min}\overline{r}}\beta\left(T, \delta\right) \leq \frac{4\beta(T,\delta)}{\left\lfloor \frac{\alpha}{K}d_i^{(\ell)}\right\rfloor},
+\end{eqnarray*}
+%
+where the last inequality follows from the fact that $\overline{r} \leq s_{\min}$ as $d_i^{(\ell)} \leq (\tau_i^{(\ell)} - \tau_i^{(\ell-1)})/2$ by Assumption~\ref{ass:6:LongPeriods}.
+Now the definition of $d_i^{(\ell)}$ readily implies that
+\[\left\lfloor \frac{\alpha}{K}d_i^{(\ell)}\right\rfloor > 4\beta(T,\delta) / \left(\Delta_i^{(\ell)}\right)^2,\]
+which yields
+\[\forall s \in \cI_k, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta_i^{(\ell)}\right)^2.\]
+Hence, the probability in the right-hand side of \eqref{eq:6:FromHere} is zero, which yields $(b) \leq \delta$.
+
 \subsection{Interpretation}
@@
 
-\section{A Unified Regret Analysis}
-\label{sec:6:RegretAnalysis}
-
-\TODOL{Rewrite, include the proof of at least one case here! No need to be that quick for the thesis, we only wrote it like this for the COLT version \cite{Besson2019GLRT}.}
-
-In this section, we sketch a unified proof for Theorem~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound}, whose detailed proofs can be found in Appendix~\ref{proof:6:mainRegretBoundGlobal} and \ref{proof:6:mainRegretBound} respectively.
-We emphasize that our approach is significantly different from those proposed by \cite{CaoZhenKvetonXie18} for \MUCB{} and by \cite{LiuLeeShroff17} for \CUSUMUCB.
-
-Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
-Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
-\begin{small}
-\begin{align*}
- &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
- & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
-\end{align*}%
-\end{small}%
-%
-Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
-%
-\begin{footnotesize}
-\begin{eqnarray*}
-R_T \leq T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} + \ \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
-\end{eqnarray*}%
-\end{footnotesize}%
-%
-Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay.
-Each regret bound then follows from upper bounds on term (A), term (B), and on the failure probability $\bP(\cE_T^c)$.
-%
-To control (A) and (B), we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
-
-The tricky part of each proof, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$.
-For example for local changes (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
-\[
-    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
-\]
-where $\hat\tau_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
-Introducing the event $\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \hat\tau_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\}$ that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition:
-\begin{align*}
-    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
-\end{align*}
-
-Term (a) is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
-Observe that under the bandit algorithm, the change point detector associated to arm $i$ is based on (possibly much) less than $t - \tau_i(t)$ samples from arm $i$, which makes false alarm even less likely to occur.
-Hence, it is easy to show that $(a) \leq \delta$.
-
-Term (b) is related to the control of the detection delay, which is more tricky to obtain under the \GLRklUCB{} adaptive sampling scheme,
-when compared to a result like Theorem~6 in \cite{Maillard2018GLR} for the change-point detector run in isolation.
-%
-More precisely, we need to leverage the forced exploration (Proposition~\ref{prop:6:EnoughSamples}) to be sure we have enough samples for detection.
-This explains why delays defined in Assumption~\ref{ass:6:LongPeriods} are scaled by $1/\alpha$.
-Using some elementary calculus and a concentration inequality given in Lemma~\ref{lem:6:Chernoff2arms}, we can finally prove that $(b) \leq \delta$.
-%
-Finally, the ``bad event'' is unlikely: $\bP(\cE_T^c) \leq 2\EffChange \delta.$
-%The detailed proof is provided in Appendix~\ref{proof:6:GoodEvent}.
-
 % ----------------------------------------------------------------------------
@@
 For simplicity, we restrict to rewards generated from Bernoulli distributions, even if \GLRklUCB{} can be applied to any bounded distributions.
-% %
-% Note that our proposal can also be applied to unbounded distributions, with a different divergence function used for the \GLR{} test (but results are yet to be obtained).
-% \cite{Maillard2018GLR} analyzed the \GLR{} test, without considering the bandit setting, for Gaussian and $\sigma$-sub Gaussian distributions.
+
+We report results obtained on five different piece-wise stationary bandit problems, illustrated in Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}, \ref{fig:6:Problem_5}, \ref{fig:6:Problem_4} and \ref{fig:6:Problem_6}, and described in more details below.
+Results are reported with regret tables and regret plots.
+For each experiment, the regret was estimated using $1000$ independent runs.
 
@@
 
-% -----------------------------------------------------------------
-\paragraph{Results.}
-\label{sub:6:NumericalResults}
-
-We report results obtained on three different piece-wise stationary bandit problems, illustrated in Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2} and \ref{fig:6:Problem_5} in Appendix~\ref{app:6:Figures}, and described in more details below.
-Additional results on two more problems can be found in Appendix~\ref{sub:6:TwoMoreProblems}, where additional regret curves can be found for all problems.
-For each experiment, the regret was estimated using $1000$ independent runs.
-
-% %
-% We give in Appendix~\ref{app:6:NumericalResults} the plots of the cumulated mean regret $R_t$ (as a function of $t$), as well as plots of histograms of the distribution of final regret $R_T$, for the considered algorithms.
+\paragraph{Illustrations of our benchmark}\label{app:6:Figures}
+
+We give below the plots of the cumulated mean regret $R_t$ (as a function of $t$), as well as plots of histograms of the distribution of final regret $R_T$, for the considered algorithms.
 
@@
 
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
+    \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).}
+    \label{fig:6:Problem_1}
+\end{figure}
+
 % - Pb 2 changes are on all arms at a time
@@
 The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$}).
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
+    \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).}
+    \label{fig:6:Problem_2}
+\end{figure}
 
@@
 
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
+    \subcaption{\textbf{Problem $3$}: $K=6$ arms with $T=20000$, $\Upsilon=8$ changes occur on most arms at a time ($C=19$).
+        % Inspired from the Yahoo! example from \cite{CaoZhenKvetonXie18}.
+    }
+    \label{fig:6:Problem_5}
+\end{figure}
+
+%     \caption{History of means of arms for three problems, $\mu_i(t)$ for $i\in\{1,\dots,K\}$ and $t\in\{1,\dots,T\}$.}
+%     \label{fig:6:Problems_123}
+% \end{figure}
+
+% - Pb 4 changes are on all or almost arms at a time, but sequences don't have same length
+\paragraph{Problem 4.}
+
+Like problem $1$, it uses $K=3$ arms, $\Upsilon=4$ change-points and $T=5000$,
+but the stationary sequences between successive change-points no longer have the same length, as illustrated in Figure~\ref{fig:6:Problem_4}.
+Classical (stationary) algorithms such as \klUCB{} can be ``tricked'' by large enough stationary sequences, as they learn with a large confidence the optimal arm, and then fail to adapt to a new optimal arm after a change-point.
+We observe below in Table~\ref{table:6:totalResults2} that they can suffer higher regret when the change-points are more spaced out, as this problem starts with a longer stationary sequence of length $T/2$.
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
+    \caption{\textbf{Problem 4}: $K=3$ arms with $T=5000$, $\Upsilon=4$ changes occur on all arms at a time (\ie, $C=12$).}
+    \label{fig:6:Problem_4}
+\end{figure}
+
+% - Pb 6 is Yahoo! example from Figure~3 in CUSUM-UCB paper \cite{LiuLeeShroff17}
+\paragraph{Problem 5.}
+
+Like problem $3$, this hard problem is inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!} (see Figure~3 from \cite{LiuLeeShroff17}).
+It is harder, with $\Upsilon=81$ change-points on $K=5$ arms for a longer horizon of $T=100000$.
+Some arms change at almost every time steps, for a total number of breakpoints $C=179$, but the optimal arm is almost always the same one (\textcolor{red}{arm $0$, with $\bullet$}).
+It is a good benchmark to see if the actively adaptive policies do not detect \emph{too many changes}, as the Oracle-Restart policy suffers higher regret in comparison to \klUCB.
+Means are also bounded in $[0.01, 0.07]$, with small gaps of amplitude ranging from $0.02$ to $0.001$,
+as shown in Figure~\ref{fig:6:Problem_6}.
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
+    \caption{\textbf{Problem 5}: $K=5$, $T=100000$, $\Upsilon=81$ changes occurring on some arms at every time ($C=179$).
+    }
+    \label{fig:6:Problem_6}
+\end{figure}
+
 % \TODOE{tu as bien tuné CUSUM-UCB avec $\epsilon = 0.02$ dans ce cas?}
@@
 
-Table \ref{table:6:totalResults1} shows the final regret $R_T$ obtained for each algorithm.
+% -----------------------------------------------------------------
+\paragraph{Results.}
+\label{sub:6:NumericalResults}
+
+% \subsubsection{Results in terms of final regret $R_T$}
+% \label{subsub:6:regretPlots}
+
+The two Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2} show the final regret $R_T$ obtained for each algorithm.
 Results highlighted in \textbf{bold} show the best non-oracle algorithm for each experiments,
@@
 
+\begin{table}[ht]
+    % \begin{small}
+    \centering
+    \begin{tabular}{c|cccccc}
+        \textbf{Algorithms} \textbackslash \textbf{Problems} & Pb $4$ ($T=5000$) & Pb $4$ ($T=10000$) & Pb $5$ \\
+        \hline
+        Oracle-Restart \klUCB{} & $\mathbf{68 \pm 40}$ & $\mathbf{86 \pm 50}$ & $126 \pm 54$ \\
+        \hline
+        \klUCB{} & $615 \pm 74$ & $1218 \pm 123$ & $106 \pm 36$ \\
+        SW-\klUCB{} & $202 \pm 33$ & $322 \pm 47$ & $228 \pm 27$ \\
+        Discounted-\klUCB{} & $911 \pm 210$ & $1741 \pm 200$ & $2085 \pm 910$ \\
+        \hline
+        Thompson sampling & $756 \pm 65$ & $1476 \pm 137$ & $\mathbf{88 \pm 39}$ \\
+        DTS & $250 \pm 39$ & $481 \pm 58$ & $238 \pm 24$ \\
+        \hline
+        \MklUCB{} & $337 \pm 46$ & $544 \pm 47$ & $116 \pm 36$ \\
+        \CUSUMklUCB{} & $267 \pm 69$ & $343 \pm 94$ & $117 \pm 34$ \\
+        \hline
+        \GLRklUCB{}(Local) & $\mathbf{99 \pm 32}$ & $\mathbf{128 \pm 42}$ & $149 \pm 34$ \\
+        \GLRklUCB{}(Global) & $128 \pm 32$ & $185 \pm 47$ & $152 \pm 32$
+    \end{tabular}
+    \caption{Mean regret $\pm$ $1$ std-dev.
+Problem $4$ use $K=3$ arms, and a first long stationary sequence.
+Problem $5$ use $K=5$, $T=20000$ and is much harder with $\Upsilon=82$ breakpoints and $C=179$ changes.}
+    \label{table:6:totalResults2}
+    % \end{small}
+\end{table}
+
 One can note that the best non-oracle strategies are actively adaptive,
@@
 
+\paragraph{Regret plots}
+\label{subsub:6:regretPlots}
+
+We also include below figures showing the simulation results for the five problems.
+The results in terms of mean regret $R_T$ are given above in Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2}, but it is also interesting to observe two plots for each experiments.
+%
+First, we show the mean regret as a function of time (\ie, $R_t$ for $1 \leq t \leq T$), for the $9$ algorithms considered (Discounted-\klUCB{} is removed as it is always the most inefficient and suffers high regret).
+%
+Efficient stationary algorithms, like TS and \klUCB, typically suffer a linear regret after a change on the optimal arm changes if they had ``too'' many samples before the change-points (\eg, on Figure~\ref{fig:6:meanRegretPb1} and even more on Figures~\ref{fig:6:meanRegretPb4} and \ref{fig:6:meanRegretPb5}).
+This illustrates the conjecture that classical algorithms can suffer linear regret even on simple piece-wise stationary problems.
+
+On very simple problems, like problem 1, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there is more changes, we clearly observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much more than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
+Our approach, with the two options of \textbf{Local} or \textbf{Global} restarts, performs very closely to the oracle for problem $4$.
+
+Finally, in the case of hard problems, like problems $3$ and $5$, that have a lot of changes but where the optimal arm barely changes, we verify in Figure~\ref{fig:6:meanRegretPb5} that \klUCB{} and TS can outperform the oracle policy.
+Indeed the oracle policy is suboptimal as it restarts as soon as one arm change but is unaware of the meaningful changes, and stationary policies which quickly identify the best arm will play it most of the times, achieving a smaller regret.
+We note that, sadly, all actively adaptive policies fail to outperform stationary policies on such hard problems, because they do not observe enough rewards from each arm between two restarts (\ie, the Assumptions~\ref{ass:6:LongPeriodsGlobal} and \ref{ass:6:LongPeriods} for our Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} are not satisfied).
+We can also verify that the two options, \textbf{Local} and \textbf{Global} restart, for \GLRklUCB, give close results, and that the \textbf{Local} option is always better.
+
+We also show the empirical distribution of the regret $R_T$, on Figure~\ref{fig:6:histogramRegretPb1}.
+It shows that all algorithms have a rather small variance on their regret, except Thompson Sampling which has a large tail due to its large mean regret on this (easy) non-stationary problems.
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
+    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 1.}
+    \label{fig:6:meanRegretPb1}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
+    \caption{Histograms of the distributions of regret $R_T$ ($T=5000$) for problem 1.}
+    \label{fig:6:histogramRegretPb1}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
+    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem $2$.}
+    \label{fig:6:meanRegretPb2}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K6_T20000_N500__9_algos_pb3/main____env1-1_8638317743626238996.pdf}
+    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 20000$) for problem 3.}
+    \label{fig:6:meanRegretPb3}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb4/main____env1-1_7471744535647872376.pdf}
+    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 4.}
+    \label{fig:6:meanRegretPb4}
+\end{figure}
+
+% -----------------------------------------------------------------
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K5_T100000_N500__9_algos_pb5/main____env1-1_950848333706043320.pdf}
+    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 100000$) for problem 5.}
+    \label{fig:6:meanRegretPb5}
+\end{figure}
+
 %----------------------------------------------------------------------------
-\section{Conclusion of Chapter 6}
-
-We proposed a new algorithm for the piece-wise stationary bandit problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
+\section{Conclusion}
+\label{sub:6:conclusion}
+
+In this chapter, we studied the piece-wise stationary bandit model, and reviewed existing work.
+We proposed a new algorithm for this problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
 This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but \emph{without any other prior knowledge on the problem}, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
@@
 
-We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this paper (and in previous work), that would require only a few ``meaningful'' changes to be detected.
+We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected.
 This interesting research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$.
@@
 
-\paragraph{Acknowledgments}
-Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.
+% \paragraph{Acknowledgments}
+% Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.
 
 \paragraph{A note on the simulation code}
+%
 The page \texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html} explains how to reproduce the experiments used for this Section.
-
-% % ----------------------------------------------------------------------------
-% \bibliographystyle{alpha}
-% \bibliography{biblioBandits}

### 2019-03-12 2-Chapters/6-Chapter/chapter6.tex
Done for now, I have to go for my teaching duties.

@@
 In this last chapter, we study a generalization of the stationary multi-armed bandit model,
-to account for possible non-stationarity of possible applications of MAB algorithms.
-Instead of dealing with changes happening possibly at every time step, like in adversarial models, we consider a finite number of \emph{change-points}, and the problem is stationary on any interval between two consecutive change-points (it is piece-wise stationary).
+to account for possible non-stationarity of MAB models.
+Instead of dealing with changes happening possibly at every time step, like in adversarial models, we consider a finite number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
 %
-We start by reviewing the existing previous work on piece-wise stationary MAB models, the main families of proposed solutions are either passively adaptive or actively adaptive.
+We start by reviewing the existing previous work on piece-wise stationary MAB models, for which the two main families of proposed solutions are either passively adaptive or actively adaptive.
 In practice, actively adaptive strategies usually greatly outperform the passively adaptive ones.
@@
 %
-We include additional details about our policy \GLRklUCB, including a sensibility analysis of its two parameters, a numerical evaluation of its time and memory costs as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
+We also include additional details about our policy \GLRklUCB, including a sensibility analysis of its two parameters, a numerical evaluation of its time and memory costs, as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
 

### 2019-03-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for now, I have to go for my teaching duties.

@@
 
-\TODOL{Be much more concise, if I already give an abstract in the beginning of the chapter...}
+% \TODOL{Be much more concise, if I already give an abstract in the beginning of the chapter...}
 
@@
 
-\TODOL{Be much more concise: in the thesis, MAB models and applications have already been introduced, motivated etc!}
+% \TODOL{Be much more concise: in the thesis, MAB models and applications have already been introduced, motivated etc!}
 
@@
 
-Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy (\cite{Thompson33}).
-More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, or parameters of a dynamically configurable radio hardware (\cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}).
+Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
+More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
 Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
@@
 
-A first possibility to cope with non-stationary is to model the decision making problem as an \emph{adversarial bandit problem} (\cite{Auer02NonStochastic}).
+A first possibility to cope with non-stationary is to model the decision making problem as an \emph{adversarial bandit problem} \cite{Auer02NonStochastic}.
 Under this model, rewards are completely arbitrary and are not assumed to follow any probability distribution.
@@
 For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
-% It is sometimes referred to as the abruptly changing MAB, or switching environments (\cite{MellorShapiro13}).
+It is sometimes referred to as the abruptly changing MAB, or switching environments \cite{MellorShapiro13}.
 %
@@
 In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB.
-Like previous approaches -- \CUSUM{} (\cite{LiuLeeShroff17}) and \MUCB{} (\cite{CaoZhenKvetonXie18}) -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector.
-For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 (\cite{Auer02}) used in previous works.
+Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector.
+For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 \cite{Auer02} used in previous works.
 For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest.
@@
 
-In this work we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
+In this chapter we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
 We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
@@
 
-\paragraph{Outline}
+\paragraph{Outline.}
 %
@@
 We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}.
-In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test (GLRT) as a Change-Point Detector (CPD) algorithm.
-We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bound on the regret of each variant.
-The unified regret analysis for these two algorithms is sketched in Section~\ref{sec:6:RegretAnalysis}.
-Numerical experiments
-are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.
+In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
+We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on the regret of both variants.
+The unified regret analysis for these two algorithms is given in Section~\ref{sec:6:RegretAnalysis}.
+Numerical experiments are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.
 
@@
 In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
-We define the $k$-th breakpoint by
+If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint by
 \begin{equation}
@@
 %
-Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, the rewards $(X_{i,t})$ associated to each arms are \iid.
+Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, that is on any stationary segments, the rewards $(X_{i,t})$ associated to each arms are \iid{} (hence the name of \emph{piece-wise stationary} problems).
 
 Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
-Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only a few arms change at each breakpoint.
-Introducing the number of change-points  on arm $i$, defined as $\NCi:=\sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
+Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
+For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fix channel or subset of channels, and on a particular day in a particular city, if the company deploys a lot of new objects, then one or some channels see their mean occupancy rates abruptly change.
+
+Introducing the number of change-points on arm $i$, defined as $\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
 $\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms.
@@
 
-\paragraph{An interesting generalization}
+\paragraph{An interesting interpolation.}
 %
@@
 Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
-Active strategies monitors the rewards of each arm by using a change detection algorithm (\cite{Basseville93}), and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
+Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
 Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.
@@
 %
-The RExp3 algorithm (\cite{Besbes14stochastic}) can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm.
+It can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm.
 Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
@@
 We distinguish two families,
-the first line of research uses frequentist change-point detectors (\cite{Basseville93}) combined with stationary policies, usually index policies like \UCB.
+the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
 When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
@@
 
-Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\alpha\in(0,1)$.
+Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
 \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
@@
 
-\paragraph{Advantages of our approach}
+\paragraph{Advantages of our approach.}
 %
@@
 Our proposal \GLRklUCB{} is inspired by both the \MUCB{} and \CUSUMUCB{} algorithms.
-Previous works focussed on using \UCB{} (\cite{LiuLeeShroff17,CaoZhenKvetonXie18}),
-but we propose to use \klUCB{} instead, as it is known to be more efficient for Bernoulli rewards as well as for a more generic case of bounded or one-dimensional exponential families (\cite{KLUCBJournal}).
+Previous works focussed on using \UCB{} \cite{LiuLeeShroff17,CaoZhenKvetonXie18},
+but we propose to use \klUCB{} instead, as it is known to be more efficient for Bernoulli rewards as well as for a more generic case of bounded or one-dimensional exponential families \cite{KLUCBJournal}.
 Theoretical results for these works were only given for \UCB, but they both suggest that extending the results to \klUCB{} (or other efficient stationary policies) should not be difficult.
-For a fair comparison, we therefore chose to compare the different Change-Point detector algorithms combined with \klUCB.
+For a fair comparison, we therefore chose to compare the different change-point detector algorithms combined with \klUCB.
 
 % -----------------------------------------------------------------
-\section{The Bernoulli GLR Change-Point Detector}
+\section{The Bernoulli GLRT Change-Point Detector}
 \label{sec:6:ChangePointDetector}
@@
 Assuming that  we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$.
-We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau \in\N^*$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
-%
-A sequential change-point detector is a stopping time $\hat\tau$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
-$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N, \bE[X_i] = \mu_0\right)$.
+We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
+%
+A sequential change-point detector is a stopping time $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
+$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N^*, \bE[X_i] = \mu_0\right)$.
 
@@
 
-\TODOL{We need to include and prove our results for false alarm, delay, etc, all results we squeezed and compressed for the COLT version \cite{Besson2019GLRT}.}
+\TODOL{We need to PROVE our results for false alarm and delay !.}
 
@@
 If the samples $(X_t)$ were all drawn from a Bernoulli distribution, our change-point detection problem would reduce to a parametric sequential test of
-$\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$
+\[\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0)),\]
 against the alternative
-$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
+\[\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1)).\]
+
 The Generalized Likelihood Ratio statistic for this test is defined by
-\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < t}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_t ; \mu_0)},\]
-%
-where $\ell(X_1, \ldots, X_t ; \mu_0)$ and $\ell(X_1, \ldots, X_t ; \mu_0,\mu_1,\tau)$ denote the likelihoods of the first $t$ observations under a model in $\cH_0$ and $\cH_1$.
-High values of this statistic tend to indicate rejection of $\cH_0$.
-Using the form of the likelihood for Bernoulli distribution, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
+\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},\]
+%
+where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
+High values of this statistic tend to indicate rejection of $\cH_0$ over $\cH_1$.
+%
+By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
 %
@@
 %
-Indeed, one can show that $\GLR(t) = \sup_{s \in [1,t]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:t}\right) + (t-s) \times \kl\left(\hat{\mu}_{s+1:t},\hat{\mu}_{1:t}\right)\right]$,
-where for $k \leq k'$, $\hat{\mu}_{k:k'}$ denotes the average of the observations collected between the instants $k$ and $k'$. This motivates the definition of the Bernoulli GLR change point detector.
+Indeed, one can show that $\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right]$,
+where for $k \leq k'$, $\widehat{\mu}_{k:k'} = \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
+%
+This motivates the following definition of the Bernoulli GLRT change point detector.
 
 \begin{definition}\label{def:6:GLRDef}
-    The Bernoulli GLR change point detector with threshold function $\beta(n,\delta)$ is
+    The Bernoulli GLRT (B-GLRT) change point detector with threshold function $\beta(n,\delta)$ is the stopping time $\widehat{\tau}_{\delta}$ defined by
     %
     \begin{equation}\label{def:6:GLR}
-        \hat\tau_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [1,t]} \left[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
+        \widehat{\tau}_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [1,t]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
     \end{equation}
@@
 
-Asymptotic properties of the GLR for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR} but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence between two elements in that exponential family that have mean $x$ and $y$.
+Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR} but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence between two elements in that exponential family that have mean $x$ and $y$.
 For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to \eqref{def:6:GLR} with $\kl(x,y) = 2(x-y)^2$ when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
 
-In this section, we provide new non-asymptotic results about the Bernoulli GLR test under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
-Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the Bernoulli GLR may stop earlier that the Gaussian GLR, based on the quadratic divergence $2(x-y)^2$.
+In this section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
+Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $2(x-y)^2$.
 
 % ----------------------------------------------------------------------------
-\subsection{Non-Asymptotics Properties of the Bernoulli GLR}\label{subsec:6:PropGLR}
+\subsection{Non-Asymptotics Properties of the B-GLRT}\label{subsec:6:PropGLR}
 
@@
 %
-We give below two Lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:DetectionDelay} which bounds the detection delay if there is enough samples before a change.
+We give below two Lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.
 
 %-----------------------------------------------------------------------------
-\paragraph{False alarm probability}
+\paragraph{False alarm probability.}
 
 In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small.
-To define $\beta$, we need to introduce the function $\cT$,
+To define $\beta$, we need to introduce the function $\cT$, defined for $x>0$ by
 %
 \begin{equation}\label{def:6:function_T}
-    \cT(x) ~:=~ 2 \tilde h\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
+    \cT(x) ~:=~ 2 \widetilde{h}\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
 \end{equation}
@@
 %
-And for any $x \ge 0$, $\tilde h(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise.
+And for any $x \ge 0$, $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise.
 The function $\cT$ is easy to compute numerically.
-The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function (\cite{Corless96}) (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
+The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96} (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
 %
@@
     Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
-    $\bP_{\mu_0}(\hat\tau_\delta < \infty) \leq \delta$, with the threshold function
+    $\bP_{\mu_0}(\widehat{\tau}_\delta < \infty) \leq \delta$, with the threshold function
     %
@@
 
+\begin{proof}
+    \TODOL{Write it!}
+\end{proof}
+
 %-----------------------------------------------------------------------------
-\paragraph{Detection delay}
+\paragraph{Detection delay.}
 
 Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from $\mu_0$ to $\mu_1$ occurs at time $\tau$.
-We already observed that from Pinsker's inequality, the Bernoulli GLR stops earlier than a Gaussian GLR.
-Hence, one can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the Bernoulli GLR.
+We already observed that from Pinsker's inequality, the B-GLRT stops earlier than a Gaussian GLR.
+Hence, one can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the B-GLRT.
 Letting $\Delta = |\mu_0 - \mu_1|$, one can essentially establish that for $\tau$ larger than $(1/\Delta^2)\ln(1/\delta)$ (\ie, enough samples before the change), the delay can be of the same magnitude (\ie, enough samples after the change).
 In our bandit analysis to follow, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and \ref{lem:6:GoodEvent}).
+
+\TODOL{Include in Appendix a formal statement as well as a proof of Pinsker's inequality!}
 
@@
 
-For a threshold $\beta$ chosen as in the Lemma~\ref{lemma:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lemma:Delay} is that if the break $\tau$ takes place after $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude.
-%
-In other words,
-if the B-GLRT test observes enough samples before rupture,
-its delay $\widehat{\tau}_\delta$ is in the order of $O(\Delta^{-2} \ln(1/\delta))$, in high probability.
+For a threshold $\beta$ chosen as in the Lemma~\ref{lem:6:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lem:6:Delay} is that if the break $\tau$ takes place after $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude.
+%
+In other words, if the B-GLRT test observes enough samples before a change-point,
+its delay $\widehat{\tau}_\delta$ is in the order of $O(\Delta^{-2} \ln(1/\delta))$, with high probability.
+
+\begin{proof}
+    \TODOL{Write it!}
+\end{proof}
 
@@
 
-Lemma~\ref{lem:6:FalseAlarm} provides the first control of false alarm for the Bernoulli GLR employed for bounded data.
+Lemma~\ref{lem:6:FalseAlarm} provides the first control of false alarm for the B-GLRT employed for bounded data.
 However, the threshold \eqref{def:6:beta} is not fully explicit as the function $\cT(x)$ can only be computed numerically.
@@
 %
-For the Bernoulli GLR, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative, and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
-
-Also note that, as any test based on scan-statistics, the GLR can be costly to implement as at every time step, it considers all previous time steps as a possible positions for a change-point.
+For the B-GLRT, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative, and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
+
+Also note that, as any test based on scan-statistics, the B-GLRT can be costly to implement as at every time step, it considers all previous time steps as a possible positions for a change-point.
 Thus, in practice the following adaptation may be interesting, based on down-sampling the possible time steps:
@@
 \begin{equation}\label{def:6:GLRTricks}
-    \tilde \tau_{\delta} = \inf \left\{ n \in \cN : \sup_{s \in \cS_n} \left[s \times d\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times d\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\},
+    \widetilde{\tau}_{\delta} = \inf \left\{ n \in \cN : \sup_{s \in \cS_n} \left[s \times d\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times d\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\},
 \end{equation}
@@
 %-----------------------------------------------------------------------------
-\paragraph{Details About the Algorithm}
+\paragraph{Details about the algorithm.}
 
@@
 %
-\GLRklUCB{} combines the \klUCB{} algorithm (\cite{KLUCBJournal}), known to be optimal for Bernoulli bandits, with the Bernoulli GLR change-point detector introduced in Section~\ref{sec:6:ChangePointDetector}.
+\GLRklUCB{} combines the \klUCB{} algorithm as it is given and analyzed by \cite{KLUCBJournal}, known to be optimal for Bernoulli bandits, with the B-GLRT change-point detector introduced in Section~\ref{sec:6:ChangePointDetector}.
 %
@@
 The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms.
-A restart happens when the Bernoulli GLR change-point detector detects a change on the arm that has been played (line $9$).
+A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
 %
@@
 We define the (\klUCB{}-like) index used by our algorithm, by denoting
-$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$, $n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)$ the number of selections of arm $i$,
-and $\hat{\mu}_i(t) = (1/n_i(t)) \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)$ their empirical mean (if $n_i(t)\neq0$).
+$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
+\[n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)\]
+the number of selections of arm $i$, and
+\[\widehat{\mu}_i(t) = (1/n_i(t)) \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
+their empirical mean (if $n_i(t)\neq0$).
 %
@@
 \begin{equation}\label{def:6:UCB}
-    \UCB_i(t) := \max \bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\hat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \bigr\}.
-\end{equation}
-
+    \UCB_i(t) := \max \bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \bigr\}.
+\end{equation}
+
+\paragraph{Two options.}
+%
 In this work, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
@@
 %
-One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{This alternative choice is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB{}: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
+One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative choice is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
 $(31)$-$(32)$.}:
@@
 %-----------------------------------------------------------------------------
-\paragraph{Forced exploration}
+\paragraph{Forced exploration.}
 
@@
 Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay.
-Each regret bound then follows from upper bounds on term (A), term (B), and on the failure probability $\bP(\cE_T^c)$.
-%
-To control (A) and (B), we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
+Each regret bound then follows from upper bounds on term $(A)$, term $(B)$, and on the failure probability $\bP(\cE_T^c)$.
+%
+To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
 
@@
 %
-where $\hat\tau_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
-Introducing the event $\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \hat\tau_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\}$ that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition:
+where $\widehat{\tau}_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
+Introducing the event $\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\}$ that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition:
 \begin{align}
-    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
+    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
 \end{align}
 
-Term (a) is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
+Term $(a)$ is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
 Observe that under the bandit algorithm, the change point detector associated to arm $i$ is based on (possibly much) less than $t - \tau_i(t)$ samples from arm $i$, which makes false alarm even less likely to occur.
@@
 
-Term (b) is related to the control of the detection delay, which is more tricky to obtain under the \GLRklUCB{} adaptive sampling scheme,
+Term $(b)$ is related to the control of the detection delay, which is more tricky to obtain under the \GLRklUCB{} adaptive sampling scheme,
 when compared to a result like Theorem~6 in \cite{Maillard2018GLR} for the change-point detector run in isolation.
@@
 
-\begin{theorem}
-    \label{thm:6:mainRegretBoundGlobal}
+\begin{theorem}\label{thm:6:mainRegretBoundGlobal}
     For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
@@
 
-\subsection{Analysis of \GLRklUCB{} with Global Changes}\label{proof:6:mainRegretBoundGlobal}
+\subsection{Proofs for \GLRklUCB{} with Global Changes}
+\label{proof:6:mainRegretBoundGlobal}
 
 We gave in Section~\ref{sec:6:RegretAnalysis} the following decomposition of the regret $R_T$,
+%
 \begin{eqnarray}\label{eq:6:GeneRegretBound}
@@
 We first introduce some notation for the proof:
-let $\hat\tau^{(k)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\hat\tau(t)$ be the last time before $t$ that the algorithm restarted.
-We denote $n_i(t) = \sum_{s=\tau(t)+1}^{t}\indic(A_s = i)$ the number of selections of arm $i$ since the last (global) restart, and $\hat{\mu}_i(t) = \frac{1}{n_i(t)}\sum_{s=\tau(t)+1}^{t}X_{i,s}\indic(A_s = i)$ their empirical average (if $n_i(t) \neq 0$).
+let $\widehat{\tau}^{(k)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
+We denote $n_i(t) = \sum_{s=\tau(t)+1}^{t}\indic(A_s = i)$ the number of selections of arm $i$ since the last (global) restart, and $\widehat{\mu}_i(t) = \frac{1}{n_i(t)}\sum_{s=\tau(t)+1}^{t}X_{i,s}\indic(A_s = i)$ their empirical average (if $n_i(t) \neq 0$).
 
@@
 
-\paragraph{Upper bound on Term $\bm{(A)}$}
-
-\begin{align*}
-    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \, \kl\left(\hat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau(t))\right)\right] \\
-     & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau(t))\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(k)} + \indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau^{(k)})\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d^{(k)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \hat\tau^{(k)})\right)\right],
+\paragraph{Upper bound on term $\bm{(A)}$.}
+
+\begin{align*}
+    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \, \kl\left(\widehat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau(t))\right)\right] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}(t))\right)\right] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(k)} + \indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}^{(k)})\right)\right] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d^{(k)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(k)})\right)\right],
 \end{align*}
@@
 \begin{equation}\label{def:6:EventCiGlobal}
-    \cC^{(k)} = \left\{\forall j \leq k, \hat\tau^{(j)} \in \{\tau^{(j)} + 1, \dots, \tau^{(j)} + d^{(j)} \} \right\}.
-\end{equation}
-%
-Clearly, $\cE_T \subseteq \cC^{(k)}$ and $\cC^{(k)}$ is $\cF_{\hat\tau^{(k)}}$-measurable.
-Observe that conditionally to $\cF_{\hat\tau^{(k)}}$, when $\cC^{(k)}$ holds, $\hat{\mu}_{k^*}(t)$ is the average of samples that have all mean $\mu_{k^*}^{(k)}$.
-%
-Thus, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_{k^*}^{(k)}$, one can write
-%
-\begin{align*}
-    &\bE\left[\left.\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau^{(k)})\right) \right| \cF_{\hat\tau^{(k)}}\right]\\
-    & = \indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\hat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \hat\tau^{(k)})\right) \;|\; \cF_{\hat\tau^{(k)}}\right] \\
-    & \leq \indic(\cC^{(k)}) \sum_{t'=1}^{\tau^{(k+1)} - \hat{\tau}^{(k)}} \bP\left(\exists s \leq t' : s \, \kl(\hat{\mu}_{s},\mu_{k^*}^{(k)}) \geq f(t')\right) \\
+    \cC^{(k)} = \left\{\forall j \leq k, \widehat{\tau}^{(j)} \in \{\tau^{(j)} + 1, \dots, \tau^{(j)} + d^{(j)} \} \right\}.
+\end{equation}
+%
+Clearly, $\cE_T \subseteq \cC^{(k)}$ and $\cC^{(k)}$ is $\cF_{\widehat{\tau}^{(k)}}$-measurable.
+Observe that conditionally to $\cF_{\widehat{\tau}^{(k)}}$, when $\cC^{(k)}$ holds, $\widehat{\mu}_{k^*}(t)$ is the average of samples that have all mean $\mu_{k^*}^{(k)}$.
+%
+Thus, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_{k^*}^{(k)}$, one can write
+%
+\begin{align*}
+    &\bE\left[\left.\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}^{(k)})\right) \right| \cF_{\widehat{\tau}^{(k)}}\right]\\
+    & = \indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}^{(k)})\right) \;|\; \cF_{\widehat{\tau}^{(k)}}\right] \\
+    & \leq \indic(\cC^{(k)}) \sum_{t'=1}^{\tau^{(k+1)} - \hat{\tau}^{(k)}} \bP\left(\exists s \leq t' : s \, \kl(\widehat{\mu}_{s},\mu_{k^*}^{(k)}) \geq f(t')\right) \\
     & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
@@
 
-\paragraph{Upper bound on Term $\bm{(B)}$.} We let $\tilde{\mu}_{i,s}^{(k)}$ denote the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
+\paragraph{Upper bound on term $\bm{(B)}$.}
+%
+We let $\tilde{\mu}_{i,s}^{(k)}$ denote the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
 Rewriting the sum in $t$ as the sum of consecutive intervals $[\tau^{(k)}+1, \tau^{(k+1)}]$,
+%
 \begin{align*}
     (B) & \leq \bE\Big[\indic(\cE_T)\sum_{k=1}^{\Upsilon_T} \sum_{t=\tau^{(k)}}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{I_t}^{(k)}\right)\indic\left(\UCB_{I_t}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
-     & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\hat{\tau}^{(k)} + \indic(\cE_T) \sum_{t=\hat\tau^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{I_t}^{(k)}\right)\indic\left(\UCB_{I_t}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\hat\tau^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\indic\left(I_t=i,\UCB_{i}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\hat{\tau}^{(k)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{I_t}^{(k)}\right)\indic\left(\UCB_{I_t}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\indic\left(I_t=i,\UCB_{i}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
      & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\times \\
-     &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\sum_{s=1}^{t-\hat\tau^{(k)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \hat\tau^{(k)})\right)\Big] \\
+     &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(k)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \widehat{\tau}^{(k)})\right)\Big] \\
     & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_i(\tau^{(k+1)})}\indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\Big]  \\
@@
 
-Conditionally to $\cF_{\hat\tau^{(k)}}$, when $\cC^{(k)}$ holds, for $s \in \{1, \ldots, n_i(\tau^{(k+1)})\}$, $\tilde{\mu}_{i,s}^{(k)}$ is the empirical mean from \iid{} observations of mean $\mu_i^{(k)}$.
-Therefore, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(k)}$, it follows from the law of total expectation that
-\begin{align*}
-    (B)& \leq \sum_{k=1}^{\Upsilon_T}d^{(k)}+\sum_{i=1}^K\sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right) \sum_{s=1}^{\tau^{(k+1)}-\tau^{(k)}}\bP\left(s \times \kl(\hat{\mu}_{s}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\cdot
+Conditionally to $\cF_{\widehat{\tau}^{(k)}}$, when $\cC^{(k)}$ holds, for $s \in \{1, \ldots, n_i(\tau^{(k+1)})\}$, $\tilde{\mu}_{i,s}^{(k)}$ is the empirical mean from \iid{} observations of mean $\mu_i^{(k)}$.
+Therefore, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(k)}$, it follows from the law of total expectation that
+\begin{align*}
+    (B)& \leq \sum_{k=1}^{\Upsilon_T}d^{(k)}+\sum_{i=1}^K\sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right) \sum_{s=1}^{\tau^{(k+1)}-\tau^{(k)}}\bP\left(s \times \kl(\widehat{\mu}_{s}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\cdot
 \end{align*}
+%
 If $\mu_{k^*}^{(k)} > \mu_i^{(k)}$ by definition, we can use the analysis of \klUCB{} from \cite{KLUCBJournal} to further upper bound the right-most part, and we obtain
+%
 \begin{align}\label{eq:6:TermBFinalGlobal}
@@
 
-% We introduce $\hat{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after $t=\tau^{(k)}+1$ and $\tilde{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
+% We introduce $\widehat{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after $t=\tau^{(k)}+1$ and $\tilde{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
 
@@
     \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T}
-    \bP\left(\left. \hat\tau^{(k)} \notin \{ \tau^{(k)}+1, \dots, \tau^{(k)} + d^{(k)} \} \right|\; \cC^{(k-1)}\right) \\
-    \\ & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\hat\tau^{(k)} \leq \tau^{(k)} \;|\; \cC^{(k-1)}\right)}_{(a)} + \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\hat\tau^{(k)} \geq \tau^{(k)} + d^{(k)} \;|\; \cC^{(k-1)}\right)}_{(b)}.
+    \bP\left(\left. \widehat{\tau}^{(k)} \notin \{ \tau^{(k)}+1, \dots, \tau^{(k)} + d^{(k)} \} \right|\; \cC^{(k-1)}\right) \\
+    \\ & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \leq \tau^{(k)} \;|\; \cC^{(k-1)}\right)}_{(a)} + \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)} \;|\; \cC^{(k-1)}\right)}_{(b)}.
 \end{align*}
@@
 %
-$\hat\tau^{(k)} \leq \tau^{(k)}$ implies that there exists an arm whose associated change point detector has experienced a false-alarm:
+$\widehat{\tau}^{(k)} \leq \tau^{(k)}$ implies that there exists an arm whose associated change point detector has experienced a false-alarm:
 %
@@
     (a) & \leq \bP\left(\exists i, \exists s < t \leq n_i(\tau_i^{(k)}) : s \, \kl\left(\tilde{\mu}_{i,1:s}^{(k-1)},\tilde{\mu}_{i,1:t}^{(k-1)}\right) + (t - s) \, \kl\left(\tilde{\mu}_{i,s+1:t}^{(k-1)},\tilde{\mu}_{i,1:t}^{(k-1)}\right) > \beta(t, \delta) \;|\; \cC^{(k-1)}\right) \\
-    & \leq \sum_{i=1}^K\bP\left(\exists s < t : s \, \kl(\hat{\mu}_{1:s},\mu_i^{(k-1)}) + (t - s) \, \kl(\hat{\mu}_{s+1:t}, \mu_{i}^{(k-1)}) > \beta(t, \delta)\right),
+    & \leq \sum_{i=1}^K\bP\left(\exists s < t : s \, \kl(\widehat{\mu}_{1:s},\mu_i^{(k-1)}) + (t - s) \, \kl(\widehat{\mu}_{s+1:t}, \mu_{i}^{(k-1)}) > \beta(t, \delta)\right),
 \end{align*}
@@
 
-\paragraph{Upper bound on term (b): controlling the delay.}
+\paragraph{Upper bound on term $(b)$: controlling the delay.}
 %
@@
 (as the mapping $t\mapsto n_i(t) - n_i(\tau^{(k)})$ is non-decreasing, is $0$ at $t=\tau^{(k)}$ and its value at $\tau^{(k)}+d^{(k)}$ is larger than $\overline{r}$).
-Using that \[\left(\hat\tau^{(k)} \geq \tau^{(k)} + d^{(k)}\right) \subseteq \left(\hat\tau^{(k)} \geq \overline{t}\right)\]
-the event $\left(\hat\tau^{(k)} \geq \tau^{(k)} + d^{(k)}\right)$ further implies that
+Using that \[\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)}\right) \subseteq \left(\widehat{\tau}^{(k)} \geq \overline{t}\right)\]
+the event $\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)}\right)$ further implies that
 \[
@@
 Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_i(\tau^{(k)})$ lies in the interval $
-\left[\left\lfloor \frac{\alpha}{K}\left(\tau^{(k)}-\hat\tau^{(k-1)}\right)\right\rfloor,\left(\tau^{(k)}-\hat\tau^{(k-1)}\right)\right]$.
+\left[\left\lfloor \frac{\alpha}{K}\left(\tau^{(k)}-\widehat{\tau}^{(k-1)}\right)\right\rfloor,\left(\tau^{(k)}-\widehat{\tau}^{(k-1)}\right)\right]$.
 Conditionally to $\cC^{(k-1)}$, one obtains furthermore using that $d^{(k-1)} \leq (\tau^{(k)} - \tau^{(k-1)})/2$ -- which follows from Assumption~\ref{ass:6:LongPeriodsGlobal} -- that
+%
 \begin{eqnarray*}
@@
 
-Introducing $\hat{\mu}_{a,s}$ (resp.
-$\hat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\hat{\mu}_i^{(k-1)}$ (resp.
-$\hat{\mu}_i^{(k)}$), such that $\hat{\mu}_{a,s}$ and $\hat{\mu}_{b,r}$ are independent, it follows that
+Introducing $\widehat{\mu}_{a,s}$ (resp.
+$\widehat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\widehat{\mu}_i^{(k-1)}$ (resp.
+$\widehat{\mu}_i^{(k)}$), such that $\widehat{\mu}_{a,s}$ and $\widehat{\mu}_{b,r}$ are independent, it follows that
 \[
-    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\hat{\mu}_{a,s},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\hat{\mu}_{b,\overline{r}},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
+    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\widehat{\mu}_{a,s},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\widehat{\mu}_{b,\overline{r}},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
 \]
@@
 Using Pinsker's inequality and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = | \Delta_i^k |$), one can write
+%
 \begin{align}
-    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
-    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,s} - \Delta_i^{(k)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
-    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
+    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
+    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,s} - \Delta_i^{(k)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
+    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
 \end{align}
@@
 For the second term, we use the observation
-\[\frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(k)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
+\[\frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(k)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
 and, using that $\Delta^{(k)} = | \Delta_i^k|$, one obtains
-\begin{equation}(b) \leq \delta + \bP\left(\exists s \in \cI_k : \Delta^{(k)} \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).\label{eq:6:FromHereGlobal}\end{equation}
+%
+\begin{equation}\label{eq:6:FromHereGlobal}
+    (b) \leq \delta + \bP\left(\exists s \in \cI_k : \Delta^{(k)} \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).
+\end{equation}
+%
 Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau^{(k)}-\tau^{(k-1)})/2\right\rfloor$.
@@
 \[\forall s \in \cI_k, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta^{(k)}\right)^2.\]
+%
 Hence, the probability in the right-hand side of \eqref{eq:6:FromHereGlobal} is zero, which yields $(b) \leq \delta$.
@@
 Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}. A first key observation is
-\[s \times \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\hat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\hat{\mu}_{s+1:n},\lambda\right)\right].\]
+\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
 Hence the probability of a false alarm occurring is upper bounded as
@@
 \begin{align*}
-    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\hat{\mu}_{1:s},\hat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\hat{\mu}_{s+1:n},\hat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
-    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\hat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\hat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\hat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\hat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
-    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\hat{\mu}_{s},\mu_0\right) + r \, \kl\left(\hat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
+    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
+    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
+    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
 \end{align*}
-where $\hat{\mu}_{s}$ and $\hat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
+%
+where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
 As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
@@
     &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu_0\right) + r \, \kl\left(\hat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu_0\right) + r \, \kl\left(\hat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
 \end{align*}
+%
 And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
@@
     Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family.
-Introducing the notation $\hat{\mu}_s = \frac{1}{s}\sum_{i=1}^s X_i$ and $\hat{\mu}'_r = \frac{1}{r}\sum_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
-    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu\right) + r \, \kl\left(\hat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
+Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
+    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
     where $\cT$ is the function defined in \eqref{def:6:function_T}.
@@
 \paragraph{Proof of Lemma~\ref{lem:6:ConcFirst}.}
+%
 Using the same construction as in the proof of Theorem~14 in \cite{KK18Martingales}, one can prove that for every $\lambda \in I$ (for an interval $I$), there exists a non-negative super-martingale $M^\lambda (s)$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ that satisfies $\bE[M^\lambda(s)] \leq 1$ and
-\[\forall s \in \N^*, \ \ M^\lambda (s) \geq e^{\lambda [s\kl(\hat{\mu}_s,\mu) - 3\ln(1+\ln(s))] - g(\lambda)}\]
+\[\forall s \in \N^*, \ \ M^\lambda (s) \geq e^{\lambda [s\kl(\widehat{\mu}_s,\mu) - 3\ln(1+\ln(s))] - g(\lambda)}\]
 for some function $g : I \rightarrow \R$. This super-martingale is of the form
@@
 Similarly, there exists an independent super-martingale $W^\lambda (r)$ w.r.t. the filtration $\cF_r' = \sigma(Y_1,\dots,Y_r)$ such that
-\[\forall r \in \N^*, \ \ W^\lambda (r) \geq e^{\lambda [r\kl(\hat{\mu}'_r,\mu) - 3\ln(1+\ln(r))] - g(\lambda)},\]
+\[\forall r \in \N^*, \ \ W^\lambda (r) \geq e^{\lambda [r\kl(\widehat{\mu}'_r,\mu) - 3\ln(1+\ln(r))] - g(\lambda)},\]
 for the same function $g(\lambda)$.
-In the terminology of \cite{KK18Martingales}, the processes $\bm{X}(s) = s \, \kl(\hat{\mu}_s,\mu) - 3\ln(1+\ln(s))$ and $\bm{Y}(s) = r \, \kl(\hat{\mu}_r,\mu) - 3\ln(1+\ln(r))$ are called $g$-DCC for Doob-Cram\'er-Chernoff, as Doob's inequality can be applied in combination with the Cram\'er-Chernoff method to obtain deviation inequalities that are uniform in time.
+In the terminology of \cite{KK18Martingales}, the processes $\bm{X}(s) = s \, \kl(\widehat{\mu}_s,\mu) - 3\ln(1+\ln(s))$ and $\bm{Y}(s) = r \, \kl(\widehat{\mu}_r,\mu) - 3\ln(1+\ln(r))$ are called $g$-DCC for Doob-Cram\'er-Chernoff, as Doob's inequality can be applied in combination with the Cram\'er-Chernoff method to obtain deviation inequalities that are uniform in time.
 
@@
 One can write
-\begin{align*}
-    &\bP\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu\right) + r \, \kl\left(\hat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \\
+%
+\begin{align*}
+    &\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \\
     & \leq \bP\left(\exists r \in \N^* : M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right) \\
@@
 Putting things together yields
-\[\bP\left(\exists r \in \N^* : s \, \kl\left(\hat{\mu}_{s},\mu\right) + r \, \kl\left(\hat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \leq e^{-\left(\lambda u - 2g_{\xi}(\lambda)\right)},\]
+\[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \leq e^{-\left(\lambda u - 2g_{\xi}(\lambda)\right)},\]
 for any function $g_{\xi}$ defined above.
@@
 
-\begin{theorem}
-    \label{thm:6:mainRegretBound}For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on Local Restart satisfies
+\begin{theorem}\label{thm:6:mainRegretBound}
+    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on Local Restart satisfies
     %
@@
 % ------------------------------------------------------------------------------------
-\subsection{Analysis of \GLRklUCB{} with Local Changes}\label{proof:6:mainRegretBound}
+\subsection{Proof for \GLRklUCB{} with Local Changes}
+\label{proof:6:mainRegretBound}
 
@@
 \end{equation}
-where $\hat\tau_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined as in Assumption~\ref{ass:6:LongPeriods}.
+where $\widehat{\tau}_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined as in Assumption~\ref{ass:6:LongPeriods}.
 Using this assumption, one can prove the following.
@@
 
-\paragraph{Upper bound on term (A).}
-%
-\begin{align*}
-    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \kl\left(\hat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau_{i_t^*}(t))\right)\right] \\
-    & \leq \sum_{i=1}^K \sum_{k=1}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic(i_t^* =i) \indic(\tau_i(t) = \hat\tau_i^{(\ell)})\indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right)\right] \\
-    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right) + d_i^{(\ell)}(T)\right] \\
-    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}d_i^{(\ell+1)}(T) + \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}}\indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right)\right],
+\paragraph{Upper bound on term $\bm{(A)}$.}
+%
+\begin{align*}
+    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \kl\left(\widehat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau_{i_t^*}(t))\right)\right] \\
+    & \leq \sum_{i=1}^K \sum_{k=1}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic(i_t^* =i) \indic(\tau_i(t) = \widehat{\tau}_i^{(\ell)})\indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right)\right] \\
+    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right) + d_i^{(\ell)}(T)\right] \\
+    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}d_i^{(\ell+1)}(T) + \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}}\indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right)\right],
 \end{align*}%
@@
 where we introduce the event $\cC_i^{(\ell)}$ that all the changes up to the $\ell$-th have been detected:
-\begin{equation}\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \hat\tau_i^{(j)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(\ell)} \right] \right\}.\label{def:6:EventCi}\end{equation}
-Clearly, $\cE_T \subseteq \cC_i^{(\ell)}$ and $\cC_i^{(\ell)}$ is $\cF_{\hat\tau_i^{(\ell)}}$-measurable.
-Observe that conditionally to $\cF_{\hat\tau_i^{(\ell)}}$, when $\indic(\cC_i^{(\ell)})$ holds, $\hat{\mu}_{i}(t)$ is the average of samples that have all mean $\mu_i^{(\ell)}$.
-Thus, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(\ell)}$, one can write
-\begin{align*}
-    &\bE\left[\left.\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right) \right| \cF_{\hat\tau_i^{(\ell)}}\right]\\
-    & = \indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \bE\left[\indic\left(n_{i}(t) \kl\left(\hat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \hat\tau_{i}^{(\ell)})\right) \;|\; \cF_{\hat\tau_i^{(\ell)}}\right]
+\begin{equation}\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(j)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(\ell)} \right] \right\}.\label{def:6:EventCi}\end{equation}
+Clearly, $\cE_T \subseteq \cC_i^{(\ell)}$ and $\cC_i^{(\ell)}$ is $\cF_{\widehat{\tau}_i^{(\ell)}}$-measurable.
+Observe that conditionally to $\cF_{\widehat{\tau}_i^{(\ell)}}$, when $\indic(\cC_i^{(\ell)})$ holds, $\widehat{\mu}_{i}(t)$ is the average of samples that have all mean $\mu_i^{(\ell)}$.
+Thus, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(\ell)}$, one can write
+\begin{align*}
+    &\bE\left[\left.\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right) \right| \cF_{\widehat{\tau}_i^{(\ell)}}\right]\\
+    & = \indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \bE\left[\indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right) \;|\; \cF_{\widehat{\tau}_i^{(\ell)}}\right]
 % \end{align*}%
 % \begin{align*}
-    & \leq \indic(\cC_i^{(\ell)}) \sum_{t=1}^{\tau_i^{(\ell+1)} - \hat{\tau}_i^{(\ell)}} \bP\left(\exists s \leq t' : s \times \kl(\hat{\mu}_{s},\mu_i^{(\ell)}) \geq f(t')\right) \\
+    & \leq \indic(\cC_i^{(\ell)}) \sum_{t=1}^{\tau_i^{(\ell+1)} - \hat{\tau}_i^{(\ell)}} \bP\left(\exists s \leq t' : s \times \kl(\widehat{\mu}_{s},\mu_i^{(\ell)}) \geq f(t')\right) \\
     & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
@@
 
-\paragraph{Upper bound on term (B).}
+\paragraph{Upper bound on term $\bm{(B)}$.}
 %
@@
     (B) & \leq \sum_{i=1}^K\bE\Big[\indic(\cE_T)\sum_{\ell=1}^{\NCi} \sum_{t=\tau_i^{(\ell)}}^{\tau_i^{(\ell+1)}}\left(\mu_{i^*_t}(t) - \overline{\mu}_{i}^{(\ell)}\right)\indic\left(I_t = i, \UCB_i(t) \geq \mu^*_{i,\ell}\right)\Big] \\
-    & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T)\hat{\tau}_i^{(\ell)} + \indic(\cE_T) \sum_{t=\hat\tau_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\indic\left(I_t = i, \UCB_i(t) \geq \mu^*_{i,\ell}\right)\Big] \\
-%     & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+ \sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\sum_{s=1}^{t-\hat\tau_i^{(\ell)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \times \kl(\tilde{\mu}_{i,s}^{(\ell)}, \mu^*_{\ell,i}) \leq f(\tau_i^{(\ell+1)} - \hat\tau_i^{(\ell)})\right)\Big] \\
+    & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T)\hat{\tau}_i^{(\ell)} + \indic(\cE_T) \sum_{t=\widehat{\tau}_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\indic\left(I_t = i, \UCB_i(t) \geq \mu^*_{i,\ell}\right)\Big] \\
+%     & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+ \sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\sum_{s=1}^{t-\widehat{\tau}_i^{(\ell)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \times \kl(\tilde{\mu}_{i,s}^{(\ell)}, \mu^*_{\ell,i}) \leq f(\tau_i^{(\ell+1)} - \widehat{\tau}_i^{(\ell)})\right)\Big] \\
     & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+ \sum_{i=1}^K\sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cC_i^{(\ell)}) \sum_{s=1}^{n_i(\tau_i^{(\ell+1)})}\indic\left(s \times \kl(\tilde{\mu}_{i,s}^{(\ell)}, \mu^*_{i,\ell}) \leq f(\tau_i^{(\ell+1)} - \tau_i^{(\ell)})\right)\Big].
@@
 The last inequality relies on introducing a sum over $\indic(n_i(t) = s)$ and swapping the sums.
-Conditionally to $\cF_{\hat\tau_i^{(\ell)}}$, when $\cC_i^{(\ell)}$ holds, for $s \in \{1, \ldots, n_i(\tau_i^{(\ell+1)})\}$, $\tilde{\mu}_{i,s}^{(\ell)}$ is the empirical mean from \iid{} observations of mean $\overline{\mu}_i^{(\ell)}$.
-Therefore, introducing $\hat{\mu}_s$ as a sequence of \iid{} random variables with mean $\overline{\mu}_i^{(\ell)}$, it follows from the law of total expectation that
+Conditionally to $\cF_{\widehat{\tau}_i^{(\ell)}}$, when $\cC_i^{(\ell)}$ holds, for $s \in \{1, \ldots, n_i(\tau_i^{(\ell+1)})\}$, $\tilde{\mu}_{i,s}^{(\ell)}$ is the empirical mean from \iid{} observations of mean $\overline{\mu}_i^{(\ell)}$.
+Therefore, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\overline{\mu}_i^{(\ell)}$, it follows from the law of total expectation that
 %
@@
     (B) & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi} d_i^{(\ell)}
-    + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \sum_{s=1}^{\tau_i^{(\ell+1)}-\tau_i^{(\ell)}} \bP\left(s \times \kl(\hat{\mu}_{s}, \mu^*_{i,\ell}) \leq f(\tau_i^{(\ell+1)} - \tau_i^{(\ell)})\right).
+    + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \sum_{s=1}^{\tau_i^{(\ell+1)}-\tau_i^{(\ell)}} \bP\left(s \times \kl(\widehat{\mu}_{s}, \mu^*_{i,\ell}) \leq f(\tau_i^{(\ell+1)} - \tau_i^{(\ell)})\right).
 \end{align*}%
@@
 \begin{align*}
-    \bP(\cE_T^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
+    \bP(\cE_T^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
 \end{align*}
@@
     (a) & \leq \bP\left(\exists s < t \leq n_i(\tau_i^{(\ell)}) : s \times \kl\left(\tilde{\mu}_{i,1:s}^{(\ell-1)},\tilde{\mu}_{i,1:t}^{(\ell-1)}\right) + (t - s) \times \kl\left(\tilde{\mu}_{i,s+1:t}^{(\ell-1)},\tilde{\mu}_{i,1:t}^{(\ell-1)}\right) > \beta(t, \delta) \;|\; \cC_i^{(\ell-1)}\right) \\
-    & \leq \bP\left(\exists s < t : s \times \kl(\hat{\mu}_{1:s},\mu_i^{(\ell-1)}) + (t - s) \times \kl(\hat{\mu}_{s+1:t}, \mu_{i}^{(\ell-1)}) > \beta(t, \delta)\right),
+    & \leq \bP\left(\exists s < t : s \times \kl(\widehat{\mu}_{1:s},\mu_i^{(\ell-1)}) + (t - s) \times \kl(\widehat{\mu}_{s+1:t}, \mu_{i}^{(\ell-1)}) > \beta(t, \delta)\right),
 \end{align*}
@@
 (as the mapping $t\mapsto n_i(t) - n_i(\tau_i^{(\ell)})$ is non-decreasing, is $0$ at $t=\tau_i^{(\ell)}$ and its value at $\tau_i^{(\ell)}+d_i^{(\ell)}$ is larger than $\overline{r}$).
-Using that $(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)}) \subseteq (\hat\tau_i^{(\ell)} \geq \overline{t})$,
-the event $(\hat\tau_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)})$ further implies that
+Using that $(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)}) \subseteq (\widehat{\tau}_i^{(\ell)} \geq \overline{t})$,
+the event $(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)})$ further implies that
 \[
@@
 Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_i(\tau_i^{(\ell)})$ lies in $
-\left\{\left\lfloor \frac{\alpha}{K}\left(\tau_i^{(\ell)}-\hat\tau_i^{(\ell-1)}\right)\right\rfloor, \dots, \tau_i^{(\ell)}-\hat\tau_i^{(\ell-1)} \right\}$.
+\left\{\left\lfloor \frac{\alpha}{K}\left(\tau_i^{(\ell)}-\widehat{\tau}_i^{(\ell-1)}\right)\right\rfloor, \dots, \tau_i^{(\ell)}-\widehat{\tau}_i^{(\ell-1)} \right\}$.
 %
@@
 
-Introducing $\hat{\mu}_{a,s}$ (resp.
-$\hat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\hat{\mu}_i^{(\ell-1)}$ (resp.
-$\hat{\mu}_i^{(\ell)}$), such that $\hat{\mu}_{a,s}$ and $\hat{\mu}_{b,r}$ are independent, it follows that
+Introducing $\widehat{\mu}_{a,s}$ (resp.
+$\widehat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\widehat{\mu}_i^{(\ell-1)}$ (resp.
+$\widehat{\mu}_i^{(\ell)}$), such that $\widehat{\mu}_{a,s}$ and $\widehat{\mu}_{b,r}$ are independent, it follows that
 \[
-    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\hat{\mu}_{a,s},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\hat{\mu}_{b,\overline{r}},\frac{s\hat{\mu}_{a,s}+\overline{r}\hat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
+    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\widehat{\mu}_{a,s},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\widehat{\mu}_{b,\overline{r}},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
 \]
@@
 \begin{align}
-    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
-    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,s} - \Delta_i^{(\ell)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
-    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
+    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
+    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,s} - \Delta_i^{(\ell)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
+    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
 \end{align}
@@
 For the second term, we use the observation
-\[\frac{2s\overline{r}}{s+\overline{r}}\left(\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\hat{\mu}_{a,s} - \hat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(\ell)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
+\[\frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(\ell)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
 and finally get
-\begin{equation}(b) \leq \delta + \bP\left(\exists s \in \cI_k : |\Delta_i^{(\ell)}| \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).\label{eq:6:FromHere}\end{equation}
+%
+\begin{equation}\label{eq:6:FromHere}
+    (b) \leq \delta + \bP\left(\exists s \in \cI_k : |\Delta_i^{(\ell)}| \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).
+\end{equation}
+%
 Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau_i^{(\ell)}-\tau_i^{(\ell-1)})/2\right\rfloor$.
@@
 
-\subsection{Interpretation}
+\subsection{Interpretation of the results}
+\label{sub:6:interpretationRegretBounds}
 
@@
 
-\paragraph{Illustrations of our benchmark}\label{app:6:Figures}
+\paragraph{Illustrations of our benchmark.}\label{app:6:Figures}
 
@@
 
-\paragraph{Regret plots}
+\paragraph{Regret plots.}
 \label{subsub:6:regretPlots}
@@
 
-% \paragraph{Acknowledgments}
+% \paragraph{Acknowledgments.}
 % Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.
 
-\paragraph{A note on the simulation code}
+\paragraph{A note on the simulation code.}
 %

### 2019-03-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Should reaaaaaallllyyyy go

@@
 It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
-
+%
 Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
-\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
-and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
-Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
+% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
+% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
+% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
 
@@
 \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
-and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
+and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
 Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
@@
 %
-\begin{eqnarray*}
-R_T \leq T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} + \ \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
-\end{eqnarray*}
+\begin{align*}
+    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
+    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
+\end{align*}
 %
@@
     \begin{equation}\label{def:6:DelayGlobal}
-        d^{(k)} = d^{(k)}(\alpha,\delta) = \lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \rceil.
+        d^{(k)} = d^{(k)}(\alpha,\delta) = \left\lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \right\rceil.
     \end{equation}
@@
 %
-\begin{eqnarray}\label{eq:6:GeneRegretBound}
-    R_T &\leq& T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
-    & & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
-\end{eqnarray}
+\begin{align}\label{eq:6:GeneRegretBound}
+    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
+    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
+\end{align}
 
@@
 As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
-
+%
 \begin{align*}

### 2019-03-12 2-Chapters/6-Chapter/chapter6.tex
Done for tonight, I moved a lot of things around for Chapter 6, and I wrote a few paragraphs detailing more things. I need to finish by Thursday evening, and maybe Friday launch new experiments to finish appendix.

@@
 In this last chapter, we study a generalization of the stationary multi-armed bandit model,
-to account for possible non-stationarity of MAB models.
-Instead of dealing with changes happening possibly at every time step, like in adversarial models, we consider a finite number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
+to account for possible non-stationarity of the problems that could be tackled with MAB algorithms.
+Instead of dealing with changes happening possibly at every time step, like in adversarial models, we rather consider a finite number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
+
+We start by reviewing the previous work on piece-wise stationary MAB models, for which the two main families of proposed solutions are either passively or actively adaptive.
+In practice, it was already observed that actively adaptive strategies usually greatly outperform the passively adaptive ones.
+The more efficient strategies from the current state-of-the-art are based on a combination of an efficient algorithm for the stationary problem with an efficient online change detection algorithm, like CUSUM, PHT or the Generalized Likelihood Ratio Test (GLRT), and so we pursue on this direction.
 %
-We start by reviewing the existing previous work on piece-wise stationary MAB models, for which the two main families of proposed solutions are either passively adaptive or actively adaptive.
-In practice, actively adaptive strategies usually greatly outperform the passively adaptive ones.
-The more efficient strategies from the current state-of-the-art are based on a combination of an efficient algorithm for the stationary problem with an efficient online change detection algorithm, like CUSUM, PHT or the Generalized Likelihood Ratio Test (GLRT).
-Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any sub-Bernoulli distributions, which include distributions of bounded support.
-Our test, B-GLRT, enjoys finite-time guarantees in terms of both its false alarm probability (in a stationary model, the test should not detect any change), and its detection delay (in a model with a change-point, the test should detect within a reasonable delay).
-%
+Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any sub-Bernoulli distributions, which include bounded distributions.
+Our test, B-GLRT, enjoys finite-time guarantees in terms of two quantities:
+its \emph{false alarm probability}, meaning that in a stationary model, the test should not detect any change,
+and its \emph{detection delay}, meaning that in a model with a change-point, the test should detect within a reasonable delay.
+
 Thanks to these two non-asymptotic properties, the B-GLRT test can be combined with an efficient bandit policy, the \klUCB{} index policy, to propose an efficient algorithm for piece-wise stationary bandit problems, that we design \GLRklUCB.
@@
 Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$ with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
-%
+
 Finally, we give extensive numerical experiments to compare our approach, in its two variants, against the current state-of-the-art policies, on different piece-wise stationary bandit problems, of increasing difficulty.
@@
 
-\vfill{}
+% \vfill{}
 
@@
 
-\newpage
+% \newpage
 % Write miniTOC just after the title
@@
 
+We start by giving some additional useful results, along with their proofs, and then we give complementary numerical experiments to justify some choices made in the presentation of our proposal \GLRklUCB.
+
 \input{2-Chapters/6-Chapter/nonstatbandits.git/NonStatB_appendix.tex}

### 2019-03-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for tonight, I moved a lot of things around for Chapter 6, and I wrote a few paragraphs detailing more things. I need to finish by Thursday evening, and maybe Friday launch new experiments to finish appendix.

@@
 % Regret minimization in stochastic bandits has been extensively studied since the works of \cite{Robbins52} and \cite{LaiRobbins85}, and several algorithms with a $\cO(\log(T))$ \emph{problem-dependent} regret upper bound have been proposed (see, \eg, \cite{LattimoreBanditAlgorithmsBook} for a survey).
-Among those, the \klUCB{} algorithm (\cite{KLUCBJournal}) has been shown to be asymptotically optimal for Bernoulli distributions (in that it exactly matches the lower bound given by \cite{LaiRobbins85}) and can also be employed when the rewards are assumed to be bounded in $[0,1]$.
+Among those, the \klUCB{} algorithm \cite{KLUCBJournal} has been shown to be asymptotically optimal for Bernoulli distributions (in that it exactly matches the lower bound given by \cite{LaiRobbins85}) and can also be employed when the rewards are assumed to be bounded in $[0,1]$.
 % \emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{Garivier18}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.
@@
 %
-This chapter is structured as follows.
+% This chapter is structured as follows.
 We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}.
@@
 Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
+We illustrate the two extreme cases in problems $1$ and $2$ presented in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} below.
 
@@
 \TODOL{Speak about the article \cite{WeiSrivastava18} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}
+
+\paragraph{Illustrations of two problems of our benchmark.}\label{par:6:benchmark1}
+
+To illustrate the setup we presented above, we give here two problems of our benchmark, used for the numerical experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
+
+% - Pb 1 changes are only on one arm at a time
+\textbf{Problem $\bm 1$.} There are $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
+The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1}.
+% (and Figure~\ref{fig:6:Problem_1} in Appendix),
+Note that changes happen on only one arm (\ie, $C_T=\Upsilon_T=4$),
+and the optimal arm changes once at $t=\tau_2^{(1)}=2000$, with a large gap $\Delta=0.6$.
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
+    \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).}
+    \label{fig:6:Problem_1}
+\end{figure}
+
+% - Pb 2 changes are on all arms at a time
+\textbf{Problem $\bm 2$.} (see Figure~\ref{fig:6:Problem_2}). This problem is close to Problem 1, with a minimum optimality gap of $0.1$.
+However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gap $\Delta=0.1$.
+The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$}).
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
+    \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).}
+    \label{fig:6:Problem_2}
+\end{figure}
 
@@
 We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
-To the authors knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
+To our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
 Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
@@
 Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
-This study is supported by both theoretical and empirical results.
-%
-Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing  much in terms of regret.
+This study is supported by both theoretical and empirical results presented in this chapter.
+%
+In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
+it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{i,t} |\mu_i(t) - \mu_i(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{i,j} |\mu_i(t) - \mu_j(t)|$.
+
+Moreover, we can use the two problems $1$ and $2$ to illustrate the expected empirical behavior of our proposal \GLRklUCB.
+Problem $1$ has only \emph{local changes}, meaning that at any change-point, only one arm sees its distribution change, while problem $2$ has only \emph{global changes}, meaning that all arms see their distribution change at each change-point.
+They illustrate both the extreme cases of having $C_T = \Upsilon_T$ (for problem $1$) and $C_T = K \Upsilon_T$ (for problem $2$).
+\begin{itemize}
+    \item
+On one hand, we expect the \emph{local restart} variant of \GLRklUCB{} to outperform the \emph{global restart} variant for problem $1$, as the intuition suggests that it is sub-optimal to reinitialize the memory of the observations of the $K-1$ arms whose distributions did not change after a change-point is detected (as the \emph{global restart} variant do).
+    \item
+On the other hand, we expect the \emph{global restart} variant to outperform the \emph{local restart} variant for problem $2$, as detecting a change one any arm is enough to know that all arms changed and to reinitialize the memory of all arms, and thus the intuition suggests that it is sub-optimal to reinitialize only the memory of the observation of the arm on which the change-point was detected (as the \emph{local restart} variant do).
+\end{itemize}
+%
+Of course, on a given problem, without additional prior knowledge on the difficulty of the problem at hand, the algorithm cannot know which situation is more likely to happen, only local changes ($C_T = \Upsilon_T$) or only global changes ($C_T = K \Upsilon_T$) or any intermediate setting between the two extreme cases.
+Thus we do not believe to be able to design a policy that could be uniformly better than the two variants of our proposal \GLRklUCB.
+%
+Surprisingly, our numerical experiments show that the \emph{local} variant of \GLRklUCB{} actually outperforms the \emph{glocal} variant for both problem $1$ and $2$, as shown below in Table~\ref{table:6:totalResults1}, and in other problems as shown in Table~\ref{table:6:totalResults2}.
+%
+Understanding this difference in terms of numerical performance of the two variants is left as future work.
+
+Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing much in terms of regret, in Appendix~\ref{sub:6:IdeasOptimizations}.
 
@@
 
-Sequential change-point detection has been extensively studied in the statistical community (see, \eg, \cite{Basseville93} for a survey).
+Sequential change-point detection has been extensively studied in the statistical community.
+We refer to the book \cite{Basseville93} for a survey.
+%
 In this section, we are interested in detecting changes on the mean of a probability distribution with bounded support.
-Assuming that  we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$.
+
+Assume that we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$.
 We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
@@
 
-\TODOL{We need to PROVE our results for false alarm and delay !.}
-
 \subsection{Presentation of the test}
@@
 %
-Indeed, one can show that $\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right]$,
+Indeed, as proved in Appendix~\ref{app:6:GLR_with_kl}, we can show that in any one-dimensional exponential family, we have
+\[\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right],\]
+%
 where for $k \leq k'$, $\widehat{\mu}_{k:k'} = \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
@@
 % ----------------------------------------------------------------------------
-\subsection{Non-Asymptotics Properties of the B-GLRT}\label{subsec:6:PropGLR}
+\subsection{Non-Asymptotic Properties of the B-GLRT}\label{subsec:6:PropGLR}
 
@@
 %
-We give below two Lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.
+We give below two lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.
+
+\TODOL{We need to PROVE our results for false alarm and delay !.}
 
@@
 Its use for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on the function $\cT$ are also given:  $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
-The proof of Lemma~\ref{lem:6:FalseAlarm}, that actually holds for any sub-Bernoulli distribution, is given in Appendix~\ref{proof:6:FalseAlarm}.
+% The proof of Lemma~\ref{lem:6:FalseAlarm}, that actually holds for any sub-Bernoulli distribution, is given below.
 
@@
 
-\begin{proof}
-    \TODOL{Write it!}
-\end{proof}
+Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
+\begin{equation}
+    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
+\end{equation}
+where $\phi\mu(\lambda)$ is the log moment generating of some one-dimensional exponential family.
+The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
+
+Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}.
+A first key observation is
+\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
+Hence the probability of a false alarm occurring is upper bounded as
+%
+\begin{align*}
+    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
+    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
+    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
+\end{align*}
+%
+where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
+As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
+%
+\begin{align*}
+    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
+\end{align*}
+%
+And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
+
+\begin{lemma}\label{lem:6:ConcFirst}
+    $(X_i)_{i \in \N}$ and $(Y_i)_{i \in \N}$ two independent \iid{} processes with resp. means $\mu$ and $\mu'$ such that
+    \[\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)} \ \ \text{and} \ \ \bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)},\]
+    where $\phi_\mu(\lambda) = \bE_{X \sim \nu^{\mu}}[e^{\lambda X}]$ is the moment generating function of the distribution $\nu^\mu$, which is the unique distribution in an exponential family that has mean $\mu$.
+    %
+    Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family.
+Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
+    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
+    where $\cT$ is the function defined in \eqref{def:6:function_T}.
+\end{lemma}
 
@@
 the number of selections of arm $i$, and
-\[\widehat{\mu}_i(t) = (1/n_i(t)) \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
+\[\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
 their empirical mean (if $n_i(t)\neq0$).
@@
 %
-In this work, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
+For this algorithm, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
 %
@@
 
-The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by (\cite{CaoZhenKvetonXie18}), whereas \CUSUMUCB{} performs randomized exploration.
+The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs randomized exploration.
 %
@@
 \begin{proposition}\label{prop:6:EnoughSamples}
-    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $i$ (\ie, for a $k\in\{1,\dots,\NCi\}$, one has $\tau_i^{(k)} = \tau_i(t) < s \leq t < \tau_i^{(k+1)}$) it holds that $n_i(t) - n_i(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor$.
+    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $i$ (\ie, for a $k\in\{1,\dots,\NCi\}$, one has $\tau_i^{(k)} = \tau_i(t) < s \leq t < \tau_i^{(k+1)}$) it holds that
+    \begin{equation}
+        n_i(t) - n_i(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor.
+    \end{equation}
 \end{proposition}
@@
 %-----------------------------------------------------------------------------
-\subsection{A Unified Regret Analysis}
-
-\TODOL{Rewrite, include the proof of at least one case here! No need to be that quick for the thesis, we only wrote it like this for the COLT version \cite{Besson2019GLRT}.}
+\subsection{Sketch of the unified regret analysis}
 
@@
 %
-\begin{align*}
+\begin{align}\label{eq:6:GeneRegretBound}
     R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
     & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
-\end{align*}
+\end{align}
 %
@@
 
+\begin{itemize}
+\item
 Term $(a)$ is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
@@
 
+\item
 Term $(b)$ is related to the control of the detection delay, which is more tricky to obtain under the \GLRklUCB{} adaptive sampling scheme,
@@
 Using some elementary calculus and a concentration inequality given in Lemma~\ref{lem:6:Chernoff2arms}, we can finally prove that $(b) \leq \delta$.
-%
+\end{itemize}
+
 Finally, the ``bad event'' is unlikely: $\bP(\cE_T^c) \leq 2\EffChange \delta$.
@@
 
-Assumption~\ref{ass:6:LongPeriodsGlobal} below is easy to interpret and standard in non-stationary bandits.
+The Assumption~\ref{ass:6:LongPeriodsGlobal} below is easy to interpret and standard in non-stationary bandits.
 It requires that the distance between two consecutive breakpoints is large enough: how large depends on the magnitude of the largest change that happen at those two breakpoints.
@@
 
-\begin{assumption}\label{ass:6:LongPeriodsGlobal}
+\begin{assumption}[Enough samples between two change-points]
+    \label{ass:6:LongPeriodsGlobal}
     Define
@@
 
-We gave in Section~\ref{sec:6:RegretAnalysis} the following decomposition of the regret $R_T$,
-%
-\begin{align}\label{eq:6:GeneRegretBound}
-    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
-    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
-\end{align}
+% Our analysis relies on the general regret decomposition~\eqref{eq:6:GeneRegretBound}, with the following appropriate good event.
+
+% We gave in Section~\ref{sec:6:RegretAnalysis} the following decomposition of the regret $R_T$,
+% %
+% \begin{align}\label{eq:6:GeneRegretBound}
+%     R_T \leq T \bP\left(\cE_T^c\right) + \alpha T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
+%     & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
+% \end{align}
 
@@
     & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\times \\
-     &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(k)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \widehat{\tau}^{(k)})\right)\Big] \\
+    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(k)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \widehat{\tau}^{(k)})\right)\Big]
+\end{align*}
+\begin{align*}
     & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_i(\tau^{(k+1)})}\indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\Big]  \\
@@
 
-\subsubsection{Controlling the probability of the good event: Proof of Lemma~\ref{lem:6:GoodEventGlobal}}
+\subsubsection{Controlling the probability of the good event: proof of Lemma~\ref{lem:6:GoodEventGlobal}}
 
@@
     \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T}
-    \bP\left(\left. \widehat{\tau}^{(k)} \notin \{ \tau^{(k)}+1, \dots, \tau^{(k)} + d^{(k)} \} \right|\; \cC^{(k-1)}\right) \\
-    \\ & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \leq \tau^{(k)} \;|\; \cC^{(k-1)}\right)}_{(a)} + \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)} \;|\; \cC^{(k-1)}\right)}_{(b)}.
+    \bP\left(\left. \widehat{\tau}^{(k)} \notin \{ \tau^{(k)}+1, \dots, \tau^{(k)} + d^{(k)} \} \right|\; \cC^{(k-1)}\right)
 \end{align*}
+And thus we have
+\begin{align*}
+    \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \leq \tau^{(k)} \;|\; \cC^{(k-1)}\right)}_{(a)} \;\; + \;\; \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)} \;|\; \cC^{(k-1)}\right)}_{(b)}.
+\end{align*}
 %
@@
 %
-$\widehat{\tau}^{(k)} \leq \tau^{(k)}$ implies that there exists an arm whose associated change point detector has experienced a false-alarm:
+We have $\widehat{\tau}^{(k)} \leq \tau^{(k)}$ and it implies that there exists an arm whose associated change point detector has experienced a false-alarm:
 %
@@
 % ----------------------------------------------------------------------------
-\subsubsection{Proof of Lemma~\ref{lem:6:FalseAlarm}} \label{proof:6:FalseAlarm}
-
-Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
-\begin{equation}
-    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
-\end{equation}
-where $\phi\mu(\lambda)$ is the log moment generating of some one-dimensional exponential family.
-The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
-
-Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}. A first key observation is
-\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
-Hence the probability of a false alarm occurring is upper bounded as
-%
-\begin{align*}
-    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
-    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
-    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
-\end{align*}
-%
-where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
-As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
-%
-\begin{align*}
-    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
-\end{align*}
-%
-And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
-
-\begin{lemma}\label{lem:6:ConcFirst}
-    $(X_i)_{i \in \N}$ and $(Y_i)_{i \in \N}$ two independent \iid{} processes with resp. means $\mu$ and $\mu'$ such that
-    \[\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)} \ \ \text{and} \ \ \bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)},\]
-    where $\phi_\mu(\lambda) = \bE_{X \sim \nu^{\mu}}[e^{\lambda X}]$ is the moment generating function of the distribution $\nu^\mu$, which is the unique distribution in an exponential family that has mean $\mu$.
-    %
-    Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family.
-Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
-    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
-    where $\cT$ is the function defined in \eqref{def:6:function_T}.
-\end{lemma}
-
-\paragraph{Proof of Lemma~\ref{lem:6:ConcFirst}.}
-%
-Using the same construction as in the proof of Theorem~14 in \cite{KK18Martingales}, one can prove that for every $\lambda \in I$ (for an interval $I$), there exists a non-negative super-martingale $M^\lambda (s)$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ that satisfies $\bE[M^\lambda(s)] \leq 1$ and
-\[\forall s \in \N^*, \ \ M^\lambda (s) \geq e^{\lambda [s\kl(\widehat{\mu}_s,\mu) - 3\ln(1+\ln(s))] - g(\lambda)}\]
-for some function $g : I \rightarrow \R$. This super-martingale is of the form
-\[M^\lambda (s) = \int e^{\eta\sum_{i=1}^s X_i - \phi_\mu(\lambda)s} d\pi(\eta)\]
-for a well-chosen probability distribution $\pi$, and the function $g$ can be chosen to be any
-\begin{eqnarray*}
-    g_{\xi} : \left[0; 1/(1+\xi)\right] & \longrightarrow & \R \\
-    \lambda & \mapsto & \lambda(1+\xi) \ln \left(\frac{\pi^2}{3(\ln(1+\xi))^2}\right) -  \ln(1 - \lambda(1 + \xi))
-\end{eqnarray*}
-for a parameter $\xi \in [0,1/2]$.
-
-Similarly, there exists an independent super-martingale $W^\lambda (r)$ w.r.t. the filtration $\cF_r' = \sigma(Y_1,\dots,Y_r)$ such that
-\[\forall r \in \N^*, \ \ W^\lambda (r) \geq e^{\lambda [r\kl(\widehat{\mu}'_r,\mu) - 3\ln(1+\ln(r))] - g(\lambda)},\]
-for the same function $g(\lambda)$.
-In the terminology of \cite{KK18Martingales}, the processes $\bm{X}(s) = s \, \kl(\widehat{\mu}_s,\mu) - 3\ln(1+\ln(s))$ and $\bm{Y}(s) = r \, \kl(\widehat{\mu}_r,\mu) - 3\ln(1+\ln(r))$ are called $g$-DCC for Doob-Cram\'er-Chernoff, as Doob's inequality can be applied in combination with the Cram\'er-Chernoff method to obtain deviation inequalities that are uniform in time.
-
-Here we have to modify the technique used in their Lemma 4 in order to take into account the two stochastic processes, and the presence of super-martingales instead of martingales (for which Doob inequality still works).
-One can write
-%
-\begin{align*}
-    &\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \\
-    & \leq \bP\left(\exists r \in \N^* : M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right) \\
-    & = \lim_{n \rightarrow \infty} \bP\left(\exists r \in \{1,\dots,n\} :  M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right) \\
-    & = \lim_{n \rightarrow \infty} \bP\left(\sup_{r \in \{1,\dots,n\}}  M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right).
-\end{align*}
-%
-Using that $\tilde{M}(r) = M^\lambda(s)W^\lambda(r)$ is a super-martingale with respect to the filtration \[\tilde{\cF}_r = \sigma(X_1,\dots,X_s, Y_1,\dots,Y_r),\] one can apply Doob's maximal inequality to obtain
-\begin{eqnarray*}
-    \bP\left(\sup_{r \in \{1,\dots,n\}}  M^\lambda(s)W^\lambda(r) > e^{\lambda u - 2g(\lambda)}\right) &\leq& e^{-(\lambda u - 2g(\lambda))}\bE[\tilde{M}(1))] \\
-    & = & e^{-(\lambda u - 2g(\lambda))}\bE[M^\lambda(s)W^\lambda(1)] \\
-    & \leq & e^{-(\lambda u - 2g(\lambda))},
-\end{eqnarray*}
-using that $M^\lambda(s)$ and $W^\lambda(1)$ are independent and have an expectation smaller than $1$.
-
-Putting things together yields
-\[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + u\right) \leq e^{-\left(\lambda u - 2g_{\xi}(\lambda)\right)},\]
-for any function $g_{\xi}$ defined above.
-%
-The conclusion follows by optimizing for both $\lambda$ and $\xi$, using Lemma~18 in \cite{KK18Martingales}.
-
-% ----------------------------------------------------------------------------
 
@@
         \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$,
-        % \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
-        % \item Choosing $\alpha = \sqrt{\frac{\EffChange\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\EffChange T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
+        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
+        \item Choosing $\alpha = \sqrt{\frac{\EffChange\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\EffChange T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
         \item Choosing $\alpha = \sqrt{\frac{K\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{K\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$.
@@
 
-\TODOL{Only two variants in this corollary, not the four variants?}
+% \TODOL{Only two variants in this corollary, not the four variants?}
 
@@
 
+\TODOL{Maybe I can give this part in Appendix, as it is quite similar to the proof for Global changes, or simply refer to the paper where we give this in details ?}
+
 \subsubsection{Proof of Theorem~\ref{thm:6:mainRegretBound}}
@@
 Our analysis relies on the general regret decomposition~\eqref{eq:6:GeneRegretBound}, with the following appropriate good event.
+%
 \begin{equation}\label{def:6:GoodEvenLocal}
@@
 To upper bound Term B, we introduce a sum over all arms and rewrite the sum in $t$ as a sum of consecutive intervals $[\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]$.
+%
 The decomposition using furthermore that on $\cE_T$, $\tau_i^{\ell} \leq \hat \tau_i^ \ell \leq \tau_i^{\ell+1}$ for all changes, by Assumption~\ref{ass:6:LongPeriods}.
@@
 
-Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} both show that there exists a tuning of $\alpha$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
-For very particular instances such that $\Upsilon_T = \EffChange$, \ie, at each break-point only one arm changes (\eg, problem $1$ in Section~\ref{sec:6:NumericalExperiments}), the theory advocates the use of local changes.
-%
-Indeed, while the regret guarantees obtained are similar, those obtained for local changes hold for a wider variety of problems as Assumption~\ref{ass:6:LongPeriods} is less stringent than \ref{ass:6:LongPeriodsGlobal}.
+The regret bounds we obtain for the two variants of our proposal \GLRklUCB{},
+Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound},
+both show that there exists a tuning of $\alpha$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
+%
+For very particular instances such that $\Upsilon_T = \EffChange$, \ie, at each break-point only one arm changes (\eg, problem $1$ from Figure~\ref{fig:6:Problem_1}), the theory advocates the use of local changes.
+%
+Indeed, while the regret guarantees obtained are similar, those obtained for local changes hold for a wider variety of problems as Assumption~\ref{ass:6:LongPeriods} is less stringent than Assumption~\ref{ass:6:LongPeriodsGlobal}.
 %
@@
 However, we only obtain regret upper bounds -- thus providing a theoretical safety net for both variants of our algorithm, and the practical story is different, as discussed in Section~\ref{sec:6:NumericalExperiments}.
-We find that \GLRklUCB{} performs better with local restarts.
-
-One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, our regret bounds hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
-Hence those guarantees are valid on ``easy'' problem instances only, with few changes of a large magnitude (\eg, not for problems $3$ or $5$).
-%
-However, this does not prevent our algorithms from performing well on more realistic instances, and numerical experiments support this claim.
-% (see problem~3 in Section~\ref{sec:6:NumericalExperiments}).
-Note that \MUCB{} (\cite{CaoZhenKvetonXie18}) is also analyzed for the same type of unrealistic assumptions, while its practical performance is illustrated beyond those.
+We find that \GLRklUCB{} performs better with local restarts, uniformly on all problems.
+
+One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, our regret bounds hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
+Hence those guarantees are valid on ``easy'' problem instances only, with few changes of a large magnitude, and in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
+%
+However, this does not prevent our algorithms from performing well on more realistic instances, and numerical experiments support this claim,
+as illustrated for instance with problem $3$ presented in the next Section~\ref{sec:6:NumericalExperiments}).
+%
+Note that \MUCB{} \cite{CaoZhenKvetonXie18} is also analyzed for the same type of unrealistic assumptions, while its practical performance is illustrated beyond those.
 
@@
 
-We include in our study two algorithms designed for the classical MAB, \klUCB{} (\cite{Garivier11KL}) and Thompson sampling (\cite{AgrawalGoyal11,Kaufmann12Thompson}),
+We include in our study two algorithms designed for the classical MAB, \klUCB{} \cite{Garivier11KL}, and Thompson sampling \cite{AgrawalGoyal11,Kaufmann12Thompson},
 as well as an ``oracle'' version of \klUCB, that we call Oracle-Restart. This algorithm knows the exact locations of the breakpoints, and restarts \klUCB{} at those locations (without any delay).
@@
 For this last algorithm, the discount factor $\gamma=0.75$ suggested by the authors was performing significantly worse on our problem instances.
-%More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).
+More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).
 
@@
 Our main goal is to compare against \emph{actively adaptive algorithms}.
-We  include \CUSUMklUCB{} (\cite{LiuLeeShroff17}), tuned with $M=150$ and $\varepsilon=0.1$ for easy problems ($1,2,4$) and $\varepsilon=0.001$ for hard problems ($3,5$),
+We include \CUSUMklUCB{} \cite{LiuLeeShroff17}, tuned with $M=150$ and $\varepsilon=0.1$ for easy problems ($1,2,4$) and $\varepsilon=0.001$ for hard problems ($3,5$),
 % based on a prior knowledge of the difficulty of the problems,
@@
 %
-Finally, we include \MklUCB{} (\cite{CaoZhenKvetonXie18}),
-% as a simple adaptation of the Monitored-UCB (\MUCB) algorithm,
+Finally, we include \MklUCB{} \cite{CaoZhenKvetonXie18},
+% as a simple adaptation of the \MUCB algorithm,
 tuned with $w=150$, based on a prior knowledge of the problems as the formula using $\delta_{\min}$ given in the paper is too large for small horizons (on all our problem instances), a threshold $b=\sqrt{w \log(2 K T)}$ and $\gamma=\sqrt{\Upsilon_T K (2 b + 3 \sqrt{w}) / (2 T)}$ as suggested by Remark~4 in the paper.
@@
 from Corollaries \ref{cor:6:Global} and \ref{cor:6:Local}.
-% Both choices of $\alpha$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
-The constant is set to $\alpha_0 = 0.05$
+Both choices of $\alpha$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
+the constant is set to $\alpha_0 = 0.05$
 (we show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\alpha\leq0.1$).
@@
 
-\paragraph{Illustrations of our benchmark.}\label{app:6:Figures}
-
-We give below the plots of the cumulated mean regret $R_t$ (as a function of $t$), as well as plots of histograms of the distribution of final regret $R_T$, for the considered algorithms.
-
-% - Pb 1 changes are only on one arm at a time
-\textbf{Problem $\bm 1$.} There are $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
-The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1}.
-% (and Figure~\ref{fig:6:Problem_1} in Appendix),
-Note that changes happen on only one arm (\ie, $C_T=\Upsilon_T=4$),
-and the optimal arm changes once at $t=\tau_2^{(1)}=2000$, with a large gap $\Delta=0.6$.
-
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
-    \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).}
-    \label{fig:6:Problem_1}
-\end{figure}
-
-% - Pb 2 changes are on all arms at a time
-\textbf{Problem $\bm 2$.} (see Figure~\ref{fig:6:Problem_2}). This problem is close to Problem 1, with a minimum optimality gap of $0.1$.
-However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gap $\Delta=0.1$.
-The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$}).
-
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
-    \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).}
-    \label{fig:6:Problem_2}
-\end{figure}
+\paragraph{Illustrations of the rest of our benchmark.}\label{par:6:benchmark2}
+
+We continue here the presentation of the other problems of our benchmark, see Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} above for the two problems $1$ and $2$.
 
@@
 Note that the assumptions of our regret upper bounds are violated, as well as the assumptions for the analysis of \MUCB{} and \CUSUMUCB.
-This problem is inspired from Figure~3 of \cite{CaoZhenKvetonXie18}, where the synthetic data was obtained from manipulations on a real-world database of clicks from \emph{Yahoo!}.
+It is interesting to say that this problem is inspired from Figure~3 of \cite{CaoZhenKvetonXie18}, where the synthetic data was obtained from manipulations on a real-world database of clicks from \emph{Yahoo!}.
 
@@
 
-%     \caption{History of means of arms for three problems, $\mu_i(t)$ for $i\in\{1,\dots,K\}$ and $t\in\{1,\dots,T\}$.}
-%     \label{fig:6:Problems_123}
-% \end{figure}
-
 % - Pb 4 changes are on all or almost arms at a time, but sequences don't have same length
@@
 
-Like problem $3$, this hard problem is inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!} (see Figure~3 from \cite{LiuLeeShroff17}).
-It is harder, with $\Upsilon=81$ change-points on $K=5$ arms for a longer horizon of $T=100000$.
+Like problem $3$, this harder problem is also inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!}, but from our other competitor paper, see Figure~3 from \cite{LiuLeeShroff17}.
+%
+It is much harder, with $\Upsilon=81$ change-points on $K=5$ arms for a longer horizon of $T=100000$.
 Some arms change at almost every time steps, for a total number of breakpoints $C=179$, but the optimal arm is almost always the same one (\textcolor{red}{arm $0$, with $\bullet$}).
@@
 
+\TODOL{Justify MORE AND WAY MORE about all these results!}
+
 \paragraph{Regret plots.}
@@
 
-On very simple problems, like problem 1, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there is more changes, we clearly observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much more than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
+On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there is more changes, we clearly observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much more than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
 Our approach, with the two options of \textbf{Local} or \textbf{Global} restarts, performs very closely to the oracle for problem $4$.
@@
 
+% -----------------------------------------------------------------
 \begin{figure}[h!]  % [htbp]
@@
 
-In this chapter, we studied the piece-wise stationary bandit model, and reviewed existing work.
+In this chapter, we studied and presented the piece-wise stationary bandit model, and reviewed existing work.
+%
 We proposed a new algorithm for this problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
@@
 We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as our proposal \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
-Finally, we would like to study in the future possible extension of our approach to the slowly varying model (\cite{WeiSrivastava18}).
-Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm (\cite{Besson2018ALT}), to efficiently tackle the piece-wise stationary distributed multi-player bandit model introduced in \cite{Wei2018}.
+Finally, we would like to study in the future possible extension of our approach to the slowly varying model, as introduced in \cite{WeiSrivastava18}.
+
+\TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-player bandit model introduced in \cite{Wei2018} by using \MCTopM{} and \GLRklUCB{} with each other? }
+
+% Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm \cite{Besson2018ALT}, to efficiently tackle the piece-wise stationary distributed multi-player bandit model introduced in \cite{Wei2018}.
 
@@
 %
-The page \texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html} explains how to reproduce the experiments used for this Section.
+All the experiments used for this chapter were performed using SMPyBandits,
+and the page \texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html} explains how to reproduce them.

### 2019-03-13 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Again lots of work on Chapter 6, see #14

@@
 For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
-It is sometimes referred to as the abruptly changing MAB, or switching environments \cite{MellorShapiro13}.
+It is sometimes referred to as the abruptly changing \cite{WeiSrivastava18Abruptly}, or switching environments \cite{MellorShapiro13}.
 %
@@
 Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector.
-For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 \cite{Auer02} used in previous works.
+For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 used in previous works.
 For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest.
@@
 A policy $\pi$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
-The performance of $\pi$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\pi$:
+The performance of $\pi$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\pi$,
 %
@@
 In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
-If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint by
+If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint recursively by
 \begin{equation}
@@
 %
-Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, that is on any stationary segments, the rewards $(X_{i,t})$ associated to each arms are \iid{} (hence the name of \emph{piece-wise stationary} problems).
+Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, that is on any stationary segments, the rewards $(X_{i,t})$ associated to each arm are \iid{} (hence the name of \emph{piece-wise stationary} problems).
 
@@
 
-Introducing the number of change-points on arm $i$, defined as $\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
+We define the \emph{number of change-points} on arm $i$,
+$\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$,
+for which it clearly holds that
 $\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms.
@@
 However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
-
-\TODOL{Speak about the article \cite{WeiSrivastava18} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}
 
@@
 Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
-This study is supported by both theoretical and empirical results presented in this chapter.
+% This study is supported by both theoretical and empirical results presented in this chapter.
 %
@@
 
+\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
+%
+In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
+we will assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
+Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
+where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
+and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
+%
+This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
+Empirically, we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
+
+% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}
+
 % -----------------------------------------------------------------
@@
 %
-where for $k \leq k'$, $\widehat{\mu}_{k:k'} = \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
+where for $k \leq k'$, $\widehat{\mu}_{k:k'} := \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
 %
@@
 
-Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR} but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence between two elements in that exponential family that have mean $x$ and $y$.
-For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to \eqref{def:6:GLR} with $\kl(x,y) = 2(x-y)^2$ when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
-
-In this section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
-Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $2(x-y)^2$.
+Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR}, but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence $d(x,y)$ between two elements in that exponential family that have mean $x$ and $y$.
+For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to the stopping time \eqref{def:6:GLR} with the choice $d(x,y) = 2(x-y)^2$, when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
+Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
+%
+In the next section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
 
 % ----------------------------------------------------------------------------
-\subsection{Non-Asymptotic Properties of the B-GLRT}\label{subsec:6:PropGLR}
+\subsection{Non-asymptotic properties of the B-GLRT}\label{subsec:6:PropGLR}
 
@@
 We give below two lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.
-
-\TODOL{We need to PROVE our results for false alarm and delay !.}
 
@@
 \end{equation}
-where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$.
-%
-And for any $x \ge 0$, $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise.
-The function $\cT$ is easy to compute numerically.
-The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96} (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
-%
-Its use for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on the function $\cT$ are also given:  $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
-% The proof of Lemma~\ref{lem:6:FalseAlarm}, that actually holds for any sub-Bernoulli distribution, is given below.
+where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$,
+%
+we define $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise, for any $x \ge 0$,
+and with the value $\zeta(2) = \pi^2 / 6$.
+%
+Even if it does not have a closed form expression, the function $\cT$ is easy to compute numerically.
+%
+The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96}, which is the inverse of $x \mapsto x \exp(x)$, as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
+
+The use of $\cT$ for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on this function $\cT$ are also given:
+$\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
 
@@
 
+\begin{proof}
 Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
@@
 \end{equation}
-where $\phi\mu(\lambda)$ is the log moment generating of some one-dimensional exponential family.
+where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family.
 The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
 
-Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}.
-A first key observation is
+Let us go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}.
+We first observe that
+% A first key observation is
 \[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
@@
     \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
-    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
+    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)
+\end{align*}%
+\begin{align*}
     & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
@@
 And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
+%
+\end{proof}
 
@@
 \end{lemma}
+
+This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
 
@@
 \begin{proof}
-    \TODOL{Write it!}
+    \TODOL{Write it ?}
 \end{proof}
@@
 with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}.
-We define the (\klUCB{}-like) index used by our algorithm, by denoting
+We define the (\klUCB{} like) index used by our algorithm, by denoting
 $\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
@@
 %
-With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ (if $t>1$ else $f(t)=0$),
+With the exploration function $f(t)$ defined as $\ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
 the index is defined as
@@
 \begin{equation}\label{def:6:UCB}
-    \UCB_i(t) := \max \bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \bigr\}.
-\end{equation}
-
-\paragraph{Two options.}
+    \UCB_i(t) := \max \Bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \Bigr\}.
+\end{equation}
+
+%-----------------------------------------------------------------------------
+\paragraph{Two options for restarts.}
 %
@@
 $f(t-\tau_i(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{i=1}^K n_i(t)$.
+
+%-----------------------------------------------------------------------------
+\paragraph{Threshold function $\beta$.}
+
+\TODOL{I should link to Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} where I present two experiments which justifies this choice of the threshold function $\beta(n,\delta)$: it works fine and it's much simpler to implement numerically than the one we chose to analyze mathematically!}
 
@@
 
+\TODOL{I should link to Appendix~\ref{sec:6:exploringDifferentForcedExplorationMechanisms} where I present two experiments which justifies this choice of the forced exploration mechanism: it works fine and it's the simplest to implement numerically and to analyze mathematically!}
+
 %-----------------------------------------------------------------------------
 %-----------------------------------------------------------------------------
-\section{Regret results and proofs for \GLRklUCB}
+\section{Finite-time upper-bounds on the regret of \GLRklUCB}
 \label{sec:6:RegretAnalysis}
@@
 This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the finite-time regret bounds for the two variants, with their proofs, which both follow the skeleton of the unified analysis.
+% We highlight that our results are all non-asymptotic, contrarily to what is given for \MUCB{} \cite{CaoZhenKvetonXie18} and \CUSUMUCB{} \cite{LiuLeeShroff17}.
 
@@
 
-In this section, we sketch a unified proof for Theorem~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound}, whose detailed proofs are given in Sections \ref{proof:6:mainRegretBoundGlobal} and \ref{proof:6:mainRegretBound} respectively.
+In this section, we sketch a unified proof for the two Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given below, whose detailed proofs are given in Sections \ref{proof:6:mainRegretBoundGlobal} and \ref{proof:6:mainRegretBound} respectively.
 We emphasize that our approach is significantly different from those proposed by \cite{CaoZhenKvetonXie18} for \MUCB{} and by \cite{LiuLeeShroff17} for \CUSUMUCB.
@@
 %
-Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
+Introducing some \emph{good event} $\cE_T$, to be specified in each case, and its complementary $\cE_T^c$, one can write the following decomposition, that highlights two important terms $(A)$ and $(B)$
 %
@@
     R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
-    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
+    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}. \nonumer
 \end{align}
@@
 where $\widehat{\tau}_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
-Introducing the event $\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\}$ that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition:
+Introducing the event
+\[\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\},\]
+that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition
+%
 \begin{align}
@@
 
-Finally, the ``bad event'' is unlikely: $\bP(\cE_T^c) \leq 2\EffChange \delta$.
+Finally, the ``bad event'' is unlikely, $\bP(\cE_T^c) \leq 2\EffChange \delta$.
+By putting together the three pieces, that are a bound on $(A)$, on $(B)$ and on $\bP(\cE_T^c)$, we obtain the desired finite-time upper-bound on the regret of our proposal \GLRklUCB.
 
@@
 
-The Assumption~\ref{ass:6:LongPeriodsGlobal} below is easy to interpret and standard in non-stationary bandits.
+\begin{assumption}[Enough samples between two global change-points]
+    \label{ass:6:LongPeriodsGlobal}
+    Define
+    \begin{equation}\label{def:6:DelayGlobal}
+        d^{(k)} = d^{(k)}(\alpha,\delta) = \left\lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \right\rceil.
+    \end{equation}
+    %
+    Then we assume that for all $k \in \{1,\dots,\Upsilon_T\}$,
+    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
+\end{assumption}
+
+Assumption~\ref{ass:6:LongPeriodsGlobal} is easy to interpret and standard in non-stationary bandits.
 It requires that the distance between two consecutive breakpoints is large enough: how large depends on the magnitude of the largest change that happen at those two breakpoints.
@@
 and the $\Delta^{(k)}$ terms expressing the hardness of the change-point detection problem.
-
-\begin{assumption}[Enough samples between two change-points]
-    \label{ass:6:LongPeriodsGlobal}
-    Define
-    \begin{equation}\label{def:6:DelayGlobal}
-        d^{(k)} = d^{(k)}(\alpha,\delta) = \left\lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \right\rceil.
-    \end{equation}
-    %
-    Then we assume that for all $k \in \{1,\dots,\Upsilon_T\}$,
-    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
-\end{assumption}
 
@@
 
-\subsection{Proofs for \GLRklUCB{} with Global Changes}
+\subsection*{Proof for \GLRklUCB{} with Global Changes}
 \label{proof:6:mainRegretBoundGlobal}
@@
 
-\subsubsection{Proof of Theorem~\ref{thm:6:mainRegretBoundGlobal}.}
+% \subsubsection{Proof of Theorem~\ref{thm:6:mainRegretBoundGlobal}.}
 
@@
 Thus, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_{k^*}^{(k)}$, one can write
-%
+
 \begin{align*}
@@
 
-Using Pinsker's inequality and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = | \Delta_i^k |$), one can write
+Using Pinsker's inequality and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = |\Delta_i^k|$), one can write
 %
@@
 %
-Using Lemma~\ref{lem:6:Chernoff2arms} (given above in Appendix~\ref{proof:6:Chernoff2arms}) and a union bound, the first term in the right hand side is upper bounded by $\delta$ (as $\beta(r+s,\delta) \geq \beta(r,\delta) \geq \log(3s\sqrt{s}/\delta)$).
+Using Lemma~\ref{lem:6:Chernoff2arms} stated in Appendix~\ref{proof:6:Chernoff2arms}, and a union bound, the first term in the right hand side is upper bounded by $\delta$ (as $\beta(r+s,\delta) \geq \beta(r,\delta) \geq \log(3s\sqrt{s}/\delta)$).
 For the second term, we use the observation
 \[\frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(k)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
-and, using that $\Delta^{(k)} = | \Delta_i^k|$, one obtains
+and, using that $\Delta^{(k)} = |\Delta_i^k|$, one obtains
 %
@@
 
-Assumption~\ref{ass:6:LongPeriods} requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
-Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{i=1}^K \NCi \leq K \Upsilon_T$.
-
-\begin{assumption}\label{ass:6:LongPeriods}
+\begin{assumption}[Enough samples between two local change-points]
+    \label{ass:6:LongPeriods}
     Define
@@
 
+Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to our knowledge our analysis is the first one to be given for such assumption on the problem difficulty.
+%
+It requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
+Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{i=1}^K \NCi \leq K \Upsilon_T$.
+
 \begin{theorem}\label{thm:6:mainRegretBound}
@@
 \end{theorem}
+
+We can directly obtain different regret upper-bounds for different choices of the two parameters $\alpha$ and $\delta$.
+We prefer to state the four cases, to highlight the modularity of the result given by Theorem~\ref{thm:6:mainRegretBound}.
 
@@
 
-% \TODOL{Only two variants in this corollary, not the four variants?}
-
-% ------------------------------------------------------------------------------------
-\subsection{Proof for \GLRklUCB{} with Local Changes}
-\label{proof:6:mainRegretBound}
-
-\TODOL{Maybe I can give this part in Appendix, as it is quite similar to the proof for Global changes, or simply refer to the paper where we give this in details ?}
-
-\subsubsection{Proof of Theorem~\ref{thm:6:mainRegretBound}}
-
-Our analysis relies on the general regret decomposition~\eqref{eq:6:GeneRegretBound}, with the following appropriate good event.
-%
-\begin{equation}\label{def:6:GoodEvenLocal}
-    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
-\end{equation}
-where $\widehat{\tau}_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined as in Assumption~\ref{ass:6:LongPeriods}.
-Using this assumption, one can prove the following.
-
-\begin{lemma}\label{lem:6:GoodEvent}
-    The ``bad event'' is highly unlikely, as it satisfies $\bP((\cE_T(\alpha,\delta))^c) \leq 2\delta \EffChange$.
-\end{lemma}
-
-We now turn our attention to upper bounding terms $(A)$ and $(B)$ in \eqref{eq:6:GeneRegretBound}.
-
-\paragraph{Upper bound on term $\bm{(A)}$.}
-%
-\begin{align*}
-    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \kl\left(\widehat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau_{i_t^*}(t))\right)\right] \\
-    & \leq \sum_{i=1}^K \sum_{k=1}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic(i_t^* =i) \indic(\tau_i(t) = \widehat{\tau}_i^{(\ell)})\indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right)\right] \\
-    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right) + d_i^{(\ell)}(T)\right] \\
-    & \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}d_i^{(\ell+1)}(T) + \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\bE\left[\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}}\indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right)\right],
-\end{align*}%
-%
-where we introduce the event $\cC_i^{(\ell)}$ that all the changes up to the $\ell$-th have been detected:
-\begin{equation}\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(j)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(\ell)} \right] \right\}.\label{def:6:EventCi}\end{equation}
-Clearly, $\cE_T \subseteq \cC_i^{(\ell)}$ and $\cC_i^{(\ell)}$ is $\cF_{\widehat{\tau}_i^{(\ell)}}$-measurable.
-Observe that conditionally to $\cF_{\widehat{\tau}_i^{(\ell)}}$, when $\indic(\cC_i^{(\ell)})$ holds, $\widehat{\mu}_{i}(t)$ is the average of samples that have all mean $\mu_i^{(\ell)}$.
-Thus, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(\ell)}$, one can write
-\begin{align*}
-    &\bE\left[\left.\indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right) \right| \cF_{\widehat{\tau}_i^{(\ell)}}\right]\\
-    & = \indic(\cC_i^{(\ell)}) \sum_{t=\hat{\tau}_i^{(\ell)}}^{\tau_i^{(\ell+1)}} \bE\left[\indic\left(n_{i}(t) \kl\left(\widehat{\mu}_{i}(t), \mu_{i}\right) \geq f(t - \widehat{\tau}_{i}^{(\ell)})\right) \;|\; \cF_{\widehat{\tau}_i^{(\ell)}}\right]
-% \end{align*}%
-% \begin{align*}
-    & \leq \indic(\cC_i^{(\ell)}) \sum_{t=1}^{\tau_i^{(\ell+1)} - \hat{\tau}_i^{(\ell)}} \bP\left(\exists s \leq t' : s \times \kl(\widehat{\mu}_{s},\mu_i^{(\ell)}) \geq f(t')\right) \\
-    & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
-\end{align*}%
-%
-where the last but one inequality relies on the concentration inequality given in Lemma 2 of \cite{KLUCBJournal}, and the fact that $f(t) = \ln(t) + 3 \ln(\ln(t))$.
-Finally, using the law of total expectation yields
-\begin{equation}
-    (A) \leq \sum_{i=1}^K \sum_{\ell=0}^{\NCi}\left[d_i^{(\ell+1)}(T) + \ln(\ln(T))\right].
-    \label{eq:6:TermAFinal}
-\end{equation}
-
-\paragraph{Upper bound on term $\bm{(B)}$.}
-%
-Recall that $\mu^*_{i,\ell}$ is defined from the statement of Theorem~\ref{thm:6:mainRegretBound} as the smallest value that still outperforms arm $i$ on the the interval between the $\ell$ and $(\ell+1)$-st change of $i$.
-We let $\tilde{\mu}_{i,s}^{(\ell)}$ denote the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}_i^{(\ell)}+1$.
-To upper bound Term B, we introduce a sum over all arms and rewrite the sum in $t$ as a sum of consecutive intervals $[\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]$.
-%
-The decomposition using furthermore that on $\cE_T$, $\tau_i^{\ell} \leq \hat \tau_i^ \ell \leq \tau_i^{\ell+1}$ for all changes, by Assumption~\ref{ass:6:LongPeriods}.
-%
-\begin{align*}
-    (B) & \leq \sum_{i=1}^K\bE\Big[\indic(\cE_T)\sum_{\ell=1}^{\NCi} \sum_{t=\tau_i^{(\ell)}}^{\tau_i^{(\ell+1)}}\left(\mu_{i^*_t}(t) - \overline{\mu}_{i}^{(\ell)}\right)\indic\left(I_t = i, \UCB_i(t) \geq \mu^*_{i,\ell}\right)\Big] \\
-    & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T)\hat{\tau}_i^{(\ell)} + \indic(\cE_T) \sum_{t=\widehat{\tau}_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\indic\left(I_t = i, \UCB_i(t) \geq \mu^*_{i,\ell}\right)\Big] \\
-%     & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+ \sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}_i^{(\ell)}+1}^{\tau_i^{(\ell+1)}}\sum_{s=1}^{t-\widehat{\tau}_i^{(\ell)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \times \kl(\tilde{\mu}_{i,s}^{(\ell)}, \mu^*_{\ell,i}) \leq f(\tau_i^{(\ell+1)} - \widehat{\tau}_i^{(\ell)})\right)\Big] \\
-    & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+ \sum_{i=1}^K\sum_{\ell=1}^{\NCi}\bE\Big[\indic(\cC_i^{(\ell)}) \sum_{s=1}^{n_i(\tau_i^{(\ell+1)})}\indic\left(s \times \kl(\tilde{\mu}_{i,s}^{(\ell)}, \mu^*_{i,\ell}) \leq f(\tau_i^{(\ell+1)} - \tau_i^{(\ell)})\right)\Big].
-\end{align*}%
-%
-The last inequality relies on introducing a sum over $\indic(n_i(t) = s)$ and swapping the sums.
-Conditionally to $\cF_{\widehat{\tau}_i^{(\ell)}}$, when $\cC_i^{(\ell)}$ holds, for $s \in \{1, \ldots, n_i(\tau_i^{(\ell+1)})\}$, $\tilde{\mu}_{i,s}^{(\ell)}$ is the empirical mean from \iid{} observations of mean $\overline{\mu}_i^{(\ell)}$.
-Therefore, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\overline{\mu}_i^{(\ell)}$, it follows from the law of total expectation that
-%
-\begin{align*}
-    (B) & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi} d_i^{(\ell)}
-    + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \sum_{s=1}^{\tau_i^{(\ell+1)}-\tau_i^{(\ell)}} \bP\left(s \times \kl(\widehat{\mu}_{s}, \mu^*_{i,\ell}) \leq f(\tau_i^{(\ell+1)} - \tau_i^{(\ell)})\right).
-\end{align*}%
-%
-As $\mu^*_{\ell,i} > \mu_i^{(\ell)}$ by definition, we can use the same analysis as in the proof of Fact~2 in Appendix~A.2 of \cite{KLUCBJournal} to show that
-% \begin{align*}
-%     (B)_i & \leq \sum_{\ell=1}^{\NCi}d_i^{(\ell)}+\sum_{\ell=1}^{\NCi}\left[ \frac{\ln(T) + 3\ln\ln(T)}{\kl(\mu_i^{(\ell)}, \mu^*_{\ell,i})} + \sqrt{\frac{2\pi(\kl'(\mu_i^{(\ell)},\mu^*_{\ell,i}))^2}{(\kl(\mu_i^{(\ell)}, \mu^*_{\ell,i}))^2}}\sqrt{\ln(T) + 3\ln\ln(T)} + 2\left(\frac{\kl'(\mu_i^{(\ell)},\mu^*_{\ell,i})}{\kl(\mu_i^{(\ell)}, \mu^*_{\ell,i})}\right)^2\right].
-% \end{align*}
-\begin{align}\label{eq:6:TermBFinal}
-    (B) & \leq \sum_{i=1}^K\sum_{\ell=1}^{\NCi}d_i^{(\ell)}+\sum_{i=1}^K\sum_{\ell=1}^{\NCi}\left[ \frac{\ln(T) + 3\ln\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},\mu^*_{i,\ell})} + O\left(\sqrt{\log(T)}\right)\right].
-\end{align}
-The result follows by combining the decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEvent} and the bounds \eqref{eq:6:TermAFinal} and \eqref{eq:6:TermBFinal}.
-
-\subsubsection{Proof of Lemma~\ref{lem:6:GoodEvent}}\label{proof:6:GoodEvent}
-
-With the event $\cC_i^{(\ell)}$ defined in \eqref{def:6:EventCi}, a simple union bound yields
-\begin{align*}
-    \bP(\cE_T^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
-\end{align*}
-
-The final result follows by proving that the terms $(a)$ and $(b)$ are both upper bounded by $\delta$.
-
-\paragraph{Upper bound on $(a)$: controlling the false alarms.}
-Under the bandit algorithm, the change point detector associated to arm $i$ is based on (possibly much) less than $t - \tau_i(t)$ samples from arm $i$, which makes false alarm even less likely to occur.
-More precisely, we upper bound term $(a)$ by
-%
-\begin{align*}
-    (a) & \leq \bP\left(\exists s < t \leq n_i(\tau_i^{(\ell)}) : s \times \kl\left(\tilde{\mu}_{i,1:s}^{(\ell-1)},\tilde{\mu}_{i,1:t}^{(\ell-1)}\right) + (t - s) \times \kl\left(\tilde{\mu}_{i,s+1:t}^{(\ell-1)},\tilde{\mu}_{i,1:t}^{(\ell-1)}\right) > \beta(t, \delta) \;|\; \cC_i^{(\ell-1)}\right) \\
-    & \leq \bP\left(\exists s < t : s \times \kl(\widehat{\mu}_{1:s},\mu_i^{(\ell-1)}) + (t - s) \times \kl(\widehat{\mu}_{s+1:t}, \mu_{i}^{(\ell-1)}) > \beta(t, \delta)\right),
-\end{align*}
-with $\hat\mu_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_i^{(\ell-1)}$.
-Indeed, conditionally to $\cC_i^{(\ell-1)}$, the $n_i(\tau_i^{(\ell)})$ successive observations of arm $i$ arm starting from $\hat \tau_i^{(\ell)}$ are \iid{} with mean $\mu_i^{(\ell-1)}$.
-Using Lemma~\ref{lem:6:ConcFirst}, term $(a)$ is upper bounded by $\delta$.
-
-\paragraph{Upper bound on term $(b)$: controlling the delay.}
-%
-Controlling the detection delay on arm $i$ under an adaptive sampling scheme can be tricky.
-Here we need to leverage the forced exploration (Proposition~\ref{prop:6:EnoughSamples}) to be sure we have enough samples to ensure detection: the effect is that delays will be scaled by the exploration parameter $\alpha$.
-
-First, it follows from Proposition~\ref{prop:6:EnoughSamples} that there exists $\overline{t} \in \left\{\tau_i^{(\ell)}, \dots, \tau_i^{(\ell)} + d_i^{(\ell)} \right\}$ such that
-$n_i(\overline{t}) - n_i(\tau_i^{(\ell)}) = \overline{r}$ where $\overline{r} = \lfloor \frac{\alpha}{K} d_i^{(\ell)}\rfloor$
-(as the mapping $t\mapsto n_i(t) - n_i(\tau_i^{(\ell)})$ is non-decreasing, is $0$ at $t=\tau_i^{(\ell)}$ and its value at $\tau_i^{(\ell)}+d_i^{(\ell)}$ is larger than $\overline{r}$).
-Using that $(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)}) \subseteq (\widehat{\tau}_i^{(\ell)} \geq \overline{t})$,
-the event $(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)})$ further implies that
-\[
-    n_i(\tau_i^{(\ell)}) \, \kl\left(\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)})},\tilde{\mu}^{\ell-1}_{i,n_i(\overline{t})}\right)
-    +  \overline{r} \, \kl\left(\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)}) : n_i(\overline{t})},\tilde{\mu}^{\ell-1}_{i,n_i(\overline{t})}\right) \leq \beta(n_i(\tau_i^{(\ell)}) + \overline{r},\delta),
-\]
-where $\tilde{\mu}^{\ell-1}_{i,s}$ denotes the empirical mean of the $s$ first observation of arm $i$ since the $(\ell-1)$-th restart $\hat{\tau}_i^{(\ell-1)}$ and  $\tilde{\mu}^{\ell-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
-Conditionally to $\cC^{(\ell-1)}_i$, $\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)})}$ is the empirical mean of $n_i(\tau_i^{(\ell)})$ \iid{} replications of mean $\mu_i^{\ell-1}$, whereas $\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)}) : n_i(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_i^{\ell}$.
-
-Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_i(\tau_i^{(\ell)})$ lies in $
-\left\{\left\lfloor \frac{\alpha}{K}\left(\tau_i^{(\ell)}-\widehat{\tau}_i^{(\ell-1)}\right)\right\rfloor, \dots, \tau_i^{(\ell)}-\widehat{\tau}_i^{(\ell-1)} \right\}$.
-%
-Conditionally to $\cC_i^{(\ell-1)}$, one obtains furthermore using that $d_i^{(\ell-1)} \leq (\tau_i^{(\ell)} - \tau_i^{(\ell-1)})/2$ -- which follows from Assumption~\ref{ass:6:LongPeriods} -- that
-\begin{eqnarray*}
-    n_i(\tau_i^{(\ell)})& \in &  \left\{ \left\lfloor \frac{\alpha}{K}\left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)} - d_i^{(\ell-1)}\right)\right\rfloor, \dots, \left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)}\right) \right\} \\
-    n_i(\tau_i^{(\ell)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K}\left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)}\right)\right\rfloor, \dots, \left(\tau_i^{(\ell)}-\tau_i^{(\ell-1)}\right) \right\} = \cI_k.
-\end{eqnarray*}
-
-Introducing $\widehat{\mu}_{a,s}$ (resp.
-$\widehat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\widehat{\mu}_i^{(\ell-1)}$ (resp.
-$\widehat{\mu}_i^{(\ell)}$), such that $\widehat{\mu}_{a,s}$ and $\widehat{\mu}_{b,r}$ are independent, it follows that
-\[
-    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\widehat{\mu}_{a,s},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\widehat{\mu}_{b,\overline{r}},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
-\]
-where we have also used that $\tilde{\mu}^{\ell-1}_{i,n_i(\overline{t})} = \frac{n_i(\tau_i^{(\ell)})\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)})} + \overline{r}\tilde{\mu}^{\ell-1}_{i,n_i(\tau_i^{(\ell)}) : n_i(\overline{t})}}{n_i(\tau_i^{(\ell)}) + \overline{r}}$.
-
-Using Pinsker's inequality and introducing the gap $\Delta_i^{(\ell)} = \mu_i^{(\ell-1)} - {\mu}_i^{(\ell)}$, one can write
-\begin{align}
-    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
-    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,s} - \Delta_i^{(\ell)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
-    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
-\end{align}
-%
-
-Using Lemma~\ref{lem:6:Chernoff2arms} stated in Appendix~\ref{proof:6:Chernoff2arms} and a union bound, the first term in the right hand side is upper bounded by $\delta$ (as $\beta(r+s,\delta) \geq \beta(r,\delta) \geq \log(3s\sqrt{s}/\delta)$).
-For the second term, we use the observation
-\[\frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(\ell)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(\ell)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
-and finally get
-%
-\begin{equation}\label{eq:6:FromHere}
-    (b) \leq \delta + \bP\left(\exists s \in \cI_k : |\Delta_i^{(\ell)}| \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).
-\end{equation}
-%
-Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau_i^{(\ell)}-\tau_i^{(\ell-1)})/2\right\rfloor$.
-Using that the mappings $s \mapsto (s+\overline{r})/s\overline{r}$ and $s \mapsto \beta(s + \overline{r},\delta)$ are respectively decreasing and increasing in $s$, one has, for all $s\in \cI_k$,
-\begin{eqnarray*}
-    2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) & \leq &
-    2\frac{s_{\min}+\overline{r}}{s_{\min}\overline{r}}\beta\left(T, \delta\right) \leq \frac{4\beta(T,\delta)}{\left\lfloor \frac{\alpha}{K}d_i^{(\ell)}\right\rfloor},
-\end{eqnarray*}
-%
-where the last inequality follows from the fact that $\overline{r} \leq s_{\min}$ as $d_i^{(\ell)} \leq (\tau_i^{(\ell)} - \tau_i^{(\ell-1)})/2$ by Assumption~\ref{ass:6:LongPeriods}.
-Now the definition of $d_i^{(\ell)}$ readily implies that
-\[\left\lfloor \frac{\alpha}{K}d_i^{(\ell)}\right\rfloor > 4\beta(T,\delta) / \left(\Delta_i^{(\ell)}\right)^2,\]
-which yields
-\[\forall s \in \cI_k, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta_i^{(\ell)}\right)^2.\]
-Hence, the probability in the right-hand side of \eqref{eq:6:FromHere} is zero, which yields $(b) \leq \delta$.
-
+The proof of \GLRklUCB{} with Local Changes is given in Appendix~\ref{proof:6:mainRegretBound}, as it is quite similar to the proof for Global changes given above.
+
+% ----------------------------------------------------------------------------
 \subsection{Interpretation of the results}
@@
 % ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
 \section{Experimental Results}
 \label{sec:6:NumericalExperiments}
+% ----------------------------------------------------------------------------
 
@@
 
-We report results obtained on five different piece-wise stationary bandit problems, illustrated in Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}, \ref{fig:6:Problem_5}, \ref{fig:6:Problem_4} and \ref{fig:6:Problem_6}, and described in more details below.
+We report results obtained on five different piece-wise stationary bandit problems, illustrated in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} above, and Figures~\ref{fig:6:Problem_5}, \ref{fig:6:Problem_4} and \ref{fig:6:Problem_6} below, and described in more details below.
 Results are reported with regret tables and regret plots.
@@
     \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
-    \subcaption{\textbf{Problem $3$}: $K=6$ arms with $T=20000$, $\Upsilon=8$ changes occur on most arms at a time ($C=19$).
-        % Inspired from the Yahoo! example from \cite{CaoZhenKvetonXie18}.
-    }
+    \caption{\textbf{Problem $3$}: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
     \label{fig:6:Problem_5}
@@
     \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
-    \caption{\textbf{Problem 4}: $K=3$ arms with $T=5000$, $\Upsilon=4$ changes occur on all arms at a time (\ie, $C=12$).}
+    \caption{\textbf{Problem $4$}: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
+    % \caption{\textbf{Problem 4}: $K=3$ arms with $T=5000$, $\Upsilon=4$ changes occur on all arms at a time (\ie, $C=12$).}
     \label{fig:6:Problem_4}
@@
     \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
-    \caption{\textbf{Problem 5}: $K=5$, $T=100000$, $\Upsilon=81$ changes occurring on some arms at every time ($C=179$).
-    }
+    \caption{\textbf{Problem $5$}: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
+    % \caption{\textbf{Problem 5}: $K=5$, $T=100000$, $\Upsilon=81$ changes occurring on some arms at every time ($C=179$).}
     \label{fig:6:Problem_6}
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
     \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 1.}
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
     \caption{Histograms of the distributions of regret $R_T$ ($T=5000$) for problem 1.}
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
     \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem $2$.}
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K6_T20000_N500__9_algos_pb3/main____env1-1_8638317743626238996.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K6_T20000_N500__9_algos_pb3/main____env1-1_8638317743626238996.pdf}
     \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 20000$) for problem 3.}
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb4/main____env1-1_7471744535647872376.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb4/main____env1-1_7471744535647872376.pdf}
     \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 4.}
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K5_T100000_N500__9_algos_pb5/main____env1-1_950848333706043320.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K5_T100000_N500__9_algos_pb5/main____env1-1_950848333706043320.pdf}
     \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 100000$) for problem 5.}
@@
 We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as our proposal \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
-Finally, we would like to study in the future possible extension of our approach to the slowly varying model, as introduced in \cite{WeiSrivastava18}.
-
-\TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-player bandit model introduced in \cite{Wei2018} by using \MCTopM{} and \GLRklUCB{} with each other? }
-
-% Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm \cite{Besson2018ALT}, to efficiently tackle the piece-wise stationary distributed multi-player bandit model introduced in \cite{Wei2018}.
+Finally, we would like to study in the future possible extension of our approach to the slowly varying model, as introduced in \cite{WeiSrivastava18Abruptly}.
+
+\TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed} by using \MCTopM{} and \GLRklUCB{} with each other?
+}
+
+% Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm \cite{Besson2018ALT}, to efficiently tackle the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed}.
 

### 2019-03-13 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for now

@@
 %
-For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not evolve over time} may be a big limitation.
+For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
 Indeed, in cognitive radio new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate; whereas in online recommendation, the popularity of items is also subject to trends.
@@
 
-\TODOL{Rename GLR, GLRT and GLR Bernoulli to B-GLRT like what we did in our GRETSI 2019 paper \cite{Besson2019Gretsi}.}
+% \TODOL{Rename GLR, GLRT and GLR Bernoulli to B-GLRT like what we did in our GRETSI 2019 paper \cite{Besson2019Gretsi}.}
 
@@
 
-\paragraph{Illustrations of two problems of our benchmark.}\label{par:6:benchmark1}
-
-To illustrate the setup we presented above, we give here two problems of our benchmark, used for the numerical experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
+\paragraph{Illustrations of two problems.}\label{par:6:benchmark1}
+
+To illustrate the setup presented above, we give here two problems of our benchmark, used for the numerical experiments presented in Section~\ref{sec:6:NumericalExperiments}.
 
 % - Pb 1 changes are only on one arm at a time
-\textbf{Problem $\bm 1$.} There are $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
+\textbf{Problem $\bm 1$.}
+We consider $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
 The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1}.
@@
 Note that changes happen on only one arm (\ie, $C_T=\Upsilon_T=4$),
-and the optimal arm changes once at $t=\tau_2^{(1)}=2000$, with a large gap $\Delta=0.6$.
+and the optimal arm changes once at $\tau_2^{(1)}=2000$, with a large gap of $0.6$.
 
@@
     \centering
-    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
+    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
     \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).}
@@
 % - Pb 2 changes are on all arms at a time
-\textbf{Problem $\bm 2$.} (see Figure~\ref{fig:6:Problem_2}). This problem is close to Problem 1, with a minimum optimality gap of $0.1$.
-However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gap $\Delta=0.1$.
-The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$}).
+\textbf{Problem $\bm 2$.}
+This problem is close to Problem 1, with a minimum optimality gap of $0.1$,
+and shown in Figure~\ref{fig:6:Problem_2}.
+However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$.
+The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$ markers}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$ markers}).
 
@@
     \centering
-    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
+    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
     \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).}
@@
 % -----------------------------------------------------------
-\section{Related Work}
+\section{Review of Related Works}
 \label{sec:6:relatedWork}
@@
 The Discounted UCB (D-UCB) algorithm was first introduced in \cite{Kocsis06}, and it is an adaptation of the \UCB{} algorithm, with a discount factor $\gamma\in(0,1)$.
-It works by decreasing all the past rewards by a multiplicative factor, when receiving a new reward from an arm, for the recent rewards to weight more in the discounted empirical mean used for the computation of the \UCB{} indexes.
+It works by decreasing all the past rewards by a multiplicative factor, when receiving a new reward from an arm, so that the recent rewards weight more in the (discounted) empirical mean used for the computation of the \UCB{} indexes.
 The regret of D-UCB was proved to be upper-bounded by $\cO(\sqrt{\Upsilon_T T }\log(T))$ in \cite{Garivier11UCBDiscount}, with a tuning $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, dependent on $\Upsilon_T$.
@@
 
-Both D-UCB and SW-UCB builds on a stationary policy,
-but for example EXP3.S from \cite{Auer02NonStochastic}
+Both D-UCB and SW-UCB builds on a stationary policy,but for example EXP3.S from \cite{Auer02NonStochastic}
 builds on the EXP3 policy, which is designed for the adversarial case,
@@
 Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
-A  $\cO(V_T^{1/3} T^{2/3})$ regret bound is proved, which is weaker than existing results in our setting.
-Hence we do no include this algorithm in our experiments.
-%
+A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proved, which is weaker than existing results in our setting. Hence we do no include this algorithm in our experiments.
 
@@
 The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley Test and the \UCB{} policy, but no theoretical guarantee was given.
-The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a a suboptimal arm is detecting to become optimal.
+The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
 A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proved, even when $\Upsilon_T$ is known.
@@
 %
-\CUSUMUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but only for Bernoulli distributions,
+\CUSUMUCB{} achieves a better regret upper bound, of the order of $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but \emph{only} for Bernoulli distributions,
 and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
 %
-\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1).
+\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in their paper).
 
@@
 Note that the regret bound for these methods are still open-problems,
+and they are computationally more costly,
 therefore we choose to not include them in our experiments.
@@
 
-Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing much in terms of regret, in Appendix~\ref{sub:6:IdeasOptimizations}.
+Finally, on the practical side, while we can note that the proposed B-GLRT test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing much in terms of regret, in Appendix~\ref{sub:6:IdeasOptimizations}.
 
@@
 and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
-%
+
 This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
@@
 Generalized Likelihood Ratio tests date back to the seminal work of \cite{Wilks1938}, and were for instance studied for change-point detection by \cite{barnard1959control,siegmund1995using}.
-Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, can be use for our problem.
+Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, can be used for our problem.
 We propose instead to exploit the fact that bounded distributions are also dominated by Bernoulli distributions.
+
+\begin{definition}[Sub-Bernoulli distributions]
     We call a \emph{sub-Bernoulli distribution} any distribution $\nu$ that satisfies
-$\ln \bE_{X\sim\nu}\left[e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu=\bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) = \ln(1-\mu + \mu e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$.
-Lemma 1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
+    $\ln \bE_{X\sim\nu}\left[e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu=\bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) = \ln(1-\mu + \mu e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$, for any $\mu\in[0,1]$.
+\end{definition}
+
+We refer to Lemma~1 of \cite{KLUCBJournal} that establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
+Our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm know before hand the values of $a,b$).
 
@@
     \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
-    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)
-\end{align*}%
-\begin{align*}
+    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
+% \end{align*}%FIXME uncomment this if needed
+% \begin{align*}
     & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
@@
 where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
-As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
+As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$, if we denote $F(x)=\ln(1+\ln(x))$,
 %
@@
     &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
-    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6F(s+r)+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
+    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3F(s)+3F(r) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
 \end{align*}
@@
 \begin{lemma}\label{lem:6:ConcFirst}
-    $(X_i)_{i \in \N}$ and $(Y_i)_{i \in \N}$ two independent \iid{} processes with resp. means $\mu$ and $\mu'$ such that
-    \[\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)} \ \ \text{and} \ \ \bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)},\]
+    $(X_i)_{i \in \N^*}$ and $(Y_i)_{i \in \N^*}$ two independent \iid{} processes with respective means $\mu$ and $\mu'$ such that
+    $\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)}$,
     where $\phi_\mu(\lambda) = \bE_{X \sim \nu^{\mu}}[e^{\lambda X}]$ is the moment generating function of the distribution $\nu^\mu$, which is the unique distribution in an exponential family that has mean $\mu$.
-    %
+
     Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family.
-Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
+Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
     \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
@@
 \end{lemma}
-
+\begin{proof}
     This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
+\end{proof}
 
@@
 
-\TODOL{Include in Appendix a formal statement as well as a proof of Pinsker's inequality!}
-
-\TODOL{PROVE the lemma on detection delay !}
+\TODOL{Include in Appendix a formal statement and a proof of Pinsker's inequality?}
 
@@
     \begin{equation}
-        \bP_{\mu_0,\mu_1,\tau} (\widehat{\tau}_{\delta} \geq \tau + u) \leq \exp( -\frac{2\tau u}{\tau + u}\left(\max[ 0, \Delta - \sqrt{\frac{\tau + u}{2\tau u} \beta(\tau + u,\delta)} ]\right)^2 ).
+        \bP_{\mu_0,\mu_1,\tau} (\widehat{\tau}_{\delta} \geq \tau + u) \leq \exp\left( -\frac{2\tau u}{\tau + u}\left(\max\left[ 0, \Delta - \sqrt{\frac{\tau + u}{2\tau u} \beta(\tau + u,\delta)} \right]\right)^2 \right).
     \end{equation}
 \end{lemma}
-
-For a threshold $\beta$ chosen as in the Lemma~\ref{lem:6:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lem:6:Delay} is that if the break $\tau$ takes place after $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude.
-%
-In other words, if the B-GLRT test observes enough samples before a change-point,
-its delay $\widehat{\tau}_\delta$ is in the order of $O(\Delta^{-2} \ln(1/\delta))$, with high probability.
 
@@
 \end{proof}
+
+For a threshold $\beta$ chosen as in the Lemma~\ref{lem:6:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lem:6:Delay} is that if the break-point $\tau$ takes place after $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude (with high probability).
+%
+In other words, if the B-GLRT test observes enough samples before a change-point,
+its delay $\widehat{\tau}_\delta$ is in the order of $O(\Delta^{-2} \ln(1/\delta))$, with high probability.
 
@@
 %
-For the B-GLRT, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative, and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
+For the B-GLRT, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative (see Appendix~\ref{sec:6:exploringDifferentThresholdFunctions}), and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
 
@@
 %
-for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant enjoys the exact same false-alarm control.
+for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant enjoys the same false-alarm control.
 However, the detection delay may be slightly increased.
@@
 %
-With the exploration function $f(t)$ defined as $\ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
-the index is defined as
+With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
+the index is defined using the binary relative entropy as
 %
@@
 %
-One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative choice is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
+One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
 $(31)$-$(32)$.}:
@@
 
-\TODOL{I should link to Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} where I present two experiments which justifies this choice of the threshold function $\beta(n,\delta)$: it works fine and it's much simpler to implement numerically than the one we chose to analyze mathematically!}
+We present in Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} numerical simulations that compare different choices of threshold functions $\beta$.
+We compare the complicated value used in Lemma~\ref{lem:6:FalseAlarm}, using a numerical approximation of the function $\cT$,
+with the simpler value $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$,
+as well as a few other choices.
+The sum-up of these simulations is that it is enough to use a simpler and more explicit threshold to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
+
+% \TODOL{I should link to Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} where I present two experiments which justifies this choice of the threshold function $\beta(n,\delta)$: it works fine and it's much simpler to implement numerically than the one we chose to analyze mathematically!}
 
@@
         & \geq \sum_{u=s+1}^{t} \indic\left(A_u = i, A_u = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
-        & \geq \sum_{u=s+1}^{t} \indic\left(u \mod \left\lceil \frac{K}{\alpha} \right\rceil = i\right) \\
+        & \geq \sum_{u=s+1}^{t} \indic\left(i = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
         & = \bigl(t - (s+1) + 1\bigr) / \left\lceil \frac{K}{\alpha} \right\rceil
@@
 
-\TODOL{I should link to Appendix~\ref{sec:6:exploringDifferentForcedExplorationMechanisms} where I present two experiments which justifies this choice of the forced exploration mechanism: it works fine and it's the simplest to implement numerically and to analyze mathematically!}
+\paragraph{Other forced exploration schemes.}
+%
+We present in Appendix~\ref{sec:6:exploringDifferentForcedExplorationMechanisms} numerical simulations that compare three different options of forced exploration schemes.
+We compare the uniformly random exploration used by \CUSUMUCB, against the deterministic scheme used in Algorithm~\ref{algo:6:GLRklUCB}, and against a more complicated scheme based on \emph{tracking}.
+The sum-up of these simulations is that it is enough to use the deterministic scheme to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
+As it is the simplest and less numerically costly one, this is the one we chose for Algorithm~\ref{algo:6:GLRklUCB}.
+
+% \TODOL{I should link to Appendix~\ref{sec:6:exploringDifferentForcedExplorationMechanisms} where I present two experiments which justifies this choice of the forced exploration mechanism: it works fine and it's the simplest to implement numerically and to analyze mathematically!}
 
@@
 \subsection{Sketch of the unified regret analysis}
+\label{sub:6:proofSkeleton}
 
@@
 \begin{align*}
- &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
+ & \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t)) \\
+ & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
  & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
@@
     For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
-    \begin{equation}
-        R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
-    \end{equation}
+    \begin{align}
+        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T \\
+        & + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
+\nonumber
+    \end{align}
 \end{theorem}
+
+We highlight that \emph{this result is finite-time} and not asymptotic, even if it uses the notation $\bigO{\sqrt{\ln(T)}}$.
+This last term does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a constant times $\sqrt{\ln(T)}$ from a certain time $T_0$.
 
@@
 
-\subsection*{Proof for \GLRklUCB{} with Global Changes}
+\subsection{Proof for \GLRklUCB{} with Global Changes}
 \label{proof:6:mainRegretBoundGlobal}
@@
 
-Combining the regret decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEventGlobal} and the two upper bounds in \eqref{eq:6:TermAFinalGlobal} and \eqref{eq:6:TermBFinalGlobal},
+Combining the regret decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEventGlobal} and the two upper bounds of $(A)$ in \eqref{eq:6:TermAFinalGlobal} and of $(B)$ in \eqref{eq:6:TermBFinalGlobal},
 \[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
@@
 
-A few new notation are needed to state a regret bound for \GLRklUCB{} using local changes.
+Some new notations are needed to state a regret bound for \GLRklUCB{} using local changes.
 We let $\tau_i^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $i$}: $\tau_i^{(\ell)} = \inf \{ t > \tau_i^{(\ell - 1)} : \mu_i(t) \neq \mu_i(t+1)\}$,
@@
 
+Like for the first Theorem~\ref{thm:6:mainRegretBoundGlobal},
+we highlight that \emph{this result is finite-time}.
+%
 We can directly obtain different regret upper-bounds for different choices of the two parameters $\alpha$ and $\delta$.
@@
 
-The proof of \GLRklUCB{} with Local Changes is given in Appendix~\ref{proof:6:mainRegretBound}, as it is quite similar to the proof for Global changes given above.
+The proof of \GLRklUCB{} with Local Changes is only given in Appendix~\ref{proof:6:mainRegretBound}, as it is quite similar to the proof for Global changes given above.
+The main difficulty relies in defining the ``good event'' $\cE_T$, and proving that it happens with high probability (by showing that the complementary event $\cE_T^c$ is highly unlikely).
 
@@
 The regret bounds we obtain for the two variants of our proposal \GLRklUCB{},
-Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound},
+Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
 both show that there exists a tuning of $\alpha$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
@@
 % \TODOE{Je l'ai déjà dit, mais vraiment on va nous accuser d'avoir optimisé un paramètre chez nous et pas chez les autres.
-Est-ce qu'avec $\alpha_0=1$ on a des résultats comparables ou bien on est complètement à côté ???}
+Est-ce qu'avec $\alpha_0=1$ on a des résultats comparables ou bien on est complètement à côté ?}
 %
@@
 
-\TODOL{Justify MORE AND WAY MORE about all these results!}
-
 \paragraph{Regret plots.}
@@
 All the experiments used for this chapter were performed using SMPyBandits,
-and the page \texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html} explains how to reproduce them.
+and we refer to the following page for details on how to reproduce them,
+\texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html}

### 2019-03-14 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for now
Added experiments about choice of threshold functions, and choice of exploration mechanism.
Sorry Nabil I couldn't be there...
Almost done for this chapter 6. Maybe I'll move all these « exploration » numerical experiments in the main text, but use smaller tables (explore less values?)

@@
 %-----------------------------------------------------------------------------
-\paragraph{False alarm probability.}
+\paragraph{False alarm probability.}\label{par:6:falseAlarm}
 
@@
 %
-For the B-GLRT, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative (see Appendix~\ref{sec:6:exploringDifferentThresholdFunctions}), and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
+For the B-GLRT, numerical simulations suggest that the threshold \eqref{def:6:beta} is a bit conservative (see Appendix~\ref{sec:6:exploringDifferentThresholdFunctions}), and in practice we recommend to keep only the leading term and use $\beta(n,\delta) = \ln(3n^{3/2}/\delta) = \ln(3) + 3/2\log(n) - \log(\delta)$.
 
@@
 To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
-\begin{equation}
+\begin{equation}\label{eq:whenDoesGLRklUCBrestart}
     \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
@@
 $\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
-\[n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)\]
+$n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)$
 the number of selections of arm $i$, and
-\[\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
+$\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)$
 their empirical mean (if $n_i(t)\neq0$).
@@
 We compare the complicated value used in Lemma~\ref{lem:6:FalseAlarm}, using a numerical approximation of the function $\cT$,
-with the simpler value $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$,
-as well as a few other choices.
-The sum-up of these simulations is that it is enough to use a simpler and more explicit threshold to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
+with the simpler value $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$, as well as two other choices.
+The sum-up of these simulations is that it is enough to use a simpler and more explicit threshold, thus we recommend in practice to use $\beta_1(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
+%  to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
 
@@
 Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
-Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
+Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}),
+we observe that we have $\mu_{i^*_t} - \mu_{I_t} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
+and thus we can write a first decomposition of the regret,
 \begin{align*}
@@
     R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
-    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}. \nonumer
+    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}. \nonumber
 \end{align}
@@
 \[\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\},\]
-that all the changes up to the $\ell$-th have been detected, a union bound yields the following decomposition
+that all the changes up to the $\ell$-th have been detected, a union bound yields this decomposition
 %
@@
 Both choices of $\alpha$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
-the constant is set to $\alpha_0 = 0.05$
-(we show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\alpha\leq0.1$).
+the constant is set to $\alpha_0 = 0.05$.
+We show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\alpha\leq0.1$.
+We do not use a constant $\delta_0$ to change the confidence level to $\delta = \delta_0 \delta_T$, as we show in Appendix~\ref{sec:6:choosingDelta} that $\delta_0=1$ is uniformly better for different problems.
 %

### 2019-03-14 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
End of simulations
Almost okay for this proof of simplified expression of the B-GLR test

@@
 
-\subsection{Presentation of the test}
+\subsection{Presentation of the test}\label{sub:6:presentationOfGLRTest}
 
@@
 The Generalized Likelihood Ratio statistic for this test is defined by
-\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},\]
+\begin{equation}\label{eq:6:firstDefGLRT}
+    \GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},
+\end{equation}
 %
@@
 %
-Indeed, as proved in Appendix~\ref{app:6:GLR_with_kl}, we can show that in any one-dimensional exponential family, we have
-\[\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right],\]
+Indeed, as proved in Appendix~\ref{app:6:GLR_with_kl} (see also equation $(1.9)$ from \cite{LaiXing10}), we can show that in any one-dimensional exponential family, we have
+\[\GLR(n) = \exp\Bigl[ \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right], \Bigr] \]
 %
@@
     \begin{equation}\label{def:6:GLR}
-        \widehat{\tau}_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [1,t]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
+        \widehat{\tau}_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
     \end{equation}
@@
 We first observe that
-% A first key observation is
-\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
+\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\lambda \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
 Hence the probability of a false alarm occurring is upper bounded as
@@
     & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
-    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
+    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^*:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
 \end{align*}
@@
     \begin{tabular}{c|cccccc}
-    \textbf{Algorithms} \textbackslash \textbf{Problems} & Pb $1$ & Pb $2$ & Pb $3$ \\
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $1$ & Pb $2$ & Pb $3$ \\
         \hline
@@
     \begin{tabular}{c|cccccc}
-        \textbf{Algorithms} \textbackslash \textbf{Problems} & Pb $4$ ($T=5000$) & Pb $4$ ($T=10000$) & Pb $5$ \\
+        \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $4$ ($T=5000$) & Pb $4$ ($T=10000$) & Pb $5$ \\
         \hline

### 2019-03-14 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for now, but I still have to do one plot that Emilie wanted to see.

@@
 %
-Indeed, as proved in Appendix~\ref{app:6:GLR_with_kl} (see also equation $(1.9)$ from \cite{LaiXing10}), we can show that in any one-dimensional exponential family, we have
-\[\GLR(n) = \exp\Bigl[ \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right], \Bigr] \]
+Indeed, we show in Appendix~\ref{app:6:GLR_with_kl} that in any one-dimensional exponential family, we have
+\begin{equation}\label{eq:6:secondDefGLRT}
+    \GLR(n) = \exp\Bigl[ \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right], \Bigr].
+\end{equation}
 %
@@
 This motivates the following definition of the Bernoulli GLRT change point detector.
+Note that the $\exp$ appearing in the \eqref{eq:6:secondDefGLRT} is absorbed as a $\log$ in the threshold function $\beta$.
 
@@
 
-\TODOL{Include in Appendix a formal statement and a proof of Pinsker's inequality?}
-
 \begin{lemma}\label{lem:6:Delay}
@@
 \end{lemma}
-
+%
 \begin{proof}
-    \TODOL{Write it ?}
+    It is actually not used as such in the proofs of the regret upper bound given below, but a similar result is used and proved in the proofs.
 \end{proof}
@@
 \begin{equation}\label{def:6:GLRTricks}
-    \widetilde{\tau}_{\delta} = \inf \left\{ n \in \cN : \sup_{s \in \cS_n} \left[s \times d\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times d\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\},
+    \widetilde{\tau}_{\delta} = \inf \left\{ n \in \cN : \sup_{s \in \cS_n} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\},
 \end{equation}
@@
 
-%-----------------------------------------------------------------------------
-\paragraph{Details about the algorithm.}
+% %-----------------------------------------------------------------------------
+% \paragraph{Details about the algorithm.}
 
@@
     Lines~$3$-$4$ state that $A_u = u \mod \lceil \frac{K}{\alpha} \rceil$
-    if $u \mod \lceil \frac{K}{\alpha} \rceil \in \{1,\dots,K\}$
-    (see Algorithm~\ref{algo:6:GLRklUCB} for details).
-    Thus we have
+    if $u \mod \lceil \frac{K}{\alpha} \rceil \in \{1,\dots,K\}$,
+    % (see Algorithm~\ref{algo:6:GLRklUCB} for details).
+    thus we directly find
     \begin{align*}
@@
     \end{align*}
-    Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
+    % Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
 \end{proof}
 
-\paragraph{Other forced exploration schemes.}
+\paragraph{Other forced exploration schemes?}
 %
@@
 
-This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the finite-time regret bounds for the two variants, with their proofs, which both follow the skeleton of the unified analysis.
-% We highlight that our results are all non-asymptotic, contrarily to what is given for \MUCB{} \cite{CaoZhenKvetonXie18} and \CUSUMUCB{} \cite{LiuLeeShroff17}.
+This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the finite-time regret bounds for the two variants, with the proof of the first one (the second one is proved in Appendix~\ref{proof:6:mainRegretBound}), which both follows the skeleton of the unified analysis.
 
@@
     %
-    Then we assume that for all $k \in \{1,\dots,\Upsilon_T\}$,
+    Then \emph{we assume} that for all change $k \in \{1,\dots,\Upsilon_T\}$,
     $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
@@
 
-We first introduce some notation for the proof:
+We first introduce some notations for the proof:
 let $\widehat{\tau}^{(k)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
@@
 
-Some new notations are needed to state a regret bound for \GLRklUCB{} using local changes.
+Some new notations are needed to state a regret bound for \GLRklUCB{} using local changes,
+and to distinguish notations between local and global changes, we denote $\ell$ instead of $k$ for the indexes of change-points.
 We let $\tau_i^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $i$}: $\tau_i^{(\ell)} = \inf \{ t > \tau_i^{(\ell - 1)} : \mu_i(t) \neq \mu_i(t+1)\}$,
@@
 We also introduce the gap $\Delta_i^{(\ell)} = \overline{\mu}_i^{\ell} - \overline{\mu}_i^{\ell-1} > 0$.
-% ($\ell$ is used instead of $k$ to distinguish notations between local and global changes)
 
@@
     %
-    We assume that for all arm $i$ and all $\ell \in \{1,\dots,\NCi\}$, $\tau_i^{(\ell)} - \tau_i^{(\ell-1)} \geq 2\max (d_i^{(\ell)},d_i^{(\ell-1)})$.
+    Then \emph{we assume} that for all arm $i$ and all change-point of that arm $\ell \in \{1,\dots,\NCi\}$, $\tau_i^{(\ell)} - \tau_i^{(\ell-1)} \geq 2\max (d_i^{(\ell)},d_i^{(\ell-1)})$.
 \end{assumption}
@@
 % ----------------------------------------------------------------------------
-\subsection{Interpretation of the results}
+\subsection{Interpretation and comparison of the results}
 \label{sub:6:interpretationRegretBounds}

### 2019-03-14 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Should be OKAY for chapter 6... Je veux l'envoyer et rentrer. Je me sens si triste aujourd'hui.

@@
 
+% -----------------------------------------------------------------
+\paragraph{Examples of detection delays \emph{on one run} for two simple problems.}
+
+Before giving larger results that compare our approach against other algorithms, we consider the two problems $1$ and $2$ presented above in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} (for $T=5000$).
+\emph{For one (random) simulation}, we give below examples of the efficiency of the change-point detection algorithm used by \GLRklUCB{} as well as \CUSUMklUCB{} and \MklUCB,
+by giving the times at which they detected a change and reinitialized their memory of observations of one or all the arms.
+We also give the number of observations (since the last restart) of the arm on which the change was detected.
+As shown below in the summary of our experiments, the three change-point detection algorithm give similar performances in terms of regret, but they have different behaviors.
+The test in \MUCB{} usually detects less changes, while the test in \CUSUMUCB{} usually detects more changes and even unnecessary changes.
+%
+On the examples presented below as well as large scale examples, all the tests appear to have no false alarm, and always a small detection delay for changes that are ``easy enough'' to be efficiently detected.
+The four algorithms are presented in the order of decreasing performance (\ie, the first one is empirically the most efficient, that is, it obtains a smaller regret in average).
+
+\TODOL{Je ne suis pas sur que cette forme soit la meilleure pour présenter ces exemples, mais je n'aime vraiment pas la forme graphique qu'on avait suggéré ensemble avec Emilie, en fait on voit rien si on met plus d'un algo...}
+
+\textbf{Problem $\bm{1}$} has only local changes, and important changes, that denote here changes on the current optimal arm, happen only at $t=2000$ when arm $2$ becomes sub-optimal ($\mu_2$ goes from $0.9$ to $0.1$) and arm $0$ becomes optimal, and at $t=3000$ when arm $0$ stays optimal but sees its mean change from $0.3$ to $0.7$.
+Other changes concern arm $1$ at $t=1000$ and $t=4000$.
+
+\begin{itemize}
+    \item
+    \GLRklUCB{} with \emph{local restarts} restarted at $2014$ from a change detected on arm $2$, after $1970$ pulls of this arm, and at time $3032$ on arm $0$ (after $840$ pulls), and did not restart then.
+It successfully detects both important changes, and obtains a good performance in terms of regret, with $R_T = 63$ (on only one run).
+    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2014 for arm 2, after 1969 pulls of that arm (giving mean reward = 0.888).
+Last restart on that arm was at tau = 0
+    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 3032 for arm 0, after 771 pulls of that arm (giving mean reward = 0.336).
+Last restart on that arm was at tau = 0
+
+    \item
+    \GLRklUCB{} with \emph{global restarts} restarted at $2009$ from a change detected on arm $2$, after $1972$ pulls of this arm, and at time $3031$ on arm $0$ (after $827$ pulls), and did not restart then.
+It successfully detects both important changes, and obtains a good performance in terms of regret, with $R_T = 71$.
+    Surprisingly, we do not observe a significant difference with the \emph{local restarts} variant.
+    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2009 for arm 2, after 1972 pulls of that arm (giving mean reward = 0.896).
+Last restart on that arm was at tau = 0
+    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 3031 for arm 0, after 827 pulls of that arm (giving mean reward = 0.32).
+Last restart on that arm was at tau = 2009
+
+    \item
+    \CUSUMklUCB{} restarted at $2011$ from a change detected on arm $2$, after $1861$ pulls of this arm, at time $2512$ on arm $1$ (after $249$ pulls), at time $3027$ on arm $0$ (after $825$ pulls) and again at time $3505$ on arm $0$ (after $466$ pulls), at time $4383$ on arm $1$ (after $123$ pulls), and a last time at $t=4914$ on arm $0$ (after $1304$ pulls), and did not restart then.
+    It restarted too many times on arm $0$ after its last change at $t=3000$, but except this weakness, it successfully detected the two important changes, and obtains a final regret $R_T = 150$.
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2011 for arm 2, after 1861 pulls of that arm (giving mean reward = 0.895). Last restart on that arm was at tau = 0
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2512 for arm 1, after 249 pulls of that arm (giving mean reward = 0.265). Last restart on that arm was at tau = 0
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3027 for arm 0, after 825 pulls of that arm (giving mean reward = 0.33). Last restart on that arm was at tau = 0
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3505 for arm 0, after 466 pulls of that arm (giving mean reward = 0.717). Last restart on that arm was at tau = 3027
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 4383 for arm 1, after 123 pulls of that arm (giving mean reward = 0.268). Last restart on that arm was at tau = 2512
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 4914 for arm 0, after 1304 pulls of that arm (giving mean reward = 0.699). Last restart on that arm was at tau = 3505
+
+    \item
+    \MklUCB{} restarted at $2055$ from a change detected on arm $2$, after $1811$ pulls of this arm, and at time $3080$ on arm $0$ (after $868$ pulls), and did not restart then.
+    It successfully detects both important changes, and obtains a rather bad performance in terms of regret, with $R_T = 280$.
+    % For a player M-klUCB($w=150$, Global) a change was detected at time 2055 for arm 2, after 1811 pulls of that arm (giving mean reward = 0.0552). Last restart on that arm was at tau = 0
+    % For a player M-klUCB($w=150$, Global) a change was detected at time 1025 for arm 0, after 868 pulls of that arm (giving mean reward = 0.0806). Last restart on that arm was at tau = 2055
+\end{itemize}
+
+\textbf{Problem $\bm{2}$} has only global changes, and important changes happen only at $t=1000$ when arm $2$ sees its mean change from $0.9$ to $0.7$, then at $t=2000$ when it becomes sub-optimal ($\mu_2$ goes from $0.7$ to $0.5$) and arm $0$ becomes optimal, and at $t=3000$ and $t=4000$ when arm $0$ stays optimal but sees its mean change from $0.6$ to $0.7$ and then from $0.7$ to $0.8$.
+
+\begin{itemize}
+    \item
+    \GLRklUCB{} with \emph{local restarts} restarted at $1070$ from a change detected on arm $2$, after $1030$ pulls of this arm, again at time $2367$ on arm $0$ (after $1006$ pulls), and at time $4203$ on arm $0$ (after $1873$ pulls), and did not restart then.
+    It successfully detects the important change, and obtains a good performance in terms of regret, with a final regret $R_T = 115$.
+    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 1070 for arm 2, after 1030 pulls of that arm (giving mean reward = 0.89).
+Last restart on that arm was at tau = 0
+    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2367 for arm 2, after 1006 pulls of that arm (giving mean reward = 0.654).
+Last restart on that arm was at tau = 1070
+    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 4203 for arm 0, after 1873 pulls of that arm (giving mean reward = 0.666).
+Last restart on that arm was at tau = 0
+
+    \item
+    \GLRklUCB{} with \emph{global restarts} restarted at time $1111$ from a change detected on arm $2$, after $1060$ pulls of this arm, and at time $2705$ on arm $0$ (after $1038$ pulls), and did not restart then.
+    It successfully detects the important changes, but with a larger delay than the \emph{local restarts} variant, and also obtains a good performance in terms of regret, with a final regret $R_T = 125$.
+    Here again we do not observe a significant difference with the \emph{local restarts} variant.
+    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 1111 for arm 2, after 1060 pulls of that arm (giving mean reward = 0.888).
+Last restart on that arm was at tau = 0
+    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2705 for arm 2, after 1038 pulls of that arm (giving mean reward = 0.637).
+Last restart on that arm was at tau = 1111
+
+    \item
+    \CUSUMklUCB{} restarted at $1055$ from a change detected on arm $2$, after $963$ pulls of this arm, at time $2051$ again on arm $2$ (after $922$ pulls), at time $2305$ on arm $0$ (after $185$ pulls), at time $2591$ on arm $0$ (after $272$ pulls), at time $3387$ on arm $0$ (after $770$ pulls), at time $3677$ on arm $1$ (after $169$ pulls), and did not restart then.
+    It missed some changes but detected the two ones, and obtained a final regret $R_T=142$ on this run.
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 1055 for arm 2, after 963 pulls of that arm (giving mean reward = 0.916). Last restart on that arm was at tau = 0
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2051 for arm 2, after 922 pulls of that arm (giving mean reward = 0.682). Last restart on that arm was at tau = 1055
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2305 for arm 0, after 185 pulls of that arm (giving mean reward = 0.53). Last restart on that arm was at tau = 0
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2591 for arm 0, after 272 pulls of that arm (giving mean reward = 0.618). Last restart on that arm was at tau = 2305
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3387 for arm 0, after 770 pulls of that arm (giving mean reward = 0.653). Last restart on that arm was at tau = 2591
+    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3677 for arm 1, after 169 pulls of that arm (giving mean reward = 0.373). Last restart on that arm was at tau = 0
+
+    \item
+    \MklUCB{} did not detect any change, thus it obtains bad performance in terms of regret, with $R_T = 570$ (on this run).
+\end{itemize}
+
+In these two examples, we compare the two variants of our proposal with the two state-of-the-art policies \CUSUMklUCB{} and \MklUCB,
+that are all based on combining \klUCB{} with a change-point detection algorithm.
+The results given above should be taken carefully, they only have the purpose of being an illustration of the possible behaviors of these algorithms, as they are the results of \emph{only one (random) simulation}!
+But it is still interesting to compare the final regret in one run, as given above, with the mean regret for $1000$ independent runs, as given below in Table~\ref{table:6:totalResults1}.
+The values are not the same, but the ranking is: our proposal outperforms both \CUSUMklUCB{} and \MklUCB, and the test based on \CUSUM{} seems more efficient than the test based of \MklUCB{} in our problems at hand.
+
+% -----------------------------------------------------------------
 \paragraph{Illustrations of the rest of our benchmark.}\label{par:6:benchmark2}
@@
 
-We also include below figures showing the simulation results for the five problems.
+We also include in Appendix~\ref{app:6:moreFigures} some figures showing the simulation results for the five problems.
 The results in terms of mean regret $R_T$ are given above in Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2}, but it is also interesting to observe two plots for each experiments.
@@
 It shows that all algorithms have a rather small variance on their regret, except Thompson Sampling which has a large tail due to its large mean regret on this (easy) non-stationary problems.
-
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
-    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 1.}
-    \label{fig:6:meanRegretPb1}
-\end{figure}
-
-% -----------------------------------------------------------------
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
-    \caption{Histograms of the distributions of regret $R_T$ ($T=5000$) for problem 1.}
-    \label{fig:6:histogramRegretPb1}
-\end{figure}
-
-% -----------------------------------------------------------------
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
-    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem $2$.}
-    \label{fig:6:meanRegretPb2}
-\end{figure}
-
-% -----------------------------------------------------------------
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K6_T20000_N500__9_algos_pb3/main____env1-1_8638317743626238996.pdf}
-    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 20000$) for problem 3.}
-    \label{fig:6:meanRegretPb3}
-\end{figure}
-
-% -----------------------------------------------------------------
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb4/main____env1-1_7471744535647872376.pdf}
-    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 4.}
-    \label{fig:6:meanRegretPb4}
-\end{figure}
-
-% -----------------------------------------------------------------
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K5_T100000_N500__9_algos_pb5/main____env1-1_950848333706043320.pdf}
-    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 100000$) for problem 5.}
-    \label{fig:6:meanRegretPb5}
-\end{figure}
 

### 2019-03-17 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Added a Lemma for Pinsker's inequality, as well as its proof.

@@
 For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to the stopping time \eqref{def:6:GLR} with the choice $d(x,y) = 2(x-y)^2$, when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
-Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
+Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$ (see Lemma~\ref{lem:6:Pinsker}), hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
 %
@@
 Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from $\mu_0$ to $\mu_1$ occurs at time $\tau$.
-We already observed that from Pinsker's inequality, the B-GLRT stops earlier than a Gaussian GLR.
+We already observed that from Pinsker's inequality (Lemma~\ref{lem:6:Pinsker}), the B-GLRT stops earlier than a Gaussian GLR.
 Hence, one can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the B-GLRT.
@@
 
-Using Pinsker's inequality and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = |\Delta_i^k|$), one can write
+Using Pinsker's inequality from Lemma~\ref{lem:6:Pinsker} and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = |\Delta_i^k|$), one can write
 %

### 2019-03-25 3-Appendices/2-Appendix/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.git/Rebuttal_COLT_Paper_177.md
Je suis bourrin mais je git TOUT pour éviter une catastrophe...

We thank our three reviewers for their careful readings and comments, and we address below specific concerns raised by reviewers 2 and 3.

The typo noted by reviewer 1 in the Algorithm was corrected.

# Response to Reviewer 2

- "In the introduction, the paper claims to answer [...]".

We changed this sentence in the introduction, to reflect what is announced in the abstract and what is actually proved in the paper, thanks for the remark.

- "An (advanced) extension [...] *Corralling a Band of Bandit Algorithms* by Agarwal et al. in COLT 2017".

That is indeed an advanced extension, and we will try to look into it in the future.

- "Lemma 2 is not correct: It assumes that the regret will increase when the last epoch is extended from time $T$ to time $T_{L_T}$. Strictly speaking, this is not even defined, because there are no rewards beyond time $T$".

Our definition of the reward streams defines $(Y_{k,t})_{t\in\mathbb{N}}$, and not just $Y_{k,0},\dots,Y_{k,T}$.
Furthermore, the upper bound in Lemma 2 refers to regret with horizon $T_i - T_{i-1}$ only for an algorithm which uses this horizon, so according to the definition of $r(t)$ (top of page 2), the reward $r(t)$ is actually well defined here.

- "But even being more liberal it is incorrect, because the regret is not monotonically increasing in T. If the result is restated in terms of a bound on the regret, however, then it can be salvaged [...]".

Thanks for mentioning this, and indeed the regret is monotonically increasing for stochastic models but could not be for generic adversarial models.
However, it is true for oblivious adversaries, which is our case of interest as we defined the reward streams $(Y_{k,t})_{t\in\mathbb{N}}$ (if they are fixed from the beginning they cannot be adaptive to the player's actions).
We will be clearer on this point in the final version, by announcing when needed that we mean "oblivious adversary" when we think of "possibly non-stochastic models".

- "The paper's interpretation of Theorem 5 is that geometric doubling does not preserve logarithmic regret, but I would view it as saying that we only pay one extra $\log(T)$ factor, which is not so bad. So I would not interpret this as a negative result".

Thanks, that is another, valid, point of view. We wanted to highlight the fact that an order-optimal algorithm (for problem-dependent bounds, ie, logarithmic regret) becomes sub-optimal when the geometric doubling trick is used.
But we added a remark stating that we only pay one extra $\log(T)$ factor (and depending on the value of $\delta$, it can be not so bad or very bad), to be clearer on this point.

- "Theorems 5 and 9 [...]".

Thank you for your suggestion, we did modify both theorems according to what you suggest.

- "In Theorems 4 and 7, referring to the constant factor overhead as a *loss* is confusing in the context of the bandit setting".

We agree that indeed loss is misleading in the context of online learning, we changed it to multiplicative *overhead*.

- "Definition 6 is ungrammatical".

Definitions 3 and 6 were slightly modified to be grammatically correct, thank you for noticing.

# Response to Reviewer 3

- "While this may be a useful reference point for future work, the results are rather unsurprising".

The necessity of using an exponential doubling trick to preserve problem-dependent bounds (ie, logarithmic) was maybe a known fact (see for instance [*UCB revisited*, Auer et al, 2010]) but, to the best of our knowledge, it was not proved before (Theorem 5).
Additionally, our lower bound in Theorem 9 showing that an extra $\log(T)$ term is paid when using an exponential doubling trick for problem-independent bounds (ie, in $\sqrt{T}$) is also new.

- "I'm especially not convinced with the usefulness of the doubling technique on lower bound".

This part of the paper should actually be understood the other way around: we do not demonstrate the usefulness of the doubling technique on lower bound (that is, to prove that it can be used to preserve a lower bound), but we do the opposite. We use a lower bound to prove that for certain doubling tricks *cannot* be used to preserve certain upper bounds.
For example, Theorem 5 shows that a geometric doubling trick is not increasing fast enough to preserve logarithmic regret bounds. Failing to prove an upper bound is not enough to support this claim, and to the best of our knowledge, proving an actual lower bound like Theorem 5 is the only possible approach.

- "Why are these results limited to the multi-armed bandits problem? All statements seem to be agnostic of the partial information setting".

Thanks for noting that. Indeed, we believe that our upper bound (Ths 4 and 7) can be extended to other feedback models (eg, full information) and we will mention it in the final version. It is a bit less clear for the lower bounds, as Lemma 2 requires a *stochastic* bandit feedback.
It might be true as well for online learning against a stochastic adversary, but the bandit case is more interesting anyways.

- "Doubling techniques often get tricky for problem-dependent bounds, because the regret scales not only in $T$ but also in other functions of the input (reward sequence). [...]".

As this paper is motivated by doubling tricks on stochastic bandit algorithms, we did not consider this type of "problem-dependent" bounds (eg, second-order bounds or bounds that depend on the loss of the best action). But we will look into this work, thanks for mentioning it.

### 2019-03-25 3-Appendices/2-Appendix/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.git/Rebuttal_COLT_Paper_177.tex
Je suis bourrin mais je git TOUT pour éviter une catastrophe...

\documentclass[anon,12pt]{colt2018} % Anonymized submission
% \documentclass[12pt]{colt2018} % Include author names
% http://www.learningtheory.org/colt2018/call.html#instructions

\title{Rebuttal for COLT 2018, paper 177}

\usepackage{macrosText}

\coltauthor{\Name{Lilian Besson}\textsuperscript{$\dagger$} \Email{{Lilian}{.}{Besson}{@}{CentraleSupelec}{.}{fr}} \\
    \addr CentraleSup\'elec (campus of Rennes), IETR, SCEE Team,\\
    Avenue de la Boulaie -- CS $47601$, F-$35576$ Cesson-S\'evign\'e, France
    \AND
    \Name{Emilie Kaufmann} \Email{{emilie}{.}{kaufmann}{@}{univ}{-}{lille1}{.}{fr}} \\
    \addr CNRS \& Universit\'e de Lille, Inria SequeL team\\
    UMR $9189$ -- CRIStAL,  F-$59000$ Lille, France
}

\begin{document}

\maketitle
% -----------------------------------------------------------------

We thank our three reviewers for their careful readings and comments,
and address below specific concerns raised by reviewer 2 and 3.
The typo noted by reviewer 1 in the Algorithm was corrected.

% -----------------------------------------------------------------
\section{Response to Reviewer 2}

\begin{itemize}
    \item
    ``In the introduction, the paper claims to answer [\ldots]''.

    We changed this sentence in the introduction, to reflect what is announced in the abstract and what is actually proved in the paper, thanks for the remark.

    \item
    ``An (advanced) extension [\ldots] \emph{Corralling a Band of Bandit Algorithms} by Agarwal et al. in COLT 2017''.

    That is indeed an advanced extension, and we will try to look into it in the future.

    \item
    ``Lemma 2 is not correct: It assumes that the regret will increase when the last epoch is extended from time $T$ to time $T_{L_T}$. Strictly speaking, this is not even defined, because there are no rewards beyond time $T$''.

    Our definition of the reward streams defines $(Y_{k,t})_{t\in\mathbb{N}}$, and not just $Y_{k,0},\dots,Y_{k,T}$.
    Furthermore, the upper bound in Lemma~2 refers to regret with horizon $T_i - T_{i-1}$ only for an algorithm which uses this horizon, so according to the definition of $r(t)$ (top of page 2), the reward $r(t)$ is actually well defined here.

    \item
    ``But even being more liberal it is incorrect, because the regret is not monotonically increasing in T. If the result is restated in terms of a bound on the regret, however, then it can be salvaged [\ldots]''.

    Thanks for mentioning this, and indeed the regret is monotonically increasing for stochastic models but could not be for generic adversarial models.
    However, it is true for oblivious adversaries, which is our case of interest as we defined the reward streams $(Y_{k,t})_{t\in\mathbb{N}}$ (if they are fixed from the beginning they cannot be adaptive to the player's actions).
    We will be clearer on this point in the final version, by announcing when needed that we mean "oblivious adversary" when we think of "possibly non-stochastic models".
    % The suggestion was a good start, and we modified accordingly the Lemma 2 and its proof, as well as the other upper bound results.

    \item
    ``The paper's interpretation of Theorem 5 is that geometric doubling does not preserve logarithmic regret, but I would view it as saying that we only pay one extra $\log(T)$ factor, which is not so bad. So I would not interpret this as a negative result''.

    Thanks, that is another, valid, point of view.
    We wanted to highlight the fact that an order-optimal algorithm (for problem-dependent bounds, \ie, logarithmic regret) becomes sub-optimal when the geometric doubling trick is used.
    But we added a remark stating that we only pay one extra $\log(T)$ factor (and depending on the value of $\delta$, it can be not so bad or very bad), to be clearer on this point.

    \item
    ``Theorems 5 and 9 [\ldots]''.

    Thank you for your suggestion, we did modify both theorems according to what you suggest.

    \item
    ``In Theorems 4 and 7, referring to the constant factor overhead as a \emph{loss} is confusing in the context of the bandit setting''.

    We agree that indeed loss is misleading in the context of online learning, we changed it to multiplicative *overhead*.

    \item
    ``Definition 6 is ungrammatical''.

    Definitions 3 and 6 were slightly modified to be grammatically correct, thank you for noticing.

\end{itemize}

\hr{}
% -----------------------------------------------------------------

% -----------------------------------------------------------------
\section{Response to Reviewer 3}

\begin{itemize}
    \item
    ``While this may be a useful reference point for future work, the results are rather unsurprising''.

    The necessity of using an exponential doubling trick to preserve problem-dependent bounds (\ie, logarithmic) was maybe a known fact (see for instance [\emph{UCB revisited}, Auer02 et al, 2010]) but, to the best of our knowledge, it was not proved before (Theorem~5).
    Additionally, our lower bound in Theorem~9 showing that an extra $\log(T)$ term is paid when using an exponential doubling trick for problem-independent bounds (\ie, in $\sqrt{T}$) is also new.

    \item
    ``I'm especially not convinced with the usefulness of the doubling technique on lower bound''.

    This part of the paper should actually be understood the other way around: we do not demonstrate the usefulness of the doubling technique on lower bound (that is, to prove that it can be used to preserve a lower bound), but we do the opposite.
    We use a lower bound to prove that for  certain doubling tricks \emph{cannot} be used to preserve certain upper bounds.
    For example, Theorem~5 shows that a geometric doubling trick is not increasing fast enough to preserve logarithmic regret bounds.
    Failing to prove an upper bound is not enough to support this claim, and to the best of our knowledge, proving an actual lower bound like Theorem~5 is the only possible approach.

    \item
    ``Why are these results limited to the multi-armed bandits problem? All statements seem to be agnostic of the partial information setting''.

    % All the results concerning regret upper bounds are indeed agnostic of the partial information hypothesis, and we could rewrite the relevant parts of the paper to take this generalization into account.
    % However, it might not be as easy for the lower bounds, as Lemma~2 is not true for any bandit model but is true with a stochastic hypothesis.

    % Under such short time constraints, the authors did not have time to check very carefully this point, but we do believe that the paper could also be extended to include the full information setting.

    Thanks for noting that.
    Indeed, we believe that our upper bound (Th 4 and 7) can be extended to other feedback models (e.g., full information) and we will mention it in the final version.
    % FIXME
    It is a bit less clear for the lower bounds, as Lemma 2 requires a *stochastic* bandit feedback.
    It might be true as well for online learning against a stochastic adversary, but the bandit case is more interesting anyways.

    \item
    ``Doubling techniques often get tricky for problem-dependent bounds, because the regret scales not only in $T$ but also in other functions of the input (reward sequence). [\ldots]''.

    As this paper is motivated by doubling tricks on stochastic bandit algorithms, we did not consider this type of "problem-dependent" bounds (e.g., second-order bounds or bounds that depend on the loss of the best action).
    But we will look into this work, thanks for mentioning it.

    % Regarding this reference, Neu said ``While one could use a standard doubling trick to overcome this difficulty, we choose to take a different path to circumvent this issue'', so we do not believe that the carefulness demonstrated in his proof can be a clue on the difficulty to use doubling techniques for problem-dependent bounds.

\end{itemize}

\end{document}

### 2019-03-25 2-Chapters/3-Chapter/SMPyBandits_paper.git/longpaper.tex
Je suis bourrin mais je git TOUT pour éviter une catastrophe...

\hypertarget{summary}{%
\section{Summary}\label{summary}}

This article presents \href{https://perso.crans.org/besson/}{my}
numerical environment \emph{SMPyBandits}, written in
\href{https://www.python.org/}{Python (2 or 3)} {[}@python{]}, for
numerical simulations on \emph{single}-player and \emph{multi}-players
\href{https://en.wikipedia.org/wiki/Multi-armed_bandit}{Multi-Armed
Bandits (MAB)} algorithms {[}@Bubeck12{]}.

\emph{SMPyBandits} is the most complete open-source implementation of
state-of-the-art algorithms tackling various kinds of sequential
learning problems referred to as Multi-Armed Bandits. It aims at being
extensive, simple to use and maintain, with a clean and perfectly
documented codebase. But most of all it allows fast prototyping of
simulations and experiments, with an easy configuration system and
command-line options to customize experiments while starting them (see
below for an example).

\emph{SMPyBandits} does not aim at being blazing fast or perfectly
memory efficient, and comes with a pure Python implementation with no
dependency except standard open-source Python packages. Even if some
critical parts are also available as a \texttt{C} Python extension, and
even by using Numba {[}@numba{]} whenever it is possible, if simulation
speed really matters, one should rather refer to less exhaustive but
faster implementations, like for example {[}@TorLibbandit{]} in
\texttt{C++} or {[}@VishMABjl{]} in Julia.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{presentation}{%
\subsection{Presentation}\label{presentation}}

\hypertarget{single-player-mab}{%
\subsubsection{Single-Player MAB}\label{single-player-mab}}

Multi-Armed Bandit (MAB) problems are well-studied sequential decision
making problems in which an agent repeatedly chooses an action (the
``arm'' of a one-armed bandit) in order to maximize some total reward
{[}@Robbins52,LaiRobbins85{]}. Initial motivation for their study came
from the modeling of clinical trials, as early as 1933 with the seminal
work of Thompson {[}@Thompson33{]}. In this example, arms correspond to
different treatments with unknown, random effect. Since then, MAB models
have been proved useful for many more applications, that range from
cognitive radio {[}@Jouini09{]} to online content optimization (news
article recommendation {[}@Li10{]}, online advertising
{[}@LiChapelle11{]} or A/B Testing {[}@Kaufmann14;Jamieson17{]}), or
portfolio optimization {[}@Sani12{]}.

This Python package is the most complete open-source implementation of
single-player (classical) bandit algorithms
(\href{https://smpybandits.github.io/docs/Policies.html}{over 65!}). We
use a well-designed hierarchical structure and
\href{https://smpybandits.github.io/uml_diagrams/README.html}{class
inheritance scheme} to minimize redundancy in the codebase, and for
instance the code specific to the UCB algorithm
{[}@LaiRobbins85;@Auer02{]} is as short as this (and fully documented),
by inheriting from the
\href{https://smpybandits.github.io/docs/Policies.IndexPolicy.html}{\texttt{IndexPolicy}}
class:

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{from}\NormalTok{ numpy }\ImportTok{import}\NormalTok{ sqrt, log}
\ImportTok{from}\NormalTok{ .IndexPolicy }\ImportTok{import}\NormalTok{ IndexPolicy}

\KeywordTok{class}\NormalTok{ UCB(IndexPolicy):}
  \CommentTok{""" The UCB policy for bounded bandits.}
\CommentTok{  Reference: [Lai & Robbins, 1985]. """}

  \KeywordTok{def}\NormalTok{ computeIndex(}\VariableTok{self}\NormalTok{, arm):}
    \CommentTok{r""" Compute the current index, at time t and}
\CommentTok{    after :math:`N_k(t)` pulls of arm k:}

\CommentTok{    .. math::}

\CommentTok{        I_k(t) = \textbackslash{}frac\{X_k(t)\}\{N_k(t)\}}
\CommentTok{        + \textbackslash{}sqrt\{\textbackslash{}frac\{2 \textbackslash{}log(t)\}\{N_k(t)\}\}.}
\CommentTok{    """}
    \ControlFlowTok{if} \VariableTok{self}\NormalTok{.pulls[arm] }\OperatorTok{<} \DecValTok{1}\NormalTok{:  }\CommentTok{# forced exploration}
      \ControlFlowTok{return} \BuiltInTok{float}\NormalTok{(}\StringTok{'+inf'}\NormalTok{)   }\CommentTok{# in the first steps}
    \ControlFlowTok{else}\NormalTok{:                    }\CommentTok{# or compute UCB index}
\NormalTok{      estimated_mean }\OperatorTok{=}\NormalTok{ (}\VariableTok{self}\NormalTok{.rewards[arm] }\OperatorTok{/} \VariableTok{self}\NormalTok{.pulls[arm])}
\NormalTok{      exploration_bias }\OperatorTok{=}\NormalTok{ sqrt((}\DecValTok{2} \OperatorTok{*}\NormalTok{ log(}\VariableTok{self}\NormalTok{.t)) }\OperatorTok{/} \VariableTok{self}\NormalTok{.pulls[arm])}
      \ControlFlowTok{return}\NormalTok{ estimated_mean }\OperatorTok{+}\NormalTok{ exploration_bias}
\end{Highlighting}
\end{Shaded}

\hypertarget{multi-players-mab}{%
\subsubsection{Multi-Players MAB}\label{multi-players-mab}}

For Cognitive Radio applications, a well-studied extension is to
consider \(M\geq2\) players, interacting on the \emph{same} \(K\) arms.
Whenever two or more players select the same arm at the same time, they
all suffer from a collision. Different collision models has been
proposed, and the simplest one consist in giving a \(0\) reward to each
colliding players. Without any centralized supervision or coordination
between players, they must learn to access the \(M\) best resources
(\emph{i.e.}, arms with highest means) without collisions.

This package implements
\href{https://smpybandits.github.io/docs/Environment.CollisionModels.html}{all
the collision models} found in the literature, as well as all the
algorithms from the last 10 years or so (including
\href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoRand.html}{\texttt{rhoRand}}
from 2009,
\href{https://smpybandits.github.io/docs/Policies.MEGA.html}{\texttt{MEGA}}
from 2015,
\href{https://smpybandits.github.io/docs/Policies.MusicalChair.html}{\texttt{MusicalChair}}
from 2016, and our state-of-the-art algorithms
\href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.RandTopM.html}{\texttt{RandTopM}}
and
\href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.MCTopM.html}{\texttt{MCTopM}})
from {[}@BessonALT2018{]}.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{purpose}{%
\subsection{Purpose}\label{purpose}}

The main goal of this package is to implement
\href{https://smpybandits.github.io/API.html}{with the same API} most of
the existing single- and multi-player multi-armed bandit algorithms.
Each algorithm comes with a clean documentation page, containing a
reference to the research article(s) that introduced it, and with
remarks on its numerical efficiency.

It is neither the first nor the only open-source implementation of
multi-armed bandits algorithms, although one can notice the absence of
any well-maintained reference implementation. I built \emph{SMPyBandits}
from a framework called \emph{pymaBandits} {[}@pymaBandits{]}, which
implemented a few algorithms and three kinds of arms, in both Python and
MATLAB. The goal was twofolds, first to implement as many algorithms as
possible to have a complete implementation of the current state of
research in MAB, and second to implement multi-players simulations with
different models.

Since November \(2016\), I follow actively the latest publications
related to Multi-Armed Bandits (MAB) research, and usually I implement
quickly any new algorithms. For instance,
\href{https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html}{Exp3++},
\href{https://smpybandits.github.io/docs/Policies.CORRAL.html}{CORRAL}
and
\href{https://smpybandits.github.io/docs/Policies.SparseUCB.html}{SparseUCB}
were each introduced by articles
(\href{https://arxiv.org/pdf/1702.06103}{for Exp3++},
\href{https://arxiv.org/abs/1612.06246v2}{for CORRAL},
\href{https://arxiv.org/abs/1706.01383}{for SparseUCB}) presented at
COLT in July 2017,
\href{https://smpybandits.github.io/docs/Policies.LearnExp.html}{LearnExp}
comes from a \href{https://arxiv.org/abs/1702.04825}{NIPS 2017 paper},
and
\href{https://smpybandits.github.io/docs/Policies.klUCBPlusPlus.html}{kl-UCB++}
from an \href{https://hal.inria.fr/hal-01475078}{ALT 2017 paper}.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{features}{%
\subsection{Features}\label{features}}

With this numerical framework, simulations can run on a single CPU or a
multi-core machine using joblib {[}@joblib{]}, and summary plots are
automatically saved as high-quality PNG, PDF and EPS (ready for being
used in research article), using matplotlib {[}@matplotlib{]} and
seaborn {[}@seaborn{]}. Making new simulations is very easy, one only
needs to write a configuration script and no knowledge of the internal
code architecture.

\hypertarget{examples-of-configuration-for-some-simulations}{%
\subsubsection{Examples of configuration for some
simulations}\label{examples-of-configuration-for-some-simulations}}

A small script
\href{https://smpybandits.github.io/docs/configuration.html}{\texttt{configuration.py}}
is used to import the
\href{https://smpybandits.github.io/docs/Arms.html}{arm classes}, the
\href{https://smpybandits.github.io/docs/Policies.html}{policy classes}
and define the problems and the experiments. For instance, we can
compare the standard anytime
\href{https://smpybandits.github.io/docs/Policies.klUCB.html}{\texttt{klUCB}}
algorithm against the non-anytime variant
\href{https://smpybandits.github.io/docs/Policies.klUCBPlusPlus.html}{\texttt{klUCBPlusPlus}}
algorithm, as well as
\href{https://smpybandits.github.io/docs/Policies.UCBalpha.html}{\texttt{UCB}}
(with \(\alpha=1\)) and
\href{https://smpybandits.github.io/docs/Policies.Thompson.html}{\texttt{Thompson}}
(with
\href{https://smpybandits.github.io/docs/Policies.Posterior.Beta.html}{Beta
posterior}). See below in Figure \ref{fig:plot1} for the result showing
the average regret for these \(4\) algorithms.

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{from}\NormalTok{ Arms }\ImportTok{import} \OperatorTok{*;}  \ImportTok{from}\NormalTok{ Policies }\ImportTok{import} \OperatorTok{*}

\NormalTok{configuration }\OperatorTok{=}\NormalTok{ \{}
  \StringTok{"horizon"}\NormalTok{: }\DecValTok{10000}\NormalTok{,     }\CommentTok{# Finite horizon of the simulation}
  \StringTok{"repetitions"}\NormalTok{: }\DecValTok{1000}\NormalTok{,  }\CommentTok{# Number of repetitions}
  \StringTok{"n_jobs"}\NormalTok{: }\DecValTok{-1}\NormalTok{,         }\CommentTok{# Max number of cores for parallelization}
  \CommentTok{# Environment configuration, you can set up more than one.}
  \StringTok{"environment"}\NormalTok{: [ \{}
      \StringTok{"arm_type"}\NormalTok{: Bernoulli,}
      \StringTok{"params"}\NormalTok{: [}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.3}\NormalTok{, }\FloatTok{0.4}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.6}\NormalTok{, }\FloatTok{0.7}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.9}\NormalTok{]}
\NormalTok{  \}],}
  \CommentTok{# Policies that should be simulated, and their parameters.}
  \StringTok{"policies"}\NormalTok{: [}
\NormalTok{      \{}\StringTok{"archtype"}\NormalTok{: klUCB, }\StringTok{"params"}\NormalTok{: \{\} \},}
\NormalTok{      \{}\StringTok{"archtype"}\NormalTok{: klUCBPlusPlus,}
       \StringTok{"params"}\NormalTok{: \{ }\StringTok{"horizon"}\NormalTok{: }\DecValTok{10000}\NormalTok{ \} \},}
\NormalTok{      \{}\StringTok{"archtype"}\NormalTok{: UCBalpha,}
       \StringTok{"params"}\NormalTok{: \{ }\StringTok{"alpha"}\NormalTok{: }\DecValTok{1}\NormalTok{ \} \},}
\NormalTok{      \{}\StringTok{"archtype"}\NormalTok{: Thompson, }\StringTok{"params"}\NormalTok{: \{\} \},}
\NormalTok{  ]\}}
\end{Highlighting}
\end{Shaded}

For a second example, this snippet is a minimal example \footnote{See
  the file
  \href{https://smpybandits.github.io/docs/configuration_multiplayers.html}{\texttt{configuration\_multiplayers.py}}
  in the code for more details.} of configuration for multiplayer
simulations, comparing different multi-player algorithms used with the
\href{https://smpybandits.github.io/docs/Policies.klUCB.html}{\texttt{klUCB}}
index policy. See below in Figure \ref{fig:plot2} for an illustration.

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{from}\NormalTok{ Arms }\ImportTok{import} \OperatorTok{*;}  \ImportTok{from}\NormalTok{ Policies }\ImportTok{import} \OperatorTok{*}
\ImportTok{from}\NormalTok{ PoliciesMultiPlayers }\ImportTok{import} \OperatorTok{*}
\NormalTok{nbPlayers }\OperatorTok{=} \DecValTok{3}

\NormalTok{configuration }\OperatorTok{=}\NormalTok{ \{}
  \StringTok{"horizon"}\NormalTok{: }\DecValTok{10000}\NormalTok{,    }\CommentTok{# Finite horizon of the simulation}
  \StringTok{"repetitions"}\NormalTok{: }\DecValTok{100}\NormalTok{,  }\CommentTok{# Number of repetitions}
  \StringTok{"n_jobs"}\NormalTok{: }\DecValTok{-1}\NormalTok{,        }\CommentTok{# Max number of cores for parallelization}
  \CommentTok{# Environment configuration, you can set up more than one.}
  \StringTok{"environment"}\NormalTok{: [ \{}
      \StringTok{"arm_type"}\NormalTok{: Bernoulli,}
      \StringTok{"params"}\NormalTok{: [}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.3}\NormalTok{, }\FloatTok{0.4}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.6}\NormalTok{, }\FloatTok{0.7}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.9}\NormalTok{]}
\NormalTok{  \} ],}
  \CommentTok{# Policies that should be simulated, and their parameters.}
  \StringTok{"successive_players"}\NormalTok{: [}
\NormalTok{      CentralizedMultiplePlay(nbPlayers, nbArms, klUCB).children,}
\NormalTok{      RandTopM(nbPlayers, nbArms, klUCB).children,}
\NormalTok{      MCTopM(nbPlayers, nbArms, klUCB).children,}
\NormalTok{      Selfish(nbPlayers, nbArms, klUCB).children,}
\NormalTok{      rhoRand(nbPlayers, nbArms, klUCB).children,}
\NormalTok{  ] \}}
\end{Highlighting}
\end{Shaded}

\hypertarget{documentation}{%
\subsubsection{Documentation}\label{documentation}}

A complete sphinx {[}@sphinx{]} documentation for each algorithms and
every piece of code, included the constants in the different
configuration files, is available here:
\href{https://smpybandits.github.io/}{\texttt{https://smpybandits.github.io}}.

\hypertarget{other-noticeable-features}{%
\subsubsection{Other noticeable
features}\label{other-noticeable-features}}

\hypertarget{single-player-policies}{%
\paragraph{\texorpdfstring{\href{https://smpybandits.github.io/docs/Policies.html}{Single-player
Policies}}{Single-player Policies}}\label{single-player-policies}}

\begin{itemize}
\tightlist
\item
  More than 65 algorithms, including all known variants of the
  \href{https://smpybandits.github.io/docs/Policies.UCB.html}{\texttt{UCB}},
  \href{https://smpybandits.github.io/docs/Policies.klUCB.html}{kl-UCB},
  \href{https://smpybandits.github.io/docs/Policies.MOSS.html}{\texttt{MOSS}}
  and
  \href{https://smpybandits.github.io/docs/Policies.Thompson.html}{Thompson
  Sampling} algorithms, as well as other less known algorithms
  (\href{https://smpybandits.github.io/docs/Policies.OCUCB.html}{\texttt{OCUCB}},
  \href{https://smpybandits.github.io/docs/Policies.OCUCB.html}{\texttt{BESA}},
  \href{https://smpybandits.github.io/docs/Policies.OSSB.html}{\texttt{OSSB}}
  etc).
\item
  Implementation of very recent Multi-Armed Bandits algorithms, e.g.,
  \href{https://smpybandits.github.io/docs/Policies.klUCBPlusPlus.html}{\texttt{kl-UCB++}},
  \href{https://smpybandits.github.io/docs/Policies.UCBdagger.html}{\texttt{UCB-dagger}},
  or
  \href{https://smpybandits.github.io/docs/Policies.MOSSAnytime.html}{\texttt{MOSS-anytime}}
  (from \href{http://proceedings.mlr.press/v48/degenne16.pdf}{this COLT
  2016 article}).
\item
  Experimental policies:
  \href{https://smpybandits.github.io/docs/Policies.BlackBoxOpt.html}{\texttt{BlackBoxOpt}}
  or
  \href{https://smpybandits.github.io/docs/Policies.UnsupervisedLearning.html}{\texttt{UnsupervisedLearning}}
  (using Gaussian processes to learn the arms distributions).
\end{itemize}

\hypertarget{arms-and-problems}{%
\paragraph{Arms and problems}\label{arms-and-problems}}

\begin{itemize}
\tightlist
\item
  The framework mainly targets stochastic bandits, with arms following
  \href{https://smpybandits.github.io/docs/Arms/Bernoulli.html}{\texttt{Bernoulli}},
  bounded (truncated) or unbounded
  \href{https://smpybandits.github.io/docs/Arms/Gaussian.html}{\texttt{Gaussian}},
  \href{https://smpybandits.github.io/docs/Arms/Exponential.html}{\texttt{Exponential}},
  \href{https://smpybandits.github.io/docs/Arms/Gamma.html}{\texttt{Gamma}}
  or
  \href{https://smpybandits.github.io/docs/Arms/Poisson.html}{\texttt{Poisson}}
  distributions.
\item
  The default configuration is to use a fixed problem for N repetitions
  (e.g.~1000 repetitions, use
  \href{https://smpybandits.github.io/docs/Environment/MAB.html}{\texttt{MAB.MAB}}),
  but there is also a perfect support for ``Bayesian'' problems where
  the mean vector \(\mu_1,\dots,\mu_K\) change \emph{at every
  repetition} (see
  \href{https://smpybandits.github.io/docs/Environment/MAB.html}{\texttt{MAB.DynamicMAB}}).
\item
  There is also a good support for Markovian problems, see
  \href{https://smpybandits.github.io/docs/Environment/MAB.html}{\texttt{MAB.MarkovianMAB}},
  even though I preferred to not implement policies specifically
  designed for Markovian problems.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{other-remarks}{%
\subsection{Other remarks}\label{other-remarks}}

\begin{itemize}
\tightlist
\item
  The framework is implemented in an imperative and object oriented
  style. Algorithm and arms are represented as classes, and the API of
  the \texttt{Arms}, \texttt{Policy} and \texttt{MultiPlayersPolicy}
  classes is \href{https://smpybandits.github.io/API.html}{clearly
  documented}.
\item
  The code is
  \href{https://smpybandits.github.io/logs/main_pylint_log.txt}{clean},
  and a special care is given to keep it compatible for both
  \href{https://smpybandits.github.io/logs/main_pylint_log.txt}{Python
  2} and
  \href{https://smpybandits.github.io/logs/main_pylint3_log.txt}{Python
  3}.
\item
  The joblib library {[}@joblib{]} is used for the
  \href{https://smpybandits.github.io/docs/Environment.Evaluator.html}{\texttt{Evaluator}}
  classes, so the simulations are easily ran in parallel on multi-core
  machines and servers \footnote{Note that \emph{SMPyBandits} does no
    need a GPU and is not optimized to run on a cluster. In particular,
    it does not take advantage of popular libraries like
    \href{https://github.com/pydata/numexpr}{numexpr},
    \href{http://www.deeplearning.net/software/theano/}{theano} or
    \href{https://www.tensorflow.org/}{tensorflow}.}.
\end{itemize}

\hypertarget{how-to-run-the-experiments}{%
\subsubsection{How to run the experiments
?}\label{how-to-run-the-experiments}}

For example, this short bash snippet \footnote{See
  \href{https://smpybandits.github.io/How_to_run_the_code.html}{this
  page of the documentation} for more details.} shows how to clone the
code, install the requirements for Python 3 (in a virtualenv
{[}@virtualenv{]}), and starts some simulation for \(N=1000\)
repetitions of the default non-Bayesian Bernoulli-distributed problem,
for \(K=9\) arms, an horizon of \(T=10000\) and on \(4\) CPUs \footnote{It
  takes about \(20\) to \(40\) minutes for each simulation, on a
  standard \(4\)-cores \(64\) bits GNU/Linux laptop.}. Using environment
variables (\texttt{N=1000}) when launching the simulation is not
required but it is convenient.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# 1. get the code in /tmp/, or wherever you want}
\BuiltInTok{cd}\NormalTok{ /tmp/}
\FunctionTok{git}\NormalTok{ clone https://GitHub.com/SMPyBandits/SMPyBandits.git}
\BuiltInTok{cd}\NormalTok{ SMPyBandits.git}
\CommentTok{# 2. just be sure you have the latest virtualenv from Python 3}
\FunctionTok{sudo}\NormalTok{ pip3 install --upgrade virtualenv}
\CommentTok{# 3. create and active the virtualenv}
\ExtensionTok{virtualenv3}\NormalTok{ venv }\KeywordTok{||} \ExtensionTok{virtualenv}\NormalTok{ venv}
\BuiltInTok{.} \ExtensionTok{venv/bin/activate}
\CommentTok{# 4. install the requirements in the virtualenv}
\ExtensionTok{pip3}\NormalTok{ install -r requirements.txt}
\CommentTok{# 5. run a single-player simulation!}
\VariableTok{N=}\NormalTok{1000 }\VariableTok{T=}\NormalTok{10000 }\VariableTok{K=}\NormalTok{9 }\VariableTok{N_JOBS=}\NormalTok{4 }\FunctionTok{make}\NormalTok{ single}
\CommentTok{# 6. run a multi-player simulation for 3 players!}
\VariableTok{N=}\NormalTok{1000 }\VariableTok{T=}\NormalTok{10000 }\VariableTok{M=}\NormalTok{3 }\VariableTok{K=}\NormalTok{9 }\VariableTok{N_JOBS=}\NormalTok{4 }\FunctionTok{make}\NormalTok{ moremulti}
\end{Highlighting}
\end{Shaded}

\hypertarget{examples-of-illustrations}{%
\subsubsection{Examples of
illustrations}\label{examples-of-illustrations}}

The two simulations above produce these plots showing the average
cumulated regret \footnote{The regret is the difference between the
  cumulated rewards of the best fixed-armed strategy (which is the
  oracle strategy for stationary bandits) and the cumulated rewards of
  the considered algorithms.} for each algorithm, which is the reference
measure of efficiency for algorithms in the multi-armed bandits
framework.

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth,height=\textheight]{../plots/paper/1.png}
\caption{Single-player simulation showing the regret of \(4\)
algorithms, and the asymptotic lower-bound from {[}@LaiRobbins85{]}.
They all perform very well, and at finite time they are empirically
\emph{below} the asymptotic lower-bound. Each algorithm is known to be
order-optimal (\emph{i.e.}, its regret is proved to match the
lower-bound up-to a constant), and each but UCB is known to be optimal
(\emph{i.e.} with the constant matching the
lower-bound).\label{fig:plot1}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth,height=\textheight]{../plots/paper/2.png}
\caption{Multi-player simulation showing the regret of \(6\) algorithms,
and the asymptotic lower-bound from {[}@BessonALT2018{]}. The best
algorithm is the centralized version, but for decentralized algorithms,
our proposals outperform the previous state-of-the-art
\href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoRand.html}{\texttt{rhoRand}}
policy.\label{fig:plot2}}
\end{figure}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{research-using-smpybandits}{%
\subsection{\texorpdfstring{Research using
\emph{SMPyBandits}}{Research using SMPyBandits}}\label{research-using-smpybandits}}

\emph{SMPyBandits} was used for the following research articles since
\(2017\) \footnote{\href{http://perso.crans.org/besson/}{I (Lilian
  Besson)} have \href{http://perso.crans.org/besson/phd/}{started my
  PhD} in October \(2016\), and this is a part of my \textbf{on going}
  research since December \(2016\). I launched the
  \href{https://smpybandits.github.io/}{documentation} on March
  \(2017\), I wrote my first research articles using this framework in
  \(2017\) and I was finally able to open-source my project in February
  \(2018\).}:

\begin{itemize}
\item
  For this first article, {[}@Bonnefoi17{]}, \emph{SMPyBandits} was not
  used to generate the main figures, but to explore on a smaller scale
  many other approaches (using
  \href{https://smpybandits.github.io/docs/Environment.EvaluatorSparseMultiPlayers.html}{\texttt{EvaluatorSparseMultiPlayers}}).
\item
  For {[}@BessonALT2018{]}, we used \emph{SMPyBandits} for all the
  simulations for multi-player bandit algorithms \footnote{More details
    and illustrations are given on the documentation page,
    \href{https://smpybandits.github.io/MultiPlayers.html}{\texttt{MultiPlayers}}.}.
  We designed the two
  \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.RandTopM.html}{\texttt{RandTopM}}
  and
  \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.MCTopM.html}{\texttt{MCTopM}}
  algorithms and proved than they enjoy logarithmic regret in the usual
  setting, and outperform significantly the previous state-of-the-art
  solutions (\emph{i.e.},
  \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoRand.html}{\texttt{rhoRand}},
  \href{https://smpybandits.github.io/docs/Policies.MEGA.html}{\texttt{MEGA}}
  and
  \href{https://smpybandits.github.io/docs/Policies.MusicalChair.html}{\texttt{MusicalChair}}).
\end{itemize}

\begin{itemize}
\tightlist
\item
  In {[}@BessonWCNC2018{]}, we used \emph{SMPyBandits} to illustrate and
  compare different aggregation algorithms \footnote{More details and
    illustrations are given on the documentation page,
    \href{https://smpybandits.github.io/Aggregation.html}{\texttt{Aggregation}}.}.
  We designed a variant of the Exp3 algorithm for online aggregation of
  experts {[}@Bubeck12{]}, called
  \href{https://smpybandits.github.io/docs/Policies.Aggregator.html}{\texttt{Aggregator}}.
  Aggregating experts is a well-studied idea in sequential learning and
  in machine learning in general. We showed that it can be used in
  practice to select on the run the best bandit algorithm for a certain
  problem from a fixed pool of experts. This idea and algorithm can have
  interesting impact for Opportunistic Spectrum Access applications
  {[}@Jouini09{]} that use multi-armed bandits algorithms for sequential
  learning and network efficiency optimization.
\end{itemize}

\begin{itemize}
\tightlist
\item
  In {[}@Besson2018c{]}, we used \emph{SMPyBandits} to illustrate and
  compare different ``doubling trick'' schemes \footnote{More details
    and illustrations are given on the documentation page,
    \href{https://smpybandits.github.io/DoublingTrick.html}{\texttt{DoublingTrick}}.}.
  In sequential learning, an algorithm is \emph{anytime} if it does not
  need to know the horizon \(T\) of the experiments. A well-known trick
  for transforming any non-anytime algorithm to an anytime variant is
  the ``Doubling Trick'': start with an horizon \(T_0\in\mathbb{N}\),
  and when \(t > T_i\), use \(T_{i+1} = 2 T_i\). We studied two generic
  sequences of growing horizons (geometric and exponential), and we
  proved two theorems that generalized previous results. A geometric
  sequence suffices to minimax regret bounds (in
  \(R_T = \mathcal{O}(\sqrt(T))\)), with a constant multiplicative loss
  \(\ell \leq 4\), but cannot be used to conserve a logarithmic regret
  bound (in \(R_T = \mathcal{O}(\log(T))\)). And an exponential sequence
  can be used to conserve logarithmic bounds, with a constant
  multiplicative loss also \(\ell \leq 4\) in the usual setting. It is
  still an open question to know if a well-tuned exponential sequence
  can conserve minimax bounds or weak minimax bounds (in
  \(R_T = \mathcal{O}(\sqrt{T \log(T)})\)).
\end{itemize}

\hypertarget{dependencies}{%
\subsection{Dependencies}\label{dependencies}}

The framework is written in Python {[}@python{]}, using matplotlib
{[}@matplotlib{]} for 2D plotting, numpy {[}@numpy{]} for data storing,
random number generations and and operations on arrays, scipy
{[}@scipy{]} for statistical and special functions, and seaborn
{[}@seaborn{]} for pretty plotting and colorblind-aware colormaps.
Optional dependencies include joblib {[}@joblib{]} for parallel
simulations, numba {[}@numba{]} for automatic speed-up on some small
functions, as well as sphinx {[}@sphinx{]} for generating the
documentations. I also acknowledge the use of virtualenv
{[}@virtualenv{]} for launching simulations in isolated environments,
and jupyter {[}@jupyter{]} used with ipython {[}@ipython{]} to
experiment with the code.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

### 2019-03-25 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex
Je suis bourrin mais je git TOUT pour éviter une catastrophe...

\begin{tikzpicture}[>=latex,line join=bevel,scale=0.5]
%%
  \node (155) at (451.19bp,41.0bp) [draw=red,rectangle,thick] {[[1/2,0/1], [1/2,0/1]]};
  \node (10) at (635.19bp,215.0bp) [draw,rectangle] {[[0/0,1/1], [0/1,0/0]]};
  \node (3) at (267.19bp,215.0bp) [draw,rectangle] {[[0/0,0/1], [1/1,0/0]]};
  \node (6) at (451.19bp,215.0bp) [draw,rectangle] {[[1/1,0/0], [0/0,0/1]]};
  \node (14) at (83.193bp,128.0bp) [draw=red,rectangle] {[[0/1,1/1], [0/1,1/1]]};
  \node (1) at (83.193bp,215.0bp) [draw,rectangle] {[[0/1,0/0], [0/0,1/1]]};
  \node (0) at (359.19bp,302.0bp) [draw=darkgreen,rectangle] {[[0/0,0/0], [0/0,0/0]]};
  \node (22) at (267.19bp,128.0bp) [draw=red,rectangle] {[[1/1,0/1], [1/1,0/1]]};
  \node (234) at (635.19bp,41.0bp) [draw=red,rectangle,thick] {[[0/1,1/2], [0/1,1/2]]};
  \node (65) at (83.193bp,41.0bp) [draw=red,rectangle,thick] {[[0/1,1/2], [0/1,1/2]]};
  \node (110) at (267.19bp,41.0bp) [draw=red,rectangle,thick] {[[1/2,0/1], [1/2,0/1]]};
  \node (43) at (635.19bp,128.0bp) [draw=red,rectangle] {[[0/1,1/1], [0/1,1/1]]};
  \node (30) at (451.19bp,128.0bp) [draw=red,rectangle] {[[1/1,0/1], [1/1,0/1]]};
  \draw [red,->] (43) ..controls (635.19bp,98.163bp) and (635.19bp,82.548bp)  .. (234);
  \draw (645.69bp,84.5bp) node {$1$};
  \draw [black,->] (0) ..controls (225.73bp,289.45bp) and (154.26bp,280.04bp)  .. (128.19bp,266.0bp) .. controls (117.33bp,260.15bp) and (107.77bp,250.61bp)  .. (1);
  \draw (199.19bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (300.86bp,281.92bp) and (290.11bp,275.07bp)  .. (282.19bp,266.0bp) .. controls (276.58bp,259.57bp) and (273.06bp,251.1bp)  .. (3);
  \draw (353.19bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [red,->] (22) ..controls (267.19bp,98.163bp) and (267.19bp,82.548bp)  .. (110);
  \draw (277.69bp,84.5bp) node {$1$};
  \draw [black,->] (0) ..controls (485.53bp,296.17bp) and (538.83bp,287.27bp)  .. (582.19bp,266.0bp) .. controls (594.44bp,259.99bp) and (605.77bp,250.18bp)  .. (10);
  \draw (674.19bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (1) ..controls (83.193bp,185.16bp) and (83.193bp,169.55bp)  .. (14);
  \draw (149.19bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (10) ..controls (635.19bp,185.16bp) and (635.19bp,169.55bp)  .. (43);
  \draw (701.19bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (6) ..controls (451.19bp,185.16bp) and (451.19bp,169.55bp)  .. (30);
  \draw (517.19bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [red,->] (14) ..controls (83.193bp,98.163bp) and (83.193bp,82.548bp)  .. (65);
  \draw (93.693bp,84.5bp) node {$1$};
  \draw [red,->] (30) ..controls (451.19bp,98.163bp) and (451.19bp,82.548bp)  .. (155);
  \draw (461.69bp,84.5bp) node {$1$};
  \draw [black,->] (0) ..controls (406.04bp,280.6bp) and (416.21bp,273.97bp)  .. (424.19bp,266.0bp) .. controls (430.81bp,259.39bp) and (436.24bp,250.81bp)  .. (6);
  \draw (507.19bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (3) ..controls (267.19bp,185.16bp) and (267.19bp,169.55bp)  .. (22);
  \draw (333.19bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
%
\end{tikzpicture}

### 2019-03-25 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__onlyfig.gv.tex
Je suis bourrin mais je git TOUT pour éviter une catastrophe...

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (18391.0bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (105) at (11123.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (39) at (31639.0bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [0/1,0/1]]};
  \node (101) at (10387.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (30) at (20228.0bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
  \node (206) at (29707.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (13) at (3211.2bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,1/1]]};
  \node (181) at (25107.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (175) at (24003.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (7) at (20978.0bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/1,0/0]]};
  \node (73) at (5235.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,1/1]]};
  \node (241) at (36147.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,0/2]]};
  \node (159) at (21059.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [0/2,0/1]]};
  \node (93) at (8915.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,1/1]]};
  \node (184) at (25659.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (123) at (14435.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,2/2]]};
  \node (219) at (32099.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (189) at (26579.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,2/2]]};
  \node (230) at (34123.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (198) at (28235.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/2,0/1]]};
  \node (22) at (11949.0bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
  \node (19) at (9285.2bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,1/1]]};
  \node (14) at (3579.2bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
  \node (8) at (26303.0bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [0/1,0/0]]};
  \node (20) at (10664.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [1/1,1/1]]};
  \node (154) at (20139.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,0/2]]};
  \node (35) at (27039.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,1/1]]};
  \node (0) at (19381.0bp,302.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (160) at (21243.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (171) at (23267.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (213) at (30995.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (16) at (5787.2bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,1/1]]};
  \node (221) at (32467.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (236) at (35227.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (109) at (11859.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,0/1]]};
  \node (107) at (11491.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (232) at (34491.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [0/2,0/1]]};
  \node (130) at (15723.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,1/2]]};
  \node (162) at (21611.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,0/2]]};
  \node (178) at (24555.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (194) at (27499.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,1/2]]};
  \node (204) at (29339.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (62) at (3211.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,1/1]]};
  \node (15) at (4407.2bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,0/1]]};
  \node (103) at (10755.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,2/2]]};
  \node (143) at (18115.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (12) at (2382.2bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,1/1]]};
  \node (115) at (12963.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (217) at (31731.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (145) at (18483.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (216) at (31547.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (57) at (2291.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,2/2]]};
  \node (112) at (12411.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (27) at (17563.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (242) at (36331.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,0/2]]};
  \node (152) at (19771.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (141) at (17747.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (244) at (36699.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,0/2]]};
  \node (214) at (31179.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,0/1]]};
  \node (64) at (3579.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (53) at (1555.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,1/1]]};
  \node (37) at (29523.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (177) at (24371.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (98) at (9835.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,1/1]]};
  \node (131) at (15907.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (50) at (1003.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (191) at (26947.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (84) at (7259.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,1/1]]};
  \node (163) at (21795.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,0/2]]};
  \node (166) at (22347.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (47) at (451.19bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,2/2]]};
  \node (91) at (8547.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,1/2]]};
  \node (170) at (23083.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (60) at (2843.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (183) at (25475.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (89) at (8179.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (106) at (11307.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,0/1]]};
  \node (146) at (18667.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,0/1]]};
  \node (111) at (12227.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (52) at (1371.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (210) at (30443.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (61) at (3027.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/1,1/2]]};
  \node (196) at (27867.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (174) at (23819.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (121) at (14067.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,1/1]]};
  \node (79) at (6339.2bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,2/2]]};
  \node (116) at (13147.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (44) at (35779.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,0/1]]};
  \node (185) at (25843.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (240) at (35963.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (234) at (34859.0bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (151) at (19587.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (164) at (21979.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (155) at (20323.0bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (87) at (7811.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (207) at (29891.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (74) at (5419.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,2/2]]};
  \node (188) at (26395.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (63) at (3395.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,1/1]]};
  \node (220) at (32283.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,0/1]]};
  \node (10) at (34309.0bp,215.0bp) [draw,ellipse] {[[0/0,1/1], [0/1,0/0]]};
  \node (129) at (15539.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,1/1]]};
  \node (38) at (30811.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,0/1]]};
  \node (137) at (17011.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (197) at (28051.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (33) at (25107.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (132) at (16091.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,2/2]]};
  \node (128) at (15355.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,2/2]]};
  \node (205) at (29523.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (6) at (19769.0bp,215.0bp) [draw,ellipse] {[[1/1,0/0], [0/0,0/1]]};
  \node (228) at (33755.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,2/2]]};
  \node (153) at (19955.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/2,0/1]]};
  \node (85) at (7443.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (158) at (20875.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/2,0/1]]};
  \node (193) at (27315.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (104) at (10939.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,1/2]]};
  \node (239) at (35779.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (148) at (19035.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,0/1]]};
  \node (113) at (12595.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (118) at (13515.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,2/2]]};
  \node (97) at (9651.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (139) at (17379.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (173) at (23635.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (86) at (7627.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,0/1]]};
  \node (225) at (33203.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (199) at (28419.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,0/2]]};
  \node (127) at (15171.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,1/2]]};
  \node (150) at (19403.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (203) at (29155.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (218) at (31915.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [0/2,0/1]]};
  \node (25) at (14711.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,1/1]]};
  \node (229) at (33939.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/1,0/2]]};
  \node (133) at (16275.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (36) at (27958.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,0/1]]};
  \node (209) at (30259.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (5) at (18994.0bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [0/0,0/1]]};
  \node (149) at (19219.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (43) at (34765.0bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
  \node (125) at (14803.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,1/1]]};
  \node (223) at (32835.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (110) at (12043.0bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (3) at (11491.0bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [1/1,0/0]]};
  \node (67) at (4131.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,0/1]]};
  \node (78) at (6155.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,1/2]]};
  \node (165) at (22163.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,0/2]]};
  \node (94) at (9099.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,1/1]]};
  \node (117) at (13331.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/2,1/1]]};
  \node (237) at (35411.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/2,0/1]]};
  \node (122) at (14251.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,1/1]]};
  \node (51) at (1187.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (49) at (819.19bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,1/1]]};
  \node (34) at (26303.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,1/1]]};
  \node (157) at (20691.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (119) at (13699.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,1/2]]};
  \node (192) at (27131.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,2/2]]};
  \node (82) at (6891.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/1,1/2]]};
  \node (24) at (13424.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [1/1,1/1]]};
  \node (32) at (23083.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [0/1,0/1]]};
  \node (68) at (4315.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,0/1]]};
  \node (18) at (7811.2bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,0/1]]};
  \node (31) at (21243.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,0/1]]};
  \node (235) at (35043.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/2,0/1]]};
  \node (172) at (23451.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (46) at (267.19bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,1/2]]};
  \node (208) at (30075.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (142) at (17931.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (224) at (33019.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,1/2]]};
  \node (66) at (3947.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (168) at (22715.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (226) at (33387.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,2/2]]};
  \node (227) at (33571.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (202) at (28971.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (42) at (34309.0bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [0/1,0/1]]};
  \node (41) at (33478.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,1/1]]};
  \node (69) at (4499.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,0/1]]};
  \node (23) at (12595.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,1/1]]};
  \node (77) at (5971.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/1,1/2]]};
  \node (135) at (16643.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (238) at (35595.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [0/2,0/1]]};
  \node (161) at (21427.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (76) at (5787.2bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,1/1]]};
  \node (138) at (17195.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (144) at (18299.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (88) at (7995.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (108) at (11675.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (179) at (24739.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (45) at (83.193bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (156) at (20507.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/2,0/1]]};
  \node (17) at (7075.2bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,1/1]]};
  \node (80) at (6523.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,1/1]]};
  \node (212) at (30811.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (140) at (17563.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (186) at (26027.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (58) at (2475.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (55) at (1923.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (211) at (30627.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,0/1]]};
  \node (147) at (18851.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,0/1]]};
  \node (102) at (10571.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/2,1/1]]};
  \node (40) at (32467.0bp,128.0bp) [draw,ellipse] {[[0/1,1/1], [1/1,0/1]]};
  \node (222) at (32651.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (65) at (3763.2bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (92) at (8731.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,2/2]]};
  \node (124) at (14619.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,1/2]]};
  \node (180) at (24923.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (72) at (5051.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,1/1]]};
  \node (1) at (3211.2bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/0,1/1]]};
  \node (83) at (7075.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,1/1]]};
  \node (59) at (2659.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,1/2]]};
  \node (233) at (34675.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (187) at (26211.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (29) at (19769.0bp,128.0bp) [draw,ellipse] {[[1/1,0/1], [0/1,0/1]]};
  \node (200) at (28603.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (81) at (6707.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (26) at (15999.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [0/1,1/1]]};
  \node (100) at (10203.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (190) at (26763.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (201) at (28787.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (28) at (18943.0bp,128.0bp) [draw,ellipse] {[[1/1,1/1], [1/1,0/1]]};
  \node (21) at (11491.0bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,0/1]]};
  \node (9) at (30811.0bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/0,0/1]]};
  \node (176) at (24187.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (2) at (7075.2bp,215.0bp) [draw,ellipse] {[[0/0,1/1], [1/1,0/0]]};
  \node (95) at (9283.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (75) at (5603.2bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,1/2]]};
  \node (11) at (1005.2bp,128.0bp) [draw,ellipse] {[[0/1,0/1], [1/1,1/1]]};
  \node (169) at (22899.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (54) at (1739.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/1,1/2]]};
  \node (126) at (14987.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/1,1/2]]};
  \node (90) at (8363.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (96) at (9467.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (70) at (4683.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (56) at (2107.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (195) at (27683.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (71) at (4867.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,1/1]]};
  \node (231) at (34307.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (99) at (10019.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/1,1/2]]};
  \node (120) at (13883.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,1/1]]};
  \node (4) at (14711.0bp,215.0bp) [draw,ellipse] {[[1/1,0/0], [0/0,1/1]]};
  \node (215) at (31363.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/1,0/2]]};
  \node (167) at (22531.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (243) at (36515.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (182) at (25291.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (114) at (12779.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,2/2]]};
  \node (48) at (635.19bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,1/1]]};
  \node (134) at (16459.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,2/2]]};
  \node (136) at (16827.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \draw [black,->] (39) ..controls (31692.0bp,107.22bp) and (31702.0bp,100.51bp)  .. (31710.0bp,92.0bp) .. controls (31716.0bp,85.469bp) and (31721.0bp,76.875bp)  .. (217);
  \draw (31790.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (36) ..controls (27838.0bp,118.26bp) and (27790.0bp,109.48bp)  .. (27749.0bp,92.0bp) .. controls (27734.0bp,85.571bp) and (27720.0bp,75.084bp)  .. (195);
  \draw (27810.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (25) ..controls (14579.0bp,114.93bp) and (14509.0bp,105.45bp)  .. (14483.0bp,92.0bp) .. controls (14472.0bp,86.08bp) and (14462.0bp,76.405bp)  .. (123);
  \draw (14554.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (37) ..controls (29459.0bp,97.451bp) and (29415.0bp,76.913bp)  .. (204);
  \draw (29458.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (18) ..controls (7916.8bp,117.38bp) and (7945.2bp,108.42bp)  .. (7967.2bp,92.0bp) .. controls (7975.2bp,86.062bp) and (7981.2bp,77.147bp)  .. (88);
  \draw (8047.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (33) ..controls (25273.0bp,122.71bp) and (25395.0bp,114.74bp)  .. (25436.0bp,92.0bp) .. controls (25446.0bp,86.408bp) and (25455.0bp,77.041bp)  .. (183);
  \draw (25469.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (16) ..controls (5885.1bp,114.44bp) and (5911.7bp,106.03bp)  .. (5933.2bp,92.0bp) .. controls (5942.5bp,85.923bp) and (5950.4bp,76.751bp)  .. (77);
  \draw (5965.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (20) ..controls (10539.0bp,121.01bp) and (10487.0bp,112.13bp)  .. (10444.0bp,92.0bp) .. controls (10431.0bp,85.909bp) and (10419.0bp,75.953bp)  .. (101);
  \draw (10526.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (2) ..controls (7284.3bp,189.86bp) and (7584.5bp,155.18bp)  .. (18);
  \draw (7564.2bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (35) ..controls (27170.0bp,114.59bp) and (27239.0bp,105.07bp)  .. (27265.0bp,92.0bp) .. controls (27277.0bp,86.012bp) and (27288.0bp,76.199bp)  .. (193);
  \draw (27367.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (33) ..controls (25043.0bp,97.451bp) and (24999.0bp,76.913bp)  .. (180);
  \draw (25042.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (11) ..controls (746.55bp,125.93bp) and (343.19bp,120.68bp)  .. (209.19bp,92.0bp) .. controls (179.53bp,85.651bp) and (147.8bp,73.021bp)  .. (45);
  \draw (270.19bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (3) ..controls (11636.0bp,187.06bp) and (11789.0bp,158.81bp)  .. (22);
  \draw (11813.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (23) ..controls (12490.0bp,117.38bp) and (12461.0bp,108.42bp)  .. (12439.0bp,92.0bp) .. controls (12431.0bp,86.062bp) and (12425.0bp,77.147bp)  .. (112);
  \draw (12515.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (35) ..controls (26918.0bp,118.94bp) and (26869.0bp,110.13bp)  .. (26827.0bp,92.0bp) .. controls (26813.0bp,85.664bp) and (26799.0bp,75.32bp)  .. (190);
  \draw (26898.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (37) ..controls (29523.0bp,98.163bp) and (29523.0bp,82.548bp)  .. (205);
  \draw (29594.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (33) ..controls (25327.0bp,120.68bp) and (25601.0bp,110.43bp)  .. (25695.0bp,92.0bp) .. controls (25730.0bp,85.143bp) and (25769.0bp,72.053bp)  .. (185);
  \draw (25830.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (18281.0bp,293.4bp) and (8245.3bp,224.08bp)  .. (2);
  \draw (14025.0bp,258.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (40) ..controls (32392.0bp,108.95bp) and (32370.0bp,101.5bp)  .. (32351.0bp,92.0bp) .. controls (32337.0bp,84.707bp) and (32322.0bp,74.303bp)  .. (220);
  \draw (32407.0bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (10) ..controls (34081.0bp,190.71bp) and (33724.0bp,154.16bp)  .. (41);
  \draw (33999.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (6) ..controls (20101.0bp,194.85bp) and (20888.0bp,149.5bp)  .. (31);
  \draw (20660.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (19) ..controls (9478.8bp,120.96bp) and (9673.2bp,111.74bp)  .. (9741.2bp,92.0bp) .. controls (9762.8bp,85.72bp) and (9785.1bp,74.102bp)  .. (98);
  \draw (9847.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (35590.0bp,124.38bp) and (35400.0bp,117.28bp)  .. (35240.0bp,92.0bp) .. controls (35192.0bp,84.334bp) and (35138.0bp,70.222bp)  .. (235);
  \draw (35256.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (33) ..controls (25266.0bp,120.21bp) and (25386.0bp,111.46bp)  .. (25488.0bp,92.0bp) .. controls (25529.0bp,84.212bp) and (25574.0bp,70.776bp)  .. (184);
  \draw (25620.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (15) ..controls (4288.5bp,117.67bp) and (4240.5bp,108.92bp)  .. (4200.2bp,92.0bp) .. controls (4184.8bp,85.536bp) and (4169.5bp,75.044bp)  .. (67);
  \draw (4261.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (29) ..controls (19877.0bp,112.58bp) and (19919.0bp,103.81bp)  .. (19933.0bp,92.0bp) .. controls (19940.0bp,86.147bp) and (19945.0bp,77.474bp)  .. (153);
  \draw (19960.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (36) ..controls (28099.0bp,118.46bp) and (28177.0bp,109.49bp)  .. (28204.0bp,92.0bp) .. controls (28213.0bp,86.28bp) and (28220.0bp,77.133bp)  .. (198);
  \draw (28234.0bp,84.5bp) node {$1/2$};
  \draw [red,->] (14) ..controls (3677.1bp,114.44bp) and (3703.7bp,106.03bp)  .. (3725.2bp,92.0bp) .. controls (3734.5bp,85.923bp) and (3742.4bp,76.751bp)  .. (65);
  \draw (3752.7bp,84.5bp) node {$1$};
  \draw [black,->] (10) ..controls (34641.0bp,194.83bp) and (35424.0bp,149.54bp)  .. (44);
  \draw (35198.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (33) ..controls (25107.0bp,98.163bp) and (25107.0bp,82.548bp)  .. (181);
  \draw (25178.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (24) ..controls (13245.0bp,120.78bp) and (13090.0bp,111.82bp)  .. (13035.0bp,92.0bp) .. controls (13019.0bp,85.991bp) and (13002.0bp,75.206bp)  .. (115);
  \draw (13106.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (40) ..controls (32573.0bp,117.38bp) and (32601.0bp,108.42bp)  .. (32623.0bp,92.0bp) .. controls (32631.0bp,86.062bp) and (32637.0bp,77.147bp)  .. (222);
  \draw (32702.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (4) ..controls (14406.0bp,193.83bp) and (13751.0bp,150.6bp)  .. (24);
  \draw (14210.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (13) ..controls (3211.2bp,98.163bp) and (3211.2bp,82.548bp)  .. (62);
  \draw (3272.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (25) ..controls (14496.0bp,124.1bp) and (14244.0bp,116.92bp)  .. (14158.0bp,92.0bp) .. controls (14137.0bp,85.931bp) and (14116.0bp,74.475bp)  .. (121);
  \draw (14229.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (35613.0bp,122.94bp) and (35490.0bp,115.06bp)  .. (35449.0bp,92.0bp) .. controls (35439.0bp,86.401bp) and (35431.0bp,77.032bp)  .. (237);
  \draw (35531.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (2) ..controls (6769.6bp,193.83bp) and (6114.5bp,150.6bp)  .. (16);
  \draw (6565.2bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (9) ..controls (31038.0bp,190.67bp) and (31393.0bp,154.25bp)  .. (39);
  \draw (31340.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (31) ..controls (21054.0bp,124.38bp) and (20864.0bp,117.28bp)  .. (20704.0bp,92.0bp) .. controls (20656.0bp,84.334bp) and (20602.0bp,70.222bp)  .. (156);
  \draw (20720.0bp,84.5bp) node {$1/4$};
  \draw [red,->] (22) ..controls (11982.0bp,97.517bp) and (12001.0bp,79.93bp)  .. (110);
  \draw (12012.0bp,84.5bp) node {$1$};
  \draw [black,->] (12) ..controls (2327.7bp,107.54bp) and (2317.1bp,100.76bp)  .. (2309.2bp,92.0bp) .. controls (2303.3bp,85.478bp) and (2299.3bp,76.886bp)  .. (57);
  \draw (2380.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (41) ..controls (33344.0bp,115.41bp) and (33264.0bp,106.15bp)  .. (33194.0bp,92.0bp) .. controls (33153.0bp,83.582bp) and (33107.0bp,70.306bp)  .. (224);
  \draw (33210.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (21) ..controls (11633.0bp,123.74bp) and (11713.0bp,115.46bp)  .. (11779.0bp,92.0bp) .. controls (11797.0bp,85.626bp) and (11815.0bp,74.665bp)  .. (109);
  \draw (11825.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (41) ..controls (33619.0bp,118.46bp) and (33697.0bp,109.49bp)  .. (33724.0bp,92.0bp) .. controls (33733.0bp,86.28bp) and (33740.0bp,77.133bp)  .. (228);
  \draw (33799.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (39) ..controls (31819.0bp,121.39bp) and (31976.0bp,112.72bp)  .. (32031.0bp,92.0bp) .. controls (32047.0bp,86.062bp) and (32062.0bp,75.426bp)  .. (219);
  \draw (32118.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (37) ..controls (29689.0bp,122.71bp) and (29811.0bp,114.74bp)  .. (29852.0bp,92.0bp) .. controls (29862.0bp,86.408bp) and (29871.0bp,77.041bp)  .. (207);
  \draw (29885.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (8) ..controls (26660.0bp,195.69bp) and (27576.0bp,148.62bp)  .. (36);
  \draw (27295.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (27) ..controls (17783.0bp,120.68bp) and (18057.0bp,110.43bp)  .. (18151.0bp,92.0bp) .. controls (18186.0bp,85.143bp) and (18225.0bp,72.053bp)  .. (144);
  \draw (18286.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (19255.0bp,273.25bp) and (19135.0bp,246.99bp)  .. (5);
  \draw (19226.0bp,258.5bp) node {$1/4$};
  \draw [black,->] (13) ..controls (3362.5bp,116.15bp) and (3466.0bp,106.08bp)  .. (3504.2bp,92.0bp) .. controls (3521.3bp,85.675bp) and (3538.4bp,74.718bp)  .. (64);
  \draw (3604.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (36) ..controls (27904.0bp,107.54bp) and (27893.0bp,100.76bp)  .. (27885.0bp,92.0bp) .. controls (27879.0bp,85.478bp) and (27875.0bp,76.886bp)  .. (196);
  \draw (27956.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (38) ..controls (30953.0bp,123.74bp) and (31033.0bp,115.46bp)  .. (31099.0bp,92.0bp) .. controls (31117.0bp,85.626bp) and (31135.0bp,74.665bp)  .. (214);
  \draw (31145.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (2) ..controls (7075.2bp,185.16bp) and (7075.2bp,169.55bp)  .. (17);
  \draw (7141.2bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (25) ..controls (14915.0bp,119.43bp) and (15139.0bp,108.76bp)  .. (15218.0bp,92.0bp) .. controls (15250.0bp,85.155bp) and (15285.0bp,72.364bp)  .. (128);
  \draw (15329.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (35) ..controls (27226.0bp,125.61bp) and (27396.0bp,119.22bp)  .. (27452.0bp,92.0bp) .. controls (27464.0bp,86.384bp) and (27474.0bp,76.573bp)  .. (194);
  \draw (27487.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (9) ..controls (31168.0bp,195.69bp) and (32085.0bp,148.62bp)  .. (40);
  \draw (31794.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (38) ..controls (30645.0bp,122.94bp) and (30522.0bp,115.06bp)  .. (30481.0bp,92.0bp) .. controls (30471.0bp,86.401bp) and (30463.0bp,77.032bp)  .. (210);
  \draw (30563.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (31) ..controls (21045.0bp,123.06bp) and (20840.0bp,115.18bp)  .. (20770.0bp,92.0bp) .. controls (20752.0bp,86.013bp) and (20734.0bp,74.976bp)  .. (157);
  \draw (20831.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (0) ..controls (20439.0bp,293.13bp) and (29686.0bp,224.37bp)  .. (9);
  \draw (25862.0bp,258.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (36) ..controls (28009.0bp,106.92bp) and (28019.0bp,100.26bp)  .. (28027.0bp,92.0bp) .. controls (28034.0bp,85.47bp) and (28039.0bp,76.913bp)  .. (197);
  \draw (28119.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (11) ..controls (893.33bp,112.15bp) and (839.38bp,103.33bp)  .. (792.19bp,92.0bp) .. controls (756.05bp,83.324bp) and (716.15bp,70.574bp)  .. (48);
  \draw (807.69bp,84.5bp) node {$1/4$};
  \draw [black,->] (8) ..controls (26512.0bp,189.86bp) and (26813.0bp,155.18bp)  .. (35);
  \draw (26781.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (44) ..controls (36002.0bp,121.59bp) and (36283.0bp,112.09bp)  .. (36379.0bp,92.0bp) .. controls (36411.0bp,85.309bp) and (36446.0bp,72.45bp)  .. (243);
  \draw (36489.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (16) ..controls (5892.6bp,110.89bp) and (5941.9bp,102.17bp)  .. (5985.2bp,92.0bp) .. controls (6024.4bp,82.796bp) and (6068.0bp,69.925bp)  .. (78);
  \draw (6123.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (28) ..controls (18995.0bp,107.03bp) and (19005.0bp,100.36bp)  .. (19013.0bp,92.0bp) .. controls (19019.0bp,85.554bp) and (19024.0bp,77.115bp)  .. (148);
  \draw (19039.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (11) ..controls (838.77bp,123.64bp) and (706.24bp,115.96bp)  .. (594.19bp,92.0bp) .. controls (560.6bp,84.818bp) and (524.07bp,71.95bp)  .. (47);
  \draw (665.19bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (27) ..controls (17337.0bp,122.9bp) and (17047.0bp,114.56bp)  .. (16948.0bp,92.0bp) .. controls (16920.0bp,85.556bp) and (16890.0bp,73.136bp)  .. (136);
  \draw (17019.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (6) ..controls (19543.0bp,190.68bp) and (19189.0bp,154.3bp)  .. (28);
  \draw (19461.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (23) ..controls (12439.0bp,118.09bp) and (12329.0bp,108.54bp)  .. (12289.0bp,92.0bp) .. controls (12275.0bp,85.987bp) and (12261.0bp,75.584bp)  .. (111);
  \draw (12355.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (4) ..controls (14711.0bp,185.16bp) and (14711.0bp,169.55bp)  .. (25);
  \draw (14767.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (31) ..controls (21398.0bp,118.73bp) and (21512.0bp,109.61bp)  .. (21609.0bp,92.0bp) .. controls (21654.0bp,83.893bp) and (21704.0bp,70.26bp)  .. (163);
  \draw (21758.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (25) ..controls (14446.0bp,122.52bp) and (14019.0bp,112.89bp)  .. (13956.0bp,92.0bp) .. controls (13939.0bp,86.313bp) and (13923.0bp,75.449bp)  .. (120);
  \draw (14038.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (42) ..controls (34309.0bp,98.163bp) and (34308.0bp,82.548bp)  .. (231);
  \draw (34379.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (26) ..controls (15931.0bp,106.74bp) and (15905.0bp,99.058bp)  .. (15882.0bp,92.0bp) .. controls (15847.0bp,81.317bp) and (15808.0bp,69.051bp)  .. (130);
  \draw (15898.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (39) ..controls (31566.0bp,110.32bp) and (31555.0bp,102.83bp)  .. (31547.0bp,92.0bp) .. controls (31543.0bp,85.441bp) and (31541.0bp,77.158bp)  .. (216);
  \draw (31629.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (41) ..controls (33529.0bp,106.92bp) and (33539.0bp,100.26bp)  .. (33547.0bp,92.0bp) .. controls (33554.0bp,85.47bp) and (33559.0bp,76.913bp)  .. (227);
  \draw (33639.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (36) ..controls (28096.0bp,116.39bp) and (28180.0bp,107.19bp)  .. (28253.0bp,92.0bp) .. controls (28292.0bp,83.853bp) and (28336.0bp,70.708bp)  .. (199);
  \draw (28379.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (13) ..controls (3293.8bp,110.32bp) and (3317.3bp,102.66bp)  .. (3337.2bp,92.0bp) .. controls (3349.8bp,85.271bp) and (3362.0bp,75.397bp)  .. (63);
  \draw (3429.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (34) ..controls (26240.0bp,108.63bp) and (26229.0bp,101.61bp)  .. (26221.0bp,92.0bp) .. controls (26216.0bp,85.695bp) and (26213.0bp,77.483bp)  .. (187);
  \draw (26297.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (4) ..controls (14475.0bp,204.63bp) and (14127.0bp,190.76bp)  .. (13834.0bp,179.0bp) .. controls (13405.0bp,161.76bp) and (12891.0bp,140.98bp)  .. (23);
  \draw (13910.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (34) ..controls (26426.0bp,120.04bp) and (26477.0bp,111.18bp)  .. (26519.0bp,92.0bp) .. controls (26533.0bp,85.815bp) and (26546.0bp,75.718bp)  .. (189);
  \draw (26598.0bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (38) ..controls (30713.0bp,114.44bp) and (30687.0bp,106.03bp)  .. (30665.0bp,92.0bp) .. controls (30656.0bp,85.923bp) and (30648.0bp,76.751bp)  .. (211);
  \draw (30736.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (12) ..controls (2432.7bp,106.92bp) and (2443.1bp,100.26bp)  .. (2451.2bp,92.0bp) .. controls (2457.6bp,85.47bp) and (2462.5bp,76.913bp)  .. (58);
  \draw (2543.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (33) ..controls (25205.0bp,114.44bp) and (25232.0bp,106.03bp)  .. (25253.0bp,92.0bp) .. controls (25263.0bp,85.923bp) and (25270.0bp,76.751bp)  .. (182);
  \draw (25351.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (21) ..controls (11325.0bp,122.94bp) and (11202.0bp,115.06bp)  .. (11161.0bp,92.0bp) .. controls (11151.0bp,86.401bp) and (11143.0bp,77.032bp)  .. (105);
  \draw (11243.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (31) ..controls (21243.0bp,98.163bp) and (21243.0bp,82.548bp)  .. (160);
  \draw (21314.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (25) ..controls (14950.0bp,122.66bp) and (15282.0bp,113.85bp)  .. (15394.0bp,92.0bp) .. controls (15429.0bp,85.331bp) and (15466.0bp,72.321bp)  .. (129);
  \draw (15462.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (32) ..controls (23270.0bp,123.4bp) and (23455.0bp,115.78bp)  .. (23611.0bp,92.0bp) .. controls (23663.0bp,84.136bp) and (23721.0bp,69.81bp)  .. (174);
  \draw (23757.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (1) ..controls (3332.2bp,186.05bp) and (3443.8bp,160.27bp)  .. (14);
  \draw (3484.2bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (35) ..controls (27004.0bp,105.35bp) and (26995.0bp,98.856bp)  .. (26987.0bp,92.0bp) .. controls (26979.0bp,84.545bp) and (26971.0bp,75.378bp)  .. (191);
  \draw (27048.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (19) ..controls (9382.8bp,114.44bp) and (9409.0bp,106.02bp)  .. (9430.2bp,92.0bp) .. controls (9439.4bp,85.915bp) and (9447.1bp,76.741bp)  .. (96);
  \draw (9517.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (17) ..controls (7157.8bp,110.32bp) and (7181.3bp,102.66bp)  .. (7201.2bp,92.0bp) .. controls (7213.8bp,85.271bp) and (7226.0bp,75.397bp)  .. (84);
  \draw (7293.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (23) ..controls (12595.0bp,98.163bp) and (12595.0bp,82.548bp)  .. (113);
  \draw (12661.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (24) ..controls (13456.0bp,97.517bp) and (13475.0bp,79.93bp)  .. (118);
  \draw (13535.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (11) ..controls (775.46bp,124.3bp) and (476.36bp,117.25bp)  .. (375.19bp,92.0bp) .. controls (350.16bp,85.753bp) and (323.85bp,73.704bp)  .. (46);
  \draw (456.69bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (18) ..controls (7736.3bp,108.95bp) and (7714.2bp,101.5bp)  .. (7695.2bp,92.0bp) .. controls (7680.6bp,84.707bp) and (7665.8bp,74.303bp)  .. (86);
  \draw (7751.2bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (33) ..controls (24939.0bp,125.15bp) and (24802.0bp,118.01bp)  .. (24687.0bp,92.0bp) .. controls (24656.0bp,85.044bp) and (24623.0bp,72.462bp)  .. (178);
  \draw (24748.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (13) ..controls (3075.8bp,119.8bp) and (3002.9bp,111.12bp)  .. (2941.2bp,92.0bp) .. controls (2919.0bp,85.136bp) and (2895.8bp,73.527bp)  .. (60);
  \draw (3022.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (6) ..controls (19915.0bp,187.06bp) and (20067.0bp,158.81bp)  .. (30);
  \draw (20091.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (16) ..controls (5593.1bp,120.89bp) and (5398.3bp,111.63bp)  .. (5330.2bp,92.0bp) .. controls (5308.4bp,85.718bp) and (5285.8bp,74.099bp)  .. (73);
  \draw (5391.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (24) ..controls (13392.0bp,97.517bp) and (13373.0bp,79.93bp)  .. (117);
  \draw (13399.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (7) ..controls (21390.0bp,197.35bp) and (22643.0bp,146.79bp)  .. (32);
  \draw (22167.0bp,171.5bp) node {$1$};
  \draw [black,->] (32) ..controls (22985.0bp,114.44bp) and (22959.0bp,106.03bp)  .. (22937.0bp,92.0bp) .. controls (22928.0bp,85.923bp) and (22920.0bp,76.751bp)  .. (169);
  \draw (23008.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (11) ..controls (894.24bp,113.39bp) and (851.7bp,104.6bp)  .. (837.19bp,92.0bp) .. controls (830.46bp,86.15bp) and (826.24bp,77.586bp)  .. (49);
  \draw (918.69bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (37) ..controls (29743.0bp,120.68bp) and (30017.0bp,110.43bp)  .. (30111.0bp,92.0bp) .. controls (30146.0bp,85.143bp) and (30185.0bp,72.053bp)  .. (209);
  \draw (30246.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (10) ..controls (34454.0bp,186.97bp) and (34605.0bp,158.83bp)  .. (43);
  \draw (34630.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (25) ..controls (14658.0bp,107.25bp) and (14647.0bp,100.55bp)  .. (14639.0bp,92.0bp) .. controls (14633.0bp,85.48bp) and (14629.0bp,76.888bp)  .. (124);
  \draw (14710.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (21) ..controls (11393.0bp,114.44bp) and (11367.0bp,106.03bp)  .. (11345.0bp,92.0bp) .. controls (11336.0bp,85.923bp) and (11328.0bp,76.751bp)  .. (106);
  \draw (11416.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (19) ..controls (9055.1bp,124.53bp) and (8754.7bp,117.7bp)  .. (8653.2bp,92.0bp) .. controls (8628.7bp,85.79bp) and (8603.0bp,73.821bp)  .. (91);
  \draw (8734.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (17) ..controls (7075.2bp,98.163bp) and (7075.2bp,82.548bp)  .. (83);
  \draw (7136.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (17) ..controls (7011.2bp,97.451bp) and (6966.8bp,76.913bp)  .. (82);
  \draw (7010.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (20) ..controls (10485.0bp,120.78bp) and (10330.0bp,111.82bp)  .. (10275.0bp,92.0bp) .. controls (10259.0bp,85.991bp) and (10242.0bp,75.206bp)  .. (100);
  \draw (10346.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (24) ..controls (13299.0bp,121.01bp) and (13247.0bp,112.13bp)  .. (13204.0bp,92.0bp) .. controls (13191.0bp,85.909bp) and (13179.0bp,75.953bp)  .. (116);
  \draw (13286.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (31) ..controls (21466.0bp,121.59bp) and (21747.0bp,112.09bp)  .. (21843.0bp,92.0bp) .. controls (21875.0bp,85.309bp) and (21910.0bp,72.45bp)  .. (164);
  \draw (21953.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (18) ..controls (7811.2bp,98.163bp) and (7811.2bp,82.548bp)  .. (87);
  \draw (7887.2bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (23) ..controls (12686.0bp,112.09bp) and (12710.0bp,104.12bp)  .. (12731.0bp,92.0bp) .. controls (12742.0bp,85.614bp) and (12752.0bp,76.034bp)  .. (114);
  \draw (12805.0bp,84.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (1) ..controls (3211.2bp,185.16bp) and (3211.2bp,169.55bp)  .. (13);
  \draw (3267.2bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (44) ..controls (35937.0bp,118.39bp) and (36048.0bp,108.95bp)  .. (36087.0bp,92.0bp) .. controls (36101.0bp,86.054bp) and (36114.0bp,75.795bp)  .. (241);
  \draw (36126.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (42) ..controls (34466.0bp,118.48bp) and (34577.0bp,109.06bp)  .. (34616.0bp,92.0bp) .. controls (34630.0bp,86.053bp) and (34643.0bp,75.794bp)  .. (233);
  \draw (34700.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (2) ..controls (6601.3bp,198.9bp) and (4912.6bp,145.1bp)  .. (15);
  \draw (5967.2bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (32) ..controls (22941.0bp,124.0bp) and (22861.0bp,115.76bp)  .. (22794.0bp,92.0bp) .. controls (22776.0bp,85.637bp) and (22758.0bp,74.676bp)  .. (168);
  \draw (22855.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (0) ..controls (18717.0bp,297.82bp) and (15204.0bp,280.94bp)  .. (12345.0bp,266.0bp) .. controls (8608.2bp,246.48bp) and (4011.3bp,220.53bp)  .. (1);
  \draw (12416.0bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (19730.0bp,282.43bp) and (20604.0bp,235.94bp)  .. (7);
  \draw (20291.0bp,258.5bp) node {$1/4$};
  \draw [black,->] (19) ..controls (9173.5bp,113.58bp) and (9130.7bp,104.8bp)  .. (9116.2bp,92.0bp) .. controls (9109.5bp,86.13bp) and (9105.5bp,77.562bp)  .. (94);
  \draw (9197.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (27) ..controls (17395.0bp,125.15bp) and (17258.0bp,118.01bp)  .. (17143.0bp,92.0bp) .. controls (17112.0bp,85.044bp) and (17079.0bp,72.462bp)  .. (137);
  \draw (17204.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [red,->] (43) ..controls (34798.0bp,97.517bp) and (34817.0bp,79.93bp)  .. (234);
  \draw (34828.0bp,84.5bp) node {$1$};
  \draw [black,->] (11) ..controls (1103.6bp,114.7bp) and (1130.0bp,106.23bp)  .. (1151.2bp,92.0bp) .. controls (1160.3bp,85.921bp) and (1167.8bp,76.748bp)  .. (51);
  \draw (1238.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (41) ..controls (33424.0bp,107.54bp) and (33413.0bp,100.76bp)  .. (33405.0bp,92.0bp) .. controls (33399.0bp,85.478bp) and (33395.0bp,76.886bp)  .. (226);
  \draw (33476.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (32) ..controls (23210.0bp,116.39bp) and (23276.0bp,107.53bp)  .. (23332.0bp,92.0bp) .. controls (23359.0bp,84.608bp) and (23388.0bp,72.613bp)  .. (172);
  \draw (23457.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (38) ..controls (30811.0bp,98.163bp) and (30811.0bp,82.548bp)  .. (212);
  \draw (30882.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (12) ..controls (2206.5bp,119.52bp) and (2056.5bp,109.98bp)  .. (2003.2bp,92.0bp) .. controls (1984.8bp,85.786bp) and (1966.2bp,74.584bp)  .. (55);
  \draw (2074.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (9) ..controls (30811.0bp,185.16bp) and (30811.0bp,169.55bp)  .. (38);
  \draw (30877.0bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (27) ..controls (17563.0bp,98.163bp) and (17563.0bp,82.548bp)  .. (140);
  \draw (17634.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (18514.0bp,291.66bp) and (12413.0bp,225.93bp)  .. (3);
  \draw (15979.0bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (10) ..controls (34309.0bp,185.16bp) and (34309.0bp,169.55bp)  .. (42);
  \draw (34385.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (44) ..controls (35934.0bp,118.73bp) and (36048.0bp,109.61bp)  .. (36145.0bp,92.0bp) .. controls (36190.0bp,83.893bp) and (36240.0bp,70.26bp)  .. (242);
  \draw (36294.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (26) ..controls (16186.0bp,125.61bp) and (16356.0bp,119.22bp)  .. (16412.0bp,92.0bp) .. controls (16424.0bp,86.384bp) and (16434.0bp,76.573bp)  .. (134);
  \draw (16492.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (16) ..controls (6009.2bp,121.27bp) and (6287.4bp,111.5bp)  .. (6383.2bp,92.0bp) .. controls (6416.2bp,85.282bp) and (6451.9bp,72.424bp)  .. (80);
  \draw (6450.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (16) ..controls (5506.1bp,123.2bp) and (5017.1bp,114.16bp)  .. (4946.2bp,92.0bp) .. controls (4927.8bp,86.266bp) and (4909.6bp,75.137bp)  .. (71);
  \draw (5027.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (15) ..controls (4231.6bp,119.42bp) and (4081.5bp,109.83bp)  .. (4028.2bp,92.0bp) .. controls (4009.6bp,85.782bp) and (3990.8bp,74.58bp)  .. (66);
  \draw (4099.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (34) ..controls (26172.0bp,114.59bp) and (26103.0bp,105.07bp)  .. (26077.0bp,92.0bp) .. controls (26065.0bp,86.012bp) and (26055.0bp,76.199bp)  .. (186);
  \draw (26143.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (1) ..controls (2983.7bp,190.67bp) and (2628.5bp,154.25bp)  .. (12);
  \draw (2923.2bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (12) ..controls (2262.8bp,118.04bp) and (2214.6bp,109.27bp)  .. (2174.2bp,92.0bp) .. controls (2159.1bp,85.558bp) and (2144.3bp,75.069bp)  .. (56);
  \draw (2235.2bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (40) ..controls (32467.0bp,98.163bp) and (32467.0bp,82.548bp)  .. (221);
  \draw (32543.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (32) ..controls (23278.0bp,121.15bp) and (23474.0bp,112.03bp)  .. (23542.0bp,92.0bp) .. controls (23564.0bp,85.73bp) and (23586.0bp,74.112bp)  .. (173);
  \draw (23592.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (31) ..controls (21401.0bp,118.39bp) and (21512.0bp,108.95bp)  .. (21551.0bp,92.0bp) .. controls (21565.0bp,86.054bp) and (21578.0bp,75.795bp)  .. (162);
  \draw (21590.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (33) ..controls (24980.0bp,116.52bp) and (24914.0bp,107.66bp)  .. (24857.0bp,92.0bp) .. controls (24831.0bp,84.628bp) and (24802.0bp,72.633bp)  .. (179);
  \draw (24928.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (37) ..controls (29682.0bp,120.21bp) and (29802.0bp,111.46bp)  .. (29904.0bp,92.0bp) .. controls (29945.0bp,84.212bp) and (29990.0bp,70.776bp)  .. (208);
  \draw (30036.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (27) ..controls (17499.0bp,97.451bp) and (17455.0bp,76.913bp)  .. (139);
  \draw (17498.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (27) ..controls (17729.0bp,122.71bp) and (17851.0bp,114.74bp)  .. (17892.0bp,92.0bp) .. controls (17902.0bp,86.408bp) and (17911.0bp,77.041bp)  .. (142);
  \draw (17925.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (11) ..controls (1004.5bp,98.163bp) and (1004.1bp,82.548bp)  .. (50);
  \draw (1076.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (31) ..controls (21077.0bp,122.94bp) and (20954.0bp,115.06bp)  .. (20913.0bp,92.0bp) .. controls (20903.0bp,86.401bp) and (20895.0bp,77.032bp)  .. (158);
  \draw (20995.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (26) ..controls (15946.0bp,107.25bp) and (15935.0bp,100.55bp)  .. (15927.0bp,92.0bp) .. controls (15921.0bp,85.48bp) and (15917.0bp,76.888bp)  .. (131);
  \draw (15998.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (29) ..controls (19627.0bp,124.08bp) and (19547.0bp,115.84bp)  .. (19481.0bp,92.0bp) .. controls (19464.0bp,85.639bp) and (19446.0bp,74.679bp)  .. (150);
  \draw (19542.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (16) ..controls (5956.2bp,125.46bp) and (6093.6bp,118.44bp)  .. (6209.2bp,92.0bp) .. controls (6239.5bp,85.065bp) and (6272.2bp,72.483bp)  .. (79);
  \draw (6318.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (19) ..controls (9284.5bp,98.163bp) and (9284.1bp,82.548bp)  .. (95);
  \draw (9355.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (28) ..controls (18890.0bp,107.25bp) and (18879.0bp,100.55bp)  .. (18871.0bp,92.0bp) .. controls (18865.0bp,85.48bp) and (18861.0bp,76.888bp)  .. (147);
  \draw (18942.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (37) ..controls (29396.0bp,116.52bp) and (29330.0bp,107.66bp)  .. (29273.0bp,92.0bp) .. controls (29247.0bp,84.628bp) and (29218.0bp,72.633bp)  .. (203);
  \draw (29344.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (39) ..controls (31562.0bp,107.84bp) and (31532.0bp,99.997bp)  .. (31506.0bp,92.0bp) .. controls (31475.0bp,82.408bp) and (31440.0bp,70.298bp)  .. (215);
  \draw (31522.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (37) ..controls (29266.0bp,125.14bp) and (28870.0bp,118.94bp)  .. (28737.0bp,92.0bp) .. controls (28706.0bp,85.566bp) and (28672.0bp,72.787bp)  .. (200);
  \draw (28798.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (12) ..controls (2523.8bp,118.67bp) and (2602.4bp,109.74bp)  .. (2629.2bp,92.0bp) .. controls (2637.8bp,86.27bp) and (2644.4bp,77.121bp)  .. (59);
  \draw (2659.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (38) ..controls (30909.0bp,114.44bp) and (30936.0bp,106.03bp)  .. (30957.0bp,92.0bp) .. controls (30967.0bp,85.923bp) and (30974.0bp,76.751bp)  .. (213);
  \draw (31034.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (27) ..controls (17661.0bp,114.44bp) and (17688.0bp,106.03bp)  .. (17709.0bp,92.0bp) .. controls (17719.0bp,85.923bp) and (17726.0bp,76.751bp)  .. (141);
  \draw (17807.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (20189.0bp,291.08bp) and (25445.0bp,226.54bp)  .. (8);
  \draw (23338.0bp,258.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (42) ..controls (34407.0bp,114.44bp) and (34433.0bp,106.02bp)  .. (34454.0bp,92.0bp) .. controls (34463.0bp,85.915bp) and (34471.0bp,76.741bp)  .. (232);
  \draw (34541.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (3) ..controls (11264.0bp,190.68bp) and (10910.0bp,154.3bp)  .. (20);
  \draw (11183.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (4) ..controls (15017.0bp,193.83bp) and (15672.0bp,150.6bp)  .. (26);
  \draw (15498.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (19) ..controls (9118.3bp,123.92bp) and (8984.9bp,116.34bp)  .. (8872.2bp,92.0bp) .. controls (8839.0bp,84.841bp) and (8803.0bp,71.973bp)  .. (92);
  \draw (8943.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (34) ..controls (26355.0bp,107.03bp) and (26365.0bp,100.36bp)  .. (26373.0bp,92.0bp) .. controls (26379.0bp,85.554bp) and (26384.0bp,77.115bp)  .. (188);
  \draw (26449.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (15) ..controls (4353.6bp,107.25bp) and (4343.1bp,100.55bp)  .. (4335.2bp,92.0bp) .. controls (4329.1bp,85.48bp) and (4324.8bp,76.888bp)  .. (68);
  \draw (4406.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (15) ..controls (4475.0bp,106.68bp) and (4500.4bp,99.009bp)  .. (4523.2bp,92.0bp) .. controls (4558.2bp,81.245bp) and (4597.5bp,68.978bp)  .. (70);
  \draw (4653.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (16) ..controls (5629.5bp,118.39bp) and (5518.8bp,108.95bp)  .. (5479.2bp,92.0bp) .. controls (5465.3bp,86.054bp) and (5452.1bp,75.795bp)  .. (74);
  \draw (5550.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (17) ..controls (7226.5bp,116.15bp) and (7330.0bp,106.08bp)  .. (7368.2bp,92.0bp) .. controls (7385.3bp,85.675bp) and (7402.4bp,74.718bp)  .. (85);
  \draw (7468.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (1) ..controls (2787.2bp,197.66bp) and (1457.8bp,146.44bp)  .. (11);
  \draw (2307.2bp,171.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (20) ..controls (10766.0bp,112.21bp) and (10806.0bp,103.88bp)  .. (10840.0bp,92.0bp) .. controls (10862.0bp,84.484bp) and (10885.0bp,73.057bp)  .. (104);
  \draw (10944.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (44) ..controls (35681.0bp,114.44bp) and (35655.0bp,106.03bp)  .. (35633.0bp,92.0bp) .. controls (35624.0bp,85.923bp) and (35616.0bp,76.751bp)  .. (238);
  \draw (35704.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (33) ..controls (24850.0bp,125.14bp) and (24454.0bp,118.94bp)  .. (24321.0bp,92.0bp) .. controls (24290.0bp,85.566bp) and (24256.0bp,72.787bp)  .. (176);
  \draw (24382.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (28) ..controls (19010.0bp,106.63bp) and (19036.0bp,98.961bp)  .. (19058.0bp,92.0bp) .. controls (19093.0bp,81.173bp) and (19133.0bp,68.905bp)  .. (149);
  \draw (19184.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (25) ..controls (14537.0bp,118.77bp) and (14390.0bp,108.88bp)  .. (14337.0bp,92.0bp) .. controls (14317.0bp,85.688bp) and (14297.0bp,74.332bp)  .. (122);
  \draw (14398.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (11) ..controls (1237.0bp,125.83bp) and (1542.0bp,120.24bp)  .. (1644.2bp,92.0bp) .. controls (1666.1bp,85.955bp) and (1688.6bp,74.348bp)  .. (54);
  \draw (1695.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (25) ..controls (14763.0bp,107.03bp) and (14773.0bp,100.36bp)  .. (14781.0bp,92.0bp) .. controls (14787.0bp,85.554bp) and (14792.0bp,77.115bp)  .. (125);
  \draw (14862.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (42) ..controls (34197.0bp,113.58bp) and (34155.0bp,104.8bp)  .. (34140.0bp,92.0bp) .. controls (34134.0bp,86.13bp) and (34129.0bp,77.562bp)  .. (230);
  \draw (34222.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (16) ..controls (5555.9bp,125.48bp) and (5251.4bp,119.55bp)  .. (5149.2bp,92.0bp) .. controls (5126.5bp,85.876bp) and (5102.9bp,74.068bp)  .. (72);
  \draw (5220.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (19) ..controls (9025.9bp,126.13bp) and (8621.4bp,121.11bp)  .. (8487.2bp,92.0bp) .. controls (8458.0bp,85.659bp) and (8426.7bp,73.029bp)  .. (90);
  \draw (8548.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (3) ..controls (11491.0bp,185.16bp) and (11491.0bp,169.55bp)  .. (21);
  \draw (11567.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (26) ..controls (16130.0bp,114.59bp) and (16199.0bp,105.07bp)  .. (16225.0bp,92.0bp) .. controls (16237.0bp,86.012bp) and (16248.0bp,76.199bp)  .. (133);
  \draw (16327.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (32) ..controls (23333.0bp,122.4bp) and (23706.0bp,113.05bp)  .. (23832.0bp,92.0bp) .. controls (23874.0bp,85.063bp) and (23919.0bp,71.375bp)  .. (175);
  \draw (23976.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (29) ..controls (19879.0bp,111.93bp) and (19933.0bp,103.1bp)  .. (19979.0bp,92.0bp) .. controls (20016.0bp,83.201bp) and (20057.0bp,70.394bp)  .. (154);
  \draw (20103.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (35) ..controls (27091.0bp,107.03bp) and (27101.0bp,100.36bp)  .. (27109.0bp,92.0bp) .. controls (27115.0bp,85.554bp) and (27120.0bp,77.115bp)  .. (192);
  \draw (27190.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (29) ..controls (19770.0bp,100.22bp) and (19770.0bp,87.936bp)  .. (19770.0bp,77.0bp) .. controls (19770.0bp,74.448bp) and (19770.0bp,71.791bp)  .. (152);
  \draw (19852.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (0) ..controls (19508.0bp,273.2bp) and (19629.0bp,246.81bp)  .. (6);
  \draw (19669.0bp,258.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (24) ..controls (13526.0bp,112.21bp) and (13566.0bp,103.88bp)  .. (13600.0bp,92.0bp) .. controls (13622.0bp,84.484bp) and (13645.0bp,73.057bp)  .. (119);
  \draw (13704.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (8) ..controls (26012.0bp,193.29bp) and (25419.0bp,151.19bp)  .. (33);
  \draw (25852.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (21) ..controls (11491.0bp,98.163bp) and (11491.0bp,82.548bp)  .. (107);
  \draw (11562.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (15) ..controls (4458.8bp,107.03bp) and (4469.2bp,100.36bp)  .. (4477.2bp,92.0bp) .. controls (4483.3bp,85.554bp) and (4487.9bp,77.115bp)  .. (69);
  \draw (4503.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (44) ..controls (35779.0bp,98.163bp) and (35779.0bp,82.548bp)  .. (239);
  \draw (35850.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (25) ..controls (14844.0bp,115.11bp) and (14923.0bp,105.82bp)  .. (14992.0bp,92.0bp) .. controls (15035.0bp,83.436bp) and (15082.0bp,70.024bp)  .. (127);
  \draw (15133.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (35581.0bp,123.06bp) and (35376.0bp,115.18bp)  .. (35306.0bp,92.0bp) .. controls (35288.0bp,86.013bp) and (35270.0bp,74.976bp)  .. (236);
  \draw (35367.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (17) ..controls (6939.8bp,119.8bp) and (6866.9bp,111.12bp)  .. (6805.2bp,92.0bp) .. controls (6783.0bp,85.136bp) and (6759.8bp,73.527bp)  .. (81);
  \draw (6886.7bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (18) ..controls (7967.7bp,118.09bp) and (8077.7bp,108.54bp)  .. (8117.2bp,92.0bp) .. controls (8131.6bp,85.987bp) and (8145.3bp,75.584bp)  .. (89);
  \draw (8208.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [black,->] (31) ..controls (21498.0bp,124.21bp) and (21888.0bp,116.9bp)  .. (22018.0bp,92.0bp) .. controls (22053.0bp,85.413bp) and (22090.0bp,72.334bp)  .. (165);
  \draw (22141.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (44) ..controls (36034.0bp,124.21bp) and (36424.0bp,116.9bp)  .. (36554.0bp,92.0bp) .. controls (36589.0bp,85.413bp) and (36626.0bp,72.334bp)  .. (244);
  \draw (36677.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (9) ..controls (30506.0bp,193.83bp) and (29850.0bp,150.6bp)  .. (37);
  \draw (30320.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (0) ..controls (18731.0bp,289.16bp) and (15405.0bp,228.63bp)  .. (4);
  \draw (17386.0bp,258.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (5) ..controls (18668.0bp,194.65bp) and (17913.0bp,149.78bp)  .. (27);
  \draw (18375.0bp,171.5bp) node {$1$};
  \draw [black,->] (42) ..controls (34197.0bp,112.3bp) and (34143.0bp,103.45bp)  .. (34095.0bp,92.0bp) .. controls (34059.0bp,83.354bp) and (34020.0bp,70.604bp)  .. (229);
  \draw (34111.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (19) ..controls (9442.4bp,118.48bp) and (9552.7bp,109.06bp)  .. (9592.2bp,92.0bp) .. controls (9605.9bp,86.053bp) and (9618.9bp,75.794bp)  .. (97);
  \draw (9676.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (31) ..controls (21341.0bp,114.44bp) and (21368.0bp,106.03bp)  .. (21389.0bp,92.0bp) .. controls (21399.0bp,85.923bp) and (21406.0bp,76.751bp)  .. (161);
  \draw (21476.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (31) ..controls (21145.0bp,114.44bp) and (21119.0bp,106.03bp)  .. (21097.0bp,92.0bp) .. controls (21088.0bp,85.923bp) and (21080.0bp,76.751bp)  .. (159);
  \draw (21168.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (25) ..controls (14842.0bp,114.59bp) and (14911.0bp,105.07bp)  .. (14937.0bp,92.0bp) .. controls (14949.0bp,86.012bp) and (14960.0bp,76.199bp)  .. (126);
  \draw (14973.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (11) ..controls (1199.7bp,121.19bp) and (1395.0bp,112.12bp)  .. (1463.2bp,92.0bp) .. controls (1484.4bp,85.733bp) and (1506.3bp,74.114bp)  .. (53);
  \draw (1569.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (11) ..controls (1163.0bp,118.64bp) and (1273.7bp,109.28bp)  .. (1313.2bp,92.0bp) .. controls (1326.8bp,86.057bp) and (1339.5bp,75.798bp)  .. (52);
  \draw (1398.2bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (44) ..controls (35877.0bp,114.44bp) and (35904.0bp,106.03bp)  .. (35925.0bp,92.0bp) .. controls (35935.0bp,85.923bp) and (35942.0bp,76.751bp)  .. (240);
  \draw (36012.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (28) ..controls (18767.0bp,119.71bp) and (18616.0bp,110.24bp)  .. (18562.0bp,92.0bp) .. controls (18544.0bp,85.852bp) and (18526.0bp,74.802bp)  .. (145);
  \draw (18633.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (32) ..controls (22855.0bp,123.82bp) and (22559.0bp,116.31bp)  .. (22459.0bp,92.0bp) .. controls (22433.0bp,85.704bp) and (22406.0bp,73.573bp)  .. (166);
  \draw (22520.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (20) ..controls (10696.0bp,97.517bp) and (10715.0bp,79.93bp)  .. (103);
  \draw (10775.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (6) ..controls (19769.0bp,185.16bp) and (19769.0bp,169.55bp)  .. (29);
  \draw (19845.0bp,171.5bp) node {$(1-\mu_{1})(1-\mu_{2})$};
  \draw [black,->] (16) ..controls (5689.3bp,114.44bp) and (5662.7bp,106.03bp)  .. (5641.2bp,92.0bp) .. controls (5631.9bp,85.923bp) and (5624.0bp,76.751bp)  .. (75);
  \draw (5712.2bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (37) ..controls (29355.0bp,125.15bp) and (29218.0bp,118.01bp)  .. (29103.0bp,92.0bp) .. controls (29072.0bp,85.044bp) and (29039.0bp,72.462bp)  .. (202);
  \draw (29164.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (27) ..controls (17436.0bp,116.52bp) and (17370.0bp,107.66bp)  .. (17313.0bp,92.0bp) .. controls (17287.0bp,84.628bp) and (17258.0bp,72.633bp)  .. (138);
  \draw (17384.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (27) ..controls (17306.0bp,125.14bp) and (16910.0bp,118.94bp)  .. (16777.0bp,92.0bp) .. controls (16746.0bp,85.566bp) and (16712.0bp,72.787bp)  .. (135);
  \draw (16838.0bp,84.5bp) node {$\mu_{1}\mu_{2}/4$};
  \draw [black,->] (41) ..controls (33346.0bp,115.21bp) and (33275.0bp,105.78bp)  .. (33249.0bp,92.0bp) .. controls (33238.0bp,86.082bp) and (33228.0bp,76.407bp)  .. (225);
  \draw (33320.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/2$};
  \draw [black,->] (39) ..controls (31770.0bp,114.59bp) and (31839.0bp,105.07bp)  .. (31865.0bp,92.0bp) .. controls (31877.0bp,86.012bp) and (31888.0bp,76.199bp)  .. (218);
  \draw (31956.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (20) ..controls (10632.0bp,97.517bp) and (10613.0bp,79.93bp)  .. (102);
  \draw (10639.0bp,84.5bp) node {$1/2$};
  \draw [black,->] (8) ..controls (26303.0bp,185.16bp) and (26303.0bp,169.55bp)  .. (34);
  \draw (26359.0bp,171.5bp) node {$\mu_{1}\mu_{2}$};
  \draw [black,->] (3) ..controls (11067.0bp,197.66bp) and (9737.8bp,146.44bp)  .. (19);
  \draw (10586.0bp,171.5bp) node {$\mu_{2}(1-\mu_{1})$};
  \draw [black,->] (32) ..controls (23083.0bp,98.163bp) and (23083.0bp,82.548bp)  .. (170);
  \draw (23099.0bp,84.5bp) node {$1/4$};
  \draw [black,->] (21) ..controls (11589.0bp,114.44bp) and (11616.0bp,106.03bp)  .. (11637.0bp,92.0bp) .. controls (11647.0bp,85.923bp) and (11654.0bp,76.751bp)  .. (108);
  \draw (11714.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (37) ..controls (29297.0bp,122.9bp) and (29007.0bp,114.56bp)  .. (28908.0bp,92.0bp) .. controls (28880.0bp,85.556bp) and (28850.0bp,73.136bp)  .. (201);
  \draw (28979.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (37) ..controls (29621.0bp,114.44bp) and (29648.0bp,106.03bp)  .. (29669.0bp,92.0bp) .. controls (29679.0bp,85.923bp) and (29686.0bp,76.751bp)  .. (206);
  \draw (29767.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})/4$};
  \draw [black,->] (29) ..controls (19671.0bp,114.7bp) and (19644.0bp,106.23bp)  .. (19623.0bp,92.0bp) .. controls (19614.0bp,85.921bp) and (19607.0bp,76.748bp)  .. (151);
  \draw (19694.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (33) ..controls (24881.0bp,122.9bp) and (24591.0bp,114.56bp)  .. (24492.0bp,92.0bp) .. controls (24464.0bp,85.556bp) and (24434.0bp,73.136bp)  .. (177);
  \draw (24563.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (16) ..controls (5787.2bp,98.163bp) and (5787.2bp,82.548bp)  .. (76);
  \draw (5858.2bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (26) ..controls (16051.0bp,107.03bp) and (16061.0bp,100.36bp)  .. (16069.0bp,92.0bp) .. controls (16075.0bp,85.554bp) and (16080.0bp,77.115bp)  .. (132);
  \draw (16150.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/2$};
  \draw [black,->] (0) ..controls (20024.0bp,297.73bp) and (23288.0bp,281.04bp)  .. (25947.0bp,266.0bp) .. controls (29355.0bp,246.73bp) and (33544.0bp,220.76bp)  .. (10);
  \draw (28546.0bp,258.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (19) ..controls (9516.6bp,125.56bp) and (9820.3bp,119.71bp)  .. (9922.2bp,92.0bp) .. controls (9944.6bp,85.906bp) and (9967.8bp,74.178bp)  .. (99);
  \draw (9973.7bp,84.5bp) node {$1/4$};
  \draw [black,->] (32) ..controls (23147.0bp,97.451bp) and (23192.0bp,76.913bp)  .. (171);
  \draw (23257.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})/4$};
  \draw [black,->] (32) ..controls (22887.0bp,121.9bp) and (22687.0bp,113.27bp)  .. (22618.0bp,92.0bp) .. controls (22598.0bp,85.864bp) and (22578.0bp,74.52bp)  .. (167);
  \draw (22689.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (40) ..controls (32624.0bp,117.91bp) and (32733.0bp,108.31bp)  .. (32772.0bp,92.0bp) .. controls (32787.0bp,85.982bp) and (32801.0bp,75.578bp)  .. (223);
  \draw (32862.0bp,84.5bp) node {$\mu_{1}(1-\mu_{2})$};
  \draw [red,->] (30) ..controls (20268.0bp,105.71bp) and (20277.0bp,99.171bp)  .. (20285.0bp,92.0bp) .. controls (20293.0bp,84.831bp) and (20301.0bp,75.826bp)  .. (155);
  \draw (20309.0bp,84.5bp) node {$1$};
  \draw [black,->] (28) ..controls (18824.0bp,117.89bp) and (18776.0bp,109.13bp)  .. (18735.0bp,92.0bp) .. controls (18720.0bp,85.55bp) and (18705.0bp,75.06bp)  .. (146);
  \draw (18796.0bp,84.5bp) node {$\mu_{1}\mu_{2}/2$};
  \draw [black,->] (13) ..controls (3147.2bp,97.451bp) and (3102.8bp,76.913bp)  .. (61);
  \draw (3146.7bp,84.5bp) node {$1/2$};
  \draw [black,->] (27) ..controls (17722.0bp,120.21bp) and (17842.0bp,111.46bp)  .. (17944.0bp,92.0bp) .. controls (17985.0bp,84.212bp) and (18030.0bp,70.776bp)  .. (143);
  \draw (18076.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})/4$};
  \draw [black,->] (19) ..controls (9173.2bp,112.3bp) and (9118.8bp,103.45bp)  .. (9071.2bp,92.0bp) .. controls (9035.3bp,83.354bp) and (8995.6bp,70.604bp)  .. (93);
  \draw (9086.7bp,84.5bp) node {$1/4$};
%
\end{tikzpicture}

### 2019-03-25 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing__leafs__onlyfig.gv.tex
Je suis bourrin mais je git TOUT pour éviter une catastrophe...

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (293.0bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish UCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (0) at (238.1bp,128.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (19) at (125.1bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (68) at (352.1bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \draw [black,->] (0) ..controls (150.85bp,113.55bp) and (136.0bp,105.2bp)  .. (126.1bp,92.0bp) .. controls (121.31bp,85.627bp) and (119.83bp,77.397bp)  .. (19);
  \draw (228.1bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1} - 1)^{2}/2$};
  \draw [black,->] (0) ..controls (306.21bp,109.06bp) and (319.69bp,101.89bp)  .. (330.1bp,92.0bp) .. controls (336.72bp,85.708bp) and (341.47bp,77.073bp)  .. (68);
  \draw (442.1bp,84.5bp) node {$\mu_{1}^{2}(\mu_{2} - 1)^{2}/2$};
%
\end{tikzpicture}

### 2019-03-25 3-Appendices/2-Appendix/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.git/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.tex
Je suis bourrin mais je git TOUT pour éviter une catastrophe...

% \documentclass[anon,12pt]{colt2018} % Anonymized submission
\documentclass[12pt]{colt2018} % Include author names
% http://www.learningtheory.org/colt2018/call.html#instructions

% The following packages will be automatically loaded:
% amsmath, amssymb, natbib, graphicx, url, algorithm2e

\title[What Doubling Tricks Can and Cannot Do for Multi-Armed Bandits]{What Doubling Tricks Can and Cannot Do for Multi-Armed Bandits}

\usepackage{macrosArticle}
\usepackage{macrosText}
\usepackage{times}
\usepackage{array}
\usepackage{float}

\coltauthor{\Name{Lilian Besson} \Email{{Lilian}{.}{Besson}{@}{CentraleSupelec}{.}{fr}} \\
    \addr CentraleSup\'elec (campus of Rennes), IETR, SCEE Team,\\
    Avenue de la Boulaie -- CS $47601$, F-$35576$ Cesson-S\'evign\'e, France
    \AND
    \Name{Emilie Kaufmann} \Email{{Emilie}{.}{Kaufmann}{@}{Univ}{-}{Lille}{.}{fr}} \\
    \addr CNRS \& Universit\'e de Lille, Inria SequeL team\\
    UMR $9189$ -- CRIStAL,  F-$59000$ Lille, France
}

\begin{document}

\maketitle

% -----------------------------------------------------------------
\begin{abstract}
    An online reinforcement learning algorithm is \emph{anytime} if it does not need to know in advance the horizon $T$ of the experiment.
    A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick''.
    In the context of adversarial or stochastic multi-armed bandits,
    the performance of an algorithm is measured by its regret,
    and we study two families of sequences of growing horizons (geometric and exponential)
    to generalize previously known results that certain doubling tricks can be used to conserve certain regret bounds.
    In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log T)$.
    We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log T)$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
    %
    % Numerical experiments also show the efficiency of this approach
    % when the appropriate doubling trick is used with
    % efficient non-anytime algorithms,
    % like the Approximated Finite-Horizons Gittins Index (\AFHG) algorithm by \cite{Lattimore16a} or \KLUCBpp{} by \cite{Menard17}.

    % DEADLINE
    % - Paper submission deadline: February 16, 2018, 11:00 PM EST
    % \verb+http://www.learningtheory.org/colt2018/call.html#instructions+
\end{abstract}

% -----------------------------------------------------------------
\begin{keywords}
    Multi-Armed Bandits;
    % Reinforcement Learning;
    % Online Learning;
    Anytime Algorithms;
    Sequential Learning;
    Doubling Trick.
\end{keywords}

% \iffalse  % XXX
\FIXME{
    Remove Table of Contents when submitting.
    Maybe bring back table of content for the HAL version?
}
\tableofcontents
% \fi % XXX

\TODO{Maths:}
\begin{enumerate}
    % \item Write a code that plots, for a function $f$,
    % \[\frac{\sum_{i=0}^{L_T} (T_i - T_{i-1})^{\gamma} \log(T_i - T_{i-1})^{\delta}}{T^{\gamma} (\log(T))^{\delta}}\]
    % (and the same without the $T_{i-1}$), just to have an idea how bad is our initial upper bound
    \item Is a function for which it is bounded for $(\gamma,\delta)=(1/2,0)$ and $(\gamma,\delta)=(0,1)$ then prove it mathematically!
    \item Extend the lower bound at the end of this document for all $\alpha \in (0,1]$
    \item Include also the previous geometric lower bound in this statement
\end{enumerate}

\TODO{Writing:}
\begin{itemize}
    \item Change introduction,
\end{itemize}
% -----------------------------------------------------------------

% -----------------------------------------------------------------
\section{Introduction}
% -----------------------------------------------------------------

\todo[inline]{begin todo}
    \begin{itemize}
        \item
        Change the motivation:
        understanding why those doubling tricks are used, and what they can preserve...
        \item
        Remove the idea of problem dependent + worse-case optimal
        \item
        Update the literature review:
        Talk about the \KLUCBswitch{} algorithm from \cite{Garivier18}.
    \end{itemize}
\todo[inline]{end todo}

Multi-Armed Bandit (MAB) problems are well-studied sequential decision making problems in which an agent repeatedly chooses an action (the ``arm'' of a one-armed bandit) in order to maximize some total reward \citep{Robbins52,LaiRobbins85}.
Initial motivation for their study came from the modeling of clinical trials, as early as 1933 with the seminal work of \cite{Thompson33}.
In this example, arms correspond to different treatments with unknown, random effect.
Since then, MAB models have been proved useful for many more applications, that range from
cognitive radio \citep{Jouini09} to
online content optimization (\eg, news article recommendation \citep{Li10}, online advertising \citep{LiChapelle11}, A/B Testing \citep{Kaufmann14,Jamieson17ABTest}),
or portfolio optimization \citep{Sani2012risk}.

While the number of patients involved in a clinical study (and thus the number of treatments to select) is often decided in advance, in other contexts the total number of decisions to make (the horizon $T$) is unknown.
It may correspond to the total number of visitors of a website optimizing its displays for a certain period of time, or to the number of attempted communications in a smart radio device.
In such cases, it is thus crucial to devise \emph{anytime algorithms}, that is algorithms that do no rely on the knowledge of this horizon $T$ to sequentially select arms.
A general way to turn any base algorithm into an anytime algorithm is the use of the so-called Doubling Trick, first proposed by \cite{auer1995gambling}, that consists in repeatedly running the base algorithm with increasing horizons.
Motivated by the frequent use of this technique and the absence of a generic study of its effect on the algorithm's efficiency,
this paper investigates in details two families of doubling sequences (geometric and exponential), and shows that the former should be avoided for stochastic problems.

More formally, a MAB model is a set of $K$ arms, each arm $k$ being associated to a (unknown) \emph{reward stream} $(Y_{k,t})_{t\in\N}$.
Fix $T$ a finite (possibly unknown) horizon.
At each time step $t \in \{1,\dots,T\}$ an agent selects an arm $A(t) \in \{1,\dots,K\}$ and receives as a reward the current value of the associated reward stream, $r(t) := Y_{A(t), t}$.
The agent's decision strategy (or \emph{bandit algorithm})
$\cA_T := (A(t), t\in \{1,\dots,T\})$ is such that $A(t)$ can only rely on the past observations
% $\cF_{t-1} = (A(1),r(1),\dots,A(t-1),r(t-1))$,
$A(1),r(1),\dots,A(t-1),r(t-1)$,
on external randomness and (possibly) on the knowledge of the horizon $T$.
The objective of the agent is to find an algorithm $\cA$ that maximizes the expected cumulated rewards, where the expectation is taken over the possible randomness used by the algorithm and the possible randomness in the generation of the rewards stream.
In the oblivious case, in which the reward streams are independent of the algorithm's choice, this is equivalent to minimizing the \emph{regret}, defined as
\begin{equation}\label{eq:regret}
    R_T(\cA_T) := \max_{k \in \{1,\dots,K\}}
    % \mu^* T - \E_{\mu}\left[\sum_{t=1}^{T} r(t)\right]
    \E\left[\sum_{t=1}^{T} \left(Y_{k,t} - Y_{A(t),t}\right)\right].
    % = \mu^* T - \sum_{k=1}^K \mu_k \E_{\mu}\left[ T_k(T) \right], \
\end{equation}

This quantity, referred to as \emph{pseudo-regret} in \cite{Bubeck12}, quantifies the difference between the expected cumulated reward of the best fixed action, and that of the strategy $\cA_T$.
For the general adversarial bandit problem \citep{Auer02NonStochastic}, in which the rewards streams are arbitrary (picked by an adversary), a \emph{worst-case} lower bound has been given.
It says that for every algorithm, there exists (stochastic) reward streams such that the regret is larger than $(1/20)\sqrt{KT}$
\citep{Auer02NonStochastic}.
% \citep[Th.5.1]{Auer02NonStochastic}.
Besides, the \ExpThree{} algorithm has been shown to have a regret of order $\sqrt{KT\log(K)}$.

Much smaller regret may be obtained in \emph{stochastic} MAB models, in which the reward stream from each arm $k$ is assumed to be \iid, from some (unknown) distribution $\nu_k$, with mean $\mu_k$.
In that case, various algorithms have been proposed with \emph{problem-dependent} regret upper bounds of the form $C(\bm\nu) \log(T),$ where $C(\bm\nu)$ is a constant that only depend on the arms distributions.
Different assumptions on the arms distributions lead to different problem-dependent constants.
In particular, under some parametric assumptions (\eg, Gaussian distributions, exponential families), \emph{asymptotically optimal} algorithms have been proposed and analyzed (\eg, \KLUCB{} \citep{KLUCBJournal} or Thompson sampling \citep{AgrawalGoyal11,Kaufmann12Thompson}), for which the constant $C(\bm\nu)$ obtained in the regret upper bound matches exactly that of the lower bound given by \cite{LaiRobbins85}.
Under the non-parametric assumption that the $\nu_k$ are bounded in $[0,1]$, the regret of the \UCBone{} algorithm \citep{Auer02} is of the above form with $C(\bm\nu) = 8 \times \sum_{k : \mu_k > \mu^*} (\mu^* - \mu_k)^{-1}$, where $\mu^* = \max_k \mu_k$ is the mean of the best arm.
%\UCBone{} is referred to as \emph{order-optimal} as its regret has, up to a multiplicative constant, the right dependency in the gaps.
Like in this last example, all the available constants $C(\bm\nu)$ become very large on ``hard'' instances, in which some arms are very close to the best arm.
On such instances,  $C(\bm\nu)\log(T)$ may be much larger than the worst-case $(1/20)\sqrt{KT}$, and distribution-independent guarantees may actually be preferred.

The \MOSS{} algorithm, proposed by \cite{Audibert2009minimax}, is the first stochastic bandit algorithm
to enjoy a problem-dependent logarithmic regret
and to be optimal in a \emph{minimax} sense, as its regret is proved to be upper bounded by $\sqrt{KT}$, for bandit models with rewards in $[0,1]$.
However the corresponding constant $C(\bm \nu)$ is proportional to $K/\Delta_{\min}$, where $\Delta_{\min} = \min_{k} (\mu^*-\mu_k)$ is the minimal gap, which worsen the constant of \UCBone{}.
Another drawback of \MOSS{} is that it is \emph{not} anytime.
%
These two shortcoming have been overcame recently in two different works.
On one hand, the \MOSSAnytime{} algorithm \citep{Degenne16} is minimax optimal and anytime, but its problem-dependent regret does not improve that of \MOSS.
On the other hand, the \KLUCBpp{} algorithm \citep{Menard17} is simultaneously minimax optimal and asymptotically optimal (\ie, it has the best problem-dependent constant $C(\bm\nu)$), but it is not anytime. A natural question is thus to know whether a Doubling Trick could overcome this limitation.

This question is the starting point of our comprehensive study of the Doubling Trick: can a single Doubling Trick be used to preserve both problem-dependent (logarithmic) regret and minimax (square-root) regret?
We answer this question partially, by showing that two different types of Doubling Trick may actually be needed.
In this paper, we investigate how algorithms enjoying regret guarantees of the generic form
\begin{equation}\label{eq:genericWithLogT}
    \forall T \geq 1,\;\;\;\; R_T(\cA_T) \leq c \; T^{\gamma} (\log T)^{\delta} + o( T^{\gamma} \left(\log T)^{\delta} \right)
\end{equation}
may be turned into an anytime algorithm enjoying \emph{similar} regret guarantees with an appropriate Doubling Trick.
%
This does not come for free, and we exhibit a ``price of Doubling Trick'', that is a constant factor larger than $1$, referred to as a \emph{constant multiplicative overhead}.

% Outline of this article
%\paragraph{Outline}
The rest of the paper is organized as follows.
The Doubling Trick is formally defined in Section~\ref{sec:genericAlgo}, along with a generic tool for its analysis.
In Section~\ref{sec:geometricDoublingTrick}, we present upper and lower bounds on the regret of algorithms to which a geometric Doubling Trick is applied.
Section~\ref{sec:exponentialDoublingTrick} investigates regret guarantees that can be obtained for a ``faster'' exponential Doubling Trick.
Experimental results are then reported in Section~\ref{sec:experiments}.
Complementary elements of proofs are deferred to the appendix.
% We give missing proofs and additional results in Appendix~\ref{sec:missingproofs}, an heuristic $\DTnr$ improving $\DTr$ in App.~\ref{sec:DTnr} along with more experimental results in App.~\ref{sec:otherExperiments}, and other technical tools are given in App.~\ref{sec:otherproofs}.

% -----------------------------------------------------------------
\section{Doubling Tricks }\label{sec:genericAlgo}

The Doubling Trick, denoted by \DT, is a general procedure to convert a (possibly non-anytime) algorithm into an anytime algorithm.
It is formally stated below as Algorithm~\ref{algo:DTr} and depends on a non-decreasing diverging \emph{doubling sequence} $(T_i)_{i\in \N}$
(\ie, $T_i\to\infty$ for $i\to\infty$).
% such that $T_i$ goes to infinity.
\DT{} fully restarts the underlying algorithm $\cA$ at the beginning of each new sequence (at $t = T_i + 1$), and run this algorithm on a sequence of length ($T_i - T_{i-1}$).
% If the algorithm $\cA$ needs to know the horizon, $\cA_{T=\square}$ denotes $\cA$ initialized with the value $\square$ for its parameter $T$,
% and if $\cA$ does not need the horizon, $\cA_{T=\square}$ is simply a new instance of $\cA$.

% \FIXME{Remove box ?}

%
%It is immediate to check that for any $t$, the decisions taken by $\cA' = \DTr(\cA, (T_i)_{i\in\N})$, $A'(t)$, are also measurable regarding the previous observations.

%
%It is a very generic\footnote{It is well known to not even be restricted to the setting of Multi-Armed Bandit.} algorithm that can be applied to any algorithm $A$.
%

% - Full restart is very generic and can work for most algorithms.
% - Non-restart only works for specific cases that only use the horizon $T$ as just a numerical parameter.

% \subsection{Doubling Trick Algorithm with Full Restart}

% Let $A' := \DTr(A, (T_i)_{i\in\N})$ denotes the following Algorithm~\ref{algo:DTr}.

\vspace*{5pt}  % XXX remove if problem
% \centering
% Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
% Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
% \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
\begin{algorithm}[H]
        % XXX Options
        \LinesNumbered  % XXX Option to number the line
        \DontPrintSemicolon
        % \RestyleAlgo{boxed}
        % XXX Input, data and output
        \KwIn{Bandit algorithm $\cA$, and doubling sequence $(T_i)_{i\in\N}$.}
        % \KwData{Data}
        % \KwResult{Result}
        \BlankLine
        % XXX Algorithm
        Let $i = 0$, and initialize algorithm $\cA^{(0)} = \cA_{T_0}$.\;
        \For{$t = 1, \dots, T-1$}{
            \If(\tcp*[f]{Next horizon $T_{i+1}$ from the sequence}){
                $t > T_i$
            }{
                Let $i = i + 1$.\;
                Initialize algorithm $\cA^{(i)} = \cA_{T_i - T_{i-1}}$.
                \tcp*[f]{Full restart}
            }
            Play with $\cA^{(i)}$: play arm $A'(t) := A^{(i)}(t - T_{i-1})$, observe reward $r(t) = Y_{A'(t), t}$. \nllabel{line:internalTimer_DTr}
        }
        \caption{The Generic Doubling Trick Algorithm, $\cA' = \DTr(\cA, (T_i)_{i\in\N})$.}
\label{algo:DTr}
\end{algorithm}
% \vspace*{-5pt}  % XXX remove if problem
\vspace*{5pt}  % XXX remove if problem

% With examples of an algorithm $A$ that could \textbf{not} be used for the full-restart case ?
% ``Explore-then-Commit'' algorithms by~\cite{Garivier2016ETC}, or the DMED algorithm by~\cite{Honda10}. Others ?
%
% With examples of an algorithm $A$ that could \textbf{not} be used for the non-restarting case?
% ``Explore-then-Commit'' algorithms by~\cite{Garivier2016ETC}, or the DMED algorithm by~\cite{Honda10}. Others ?

\paragraph{Related work.}
The Doubling Trick is a well known idea in online learning, that can be traced back to \cite{auer1995gambling}.
In the literature, the term Doubling Trick usually refers to the geometric sequence $T_i = 2^i$, in which the horizon is actually \emph{doubling}, that was popularized by
\cite{CesaLugosi06} in the context of adversarial bandits. Specific doubling tricks have also been used for stochastic bandits, for example in the work of \cite{Auer10}, which uses
the doubling sequence $T_i = 2^{2^{i}}$ to turn the \UCBR{} algorithm into an anytime algorithm.

\paragraph{Elements of regret analysis.}
% \subsection{Growing Sequences $(T_i)_{i\in\N}$ of Successive Horizons}\label{sub:growingSequences}
For a sequence $(T_i)_{i\in\N}$, with $T_i\in\N$ for all $i$, we denote $T_{-1} = 0$,
and $T_0$ is always taken non-zero, $T_0 > 0$ (\ie, $T_0\in\N^*$).
We only consider \emph{non-decreasing} and \emph{diverging} sequences
(that is, $\forall i,\; T_{i+1} \geq T_i$, and $T_i \to \infty$ for $i \to \infty$).

% We study two doubling sequence and give basic results for them.
% The first sequence is denoted as the ``geometric growth'' of parameters $T_0, b$,
% and the second sequence is denoted as the ``exponential growth'' of parameters $T_0, a, b$.

\begin{definition}[Last Term $L_T$]\label{def:lastTerm}

    For a non-decreasing diverging sequence $(T_i)_{i\in\N}$
    and $T\in\N$,
    we can define $L_T\bigl((T_i)_{i\in\N}\bigr)$ by
    \begin{equation}
        \forall T\geq1,\;\;\;\;
        L_T\bigl((T_i)_{i\in\N}\bigr) := \min\left\{ i \in \N : T_i > T \right\}.
    \end{equation}

    It is simply denoted $L_T$ when there is no ambiguity (\eg, if the doubling sequence is chosen).
\end{definition}
% \begin{proof}\label{proof:lastTerm}
%     As $T_i \to \infty$ for $i \to \infty$,
%     the minimum exists for any $T\in\N$.
%     %
%     As the sequence is non-decreasing,
%     $L_T \geq 0$ is also the only integer $k\in\N$ satisfying
%     $T_{k - 1} < T \leq T_{k}$
%     (as $T_{-1} = 0$).
% \end{proof}

% \subsection{Decomposing the Regret on the Sub-Sequences}

$\DTr(\cA)$ reinitializes its underlying algorithm $\cA$ at each time $T_i$,
and in generality the total regret is upper bounded by the regret on
each sequence $\{T_i,\dots, T_{i+1} - 1\}$.
By considering the last partial sequence $\{T_{L_T-1}, \dots, T - 1\}$, this splitting can be used to get
% both a lower bound \eqref{eq:decompositionIneq_LB} by excluding the last sequence,
a generic upper bound~\eqref{eq:decompositionIneq_UB} by taking into account a larger last sequence (up to $T_{L_T} - 1$).
And for stochastic bandit models, the \iid{} hypothesis on the rewards streams makes the splitting on each sequence an equality, so we can also get the lower bound \eqref{eq:decompositionIneq_LB} by excluding the last partial sequence.
Lemma~\ref{lem:decomposition} is
% easy from the \iid{} hypothesis, and is
proved in Appendix~\ref{sub:proof_decompositionDTr}.

% \begin{lemma}[Regret Decomposition for $\DTr$]\label{lem:decomposition}
%     For a fixed (but unknown) horizon $T$ and a sequence $(T_i)_{i\in\N}$,
%     the regret of a sequence-based algorithm satisfies this identity \eqref{eq:decompositionEq} and the inequalities \eqref{eq:decompositionIneq},
%     \vspace*{-5pt}  % XXX remove if problem
%     \begin{align}
%         R_T(\DTr(A, (T_i)_{i\in\N}))
%         =
%         \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}})
%         + R_{T - T_{L_T-1}}(A_{T = T_{L_T} - T_{L_T-1}}) \label{eq:decompositionEq} \\
%         \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}})
%         \underset{LB}{\leq}
%         R_T(\DTr(A, (T_i)_{i\in\N}))
%         \underset{UP}{\leq}
%         \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}). \label{eq:decompositionIneq}
%     \end{align}
% \end{lemma}

% \FIXME{A new version of this result. Ajust the explanation above and the proof}
% \FIXME{Exemple of the alternative notations are in the result}

\begin{lemma}[Regret Lower and Upper Bounds for $\DTr$]\label{lem:decomposition}

    For \emph{any bandit model} and algorithm $\cA$ and horizon $T$, one has the generic upper bound
    \vspace*{-5pt}  % XXX remove if problem
    \begin{align}
        R_T(\DTr(\cA, (T_i)_{i\in\N}))
        \leq \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(\cA_{T_i - T_{i-1}}). \tag{LB}\label{eq:decompositionIneq_LB}
    \end{align}

    Under a \emph{stochastic bandit model}, one has furthermore the lower bound
    \begin{align}
        R_T(\DTr(\cA, (T_i)_{i\in\N}))
        \geq \sum_{i=0}^{L_{T-1}} R_{T_i - T_{i-1}}(\cA_{T_i - T_{i-1}}). \tag{UB}\label{eq:decompositionIneq_UB}
    \end{align}
\end{lemma}

As expected, the key to obtain regret guarantees for a Doubling Trick algorithm is to carefully choose the doubling sequence $(T_i)_{i\in\N}$.
%
Empirically, one can verify that sequences with slow growth gives terrible results, and for example using an arithmetic progression typically gives a linear regret.
% XXX in fact for any algorithm that samples each arm once in its $K$ first steps, any arithmetic progression (\ie, $T_i - T_{i-1}$ is bounded) induces a linear regret.
%
Building on this result, we will prove that if $\cA$ satisfies a certain regret bound ($R_T = \cO(T^{\gamma})$, $\cO((\log T)^{\delta})$, or $\cO(T^{\gamma} (\log T)^{\delta})$) then an appropriate anytime version of $\cA$ with a certain doubling trick can conserve the regret bound, with an explicit constant multiplicative overhead $\ell > 1$.
%
In this paper, we study in depth two families of sequences: first geometric and then exponential growths.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{A Generic Analysis of the Regret of a Doubling-Trick Anytime Algorithm}
\TODO{Finish this (Lilian) :
    OK for presentation,
    OK for upper-bound of sum by last term,
    TODO for last term bound by term in $T$
}

In this first Section,
we start by giving an overview of our technique to conserve regret upper bounds for doubling-trick algorithms.
We present a generic way of defining doubling sequences, and two examples to find back the two cases commonly used in the literature (\ie, geometric and exponential doubling sequences).
Then we explain how to bound the regret of a doubling-trick anytime algorithm, for any sequence of a certain parametric form, to obtain a unified analysis of our upper bounds.

% -----------------------------------------------------------------
\subsection{Overview of our proof}

We propose a generic analysis of a Doubling Trick based algorithm.
Assume $\cA_T$ has bounded regret in the form $R_T(\cA_T) \leq c T^{\gamma} (\log T)^{\delta}$.
First note that the regret of the anytime algorithm $\DT(\cA)$ is upper bounded by
$c\sum_{i=0}^{L_T} (T_i)^{\gamma} (\log(T_i))^{\delta}$ (by Lemma~\ref{lem:decomposition}).

\begin{itemize}
    \item
    The first requirement for a good regret bound is to find a suitable choice of sequence $(T_i)_i$,
    such that the last term in this sum, \ie, $(T_{L_T})^{\gamma} (\log(T_{L_T}))^{\delta}$, is of order $T^{\gamma} \log(T)^{\delta}$, up-to some multiplicative constants.
    % % XXX Remove this itemize
    % \begin{itemize}
    %     \item $f(t) = \log(t)$: true for any values,
    %     \item $f(t) = c t$: true for $\gamma = 0$, \emph{almost} true for any values,
    %     \item $f(t) = t^\alpha$: true for $\gamma =0$, \emph{almost} true for any values. (check!)
    % \end{itemize}

    \item
    The second step of our analysis is to use that the doubling sequence grows \emph{fast enough} so that the sum is indeed dominated by its last term.
    % % XXX Remove this itemize
    % \begin{itemize}
    %     \item $f(t) = \log(t)$: true for $\gamma > 0$,
    %     \item $f(t) = c t$: true for any values,
    %     \item $f(t) = t^\alpha$: true for $\gamma > 0$ (CONJ: does not work for $\gamma=0$)
    % \end{itemize}
\end{itemize}

\paragraph{Lower bounds}

A geometric doubling trick \emph{cannot} preserve a logarithmic regret and neither can the faster sequences defined with $f(i) = i^\alpha$ for any $\alpha \in (0,1)$.
This shows that we \textbf{need} a sequence growing at least as fast as the one for $f(i) = i$, \ie, an exponential sequence.

\TODO{Last claim only proved for $\delta=1$ and $f(t)=\sqrt{t}$}
\TODO{Can we find \emph{another} function $f$ that work? worst-case check numerically the second step}
\TODO{How about this form, with $\beta<0$ ? $f(i) = i^{\alpha} \left(\log(i)\right)^{\beta}$}
\TODOE{given that $i^{\alpha} \left(\log(i)\right)^{\beta}$ is squeezed between $i^\alpha$ and $i^{\alpha+\epsilon}$ I don't expect a different behavior}

% -----------------------------------------------------------------
\subsection{A generic doubling sequence}

Let us consider an increasing function $f: \N \to \N$, and a constant $c>0$.
We define an increasing\footnote{~Note that if $(T_i)$ is not increasing, but diverging, one can always define a new sequence $(T_i')$ that is also increasing and grows faster than $(T_i)$ with this trick: $\forall j, T_{j+1}' := T_{\min\{i > j, T_i > T_j \}}$. The sequence $(T_i')$ always satisfies $T_i' > T_{i-1}'$, and it is the slowest sequence that satisfies this and is faster than $(T_i)$.}
% FIXME remove this footnote? We added hypothesis such that it is no longer required
and diverging sequence $(T_i)_{i\in\N}$ by
$\forall i\in\N, T_i := \lfloor c_0 \exp(c \exp(f(i))) \rfloor$.
The function $f$ has to grow fast enough so that $T_i - T_{i-1}$ increases and diverges, and so that $(T_i)_i$ is increasing.

\paragraph{Examples}
Possible functions $f$ include:
\begin{itemize}
    \item
    The function $f(i) = \log(i)$ gives a geometric doubling,
    with $c_0 = T_0$, $b = \exp(c)$,
    $$\forall i\in\N, \; T_i = \lfloor c_0 \exp(c \exp(f(i))) \rfloor = \lfloor T_0 b^i \rfloor.$$
    \item
    The function $f(i) = \alpha i$ gives an exponential doubling,
    with $c_0 = \frac{T_0}{\exp(c)}$, $a = \exp(c)$, $b = \exp(\alpha)$,
    $$\forall i\in\N, \; T_i = \lfloor c_0 \exp(c \exp(f(i))) \rfloor = \lfloor T_0 a^{b^i} \rfloor.$$
    \item
    And choosing a more generic form $f(i) = c_2 i^d (\log i)^e$, for $c_2,d>0$ and $e\in\R$, which is increasing from $t \geq \exp(\frac{-e}{d})$ (so increasing from a certain value, which is enough for us).
    As soon as $d>0$, any value of $e$ gives a doubling sequence growing \emph{faster than any geometric sequences}, and as soon as $d<1$, it gives a doubling sequence growing \emph{slower than any exponential sequences}.
    As such, this generic form is the best candidate to unify the previously known results about both the geometric sequences used to conserve $\bigO{\sqrt{T}}$ regret bounds, and the exponential sequences used to conserve $\bigO{\log(T)}$ regret bounds.
\end{itemize}

% -----------------------------------------------------------------
\subsection{Unified analysis of the upper-bound}

Let $\gamma,\delta\in\R$, both non-negative but not zero simultaneously.
Fix the horizon $T$ and let $\cA_T$ denote a non-anytime algorithm.
Assume that it satisfies
\[R_T(\cA_T) \leq g_1(T) + c_1 T^{\gamma} (\log T)^{\delta}\]
for some constant $c_1 > 0$ and $g_1(t) = \smallO{T^{\gamma} (\log T)^{\delta}}$ (uniformly for any $T$).

We consider $\DT(\cA)$ the doubling-trick anytime version of $\cA_T$,
for which Lemma~\ref{lem:decomposition} gives that
its regret is upper bounded by this sum,
where $g_2(t) = \smallO{T^{\gamma} (\log T)^{\delta}}$
thanks to Lemma~\ref{lem:ControlLemmaSumOfSmallO}.

\begin{equation}
    R_T(\DT(\cA)) \leq g_2(T) +  c_1 \sum_{i=0}^{L_T} (T_i - T_{i-1})^{\gamma} (\log(T_i - T_{i-1}))^{\delta}
\end{equation}

Our first trick is to directly bound $T_i - T_{i-1} \leq T_i$, and it could seem to be a really crude bound.
But as soon as $f$ is increasing, $T_i - T_{i-1}$ is diverging, and for any $f$ of the form we consider, $T_i - T_{i-1}$ is actually equivalent to $T_i$.
We give in Appendix~\ref{appendix:equivalence_Ti_and_TimTim1} a result to justify that this is not suboptimal, and as it eases the proofs we prefer to simply the presentation as much as possible.

\TODOE{so far I proved the equivalence only for the exponential doubling, is it true in general?}

\begin{equation}
    R_T(\DT(\cA)) \leq g_2(T) +  c_1 \sum_{i=0}^{L_T} T_i^{\gamma} (\log(T_i))^{\delta}.
\end{equation}

If $f$ is increasing only from a constant value $i_{\min} \geq 0$, we can consider the first terms as constants and include them in the function $g_2$.
So we can suppose that $f$ is increasing from a certain rank $i_{\min}$, and assume $T$ to be large enough so that $L_T - 1 \geq i_{\min}$ (as $L_T \to+\infty$ when $T\to+\infty$).
For instance, $f(i) = c_2 i^d (\log i)^e$ is increasing from $i \geq i_{\min} = \lceil \exp(\frac{-e}{d})\rceil$.

We focus on this sum from $i=i_{\min}$ to $i=L_T$, \ie, $\sum_{i=i_{\min}}^{L_T} T_i^{\gamma} (\log(T_i))^{\delta}$.
The second trick is to show that we can bound the sum from $i_{\min}$ to $L_T-1$ as a constant times the last term for $i=L_T$, so that the whole sum is of the order of its last term.
We have the next lemma and prove it below.

\begin{lemma}\label{lem:unifiedUpperBound}
    \textbf{(Case 1)}
    For any $\gamma > 0,\delta$ and any choice of $f(i) = c_2 i^d (\log i)^e$ ($c_2>0$, $d\geq0$, $e\in\R$),
    there exists a constant $C$ such that the following inequality holds,
    for any $T$ large enough,
    \begin{equation}\label{eq:unifiedUpperBound1}
        \sum_{i=i_{\min}}^{L_T - 1} T_i^{\gamma} (\log(T_i))^{\delta}
        \leq C \; \times \; T_{L_T}^{\gamma} (\log(T_{L_T}))^{\delta}.
    \end{equation}

    \textbf{(Case 2)}
    If $\gamma = 0$ and $d > 0$,
    % a more careful analysis is needed,
    there exists $g(T) = \smallO{(\log(T_{L_T}))^{\delta}}$ such that
    \begin{equation}\label{eq:unifiedUpperBound2}
        \sum_{i=i_{\min}}^{L_T - 1} (\log(T_i))^{\delta}
        \leq g(T) + C \; \times \; (\log(T_{L_T}))^{\delta}.
    \end{equation}

    \textbf{(Case 3)}
    Finally, if $\gamma = 0$ and $d = 0$, \textcolor{red}{sadly we cannot obtain anything else but}
    \begin{equation}\label{eq:unifiedUpperBound3}
        \sum_{i=i_{\min}}^{L_T - 1} (\log(T_i))^{\delta}
        \leq C \; \times \; (\log(T_{L_T}))^{\textcolor{red}{1+}\delta}.
    \end{equation}
\end{lemma}

\TODOE{Not so bad, we don't need the $\gamma = 0$ and $d=0$, we have a lower bound in that case}

\hr{}

\begin{proof}
    \textbf{(Case 3)}
    First of all, let us tackle the easiest case when $\gamma=0$ and $d = 0$, \ie, the sequence is a geometric one.
    We have a lower-bound that shows that $R_T(\cA_T) = \Theta((\log T)^{\delta})$ is lost by a geometric doubling trick, as $R_T(\DT(\cA)) = \Theta((\log T)^{1+\delta})$ (see Theorem~\ref{thm:LowerBoundGeom_log}).
    We can quickly show our upper-bound easily in this case,
    as $\sum_{i=i_{\min}}^{L_T} (\log T_i)^{\delta} \leq L_T (\log T_{L_T})^{\delta}$,
    $L_T = \bigO{\log T}$ and so $\log T_{L_T} = \bigO{\log T}$ too.
    So $\sum_{i=i_{\min}}^{L_T} (\log T_i)^{\delta} = \bigO{(\log T)^{1+\delta}}$.

    \hr{}
    \textbf{(Case 2)}
    If $\gamma=0$ but for other sequences (\ie, $d>0$), a more careful analysis is needed.
    % Actually I think it is quite easy.
    In a nutshell, we have
    $T_i = \bigO{a^{b^{i^d}}}$ if $d>0$ and so $\log T_i = \bigO{b^{i^d}}$,
    for which by a sum-integral and then a change of variable we can bound the sum by what we want.
    % (it is a generalization of what we know to do correctly for bounding the geometric sum by a constant times its last term).

    For $i \leq L_T - 1$, we can bound $f(i) \leq c_2 (\log L_T)^e i^d$,
    so if we call $c_3(e,T) := c_2 (\log L_T)^e$, we have
    $\log T_i \leq \text{constant} + c \exp(f(i))$.
    Using Lemma~\ref{lem:expoInequality} and $\delta<1$, it gives,
    $(\log T_i)^{\delta} \leq \text{constant} + c^{\delta} \exp(\delta f(i)) \leq \text{constant} + c^{\delta} \exp(\delta c_3(e,T) i^d)$.
    Thus when controlling the sum from $i=i_{\min}$ to $L_T - 1$,
    constant term will give a $\bigO{L_T}$, which itself can be neglected against the right hand side (\ie, is a $\smallO{}$ of the second term).

    Let focus on $\sum_{i=i_{\min}}^{L_T - 1} \exp(\alpha i^d)$ for $\alpha := \delta c_3(e,T) > 0$.
    By using a sum-integral comparison, we bound it by $\int_{i_{\min}}^{L_T} \exp(\alpha t^d) \dt$.
    The change of variable $u = t^d$ (of class $\cC^1$), and the observation that $u^{\frac{1-d}{d}} \leq 1$ gives directly that
    $\int_{i_{\min}}^{L_T} \exp(\alpha t^d) \dt \leq \frac{1}{\alpha d} \exp(\alpha L_T^d)$.

    Finally, with an appropriate $g(T) = \smallO{(\log T)^{\delta}}$,
    we obtain that
    $\sum_{i_{\min}}^{L_T - 1} (\log T_i)^{\delta} \leq g(T) + c^{\delta} \frac{1}{c_2 \delta d} \exp(\delta f(L_T)) \leq g(T) + \frac{1}{c_2 \delta d} (\log(T_{L_T}))^{\delta}$ as wanted.

    \hr{}
    \textbf{(Case 1)}
    Now, suppose $\gamma>0$. A quick trick we can use is to bound all $\log(T_i)$ terms by the last one $\log(T_{L_T-1})$ and factor them.
    We can thus focus on $\sum_{i=i_{\min}}^{L_T-1} T_i^{\gamma}$.
    Bounding $T_i^{\gamma}$ by $c_0 \exp(c \gamma \exp(f(i)))$,
    and using a sum-integral inequality, we obtain,
    \begin{align*}
        \sum_{i=i_{\min}}^{L_T - 1} T_i^{\gamma} (\log(T_i))^{\delta}
        & \leq
        (\log T_{L_T-1})^{\delta} c_0 \int_{i_{\min}}^{L_T} \exp(c \gamma \exp(f(t))) \dt.
    \end{align*}

    The goal is now to bound this integral.
    Let $c_3(e,T) := c_2 \max(i_{\min}^e,(\log L_T)^e) > 0$,
    and first observe that $f(t) \leq c_3(e,T) t^d$ for any $t\in[i_{\min}, L_T]$,
    as $t \mapsto (\log t)^e$ is monotonic, no matter if $e<0$, $e>0$ or $e=0$.
    \TODO{Everything is simpler if we restrict to $e\geq0$ I think... But we forget about the case of $f(i) = \frac{i}{\log i}$. Maybe it's simpler to say from the beginning that $d,e\geq0$ and $de > 0$ (\ie, not both can be $0$ simultaneously).}
    We can thus bound the integral by
    $\int_{i_{\min}}^{L_T} \exp(c \gamma \exp(c_3(e,T) t^d)) \dt$.
    The change of variable $u := \exp(c_3(e,T) t^d)$ is of class $\cC^1$, and
    has a Jacobian $\frac{\du}{\dt} = c_3(e,T) \times d \times t^{d-1} \times u$,
    and so $\dt = \frac{1}{d c_3(e,T)} \frac{(\log u)^{\frac{1-d}{d}}}{u} \du$ after simplification.
    Let $\kappa = \frac{1-d}{d}$.
    Let $h(u) = \frac{(\log u)^{\kappa}}{u}$ (well defined for $u > 1$).
    We can bound the integral, do this change of variable and obtain
    \begin{align*}
        \int_{i_{\min}}^{L_T} \exp(c \gamma \exp(f(t))) \dt
        & \leq \int_{i_{\min}}^{L_T} \exp(c \gamma \exp(c_3(e,T) t^d)) \dt \\
        & \leq \frac{1}{d c_3(e,T)} \int_{\exp(c_3(e,T) i_{\min}^d)}^{\exp(c_3(e,T) L_T^d)} \exp(c \gamma u) h(u) \du\\
        \intertext{
        We can do a simple study of $h$.
        It is of class $\cC^1$, and $h'(u)$ has the sign of $\frac{\kappa}{\log u} - 1$, so $h$ is increasing then decreasing, and attains its global maximum at $u = \exp(\kappa)$.
        So we are sure that $h(u)$ is bounded for $u\in[ \exp(c_3(e,T) i_{\min}^d), \exp(c_3e,(T) L_T^d) ]$.
        Its maximum on this interval is bounded by its global maximum, \ie,
        $h(\exp(\kappa)) = \frac{\kappa^{\kappa}}{\exp(\kappa)}$, a constant w.r.t. $T$.
        Let us call $c_4(d,e,T) := \frac{1}{d c_3(e,T)} \times \frac{(\frac{1-d}{d})^{(\frac{1-d}{d})}}{\exp((\frac{1-d}{d}))} > 0$.
        Thus we have
        }
        & \leq c_4(d,e,T) \int_{\exp(c_3(e,T) i_{\min}^d)}^{\exp(c_3(e,T) L_T^d)} \exp(c \gamma u) \du.\\
        \intertext{
            Let $c_5(c,d,e,\gamma,T) := \frac{1}{c \gamma} c_4(d,e,T)$,
            we can now be bound the integral and conclude.
        }
        \int_{i_{\min}}^{L_T} \exp(c \gamma \exp(f(t))) \dt
        & \leq c_5(c,d,e,\gamma,T) \times \Bigl[ \exp(c \gamma u) \Bigr]_{\exp(c_3(e,T) i_{\min}^d)}^{\exp(c_3(e,T) L_T^d)}\\
        & \leq c_5(c,d,e,\gamma,T) \exp(c \gamma \exp(c_3(e,T) L_T^d)).
    \end{align*}

    To finish, the initial sum is now bounded by
    \begin{align*}
        \sum_{i=i_{\min}}^{L_T - 1} T_i^{\gamma} (\log(T_i))^{\delta}
        & \leq
        (\log T_{L_T-1})^{\delta} c_0 c_5(c,d,e,\gamma,T) \exp(c \gamma \exp(c_3(e,T) L_T^d)).
    \end{align*}
    with three values
    $c_5(c,d,e,\gamma,T) := c_4(d,e,T) \frac{1}{c \gamma}$,
    $c_4(d,e,T) := \frac{1}{d c_3(e,T)} \times \frac{(\frac{1-d}{d})^{(\frac{1-d}{d})}}{\exp((\frac{1-d}{d}))}$,
    and
    $c_3(e,T) := c_2 \max(i_{\min}^e,(\log L_T)^e)$.

    \TODO{Just force $e>0$, so $c_3(e,T)$ is just $c_2 (\log L_T)^e$.}

    If $e>0$, then $c_3(e,T)$ is just $c_2 (\log L_T)^e$.
    No matter the value of $e$, we clearly have
    $c_3(e,T) \geq c_2$ and thus $c_4$ and $c_5$ are constant w.r.t. $T$.
    Moreover, remember that there is an integer part (\ie, $\lfloor \dots \rfloor$) in the definition of $T_i$,
    so $\exp(c \gamma \exp(c_2 L_T^d (\log L_T)^e)) \leq (T_{L_T} + 1)^{\gamma} \leq T_{L_T}^{\gamma} + 1$ (as $\gamma \leq 1$, see Lemma~\ref{lem:expoInequality}).
    Thus, have two constants $c_6>0$ and $C>0$ such that
    \begin{align*}
        \sum_{i=i_{\min}}^{L_T - 1} T_i^{\gamma} (\log(T_i))^{\delta}
        & \leq
        c_6 + C T_{L_T}^{\gamma} (\log T_{L_T})^{\delta}.
    \end{align*}

\end{proof}

\TODO{Now we have to do the last part: show that this term as function of $T_{L_T}$ is bounded by $T^{\gamma} (\log T)^{\delta}$ and conclude.}

Now to finish the proof, we need to show that this last term, $T_{L_T}^{\gamma} (\log T_{L_T})^{\delta}$, is always bounded by a constant times $T^{\gamma} (\log T)^{\delta}$,
for any of the considered doubling sequences.

\TODO{I cannot find a satisfactory way to write all this, in a clean and unified way!}

% -----------------------------------------------------------------
\section{What the Geometric Doubling Trick Can and Cannot Do}\label{sec:geometricDoublingTrick}

We define geometric doubling sequences, and prove that they can be used
to conserve bounds in $\bigO{T^{\gamma}(\log T)^{\delta}}$ with $\gamma>0$ but cannot be used to conserve bounds in $\bigO{(\log T)^{\delta}}$.

% \subsection{The Geometric Sequence of Growing Horizons}

\begin{definition}[Geometric Growth]\label{def:geomSequence}
    For $b\in\R$, $b>1$ and $T_0\in\N^*$,
    the geometric sequence $T_i = \lfloor T_0 b^i \rfloor, \forall i \in \N$
    is non-decreasing and diverging,
    and satisfies furthermore
    \vspace*{-10pt}  % XXX remove if problem
    \begin{align}
        \forall T < T_0,\;
        L_{T} = 0,
        \;\;\text{and}\;\;
        \forall T \geq T_0,\;
        L_T = \left\lceil \log_b\left( \frac{T}{T_0} \right) \right\rceil, \label{eq:geomSequence2} \\
        \forall i > 0,\;\;
        T_0 (b - 1) b^{i-1} - 1
        \leq T_i - T_{i-1}
        \leq 1 + T_0 (b - 1) b^{i-1}. \label{eq:geomSequence1}
    \end{align}
    Asymptotically for $i$ and $T\to\infty$,
    $T_i = \bigO{b^i}$ and
    $L_T \sim \log_b(T) = \bigO{\log T}$.
\end{definition}
% \begin{proof}\label{proof:geomSequence}
%     The sequence is non-decreasing as $b>1$,
%     however it can be non-increasing for small values of $i$, \eg, $T_0=1$ and $b=1.1$ gives $T_i=1$ for $i=0$ to $7$.
%     %
%     But $b^i\to\infty$ and so $\lfloor T_0 b^i \rfloor \to \infty$ as well.

%     If $T > T_0$, the formula~\eqref{eq:geomSequence2} for $L_T$ comes from its definition as the smallest $i$ such that $T \leq T_i$,
%     \begin{align*}
%         \forall i > 0,\;\;\;\;
%         T \leq T_i
%         &\Leftrightarrow T \leq \lfloor T_0 b^i \rfloor
%         \underset{\text{as } T\in\N}{\Leftrightarrow} T \leq T_0 b^i
%         \Leftrightarrow \frac{T}{T_0} \leq b^i \\
%         &\Leftrightarrow \log_b\left(\frac{T}{T_0}\right) \leq i
%         \underset{\text{as } i\in\N}{\Leftrightarrow} i \geq \left\lceil \log_b\left(\frac{T}{T_0}\right) \right\rceil.
%     \end{align*}

%     For the inequality~\eqref{eq:geomSequence1}, we have $T_0 b^i - 1\leq T_i \leq T_0 b^i = T_0 b b^{i-1}$,
%     and $T_0 b^{i-1} - 1 \leq T_{i-1} \leq T_0 b^{i-1}$ (as $\lfloor x \rfloor \geq x - 1$),
%     so clearly $T_0 (b - 1) b^{i-1} - 1 \leq T_i - T_{i-1} \leq 1 + T_0 (b - 1) b^{i-1}$.
% \end{proof}

\subsection{Conserving a Regret Upper Bound with Geometric Horizons}

A geometric doubling sequence allows to conserve a minimax bound (\ie, $R_T = \mathcal{O}(\sqrt{T})$).
It was suggested, for instance, in~\cite[Ex.2.9]{CesaLugosi06}.
%
We generalize this result in the following theorem, proved in Appendix~\ref{sub:proof_keep45}, by extending it from $\sqrt{T}$ bounds to bounds of the form $T^{\gamma} (\log T)^{\delta}$ for \emph{any} $0<\gamma<1$ and $\delta \geq 0$.
%
Note that no distinction is done on the case $\delta=0$ neither in the expression of the constant overhead, nor in the proof.

% \begin{theorem}\label{thm:keep4}
%     If an algorithm $A$ satisfies
%     $ R_T(A) \leq c T^{\gamma} + f(T)$,
%     for $0<\gamma<1$, $c > 0$, and an increasing function $f(t) = \smallO{t^{\gamma}}$ (at $t\to\infty$),
%     then the anytime version $A':=\DTr(A, (T_i)_{i\in\N})$ with the geometric sequence $(T_i)_{i\in\N}$ of parameters $T_0\in\N^*,b>1$ (\ie, $T_i = \lfloor T_0 b^i \rfloor$) satisfies
%     \begin{equation}\label{eq:keep4_1}
%         \forall T \geq 0,\;\;\;\;
%         R_T(A') \leq \ell(\gamma, b) \; c \; T^{\gamma} + g(T).
%     \end{equation}
%     with an increasing function $g(t) = \smallO{t^{\gamma}}$,
%     and a \emph{constant overhead} $\ell(\gamma, b) > 1$,
%     % depending only on $\gamma$ and $b$,
%     \begin{equation}\label{eq:keep5_2}
%         \ell(\gamma, b) := \frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}.
%     \end{equation}

%     For a fixed $\gamma$, minimizing $\ell(\gamma, b)$ does not always give a unique solution.
%     If $\gamma \gtrsim 0.384$, there is a unique solution $b^*(\gamma) > 1$, solution of $b^{\gamma+1} - 2b + 1 = 0$, with no explicit form if $\gamma$ is unknown.
%     % $\ell(\gamma, b) \geq 4$ with the minimum attained for $b^*(\gamma) = 2^{\frac{1}{\gamma}}$.
%     % \begin{equation}
%     %     R_T(\DTr(A, (T_0 2^{\frac{i}{\gamma}})_{i\in\N})) \leq 4 c T^{\gamma} + g(T).
%     % \end{equation}
% \end{theorem}

% This first result can be applied to bounds of the form $R_T = \mathcal{O}(\sqrt{T})$ (with $\gamma=\frac{1}{2}$),
% % that are satisfied by minimax optimal algorithms (like kl-UCB$^++$ by~\cite{Menard17}),
% and the next result generalizes to the case
% of bounds of the form $R_T = \bigO{\sqrt{T \log T}}$ (with $\gamma=\frac{1}{2}, \delta=\frac{1}{2}$).
% % that are satisfied by minimax sub-optimal algorithms (like AFGH by~\cite{Lattimore16a}).

% \FIXME{NO we do not ``have to be less subtle'', I can obtain a similar overhead for this second theorem, TODO write it!}
%
% Note that the constant overhead in Theorem~\ref{thm:keep45} is larger than the one in Theorem~\ref{thm:keep4} because we have to be less subtle in our control of the inequalities to handle the $\log(T_i)$ terms easily (we use $T_i - T_{i-1} \leq T_i$ instead of $\leq 1 + T_0 (b - 1) b^{i-1}$.
% %
% Theorem~\ref{thm:keep45} is proved in Appendix~\ref{sub:proof_keep45}.

\begin{theorem}\label{thm:keep45}
    If an algorithm $\cA$ satisfies
    $ R_T(\cA_T) \leq c \; T^{\gamma} (\log T)^{\delta} + f(T)$,
    for $0<\gamma<1$, $\delta \geq 0$
    and for $c > 0$, and an increasing function $f(t) = \smallO{t^{\gamma} (\log t)^{\delta}}$ (at $t\to\infty$),
    then the anytime version $\cA':=\DTr(\cA, (T_i)_{i\in\N})$ with the geometric sequence $(T_i)_{i\in\N}$ of parameters $T_0\in\N^*,b>1$ (\ie, $T_i = \lfloor T_0 b^i\rfloor$)
    \emph{with the condition} $T_0 (b-1) > 1$ if $\delta>0$,
    satisfies,
    \begin{equation}\label{eq:keep45_1}
        % \forall T \geq 1,\;\;\;\;
        R_T(\cA') \leq \ell(\gamma, \delta, T_0, b) \; c \; T^{\gamma} \left(\log T\right)^{\delta} + g(T),
    \end{equation}
    with a increasing function $g(t) = \smallO{t^{\gamma} (\log t)^{\delta}}$,
    and a \emph{constant overhead} $\ell(\gamma, \delta, T_0, b) > 1$,
    \begin{equation}\label{eq:keep45_2}
        \ell(\gamma, \delta, T_0, b) := \left(\frac{\log(T_0(b-1) + 1)}{\log(T_0(b-1))}\right)^{\delta} \; \times \; \frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}.
    \end{equation}
    % % Where $\log^(x) := \max(1, \log(x))$.
\end{theorem}

For a fixed $\gamma$ and $\delta$, minimizing $\ell(\gamma, \delta, T_0, b)$ does not always give a unique solution.
On one hand, if $\gamma \gtrsim 0.384$, there is a unique solution $b^*(\gamma) > 1$ minimizing the $\frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}$ term, solution of $b^{\gamma+1} - 2b + 1 = 0$, but without a closed form if $\gamma$ is unknown.
%
On the other hand, for any $\gamma$, the term depending on $\delta$ tends quickly to $1$ when $T_0$ increases.

\paragraph{Practical considerations.}
%
Empirically, when $\gamma$ and $\delta$ are fixed and known, there is no need to minimize $\ell$ jointly.
%
It can be minimized separately by first minimizing $\frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}$,
that is by solving $b^{\gamma+1} - 2b + 1 = 0$ numerically (\eg, with Newton's method), and then by taking $T_0$ large enough so that the other term is close enough to $1$.

For the usual case of $\gamma=1/2$ and $\delta=0$ (\ie, bounds in $\sqrt{T}$), the optimal choice of $b$ is $\frac{3+\sqrt{5}}{2}$ leading to $\ell \simeq 3.33$,
and the usual choice of $b=2$ gives $\ell \simeq 3.41$ (see Corollary~\ref{cor:keep4} in appendix).
%
Any large enough $T_0$ gives similar performance, and empirically $T_0 \gg K$ is preferred, as most algorithms explore each arm once in their first steps
(\eg, $T_0 = 200$ for $K=9$ for our experiments).

\subsection{A Regret Lower Bound with Geometric Horizons}\label{sec:lowerboundsGeom}

We observe that the constant overhead in Eq.~\eqref{eq:keep45_2} from the previous Theorem~\ref{thm:keep45} blows up when $\gamma$ goes to zero,
giving the intuition that no geometric doubling trick could be used to preserve a logarithmic bound (\ie, with $\gamma=0$, $\delta>0$).
This is confirmed by the lower bound given below.

\begin{theorem}\label{thm:LowerBoundGeom_log}
    If $\cA$ satisfies
    $ R_T(\cA_T) \geq c \; (\log T)^{\delta}$,
    for $c > 0$ and $\delta>0$,
    and under a certain \emph{stochastic model},
    % if $T_0 (b-1) > 1$
    then the anytime version $\cA':=\DTr(\cA, (T_i)_{i\in\N})$ with the geometric sequence $(T_i)_{i\in\N}$ of parameters $T_0\in\N^*,b>1$ (\ie, $T_i = \lfloor T_0 b^i \rfloor$) satisfies this lower bound for a certain constant $c' > 0$,
    under the same model,
    \begin{equation}\label{eq:LowerBoundGeom_log_1}
        \forall T \geq 1,
        L_T \geq 2 \implies
        R_T(\cA') \geq
        % c_1 +
        % c_2 (\log T)^{\delta} +
        c' \; \left(\log T\right)^{\delta+1}.
    \end{equation}
    % for a constant $c' > 0$.
    %
    This implies that $R_T(\cA') = \Omega((\log T)^{\delta+1})$, which proves that a geometric sequence \emph{cannot} be used to conserve a logarithmic regret bound.
\end{theorem}

Theorem~\ref{thm:LowerBoundGeom_log} implies that a geometric sequence \emph{cannot}\footnote{Note that we only pay one extra $\log(T)$ factor, which is bad if $\delta$ is small, but not so bad for large $\delta$.} be used to conserve a finite-horizon lower bound like $R_T(\cA_T) \geq c \log(T)$. A complementary lower bound, stated as Theorem~\ref{thm:LowerBoundGeom_sqrt} in Appendix~\ref{sub:LowerBoundGeom_sqrt}, shows that if the regret is lower bounded at finite horizon by $R_T(\cA_T) \geq c \sqrt{T}$, then a comparable lower bound for the Doubling Trick algorithm $\DTr(\cA)$, possibly with a larger constant.
%

This special case ($\delta=1$) is indeed the most interesting, as in the stochastic case the regret of any uniformly efficient algorithm is at least logarithmic (\cite{LaiRobbins85}), and efficient algorithms with logarithmic regret have been exhibited.
%as efficient algorithms for stationary bandits should match the lower bound from \cite{LaiRobbins85}:
%  gives that $\lim\inf \frac{R_T(\cA)}{\log(T)}$ is a constant for any uniformly efficient algorithm, and an order-optimal algorithm have $R_T(\cA) = \bigO{\log(T)}$ so it has a (small enough) constant $c$ such that $R_T(\cA_T) \geq c \log(T)$ (maybe from a certain rank).
%
%
% XXX J'aime pas cette phrase "was maybe known", on peut pas dire "was intuited" ?
% This result was maybe known for the special case $\delta=1$ by \cite{Auer10}, as they used an exponential doubling trick ($T_i = 2^{2^i}$), but as far as the authors know, it has never been explicitly stated before.
%
% We interpret this result as announced above:
If $R_T(\cA_T)/{\log T}$ is bounded, then using a geometric sequence in the doubling trick algorithm is a bad idea as it guarantees a blow up in the regret, that is $R_T(\DTr(\cA, (T_i)_{i\in\N})) = \Omega((\log T)^2)$.
%
This result is the reason we need to consider successive horizons growing faster than a geometric sequence (\ie, such that $\log(T_i) \gg i$),
like the exponential sequence, which is studied in Section~\ref{sec:exponentialDoublingTrick}.

\subsection{Proof of Theorem~\ref{thm:LowerBoundGeom_log}}\label{proof:LowerBoundGeom_log}

    Let $\cA' := \DTr(\cA, (T_i)_{i\in\N})$ and consider a fixed \emph{stochastic} bandit problem.
    The lower bound \eqref{eq:decompositionIneq_LB} from Lemma~\ref{lem:decomposition} gives
    % Using Lemma~\ref{lem:decomposition} we get a decomposition on the regret of $\cA$ for time $T$ as a sum of regrets on sequences of successive lengths $T_i - T_{i-1}$, without the last sequence to have a lower bound, and so if $L_T \geq 2$, we have
    \begin{align*}
        R_T(\cA')
        &\geq \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(\cA_{T = T_i - T_{i-1}}) \\
        \intertext{
            We bound $T_i - T_{i-1} \geq T_0 (b - 1) b^{i-1} - 1$ for any $i>0$, thanks to Definition~\ref{def:geomSequence},
            and we can use the hypothesis on $\cA$ for each regret term.
        }
        &\geq \sum_{i=0}^{L_T-1} c (\log(T_i - T_{i-1}))^{\delta}
        \geq c \sum_{i=1}^{L_T-1} \bigl(\log\left(T_0 (b-1) b^{i-1} - 1\right)\bigr)^{\delta} \\
        &= c \sum_{i=0}^{L_T-2} \bigl(\log\left(T_0 (b-1) b^i - 1\right)\bigr)^{\delta} \;\;\;\;\;\;(\text{with } i := i - 1)\\
        \intertext{
            Let $x_i := T_0 (b-1) b^i > 0$.
            If we have $T_0 (b-1) > 1$ (see below ($\clubsuit$) in Page~\pageref{eq:clubsuit} for a discussion on the other case),
            then Lemma~\ref{lem:logxm1_logx} (Eq.~\eqref{eq:logxm1_logx_2}) gives
            $\log(x_i - 1) \geq \frac{\log(T_0 (b-1) - 1)}{\log(T_0 (b-1))} \log(x_i)$ as $x_i > 1$.
            % Moreover, $x_i \geq T_0 (b-1)$ for all $i\geq0$, and so $-\frac{1}{x_i} \geq -\frac{1}{T_0 (b-1)}$,
            % which gives $\log\left(1 - \frac{1}{x_i}\right) \geq \log\left(1 - \frac{1}{T_0 (b-1)}\right)$.
            % We can distribute the power on $\delta$ by a geometric sum, and so
            For lower bounds, there is no need to handle the constants tightly, and we have $x_i \geq b^i$ by hypothesis, so let call this constant $c' = c \left(\frac{\log(T_0 (b-1) - 1)}{\log(T_0 (b-1))}\right)^{\delta} > 0$, and thus it simplifies to
        }
        &\geq c' \sum_{i=0}^{L_T-2} \left(\log(b^i)\right)^{\delta} \\
        \intertext{
            % For this sum, as $\delta>0$,
            A sum-integral minoration for the increasing function $t \mapsto t^{\delta}$ (as $\delta>0$)
            % Lemma~\ref{lem:sumIntegralIneq} (sum-integral minoration),
            gives $\sum_{i=0}^{L_T-2} \left(\log(b^i)\right)^{\delta} = (\log b)^{\delta} \sum_{i=1}^{L_T-2} i^{\delta} \geq (\log b)^{\delta} \int_{0}^{L_T-2} t^{\delta} \dt = \frac{(\log b)^{\delta}}{\delta+1} \left(L_T-2\right)^{\delta+1}$ (if $L_T \geq 2$), and so
        }
        R_T(\cA') &\geq c' \frac{(\log b)^{\delta}}{\delta+1} (L_T-2)^{\delta+1} \\
        \intertext{
            For the geometric sequence, we know that $L_T \geq \log_b\left(\frac{T}{T_0}\right) \geq \log_b(T)$,
            and $\log_b(T) - 2 \sim \log_b(T)$ at $T\to\infty$ so there exists a constant $0 < c'' < 1$ such that $L_T - 2 \geq c'' \log_b(T)$ for $T$ large enough ($\geq \frac{2}{b-1}$), and such that $L_T \geq 2$.
            And thus we just proved that there is a constant $c''' > 0$ such that
        }
        R_T(\cA') &\geq c''' (\log T)^{\delta+1} =: g(T).
    \end{align*}
    % Even if $c_2$ is not positive, this does not change the result, as in any case $g(T) = c_3 (\log T)^{\delta+1} + \smallO{(\log T)^{\delta+1}} = \bigO{(\log T)^{\delta+1}}$.
    So this proves that for $T$ large enough, $R_T(\cA') \geq g(T)$ with $g(T) = \bigO{(\log T)^{\delta+1}}$,
    and so $R_T(\cA') = \Omega((\log T)^{\delta+1})$,
    which also implies that $R_T(\cA')$ \emph{cannot} be a $\bigO{(\log T)^{\delta}}$.

    \label{eq:clubsuit}
    ($\clubsuit$) If we do not have the hypothesis $T_0(b-1) > 1$, the same proof could be done,
    by observing that from $i \geq i_0$ large enough, we have $x_i \geq b^{i-i_0}$ (as soon as $b^{i_0} \geq \frac{1}{T_0 (b-1)} > 0$, \ie, $i_0 \geq \lceil -\log_b(T_0 (b-1)) \rceil \geq 1$),
    and so the same arguments can be used, to obtain a sum from $i = i_0 + 1$ instead of from $i = 1$.
    For a fixed $i_0$, we also have $L_T - 2 - i_0 \geq c'' \log(T)$ for a (small enough) constant $c''$, and thus we obtain the same result.
    %
    \hfill{}
    $\blacksquare$

% -----------------------------------------------------------------
\section{What Can the Exponential Doubling Trick Do?}\label{sec:exponentialDoublingTrick}

We define exponential doubling sequences, and prove that they can be used to conserve bounds in $\bigO{(\log T)^{\delta}}$, unlike the previously studied geometric sequences.
%
Furthermore, we provide elements showing that they may also conserve bounds in $\bigO{T^{\gamma}}$ or $\bigO{T^{\gamma}(\log T)^{\delta}}$.

% \subsection{The Exponential Sequence of Growing Horizons}

\begin{definition}[Exponential Growth]\label{def:expSequence}
    For $a,b\in\R$, $a,b>1$, $T_0\in\N^*$,
    and $\tau := \frac{T_0}{a} \in \R$,
    % , \geq 0$,
    the exponential sequence
    $T_i := \lfloor \tau a^{b^{i}} \rfloor, \forall i\in\N$
    is non-decreasing and diverging,
    and satisfies furthermore
    \begin{equation}\label{eq:expSequence}
        \forall T < T_0,\;
        L_{T} = 0,
        \;\;\text{and}\;\;
        \forall T \geq T_0,\;
        L_T = \left\lceil \log_b\left( \log_a\left( \frac{T}{\tau}\right)\right) \right\rceil.
    \end{equation}
    % \vspace*{-15pt}  % XXX remove if problem
    Asymptotically for $i$ and $T\to\infty$,
    $T_i = \bigO{a^{b^i}}$ and
    $L_T \sim \log_b(\log_a(\frac{T}{\tau})) = \bigO{\log \log T}$.
\end{definition}
% \begin{proof}\label{proof:expSequence}
%     The constant $\tau$ is not uniquely defined, as we only impose $T_0 = \lfloor \tau a^{b^0} \rfloor = \lfloor \tau a \rfloor$,
%     so any $\tau$ in $[\frac{T_0}{a}, \frac{T_0 + 1}{a})$ is correct.
%     But choosing $\tau = \frac{T_0}{a}$ is the simplest choice.

%     We ask $b>1$ and $a>1$ so $\tau a^{b^{i+1}} > \tau a^{b^i}$, so $\forall i>0, T_{i+1} \geq T_i$ as $x\mapsto \lfloor x \rfloor$ is non-decreasing,
%     thus the sequence is non-decreasing
%     (even if it can be non-increasing, for small values of $i$).

%     If $T>T_0$, the formula~\ref{eq:expSequence} for $L_T$ comes from its definition as the smallest $i$ such that $T \leq T_i$,
%     \begin{align*}
%         T \leq T_i
%         & \Leftrightarrow T \leq \lfloor \tau a^{b^i} \rfloor
%         \underset{\text{as } T\in\N}{\Leftrightarrow} T \leq \tau a^{b^i}
%         \Leftrightarrow \frac{T}{\tau} \leq a^{b^i} \\
%         & \Leftrightarrow \log_b\left(\log_a\left(\frac{T}{\tau}\right)\right) \leq i
%         \Leftrightarrow i \geq \left\lceil \log_b\left(\log_a\left(\frac{T}{\tau}\right)\right) \right\rceil.
%     \end{align*}
%     This sequence is simpler to work with if $a=T_0$, \ie, $\tau=1$, in which case $T_i$ is simply $\lfloor T_0^{b^i} \rfloor$.
%     %  for all $i\in\N$.
% \end{proof}

\subsection{Conserving a Regret Upper Bound with Exponential Horizons}

An exponential doubling sequence allows to conserve a problem-dependent bound on regret (\ie, $R_T = \bigO{\log T}$). This was already used in particular cases by ~\cite{Auer10} and more recently by \cite{Liauetal2017}.
We generalize this result in the following theorem.

% (with $\gamma=0, \delta=1$).
% that are satisfied by minimax optimal algorithms (like kl-UCB$^++$ by~\cite{Menard17}),
% and the second one is looser but more general, for instance to algorithms with bounds of the form $R_T = \bigO{\sqrt{T \log T}}$ (with $\gamma=1/2, \delta=1/2$).
% that are satisfied by minimax sub-optimal algorithms (like AFGH by~\cite{Lattimore16a}).
% The constants from Theorems~\ref{thm:keep45} and \ref{thm:keep5} are similar: $\ell(\gamma, b) = \ell(\delta, b)$.

% \FIXME{Why should the overhead satisfy $\ell(\gamma, \delta, T_0, a, b) >1$ here? We have no condition on $a > T_0$ so choosing a large $T_0$ can reduce the constant as much as we want!}

\begin{theorem}\label{thm:keep52}
    If an algorithm $\cA$ satisfies
    $ R_T(\cA_T) \leq c \; T^{\gamma} (\log T)^{\delta}+ f(T)$,
    for $0\leq\gamma<1$, $\delta \geq 0$,
    and for $c > 0$, and an increasing function $f(t) = \smallO{t^{\gamma} (\log t)^{\delta}}$ (at $t\to\infty$),
    then the anytime version $\cA':=\DTr(\cA, (T_i)_{i\in\N})$ with the exponential sequence $(T_i)_{i\in\N}$ of parameters $T_0\in\N^*,a,b>1$ (\ie, $T_i = \lfloor \frac{T_0}{a} a^{b^i}\rfloor$),
    satisfies the following inequality,
    % at the condition that $\Ei\left(\gamma\log(a)\right) \geq 0$ if $\gamma>0$ (\ie, $a > \e^{x_0/\gamma}$ where $x_0 \simeq 0.373$, cf. Lemma~\ref{lem:SumInequalityUsingEi}).
    \begin{equation}\label{eq:keep52_1}
        % \forall T \geq 1,\;\;\;\;
        R_T(\cA') \leq \ell(\gamma, \delta, T_0, a, b) \; c \left(T^b\right)^{\gamma} \left(\log T\right)^{\delta} + g(T).
    \end{equation}
    with an increasing function $g(t) = \smallO{(t^b)^{\gamma} (\log t)^{\delta}}$,
    and a \emph{constant overhead} $\ell(\gamma, \delta, T_0, a, b) > 0$,
    \begin{equation}\label{eq:keep52_2}
        \ell(\gamma, \delta, T_0, a, b) :=
        \begin{cases}
            \left(\frac{a}{T_0}\right)^{(b-1)\gamma} \; \frac{b^{2\delta}}{b^{\delta} - 1} > 0
            & \text{ if } \delta > 0 \\
            1 + \frac{1}{(\log(a)) (\log(b^{\gamma}))} > 1
            & \text{ if } \delta = 0
        \end{cases}
    \end{equation}
\end{theorem}

This result first shows that an exponential doubling trick can preserve a logarithmic regret bound ($\bigO(\log T)$, which corresponds to $\gamma=0$ and $\delta=1$),
% This result can be applied
% to bounds of the form $R_T = \bigO{(\log T)^{\delta}}$ (\ie, with $\gamma=0$)
with a multiplicative constant overhead $\ell \geq 4$.
%
%
It can further be applied to bounds of the generic form $R_T = \bigO{T^{\gamma} (\log T)^{\delta}}$, but with a \emph{significant overhead} as $T^{\gamma}$ becomes $T^{b\gamma}$, additionally to the constant multiplicative overhead $\ell > 0$.
However, it is important to notice that for $\gamma>0$, the constant $\ell$ can be made arbitrarily small
% (by increasing $T_0$).
%
% The fact that the constant overhead is not lower bounded and can be as close to $0$ as we want
(with a large enough first step $T_0$).
This observation is encouraging,
and let the authors think that a tighter upper bound could be proved.
% Even though we are not (yet) able to match th $T^{\gamma}$, we believe that it is possible  for this very reason.
%
% generalize this result for bounds of the form $R_T = \bigO{T^{\gamma} (\log T)^{\delta}}$, and the best result the authors obtained is the following.

\begin{remark}\label{cor:Keep5}
%     In particular, order-optimal and optimal algorithms for problem-dependent bound
%     have $\gamma=0$ and $\delta=1$,
%     as well as $f(t)=0$,
%     in which case Theorem~\ref{thm:keep52} gives a simpler bound,
%     as for any $b>1$, $\ell(\gamma=0, b) = \frac{b^2}{b-1} \geq 4$, and so
%     \vspace*{-10pt}  % XXX remove if problem
%
An interesting particular case of Theorem~\ref{thm:keep52} is the following ($\gamma =0, \delta=1$ and $f(t) = 0$).
        \begin{equation}
        % \forall T \geq 1,\;\;\;\;
        R_T(\cA_T) \leq c \; \log(T)
        \implies
        R_T(\DTr(\cA, (\lfloor T_0^{b^i} \rfloor)_{i\in\N})) \leq \frac{b^2}{b-1} \; c \; \log(T).
    \end{equation}
    %
    In this upper bound, the optimal choice of $b$ is $b=2$, which yields a constant multiplicative overhead of $\ell(\gamma=0, b) = 4$. It can be observed that this overhead is twice smaller as the overhead of $8.0625$ obtained by~\cite[Sec.4]{Auer10}\footnote{In \cite{Auer10}, the authors obtained an overhead of $258/32 = 8.0625 \geq 8$, as the ratio between the constants for the $\log(T)$ terms, respectively $258$ in Th.4.1 and $32$ in Th.3.1.}.
    % Also note that their Th.4.1 is not as clean as simply $R_T \leq \ell \; c \log(T)$, and in particular it seems that selecting a small enough $\lambda$ requires to know the horizon and cannot be tuned with a doubling trick.
\end{remark}

% We observe that the constant overhead from Theorem~\ref{thm:keep52} blows up for $\delta=0$ (as $b^{\delta}-1 \to 0$),
% % \ie, $\ell(\delta, b) \to +\infty$ for $\delta\to0$,
% and we give below in Theorem~\ref{thm:LowerBoundExpo_sqrt} another result toward a more precise lower bound.

\subsection{A Regret Lower Bound with Exponential Horizons}\label{sec:lowerboundsExpo}

% The previous Section~\ref{sec:geometricDoublingTrick} concluded on a clear take home message: geometric doubling tricks cannot conserve logarithmic regret guarantees.
% Proving or disproving the converse result for exponential doubling tricks and minimax regret bounds is not yet possible.

Assuming the upper bound of Theorem~\ref{thm:keep52} obtained for $\gamma > 0$ are tight would lead to think that exponential doubling tricks cannot preserve minimax regret bounds of the form $O(\sqrt{T}))$. If true, such a conjecture would need to be supported by a lower bound (a counterpart of Theorem~\ref{thm:LowerBoundGeom_log}). Theorem~\ref{thm:LowerBoundExpo_sqrt} provides such a lower bound, but as discussed below, combined with Theorem~\ref{thm:keep52}, this result rather advocates the use of exponential doubling tricks.  Theorem~\ref{thm:LowerBoundExpo_sqrt} is proved in Appendix~\ref{sub:proof_LowerBoundExpo_sqrt}.

% The lower bound from Theorem~\ref{thm:LowerBoundExpo_sqrt} below is a second step towards proving or disproving the converse of the lower bound of Theorem~\ref{thm:LowerBoundGeom_log}, for bounds of the form $R_T = \bigO{T^{\gamma} (\log T)^{\delta}}$ (with or without $\delta=0$).
% and an exponential sequence.
% This result, along with the previous one Theorem~\ref{thm:keep52}, starts to answer the question of whether an exponential sequence is sufficient to conserve both minimax and problem-dependent guarantees with Doubling Tricks.
%

\begin{theorem}\label{thm:LowerBoundExpo_sqrt}
    If $\cA$ satisfies
    $ R_T(\cA_T) \geq c \; T^{\gamma}$,
    % $(\log T)^{\delta}$
    for $c > 0$ and $0 < \gamma \leq 1$,
    and under a certain \emph{stochastic model},
    % and $\delta \geq 0$,
    then the anytime version $\cA':=\DTr(\cA, (T_i)_{i\in\N})$ with the exponential sequence $(T_i)_{i\in\N}$ of parameters $T_0\in\N^*$, $a>1$, $b>1$ (\ie, $T_i = \lfloor \frac{T_0}{a} a^{b^i} \rfloor$), satisfies this lower bound for a certain constant $c' > 0$,
    under the same model,
    \begin{equation}
        \forall T \geq 1,\;\;\;\;
        R_T(\cA') \geq c' \; \left(T^{\frac{1}{b}}\right)^{\gamma}.
        % \left(\log T\right)^{\delta}.
    \end{equation}
    % for a constant $c' > 0$.
\end{theorem}

% This result, along with the previous one Theorem~\ref{thm:keep52}, starts to answer the question of whether an exponential sequence is sufficient to conserve both minimax and problem-dependent guarantees with Doubling Tricks.

We already saw that any exponential doubling trick can conserve logarithmic problem-dependent regret bounds.
If we could take $b\to1$ in the two previous Theorems~\ref{thm:keep52} and \ref{thm:LowerBoundExpo_sqrt},
both results would match and prove that there exists an exponential doubling trick that can also be used to conserve minimax regret bounds.
%
This argument is not so easy to formulate, as $b$ cannot depend on $T$, but it supports our belief that exponential doubling tricks are good candidates for (asymptotically) preserving both problem-dependent and minimax regret bounds.

%and even if it can be taken as small as we want, it cannot be taken to $1$ (it has to be constant when $T\to\infty$).

% We interpret this result as announced above:
% if $R_T(\cA_T)$ satisfies both finite-time square-root lower  and upper bound (\ie, $\frac{R_T(\cA_T)}{\sqrt{T}}$ is bounded),
% then using an exponential sequence for the doubling trick algorithm is a bad idea as it guarantees a blow up in the regret lower bound and prove that $R_T(\DTr(A, (T_i)_{i\in\N})) = \Omega(T)$.

\subsection{Proof of Theorem~\ref{thm:keep52}} \label{proof:Keep52}

    Let $\cA' := \DTr(\cA, (T_i)_{i\in\N})$, and consider a fixed bandit problem.
    We first consider the harder case of $\delta>0$, see below in Page~\pageref{eq:spadesuit} in $(\spadesuit)$ for the other case.
    The lower bound \eqref{eq:decompositionIneq_LB} from Lemma~\ref{lem:decomposition} gives
    % We begin like in the proof of Theorem~\ref{thm:keep45} (in App.~\ref{sub:proof_keep45}).
    \begin{align*}
        R_T(\cA')
        &\leq \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(\cA_{T = T_i - T_{i-1}}) \\
        \intertext{
            We bound naively\footnotemark{} $T_i - T_{i-1} \leq T_i \leq \frac{T_0}{a} a^{b^i}$,
            and we can use the hypothesis on $\cA$ for each regret term,
            as $f$ and $t \mapsto c t^{\gamma} (\log t)^{\delta}$ are non-decreasing for $t\geq 1$ (by hypothesis for $f$ and by Lemma~\ref{lem:tgammalogtdeltaIncreasing}).
        }
        &\leq \sum_{i=0}^{L_T} f(T_i) + c \sum_{i=0}^{L_T} \left(T_i\right)^{\gamma} \left(\log\left(T_i\right)\right)^{\delta}
        \leq g_1(T) + c \sum_{i=0}^{L_T} \left(\frac{T_0}{a} a^{b^i}\right)^{\gamma} \left(\log\left(\frac{T_0}{a} a^{b^i}\right)\right)^{\delta} \\
        \intertext{
            The first part is denoted $g_1(T):= \sum_{i=0}^{L_T} f(T_i)$ and is dealt with Lemma~\ref{lem:ControlLemmaSumOfSmallO}:
            the sum of $f(T_i)$ is a $\smallO{\sum_{i=0}^{L_T} T_i^{\gamma} (\log(T_i))^{\delta}}$, as $f(t) = \smallO{t^{\gamma} (\log t)^{\delta}}$ by hypothesis, and this sum of $T_i^{\gamma} (\log(T_i))^{\delta}$ is proved below to be bounded by $T^{b\gamma} (\log T)^{\delta}$.
            So $g_1(T) = \smallO{T^{b\gamma} (\log T)^{\delta}}$.
            % for a certain constant $c'>0$.
            %
            The second part is $c \left(\frac{T_0}{a}\right)^{\gamma} \sum_{i=0}^{L_T} \left(a^{b^i}\right)^{\gamma} \left(\log\left(\frac{T_0}{a} a^{b^i}\right)\right)^{\delta}$.
            Define $\log^+(x) := \max(\log(x), 0) \geq 0$, so whether $\frac{T_0}{a} \leq 1$ or $ >1$, we always have $\log\left(\frac{T_0}{a} a^{b^i}\right) \leq \log^+\left(\frac{T_0}{a}\right) + \log\left(a^{b^i}\right)$.
            %
            Then we can use Lemma~\ref{lem:expoInequality} (Eq.~\eqref{eq:expoInequality1}) to distribute the power on $\delta$ (as it is $<1$). So $\left(\log\left(\frac{T_0}{a} a^{b^i}\right)\right)^{\delta} \leq \left(\log^+\left(\frac{T_0}{a}\right)\right)^{\delta} + \left(\log(a)\right)^{\delta} \left(b^i\right)^{\delta}$ with the convention that $0^{\delta} = 0$ (even if $\delta=0$),
            and so this gives
        }
        &\leq g_1(T) + c \left(\frac{T_0}{a}\right)^{\gamma} \left[
            \left(\log^+\left(\frac{T_0}{a}\right)\right)^{\delta} \sum_{i=0}^{L_T} (a^{b^i})^{\gamma}
            + \left(\log(a)\right)^{\delta} \sum_{i=0}^{L_T} (a^{b^i})^{\gamma} \left(b^i\right)^{\delta}
        \right] \\
        \intertext{
            % For the two sums,
            % we use two results proved in Appendix~\ref{sec:otherproofs}.
            %
            If $\gamma=0$ then the first sum is just $L_T+1 = \bigO{\log(\log T)}$ which can be included in $g_1(T) = \smallO{(\log T)^{\delta}}$ (still increasing),
            and so only the second sum has to be bounded, and a geometric sum gives $\sum_{i=0}^{L_T} \left(b^i\right)^{\delta} \leq \frac{b^{\delta}}{b^{\delta} - 1} (b^{L_T})^{\delta}$.
            %
            But if $\gamma > 0$,
            % then first we use
            % Lemma~\ref{lem:SumInequalityUsingEi} (Eq.~\eqref{eq:SumInequalityUsingEi_2}) to
            we can naively
            bound the first sum by $\sum_{i=0}^{L_T} (a^{b^i})^{\gamma} \leq (L_T+1) (a^{b^{L_T}})^{\gamma}$
            % (if $\Ei\left(\gamma\log(a)\right) \geq 0$),
            % and as $\delta>0$.
            Observe that $a^{b^{L_T}} = (a^{b^{L_T}-1})^b \leq (a\frac{T}{T_0})^b$.
            So $a^{b^{L_T}} = \bigO{T^b}$ and $L_T + 1 = \bigO{\log(\log T)}$, thus the first sum is a $\bigO{T^{b\gamma} \log(\log T)} = \smallO{T^{b\gamma} (\log T)^{\delta}}$ (as $\delta>0$).
            In both cases, the first sum can be included in $g_2(T)$ which is still a $\smallO{T^{b\gamma} (\log T)^{\delta}}$
            %
            Another geometric sum bounds the second sum by $\sum_{i=0}^{L_T} (a^{b^i})^{\gamma} \left(b^i\right)^{\delta} \leq (a^{b^{L_T}})^{\gamma} \sum_{i=0}^{L_T} \left(b^i\right)^{\delta} \leq \frac{b^{\delta}}{b^{\delta} - 1}  (b^{L_T})^{\delta}$.
        }
        &\leq g_1(T)
        % + c_1 (a^{b^{L_T}})^{\gamma}
        + c_1 (a^{b^{L_T}})^{\gamma} (b^{L_T})^{\delta}
        \intertext{
            We identify a constant multiplicative overhead
            % $c_1$,
            % depending on $T_0$, $a$, $b$ and $\gamma$, $\delta$.
            % $c_1 := c \left(\frac{T_0}{a}\right)^{\gamma} \left(\log^+\left(\frac{T_0}{a}\right)\right)^{\delta} \left(1 + \frac{1}{\log(b)}\right) \mathbbm{1}(\gamma > 0) \geq 0$,
            % and
            $c_1 := c \left(\frac{T_0}{a}\right)^{\gamma} \frac{b^{\delta}}{b^{\delta} - 1} \left(\log a\right)^{\delta} > 0$.
            %
            % The term $c_1 (a^{b^{L_T}})^{\gamma}$ is $\bigO{T^{b\gamma}}$ which can be included in $g_2(T) := g(T) + c_1 (a^{b^{L_T}})^{\gamma}$ which is still a $\smallO{T^{b\gamma} (\log T)^{\delta}}$.
            %
            The only term left which depends on $L_T$ is $(a^{b^{L_T}})^{\gamma} (b^{L_T})^{\delta}$,
            and it can be bounded by using
            $b^{L_T} = b b^{L_T-1} \leq b \log_a(a \frac{T}{T_0}) = b + b \log^+_a(\frac{T}{T_0}) \leq b + b \log_a(T)$ (as $T\geq1$),
            and
            again with $a^{b^{L_T}} \leq (a\frac{T}{T_0})^b$.
            The constant part of $b^{L_T}$ also gives a $\bigO{T^{b\gamma}}$ term, that can be included in $g(T) := g_2(T) + (a \frac{T}{T_0})^{b\gamma}$ which is still a $\smallO{T^{b\gamma} (\log T)^{\delta}}$, and is still increasing as sum of increasing functions.
            So we can focus on the last term, and we obtain
        }
        &\leq g(T) + c_1 \left(\frac{b}{\log(a)}\right)^{\delta} \left[ \left(\frac{a}{T_0}\right)^{b} \; T^{b} \right]^{\gamma} \left(\log T\right)^{\delta} \\
        \implies R_T(\cA') &\leq g(T) + \ell(\gamma, \delta, T_0, a, b) \; c T^{b\gamma} \left(\log T\right)^{\delta} \text{ with an increasing } g(t) = \smallO{t^{b\gamma} \left(\log t\right)^{\delta}}.
    \end{align*}
    \footnotetext{Here, using the more subtle bound $T_i - T_i \leq \frac{T_0}{a} a^{b^{i-1}} (\alpha^{b^{i-1}}) + 1$, with $\alpha=a^{b-1}$, from Definition~\ref{def:expSequence}, does not seem to help as it becomes too complicated to handle clearly in the $\log$ terms.}

    So the constant multiplicative overhead $\ell$ depends on $\gamma$ and $\delta$ as well as on $T_0$, $a$ and $b$ and is
    \begin{align}
        \ell(\gamma, \delta, T_0, a, b) := \left(\frac{a}{T_0}\right)^{(b-1)\gamma} \;\times\; \frac{b^{2\delta}}{b^{\delta} - 1} > 0,
        \;\;\;\;\text{if}\; \delta > 0.
    \end{align}

    If $T_0 = a$, the overhead
    $\ell(\gamma, \delta, T_0, a, b)$ is minimal at $b^*(\delta) = 2^{1/\delta} > 1$
    and for a minimal value of $\min_{b > 1} \ell(\gamma, \delta, T_0, a, b) = 4$ (\textbf{for any} $\delta$ and $\gamma$).
    %
    Finally, the $a/T_0$ part tends to $0$ if $T_0\to\infty$ so the overhead can be made as small as we want, simply by choosing a $T_0$ large enough (but constant \emph{w.r.t.} $T$).

    $(\spadesuit)$ Now for the other case of $\delta=0$, we can start similarly,
    but instead of bounding naively
    $\sum_{i=0}^{L_T} (a^{b^i})^{\gamma}$
    by $(L_T+1) (a^{b^{L_T}})^{\gamma}$
    we use Lemma~\ref{lem:DoubleExpSumIneq}
    to get a more subtle bound:
    $\sum_{i=0}^{L_T} (a^{b^i})^{\gamma} \leq a^{\gamma} + (1 + \frac{1}{(\log(a)) (\log(b^{\gamma}))}) (a^{b^{L_T}})^{\gamma}$.
    The constant term gets included in $g(T)$, and for the non-constant part,
    $(a^{b^{L_T}})^{\gamma}$ is handled similarly.
    Finally we obtain the overhead
    \begin{align}\label{eq:spadesuit}
        \ell(\gamma, 0, T_0, a, b) := 1 + \frac{1}{(\log(a)) (\log(b^{\gamma}))} > 1.
    \end{align}

\vspace*{-10pt}  % XXX remove if problem
\hfill{}
$\blacksquare$

% -----------------------------------------------------------------
\section{Numerical Experiments}\label{sec:experiments}

We illustrate here the practical cost of using Doubling Trick, for two interesting non-anytime algorithms that have recently been proposed in the literature:
\emph{Approximated Finite-Horizon Gittins indexes}, that we refer to as \AFHG, by \cite{Lattimore16a} (for Gaussian bandits with known variance)
and \KLUCBpp{} by \cite{Menard17} (for Bernoulli bandits).

We first provide some details on these two algorithms, and then illustrate the behavior of Doubling Tricks applied to these algorithms with different doubling sequences.
% \cite{Garivier18}

\subsection{Two Index-Based Algorithms}\label{sub:twoIndexBasedAlgorithms}

% We also give two examples of recent non-anytime index-based algorithms (\KLUCBpp{} and \AFHG)
%For the sake of completeness, we give here the index used by the algorithms \AFHG{} \citep{Lattimore16a} and \KLUCBpp{} \citep{Menard17}.
% that are known to be very efficient both practically and theoretically, and have been proved order-optimal for both problem-dependent and minimax bounds.
%
We denote by
$X_k(t) := \sum_{s < t} Y_{A(s), s} \mathbbm{1}(A(s) = k)$ the accumulated rewards from arm $k$,
and $N_k(t) := \sum_{s < t} \mathbbm{1}(A(s) = k)$ the number of times arm $k$ was sampled.
%
Both algorithms $\cA$ assume to know the horizon $T$.
They compute an index $I_k^{\cA}(t)\in\R$ for each arm $k\in\{1,\dots,K\}$ at each time step $t\in\{1,\dots,T\}$,
and use the indexes to choose the arm with highest index, \ie, $A(t) := \arg\max_{k\in\{1,\dots,K\}} I_k(t)$ (ties are broken uniformly at random).

\begin{itemize}
    \item The algorithm \AFHG{} can be applied for Gaussian bandits with variance $V$ ($=1$ for our experiments).
    % with parameter $\alpha>0$ (default to $1/2$),
    Let $m(T, t) = T - t + 1 \geq 1$, and let
    \begin{equation}\label{eq:index_AFHG}
        I_k^{\AFHG}(t) := \frac{X_k(t)}{N_k(t)} + \sqrt{\frac{V}{N_k(t)} \log\left( \frac{m(T, t)}{N_k(t) \log^{1/2}\left( \frac{m(T, t)}{N_k(t)} \right)} \right)}.
    \end{equation}

    \item The algorithm \KLUCBpp{} can be applied for bounded rewards in $[0,1]$, and in particular for Bernoulli bandits.
    The binary Kullback-Leibler divergence is $\kl(x,y) := x \log(x/y) + (1-x) \log((1-x)/(1-y))$ (for $0<x,y<1$),
    % with parameter $c\geq0$ (default to $c=1$),
    and let $\log^+(x) := \max(0, \log(x)) \geq 0$.
    Let the function $g(n, T) := \log^+\left(\frac{T}{K n} (1 + \left(\log^+(\frac{T}{K n})\right)^2)\right)$ for $n \leq T$,
    and finally let
    \begin{equation}\label{eq:index_KLUCBpp}
        I_k^{\KLUCBpp}(t) := \sup\limits_{q \in [0, 1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T)}{N_k(t)} \right\}.
    \end{equation}
\end{itemize}

% Explain quickly an example of a simple horizon-dependent algorithms, maybe give the explicit formula for the indexes of \KLUCBpp{} and \AFHG{} (good idea to include them, as they are not stated explicitly as such in the article but only in the video of Lattimore's COLT 2016 talk, cf \url{https://youtu.be/p8AwKiudhZ4?t=276}, as the two are used for the experiments?}

\subsection{Experiments}\label{sub:experiments}

% \FIXME{Adapt the text after deciding which figures to include, and WHY INCLUDE WHAT (on doit avoir un message UTILE avec chaque figure, sinon c'est inutile !)}

% \FIXME{Donner des take home message plus propre et plus utiles !}

We present some results from numerical experiments on Bernoulli and Gaussian bandits. More results are presented in Appendix~\ref{sec:otherExperiments}. %, as well as other families (truncated Exponential, Poisson, Gamma arms) which are not included for the sake of conciseness.}.
%
%For the sake of simplicity,
We present in pages~\pageref{fig:bernoulliBandits_DoublingTrick_Restart_BayesianProblem}
and \pageref{fig:bernoulliBandits_DoublingTrick_Restart_fixedProblem}
results for $K=9$ arms
% Different values of $K$ are used, $K=4$ then $K=9$ and $K=16$,
and horizon $T=45678$ (to ensure that no choice of sequence were lucky and had $T_{L_T-1} = T$ or too close to it).
%
We ran $n=1000$ repetitions of the random experiment, either on the same  ``easy'' bandit problem $\bmu$ (with evenly spaced means),
or on $n$ different random instances $\bmu$ sampled uniformly in $[0,1]^K$,
% (this is referred to as ``Bayesian MAB'' in the titles),
and we plot the average regret on $n$ simulations.
%
The black line without markers is the (asymptotic) lower bound in $\sum_{k \neq k^*} (\kl(\mu_k, \mu^*))^{-1} \log T$, from \cite{LaiRobbins85}.
% For the algorithms $\cA$,
We consider
\KLUCBpp{} for Bernoulli bandits (Figures~\ref{fig:bernoulliBandits_DoublingTrick_Restart_BayesianProblem}, \ref{fig:bernoulliBandits_DoublingTrick_Restart_fixedProblem})
% (with $c=1$)
or \AFHG{} for Gaussian bandits (Figures~\ref{fig:gaussianBandits_DoublingTrick_Restart_fixedProblem}, \ref{fig:gaussianBandits_DoublingTrick_Restart_fixedProblem2}),

Each doubling trick algorithm uses the same $T_0=200$ as a first guess for the horizon.
%
% (with $\alpha=0.5$),
We include both the non-anytime version that knows the horizon $T$, and different anytime versions to compare the choice of doubling trick.
To compare against an algorithm that does not need the horizon,
% (but suffers from the successive restart),
we also include \KLUCB{}~\citep{KLUCBJournal} as a baseline for Bernoulli bandits and for Gaussian bandits (in the Gaussian version, the divergence used is $\mathrm{kl}(x,y)=(x-y)^2/2$, and the algorithm is referred to as \UCB{}).
%
We consider are a geometric doubling sequence with $b=2$, and two different exponential doubling sequences: the ``classical'' $b=2$ and a ``slower'' one with $b=1.1$. Both use $a=T_0=200$, and the last one is using $a=2$, $b=2$.
%
Despite what was proved theoretically in Theorem~\ref{thm:keep52}, using $a=T_0$ and a large enough $T_0$ improves regarding to using $a=2$ and a leading $(T_0/a)$ factor.

% See below in Figures~\ref{fig:bernoulliBandits_DoublingTrick_Restart_BayesianProblem} and \ref{fig:bernoulliBandits_DoublingTrick_Restart_fixedProblem} where the algorithms are fully restarted at each breakpoint (using Algorithm~\ref{algo:DTr}).
% % and \ref{fig:bernoulliBandits_DoublingTrick_noRestart} where the algorithms are not fully restarted (using Algorithm~\ref{algo:DTr}).

% \FIXME{I would keep the version with no restart for later, \eg, the experimental/discussion section}

Another version of the Doubling Trick with ``no restart'', denoted $\DTnr$, is presented in Appendix~\ref{sec:DTnr}, but it is only an heuristic and cannot be applied to any algorithm $\cA$.
Algorithm~\ref{algo:DTnr} can be applied to \KLUCBpp{} or \AFHG{} for instance, as they use $T$ just as a numerical parameter (see Eqs.~\ref{eq:index_AFHG} and \ref{eq:index_KLUCBpp}), but its first limitation is that it cannot be applied to \DMEDp{} \citep{Honda10} or \Exppp{} \citep{Seldin17}, or any algorithms based on arm eliminations, for example.
%
A second limitation is the difficulty to analyze this ``no restart'' variant,
due to the unpredictable effect on regret of giving non-uniform prior information to the underlying algorithm $\cA$ on each successive sequence.
An interesting future work would be to analyze it, either in general or for a specific algorithm like \KLUCBpp.
%
Despite its limitations, this heuristic exhibits as expected better empirical performance than \DT, as can be observed in Appendix~\ref{sec:otherExperiments}.

% -----------------------------------------------------------------
\section{Conclusion}\label{sec:conclusion}

We formalized and studied the well-known ``Doubling Trick'' for generic multi-armed bandit problems, that is used to automatically obtain an anytime algorithm from any non-anytime algorithm. Our results are summarized in Table~\ref{fig:array_conclusion}. We show that a geometric doubling can be used to conserve minimax regret bounds (in $\sqrt{T}$), with a constant overhead (typically $\geq 3.33$),
but cannot be used to conserve problem-dependent bounds (in $\log T$),
for which a faster doubling sequence is needed.
An exponential doubling sequence can conserve logarithmic regret bounds also with a constant overhead,
% typically $\geq 3.33$,
but it is still an open question to know if minimax bounds can be obtained for this faster growing sequence.
%
Partial results of both a lower and an upper bound, for bounds of the generic form $T^{\gamma} (\log T)^{\delta}$, let use believe in a positive answer.

It is still an open problem to know if an anytime algorithm
can be both asymptotically optimal for the problem-dependent regret (\ie, with the exact constant) and optimal in a minimax regret (\ie, have a $\sqrt{KT}$ regret),
%
but we believe that using a doubling trick on non-anytime algorithms like \KLUCBpp{} cannot be the solution.
%
We showed that it cannot work with a geometric doubling sequence, and conjecture that exponential doubling trick would never bring the right constant either.

% \FIXME{More future work ?}

%A more promising direction would be to analyze manually a specific algorithm, in line of the work on MOSS-anytime \citep{Degenne16}.

% Cf. https://fr.wikibooks.org/wiki/LaTeX/Tableaux#Alignement_du_texte
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}

\begin{figure}[!t]
    \centering
    \begin{tabular}{M{3cm}|M{5cm}|M{6cm}}
        Bound $\backslash$ Doubling
        % |
        & Geometric, $T_i = \lfloor T_0 b^i\rfloor$
        % |
        & Exponential, $T_i = \lfloor \frac{T_0}{a} a^{b^i}\rfloor$
        \tabularnewline
        \hline
        \hline
        % ----------------------------
        $(\log T)^{\delta}$
        % |
        & \textcolor{red}{$\times$ \emph{Known to fail}}
        % lower bound in
        \newline
        $R_T(\DT) \geq c' (\log T)^{1+\delta}$ if $R_T(\cA_T) \geq c (\log T)^{\delta}$.
        \newline
        (Theorem~\ref{thm:LowerBoundGeom_log})
        % |
        & \textcolor{blue}{$\checkmark$ \emph{Known to work}}, with overhead $\ell(\delta, b) = \frac{b^{2\delta}}{b^{\delta}-1} > 1$.
        \newline
        (Theorem~\ref{thm:keep52})
        \tabularnewline
        \hline
        % ----------------------------
        $T^{\gamma}$
        % |
        & \textcolor{blue}{$\checkmark$ \emph{Known to work}}, with overhead $\ell(\gamma, b) = \frac{b^{\gamma}(b-1)^{\gamma}}{b^{\gamma}-1} > 1$.
        \newline
        (Theorem~\ref{thm:keep45})
        % |
        & \textcolor{darkgreen}{? \emph{Partial}}, best known bound is $c_0' (T^{\frac{1}{b}})^{\gamma} \leq R_T (\DTr \leq \ell c (T^b)^{\gamma}$
        with overhead $\ell>1$,
        if $c_0 T^{\gamma} \leq R_T(\cA_T) \leq c T^{\gamma}$.
        \newline
        (Theorems~\ref{thm:keep52}, \ref{thm:LowerBoundExpo_sqrt})
        \tabularnewline
        \hline
        % \cline{1-2}
        % ----------------------------
        $T^{\gamma} \; (\log T)^{\delta}$
        \newline
        for both
        \newline
        $\gamma>0$, $\delta>0$
        % |
        & \textcolor{blue}{$\checkmark$ \emph{Known to work}}, with overhead $\ell(\gamma, \delta, T_0, b) = \left(\frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \right)^{\delta} \frac{b^{\gamma}(b-1)^{\gamma}}{b^{\gamma}-1} > 1$.
        \newline
        (Theorem~\ref{thm:keep45} )
        % |
        & \textcolor{darkgreen}{? \emph{Partial}}, best known bound is $R_T (\DTr) \leq \ell c (T^b)^{\gamma} (\log T)^{\delta}$ if $R_T(\cA_T) \leq c T^{\gamma} (\log T)^{\delta}$,
        % with overhead $\ell\to0$ for $T_0\to\infty$.
        \newline
        (Theorem~\ref{thm:keep52})
        % \tabularnewline
        % \hline
        % ----------------------------
    \end{tabular}
    \caption{Summary of \emph{known \textcolor{blue}{positive $\checkmark$} and \textcolor{red}{negative $\times$} and \textcolor{darkgreen}{partial results ?}}.}
    \label{fig:array_conclusion}
\end{figure}

% ------------------------------------
% XXX Before first plot
% \hr{}

%
% Regular plots of centralized regrets
%
\begin{figure}[H]
    \centering
    % \begin{subfigure}[!h]{0.49\textwidth}
    \includegraphics[width=0.95\textwidth]{figures/main____env1-1_1217677871459230631.pdf}
    % \end{subfigure}
    % % ~
    % \begin{subfigure}[!h]{0.49\textwidth}
    %   \includegraphics[width=10092textwidth]{XXX.pdf}
    % \end{subfigure}
    \caption{Regret for \DT, for $K=9$ Bernoulli arms, horizon $T=45678$, $n=1000$ repetitions and $\boldsymbol{\mu}$ taken uniformly in $[0,1]^K$. Geometric doubling ($b=2$) and slow exponential doubling ($b=1.1$) are too slow, and short first sequences make the regret blow up in the beginning of the experiment. At $t=40000$ we see clearly the effect of a new sequence for the best doubling trick ($T_i = 200 \times 2^i$). As expected, \KLUCBpp{} outperforms \KLUCB, and if the doubling sequence is growing fast enough then $\DT(\KLUCBpp)$ can perform as well as \KLUCB{} (see for $t < 40000$).}
    \label{fig:bernoulliBandits_DoublingTrick_Restart_BayesianProblem}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

%
% Regular plots of centralized regrets
%
\begin{figure}[H]
    \centering
    % \begin{subfigure}[!h]{0.49\textwidth}
    \includegraphics[width=0.93\textwidth]{figures/main____env1-1_3633169128724378553.pdf}
    % \end{subfigure}
    % % ~
    % \begin{subfigure}[!h]{0.49\textwidth}
    %   \includegraphics[width=10092textwidth]{XXX.pdf}
    % \end{subfigure}
    \caption{Similarly but for $\boldsymbol{\mu}$ evenly spaced in $[0,1]^K$ ($\{0.1,\dots,0.9\}$). Both \KLUCB{} and \KLUCBpp{} are very efficient on ``easy'' problems like this one, and we can check visually that they match the lower bound from \cite{LaiRobbins85}. As before we check that slow doubling are too slow to give reasonable performance.}
    \label{fig:bernoulliBandits_DoublingTrick_Restart_fixedProblem}
    \vspace*{-15pt}  % XXX remove if problem
\end{figure}
%
%
% Regular plots of centralized regrets
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{0.49\textwidth}
    \includegraphics[width=0.93\textwidth]{figures/main____env1-1_2223860464453456415.pdf}
    % \end{subfigure}
    % % ~
    % \begin{subfigure}[!h]{0.49\textwidth}
    %   \includegraphics[width=1.00\textwidth]{XXX.pdf}
    % \end{subfigure}
    \caption{Regret for $K=9$ Gaussian arms $\cN(\mu, 1)$, horizon $T=45678$, $n=1000$ repetitions and $\boldsymbol{\mu}$ taken uniformly in $[-5,5]^K$ and variance $V=1$. On ``hard'' problems like this one, both \UCB{} and \AFHG{} perform similarly and poorly \emph{w.r.t.} to the lower bound from \cite{LaiRobbins85}. As before we check that geometric doubling ($b=2$) and slow exponential doubling ($b=1.1$) are too slow, but a fast enough doubling sequence does give reasonable performance for the anytime \AFHG{} obtained by Doubling Trick.}
    \label{fig:gaussianBandits_DoublingTrick_Restart_fixedProblem}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

% % \FIXME{For a newpage to not have the references starting in a ugly place}
% \clearpage
% \newpage

% -----------------------------------------------------------------
% Acknowledgments---Will not appear in anonymized version
\acks{
    This work is supported
    by the French National Research Agency (ANR), under the project BADASS (grant coded: \texttt{N ANR-16-CE40-0002}),
    % by CentraleSupélec,
    % by the CNRS, under the PEPS project BIO,
    by the French Ministry of Higher Education and Research (MENESR) and {\'E}cole Normale Sup{\'e}rieure Paris-Saclay.
    %
    %Thanks to Odalric-Ambrym Maillard and Florian Strub at Inria Lille for useful discussions.
    %
    The authors wish to thank the anonymous reviewers for their valuable comments that helped to improve the paper.
}

% -----------------------------------------------------------------
% \phantom{\cite{Lattimore16b}}

\bibliography{doublingTrick}
% \nocite{*}  % XXX remove at the end, to only include quoted references!

% \iffalse
\vskip 0.2in
% XXX Remove if problem?
\hr{}
% \vfill{}
\emph{Note}:
    the simulation code used for the experiments is using Python 3.
    % https://www.Python.org/
    It is open-sourced at \verb|https://GitHub.com/SMPyBandits/SMPyBandits|
    %  but is available upon request,
    and fully documented at \newline
    \verb|https://SMPyBandits.GitHub.io|.
    %~\citep{SMPyBandits}.
% \fi
% \hr{}

\clearpage
\newpage

% -----------------------------------------------------------------
\appendix

% -----------------------------------------------------------------
\section{Omitted Proofs}\label{sec:missingproofs}

We include here the proofs omitted in the main document.

\subsection{Proof of Lemma~\ref{lem:decomposition}, ``Regret Lower and Upper Bounds for \DTr''}\label{sub:proof_decompositionDTr}

    Let $\cA'$ denote $\DTr(\cA, (T_i)_{i\in\N})$. For every $k\in \{1,\dots, K\}$,
    \begin{align*}
        & \E\left[\sum_{t=1}^T (X_{k,t} - X_{A(t),t})\right] = \sum_{i=0}^{L_T -1} \E\left[\sum_{t=T_{i-1}}^{T_i}(X_{k,t} - X_{A(t),t})\right] +  \E\left[ \sum_{t=T_{L_T-1}}^{T}(X_{k,t} - X_{A(t),t})\right] \\
        & \hspace{0.7cm} \leq
        \sum_{i=0}^{L_T -1} \max_{k \in \{1,\dots,K\}}\E\left[\sum_{t=T_{i-1}}^{T_i} (X_{k,t} - X_{A(t),t})\right] + \max_{k \in \{1,\dots,K\}}\E\left[ \sum_{t=T_{L_T-1}}^{T}(X_{k,t} - X_{A(t),t})\right] \\
        & \hspace{0.7cm} \leq
        \sum_{i=0}^{L_T -1}R_{T_i-T_{i-1}}\left(\cA'\right) + R_{T - T_{L_T-1}}\left(\cA'\right).
    \end{align*}

    Thus, by definition of the regret
    \begin{align*}
        R_T(\cA')
        &\leq
        \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(\cA')+ R_{T - T_{L_T-1}}(\cA') \\
        &=
        \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}\left(\cA_{T_{i} - T_{i-1}}\right)
        + R_{ \underbrace{T - T_{L_T-1}}_{\leq T_{L_T} - T_{L_T-1}} }\left(\cA_{T_{L_T} - T_{L_{T-1}}}\right) \\
        &\leq
        \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}\left(\cA_{T_{i} - T_{i-1}}\right).
    \end{align*}

    In the stochastic case, it is well known that the regret can be rewritten in the following way, introducing $\mu_k$ the mean of arm $k$ and $\mu^*$ the mean of the best arm:
    \begin{align*}
        R_T(\cA') &=
        \E\left[\sum_{t=1}^{T}(\mu^* - \mu_{A(t)})\right] \\
        &=
        \sum_{i=0}^{L_{T-1}} \E\left[\sum_{t=T_{i-1}}^{T_i}(\mu^* - \mu_{A(t)})\right] + \E\left[\sum_{t=T_{L_T-1}}^{T}(\mu^* - \mu_{A(t)})\right] \\
        &=
        \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}\left(\cA_{T_i - T_{i-1}}\right)) + \underbrace{R_{T - T_{L_T-1}}(\cA')}_{\geq 0}.
    \end{align*}
    and the lower bound follows.
\hfill{}
$\blacksquare$

\newpage

%     By definition of the regret (Eq.\ref{eq:regret}),
%     and without any \iid{} hypothesis on the rewards,
%     let $\cA'$ denotes $\DTr(\cA, (T_i)_{i\in\N})$.
%     As the $\max$ of a sum is bounded by the sum of the $\max$,
%     we can split the sum on $t\in\{1,\dots,T\}$ on the discrete intervals $\{T_{i-1}, \dots, T_i-1\}$ for $i=0,\dots,L_T-1$ and for $\{T_{L_T-1}, \dots, T-1\}$ for the last sequence\footnote{Beware that it is of course $\cA_{T = T_{L_T} - T_{L_T-1}}$ that is used on the last sequence, and not $\cA_{T = T - T_{L_T-1}}$ as the later would need to use the (unknown) horizon $T$.}, and so
%     % then if
%     % $\cR(t)$ denote the instantaneous regret at time $t$,
%     % $\cR(t) := \sum_{k=1}^K (\max_j \mu_j^{(t)} - \mu_k^{(t)}) \Pr_{\cF_{t-1}}\left[ A'(t) = k \right]$,
%     % We have
%     \begin{align*}
%         R_T(\cA')
%         % &\leq \sum_{t=1}^{T} \cR(t),
%         % &\leq \sum_{i=0}^{L_T-1} \sum_{t=T_{i-1}}^{T_i - 1} \cR(t) + \sum_{t=T_{L_T-1}}^{T} \cR(t)
%         &\leq
%         \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(\cA')
%         + \underbrace{R_{T - T_{L_T-1}}(\cA')}_{:= S \geq 0}.
%         % \intertext{
%         %     So to get the identity leading to Eq.~\eqref{eq:decompositionIneq_UB}, one simply has to observe that, by definition of the algorithm $\DTr$, on each of these sequences
%         %     (for $i\in\{0,\dots,L_T\}$),
%         %     $\cA'$ is exactly $\cA_{T_i-T_{i-1}}$ with its internal timer using $t-T_{i-1}$ instead of $t$ (cf. Algorithm~\ref{algo:DTr}, Line~\ref{line:internalTimer_DTr}).
%         %     % The \iid{} hypothesis on the rewards and the fact that we fully restart the algorithm gives
%         %     % $\Pr_{\bmu}\left[ A'(t) = k \right] = \Pr_{\bmu}\left[ A_{T_i-T_{i-1}}(t-T_{i-1}) = k \right]$
%         %     % if $t\in\{T_{i-1}, \dots, T_i-1\}$.
%         % }
%         % R_T(\cA') &\leq
%         % \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(\cA_{T = T_i - T_{i-1}})
%         % + \underbrace{R_{T - T_{L_T-1}}(\cA_{T = T_{L_T} - T_{L_T-1}})}_{:= S \geq 0}
%     \end{align*}
%     So to get Eq.~\eqref{eq:decompositionIneq_UB}, one simply has to observe that, by definition of the algorithm $\DTr$, on each of these sequences
%     (for $i\in\{0,\dots,L_T\}$),
%     $\cA'$ is exactly $\cA_{T_i-T_{i-1}}$ with its internal timer using $t-T_{i-1}$ instead of $t$ (cf. Algorithm~\ref{algo:DTr}, Line~\ref{line:internalTimer_DTr}).
%     % First, we lower bound the last sum $S$ by $0$ to get the lower bound $(LB)$.
%     % \implies R_T(A') &\underset{LB}{\geq} \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}).
%     We bound $S$ by a sum on a \emph{longer} interval, up to $T_{L_T} > T$, as each term in $S$ is non-negative.
%     % ($(\max_j \mu_j^{(t)} - \mu_k^{(t)}) \Pr_{\bmu}\left[ A'(t) = k \right] \geq 0$).
%     % R_T(A')
%     % &\leq \sum_{i=0}^{L_T} \sum_{t=0}^{T_i - T_{i-1}} \sum_{k=1}^K (\max_j \mu_j^{(t)} - \mu_k^{(t)}) \Pr_{\bmu}\left[ A_{T_i-T_{i-1}}(t) = k \right] \\
%     % \implies R_T(A') &\underset{UP}{\leq} \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}).
%
%     Now for the stochastic case, the \iid{} hypothesis on the reward make the first inequality above an equality, as the $\max$ and the expectation can be exchanged.
%     For the lower bound Eq.~\eqref{eq:decompositionIneq_LB}, we simply have to observe that $S \geq 0$.
% \end{proof}

% \paragraph{Remark.}
% Note that the lower bound becomes looser when $T$ gets closer to $T_{L_T}-1$, and conversely the upper bound becomes looser when $T$ gets closer to $T_{L_T-1}$.

% \subsection{Proof of Theorem~\ref{thm:keep4}}\label{sub:proof_keep4}

% \begin{proof}\label{proof:Keep4}
%     Let $A' := \DTr(A, (T_i)_{i\in\N})$.
%     We start by using Lemma~\ref{lem:decomposition},
%     % we get a decomposition on the regret of $A$ for time $T$ as a sum of regrets on sequences of successive lengths $T_i - T_{i-1}$, so
%     \begin{align*}
%         R_T(A')
%         &\leq \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}) \\
%         \intertext{
%             We bound $T_i - T_{i-1} \leq 1 + T_0 (b - 1) b^{i-1}$ for $i>0$, thanks to Definition~\ref{def:geomSequence},
%             and we can use the hypothesis on $A$ for each regret term,
%             as $t \mapsto c t^{\gamma}$ and $f$ are non-decreasing for $t\geq 1$ (as $\gamma>0$).
%             Additionally,
%             $(T_i - T_{i-1})^{\gamma} \leq 1 + T_0^{\gamma} (b - 1)^{\gamma} (b^{i-1})^{\gamma}$ by Lemma~\ref{lem:expoInequality} (Eq.~\eqref{eq:expoInequality1}, as $\gamma<1$), thus
%         }
%         &\leq \sum_{i=0}^{L_T} c (T_i - T_{i-1})^{\gamma} + f(T_i - T_{i-1})
%         \leq \sum_{i=0}^{L_T} f(T_i) + c T_0^{\gamma} \sum_{i=1}^{L_T} c (T_i - T_{i-1})^{\gamma} \\
%         &\leq \left[\sum_{i=0}^{L_T} f(T_i) + c (T_0^{\gamma} + L_T)\right] + c T_0^{\gamma} (b - 1)^{\gamma} \sum_{i=1}^{L_T} (b^{i-1})^{\gamma} \\
%         \intertext{
%             The first part is denoted $g(T):=\sum_{i=0}^{L_T} f(T_i) + c (T_0^{\gamma} + L_T)$, and is an increasing function as a sum of increasing functions.
%             By Lemma~\ref{lem:ControlLemmaSumOfSmallO}, the sum of $f(T_i)$ is a $\smallO{\sum_{i=0}^{L_T} T_i^{\gamma}}$, as $f(t) = \smallO{t^{\gamma}}$ by hypothesis, and this sum of $T_i^{\gamma}$ is proved below to be bounded by $c' T^{\gamma}$ for a certain constant $c'>0$.
%             Moreover, $L_T = \bigO{\log T} = \smallO{T^{\gamma}}$ (for any $\gamma>0$),
%             so $g(t)$ is $\smallO{t^{\gamma}}$ as wanted.
%             %
%             We bound the second part by $\sum_{i=0}^{L_T-1} (b^i)^{\gamma} \leq \frac{1}{b^{\gamma} - 1} (b^{L_T})^{\gamma}$ thanks to a geometric sum and thus
%         }
%         &\leq g(T) + c T_0^{\gamma} \frac{(b - 1)^{\gamma}}{b^{\gamma} - 1} (b^{L_T})^{\gamma}
%         \intertext{
%             We see that $b^{L_T} = b b^{L_T - 1}$, and $b^{L_T - 1} \leq \frac{T}{T_0}$ also thanks to Definition~\ref{def:geomSequence}, so
%         }
%         &\leq g(T) + c T_0^{\gamma} \frac{b^{\gamma} (b - 1)^{\gamma}}{b^{\gamma} - 1} \frac{T^{\gamma}}{T_0^{\gamma}}
%         = g(T) + \frac{b^{\gamma} (b - 1)^{\gamma}}{b^{\gamma} - 1} \; c T^{\gamma} \\
%         \implies R_T(A') &\leq g(T) + \ell(\gamma, b) \; c T^{\gamma} \text{ with an increasing } g(T) = \smallO{T^{\gamma}}.
%     \end{align*}

%     To minimize this constant overhead $\ell(\gamma, b) = \frac{b^{\gamma} (b - 1)^{\gamma}}{b^{\gamma} - 1} > 1$, we fix $\gamma$ and study $h: b \mapsto \ell(\gamma, b)$.
%     The function $h$ is of class $\cC^1$ on $(1, \infty)$ and $h(b) \to +\infty$ for $b\to 1^+$ and $b\to\infty$,
%     so $h$ has a (possibly non-unique) global minimum and attains it.
%     Moreover $h'(b) = \frac{\gamma b^{\gamma-1} (b-1)^{\gamma-1}}{(b^{\gamma}-1)^2} \left( b^{\gamma+1} - 2 b + 1 \right)$ has the sign of $b^{\gamma+1} - 2 b + 1$,
%     which does not have a constant sign and does not have explicit root(s) for a generic $\gamma$.
% \end{proof}

\subsection{Proof of Theorem~\ref{thm:keep45}, ``Conserving a Regret Upper Bound with Geometric Horizons''}\label{sub:proof_keep45}

It is interesting to note that
the proof is valid for both the easiest case when $\delta=0$ (as it was known in \cite{CesaLugosi06} for $\gamma=1/2$)
and the generic case when $\delta\geq0$,
with no distinction.

As far as the authors know, this result in its generality with $\delta\geq0$ is new.

\vspace*{5pt}  % XXX remove if problem
\begin{proof}\label{proof:Keep45}
    Let $\cA' := \DTr(\cA, (T_i)_{i\in\N})$, and consider a fixed bandit problem.
    % The proof starts exactly like for the previous one for Theorem~\ref{thm:keep4}.
    The upper bound \eqref{eq:decompositionIneq_UB} from Lemma~\ref{lem:decomposition} gives
    \begin{align*}
        R_T(\cA')
        &\leq \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}) \\
        \intertext{
            We can use the hypothesis on $\cA$ for each regret term,
            as $f$ and $t \mapsto c t^{\gamma} (\log t)^{\delta}$ are non-decreasing for $t\geq 1$ (by hypothesis for $f$ and by Lemma~\ref{lem:tgammalogtdeltaIncreasing}).
        }
        &\leq \sum_{i=0}^{L_T} f(T_i - T_{i-1}) + c T_0^{\gamma} (\log T_0)^{\delta}
        + c \sum_{i=1}^{L_T} \left(T_i - T_{i-1}\right)^{\gamma} \left(\log\left(T_i - T_{i-1}\right)\right)^{\delta}
        \intertext{
            The first part is denoted $g_1(T):= \sum_{i=0}^{L_T} f(T_i - T_{i-1}) + c T_0^{\gamma} (\log T_0)^{\delta}$, it is an increasing function as a sum of increasing functions, and it is dealt with by using Lemma~\ref{lem:ControlLemmaSumOfSmallO}:
            the sum of $f(T_i - T_{i-1})$ is a $\smallO{\sum_{i=0}^{L_T} (T_i - T_{i-1})^{\gamma} (\log(T_i - T_{i-1}))^{\delta}}$, as $f(t) = \smallO{t^{\gamma} (\log t)^{\delta}}$ by hypothesis, and this sum of $(T_i - T_{i-1})^{\gamma} (\log(T_i - T_{i-1}))^{\delta}$ is proved below to be bounded by $c' T^{\gamma} (\log T)^{\delta}$ for a certain constant $c'>0$,
            which gives $g_1(T) = \smallO{T^{\gamma} (\log T)^{\delta}}$.
            %
            For the second part,
            we bound $T_i - T_{i-1} \leq T_0 (b-1) b^{i-1} + 1$ thanks to Definition~\ref{def:geomSequence}.
            Moreover, as $\gamma<1$
            we can use Lemma~\ref{lem:expoInequality} (Eq.~\eqref{eq:expoInequality1}) to distribute the power on $\gamma$,
            so $(T_0 (b-1) b^{i-1} + 1)^{\gamma} \leq (T_0 (b-1) b^{i-1})^{\gamma} + \mathbbm{1}(\gamma\neq0)$ (indeed if $\gamma=0$ both sides are equal to $1$).
            This gives
        }
        &\leq g_1(T) + c
            (T_0 (b-1))^{\gamma} \sum_{i=1}^{L_T} (b^{i-1})^{\gamma} \left(\log(T_i - T_{i-1})\right)^{\delta}
            + c \mathbbm{1}(\gamma\neq0) \sum_{i=1}^{L_T} \left(\log(T_i - T_{i-1})\right)^{\delta} \\
        \intertext{
            If $\gamma\neq 0$,
            the last sum is bounded by $\sum_i^{L_T} (\log T_i)^{\delta} \leq (\log T_0)^{\delta}(L_T + 1) + (\log b)^{\delta} \sum_i^{L_T} i^{\delta}$ which is a $\bigO{L_T^{\delta+1}} = \bigO{(\log T)^{\delta+1}} = \smallO{T^{\gamma} (\log T)^{\delta}}$ (as $\gamma>0$, thanks to a geometric sum), and so it can be included in $g_2(T) = \smallO{T^{\gamma} (\log T)^{\delta}}$.
            If $\gamma=0$, there is only the first sum.
            %
            We bound again $T_i - T_{i-1} \leq T_0 (b-1) b^{i-1} + 1$
            and use Lemma~\ref{lem:logxm1_logx} to bound $\log(T_0 (b-1) b^{i-1} + 1)$
            by $\frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \log(T_0 (b-1) b^{i-1}) $ term (as $T_0 (b-1) > 1$ by hypothesis).
        }
        &\leq g_2(T) + c
            (T_0 (b-1))^{\gamma} \sum_{i=1}^{L_T} (b^{i-1})^{\gamma} \left( \frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \log(T_0 (b-1) b^{i-1})\right)^{\delta} \\
        \intertext{
            We split the $\log(T_0(b-1)b^{i-1})$ term in two, and
            once again, the term with $\log(T_0(b-1))$ gives a $\bigO{{b^{L_T-1}}}$ (by a geometric sum), which gets included in $g_3(T) = \smallO{T^{\gamma} (\log T)^{\delta}}$.
            We focus on the fastest term, and we can now rewrite the sum from $i=0$ to $L_T-1$,
        }
        &\leq g_3(T) + c
            (T_0 (b-1))^{\gamma} \left(\log(b) \frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \right)^{\delta} \sum_{i=0}^{L_T-1} (b^i)^{\gamma} i^{\delta} \\
        \intertext{
            We naively bound $i^{\delta}$ by $(L_T-1)^{\delta}$,
            and use a geometric sum
            % Corollary~\ref{cor:SumIneq3} (Eq.~\eqref{eq:sum_igamma_bidelta})
            to have
        }
        &\leq g_3(T) + c
            (T_0 (b-1))^{\gamma} \left(\log(b) \frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \right)^{\delta} (L_T-1)^{\delta} \frac{b^{\gamma}}{b^{\gamma}-1} (b^{L_T-1})^{\gamma} \\
        \intertext{
            Finally, observe that $L_T-1 \leq \log_b(\frac{T}{T_0}) \leq \log_b(T)$, so the $(\log b)^{\delta}$ term simplifies,
            and observe that $b^{L_T-1} \leq \frac{T}{T_0}$ so the $T_0^{\gamma}$ term also simplifies.
            Thus we get
        }
        &\leq g_3(T) + c
            \left(\frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \right)^{\delta} \frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma}-1} T^{\gamma} (\log T)^{\delta}.
    \end{align*}

    The constant multiplicative overhead $\ell$ depends on $\gamma$ and $\delta$ as well as on $T_0$ and $b$,
    and is
    $\ell(\gamma, \delta, T_0, b) := \left(\frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \right)^{\delta} \frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma}-1} > 1$.
\end{proof}

\paragraph{Minimizing the constant overhead?}
%
This constant overhead has two distinct part,
$\ell(\gamma, \delta, T_0, b) = \ell_1(\delta, T_0, b)$ $\ell_2(\gamma, b)$,
with $\ell_1$ depending on $\delta$, $T_0$ and $b$ (equal to $1$ if $\delta=0$),
and $\ell_2$ depending on $\gamma$ and $b$.
\begin{itemize}
    \item
Minimizing this constant overhead $\ell_1(\delta, T_0, b) := \left(\frac{\log(T_0 (b-1) + 1)}{\log(T_0 (b-1))} \right)^{\delta}\geq 1$ is independent of $\delta$ (even if it is $0$).
If we assume $b$ to be fixed,
% (as a value is suggested by minimizing $\ell_2(\gamma, b)$),
$\ell_1(\delta, T_0, b) \to 1$ when $T_0\to\infty$.
Moreover, for any $\delta$ and $b>1$,
$\ell_1(\delta, T_0, b)$ goes to $1$ very quickly when $T_0$ is large enough.
For instance, for $\gamma=\delta=\frac{1}{2}$ and $b=\frac{3+\sqrt{5}}{2}$ (see Corollary~\ref{cor:keep4}),
then $\ell_1(\delta, T_0, b) \simeq 1.109$ for $T_0=2$, $\simeq 1.01$ for $T_0=10$ and $\simeq 1.0004$ for $T_0=100$.
    \item
To minimize this constant overhead $\ell_2(\gamma, b) := \frac{b^{\gamma} (b - 1)^{\gamma}}{b^{\gamma} - 1} > 1$, we fix $\gamma$ and study $h: b \mapsto \ell_2(\gamma, b)$.
The function $h$ is of class $\cC^1$ on $(1, \infty)$ and $h(b) \to +\infty$ for $b\to 1^+$ and $b\to\infty$,
so $h$ has a (possibly non-unique) global minimum and attains it.
Moreover $h'(b) = \frac{\gamma b^{\gamma-1} (b-1)^{\gamma-1}}{(b^{\gamma}-1)^2} \left( b^{\gamma+1} - 2 b + 1 \right)$ has the sign of $b^{\gamma+1} - 2 b + 1$,
which does not have a constant sign and does not have explicit root(s) for a generic $\gamma$.
%
However, it is easy to minimizing $\ell_2(\gamma, b)$ for $b$ numerically when $\gamma$ is known and fixed (with, \eg, Newton's method).
\end{itemize}

% XXX To minimize it, we fix $\gamma$ and study $h: x \mapsto \frac{x^2}{x - 1}$.
% The function $h$ is of class $\cC^1$ on $(1, \infty)$ and $h(b) \to +\infty$ for $b\to 1^+$ and $b\to\infty$,
% so $h$ has a (possibly non-unique) global minimum and attains it.
% Moreover $h'(x) = \frac{x}{(x-1)^2}(x-2)$ has the sign of $x-2$, so $h$ is convex
% and has a unique global minimum attained at $x^*=2$.
% At this point, $h(x^*) = \frac{(x^*)^2}{x^* - 1} = 4$.
% % Moreover $h'(b) = \frac{b^{\gamma - 1} \gamma}{(\log(b^{\gamma}))^2} \left( (\log(b^{\gamma}))^2 + \log(b^{\gamma}) - 1 \right)$ has the sign of $y^2 + y - 1$ (if $y := \log(b^{\gamma}) > 0$).
% % The polynomial $y^2 - y + 1$ has a unique root on $(0, +\infty)$, $y_0 := \frac{\sqrt{5} - 1}{2} \simeq 0.618$.
% % Thus, as $x_0 := \exp(y_0) > 1$, $h$ has a unique global minimum, attained for $b^*(\gamma) = x_0^{1/\gamma} = \exp(\frac{y_0}{\gamma}) > 1$.
% % At this point, $h(b^*(\gamma)) = x_0 (1 + \frac{1}{x_0}) = \exp\left(\frac{\sqrt{5} - 1}{2}\right) \left( \frac{\sqrt{5} + 1}{\sqrt{5} - 1}\right) \simeq 4.86 > 1$.
% %
% % Note that this constant is larger than the minimum found in the previous Theorem~\ref{thm:keep4}, as we were less subtle in the inequalities.

The result from Theorem~\ref{thm:keep45} of course implies the result from \cite[Ex.2.9]{CesaLugosi06}, in the special case of $\delta=0$ and $\gamma=\frac{1}{2}$ (for minimax bounds),
as stated numerically in the following Corollary~\ref{cor:keep4}.

\begin{corollary}\label{cor:keep4}
    If $\gamma=\frac{1}{2}$ and $\delta=0$,
    the multiplicative overhead $\ell(\frac{1}{2}, 0, T_0, b)$ does not depend on $T_0$.
    It is then minimal for $b^*(\frac{1}{2}) = \frac{3 + \sqrt{5}}{2} \simeq 2.62$ and its minimum is $\sqrt{\frac{11 + 5 \sqrt{5}}{2}} \simeq 3.33$.
    Usually $b = 2$ is used, which gives an overhead of $\frac{\sqrt{2}}{\sqrt{2} - 1} \simeq 3.41$, close to the optimal value.

    % If $c_1 \neq 0$, the optimal value $b^*$ for the parameter $b$ depends on $c_1$ and $c_1'$ but not on $T_0$, so it can still be computed for known values of $c_1,c_1'$.
    % There is a tradeoff between the term $\frac{2}{\e \log(b)}$ in front of $c_1$ (which goes to $\infty$ for $b\to 1$ and $0$ for $b\to\infty$) and the term in front of $c_1'$ (which goes to $\infty$ for $b\to\infty$), and a global minimum $b^*$ always exist and is unique, but a closed form is less helpful.

    In particular, order-optimal and optimal algorithms for the minimax bound
    have $\gamma=\frac{1}{2}$ and $f(t)=0$,
    for which Theorem~\ref{thm:keep52} gives a simpler bound
    \begin{equation}\label{eq:corkeep4_1}
        % \forall T \geq 1,\;\;\;\;
        R_T(\cA_T) \leq c \sqrt{T}
        \implies
        R_T(\DTr(\cA, (T_0 2^i)_{i\in\N})) \leq \frac{\sqrt{2}}{\sqrt{2} - 1} c \sqrt{T}.
    \end{equation}
\end{corollary}

% \subsection{XXX}

% % \FIXME{Celui là est bon est intéressant. Mais ça ne révolutionne rien. Par contre il est très propre puisqu'on trouve la même overhead que dans le théorème \ref{thm:keep4} avec geometric sequences for $\sqrt{T}$ bounds.}

% \begin{theorem}\label{thm:keep5}
%     If an algorithm $A$ satisfies
%     $ R_T(A) \leq c (\log T)^{\delta}+ f(t)$,
%     for $\delta > 0$,
%     and for $c > 0$, and an increasing function $f(t) = \smallO{(\log t)^{\delta}}$ (at $t\to\infty$),
%     then the anytime version $A':=\DTr(A, (T_i)_{i\in\N})$ with the exponential sequence $(T_i)_{i\in\N}$ of parameters $T_0\in\N^*,a,b>1$ (\ie, $T_i = \lfloor \frac{T_0}{a} a^{b^i}\rfloor$),
%     satisfies the following inequality
%     \begin{equation}\label{eq:keep5_1}
%         \forall T \geq 1,\;\;\;\;
%         R_T(A') \leq \ell(\delta, b) \; c \; (\log T)^{\delta} + g(T).
%     \end{equation}
%     with an increasing function $g(t) = \smallO{(\log T)^{\delta}}$,
%     and a \emph{constant overhead} $\ell(\delta, b) > 1$
%     \begin{equation}\label{eq:keep5_2}
%         \ell(\delta, b) :=
%         \frac{b^{2\delta}}{b^{\delta} - 1} \geq 4.
%     \end{equation}

%     $\ell(\delta, b)$ is minimal at $b^*(\delta) = 2^{1/\delta} > 1$
%     and for a minimal value of $\min\limits_{b > 1} \ell(\delta, b) = 4$ (\textbf{for any} $\delta$).
% \end{theorem}

% \FIXME{Move the proof in Appendix? It's not that different from the previous one...}
% \begin{proof}\label{proof:Keep5}
%     Let $A' := \DTr(A, (T_i)_{i\in\N})$.
%     The proof starts like for the previous ones for Theorem~\ref{thm:keep45}.
%     \begin{align*}
%         R_T(A')
%         &\leq \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}) \\
%         \intertext{
%             We bound naively $T_i - T_{i-1} \leq T_i \leq \frac{T_0}{a} a^{b^i}$,
%             and we can use the hypothesis on $A$ for each regret term,
%             as $f$ and $t \mapsto c  (\log t)^{\delta}$ are non-decreasing for $t\geq 1$ (by hypothesis for $f$ and by Lemma~\ref{lem:tgammalogtdeltaIncreasing}).
%         }
%         &\leq \sum_{i=0}^{L_T} f(T_i) + c \sum_{i=0}^{L_T} \left(\log\left(T_i\right)\right)^{\delta}
%         \leq g_1(T) + c \sum_{i=0}^{L_T} \left(\log\left(\frac{T_0}{a} a^{b^i}\right)\right)^{\delta} \\
%         \intertext{
%             The first part is denoted $g_1(T):= \sum_{i=0}^{L_T} f(T_i)$ and is dealt with Lemma~\ref{lem:ControlLemmaSumOfSmallO}.
%             The sum of $f(T_i)$ is a $\smallO{\sum_{i=0}^{L_T} (\log(T_i))^{\delta}}$, as $f(t) = \smallO{(\log t)^{\delta}}$ by hypothesis, and this sum of $(\log(T_i))^{\delta}$ is proved below to be bounded by $(\log T)^{\delta}$ for a certain constant $c'>0$.
%             %
%             The second part is $c \sum_{i=0}^{L_T}  \left(\log\left(\frac{T_0}{a} a^{b^i}\right)\right)^{\delta}$.
%             Define $\log^+(x) := \max(\log(x), 0) \geq 0$, so whether $\frac{T_0}{a} \leq 1$ or $ >1$, we always have $\log\left(\frac{T_0}{a} a^{b^i}\right) \leq \log^+\left(\frac{T_0}{a}\right) + \log\left(a^{b^i}\right)$.
%             %
%             Thus we can use Lemma~\ref{lem:expoInequality} (Eq.~\eqref{eq:expoInequality1}) to distribute the power on $\delta$ (as it is $<1$). So $\left(\log\left(\frac{T_0}{a} a^{b^i}\right)\right)^{\delta} \leq \left(\log^+\left(\frac{T_0}{a}\right)\right)^{\delta} + \left(\log(a)\right)^{\delta} \left(b^i\right)^{\delta}$ with $0^{\delta} = 0$ (even if $\delta=0$),
%             and so this gives
%         }
%         &\leq g_1(T) + c \left[
%             \left(\log^+\left(\frac{T_0}{a}\right)\right)^{\delta} (L_T + 1)
%             + \left(\log(a)\right)^{\delta} \sum_{i=0}^{L_T} \left(b^i\right)^{\delta}
%         \right] \\
%         \intertext{
%             For the two terms,
%             we use two results proved in Appendix~\ref{sec:otherproofs}.
%             We know that for the exponential sequence, $L_T = \bigO{\log(\log T)}$ which can be included in $g_1(T) = \smallO{(\log T)^{\delta}}$ (still increasing),
%             and so only the second sum has to be bounded.
%             a geometric sum gives $\sum_{i=0}^{L_T} \left(b^i\right)^{\delta} \leq \left(\frac{b^{\delta}}{b^{\delta} - 1}\right) (b^{L_T})^{\delta} = \left(\frac{b^{2\delta}}{b^{\delta} - 1}\right) (b^{L_T-1})^{\delta}$.
%         }
%         &\leq g_1(T) + c c' (b^{L_T-1})^{\delta} \\
%         \intertext{
%             We identify a constant multiplicative overhead $c' := (\log a)^{\delta} \frac{b^{2\delta}}{b^{\delta} - 1}$.
%             We know that
%             $b^{L_T-1} \leq \log_a(\frac{T}{T_0}) \leq \log_a(T)$ (as $T_0 \geq 1$),
%             and so the $(\log a)^{\delta}$ get cancelled and thus
%         }
%         &\leq g(T) + \frac{b^{2\delta}}{b^{\delta} - 1} c \left(\log T\right)^{\delta} \\
%         \implies R_T(A') &\leq g(T) + \ell(\delta, b) \; c \left(\log T\right)^{\delta} \text{ with an increasing } g(T) = \smallO{\left(\log T\right)^{\delta}}.
%     \end{align*}

%     So the constant multiplicative overhead $\ell$ depends only on $\delta$ and $b$ and is
%     \begin{align*}
%         \ell(\delta, b) :=
%         \frac{b^{2\delta}}{b^{\delta} - 1} \geq 4
%         % \begin{cases}
%         %     1 + \frac{1}{\delta\log(b)} > 1 & \text{~if~} \gamma > 0, \\
%         %     1 + \frac{1}{b^{\delta} - 1} > 1 & \text{~if~} \gamma = 0.
%         % \end{cases}
%     \end{align*}
%     The constant is always larger than its minimum $\min_{b>1} \ell(\delta, b) = \ell(\delta, b^*(\delta)) = 4$, attained for $b = b^*(\delta) = 2^{\frac{1}{\delta}}$.
% \end{proof}

\subsection{Proof of Theorem~\ref{thm:LowerBoundExpo_sqrt}, ``Minimax Regret Lower Bound with Exponential Horizons''}\label{sub:proof_LowerBoundExpo_sqrt}

\begin{proof}\label{proof:LowerBoundExpo_sqrt}
    Let $\cA' := \DTr(\cA, (T_i)_{i\in\N})$, and consider a fixed \emph{stochastic} bandit problem.
    The lower bound \eqref{eq:decompositionIneq_LB} from Lemma~\ref{lem:decomposition} gives
    % The proof starts similarly to the proof of Theorem~\ref{thm:LowerBoundGeom_log}.
    \begin{align*}
        R_T(\cA')
        &\geq \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}) \\
        \intertext{
            We can use the hypothesis on $\cA$ for each regret term,
            and as $0 < \gamma \leq 1$, we can use Lemma~\ref{lem:expoInequality} (Eq.~\eqref{eq:expoInequality3}) to distribute the power on $\gamma$ to ease the proof and obtain
        }
        &\geq c T_0^{\gamma} + c \sum_{i=1}^{L_T-1} (T_i - T_{i-1})^{\gamma} \\
        &\geq c T_0^{\gamma} + c \sum_{i=1}^{L_T-1} \left(T_i^{\gamma} - T_{i-1}^{\gamma}\right) \;\;\;\;\text{(it is a telescopic sum and simplifies)} \\
        &\geq c T_{L_T-1}^{\gamma} \\
        \intertext{
            % $(\log (T_{L_T-1}))^{\delta}$
            % \FIXME{Maybe justify more how to handle the $(\log(...))^{\delta}$ term? Je crois que c'est trop rapide en l'état, on doit dire mieux (mais le théorème marche, j'en suis sur).}
            % As the telescopic sum simplifies.
            % and because we bound $\log (T_i - T_{i-1}) \geq \log (T_{L_T-1})$.
            % We bound naively $T_i \geq \frac{T_0}{a} a^{b^i} - \frac{T_0}{a} a^{b^{i-1}} - 1$.
            Observe that
            $T_{L_T-1} \geq (T_{L_T})^{\frac{1}{b}}$ by definition of the exponential sequence (Def.~\ref{def:expSequence}), and $T_{L_T} \geq T$ (Def.~\ref{def:lastTerm}).
            For the $\log (T_{L_T-1})$ term, we simply have $\log (T_{L_T-1}) \geq \frac{1}{b} \log(T)$ so if $c' = c / b$,
            then we obtain what we want
        }
        R_T(\cA')
        &\geq c' \; T^{\frac{\gamma}{b}}.
    \end{align*}
    This lower bound goes from $R_T(\cA_T) = \Omega(T^{\gamma})$ to $R_T(\DT(\cA)) = \Omega(T^{\frac{\gamma}{b}})$,
    and it looks very similar to the upper bound from Theorem~\ref{thm:keep52} where
    $R_T(\DT(\cA)) = \bigO{T^{b\gamma}}$
    was obtained from
    $R_T(\cA_T) = \bigO{T^{\gamma}}$.
\end{proof}

\paragraph{Remark}
It does seem sub-optimal to lower bound $T_{L_T-1}$ like this ($T_{L_T-1} \geq (T_{L_T})^{\frac{1}{b}}$), but we remind that $T$ can be located anywhere in the discrete interval $\{T_{L_T-1}, \dots, T_{L_T} - 1\}$, so in the worst case when $T$ is very close to $T_{L_T}$ (and for large enough $T$), we indeed have $T_{L_T-1}^b \sim T_{L_T}$ and $T_{L_T} \sim T$, so with this approach, the lower bound $T_{L_T-1} \geq T^{\frac{1}{b}}$ cannot be improved.

% -----------------------------------------------------------------

% \hr{}

% \section{Minimax Regret Lower Bound with Geometric Horizons}\label{sub:LowerBoundGeom_sqrt}

% \TODO{Remove, weird statement hard to get for the reader}

% We include here a last result that partly replies to Theorem~\ref{thm:keep45}.
% It is more subtle that the lower bound in Theorem~\ref{thm:LowerBoundGeom_log} but still provides an interesting insight:
% if $b$ is not chosen carefully (\ie, if $\ell_0(\gamma,b) > 1$), then the anytime version of $\cA_T$ using a geometric Doubling Trick suffers a non-improvable constant multiplicative overhead compared to $\cA_T$.

% \begin{theorem}\label{thm:LowerBoundGeom_sqrt}
%     For \emph{stochastic models},
%     if $\cA$ satisfies
%     $ R_T(\cA_T) \geq c \; T^{\gamma}$,
%     for $0<\gamma<1$ and $c > 0$,
%     then the anytime version $\cA':=\DTr(\cA, (T_i)_{i\in\N})$ with the geometric sequence $(T_i)_{i\in\N}$ of parameters $T_0\in\N^*,b>1$ (\ie, $T_i = \lfloor T_0 b^i \rfloor$) satisfies
%     \begin{equation}
%         % \forall T \geq 1,\;\;\;\;
%         L_T \geq 2 \implies
%         R_T(\cA') \geq \ell_0(\gamma, b) \; c \; T^{\gamma} + g_0(T).
%     \end{equation}
%     with $g_0(t) = \bigO{\log t} = \smallO{t^{\gamma}}$,
%     and a \emph{constant overhead} $\ell_0(\gamma, b)$
%     depending only on $\gamma$ and $b$,
%     \begin{align}
%         \ell_0(\gamma, b) &= \frac{(b-1)^{\gamma}}{b^{\gamma} (b^{\gamma} - 1)} > 0.
%     \end{align}
%     $\ell_0(\gamma, b)$ is always $> 0$ and tends to $0$ for $b\to\infty$,
%     and some choice of $b$ gives $\ell_0(\gamma, b) > 1$.
%     % For instance for $\gamma=\frac{1}{2}$, $b=2$ gives $\ell_0(\frac{1}{2}, 2) = 1+\frac{\sqrt{2}}{2} \simeq 1.71$.
% \end{theorem}
% %
% \begin{proof}\label{proof:LowerBoundGeom_sqrt}
%     Let $\cA' := \DTr(\cA, (T_i)_{i\in\N})$, and consider a fixed \emph{stochastic} bandit problem.
%     Assume $L_T \geq 2$.
%     The lower bound \eqref{eq:decompositionIneq_LB} from Lemma~\ref{lem:decomposition} gives
%     % we get a decomposition on the regret of $\cA$ for time $T$ as a sum of regrets on sequences of successive lengths $T_i - T_{i-1}$, without the last sequence to have a lower bound, and
%     % so if $L_T \geq 2$ we have
%     \begin{align*}
%         R_T(\cA')
%         &\geq \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(A_{T = T_i - T_{i-1}}) \\
%         \intertext{
%             We bound $T_i - T_{i-1} \geq T_0 (b - 1) b^{i-1} - 1$ for $i>0$, thanks to Definition~\ref{def:geomSequence},
%             and we can use the hypothesis on $\cA$ for each regret term.
%             Additionally, we have
%             $(T_i - T_{i-1})^{\gamma} \geq (T_0 (b - 1) b^{i-1} - 1)^{\gamma} \geq (T_0 (b-1))^{\gamma} (b^{i-1})^{\gamma} - 1$ by Lemma~\ref{lem:expoInequality} (Eq.~\eqref{eq:expoInequality3}, as $b>1$ and $0<\gamma<1$), thus
%         }
%         &\geq \sum_{i=0}^{L_T-1} c (T_i - T_{i-1})^{\gamma}
%         \geq c T_0^{\gamma} + c T_0^{\gamma} (b-1)^{\gamma} \sum_{i=1}^{L_T-1} (b^{i-1})^{\gamma} - c \sum_{i=1}^{L_T-1} 1 \\
%         &\geq c T_0^{\gamma} + c T_0^{\gamma} (b-1)^{\gamma} \sum_{i=0}^{L_T-2} (b^i)^{\gamma} - c (L_T - 1) \\
%         \intertext{
%             We have $\sum_{i=0}^{L_T-1} (b^i)^{\gamma} = \frac{(b^{L_T-1})^{\gamma} - 1}{b^{\gamma} - 1}$ thanks to a geometric sum (with $\gamma>0$) and thus
%         }
%         &\geq c T_0^{\gamma} + c T_0^{\gamma} (b-1)^{\gamma} \frac{(b^{L_T-1})^{\gamma} - 1}{b^{\gamma} - 1} + c (1 - L_T)
%         \intertext{
%             Thanks to Definition~\ref{def:geomSequence}, $b^{L_T - 1}$ satisfies $b^{L_T - 1} \geq \frac{1}{b} \frac{T}{T_0}$.
%             Let $g_0(T) := c T_0^{\gamma} \left(\frac{(b - 1)^{\gamma}}{b^{\gamma} - 1} - 1\right) + c (L_T - 1) = \bigO{1} + \bigO{\log_b(\frac{T}{T_0})} = \bigO{\log T} = \smallO{T^{\gamma}}$ and $g_0(T) > 0$, then we have
%         }
%         &\geq c \frac{(b - 1)^{\gamma}}{b^{\gamma}(b^{\gamma} - 1)} T^{\gamma} - \left[ c T_0^{\gamma} \left( \frac{(b - 1)^{\gamma}}{b^{\gamma} - 1} - 1 \right) + c (L_T - 1) \right].
%     \end{align*}
%     We obtain as announced,
%     $R_T(\cA') \geq \ell_0(b) \; c T^{\gamma} + g_0(T)$ with $g_0(T) = \bigO{\log T} = \smallO{T^{\gamma}}$.
% \end{proof}

% \paragraph{Maximizing the constant overhead?}
% %
% To maximize\footnote{For the largest possible lower bound, we try to maximize the constant overhead in the lower bound.} $\ell_0(\gamma, b) := \frac{(b - 1)^{\gamma}}{b^{\gamma}(b^{\gamma} - 1)} > 0$, we fix $\gamma$ and study the function $h: b \mapsto \ell_0(\gamma, b)$.
% The function $h$ is of class $\cC^1$ on $(1, \infty)$ and $h(b) \to +\infty$ for $b\to 1^+$ and $h(b) \to 0$ for $b\to\infty$.
% Moreover $h'(b) = - \gamma \frac{(b - 1)^{\gamma - 1}}{b^{\gamma + 1}\left(b^{\gamma} - 1\right)^{2}}  \left(- 2 b^{\gamma} + b^{\gamma + 1} + 1\right)$
% has the same sign as $f(b) := - \left(- 2 b^{\gamma} + b^{\gamma + 1} + 1\right)$.
% The function $f$ is of class $\cC^1$, with $f(1) = 0$ and  $f'(b) = - (\gamma+1)(b - \frac{2\gamma}{\gamma+1}) b^{\gamma-1}$, and as $0 < \gamma < 1$, $\frac{2\gamma}{\gamma+1} < 1$ so $f'(b) < 0$ for all $b>1$.
% Thus $f$ is decreasing, and $\forall b>1, f(b) < f(1) = 0$.
% So $h'$ has a negative sign, and this allows to conclude that $h$ is decreasing,
% and so $b\mapsto \ell_0(\gamma, b)$ has no global maximum at fixed $\gamma$,
% and $\ell_0\to\infty$ if $b \to 1^+$.
% %
% % is negative for any $b>1$ as $b + \sqrt{b} - 1 > 1 > 0$, so $h$ is increasing.

% \paragraph{Relationship with the upper bound.}
% %
% For any $b>1$, we compare $\ell_0(\gamma, b)$ with $\ell(\gamma, b)$
% and we see that, interestingly, $\ell_0(\gamma, b) = \ell_2(\gamma, b) / b^{2\gamma}$,
% with $\ell_2(\gamma, b)$ from Theorem~\ref{thm:keep45}.
% %
% % For example, $b=2$ gives $\ell_0(2) = \frac{\sqrt{2}}{2}+1 \simeq 1.71$,
% % that has to be compared to the constant overhead $\ell(\frac{1}{2}, 2) = \frac{\sqrt{2}}{\sqrt{2}-1} = 2 + \sqrt{2} = 2 \ell_0(2) \simeq 3.41$ obtained for the regret upper bound in Proposition~\ref{cor:keep4}.
% %
% For the particular case of $\gamma=\frac{1}{2}$, this lower bound also leads to another interesting remark:
% if $b$ is chosen to minimize the overhead in the upper bound (Theorem~\ref{thm:keep45}, $b^*(\frac{1}{2})=\frac{3+\sqrt{5}}{2}$), then this lower bound gives $\ell_0(\frac{1}{2}, b^*(\frac{1}{2})) = 1 + \frac{\sqrt{2}}{2} \simeq 1.71 > 1$, which proves that this choice of geometric doubling trick cannot be used to conserve an optimal algorithm,
% \ie, the constant overhead cannot be made as close to $1$ as we want.

\hr{}

\section{An Efficient Heuristic, the Doubling Trick with ``No Restart''}\label{sec:DTnr}

\TODO{Should we keep it? Yes I think, because it's important to show that we thought about it!}

Let $\cA' := \DTnr(\cA, (T_i)_{i\in\N})$ denotes the following Algorithm~\ref{algo:DTnr}.
The only difference with $\DTr$ (Algorithm~\ref{algo:DTr}) is that the history from all the steps from $t=1$ to $t=T_i$ is used to reinitialize the new algorithm $\cA^{(i)}$.
To be more precise, this means that a fresh algorithm $\cA^{(i)}$ is created, and then fed with successive observations $(A'(s), Y_{A'(s), s})$ for all $1 \leq s < t$, like if it was playing \emph{from the beginning}.
Note that $A^{(i)}$ could have chosen a different path of actions, but we give it the observations from the previous plays of $\cA'$.

This obviously cannot be applied to any kind of algorithm $\cA$, and for instance any algorithm based on arm elimination (\eg, Explore-Then-Commit approaches like in \cite{Garivier2016ETC}) will most surely fail with this approach.
%
This second doubling trick algorithm $\DTnr$ can be applied in practice if $\cA$ is index-based and uses the horizon $T$ as a simple numerical parameter in its indexes,
like it is the case for instance for
\AFHG{} (cf. Eq.~\eqref{eq:index_AFHG}).
or \KLUCBpp{} (cf. Eq.~\eqref{eq:index_KLUCBpp}).

\vspace*{5pt}  % XXX remove if problem
% \centering
% Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
% Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
% \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
\begin{algorithm}[H]
        % XXX Options
        \LinesNumbered  % XXX Option to number the line
        \DontPrintSemicolon
        % \RestyleAlgo{boxed}
        % XXX Input, data and output
        \KwIn{Bandit algorithm $\cA$, and sequence $(T_i)_{i\in\N}$\;}
        % \KwData{Data}
        % \KwResult{Result}
        \BlankLine
        % XXX Algorithm
        Let $i = 0$, and initialize algorithm $\cA^{(0)} = \cA_{T_0}$.\\
        \For{$t = 1, \dots, T$}{
            \If(\tcp*[f]{Next horizon $T_{i+1}$ from the sequence}){
                $t > T_i$
            }{
                Let $i = i + 1$.\;
                Initialize algorithm $\cA^{(i)} = A_{T_i - T_{i-1}}$.\;
                Update internal memory of $\cA^{(i)}$ with the history of plays and rewards from $\cA_{i-1}$.
            }
            Play with $\cA^{(i)}$: play arm $A'(t) := A^{(i)}(t - T_i)$, observe reward $r(t) = Y_{A'(t), t}$. \nllabel{line:internalTimer_DTnr}
        }
        \caption{The Non-Restarting Doubling Trick Algorithm, $\cA' = \DTnr(\cA, (T_i)_{i\in\N})$.}
\label{algo:DTnr}
\end{algorithm}
% \vspace*{-5pt}  % XXX remove if problem
\vspace*{10pt}  % XXX remove if problem

% We also have the same result for the second algorithm with no fresh restart.
% XXX That's one of the non-trivial part, to justify properly the following decomposition.
% XXX Actually I do not think we can prove this, for the first sequences, there is no argument to say that a non-fresh restart will perform well...

However, it is much harder to state any  theoretical result on this heuristic $\DTnr$.
We conjecture that a regret upper bound similar to \eqref{eq:decompositionIneq_UB} from Lemma~\ref{lem:decomposition} could still be obtained, but it is still an open problem that the authors do not know how to tackle for a generic algorithm.
%
The intuition is that starting $\cA^{(i)}$ with some previous observations from the (same) \iid{} process $(Y_{k,s})_{k\in\{1,\dots,K\},s\in\N}$ can only improve the performance of $\cA^{(i)}$,
and lead to a smaller regret on the interval $\{T_{i},\dots,T_{i+1}-1\}$.

% \begin{conjecture}[Regret Decomposition for $\DTnr$]\label{conj:decompositionDTnr}
%
%     For a fixed (but unknown) horizon $T$ and a sequence $(T_i)_{i\in\N}$, \emph{we expect to also have}
%     % \label{eq:decompositionDTnr}
%     \begin{align*}
%         \sum_{i=0}^{L_T-1} R_{T_i - T_{i-1}}(\cA_{T = T_i - T_{i-1}})
%         &\underset{(LB)}{\leq}
%         R_T(\DTnr(\cA, (T_i)_{i\in\N}))
%         &\underset{(UB)}{\leq}
%         \sum_{i=0}^{L_T} R_{T_i - T_{i-1}}(\cA_{T = T_i - T_{i-1}}).
%     \end{align*}
% \end{conjecture}
% % \begin{proof}\label{proof:decompositionDTnr}
% %     \FIXME{DO IT}
% % \end{proof}

\hr{}

% -----------------------------------------------------------------
\section{Basic but Useful Results}\label{sec:otherproofs}

All the logarithm $\log$ are taken in base $\e = \exp(1)$ (\ie, natural logarithm $\log$, but $\log$ is preferred for readability).
Logarithms in a basis $b > 1$ are denoted $\log_b(x) := \frac{\log x}{\log b}$, for any $x\in\R$, $x>0$.

We remind that $\lfloor x \rfloor$ denotes the integer part of $x\in\R$, and for $x>0$, that is
the unique integer $i$ such that $i \leq x < i + 1$.
The only property we use is its definition and the fact that $\lfloor x \rfloor \leq x$.
We also define $\lceil x \rceil := 1 + \lfloor x \rfloor$ for $x \geq 0$, which is the unique integer $j$ such that $j-1 \leq x < j$.

\subsection{Weighted Geometric Inequality}

% We prove some inequalities that control a sum of terms by a constant factor times the next term of the sum, \ie,
% %
% $\sum_{i=0}^{n-1} f(i) \leq C f(n)$ for a function $f$.
% % and we try to be as tight as possible on the constant $C$.
% %
% We focus on inequalities with this form,
% as it directly gives $\sum_{i=0}^{n} f(i) \leq (1 + C) f(n)$.

% \begin{lemma}[Geometric Inequality]\label{lem:geomIneq}
%     For any $n \in \N^*$, $b > 1$ and $0<\delta$,
%     \begin{equation}
%         \sum_{i=0}^{n-1} (b^i)^{\delta} \leq
%         \frac{1}{b^{\delta} - 1} (b^n)^{\delta}
%         = \frac{b^{\delta}}{b^{\delta} - 1} (b^{n-1})^{\delta}.
%     \end{equation}

%     % can we say this constant is the best possible one? Yes for this one
%     And note that the leading constant $1/(b^{\delta} - 1)$ is optimal.
% \end{lemma}
% \begin{proof}\label{proof:geomIneq}
%     It is simply the sum of a geometric sequence, for $q := b^{\delta} > 1$, as $(b^i)^{\delta} = (b^{\delta})^i$.
%     % \begin{equation*}
%     %     \sum_{i=0}^{n-1} (b^i)^{\delta}
%     %     = \sum_{i=0}^{n-1} q^i = \frac{q^n - 1}{q - 1} \leq \frac{1}{q - 1} q^n
%     %     = \frac{1}{b^{\delta} - 1} (b^n)^{\delta}.
%     % \end{equation*}
%     The leading constant is optimal, as when $n\to\infty$
%     the two sides are equivalent (\ie, $\frac{q^n - 1}{q^n} \to 1$).
%     % because $-1/q^n \to 0$.
% \end{proof}

% This second inequality is more generic,
% but we can no longer state that the constant on the right hand side is the best possible one.
% XXX Interesting remark
% Also note that it is looser than a geometric sum, as for $b>1$, $\delta>0$, $b^{\delta}>1$ and for any $x>1$, $0 < \log(x) \leq x - 1$ so $0 \leq \frac{1}{x-1} \leq \frac{1}{\log(x)}$.

\begin{lemma}[\emph{Weighted} Geometric Inequality]\label{lem:GeomWeightedSumIneq}
    For any $n \in \N^*$, $b > 1$ and $\delta > 0$,
    and if $f$ is a function of class $\cC^1$, non-decreasing and non-negative on $[0, \infty)$,
    we have
    \begin{equation}
        \sum_{i=0}^{n} f(i) (b^i)^{\delta} \leq
        \frac{b^{\delta}}{b^{\delta} - 1} f(n) (b^n)^{\delta}.
    \end{equation}

    % \FIXME{can we say this constant is the best possible one? Not sure for this one}
\end{lemma}
\begin{proof}\label{proof:ineq2}
    % Let $g(t) := f(t) (b^t)^{\delta} = u'(t) f(t)$,
    % for $u'(t) = (b^{\delta})^t$.
    % $f$ and $u'$ are both non-decreasing and non-negative, so is $g$, and thus
    % Lemma~\ref{lem:sumIntegralIneq} gives
    % $\sum\limits_{i=0}^{n-1} g(i) \leq \int_0^n g(t) \dt$.
    % %
    % And $u(t) = \frac{1}{\log(b^{\delta})} (b^{\delta})^t = \frac{1}{\delta \log b} (b^{\delta})^t$
    % and $f'(t)$ exists and is continuous, by assumption on $f$.
    % We verify that $u$ and $f$ are both non-negative and increasing (because $b^{\delta}>1$ as $b>1$ and $\delta>0$ so $\frac{1}{\log(b^{\delta}} > 0$, and by assumption for $f$),
    % so we can apply Lemma~\ref{lem:IPI} and conclude
    % $$ \sum\limits_{i=0}^{n-1} f(i) (b^i)^{\delta} \leq \int_0^n g(t) \dt \leq u(n) f(n) = \frac{1}{\delta \log(b)} f(n) (b^n)^{\delta}.$$
    %
    By hypothesis, $f$ is non-decreasing, so $\forall i\in\{0,\dots,n\}, f(i) \leq f(n)$,
    and so by using the sum of a geometric sequence, we have
    \begin{align*}
        \sum_{i=0}^{n} f(i) (b^i)^{\delta}
        \leq f(n) \left( \sum_{i=0}^{n} (b^i)^{\delta} \right)
        \leq f(n) \frac{1}{b^{\delta} - 1} (b^{n+1})^{\delta}
        = \frac{b^{\delta}}{b^{\delta} - 1} \left( f(n) (b^n)^{\delta} \right).
    \end{align*}
    %
    $f(i) = 1$ gives the geometric inequality.
    %
    Note that if we make $\delta\to0$, the left sum converges to $\sum_{i=0}^{n-1} f(i)$ and the right term diverges to $+\infty$, making this inequality completely uninformative.
\end{proof}

% \vspace*{-10pt}
% \begin{corollary}\label{cor:SumIneq3}
%     For any $\delta>0$ and $0 < \gamma < 1$,
%     with
%     $f(i)=i^{\gamma}$ or
%     $f(i)=(a^{b^i})^{\gamma}$ (for $a>1$, $b>1$),
%     % gives respectively
%     \begin{align}
%         \sum_{i=0}^{n} (b^i)^{\gamma} i^{\delta} &\leq \frac{b^{\gamma}}{b^{\gamma} - 1} (b^n)^{\gamma} n^{\delta}, \label{eq:sum_igamma_bidelta}\\
%         \sum_{i=0}^{n} (a^{b^i})^{\gamma} (b^i)^{\delta} &\leq \frac{b^{\delta}}{b^{\delta} - 1} (a^{b^n})^{\gamma} (b^n)^{\delta}.\label{eq:sum_abigamma_bidelta}
%     \end{align}
%     % And by isolating the last term, we also have
%     % \begin{align}
%     %     \sum_{i=0}^{n} (a^{b^i})^{\gamma} (b^i)^{\delta} &\leq \left(1 + \frac{1}{\delta \log b}\right) (a^{b^n})^{\gamma} (b^n)^{\delta}.\label{eq:sum_abigamma_bidelta2}
%     % \end{align}
%     % \FIXME{Optimal constant ? Numerically I observe not, but I do not know how to prove anything better!}
% \end{corollary}

% Note that the $(b^i)^{\delta}$ term is needed in
% Eqs.~\eqref{eq:sum_igamma_bidelta}
% and \eqref{eq:sum_abigamma_bidelta}
% (\ie, $\delta>0$ is necessary), otherwise no inequality holds for a generic term $f(i)$.
% For instance, $\sum_{i=0}^{n-1} \sqrt{i} \underset{n\to\infty}{\sim} 2/3 \; n^{3/2}$ so no constant $C$ exists\footnote{In fact, the sequence $(\sqrt{i})_{i\in\N}$ does not grow fast enough for this $C$ to exist.} such that $\sum_{i=0}^{n-1} \sqrt{i} \leq C \sqrt{n}$.
% % https://www.wolframalpha.com/input/?i=HarmonicNumber%5Bn,-1%2F2%5D
% % Lemma~\ref{lem:SumInequalityUsingEi}, proved below, extends~\eqref{eq:sum_abigamma_bidelta} to the case of $\delta=0$.

\subsection{Another Sum Inequality}

This second result is similar to the previous one but for a ``doubly exponential'' sequence, \ie, $a^{b^i}$,
as it also bounds a sum of increasing terms by a constant times its last term.

\begin{lemma}\label{lem:DoubleExpSumIneq}
    For any $n \in \N^*$, $a > 1$, $b > 1$ and $\gamma > 0$,
    we have
    \begin{equation}
        \sum_{i=0}^{n} (a^{b^i})^{\gamma} \leq
        a^{\gamma} + \left(1 + \frac{1}{(\log(a)) (\log(b^{\gamma}))}\right)
        (a^{b^n})^{\gamma}
        = \bigO{(a^{b^n})^{\gamma}}.
    \end{equation}
\end{lemma}
\begin{proof}\label{proof:DoubleExpSumIneq}
    We first isolate both the first and last term in the sum and focus on the from $i=1$ sum up to $i=n-1$.
    As the function $t \mapsto (a^{b^t})^{\gamma}$ is increasing for $t \geq 1$,
    we use a sum-integral inequality,
    and then the change of variable $u := \gamma b^t$ is of class $\cC^1$,
    of Jacobian $\dt = \frac{1}{\log b} \frac{\du}{u}$,
    gives
    \begin{align*}
        \sum_{i=1}^{n-1} (a^{b^i})^{\gamma}
        &\leq \int_{1}^{n} a^{\gamma b^t} \dt
        \leq \frac{1}{\log(b^{\gamma})} \int_{\gamma b}^{\gamma b^n} \frac{a^{u}}{u} \du \\
        \intertext{
            Now for $u\geq1$, observe that $\frac{a^{u}}{u} \leq a^{u}$,
            and as $\gamma b > 1$, we have
        }
        &\leq \frac{1}{\log(b^{\gamma})} \int_{\gamma b}^{\gamma b^n} a^{u} \du
        \leq \frac{1}{\log(b^{\gamma})} \frac{1}{\log(a)} a^{\gamma b^n}
        = \frac{1}{(\log(a)) (\log(b^{\gamma}))} {(a^{b^n})}^{\gamma}.
    \end{align*}
    Finally, we obtain as desired,
    $\sum\limits_{i=0}^{n} (a^{b^i})^{\gamma} \leq
    a^{\gamma} + (a^{b^n})^{\gamma} + \frac{1}{(\log(a)) (\log(b^{\gamma}))} {(a^{b^n})}^{\gamma}$.
\end{proof}

\subsection{Basic Functional Inequalities}

These functional inequalities are used in the proof of the main theorems.

% \begin{lemma}[Square Root Inequality]\label{lem:expoInequality}
%     For any $x,y \geq 0$,
%     \begin{equation}
%         \sqrt{x + y} \leq \sqrt{x} + \sqrt{y}.
%     \end{equation}
% \end{lemma}
% \begin{proof}\label{proof:expoInequality}
%     As $x,y\geq0$, $2\sqrt{x}\sqrt{y} \geq 0$,
%     so $x + y \leq x + y + 2\sqrt{x}\sqrt{y} = (\sqrt{x} + \sqrt{y})^2$.
%     The function $t\mapsto \sqrt{t}$ is non-decreasing, so
%     taking the square root on both sides gives,
%     $\sqrt{x + y} \leq \sqrt{x} + \sqrt{y}$.
% \end{proof}

% We also have a more generic result.

\begin{lemma}[Generalized Square Root Inequalities]\label{lem:expoInequality}
    For any $x,y \geq 0$ and $0 < \delta < 1$,
    \begin{equation}\label{eq:expoInequality1}
        (x + y)^{\delta} \leq x^{\delta} + y^{\delta}.
    \end{equation}
    % Conversely, if $\delta>1$,
    % \begin{equation}\label{eq:expoInequality2}
    %     (x + y)^{\delta} \geq x^{\delta} + y^{\delta}.
    % \end{equation}
    % \FIXME{I'm not sure I use this second inequality, just remove it if not used...}
    And conversely for any $0 < \delta < 1$, and $x,y \geq 0$, if $x \geq y$ then
    \begin{equation}\label{eq:expoInequality3}
        (x - y)^{\delta} \geq x^{\delta} - y^{\delta}.
    \end{equation}
\end{lemma}
\begin{proof}\label{proof:expoInequality}
    Fix $y\geq0$. Let $f(x):=(x+y)^{\delta} - (x^{\delta} + y^{\delta})$ for $x\geq0$.
    First, $f(0)= y^{\delta} - y^{\delta} = 0$,
    and as $\delta>0$, $f$ is differentiable on $[0, \infty)$, with
    $f'(x) = (\log \delta)(x+y)^{\delta} - (\log \delta)x^{\delta} = (\log \delta)( (x+y)^{\delta} - x^{\delta})$,
    and as $\delta < 1$, $\log\delta < 0$, and $(x+y)^{\delta} \geq x^{\delta}$, so $f'(x) \leq 0$ for any $x\geq0$.
    %
    Therefore, $f$ is non-increasing, and so $\forall x\geq0, f(x) \leq f(0) = 0$, so $f$ is non-positive, giving the desired inequality (for any $y\geq0$ and any $x\geq0$).

    % The same proof works for $\delta \geq 1$, but with $\log \delta \geq 0$ so it gives the converse inequality.

    The second inequality is a direct application of the first one.
    Assume $x \geq y$, and let $x' = x - y \geq 0$, then
    $(x' + y)^{\delta} \leq (x')^{\delta} + y^{\delta}$.
    This gives $(x-y)^{\delta} = (x')^{\delta} \leq (x'+y)^{\delta} - y^{\delta} = x^{\delta} - y^{\delta}$.
    %  as desired for Eq.~\eqref{eq:expoInequality3}.
    %
    % For the third inequality, we first prove it for $y=1$. Fix $\delta>0$, and let $g(x):=(x-1)^{\delta} - (x^{\delta} - 1)$ for $x\geq1$.
    % $g$ is of class $\cC^1$ on $[1, \infty)$, and $g'(x) = \delta \left[ (x-1)^{\delta-1} - x^{\delta-1}\right] \geq 0$ because $t \mapsto t^{\delta-1}$ is decreasing on $(1, \infty)$ as $\delta-1 < 0$.
    % Thus $g$ is non-decreasing, and so $\forall x\geq1, g(x) \geq g(1)=0$.
    % So this gives $(x - 1)^{\delta} \geq x^{\delta} - 1.$
    % Now if $y\neq1$, just write $(x - y)^{\delta} = y^{\delta} (\frac{x}{y} - 1)^{\delta}$ and $\frac{x}{y} \geq 1$ as $x \geq y$, so applying the special case of $y=1$ gives the desired inequality.
\end{proof}

% \FIXME{I'm not sure I use this inequality anymore... Remove if it is not needed!}

% \begin{lemma}[Log and Exponential Inequality]\label{lem:logAndSqrtInequality}
%     % For any $x > 0$ and $b>1$,
%     % \begin{equation}
%     %     \log(x) \leq \frac{2}{\e} \sqrt{x},
%     %     \;\;\text{and}\;\;
%     %     \log_b(x) \leq \frac{2}{\e \log(b)} \sqrt{x}.
%     % \end{equation}
%     For any $x > 0$, $b>1$
%     and $\gamma > 0$,
%     \begin{equation}
%         \log(x) \leq \frac{1}{\gamma \e} x^{\gamma},
%         \;\;\text{and}\;\;
%         \log_b(x) \leq \frac{1}{\gamma \e \log(b)} x^{\gamma}.
%     \end{equation}
%     Note that the even though the constant is the best possible one, this inequality is \emph{very} loose.
% \end{lemma}
% \begin{proof}\label{proof:logAndSqrtInequality}
%     Let $f(x) := \frac{\log x}{x^{\gamma}}$, then $f$ is of class $\cC^1$ on $(0, +\infty)$,
%     $f$ is negative on $(0, 1)$,
%     $f(1) = 0 = \lim_{x\to\infty} f(x)$,
%     and $f$ is positive on $(1, +\infty)$.
%     Moreover,
%     $f'(x) = \frac{1 - \gamma\log(x)}{x^{\gamma + 1}}$
%     has a monotonic sign, non-increasing, so $f$ is actually concave.
%     Therefore, it has a unique global maximum,
%     attained where $f'(x) = 0$.
%     The unique solution to this equation is $x = x_0 = \e^{1/\gamma}$,
%     so $f$ attains its global maximum at $x_0$,
%     for which $f(x_0) = \frac{\log(\e^{1/\gamma})}{(\e^{1/\gamma})^{\gamma}} = \frac{1}{\gamma\e}$.
%     To conclude, $\forall x>0$, $f(x) \leq f(x_0)$ gives $\log(x) \leq \frac{1}{\gamma\e} x^{\gamma}$.

%     Dividing both sides by $\log(b)$ gives the second inequality for $\log_b(x)$.
% \end{proof}

\vspace*{5pt}

\begin{lemma}[Bounding $\log(x-\Delta)$]\label{lem:logxm1_logx}
    Let $x_0 > 1$ and $0 < \Delta < x_0$ (\eg, $\Delta \leq 1$), then
    \begin{equation}\label{eq:logxm1_logx_1}
        \forall x \geq x_0,\;\;
        \frac{\log(x_0-\Delta)}{\log(x_0)} \log(x)
        \leq \log(x-\Delta)
        \leq \log(x).
    \end{equation}
    With $\Delta=1$, it implies that if $T_0>1$, $b>1$ satisfy $T_0(b-1)>1$,
    then for any $i\in\N$, we have
    \begin{equation}\label{eq:logxm1_logx_2}
        \log(T_0(b-1)b^i-1) \geq \frac{\log(T_0(b-1)-1)}{\log(T_0(b-1))} \log(T_0(b-1)b^i).
    \end{equation}
\end{lemma}
\begin{proof}\label{proof:logxm1_logx}
    Let $f(x) := \frac{\log(x-\Delta)}{\log(x)}$, defined for $x\geq x_0$.
    It is of class $\cC^1$, and by differentiating, we have $f'(x) = \frac{\log(x) - \log(x-\Delta)}{x (\log x)^2} > 0$ as $\log$ is increasing.
    So $f$ is increasing, and its minimum is attained at $x = x_0$,
    \ie, $\forall x\geq x_0, f(x) \geq f(x_0) = \frac{\log(x_0 - \Delta)}{\log(x_0)} > 0$, which gives Eq.~\eqref{eq:logxm1_logx_1}.

    The corollary is immediate but stated explicitly for clarity when used in page~\pageref{proof:LowerBoundGeom_log}.
\end{proof}

\vspace*{5pt}

% [$x \mapsto x^{\gamma}(\log x)^{\delta}$ is increasing]
\begin{lemma}\label{lem:tgammalogtdeltaIncreasing}
    For any $\gamma > 0$ and $\delta>0$,
    the function $f: x \mapsto x^{\gamma}(\log x)^{\delta}$ is increasing on $[1, \infty)$.
\end{lemma}
\begin{proof}\label{proof:tgammalogtdeltaIncreasing}
    $f$ is of class $\cC^1$ on $[1, \infty)$.
    %
    First, if $\gamma > 0$,
    we have $f'(x) = x^{\gamma-1}(\log x)^{\delta-1} \left( \delta + \gamma \log(x)\right)$.
    So $f'(x) > 0$ if and only if $\delta + \gamma \log(x) \geq 0$, that it $x \geq \exp(- \frac{\delta}{\gamma})$.
    But $x \geq 1$ and $ < 0$ so $f'(x)$ is always positive, and thus $f$ is increasing.
    %
    Then, if $\gamma = 0$,
    we have $f'(x) = \delta \frac{1}{x}(\log x)^{\delta-1} > 0$ as $x > 1$ gives $\log x > 0$
    and so $(\log x)^{\delta-1} > 0$.

    It is also true if $\gamma \geq 0$, $\delta \geq 0$ if not both are zero simultaneously.
\end{proof}

\subsection{Controlling an Unbounded Sum of Dominated Terms}

This Lemma is used in the proofs of our upper bounds (Theorems~\ref{thm:keep45} and \ref{thm:keep52}),
to handle the sum of $f(T_i)$ terms.
%
In particular, it can be applied to $(T_i)$ the geometric sequence and $g(t) = h(t) = t^{\gamma}$ or $g(t) = h(t) = t^{\gamma} (\log t)^{\delta}$ (for Theorem~\ref{thm:keep45})
for $\gamma>0$ and $\delta\geq0$;
or $(T_i)$ the exponential sequence, $g(t) = t^{\gamma} (\log t)^{\delta}$ and $h(t) =  t^{b\gamma} (\log t)^{\delta}$ (for Theorem~\ref{thm:keep52})
for $\gamma\geq0$ and $\delta\geq0$.
%
Note that it would be obvious if $L_T$ was bounded for $T\to\infty$, but a more careful analysis has to be given as $L_T \to \infty$.

\begin{lemma}\label{lem:ControlLemmaSumOfSmallO}
    Let $f$, $g$ and $h$ be three positive, diverging and non-decreasing functions on $[1, \infty)$, such that
    $f(t) = \smallO{g(t)}$ for $t\to\infty$.
    Let a non-decreasing diverging sequence $(T_i)_{i\in\N}$,
    and a diverging sequence $(L_T)_{T\in\N}$ (\ie, $T_i\to\infty$ for $i\to\infty$ and $L_T\to\infty$ if $T\to\infty)$,
    such that there exists a constant $c\geq0$ satisfying
    $\forall T\geq1, \sum\limits_{i=0}^{L_T} g(T_i) \leq c \times h(T)$.\\
    Then the (unbounded) sum of dominated terms $f(T_i)$ is still dominated by $h(T)$, \ie,
    \begin{align}\label{eq:ControlLemmaSumOfSmallO}
        f(t) \underset{T\to\infty}{=} \smallO{g(t)}
        \;\text{and}\;
        \exists c\geq0, \sum_{i=0}^{L_T} g(T_i) \underset{\forall T\geq1}{\leq} c \times h(T)
        \implies
        \sum_{i=0}^{L_T} f(T_i) \underset{T\to\infty}{=} \smallO{h(T)}.
    \end{align}
\end{lemma}
\begin{proof}\label{proof:ControlLemmaSumOfSmallO}
    By hypothesis, if $f$ is dominated by $g$, formally $f(t) = \smallO{g(t)}$, then there exists a positive function $\eps(t)$ such that $f(t) = g(t) \eps(t)$ and $\eps(t) \to 0$ for $t\to0$.
    Fix $\eta>0$, as small as we want, then there exists $T_{\eta} \geq 1$ such that
    $\forall t \geq T_{\eta}, \eps(t) \leq \eta$.
    %
    Let $i_{\eta}$ such that $T_{i_{\eta}-1} < T_{\eta} \leq T_{i_{\eta}}$.
    Now for any $T \geq T_{\eta}$ and large enough so that $L_T \geq T_{\eta}$,
    we can start to split the sum
    \begin{align*}
        \sum_{i=0}^{L_T} f(T_i)
        &= \sum_{i=0}^{i_{\eta} - 1} f(T_i) + \sum_{i=i_{\eta}}^{L_T} f(T_i) \\
        \intertext{
            The first sum is naively bounded by $i_{\eta} \times f(T_{i_{\eta} - 1})$ as $f$ is increasing,
            and for the second sum, for any $i\geq i_{\eta}$, $T_i \geq T_{\eta}$ and so $f(T_i) = \eps(T_i) g(T_i) \leq \eta g(T_i)$, thus
        }
        &\leq i_{\eta} f(T_{i_{\eta} - 1}) + \eta \times \left( \sum_{i=i_{\eta}}^{L_T} g(T_i) \right) \\
        \intertext{
            The sum is smaller than a sum on a larger interval, as $g(T_i) \geq 0$ for any $i$,
            and $f$ is increasing so
        }
        &\leq i_{\eta} f(T_{\eta}) + \eta \left( \sum_{i=0}^{L_T} g(T_i) \right) \\
        \intertext{
            But now, $f(T_{\eta}) \leq \eta g(T_{\eta})$ by hypothesis, and this sum is smaller than $c \times h(T)$ also by hypothesis
        }
        &\leq i_{\eta} \eta \times g(T_{\eta}) + \eta c\times  h(T)
        = \eta \left( i_{\eta}  g(T_{\eta}) + c\times  h(T) \right) \\
        \intertext{
            Finally, we use the hypothesis that $h(T)$ is diverging and as $\eta$ and $T_{\eta}$ are both fixed,
            there exists a $\widetilde{T_{\eta}} \geq T_{\eta}$ large enough so that
            $i_{\eta} g(T_{\eta}) \leq h(T)$ for all $T \geq \widetilde{T_{\eta}}$.
            And so we have finally proved that
        }
        &\forall \eta > 0,\;\;
        \exists \widetilde{T_{\eta}} \geq 1,\;\;
        \forall T \geq \widetilde{T_{\eta}},\;\;
        \sum_{i=0}^{L_T} f(T_i) \leq \eta (c+1) \times h(T).
    \end{align*}
    This concludes the proof and shows that $\sum\limits_{i=0}^{L_T} f(T_i) = \smallO{h(T)}$ as wanted.
\end{proof}

% -----------------------------------------------------------------
\section{Additional Experiments}\label{sec:otherExperiments}

We presents additional experiments, for Gaussian bandits and for the heuristic \DTnr{}.

\subsection{Experiments with Gaussian Bandits (with Known Variance)}\label{sub:experimentsGaussianBandits}

We include here another figure for experiments on Gaussian bandits, see Figure~\ref{fig:gaussianBandits_DoublingTrick_Restart_fixedProblem2}.
% \FIXME{Give some conclusion and message for these plots!}

%
% Regular plots of centralized regrets
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{0.49\textwidth}
    \includegraphics[width=1.00\textwidth]{figures/main____env1-1_6979515539977716717.pdf}
    % \end{subfigure}
    % % ~
    % \begin{subfigure}[!h]{0.49\textwidth}
    %   \includegraphics[width=1.00\textwidth]{XXX.pdf}
    % \end{subfigure}
    \caption{Regret for \DT, for $K=9$ Gaussian arms $\cN(\mu, 1)$, horizon $T=45678$, $n=1000$ repetitions and $\boldsymbol{\mu}$ uniformly spaced in $[-5,5]^K$. On ``easy'' problems like this one, both \UCB{} and \AFHG{} perform similarly and attain near constant regret (identifying the best Gaussian arm is very easy here as they are sufficiently distinct). Each doubling trick also appear to attain near constant regret, but geometric doubling ($b=2$) and slow exponential doubling ($b=1.1$) are slower to converge and thus less efficient.}
    \label{fig:gaussianBandits_DoublingTrick_Restart_fixedProblem2}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

% We include here some experiments on truncated Gaussian Bandits, \ie, bandit problems were each arm has a Gaussian distribution
% % ${\mathcal{N}}_{[a,b]}(\mu, \sigma)$,
% ${\mathcal{N}}(\mu, \sigma)$,
% with known and fixed $\sigma$.
% % and truncated to the interval $[a, b]$.
% In practice, we use
% % $[a, b] = [0, 1]$, $\sigma=0.05$ and $\mu\in[0,1]$.
% $\sigma=1$ and $\mu\in[-5, 5]$.
% %
% We consider a similar setup as the one used for the experiments presented above in Section~\ref{sec:experiments}.

% \FIXME{Copy some figures, include them, talk about it quickly}

\subsection{Experiments with \DTnr}\label{sub:experimentsDTnr}

As mentioned previously, the \DTnr{} algorithm (Algorithm~\ref{algo:DTnr})
is only an heuristic so far, as no theoretical guarantee was proved for it.
%
For the sake of completeness, we also include results from numerical experiments with it, to compare its performance against the ``with restart'' version \DTr.

As expected, \DTnr{} enjoys much better empirical performance,
and in Figs.~\ref{fig:bernoulliBandits_DoublingTrick_noRestart_BayesianProblem} and \ref{fig:bernoulliBandits_DoublingTrick_noRestart_fixedProblem} we see that a geometric or a slow exponential doubling trick with no restart with \KLUCBpp{} can outperform \KLUCB{} and perform similarly to the non-anytime \KLUCBpp.
%
But as observed before, the regret blows up after the beginning of each new sequence if the doubling sequence increase too fast (\eg, exponential doubling).
Despite what is proved theoretically in Theorem~\ref{thm:LowerBoundGeom_log}, here we observe that the geometric doubling is the only one to be slow enough to be efficient.

%
% Regular plots of centralized regrets
%
\begin{figure}[H]
    \centering
    % \begin{subfigure}[H]{0.49\textwidth}
    \includegraphics[width=0.92\textwidth]{figures/main____env1-1_5964629015089571121.pdf}
    % \end{subfigure}
    % % ~
    % \begin{subfigure}[H]{0.49\textwidth}
    %   \includegraphics[width=0.92\textwidth]{XXX.pdf}
    % \end{subfigure}
    \caption{Regret for $K=9$ Bernoulli arms, horizon $T=45678$, $n=1000$ repetitions and $\boldsymbol{\mu}$ taken uniformly in $[0,1]^K$, for \DTnr. Geometric doubling (\eg, $b=2$) and slow exponential doubling (\eg, $b=1.1$) are too slow, and short first sequences make the regret blow up in the beginning of the experiment. At $t=40000$ we see clearly the effect of a new sequence for the best doubling trick ($T_i = 200 \times 2^i$). As expected, \KLUCBpp{} outperforms \KLUCB, and if the doubling sequence is growing fast enough then $\DTnr(\KLUCBpp)$ can perform as well as \KLUCB.}
    \label{fig:bernoulliBandits_DoublingTrick_noRestart_BayesianProblem}
    \vspace*{-15pt}  % XXX remove if problem
\end{figure}
%
%
% Regular plots of centralized regrets
%
\begin{figure}[H]
    \centering
    % \begin{subfigure}[H]{0.49\textwidth}
    \includegraphics[width=0.92\textwidth]{figures/main____env1-1_5972568793654673752.pdf}
    % \end{subfigure}
    % % ~
    % \begin{subfigure}[H]{0.49\textwidth}
    %   \includegraphics[width=0.92\textwidth]{XXX.pdf}
    % \end{subfigure}
    \caption{$K=9$ Bernoulli arms with $\boldsymbol{\mu}$ evenly spaced in $[0,1]^K$. On easy problems like this one, both \KLUCB{} and \KLUCBpp{} are very efficient, and here the geometric allows the \DTnr{} anytime version of \KLUCBpp{} to outperform both \KLUCB{} and \KLUCBpp.}
    \label{fig:bernoulliBandits_DoublingTrick_noRestart_fixedProblem}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

\hr{}

\section{New Lower Bound in the Intermediate Regime}

\TODO{Lilian: remove all the explanations that are now useless as I extended them in a section above.}
\TODO{Q : for all choices of $f$, do we have that $T_i - T_{i-1}$ (which is increasing) tends to infinity?}

Defining $T_i = \lfloor \exp(\alpha \exp(f(i)))\rfloor$ where $f$ is some increasing function that tends to infinity. The last term operator is
\begin{eqnarray*}
    L_T & = & \min \{ i \in \N : T_{i-1} < T \leq T_i \} \\
    &= & \min \{ i \in \N : f(i) \geq \log\left(\log(T)/\alpha\right)\}.
\end{eqnarray*}
Therefore, it holds that $L_T \geq f^{-1}\left(\log\left(\log(T)/\alpha\right)\right)$.

The usual geometric doubling trick corresponds to $f(i) = \log(i)$, while the exponential doubling trick corresponds to $f$ being a linear function. We provide lower bounds in this section showing that some type of doubling sequence cannot preserve logarithmic bounds.

\begin{theorem}\label{thm:LowerBounds}
    Fix $c$ some constant and let $\cA$ be an online algorithm against a stochastic adversary satisfying
    \[ \forall T \in \N,  R_T(\cA_T) \geq c \; (\log T).\]
    Fix $\alpha \in \R^+$, $b \in (0,1]$ and let $f_b(t) = t^b$. Let $(T_i)_{i\in\N}$ be the doubling sequence associated to $f_b$ and $\alpha$ and denote by $\cA'$ the corresponding doubling algorithm. There exists $\tilde{T}_b$ such that
    \begin{equation}\label{eq:LowerBoundGeom_log_1}
        \forall T \geq \tilde{T}_b, \ \
        R_T(\cA') \geq
        c' \; \left(\log T\right)(\log \log T)^{\frac{1-b}{b}}.
    \end{equation}
\end{theorem}

\TODO{What if we have $\log(T)^{\delta}$ in place of $\log(T)$? the initial common lower bound may not be valid any more}

\TODO{Either fit the $f(t) =\log(t)$ bound in this theorem or state a specific result}

\begin{proof}

We will repeatedly use the fact that for all $x \geq 3$, $\log (x-1) \geq (1/2)\log(x)$. Let $n_1 \in \N^*$ such that for all $i \geq n_1$,
\[\exp\left(\alpha e^{f_b(i)}\right) - \exp\left(\alpha e^{f_b(i-1)}\right) \geq 3 \ \ \text{and} \ \ e^{f_b(i)} - e^{f_b(i-1)} \geq \frac{\log 3}{\alpha}\]
(this is possible as in the two expressions the left hand-side tends to infinity as $i$ tends to infinity).
Note that for all $i\geq 0$,
\begin{eqnarray*}
    T_{i} - T_{i-1} & \geq & \exp\left(\alpha e^{f_b(i)}\right) - \exp\left(\alpha e^{f_b(i-1)}\right) - 1.
\end{eqnarray*}

Using the lower bound \eqref{eq:decompositionIneq_LB} from Lemma~\ref{lem:decomposition} together with the assumption on algorithm $\cA$ yields, for all $T$ such that $L_T \geq n_1 + 1$,
\begin{eqnarray*}
    R_T(\cA') & \geq & c\sum_{i=0}^{L_T-1} (\log(T_i - T_{i-1})) \\
    & \geq & c\sum_{i=0}^{L_T-1} \log\left(\exp\left(\alpha e^{f_b(i)}\right) - \exp\left(\alpha e^{f_b(i-1)}\right) - 1\right)\\
    & \geq & \frac{c}{2}\sum_{i=n_1}^{L_T-1} \log\left(\exp\left(\alpha e^{f_b(i)}\right) - \exp\left(\alpha e^{f_b(i-1)}\right)\right)\\
    & \geq & \frac{c}{2}\sum_{i=n_1}^{L_T-1}\left[\alpha e^{f_b(i-1)} + \log \left(\exp\left(\alpha \left[e^{f_b(i)} - e^{f_b(i-1)}\right]\right) - 1\right)\right] \\
    & \geq & \frac{c}{2}\sum_{i=n_1}^{L_T-1}\left[\alpha e^{f_b(i-1)} + \frac{1}{2} \alpha \left(e^{f_b(i)} - e^{f_b(i-1)}\right)\right] \\
    & = & \frac{c}{2} \left[\sum_{i=n_1}^{L_T-1}\alpha e^{f_b(i-1)} + \frac{1}{2} \alpha \left(e^{f_b(L_T-1)} - e^{f_b(T_1-1)}\right)\right] \\
    & \geq & \frac{\alpha c}{4}\sum_{i=n_1-1}^{L_T-1} e^{f_b(i)} - \frac{c}{4}\alpha e^{f_b(n_1-1)} \\
    & \geq & \frac{\alpha c}{4}\underbrace{\int_{n_1}^{L_T-1} e^{f_b(t)} \dt}_{:=\cI_T} - \frac{c}{4}\alpha e^{f_b(n_1-1)}.
\end{eqnarray*}
In order to prove a lower bound for a specific sequence characterized by $\alpha$ and $f_b$, it is thus sufficient to lower bound the integral $\cI_T$ defined above.

Note that $L_T \geq \left(\log\left(\frac{\log(T)}{\alpha}\right)\right)^{\frac{1}{b}}$ hence
\begin{eqnarray*}
    \cI_T & \geq & \int_{n_1}^{\left(\log\left(\frac{\log(T)}{\alpha}\right)\right)^{\frac{1}{b}} - 1} e^{t^b} \dt =  \frac{1}{b}\int_{n_1^b}^{B_{T,b}}u^{\frac{1-b}{b}}e^u \du,
\end{eqnarray*}
where $B_{T,b}:=\left[\left(\log\left(\frac{\log(T)}{\alpha}\right)\right)^{\frac{1}{b}} - 1\right]^b$, using the change of variable $u = t^b$ (of class $\cC^1$).
Assume now that $T$ large enough so that $B_{T,b} \geq n_1+1$.
Using that $u \mapsto u^{\frac{1-b}{b}}e^u$ is non-decreasing and positive yields
\begin{eqnarray*}
    \cI_T & \geq & \frac{1}{b}\int_{B_{T,b}-1}^{B_{T,b}} u^{\frac{1-b}{b}}e^u \du \geq \frac{1}{b}(B_{T,b} - 1)^{\frac{1-b}{b}}e^{B_{T,b} - 1} \\
    & = & \frac{1}{b e \alpha} (\log T) \left(\log\left({\log T}/{\alpha}\right)\right)^{\frac{1-b}{b}} \times \underbrace{\left(\frac{B_{T,b} - 1}{\log (\log(T)/\alpha)}\right)^{\frac{1-b}{b}}}_{\cE_1(T)} \times \underbrace{e^{B_{T,b}  - \log (\log(T)/\alpha)}}_{\cE_2(T)}.
\end{eqnarray*}

As $B_{T,b} \sim \log (\log(T)/\alpha)$ when $T \rightarrow \infty$, $\cE_1(T) \sim 1$ and there exists some constant $\tilde{T}_1$
such that if $T$ is larger than $T_1$, $\cE_1(T) \geq \frac{1}{2}$.
As for $\cE_2(T)$, it can be rewritten
\[\cE_2(T) = \exp\left(h(\log(\log(T)/\alpha))\right), \ \ \text{with} \ \ h(u) = (u^{1/b} - 1)^b - u.\]

It can be shown that $h(u) \leq 0$ for all $u \geq 1$ and that $h$ is increasing.
Therefore, there exists a constant $c_2 < 0$ such that $h(u) \geq c_2$ for all $u \geq 1$.
Hence, for $T \geq \tilde{T}_2$ such that $\log (\log(T)/\alpha) \geq 1$, $\cE_2(T) \geq \exp(c_2)>0$.
It follows that for $T \geq \tilde{T}_0 = \max(\tilde{T}_1 ,\tilde{T}_2)$,
\begin{eqnarray*}
    \cI_T \geq \frac{e^{c_2}}{2b e \alpha} (\log T) \left(\log\left({\log T}/{\alpha}\right)\right)^{\frac{1-b}{b}}.
\end{eqnarray*}

% for $b=2$ integrating by parts yields
% \begin{eqnarray*}
%  \cI_T & = & 2\int_{\sqrt{T_1}}^{\sqrt{\left(\log\left(\frac{\log(T)}{\alpha}\right)\right)^2 - 1}} ue^{u} \dt\\
%  & = & 2\exp\left(\sqrt{\left(\log\left(\frac{\log(T)}{\alpha}\right)\right)^2 - 1}\right)\left[\sqrt{\left(\log\left(\frac{\log(T)}{\alpha}\right)\right)^2 - 1} -1\right]- 2(\sqrt{T_1} - 1) e^{\sqrt{T_1}}.
% \end{eqnarray*}

\end{proof}

\hr{}

\section{Some thoughts about an upper bound for the exponential doubling}\label{appendix:equivalence_Ti_and_TimTim1}

We study $f(t) = c t$ in this section and study what happens with an algorithm having a $T^{\gamma}$ regret. Note that setting $c=\log (b)$ and $\alpha = \log(a)$ we recovers our previous $T_i  = \lfloor a^{b^i}\rfloor$.

We convinced ourselves in Toulouse that we cannot improve the upper bound with our current technique of upper bounding everything by the last term and evaluating the last term, as this last term $(T_{L_T})^{\gamma}$ can be shown to be ``of order'' to $T^{\gamma e^c}$ (see details below).

One may think that this approach was sub-optimal and study instead the less crude upper bound
\[\sum_{t=0}^{L_T} (T_{i} - T_{i-1})^{\gamma}\]
We start by providing an equivalent of each term in this sum. First note that we will sometimes get rid of some floors when doing asymptotic developments as $\lfloor x_i \rfloor \sim x_i$ for any sequence $x_i \rightarrow \infty$. As $T_i - T_{i-1}  =  T_{i-1}\left(T_i/T_{i-1} - 1 \right) $, we first study
\begin{eqnarray*}
\frac{T_i}{T_{i-1}} & \sim & \exp\left(\alpha (e^{ci} - e^{c(i-1)}\right) = \exp\left(\alpha e^{c(i-1)}(e^c - 1)\right) \\
& = & \left(\exp\left(\alpha e^{c(i-1)}\right)\right)^{e^c -1} \sim (T_{i-1})^{e^c-1}
\end{eqnarray*}
Hence $T_{i}/T_{i-1} - 1 \sim (T_{i-1})^{e^{c}-1}$ and producting two equivalents yields
\[T_{i} - T_{i-1} \sim (T_{i-1})^{e^c} \sim \exp(\alpha e^c \exp(c(i-1)) = \exp(\alpha \exp(ci)) \sim T_i.\]
This actually shows that there is nothing to gain with using the tighter sum (the ``plot'' should thus have been increasing...)

More precisely, our analysis is tight in the following sense

\begin{lemma} For any exponential doubling sequence,
\[\sum_{i=0}^{n}T_i^{\gamma} \underset{n \rightarrow \infty}{=} \Theta(T_n^{\gamma})\]
\end{lemma}

\TODO{E: I am pretty sure we proved that in Toulouse}

If this is true, then I can prove the following:
\[\limsup_{T\rightarrow \infty} \frac{\log\left[\sum_{i=0}^{L_T} (T_i - T_{i-1})^{\gamma}\right]}{\log [T^{\gamma}]} = e^c\]
and
\[\liminf_{T \rightarrow \infty} \frac{\log\left[\sum_{i=0}^{L_{T}-1} (T_i - T_{i-1})^{\gamma}\right]}{\log [T^{\gamma}]} = e^{-c}\]
which shows that we cannot hope for better than our previous upper and lower bounds if we start from the Lemma~\ref{lem:decomposition}. And as for a non-anytime algorithm, we can never control the intermediate regret $R_t$ for $t\leq T$, we have not hope to be able to use what happens between $L_{T} - 1$ and $L_T$ hence Lemma~\ref{lem:decomposition} was not such a bad start...

\TODO{E: if needeed, I will imput the proof}

Maybe we need something slightly slower, like $f(i) = ci/\log(i)$, I will try to check numerically whether $T_{L_T} = \Theta(T)$ in that case (which is what would make the upper bound work...)

\end{document}

### 2019-03-25 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/plot_datafile_append_Uniform_vs_UCB_vs_TS.py

@@ -1,0 +1,167 @@
+#!/usr/bin/env python3
+# -*- coding: utf8 -*-
+
+""" Plot the datafile 'append' for an Object which uses the Uniform, or UCB, or TS algorithm.
+
+Quite hacky.
+"""
+
+import sys, os, os.path, glob
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.ticker import PercentFormatter
+import seaborn as sns
+
+
+def palette(nb):
+    """ Use a smart palette from seaborn, for nb different plots on the same figure.
+
+    - Ref: http://seaborn.pydata.org/generated/seaborn.hls_palette.html#seaborn.hls_palette
+    """
+    return sns.hls_palette(nb + 1)[:nb]
+
+
+def makemarkers(nb):
+    """ Give a list of cycling markers. See http://matplotlib.org/api/markers_api.html
+
+    .. note:: This what I consider the *optimal* sequence of markers, they are clearly differentiable one from another and all are pretty.
+
+    Examples:
+
+    >>> makemarkers(7)
+    ['o', 'D', 'v', 'p', '<', 's', '^']
+    >>> makemarkers(12)
+    ['o', 'D', 'v', 'p', '<', 's', '^', '*', 'h', '>', 'o', 'D']
+    """
+    allmarkers = ['o', 'D', 'v', 'p', '<', 's', '^', '*', 'h', '>']
+    longlist = allmarkers * (1 + int(nb / float(len(allmarkers))))  # Cycle the good number of time
+    return longlist[:nb]  # Truncate
+
+
+def parse_one_line(line):
+    """ Get one line of a file in this format
+        16:17:266:2864    3:4:194:2443
+
+    and return a couple of lists
+
+        [16,17,266,2864], [3,4,194,2443]
+    """
+    line_trial, line_success = line.split('    ')
+    one_trial = [int(s) for s in line_trial.split(':')]
+    one_success = [int(s) for s in line_success.split(':')]
+    return one_trial, one_success
+
+
+def read_data(filename):
+    with open(filename, 'r') as f:
+        lines = f.readlines()
+    horizon = len(lines)
+    # hack to read number of lines
+    nb_arm = lines[0].split('    ')[0].count(':') + 1
+    # create data
+    trials = np.zeros((horizon, nb_arm), dtype=int)
+    successes = np.zeros((horizon, nb_arm), dtype=int)
+    # read it and fill them
+    print("For file", filename)  # DEBUG
+    for i, line in enumerate(lines):
+        try:
[…]

### 2019-03-25 2-Chapters/6-Chapter/nonstatbandits.git/simulateForcedExploration.py

@@ -1,0 +1,133 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+import tqdm
+
+means_0 = [0.1, 0.5, 0.9]
+means_1 = [0.9, 0.5, 0.1]
+maxMean = max(max(means_0), max(means_1))
+nbArms = len(means_0)
+horizon = 1000
+
+def reward(arm, t):
+    if t <= horizon/2:
+        return np.random.binomial(1, means_0[arm])
+    else:
+        return np.random.binomial(1, means_1[arm])
+
+def simulateBandit(bandit, detect):
+    pulls = np.zeros(nbArms, dtype=int)
+    all_pulls = [ ]
+    rewards = np.zeros(nbArms)
+    all_rewards = [ [] for i in range(nbArms) ]
+    all_rewards_shared = [ ]
+    last_restart = 0
+    for t in range(horizon):
+        choice = bandit(pulls, rewards, t - last_restart)
+        pulls[choice] += 1
+        all_pulls.append(choice)
+        r = reward(choice, t)
+        rewards[choice] += r
+        all_rewards[choice].append(r)
+        all_rewards_shared.append(r)
+        # print(f"Time {t}, bandit chose {choice}, saw reward {r} (pulls = {pulls}, rewards = {rewards})")
+        if detect(pulls, rewards, all_rewards, choice, r, t, last_restart):
+            print(f"  CDP algorithm {detect.__name__} chose to restart at time {t}, previous restart was {last_restart}")
+            last_restart = t
+            pulls[:] = 0
+            rewards[:] = 0
+            all_rewards = [ [] for i in range(nbArms) ]
+    # now compute the history of regret
+    maxMeans = np.array(([max(means_0)] * (horizon//2)) + ([max(means_1)] * (horizon//2)))
+    all_rewards_shared = np.array(all_rewards_shared)
+    return np.cumsum(maxMeans - all_rewards_shared)
+    return maxMean * horizon - np.sum(rewards)
+
+
+def ucb(pulls, rewards, t, *args, **kwargs):
+    if np.any(pulls <= 0):
+        return np.random.choice(np.where(pulls <= 0)[0])
+    indexes = rewards/pulls + np.sqrt(np.log(t) / pulls)
+    return np.random.choice(np.where(np.isclose(indexes, np.max(indexes)))[0])
+
+
[…]

### 2019-03-29 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Update bibtex

@@
 We will continue investigating regret bounds for \Aggr,
-and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo17}).
+and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).
 

### 2019-03-29 2-Chapters/4-Chapter/chapter4.tex
Update bibtex

@@
 It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
-It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo17} or structured MAB \cite{Combes17} models and algorithms for this extension.
+It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
 

### 2019-03-29 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand

@@
 We assume that this network is constituted by two types of devices:
-on one hand, we have \emph{static} devices that operate in one channel\footnote{~Note that, for unlicensed bands, this definition also encompasses any device following a different standard or trying to establish communication with gateways of other networks.} in order to communicate with the gateway.
+On the one hand, we have \emph{static} devices that operate in one channel\footnote{~Note that, for unlicensed bands, this definition also encompasses any device following a different standard or trying to establish communication with gateways of other networks.} in order to communicate with the gateway.
 %
@@
 	\item
-	On one hand, in a congested network, when devices suffer from a large probability of collision on their first transmission (\ie, $p_c$ is not so small), then $p_{c1}\simeq p_c$ and so devices cannot really hope to reduce their collision probabilities even if the use a different channel for retransmission.
+	On the one hand, in a congested network, when devices suffer from a large probability of collision on their first transmission (\ie, $p_c$ is not so small), then $p_{c1}\simeq p_c$ and so devices cannot really hope to reduce their collision probabilities even if the use a different channel for retransmission.
 	\item

### 2019-03-29 2-Chapters/4-Chapter/chapter4.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand

@@
 %
-On one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
+On the one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
 On the other hand, it is quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.

### 2019-03-29 3-Appendices/2-Appendix/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.git/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand

@@
 These two shortcoming have been overcame recently in two different works.
-On one hand, the \MOSSAnytime{} algorithm \citep{Degenne16} is minimax optimal and anytime, but its problem-dependent regret does not improve that of \MOSS.
+On the one hand, the \MOSSAnytime{} algorithm \citep{Degenne16} is minimax optimal and anytime, but its problem-dependent regret does not improve that of \MOSS.
 On the other hand, the \KLUCBpp{} algorithm \citep{Menard17} is simultaneously minimax optimal and asymptotically optimal (\ie, it has the best problem-dependent constant $C(\bm\nu)$), but it is not anytime.
@@
 For a fixed $\gamma$ and $\delta$, minimizing $\ell(\gamma, \delta, T_0, b)$ does not always give a unique solution.
-On one hand, if $\gamma \gtrsim 0.384$, there is a unique solution $b^*(\gamma) > 1$ minimizing the $\frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}$ term, solution of $b^{\gamma+1} - 2b + 1 = 0$, but without a closed form if $\gamma$ is unknown.
+On the one hand, if $\gamma \gtrsim 0.384$, there is a unique solution $b^*(\gamma) > 1$ minimizing the $\frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}$ term, solution of $b^{\gamma+1} - 2b + 1 = 0$, but without a closed form if $\gamma$ is unknown.
 %

### 2019-03-29 2-Chapters/6-Chapter/chapter6.tex
Again lots of work on Chapter 6, see #14
Almost done for this chapter 6. Maybe I'll move all these « exploration » numerical experiments in the main text, but use smaller tables (explore less values?)
Using \begin{framed}...\end{framed} for all algorithms.

@@
 
-\paragraph{Abstract}
+\paragraph{Abstract.}
 
 In this last chapter, we study a generalization of the stationary multi-armed bandit model,
-to account for possible non-stationarity of the problems that could be tackled with MAB algorithms.
-Instead of dealing with changes happening possibly at every time step, like in adversarial models, we rather consider a finite number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
+to account for possible non-stationarity of the rewards.
+% that is, for problems that could be tackled with MAB algorithms.
+Instead of dealing with changes happening possibly at every time step, like in adversarial models, we rather consider a finite small number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
 
@@
 
-Thanks to these two non-asymptotic properties, the B-GLRT test can be combined with an efficient bandit policy, the \klUCB{} index policy, to propose an efficient algorithm for piece-wise stationary bandit problems, that we design \GLRklUCB.
+Thanks to these two non-asymptotic properties, the B-GLRT test can be combined with an efficient bandit policy, the \klUCB{} index policy, to propose an efficient algorithm for piece-wise stationary bandit problems, \GLRklUCB.
 Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
@@
 We give the first unified analysis of a piece-wise stationary bandit algorithm for the two variants, and our analysis is completely non-asymptotic and yield regret bounds with explicit constants.
-Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$ with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
+Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$, when the algorithm runs with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
 
@@
 
-\paragraph{Publications}
-
+\paragraph{Publications.}
+%
 This chapter is based on the following publications: \cite{Besson2019GLRT,Besson2019Gretsi}.
@@
 
-\TODOL{Rewrite and reorder most of this chapter, with more subsections, less paragraphs etc, and less copy-paste of our paper \cite{Besson2019GLRT}.}
+% \TODOL{Rewrite and reorder most of this chapter, with more subsections, less paragraphs etc, and less copy-paste of our paper \cite{Besson2019GLRT}.}
 
@@
 % % ----------------------------------------------------------------------------
-% \section{Conclusion of Chapter 6}
+% \section{Conclusion}
 % \label{sec:6:conclusion}

### 2019-03-29 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Using \begin{framed}...\end{framed} for all algorithms.

@@
 % -----------------------------------------------------------------
-\section{Introduction}
+\section{Motivation}
 \label{sec:6:Introduction}
@@
 
-Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
-More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
-Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
+% Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
+% More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
+% Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
 %
@@
 
-In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB.
+\TODOL{I need to write a paragraph about using the ``adversarial setting'' for piece-wise stationary.
+It is very important to state that EXP3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with prior knowledge of $T$ and $\Upsilon_T$.}
+
+% In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB.
 Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector.
@@
 
-In this chapter we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
-We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
-Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.
+% In this chapter we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
+% We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
+% Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.
 
@@
 Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
-For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fix channel or subset of channels, and on a particular day in a particular city, if the company deploys a lot of new objects, then one or some channels see their mean occupancy rates abruptly change.
+For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fixed channel or subset of channels, and on a particular day in a particular city, if the company deploys a lot of new objects, then one or some channels see their mean occupancy rates abruptly change.
 
@@
 %
-The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while the adversarial model can be considered as a special (worst) case, with $\Upsilon_T = T$.
+The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while some adversarial models can be considered as a special (worst) case, with $\Upsilon_T = T$ (when considering an adaptive or an oblivious adversary).
 However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
+Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the EXP3.S algorithm for the piece-wise stationary problem.
+\TODOL{Explain here that we should have been more careful regarding this sentence?}
 
@@
 
-To illustrate the setup presented above, we give here two problems of our benchmark, used for the numerical experiments presented in Section~\ref{sec:6:NumericalExperiments}.
+To illustrate the setup presented above, we give here two examples of piece-wise stationary problems.
+% problems of our benchmark,
+used for the numerical experiments presented in Section~\ref{sec:6:NumericalExperiments}.
 
@@
 Note that changes happen on only one arm (\ie, $C_T=\Upsilon_T=4$),
-and the optimal arm changes once at $\tau_2^{(1)}=2000$, with a large gap of $0.6$.
+and the optimal arm changes once at $\tau_2^{(1)}=2000$.
+%  with a large gap of $0.6$.
 
@@
 \textbf{Problem $\bm 2$.}
-This problem is close to Problem 1, with a minimum optimality gap of $0.1$,
+This problem is close to Problem 1, with a minimum optimality gap of $0.1$ (at any time, the smallest difference between two means is at least $0.1$),
 and shown in Figure~\ref{fig:6:Problem_2}.
-However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$.
+However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$ for arms $0$ and $1$, and of $0.2$ for arm $2$ (between two breakpoints, the mean change of $+0.1$ for arm $0$ and $-0.1$ for arm $1$ and $-0.2$ for arm $2$).
 The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$ markers}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$ markers}).
@@
 Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
-Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.
+Passive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
+% , while active strategies are usually tuned with a threshold for the statistical test.
 
@@
 %
-Their common idea is to adapt a classical policy into forgetting old rewards.
+Their common idea is to adapt a classical policy into forgetting old rewards, .
 If the forgetting behavior is done efficiently, then the policy can efficiently focus mostly on the most recent rewards, and passively adapt to changes when they happen.
@@
 and it was proposed by \cite{Garivier11UCBDiscount}.
-They proved that tuning its window-size to $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$, gives a bound on its regret of the form $\cO(\sqrt{\Upsilon_T T \log(T)})$.
-
-Both D-UCB and SW-UCB builds on a stationary policy,but for example EXP3.S from \cite{Auer02NonStochastic}
-builds on the EXP3 policy, which is designed for the adversarial case,
-and achieves a larger regret upper-bound.
-We also found in our experiments that it always performed worse that the reference stationary algorithms (\eg, \klUCB), so we do not include it in Section~\ref{sec:6:NumericalExperiments}.
+They prove that tuning its window-size to $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$, gives a bound on the regret of SW-UCB of the form $\cO(\sqrt{\Upsilon_T T \log(T)})$.
+
+Both D-UCB and SW-UCB builds on a stationary policy, but for example EXP3.S from \cite{Auer02NonStochastic}
+builds on the EXP3 policy, which is designed for the adversarial case.
+EXP3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
+It constitues a good baseline for our numerical experiments in ,
+even if we did find that EXP3.S performs worse that the most of the other approaches based on extensions of stationary bandit algorithms (\eg, SW-UCB).
+% , so we do not include it in .
 Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proved to be less efficient empirically, or are designed for more specific settings.
-%
-For instance, Windowed-Mean Shift from \cite{YuMannor09} is generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting as they consider side observations.
-%
-A different setting where the quantity of interest is not $\Upsilon_T$ but $V_T$ the total variational budget, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
-%
-It can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm.
-Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
-A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proved, which is weaker than existing results in our setting. Hence we do no include this algorithm in our experiments.
 
 The idea of using a simple discount factor, as for D-UCB, was recently adapted to a Bayesian policy, with the Discounted Thompson sampling (DTS) algorithm presented by \cite{RajKalyani17}.
-Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust in the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in their paper.
-The DTS algorithm can perform well in practice, for instance with $\gamma=0.75$ on our Problems $1$ and $2$ (see Section~\ref{sec:6:NumericalExperiments}).
+Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust in the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in the paper.
+The DTS algorithm can perform well in practice, for instance with $\gamma=0.75$ on Problems $1$ and $2$ (see Section~\ref{sec:6:NumericalExperiments}).
 However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
@@
 In general, we found that passively adaptive approaches can be efficient when their parameters are well tuned, but our experiments show that actively adaptive algorithms perform significantly better.
+
+Finally, XXX use two discount factors at the same time, see the NEWMA algorithm from \cite{Keriven2018}.
+
+\TODOL{Talk about NEWMA and its possible application to generalize/improve D-UCB, \cite{Keriven2018}.}
 
@@
 %
-The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley Test and the \UCB{} policy, but no theoretical guarantee was given.
+The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
+%
+The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
+%
 The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
@@
 %
-On the first hand,
-the \CUSUM{} test is rather complicated and parametric, as it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm.
+The \CUSUM{} test is rather complicated and parametric: it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm.
 A change is detected when either random walks cross a threshold $h$.
 It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
-The PH test is similarly complex, and it is also studied in \cite{LiuLeeShroff17},
-and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments,
-and even if it has the advantage of not requiring a parameter $M$,
+The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
+and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments.
+Even if it has the advantage of not requiring a parameter $M$,
 it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
 %
-On the other hand, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the different of the sum of rewards for the first half and the second half.
+In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half.
 It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
@@
 
-When all their parameters are tuned correctly, both approaches are proved to achieve a good regret upper-bound.
+When all their parameters are tuned correctly, both approaches are proved to achieve a good regret upper-bound, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
 As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
-As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$, but most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the time between changes (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
+As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$, but most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
 %
@@
 %
-\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in their paper).
-
-\CUSUMUCB{} is based on a rather complicated two-sided \CUSUM{} test, that uses the first $M$ samples from one arm to compute an initial average, and then detects whether a drift of size larger than $\varepsilon$ occurred from this value by checking whether a random walk based on the remaining observations crosses a threshold $h$.
+\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).
+
+% \CUSUMUCB{} is based on a rather complicated two-sided \CUSUM{} test, that uses the first $M$ samples from one arm to compute an initial average, and then detects whether a drift of size larger than $\varepsilon$ occurred from this value by checking whether a random walk based on the remaining observations crosses a threshold $h$.
 Thus it requires the tuning of three parameters, $M$, $\varepsilon$ and $h$.
-\CUSUMUCB{} performs \emph{local restarts} using this test, to reset the history of \emph{one arm} for which the test detects a change.
-\MUCB{} uses a much simpler test, based on the $w$ most recent observations from an arm: a change is detected if the absolute difference between the empirical means of the first and second halves of those $w$ observations exceeds a threshold $h$.
+On the one hand, \CUSUMUCB{} performs \emph{local restarts} using this test, to reset the history of \emph{one arm} for which the test detects a change.
+% \MUCB{} uses a much simpler test, based on the $w$ most recent observations from an arm: a change is detected if the absolute difference between the empirical means of the first and second halves of those $w$ observations exceeds a threshold $h$.
 So it requires the tuning of two parameters, $w$ and $h$.
-\MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
+On the other hand, \MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
 Compared to \CUSUMUCB{}, note that \MUCB{} is numerically much simpler as it only uses a bounded memory, of order $\bigO{K w}$ for $K$ arms.
@@
 %
-The second line of research of actively adaptive algorithms uses a Bayesian point of view.
+Another of research on actively adaptive algorithms uses a Bayesian point of view.
 A Bayesian CD algorithm is combined with Thompson Sampling in \cite{MellorShapiro13},
 and more recently \cite{Alami17} introduced the Memory Bandit algorithm.
-It is presented as efficient empirically, but no theoretical guarantee are given, and due to its complexity, we do not include it in our experiments.
+It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
 The idea behind Memory Bandit is to use an expert aggregation algorithm, like EXP4 from \cite{Auer02}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
@@
 
-\paragraph{Advantages of our approach.}
+\paragraph{Slowly-varying model.}
+
+A different setting where the quantity of interest is not $\Upsilon_T$ but the total variational budget $V_T$, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
+The total variation budget $V_T$ is defined as $\sum_{t=1}^{T-1} \sum_{i=1}^K D(\nu_i(t), \nu_i(t+1))$, for a certain measure of difference $D$, which measures how much the two distributions of any arm $i$ have changed from time $t$ to time $t+1$.
+Two examples can be the total variation distance $\|\cdot\|_{TV}$ or the Kullback-Leibler divergence.
+%
+The RExp3 algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm.
+Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
+A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proved, which is weaker than existing results in our setting.
+
+Three recent works proposed algorithms that achieve a small regret under the two models.
+For simple bandits, \cite{WeiSrivastava18Abruptly} proposed the SW-UCB\# and LM-DSEE algorithms FIXME.
+For contextual bandits, \cite{Luo18} and \cite{ChenLeeLuoWei2019} FIXME.
+\TODOL{Parler de ces travaux là, qui sont très forts mais pas non plus encore parfaits.}
+
+This setting is quite different from the setting we are interested by in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
+hence we do no include any of these algorithms in the experiments presented in Section~\ref{sec:6:NumericalExperiments}.
+
+\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
+%
+In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
+we assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
+Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
+where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
+and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
+%
+This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
+Empirically, we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
+
+% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}
+
+\paragraph{Positionning of our results.}
 %
@@
     \item
-On one hand, we expect the \emph{local restart} variant of \GLRklUCB{} to outperform the \emph{global restart} variant for problem $1$, as the intuition suggests that it is sub-optimal to reinitialize the memory of the observations of the $K-1$ arms whose distributions did not change after a change-point is detected (as the \emph{global restart} variant do).
+On the one hand, we expect the \emph{local restart} variant of \GLRklUCB{} to outperform the \emph{global restart} variant for problem $1$, as the intuition suggests that it is sub-optimal to reinitialize the memory of the observations of the $K-1$ arms whose distributions did not change after a change-point is detected (as the \emph{global restart} variant do).
     \item
@@
 Of course, on a given problem, without additional prior knowledge on the difficulty of the problem at hand, the algorithm cannot know which situation is more likely to happen, only local changes ($C_T = \Upsilon_T$) or only global changes ($C_T = K \Upsilon_T$) or any intermediate setting between the two extreme cases.
-Thus we do not believe to be able to design a policy that could be uniformly better than the two variants of our proposal \GLRklUCB.
-%
-Surprisingly, our numerical experiments show that the \emph{local} variant of \GLRklUCB{} actually outperforms the \emph{glocal} variant for both problem $1$ and $2$, as shown below in Table~\ref{table:6:totalResults1}, and in other problems as shown in Table~\ref{table:6:totalResults2}.
+Thus we do not believe to be able to design a policy that could be uniformly better than the two variants of \GLRklUCB.
+%
+Surprisingly, numerical experiments show that the \emph{local} variant of \GLRklUCB{} actually outperforms the \emph{glocal} variant for both problem $1$ and $2$, as shown below in Table~\ref{table:6:totalResults1}, and in other problems as shown in Table~\ref{table:6:totalResults2}.
 %
@@
 
-\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
-%
-In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
-we will assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
-Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
-where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
-and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
-
-This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
-Empirically, we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
-
-% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}
-
 % -----------------------------------------------------------------
@@
 Generalized Likelihood Ratio tests date back to the seminal work of \cite{Wilks1938}, and were for instance studied for change-point detection by \cite{barnard1959control,siegmund1995using}.
-Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, can be used for our problem.
+Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, can be used for this problem.
 We propose instead to exploit the fact that bounded distributions are also dominated by Bernoulli distributions.
@@
 
-We refer to Lemma~1 of \cite{KLUCBJournal} that establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
+Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
 Our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm know before hand the values of $a,b$).
@@
 
-If the samples $(X_t)$ were all drawn from a Bernoulli distribution, our change-point detection problem would reduce to a parametric sequential test of
+If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem introduced above would reduce to a parametric sequential test of
 \[\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0)),\]
@@
 %
-where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
-High values of this statistic tend to indicate rejection of $\cH_0$ over $\cH_1$.
+where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the \emph{likelihoods} of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
+High values of this statistic tend to indicate rejection of $\cH_0$.
 %
@@
 %
-Indeed, we show in Appendix~\ref{app:6:GLR_with_kl} that in any one-dimensional exponential family, we have
+Indeed, we show below
+% in Appendix~\ref{app:6:GLR_with_kl}
+that in any one-dimensional exponential family, we have
 \begin{equation}\label{eq:6:secondDefGLRT}
-    \GLR(n) = \exp\Bigl[ \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right], \Bigr].
+    \log \GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right].
 \end{equation}
@@
 where for $k \leq k'$, $\widehat{\mu}_{k:k'} := \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
-%
+
+\begin{proof}
+    First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
+    We have $\ell(X_1, \ldots, X_n ; \mu_0) = \prod_{i=1}^n \ell(X_i ; \mu_0)$ by independence of the observations $X_i$,
+    and $\ell(X_i ; \mu_0) = \mu_0^{X_i} (1-\mu_0)^{1-X_i}$ for Bernoulli distributions.
+    Therefore, taking the logarithm gives
+    \begin{align*}
+        \log\ell(X_1, \ldots, X_n ; \mu_0)
+        &= \sum_{i=1}^n X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) \\
+        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
+        &= n \left[ \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right].
+    \end{align*}
+    %
+    For a constant $a\in[0,1]$, let $h(x) := a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
+    and we are trying to solve $\sup_{x\in[0,1]} h(x)$.
+    If $a=0$ or $a=1$, $h$ is maximum at $x=a$.
+    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative:
+    $h'(x) = a/x - (1-a)/(1-x)$, so $h'(x) = 0$ if and only if $x=a$.
+    Thus in all cases, $\sup_{x\in[0,1]} h(x) = h(a)$.
+    %
+    Here, we have $a = \widehat{\mu}_{1:n}$, and thus we solve the $\sup_{\mu_0}$ optimization problem found in the denominator of the $\GLR(n)$ expression.
+    By replacing $\mu_0 = \widehat{\mu}_{1:n}$, we obtained the following (unique) solution,
+    \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
+        \sup_{\mu_0} \log\ell(X_1, \ldots, X_n ; \mu_0) = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
+    \end{equation}
+
+    Now let us consider the nominator of the $\GLR(n)$ expression.
+    We can again work with log-likelihoods, and so we have
+    \begin{align*}
+        \log\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)
+        &= \sum_{i=1}^{\tau} X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_i \log(\mu_1) + (1-X_i) \log(1-\mu_1) \\
+        &= s \bigl( \widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0) \bigr) \\
+        & \;\;\;\; + (n-s) \bigl( \widehat{\mu}_{s+1:n} \log(\mu_1) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\mu_1) \bigr).
+    \end{align*}
+    Because we solved in \eqref{eq:6:solutionPbOpt_for_1n} the $\sup$ for the denominator,
+    we have
+    \begin{align*}
+        \GLR(n)
+        % &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)}, \\
+        &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right) \right] }, \\
+        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] }, \\
+        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr] \Bigr],  \\
+        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1, 1 < s < n} \Bigl[
+            s \bigl( \widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0) \bigr) \\
+            & \hspace{35pt} + (n-s) \bigl( \widehat{\mu}_{s+1:n} \log(\mu_1) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\mu_1) \bigr) \\
+            & \hspace{35pt} - n \bigl( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \bigr)
+        \Bigr] \Bigr].
+    \end{align*}
+    By linearity and independence, we can separate the joint optimization problem on $\mu_0,\mu_1,s$ in two optimizations problems for $\mu_0,s$ and $\mu_1,s$, that can first be solved explicitly for $\mu_0$ (resp.
+$\mu_1$) and then left to be solved for $s$.
+    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_i = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,
+    which are both handled similarly.
+    For instance for $\mu_0$, we use the same computation as above with the function $h$ to find the optimum:
+    $\widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0)$
+    is optimum for $\mu_0 = \widehat{\mu}_{1:s}$.
+    Similarly, the term for $\mu_1$ gives that $\sup_{\mu_1} \widehat{\mu}_{s+1:n} \log(\mu_1) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\mu_1)$
+    is attained for $\mu_1 = \widehat{\mu}_{s+1:n}$.
+    Finally, by replacing the two expressions of the solutions for $\mu_0$ and $\mu_1$,
+    we obtain
+    \begin{align*}
+        \GLR(n)
+        = \exp \Bigl[ \sup\limits_{1 < s < n} \Bigl[
+            s \times \Bigl(
+                    & \widehat{\mu}_{1:s} \log(\widehat{\mu}_{1:s}) + (1 - \widehat{\mu}_{1:s}) \log(1-\widehat{\mu}_{1:s}) \\
+                    & - \widehat{\mu}_{1:s} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:s}) \log(1-\widehat{\mu}_{1:n})
+            \Bigr) \\
+            + (n-s) \times \Bigl(
+                & \widehat{\mu}_{s+1:n} \log(\widehat{\mu}_{s+1:n}) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\widehat{\mu}_{s+1:n}) \\
+                & - \widehat{\mu}_{s+1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\widehat{\mu}_{1:n})
+            \Bigr)
+        \Bigr] \Bigr].
+    \end{align*}
+    We conclude by recognizing the expressions of $s \times \kl(\widehat{\mu}_{1:s}, \widehat{\mu}_{1:n})$
+    and $(n-s) \times \kl(\widehat{\mu}_{s+1:n}, \widehat{\mu}_{1:n})$.
+\end{proof}
+
 This motivates the following definition of the Bernoulli GLRT change point detector.
-Note that the $\exp$ appearing in the \eqref{eq:6:secondDefGLRT} is absorbed as a $\log$ in the threshold function $\beta$.
+% Note that the $\exp$ appearing in the \eqref{eq:6:secondDefGLRT} is absorbed as a $\log$ in the threshold function $\beta$.
 
@@
 \begin{lemma}\label{lem:6:FalseAlarm}
-    Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
+    Let be $\bP_{\mu_0}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t$.
+    % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
     Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
@@
 Letting $\Delta = |\mu_0 - \mu_1|$, one can essentially establish that for $\tau$ larger than $(1/\Delta^2)\ln(1/\delta)$ (\ie, enough samples before the change), the delay can be of the same magnitude (\ie, enough samples after the change).
-In our bandit analysis to follow, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and \ref{lem:6:GoodEvent}).
+In the bandit analysis to follow, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and \ref{lem:6:GoodEvent}).
 
@@
     It is actually not used as such in the proofs of the regret upper bound given below, but a similar result is used and proved in the proofs.
+    \TODOL{We have to give the proof? I was fine with this sentence...}
 \end{proof}
@@
 
-Lemma~\ref{lem:6:FalseAlarm} provides the first control of false alarm for the B-GLRT employed for bounded data.
+Lemma~\ref{lem:6:FalseAlarm} provides the first non-asymptotic control of false alarm for the B-GLRT employed for bounded data.
 However, the threshold \eqref{def:6:beta} is not fully explicit as the function $\cT(x)$ can only be computed numerically.
@@
 However, the detection delay may be slightly increased.
-In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups has little impact on the regret.
+In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups
+has little impact on the regret of the bandit strategy introduced below.
 
@@
 
-We start by giving the pseudo-code of our algorithm, by explaining every part, then we give our finite-time regret upper bounds, for the two variants of using local or global restarts.
-The analysis in both cases is using a unified proof technique, that we present before giving more details about the two proofs.
+We start by giving the pseudo-code of our algorithm, by explaining every part, then we give its finite-time regret upper bounds, for the two variants of using local or global restarts.
+The analysis in both cases is using a unified proof technique, that we present before giving more details about the proof of half of the results (the other proof can be found in \cite{Besson2019GLRT}).
 
@@
 
-Our proposed algorithm, \GLRklUCB{}, combines a bandit algorithm with a change-point detector running on each arm.
+We now present the \GLRklUCB{} algorithm, which combines a bandit algorithm with a change-point detector running on each arm.
 It also needs a third ingredient, some forced exploration parameterized by $\alpha\in(0,1)$ to ensure each arm is sampled enough and changes can also be detected on arms currently under-sampled by the bandit algorithm.
@@
 %
-This algorithm, formally stated as Algorithm~\ref{algo:6:GLRklUCB}, can be used in any bandit model with bounded rewards, and is expected to be very efficient for Bernoulli distributions, which are relevant for practical applications.
+This algorithm, formally stated as Algorithm~\ref{algo:6:GLRklUCB}, can be used in any bandit model with bounded rewards, and we expect it to be very efficient for Bernoulli distributions, which are relevant for practical applications.
 
@@
     \centering
+    \begin{framed}
     % \begin{small}
@@
     % \end{small}
+    \end{framed}
 \end{figure}
@@
 We present in Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} numerical simulations that compare different choices of threshold functions $\beta$.
-We compare the complicated value used in Lemma~\ref{lem:6:FalseAlarm}, using a numerical approximation of the function $\cT$,
+We compare the non-explicit function used in Lemma~\ref{lem:6:FalseAlarm} (that make use of a numerical approximation of the function $\cT$),
 with the simpler value $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$, as well as two other choices.
-The sum-up of these simulations is that it is enough to use a simpler and more explicit threshold, thus we recommend in practice to use $\beta_1(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
+To sum-up these simulations, they validate the use of a simpler and more explicit threshold, thus we recommend in practice to use $\beta_1(n,\delta) = \ln(3n\sqrt{n}/\delta)$.
 %  to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
@@
 
-The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs randomized exploration.
-%
-We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
-Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
-\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
-and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
+The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
+Both approaches are parameterized by $\alpha\in(0,1)$.
+More precisely, \CUSUMUCB{} samples arm $i$ with probability $\alpha/K$ at each time step,
+and \MUCB{} uses a deterministic scheme, to force exploring the $K$ arms: when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$.
 Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
 %
-A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples}, proved below.
+A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples} below.
 
@@
 We compare the uniformly random exploration used by \CUSUMUCB, against the deterministic scheme used in Algorithm~\ref{algo:6:GLRklUCB}, and against a more complicated scheme based on \emph{tracking}.
-The sum-up of these simulations is that it is enough to use the deterministic scheme to have an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
-As it is the simplest and less numerically costly one, this is the one we chose for Algorithm~\ref{algo:6:GLRklUCB}.
+To sum-up these simulations, the deterministic scheme gives an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
+As it is the simplest one to handle in our proofs, this is the one we chose for Algorithm~\ref{algo:6:GLRklUCB}.
 
@@
 
-Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to our knowledge our analysis is the first one to be given for such assumption on the problem difficulty.
+Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to the best of the authors knowledge,our analysis is the first one to be given for such assumption on the problem difficulty.
 %
@@
 
-One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, our regret bounds hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
+One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
 Hence those guarantees are valid on ``easy'' problem instances only, with few changes of a large magnitude, and in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
@@
 
-We include in our study two algorithms designed for the classical MAB, \klUCB{} \cite{Garivier11KL}, and Thompson sampling \cite{AgrawalGoyal11,Kaufmann12Thompson},
+We include in this study two algorithms designed for the classical MAB, \klUCB{} \cite{Garivier11KL}, and Thompson sampling \cite{AgrawalGoyal11,Kaufmann12Thompson},
 as well as an ``oracle'' version of \klUCB, that we call Oracle-Restart. This algorithm knows the exact locations of the breakpoints, and restarts \klUCB{} at those locations (without any delay).
@@
 Discounted \klUCB{} (D-\klUCB, \cite{Kocsis06}), with discount factor $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$; Sliding-Window \klUCB{} (SW-\klUCB,  \cite{Garivier11UCBDiscount}) using window-size $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$ and Discounted Thompson sampling (DTS, \cite{RajKalyani17}) with discount factor $\gamma = 0.95$.
-For this last algorithm, the discount factor $\gamma=0.75$ suggested by the authors was performing significantly worse on our problem instances.
+For this last algorithm, the discount factor $\gamma=0.75$ suggested by the authors was performing significantly worse on the problem instances we tried.
 More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).
@@
 %
-To speed up our simulations, two optimizations are used, with $\Delta n = \Delta s = 10$,
+To speed up the simulations, two optimizations are used, with $\Delta n = \Delta s = 10$,
 and \CUSUM{} also uses the first trick with $\Delta n = 10$ (see Appendix~\ref{sub:6:IdeasOptimizations} for more details).
@@
 We also give the number of observations (since the last restart) of the arm on which the change was detected.
-As shown below in the summary of our experiments, the three change-point detection algorithm give similar performances in terms of regret, but they have different behaviors.
+As shown below, in the experiments summary, the three change-point detection algorithm give similar performances in terms of regret, but they have different behaviors.
 The test in \MUCB{} usually detects less changes, while the test in \CUSUMUCB{} usually detects more changes and even unnecessary changes.
@@
 But it is still interesting to compare the final regret in one run, as given above, with the mean regret for $1000$ independent runs, as given below in Table~\ref{table:6:totalResults1}.
-The values are not the same, but the ranking is: our proposal outperforms both \CUSUMklUCB{} and \MklUCB, and the test based on \CUSUM{} seems more efficient than the test based of \MklUCB{} in our problems at hand.
+The values are not the same, but the ranking is: our proposal outperforms both \CUSUMklUCB{} and \MklUCB, and the test based on \CUSUM{} seems more efficient than the test based of \MklUCB{} in the problems at hand.
 
 % -----------------------------------------------------------------
-\paragraph{Illustrations of the rest of our benchmark.}\label{par:6:benchmark2}
+\paragraph{Illustrations of the rest of the benchmark.}\label{par:6:benchmark2}
 
@@
 The gaps $\Delta$ are much smaller than for the first problems, with amplitudes ranging from $0.02$ to $0.001$.
-Note that the assumptions of our regret upper bounds are violated, as well as the assumptions for the analysis of \MUCB{} and \CUSUMUCB.
+Note that the assumptions of the regret upper bounds for \GLRklUCB{} are violated, as well as the assumptions for the analysis of \MUCB{} and \CUSUMUCB.
 It is interesting to say that this problem is inspired from Figure~3 of \cite{CaoZhenKvetonXie18}, where the synthetic data was obtained from manipulations on a real-world database of clicks from \emph{Yahoo!}.
@@
 
-Like problem $3$, this harder problem is also inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!}, but from our other competitor paper, see Figure~3 from \cite{LiuLeeShroff17}.
+Like problem $3$, this harder problem is also inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!}, but from another competitor paper, see Figure~3 from \cite{LiuLeeShroff17}.
 %
@@
 %
-On these problems, our proposal with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts.
-Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet our experiments show the superiority of local restarts on every instances.
+On these problems, \GLRklUCB{} with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts.
+Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instances.
 
@@
 One can note that the best non-oracle strategies are actively adaptive,
-thus our experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.
+thus the experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.
 
@@
 Indeed the oracle policy is suboptimal as it restarts as soon as one arm change but is unaware of the meaningful changes, and stationary policies which quickly identify the best arm will play it most of the times, achieving a smaller regret.
-We note that, sadly, all actively adaptive policies fail to outperform stationary policies on such hard problems, because they do not observe enough rewards from each arm between two restarts (\ie, the Assumptions~\ref{ass:6:LongPeriodsGlobal} and \ref{ass:6:LongPeriods} for our Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} are not satisfied).
+We note that, sadly, all actively adaptive policies fail to outperform stationary policies on such hard problems, because they do not observe enough rewards from each arm between two restarts (\ie, the Assumptions~\ref{ass:6:LongPeriodsGlobal} and \ref{ass:6:LongPeriods} for the two Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} are not satisfied).
 We can also verify that the two options, \textbf{Local} and \textbf{Global} restart, for \GLRklUCB, give close results, and that the \textbf{Local} option is always better.
@@
 We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
-We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as our proposal \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
-Finally, we would like to study in the future possible extension of our approach to the slowly varying model, as introduced in \cite{WeiSrivastava18Abruptly}.
+We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
+Finally, we would like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,WeiSrivastava18Abruptly}.
 

### 2019-03-29 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Presque bon pour corriger selon les modifications suggérées par Emilie.

@@
 
-This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the finite-time regret bounds for the two variants, with the proof of the first one (the second one is proved in Appendix~\ref{proof:6:mainRegretBound}), which both follows the skeleton of the unified analysis.
+This section give the finite-time regret bounds for the two variants, and interpretations of the different results.
+
+% ----------------------------------------------------------------------------
+\subsection{Results for \GLRklUCB{} using Global Changes}
+
+Recall that $\tau^{(k)}$ denotes the position of the $k$-th break-point and let $\mu_i^{(k)}$ be the mean of arm $i$ on the segment between the $k$ and $(k+1)$-th breakpoint:
+$\forall t \in \{ \tau^{(k-1)}+1, \dots, \tau^{(k)} \}, \mu_{i}(t) = \mu_i^{(k)}$.
+We also introduce $k^* = \arg\max_i \mu_i^{(k)} $ and the largest gap at break-point $k$ as $\Delta^{(k)} := \max_{i=1,\dots,K} |\mu_i^{(k)} - \mu_i^{(k-1)}| >0$.
+
+\begin{assumption}[Enough samples between two global change-points]
+    \label{ass:6:LongPeriodsGlobal}
+    Define
+    \begin{equation}\label{def:6:DelayGlobal}
+        d^{(k)} = d^{(k)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
+    \end{equation}
+    %
+    Then \emph{we assume} that for all change $k \in \{1,\dots,\Upsilon_T\}$,
+    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
+\end{assumption}
+
+Assumption~\ref{ass:6:LongPeriodsGlobal} is easy to interpret and standard in non-stationary bandits.
+It requires that the distance between two consecutive breakpoints is large enough: how large depends on the magnitude of the largest change that happen at those two breakpoints.
+Under this assumption, we provide in Theorem~\ref{thm:6:mainRegretBoundGlobal} a finite time problem-dependent regret upper bound.
+It features the parameters $\alpha$ and $\delta$,
+the KL-divergence terms $\kl(\mu_{i}^{(k)},\mu_{k^*}^{(k)})$ expressing the hardness of the (stationary) MAB problem between two breakpoints,
+and the $\Delta^{(k)}$ terms expressing the hardness of the change-point detection problem.
+
+\begin{theorem}\label{thm:6:mainRegretBoundGlobal}
+    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
+    \begin{align}
+        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T \\
+        & + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
+\nonumber
+    \end{align}
+\end{theorem}
+
+We highlight that \emph{this result is finite-time} and not asymptotic, even if it uses the notation $\bigO{\sqrt{\ln(T)}}$.
+This last term does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a constant times $\sqrt{\ln(T)}$ from a certain time $T_0$.
+
+\begin{corollary}\label{cor:6:Global}
+    For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriodsGlobal},
+    let $\Delta^{\text{opt}}$ denote the smallest value of a sub-optimality gap on one of the stationary segments, and $\Delta^{\text{change}}$ be the smallest magnitude of any change point on any arm.
+    \begin{enumerate}
+        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \Upsilon_T\sqrt{T\ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$,
+        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2}\sqrt{\Upsilon_T T \ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$.
+    \end{enumerate}
+\end{corollary}
+
+% ----------------------------------------------------------------------------
+\subsection{Results for \GLRklUCB{} using Local Changes}
+
+Some new notations are needed to state a regret bound for \GLRklUCB{} using local changes,
+and to distinguish notations between local and global changes, we denote $\ell$ instead of $k$ for the indexes of change-points.
+We let $\tau_i^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $i$}: $\tau_i^{(\ell)} = \inf \{ t > \tau_i^{(\ell - 1)} : \mu_i(t) \neq \mu_i(t+1)\}$,
+with the convention $\tau_i^{(0)}=0$, and let $\overline{\mu}_i^{(\ell)}$ be the $\ell$-th value for the mean of arm $i$, such that $\forall t \in [\tau_i^{(\ell-1)}+1, \tau_i^{(\ell)}], \ \ \mu_i(t) = \overline{\mu}_i^{(\ell)}$.
+We also introduce the gap $\Delta_i^{(\ell)} = \overline{\mu}_i^{\ell} - \overline{\mu}_i^{\ell-1} > 0$.
+
+\begin{assumption}[Enough samples between two local change-points]
+    \label{ass:6:LongPeriods}
+    Define
+    \begin{equation}\label{def:6:DelayLocal}
+        d_i^{(\ell)} = d_i^{(\ell)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
+    \end{equation}
+    %
+    Then \emph{we assume} that for all arm $i$ and all change-point of that arm $\ell \in \{1,\dots,\NCi\}$, $\tau_i^{(\ell)} - \tau_i^{(\ell-1)} \geq 2\max (d_i^{(\ell)},d_i^{(\ell-1)})$.
+\end{assumption}
+
+Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to the best of the authors knowledge,our analysis is the first one to be given for such assumption on the problem difficulty.
+%
+It requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
+Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{i=1}^K \NCi \leq K \Upsilon_T$.
+
+\begin{theorem}\label{thm:6:mainRegretBound}
+    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on Local Restart satisfies
+    %
+    \begin{equation}
+        R_T \leq 2\sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{4K}{\alpha \left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
+    \end{equation}
+    where
+    \begin{equation}
+        {\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}.
+    \end{equation}
+\end{theorem}
+
+Like for the first Theorem~\ref{thm:6:mainRegretBoundGlobal},
+we highlight that \emph{this result is finite-time}.
+%
+We can directly obtain different regret upper-bounds for different choices of the two parameters $\alpha$ and $\delta$.
+We prefer to state the four cases, to highlight the modularity of the result given by Theorem~\ref{thm:6:mainRegretBound}.
+
+\begin{corollary}\label{cor:6:Local}
+    For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriods},
+    with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, the following holds.
+    \begin{enumerate}
+        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$,
+        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
+        \item Choosing $\alpha = \sqrt{\frac{\EffChange\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\EffChange T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
+        \item Choosing $\alpha = \sqrt{\frac{K\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{K\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$.
+    \end{enumerate}
+\end{corollary}
+
+The proof of \GLRklUCB{} with Local Changes is only given in Appendix~\ref{proof:6:mainRegretBound}, as it is quite similar to the proof for Global changes given above.
+The main difficulty relies in defining the ``good event'' $\cE_T$, and proving that it happens with high probability (by showing that the complementary event $\cE_T^c$ is highly unlikely).
+
+% ----------------------------------------------------------------------------
+\subsection{Interpretation and comparison of the results}
+\label{sub:6:interpretationRegretBounds}
+
+The regret bounds we obtain for the two variants of our proposal \GLRklUCB{},
+Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
+both show that there exists a tuning of $\alpha$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
+%
+For very particular instances such that $\Upsilon_T = \EffChange$, \ie, at each break-point only one arm changes (\eg, problem $1$ from Figure~\ref{fig:6:Problem_1}), the theory advocates the use of local changes.
+%
+Indeed, while the regret guarantees obtained are similar, those obtained for local changes hold for a wider variety of problems as Assumption~\ref{ass:6:LongPeriods} is less stringent than Assumption~\ref{ass:6:LongPeriodsGlobal}.
+%
+Besides those specific instances, our results are essentially worse for local than global changes.
+However, we only obtain regret upper bounds -- thus providing a theoretical safety net for both variants of our algorithm, and the practical story is different, as discussed in Section~\ref{sec:6:NumericalExperiments}.
+We find that \GLRklUCB{} performs better with local restarts, uniformly on all problems.
+
+One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
+Hence those guarantees are valid on ``easy'' problem instances only, with few changes of a large magnitude, and in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
+%
+However, this does not prevent our algorithms from performing well on more realistic instances, and numerical experiments support this claim,
+as illustrated for instance with problem $3$ presented in the next Section~\ref{sec:6:NumericalExperiments}).
+%
+Note that \MUCB{} \cite{CaoZhenKvetonXie18} is also analyzed for the same type of unrealistic assumptions, while its practical performance is illustrated beyond those.
+
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
+
+%-----------------------------------------------------------------------------
+\subsection{Proof of the regret upper-bounds}
+\label{sec:6:proofRegret}
+
+\TODOL{Write intro}
+
+This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the proof of the finite-time regret bound for the variant using ``global restarts''.
+The proof of the other variant using ``local restarts'' follows the skeleton of the unified analysis, and can be found in Appendix~E of \cite{Besson2019GLRT}.
 
@@
 %-----------------------------------------------------------------------------
-\subsection{Results for \GLRklUCB{} using Global Changes}
-
-Recall that $\tau^{(k)}$ denotes the position of the $k$-th break-point and let $\mu_i^{(k)}$ be the mean of arm $i$ on the segment between the $k$ and $(k+1)$-th breakpoint:
-$\forall t \in \{ \tau^{(k-1)}+1, \dots, \tau^{(k)} \}, \mu_{i}(t) = \mu_i^{(k)}$.
-We also introduce $k^* = \arg\max_i \mu_i^{(k)} $ and the largest gap at break-point $k$ as $\Delta^{(k)} := \max_{i=1,\dots,K} |\mu_i^{(k)} - \mu_i^{(k-1)}| >0$.
-
-\begin{assumption}[Enough samples between two global change-points]
-    \label{ass:6:LongPeriodsGlobal}
-    Define
-    \begin{equation}\label{def:6:DelayGlobal}
-        d^{(k)} = d^{(k)}(\alpha,\delta) = \left\lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \right\rceil.
-    \end{equation}
-    %
-    Then \emph{we assume} that for all change $k \in \{1,\dots,\Upsilon_T\}$,
-    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
-\end{assumption}
-
-Assumption~\ref{ass:6:LongPeriodsGlobal} is easy to interpret and standard in non-stationary bandits.
-It requires that the distance between two consecutive breakpoints is large enough: how large depends on the magnitude of the largest change that happen at those two breakpoints.
-Under this assumption, we provide in Theorem~\ref{thm:6:mainRegretBoundGlobal} a finite time problem-dependent regret upper bound.
-It features the parameters $\alpha$ and $\delta$,
-the KL-divergence terms $\kl(\mu_{i}^{(k)},\mu_{k^*}^{(k)})$ expressing the hardness of the (stationary) MAB problem between two breakpoints,
-and the $\Delta^{(k)}$ terms expressing the hardness of the change-point detection problem.
-
-\begin{theorem}\label{thm:6:mainRegretBoundGlobal}
-    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
-    \begin{align}
-        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T \\
-        & + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
-\nonumber
-    \end{align}
-\end{theorem}
-
-We highlight that \emph{this result is finite-time} and not asymptotic, even if it uses the notation $\bigO{\sqrt{\ln(T)}}$.
-This last term does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a constant times $\sqrt{\ln(T)}$ from a certain time $T_0$.
-
-\begin{corollary}\label{cor:6:Global}
-    For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriodsGlobal},
-    let $\Delta^{\text{opt}}$ denote the smallest value of a sub-optimality gap on one of the stationary segments, and $\Delta^{\text{change}}$ be the smallest magnitude of any change point on any arm.
-    \begin{enumerate}
-        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \Upsilon_T\sqrt{T\ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2}\sqrt{\Upsilon_T T \ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$.
-    \end{enumerate}
-\end{corollary}
 
@@
 
-\subsection{Results for \GLRklUCB{} using Local Changes}
-
-Some new notations are needed to state a regret bound for \GLRklUCB{} using local changes,
-and to distinguish notations between local and global changes, we denote $\ell$ instead of $k$ for the indexes of change-points.
-We let $\tau_i^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $i$}: $\tau_i^{(\ell)} = \inf \{ t > \tau_i^{(\ell - 1)} : \mu_i(t) \neq \mu_i(t+1)\}$,
-with the convention $\tau_i^{(0)}=0$, and let $\overline{\mu}_i^{(\ell)}$ be the $\ell$-th value for the mean of arm $i$, such that $\forall t \in [\tau_i^{(\ell-1)}+1, \tau_i^{(\ell)}], \ \ \mu_i(t) = \overline{\mu}_i^{(\ell)}$.
-We also introduce the gap $\Delta_i^{(\ell)} = \overline{\mu}_i^{\ell} - \overline{\mu}_i^{\ell-1} > 0$.
-
-\begin{assumption}[Enough samples between two local change-points]
-    \label{ass:6:LongPeriods}
-    Define
-    \begin{equation}\label{def:6:DelayLocal}
-        d_i^{(\ell)} = d_i^{(\ell)}(\alpha,\delta) = \lceil \frac{4K}{\alpha\left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha}\rceil.
-    \end{equation}
-    %
-    Then \emph{we assume} that for all arm $i$ and all change-point of that arm $\ell \in \{1,\dots,\NCi\}$, $\tau_i^{(\ell)} - \tau_i^{(\ell-1)} \geq 2\max (d_i^{(\ell)},d_i^{(\ell-1)})$.
-\end{assumption}
-
-Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to the best of the authors knowledge,our analysis is the first one to be given for such assumption on the problem difficulty.
-%
-It requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
-Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{i=1}^K \NCi \leq K \Upsilon_T$.
-
-\begin{theorem}\label{thm:6:mainRegretBound}
-    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on Local Restart satisfies
-    %
-    \begin{equation}
-        R_T \leq 2\sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{4K}{\alpha \left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
-    \end{equation}
-    where
-    \begin{equation}
-        {\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}.
-    \end{equation}
-\end{theorem}
-
-Like for the first Theorem~\ref{thm:6:mainRegretBoundGlobal},
-we highlight that \emph{this result is finite-time}.
-%
-We can directly obtain different regret upper-bounds for different choices of the two parameters $\alpha$ and $\delta$.
-We prefer to state the four cases, to highlight the modularity of the result given by Theorem~\ref{thm:6:mainRegretBound}.
-
-\begin{corollary}\label{cor:6:Local}
-    For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriods},
-    with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, the following holds.
-    \begin{enumerate}
-        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{\EffChange\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\EffChange T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{K\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{K\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$.
-    \end{enumerate}
-\end{corollary}
-
-The proof of \GLRklUCB{} with Local Changes is only given in Appendix~\ref{proof:6:mainRegretBound}, as it is quite similar to the proof for Global changes given above.
-The main difficulty relies in defining the ``good event'' $\cE_T$, and proving that it happens with high probability (by showing that the complementary event $\cE_T^c$ is highly unlikely).
-
-% ----------------------------------------------------------------------------
-\subsection{Interpretation and comparison of the results}
-\label{sub:6:interpretationRegretBounds}
-
-The regret bounds we obtain for the two variants of our proposal \GLRklUCB{},
-Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
-both show that there exists a tuning of $\alpha$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
-%
-For very particular instances such that $\Upsilon_T = \EffChange$, \ie, at each break-point only one arm changes (\eg, problem $1$ from Figure~\ref{fig:6:Problem_1}), the theory advocates the use of local changes.
-%
-Indeed, while the regret guarantees obtained are similar, those obtained for local changes hold for a wider variety of problems as Assumption~\ref{ass:6:LongPeriods} is less stringent than Assumption~\ref{ass:6:LongPeriodsGlobal}.
-%
-Besides those specific instances, our results are essentially worse for local than global changes.
-However, we only obtain regret upper bounds -- thus providing a theoretical safety net for both variants of our algorithm, and the practical story is different, as discussed in Section~\ref{sec:6:NumericalExperiments}.
-We find that \GLRklUCB{} performs better with local restarts, uniformly on all problems.
-
-One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
-Hence those guarantees are valid on ``easy'' problem instances only, with few changes of a large magnitude, and in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
-%
-However, this does not prevent our algorithms from performing well on more realistic instances, and numerical experiments support this claim,
-as illustrated for instance with problem $3$ presented in the next Section~\ref{sec:6:NumericalExperiments}).
-%
-Note that \MUCB{} \cite{CaoZhenKvetonXie18} is also analyzed for the same type of unrealistic assumptions, while its practical performance is illustrated beyond those.
-
-% ----------------------------------------------------------------------------
-% ----------------------------------------------------------------------------
 \section{Experimental Results}
@@
 
-We report results obtained on five different piece-wise stationary bandit problems, illustrated in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} above, and Figures~\ref{fig:6:Problem_5}, \ref{fig:6:Problem_4} and \ref{fig:6:Problem_6} below, and described in more details below.
-Results are reported with regret tables and regret plots.
-For each experiment, the regret was estimated using $1000$ independent runs.
+We report results obtained on five different piece-wise stationary bandit problems, illustrated in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} above, and Figures~\ref{fig:6:Problem_5}, \ref{fig:6:Problem_4} and \ref{fig:6:Problem_6} below.
+% , and described in more details below.
+We present regret tables and regret plots,
+for which the regret was estimated using $1000$ independent runs.
 
@@
 \emph{For one (random) simulation}, we give below examples of the efficiency of the change-point detection algorithm used by \GLRklUCB{} as well as \CUSUMklUCB{} and \MklUCB,
-by giving the times at which they detected a change and reinitialized their memory of observations of one or all the arms.
-We also give the number of observations (since the last restart) of the arm on which the change was detected.
+by showing the times at which they detect a change and reinitialize their memory of observations of one or all the arms.
+We also display the number of observations (since the last restart) of the arm on which the change was detected.
 As shown below, in the experiments summary, the three change-point detection algorithm give similar performances in terms of regret, but they have different behaviors.
@@
 \end{itemize}
+
+\TODOL{Conclusion à tirer de cet exemple !}
 

### 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
Add Figures for Appendix of Chapter4

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_PU__v1__simple_grc.png}
    \caption{The \textbf{random traffic generator}. \texttt{Generator} is the only hand-written block, which is configured with a list of active channels, a list of occupancy rate, and constants about the PHY layer (preamble length and data length). It uses one USRP only for emitting data (``Sink'' mode).}
    \label{fig:4app:USRP_TX_PU__v1__simple_grc}
\end{figure}

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_RX_BTS__v1__simple_grc.png}
    \caption{The \textbf{IoT gateway}. The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length). All blocks need to know constants about the PHY layer (preamble length and data length). The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
    \label{fig:4app:USRP_RX_BTS__v1__simple_grc}
\end{figure}

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_SU__v1__simple_grc.png}
    \caption{The \textbf{IoT dynamic object}. The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels. Most blocks need to know constants about the PHY layer (preamble length and data length). Any of the IoT dynamic objects use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
    \label{fig:4app:USRP_TX_SU__v1__simple_grc}
\end{figure}

### 2019-04-01 2-Chapters/4-Chapter/chapter4.tex
Add Figures for Appendix of Chapter4

@@
 Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
-We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time.
+We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every instant.
 If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
@@
 Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
+
+On the first hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT objects, at the cost of a lower Quality of Service (QoS).
+On the other hand, the second model is more interesting if the main target is an improvement of the QoS of the IoT networks rather than a longer battery life.
 
@@
 \label{sec:4:appendix}
-% \section{Proof of Proposition \ref{prop:41:Lagrangian}}
-\label{sec:5:proofLagrangian}
+
+\subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
+\label{sec:4:proofLagrangian}
 % ----------------------------------------------------------------------------
@@
 \input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex}
+
+\newpage
+\subsection{Illustration of the GNU Radio Companion Flowcharts}
+\label{sec:4:IllustrationFlowcharts}
+% ----------------------------------------------------------------------------
+
+\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex}

### 2019-04-01 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Updated bibliography for one reference.
Add Figures for Appendix of Chapter4
Done for now for Chapter4

@@
 
-As explained before, setting up the future Internet of Things (IoT) networks will require to support more and more communicating devices.
-In this section, we show that intelligent devices in unlicensed bands can use Multi-Armed Bandit (MAB) learning algorithms to improve resource exploitation.
+As explained before, the future Internet of Things (IoT) networks will require to support more and more communicating devices.
+In this section, we show that intelligent devices in unlicensed bands can use Multi-Armed Bandit (MAB) learning algorithms to improve spectral resource exploitation.
 %
 We evaluate the performance of two classical MAB learning algorithms, \UCB{} and Thomson Sampling, to handle the decentralized decision-making of Spectrum Access, applied to IoT networks.
+We mean by \emph{decentralized} that learning is performed on the device side, and is spread on the totality of the devices in a network.
 We also evaluate the learning performance when the number of intelligent end-devices grows.
 %
-We show that using learning algorithms does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
-In the studied scenario, stochastic MAB learning provides an up to $16\%$ gain in term of successful transmission probabilities, and has near optimal performance even in non-stationary settings with a majority of intelligent devices.
+We show that using learning algorithms
+firstly extends devices battery autonomy (by reducing the packet-loss-ratio or equivalently, by improving the successful-communication rate),
+and secondly does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
+In the studied settings,
+Stochastic MAB learning
+% provides an up to $16\%$ gain in terms of successful transmission probabilities, and
+has near optimal performance even in non-stationary settings with a majority of intelligent devices.
 
@@
 To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
-Static devices form an interfering traffic, which could have been generated by devices using other standards as well.
+Static devices form an interfering traffic, which could be generated by devices using other standards as well.
 Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel.
 So our first hypothesis is actually not constraining.
-We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller optimally distributes all of them in the channels at the beginning of the scenario (that is, an ideal approach).
+We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
+This second approach is ideal, but not realistic in decentralized IoT networks, and it is just used here as a reference.
 Then, these three reference scenarios allow to evaluate the performance of \UCB{} and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
@@
 %
-We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of objects all sends uplink packets to the (unique) network gateway.
-Messages can be sent in a fixed number $K\geq2$ of wireless channels (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
+We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of devices all sends uplink packets to a (unique) network gateway.
+Messages can be sent in a fixed number $K\geq2$ of wireless channels,
+that are assumed to be orthogonal, that is, a message in channel $i$ will not cause collision to a message in another channel $j \neq i$.
+% (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
 The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
 %
-Note that this choice of number of channels and of frequency bands is for instance the one of the current LoRa standard in Europe, but of course the precise choice does not really matter, and future standards can choose different parameters.
+% Note that this choice of number of channels and of frequency bands is for instance the one of the current LoRa standard in Europe, but of course the precise choice does not really matter, and future standards can choose different parameters.
 
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering
@@
 The devices can transmit their packets in one of the $K$ channels.
-In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1\;\mathrm{s}$ in our implementation).
-In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the object by letting only one gateway sends back an acknowledgement, and the natural choice is for the network to send this acknowledgement by the gateway which received the uplink message.
+In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
+This is realistic in IoT networks such as networks based on LoRa.
+% (of $1\;\mathrm{s}$ in our implementation).
+In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the device by letting only one gateway sends back an acknowledgement.
+The natural choice is for the network to send this acknowledgement by the gateway which received the uplink message.
 For simplicity but without loss of generality, we consider a network with only one gateway,
-but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming uplink messages from the IoT objects being paired in the network.
+but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming uplink messages from the IoT devices being paired in the network.
 %
 
-These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated neighboring networks (in \textcolor{orange}{orange} in Figure~\ref{fig:41:system_model1}).
-This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels.
+These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated co-localized and/or neighboring networks (in \textcolor{orange}{orange} in Figure~\ref{fig:41:system_model1}).
+This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels,
+as each IoT protocol or IoT networks can choose to use a different subset of channels.
 %
-To simulate networks designed for the Internet of Things (IoT), we consider a protocol \textbf{with no sensing} and no repetition of uplink messages.
-The gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) who succeeded in sending an uplink packet.
+In order to simulate networks designed for the IoT, we consider a protocol \textbf{with no sensing} and no repetition of uplink messages.
+The gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) who succeeded in sending an uplink packet.
 %
-By considering a small number of wireless channels, and a unique PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
-the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use, that is, the most vacant one.
+By considering a small number of wireless channels, and a unique PHY layer configuration\footnote{~But MAB learning to also optimize on the PHY layer parameters have started to be explored, after our work \cite{Bonnefoi17}, see for instance \cite{KerkoucheAlami18}.} (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
+the device can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use, that is, the most vacant one.
 Indeed, in this model, the \emph{quality} of the channels (\ie, the ressources) are identified by their vacancy, or one minus their occupancy rates.
@@
     \caption{The considered time-frequency slotted protocol.
-Each frame is composed by a fix-duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
+Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
 If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
@@
     \item
-    \emph{Static} end-devices are identical, and each of them uses only one channel to communicate with the gateway.
-Their choice is assumed to be fixed in time (\ie, stationary).
-The traffic generated by these devices is considered as an interfering traffic for other devices.
+    \emph{Static} end-devices are identical, and each of them uses only \emph{one} channel to communicate with the gateway, with no loss of generality.
+    %  if devices random access the channels.
+    Their choice is assumed to be fixed in time (\ie, stationary). The traffic generated by these devices is considered as an interfering traffic for other devices.
     \item
-    \emph{Dynamic} (or \emph{smart}) end-devices can use all the available channels, by quickly changing their communication channel at any time.
-They can also store communication successes or failures they experienced in each channel, in order to change channel, possibly at every time slot.
+    \emph{Dynamic} (or \emph{smart}) end-devices can use all the available channels, by quickly changing their communication channel at any time slot, following a Machine Learning policy.
+    For that end, they use their history of past communication successes or failures they experienced in each channel, in order to learn about channel availability.
     We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
-    Note that of course we assume a limited storage capacity, so no object stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bits IEEE float number for each of the $K$ channels).
-    Our goal is to propose a learning algorithm than can be implemented by each dynamic devices, in a decentralized and independent manner, to improve their successful communication rate automatically.
+    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bits IEEE float number for each of the $K$ channels).
+    Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, to improve their successful communication rate automatically.
 \end{itemize}
@@
 all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{~In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $K / (S + D)$ for all devices to communicate successfully (in average).}.
-The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed (\ie, real time during two messages), and $1/p$ is proportional to the \emph{duty cycle}.
-For instance for IoT objects sending messages with a duration of $1$ second exactly, a ``smart'' refrigerator could have to send a daily message, and if a message can be sent every second, then $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.
+The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed, and $1/p$ is proportional to the \emph{duty cycle}.
+% (\ie, real time during two messages)
+For instance, for IoT devices sending messages with a duration of $1$ second and possibly at every second, then on a daily basis we have $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.
 
@@
 As this problem is only interesting if devices are able to communicate reasonably efficiently with the gateway, we assume that devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
-Note that even unlicensed bands have such limitations, as for instance this is a strict requirement for any object using the $867\;\mathrm{MHz}$ ISM band (and its regulation also enforces a maximum transmission power, but we do not consider power transmission in this work).
+Note that even unlicensed bands have such limitations, as for instance this is a strict requirement for any device using the $867\;\mathrm{MHz}$ ISM band in Europe (and its regulation also enforces a maximum transmission power, but we do not consider power transmission in this work).
 %
@@
 % \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
-Indeed, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ objects, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall i, S_i > 0$).
+Indeed, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall i, S_i > 0$).
 Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
-Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$ to have an average of active objects not larger than the number of channels.
+Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to have an average of active devices not larger than the number of channels.
 
@@
 it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
-Then, the dynamic device waits in this channel $C(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example of a ``smart fridge'' considered above).
+Then, the dynamic device waits in this channel $C(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example considered above).
 Before sending another message (\ie, at time $t+1$), the dynamic device knows if it received or not this \Ack{} message.
@@
 are $S_1, \dots, S_K$ the (stationary) occupancy of the channels by the static devices,
-and they are of course unknown to each dynamic device, so to maximize their cumulated rewards, they must learn the distributions of the channels, in order to be able to progressively improve its successful communication rate.
+and they are unknown to each dynamic device, so to maximize their cumulated rewards, they must learn the distributions of the channels, in order to be able to progressively improve their respective successful communication rate.
 %
@@
 
-We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of $(S_i)_i$ the repartition of static devices, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.} is possible.
+We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of $(S_i)_i$ the repartition of static devices, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
+This is not possible in IoT contexts as several completely independent networks may operate in a single place.} is possible.
 We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.
@@
 For such values for $D_i$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
-    \footnote{~It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
+    \footnote{~It is worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
     \item
@@
     \item
-    More details are given in Section~\ref{sec:5:proofLagrangian} in the Appendix of this Chapter.
+    More details are given in Section~\ref{sec:4:proofLagrangian} in the Appendix of this Chapter.
     % \hfill{}$\square$
@@
 
-While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused partly by the learning policy used by dynamic objects, but mainly by their random activation processes.
+While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused partly by the learning policy used by dynamic devices, but mainly by their random activation processes.
 
 In our model, every dynamic device implements its own learning algorithm, \emph{independently}.
-For one device, the time $t$ refers to the number of time it accessed the network (following its Bernoulli transmission process, \ie, its duty cycle), \emph{not} the total number of time slots from the beginning, as rewards are only obtained after a transmission, and IoT objects only transmit sporadically, due to low transmission duty cycles.
+For one device, the time $t$ refers to the number of time it accessed the network (following its Bernoulli transmission process, \ie, its duty cycle), \emph{not} the total number of time slots from the beginning, as rewards are only obtained after a transmission, and IoT devices only transmit sporadically, due to low transmission duty cycles.
 
@@
 Besides the modified time scale (rewards are no longer collected at every time step), this looks like a usual bandit problem.
-However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning and random activation of each devices).
+However, it cannot be modeled as a stochastic MAB, as the rewards are (unfortunately) \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning and random activation of each devices).
 %
@@
 %
-In our model, every dynamic dynamic device implements its own \UCB{} algorithm, \emph{independently}.
+In our model, every dynamic device implements its own \UCB{} algorithm, \emph{independently}.
 For one dynamic device, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
@@
 The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
-It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
+It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
 
@@
 Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
-The \emph{adversarial MAB} algorithms is a broader family, of which a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Bubeck12}.
+The \emph{adversarial MAB} algorithms is a broader family, of which a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Auer02,Bubeck12}.
 Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform significantly worse than both \UCB{} and TS in the same experiments,
@@
 % ------------------------------------------------------------------------
-\subsection{Numerical results on an example of IoT network}\label{sub:41:numericalResults}
+\subsection{Numerical results}\label{sub:41:numericalResults}
 
@@
 
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
@@
 
-\paragraph{Successful transmissions rate as a function of the dynamic devices}
-
-To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure~\ref{fig:41:perf_learning} the evolution of the gain, in term of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each object.
-We can see that when the proportion of end-devices is low (\eg, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
-The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a lowest rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$).
-
-\begin{figure}[!t]
+\begin{figure}[!h]
     \centering
@@
 
+\paragraph{Successful transmissions rate as a function of the dynamic devices}
+
+To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure~\ref{fig:41:perf_learning} the evolution of the gain, in terms of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each IoT device.
+We can see that when the proportion of end-devices is low (\eg, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
+The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a lowest rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$).
+
 Figure \ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
@@
 
-\paragraph{A sanity check}
-
-We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall i, S_i = S/K$), to check that learning approach (here, only \UCB)
-also gives interesting gain of performance, and achieve a close-to optimal successful transmission rate.
-Except for the limit case of $100\%$ of dynamic devices (where the uniform access performs as well as the optimal oracle solution), the MAB-based approach almost instantly outperforms the baseline.
-
-\begin{figure}[!t]
+\begin{figure}[!h]
     \centering
@@
 
+\paragraph{A safety check}
+
+We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall i, S_i = S/K$), to check that learning approach (here, only \UCB)
+also gives interesting gain of performance, and achieve a close-to optimal successful transmission rate.
+Except for the limit case of $100\%$ of dynamic devices, which corresponds to Figures~\ref{fig:41:100intelligent} and \ref{fig:41:ch2_100}, where the uniform access performs as well as the optimal oracle solution, the MAB-based approach almost instantly outperforms the baseline.
+
 \paragraph{Note on the simulation code}

### 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Add Figures for Appendix of Chapter4
Done for now for Chapter4

@@
 
-In this section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept of the model introduced in Section~\ref{sec:4:firstModel}.
+In this section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept (PoC) of the model introduced in Section~\ref{sec:4:firstModel}.
 
-We implement an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) objects, embedding the proposed solution,
-and a traffic generator that emulates radio interference from many other objects.
-Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol with retransmissions, which does not require any specific overhead for the learning.
+This PoC implements an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) devices, embedding the proposed solution,
+and a traffic generator that emulates radio interference from many other devices.
+Intelligent devices communicate with the gateway with a wireless ALOHA-based protocol with acknowledgements, which does not require any specific overhead for the learning.
 %
-Similarly to the previous section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
+Similarly to the previous section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent devices can improve their access rate to the network, by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
 %
-This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature at the higher level of the MAC layer, in all or only some of the devices, without any modification on the network side.
+This solution could be added in a straightforward and costless manner in most IoT networks, such as LoRaWAN networks, just by adding this feature at the higher level of the MAC layer, in all or only some of the devices, without any modification on the network side, and no signalling overhead for the devices.
 
@@
 %
-We use a TestBed designed in 2017 by the SCEE team at the Rennes campus of CentraleSupélec \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
-and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.
+We use a TestBed designed in 2017 by the SCEE team at the Rennes campus of CentraleSupélec \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop running the GNU Radio software \cite{GNURadioDocumentation},
+and where the intelligence of each device corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.
 
 In our demo, we consider a simple wireless network, consisting of one gateway (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
-Some dynamic intelligent objects (end-user or autonomous objects) try to communicate with the gateway, with a low-overhead protocol.
+Some dynamic intelligent devices (end-user or autonomous devices) try to communicate with the gateway, with a low-overhead protocol.
 This communication can be done in different channels which are also shared by devices using other networks.
@@
 %
-This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent objects at least do not reduce their successful communication rate in comparison to the naive objects.
-This also shows that in case of non-uniform stationary traffic, MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent objects.
+This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent devices at least do not reduce their successful communication rate in comparison to the naive devices.
+This also shows that in case of non-uniform stationary traffic, MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent devices.
+%
+The benefits are twice and of primary importance for IoT networks:
+the proposed approach can mitigate RF collisions,
+and enhance intelligent device battery lifetime if they do retransmissions.
 
@@
 
-We implement a PHY/MAC layers solution in order to improve the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
-We choose to implement our own physical layer and protocol, for both clarity and conciseness.
+We implement a simple PHY/MAC layers solution, in order to demonstrate the possibility of improvement of the performance of IoT communications in unlicensed bands.
+We could have used any physical layer and any ALOHA-based protocol.
+We choose to implement our own physical layer and protocol, for both clarity and conciseness, and because developing a complete IoT network protocol stack is no more a research work and would have fall outside of the scope of this thesis.
 
@@
 Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
-For example, this index can be a simple QPSK symbol ($\pm1\pm1j$).
+For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a longer sequence if the PoC have to consider more users\footnote{~With $\ell$ QPSK symbols, we can fit at most $4^{\ell}/2$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway.}.
 Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
 Thanks to this index, we can have several devices communicating with the same gateway.
+
+\begin{figure}[!b]
+    \centering
+    \includegraphics[width=0.70\linewidth]{our-demo.eps}
+    \caption{Schematic of our implementation that presents the role of each USRP platform.}
+    \label{fig:42:our_demo}
+\end{figure}
 
@@
 As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP.
-The gateway, a USRP which emulates the interfering traffic, and at least one dynamic device.
+The gateway, a USRP which emulates the interfering traffic, and at least one dynamic device, each of them having its own USRP and its own learning algorithm.
 % Cables and laptop
@@
 
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
-    \includegraphics[width=0.70\linewidth]{our-demo.eps}
-    \caption{Schematic of our implementation that presents the role of each USRP card.}
-    \label{fig:42:our_demo}
+    \includegraphics[width=0.70\linewidth]{SCEE_TestBed1.jpg}
+    \vspace*{20pt}
+    \includegraphics[width=0.70\linewidth]{SCEE_TestBed2.jpg}
+    \caption{Two pictures showing the SCEE testbed \cite[Appendix~3]{Bodinier17}, taken in $2018$.}
+    \label{fig:42:photosSCEETestBed}
 \end{figure}
@@
 The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
+The pictures in Figure~\ref{fig:42:photosSCEETestBed} shows the testbed used for these experiments.
 % Octoclock
-To ease the synchronization in both time and frequency between the boards representing the dynamic objects and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
+To ease the synchronization implementation in both time and frequency between the boards representing the dynamic devices and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
 % https://www.ettus.com/product/details/OctoClock-G
-and coaxial cables connecting every card to the Octoclock for time (PPS) and frequency synchronization, but this is not mandatory.
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=0.75\linewidth]{SCEE_TestBed1.jpg}
-    \vspace*{20pt}
-    \includegraphics[width=0.75\linewidth]{SCEE_TestBed2.jpg}
-    \caption{Two pictures showing the SCEE testbed \cite[Appendix~3]{Bodinier17}.}
-    \label{fig:42:photosSCEETestBed}
-\end{figure}
+and coaxial cables connecting every platform to the Octoclock for time (PPS) and frequency synchronization, but this is not mandatory.
 
@@
 
-We used the GNU Radio Companion software (GRC, version $3.7$, $2017$),
-and for the demonstration the laptop runs
-one GRC design to configure and control each USRP card.
-As such, a single laptop can run in parallel the control program of any number of boards\footnote{~Even if in practice, maximum efficiency is kept as long as there is not more than one GRC design by CPU core.}.
+We used the GNU Radio Companion software (GRC, version $3.7$ in $2017$),
+and a laptop runs a GRC design to configure and control each USRP platform.
+As such, one laptop can run in parallel the control program of any number of boards\footnote{~Even if in practice, maximum efficiency is kept as long as there is not more than one GRC design by CPU core.}.
 
@@
 The GNU Radio infrastructure is written entirely in \texttt{C++}, and many of the user tools are written in Python.
-GNU Radio Companion is a graphical UI used to develop GNU Radio applications:
+GNU Radio Companion is a graphical user interface (UI) used to develop GNU Radio applications:
 GRC is effectively a Python code-generation tool.
-when a flow-graph is compiled in GRC, a Python code is produced, which can be executed to connect to the USRP,
+When a flow-graph is compiled in GRC, a Python code is produced, which can be executed to connect to the USRP,
 create the desired GUI windows and widgets, and create and connect the blocks in the flow-graph.
+
+Illustrations of the flow-graph for the three components of the presented demonstration are included in Appendix~\ref{sec:4:IllustrationFlowcharts}, in Figures~\ref{fig:4app:USRP_TX_PU__v1__simple_grc}, \ref{fig:4app:USRP_RX_BTS__v1__simple_grc} and \ref{fig:4app:USRP_TX_SU__v1__simple_grc}.
 
@@
 We have designed a user interface in order to visualize the results obtained  with our experimental demonstration. This user interface is shown in Figure~\ref{fig:42:UI}.
-We can see that it is made of three parts, one for each USRP, as highlighted in \textcolor{darkred}{red}:
+We can see that it is made of three parts, one for each USRP, as highlighted in \textcolor{darkred}{circled red numbers}.
 
-\begin{figure}[!t]
+\begin{figure}[!h]
     \centering
@@
 The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.
+Messages of the random traffic are shorter in time by purpose (they could be coming from othe rIoT standard), in order to distinguish them from an intelligent device traffic on the ``waterfall'' visualizations of the traffic.
 
@@
 At the top left, we observe the constellation of the transmitted packet \emph{(a)}.
-At the bottom left, we have a time/frequency view of the lasts packets transmitted by the object \emph{(b)}.
-We can see, in this view that the object transmitted its last $9$ packets in the two best channels (channel $\#3$ and $\#4$).
-Then, at the top right of this interface \emph{(c)}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this object (\textcolor{darkred}{red}) and the acknowledgements sent by the gateway (\textcolor{darkblue}{blue}).
+At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \emph{(b)}.
+We can see, in this view that the device transmitted its last $9$ packets in the two best channels (channel $\#3$ and $\#4$).
+Then, at the top right of this interface \emph{(c)}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the gateway (\textcolor{darkblue}{blue}).
+Colors in the ``waterfall'' represent the RF power level received at the device antenna.
+Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (emission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \emph{(c)} in part (2)).
 Finally, at the bottom right \emph{(d)}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).
@@
 % ----------------------------------------------------------------------
-\subsection{Empirical results}
+\subsection{Experimental results}
 \label{sub:42:results}
@@
 
-We compare the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
-For one dynamic object, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
+We compare in this PoC the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
+For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
 Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} shows the results averaged on $10$ repetitions using the same conditions.
@@
 Each experiment takes about half a day,
-as we make objects generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
+as we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
 Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
-Each object was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_i)_i = [15\%, 10\%, 2\%, 1\%]$.
-The occupancy rate of a channel denote the mean occupancy, and it is implemented using one USRP card that runs the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).
+Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_i)_i = [15\%, 10\%, 2\%, 1\%]$.
+The occupancy rate of a channel, which denotes the mean occupancy, is implemented using the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).
 % They did not overlap each other, as
-When facing the same stationary background traffic, we see that the learning objects are both very quickly more efficient than the naive uniform object.
+When facing the same stationary background traffic, we see that the learning devices are both very quickly more efficient than the naive uniform device.
 We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
@@
 Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
+However, our experimental testbed can not run hundreds of intelligent devices.
 Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
@@
     \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
-    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success.
+    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
 Similar gains of performance were obtained in many different scenarios.}
@@
 %         \includegraphics[height=7cm]{Object2_UniformAccess__UI.png}
-%         %\caption{Object using uniform access.}
+%         %\caption{Device using uniform access.}
 %         \label{fig:42:comparing_UniformAccess_and_UCB__1}
@@
 %         \includegraphics[height=7cm]{Object1_UCB__UI.png}
-%         %\caption{Object using UCB.}
+%         %\caption{Device using UCB.}
 %         \label{fig:42:comparing_UniformAccess_and_UCB__2}
 %     \end{subfigure}
-% \caption{Comparing the success rate of an object using uniform access (left, $41\%$ of success here) and an object using the UCB algorithm (right, already $60\%$ of success).
-We see the learning object being much more present in the channels $\#4$ and $\#3$ (\textcolor{darkblue}{blue chart}, bottom right corner), as these channels are the less perturbed by the random interfering traffic, in this example.}
+% \caption{Comparing the success rate of an device using uniform access (left, $41\%$ of success here) and an device using the UCB algorithm (right, already $60\%$ of success).
+We see the learning device being much more present in the channels $\#4$ and $\#3$ (\textcolor{darkblue}{blue chart}, bottom right corner), as these channels are the less perturbed by the random interfering traffic, in this example.}
 % \label{fig:42:comparing_UniformAccess_and_UCB}
@@
 The demonstration only requires a laptop and free tools,
-as the laptop should run a GNU/Linux distribution (like Ubuntu or Debian).
+as the laptop should run a GNU/Linux distribution (like Ubuntu or Debian),
+in addition to USRP platforms from Ettus Research.
 
@@
 It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
-It also shows the connections between the USRP boards, the Octoclock, the master laptop etc.
+It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
 

### 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Using \begin{framed}...\end{framed} for all algorithms.
Done for now for Chapter4

@@
 In this section, we extend the model presented before in order to take into account the possibility for retransmission of a message after a collision.
-This extension is of major importance as most protocols for real-world IoT networks use retransmissions, and it is for instance the case of the LoRa standard.
+This extension is of major importance as most protocols for real-world IoT networks can use retransmissions, and it is for instance the case of the LoRa standard.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.
+However, the price to be paid is a shorter battery lifetime for IoT devices.
+So this approach would be used only for devices with strong delivery constraints (\eg, for health-care applications), or that can refuel their energy over time (\eg, for robotic applications).
 
@@
 %
-None of the above mentioned works discusses in detail the impact of retransmissions on the performance of MAB learning algorithms as we do now.
+None of the mentioned works discuss in detail the impact of retransmissions on the performance of MAB learning algorithms as we do now.
 
@@
 regardless the type of devices, each of them follows a slotted ALOHA protocol \cite{Roberts75}, and has a probability $p>0$ to transmit a packet in a time slot.
-We make the hypothesis that the transmission is successful if the channel is available, otherwise upon radio collision, these devices will attempt to retransmit their packet up-to $M$ times, with $M \in\mathbb{N}$.
+We make the hypothesis that the transmission is successful if the channel is available, otherwise upon radio collision.
+The novelty compare to the previous model is that
+in case of RF (uplink or downlink) collision that prevent the device to receive the \emph{Ack} from the gateway,
+these devices will attempt to retransmit their packet up-to $M$ times, with $M \in\mathbb{N}$.
 It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\llbracket 0, m-1 \rrbracket$, where $m \in\mathbb{N}, m>0$ is the length of the back-off interval
@@
 The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Figure~\ref{fig:43:Markov_model}.
-When a device have a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
+When a device has a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
 At each time slot, a transition from an idle state to a transmission state (denoted as \texttt{Trans.}) occurs if a packet transmission is required, while waiting states (denoted as \texttt{Wait}), correspond to a $m$ back-off interval.
 
-A device aims to select a channel with the highest probability of successful transmissions, for which it resorts to a reinforcement learning approach, formulated as a MAB problem.
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.70\linewidth]{Markov_model.eps}
+	% \includegraphics[width=0.70\linewidth]{Markov_model-eps-converted-to.pdf}
+	\caption{The Markov model of the behavior of all devices paired to the considered IoT network using the ALOHA protocol.}
+	\label{fig:43:Markov_model}
+\end{figure}
+
+A device aims to select a channel with the highest probability of successful transmissions, for which it uses a reinforcement learning approach, again formulated as a MAB problem.
 Contrarily to what may appear at first sight,
-the goal is not to minimize the number of retransmission, but to maximize the probability of successful transmission, considering both the first transmission of a message and the retransmissions of the same message, because the objective of each device is to maximize its battery life by minimizing its number of failed transmissions.
+the goal is not to minimize the number of retransmissions, but to maximize the probability of successful transmission, considering both the first transmission of a message and the retransmissions of the same message, because the objective of each device is to maximize its battery life by minimizing its number of transmissions.
 % , where each channel (also called arms) is viewed as a gambling machine (bandit), and each bandit has a \emph{reward}.
@@
 % to address several MAB algorithms.
-
-\begin{figure}[htp!]  % [htbp]
-	\centering
-	% \includegraphics[width=0.65\linewidth]{Markov_model.eps}
-	\includegraphics[width=0.70\linewidth]{Markov_model-eps-converted-to.pdf}
-	\caption{The Markov model of the behavior of all devices paired to the considered IoT network using the ALOHA protocol.}
-	\label{fig:43:Markov_model}
-\end{figure}
 %
@@
 
-When a device experiments a collision, it goes in a back-off state to retransmit the same packet on a channel.
+We consider IoT devices with a constraint on their QoS, imposing the successful delivery of their messages.
+When such an IoT device experiments a collision, it goes in a back-off state to retransmit the same packet on the same or another channel.
 If all devices remain in the same channel for retransmissions, it could result in a sequence of successive collisions with the same packets' devices that previously collided.
@@
 
-As it is well known \cite{Abramson1970,Roberts75}, having a collision during an access time can be overcome by a retransmission procedure (this can take several retransmission attempts).
-What interest us, here, is to obtain a mathematical approximation of the collision probability at the second transmission slot $p_{c1}$, as a function of the first collision probability $p_{c}$.
+It is well known \cite{Abramson1970,Roberts75}
+that having a collision during an access time can be overcome by a retransmission procedure (this can take several retransmission attempts).
+Our goal here is to obtain a mathematical approximation of the collision probability at the second transmission slot $p_{c1}$, as a function of the first collision probability $p_{c}$.
 
-We make two approximations $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as,
+We make two approximations $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as (they are hypotheses on which the rest of this section is built),
 \begin{itemize}
@@
 Moreover, we suppose that the number of devices involved in a collision is small in comparison to the total number of devices.
+	This is very realistic as a very small proportion of devices transmit at the same period, due to their low duty cycle.
+
 	\item $\mathcal{H}_{2}$:
-	The total number of back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
+	The total number of back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions
+	(otherwise, its battery life can be threatened if it does not have re-fueling capabilities).
 \end{itemize}
@@
 Considering one device and one channel,
-we denote $x_t^i$ the probability that it is transmitting a packet for the $i+1$ time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$),
-and let $x_t = \sum_{i=0}^{M-1}x_t^i$ be the probability that it transmits a packet.
+we denote $x_t^i$ the probability that it is transmitting a packet for the $(i+1)$-th time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$),
+and we denote $x_t = \sum_{i=0}^{M-1}x_t^i$ the probability that it transmits a packet (\ie, just the sum on $i$ of $x_t^i$).
 We consider $N > 0$ active devices following the same policy.
@@
 	\centering
+    \begin{framed}
 	\begin{algorithm}[H]
 		% \begin{small} % XXX remove if needed
-		\For(){$t = 0, \dots, T$}{
+		\For(){$t = 1, \dots, T$}{
 			Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
@@
 	\end{algorithm}
+	\end{framed}
 \end{figure}
@@
 	\centering
+    \begin{framed}
 	\begin{algorithm}[H]
 	% \begin{small} % XXX remove if needed
-	\For(){$t = 0, \dots, T$}{
+	\For(){$t = 1, \dots, T$}{
 			\uIf{First packet transmission}{
@@
 It is described in Algorithm~\ref{algo:43:TwoUCB}, and it is still a practical approach, since the storage requirements and time complexity remains linear w.r.t.
-the number of channels $K$ (\ie, of order $\mathcal{O}(K)$).
-
+the number of channels $K$ (\ie, $\mathcal{O}(K)$).
+%
 Note that, we use the subscript $({}^r)$ to denote the variables
@@
 	\centering
+    \begin{framed}
 	\begin{algorithm}[H]
 		% \begin{small} % XXX remove if needed
-		\For(){$t = 0, \dots, T$}{
+		\For(){$t = 1, \dots, T$}{
 			\uIf{First packet transmission}{
@@
 	\centering
+\begin{framed}
 	\begin{algorithm}[H]
@@
 		\For(\tcp*[f]{At every time step})
-			{$t = 0, \dots, T$}{
+		{$t = 1, \dots, T$}{
 			\uIf{First packet transmission}{
@@
 As such, we only consider a delay of $\Delta=100$.}.
-
+%
 Note that, we use the subscript $({}^d)$ to denote the variables
@@
 	\centering
+	\begin{framed}
 	\begin{algorithm}[H]
@@
 			}
-				\Else(\tcp*[f]{retransmission with delayed $\mathrm{UCB}^d$}){
+			\Else(\tcp*[f]{Packet retransmission with delayed $\mathrm{UCB}^d$}){
 				Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + B^d_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
@@
 
-A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve).
+A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve), which is (still) the current state-of-the-art in IoT networks.
 The improvement of the heuristics over the non-learning approach is clear, and for every heuristic that uses a kind of learning mechanism it can observed the successful transmission rate increases rapidly (or equivalently the PLR decreases).
 Moreover, all of these approaches show a fast convergence making them suitable for the targeted application.
-It is also worths mentioning that the employment of the same \UCB{} algorithm for retransmissions, denoted here as ``Only \UCB{}'', achieves a better performance, while a ``Random'' retransmission features a slight degradation.
+It is also worth mentioning that the employment of the same \UCB{} algorithm for retransmissions, denoted here as ``Only \UCB{}'', achieves a better performance, while a ``Random'' retransmission features a slight degradation.
 This result can be explained as follows: the loss of performance related to the separation of information for several algorithms is greater than the gain obtained by considering the first transmissions and retransmissions separately.
@@
 
-The conclusions we can draw from depicted results are twofold.
-First, MAB learning algorithms are very useful to reduce the collision rate in LPWA networks, a gain of up-to $30\%$ of successful transmission rate is observed after convergence.
-A second conclusion that can be highlighted is that, using learning mechanisms for retransmissions can be an interesting way to reduce collisions in networks with massive deployments of IoT as this can be checked in Figure~\ref{fig:43:mainExperiment2}, where the random retransmission heuristic is not very advantageous in front of the  \UCB-based approaches that use learning for channel selection during the retransmission procedure.
+The conclusions we can draw from these results are twofold.
+Firstly, MAB learning algorithms are very useful to reduce the collision rate in LPWA networks, a gain of up-to $30\%$ of successful transmission rate is observed after convergence.
+Secondly, using learning mechanisms for retransmissions can be a simple yet efficient and interesting way to reduce collisions in IoT networks, even in networks with massive deployments of IoT as this can be checked in Figure~\ref{fig:43:mainExperiment2}, where the random retransmission heuristic is greatly outperformed by the any of the \UCB-based approaches, that use learning for channel selection during the retransmission procedure.
+With $10\%$ to $30\%$ occupancy rates the considered example of IoT network can indeed be considered as an IoT network with a massive deployment of devices.
 

### 2019-04-01 2-Chapters/4-Chapter/chapter4.tex
Done for now for Chapter4

@@
 
-On the first hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT objects, at the cost of a lower Quality of Service (QoS).
+On the first hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
 On the other hand, the second model is more interesting if the main target is an improvement of the QoS of the IoT networks rather than a longer battery life.
@@
 unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
-This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.
+This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.
 
@@
 As end-devices battery life is a key constraint of IoT networks,
+and as IoT networks are decentralized (\ie, the devices initiate transmissions),
 this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
 %
-In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
+In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
+Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
 We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
 Kaufmann12Thompson}, for the first model.
-For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for solve the problem of learning in order to retransmit efficiently.
+For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.
 
 As detailed in Chapter~\ref{chapter:1},
-MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
-% For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
+For example,
+proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
+and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
+Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
 Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
-Many recent works show that stationary MAB algorithms work well for real-world radio signal.
-However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
+% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
 
@@
 Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
-As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary, making it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
+As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary.
+This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
 We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
@@
 
-We give in this section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.
+We give in this section conclusions about the two models presented above, one for long-life IoT devices (without retransmission) and for improved QoS devices (\ie, with retransmissions), and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.
 
@@
 with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
-Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
+Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
+This approach can also extend IoT devices battery life, as they suffer less from failed transmissions.
 We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the presences many ``intelligent'' end-devices which follow a random activation process.
@@
 %
-By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.
+By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.
 
@@
 In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
+If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more interesting.
 The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
@@
 Possible future extensions of our demonstration include the following points.
-We could consider more dynamic objects (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
+We could consider more dynamic devices (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more devices with the same card.
 We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
@@
 In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our demonstration \cite{Besson2019WCNC}.
-Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes, and Christophe Moy at IETR and University Rennes 1.
+Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes.
+% and Christophe Moy at IETR and University Rennes 1.
 %

### 2019-04-01 2-Chapters/6-Chapter/chapter6.tex
WOrking on the whole document but not yet Chapter5

@@
 %
-We also include additional details about our policy \GLRklUCB, including a sensibility analysis of its two parameters, a numerical evaluation of its time and memory costs, as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
+We also include additional details about our policy \GLRklUCB, including a sensitivity analysis of its two parameters, a numerical evaluation of its time and memory costs, as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
 

### 2019-04-01 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Minitoc shows only Section, it's better.
Add Figures for Appendix of Chapter4
WOrking on the whole document but not yet Chapter5

@@
     \begin{align*}
-        \log\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)
+        & \log\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau) \\
         &= \sum_{i=1}^{\tau} X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_i \log(\mu_1) + (1-X_i) \log(1-\mu_1) \\
@@
         \KwIn{\emph{Problem parameters}: $T\in\N^*$, $K\in\N^*$\;}
-        \KwIn{\emph{Algorithm parameters}: exploration probability $\alpha \in (0, 1)$, confidence level $\delta>0$\;}
+        \KwIn{\emph{Parameters}: exploration probability $\alpha \in (0, 1)$, confidence level $\delta>0$\;}
         \KwIn{\emph{Option}: \textbf{Local} or \textbf{Global} restart\;}
@@
 With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
-the index is defined using the binary relative entropy as
+the index is defined using the binary relative entropy $\kl$ as
 %
@@
 %-----------------------------------------------------------------------------
-\subsection{Proof of the regret upper-bounds}
+\section{Proof of the regret upper-bounds}
 \label{sec:6:proofRegret}
-
-\TODOL{Write intro}
 
@@
 
-\paragraph{Upper bound on term $\bm{(A)}$.}
+\paragraph{Upper bound on term $(A)$.}
 
@@
 
-\paragraph{Upper bound on term $\bm{(B)}$.}
+\paragraph{Upper bound on term $(B)$.}
 %
@@
 Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instances.
+
+\TODOL{I could add the following algorithms, just to illustrate more:
+
+- OracleRestart with UCB
+- M- and CUSUM- with UCB (as defined in the papers)
+- Exp3.S
+}
 

### 2019-04-01 4-Listofpublications/Listofpublications.tex
Show change.
WOrking on the whole document but not yet Chapter5

@@
 \begin{etaremune}
+\item
+	\href{https://perso.crans.org/besson/articles/BMBBM__IEEE_WCNC__2019.pdf}{Upper-Confidence
+	Bound for Channel Selection in LPWA Networks with Retransmissions},
+	by \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L. Besson}, J.
+	Manco-Vasquez,
+  \href{https://moychris.wordpress.com/}{C. Moy}.
+  \href{https://sites.google.com/view/wcncworkshop-motion2019/}{MoTION Workshop} at
+  \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
+  Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02049824}{HAL-02049824}.
 \item
@@
   \href{https://hal.inria.fr/hal-01575419/document}{Multi-Armed Bandit
-  Learning in IoT Networks and non-stationary settings}, par
+  Learning in IoT Networks and non-stationary settings}, by
   \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L.
@@
   (\emph{Conference on Cognitive Radio Oriented Wireless Networks}),
-  Lisbonne, Portugal, septembre 2017,
-  \href{https://hal.inria.fr/hal-01575419}{HAL-01575419}, \textbf{prix
-  du meilleur article}.
+  Lisbonne, Portugal, September 2017,
+  \href{https://hal.inria.fr/hal-01575419}{HAL-01575419}, \textbf{Best Paper Award}.
 \end{etaremune}
@@
 Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}
-	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
-	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, Mar 2017
+	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, March 2017
 	\item	R.
@@
 \item
-	\href{https://perso.crans.org/besson/articles/BMBBM__IEEE_WCNC__2019.pdf}{Upper-Confidence
-	Bound for Channel Selection in LPWA Networks with Retransmissions},
-	par \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi}, \textbf{L. Besson}, J.
-	Manco-Vasquez,
-	\href{https://moychris.wordpress.com/}{C. Moy}. Septembre 2018.
+  \href{https://hal.inria.fr/hal-XXX/document}{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation}, by \href{https://moychris.wordpress.com/}{C.
+Moy} \& \textbf{L.
+Besson}, March 2019.
+  % \href{https://hal.inria.fr/hal-XXX}{HAL-XXX}.
+\item
+  \href{https://hal.inria.fr/hal-02006471/document}{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires}, by \textbf{L. Besson} \&
+  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
+  Kaufmann}, March 2019,
+  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
+\item
+  \href{https://hal.inria.fr/hal-02006471/document}{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits}, by \textbf{L. Besson} \&
+  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
+  Kaufmann}, February 2019,
+  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
 \item
@@
 	Bandits (MAB) Algorithms in Python}, by \textbf{L. Besson},
-	développement depuis octobre 2016 et encore actif,
-	\href{https://hal.inria.fr/hal-01840022}{HAL-01840022}. Code sur
+	active development since 2016,
+	\href{https://hal.inria.fr/hal-01840022}{HAL-01840022}. Code on
 	\href{https://GitHub.com/SMPyBandits/SMPyBandits}{GitHub.com/SMPyBandits}
-	(environ 40000 lignes), documentation sur
+	(about 40000 lines), documentation and tutorials on
 	\href{https://SMPyBandits.rtfd.io}{SMPyBandits.rtfd.io}.
@@
 	\href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
-	Kaufmann}, septembre 2018,
+	Kaufmann}, September 2018,
 	\href{https://hal.inria.fr/hal-01736357}{HAL-01736357}.
@@
 	\href{https://hal.inria.fr/hal-01847480/document}{A Note on the Ei
-	Function and a Useful Sum-Inequality}, by \textbf{L. Besson}, février
+	Function and a Useful Sum-Inequality}, by \textbf{L. Besson}, February
 	2018 \href{https://hal.inria.fr/hal-01847480}{HAL-01847480}.

### 2019-04-01 2-Chapters/2-Chapter/chapter2.tex
Still fighting against my template...
Okay for chapters and parts
Small modifications on all chapters

@@
 \minitoc
+
+Abstract for Chapter 2.
+
 \newpage
@@
 % ----------------------------------------------------------------------------
-\section{Different approaches on algorithm selection, or ``How to choose your preferred bandit algorithm?''}
+\section{Different approaches on algorithm selection}
 \label{sec:2:chooseYourPreferredBanditAlgorithm}
@@
 
-A last solution is online algorithm selection, inspired from expert aggregation.
-Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292
+% A last solution is online algorithm selection, inspired from expert aggregation.
+% Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292
 

### 2019-04-01 2-Chapters/3-Chapter/chapter3.tex
Small modifications on all chapters

@@
 %
-More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markovian or non-stationary).
+More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
 
@@
 Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
+\TODOL{Do we really have time/space to talk about Markov models?}
 Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.
 
-\newpage
+% \newpage
 % Write miniTOC just after the title
@@
 
-SMPyBandits does not aim at being blazing fast or perfectly memory efficient, and comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
+SMPyBandits does not aim at being blazing fast or optimal in terms of memory usage, and it comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
 Even if some critical parts are also available as a \texttt{C} Python extension, and even by using Numba \cite{numba} whenever it is possible, if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia.
@@
 
-In \cite{Besson2018WCNC} (see Chapter~\ref{chapter:2}), we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation.}.
-We designed a variant of the Exp3 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
-Aggregating experts is a well-studied idea in sequential learning and in machine learning in general.
+In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2}, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
+We designed a variant of the Exp4 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
+% Aggregating experts is a well-studied idea in sequential learning and in machine learning in general.
 We showed that it can be used in practice to select on the run the best bandit algorithm for a certain problem from a fixed pool of experts.
@@
 
-- For \cite{Besson2018ALT} (see Chapter~\ref{chapter:5}), we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5}, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
@@
 In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
-In sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
+and lots of bibliographic referencesIn sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
 A well-known trick for transforming any non-anytime algorithm to an anytime variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
@@
 
-In \cite{Besson2019GLRT} (see Chapter~\ref{chapter:6}), we used SMPyBandits for piece-wise stationary MAB models (single-player).
+In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6}, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
 We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
-XXX
-
-\TODOL{Write a better explanation of this last paper.}
+We designed the \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.GLR_UCB.html}{GLRklUCB}} policy, and implemented most of the policies from state-of-the-art research on passively or active adaptive policies, both adversarial- or stochastic-based, designed to tackle the piece-wise stationary MAB problems.
+We also implemented a large benchmark of problems.
 

### 2019-04-01 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Small modifications on all chapters

@@
     \label{fig:41:100intelligent}}
-    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
+    % FIXME use a shorter caption for the list of figures in Appendix of the thesis: it takes lots of lines to have these (a)...(d) shown in list of figures
+    \caption[Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:from10to100}

### 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Small modifications on all chapters

@@
     \centering
-    \includegraphics[width=0.85\textwidth]{UI.eps}
+    \includegraphics[width=1.00\textwidth]{UI.eps}
     \caption{User interface of our demonstration.}
@@
 %
-A $6$-minute \textbf{video} showing our demonstration is available online at \texttt{\url{youtu.be/HospLNQhcMk}}.
-It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
+As shown below in Figure~\ref{fig:42:screenshotDemoYouTube},
+we made a $6$-minute \textbf{video} to show our demonstration online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
 It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
+
+\begin{figure}[!h]
+	\centering
+    % \includegraphics[width=0.90\textwidth]{Images/screenshotDemoYouTube.png}
+    \includegraphics[width=0.90\textwidth]{2-Chapters/4-Chapter/Images/screenshotDemoYouTube.png}
+    \caption{Screenshot of the \textbf{video} of our demonstration, \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.}
+    \label{fig:42:screenshotDemoYouTube}
+\end{figure}
 

### 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
Done for now for Chapter4
Small modifications on all chapters

@@
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_PU__v1__simple_grc.png}
-    \caption{The \textbf{random traffic generator}.
+    \caption[The \textbf{random traffic generator} flow-graph.]{The \textbf{random traffic generator}.
 \texttt{Generator} is the only hand-written block, which is configured with a list of active channels, a list of occupancy rate, and constants about the PHY layer (preamble length and data length).
@@
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_RX_BTS__v1__simple_grc.png}
-    \caption{The \textbf{IoT gateway}.
+    \caption[The \textbf{IoT gateway} flow-graph.]{The \textbf{IoT gateway}.
 The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length).
@@
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_SU__v1__simple_grc.png}
-    \caption{The \textbf{IoT dynamic object}.
+    \caption[The \textbf{IoT dynamic device} flow-graph.]{The \textbf{IoT dynamic device}.
 The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels.
 Most blocks need to know constants about the PHY layer (preamble length and data length).
-Any of the IoT dynamic objects use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
+Any of the IoT dynamic devices use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
     \label{fig:4app:USRP_TX_SU__v1__simple_grc}

### 2019-04-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand
Using \begin{framed}...\end{framed} for all algorithms.
Moved a lot of appendix in main text for Chapter 5. See #15

@@
 
-Multi-player Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
-Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms.
-Existing works assume that \emph{sensing information} is available to the algorithm.
-Under this assumption, we improve the state-of-the-art lower bound for the regret of any decentralized algorithms of a certain class.
-We introduce two algorithms, \RandTopM{} and \MCTopM{}, that are shown to empirically outperform existing algorithms. Moreover, we provide strong theoretical guarantees for these algorithms, including
-a notion of asymptotic optimality in terms of the number of selections of bad arms.
-We then introduce a promising heuristic, called \Selfish{}, that can operate
-without sensing information, which is crucial for emerging applications to Internet of Things networks. We investigate the empirical performance of this algorithm
-and provide some first theoretical elements for the understanding of its behavior.
+% Multi-player Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
+% Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms.
+% Existing works assume that \emph{sensing information} is available to the algorithm.
+Under this assumption, we improve the state-of-the-art lower bound for the regret of any decentralized algorithms (of a certain class).
+% We introduce two algorithms, \RandTopM{} and \MCTopM{}, that are shown to empirically outperform existing algorithms.
+Moreover, we provide strong theoretical guarantees for these algorithms, including
+% a notion of asymptotic optimality in terms of the number of selections of bad arms.
+% We then introduce a promising heuristic, called \Selfish{}, that can operate
+% without sensing information, which is crucial for emerging applications to Internet of Things networks. We investigate the empirical performance of this algorithm
+% and provide some first theoretical elements for the understanding of its behavior.
 
@@
 
-For the development of CR, a crucial step is to insert \emph{multiple} $M \geq 2$ smart devices in the \emph{same} background traffic.
+For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
 With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
-%who essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
+They essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
 % In the context of CR, this first extension can model a centralized agent taking decisions for end-devices, for instance a base station affecting devices to channels.
@@
 
-In this section, we take a step back and present the different feedback levels possible for multi-player MAB algorithms.
+In this chapter, we take a step back and present the different feedback levels possible for multi-player MAB algorithms.
 For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees.
@@
 
-The rest of the section is organized as follows.
+\paragraph{Outline.}
+%
+The rest of the chapter is organized as follows.
 We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
-The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms}, with the result of our experimental study reported in
-Section~\ref{sec:5:experiments}. Theoretical elements are then presented in Section~\ref{sec:5:upperbounds}.
+The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms},
+for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
+Finally, the result of our experimental study reported in Section~\ref{sec:5:experiments}.
 % and finally Section~\ref{sec:5:conclusion} concludes.
@@
 At time step $t \in\mathbb{N},t\geq1$, player $j$ selects an arm $A^j(t)$, independently from the other players' selections.
+
+\begin{definition}
   A \emph{collision} occurs at time $t$ if at least two players choose the same arm.
@@
   that respectively indicate that a collision occurs at time $t$ for player $j$ and that a collision occurs at time $t$ on arm $k$.
+\end{definition}
+
 Each player $j$ then receives (and observes) the \emph{binary rewards}
@@
 \end{equation}
-In words, she receives the reward of the selected arm if she is the only one to select this arm, and a reward zero otherwise\footnote{This provides another reason to focus on the Bernoulli model.
+In words, she receives the reward of the selected arm if she is the only one to select this arm, and a reward zero otherwise.
+This provides another reason to focus on the Bernoulli model.
 It is the hardest model, in the sense that receiving a reward zero is not enough to detect collisions.
 For other models, the data streams $(Y_{k,s})_s$ are usually continuously distributed, with no mass at zero.
-Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.}.
-Other models for rewards loss have been proposed in the literature (\eg, the reward is randomly allocated to one of the players selecting it), but we focus on full reward occlusion in this Section.
+Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.
+Other models for rewards loss have been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it, but we focus on full reward occlusion in this chapter.
 
-A multi-player MAB strategy is a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for each player, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
+A multi-player MAB strategy is a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
 First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
@@
 More precisely, depending on the application, one may consider the following three observation models, \modelun, \modeldeux{} and \modeltrois.
-%
-% If the arms have continuous distributions (or are such that  $\Pr(Y_{k,s}=0)=0$), the \emph{sensing information} $Y_{A^j(t),t}$ and \emph{collision information} $C^j(t)$ can always be extracted from the reward information.
-But for Bernoulli distribution, one may consider the following three observation models, that are not equivalent:
+
+If the arms have continuous distributions (or are such that  $\Pr(Y_{k,s}=0)=0$), the \emph{sensing information} $Y_{A^j(t),t}$ and \emph{collision information} $C^j(t)$ can always be extracted from the reward information.
+But for Bernoulli distributions, one may consider the following three observation models, that are not equivalent:
+
 \begin{itemize}
@@
 \end{itemize}
-\noindent Under each of these three models, we define $\cF^j_t$ to be the filtration generated by the observations gathered by player $j$ up to time $t$ (which contains different information under models \modelun, \modeldeux{} and \modeltrois).
+
+Under each of these three models, we define $\cF^j_t$ to be the filtration generated by the observations gathered by player $j$ up to time $t$ (which contains different information under models \modelun, \modeldeux{} and \modeltrois).
 While a \emph{centralized} algorithm may select the vector of actions for all players $(A^1(t),\dots,A^M(t))$ based on all the observations from $\bigcup_j \cF^j_{t-1}$, under a \emph{decentralized} algorithm the arm selected at time $t$ by player $j$ only depends on the past observation of this player.
@@
   $\{1, \dots, K\} \setminus \Mbest$.
-  %
+
   Note that they are both uniquely defined if $\mu_M^* > \mu_{M+1}^*$.
@@
 and the cumulated reward of strategy $\rho$:
+
+\begin{definition}
+  The excepted centralized multi-player regret is defined by
   \begin{equation}\label{eq:5:regret}
@@
   \end{equation}
-Maximizing the expected sum of global reward of the system is indeed equivalent to minimizing the regret, and we now investigate the best possible \emph{regret rate} of a decentralized multi-player algorithm.
+\end{definition}
+
+With this definition, maximizing the expected sum of global reward of the system is indeed equivalent to minimizing the regret, and we investigate the best possible \emph{regret rate} of a decentralized multi-player algorithm in the next section.
 
@@
 % \begin{definition}[]\label{def:5:nbCollisions}
-  Let $\cC_k(T)$ be the \emph{number of colliding players}\footnote{When $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one.
-So $\cC_k(T)$ counts the total \emph{number of colliding players} rather than the number of collision events.
-Hence there is small abuse of notation when calling it a number of collisions.}
-  on arm $k\in\{1,\dots,K\}$ up to horizon $T$:
-  \vspace*{-5pt}  % XXX
+  Let $\cC_k(T)$ be the \emph{number of colliding players} on arm $k\in\{1,\dots,K\}$ up to horizon $T$:
+  % \vspace*{-5pt}  % XXX
   \begin{equation}
@@
 
+Note that when $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one.
+So $\cC_k(T)$ counts the total \emph{number of colliding players} rather than the number of collision events.
+Hence there is small abuse of notation when calling it a number of collisions.
+
 Letting $\cP_M = \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
@@
   For any bandit instance $\boldsymbol{\mu}\in\cP_M$ such that $\mu_M^* > \mu_{M+1}^*$, it holds that
-  \hfill{}
-  (Proved in App.\ref{proof:5:DecompositionRegret})
-  \vspace*{-5pt}  % XXX
-  \begin{small}
-    \begin{equation*}\label{eq:5:termeundeuxtermetrois}
-      R_T(\boldsymbol{\mu}, M, \rho) =
-      \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[T_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}}
-      + \underbrace{\sum_{k \in \Mbest} (\mu_k -  \mu_M^*) (T - \E_{\mu}[T_k(T)])}_{\emph{\mytag{(b)}{eq:5:term2}}}
-      + \underbrace{\sum_{k=1}^{K} \mu_k \E_{\mu}[\cC_k(T)]}_{\emph{\mytag{(c)}{eq:5:term3}}}.
-    \end{equation*}
-  \end{small}
+  % \vspace*{-5pt}  % XXX
+  % \begin{small}
+    \begin{align}\label{eq:5:termeundeuxtermetrois}
+      R_T(\boldsymbol{\mu}, M, \rho) &=
+      \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[T_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}} \\
+      & \;\;\;\;\;\;
+      + \;\; \underbrace{\sum_{k \in \Mbest} (\mu_k -  \mu_M^*) (T - \E_{\mu}[T_k(T)])}_{\emph{\mytag{(b)}{eq:5:term2}}}
+      + \;\; \underbrace{\sum_{k=1}^{K} \mu_k \E_{\mu}[\cC_k(T)]}_{\emph{\mytag{(c)}{eq:5:term3}}}. \nonumber
+    \end{align}
+  % \end{small}
 \end{lemma}
@@
 In this decomposition, term \ref{eq:5:term1} counts the lost rewards due to \emph{sub-optimal arms} selections ($k \in \Mworst$), term \ref{eq:5:term2} counts the number of times the \emph{best arms} were \emph{not} selected ($k \in \Mbest$), and term \ref{eq:5:term3} counts the weighted number of collisions, on \emph{all arms}.
-It is valid for both centralized and decentralized algorithms.
+
+\begin{proof}
+  Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
+  \begin{align*}
+    R(T)
+    &= \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M Y_{A^j(t),t} \eta^j(t) \right]
+     = \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M \mu_{A^j(t)} \eta^j(t)\right]
+    \intertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence from $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$).
+So $\E_{\mu}[Y_{A^j(t),t} \eta^j(t)] = \sum_{k} \E_{\mu}[\mu_k \indic(A^j(t),t) \eta^j(t)] = \E_{\mu}[\mu_{A^j(t)} \eta^j(t)]$.
+And so}
+    R(T)
+    &= \E_{\mu}\left[ \sum_{t=1}^{T}\sum_{j \in \Mbest} \mu_j
+      - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
+      % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
+    &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
+      - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ T^j_k(T) \right]
+      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+    %
+    \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
+    %
+    &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ T^j_k(T) \right]
+      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+    %
+    \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
+    %
+    &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
+      + \overline{\Delta_{M^*}} \E_{\mu}\left[ T_{M^*}(T) \right] \\
+      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
+      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+    %
+    \intertext{But for $k = M^*$, $T_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ T_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ T_k(T) \right]$, so by recombining the terms, we obtain,}
+    %
+    &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
+      + \overline{\Delta_{M^*}} T M^* \\
+      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
+      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+  \end{align*}
+  %
+  The term $\overline{\Delta_k} - \overline{\Delta_{M^*}}$ simplifies to $\mu_{M^*} - \mu_k$, and so $\overline{\Delta_{M^*}} = \frac{1}{M} \sum_{k=1}^{M} \mu_k - \mu_{M^*}$ by definition of $\overline{\mu}^*$.
+And for $k=M^*$, $\mu_{M^*} - \mu_k = 0$, so the first sum can be written for $k = 1,\dots,M$ only, soIt
+  %
+  % \vspace*{5pt}
+  \begin{align*}
+    R(T)
+    &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+      + \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) T \\
+      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right]
+  \end{align*}
+  And so we obtain the decomposition with three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}.
+  \begin{align*}
+  R(T)
+    & = \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ T_k(T) \right]\right) \\
+      & \;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+      + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+  \end{align*}
+  Which is exactly the decomposition we wanted to prove.
+\end{proof}
+
+The regret decomposition in Lemma~\ref{lem:5:DecompositionRegret} is valid for both centralized and decentralized algorithms.
 For centralized algorithms, due to the absence of collisions, \ref{eq:5:term3} is obviously zero, and \ref{eq:5:term2} is non-negative, as $T_k(T) \leq T$.
 For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $T_k(T) > T$.
-However, a careful manipulation of this decomposition (see Appendix~\ref{proof:5:1stLowerBound}) shows that the regret is always lower bounded by term \ref{eq:5:term1}.
+However, a careful manipulation of this decomposition shows that the regret is always lower bounded by term \ref{eq:5:term1}.
 
@@
     For any strategy $\rho$ and $\boldsymbol{\mu}\in\cP_M$, it holds that
-    $ R_T(\boldsymbol{\mu}, M, \rho)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)]$.
+    \begin{align*}
+        R_T(\boldsymbol{\mu}, M, \rho)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)].
+    \end{align*}
 \end{lemma}
+
+\begin{proof}
+  Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
+  but it is not obvious for \ref{eq:5:term2} as there is no reason for $T_k(T)$ to be upper bounded by $T$.
+  %
+  Let $T_k^{!}(T) := \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
+  where the notation $\exists!$ stands for ``there exists a unique''.
+  Then $T_k(T) = \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(A^j(t) = k)$ can be decomposed as
+  \begin{equation*}
+    T_k(T) = \sum_{t=1}^{T} \indic(\exists! j, A^j(t) = k) + \sum_{t=1}^{T} \sum_{j=1}^{M} c_{k,t} \indic(A^j(t) = k)
+    = T_k^{!}(T) + C_k(T).
+  \end{equation*}
+  By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \rho)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
+  \begin{align*}
+    \ref{eq:5:term2} + \ref{eq:5:term3} &=
+    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+    + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)] \\
+    & \;\;\;\;\;\; + \sum_{k=1}^{M} \mu_k \E_{\mu}[C_k(T)]
+    - \sum_{k \in \Mbest} \mu_k \E_{\mu}[C_k(T)] \\
+    &=
+    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+    + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)]
+    + \sum_{k \in \Mworst} \mu_k \E_{\mu}[C_k(T)] \\
+    &=
+    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+    + \sum_{k=1}^{M} \min(\mu_M^*, \mu_k) \E_{\mu}[C_k(T)].
+  \end{align*}
+  And now both terms are non-negative, as $T_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
+  which proves that $R_T(\boldsymbol{\mu}, M, \rho) \geq \ref{eq:5:term1}$, as wanted.
+\end{proof}
 
@@
 
-To express our lower bound, we need to introduce $\kl(x,y)$ as the Kullback-Leibler divergence between the Bernoulli distribution of mean $x \neq 0,1$ and that of mean $y \neq 0,1$, so that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$.
+To express our lower bound, we need to introduce $\kl(x,y)$ as the Kullback-Leibler divergence between the Bernoulli distribution of mean $x$ and that of mean $y$, so that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$.
 %
@@
   A strategy $\rho$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$ and for all $\alpha \in (0,1)$,
-  \vspace*{-5pt}  % XXX
-  \begin{equation}
-    R_T(\boldsymbol{\mu},M,\rho) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \ \ \ \text{and } \ \
+  % \vspace*{-5pt}  % XXX
+  \begin{align}
+    & R_T(\boldsymbol{\mu},M,\rho) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
+    \text{and} & \;
     \forall j \in \{1,\dots,M\}, k \in \Mbest, \;\;
@@
     \label{eq:5:SUE}
-  \end{equation}
+  \end{align}
 \end{definition}
 
-Having a small regret on every problem instance,
-\ie, uniform efficiency,
+Having a small regret on every problem instance, \ie, being uniformly efficient,
 is a natural assumption for algorithms,
@@
 \begin{equation}\label{eq:5:intermediate}
-  \vspace*{-5pt}  % XXX
+  % \vspace*{-5pt}  % XXX
   \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[T^j_k(T)]\right) = o(T^{\alpha}).
@@
   Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\rho$ and $\boldsymbol{\mu}\in\cP_M$,
-  \vspace*{-10pt}  % XXX
+  % \vspace*{-10pt}  % XXX
   \begin{equation}\label{eq:5:LBDraws}
@@
   \noindent From Lemma~\ref{lem:5:1stLowerBound}, it follows that
-  \vspace*{-5pt}  % XXX
+  % \vspace*{-5pt}  % XXX
   \begin{equation}\label{eq:5:ourLowerBound}
@@
 
-A lower bound on the minimal number of collisions experienced by any strongly uniformly efficient decentralized algorithm would thus be a nice complement to our Theorem~\ref{thm:5:BetterLowerBound}, and it is left as future work.
+Since our article \cite{Besson2018ALT}, we kept as a future work the following question.
+A lower bound on the minimal number of collisions experienced by any strongly uniformly efficient decentralized algorithm would thus be a nice complement to our Theorem~\ref{thm:5:BetterLowerBound}.
 
@@
 A natural approach to obtain an upper bound on the regret of an algorithm is to upper bound separately each of the three terms defined in Lemma~\ref{lem:5:DecompositionRegret}.
-The following result shows that term $(b)$ can be related to the number of sub-optimal selections and the number of collisions that occurs on the $M$ best arms.
+The following result shows that term \ref{eq:5:term2} can be related to the number of sub-optimal selections and the number of collisions that occurs on the $M$ best arms.
 
@@
 \begin{lemma}\label{lem:5:1stUpperBound}
-  The term $(b)$ in Lemma~\ref{lem:5:DecompositionRegret} is upper bounded as
-  \hfill{}
-  (Proved in Appendix~\ref{proof:5:1stUpperBound})
+  The term \ref{eq:5:term2} in Lemma~\ref{lem:5:DecompositionRegret} is upper bounded as
+  % \hfill{}
+  % (Proved in Appendix~\ref{proof:5:1stUpperBound})
   %
   \begin{equation}\label{eq:5:1stUpperBound}
-    (b) \leq (\mu_1^* - \mu_M^*) \Bigl(    \sum_{k \in \Mworst} \E_{\mu}[T_k(T)]
+    \ref{eq:5:term2} \leq (\mu_1^* - \mu_M^*) \Bigl(    \sum_{k \in \Mworst} \E_{\mu}[T_k(T)]
     + \sum_{k \in \Mbest} \E_{\mu}[C_{k}(T)]
@@
   \end{equation}
-  \vspace*{-5pt}  % XXX
+  % \vspace*{-5pt}  % XXX
 \end{lemma}
 
-This result can also be used to recover Proposition 1 from \cite{Anandkumar11}, giving an upper bound on the regret that only depends on
+\begin{proof}
+  Recall that we want to upper bound
+  $ \ref{eq:5:term2} : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[T_k(T)]\right)$.
+  First, we observe that, for all $k\in \Mbest$,
+  \begin{eqnarray*}
+    T - \E_{\mu}[T_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
+    & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
+  \end{eqnarray*}
+  where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
+  \begin{eqnarray*}
+  \ref{eq:5:term2} & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[T_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
+  & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
+  \end{eqnarray*}
+  The quantity $\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)$ counts the number of optimal arms that have not been selected at time $t$.
+For each mis-selection of an optimal arm, there either exists a sub-optimal arm that has been selected, or an arm in $\Mbest$ on which a collision occurs.
+Hence
+  \[\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t) = \sum_{k \in \Mbest}\indic(C_k(t)) + \sum_{k \in \Mworst} \indic(\exists j : A^j(t) = k),\]
+  which yields
+  \[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[T_k(T)\right]\]
+  and Lemma~\ref{lem:5:1stUpperBound} follows.
+\end{proof}
+
+This result can also be used to recover Proposition~1 from \cite{Anandkumar11}, giving an upper bound on the regret that only depends on
 the \emph{expected number of sub-optimal selections} -- $\E_{\mu}[T_k(T)]$ for $k \in \Mworst$ --
@@
 
-While the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices  (even though some alternative without pre-agreement are proposed),
-the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}.
+On the first hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
+Even though some alternative without pre-agreement are proposed, they are quite complicated and we prefer to focus on other approaches.
+%
+On the other hand, the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}.
 %\footnote{\rhoRand{} only works if $M$ is known, and \cite{Anandkumar11} extended it to the case of an unknown $U$, with \rhoRandEst.
@@
 \end{equation}
-If a collision occurs, a new rank is drawn uniformly at random: $R^j(t+1) \sim \cU(\{1,\dots,M\})$.
+If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU(\{1,\dots,M\})$.
 
@@
 
-We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices:
+We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
 \begin{equation}
@@
 
-Our proposal \MCTopM{} is stated below as Algorithm~\ref{algo:5:MCTopM},
-while a simpler variant, called \RandTopM{},
-is stated as Algorithm~\ref{algo:5:RandTopM} in Appendix~\ref{app:5:RandTopM}.
-We focus on \MCTopM{} as it is easier to analyze and performs better.
+% \vspace*{-5pt}  % XXX remove if problem
+% \begin{small}  % XXX remove if problem
+  \begin{figure}[h!]
+      \begin{framed}  % XXX remove if problem
+      % \begin{small}  % XXX remove if problem
+      \centering
+      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
+      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
+      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
+      \begin{algorithm}[H]
+          % XXX Options
+          % \LinesNumbered  % XXX Option to number the line
+          % \RestyleAlgo{boxed}
+          % XXX Input, data and output
+          % \KwIn{$K$ and policy $P^j$ for arms set $\{1,\dots,K\}$\;}
+          % \KwData{Data}
+          % \KwResult{Result}
+          % XXX Algorithm
+              Let $A^j(1) \sim \cU(\{1,\dots,K\})$ and $C^j(1)=\mathrm{False}$ \\
+              \For{$t = 0, \dots, T - 1$}{
+                  %
+                  \eIf{$A^j(t) \notin \TopM(t)$}{
+                    \eIf(\tcp*[f]{collision}){$C^j(t)$}{
+                      $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
+                      \tcp*[f]{randomly switch}
+                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}){
+                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
+                      }
+                    }{
+                      $A^j(t+1) = A^j(t)$
+                      \tcp*[f]{stays on the same arm}
+                    }
+                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
+                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
+              }
+              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for a fixed underlying index policy $g^j$).}
+          \label{algo:5:RandTopM}
+      \end{algorithm}
+      % \end{small}  % XXX remove if problem
+      \end{framed}  % XXX remove if problem
+  \end{figure}
+% \end{small}  % XXX remove if problem
+% \vspace*{-5pt}  % XXX remove if problem
+
+\paragraph{The \RandTopM{} algorithm.}
+%
+First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
+It is essentially a refinement over the \RhoRand{} algorithm, and a simpler version of \MCTopM.
+The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
+%
+In \RandTopM{}, player $j$ is always considered ``not fixed'',
+and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.
+
+\paragraph{The \MCTopM{} algorithm.}
+%
+Our second proposal \MCTopM{} is stated below as Algorithm~\ref{algo:5:MCTopM},
+it is a slightly more complex extension of the \RandTopM{} algorithm.
+From there on, we focus on \MCTopM{} as it is easier to analyze and performs better.
 %
@@
 selects at time $t+1$ an arm from $\TopM(t)$.
-When a collision occurs \RandTopM{} randomly switches arm within $\TopM$, while \MCTopM{} uses a more sophisticated mechanism, that is reminiscent of ``Musical Chair'' (MC) and inspired by the work of \cite{Rosenski16}: players tend to fix themselves on arms (``chairs'') and ignore future collision when this happens.
+When a collision occurs for a player implementing the \RandTopM{} algorithm, that player randomly switches arm within $\TopM$, while \MCTopM{} uses a more sophisticated mechanism, that is reminiscent of ``Musical Chair'' (MC) and inspired by the work of \cite{Rosenski16}: players tend to fix themselves on arms (``chairs'') and ignore future collision when this happens.
 
@@
   \begin{figure}[h!]
-      % \begin{framed}  % XXX remove if problem
+      \begin{framed}  % XXX remove if problem
       % \begin{small}  % XXX remove if problem
@@
       % \end{small}  % XXX remove if problem
-      % \end{framed}  % XXX remove if problem
+      \end{framed}  % XXX remove if problem
   \end{figure}
 % \end{small}  % XXX remove if problem
-\vspace*{-5pt}  % XXX remove if problem
+% \vspace*{-5pt}  % XXX remove if problem
 
@@
 %
-The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM{} (see Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in Appendix~\ref{proof:5:collisionsMCTopM}).
+The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM, and they are illustrated in Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below.
+
+\begin{figure}[h!]
+  \resizebox{1.00\textwidth}{!}{
+  \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
+      %
+      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
+      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
+      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
+      %
+      \draw [black,->] (start) -> (notfixed.20);
+      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
+      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
+      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
+      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
+      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
+      %
+  \end{tikzpicture}
+  }
+  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
+  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
+  \label{fig:5:StateMachineAlgorithm_MCTopM}
+\end{figure}
 
@@
 %
-As such, practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or not considered.
+Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or not considered.
 
@@
 % -----------------------------------------------------------------
-\section{Numerical simulations}
-\label{sec:5:experiments}
-% -----------------------------------------------------------------
-
-% Select one problems, max two cases ($M=K$ and $M < K$), and illustrate what we want to discuss in plots with regret in normal scale, one with semi-$\log x$ scale, and at least one histogram showing the bad luck for \Selfish...
-
-% Selfish is awesome... unless $K=M=2$! Not only, we found other case of failure.
-
-% Chose one problem $\boldsymbol{\mu}$, and vary $M \leq K$ for let say $K=3$ and $K=9$. Not more figures.
-% Put most of them in appendix.
-
-We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \klUCB{} indices.
-Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.
-
-\TODOL{Rewrite appendix to include all simulations here!}
-
-\TODOL{Include here some plots showing \RhoRand{} with \UCB{} and \klUCB{} vs \MCTopM{} with \UCB{} and \klUCB{}.}
-
-In experiments that are not reported here, we could observe that using \klUCB{} rather than \UCB{} indices always yield better practical performance.
-%, as the theoretical guarantees for single player suggested,
-As the purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
-% And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
-we only report results for \klUCB.
-In a first set of experiments, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
-We also include a centralized multiple-play \klUCB{} algorithm,
-%as defined by \cite{Anantharam87a},
-essentially to check that the \emph{``price of decentralized learning''} is not too large.
-
-%
-
-We present results for two bandit instance: one with $K=3$ arms and means %\footnote{But of course the means are unknown to the algorithms, and their order is not important.}
-$\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}.
-For the second instance $K=9$ and $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
-three cases are presented: $M=6$ in Figure~\ref{fig:5:MP__K9_M6_T10000_N1000__4_algos},
-and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
-% Note the different algorithms know the number of players $M$.
-Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
-%, from $t=1$ to the horizon $T=10000$ (also unknown by the algorithms),
-We also include histograms showing the \emph{distribution} of regret at $t=T$,
-as this allows to check if the regret is indeed small for \emph{each} run of the simulation.
-%
-For the plots showing the regret, our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} is displayed.
-% and it should be surprising to see it being larger than some regret values: it is only asymptotic.
-
-Experiments with a different problem for each repetition (uniformly sampled $\boldsymbol{\mu} \sim \cU([0,1]^K)$),
-are also considered, in Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \ref{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}.
-This helps to check that no matter the \emph{complexity} of the considered problem (one measure of complexity being the constant in our lower bound),
-\MCTopM{} performs similarly or better than all the other algorithms,
-and \Selfish{} outperforms \rhoRand{} in most cases.
-%Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} is a good example
-%of outstanding performances of \MCTopM{} and \Selfish{} in comparison to \rhoRand{}.
-Empirically, our proposals were found to almost always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
-we verified that \MCTopM{} outperforms the state-of-the-art algorithms in many different problems, and is more and more efficient as $M$ and $K$ grows.
-
-%
-% Regular plots of centralized regrets
-%
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{0.49\textwidth}
-  %   \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
-  % \end{subfigure}
-  % % ~
-  % \begin{subfigure}[!h]{0.49\textwidth}
-    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
-  % \end{subfigure}
-  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-\rhoRand{} (top \textcolor{blue}{blue} curve) is outperformed by the other algorithms (and the gain increases with $M$).
-\MCTopM{} (bottom \textcolor{gold}{yellow}) outperforms all the other algorithms is most cases.}
-  % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
-  \label{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
-  % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-% Dynamic settings, when $M$ can change in time,
-% were considered in the analysis of both
-In the presence of sensing (observation model \modeldeux), we also compared our algorithms to  %multi-player multi-armed bandits algorithms
-with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}. Yet these two algorithms were found hard to use efficiently in practice and we show in
-% that were also proposed
-%
-Figure~\ref{fig:5:MP__K9_M3_T123456_N100__8_algos} that they perform poorly in comparison to \rhoRand, \RandTopM{} and \MCTopM.
-%
-\MEGA{} needs a careful tuning of \emph{five} parameters ($c$, $d$, $p_0$, $\alpha$ and $\beta$) to attain reasonable performances.
-No good guideline for tuning them is provided and using cross validation, as suggested,
-can be considered out of the scope of online sequential learning.
-%In practice, on a fixed instance, the authors do not indicate how to select the parameters, even with a perfect knowledge of the parameters ($\boldsymbol{\mu}$ and $T$).
-%
-\MusicalChair{} consists of a random exploration phase of length $T_0$ after which the players quickly converge to orthogonal strategies targeting the $M$ best arms.
-With probability $1-\delta$, its regret is proved to be ``constant'' (of order $\log(1/\delta)$).
-The theoretical minimal value for $T_0$ depends on $\delta$, on the horizon $T$ and on a lower bound $\epsilon$ on the gap $\Delta = \mu^*_M - \mu^*_{M+1}$, and the practical tuning is hard too.
-%% which are both unavailable in our setting.
-
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{0.85\textwidth}
-    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{0.85\textwidth}
-    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_HistogramsRegret____env1-1_8200873569864822246.pdf}
-  % \end{subfigure}
-  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-\RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand.
-The regret of \MCTopM{} is logarithmic, empirically with the same slope as the lower bound.
-The $x$ axis on the regret histograms have different scale for each algorithm.}
-  \label{fig:5:MP__K9_M6_T10000_N1000__4_algos}
-\end{figure}
-
-% -----------------------------------------------------------------
-\section{Theoretical elements}
+\section{Theoretical elements and order optimal regret upper bound for \MCTopM}
 \label{sec:5:upperbounds}
@@
 an asymptotically optimal analysis of the expected number of sub-optimal draws
-for \RandTopM, \MCTopM{} and \rhoRand{} combined with \klUCB{} indices,
-and Section~\ref{sub:5:UpperBoundCollisions} proves that the number of  collisions, hence the regret of \MCTopM{} are also logarithmic.
+for our two proposals \RandTopM{} and \MCTopM{} as well as for \rhoRand{}, when they are combined with \klUCB{} indices,
+and Section~\ref{sub:5:UpperBoundCollisions} proves that the number of collisions are logarithmic, and hence the regret of \MCTopM{} is also logarithmic.
 %
-Section~\ref{sub:5:SelfishFails} shortly discusses a disappointing result regarding \Selfish{}, with more insights provided in Appendix~\ref{app:5:SelfishFails}.
+Finally, Section~\ref{sub:5:SelfishFails} shortly discusses a disappointing result regarding \Selfish{}, with more insights provided in Appendix~\ref{app:5:SelfishFails}.
 
@@
 
-% For \rhoRand, \cite{Anandkumar11} essentially proved on one hand that conditionnaly to a certain ``good'' event
+% For \rhoRand, \cite{Anandkumar11} essentially proved On the one hand that conditionnaly to a certain ``good'' event
 % (all players have a correct ordering of all the $M$ best arms)
@@
 and this additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
-Even though minimizing switching was not a goal\footnote{Introducing \emph{switching costs}, like it was done in previous works, \eg, \cite{Koren17}, could be an interesting future work.},
+Even though minimizing the number of arms switching was not a goal,
 this guarantee is interesting for Cognitive Radio applications,
 where switching arms means re-configuring a radio hardware, an operation that costs energy.
-%and thus an algorithm guaranting a small number of switches is interesting.
+An algorithm guaranteeing a small number of switches is interesting.
+%
+We note that introducing \emph{switching costs}, like it was done in previous works like \cite{Koren17}, could be an interesting future work.
 
@@
 % -----------------------------------------------------------------
+\section{Numerical simulations}
+\label{sec:5:experiments}
+% -----------------------------------------------------------------
+
+% Select one problems, max two cases ($M=K$ and $M < K$), and illustrate what we want to discuss in plots with regret in normal scale, one with semi-$\log x$ scale, and at least one histogram showing the bad luck for \Selfish...
+
+% Selfish is awesome... unless $K=M=2$! Not only, we found other case of failure.
+
+% Chose one problem $\boldsymbol{\mu}$, and vary $M \leq K$ for let say $K=3$ and $K=9$. Not more figures.
+% Put most of them in appendix.
+
+We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \klUCB{} indices.
+Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.
+
+\TODOL{Rewrite appendix to include all simulations here!}
+
+\TODOL{Include here some plots showing \RhoRand{} with \UCB{} and \klUCB{} vs \MCTopM{} with \UCB{} and \klUCB{}.}
+
+In experiments that are not reported here, we could observe that using \klUCB{} rather than \UCB{} indices always yield better practical performance.
+%, as the theoretical guarantees for single player suggested,
+As the purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
+% And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
+we only report results for \klUCB.
+In a first set of experiments, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
+We also include a centralized multiple-play \klUCB{} algorithm,
+%as defined by \cite{Anantharam87a},
+essentially to check that the \emph{``price of decentralized learning''} is not too large.
+
+%
+
+We present results for two bandit instance: one with $K=3$ arms and means %\footnote{But of course the means are unknown to the algorithms, and their order is not important.}
+$\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}.
+For the second instance $K=9$ and $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
+three cases are presented: $M=6$ in Figure~\ref{fig:5:MP__K9_M6_T10000_N1000__4_algos},
+and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
+% Note the different algorithms know the number of players $M$.
+Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
+%, from $t=1$ to the horizon $T=10000$ (also unknown by the algorithms),
+We also include histograms showing the \emph{distribution} of regret at $t=T$,
+as this allows to check if the regret is indeed small for \emph{each} run of the simulation.
+%
+For the plots showing the regret, our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} is displayed.
+% and it should be surprising to see it being larger than some regret values: it is only asymptotic.
+
+Experiments with a different problem for each repetition,
+that is uniformly sampled $\boldsymbol{\mu} \sim \cU([0,1]^K)$,
+are also considered, in Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \ref{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}.
+This helps to check that no matter the \emph{complexity} of the considered problem (one measure of complexity being the constant in our lower bound),
+\MCTopM{} performs similarly or better than all the other algorithms,
+and \Selfish{} outperforms \rhoRand{} in most cases.
+%Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} is a good example
+%of outstanding performances of \MCTopM{} and \Selfish{} in comparison to \rhoRand{}.
+Empirically, our proposals were found to almost always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
+we verified that \MCTopM{} outperforms the state-of-the-art algorithms in many different problems, and is more and more efficient as $M$ and $K$ grows.
+
+%
+% Regular plots of centralized regrets
+%
+\begin{figure}[!h]
+  % \centering
+  % \begin{subfigure}[!h]{0.49\textwidth}
+  %   \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
+  % \end{subfigure}
+  % % ~
+  % \begin{subfigure}[!h]{0.49\textwidth}
+    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+\rhoRand{} (top \textcolor{blue}{blue} curve) is outperformed by the other algorithms (and the gain increases with $M$).
+\MCTopM{} (bottom \textcolor{gold}{yellow}) outperforms all the other algorithms is most cases.}
+  % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
+  \label{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
+  % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
+% Dynamic settings, when $M$ can change in time,
+% were considered in the analysis of both
+In the presence of sensing (observation model \modeldeux), we also compared our algorithms to  %multi-player multi-armed bandits algorithms
+with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}. Yet these two algorithms were found hard to use efficiently in practice and we show in
+% that were also proposed
+%
+Figure~\ref{fig:5:MP__K9_M3_T123456_N100__8_algos} that they perform poorly in comparison to \rhoRand, \RandTopM{} and \MCTopM.
+%
+\MEGA{} needs a careful tuning of \emph{five} parameters ($c$, $d$, $p_0$, $\alpha$ and $\beta$) to attain reasonable performances.
+No good guideline for tuning them is provided and using cross validation, as suggested,
+can be considered out of the scope of online sequential learning.
+%In practice, on a fixed instance, the authors do not indicate how to select the parameters, even with a perfect knowledge of the parameters ($\boldsymbol{\mu}$ and $T$).
+%
+\MusicalChair{} consists of a random exploration phase of length $T_0$ after which the players quickly converge to orthogonal strategies targeting the $M$ best arms.
+With probability $1-\delta$, its regret is proved to be ``constant'' (of order $\log(1/\delta)$).
+The theoretical minimal value for $T_0$ depends on $\delta$, on the horizon $T$ and on a lower bound $\epsilon$ on the gap $\Delta = \mu^*_M - \mu^*_{M+1}$, and the practical tuning is hard too.
+%% which are both unavailable in our setting.
+
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{0.85\textwidth}
+    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{0.85\textwidth}
+    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_HistogramsRegret____env1-1_8200873569864822246.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+\RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand.
+The regret of \MCTopM{} is logarithmic, empirically with the same slope as the lower bound.
+The $x$ axis on the regret histograms have different scale for each algorithm.}
+  \label{fig:5:MP__K9_M6_T10000_N1000__4_algos}
+\end{figure}
+
+% -----------------------------------------------------------------
 \section{Conclusion and future work}

### 2019-04-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Using \begin{framed}...\end{framed} for all algorithms.
Moved a lot of appendix in main text for Chapter 5. See #15

@@
-% -----------------------------------------------------------------
-\section{Regret Decompositions}
-\label{app:5:regretdec}
-
-% -----------------------------------------------------------------
-\subsection{Proof of Lemma~\ref{lem:5:DecompositionRegret}}
-\label{proof:5:DecompositionRegret}
-
-    Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
-    \begin{align*}
-      R(T)
-      &= \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M Y_{A^j(t),t} \eta^j(t) \right]
-       = \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M \mu_{A^j(t)} \eta^j(t)\right]
-      \shortintertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence from $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$).
+% % -----------------------------------------------------------------
+% \section{Regret Decompositions}
+% \label{app:5:regretdec}
+
+% % -----------------------------------------------------------------
+% \subsection{Proof of Lemma~\ref{lem:5:DecompositionRegret}}
+% \label{proof:5:DecompositionRegret}
+
+%     Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
+%     \begin{align*}
+%       R(T)
+%       &= \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M Y_{A^j(t),t} \eta^j(t) \right]
+%        = \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M \mu_{A^j(t)} \eta^j(t)\right]
+%       \intertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence from $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$).
 So $\E_{\mu}[Y_{A^j(t),t} \eta^j(t)] = \sum_{k} \E_{\mu}[\mu_k \indic(A^j(t),t) \eta^j(t)] = \E_{\mu}[\mu_{A^j(t)} \eta^j(t)]$.
 And so}
-      R(T)
-      &= \E_{\mu}\left[ \sum_{t=1}^{T}\sum_{j \in \Mbest} \mu_j
-        - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
-        % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
-      &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
-        - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ T^j_k(T) \right]
-        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
-      %
-      \shortintertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
-      %
-      &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ T^j_k(T) \right]
-        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
-      %
-      \shortintertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
-      %
-      &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
-        + \overline{\Delta_{M^*}} \E_{\mu}\left[ T_{M^*}(T) \right] \\
-        &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
-        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
-      %
-      \shortintertext{But for $k = M^*$, $T_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ T_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ T_k(T) \right]$, so by recombining the terms, we obtain,}
-      %
-      &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
-        + \overline{\Delta_{M^*}} T M^* \\
-        &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
-        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
-    \end{align*}
-    %
-    The term $\overline{\Delta_k} - \overline{\Delta_{M^*}}$ simplifies to $\mu_{M^*} - \mu_k$, and so $\overline{\Delta_{M^*}} = \frac{1}{M} \sum_{k=1}^{M} \mu_k - \mu_{M^*}$ by definition of $\overline{\mu}^*$.
+%       R(T)
+%       &= \E_{\mu}\left[ \sum_{t=1}^{T}\sum_{j \in \Mbest} \mu_j
+%         - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
+%         % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
+%       &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
+%         - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ T^j_k(T) \right]
+%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+%       %
+%       \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
+%       %
+%       &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ T^j_k(T) \right]
+%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+%       %
+%       \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
+%       %
+%       &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
+%         + \overline{\Delta_{M^*}} \E_{\mu}\left[ T_{M^*}(T) \right] \\
+%         &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
+%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+%       %
+%       \intertext{But for $k = M^*$, $T_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ T_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ T_k(T) \right]$, so by recombining the terms, we obtain,}
+%       %
+%       &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
+%         + \overline{\Delta_{M^*}} T M^* \\
+%         &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
+%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+%     \end{align*}
+%     %
+%     The term $\overline{\Delta_k} - \overline{\Delta_{M^*}}$ simplifies to $\mu_{M^*} - \mu_k$, and so $\overline{\Delta_{M^*}} = \frac{1}{M} \sum_{k=1}^{M} \mu_k - \mu_{M^*}$ by definition of $\overline{\mu}^*$.
 And for $k=M^*$, $\mu_{M^*} - \mu_k = 0$, so the first sum can be written for $k = 1,\dots,M$ only, so
-    %
-    % \vspace*{5pt}
-    \begin{align*}
-      R(T)
-      &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
-        + \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) T \\
-        &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
-        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right]
-    \end{align*}
-    And so we obtain the decomposition with three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}.
-    \begin{align*}
-    R(T)
-      &= \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ T_k(T) \right]\right)
-        + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
-        + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
-    \end{align*}
-    % Which is exactly the decomposition we wanted to prove.
-
-% -----------------------------------------------------------------
-\subsection{Proof of Lemma~\ref{lem:5:1stLowerBound}}
-\label{proof:5:1stLowerBound}
-
-  Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
-  but it is not obvious for \ref{eq:5:term2} as there is no reason for $T_k(T)$ to be upper bounded by $T$.
-  %
-  Let $T_k^{!}(T) := \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
-  where the notation $\exists!$ stands for ``there exists a unique''.
-  Then $T_k(T) = \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(A^j(t) = k)$ can be decomposed as
-  \begin{equation*}
-    T_k(T) = \sum_{t=1}^{T} \indic(\exists! j, A^j(t) = k) + \sum_{t=1}^{T} \sum_{j=1}^{M} c_{k,t} \indic(A^j(t) = k)
-    = T_k^{!}(T) + C_k(T).
-  \end{equation*}
-  By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \rho)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
-  \begin{align*}
-    \ref{eq:5:term2} + \ref{eq:5:term3} &=
-    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
-    + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)] \\
-    & \;\;\;\;\;\; + \sum_{k=1}^{M} \mu_k \E_{\mu}[C_k(T)]
-    - \sum_{k \in \Mbest} \mu_k \E_{\mu}[C_k(T)] \\
-    &=
-    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
-    + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)]
-    + \sum_{k \in \Mworst} \mu_k \E_{\mu}[C_k(T)] \\
-    &=
-    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
-    + \sum_{k=1}^{M} \min(\mu_M^*, \mu_k) \E_{\mu}[C_k(T)].
-  \end{align*}
-  And now both terms are non-negative, as $T_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
-  which proves that $R_T(\boldsymbol{\mu}, M, \rho) \geq \ref{eq:5:term1}$, as wanted.
-
-% -----------------------------------------------------------------
-\subsection{Proof of Lemma~\ref{lem:5:1stUpperBound}}
-\label{proof:5:1stUpperBound}
-
-Recall that we want to upper bound
-$ (b) : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[T_k(T)]\right)$.
-First, we observe that, for all $k\in \Mbest$,
-\begin{eqnarray*}
-  T - \E_{\mu}[T_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
-  & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
-\end{eqnarray*}
-where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
-\begin{eqnarray*}
- (b) & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[T_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
- & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
-\end{eqnarray*}
-The quantity $\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)$ counts the number of optimal arms that have not been selected at time $t$.
+%     %
+%     % \vspace*{5pt}
+%     \begin{align*}
+%       R(T)
+%       &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+%         + \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) T \\
+%         &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right]
+%     \end{align*}
+%     And so we obtain the decomposition with three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}.
+%     \begin{align*}
+%     R(T)
+%       &= \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ T_k(T) \right]\right)
+%         + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+%         + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
+%     \end{align*}
+%     % Which is exactly the decomposition we wanted to prove.
+
+% % -----------------------------------------------------------------
+% \subsection{Proof of Lemma~\ref{lem:5:1stLowerBound}}
+% \label{proof:5:1stLowerBound}
+
+%   Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
+%   but it is not obvious for \ref{eq:5:term2} as there is no reason for $T_k(T)$ to be upper bounded by $T$.
+%   %
+%   Let $T_k^{!}(T) := \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
+%   where the notation $\exists!$ stands for ``there exists a unique''.
+%   Then $T_k(T) = \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(A^j(t) = k)$ can be decomposed as
+%   \begin{equation*}
+%     T_k(T) = \sum_{t=1}^{T} \indic(\exists! j, A^j(t) = k) + \sum_{t=1}^{T} \sum_{j=1}^{M} c_{k,t} \indic(A^j(t) = k)
+%     = T_k^{!}(T) + C_k(T).
+%   \end{equation*}
+%   By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \rho)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
+%   \begin{align*}
+%     \ref{eq:5:term2} + \ref{eq:5:term3} &=
+%     \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+%     + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)] \\
+%     & \;\;\;\;\;\; + \sum_{k=1}^{M} \mu_k \E_{\mu}[C_k(T)]
+%     - \sum_{k \in \Mbest} \mu_k \E_{\mu}[C_k(T)] \\
+%     &=
+%     \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+%     + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)]
+%     + \sum_{k \in \Mworst} \mu_k \E_{\mu}[C_k(T)] \\
+%     &=
+%     \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+%     + \sum_{k=1}^{M} \min(\mu_M^*, \mu_k) \E_{\mu}[C_k(T)].
+%   \end{align*}
+%   And now both terms are non-negative, as $T_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
+%   which proves that $R_T(\boldsymbol{\mu}, M, \rho) \geq \ref{eq:5:term1}$, as wanted.
+
+% % -----------------------------------------------------------------
+% \subsection{Proof of Lemma~\ref{lem:5:1stUpperBound}}
+% \label{proof:5:1stUpperBound}
+
+% Recall that we want to upper bound
+% $ (b) : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[T_k(T)]\right)$.
+% First, we observe that, for all $k\in \Mbest$,
+% \begin{eqnarray*}
+%   T - \E_{\mu}[T_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
+%   & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
+% \end{eqnarray*}
+% where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
+% \begin{eqnarray*}
+%  (b) & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[T_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
+%  & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
+% \end{eqnarray*}
+% The quantity $\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)$ counts the number of optimal arms that have not been selected at time $t$.
 For each mis-selection of an optimal arm, there either exists a sub-optimal arm that has been selected, or an arm in $\Mbest$ on which a collision occurs.
 Hence
-\[\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t) = \sum_{k \in \Mbest}\indic(C_k(t)) + \sum_{k \in \Mworst} \indic(\exists j : A^j(t) = k),\]
-which yields
-\[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[T_k(T)\right]\]
-and Lemma~\ref{lem:5:1stUpperBound} follows.
+% \[\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t) = \sum_{k \in \Mbest}\indic(C_k(t)) + \sum_{k \in \Mworst} \indic(\exists j : A^j(t) = k),\]
+% which yields
+% \[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[T_k(T)\right]\]
+% and Lemma~\ref{lem:5:1stUpperBound} follows.
 
@@
 % -----------------------------------------------------------------
-\section{The \RandTopM{} algorithm}
-\label{app:5:RandTopM}
-
-\TODOL{I should include this in the main text, it makes more sense to present RandTopM and then MCTopM !}
-
-We now state precisely the \RandTopM{} algorithm below in Algorithm~\ref{algo:5:RandTopM} (page \pageref{algo:5:RandTopM}).
-It is essentially the same algorithm as \MCTopM,
-but in a simpler version as the ``Chair'' aspect is removed, that is, there is no notion of state $s^j(t)$ (\cf{} Algorithm~\ref{algo:5:MCTopM}).
-%
-Player $j$ is always considered ``not fixed'',
-and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$
-in the case of \RandTopM.
-
-\vspace*{-5pt}  % XXX remove if problem
-% \begin{small}  % XXX remove if problem
-  \begin{figure}[h!]
-      % \begin{framed}  % XXX remove if problem
-      % \begin{small}  % XXX remove if problem
-      \centering
-      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
-      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
-      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
-      \begin{algorithm}[H]
-          % XXX Options
-          % \LinesNumbered  % XXX Option to number the line
-          % \RestyleAlgo{boxed}
-          % XXX Input, data and output
-          % \KwIn{$K$ and policy $P^j$ for arms set $\{1,\dots,K\}$\;}
-          % \KwData{Data}
-          % \KwResult{Result}
-          % XXX Algorithm
-              Let $A^j(1) \sim \cU(\{1,\dots,K\})$ and $C^j(1)=\mathrm{False}$ \\
-              \For{$t = 0, \dots, T - 1$}{
-                  %
-                  \eIf{$A^j(t) \notin \TopM(t)$}{
-                    \eIf(\tcp*[f]{collision}){$C^j(t)$}{
-                      $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
-                      \tcp*[f]{randomly switch}
-                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}){
-                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
-                      }
-                    }{
-                      $A^j(t+1) = A^j(t)$
-                      \tcp*[f]{stays on the same arm}
-                    }
-                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
-                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
-              }
-              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for a fixed underlying index policy $g^j$).}
-          \label{algo:5:RandTopM}
-      \end{algorithm}
-      % \end{small}  % XXX remove if problem
-      % \end{framed}  % XXX remove if problem
-  \end{figure}
-% \end{small}  % XXX remove if problem
-\vspace*{-5pt}  % XXX remove if problem
-
-% -----------------------------------------------------------------
-% -----------------------------------------------------------------
 
@@
 %
-\shortintertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
+\intertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
 \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\

### 2019-04-04 2-Chapters/5-Chapter/chapter5.tex
Moved a lot of appendix in main text for Chapter 5. See #15

@@
 
-Finally, another interesting extension is to consider different channels utility for each devices.
-It is a very interesting extension of our model, for instance the devices can be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
-This extension is not studied per say but we quickly review the existing literature \cite{Bistritz18,KaufmannAbbas19} and discuss the possible real-world usage of the existing algorithms.
+Finally, another interesting extension of the studied models is to consider different channels utility for each devices.
+It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
+This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.
 
-\TODOL{Get bibtex of the article by Sumit J. Darak}
+\TODOL{Get bibtex of the articles by Sumit J. Darak, and talk correctly about them!}
+
+\vfill{}
+
+\paragraph{Publications}
+
+This chapter is mainly based on the following publication: \cite{Besson2018ALT}.
 
@@
 % ----------------------------------------------------------------------------
-\section{Conclusion of Chapter~\ref{chapter:5}}
+\section{Conclusion}
 \label{sec:5:conclusion}
@@
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter~\ref{chapter:5}}
+\section{Appendix}
 \label{sec:5:appendix}

### 2019-04-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Galère ce chapitre 5, je sais pas trop quoi faire.

@@
 Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.
-Other models for rewards loss have been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it, but we focus on full reward occlusion in this chapter.
+
+Note that other models for rewards loss have also been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it.
+To stay consistent with the model presented in the previous Chapter~\ref{chapter:4},
+we preferred to focus on full reward occlusion in this chapter.
 
@@
 In this section, we provide a useful decomposition of the regret (Lemma~\ref{lem:5:DecompositionRegret}) that permits to establish a new problem-dependent lower bound on the regret (Theorem~\ref{thm:5:BetterLowerBound}), and also provides key insights on the derivation of regret upper bounds (Lemma~\ref{lem:5:1stUpperBound}).
+
+% FIXME
+\begin{framed}
+  \paragraph{\textcolor{red}{Warning}.}
+  The regret lower bound given in \cite{Besson2018ALT} is not applicable to any algorithm,
+  as it was discovered by Etienne Boursier and Vianney Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
+  The proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step,
+  but the results stated in the section are now correct,
+  thanks to the modifications proposed by Emilie Kaufmann and Abbas Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
+  The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which contains both \RhoRand{} and our two proposals \RandTopM{} and \MCTopM{} but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
+\end{framed}
+% FIXME
 
@@
 
-\subsection{Two new strategies based on indices and orthogonalization: \RandTopM{} and \MCTopM}
+% \subsection{Two new strategies based on indices and orthogonalization: \RandTopM{} and \MCTopM}
+\subsection{Two new strategies based on indices and orthogonalization}
 \label{sub:5:RandTopM_and_MCTopM}
@@
 %
-However, we found out empirically that \Selfish{} is doing surprisingly well, as already noted by \cite{Bonnefoi17}, who did some experiments in the context of IoT applications.
-%
+However, we found out empirically that \Selfish{} is doing surprisingly well,
+like what we found in Chapter~\ref{chapter:4} in harder settings.
+% as already noted by \cite{Bonnefoi17}, who did some experiments in the context of IoT applications.
+
 We show in Section~\ref{sec:5:upperbounds} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
 which precludes the possibility of a logarithmic regret for any problem.
-However, in most cases it empirically performs similarly to all the algorithms described before,
+In most cases, it empirically performs similarly to all the algorithms described before,
 and usually outperforms \rhoRand,
@@
 %
-Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or not considered.
-
-% -----------------------------------------------------------------
-
-% -----------------------------------------------------------------
-\section{Theoretical elements and order optimal regret upper bound for \MCTopM}
+Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or cannot be considered.
+
+% -----------------------------------------------------------------
+
+% -----------------------------------------------------------------
+\section{Theoretical elements, and order optimal regret for \MCTopM}
 \label{sec:5:upperbounds}
@@
 
+We now focus on obtaining positive results for the algorithms we proposed in Section~\ref{sec:5:algorithms} above.
 Section~\ref{sub:5:UpperBoundSelections} gives
@@
 Our improved analysis also applies to \rhoRand{}.
-Explicit expressions for $C_{\boldsymbol{\mu}}$, $D_{\boldsymbol{\mu}}$ can be found in the proof given in Appendix~\ref{proof:5:SubOptimalSelections}.
+Explicit expressions for $C_{\boldsymbol{\mu}}$, $D_{\boldsymbol{\mu}}$ can be found in the proof given below.
 
@@
 
+\begin{proof}
+  %
+  %     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
+  %     As only user $j$ is considered here, the superscript $j$ is dropped.
+  %
+  Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
+  The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
+  \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
+  Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
+  If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
+  In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.
+
+  Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
+  \begin{align*}
+  \E_{\mu}[T_k^j(T)]
+  &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
+  = \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
+  %&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
+  &= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
+  %
+  \end{align*}
+  Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
+
+  \begin{align*}
+    \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
+    &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
+    &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
+    %
+    & \leq \sum_{t=1}^T \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+    + \sum_{m=1}^M \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right)
+    \hspace*{100pt} (\heartsuit)
+  \end{align*}
+  where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.
+
+  Now each of the two terms in the right hand side of $(\heartsuit)$ can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
+  The rightmost term in $(\heartsuit)$ can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact~1 in Appendix~A of \cite{KLUCBJournal}.
+
+  \begin{lemma}\label{lem:5:Fact1KLUCB}
+      For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
+      \begin{equation}
+        \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
+      \end{equation}
+  \end{lemma}
+
+  The leftmost term in $(\heartsuit)$ can be controlled using Lemma~\ref{lem:5:Fact2KLUCB}, that is a direct consequence of the proof of Fact~2 in Appendix~A of \cite{KLUCBJournal}.
+  Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).
+
+  \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
+  \begin{align*}
+    &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
+    + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
+  \end{lemma}
+
+  Putting things together, one obtains the non-asymptotic upper bound
+  \begin{align}\label{eq:5:UBprecise}
+    \E_{\mu}\left[T_k^j(T)\right]
+    & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
+    & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
+  \end{align}
+  which yields Lemma~\ref{lem:5:SubOptimalSelections},
+  with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.
+\end{proof}
+
 % -----------------------------------------------------------------
@@
 for both \RandTopM{} and \MCTopM{}. It essentially proves that Lines~$3$-$5$ in Algorithm~\ref{algo:5:MCTopM} (when a new arm is sampled from the non-empty subset of $\TopM(t)$)
-happen a logarithmic number of times. The proof of this result is given in Appendix~\ref{proof:5:elementaryLemma_RandTopM_MCTopM}.
+happen a logarithmic number of times. The proof of this result is given below.
+% in Appendix~\ref{proof:5:elementaryLemma_RandTopM_MCTopM}.
 
@@
 
+\begin{proof}
+  Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
+  \begin{align*}
+    & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
+    & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
+    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
+  \end{align*}
+
+  Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
+  \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
+  The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
+To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
+  Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
+  and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
+  \begin{align}
+    \sum_{t=1}^T & \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
+    & =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
+    & \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
+    & \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
+  \end{align}
+  where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
+  From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
+  \[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]
+
+  If $\mu_k > \mu_{k'}$, we rather use that
+  \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
+  and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
+  \[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
+  Summing over $k'$ yields the result.
+\end{proof}
+
 % -----------------------------------------------------------------
@@
 
-For \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
+For our proposal \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
 First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
-that we can then bound by the number of changes of arms
-(\cf{} proof in Appendix~\ref{proof:5:collisionsMCTopM}).
+that we can then bound by the number of changes of arms.
 %
@@
 %
-We note that introducing \emph{switching costs}, like it was done in previous works like \cite{Koren17}, could be an interesting future work.
+We note that introducing \emph{switching costs}, like it was done in previous works like \cite{Koren17}, is an interesting future work.
 
@@
   is upper-bounded by
-  \hfill{}
-  (Proved in Appendix~\ref{proof:5:collisionsMCTopM})
+  % \hfill{}
+  % (Proved in Appendix~\ref{proof:5:collisionsMCTopM})
   \begin{equation}
@@
 %, due to our trick of focusing on collisions for non-sitted players.
-However, unlike \MusicalChair{} our algorithm does not need the knowledge of $\mu^*_{M}-\mu^*_{M+1}$.
+However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity.
+It is indeed not very satisfying, from an applicative point of view, the require a prior knowledge
+of $\mu^*_{M}-\mu^*_{M+1}$ if one wants to run the \MusicalChair{} algorithm.
+
+\begin{proof}
+  A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small.
+Its proof is postponed to the end of this section.
+
+  %
+  Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} presented above provides a schematic representation of the execution of the \MCTopM{} algorithm, that has to be exploited in order to properly control the number of collisions.
+  %
+  The sketch of the proof is the following: by focusing only on collisions in the ``not fixed'' state, bounding the number of transitions $(2)$ and $(3)$ is enough.
+  Then, we show that both the number of transitions $(3)$ and $(5)$ are small: as a consequence of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, the average number of these transitions is $\bigO{\log T}$.
+  Finally, we use that the length of a sequence of consecutive transitions $(2)$ is also small (on average smaller than $M$), and except for possibly the first one, starting a new sequence implies a previous transition $(3)$ or $(5)$ to arrive in the state ``not fixed''.
+This gives a logarithmic number of transitions $(2)$ and $(3)$, and so gives $\E_{\mu}[\sum_k\cC_k(T)] = \bigO{\log T}$,
+  with explicit constants depending on $\boldsymbol{\mu}$ and $M$.
+
+  % \begin{figure}[h!]
+  %   \resizebox{0.98\textwidth}{!}{
+  %   \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
+  %       %
+  %       \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
+  %       \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
+  %       \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
+  %       %
+  %       \draw [black,->] (start) -> (notfixed.20);
+  %       \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
+  %       \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
+  %       \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
+  %       \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
+  %       \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
+  %       %
+  %   \end{tikzpicture}
+  %   }
+  %   \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
+  %   Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
+  %   \label{fig:5:StateMachineAlgorithm_MCTopM}
+  % \end{figure}
+
+  % Explanation of the figure
+
+  As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
+  Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
+  % following the algorithm of \MCTopM:
+  \begin{equation}
+      s^j(t+1) =
+      \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
+      \cap \left( A^j(t) \in \TopM(t) \right).
+  \end{equation}
+
+  For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words.
+At step $t$, if player $j$ is not fixed ($\overline{s^j(t)}$), she can have three behaviors when executing \MCTopM.
+  She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
+  or she stays in state $\overline{s^j(t)}$,
+  with two cases.
+  Either she sampled $A^j(t+1)$ uniformly
+  from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
+  with transition $(3)$,
+  in case of collision and if $A^j(t+1) \in \TopM(t)$,
+  or she sampled $A^j(t+1)$ uniformly
+  from $\TopM(t)$ with transition $(2)$,
+  if $A^j(t+1) \notin \TopM(t)$.
+  In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
+  %
+  % The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
+  % and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
+  % and experienced a collision at time $t-1$
+  Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).
+
+  For player $j$ and round $t$, we now introduce a few events that are useful in the proof.
+First, for every $x=1,2,3,4,5$, we denote $I_x^j(t)$ the event that a transition of type $(x)$ occurs for player $j$ after the first $t$ observations (\ie, between round $t$ and round $t+1$, to decide $A^j(t+1)$).
+  Formally they are defined by
+  \begin{align*}
+    I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
+    % go sitted
+    I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
+    \ \ \ &\text{and} \ \ \
+    I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
+    % if sitted
+    I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
+    \ \ \ &\text{and} \ \ \
+    I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
+  \end{align*}
+  %
+  Then, we introduce $\widetilde{C^j}(t)$ as the event that a collision occurs for player $j$ at round $t$ if she is not yet fixed on her arm, that is
+  \begin{equation}
+      \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
+  \end{equation}
+
+  A key observation is that $C^j(t)$ implies $\bigcup_{j'=1}^M \widetilde{C^{j'}}(t)$, as a collision necessarily involves at least one player not yet fixed on her arm ($\overline{s^{j'(t)}}$).
+  Otherwise, if they are all fixed, \ie, for all $j$, $s^j(t)$, then by definition of $s^j(t)$, none of the player changed their arm from $t-1$ to $t$, and none experienced any collision at time $t-1$ so by induction there is no collision at time $t$.
+  %
+  Thus, $\sum_{j=1}^M \Pr(C^j(t))$ can be upper bounded by $M \sum_{j=1}^M \Pr(\widetilde{C^j}(t))$ (union bound),
+  and it follows that if $\cC(T) := \sum_{k=1}^K \cC_k(T)$ then
+  \[\E_{\mu}[\cC(T)] \leq M \sum_{j=1}^M \sum_{t=1}^T \Pr(\widetilde{C^j}(t)).\]
+  We can further  observe that $\widetilde{C^j}(t)$ implies a transition $(2)$ or $(3)$, as a transition $(1)$ cannot happen in case of collision. Thus another union bound gives
+  %
+  \begin{align}
+    \sum_{t=1}^T \Pr(\widetilde{C^j}(t))
+    &\leq \sum_{t=1}^T \Pr(I_2^j(t))   + \sum_{t=1}^T \Pr(I_3^j(t)).\label{eq:5:UBTilde}
+  \end{align}
+  In the rest of the proof we focus on bounding the number of transitions $(2)$ and $(3)$.
+
+  Let $N_x^j(T)$ be the random variable denoting the number of transitions of type $(x)$.
+  Neglecting the event $\overline{s^j(t)}$ for $x=3$ and $s^j(t)$ for $x=5$, one has
+  \begin{equation}
+      \E_{\mu}[N_x^j(t)]
+      = \sum_{t=1}^T \Pr(I_x^j(t))
+      \leq \sum_{t=1}^T \Pr \left(A^j(t) \notin \TopM(t)\right)
+      \leq \sum_{t=1}^T \sum_{k=1}^K \Pr \left(A^j(t)=k, k \notin \TopM(t)\right),
+  \end{equation}
+  which is $\bigO{\log T}$ (with known constants) by Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}. In particular, this controls the second term in the right hand side of \eqref{eq:5:UBTilde}.
+
+  To control the first term $\sum_{t=1}^T \Pr(I_2^j(t))$
+  we introduce three sequences of random variables,
+  the starting times $(\theta_i)_{i \geq 1}$
+  and the ending times $(\tau_i)_{i \geq 1}$
+  (possibly larger than $T$),
+  of sequences during which $I_2(s)$ is true for all $s=\theta_i,\dots,\tau_i-1$ but not before and after,
+  that is
+  $\forall i \in \{1,\dots,n(T)\},
+  \overline{I_2^j(\theta_i - 1)}
+  \cap \bigcap_{t=\theta_i}^{\tau_i-1} I_2^j(t)
+  \cap \overline{I_2^j(\tau_i)}
+  $
+  with $n(T)$ the number of such sequences,
+  \ie,
+  $n(T) := \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
+  (or $0$ if $\theta_1$ does not exist).
+  %
+  If $\theta_i = 1$, the first sequence does not have term $\overline{I_2^j(\theta_i - 1)}$.
+
+  Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
+  \begin{align}
+    \E_{\mu}[N_2^j(t)]
+    &= \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
+    =
+    \E_{\mu}\left[ \sum_{i=1}^{n(T)} \left( \sum_{t=\theta_i}^{\tau_i - 1} 1 + \sum_{t=\tau_i}^{\theta_{i+1} - 1} 0 \right) \right]
+    =
+    \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
+    \intertext{Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$), and so we can obtain a decomposition with two terms $(\alpha)$ and $(\beta)$:}
+    % so we can use Wald's Lemma \citep{Wald45}
+    &\leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{\beta}.
+    % = (\alpha) \times (\beta).
+  \end{align}
+
+  $(\alpha)$ To control $\E_{\mu}[n(T)]$, we can observe that
+  the number of sequences $n(T)$ is smaller than $1$ plus the number of times when \emph{a sequence begins} ($1$ plus because maybe the game starts in a sequence).
+  %
+  And beginning a sequence
+  % \footnote{Similarly, one could focus on the number of times \emph{a sequence finishes}, and $I_2^j(\theta_i) \cap \overline{I_2^j(\theta_i+1)}$ implies a transition of type $(3)$ or $(1)$.
+  % The key is that after a transition $(1)$, a new sequence of transitions $(2)$ is possible only after a transition $(5)$, which total number is also controlled as $\bigO{\log T}$.}
+  at time $\theta_i$ implies
+  $\overline{I_2^j(\theta_i-1)} \cap I_2^j(\theta_i)$,
+  which implies a transition of type $(3)$ or $(5)$ at time $\theta_i - 1$, as player j is in state ``not fixed'' at time $\theta_i$ (transitions $(1)$ and $(4)$ are impossible).
+  %
+  As stated above, $\E_{\mu}[N_x^j(T)] = \bigO{\log T}$ for both $x=3$ and $x=5$,
+  and so $\E_{\mu}[n(T)] = \bigO{\log T}$ also.
+
+  $(\beta)$ To control $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
+  a simple argument can be used.
+  $\bigcup_{t=\theta_i}^{\tau_i-1} I_2^j(t)$
+  implies $C^j(t)$ for $\tau_i - \theta_i$ consecutive times.
+  %
+  The very structure of \RandTopM{} gives that in this sequence of transitions $(2)$,
+  the successive collisions (\ie, $C^j(t-1) \cap C^j(t)$)
+  implies that each new arm $A^j(t+1)$ for $t \in \{\theta_i, \tau_i-1\}$ is selected uniformly from
+  $\TopM(t+1)$,
+  a set of size $M$ with at least one available arm.
+  %
+  Indeed, as there is $M-1$ other players, at time $t+1$ \emph{at least} one arm in $\TopM(t+1)$ is not selected by any player $k'\neq k$,
+  and so player $j$ has \emph{at least} a probability $1/M$ to select
+  a free arm, which implies $\overline{C^j(t+1)}$, and so implies the end of the sequence.
+  %
+  In other words, the average length of sequences of transitions $(2)$,
+  $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
+  is bounded by the expected number of failed trial of a repeated Bernoulli experiment, with probability of success larger than $1/M$ (by the uniform choice of $A^j(t+1)$ in a set of size $M$ with at least one available arm).
+  We recognize the mean of a geometric random variable, of parameter $\lambda \geq 1/M$, and so $\E_{\mu}\left[ \tau_i - \theta_i \right] = \frac{1}{\lambda} \leq \frac{1}{1/M} = M$.
+
+  This finishes the proof as $\E_{\mu}[N_2^j(T)] = \sum_{t=1}^T \Pr(I_2^j(t)) = \bigO{\log T}$ and so
+  $\sum_{t=1}^T \Pr(\widetilde{C^j}(t) \cap (A^j(t) = k)) = \bigO{\log T}$
+  and finally
+  $\E_{\mu}[\cC(T)] = \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \bigO{\log T}$ also.
+
+  We can be more precise about the constants, all the previous arguments can be used successively:
+  \begin{align}
+    \E_{\mu}[\cC(T)]
+    &\leq M \sum_{j=1}^M \left(\sum_{t=1}^T \Pr(I_2^j(t)) + \sum_{t=1}^T \Pr(I_3^j(t))\right)
+    = M \left(\sum_{j=1}^M \E_{\mu}[N_2^j(T)] + \E_{\mu}[N_3^j(T)]\right) \\
+    &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag  \\
+    &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
+    &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
+    &\leq \left(2 M^3 + M^2\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
+  \end{align}
+
+  And so we obtain the desired inequality, with explicit constants, that depend only on $\boldsymbol{\mu}$ and $M$.
+  %
+  \begin{align}
+    \sum_{k=1}^K \E_{\mu}[\cC^k(T)] & = \E_{\mu}[\cC(T)] \nonumber\\
+    & \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
+  \end{align}
+\end{proof}
 
@@
 
+\paragraph{Logarithmic Regret for \MCTopM}
+%
 Now that the sub-optimal arms selections and the collisions
@@
   \end{equation}
+  Moreover, the dependency of the constant regarding the number of players is $G_{M,\boldsymbol{\mu}} = \bigO{M^3}$.
 \end{theorem}
+
+\paragraph{Number of switches}\label{app:5:NumberSwitches}
+%
+Note that we controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
+which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
+Thus, the total number of arm switches is also proved to be logarithmic, if all players uses
+the \MCTopM-\klUCB{} algorithm.
+
+\paragraph{Strong uniform efficiency}\label{app:5:JustifyingDefinition5}
+%
+As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proved
+to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
+%
+And as justified after Definition~\ref{def:5:DecentralizedUniformEfficiency} (page~\pageref{def:5:DecentralizedUniformEfficiency}), uniform efficiency and invariance under permutations of the users implies strong uniform efficiency, and so \MCTopM{} satisfies Definition~\ref{def:5:DecentralizedUniformEfficiency}.
+%
+This is a sanity check: the lower-bound of
+Theorem~\ref{thm:5:BetterLowerBound} indeed applies to our algorithm \MCTopM,
+and finally this highlights that it is order-optimal for the regret, in the sense that it matches the lower-bound up-to a multiplicative constant,
+and optimal for the term \ref{eq:5:term1}.
 
@@
   % \begin{subfigure}[!h]{0.85\textwidth}
-    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
+    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
   % \end{subfigure}
@@
   % \begin{subfigure}[!h]{0.85\textwidth}
-    \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_HistogramsRegret____env1-1_8200873569864822246.pdf}
+    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_HistogramsRegret____env1-1_8200873569864822246.pdf}
   % \end{subfigure}
@@
 
-To summarize, we presented three variants of Multi-Player Multi-Arm Bandits,
+To sum up, we presented three variants of Multi-Player Multi-Arm Bandits,
 with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
@@
 This work suggests several interesting further research directions.
-First, we want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
+First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
@@
 Second, it remains an open question to know if a simple decentralized algorithm can be as efficient as \MCTopM{} without knowing $M$ in advance, or in dynamic settings (when $M$ can change in time).
-We shall start by proposing variants of our algorithm that are inspired by the \rhoRandEst{} variant of \rhoRand{} proposed by \cite{Anandkumar11}.
-Finally, we want to strengthen the guarantees obtained in the absence of sensing, that is to know whether logarithmic regret is achievable and to have a better analysis of the \Selfish{} approach.
+One could start by proposing variants of our algorithm that are inspired by the \rhoRandEst{} variant of \rhoRand{} proposed by \cite{Anandkumar11}.
+Finally, one could want to strengthen the guarantees obtained in the absence of sensing, that is to know whether logarithmic regret is achievable and to have a better analysis of the \Selfish{} approach.
 Indeed, in most cases, it performs comparably to \RandTopM{} even with limited feedback and without knowing the number of players $M$, which makes it a good candidate for applications to Internet of Things networks.
@@
 
-\paragraph{A note on the simulation code}
-The page \texttt{SMPyBandits.GitHub.io/MultiPlayers.html} explains how to reproduce the experiments used for this Section.
+\paragraph{Reproducibility.}
+The page \texttt{SMPyBandits.GitHub.io/MultiPlayers.html} explains how to reproduce the experiments used for this Chapter.
 

### 2019-04-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Galère ce chapitre 5, je sais pas trop quoi faire.

@@
 % -----------------------------------------------------------------
-\section{Lower Bound: Proof of Theorem~\ref{thm:5:BetterLowerBound}}
+\subsection{Lower Bound: Proof of Theorem~\ref{thm:5:BetterLowerBound}}
 \label{proof:5:BetterLowerBound}
@@
 % -----------------------------------------------------------------
-\subsection{Proof of Theorem~\ref{thm:5:BetterLowerBound}}
+\subsubsection{Proof of Theorem~\ref{thm:5:BetterLowerBound}}
 
@@
 
-\subsection{Proof of Lemma~\ref{lem:5:CD}}
+\subsubsection{Proof of Lemma~\ref{lem:5:CD}}
 
@@
 
-\section{Proofs Elements Related to Regret Upper Bounds}
-\label{proof:5:RegretUpperBounds}
+% \section{Proofs Elements Related to Regret Upper Bounds}
+% \label{proof:5:RegretUpperBounds}
 
-This Appendix includes the main proofs, missing from the content of the article,
-that yield the regret upper bound.
-We start by controlling the sub-optimal draws when the \klUCB{} indices are used (instead of \UCB),
-with any of our proposed algorithms (\MCTopM, \RandTopM) or \rhoRand{}. Then we focus on controlling collisions for \MCTopM-\klUCB.
+% This Appendix includes the main proofs, missing from the content of the article,
+% that yield the regret upper bound.
+% We start by controlling the sub-optimal draws when the \klUCB{} indices are used (instead of \UCB),
+% with any of our proposed algorithms (\MCTopM, \RandTopM) or \rhoRand{}. Then we focus on controlling collisions for \MCTopM-\klUCB.
 
-\subsection{Control of the Sub-optimal Draws for \klUCB: Proof of Lemma~\ref{lem:5:SubOptimalSelections}}
-\label{proof:5:SubOptimalSelections}
+% \subsection{Control of the Sub-optimal Draws for \klUCB: Proof of Lemma~\ref{lem:5:SubOptimalSelections}}
+% \label{proof:5:SubOptimalSelections}
 
-%
-%     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
-%     As only user $j$ is considered here, the superscript $j$ is dropped.
-%
-Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
-The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
-\begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
-Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
-If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
-In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.
+% %
+% %     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
+% %     As only user $j$ is considered here, the superscript $j$ is dropped.
+% %
+% Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
+% The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
+% \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
+% Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
+% If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
+% In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.
 
-Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
-\begin{align*}
-\E_{\mu}[T_k^j(T)]
-&= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
-= \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
-%&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
-&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
-%
-\intertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
-\E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
-&\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
-&\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
-%
-&\leq
-\sum_{t=1}^T
-  \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
-+
-\sum_{m=1}^M
-  \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right),
-\end{align*}
-where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.
+% Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
+% \begin{align*}
+% \E_{\mu}[T_k^j(T)]
+% &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
+% = \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
+% %&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
+% &= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
+% %
+% \intertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
+% \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
+% &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
+% &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
+% %
+% &\leq
+% \sum_{t=1}^T
+%   \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+% +
+% \sum_{m=1}^M
+%   \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right),
+% \end{align*}
+% where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.
 
-Now each of the two terms in the right hand side can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
-The rightmost term can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact 1 in Appendix A of \cite{KLUCBJournal}.
-The leftmost term can be controlled using Lemma~\ref{lem:5:Fact2KLUCB} stated below, that is a direct consequence of the proof of Fact 2 in Appendix A of \cite{KLUCBJournal}.
+% Now each of the two terms in the right hand side can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
+% The rightmost term can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact 1 in Appendix A of \cite{KLUCBJournal}.
+% The leftmost term can be controlled using Lemma~\ref{lem:5:Fact2KLUCB} stated below, that is a direct consequence of the proof of Fact 2 in Appendix A of \cite{KLUCBJournal}.
 
-\begin{lemma}\label{lem:5:Fact1KLUCB}
-    For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
-    \begin{equation}
-      \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
-    \end{equation}
-\end{lemma}
+% \begin{lemma}\label{lem:5:Fact1KLUCB}
+%     For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
+%     \begin{equation}
+%       \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
+%     \end{equation}
+% \end{lemma}
 
-Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).
+% Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).
 
-\begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
-\begin{align*}
-  &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
-  + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
-\end{lemma}
+% \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
+% \begin{align*}
+%   &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
+%   + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
+% \end{lemma}
 
-Putting things together, one obtains the non-asymptotic upper bound
-\begin{align}\label{eq:5:UBprecise}
-  \E_{\mu}\left[T_k^j(T)\right]
-  & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
-  & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
-\end{align}
-which yields Lemma~\ref{lem:5:SubOptimalSelections},
-with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.
+% Putting things together, one obtains the non-asymptotic upper bound
+% \begin{align}\label{eq:5:UBprecise}
+%   \E_{\mu}\left[T_k^j(T)\right]
+%   & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
+%   & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
+% \end{align}
+% which yields Lemma~\ref{lem:5:SubOptimalSelections},
+% with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.
 
-% --------------------------------------------------------------------------------
-\subsection{Proof of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}}\label{proof:5:elementaryLemma_RandTopM_MCTopM}
+% % --------------------------------------------------------------------------------
+% \subsection{Proof of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}}\label{proof:5:elementaryLemma_RandTopM_MCTopM}
 
-Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
-\begin{align*}
-  & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
-  & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
-  & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
-\end{align*}
+% Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
+% \begin{align*}
+%   & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
+%   & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
+%   & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
+% \end{align*}
 
-Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
-\[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
-The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
+% Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
+% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
+% The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
 To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
-Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
-and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
-\begin{align}
-  & \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)
-  =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
-  & \hspace{1cm} \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
-  & \hspace{1cm}  \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
-\end{align}
-where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
-From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
-\[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]
+% Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
+% and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
+% \begin{align}
+%   & \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)
+%   =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
+%   & \hspace{1cm} \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
+%   & \hspace{1cm}  \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
+% \end{align}
+% where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
+% From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
+% \[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]
 
-If $\mu_k > \mu_{k'}$, we rather use that
-\[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
-and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
-\[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
-Summing over $k'$ yields the result.
+% If $\mu_k > \mu_{k'}$, we rather use that
+% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
+% and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
+% \[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
+% Summing over $k'$ yields the result.
 
@@
 
-% --------------------------------------------------------------------------------
-\subsection{Controlling Collisions for \MCTopM: Proof of Lemma~\ref{lem:5:collisionsMCTopM}}
-\label{proof:5:collisionsMCTopM}
+% % --------------------------------------------------------------------------------
+% \subsection{Controlling Collisions for \MCTopM: Proof of Lemma~\ref{lem:5:collisionsMCTopM}}
+% \label{proof:5:collisionsMCTopM}
 
-A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small.
+% A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small.
 Its proof is postponed to the end of this section.
 
-%
-Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below provides a schematic representation of the execution of the \MCTopM{} algorithm, that has to be exploited in order to properly control the number of collisions.
-%
-The sketch of the proof is the following: by focusing only on collisions in the ``not fixed'' state, bounding the number of transitions $(2)$ and $(3)$ is enough.
-Then, we show that both the number of transitions $(3)$ and $(5)$ are small: as a consequence of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, the average number of these transitions is $\bigO{\log T}$.
-Finally, we use that the length of a sequence of consecutive transitions $(2)$ is also small (on average smaller than $M$), and except for possibly the first one, starting a new sequence implies a previous transition $(3)$ or $(5)$ to arrive in the state ``not fixed''.
+% %
+% Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below provides a schematic representation of the execution of the \MCTopM{} algorithm, that has to be exploited in order to properly control the number of collisions.
+% %
+% The sketch of the proof is the following: by focusing only on collisions in the ``not fixed'' state, bounding the number of transitions $(2)$ and $(3)$ is enough.
+% Then, we show that both the number of transitions $(3)$ and $(5)$ are small: as a consequence of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, the average number of these transitions is $\bigO{\log T}$.
+% Finally, we use that the length of a sequence of consecutive transitions $(2)$ is also small (on average smaller than $M$), and except for possibly the first one, starting a new sequence implies a previous transition $(3)$ or $(5)$ to arrive in the state ``not fixed''.
 This gives a logarithmic number of transitions $(2)$ and $(3)$, and so gives $\E_{\mu}[\sum_k\cC_k(T)] = \bigO{\log T}$,
-with explicit constants depending on $\boldsymbol{\mu}$ and $M$.
+% with explicit constants depending on $\boldsymbol{\mu}$ and $M$.
 
-\begin{figure}[h!]
-  \resizebox{0.98\textwidth}{!}{
-  \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
-      %
-      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
-      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
-      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
-      %
-      \draw [black,->] (start) -> (notfixed.20);
-      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
-      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
-      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
-      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
-      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
-      %
-  \end{tikzpicture}
-  }
-  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
-  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
-  \label{fig:5:StateMachineAlgorithm_MCTopM}
-\end{figure}
+% \begin{figure}[h!]
+%   \resizebox{0.98\textwidth}{!}{
+%   \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
+%       %
+%       \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
+%       \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
+%       \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
+%       %
+%       \draw [black,->] (start) -> (notfixed.20);
+%       \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
+%       \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
+%       \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
+%       \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
+%       \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
+%       %
+%   \end{tikzpicture}
+%   }
+%   \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
+%   Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
+%   \label{fig:5:StateMachineAlgorithm_MCTopM}
+% \end{figure}
 
-% Explanation of the figure
+% % Explanation of the figure
 
-As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
-Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
-% following the algorithm of \MCTopM:
-\begin{equation}
-    s^j(t+1) =
-    \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
-    \cap \left( A^j(t) \in \TopM(t) \right).
-\end{equation}
+% As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
+% Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
+% % following the algorithm of \MCTopM:
+% \begin{equation}
+%     s^j(t+1) =
+%     \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
+%     \cap \left( A^j(t) \in \TopM(t) \right).
+% \end{equation}
 
-For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words.
+% For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words.
 At step $t$, if player $j$ is not fixed ($\overline{s^j(t)}$), she can have three behaviors when executing \MCTopM.
-She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
-or she stays in state $\overline{s^j(t)}$,
-with two cases.
-Either she sampled $A^j(t+1)$ uniformly
-from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
-with transition $(3)$,
-in case of collision and if $A^j(t+1) \in \TopM(t)$,
-or she sampled $A^j(t+1)$ uniformly
-from $\TopM(t)$ with transition $(2)$,
-if $A^j(t+1) \notin \TopM(t)$.
-In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
-%
-% The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
-% and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
-% and experienced a collision at time $t-1$
-Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).
+% She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
+% or she stays in state $\overline{s^j(t)}$,
+% with two cases.
+% Either she sampled $A^j(t+1)$ uniformly
+% from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
+% with transition $(3)$,
+% in case of collision and if $A^j(t+1) \in \TopM(t)$,
+% or she sampled $A^j(t+1)$ uniformly
+% from $\TopM(t)$ with transition $(2)$,
+% if $A^j(t+1) \notin \TopM(t)$.
+% In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
+% %
+% % The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
+% % and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
+% % and experienced a collision at time $t-1$
+% Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).
 
-For player $j$ and round $t$, we now introduce a few events that are useful in the proof.
+% For player $j$ and round $t$, we now introduce a few events that are useful in the proof.
 First, for every $x=1,2,3,4,5$, we denote $I_x^j(t)$ the event that a transition of type $(x)$ occurs for player $j$ after the first $t$ observations (\ie, between round $t$ and round $t+1$, to decide $A^j(t+1)$).
-Formally they are defined by
-\begin{align*}
-  I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
-  % go sitted
-  I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
-  \ \ \ &\text{and} \ \ \
-  I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
-  % if sitted
-  I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
-  \ \ \ &\text{and} \ \ \
-  I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
-\end{align*}
-%
-Then, we introduce $\widetilde{C^j}(t)$ as the event that a collision occurs for player $j$ at round $t$ if she is not yet fixed on her arm, that is
-\begin{equation}
-    \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
-\end{equation}
+% Formally they are defined by
+% \begin{align*}
+%   I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
+%   % go sitted
+%   I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
+%   \ \ \ &\text{and} \ \ \
+%   I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
+%   % if sitted
+%   I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
+%   \ \ \ &\text{and} \ \ \
+%   I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
+% \end{align*}
+% %
+% Then, we introduce $\widetilde{C^j}(t)$ as the event that a collision occurs for player $j$ at round $t$ if she is not yet fixed on her arm, that is
+% \begin{equation}
+%     \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
+% \end{equation}
 
-A key observation is that $C^j(t)$ implies $\bigcup_{j'=1}^M \widetilde{C^{j'}}(t)$, as a collision necessarily involves at least one player not yet fixed on her arm ($\overline{s^{j'(t)}}$).
-Otherwise, if they are all fixed, \ie, for all $j$, $s^j(t)$, then by definition of $s^j(t)$, none of the player changed their arm from $t-1$ to $t$, and none experienced any collision at time $t-1$ so by induction there is no collision at time $t$.
-%
-Thus, $\sum_{j=1}^M \Pr(C^j(t))$ can be upper bounded by $M \sum_{j=1}^M \Pr(\widetilde{C^j}(t))$ (union bound),
-and it follows that if $\cC(T) := \sum_{k=1}^K \cC_k(T)$ then
-\[\E_{\mu}[\cC(T)] \leq M \sum_{j=1}^M \sum_{t=1}^T \Pr(\widetilde{C^j}(t)).\]
-We can further  observe that $\widetilde{C^j}(t)$ implies a transition $(2)$ or $(3)$, as a transition $(1)$ cannot happen in case of collision. Thus another union bound gives
-%
-\begin{align}
-  \sum_{t=1}^T \Pr(\widetilde{C^j}(t))
-  &\leq \sum_{t=1}^T \Pr(I_2^j(t))   + \sum_{t=1}^T \Pr(I_3^j(t)).\label{eq:5:UBTilde}
-\end{align}
-In the rest of the proof we focus on bounding the number of transitions $(2)$ and $(3)$.
+% A key observation is that $C^j(t)$ implies $\bigcup_{j'=1}^M \widetilde{C^{j'}}(t)$, as a collision necessarily involves at least one player not yet fixed on her arm ($\overline{s^{j'(t)}}$).
+% Otherwise, if they are all fixed, \ie, for all $j$, $s^j(t)$, then by definition of $s^j(t)$, none of the player changed their arm from $t-1$ to $t$, and none experienced any collision at time $t-1$ so by induction there is no collision at time $t$.
+% %
+% Thus, $\sum_{j=1}^M \Pr(C^j(t))$ can be upper bounded by $M \sum_{j=1}^M \Pr(\widetilde{C^j}(t))$ (union bound),
+% and it follows that if $\cC(T) := \sum_{k=1}^K \cC_k(T)$ then
+% \[\E_{\mu}[\cC(T)] \leq M \sum_{j=1}^M \sum_{t=1}^T \Pr(\widetilde{C^j}(t)).\]
+% We can further  observe that $\widetilde{C^j}(t)$ implies a transition $(2)$ or $(3)$, as a transition $(1)$ cannot happen in case of collision. Thus another union bound gives
+% %
+% \begin{align}
+%   \sum_{t=1}^T \Pr(\widetilde{C^j}(t))
+%   &\leq \sum_{t=1}^T \Pr(I_2^j(t))   + \sum_{t=1}^T \Pr(I_3^j(t)).\label{eq:5:UBTilde}
+% \end{align}
+% In the rest of the proof we focus on bounding the number of transitions $(2)$ and $(3)$.
 
-Let $N_x^j(T)$ be the random variable denoting the number of transitions of type $(x)$.
-Neglecting the event $\overline{s^j(t)}$ for $x=3$ and $s^j(t)$ for $x=5$, one has
-\begin{equation}
-    \E_{\mu}[N_x^j(t)]
-    = \sum_{t=1}^T \Pr(I_x^j(t))
-    \leq \sum_{t=1}^T \Pr \left(A^j(t) \notin \TopM(t)\right)
-    \leq \sum_{t=1}^T \sum_{k=1}^K \Pr \left(A^j(t)=k, k \notin \TopM(t)\right),
-\end{equation}
-which is $\bigO{\log T}$ (with known constants) by Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}. In particular, this controls the second term in the right hand side of \eqref{eq:5:UBTilde}.
+% Let $N_x^j(T)$ be the random variable denoting the number of transitions of type $(x)$.
+% Neglecting the event $\overline{s^j(t)}$ for $x=3$ and $s^j(t)$ for $x=5$, one has
+% \begin{equation}
+%     \E_{\mu}[N_x^j(t)]
+%     = \sum_{t=1}^T \Pr(I_x^j(t))
+%     \leq \sum_{t=1}^T \Pr \left(A^j(t) \notin \TopM(t)\right)
+%     \leq \sum_{t=1}^T \sum_{k=1}^K \Pr \left(A^j(t)=k, k \notin \TopM(t)\right),
+% \end{equation}
+% which is $\bigO{\log T}$ (with known constants) by Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}. In particular, this controls the second term in the right hand side of \eqref{eq:5:UBTilde}.
 
-To control the first term $\sum_{t=1}^T \Pr(I_2^j(t))$
-we introduce three sequences of random variables,
-the starting times $(\theta_i)_{i \geq 1}$
-and the ending times $(\tau_i)_{i \geq 1}$
-(possibly larger than $T$),
-of sequences during which $I_2(s)$ is true for all $s=\theta_i,\dots,\tau_i-1$ but not before and after,
-that is
-$\forall i \in \{1,\dots,n(T)\},
-\overline{I_2^j(\theta_i - 1)}
-\cap \bigcap_{t=\theta_i}^{\tau_i-1} I_2^j(t)
-\cap \overline{I_2^j(\tau_i)}
-$
-with $n(T)$ the number of such sequences,
-\ie,
-$n(T) := \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
-(or $0$ if $\theta_1$ does not exist).
-%
-If $\theta_i = 1$, the first sequence does not have term $\overline{I_2^j(\theta_i - 1)}$.
+% To control the first term $\sum_{t=1}^T \Pr(I_2^j(t))$
+% we introduce three sequences of random variables,
+% the starting times $(\theta_i)_{i \geq 1}$
+% and the ending times $(\tau_i)_{i \geq 1}$
+% (possibly larger than $T$),
+% of sequences during which $I_2(s)$ is true for all $s=\theta_i,\dots,\tau_i-1$ but not before and after,
+% that is
+% $\forall i \in \{1,\dots,n(T)\},
+% \overline{I_2^j(\theta_i - 1)}
+% \cap \bigcap_{t=\theta_i}^{\tau_i-1} I_2^j(t)
+% \cap \overline{I_2^j(\tau_i)}
+% $
+% with $n(T)$ the number of such sequences,
+% \ie,
+% $n(T) := \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
+% (or $0$ if $\theta_1$ does not exist).
+% %
+% If $\theta_i = 1$, the first sequence does not have term $\overline{I_2^j(\theta_i - 1)}$.
 
-Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
-\begin{align}
-  \E_{\mu}[N_2^j(t)]
-  &= \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
-  =
-  \E_{\mu}\left[ \sum_{i=1}^{n(T)} \left( \sum_{t=\theta_i}^{\tau_i - 1} 1 + \sum_{t=\tau_i}^{\theta_{i+1} - 1} 0 \right) \right]
-  =
-  \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
-  \intertext{Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$), and so we can obtain}
-  % so we can use Wald's Lemma \citep{Wald45}
-  &\leq \E_{\mu}\left[ n(T) \right] \times \max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]
-  = (\alpha) \times (\beta).
-\end{align}
+% Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
+% \begin{align}
+%   \E_{\mu}[N_2^j(t)]
+%   &= \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
+%   =
+%   \E_{\mu}\left[ \sum_{i=1}^{n(T)} \left( \sum_{t=\theta_i}^{\tau_i - 1} 1 + \sum_{t=\tau_i}^{\theta_{i+1} - 1} 0 \right) \right]
+%   =
+%   \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
+%   \intertext{Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$), and so we can obtain}
+%   % so we can use Wald's Lemma \citep{Wald45}
+%   &\leq \E_{\mu}\left[ n(T) \right] \times \max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]
+%   = (\alpha) \times (\beta).
+% \end{align}
 
-$(\alpha)$ To control $\E_{\mu}[n(T)]$, we can observe that
-the number of sequences $n(T)$ is smaller than $1$ plus the number of times when \emph{a sequence begins} ($1$ plus because maybe the game starts in a sequence).
-%
-And beginning a sequence
-% \footnote{Similarly, one could focus on the number of times \emph{a sequence finishes}, and $I_2^j(\theta_i) \cap \overline{I_2^j(\theta_i+1)}$ implies a transition of type $(3)$ or $(1)$.
-% The key is that after a transition $(1)$, a new sequence of transitions $(2)$ is possible only after a transition $(5)$, which total number is also controlled as $\bigO{\log T}$.}
-at time $\theta_i$ implies
-$\overline{I_2^j(\theta_i-1)} \cap I_2^j(\theta_i)$,
-which implies a transition of type $(3)$ or $(5)$ at time $\theta_i - 1$, as player j is in state ``not fixed'' at time $\theta_i$ (transitions $(1)$ and $(4)$ are impossible).
-%
-As stated above, $\E_{\mu}[N_x^j(T)] = \bigO{\log T}$ for both $x=3$ and $x=5$,
-and so $\E_{\mu}[n(T)] = \bigO{\log T}$ also.
+% $(\alpha)$ To control $\E_{\mu}[n(T)]$, we can observe that
+% the number of sequences $n(T)$ is smaller than $1$ plus the number of times when \emph{a sequence begins} ($1$ plus because maybe the game starts in a sequence).
+% %
+% And beginning a sequence
+% % \footnote{Similarly, one could focus on the number of times \emph{a sequence finishes}, and $I_2^j(\theta_i) \cap \overline{I_2^j(\theta_i+1)}$ implies a transition of type $(3)$ or $(1)$.
+% % The key is that after a transition $(1)$, a new sequence of transitions $(2)$ is possible only after a transition $(5)$, which total number is also controlled as $\bigO{\log T}$.}
+% at time $\theta_i$ implies
+% $\overline{I_2^j(\theta_i-1)} \cap I_2^j(\theta_i)$,
+% which implies a transition of type $(3)$ or $(5)$ at time $\theta_i - 1$, as player j is in state ``not fixed'' at time $\theta_i$ (transitions $(1)$ and $(4)$ are impossible).
+% %
+% As stated above, $\E_{\mu}[N_x^j(T)] = \bigO{\log T}$ for both $x=3$ and $x=5$,
+% and so $\E_{\mu}[n(T)] = \bigO{\log T}$ also.
 
-$(\beta)$ To control $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
-a simple argument can be used.
-$\bigcup_{t=\theta_i}^{\tau_i-1} I_2^j(t)$
-implies $C^j(t)$ for $\tau_i - \theta_i$ consecutive times.
-%
-The very structure of \RandTopM{} gives that in this sequence of transitions $(2)$,
-the successive collisions (\ie, $C^j(t-1) \cap C^j(t)$)
-implies that each new arm $A^j(t+1)$ for $t \in \{\theta_i, \tau_i-1\}$ is selected uniformly from
-$\TopM(t+1)$,
-a set of size $M$ with at least one available arm.
-%
-Indeed, as there is $M-1$ other players, at time $t+1$ \emph{at least} one arm in $\TopM(t+1)$ is not selected by any player $k'\neq k$,
-and so player $j$ has \emph{at least} a probability $1/M$ to select
-a free arm, which implies $\overline{C^j(t+1)}$, and so implies the end of the sequence.
-%
-In other words, the average length of sequences of transitions $(2)$,
-$\E_{\mu}\left[ \tau_i - \theta_i \right]$,
-is bounded by the expected number of failed trial of a repeated Bernoulli experiment, with probability of success larger than $1/M$ (by the uniform choice of $A^j(t+1)$ in a set of size $M$ with at least one available arm).
-We recognize the mean of a geometric random variable, of parameter $\lambda \geq 1/M$, and so $\E_{\mu}\left[ \tau_i - \theta_i \right] = \frac{1}{\lambda} \leq \frac{1}{1/M} = M$.
+% $(\beta)$ To control $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
+% a simple argument can be used.
+% $\bigcup_{t=\theta_i}^{\tau_i-1} I_2^j(t)$
+% implies $C^j(t)$ for $\tau_i - \theta_i$ consecutive times.
+% %
+% The very structure of \RandTopM{} gives that in this sequence of transitions $(2)$,
+% the successive collisions (\ie, $C^j(t-1) \cap C^j(t)$)
+% implies that each new arm $A^j(t+1)$ for $t \in \{\theta_i, \tau_i-1\}$ is selected uniformly from
+% $\TopM(t+1)$,
+% a set of size $M$ with at least one available arm.
+% %
+% Indeed, as there is $M-1$ other players, at time $t+1$ \emph{at least} one arm in $\TopM(t+1)$ is not selected by any player $k'\neq k$,
+% and so player $j$ has \emph{at least} a probability $1/M$ to select
+% a free arm, which implies $\overline{C^j(t+1)}$, and so implies the end of the sequence.
+% %
+% In other words, the average length of sequences of transitions $(2)$,
+% $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
+% is bounded by the expected number of failed trial of a repeated Bernoulli experiment, with probability of success larger than $1/M$ (by the uniform choice of $A^j(t+1)$ in a set of size $M$ with at least one available arm).
+% We recognize the mean of a geometric random variable, of parameter $\lambda \geq 1/M$, and so $\E_{\mu}\left[ \tau_i - \theta_i \right] = \frac{1}{\lambda} \leq \frac{1}{1/M} = M$.
 
-This finishes the proof as $\E_{\mu}[N_2^j(T)] = \sum_{t=1}^T \Pr(I_2^j(t)) = \bigO{\log T}$ and so
-$\sum_{t=1}^T \Pr(\widetilde{C^j}(t) \cap (A^j(t) = k)) = \bigO{\log T}$
-and finally
-$\E_{\mu}[\cC(T)] = \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \bigO{\log T}$ also.
+% This finishes the proof as $\E_{\mu}[N_2^j(T)] = \sum_{t=1}^T \Pr(I_2^j(t)) = \bigO{\log T}$ and so
+% $\sum_{t=1}^T \Pr(\widetilde{C^j}(t) \cap (A^j(t) = k)) = \bigO{\log T}$
+% and finally
+% $\E_{\mu}[\cC(T)] = \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \bigO{\log T}$ also.
 
-We can be more precise about the constants, all the previous arguments can be used successively:
-\begin{align}
-  \E_{\mu}[\cC(T)]
-  &\leq M \sum_{j=1}^M \left(\sum_{t=1}^T \Pr(I_2^j(t)) + \sum_{t=1}^T \Pr(I_3^j(t))\right)
-  = M \left(\sum_{j=1}^M \E_{\mu}[N_2^j(T)] + \E_{\mu}[N_3^j(T)]\right) \\
-  &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag  \\
-  &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
-  &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
-  &\leq \left(2 M^3 + M^2\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
-\end{align}
+% We can be more precise about the constants, all the previous arguments can be used successively:
+% \begin{align}
+%   \E_{\mu}[\cC(T)]
+%   &\leq M \sum_{j=1}^M \left(\sum_{t=1}^T \Pr(I_2^j(t)) + \sum_{t=1}^T \Pr(I_3^j(t))\right)
+%   = M \left(\sum_{j=1}^M \E_{\mu}[N_2^j(T)] + \E_{\mu}[N_3^j(T)]\right) \\
+%   &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag  \\
+%   &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
+%   &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
+%   &\leq \left(2 M^3 + M^2\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
+% \end{align}
 
-And so we obtain the desired inequality, with explicit constants, that depend only on $\boldsymbol{\mu}$ and $M$.
-%
-\begin{equation}
-  \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \E_{\mu}[\cC(T)]
-  \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
-\end{equation}
+% And so we obtain the desired inequality, with explicit constants, that depend only on $\boldsymbol{\mu}$ and $M$.
+% %
+% \begin{equation}
+%   \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \E_{\mu}[\cC(T)]
+%   \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
+% \end{equation}
 
-\paragraph{Number of switches}\label{app:5:NumberSwitches}
-%
-Note that we controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
-which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
-Thus, the total number of arm switches is also proved to be logarithmic, if all players uses
-the \MCTopM-\klUCB{} algorithm.
+% \paragraph{Number of switches}\label{app:5:NumberSwitches}
+% %
+% Note that we controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
+% which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
+% Thus, the total number of arm switches is also proved to be logarithmic, if all players uses
+% the \MCTopM-\klUCB{} algorithm.
 
-\paragraph{Strong uniform efficiency}\label{app:5:JustifyingDefinition5}
-%
-As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proved
-to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
-%
-And as justified after Definition~\ref{def:5:DecentralizedUniformEfficiency} (page~\pageref{def:5:DecentralizedUniformEfficiency}), uniform efficiency and invariance under permutations of the users implies strong uniform efficiency, and so \MCTopM{} satisfies Definition~\ref{def:5:DecentralizedUniformEfficiency}.
-%
-This is a sanity check: the lower-bound of
-Theorem~\ref{thm:5:BetterLowerBound} indeed applies to our algorithm \MCTopM,
-and finally this highlights that it is order-optimal for the regret, in the sense that it matches the lower-bound up-to a multiplicative constant,
-and optimal for the term \ref{eq:5:term1}.
+% \paragraph{Strong uniform efficiency}\label{app:5:JustifyingDefinition5}
+% %
+% As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proved
+% to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
+% %
+% And as justified after Definition~\ref{def:5:DecentralizedUniformEfficiency} (page~\pageref{def:5:DecentralizedUniformEfficiency}), uniform efficiency and invariance under permutations of the users implies strong uniform efficiency, and so \MCTopM{} satisfies Definition~\ref{def:5:DecentralizedUniformEfficiency}.
+% %
+% This is a sanity check: the lower-bound of
+% Theorem~\ref{thm:5:BetterLowerBound} indeed applies to our algorithm \MCTopM,
+% and finally this highlights that it is order-optimal for the regret, in the sense that it matches the lower-bound up-to a multiplicative constant,
+% and optimal for the term \ref{eq:5:term1}.
 
@@
 % -----------------------------------------------------------------
-\section{Additional Discussions on \Selfish}
+\subsection{Additional Discussions on \Selfish}
 \label{app:5:SelfishFails}
@@
 % -----------------------------------------------------------------
-\section{Additional Figures}
+\subsection{Additional Figures}
 \label{app:5:moreplots}
@@
 % -----------------------------------------------------------------
-\subsection{Illustration of the lower bound}
+\subsubsection{Illustration of the lower bound}
 \label{app:5:illustrationLowerBound}
@@
 
-\subsection{Figures from Section~\ref{sec:5:experiments}}
+\subsubsection{Figures from Section~\ref{sec:5:experiments}}
 \label{app:5:plotsFromSec5}

### 2019-04-04 4-Listofpublications/Listofpublications.tex
Typo

@@
 Besson}, E.
-Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}
-	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, March 2017
+Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}.
+	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, March 2017.
 	\item	R.
 Bonnefoi, \textbf{L.
-Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016, \href{https://hal.inria.fr/hal-02013839}{HAL-02013839}
+Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016, \href{https://hal.inria.fr/hal-02013839}{HAL-02013839}.
 \end{etaremune}

### 2019-04-05 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Small proofreading of Chapter 4.

--- 2019-04-01 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
@@
 
-\paragraph{Naive policy: Random Channel Selection}
+\paragraph{$1^{\text{st}}$ - Naive policy: Random Channel Selection}
 
@@
 
-\paragraph{(Unachievable) Optimal oracle policy}
+\paragraph{$2^{\text{nd}}$ - (Unachievable) Optimal oracle policy}
 
@@
 
-\paragraph{A greedy approach of the oracle strategy}
+\paragraph{$3^{\text{rd}}$ - A greedy approach of the oracle strategy}
 
@@
     \centering
-    \subfloat[10\% of dynamic devices]{\includegraphics[scale=0.5]{10intelligent.eps}
+    \subfloat[10\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{10intelligent.eps}
     \label{fig:41:10intelligent}}
-    \hfill
-    \subfloat[30\% of dynamic devices]{\includegraphics[scale=0.5]{30intelligent.eps}
+    %
+    \subfloat[30\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{30intelligent.eps}
     \label{fig:41:30intelligent}}
-    \vskip\baselineskip
-    \vspace*{-20pt}
-    \subfloat[50\% of dynamic devices]{\includegraphics[scale=0.5]{50intelligent.eps}
+    %
+    \vspace*{-10pt}
+    \subfloat[50\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{50intelligent.eps}
     \label{fig:41:50intelligent}}
-    \hfill
-    \subfloat[100\% of dynamic devices]{\includegraphics[scale=0.5]{100intelligent.eps}
+    %
+    \subfloat[100\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{100intelligent.eps}
     \label{fig:41:100intelligent}}
-    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
+    % FIXME use a shorter caption for the list of figures in Appendix of the thesis: it takes lots of lines to have these (a)...(d) shown in list of figures
+    \caption[Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:from10to100}
@@
 Figure \ref{fig:41:from10to100} presents the evolution of the successful transmission rate, as a function of time.
-The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy from below, and the two oracle policies (optimal and greedy) from above.
+The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
 The results are displayed when $10\%$, $30\%$, $50\%$ and $100\%$ of the traffic is generated by dynamic devices.
@@
     \centering
-    \includegraphics[scale=0.60]{perf_learning.eps}
+    \includegraphics[scale=0.65]{perf_learning.eps}
     \caption{Learning with \UCB{} and Thomson Sampling, with many dynamic devices. Both algorithms achieve close to optimal performance after a reasonable learning time.}
@@
 
+\begin{figure}[!b]
+    \centering
+    \subfloat[10\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_10.eps}
+    \label{fig:41:ch2_10}}
+    %
+    \subfloat[30\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_30.eps}
+    \label{fig:41:ch2_30}}
+    %
+    \vspace*{-10pt}
+    \subfloat[50\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_50.eps}
+    \label{fig:41:ch2_50}}
+    %
+    \subfloat[100\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_100.eps}
+    \label{fig:41:ch2_100}}
+    \caption{Performance of the UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
+    \label{fig:41:figure4appendix}
+\end{figure}
+
 Figure \ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
@@
 Hence, it is surprising to have near optimal performance with stochastic MAB algorithms applied to partly or fully dynamic scenarios.
-
-\begin{figure}[!h]
-    \centering
-    \subfloat[10\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_10.eps}
-    \label{fig:41:ch2_10}}
-    \hfill
-    \subfloat[30\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_30.eps}
-    \label{fig:41:ch2_30}}
-    \vskip\baselineskip
-    \vspace*{-20pt}
-    \subfloat[50\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_50.eps}
-    \label{fig:41:ch2_50}}
-    \hfill
-    \subfloat[100\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_100.eps}
-    \label{fig:41:ch2_100}}
-    \caption{Performance of the UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
-    \label{fig:41:figure4appendix}
-\end{figure}
 

### 2019-04-05 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Small proofreading of Chapter 4.

@@
 Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
-For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a longer sequence if the PoC have to consider more users\footnote{~With $\ell$ QPSK symbols, we can fit at most $4^{\ell}/2$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway.}.
+For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a longer sequence if the PoC have to consider more users.
+With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway
 Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
@@
     \centering
-    \includegraphics[width=0.70\linewidth]{our-demo.eps}
+    \includegraphics[width=0.75\linewidth]{our-demo.eps}
     \caption{Schematic of our implementation that presents the role of each USRP platform.}
@@
 As shown below in Figure~\ref{fig:42:screenshotDemoYouTube},
-we made a $6$-minute \textbf{video} to show our demonstration online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+we realized and released a $6$-minute \textbf{video} to show our demonstration online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
 The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
@@
     % \includegraphics[width=0.90\textwidth]{Images/screenshotDemoYouTube.png}
-    \includegraphics[width=0.90\textwidth]{2-Chapters/4-Chapter/Images/screenshotDemoYouTube.png}
+    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/screenshotDemoYouTube.png}
     \caption{Screenshot of the \textbf{video} of our demonstration, \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.}

### 2019-04-05 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Small modifications on all chapters
Small proofreading of Chapter 4.

@@
 
-In this section, we extend the model presented before in order to take into account the possibility for retransmission of a message after a collision.
-This extension is of major importance as most protocols for real-world IoT networks can use retransmissions, and it is for instance the case of the LoRa standard.
+In this section, we extend the model presented before, to take into account the possibility for retransmissions of a message after a collision.
+This extension is of major importance: most protocols for real-world IoT networks can use retransmissions, it is for instance the case of the LoRa standard.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.
@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{ResultsUCB.eps}
+	\includegraphics[width=0.80\linewidth]{ResultsUCB.eps}
 	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB'']{
@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{ResultsUCB2.eps}
+	\includegraphics[width=0.80\linewidth]{ResultsUCB2.eps}
 	\caption[First comparison between the exposed heuristics for the retransmission: Only \UCB, Random, \UCB, $K$ \UCB, and Delayed \UCB]{

### 2019-04-05 2-Chapters/5-Chapter/chapter5.tex
Galère ce chapitre 5, je sais pas trop quoi faire.
Started to write the « related works for other models » section in Chapter:5

@@
 \minitoc
+
+\paragraph{Abstract}
 
@@
 % ----------------------------------------------------------------------------
-\section{Literature review of algorithms solving the ``no sensing'' case}
-\label{sec:5:noSensingCase}
+\section{Many extensions of our models: literature review and numerical experiments}
+\label{sec:5:literatureReviewOtherModels}
 
 \TODOL{I also need to include additional simulations from the latest research articles, and discussions about the new works that took inspiration in our paper.}
+
+\TODOL{Just a small sentence saying that we present here all the extensions of our models that we are aware of blabla}
+
+\subsection{Unknown (fixed) number of players}
+\label{sub:5:unknownNumberOfPlayers}
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\subsection{Without sensing information}
+\label{sub:5:withoutSensing}
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\subsubsection{Additionnal numerical simulations}
+
+\TODOL{Do some basic simulations and show !}
+
+\subsection{With communication or coordination between players}
+\label{sub:5:withCommunicationOrCoordination}
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\subsection{Arrival and departures of players}
+\label{sub:5:arrivalDepartures}
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\subsection{With different arm utilities among players}
+\label{sub:5:withDifferentMeansAmongPlayers}
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\subsection{With malicious jammers}
+\label{sub:5:withDifferentMeansAmongPlayers}
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\subsection{Real-world demonstrations to validate some models}
+\label{sub:5:USRPdemos}
+
+Similarly to our demonstration of multi-armed bandit learning in an IoT network that we presented in the previous Chapter~\ref{chapter:4} in Section~\ref{sec:4:gnuradio},
+we review in this section some related works, dating back from $2016$, who proposed a similar approach.
+All the similar works that we are aware of also used USRP boards, and implemented the demonstration using either the Simulink and MATLAB softwares, or the GNU Radio software like we did \cite{Besson2018ICT}.
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\subsection{Towards non-stationary multi-players MAB models}
+\label{sub:5:towardsNonStationaryModels}
+
+\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 

### 2019-04-05 2-Chapters/5-Chapter/chapter5.tex
Added all the references that I am aware of, regarding extensions of our model in Chapter:5

@@
 % ----------------------------------------------------------------------------
-\section{Many extensions of our models: literature review and numerical experiments}
+% ----------------------------------------------------------------------------
+\section{Literature review of many extensions of our models}
 \label{sec:5:literatureReviewOtherModels}
+% ----------------------------------------------------------------------------
+
+We conclude this chapter with a literature review of the many extensions of the three models presented above in Section~\ref{sec:5:model}.
+We wanted to highlight that the community was quite active on research on multi-player bandits in the last $10$ years, but especially active since last spring $2018$.
+For instance, our article \cite{Besson2018ALT} was the first to propose the ``no sensing'' case, and it was studied by (at least) three independent group of researchers since its publication in April $2018$ \cite{BoursierPerchet18,LugosiMehrabian18,AvnerMannor18}.
+Another example is the model with different ear means among players, who was studied first by \cite{Bistritz18} and shortly after by \cite{KaufmannAbbas19}.
+%
+By following actively the research on multi-player bandit models since the last three years, we believe to be able to give an exhaustive literature review, but it is of course possible that we missed an important research work.
+In this case, it is our mistake and not an intentional choice as we included all the different extensions of our models that we were aware of.
+
+% \TODOL{Just a small sentence saying that we present here all the extensions of our models that we are aware of blabla}
+
+Many extensions of the simpler model of multi-player bandits have been considered.
+The number of players $M$ can be fixed but initially unknown to the players, the sensing information can be absent (like our \modeltrois{} presented above), there could be communications between players (happening at each time step or only occasionally), or $M$ could evolve over time to model the possibility of arrivals or departures of players (\ie, connexions and disconnections of devices in a wireless networks).
+Also inspired by real-world wireless networks, the arm means may vary among players, for instance to model players located at different distance of the gateway, and we can also consider networks with some ``jammers'' whose purpose is not to communicate to the gateway in a collaborative way, like the $M$ devices, but to interfere with the communications of the $M$ devices.
+
+Finally, we can also be interested by models where $M$ stays fixed, but the environment evolves, for instance with abrupt changes in the means of arms.
+This last model makes a good connexion between this chapter and the next one, and an exciting future work is to further study this model in order to tackle this kind of problems by merging our proposals for stationary multi-player bandits and for non-stationary single-player bandits.
 
@@
 
-\TODOL{Just a small sentence saying that we present here all the extensions of our models that we are aware of blabla}
-
+% ----------------------------------------------------------------------------
 \subsection{Unknown (fixed) number of players}
@@
 
+In the model we presented above, we assumed the number of players $M$ to be fixed in time during the learning process.
+Without removing this hypothesis (we study this case below in Section~\ref{sub:5:arrivalDepartures}), it is interesting to remark that we made no hypothesis on whether the players can known this value $M$ or not.
+Our proposals, \RandTopM{} and \MCTopM, both assume to know $M$ beforehand, like it was done for our inspirational algorithms \RhoRand{} and \MusicalChair.
+
+\paragraph{Performance of our proposals for a wrong value of $M$?}
+%
+We can start by asking whether our efficient proposal, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different than the real number of players $M$.
+Even though we did not include numerical simulations for this case in Section~\ref{sec:5:experiments} above, we did some tests, and we observed two disappointing results.
+First, the three algorithms (\RhoRand, \RandTopM{} and \MCTopM) give linear regret if they are using a value $M'$ strictly smaller than $M$ (as soon as $\mu^*_{M-1} > \mu^*_M$), as players will converge to play ($T-\bigO{\log(T)}$ times) the $M-1$ best arms, leading to at least one collision and thus a linear regret.
+Second, if they are using $M' > M$, then there is also a certain probability to achieve a linear regret.
+Indeed, imagine one player ($M=1$) running \MCTopM{} with the false knowledge of $M'=2$, and if it learned to accurately identify the set of the $2$ best arms, then the \MCTopM{} orthogonalization scheme can make it play the worst of the two arms, and as $M=1$ this player will never encounter any collision, thus playing this suboptimal arm for $T-\bigO{\log(T)}$ times, leading to linear regret.
+
+\paragraph{Physical interpretation of the hypothesis of knowning $M$.}
+%
+% \TODOL{Explain that it does make less sense to know $M$}
+One criticism that could be made about our approach is the fact that if we study a wireless networks with no central coordination from the gateway,
+in particular where the devices cannot be assigned to a channel or be assigned a unique ID by the gateway when they first log in the network.
+Then in such networks, if one wants to apply an algorithm like \MCTopM-\klUCB,
+the $M$ devices need to known their number $M$, and have to receive it from the gateway, as it is the only part of this example of network which knowns $M$.
+It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in\{1,\dots,K\}$.
+
+\TODOL{Explain why the problem is MUCH easier if we have before hand these unique ID!}
+
+\paragraph{Literature review}
+
+Some works studied the same model as our model \modeldeux{} ``with sensing'', under the hypothesis
+
+\paragraph{Extension of our proposals to learn the value of $M$?}
+
+Similarly to what was proposed for the \rhoRandEst{} algorithm, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
+Two possible directions exist to estimate on the fly this number.
+The first idea comes from \cite{Anandkumar11}, and is quite simple.
+All players will build an estimate $\hat{M}^j(t)$ of the number of player, that start by $\hat{M}^j(0)=1$.
+As soon as one collision is observed, a player knows that $M\geq2$, so $\hat{M}^j(t+1) = 2$.
+Then,
+
 \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 
+% ----------------------------------------------------------------------------
 \subsection{Without sensing information}
@@
 
+\cite{BoursierPerchet18}
+
+Distributed Learning and Optimal Assignment in Multiplayer Heterogeneous Networks
+H Tibrewal, S Patchala, MK Hanawal - arXiv preprint arXiv …, 2019 - arxiv.org
+https://arxiv.org/pdf/1901.03868.pdf
+\cite{Tibrewal2019}
+``communication/collision trick''
+
 \subsubsection{Additionnal numerical simulations}
@@
 
+% ----------------------------------------------------------------------------
 \subsection{With communication or coordination between players}
@@
 
+Multi-user Communication Networks: A Coordinated Multi-armed Bandit Approach
+O Avner, S Mannor - arXiv preprint arXiv:1808.04875, 2018 - arxiv.org
+\cite{AvnerMannor18}
+
+% ----------------------------------------------------------------------------
 \subsection{Arrival and departures of players}
@@
 
+\cite{KumarYadav2018}
+
+Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
+Sumit J Darak, Manjesh K. Hanawal
+https://arxiv.org/pdf/1812.11651
+\cite{DarakHanawal18}
+
+% ----------------------------------------------------------------------------
 \subsection{With different arm utilities among players}
@@
 
+\cite{Bistritz18}
+
+\cite{KaufmannAbbas19}
+
+Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
+Sumit J Darak, Manjesh K. Hanawal
+https://arxiv.org/pdf/1812.11651
+\cite{DarakHanawal18}
+
+Distributed Learning and Optimal Assignment in Multiplayer Heterogeneous Networks
+H Tibrewal, S Patchala, MK Hanawal
+https://arxiv.org/pdf/1901.03868.pdf
+\cite{Tibrewal2019}
+``The channel characteristics are unknown  and  could  be  different  for  each  user  (heterogeneous)''
+
+% ----------------------------------------------------------------------------
 \subsection{With malicious jammers}
@@
 
+Learning to Coordinate in a Decentralized Cognitive Radio Network in Presence of Jammers
+Suneet Sawant, Rohit Kumar, Manjesh K. Hanawal, Sumit J. Darak 
+https://arxiv.org/abs/1803.06810
+\cite{SawantKumar2018}
+
+% ----------------------------------------------------------------------------
+\subsection{Modeling more closely a real wireless network}
+\label{sub:5:moreRealisticModels}
+
+\cite{NaparstekCohen17}
+
+Distributed Learning for Channel Allocation Over a Shared Spectrum
+S.M. Zafaruddin, Ilai Bistritz, Amir Leshem, Dusit Niyato
+https://arxiv.org/pdf/1902.06353
+\cite{Zafaruddin2019}
+
+% ----------------------------------------------------------------------------
 \subsection{Real-world demonstrations to validate some models}
@@
 
+\cite{Darak16}
+
+\cite{modiDemo2016}
+
+\cite{kumar2016two}
+
+\cite{KumarYadav2018}
+
+\cite{SawantKumar2018}
+
+\cite{JoshiKumar2018}
+
+% ----------------------------------------------------------------------------
 \subsection{Towards non-stationary multi-players MAB models}
@@
 \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+\cite{WeiSrivastava18Distributed}
+
+\cite{AlaturLevyKrause19}
 

### 2019-04-05 2-Chapters/5-Chapter/chapter5.tex
This review of existing literature on multi-player bandit will take so long... I can spend at least 1 hour on each paper, there is like 12

@@
 
+\TODOL{
+    Pour chaque extension au modèle ci dessous, je veux suivre la même organisation :
+
+    1. j'explique l'extension, et en quoi c'est difficile (= en quoi notre approche échoue si on fait rien de plus),
+    2. j'explique les travaux existants, et les résultats qu'ils obtiennent e.g. en terme de bornes de regret
+    3. si j'ai déjà le code pour, je peux montrer UNE simulation
+}
+
 % ----------------------------------------------------------------------------
@@
 
-Some works studied the same model as our model \modeldeux{} ``with sensing'', under the hypothesis
+Some works studied the same model as our model \modeldeux{} ``with sensing'', under the hypothesis that players do not know in the value of $M$.
 
@@
 
-Similarly to what was proposed for the \rhoRandEst{} algorithm, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
+Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
 Two possible directions exist to estimate on the fly this number.
+
 The first idea comes from \cite{Anandkumar11}, and is quite simple.
@@
 As soon as one collision is observed, a player knows that $M\geq2$, so $\hat{M}^j(t+1) = 2$.
-Then,
+Then, based on probabilistic computations on the expected number of collisions if there were $m$ players, all following the same strategy (\eg, \RhoRand{} in the case of ), and because the formula is simple to compute for different $m$, the authors proposed a statistical test of the hypothesis $M \leq \hat{M}^j(t)$ against $M > \hat{M}^j(t)$ which is expressed as a simple comparison of the current number of collisions (since last update of $\hat{M}^j(t)$) against a threshold.
+If a player observed ``too many'' collisions (that are unlikely to be caused by only $\hat{M}^j(t) - 1$ other players), then she increases her current estimate $\hat{M}^j(t+1) = \hat{M}^j(t) + 1$.
+We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for our proposals \RandTopM{} and \MCTopM.
 
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+The second idea comes from \cite{Rosenski16}, and suppose to know beforehand both the horizon $T$ and a certain measure of the difficulty of the problem (\ie, a lower bound on the minimal gap between two means).
+If all the $M$ players start to play for a long enough time $T_0$ uniformly at random among all the $K$ arms, then the (expected) number of collisions observed $\E[\cC^j_{T_0}]$ by any player $j$ is a (relatively simple) function of $M$.
+By knowing $K$ and $T_0$ and if all players use the same mechanism, then they can invert the formula to obtain the most likely estimate of $M$ which explained the observations of $\cC^j_{T_0}$ collisions.
+This second approach works empirically very fine, but it requires a fine tuning of the stopping time $T_0$.
+Rosenski et al. studies the performance of their algorithm with high-probability bounds \cite{Rosenski16}, and the tuning of $T_0$ they propose depends on prior knowledge on the problem.
+
+\TODOL{Include a simulation to show the behavior of \rhoRand{} vs \rhoRandEst, and \RandTopM{} and \MCTopM{} with prior knowledge of $M$ vs the ``estimate'' versions.}
 
@@
 
+We presented in Section~\ref{sec:5:model} the model \modeltrois, without sensing information.
+Players can only observe $Y_k(t)$, the product of the \iid{} random sensing information from arm $k$ and of the no-collision indicator.
+This model is clearly harder than the model with sensing information, and even though our proposal \Selfish{} works fine in simulations (in terms of average regret), we proved that it can yield linear regret.
+In our paper \cite{Besson2018ALT} as well as the previous sections, it was left as a future work to know if an algorithm can achieve sub-linear regret in this harder model without sensing information.
+
 \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 
-\cite{BoursierPerchet18}
+Inspired by our article \cite{Besson2018ALT}, Boursier and Perchet studied this question in the summer $2018$ following its publication \cite{BoursierPerchet18}.
+Their answer is that the model without sensing information is essentially not harder than with sensing information.
 
-Distributed Learning and Optimal Assignment in Multiplayer Heterogeneous Networks
-H Tibrewal, S Patchala, MK Hanawal - arXiv preprint arXiv …, 2019 - arxiv.org
-https://arxiv.org/pdf/1901.03868.pdf
-\cite{Tibrewal2019}
-``communication/collision trick''
+\paragraph{About \Selfish-\UCB{} inefficiency.}
+%
+They also prove in Appendix~A \cite{BoursierPerchet18} that \Selfish-\UCB{} has a linear regret, in the theoretical case, with a very neat argument from number theory (using Lindemann-Weierstrass theorem).
+As we conjectured, there is a gap between the theoretical result and its practical consequence, as the Theorem~4 they gave is only valid for real-valued number, and not for hardware-represented floating point number.
+Their proof is supporting what we illustrated in Figure~\ref{fig:5:oneGameTree_SelfishKLUCB}.
+Their argument is essentially to prove that for \Selfish-powered players using the simple \UCB-indexes, with a probability $p$ at time $t$ (both independent from $T$), two players might ahve the same number of pulls and the same observed rewards for each arm.
+In that case, the two players would pull the exact same arms and thus collide for a long time, until they reach a ``tie breaking point'' where they could choose different arms thanks to a random tie breaking rule (\eg, if two values of their \UCB{} indexes are the same, the $\argmax$ is a uniform random choice among the two arms).
+They prove that it is unlikely to encounter any of these ``tie breaking points'', in theory if the \UCB{} indexes are real-valued number (that can be rational or irrational).
 

### 2019-04-06 2-Chapters/5-Chapter/chapter5.tex
Keep working seriously on end of Chapter 5

@@
 
-We conclude this chapter with a literature review of the many extensions of the three models presented above in Section~\ref{sec:5:model}.
+Before concluding this chapter, we give a literature review of the many extensions of the three models presented above in Section~\ref{sec:5:model}.
 We wanted to highlight that the community was quite active on research on multi-player bandits in the last $10$ years, but especially active since last spring $2018$.
 For instance, our article \cite{Besson2018ALT} was the first to propose the ``no sensing'' case, and it was studied by (at least) three independent group of researchers since its publication in April $2018$ \cite{BoursierPerchet18,LugosiMehrabian18,AvnerMannor18}.
-Another example is the model with different ear means among players, who was studied first by \cite{Bistritz18} and shortly after by \cite{KaufmannAbbas19}.
+Another example is the model with different arm means among players, who was studied in \cite{Anandkumar10,Kalathil12} and then more recently in \cite{Bistritz18,KaufmannAbbas19}.
 %
@@
 In this case, it is our mistake and not an intentional choice as we included all the different extensions of our models that we were aware of.
-
-% \TODOL{Just a small sentence saying that we present here all the extensions of our models that we are aware of blabla}
 
@@
 
-\TODOL{I also need to include additional simulations from the latest research articles, and discussions about the new works that took inspiration in our paper.}
-
 \TODOL{
-    Pour chaque extension au modèle ci dessous, je veux suivre la même organisation :
+    Pour chaque extension présentée ci dessous, je veux suivre la même organisation :
 
     1. j'explique l'extension, et en quoi c'est difficile (= en quoi notre approche échoue si on fait rien de plus),
+
     2. j'explique les travaux existants, et les résultats qu'ils obtiennent e.g. en terme de bornes de regret
+
     3. si j'ai déjà le code pour, je peux montrer UNE simulation
@@
 Without removing this hypothesis (we study this case below in Section~\ref{sub:5:arrivalDepartures}), it is interesting to remark that we made no hypothesis on whether the players can known this value $M$ or not.
-Our proposals, \RandTopM{} and \MCTopM, both assume to know $M$ beforehand, like it was done for our inspirational algorithms \RhoRand{} and \MusicalChair.
+Our proposals, \RandTopM{} and \MCTopM, both assume to know $M$ beforehand, like it was done for our main inspiration \RhoRand.
 
@@
 %
-We can start by asking whether our efficient proposal, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different than the real number of players $M$.
-Even though we did not include numerical simulations for this case in Section~\ref{sec:5:experiments} above, we did some tests, and we observed two disappointing results.
-First, the three algorithms (\RhoRand, \RandTopM{} and \MCTopM) give linear regret if they are using a value $M'$ strictly smaller than $M$ (as soon as $\mu^*_{M-1} > \mu^*_M$), as players will converge to play ($T-\bigO{\log(T)}$ times) the $M-1$ best arms, leading to at least one collision and thus a linear regret.
-Second, if they are using $M' > M$, then there is also a certain probability to achieve a linear regret.
-Indeed, imagine one player ($M=1$) running \MCTopM{} with the false knowledge of $M'=2$, and if it learned to accurately identify the set of the $2$ best arms, then the \MCTopM{} orthogonalization scheme can make it play the worst of the two arms, and as $M=1$ this player will never encounter any collision, thus playing this suboptimal arm for $T-\bigO{\log(T)}$ times, leading to linear regret.
+We can start by asking whether our most efficient proposal, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different than the real number of players $M$.
+Even though we did not include numerical simulations for this case in Section~\ref{sec:5:experiments} above, we did some tests, which confirmed two disappointing results that were also proved analytically.
+First, the three algorithms (\RhoRand, \RandTopM{} and \MCTopM) give linear regret if they are using a value $M'$ strictly smaller than $M$ (and if $\mu^*_{M-1} > \mu^*_M$), as players will converge to play about $T-\bigO{\log(T)}$ times the $M-1$ best arms, leading to at least one collision most of the time, and thus a linear regret.
+Second, if the $M$ players falsely use a value $M' > M$, then there is also a certain (fixed) probability to achieve a linear regret.
+Indeed, imagine one player ($M=1$) running \MCTopM{} with the false knowledge of $M'=2$, and if it learned to accurately identify the set of the $2$ best arms, then the \MCTopM{} orthogonalization scheme can make it play the worst of the two arms, and as $M=1$ this player will never encounter any collision, thus playing this suboptimal arm for about $T-\bigO{\log(T)}$ times, also leading to linear regret.
 
-\paragraph{Physical interpretation of the hypothesis of knowning $M$.}
+\paragraph{Interpretation of the hypothesis of knowning $M$ for real-world networks.}
 %
 % \TODOL{Explain that it does make less sense to know $M$}
-One criticism that could be made about our approach is the fact that if we study a wireless networks with no central coordination from the gateway,
+One could criticize our approach if we study a wireless networks with no central coordination from the gateway,
 in particular where the devices cannot be assigned to a channel or be assigned a unique ID by the gateway when they first log in the network.
@@
 It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in\{1,\dots,K\}$.
-
-\TODOL{Explain why the problem is MUCH easier if we have before hand these unique ID!}
+%
+% \TODOL{Explain why the problem is MUCH easier if we have before hand these unique ID!}
+If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm running on top of a round-Robin phase can achieve order-optimal regret.
+Without giving more details, we let the interested refer to what is is explained for the RR-SW-UCB\# algorithm in \cite{WeiSrivastava18Distributed} and \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
 
@@
 Some works studied the same model as our model \modeldeux{} ``with sensing'', under the hypothesis that players do not know in the value of $M$.
-
-\paragraph{Extension of our proposals to learn the value of $M$?}
-
-Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
+As illustrated above, if we consider algorithms building on the same ideas as \RhoRand{} or \MCTopM, it seems mandatory to first build an estimate of the value of $M$ then run the initial algorithm that assumed a perfect knowledge of $M$.
 Two possible directions exist to estimate on the fly this number.
 
-The first idea comes from \cite{Anandkumar11}, and is quite simple.
-All players will build an estimate $\hat{M}^j(t)$ of the number of player, that start by $\hat{M}^j(0)=1$.
-As soon as one collision is observed, a player knows that $M\geq2$, so $\hat{M}^j(t+1) = 2$.
+The first idea comes from \cite{Anandkumar11}, and the intuition behind it is quite simple, even if the mathematical derivations are not.
+All players will build an estimate $\hat{M}^j(t)$ of the number of player, that start by $\hat{M}^j(0)=1$. As soon as one collision is observed, a player knows that $M\geq2$, so $\hat{M}^j(t+1) = 2$.
 Then, based on probabilistic computations on the expected number of collisions if there were $m$ players, all following the same strategy (\eg, \RhoRand{} in the case of ), and because the formula is simple to compute for different $m$, the authors proposed a statistical test of the hypothesis $M \leq \hat{M}^j(t)$ against $M > \hat{M}^j(t)$ which is expressed as a simple comparison of the current number of collisions (since last update of $\hat{M}^j(t)$) against a threshold.
 If a player observed ``too many'' collisions (that are unlikely to be caused by only $\hat{M}^j(t) - 1$ other players), then she increases her current estimate $\hat{M}^j(t+1) = \hat{M}^j(t) + 1$.
-We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for our proposals \RandTopM{} and \MCTopM.
 
@@
 Rosenski et al. studies the performance of their algorithm with high-probability bounds \cite{Rosenski16}, and the tuning of $T_0$ they propose depends on prior knowledge on the problem.
+%
+For this reason, we are not fond of this approach, as this hypothesis is quite unrealistic if one wants to apply this kind of algorithms for real-world wireless networks.
+We note that all the following works assume some sort of prior knowledge on the difficulty of the problem:
+\cite{kumar2017channel,KumarYadav2018,SawantKumar2018,JoshiKumar2018,DarakHanawal18,KumarDarak2019,Tibrewal2019}.
+
+\paragraph{Extension of our proposals to learn the value of $M$?}
+
+Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
+We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for our proposals \RandTopM{} and \MCTopM.
 
@@
 Players can only observe $Y_k(t)$, the product of the \iid{} random sensing information from arm $k$ and of the no-collision indicator.
-This model is clearly harder than the model with sensing information, and even though our proposal \Selfish{} works fine in simulations (in terms of average regret), we proved that it can yield linear regret.
+This model seems harder than the model with sensing information, and even though our proposal \Selfish{} works fine in simulations (in terms of average regret), we proved that it can yield linear regret.
 In our paper \cite{Besson2018ALT} as well as the previous sections, it was left as a future work to know if an algorithm can achieve sub-linear regret in this harder model without sensing information.
@@
 
-Inspired by our article \cite{Besson2018ALT}, Boursier and Perchet studied this question in the summer $2018$ following its publication \cite{BoursierPerchet18}.
-Their answer is that the model without sensing information is essentially not harder than with sensing information.
+Inspired by our article \cite{Besson2018ALT},
+Boursier and Perchet studied this question in the summer $2018$ following its publication \cite{BoursierPerchet18}.
+Their answer is that the model without sensing information is essentially not harder than the model with sensing information.
+%
+Similarly, Lugosi and Mehrabian studied the same problem \cite{LugosiMehrabian18}.
+In both works, the authors detail algorithms that are unfortunately suboptimal empirically, but give in both cases a logarithmic regret if $M\leqK$ players all independently implement the proposed algorithm.
+
+\paragraph{The ``communication trick''}
+%
+\TODOL{Write about half a page on the article \cite{BoursierPerchet18}.}
+
+\paragraph{Estimating collisions through uniform exploration}
+%
+\cite{LugosiMehrabian18} gives a first algorithm with expected regret of $\bigO{M K \log(T) / \Delta^2}$, if $\Delta = \mu^*_{M} - \mu^*_{M+1} \neq 0$, achieving a better dependency regarding $M$ when compared to our result (our bound is $\bigO{M^3}$) but at the cost of (much) larger constants hidden in the $\cO$ notation, and a non-fully explicit algorithms which usually obtain bad (or worse) empirical performance.
+Then they study an interesting extension that works also if $\Delta = 0$ or if $\Delta$ is so close than the bound in $1/\Delta^2$ becomes useless for ``small horizons''.
+This behavior is well known for classical bandits, where bounds of the form $\log(T)/\Delta^2$ become worse than a linear regret $T$ if $\Delta$ is very small (\ie, $\Delta\ll \sqrt{\log(T)/T}$).
+For this extension, their proposed algorithm is proved to achieve $\bigO{K^2 M (\log(T))^2 / \mu + K M \min(\sqrt{T \log(T)}, \log(T)/\Delta')}$ regret, if $\mu$ is a lower-bound on $\mu^*_{M}$ and $\Delta' = \max(\Delta, \min\{|\mu^*_M-\mu^*_i| : \mu^*_M > \mu^*_i \})$ (that both have to be known beforehand by the algorithm).
+%
+The proposed algorithms in \cite{LugosiMehrabian18} are all based on the Musical Chair algorithm from \cite{Rosenski16}, and the curious reader should read their Algorithm~2 (page 15) for more details.
+
+The results presented in \cite{LugosiMehrabian18} are based on some hypotheses, for instance the tuning they propose for the length $g$ of the uniform exploration phase in the beginning of their ``Musical Chair''-like algorithm is based on a prior knowledge of the horizon $T$.
+In Section~4 they explain how to relax the assumptions of their results.
+For the same example, the ``doubling trick'' technique can be used to obtain a regret upper bound for an algorithm unaware of the value of $T$, within a constant multiplicative factor of the upper bound given for the algorithm aware of $T$ (Section~4.1).
+Because the regret bound is of the form $\bigO{\sqrt{T} \log(T)}$, a simple ``doubling trick'' of increasing horizons of lengths $T_i = 2^i$ works well, as proposed in \cite{CesaLugosi06} and as studied in depth in our article \cite{Besson2018DoublingTricks}.
+They also study in Section~4.3 an extension of the model for the case with more players than arms, but we do not give more details on this aspect.
+Their definition of a centralized regret for this case is an interesting and natural generalization of our definition, and they also proposed an algorithm achieving $\bigO{M K \log(T) \exp(4M/K) / \Delta^2}$ regret in this case.
+Finally, they also proposed in Section~4.4 an extension of their algorithm to estimate the number of players $M$, which is analyzed as for the Musical Chair algorithm in \cite{Rosenski16}, and also achieve logarithmic regret.
+Note that it is significantly harder to estimate $M$ without collision information, and Algorithm~3 in page 24 is quite complex.
+We did not implement it, and it is an interesting future work to run some numerical simulations in order to validate it empirically.
 
@@
 %
-They also prove in Appendix~A \cite{BoursierPerchet18} that \Selfish-\UCB{} has a linear regret, in the theoretical case, with a very neat argument from number theory (using Lindemann-Weierstrass theorem).
+It is also proved in Appendix~A of \cite{BoursierPerchet18} that \Selfish-\UCB{} has a linear regret, in the theoretical case, with a very neat argument from number theory (using Lindemann-Weierstrass theorem).
 As we conjectured, there is a gap between the theoretical result and its practical consequence, as the Theorem~4 they gave is only valid for real-valued number, and not for hardware-represented floating point number.
 Their proof is supporting what we illustrated in Figure~\ref{fig:5:oneGameTree_SelfishKLUCB}.
-Their argument is essentially to prove that for \Selfish-powered players using the simple \UCB-indexes, with a probability $p$ at time $t$ (both independent from $T$), two players might ahve the same number of pulls and the same observed rewards for each arm.
+Their argument is essentially to prove that for \Selfish-powered players using the simple \UCB-indexes, with a probability $p$ at time $t$ (both independent from $T$), two players might have the same number of pulls and the same observed rewards for each arm.
 In that case, the two players would pull the exact same arms and thus collide for a long time, until they reach a ``tie breaking point'' where they could choose different arms thanks to a random tie breaking rule (\eg, if two values of their \UCB{} indexes are the same, the $\argmax$ is a uniform random choice among the two arms).
@@
 
+In all the models of IoT networks considered in this thesis, we assume that the different IoT devices \emph{cannot communicate} with each other, and can only communicate with a unique gateway.
+Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between players are possible.
+Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
+What we are interested here is an extension of our model where players can send some bits to one or all the other players, at every time steps or some time steps.
+Clearly, the problem is easier by allowing communication, and it is quite immediate to see that the communication capacity between players must be limited otherwise the problem becomes trivial to solve.
+Indeed, imagine that at each time step $t$, all the $M$ players could share an unbounded number of bits with the other players, at no cost, even if this hypothesis is clearly unrealistic for wireless networks.
+Then they can share all their observations, and they can all run the same multi-plays MAB algorithm \cite{Anantharam87a} (in a deterministic way), like for instance the extension of Thompson sampling for multi-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
+In this setting, a logarithmic regret is obtained easily, and the regret upper-bound of the two aforementioned algorithms asymptotically achieve the lower-bound from \cite{Anantharam87a}.
+
+Thus an interesting extension is to limit the communication between players, either to a small number of bits, or just one bit, 
+This extension is studied in \cite{Avner16,AvnerMannor18}, where they essentially propose an efficient algorithm for this problem, that also reach a logarithmic regret when $M \leq K$ players are running independently the same algorithm.
+
 \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
-
-Multi-user Communication Networks: A Coordinated Multi-armed Bandit Approach
-O Avner, S Mannor - arXiv preprint arXiv:1808.04875, 2018 - arxiv.org
-\cite{AvnerMannor18}
 

### 2019-04-07 2-Chapters/5-Chapter/chapter5.tex
Okay, starting to write the result Tables for the two extensions

@@
 We did not propose any other efficient algorithm, but our work presented in April~$2018$ inspired some other articles \cite{LugosiMehrabian18,BoursierPerchet18}.
-Two articles confirmed our finding that \Selfish{} has a linear regret, and they all proposed new algorithms.
-We do not include numerical simulations to compare all of them, but we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
+Two articles confirmed our finding that \Selfish{} can have a linear regret, and they all proposed new algorithms.
+We not include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
 
-Finally, another interesting extension of the studied models is to consider different channels utility for each devices.
+Another interesting extension of the studied model is to consider different channels utility for each devices.
 It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
@@
 
-\TODOL{Get bibtex of the articles by Sumit J. Darak, and talk correctly about them!}
+Finally, we review other extensions of our model, and give a overview of the current state-of-the-art as exhaustive as possible.
 
@@
 We wanted to highlight that the community was quite active on research on multi-player bandits in the last $10$ years, but especially active since last spring $2018$.
-For instance, our article \cite{Besson2018ALT} was the first to propose the ``no sensing'' case, and it was studied by (at least) three independent group of researchers since its publication in April $2018$ \cite{BoursierPerchet18,LugosiMehrabian18,AvnerMannor18}.
+For instance, our article \cite{Besson2018ALT} was the first to propose the ``no sensing'' case, and it was studied by (at least) two independent group of researchers since its publication in April $2018$ \cite{BoursierPerchet18,LugosiMehrabian18}.
 Another example is the model with different arm means among players, who was studied in \cite{Anandkumar10,Kalathil12} and then more recently in \cite{Bistritz18,KaufmannAbbas19}.
@@
 
+\paragraph{Simulations.}
+
+We consider a bandit problem with $K=9$ arms, of means $0.1,0.2,\dots,0.8,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
+As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require the know $M$.
+Then we compare the \RhoRand, \RandTopM{} and \MCTopM{} algorithms, using \klUCB, which know $M$ beforehand, with their extensions implementing the same algorithm as \rhoRandEst{} to estimate $M$ on the fly.
+
 \TODOL{Include a simulation to show the behavior of \rhoRand{} vs \rhoRandEst, and \RandTopM{} and \MCTopM{} with prior knowledge of $M$ vs the ``estimate'' versions.}
+
+\begin{table}[ht]
+    % \begin{footnotesize}
+        \centering
+        \begin{tabular}{c|ccc}
+        \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Number of players} & $M=3$ & $M=6$ & $M=9$ \\
+            \hline
+            Centralized multiple-play \klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+            \RhoRand-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \RhoRand-\klUCB{} + Estimate $M$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+            \RandTopM-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \RandTopM-\klUCB{} + Estimate $M$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+            \MCTopM-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \MCTopM-\klUCB{} + Estimate $M$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+            \Selfish-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+        \end{tabular}
+        \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, comparing algorithms which knows $M$ against algorithms which estimate $M$ on the fly.}
+        \label{table:5:meanRegretSimulationsEstimatingM}
+    % \end{footnotesize}
+    \end{table}
+
+\TODOL{Give interpretation from these results!}
 
@@
 In our paper \cite{Besson2018ALT} as well as the previous sections, it was left as a future work to know if an algorithm can achieve sub-linear regret in this harder model without sensing information.
-
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 
@@
 
-\TODOL{Do some basic simulations and show !}
+\TODOL{Do some basic simulations and show :
+
+    - Selfish-UCB from \cite{Besson2018ALT}.
+    - Improved Musical Chair from \cite{LugosiMehrabian18}.
+    - SIC-MMAB from \cite{BoursierPerchet18}.
+}
+
+As above, we consider a bandit problem with $K=9$ arms, of means $0.1,0.2,\dots,0.8,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
+We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
+
+\begin{table}[ht]
+    % \begin{footnotesize}
+        \centering
+        \begin{tabular}{c|ccc}
+        \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Number of players} & $M=3$ & $M=6$ & $M=9$ \\
+            \hline
+            \Selfish-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+            Improved \MusicalChair{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+            \textsc{Sic-MMAB} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \hline
+        \end{tabular}
+        \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, for the ``no sensing'' case.}
+        \label{table:5:meanRegretSimulationsNoSensing}
+    % \end{footnotesize}
+    \end{table}
+
+\TODOL{Give interpretation from these results!}
 
@@
 
-In all the models of IoT networks considered in this thesis, we assume that the different IoT devices \emph{cannot communicate} with each other, and can only communicate with a unique gateway.
+In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
 Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between players are possible.
@@
 
-Thus an interesting extension is to limit the communication between players, either to a small number of bits, or just one bit, 
-This extension is studied in \cite{Avner16,AvnerMannor18}, where they essentially propose an efficient algorithm for this problem, that also reach a logarithmic regret when $M \leq K$ players are running independently the same algorithm.
+An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit.
+We do not review papers that study this model, because the recent work \cite{BoursierPerchet18} introduced the ``communication trick'', that essentially allow any player to send one bit to all the others, at some pre-agreed time steps, with no modification on the model.
+Thanks to this ``communication trick'', recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
+This trick essentially use the fact that the players share a synchronized time, thus they can use a collision as a way to directly exchange one bit of information between two players.
+This process is slow and not efficient, as all but two players must do nothing when player $i$ is sending a bit to player $j$, but it does build a communication protocol in the model without explicit communication between players.
+This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
+We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.
+
+Another line of research is to consider models that are closer to realistic wireless communication networks, as it is done in \cite{Avner16,AvnerMannor18}.
+Explaining in details their model and proposed solutions would be quite lengthy, and we rather prefer to let the interested reader refer to the later work \cite{AvnerMannor18}.
+% where they essentially propose an efficient algorithm for this problem, that also reach a logarithmic regret when $M \leq K$ players are running independently the same algorithm.
+% FIXME don't copy/paste their abstract...
+We sum-up its contributions quickly.
+They we address several aspects of the challenge of communication networks shared by many users simultaneously: learning unknown stochastic network characteristics, sharing resources with other users while keeping coordination overhead to a minimum.
+The proposed solution combines Multi-Armed Bandit learning with a lightweight signalling-based coordination scheme, and ensures convergence to a stable allocation of resources.
+They work considers single-user level algorithms for two scenarios:an unknown fixed number of users, and a dynamic number of users.
+Analytic performance guarantees, proving convergence to stable marriage configurations, are presented for both setups.
+The algorithms are designed based on a system-wide perspective, rather than focusing on single user welfare.
+Thus, maximal re-source utilization is ensured.
+
+% ----------------------------------------------------------------------------
+\subsection{Arrival and departures of players: the ``dynamic setting''}
+\label{sub:5:arrivalDepartures}
+
+As reminded above, our model assumes the number of players $M$ remain fixed during all the bandit game.
+However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
+The existing previous work on multi-player bandit models are all motivated by possible applications to wireless networks, but most of them assume $M$ to be fixed.
+The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
+If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the rewards (\ie, sensing) as well as in the players' decisions (\ie, collisions):
+\begin{equation}\label{eq:5:defRegretDynamicSetting}
+    R_T^{\text{dyn}} = \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
+\end{equation}
+
+In \cite{Rosenski16}, the authors explain that if the model allows arrival or departures of players at \emph{any time step}, then the game is much harder, and sub-linear regret is most likely un-achievable, if we consider the natural extension of the definition of regret as in our model presented in Section~\ref{sec:5:model} above.
+However, this negative results depends on how arrival of players are modeled:
+if an arriving player has no prior memory of the previous observations, \ie, if it model a new device, then if there is no restriction on the number or frequency of arrival or departures, we can most likely prove that sub-linear regret is not achievable in general.
+A simple but extreme example shows that regret has to be linear for any bandit strategy (in the observation model with sensing).
+Consider $K=2$ Bernoulli-distributed arms of means $\mu_1=\mu_2=1/2$, and one player is always active and play any bandit strategy (\eg, \MCTopM-\klUCB).
+Every time step, another player is either arriving, without any knowledge of the problem (\eg, it is a fresh and new IoT device).
+Then no matter the strategy of these news players, even if they all play \MCTop-\klUCB{} for instance, if they player a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
+
+In \cite{Rosenski16}, it is assumed that the players all know a lower-bound $L$ on the length of all the intervals during which arrival or departures of players are allowed.
+A naive idea, if $L$ is large enough, is simply to restart the underlying algorithm every $L$ time steps, in order to directly benefit from the theoretical guarantees of the ``static setting'' (where $M$ stays constant).
+However, this idea yields a large regret if the length of ``static'' intervals $L$ is too small.
+Applied to the \MusicalChair{} algorithm, the authors in \cite{Rosenski16} call this basic extension Dynamic \MusicalChair
+Under the simple hypothesis that the overall number of players entering and leaving the game is sub-linear in $T$ (\ie, a $\smallO{T}$), they analyze the regret of Dynamic \MusicalChair{} and prove it is also sub-linear (see Theorem~2 in Section~3.4).
+%
+Under the same hypothesis, and if the lower-bound $L$ is known before-hand and is large enough, then we could also apply the same idea as from \cite{Rosenski16} to our approach.
+We believe that we could easily prove that a sub-linear regret is achievable, if all players run \MCTopM-\klUCB{} (and if currently active players restart their memory of the past observations every $L$ time steps).
+As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, using the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to our proposal \MCTopM{} would also give a small regret upper-bound.
+
+After being introduced in \cite{Rosenski16}, some more recent works studied the case of arrival or departures of players, usually referred to as ``dynamic setting'' or ``dynamic case''.
+For instance, \cite{BoursierPerchet18} studies in Section~4 the first algorithm proposed for the dynamic case under the no-sensing model.
+They study the same notion of regret as the one proposed in \cite{Rosenski16} and given above in \eqref{eq:5:defRegretDynamicSetting}, see Equation $(10)$ in Section~4.2.1.
+With the assumption of a non-decreasing number of players, \ie, if only arrivals of players are considered,
+they prove a regret upper-bound of the \textsc{Dyn-MMAB} algorithm in Theorem~3 .
+% Denote $M(t)$ the number of players at each time $t$, then they propose to study $R_T$ defined as
+% \[ R_T = \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu_^*{k} - \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t). \]
+If all arriving players are using their \textsc{Dyn-MMAB} algorithm, the ``dynamic'' regret is bounded by $\bigO{M^2 K \log(T) / \mu^*_{M}}$, if $M=M(T)$ is the total number of players involved in the problem.
 
@@
 
-% ----------------------------------------------------------------------------
-\subsection{Arrival and departures of players}
-\label{sub:5:arrivalDepartures}
-
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+Note that we have not yet implemented this extension in SMPyBandits \cite{SMPyBandits}, but we are interested to do it, and it is one of the major future work left on our library\footnote{See the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/124}}}.
 
@@
 \cite{KaufmannAbbas19}
+
+Note that we have not yet implemented this extension in SMPyBandits \cite{SMPyBandits}, but we are interested to do it, and it is one of the major future work left on our library\footnote{See the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/185}}}.
 
@@
 \cite{SawantKumar2018}
+
+Note that we have not yet implemented this extension in SMPyBandits \cite{SMPyBandits}, but we are interested to do it, and it is one of the major future work left on our library\footnote{See the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/120}}}.
 

### 2019-04-07 2-Chapters/5-Chapter/chapter5.tex
End of work for Sunday afternoon. So boring.

@@
 %
-% \TODOL{Explain that it does make less sense to know $M$}
 One could criticize our approach if we study a wireless networks with no central coordination from the gateway,
@@
 If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm running on top of a round-Robin phase can achieve order-optimal regret.
-Without giving more details, we let the interested refer to what is is explained for the RR-SW-UCB\# algorithm in \cite{WeiSrivastava18Distributed} and \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
+Without giving more details, we let the interested refer to what is is explained for the RR-SW-UCB\# algorithm in \cite{WeiSrivastava18Distributed} and the algorithms presented in \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
 
-\paragraph{Literature review}
+\paragraph{Two ideas to estimate $M$.}
 
@@
 
-\paragraph{Extension of our proposals to learn the value of $M$?}
+\paragraph{Extension of our proposals to learn the value of $M$.}
 
@@
 We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for our proposals \RandTopM{} and \MCTopM.
+Building from the theoretical analysis given for \rhoRandEst{} in \cite{Anandkumar11}, and our analysis of \MCTopM-\klUCB, we believe it is possible to show that the aforementioned extension of our proposal also achieves sub-linear regret without requiring players to know $M$ in the beginning of the bandit game.
+More precisely, we believe that \MCTopM-\klUCB{} can still give order-optimal logarithmic regret if $M$ players use it, and this extension of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is not included due to space constraints.
+Writing its proof and performing more simulations are left as possible future work.
 
@@
 Similarly, Lugosi and Mehrabian studied the same problem \cite{LugosiMehrabian18}.
-In both works, the authors detail algorithms that are unfortunately suboptimal empirically, but give in both cases a logarithmic regret if $M\leqK$ players all independently implement the proposed algorithm.
+In both works, the authors detail algorithms that are unfortunately suboptimal empirically, but give in both cases a logarithmic regret if $M \leq K$ players all independently implement the proposed algorithm.
 
@@
 % where they essentially propose an efficient algorithm for this problem, that also reach a logarithmic regret when $M \leq K$ players are running independently the same algorithm.
-% FIXME don't copy/paste their abstract...
+\TODOL{FIXME don't copy/paste their abstract...}
 We sum-up its contributions quickly.
@@
 Thus, maximal re-source utilization is ensured.
+
+\TODOL{A conclusion from this sub-section presenting this model?}
 
@@
 Every time step, another player is either arriving, without any knowledge of the problem (\eg, it is a fresh and new IoT device).
-Then no matter the strategy of these news players, even if they all play \MCTop-\klUCB{} for instance, if they player a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
+Then no matter the strategy of these news players, even if they all play \MCTopM-\klUCB{} for instance, if they player a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
 

### 2019-04-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Started to write the « related works for other models » section in Chapter:5
J'avance. En train de sauver notre borne inférieure.

@@
 % -----------------------------------------------------------------
-\section{Motivation for multi-player MAB models}
+\section{Motivations for multi-player MAB models}
 \label{sec:5:introduction}
@@
 
-For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
+For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1,chapter:4}.
 With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
@@
 Due to the communication cost implied by a central controller, a more relevant model is the
-\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
+\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and further studied shortly after in \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
 Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain)
@@
 The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
-A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all arms to estimate their means while trying to focus on the best arms to gain as much rewards as possible.
-The decentralized setting considers no exchange of information between players, that only know $K$ and $M$, and to avoid collisions, players should furthermore find orthogonal configurations (\ie, the $M$ players use the $M$ best arms without any collision), without communicating.
+A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all the arms to estimate their means, while trying to focus on the best arms to gain as much rewards as possible.
+The decentralized setting considers wireless protocol with no direct or explicit exchange of information between players, and assumes that the players only know $K$ and $M$.
+To avoid collisions, players should furthermore find orthogonal configurations, \ie, the $M$ players use the $M$ best arms without any collision, without explicitly communicating\footnote{One must now be careful about this aspect when stating that ``no explicit communications are allowed between players'' in the model, as \cite{BoursierPerchet18} proved that even in the no sensing case, the ``communication trick'' can be used to exchange information between players, by generating collisions at some pre-agreed times.
+We discuss this work more in details at the end of this chapter, in Section~\ref{sub:5:withoutSensing}.} with each other.
 Hence, in that case the trade-off is to be found between exploration, exploitation \emph{and} low collisions.
@@
 All these above-mentioned works are motivated by the OSA problem, in which it is assumed that \emph{sensing} occurs, that is each smart device observes the availability of a channel (sample from the arm) \emph{before} trying to transmit and possibly experiment a collision with other smart devices.
-However some real radio networks do not use sensing at all, \eg, emerging standards developed for \emph{Internet of Things} (IoT) networks such as LoRaWAN.
+However some real radio networks do not use sensing at all, \eg, emerging standards currently or recently developed for \emph{Internet of Things} (IoT) networks, such as LoRaWAN.
 Thus, to take into account these new applications, algorithms with additional constraints on the available feedback have to be proposed within the multiple-player MAB model.
 Especially, the typical approach that combines a (single-player) bandit algorithm based on the sensing information --to learn the quality of the channels while targeting the best ones-- with a low-complexity decentralized collision avoidance protocol, is no longer possible.
+Our article \cite{Besson2018ALT} was the first to study this other model of multi-player bandits for the ``no sensing'' case.
 
@@
 In the absence of sensing information, we propose the \Selfish{} heuristic and investigate its performance.
-Our study of this algorithm is supported by (promising) empirical performance and some first (disappointing) theoretical elements.
+Our study of this algorithm was supported in \cite{Besson2018ALT} by (promising) empirical performance and some first (disappointing) theoretical elements, and it was proved to be inefficient since then.
 
@@
 % -----------------------------------------------------------------
-\section{Multi-player bandit model with different feedback levels}
+\section{Three feedback levels for the multi-player bandit model}
 \label{sec:5:model}
@@
 % describe our stochastic assumptions
-We consider a $K$-armed Bernoulli bandit model, % (for $K\geq2$),
+Similarly to what is presented in Chapter~\ref{chapter:2},
+we consider a $K$-armed Bernoulli bandit model, % (for $K\geq2$),
 in which arm $k$ is a Bernoulli distribution with mean $\mu_k\in[0,1]$.
 We denote $(Y_{k,t})_{t\in\N}$ the \iid{} (binary) \emph{reward stream} for arm $k$, that satisfies $\Pr(Y_{k,t}=1) = \mu_k$ and that is independent from the other rewards streams.
+
+\textbf{Only Bernoulli?}
 However we mention that our lower bound and all our algorithms (and their analysis) can be easily extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}).
@@
 Note that other models for rewards loss have also been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it.
-To stay consistent with the model presented in the previous Chapter~\ref{chapter:4},
+To stay consistent and for simplicity, with the model presented in the previous Chapter~\ref{chapter:4},
 we preferred to focus on full reward occlusion in this chapter.
 
-A multi-player MAB strategy is a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
+A multi-player MAB strategy is formally defined as a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
 First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
@@
   \item[\modelun]
-    \textbf{Simultaneous sensing and collision}: player $j$ observes  $Y_{A^j(t),t}$ \emph{and} $C^j(t)$ (not previously studied).
+    \textbf{Simultaneous sensing and collision}: player $j$ observes  $Y_{A^j(t),t}$ \emph{and} $C^j(t)$.
+    We note that this first model was never previously studied, but we do not focus on it because of its unrealistic aspect, and its simplicity (it is easier than the following model, for which we obtain good theoretical results).
   \item[\modeldeux]
@@
     However, the sensing and collision information are censored.
-    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications.
+    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications, but as explained above we did not present theoretical results for this third model (yet).
     %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
@@
 \begin{framed}
-  \paragraph{\textcolor{red}{Warning}.}
+  \textbf{\textcolor{red}{Warning}.}
   The regret lower bound given in \cite{Besson2018ALT} is not applicable to any algorithm,
@@
 We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
+Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
+This class contains our proposals, \RandTopM{} and \MCTopM, but rule out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
 
@@
 
-We now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections under a \emph{decentralized strategy} that has access to the sensing information.
+As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the collision information term is zero for any algorithm $\rho$.
+However, we wanted to be still able to present our derivations and results giving the desired lower-bound, mainly because its proof and its form give insights on how to design an asymptotically optimal algorithm $\rho$, and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
+The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
+% FIXME
+% We let the interested reader fetch the details on their paper.
+Fix a player $j\in\{1,\dots,M\}$, and we consider the observations $\cO_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
+\[ \cO_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
+where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).
+Fix a problem $\bm{\mu}$, and as we do in Appendix~\ref{proof:5:BetterLowerBound}, introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
+
+For the algorithm $\rho$, we introduce the \emph{collision information term} as:
+\begin{equation}\label{eq:5:collisionInformationTerm}
+  \cI_{\bm{\mu},\bm{\lambda}}(\rho,T) := \sum_{t=1}^{T} \mathrm{KL}(\P_{\bm{\mu}}^{C_t | \cO_{t-1}}, \P_{\bm{\lambda}}^{C_t | \cO_{t-1}}).
+\end{equation}
+
+We now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections under a \emph{decentralized strategy} $\rho$ that has access to the sensing information and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T) = \smallO{\log(T)}$.
+This additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
 This result, proved in Appendix~\ref{proof:5:BetterLowerBound}, yields an asymptotic logarithmic lower bound on the regret, also given in Theorem~\ref{thm:5:BetterLowerBound}.
@@
 \begin{theorem}\label{thm:5:BetterLowerBound}
-  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\rho$ and $\boldsymbol{\mu}\in\cP_M$,
+  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\rho$ and $\boldsymbol{\mu}\in\cP_M$.
+  Furthermore, if $\rho$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$, then
   % \vspace*{-10pt}  % XXX

### 2019-04-08 2-Chapters/5-Chapter/chapter5.tex
J'avance. En train de sauver notre borne inférieure.

@@
 Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
-Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices, Even if there is many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step even the number of communicating devices is random and varies a lot.
+Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices.
+Even if there are many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step the number of communicating devices is random and unpredictable.
 
 Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider the easier case of at most $M \leq K$ devices, communicating at each time step.
-Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward at each time step.
+Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
 %
@@
 Like in previous works, we consider the centralized system regret (multi-player regret), or simply referred to as regret.
-We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions..
-We then analyze our \MCTopM{} algorithm and we show that its regret upper-bound asymptotically matches the improved lower-bound, with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
+We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
+We then analyze our \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
+%  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
 We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M$ accessing $K \geq M$ channels.
 
-We are also interested to consider the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
+We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
 Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
-Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we find on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear regret and fails to solve the no sensing multi-player MAB problem).
+Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-player MAB problems).
 We did not propose any other efficient algorithm, but our work presented in April~$2018$ inspired some other articles \cite{LugosiMehrabian18,BoursierPerchet18}.
-Two articles confirmed our finding that \Selfish{} can have a linear regret, and they all proposed new algorithms.
-We not include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
+These two articles confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
+We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
 

### 2019-04-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
J'avance, meilleure présentation de notre lower-bound et upper-bound

@@
 % -----------------------------------------------------------------
-\section{An improved asymptotic regret lower bound for a certain class of algorithms}
+\section{An improved asymptotic regret lower bound (for some algorithms)}
 \label{sec:5:lowerbound}
 % -----------------------------------------------------------------
-
-\TODOL{Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
 
@@
 For centralized algorithms, due to the absence of collisions, \ref{eq:5:term3} is obviously zero, and \ref{eq:5:term2} is non-negative, as $T_k(T) \leq T$.
-For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $T_k(T) > T$.
+For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $T_k(T) > T$ (which is counter intuitive with such notations).
 However, a careful manipulation of this decomposition shows that the regret is always lower bounded by term \ref{eq:5:term1}.
@@
 
-To express our lower bound, we need to introduce $\kl(x,y)$ as the Kullback-Leibler divergence between the Bernoulli distribution of mean $x$ and that of mean $y$, so that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$.
+Similarly to what we present above in Chapter~\ref{chapter:2}, Section~\ref{sec:2:lowerUpperBoundsRegret},
+we use the Kullback-Leibler divergence to express our lower bound.
+Let $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ be KL divergence between the Bernoulli distribution of mean $x$ and that of mean $y$.
 %
@@
 
-As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the collision information term is zero for any algorithm $\rho$.
-However, we wanted to be still able to present our derivations and results giving the desired lower-bound, mainly because its proof and its form give insights on how to design an asymptotically optimal algorithm $\rho$, and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
+As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the ``collision information term'' is zero for any algorithm $\rho$.
+However, we wanted to be still able to present the derivations of our lower-bound, mainly because its proof and its form give insights on how to design an asymptotically optimal algorithm $\rho$, and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
+
+% FIXME
 The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
-% FIXME
-% We let the interested reader fetch the details on their paper.
 Fix a player $j\in\{1,\dots,M\}$, and we consider the observations $\cO_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
@@
 where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).
+
 Fix a problem $\bm{\mu}$, and as we do in Appendix~\ref{proof:5:BetterLowerBound}, introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
-
-For the algorithm $\rho$, we introduce the \emph{collision information term} as:
+Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
+We denote $\Pr_{\bm{\mu}}^{\cO_t}$ the distribution of the vector $\cO_t$ under the model $\bm{\mu}$ when using algorithm $\rho$.
+Then we introduce the \emph{collision information term} as:
 \begin{equation}\label{eq:5:collisionInformationTerm}
-  \cI_{\bm{\mu},\bm{\lambda}}(\rho,T) := \sum_{t=1}^{T} \mathrm{KL}(\P_{\bm{\mu}}^{C_t | \cO_{t-1}}, \P_{\bm{\lambda}}^{C_t | \cO_{t-1}}).
-\end{equation}
-
-We now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections under a \emph{decentralized strategy} $\rho$ that has access to the sensing information and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T) = \smallO{\log(T)}$.
-This additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
-This result, proved in Appendix~\ref{proof:5:BetterLowerBound}, yields an asymptotic logarithmic lower bound on the regret, also given in Theorem~\ref{thm:5:BetterLowerBound}.
+  \cI_{\bm{\mu},\bm{\lambda}}(\rho,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}).
+\end{equation}
+
+For a \emph{decentralized strategy} $\rho$ that has access to the sensing information (\ie, ruling out model \modeltrois), and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T) = \smallO{\log(T)}$,
+we now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections.
+The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
+The theorem stated below is proved in Appendix~\ref{proof:5:BetterLowerBound}, and it also yields an asymptotic logarithmic lower bound on the regret.
 
@@
 for every $M \in \{2,\dots,K\}$ and $\boldsymbol{\mu} \in \cP_M$.
-
+%
 Since our article \cite{Besson2018ALT}, we kept as a future work the following question.
 A lower bound on the minimal number of collisions experienced by any strongly uniformly efficient decentralized algorithm would thus be a nice complement to our Theorem~\ref{thm:5:BetterLowerBound}.
-
-\TODOL{Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
 
@@
 
+\paragraph{Index policies.}
+%
 The approaches we now describe for multi-player bandits can be used in combination with any index policy, but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
@@
 
-% RhoRand
-
+\paragraph{Two ideas for orthogonalization?}
+%
 On the first hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
@@
 % Now, our algorithms !!
-
+\paragraph{Our two proposals.}
 We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
@@
 \end{equation}
+
+\paragraph{The \RandTopM{} algorithm.}
+%
+First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
+It is essentially a refinement over the \RhoRand{} algorithm, and a simpler version of \MCTopM.
+The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
+%
+In \RandTopM{}, player $j$ is always considered ``not fixed'',
+and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.
 
@@
 % \vspace*{-5pt}  % XXX remove if problem
-
-\paragraph{The \RandTopM{} algorithm.}
-%
-First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
-It is essentially a refinement over the \RhoRand{} algorithm, and a simpler version of \MCTopM.
-The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
-%
-In \RandTopM{}, player $j$ is always considered ``not fixed'',
-and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.
 
@@
 % -----------------------------------------------------------------
-\subsection{The \Selfish{} approach}
+\subsection{The \Selfish{} approach: a heuristic for the ``no sensing case''}
 \label{sub:5:Selfish}
@@
   For any $\boldsymbol{\mu}\in\cP_M$,
-  if player $j\in\{1,\dots,M\}$ uses the \RandTopM-, \MCTopM- or \rhoRand-\klUCB{}
-  decentralized policy with exploration function $f(t) = \log(t) + 3 \log\log(t)$,
-  then for any sub-optimal arm $k \in \Mworst$, there exists two problem depend constants $C_{\boldsymbol{\mu}}$, $D_{\boldsymbol{\mu}}$ such that
+  let player $j\in\{1,\dots,M\}$ use the \RandTopM-, \MCTopM- or \rhoRand-\klUCB{}
+  decentralized policy with exploration function $f(t) = \log(t) + 3 \log\log(t)$.
+  Then for any sub-optimal arm $k \in \Mworst$, there exists two problem-dependent positive constants $C_{\boldsymbol{\mu}}, D_{\boldsymbol{\mu}} > 0$ such that
   % %
@@
   \begin{equation}\label{eq:5:SubOptimalSelections}
-      \E_{\mu}[T_k^j(T)] \leq \frac{\log(T)}{\kl(\mu_k,\mu_{M}^*)}+ C_{\boldsymbol{\mu}} \sqrt{\log(T)} + D_{\boldsymbol{\mu}}\log\log(T) + 3M + 1.
+      \E_{\mu}[T_k^j(T)] \leq
+      \frac{\log(T)}{\kl(\mu_k,\mu_{M}^*)}
+      + \underbrace{C_{\boldsymbol{\mu}} \sqrt{\log(T)} + D_{\boldsymbol{\mu}}\log\log(T) + 3M + 1}_{= \smallO{\log(T)}}.
   \end{equation}
@@
 
-  Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
+  Thanks to \eqref{eq:5:KeyInclusion} and if $\Pr_{\bm{\mu}} $ denote here the probability under model $\bm{\mu}$, it is easy to decompose the number of selections of arm $k$ by user $j$ up to round $T$ as
   \begin{align*}
@@
   &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
-  = \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
-  %&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
-  &= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
+  = \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k \right).\\
+  %&= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
+  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
   %
@@
   Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
-
   \begin{align*}
-    \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
-    &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
-    &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
-    %
-    & \leq \sum_{t=1}^T \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
-    + \sum_{m=1}^M \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right)
-    \hspace*{100pt} (\heartsuit)
+    \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
+    &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
+    &\leq \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(g_{m^*}(t) < \mu_m^* \right) \\
+    %
+    & \leq \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+    + \sum_{m=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( g_{m^*}(t) < \mu_m^* \right)
+    \hspace*{80pt} (\heartsuit)  % TODO cleaner?
   \end{align*}
@@
       \begin{equation}
-        \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
+        \sum_{t=1}^T \Pr_{\bm{\mu}} \left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
       \end{equation}
@@
   \begin{align*}
-    &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
+    &\sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
     + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
@@
   \begin{align*}
-    & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
-    & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
-    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
+    & \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
+    & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
+    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
   \end{align*}
@@
   Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
-  \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
+  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
   The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
@@
   \begin{align}
-    \sum_{t=1}^T & \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
+    \sum_{t=1}^T & \Pr\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
     & =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
     & \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
-    & \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
+    & \leq \sum_{s=1}^T \Pr\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
   \end{align}
@@
   If $\mu_k > \mu_{k'}$, we rather use that
-  \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
+  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \Pr\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
   and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields

### 2019-04-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Figure de l'annexe dans le corps du chapitre 5, pour l'instant le placement est chelou...

@@
   \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
+  \begin{small} % XXX
   \begin{align*}
@@
   \end{lemma}
+  \end{small} % XXX
 
@@
     = M \left(\sum_{j=1}^M \E_{\mu}[N_2^j(T)] + \E_{\mu}[N_3^j(T)]\right) \\
-    &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag  \\
+    &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag
+  \end{align}%
+  \begin{align}
     &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
@@
 
-The analysis of \Selfish{} is harder, but we tried our best to obtain some understanding of the behavior of this algorithm, that seems to be doing surprisingly well in many contexts, as in our experiments with $K=9$ arms and in extensive experiments not reported in this section.
+The analysis of \Selfish{} is harder, but we obtained some understanding of the behavior of this algorithm, that seems to be doing surprisingly well in many contexts, as in our experiments with $K=9$ arms and in extensive experiments not reported in this section.
 However, a disappointing result is that we found simple problems, usually with small number of arms, for which the algorithm may fail.
@@
 This result would then prevent \Selfish{} from having a logarithmic regret.
-However, it is to be noted that the lower bound of Theorem~\ref{thm:5:BetterLowerBound} does not apply to the censored observation model \modeltrois{} under which \Selfish{} operates, and it is not known yet whether logarithmic regret is at all possible.
+However, it is to be noted that the lower bound of Theorem~\ref{thm:5:BetterLowerBound} does not apply to the censored observation model \modeltrois{} under which \Selfish{} operates.
+%
+The \textsc{Sic-MMAB} algorithm proposed in \cite{BoursierPerchet18} answers the question we left open last year in \cite{Besson2018ALT},
+of whether logarithmic regret is at all possible for the ``no sensing'' case (model \modeltrois).
 
@@
 We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \klUCB{} indices.
-Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.
-
-\TODOL{Rewrite appendix to include all simulations here!}
+% Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.
 
@@
 We also include a centralized multiple-play \klUCB{} algorithm,
-%as defined by \cite{Anantharam87a},
+as defined by \cite{Anantharam87a},
 essentially to check that the \emph{``price of decentralized learning''} is not too large.
@@
 
-We present results for two bandit instance: one with $K=3$ arms and means %\footnote{But of course the means are unknown to the algorithms, and their order is not important.}
+We present results for two bandit instance: one with $K=3$ arms and means
 $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}.
@@
 and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
-% Note the different algorithms know the number of players $M$.
+Note that here, we consider different algorithms that know the number of players $M$.
 Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
-%, from $t=1$ to the horizon $T=10000$ (also unknown by the algorithms),
+%
 We also include histograms showing the \emph{distribution} of regret at $t=T$,
@@
 For the plots showing the regret, our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} is displayed.
-% and it should be surprising to see it being larger than some regret values: it is only asymptotic.
 
@@
 and \Selfish{} outperforms \rhoRand{} in most cases.
-%Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} is a good example
-%of outstanding performances of \MCTopM{} and \Selfish{} in comparison to \rhoRand{}.
+Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} is a good example
+of outstanding performances of \MCTopM{} and \Selfish{} in comparison to \rhoRand{}.
+%
 Empirically, our proposals were found to almost always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
@@
 
-% -----------------------------------------------------------------
-\section{Conclusion and future work}
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{0.85\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized____env1-1_7803645526012310577.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{0.85\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized_semilogy____env1-1_7803645526012310577.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem]{Regret (in log-$y$ scale on the right) for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
+With a perfect knowledge on the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly in this simple setting, even with the knowledge of the horizon $T$ for \MusicalChair{}.
+The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can bounded with probability $0.5$ and $0.9$ respectively.
+The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.}
+  \label{fig:5:MP__K9_M3_T123456_N100__8_algos}
+\end{figure}
+
+%
+% Regular plots of centralized regrets
+%
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{0.49\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_5016720151160452442.pdf}
+  % \end{subfigure}
+  % % ~
+  % \begin{subfigure}[!h]{0.49\textwidth}
+  %   \includegraphics[width=1.10\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
+  % \end{subfigure}
+  \caption[Failure case of \Selfish]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
+Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$).
+The regret for the three other algorithms is very small for this problem, always smaller than $100$ here.}
+  \label{fig:5:selfish_fail1}
+  % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
+%
+% Regular plots of centralized regrets
+%
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{0.49\textwidth}
+      \includegraphics[width=1.05\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_2643560344649862285.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{0.49\textwidth}
+      \includegraphics[width=1.05\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_4683123079851881812.pdf}
+  % \end{subfigure}
+  \caption[Second failure case of \Selfish]{Regret, $M=2$ and $M=3$ players, $K=3$ arms, horizon $T=5000$, against $1000$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+\Selfish{} (top curve in \textcolor{darkgreen}{green}) clearly fails in such setting with small $K$.}
+  \label{fig:5:selfish_fail2}
+  % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
+%
+% Regular plots of centralized regrets
+%
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      % \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
+      \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_1035303196230283176.pdf}
+      % \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_5016720151160452442.pdf}
+  % \end{subfigure}
+  ~
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
+  % \end{subfigure}
+  \caption[Third failure case of \Selfish]{Regret for $M=3$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
+Axis $x$ is for regret (different scale for each), and the top \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($11$ cases of $R_T \geq T$, out of $1000$).
+The regret for the three other algorithms is very small for this problem, and even appears constant.}
+  \label{fig:5:selfish_fail3}
+  % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
+%
+% Regular plots of centralized regrets
+%
+
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      % \includegraphics[width=1.00\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3251433209347345969.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+\rhoRand{} (top \textcolor{blue}{blue}) is outperformed by the other algorithms (and the gain increases when $M$ increases), which all perform similarly in such configurations.
+Note that the (small) tail of the histograms come from complicated problems $\boldsymbol{\mu}$ and not failure cases.}
+  \label{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
+  % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_RegretCentralized____env1-1_3892966382091165662.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3892966382091165662.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+This extreme case $M=K$ shows the drastic difference of behavior between \RandTopM{} and \MCTopM, having constant regret, and \rhoRand{} and \Selfish, having large regret.}
+  \label{fig:5:MP__K9_M9_T5000_N500__4_algos__all_HistogramsRegret}
+  % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M2_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_2643359116089264295.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{1.00\textwidth}
+  %   \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M9_T10000_N200__4_algos/all_RegretCentralized_loglog____env1-1_2306423191427933958.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$,Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand.
+\MCTopM{} is always among the best algorithms, and for $M$ not too small, its regret seems logarithmic with a constant matching the lower bound.}
+  \label{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  % \begin{subfigure}[!h]{0.75\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{0.75\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{0.75\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$]{Regret (in log-log scale), for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$.
+The same observation as in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} can be made.
+\Selfish{} outperforms \MCTopM{} for $M=2$ here.
+Additionally, \MCTopM{} is the only algorithm to not fail dramatically when $M=K$ here.}
+  \label{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}
+\end{figure}
+
+% -----------------------------------------------------------------
+\section{Conclusion of our contributions}
 \label{sec:5:conclusion}
@@
 
-% biblio usually before the supplementary material
-% \bibliography{mpBandits}
+\begin{framed}
+  \textcolor{red}{\textbf{FIXME}}
+  Je ne sais pas trop comment inclure toutes ces figures ici, avant la dernière section où je parle des extensions de nos modèles.
+  J'aimerai bien avoir quelques pages simplement remplies des figures, sans avoir le texte de Section~\ref{sec:5:literatureReviewOtherModels} qui commence au milieu...
+  Une idée ?
+\end{framed}

### 2019-04-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
End of work for Sunday afternoon. So boring.
Figure de l'annexe dans le corps du chapitre 5, pour l'instant le placement est chelou...

@@
     \centering
-    \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex}
-    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing.gv.pdf}
-    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing__leafs.gv.pdf}
-    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing.gv.pdf}
-    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs.gv.pdf}
+    % \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex}
+    \includegraphics[width=0.70\textwidth]{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing.pdf}
+    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing__leafs.pdf}
+    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing.pdf}
+    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs.pdf}
     \caption[Trajectories of $M=2$ players using \Selfish-\UCB{} for $K=2$ arms.]{For $K=2$ arms and $M=2$ players using \Selfish-\UCB, for depth$=3$: $2$ \textbf{\textcolor{red}{absorbing configurations}}.
@@
 
-\subsubsection{Figures from Section~\ref{sec:5:experiments}}
-\label{app:5:plotsFromSec5}
+% \subsubsection{Figures from Section~\ref{sec:5:experiments}}
+% \label{app:5:plotsFromSec5}
 
-This last Appendix includes the figures used in Section~\ref{sec:5:experiments},
-with additional comments.
-
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{0.85\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized____env1-1_7803645526012310577.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{0.85\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized_semilogy____env1-1_7803645526012310577.pdf}
-    % \end{subfigure}
-    \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem]{Regret (in log-$y$ scale on the right) for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
-With a perfect knowledge on the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly in this simple setting, even with the knowledge of the horizon $T$ for \MusicalChair{}.
-The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can bounded with probability $0.5$ and $0.9$ respectively.
-The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.}
-    \label{fig:5:MP__K9_M3_T123456_N100__8_algos}
-\end{figure}
-
-%
-% Regular plots of centralized regrets
-%
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{0.49\textwidth}
-        \includegraphics[width=0.90\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_5016720151160452442.pdf}
-    % \end{subfigure}
-    % % ~
-    % \begin{subfigure}[!h]{0.49\textwidth}
-    %   \includegraphics[width=1.10\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
-    % \end{subfigure}
-    \caption[Failure case of \Selfish]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
-Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$).
-The regret for the three other algorithms is very small for this problem, always smaller than $100$ here.}
-    \label{fig:5:selfish_fail1}
-    % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-%
-% Regular plots of centralized regrets
-%
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{0.49\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_2643560344649862285.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{0.49\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_4683123079851881812.pdf}
-    % \end{subfigure}
-    \caption[Second failure case of \Selfish]{Regret, $M=2$ and $M=3$ players, $K=3$ arms, horizon $T=5000$, against $1000$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-\Selfish{} (top curve in \textcolor{darkgreen}{green}) clearly fails in such setting with small $K$.}
-    \label{fig:5:selfish_fail2}
-    % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-%
-% Regular plots of centralized regrets
-%
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        % \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
-        \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_1035303196230283176.pdf}
-        % \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_5016720151160452442.pdf}
-    % \end{subfigure}
-    ~
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
-    % \end{subfigure}
-    \caption[Third failure case of \Selfish]{Regret for $M=3$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
-Axis $x$ is for regret (different scale for each), and the top \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($11$ cases of $R_T \geq T$, out of $1000$).
-The regret for the three other algorithms is very small for this problem, and even appears constant.}
-    \label{fig:5:selfish_fail3}
-    % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-%
-% Regular plots of centralized regrets
-%
-
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        % \includegraphics[width=1.00\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3251433209347345969.pdf}
-    % \end{subfigure}
-    \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-\rhoRand{} (top \textcolor{blue}{blue}) is outperformed by the other algorithms (and the gain increases when $M$ increases), which all perform similarly in such configurations.
-Note that the (small) tail of the histograms come from complicated problems $\boldsymbol{\mu}$ and not failure cases.}
-    \label{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
-    % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_RegretCentralized____env1-1_3892966382091165662.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3892966382091165662.pdf}
-    % \end{subfigure}
-    \caption[Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-This extreme case $M=K$ shows the drastic difference of behavior between \RandTopM{} and \MCTopM, having constant regret, and \rhoRand{} and \Selfish, having large regret.}
-    \label{fig:5:MP__K9_M9_T5000_N500__4_algos__all_HistogramsRegret}
-    % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M2_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_2643359116089264295.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{1.00\textwidth}
-    %   \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K9_M9_T10000_N200__4_algos/all_RegretCentralized_loglog____env1-1_2306423191427933958.pdf}
-    % \end{subfigure}
-    \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$,Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand.
-\MCTopM{} is always among the best algorithms, and for $M$ not too small, its regret seems logarithmic with a constant matching the lower bound.}
-    \label{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}
-\end{figure}
-
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{0.75\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{0.75\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{0.75\textwidth}
-        \includegraphics[width=1.00\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
-    % \end{subfigure}
-    \caption[Regret for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$]{Regret (in log-log scale), for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$.
-The same observation as in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} can be made.
-\Selfish{} outperforms \MCTopM{} for $M=2$ here.
-Additionally, \MCTopM{} is the only algorithm to not fail dramatically when $M=K$ here.}
-    \label{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}
-\end{figure}
+% This last Appendix includes the figures used in Section~\ref{sec:5:experiments},
+% with additional comments.

### 2019-04-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Oulala la thèse commence à être bien trop longue...

@@
 
-\TODOL{Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
-
-% -----------------------------------------------------------------
-\subsubsection{Proof of Theorem~\ref{thm:5:BetterLowerBound}}
-
 The lower bound that we present relies on the following \emph{change-of-distribution} lemma that we prove in the next section, following recent arguments from \cite{Garivier16TrueShape} that have to be adapted to incorporate the collision information.
 
-\begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-player bandit models denoted by $\mu$ and $\lambda$ respectively, it holds that
+\begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-player bandit models denoted by $\bm{\mu}$ and $\bm{\lambda}$ respectively, it holds that
 \begin{equation}
-  \sum_{k=1}^{K}\E_{\mu}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\mu}(A),\Pr_{\lambda}(A)\right).
+  \sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).
 \end{equation}
@@
 
-Let $k$ be a sub-optimal arm under $\mu$,
+Let $k$ be a sub-optimal arm under $\bm{\mu}$,
 fix $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$,
-% We define the bandit instance $\lambda$ such that
-and let $\lambda$ be the bandit instance such that
+% We define the bandit instance $\bm{\lambda}$ such that
+and let $\bm{\lambda}$ be the bandit instance such that
 \[\left\{\begin{array}{ccl}
@@
 \right.\]
-Clearly, $\lambda\in\cP_M$ also,
-and the set of $M$ best arms under $\mu$ and $\lambda$ differ by one arm: if \Mbest$_{\mu} = \{1^*,\dots,M^*\}$ then $\Mbest_{\lambda} = \{1^*,\dots,(M-1)^*,k\}$.
+Clearly, $\bm{\lambda}\in\cP_M$ also,
+and the set of $M$ best arms under $\bm{\mu}$ and $\bm{\lambda}$ differ by one arm: if \Mbest$_{\bm{\mu}} = \{1^*,\dots,M^*\}$ then $\Mbest_{\bm{\lambda}} = \{1^*,\dots,(M-1)^*,k\}$.
 Thus, one expects the ($\cF^j_T$-mesurable) event
 \[A_T = \left(T_k^j(T) > \frac{T}{2M}\right)\]
-to have a small probability under $\mu$ (under which $k$ is sub-optimal) and a large probability under $\lambda$ (under which $k$ is one of the optimal arms, and is likely to be drawn a lot).
+to have a small probability under $\bm{\mu}$ (under which $k$ is sub-optimal) and a large probability under $\bm{\lambda}$ (under which $k$ is one of the optimal arms, and is likely to be drawn a lot).
 
@@
 \begin{eqnarray*}
-  \E_{\mu}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \kl \left(\Pr_{\mu}(A_T),\Pr_{\lambda}(A_T)\right),\nonumber \\
-  \E_{\mu}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - \Pr_{\mu}(A_T)\right) \log \left(\frac{1}{\Pr_{\lambda}(\overline{A_T})}\right) - \log(2),
+  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \kl \left(\Pr_{\bm{\mu}}(A_T),\Pr_{\bm{\lambda}}(A_T)\right),\nonumber \\
+  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - \Pr_{\bm{\mu}}(A_T)\right) \log \left(\frac{1}{\Pr_{\bm{\lambda}}(\overline{A_T})}\right) - \log(2),
 \end{eqnarray*}
@@
 \begin{align*}
-  \Pr_{\mu}\left(A_T\right)
-    & \leq 2M\frac{\E_\mu\left[T^j_k(T)\right]}{T} =: x_T, \\
-  \Pr_{\lambda}\left(\overline{A_T}\right)
-    & = \Pr_{\lambda}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_\lambda\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
+  \Pr_{\bm{\mu}}\left(A_T\right)
+    & \leq 2M\frac{\E_{\bm{\mu}}\left[T^j_k(T)\right]}{T} =: x_T, \\
+  \Pr_{\bm{\lambda}}\left(\overline{A_T}\right)
+    & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_\bm{\lambda}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
 \end{align*}
@@
 \begin{eqnarray}\label{eq:5:CqCD}
-  \E_{\mu}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - x_T\right) \log \left(\frac{T}{y_T}\right) - \log(2).
+  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - x_T\right) \log \left(\frac{T}{y_T}\right) - \log(2).
 \end{eqnarray}
-The strong uniform efficiency assumption  (see Definition~\ref{def:5:DecentralizedUniformEfficiency}) further tells us that $x_T \rightarrow 0$ (as $\E_\mu[T_k^j(T)] = o(T^\alpha)$ for all $\alpha$) and $y_T = o(T^\alpha)$ when $T \to \infty$, for all $\alpha \in (0,1)$.
+The strong uniform efficiency assumption  (see Definition~\ref{def:5:DecentralizedUniformEfficiency}) further tells us that $x_T \rightarrow 0$ (as $\E_{\bm{\mu}}[T_k^j(T)] = o(T^\alpha)$ for all $\alpha$) and $y_T = o(T^\alpha)$ when $T \to \infty$, for all $\alpha \in (0,1)$.
 As a consequence, observe that $\log(y_T)/\log(T) \rightarrow 0$ when $T$ tends to infinity and
@@
 \begin{equation}\label{eq:5:CqCD2}
-  \liminf_{T\to\infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k,\mu_{M^*}+\varepsilon)},
+  \liminf_{T\to\infty} \frac{\E_{\bm{\mu}}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k,\mu_{M^*}+\varepsilon)},
 \end{equation}
@@
 
-Under observation model \modelun{} and \modeldeux, the strategy $A^j(t)$ decides which arm to play based on the information contained in $I_{t-1}$, where $I_0 = U_0$ and
-\[\forall t > 0, \ \ I_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
+Under observation model \modelun{} and \modeldeux, the strategy $A^j(t)$ decides which arm to play based on the information contained in $\cO_{t-1}$, where $\cO_0 = U_0$ and
+\[\forall t > 0, \ \ \cO_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
 where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $\{1,\dots,M\}$ for new ranks or arms.} useful to select $A^j(t)$.
-Formally, one can say that $A^j(t)$ is $\sigma(I_{t-1})$ measurable\footnote{$\sigma(I_t)$ denotes the sigma-Algebra generated by the observations $I_t$.} (as $\cF^j(t) \subseteq \sigma(I_{t})$, with an equality under observation model \modelun).
+Formally, one can say that $A^j(t)$ is $\sigma(\cO_{t-1})$ measurable, if $\sigma(\cO_t)$ denotes the sigma-Algebra generated by the observations $\cO_t$, and as $\cF^j(t) \subseteq \sigma(\cO_{t})$, with an equality under observation model \modelun.
 
-Under two bandit models $\mu$ and $\lambda$, we let $\Pr_{\mu}^{I_t}$ (resp.
-$\Pr_{\lambda}^{I_t}$) be the distribution of the observations under model $\mu$ (resp.
-$\lambda$), given a fixed algorithm.
-Using the exact same technique as \cite{Garivier16TrueShape} (the contraction of entropy principle), one can establish that for any event $A$ that is $\sigma(I_t)$-measurable\footnote{In the work of \cite{Garivier16TrueShape}, the statement is more general and the probability of an event $A$ is replaced by the expectation of any $\cF_T$-measurable random variable $Z$ bounded in $[0,1]$.},
-\[\KL\left(\Pr_{\mu}^{I_t},\Pr_{\lambda}^{I_t}\right) \geq \kl\left(\Pr_{\mu}(A),\Pr_{\lambda}(A)\right).\]
+Under two bandit models $\bm{\mu}$ and $\bm{\lambda}$, we let $\Pr_{\bm{\mu}}^{\cO_t}$ (resp.
+$\Pr_{\bm{\lambda}}^{\cO_t}$) be the distribution of the observations under model $\bm{\mu}$ (resp.
+$\bm{\lambda}$), given a fixed algorithm.
+Using the exact same technique as \cite{Garivier16TrueShape}, called the ``contraction of entropy'' principle, one can establish that for any event $A$ that is $\sigma(\cO_t)$-measurable\footnote{In the work of \cite{Garivier16TrueShape}, the statement is more general and the probability of an event $A$ is replaced by the expectation of any $\cF_T$-measurable random variable $Z$ bounded in $[0,1]$.},
+\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) \geq \kl\left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).\]
 
-The next step is to relate the complicated KL-divergence $\KL\left(\Pr_{\mu}^{I_t},\Pr_{\lambda}^{I_t}\right)$ to the number of arm selections.
+The next step is to relate the complicated KL-divergence $\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right)$ to the number of arm selections.
 Proceeding similarly as \cite{Garivier16TrueShape}, one can write, using the chain rule for KL-divergence, that
-\begin{equation}\label{eq:5:Iterative}\KL\left(\Pr_{\mu}^{I_{t}},\Pr_{\lambda}^{I_{t}}\right) =  \KL\left(\Pr_{\mu}^{I_{t-1}},\Pr_{\lambda}^{I_{t-1}}\right) + \KL\left(\Pr_{\mu}^{Y_{t},C_t,U_t | I_{t-1}},\Pr_{\lambda}^{Y_{t},C_t,U_t | I_{t-1}}\right).\end{equation}
-Now observe that conditionally to $I_{t-1}$, $U_t$, $Y_t$ and $C_t$ are independent, as once the selected arm is known, the value of the sensing $Y_t$ does not influence the other players selecting that arm, and $U_t$ is some exogenous randomness.
-Using further that the distribution of $U_t$ is the same under $\mu$ and $\lambda$, one obtains
-\begin{equation}\KL\left(\Pr_{\mu}^{Y_{t},C_t | I_{t-1}},\Pr_{\lambda}^{Y_{t},C_t | I_{t-1}}\right) =
- \KL\left(\Pr_{\mu}^{Y_{t}| I_{t-1}},\Pr_{\lambda}^{Y_{t}| I_{t-1}}\right) + \KL\left(\Pr_{\mu}^{C_t | I_{t-1}},\Pr_{\lambda}^{C_t | I_{t-1}}\right).\label{eq:5:DecKL}
+\begin{equation}\label{eq:5:Iterative}\KL\left(\Pr_{\bm{\mu}}^{\cO_{t}},\Pr_{\bm{\lambda}}^{\cO_{t}}\right) =  \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t,U_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t,U_t | \cO_{t-1}}\right).\end{equation}
+Now observe that conditionally to $\cO_{t-1}$, $U_t$, $Y_t$ and $C_t$ are independent, as once the selected arm is known, the value of the sensing $Y_t$ does not influence the other players selecting that arm, and $U_t$ is some exogenous randomness.
+Using further that the distribution of $U_t$ is the same under $\bm{\mu}$ and $\bm{\lambda}$, one obtains
+\begin{equation}\KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t | \cO_{t-1}}\right) =
+ \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right).\label{eq:5:DecKL}
 \end{equation}
 
-The first term in \eqref{eq:5:DecKL} can be rewritten using the same argument as \cite{Garivier16TrueShape}, that relies on the fact that conditionally to $I_{t-1}$, $Y_t$ is a Bernoulli distribution with mean $\mu_{A^j(t)}$ under the instance $\mu$ and $\lambda_{A^j(t)}$ under the instance $\lambda$:
+The first term in \eqref{eq:5:DecKL} can be rewritten using the same argument as \cite{Garivier16TrueShape}, that relies on the fact that conditionally to $\cO_{t-1}$, $Y_t$ is a Bernoulli distribution with mean $\mu_{A^j(t)}$ under the instance $\bm{\mu}$ and $\lambda_{A^j(t)}$ under the instance $\bm{\lambda}$:
 \begin{eqnarray*}
-\KL\left(\Pr_{\mu}^{Y_{t}| I_{t-1}},\Pr_{\lambda}^{Y_{t}| I_{t-1}}\right)  & = & \E_\mu\left[\E_\mu\left[ \KL\left(\Pr_{\mu}^{Y_{t}| I_{t-1}},\Pr_{\lambda}^{Y_{t}| I_{t-1}}\right) | I_{t-1} \right]\right] \\
-& = & \E_\mu \left[\kl \left(\mu_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
-& = & \E_\mu  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right],
+\KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right)  & = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
+& = & \E_{\bm{\mu}} \left[\kl \left(\mu}_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
+& = & \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu}_{k},\lambda_{k}\right)\right],
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-\KL\left(\Pr_{\mu}^{C_t | I_{t-1}},\Pr_{\lambda}^{C_t | I_{t-1}}\right)
-& = & \E_\mu\left[\E_\mu\left[ \KL\left(\Pr_{\mu}^{C_t| I_{t-1}},\Pr_{\lambda}^{C_t| I_{t-1}}\right) | I_{t-1} \right]\right] \\
-& = & \E_\mu\left[\E_\mu\left[\E_\mu\left[ \KL\left(\Pr_{\mu}^{C_t| I_{t-1}},\Pr_{\lambda}^{C_t| I_{t-1}}\right) | \bigcup_{j' \neq j} I_{t-1}^{j'}\right]| I_{t-1} \right]\right], \\
+\KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right)
+& = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
+& = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \bigcup_{j' \neq j} \cO_{t-1}^{j'}\right]| \cO_{t-1} \right]\right], \\
 \end{eqnarray*}
-where $I_{t-1}^{j'}$ denote the information available to player $j' \neq j$.
-Knowing the information available to all other players player $C_t$ is an almost surely constant random variable, whose distribution is the same under $\mu$ and $\lambda$.
-Hence the inner expectation is zero and so does $\KL\left(\Pr_{\mu}^{C_t | I_{t-1}},\Pr_{\lambda}^{C_t | I_{t-1}}\right)$.
-
-\TODOL{This is wrong as explained in \cite{BoursierPerchet18}! Rewrite this to include new knowledge presented in Appendix of \cite{KaufmannAbbas19}.}
+where $\cO_{t-1}^{j'}$ denote the information available to player $j' \neq j$.
+\textcolor{red}{By hypothesis}, the ``collision information term'' is assumed to be a $\smallO{\log(T)}$,
+and thus $\KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right) = \smallO{\log(T)}$.
 
 Putting things together, we showed that
-\[\KL\left(\Pr_{\mu}^{I_t},\Pr_{\lambda}^{I_t}\right) = \KL\left(\Pr_{\mu}^{I_{t-1}},\Pr_{\lambda}^{I_{t-1}}\right) + \E_\mu  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right].\]
-Iterating this equality and using that $\KL\left(\Pr_{\mu}^{I_{0}},\Pr_{\lambda}^{I_{0}}\right)=0$ yields that
-\[\sum_{k=1}^{K}\E_{\mu}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\mu}(A),\Pr_{\lambda}(A)\right),\]
-for all $A \in \sigma(I_T)$, in particular for all $A \in \cF^j_T$.
+\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) = \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu}_{k},\lambda_{k}\right)\right] + \smallO{\log(T)}.\]
+Iterating this equality and using that $\KL\left(\Pr_{\bm{\mu}}^{\cO_{0}},\Pr_{\bm{\lambda}}^{\cO_{0}}\right)=0$ yields that
+\[\sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu}_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right) + \smallO{\log(T)},\]
+for all $A \in \sigma(\cO_T)$, in particular for all $A \in \cF^j_T$.
 

### 2019-04-08 2-Chapters/5-Chapter/chapter5.tex
Oulala la thèse commence à être bien trop longue...

@@
 As illustrated above, if we consider algorithms building on the same ideas as \RhoRand{} or \MCTopM, it seems mandatory to first build an estimate of the value of $M$ then run the initial algorithm that assumed a perfect knowledge of $M$.
-Two possible directions exist to estimate on the fly this number.
+Two possible directions exist to estimate $M$ on the fly.
 
@@
 
-\TODOL{Include a simulation to show the behavior of \rhoRand{} vs \rhoRandEst, and \RandTopM{} and \MCTopM{} with prior knowledge of $M$ vs the ``estimate'' versions.}
+\TODOL{Include a simulation to show the behavior of \rhoRand{} vs \rhoRandEst, and \RandTopM{} and \MCTopM{} with prior knowledge of $M$ vs the ``estimate'' versions.
+
+FIXME Je dois refaire ces simulations, les chiffres ici sont avec $T=1000$ et $N=4$ répétitions (juste faites vite fait dans le train, je dois les lancer sur la machine du bureau).
+}
 
@@
             \hline
-            Centralized multiple-play \klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            Centralized multiple-play \klUCB{} & $\mathbf{52 \pm 15}$ & $\mathbf{35 \pm 15}$ & $\mathbf{0}$ \\
             \hline
-            \RhoRand-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
-            \RhoRand-\klUCB{} + Estimate $M$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \RhoRand-\klUCB{} & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
+            \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$326 \pm 86$} & \textcolor{red}{$1213 \pm 168$} & \textcolor{red}{$2403 \pm 121$} \\
             \hline
-            \RandTopM-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
-            \RandTopM-\klUCB{} + Estimate $M$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \RandTopM-\klUCB{} & $188 \pm 18$ & $671 \pm 159$ & $790 \pm 470$ \\
+            \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$238 \pm 27$} & \textcolor{red}{$778 \pm 156$} & \textcolor{red}{$1098 \pm 153$} \\
             \hline
-            \MCTopM-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
-            \MCTopM-\klUCB{} + Estimate $M$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \MCTopM-\klUCB{} & $134 \pm 6$ & $304 \pm 19$ & $42 \pm 5$ \\
+            \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$210 \pm 66$} & \textcolor{red}{$465 \pm 39$} & \textcolor{red}{$602 \pm 54$} \\
             \hline
-            \Selfish-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+            \Selfish-\klUCB{} & $136 \pm 17$ & $468 \pm 87$ & $1379 \pm 16$ \\
             \hline
         \end{tabular}
-        \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, comparing algorithms which knows $M$ against algorithms which estimate $M$ on the fly.}
+        \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, comparing algorithms which knows $M$ against algorithms which estimate $M$ on the fly.
+The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar costs on their regret when they have to estimate $M$ on the fly.
+Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.}
         \label{table:5:meanRegretSimulationsEstimatingM}

### 2019-04-08 2-Chapters/3-Chapter/chapter3.tex
Small proofreading of Chapter 4.
Hard to focus

@@
 Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
-\TODOL{Do we really have time/space to talk about Markov models?}
-Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.
+% \TODOL{Do we really have time/space to talk about Markov models?}
+% Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.
 
@@
 \begin{figure}[h!]  % [htbp]
-	\centering
-	\includegraphics[width=0.95\linewidth]{3.pdf}
+	% \centering
+	\includegraphics[width=1.05\linewidth]{3.pdf}
 	\caption[Example of a single-player simulation showing the average regret and histogram of regrets of $4$ algorithms]{
@@
 \begin{figure}[h!]  % [htbp]
-	\centering
-	\includegraphics[width=0.95\linewidth]{3_hist.pdf}
+	% \centering
+	\includegraphics[width=1.05\linewidth]{3_hist.pdf}
 	\caption{Histogram of regret for the same experiment.}
@@
 
-% ----------------------------------------------------------------------------
-\section{An example of another model: rested or restless Markov models}
-\label{sec:3:markovModels}
+% % ----------------------------------------------------------------------------
+% \section{An example of another model: rested or restless Markov models}
+% \label{sec:3:markovModels}
 
-%  - FIXME maybe not included?
+% %  - FIXME maybe not included?
 
-Maybe if I have space at the end of this chapter, I can show another part of the library, maybe introduce the rested/restless Markov models (with maybe just one example) and show simulations in this model?
+% Maybe if I have space at the end of this chapter, I can show another part of the library, maybe introduce the rested/restless Markov models (with maybe just one example) and show simulations in this model?
 
-SMPyBandits also implemented sparse MAB models (with $s<K$ arms with a mean $\mu_i>0$ and $K-s$ arms with a mean $\mu_i \leq 0$), I could also present this model, the various algorithms designed to tackle it, and some numerical experiments.
+% SMPyBandits also implemented sparse MAB models (with $s<K$ arms with a mean $\mu_i>0$ and $K-s$ arms with a mean $\mu_i \leq 0$), I could also present this model, the various algorithms designed to tackle it, and some numerical experiments.
 
@@
 
-\TODOL{Write it!}
+Nowadays, parallelism is everywhere in the computation world, and any serious framework for numerical simulations must try at least one of the three approaches to gain from parallelism.
+%
+For all the different numerical simulations presented in this thesis, except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4}, the setting is the same: we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs.
+On the first hand, Because of the sequential learning aspect, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
+On the other hand, parallelism can help greatly for the two other variables: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
+Similarly, if we want to simulate $1000$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $1000/4=250$ repetitions on each process also bring a speed-up factor of $4$.
+%
+In this last section, we quickly review the chosen approach for SMPyBandits (multi-core on one machine), and we explain why the two other approaches were less appropriate for our study of multi-armed bandit problems.
+
+The first approach is to use multiple cores of the same machines, and because it is both the simplest and the less financially as well as ecologically costly, this is the approach implemented in SMPyBandits.
+The machines I had access to during my thesis, either my own laptop or a workstation hosted the SCEE team in CentraleSupélec campus, were equipped with $i5$ or $i7$ Intel CPU with $4$ or $12$ cores.
 
@@
 
-Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).
+% Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).
+%
+The two other approaches we could have consider is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
+Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.
+Since $2016$, we saw a large press coverage as well as an extensive use in research of deep learning libraries that make general-purpose machine learning algorithms train on the GPU of a user's laptop or a cluster of GPU.
+This success is mainly possible because of the heavy parallelism of such training algorithms, and the parallel nature of GPU.
+To the best of the author knowledge, nobody has tried to implement high performance MAB simulations by using the ``parallelism power'' of a GPU (at least, no code for such experiments were made public in $2019$).
+I worked on a GPU, implementing fluid dynamic simulations in an internship in $2012$, and I have since then kept a curiosity on how to use GPU-powered libraries and code.
+The author has contributed to famous libraries like Theano (\href{http://deeplearning.net/software/theano/}{\texttt{DeepLearning.net/software/theano}}) or Keras (\href{https://keras.io/}{\texttt{keras.io}}), and my limited knowledge on such libraries made me believe that it was not easy to use a GPU for bandit simulations, and most surely it would not have been worth the time.
+
+I did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
+It is partly due to time constraint, as I would have been curious to try, but also because we found that it would not have helped us much to use a large scale cluster.
+The main reason is that in the multi-armed bandit and sequential learning literature, most research papers do not even include an experimental section, and for the papers who do take the time to implement and test their proposed algorithms, it is almost done on just a few problems and for short- or medium- duration experiments.
+For instance, the papers we consider to be the best ones regarding their empirical sections are \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, for piece-wise stationary bandits, and they mainly consider reasonable problems of horizon $T=10000$ and no more than $1000$ independent repetitions.
+Each paper considers one harder problem, of horizon $T=1000000$ and less repetitions.
+%
+In each article written during my thesis, we included extensive numerical simulations, and even the longest ones (for \cite{Besson2019GLRT}) were short enough to run in less than $12$ hours on a $12$-core workstation, so we could run a few large-scale simulations over night.
+
+\TODOL{Finish this section}

### 2019-04-08 2-Chapters/3-Chapter/chapter3.tex
Almost done for this last note finishing Chapter:3 about parallel computations for SMPyBandits.

@@
 
-\subsection{Using multi-core to speed-up simulations}
+\subsection{A final note about parallelism in SMPyBandits}
+\label{sub:3:parallelSimulations}
+% Using multi-core to speed-up simulations
 
-Nowadays, parallelism is everywhere in the computation world, and any serious framework for numerical simulations must try at least one of the three approaches to gain from parallelism.
+Nowadays, parallelism is everywhere in the computational world, and any serious framework for numerical simulations must explore at least one of the three main approaches to (try to) gain performance from parallelism.
 %
-For all the different numerical simulations presented in this thesis, except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4}, the setting is the same: we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs.
-On the first hand, Because of the sequential learning aspect, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
-On the other hand, parallelism can help greatly for the two other variables: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
-Similarly, if we want to simulate $1000$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $1000/4=250$ repetitions on each process also bring a speed-up factor of $4$.
+For all the different numerical simulations presented in this thesis, except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4}, the setting is the same: we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=100$ like in Chapter~\ref{chapter:6}).
+On the first hand, because of the fundamentally sequential nature of bandit games, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
+On the other hand, parallelism can help greatly for the two other axes: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
+Similarly, if we want to run $100$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $100/4=25$ repetitions on each process also bring a speed-up factor of $4$.
 %
@@
 
+\paragraph{What we did implement: \tettt{Joblib} for multi-core simulations.}
+%
 The first approach is to use multiple cores of the same machines, and because it is both the simplest and the less financially as well as ecologically costly, this is the approach implemented in SMPyBandits.
 The machines I had access to during my thesis, either my own laptop or a workstation hosted the SCEE team in CentraleSupélec campus, were equipped with $i5$ or $i7$ Intel CPU with $4$ or $12$ cores.
+%
+As explained in the page \href{https://smpybandits.github.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}, we implemented in SMPyBandits an easy way to run any simulations on $n$ cores of a machine, using the \texttt{Joblib} \cite{joblib} library.
+It is implemented in a completely transparent way, and if someone uses the command-line variable to configure experiments, using one core or all the cores of the machine one changes \texttt{N\_JOBS=1} to \texttt{N\_JOBS=-1}, like in this example.
 
-Quickly explain the advantages and drawbacks of using \texttt{Joblib} and \texttt{joblib.Parallel} to run simulations on multiple cores of the same computer
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+    \begin{listing}[h!]
+        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
+$ BAYES=False ARM_TYPE=Bernoulli N=100 T=10000 K=9 N_JOBS=1 \
+  python3 main.py configuration.py
+        \end{minted}
+        \caption{To run such simulation using the maximum number of cores, use \texttt{N\_JOBS=-1} instead.}
+        % \captionof{listing}{Run a simulation using 1 or \texttt{N} cores \label{lst:3:runOneCoreOrMore}.}
+        \label{lst:3:runOneCoreOrMore}
+    \end{listing}
+\end{small}
+
+As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal, dividing the running time by \texttt{N\_JOBS}.
+But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is highly sub-optimal, as the CPU will essentially spend all his time (and memory) managing the different jobs, and not actually doing the simulations.
+Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits.
+We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $20$ jobs.
+We give in Table~\ref{table:3:speedUpTimeParallelComputations}
+
+\TODOL{Je veux terminer cet exemple!}
+
+% FIXME run
+% for N in 1 10 100; do for NJOBS in 1 4 100; do DEBUG=True SAVEALL=False NOPLOTS=True N=$N T=10000 N_JOBS=$NJOBS make single; read; done
+
+\begin{table}[ht]
+% \begin{footnotesize}
+    \centering
+    \begin{tabular}{c|ccc}
+    \textbf{Repetitions} $\;$ \textbackslash $\;$ \texttt{N\_JOBS} & $1$ & $4$ & $20$ \\
+        \hline
+        $1$ repetition    & $XXX \;\text{s}$ & $XXX \;\text{s}$ & $XXX \;\text{s}$ \\
+        $10$ repetitions  & $XXX \;\text{s}$ & $XXX \;\text{s}$ & $XXX \;\text{s}$ \\
+        $100$ repetitions & $59 \;\text{s}$ & $37 \;\text{s}$ & $XXX \;\text{s}$ \\
+        \hline
+    \end{tabular}
+    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.}
+    \label{table:3:speedUpTimeParallelComputations}
+% \end{footnotesize}
+\end{table}
 
 % Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).
+\paragraph{Approaches we did not try.}
 %
 The two other approaches we could have consider is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
+
+On the first hand, I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
 Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.
@@
 I worked on a GPU, implementing fluid dynamic simulations in an internship in $2012$, and I have since then kept a curiosity on how to use GPU-powered libraries and code.
-The author has contributed to famous libraries like Theano (\href{http://deeplearning.net/software/theano/}{\texttt{DeepLearning.net/software/theano}}) or Keras (\href{https://keras.io/}{\texttt{keras.io}}), and my limited knowledge on such libraries made me believe that it was not easy to use a GPU for bandit simulations, and most surely it would not have been worth the time.
+I have contributed to and used famous deep-learning libraries, like Theano (\href{http://deeplearning.net/software/theano/}{\texttt{DeepLearning.net/software/theano}}) or Keras (\href{https://keras.io/}{\texttt{keras.io}}), and my limited knowledge on such libraries made me believe that it was not easy to use a GPU for bandit simulations, and most surely it would not have been worth the time.
+I would be very curious to understand how a GPU could be used to implement highly efficient simulations for sequential learning problems, because it seemed hard whenever I thought about it.
 
-I did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
-It is partly due to time constraint, as I would have been curious to try, but also because we found that it would not have helped us much to use a large scale cluster.
-The main reason is that in the multi-armed bandit and sequential learning literature, most research papers do not even include an experimental section, and for the papers who do take the time to implement and test their proposed algorithms, it is almost done on just a few problems and for short- or medium- duration experiments.
+On the other hand, I also did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
+It is partly due to time constraint, as I would have been curious to try, but mainly because we found that it would not have helped us much to use a large scale cluster.
+The main reason is that in the multi-armed bandit and sequential learning literature, most research papers do not even include an experimental section, and for the papers who did take the time to implement and test their proposed algorithms, it is almost done on just a few problems and for short- or medium- duration experiments.
 For instance, the papers we consider to be the best ones regarding their empirical sections are \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, for piece-wise stationary bandits, and they mainly consider reasonable problems of horizon $T=10000$ and no more than $1000$ independent repetitions.
@@
 In each article written during my thesis, we included extensive numerical simulations, and even the longest ones (for \cite{Besson2019GLRT}) were short enough to run in less than $12$ hours on a $12$-core workstation, so we could run a few large-scale simulations over night.
-
-\TODOL{Finish this section}
+For such reasons, we prefer to not try to run simulations on a cluster.

### 2019-04-08 2-Chapters/3-Chapter/chapter3.tex
Done

@@
 This chapter is based on the following publication: \cite{SMPyBanditsJMLR}.
-The code code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is publicly released under the MIT open-source License.
+The code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is publicly released under the MIT open-source License.
 
@@
 Note that the second step should of course be repeated a large number of times if we want to study \emph{mean} regret and not only the regret in one single trajectory.
-
-% FIXME finish this paragraph
-If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and ensure XXX.
+If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms.
 
@@
 
-SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms from the last 10 years (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}).
+SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}).
 %
@@
 which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$).
-
-The default is to use the same distribution for the $K$ arms, but it also possible to mix them.
-
-% FIXME
-\TODOL{point to the notebook!}
-% https://perso.crans.org/besson/phd/notebooks/Easily_creating_MAB_problems.html
+%
+The default is to use the same distribution for the $K$ arms, but it also possible to mix them (see for instance Figure~\ref{fig:25:HarderMixed}).
+%
+We do not give more details, but the interested reader can refer to the following Jupyter notebook \cite{jupyter} document hosted at:\\
+\begin{small}
+    \href{https://smpybandits.github.io/notebooks/Easily_creating_MAB_problems.html}{\texttt{SMPyBandits.GitHub.io/notebooks/Easily\_creating\_MAB\_problems.html}}
+\end{small}
 
@@
 Choosing the algorithms is easy by customizing the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
-For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{Policies.klUCB}}) against the non-anytime variant $\klUCB^{++}$ algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{Policies.klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{Policies.UCB}}) and Thompson sampling (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Policies.Thompson}}) with a Beta posterior (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Policies.Posterior.Beta}})).
+For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{Policies.klUCB}}) against the non-anytime variant $\klUCB^{++}$ algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{Policies.klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{Policies.UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Policies.Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Policies.Posterior.Beta}})).
 
@@
 
-Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in a HDF5 file (\cite{h5py}, \texttt{\href{http://docs.h5py.org/en/stable/high/file.html}{docs.h5py.org/en/stable/high/file.html}})\footnote{~For example, this simulation produces this HDF5 file\\\texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}}.
-Figure~\ref{fig:3:firstPlot} below shows the average regret for these $4$ algorithms.
-The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.
-The Figure~\ref{fig:3:firstPlot_hist} shows the histogram of regret for the same experiment.
-
 \begin{figure}[h!]  % [htbp]
@@
 \end{figure}
+
+Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in a HDF5 file\footnote{~For example, this simulation produces this HDF5 file\\\texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}}
+\cite{h5py}.
+% \texttt{\href{http://docs.h5py.org/en/stable/high/file.html}{docs.h5py.org/en/stable/high/file.html}}).
+Figure~\ref{fig:3:firstPlot} above shows the average regret for these $4$ algorithms.
+The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.
+The Figure~\ref{fig:3:firstPlot_hist} below shows the histogram of regret for the same experiment.
 
@@
 
-\subsection{Research papers using SMPyBandits}
+\subsection{Research works using SMPyBandits}
 
-SMPyBandits was used for the following research articles since $2017$.
+SMPyBandits was used for the following research articles since $2017$, and it is used for the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
 
-In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2}, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
+In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
 We designed a variant of the Exp4 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
@@
 
-In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5}, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
@@
 
-In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6}, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
+In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
 We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.

### 2019-04-08 2-Chapters/5-Chapter/chapter5.tex
Done
Expérience en cours pour la fin du Chapitre:3.

@@
 FIXME Je dois refaire ces simulations, les chiffres ici sont avec $T=1000$ et $N=4$ répétitions (juste faites vite fait dans le train, je dois les lancer sur la machine du bureau).
+
+% for M in 3 6 9; do DEBUG=False SAVEALL=True NOPLOTS=True M=$M K=9 N_JOBS=-1 N=100 T=10000 make moremultiplayers; read; done
 }
@@
     \end{tabular}
-        \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, comparing algorithms which knows $M$ against algorithms which estimate $M$ on the fly.
-The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar costs on their regret when they have to estimate $M$ on the fly.
-Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.}
+    \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, comparing algorithms which knows $M$ against algorithms which estimate $M$ on the fly.}
     \label{table:5:meanRegretSimulationsEstimatingM}
@@
 \TODOL{Give interpretation from these results!}
+One can observe in Table~\ref{table:5:meanRegretSimulationsEstimatingM} the empirical performances of different algorithms in an example problem, in each case of low, medium and maximum number of players ($M=3,6,9$ for $K=9$ arms).
+The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar costs on their regret when they have to estimate $M$ on the fly.
+Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.
+For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large (on the three algorithms and two cases, about $+30\%$), and consistent on the different algorithm.
+The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proved to be $\smallO{\log(T)}$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret (written ``Estimate $M$'' in Table~\ref{table:5:meanRegretSimulationsEstimatingM}).
+In this extreme case, \RhoRand{} and its extension both perform very closely, as expected.
 

### 2019-04-08 2-Chapters/5-Chapter/chapter5.tex
Done for today...

@@
 
-\TODOL{Give interpretation from these results!}
 One can observe in Table~\ref{table:5:meanRegretSimulationsEstimatingM} the empirical performances of different algorithms in an example problem, in each case of low, medium and maximum number of players ($M=3,6,9$ for $K=9$ arms).
@@
 In this extreme case, \RhoRand{} and its extension both perform very closely, as expected.
+%
+Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension \MCTopM-\klUCB.
 
@@
 Explaining in details their model and proposed solutions would be quite lengthy, and we rather prefer to let the interested reader refer to the later work \cite{AvnerMannor18}.
-% where they essentially propose an efficient algorithm for this problem, that also reach a logarithmic regret when $M \leq K$ players are running independently the same algorithm.
-\TODOL{FIXME don't copy/paste their abstract...}
 We sum-up its contributions quickly.
-They we address several aspects of the challenge of communication networks shared by many users simultaneously: learning unknown stochastic network characteristics, sharing resources with other users while keeping coordination overhead to a minimum.
-The proposed solution combines Multi-Armed Bandit learning with a lightweight signalling-based coordination scheme, and ensures convergence to a stable allocation of resources.
-They work considers single-user level algorithms for two scenarios:an unknown fixed number of users, and a dynamic number of users.
+They address several aspects of the challenge of communication networks shared by many users simultaneously: learning unknown stochastic network characteristics, sharing resources with other users while keeping coordination overhead to a minimum.
+The solution they proposed combines Multi-Armed Bandit learning with a lightweight signalling-based coordination scheme, and ensures convergence to a stable allocation of resources.
+Their work considers single-user level algorithms for two scenarios: an unknown fixed number of users, and a dynamic number of users, both for different arms means for each users (players).
 Analytic performance guarantees, proving convergence to stable marriage configurations, are presented for both setups.
-The algorithms are designed based on a system-wide perspective, rather than focusing on single user welfare.
-Thus, maximal re-source utilization is ensured.
+The algorithms are based on a system-wide perspective, rather than focusing on single user welfare.
 
@@
 We believe that we could easily prove that a sub-linear regret is achievable, if all players run \MCTopM-\klUCB{} (and if currently active players restart their memory of the past observations every $L$ time steps).
-As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, using the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to our proposal \MCTopM{} would also give a small regret upper-bound.
+As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, applying the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to our proposal \MCTopM{} would also give a small regret upper-bound.
 
@@
 
-\cite{NaparstekCohen17}
+Due to space constraint, we simply quote here three recent articles which proposed a similar approach of using decentralized reinforcement learning algorithms on the device-side on wireless networks,
+but proposed models closer to the reality of wireless networks.
+%
+The first work is \cite{NaparstekCohen17}, which proposes to use a decentralized learning based on deep learning, in the ``no sensing'' case, but in a model where the users have to learn not only the channels to use for their uplink messages (only from the feedback through the acknowledgements, like in Section~\ref{sec:4:firstModel}), but where they have to learn the whole ``spectrum access actions''.
+% Deep Multi-User Reinforcement Learning for Dynamic Spectrum Access in Multichannel Wireless Networks
+% FIXME continue!
 
-Distributed Learning for Channel Allocation Over a Shared Spectrum
-S.M. Zafaruddin, Ilai Bistritz, Amir Leshem, Dusit Niyato
-https://arxiv.org/pdf/1902.06353
-\cite{Zafaruddin2019}
+%
+The work of \cite{AvnerMannor18} is discussed above for the case with communication between players, and their model is of high interest from the point-of-view of real-world wireless communication protocols.
+
+Finally, the very recent work of \cite{Zafaruddin2019} is the first one to present experiments done on simulated LTE and 5G channels.
+The mathematics behind their model are actually quite close to our model, but they explain it in terms of OFDMA and Quality-of-Service (QoS).
+Their algorithm is essentially based on a pre-agreement of the $M$ players, that will deterministically run an alternance of exploration phase, auction phase and exploitation phase, of (exponentially) increasing durations.
+By diving into the details of the modulation and giving explicitly the form of the uplink messages sent by the devices, the authors are able to set-up two different uplink packets, to efficiently perform the auction phase (see Figure~2).
+They prove a regret upper-bound of the order $R_T = \bigO{\log(T)}$, with no special care regarding the constants, and we can also note that their algorithm require a prior knowledge of $\Delta_{\min}$ a problem-dependent constant (Theorem~3).
 
@@
 Similarly to our demonstration of multi-armed bandit learning in an IoT network that we presented in the previous Chapter~\ref{chapter:4} in Section~\ref{sec:4:gnuradio},
-we review in this section some related works, dating back from $2016$, who proposed a similar approach.
+we list in this section some related works, dating back from $2016$, who proposed a similar approach.
 All the similar works that we are aware of also used USRP boards, and implemented the demonstration using either the Simulink and MATLAB softwares, or the GNU Radio software like we did \cite{Besson2018ICT}.
 
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+In \cite{darak2016bayesian,Darak16}, the authors study the same model and focussed on the \RhoRand{} policy, which was the state-of-the-art back in $2016$, combined with \UCB{} or \klUCB, as well as an extension using Bayes-UCB from \cite{Kaufmann12BUCB}.
+Their idea is to use the same skeleton as \RhoRand, that is to assign a \emph{rank} to each player, and make players change their rank after any collision.
+But instead of selecting a new rank uniformly at random among $\{1,\dots,M\}$, they proposed to use a ``second-stage'' learning policy\footnote{This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.}, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
+This algorithm is implemented in SMPyBandits and named \texttt{rhoLearn}\footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.}, where it can use Bayes-UCB or any of the $65$ or more bandit algorithms available in the library.
+Even if no theoretical guarantee backs up this idea of second-stage learning with Bayes-UCB, empirical simulations found that it can be efficient.
+We illustrate this in Table~\ref{table:5:comparisonRhoRandRhoLearn} below, where we consider the same problem as described above in Section~\ref{sub:5:withoutSensing} (Table~\ref{table:5:meanRegretSimulationsNoSensing}).
+The first \texttt{rhoLearn} uses Bayes-UCB for the second-stage learning, while the second uses Exp3,
+and we include both the centralized multi-play version of \klUCB{} and our proposal \MCTopM-\klUCB{} for comparison.
 
-\cite{Darak16}
+% FIXME run the simulations!
+\begin{table}[ht]
+% \begin{footnotesize}
+    \centering
+    \begin{tabular}{c|ccc}
+    \textbf{Algorithm} $\;$ \textbackslash $\;$ Number of players & $M=3$ & $M=6$ & $M=9$ \\
+        \hline
+        Centralized multiple-play \klUCB{} & $\mathbf{52 \pm 15}$ & $\mathbf{35 \pm 15}$ & $\mathbf{0}$ \\
+        \hline
+        \RhoRand-\klUCB{} & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
+        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
+        \texttt{rhoLearn}-\klUCB{} + Exp3 & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
+        \hline
+        \MCTopM-\klUCB{} & $134 \pm 6$ & $304 \pm 19$ & $42 \pm 5$ \\
+        \hline
+    \end{tabular}
+    \caption{Comparison of \RhoRand{} and \texttt{rhoLearn} for a simple multi-player bandit problem with $K=9$ arms.}
+    \label{table:5:comparisonRhoRandRhoLearn}
+% \end{footnotesize}
+\end{table}
 
-\cite{modiDemo2016}
+% FIXME mcuh more details?
 
-\cite{kumar2016two}
+Other works in the same line of research included \cite{modiDemo2016} implemented using the same USRP testbed as for our demonstration in Chapter~\ref{chapter:4} (Section~\ref{sec:4:gnuradio}),
+and the first work of a research team based in IIT Dehli, \cite{kumar2016two}.
+%
+More recent works by the same team, \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}, all follow the same approach and include an experimental section that shows the results of an implementation of the proposed algorithms on USRP boards.
 
-\cite{KumarYadav2018}
-
-\cite{SawantKumar2018}
-
-\cite{JoshiKumar2018}
+\TODOL{I should explain more, no? Who does what and why? So long...}
 
@@
 
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+Since the beginning of this thesis, all the studied bandit problems were stationary: the rewards coming from choosing arm $k$ are all \iid{} and follow the same distribution $\nu_k$.
+The next Chapter~\ref{chapter:6} is focussed on a relaxation of this hypothesis, the piece-wise stationary bandit model.
+When we were working on multi-players bandits in Autumn $2017$ \cite{Besson2018ALT}, we left the study of the piece-wise stationary case as a future work, and shortly after we were excited to see that two independent works tackling this question were published, in December $2018$ and in February $2019$.
+Without diving too much into the details, we review here these two works.
+Without a more serious analysis, we conjecture that it should not be difficult to merge the regret lower-bound for stationary \emph{multi-players} MAB (Theorem~\ref{thm:5:BetterLowerBound}) and the lower-bound for \emph{piece-wise stationary} single-player MAB from \cite{Garivier11UCBDiscount}.
+We conjecture that any reasonable decentralized bandit algorithm must suffer a regret at least $M \sqrt{K \Upsilon_T T}$ for $M \leq K$ players, $K$ arms, and $\Upsilon_T$ stationary intervals.
 
-\cite{WeiSrivastava18Distributed}
+On the first hand, piece-wise stationary multi-players MAB can be tackled by extending algorithms developed for the single-player case.
+In \cite{WeiSrivastava18Distributed}, the authors study exactly the same multi-players bandit model as our model, and they propose two distributed algorithms that can efficiently be used by $M \leq K$ players to achieve sub-linear regret for piece-wise stationary problems, essentially by considering it as a harder case of a stationary problem.
+\\
+If $\Upsilon_T$ is bounded by $\Upsilon_T = \bigO{T^{\gamma}}$, for a known $\gamma$ but an unknown $\Upsilon_T$, they prove for instance in Theorem~2 that the expected cumulative regret of their RR-SW-UCB\# algorithm is bounded by $R_T = \bigO{T^{\frac{1+\gamma}{2}} \log(T)}$ (what we call the centralized system regret).
+This bound is of the same order as the bound obtained by the same authors for the SW-UCB\# algorithm in \cite{WeiSrivastava18Abruptly} for the single-player case.
+If $\Upsilon_T = \bigO{T^{\gamma}}$, this bound is comparable to the results obtained for most of the research literature on piece-wise stationary bandits (earlier works like D-UCB in \cite{Garivier11UCBDiscount} have the $\log(T)$ outside the square root, while more recent works all improved this aspect and have the $\log(T)$ in the square root, like for instance our proposals \GLRklUCB{} in \cite{Besson2019GLRT} and presented in Chapter~\ref{chapter:6}).
+Even if their analysis is not explicit regarding the constant, in the case where $\Upsilon_T = \bigO{T^{\gamma}}$ for a known $\gamma$, their regret upper-bound actually matches our conjectured lower-bound, up to a logarithmic factor $\log(T)$.
+\\
+%
+And finally, even though numerical simulations in their papers \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} are interesting and confirm the regret upper-bounds, they do not compare with any other algorithm, and it is left as an interesting future work to study this direction in more details.
+Sadly, a major drawback of their work is that assume that player $j\in\{1,\dots,M\}$ knows its index $j$, and as we explained above this small hypothesis is actually quite strong, as it allows players to be already orthogonal, and it reduces greatly the difficulty of the decentralized bandit problem.
 
-\cite{AlaturLevyKrause19}
+On the other hand, another possible is to extend ideas developed for the adversarial setting.
+In \cite{AlaturLevyKrause19}, the authors essentially consider the piece-wise stationary as an easier case of an adversarial problem. \\
+They build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf.
+Algorithm 1), and on they apparently rediscovered the ``communication trick'' independently from \cite{BoursierPerchet18} to use communications between players to set up a ``coordinator'' player and $M-1$ ``followers'' (running respectively their Algorithms 2 and 3).
+%
+Their key algorithmic technique is to imitate the idealized case where there is full communication between the players.
+Then, to address the no-communication constraint, we enforce the players to keep the same decisions (arms) within long periods of time (blocks).
+This gives them the chance to coordinate between themselves via a simple protocol that uses collisions as a primitive, yet effective manner of communication.\\
+%
+Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
+As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as showed in Theorem~4.1
+(their notation uses $K$ for $M$ the number of players and $N$ for $K$ the number of arms).
+This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for Exp3 for the single-player adversarial setting in \cite{Auer02NonStochastic}.
+Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
+However, we note that the dependency in $M$ is surprisingly better for their result than for our regret upper-bound for \MCTopM-\klUCB, as its regret scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}.\\
+Moreover, at first sight, this idea of ``meta'' arms implies an exponential blow up in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
+the authors provide in Section~4.1 and Lemma~4.1 an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} for a good review).
+%
+In this second work also, we can note the poor experimental section, as the proposed algorithm is only compared against the \MusicalChair{} algorithm, in ``easy'' problems (small number of break-points $\Upsilon_T$).
+Despite being more complicated than other approaches, their algorithm ``\texttt{C \& P}'' should not be too hard to implement ourself, and studying its empirical performances is an interesting future work.
 
+\newpage  % FIXME
 % ----------------------------------------------------------------------------
@@
 
+\newpage  % FIXME
 % ----------------------------------------------------------------------------

### 2019-04-09 2-Chapters/5-Chapter/chapter5.tex
...

@@
 It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in\{1,\dots,K\}$.
-%
-% \TODOL{Explain why the problem is MUCH easier if we have before hand these unique ID!}
+
+\TODOL{Explain why the problem is MUCH easier if we have before hand these unique ID!}
 If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm running on top of a round-Robin phase can achieve order-optimal regret.
@@
 
-Finally, the very recent work of \cite{Zafaruddin2019} is the first one to present experiments done on simulated LTE and 5G channels.
+Finally, the very recent work of \cite{Zafaruddin2019} is the first one to present experiments of reinforcement learning done on simulated LTE and 5G channels.
 The mathematics behind their model are actually quite close to our model, but they explain it in terms of OFDMA and Quality-of-Service (QoS).
@@
 We illustrate this in Table~\ref{table:5:comparisonRhoRandRhoLearn} below, where we consider the same problem as described above in Section~\ref{sub:5:withoutSensing} (Table~\ref{table:5:meanRegretSimulationsNoSensing}).
-The first \texttt{rhoLearn} uses Bayes-UCB for the second-stage learning, while the second uses Exp3,
+The first \texttt{rhoLearn} uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
 and we include both the centralized multi-play version of \klUCB{} and our proposal \MCTopM-\klUCB{} for comparison.
 
-% FIXME run the simulations!
+% FIXME run the simulations with N=100 and not N=4
+% for M in 3 6 9; do DEBUG=False SAVEALL=False NOPLOTS=True M=$M K=9 N_JOBS=-1 N=100 T=10000 make moremultiplayers; read; done
 \begin{table}[ht]
@@
         \hline
-        Centralized multiple-play \klUCB{} & $\mathbf{52 \pm 15}$ & $\mathbf{35 \pm 15}$ & $\mathbf{0}$ \\
+        Centralized multiple-play \klUCB{} & $\mathbf{93 \pm 15}$ & $\mathbf{72 ± 19}$ & $\mathbf{0}$ \\
         \hline
-        \RhoRand-\klUCB{} & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
-        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
-        \texttt{rhoLearn}-\klUCB{} + Exp3 & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
+        \RhoRand-\klUCB{} & $429 \pm 134$ & $2451 ± 338$ & $6044 ± 744$ \\
+        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $473 \pm 141$ & $1277 ± 535$ & $1193 ± 303$ \\
+        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $473 \pm 141$ & $1277 ± 535$ & $1158 ± 133$ \\
+        \texttt{rhoLearn}-\klUCB{} + Exp3 & $548 \pm 176$ & $1880 ± 389$ & $5222 ± 813$ \\
         \hline
-        \MCTopM-\klUCB{} & $134 \pm 6$ & $304 \pm 19$ & $42 \pm 5$ \\
+        \MCTopM-\klUCB{} & $269 \pm 34$ & $401 ± 36$ & $39 ± 3$ \\
         \hline
@@
 
-% FIXME mcuh more details?
-
+% FIXME much more details?
 Other works in the same line of research included \cite{modiDemo2016} implemented using the same USRP testbed as for our demonstration in Chapter~\ref{chapter:4} (Section~\ref{sec:4:gnuradio}),
@@
 More recent works by the same team, \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}, all follow the same approach and include an experimental section that shows the results of an implementation of the proposed algorithms on USRP boards.
-
 \TODOL{I should explain more, no? Who does what and why? So long...}

### 2019-04-09 2-Chapters/5-Chapter/chapter5.tex
Hard to work on this while doing so many other things...

@@
         \hline
-        Centralized multiple-play \klUCB{} & $\mathbf{52 \pm 15}$ & $\mathbf{35 \pm 15}$ & $\mathbf{0}$ \\
+        Centralized multiple-play \klUCB{} & $\mathbf{92 \pm 20}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0}$ \\
         \hline
-        \RhoRand-\klUCB{} & $297 \pm 18$ & $1411 \pm 144$ & $2351 \pm 55$ \\
-        \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$326 \pm 86$} & \textcolor{red}{$1213 \pm 168$} & \textcolor{red}{$2403 \pm 121$} \\
+        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $2351 \pm 55$ \\
+        \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$1422 \pm 1051$} & \textcolor{red}{$9030 \pm 1922$} & \textcolor{red}{$2403 \pm 121$} \\
         \hline
-        \RandTopM-\klUCB{} & $188 \pm 18$ & $671 \pm 159$ & $790 \pm 470$ \\
-        \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$238 \pm 27$} & \textcolor{red}{$778 \pm 156$} & \textcolor{red}{$1098 \pm 153$} \\
+        \RandTopM-\klUCB{} & $268 \pm 45$ & $941 \pm 217$ & $790 \pm 470$ \\
+        \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$688 \pm 614$} & \textcolor{red}{$4256 \pm 2701$} & \textcolor{red}{$1098 \pm 153$} \\
         \hline
-        \MCTopM-\klUCB{} & $134 \pm 6$ & $304 \pm 19$ & $42 \pm 5$ \\
-        \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$210 \pm 66$} & \textcolor{red}{$465 \pm 39$} & \textcolor{red}{$602 \pm 54$} \\
+        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $42 \pm 5$ \\
+        \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$602 \pm 54$} \\
         \hline
-        \Selfish-\klUCB{} & $136 \pm 17$ & $468 \pm 87$ & $1379 \pm 16$ \\
+        \Selfish-\klUCB{} & $250 \pm 34$ & $735 \pm 85$ & $1379 \pm 16$ \\
         \hline
@@
 
+In all this chapter and the previous research literature, another common hypothesis is that all the $M$ players are pursuing the same goal, and are all behaving nicely by following the same algorithm.
+% Efficient utilization of licensed spectrum in the cognitive radio network is challenging due to lack of coordination among the Secondary Users (SUs).
+Distributed algorithms proposed in the literature aim to maximize the network throughput by ensuring orthogonal channel allocation for the SUs.
+However, these algorithms work under the assumption that all the SUs faithfully follow the algorithms which may not always hold due to the decentralized nature of the network.
+In the paper \cite{SawantKumar2018}, the authors study for the first time distributed algorithms that are robust against malicious behavior, also called \emph{jamming attack}.
+They consider both the cases of jammers launching coordinated and uncoordinated attacks, and consider a set of $J$ jammers.
+In the coordinated attack, the jammers select non-overlapping channels to attack in each time slot and can significantly increase the number of collisions for SUs.
+They setup the problem in each scenario as a multi-player bandit and develop algorithm,
+and their analysis shows that when the SUs faithfully implement proposed algorithms, the regret is constant with high probability.
+They validate their claims through exhaustive synthetic experiments and also through a realistic USRP.
+Their synthetic experiments consider different cases, for different values of $K$ the number of channels, $M$ the number of players and $J$ the number of jammers.
+
 \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 
-Learning to Coordinate in a Decentralized Cognitive Radio Network in Presence of Jammers
-Suneet Sawant, Rohit Kumar, Manjesh K. Hanawal, Sumit J. Darak
-https://arxiv.org/abs/1803.06810
-\cite{SawantKumar2018}
-
 Note that we have not yet implemented this extension in SMPyBandits \cite{SMPyBandits}, but we are interested to do it, and it is one of the major future work left on our library\footnote{See the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/120}}}.
+It would not be too difficult to do, we just lacked some time.
 
@@
         \hline
-        Centralized multiple-play \klUCB{} & $\mathbf{93 \pm 15}$ & $\mathbf{72 ± 19}$ & $\mathbf{0}$ \\
+        Centralized multiple-play \klUCB{} & $\mathbf{92 \pm 20}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0}$ \\
         \hline
-        \RhoRand-\klUCB{} & $429 \pm 134$ & $2451 ± 338$ & $6044 ± 744$ \\
-        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $473 \pm 141$ & $1277 ± 535$ & $1193 ± 303$ \\
-        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $473 \pm 141$ & $1277 ± 535$ & $1158 ± 133$ \\
-        \texttt{rhoLearn}-\klUCB{} + Exp3 & $548 \pm 176$ & $1880 ± 389$ & $5222 ± 813$ \\
+        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6044 \pm 744$ \\
         \hline
-        \MCTopM-\klUCB{} & $269 \pm 34$ & $401 ± 36$ & $39 ± 3$ \\
+        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1193 \pm 303$ \\
+        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1158 \pm 133$ \\
+        \texttt{rhoLearn}-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5222 \pm 813$ \\
+        \hline
+        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $39 \pm 3$ \\
         \hline

### 2019-04-09 2-Chapters/5-Chapter/chapter5.tex
Done for two simulations for Chapter 5

@@
 
-\TODOL{Include a simulation to show the behavior of \rhoRand{} vs \rhoRandEst, and \RandTopM{} and \MCTopM{} with prior knowledge of $M$ vs the ``estimate'' versions.
-
-FIXME Je dois refaire ces simulations, les chiffres ici sont avec $T=1000$ et $N=4$ répétitions (juste faites vite fait dans le train, je dois les lancer sur la machine du bureau).
-
+% \TODOL{Include a simulation to show the behavior of \rhoRand{} vs \rhoRandEst, and \RandTopM{} and \MCTopM{} with prior knowledge of $M$ vs the ``estimate'' versions.}
+% Je dois refaire ces simulations, les chiffres ici sont avec $T=1000$ et $N=4$ répétitions (juste faites vite fait dans le train, je dois les lancer sur la machine du bureau).
 % for M in 3 6 9; do DEBUG=False SAVEALL=True NOPLOTS=True M=$M K=9 N_JOBS=-1 N=100 T=10000 make moremultiplayers; read; done
-}
 
@@
         \hline
-        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $2351 \pm 55$ \\
-        \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$1422 \pm 1051$} & \textcolor{red}{$9030 \pm 1922$} & \textcolor{red}{$2403 \pm 121$} \\
+        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 ± 1035$ \\
+        \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$1422 \pm 1051$} & \textcolor{red}{$9030 \pm 1922$} & \textcolor{red}{$7264 ± 1009 $} \\
         \hline
-        \RandTopM-\klUCB{} & $268 \pm 45$ & $941 \pm 217$ & $790 \pm 470$ \\
-        \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$688 \pm 614$} & \textcolor{red}{$4256 \pm 2701$} & \textcolor{red}{$1098 \pm 153$} \\
+        \RandTopM-\klUCB{} & $268 \pm 45$ & $941 \pm 217$ & $437 ± 367$ \\
+        \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$688 \pm 614$} & \textcolor{red}{$4256 \pm 2701$} & \textcolor{red}{$1155 ± 544$} \\
         \hline
-        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $42 \pm 5$ \\
-        \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$602 \pm 54$} \\
+        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 ± 13$ \\
+        \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$618 ± 29$} \\
         \hline
-        \Selfish-\klUCB{} & $250 \pm 34$ & $735 \pm 85$ & $1379 \pm 16$ \\
+        \Selfish-\klUCB{} & $250 \pm 34$ & $735 \pm 85$ & $3010 ± 472$ \\
         \hline
@@
 
-% FIXME run the simulations with N=100 and not N=4
+% DONE run the simulations with N=100 and not N=4
 % for M in 3 6 9; do DEBUG=False SAVEALL=False NOPLOTS=True M=$M K=9 N_JOBS=-1 N=100 T=10000 make moremultiplayers; read; done
@@
         \hline
-        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6044 \pm 744$ \\
+        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 ± 1035$ \\
         \hline
-        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1193 \pm 303$ \\
-        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1158 \pm 133$ \\
-        \texttt{rhoLearn}-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5222 \pm 813$ \\
+        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1416 ± 347$ \\
+        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 ± 331$ \\
+        \texttt{rhoLearn}-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5134 ± 976$ \\
         \hline
-        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $39 \pm 3$ \\
+        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 ± 13$ \\
         \hline

### 2019-04-10 2-Chapters/5-Chapter/chapter5.tex
Done for now. Oh boy what a messy day

@@
         Centralized multiple-play \klUCB{} & $\mathbf{92 \pm 20}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0}$ \\
+        \Selfish-\klUCB{} & $250 \pm 34$ & $735 \pm 85$ & $3010 \pm 472$ \\
         \hline
-        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 ± 1035$ \\
-        \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$1422 \pm 1051$} & \textcolor{red}{$9030 \pm 1922$} & \textcolor{red}{$7264 ± 1009 $} \\
+        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 \pm 1035$ \\
+        \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$1422 \pm 1051$} & \textcolor{red}{$9030 \pm 1922$} & \textcolor{red}{$7264 \pm 1009 $} \\
         \hline
-        \RandTopM-\klUCB{} & $268 \pm 45$ & $941 \pm 217$ & $437 ± 367$ \\
-        \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$688 \pm 614$} & \textcolor{red}{$4256 \pm 2701$} & \textcolor{red}{$1155 ± 544$} \\
+        \RandTopM-\klUCB{} & $268 \pm 45$ & $941 \pm 217$ & $437 \pm 367$ \\
+        \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$688 \pm 614$} & \textcolor{red}{$4256 \pm 2701$} & \textcolor{red}{$1155 \pm 544$} \\
         \hline
-        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 ± 13$ \\
-        \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$618 ± 29$} \\
-        \hline
-        \Selfish-\klUCB{} & $250 \pm 34$ & $735 \pm 85$ & $3010 ± 472$ \\
+        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 \pm 13$ \\
+        \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$618 \pm 29$} \\
         \hline
@@
 One can observe in Table~\ref{table:5:meanRegretSimulationsEstimatingM} the empirical performances of different algorithms in an example problem, in each case of low, medium and maximum number of players ($M=3,6,9$ for $K=9$ arms).
-The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar costs on their regret when they have to estimate $M$ on the fly.
+The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar costs on their regret when they have to estimate $M$ on the fly (written ``Estimate $M$'' in red, in Table~\ref{table:5:meanRegretSimulationsEstimatingM}).
 Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.
-For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large (on the three algorithms and two cases, about $+30\%$), and consistent on the different algorithm.
-The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proved to be $\smallO{\log(T)}$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret (written ``Estimate $M$'' in Table~\ref{table:5:meanRegretSimulationsEstimatingM}).
+For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large (on the three algorithms and two cases, about $\times 3.4$), and consistent on the different algorithm.
+The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proved to be $\smallO{\log(T)}$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
 In this extreme case, \RhoRand{} and its extension both perform very closely, as expected.
 %
-Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension \MCTopM-\klUCB.
+Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension of our proposal \MCTopM-\klUCB.
 
@@
 
-\paragraph{The ``communication trick''}
+\paragraph{The ``communication trick''.}
 %
@@
 
-\paragraph{Estimating collisions through uniform exploration}
+\paragraph{Estimating collisions through uniform exploration.}
 %
@@
 
-\subsubsection{Additionnal numerical simulations}
+\paragraph{Additionnal numerical simulations.}
 
-\TODOL{Do some basic simulations and show :
-
-    - Selfish-UCB from \cite{Besson2018ALT}.
-    - Improved Musical Chair from \cite{LugosiMehrabian18}.
-    - SIC-MMAB from \cite{BoursierPerchet18}.
-}
-
+To illustrate the difficulty of the ``no sensing'' case, we illustrate here the performances of different algorithms designed specifically for this setting.
 As above, we consider a bandit problem with $K=9$ arms, of means $0.1,0.2,\dots,0.8,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
 We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
+The centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, but it is interesting to also add it here, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, because it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
+%
+For the two other algorithms, we use the parameters advised in all papers: for \textsc{Sic-MMAB}, we used $T_0 = \lceil K \log(T) \rceil$ as advised by the authors.
+For Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis.
+We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear (\eg, $12150$ for $T=10000$).
+
+% for T in 10000; do for K in 9; do for N in 4 100; do for M in 3 6 9; do DEBUG=False N_JOBS=4 N=$N K=$K M=$M T=$T make --noFreeSMS moremulti && cp logs/main_multiplayers_more_py3_log.txt logs/main_multiplayers_more_py3_log__K${K}_N${N}_M${M}_T${T}_$(date +"%d-%m-%Y_%H-%M-%S").txt && FreeSMS.py "Done for simulations for $K arms, $M players, T=$T N=$N and lots algorithms for experiments for Chapter 5 of my PhD.
+Love from Rennes."; done; done; done; done
+
+We give in Table~\ref{table:5:meanRegretSimulationsNoSensing}, we give the mean regret obtained for these different algorithms, and we observe a drastic difference between the unrealistic centralized algorithm, which achieves a very small regret, the heuristic \Selfish-\klUCB{} which achieve small mean regret (but is small to fail in theory and in some instances), and the two other algorithms.
+We were unable to find any bug in our implementation of Improved \MusicalChair{} from \cite{LugosiMehrabian18} and all the different tuning of $c$ (or $g$) explored gave the same disappointing result (linear regret).
+The \textsc{Sic-MMAB} algorithm performs better than the \Selfish{} heuristic for the extreme case of $M=K$, but its large regret in this case shows that the orthogonalization protocol developed by \cite{BoursierPerchet18} is not fast to converge (for illustration, for the same problem for $M=K=9$ for the ``sensing'' case, \MCTopM-\klUCB{} achieves a mean regret of about $40$, two orders of magnitude smaller!).
+%
+Further empirical evaluation would be needed to fully understand the situation, and a first future work would be to fix the algorithm proposed by \cite{LugosiMehrabian18} and illustrate in some problems that it can indeed achieve sub-linear regret (maybe it does but only for large horizons, even if we did try larger values of $T$ like up-to $T=200000$).
 
@@
         \hline
-            \Selfish-\klUCB{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+        Centralized Multiple play \klUCB{} & $\mathbf{91 \pm 21}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0 \pm 0}$ \\
+        \Selfish-\klUCB{} & $249 \pm 35$ & $728 \pm 94$ & $2953 \pm 501$ \\
         \hline
-            Improved \MusicalChair{} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
-            \hline
-            \textsc{Sic-MMAB} & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ & $\mathbf{XXX \pm XXX}$ \\
+        \textsc{Sic-MMAB} & $1705 \pm 340$ & $3915 \pm 300$ & $1713 \pm 52$ \\
+        Improved \MusicalChair{} & $\textcolor{red}{12149 \pm 60}$ & $\textcolor{red}{22341 \pm 77}$ & $\textcolor{red}{27462 \pm 85}$ \\
         \hline
@@
 
-\TODOL{Give interpretation from these results!}
+% % M=3
+% SIC-MMAB(kl-UCB, $T_0=226$) : 1705 ± 340
+% TSN($T_{RH} = 140, T_{SH} = 1.53e+04$) : 9001 ± 2
+% Selfish-kl-UCB : 249 ± 35
+% CentralizedMultiplePlay(kl-UCB) : 91 ± 21
+% MusicalChair($T_0=1000$) : 1410 ± 641
+% MusicalChair($T_0=450$) : 1268 ± 1110
+% MusicalChair($T_0=900$) : 1355 ± 681
+% MusicalChair($T_0=1350$) : 1791 ± 561
+% MusicalChair($T_0=95117$) : 12145 ± 59
+% MusicalChair($T_0=118764$) : 12142 ± 59
+% MusicalChair($T_0=220257$) : 12146 ± 59
+% MCNoSensing($M=3$, $T=10000$, $c=1$, $g=235$) : 12149 ± 60
+% MCNoSensing($M=3$, $T=10000$, $c=10$, $g=2.35e+03$) : 12148 ± 66
+% MCNoSensing($M=3$, $T=10000$, $c=128$, $g=3.01e+04$) : 12149 ± 56
+
+% % M=6
+
+% SIC-MMAB(kl-UCB, $T_0=226$) : 3915 ± 300
+% TSN($T_{RH} = 140, T_{SH} = 1.53e+04$) : 9008 ± 5
+% Selfish-kl-UCB'  : 728 ± 94
+% CentralizedMultiplePlay(kl-UCB)' : 70 ± 16
+% MusicalChair($T_0=1000$) : 3130 ± 1583
+% MusicalChair($T_0=450$) : 2690 ± 2473
+% MusicalChair($T_0=900$) : 3015 ± 1839
+% MusicalChair($T_0=1350$) : 3403 ± 1202
+% MusicalChair($T_0=95117$) : 22363 ± 89
+% MusicalChair($T_0=118764$) : 22352 ± 74
+% MusicalChair($T_0=220257$) : 22352 ± 75
+% MCNoSensing($M=6$, $T=10000$, $c=1$, $g=235$) : 22341 ± 77
+% MCNoSensing($M=6$, $T=10000$, $c=10$, $g=2.35e+03$) : 22365 ± 80
+% MCNoSensing($M=6$, $T=10000$, $c=128$, $g=3.01e+04$) : 22349 ± 84
+
+% % M=9
+% SIC-MMAB(kl-UCB, $T_0=226$) : 1713 ± 52
+% TSN($T_{RH} = 140, T_{SH} = 1.53e+04$) : 39 ± 18
+% Selfish-kl-UCB : 2953 ± 501
+% CentralizedMultiplePlay(kl-UCB) : 0 ± 0
+% MusicalChair($T_0=1000$) : 2775 ± 27
+% MusicalChair($T_0=450$) : 1455 ± 1059
+% MusicalChair($T_0=900$) : 2506 ± 30
+% MusicalChair($T_0=1350$) : 3737 ± 36
+% MusicalChair($T_0=95117$) : 27466 ± 90
+% MusicalChair($T_0=118764$) : 27480 ± 80
+% MusicalChair($T_0=220257$) : 27467 ± 95
+% MCNoSensing($M=9$, $T=10000$, $c=1$, $g=235$) : 27462 ± 85
+% MCNoSensing($M=9$, $T=10000$, $c=10$, $g=2.35e+03$) : 27463 ± 101
+% MCNoSensing($M=9$, $T=10000$, $c=128$, $g=3.01e+04$) : 27462 ± 76
 
@@
         \hline
-        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 ± 1035$ \\
+        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 \pm 1035$ \\
         \hline
-        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1416 ± 347$ \\
-        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 ± 331$ \\
-        \texttt{rhoLearn}-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5134 ± 976$ \\
+        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1416 \pm 347$ \\
+        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 \pm 331$ \\
+        \texttt{rhoLearn}-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5134 \pm 976$ \\
         \hline
-        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 ± 13$ \\
+        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 \pm 13$ \\
         \hline

### 2019-04-16 1-Preamble/4-Acknowledgement/acknowledgement.tex
Almost finished the acknowledgements...

@@
 
-\TODO{To finish}
+\TODOE{FIXME à finir, en septembre !}
 % See https://perso.crans.org/besson/phd/thanks/
 
-And I would like to acknowledge and thank the following people.
+I would like to acknowledge and thank the following people, and also acknowledge the various funding who supported our works during the last three years.
 
-My PhD advisors, \href{https://moychristophe.wordpress.com/}{Christophe Moy} and \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}.
+My first thoughts go to my two PhD advisors.
+%
+Firstly, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in the IETR lab and in CentraleSupélec campus of Rennes, a very nice and warm environment not only for research but also for day-to-day life.
+Christophe was very supportive, on all aspects of the PhD work and he was also helpful during difficult moments in the last three years.
+Many thanks for being so nice and dynamic, and always comprehensive.
+Christophe was immensely helpful for his expertise on wireless communications and engineering, and especially on cognitive radio.
+Regardless of what we did together, debugging a difficult GNU Radio block, writing a paper together by night in a plane coming back from India, chatting around a friendly barbecue or playing ``palets'' together along the ocean in Vannes, I was extremely lucky to collaborate with and be around Christophe.
+We had a nice first contact when we met in Spring 2015 in Mahindra Ecole Centrale, in Hyderabad, India, where I taught for one year, and I am honored to have pursued my PhD under his supervision and to have worked with him, and I am also happy to have lived great human moments along with Christophe since 2016.
+%
+Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years.
+Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.
+Emilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep the highest possible rigour in all parts of our exchanges.
+I was lucky to be able to visit the SequeL team at Inria Lille on regular occasions, and to work in close collaboration with Emilie during these visits.
+Intensive white board sessions proved to be fruitful, giving the research work that I am the most proud of \cite{Besson2018ALT}, and rich research papers that join her rigorous expertise in statistical learning and mathematical analysis of bandit algorithms with my passion for numerical experiments, simulations and algorithm designs \cite{Besson2018ALT,Besson2018DoublingTricks,Besson2019GLRT}.
 
-My family: my parents, Christophe and Patricia, my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian} and my younger brother Fabian, Marine, my grand-parents, Jeannine, Jean-Marc and Jacqueline, my uncle and aunt Jacques, Bernadette, and my cousin Lara.
+Then I would like to thank the dear member of my PhD jury.
+Many thanks to \href{https://sites.google.com/site/vianneyperchet/}{Vianney Perchet} and \href{http://researchers.lille.inria.fr/~mitton/}{Nathalie Mitton} for accepting to review my thesis during the summer 2019, and for their very valuable comments.
+Thanks to Patrick Maillé and Raphaël Féraud for following our work since Spring 2017,
+and to Richard Combes as well for being part of my jury.
 
-My closest friends, sorted in an absurd chronological list: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia, Julia, Pierre, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément, Jill-Jênn, Alain, Claire, Édouard, Loïc, Damien, Kumudham, Priscilla, Thibault, Sébastien, Sélim, Valentin, Marine, Johanne, Lola.
+Of course, my thought then go to my family: my parents, Christophe and Patricia, my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian} and my younger brother Fabian, Marine, my grand-parents, Jeannine, Jean-Marc and Jacqueline, my uncle and aunt Jacques, Bernadette, and my cousin Lara.
+\TODOL{Une belle phrase sur leur soutien ? En particulier les parents et l'éducatin qu'ils nous ont donnés, Florian et son soutien inconditionnel et son rôle de modèle, Fabian pour ?}
 
-My wonderful colleagues during these three years: Rémi, Pascal, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin, Vincent, Rami, Muhammad, Cristo, Adrien, Esteban, Ali, Éloïse, Bastien, Corentin, Georgios, in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas in Lille.
+I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{Sorry if I forgot you, }: Florian, Fabian, Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia from Briançon, Mayotte and Pierre from Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien from Cachan, Kumudham and Priscilla from India, Thibault, Sébastien, Sélim, Valentin from Cachan again, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola from Rennes.
+In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in Hungary/Ausria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Nantes\dots I loved traveling with you and hope to continue!
 
-Thanks to some people for their support: Karine, Jeannine, Grégory, Anne and others at CentraleSupélec Rennes; François, David, David at ENS de Rennes, Romaric at ENSAI, Amélie at Inria Lille, Amélie at Lycée Chateaubriand.
-During my studies at ENS Cachan (now Paris-Saclay), I am thankful to all the great professors who taught me, including Frédéric, Alain and Claudine in the maths department, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
+Thanks to my wonderful colleagues during these three years: Rémi, Pascal, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin, Vincent, Rami, Muhammad, Cristo, Adrien, Esteban, Ali, Éloïse, Nabil, Bastien, Corentin, Georgios, Morgane in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas in Lille.
+%
+Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile and others at CentraleSupélec campus of Rennes; François, David, David at ENS de Rennes, Romaric at ENSAI, Amélie at Inria Lille, Amélie at Lycée Chateaubriand, and Gilbert at Lycée Joliot-Curie.
+Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially Martine and J-B, your daily smile is priceless.
+%
+During my studies at ENS de Cachan, I am thankful to all the great professors who taught me and pushed me into studying statistical learning and computer science, including Frédéric, Alain and Claudine in the maths department, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
 
 % TODO financial thanks
-
 This work is supported by
@@
 
-Some friends thanked me in their own PhD thesis, and it is a pleasure to do the same.
+Some friends thanked me in their own PhD thesis, and it is a pleasure to do the same, and cross-link to their own PhD thesis.
 In a chronological order, I thank
 %,
-\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{http://jiji.cat/cat.pdf}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://drive.google.com/open?id=17CBibM8ZlzIbyVxpxktK3j5rNePALUs8}{thesis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.
+\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{http://jiji.cat/cat.pdf}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.
 
-% Je souhaite terminer ces remerciements par une pensée toute particulière pour Lola, qui m'a soutenu, qui m'a émerveillé par son enthousiasme et qui m'a réconforté par son amour durant la dernière année.
-Je t'aime chaton, impressionnant !
-Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó, me sorprendió con su entusiasmo y me consoló con su amor durante el año pasado.
-\textcolor{white}{Yo te quiero gatita, ¡impresivo!}
+% Je souhaite terminer ces remerciements par une pensée toute particulière pour Lola, qui m'a soutenu quotidiennement, qui m'a émerveillé par son enthousiasme et qui m'a réconforté par son amour, durant la dernière année.
+Je t'aime Lola, veut-tu vivre avec moi ?
+Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con su amor, durante el año pasado.
+\textcolor{white}{Je t'aime Lola, veux-tu vivre avec moi ?}
 

### 2019-04-16 1-Preamble/4-Acknowledgement/acknowledgement.tex
Added my latest publications (just quote it, in the introduction or in the conclusion?)

--- 2019-02-12 1-Preamble/4-Acknowledgement/acknowledgement.tex
@@
 
-\TODO{To finish}
+% \TODOL{FIXME à finir, en septembre !}
 % See https://perso.crans.org/besson/phd/thanks/
 
-And I would like to acknowledge and thank the following people.
+I would like to acknowledge and thank the following people, and also acknowledge the various funding who supported our works during the last three years.
 
-My PhD advisors, \href{https://moychristophe.wordpress.com/}{Christophe Moy} and \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}.
+My first thoughts go to my two PhD advisors.
+%
+Firstly, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in IETR lab and in CentraleSupélec campus of Rennes, a very nice and warm environment for research and day-to-day life.
+Christophe was very supportive on all aspects of the PhD work, and he was also helpful during difficult moments in the last three years.
+Many thanks for being so nice and dynamic\footnote{We had a nice first contact when we met in Spring 2015 in Mahindra Ecole Centrale, in Hyderabad, India, where I taught for one year, and I am honored to have pursued my PhD under his supervision and to have worked with him, and I am also happy to have lived great human moments along with Christophe since 2016.}, and always comprehensive.
+Christophe was helpful for his expertise on wireless communications and engineering, and especially on cognitive radio.
+Regardless of what we did together, whether it was debugging a difficult GNU Radio block \cite{Besson2018ICT,Besson2019WCNC}, writing a paper together by night in a plane coming back from India \cite{MoyBesson2019}, chatting around a friendly barbecue or playing ``palets'' together along the ocean in Vannes, I was lucky to collaborate with and be around Christophe.\\
+%
+Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years\footnote{Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.}.
+Émilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep the highest possible rigour in all parts of our exchanges.
+I was lucky to be able to visit the SequeL team at Inria Lille on regular occasions, and to work in close collaboration with Émilie during these visits\footnote{Merci à Romain et toi pour votre accueil chaleureux, et aussi merci à Florian, Nicolas et Édouard.}.
+Intensive white board sessions proved to be fruitful, giving the research work that I am the most proud of \cite{Besson2018ALT}, and rich research papers that join her rigorous expertise in statistical learning and mathematical analysis of bandit algorithms with my passion for numerical experiments, simulations and algorithm designs \cite{Besson2018ALT,Besson2018DoublingTricks,Besson2019GLRT}.
 
-My family: my parents, Christophe and Patricia, my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian} and my younger brother Fabian, Marine, my grand-parents, Jeannine, Jean-Marc and Jacqueline, my uncle and aunt Jacques, Bernadette, and my cousin Lara.
+Then I would like to thank the dear members of my PhD jury.
+I was lucky to be able to discuss with them on regular occasions, and these discussions have been very useful for some of our works.
+Many thanks to \href{https://sites.google.com/site/vianneyperchet/}{Vianney Perchet} and \href{http://researchers.lille.inria.fr/~mitton/}{Nathalie Mitton} for accepting to review my thesis during the summer 2019, and for their very valuable comments.
+Thanks to Patrick Maillé and Raphaël Féraud for following our work since Spring 2017,
+and to Richard Combes as well for being part of my jury.
+Some of their articles were strong inspirations for our own works, and I wish to also thank their co-authors, and in particular the Masters or PhD students, and useful discussions we had whenever we met, especially thanks to Claire Vernade, Réda Alami and Étienne Boursier, and others I could not meet (\eg, Viktor Toldov, Robin Allesiardo).
 
-My closest friends, sorted in an absurd chronological list: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia, Julia, Pierre, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément, Jill-Jênn, Alain, Claire, Édouard, Loïc, Damien, Kumudham, Priscilla, Thibault, Sébastien, Sélim, Valentin, Marine, Johanne, Lola.
+Of course on a more personal level, my thoughts are strongly dedicated to my family.
+I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model, who has been of a constant support during all my studies, and whose serious work kept pushing me further.
+I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life.
+Very warm thanks then go to Marine, to my grand-parents, Jeannine, Jean-Marc and Jacqueline, my uncle Jacques and aunt Bernadette, and my cousins.
+Finally, I thank sincerely my parents, Christophe and Patricia, for being the best parents since the last 30 years, and for their vital moral support.
+\emph{Merci de nous avoir inculqué une si forte curiosité et d'être si gentils !}
 
-My wonderful colleagues during these three years: Rémi, Pascal, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin, Vincent, Rami, Muhammad, Cristo, Adrien, Esteban, Ali, Éloïse, Bastien, Corentin, Georgios, in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas in Lille.
+I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
+And due to space constraints, I had to chose quite arbitrarily who to include\dots}: Florian, Fabian, Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia from Briançon, Mayotte and Pierre from Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément from Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also from Cachan, Kumudham and Priscilla from India, Thibault, Sébastien, Sélim, Valentin again from Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola from Rennes.
+In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
+I loved traveling with you and hope to continue!
 
-Thanks to some people for their support: Karine, Jeannine, Grégory, Anne and others at CentraleSupélec Rennes; François, David, David at ENS de Rennes, Romaric at ENSAI, Amélie at Inria Lille, Amélie at Lycée Chateaubriand.
-During my studies at ENS Cachan (now Paris-Saclay), I am thankful to all the great professors who taught me, including Frédéric, Alain and Claudine in the maths department, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
+Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas in Lille.
+I want to thank especially the ``coffee club'' members$^*$ in Rennes, for so many interesting discussions who where sometimes the highlights of our day-to-day life at the office.
+
+Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline and others at CRIStAL in Lille.
+Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially Martine and J-B, your daily smile is priceless!
+I warmly thank David P.
+for his trust, and François, David P.
+and David C.
+at ENS de Rennes, and Romaric at ENSAI : I loved every single of the $3 \times 64$ hours I spent teaching with you in the last three years.
+I am grateful to Amélie at Lycée Chateaubriand and Gilbert at Lycée Joliot-Curie for useful exchanges.
+%
+I am thankful to all the great professors who taught me during my studies at ENS de Cachan, between 2011 and 2016, and for some who pushed me into teaching, and studying statistical learning and computer science, including Frédéric, Alain and Claudine in the maths department, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
 
 % TODO financial thanks
-
-This work is supported by
+This work and my PhD is supported by
 the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{N ANR-16-CE40-0002}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
 by R\'egion Bretagne, France,
-the French Ministry of Higher Education and Research (MENESR),
+by CNRS under the PEPS project BIO (in 2017),
+by the French Ministry of Higher Education and Research (MENESR),
 by \'Ecole Normale Sup\'erieure de Paris-Saclay.
@@
 and by Ministry of Higher Education and Research, Brittany and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
+I was also lucky to be well received very frequently at Inria Lille Nord Europe, and to be hosted at CentraleSupélec campus of Rennes.
 
-I would also like to thank some things. First, the Earth, then my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents.
-My \href{https://perso.crans.org/besson/cuisine/}{culinary friends}: cheese, chocolate, coffee, vegetables etc!
-My \href{https://perso.crans.org/besson/}{technological friends}: Bash, Python, XUbuntu, Firefox, Konsole, Visual Studio Code etc.
+I would also like to thank some unusual things.
+Some people thank God in their thesis, I feel that I prefer to thank the Earth, and apologize for the too many times I took a plane for a conference abroad, and in general about the environment impact of this research work.
+I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} who gave me energy (\eg, cheese, chocolate, coffee, vegetables etc), and finally my \href{https://perso.crans.org/besson/}{favorite technological tools}: git, GitHub and Bitbucket, \LaTeX, Bash and Python, GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other tools.
 
-Some friends thanked me in their own PhD thesis, and it is a pleasure to do the same.
+In the last few years, some friends thanked me in their own PhD thesis, and it is a pleasure to do the same, and cross-link to their own PhD thesis.
 In a chronological order, I thank
 %,
-\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{http://jiji.cat/cat.pdf}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://drive.google.com/open?id=17CBibM8ZlzIbyVxpxktK3j5rNePALUs8}{thesis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.
+\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{Que je remercie deux fois plus que ce qu'il me remerciant dans sa propre thèse !
+Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.
 
-% Je souhaite terminer ces remerciements par une pensée toute particulière pour Lola, qui m'a soutenu, qui m'a émerveillé par son enthousiasme et qui m'a réconforté par son amour durant la dernière année.
-Je t'aime chaton, impressionnant !
-Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó, me sorprendió con su entusiasmo y me consoló con su amor durante el año pasado.
-Yo te quiero gatita, ¡impresivo!
+% I would like to conclude these acknowledgements by a special thought toward Lola, who was of great support on a daily basis, who marvelled me by her enthusiasm and dynamism, and who comforted me with her laughs and love, in the last year.
+% \textcolor{white}{I love you Lola, would you like to live with me?}
+%
+% Je souhaite terminer ces remerciements par une pensée toute particulière pour Lola, qui m'a soutenu quotidiennement, qui m'a émerveillé par son enthousiasme et qui m'a réconforté par son amour, durant la dernière année.
+% \textcolor{white}{Je t'aime Lola, veut-tu vivre avec moi ?}
+%
+\emph{Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con sus risas y su amor, durante el año pasado.}
+\textcolor{white}{Te quiero, Lola, ¿quieres vivir conmigo?}
 

### 2019-04-16 1-Preamble/4-Acknowledgement/acknowledgement.tex
Small changes on Chapter:4 after the talk of yesterday.

@@
 I would also like to thank some unusual things.
-Some people thank God in their thesis, I feel that I prefer to thank the Earth, and apologize for the too many times I took a plane for a conference abroad, and in general about the environment impact of this research work.
-I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} who gave me energy (\eg, cheese, chocolate, coffee, vegetables etc), and finally my \href{https://perso.crans.org/besson/}{favorite technological tools}: git, GitHub and Bitbucket, \LaTeX, Bash and Python, GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other tools.
+Some people thank God in their thesis, I feel that I prefer to thank the Earth, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
+I'll try hard to do better from now on!
+I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} who gave me energy (\eg, cheese, chocolate, coffee, vegetables etc), and finally my \href{https://perso.crans.org/besson/}{favorite technological tools}: git, GitHub and Bitbucket, \LaTeX, Bash and Python, GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other free and open-source tools.
 

### 2019-04-16 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Small changes on Chapter:4 after the talk of yesterday.

@@
 This solution could be added in a straightforward and costless manner in most IoT networks, such as LoRaWAN networks, just by adding this feature at the higher level of the MAC layer, in all or only some of the devices, without any modification on the network side, and no signalling overhead for the devices.
+
+\paragraph{Related work.}
+% FIXME
+\TODOL{Est-ce que je fais ça ici ou à la fin du chapitre 5 ? Ou peut-être en chapitre 1, ça ferait du contenu au chapitre 1...}
+We evoke quickly here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
+\cite{darak2016bayesian}
+\cite{Darak16}
+\cite{modiDemo2016}
+\cite{kumar2016two}
+\cite{KumarYadav2018}
+\cite{SawantKumar2018}
+\cite{JoshiKumar2018}
 
@@
 %
-As shown below in Figure~\ref{fig:42:screenshotDemoYouTube},
-we realized and released a $6$-minute \textbf{video} to show our demonstration online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+As depicted in Figure~\ref{fig:42:screenshotDemoYouTube},
+we realized a $6$-minute \textbf{video} to sum-up our demonstration and advertise our work online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
 The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
 It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
+% \textcolor{white}{Of course, the video was not seen much, because our work is useless! Yay!}
 

### 2019-04-16 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
Small changes on Chapter:4 after the talk of yesterday.

--- 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
@@
-\begin{figure}[!b]
+For the curious reader, we wanted to include here an illustration and a description of the three components of our demonstration, presented above in Section~\ref{sec:4:gnuradio}.
+
+Figure~\ref{fig:4app:USRP_TX_PU__v1__simple_grc} shows the \textbf{random traffic generator} flow-graph.
+\texttt{Generator} is the only hand-written block, which is configured with a list of active channels, a list of occupancy rate, and constants about the PHY layer (preamble length and data length).
+It uses one USRP only for emitting data (``Sink'' mode).
+
+\begin{figure}[!h]
 	\centering
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_PU__v1__simple_grc.png}
-    \caption{The \textbf{random traffic generator}.
-\texttt{Generator} is the only hand-written block, which is configured with a list of active channels, a list of occupancy rate, and constants about the PHY layer (preamble length and data length).
-It uses one USRP only for emitting data (``Sink'' mode).}
+    \caption{The \textbf{random traffic generator} flow-graph.}
     \label{fig:4app:USRP_TX_PU__v1__simple_grc}
@@
 
-\begin{figure}[!b]
+The \textbf{IoT gateway} is presented in Figure~\ref{fig:4app:USRP_TX_PU__v1__simple_grc}.
+%
+The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length).
+All blocks need to know constants about the PHY layer (preamble length and data length).
+The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).
+
+\begin{figure}[!h]
 	\centering
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_RX_BTS__v1__simple_grc.png}
-    \caption{The \textbf{IoT gateway}.
-The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length).
-All blocks need to know constants about the PHY layer (preamble length and data length).
-The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
+    \caption{The \textbf{IoT gateway} flow-graph.}
     \label{fig:4app:USRP_RX_BTS__v1__simple_grc}
@@
 
-\begin{figure}[!b]
+Figure~\ref{fig:4app:USRP_TX_SU__v1__simple_grc} below shows the \textbf{IoT dynamic device} flow-graph.
+The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels.
+Most blocks need to know constants about the PHY layer (preamble length and data length).
+Any of the IoT dynamic devices use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes)
+
+\begin{figure}[!h]
 	\centering
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_SU__v1__simple_grc.png}
-    \caption{The \textbf{IoT dynamic object}.
-The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels.
-Most blocks need to know constants about the PHY layer (preamble length and data length).
-Any of the IoT dynamic objects use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
+    \caption{The \textbf{IoT dynamic device} flow-graph.}
     \label{fig:4app:USRP_TX_SU__v1__simple_grc}

### 2019-04-16 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Small changes on Chapter:4 after the talk of yesterday.

@@
 
-\TODOL{Remove this part or compress: I will remove as soon as I have clear equations in Chapter~\ref{chapter:2} as references for the \UCB{} algorithm.}
+% %-----------------------------------------------------------------
+% \subsection{The \UCB{} algorithm as a building block for different heuristics}
+% \label{sub:43:MABalgo}
+% % ----------------------------------------------------------------------
 
-%-----------------------------------------------------------------
-\subsection{The \UCB{} algorithm as a building block for different heuristics}
-\label{sub:43:MABalgo}
+% Before presenting our proposed heuristics, we remind here the \UCB{} bandit algorithm \cite{Auer02}.
+% We include it again to facilitate the understanding of the heuristics, which use \UCB{} as a building block.
+
+% % % --- --- --- --- --- --- --- --- --- ---
+% \paragraph{The \UCB{} algorithm}\label{sub:43:algoUCB}
+
+% More formally, for one device, let $N_k(t)$ be the number of times the channel $k$ (for $k\in \llbracket 1, K \rrbracket$) was selected up-to time $t-1$, for $t\geq 0$
+% for any $t\in\mathbb{N}$,
+% \begin{equation}\label{eq:43:Nkt}
+% 	N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(C(\tau) = k),
+% \end{equation}
+% The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
+% \begin{equation}\label{eq:43:mukt}
+% 	\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(C(\tau) = k).
+% \end{equation}
+% where $r_{k}(t)$ is the reward obtained after transmission in channel $k$ at time $t$ ($1$ for a successful transmission, and $0$ otherwise)
+% %
+% A \emph{confidence} term $B_k(t)$ is given by \cite{Auer02},
+% \begin{equation}\label{eq:43:Bkt}
+% 	B_k(t) = \sqrt{\alpha \log(t) / N_k(t)},
+% \end{equation}
+% where $\alpha$ refers to an exploration coefficient,
+% % \footnote{~In fact, the larger this coefficient is, the longer the exploration, while the \UCB{} algorithm is proven to be order optimal for $\alpha>0.5$ \cite{Bubeck12}, and has reported a good performance for lower values of $\alpha>0$.},
+% that we chose equal to $1/2$, as suggested in \cite{Audibert07} and as done in previous works \cite{Bonnefoi18,Bonnefoi17}.
+% Then, an upper confidence bound in each channel $k$ is defined as
+% \begin{equation}\label{eq:43:ucb}
+% 	U_k(t) = \widehat{\mu_k}(t) + B_k(t).
+% \end{equation}
+% Finally, the transmission channel at time step $t$
+% is the one maximizing this \UCB{} index $U_k(t)$,
+% as it is the one expected to be the best one at the current time step $t$,
+% \begin{equation}\label{eq:43:maxucb}
+% 	C(t) = \arg\max_{1\leq k \leq K} U_k(t).
+% \end{equation}
+
+% The \UCB{} algorithm is implemented independently by each device, and we present it in Algorithm~\ref{algo:43:UCB}.
+% Note that a device using this first approach is only able to select a channel for the first and all the corresponding retransmissions of a packet.
+
+% % \begin{small} % XXX remove if needed
+% \begin{figure}[h!]
+% 	\centering
+%     \begin{framed}
+% 	\begin{algorithm}[H]
+% 		% \begin{small} % XXX remove if needed
+% 		\For(){$t = 1, \dots, T$}{
+% 			Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$\;
+% 			Transmit in channel $C(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
+% 			Reward $r_{C(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
+% 		}
+% 		\caption{The \UCB{} algorithm for channel selection (base building block).}
+% 		\label{algo:43:UCB}
+% 		% \end{small} % XXX remove if needed
+% 	\end{algorithm}
+% 	\end{framed}
+% \end{figure}
+% % \end{small} % XXX remove if needed
+
+% ----------------------------------------------------------------------
+\subsection{The first heuristics: \UCB{} unaware of retransmissions}
+\label{sub:43:heuristics}
 % ----------------------------------------------------------------------
 
-Before presenting our proposed heuristics, we remind here the \UCB{} bandit algorithm \cite{Auer02}.
-We include it again to facilitate the understanding of the heuristics, which use \UCB{} as a building block.
+% \paragraph{First stage \UCB}\label{sub:43:firstStageUCB}
 
-% % --- --- --- --- --- --- --- --- --- ---
-\paragraph{The \UCB{} algorithm}\label{sub:43:algoUCB}
+This first heuristic we propose is unaware of retransmission: the same channel is used for retransmissions.
+The \UCB{} algorithm is implemented independently by each device, and we present it in Algorithm~\ref{algo:43:UCB}.
+Note that a device using this first approach is only able to select a channel for the first transmission (using \UCB), and it uses the same channel for all the corresponding retransmissions of a packet (if retransmissions happen).
 
@@
 for any $t\in\mathbb{N}$,
-\begin{equation}\label{eq:43:Nkt}
-	N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(C(\tau) = k),
-\end{equation}
+$N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(C(\tau) = k)$.
 The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
-\begin{equation}\label{eq:43:mukt}
-	\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(C(\tau) = k).
-\end{equation}
-where $r_{k}(t)$ is the reward obtained after transmission in channel $k$ at time $t$ ($1$ for a successful transmission, and $0$ otherwise)
-%
-A \emph{confidence} term $B_k(t)$ is given by \cite{Auer02},
-\begin{equation}\label{eq:43:Bkt}
-	B_k(t) = \sqrt{\alpha \log(t) / N_k(t)},
-\end{equation}
-where $\alpha$ refers to an exploration coefficient,
-% \footnote{~In fact, the larger this coefficient is, the longer the exploration, while the \UCB{} algorithm is proven to be order optimal for $\alpha>0.5$ \cite{Bubeck12}, and has reported a good performance for lower values of $\alpha>0$.},
-that we chose equal to $1/2$, as suggested in \cite{Audibert07} and as done in previous works \cite{Bonnefoi18,Bonnefoi17}.
-Then, an upper confidence bound in each channel $k$ is defined as
-\begin{equation}\label{eq:43:ucb}
-	U_k(t) = \widehat{\mu_k}(t) + B_k(t).
-\end{equation}
+$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(C(\tau) = k)$,
+where $r_{k}(t)$ is the reward obtained after transmission in channel $k$ at time $t$.
+The upper confidence bound in each channel $k$ is defined as
+$U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$.
 Finally, the transmission channel at time step $t$
@@
 as it is the one expected to be the best one at the current time step $t$,
-\begin{equation}\label{eq:43:maxucb}
-	C(t) = \arg\max_{1\leq k \leq K} U_k(t).
-\end{equation}
-
-The \UCB{} algorithm is implemented independently by each device, and we present it in Algorithm~\ref{algo:43:UCB}.
-Note that a device using this first approach is only able to select a channel for the first and all the corresponding retransmissions of a packet.
+$C(t) = \arg\max_{1\leq k \leq K} U_k(t)$.
 
@@
 		\For(){$t = 1, \dots, T$}{
-			Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+			\uIf{First packet transmission}{
+				Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$\;
 				Transmit in channel $C(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
@@
 			}
-		\caption{The \UCB{} algorithm for channel selection (base building block).}
+			\Else(\tcp*[f]{Retransmit in same channel}){
+				j $\leftarrow$ last channel selected by first-stage \UCB\;
+				Transmit in channel $C(t) = j$\;
+			}
+		}
+		\caption{Heuristic: first-stage \UCB{} and retransmission in same channel.}
 		\label{algo:43:UCB}
@@
 
+\paragraph{A warning about notations: local (device) time vs global (gateway) time.}
+%
+In all the algorithms and notations presented in this Section, the time steps $t$ do not refer to the global time steps (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
+In other words, each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $M$ times (\eg, $M=10$) until it receives an \Ack.
+To be clear, once again, we repeat that the time steps $t$ in the following algorithms denote the number of time the device sent an uplink packet, and these uplink (re)transmissions can happen at different (real-world) times, more or less spread out in time, it does not matter to the learning part of this model and work.
+
 % ----------------------------------------------------------------------
@@
 A device that implements the UCB algorithm is led to focus its transmissions and retransmissions in the channel which is currently identified as the best.
-As explained before, focusing in one channel increases the collision probability in retransmissions.
-We describe here the proposed heuristics for the channel selection in a retransmission. It is carried out taking
-into account that a device can incorporate a different channel selection strategy while being in a back-off state.
+As explained above in Section~\ref{sub:43:motivations}, focusing in one channel could increase the collision probability in retransmissions.
+We describe here the proposed heuristics for the channel selection in a retransmission.
+It is carried out taking into account that a device can incorporate a different channel selection strategy while being in a back-off state.
 Hence, a natural question is to evaluate whether using this additional contextual information can improve the performance of a learning policy.
@@
 			\Else(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^r$}){
-				Compute $\forall k, U^r_k(t) = \widehat{\mu^r_k}(t) + B^r_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+				Compute $\forall k, U^r_k(t) = \widehat{\mu^r_k}(t) + \sqrt{\alpha \log(t) / N_k^r(t)}$\;
 				Transmit in channel $C^r(t) = \arg\max_{1\leq k \leq K} U^r_k(t)$\;
@@
 				j $\leftarrow$ last channel selected by first-stage \UCB\;
-				Compute $\forall k, U_k^j(t) = \widehat{\mu_k}^j(t) + B_k^j(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+				Compute $\forall k, U_k^j(t) = \widehat{\mu_k}^j(t) + \sqrt{\alpha \log(t) / N_k^j(t)}$\;
 				Transmit in channel $C^j(t) = \arg\max_{1\leq k \leq K} U^j_k(t)$\;
@@
 The idea behind this delay is to allow the first stage \UCB{} to start learning the best channel, before starting the second stage \UCB{} (see details in Algorithm~\ref{algo:43:UCBwithDelay}).
-The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to be able to simulate the problem in advance in order to find a good value of some internal parameter.
+The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to simulate the problem before-hand to find a good value of some internal parameter.
 As such, we only consider a delay of $\Delta=100$.}.
@@
 			\Else(\tcp*[f]{Packet retransmission with delayed $\mathrm{UCB}^d$}){
-				Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + B^d_k(t)$ following Eqs.~\eqref{eq:43:Nkt},~\eqref{eq:43:mukt}, and~\eqref{eq:43:Bkt}\;
+				Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + \sqrt{\alpha \log(t) / N_k^d(t)}$\;
 				Transmit in channel $C^d(t) = \arg\max_{1\leq k \leq K} U^d_k(t)$\;
@@
 
-\paragraph{A last idea that we did not explore}
+\paragraph{Other ideas that we did not explore}
 %
 Instead of considering a first-stage \UCB{} followed by a random channel retransmission, we could have considered a random first-stage channel retransmission followed by a second-stage \UCB.
-This was not considered in our study \cite{Bonnefoi2019WCNC}, and we preferred to not add it to the experiments presented below, for two reasons: to avoid clutter in the plots, and to win some time as these experiments had to run for quite a long time.
+This was not considered in our study \cite{Bonnefoi2019WCNC}, and we preferred to not add it to the experiments presented below, for three reasons: to avoid clutter in the plots, to win some time as these experiments had to run for quite a long time, but mainly because the first-stage channel selection is most important one as we illustrate by the large difference in terms of performance between the uniform channel selection and the ``only UCB'' heuristic, below in Figures~\ref{fig:43:mainExperiment1} and \ref{fig:43:mainExperiment2}.
+
+Similarly, Philippe Mary asked me at the MoTION workshop where I presented \cite{Bonnefoi2019WCNC}, why didn't we rather used Bayes-UCB instead of \UCB{} (from \cite{Kaufmann12BUCB}), as it is known to empirically outperform \UCB{} (as illustrated for instance in \cite{darak2016bayesian,kumar2016two}).
+We could have chose any bandit algorithm for the ``first stage'' component used in this Section, and for simplicity and clarity we focussed on a simple but efficient one (\UCB).
+We believe that the same empirical results, but more importantly, the same conclusions, could be given if we used Bayes-UCB or other bandit algorithm for the base building block in the different heuristics proposed in Section~\ref{sub:43:heuristics}.
 

### 2019-04-16 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
J'arriverai jamais à finir... trop chiant.

@@
 for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
-Finally, the result of our experimental study reported in Section~\ref{sec:5:experiments}.
+The result of our experimental study reported in Section~\ref{sec:5:experiments}.
+Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
 % and finally Section~\ref{sec:5:conclusion} concludes.
@@
   The proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step,
-  but the results stated in the section are now correct,
+  but the results stated in this section are now correct,
   thanks to the modifications proposed by Emilie Kaufmann and Abbas Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
@@
     % so we can use Wald's Lemma \citep{Wald45}
-    &\leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{\beta}.
+    &\leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}.
     % = (\alpha) \times (\beta).

### 2019-04-16 2-Chapters/5-Chapter/chapter5.tex
J'arriverai jamais à finir... trop chiant.

@@
 \TODOL{
-    Pour chaque extension présentée ci dessous, je veux suivre la même organisation :
+    Pour les extension présentées ici, je veux suivre la même organisation :
 
@@
 
-\TODOL{Explain why the problem is MUCH easier if we have before hand these unique ID!}
 If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm running on top of a round-Robin phase can achieve order-optimal regret.
-Without giving more details, we let the interested refer to what is is explained for the RR-SW-UCB\# algorithm in \cite{WeiSrivastava18Distributed} and the algorithms presented in \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
+If the players know the time horizon $T$, then fix a confidence level $\delta$.
+For the first $T_0$ time steps, the $M$ players will use a simple round-Robin game, using their (unique) ID to stay orthogonal: player $j$ starts at arm $j$, then $j+1$, then cycle in $\{1,\dots,K\}$.
+They encounter no collision in these time steps, and then user $j$ targets the $j$-th best arm among the set of $M$-best arm.
+If $T_0$ is large enough, all players have built the same estimate of the ranking of the arms (and not only correctly identified the set of $M$-best arms), with high probability (at least $\delta$), and thus they will also encounter no collision and no regret from after time $T_0$.
+The mean regret of such approach is easily bounded by $R_T \leq M T_0 + (1 - \delta) (T - T_0)$, so by using $\delta = 1-1/T$, $R_T = \bigO{T_0}$.
+By calibrating $T_0$ based on $\delta$ (which needs prior knowledge of $T$) and the minimal gap $\Delta$ between \Mbest{} and \Mworst, one can show, using similar arguments as used by \cite{Rosenski16} for the Musical Chair policy, that with large probability a constant regret is obtained.
+A logarithmic regret can be obtained from the same bound, proving the order-optimality of this simple approach, if we assume that user $j$ knows it is user number $j$.
+%
+Without giving more details, we let the interested refer to what is is explained for the algorithms presented in \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
+%
+These works assume a prior knowledge of $\Delta$, or other measures of the difficulty of the problem, and as such we do not.
+FIXME
+\TODOL{FIXME écrire ça sans être fini}
 
@@
 \TODOL{Write about half a page on the article \cite{BoursierPerchet18}.}
+
+The recent work \cite{BoursierPerchet18} proposed an idea they called the ``communication trick''.
+It essentially allow any player to send one bit to all the others, at some pre-agreed time steps, with no modification on the model.
+Thanks to this ``communication trick'', recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
+This trick essentially use the fact that the players share a synchronized time, thus they can use a collision as a way to directly exchange one bit of information between two players.
+This process is slow and not efficient, as all but two players must do nothing when player $i$ is sending a bit to player $j$, but it does build a communication protocol in the model without explicit communication between players.
+This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
+We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained
 
@@
 An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit.
-We do not review papers that study this model, because the recent work \cite{BoursierPerchet18} introduced the ``communication trick'', that essentially allow any player to send one bit to all the others, at some pre-agreed time steps, with no modification on the model.
-Thanks to this ``communication trick'', recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
-This trick essentially use the fact that the players share a synchronized time, thus they can use a collision as a way to directly exchange one bit of information between two players.
-This process is slow and not efficient, as all but two players must do nothing when player $i$ is sending a bit to player $j$, but it does build a communication protocol in the model without explicit communication between players.
+We do not review papers that study this model, because the recent work \cite{BoursierPerchet18} introduced the ``communication trick''.
+Recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
 This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
-We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.
+Due to space constraint, we let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.
 

### 2019-04-16 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
More on discussions to finish chapter 5

@@
 
-We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \klUCB{} indices.
-% Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.
-
-\TODOL{Include here some plots showing \RhoRand{} with \UCB{} and \klUCB{} vs \MCTopM{} with \UCB{} and \klUCB{}.}
-
-In experiments that are not reported here, we could observe that using \klUCB{} rather than \UCB{} indices always yield better practical performance.
-%, as the theoretical guarantees for single player suggested,
-As the purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
-% And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
-we only report results for \klUCB.
-In a first set of experiments, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
+We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \UCB{} or \klUCB{} indices.
+The analysis given in Section~\ref{sec:5:upperbounds} was focussing on \klUCB, because it is well known that it is more efficient both in practice and in theory.
+To illustrate this, we first show below the result of some experiments comparing different algorithms that use either the \UCB{} or the \klUCB{} indexes.
+%
+Our proposed algorithms, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
 We also include a centralized multiple-play \klUCB{} algorithm,
@@
 essentially to check that the \emph{``price of decentralized learning''} is not too large.
+
+The first experiment is considering $K=9$ arms, with means $\bm{\mu}=[0.1,0.2,\dots,0.9]$, and $M=3$ then $6$ then $9$ players.
+Note that here and as in all this section we consider different algorithms that know the number of players $M$ (see Section~\ref{sub:5:unknownNumberOfPlayers} below for a discussion on the case when $M$ could be unknown).
+%
+Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
+%
+We report in Table~\ref{table:5:comparisonUCB_klUCB} the results in terms of mean regret, $\pm$ one standard deviation.
+The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
+Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
+Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved (the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is $\cO(1)$ if there is no arms in \Mworst{} !), and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous (state-of-the-art) approaches using ranks.
+%
+Additional experiments and illustrations are given below.
+We illustrate below the (empirical mean) regret $R_t$ as a function of time and not only results in terms of empirical mean regret $R_T$ at the end of the bandit game.
+
+% for N in 4 100; do for M in 3 6 9; do DEBUG=True SAVEALL=False NOPLOTS=True M=$M K=30 N_JOBS=-1 N=$N T=10000 make moremultiplayers; echo "Done for N=$N and M=$M"; read; done; echo "Done for N=$N"; read; done
+% TODO run the simulations with N=100 and not N=4
+\begin{table}[ht]
+  % \begin{footnotesize}
+      \centering
+      \begin{tabular}{c|ccc}
+      \textbf{Algorithm} $\;$ \textbackslash $\;$ Number of players & $M=3$ & $M=6$ & $M=9$ \\
+          \hline
+          Centralized multiple-play \UCB{} & $317 \pm 37$ & $219 \pm 17$ & $0$ \\
+          Centralized multiple-play \klUCB{} & $94 \pm 32$ & $77 \pm 7$ & $0$ \\
+          \hline
+          \Selfish-\UCB{} & $1175 \pm 70$ & $3705 \pm 379$ & $12709 \pm 246$ \\
+          \Selfish-\klUCB{} & $230 \pm 34$ & $673 \pm 48$ & $2689 \pm 230$ \\
+          \hline
+          \RhoRand-\UCB{} & $1534 \pm 145$ & $4797 \pm 319$ & $12047 \pm 1056$ \\
+          \RhoRand-\klUCB{} & $349 \pm 58$ & $2688 \pm 469$ & $6709 \pm 1186$ \\
+          \hline
+          \RandTopM-\UCB{} & $1074 \pm 36$ & $2751 \pm 229$ & $252 \pm 150$ \\
+          \RandTopM-\klUCB{} & $245 \pm 30$ & $1181 \pm 100$ & $617 \pm 592$ \\
+          \hline
+          \MCTopM-\UCB{} & $941 \pm 23$ & $1368 \pm 86$ & $46 \pm 5$ \\
+          \MCTopM-\klUCB{} & $\mathbf{277 \pm 12}$ & $\mathbf{391 \pm 70}$ & $\mathbf{50 \pm 11}$ \\
+          \hline
+      \end{tabular}
+      \caption{Using \klUCB{} is much more efficient than using \UCB{}, for multi-player bandit (here in a simple problem with $K=9$ arms).}
+      \label{table:5:comparisonUCB_klUCB}
+  % \end{footnotesize}
+  \end{table}
+
+The purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
+% And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
+and as we observed that using \klUCB{} rather than \UCB{} indices always yield better practical performance,
+from now on we only report results for \klUCB.
+%, as the theoretical guarantees for single player suggested,
 
@@
 and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
-Note that here, we consider different algorithms that know the number of players $M$.
-Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
 %

### 2019-04-16 2-Chapters/5-Chapter/chapter5.tex
More on discussions to finish chapter 5

@@
 FIXME
-\TODOL{FIXME écrire ça sans être fini}
+\TODOL{FIXME écrire ça sans être méchant envers leurs travaux !}
 
@@
 This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
-We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained
+We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.
 
@@
 What we are interested here is an extension of our model where players can send some bits to one or all the other players, at every time steps or some time steps.
-Clearly, the problem is easier by allowing communication, and it is quite immediate to see that the communication capacity between players must be limited otherwise the problem becomes trivial to solve.
+
+Clearly, the problem is easier by allowing communication, and it is quite immediate to see that the communication capacity between players must be limited otherwise the problem is already known and solved.
 Indeed, imagine that at each time step $t$, all the $M$ players could share an unbounded number of bits with the other players, at no cost, even if this hypothesis is clearly unrealistic for wireless networks.
-Then they can share all their observations, and they can all run the same multi-plays MAB algorithm \cite{Anantharam87a} (in a deterministic way), like for instance the extension of Thompson sampling for multi-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
+Then they can share all their observations, and they can all run the same multiple-plays MAB algorithm \cite{Anantharam87a} (in a deterministic way), like for instance the extension of Thompson sampling for multiple-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
 In this setting, a logarithmic regret is obtained easily, and the regret upper-bound of the two aforementioned algorithms asymptotically achieve the lower-bound from \cite{Anantharam87a}.
 
-An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit.
-We do not review papers that study this model, because the recent work \cite{BoursierPerchet18} introduced the ``communication trick''.
-Recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
+An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit, at every or only some time steps.
+We identified that the state-of-the-art on this direction of research consists in the two very recent works published in April 2019, while I was finishing this chapter: \cite{tao2019collaborative} and \cite{wang2019distributed}.
+Distributed pure exploration is studied in \cite{tao2019collaborative}, where the proposed new lower bounds for the regret of any algorithm for the distributed best arm identification problem, under the fixed time or fixed confidence settings.
+The also propose effective algorithms that asymptotically match their lower-bounds (up-to logarithmic factors, in some cases).
+The second work studies distributed learning for (not necessarily stochastic) multi-armed bandits as well as linear bandits \cite{tao2019collaborative}.
+For a distributed $K$-armed bandit with $M$ agents, they developed two protocols achieving near-optimal regret $\cO(\sqrt{M K T \log(T)})$,
+and requiring little communication cost,
+one is independent of the time horizon $T$ and use $\cO(M \log(T))$ communications,
+and the other is independent of the number of arms $K$ and use $\cO(M K \log(M))$ communications.
+Their model and algorithms fit in the distributed, decentralized framework we advertise in all this thesis, and this last work is very interesting.
+
+Additionally, after the recent work \cite{BoursierPerchet18} introduced the ``communication trick'',
+some recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
 This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.

### 2019-04-17 2-Chapters/1-Chapter/chapter1.tex
Okay for chapters and parts
Small modifications on all chapters
Work work work...

@@
 % First chapter begins here
-\chapter{Introduction}
+\chapter{Setting of This Thesis}
 \label{chapter:1}
 \minitoc
+
+Abstract for Chapter 1.
+
 \newpage
@@
     \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
-    \texttt{hal.inria.fr/hal-02006825}.
+    \texttt{HAL.Inria.fr/hal-02006825}.
 
@@
     \emph{Algorithmic Learning Theory} (ALT, \texttt{www.cs.cornell.edu/conferences/alt2018}), Lanzarote, Spain, April 2018,
-    \texttt{hal.inria.fr/hal-01629733}.
+    \texttt{HAL.Inria.fr/hal-01629733}.
 
@@
     \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2018.ieee-wcnc.org}), Barcelone, Spain, April 2018,
-    \texttt{hal.inria.fr/hal-01705292}.
+    \texttt{HAL.Inria.fr/hal-01705292}.
 
@@
     \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \texttt{crowncom.org/2017}), Lisboa, Portugal, Septembre $2017$,
-    \texttt{hal.inria.fr/hal-01575419},
+    \texttt{HAL.Inria.fr/hal-01575419},
     \textbf{Best Paper Award}.
@@
 % =============================================================================
-\subsection{Submitted works or in progress works waiting for a new submission}
+\subsection{Publications in international conferences without proceedings}
 
 \begin{itemize}
-\item
-    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
-    \texttt{hal.inria.fr/hal-02006471}.
 
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
-    Submitted for MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}), February $2019$.
-
-\item
-    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
-    by \textbf{Lilian Besson}, active development since October $2016$,
-    \texttt{hal.inria.fr/hal-01840022}.
-    About $40000$ lines of code, hosted on \texttt{GitHub.com/SMPyBandits},
-    with a complete documentation on \texttt{SMPyBandits.rtfd.io}.
-
-\item
-    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
-    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
-    \texttt{hal.inria.fr/hal-01736357}.
+    Presented at the 1st International MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}),
+    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
 
@@
 
+\end{itemize}
+
+% =============================================================================
+\subsection{Submitted works}
+
+\begin{itemize}
+\item
+    \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
+    submitted at GRETSI 2019, \texttt{GRETSI.fr/colloque2019},
+    by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
+    \texttt{HAL.Inria.fr/hal-02006471}.
+
+\item
+    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
+    submitted at COLT 2019, \texttt{LearningTheory.org/colt2019/},
+    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
+    \texttt{HAL.Inria.fr/hal-02006471}.
+
+\end{itemize}
+
+% =============================================================================
+\subsection{In progress works waiting for a new submission}
+
+\begin{itemize}
+\item
+    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
+    by \textbf{Lilian Besson}, active development since October $2016$,
+    \texttt{HAL.Inria.fr/hal-01840022}.
+    About $40000$ lines of code, hosted on \texttt{GitHub.com/SMPyBandits},
+    with a complete documentation on \texttt{SMPyBandits.rtfd.io}.
+
+\item
+    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
+    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
+    \texttt{HAL.Inria.fr/hal-01736357}.
+
+\end{itemize}
+
+% =============================================================================
+\subsection{Other works}
+
+\begin{itemize}
 \item
@@
     February $2018$,
-    \texttt{hal.inria.fr/hal-01847480}.
+    \texttt{HAL.Inria.fr/hal-01847480}.
 
@@
 
-\paragraph{Seminars} I gave some presentations in the following events:
+\paragraph{Seminars}
+    I gave some presentations in the following events:
     SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
@@
     for the brown-bag seminar at ENSAI in Bruz in January $2018$,
-    for the weekly CMAP seminar at École Polytechnique in October $2018$.
+    for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
+    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in May $2019$.
 
@@
     We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
-    and I gave about $10$ training, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.
+    and I gave about $12$ training sessions, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.
 
@@
 
-\paragraph{Conferences} I attended the following conferences:
+\paragraph{Conferences}
+    I attended the following conferences:
 	\emph{International Conference on Communication} ICC (Paris), May $2017$,
@@
 
-\paragraph{Seminars} I attended the following seminars:
+\paragraph{Seminars}
+    I attended the following seminars:
 	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
@@
 \paragraph{Responsabilities}
-    president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
+    I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
     I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{Cf.
-\texttt{hal.inria.fr/hal-02013839/document}}.
+\texttt{HAL.Inria.fr/hal-02013839/document}}.
     I also helped for the organization and presented another poster\footnote{Cf.
-\texttt{hal.inria.fr/hal-02013847/document}} at IETR seminar (\texttt{seminar-ietr-18.sciencesconf.org}), held in Vannes in June $2018$.
+\texttt{HAL.Inria.fr/hal-02013847/document}} at IETR seminar (\texttt{seminar-ietr-18.sciencesconf.org}), held in Vannes in June $2018$.
 
@@
 	for \emph{European Workshop on Reinforcement Learning} (Lille) \texttt{ewrl.wordpress.com/ewrl14-2018}, in October $2018$, I reviewed $5$ papers.
-	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \href{https://nips.cc/Conferences/2018/}{\emph{NeurIPS}} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.
+	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.
 
@@
 	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
-	from January $2017$ to June $2019$.
+	from January $2017$ to September $2018$.

### 2019-04-17 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Work work work...

@@
 % for N in 4 100; do for M in 3 6 9; do DEBUG=True SAVEALL=False NOPLOTS=True M=$M K=30 N_JOBS=-1 N=$N T=10000 make moremultiplayers; echo "Done for N=$N and M=$M"; read; done; echo "Done for N=$N"; read; done
-% TODO run the simulations with N=100 and not N=4
+%
 \begin{table}[ht]
@@
           \hline
-          Centralized multiple-play \UCB{} & $317 \pm 37$ & $219 \pm 17$ & $0$ \\
-          Centralized multiple-play \klUCB{} & $94 \pm 32$ & $77 \pm 7$ & $0$ \\
+          Centralized multiple-play \UCB{} & $321 \pm 30$ & $233 \pm 25$ & $0$ \\
+          Centralized multiple-play \klUCB{} & $94 \pm 17$ & $68 \pm 18$ & $0$ \\
           \hline
-          \Selfish-\UCB{} & $1175 \pm 70$ & $3705 \pm 379$ & $12709 \pm 246$ \\
-          \Selfish-\klUCB{} & $230 \pm 34$ & $673 \pm 48$ & $2689 \pm 230$ \\
+          \Selfish-\UCB{} & $1263 \pm 157$ & $3694 \pm 387$ & $12420 \pm 404$ \\
+          \Selfish-\klUCB{} & $243 \pm 31$ & $743 \pm 113$ & $3005 \pm 492$ \\
           \hline
-          \RhoRand-\UCB{} & $1534 \pm 145$ & $4797 \pm 319$ & $12047 \pm 1056$ \\
-          \RhoRand-\klUCB{} & $349 \pm 58$ & $2688 \pm 469$ & $6709 \pm 1186$ \\
+          \RhoRand-\UCB{} & $1455 \pm 208$ & $4775 \pm 463$ & $11794 \pm 1083$ \\
+          \RhoRand-\klUCB{} & $394 \pm 96$ & $2385 \pm 412$ & $7057 \pm 1053$ \\
           \hline
-          \RandTopM-\UCB{} & $1074 \pm 36$ & $2751 \pm 229$ & $252 \pm 150$ \\
-          \RandTopM-\klUCB{} & $245 \pm 30$ & $1181 \pm 100$ & $617 \pm 592$ \\
+          \RandTopM-\UCB{} & $1020 \pm 92$ & $2899 \pm 418$ & $470 \pm 346$ \\
+          \RandTopM-\klUCB{} & $258 \pm 43$ & $902 \pm 234$ & $551 \pm 520$ \\
           \hline
-          \MCTopM-\UCB{} & $941 \pm 23$ & $1368 \pm 86$ & $46 \pm 5$ \\
-          \MCTopM-\klUCB{} & $\mathbf{277 \pm 12}$ & $\mathbf{391 \pm 70}$ & $\mathbf{50 \pm 11}$ \\
+          \MCTopM-\UCB{} & $980 \pm 76$ & $1466 \pm 132$ & $43 \pm 13$ \\
+          \MCTopM-\klUCB{} & $\mathbf{248 \pm 40}$ & $\mathbf{410 \pm 54}$ & $\mathbf{42 \pm 10}$ \\
           \hline
@@
 %
-Empirically, our proposals were found to almost always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
+Empirically, our proposals were found to always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
 we verified that \MCTopM{} outperforms the state-of-the-art algorithms in many different problems, and is more and more efficient as $M$ and $K$ grows.

### 2019-04-17 2-Chapters/5-Chapter/chapter5.tex
Work work work...

@@
 One can observe in Table~\ref{table:5:meanRegretSimulationsEstimatingM} the empirical performances of different algorithms in an example problem, in each case of low, medium and maximum number of players ($M=3,6,9$ for $K=9$ arms).
-The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar costs on their regret when they have to estimate $M$ on the fly (written ``Estimate $M$'' in red, in Table~\ref{table:5:meanRegretSimulationsEstimatingM}).
+The three algorithms \RhoRand, \RandTopM{} and \MCTopM{} all suffer from similar increase on their regret when they have to estimate $M$ on the fly (written ``Estimate $M$'' in red, in Table~\ref{table:5:meanRegretSimulationsEstimatingM}).
 Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.
 For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large (on the three algorithms and two cases, about $\times 3.4$), and consistent on the different algorithm.
-The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proved to be $\smallO{\log(T)}$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
-In this extreme case, \RhoRand{} and its extension both perform very closely, as expected.
+The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proved to be $\cO(1)$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
+In this extreme case, \RhoRand{} and its extension both perform very closely, as expected, and much worse than \MCTopM.
 %

### 2019-04-17 3-Appendices/2-Appendix/appendix2.tex
Ok for tex root
Still fighting against my template...
Work work work...

--- 2019-01-17 3-Appendices/2-Appendix/appendix2.tex
@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 % ******************************* Thesis Appendix B ****************************
-\chapter{Appendix for Chapter \ref{chapter:2}}
-\label{app:chapt2}
+\chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
+\label{app:2:DoublingTricks}
 
-In this appendix, blabla.
+This appendix quickly presents the contributions of another publication that was not presented in the main text of this thesis, \cite{Besson2018DoublingTricks}.
+
+% \TODOL{
+% I cannot include it raw without more work on the maths, we have to write only the best versions of our results.
+
+% But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!
+
+% I need to try DoublingTrick + GLRklUCB, at least experimentally!
+% For the theoretical analysis, it will probably be a huge failure.
+% }
+
+As introduced in Chapter~\ref{chapter:2},
+an online reinforcement learning algorithm is \emph{anytime} if it does not need to know in advance the horizon $T$ of the experiment.
+%
+Depending on the context of the practical application of interest, it might be unrealistic to assume to know in advance $T$.
+We give two examples to illustrate where this prior knowledge can be realistic or not.
+On the first hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for instance we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
+On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices: usually a learning step corresponds to an uplink and then downlink message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages.
+While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies), many kinds of applications such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.
+
+For this later range of applications, it is of highest interest to be able to use low-cost MAB algorithms that do not rely on prior knowledge of the problem for which they will be applied, and especially do not need to know the horizon $T$.
+While we proposed in Chapter~\ref{chapter:5} any-time algorithms solving the presented problem of stationary multi-player bandit, our solution for the problem of non-stationary MAB problem studied in Chapter~\ref{chapter:6} does rely on a prior knowledge of the horizon $T$.
+It is interesting to know what is the best approach to fix this weakness: work more and design algorithms that are inherently any-time, or use a generic technique to avoid depending on a prior knowledge of $T$, and automatically transform our non-anytime approach to make it anytime.
+
+A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06}.
+We studied doubling tricks between fall 2017 and spring 2018, and we wrote an article \cite{Besson2018DoublingTricks} which was rejected at the COLT 2018 conference, and unfortunately we lacked the time to improve it and resubmit it to another conference.
+%
+In the context of adversarial or stochastic multi-armed bandits,
+the performance of an algorithm is measured by its regret,
+and we study two families of sequences of growing horizons (geometric and exponential)
+to generalize previously known results that certain doubling tricks can be used to conserve certain regret bounds.
+In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log T)$.
+We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log T)$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
+Interestingly, we prove that they conserve bounds of the mixed form $R_T = \cO(T^{\gamma} (\log(T))^{\delta})$, for $0<\gamma<1$ and $\delta>0$, and so an exponential doubling trick could be used to obtain an anytime version of the algorithm we propose for non-stationary bandits, \GLRklUCB{} in Chapter~\ref{chapter:6}.
+However, our study was only focussing on stationary bandit, and it is left as a future work to explore the harder case of piece-wise stationary problems.
+
+In our article \cite{Besson2018DoublingTricks}, we also present numerical experiments in the case of stationary MAB problems, to compare the performance of efficient anytime algorithms, like \klUCB, against efficient non-anytime algorithms, like \KLUCBpp, made anytime with different doubling-trick.
+We conclude that for such problems, if $T$ is not known before, it is always more efficient to use policies that were designed to be anytime that to use a doubling trick.
+For example, an applicative paper written in 2018, \cite{li2019useDoublingTrick}, only tested the use of an exponential doubling trick combined with \KLUCBpp, but most surely the \klUCB{} algorithm would have given better empirical performance as well as more robust results.
+
+To conclude this work, we would need to complete the study of the doubling tricks, and instead of focussing on two families of doubling tricks (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bound.
+Such doubling is given either by the function mapping the current time $t$ to the current estimate of the horizon $T(t)$, or the sequence $(T_i)_{i\in\N^*}$ of successive estimates of the horizons.
+We are interested in pursuing this work, in the hope of finding an intermediate sequence, growing faster than a geometric but slower than an exponential doubling sequence, that can preserve both worlds, problem-dependent bounds in $\cO(\log(T))$ and problem-independent bounds in $\cO(\sqrt{T})$.

### 2019-04-17 2-Chapters/1-Chapter/chapter1.tex
Cette thèse me semble interminable...

@@
 
+\TODOL{Write this!}
+
 \newpage
@@
 
+\TODOL{Write this!}
+
 \subsection{Spectrum scarcity}
@@
 \label{sec:1:solutions}
+
+\TODOL{Write this!}
 
@@
 \label{sec:1:contributions}
+
+\TODOL{Write this!}
 
@@
     \texttt{HAL.Inria.fr/hal-02006825}.
+    \cite{Besson2019WCNC}
 
@@
     \texttt{HAL.Inria.fr/hal-01629733}.
+    \cite{Besson2018ALT}
 
@@
     \texttt{HAL.Inria.fr/hal-01705292}.
+    \cite{Besson2018WCNC}
 
@@
     \textbf{Best Paper Award}.
+    \cite{Bonnefoi17}
 
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
-    Presented at the 1st International MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}),
-    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
+    1st International MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}),
+    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019.
+    \cite{Bonnefoi2019WCNC}
 
@@
     See \texttt{YouTu.be/HospLNQhcMk} for a $6$-minutes presentation video.
+    \cite{Besson2018ICT}
 
@@
     \texttt{HAL.Inria.fr/hal-02006471}.
+    \cite{Besson2019Gretsi}
 
@@
     \texttt{HAL.Inria.fr/hal-02006471}.
+    \cite{Besson2019GLRT}
 
@@
     with a complete documentation on \texttt{SMPyBandits.rtfd.io}.
+    \cite{SMPyBandits,SMPyBanditsJMLR}
 
@@
     \texttt{HAL.Inria.fr/hal-01736357}.
+    \cite{Besson2018DoublingTricks}
 

### 2019-04-17 2-Chapters/3-Chapter/chapter3.tex
Expérience en cours pour la fin du Chapitre:3.
Done for today...
Done for this simulation
Cette thèse me semble interminable...

@@
         \end{minted}
-        \caption{Small snippet of Bash to download and install dependencies of SMPyBandits}
-        % \captionof{listing}{Small snippet of Bash to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
+        \caption{Small snippet of Bash code to download and install dependencies of SMPyBandits}
+        % \captionof{listing}{Small snippet of Bash code to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
         \label{lst:3:howToInstallLibrary}
@@
     \end{minted}
-    \caption{Small snippet of Bash to run a simple experiment with SMPyBandits}
-    % \captionof{listing}{Small snippet of Bash to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
+    \caption{Small snippet of Bash code to run a simple experiment with SMPyBandits}
+    % \captionof{listing}{Small snippet of Bash code to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
     \label{lst:3:howToRunBasicLibrary}
@@
     \end{minted}
-    \caption{Small snippet of Bash to run an experiment}
-    % \captionof{listing}{Small snippet of Bash to run an experiment \label{lst:3:howToRunExperiment2}.}
+    \caption{Small snippet of Bash code to run an experiment}
+    % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}
     \label{lst:3:howToRunExperiment2}
@@
 
-\TODOL{Je veux terminer cet exemple!}
-
-% FIXME run
 % for N in 1 10 100; do for NJOBS in 1 4 100; do DEBUG=True SAVEALL=False NOPLOTS=True N=$N T=10000 N_JOBS=$NJOBS make single; read; done
@@
         \hline
-        $1$ repetition    & $XXX \;\text{s}$ & $XXX \;\text{s}$ & $XXX \;\text{s}$ \\
-        $10$ repetitions  & $XXX \;\text{s}$ & $XXX \;\text{s}$ & $XXX \;\text{s}$ \\
-        $100$ repetitions & $59 \;\text{s}$ & $37 \;\text{s}$ & $XXX \;\text{s}$ \\
+        $1$ repetition    & $15 \;\text{s}$ & $26 \;\text{s}$ & $43 \;\text{s}$ \\
+        $10$ repetitions  & $87 \;\text{s}$ & $51 \;\text{s}$ & $76 \;\text{s}$ \\
+        $100$ repetitions & $749 \;\text{s}$ & $272 \;\text{s}$ & $308 \;\text{s}$ \\
+        $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1346 \;\text{s}$ \\
         \hline

### 2019-04-17 3-Appendices/1-Appendix/appendix1.tex
Ok for tex root
WOrking on the whole document but not yet Chapter5
Cette thèse me semble interminable...

--- 2019-01-17 3-Appendices/1-Appendix/appendix1.tex
@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 % ******************************* Thesis Appendix A ****************************
-\chapter{Appendix for Chapter \ref{chapter:1}}
-\label{app:chapt1}
+\chapter{Additional Mathematical Tools}
+\label{app:1:mathsTools}
 
-In this appendix, blabla.
+In this appendix, we present some mathematical tools that are used in various proofs in this thesis.
+
+% ----------------------------------------------------------------------------
+\section{Concentration Inequalities}
+\label{sec:app1:ConcentrationInequalities}
+
+\TODOL{I probably need to include Hoeffding's inequality, Chernoff's method etc?}
+
+Reference: \cite{Boucheron2013} and maybe the original paper by Hoeffding, Chernoff etc.
+
+% ----------------------------------------------------------------------------
+\section{Useful lemmas}
+\label{sec:app1:UsefulLemmas}
+
+\TODOL{I probably need to include a bunch of useful lemmas, used in various places in the thesis, like Pinsker's inequality, etc.}
+
+% ----------------------------------------------------------------------------
+\section{Results about klUCB}
+\label{sec:app1:resultsklUCB}
+
+\TODOL{I probably need to include some results about klUCB, mainly coming from \cite{KLUCBJournal}, used in various places in the thesis?}

### 2019-04-17 3-Appendices/2-Appendix/appendix2.tex
Cette thèse me semble interminable...

@@
 % ******************************* Thesis Appendix B ****************************
-\chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
+% \chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
+\chapter{About Doubling-Trick for Multi-Armed Bandits}
 \label{app:2:DoublingTricks}
 
-This appendix quickly presents the contributions of another publication that was not presented in the main text of this thesis, \cite{Besson2018DoublingTricks}.
+This appendix quickly presents the contributions of another publication that was not presented in the main text of this thesis.
+We studied doubling tricks and their possible use for multi-armed bandits, between fall 2017 and spring 2018, and we wrote an article \cite{Besson2018DoublingTricks} which was rejected at the COLT 2018 conference, and unfortunately we lacked the time to improve it and resubmit it to another conference.
 
-% \TODOL{
-% I cannot include it raw without more work on the maths, we have to write only the best versions of our results.
+\TODOL{
+I cannot include it raw without more work on the maths, we have to write only the best versions of our results.
 
-% But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!
+But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!
 
-% I need to try DoublingTrick + GLRklUCB, at least experimentally!
-% For the theoretical analysis, it will probably be a huge failure.
-% }
+I need to try DoublingTrick + GLRklUCB, at least experimentally!
+For the theoretical analysis, it will probably be a huge failure.
+}
 
@@
 A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06}.
-We studied doubling tricks between fall 2017 and spring 2018, and we wrote an article \cite{Besson2018DoublingTricks} which was rejected at the COLT 2018 conference, and unfortunately we lacked the time to improve it and resubmit it to another conference.
 %

### 2019-04-17 4-Listofpublications/Listofpublications.tex
Cette thèse me semble interminable...

@@
 \addcontentsline{toc}{chapter}{List of Publications}
+
+\TODOL{J'ai déjà inclus une liste de publication à la fin du Chapitre 1, nécessaire ici aussi ?}
 
@@
   Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02049824}{HAL-02049824}.
+  \cite{Bonnefoi2019WCNC}
 \item
@@
   Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02006825}{HAL-02006825}.
+  \cite{Besson2019WCNC}
 \item
@@
   presented at ICT (\emph{International Conference on Communication}),
-  Saint-Malo, France, June 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}, \href{https://hal.inria.fr/hal-02013866}{HAL-02013866}
+  Saint-Malo, France, June 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}, \href{https://hal.inria.fr/hal-02013866}{HAL-02013866}.
+  \cite{Besson2018ICT}
 \item
@@
   \href{https://hal.inria.fr/hal-01629733}{HAL-01629733}.
+  \cite{Besson2018ALT}
 \item
@@
   avril 2018, \href{https://hal.inria.fr/hal-01705292}{HAL-01705292}.
+  \cite{Besson2018WCNC}
 \item
@@
   \href{https://hal.inria.fr/hal-01575419}{HAL-01575419}, \textbf{Best Paper Award}.
+  \cite{Bonnefoi17}
 \end{etaremune}
@@
 Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}.
-	\item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, March 2017.
+	% \item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, March 2017.
 	\item	R.
@@
   % \href{https://hal.inria.fr/hal-XXX}{HAL-XXX}.
+  \cite{MoyBesson2019}
 \item
@@
   \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
+  \cite{Besson2019Gretsi}
 \item
@@
   \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
+  \cite{Besson2019GLRT}
 \item
@@
   \href{https://SMPyBandits.rtfd.io}{SMPyBandits.rtfd.io}.
+  \cite{SMPyBandits,SMPyBanditsJMLR}
 \item
@@
   \href{https://hal.inria.fr/hal-01736357}{HAL-01736357}.
+  \cite{Besson2018DoublingTricks}
 \item

### 2019-04-17 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
So long

@@
 % ----------------------------------------------------------------------
-
 We compare in this PoC the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
 For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
-Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} shows the results averaged on $10$ repetitions using the same conditions.
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
+    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
+Similar gains of performance were obtained in many different scenarios.}
+    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
+\end{figure}
+
+In Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} above, we show the results averaged on $10$ repetitions using the same conditions.
 %
@@
 Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.
-
-\begin{figure}[!t]
-	\centering
-    \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
-    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
-Similar gains of performance were obtained in many different scenarios.}
-    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
-\end{figure}
 
@@
 
+\begin{figure}[!h]
+	\centering
+    % \includegraphics[width=0.90\textwidth]{Images/screenshotDemoYouTube.png}
+    \includegraphics[width=0.90\textwidth]{2-Chapters/4-Chapter/Images/screenshotDemoYouTube.png}
+    \caption{Screenshot of the \textbf{video} of our demonstration, \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.}
+    \label{fig:42:screenshotDemoYouTube}
+\end{figure}
+
 \paragraph{Video}
@@
 
-\begin{figure}[!h]
-	\centering
-    % \includegraphics[width=0.90\textwidth]{Images/screenshotDemoYouTube.png}
-    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/screenshotDemoYouTube.png}
-    \caption{Screenshot of the \textbf{video} of our demonstration, \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.}
-    \label{fig:42:screenshotDemoYouTube}
-\end{figure}
-
 % \paragraph{Special acknowledgment}

### 2019-04-17 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
So long

@@
 %
-The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length).
+We wrote the following blocks
+1) the \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power),
+2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), 
+and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length).
 All blocks need to know constants about the PHY layer (preamble length and data length).
@@
 Figure~\ref{fig:4app:USRP_TX_SU__v1__simple_grc} below shows the \textbf{IoT dynamic device} flow-graph.
-The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels.
+The hand-written blocks are the
+1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack),
+2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway,
+and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels.
 Most blocks need to know constants about the PHY layer (preamble length and data length).

### 2019-04-17 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
So long

@@
 
-For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1,chapter:4}.
+For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1},\ref{chapter:4}.
 With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
@@
 % -----------------------------------------------------------------
-\subsection{The \Selfish{} approach: a heuristic for the ``no sensing case''}
+\subsection{The \Selfish{} heuristic for the ``no sensing case''}
 \label{sub:5:Selfish}
@@
 \MEGA{} needs a careful tuning of \emph{five} parameters ($c$, $d$, $p_0$, $\alpha$ and $\beta$) to attain reasonable performances.
-No good guideline for tuning them is provided and using cross validation, as suggested,
-can be considered out of the scope of online sequential learning.
+No good guideline for tuning them is provided and using \emph{cross validation}, as suggested by \cite{Avner15},
+can be considered out of the scope of \emph{online} sequential learning.
 %In practice, on a fixed instance, the authors do not indicate how to select the parameters, even with a perfect knowledge of the parameters ($\boldsymbol{\mu}$ and $T$).
@@
 
-% -----------------------------------------------------------------
-\section{Conclusion of our contributions}
-\label{sec:5:conclusion}
-% -----------------------------------------------------------------
-
-To sum up, we presented three variants of Multi-Player Multi-Arm Bandits,
-with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
-For the two easiest models --with sensing--, our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret.
-In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
-
-This work suggests several interesting further research directions.
-First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
-So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
-However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
-Second, it remains an open question to know if a simple decentralized algorithm can be as efficient as \MCTopM{} without knowing $M$ in advance, or in dynamic settings (when $M$ can change in time).
-One could start by proposing variants of our algorithm that are inspired by the \rhoRandEst{} variant of \rhoRand{} proposed by \cite{Anandkumar11}.
-Finally, one could want to strengthen the guarantees obtained in the absence of sensing, that is to know whether logarithmic regret is achievable and to have a better analysis of the \Selfish{} approach.
-Indeed, in most cases, it performs comparably to \RandTopM{} even with limited feedback and without knowing the number of players $M$, which makes it a good candidate for applications to Internet of Things networks.
-
-% -----------------------------------------------------------------
-
-% \paragraph{Acknowledgments}
-% Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.
-
-\paragraph{Reproducibility.}
-The page \texttt{SMPyBandits.GitHub.io/MultiPlayers.html} explains how to reproduce the experiments used for this Chapter.
-
 \begin{framed}

### 2019-04-17 2-Chapters/5-Chapter/chapter5.tex
So long

@@
 
-Finally, we review other extensions of our model, and give a overview of the current state-of-the-art as exhaustive as possible.
+Finally, we review other extensions of our model, and give a overview of the current state-of-the-art.
+For some extensions, we discuss how to adapt our proposal and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
 
@@
 %
-These works assume a prior knowledge of $\Delta$, or other measures of the difficulty of the problem, and as such we do not.
-FIXME
-\TODOL{FIXME écrire ça sans être méchant envers leurs travaux !}
+These works assume a prior knowledge of $\Delta$, or other measures of the difficulty of the problem, and as such we do not find them comparable with the approach chosen here.
 
@@
 %
-The work of \cite{AvnerMannor18} is discussed above for the case with communication between players, and their model is of high interest from the point-of-view of real-world wireless communication protocols.
+The work of \cite{AvnerMannor18} is discussed above for the case with communicating players, and their model is very interesting from the point-of-view of real-world wireless communication protocols.
 
@@
 Similarly to our demonstration of multi-armed bandit learning in an IoT network that we presented in the previous Chapter~\ref{chapter:4} in Section~\ref{sec:4:gnuradio},
-we list in this section some related works, dating back from $2016$, who proposed a similar approach.
+we list here some related works, dating back from $2016$, who proposed a similar approach.
 All the similar works that we are aware of also used USRP boards, and implemented the demonstration using either the Simulink and MATLAB softwares, or the GNU Radio software like we did \cite{Besson2018ICT}.
@@
 \label{sec:5:conclusion}
+% -----------------------------------------------------------------
 
-In this chapter, we saw...
+To sum up, we presented in this Chapter three variants of Multi-Player Multi-Arm Bandits,
+with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
+For the two easiest models --with sensing--, our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret.
+In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
+%
+We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proved to be very active in the last two years.
+For some models, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt our proposal \MCTopM{} to other settings.
+For example, it assumes to know the number of players $M$ before-hand, but we illustrated that previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
+Further works would be required to adapt our theoretical analysis to these various extensions.
 
-Future works include...
+This Chapter suggests several interesting further research directions.
+First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
+So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
+However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
+Similarly to what was done very recently in \cite{wang2019distributed} for communication between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
 
-\newpage  % FIXME
+% Second, it remains an open question to know if a simple decentralized algorithm can be as efficient as \MCTopM{} without knowing $M$ in advance, or in dynamic settings (when $M$ can change in time).
+One could start by proposing variants of our algorithm that are inspired by the \rhoRandEst{} variant of \rhoRand{} proposed by \cite{Anandkumar11}.
+% Finally, one could want to strengthen the guarantees obtained in the absence of sensing, that is to know whether logarithmic regret is achievable.
+
+% -----------------------------------------------------------------
+
+% \paragraph{Acknowledgments}
+% Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.
+
+\paragraph{A note on the simulation code.}
+%
+All the experiments used for this chapter were performed using SMPyBandits,
+and we refer to the following page for details on how to reproduce them,
+\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}
+
+\newpage
 % ----------------------------------------------------------------------------

### 2019-04-17 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Hard to focus
So long

@@
 \TODOL{I need to write a paragraph about using the ``adversarial setting'' for piece-wise stationary.
-It is very important to state that EXP3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with prior knowledge of $T$ and $\Upsilon_T$.}
+It is very important to state that EXP3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with NO prior knowledge of $T$ and $\Upsilon_T$.}
 
@@
 % -----------------------------------------------------------------
-\section{The Piece-Wise Stationary Bandit Model}
+\section{The piece-wise stationary bandit model}
 \label{sec:6:BanditSetting}
@@
 % -----------------------------------------------------------
-\section{Review of Related Works}
+\section{Review of related works}
 \label{sec:6:relatedWork}
@@
 
-\TODOL{Talk about NEWMA and its possible application to generalize/improve D-UCB, \cite{Keriven2018}.}
+\TODOL{Talk about NEWMA and its possible application to generalize/improve D-UCB, \cite{Keriven2018}, if I have the time.}
 
@@
     It is actually not used as such in the proofs of the regret upper bound given below, but a similar result is used and proved in the proofs.
-    \TODOL{We have to give the proof? I was fine with this sentence...}
+    \TODOL{We have to give the proof, or remove this result!}
 \end{proof}
@@
 %-----------------------------------------------------------------------------
-\section{The \GLRklUCB{} Algorithm}
+\section{The \GLRklUCB{} algorithm}
 \label{sec:6:GLRklUCB_Algorithm}
@@
 
-\subsection{Proof for \GLRklUCB{} with Global Changes}
+\subsection{Proof for \GLRklUCB{} with global changes}
 \label{proof:6:mainRegretBoundGlobal}
@@
 
-\section{Experimental Results}
+\section{Experimental results}
 \label{sec:6:NumericalExperiments}

### 2019-04-17 2-Chapters/Conclusion/conclusion.tex
So long

@@
 Abstract for Conclusion.
+\TODOL{Write this!}
 
@@
 
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
 \section{General Conclusion}
+
 In this PhD thesis, blabla.
+
+\TODOL{Write this!}
+
+In \textbf{Chapter \ref{chapter:1}}, blabla.
+
+In \textbf{Chapter \ref{chapter:2}}, blabla.
+
+In \textbf{Chapter \ref{chapter:3}}, blabla.
 
@@
 
+In \textbf{Chapter \ref{chapter:5}}, blabla.
+
+In \textbf{Chapter \ref{chapter:6}}, blabla.
+
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
 \section{Perspectives}
+
 We present in the following the possible areas of future studies.
 
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
 \section{Personal Summary}
@@
 
-J'aimerai conclure en dédicaçant cette thèse à mes deux frères, qui ont été ma plus grande source d'inspiration depuis mon plus jeune âge ; ainsi qu'à la mémoire de XXX XXX qui nous manquera cruellement.
-\textcolor{white}{Je souhaite aussi remercier une fois encore mes amis et amies les plus proches, Hélène, Pierre, Lola, Sélim, Angèle}
+% J'aimerai conclure en dédicaçant cette thèse à mes deux frères, qui ont été ma plus grande source d'inspiration depuis mon plus jeune âge ; ainsi qu'à la mémoire de XXX XXX qui nous manquera cruellement.
+% \textcolor{white}{Je souhaite aussi remercier une fois encore mes amis et amies les plus proches, Hélène, Pierre, Lola, Sélim, Angèle}

### 2019-04-18 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for now.

@@
 For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1},\ref{chapter:4}.
+%
+In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
+We consider $M \leq K$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
+A perfect time and frequency synchronization is assumed,
+and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
+As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
+%
+Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
+Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices.
+Even if there are many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step the number of communicating devices is random and unpredictable.
+
+Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider the easier case of at most $M \leq K$ devices, communicating at each time step.
+Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
+%
 With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
@@
 \emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and further studied shortly after in \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
-Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain)
-and by \cite{Avner15,Avner16} and \cite{Rosenski16} (for \iid{} or piece-wise \iid{} arms).
-%
+Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain),
+and by \cite{Avner15,Avner16} and \cite{Rosenski16} for \iid{} arms.
+%
+In the point of view of the wireless protocol, each time frame is separated as before in a \emph{sensing} phase (during which the device senses for the background traffic),
+an \emph{uplink} phase (during which the device sends a packet to the gateway if it sensed the chosen channel to be free),
+and a \emph{downlink} phase (during which it waits for an \Ack{} from the gateway).
+In this first model, the binary reward is $1$ only if the channel was sensed to be free of background traffic \textbf{and} if \Ack{} was received, and the device has access to both information.
+We present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
+Like in previous works, we consider the centralized system regret (multi-player regret), or simply referred to as regret.
+We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
+We then analyze our \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
+%  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
+We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M$ accessing $K \geq M$ channels.
+
 The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
@@
 Especially, the typical approach that combines a (single-player) bandit algorithm based on the sensing information --to learn the quality of the channels while targeting the best ones-- with a low-complexity decentralized collision avoidance protocol, is no longer possible.
-Our article \cite{Besson2018ALT} was the first to study this other model of multi-player bandits for the ``no sensing'' case.
-
-In this chapter, we take a step back and present the different feedback levels possible for multi-player MAB algorithms.
+%
+Our article \cite{Besson2018ALT} was the first to study this other model of multi-player bandits for the ``no sensing'' case,
+% We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
+% Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, 
+for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
+Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-player MAB problems).
+We do not propose any other efficient algorithm, but our work presented in April~$2018$ inspired two articles shorlty after \cite{LugosiMehrabian18,BoursierPerchet18}.
+They confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
+We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
+
+% In this chapter, we take a step back and present the different feedback levels possible for multi-player MAB algorithms.
 For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees.
@@
 Our study of this algorithm was supported in \cite{Besson2018ALT} by (promising) empirical performance and some first (disappointing) theoretical elements, and it was proved to be inefficient since then.
+
+% Another interesting extension of the studied model is to consider different channels utility for each devices.
+It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
+% This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.
 
@@
 % FIXME
-\begin{framed}
+% \begin{framed}
   \textbf{\textcolor{red}{Warning}.}
@@
   The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which contains both \RhoRand{} and our two proposals \RandTopM{} and \MCTopM{} but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
-\end{framed}
+% \end{framed}
 % FIXME
@@
 
+Regardless of whether sensing is or not possible, we present formally in Section~\ref{sub:5:Selfish} the \Selfish{} policy, as used by all the IoT devices in the first model presented in Chapter~\ref{chapter:4}.
+It does not use an orthogonalization strategy as the collisions are directly accounted for in the \UCB-like indices that are used by each device to select its channel.
+\Selfish{} can also be used under observation model \modeltrois{} --\emph{without sensing}--, and without the knowledge of $M$.
+It was conjectured in \cite{Besson2018ALT} that \Selfish{} can suffer linear regret, and later confirmed in \cite{LugosiMehrabian18,BoursierPerchet18}.
+%
 When sensing is possible, that is under observation models \modelun{} and \modeldeux, most existing strategies build on a \emph{single-player bandit algorithm} (usually an \emph{index policy}) that relies on the sensing information, together with an \emph{orthogonalization strategy} to deal with collisions.
-We present this approach in more details in Section~\ref{sub:5:RandTopM_and_MCTopM} and introduce two new algorithms of this kind, \RandTopM{} and \MCTopM.
-%
-Then, we suggest in Section~\ref{sub:5:Selfish} a completely different approach, called \Selfish{}, that no longer requires an orthogonalization strategy as the collisions are directly accounted for in the indices that are used.
-\Selfish{} can also be used under observation model \modeltrois{} --\emph{without sensing}--, and without the knowledge of $M$.
+Following this approach, we introduce two new algorithms, \RandTopM{} and \MCTopM, in Section~\ref{sub:5:RandTopM_and_MCTopM}.
+
+% -----------------------------------------------------------------
+\subsection{The \Selfish{} heuristic for the ``no sensing case''}
+\label{sub:5:Selfish}
+
+Under observation model \modeltrois{} no sensing information is available and the previous algorithms cannot be used, as the sum of sensing information $S_k^j(t)$ and thus the empirical mean $\widehat{\mu}_k^j(t)$ cannot be computed, hence neither the indices $g_k^j(t)$.
+However, one can still define a notion of \emph{empirical reward} received from arm $k$ by player $j$, by introducing
+%
+\vspace*{-5pt}
+\begin{equation}
+  \widetilde{S_k}^j(t) = \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
+  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) := \widetilde{S_k}^j(t) \;/\; T_k^j(t).
+\end{equation}
+
+Note that $\widetilde{\mu_k}^j(t)$ is no longer meant to be an unbiased estimate of $\mu_k$ as it also takes into account the collision information, that is present in the reward.
+Based on this empirical reward, one can similarly defined modified indices as
+%
+\begin{equation}\label{eq:5:indexTilde}
+  \widetilde{g_k}^j(t) = \begin{cases}
+      \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
+      &\text{for } \UCB, \\
+      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widetilde{\mu_k}^j(t), q) \leq f(t) \right\}
+      &\text{for } \klUCB.
+  \end{cases}
+\end{equation}
+
+Given any of these two index policies (\UCB{} or \klUCB), the \Selfish{} algorithm is defined by,
+%
+\begin{equation}
+  A^j(t) = \argmax{ k \in \{1,\dots,K\}} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
+\end{equation}
+The name comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
+The reason that this may work precisely comes from the fact that $\widetilde{g_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
+but some hybrid index that simultaneously increases when a transmission occurs and decreases when a collision occurs.
+
+This behavior is easier to be understood for the case of \Selfish-\UCB{} in which, letting $N_k^{j,C}(t) = \sum_{s=1}^t \indic(C^j(t))$ be the number of collisions on arm $k$, one can show that the hybrid \Selfish{} index induces a penalty proportional to the fraction of collision on this arm and the quality of the arm itself:
+\begin{equation}
+  \widetilde{g_k}^j(t) = g_k^j(t) -
+  \underbrace{\left(\frac{N_k^{j,C}(t)}{N_k^j(t)}\right)}_{\text{fraction of collisions}}
+  \underbrace{\left(\frac{1}{N_k^{j,C}(t)}\sum_{t=1}^{T}Y_{A^j(t),t} \indic(C^j(t)) \indic(A^j(t) = k)\right)}_{\text{estimate of } \mu_k }.
+\end{equation}
+
+From a bandit perspective, it looks like each player is using a stochastic bandit algorithm (\UCB{} or \klUCB) when interacting with $K$ arms that give a feedback (the reward, and not the sensing information) that is far from being \iid{} from some distribution, due to the collisions.
+%
+As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like $\mathrm{EXP3}$ \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
+%
+However, we found out empirically that \Selfish{} is doing surprisingly well when using \UCB-like indexes, greatly outperforming \Selfish{} based on $\mathrm{EXP3}$,
+like what we found in Chapter~\ref{chapter:4} in harder settings.
+
+We show in Section~\ref{sub:5:SelfishFails} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
+which precludes the possibility of a logarithmic regret for any problem.
+In most cases, it empirically performs similarly to all the algorithms described before,
+and usually outperforms \rhoRand,
+even if it neither exploits the sensing information, nor the knowledge of the number of players $M$.
+%
+Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or cannot be considered.
+Thus we propose next our main contribution, the \MCTopM{} algorithm, proved to be asymptotically optimal for the identification of suboptimal arms (when using \klUCB), attaining order-optimal logarithmic regret, and outperforming all the other algorithms for the ``sensing case''.
+
+% -----------------------------------------------------------------
 
@@
 % -----------------------------------------------------------------
-\subsection{The \Selfish{} heuristic for the ``no sensing case''}
-\label{sub:5:Selfish}
-
-Under observation model \modeltrois{} no sensing information is available and the previous algorithms cannot be used, as the sum of sensing information $S_k^j(t)$ and thus the empirical mean $\widehat{\mu}_k^j(t)$ cannot be computed, hence neither the indices $g_k^j(t)$.
-However, one can still define a notion of \emph{empirical reward} received from arm $k$ by player $j$, by introducing
-%
-\vspace*{-5pt}
-\begin{equation}
-  \widetilde{S_k}^j(t) = \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
-  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) := \widetilde{S_k}^j(t) \;/\; T_k^j(t).
-\end{equation}
-
-Note that $\widetilde{\mu_k}^j(t)$ is no longer meant to be an unbiased estimate of $\mu_k$ as it also takes into account the collision information, that is present in the reward.
-Based on this empirical reward, one can similarly defined modified indices as
-%
-\begin{equation}\label{eq:5:indexTilde}
-  \widetilde{g_k}^j(t) = \begin{cases}
-      \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
-      &\text{for } \UCB, \\
-      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widetilde{\mu_k}^j(t), q) \leq f(t) \right\}
-      &\text{for } \klUCB.
-  \end{cases}
-\end{equation}
-
-Given any of these two index policies (\UCB{} or \klUCB), the \Selfish{} algorithm is defined by,
-%
-\begin{equation}
-  A^j(t) = \argmax{ k \in \{1,\dots,K\}} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
-\end{equation}
-The name comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
-The reason that this may work precisely comes from the fact that $\widetilde{g_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
-but some hybrid index that simultaneously increases when a transmission occurs and decreases when a collision occurs.
-
-This behavior is easier to be understood for the case of \Selfish-\UCB{} in which, letting $N_k^{j,C}(t) = \sum_{s=1}^t \indic(C^j(t))$ be the number of collisions on arm $k$, one can show that the hybrid \Selfish{} index induces a penalty proportional to the fraction of collision on this arm and the quality of the arm itself:
-\begin{equation}
-  \widetilde{g_k}^j(t) = g_k^j(t) -
-  \underbrace{\left(\frac{N_k^{j,C}(t)}{N_k^j(t)}\right)}_{\text{fraction of collisions}}
-  \underbrace{\left(\frac{1}{N_k^{j,C}(t)}\sum_{t=1}^{T}Y_{A^j(t),t} \indic(C^j(t)) \indic(A^j(t) = k)\right)}_{\text{estimate of } \mu_k }.
-\end{equation}
-
-From a bandit perspective, it looks like each player is using a stochastic bandit algorithm (\UCB{} or \klUCB) when interacting with $K$ arms that give a feedback (the reward, and not the sensing information) that is far from being \iid{} from some distribution, due to the collisions.
-%
-As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like $\mathrm{EXP3}$ \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
-%
-However, we found out empirically that \Selfish{} is doing surprisingly well,
-like what we found in Chapter~\ref{chapter:4} in harder settings.
-% as already noted by \cite{Bonnefoi17}, who did some experiments in the context of IoT applications.
-
-We show in Section~\ref{sec:5:upperbounds} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
-which precludes the possibility of a logarithmic regret for any problem.
-In most cases, it empirically performs similarly to all the algorithms described before,
-and usually outperforms \rhoRand,
-even if it neither exploits the sensing information, nor the knowledge of the number of players $M$.
-%
-Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or cannot be considered.
-
-% -----------------------------------------------------------------
-
-% -----------------------------------------------------------------
 \section{Theoretical elements, and order optimal regret for \MCTopM}
@@
 with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, the histograms in Figure~\ref{fig:5:selfish_fail1} suggests that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} can be very large.
-We provide a discussion in Appendix~\ref{app:5:SelfishFails} about when such situations may happen, including a conjectured (constant, but small) lower bound on the probability that \Selfish{} experience collision almost at every round.
+
+In the Appendix~E of \cite{Besson2018ALT},
+we discussed about when such situations may happen, including a conjectured (constant, but small) lower bound on the probability that \Selfish{} experience collision almost at every round.
 This result would then prevent \Selfish{} from having a logarithmic regret.
@@
 of whether logarithmic regret is at all possible for the ``no sensing'' case (model \modeltrois).
+They confirmed that \Selfish-\UCB{} can indeed suffer linear regret, and proposed an algorithm based on a ``communication trick'' to achieve logarithmic regret for this harder model \modeltrois.
 
@@
           \hline
-          Centralized multiple-play \UCB{} & $321 \pm 30$ & $233 \pm 25$ & $0$ \\
-          Centralized multiple-play \klUCB{} & $94 \pm 17$ & $68 \pm 18$ & $0$ \\
+          Centralized multiple-play with \UCB{} & $321 \pm 30$ & $233 \pm 25$ & $0$ \\
+          Centralized multiple-play with \klUCB{} & $94 \pm 17$ & $68 \pm 18$ & $0$ \\
           \hline
-          \Selfish-\UCB{} & $1263 \pm 157$ & $3694 \pm 387$ & $12420 \pm 404$ \\
-          \Selfish-\klUCB{} & $243 \pm 31$ & $743 \pm 113$ & $3005 \pm 492$ \\
+          \Selfish{} with \UCB{} & $1263 \pm 157$ & $3694 \pm 387$ & $12420 \pm 404$ \\
+          \Selfish{} with \klUCB{} & $243 \pm 31$ & $743 \pm 113$ & $3005 \pm 492$ \\
           \hline
-          \RhoRand-\UCB{} & $1455 \pm 208$ & $4775 \pm 463$ & $11794 \pm 1083$ \\
-          \RhoRand-\klUCB{} & $394 \pm 96$ & $2385 \pm 412$ & $7057 \pm 1053$ \\
+          \RhoRand{} with \UCB{} & $1455 \pm 208$ & $4775 \pm 463$ & $11794 \pm 1083$ \\
+          \RhoRand{} with \klUCB{} & $394 \pm 96$ & $2385 \pm 412$ & $7057 \pm 1053$ \\
           \hline
-          \RandTopM-\UCB{} & $1020 \pm 92$ & $2899 \pm 418$ & $470 \pm 346$ \\
-          \RandTopM-\klUCB{} & $258 \pm 43$ & $902 \pm 234$ & $551 \pm 520$ \\
+          \RandTopM{} with \UCB{} & $1020 \pm 92$ & $2899 \pm 418$ & $470 \pm 346$ \\
+          \RandTopM{} with \klUCB{} & $258 \pm 43$ & $902 \pm 234$ & $551 \pm 520$ \\
           \hline
-          \MCTopM-\UCB{} & $980 \pm 76$ & $1466 \pm 132$ & $43 \pm 13$ \\
-          \MCTopM-\klUCB{} & $\mathbf{248 \pm 40}$ & $\mathbf{410 \pm 54}$ & $\mathbf{42 \pm 10}$ \\
+          \MCTopM{} with \UCB{} & $980 \pm 76$ & $1466 \pm 132$ & $43 \pm 13$ \\
+          \MCTopM{} with \klUCB{} & $\mathbf{248 \pm 40}$ & $\mathbf{410 \pm 54}$ & $\mathbf{42 \pm 10}$ \\
           \hline
@@
   % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
+      \includegraphics[width=0.90\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
   % \end{subfigure}
@@
   % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
+      \includegraphics[width=0.90\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
   % \end{subfigure}
@@
   % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
+      \includegraphics[width=0.90\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
   % \end{subfigure}

### 2019-04-18 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Hard to focus
Done for now.

@@
 % -----------------------------------------------------------------
-\subsection{Lower Bound: Proof of Theorem~\ref{thm:5:BetterLowerBound}}
+\subsection{Lower bound: proof of Theorem~\ref{thm:5:BetterLowerBound}}
 \label{proof:5:BetterLowerBound}
@@
 
-% -----------------------------------------------------------------
-% -----------------------------------------------------------------
-\subsection{Additional Discussions on \Selfish}
-\label{app:5:SelfishFails}
+% % -----------------------------------------------------------------
+% % -----------------------------------------------------------------
+% \subsection{Additional discussions on \Selfish}
+% \label{app:5:SelfishFails}
 
-\TODOL{Maybe we can completely remove this, and just points to the paper \cite{Besson2018ALT}?}
+% \TODOL{Maybe we can completely remove this, and just points to the paper \cite{Besson2018ALT}?}
 
-% OBSERVATION
-% Empirically it works fine in average of regrets on lots of simulations, but on worst case regret (or in a histogram) we can see it does not have a logarithmic average regret
-%
-% BAD
-% Explain on a small example ($K=2, M=3$ or $K=M=2$) that we found both manually, empirically and with a formal game tree exploration a small probability of failure, for both \Selfish-\UCB{} and \Selfish-\klUCB...
-%
-% a small probability of failure... $K=2, M=3$ or $K=M=2$. (idée: comparer \Selfish{} et \rhoRand{} en terme d'histogramme de la distribution de $R_T$)
-%
-% DONE estimer la proba ? ou la calculer avec le calcul formel ? ca c'est fait c'est bon
-%
-% a curiosity: when $K=M=2$ and $\mu_1 = \mu_2$, using \Selfish{} KL-UCB does orthogonalize !
+% % OBSERVATION
+% % Empirically it works fine in average of regrets on lots of simulations, but on worst case regret (or in a histogram) we can see it does not have a logarithmic average regret
+% %
+% % BAD
+% % Explain on a small example ($K=2, M=3$ or $K=M=2$) that we found both manually, empirically and with a formal game tree exploration a small probability of failure, for both \Selfish-\UCB{} and \Selfish-\klUCB...
+% %
+% % a small probability of failure... $K=2, M=3$ or $K=M=2$. (idée: comparer \Selfish{} et \rhoRand{} en terme d'histogramme de la distribution de $R_T$)
+% %
+% % DONE estimer la proba ? ou la calculer avec le calcul formel ? ca c'est fait c'est bon
+% %
+% % a curiosity: when $K=M=2$ and $\mu_1 = \mu_2$, using \Selfish{} KL-UCB does orthogonalize !
 
-As said before, analyzing \Selfish{} is harder,
-but for instance one can prove that it yields constant
-collisions and regret for the trivial case of $\mu_1=\mu_2=1$ and $M=2$.
-%
-Empirically, when \Selfish{} is compared to the other algorithms,
-% with \UCB{} or \klUCB{} indices,
-% and the average regret is considered (for let say $1000$ repetitions),
-it is hard to find a case when \Selfish{} performs badly, as its (empirical average) regret always appeared logarithmic.
-%
-But an issue of only visualizing the empirical average regret for a certain number of repetitions is
-that if a certain ``bad'' run happens only with small probability, it is possible that it never happened
-in a simulation, or that it happened a few times but not enough to make the average regret look non-logarithmic\footnote{For instance, $0.999 \log(T) + 0.001 T$ looks more like $\log(T)$ than $T$ so an event yielding linear regret with ``small'' probability $10^{-3}$ cannot be observed from a plot showing the average regret.}.
-%
-This is why the distribution of regret, at the end of the simulations, $R_T$, is also displayed (in Appendix~\ref{app:5:moreplots}).
-%
-In a simple problem, with $M=2$ or $M=3$ players competing for $K=3$ arms,
-for instance with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
-the histogram in Figure~\ref{fig:5:selfish_fail1} shows that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} is not small (and appears linear).
-%
-%
-Additionally, Figure~\ref{fig:5:selfish_fail2} shows that \Selfish-\klUCB{} also has bad performance against random uniform problems $\boldsymbol{\mu}\in[0,1]^K$,
-for $M=2$ or $3$ and $K=3$, in a lot of cases.
-In comparison, the others algorithms seem to have a logarithmic regret (for $M=2$ and all algorithms)
-or even a constant regret (for $M=3$ and \RandTopM{} and \MCTopM).
+% As said before, analyzing \Selfish{} is harder,
+% but for instance one can prove that it yields constant
+% collisions and regret for the trivial case of $\mu_1=\mu_2=1$ and $M=2$.
+% %
+% Empirically, when \Selfish{} is compared to the other algorithms,
+% % with \UCB{} or \klUCB{} indices,
+% % and the average regret is considered (for let say $1000$ repetitions),
+% it is hard to find a case when \Selfish{} performs badly, as its (empirical average) regret always appeared logarithmic.
+% %
+% But an issue of only visualizing the empirical average regret for a certain number of repetitions is
+% that if a certain ``bad'' run happens only with small probability, it is possible that it never happened
+% in a simulation, or that it happened a few times but not enough to make the average regret look non-logarithmic\footnote{For instance, $0.999 \log(T) + 0.001 T$ looks more like $\log(T)$ than $T$ so an event yielding linear regret with ``small'' probability $10^{-3}$ cannot be observed from a plot showing the average regret.}.
+% %
+% This is why the distribution of regret, at the end of the simulations, $R_T$, is also displayed (in Appendix~\ref{app:5:moreplots}).
+% %
+% In a simple problem, with $M=2$ or $M=3$ players competing for $K=3$ arms,
+% for instance with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
+% the histogram in Figure~\ref{fig:5:selfish_fail1} shows that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} is not small (and appears linear).
+% %
+% %
+% Additionally, Figure~\ref{fig:5:selfish_fail2} shows that \Selfish-\klUCB{} also has bad performance against random uniform problems $\boldsymbol{\mu}\in[0,1]^K$,
+% for $M=2$ or $3$ and $K=3$, in a lot of cases.
+% In comparison, the others algorithms seem to have a logarithmic regret (for $M=2$ and all algorithms)
+% or even a constant regret (for $M=3$ and \RandTopM{} and \MCTopM).
 
-The intuition behind these configurations when \Selfish{} performs poorly is the following,
-if all players use the same algorithm and the same indices.
-If two players $i$ and $j$ have exactly the same vectors $[\widetilde{S_k}^i(t)] = [\widetilde{S_k}^j(t)]$
-and $[T_k^i(t)] = [T_k^j(t)]$ at some time step $t \geq 1$,
-with different values for each $k$
-and so that the index vectors $\mathbf{g}^j(t) = \mathbf{g}^{i}(t)$ have different values for each arm $k$,
-then both players will take the same decisions at time $t$, and collide.
-Colliding does not change $[\widetilde{S_k}^j(t+1)]$ but increase one value in $[T_k^j(t+1)]$ by $1$.
-Then at the next step the same conditions on $\widetilde{S}^j$ and $N^j$ are preserved,
-and if the same condition on $\mathbf{g}^j(t+1)$ are also preserved,
-the two players will continue to collide.
-%
-We did not succeed in proving mathematically
-that the preservation of the first hypothesis on $\widetilde{S}^j$ and $N^j$
-implies the preservation of the hypothesis on index $\mathbf{g}^j$,
-but numerically it turns out to be always the case:
-and so two players colliding in such a setting will continue to do so infinitely:
-we denote such configurations as \emph{absorbing}.
+% The intuition behind these configurations when \Selfish{} performs poorly is the following,
+% if all players use the same algorithm and the same indices.
+% If two players $i$ and $j$ have exactly the same vectors $[\widetilde{S_k}^i(t)] = [\widetilde{S_k}^j(t)]$
+% and $[T_k^i(t)] = [T_k^j(t)]$ at some time step $t \geq 1$,
+% with different values for each $k$
+% and so that the index vectors $\mathbf{g}^j(t) = \mathbf{g}^{i}(t)$ have different values for each arm $k$,
+% then both players will take the same decisions at time $t$, and collide.
+% Colliding does not change $[\widetilde{S_k}^j(t+1)]$ but increase one value in $[T_k^j(t+1)]$ by $1$.
+% Then at the next step the same conditions on $\widetilde{S}^j$ and $N^j$ are preserved,
+% and if the same condition on $\mathbf{g}^j(t+1)$ are also preserved,
+% the two players will continue to collide.
+% %
+% We did not succeed in proving mathematically
+% that the preservation of the first hypothesis on $\widetilde{S}^j$ and $N^j$
+% implies the preservation of the hypothesis on index $\mathbf{g}^j$,
+% but numerically it turns out to be always the case:
+% and so two players colliding in such a setting will continue to do so infinitely:
+% we denote such configurations as \emph{absorbing}.
 
-We wrote a script
-\footnote{~See online \texttt{\url{smpybandits.github.io/docs/complete_tree_exploration_for_MP_bandits.html}}}
-that explores formally all the possible runs,
-up-to a certain small time horizon,
-by exploring the complete \emph{game tree},
-of possible (random) rewards from the $K$ arms
-and (random) actions from the $M$ players, up-to a small depth of let say $T=8$.
-%
-Such game tree becomes quickly very large, but this was enough to confirm that with
-a certain small probability, function of $\mu_1,\dots,\mu_K$,
-two players can arrive in just a few steps in a ``bad'' absorbing configuration.
-%
-For instance, for only $K=2$ arms, the following game tree in Figure~\ref{fig:5:oneGameTree_SelfishKLUCB} illustrates the first $3$ steps that can lead to $2$ absorbing configurations.
-Using symbolic computations, the probability of reaching any of them was found to be
-$\geq \mu_1^2(1-\mu_2)^2/2 + \mu_2^2(1-\mu_1)^2/2$ for \Selfish-\UCB.
-That is far from being negligible, as it evaluates to $0.328$ for $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
-and a numerical simulation on $1000$ runs found $325$ cases of bad performance.
-%
-The same game tree exploration can be made for \Selfish-\klUCB,
-but so far we were not able to justify why it experiences fewer cases of bad performances even though our software found the same (lower bound on) failure probability.
+% We wrote a script
+% \footnote{~See online \texttt{\url{smpybandits.github.io/docs/complete_tree_exploration_for_MP_bandits.html}}}
+% that explores formally all the possible runs,
+% up-to a certain small time horizon,
+% by exploring the complete \emph{game tree},
+% of possible (random) rewards from the $K$ arms
+% and (random) actions from the $M$ players, up-to a small depth of let say $T=8$.
+% %
+% Such game tree becomes quickly very large, but this was enough to confirm that with
+% a certain small probability, function of $\mu_1,\dots,\mu_K$,
+% two players can arrive in just a few steps in a ``bad'' absorbing configuration.
+% %
+% For instance, for only $K=2$ arms, the following game tree in Figure~\ref{fig:5:oneGameTree_SelfishKLUCB} illustrates the first $3$ steps that can lead to $2$ absorbing configurations.
+% Using symbolic computations, the probability of reaching any of them was found to be
+% $\geq \mu_1^2(1-\mu_2)^2/2 + \mu_2^2(1-\mu_1)^2/2$ for \Selfish-\UCB.
+% That is far from being negligible, as it evaluates to $0.328$ for $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$,
+% and a numerical simulation on $1000$ runs found $325$ cases of bad performance.
+% %
+% The same game tree exploration can be made for \Selfish-\klUCB,
+% but so far we were not able to justify why it experiences fewer cases of bad performances even though our software found the same (lower bound on) failure probability.
 
-\begin{figure}[h!]
-    \centering
-    % \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex}
-    \includegraphics[width=0.70\textwidth]{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing.pdf}
-    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing__leafs.pdf}
-    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing.pdf}
-    % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs.pdf}
-    \caption[Trajectories of $M=2$ players using \Selfish-\UCB{} for $K=2$ arms.]{For $K=2$ arms and $M=2$ players using \Selfish-\UCB, for depth$=3$: $2$ \textbf{\textcolor{red}{absorbing configurations}}.
+% \begin{figure}[h!]
+%     \centering
+%     % \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex}
+%     \includegraphics[width=0.70\textwidth]{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing.pdf}
+%     % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_UCB__absorbing__leafs.pdf}
+%     % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing.pdf}
+%     % \includegraphics[width=0.70\textwidth]{Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs.pdf}
+%     \caption[Trajectories of $M=2$ players using \Selfish-\UCB{} for $K=2$ arms.]{For $K=2$ arms and $M=2$ players using \Selfish-\UCB, for depth$=3$: $2$ \textbf{\textcolor{red}{absorbing configurations}}.
 Each rectangle represents a configuration, as the matrix $[[\widetilde{S_k}^j(t) / T_k^j(t) ]_j]_k$.
@@
 Transitions are labeled with their probabilities.}
-    \label{fig:5:oneGameTree_SelfishKLUCB}
-\end{figure}
+%     \label{fig:5:oneGameTree_SelfishKLUCB}
+% \end{figure}
 
-From the structure of such game tree, we conjecture that the probability of reaching absorbing configurations (before a certain time $t$) is always lower-bounded
-by a polynomial function of $\mu_1,\dots,\mu_K$ and $1-\mu_1,\dots,1-\mu_K$,
-of degree at most $t$ in each variable.
-% with coefficients smaller than $1$.
-% and without simple factors (\ie, $\mu_k^d$).
-%
-As such, the lower bound on probability of failures should decrease when $K$ and $M$ increase, and this is coherent with the experiments for $K=9$ or $K=17$ (see Figures~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} and \ref{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}),
-where \Selfish{} is shown to be uniformly more efficient than \rhoRand.
-%
-Of course, one cannot run an infinite number of simulations, and the smaller the probability of failure, the less likely it is to observe a failure
-in a finite number of runs.
+% From the structure of such game tree, we conjecture that the probability of reaching absorbing configurations (before a certain time $t$) is always lower-bounded
+% by a polynomial function of $\mu_1,\dots,\mu_K$ and $1-\mu_1,\dots,1-\mu_K$,
+% of degree at most $t$ in each variable.
+% % with coefficients smaller than $1$.
+% % and without simple factors (\ie, $\mu_k^d$).
+% %
+% As such, the lower bound on probability of failures should decrease when $K$ and $M$ increase, and this is coherent with the experiments for $K=9$ or $K=17$ (see Figures~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} and \ref{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}),
+% where \Selfish{} is shown to be uniformly more efficient than \rhoRand.
+% %
+% Of course, one cannot run an infinite number of simulations, and the smaller the probability of failure, the less likely it is to observe a failure
+% in a finite number of runs.
 
-% IDEA
-\paragraph{Ideas to fix \Selfish{} ?}
-%
-It could be possible to change the \Selfish{} algorithm to add a way to escape such absorbing trajectories.
-For instance one could imagine that after seen seeing, \eg, $x=10$ collisions in a row,
-a certain random action could be taken by the players.
-These tricks can work empirically in some cases,
-but they are harder to analyze formally,
-and it is hard to tune the parameters (here $x$, but possibly more),
-and we do not find such tricks to be promising from a theoretical point-of-view.
+% % IDEA
+% \paragraph{Ideas to fix \Selfish{} ?}
+% %
+% It could be possible to change the \Selfish{} algorithm to add a way to escape such absorbing trajectories.
+% For instance one could imagine that after seen seeing, \eg, $x=10$ collisions in a row,
+% a certain random action could be taken by the players.
+% These tricks can work empirically in some cases,
+% but they are harder to analyze formally,
+% and it is hard to tune the parameters (here $x$, but possibly more),
+% and we do not find such tricks to be promising from a theoretical point-of-view.
 
@@
 % -----------------------------------------------------------------
-\subsection{Additional Figures}
+\subsection{Additional figures}
 \label{app:5:moreplots}
 
-The plots missing from Section~\ref{sec:5:experiments} are included here,
-as well as some additional numerical results.
+We include here plots missing from Section~\ref{sec:5:experiments},
+and some additional numerical results.
 
@@
   \centering
-  \includegraphics[width=0.70\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
+  \includegraphics[width=0.85\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
   \caption[Comparison of our lower bound against the one from \cite{Zhao10}]{Comparison of our lower bound against the one from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
@@
 The lower bounds (in black) are $C(\boldsymbol{\mu}, M) \log t$, the dashed line
-for \citeauthor{Zhao10}'s lower bound, and the continuous line is our lower bound.
+for the lower bound from \cite{Zhao10}, and the continuous line is our lower bound.
 %

### 2019-04-18 2-Chapters/5-Chapter/chapter5.tex
Done for now.

@@
 
+\newpage
 \paragraph{Abstract}
@@
 In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
-We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
-A perfect time and frequency synchronization is assumed,
-and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
-As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
-%
-Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
-Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices.
-Even if there are many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step the number of communicating devices is random and unpredictable.
-
-Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider the easier case of at most $M \leq K$ devices, communicating at each time step.
-Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
+We restrict to the easier case of at most $M \leq K$ devices in a network with $K$ channels, because it was found very hard to analyze the aforementioned IoT network model.
 %
 We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
-Each time frame is separated as before in a \emph{sensing} phase (during which the device senses for the background traffic),
-an \emph{uplink} phase (during which the device sends a packet to the gateway if it sensed the chosen channel to be free),
-and a \emph{downlink} phase (during which it waits for an \Ack{} from the gateway).
-In this first model, the binary reward is $1$ only if the channel was sensed to be free of background traffic \textbf{and} if \Ack{} was received, and the device has access to both information.
 For this first model, we present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
-Like in previous works, we consider the centralized system regret (multi-player regret), or simply referred to as regret.
+Like in previous works, we consider the centralized system regret (multi-player regret).
 We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
@@
 %  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
-We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M$ accessing $K \geq M$ channels.
+We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M \leq K$ accessing $K$ channels.
 
@@
 Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
-Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-player MAB problems).
-We did not propose any other efficient algorithm, but our work presented in April~$2018$ inspired some other articles \cite{LugosiMehrabian18,BoursierPerchet18}.
-These two articles confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
-We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
-
-Another interesting extension of the studied model is to consider different channels utility for each devices.
-It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
-This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.
+Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, we conjectured that it suffers from linear regret, and it was later confirmed by two works inspired by our article \cite{LugosiMehrabian18,BoursierPerchet18}.
+They both proposed new algorithms, proved to achieve logarithmic regret under the no sensing model.
+We include numerical simulations to compare their proposals, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
 
@@
 
-Before concluding this chapter, we give a literature review of the many extensions of the three models presented above in Section~\ref{sec:5:model}.
+Before concluding this chapter, we give a literature review of some of the many of the three models presented above in Section~\ref{sec:5:model}.
 We wanted to highlight that the community was quite active on research on multi-player bandits in the last $10$ years, but especially active since last spring $2018$.
 For instance, our article \cite{Besson2018ALT} was the first to propose the ``no sensing'' case, and it was studied by (at least) two independent group of researchers since its publication in April $2018$ \cite{BoursierPerchet18,LugosiMehrabian18}.
-Another example is the model with different arm means among players, who was studied in \cite{Anandkumar10,Kalathil12} and then more recently in \cite{Bistritz18,KaufmannAbbas19}.
+Another example is the model with different arm means among players, who was initially studied in \cite{Anandkumar10,Kalathil12} and then more recently in \cite{Bistritz18,KaufmannAbbas19}.
 %
-By following actively the research on multi-player bandit models since the last three years, we believe to be able to give an exhaustive literature review, but it is of course possible that we missed an important research work.
+% FIXME écrit pas ça c'est inutile !
+% By following actively the research on multi-player bandit models since the last three years, we believe to be able to give an exhaustive literature review, but it is of course possible that we missed an important research work.
 In this case, it is our mistake and not an intentional choice as we included all the different extensions of our models that we were aware of.
@@
 Finally, we can also be interested by models where $M$ stays fixed, but the environment evolves, for instance with abrupt changes in the means of arms.
-This last model makes a good connexion between this chapter and the next one, and an exciting future work is to further study this model in order to tackle this kind of problems by merging our proposals for stationary multi-player bandits and for non-stationary single-player bandits.
+This last model makes a good connexion between this chapter and the next one, and an exciting future work is to further study this model in order to tackle this kind of problems by merging our contributions for stationary multi-player bandits and for non-stationary single-player bandits.
 
@@
 %
-\TODOL{Write about half a page on the article \cite{BoursierPerchet18}.}
+% \TODOL{Write about half a page on the article \cite{BoursierPerchet18}.}
 
@@
 For the same example, the ``doubling trick'' technique can be used to obtain a regret upper bound for an algorithm unaware of the value of $T$, within a constant multiplicative factor of the upper bound given for the algorithm aware of $T$ (Section~4.1).
-Because the regret bound is of the form $\bigO{\sqrt{T} \log(T)}$, a simple ``doubling trick'' of increasing horizons of lengths $T_i = 2^i$ works well, as proposed in \cite{CesaLugosi06} and as studied in depth in our article \cite{Besson2018DoublingTricks}.
+Because the regret bound is of the form $\bigO{\sqrt{T} \log(T)}$, a simple ``doubling trick'' of increasing horizons of lengths $T_i = 2^i$ works well, as proposed in \cite{CesaLugosi06} and as studied in depth in our article \cite{Besson2018DoublingTricks}, and quickly presented in Appendix~\ref{app:2:DoublingTricks}.
+%
 They also study in Section~4.3 an extension of the model for the case with more players than arms, but we do not give more details on this aspect.
@@
 As we conjectured, there is a gap between the theoretical result and its practical consequence, as the Theorem~4 they gave is only valid for real-valued number, and not for hardware-represented floating point number.
-Their proof is supporting what we illustrated in Figure~\ref{fig:5:oneGameTree_SelfishKLUCB}.
+Their proof is supporting what we illustrated in Figure~4 in \cite{Besson2018ALT}.
 Their argument is essentially to prove that for \Selfish-powered players using the simple \UCB-indexes, with a probability $p$ at time $t$ (both independent from $T$), two players might have the same number of pulls and the same observed rewards for each arm.
@@
 
-\TODOL{A conclusion from this sub-section presenting this model?}
+% \TODOL{A conclusion from this sub-section presenting this model?}
 
@@
 
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+% \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 
-Note that we have not yet implemented this extension in SMPyBandits \cite{SMPyBandits}, but we are interested to do it, and it is one of the major future work left on our library\footnote{See the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/124}}}.
+% \cite{KumarYadav2018}
 
-\cite{KumarYadav2018}
-
-Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
-Sumit J Darak, Manjesh K. Hanawal
-https://arxiv.org/pdf/1812.11651
-\cite{DarakHanawal18}
+% Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
+% Sumit J Darak, Manjesh K. Hanawal
+% https://arxiv.org/pdf/1812.11651
+% \cite{DarakHanawal18}
 
@@
 \cite{KaufmannAbbas19}
-
-Note that we have not yet implemented this extension in SMPyBandits \cite{SMPyBandits}, but we are interested to do it, and it is one of the major future work left on our library\footnote{See the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/185}}}.
 
@@
 
-Note that we have not yet implemented this extension in SMPyBandits \cite{SMPyBandits}, but we are interested to do it, and it is one of the major future work left on our library\footnote{See the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/120}}}.
-It would not be too difficult to do, we just lacked some time.
-
 % ----------------------------------------------------------------------------
@@
 % FIXME continue!
-
+%
 %
@@
 By diving into the details of the modulation and giving explicitly the form of the uplink messages sent by the devices, the authors are able to set-up two different uplink packets, to efficiently perform the auction phase (see Figure~2).
-They prove a regret upper-bound of the order $R_T = \bigO{\log(T)}$, with no special care regarding the constants, and we can also note that their algorithm require a prior knowledge of $\Delta_{\min}$ a problem-dependent constant (Theorem~3).
+They prove a regret upper-bound of the order $R_T = \bigO{\log(T)}$, with no special care regarding the constants, but we can also note that their algorithm require a prior knowledge of $\Delta_{\min}$ a problem-dependent constant (Theorem~3).
 
@@
 Without a more serious analysis, we conjecture that it should not be difficult to merge the regret lower-bound for stationary \emph{multi-players} MAB (Theorem~\ref{thm:5:BetterLowerBound}) and the lower-bound for \emph{piece-wise stationary} single-player MAB from \cite{Garivier11UCBDiscount}.
-We conjecture that any reasonable decentralized bandit algorithm must suffer a regret at least $M \sqrt{K \Upsilon_T T}$ for $M \leq K$ players, $K$ arms, and $\Upsilon_T$ stationary intervals.
+We conjecture that any reasonable decentralized bandit algorithm must suffer a regret at least $\Omega(M \sqrt{K \Upsilon_T T})$ for $M \leq K$ players, $K$ arms, and $\Upsilon_T$ stationary intervals.
 
@@
 \\
+\indent
 If $\Upsilon_T$ is bounded by $\Upsilon_T = \bigO{T^{\gamma}}$, for a known $\gamma$ but an unknown $\Upsilon_T$, they prove for instance in Theorem~2 that the expected cumulative regret of their RR-SW-UCB\# algorithm is bounded by $R_T = \bigO{T^{\frac{1+\gamma}{2}} \log(T)}$ (what we call the centralized system regret).
@@
 %
+\indent
 And finally, even though numerical simulations in their papers \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} are interesting and confirm the regret upper-bounds, they do not compare with any other algorithm, and it is left as an interesting future work to study this direction in more details.
@@
 In \cite{AlaturLevyKrause19}, the authors essentially consider the piece-wise stationary as an easier case of an adversarial problem. \\
+\indent
 They build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf.
@@
 %
+\indent
 Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
@@
 However, we note that the dependency in $M$ is surprisingly better for their result than for our regret upper-bound for \MCTopM-\klUCB, as its regret scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}.\\
+\indent
 Moreover, at first sight, this idea of ``meta'' arms implies an exponential blow up in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
@@
 Despite being more complicated than other approaches, their algorithm ``\texttt{C \& P}'' should not be too hard to implement ourself, and studying its empirical performances is an interesting future work.
+
+% ----------------------------------------------------------------------------
+\subsection{A note on SMPyBandits}
+
+We have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
+We have not yet implemented the other extensions of our model in SMPyBandits, but we are interested to do it, and it is one of the major future work left on our library.
+For more details, see the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}: \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
 

### 2019-04-18 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
HARD

@@
 
-Regardless of whether sensing is or not possible, we present formally in Section~\ref{sub:5:Selfish} the \Selfish{} policy, as used by all the IoT devices in the first model presented in Chapter~\ref{chapter:4}.
+Regardless of whether sensing is or not possible, we start by presenting formally in Section~\ref{sub:5:Selfish} the \Selfish{} heuristic, as it was used by all the IoT devices in the first model presented in Chapter~\ref{chapter:4}.
 It does not use an orthogonalization strategy as the collisions are directly accounted for in the \UCB-like indices that are used by each device to select its channel.
@@
 % -----------------------------------------------------------------
-\subsection{The \Selfish{} heuristic for the ``no sensing case''}
+\subsection{The \Selfish{} heuristic, with or without ``sensing''}
 \label{sub:5:Selfish}
@@
 % -----------------------------------------------------------------
-\section{Numerical simulations}
+\section{Numerical simulations in presence of sensing}
 \label{sec:5:experiments}
@@
   % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=0.90\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
+      \includegraphics[width=0.75\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
   % \end{subfigure}
@@
   % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=0.90\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
+      \includegraphics[width=0.75\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
   % \end{subfigure}
@@
   % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=0.90\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
+      \includegraphics[width=0.75\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
   % \end{subfigure}

### 2019-04-18 2-Chapters/5-Chapter/chapter5.tex
HARD

@@
 But instead of selecting a new rank uniformly at random among $\{1,\dots,M\}$, they proposed to use a ``second-stage'' learning policy\footnote{This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.}, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
-This algorithm is implemented in SMPyBandits and named \texttt{rhoLearn}\footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.}, where it can use Bayes-UCB or any of the $65$ or more bandit algorithms available in the library.
+This algorithm is implemented in SMPyBandits and named \rhoLearn\footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.}, where it can use Bayes-UCB or any of the $65$ or more bandit algorithms available in the library.
 Even if no theoretical guarantee backs up this idea of second-stage learning with Bayes-UCB, empirical simulations found that it can be efficient.
 We illustrate this in Table~\ref{table:5:comparisonRhoRandRhoLearn} below, where we consider the same problem as described above in Section~\ref{sub:5:withoutSensing} (Table~\ref{table:5:meanRegretSimulationsNoSensing}).
-The first \texttt{rhoLearn} uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
+The first \rhoLearn uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
 and we include both the centralized multi-play version of \klUCB{} and our proposal \MCTopM-\klUCB{} for comparison.
@@
         \hline
-        \texttt{rhoLearn}-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1416 \pm 347$ \\
-        \texttt{rhoLearn}-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 \pm 331$ \\
-        \texttt{rhoLearn}-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5134 \pm 976$ \\
+        \rhoLearn-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1416 \pm 347$ \\
+        \rhoLearn-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 \pm 331$ \\
+        \rhoLearn-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5134 \pm 976$ \\
         \hline
@@
     \end{tabular}
-    \caption{Comparison of \RhoRand{} and \texttt{rhoLearn} for a simple multi-player bandit problem with $K=9$ arms.}
+    \caption{Comparing \RhoRand{} and \rhoLearn on a simple MP bandit problem with $K=9$ arms.}
     \label{table:5:comparisonRhoRandRhoLearn}
@@
 
-Since the beginning of this thesis, all the studied bandit problems were stationary: the rewards coming from choosing arm $k$ are all \iid{} and follow the same distribution $\nu_k$.
-The next Chapter~\ref{chapter:6} is focussed on a relaxation of this hypothesis, the piece-wise stationary bandit model.
+Since the beginning of this thesis, all the studied bandit problems were stationary: the rewards coming from choosing arm $k$ are \iid{} and follow the same distribution $\nu_k$.
+The next Chapter~\ref{chapter:6} is focussed on the piece-wise stationary bandit model, a relaxation of this hypothesis.
+
 When we were working on multi-players bandits in Autumn $2017$ \cite{Besson2018ALT}, we left the study of the piece-wise stationary case as a future work, and shortly after we were excited to see that two independent works tackling this question were published, in December $2018$ and in February $2019$.
@@
 \indent
-If $\Upsilon_T$ is bounded by $\Upsilon_T = \bigO{T^{\gamma}}$, for a known $\gamma$ but an unknown $\Upsilon_T$, they prove for instance in Theorem~2 that the expected cumulative regret of their RR-SW-UCB\# algorithm is bounded by $R_T = \bigO{T^{\frac{1+\gamma}{2}} \log(T)}$ (what we call the centralized system regret).
+The proposed the RR-SW-UCB\# algorithm, which combines their SW-UCB\# algorithm, previously proposed in \cite{WeiSrivastava18Abruptly}, and a Round-Robin hoping, under the assumption that each user $j\in\{1,\dots,M\}$ knows its ID $j$ (see above the discussion we give on this unrealistic assumption).
+The SW-UCB\# algorithm is discussed more in details in the literature review of Chapter~\ref{chapter:6} below.
+%
+If $\Upsilon_T$ is bounded by $\Upsilon_T = \bigO{T^{\gamma}}$, for a known $\gamma$ but an unknown $\Upsilon_T$, they prove for instance in Theorem~2 \cite{WeiSrivastava18Distributed} that the expected cumulative regret of their RR-SW-UCB\# algorithm is bounded by $R_T = \bigO{T^{\frac{1+\gamma}{2}} \log(T)}$ (what we call the centralized system regret).
 This bound is of the same order as the bound obtained by the same authors for the SW-UCB\# algorithm in \cite{WeiSrivastava18Abruptly} for the single-player case.
@@
 
-On the other hand, another possible is to extend ideas developed for the adversarial setting.
+On the other hand, another possibility is to extend ideas developed for the adversarial setting.
 In \cite{AlaturLevyKrause19}, the authors essentially consider the piece-wise stationary as an easier case of an adversarial problem. \\
@@
 They build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf.
-Algorithm 1), and on they apparently rediscovered the ``communication trick'' independently from \cite{BoursierPerchet18} to use communications between players to set up a ``coordinator'' player and $M-1$ ``followers'' (running respectively their Algorithms 2 and 3).
+Algorithm 1), and they apparently discovered the ``communication trick'' independently from \cite{BoursierPerchet18} to use (virtual) communications between players to set up a ``coordinator'' player and $M-1$ ``followers'' (running respectively their Algorithms 2 and 3).
 %
@@
 
+\TODOL{Explain that we think the following approach can work very efficiently:
+
+    - Combine \MCTopM{} with \GLRklUCB{} instead of \klUCB,
+    - and incorporate also the detected change-points by B-\GLR{} test in the orthogonalization procedure: after any change-point, the player is not sitting anymore.
+    - Modification: only a change-point on the arm on which player is sitting can make him move?
+It's already the case: the \GLR{} test can only detect change-point on an arm that was played, and the \MCTopM{} orthogonalization procedure forces to play the same arm as long as the player is sitting.
+}
+
 % ----------------------------------------------------------------------------
@@
 We have not yet implemented the other extensions of our model in SMPyBandits, but we are interested to do it, and it is one of the major future work left on our library.
-For more details, see the issue ticket open at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}: \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/X}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
 
@@
 and we refer to the following page for details on how to reproduce them,
-\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}
+\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}}
 

### 2019-04-18 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
I don't know how to talk about Sumit et al work in limited space

@@
 % And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
-and as we observed that using \klUCB{} rather than \UCB{} indices always yield better practical performance,
+and as we observed in Table~\ref{table:5:comparisonUCB_klUCB} that using \klUCB{} rather than \UCB{} indices always yield better practical performance,
 from now on we only report results for \klUCB.
@@
 
-We present results for two bandit instance: one with $K=3$ arms and means
+We present results for two bandit instances: one with $K=3$ arms and means
 $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}.
@@
 % were considered in the analysis of both
-In the presence of sensing (observation model \modeldeux), we also compared our algorithms to  %multi-player multi-armed bandits algorithms
-with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}. Yet these two algorithms were found hard to use efficiently in practice and we show in
+We also compared our algorithms to  %multi-player multi-armed bandits algorithms
+with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}, in the presence of sensing, \ie, observation model \modeldeux, for which they were developed.
+Yet these two algorithms were found hard to use efficiently in practice and we show in
 % that were also proposed

### 2019-04-18 2-Chapters/5-Chapter/chapter5.tex
I don't know how to talk about Sumit et al work in limited space

@@
 Similarly, Lugosi and Mehrabian studied the same problem \cite{LugosiMehrabian18}.
-In both works, the authors detail algorithms that are unfortunately suboptimal empirically, but give in both cases a logarithmic regret if $M \leq K$ players all independently implement the proposed algorithm.
+In both works, the authors detail algorithms that give in both cases a logarithmic regret if $M \leq K$ players all independently implement the proposed algorithm.
 
@@
 Note that it is significantly harder to estimate $M$ without collision information, and Algorithm~3 in page 24 is quite complex.
-We did not implement it, and it is an interesting future work to run some numerical simulations in order to validate it empirically.
+We did not implement it, and it would be interesting to run some numerical simulations in order to validate it empirically.
 
@@
 We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
-The centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, but it is interesting to also add it here, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, because it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
+The centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, but it is interesting to also add it here, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, as it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
 %
@@
 For Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis.
-We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear (\eg, $12150$ for $T=10000$).
+We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, however we also experimented with larger values of $T$ and found a same linear behavior.}.
 
@@
 %
-Further empirical evaluation would be needed to fully understand the situation, and a first future work would be to fix the algorithm proposed by \cite{LugosiMehrabian18} and illustrate in some problems that it can indeed achieve sub-linear regret (maybe it does but only for large horizons, even if we did try larger values of $T$ like up-to $T=200000$).
+Further empirical evaluation would be needed to fully understand the situation, and a first future work would be to either fix our implementation of the algorithm proposed by \cite{LugosiMehrabian18} or propose an efficient modification, and illustrate in some problems that it can indeed achieve sub-linear regret (maybe it does but only for large horizons, even if we did try larger values of $T$ like up-to $T=200000$).
 
@@
 Indeed, imagine that at each time step $t$, all the $M$ players could share an unbounded number of bits with the other players, at no cost, even if this hypothesis is clearly unrealistic for wireless networks.
-Then they can share all their observations, and they can all run the same multiple-plays MAB algorithm \cite{Anantharam87a} (in a deterministic way), like for instance the extension of Thompson sampling for multiple-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
+Then they can share all their observations, and they can all run the same multiple-plays MAB algorithm \cite{Anantharam87a}, like for instance the extension of Thompson sampling for multiple-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
 In this setting, a logarithmic regret is obtained easily, and the regret upper-bound of the two aforementioned algorithms asymptotically achieve the lower-bound from \cite{Anantharam87a}.
@@
 Every time step, another player is either arriving, without any knowledge of the problem (\eg, it is a fresh and new IoT device).
-Then no matter the strategy of these news players, even if they all play \MCTopM-\klUCB{} for instance, if they player a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
+Then no matter the strategy of these news players, even if they all play \MCTopM-\klUCB{} for instance, if they play a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
 

### 2019-04-18 2-Chapters/5-Chapter/chapter5.tex
Go

@@
 The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
-If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the rewards (\ie, sensing) as well as in the players' decisions (\ie, collisions):
+If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the sensing information, as well as in the players' decisions (\ie, collisions):
 \begin{equation}\label{eq:5:defRegretDynamicSetting}
@@
 
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+In the multi-player model presented in this chapter, we assume that the arm distributions are the same for all the players.
+For cognitive radio applications, where arms model channels and players are radio devices, it can be unrealistic to consider that two players, maybe located at different distances from the gateway or equipped with different hardwares, encounter the same mean quality when accessing the same channel.
+%
+Motivated by this weakness, some researchers studied an interesting extension of our model of interest, considering multi-player MAB models with different arms distributions among players.
+%
+Starting in 2012 by \cite{Kalathil12} where arms are Markov chain, this model was studied more actively recently, \cite{DarakHanawal18,Bistritz18,KaufmannAbbas19,Tibrewal2019}.
 
-\cite{Bistritz18}
+In such models, instead of considering $K$ arms characterized by a vector of distributions, $(\nu_k)_{1\leq k \leq K}$, if there is $M$ players we consider a \emph{matrix of distributions}, $(\nu_k^j)_{1 \leq k \leq K, 1 \leq j \leq M}$.
+Two users $j$ and $j'$ can experience different utilities for the same arm $k$, \ie, $\nu_k^{j} \neq \nu_k^{j'}$.
+The goal stays the same, each player wants to maximize its cumulated reward, with or without explicit communications between players, with or without sensing information, but always without central supervision.
+%
+Like before, maximizing the rewards of each player simultaneously is maybe not possible.
+As soon as the matrix is not invariant under permutation of the users, the problem nature changes fundamentally:
+instead of finding an optimal orthogonal assignment of the $M$ players to the \Mbest{} arms,
+the goal of the system is now to reach an equilibrium position, also referred to as a stable marriage.
+Assignment are also called matching, and total (mean) reward of an assignment is its utility.
+Such equilibrium position means that the utility obtained by the $M$ player cannot be increased by swapping two users.
+Indeed imagine just $M=3$ players and $K=3$ arms, and Bernoulli distributions of means $[0.1, 0.5, 0.5]$, $[0.1, 0.5, 0.5]$ and $[0.9, 0.5, 0.1]$ for player $1$, $2$ and $3$.
+Then two optimal affectations of players to arms are $[3,2,1]$ or $[2,3,1]$, both giving the same utility of $1.9$.
 
-\cite{KaufmannAbbas19}
+In this extension, performance is still evaluated by a system regret, now defined as the difference between the sum of the cumulated rewards by the $M$ players, and the utility of any matching.
+The question is to possible to obtain a logarithmic --or even sub-linear-- regret in this problem is more difficult that for the model presented in Section~\ref{sec:5:model}.
+This question was first answered in \cite{Bistritz18}, and proposed an algorithm based on alternating three phases, and increasing their lengths after the end of each epoch ($2^p$ for the $p$-th epoch).
+First, players explore in order to estimate the expectations of the arm rewards ; then players use their ``Game of Thrones'' dynamics (inspired by Musical Chair \cite{Rosenski16}) and play the optimal solution most of the time ; finally players play the action they played most of the time in the recent GoT phases.
+% combining Musical Chair with an epoch-based alternative phases of forced exploration and pure exploitation.
+They analyse their ``Game of Thrones'' algorithm and proved that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
 
-Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
-Sumit J Darak, Manjesh K. Hanawal
-https://arxiv.org/pdf/1812.11651
-\cite{DarakHanawal18}
+Until the very recent work of \cite{KaufmannAbbas19}, it was unknown if a logarithmic regret was possible.
+They proposed an algorithm that is based on two phases: first, players will learn $M$ and learn orthogonal ranks, using the ``communication trick'' by \cite{BoursierPerchet18}, and then one player is elected as a leader and the others are followers.
+This second step, after initialization, is also using epoch of increasing lengths $2^p$.
+Until the optimal solution, each epoch is alternating three phases.
+Followers start by a Round-Robin uniform exploration with no collision, then they use the ``communication trick'' to communicate their samples to the leader, and finally they start an exploitation phase until the end of the game if the leader player tells them so.
+The leader can thus collect enough samples from all arms and all players, and is able to solve iteratively the stable marriage problem, and sends back the successive estimate of the solution to the players.
+%
+In the easier case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO(M K \log(K T) + K M^3 \log(M K T) / \Delta) + K M^2 (\log(M \log(K T) / \Delta))^2$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
+In the generic case, their algorithm M-ETC with Elimination achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$.
 
-Distributed Learning and Optimal Assignment in Multiplayer Heterogeneous Networks
-H Tibrewal, S Patchala, MK Hanawal
-https://arxiv.org/pdf/1901.03868.pdf
-\cite{Tibrewal2019}
-``The channel characteristics are unknown  and  could  be  different  for  each  user  (heterogeneous)''
+% \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+Finally, we note that previous works all focussed on the ``sensing'' case, but most likely sub-linear regret can be achieve by decentralized algorithms that leverage the same techniques as introduced by \cite{BoursierPerchet18,LugosiMehrabian18} for the ``no sensing case''.
+%
+We conclude by noting that this model was recently studied by two very recent other works, \cite{DarakHanawal18,Tibrewal2019}, who obtained results comparable to the results from the two works presented above.
+They both also study the case of dynamic settings, with arrival or departures of players, as presented in Section~\ref{sub:5:arrivalDepartures}.
+Both articles \cite{DarakHanawal18,Tibrewal2019} use the terminology of ad-hoc networks, and they compare empirically their proposal with some previous works, while \cite{KaufmannAbbas19} do not include numerical simulations, and while \cite{Bistritz18} illustrate the performance of their GoT algorithm on a simple example, they do not compare with other algorithms.
+%
+It would be interesting to compare empirically all the different approaches.
+Another interesting directions are to study the possible extension of this model with different arm utilities by players to the non-stationary case, or real-world validation of such decentralized algorithms in real IoT networks.
+
+% Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
+% Sumit J Darak, Manjesh K. Hanawal
+% https://arxiv.org/pdf/1812.11651
+% \cite{DarakHanawal18}
+
+% Distributed Learning and Optimal Assignment in Multiplayer Heterogeneous Networks
+% H Tibrewal, S Patchala, MK Hanawal
+% https://arxiv.org/pdf/1901.03868.pdf
+% \cite{Tibrewal2019}
+% ``The channel characteristics are unknown  and  could  be  different  for  each  user  (heterogeneous)''
 
@@
 
-\TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+% \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 
@@
 
+\TODOL{Je peux faire des expériences pour cette approche, est-ce que j'en fais ? Ici, ou à la fin du chapitre 6 ? Ou en conclusion ?}
+
 % ----------------------------------------------------------------------------

### 2019-04-19 2-Chapters/3-Chapter/chapter3.tex
Done for now, I'm almost boarding my plane WHOO, back to France in a few hours.
- Love you kitty

@@
 
+\paragraph{Abstract.}
+%
 SMPyBandits is a package I have developed since the beginning of my PhD.
@@
 
-SMPyBandits does not aim at being blazing fast or optimal in terms of memory usage, and it comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
-Even if some critical parts are also available as a \texttt{C} Python extension, and even by using Numba \cite{numba} whenever it is possible, if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia.
+SMPyBandits does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
+Some critical parts are also available as a \texttt{C} Python extension, and the just-in-time Numba compiler \cite{numba} is used whenever it is possible, so we can note that we optimized the time efficiency of what could be (easily) optimized.
+However if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia.
+Note that both are not maintained anymore, and contain just a few algorithms for the simple stationary MAB model.
 
@@
 then we present in Section~\ref{sec:3:timeAndMemoryCosts} additional numerical simulations to discuss about the time efficient and memory costs of the some MAB algorithms (we focus on the ones being used in the rest of this thesis, \UCB, \klUCB, and Thompson Sampling).
-To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms, but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
+To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
+% but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
 %
-Finally, in Appendix~\ref{sec:3:appendix} we include some minimalist files showing how to use the library, as well as details regarding the possible use of multiple cores to speed-up simulations.
+Finally, in Appendix~\ref{sec:3:appendix} we include some minimalist files showing how to use the library, as well as details regarding the use of parallel computing to speed-up simulations.
 
@@
 %
+For instance the following code creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate the problem.
+We give another example for truncated Gaussian distributions, and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
+
+% https://tex.stackexchange.com/a/12430/
+\begin{small}
+\begin{listing}[h!]
+    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python}
+from SMPyBandits.Arms import *
+arm1 = Bernoulli(0.1)
+arm2 = Bernoulli(0.5)
+arm3 = Bernoulli(0.9)
+
+means = [0.1, 0.5, 0.9]
+from SMPyBandits.Environment import MAB
+bernoulli_problem = MAB([arm1, arm2, arm3])
+# but also
+bernoulli_problem = MAB({'arm_type': Bernoulli, 'params': means})
+
+gaussian_problem = MAB({'arm_type': Gaussian, 'params': means})
+gaussian_problem.plotHistogram()  # display the histogram showed below
+    \end{minted}
+    \caption{Small snippet of Bash code to run a simple experiment with SMPyBandits}
+    % \captionof{listing}{Small snippet of Bash code to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
+    \label{lst:3:howToRunBasicLibrary}
+\end{listing}
+\end{small}
+
+\begin{figure}[h!]  % [htbp]
+	\centering
+	\includegraphics[width=0.95\linewidth]{exampleOfRewards.pdf}
+	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$, $0.5$ and $0.9$.}
+	\label{fig:3:exampleOfRewards}
+\end{figure}
+
 We do not give more details, but the interested reader can refer to the following Jupyter notebook \cite{jupyter} document hosted at:\\
@@
 
-With this numerical framework, simulations can run on a single CPU or a multi-core machine using joblib \cite{joblib},
+With this numerical framework, simulations can run on a single CPU or a multi-core machine using joblib \cite{joblib} (see Appendix~\ref{sub:3:parallelSimulations}),
 and summary plots are automatically saved as high-quality PNG, PDF and EPS, using matplotlib \cite{matplotlib} and seaborn \cite{seaborn}.
@@
         \end{minted}
-        \caption{Small snippet of Bash code to download and install dependencies of SMPyBandits}
+        \caption{Small snippet of Bash code to download and install dependencies of SMPyBandits.}
         % \captionof{listing}{Small snippet of Bash code to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
@@
     \end{minted}
-    \caption{Small snippet of Bash code to run a simple experiment with SMPyBandits}
+    \caption{Small snippet of Bash code to run a simple experiment with SMPyBandits.}
     % \captionof{listing}{Small snippet of Bash code to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
@@
   {"archtype": klUCB, "params": {"klucb": klucbBern}},
-  {"archtype": klUCBPlusPlus, "params": {"horizon": HORIZON, "klucb": klucbBern}},
+  {"archtype": klUCBPlusPlus, "params": {"horizon": T, "klucb": klucbBern}},
   {"archtype": UCBalpha, "params": {"alpha": 1}},
@@
 
-In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
+-- In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
 We designed a variant of the Exp4 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
@@
 
-In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+-- In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
 
-In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
+-- In \cite{Besson2018DoublingTricks} (quickly presented in Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
 and lots of bibliographic referencesIn sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
@@
 
-In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
+-- In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
 We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
 We designed the \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.GLR_UCB.html}{GLRklUCB}} policy, and implemented most of the policies from state-of-the-art research on passively or active adaptive policies, both adversarial- or stochastic-based, designed to tackle the piece-wise stationary MAB problems.
-We also implemented a large benchmark of problems.
+We also implemented a large benchmark of different problems.
 
@@
 \label{sec:3:reviewSPAlgorithms}
+
+\TODOL{Write text for this Section, write configuration.py file, run short-term experiment, include them, if happy about it, run longer ang larger experiments}
 
@@
 \label{sec:3:timeAndMemoryCosts}
+
+\TODOL{Write text for this Section, write configuration.py file, run short-term experiment, include them, if happy about it, run longer ang larger experiments}
 
@@
 % ----------------------------------------------------------------------------
-\section{Conclusion of Chapter 3}
+\section{Conclusion}
 \label{sec:3:conclusion}
@@
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter 3}
+\section{Appendix}
 \label{sec:3:appendix}
@@
 We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $20$ jobs.
-We give in Table~\ref{table:3:speedUpTimeParallelComputations}
+We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=1000$, and different number of repetitions and number of jobs.
+It clearly illustrates that using more jobs than the number of CPU is suboptimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
+Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
 

### 2019-04-19 2-Chapters/2-Chapter/chapter2.tex
Cette thèse me semble interminable...
One new ref, and lots of work on chapter 3

@@
 % First chapter begins here
-\chapter{Single Player and Stationary Multi-Armed Bandit Models}
+\chapter{The Single Player and Stationary Multi-Armed Bandit Model}
 \label{chapter:2}
@@
 
-Abstract for Chapter 2.
+In this chapter, we present the first model of multi-armed bandit (MAB) studied in this thesis,
+the single-player and stationary stochastic MAB model.
+MAB models a decision-making game, where a \emph{player} has to sequentially select one action in a (usually finite) set of actions, and only receives a (random) feedback about the action she selected, also called \emph{reward}.
+Her goal is to maximize its sum of received rewards, from time $t=1$ to $t=T$ if the game is played at an horizon $T$.
+The difficulty comes from balancing the trade-off between \emph{exploration}, because she has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
+%
+We focus on the model with finite arms, and binary or real-valued stochastic rewards.
+After introducing formally the notations used in all this thesis, for arms, distributions, rewards etc, we review shortly various possible applications of multi-armed bandits.
+The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication of a device to a gateway (\eg, spreading factor for LoRa, power allocation for NOMA etc).
+
+We explain the measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and horizon $T$.
+The regret compares the performance of the decision-making player and the oracle who knows in advance the best arm and always selects it.
+Other measures of performance, such as best-arm identification, are not used later on in this document, but they are also shortly discussed.
+%
+We give an overview of the main families of algorithms designed to tackle the stochastic MAB model, and we mainly focus on the \UCB{} and \klUCB{} algorithms, as they are the two we used in the next chapters.
+Part of the work presented in Chapter~\ref{chapter:4} also used the Thompson sampling algorithm, but no theoretical analyses were done about it.
+When presenting our library SMPyBandits in Chapter~\ref{chapter:3} our library SMPyBandits, designed to perform numerical simulations of MAB problem, we use it to compare the empirical performances of the most famous and most efficient algorithms, on a few different problems.
+%
+In the point-of-view of practical simulations, one must also take care of two other measures of performance: time and memory complexity, which can vary greatly between different families of algorithms.
+We also give an empirical comparison of different algorithms for these two measures in Chapter~\ref{chapter:3}.
+
+We then review the main theoretical results from famous previous works, that give a lower-bound on the regret of any algorithm, essentially showing that we cannot find an algorithm achieving a sub-logarithmic regret, in this stochastic MAB model.
+We also gives regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
+They both attain a logarithmic regret, proving their order optimality (\ie, their regret upper-bound asymptotically matches the lower-bound), moreover \klUCB{} is known to be optimal for a wide range of arms distributions.
+A sketch of the proof of both results is given in Appendix.
+
+Finally, we explore the problem of algorithm selection: as there is many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
+We present one of the first contribution made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms.
+We illustrate its performance empirically, but no satisfactory theoretical result is given, as the bound we hoped for have been proved unachievable since then.
+
+\paragraph{Publications.}
+%
+Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on the following publication: \cite{Besson2018WCNC}.
+The notations used in this thesis are the usual notations in all the MAB literature, and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
+
+\paragraph{Main references for the curious reader.}
+%
+For more details and a more pedagogical introduction to multi-armed bandits, we suggest to the interested reader to read the short text-book \cite{Slivkins2019}.
+Another excellent but reasonably short survey is \cite{Bubeck12}, while the more recent book \cite{LattimoreBanditAlgorithmsBook} is the most complete resource about bandit algorithms.
 
@@
 No linear bandits, no contextual bandits etc.
+
+\TODOL{Write this!}
 
@@
 % ----------------------------------------------------------------------------
-\section{Conclusion of Chapter 2}
+\section{Conclusion}
 \label{sec:2:conclusion}
@@
 
-Future works include...
+\TODOL{Write this!}
 
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter 2}
+\section{Appendix}
 \label{sec:2:appendix}
 
-\subsection{Eficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
+\TODOL{Write this!}
+
+\subsection{Efficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
 

### 2019-04-19 2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py

@@ -1,8 +1,7 @@
 """ An example of a configuration file to launch some the simulations
 For the single-player case.
 """
-CPU_COUNT = 4
-from os import getenv
+from os import getenv  # to get command line variables
 
 # Import arms
 from SMPyBandits.Arms import *
@@ -23,7 +22,7 @@ REPETITIONS = int(getenv('N', 100))
 # -1 means all the CPU cores, 1 means no parallelization.
 N_JOBS = int(getenv('N_JOBS', -1))
 
-# Number of arms for non-hard-coded problems (Bayesian problems)
+# Number of arms for non-hard-coded problems
 NB_ARMS = int(getenv('K', 3))
 
 # Lower value of means
@@ -31,21 +30,13 @@ LOWER = float(getenv('LOWER', 0))
 # Amplitude value of means
 AMPLITUDE = float(getenv('AMPLITUDE', 1))
 
-# Type of arms for non-hard-coded problems (Bayesian problems)
+# Type of arms for non-hard-coded problems
 ARM_TYPE = mapping_ARM_TYPE[str(getenv('ARM_TYPE', "Bernoulli"))]
 
 # Means of arms for non-hard-coded problems
 MEANS = uniformMeans(nbArms=NB_ARMS, delta=0.05,
     lower=LOWER, amplitude=AMPLITUDE, isSorted=True
 )
-
-import numpy as np
-# more parametric? Read from cli?
-MEANS_STR = getenv('MEANS', '')
-if MEANS_STR != '':
-    MEANS_STR = MEANS_STR.replace('[','').replace(']','').split(',')
-    MEANS = [ float(m) for m in MEANS_STR ]
-    print("Using cli env variable to use MEANS = {}.".format(MEANS))  # DEBUG
 
 # This dictionary configures the experiments
 configuration = {
@@ -71,16 +62,14 @@ configuration = {
             "archtype": Uniform,   # The stupidest policy, fully uniform
             "params": {}
         },
-        # # --- Fixed arm algorithm
+        # --- Fixed arm algorithm
         {
             "archtype": TakeFixedArm,
             "params": { "armIndex": 0 },  # Take worse arm!
-        },
-        {
+        }, {
             "archtype": TakeFixedArm,
             "params": { "armIndex": 1 },  # Take second worse arm!
-        },
-        {
+        }, {
             "archtype": TakeFixedArm,
             "params": { "armIndex": min(2, nbArms - 1) },  # Take third worse arm!
         },

### 2019-04-19 2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py

@@ -1,5 +1,5 @@
 """ Example of use of SMPyBandits.
-See https://smpybandits.github.io/API.html for more"""
+See https://SMPyBandits.GitHub.io/API.html for more details!"""
 import numpy as np
 np.random.seed(0)  # for reproducibility
 
@@ -8,19 +8,19 @@ arms = [Bernoulli(0.1), Bernoulli(0.9)]
 
 from SMPyBandits.Environment import MAB
 my_MAB_problem = MAB(arms)
-nbArms = my_MAB_problem.nbArms
+nbArms = my_MAB_problem.nbArms  # 2 arms !
 
 from SMPyBandits.Policies import UCB
 my_UCB_algo = UCB(nbArms)
 my_UCB_algo.startGame()  # reset internal memory
 
 horizon = 1000
-for t in range(horizon):
+for t in range(horizon):  # simulation loop
     chosen_arm = my_UCB_algo.choice()
     observed_reward = my_MAB_problem.draw(chosen_arm)
     my_UCB_algo.getReward(chosen_arm, observed_reward)
 
-cumulated_reward = sum(my_UCB_algo.rewards)
-number_of_plays = sum(my_UCB_algo.pulls)  # = 1000
+cumulated_reward = sum(my_UCB_algo.rewards)  # random!
+number_of_plays  = sum(my_UCB_algo.pulls)    # horizon = 1000
 mean_reward = cumulated_reward / number_of_plays
 print("The UCB algorithm obtains here a mean reward =", mean_reward)

### 2019-04-19 2-Chapters/3-Chapter/chapter3.tex
One new ref, and lots of work on chapter 3
Still working on Chapter 2 and 3 simulatneously.

@@
 
-\paragraph{Publications}
+\paragraph{Publications.}
 %
@@
 
+\subsection{Continuous integration using Read the Docs and Travis CI}
+
+% FIXME
+\TODOL{I need to quickly explain that SMPyBandits is using Read the Docs since $2017$ and Travis Ci since $2018$ to:
+
+1.
+automatically build the documentation on Read the Docs after every commit.
+This allows to regularly check that the library is well formatted and can be imported correctly, as well as keeping the online documentation up-to-date.
+
+2.
+automatically run short numerical simulations on Travis CI after every commit.
+This allows to check that no modification on any part of the codebase breaks anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
+The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proved very useful to quickly find and fix new bugs.
+}
+
 \subsection{Research works using SMPyBandits}
@@
 
+\newpage  % WARNING
 % ----------------------------------------------------------------------------
@@
 
-In this chapter, we saw...
+In this chapter, we presented our Python library, named SMPyBandits, by detailing three main points:
+$(i)$ its purpose and qualities,
+$(ii)$ the structure of the package,
+and $(iii)$ an overview of its usage:
+%
 
-Future works include...
+$(i)$
+Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic (see Chapter~\ref{chapter:6}) problems of single- or multi-player multi-armed bandits.
+SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
+Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.
 
+$(ii)$
+We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
+As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
+Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
+One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
+More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.
+All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
+%
+When comparing algorithms on a problem, the main performance measure is the regret, but the library also compute, store and visualize other measures such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
+
+$(iii)$
+Finally, we presented how to use SMPyBandits, if one wants to run some pre-designed simulations or design new simulations.
+Running a simulation is very easy, and different examples showed that the main parameters such as the time horizon $T$ or the number of repetitions can be configured directly from the command line when running the Python script, or by modifying the code.
+Designing a new simulation requires to have a basic knowledge of Python, but not to dive into the implementation of the library.
+We give below in Appendix~\ref{sec:3:appendix} a small but complete example of the two files that has to be written for a new simulation: a \texttt{main} file which essentially defines the desired visualizations after the simulation, and a \texttt{configuration} file which defines the problems to simulate and their parameters, as well as the algorithms that will be compared.
+
+% WARNING
+\paragraph{Remark.}
+We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of SMPyBandits, and as such the library, its documentation and this chapter  are considered one of the main contribution of this thesis.
+%
+Our library is used for the numerical simulations in all this thesis, except Chapter~\ref{chapter:4}.
+We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
+We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} that we shortly present in \ref{app:2:DoublingTricks}.
+
+\paragraph{Future works.}
+%
+An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models.
+For instance, adding support for non-finite arms is interesting, \eg, linear or contextual bandit.
+
+In Chapter~\ref{chapter:5}, we use SMPyBandits to simulate multi-player bandit problems, and some we present some extensions of this model.
+Some have already been implemented, but other extensions are yet to be added.
+It is one of the major future work left on our library.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/X}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
+
+Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity.
+We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted.
+%
+I would personally love to be able in the next to teach a graduate or under-graduate course on multi-armed bandits, and use my library as a support to illustrate the course, as well as practical sessions.
+
+\newpage  % WARNING
 % ----------------------------------------------------------------------------
@@
 
-\subsection{A final note about parallelism in SMPyBandits}
+\subsection{Using parallelism in SMPyBandits}
 \label{sub:3:parallelSimulations}
@@
 
-As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal, dividing the running time by \texttt{N\_JOBS}.
-But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is highly sub-optimal, as the CPU will essentially spend all his time (and memory) managing the different jobs, and not actually doing the simulations.
+As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
+%  dividing the running time by \texttt{N\_JOBS}.
+But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is sub-optimal, as the CPU will essentially spend all his time (and memory) managing the different jobs, and not actually doing the simulations.
 Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits.
@@
 We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=1000$, and different number of repetitions and number of jobs.
-It clearly illustrates that using more jobs than the number of CPU is suboptimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
+It clearly illustrates that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
 Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
@@
     \begin{tabular}{c|ccc}
-    \textbf{Repetitions} $\;$ \textbackslash $\;$ \texttt{N\_JOBS} & $1$ & $4$ & $20$ \\
+    \textbf{Repetitions} $\;$ \textbackslash $\;$ number of jobs \texttt{N\_JOBS} & $1$ & $4$ ($=$ nb CPU cores) & $20$ ($>$ nb CPU cores) \\
         \hline
@@
         $100$ repetitions & $749 \;\text{s}$ & $272 \;\text{s}$ & $308 \;\text{s}$ \\
-        $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1346 \;\text{s}$ \\
+        $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1846 \;\text{s}$ \\
         \hline
@@
 
-On the first hand, I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
+$(i)$ I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
 Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.
@@
 
-On the other hand, I also did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
+$(ii)$ I also did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
 It is partly due to time constraint, as I would have been curious to try, but mainly because we found that it would not have helped us much to use a large scale cluster.

### 2019-04-19 2-Chapters/3-Chapter/chapter3.tex
Almost done for text of Chapter 3
- , except 3.2 and 3.3 which will be numerical experiments (to write, design, run and include)

@@
 %
-SMPyBandits is a package I have developed since the beginning of my PhD.
-It is designed to allow easy and efficient numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms, and SMPyBandits is written in Python.
-This library is by far the most complete open-source implementation of state-of-the-art algorithms tackling various kinds of MAB sequential learning problems.
-It aims at being extensive, simple to use and maintain, with a clean and well documented codebase.
+SMPyBandits is a Python package I have developed since the beginning of my PhD.
+It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
+%  and SMPyBandits is written in Python.
+This library is by far the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
+It aims at being extensive, simple to use and maintain, and has a clean and well documented codebase.
 It allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments.
@@
 More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
+It uses the most recent technologies, by being hosted on GitHub, and uses two online services of continuous integration to run automated tests and build the documentation.
 
@@
 
-SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms tackling various kinds of sequential learning problems referred to as Multi-Armed Bandits.
+SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms and models of Multi-Armed Bandits.
 It aims at being extensive, simple to use and maintain, with a clean and perfectly documented codebase.
@@
 We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
-then we present in Section~\ref{sec:3:timeAndMemoryCosts} additional numerical simulations to discuss about the time efficient and memory costs of the some MAB algorithms (we focus on the ones being used in the rest of this thesis, \UCB, \klUCB, and Thompson Sampling).
-To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
+then we present in Section~\ref{sec:3:timeAndMemoryCosts} additional numerical simulations to discuss about the time efficient and memory costs of the some MAB algorithms.
+We focus on the ones being used in the rest of this thesis, \UCB, \klUCB, and Thompson Sampling.
+% To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
 % but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
@@
 This chapter is based on the following publication: \cite{SMPyBanditsJMLR}.
-The code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is publicly released under the MIT open-source License.
+The code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is freely publicly released under the MIT open-source License.
 
@@
 We start by explaining how SMPyBandits simulates MAB problems, by detailing the different parts.
-We then detail how SMPyBandits implements MAB algorithms (\eg, \UCB),
+We then detail how it implements MAB algorithms (\eg, \UCB),
 and what kind of information is displayed, saved and plotted after a batch of simulations of bandit problems.
@@
 Figure~\ref{fig:3:firstPlot} above shows the average regret for these $4$ algorithms.
-The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.
-The Figure~\ref{fig:3:firstPlot_hist} below shows the histogram of regret for the same experiment.
+% The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.
+The Figure~\ref{fig:3:firstPlot_hist} below shows the histogram of regret obtained at the end of the experiment (\ie, $R_T$) for the same example.
 
@@
 
-Optional dependencies include \texttt{joblib} \cite{joblib} for parallel simulations, \texttt{numba} \cite{numba} for automatic speed-up on small functions, \texttt{sphinx} \cite{sphinx} for generating the documentation, \texttt{virtualenv} \cite{virtualenv} for launching simulations in isolated environments, and \texttt{jupyter} \cite{jupyter} used with \texttt{ipython} \cite{ipython} to experiment with the code.
+Optional dependencies include \texttt{joblib} \cite{joblib} for parallel simulations, \texttt{numba} \cite{numba} for automatic speed-up on small functions using a Just-in-Time compiler, \texttt{sphinx} \cite{sphinx} for generating the documentation, \texttt{virtualenv} \cite{virtualenv} for launching simulations in isolated environments, and \texttt{jupyter} \cite{jupyter} used with \texttt{ipython} \cite{ipython} to experiment with the code.
 
 All the quoted libraries are free and open-source and can be installed in one command using the \texttt{pip} (\texttt{\href{https://pip.pypa.io/}{pip.PyPa.io}}) or \texttt{conda} (\texttt{\href{http://conda.anaconda.org/}{conda.anaconda.org}}) package manager.
+%
+When SMPyBandits is installed using \texttt{pip install SMPyBandits}, the dependencies are of course automatically installed if not already present.
 
@@
 
-SMPyBandits was used for the following research articles since $2017$, and it is used for the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
+SMPyBandits was used for the following research articles since $2017$, and it is used for the previous and the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
 
@@
 % ----------------------------------------------------------------------------
-\section{Some numerical experiments to compare time and memory costs of various algorithms}
+\section{Comparing the time and memory costs of various algorithms}
 \label{sec:3:timeAndMemoryCosts}
@@
 %
-An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models.
-For instance, adding support for non-finite arms is interesting, \eg, linear or contextual bandit.
+An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
+For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
+or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+%
+In Chapter~\ref{chapter:5}, we use SMPyBandits to simulate multi-player bandit problems, and we also present many extensions of this model.
+Some have already been implemented, but other extensions are yet to be added.
+This direction is one of the major future work left on our library
+(see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
 
-In Chapter~\ref{chapter:5}, we use SMPyBandits to simulate multi-player bandit problems, and some we present some extensions of this model.
-Some have already been implemented, but other extensions are yet to be added.
-It is one of the major future work left on our library.
-For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/X}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
+Since $2017$, I also received by email or by GitHub issues some requests to implement new features or models.
+Most of them were implemented quickly, but some would require more time and are left as possible future works.
+Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).
 
-Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity.
-We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted.
+Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
+Its documentation has seen already about $20000$ visits, the project on GitHub had $74$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
+While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
 %
-I would personally love to be able in the next to teach a graduate or under-graduate course on multi-armed bandits, and use my library as a support to illustrate the course, as well as practical sessions.
+We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.
+
+As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
+I would also love to be able in the next to teach a graduate or under-graduate course on multi-armed bandits, and use my library as a support to illustrate the course, as well as practical sessions.
 

### 2019-04-19 2-Chapters/3-Chapter/chapter3.tex
Ca avance bien YAS YAS YAS

@@
 
-In this chapter, we presented our Python library, named SMPyBandits, by detailing three main points:
+In this chapter, we presented our Python library SMPyBandits, by detailing three main points:
 $(i)$ its purpose and qualities,
-$(ii)$ the structure of the package,
+$(ii)$ its organization,
 and $(iii)$ an overview of its usage:
@@
 % WARNING
-\paragraph{Remark.}
+% \paragraph{Remark.}
 We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of SMPyBandits, and as such the library, its documentation and this chapter  are considered one of the main contribution of this thesis.
@@
 We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
-We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} that we shortly present in \ref{app:2:DoublingTricks}.
+We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} that we shortly present in Appendix~\ref{app:2:DoublingTricks}.
 
@@
 %
-In Chapter~\ref{chapter:5}, we use SMPyBandits to simulate multi-player bandit problems, and we also present many extensions of this model.
-Some have already been implemented, but other extensions are yet to be added.
-This direction is one of the major future work left on our library
-(see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
+At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-player bandit model,
+and even if some have already been implemented, a major future work is to implement the most interesting ones
+(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
 
@@
     \begin{tabular}{c|ccc}
-    \textbf{Repetitions} $\;$ \textbackslash $\;$ number of jobs \texttt{N\_JOBS} & $1$ & $4$ ($=$ nb CPU cores) & $20$ ($>$ nb CPU cores) \\
+    \textbf{Repetitions} $\;$ \textbackslash $\;$ number of jobs \texttt{N\_JOBS} & $1$ & $4$ ($=$ nb cores) & $20$ ($>$ nb cores) \\
         \hline
@@
 \end{table}
+
+\TODOL{Pas sûr que ça vaille la peine d'inclure ce passage, c'est vraiment optionnel. Dans tous les cas, je devrais rajouter tout cette sous section dans la documentation, c'est intéressant !}
 

### 2019-04-19 1-Preamble/4-Acknowledgement/acknowledgement.tex
Done

@@
 Christophe was helpful for his expertise on wireless communications and engineering, and especially on cognitive radio.
-Regardless of what we did together, whether it was debugging a difficult GNU Radio block \cite{Besson2018ICT,Besson2019WCNC}, writing a paper together by night in a plane coming back from India \cite{MoyBesson2019}, chatting around a friendly barbecue or playing ``palets'' together along the ocean in Vannes, I was lucky to collaborate with and be around Christophe.\\
+Regardless of what we did together, whether it was debugging a difficult GNU Radio block \cite{Besson2018ICT,Besson2019WCNC}, writing a paper by night while Christophe was in a plane coming back from India \cite{MoyBesson2019}, chatting at a barbecue or playing ``palets'' together along the ocean in Vannes, I was lucky to collaborate with and be around Christophe.\\
 %
+\indent
 Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years\footnote{Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.}.
-Émilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep the highest possible rigour in all parts of our exchanges.
+Émilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep the highest possible rigour in all parts of our exchanges and productions.
 I was lucky to be able to visit the SequeL team at Inria Lille on regular occasions, and to work in close collaboration with Émilie during these visits\footnote{Merci à Romain et toi pour votre accueil chaleureux, et aussi merci à Florian, Nicolas et Édouard.}.
@@
 and to Richard Combes as well for being part of my jury.
-Some of their articles were strong inspirations for our own works, and I wish to also thank their co-authors, and in particular the Masters or PhD students, and useful discussions we had whenever we met, especially thanks to Claire Vernade, Réda Alami and Étienne Boursier, and others I could not meet (\eg, Viktor Toldov, Robin Allesiardo).
+Some of their articles were strong inspirations for our own works, and I wish to also thank their co-authors, and in particular the Masters or PhD students, and useful discussions we had whenever we met, especially thanks to Claire Vernade, Réda Alami, Étienne Boursier, Frédéric Loge-Munerel and Pratik Gajane, and others I could not meet (\eg, Viktor Toldov or Robin Allesiardo).
 
 Of course on a more personal level, my thoughts are strongly dedicated to my family.
-I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model, who has been of a constant support during all my studies, and whose serious work kept pushing me further.
+%
+I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my studies.
+His whose solid and serious work kept pushing me further, some of our board games nights were brain teasers comparable with some maths problems solved for this thesis,
+and I was happy to help him for the organization of the research colloque\footnote{Si vous l'avez manqué, un livre a été publié suite au colloque, \href{https://www.editions-vendemiaire.com/catalogue/a-paraitre/kaamelott-un-livre-d-histoire-florian-besson-et-justine-breton-dir/}{``Kaamelott, un livre d'histoire'', aux éditions Vandémiaire}.} on the \emph{Kaamelott} french TV show in February 2018, it was the funnier and most exciting research event I have been to in the last four years!
+%
 I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life.

### 2019-04-19 5-Nomenclature/Nomenclature.tex
Finishing Chapter:4
Almost finished the acknowledgements...
One new ref, and lots of work on chapter 3
Done

--- 2019-02-13 Nomenclature/Nomenclature.tex
@@
+% FIXME Est-ce que je dois mettre dans la nomenclature symboles aussi ?
+
+% --------------------------------
+% matches Abbreviations        > A
+
 \nomenclature[A]{wrt}{\textit{with respect to}}
 \nomenclature[A]{wlog}{\textit{with loss of generality}} % TODO not used?
-\nomenclature[A]{iid}{\textit{identically and independently distributed}}
+\nomenclature[A]{i.i.d.}{\textit{identically and independently distributed}}
 
@@
 \nomenclature[A]{LPWAN}{\textit{Low-Power Wide-Area Network}}
+\nomenclature[A]{NOMA}{\textit{Non-Orthogonal Multiple Access}}
+\nomenclature[A]{LAN}{\textit{Local Area Network}}
+\nomenclature[A]{WLAN}{\textit{Wireless Local Area Network}}
+\nomenclature[A]{ISM}{\textit{Industrial, Scientific and Medical} (bands)}
 \nomenclature[A]{CR}{\textit{Cognitive Radio}}
+\nomenclature[A]{Ack}{\textit{Acknowledgement}}
+\nomenclature[A]{SNR}{\textit{Signal to Noise Ratio}}
+\nomenclature[A]{RF}{\textit{Radio Frequency}}
+\nomenclature[A]{DSA}{\textit{Dynamic Spectrum Access}}
+\nomenclature[A]{PHY}{\textit{Physical} (layer)}
+\nomenclature[A]{MAC}{\textit{Medium ACcess} (layer)}
+\nomenclature[A]{ALOHA}{\textit{ALOHA} (not an acronym)}
 \nomenclature[A]{OSA}{\textit{Opportunistic Spectrum Access}}
@@
 \nomenclature[A]{MCTopM}{\textit{Musical-Chair on Top-M} (algorithm)}
-\nomenclature[A]{RandTopM}{\textit{Random Hoping Musical-Chair on Top-M} (algorithm)}
+\nomenclature[A]{RandTopM}{\textit{Random Hoping on Top-M} (algorithm)}
 \nomenclature[A]{GLR}{\textit{Generalized Likelihood Ratio}}
+\nomenclature[A]{MAB}{\textit{Multi-Armed Bandit}}
+\nomenclature[A]{ML}{\textit{Machine Learning}}
+\nomenclature[A]{RL}{\textit{Reinforcement Learning}}
 \nomenclature[A]{GLRT}{\textit{Generalized Likelihood Ratio Test}}
+
+% --------------------------------
+% matches Greek Symbols        > G
+\nomenclature[G]{$\mu_k,\mu_k(t),\mu_k^j$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$.}
+
+% --------------------------------
+% matches Roman Symbols        > L
+\nomenclature[L]{$t$}{Time step, $t\in\{1,\dots,T\}$}
+\nomenclature[L]{$T$}{Time horizon, the duration of the bandit game}
+
+% --------------------------------
+% matches mathematical symbols > O
+
+% --------------------------------
+% matches Subscripts           > I
+
+% --------------------------------
+% matches Superscripts         > E

### 2019-04-23 1-Preamble/4-Acknowledgement/acknowledgement.tex
Je suis tellement inefficace c'est triste

@@
 
-I would like to acknowledge and thank the following people, and also acknowledge the various funding who supported our works during the last three years.
+I would like to acknowledge and thank the following people, and also acknowledge the various funding who supported our works during the last three years and made this thesis possible.
 
@@
 Many thanks to \href{https://sites.google.com/site/vianneyperchet/}{Vianney Perchet} and \href{http://researchers.lille.inria.fr/~mitton/}{Nathalie Mitton} for accepting to review my thesis during the summer 2019, and for their very valuable comments.
-Thanks to Patrick Maillé and Raphaël Féraud for following our work since Spring 2017,
-and to Richard Combes as well for being part of my jury.
-Some of their articles were strong inspirations for our own works, and I wish to also thank their co-authors, and in particular the Masters or PhD students, and useful discussions we had whenever we met, especially thanks to Claire Vernade, Réda Alami, Étienne Boursier, Frédéric Loge-Munerel and Pratik Gajane, and others I could not meet (\eg, Viktor Toldov or Robin Allesiardo).
+Thanks to \href{http://perso.telecom-bretagne.eu/patrickmaille/}{Patrick Maillé} and \href{https://www.researchgate.net/profile/Raphael_Feraud/}{Raphaël Féraud} for following our work since Spring 2017,
+and to \href{http://rcombes.supelec.free.fr/}{Richard Combes} as well for being part of my jury.
+Some of their articles were strong inspirations for our own works, and I wish to also thank their co-authors, and in particular the Masters or PhD students, and useful discussions we had whenever we met, especially thanks to \href{https://www.cvernade.com/}{Claire Vernade}, \href{https://www.linkedin.com/in/reda-alami-067304109}{Réda Alami}, \href{https://www.researchgate.net/profile/Etienne_Boursier}{Étienne Boursier}, \href{https://www.researchgate.net/profile/Frederic_Loge_Munerel}{Frédéric Loge-Munerel} and \href{https://pratikgajane.wordpress.com/}{Pratik Gajane}, and others I could not meet (\eg, Viktor Toldov or \href{https://fr.linkedin.com/in/robin-allesiardo-phd-39221a92}{Robin Allesiardo}).
 
@@
 I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my studies.
-His whose solid and serious work kept pushing me further, some of our board games nights were brain teasers comparable with some maths problems solved for this thesis,
-and I was happy to help him for the organization of the research colloque\footnote{Si vous l'avez manqué, un livre a été publié suite au colloque, \href{https://www.editions-vendemiaire.com/catalogue/a-paraitre/kaamelott-un-livre-d-histoire-florian-besson-et-justine-breton-dir/}{``Kaamelott, un livre d'histoire'', aux éditions Vandémiaire}.} on the \emph{Kaamelott} french TV show in February 2018, it was the funnier and most exciting research event I have been to in the last four years!
+His solid and serious work kept pushing me further, some of our board games nights were brain teasers comparable with some maths problems solved for this thesis,
+and I was happy to help him for the organization of the \href{http://lettres.sorbonne-universite.fr/Kaamelott}{research colloque}\footnote{Un livre a aussi été publié suite \href{http://lettres.sorbonne-universite.fr/Kaamelott}{au colloque}, \href{https://www.editions-vendemiaire.com/catalogue/a-paraitre/kaamelott-un-livre-d-histoire-florian-besson-et-justine-breton-dir/}{``Kaamelott, un livre d'histoire'', aux éditions Vandémiaire}.} on the \emph{Kaamelott} french TV show in February 2017, it was \href{https://www.fabula.org/actualites/kaamelott-ou-la-relecture-de-l-histoire_75490.php}{the funniest} and \href{https://www.vousnousils.fr/2017/03/22/kaamelott-un-tremplin-vers-lhistoire-et-la-litterature-601428}{most exciting} \href{https://lpcm.hypotheses.org/10718}{research event} I have been to in the last four years!
 %
@@
 
-Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas in Lille.
-I want to thank especially the ``coffee club'' members$^*$ in Rennes, for so many interesting discussions who where sometimes the highlights of our day-to-day life at the office.
+Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Edouard in Lille.
+I want to thank especially the ``coffee friends$^*$'' in Rennes, for the daily interesting discussions, which where sometimes the highlights of our life at the office.
+%
+Thanks to everybody who played \emph{futsal} on Wednesday lunch, I still have to progress immensely before I could pretend to be useful, but it's been a lot of fun to play with you guys!
 
@@
 %
-\emph{Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con sus risas y su amor, durante el año pasado.}
-\textcolor{white}{Te quiero, Lola, ¿quieres vivir conmigo?}
+\emph{Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con sus risas y su amor, durante el año pasado.}\\
+\indent
+\textcolor{white}{$\heartsuit$ Te quiero, Lola, ¿quieres vivir conmigo?}
 

### 2019-04-23 2-Chapters/3-Chapter/chapter3.tex
Je suis tellement inefficace c'est triste

@@
 The
-    sample (random) feedback $r_k(t)$ (\iid) from distributions $\nu_{k}$, and compute the rewards $Y^{(j)}(t)$ from the random feedback and the collision model.
-Finally feeds the observation $(A^{(j)}(t),Y^{(j)}(t))$ to the $M$ players,
+    sample (random) feedback $Y_k(t)$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^{(j)}(t)$ from the random feedback and the collision model.
+Finally feeds the observation $(A^{(j)}(t), Y_{A^{(j)}(t)}(t), r^{(j)}(t))$ to player $j$, for all the $M$ players,
     \item At the end, compute the accumulated reward, the regret, plot visualizations etc.
@@
 
-\paragraph{What we did implement: \tettt{Joblib} for multi-core simulations.}
+\paragraph{What we did implement: \texttt{Joblib} for multi-core simulations.}
 %

### 2019-04-23 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Je suis tellement inefficace c'est triste

@@
 Messages can be sent in a fixed number $K\geq2$ of wireless channels,
-that are assumed to be orthogonal, that is, a message in channel $i$ will not cause collision to a message in another channel $j \neq i$.
+that are assumed to be orthogonal, that is, a message in channel $k$ will not cause collision to a message in another channel $j \neq i$.
 % (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
@@
 We further assume that there are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
-Furthermore, in channel $i \in \llbracket 1; K \rrbracket$ there are $0 \leq S_i \leq S$ static devices (so $S = \sum_{i=1}^{K} S_i$).
+Furthermore, in channel $k \in \llbracket 1; K \rrbracket$ there are $0 \leq S_k \leq S$ static devices (so $S = \sum_{k=1}^{K} S_k$).
 
@@
 % \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
-Indeed, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall i, S_i > 0$).
+Indeed, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$).
 Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
@@
 %
-They are however useful for our numerical simulations, and will be used as references, in order to show that our MAB-based approaches quickly learn to perform almost optimally.
+They are however useful for our numerical simulations, and are used as references, to show that our MAB-based approaches quickly learn to perform almost optimally.
 
@@
 
-In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $i$, at that time slot.
-The probability of successful transmission is computed as follows, because the $S_i$ static devices in each channel $i$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$:
+In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $k$, at that time slot.
+The probability of successful transmission is computed as follows, because the $S_k$ static devices in each channel $k$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$:
 \begin{equation}
-    \Pr(\text{success}|\text{sent}) = \sum_{i=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;i)}_{\text{No one else sent in channel}\; i} \; \underbrace{\Pr(\text{sent in channel}\,i)}_{= 1/K, \text{ by uniform choice}}
+    \Pr(\text{success}|\text{sent}) = \sum_{k=1}^{K} \underbrace{\Pr(\text{success}|\text{sent in channel}\;k)}_{\text{No one else sent in channel}\; k} \; \underbrace{\Pr(\text{sent in channel}\,k)}_{= 1/K, \text{ by uniform choice}}
 \end{equation}
 
-All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $i$ for any dynamic device is $p / K$, and there are $D-1$ other dynamic devices.
+All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $k$ for any dynamic device is $p / K$, and there are $D-1$ other dynamic devices.
 As they are independent, the probability that no other dynamic device sent in $i$
@@
     \Pr(\text{success}|\text{sent})
-    & = \sum_{i=1}^{K} \underbrace{(1 - p / K)^{D-1}}_{\text{No other dynamic device}} \times \underbrace{(1-p)^{S_i}}_{\text{No static device}} \times\; \frac{1}{K} \nonumber \\
-    & = \frac{1}{K} \left(1-\frac{p}{K}\right)^{D-1} \sum_{i=1}^{K} (1-p)^{S_i} .
+    & = \sum_{k=1}^{K} \underbrace{(1 - p / K)^{D-1}}_{\text{No other dynamic device}} \times \underbrace{(1-p)^{S_k}}_{\text{No static device}} \times\; \frac{1}{K} \nonumber \\
+    & = \frac{1}{K} \left(1-\frac{p}{K}\right)^{D-1} \sum_{k=1}^{K} (1-p)^{S_k} .
 \end{align}
@@
 
-We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of $(S_i)_i$ the repartition of static devices, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
+We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of $(S_k)_k$ the repartition of static devices, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
 This is not possible in IoT contexts as several completely independent networks may operate in a single place.} is possible.
@@
 
-If the oracle chooses a fixed configuration of dynamic devices, with $D_i$ devices affected to channel $i$ is fixed (in time, \ie, stationary),
+If the oracle chooses a fixed configuration of dynamic devices, with $D_k$ devices affected to channel $k$ is fixed (in time, \ie, stationary),
 then this probability is computed as before:
@@
     \Pr(\text{success}|\text{sent})
-    & = \sum_{i=1}^{K} \Pr(\text{success}|\text{sent in channel}\;i) \; \Pr(\text{sent in channel}\;i) \nonumber \\
-    & = \sum_{i=1}^{K} \underbrace{(1 - p)^{D_i - 1}}_{\;\;D_i - 1 \;\text{others}\;\;} \times \underbrace{(1 - p)^{S_i}}_{\;\;\text{No static device}\;\;} \times \underbrace{ D_i / D }_{\;\;\text{Sent in channel}\; i\;\;}.
+    & = \sum_{k=1}^{K} \Pr(\text{success}|\text{sent in channel}\;k) \; \Pr(\text{sent in channel}\;k) \nonumber \\
+    & = \sum_{k=1}^{K} \underbrace{(1 - p)^{D_k - 1}}_{\;\;D_k - 1 \;\text{others}\;\;} \times \underbrace{(1 - p)^{S_k}}_{\;\;\text{No static device}\;\;} \times \underbrace{ D_k / D }_{\;\;\text{Sent in channel}\; k\;\;}.
 \end{align}
@@
     % \begin{cases}
-    \underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
-    \text{such that}\;\; & \sum_{i=1}^{K} D_i = D, \label{eq:41:eqCstr}\\
-    & D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
+    \underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{k=1}^{K} D_k (1 - p)^{S_k + D_k -1}, \label{eq:41:optPb}\\
+    \text{such that}\;\; & \sum_{k=1}^{K} D_k = D, \label{eq:41:eqCstr}\\
+    & D_k \geq 0 \qquad \forall k\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
     % \end{cases}
@@
 
-    It gives a closed form expression for the (unique) optimal solution $D_i^*(\lambda)$, depending on the system parameters, and the unknown Lagrange multiplier $\lambda \in \mathbb{R}$.
+    It gives a closed form expression for the (unique) optimal solution $D_k^*(\lambda)$, depending on the system parameters, and the unknown Lagrange multiplier $\lambda \in \mathbb{R}$.
 
     \begin{equation}\label{eq:41:Dilambda}
-        D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{+} .
+        D_k^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_k-1}} \right)-1 \right]\right)^{+} .
     \end{equation}
@@
     \item
-    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;K \rrbracket$.
-For such values for $D_i$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
+    In a realistic scenario, we can assume that $D_k\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall k\in\llbracket 1;K \rrbracket$.
+For such values for $D_k$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{k=1}^{K} D_k (1 - p)^{S_k + D_k -1}$ is concave as the sum of concave functions
     \footnote{~It is worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
@@
 \begin{equation}\label{eq:41:constraintLambda}
- \sum_{i=1}^{K} D_i^*(\lambda) = D.
+    \sum_{k=1}^{K} D_k^*(\lambda) = D.
 \end{equation}
@@
 Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
-Solving the optimization problem provides the optimal real number value for $D_i^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $i$:
+Solving the optimization problem provides the optimal real number value for $D_k^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $k$:
 %
-$\widehat{D_i} = \lfloor D_i^* \rfloor$ for $1 \leq i < K$, and $\widehat{D_{K}} = D - \sum_{i=1}^{K - 1} \widehat{D_i}$.
+$\widehat{D_k} = \lfloor D_k^* \rfloor$ for $1 \leq k < K$, and $\widehat{D_{K}} = D - \sum_{k=1}^{K - 1} \widehat{D_k}$.
 
@@
 the third solution is a sub-optimal naive policy, simple to set up, but also unpractical as it also needs an oracle.
-End-devices are iteratively inserted in the channels with the lowest load (\ie, the index $i$ minimizing $S_i + D_i(\tau)$ at global time step $\tau$).
+End-devices are iteratively inserted in the channels with the lowest load (\ie, the index $k$ minimizing $S_k + D_k(\tau)$ at global time step $\tau$).
 Once the number of devices in each channel is computed, the probability of sending successfully a message is also given by equation \eqref{eq:41:prob_col}.
@@
 
-We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall i, S_i = S/K$), to check that learning approach (here, only \UCB)
+We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall k, S_k = S/K$), to check that learning approach (here, only \UCB)
 also gives interesting gain of performance, and achieve a close-to optimal successful transmission rate.

### 2019-04-23 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
So long
Je suis tellement inefficace c'est triste

@@
 regardless the type of devices, each of them follows a slotted ALOHA protocol \cite{Roberts75}, and has a probability $p>0$ to transmit a packet in a time slot.
-We make the hypothesis that the transmission is successful if the channel is available, otherwise upon radio collision.
+We make the hypothesis that the transmission is successful if the channel is available, otherwise it fails upon radio collision.
 The novelty compare to the previous model is that
@@
 % \begin{equation}\label{eq:43:Nkt}
-% 	N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(C(\tau) = k),
+% 	N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(A(\tau) = k),
 % \end{equation}
@@
 % \begin{equation}\label{eq:43:mukt}
-% 	\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(C(\tau) = k).
+% 	\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(A(\tau) = k).
 % \end{equation}
@@
 % \begin{equation}\label{eq:43:maxucb}
-% 	C(t) = \arg\max_{1\leq k \leq K} U_k(t).
+% 	A(t) = \arg\max_{1\leq k \leq K} U_k(t).
 % \end{equation}
@@
 % 			Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$\;
-% 			Transmit in channel $C(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
-% 			Reward $r_{C(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
+% 			Transmit in channel $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
+% 			Reward $r_{A(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
 % 		}
@@
 for any $t\in\mathbb{N}$,
-$N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(C(\tau) = k)$.
+$N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(A(\tau) = k)$.
 The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
-$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(C(\tau) = k)$,
+$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(A(\tau) = k)$,
 where $r_{k}(t)$ is the reward obtained after transmission in channel $k$ at time $t$.
@@
 as it is the one expected to be the best one at the current time step $t$,
-$C(t) = \arg\max_{1\leq k \leq K} U_k(t)$.
+$A(t) = \arg\max_{1\leq k \leq K} U_k(t)$.
 
@@
 				Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$\;
-				Transmit in channel $C(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
-				Reward $r_{C(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
+				Transmit in channel $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
+				Reward $r_{A(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
 			}
@@
 				j $\leftarrow$ last channel selected by first-stage \UCB\;
-				Transmit in channel $C(t) = j$\;
+				Transmit in channel $A(t) = j$\;
 			}
@@
 			\Else(\tcp*[f]{Random retransmission}){
-				Transmit in channel $C(t) \sim \mathcal{U}(1,\ldots,K)$\;
+				Transmit in channel $A(t) \sim \mathcal{U}(1,\ldots,K)$\;
 			}
@@
 
-Although, this approach increases the complexity and storage requirements (now, of order $\mathcal{O}(K^2)$).
-For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
-For instance, for $K=4$ channels,
-the memory to store $K+1=5$ algorithms is of the order of the requirements to store one.
-
-Note that for other networks this heuristic could not be practical.
-The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels (\eg, $K=128$).
-But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.
-
 \vspace*{-3pt}
@@
 				Compute $\forall k, U_k^j(t) = \widehat{\mu_k}^j(t) + \sqrt{\alpha \log(t) / N_k^j(t)}$\;
-				Transmit in channel $C^j(t) = \arg\max_{1\leq k \leq K} U^j_k(t)$\;
-				Reward $r^j_{C^j(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
+				Transmit in channel $A^j(t) = \arg\max_{1\leq k \leq K} U^j_k(t)$\;
+				Reward $r^j_{A^j(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
 			}
@@
 % \end{small} % XXX remove if needed
+
+Although, this approach increases the complexity and storage requirements (now, of order $\mathcal{O}(K^2)$).
+For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
+For instance, for $K=4$ channels,
+the memory to store $K+1=5$ algorithms is of the order of the requirements to store one.
+
+Note that for other networks this heuristic could not be practical.
+The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels (\eg, $K=128$).
+But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.
 
@@
 			\uElseIf(\tcp*[f]{Random retransmission}){$t \leq \Delta$}{
-				Transmit randomly in a channel $C(t) \sim \mathcal{U}(1,\ldots,K)$.
+				Transmit randomly in a channel $A(t) \sim \mathcal{U}(1,\ldots,K)$.
 			}
@@
 				Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + \sqrt{\alpha \log(t) / N_k^d(t)}$\;
-				Transmit in channel $C^d(t) = \arg\max_{1\leq k \leq K} U^d_k(t)$\;
-				Reward $r^d_{C^d(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
+				Transmit in channel $A^d(t) = \arg\max_{1\leq k \leq K} U^d_k(t)$\;
+				Reward $r^d_{A^d(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
 			}
@@
 
+\textbf{Remark.}
+%
 Similarly, Philippe Mary asked me at the MoTION workshop where I presented \cite{Bonnefoi2019WCNC}, why didn't we rather used Bayes-UCB instead of \UCB{} (from \cite{Kaufmann12BUCB}), as it is known to empirically outperform \UCB{} (as illustrated for instance in \cite{darak2016bayesian,kumar2016two}).

### 2019-04-23 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Je suis tellement inefficace c'est triste

@@
 \subsection{Towards regret upper bounds}
+\label{sub:5:towardsRegretUpperBounds}
 
@@
 %
-\vspace*{-5pt}
+% \vspace*{-5pt}
 \begin{equation}
@@
 \paragraph{Two ideas for orthogonalization?}
+\label{par:5:twoIdeasOrthogonalization}
 %
@@
 
-  \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
+  \begin{lemma}\label{lem:5:Fact2KLUCB}
+    For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
     \begin{small} % XXX
@@
         &\sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
-    + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
+        + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.
+      \end{align*}
+    \end{small} % XXX
 \end{lemma}
-  \end{small} % XXX
 
@@
 In the Appendix~E of \cite{Besson2018ALT},
-we discussed about when such situations may happen, including a conjectured (constant, but small) lower bound on the probability that \Selfish{} experience collision almost at every round.
+we explained when such situations may happen, and we included a conjectured (constant, but small) lower bound on the probability that \Selfish{} experience collision almost at every round.
 This result would then prevent \Selfish{} from having a logarithmic regret.

### 2019-04-23 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
HARD
Je suis tellement inefficace c'est triste

@@
   \Pr_{\bm{\lambda}}\left(\overline{A_T}\right)
-    & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_\bm{\lambda}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
+    & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_{\bm{\lambda}}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
 \end{align*}
@@
 \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right)  & = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
-& = & \E_{\bm{\mu}} \left[\kl \left(\mu}_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
-& = & \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu}_{k},\lambda_{k}\right)\right],
+& = & \E_{\bm{\mu}} \left[\kl \left(\mu_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
+& = & \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right],
 \end{eqnarray*}
@@
 Putting things together, we showed that
-\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) = \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu}_{k},\lambda_{k}\right)\right] + \smallO{\log(T)}.\]
+\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) = \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right] + \smallO{\log(T)}.\]
 Iterating this equality and using that $\KL\left(\Pr_{\bm{\mu}}^{\cO_{0}},\Pr_{\bm{\lambda}}^{\cO_{0}}\right)=0$ yields that
@@
 % -----------------------------------------------------------------
-\subsubsection{Illustration of the lower bound}
+\subsubsection{An illustration of the lower bound}
 \label{app:5:illustrationLowerBound}
 
-\TODOL{Maybe we can not include this part and just points to the paper \cite{Besson2018ALT}?}
+% \TODOL{Maybe we can not include this part and just points to the paper \cite{Besson2018ALT}?}
 
@@
   \centering
-  \includegraphics[width=0.85\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
+  \includegraphics[width=0.95\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
   \caption[Comparison of our lower bound against the one from \cite{Zhao10}]{Comparison of our lower bound against the one from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
@@
 
-%
-% System regret and three terms
-%
-\begin{figure}[!h]
-    \centering
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_2092905764868974160.pdf}
-    % \end{subfigure}
-    % ~
-    % \begin{subfigure}[!h]{1.00\textwidth}
-        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_1699475021767911583.pdf}
-    % \end{subfigure}
-    \caption[Multi-player regret with its three terms]{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
-    \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
-    % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
+\paragraph{Additional figures for numerical simulations.}
 
@@
 
+%
+% System regret and three terms
+%
+\begin{figure}[!h]
+    \centering
+    % \begin{subfigure}[!h]{1.00\textwidth}
+        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_2092905764868974160.pdf}
+    % \end{subfigure}
+    % ~
+    % \begin{subfigure}[!h]{1.00\textwidth}
+        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_1699475021767911583.pdf}
+    % \end{subfigure}
+    \caption[Multi-player regret with its three terms]{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
+    \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
+    % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
 % \subsubsection{Figures from Section~\ref{sec:5:experiments}}

### 2019-04-23 2-Chapters/5-Chapter/chapter5.tex
Je suis tellement inefficace c'est triste

@@
 \paragraph{Performance of our proposals for a wrong value of $M$?}
+\label{par:5:usingWrongValueofM}
 %
@@
 %
-The proposed algorithms in \cite{LugosiMehrabian18} are all based on the Musical Chair algorithm from \cite{Rosenski16}, and the curious reader should read their Algorithm~2 (page 15) for more details.
+The proposed algorithms in \cite{LugosiMehrabian18} are all based on the Musical Chair algorithm from \cite{Rosenski16}, and the curious reader should read for instance their Algorithm~2 (page 15) for more details.
 
@@
 In this extension, performance is still evaluated by a system regret, now defined as the difference between the sum of the cumulated rewards by the $M$ players, and the utility of any matching.
-The question is to possible to obtain a logarithmic --or even sub-linear-- regret in this problem is more difficult that for the model presented in Section~\ref{sec:5:model}.
+The question is to know if it is possible to obtain a logarithmic --or even sub-linear-- regret in this problem is more difficult that for the model presented in Section~\ref{sec:5:model}.
 This question was first answered in \cite{Bistritz18}, and proposed an algorithm based on alternating three phases, and increasing their lengths after the end of each epoch ($2^p$ for the $p$-th epoch).
@@
 On the other hand, another possibility is to extend ideas developed for the adversarial setting.
-In \cite{AlaturLevyKrause19}, the authors essentially consider the piece-wise stationary as an easier case of an adversarial problem. \\
+In \cite{AlaturLevyKrause19}, the authors essentially consider the piece-wise stationary as an easier case of an adversarial problem.
+The recent work \cite{bande2019adversarial} also proposes a decentralized algorithm that can achieve sub-linear regret ($\cO(T^{3/4})$) under an adversarial multi-players bandit model.
+\\
 \indent
-They build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf.
+In \cite{AlaturLevyKrause19}, the authors build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf.
 Algorithm 1), and they apparently discovered the ``communication trick'' independently from \cite{BoursierPerchet18} to use (virtual) communications between players to set up a ``coordinator'' player and $M-1$ ``followers'' (running respectively their Algorithms 2 and 3).

### 2019-04-23 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
HARD
Je suis tellement inefficace c'est triste

@@
 
-Finally, XXX use two discount factors at the same time, see the NEWMA algorithm from \cite{Keriven2018}.
-
-\TODOL{Talk about NEWMA and its possible application to generalize/improve D-UCB, \cite{Keriven2018}, if I have the time.}
+% Finally, XXX use two discount factors at the same time, see the NEWMA algorithm from \cite{Keriven2018}.
+
+% \TODOL{Talk about NEWMA and its possible application to generalize/improve D-UCB, \cite{Keriven2018}, if I have the time.}
 
@@
 
-In this section, we sketch a unified proof for the two Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given below, whose detailed proofs are given in Sections \ref{proof:6:mainRegretBoundGlobal} and \ref{proof:6:mainRegretBound} respectively.
+In this section, we sketch a unified proof for the two Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given below, whose detailed proofs are given in Section~\ref{proof:6:mainRegretBoundGlobal} and in Appendix~E of our paper \cite{Besson2019GLRT} respectively.
 We emphasize that our approach is significantly different from those proposed by \cite{CaoZhenKvetonXie18} for \MUCB{} and by \cite{LiuLeeShroff17} for \CUSUMUCB.
@@
 and we refer to the following page for details on how to reproduce them,
-\texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html}
+\href{https://SMPyBandits.GitHub.io/NonStationaryBandits.html}{\texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html}}

### 2019-04-23 5-Nomenclature/Nomenclature.tex
Je suis tellement inefficace c'est triste

@@
 
+% Computer science related terms
+\nomenclature[A]{CPU}{\textit{Central Processing Unit} (the electronic circuitry within a computer that carries out the instructions of a computer program by performing the basic arithmetic, logical, control and input/output (I/O) operations)}
+\nomenclature[A]{RAM}{\textit{Random Access Memory} (efficient working memory of a modern computer)}
+
 % Algorithms or tools
@@
 % matches Greek Symbols        > G
-\nomenclature[G]{$\mu_k,\mu_k(t),\mu_k^j$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$.}
+\nomenclature[G]{$\nu_k$,$\nu_k(t)$,$\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
+\nomenclature[G]{$\mu_k$,$\mu_k(t)$,$\mu_k^j$,$\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$ in Chapter~\ref{chapter:5}, arm $i$ in the $k$-th stationary interval in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\mu^*$,$\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
+\nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda},$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
+\nomenclature[G]{$\Delta$}{Usually used to denote the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}.
+In Chapter~\ref{chapter:4} in Algorithm~\ref{algo:43:UCBwithDelay} it denotes a delay.
+In Chapter~\ref{chapter:6}, it can denote different gaps}
+\nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\delta$}{Usually used to denote a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}.
+Also denote a confidence level of an algorithm in Section~\ref{sec:6:ChangePointDetector}}
+\nomenclature[G]{$\Upsilon_T$}{Number of break-points in a piece-wise stationary MAB problem in Chapter~\ref{chapter:6}.
+$\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
+\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Usually used to denote the location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\tau$}{Usually used to denote the length of a sliding-window algorithm like Sliding-Window-\UCB{} in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\gamma$}{Usually used to denote the discount factor of an algorithm like Discounted-\UCB{} in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\varepsilon$}{Usually used to denote a small positive real value, \eg, $\lambda_k = \mu_k + \varepsilon$ in Chapter~\ref{chapter:5} or a drift correction parameter in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\alpha$}{Usually used to denote the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\beta$}{Usually used to denote a threshold of GLR tests in Chapter~\ref{chapter:6}}
+
+\nomenclature[G]{$\alpha_0$, $\delta_0$}{Usually used to denote a constant scaling of a parameter of an algorithm, for instance $\alpha = \alpha_0 \alpha_T$ is used in Chapter~\ref{chapter:6}, where $\alpha_T$ is the formula suggested by the theoretical analysis and $\alpha_0 > 0$, \eg, $\alpha_0 = 0.01$}
 
@@
 % matches Roman Symbols        > L
-\nomenclature[L]{$t$}{Time step, $t\in\{1,\dots,T\}$}
-\nomenclature[L]{$T$}{Time horizon, the duration of the bandit game}
+\nomenclature[L]{$XXX$}{XXX}
+\nomenclature[L]{$t$, $s$, $n$, $r$}{Time step, $t\in\{1,\dots,T\}$. Chapter~\ref{chapter:6} also uses $s$, $n$ and $r$, \eg, in the $\sup$ of the stopping times or some technical lemmas}
+\nomenclature[L]{$T$}{Time horizon, the duration of the bandit game (always $T\geq1$)}
+\nomenclature[L]{$T_0$, $T_1$}{Fixed durations of some algorithms based on different phases, \eg, the Musical Chair algorithm from \cite{Rosenski16}}
+\nomenclature[L]{$\cA$, $\pi$}{A algorithm, a policy, designing the algorithm (\eg, \UCB) rather than the strategy.
+$\pi$ can also denote a permutation in Section~\ref{par:5:twoIdeasOrthogonalization}}
+\nomenclature[L]{$\rho$, $\rho^k$}{A bandit strategy, denoting the mathematical object which maps past observations to future decisions (mainly in Chapter~\ref{chapter:5})}
+\nomenclature[L]{$A(t)$, $A^j(t)$}{Decision of algorithm $\cA$ at time $t$, $A(t)\in\{1,\dots,K\}$. Decision for user $j$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$Y_{k,t}$}{Random sample from the $k$ arm at time $t$}
+\nomenclature[L]{$\cO_t$}{Vector of observations until time $t$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$r(t)$, $r^j(t)$}{Reward obtained at time $t$, for user $j$ at time $t$}
+\nomenclature[L]{$K$}{Number of arms for multi-armed bandit games}
+\nomenclature[L]{$k$}{Usually used to denote the $k$-th arm, $k\in\{1,\dots,K\}$}
+\nomenclature[L]{$M$}{In Chapter~\ref{chapter:4}, denote the maximum number of retransmission of a packet in the ALOHA protocol, and in Chapter~\ref{chapter:5}, denote the number of player for multi-players bandit games}
+\nomenclature[L]{$m$}{In Chapter~\ref{chapter:4}, denote the maximum length of the back-off interval after a collision}
+\nomenclature[L]{$i$, $j$}{Usually used to denote the $i$-th or $j$-th player, $i,j\in\{1,\dots,M\}$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$N$}{Usually used to denote the number of independent repetitions of the same numerical experiments (\eg, $N=1000$).
+In Chapter~\ref{chapter:4}, denote the number of IoT (dynamic) devices in the studied network}
+
+\nomenclature[L]{$R_T$, $R_T^{\cA}$}{Regret (of an algorithm $\cA$) for horizon $T$}
+
+\nomenclature[L]{$N_k(t)$ or $T_k(t)$}{Number of samples obtained for arm $k$ at time $t$ ($N_k(t)$ in Chapter~\ref{chapter:4}, $T_k(t)$ in Chapter~\ref{chapter:5})}
+\nomenclature[L]{$\widehat{\mu}_k(t)$}{Empirical mean of rewards obtained for arm $k$ at time $t$}
+\nomenclature[L]{$I_k(t)$}{Index of arm $k$ at time $t$ for an index policy}
+\nomenclature[L]{$\UCB_k(t)$}{Upper-Confidence Bound (\UCB) of arm $k$ at time $t$ for an index policy}
+\nomenclature[L]{$g_k^j(t)$,$g^j(t)$}{Index of (arm $k$) at time $t$ of user $j$ for an index policy in Chapter~\ref{chapter:5}}
+
+\nomenclature[L]{$p$}{In Chapter~\ref{chapter:4}, probability of transmission for the devices following a Bernoulli random emission pattern (\eg, $p=10^{-5}$)}
+\nomenclature[L]{$S$, $S_k$}{In Chapter~\ref{chapter:4}, total number of \emph{static} devices in the network, resp. in channel $k$}
+\nomenclature[L]{$D$, $D_k$}{In Chapter~\ref{chapter:4}, total number of \emph{dynamic} devices in the network, resp. in channel $k$}
+
+\nomenclature[L]{$C_k(t)$, $C^j(t)$}{In Chapters~\ref{chapter:4} and \ref{chapter:5}, collision indicator on arm $k$ or for user $j$ at time $j$}
+
+\nomenclature[L]{$C_{\bm{\mu}}$, $D_{\bm{\mu}}$, $G_{M,\bm{\mu}}$}{In Chapter~\ref{chapter:5}, constants depending on the problem parameters only (or also on $M$)}
+
+\nomenclature[L]{$\cH_0$, $\cH_1$, $\cH_2$}{Hypothesis, in Chapters~\ref{chapter:4} and \ref{chapter:6}}
+\nomenclature[L]{$d^{(k)}$, $d^{(\ell)}$}{Lower-bound on number of samples for accurate detection of change-points in Chapter~\ref{chapter:6}}
+\nomenclature[L]{$\cT$}{In Chapter~\ref{chapter:6}, denote a function in \eqref{def:6:function_T} and a set of time steps in Section~\ref{sub:6:IdeasOptimizations}}
 
@@
 % matches mathematical symbols > O
+\nomenclature[O]{$\kl$}{Binary relative entropy, KullBack-Leibler divergence between two Bernoulli distribution: $\kl(x,y) = x\log(x/y) + (1-x) \log((1-x)/(1-y))$}
+\nomenclature[O]{$\kl'$}{Derivative of the binary relative entropy function $\kl(x,y)$, with respect to $x$, for a fixed $y$, in Chapter~\ref{chapter:5}, see Lemma~\ref{lem:5:Fact2KLUCB}}
+\nomenclature[O]{$d(x,y)$}{Divergence function between two distributions of parameters $x$ and $y$, in a one-dimensional exponential family, in Chapter~\ref{chapter:6}}
+
+\nomenclature[O]{$\cW$}{Lambert $\cW$ function, \cite{Corless96}}
+\nomenclature[O]{$f$,$g$,$h$}{Real-valued functions, \eg, exploration function used for \klUCB{} indexes, $f(t) = \log(t) + 3 \log(\log(t))$}
+
+\nomenclature[O]{$\Pr$}{Probability measure under a probabilistic model}
+\nomenclature[O]{$\E$}{Expectancy under a probabilistic model}
+\nomenclature[O]{$\mathbbm{1}(E)$}{Indicator function of an event $E$ ($= 1$ if and only if the event $E$ is true)}
+\nomenclature[O]{$\cF$, $\cF_t$}{Filtration in a probabilistic model, after $t-1$ prior observations}
+\nomenclature[O]{$I_1,\dots,I_5$}{Used in Chapter~\ref{chapter:5} to denote various events}
+\nomenclature[O]{$\cE$, $\cE_T$}{Used in Chapter~\ref{chapter:6} to denote ``good events'' that happen most of the time}
+
+\nomenclature[O]{$\lfloor \bullet \rfloor$, $\lceil \bullet \rceil$}{Floor and ceiling funtions, $\lfloor x \rfloor = \sup\{n\in\Z, x \leq n\}$ and $\lceil x \rceil = \sup\{n\in\Z, x < n\}$}
+
+\nomenclature[O]{$o(\bullet)$, $\cO(\bullet)$, $\Omega(\bullet)$}{Landau notations: $f(x)=o(g(x)$ means that $g(x)\neq0$ and $f(x)/g(x)\to0$ for $x\to\infty$, $f(x)=\cO(g(x))$ means that there exists $x_0,K>0$ such that $f(x)\leq K g(x)$ from $x\geq x_0$, and $f(x)=\Omega(g(x))$ means $g(x)=\cO(f(x))$}
 
@@
 % matches Subscripts           > I
+\nomenclature[I]{$x_k$}{Usually the subscript index $k$ is used to denote a variable depending on an arm, for $k\in\{1,\dots,K\}$}
+\nomenclature[I]{$\overline{E}$}{Complement of an event $E$}
+\nomenclature[I]{$X'$}{Usually used to denote a ``wrong value'' of a variable $X$, for instance $M'$ denotes in Section~\ref{par:5:usingWrongValueofM}}
+\nomenclature[I]{$\widehat{X}$}{Usually used to denote an ``empirical value'', a mean or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{\mu_k}$ the empirical mean of arm $k$ or $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:towardsRegretUpperBounds}}
+\nomenclature[I]{$\widetilde{X}$}{Usually used to denote another ``empirical value'' or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:Selfish}.
+Also denote a surrogate for a function without a closed form, \eg, $\widetilde{T}$ in Section~\ref{sec:6:exploringDifferentThresholdFunctions}}
 
@@
 % matches Superscripts         > E
+\nomenclature[E]{$y^j$}{Usually the superscript index $j$ is used to denote a variable depending on a player, for $j\in\{1,\dots,M\}$}
+\nomenclature[E]{$y^{(k)}$}{Usually the superscript index $(k)$ is used to denote a variable in the $k$-th stationary interval, under the \emph{global} changes point-of-view, in Chapter~\ref{chapter:6}}
+\nomenclature[E]{$y^{(\ell)}$}{Usually the superscript index $(\ell)$ is used to denote a variable in the $\ell$-th stationary interval, under the \emph{local} changes point-of-view, in Chapter~\ref{chapter:6}}

### 2019-04-23 5-Nomenclature/Nomenclature.tex
Typo in nomenclature

@@
 \nomenclature[A]{DSA}{\textit{Dynamic Spectrum Access}}
-\nomenclature[A]{PHY}{\textit{Physical} (layer)}
+\nomenclature[A]{PHY}{\textit{PHYsical} (layer)}
 \nomenclature[A]{MAC}{\textit{Medium ACcess} (layer)}
@@
 % Computer science related terms
-\nomenclature[A]{CPU}{\textit{Central Processing Unit} (the electronic circuitry within a computer that carries out the instructions of a computer program by performing the basic arithmetic, logical, control and input/output (I/O) operations)}
-\nomenclature[A]{RAM}{\textit{Random Access Memory} (efficient working memory of a modern computer)}
+\nomenclature[A]{CPU}{\textit{Central Processing Unit}}
+%  (the electronic circuitry within a computer that carries out the instructions of a computer program by performing the basic arithmetic, logical, control and input/output (I/O) operations)
+\nomenclature[A]{RAM}{\textit{Random Access Memory}}
+%  (efficient working memory of a modern computer)
 
@@
 % matches Greek Symbols        > G
-\nomenclature[G]{$\nu_k$,$\nu_k(t)$,$\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
-\nomenclature[G]{$\mu_k$,$\mu_k(t)$,$\mu_k^j$,$\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$ in Chapter~\ref{chapter:5}, arm $i$ in the $k$-th stationary interval in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\mu^*$,$\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
+\nomenclature[G]{$\nu_k$, $\nu_k(t)$, $\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
+\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$ in Chapter~\ref{chapter:5}, arm $i$ in the $k$-th stationary interval in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
 \nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda},$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
-\nomenclature[G]{$\Delta$}{Usually used to denote the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}.
+\nomenclature[G]{$\Delta$}{Usually denotes the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}.
 In Chapter~\ref{chapter:4} in Algorithm~\ref{algo:43:UCBwithDelay} it denotes a delay.
@@
 \nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\delta$}{Usually used to denote a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}.
+\nomenclature[G]{$\delta$}{Usually denotes a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}.
 Also denote a confidence level of an algorithm in Section~\ref{sec:6:ChangePointDetector}}
@@
 $\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
-\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Usually used to denote the location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\tau$}{Usually used to denote the length of a sliding-window algorithm like Sliding-Window-\UCB{} in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\gamma$}{Usually used to denote the discount factor of an algorithm like Discounted-\UCB{} in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\varepsilon$}{Usually used to denote a small positive real value, \eg, $\lambda_k = \mu_k + \varepsilon$ in Chapter~\ref{chapter:5} or a drift correction parameter in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\alpha$}{Usually used to denote the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\beta$}{Usually used to denote a threshold of GLR tests in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Usually denotes the location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\tau$}{Usually denotes the length of a sliding-window algorithm like Sliding-Window-\UCB{} in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\gamma$}{Usually denotes the discount factor of an algorithm like Discounted-\UCB{} in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, $\lambda_k = \mu_k + \varepsilon$ in Chapter~\ref{chapter:5} or a drift correction parameter in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\beta$}{Usually denotes a threshold of GLR tests in Chapter~\ref{chapter:6}}
 
-\nomenclature[G]{$\alpha_0$, $\delta_0$}{Usually used to denote a constant scaling of a parameter of an algorithm, for instance $\alpha = \alpha_0 \alpha_T$ is used in Chapter~\ref{chapter:6}, where $\alpha_T$ is the formula suggested by the theoretical analysis and $\alpha_0 > 0$, \eg, $\alpha_0 = 0.01$}
+\nomenclature[G]{$\alpha_0$, $\delta_0$}{Usually denotes a constant scaling of a parameter of an algorithm, for instance $\alpha = \alpha_0 \alpha_T$ is used in Chapter~\ref{chapter:6}, where $\alpha_T$ is the formula suggested by the theoretical analysis and $\alpha_0 > 0$, \eg, $\alpha_0 = 0.01$}
 
@@
 % matches Roman Symbols        > L
-\nomenclature[L]{$XXX$}{XXX}
 \nomenclature[L]{$t$, $s$, $n$, $r$}{Time step, $t\in\{1,\dots,T\}$. Chapter~\ref{chapter:6} also uses $s$, $n$ and $r$, \eg, in the $\sup$ of the stopping times or some technical lemmas}
@@
 \nomenclature[L]{$K$}{Number of arms for multi-armed bandit games}
-\nomenclature[L]{$k$}{Usually used to denote the $k$-th arm, $k\in\{1,\dots,K\}$}
+\nomenclature[L]{$k$}{Usually denotes the $k$-th arm, $k\in\{1,\dots,K\}$, mainly used in subscripts}
 \nomenclature[L]{$M$}{In Chapter~\ref{chapter:4}, denote the maximum number of retransmission of a packet in the ALOHA protocol, and in Chapter~\ref{chapter:5}, denote the number of player for multi-players bandit games}
 \nomenclature[L]{$m$}{In Chapter~\ref{chapter:4}, denote the maximum length of the back-off interval after a collision}
-\nomenclature[L]{$i$, $j$}{Usually used to denote the $i$-th or $j$-th player, $i,j\in\{1,\dots,M\}$ in Chapter~\ref{chapter:5}}
-\nomenclature[L]{$N$}{Usually used to denote the number of independent repetitions of the same numerical experiments (\eg, $N=1000$).
+\nomenclature[L]{$i$, $j$}{Usually denotes the $i$-th or $j$-th player, $i,j\in\{1,\dots,M\}$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$N$}{Usually denotes the number of independent repetitions of the same numerical experiments (\eg, $N=1000$).
 In Chapter~\ref{chapter:4}, denote the number of IoT (dynamic) devices in the studied network}
@@
 \nomenclature[L]{$\UCB_k(t)$}{Upper-Confidence Bound (\UCB) of arm $k$ at time $t$ for an index policy}
-\nomenclature[L]{$g_k^j(t)$,$g^j(t)$}{Index of (arm $k$) at time $t$ of user $j$ for an index policy in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$g_k^j(t)$, $g^j(t)$}{Index of (arm $k$) at time $t$ of user $j$ for an index policy in Chapter~\ref{chapter:5}}
 
@@
 
-\nomenclature[O]{$\cW$}{Lambert $\cW$ function, \cite{Corless96}}
-\nomenclature[O]{$f$,$g$,$h$}{Real-valued functions, \eg, exploration function used for \klUCB{} indexes, $f(t) = \log(t) + 3 \log(\log(t))$}
+\nomenclature[O]{$\cW$}{Lambert $\cW$ function, the first branch of the inverse of $x\mapsto x \exp(x)$, cf. \cite{Corless96}}
+\nomenclature[O]{$f$, $g$, $h$}{Real-valued functions, \eg, exploration function used for \klUCB{} indexes, $f(t) = \log(t) + 3 \log(\log(t))$}
 
@@
 
-% --------------------------------
-% matches Subscripts           > I
-\nomenclature[I]{$x_k$}{Usually the subscript index $k$ is used to denote a variable depending on an arm, for $k\in\{1,\dots,K\}$}
-\nomenclature[I]{$\overline{E}$}{Complement of an event $E$}
-\nomenclature[I]{$X'$}{Usually used to denote a ``wrong value'' of a variable $X$, for instance $M'$ denotes in Section~\ref{par:5:usingWrongValueofM}}
-\nomenclature[I]{$\widehat{X}$}{Usually used to denote an ``empirical value'', a mean or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{\mu_k}$ the empirical mean of arm $k$ or $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:towardsRegretUpperBounds}}
-\nomenclature[I]{$\widetilde{X}$}{Usually used to denote another ``empirical value'' or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:Selfish}.
+\nomenclature[O]{$\overline{E}$}{Complement of an event $E$}
+\nomenclature[O]{$X'$}{Usually denotes a ``wrong value'' of a variable $X$, for instance $M'$ denotes in Section~\ref{par:5:usingWrongValueofM}}
+\nomenclature[O]{$\widehat{X}$}{Usually denotes an ``empirical value'', a mean or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{\mu_k}$ the empirical mean of arm $k$ or $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:towardsRegretUpperBounds}}
+\nomenclature[O]{$\widetilde{X}$}{Usually denotes another ``empirical value'' or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:Selfish}.
 Also denote a surrogate for a function without a closed form, \eg, $\widetilde{T}$ in Section~\ref{sec:6:exploringDifferentThresholdFunctions}}
@@
 % --------------------------------
+% matches Subscripts           > I
+\nomenclature[I]{$x_k$}{Usually the subscript index $k$ is denotes a variable depending on an arm, for $k\in\{1,\dots,K\}$}
+
+% --------------------------------
 % matches Superscripts         > E
-\nomenclature[E]{$y^j$}{Usually the superscript index $j$ is used to denote a variable depending on a player, for $j\in\{1,\dots,M\}$}
-\nomenclature[E]{$y^{(k)}$}{Usually the superscript index $(k)$ is used to denote a variable in the $k$-th stationary interval, under the \emph{global} changes point-of-view, in Chapter~\ref{chapter:6}}
-\nomenclature[E]{$y^{(\ell)}$}{Usually the superscript index $(\ell)$ is used to denote a variable in the $\ell$-th stationary interval, under the \emph{local} changes point-of-view, in Chapter~\ref{chapter:6}}
+\nomenclature[E]{$y^j$}{Usually the superscript index $j$ is denotes a variable depending on a player, for $j\in\{1,\dots,M\}$}
+\nomenclature[E]{$y^{(k)}$, $y^{(\ell)}$}{Usually the superscript index $(k)$ (resp.
+$(\ell)$) is denotes a variable in the $k$-th stationary interval, under the \emph{global} (resp.
+\emph{local}) changes point-of-view, in Chapter~\ref{chapter:6}}

### 2019-04-23 2-Chapters/1-Chapter/chapter1.tex
Talk to Emilie

@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
-    \texttt{HAL.Inria.fr/hal-02006825}.
+    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019,
+    \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
     \cite{Besson2019WCNC}
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann.
-    \emph{Algorithmic Learning Theory} (ALT, \texttt{www.cs.cornell.edu/conferences/alt2018}), Lanzarote, Spain, April 2018,
-    \texttt{HAL.Inria.fr/hal-01629733}.
+    \emph{Algorithmic Learning Theory} (ALT, \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}), Lanzarote, Spain, April 2018,
+    \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
     \cite{Besson2018ALT}
@@
     by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2018.ieee-wcnc.org}), Barcelone, Spain, April 2018,
-    \texttt{HAL.Inria.fr/hal-01705292}.
+    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}), Barcelone, Spain, April 2018,
+    \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
     \cite{Besson2018WCNC}
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
-    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \texttt{crowncom.org/2017}), Lisboa, Portugal, Septembre $2017$,
-    \texttt{HAL.Inria.fr/hal-01575419},
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}), Lisboa, Portugal, Septembre $2017$,
+    \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
     \textbf{Best Paper Award}.
@@
 \item
+    \emph{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation},\\
+    by Christophe Moy \& \textbf{Lilian Besson}.
+    1st International ISIoT workshop (\href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}),
+    at \emph{Conference on Distributed Computing in Sensor Systems} (IEEE DCOSS, \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}), Santorini, Greece, May 2019.
+    \cite{MoyBesson2019}
+
+\item
     \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
     by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
-    1st International MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}),
-    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019.
+    1st International MOTIoN workshop (\href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}),
+    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019.
     \cite{Bonnefoi2019WCNC}
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
-    Demonstration presented at \emph{International Conference on Communication} (ICT, \texttt{ict-2018.org/demos}), in Saint-Malo, France in June $2018$.
-    See \texttt{YouTu.be/HospLNQhcMk} for a $6$-minutes presentation video.
+    Demonstration presented at \emph{International Conference on Communication} (ICT, \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}), in Saint-Malo, France in June $2018$.
+    See \href{https://YouTu.be/HospLNQhcMk}{\texttt{YouTu.be/HospLNQhcMk}} for a $6$-minutes presentation video.
     \cite{Besson2018ICT}
@@
     \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
-    submitted at GRETSI 2019, \texttt{GRETSI.fr/colloque2019},
+    submitted at GRETSI 2019, \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}},
     by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
-    \texttt{HAL.Inria.fr/hal-02006471}.
+    \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
     \cite{Besson2019Gretsi}
-
-\item
-    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
-    submitted at COLT 2019, \texttt{LearningTheory.org/colt2019/},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
-    \texttt{HAL.Inria.fr/hal-02006471}.
-    \cite{Besson2019GLRT}
 
@@
 \begin{itemize}
+
+\item
+    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
+    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
+    \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
+    \cite{Besson2019GLRT}
+
 \item
@@
     by \textbf{Lilian Besson}, active development since October $2016$,
-    \texttt{HAL.Inria.fr/hal-01840022}.
-    About $40000$ lines of code, hosted on \texttt{GitHub.com/SMPyBandits},
-    with a complete documentation on \texttt{SMPyBandits.rtfd.io}.
+    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
+    About $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
+    with a complete documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}}.
     \cite{SMPyBandits,SMPyBanditsJMLR}
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
-    \texttt{HAL.Inria.fr/hal-01736357}.
+    \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.
     \cite{Besson2018DoublingTricks}
@@
     February $2018$,
-    \texttt{HAL.Inria.fr/hal-01847480}.
+    \href{https://HAL.Inria.fr/hal-01847480}{\texttt{HAL.Inria.fr/hal-01847480}}.
 
@@
     I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{Cf.
-\texttt{HAL.Inria.fr/hal-02013839/document}}.
+\href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
     I also helped for the organization and presented another poster\footnote{Cf.
-\texttt{HAL.Inria.fr/hal-02013847/document}} at IETR seminar (\texttt{seminar-ietr-18.sciencesconf.org}), held in Vannes in June $2018$.
+\href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Evaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.
 
 \paragraph{Reviews}
-	for \emph{European Workshop on Reinforcement Learning} (Lille) \texttt{ewrl.wordpress.com/ewrl14-2018}, in October $2018$, I reviewed $5$ papers.
+	for \emph{European Workshop on Reinforcement Learning} (Lille) \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}, in October $2018$, I reviewed $5$ papers.
 	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

### 2019-04-23 2-Chapters/3-Chapter/chapter3.tex
Talk to Emilie

@@
 %
-Finally, in Appendix~\ref{sec:3:appendix} we include some minimalist files showing how to use the library, as well as details regarding the use of parallel computing to speed-up simulations.
+Finally, in Appendix~\ref{sec:3:appendix} we include some small files showing how to use the library, as well as details regarding the use of parallel computing to speed-up simulations.
 
@@
 
+We have not yet tried to add support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
+%
+However, note that our library does support finite-state real-valued Markov MAB models, where arms do not correspond to distributions but to Markov chains, in the two variants of \emph{rested} or \emph{restless} Markov chains, as introduced by \cite{Anantharam87b}.
+
 \subsection{MAB algorithms}
@@
 
-\subsubsection{Features}
+\subsection{Features}
 
@@
 
-\subsubsection{How to run experiments?}
+\subsection{How to run experiments?}
 
@@
 
-\subsection{Continuous integration using Read the Docs and Travis CI}
+\subsection{Continuous integration with Read the Docs and Travis CI}
 
-% FIXME
-\TODOL{I need to quickly explain that SMPyBandits is using Read the Docs since $2017$ and Travis Ci since $2018$ to:
+% \TODOL{I need to quickly explain that SMPyBandits is using Read the Docs since $2017$ and Travis Ci since $2018$ to: }
 
-1.
-automatically build the documentation on Read the Docs after every commit.
+Since $2017$ and $2018$, SMPyBandits is using the free version of two online continuous integration (CI) service, to automatically build and host the documentation, and to automatically test the library on some numerical simulations.
+%
+These CI services are free for open-source work, and are both triggered whenever a modification of the codebase is sent to the hosting platform (\ie, GitHub).
+
+\paragraph{Read the Docs.}
+Since $2017$, we have been using the free web service provided by Read the Docs (\href{https://readthedocs.org/}{\texttt{ReadTheDocs.org}}).
+Read the Docs allows to automatically build the documentation after every commit.
 This allows to regularly check that the library is well formatted and can be imported correctly, as well as keeping the online documentation up-to-date.
+Moreover, they offer to host the documentation online, at \href{https://smpybandits.rtfd.io/}{\texttt{SMPyBandits.ReadTheDocs.io}}.
+For more details, see \href{https://readthedocs.org/projects/smpybandits/}{\texttt{https://ReadTheDocs.org/projects/SMPyBandits}}.
+%
+Since its first use, the service did about 180 builds, and about $10\%$ of them were useful to detect newly introduced issues or bugs in the code.
+The build is configured using the \texttt{.readthedocs.yml} file in SMPyBandits main folder, and it builds the documentation in about two minutes.
+I have also been building the documentation manually on a weekly basis, to also host it on \href{https://smpybandits.github.io/}{\texttt{SMPyBandits.GitHub.io}} thanks to GitHub pages.
+In the last two years, the documentation saw about 8500 unique visits, while the GitHub project have been visited about 3500 unique visits.
 
-2.
-automatically run short numerical simulations on Travis CI after every commit.
-This allows to check that no modification on any part of the codebase breaks anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
+\paragraph{Travis CI.}
+Since $2018$, we have been using the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
+Travis CI allows to automatically run short numerical simulations after every commit.
+This allows to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
 The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proved very useful to quickly find and fix new bugs.
-}
+For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
+%
+Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
+The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-player model).
+Builds typically run for $15$ minutes, and Travis CI have been proved to be very useful in our development process.
 
@@
 $(i)$
-Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic (see Chapter~\ref{chapter:6}) problems of single- or multi-player multi-armed bandits.
+Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-player multi-armed bandits.
 SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
@@
 %
-When comparing algorithms on a problem, the main performance measure is the regret, but the library also compute, store and visualize other measures such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
+When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
 
@@
 % \paragraph{Remark.}
-We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of SMPyBandits, and as such the library, its documentation and this chapter  are considered one of the main contribution of this thesis.
+We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of SMPyBandits, and as such the library, its documentation and this chapter  are considered an important contribution of this thesis.
 %
 Our library is used for the numerical simulations in all this thesis, except Chapter~\ref{chapter:4}.
-We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
+% We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
 We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} that we shortly present in Appendix~\ref{app:2:DoublingTricks}.
@@
 and even if some have already been implemented, a major future work is to implement the most interesting ones
-(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
-
+(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).\\
+%
+\indent
 Since $2017$, I also received by email or by GitHub issues some requests to implement new features or models.
@@
 Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
-Its documentation has seen already about $20000$ visits, the project on GitHub had $74$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
+Its documentation has seen already about $20000$ visits, the project on GitHub had $76$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
 %
-We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.
-
+We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.\\
+%
+\indent
 As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
@@
 It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
-It uses environment variables, so a user can configure some parameters of the simulation when launching them:
+It uses environment variables, so a user can configure some parameters of the simulation when launching them from a terminal:
 

### 2019-04-23 2-Chapters/4-Chapter/chapter4.tex
Go
Talk to Emilie

@@
 We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
-Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card is enough to simulate another gateway).
+Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card to simulate another gateway).
 %

### 2019-04-23 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Talk to Emilie

@@
 Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
-This class contains our proposals, \RandTopM{} and \MCTopM, but rule out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
+This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
 
@@
 \end{equation}
-
-\paragraph{The \RandTopM{} algorithm.}
-%
-First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
-It is essentially a refinement over the \RhoRand{} algorithm, and a simpler version of \MCTopM.
-The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
-%
-In \RandTopM{}, player $j$ is always considered ``not fixed'',
-and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.
 
@@
 % \vspace*{-5pt}  % XXX remove if problem
+
+\paragraph{The \RandTopM{} algorithm.}
+%
+First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
+It is essentially a refinement over the \RhoRand{} algorithm, to not use the indirection of ranks, and a simpler version of \MCTopM.
+The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
+%
+In \RandTopM{}, player $j$ is always considered ``not fixed'',
+and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.
 
@@
   \resizebox{1.00\textwidth}{!}{
-  \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
+  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.5]
       %
@@
 % -----------------------------------------------------------------
-\section{Theoretical elements, and order optimal regret for \MCTopM}
+\section{Theoretical elements, and regret upper bound for \MCTopM}
 \label{sec:5:upperbounds}
@@
 Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
-%
+
 We report in Table~\ref{table:5:comparisonUCB_klUCB} the results in terms of mean regret, $\pm$ one standard deviation.

### 2019-04-23 2-Chapters/5-Chapter/chapter5.tex
Talk to Emilie

@@
 
-We consider a bandit problem with $K=9$ arms, of means $0.1,0.2,\dots,0.8,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
+We consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
 As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require the know $M$.
-Then we compare the \RhoRand, \RandTopM{} and \MCTopM{} algorithms, using \klUCB, which know $M$ beforehand, with their extensions implementing the same algorithm as \rhoRandEst{} to estimate $M$ on the fly.
+Then we compare the \RhoRand, \RandTopM{} and \MCTopM{} algorithms, using \klUCB, which know $M$ beforehand, with their extensions implementing the same algorithm as \rhoRandEst, to estimate $M$ on the fly.
 
@@
 To illustrate the difficulty of the ``no sensing'' case, we illustrate here the performances of different algorithms designed specifically for this setting.
-As above, we consider a bandit problem with $K=9$ arms, of means $0.1,0.2,\dots,0.8,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
+As above, we consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
 We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
@@
 In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
-Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between players are possible.
+Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between devices are possible.
 Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
@@
 An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit, at every or only some time steps.
-We identified that the state-of-the-art on this direction of research consists in the two very recent works published in April 2019, while I was finishing this chapter: \cite{tao2019collaborative} and \cite{wang2019distributed}.
+We identified that the state-of-the-art on this direction of research consists in the two very recent works
+% published in April 2019, while I was finishing this chapter:
+\cite{tao2019collaborative} and \cite{wang2019distributed}.
 Distributed pure exploration is studied in \cite{tao2019collaborative}, where the proposed new lower bounds for the regret of any algorithm for the distributed best arm identification problem, under the fixed time or fixed confidence settings.

### 2019-04-23 3-Appendices/2-Appendix/appendix2.tex
So long
Talk to Emilie

@@
 % \chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
-\chapter{About Doubling-Trick for Multi-Armed Bandits}
+\chapter{About Doubling Tricks for Multi-Armed Bandits}
 \label{app:2:DoublingTricks}
@@
 
-\TODOL{
-I cannot include it raw without more work on the maths, we have to write only the best versions of our results.
+% \TODOL{
+% I cannot include it raw without more work on the maths, we have to write only the best versions of our results.
 
-But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!
+% But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!
 
-I need to try DoublingTrick + GLRklUCB, at least experimentally!
-For the theoretical analysis, it will probably be a huge failure.
-}
+% I need to try DoublingTrick + GLRklUCB, at least experimentally!
+% For the theoretical analysis, it will probably be a huge failure.
+% }
 

### 2019-04-23 5-Nomenclature/Nomenclature.tex
Talk to Emilie

@@
 \nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Usually denotes the location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\tau$}{Usually denotes the length of a sliding-window algorithm like Sliding-Window-\UCB{} in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\gamma$}{Usually denotes the discount factor of an algorithm like Discounted-\UCB{} in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\tau$}{Usually denotes the length of a sliding-window, \eg, in SW-\UCB{} in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\gamma$}{Usually denotes a discount factor, \eg, in D-\UCB{} or D-TS in Chapter~\ref{chapter:6}}
 \nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, $\lambda_k = \mu_k + \varepsilon$ in Chapter~\ref{chapter:5} or a drift correction parameter in Chapter~\ref{chapter:6}}
@@
 \nomenclature[O]{$\kl$}{Binary relative entropy, KullBack-Leibler divergence between two Bernoulli distribution: $\kl(x,y) = x\log(x/y) + (1-x) \log((1-x)/(1-y))$}
-\nomenclature[O]{$\kl'$}{Derivative of the binary relative entropy function $\kl(x,y)$, with respect to $x$, for a fixed $y$, in Chapter~\ref{chapter:5}, see Lemma~\ref{lem:5:Fact2KLUCB}}
+% \nomenclature[O]{$\kl'$}{Derivative of the binary relative entropy function $\kl(x,y)$, with respect to $x$, for a fixed $y$, in Chapter~\ref{chapter:5}, see Lemma~\ref{lem:5:Fact2KLUCB}}
 \nomenclature[O]{$d(x,y)$}{Divergence function between two distributions of parameters $x$ and $y$, in a one-dimensional exponential family, in Chapter~\ref{chapter:6}}
@@
 % matches Subscripts           > I
-\nomenclature[I]{$x_k$}{Usually the subscript index $k$ is denotes a variable depending on an arm, for $k\in\{1,\dots,K\}$}
+\nomenclature[I]{$x_k$}{Usually denotes a variable depending on an arm, for $k\in\{1,\dots,K\}$}
 
@@
 % matches Superscripts         > E
-\nomenclature[E]{$y^j$}{Usually the superscript index $j$ is denotes a variable depending on a player, for $j\in\{1,\dots,M\}$}
+\nomenclature[E]{$y^j$}{Usually denotes a variable depending on a player, for $j\in\{1,\dots,M\}$}
 \nomenclature[E]{$y^{(k)}$, $y^{(\ell)}$}{Usually the superscript index $(k)$ (resp.
-$(\ell)$) is denotes a variable in the $k$-th stationary interval, under the \emph{global} (resp.
+$(\ell)$) denotes a variable in the $k$-th stationary interval, under the \emph{global} (resp.
 \emph{local}) changes point-of-view, in Chapter~\ref{chapter:6}}

### 2019-04-23 2-Chapters/3-Chapter/chapter3.tex
Done for now

@@
 %
-This library tackles one dimensional distributions,
+We focus on one dimensional distributions,
 and supports \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Arms.Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Arms.Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Arms.Poisson}}), and a generic \emph{discrete} (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{Arms.DiscreteArm}}) distributions,
@@
 
-A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here: \texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
+A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here:
+\texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
 
@@
 
+\TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ?}
+
 Nowadays, parallelism is everywhere in the computational world, and any serious framework for numerical simulations must explore at least one of the three main approaches to (try to) gain performance from parallelism.

### 2019-04-23 2-Chapters/5-Chapter/chapter5.tex
Done for now

@@
 Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension of our proposal \MCTopM-\klUCB.
+
+% ----------------------------------------------------------------------------
+\subsection{Arrival and departures of players: the ``dynamic setting''}
+\label{sub:5:arrivalDepartures}
+
+As reminded above, our model assumes the number of players $M$ remain fixed during all the bandit game.
+However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
+The existing previous work on multi-player bandit models are all motivated by possible applications to wireless networks, but most of them assume $M$ to be fixed.
+The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
+If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the sensing information, as well as in the players' decisions (\ie, collisions):
+\begin{equation}\label{eq:5:defRegretDynamicSetting}
+    R_T^{\text{dyn}} = \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
+\end{equation}
+
+In \cite{Rosenski16}, the authors explain that if the model allows arrival or departures of players at \emph{any time step}, then the game is much harder, and sub-linear regret is most likely un-achievable, if we consider the natural extension of the definition of regret as in our model presented in Section~\ref{sec:5:model} above.
+However, this negative results depends on how arrival of players are modeled:
+if an arriving player has no prior memory of the previous observations, \ie, if it model a new device, then if there is no restriction on the number or frequency of arrival or departures, we can most likely prove that sub-linear regret is not achievable in general.
+A simple but extreme example shows that regret has to be linear for any bandit strategy (in the observation model with sensing).
+Consider $K=2$ Bernoulli-distributed arms of means $\mu_1=\mu_2=1/2$, and one player is always active and play any bandit strategy (\eg, \MCTopM-\klUCB).
+Every time step, another player is either arriving, without any knowledge of the problem (\eg, it is a fresh and new IoT device).
+Then no matter the strategy of these news players, even if they all play \MCTopM-\klUCB{} for instance, if they play a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
+
+In \cite{Rosenski16}, it is assumed that the players all know a lower-bound $L$ on the length of all the intervals during which arrival or departures of players are allowed.
+A naive idea, if $L$ is large enough, is simply to restart the underlying algorithm every $L$ time steps, in order to directly benefit from the theoretical guarantees of the ``static setting'' (where $M$ stays constant).
+However, this idea yields a large regret if the length of ``static'' intervals $L$ is too small.
+Applied to the \MusicalChair{} algorithm, the authors in \cite{Rosenski16} call this basic extension Dynamic \MusicalChair
+Under the simple hypothesis that the overall number of players entering and leaving the game is sub-linear in $T$ (\ie, a $\smallO{T}$), they analyze the regret of Dynamic \MusicalChair{} and prove it is also sub-linear (see Theorem~2 in Section~3.4).
+%
+Under the same hypothesis, and if the lower-bound $L$ is known before-hand and is large enough, then we could also apply the same idea as from \cite{Rosenski16} to our approach.
+We believe that we could easily prove that a sub-linear regret is achievable, if all players run \MCTopM-\klUCB{} (and if currently active players restart their memory of the past observations every $L$ time steps).
+As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, applying the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to our proposal \MCTopM{} would also give a small regret upper-bound.
+
+After being introduced in \cite{Rosenski16}, some more recent works studied the case of arrival or departures of players, usually referred to as ``dynamic setting'' or ``dynamic case''.
+For instance, \cite{BoursierPerchet18} studies in Section~4 the first algorithm proposed for the dynamic case under the no-sensing model.
+They study the same notion of regret as the one proposed in \cite{Rosenski16} and given above in \eqref{eq:5:defRegretDynamicSetting}, see Equation $(10)$ in Section~4.2.1.
+With the assumption of a non-decreasing number of players, \ie, if only arrivals of players are considered,
+they prove a regret upper-bound of the \textsc{Dyn-MMAB} algorithm in Theorem~3 .
+% Denote $M(t)$ the number of players at each time $t$, then they propose to study $R_T$ defined as
+% \[ R_T = \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu_^*{k} - \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t). \]
+If all arriving players are using their \textsc{Dyn-MMAB} algorithm, the ``dynamic'' regret is bounded by $\bigO{M^2 K \log(T) / \mu^*_{M}}$, if $M=M(T)$ is the total number of players involved in the problem.
+
+% \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
+
+% \cite{KumarYadav2018}
+
+% Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
+% Sumit J Darak, Manjesh K. Hanawal
+% https://arxiv.org/pdf/1812.11651
+% \cite{DarakHanawal18}
 
@@
 % ----------------------------------------------------------------------------
-\subsection{Arrival and departures of players: the ``dynamic setting''}
-\label{sub:5:arrivalDepartures}
-
-As reminded above, our model assumes the number of players $M$ remain fixed during all the bandit game.
-However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
-The existing previous work on multi-player bandit models are all motivated by possible applications to wireless networks, but most of them assume $M$ to be fixed.
-The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
-If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the sensing information, as well as in the players' decisions (\ie, collisions):
-\begin{equation}\label{eq:5:defRegretDynamicSetting}
-    R_T^{\text{dyn}} = \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
-\end{equation}
-
-In \cite{Rosenski16}, the authors explain that if the model allows arrival or departures of players at \emph{any time step}, then the game is much harder, and sub-linear regret is most likely un-achievable, if we consider the natural extension of the definition of regret as in our model presented in Section~\ref{sec:5:model} above.
-However, this negative results depends on how arrival of players are modeled:
-if an arriving player has no prior memory of the previous observations, \ie, if it model a new device, then if there is no restriction on the number or frequency of arrival or departures, we can most likely prove that sub-linear regret is not achievable in general.
-A simple but extreme example shows that regret has to be linear for any bandit strategy (in the observation model with sensing).
-Consider $K=2$ Bernoulli-distributed arms of means $\mu_1=\mu_2=1/2$, and one player is always active and play any bandit strategy (\eg, \MCTopM-\klUCB).
-Every time step, another player is either arriving, without any knowledge of the problem (\eg, it is a fresh and new IoT device).
-Then no matter the strategy of these news players, even if they all play \MCTopM-\klUCB{} for instance, if they play a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
-
-In \cite{Rosenski16}, it is assumed that the players all know a lower-bound $L$ on the length of all the intervals during which arrival or departures of players are allowed.
-A naive idea, if $L$ is large enough, is simply to restart the underlying algorithm every $L$ time steps, in order to directly benefit from the theoretical guarantees of the ``static setting'' (where $M$ stays constant).
-However, this idea yields a large regret if the length of ``static'' intervals $L$ is too small.
-Applied to the \MusicalChair{} algorithm, the authors in \cite{Rosenski16} call this basic extension Dynamic \MusicalChair
-Under the simple hypothesis that the overall number of players entering and leaving the game is sub-linear in $T$ (\ie, a $\smallO{T}$), they analyze the regret of Dynamic \MusicalChair{} and prove it is also sub-linear (see Theorem~2 in Section~3.4).
-%
-Under the same hypothesis, and if the lower-bound $L$ is known before-hand and is large enough, then we could also apply the same idea as from \cite{Rosenski16} to our approach.
-We believe that we could easily prove that a sub-linear regret is achievable, if all players run \MCTopM-\klUCB{} (and if currently active players restart their memory of the past observations every $L$ time steps).
-As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, applying the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to our proposal \MCTopM{} would also give a small regret upper-bound.
-
-After being introduced in \cite{Rosenski16}, some more recent works studied the case of arrival or departures of players, usually referred to as ``dynamic setting'' or ``dynamic case''.
-For instance, \cite{BoursierPerchet18} studies in Section~4 the first algorithm proposed for the dynamic case under the no-sensing model.
-They study the same notion of regret as the one proposed in \cite{Rosenski16} and given above in \eqref{eq:5:defRegretDynamicSetting}, see Equation $(10)$ in Section~4.2.1.
-With the assumption of a non-decreasing number of players, \ie, if only arrivals of players are considered,
-they prove a regret upper-bound of the \textsc{Dyn-MMAB} algorithm in Theorem~3 .
-% Denote $M(t)$ the number of players at each time $t$, then they propose to study $R_T$ defined as
-% \[ R_T = \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu_^*{k} - \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t). \]
-If all arriving players are using their \textsc{Dyn-MMAB} algorithm, the ``dynamic'' regret is bounded by $\bigO{M^2 K \log(T) / \mu^*_{M}}$, if $M=M(T)$ is the total number of players involved in the problem.
-
-% \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
-
-% \cite{KumarYadav2018}
-
-% Distributed Learning and Stable Orthogonalization in Ad-Hoc Networks with Heterogeneous Channels
-% Sumit J Darak, Manjesh K. Hanawal
-% https://arxiv.org/pdf/1812.11651
-% \cite{DarakHanawal18}
-
-% ----------------------------------------------------------------------------
 \subsection{With different arm utilities among players}
@@
 In the easier case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO(M K \log(K T) + K M^3 \log(M K T) / \Delta) + K M^2 (\log(M \log(K T) / \Delta))^2$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
-In the generic case, their algorithm M-ETC with Elimination achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$.
+In the generic case, their algorithm M-ETC with Elimination achieves a regret of $\cO((\log(T))^{1+\kappa})$ for any positive constant $\kappa$.
 
@@
 % ----------------------------------------------------------------------------
-\subsection{With malicious jammers}
-\label{sub:5:withDifferentMeansAmongPlayers}
-
-In all this chapter and the previous research literature, another common hypothesis is that all the $M$ players are pursuing the same goal, and are all behaving nicely by following the same algorithm.
-% Efficient utilization of licensed spectrum in the cognitive radio network is challenging due to lack of coordination among the Secondary Users (SUs).
-Distributed algorithms proposed in the literature aim to maximize the network throughput by ensuring orthogonal channel allocation for the SUs.
-However, these algorithms work under the assumption that all the SUs faithfully follow the algorithms which may not always hold due to the decentralized nature of the network.
-In the paper \cite{SawantKumar2018}, the authors study for the first time distributed algorithms that are robust against malicious behavior, also called \emph{jamming attack}.
-They consider both the cases of jammers launching coordinated and uncoordinated attacks, and consider a set of $J$ jammers.
-In the coordinated attack, the jammers select non-overlapping channels to attack in each time slot and can significantly increase the number of collisions for SUs.
-They setup the problem in each scenario as a multi-player bandit and develop algorithm,
-and their analysis shows that when the SUs faithfully implement proposed algorithms, the regret is constant with high probability.
-They validate their claims through exhaustive synthetic experiments and also through a realistic USRP.
-Their synthetic experiments consider different cases, for different values of $K$ the number of channels, $M$ the number of players and $J$ the number of jammers.
-
-% \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
-
-% ----------------------------------------------------------------------------
 \subsection{Modeling more closely a real wireless network}
@@
 
-Due to space constraint, we simply quote here three recent articles which proposed a similar approach of using decentralized reinforcement learning algorithms on the device-side on wireless networks,
+We simply quote here three recent articles which proposed a similar approach of using decentralized reinforcement learning algorithms on the device-side on wireless networks,
 but proposed models closer to the reality of wireless networks.
@@
 % ----------------------------------------------------------------------------
-\subsection{Real-world demonstrations to validate some models}
+\subsection{Inspirations from real-world demonstrations to validate some models}
 \label{sub:5:USRPdemos}
@@
 \TODOL{I should explain more, no? Who does what and why? So long...}
+
+% ----------------------------------------------------------------------------
+\subsection{With malicious jammers}
+\label{sub:5:withDifferentMeansAmongPlayers}
+
+In all this chapter and the previous research literature, another common hypothesis is that all the $M$ players are pursuing the same goal, and are all behaving nicely by following the same algorithm.
+% Efficient utilization of licensed spectrum in the cognitive radio network is challenging due to lack of coordination among the Secondary Users (SUs).
+Distributed algorithms proposed in the literature aim to maximize the network throughput by ensuring orthogonal channel allocation for the SUs.
+However, these algorithms work under the assumption that all the SUs faithfully follow the algorithms which may not always hold due to the decentralized nature of the network.
+In the paper \cite{SawantKumar2018}, the authors study for the first time distributed algorithms that are robust against malicious behavior, also called \emph{jamming attack}.
+They consider both the cases of jammers launching coordinated and uncoordinated attacks, and consider a set of $J$ jammers.
+In the coordinated attack, the jammers select non-overlapping channels to attack in each time slot and can significantly increase the number of collisions for SUs.
+They setup the problem in each scenario as a multi-player bandit and develop algorithm,
+and their analysis shows that when the SUs faithfully implement proposed algorithms, the regret is constant with high probability.
+They validate their claims through exhaustive synthetic experiments and also through a realistic USRP.
+Their synthetic experiments consider different cases, for different values of $K$ the number of channels, $M$ the number of players and $J$ the number of jammers.
+
+% \TODOL{Present the model, quote the main papers, explain the difficulty and how XXX algorithm solves it!}
 

### 2019-05-02 2-Chapters/3-Chapter/chapter3.tex
Almost done for text for chapter 3. I'm still not sure if I want to launch, copy-paste and comment on these simulations... What do they really give for the reader??

@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{exampleOfRewards.pdf}
+	\includegraphics[width=0.75\linewidth]{exampleOfRewards.pdf}
 	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$, $0.5$ and $0.9$.}
@@
 \subsection{Dependencies}
+\label{sub:3:dependencies}
 
@@
 
-Since $2017$ and $2018$, SMPyBandits is using the free version of two online continuous integration (CI) service, to automatically build and host the documentation, and to automatically test the library on some numerical simulations.
-%
-These CI services are free for open-source work, and are both triggered whenever a modification of the codebase is sent to the hosting platform (\ie, GitHub).
+Since $2017$ and $2018$, SMPyBandits is using two online continuous integration (CI) service, to automatically build and host the documentation, and to automatically test the library on some numerical simulations.
+%
+These CI services are free for open-source works, and are both triggered whenever a modification of the codebase is sent to the hosting platform (\ie, GitHub).
 
@@
 
--- In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+-- In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
 
--- In \cite{Besson2018DoublingTricks} (quickly presented in Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
+-- In \cite{Besson2018DoublingTricks} (quickly presented in Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
 and lots of bibliographic referencesIn sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
@@
 -- In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
-We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
+We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
 We designed the \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.GLR_UCB.html}{GLRklUCB}} policy, and implemented most of the policies from state-of-the-art research on passively or active adaptive policies, both adversarial- or stochastic-based, designed to tackle the piece-wise stationary MAB problems.
@@
 
+In this section, we use the SMPyBandits library to compare experimentally various state-of-the-art single-player algorithms on different kinds of multi-armed bandit problems.
+First, we detail the list of $9$ different algorithms that we compare, and the different problems on which they are all tested.
+We give the results of the numerical experiments, in terms of mean regret at the end of the experiment of different horizons $T$.
+Additional results in terms of real measurements of time and memory are given and discussed in the next Section~\ref{sec:3:timeAndMemoryCosts}.
+
+Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on using the simple \UCB{} algorithm when simplicity is favored, \ie, in the more applied Chapter~\ref{chapter:4}, and the \klUCB{} algorithm when efficiency is favored, \ie, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}.
+
+\subsection{Experimental setup: algorithms and problems}
+
+\paragraph{List of algorithms}
+
+We consider the nine following algorithms. For each of them, we give a bibliographic reference, that corresponds to the most recent article studying it and not the first one which introduced it.
+We give the chosen tuning of the parameters, for parametric algorithms.
+Algorithms that are not anytime use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, of course they use the correct successive values.
+We also give the name of the corresponding class in the \texttt{Policies} module in SMPyBandits.
+
+% FIXME order from oldest to most recent, and simplest to more complex
+
+\begin{itemize}
+    \item Algorithm 1 is
+    an $\varepsilon$-greedy algorithm, as described in \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
+
+    \item Algorithm 2 is
+    an Explore-then-Commit algorithm from \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
+
+    \item Algorithm 3 is
+    $\mathrm{Exp}3^{++}$ from \cite{Seldin17}, using $\alpha=3$ and $\beta=256$ as advised (\texttt{Exp3PlusPlus}).
+
+    \item Algorithm 4 is
+    $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCB}).
+
+    \item Algorithm 5 is
+    \klUCB{} from \cite{KLUCBJournal}, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern} (\texttt{klUCB}).
+
+    \item Algorithm 6 is
+    Thompson sampling from \cite{Kaufmann12Thompson}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{Thompson}).
+
+    \item Algorithm 7 is
+    Bayes-UCB from \cite{Kaufmann12BUCB}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
+
+    \item Algorithm 8 is
+    AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
+
+    \item Algorithm 9 is
+    BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, using a binary tournament implement in a naive but efficient way (not using recursive functions) (\texttt{BESA} with \texttt{''non\_recursive''=True}).
+\end{itemize}
+
+In the results presented in the main text of this chapter, we focus on the most well known algorithms, and two others that are efficient but less known (AdBandits and BESA).
+We present in Appendix~\ref{sub:3:additionalExperiments} seven other algorithms, being more recent (from $2016$, $3$ from $2018$ and $2$ from $2019$) but are usually more complex, or are based on a different point-of-view.
+We include numerical results for these other algorithms as well, to justify empirically that despite their respective qualities, and despite being efficient, it is reasonable to focus on \UCB{} and \klUCB{} algorithms in the rest of this thesis.
+
+\paragraph{Two different problems}
+
+We consider two different problems, that are described here for a generic value of $K \geq 2$ the number of arms.
+For simplicity, we focus on Bernoulli distributed arms, even if similar results were observed for other distributions, in particular for truncated or untruncated Gaussian distributions.
+
+\begin{itemize}
+    \item \textbf{Problem 1} considers $K$ arms of equally spaced means, in $[0,1]$.
+    It uses a minimum mean of $0.05$ and a maximum mean of $0.95$.
+    For instance, for $K=2$, $\bm{\mu}=[0.05, 0.95]$ and for $K=5$, $\bm{\mu} = [0.05, 0.275, 0.5, 0.725, 0.95]$.
+
+    \item \textbf{Problem 2} takes another approach. Instead of simulating $1000$ times the same problem, it generates a new random problem for each independent run.
+    It considers a randomly generated vector of means, each being sampled uniformly at random in $[0,1]$, with a minimum gap of $\Delta^{\min}$, $\bm{\mu} \sim \cE(\Delta^{\min},\mu_{\min},\mu_{\max})$, as well as minimum and maximum values of $\mu_{\min}$ and $\mu_{\max}$.
+    This space $\cE_{\Delta^{\min}}$ is defined as $\{ \bm{\mu} \in [\mu_{\min}, \mu_{\max}]^K : \min_{k\neq k'} |\mu_k - \mu_{k'}| \geq \Delta^{\min} \}$.
+    We use $\mu_{\min} = \Delta^{\min}$ and $\mu_{\max} = 1 - \Delta^{\min}$,
+    and the value used for $\Delta^{\min}$ is $0.1$ for $K \leq 3$ (for ``easy'' problems), and $\frac{1}{3 K}$ for $K \geq 5$ (for ``harder'' problems).
+    % 0.1 if NB_ARMS <= 5 else 1. / (3 * NB_ARMS)
+\end{itemize}
+
+\paragraph{Other parameters of the experiments}
+
+We consider increasing values for the horizon, $T=1000$, $T=5000$, $T=10000$ and $T=50000$.  % FIXME update if needed?
+We consider as well an increasing number of arms, $K=2$, $K=5$, $K=9$, $K=20$ and $K=50$.  % FIXME update if needed?
+For all experiments, we run $N=1000$ independent simulations.  % FIXME update if needed?
+
+\subsection{Experiments results}
+
+We give in the tables below the results of the experiments.
+For the two different problems and different values of $T$ and $K$, we give the mean $\pm 1$ standard-variation of the empirical regret $R_T$ obtained by the different algorithms at the end of the experiments.
+Tables~\ref{table:3:meanRegret_problem1} is for the fixed-gap \textbf{problem 1}
+and \ref{table:3:meanRegret_problem2}  is for the randomly generated \textbf{problem 2}.
+Each column corresponds to one value of $T$, and each line corresponds to one algorithm.
+For each cell, we report successive values that correspond to the successive values of the number of arms $K$.
+
+\TODOL{FIXME launch these experiments, for a very short number of repetitions N=4, and if I'm happy about the results and if they are easily copied-and-pasted here, then launch with N=100 or N=1000 depending on the running
+
+% DEBUG=True; NOPLOTS=False; SAVEALL=False; for N in 4 100 1000; do for BAYES in False True; do for T in 1000 5000 10000 50000; do for K in 2 5 10 20 50; do debut=$(date); DEBUG=$DEBUG NOPLOTS=$NOPLOTS SAVEALL=$SAVEALL N=$N N_JOBS=-1 T=$T K=$K BAYES=$BAYES make --noFreeSMS single && cp logs/main_py3_log.txt logs/main_py3_log__N${N}_BAYES${BAYES}_T{T}_K{K}.txt && FreeSMS.py "Expérience commencée ${debut} terminée, avec N=${N}, BAYES=${BAYES}, T=${T}, K=${K}." ; done; done; done; done time.}
+
+\begin{table}[!t]
+\begin{footnotesize}
+    \centering
+    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+        \hline
+        $\varepsilon$-greedy &
+            $R^{1}_{T_1,K_1} = 0 \pm 0$
+                $R^{1}_{T_1,K_2} = 0 \pm 0$
+                $R^{1}_{T_1,K_3} = 0 \pm 0$
+                $R^{1}_{T_1,K_4} = 0 \pm 0$
+                $R^{1}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{1}_{T_2,K_1} = 0 \pm 0$
+                $R^{1}_{T_2,K_2} = 0 \pm 0$
+                $R^{1}_{T_2,K_3} = 0 \pm 0$
+                $R^{1}_{T_2,K_4} = 0 \pm 0$
+                $R^{1}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{1}_{T_3,K_1} = 0 \pm 0$
+                $R^{1}_{T_3,K_2} = 0 \pm 0$
+                $R^{1}_{T_3,K_3} = 0 \pm 0$
+                $R^{1}_{T_3,K_4} = 0 \pm 0$
+                $R^{1}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{1}_{T_4,K_1} = 0 \pm 0$
+                $R^{1}_{T_4,K_2} = 0 \pm 0$
+                $R^{1}_{T_4,K_3} = 0 \pm 0$
+                $R^{1}_{T_4,K_4} = 0 \pm 0$
+                $R^{1}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        Explore-then-Commit &
+            $R^{2}_{T_1,K_1} = 0 \pm 0$
+                $R^{2}_{T_1,K_2} = 0 \pm 0$
+                $R^{2}_{T_1,K_3} = 0 \pm 0$
+                $R^{2}_{T_1,K_4} = 0 \pm 0$
+                $R^{2}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{2}_{T_2,K_1} = 0 \pm 0$
+                $R^{2}_{T_2,K_2} = 0 \pm 0$
+                $R^{2}_{T_2,K_3} = 0 \pm 0$
+                $R^{2}_{T_2,K_4} = 0 \pm 0$
+                $R^{2}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{2}_{T_3,K_1} = 0 \pm 0$
+                $R^{2}_{T_3,K_2} = 0 \pm 0$
+                $R^{2}_{T_3,K_3} = 0 \pm 0$
+                $R^{2}_{T_3,K_4} = 0 \pm 0$
+                $R^{2}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{2}_{T_4,K_1} = 0 \pm 0$
+                $R^{2}_{T_4,K_2} = 0 \pm 0$
+                $R^{2}_{T_4,K_3} = 0 \pm 0$
+                $R^{2}_{T_4,K_4} = 0 \pm 0$
+                $R^{2}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        $\mathrm{Exp}3^{++}$ &
+            $R^3_{T_1,K_1} = 0 \pm 0$
+                $R^3_{T_1,K_2} = 0 \pm 0$
+                $R^3_{T_1,K_3} = 0 \pm 0$
+                $R^3_{T_1,K_4} = 0 \pm 0$
+                $R^3_{T_1,K_5} = 0 \pm 0$ &
+            $R^3_{T_2,K_1} = 0 \pm 0$
+                $R^3_{T_2,K_2} = 0 \pm 0$
+                $R^3_{T_2,K_3} = 0 \pm 0$
+                $R^3_{T_2,K_4} = 0 \pm 0$
+                $R^3_{T_2,K_5} = 0 \pm 0$ &
+            $R^3_{T_3,K_1} = 0 \pm 0$
+                $R^3_{T_3,K_2} = 0 \pm 0$
+                $R^3_{T_3,K_3} = 0 \pm 0$
+                $R^3_{T_3,K_4} = 0 \pm 0$
+                $R^3_{T_3,K_5} = 0 \pm 0$ &
+            $R^3_{T_4,K_1} = 0 \pm 0$
+                $R^3_{T_4,K_2} = 0 \pm 0$
+                $R^3_{T_4,K_3} = 0 \pm 0$
+                $R^3_{T_4,K_4} = 0 \pm 0$
+                $R^3_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        $\UCB_1$ &
+            $R^{4}_{T_1,K_1} = 0 \pm 0$
+                $R^{4}_{T_1,K_2} = 0 \pm 0$
+                $R^{4}_{T_1,K_3} = 0 \pm 0$
+                $R^{4}_{T_1,K_4} = 0 \pm 0$
+                $R^{4}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{4}_{T_2,K_1} = 0 \pm 0$
+                $R^{4}_{T_2,K_2} = 0 \pm 0$
+                $R^{4}_{T_2,K_3} = 0 \pm 0$
+                $R^{4}_{T_2,K_4} = 0 \pm 0$
+                $R^{4}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{4}_{T_3,K_1} = 0 \pm 0$
+                $R^{4}_{T_3,K_2} = 0 \pm 0$
+                $R^{4}_{T_3,K_3} = 0 \pm 0$
+                $R^{4}_{T_3,K_4} = 0 \pm 0$
+                $R^{4}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{4}_{T_4,K_1} = 0 \pm 0$
+                $R^{4}_{T_4,K_2} = 0 \pm 0$
+                $R^{4}_{T_4,K_3} = 0 \pm 0$
+                $R^{4}_{T_4,K_4} = 0 \pm 0$
+                $R^{4}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        \klUCB{} &
+            $R^{5}_{T_1,K_1} = 0 \pm 0$
+                $R^{5}_{T_1,K_2} = 0 \pm 0$
+                $R^{5}_{T_1,K_3} = 0 \pm 0$
+                $R^{5}_{T_1,K_4} = 0 \pm 0$
+                $R^{5}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{5}_{T_2,K_1} = 0 \pm 0$
+                $R^{5}_{T_2,K_2} = 0 \pm 0$
+                $R^{5}_{T_2,K_3} = 0 \pm 0$
+                $R^{5}_{T_2,K_4} = 0 \pm 0$
+                $R^{5}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{5}_{T_3,K_1} = 0 \pm 0$
+                $R^{5}_{T_3,K_2} = 0 \pm 0$
+                $R^{5}_{T_3,K_3} = 0 \pm 0$
+                $R^{5}_{T_3,K_4} = 0 \pm 0$
+                $R^{5}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{5}_{T_4,K_1} = 0 \pm 0$
+                $R^{5}_{T_4,K_2} = 0 \pm 0$
+                $R^{5}_{T_4,K_3} = 0 \pm 0$
+                $R^{5}_{T_4,K_4} = 0 \pm 0$
+                $R^{5}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        Thompson sampling &
+            $R^{6}_{T_1,K_1} = 0 \pm 0$
+                $R^{6}_{T_1,K_2} = 0 \pm 0$
+                $R^{6}_{T_1,K_3} = 0 \pm 0$
+                $R^{6}_{T_1,K_4} = 0 \pm 0$
+                $R^{6}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{6}_{T_2,K_1} = 0 \pm 0$
+                $R^{6}_{T_2,K_2} = 0 \pm 0$
+                $R^{6}_{T_2,K_3} = 0 \pm 0$
+                $R^{6}_{T_2,K_4} = 0 \pm 0$
+                $R^{6}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{6}_{T_3,K_1} = 0 \pm 0$
+                $R^{6}_{T_3,K_2} = 0 \pm 0$
+                $R^{6}_{T_3,K_3} = 0 \pm 0$
+                $R^{6}_{T_3,K_4} = 0 \pm 0$
+                $R^{6}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{6}_{T_4,K_1} = 0 \pm 0$
+                $R^{6}_{T_4,K_2} = 0 \pm 0$
+                $R^{6}_{T_4,K_3} = 0 \pm 0$
+                $R^{6}_{T_4,K_4} = 0 \pm 0$
+                $R^{6}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        Thompson sampling &
+            $R^{7}_{T_1,K_1} = 0 \pm 0$
+                $R^{7}_{T_1,K_2} = 0 \pm 0$
+                $R^{7}_{T_1,K_3} = 0 \pm 0$
+                $R^{7}_{T_1,K_4} = 0 \pm 0$
+                $R^{7}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{7}_{T_2,K_1} = 0 \pm 0$
+                $R^{7}_{T_2,K_2} = 0 \pm 0$
+                $R^{7}_{T_2,K_3} = 0 \pm 0$
+                $R^{7}_{T_2,K_4} = 0 \pm 0$
+                $R^{7}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{7}_{T_3,K_1} = 0 \pm 0$
+                $R^{7}_{T_3,K_2} = 0 \pm 0$
+                $R^{7}_{T_3,K_3} = 0 \pm 0$
+                $R^{7}_{T_3,K_4} = 0 \pm 0$
+                $R^{7}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{7}_{T_4,K_1} = 0 \pm 0$
+                $R^{7}_{T_4,K_2} = 0 \pm 0$
+                $R^{7}_{T_4,K_3} = 0 \pm 0$
+                $R^{7}_{T_4,K_4} = 0 \pm 0$
+                $R^{7}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        AdBandits &
+            $R^{8}_{T_1,K_1} = 0 \pm 0$
+                $R^{8}_{T_1,K_2} = 0 \pm 0$
+                $R^{8}_{T_1,K_3} = 0 \pm 0$
+                $R^{8}_{T_1,K_4} = 0 \pm 0$
+                $R^{8}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{8}_{T_2,K_1} = 0 \pm 0$
+                $R^{8}_{T_2,K_2} = 0 \pm 0$
+                $R^{8}_{T_2,K_3} = 0 \pm 0$
+                $R^{8}_{T_2,K_4} = 0 \pm 0$
+                $R^{8}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{8}_{T_3,K_1} = 0 \pm 0$
+                $R^{8}_{T_3,K_2} = 0 \pm 0$
+                $R^{8}_{T_3,K_3} = 0 \pm 0$
+                $R^{8}_{T_3,K_4} = 0 \pm 0$
+                $R^{8}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{8}_{T_4,K_1} = 0 \pm 0$
+                $R^{8}_{T_4,K_2} = 0 \pm 0$
+                $R^{8}_{T_4,K_3} = 0 \pm 0$
+                $R^{8}_{T_4,K_4} = 0 \pm 0$
+                $R^{8}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        BESA &
+            $R^{9}_{T_1,K_1} = 0 \pm 0$
+                $R^{9}_{T_1,K_2} = 0 \pm 0$
+                $R^{9}_{T_1,K_3} = 0 \pm 0$
+                $R^{9}_{T_1,K_4} = 0 \pm 0$
+                $R^{9}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{9}_{T_2,K_1} = 0 \pm 0$
+                $R^{9}_{T_2,K_2} = 0 \pm 0$
+                $R^{9}_{T_2,K_3} = 0 \pm 0$
+                $R^{9}_{T_2,K_4} = 0 \pm 0$
+                $R^{9}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{9}_{T_3,K_1} = 0 \pm 0$
+                $R^{9}_{T_3,K_2} = 0 \pm 0$
+                $R^{9}_{T_3,K_3} = 0 \pm 0$
+                $R^{9}_{T_3,K_4} = 0 \pm 0$
+                $R^{9}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{9}_{T_4,K_1} = 0 \pm 0$
+                $R^{9}_{T_4,K_2} = 0 \pm 0$
+                $R^{9}_{T_4,K_3} = 0 \pm 0$
+                $R^{9}_{T_4,K_4} = 0 \pm 0$
+                $R^{9}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+    \end{tabular}
+    \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 1} with different values of time $T$ and number of arm $K$.
+    }
+    \label{table:3:meanRegret_problem1}
+\end{footnotesize}
+\end{table}
+
+\begin{table}[!t]
+\begin{footnotesize}
+    \centering
+    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+        \hline
+        $\varepsilon$-greedy &
+            $R^{1}_{T_1,K_1} = 0 \pm 0$
+                $R^{1}_{T_1,K_2} = 0 \pm 0$
+                $R^{1}_{T_1,K_3} = 0 \pm 0$
+                $R^{1}_{T_1,K_4} = 0 \pm 0$
+                $R^{1}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{1}_{T_2,K_1} = 0 \pm 0$
+                $R^{1}_{T_2,K_2} = 0 \pm 0$
+                $R^{1}_{T_2,K_3} = 0 \pm 0$
+                $R^{1}_{T_2,K_4} = 0 \pm 0$
+                $R^{1}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{1}_{T_3,K_1} = 0 \pm 0$
+                $R^{1}_{T_3,K_2} = 0 \pm 0$
+                $R^{1}_{T_3,K_3} = 0 \pm 0$
+                $R^{1}_{T_3,K_4} = 0 \pm 0$
+                $R^{1}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{1}_{T_4,K_1} = 0 \pm 0$
+                $R^{1}_{T_4,K_2} = 0 \pm 0$
+                $R^{1}_{T_4,K_3} = 0 \pm 0$
+                $R^{1}_{T_4,K_4} = 0 \pm 0$
+                $R^{1}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        Explore-then-Commit &
+            $R^{2}_{T_1,K_1} = 0 \pm 0$
+                $R^{2}_{T_1,K_2} = 0 \pm 0$
+                $R^{2}_{T_1,K_3} = 0 \pm 0$
+                $R^{2}_{T_1,K_4} = 0 \pm 0$
+                $R^{2}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{2}_{T_2,K_1} = 0 \pm 0$
+                $R^{2}_{T_2,K_2} = 0 \pm 0$
+                $R^{2}_{T_2,K_3} = 0 \pm 0$
+                $R^{2}_{T_2,K_4} = 0 \pm 0$
+                $R^{2}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{2}_{T_3,K_1} = 0 \pm 0$
+                $R^{2}_{T_3,K_2} = 0 \pm 0$
+                $R^{2}_{T_3,K_3} = 0 \pm 0$
+                $R^{2}_{T_3,K_4} = 0 \pm 0$
+                $R^{2}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{2}_{T_4,K_1} = 0 \pm 0$
+                $R^{2}_{T_4,K_2} = 0 \pm 0$
+                $R^{2}_{T_4,K_3} = 0 \pm 0$
+                $R^{2}_{T_4,K_4} = 0 \pm 0$
+                $R^{2}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        $\mathrm{Exp}3^{++}$ &
+            $R^3_{T_1,K_1} = 0 \pm 0$
+                $R^3_{T_1,K_2} = 0 \pm 0$
+                $R^3_{T_1,K_3} = 0 \pm 0$
+                $R^3_{T_1,K_4} = 0 \pm 0$
+                $R^3_{T_1,K_5} = 0 \pm 0$ &
+            $R^3_{T_2,K_1} = 0 \pm 0$
+                $R^3_{T_2,K_2} = 0 \pm 0$
+                $R^3_{T_2,K_3} = 0 \pm 0$
+                $R^3_{T_2,K_4} = 0 \pm 0$
+                $R^3_{T_2,K_5} = 0 \pm 0$ &
+            $R^3_{T_3,K_1} = 0 \pm 0$
+                $R^3_{T_3,K_2} = 0 \pm 0$
+                $R^3_{T_3,K_3} = 0 \pm 0$
+                $R^3_{T_3,K_4} = 0 \pm 0$
+                $R^3_{T_3,K_5} = 0 \pm 0$ &
+            $R^3_{T_4,K_1} = 0 \pm 0$
+                $R^3_{T_4,K_2} = 0 \pm 0$
+                $R^3_{T_4,K_3} = 0 \pm 0$
+                $R^3_{T_4,K_4} = 0 \pm 0$
+                $R^3_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        $\UCB_1$ &
+            $R^{4}_{T_1,K_1} = 0 \pm 0$
+                $R^{4}_{T_1,K_2} = 0 \pm 0$
+                $R^{4}_{T_1,K_3} = 0 \pm 0$
+                $R^{4}_{T_1,K_4} = 0 \pm 0$
+                $R^{4}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{4}_{T_2,K_1} = 0 \pm 0$
+                $R^{4}_{T_2,K_2} = 0 \pm 0$
+                $R^{4}_{T_2,K_3} = 0 \pm 0$
+                $R^{4}_{T_2,K_4} = 0 \pm 0$
+                $R^{4}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{4}_{T_3,K_1} = 0 \pm 0$
+                $R^{4}_{T_3,K_2} = 0 \pm 0$
+                $R^{4}_{T_3,K_3} = 0 \pm 0$
+                $R^{4}_{T_3,K_4} = 0 \pm 0$
+                $R^{4}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{4}_{T_4,K_1} = 0 \pm 0$
+                $R^{4}_{T_4,K_2} = 0 \pm 0$
+                $R^{4}_{T_4,K_3} = 0 \pm 0$
+                $R^{4}_{T_4,K_4} = 0 \pm 0$
+                $R^{4}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        \klUCB{} &
+            $R^{5}_{T_1,K_1} = 0 \pm 0$
+                $R^{5}_{T_1,K_2} = 0 \pm 0$
+                $R^{5}_{T_1,K_3} = 0 \pm 0$
+                $R^{5}_{T_1,K_4} = 0 \pm 0$
+                $R^{5}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{5}_{T_2,K_1} = 0 \pm 0$
+                $R^{5}_{T_2,K_2} = 0 \pm 0$
+                $R^{5}_{T_2,K_3} = 0 \pm 0$
+                $R^{5}_{T_2,K_4} = 0 \pm 0$
+                $R^{5}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{5}_{T_3,K_1} = 0 \pm 0$
+                $R^{5}_{T_3,K_2} = 0 \pm 0$
+                $R^{5}_{T_3,K_3} = 0 \pm 0$
+                $R^{5}_{T_3,K_4} = 0 \pm 0$
+                $R^{5}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{5}_{T_4,K_1} = 0 \pm 0$
+                $R^{5}_{T_4,K_2} = 0 \pm 0$
+                $R^{5}_{T_4,K_3} = 0 \pm 0$
+                $R^{5}_{T_4,K_4} = 0 \pm 0$
+                $R^{5}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        Thompson sampling &
+            $R^{6}_{T_1,K_1} = 0 \pm 0$
+                $R^{6}_{T_1,K_2} = 0 \pm 0$
+                $R^{6}_{T_1,K_3} = 0 \pm 0$
+                $R^{6}_{T_1,K_4} = 0 \pm 0$
+                $R^{6}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{6}_{T_2,K_1} = 0 \pm 0$
+                $R^{6}_{T_2,K_2} = 0 \pm 0$
+                $R^{6}_{T_2,K_3} = 0 \pm 0$
+                $R^{6}_{T_2,K_4} = 0 \pm 0$
+                $R^{6}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{6}_{T_3,K_1} = 0 \pm 0$
+                $R^{6}_{T_3,K_2} = 0 \pm 0$
+                $R^{6}_{T_3,K_3} = 0 \pm 0$
+                $R^{6}_{T_3,K_4} = 0 \pm 0$
+                $R^{6}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{6}_{T_4,K_1} = 0 \pm 0$
+                $R^{6}_{T_4,K_2} = 0 \pm 0$
+                $R^{6}_{T_4,K_3} = 0 \pm 0$
+                $R^{6}_{T_4,K_4} = 0 \pm 0$
+                $R^{6}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        Bayes-UCB &
+            $R^{7}_{T_1,K_1} = 0 \pm 0$
+                $R^{7}_{T_1,K_2} = 0 \pm 0$
+                $R^{7}_{T_1,K_3} = 0 \pm 0$
+                $R^{7}_{T_1,K_4} = 0 \pm 0$
+                $R^{7}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{7}_{T_2,K_1} = 0 \pm 0$
+                $R^{7}_{T_2,K_2} = 0 \pm 0$
+                $R^{7}_{T_2,K_3} = 0 \pm 0$
+                $R^{7}_{T_2,K_4} = 0 \pm 0$
+                $R^{7}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{7}_{T_3,K_1} = 0 \pm 0$
+                $R^{7}_{T_3,K_2} = 0 \pm 0$
+                $R^{7}_{T_3,K_3} = 0 \pm 0$
+                $R^{7}_{T_3,K_4} = 0 \pm 0$
+                $R^{7}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{7}_{T_4,K_1} = 0 \pm 0$
+                $R^{7}_{T_4,K_2} = 0 \pm 0$
+                $R^{7}_{T_4,K_3} = 0 \pm 0$
+                $R^{7}_{T_4,K_4} = 0 \pm 0$
+                $R^{7}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        AdBandits &
+            $R^{8}_{T_1,K_1} = 0 \pm 0$
+                $R^{8}_{T_1,K_2} = 0 \pm 0$
+                $R^{8}_{T_1,K_3} = 0 \pm 0$
+                $R^{8}_{T_1,K_4} = 0 \pm 0$
+                $R^{8}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{8}_{T_2,K_1} = 0 \pm 0$
+                $R^{8}_{T_2,K_2} = 0 \pm 0$
+                $R^{8}_{T_2,K_3} = 0 \pm 0$
+                $R^{8}_{T_2,K_4} = 0 \pm 0$
+                $R^{8}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{8}_{T_3,K_1} = 0 \pm 0$
+                $R^{8}_{T_3,K_2} = 0 \pm 0$
+                $R^{8}_{T_3,K_3} = 0 \pm 0$
+                $R^{8}_{T_3,K_4} = 0 \pm 0$
+                $R^{8}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{8}_{T_4,K_1} = 0 \pm 0$
+                $R^{8}_{T_4,K_2} = 0 \pm 0$
+                $R^{8}_{T_4,K_3} = 0 \pm 0$
+                $R^{8}_{T_4,K_4} = 0 \pm 0$
+                $R^{8}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        BESA &
+            $R^{9}_{T_1,K_1} = 0 \pm 0$
+                $R^{9}_{T_1,K_2} = 0 \pm 0$
+                $R^{9}_{T_1,K_3} = 0 \pm 0$
+                $R^{9}_{T_1,K_4} = 0 \pm 0$
+                $R^{9}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{9}_{T_2,K_1} = 0 \pm 0$
+                $R^{9}_{T_2,K_2} = 0 \pm 0$
+                $R^{9}_{T_2,K_3} = 0 \pm 0$
+                $R^{9}_{T_2,K_4} = 0 \pm 0$
+                $R^{9}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{9}_{T_3,K_1} = 0 \pm 0$
+                $R^{9}_{T_3,K_2} = 0 \pm 0$
+                $R^{9}_{T_3,K_3} = 0 \pm 0$
+                $R^{9}_{T_3,K_4} = 0 \pm 0$
+                $R^{9}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{9}_{T_4,K_1} = 0 \pm 0$
+                $R^{9}_{T_4,K_2} = 0 \pm 0$
+                $R^{9}_{T_4,K_3} = 0 \pm 0$
+                $R^{9}_{T_4,K_4} = 0 \pm 0$
+                $R^{9}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+    \end{tabular}
+    \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 2} with different values of time $T$ and number of arm $K$.
+    }
+    \label{table:3:meanRegret_problem2}
+\end{footnotesize}
+\end{table}
+
+% Include in appendix of this chapter the results for 7 more algorithms, the most recent and less important ones (MOSSAnytime, ApproximatedFHGittins, UCBdagger, klUCBswitchAnytime, PHE, RCB, Tsallis-Inf)
+
+\subsection{Observations and summary of the experiments}
+
+\TODOL{Use UCB and klUCB}
+\TODOL{Regret usually scales as $\bigO{K}$ and $\bigO{\log T}$}
+
 I want to show that SMPyBandits contains highly documented implementations of all the most important algorithms designed for stationary single-player multi-armed bandit.
@@
 \TODOL{Write text for this Section, write configuration.py file, run short-term experiment, include them, if happy about it, run longer ang larger experiments}
+
+In this section, we report additional experiment results from the simulations described in the previous section.
+Instead of studying on the efficiency of the algorithms (\ie, their regret), we report results of real measurements in terms of computational time as well as memory storage.
+%
+The goal of this section is to highlight that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in\{1,\dots,T\}$ bounded by $\bigO{K}$, independent of $t$.
+%
+While the results reported in the previous section should not depend on the implementation of the different algorithms, the results in this section concern real measurements of both time and memory consumptions of the simulation software used for these simulations.
+Hence, the reported results highly depend on many factors, including how the code is written, and where and when it is run.
+
+\begin{itemize}
+    \item Quality and optimization of the code:
+    The same level of optimization is used in all the codebase of SMPyBandits, and most of the time, the code is pure and naive Python and was not optimized.
+    Mathematical functions (\eg, $\sqrt{\bullet}, \log$) and random numbers used are using \texttt{numpy} and \texttt{numpy.random} modules \cite{numpy}, which are based on C and Cython code and can be considered highly efficient.
+    Computations of Kullback-Leibler and indexes for \klUCB{} algorithms are based on a compiled CPython extension written in C and can also be considered highly efficient.
+    As we can safely affirm that the code of the different algorithms has the same level of optimization, the comparison is fair, and we do not favor any algorithm in their implementation in SMPyBandits.
+
+    \item About the experimental environment:
+    the exact measurements displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
+    As long as the different algorithms and simulations are performed on the same environment, the comparison and conclusions we make from them are fair and do make sense.
+\end{itemize}
+
+But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
+Relative difference for costs do not depend much on the code and the machine used for the simulation.
+
+The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
+The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
+%
+Thus, we can check that efficient algorithms have a running time independent on the horizon $T$, instead of just observing a linear dependency, and we can check that efficient algorithms have a memory cost independent on the horizon.
+All the considered algorithms in this Chapter are efficient in this aspect, but in Chapter~\ref{chapter:6} we study algorithms that essentially have to store, and run computations, on an increasing number of observations (\eg, \CUSUMUCB), so their computation cost at time $t$ increases (linearly) with $t$, and thus they are more costly.
+%
+We can also check that efficient algorithms have a running time and a memory cost linear in the number of arms $K$, while more complex algorithms like BESA suffer from an exponential blow-up of their complexity when $K$ increases.
+
+\subsection{Computational time}
+
+\TODOL{Just one long table for problem 1, not necessary to include other problems}
+
+\begin{table}[!t]
+\begin{footnotesize}
+    \centering
+    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+        \hline
+        $\varepsilon$-greedy &
+            $\cT^{1}_{T_1,K_1} = 0$
+                $\cT^{1}_{T_1,K_2} = 0$
+                $\cT^{1}_{T_1,K_3} = 0$
+                $\cT^{1}_{T_1,K_4} = 0$
+                $\cT^{1}_{T_1,K_5} = 0$ &
+            $\cT^{1}_{T_2,K_1} = 0$
+                $\cT^{1}_{T_2,K_2} = 0$
+                $\cT^{1}_{T_2,K_3} = 0$
+                $\cT^{1}_{T_2,K_4} = 0$
+                $\cT^{1}_{T_2,K_5} = 0$ &
+            $\cT^{1}_{T_3,K_1} = 0$
+                $\cT^{1}_{T_3,K_2} = 0$
+                $\cT^{1}_{T_3,K_3} = 0$
+                $\cT^{1}_{T_3,K_4} = 0$
+                $\cT^{1}_{T_3,K_5} = 0$ &
+            $\cT^{1}_{T_4,K_1} = 0$
+                $\cT^{1}_{T_4,K_2} = 0$
+                $\cT^{1}_{T_4,K_3} = 0$
+                $\cT^{1}_{T_4,K_4} = 0$
+                $\cT^{1}_{T_4,K_5} = 0$ \\
+        \hline
+        Explore-then-Commit &
+            $\cT^{2}_{T_1,K_1} = 0$
+                $\cT^{2}_{T_1,K_2} = 0$
+                $\cT^{2}_{T_1,K_3} = 0$
+                $\cT^{2}_{T_1,K_4} = 0$
+                $\cT^{2}_{T_1,K_5} = 0$ &
+            $\cT^{2}_{T_2,K_1} = 0$
+                $\cT^{2}_{T_2,K_2} = 0$
+                $\cT^{2}_{T_2,K_3} = 0$
+                $\cT^{2}_{T_2,K_4} = 0$
+                $\cT^{2}_{T_2,K_5} = 0$ &
+            $\cT^{2}_{T_3,K_1} = 0$
+                $\cT^{2}_{T_3,K_2} = 0$
+                $\cT^{2}_{T_3,K_3} = 0$
+                $\cT^{2}_{T_3,K_4} = 0$
+                $\cT^{2}_{T_3,K_5} = 0$ &
+            $\cT^{2}_{T_4,K_1} = 0$
+                $\cT^{2}_{T_4,K_2} = 0$
+                $\cT^{2}_{T_4,K_3} = 0$
+                $\cT^{2}_{T_4,K_4} = 0$
+                $\cT^{2}_{T_4,K_5} = 0$ \\
+        \hline
+        $\mathrm{Exp}3^{++}$ &
+            $\cT^3_{T_1,K_1} = 0$
+                $\cT^3_{T_1,K_2} = 0$
+                $\cT^3_{T_1,K_3} = 0$
+                $\cT^3_{T_1,K_4} = 0$
+                $\cT^3_{T_1,K_5} = 0$ &
+            $\cT^3_{T_2,K_1} = 0$
+                $\cT^3_{T_2,K_2} = 0$
+                $\cT^3_{T_2,K_3} = 0$
+                $\cT^3_{T_2,K_4} = 0$
+                $\cT^3_{T_2,K_5} = 0$ &
+            $\cT^3_{T_3,K_1} = 0$
+                $\cT^3_{T_3,K_2} = 0$
+                $\cT^3_{T_3,K_3} = 0$
+                $\cT^3_{T_3,K_4} = 0$
+                $\cT^3_{T_3,K_5} = 0$ &
+            $\cT^3_{T_4,K_1} = 0$
+                $\cT^3_{T_4,K_2} = 0$
+                $\cT^3_{T_4,K_3} = 0$
+                $\cT^3_{T_4,K_4} = 0$
+                $\cT^3_{T_4,K_5} = 0$ \\
+        \hline
+        $\UCB_1$ &
+            $\cT^{4}_{T_1,K_1} = 0$
+                $\cT^{4}_{T_1,K_2} = 0$
+                $\cT^{4}_{T_1,K_3} = 0$
+                $\cT^{4}_{T_1,K_4} = 0$
+                $\cT^{4}_{T_1,K_5} = 0$ &
+            $\cT^{4}_{T_2,K_1} = 0$
+                $\cT^{4}_{T_2,K_2} = 0$
+                $\cT^{4}_{T_2,K_3} = 0$
+                $\cT^{4}_{T_2,K_4} = 0$
+                $\cT^{4}_{T_2,K_5} = 0$ &
+            $\cT^{4}_{T_3,K_1} = 0$
+                $\cT^{4}_{T_3,K_2} = 0$
+                $\cT^{4}_{T_3,K_3} = 0$
+                $\cT^{4}_{T_3,K_4} = 0$
+                $\cT^{4}_{T_3,K_5} = 0$ &
+            $\cT^{4}_{T_4,K_1} = 0$
+                $\cT^{4}_{T_4,K_2} = 0$
+                $\cT^{4}_{T_4,K_3} = 0$
+                $\cT^{4}_{T_4,K_4} = 0$
+                $\cT^{4}_{T_4,K_5} = 0$ \\
+        \hline
+        \klUCB{} &
+            $\cT^{5}_{T_1,K_1} = 0$
+                $\cT^{5}_{T_1,K_2} = 0$
+                $\cT^{5}_{T_1,K_3} = 0$
+                $\cT^{5}_{T_1,K_4} = 0$
+                $\cT^{5}_{T_1,K_5} = 0$ &
+            $\cT^{5}_{T_2,K_1} = 0$
+                $\cT^{5}_{T_2,K_2} = 0$
+                $\cT^{5}_{T_2,K_3} = 0$
+                $\cT^{5}_{T_2,K_4} = 0$
+                $\cT^{5}_{T_2,K_5} = 0$ &
+            $\cT^{5}_{T_3,K_1} = 0$
+                $\cT^{5}_{T_3,K_2} = 0$
+                $\cT^{5}_{T_3,K_3} = 0$
+                $\cT^{5}_{T_3,K_4} = 0$
+                $\cT^{5}_{T_3,K_5} = 0$ &
+            $\cT^{5}_{T_4,K_1} = 0$
+                $\cT^{5}_{T_4,K_2} = 0$
+                $\cT^{5}_{T_4,K_3} = 0$
+                $\cT^{5}_{T_4,K_4} = 0$
+                $\cT^{5}_{T_4,K_5} = 0$ \\
+        \hline
+        Thompson sampling &
+            $\cT^{6}_{T_1,K_1} = 0$
+                $\cT^{6}_{T_1,K_2} = 0$
+                $\cT^{6}_{T_1,K_3} = 0$
+                $\cT^{6}_{T_1,K_4} = 0$
+                $\cT^{6}_{T_1,K_5} = 0$ &
+            $\cT^{6}_{T_2,K_1} = 0$
+                $\cT^{6}_{T_2,K_2} = 0$
+                $\cT^{6}_{T_2,K_3} = 0$
+                $\cT^{6}_{T_2,K_4} = 0$
+                $\cT^{6}_{T_2,K_5} = 0$ &
+            $\cT^{6}_{T_3,K_1} = 0$
+                $\cT^{6}_{T_3,K_2} = 0$
+                $\cT^{6}_{T_3,K_3} = 0$
+                $\cT^{6}_{T_3,K_4} = 0$
+                $\cT^{6}_{T_3,K_5} = 0$ &
+            $\cT^{6}_{T_4,K_1} = 0$
+                $\cT^{6}_{T_4,K_2} = 0$
+                $\cT^{6}_{T_4,K_3} = 0$
+                $\cT^{6}_{T_4,K_4} = 0$
+                $\cT^{6}_{T_4,K_5} = 0$ \\
+        \hline
+        Thompson sampling &
+            $\cT^{7}_{T_1,K_1} = 0$
+                $\cT^{7}_{T_1,K_2} = 0$
+                $\cT^{7}_{T_1,K_3} = 0$
+                $\cT^{7}_{T_1,K_4} = 0$
+                $\cT^{7}_{T_1,K_5} = 0$ &
+            $\cT^{7}_{T_2,K_1} = 0$
+                $\cT^{7}_{T_2,K_2} = 0$
+                $\cT^{7}_{T_2,K_3} = 0$
+                $\cT^{7}_{T_2,K_4} = 0$
+                $\cT^{7}_{T_2,K_5} = 0$ &
+            $\cT^{7}_{T_3,K_1} = 0$
+                $\cT^{7}_{T_3,K_2} = 0$
+                $\cT^{7}_{T_3,K_3} = 0$
+                $\cT^{7}_{T_3,K_4} = 0$
+                $\cT^{7}_{T_3,K_5} = 0$ &
+            $\cT^{7}_{T_4,K_1} = 0$
+                $\cT^{7}_{T_4,K_2} = 0$
+                $\cT^{7}_{T_4,K_3} = 0$
+                $\cT^{7}_{T_4,K_4} = 0$
+                $\cT^{7}_{T_4,K_5} = 0$ \\
+        \hline
+        AdBandits &
+            $\cT^{8}_{T_1,K_1} = 0$
+                $\cT^{8}_{T_1,K_2} = 0$
+                $\cT^{8}_{T_1,K_3} = 0$
+                $\cT^{8}_{T_1,K_4} = 0$
+                $\cT^{8}_{T_1,K_5} = 0$ &
+            $\cT^{8}_{T_2,K_1} = 0$
+                $\cT^{8}_{T_2,K_2} = 0$
+                $\cT^{8}_{T_2,K_3} = 0$
+                $\cT^{8}_{T_2,K_4} = 0$
+                $\cT^{8}_{T_2,K_5} = 0$ &
+            $\cT^{8}_{T_3,K_1} = 0$
+                $\cT^{8}_{T_3,K_2} = 0$
+                $\cT^{8}_{T_3,K_3} = 0$
+                $\cT^{8}_{T_3,K_4} = 0$
+                $\cT^{8}_{T_3,K_5} = 0$ &
+            $\cT^{8}_{T_4,K_1} = 0$
+                $\cT^{8}_{T_4,K_2} = 0$
+                $\cT^{8}_{T_4,K_3} = 0$
+                $\cT^{8}_{T_4,K_4} = 0$
+                $\cT^{8}_{T_4,K_5} = 0$ \\
+        \hline
+        BESA &
+            $\cT^{9}_{T_1,K_1} = 0$
+                $\cT^{9}_{T_1,K_2} = 0$
+                $\cT^{9}_{T_1,K_3} = 0$
+                $\cT^{9}_{T_1,K_4} = 0$
+                $\cT^{9}_{T_1,K_5} = 0$ &
+            $\cT^{9}_{T_2,K_1} = 0$
+                $\cT^{9}_{T_2,K_2} = 0$
+                $\cT^{9}_{T_2,K_3} = 0$
+                $\cT^{9}_{T_2,K_4} = 0$
+                $\cT^{9}_{T_2,K_5} = 0$ &
+            $\cT^{9}_{T_3,K_1} = 0$
+                $\cT^{9}_{T_3,K_2} = 0$
+                $\cT^{9}_{T_3,K_3} = 0$
+                $\cT^{9}_{T_3,K_4} = 0$
+                $\cT^{9}_{T_3,K_5} = 0$ &
+            $\cT^{9}_{T_4,K_1} = 0$
+                $\cT^{9}_{T_4,K_2} = 0$
+                $\cT^{9}_{T_4,K_3} = 0$
+                $\cT^{9}_{T_4,K_4} = 0$
+                $\cT^{9}_{T_4,K_5} = 0$ \\
+        \hline
+    \end{tabular}
+    \caption{Normalized running time, rounded and in milli-seconds, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+    \label{table:3:time_problem1}
+\end{footnotesize}
+\end{table}
+
+\subsection{Memory cost}
+
+\TODOL{Just one long table for problem 1, not necessary to include other problems}
+
+\begin{table}[!t]
+\begin{footnotesize}
+    \centering
+    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+        \hline
+        $\varepsilon$-greedy &
+            $\cM^{1}_{T_1,K_1} = 0$
+                $\cM^{1}_{T_1,K_2} = 0$
+                $\cM^{1}_{T_1,K_3} = 0$
+                $\cM^{1}_{T_1,K_4} = 0$
+                $\cM^{1}_{T_1,K_5} = 0$ &
+            $\cM^{1}_{T_2,K_1} = 0$
+                $\cM^{1}_{T_2,K_2} = 0$
+                $\cM^{1}_{T_2,K_3} = 0$
+                $\cM^{1}_{T_2,K_4} = 0$
+                $\cM^{1}_{T_2,K_5} = 0$ &
+            $\cM^{1}_{T_3,K_1} = 0$
+                $\cM^{1}_{T_3,K_2} = 0$
+                $\cM^{1}_{T_3,K_3} = 0$
+                $\cM^{1}_{T_3,K_4} = 0$
+                $\cM^{1}_{T_3,K_5} = 0$ &
+            $\cM^{1}_{T_4,K_1} = 0$
+                $\cM^{1}_{T_4,K_2} = 0$
+                $\cM^{1}_{T_4,K_3} = 0$
+                $\cM^{1}_{T_4,K_4} = 0$
+                $\cM^{1}_{T_4,K_5} = 0$ \\
+        \hline
+        Explore-then-Commit &
+            $\cM^{2}_{T_1,K_1} = 0$
+                $\cM^{2}_{T_1,K_2} = 0$
+                $\cM^{2}_{T_1,K_3} = 0$
+                $\cM^{2}_{T_1,K_4} = 0$
+                $\cM^{2}_{T_1,K_5} = 0$ &
+            $\cM^{2}_{T_2,K_1} = 0$
+                $\cM^{2}_{T_2,K_2} = 0$
+                $\cM^{2}_{T_2,K_3} = 0$
+                $\cM^{2}_{T_2,K_4} = 0$
+                $\cM^{2}_{T_2,K_5} = 0$ &
+            $\cM^{2}_{T_3,K_1} = 0$
+                $\cM^{2}_{T_3,K_2} = 0$
+                $\cM^{2}_{T_3,K_3} = 0$
+                $\cM^{2}_{T_3,K_4} = 0$
+                $\cM^{2}_{T_3,K_5} = 0$ &
+            $\cM^{2}_{T_4,K_1} = 0$
+                $\cM^{2}_{T_4,K_2} = 0$
+                $\cM^{2}_{T_4,K_3} = 0$
+                $\cM^{2}_{T_4,K_4} = 0$
+                $\cM^{2}_{T_4,K_5} = 0$ \\
+        \hline
+        $\mathrm{Exp}3^{++}$ &
+            $\cM^3_{T_1,K_1} = 0$
+                $\cM^3_{T_1,K_2} = 0$
+                $\cM^3_{T_1,K_3} = 0$
+                $\cM^3_{T_1,K_4} = 0$
+                $\cM^3_{T_1,K_5} = 0$ &
+            $\cM^3_{T_2,K_1} = 0$
+                $\cM^3_{T_2,K_2} = 0$
+                $\cM^3_{T_2,K_3} = 0$
+                $\cM^3_{T_2,K_4} = 0$
+                $\cM^3_{T_2,K_5} = 0$ &
+            $\cM^3_{T_3,K_1} = 0$
+                $\cM^3_{T_3,K_2} = 0$
+                $\cM^3_{T_3,K_3} = 0$
+                $\cM^3_{T_3,K_4} = 0$
+                $\cM^3_{T_3,K_5} = 0$ &
+            $\cM^3_{T_4,K_1} = 0$
+                $\cM^3_{T_4,K_2} = 0$
+                $\cM^3_{T_4,K_3} = 0$
+                $\cM^3_{T_4,K_4} = 0$
+                $\cM^3_{T_4,K_5} = 0$ \\
+        \hline
+        $\UCB_1$ &
+            $\cM^{4}_{T_1,K_1} = 0$
+                $\cM^{4}_{T_1,K_2} = 0$
+                $\cM^{4}_{T_1,K_3} = 0$
+                $\cM^{4}_{T_1,K_4} = 0$
+                $\cM^{4}_{T_1,K_5} = 0$ &
+            $\cM^{4}_{T_2,K_1} = 0$
+                $\cM^{4}_{T_2,K_2} = 0$
+                $\cM^{4}_{T_2,K_3} = 0$
+                $\cM^{4}_{T_2,K_4} = 0$
+                $\cM^{4}_{T_2,K_5} = 0$ &
+            $\cM^{4}_{T_3,K_1} = 0$
+                $\cM^{4}_{T_3,K_2} = 0$
+                $\cM^{4}_{T_3,K_3} = 0$
+                $\cM^{4}_{T_3,K_4} = 0$
+                $\cM^{4}_{T_3,K_5} = 0$ &
+            $\cM^{4}_{T_4,K_1} = 0$
+                $\cM^{4}_{T_4,K_2} = 0$
+                $\cM^{4}_{T_4,K_3} = 0$
+                $\cM^{4}_{T_4,K_4} = 0$
+                $\cM^{4}_{T_4,K_5} = 0$ \\
+        \hline
+        \klUCB{} &
+            $\cM^{5}_{T_1,K_1} = 0$
+                $\cM^{5}_{T_1,K_2} = 0$
+                $\cM^{5}_{T_1,K_3} = 0$
+                $\cM^{5}_{T_1,K_4} = 0$
+                $\cM^{5}_{T_1,K_5} = 0$ &
+            $\cM^{5}_{T_2,K_1} = 0$
+                $\cM^{5}_{T_2,K_2} = 0$
+                $\cM^{5}_{T_2,K_3} = 0$
+                $\cM^{5}_{T_2,K_4} = 0$
+                $\cM^{5}_{T_2,K_5} = 0$ &
+            $\cM^{5}_{T_3,K_1} = 0$
+                $\cM^{5}_{T_3,K_2} = 0$
+                $\cM^{5}_{T_3,K_3} = 0$
+                $\cM^{5}_{T_3,K_4} = 0$
+                $\cM^{5}_{T_3,K_5} = 0$ &
+            $\cM^{5}_{T_4,K_1} = 0$
+                $\cM^{5}_{T_4,K_2} = 0$
+                $\cM^{5}_{T_4,K_3} = 0$
+                $\cM^{5}_{T_4,K_4} = 0$
+                $\cM^{5}_{T_4,K_5} = 0$ \\
+        \hline
+        Thompson sampling &
+            $\cM^{6}_{T_1,K_1} = 0$
+                $\cM^{6}_{T_1,K_2} = 0$
+                $\cM^{6}_{T_1,K_3} = 0$
+                $\cM^{6}_{T_1,K_4} = 0$
+                $\cM^{6}_{T_1,K_5} = 0$ &
+            $\cM^{6}_{T_2,K_1} = 0$
+                $\cM^{6}_{T_2,K_2} = 0$
+                $\cM^{6}_{T_2,K_3} = 0$
+                $\cM^{6}_{T_2,K_4} = 0$
+                $\cM^{6}_{T_2,K_5} = 0$ &
+            $\cM^{6}_{T_3,K_1} = 0$
+                $\cM^{6}_{T_3,K_2} = 0$
+                $\cM^{6}_{T_3,K_3} = 0$
+                $\cM^{6}_{T_3,K_4} = 0$
+                $\cM^{6}_{T_3,K_5} = 0$ &
+            $\cM^{6}_{T_4,K_1} = 0$
+                $\cM^{6}_{T_4,K_2} = 0$
+                $\cM^{6}_{T_4,K_3} = 0$
+                $\cM^{6}_{T_4,K_4} = 0$
+                $\cM^{6}_{T_4,K_5} = 0$ \\
+        \hline
+        Bayes-UCB &
+            $\cM^{7}_{T_1,K_1} = 0$
+                $\cM^{7}_{T_1,K_2} = 0$
+                $\cM^{7}_{T_1,K_3} = 0$
+                $\cM^{7}_{T_1,K_4} = 0$
+                $\cM^{7}_{T_1,K_5} = 0$ &
+            $\cM^{7}_{T_2,K_1} = 0$
+                $\cM^{7}_{T_2,K_2} = 0$
+                $\cM^{7}_{T_2,K_3} = 0$
+                $\cM^{7}_{T_2,K_4} = 0$
+                $\cM^{7}_{T_2,K_5} = 0$ &
+            $\cM^{7}_{T_3,K_1} = 0$
+                $\cM^{7}_{T_3,K_2} = 0$
+                $\cM^{7}_{T_3,K_3} = 0$
+                $\cM^{7}_{T_3,K_4} = 0$
+                $\cM^{7}_{T_3,K_5} = 0$ &
+            $\cM^{7}_{T_4,K_1} = 0$
+                $\cM^{7}_{T_4,K_2} = 0$
+                $\cM^{7}_{T_4,K_3} = 0$
+                $\cM^{7}_{T_4,K_4} = 0$
+                $\cM^{7}_{T_4,K_5} = 0$ \\
+        \hline
+        AdBandits &
+            $\cM^{8}_{T_1,K_1} = 0$
+                $\cM^{8}_{T_1,K_2} = 0$
+                $\cM^{8}_{T_1,K_3} = 0$
+                $\cM^{8}_{T_1,K_4} = 0$
+                $\cM^{8}_{T_1,K_5} = 0$ &
+            $\cM^{8}_{T_2,K_1} = 0$
+                $\cM^{8}_{T_2,K_2} = 0$
+                $\cM^{8}_{T_2,K_3} = 0$
+                $\cM^{8}_{T_2,K_4} = 0$
+                $\cM^{8}_{T_2,K_5} = 0$ &
+            $\cM^{8}_{T_3,K_1} = 0$
+                $\cM^{8}_{T_3,K_2} = 0$
+                $\cM^{8}_{T_3,K_3} = 0$
+                $\cM^{8}_{T_3,K_4} = 0$
+                $\cM^{8}_{T_3,K_5} = 0$ &
+            $\cM^{8}_{T_4,K_1} = 0$
+                $\cM^{8}_{T_4,K_2} = 0$
+                $\cM^{8}_{T_4,K_3} = 0$
+                $\cM^{8}_{T_4,K_4} = 0$
+                $\cM^{8}_{T_4,K_5} = 0$ \\
+        \hline
+        BESA &
+            $\cM^{9}_{T_1,K_1} = 0$
+                $\cM^{9}_{T_1,K_2} = 0$
+                $\cM^{9}_{T_1,K_3} = 0$
+                $\cM^{9}_{T_1,K_4} = 0$
+                $\cM^{9}_{T_1,K_5} = 0$ &
+            $\cM^{9}_{T_2,K_1} = 0$
+                $\cM^{9}_{T_2,K_2} = 0$
+                $\cM^{9}_{T_2,K_3} = 0$
+                $\cM^{9}_{T_2,K_4} = 0$
+                $\cM^{9}_{T_2,K_5} = 0$ &
+            $\cM^{9}_{T_3,K_1} = 0$
+                $\cM^{9}_{T_3,K_2} = 0$
+                $\cM^{9}_{T_3,K_3} = 0$
+                $\cM^{9}_{T_3,K_4} = 0$
+                $\cM^{9}_{T_3,K_5} = 0$ &
+            $\cM^{9}_{T_4,K_1} = 0$
+                $\cM^{9}_{T_4,K_2} = 0$
+                $\cM^{9}_{T_4,K_3} = 0$
+                $\cM^{9}_{T_4,K_4} = 0$
+                $\cM^{9}_{T_4,K_5} = 0$ \\
+        \hline
+    \end{tabular}
+    \caption{Non-normalized memory cost, rounded and in Bytes, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+    \label{table:3:memory_problem1}
+\end{footnotesize}
+\end{table}
+
+\subsection{Observations and summary of the experiments}
+
+\TODOL{Use UCB and klUCB}
+\TODOL{Regret usually scales as $\bigO{K}$ and $\bigO{\log T}$}
 
@@
 
-\subsection{Using parallelism in SMPyBandits}
+% --------------------------------------------------------------
+\subsection{Additional experimental results: more recent algorithms}
+\label{sub:3:additionalExperiments}
+
+This Appendix section describes seven more algorithms that are also compared with the nine presented above in Section~\ref{sec:3:reviewSPAlgorithms}.
+
+\TODOL{Si on a la place, j'aimerai bien inclure des résultats pour 7 algorithmes de plus, que voici :}
+
+\begin{itemize}
+    \item Algorithm 10 is
+    $\mathrm{MOSS}$-$\mathrm{Anytime}$ from \cite{Degenne16}, using $\alpha=1$ (\texttt{MOSSAnytime}).
+
+    \item Algorithm 11 is
+    the approximated Finite-Horizon Gittins from \cite{Lattimore16a}, using $\alpha=1$ (\texttt{ApproximatedFHGittins}).
+
+    \item Algorithm 12 is
+    $\UCB\dagger$ from \cite{Lattimore2018refining}, using $\alpha=1$ (\texttt{UCBdagger}).
+
+    \item Algorithm 13 is
+    the anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018}, using $\alpha=1$ (\texttt{klUCBswitchAnytime}).
+
+    \item Algorithm 14 is
+    \textsc{Tsallis-Inf} from \cite{Zimmert2018}, using $\alpha=0.5$ (it seems to be the most efficient of the algorithms described in the article) (\texttt{TsallisInf}).
+
+    \item Algorithm 15 is
+    $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019}, using $\alpha=1$ and permutations uniformly sampled in $[0,1]$ (it is the simplest of the algorithms described in the article) (\texttt{RCB}).
+
+    \item Algorithm 16 is
+    $\mathrm{PHE}$ (Perturbed-History Exploration) from \cite{KvetonSzepesvari2019}, using a perturbation scale of $0.5$ as advised (\texttt{PHE}).
+
+\end{itemize}
+\TODOL{FIXME order correctly in configuration.py}
+
+We give below the results for these additional algorithms, only for the problem 1,
+first in terms of regret in Table~\ref{table:3:meanRegret_problem1_otherAlgorithms},
+then in terms of computation times in Table~\ref{table:3:time_problem1_otherAlgorithms}
+and for storage costs in Table~\ref{table:3:memory_problem1_otherAlgorithms}.
+
+\paragraph{Mean regret}
+
+\begin{table}[!t]
+\begin{footnotesize}
+    \centering
+    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+        $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
+            $R^{10}_{T_1,K_1} = 0 \pm 0$
+                $R^{10}_{T_1,K_2} = 0 \pm 0$
+                $R^{10}_{T_1,K_3} = 0 \pm 0$
+                $R^{10}_{T_1,K_4} = 0 \pm 0$
+                $R^{10}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{10}_{T_2,K_1} = 0 \pm 0$
+                $R^{10}_{T_2,K_2} = 0 \pm 0$
+                $R^{10}_{T_2,K_3} = 0 \pm 0$
+                $R^{10}_{T_2,K_4} = 0 \pm 0$
+                $R^{10}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{10}_{T_3,K_1} = 0 \pm 0$
+                $R^{10}_{T_3,K_2} = 0 \pm 0$
+                $R^{10}_{T_3,K_3} = 0 \pm 0$
+                $R^{10}_{T_3,K_4} = 0 \pm 0$
+                $R^{10}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{10}_{T_4,K_1} = 0 \pm 0$
+                $R^{10}_{T_4,K_2} = 0 \pm 0$
+                $R^{10}_{T_4,K_3} = 0 \pm 0$
+                $R^{10}_{T_4,K_4} = 0 \pm 0$
+                $R^{10}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        Appr. F-H Gittins &
+            $R^{11}_{T_1,K_1} = 0 \pm 0$
+                $R^{11}_{T_1,K_2} = 0 \pm 0$
+                $R^{11}_{T_1,K_3} = 0 \pm 0$
+                $R^{11}_{T_1,K_4} = 0 \pm 0$
+                $R^{11}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{11}_{T_2,K_1} = 0 \pm 0$
+                $R^{11}_{T_2,K_2} = 0 \pm 0$
+                $R^{11}_{T_2,K_3} = 0 \pm 0$
+                $R^{11}_{T_2,K_4} = 0 \pm 0$
+                $R^{11}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{11}_{T_3,K_1} = 0 \pm 0$
+                $R^{11}_{T_3,K_2} = 0 \pm 0$
+                $R^{11}_{T_3,K_3} = 0 \pm 0$
+                $R^{11}_{T_3,K_4} = 0 \pm 0$
+                $R^{11}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{11}_{T_4,K_1} = 0 \pm 0$
+                $R^{11}_{T_4,K_2} = 0 \pm 0$
+                $R^{11}_{T_4,K_3} = 0 \pm 0$
+                $R^{11}_{T_4,K_4} = 0 \pm 0$
+                $R^{11}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        $\UCB\dagger$ &
+            $R^{12}_{T_1,K_1} = 0 \pm 0$
+                $R^{12}_{T_1,K_2} = 0 \pm 0$
+                $R^{12}_{T_1,K_3} = 0 \pm 0$
+                $R^{12}_{T_1,K_4} = 0 \pm 0$
+                $R^{12}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{12}_{T_2,K_1} = 0 \pm 0$
+                $R^{12}_{T_2,K_2} = 0 \pm 0$
+                $R^{12}_{T_2,K_3} = 0 \pm 0$
+                $R^{12}_{T_2,K_4} = 0 \pm 0$
+                $R^{12}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{12}_{T_3,K_1} = 0 \pm 0$
+                $R^{12}_{T_3,K_2} = 0 \pm 0$
+                $R^{12}_{T_3,K_3} = 0 \pm 0$
+                $R^{12}_{T_3,K_4} = 0 \pm 0$
+                $R^{12}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{12}_{T_4,K_1} = 0 \pm 0$
+                $R^{12}_{T_4,K_2} = 0 \pm 0$
+                $R^{12}_{T_4,K_3} = 0 \pm 0$
+                $R^{12}_{T_4,K_4} = 0 \pm 0$
+                $R^{12}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        KL-UCB-Switch &
+            $R^{13}_{T_1,K_1} = 0 \pm 0$
+                $R^{13}_{T_1,K_2} = 0 \pm 0$
+                $R^{13}_{T_1,K_3} = 0 \pm 0$
+                $R^{13}_{T_1,K_4} = 0 \pm 0$
+                $R^{13}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{13}_{T_2,K_1} = 0 \pm 0$
+                $R^{13}_{T_2,K_2} = 0 \pm 0$
+                $R^{13}_{T_2,K_3} = 0 \pm 0$
+                $R^{13}_{T_2,K_4} = 0 \pm 0$
+                $R^{13}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{13}_{T_3,K_1} = 0 \pm 0$
+                $R^{13}_{T_3,K_2} = 0 \pm 0$
+                $R^{13}_{T_3,K_3} = 0 \pm 0$
+                $R^{13}_{T_3,K_4} = 0 \pm 0$
+                $R^{13}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{13}_{T_4,K_1} = 0 \pm 0$
+                $R^{13}_{T_4,K_2} = 0 \pm 0$
+                $R^{13}_{T_4,K_3} = 0 \pm 0$
+                $R^{13}_{T_4,K_4} = 0 \pm 0$
+                $R^{13}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        \textsc{Tsallis-Inf} &
+            $R^{14}_{T_1,K_1} = 0 \pm 0$
+                $R^{14}_{T_1,K_2} = 0 \pm 0$
+                $R^{14}_{T_1,K_3} = 0 \pm 0$
+                $R^{14}_{T_1,K_4} = 0 \pm 0$
+                $R^{14}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{14}_{T_2,K_1} = 0 \pm 0$
+                $R^{14}_{T_2,K_2} = 0 \pm 0$
+                $R^{14}_{T_2,K_3} = 0 \pm 0$
+                $R^{14}_{T_2,K_4} = 0 \pm 0$
+                $R^{14}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{14}_{T_3,K_1} = 0 \pm 0$
+                $R^{14}_{T_3,K_2} = 0 \pm 0$
+                $R^{14}_{T_3,K_3} = 0 \pm 0$
+                $R^{14}_{T_3,K_4} = 0 \pm 0$
+                $R^{14}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{14}_{T_4,K_1} = 0 \pm 0$
+                $R^{14}_{T_4,K_2} = 0 \pm 0$
+                $R^{14}_{T_4,K_3} = 0 \pm 0$
+                $R^{14}_{T_4,K_4} = 0 \pm 0$
+                $R^{14}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        $\mathrm{RCB}$ &
+            $R^{15}_{T_1,K_1} = 0 \pm 0$
+                $R^{15}_{T_1,K_2} = 0 \pm 0$
+                $R^{15}_{T_1,K_3} = 0 \pm 0$
+                $R^{15}_{T_1,K_4} = 0 \pm 0$
+                $R^{15}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{15}_{T_2,K_1} = 0 \pm 0$
+                $R^{15}_{T_2,K_2} = 0 \pm 0$
+                $R^{15}_{T_2,K_3} = 0 \pm 0$
+                $R^{15}_{T_2,K_4} = 0 \pm 0$
+                $R^{15}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{15}_{T_3,K_1} = 0 \pm 0$
+                $R^{15}_{T_3,K_2} = 0 \pm 0$
+                $R^{15}_{T_3,K_3} = 0 \pm 0$
+                $R^{15}_{T_3,K_4} = 0 \pm 0$
+                $R^{15}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{15}_{T_4,K_1} = 0 \pm 0$
+                $R^{15}_{T_4,K_2} = 0 \pm 0$
+                $R^{15}_{T_4,K_3} = 0 \pm 0$
+                $R^{15}_{T_4,K_4} = 0 \pm 0$
+                $R^{15}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+        $\mathrm{PHE}$ &
+            $R^{16}_{T_1,K_1} = 0 \pm 0$
+                $R^{16}_{T_1,K_2} = 0 \pm 0$
+                $R^{16}_{T_1,K_3} = 0 \pm 0$
+                $R^{16}_{T_1,K_4} = 0 \pm 0$
+                $R^{16}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{16}_{T_2,K_1} = 0 \pm 0$
+                $R^{16}_{T_2,K_2} = 0 \pm 0$
+                $R^{16}_{T_2,K_3} = 0 \pm 0$
+                $R^{16}_{T_2,K_4} = 0 \pm 0$
+                $R^{16}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{16}_{T_3,K_1} = 0 \pm 0$
+                $R^{16}_{T_3,K_2} = 0 \pm 0$
+                $R^{16}_{T_3,K_3} = 0 \pm 0$
+                $R^{16}_{T_3,K_4} = 0 \pm 0$
+                $R^{16}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{16}_{T_4,K_1} = 0 \pm 0$
+                $R^{16}_{T_4,K_2} = 0 \pm 0$
+                $R^{16}_{T_4,K_3} = 0 \pm 0$
+                $R^{16}_{T_4,K_4} = 0 \pm 0$
+                $R^{16}_{T_4,K_5} = 0 \pm 0$ \\
+        \hline
+    \end{tabular}
+    \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on problem $1$ with different values of time $T$ and number of arm $K$.
+    }
+    \label{table:3:meanRegret_problem1_otherAlgorithms}
+\end{footnotesize}
+\end{table}
+
+\paragraph{Computation time}
+
+\TODOL{A faire}
+
+\begin{table}[!t]
+\begin{footnotesize}
+    \centering
+    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+        $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
+            $\cT^{10}_{T_1,K_1} = 0$
+                $\cT^{10}_{T_1,K_2} = 0$
+                $\cT^{10}_{T_1,K_3} = 0$
+                $\cT^{10}_{T_1,K_4} = 0$
+                $\cT^{10}_{T_1,K_5} = 0$ &
+            $\cT^{10}_{T_2,K_1} = 0$
+                $\cT^{10}_{T_2,K_2} = 0$
+                $\cT^{10}_{T_2,K_3} = 0$
+                $\cT^{10}_{T_2,K_4} = 0$
+                $\cT^{10}_{T_2,K_5} = 0$ &
+            $\cT^{10}_{T_3,K_1} = 0$
+                $\cT^{10}_{T_3,K_2} = 0$
+                $\cT^{10}_{T_3,K_3} = 0$
+                $\cT^{10}_{T_3,K_4} = 0$
+                $\cT^{10}_{T_3,K_5} = 0$ &
+            $\cT^{10}_{T_4,K_1} = 0$
+                $\cT^{10}_{T_4,K_2} = 0$
+                $\cT^{10}_{T_4,K_3} = 0$
+                $\cT^{10}_{T_4,K_4} = 0$
+                $\cT^{10}_{T_4,K_5} = 0$ \\
+        \hline
+        Appr. F-H Gittins &
+            $\cT^{11}_{T_1,K_1} = 0$
+                $\cT^{11}_{T_1,K_2} = 0$
+                $\cT^{11}_{T_1,K_3} = 0$
+                $\cT^{11}_{T_1,K_4} = 0$
+                $\cT^{11}_{T_1,K_5} = 0$ &
+            $\cT^{11}_{T_2,K_1} = 0$
+                $\cT^{11}_{T_2,K_2} = 0$
+                $\cT^{11}_{T_2,K_3} = 0$
+                $\cT^{11}_{T_2,K_4} = 0$
+                $\cT^{11}_{T_2,K_5} = 0$ &
+            $\cT^{11}_{T_3,K_1} = 0$
+                $\cT^{11}_{T_3,K_2} = 0$
+                $\cT^{11}_{T_3,K_3} = 0$
+                $\cT^{11}_{T_3,K_4} = 0$
+                $\cT^{11}_{T_3,K_5} = 0$ &
+            $\cT^{11}_{T_4,K_1} = 0$
+                $\cT^{11}_{T_4,K_2} = 0$
+                $\cT^{11}_{T_4,K_3} = 0$
+                $\cT^{11}_{T_4,K_4} = 0$
+                $\cT^{11}_{T_4,K_5} = 0$ \\
+        \hline
+        $\UCB\dagger$ &
+            $\cT^{12}_{T_1,K_1} = 0$
+                $\cT^{12}_{T_1,K_2} = 0$
+                $\cT^{12}_{T_1,K_3} = 0$
+                $\cT^{12}_{T_1,K_4} = 0$
+                $\cT^{12}_{T_1,K_5} = 0$ &
+            $\cT^{12}_{T_2,K_1} = 0$
+                $\cT^{12}_{T_2,K_2} = 0$
+                $\cT^{12}_{T_2,K_3} = 0$
+                $\cT^{12}_{T_2,K_4} = 0$
+                $\cT^{12}_{T_2,K_5} = 0$ &
+            $\cT^{12}_{T_3,K_1} = 0$
+                $\cT^{12}_{T_3,K_2} = 0$
+                $\cT^{12}_{T_3,K_3} = 0$
+                $\cT^{12}_{T_3,K_4} = 0$
+                $\cT^{12}_{T_3,K_5} = 0$ &
+            $\cT^{12}_{T_4,K_1} = 0$
+                $\cT^{12}_{T_4,K_2} = 0$
+                $\cT^{12}_{T_4,K_3} = 0$
+                $\cT^{12}_{T_4,K_4} = 0$
+                $\cT^{12}_{T_4,K_5} = 0$ \\
+        \hline
+        KL-UCB-Switch &
+            $\cT^{13}_{T_1,K_1} = 0$
+                $\cT^{13}_{T_1,K_2} = 0$
+                $\cT^{13}_{T_1,K_3} = 0$
+                $\cT^{13}_{T_1,K_4} = 0$
+                $\cT^{13}_{T_1,K_5} = 0$ &
+            $\cT^{13}_{T_2,K_1} = 0$
+                $\cT^{13}_{T_2,K_2} = 0$
+                $\cT^{13}_{T_2,K_3} = 0$
+                $\cT^{13}_{T_2,K_4} = 0$
+                $\cT^{13}_{T_2,K_5} = 0$ &
+            $\cT^{13}_{T_3,K_1} = 0$
+                $\cT^{13}_{T_3,K_2} = 0$
+                $\cT^{13}_{T_3,K_3} = 0$
+                $\cT^{13}_{T_3,K_4} = 0$
+                $\cT^{13}_{T_3,K_5} = 0$ &
+            $\cT^{13}_{T_4,K_1} = 0$
+                $\cT^{13}_{T_4,K_2} = 0$
+                $\cT^{13}_{T_4,K_3} = 0$
+                $\cT^{13}_{T_4,K_4} = 0$
+                $\cT^{13}_{T_4,K_5} = 0$ \\
+        \hline
+        \textsc{Tsallis-Inf} &
+            $\cT^{14}_{T_1,K_1} = 0$
+                $\cT^{14}_{T_1,K_2} = 0$
+                $\cT^{14}_{T_1,K_3} = 0$
+                $\cT^{14}_{T_1,K_4} = 0$
+                $\cT^{14}_{T_1,K_5} = 0$ &
+            $\cT^{14}_{T_2,K_1} = 0$
+                $\cT^{14}_{T_2,K_2} = 0$
+                $\cT^{14}_{T_2,K_3} = 0$
+                $\cT^{14}_{T_2,K_4} = 0$
+                $\cT^{14}_{T_2,K_5} = 0$ &
+            $\cT^{14}_{T_3,K_1} = 0$
+                $\cT^{14}_{T_3,K_2} = 0$
+                $\cT^{14}_{T_3,K_3} = 0$
+                $\cT^{14}_{T_3,K_4} = 0$
+                $\cT^{14}_{T_3,K_5} = 0$ &
+            $\cT^{14}_{T_4,K_1} = 0$
+                $\cT^{14}_{T_4,K_2} = 0$
+                $\cT^{14}_{T_4,K_3} = 0$
+                $\cT^{14}_{T_4,K_4} = 0$
+                $\cT^{14}_{T_4,K_5} = 0$ \\
+        \hline
+        $\mathrm{RCB}$ &
+            $\cT^{15}_{T_1,K_1} = 0$
+                $\cT^{15}_{T_1,K_2} = 0$
+                $\cT^{15}_{T_1,K_3} = 0$
+                $\cT^{15}_{T_1,K_4} = 0$
+                $\cT^{15}_{T_1,K_5} = 0$ &
+            $\cT^{15}_{T_2,K_1} = 0$
+                $\cT^{15}_{T_2,K_2} = 0$
+                $\cT^{15}_{T_2,K_3} = 0$
+                $\cT^{15}_{T_2,K_4} = 0$
+                $\cT^{15}_{T_2,K_5} = 0$ &
+            $\cT^{15}_{T_3,K_1} = 0$
+                $\cT^{15}_{T_3,K_2} = 0$
+                $\cT^{15}_{T_3,K_3} = 0$
+                $\cT^{15}_{T_3,K_4} = 0$
+                $\cT^{15}_{T_3,K_5} = 0$ &
+            $\cT^{15}_{T_4,K_1} = 0$
+                $\cT^{15}_{T_4,K_2} = 0$
+                $\cT^{15}_{T_4,K_3} = 0$
+                $\cT^{15}_{T_4,K_4} = 0$
+                $\cT^{15}_{T_4,K_5} = 0$ \\
+        \hline
+        $\mathrm{PHE}$ &
+            $\cT^{16}_{T_1,K_1} = 0$
+                $\cT^{16}_{T_1,K_2} = 0$
+                $\cT^{16}_{T_1,K_3} = 0$
+                $\cT^{16}_{T_1,K_4} = 0$
+                $\cT^{16}_{T_1,K_5} = 0$ &
+            $\cT^{16}_{T_2,K_1} = 0$
+                $\cT^{16}_{T_2,K_2} = 0$
+                $\cT^{16}_{T_2,K_3} = 0$
+                $\cT^{16}_{T_2,K_4} = 0$
+                $\cT^{16}_{T_2,K_5} = 0$ &
+            $\cT^{16}_{T_3,K_1} = 0$
+                $\cT^{16}_{T_3,K_2} = 0$
+                $\cT^{16}_{T_3,K_3} = 0$
+                $\cT^{16}_{T_3,K_4} = 0$
+                $\cT^{16}_{T_3,K_5} = 0$ &
+            $\cT^{16}_{T_4,K_1} = 0$
+                $\cT^{16}_{T_4,K_2} = 0$
+                $\cT^{16}_{T_4,K_3} = 0$
+                $\cT^{16}_{T_4,K_4} = 0$
+                $\cT^{16}_{T_4,K_5} = 0$ \\
+        \hline
+    \end{tabular}
+    \caption{Normalized running time, rounded and in milli-seconds, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+    \label{table:3:time_problem1_otherAlgorithms}
+\end{footnotesize}
+\end{table}
+
+\paragraph{Storage cost}
+
+\TODOL{A faire}
+
+\begin{table}[!t]
+\begin{footnotesize}
+    \centering
+    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+        $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
+            $\cM^{10}_{T_1,K_1} = 0$
+                $\cM^{10}_{T_1,K_2} = 0$
+                $\cM^{10}_{T_1,K_3} = 0$
+                $\cM^{10}_{T_1,K_4} = 0$
+                $\cM^{10}_{T_1,K_5} = 0$ &
+            $\cM^{10}_{T_2,K_1} = 0$
+                $\cM^{10}_{T_2,K_2} = 0$
+                $\cM^{10}_{T_2,K_3} = 0$
+                $\cM^{10}_{T_2,K_4} = 0$
+                $\cM^{10}_{T_2,K_5} = 0$ &
+            $\cM^{10}_{T_3,K_1} = 0$
+                $\cM^{10}_{T_3,K_2} = 0$
+                $\cM^{10}_{T_3,K_3} = 0$
+                $\cM^{10}_{T_3,K_4} = 0$
+                $\cM^{10}_{T_3,K_5} = 0$ &
+            $\cM^{10}_{T_4,K_1} = 0$
+                $\cM^{10}_{T_4,K_2} = 0$
+                $\cM^{10}_{T_4,K_3} = 0$
+                $\cM^{10}_{T_4,K_4} = 0$
+                $\cM^{10}_{T_4,K_5} = 0$ \\
+        \hline
+        Appr. F-H Gittins &
+            $\cM^{11}_{T_1,K_1} = 0$
+                $\cM^{11}_{T_1,K_2} = 0$
+                $\cM^{11}_{T_1,K_3} = 0$
+                $\cM^{11}_{T_1,K_4} = 0$
+                $\cM^{11}_{T_1,K_5} = 0$ &
+            $\cM^{11}_{T_2,K_1} = 0$
+                $\cM^{11}_{T_2,K_2} = 0$
+                $\cM^{11}_{T_2,K_3} = 0$
+                $\cM^{11}_{T_2,K_4} = 0$
+                $\cM^{11}_{T_2,K_5} = 0$ &
+            $\cM^{11}_{T_3,K_1} = 0$
+                $\cM^{11}_{T_3,K_2} = 0$
+                $\cM^{11}_{T_3,K_3} = 0$
+                $\cM^{11}_{T_3,K_4} = 0$
+                $\cM^{11}_{T_3,K_5} = 0$ &
+            $\cM^{11}_{T_4,K_1} = 0$
+                $\cM^{11}_{T_4,K_2} = 0$
+                $\cM^{11}_{T_4,K_3} = 0$
+                $\cM^{11}_{T_4,K_4} = 0$
+                $\cM^{11}_{T_4,K_5} = 0$ \\
+        \hline
+        $\UCB\dagger$ &
+            $\cM^{12}_{T_1,K_1} = 0$
+                $\cM^{12}_{T_1,K_2} = 0$
+                $\cM^{12}_{T_1,K_3} = 0$
+                $\cM^{12}_{T_1,K_4} = 0$
+                $\cM^{12}_{T_1,K_5} = 0$ &
+            $\cM^{12}_{T_2,K_1} = 0$
+                $\cM^{12}_{T_2,K_2} = 0$
+                $\cM^{12}_{T_2,K_3} = 0$
+                $\cM^{12}_{T_2,K_4} = 0$
+                $\cM^{12}_{T_2,K_5} = 0$ &
+            $\cM^{12}_{T_3,K_1} = 0$
+                $\cM^{12}_{T_3,K_2} = 0$
+                $\cM^{12}_{T_3,K_3} = 0$
+                $\cM^{12}_{T_3,K_4} = 0$
+                $\cM^{12}_{T_3,K_5} = 0$ &
+            $\cM^{12}_{T_4,K_1} = 0$
+                $\cM^{12}_{T_4,K_2} = 0$
+                $\cM^{12}_{T_4,K_3} = 0$
+                $\cM^{12}_{T_4,K_4} = 0$
+                $\cM^{12}_{T_4,K_5} = 0$ \\
+        \hline
+        KL-UCB-Switch &
+            $\cM^{13}_{T_1,K_1} = 0$
+                $\cM^{13}_{T_1,K_2} = 0$
+                $\cM^{13}_{T_1,K_3} = 0$
+                $\cM^{13}_{T_1,K_4} = 0$
+                $\cM^{13}_{T_1,K_5} = 0$ &
+            $\cM^{13}_{T_2,K_1} = 0$
+                $\cM^{13}_{T_2,K_2} = 0$
+                $\cM^{13}_{T_2,K_3} = 0$
+                $\cM^{13}_{T_2,K_4} = 0$
+                $\cM^{13}_{T_2,K_5} = 0$ &
+            $\cM^{13}_{T_3,K_1} = 0$
+                $\cM^{13}_{T_3,K_2} = 0$
+                $\cM^{13}_{T_3,K_3} = 0$
+                $\cM^{13}_{T_3,K_4} = 0$
+                $\cM^{13}_{T_3,K_5} = 0$ &
+            $\cM^{13}_{T_4,K_1} = 0$
+                $\cM^{13}_{T_4,K_2} = 0$
+                $\cM^{13}_{T_4,K_3} = 0$
+                $\cM^{13}_{T_4,K_4} = 0$
+                $\cM^{13}_{T_4,K_5} = 0$ \\
+        \hline
+        \textsc{Tsallis-Inf} &
+            $\cM^{14}_{T_1,K_1} = 0$
+                $\cM^{14}_{T_1,K_2} = 0$
+                $\cM^{14}_{T_1,K_3} = 0$
+                $\cM^{14}_{T_1,K_4} = 0$
+                $\cM^{14}_{T_1,K_5} = 0$ &
+            $\cM^{14}_{T_2,K_1} = 0$
+                $\cM^{14}_{T_2,K_2} = 0$
+                $\cM^{14}_{T_2,K_3} = 0$
+                $\cM^{14}_{T_2,K_4} = 0$
+                $\cM^{14}_{T_2,K_5} = 0$ &
+            $\cM^{14}_{T_3,K_1} = 0$
+                $\cM^{14}_{T_3,K_2} = 0$
+                $\cM^{14}_{T_3,K_3} = 0$
+                $\cM^{14}_{T_3,K_4} = 0$
+                $\cM^{14}_{T_3,K_5} = 0$ &
+            $\cM^{14}_{T_4,K_1} = 0$
+                $\cM^{14}_{T_4,K_2} = 0$
+                $\cM^{14}_{T_4,K_3} = 0$
+                $\cM^{14}_{T_4,K_4} = 0$
+                $\cM^{14}_{T_4,K_5} = 0$ \\
+        \hline
+        $\mathrm{RCB}$ &
+            $\cM^{15}_{T_1,K_1} = 0$
+                $\cM^{15}_{T_1,K_2} = 0$
+                $\cM^{15}_{T_1,K_3} = 0$
+                $\cM^{15}_{T_1,K_4} = 0$
+                $\cM^{15}_{T_1,K_5} = 0$ &
+            $\cM^{15}_{T_2,K_1} = 0$
+                $\cM^{15}_{T_2,K_2} = 0$
+                $\cM^{15}_{T_2,K_3} = 0$
+                $\cM^{15}_{T_2,K_4} = 0$
+                $\cM^{15}_{T_2,K_5} = 0$ &
+            $\cM^{15}_{T_3,K_1} = 0$
+                $\cM^{15}_{T_3,K_2} = 0$
+                $\cM^{15}_{T_3,K_3} = 0$
+                $\cM^{15}_{T_3,K_4} = 0$
+                $\cM^{15}_{T_3,K_5} = 0$ &
+            $\cM^{15}_{T_4,K_1} = 0$
+                $\cM^{15}_{T_4,K_2} = 0$
+                $\cM^{15}_{T_4,K_3} = 0$
+                $\cM^{15}_{T_4,K_4} = 0$
+                $\cM^{15}_{T_4,K_5} = 0$ \\
+        \hline
+        $\mathrm{RCB}$ &
+            $\cM^{16}_{T_1,K_1} = 0$
+                $\cM^{16}_{T_1,K_2} = 0$
+                $\cM^{16}_{T_1,K_3} = 0$
+                $\cM^{16}_{T_1,K_4} = 0$
+                $\cM^{16}_{T_1,K_5} = 0$ &
+            $\cM^{16}_{T_2,K_1} = 0$
+                $\cM^{16}_{T_2,K_2} = 0$
+                $\cM^{16}_{T_2,K_3} = 0$
+                $\cM^{16}_{T_2,K_4} = 0$
+                $\cM^{16}_{T_2,K_5} = 0$ &
+            $\cM^{16}_{T_3,K_1} = 0$
+                $\cM^{16}_{T_3,K_2} = 0$
+                $\cM^{16}_{T_3,K_3} = 0$
+                $\cM^{16}_{T_3,K_4} = 0$
+                $\cM^{16}_{T_3,K_5} = 0$ &
+            $\cM^{16}_{T_4,K_1} = 0$
+                $\cM^{16}_{T_4,K_2} = 0$
+                $\cM^{16}_{T_4,K_3} = 0$
+                $\cM^{16}_{T_4,K_4} = 0$
+                $\cM^{16}_{T_4,K_5} = 0$ \\
+        \hline
+    \end{tabular}
+    \caption{Non-normalized memory cost, rounded and in Bytes, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+    \label{table:3:memory_problem1_otherAlgorithms}
+\end{footnotesize}
+\end{table}
+
+\paragraph{Interpretation}
+
+\TODOL{A faire}
+
+\paragraph{Bash command to run these experiments}
+
+Finally, we give below the Bash command that was used to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
+It uses \texttt{for} loops from Bash and not Python and environment variables, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $N$, $T$ and $K$.
+
+% https://tex.stackexchange.com/a/12430/
+\begin{listing}[h!]
+    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
+$ for N in 4 100 1000; do \
+  for BAYES in False True; do\
+  for T in 1000 5000 10000 50000; do \
+  for K in 2 5 10 20 50; do \
+    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
+             N=$N BAYES=$BAYES T=$T K=$K make single \
+    && cp logs/main_py3_log.txt logs/main_py3_log__N${N}_BAYES${BAYES}_T{T}_K{K}.txt \
+  done; done; done; done
+    \end{minted}
+    \caption{Snippet of Bash code to run the large experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
+    % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:BashCodeToLaunchLargeExperiments}.}
+    \label{lst:3:BashCodeToLaunchLargeExperiments}
+\end{listing}
+
+% --------------------------------------------------------------
+\subsection{Additional remark: about parallelism in SMPyBandits}
 \label{sub:3:parallelSimulations}

### 2019-05-02 2-Chapters/3-Chapter/chapter3.tex
Fixed typos

@@
 In this section, we use the SMPyBandits library to compare experimentally various state-of-the-art single-player algorithms on different kinds of multi-armed bandit problems.
-First, we detail the list of $9$ different algorithms that we compare, and the different problems on which they are all tested.
+First, we detail the list of different algorithms that we compare, and the different problems on which they are all tested.
 We give the results of the numerical experiments, in terms of mean regret at the end of the experiment of different horizons $T$.
@@
 
-\TODOL{FIXME launch these experiments, for a very short number of repetitions N=4, and if I'm happy about the results and if they are easily copied-and-pasted here, then launch with N=100 or N=1000 depending on the running
-
-% DEBUG=True; NOPLOTS=False; SAVEALL=False; for N in 4 100 1000; do for BAYES in False True; do for T in 1000 5000 10000 50000; do for K in 2 5 10 20 50; do debut=$(date); DEBUG=$DEBUG NOPLOTS=$NOPLOTS SAVEALL=$SAVEALL N=$N N_JOBS=-1 T=$T K=$K BAYES=$BAYES make --noFreeSMS single && cp logs/main_py3_log.txt logs/main_py3_log__N${N}_BAYES${BAYES}_T{T}_K{K}.txt && FreeSMS.py "Expérience commencée ${debut} terminée, avec N=${N}, BAYES=${BAYES}, T=${T}, K=${K}." ; done; done; done; done time.}
+\TODOL{FIXME launch these experiments, for a very short number of repetitions N=4, and if I'm happy about the results and if they are easily copied-and-pasted here, then launch with N=100 or N=1000 depending on the running time of the whole simulations}
+
+% DEBUG=True; NOPLOTS=False; SAVEALL=False; for N in 4 100 1000; do for BAYES in False True; do for T in 1000 5000 10000 50000; do for K in 2 5 10 20 50; do debut=$(date); DEBUG=$DEBUG NOPLOTS=$NOPLOTS SAVEALL=$SAVEALL N=$N N_JOBS=-1 T=$T K=$K BAYES=$BAYES make --noFreeSMS single && cp logs/main_py3_log.txt logs/main_py3_log__N${N}_BAYES${BAYES}_T{T}_K{K}.txt && FreeSMS.py "Expérience commencée ${debut} terminée, avec N=${N}, BAYES=${BAYES}, T=${T}, K=${K}." ; done; done; done; done time.
 
 \begin{table}[!t]
-\begin{footnotesize}
+\begin{footnotesize}  % WARNING
     \centering
@@
     \label{table:3:meanRegret_problem1}
-\end{footnotesize}
+\end{footnotesize}  % WARNING
 \end{table}
@@
 \begin{table}[!t]
-\begin{footnotesize}
+\begin{footnotesize}  % WARNING
     \centering
@@
     \label{table:3:meanRegret_problem2}
-\end{footnotesize}
+\end{footnotesize}  % WARNING
 \end{table}
@@
     \label{table:3:time_problem1}
-\end{footnotesize}
+\end{footnotesize}  % WARNING
 \end{table}
@@
     \label{table:3:memory_problem1}
-\end{footnotesize}
+\end{footnotesize}  % WARNING
 \end{table}
@@
 \label{sec:3:appendix}
+
+In this Appendix, we start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
+We conclude by a discussion on parallel computations and the use of multi-core CPU to speed-up simulations in SMPyBandits.
 
@@
 \end{itemize}
-\TODOL{FIXME order correctly in configuration.py}
 
@@
     \label{table:3:meanRegret_problem1_otherAlgorithms}
-\end{footnotesize}
+\end{footnotesize}  % WARNING
 \end{table}
@@
     \label{table:3:time_problem1_otherAlgorithms}
-\end{footnotesize}
+\end{footnotesize}  % WARNING
 \end{table}
@@
     \label{table:3:memory_problem1_otherAlgorithms}
-\end{footnotesize}
+\end{footnotesize}  % WARNING
 \end{table}
@@
 
-Finally, we give below the Bash command that was used to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
+Finally, we give below in Listing~\ref{lst:3:BashCodeToLaunchLargeExperiments} the Bash command that was used to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
 It uses \texttt{for} loops from Bash and not Python and environment variables, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $N$, $T$ and $K$.
@@
 $ for N in 4 100 1000;            do \
-  for BAYES in False True; do\
+  for BAYES in False True;        do \
   for T in 1000 5000 10000 50000; do \
@@
     DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
-             N=$N BAYES=$BAYES T=$T K=$K make single \
-    && cp logs/main_py3_log.txt logs/main_py3_log__N${N}_BAYES${BAYES}_T{T}_K{K}.txt \
+                         N=$N BAYES=$BAYES T=$T K=$K \
+                    python3 main.py configuration.py \
+    && cp logs/main_py3_log.txt \
+          logs/main_py3_log__N${N}_BAYES${BAYES}_T{T}_K{K}.txt \
   done; done; done; done
     \end{minted}
-    \caption{Snippet of Bash code to run the large experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
+    \caption{Snippet of Bash code to run the large experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.}
     % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:BashCodeToLaunchLargeExperiments}.}

### 2019-05-02 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Fixed typos

@@
 Iterating this equality and using that $\KL\left(\Pr_{\bm{\mu}}^{\cO_{0}},\Pr_{\bm{\lambda}}^{\cO_{0}}\right)=0$ yields that
-\[\sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu}_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right) + \smallO{\log(T)},\]
+\[\sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right) + \smallO{\log(T)},\]
 for all $A \in \sigma(\cO_T)$, in particular for all $A \in \cF^j_T$.

### 2019-05-02 2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py

@@ -61,36 +61,27 @@ configuration = {
         {
             "archtype": Uniform,   # The stupidest policy, fully uniform
             "params": {}
-        },
-        # --- Fixed arm algorithm
-        {
-            "archtype": TakeFixedArm,
-            "params": { "armIndex": 0 },  # Take worse arm!
+        }, {  # --- Fixed arm algorithm
+            "archtype": TakeFixedArm,   # Take worse arm!
+            "params": { "armIndex": 0 },
         }, {
-            "archtype": TakeFixedArm,
-            "params": { "armIndex": 1 },  # Take second worse arm!
+            "archtype": TakeFixedArm,   # Take second worse arm!
+            "params": { "armIndex": 1 },
         }, {
-            "archtype": TakeFixedArm,
-            "params": { "armIndex": min(2, nbArms - 1) },  # Take third worse arm!
-        },
-        # --- UCB algorithm
-        {
+            "archtype": TakeFixedArm,   # Take third worse arm!
+            "params": { "armIndex": min(2, nbArms - 1) },
+        }, {  # --- UCB algorithm
             "archtype": UCBalpha,   # UCB with alpha=1 parameter
             "params": { "alpha": 1.0 }
-        },
-        # --- Thompson algorithm
-        {
+        }, {  # --- Thompson algorithm
             "archtype": Thompson,
             "params": { "posterior": Beta }
-        },
-        # --- KL UCB algorithm
-        {
+        }, {  # --- KL UCB algorithm
             "archtype": klUCB,
             "params": { "klucb": klBern }
         },
     ]
-}
+}  # DONE
 
-# DONE
 print("Loaded experiments configuration:")
 print("configuration =", configuration)  # DEBUG

### 2019-05-06 2-Chapters/3-Chapter/chapter3.tex
Almost done for describing all this text about algorithms included in the simulations for end of Chapter 3.

@@
 % ----------------------------------------------------------------------------
-\section{Experimental comparisons of state-of-the-art single-player algorithms}
+\section{Experimental comparisons of state-of-the-art algorithms}
 \label{sec:3:reviewSPAlgorithms}
@@
     $\mathrm{Exp}3^{++}$ from \cite{Seldin17}, using $\alpha=3$ and $\beta=256$ as advised (\texttt{Exp3PlusPlus}).
+    It is a recent variant of the $\mathrm{Exp3}$ algorithm, using an adaptive tuning of its parameter.
+It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
 
@@
     $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCB}).
+    It was proved to achieve order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
 
@@
     \klUCB{} from \cite{KLUCBJournal}, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern} (\texttt{klUCB}).
+    It was proved to achieve asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
+    It is computationally more costly that \UCB, as at each time step $t$, to compute the $K$ indexes, a convex optimization problem must be solved approximately, for each arm.
+    In practice, a few steps of a Newton method are enough to obtain a precision of more than 10 digits, and empirically we found that even in the worst cases \klUCB{} is an order of magnitude slower than \UCB.
 
@@
     Thompson sampling from \cite{Kaufmann12Thompson}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{Thompson}).
+    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+    It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on C or Cython code, highly optimized).
 
@@
     Bayes-UCB from \cite{Kaufmann12BUCB}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
+    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
+    It is comparable to Thompson sampling in terms of memory complexity, but slower in terms of time complexity as computing a $1-1/t$ quantile is more costly than sampling from a distribution.
+    But in particular for Bernoulli distributed arms and for Beta posteriors and priors, Bayes-UCB is only about twice as slow as Thompson sampling.
 
@@
     AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
+    It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posteriors on each arm.
+It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson decision rule (\ie, choose the arm maximizing one draw of random samples from the posteriors), otherwise it uses the Bayes-UCB decision rule (\ie, choose the arm maximizing the $1-1/t$ quantile).
+    It was proposed in this article and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
+    In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, is comparable in terms of computation times, but costs more memory.
 
     \item Algorithm 9 is
-    BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, using a binary tournament implement in a naive but efficient way (not using recursive functions) (\texttt{BESA} with \texttt{''non\_recursive''=True}).
+    BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, and it consists in trying to fix a simple observation.
+    For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute any statistics for each arms (\eg, mean).
+    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arm $1,2$ has $n_1,n_2$ samples and if $n_1>n_2$ (for instance), first sub-sample $n_2$ observations from $n_1$, then compute the mean of these observations, and play the arm maximizing this mean (now that there is the same number for both arms, intuitively it makes more sense to compare these quantities).
+    It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
+    Analyzing it in the generic case was found to be difficult, and is still an open problem.
+    We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions) (\texttt{BESA} with \texttt{''non\_recursive''=True}).
+    Due to its binary nature, the extension to more than $K=2$ arms is costly in terms of both time and memory, as illustrated below.
 \end{itemize}
@@
 
-We consider increasing values for the horizon, $T=1000$, $T=5000$, $T=10000$ and $T=50000$.  % FIXME update if needed?
-We consider as well an increasing number of arms, $K=2$, $K=5$, $K=9$, $K=20$ and $K=50$.  % FIXME update if needed?
-For all experiments, we run $N=1000$ independent simulations.  % FIXME update if needed?
+% FIXME update if needed?
+We consider increasing values for the horizon, $T=1000$, $T=2500$, $T=5000$, $T=10000$, $T=20000$, $T=25000$ and $T=50000$.
+We consider as well an increasing number of arms, $K=2$, $K=4$, $K=8$, $K=12$, $K=16$, $K=20$ and $K=32$.
+For all experiments, we run $N=100$ independent simulations.
 
@@
         $\varepsilon$-greedy &
-            $R^{1}_{T_1,K_1} = 0 \pm 0$
-                $R^{1}_{T_1,K_2} = 0 \pm 0$
-                $R^{1}_{T_1,K_3} = 0 \pm 0$
-                $R^{1}_{T_1,K_4} = 0 \pm 0$
-                $R^{1}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{1}_{T_2,K_1} = 0 \pm 0$
-                $R^{1}_{T_2,K_2} = 0 \pm 0$
-                $R^{1}_{T_2,K_3} = 0 \pm 0$
-                $R^{1}_{T_2,K_4} = 0 \pm 0$
-                $R^{1}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{1}_{T_3,K_1} = 0 \pm 0$
-                $R^{1}_{T_3,K_2} = 0 \pm 0$
-                $R^{1}_{T_3,K_3} = 0 \pm 0$
-                $R^{1}_{T_3,K_4} = 0 \pm 0$
-                $R^{1}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{1}_{T_4,K_1} = 0 \pm 0$
-                $R^{1}_{T_4,K_2} = 0 \pm 0$
-                $R^{1}_{T_4,K_3} = 0 \pm 0$
-                $R^{1}_{T_4,K_4} = 0 \pm 0$
-                $R^{1}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'1}_{T_1,K_1} = 0 \pm 0$
+                $R^{'1}_{T_1,K_2} = 0 \pm 0$
+                $R^{'1}_{T_1,K_3} = 0 \pm 0$
+                $R^{'1}_{T_1,K_4} = 0 \pm 0$
+                $R^{'1}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'1}_{T_2,K_1} = 0 \pm 0$
+                $R^{'1}_{T_2,K_2} = 0 \pm 0$
+                $R^{'1}_{T_2,K_3} = 0 \pm 0$
+                $R^{'1}_{T_2,K_4} = 0 \pm 0$
+                $R^{'1}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'1}_{T_3,K_1} = 0 \pm 0$
+                $R^{'1}_{T_3,K_2} = 0 \pm 0$
+                $R^{'1}_{T_3,K_3} = 0 \pm 0$
+                $R^{'1}_{T_3,K_4} = 0 \pm 0$
+                $R^{'1}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'1}_{T_4,K_1} = 0 \pm 0$
+                $R^{'1}_{T_4,K_2} = 0 \pm 0$
+                $R^{'1}_{T_4,K_3} = 0 \pm 0$
+                $R^{'1}_{T_4,K_4} = 0 \pm 0$
+                $R^{'1}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         Explore-then-Commit &
-            $R^{2}_{T_1,K_1} = 0 \pm 0$
-                $R^{2}_{T_1,K_2} = 0 \pm 0$
-                $R^{2}_{T_1,K_3} = 0 \pm 0$
-                $R^{2}_{T_1,K_4} = 0 \pm 0$
-                $R^{2}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{2}_{T_2,K_1} = 0 \pm 0$
-                $R^{2}_{T_2,K_2} = 0 \pm 0$
-                $R^{2}_{T_2,K_3} = 0 \pm 0$
-                $R^{2}_{T_2,K_4} = 0 \pm 0$
-                $R^{2}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{2}_{T_3,K_1} = 0 \pm 0$
-                $R^{2}_{T_3,K_2} = 0 \pm 0$
-                $R^{2}_{T_3,K_3} = 0 \pm 0$
-                $R^{2}_{T_3,K_4} = 0 \pm 0$
-                $R^{2}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{2}_{T_4,K_1} = 0 \pm 0$
-                $R^{2}_{T_4,K_2} = 0 \pm 0$
-                $R^{2}_{T_4,K_3} = 0 \pm 0$
-                $R^{2}_{T_4,K_4} = 0 \pm 0$
-                $R^{2}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'2}_{T_1,K_1} = 0 \pm 0$
+                $R^{'2}_{T_1,K_2} = 0 \pm 0$
+                $R^{'2}_{T_1,K_3} = 0 \pm 0$
+                $R^{'2}_{T_1,K_4} = 0 \pm 0$
+                $R^{'2}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'2}_{T_2,K_1} = 0 \pm 0$
+                $R^{'2}_{T_2,K_2} = 0 \pm 0$
+                $R^{'2}_{T_2,K_3} = 0 \pm 0$
+                $R^{'2}_{T_2,K_4} = 0 \pm 0$
+                $R^{'2}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'2}_{T_3,K_1} = 0 \pm 0$
+                $R^{'2}_{T_3,K_2} = 0 \pm 0$
+                $R^{'2}_{T_3,K_3} = 0 \pm 0$
+                $R^{'2}_{T_3,K_4} = 0 \pm 0$
+                $R^{'2}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'2}_{T_4,K_1} = 0 \pm 0$
+                $R^{'2}_{T_4,K_2} = 0 \pm 0$
+                $R^{'2}_{T_4,K_3} = 0 \pm 0$
+                $R^{'2}_{T_4,K_4} = 0 \pm 0$
+                $R^{'2}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         $\mathrm{Exp}3^{++}$ &
-            $R^3_{T_1,K_1} = 0 \pm 0$
-                $R^3_{T_1,K_2} = 0 \pm 0$
-                $R^3_{T_1,K_3} = 0 \pm 0$
-                $R^3_{T_1,K_4} = 0 \pm 0$
-                $R^3_{T_1,K_5} = 0 \pm 0$ &
-            $R^3_{T_2,K_1} = 0 \pm 0$
-                $R^3_{T_2,K_2} = 0 \pm 0$
-                $R^3_{T_2,K_3} = 0 \pm 0$
-                $R^3_{T_2,K_4} = 0 \pm 0$
-                $R^3_{T_2,K_5} = 0 \pm 0$ &
-            $R^3_{T_3,K_1} = 0 \pm 0$
-                $R^3_{T_3,K_2} = 0 \pm 0$
-                $R^3_{T_3,K_3} = 0 \pm 0$
-                $R^3_{T_3,K_4} = 0 \pm 0$
-                $R^3_{T_3,K_5} = 0 \pm 0$ &
-            $R^3_{T_4,K_1} = 0 \pm 0$
-                $R^3_{T_4,K_2} = 0 \pm 0$
-                $R^3_{T_4,K_3} = 0 \pm 0$
-                $R^3_{T_4,K_4} = 0 \pm 0$
-                $R^3_{T_4,K_5} = 0 \pm 0$ \\
+            $R^3'_{T_1,K_1} = 0 \pm 0$
+                $R^3'_{T_1,K_2} = 0 \pm 0$
+                $R^3'_{T_1,K_3} = 0 \pm 0$
+                $R^3'_{T_1,K_4} = 0 \pm 0$
+                $R^3'_{T_1,K_5} = 0 \pm 0$ &
+            $R^3'_{T_2,K_1} = 0 \pm 0$
+                $R^3'_{T_2,K_2} = 0 \pm 0$
+                $R^3'_{T_2,K_3} = 0 \pm 0$
+                $R^3'_{T_2,K_4} = 0 \pm 0$
+                $R^3'_{T_2,K_5} = 0 \pm 0$ &
+            $R^3'_{T_3,K_1} = 0 \pm 0$
+                $R^3'_{T_3,K_2} = 0 \pm 0$
+                $R^3'_{T_3,K_3} = 0 \pm 0$
+                $R^3'_{T_3,K_4} = 0 \pm 0$
+                $R^3'_{T_3,K_5} = 0 \pm 0$ &
+            $R^3'_{T_4,K_1} = 0 \pm 0$
+                $R^3'_{T_4,K_2} = 0 \pm 0$
+                $R^3'_{T_4,K_3} = 0 \pm 0$
+                $R^3'_{T_4,K_4} = 0 \pm 0$
+                $R^3'_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         $\UCB_1$ &
-            $R^{4}_{T_1,K_1} = 0 \pm 0$
-                $R^{4}_{T_1,K_2} = 0 \pm 0$
-                $R^{4}_{T_1,K_3} = 0 \pm 0$
-                $R^{4}_{T_1,K_4} = 0 \pm 0$
-                $R^{4}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{4}_{T_2,K_1} = 0 \pm 0$
-                $R^{4}_{T_2,K_2} = 0 \pm 0$
-                $R^{4}_{T_2,K_3} = 0 \pm 0$
-                $R^{4}_{T_2,K_4} = 0 \pm 0$
-                $R^{4}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{4}_{T_3,K_1} = 0 \pm 0$
-                $R^{4}_{T_3,K_2} = 0 \pm 0$
-                $R^{4}_{T_3,K_3} = 0 \pm 0$
-                $R^{4}_{T_3,K_4} = 0 \pm 0$
-                $R^{4}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{4}_{T_4,K_1} = 0 \pm 0$
-                $R^{4}_{T_4,K_2} = 0 \pm 0$
-                $R^{4}_{T_4,K_3} = 0 \pm 0$
-                $R^{4}_{T_4,K_4} = 0 \pm 0$
-                $R^{4}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'4}_{T_1,K_1} = 0 \pm 0$
+                $R^{'4}_{T_1,K_2} = 0 \pm 0$
+                $R^{'4}_{T_1,K_3} = 0 \pm 0$
+                $R^{'4}_{T_1,K_4} = 0 \pm 0$
+                $R^{'4}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'4}_{T_2,K_1} = 0 \pm 0$
+                $R^{'4}_{T_2,K_2} = 0 \pm 0$
+                $R^{'4}_{T_2,K_3} = 0 \pm 0$
+                $R^{'4}_{T_2,K_4} = 0 \pm 0$
+                $R^{'4}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'4}_{T_3,K_1} = 0 \pm 0$
+                $R^{'4}_{T_3,K_2} = 0 \pm 0$
+                $R^{'4}_{T_3,K_3} = 0 \pm 0$
+                $R^{'4}_{T_3,K_4} = 0 \pm 0$
+                $R^{'4}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'4}_{T_4,K_1} = 0 \pm 0$
+                $R^{'4}_{T_4,K_2} = 0 \pm 0$
+                $R^{'4}_{T_4,K_3} = 0 \pm 0$
+                $R^{'4}_{T_4,K_4} = 0 \pm 0$
+                $R^{'4}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         \klUCB{} &
-            $R^{5}_{T_1,K_1} = 0 \pm 0$
-                $R^{5}_{T_1,K_2} = 0 \pm 0$
-                $R^{5}_{T_1,K_3} = 0 \pm 0$
-                $R^{5}_{T_1,K_4} = 0 \pm 0$
-                $R^{5}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{5}_{T_2,K_1} = 0 \pm 0$
-                $R^{5}_{T_2,K_2} = 0 \pm 0$
-                $R^{5}_{T_2,K_3} = 0 \pm 0$
-                $R^{5}_{T_2,K_4} = 0 \pm 0$
-                $R^{5}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{5}_{T_3,K_1} = 0 \pm 0$
-                $R^{5}_{T_3,K_2} = 0 \pm 0$
-                $R^{5}_{T_3,K_3} = 0 \pm 0$
-                $R^{5}_{T_3,K_4} = 0 \pm 0$
-                $R^{5}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{5}_{T_4,K_1} = 0 \pm 0$
-                $R^{5}_{T_4,K_2} = 0 \pm 0$
-                $R^{5}_{T_4,K_3} = 0 \pm 0$
-                $R^{5}_{T_4,K_4} = 0 \pm 0$
-                $R^{5}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'5}_{T_1,K_1} = 0 \pm 0$
+                $R^{'5}_{T_1,K_2} = 0 \pm 0$
+                $R^{'5}_{T_1,K_3} = 0 \pm 0$
+                $R^{'5}_{T_1,K_4} = 0 \pm 0$
+                $R^{'5}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'5}_{T_2,K_1} = 0 \pm 0$
+                $R^{'5}_{T_2,K_2} = 0 \pm 0$
+                $R^{'5}_{T_2,K_3} = 0 \pm 0$
+                $R^{'5}_{T_2,K_4} = 0 \pm 0$
+                $R^{'5}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'5}_{T_3,K_1} = 0 \pm 0$
+                $R^{'5}_{T_3,K_2} = 0 \pm 0$
+                $R^{'5}_{T_3,K_3} = 0 \pm 0$
+                $R^{'5}_{T_3,K_4} = 0 \pm 0$
+                $R^{'5}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'5}_{T_4,K_1} = 0 \pm 0$
+                $R^{'5}_{T_4,K_2} = 0 \pm 0$
+                $R^{'5}_{T_4,K_3} = 0 \pm 0$
+                $R^{'5}_{T_4,K_4} = 0 \pm 0$
+                $R^{'5}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         Thompson sampling &
-            $R^{6}_{T_1,K_1} = 0 \pm 0$
-                $R^{6}_{T_1,K_2} = 0 \pm 0$
-                $R^{6}_{T_1,K_3} = 0 \pm 0$
-                $R^{6}_{T_1,K_4} = 0 \pm 0$
-                $R^{6}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{6}_{T_2,K_1} = 0 \pm 0$
-                $R^{6}_{T_2,K_2} = 0 \pm 0$
-                $R^{6}_{T_2,K_3} = 0 \pm 0$
-                $R^{6}_{T_2,K_4} = 0 \pm 0$
-                $R^{6}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{6}_{T_3,K_1} = 0 \pm 0$
-                $R^{6}_{T_3,K_2} = 0 \pm 0$
-                $R^{6}_{T_3,K_3} = 0 \pm 0$
-                $R^{6}_{T_3,K_4} = 0 \pm 0$
-                $R^{6}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{6}_{T_4,K_1} = 0 \pm 0$
-                $R^{6}_{T_4,K_2} = 0 \pm 0$
-                $R^{6}_{T_4,K_3} = 0 \pm 0$
-                $R^{6}_{T_4,K_4} = 0 \pm 0$
-                $R^{6}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'6}_{T_1,K_1} = 0 \pm 0$
+                $R^{'6}_{T_1,K_2} = 0 \pm 0$
+                $R^{'6}_{T_1,K_3} = 0 \pm 0$
+                $R^{'6}_{T_1,K_4} = 0 \pm 0$
+                $R^{'6}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'6}_{T_2,K_1} = 0 \pm 0$
+                $R^{'6}_{T_2,K_2} = 0 \pm 0$
+                $R^{'6}_{T_2,K_3} = 0 \pm 0$
+                $R^{'6}_{T_2,K_4} = 0 \pm 0$
+                $R^{'6}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'6}_{T_3,K_1} = 0 \pm 0$
+                $R^{'6}_{T_3,K_2} = 0 \pm 0$
+                $R^{'6}_{T_3,K_3} = 0 \pm 0$
+                $R^{'6}_{T_3,K_4} = 0 \pm 0$
+                $R^{'6}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'6}_{T_4,K_1} = 0 \pm 0$
+                $R^{'6}_{T_4,K_2} = 0 \pm 0$
+                $R^{'6}_{T_4,K_3} = 0 \pm 0$
+                $R^{'6}_{T_4,K_4} = 0 \pm 0$
+                $R^{'6}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         Bayes-UCB &
-            $R^{7}_{T_1,K_1} = 0 \pm 0$
-                $R^{7}_{T_1,K_2} = 0 \pm 0$
-                $R^{7}_{T_1,K_3} = 0 \pm 0$
-                $R^{7}_{T_1,K_4} = 0 \pm 0$
-                $R^{7}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{7}_{T_2,K_1} = 0 \pm 0$
-                $R^{7}_{T_2,K_2} = 0 \pm 0$
-                $R^{7}_{T_2,K_3} = 0 \pm 0$
-                $R^{7}_{T_2,K_4} = 0 \pm 0$
-                $R^{7}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{7}_{T_3,K_1} = 0 \pm 0$
-                $R^{7}_{T_3,K_2} = 0 \pm 0$
-                $R^{7}_{T_3,K_3} = 0 \pm 0$
-                $R^{7}_{T_3,K_4} = 0 \pm 0$
-                $R^{7}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{7}_{T_4,K_1} = 0 \pm 0$
-                $R^{7}_{T_4,K_2} = 0 \pm 0$
-                $R^{7}_{T_4,K_3} = 0 \pm 0$
-                $R^{7}_{T_4,K_4} = 0 \pm 0$
-                $R^{7}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'7}_{T_1,K_1} = 0 \pm 0$
+                $R^{'7}_{T_1,K_2} = 0 \pm 0$
+                $R^{'7}_{T_1,K_3} = 0 \pm 0$
+                $R^{'7}_{T_1,K_4} = 0 \pm 0$
+                $R^{'7}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'7}_{T_2,K_1} = 0 \pm 0$
+                $R^{'7}_{T_2,K_2} = 0 \pm 0$
+                $R^{'7}_{T_2,K_3} = 0 \pm 0$
+                $R^{'7}_{T_2,K_4} = 0 \pm 0$
+                $R^{'7}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'7}_{T_3,K_1} = 0 \pm 0$
+                $R^{'7}_{T_3,K_2} = 0 \pm 0$
+                $R^{'7}_{T_3,K_3} = 0 \pm 0$
+                $R^{'7}_{T_3,K_4} = 0 \pm 0$
+                $R^{'7}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'7}_{T_4,K_1} = 0 \pm 0$
+                $R^{'7}_{T_4,K_2} = 0 \pm 0$
+                $R^{'7}_{T_4,K_3} = 0 \pm 0$
+                $R^{'7}_{T_4,K_4} = 0 \pm 0$
+                $R^{'7}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         AdBandits &
-            $R^{8}_{T_1,K_1} = 0 \pm 0$
-                $R^{8}_{T_1,K_2} = 0 \pm 0$
-                $R^{8}_{T_1,K_3} = 0 \pm 0$
-                $R^{8}_{T_1,K_4} = 0 \pm 0$
-                $R^{8}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{8}_{T_2,K_1} = 0 \pm 0$
-                $R^{8}_{T_2,K_2} = 0 \pm 0$
-                $R^{8}_{T_2,K_3} = 0 \pm 0$
-                $R^{8}_{T_2,K_4} = 0 \pm 0$
-                $R^{8}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{8}_{T_3,K_1} = 0 \pm 0$
-                $R^{8}_{T_3,K_2} = 0 \pm 0$
-                $R^{8}_{T_3,K_3} = 0 \pm 0$
-                $R^{8}_{T_3,K_4} = 0 \pm 0$
-                $R^{8}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{8}_{T_4,K_1} = 0 \pm 0$
-                $R^{8}_{T_4,K_2} = 0 \pm 0$
-                $R^{8}_{T_4,K_3} = 0 \pm 0$
-                $R^{8}_{T_4,K_4} = 0 \pm 0$
-                $R^{8}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'8}_{T_1,K_1} = 0 \pm 0$
+                $R^{'8}_{T_1,K_2} = 0 \pm 0$
+                $R^{'8}_{T_1,K_3} = 0 \pm 0$
+                $R^{'8}_{T_1,K_4} = 0 \pm 0$
+                $R^{'8}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'8}_{T_2,K_1} = 0 \pm 0$
+                $R^{'8}_{T_2,K_2} = 0 \pm 0$
+                $R^{'8}_{T_2,K_3} = 0 \pm 0$
+                $R^{'8}_{T_2,K_4} = 0 \pm 0$
+                $R^{'8}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'8}_{T_3,K_1} = 0 \pm 0$
+                $R^{'8}_{T_3,K_2} = 0 \pm 0$
+                $R^{'8}_{T_3,K_3} = 0 \pm 0$
+                $R^{'8}_{T_3,K_4} = 0 \pm 0$
+                $R^{'8}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'8}_{T_4,K_1} = 0 \pm 0$
+                $R^{'8}_{T_4,K_2} = 0 \pm 0$
+                $R^{'8}_{T_4,K_3} = 0 \pm 0$
+                $R^{'8}_{T_4,K_4} = 0 \pm 0$
+                $R^{'8}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
         BESA &
-            $R^{9}_{T_1,K_1} = 0 \pm 0$
-                $R^{9}_{T_1,K_2} = 0 \pm 0$
-                $R^{9}_{T_1,K_3} = 0 \pm 0$
-                $R^{9}_{T_1,K_4} = 0 \pm 0$
-                $R^{9}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{9}_{T_2,K_1} = 0 \pm 0$
-                $R^{9}_{T_2,K_2} = 0 \pm 0$
-                $R^{9}_{T_2,K_3} = 0 \pm 0$
-                $R^{9}_{T_2,K_4} = 0 \pm 0$
-                $R^{9}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{9}_{T_3,K_1} = 0 \pm 0$
-                $R^{9}_{T_3,K_2} = 0 \pm 0$
-                $R^{9}_{T_3,K_3} = 0 \pm 0$
-                $R^{9}_{T_3,K_4} = 0 \pm 0$
-                $R^{9}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{9}_{T_4,K_1} = 0 \pm 0$
-                $R^{9}_{T_4,K_2} = 0 \pm 0$
-                $R^{9}_{T_4,K_3} = 0 \pm 0$
-                $R^{9}_{T_4,K_4} = 0 \pm 0$
-                $R^{9}_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'9}_{T_1,K_1} = 0 \pm 0$
+                $R^{'9}_{T_1,K_2} = 0 \pm 0$
+                $R^{'9}_{T_1,K_3} = 0 \pm 0$
+                $R^{'9}_{T_1,K_4} = 0 \pm 0$
+                $R^{'9}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'9}_{T_2,K_1} = 0 \pm 0$
+                $R^{'9}_{T_2,K_2} = 0 \pm 0$
+                $R^{'9}_{T_2,K_3} = 0 \pm 0$
+                $R^{'9}_{T_2,K_4} = 0 \pm 0$
+                $R^{'9}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'9}_{T_3,K_1} = 0 \pm 0$
+                $R^{'9}_{T_3,K_2} = 0 \pm 0$
+                $R^{'9}_{T_3,K_3} = 0 \pm 0$
+                $R^{'9}_{T_3,K_4} = 0 \pm 0$
+                $R^{'9}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'9}_{T_4,K_1} = 0 \pm 0$
+                $R^{'9}_{T_4,K_2} = 0 \pm 0$
+                $R^{'9}_{T_4,K_3} = 0 \pm 0$
+                $R^{'9}_{T_4,K_4} = 0 \pm 0$
+                $R^{'9}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
@@
     $\mathrm{MOSS}$-$\mathrm{Anytime}$ from \cite{Degenne16}, using $\alpha=1$ (\texttt{MOSSAnytime}).
+    It is a recent variant of the $\mathrm{MOSS}$ algorithm \cite{BubeckSlivkins12}, using an adaptive tuning of its parameter.
+It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+    Empirically, it performs usually similarly to \UCB, and is slightly more costly in terms of both time and memory.
 
@@
     the approximated Finite-Horizon Gittins from \cite{Lattimore16a}, using $\alpha=1$ (\texttt{ApproximatedFHGittins}).
+    It mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
+    It was proved to achieve order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+    Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in terms of both time and memory.
 
@@
     $\UCB\dagger$ from \cite{Lattimore2018refining}, using $\alpha=1$ (\texttt{UCBdagger}).
+    It mimics the \UCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
+    It was also proved to achieve order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+    Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in computation time, but it is much more costly in terms of memory.
 
     \item Algorithm 13 is
-    the anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018}, using $\alpha=1$ (\texttt{klUCBswitchAnytime}).
+    the anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018} (\texttt{klUCBswitchAnytime}).
+    It is a very recent variant of the \klUCB{} algorithm \cite{KLUCBJournal},
+    and it mimics the \klUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
+    It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+    Empirically, it usually outperforms \klUCB{} and costs about the same memory, but it is more costly in terms of computation time.
 
     \item Algorithm 14 is
-    \textsc{Tsallis-Inf} from \cite{Zimmert2018}, using $\alpha=0.5$ (it seems to be the most efficient of the algorithms described in the article) (\texttt{TsallisInf}).
+    \textsc{Tsallis-Inf} from \cite{Zimmert2018}, using $\alpha=1/2$ (it seems to be the most efficient of the algorithms described in the article) (\texttt{TsallisInf}).
+    It is based on the online mirror descent algorithm with a Tsallis entropy regularizer, and it is anytime.
+    It was also proved to to obtain good ``best of both world'' performances.
+    Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
+Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction},Hazan2016introduction and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
+    \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as KL-UCB-Switch), and costs an-order-of-magnitude more memory.
 
     \item Algorithm 15 is
-    $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019}, using $\alpha=1$ and permutations uniformly sampled in $[0,1]$ (it is the simplest of the algorithms described in the article) (\texttt{RCB}).
+    $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019}, using $\alpha=1$ and perturbations uniformly sampled in $[0,1]$.
+We prefer to use the simplest of the algorithms described in the article, and other algorithms in the same article used other families of distributions for the perturbations (\texttt{RCB}).
+    It was analyzed and found to be order-optimal for problem-dependent bounds, and empirically we found that it is usually performs slightly worse than \UCB, in terms of regret, time and memory.
+    It is included because it remains comparably efficient with the other state-of-the-art algorithms, and because the key message from \cite{KimTewari2019} is very interesting:
+    ``the optimism embedded in UCB can be replaced by simple randomization''.
 
@@
     $\mathrm{PHE}$ (Perturbed-History Exploration) from \cite{KvetonSzepesvari2019}, using a perturbation scale of $0.5$ as advised (\texttt{PHE}).
-
+    The algorithm adds $\cO(t)$ \iid{} pseudo-rewards to its history in round $t$, and then pulls the arm with the highest estimated value in its perturbed history.
+    It was also analyzed and found to be comparable with \UCB.
+    Like $\mathrm{RCB}$, $\mathrm{PHE}$ was found to be empirically slightly worse than other state-of-the-art algorithms.
+    Interestingly, its time and memory consumption stay constant with respect to the step number $t$ and the horizon $T$, because generating and summing $t$ pseudo-rewards from a Bernoulli distribution can actually be done in $\cO(1)$ time and space, by using efficient sampling methods for sampling from Binomial distributions.
 \end{itemize}

### 2019-05-06 2-Chapters/3-Chapter/logs/extract_data_from_hdf5.py

@@ -1,0 +1,19 @@
+#!/usr/bin/env python3
+""" Read a bunch of data from all the HDF5 files in the current folder, and plot aggregated view of the simulation data."""
+
+import glob
+import numpy as np
+import matplotlib.pyplot as plt
+import h5py
+
+
+def main():
+    list_of_files = glob.glob("*.hdf5")
+    dict_of_envs = {
+        itspath: hdf5.File(itspath)['env_0']
+        for itspath in list_of_files
+    }
+    
+
+if __name__ == "__main__":
+    main()

### 2019-05-06 2-Chapters/3-Chapter/logs/extract_data_from_logs.sh

@@ -1,0 +1,47 @@
+#!/usr/bin/env bash
+set -euo pipefail
+
+texoutput=chapter3.tex
+cp -vf ../$texoutput ./$texoutput
+
+DEBUG=True
+NOPLOTS=False
+SAVEALL=False
+N_JOBS=4
+nbplayer=16
+
+declare -i n_i
+declare -i n_b
+declare -i n_t
+declare -i n_k
+
+n_i=0
+for N in 40 100 1000; do
+    n_i+=1
+    n_b=0
+    for BAYES in False True; do
+        n_b+=1
+        n_t=0
+        for T in 1000 5000 10000 50000; do
+            n_t+=1
+            n_k=0
+            for K in 2 5 10 20 50; do
+                n_k+=1
+                log_file=main_py3_log__N${N}_BAYES${BAYES}_T${T}_K${K}__$(date +"%d-%m-%Y_%H-%M-%S").txt
+                for n in $(seq 1 ${nbplayer}); do
+                    grep "R^{${n}} = [0-9\.]* pm [0-9\.]*" $log_file
+                    echo sed -i.backup s/"R^{${n}} = [0-9\.]* pm [0-9\.]*"/"R^{${n}} = [0-9\.]* pm [0-9\.]*"/g ./$texoutput
+                    grep "M^{${n}} = [0-9\.]* pm [0-9\.]*" $log_file
+                    grep "T^{${n}} = [0-9\.]* pm [0-9\.]*" $log_file
+                done;
+                # R^{[0-9\.]*}_{T=${T},K=${T}} = [0-9\.]* pm [0-9\.]*
+                # R^{9}_{T_${n_t},K_${n_k}} = 0 \pm 0
+                # ==>
+                # R^{9}_{T_3,K_1} = 0 \pm 0
+                # sed -i.backup s/''/''/g ./$texoutput
+            done
+        done
+    done
+done
+
+cp -vf ./$texoutput ../$texoutput

### 2019-05-06 2-Chapters/3-Chapter/logs/launch_simulations.sh

@@ -1,0 +1,31 @@
+    #!/usr/bin/env bash
+set -euo pipefail
+
+DEBUG=False
+NOPLOTS=True
+SAVEALL=True
+N_JOBS=12
+
+declare -i nbexperiment
+nbexperiment=0
+
+for N in 100; do
+# for N in $N_JOBS; do
+    for BAYES in False True; do
+    # for BAYES in False; do
+        for T in 1000 2500 5000 10000 20000 25000 50000; do
+        # for T in 1000 2500; do
+            for K in 2 4 8 12 16 20 32; do
+            # for K in 2 4 8; do
+                debut=$(date)
+                nbexperiment+=1
+
+                DEBUG=$DEBUG NOPLOTS=$NOPLOTS SAVEALL=$SAVEALL N_JOBS=$N_JOBS \
+                N=$N T=$T K=$K BAYES=$BAYES mymake.sh --noFreeSMS single \
+                && cp -f logs/main_py3_log.txt \
+                            logs/main_py3_log__N${N}_BAYES${BAYES}_T${T}_K${K}__$(date +"%d-%m-%Y_%H-%M-%S").txt \
+                && FreeSMS.py "Expérience #${nbexperiment}, commencée ${debut}, terminée ! Avec N=${N}, BAYES=${BAYES}, T=${T}, K=${K}."
+            done
+        done
+    done
+done

### 2019-05-06 2-Chapters/3-Chapter/chapter3.tex
SO NOT SURE HOW TO CONTINUE THIS CHAPTER

@@
 
-\TODOL{Write text for this Section, write configuration.py file, run short-term experiment, include them, if happy about it, run longer ang larger experiments}
+% \TODOL{Write text for this Section, write configuration.py file, run short-term experiment, include them, if happy about it, run longer ang larger experiments}
 
@@
     an $\varepsilon$-greedy algorithm, as described in \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
+    It has a very small cost in terms of time and memory, but it usually achieves linear regret.
 
@@
     an Explore-then-Commit algorithm from \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
+    Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but only obtain large (linear) regret.
 
@@
 % FIXME update if needed?
-We consider increasing values for the horizon, $T=1000$, $T=2500$, $T=5000$, $T=10000$, $T=20000$, $T=25000$ and $T=50000$.
-We consider as well an increasing number of arms, $K=2$, $K=4$, $K=8$, $K=12$, $K=16$, $K=20$ and $K=32$.
+We consider increasing values for the horizon, and we focus on $7$ exponentially-spaced values: $T=1000$, $T=2500$, $T=5000$, $T=10000$, $T=20000$, $T=25000$ and $T=50000$.
+We consider as well an increasing number of arms, and we focus again on $7$ values: $K=2$, $K=4$, $K=8$, $K=12$, $K=16$, $K=20$ and $K=32$.
 For all experiments, we run $N=100$ independent simulations.
@@
 
-\TODOL{FIXME launch these experiments, for a very short number of repetitions N=4, and if I'm happy about the results and if they are easily copied-and-pasted here, then launch with N=100 or N=1000 depending on the running time of the whole simulations}
+% \TODOL{FIXME launch these experiments, for a very short number of repetitions N=4, and if I'm happy about the results and if they are easily copied-and-pasted here, then launch with N=100 or N=1000 depending on the running time of the whole simulations}
 
@@
 
+\paragraph{Normalizing data?}
+
+The two figures below regret normalized data, in the following way.
+For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
+Simply considering the average of such measurements make no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
+%
+We know that efficient algorithms are expected to follow these patterns:
+\begin{itemize}
+    \item Final regret $R_T$ should scale as $\cO(K \log(T))$,
+    \item Total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
+    \item Total memory cost should scale $\cM_T$ as $\cO(K \log(T))$.
+\end{itemize}
+Thus, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
+
 \subsection{Computational time}
@@
 \TODOL{Just one long table for problem 1, not necessary to include other problems}
+
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.05\linewidth]{16_different_algorithms__lognormregret_vs_lognormtime__56pb__7KS_7Ts.pdf}
+	\caption[Normalized mean regret vs normalized running time (in micro-seconds).]{
+        Normalized mean regret vs normalized running time (in micro-seconds),
+        aggregating the results from different values of $K$ and $T$, for $16$ algorithms.
+        Both the $x$-axis and $y$-axis are in log-scale.
+        \UCB{} and \klUCB{} are among the best algorithms, while AdBandits, Thompson sampling and Bayes-UCB slightly outperform them in terms of regret, and have similar running times.
+	}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormtime__56pb__7KS_7Ts}
+\end{figure}
 
@@
 \TODOL{Just one long table for problem 1, not necessary to include other problems}
+
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.05\linewidth]{16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7KS_7Ts.pdf}
+	\caption[Normalized mean regret vs normalized memory costs (in bytes).]{
+        Normalized mean regret vs normalized memory costs (in bytes),
+        aggregating the results from different values of $K$ and $T$, for $16$ algorithms.
+        Both the $x$-axis and $y$-axis are in log-scale.
+        Thompson sampling appears as the best algorithm in this visualization, while \UCB{} has the advantage of being memory efficient, and \klUCB{} obtains similar performances.
+	}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7KS_7Ts}
+\end{figure}
 

### 2019-05-06 1-Preamble/4-Acknowledgement/acknowledgement.tex
So boring

@@
 
-I would like to acknowledge and thank the following people, and also acknowledge the various funding who supported our works during the last three years and made this thesis possible.
+I would like to acknowledge and thank the following people, and also acknowledge the various fundings who supported our works during the last three years and made this thesis possible.
 
-My first thoughts go to my two PhD advisors.
+My first thoughts go to my two PhD advisors.\\
 %
+\indent
 Firstly, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in IETR lab and in CentraleSupélec campus of Rennes, a very nice and warm environment for research and day-to-day life.

### 2019-05-06 2-Chapters/1-Chapter/chapter1.tex
So boring

@@
 \paragraph{Tutorial}
-    I gave a tutorial\footnote{Cf. \texttt{hal.archives-ouvertes.fr/cel-01830248}} on the Julia language, at IETR seminar in Vannes in June $2018$,
+    I gave a tutorial\footnote{Cf. \href{HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}} on the Julia language, at IETR seminar in Vannes in June $2018$,
     with Pierre Haessig (\texttt{pierreh.eu}).

### 2019-05-06 2-Chapters/3-Chapter/chapter3.tex
So boring

@@
 
-\paragraph{List of algorithms}
+% \paragraph{List of algorithms}
 
@@
     For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute any statistics for each arms (\eg, mean).
-    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arm $1,2$ has $n_1,n_2$ samples and if $n_1>n_2$ (for instance), first sub-sample $n_2$ observations from $n_1$, then compute the mean of these observations, and play the arm maximizing this mean (now that there is the same number for both arms, intuitively it makes more sense to compare these quantities).
+    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arm $1,2$ has $n_1,n_2$ samples and if $n_1>n_2$ (for instance), first sub-sample $n_2$ observations from $n_1$, then compute the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, intuitively it makes more sense to compare these quantities).
     It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
@@
 For all experiments, we run $N=100$ independent simulations.
+See the snippet of code~\ref{lst:3:BashCodeToLaunchLargeExperiments} in the Appendix below.
 
@@
         $\mathrm{Exp}3^{++}$ &
-            $R^3'_{T_1,K_1} = 0 \pm 0$
-                $R^3'_{T_1,K_2} = 0 \pm 0$
-                $R^3'_{T_1,K_3} = 0 \pm 0$
-                $R^3'_{T_1,K_4} = 0 \pm 0$
-                $R^3'_{T_1,K_5} = 0 \pm 0$ &
-            $R^3'_{T_2,K_1} = 0 \pm 0$
-                $R^3'_{T_2,K_2} = 0 \pm 0$
-                $R^3'_{T_2,K_3} = 0 \pm 0$
-                $R^3'_{T_2,K_4} = 0 \pm 0$
-                $R^3'_{T_2,K_5} = 0 \pm 0$ &
-            $R^3'_{T_3,K_1} = 0 \pm 0$
-                $R^3'_{T_3,K_2} = 0 \pm 0$
-                $R^3'_{T_3,K_3} = 0 \pm 0$
-                $R^3'_{T_3,K_4} = 0 \pm 0$
-                $R^3'_{T_3,K_5} = 0 \pm 0$ &
-            $R^3'_{T_4,K_1} = 0 \pm 0$
-                $R^3'_{T_4,K_2} = 0 \pm 0$
-                $R^3'_{T_4,K_3} = 0 \pm 0$
-                $R^3'_{T_4,K_4} = 0 \pm 0$
-                $R^3'_{T_4,K_5} = 0 \pm 0$ \\
+            $R^{'3}_{T_1,K_1} = 0 \pm 0$
+                $R^{'3}_{T_1,K_2} = 0 \pm 0$
+                $R^{'3}_{T_1,K_3} = 0 \pm 0$
+                $R^{'3}_{T_1,K_4} = 0 \pm 0$
+                $R^{'3}_{T_1,K_5} = 0 \pm 0$ &
+            $R^{'3}_{T_2,K_1} = 0 \pm 0$
+                $R^{'3}_{T_2,K_2} = 0 \pm 0$
+                $R^{'3}_{T_2,K_3} = 0 \pm 0$
+                $R^{'3}_{T_2,K_4} = 0 \pm 0$
+                $R^{'3}_{T_2,K_5} = 0 \pm 0$ &
+            $R^{'3}_{T_3,K_1} = 0 \pm 0$
+                $R^{'3}_{T_3,K_2} = 0 \pm 0$
+                $R^{'3}_{T_3,K_3} = 0 \pm 0$
+                $R^{'3}_{T_3,K_4} = 0 \pm 0$
+                $R^{'3}_{T_3,K_5} = 0 \pm 0$ &
+            $R^{'3}_{T_4,K_1} = 0 \pm 0$
+                $R^{'3}_{T_4,K_2} = 0 \pm 0$
+                $R^{'3}_{T_4,K_3} = 0 \pm 0$
+                $R^{'3}_{T_4,K_4} = 0 \pm 0$
+                $R^{'3}_{T_4,K_5} = 0 \pm 0$ \\
         \hline
@@
 	% \centering
-	\includegraphics[width=1.05\linewidth]{16_different_algorithms__lognormregret_vs_lognormtime__56pb__7KS_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormtime__56pb__7KS_7Ts.pdf}
 	\caption[Normalized mean regret vs normalized running time (in micro-seconds).]{
@@
 	% \centering
-	\includegraphics[width=1.05\linewidth]{16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7KS_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7KS_7Ts.pdf}
 	\caption[Normalized mean regret vs normalized memory costs (in bytes).]{
@@
     It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
-    Empirically, it usually outperforms \klUCB{} and costs about the same memory, but it is more costly in terms of computation time.
+    Empirically, it usually outperforms \klUCB{} in terms of regret, and costs about the same memory, but it is slower.
 
@@
   for BAYES in False True;        do \
-  for T in 1000 5000 10000 50000; do \
-  for K in 2 5 10 20 50;          do \
+  for T in 1000 2500 5000 10000 20000 50000; do \
+  for K     in 2 4 8 12 16 20 32; do \
     DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \

### 2019-05-06 3-Appendices/2-Appendix/appendix2.tex
So boring

@@
 This appendix quickly presents the contributions of another publication that was not presented in the main text of this thesis.
-We studied doubling tricks and their possible use for multi-armed bandits, between fall 2017 and spring 2018, and we wrote an article \cite{Besson2018DoublingTricks} which was rejected at the COLT 2018 conference, and unfortunately we lacked the time to improve it and resubmit it to another conference.
+We studied doubling tricks and their possible uses for multi-armed bandits, between fall 2017 and spring 2018, and we wrote an article \cite{Besson2018DoublingTricks} which was rejected at the COLT 2018 conference, and unfortunately we lacked the time to improve it and resubmit it to another conference.
 
@@
 
-To conclude this work, we would need to complete the study of the doubling tricks, and instead of focussing on two families of doubling tricks (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bound.
+To conclude this work, we would need to complete the study of the doubling tricks, and instead of focussing on two families (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bound.
 Such doubling is given either by the function mapping the current time $t$ to the current estimate of the horizon $T(t)$, or the sequence $(T_i)_{i\in\N^*}$ of successive estimates of the horizons.

### 2019-05-06 2-Chapters/3-Chapter/chapter3.tex
Lots of new plots.

@@
 
-We consider the nine following algorithms. For each of them, we give a bibliographic reference, that corresponds to the most recent article studying it and not the first one which introduced it.
+We consider the nine following algorithms, and seven more are described in Appendix~\ref{sub:3:additionalExperiments}.
+For each of them, we give a bibliographic reference, that corresponds to the most recent article studying it and not the first one which introduced it.
 We give the chosen tuning of the parameters, for parametric algorithms.
-Algorithms that are not anytime use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, of course they use the correct successive values.
-We also give the name of the corresponding class in the \texttt{Policies} module in SMPyBandits.
+Algorithms that are not anytime use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, of course they use the correct successive values. \\
+%
+\indent
+For reproducibility, we also give in ``\texttt{typewriter font}'' the name of the corresponding class in the \texttt{Policies} module in SMPyBandits.
 
@@
 
-\begin{table}[!t]
-\begin{footnotesize}  % WARNING
-    \centering
-    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-        \hline
-        $\varepsilon$-greedy &
-            $R^{1}_{T_1,K_1} = 0 \pm 0$
-                $R^{1}_{T_1,K_2} = 0 \pm 0$
-                $R^{1}_{T_1,K_3} = 0 \pm 0$
-                $R^{1}_{T_1,K_4} = 0 \pm 0$
-                $R^{1}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{1}_{T_2,K_1} = 0 \pm 0$
-                $R^{1}_{T_2,K_2} = 0 \pm 0$
-                $R^{1}_{T_2,K_3} = 0 \pm 0$
-                $R^{1}_{T_2,K_4} = 0 \pm 0$
-                $R^{1}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{1}_{T_3,K_1} = 0 \pm 0$
-                $R^{1}_{T_3,K_2} = 0 \pm 0$
-                $R^{1}_{T_3,K_3} = 0 \pm 0$
-                $R^{1}_{T_3,K_4} = 0 \pm 0$
-                $R^{1}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{1}_{T_4,K_1} = 0 \pm 0$
-                $R^{1}_{T_4,K_2} = 0 \pm 0$
-                $R^{1}_{T_4,K_3} = 0 \pm 0$
-                $R^{1}_{T_4,K_4} = 0 \pm 0$
-                $R^{1}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        Explore-then-Commit &
-            $R^{2}_{T_1,K_1} = 0 \pm 0$
-                $R^{2}_{T_1,K_2} = 0 \pm 0$
-                $R^{2}_{T_1,K_3} = 0 \pm 0$
-                $R^{2}_{T_1,K_4} = 0 \pm 0$
-                $R^{2}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{2}_{T_2,K_1} = 0 \pm 0$
-                $R^{2}_{T_2,K_2} = 0 \pm 0$
-                $R^{2}_{T_2,K_3} = 0 \pm 0$
-                $R^{2}_{T_2,K_4} = 0 \pm 0$
-                $R^{2}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{2}_{T_3,K_1} = 0 \pm 0$
-                $R^{2}_{T_3,K_2} = 0 \pm 0$
-                $R^{2}_{T_3,K_3} = 0 \pm 0$
-                $R^{2}_{T_3,K_4} = 0 \pm 0$
-                $R^{2}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{2}_{T_4,K_1} = 0 \pm 0$
-                $R^{2}_{T_4,K_2} = 0 \pm 0$
-                $R^{2}_{T_4,K_3} = 0 \pm 0$
-                $R^{2}_{T_4,K_4} = 0 \pm 0$
-                $R^{2}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        $\mathrm{Exp}3^{++}$ &
-            $R^3_{T_1,K_1} = 0 \pm 0$
-                $R^3_{T_1,K_2} = 0 \pm 0$
-                $R^3_{T_1,K_3} = 0 \pm 0$
-                $R^3_{T_1,K_4} = 0 \pm 0$
-                $R^3_{T_1,K_5} = 0 \pm 0$ &
-            $R^3_{T_2,K_1} = 0 \pm 0$
-                $R^3_{T_2,K_2} = 0 \pm 0$
-                $R^3_{T_2,K_3} = 0 \pm 0$
-                $R^3_{T_2,K_4} = 0 \pm 0$
-                $R^3_{T_2,K_5} = 0 \pm 0$ &
-            $R^3_{T_3,K_1} = 0 \pm 0$
-                $R^3_{T_3,K_2} = 0 \pm 0$
-                $R^3_{T_3,K_3} = 0 \pm 0$
-                $R^3_{T_3,K_4} = 0 \pm 0$
-                $R^3_{T_3,K_5} = 0 \pm 0$ &
-            $R^3_{T_4,K_1} = 0 \pm 0$
-                $R^3_{T_4,K_2} = 0 \pm 0$
-                $R^3_{T_4,K_3} = 0 \pm 0$
-                $R^3_{T_4,K_4} = 0 \pm 0$
-                $R^3_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        $\UCB_1$ &
-            $R^{4}_{T_1,K_1} = 0 \pm 0$
-                $R^{4}_{T_1,K_2} = 0 \pm 0$
-                $R^{4}_{T_1,K_3} = 0 \pm 0$
-                $R^{4}_{T_1,K_4} = 0 \pm 0$
-                $R^{4}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{4}_{T_2,K_1} = 0 \pm 0$
-                $R^{4}_{T_2,K_2} = 0 \pm 0$
-                $R^{4}_{T_2,K_3} = 0 \pm 0$
-                $R^{4}_{T_2,K_4} = 0 \pm 0$
-                $R^{4}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{4}_{T_3,K_1} = 0 \pm 0$
-                $R^{4}_{T_3,K_2} = 0 \pm 0$
-                $R^{4}_{T_3,K_3} = 0 \pm 0$
-                $R^{4}_{T_3,K_4} = 0 \pm 0$
-                $R^{4}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{4}_{T_4,K_1} = 0 \pm 0$
-                $R^{4}_{T_4,K_2} = 0 \pm 0$
-                $R^{4}_{T_4,K_3} = 0 \pm 0$
-                $R^{4}_{T_4,K_4} = 0 \pm 0$
-                $R^{4}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        \klUCB{} &
-            $R^{5}_{T_1,K_1} = 0 \pm 0$
-                $R^{5}_{T_1,K_2} = 0 \pm 0$
-                $R^{5}_{T_1,K_3} = 0 \pm 0$
-                $R^{5}_{T_1,K_4} = 0 \pm 0$
-                $R^{5}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{5}_{T_2,K_1} = 0 \pm 0$
-                $R^{5}_{T_2,K_2} = 0 \pm 0$
-                $R^{5}_{T_2,K_3} = 0 \pm 0$
-                $R^{5}_{T_2,K_4} = 0 \pm 0$
-                $R^{5}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{5}_{T_3,K_1} = 0 \pm 0$
-                $R^{5}_{T_3,K_2} = 0 \pm 0$
-                $R^{5}_{T_3,K_3} = 0 \pm 0$
-                $R^{5}_{T_3,K_4} = 0 \pm 0$
-                $R^{5}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{5}_{T_4,K_1} = 0 \pm 0$
-                $R^{5}_{T_4,K_2} = 0 \pm 0$
-                $R^{5}_{T_4,K_3} = 0 \pm 0$
-                $R^{5}_{T_4,K_4} = 0 \pm 0$
-                $R^{5}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        Thompson sampling &
-            $R^{6}_{T_1,K_1} = 0 \pm 0$
-                $R^{6}_{T_1,K_2} = 0 \pm 0$
-                $R^{6}_{T_1,K_3} = 0 \pm 0$
-                $R^{6}_{T_1,K_4} = 0 \pm 0$
-                $R^{6}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{6}_{T_2,K_1} = 0 \pm 0$
-                $R^{6}_{T_2,K_2} = 0 \pm 0$
-                $R^{6}_{T_2,K_3} = 0 \pm 0$
-                $R^{6}_{T_2,K_4} = 0 \pm 0$
-                $R^{6}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{6}_{T_3,K_1} = 0 \pm 0$
-                $R^{6}_{T_3,K_2} = 0 \pm 0$
-                $R^{6}_{T_3,K_3} = 0 \pm 0$
-                $R^{6}_{T_3,K_4} = 0 \pm 0$
-                $R^{6}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{6}_{T_4,K_1} = 0 \pm 0$
-                $R^{6}_{T_4,K_2} = 0 \pm 0$
-                $R^{6}_{T_4,K_3} = 0 \pm 0$
-                $R^{6}_{T_4,K_4} = 0 \pm 0$
-                $R^{6}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        Thompson sampling &
-            $R^{7}_{T_1,K_1} = 0 \pm 0$
-                $R^{7}_{T_1,K_2} = 0 \pm 0$
-                $R^{7}_{T_1,K_3} = 0 \pm 0$
-                $R^{7}_{T_1,K_4} = 0 \pm 0$
-                $R^{7}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{7}_{T_2,K_1} = 0 \pm 0$
-                $R^{7}_{T_2,K_2} = 0 \pm 0$
-                $R^{7}_{T_2,K_3} = 0 \pm 0$
-                $R^{7}_{T_2,K_4} = 0 \pm 0$
-                $R^{7}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{7}_{T_3,K_1} = 0 \pm 0$
-                $R^{7}_{T_3,K_2} = 0 \pm 0$
-                $R^{7}_{T_3,K_3} = 0 \pm 0$
-                $R^{7}_{T_3,K_4} = 0 \pm 0$
-                $R^{7}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{7}_{T_4,K_1} = 0 \pm 0$
-                $R^{7}_{T_4,K_2} = 0 \pm 0$
-                $R^{7}_{T_4,K_3} = 0 \pm 0$
-                $R^{7}_{T_4,K_4} = 0 \pm 0$
-                $R^{7}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        AdBandits &
-            $R^{8}_{T_1,K_1} = 0 \pm 0$
-                $R^{8}_{T_1,K_2} = 0 \pm 0$
-                $R^{8}_{T_1,K_3} = 0 \pm 0$
-                $R^{8}_{T_1,K_4} = 0 \pm 0$
-                $R^{8}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{8}_{T_2,K_1} = 0 \pm 0$
-                $R^{8}_{T_2,K_2} = 0 \pm 0$
-                $R^{8}_{T_2,K_3} = 0 \pm 0$
-                $R^{8}_{T_2,K_4} = 0 \pm 0$
-                $R^{8}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{8}_{T_3,K_1} = 0 \pm 0$
-                $R^{8}_{T_3,K_2} = 0 \pm 0$
-                $R^{8}_{T_3,K_3} = 0 \pm 0$
-                $R^{8}_{T_3,K_4} = 0 \pm 0$
-                $R^{8}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{8}_{T_4,K_1} = 0 \pm 0$
-                $R^{8}_{T_4,K_2} = 0 \pm 0$
-                $R^{8}_{T_4,K_3} = 0 \pm 0$
-                $R^{8}_{T_4,K_4} = 0 \pm 0$
-                $R^{8}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        BESA &
-            $R^{9}_{T_1,K_1} = 0 \pm 0$
-                $R^{9}_{T_1,K_2} = 0 \pm 0$
-                $R^{9}_{T_1,K_3} = 0 \pm 0$
-                $R^{9}_{T_1,K_4} = 0 \pm 0$
-                $R^{9}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{9}_{T_2,K_1} = 0 \pm 0$
-                $R^{9}_{T_2,K_2} = 0 \pm 0$
-                $R^{9}_{T_2,K_3} = 0 \pm 0$
-                $R^{9}_{T_2,K_4} = 0 \pm 0$
-                $R^{9}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{9}_{T_3,K_1} = 0 \pm 0$
-                $R^{9}_{T_3,K_2} = 0 \pm 0$
-                $R^{9}_{T_3,K_3} = 0 \pm 0$
-                $R^{9}_{T_3,K_4} = 0 \pm 0$
-                $R^{9}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{9}_{T_4,K_1} = 0 \pm 0$
-                $R^{9}_{T_4,K_2} = 0 \pm 0$
-                $R^{9}_{T_4,K_3} = 0 \pm 0$
-                $R^{9}_{T_4,K_4} = 0 \pm 0$
-                $R^{9}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-    \end{tabular}
-    \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 1} with different values of time $T$ and number of arm $K$.
+% \begin{table}[!t]
+% \begin{footnotesize}  % WARNING
+%     \centering
+%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+%         \hline
+%         $\varepsilon$-greedy &
+%             $R^{1}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{1}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{1}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{1}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{1}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{1}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{1}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{1}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{1}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{1}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{1}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{1}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{1}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{1}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{1}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{1}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{1}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{1}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{1}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{1}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         Explore-then-Commit &
+%             $R^{2}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{2}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{2}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{2}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{2}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{2}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{2}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{2}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{2}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{2}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{2}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{2}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{2}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{2}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{2}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{2}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{2}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{2}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{2}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{2}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         $\mathrm{Exp}3^{++}$ &
+%             $R^3_{T_1,K_1} = 0 \pm 0$
+%                 $R^3_{T_1,K_2} = 0 \pm 0$
+%                 $R^3_{T_1,K_3} = 0 \pm 0$
+%                 $R^3_{T_1,K_4} = 0 \pm 0$
+%                 $R^3_{T_1,K_5} = 0 \pm 0$ &
+%             $R^3_{T_2,K_1} = 0 \pm 0$
+%                 $R^3_{T_2,K_2} = 0 \pm 0$
+%                 $R^3_{T_2,K_3} = 0 \pm 0$
+%                 $R^3_{T_2,K_4} = 0 \pm 0$
+%                 $R^3_{T_2,K_5} = 0 \pm 0$ &
+%             $R^3_{T_3,K_1} = 0 \pm 0$
+%                 $R^3_{T_3,K_2} = 0 \pm 0$
+%                 $R^3_{T_3,K_3} = 0 \pm 0$
+%                 $R^3_{T_3,K_4} = 0 \pm 0$
+%                 $R^3_{T_3,K_5} = 0 \pm 0$ &
+%             $R^3_{T_4,K_1} = 0 \pm 0$
+%                 $R^3_{T_4,K_2} = 0 \pm 0$
+%                 $R^3_{T_4,K_3} = 0 \pm 0$
+%                 $R^3_{T_4,K_4} = 0 \pm 0$
+%                 $R^3_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         $\UCB_1$ &
+%             $R^{4}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{4}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{4}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{4}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{4}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{4}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{4}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{4}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{4}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{4}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{4}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{4}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{4}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{4}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{4}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{4}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{4}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{4}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{4}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{4}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         \klUCB{} &
+%             $R^{5}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{5}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{5}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{5}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{5}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{5}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{5}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{5}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{5}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{5}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{5}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{5}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{5}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{5}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{5}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{5}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{5}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{5}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{5}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{5}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         Thompson sampling &
+%             $R^{6}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{6}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{6}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{6}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{6}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{6}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{6}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{6}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{6}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{6}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{6}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{6}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{6}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{6}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{6}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{6}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{6}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{6}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{6}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{6}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         Thompson sampling &
+%             $R^{7}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{7}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{7}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{7}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{7}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{7}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{7}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{7}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{7}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{7}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{7}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{7}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{7}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{7}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{7}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{7}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{7}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{7}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{7}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{7}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         AdBandits &
+%             $R^{8}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{8}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{8}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{8}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{8}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{8}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{8}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{8}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{8}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{8}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{8}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{8}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{8}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{8}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{8}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{8}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{8}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{8}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{8}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{8}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         BESA &
+%             $R^{9}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{9}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{9}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{9}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{9}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{9}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{9}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{9}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{9}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{9}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{9}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{9}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{9}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{9}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{9}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{9}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{9}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{9}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{9}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{9}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%     \end{tabular}
+%     \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 1} with different values of time $T$ and number of arm $K$.
+%     }
+%     \label{table:3:meanRegret_problem1}
+% \end{footnotesize}  % WARNING
+% \end{table}
+
+% \begin{table}[!t]
+% \begin{footnotesize}  % WARNING
+%     \centering
+%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+%         \hline
+%         $\varepsilon$-greedy &
+%             $R^{'1}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'1}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'1}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'1}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'1}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'1}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'1}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'1}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'1}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'1}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'1}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'1}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'1}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'1}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'1}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'1}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'1}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'1}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'1}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'1}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         Explore-then-Commit &
+%             $R^{'2}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'2}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'2}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'2}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'2}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'2}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'2}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'2}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'2}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'2}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'2}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'2}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'2}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'2}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'2}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'2}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'2}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'2}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'2}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'2}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         $\mathrm{Exp}3^{++}$ &
+%             $R^{'3}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'3}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'3}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'3}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'3}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'3}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'3}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'3}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'3}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'3}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'3}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'3}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'3}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'3}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'3}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'3}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'3}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'3}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'3}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'3}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         $\UCB_1$ &
+%             $R^{'4}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'4}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'4}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'4}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'4}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'4}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'4}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'4}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'4}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'4}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'4}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'4}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'4}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'4}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'4}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'4}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'4}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'4}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'4}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'4}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         \klUCB{} &
+%             $R^{'5}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'5}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'5}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'5}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'5}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'5}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'5}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'5}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'5}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'5}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'5}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'5}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'5}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'5}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'5}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'5}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'5}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'5}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'5}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'5}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         Thompson sampling &
+%             $R^{'6}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'6}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'6}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'6}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'6}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'6}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'6}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'6}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'6}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'6}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'6}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'6}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'6}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'6}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'6}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'6}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'6}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'6}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'6}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'6}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         Bayes-UCB &
+%             $R^{'7}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'7}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'7}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'7}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'7}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'7}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'7}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'7}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'7}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'7}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'7}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'7}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'7}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'7}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'7}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'7}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'7}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'7}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'7}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'7}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         AdBandits &
+%             $R^{'8}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'8}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'8}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'8}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'8}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'8}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'8}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'8}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'8}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'8}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'8}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'8}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'8}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'8}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'8}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'8}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'8}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'8}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'8}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'8}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         BESA &
+%             $R^{'9}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{'9}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{'9}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{'9}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{'9}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{'9}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{'9}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{'9}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{'9}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{'9}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{'9}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{'9}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{'9}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{'9}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{'9}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{'9}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{'9}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{'9}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{'9}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{'9}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%     \end{tabular}
+%     \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 2} with different values of time $T$ and number of arm $K$.
+%     }
+%     \label{table:3:meanRegret_problem2}
+% \end{footnotesize}  % WARNING
+% \end{table}
+
+% Include in appendix of this chapter the results for 7 more algorithms, the most recent and less important ones (MOSSAnytime, ApproximatedFHGittins, UCBdagger, klUCBswitchAnytime, PHE, RCB, Tsallis-Inf)
+
+% FIXME generate from the notebook then copy from WS3
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_arms__56pb__7Ks_T50000.pdf}
+	\caption[Regret vs different values of $K$.]{
+        Regret vs different values of $K$ ($R_T / \log(T)$),
+        for $T=50000$ and for $16$ algorithms.
+        The $y$-axis is in log-scale.
+        All algorithms appear to have a regret slowly growing with respect to $K$, as predicted by the regret bounds which are linear with respect to the number of arms, as predicted.
+        \textsc{Tsallis-Inf} is the only state-of-the-art algorithm which appears to be very inefficient for problems with a small number of arms.
+        Both the $\varepsilon$-greedy and Explore-then-Commit algorithms performed poorly, achieving linear regret (not displayed).
 	}
-    \label{table:3:meanRegret_problem1}
-\end{footnotesize}  % WARNING
-\end{table}
-
-\begin{table}[!t]
-\begin{footnotesize}  % WARNING
-    \centering
-    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-        \hline
-        $\varepsilon$-greedy &
-            $R^{'1}_{T_1,K_1} = 0 \pm 0$
-                $R^{'1}_{T_1,K_2} = 0 \pm 0$
-                $R^{'1}_{T_1,K_3} = 0 \pm 0$
-                $R^{'1}_{T_1,K_4} = 0 \pm 0$
-                $R^{'1}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'1}_{T_2,K_1} = 0 \pm 0$
-                $R^{'1}_{T_2,K_2} = 0 \pm 0$
-                $R^{'1}_{T_2,K_3} = 0 \pm 0$
-                $R^{'1}_{T_2,K_4} = 0 \pm 0$
-                $R^{'1}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'1}_{T_3,K_1} = 0 \pm 0$
-                $R^{'1}_{T_3,K_2} = 0 \pm 0$
-                $R^{'1}_{T_3,K_3} = 0 \pm 0$
-                $R^{'1}_{T_3,K_4} = 0 \pm 0$
-                $R^{'1}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'1}_{T_4,K_1} = 0 \pm 0$
-                $R^{'1}_{T_4,K_2} = 0 \pm 0$
-                $R^{'1}_{T_4,K_3} = 0 \pm 0$
-                $R^{'1}_{T_4,K_4} = 0 \pm 0$
-                $R^{'1}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        Explore-then-Commit &
-            $R^{'2}_{T_1,K_1} = 0 \pm 0$
-                $R^{'2}_{T_1,K_2} = 0 \pm 0$
-                $R^{'2}_{T_1,K_3} = 0 \pm 0$
-                $R^{'2}_{T_1,K_4} = 0 \pm 0$
-                $R^{'2}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'2}_{T_2,K_1} = 0 \pm 0$
-                $R^{'2}_{T_2,K_2} = 0 \pm 0$
-                $R^{'2}_{T_2,K_3} = 0 \pm 0$
-                $R^{'2}_{T_2,K_4} = 0 \pm 0$
-                $R^{'2}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'2}_{T_3,K_1} = 0 \pm 0$
-                $R^{'2}_{T_3,K_2} = 0 \pm 0$
-                $R^{'2}_{T_3,K_3} = 0 \pm 0$
-                $R^{'2}_{T_3,K_4} = 0 \pm 0$
-                $R^{'2}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'2}_{T_4,K_1} = 0 \pm 0$
-                $R^{'2}_{T_4,K_2} = 0 \pm 0$
-                $R^{'2}_{T_4,K_3} = 0 \pm 0$
-                $R^{'2}_{T_4,K_4} = 0 \pm 0$
-                $R^{'2}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        $\mathrm{Exp}3^{++}$ &
-            $R^{'3}_{T_1,K_1} = 0 \pm 0$
-                $R^{'3}_{T_1,K_2} = 0 \pm 0$
-                $R^{'3}_{T_1,K_3} = 0 \pm 0$
-                $R^{'3}_{T_1,K_4} = 0 \pm 0$
-                $R^{'3}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'3}_{T_2,K_1} = 0 \pm 0$
-                $R^{'3}_{T_2,K_2} = 0 \pm 0$
-                $R^{'3}_{T_2,K_3} = 0 \pm 0$
-                $R^{'3}_{T_2,K_4} = 0 \pm 0$
-                $R^{'3}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'3}_{T_3,K_1} = 0 \pm 0$
-                $R^{'3}_{T_3,K_2} = 0 \pm 0$
-                $R^{'3}_{T_3,K_3} = 0 \pm 0$
-                $R^{'3}_{T_3,K_4} = 0 \pm 0$
-                $R^{'3}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'3}_{T_4,K_1} = 0 \pm 0$
-                $R^{'3}_{T_4,K_2} = 0 \pm 0$
-                $R^{'3}_{T_4,K_3} = 0 \pm 0$
-                $R^{'3}_{T_4,K_4} = 0 \pm 0$
-                $R^{'3}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        $\UCB_1$ &
-            $R^{'4}_{T_1,K_1} = 0 \pm 0$
-                $R^{'4}_{T_1,K_2} = 0 \pm 0$
-                $R^{'4}_{T_1,K_3} = 0 \pm 0$
-                $R^{'4}_{T_1,K_4} = 0 \pm 0$
-                $R^{'4}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'4}_{T_2,K_1} = 0 \pm 0$
-                $R^{'4}_{T_2,K_2} = 0 \pm 0$
-                $R^{'4}_{T_2,K_3} = 0 \pm 0$
-                $R^{'4}_{T_2,K_4} = 0 \pm 0$
-                $R^{'4}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'4}_{T_3,K_1} = 0 \pm 0$
-                $R^{'4}_{T_3,K_2} = 0 \pm 0$
-                $R^{'4}_{T_3,K_3} = 0 \pm 0$
-                $R^{'4}_{T_3,K_4} = 0 \pm 0$
-                $R^{'4}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'4}_{T_4,K_1} = 0 \pm 0$
-                $R^{'4}_{T_4,K_2} = 0 \pm 0$
-                $R^{'4}_{T_4,K_3} = 0 \pm 0$
-                $R^{'4}_{T_4,K_4} = 0 \pm 0$
-                $R^{'4}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        \klUCB{} &
-            $R^{'5}_{T_1,K_1} = 0 \pm 0$
-                $R^{'5}_{T_1,K_2} = 0 \pm 0$
-                $R^{'5}_{T_1,K_3} = 0 \pm 0$
-                $R^{'5}_{T_1,K_4} = 0 \pm 0$
-                $R^{'5}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'5}_{T_2,K_1} = 0 \pm 0$
-                $R^{'5}_{T_2,K_2} = 0 \pm 0$
-                $R^{'5}_{T_2,K_3} = 0 \pm 0$
-                $R^{'5}_{T_2,K_4} = 0 \pm 0$
-                $R^{'5}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'5}_{T_3,K_1} = 0 \pm 0$
-                $R^{'5}_{T_3,K_2} = 0 \pm 0$
-                $R^{'5}_{T_3,K_3} = 0 \pm 0$
-                $R^{'5}_{T_3,K_4} = 0 \pm 0$
-                $R^{'5}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'5}_{T_4,K_1} = 0 \pm 0$
-                $R^{'5}_{T_4,K_2} = 0 \pm 0$
-                $R^{'5}_{T_4,K_3} = 0 \pm 0$
-                $R^{'5}_{T_4,K_4} = 0 \pm 0$
-                $R^{'5}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        Thompson sampling &
-            $R^{'6}_{T_1,K_1} = 0 \pm 0$
-                $R^{'6}_{T_1,K_2} = 0 \pm 0$
-                $R^{'6}_{T_1,K_3} = 0 \pm 0$
-                $R^{'6}_{T_1,K_4} = 0 \pm 0$
-                $R^{'6}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'6}_{T_2,K_1} = 0 \pm 0$
-                $R^{'6}_{T_2,K_2} = 0 \pm 0$
-                $R^{'6}_{T_2,K_3} = 0 \pm 0$
-                $R^{'6}_{T_2,K_4} = 0 \pm 0$
-                $R^{'6}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'6}_{T_3,K_1} = 0 \pm 0$
-                $R^{'6}_{T_3,K_2} = 0 \pm 0$
-                $R^{'6}_{T_3,K_3} = 0 \pm 0$
-                $R^{'6}_{T_3,K_4} = 0 \pm 0$
-                $R^{'6}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'6}_{T_4,K_1} = 0 \pm 0$
-                $R^{'6}_{T_4,K_2} = 0 \pm 0$
-                $R^{'6}_{T_4,K_3} = 0 \pm 0$
-                $R^{'6}_{T_4,K_4} = 0 \pm 0$
-                $R^{'6}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        Bayes-UCB &
-            $R^{'7}_{T_1,K_1} = 0 \pm 0$
-                $R^{'7}_{T_1,K_2} = 0 \pm 0$
-                $R^{'7}_{T_1,K_3} = 0 \pm 0$
-                $R^{'7}_{T_1,K_4} = 0 \pm 0$
-                $R^{'7}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'7}_{T_2,K_1} = 0 \pm 0$
-                $R^{'7}_{T_2,K_2} = 0 \pm 0$
-                $R^{'7}_{T_2,K_3} = 0 \pm 0$
-                $R^{'7}_{T_2,K_4} = 0 \pm 0$
-                $R^{'7}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'7}_{T_3,K_1} = 0 \pm 0$
-                $R^{'7}_{T_3,K_2} = 0 \pm 0$
-                $R^{'7}_{T_3,K_3} = 0 \pm 0$
-                $R^{'7}_{T_3,K_4} = 0 \pm 0$
-                $R^{'7}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'7}_{T_4,K_1} = 0 \pm 0$
-                $R^{'7}_{T_4,K_2} = 0 \pm 0$
-                $R^{'7}_{T_4,K_3} = 0 \pm 0$
-                $R^{'7}_{T_4,K_4} = 0 \pm 0$
-                $R^{'7}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        AdBandits &
-            $R^{'8}_{T_1,K_1} = 0 \pm 0$
-                $R^{'8}_{T_1,K_2} = 0 \pm 0$
-                $R^{'8}_{T_1,K_3} = 0 \pm 0$
-                $R^{'8}_{T_1,K_4} = 0 \pm 0$
-                $R^{'8}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'8}_{T_2,K_1} = 0 \pm 0$
-                $R^{'8}_{T_2,K_2} = 0 \pm 0$
-                $R^{'8}_{T_2,K_3} = 0 \pm 0$
-                $R^{'8}_{T_2,K_4} = 0 \pm 0$
-                $R^{'8}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'8}_{T_3,K_1} = 0 \pm 0$
-                $R^{'8}_{T_3,K_2} = 0 \pm 0$
-                $R^{'8}_{T_3,K_3} = 0 \pm 0$
-                $R^{'8}_{T_3,K_4} = 0 \pm 0$
-                $R^{'8}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'8}_{T_4,K_1} = 0 \pm 0$
-                $R^{'8}_{T_4,K_2} = 0 \pm 0$
-                $R^{'8}_{T_4,K_3} = 0 \pm 0$
-                $R^{'8}_{T_4,K_4} = 0 \pm 0$
-                $R^{'8}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        BESA &
-            $R^{'9}_{T_1,K_1} = 0 \pm 0$
-                $R^{'9}_{T_1,K_2} = 0 \pm 0$
-                $R^{'9}_{T_1,K_3} = 0 \pm 0$
-                $R^{'9}_{T_1,K_4} = 0 \pm 0$
-                $R^{'9}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{'9}_{T_2,K_1} = 0 \pm 0$
-                $R^{'9}_{T_2,K_2} = 0 \pm 0$
-                $R^{'9}_{T_2,K_3} = 0 \pm 0$
-                $R^{'9}_{T_2,K_4} = 0 \pm 0$
-                $R^{'9}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{'9}_{T_3,K_1} = 0 \pm 0$
-                $R^{'9}_{T_3,K_2} = 0 \pm 0$
-                $R^{'9}_{T_3,K_3} = 0 \pm 0$
-                $R^{'9}_{T_3,K_4} = 0 \pm 0$
-                $R^{'9}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{'9}_{T_4,K_1} = 0 \pm 0$
-                $R^{'9}_{T_4,K_2} = 0 \pm 0$
-                $R^{'9}_{T_4,K_3} = 0 \pm 0$
-                $R^{'9}_{T_4,K_4} = 0 \pm 0$
-                $R^{'9}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-    \end{tabular}
-    \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 2} with different values of time $T$ and number of arm $K$.
+	\label{fig:3:16_different_algorithms__lognormregret_vs_arms__56pb__7Ks_T50000}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normregret_vs_horizons__56pb__K32_7Ts.pdf}
+	\caption[Regret vs different values of $T$.]{
+        Regret vs different values of $T$ ($R_T / K$),
+        for $K=32$ and for $16$ algorithms.
+        All efficient algorithms appear to have a logarithmic regret, meaning that their regret bound is indeed logarithmic with respect to the horizon, as predicted.
+        The respective ranking of all the algorithms also appears to remain preserved for different values of $T$, which is also backed-up by theoretical results: if two algorithms $\cA$ and $\cA'$ has a regret close to their regret bounds, of the form $R_T \leq \cO(K \log(T))$, and the bound for $\cA$ use a smaller constant than the bound for $\cA'$, $\cA$ should obtain a smaller regret than $\cA'$ no matter the horizon.
+        Finally, we also observe once more than both the $\varepsilon$-greedy and Explore-then-Commit algorithms performed poorly, achieving linear regret (not displayed after $T=100000$).
 	}
-    \label{table:3:meanRegret_problem2}
-\end{footnotesize}  % WARNING
-\end{table}
-
-% Include in appendix of this chapter the results for 7 more algorithms, the most recent and less important ones (MOSSAnytime, ApproximatedFHGittins, UCBdagger, klUCBswitchAnytime, PHE, RCB, Tsallis-Inf)
+	\label{fig:3:16_different_algorithms__normregret_vs_horizons__56pb__K32_7Ts}
+\end{figure}
 
@@
 
-\begin{table}[!t]
-\begin{footnotesize}  % WARNING
-    \centering
-    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-        \hline
-        $\varepsilon$-greedy &
-            $\cT^{1}_{T_1,K_1} = 0$
-                $\cT^{1}_{T_1,K_2} = 0$
-                $\cT^{1}_{T_1,K_3} = 0$
-                $\cT^{1}_{T_1,K_4} = 0$
-                $\cT^{1}_{T_1,K_5} = 0$ &
-            $\cT^{1}_{T_2,K_1} = 0$
-                $\cT^{1}_{T_2,K_2} = 0$
-                $\cT^{1}_{T_2,K_3} = 0$
-                $\cT^{1}_{T_2,K_4} = 0$
-                $\cT^{1}_{T_2,K_5} = 0$ &
-            $\cT^{1}_{T_3,K_1} = 0$
-                $\cT^{1}_{T_3,K_2} = 0$
-                $\cT^{1}_{T_3,K_3} = 0$
-                $\cT^{1}_{T_3,K_4} = 0$
-                $\cT^{1}_{T_3,K_5} = 0$ &
-            $\cT^{1}_{T_4,K_1} = 0$
-                $\cT^{1}_{T_4,K_2} = 0$
-                $\cT^{1}_{T_4,K_3} = 0$
-                $\cT^{1}_{T_4,K_4} = 0$
-                $\cT^{1}_{T_4,K_5} = 0$ \\
-        \hline
-        Explore-then-Commit &
-            $\cT^{2}_{T_1,K_1} = 0$
-                $\cT^{2}_{T_1,K_2} = 0$
-                $\cT^{2}_{T_1,K_3} = 0$
-                $\cT^{2}_{T_1,K_4} = 0$
-                $\cT^{2}_{T_1,K_5} = 0$ &
-            $\cT^{2}_{T_2,K_1} = 0$
-                $\cT^{2}_{T_2,K_2} = 0$
-                $\cT^{2}_{T_2,K_3} = 0$
-                $\cT^{2}_{T_2,K_4} = 0$
-                $\cT^{2}_{T_2,K_5} = 0$ &
-            $\cT^{2}_{T_3,K_1} = 0$
-                $\cT^{2}_{T_3,K_2} = 0$
-                $\cT^{2}_{T_3,K_3} = 0$
-                $\cT^{2}_{T_3,K_4} = 0$
-                $\cT^{2}_{T_3,K_5} = 0$ &
-            $\cT^{2}_{T_4,K_1} = 0$
-                $\cT^{2}_{T_4,K_2} = 0$
-                $\cT^{2}_{T_4,K_3} = 0$
-                $\cT^{2}_{T_4,K_4} = 0$
-                $\cT^{2}_{T_4,K_5} = 0$ \\
-        \hline
-        $\mathrm{Exp}3^{++}$ &
-            $\cT^3_{T_1,K_1} = 0$
-                $\cT^3_{T_1,K_2} = 0$
-                $\cT^3_{T_1,K_3} = 0$
-                $\cT^3_{T_1,K_4} = 0$
-                $\cT^3_{T_1,K_5} = 0$ &
-            $\cT^3_{T_2,K_1} = 0$
-                $\cT^3_{T_2,K_2} = 0$
-                $\cT^3_{T_2,K_3} = 0$
-                $\cT^3_{T_2,K_4} = 0$
-                $\cT^3_{T_2,K_5} = 0$ &
-            $\cT^3_{T_3,K_1} = 0$
-                $\cT^3_{T_3,K_2} = 0$
-                $\cT^3_{T_3,K_3} = 0$
-                $\cT^3_{T_3,K_4} = 0$
-                $\cT^3_{T_3,K_5} = 0$ &
-            $\cT^3_{T_4,K_1} = 0$
-                $\cT^3_{T_4,K_2} = 0$
-                $\cT^3_{T_4,K_3} = 0$
-                $\cT^3_{T_4,K_4} = 0$
-                $\cT^3_{T_4,K_5} = 0$ \\
-        \hline
-        $\UCB_1$ &
-            $\cT^{4}_{T_1,K_1} = 0$
-                $\cT^{4}_{T_1,K_2} = 0$
-                $\cT^{4}_{T_1,K_3} = 0$
-                $\cT^{4}_{T_1,K_4} = 0$
-                $\cT^{4}_{T_1,K_5} = 0$ &
-            $\cT^{4}_{T_2,K_1} = 0$
-                $\cT^{4}_{T_2,K_2} = 0$
-                $\cT^{4}_{T_2,K_3} = 0$
-                $\cT^{4}_{T_2,K_4} = 0$
-                $\cT^{4}_{T_2,K_5} = 0$ &
-            $\cT^{4}_{T_3,K_1} = 0$
-                $\cT^{4}_{T_3,K_2} = 0$
-                $\cT^{4}_{T_3,K_3} = 0$
-                $\cT^{4}_{T_3,K_4} = 0$
-                $\cT^{4}_{T_3,K_5} = 0$ &
-            $\cT^{4}_{T_4,K_1} = 0$
-                $\cT^{4}_{T_4,K_2} = 0$
-                $\cT^{4}_{T_4,K_3} = 0$
-                $\cT^{4}_{T_4,K_4} = 0$
-                $\cT^{4}_{T_4,K_5} = 0$ \\
-        \hline
-        \klUCB{} &
-            $\cT^{5}_{T_1,K_1} = 0$
-                $\cT^{5}_{T_1,K_2} = 0$
-                $\cT^{5}_{T_1,K_3} = 0$
-                $\cT^{5}_{T_1,K_4} = 0$
-                $\cT^{5}_{T_1,K_5} = 0$ &
-            $\cT^{5}_{T_2,K_1} = 0$
-                $\cT^{5}_{T_2,K_2} = 0$
-                $\cT^{5}_{T_2,K_3} = 0$
-                $\cT^{5}_{T_2,K_4} = 0$
-                $\cT^{5}_{T_2,K_5} = 0$ &
-            $\cT^{5}_{T_3,K_1} = 0$
-                $\cT^{5}_{T_3,K_2} = 0$
-                $\cT^{5}_{T_3,K_3} = 0$
-                $\cT^{5}_{T_3,K_4} = 0$
-                $\cT^{5}_{T_3,K_5} = 0$ &
-            $\cT^{5}_{T_4,K_1} = 0$
-                $\cT^{5}_{T_4,K_2} = 0$
-                $\cT^{5}_{T_4,K_3} = 0$
-                $\cT^{5}_{T_4,K_4} = 0$
-                $\cT^{5}_{T_4,K_5} = 0$ \\
-        \hline
-        Thompson sampling &
-            $\cT^{6}_{T_1,K_1} = 0$
-                $\cT^{6}_{T_1,K_2} = 0$
-                $\cT^{6}_{T_1,K_3} = 0$
-                $\cT^{6}_{T_1,K_4} = 0$
-                $\cT^{6}_{T_1,K_5} = 0$ &
-            $\cT^{6}_{T_2,K_1} = 0$
-                $\cT^{6}_{T_2,K_2} = 0$
-                $\cT^{6}_{T_2,K_3} = 0$
-                $\cT^{6}_{T_2,K_4} = 0$
-                $\cT^{6}_{T_2,K_5} = 0$ &
-            $\cT^{6}_{T_3,K_1} = 0$
-                $\cT^{6}_{T_3,K_2} = 0$
-                $\cT^{6}_{T_3,K_3} = 0$
-                $\cT^{6}_{T_3,K_4} = 0$
-                $\cT^{6}_{T_3,K_5} = 0$ &
-            $\cT^{6}_{T_4,K_1} = 0$
-                $\cT^{6}_{T_4,K_2} = 0$
-                $\cT^{6}_{T_4,K_3} = 0$
-                $\cT^{6}_{T_4,K_4} = 0$
-                $\cT^{6}_{T_4,K_5} = 0$ \\
-        \hline
-        Thompson sampling &
-            $\cT^{7}_{T_1,K_1} = 0$
-                $\cT^{7}_{T_1,K_2} = 0$
-                $\cT^{7}_{T_1,K_3} = 0$
-                $\cT^{7}_{T_1,K_4} = 0$
-                $\cT^{7}_{T_1,K_5} = 0$ &
-            $\cT^{7}_{T_2,K_1} = 0$
-                $\cT^{7}_{T_2,K_2} = 0$
-                $\cT^{7}_{T_2,K_3} = 0$
-                $\cT^{7}_{T_2,K_4} = 0$
-                $\cT^{7}_{T_2,K_5} = 0$ &
-            $\cT^{7}_{T_3,K_1} = 0$
-                $\cT^{7}_{T_3,K_2} = 0$
-                $\cT^{7}_{T_3,K_3} = 0$
-                $\cT^{7}_{T_3,K_4} = 0$
-                $\cT^{7}_{T_3,K_5} = 0$ &
-            $\cT^{7}_{T_4,K_1} = 0$
-                $\cT^{7}_{T_4,K_2} = 0$
-                $\cT^{7}_{T_4,K_3} = 0$
-                $\cT^{7}_{T_4,K_4} = 0$
-                $\cT^{7}_{T_4,K_5} = 0$ \\
-        \hline
-        AdBandits &
-            $\cT^{8}_{T_1,K_1} = 0$
-                $\cT^{8}_{T_1,K_2} = 0$
-                $\cT^{8}_{T_1,K_3} = 0$
-                $\cT^{8}_{T_1,K_4} = 0$
-                $\cT^{8}_{T_1,K_5} = 0$ &
-            $\cT^{8}_{T_2,K_1} = 0$
-                $\cT^{8}_{T_2,K_2} = 0$
-                $\cT^{8}_{T_2,K_3} = 0$
-                $\cT^{8}_{T_2,K_4} = 0$
-                $\cT^{8}_{T_2,K_5} = 0$ &
-            $\cT^{8}_{T_3,K_1} = 0$
-                $\cT^{8}_{T_3,K_2} = 0$
-                $\cT^{8}_{T_3,K_3} = 0$
-                $\cT^{8}_{T_3,K_4} = 0$
-                $\cT^{8}_{T_3,K_5} = 0$ &
-            $\cT^{8}_{T_4,K_1} = 0$
-                $\cT^{8}_{T_4,K_2} = 0$
-                $\cT^{8}_{T_4,K_3} = 0$
-                $\cT^{8}_{T_4,K_4} = 0$
-                $\cT^{8}_{T_4,K_5} = 0$ \\
-        \hline
-        BESA &
-            $\cT^{9}_{T_1,K_1} = 0$
-                $\cT^{9}_{T_1,K_2} = 0$
-                $\cT^{9}_{T_1,K_3} = 0$
-                $\cT^{9}_{T_1,K_4} = 0$
-                $\cT^{9}_{T_1,K_5} = 0$ &
-            $\cT^{9}_{T_2,K_1} = 0$
-                $\cT^{9}_{T_2,K_2} = 0$
-                $\cT^{9}_{T_2,K_3} = 0$
-                $\cT^{9}_{T_2,K_4} = 0$
-                $\cT^{9}_{T_2,K_5} = 0$ &
-            $\cT^{9}_{T_3,K_1} = 0$
-                $\cT^{9}_{T_3,K_2} = 0$
-                $\cT^{9}_{T_3,K_3} = 0$
-                $\cT^{9}_{T_3,K_4} = 0$
-                $\cT^{9}_{T_3,K_5} = 0$ &
-            $\cT^{9}_{T_4,K_1} = 0$
-                $\cT^{9}_{T_4,K_2} = 0$
-                $\cT^{9}_{T_4,K_3} = 0$
-                $\cT^{9}_{T_4,K_4} = 0$
-                $\cT^{9}_{T_4,K_5} = 0$ \\
-        \hline
-    \end{tabular}
-    \caption{Normalized running time, rounded and in milli-seconds, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
-    \label{table:3:time_problem1}
-\end{footnotesize}  % WARNING
-\end{table}
+% FIXME generate from the notebook then copy from WS3
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormtime_vs_arms__56pb__7Ks_T50000.pdf}
+	\caption[Normalized running time vs different values of $K$.]{
+        Normalized running time vs different values of $K$ ($\cT_T / T$),
+        for $T=50000$ and for $16$ algorithms.
+        $y$-axis is in log-scale.
+        All algorithms except BESA has a linear normalized running time, meaning that for $K$ arms they use a computation time proportional to $K$, as predicted.
+	}
+	\label{fig:3:16_different_algorithms__lognormtime_vs_arms__56pb__7Ks_T50000}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normtime_vs_horizons__56pb__K32_7Ts.pdf}
+	\caption[Normalized running time vs different values of $T$.]{
+        Normalized running time vs different values of $T$ ($\cT_T / T$),
+        for $K=32$ and for $16$ algorithms.
+        All algorithms except BESA has a constant normalized running time, meaning that for $T$ rounds they use a computation time proportional to $T$, as predicted.
+	}
+	\label{fig:3:16_different_algorithms__normtime_vs_horizons__56pb__K32_7Ts}
+\end{figure}
+
+% \begin{table}[!t]
+% \begin{footnotesize}  % WARNING
+%     \centering
+%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+%         \hline
+%         $\varepsilon$-greedy &
+%             $\cT^{1}_{T_1,K_1} = 0$
+%                 $\cT^{1}_{T_1,K_2} = 0$
+%                 $\cT^{1}_{T_1,K_3} = 0$
+%                 $\cT^{1}_{T_1,K_4} = 0$
+%                 $\cT^{1}_{T_1,K_5} = 0$ &
+%             $\cT^{1}_{T_2,K_1} = 0$
+%                 $\cT^{1}_{T_2,K_2} = 0$
+%                 $\cT^{1}_{T_2,K_3} = 0$
+%                 $\cT^{1}_{T_2,K_4} = 0$
+%                 $\cT^{1}_{T_2,K_5} = 0$ &
+%             $\cT^{1}_{T_3,K_1} = 0$
+%                 $\cT^{1}_{T_3,K_2} = 0$
+%                 $\cT^{1}_{T_3,K_3} = 0$
+%                 $\cT^{1}_{T_3,K_4} = 0$
+%                 $\cT^{1}_{T_3,K_5} = 0$ &
+%             $\cT^{1}_{T_4,K_1} = 0$
+%                 $\cT^{1}_{T_4,K_2} = 0$
+%                 $\cT^{1}_{T_4,K_3} = 0$
+%                 $\cT^{1}_{T_4,K_4} = 0$
+%                 $\cT^{1}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Explore-then-Commit &
+%             $\cT^{2}_{T_1,K_1} = 0$
+%                 $\cT^{2}_{T_1,K_2} = 0$
+%                 $\cT^{2}_{T_1,K_3} = 0$
+%                 $\cT^{2}_{T_1,K_4} = 0$
+%                 $\cT^{2}_{T_1,K_5} = 0$ &
+%             $\cT^{2}_{T_2,K_1} = 0$
+%                 $\cT^{2}_{T_2,K_2} = 0$
+%                 $\cT^{2}_{T_2,K_3} = 0$
+%                 $\cT^{2}_{T_2,K_4} = 0$
+%                 $\cT^{2}_{T_2,K_5} = 0$ &
+%             $\cT^{2}_{T_3,K_1} = 0$
+%                 $\cT^{2}_{T_3,K_2} = 0$
+%                 $\cT^{2}_{T_3,K_3} = 0$
+%                 $\cT^{2}_{T_3,K_4} = 0$
+%                 $\cT^{2}_{T_3,K_5} = 0$ &
+%             $\cT^{2}_{T_4,K_1} = 0$
+%                 $\cT^{2}_{T_4,K_2} = 0$
+%                 $\cT^{2}_{T_4,K_3} = 0$
+%                 $\cT^{2}_{T_4,K_4} = 0$
+%                 $\cT^{2}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\mathrm{Exp}3^{++}$ &
+%             $\cT^3_{T_1,K_1} = 0$
+%                 $\cT^3_{T_1,K_2} = 0$
+%                 $\cT^3_{T_1,K_3} = 0$
+%                 $\cT^3_{T_1,K_4} = 0$
+%                 $\cT^3_{T_1,K_5} = 0$ &
+%             $\cT^3_{T_2,K_1} = 0$
+%                 $\cT^3_{T_2,K_2} = 0$
+%                 $\cT^3_{T_2,K_3} = 0$
+%                 $\cT^3_{T_2,K_4} = 0$
+%                 $\cT^3_{T_2,K_5} = 0$ &
+%             $\cT^3_{T_3,K_1} = 0$
+%                 $\cT^3_{T_3,K_2} = 0$
+%                 $\cT^3_{T_3,K_3} = 0$
+%                 $\cT^3_{T_3,K_4} = 0$
+%                 $\cT^3_{T_3,K_5} = 0$ &
+%             $\cT^3_{T_4,K_1} = 0$
+%                 $\cT^3_{T_4,K_2} = 0$
+%                 $\cT^3_{T_4,K_3} = 0$
+%                 $\cT^3_{T_4,K_4} = 0$
+%                 $\cT^3_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\UCB_1$ &
+%             $\cT^{4}_{T_1,K_1} = 0$
+%                 $\cT^{4}_{T_1,K_2} = 0$
+%                 $\cT^{4}_{T_1,K_3} = 0$
+%                 $\cT^{4}_{T_1,K_4} = 0$
+%                 $\cT^{4}_{T_1,K_5} = 0$ &
+%             $\cT^{4}_{T_2,K_1} = 0$
+%                 $\cT^{4}_{T_2,K_2} = 0$
+%                 $\cT^{4}_{T_2,K_3} = 0$
+%                 $\cT^{4}_{T_2,K_4} = 0$
+%                 $\cT^{4}_{T_2,K_5} = 0$ &
+%             $\cT^{4}_{T_3,K_1} = 0$
+%                 $\cT^{4}_{T_3,K_2} = 0$
+%                 $\cT^{4}_{T_3,K_3} = 0$
+%                 $\cT^{4}_{T_3,K_4} = 0$
+%                 $\cT^{4}_{T_3,K_5} = 0$ &
+%             $\cT^{4}_{T_4,K_1} = 0$
+%                 $\cT^{4}_{T_4,K_2} = 0$
+%                 $\cT^{4}_{T_4,K_3} = 0$
+%                 $\cT^{4}_{T_4,K_4} = 0$
+%                 $\cT^{4}_{T_4,K_5} = 0$ \\
+%         \hline
+%         \klUCB{} &
+%             $\cT^{5}_{T_1,K_1} = 0$
+%                 $\cT^{5}_{T_1,K_2} = 0$
+%                 $\cT^{5}_{T_1,K_3} = 0$
+%                 $\cT^{5}_{T_1,K_4} = 0$
+%                 $\cT^{5}_{T_1,K_5} = 0$ &
+%             $\cT^{5}_{T_2,K_1} = 0$
+%                 $\cT^{5}_{T_2,K_2} = 0$
+%                 $\cT^{5}_{T_2,K_3} = 0$
+%                 $\cT^{5}_{T_2,K_4} = 0$
+%                 $\cT^{5}_{T_2,K_5} = 0$ &
+%             $\cT^{5}_{T_3,K_1} = 0$
+%                 $\cT^{5}_{T_3,K_2} = 0$
+%                 $\cT^{5}_{T_3,K_3} = 0$
+%                 $\cT^{5}_{T_3,K_4} = 0$
+%                 $\cT^{5}_{T_3,K_5} = 0$ &
+%             $\cT^{5}_{T_4,K_1} = 0$
+%                 $\cT^{5}_{T_4,K_2} = 0$
+%                 $\cT^{5}_{T_4,K_3} = 0$
+%                 $\cT^{5}_{T_4,K_4} = 0$
+%                 $\cT^{5}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Thompson sampling &
+%             $\cT^{6}_{T_1,K_1} = 0$
+%                 $\cT^{6}_{T_1,K_2} = 0$
+%                 $\cT^{6}_{T_1,K_3} = 0$
+%                 $\cT^{6}_{T_1,K_4} = 0$
+%                 $\cT^{6}_{T_1,K_5} = 0$ &
+%             $\cT^{6}_{T_2,K_1} = 0$
+%                 $\cT^{6}_{T_2,K_2} = 0$
+%                 $\cT^{6}_{T_2,K_3} = 0$
+%                 $\cT^{6}_{T_2,K_4} = 0$
+%                 $\cT^{6}_{T_2,K_5} = 0$ &
+%             $\cT^{6}_{T_3,K_1} = 0$
+%                 $\cT^{6}_{T_3,K_2} = 0$
+%                 $\cT^{6}_{T_3,K_3} = 0$
+%                 $\cT^{6}_{T_3,K_4} = 0$
+%                 $\cT^{6}_{T_3,K_5} = 0$ &
+%             $\cT^{6}_{T_4,K_1} = 0$
+%                 $\cT^{6}_{T_4,K_2} = 0$
+%                 $\cT^{6}_{T_4,K_3} = 0$
+%                 $\cT^{6}_{T_4,K_4} = 0$
+%                 $\cT^{6}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Thompson sampling &
+%             $\cT^{7}_{T_1,K_1} = 0$
+%                 $\cT^{7}_{T_1,K_2} = 0$
+%                 $\cT^{7}_{T_1,K_3} = 0$
+%                 $\cT^{7}_{T_1,K_4} = 0$
+%                 $\cT^{7}_{T_1,K_5} = 0$ &
+%             $\cT^{7}_{T_2,K_1} = 0$
+%                 $\cT^{7}_{T_2,K_2} = 0$
+%                 $\cT^{7}_{T_2,K_3} = 0$
+%                 $\cT^{7}_{T_2,K_4} = 0$
+%                 $\cT^{7}_{T_2,K_5} = 0$ &
+%             $\cT^{7}_{T_3,K_1} = 0$
+%                 $\cT^{7}_{T_3,K_2} = 0$
+%                 $\cT^{7}_{T_3,K_3} = 0$
+%                 $\cT^{7}_{T_3,K_4} = 0$
+%                 $\cT^{7}_{T_3,K_5} = 0$ &
+%             $\cT^{7}_{T_4,K_1} = 0$
+%                 $\cT^{7}_{T_4,K_2} = 0$
+%                 $\cT^{7}_{T_4,K_3} = 0$
+%                 $\cT^{7}_{T_4,K_4} = 0$
+%                 $\cT^{7}_{T_4,K_5} = 0$ \\
+%         \hline
+%         AdBandits &
+%             $\cT^{8}_{T_1,K_1} = 0$
+%                 $\cT^{8}_{T_1,K_2} = 0$
+%                 $\cT^{8}_{T_1,K_3} = 0$
+%                 $\cT^{8}_{T_1,K_4} = 0$
+%                 $\cT^{8}_{T_1,K_5} = 0$ &
+%             $\cT^{8}_{T_2,K_1} = 0$
+%                 $\cT^{8}_{T_2,K_2} = 0$
+%                 $\cT^{8}_{T_2,K_3} = 0$
+%                 $\cT^{8}_{T_2,K_4} = 0$
+%                 $\cT^{8}_{T_2,K_5} = 0$ &
+%             $\cT^{8}_{T_3,K_1} = 0$
+%                 $\cT^{8}_{T_3,K_2} = 0$
+%                 $\cT^{8}_{T_3,K_3} = 0$
+%                 $\cT^{8}_{T_3,K_4} = 0$
+%                 $\cT^{8}_{T_3,K_5} = 0$ &
+%             $\cT^{8}_{T_4,K_1} = 0$
+%                 $\cT^{8}_{T_4,K_2} = 0$
+%                 $\cT^{8}_{T_4,K_3} = 0$
+%                 $\cT^{8}_{T_4,K_4} = 0$
+%                 $\cT^{8}_{T_4,K_5} = 0$ \\
+%         \hline
+%         BESA &
+%             $\cT^{9}_{T_1,K_1} = 0$
+%                 $\cT^{9}_{T_1,K_2} = 0$
+%                 $\cT^{9}_{T_1,K_3} = 0$
+%                 $\cT^{9}_{T_1,K_4} = 0$
+%                 $\cT^{9}_{T_1,K_5} = 0$ &
+%             $\cT^{9}_{T_2,K_1} = 0$
+%                 $\cT^{9}_{T_2,K_2} = 0$
+%                 $\cT^{9}_{T_2,K_3} = 0$
+%                 $\cT^{9}_{T_2,K_4} = 0$
+%                 $\cT^{9}_{T_2,K_5} = 0$ &
+%             $\cT^{9}_{T_3,K_1} = 0$
+%                 $\cT^{9}_{T_3,K_2} = 0$
+%                 $\cT^{9}_{T_3,K_3} = 0$
+%                 $\cT^{9}_{T_3,K_4} = 0$
+%                 $\cT^{9}_{T_3,K_5} = 0$ &
+%             $\cT^{9}_{T_4,K_1} = 0$
+%                 $\cT^{9}_{T_4,K_2} = 0$
+%                 $\cT^{9}_{T_4,K_3} = 0$
+%                 $\cT^{9}_{T_4,K_4} = 0$
+%                 $\cT^{9}_{T_4,K_5} = 0$ \\
+%         \hline
+%     \end{tabular}
+%     \caption{Normalized running time, rounded and in milli-seconds, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+%     \label{table:3:time_problem1}
+% \end{footnotesize}  % WARNING
+% \end{table}
 
@@
 
-\begin{table}[!t]
-\begin{footnotesize}  % WARNING
-    \centering
-    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-        \hline
-        $\varepsilon$-greedy &
-            $\cM^{1}_{T_1,K_1} = 0$
-                $\cM^{1}_{T_1,K_2} = 0$
-                $\cM^{1}_{T_1,K_3} = 0$
-                $\cM^{1}_{T_1,K_4} = 0$
-                $\cM^{1}_{T_1,K_5} = 0$ &
-            $\cM^{1}_{T_2,K_1} = 0$
-                $\cM^{1}_{T_2,K_2} = 0$
-                $\cM^{1}_{T_2,K_3} = 0$
-                $\cM^{1}_{T_2,K_4} = 0$
-                $\cM^{1}_{T_2,K_5} = 0$ &
-            $\cM^{1}_{T_3,K_1} = 0$
-                $\cM^{1}_{T_3,K_2} = 0$
-                $\cM^{1}_{T_3,K_3} = 0$
-                $\cM^{1}_{T_3,K_4} = 0$
-                $\cM^{1}_{T_3,K_5} = 0$ &
-            $\cM^{1}_{T_4,K_1} = 0$
-                $\cM^{1}_{T_4,K_2} = 0$
-                $\cM^{1}_{T_4,K_3} = 0$
-                $\cM^{1}_{T_4,K_4} = 0$
-                $\cM^{1}_{T_4,K_5} = 0$ \\
-        \hline
-        Explore-then-Commit &
-            $\cM^{2}_{T_1,K_1} = 0$
-                $\cM^{2}_{T_1,K_2} = 0$
-                $\cM^{2}_{T_1,K_3} = 0$
-                $\cM^{2}_{T_1,K_4} = 0$
-                $\cM^{2}_{T_1,K_5} = 0$ &
-            $\cM^{2}_{T_2,K_1} = 0$
-                $\cM^{2}_{T_2,K_2} = 0$
-                $\cM^{2}_{T_2,K_3} = 0$
-                $\cM^{2}_{T_2,K_4} = 0$
-                $\cM^{2}_{T_2,K_5} = 0$ &
-            $\cM^{2}_{T_3,K_1} = 0$
-                $\cM^{2}_{T_3,K_2} = 0$
-                $\cM^{2}_{T_3,K_3} = 0$
-                $\cM^{2}_{T_3,K_4} = 0$
-                $\cM^{2}_{T_3,K_5} = 0$ &
-            $\cM^{2}_{T_4,K_1} = 0$
-                $\cM^{2}_{T_4,K_2} = 0$
-                $\cM^{2}_{T_4,K_3} = 0$
-                $\cM^{2}_{T_4,K_4} = 0$
-                $\cM^{2}_{T_4,K_5} = 0$ \\
-        \hline
-        $\mathrm{Exp}3^{++}$ &
-            $\cM^3_{T_1,K_1} = 0$
-                $\cM^3_{T_1,K_2} = 0$
-                $\cM^3_{T_1,K_3} = 0$
-                $\cM^3_{T_1,K_4} = 0$
-                $\cM^3_{T_1,K_5} = 0$ &
-            $\cM^3_{T_2,K_1} = 0$
-                $\cM^3_{T_2,K_2} = 0$
-                $\cM^3_{T_2,K_3} = 0$
-                $\cM^3_{T_2,K_4} = 0$
-                $\cM^3_{T_2,K_5} = 0$ &
-            $\cM^3_{T_3,K_1} = 0$
-                $\cM^3_{T_3,K_2} = 0$
-                $\cM^3_{T_3,K_3} = 0$
-                $\cM^3_{T_3,K_4} = 0$
-                $\cM^3_{T_3,K_5} = 0$ &
-            $\cM^3_{T_4,K_1} = 0$
-                $\cM^3_{T_4,K_2} = 0$
-                $\cM^3_{T_4,K_3} = 0$
-                $\cM^3_{T_4,K_4} = 0$
-                $\cM^3_{T_4,K_5} = 0$ \\
-        \hline
-        $\UCB_1$ &
-            $\cM^{4}_{T_1,K_1} = 0$
-                $\cM^{4}_{T_1,K_2} = 0$
-                $\cM^{4}_{T_1,K_3} = 0$
-                $\cM^{4}_{T_1,K_4} = 0$
-                $\cM^{4}_{T_1,K_5} = 0$ &
-            $\cM^{4}_{T_2,K_1} = 0$
-                $\cM^{4}_{T_2,K_2} = 0$
-                $\cM^{4}_{T_2,K_3} = 0$
-                $\cM^{4}_{T_2,K_4} = 0$
-                $\cM^{4}_{T_2,K_5} = 0$ &
-            $\cM^{4}_{T_3,K_1} = 0$
-                $\cM^{4}_{T_3,K_2} = 0$
-                $\cM^{4}_{T_3,K_3} = 0$
-                $\cM^{4}_{T_3,K_4} = 0$
-                $\cM^{4}_{T_3,K_5} = 0$ &
-            $\cM^{4}_{T_4,K_1} = 0$
-                $\cM^{4}_{T_4,K_2} = 0$
-                $\cM^{4}_{T_4,K_3} = 0$
-                $\cM^{4}_{T_4,K_4} = 0$
-                $\cM^{4}_{T_4,K_5} = 0$ \\
-        \hline
-        \klUCB{} &
-            $\cM^{5}_{T_1,K_1} = 0$
-                $\cM^{5}_{T_1,K_2} = 0$
-                $\cM^{5}_{T_1,K_3} = 0$
-                $\cM^{5}_{T_1,K_4} = 0$
-                $\cM^{5}_{T_1,K_5} = 0$ &
-            $\cM^{5}_{T_2,K_1} = 0$
-                $\cM^{5}_{T_2,K_2} = 0$
-                $\cM^{5}_{T_2,K_3} = 0$
-                $\cM^{5}_{T_2,K_4} = 0$
-                $\cM^{5}_{T_2,K_5} = 0$ &
-            $\cM^{5}_{T_3,K_1} = 0$
-                $\cM^{5}_{T_3,K_2} = 0$
-                $\cM^{5}_{T_3,K_3} = 0$
-                $\cM^{5}_{T_3,K_4} = 0$
-                $\cM^{5}_{T_3,K_5} = 0$ &
-            $\cM^{5}_{T_4,K_1} = 0$
-                $\cM^{5}_{T_4,K_2} = 0$
-                $\cM^{5}_{T_4,K_3} = 0$
-                $\cM^{5}_{T_4,K_4} = 0$
-                $\cM^{5}_{T_4,K_5} = 0$ \\
-        \hline
-        Thompson sampling &
-            $\cM^{6}_{T_1,K_1} = 0$
-                $\cM^{6}_{T_1,K_2} = 0$
-                $\cM^{6}_{T_1,K_3} = 0$
-                $\cM^{6}_{T_1,K_4} = 0$
-                $\cM^{6}_{T_1,K_5} = 0$ &
-            $\cM^{6}_{T_2,K_1} = 0$
-                $\cM^{6}_{T_2,K_2} = 0$
-                $\cM^{6}_{T_2,K_3} = 0$
-                $\cM^{6}_{T_2,K_4} = 0$
-                $\cM^{6}_{T_2,K_5} = 0$ &
-            $\cM^{6}_{T_3,K_1} = 0$
-                $\cM^{6}_{T_3,K_2} = 0$
-                $\cM^{6}_{T_3,K_3} = 0$
-                $\cM^{6}_{T_3,K_4} = 0$
-                $\cM^{6}_{T_3,K_5} = 0$ &
-            $\cM^{6}_{T_4,K_1} = 0$
-                $\cM^{6}_{T_4,K_2} = 0$
-                $\cM^{6}_{T_4,K_3} = 0$
-                $\cM^{6}_{T_4,K_4} = 0$
-                $\cM^{6}_{T_4,K_5} = 0$ \\
-        \hline
-        Bayes-UCB &
-            $\cM^{7}_{T_1,K_1} = 0$
-                $\cM^{7}_{T_1,K_2} = 0$
-                $\cM^{7}_{T_1,K_3} = 0$
-                $\cM^{7}_{T_1,K_4} = 0$
-                $\cM^{7}_{T_1,K_5} = 0$ &
-            $\cM^{7}_{T_2,K_1} = 0$
-                $\cM^{7}_{T_2,K_2} = 0$
-                $\cM^{7}_{T_2,K_3} = 0$
-                $\cM^{7}_{T_2,K_4} = 0$
-                $\cM^{7}_{T_2,K_5} = 0$ &
-            $\cM^{7}_{T_3,K_1} = 0$
-                $\cM^{7}_{T_3,K_2} = 0$
-                $\cM^{7}_{T_3,K_3} = 0$
-                $\cM^{7}_{T_3,K_4} = 0$
-                $\cM^{7}_{T_3,K_5} = 0$ &
-            $\cM^{7}_{T_4,K_1} = 0$
-                $\cM^{7}_{T_4,K_2} = 0$
-                $\cM^{7}_{T_4,K_3} = 0$
-                $\cM^{7}_{T_4,K_4} = 0$
-                $\cM^{7}_{T_4,K_5} = 0$ \\
-        \hline
-        AdBandits &
-            $\cM^{8}_{T_1,K_1} = 0$
-                $\cM^{8}_{T_1,K_2} = 0$
-                $\cM^{8}_{T_1,K_3} = 0$
-                $\cM^{8}_{T_1,K_4} = 0$
-                $\cM^{8}_{T_1,K_5} = 0$ &
-            $\cM^{8}_{T_2,K_1} = 0$
-                $\cM^{8}_{T_2,K_2} = 0$
-                $\cM^{8}_{T_2,K_3} = 0$
-                $\cM^{8}_{T_2,K_4} = 0$
-                $\cM^{8}_{T_2,K_5} = 0$ &
-            $\cM^{8}_{T_3,K_1} = 0$
-                $\cM^{8}_{T_3,K_2} = 0$
-                $\cM^{8}_{T_3,K_3} = 0$
-                $\cM^{8}_{T_3,K_4} = 0$
-                $\cM^{8}_{T_3,K_5} = 0$ &
-            $\cM^{8}_{T_4,K_1} = 0$
-                $\cM^{8}_{T_4,K_2} = 0$
-                $\cM^{8}_{T_4,K_3} = 0$
-                $\cM^{8}_{T_4,K_4} = 0$
-                $\cM^{8}_{T_4,K_5} = 0$ \\
-        \hline
-        BESA &
-            $\cM^{9}_{T_1,K_1} = 0$
-                $\cM^{9}_{T_1,K_2} = 0$
-                $\cM^{9}_{T_1,K_3} = 0$
-                $\cM^{9}_{T_1,K_4} = 0$
-                $\cM^{9}_{T_1,K_5} = 0$ &
-            $\cM^{9}_{T_2,K_1} = 0$
-                $\cM^{9}_{T_2,K_2} = 0$
-                $\cM^{9}_{T_2,K_3} = 0$
-                $\cM^{9}_{T_2,K_4} = 0$
-                $\cM^{9}_{T_2,K_5} = 0$ &
-            $\cM^{9}_{T_3,K_1} = 0$
-                $\cM^{9}_{T_3,K_2} = 0$
-                $\cM^{9}_{T_3,K_3} = 0$
-                $\cM^{9}_{T_3,K_4} = 0$
-                $\cM^{9}_{T_3,K_5} = 0$ &
-            $\cM^{9}_{T_4,K_1} = 0$
-                $\cM^{9}_{T_4,K_2} = 0$
-                $\cM^{9}_{T_4,K_3} = 0$
-                $\cM^{9}_{T_4,K_4} = 0$
-                $\cM^{9}_{T_4,K_5} = 0$ \\
-        \hline
-    \end{tabular}
-    \caption{Non-normalized memory cost, rounded and in Bytes, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
-    \label{table:3:memory_problem1}
-\end{footnotesize}  % WARNING
-\end{table}
+% FIXME generate from the notebook then copy from WS3
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normmemory_vs_arms__56pb__7Ks_T50000.pdf}
+	\caption[Normalized memory cost vs different values of $K$.]{
+        Normalized memory cost vs different values of $K$ ($\cM_T / K$),
+        for $T=50000$ and for $16$ algorithms.
+        % $y$-axis is in log-scale.
+        All algorithms (except BESA) has an almost constant memory cost, meaning that for $K$ arms they use a storage proportional to $K$, as predicted.
+	}
+	\label{fig:3:16_different_algorithms__normmemory_vs_arms__56pb__7Ks_T50000}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normmemory_vs_horizons__56pb__K32_7Ts.pdf}
+	\caption[Normalized memory cost vs different values of $T$.]{
+        Normalized memory cost vs different values of $T$ ($\cM_T / K$),
+        for $K=32$ and for $16$ algorithms.
+        All algorithms (except BESA) has a very slowly growing normalized memory cost, meaning that for $T$ rounds they use a computation time almost independent of $T$, as predicted.
+        There is a clear separation between three groups.
+        First, BESA is very costly in terms of memory, and $\UCB\dagger$ and \textsc{Tsallis-Inf} are also more costly than others, and both experiences a linear growth of their memory cost.
+        Then, we see PHE, $\mathrm{Exp}3++$, Bayes-UCB and AdBandits having a small but linear growth of memory cost, and it can be explained for each of them (\eg, for Bayes-UCB, computing the $1-1/t$ quantile at time $t$ becomes more and more costly as $t$ grows, and thus as $T$ grows,  ).
+        Finally, the other algorithms (including \UCB, \klUCB{} and Thompson sampling) have a very small footprint and their memory cost do not grow much when $T$ increases.
+	}
+	\label{fig:3:16_different_algorithms__normmemory_vs_horizons__56pb__K32_7Ts}
+\end{figure}
+
+% \begin{table}[!t]
+% \begin{footnotesize}  % WARNING
+%     \centering
+%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+%         \hline
+%         $\varepsilon$-greedy &
+%             $\cM^{1}_{T_1,K_1} = 0$
+%                 $\cM^{1}_{T_1,K_2} = 0$
+%                 $\cM^{1}_{T_1,K_3} = 0$
+%                 $\cM^{1}_{T_1,K_4} = 0$
+%                 $\cM^{1}_{T_1,K_5} = 0$ &
+%             $\cM^{1}_{T_2,K_1} = 0$
+%                 $\cM^{1}_{T_2,K_2} = 0$
+%                 $\cM^{1}_{T_2,K_3} = 0$
+%                 $\cM^{1}_{T_2,K_4} = 0$
+%                 $\cM^{1}_{T_2,K_5} = 0$ &
+%             $\cM^{1}_{T_3,K_1} = 0$
+%                 $\cM^{1}_{T_3,K_2} = 0$
+%                 $\cM^{1}_{T_3,K_3} = 0$
+%                 $\cM^{1}_{T_3,K_4} = 0$
+%                 $\cM^{1}_{T_3,K_5} = 0$ &
+%             $\cM^{1}_{T_4,K_1} = 0$
+%                 $\cM^{1}_{T_4,K_2} = 0$
+%                 $\cM^{1}_{T_4,K_3} = 0$
+%                 $\cM^{1}_{T_4,K_4} = 0$
+%                 $\cM^{1}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Explore-then-Commit &
+%             $\cM^{2}_{T_1,K_1} = 0$
+%                 $\cM^{2}_{T_1,K_2} = 0$
+%                 $\cM^{2}_{T_1,K_3} = 0$
+%                 $\cM^{2}_{T_1,K_4} = 0$
+%                 $\cM^{2}_{T_1,K_5} = 0$ &
+%             $\cM^{2}_{T_2,K_1} = 0$
+%                 $\cM^{2}_{T_2,K_2} = 0$
+%                 $\cM^{2}_{T_2,K_3} = 0$
+%                 $\cM^{2}_{T_2,K_4} = 0$
+%                 $\cM^{2}_{T_2,K_5} = 0$ &
+%             $\cM^{2}_{T_3,K_1} = 0$
+%                 $\cM^{2}_{T_3,K_2} = 0$
+%                 $\cM^{2}_{T_3,K_3} = 0$
+%                 $\cM^{2}_{T_3,K_4} = 0$
+%                 $\cM^{2}_{T_3,K_5} = 0$ &
+%             $\cM^{2}_{T_4,K_1} = 0$
+%                 $\cM^{2}_{T_4,K_2} = 0$
+%                 $\cM^{2}_{T_4,K_3} = 0$
+%                 $\cM^{2}_{T_4,K_4} = 0$
+%                 $\cM^{2}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\mathrm{Exp}3^{++}$ &
+%             $\cM^3_{T_1,K_1} = 0$
+%                 $\cM^3_{T_1,K_2} = 0$
+%                 $\cM^3_{T_1,K_3} = 0$
+%                 $\cM^3_{T_1,K_4} = 0$
+%                 $\cM^3_{T_1,K_5} = 0$ &
+%             $\cM^3_{T_2,K_1} = 0$
+%                 $\cM^3_{T_2,K_2} = 0$
+%                 $\cM^3_{T_2,K_3} = 0$
+%                 $\cM^3_{T_2,K_4} = 0$
+%                 $\cM^3_{T_2,K_5} = 0$ &
+%             $\cM^3_{T_3,K_1} = 0$
+%                 $\cM^3_{T_3,K_2} = 0$
+%                 $\cM^3_{T_3,K_3} = 0$
+%                 $\cM^3_{T_3,K_4} = 0$
+%                 $\cM^3_{T_3,K_5} = 0$ &
+%             $\cM^3_{T_4,K_1} = 0$
+%                 $\cM^3_{T_4,K_2} = 0$
+%                 $\cM^3_{T_4,K_3} = 0$
+%                 $\cM^3_{T_4,K_4} = 0$
+%                 $\cM^3_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\UCB_1$ &
+%             $\cM^{4}_{T_1,K_1} = 0$
+%                 $\cM^{4}_{T_1,K_2} = 0$
+%                 $\cM^{4}_{T_1,K_3} = 0$
+%                 $\cM^{4}_{T_1,K_4} = 0$
+%                 $\cM^{4}_{T_1,K_5} = 0$ &
+%             $\cM^{4}_{T_2,K_1} = 0$
+%                 $\cM^{4}_{T_2,K_2} = 0$
+%                 $\cM^{4}_{T_2,K_3} = 0$
+%                 $\cM^{4}_{T_2,K_4} = 0$
+%                 $\cM^{4}_{T_2,K_5} = 0$ &
+%             $\cM^{4}_{T_3,K_1} = 0$
+%                 $\cM^{4}_{T_3,K_2} = 0$
+%                 $\cM^{4}_{T_3,K_3} = 0$
+%                 $\cM^{4}_{T_3,K_4} = 0$
+%                 $\cM^{4}_{T_3,K_5} = 0$ &
+%             $\cM^{4}_{T_4,K_1} = 0$
+%                 $\cM^{4}_{T_4,K_2} = 0$
+%                 $\cM^{4}_{T_4,K_3} = 0$
+%                 $\cM^{4}_{T_4,K_4} = 0$
+%                 $\cM^{4}_{T_4,K_5} = 0$ \\
+%         \hline
+%         \klUCB{} &
+%             $\cM^{5}_{T_1,K_1} = 0$
+%                 $\cM^{5}_{T_1,K_2} = 0$
+%                 $\cM^{5}_{T_1,K_3} = 0$
+%                 $\cM^{5}_{T_1,K_4} = 0$
+%                 $\cM^{5}_{T_1,K_5} = 0$ &
+%             $\cM^{5}_{T_2,K_1} = 0$
+%                 $\cM^{5}_{T_2,K_2} = 0$
+%                 $\cM^{5}_{T_2,K_3} = 0$
+%                 $\cM^{5}_{T_2,K_4} = 0$
+%                 $\cM^{5}_{T_2,K_5} = 0$ &
+%             $\cM^{5}_{T_3,K_1} = 0$
+%                 $\cM^{5}_{T_3,K_2} = 0$
+%                 $\cM^{5}_{T_3,K_3} = 0$
+%                 $\cM^{5}_{T_3,K_4} = 0$
+%                 $\cM^{5}_{T_3,K_5} = 0$ &
+%             $\cM^{5}_{T_4,K_1} = 0$
+%                 $\cM^{5}_{T_4,K_2} = 0$
+%                 $\cM^{5}_{T_4,K_3} = 0$
+%                 $\cM^{5}_{T_4,K_4} = 0$
+%                 $\cM^{5}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Thompson sampling &
+%             $\cM^{6}_{T_1,K_1} = 0$
+%                 $\cM^{6}_{T_1,K_2} = 0$
+%                 $\cM^{6}_{T_1,K_3} = 0$
+%                 $\cM^{6}_{T_1,K_4} = 0$
+%                 $\cM^{6}_{T_1,K_5} = 0$ &
+%             $\cM^{6}_{T_2,K_1} = 0$
+%                 $\cM^{6}_{T_2,K_2} = 0$
+%                 $\cM^{6}_{T_2,K_3} = 0$
+%                 $\cM^{6}_{T_2,K_4} = 0$
+%                 $\cM^{6}_{T_2,K_5} = 0$ &
+%             $\cM^{6}_{T_3,K_1} = 0$
+%                 $\cM^{6}_{T_3,K_2} = 0$
+%                 $\cM^{6}_{T_3,K_3} = 0$
+%                 $\cM^{6}_{T_3,K_4} = 0$
+%                 $\cM^{6}_{T_3,K_5} = 0$ &
+%             $\cM^{6}_{T_4,K_1} = 0$
+%                 $\cM^{6}_{T_4,K_2} = 0$
+%                 $\cM^{6}_{T_4,K_3} = 0$
+%                 $\cM^{6}_{T_4,K_4} = 0$
+%                 $\cM^{6}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Bayes-UCB &
+%             $\cM^{7}_{T_1,K_1} = 0$
+%                 $\cM^{7}_{T_1,K_2} = 0$
+%                 $\cM^{7}_{T_1,K_3} = 0$
+%                 $\cM^{7}_{T_1,K_4} = 0$
+%                 $\cM^{7}_{T_1,K_5} = 0$ &
+%             $\cM^{7}_{T_2,K_1} = 0$
+%                 $\cM^{7}_{T_2,K_2} = 0$
+%                 $\cM^{7}_{T_2,K_3} = 0$
+%                 $\cM^{7}_{T_2,K_4} = 0$
+%                 $\cM^{7}_{T_2,K_5} = 0$ &
+%             $\cM^{7}_{T_3,K_1} = 0$
+%                 $\cM^{7}_{T_3,K_2} = 0$
+%                 $\cM^{7}_{T_3,K_3} = 0$
+%                 $\cM^{7}_{T_3,K_4} = 0$
+%                 $\cM^{7}_{T_3,K_5} = 0$ &
+%             $\cM^{7}_{T_4,K_1} = 0$
+%                 $\cM^{7}_{T_4,K_2} = 0$
+%                 $\cM^{7}_{T_4,K_3} = 0$
+%                 $\cM^{7}_{T_4,K_4} = 0$
+%                 $\cM^{7}_{T_4,K_5} = 0$ \\
+%         \hline
+%         AdBandits &
+%             $\cM^{8}_{T_1,K_1} = 0$
+%                 $\cM^{8}_{T_1,K_2} = 0$
+%                 $\cM^{8}_{T_1,K_3} = 0$
+%                 $\cM^{8}_{T_1,K_4} = 0$
+%                 $\cM^{8}_{T_1,K_5} = 0$ &
+%             $\cM^{8}_{T_2,K_1} = 0$
+%                 $\cM^{8}_{T_2,K_2} = 0$
+%                 $\cM^{8}_{T_2,K_3} = 0$
+%                 $\cM^{8}_{T_2,K_4} = 0$
+%                 $\cM^{8}_{T_2,K_5} = 0$ &
+%             $\cM^{8}_{T_3,K_1} = 0$
+%                 $\cM^{8}_{T_3,K_2} = 0$
+%                 $\cM^{8}_{T_3,K_3} = 0$
+%                 $\cM^{8}_{T_3,K_4} = 0$
+%                 $\cM^{8}_{T_3,K_5} = 0$ &
+%             $\cM^{8}_{T_4,K_1} = 0$
+%                 $\cM^{8}_{T_4,K_2} = 0$
+%                 $\cM^{8}_{T_4,K_3} = 0$
+%                 $\cM^{8}_{T_4,K_4} = 0$
+%                 $\cM^{8}_{T_4,K_5} = 0$ \\
+%         \hline
+%         BESA &
+%             $\cM^{9}_{T_1,K_1} = 0$
+%                 $\cM^{9}_{T_1,K_2} = 0$
+%                 $\cM^{9}_{T_1,K_3} = 0$
+%                 $\cM^{9}_{T_1,K_4} = 0$
+%                 $\cM^{9}_{T_1,K_5} = 0$ &
+%             $\cM^{9}_{T_2,K_1} = 0$
+%                 $\cM^{9}_{T_2,K_2} = 0$
+%                 $\cM^{9}_{T_2,K_3} = 0$
+%                 $\cM^{9}_{T_2,K_4} = 0$
+%                 $\cM^{9}_{T_2,K_5} = 0$ &
+%             $\cM^{9}_{T_3,K_1} = 0$
+%                 $\cM^{9}_{T_3,K_2} = 0$
+%                 $\cM^{9}_{T_3,K_3} = 0$
+%                 $\cM^{9}_{T_3,K_4} = 0$
+%                 $\cM^{9}_{T_3,K_5} = 0$ &
+%             $\cM^{9}_{T_4,K_1} = 0$
+%                 $\cM^{9}_{T_4,K_2} = 0$
+%                 $\cM^{9}_{T_4,K_3} = 0$
+%                 $\cM^{9}_{T_4,K_4} = 0$
+%                 $\cM^{9}_{T_4,K_5} = 0$ \\
+%         \hline
+%     \end{tabular}
+%     \caption{Non-normalized memory cost, rounded and in Bytes, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+%     \label{table:3:memory_problem1}
+% \end{footnotesize}  % WARNING
+% \end{table}
 
@@
 
-\paragraph{Mean regret}
-
-\begin{table}[!t]
-\begin{footnotesize}  % WARNING
-    \centering
-    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-        $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
-            $R^{10}_{T_1,K_1} = 0 \pm 0$
-                $R^{10}_{T_1,K_2} = 0 \pm 0$
-                $R^{10}_{T_1,K_3} = 0 \pm 0$
-                $R^{10}_{T_1,K_4} = 0 \pm 0$
-                $R^{10}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{10}_{T_2,K_1} = 0 \pm 0$
-                $R^{10}_{T_2,K_2} = 0 \pm 0$
-                $R^{10}_{T_2,K_3} = 0 \pm 0$
-                $R^{10}_{T_2,K_4} = 0 \pm 0$
-                $R^{10}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{10}_{T_3,K_1} = 0 \pm 0$
-                $R^{10}_{T_3,K_2} = 0 \pm 0$
-                $R^{10}_{T_3,K_3} = 0 \pm 0$
-                $R^{10}_{T_3,K_4} = 0 \pm 0$
-                $R^{10}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{10}_{T_4,K_1} = 0 \pm 0$
-                $R^{10}_{T_4,K_2} = 0 \pm 0$
-                $R^{10}_{T_4,K_3} = 0 \pm 0$
-                $R^{10}_{T_4,K_4} = 0 \pm 0$
-                $R^{10}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        Appr. F-H Gittins &
-            $R^{11}_{T_1,K_1} = 0 \pm 0$
-                $R^{11}_{T_1,K_2} = 0 \pm 0$
-                $R^{11}_{T_1,K_3} = 0 \pm 0$
-                $R^{11}_{T_1,K_4} = 0 \pm 0$
-                $R^{11}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{11}_{T_2,K_1} = 0 \pm 0$
-                $R^{11}_{T_2,K_2} = 0 \pm 0$
-                $R^{11}_{T_2,K_3} = 0 \pm 0$
-                $R^{11}_{T_2,K_4} = 0 \pm 0$
-                $R^{11}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{11}_{T_3,K_1} = 0 \pm 0$
-                $R^{11}_{T_3,K_2} = 0 \pm 0$
-                $R^{11}_{T_3,K_3} = 0 \pm 0$
-                $R^{11}_{T_3,K_4} = 0 \pm 0$
-                $R^{11}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{11}_{T_4,K_1} = 0 \pm 0$
-                $R^{11}_{T_4,K_2} = 0 \pm 0$
-                $R^{11}_{T_4,K_3} = 0 \pm 0$
-                $R^{11}_{T_4,K_4} = 0 \pm 0$
-                $R^{11}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        $\UCB\dagger$ &
-            $R^{12}_{T_1,K_1} = 0 \pm 0$
-                $R^{12}_{T_1,K_2} = 0 \pm 0$
-                $R^{12}_{T_1,K_3} = 0 \pm 0$
-                $R^{12}_{T_1,K_4} = 0 \pm 0$
-                $R^{12}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{12}_{T_2,K_1} = 0 \pm 0$
-                $R^{12}_{T_2,K_2} = 0 \pm 0$
-                $R^{12}_{T_2,K_3} = 0 \pm 0$
-                $R^{12}_{T_2,K_4} = 0 \pm 0$
-                $R^{12}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{12}_{T_3,K_1} = 0 \pm 0$
-                $R^{12}_{T_3,K_2} = 0 \pm 0$
-                $R^{12}_{T_3,K_3} = 0 \pm 0$
-                $R^{12}_{T_3,K_4} = 0 \pm 0$
-                $R^{12}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{12}_{T_4,K_1} = 0 \pm 0$
-                $R^{12}_{T_4,K_2} = 0 \pm 0$
-                $R^{12}_{T_4,K_3} = 0 \pm 0$
-                $R^{12}_{T_4,K_4} = 0 \pm 0$
-                $R^{12}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        KL-UCB-Switch &
-            $R^{13}_{T_1,K_1} = 0 \pm 0$
-                $R^{13}_{T_1,K_2} = 0 \pm 0$
-                $R^{13}_{T_1,K_3} = 0 \pm 0$
-                $R^{13}_{T_1,K_4} = 0 \pm 0$
-                $R^{13}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{13}_{T_2,K_1} = 0 \pm 0$
-                $R^{13}_{T_2,K_2} = 0 \pm 0$
-                $R^{13}_{T_2,K_3} = 0 \pm 0$
-                $R^{13}_{T_2,K_4} = 0 \pm 0$
-                $R^{13}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{13}_{T_3,K_1} = 0 \pm 0$
-                $R^{13}_{T_3,K_2} = 0 \pm 0$
-                $R^{13}_{T_3,K_3} = 0 \pm 0$
-                $R^{13}_{T_3,K_4} = 0 \pm 0$
-                $R^{13}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{13}_{T_4,K_1} = 0 \pm 0$
-                $R^{13}_{T_4,K_2} = 0 \pm 0$
-                $R^{13}_{T_4,K_3} = 0 \pm 0$
-                $R^{13}_{T_4,K_4} = 0 \pm 0$
-                $R^{13}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        \textsc{Tsallis-Inf} &
-            $R^{14}_{T_1,K_1} = 0 \pm 0$
-                $R^{14}_{T_1,K_2} = 0 \pm 0$
-                $R^{14}_{T_1,K_3} = 0 \pm 0$
-                $R^{14}_{T_1,K_4} = 0 \pm 0$
-                $R^{14}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{14}_{T_2,K_1} = 0 \pm 0$
-                $R^{14}_{T_2,K_2} = 0 \pm 0$
-                $R^{14}_{T_2,K_3} = 0 \pm 0$
-                $R^{14}_{T_2,K_4} = 0 \pm 0$
-                $R^{14}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{14}_{T_3,K_1} = 0 \pm 0$
-                $R^{14}_{T_3,K_2} = 0 \pm 0$
-                $R^{14}_{T_3,K_3} = 0 \pm 0$
-                $R^{14}_{T_3,K_4} = 0 \pm 0$
-                $R^{14}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{14}_{T_4,K_1} = 0 \pm 0$
-                $R^{14}_{T_4,K_2} = 0 \pm 0$
-                $R^{14}_{T_4,K_3} = 0 \pm 0$
-                $R^{14}_{T_4,K_4} = 0 \pm 0$
-                $R^{14}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        $\mathrm{RCB}$ &
-            $R^{15}_{T_1,K_1} = 0 \pm 0$
-                $R^{15}_{T_1,K_2} = 0 \pm 0$
-                $R^{15}_{T_1,K_3} = 0 \pm 0$
-                $R^{15}_{T_1,K_4} = 0 \pm 0$
-                $R^{15}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{15}_{T_2,K_1} = 0 \pm 0$
-                $R^{15}_{T_2,K_2} = 0 \pm 0$
-                $R^{15}_{T_2,K_3} = 0 \pm 0$
-                $R^{15}_{T_2,K_4} = 0 \pm 0$
-                $R^{15}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{15}_{T_3,K_1} = 0 \pm 0$
-                $R^{15}_{T_3,K_2} = 0 \pm 0$
-                $R^{15}_{T_3,K_3} = 0 \pm 0$
-                $R^{15}_{T_3,K_4} = 0 \pm 0$
-                $R^{15}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{15}_{T_4,K_1} = 0 \pm 0$
-                $R^{15}_{T_4,K_2} = 0 \pm 0$
-                $R^{15}_{T_4,K_3} = 0 \pm 0$
-                $R^{15}_{T_4,K_4} = 0 \pm 0$
-                $R^{15}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-        $\mathrm{PHE}$ &
-            $R^{16}_{T_1,K_1} = 0 \pm 0$
-                $R^{16}_{T_1,K_2} = 0 \pm 0$
-                $R^{16}_{T_1,K_3} = 0 \pm 0$
-                $R^{16}_{T_1,K_4} = 0 \pm 0$
-                $R^{16}_{T_1,K_5} = 0 \pm 0$ &
-            $R^{16}_{T_2,K_1} = 0 \pm 0$
-                $R^{16}_{T_2,K_2} = 0 \pm 0$
-                $R^{16}_{T_2,K_3} = 0 \pm 0$
-                $R^{16}_{T_2,K_4} = 0 \pm 0$
-                $R^{16}_{T_2,K_5} = 0 \pm 0$ &
-            $R^{16}_{T_3,K_1} = 0 \pm 0$
-                $R^{16}_{T_3,K_2} = 0 \pm 0$
-                $R^{16}_{T_3,K_3} = 0 \pm 0$
-                $R^{16}_{T_3,K_4} = 0 \pm 0$
-                $R^{16}_{T_3,K_5} = 0 \pm 0$ &
-            $R^{16}_{T_4,K_1} = 0 \pm 0$
-                $R^{16}_{T_4,K_2} = 0 \pm 0$
-                $R^{16}_{T_4,K_3} = 0 \pm 0$
-                $R^{16}_{T_4,K_4} = 0 \pm 0$
-                $R^{16}_{T_4,K_5} = 0 \pm 0$ \\
-        \hline
-    \end{tabular}
-    \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on problem $1$ with different values of time $T$ and number of arm $K$.
-    }
-    \label{table:3:meanRegret_problem1_otherAlgorithms}
-\end{footnotesize}  % WARNING
-\end{table}
-
-\paragraph{Computation time}
+% \paragraph{Mean regret}
+
+% \begin{table}[!t]
+% \begin{footnotesize}  % WARNING
+%     \centering
+%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+%         $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
+%             $R^{10}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{10}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{10}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{10}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{10}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{10}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{10}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{10}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{10}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{10}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{10}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{10}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{10}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{10}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{10}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{10}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{10}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{10}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{10}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{10}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         Appr. F-H Gittins &
+%             $R^{11}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{11}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{11}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{11}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{11}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{11}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{11}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{11}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{11}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{11}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{11}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{11}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{11}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{11}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{11}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{11}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{11}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{11}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{11}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{11}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         $\UCB\dagger$ &
+%             $R^{12}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{12}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{12}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{12}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{12}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{12}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{12}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{12}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{12}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{12}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{12}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{12}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{12}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{12}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{12}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{12}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{12}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{12}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{12}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{12}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         KL-UCB-Switch &
+%             $R^{13}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{13}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{13}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{13}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{13}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{13}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{13}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{13}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{13}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{13}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{13}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{13}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{13}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{13}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{13}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{13}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{13}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{13}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{13}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{13}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         \textsc{Tsallis-Inf} &
+%             $R^{14}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{14}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{14}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{14}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{14}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{14}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{14}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{14}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{14}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{14}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{14}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{14}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{14}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{14}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{14}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{14}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{14}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{14}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{14}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{14}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         $\mathrm{RCB}$ &
+%             $R^{15}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{15}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{15}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{15}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{15}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{15}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{15}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{15}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{15}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{15}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{15}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{15}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{15}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{15}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{15}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{15}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{15}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{15}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{15}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{15}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%         $\mathrm{PHE}$ &
+%             $R^{16}_{T_1,K_1} = 0 \pm 0$
+%                 $R^{16}_{T_1,K_2} = 0 \pm 0$
+%                 $R^{16}_{T_1,K_3} = 0 \pm 0$
+%                 $R^{16}_{T_1,K_4} = 0 \pm 0$
+%                 $R^{16}_{T_1,K_5} = 0 \pm 0$ &
+%             $R^{16}_{T_2,K_1} = 0 \pm 0$
+%                 $R^{16}_{T_2,K_2} = 0 \pm 0$
+%                 $R^{16}_{T_2,K_3} = 0 \pm 0$
+%                 $R^{16}_{T_2,K_4} = 0 \pm 0$
+%                 $R^{16}_{T_2,K_5} = 0 \pm 0$ &
+%             $R^{16}_{T_3,K_1} = 0 \pm 0$
+%                 $R^{16}_{T_3,K_2} = 0 \pm 0$
+%                 $R^{16}_{T_3,K_3} = 0 \pm 0$
+%                 $R^{16}_{T_3,K_4} = 0 \pm 0$
+%                 $R^{16}_{T_3,K_5} = 0 \pm 0$ &
+%             $R^{16}_{T_4,K_1} = 0 \pm 0$
+%                 $R^{16}_{T_4,K_2} = 0 \pm 0$
+%                 $R^{16}_{T_4,K_3} = 0 \pm 0$
+%                 $R^{16}_{T_4,K_4} = 0 \pm 0$
+%                 $R^{16}_{T_4,K_5} = 0 \pm 0$ \\
+%         \hline
+%     \end{tabular}
+%     \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on problem $1$ with different values of time $T$ and number of arm $K$.
+%     }
+%     \label{table:3:meanRegret_problem1_otherAlgorithms}
+% \end{footnotesize}  % WARNING
+% \end{table}
+
+% \paragraph{Computation time}
+
+% \TODOL{A faire}
+
+% \begin{table}[!t]
+% \begin{footnotesize}  % WARNING
+%     \centering
+%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+%         $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
+%             $\cT^{10}_{T_1,K_1} = 0$
+%                 $\cT^{10}_{T_1,K_2} = 0$
+%                 $\cT^{10}_{T_1,K_3} = 0$
+%                 $\cT^{10}_{T_1,K_4} = 0$
+%                 $\cT^{10}_{T_1,K_5} = 0$ &
+%             $\cT^{10}_{T_2,K_1} = 0$
+%                 $\cT^{10}_{T_2,K_2} = 0$
+%                 $\cT^{10}_{T_2,K_3} = 0$
+%                 $\cT^{10}_{T_2,K_4} = 0$
+%                 $\cT^{10}_{T_2,K_5} = 0$ &
+%             $\cT^{10}_{T_3,K_1} = 0$
+%                 $\cT^{10}_{T_3,K_2} = 0$
+%                 $\cT^{10}_{T_3,K_3} = 0$
+%                 $\cT^{10}_{T_3,K_4} = 0$
+%                 $\cT^{10}_{T_3,K_5} = 0$ &
+%             $\cT^{10}_{T_4,K_1} = 0$
+%                 $\cT^{10}_{T_4,K_2} = 0$
+%                 $\cT^{10}_{T_4,K_3} = 0$
+%                 $\cT^{10}_{T_4,K_4} = 0$
+%                 $\cT^{10}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Appr. F-H Gittins &
+%             $\cT^{11}_{T_1,K_1} = 0$
+%                 $\cT^{11}_{T_1,K_2} = 0$
+%                 $\cT^{11}_{T_1,K_3} = 0$
+%                 $\cT^{11}_{T_1,K_4} = 0$
+%                 $\cT^{11}_{T_1,K_5} = 0$ &
+%             $\cT^{11}_{T_2,K_1} = 0$
+%                 $\cT^{11}_{T_2,K_2} = 0$
+%                 $\cT^{11}_{T_2,K_3} = 0$
+%                 $\cT^{11}_{T_2,K_4} = 0$
+%                 $\cT^{11}_{T_2,K_5} = 0$ &
+%             $\cT^{11}_{T_3,K_1} = 0$
+%                 $\cT^{11}_{T_3,K_2} = 0$
+%                 $\cT^{11}_{T_3,K_3} = 0$
+%                 $\cT^{11}_{T_3,K_4} = 0$
+%                 $\cT^{11}_{T_3,K_5} = 0$ &
+%             $\cT^{11}_{T_4,K_1} = 0$
+%                 $\cT^{11}_{T_4,K_2} = 0$
+%                 $\cT^{11}_{T_4,K_3} = 0$
+%                 $\cT^{11}_{T_4,K_4} = 0$
+%                 $\cT^{11}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\UCB\dagger$ &
+%             $\cT^{12}_{T_1,K_1} = 0$
+%                 $\cT^{12}_{T_1,K_2} = 0$
+%                 $\cT^{12}_{T_1,K_3} = 0$
+%                 $\cT^{12}_{T_1,K_4} = 0$
+%                 $\cT^{12}_{T_1,K_5} = 0$ &
+%             $\cT^{12}_{T_2,K_1} = 0$
+%                 $\cT^{12}_{T_2,K_2} = 0$
+%                 $\cT^{12}_{T_2,K_3} = 0$
+%                 $\cT^{12}_{T_2,K_4} = 0$
+%                 $\cT^{12}_{T_2,K_5} = 0$ &
+%             $\cT^{12}_{T_3,K_1} = 0$
+%                 $\cT^{12}_{T_3,K_2} = 0$
+%                 $\cT^{12}_{T_3,K_3} = 0$
+%                 $\cT^{12}_{T_3,K_4} = 0$
+%                 $\cT^{12}_{T_3,K_5} = 0$ &
+%             $\cT^{12}_{T_4,K_1} = 0$
+%                 $\cT^{12}_{T_4,K_2} = 0$
+%                 $\cT^{12}_{T_4,K_3} = 0$
+%                 $\cT^{12}_{T_4,K_4} = 0$
+%                 $\cT^{12}_{T_4,K_5} = 0$ \\
+%         \hline
+%         KL-UCB-Switch &
+%             $\cT^{13}_{T_1,K_1} = 0$
+%                 $\cT^{13}_{T_1,K_2} = 0$
+%                 $\cT^{13}_{T_1,K_3} = 0$
+%                 $\cT^{13}_{T_1,K_4} = 0$
+%                 $\cT^{13}_{T_1,K_5} = 0$ &
+%             $\cT^{13}_{T_2,K_1} = 0$
+%                 $\cT^{13}_{T_2,K_2} = 0$
+%                 $\cT^{13}_{T_2,K_3} = 0$
+%                 $\cT^{13}_{T_2,K_4} = 0$
+%                 $\cT^{13}_{T_2,K_5} = 0$ &
+%             $\cT^{13}_{T_3,K_1} = 0$
+%                 $\cT^{13}_{T_3,K_2} = 0$
+%                 $\cT^{13}_{T_3,K_3} = 0$
+%                 $\cT^{13}_{T_3,K_4} = 0$
+%                 $\cT^{13}_{T_3,K_5} = 0$ &
+%             $\cT^{13}_{T_4,K_1} = 0$
+%                 $\cT^{13}_{T_4,K_2} = 0$
+%                 $\cT^{13}_{T_4,K_3} = 0$
+%                 $\cT^{13}_{T_4,K_4} = 0$
+%                 $\cT^{13}_{T_4,K_5} = 0$ \\
+%         \hline
+%         \textsc{Tsallis-Inf} &
+%             $\cT^{14}_{T_1,K_1} = 0$
+%                 $\cT^{14}_{T_1,K_2} = 0$
+%                 $\cT^{14}_{T_1,K_3} = 0$
+%                 $\cT^{14}_{T_1,K_4} = 0$
+%                 $\cT^{14}_{T_1,K_5} = 0$ &
+%             $\cT^{14}_{T_2,K_1} = 0$
+%                 $\cT^{14}_{T_2,K_2} = 0$
+%                 $\cT^{14}_{T_2,K_3} = 0$
+%                 $\cT^{14}_{T_2,K_4} = 0$
+%                 $\cT^{14}_{T_2,K_5} = 0$ &
+%             $\cT^{14}_{T_3,K_1} = 0$
+%                 $\cT^{14}_{T_3,K_2} = 0$
+%                 $\cT^{14}_{T_3,K_3} = 0$
+%                 $\cT^{14}_{T_3,K_4} = 0$
+%                 $\cT^{14}_{T_3,K_5} = 0$ &
+%             $\cT^{14}_{T_4,K_1} = 0$
+%                 $\cT^{14}_{T_4,K_2} = 0$
+%                 $\cT^{14}_{T_4,K_3} = 0$
+%                 $\cT^{14}_{T_4,K_4} = 0$
+%                 $\cT^{14}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\mathrm{RCB}$ &
+%             $\cT^{15}_{T_1,K_1} = 0$
+%                 $\cT^{15}_{T_1,K_2} = 0$
+%                 $\cT^{15}_{T_1,K_3} = 0$
+%                 $\cT^{15}_{T_1,K_4} = 0$
+%                 $\cT^{15}_{T_1,K_5} = 0$ &
+%             $\cT^{15}_{T_2,K_1} = 0$
+%                 $\cT^{15}_{T_2,K_2} = 0$
+%                 $\cT^{15}_{T_2,K_3} = 0$
+%                 $\cT^{15}_{T_2,K_4} = 0$
+%                 $\cT^{15}_{T_2,K_5} = 0$ &
+%             $\cT^{15}_{T_3,K_1} = 0$
+%                 $\cT^{15}_{T_3,K_2} = 0$
+%                 $\cT^{15}_{T_3,K_3} = 0$
+%                 $\cT^{15}_{T_3,K_4} = 0$
+%                 $\cT^{15}_{T_3,K_5} = 0$ &
+%             $\cT^{15}_{T_4,K_1} = 0$
+%                 $\cT^{15}_{T_4,K_2} = 0$
+%                 $\cT^{15}_{T_4,K_3} = 0$
+%                 $\cT^{15}_{T_4,K_4} = 0$
+%                 $\cT^{15}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\mathrm{PHE}$ &
+%             $\cT^{16}_{T_1,K_1} = 0$
+%                 $\cT^{16}_{T_1,K_2} = 0$
+%                 $\cT^{16}_{T_1,K_3} = 0$
+%                 $\cT^{16}_{T_1,K_4} = 0$
+%                 $\cT^{16}_{T_1,K_5} = 0$ &
+%             $\cT^{16}_{T_2,K_1} = 0$
+%                 $\cT^{16}_{T_2,K_2} = 0$
+%                 $\cT^{16}_{T_2,K_3} = 0$
+%                 $\cT^{16}_{T_2,K_4} = 0$
+%                 $\cT^{16}_{T_2,K_5} = 0$ &
+%             $\cT^{16}_{T_3,K_1} = 0$
+%                 $\cT^{16}_{T_3,K_2} = 0$
+%                 $\cT^{16}_{T_3,K_3} = 0$
+%                 $\cT^{16}_{T_3,K_4} = 0$
+%                 $\cT^{16}_{T_3,K_5} = 0$ &
+%             $\cT^{16}_{T_4,K_1} = 0$
+%                 $\cT^{16}_{T_4,K_2} = 0$
+%                 $\cT^{16}_{T_4,K_3} = 0$
+%                 $\cT^{16}_{T_4,K_4} = 0$
+%                 $\cT^{16}_{T_4,K_5} = 0$ \\
+%         \hline
+%     \end{tabular}
+%     \caption{Normalized running time, rounded and in milli-seconds, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+%     \label{table:3:time_problem1_otherAlgorithms}
+% \end{footnotesize}  % WARNING
+% \end{table}
+
+% \paragraph{Storage cost}
+
+% \TODOL{A faire}
+
+% \begin{table}[!t]
+% \begin{footnotesize}  % WARNING
+%     \centering
+%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
+%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
+%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
+%         $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
+%             $\cM^{10}_{T_1,K_1} = 0$
+%                 $\cM^{10}_{T_1,K_2} = 0$
+%                 $\cM^{10}_{T_1,K_3} = 0$
+%                 $\cM^{10}_{T_1,K_4} = 0$
+%                 $\cM^{10}_{T_1,K_5} = 0$ &
+%             $\cM^{10}_{T_2,K_1} = 0$
+%                 $\cM^{10}_{T_2,K_2} = 0$
+%                 $\cM^{10}_{T_2,K_3} = 0$
+%                 $\cM^{10}_{T_2,K_4} = 0$
+%                 $\cM^{10}_{T_2,K_5} = 0$ &
+%             $\cM^{10}_{T_3,K_1} = 0$
+%                 $\cM^{10}_{T_3,K_2} = 0$
+%                 $\cM^{10}_{T_3,K_3} = 0$
+%                 $\cM^{10}_{T_3,K_4} = 0$
+%                 $\cM^{10}_{T_3,K_5} = 0$ &
+%             $\cM^{10}_{T_4,K_1} = 0$
+%                 $\cM^{10}_{T_4,K_2} = 0$
+%                 $\cM^{10}_{T_4,K_3} = 0$
+%                 $\cM^{10}_{T_4,K_4} = 0$
+%                 $\cM^{10}_{T_4,K_5} = 0$ \\
+%         \hline
+%         Appr. F-H Gittins &
+%             $\cM^{11}_{T_1,K_1} = 0$
+%                 $\cM^{11}_{T_1,K_2} = 0$
+%                 $\cM^{11}_{T_1,K_3} = 0$
+%                 $\cM^{11}_{T_1,K_4} = 0$
+%                 $\cM^{11}_{T_1,K_5} = 0$ &
+%             $\cM^{11}_{T_2,K_1} = 0$
+%                 $\cM^{11}_{T_2,K_2} = 0$
+%                 $\cM^{11}_{T_2,K_3} = 0$
+%                 $\cM^{11}_{T_2,K_4} = 0$
+%                 $\cM^{11}_{T_2,K_5} = 0$ &
+%             $\cM^{11}_{T_3,K_1} = 0$
+%                 $\cM^{11}_{T_3,K_2} = 0$
+%                 $\cM^{11}_{T_3,K_3} = 0$
+%                 $\cM^{11}_{T_3,K_4} = 0$
+%                 $\cM^{11}_{T_3,K_5} = 0$ &
+%             $\cM^{11}_{T_4,K_1} = 0$
+%                 $\cM^{11}_{T_4,K_2} = 0$
+%                 $\cM^{11}_{T_4,K_3} = 0$
+%                 $\cM^{11}_{T_4,K_4} = 0$
+%                 $\cM^{11}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\UCB\dagger$ &
+%             $\cM^{12}_{T_1,K_1} = 0$
+%                 $\cM^{12}_{T_1,K_2} = 0$
+%                 $\cM^{12}_{T_1,K_3} = 0$
+%                 $\cM^{12}_{T_1,K_4} = 0$
+%                 $\cM^{12}_{T_1,K_5} = 0$ &
+%             $\cM^{12}_{T_2,K_1} = 0$
+%                 $\cM^{12}_{T_2,K_2} = 0$
+%                 $\cM^{12}_{T_2,K_3} = 0$
+%                 $\cM^{12}_{T_2,K_4} = 0$
+%                 $\cM^{12}_{T_2,K_5} = 0$ &
+%             $\cM^{12}_{T_3,K_1} = 0$
+%                 $\cM^{12}_{T_3,K_2} = 0$
+%                 $\cM^{12}_{T_3,K_3} = 0$
+%                 $\cM^{12}_{T_3,K_4} = 0$
+%                 $\cM^{12}_{T_3,K_5} = 0$ &
+%             $\cM^{12}_{T_4,K_1} = 0$
+%                 $\cM^{12}_{T_4,K_2} = 0$
+%                 $\cM^{12}_{T_4,K_3} = 0$
+%                 $\cM^{12}_{T_4,K_4} = 0$
+%                 $\cM^{12}_{T_4,K_5} = 0$ \\
+%         \hline
+%         KL-UCB-Switch &
+%             $\cM^{13}_{T_1,K_1} = 0$
+%                 $\cM^{13}_{T_1,K_2} = 0$
+%                 $\cM^{13}_{T_1,K_3} = 0$
+%                 $\cM^{13}_{T_1,K_4} = 0$
+%                 $\cM^{13}_{T_1,K_5} = 0$ &
+%             $\cM^{13}_{T_2,K_1} = 0$
+%                 $\cM^{13}_{T_2,K_2} = 0$
+%                 $\cM^{13}_{T_2,K_3} = 0$
+%                 $\cM^{13}_{T_2,K_4} = 0$
+%                 $\cM^{13}_{T_2,K_5} = 0$ &
+%             $\cM^{13}_{T_3,K_1} = 0$
+%                 $\cM^{13}_{T_3,K_2} = 0$
+%                 $\cM^{13}_{T_3,K_3} = 0$
+%                 $\cM^{13}_{T_3,K_4} = 0$
+%                 $\cM^{13}_{T_3,K_5} = 0$ &
+%             $\cM^{13}_{T_4,K_1} = 0$
+%                 $\cM^{13}_{T_4,K_2} = 0$
+%                 $\cM^{13}_{T_4,K_3} = 0$
+%                 $\cM^{13}_{T_4,K_4} = 0$
+%                 $\cM^{13}_{T_4,K_5} = 0$ \\
+%         \hline
+%         \textsc{Tsallis-Inf} &
+%             $\cM^{14}_{T_1,K_1} = 0$
+%                 $\cM^{14}_{T_1,K_2} = 0$
+%                 $\cM^{14}_{T_1,K_3} = 0$
+%                 $\cM^{14}_{T_1,K_4} = 0$
+%                 $\cM^{14}_{T_1,K_5} = 0$ &
+%             $\cM^{14}_{T_2,K_1} = 0$
+%                 $\cM^{14}_{T_2,K_2} = 0$
+%                 $\cM^{14}_{T_2,K_3} = 0$
+%                 $\cM^{14}_{T_2,K_4} = 0$
+%                 $\cM^{14}_{T_2,K_5} = 0$ &
+%             $\cM^{14}_{T_3,K_1} = 0$
+%                 $\cM^{14}_{T_3,K_2} = 0$
+%                 $\cM^{14}_{T_3,K_3} = 0$
+%                 $\cM^{14}_{T_3,K_4} = 0$
+%                 $\cM^{14}_{T_3,K_5} = 0$ &
+%             $\cM^{14}_{T_4,K_1} = 0$
+%                 $\cM^{14}_{T_4,K_2} = 0$
+%                 $\cM^{14}_{T_4,K_3} = 0$
+%                 $\cM^{14}_{T_4,K_4} = 0$
+%                 $\cM^{14}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\mathrm{RCB}$ &
+%             $\cM^{15}_{T_1,K_1} = 0$
+%                 $\cM^{15}_{T_1,K_2} = 0$
+%                 $\cM^{15}_{T_1,K_3} = 0$
+%                 $\cM^{15}_{T_1,K_4} = 0$
+%                 $\cM^{15}_{T_1,K_5} = 0$ &
+%             $\cM^{15}_{T_2,K_1} = 0$
+%                 $\cM^{15}_{T_2,K_2} = 0$
+%                 $\cM^{15}_{T_2,K_3} = 0$
+%                 $\cM^{15}_{T_2,K_4} = 0$
+%                 $\cM^{15}_{T_2,K_5} = 0$ &
+%             $\cM^{15}_{T_3,K_1} = 0$
+%                 $\cM^{15}_{T_3,K_2} = 0$
+%                 $\cM^{15}_{T_3,K_3} = 0$
+%                 $\cM^{15}_{T_3,K_4} = 0$
+%                 $\cM^{15}_{T_3,K_5} = 0$ &
+%             $\cM^{15}_{T_4,K_1} = 0$
+%                 $\cM^{15}_{T_4,K_2} = 0$
+%                 $\cM^{15}_{T_4,K_3} = 0$
+%                 $\cM^{15}_{T_4,K_4} = 0$
+%                 $\cM^{15}_{T_4,K_5} = 0$ \\
+%         \hline
+%         $\mathrm{RCB}$ &
+%             $\cM^{16}_{T_1,K_1} = 0$
+%                 $\cM^{16}_{T_1,K_2} = 0$
+%                 $\cM^{16}_{T_1,K_3} = 0$
+%                 $\cM^{16}_{T_1,K_4} = 0$
+%                 $\cM^{16}_{T_1,K_5} = 0$ &
+%             $\cM^{16}_{T_2,K_1} = 0$
+%                 $\cM^{16}_{T_2,K_2} = 0$
+%                 $\cM^{16}_{T_2,K_3} = 0$
+%                 $\cM^{16}_{T_2,K_4} = 0$
+%                 $\cM^{16}_{T_2,K_5} = 0$ &
+%             $\cM^{16}_{T_3,K_1} = 0$
+%                 $\cM^{16}_{T_3,K_2} = 0$
+%                 $\cM^{16}_{T_3,K_3} = 0$
+%                 $\cM^{16}_{T_3,K_4} = 0$
+%                 $\cM^{16}_{T_3,K_5} = 0$ &
+%             $\cM^{16}_{T_4,K_1} = 0$
+%                 $\cM^{16}_{T_4,K_2} = 0$
+%                 $\cM^{16}_{T_4,K_3} = 0$
+%                 $\cM^{16}_{T_4,K_4} = 0$
+%                 $\cM^{16}_{T_4,K_5} = 0$ \\
+%         \hline
+%     \end{tabular}
+%     \caption{Non-normalized memory cost, rounded and in Bytes, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
+%     \label{table:3:memory_problem1_otherAlgorithms}
+% \end{footnotesize}  % WARNING
+% \end{table}
+
+\paragraph{Interpretation}
 
@@
 
-\begin{table}[!t]
-\begin{footnotesize}  % WARNING
-    \centering
-    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-        $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
-            $\cT^{10}_{T_1,K_1} = 0$
-                $\cT^{10}_{T_1,K_2} = 0$
-                $\cT^{10}_{T_1,K_3} = 0$
-                $\cT^{10}_{T_1,K_4} = 0$
-                $\cT^{10}_{T_1,K_5} = 0$ &
-            $\cT^{10}_{T_2,K_1} = 0$
-                $\cT^{10}_{T_2,K_2} = 0$
-                $\cT^{10}_{T_2,K_3} = 0$
-                $\cT^{10}_{T_2,K_4} = 0$
-                $\cT^{10}_{T_2,K_5} = 0$ &
-            $\cT^{10}_{T_3,K_1} = 0$
-                $\cT^{10}_{T_3,K_2} = 0$
-                $\cT^{10}_{T_3,K_3} = 0$
-                $\cT^{10}_{T_3,K_4} = 0$
-                $\cT^{10}_{T_3,K_5} = 0$ &
-            $\cT^{10}_{T_4,K_1} = 0$
-                $\cT^{10}_{T_4,K_2} = 0$
-                $\cT^{10}_{T_4,K_3} = 0$
-                $\cT^{10}_{T_4,K_4} = 0$
-                $\cT^{10}_{T_4,K_5} = 0$ \\
-        \hline
-        Appr. F-H Gittins &
-            $\cT^{11}_{T_1,K_1} = 0$
-                $\cT^{11}_{T_1,K_2} = 0$
-                $\cT^{11}_{T_1,K_3} = 0$
-                $\cT^{11}_{T_1,K_4} = 0$
-                $\cT^{11}_{T_1,K_5} = 0$ &
-            $\cT^{11}_{T_2,K_1} = 0$
-                $\cT^{11}_{T_2,K_2} = 0$
-                $\cT^{11}_{T_2,K_3} = 0$
-                $\cT^{11}_{T_2,K_4} = 0$
-                $\cT^{11}_{T_2,K_5} = 0$ &
-            $\cT^{11}_{T_3,K_1} = 0$
-                $\cT^{11}_{T_3,K_2} = 0$
-                $\cT^{11}_{T_3,K_3} = 0$
-                $\cT^{11}_{T_3,K_4} = 0$
-                $\cT^{11}_{T_3,K_5} = 0$ &
-            $\cT^{11}_{T_4,K_1} = 0$
-                $\cT^{11}_{T_4,K_2} = 0$
-                $\cT^{11}_{T_4,K_3} = 0$
-                $\cT^{11}_{T_4,K_4} = 0$
-                $\cT^{11}_{T_4,K_5} = 0$ \\
-        \hline
-        $\UCB\dagger$ &
-            $\cT^{12}_{T_1,K_1} = 0$
-                $\cT^{12}_{T_1,K_2} = 0$
-                $\cT^{12}_{T_1,K_3} = 0$
-                $\cT^{12}_{T_1,K_4} = 0$
-                $\cT^{12}_{T_1,K_5} = 0$ &
-            $\cT^{12}_{T_2,K_1} = 0$
-                $\cT^{12}_{T_2,K_2} = 0$
-                $\cT^{12}_{T_2,K_3} = 0$
-                $\cT^{12}_{T_2,K_4} = 0$
-                $\cT^{12}_{T_2,K_5} = 0$ &
-            $\cT^{12}_{T_3,K_1} = 0$
-                $\cT^{12}_{T_3,K_2} = 0$
-                $\cT^{12}_{T_3,K_3} = 0$
-                $\cT^{12}_{T_3,K_4} = 0$
-                $\cT^{12}_{T_3,K_5} = 0$ &
-            $\cT^{12}_{T_4,K_1} = 0$
-                $\cT^{12}_{T_4,K_2} = 0$
-                $\cT^{12}_{T_4,K_3} = 0$
-                $\cT^{12}_{T_4,K_4} = 0$
-                $\cT^{12}_{T_4,K_5} = 0$ \\
-        \hline
-        KL-UCB-Switch &
-            $\cT^{13}_{T_1,K_1} = 0$
-                $\cT^{13}_{T_1,K_2} = 0$
-                $\cT^{13}_{T_1,K_3} = 0$
-                $\cT^{13}_{T_1,K_4} = 0$
-                $\cT^{13}_{T_1,K_5} = 0$ &
-            $\cT^{13}_{T_2,K_1} = 0$
-                $\cT^{13}_{T_2,K_2} = 0$
-                $\cT^{13}_{T_2,K_3} = 0$
-                $\cT^{13}_{T_2,K_4} = 0$
-                $\cT^{13}_{T_2,K_5} = 0$ &
-            $\cT^{13}_{T_3,K_1} = 0$
-                $\cT^{13}_{T_3,K_2} = 0$
-                $\cT^{13}_{T_3,K_3} = 0$
-                $\cT^{13}_{T_3,K_4} = 0$
-                $\cT^{13}_{T_3,K_5} = 0$ &
-            $\cT^{13}_{T_4,K_1} = 0$
-                $\cT^{13}_{T_4,K_2} = 0$
-                $\cT^{13}_{T_4,K_3} = 0$
-                $\cT^{13}_{T_4,K_4} = 0$
-                $\cT^{13}_{T_4,K_5} = 0$ \\
-        \hline
-        \textsc{Tsallis-Inf} &
-            $\cT^{14}_{T_1,K_1} = 0$
-                $\cT^{14}_{T_1,K_2} = 0$
-                $\cT^{14}_{T_1,K_3} = 0$
-                $\cT^{14}_{T_1,K_4} = 0$
-                $\cT^{14}_{T_1,K_5} = 0$ &
-            $\cT^{14}_{T_2,K_1} = 0$
-                $\cT^{14}_{T_2,K_2} = 0$
-                $\cT^{14}_{T_2,K_3} = 0$
-                $\cT^{14}_{T_2,K_4} = 0$
-                $\cT^{14}_{T_2,K_5} = 0$ &
-            $\cT^{14}_{T_3,K_1} = 0$
-                $\cT^{14}_{T_3,K_2} = 0$
-                $\cT^{14}_{T_3,K_3} = 0$
-                $\cT^{14}_{T_3,K_4} = 0$
-                $\cT^{14}_{T_3,K_5} = 0$ &
-            $\cT^{14}_{T_4,K_1} = 0$
-                $\cT^{14}_{T_4,K_2} = 0$
-                $\cT^{14}_{T_4,K_3} = 0$
-                $\cT^{14}_{T_4,K_4} = 0$
-                $\cT^{14}_{T_4,K_5} = 0$ \\
-        \hline
-        $\mathrm{RCB}$ &
-            $\cT^{15}_{T_1,K_1} = 0$
-                $\cT^{15}_{T_1,K_2} = 0$
-                $\cT^{15}_{T_1,K_3} = 0$
-                $\cT^{15}_{T_1,K_4} = 0$
-                $\cT^{15}_{T_1,K_5} = 0$ &
-            $\cT^{15}_{T_2,K_1} = 0$
-                $\cT^{15}_{T_2,K_2} = 0$
-                $\cT^{15}_{T_2,K_3} = 0$
-                $\cT^{15}_{T_2,K_4} = 0$
-                $\cT^{15}_{T_2,K_5} = 0$ &
-            $\cT^{15}_{T_3,K_1} = 0$
-                $\cT^{15}_{T_3,K_2} = 0$
-                $\cT^{15}_{T_3,K_3} = 0$
-                $\cT^{15}_{T_3,K_4} = 0$
-                $\cT^{15}_{T_3,K_5} = 0$ &
-            $\cT^{15}_{T_4,K_1} = 0$
-                $\cT^{15}_{T_4,K_2} = 0$
-                $\cT^{15}_{T_4,K_3} = 0$
-                $\cT^{15}_{T_4,K_4} = 0$
-                $\cT^{15}_{T_4,K_5} = 0$ \\
-        \hline
-        $\mathrm{PHE}$ &
-            $\cT^{16}_{T_1,K_1} = 0$
-                $\cT^{16}_{T_1,K_2} = 0$
-                $\cT^{16}_{T_1,K_3} = 0$
-                $\cT^{16}_{T_1,K_4} = 0$
-                $\cT^{16}_{T_1,K_5} = 0$ &
-            $\cT^{16}_{T_2,K_1} = 0$
-                $\cT^{16}_{T_2,K_2} = 0$
-                $\cT^{16}_{T_2,K_3} = 0$
-                $\cT^{16}_{T_2,K_4} = 0$
-                $\cT^{16}_{T_2,K_5} = 0$ &
-            $\cT^{16}_{T_3,K_1} = 0$
-                $\cT^{16}_{T_3,K_2} = 0$
-                $\cT^{16}_{T_3,K_3} = 0$
-                $\cT^{16}_{T_3,K_4} = 0$
-                $\cT^{16}_{T_3,K_5} = 0$ &
-            $\cT^{16}_{T_4,K_1} = 0$
-                $\cT^{16}_{T_4,K_2} = 0$
-                $\cT^{16}_{T_4,K_3} = 0$
-                $\cT^{16}_{T_4,K_4} = 0$
-                $\cT^{16}_{T_4,K_5} = 0$ \\
-        \hline
-    \end{tabular}
-    \caption{Normalized running time, rounded and in milli-seconds, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
-    \label{table:3:time_problem1_otherAlgorithms}
-\end{footnotesize}  % WARNING
-\end{table}
-
-\paragraph{Storage cost}
-
-\TODOL{A faire}
-
-\begin{table}[!t]
-\begin{footnotesize}  % WARNING
-    \centering
-    \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-        & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-        $\mathrm{MOSS}$-$\mathrm{Anytime}$ &
-            $\cM^{10}_{T_1,K_1} = 0$
-                $\cM^{10}_{T_1,K_2} = 0$
-                $\cM^{10}_{T_1,K_3} = 0$
-                $\cM^{10}_{T_1,K_4} = 0$
-                $\cM^{10}_{T_1,K_5} = 0$ &
-            $\cM^{10}_{T_2,K_1} = 0$
-                $\cM^{10}_{T_2,K_2} = 0$
-                $\cM^{10}_{T_2,K_3} = 0$
-                $\cM^{10}_{T_2,K_4} = 0$
-                $\cM^{10}_{T_2,K_5} = 0$ &
-            $\cM^{10}_{T_3,K_1} = 0$
-                $\cM^{10}_{T_3,K_2} = 0$
-                $\cM^{10}_{T_3,K_3} = 0$
-                $\cM^{10}_{T_3,K_4} = 0$
-                $\cM^{10}_{T_3,K_5} = 0$ &
-            $\cM^{10}_{T_4,K_1} = 0$
-                $\cM^{10}_{T_4,K_2} = 0$
-                $\cM^{10}_{T_4,K_3} = 0$
-                $\cM^{10}_{T_4,K_4} = 0$
-                $\cM^{10}_{T_4,K_5} = 0$ \\
-        \hline
-        Appr. F-H Gittins &
-            $\cM^{11}_{T_1,K_1} = 0$
-                $\cM^{11}_{T_1,K_2} = 0$
-                $\cM^{11}_{T_1,K_3} = 0$
-                $\cM^{11}_{T_1,K_4} = 0$
-                $\cM^{11}_{T_1,K_5} = 0$ &
-            $\cM^{11}_{T_2,K_1} = 0$
-                $\cM^{11}_{T_2,K_2} = 0$
-                $\cM^{11}_{T_2,K_3} = 0$
-                $\cM^{11}_{T_2,K_4} = 0$
-                $\cM^{11}_{T_2,K_5} = 0$ &
-            $\cM^{11}_{T_3,K_1} = 0$
-                $\cM^{11}_{T_3,K_2} = 0$
-                $\cM^{11}_{T_3,K_3} = 0$
-                $\cM^{11}_{T_3,K_4} = 0$
-                $\cM^{11}_{T_3,K_5} = 0$ &
-            $\cM^{11}_{T_4,K_1} = 0$
-                $\cM^{11}_{T_4,K_2} = 0$
-                $\cM^{11}_{T_4,K_3} = 0$
-                $\cM^{11}_{T_4,K_4} = 0$
-                $\cM^{11}_{T_4,K_5} = 0$ \\
-        \hline
-        $\UCB\dagger$ &
-            $\cM^{12}_{T_1,K_1} = 0$
-                $\cM^{12}_{T_1,K_2} = 0$
-                $\cM^{12}_{T_1,K_3} = 0$
-                $\cM^{12}_{T_1,K_4} = 0$
-                $\cM^{12}_{T_1,K_5} = 0$ &
-            $\cM^{12}_{T_2,K_1} = 0$
-                $\cM^{12}_{T_2,K_2} = 0$
-                $\cM^{12}_{T_2,K_3} = 0$
-                $\cM^{12}_{T_2,K_4} = 0$
-                $\cM^{12}_{T_2,K_5} = 0$ &
-            $\cM^{12}_{T_3,K_1} = 0$
-                $\cM^{12}_{T_3,K_2} = 0$
-                $\cM^{12}_{T_3,K_3} = 0$
-                $\cM^{12}_{T_3,K_4} = 0$
-                $\cM^{12}_{T_3,K_5} = 0$ &
-            $\cM^{12}_{T_4,K_1} = 0$
-                $\cM^{12}_{T_4,K_2} = 0$
-                $\cM^{12}_{T_4,K_3} = 0$
-                $\cM^{12}_{T_4,K_4} = 0$
-                $\cM^{12}_{T_4,K_5} = 0$ \\
-        \hline
-        KL-UCB-Switch &
-            $\cM^{13}_{T_1,K_1} = 0$
-                $\cM^{13}_{T_1,K_2} = 0$
-                $\cM^{13}_{T_1,K_3} = 0$
-                $\cM^{13}_{T_1,K_4} = 0$
-                $\cM^{13}_{T_1,K_5} = 0$ &
-            $\cM^{13}_{T_2,K_1} = 0$
-                $\cM^{13}_{T_2,K_2} = 0$
-                $\cM^{13}_{T_2,K_3} = 0$
-                $\cM^{13}_{T_2,K_4} = 0$
-                $\cM^{13}_{T_2,K_5} = 0$ &
-            $\cM^{13}_{T_3,K_1} = 0$
-                $\cM^{13}_{T_3,K_2} = 0$
-                $\cM^{13}_{T_3,K_3} = 0$
-                $\cM^{13}_{T_3,K_4} = 0$
-                $\cM^{13}_{T_3,K_5} = 0$ &
-            $\cM^{13}_{T_4,K_1} = 0$
-                $\cM^{13}_{T_4,K_2} = 0$
-                $\cM^{13}_{T_4,K_3} = 0$
-                $\cM^{13}_{T_4,K_4} = 0$
-                $\cM^{13}_{T_4,K_5} = 0$ \\
-        \hline
-        \textsc{Tsallis-Inf} &
-            $\cM^{14}_{T_1,K_1} = 0$
-                $\cM^{14}_{T_1,K_2} = 0$
-                $\cM^{14}_{T_1,K_3} = 0$
-                $\cM^{14}_{T_1,K_4} = 0$
-                $\cM^{14}_{T_1,K_5} = 0$ &
-            $\cM^{14}_{T_2,K_1} = 0$
-                $\cM^{14}_{T_2,K_2} = 0$
-                $\cM^{14}_{T_2,K_3} = 0$
-                $\cM^{14}_{T_2,K_4} = 0$
-                $\cM^{14}_{T_2,K_5} = 0$ &
-            $\cM^{14}_{T_3,K_1} = 0$
-                $\cM^{14}_{T_3,K_2} = 0$
-                $\cM^{14}_{T_3,K_3} = 0$
-                $\cM^{14}_{T_3,K_4} = 0$
-                $\cM^{14}_{T_3,K_5} = 0$ &
-            $\cM^{14}_{T_4,K_1} = 0$
-                $\cM^{14}_{T_4,K_2} = 0$
-                $\cM^{14}_{T_4,K_3} = 0$
-                $\cM^{14}_{T_4,K_4} = 0$
-                $\cM^{14}_{T_4,K_5} = 0$ \\
-        \hline
-        $\mathrm{RCB}$ &
-            $\cM^{15}_{T_1,K_1} = 0$
-                $\cM^{15}_{T_1,K_2} = 0$
-                $\cM^{15}_{T_1,K_3} = 0$
-                $\cM^{15}_{T_1,K_4} = 0$
-                $\cM^{15}_{T_1,K_5} = 0$ &
-            $\cM^{15}_{T_2,K_1} = 0$
-                $\cM^{15}_{T_2,K_2} = 0$
-                $\cM^{15}_{T_2,K_3} = 0$
-                $\cM^{15}_{T_2,K_4} = 0$
-                $\cM^{15}_{T_2,K_5} = 0$ &
-            $\cM^{15}_{T_3,K_1} = 0$
-                $\cM^{15}_{T_3,K_2} = 0$
-                $\cM^{15}_{T_3,K_3} = 0$
-                $\cM^{15}_{T_3,K_4} = 0$
-                $\cM^{15}_{T_3,K_5} = 0$ &
-            $\cM^{15}_{T_4,K_1} = 0$
-                $\cM^{15}_{T_4,K_2} = 0$
-                $\cM^{15}_{T_4,K_3} = 0$
-                $\cM^{15}_{T_4,K_4} = 0$
-                $\cM^{15}_{T_4,K_5} = 0$ \\
-        \hline
-        $\mathrm{RCB}$ &
-            $\cM^{16}_{T_1,K_1} = 0$
-                $\cM^{16}_{T_1,K_2} = 0$
-                $\cM^{16}_{T_1,K_3} = 0$
-                $\cM^{16}_{T_1,K_4} = 0$
-                $\cM^{16}_{T_1,K_5} = 0$ &
-            $\cM^{16}_{T_2,K_1} = 0$
-                $\cM^{16}_{T_2,K_2} = 0$
-                $\cM^{16}_{T_2,K_3} = 0$
-                $\cM^{16}_{T_2,K_4} = 0$
-                $\cM^{16}_{T_2,K_5} = 0$ &
-            $\cM^{16}_{T_3,K_1} = 0$
-                $\cM^{16}_{T_3,K_2} = 0$
-                $\cM^{16}_{T_3,K_3} = 0$
-                $\cM^{16}_{T_3,K_4} = 0$
-                $\cM^{16}_{T_3,K_5} = 0$ &
-            $\cM^{16}_{T_4,K_1} = 0$
-                $\cM^{16}_{T_4,K_2} = 0$
-                $\cM^{16}_{T_4,K_3} = 0$
-                $\cM^{16}_{T_4,K_4} = 0$
-                $\cM^{16}_{T_4,K_5} = 0$ \\
-        \hline
-    \end{tabular}
-    \caption{Non-normalized memory cost, rounded and in Bytes, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
-    \label{table:3:memory_problem1_otherAlgorithms}
-\end{footnotesize}  % WARNING
-\end{table}
-
-\paragraph{Interpretation}
-
-\TODOL{A faire}
-
 \paragraph{Bash command to run these experiments}
@@
 Finally, we give below in Listing~\ref{lst:3:BashCodeToLaunchLargeExperiments} the Bash command that was used to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
-It uses \texttt{for} loops from Bash and not Python and environment variables, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $N$, $T$ and $K$.
+It uses \texttt{for} loops and environment variables from Bash, and not Python, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $N$, $T$ and $K$.
 

### 2019-05-06 2-Chapters/3-Chapter/chapter3.tex
Done for today. I don't know if Xtof and Mli will be happy about these plots, but I am :-)

@@
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormtime__56pb__7KS_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormtime__56pb__7Ks_7Ts.pdf}
 	\caption[Normalized mean regret vs normalized running time (in micro-seconds).]{
@@
 	}
-	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormtime__56pb__7KS_7Ts}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormtime__56pb__7Ks_7Ts}
 \end{figure}
@@
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7KS_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7Ks_7Ts.pdf}
 	\caption[Normalized mean regret vs normalized memory costs (in bytes).]{
@@
 	}
-	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7KS_7Ts}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7Ks_7Ts}
 \end{figure}

### 2019-05-07 2-Chapters/3-Chapter/chapter3.tex
I want to relaunch simulations, for other values of K (but just aggregate data).

@@
 We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
-then we present in Section~\ref{sec:3:timeAndMemoryCosts} additional numerical simulations to discuss about the time efficient and memory costs of the some MAB algorithms.
-We focus on the ones being used in the rest of this thesis, \UCB, \klUCB, and Thompson Sampling.
+then we discuss in Section~\ref{sec:3:timeAndMemoryCosts} about the time efficient and memory costs of the some MAB algorithms.
+The take away messages are two-fold:
+from a practical point-of-view, it is usually advised to use the simplest algorithm and we advise to use \UCB{} (as we do in the next Chapter~\ref{chapter:4}),
+and that for theoretical works where optimality of the MAB algorithm matters, we advise to use \klUCB{} (as we do in the next Chapters~\ref{chapter:5} and \ref{chapter:6}).
 % To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
@@
 %
-Finally, in Appendix~\ref{sec:3:appendix} we include some small files showing how to use the library, as well as details regarding the use of parallel computing to speed-up simulations.
+Finally, in Appendix~\ref{sec:3:appendix} we include some small files showing how to use the library, and we conclude with some details regarding the use of parallel computing to speed-up simulations.
 
@@
 
-SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}).
+SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.html}{PoliciesMultiPlayers}}).
 %
 It includes
-\texttt{rhoRand} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{PoliciesMultiPlayers.rhoRand}}) from \cite{Anandkumar11},
-\texttt{MEGA} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}}) from \cite{Avner15},
-\texttt{MusicalChair} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{Policies.MusicalChair}}) from \cite{Rosenski16},
+\texttt{rhoRand}
+% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{PoliciesMultiPlayers.rhoRand}})
+from \cite{Anandkumar11},
+\texttt{MEGA}
+% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}})
+from \cite{Avner15},
+\texttt{MusicalChair}
+% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{Policies.MusicalChair}})
+from \cite{Rosenski16},
 and our state-of-the-art algorithms
-\texttt{RandTopM} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{PoliciesMultiPlayers.RandTopM}}) and
-\texttt{MCTopM} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}}) from \cite{Besson2018ALT}.
+\texttt{RandTopM}
+% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{PoliciesMultiPlayers.RandTopM}})
+and \texttt{MCTopM}
+% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}})
+from \cite{Besson2018ALT}.
 For comparison, realistic (\eg, \UCB{} for multiple play) or full-knowledge centralized algorithms are also implemented.
@@
 Any simulation library targeting multi-player bandit problems must implement at least another component:
-a collision model, and a simulation loop that essentially looks like this.
+% a collision model, and
+a simulation loop that essentially looks like this.
 \begin{itemize}
@@
 
-\subsection{Features}
+\subsection{Summary of the features}
 
@@
 
-A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here:
+A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files:
 \texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
@@
 They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proved to match the lower-bound up-to a constant), and each but UCB is known to be optimal (\ie, with the constant matching the lower-bound).
-For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).
 	}
@@
 
-Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in a HDF5 file\footnote{~For example, this simulation produces this HDF5 file\\\texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}}
+Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards, regret and other data) in a HDF5 file\footnote{~For example, this simulation produced this HDF5 file\\\texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}}
 \cite{h5py}.
@@
 	\includegraphics[width=1.05\linewidth]{3_hist.pdf}
-	\caption{Histogram of regret for the same experiment.}
+	\caption{Histogram of regret for the same experiment.
+        For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).
+    }
 	\label{fig:3:firstPlot_hist}
@@
 Moreover, they offer to host the documentation online, at \href{https://smpybandits.rtfd.io/}{\texttt{SMPyBandits.ReadTheDocs.io}}.
-For more details, see \href{https://readthedocs.org/projects/smpybandits/}{\texttt{https://ReadTheDocs.org/projects/SMPyBandits}}.
+For more details, see \href{https://readthedocs.org/projects/smpybandits/}{\texttt{ReadTheDocs.org/projects/SMPyBandits}}.
 %
@@
 
-\subsection{Research works using SMPyBandits}
+\subsection{List of research works using SMPyBandits}
 
@@
 
-In this section, we use the SMPyBandits library to compare experimentally various state-of-the-art single-player algorithms on different kinds of multi-armed bandit problems.
+In this section, we use the SMPyBandits library to compare experimentally various state-of-the-art single-player algorithms on some multi-armed bandit problems.
 First, we detail the list of different algorithms that we compare, and the different problems on which they are all tested.
@@
 \indent
-For reproducibility, we also give in ``\texttt{typewriter font}'' the name of the corresponding class in the \texttt{Policies} module in SMPyBandits.
+For reproducibility, we also give in ``\texttt{typewriter font}'' the name of the corresponding class in the \texttt{Policies} module in SMPyBandits
+(\eg, $\UCB_1$ is \texttt{Policies.UCBalpha}).
 
@@
     \item Algorithm 4 is
-    $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCB}).
+    $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
     It was proved to achieve order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
@@
     In practice, a few steps of a Newton method are enough to obtain a precision of more than 10 digits, and empirically we found that even in the worst cases \klUCB{} is an order of magnitude slower than \UCB.
+    \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}
 
@@
 
-In the results presented in the main text of this chapter, we focus on the most well known algorithms, and two others that are efficient but less known (AdBandits and BESA).
-We present in Appendix~\ref{sub:3:additionalExperiments} seven other algorithms, being more recent (from $2016$, $3$ from $2018$ and $2$ from $2019$) but are usually more complex, or are based on a different point-of-view.
-We include numerical results for these other algorithms as well, to justify empirically that despite their respective qualities, and despite being efficient, it is reasonable to focus on \UCB{} and \klUCB{} algorithms in the rest of this thesis.
+We focussed on the most well known algorithms, and two others that are efficient but less known (AdBandits and BESA),
+but to illustrate that SMPyBandits implements many other solutions to the MAB problem proposed in very recent works, we explain in Appendix~\ref{sub:3:additionalExperiments} seven other algorithms.
+They are more recent (some from $2016$, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or are based on a different point-of-view.
+The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} algorithms in the rest of this thesis, as they stay comparable with the state-of-the-art algorithms but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to PHE or RCB)
 
@@
 For all experiments, we run $N=100$ independent simulations.
-See the snippet of code~\ref{lst:3:BashCodeToLaunchLargeExperiments} in the Appendix below.
+We show in the snippet of code~\ref{lst:3:BashCodeToLaunchLargeExperiments} in the Appendix below more details about how to run these experiments.
 
@@
 
-We give in the tables below the results of the experiments.
-For the two different problems and different values of $T$ and $K$, we give the mean $\pm 1$ standard-variation of the empirical regret $R_T$ obtained by the different algorithms at the end of the experiments.
-Tables~\ref{table:3:meanRegret_problem1} is for the fixed-gap \textbf{problem 1}
-and \ref{table:3:meanRegret_problem2} is for the randomly generated \textbf{problem 2}.
-Each column corresponds to one value of $T$, and each line corresponds to one algorithm.
-For each cell, we report successive values that correspond to the successive values of the number of arms $K$.
+We give in the figures below the results of the experiments.
+% For the two different problems and different values of $T$ and $K$, we give the mean $\pm 1$ standard-variation of the empirical regret $R_T$ obtained by the different algorithms at the end of the experiments.
+% Tables~\ref{table:3:meanRegret_problem1} is for the fixed-gap \textbf{problem 1}
+% and \ref{table:3:meanRegret_problem2} is for the randomly generated \textbf{problem 2}.
+% Each column corresponds to one value of $T$, and each line corresponds to one algorithm.
+% For each cell, we report successive values that correspond to the successive values of the number of arms $K$.
 
@@
 
-I want to show that SMPyBandits contains highly documented implementations of all the most important algorithms designed for stationary single-player multi-armed bandit.
-
-I will include here formulas about the main algorithms (klUCB, MOSS, etc).
-I will pick some problems (Bernoulli, Gaussian, others), with fixed means or ``Bayesian'' means (\ie, changing at each time) and I will run extensive numerical simulations to compare the main algorithms.
-
-The take away message is: klUCB beats UCB, Thompson sampling is great too, all the others algorithms are less efficient or comparable, so we only use klUCB for all the rest of this thesis.
-
-I will include graphs showing the mean regret $R_t$ as a function of time $t$, as well as histograms or boxplots showing the distribution of $R_T$.
+% I want to show that SMPyBandits contains highly documented implementations of all the most important algorithms designed for stationary single-player multi-armed bandit.
+
+% I will include here formulas about the main algorithms (klUCB, MOSS, etc).
+% I will pick some problems (Bernoulli, Gaussian, others), with fixed means or ``Bayesian'' means (\ie, changing at each time) and I will run extensive numerical simulations to compare the main algorithms.
+
+% The take away message is: klUCB beats UCB, Thompson sampling is great too, all the others algorithms are less efficient or comparable, so we only use klUCB for all the rest of this thesis.
+
+% I will include graphs showing the mean regret $R_t$ as a function of time $t$, as well as histograms or boxplots showing the distribution of $R_T$.
 
@@
 %
-The goal of this section is to highlight that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in\{1,\dots,T\}$ bounded by $\bigO{K}$, independent of $t$.
-%
+The goal of this section is to verify that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in\{1,\dots,T\}$ bounded by $\bigO{K}$, independent of $t$.
+
 While the results reported in the previous section should not depend on the implementation of the different algorithms, the results in this section concern real measurements of both time and memory consumptions of the simulation software used for these simulations.
@@
 Relative difference for costs do not depend much on the code and the machine used for the simulation.
-
-The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
+%
+% The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
 The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
@@
     \item Total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
-    \item Total memory cost should scale $\cM_T$ as $\cO(K \log(T))$.
+    \item Total memory cost $\cM_T$ should scale as $\cO(K \log(T))$.
 \end{itemize}
-Thus, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
+Thus, on the first hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
+%
+On the other hand, when we plot a quantity ($R_T$, $\cT_T$, $\cM_T$) as a function of $T$ (resp. of $K$) we only normalize by $K$ (resp. by $T$).
 
@@
 
-\TODOL{Just one long table for problem 1, not necessary to include other problems}
+% \TODOL{Just one long table for problem 1, not necessary to include other problems}
 
@@
 
-I want to show the real numerical costs in terms of computation time and storage requirement of the various algorithms for single-player bandits.
-
-Maybe I can pick one simple problem family (\eg, evenly spaced in $[\varepsilon,1-\varepsilon]$), and run experiments for $T=1000,5000,10000$ etc, and $K=2,3,5,9,20,30,50,100,1000$ and show the evolution of time/memory as a function of $T$ and $K$.
-
-The goal is to highlight that the simplest (but most efficient) algorithms have a memory cost bounded by $\bigO{K}$ and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
+% I want to show the real numerical costs in terms of computation time and storage requirement of the various algorithms for single-player bandits.
+
+% Maybe I can pick one simple problem family (\eg, evenly spaced in $[\varepsilon,1-\varepsilon]$), and run experiments for $T=1000,5000,10000$ etc, and $K=2,3,5,9,20,30,50,100,1000$ and show the evolution of time/memory as a function of $T$ and $K$.
+
+% The goal is to highlight that the simplest (but most efficient) algorithms have a memory cost bounded by $\bigO{K}$ and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
 
@@
 % --------------------------------------------------------------
-\subsection{Additional experimental results: more recent algorithms}
+\subsection{Description of seven more recent algorithms for experiments in Section~\ref{sec:3:reviewSPAlgorithms}}
 \label{sub:3:additionalExperiments}
@@
 
-We give below the results for these additional algorithms, only for the problem 1,
-first in terms of regret in Table~\ref{table:3:meanRegret_problem1_otherAlgorithms},
-then in terms of computation times in Table~\ref{table:3:time_problem1_otherAlgorithms}
-and for storage costs in Table~\ref{table:3:memory_problem1_otherAlgorithms}.
+% We give below the results for these additional algorithms, only for the problem 1,
+% first in terms of regret in Table~\ref{table:3:meanRegret_problem1_otherAlgorithms},
+% then in terms of computation times in Table~\ref{table:3:time_problem1_otherAlgorithms}
+% and for storage costs in Table~\ref{table:3:memory_problem1_otherAlgorithms}.
 

### 2019-05-07 2-Chapters/3-Chapter/chapter3.tex
Longer simulations are in progress. The notebook to analyze simulations is nice

@@
 We show in the snippet of code~\ref{lst:3:BashCodeToLaunchLargeExperiments} in the Appendix below more details about how to run these experiments.
+
+\TODOL{Dans les figures où je montre une seule valeur de $T$ ou de $K$, et $R_T$, $\cT_T$ ou $\cM_T$ en fonction de $K$ ou de $T$, il faudrait avoir plus de valeurs à montrer, et pas une progression exponentielle...
+J'ai relancé avec :
+    - $T=10000$ et $K = 6, 10, 14, 18, 22, 24, 26, 28, 32$ pour couvrir $2 \dots 32$ avec un pas de $2$,
+    - $K=32$ et $T = 15000, 30000, 35000, 40000, 45000$ pour couvrir $5000 \dots 50000$ avec un pas de $5000$,
+}
 
@@
 
-\TODOL{Use UCB and klUCB}
-\TODOL{Regret usually scales as $\bigO{K}$ and $\bigO{\log T}$}
+\TODOL{Au lieu d'avoir des longues captions pour les figures, je vais écrire quelques paragraphes donnant les mêmes conclusion et faisant référence aux figures.
+Je ferai ça très vite si on est contents des figures.}
+
+\TODOL{Use UCB and klUCB, it's enough to be efficient!}
+\TODOL{Regret usually scales as $\bigO{K}$ and $\bigO{\log T}$ !}
 
@@
 % ----------------------------------------------------------------------------
-\section{Comparing the time and memory costs of various algorithms}
+\section{Comparing real measurements of time and memory costs}
 \label{sec:3:timeAndMemoryCosts}
@@
 
-In this section, we report additional experiment results from the simulations described in the previous section.
+In this section, we report additional experiment results from the simulations described in the previous Section~\ref{sec:3:reviewSPAlgorithms}.
 Instead of studying on the efficiency of the algorithms (\ie, their regret), we report results of real measurements in terms of computational time as well as memory storage.
@@
     \item About the experimental environment:
-    the exact measurements displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
+    the exact measurements used for the figures displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
     As long as the different algorithms and simulations are performed on the same environment, the comparison and conclusions we make from them are fair and do make sense.
+
 \end{itemize}
 
+\pagagraph{About the measurement protocol.}
+%
+To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.
+\begin{itemize}
+    \item For time, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
+    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in\{1,\dots,T\}$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
+    This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.
+    \item For memory, we use two different approaches whether the code runs on a UNIX or a Windows system.
+    On UNIX, the \texttt{resource} module allows to measure the memory of the current process, and by counting the difference in memory used by the \texttt{Evaluator} object before and after the ``for'' loop, we can track the memory used by the algorithm to store all its interior variables.
+\end{itemize}
+In both cases, the ``for'' loop takes some time and stores many variables, but the only difference in terms of both time or memory between two algorithms is explained by the difference in the implementation of the algorithms.
+%     def time()
+% time() -> floating point number
+
+% Return the current time in seconds since the Epoch. Fractions of a second may be present if the system clock provides them.
+
+% sys.getsizeof(pickle.dumps(policy))
+% return resource.getrusage(resource.RUSAGE_SELF).ru_maxrss
+
+\paragraph{Comparing measurements.}
 But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
@@
 \paragraph{Normalizing data?}
-
+%
 The two figures below regret normalized data, in the following way.
@@
 \begin{itemize}
-    \item Final regret $R_T$ should scale as $\cO(K \log(T))$,
-    \item Total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
-    \item Total memory cost $\cM_T$ should scale as $\cO(K \log(T))$.
+    \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
+    \item The total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
+    \item The total memory cost $\cM_T$ should scale as $\cO(K)$, independent of the horizon.
 \end{itemize}
@@
 
+\TODOL{Au lieu d'avoir des longues captions pour les figures, je vais écrire quelques paragraphes donnant les mêmes conclusion et faisant référence aux figures.
+Je ferai ça très vite si on est contents des figures.}
+
 \TODOL{Use UCB and klUCB}
-\TODOL{Regret usually scales as $\bigO{K}$ and $\bigO{\log T}$}
+\TODOL{Time complexity usually scales as $\bigO{KT}$}
+\TODOL{Memory complexity usually scales as $\bigO{K}$}
 
@@
 
-\TODOL{Si on a la place, j'aimerai bien inclure des résultats pour 7 algorithmes de plus, que voici :}
+% \TODOL{Si on a la place, j'aimerai bien inclure des résultats pour 7 algorithmes de plus, que voici :}
 
@@
     \item Algorithm 13 is
-    the anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018} (\texttt{klUCBswitchAnytime}).
-    It is a very recent variant of the \klUCB{} algorithm \cite{KLUCBJournal},
-    and it mimics the \klUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
+    the anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018},
+    is a very recent variant of the \klUCB{} algorithm \cite{KLUCBJournal}
+    (\texttt{klUCBswitchAnytime}).
+    It mimics the \klUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
     It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.

### 2019-05-07 2-Chapters/3-Chapter/logs/launch_simulations.sh

@@ -11,11 +11,28 @@ nbexperiment=0
 
 for N in 100; do
 # for N in $N_JOBS; do
-    for BAYES in False True; do
-    # for BAYES in False; do
-        for T in 1000 2500 5000 10000 20000 25000 50000; do
-        # for T in 1000 2500; do
-            for K in 2 4 8 12 16 20 32; do
+    # for BAYES in False True; do
+    for BAYES in False; do
+        # for T in 1000 2500 5000 10000 20000 25000 50000; do
+        for T in 10000; do
+            # for K in 2 4 8 12 16 20 32; do
+            for K in 6 10 14 18 22 24 26 28 32; do
+            # for K in 2 4 8; do
+                debut=$(date)
+                nbexperiment+=1
+
+                DEBUG=$DEBUG NOPLOTS=$NOPLOTS SAVEALL=$SAVEALL N_JOBS=$N_JOBS \
+                N=$N T=$T K=$K BAYES=$BAYES mymake.sh --noFreeSMS single \
+                && cp -f logs/main_py3_log.txt \
+                            logs/main_py3_log__N${N}_BAYES${BAYES}_T${T}_K${K}__$(date +"%d-%m-%Y_%H-%M-%S").txt \
+                && FreeSMS.py "Expérience #${nbexperiment}, commencée ${debut}, terminée ! Avec N=${N}, BAYES=${BAYES}, T=${T}, K=${K}."
+            done
+        done
+        for T in 15000 30000 35000 40000 45000; do
+        # for T in 10000; do
+            # for K in 2 4 8 12 16 20 32; do
+            for K in 32; do
+            # for K in 6 10 14 18 22 24 26 28 32; do
             # for K in 2 4 8; do
                 debut=$(date)
                 nbexperiment+=1

### 2019-05-07 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done... So tired.

@@
 %
-Finally, Section~\ref{sub:5:SelfishFails} shortly discusses a disappointing result regarding \Selfish{}, with more insights provided in Appendix~\ref{app:5:SelfishFails}.
+Finally, Section~\ref{sub:5:SelfishFails} shortly discusses a disappointing result regarding \Selfish.
+%  with more insights provided in Appendix~\ref{app:5:SelfishFails}.
 
@@
 \begin{proof}
-  Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
+  We analyze the case when the current arm leaves the set $\TopM$ (Line 4):
   \begin{align*}
@@
 However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity.
-It is indeed not very satisfying, from an applicative point of view, the require a prior knowledge
+It is indeed not very satisfying, from an applicative point of view, to require a prior knowledge
 of $\mu^*_{M}-\mu^*_{M+1}$ if one wants to run the \MusicalChair{} algorithm.
@@
   Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
-  \begin{align}
+  \[
     \E_{\mu}[N_2^j(t)]
-    &= \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
+    = \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
     =
@@
     \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
-    \intertext{Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$), and so we can obtain a decomposition with two terms $(\alpha)$ and $(\beta)$:}
-    % so we can use Wald's Lemma \citep{Wald45}
-    &\leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}.
-    % = (\alpha) \times (\beta).
-  \end{align}
+  \]
+  Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$),
+  thus we can use Wald's Lemma \citep{Wald45},
+  to obtain a decomposition with two terms $(\alpha)$ and $(\beta)$:
+  \[
+    \E_{\mu}[N_2^j(t)] \leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}.
+  \]
 
@@
 Our proposed algorithms, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
-We also include a centralized multiple-play \klUCB{} algorithm,
+We also include a (unrealistic) centralized multiple-play \klUCB{} algorithm,
 as defined by \cite{Anantharam87a},
@@
 
+\paragraph{First experiment: \UCB{} vs \klUCB{}.}
+%
 The first experiment is considering $K=9$ arms, with means $\bm{\mu}=[0.1,0.2,\dots,0.9]$, and $M=3$ then $6$ then $9$ players.
@@
 
-We report in Table~\ref{table:5:comparisonUCB_klUCB} the results in terms of mean regret, $\pm$ one standard deviation.
+We report in Table~\ref{table:5:comparisonUCB_klUCB} below the results in terms of mean regret, $\pm$ one standard deviation.
 The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
@@
 
-%
-
-We present results for two bandit instances: one with $K=3$ arms and means
-$\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}.
-For the second instance $K=9$ and $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
+\paragraph{Other experiments.}
+%
+We now present more results for two bandit instances.
+\begin{itemize}
+  \item
+  The first instance is a small problem one with $K=3$ arms and means
+  $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for the case of $M=2$ players.
+  It is used to illustrate that in some unlucky runs, \Selfish{} can suffer a linear regret, as illustrated in Figure~\ref{fig:5:selfish_fail1}.
+  \item
+  The second instance considers $K=9$ arms with means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
   three cases are presented: $M=6$ in Figure~\ref{fig:5:MP__K9_M6_T10000_N1000__4_algos},
   and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
-%
-We also include histograms showing the \emph{distribution} of regret at $t=T$,
+\end{itemize}
+%
+We also include histograms showing the \emph{distribution} of the final regret $R_T$,
 as this allows to check if the regret is indeed small for \emph{each} run of the simulation.
@@
 
-Experiments with a different problem for each repetition,
-that is uniformly sampled $\boldsymbol{\mu} \sim \cU([0,1]^K)$,
-are also considered, in Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \ref{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}.
-This helps to check that no matter the \emph{complexity} of the considered problem (one measure of complexity being the constant in our lower bound),
-\MCTopM{} performs similarly or better than all the other algorithms,
-and \Selfish{} outperforms \rhoRand{} in most cases.
-Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} is a good example
-of outstanding performances of \MCTopM{} and \Selfish{} in comparison to \rhoRand{}.
-%
-Empirically, our proposals were found to always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
-we verified that \MCTopM{} outperforms the state-of-the-art algorithms in many different problems, and is more and more efficient as $M$ and $K$ grows.
-
 %
@@
 %
-\begin{figure}[!h]
+\begin{figure}[!t]
+  \centering
+  % \begin{subfigure}[!h]{0.49\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_5016720151160452442.pdf}
+  % \end{subfigure}
+  % % ~
+  % \begin{subfigure}[!h]{0.49\textwidth}
+  %   \includegraphics[width=1.10\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
+  % \end{subfigure}
+  \caption[Failure case of \Selfish]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
+Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$).
+The regret for the three other algorithms is very small for this problem, always smaller than $100$ here.}
+  \label{fig:5:selfish_fail1}
+  % \vspace*{-15pt}  % XXX remove if problem
+\end{figure}
+
+% %
+% % Regular plots of centralized regrets
+% %
+% \begin{figure}[!t]
+%   \centering
+%   % \begin{subfigure}[!h]{1.00\textwidth}
+%       % \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
+%       \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_1035303196230283176.pdf}
+%       % \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_5016720151160452442.pdf}
+%   % \end{subfigure}
+%   ~
+%   % \begin{subfigure}[!h]{1.00\textwidth}
+%       \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
+%   % \end{subfigure}
+%   \caption[Third failure case of \Selfish]{Regret for $M=3$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
+Axis $x$ is for regret (different scale for each), and the top \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($11$ cases of $R_T \geq T$, out of $1000$).
+The regret for the three other algorithms is very small for this problem, and even appears constant.}
+%   \label{fig:5:selfish_fail3}
+%   % \vspace*{-15pt}  % XXX remove if problem
+% \end{figure}
+
+\begin{figure}[!b]
+  \centering
+  % \begin{subfigure}[!h]{0.85\textwidth}
+    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{0.85\textwidth}
+    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_HistogramsRegret____env1-1_8200873569864822246.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+\RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand.
+The regret of \MCTopM{} is logarithmic, empirically with the same slope as the lower bound.
+The $x$ axis on the regret histograms have different scale for each algorithm.}
+  \label{fig:5:MP__K9_M6_T10000_N1000__4_algos}
+\end{figure}
+
+\begin{figure}[!t]
+  \centering
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.10\textwidth]{MP__K9_M2_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_2643359116089264295.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{1.00\textwidth}
+  %   \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{1.00\textwidth}
+      \includegraphics[width=1.10\textwidth]{MP__K9_M9_T10000_N200__4_algos/all_RegretCentralized_loglog____env1-1_2306423191427933958.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand.
+\MCTopM{} is always among the best algorithms, and for $M$ not too small, its regret seems logarithmic with a constant matching the lower bound.}
+  \label{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}
+\end{figure}
+
+% \begin{figure}[!t]
+%   \centering
+%   % \begin{subfigure}[!h]{0.75\textwidth}
+%       \includegraphics[width=0.75\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
+%   % \end{subfigure}
+%   % ~
+%   % \begin{subfigure}[!h]{0.75\textwidth}
+%       \includegraphics[width=0.75\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
+%   % \end{subfigure}
+%   % ~
+%   % \begin{subfigure}[!h]{0.75\textwidth}
+%       \includegraphics[width=0.75\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
+%   % \end{subfigure}
+%   \caption[Regret for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$]{Regret (in log-log scale), for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$.
+The same observation as in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} can be made.
+\Selfish{} outperforms \MCTopM{} for $M=2$ here.
+Additionally, \MCTopM{} is the only algorithm to not fail dramatically when $M=K$ here.}
+%   \label{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}
+% \end{figure}
+
+\begin{figure}[!t]
+  \centering
+  % \begin{subfigure}[!h]{0.85\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized____env1-1_7803645526012310577.pdf}
+  % \end{subfigure}
+  % ~
+  % \begin{subfigure}[!h]{0.85\textwidth}
+      \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized_semilogy____env1-1_7803645526012310577.pdf}
+  % \end{subfigure}
+  \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem]{Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
+With a perfect knowledge on the horizon and the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly in this simple setting.
+The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can bounded with probability $0.5$ and $0.9$ respectively.
+The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.
+The log-$y$ scale is used to easily differentiate between the different algorithms, and highlight that our proposal outperform both \MEGA{} and \MusicalChair{} by three orders of magnitudes!}
+  \label{fig:5:MP__K9_M3_T123456_N100__8_algos}
+\end{figure}
+
+%
+% Regular plots of centralized regrets
+%
+\begin{figure}[!b]
   % \centering
@@
 
-% Dynamic settings, when $M$ can change in time,
-% were considered in the analysis of both
-We also compared our algorithms to  %multi-player multi-armed bandits algorithms
-with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}, in the presence of sensing, \ie, observation model \modeldeux, for which they were developed.
-Yet these two algorithms were found hard to use efficiently in practice and we show in
-% that were also proposed
-%
-Figure~\ref{fig:5:MP__K9_M3_T123456_N100__8_algos} that they perform poorly in comparison to \rhoRand, \RandTopM{} and \MCTopM.
-%
-\MEGA{} needs a careful tuning of \emph{five} parameters ($c$, $d$, $p_0$, $\alpha$ and $\beta$) to attain reasonable performances.
-No good guideline for tuning them is provided and using \emph{cross validation}, as suggested by \cite{Avner15},
-can be considered out of the scope of \emph{online} sequential learning.
-%In practice, on a fixed instance, the authors do not indicate how to select the parameters, even with a perfect knowledge of the parameters ($\boldsymbol{\mu}$ and $T$).
-%
-\MusicalChair{} consists of a random exploration phase of length $T_0$ after which the players quickly converge to orthogonal strategies targeting the $M$ best arms.
-With probability $1-\delta$, its regret is proved to be ``constant'' (of order $\log(1/\delta)$).
-The theoretical minimal value for $T_0$ depends on $\delta$, on the horizon $T$ and on a lower bound $\epsilon$ on the gap $\Delta = \mu^*_M - \mu^*_{M+1}$, and the practical tuning is hard too.
-%% which are both unavailable in our setting.
-
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{0.85\textwidth}
-    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{0.85\textwidth}
-    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_HistogramsRegret____env1-1_8200873569864822246.pdf}
-  % \end{subfigure}
-  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-\RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand.
-The regret of \MCTopM{} is logarithmic, empirically with the same slope as the lower bound.
-The $x$ axis on the regret histograms have different scale for each algorithm.}
-  \label{fig:5:MP__K9_M6_T10000_N1000__4_algos}
-\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{0.85\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized____env1-1_7803645526012310577.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{0.85\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M3_T123456_N100__8_algos/all_RegretCentralized_semilogy____env1-1_7803645526012310577.pdf}
-  % \end{subfigure}
-  \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem]{Regret (in log-$y$ scale on the right) for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
-With a perfect knowledge on the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly in this simple setting, even with the knowledge of the horizon $T$ for \MusicalChair{}.
-The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can bounded with probability $0.5$ and $0.9$ respectively.
-The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.}
-  \label{fig:5:MP__K9_M3_T123456_N100__8_algos}
-\end{figure}
-
 %
@@
 %
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{0.49\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_5016720151160452442.pdf}
-  % \end{subfigure}
+% \begin{figure}[!t]
+%   \centering
+%   % \begin{subfigure}[!h]{0.49\textwidth}
+%       \includegraphics[width=1.05\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_2643560344649862285.pdf}
+%   % \end{subfigure}
 %   % ~
-  % \begin{subfigure}[!h]{0.49\textwidth}
-  %   \includegraphics[width=1.10\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
-  % \end{subfigure}
-  \caption[Failure case of \Selfish]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
-Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$).
-The regret for the three other algorithms is very small for this problem, always smaller than $100$ here.}
-  \label{fig:5:selfish_fail1}
-  % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
+%   % \begin{subfigure}[!h]{0.49\textwidth}
+%       \includegraphics[width=1.05\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_4683123079851881812.pdf}
+%   % \end{subfigure}
+%   \caption[Second failure case of \Selfish]{Regret, $M=2$ and $M=3$ players, $K=3$ arms, horizon $T=5000$, against $1000$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+\Selfish{} (top curve in \textcolor{darkgreen}{green}) clearly fails in such setting with small $K$.}
+%   \label{fig:5:selfish_fail2}
+%   % \vspace*{-15pt}  % XXX remove if problem
+% \end{figure}
 
@@
 %
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{0.49\textwidth}
-      \includegraphics[width=1.05\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_2643560344649862285.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{0.49\textwidth}
-      \includegraphics[width=1.05\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_4683123079851881812.pdf}
-  % \end{subfigure}
-  \caption[Second failure case of \Selfish]{Regret, $M=2$ and $M=3$ players, $K=3$ arms, horizon $T=5000$, against $1000$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-\Selfish{} (top curve in \textcolor{darkgreen}{green}) clearly fails in such setting with small $K$.}
-  \label{fig:5:selfish_fail2}
-  % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-%
-% Regular plots of centralized regrets
-%
-\begin{figure}[!h]
+
+\begin{figure}[!t]
   \centering
   % \begin{subfigure}[!h]{1.00\textwidth}
-      % \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
-      \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_RegretCentralized____env1-1_1035303196230283176.pdf}
-      % \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_RegretCentralized____env1-1_5016720151160452442.pdf}
-  % \end{subfigure}
-  ~
-  % \begin{subfigure}[!h]{1.00\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K3_M3_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_1035303196230283176.pdf}
-  % \end{subfigure}
-  \caption[Third failure case of \Selfish]{Regret for $M=3$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
-Axis $x$ is for regret (different scale for each), and the top \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($11$ cases of $R_T \geq T$, out of $1000$).
-The regret for the three other algorithms is very small for this problem, and even appears constant.}
-  \label{fig:5:selfish_fail3}
-  % \vspace*{-15pt}  % XXX remove if problem
-\end{figure}
-
-%
-% Regular plots of centralized regrets
-%
-
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{1.00\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
+      \includegraphics[width=1.10\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
   % \end{subfigure}
@@
       % \includegraphics[width=1.00\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3251433209347345969.pdf}
+      \includegraphics[width=1.10\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3251433209347345969.pdf}
   % \end{subfigure}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!t]
   \centering
   % \begin{subfigure}[!h]{1.00\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_RegretCentralized____env1-1_3892966382091165662.pdf}
+      \includegraphics[width=1.10\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_RegretCentralized____env1-1_3892966382091165662.pdf}
   % \end{subfigure}
@@
   % \begin{subfigure}[!h]{1.00\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3892966382091165662.pdf}
+      \includegraphics[width=1.10\textwidth]{MP__K9_M9_T5000_N500__4_algos/all_HistogramsRegret____env1-1_3892966382091165662.pdf}
   % \end{subfigure}
-  \caption[Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=9$ players for $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+  \caption[Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
 This extreme case $M=K$ shows the drastic difference of behavior between \RandTopM{} and \MCTopM, having constant regret, and \rhoRand{} and \Selfish, having large regret.}
@@
 
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{1.00\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M2_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_2643359116089264295.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{1.00\textwidth}
-  %   \includegraphics[width=1.00\textwidth]{MP__K9_M6_T10000_N1000__4_algos/all_RegretCentralized_loglog____env1-1_8200873569864822246.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{1.00\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K9_M9_T10000_N200__4_algos/all_RegretCentralized_loglog____env1-1_2306423191427933958.pdf}
-  % \end{subfigure}
-  \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$,Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand.
-\MCTopM{} is always among the best algorithms, and for $M$ not too small, its regret seems logarithmic with a constant matching the lower bound.}
-  \label{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}
-\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=0.75\textwidth]{MP__K17_M6_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_4163066365888233475.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=0.75\textwidth]{MP__K17_M12_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_3856003705095179548.pdf}
-  % \end{subfigure}
-  % ~
-  % \begin{subfigure}[!h]{0.75\textwidth}
-      \includegraphics[width=0.75\textwidth]{MP__K17_M17_T10000_N100__4_algos/all_RegretCentralized_loglog____env1-1_8969236287861113966.pdf}
-  % \end{subfigure}
-  \caption[Regret for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$]{Regret (in log-log scale), for $M=6, 12, 17$ players for a ``difficult'' problem with $K=17$, and $T=5000$.
-The same observation as in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos} can be made.
-\Selfish{} outperforms \MCTopM{} for $M=2$ here.
-Additionally, \MCTopM{} is the only algorithm to not fail dramatically when $M=K$ here.}
-  \label{fig:5:MP__K17_M6-12-17_T10000_N100__4_algos}
-\end{figure}
-
-\begin{framed}
-  \textcolor{red}{\textbf{FIXME}}
-  Je ne sais pas trop comment inclure toutes ces figures ici, avant la dernière section où je parle des extensions de nos modèles.
-  J'aimerai bien avoir quelques pages simplement remplies des figures, sans avoir le texte de Section~\ref{sec:5:literatureReviewOtherModels} qui commence au milieu...
-  Une idée ?
-\end{framed}
+% Dynamic settings, when $M$ can change in time,
+% were considered in the analysis of both
+We also compared our algorithms to  %multi-player multi-armed bandits algorithms
+with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}, in the presence of sensing, \ie, observation model \modeldeux, for which they were developed.
+Yet these two algorithms were found hard to use efficiently in practice and we show in
+% that were also proposed
+%
+Figure~\ref{fig:5:MP__K9_M3_T123456_N100__8_algos} that they perform poorly in comparison to \rhoRand, \RandTopM{} and \MCTopM.
+%
+\MEGA{} needs a careful tuning of \emph{five} parameters ($c$, $d$, $p_0$, $\alpha$ and $\beta$) to attain reasonable performances.
+No good guideline for tuning them is provided and using \emph{cross validation}, as suggested by \cite{Avner15},
+can be considered out of the scope of \emph{online} sequential learning.
+%In practice, on a fixed instance, the authors do not indicate how to select the parameters, even with a perfect knowledge of the parameters ($\boldsymbol{\mu}$ and $T$).
+%
+\MusicalChair{} consists of a random exploration phase of length $T_0$ after which the players quickly converge to orthogonal strategies targeting the $M$ best arms.
+With probability $1-\delta$, its regret is proved to be ``constant'' (of order $\log(1/\delta)$).
+The theoretical minimal value for $T_0$ depends on $\delta$, on the horizon $T$ and on a lower bound $\epsilon$ on the gap $\Delta = \mu^*_M - \mu^*_{M+1}$, and the practical tuning is hard too.
+%% which are both unavailable in our setting.
+
+\paragraph{Uniformly sampled problems.}
+%
+Experiments with a different problem for each repetition,
+that is uniformly sampled $\boldsymbol{\mu} \sim \cU([0,1]^K)$,
+are also considered, in Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \ref{fig:5:MP__K9_M2_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}.
+This helps to check that no matter the \emph{complexity} of the considered problem (one measure of complexity being the constant in our lower bound),
+\MCTopM{} performs similarly or better than all the other algorithms,
+and \Selfish{} outperforms \rhoRand{} in most cases.
+Figure~\ref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} is a good example
+of outstanding performances of \MCTopM{} and \Selfish{} in comparison to \rhoRand{}.
+%
+Empirically, our proposals were found to always outperform \rhoRand{}, and except for \Selfish{} that can fail badly on problems with small $K$,
+we verified that \MCTopM{} outperforms the state-of-the-art algorithms in many different problems, and is more and more efficient as $M$ and $K$ grows.
+
+Note that more numerical experiments were conducted for the article \cite{Besson2018ALT}, and in particular the last pages of the paper show more figures.
+
+% \begin{framed}
+%   \textcolor{red}{\textbf{FIXME}}
+%   Je ne sais pas trop comment inclure toutes ces figures ici, avant la dernière section où je parle des extensions de nos modèles.
+%   J'aimerai bien avoir quelques pages simplement remplies des figures, sans avoir le texte de Section~\ref{sec:5:literatureReviewOtherModels} qui commence au milieu...
+%   Une idée ?
+% \end{framed}

### 2019-05-07 2-Chapters/5-Chapter/chapter5.tex
Done... So tired.

@@
 
-\TODOL{
-    Pour les extension présentées ici, je veux suivre la même organisation :
+% \TODOL{
+%     Pour les extension présentées ici, je veux suivre la même organisation :
 
-    1. j'explique l'extension, et en quoi c'est difficile (= en quoi notre approche échoue si on fait rien de plus),
+%     1. j'explique l'extension, et en quoi c'est difficile (= en quoi notre approche échoue si on fait rien de plus),
 
-    2. j'explique les travaux existants, et les résultats qu'ils obtiennent e.g. en terme de bornes de regret
+%     2. j'explique les travaux existants, et les résultats qu'ils obtiennent e.g. en terme de bornes de regret
 
-    3. si j'ai déjà le code pour, je peux montrer UNE simulation
-}
+%     3. si j'ai déjà le code pour, je peux montrer UNE simulation
+% }
 
@@
 It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in\{1,\dots,K\}$.
+
+\TODOL{Expliquer pourquoi ce n'est pas très réaliste ? Le paragraphe ci dessous n'est pas très simple à comprendre, je peux faire mieux!}
 
@@
 Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
-What we are interested here is an extension of our model where players can send some bits to one or all the other players, at every time steps or some time steps.
+What we are interested here is an extension of the model where players can send some bits to one or all the other players, at every time steps or some time steps.
 
@@
 %
-Motivated by this weakness, some researchers studied an interesting extension of our model of interest, considering multi-player MAB models with different arms distributions among players.
+Motivated by this weakness, some researchers studied an interesting extension of the model of interest, considering multi-player MAB models with different arms distributions among players.
 %
@@
 We have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
-We have not yet implemented the other extensions of our model in SMPyBandits, but we are interested to do it, and it is one of the major future work left on our library.
+We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future work left on our library.
 For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/X}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.

### 2019-05-08 2-Chapters/3-Chapter/chapter3.tex
...

@@
 It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson decision rule (\ie, choose the arm maximizing one draw of random samples from the posteriors), otherwise it uses the Bayes-UCB decision rule (\ie, choose the arm maximizing the $1-1/t$ quantile).
-    It was proposed in this article and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
+    It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
     In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, is comparable in terms of computation times, but costs more memory.
@@
     BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, and it consists in trying to fix a simple observation.
-    For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute any statistics for each arms (\eg, mean).
-    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arm $1,2$ has $n_1,n_2$ samples and if $n_1>n_2$ (for instance), first sub-sample $n_2$ observations from $n_1$, then compute the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, intuitively it makes more sense to compare these quantities).
+    For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
+    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
     It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
     Analyzing it in the generic case was found to be difficult, and is still an open problem.
-    We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions) (\texttt{BESA} with \texttt{''non\_recursive''=True}).
-    Due to its binary nature, the extension to more than $K=2$ arms is costly in terms of both time and memory, as illustrated below.
+    We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions, \ie, using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
+    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffer from an exponential blow-up when $K$ increases.
 \end{itemize}
 
-We focussed on the most well known algorithms, and two others that are efficient but less known (AdBandits and BESA),
+We focussed here on the most well known algorithms, and two others that are efficient but less known (AdBandits and BESA),
 but to illustrate that SMPyBandits implements many other solutions to the MAB problem proposed in very recent works, we explain in Appendix~\ref{sub:3:additionalExperiments} seven other algorithms.

### 2019-05-09 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
I'm happy about the modifications for Chapter 5:
- no proof for lower-bound
- less figures
- shorter Appendix
- few more details in a proof
- less clutter in some equations

@@
 
-\paragraph{Related work.}
-% FIXME
-\TODOL{Est-ce que je fais ça ici ou à la fin du chapitre 5 ? Ou peut-être en chapitre 1, ça ferait du contenu au chapitre 1...}
+\paragraph{Related works.}
+% FIXME?
+% \TODOL{Est-ce que je fais ça ici ou à la fin du chapitre 5 ? Ou peut-être en chapitre 1, ça ferait du contenu au chapitre 1...}
+%
 We evoke quickly here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
-\cite{darak2016bayesian}
-\cite{Darak16}
-\cite{modiDemo2016}
-\cite{kumar2016two}
-\cite{KumarYadav2018}
-\cite{SawantKumar2018}
-\cite{JoshiKumar2018}
+In $2015$ and $2016$, a PhD student (Navikkumar Modi) and post-doc (Sumit J. Darak) in our team SCEE worked with Christophe Moy and others for the following publications,
+\cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
+Since $2017$, Sumit J.
+Darak and his team at IIIT Delhi have been very active in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+\cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
 
@@
 
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering

### 2019-05-09 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
HARD
I'm happy about the modifications for Chapter 5:
- no proof for lower-bound
- less figures
- shorter Appendix
- few more details in a proof
- less clutter in some equations

--- 2019-04-16 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
@@
-For the curious reader, we wanted to include here an illustration and a description of the three components of our demonstration, presented above in Section~\ref{sec:4:gnuradio}.
+For the curious reader, we wanted to include here an illustration and a description of the three components of the demonstration we presented above in Section~\ref{sec:4:gnuradio}.
+%
+The code corresponding to the following components is written in \texttt{C++} and \texttt{Python} for the hand-written blocks, and using the GUI of the GNU Radio Companion software.
+The flow-graphs are saved as XML files, and the complete code of our demo is fully available online, open-sourced under GPLv3 license, at
+\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit-learning}\\\texttt{-for-iot-networks-with-grc/}}.
+
+\subsubsection*{Random traffic generator}
 
@@
 \begin{figure}[!h]
-	\centering
-    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_PU__v1__simple_grc.png}
+	% \centering
+    \includegraphics[width=1.02\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_PU__v1__simple_grc.png}
     \caption{The \textbf{random traffic generator} flow-graph.}
@@
 
-The \textbf{IoT gateway} is presented in Figure~\ref{fig:4app:USRP_TX_PU__v1__simple_grc}.
+\subsubsection*{IoT gateway}
+
+The \textbf{IoT gateway} is presented in Figure~\ref{fig:4app:USRP_RX_BTS__v1__simple_grc} below.
 %
-The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length).
-All blocks need to know constants about the PHY layer (preamble length and data length).
-The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).
+We wrote the following blocks:
+\begin{itemize}
+    \item
+    1) the \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power),
+    \item
+    2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement),
+    \item
+    and finally
+    3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length).
+\end{itemize}
+All blocks need to know constants about the PHY layer (preamble length and data length). The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).
 
 \begin{figure}[!h]
-	\centering
+	% \centering
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_RX_BTS__v1__simple_grc.png}
@@
 
+\subsubsection*{IoT dynamic device}
+
 Figure~\ref{fig:4app:USRP_TX_SU__v1__simple_grc} below shows the \textbf{IoT dynamic device} flow-graph.
-The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels.
+The hand-written blocks are the
+\begin{itemize}
+    \item
+    1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack),
+    \item
+    2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway,
+    \item
+    and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels.
+\end{itemize}
 Most blocks need to know constants about the PHY layer (preamble length and data length).
@@
 \begin{figure}[!h]
-	\centering
+	% \centering
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_SU__v1__simple_grc.png}

### 2019-05-09 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Talk to Emilie
I'm happy about the modifications for Chapter 5:
- no proof for lower-bound
- less figures
- shorter Appendix
- few more details in a proof
- less clutter in some equations

@@
 	% \includegraphics[width=0.65\linewidth]{Approximation_m10.eps}
-	\includegraphics[width=0.65\linewidth]{Approximation_m10-eps-converted-to.pdf}
+	\includegraphics[width=0.80\linewidth]{Approximation_m10-eps-converted-to.pdf}
 	\caption[Proposed approximation for the probability of collision at the second transmission]{Proposed approximation for the probability of collision at the second transmission.
@@
 		}
-		\caption{Heuristic: first-stage \UCB{} and retransmission in same channel.}
+		\caption{First-stage \UCB{} and retransmission in same channel.}
 		\label{algo:43:UCB}
@@
 
-\paragraph{A warning about notations: local (device) time vs global (gateway) time.}
+\paragraph{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}.}
 %
@@
 In other words, each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $M$ times (\eg, $M=10$) until it receives an \Ack.
+
 To be clear, once again, we repeat that the time steps $t$ in the following algorithms denote the number of time the device sent an uplink packet, and these uplink (re)transmissions can happen at different (real-world) times, more or less spread out in time, it does not matter to the learning part of this model and work.
@@
 
-\vspace*{-3pt}
+% \vspace*{-3pt}
 \begin{figure}[h!]
@@
 		}
-		\caption{Heuristic: uniform random retransmission.}    % A naive
+		\caption{Uniform random retransmission.}    % A naive
 		\label{algo:43:UCBthenRandom}
@@
 
-\vspace*{-3pt}
+% \vspace*{-3pt}
 % \begin{small} % XXX remove if needed
@@
 		}
-		\caption{Heuristic: \UCB{} for retransmission.}    % A naive
+		\caption{\UCB{} for retransmission.}    % A naive
 		\label{algo:43:TwoUCB}
@@
 		}
-		\caption{Heuristic: $K$ different {\UCB}s for retransmission.}
+		\caption{$K$ different {\UCB}s for retransmission.}
 		\label{algo:43:UCBthenKp1}
@@
 The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to simulate the problem before-hand to find a good value of some internal parameter.
-As such, we only consider a delay of $\Delta=100$.}.
+As such, we only consider a delay of $\Delta=100$ in our experiments, and we did not try to optimize it.}.
 %
@@
 		}
-		\caption{Heuristic: Delayed \UCB{} for retransmission.}
+		\caption{Delayed \UCB{} for retransmission.}
 		\label{algo:43:UCBwithDelay}
@@
 
-\paragraph{Other ideas that we did not explore}
+\paragraph{Other ideas that we did not explore.}
 %
@@
 
-\textbf{Remark.}
+\paragraph{Using another bandit policy?}
 %
@@
 We believe that the same empirical results, but more importantly, the same conclusions, could be given if we used Bayes-UCB or other bandit algorithm for the base building block in the different heuristics proposed in Section~\ref{sub:43:heuristics}.
+The goal of this Section was not to optimize on the bandit policy (as we presented in Section~\ref{sec:3:reviewSPAlgorithms} some numerical simulations doing precisely this), but rather to compare the different heuristics, for a fixed bandit policy (for which we preferred to use \UCB{} for simplicity).
 

### 2019-05-09 2-Chapters/4-Chapter/chapter4.tex
I'm happy about the modifications for Chapter 5:
- no proof for lower-bound
- less figures
- shorter Appendix
- few more details in a proof
- less clutter in some equations

@@
 
+% \newpage  % WARNING ?
 % ----------------------------------------------------------------------------
@@
 
-We give in this section conclusions about the two models presented above, one for long-life IoT devices (without retransmission) and for improved QoS devices (\ie, with retransmissions), and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.
+% We give in this section conclusions about the two models presented above, one for long-life IoT devices (without retransmission) and for improved QoS devices (\ie, with retransmissions), and about our demonstration, as well as directions of future works, some of which are currently being studied by researchers of our team.
 
-\TODOL{Je devrais réécrire toute cette conclusion, si j'ai le temps, pour faire mieux qu'une suite de 3 conclusions et 3 future works, et être plus concis et moins ``short sighted'' comme disait Emilie.}
+% \TODOL{Je devrais réécrire toute cette conclusion, si j'ai le temps, pour faire mieux qu'une suite de 3 conclusions et 3 future works, et être plus concis et moins ``short sighted'' comme disait Emilie.}
 
-\subsection*{Conclusions}
+% \subsection*{Conclusions}
 
@@
 Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
-This approach can also extend IoT devices battery life, as they suffer less from failed transmissions.
-We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the presences many ``intelligent'' end-devices which follow a random activation process.
-
-This is both a surprising and a very encouraging result, showing that application of bandit algorithms tailored for a stochastic model is still useful in broader settings.
-The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{can not} communicate with each others, and \emph{do not} know the system parameters.
+Similarly, if the number of devices remain constant, increasing the successful transmission rate directly extends the IoT devices battery life, as they suffer less from failed transmissions.
+We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
+%
+This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
+The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each others, and \emph{do not} know the system parameters.
 
@@
 
-In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
-If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more interesting.
+In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency.
+% , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
+If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
 The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
-We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
-
+We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
+%
 We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
@@
 
-Future works related to this chapter include the following directions.
+% Future works related to this chapter include the following directions.
 

### 2019-05-09 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
I'm happy about the modifications for Chapter 5:
- no proof for lower-bound
- less figures
- shorter Appendix
- few more details in a proof
- less clutter in some equations

@@
 % -----------------------------------------------------------------
-\section{An improved asymptotic regret lower bound (for some algorithms)}
+% \section{An improved asymptotic regret lower bound (for some algorithms)}
+\section{Decomposing the regret to get an intuition for the design of efficient decentralized algorithms}
 \label{sec:5:lowerbound}
@@
   thanks to the modifications proposed by Emilie Kaufmann and Abbas Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
-  The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which contains both \RhoRand{} and our two proposals \RandTopM{} and \MCTopM{} but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
+  The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which \emph{should} contain \RhoRand, \RandTopM{} and \MCTopM, but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
+  Proving that the proposed algorithms are in this class is left as an open question.
 % \end{framed}
@@
 However, a careful manipulation of this decomposition shows that the regret is always lower bounded by term \ref{eq:5:term1}.
+The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} in Appendix illustrate two cases of $M<K$ and $M=K$ and the different impact of the three terms (\ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}) on the regret.
 
@@
 We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
-Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
-This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
+% Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
+% This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
 
@@
 As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the ``collision information term'' is zero for any algorithm $\rho$.
-However, we wanted to be still able to present the derivations of our lower-bound, mainly because its proof and its form give insights on how to design an asymptotically optimal algorithm $\rho$, and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
+However, we wanted to be still able to present the derivations of our lower-bound, mainly because its proof and its form give important insights on how to design an asymptotically optimal algorithm $\rho$.
+% and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
 
@@
 
-Fix a problem $\bm{\mu}$, and as we do in Appendix~\ref{proof:5:BetterLowerBound}, introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
+Fix a problem $\bm{\mu}$,
+% and as we do in Appendix~\ref{proof:5:BetterLowerBound},
+then introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
 Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
@@
 The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
-The theorem stated below is proved in Appendix~\ref{proof:5:BetterLowerBound}, and it also yields an asymptotic logarithmic lower bound on the regret.
+The theorem stated below is proved in the Appendix of \cite{Besson2018ALT},
+% \ref{proof:5:BetterLowerBound},
+and it also yields an asymptotic logarithmic lower bound on the regret.
 
@@
 
+\paragraph{Price of decentralized learning?}
+%
 Interestingly, our lower bound is exactly a multiplicative constant factor $M$ away from the lower bound given by \cite{Anantharam87a} for centralized algorithms (which is clearly a simpler setting).
@@
 \begin{equation}
-  A^j(t) = \argmax{ k \in \{1,\dots,K\}} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
+  A^j(t) = \argmax_{ k \in \{1,\dots,K\}} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
 \end{equation}
@@
 
-We show in Section~\ref{sub:5:SelfishFails} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
+We illustrate in Section~\ref{sub:5:SelfishFails} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
 which precludes the possibility of a logarithmic regret for any problem.
@@
 \end{equation}
+
+\paragraph{The \RandTopM{} algorithm.}
+%
+We precisely state our first proposal below in Algorithm~\ref{algo:5:RandTopM}.
 
@@
 
-\paragraph{The \RandTopM{} algorithm.}
-%
-First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
-It is essentially a refinement over the \RhoRand{} algorithm, to not use the indirection of ranks, and a simpler version of \MCTopM.
-The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
+\RandTopM{} is essentially a refinement over \RhoRand, to not use the indirection of ranks, and a simpler version of \MCTopM.
+The difference with \MCTopM{} is that the later introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
 %
@@
   \resizebox{1.00\textwidth}{!}{
-  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.5]
+  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.8]
       %
@@
   Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
-  \begin{align*}
-    \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
-    &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
-    &\leq \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(g_{m^*}(t) < \mu_m^* \right) \\
-    %
-    & \leq \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
-    + \sum_{m=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( g_{m^*}(t) < \mu_m^* \right)
-    \hspace*{80pt} (\heartsuit)  % TODO cleaner?
-  \end{align*}
+  \begin{align}
+    \E_{\mu}[T_k^j(T)] &\leq
+    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in \{1,\dots,M\}: \; g_{m_1^*}(t) < \mu_m^* \right) + \notag \\
+    \;\;\;\;
+    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in \{1,\dots,M\}:\; g_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in \{1,\dots,M\}: \;  g_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
+    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(g_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in \{1,\dots,M\}:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
+    %
+    & \leq \sum_{m_1=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( g_{m_1^*}(t) < \mu_{m_1}^* \right)
+    + \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+    \label{eq:5:sumOfProbaForTkj}
+  \end{align}
   where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.
 
-  Now each of the two terms in the right hand side of $(\heartsuit)$ can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
-  The rightmost term in $(\heartsuit)$ can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact~1 in Appendix~A of \cite{KLUCBJournal}.
+  Now each of the two terms in the right hand side of \eqref{eq:5:sumOfProbaForTkj} can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
+  The leftmost term in \eqref{eq:5:sumOfProbaForTkj} can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact~1 in Appendix~A of \cite{KLUCBJournal}.
 
@@
 
-  The leftmost term in $(\heartsuit)$ can be controlled using Lemma~\ref{lem:5:Fact2KLUCB}, that is a direct consequence of the proof of Fact~2 in Appendix~A of \cite{KLUCBJournal}.
+  The rightmost term in \eqref{eq:5:sumOfProbaForTkj} can be controlled using Lemma~\ref{lem:5:Fact2KLUCB}, that is a direct consequence of the proof of Fact~2 in Appendix~A of \cite{KLUCBJournal}.
   Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).

### 2019-05-09 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
I'm happy about the modifications for Chapter 5:
- no proof for lower-bound
- less figures
- shorter Appendix
- few more details in a proof
- less clutter in some equations

@@
 
-% -----------------------------------------------------------------
-% -----------------------------------------------------------------
-\subsection{Lower bound: proof of Theorem~\ref{thm:5:BetterLowerBound}}
-\label{proof:5:BetterLowerBound}
+% % -----------------------------------------------------------------
+% % -----------------------------------------------------------------
+% \subsection{Lower bound: proof of Theorem~\ref{thm:5:BetterLowerBound}}
+% \label{proof:5:BetterLowerBound}
 
-The lower bound that we present relies on the following \emph{change-of-distribution} lemma that we prove in the next section, following recent arguments from \cite{Garivier16TrueShape} that have to be adapted to incorporate the collision information.
+% The lower bound that we present relies on the following \emph{change-of-distribution} lemma that we prove in the next section, following recent arguments from \cite{Garivier16TrueShape} that have to be adapted to incorporate the collision information.
 
-\begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-player bandit models denoted by $\bm{\mu}$ and $\bm{\lambda}$ respectively, it holds that
-\begin{equation}
-  \sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).
-\end{equation}
-\end{lemma}
+% \begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-player bandit models denoted by $\bm{\mu}$ and $\bm{\lambda}$ respectively, it holds that
+% \begin{equation}
+%   \sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).
+% \end{equation}
+% \end{lemma}
 
-Let $k$ be a sub-optimal arm under $\bm{\mu}$,
-fix $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$,
-% We define the bandit instance $\bm{\lambda}$ such that
-and let $\bm{\lambda}$ be the bandit instance such that
-\[\left\{\begin{array}{ccl}
-          \lambda_\ell & = & \mu_\ell \ \ \  \text{for all } \ell \neq k, \\
-          \lambda_k & = & \mu_{M^*} + \varepsilon.
-         \end{array}
-\right.\]
-Clearly, $\bm{\lambda}\in\cP_M$ also,
-and the set of $M$ best arms under $\bm{\mu}$ and $\bm{\lambda}$ differ by one arm: if \Mbest$_{\bm{\mu}} = \{1^*,\dots,M^*\}$ then $\Mbest_{\bm{\lambda}} = \{1^*,\dots,(M-1)^*,k\}$.
-Thus, one expects the ($\cF^j_T$-mesurable) event
-\[A_T = \left(T_k^j(T) > \frac{T}{2M}\right)\]
-to have a small probability under $\bm{\mu}$ (under which $k$ is sub-optimal) and a large probability under $\bm{\lambda}$ (under which $k$ is one of the optimal arms, and is likely to be drawn a lot).
+% Let $k$ be a sub-optimal arm under $\bm{\mu}$,
+% fix $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$,
+% % We define the bandit instance $\bm{\lambda}$ such that
+% and let $\bm{\lambda}$ be the bandit instance such that
+% \[\left\{\begin{array}{ccl}
+%           \lambda_\ell & = & \mu_\ell \ \ \  \text{for all } \ell \neq k, \\
+%           \lambda_k & = & \mu_{M^*} + \varepsilon.
+%          \end{array}
+% \right.\]
+% Clearly, $\bm{\lambda}\in\cP_M$ also,
+% and the set of $M$ best arms under $\bm{\mu}$ and $\bm{\lambda}$ differ by one arm: if \Mbest$_{\bm{\mu}} = \{1^*,\dots,M^*\}$ then $\Mbest_{\bm{\lambda}} = \{1^*,\dots,(M-1)^*,k\}$.
+% Thus, one expects the ($\cF^j_T$-mesurable) event
+% \[A_T = \left(T_k^j(T) > \frac{T}{2M}\right)\]
+% to have a small probability under $\bm{\mu}$ (under which $k$ is sub-optimal) and a large probability under $\bm{\lambda}$ (under which $k$ is one of the optimal arms, and is likely to be drawn a lot).
 
-Applying the inequality in Lemma~\ref{lem:5:CD}, and noting that the sum in the left-hand side reduces to on term as there is a single arm whose distribution is changed, one obtains
-\begin{eqnarray*}
-  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \kl \left(\Pr_{\bm{\mu}}(A_T),\Pr_{\bm{\lambda}}(A_T)\right),\nonumber \\
-  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - \Pr_{\bm{\mu}}(A_T)\right) \log \left(\frac{1}{\Pr_{\bm{\lambda}}(\overline{A_T})}\right) - \log(2),
-\end{eqnarray*}
-using the fact that the binary KL-divergence satisfies $\kl(x,y) = \kl(1-x,1-y)$ as well as the inequality  $\kl(x, y) \geq x \log\left(1/y\right) - \log(2)$, proved by \cite{Garivier16TrueShape}.
-Now, using Markov inequality yields
-\begin{align*}
-  \Pr_{\bm{\mu}}\left(A_T\right)
-    & \leq 2M\frac{\E_{\bm{\mu}}\left[T^j_k(T)\right]}{T} =: x_T, \\
-  \Pr_{\bm{\lambda}}\left(\overline{A_T}\right)
-    & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_{\bm{\lambda}}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
-\end{align*}
-which defines two sequences $x_T$ and $y_T$, such that
-\begin{eqnarray}\label{eq:5:CqCD}
-  \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - x_T\right) \log \left(\frac{T}{y_T}\right) - \log(2).
-\end{eqnarray}
-The strong uniform efficiency assumption (see Definition~\ref{def:5:DecentralizedUniformEfficiency}) further tells us that $x_T \rightarrow 0$ (as $\E_{\bm{\mu}}[T_k^j(T)] = o(T^\alpha)$ for all $\alpha$) and $y_T = o(T^\alpha)$ when $T \to \infty$, for all $\alpha \in (0,1)$.
-As a consequence, observe that $\log(y_T)/\log(T) \rightarrow 0$ when $T$ tends to infinity and
-\[
-  \frac{\left(1 - x_T\right) \log \left({T}/{y_T}\right) - \log(2)}{\log(T)} = 1-x_T - (1-x_T) \frac{\log(y_T)}{\log(T)} - \frac{\log(2)}{\log(T)}
-\]
-tends to one when $T$ tends to infinity.
-From Equation~\eqref{eq:5:CqCD}, this yields
-\begin{equation}\label{eq:5:CqCD2}
-  \liminf_{T\to\infty} \frac{\E_{\bm{\mu}}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k,\mu_{M^*}+\varepsilon)},
-\end{equation}
-for all $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$.
-Letting $\varepsilon$ go to zero gives the conclusion (as $\kl$ is continuous).
+% Applying the inequality in Lemma~\ref{lem:5:CD}, and noting that the sum in the left-hand side reduces to on term as there is a single arm whose distribution is changed, one obtains
+% \begin{eqnarray*}
+%   \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \kl \left(\Pr_{\bm{\mu}}(A_T),\Pr_{\bm{\lambda}}(A_T)\right),\nonumber \\
+%   \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - \Pr_{\bm{\mu}}(A_T)\right) \log \left(\frac{1}{\Pr_{\bm{\lambda}}(\overline{A_T})}\right) - \log(2),
+% \end{eqnarray*}
+% using the fact that the binary KL-divergence satisfies $\kl(x,y) = \kl(1-x,1-y)$ as well as the inequality  $\kl(x, y) \geq x \log\left(1/y\right) - \log(2)$, proved by \cite{Garivier16TrueShape}.
+% Now, using Markov inequality yields
+% \begin{align*}
+%   \Pr_{\bm{\mu}}\left(A_T\right)
+%     & \leq 2M\frac{\E_{\bm{\mu}}\left[T^j_k(T)\right]}{T} =: x_T, \\
+%   \Pr_{\bm{\lambda}}\left(\overline{A_T}\right)
+%     & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_{\bm{\lambda}}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
+% \end{align*}
+% which defines two sequences $x_T$ and $y_T$, such that
+% \begin{eqnarray}\label{eq:5:CqCD}
+%   \E_{\bm{\mu}}\left[T_k^j(T)\right] \kl (\mu_k, \mu_{M^*}+\varepsilon) &\geq & \left(1 - x_T\right) \log \left(\frac{T}{y_T}\right) - \log(2).
+% \end{eqnarray}
+% The strong uniform efficiency assumption (see Definition~\ref{def:5:DecentralizedUniformEfficiency}) further tells us that $x_T \rightarrow 0$ (as $\E_{\bm{\mu}}[T_k^j(T)] = o(T^\alpha)$ for all $\alpha$) and $y_T = o(T^\alpha)$ when $T \to \infty$, for all $\alpha \in (0,1)$.
+% As a consequence, observe that $\log(y_T)/\log(T) \rightarrow 0$ when $T$ tends to infinity and
+% \[
+%   \frac{\left(1 - x_T\right) \log \left({T}/{y_T}\right) - \log(2)}{\log(T)} = 1-x_T - (1-x_T) \frac{\log(y_T)}{\log(T)} - \frac{\log(2)}{\log(T)}
+% \]
+% tends to one when $T$ tends to infinity.
+% From Equation~\eqref{eq:5:CqCD}, this yields
+% \begin{equation}\label{eq:5:CqCD2}
+%   \liminf_{T\to\infty} \frac{\E_{\bm{\mu}}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k,\mu_{M^*}+\varepsilon)},
+% \end{equation}
+% for all $\varepsilon \in \left(0, \mu_{M-1}^* - \mu_M^*\right)$.
+% Letting $\varepsilon$ go to zero gives the conclusion (as $\kl$ is continuous).
 
-% -----------------------------------------------------------------
+% % -----------------------------------------------------------------
 
-\subsubsection{Proof of Lemma~\ref{lem:5:CD}}
+% \subsubsection{Proof of Lemma~\ref{lem:5:CD}}
 
-Under observation model \modelun{} and \modeldeux, the strategy $A^j(t)$ decides which arm to play based on the information contained in $\cO_{t-1}$, where $\cO_0 = U_0$ and
-\[\forall t > 0, \ \ \cO_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
-where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $\{1,\dots,M\}$ for new ranks or arms.} useful to select $A^j(t)$.
-Formally, one can say that $A^j(t)$ is $\sigma(\cO_{t-1})$ measurable, if $\sigma(\cO_t)$ denotes the sigma-Algebra generated by the observations $\cO_t$, and as $\cF^j(t) \subseteq \sigma(\cO_{t})$, with an equality under observation model \modelun.
+% Under observation model \modelun{} and \modeldeux, the strategy $A^j(t)$ decides which arm to play based on the information contained in $\cO_{t-1}$, where $\cO_0 = U_0$ and
+% \[\forall t > 0, \ \ \cO_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
+% where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $\{1,\dots,M\}$ for new ranks or arms.} useful to select $A^j(t)$.
+% Formally, one can say that $A^j(t)$ is $\sigma(\cO_{t-1})$ measurable, if $\sigma(\cO_t)$ denotes the sigma-Algebra generated by the observations $\cO_t$, and as $\cF^j(t) \subseteq \sigma(\cO_{t})$, with an equality under observation model \modelun.
 
-Under two bandit models $\bm{\mu}$ and $\bm{\lambda}$, we let $\Pr_{\bm{\mu}}^{\cO_t}$ (resp.
+% Under two bandit models $\bm{\mu}$ and $\bm{\lambda}$, we let $\Pr_{\bm{\mu}}^{\cO_t}$ (resp.
 $\Pr_{\bm{\lambda}}^{\cO_t}$) be the distribution of the observations under model $\bm{\mu}$ (resp.
 $\bm{\lambda}$), given a fixed algorithm.
-Using the exact same technique as \cite{Garivier16TrueShape}, called the ``contraction of entropy'' principle, one can establish that for any event $A$ that is $\sigma(\cO_t)$-measurable\footnote{In the work of \cite{Garivier16TrueShape}, the statement is more general and the probability of an event $A$ is replaced by the expectation of any $\cF_T$-measurable random variable $Z$ bounded in $[0,1]$.},
-\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) \geq \kl\left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).\]
+% Using the exact same technique as \cite{Garivier16TrueShape}, called the ``contraction of entropy'' principle, one can establish that for any event $A$ that is $\sigma(\cO_t)$-measurable\footnote{In the work of \cite{Garivier16TrueShape}, the statement is more general and the probability of an event $A$ is replaced by the expectation of any $\cF_T$-measurable random variable $Z$ bounded in $[0,1]$.},
+% \[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) \geq \kl\left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right).\]
 
-The next step is to relate the complicated KL-divergence $\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right)$ to the number of arm selections.
-Proceeding similarly as \cite{Garivier16TrueShape}, one can write, using the chain rule for KL-divergence, that
-\begin{equation}\label{eq:5:Iterative}\KL\left(\Pr_{\bm{\mu}}^{\cO_{t}},\Pr_{\bm{\lambda}}^{\cO_{t}}\right) =  \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t,U_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t,U_t | \cO_{t-1}}\right).\end{equation}
-Now observe that conditionally to $\cO_{t-1}$, $U_t$, $Y_t$ and $C_t$ are independent, as once the selected arm is known, the value of the sensing $Y_t$ does not influence the other players selecting that arm, and $U_t$ is some exogenous randomness.
-Using further that the distribution of $U_t$ is the same under $\bm{\mu}$ and $\bm{\lambda}$, one obtains
-\begin{equation}\KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t | \cO_{t-1}}\right) =
- \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right).\label{eq:5:DecKL}
-\end{equation}
+% The next step is to relate the complicated KL-divergence $\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right)$ to the number of arm selections.
+% Proceeding similarly as \cite{Garivier16TrueShape}, one can write, using the chain rule for KL-divergence, that
+% \begin{equation}\label{eq:5:Iterative}\KL\left(\Pr_{\bm{\mu}}^{\cO_{t}},\Pr_{\bm{\lambda}}^{\cO_{t}}\right) =  \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t,U_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t,U_t | \cO_{t-1}}\right).\end{equation}
+% Now observe that conditionally to $\cO_{t-1}$, $U_t$, $Y_t$ and $C_t$ are independent, as once the selected arm is known, the value of the sensing $Y_t$ does not influence the other players selecting that arm, and $U_t$ is some exogenous randomness.
+% Using further that the distribution of $U_t$ is the same under $\bm{\mu}$ and $\bm{\lambda}$, one obtains
+% \begin{equation}\KL\left(\Pr_{\bm{\mu}}^{Y_{t},C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t},C_t | \cO_{t-1}}\right) =
+%  \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) + \KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right).\label{eq:5:DecKL}
+% \end{equation}
 
-The first term in \eqref{eq:5:DecKL} can be rewritten using the same argument as \cite{Garivier16TrueShape}, that relies on the fact that conditionally to $\cO_{t-1}$, $Y_t$ is a Bernoulli distribution with mean $\mu_{A^j(t)}$ under the instance $\bm{\mu}$ and $\lambda_{A^j(t)}$ under the instance $\bm{\lambda}$:
-\begin{eqnarray*}
-\KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right)  & = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
-& = & \E_{\bm{\mu}} \left[\kl \left(\mu_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
-& = & \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right],
-\end{eqnarray*}
-We now show that second term in \eqref{eq:5:DecKL} is zero:
-\begin{eqnarray*}
-\KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right)
-& = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
-& = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \bigcup_{j' \neq j} \cO_{t-1}^{j'}\right]| \cO_{t-1} \right]\right], \\
-\end{eqnarray*}
-where $\cO_{t-1}^{j'}$ denote the information available to player $j' \neq j$.
-\textcolor{red}{By hypothesis}, the ``collision information term'' is assumed to be a $\smallO{\log(T)}$,
-and thus $\KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right) = \smallO{\log(T)}$.
+% The first term in \eqref{eq:5:DecKL} can be rewritten using the same argument as \cite{Garivier16TrueShape}, that relies on the fact that conditionally to $\cO_{t-1}$, $Y_t$ is a Bernoulli distribution with mean $\mu_{A^j(t)}$ under the instance $\bm{\mu}$ and $\lambda_{A^j(t)}$ under the instance $\bm{\lambda}$:
+% \begin{eqnarray*}
+% \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right)  & = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{Y_{t}| \cO_{t-1}},\Pr_{\bm{\lambda}}^{Y_{t}| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
+% & = & \E_{\bm{\mu}} \left[\kl \left(\mu_{A^j(t)},\lambda_{A^j(t)}\right)\right] \\
+% & = & \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right],
+% \end{eqnarray*}
+% We now show that second term in \eqref{eq:5:DecKL} is zero:
+% \begin{eqnarray*}
+% \KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right)
+% & = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \cO_{t-1} \right]\right] \\
+% & = & \E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[\E_{\bm{\mu}}\left[ \KL\left(\Pr_{\bm{\mu}}^{C_t| \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t| \cO_{t-1}}\right) | \bigcup_{j' \neq j} \cO_{t-1}^{j'}\right]| \cO_{t-1} \right]\right], \\
+% \end{eqnarray*}
+% where $\cO_{t-1}^{j'}$ denote the information available to player $j' \neq j$.
+% \textcolor{red}{By hypothesis}, the ``collision information term'' is assumed to be a $\smallO{\log(T)}$,
+% and thus $\KL\left(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}},\Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}\right) = \smallO{\log(T)}$.
 
-Putting things together, we showed that
-\[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) = \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right] + \smallO{\log(T)}.\]
-Iterating this equality and using that $\KL\left(\Pr_{\bm{\mu}}^{\cO_{0}},\Pr_{\bm{\lambda}}^{\cO_{0}}\right)=0$ yields that
-\[\sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right) + \smallO{\log(T)},\]
-for all $A \in \sigma(\cO_T)$, in particular for all $A \in \cF^j_T$.
+% Putting things together, we showed that
+% \[\KL\left(\Pr_{\bm{\mu}}^{\cO_t},\Pr_{\bm{\lambda}}^{\cO_t}\right) = \KL\left(\Pr_{\bm{\mu}}^{\cO_{t-1}},\Pr_{\bm{\lambda}}^{\cO_{t-1}}\right) + \E_{\bm{\mu}}  \left[\sum_{k=1}^K \indic{(A^j(t)=k)} \kl \left(\mu_{k},\lambda_{k}\right)\right] + \smallO{\log(T)}.\]
+% Iterating this equality and using that $\KL\left(\Pr_{\bm{\mu}}^{\cO_{0}},\Pr_{\bm{\lambda}}^{\cO_{0}}\right)=0$ yields that
+% \[\sum_{k=1}^{K}\E_{\bm{\mu}}\left[T_k^j(T)\right]\kl(\mu_k,\lambda_k) \geq \kl \left(\Pr_{\bm{\mu}}(A),\Pr_{\bm{\lambda}}(A)\right) + \smallO{\log(T)},\]
+% for all $A \in \sigma(\cO_T)$, in particular for all $A \in \cF^j_T$.
 
@@
 % -----------------------------------------------------------------
-\subsection{Additional figures}
+% \subsection{Additional figures}
 \label{app:5:moreplots}
 
-We include here plots missing from Section~\ref{sec:5:experiments},
-and some additional numerical results.
+% We include here plots missing from Section~\ref{sec:5:experiments},
+% and some additional numerical results.
 
 % -----------------------------------------------------------------
-\subsubsection{An illustration of the lower bound}
+\subsection*{Illustrating the lower bound}
 \label{app:5:illustrationLowerBound}
@@
 
-\paragraph{Additional figures for numerical simulations.}
+% -----------------------------------------------------------------
+\subsection*{Illustrating the regret decomposition}
+\label{app:5:illustrationRegretDecomposition}
 
-Figures~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
-show the regret $R_T(\boldsymbol{\mu}, M, \rho)$
+The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} below
+shows the regret $R_T(\boldsymbol{\mu}, M, \rho)$
 on the same example problem $\boldsymbol{\mu}$, with $K = 9$ arms and respectively $M = 6$, or $9$ players, for \Selfish-\klUCB.

### 2019-05-09 2-Chapters/5-Chapter/chapter5.tex
I'm happy about the modifications for Chapter 5:
- no proof for lower-bound
- less figures
- shorter Appendix
- few more details in a proof
- less clutter in some equations

@@
 We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
+
 For this first model, we present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
@@
 We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
-We then analyze our \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
+We then analyze the \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
 %  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
@@
 
-Finally, we review other extensions of our model, and give a overview of the current state-of-the-art.
-For some extensions, we discuss how to adapt our proposal and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
+Finally, we review other extensions of the models, and give a overview of the current state-of-the-art.
+For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
 
@@
 % ----------------------------------------------------------------------------
-\section{Literature review of many extensions of our models}
+\section{Literature review of many extensions of the models}
 \label{sec:5:literatureReviewOtherModels}
@@
 Many extensions of the simpler model of multi-player bandits have been considered.
-The number of players $M$ can be fixed but initially unknown to the players, the sensing information can be absent (like our \modeltrois{} presented above), there could be communications between players (happening at each time step or only occasionally), or $M$ could evolve over time to model the possibility of arrivals or departures of players (\ie, connexions and disconnections of devices in a wireless networks).
+The number of players $M$ can be fixed but initially unknown to the players, the sensing information can be absent (like the model \modeltrois{} presented above), there could be communications between players (happening at each time step or only occasionally), or $M$ could evolve over time to model the possibility of arrivals or departures of players (\ie, connexions and disconnections of devices in a wireless networks).
 Also inspired by real-world wireless networks, the arm means may vary among players, for instance to model players located at different distance of the gateway, and we can also consider networks with some ``jammers'' whose purpose is not to communicate to the gateway in a collaborative way, like the $M$ devices, but to interfere with the communications of the $M$ devices.
@@
 Without removing this hypothesis (we study this case below in Section~\ref{sub:5:arrivalDepartures}), it is interesting to remark that we made no hypothesis on whether the players can known this value $M$ or not.
-Our proposals, \RandTopM{} and \MCTopM, both assume to know $M$ beforehand, like it was done for our main inspiration \RhoRand.
+Our proposals, \RandTopM{} and \MCTopM, both assume to know $M$ beforehand, like it was done for their main inspiration, \RhoRand.
 
@@
 %
-We can start by asking whether our most efficient proposal, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different than the real number of players $M$.
+We can start by asking whether the most efficient algorithm, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different than the real number of players $M$.
 Even though we did not include numerical simulations for this case in Section~\ref{sec:5:experiments} above, we did some tests, which confirmed two disappointing results that were also proved analytically.
@@
 Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
-We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for our proposals \RandTopM{} and \MCTopM.
-Building from the theoretical analysis given for \rhoRandEst{} in \cite{Anandkumar11}, and our analysis of \MCTopM-\klUCB, we believe it is possible to show that the aforementioned extension of our proposal also achieves sub-linear regret without requiring players to know $M$ in the beginning of the bandit game.
+We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for \RandTopM{} and \MCTopM.
+Building from the theoretical analysis given for \rhoRandEst{} in \cite{Anandkumar11}, and the analysis of \MCTopM-\klUCB, we believe it is possible to show that the aforementioned extension also achieves sub-linear regret without requiring players to know $M$ in the beginning of the bandit game.
 More precisely, we believe that \MCTopM-\klUCB{} can still give order-optimal logarithmic regret if $M$ players use it, and this extension of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is not included due to space constraints.
@@
 %
-Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension of our proposal \MCTopM-\klUCB.
+Additional simulations, for increasing time horizons $T$, confirmed the expected order-optimal regret of this extension of \MCTopM-\klUCB.
 
@@
 
-As reminded above, our model assumes the number of players $M$ remain fixed during all the bandit game.
+As reminded above, the model assumes the number of players $M$ remain fixed during all the bandit game.
 However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
@@
 We believe that we could easily prove that a sub-linear regret is achievable, if all players run \MCTopM-\klUCB{} (and if currently active players restart their memory of the past observations every $L$ time steps).
-As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, applying the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to our proposal \MCTopM{} would also give a small regret upper-bound.
+As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, applying the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to \MCTopM{} would also give a small regret upper-bound.
 
@@
 Players can only observe $Y_k(t)$, the product of the \iid{} random sensing information from arm $k$ and of the no-collision indicator.
-This model seems harder than the model with sensing information, and even though our proposal \Selfish{} works fine in simulations (in terms of average regret), we proved that it can yield linear regret.
+This model seems harder than the model with sensing information, and even though the proposal \Selfish{} works fine in simulations (in terms of average regret), we proved that it can yield linear regret.
 In our paper \cite{Besson2018ALT} as well as the previous sections, it was left as a future work to know if an algorithm can achieve sub-linear regret in this harder model without sensing information.
 
-Inspired by our article \cite{Besson2018ALT},
+Inspired by \cite{Besson2018ALT},
 Boursier and Perchet studied this question in the summer $2018$ following its publication \cite{BoursierPerchet18}.
@@
 They also study in Section~4.3 an extension of the model for the case with more players than arms, but we do not give more details on this aspect.
-Their definition of a centralized regret for this case is an interesting and natural generalization of our definition, and they also proposed an algorithm achieving $\bigO{M K \log(T) \exp(4M/K) / \Delta^2}$ regret in this case.
+Their definition of a centralized regret for this case is an interesting and natural generalization of the definition, and they also proposed an algorithm achieving $\bigO{M K \log(T) \exp(4M/K) / \Delta^2}$ regret in this case.
 Finally, they also proposed in Section~4.4 an extension of their algorithm to estimate the number of players $M$, which is analyzed as for the Musical Chair algorithm in \cite{Rosenski16}, and also achieve logarithmic regret.
@@
 Finally, the very recent work of \cite{Zafaruddin2019} is the first one to present experiments of reinforcement learning done on simulated LTE and 5G channels.
-The mathematics behind their model are actually quite close to our model, but they explain it in terms of OFDMA and Quality-of-Service (QoS).
+The mathematics behind their model are actually quite close to the model, but they explain it in terms of OFDMA and Quality-of-Service (QoS).
 Their algorithm is essentially based on a pre-agreement of the $M$ players, that will deterministically run an alternance of exploration phase, auction phase and exploitation phase, of (exponentially) increasing durations.
@@
 The first \rhoLearn uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
-and we include both the centralized multi-play version of \klUCB{} and our proposal \MCTopM-\klUCB{} for comparison.
+and we include both the centralized multi-play version of \klUCB{} and \MCTopM-\klUCB{} for comparison.
 
@@
 More recent works by the same team, \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}, all follow the same approach and include an experimental section that shows the results of an implementation of the proposed algorithms on USRP boards.
-\TODOL{I should explain more, no? Who does what and why? So long...}
+% \TODOL{I should explain more, no? Who does what and why? So long...}
 
@@
 This bound is of the same order as the bound obtained by the same authors for the SW-UCB\# algorithm in \cite{WeiSrivastava18Abruptly} for the single-player case.
-If $\Upsilon_T = \bigO{T^{\gamma}}$, this bound is comparable to the results obtained for most of the research literature on piece-wise stationary bandits (earlier works like D-UCB in \cite{Garivier11UCBDiscount} have the $\log(T)$ outside the square root, while more recent works all improved this aspect and have the $\log(T)$ in the square root, like for instance our proposals \GLRklUCB{} in \cite{Besson2019GLRT} and presented in Chapter~\ref{chapter:6}).
-Even if their analysis is not explicit regarding the constant, in the case where $\Upsilon_T = \bigO{T^{\gamma}}$ for a known $\gamma$, their regret upper-bound actually matches our conjectured lower-bound, up to a logarithmic factor $\log(T)$.
+If $\Upsilon_T = \bigO{T^{\gamma}}$, this bound is comparable to the results obtained for most of the research literature on piece-wise stationary bandits (earlier works like D-UCB in \cite{Garivier11UCBDiscount} have the $\log(T)$ outside the square root, while more recent works all improved this aspect and have the $\log(T)$ in the square root, like for instance our algorithm \GLRklUCB{} in \cite{Besson2019GLRT} and presented in Chapter~\ref{chapter:6}).
+Even if their analysis is not explicit regarding the constant, in the case where $\Upsilon_T = \bigO{T^{\gamma}}$ for a known $\gamma$, their regret upper-bound actually matches the conjectured lower-bound, up to a logarithmic factor $\log(T)$.
 \\
@@
 Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
-However, we note that the dependency in $M$ is surprisingly better for their result than for our regret upper-bound for \MCTopM-\klUCB, as its regret scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}.\\
+However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB, as its regret scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}.\\
 \indent
@@
 
-\TODOL{Je peux faire des expériences pour cette approche, est-ce que j'en fais ? Ici, ou à la fin du chapitre 6 ? Ou en conclusion ?}
+% \TODOL{Je peux faire des expériences pour cette approche, est-ce que j'en fais ? Ici, ou à la fin du chapitre 6 ? Ou en conclusion ?}
 
@@
 We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proved to be very active in the last two years.
-For some models, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt our proposal \MCTopM{} to other settings.
+For some models, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
 For example, it assumes to know the number of players $M$ before-hand, but we illustrated that previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
-Further works would be required to adapt our theoretical analysis to these various extensions.
+Further works would be required to adapt the theoretical analysis to these various extensions.
 
@@
 
-We include here some missing proofs and additional simulation results for this Chapter.
+% We include here some missing proofs and additional simulation results for this Chapter.
 

### 2019-05-09 2-Chapters/6-Chapter/chapter6.tex
So long
Super chaud de défendre notre boulot du chapitre 6 avec des graphes comme ceux des problèmes 3 et 5... on est nul en fait, meme face à Thomspon sampling normal

@@
 
+\newpage
 \paragraph{Abstract.}
@@
 Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
-\MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
+% \MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
 %
@@
 
-We start by giving some additional useful results, along with their proofs, and then we give complementary numerical experiments to justify some choices made in the presentation of our proposal \GLRklUCB.
+We start by including figures as complementary illustrations to the numerical results presented in Section~\ref{sec:6:NumericalExperiments} above.
+We then give some additional useful results, along with their proofs, and then we give complementary numerical experiments to justify some choices made in the presentation of our proposal \GLRklUCB.
 

### 2019-05-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Super chaud de défendre notre boulot du chapitre 6 avec des graphes comme ceux des problèmes 3 et 5... on est nul en fait, meme face à Thomspon sampling normal

@@
 % -----------------------------------------------------------------
-\section{Motivation}
+\section{Motivation for non stationary MAB models}
 \label{sec:6:Introduction}
@@
 
-\section{Experimental results}
+\section{Experimental results for piece-wise stationary bandit problems}
 \label{sec:6:NumericalExperiments}
@@
 It is a good benchmark to see if the actively adaptive policies do not detect \emph{too many changes}, as the Oracle-Restart policy suffers higher regret in comparison to \klUCB.
-Means are also bounded in $[0.01, 0.07]$, with small gaps of amplitude ranging from $0.02$ to $0.001$,
+Means are also bounded in $[0.01, 0.07]$, with small gaps of amplitude in $[0.001, 0.02]$,
 as shown in Figure~\ref{fig:6:Problem_6}.
@@
     \centering
-    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
-    \caption{\textbf{Problem $5$}: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
+    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
+    \caption{\textbf{Pb $5$}: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
     % \caption{\textbf{Problem 5}: $K=5$, $T=100000$, $\Upsilon=81$ changes occurring on some arms at every time ($C=179$).}
@@
 % \TODOE{La notion $\Delta^{\text{change}}$ est déjà définie dans la section 4, et ici elle ne veut pas dire la même chose...}
+
+\paragraph{First experiment: \UCB{} vs \klUCB{}.}
+%
+% The first experiment is considering $K=9$ arms, with means $\bm{\mu}=[0.1,0.2,\dots,0.9]$, and $M=3$ then $6$ then $9$ players.
+% Note that here and as in all this section we consider different algorithms that know the number of players $M$ (see Section~\ref{sub:5:unknownNumberOfPlayers} below for a discussion on the case when $M$ could be unknown).
+% %
+% Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
+
+% In our numerical experiments, we are indeed comparing the two variants of our approach to *modified* versions of M-UCB and CUSUM-UCB, that are based on klUCB instead of UCB.
+As we explained, klUCB is known to perform uniformly better than UCB in stochastic bandits with bounded rewards, and so we are actually helping our competitors by allowing them to use klUCB.
+The reason for this is that we don’t want the paper to be another illustration of "klUCB is better than UCB".
+Our comparison therefore gives a better idea of which change-point detector should be used in combination with a good bandit algorithm.
+
+Similarly to what is presented in Chapter~\ref{chapter:5} in Table~\ref{table:5:comparisonUCB_klUCB}, we start by some experiments that justify the focus on the \klUCB{} index policy.
+The purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices for piece-wise stationary problems.
+%
+Consider two experiments on \textbf{problems 1} and \textbf{2}, with a horizon of $T=20000$ and $1000$ independent repetitions,
+for which we compare \CUSUMUCB{} against \CUSUMklUCB, \MUCB{} against \MklUCB, and the oracle policy using \UCB{} against \klUCB.
+We report in the Table~\ref{table:6:comparisonUCB_klUCB} below the results (in terms of mean regret),
+and we observe that using \klUCB{} rather than \UCB{} indices always yield better practical performance.
+%
+Thus it is fair to compare our proposal against the modified versions of the two algorithms CUSUM and M- that use the \klUCB{} index policy instead of \UCB, as they perform uniformly better with \klUCB{} than with \UCB{} (the same tendency was observed on all other problems).
+We also note that, without surprise, the oracle policy also performs (much) better with \klUCB{} than with \UCB.
+Consequently, from now on we only report results for \klUCB.
+
+% for N in 4 100; do for PROBLEMS in 1 2; do DEBUG=True SAVEALL=False NOPLOTS=True M=$M PROBLEMS=$PROBLEMS N_JOBS=-1 N=$N T=10000 make nonstationary; echo "Done for N=$N and M=$M"; read; done; echo "Done for N=$N"; read; done
+%
+\begin{table}[ht]
+  % \begin{footnotesize}
+      \centering
+      \begin{tabular}{c|cc}
+        \textbf{Algorithm} $\;$ \textbackslash $\;$ Problems & Pb 1 & Pb 2 \\
+        \hline
+        Oracle-Restart \UCB{} & $216$ & $67$ \\
+        Oracle-Restart \klUCB{} & $\mathbf{54}$ & $\mathbf{219}$ \\
+        \hline
+        \MUCB{} & $878$ & $2040$ \\
+        \MklUCB{} & $817$ & $1485$ \\
+        \hline
+        \CUSUMUCB{} & $439$ & $381$ \\
+        \CUSUMklUCB{} & $304$ & $316$ \\
+        \hline
+        \GLRklUCB{}(Local) & $\mathbf{132}$ & $\mathbf{186}$
+      \end{tabular}
+      \caption{Using \klUCB{} is much more efficient than using \UCB{}, for non-stationary bandit.}
+      \label{table:6:comparisonUCB_klUCB}
+  % \end{footnotesize}
+  \end{table}
+
+%   - For Problem 1: CUSUM-klUCB got a mean regret of 304, CUSUM-UCB got 439, M-klUCB got 817 and M-UCB 878, where OracleRestart-klUCB obtained 54 and OracleRestart-UCB got 216, and our approach GLR-klUCB got 132.
+%   - For Problem 2: CUSUM-klUCB got a mean regret of 316, CUSUM-UCB got 381, M-klUCB got 1485 and M-UCB 2040, where OracleRestart-klUCB obtained 67 and OracleRestart-UCB got 219, and our approach GLR-klUCB got 186.
 
@@
 
-- OracleRestart with UCB
-- M- and CUSUM- with UCB (as defined in the papers)
+% - OracleRestart with UCB
+% - M- and CUSUM- with UCB (as defined in the papers)
 - Exp3.S
@@
 
-\TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed} by using \MCTopM{} and \GLRklUCB{} with each other?
+As mentioned at the end of Chapter~\ref{chapter:5}, another interesting direction of future work is to study non-stationary distributed multi-player bandits.
+A natural extension of the non-stationary model presented in this Chapter is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
+The authors build on their recent work \cite{WeiSrivastava18Abruptly} and show that a regret bounded by $\cO(\sqrt{T^{\frac{1+\nu}{2}} \log(T)})$ can still be achieved by $M$ players, in the number of breaking points is $\Upsilon_T = \cO(T^{\nu})$.
+Their algorithm assume that player $j$ knows its ID $j\in\{1,\dots,M\}$ as well as $\nu$, and removing these hypotheses is an interesting direction of future work.
+% \TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed} by using \MCTopM{} and \GLRklUCB{} with each other?
 }
-
+Furthermore, a promising direction is to directly try to join our contributions from Chapter~\ref{chapter:5} and \ref{chapter:6}, and propose an efficient algorithm using three parts:
+\klUCB{} indexes for arm selection,
+\MCTopM{} for orthogonalization (\ie, dealing with collisions),
+and \GLRklUCB{} for non-stationarity (\ie, dealing with abrupt changes).
+%
 % Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm \cite{Besson2018ALT}, to efficiently tackle the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed}.

### 2019-05-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Chapter 6: talk about [Luo et al, COLT, 2018]

@@
 
+Another interesting recent work is \cite{AuerGajaneOrtner18}, where the authors propose an efficient algorithm, AdSwitch, for the two-armed case ($K=2$).
+It proceeds in episodes, starting a new episode when the algorithm detects a change in one of the arms.
+Each episode is an estimation phase, playing both arms alternatingly until their means can be distinguished, then an exploitation phase.
+With some low probability, an exploration phase is started that checks whether a change has occurred.
+Similarly to what previous approaches did (\eg, \CUSUM{} uses a threshold $\varepsilon$), this phase checks for changes of a certain minimum magnitude, but the innovation in this work is that algorithm does not need any prior knowledge of the minimum magnitude.
+Instead, it uses geometrically decreasing magnitudes $d_i = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only consider a few different values of $d_i$.
+% Our  algorithmAdSwitch(shown  as  Algorithm  1)  proceeds  in  episodes,  where  a  newepisode starts when the algorithm detects a change in one of the arms.
+Each episode startswith an estimation phase (Step 2), in which both arms are chosen alternatingly until theirmeans can be distinguished (Step 3).
+After the estimation phase,  the algorithm exploitsthe empirical best arm (Step 5).
+However,  in order to detect changes,  at each step withsome probability (Step 4) an exploration phase is started that checks whether a change hasoccurred (comparing the resulting empirical means to that obtained from the estimationphase).
+During such exploration phases the algorithm checks for changes of certain magni-tudes (denoted bydi).
+Note that our algorithm does not require knowledge of the numberof changesL.
+The following theorem provides an upper bound on the expected regret ofour algorithm
+They prove it achieves a regret bound of $\cO(\log(T) \sqrt{\Upsilon_T T})$, without a prior knowledge of $\Upsilon_T$.
+However, this work as two drawbacks:
+first, the algorithm currently only applies to two arms, and the generalization does not seem straightforward.
+Then, the theoretical result is valid for ``sufficiently large constants $C_1$ and $C_2$'', with a large constant $C$ hiding in the $\cO(\bullet)$ notation.
+We did some experiments with AdSwitch, even if it is not included in the benchmark presented below in Section~\ref{sec:6:NumericalExperiments}.
+It uses two parameters $C_1,C_2$ that are set at large values for the theoretical analysis,
+and while we explored numerically different choices, it seemed that setting $C_1=C_2=1$ gave the best performance (which is outside of the comfort zone of the theoretical analysis).
+Even for this tuning, and for problems with just $K=2$ arms, empirically   this policy is performing poorly in comparison to other policies based on active CD detection and \klUCB{} indexes.
+%  We did not find a practical tuning outperforming GLR-klUCB yet.
+% It would be interesting to compare
+
 \paragraph{Expert aggregation.}
@@
 
-Three recent works proposed algorithms that achieve a small regret under the two models.
-For simple bandits, \cite{WeiSrivastava18Abruptly} proposed the SW-UCB\# and LM-DSEE algorithms FIXME.
-For contextual bandits, \cite{Luo18} and \cite{ChenLeeLuoWei2019} FIXME.
-\TODOL{Parler de ces travaux là, qui sont très forts mais pas non plus encore parfaits.}
-
 This setting is quite different from the setting we are interested by in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
@@
 %
-This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
-Empirically, we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
-
+This interpretation is interesting too,
+% but due to time and space constraints, we preferred to not explore it further.
+but empirically we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
+%
 % \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}
+
+\paragraph{Adversarial or non-stationary contextual bandits.}
+%
+For contextual bandits, two very recent works are \cite{Luo18} and \cite{ChenLeeLuoWei2019}.
+While both works target a more general setting, their algorithms are also applicable to non-contextual bandits, \ie, classical bandits like the model in this chapter.
+From Table~1 in \cite{Luo18}, one can see that their two algorithms Exp4.S and Ada-ILTCB recover the same regret guarantees as we do, in the non-contextual case:
+that is, a $\cO(\Upsilon_T \sqrt{T}$ regret without knowing the number of changes and $\cO(\sqrt{\Upsilon_T T})$ with this knowledge.
+We discuss both these algorithms:
+%
+first, Exp4.S actually reduces to Exp3.S in the non-contextual case, and Exp3.S is discussed above.
+% As already acknowledged in the answer to  Reviewer 1,  we were indeed unaware of the very good theoretical properties of this algorithm in the piece-wise non-stationary setting.
+We will revise our state-of-the-art section and include the Exp3.S algorithm as a baseline in our experiments.
+However, preliminary experiments reveal that this algorithm is much less efficient that GLR-klUCB.
+%
+Then, the Ada-ILTCB algorithm is however a less appealing candidate for our setting, for the following reason.
+This algorithm requires as input a parameter $L$ which needs to be an upper bound on the largest stationary sequence in the problem, in order for its regret to scale as $\Upsilon_T \sqrt{L}$ (neglecting $K$ and $\log$ factors).
+It is claimed that the tuning $L=T/\Upsilon_T$ yields the optimal bound $\sqrt{\Upsilon_T T}$.
+However, this tuning is only possible in a ``balanced'' instance without stationary sequence longer than $T/\Upsilon_T$.
+In a piece-wise stationary instance with a stationary sequence of length $T/2$, one cannot get better than $\cO(\Upsilon_T \sqrt{T})$ regret, by setting $L=T/2$ (see for instance the problem 4 in our benchmark below in Section~\ref{sec:6:NumericalExperiments}), and as such the regret bound of Ada-ILTCB appear much worse than the announced $\cO(\sqrt{\Upsilon_T T})$ bound.
+%
+Besides, no experiments are reported by \cite{Luo18}, and the Ada-ILTCB algorithm seems much more complicated to implement than other alternatives,
+so we prefer to not include it in the experiments presented later.
+% with a valid tuning of L in each case.
+
+\TODOL{Okay ça suffit pour ces travaux ?}
 
@@
 More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).
-
+%
 % \TODOE{Tu confirmes bien que $\gamma=0.75$ ne marchait sur AUCUN problème ? Car ta phrase actuelle n'exclut pas que ce choix marche pour les probèmes courts}
+%
+Additionally, we include the Exp3.S algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\alpha=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \mathrm{e}) / ((\mathrm{e}-1)T)})$.
 
@@
 Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instances.
-
-\TODOL{I could add the following algorithms, just to illustrate more:
-
-% - OracleRestart with UCB
-% - M- and CUSUM- with UCB (as defined in the papers)
-- Exp3.S
-}
+The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including Exp3, and recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17},
+and it usually performs similarly to \MklUCB, however it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.
+
+\TODOL{I HAVE TO add the following algorithms, just to illustrate more Exp3.S}
 
@@
         Oracle-Restart \klUCB{} & $\mathbf{37 \pm 37}$ & $\mathbf{45 \pm 34}$ & $\mathbf{257 \pm 86}$ \\
+        \hline
+        Exp3.S & $XXX$ & $XXX$ & $XXX$ \\
         \hline
@@
     \end{tabular}
-    \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on problems $1$, $2$ (with $T=5000$) and $3$ ($T=20000$).
+    \caption{Mean regret $\pm$ $1$ std-dev, on problems $1$, $2$ (with $T=5000$) and $3$ ($T=20000$).
     }
@@
         Oracle-Restart \klUCB{} & $\mathbf{68 \pm 40}$ & $\mathbf{86 \pm 50}$ & $126 \pm 54$ \\
+        \hline
+        Exp3.S & $XXX$ & $XXX$ & $XXX$ \\
         \hline
@@
 
+\newpage
 %----------------------------------------------------------------------------

### 2019-05-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Chapter 6:
- [ ] visu localisation changements (à faire à la main dans un notebook)
- [ ] expériences rajouter Exp3.S vite fait (100 répét et pas 1000)

@@
 However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round.
-For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
-It is sometimes referred to as the abruptly changing \cite{WeiSrivastava18Abruptly}, or switching environments \cite{MellorShapiro13}.
+For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced in Section~8 of the seminal paper \cite{Auer02NonStochastic}.
+They propose a simple extension of the EXP3 policy, referred to as Exp3.S, that uses exponential weights like EXP3 tweaked with an adaptive forced exploration probability to passively adapt to changes.
+It was shown that Exp3.S attains a regret of $\cO(\sqrt{K T S \log(K T)})$, when comparing with an arbitrary sequence of comparators that does not switch for more than $S-1$ times.
+Running their algorithm over the non-stationary problem, and picking their comparators sequence as the best arm in each stationary interval, one can already get $\sqrt(\Upsilon_T T \log(T))$ regret, and the only prior knowledge needed to run the Exp3.S algorithm is $T$ and $\Upsilon_T$.
+Moreover, Exp3.S is efficient both in term of storage and computation time, and simple to implement.
+
+It could seem that the problem is considered as solved by this simple Exp3.S algorithm, but it was actively studied since the seminal paper \cite{Auer02NonStochastic}.
+Two main reasons explain the dynamic research on this problem:
+first, it is well known that despite their qualities, exponential-weights algorithms like Exp3 can usually be greatly outperformed by UCB-based algorithms, for stochastic problems, and so it is expected that Exp3.S can be outperformed by other algorithms, thus any practical application might benefit from algorithms more efficient than Exp3.S.
+%
+The second reason is that passively adaptive algorithms function as black-box models: they do not try to detect changes and do not explain why they changed from one arm to another.
+A lot of the recent research has been focussed on actively adaptive algorithms, precisely because they allow to interpret their decisions, by detecting changes on arms with statistical tests.
+The piece-wise stationary MAB model was then studied by \cite{Kocsis06} and \cite{YuMannor09}.
+%
+This model is sometimes referred to as the abruptly changing \cite{WeiSrivastava18Abruptly}, or switching environments \cite{MellorShapiro13}.
 %
@@
 However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
-Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the EXP3.S algorithm for the piece-wise stationary problem.
-\TODOL{Explain here that we should have been more careful regarding this sentence?}
-
-\paragraph{Illustrations of two problems.}\label{par:6:benchmark1}
-
-To illustrate the setup presented above, we give here two examples of piece-wise stationary problems.
+Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the EXP3.S algorithm for the piece-wise stationary problem, as explained in Section~\ref{sec:6:Introduction} above.
+% \TODOL{Explain here that we should have been more careful regarding this sentence?}
+
+\paragraph{Two example of problems.}\label{par:6:benchmark1}
+
+To illustrate the model, we give here two examples of piece-wise stationary problems.
 % problems of our benchmark,
@@
     \centering
-    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
-    \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
+    \subcaption{\textbf{Pb $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).}
     \label{fig:6:Problem_1}
@@
     \centering
-    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
     \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).}
@@
 For a fair comparison, we therefore chose to compare the different change-point detector algorithms combined with \klUCB.
+We also include in Table~\ref{table:6:comparisonUCB_klUCB} a comparison of the performance of different algorithms when using \UCB{} or \klUCB{} indexes, to illustrate that using a more efficient index policy always improves performance, as predicted.
 

### 2019-05-09 2-Chapters/3-Chapter/chapter3.tex
Added screenshots from the documentation.

@@
 This library is by far the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
-It aims at being extensive, simple to use and maintain, and has a clean and well documented codebase.
-It allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments.
+It aims at being exhaustive, simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
+It allows fast prototyping of simulations, with an easy configuration system and command-line options to customize experiments.
 Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
@@
 More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
-It uses the most recent technologies, by being hosted on GitHub, and uses two online services of continuous integration to run automated tests and build the documentation.
+It is hosted on GitHub, and uses the most recent technologies, by using two online services of continuous integration to run automated tests and build the documentation.
 
@@
 
-SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms and models of Multi-Armed Bandits.
-It aims at being extensive, simple to use and maintain, with a clean and perfectly documented codebase.
+% SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms and models of Multi-Armed Bandits.
+% It aims at being extensive, simple to use and maintain, with a clean and perfectly documented codebase.
 But most of all it allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments while starting them (see below for an example).
 
-SMPyBandits does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
+SMPyBandits stands for \emph{S}ingle and \emph{M}ulti \emph{P}layer \emph{Bandits} in P\emph{y}thon.
+The library does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a pure Python implementation \cite{python}, and uses only standard open-source Python packages (\ie, no hard to install dependencies).
 Some critical parts are also available as a \texttt{C} Python extension, and the just-in-time Numba compiler \cite{numba} is used whenever it is possible, so we can note that we optimized the time efficiency of what could be (easily) optimized.
@@
 We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
-then we discuss in Section~\ref{sec:3:timeAndMemoryCosts} about the time efficient and memory costs of the some MAB algorithms.
+then we discuss in Section~\ref{sec:3:timeAndMemoryCosts} about the time and memory costs of MAB algorithms.
 The take away messages are two-fold:
@@
 
-We start by explaining how SMPyBandits simulates MAB problems, by detailing the different parts.
+We start by explaining how SMPyBandits simulates MAB problems, by detailing its components:
+MAB problems (single- or multi-players),
+and reward distributions.
+%
 We then detail how it implements MAB algorithms (\eg, \UCB),
@@
 
-The main goal of the library is to be easily able to simulate a small number of algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-player) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
+The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-player) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
 After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithms, as well as regret and other visualizations.
@@
 used to sample \iid{} rewards $r_k(t) \sim \nu_k, \forall t$.
-An agent choose arm $A(t)\in\{1,\dots,K\}$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
+An agent chooses arm $A(t)\in\{1,\dots,K\}$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
 Her goal is to maximize $\sum_{t=1}^T r_{A(t)}(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.
@@
 Any simulation library targeting single-player bandit problems must implement at least three components:
-reward distributions, MAB algorithm, and a simulation loop that essentially looks like this.
+reward distributions, MAB algorithm, and a simulation loop that essentially looks like this:
 \begin{itemize}
@@
     \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
-Ask algorithm $\cA$ his chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm,
-    \item At the end, compute the accumulated reward, the regret, plot visualizations etc.
+Ask algorithm $\cA$ her chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm,
+    \item At the end, compute the cumulated reward, the regret, plot visualizations etc.
 \end{itemize}
 %
-Note that the second step should of course be repeated a large number of times if we want to study \emph{mean} regret and not only the regret in one single trajectory.
+Note that the second step is repeated a large number of times (\eg, $N=1000$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
 If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms.
@@
 
-For Cognitive Radio and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
-Whenever two or more players select the same arm at the same time, they all suffer from a collision.
-%
-Different collision models has been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
+For Cognitive Radio dealing with OSA problems and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
+Whenever two or more players select the same arm at the same time (\eg, for OSA, the same frequency channel), they all suffer from a radio collision.
+%
+Different collision models have been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
 Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
@@
 % a collision model, and
-a simulation loop that essentially looks like this.
+a simulation loop that essentially looks like this:
 \begin{itemize}
@@
     \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
-Ask each player $\cA^{(j)}$ his chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
+Ask each player $\cA^{(j)}$ her chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
 The
@@
 %
-We focus on one dimensional distributions,
-and supports \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Arms.Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Arms.Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Arms.Poisson}}), and a generic \emph{discrete} (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{Arms.DiscreteArm}}) distributions,
-as well as \emph{exponential} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Exponential.html}{Arms.Exponential}}), \emph{gamma} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gamma.html}{Arms.Gamma}}), \emph{Gaussian} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gaussian.html}{Arms.Gaussian}}) and \emph{uniform} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Uniform.html}{Arms.Uniform}}) continuous distributions,
-which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$).
+We focus on one dimensional distributions, implemented in the \texttt{\href{https://smpybandits.github.io/docs/Arms.html}{Arms}} module.
+The library supports discrete distributions: \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Poisson}}), and a generic \emph{discrete} distribution (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{DiscreteArm}}),
+as well as continuous distributions,
+which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$):
+\emph{exponential} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Exponential.html}{Exponential}}), \emph{gamma} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gamma.html}{Gamma}}), \emph{Gaussian} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gaussian.html}{Gaussian}}) and \emph{uniform} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Uniform.html}{Uniform}}).
 %
@@
 %
-For instance the following code creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate the problem.
+For instance the following code creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate sthe problem.
 We give another example for truncated Gaussian distributions, and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
@@
 
-We do not give more details, but the interested reader can refer to the following Jupyter notebook \cite{jupyter} document hosted at:\\
+For more details, an interested reader can refer to the following Jupyter notebook \cite{jupyter}:
 \begin{small}
@@
 SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
-containing over 65 (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}) algorithms.
+containing over 65 algorithms (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}).
 It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams showed on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
@@
 
-A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files:
+\subsection{Documentation}
+
+A complete documentation, for each algorithm and the entire codebase, is available online at
+% even including the constants in the different configuration files:
 \texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
+It uses the Sphinx software \cite{sphinx}, and the content is directly written in the Python files as docstring, to allow users to have access to the documentation from within their IDE or the Python console.
+The most interesting component of the library is the many algorithms being implemented: for each of them, the documentation gives at least a reference to a research paper (\eg, \cite{Kaufmann12BUCB} for \texttt{BayesUCB}), as well as a bird-eye view of its behavior.
+For most algorithms, in particular for index policies, the internal variables of the implementation are carefully linked with the notations of each paper, and formulas explaining the way the algorithm selects its arm are usually given.
+Whenever it is appropriate, the documentation also includes warning or information about the empirical performance of costly algorithms.
+The documentation also contains extensive examples of any intermediate numerical functions, that are also used as tests (using \texttt{doctest.testmod} from the standard library).
+For example, Figures~\ref{fig:3:twoScreenshotsOfTheDocumentation} and \ref{fig:3:twoMoreScreenshotsOfTheDocumentation} below show four screenshots from the documentation.
+We show the main page, the list of algorithms in the \texttt{Policies} module, an example of the documentation of one algorithm (for the \klUCB-Switch algorithm from \cite{Garivier18}), and a page detailing the API and organization of the library as well as instructions to follow if someone wants to implement a new model or a new algorithm.
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+	\includegraphics[width=0.49\linewidth]{overview_documentation_1.png}
+    \includegraphics[width=0.49\linewidth]{overview_documentation_2.png}
+	\caption{Screenshots from two pages of the documentation.}
+	\label{fig:3:twoScreenshotsOfTheDocumentation}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+	\includegraphics[width=0.49\linewidth]{overview_documentation_3.png}
+	\includegraphics[width=0.49\linewidth]{overview_documentation_4.png}
+	\caption{Screenshots from two other pages of the documentation.}
+	\label{fig:3:twoMoreScreenshotsOfTheDocumentation}
+\end{figure}
 
@@
 
-A small script \texttt{configuration.py} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/configuration.html}{SMPyBandits.GitHub.io/docs/configuration.html}}) is used to import the arm classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{SMPyBandits.GitHub.io/docs/Arms.html}}), the policy classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{SMPyBandits.GitHub.io/docs/Policies.html}}) and define the problems and the experiments.
-Choosing the algorithms is easy by customizing the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
-For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{Policies.klUCB}}) against the non-anytime variant $\klUCB^{++}$ algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{Policies.klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{Policies.UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Policies.Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Policies.Posterior.Beta}})).
+A small script \texttt{configuration.py}
+% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/configuration.html}{SMPyBandits.GitHub.io/docs/configuration.html}})
+is used to import the arm classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{Arms}} module), the policy classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}} module) and define the problems and the experiments.
+The \texttt{configuration["policies"]} list in the \texttt{configuration.py} file is used to choose the algorithms to include.
+For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{klUCB}} in \texttt{Policies} module) against the non-anytime variant $\klUCB^{++}$ algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Posterior.Beta}})).
 
@@
 \begin{figure}[h!]  % [htbp]
-	% \centering
+	\centering
 	\includegraphics[width=1.05\linewidth]{3.pdf}
-	\caption[Example of a single-player simulation showing the average regret and histogram of regrets of $4$ algorithms]{
-		Example of a single-player simulation showing the average regret and histogram of regrets of four algorithms.
+	\caption[Example of a single-player simulation showing the average regret of $4$ algorithms]{
+		Example of a single-player simulation showing the average regret of four algorithms.
 They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proved to match the lower-bound up-to a constant), and each but UCB is known to be optimal (\ie, with the constant matching the lower-bound).
@@
 \begin{figure}[h!]  % [htbp]
-	% \centering
+	\centering
 	\includegraphics[width=1.05\linewidth]{3_hist.pdf}
-	\caption{Histogram of regret for the same experiment.
-        For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).
-    }
+	\caption{Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}.
+For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).}
 	\label{fig:3:firstPlot_hist}
@@
 This allows to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
-The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proved very useful to quickly find and fix new bugs.
+The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
 For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
@@
 The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-player model).
-Builds typically run for $15$ minutes, and Travis CI have been proved to be very useful in our development process.
+Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.
 
@@
 
-SMPyBandits was used for the following research articles since $2017$, and it is used for the previous and the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
+SMPyBandits has been used for the following research articles since $2017$, and it is used for the previous and the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
 
@@
 -- In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
-We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
+We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they obtain logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
 
 -- In \cite{Besson2018DoublingTricks} (quickly presented in Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
-and lots of bibliographic referencesIn sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments.
-A well-known trick for transforming any non-anytime algorithm to an anytime variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
+In sequential learning, an algorithm is said to be ``anytime'' if it does not need to know the horizon $T$ of the experiments.
+A well-known trick for transforming any ``non-anytime'' algorithm to an ``anytime'' variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
 We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results.
-A geometric sequence suffices to conserve minimax regret bounds (in $R_T = \mathcal{O}(\sqrt{T})$), with a constant multiplicative loss $\ell \leq 4$, but cannot be used to conserve a logarithmic regret bound (in $R_T = \mathcal{O}(\log(T))$).
-And an exponential sequence can be used to conserve logarithmic bounds, with a constant multiplicative loss also $\ell \leq 4$ in the usual setting.
-It is still an open question to know if a well-tuned exponential sequence can conserve minimax bounds, or "weak" minimax bounds (in $R_T = \mathcal{O}(\sqrt{T \log(T)})$).
+A geometric sequence suffices to keep minimax regret bounds (in $R_T = \mathcal{O}(\sqrt{T})$), with a constant multiplicative loss $\ell \leq 4$, but cannot be used to keep a logarithmic regret bound (in $R_T = \mathcal{O}(\log(T))$).
+And an exponential sequence can be used to keep logarithmic bounds, with a constant multiplicative loss also $\ell \leq 4$ in the usual setting.
+It is still an open question to know if a well-tuned exponential sequence can keep minimax bounds, or "weak" minimax bounds (in $R_T = \mathcal{O}(\sqrt{T \log(T)})$).
 
@@
 
-Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on using the simple \UCB{} algorithm when simplicity is favored, \ie, in the more applied Chapter~\ref{chapter:4}, and the \klUCB{} algorithm when efficiency is favored, \ie, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}.
+Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on using
+\begin{itemize}
+    \item
+    the simple \UCB{} algorithm when simplicity is favored, \ie, in the more applied Chapter~\ref{chapter:4},
+    \item
+    and the \klUCB{} algorithm when efficiency is favored, \ie, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}.
+\end{itemize}
 
@@
 We consider the nine following algorithms, and seven more are described in Appendix~\ref{sub:3:additionalExperiments}.
-For each of them, we give a bibliographic reference, that corresponds to the most recent article studying it and not the first one which introduced it.
+For each of them, we give a bibliographic reference, that corresponds to a recent article studying it and not the first one which introduced it.
 We give the chosen tuning of the parameters, for parametric algorithms.
-Algorithms that are not anytime use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, of course they use the correct successive values. \\
+Algorithms that are ``not anytime'' use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, they use the correct successive values. \\
 %
@@
     It is a recent variant of the $\mathrm{Exp3}$ algorithm, using an adaptive tuning of its parameter.
-It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
 
@@
     $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
-    It was proved to achieve order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+    It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
 
@@
     \klUCB{} from \cite{KLUCBJournal}, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern} (\texttt{klUCB}).
-    It was proved to achieve asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
+    It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
     It is computationally more costly that \UCB, as at each time step $t$, to compute the $K$ indexes, a convex optimization problem must be solved approximately, for each arm.
@@
     It is a recent variant of the $\mathrm{MOSS}$ algorithm \cite{BubeckSlivkins12}, using an adaptive tuning of its parameter.
-It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+It is anytime, and obtains good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
     Empirically, it performs usually similarly to \UCB, and is slightly more costly in terms of both time and memory.
@@
     It mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
-    It was proved to achieve order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+    It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
     Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in terms of both time and memory.
@@
     It mimics the \UCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
-    It was also proved to achieve order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+    It also achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
     Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in computation time, but it is much more costly in terms of memory.
@@
     It mimics the \klUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
-    It is anytime, and was proved to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+    It is anytime, and was prove to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
     Empirically, it usually outperforms \klUCB{} in terms of regret, and costs about the same memory, but it is slower.
@@
 %  dividing the running time by \texttt{N\_JOBS}.
-But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is sub-optimal, as the CPU will essentially spend all his time (and memory) managing the different jobs, and not actually doing the simulations.
+But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is sub-optimal, as the CPU will essentially spend all its time (and memory) managing the different jobs, and not actually doing the simulations.
 Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits.

### 2019-05-09 2-Chapters/3-Chapter/chapter3.tex
SO LONG

@@
 After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithms, as well as regret and other visualizations.
-The same problem is simulated for $N$ independent repetitions (\eg, $N=1000$) in order to show mean results with a low variance.
+The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with a low variance.
 
@@
 %
-Note that the second step is repeated a large number of times (\eg, $N=1000$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
+Note that the second step is repeated a large number of times (\eg, $N=100$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
 If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms.
@@
 %
-The default is to use the same distribution for the $K$ arms, but it also possible to mix them (see for instance Figure~\ref{fig:25:HarderMixed}).
+The default is to use the same distribution for the $K$ arms, as this is the setting studied in the literature,
+but it also possible to mix them.
+%  (\eg, one arm is Bernoulli-distributed and another).
+%  (see for instance Figure~\ref{fig:25:HarderMixed}).
 %
@@
 is used to import the arm classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{Arms}} module), the policy classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}} module) and define the problems and the experiments.
-The \texttt{configuration["policies"]} list in the \texttt{configuration.py} file is used to choose the algorithms to include.
+Choosing the algorithms to include is done by changing
+the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
 For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{klUCB}} in \texttt{Policies} module) against the non-anytime variant $\klUCB^{++}$ algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Posterior.Beta}})).
@@
   {"archtype": klUCB, "params": {"klucb": klucbBern}},
-  {"archtype": klUCBPlusPlus, "params": {"horizon": T, "klucb": klucbBern}},
+  {"archtype": klUCBPlusPlus,
+   "params": {"horizon": T, "klucb": klucbBern}},
   {"archtype": UCBalpha, "params": {"alpha": 1}},
@@
 		Example of a single-player simulation showing the average regret of four algorithms.
-They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proved to match the lower-bound up-to a constant), and each but UCB is known to be optimal (\ie, with the constant matching the lower-bound).
+They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proved to match the lower-bound up-to a constant), and each but UCB is known to be asymptotically optimal (\ie, with the constant matching the lower-bound).
 	}
@@
 
-Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on using
+Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on two algorithms depending if simplicity or efficiency is favored:
 \begin{itemize}
     \item
-    the simple \UCB{} algorithm when simplicity is favored, \ie, in the more applied Chapter~\ref{chapter:4},
+    \UCB{} is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
     \item
-    and the \klUCB{} algorithm when efficiency is favored, \ie, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}.
+    \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when efficiency is favored.
 \end{itemize}
@@
 
-% FIXME order from oldest to most recent, and simplest to more complex
-
 \begin{itemize}
@@
     It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
-    It is computationally more costly that \UCB, as at each time step $t$, to compute the $K$ indexes, a convex optimization problem must be solved approximately, for each arm.
-    In practice, a few steps of a Newton method are enough to obtain a precision of more than 10 digits, and empirically we found that even in the worst cases \klUCB{} is an order of magnitude slower than \UCB.
+    It is computationally more costly that \UCB, as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
+    In practice, a few steps of a simple Newton method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than an order of magnitude slower than \UCB.
+    % of more than 10 digits
     \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}
@@
     AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
-    It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posteriors on each arm.
+    It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posterior on each arm.
 It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson decision rule (\ie, choose the arm maximizing one draw of random samples from the posteriors), otherwise it uses the Bayes-UCB decision rule (\ie, choose the arm maximizing the $1-1/t$ quantile).
     It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
-    In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, is comparable in terms of computation times, but costs more memory.
+    In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
 
@@
     The same level of optimization is used in all the codebase of SMPyBandits, and most of the time, the code is pure and naive Python and was not optimized.
-    Mathematical functions (\eg, $\sqrt{\bullet}, \log$) and random numbers used are using \texttt{numpy} and \texttt{numpy.random} modules \cite{numpy}, which are based on C and Cython code and can be considered highly efficient.
-    Computations of Kullback-Leibler and indexes for \klUCB{} algorithms are based on a compiled CPython extension written in C and can also be considered highly efficient.
+    Mathematical functions (\eg, $\sqrt{\bullet}, \log$) and random numbers used are using \texttt{numpy} and \texttt{numpy.random} modules \cite{numpy}, which are based on C and Cython code \cite{cython} and can be considered highly efficient.
+    Computations of Kullback-Leibler and indexes for \klUCB{} algorithms are based on a compiled CPython extension written in C \cite{python} and can also be considered highly efficient.
     As we can safely affirm that the code of the different algorithms has the same level of optimization, the comparison is fair, and we do not favor any algorithm in their implementation in SMPyBandits.
@@
 
-In this Appendix, we start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
-We conclude by a discussion on parallel computations and the use of multi-core CPU to speed-up simulations in SMPyBandits.
+We start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
+We conclude by a discussion on parallel multi-core CPU computations to speed-up SMPyBandits.
 
@@
 \end{verbatim}
-
+% \vspace*{-10pt} % WARNING
+%
 % https://tex.stackexchange.com/a/12430/
@@
     Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
-Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction},Hazan2016introduction and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
+Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
     \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as KL-UCB-Switch), and costs an-order-of-magnitude more memory.

### 2019-05-09 2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py

@@ -2,14 +2,11 @@
 See https://SMPyBandits.GitHub.io/API.html for more details!"""
 import numpy as np
 np.random.seed(0)  # for reproducibility
-
 from SMPyBandits.Arms import Bernoulli
 arms = [Bernoulli(0.1), Bernoulli(0.9)]
-
 from SMPyBandits.Environment import MAB
 my_MAB_problem = MAB(arms)
 nbArms = my_MAB_problem.nbArms  # 2 arms !
-
 from SMPyBandits.Policies import UCB
 my_UCB_algo = UCB(nbArms)
 my_UCB_algo.startGame()  # reset internal memory

### 2019-05-09 2-Chapters/3-Chapter/chapter3.tex
Break, checking Ch4, Ch5

@@
 
+\TODOL{Tout ce blabla est un peu trop long, je devrais couper et être plus concis !}
+
 % \TODOL{Write text for this Section, write configuration.py file, run short-term experiment, include them, if happy about it, run longer ang larger experiments}
@@
 
-\paragraph{Normalizing data?}
+\paragraph{Normalizing data.}
 %
@@
 For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
-Simply considering the average of such measurements make no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
+Simply considering the average of such measurements makes no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
 %
@@
 
-\TODOL{Pas sûr que ça vaille la peine d'inclure ce passage, c'est vraiment optionnel. Dans tous les cas, je devrais rajouter tout cette sous section dans la documentation, c'est intéressant !}
+\TODOL{Pas sûr que ça vaille la peine d'inclure ce passage, c'est vraiment optionnel !}
 

### 2019-05-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Added screenshots from the documentation.
SO LONG
OK for Ch5

@@
 
-\TODOL{I need to write a paragraph about using the ``adversarial setting'' for piece-wise stationary.
+% \TODOL{I need to write a paragraph about using the ``adversarial setting'' for piece-wise stationary.
 It is very important to state that EXP3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with NO prior knowledge of $T$ and $\Upsilon_T$.}
@@
 We consider $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
-The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1}.
+The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1} below.
 % (and Figure~\ref{fig:6:Problem_1} in Appendix),
@@
     \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
-    \subcaption{\textbf{Pb $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).}
+    \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).
+        The means are in $[0,1]$, and there are $C+1=5$ stationary intervals of equal lengths.
+        Some changes do not modify the optimal arm (\eg, at $T=1000$ and $T=4000$) and others do.
+    }
     \label{fig:6:Problem_1}
@@
     \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
-    \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).}
+    \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).
+        The means are again in $[0,1]$, and there are also $C+1=5$ stationary intervals of equal lengths.
+        % Some changes do not modify the optimal arm, but the mean of each arm does change at all change points.
+    }
     \label{fig:6:Problem_2}
@@
     It is actually not used as such in the proofs of the regret upper bound given below, but a similar result is used and proved in the proofs.
-    \TODOL{We have to give the proof, or remove this result!}
+    \TODOL{We have to give the proof, or remove this result?
+Je ne vois pas le problème de le laisser là, ça formalise le résultat mais on a pas besoin de celui là, ça permet de bien formaliser ce qu'on entend par contrôler le délai, et dans la preuve plus bas on prouve ce qu'on souhaite.}
 \end{proof}
@@
 %
-One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
-$(31)$-$(32)$.}:
+One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative is currently not correctly supported by theory, as we found mistakes in the analysis of \CUSUMUCB: the main problem resides in the use of Hoeffding's inequality with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
+$(31)$-$(32)$ in the paper \cite{LiuLeeShroff17}.}:
 $f(t-\tau_i(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{i=1}^K n_i(t)$.
@@
 
-\TODOL{I HAVE TO add the following algorithms, just to illustrate more Exp3.S}
+% \TODOL{I HAVE TO add the following algorithms, just to illustrate more Exp3.S}
 
@@
         \hline
-        Exp3.S & $XXX$ & $XXX$ & $XXX$ \\
+        Exp3.S & $352 \pm 51$ & $310 \pm 62$ & $665 \pm 93$ \\
         \hline
@@
         \hline
-        Exp3.S & $XXX$ & $XXX$ & $XXX$ \\
+        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & $1723 \pm 421$ \\
         \hline
@@
 Problem $4$ use $K=3$ arms, and a first long stationary sequence.
-Problem $5$ use $K=5$, $T=20000$ and is much harder with $\Upsilon=82$ breakpoints and $C=179$ changes.}
+Problem $5$ use $K=5$, $T=100000$ and is much harder with $\Upsilon=82$ breakpoints and $C=179$ changes.}
     \label{table:6:totalResults2}

### 2019-05-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Trying to fix last issue on Chapter6

@@
 
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
+    \caption{Location of the detected change-points for four algorithms on Problem 1.}
+    \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1}
+\end{figure}
+
 \TODOL{Conclusion à tirer de cet exemple !}
@@
 \TODOL{Conclusion à tirer de cet exemple !}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
+    \caption{Location of the detected change-points for four algorithms on Problem 2.}
+    \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2}
+\end{figure}
 

### 2019-05-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
i should die ?

@@
 
-The proof of \GLRklUCB{} with Local Changes is only given in Appendix~\ref{proof:6:mainRegretBound}, as it is quite similar to the proof for Global changes given above.
+The proof of \GLRklUCB{} with Local Changes is not included, but it can be found in the Appendix of \cite{Besson2019GLRT}, as it is quite similar to the proof for Global changes given above.
 The main difficulty relies in defining the ``good event'' $\cE_T$, and proving that it happens with high probability (by showing that the complementary event $\cE_T^c$ is highly unlikely).
@@
     &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(k)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \widehat{\tau}^{(k)})\right)\Big]
-\end{align*}
+\end{align*}%
 \begin{align*}
@@
 
-\TODOL{Je ne suis pas sur que cette forme soit la meilleure pour présenter ces exemples, mais je n'aime vraiment pas la forme graphique qu'on avait suggéré ensemble avec Emilie, en fait on voit rien si on met plus d'un algo...}
+% \TODOL{Je ne suis pas sur que cette forme soit la meilleure pour présenter ces exemples, mais je n'aime vraiment pas la forme graphique qu'on avait suggéré ensemble avec Emilie, en fait on voit rien si on met plus d'un algo...}
 
@@
 
-\begin{itemize}
-    \item
-    \GLRklUCB{} with \emph{local restarts} restarted at $2014$ from a change detected on arm $2$, after $1970$ pulls of this arm, and at time $3032$ on arm $0$ (after $840$ pulls), and did not restart then.
+Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1} below displays the locations of the successively detected change-points by four algorithms,
+$1)$ \MklUCB,
+$2)$ \CUSUMklUCB,
+$3)$ \GLRklUCB{} with \emph{local restart},
+$4)$ \GLRklUCB{} with \emph{global restart}.
+In this ``easy'' problem, all algorithms correctly detect the important changes, except \CUSUM{} which detected a change on arm $1$ (in \textcolor{red}{red}) twice after the change-point located at $t=3000$.
+The three other algorithms have a very small delay, for instance only $9$ samples from arm $0$ (in \textcolor{blue}{blue}) after its change at time $t=2000$ are enough for the \GLR{} test to detect a change.
+Very small delays are possible only after having collected a lot of samples of the arm which changed, and for instance the same algorithm obtains a delay of $32$ samples from \textcolor{red}{arm $1$} for the next change at $t=3000$, because this arm was sampled less.
+On this problem, while \MklUCB{} detects the same changes as \GLRklUCB, with larger but comparable delays, it is seen to obtain a larger regret than the two \GLR{} variants, because the tuning of its forced exploration probability makes it explore suboptimal arms more often.
+
+% \begin{itemize}
+%     \item
+%     \GLRklUCB{} with \emph{local restarts} restarted at $2014$ from a change detected on arm $2$, after $1970$ pulls of this arm, and at time $3032$ on arm $0$ (after $840$ pulls), and did not restart then.
 It successfully detects both important changes, and obtains a good performance in terms of regret, with $R_T = 63$ (on only one run).
-    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2014 for arm 2, after 1969 pulls of that arm (giving mean reward = 0.888).
+%     % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2014 for arm 2, after 1969 pulls of that arm (giving mean reward = 0.888).
 Last restart on that arm was at tau = 0
-    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 3032 for arm 0, after 771 pulls of that arm (giving mean reward = 0.336).
+%     % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 3032 for arm 0, after 771 pulls of that arm (giving mean reward = 0.336).
 Last restart on that arm was at tau = 0
 
-    \item
-    \GLRklUCB{} with \emph{global restarts} restarted at $2009$ from a change detected on arm $2$, after $1972$ pulls of this arm, and at time $3031$ on arm $0$ (after $827$ pulls), and did not restart then.
+%     \item
+%     \GLRklUCB{} with \emph{global restarts} restarted at $2009$ from a change detected on arm $2$, after $1972$ pulls of this arm, and at time $3031$ on arm $0$ (after $827$ pulls), and did not restart then.
 It successfully detects both important changes, and obtains a good performance in terms of regret, with $R_T = 71$.
-    Surprisingly, we do not observe a significant difference with the \emph{local restarts} variant.
-    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2009 for arm 2, after 1972 pulls of that arm (giving mean reward = 0.896).
+%     Surprisingly, we do not observe a significant difference with the \emph{local restarts} variant.
+%     % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2009 for arm 2, after 1972 pulls of that arm (giving mean reward = 0.896).
 Last restart on that arm was at tau = 0
-    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 3031 for arm 0, after 827 pulls of that arm (giving mean reward = 0.32).
+%     % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 3031 for arm 0, after 827 pulls of that arm (giving mean reward = 0.32).
 Last restart on that arm was at tau = 2009
 
-    \item
-    \CUSUMklUCB{} restarted at $2011$ from a change detected on arm $2$, after $1861$ pulls of this arm, at time $2512$ on arm $1$ (after $249$ pulls), at time $3027$ on arm $0$ (after $825$ pulls) and again at time $3505$ on arm $0$ (after $466$ pulls), at time $4383$ on arm $1$ (after $123$ pulls), and a last time at $t=4914$ on arm $0$ (after $1304$ pulls), and did not restart then.
-    It restarted too many times on arm $0$ after its last change at $t=3000$, but except this weakness, it successfully detected the two important changes, and obtains a final regret $R_T = 150$.
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2011 for arm 2, after 1861 pulls of that arm (giving mean reward = 0.895). Last restart on that arm was at tau = 0
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2512 for arm 1, after 249 pulls of that arm (giving mean reward = 0.265). Last restart on that arm was at tau = 0
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3027 for arm 0, after 825 pulls of that arm (giving mean reward = 0.33). Last restart on that arm was at tau = 0
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3505 for arm 0, after 466 pulls of that arm (giving mean reward = 0.717). Last restart on that arm was at tau = 3027
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 4383 for arm 1, after 123 pulls of that arm (giving mean reward = 0.268). Last restart on that arm was at tau = 2512
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 4914 for arm 0, after 1304 pulls of that arm (giving mean reward = 0.699). Last restart on that arm was at tau = 3505
-
-    \item
-    \MklUCB{} restarted at $2055$ from a change detected on arm $2$, after $1811$ pulls of this arm, and at time $3080$ on arm $0$ (after $868$ pulls), and did not restart then.
-    It successfully detects both important changes, and obtains a rather bad performance in terms of regret, with $R_T = 280$.
-    % For a player M-klUCB($w=150$, Global) a change was detected at time 2055 for arm 2, after 1811 pulls of that arm (giving mean reward = 0.0552). Last restart on that arm was at tau = 0
-    % For a player M-klUCB($w=150$, Global) a change was detected at time 1025 for arm 0, after 868 pulls of that arm (giving mean reward = 0.0806). Last restart on that arm was at tau = 2055
-\end{itemize}
+%     \item
+%     \CUSUMklUCB{} restarted at $2011$ from a change detected on arm $2$, after $1861$ pulls of this arm, at time $2512$ on arm $1$ (after $249$ pulls), at time $3027$ on arm $0$ (after $825$ pulls) and again at time $3505$ on arm $0$ (after $466$ pulls), at time $4383$ on arm $1$ (after $123$ pulls), and a last time at $t=4914$ on arm $0$ (after $1304$ pulls), and did not restart then.
+%     It restarted too many times on arm $0$ after its last change at $t=3000$, but except this weakness, it successfully detected the two important changes, and obtains a final regret $R_T = 150$.
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2011 for arm 2, after 1861 pulls of that arm (giving mean reward = 0.895). Last restart on that arm was at tau = 0
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2512 for arm 1, after 249 pulls of that arm (giving mean reward = 0.265). Last restart on that arm was at tau = 0
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3027 for arm 0, after 825 pulls of that arm (giving mean reward = 0.33). Last restart on that arm was at tau = 0
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3505 for arm 0, after 466 pulls of that arm (giving mean reward = 0.717). Last restart on that arm was at tau = 3027
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 4383 for arm 1, after 123 pulls of that arm (giving mean reward = 0.268). Last restart on that arm was at tau = 2512
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 4914 for arm 0, after 1304 pulls of that arm (giving mean reward = 0.699). Last restart on that arm was at tau = 3505
+
+%     \item
+%     \MklUCB{} restarted at $2055$ from a change detected on arm $2$, after $1811$ pulls of this arm, and at time $3080$ on arm $0$ (after $868$ pulls), and did not restart then.
+%     It successfully detects both important changes, and obtains a rather bad performance in terms of regret, with $R_T = 280$.
+%     % For a player M-klUCB($w=150$, Global) a change was detected at time 2055 for arm 2, after 1811 pulls of that arm (giving mean reward = 0.0552). Last restart on that arm was at tau = 0
+%     % For a player M-klUCB($w=150$, Global) a change was detected at time 1025 for arm 0, after 868 pulls of that arm (giving mean reward = 0.0806). Last restart on that arm was at tau = 2055
+% \end{itemize}
 
@@
 
-\TODOL{Conclusion à tirer de cet exemple !}
+% \TODOL{Conclusion à tirer de cet exemple !}
 
@@
 
-\begin{itemize}
-    \item
-    \GLRklUCB{} with \emph{local restarts} restarted at $1070$ from a change detected on arm $2$, after $1030$ pulls of this arm, again at time $2367$ on arm $0$ (after $1006$ pulls), and at time $4203$ on arm $0$ (after $1873$ pulls), and did not restart then.
-    It successfully detects the important change, and obtains a good performance in terms of regret, with a final regret $R_T = 115$.
-    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 1070 for arm 2, after 1030 pulls of that arm (giving mean reward = 0.89).
+Like for the first problem, we illustrate the behavior of the same four algorithms in Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2} below.
+In this other problem, the difference between \MklUCB{} and \CUSUMklUCB{} is clear:
+the first algorithm fails to detect any change, leading to a large regret,
+while the second one detects the changes that happen on the currently optimal arm and some other changes.
+Drawing conclusions on their behavior from a single simulation is tricky, as this example of behavior is counter intuitive:
+On the first hand, \CUSUM{} uses local restarts, and it should be less efficient than global restarts for this problem, as changes are all global.
+On the other hand, \MklUCB{} uses global restarts, but here it fails to detect any changes.
+%
+The two \GLRklUCB{} variants correctly detect the important changes, and we observe that their delays to detect a change can vary, mainly due to the randomness (we remind that we illustrate only one repetition of the simulation here).
+For instance, after the change on arm $1$ (in \textcolor{red}{red}), the \emph{local restart} detects it at time $t=2376$ and the \emph{global restart} detects it at time $t=2701$.
+
+% \begin{itemize}
+%     \item
+%     \GLRklUCB{} with \emph{local restarts} restarted at $1070$ from a change detected on arm $2$, after $1030$ pulls of this arm, again at time $2367$ on arm $0$ (after $1006$ pulls), and at time $4203$ on arm $0$ (after $1873$ pulls), and did not restart then.
+%     It successfully detects the important change, and obtains a good performance in terms of regret, with a final regret $R_T = 115$.
+%     % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 1070 for arm 2, after 1030 pulls of that arm (giving mean reward = 0.89).
 Last restart on that arm was at tau = 0
-    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2367 for arm 2, after 1006 pulls of that arm (giving mean reward = 0.654).
+%     % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2367 for arm 2, after 1006 pulls of that arm (giving mean reward = 0.654).
 Last restart on that arm was at tau = 1070
-    % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 4203 for arm 0, after 1873 pulls of that arm (giving mean reward = 0.666).
+%     % For a player GLR-klUCB_forGLR(Local, $\Delta n=1$, $\Delta s=1$) a change was detected at time 4203 for arm 0, after 1873 pulls of that arm (giving mean reward = 0.666).
 Last restart on that arm was at tau = 0
 
-    \item
-    \GLRklUCB{} with \emph{global restarts} restarted at time $1111$ from a change detected on arm $2$, after $1060$ pulls of this arm, and at time $2705$ on arm $0$ (after $1038$ pulls), and did not restart then.
-    It successfully detects the important changes, but with a larger delay than the \emph{local restarts} variant, and also obtains a good performance in terms of regret, with a final regret $R_T = 125$.
-    Here again we do not observe a significant difference with the \emph{local restarts} variant.
-    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 1111 for arm 2, after 1060 pulls of that arm (giving mean reward = 0.888).
+%     \item
+%     \GLRklUCB{} with \emph{global restarts} restarted at time $1111$ from a change detected on arm $2$, after $1060$ pulls of this arm, and at time $2705$ on arm $0$ (after $1038$ pulls), and did not restart then.
+%     It successfully detects the important changes, but with a larger delay than the \emph{local restarts} variant, and also obtains a good performance in terms of regret, with a final regret $R_T = 125$.
+%     Here again we do not observe a significant difference with the \emph{local restarts} variant.
+%     % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 1111 for arm 2, after 1060 pulls of that arm (giving mean reward = 0.888).
 Last restart on that arm was at tau = 0
-    % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2705 for arm 2, after 1038 pulls of that arm (giving mean reward = 0.637).
+%     % For a player GLR-klUCB_forGLR(Global, $\Delta n=1$, $\Delta s=1$) a change was detected at time 2705 for arm 2, after 1038 pulls of that arm (giving mean reward = 0.637).
 Last restart on that arm was at tau = 1111
 
-    \item
-    \CUSUMklUCB{} restarted at $1055$ from a change detected on arm $2$, after $963$ pulls of this arm, at time $2051$ again on arm $2$ (after $922$ pulls), at time $2305$ on arm $0$ (after $185$ pulls), at time $2591$ on arm $0$ (after $272$ pulls), at time $3387$ on arm $0$ (after $770$ pulls), at time $3677$ on arm $1$ (after $169$ pulls), and did not restart then.
-    It missed some changes but detected the two ones, and obtained a final regret $R_T=142$ on this run.
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 1055 for arm 2, after 963 pulls of that arm (giving mean reward = 0.916). Last restart on that arm was at tau = 0
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2051 for arm 2, after 922 pulls of that arm (giving mean reward = 0.682). Last restart on that arm was at tau = 1055
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2305 for arm 0, after 185 pulls of that arm (giving mean reward = 0.53). Last restart on that arm was at tau = 0
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2591 for arm 0, after 272 pulls of that arm (giving mean reward = 0.618). Last restart on that arm was at tau = 2305
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3387 for arm 0, after 770 pulls of that arm (giving mean reward = 0.653). Last restart on that arm was at tau = 2591
-    % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3677 for arm 1, after 169 pulls of that arm (giving mean reward = 0.373). Last restart on that arm was at tau = 0
-
-    \item
-    \MklUCB{} did not detect any change, thus it obtains bad performance in terms of regret, with $R_T = 570$ (on this run).
-\end{itemize}
-
-\TODOL{Conclusion à tirer de cet exemple !}
+%     \item
+%     \CUSUMklUCB{} restarted at $1055$ from a change detected on arm $2$, after $963$ pulls of this arm, at time $2051$ again on arm $2$ (after $922$ pulls), at time $2305$ on arm $0$ (after $185$ pulls), at time $2591$ on arm $0$ (after $272$ pulls), at time $3387$ on arm $0$ (after $770$ pulls), at time $3677$ on arm $1$ (after $169$ pulls), and did not restart then.
+%     It missed some changes but detected the two ones, and obtained a final regret $R_T=142$ on this run.
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 1055 for arm 2, after 963 pulls of that arm (giving mean reward = 0.916). Last restart on that arm was at tau = 0
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2051 for arm 2, after 922 pulls of that arm (giving mean reward = 0.682). Last restart on that arm was at tau = 1055
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2305 for arm 0, after 185 pulls of that arm (giving mean reward = 0.53). Last restart on that arm was at tau = 0
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 2591 for arm 0, after 272 pulls of that arm (giving mean reward = 0.618). Last restart on that arm was at tau = 2305
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3387 for arm 0, after 770 pulls of that arm (giving mean reward = 0.653). Last restart on that arm was at tau = 2591
+%     % For a player CUSUM-klUCB(lazy detect 1) a change was detected at time 3677 for arm 1, after 169 pulls of that arm (giving mean reward = 0.373). Last restart on that arm was at tau = 0
+
+%     \item
+%     \MklUCB{} did not detect any change, thus it obtains bad performance in terms of regret, with $R_T = 570$ (on this run).
+% \end{itemize}
+
+% \TODOL{Conclusion à tirer de cet exemple !}
 
@@
         \hline
-        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & $1723 \pm 421$ \\
+        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & $723 \pm 121$ \\
         \hline
@@
 
-We also include in Appendix~\ref{app:6:moreFigures} some figures showing the simulation results for the five problems.
+We show below on Figure~\ref{fig:6:meanRegretPb1} and additional in Appendix~\ref{app:6:moreFigures} the simulation results for the five problems.
 The results in terms of mean regret $R_T$ are given above in Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2}, but it is also interesting to observe two plots for each experiments.
@@
 This illustrates the conjecture that classical algorithms can suffer linear regret even on simple piece-wise stationary problems.
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.20\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
+    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 1.}
+    \label{fig:6:meanRegretPb1}
+\end{figure}
 

### 2019-05-10 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Using \begin{framed}...\end{framed} for all algorithms.
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

@@
 one arm is chosen with probability $1$ and the others with probability $0$.
-However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $\{1,\dots,K\}$ at each time step, and draw the next selected arm according to this distribution.
+However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $[K]$ at each time step, and draw the next selected arm according to this distribution.
 
@@
 $\pi^t_a$ is the probability of trusting the decision made by algorithm $\Alg_a$ at time $t$.
-$\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it decides whom to trust by sampling $a \in \{1,\dots,N\}$ from $\pi^t$, then follows $\Alg_a$'s decision.
+$\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it decides whom to trust by sampling $a \in [N]$ from $\pi^t$, then follows $\Alg_a$'s decision.
 The main questions are then to know what observations (\ie, arms and rewards) should be given as feedback to which algorithms,
@@
 		\centering
+		\begin{framed}
 		% Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
@@
 			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$}
-			\KwData{Initial uniform distribution, $\pi^{0} = U(\{1, \dots, N\})$}
+			\KwData{Initial uniform distribution, $\pi^{0} = U([N])$}
 			\KwResult{$\Alg_{\mathrm{aggr}}=\Aggr\left[\Alg_1, \dots, \Alg_N\right]$}
@@
 					$\Alg_a$ updates its indexes (e.g., \UCB{} indexes)\;
-					It chooses $A^{t+1}_{a} \in \{1, \dots, K\}$.
+					It chooses $A^{t+1}_{a} \in [K]$.
 				}
@@
 		\end{algorithm}
+		\end{framed}
 	\end{figure}

### 2019-05-10 2-Chapters/3-Chapter/chapter3.tex
...
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

@@
 used to sample \iid{} rewards $r_k(t) \sim \nu_k, \forall t$.
-An agent chooses arm $A(t)\in\{1,\dots,K\}$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
+An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
 Her goal is to maximize $\sum_{t=1}^T r_{A(t)}(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.
@@
 %
-The goal of this section is to verify that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in\{1,\dots,T\}$ bounded by $\bigO{K}$, independent of $t$.
+The goal of this section is to verify that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
 
@@
 
-\pagagraph{About the measurement protocol.}
+\paragraph{About the measurement protocol.}
 %
@@
     \item For time, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
-    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in\{1,\dots,T\}$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
+    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in[T]$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
     This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.

### 2019-05-10 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Talk to Emilie
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

@@
 Following a Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
-it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
+it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in[K]$.
 Then, the dynamic device waits in this channel $C(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example considered above).
@@
 
-Figure \ref{fig:41:from10to100} presents the evolution of the successful transmission rate, as a function of time.
+Figure \ref{fig:41:from10to100} presents the successful transmission rate as a function of time.
 The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).

### 2019-05-10 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
OK for Ch5
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

@@
 
+% FIXME
+\TODOL{Est-ce qu'on enlève toute mention de la lower-bound et de "order-optimal" du coup ?!}
+% FIXME
+
 The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
@@
 % explain the interaction protocol and the GOAL
-In the multi-player MAB setting, there are $M \in \{1,\dots,K\}$ players (or agents),
+In the multi-player MAB setting, there are $M \in [K]$ players (or agents),
 that have to make decisions at some pre-specified time instants.
@@
   A \emph{collision} occurs at time $t$ if at least two players choose the same arm.
-  We introduce the two events, for $j\in\{1,\dots,M\}$ and $k\in\{1,\dots,K\}$,
+  We introduce the two events, for $j\in[M]$ and $k\in[K]$,
   \begin{equation}
@@
   Similarly, \Mworst{} denotes the set of indices of the $K-M$ arms with smallest means (\emph{worst arms}),
-  $\{1, \dots, K\} \setminus \Mbest$.
+  $[K] \setminus \Mbest$.
 
@@
   Let $T^j_k(T) := \sum_{t=1}^T \indic(A^j(t) = k)$,
-  and denote $T_k(T) := \sum_{j=1}^M T^j_k(T)$ the \emph{number of selections} of arm $k\in\{1,\dots,K\}$ by any player $j\in\{1,\dots,M\}$, up to time $T$.
+  and denote $T_k(T) := \sum_{j=1}^M T^j_k(T)$ the \emph{number of selections} of arm $k\in[K]$ by any player $j\in[M]$, up to time $T$.
   % and $\cC_k(T)$ be the number of collisions on arm $k$ up-to time $T$.
@@
 % \begin{definition}[]\label{def:5:nbCollisions}
-  Let $\cC_k(T)$ be the \emph{number of colliding players} on arm $k\in\{1,\dots,K\}$ up to horizon $T$:
+  Let $\cC_k(T)$ be the \emph{number of colliding players} on arm $k\in[K]$ up to horizon $T$:
   % \vspace*{-5pt}  % XXX
@@
     %
-    \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
+    \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
     %
@@
     \text{and} & \;
-    \forall j \in \{1,\dots,M\}, k \in \Mbest, \;\;
+    \forall j \in [M], k \in \Mbest, \;\;
     % a_{j,k}
@@
 Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\rho^{\gamma} = (\rho^{\gamma(1)},\dots,\rho^{\gamma(M)})$ is independent from the choice of permutation $\gamma \in \Sigma_M$.
-In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in \{1,\dots,M\}$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
+In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
 Note that all our proposed algorithms are permutation invariant and \MCTopM{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.
@@
 The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
-Fix a player $j\in\{1,\dots,M\}$, and we consider the observations $\cO_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
+Fix a player $j\in[M]$, and we consider the observations $\cO_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
 \[ \cO_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
@@
   \begin{equation}\label{eq:5:LBDraws}
-    \forall j \in \{1,\dots,M\}, \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
+    \forall j \in [M], \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
   \end{equation}
@@
 \end{equation}
-as for every $k \in \Mworst$ and $j \in \{1,\dots,M\}$, $\kl(\mu_k, \mu_j^*) \geq \kl(\mu_k, \mu_M^*)$
+as for every $k \in \Mworst$ and $j \in [M]$, $\kl(\mu_k, \mu_j^*) \geq \kl(\mu_k, \mu_M^*)$
 (see Figure~\ref{fig:5:CompLowerBounds} in Appendix~\ref{app:5:illustrationLowerBound}).
@@
   \end{eqnarray*}
-  where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
+  where we denote by $\widehat{S}_t = \{A^j(t), j \in [M]\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
   \begin{eqnarray*}
@@
 \begin{equation}
-  A^j(t) = \argmax_{ k \in \{1,\dots,K\}} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
+  A^j(t) = \argmax_{ k \in [K]} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
 \end{equation}
@@
 More formally, letting $\pi(k,\mathbf{g})$ be the index of the $k$-th largest entry in a vector $\mathbf{g}$,
-% (for $k\in\{1,\dots,K\}$),
-in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in\{1,\dots,M\}$
+% (for $k\in[K]$),
+in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in[M]$
 and selects at time $t$,
@@
 \end{equation}
-If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU(\{1,\dots,M\})$.
+If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU([M])$.
 
@@
           % XXX Input, data and output
-          % \KwIn{$K$ and policy $P^j$ for arms set $\{1,\dots,K\}$\;}
+          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
           % \KwData{Data}
@@
           % XXX Algorithm
-              Let $A^j(1) \sim \cU(\{1,\dots,K\})$ and $C^j(1)=\mathrm{False}$ \\
+              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ \\
               \For{$t = 0, \dots, T - 1$}{
@@
           % XXX Algorithm
-              Let $A^j(1) \sim \cU(\{1,\dots,K\})$ and $C^j(1)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
+              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
               \For{$t = 0, \dots, T-1$}{
@@
   For any $\boldsymbol{\mu}\in\cP_M$,
-  let player $j\in\{1,\dots,M\}$ use the \RandTopM-, \MCTopM- or \rhoRand-\klUCB{}
+  let player $j\in[M]$ use the \RandTopM-, \MCTopM- or \rhoRand-\klUCB{}
   decentralized policy with exploration function $f(t) = \log(t) + 3 \log\log(t)$.
@@
   %
-  Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
+  Fix $k\in\Mworst$ and a player $j \in [M]$.
   The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
@@
   %&= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
-  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
+  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in[M]:\; g_{m^*}^j(t) < g_k^j(t) \right).
   %
@@
     \E_{\mu}[T_k^j(T)] &\leq
-    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in \{1,\dots,M\}: \; g_{m_1^*}(t) < \mu_m^* \right) + \notag \\
+    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; g_{m_1^*}(t) < \mu_m^* \right) + \notag \\
     \;\;\;\;
-    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in \{1,\dots,M\}:\; g_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in \{1,\dots,M\}: \;  g_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
-    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(g_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in \{1,\dots,M\}:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
+    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; g_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in [M]: \;  g_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
+    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(g_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
     %
@@
   \end{align}
-  where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.
+  where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in [M]$.
 
@@
   For any $\boldsymbol{\mu}\in\cP_M$,
-  any player $j \in \{1, \dots, M\}$ using
+  any player $j \in [M]$ using
   \RandTopM- or \MCTopM-\klUCB,

### 2019-05-10 2-Chapters/5-Chapter/chapter5.tex
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

@@
 the $M$ devices need to known their number $M$, and have to receive it from the gateway, as it is the only part of this example of network which knowns $M$.
-It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in\{1,\dots,K\}$.
+It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in[K]$.
 
@@
 If the players know the time horizon $T$, then fix a confidence level $\delta$.
-For the first $T_0$ time steps, the $M$ players will use a simple round-Robin game, using their (unique) ID to stay orthogonal: player $j$ starts at arm $j$, then $j+1$, then cycle in $\{1,\dots,K\}$.
+For the first $T_0$ time steps, the $M$ players will use a simple round-Robin game, using their (unique) ID to stay orthogonal: player $j$ starts at arm $j$, then $j+1$, then cycle in $[K]$.
 They encounter no collision in these time steps, and then user $j$ targets the $j$-th best arm among the set of $M$-best arm.
@@
 Their idea is to use the same skeleton as \RhoRand, that is to assign a \emph{rank} to each player, and make players change their rank after any collision.
-But instead of selecting a new rank uniformly at random among $\{1,\dots,M\}$, they proposed to use a ``second-stage'' learning policy\footnote{This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.}, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
+But instead of selecting a new rank uniformly at random among $[M]$, they proposed to use a ``second-stage'' learning policy\footnote{This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.}, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
 This algorithm is implemented in SMPyBandits and named \rhoLearn\footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.}, where it can use Bayes-UCB or any of the $65$ or more bandit algorithms available in the library.
@@
 \indent
-The proposed the RR-SW-UCB\# algorithm, which combines their SW-UCB\# algorithm, previously proposed in \cite{WeiSrivastava18Abruptly}, and a Round-Robin hoping, under the assumption that each user $j\in\{1,\dots,M\}$ knows its ID $j$ (see above the discussion we give on this unrealistic assumption).
+The proposed the RR-SW-UCB\# algorithm, which combines their SW-UCB\# algorithm, previously proposed in \cite{WeiSrivastava18Abruptly}, and a Round-Robin hoping, under the assumption that each user $j\in[M]$ knows its ID $j$ (see above the discussion we give on this unrealistic assumption).
 The SW-UCB\# algorithm is discussed more in details in the literature review of Chapter~\ref{chapter:6} below.
@@
 And finally, even though numerical simulations in their papers \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} are interesting and confirm the regret upper-bounds, they do not compare with any other algorithm, and it is left as an interesting future work to study this direction in more details.
-Sadly, a major drawback of their work is that assume that player $j\in\{1,\dots,M\}$ knows its index $j$, and as we explained above this small hypothesis is actually quite strong, as it allows players to be already orthogonal, and it reduces greatly the difficulty of the decentralized bandit problem.
+Sadly, a major drawback of their work is that assume that player $j\in[M]$ knows its index $j$, and as we explained above this small hypothesis is actually quite strong, as it allows players to be already orthogonal, and it reduces greatly the difficulty of the decentralized bandit problem.
 

### 2019-05-10 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

@@
 
-% Multi-Armed Bandit (MAB) problems form a well-studied class of sequential decision making problems, in which an agent repeatedly chooses an action $I_t \in\{1,\dots,K\}$ or ``arm'' --in reference to the arm of a one-armed bandit-- among a set of $K$ arms.
+% Multi-Armed Bandit (MAB) problems form a well-studied class of sequential decision making problems, in which an agent repeatedly chooses an action $I_t \in[K]$ or ``arm'' --in reference to the arm of a one-armed bandit-- among a set of $K$ arms.
 In the most standard version of the stochastic bandit model, each arm $i$ is associated with an \iid{} sequence of rewards $(r_{i,t})$ that follow some distribution of mean $\mu_i$.
@@
 A \emph{piece-wise stationary bandit model} is characterized by a set of $K$ arms.
-A (random) stream of rewards $(X_{i,t})_{t\in\N^*}$ is associated to each arm $i \in \{1,\dots,K\}$.
+A (random) stream of rewards $(X_{i,t})_{t\in\N^*}$ is associated to each arm $i \in [K]$.
 We assume that the rewards are bounded, and without loss of generality we assume that $X_{i,t} \in [0,1]$.
 We denote by $\mu_{i}(t) :=  \bE[X_{i,t}]$ the mean reward of arm $i$ at round $t$.
-At each round $t$, a decision maker has to select an arm $I_t\in\{1,\dots,K\}$, based on past observation and receives the corresponding reward $r(t) = X_{I_t,t}$.
+At each round $t$, a decision maker has to select an arm $I_t\in[K]$, based on past observation and receives the corresponding reward $r(t) = X_{I_t,t}$.
 At time $t$, we denote by $i_t^*$ an arm with maximal expected reward, \ie, $\mu_{i_t^*}(t) = \max_i \mu_i(t)$, called an optimal arm (possibly not unique).
@@
 
-In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
+In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in[K]: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
 If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint recursively by
@@
 EXP3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
-It constitues a good baseline for our numerical experiments in ,
+It constitues a good baseline for our numerical experiments in Section~\ref{sec:6:NumericalExperiments},
 even if we did find that EXP3.S performs worse that the most of the other approaches based on extensions of stationary bandit algorithms (\eg, SW-UCB).
@@
 % \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
-% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
+% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\alpha \right\rceil$).
 % Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
@@
         % XXX Algorithm
-        \textbf{Initialization: } $\forall i \in \{1,\dots,K\}$, $\tau_i \leftarrow 0$ and $n_i \leftarrow 0$ \\
+        \textbf{Initialization: } $\forall i \in [K]$, $\tau_i \leftarrow 0$ and $n_i \leftarrow 0$ \\
         \For{$t=1,2,\ldots, T$}{
             \uIf{
-                $\alpha>0$ and $t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor \in \{1,\dots,K\}$
+                $\alpha>0$ and $t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor \in [K]$
             }{
@@
             \Else{
-                $A_t \leftarrow \arg\max\limits_{i\in \{1,\dots,K\}} \mbox{UCB}_i(t)$ as defined in \eqref{def:6:UCB}
+                $A_t \leftarrow \arg\max\limits_{i\in [K]} \mbox{UCB}_i(t)$ as defined in \eqref{def:6:UCB}
             }
@@
                 \uIf{Global restart}{
-                    $\forall i\in \{1,\dots,K\}, \tau_i \leftarrow t$ and $n_i \leftarrow 0 $.
+                    $\forall i\in [K], \tau_i \leftarrow t$ and $n_i \leftarrow 0 $.
                     \tcp*[f]{global restart}
@@
 More precisely, \CUSUMUCB{} samples arm $i$ with probability $\alpha/K$ at each time step,
-and \MUCB{} uses a deterministic scheme, to force exploring the $K$ arms: when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$.
+and \MUCB{} uses a deterministic scheme, to force exploring the $K$ arms: when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\alpha \right\rceil$.
 Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
@@
 \begin{proof}
-    We consider one arm $i\in\{1,\dots,K\}$, and when the \GLRklUCB{} algorithm is running,
+    We consider one arm $i\in[K]$, and when the \GLRklUCB{} algorithm is running,
     we consider two time steps $s\leq t\in\N^*$, chosen between two restart times for that arm $i$.
     Lines~$3$-$4$ state that $A_u = u \mod \lceil \frac{K}{\alpha} \rceil$
-    if $u \mod \lceil \frac{K}{\alpha} \rceil \in \{1,\dots,K\}$,
+    if $u \mod \lceil \frac{K}{\alpha} \rceil \in [K]$,
     % (see Algorithm~\ref{algo:6:GLRklUCB} for details).
@@
 Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instances.
-The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including Exp3, and recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17},
+The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17},
 and it usually performs similarly to \MklUCB, however it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.
@@
         \hline
-        Exp3.S & $352 \pm 51$ & $310 \pm 62$ & $665 \pm 93$ \\
+        Exp3.S & $352 \pm 51$ & $310 \pm 62$ & XXX \\
+        % XXX for pb3 = $665 \pm 93$
         \hline
@@
         \hline
-        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & $723 \pm 121$ \\
+        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & XXX \\
+        % XXX for pb5 = $723 \pm 121$ ?
         \hline
@@
 The authors build on their recent work \cite{WeiSrivastava18Abruptly} and show that a regret bounded by $\cO(\sqrt{T^{\frac{1+\nu}{2}} \log(T)})$ can still be achieved by $M$ players, in the number of breaking points is $\Upsilon_T = \cO(T^{\nu})$.
-Their algorithm assume that player $j$ knows its ID $j\in\{1,\dots,M\}$ as well as $\nu$, and removing these hypotheses is an interesting direction of future work.
+Their algorithm assume that player $j$ knows its ID $j\in[M]$ as well as $\nu$, and removing these hypotheses is an interesting direction of future work.
 % \TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed} by using \MCTopM{} and \GLRklUCB{} with each other?

### 2019-05-10 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Added two references, as suggested by Christophe.

@@
 We evoke quickly here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
-In $2015$ and $2016$, a PhD student (Navikkumar Modi) and post-doc (Sumit J. Darak) in our team SCEE worked with Christophe Moy and others for the following publications,
+Starting from $2010$, the works of Wassim Jouini \cite{Jouini09,Jouini10,Jouini12} were the first ones to propose to use Reinforcement Learning for Cognitive Radio, especially MAB and the \UCB{} algorithm, and from $2014$, proof-of-concepts were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
+In $2015$ and $2016$, Christophe Moy continued to work on this direction, with a PhD student (Navikkumar Modi) and post-doc (Sumit J. Darak) in our team SCEE, for the following publications
 \cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.

### 2019-05-10 2-Chapters/5-Chapter/chapter5.tex
Added two references, as suggested by Christophe.

@@
 
-% FIXME much more details?
-Other works in the same line of research included \cite{modiDemo2016} implemented using the same USRP testbed as for our demonstration in Chapter~\ref{chapter:4} (Section~\ref{sec:4:gnuradio}),
-and the first work of a research team based in IIT Dehli, \cite{kumar2016two}.
-%
-More recent works by the same team, \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}, all follow the same approach and include an experimental section that shows the results of an implementation of the proposed algorithms on USRP boards.
-% \TODOL{I should explain more, no? Who does what and why? So long...}
+% % FIXME much more details?
+% Other works in the same line of research included \cite{modiDemo2016} implemented using the same USRP testbed as for our demonstration in Chapter~\ref{chapter:4} (Section~\ref{sec:4:gnuradio}),
+% and the first work of a research team based in IIT Dehli, \cite{kumar2016two}.
+% %
+% More recent works by the same team, \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}, all follow the same approach and include an experimental section that shows the results of an implementation of the proposed algorithms on USRP boards.
+% % \TODOL{I should explain more, no? Who does what and why? So long...}
 

### 2019-05-10 2-Chapters/1-Chapter/chapter1.tex
:-( it took me 40 minutes for a useless figure. I'm stupid.

@@
 \subsection{Example: decentralized learning in unlicensed bands with selfish or orthogonalization-based MAB algorithms}
+
+\subsection{Reinforcement learning}
+
+\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.85\textwidth}{!}{
+        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=2]
+            %
+            % We start by placing the blocks
+            \node [block] (player) at (0,0) {Player};
+            % We draw an edge between the player and system block to
+            \node [block] (environment) at (2,0) {Environment};
+            % Once the nodes are placed, connecting them is easy.
+            \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Reward} (player);
+            %
+        \end{tikzpicture}
+    }
+\caption{Reinforcement learning cycle.}
+\label{fig:1:ReinforcementLearningCycle}
+\end{figure}
 

### 2019-05-10 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
...

@@
 %
-Thus it is fair to compare our proposal against the modified versions of the two algorithms CUSUM and M- that use the \klUCB{} index policy instead of \UCB, as they perform uniformly better with \klUCB{} than with \UCB{} (the same tendency was observed on all other problems).
+Thus it is fair to compare our proposal against the modified versions of the two algorithms \CUSUM{} and $\mathrm{M}$- that use the \klUCB{} index policy instead of \UCB, as they perform uniformly better with \klUCB{} than with \UCB{} (the same tendency was observed on all other problems).
 We also note that, without surprise, the oracle policy also performs (much) better with \klUCB{} than with \UCB.
@@
       \centering
-      \begin{tabular}{c|cc}
-        \textbf{Algorithm} $\;$ \textbackslash $\;$ Problems & Pb 1 & Pb 2 \\
+      \begin{tabular}{cc|cc}
+        \textbf{Algorithm} & \textbf{Index policy} & \textbf{Problem 1} & \textbf{Problem 2} \\
         \hline
-        Oracle-Restart \UCB{} & $216$ & $67$ \\
-        Oracle-Restart \klUCB{} & $\mathbf{54}$ & $\mathbf{219}$ \\
+        \multirow{2}{*}{Oracle-Restart}
+            & \UCB{} & $216$ & $219$ \\
+            & \klUCB{} & $\mathbf{54}$ & $\mathbf{67}$ \\
         \hline
-        \MUCB{} & $878$ & $2040$ \\
-        \MklUCB{} & $817$ & $1485$ \\
+        \multirow{2}{*}{$\mathrm{M}$-}
+            & \UCB{} & $878$ & $2040$ \\
+            & \klUCB{} & $817$ & $1485$ \\
         \hline
-        \CUSUMUCB{} & $439$ & $381$ \\
-        \CUSUMklUCB{} & $304$ & $316$ \\
+        \multirow{2}{*}{\CUSUM}
+            & \UCB{} & $439$ & $381$ \\
+            & \klUCB{} & $304$ & $316$ \\
         \hline
-        \GLRklUCB{}(Local) & $\mathbf{132}$ & $\mathbf{186}$
+        \multirow{2}{*}{\GLR{} (Local)}
+            &\UCB{} & XXX & XXX \\  % FIXME do it!
+            & \klUCB{} & $\mathbf{132}$ & $\mathbf{186}$
       \end{tabular}
@@
         \hline
-        Exp3.S & $352 \pm 51$ & $310 \pm 62$ & XXX \\
-        % XXX for pb3 = $665 \pm 93$
+        Exp3.S & $352 \pm 51$ & $310 \pm 62$ & $665 \pm 93$ \\
         \hline
@@
         \hline
-        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & XXX \\
-        % XXX for pb5 = $723 \pm 121$ ?
+        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & $723 \pm 121$ \\
         \hline
@@
 %
-First, we show the mean regret as a function of time (\ie, $R_t$ for $1 \leq t \leq T$), for the $9$ algorithms considered (Discounted-\klUCB{} is removed as it is always the most inefficient and suffers high regret).
+First, we show the mean regret as a function of time (\ie, $R_t$ for $1 \leq t \leq T$), for $9$ of the considered algorithms (as they are outperformed by the others, Exp3.S and Discounted-\klUCB{} are not included to avoid clutter).
 %

### 2019-05-12 2-Chapters/3-Chapter/logs/launch_simulations.sh
I launched large scale simulations (for one K, many T + for one T, many K), I'll explore them tomorrow.

@@ -1,37 +1,39 @@
-    #!/usr/bin/env bash
+#!/usr/bin/env bash
 set -euo pipefail
 
 DEBUG=False
-NOPLOTS=True
+NOPLOTS=False
 SAVEALL=True
-N_JOBS=12
+N_JOBS=-1
 
 declare -i nbexperiment
 nbexperiment=0
 
-for N in 100; do
+for N in 1000; do
 # for N in $N_JOBS; do
     # for BAYES in False True; do
-    for BAYES in False; do
+    # for BAYES in False; do
+    for BAYES in True; do
         # for T in 1000 2500 5000 10000 20000 25000 50000; do
-        for T in 10000; do
+        for T in 5000; do
             # for K in 2 4 8 12 16 20 32; do
-            for K in 6 10 14 18 22 24 26 28 32; do
             # for K in 2 4 8; do
+            for K in 2 4 6 8 10 12 14 16 18 20 22 24 26 28 30 32; do
                 debut=$(date)
                 nbexperiment+=1
 
                 DEBUG=$DEBUG NOPLOTS=$NOPLOTS SAVEALL=$SAVEALL N_JOBS=$N_JOBS \
                 N=$N T=$T K=$K BAYES=$BAYES mymake.sh --noFreeSMS single \
                 && cp -f logs/main_py3_log.txt \
-                            logs/main_py3_log__N${N}_BAYES${BAYES}_T${T}_K${K}__$(date +"%d-%m-%Y_%H-%M-%S").txt \
+                    logs/main_py3_log__N${N}_BAYES${BAYES}_T${T}_K${K}__$(date +"%d-%m-%Y_%H-%M-%S").txt \
                 && FreeSMS.py "Expérience #${nbexperiment}, commencée ${debut}, terminée ! Avec N=${N}, BAYES=${BAYES}, T=${T}, K=${K}."
             done
         done
-        for T in 15000 30000 35000 40000 45000; do
+        # for T in 15000 30000 35000 40000 45000; do
         # for T in 10000; do
+        for T in 1000 2000 5000 10000 15000 20000 25000 30000 35000 40000 45000 50000; do
             # for K in 2 4 8 12 16 20 32; do
-            for K in 32; do
+            for K in 8; do
             # for K in 6 10 14 18 22 24 26 28 32; do
             # for K in 2 4 8; do
                 debut=$(date)
[…]

### 2019-05-15 2-Chapters/2-Chapter/chapter2.tex
Still working on Chapter 2 and 3 simulatneously.
Ca avance bien YAS YAS YAS
So boring
Done... So tired.
Added screenshots of the demo now hosted at https://perso.crans.org/besson/phd/MAB_interactive_demo/

--- 2019-04-01 2-Chapters/2-Chapter/chapter2.tex
@@
 % First chapter begins here
-\chapter{Single Player and Stationary Multi-Armed Bandit Models}
+\chapter{The Single Player and Stationary Multi-Armed Bandits Model}
 \label{chapter:2}
@@
 
-Abstract for Chapter 2.
+\TODOL{L'abstract ici est trop long, mais je l'ai écrit avant de commencer le chapitre, pour avoir une idée de ce que je devais faire ! Je couperai à la fin !}
+
+\paragraph{Abstract.}
+%
+In this chapter, we present the common base of all the models studied in this thesis:
+the single-player and stationary stochastic MAB model.
+We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a one-dimensional distribution.
+In all this thesis, we consider one-dimensional exponential families, and in particular Bernoulli distributions.
+
+We review quickly the possible applications of bandits, that range from recommender systems and information retrieval to healthcare, finance.
+In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits can be applied with success.
+
+We then explain the measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and horizon $T$.
+The regret compares the performance of the decision-making player and the oracle who knows in advance the best arm and always selects it.
+Other measures of performance, such as best-arm identification, are not used later on in this document, but they are also shortly discussed.
+%
+We review the main theoretical results from famous previous works, that give a lower-bound on the regret of any algorithm, essentially showing that we cannot find an algorithm achieving a sub-logarithmic regret, in this stochastic MAB model.
+
+We give an overview of the main families of algorithms designed to tackle the stochastic MAB model, and we mainly focus on the \UCB, \klUCB{} and Thompson sampling algorithms, as they are the two we used in the next chapters.
+% Part of the work presented in Chapter~\ref{chapter:4} also used the Thompson sampling algorithm, but no theoretical analyses were done about it.
+When presenting our library SMPyBandits in Chapter~\ref{chapter:3}, designed to perform numerical simulations of MAB problem, we use it to compare the empirical performances of the most famous and most efficient algorithms, on different problems.
+%
+In the point-of-view of numerical simulations or real-world applications of bandits on embedded systems, one must also take care of two other measures of performance: time and memory complexities, which can vary greatly between different families of algorithms.
+We also give an empirical comparison of different algorithms for these two measures in Chapter~\ref{chapter:3}.
+
+We also gives regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
+They both attain a logarithmic regret, proving their order optimality (\ie, their regret upper-bound asymptotically matches the lower-bound), moreover \klUCB{} is known to be optimal for a wide range of arms distributions.
+% A sketch of the proof of both results is given in Appendix.
+
+Finally, we explore the problem of algorithm selection: as there are many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
+We present one of the first contribution made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms.
+We illustrate its performance empirically, but no satisfactory theoretical result is given, as the bound we had hoped for have been proved unachievable since then.
+
+\paragraph{Publication.}
+%
+Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on \cite{Besson2018WCNC}.
+The notations used in this thesis are the usual notations in all the MAB literature, and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
+
+\paragraph{Main references for the curious reader.}
+%
+For more details and a more pedagogical introduction to multi-armed bandits, we suggest to the interested reader to read the short text-book \cite{Slivkins2019}.
+Another excellent but reasonably short survey is \cite{Bubeck12}, while the more recent book \cite{LattimoreBanditAlgorithmsBook} is the most complete resource about bandit algorithms.
+Finally, we recommend \cite{bouneffouf2019survey} for a short but good survey on applications of MAB.
 
@@
 
+Multi-Armed Bandits models were first introduced by Thompson as early as 1933 in \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52}.
+This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
+The name \emph{multi-armed bandits} comes from the one-armed bandit found in casinos, see Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step0} below.
+% The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
+% Since then, many efficient algorithms were proposed, XXX
+
+%
+A MAB is a model of a decision-making game, where a \emph{player} has to sequentially select one action in a (usually finite) set of actions, and only receives a (random) feedback about the action she selected, also called \emph{reward}.
+More formally, we consider $K\in\N, K\geq2$ arms, and the player's decision at time $t$ is denoted $A(t)\in\{1,\dots,K\}$, after what she receives a reward $r_{A(t)}(t)\in\R$, and so on for $t\in\N^*$.
+The goal of the player is to maximize its sum of received rewards, from time $t=1$ to $t=T$ if the game is played at an horizon $T$.
+The difficulty comes from balancing the trade-off between \emph{exploration}, because she has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
+The MAB model is a famous example of a reinforcement learning model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
+We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
+
+\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.40\textwidth}{!}{
+        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=2]
+            %
+            % We start by placing the blocks
+            \node [block] (player) at (0,0) {Player};
+            % We draw an edge between the player and system block to
+            \node [block] (environment) at (2,0) {MAB problem};
+            % Once the nodes are placed, connecting them is easy.
+            \draw [->] (player) to[bend left=90] node[pos=0.5] {Discrete action $k=A(t) \in\{1,\dots,K\}$} (environment);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Real-valued reward $r_k(t) \in \R$} (player);
+            %
+        \end{tikzpicture}
+    }
+\caption{Reinforcement learning cycle in a MAB model, for time steps $t=1,\dots,T$.}
+\label{fig:2:ReinforcementLearningCycleMABmodel}
+\end{figure}
+
+The first application of MAB models was for clinical trials, were an arm represent a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
+The mean of an arm, in this application, represents the mean success rate of a treatment, and the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
+%
+The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication of a device to a gateway (\eg, spreading factor for LoRa, power allocation for NOMA etc).
+%
+After introducing formally the notations used in all this thesis, we review below possible applications of multi-armed bandits.
+
 % ----------------------------------------------------------------------------
@@
 
-In this thesis, we only consider models with finite arms.
-No linear bandits, no contextual bandits etc.
+In all this thesis, we focus on the model with finite arms, and binary or real-valued stochastic rewards, that is $A(t)\in\{1,\dots,K\}$ for a fixed and known number of arms $K\in\N,K\geq2$, and $r_{A(t)}(t)\in\R$.
+%
+Restricting to models with a finite number of arms
+rules out an entire domain of the recent research on MAB, in particular we do not consider linear nor contextual bandits (for more details, see Parts V and VI of \cite{LattimoreBanditAlgorithmsBook}).
 
-For now, we only consider stochastic rewards. Piece-wise stochastic models are studied in Chapter~\ref{chapter:6}.
+We also restrict to stochastic rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in\{1,\dots,\}=[K]$.
+Rewards are stochastic and stationary, meaning that $(r_k(t))_{t\in\N^*}$ is independent and identically distributed (\iid), $r_k(t) \sim \nu_k$ for any $t\geq1$.
+
+\paragraph{Interactive demo for the novice.}
+%
+If you are discovering the concept of bandit model here, I would like to recommend you to go online and play a few times with an interactive demonstration.
+On this demo, you are playing against a MAB problem with $K=5$ arms, when you have $T=100$ decisions to make.
+The demo is hosted online, on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}.
+
+The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
+The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r_{A(t)(t)$.
+Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r_{A(t)}(t)\in\{0,1}$, that is sampled \iid{} from $\nu_k$.
+%
+The UI of the demo also shows the current value of $t$ (``total plays'') and $\sum_{s=1}^t r_{A(s)}(s)$ the ``total reward''.
+For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
+%
+The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated probs'' line.
+
+We first illustrate in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} the current state of the game at time $t=24$.
+% Currently at time $t=24$ (out of $T=100$ total time steps),
+At this step, the player has collected a sum of rewards of $14$, by observing $X(t) = [6,2,2,2,2]$ rewards of value $1$ in the $K=5$ different arms.
+Arms were sampled $N(t) = [8,4,4,4,4]$ times, meaning that the value $0$ was seen respectively $[2,2,2,2,2]$ times, and currently arm $1$ appears to be the best one.
+The true means of the arms are $\bm{\mu}=[0.6, 0.2, 0.55, 0.7, 0.5]$, and (much) more samples are needed before the player can accurately identify arm $4$ as the best arm.
+%
+We also illustrate in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} the result of an example of run, when the player was following the \UCB{} algorithm from \cite{Auer02}.
+After $T=100$ steps, the player obtained a cumulated reward of $56$, by playing mostly arms $4,3,5,1,0$ (in decreasing order of number of plays).
+The empirical means $\widehat{\mu_k}(T)$ correctly identify the best arm (arm $4$), but do not correctly rank the arms as arms $1$ and $3$ obtained means of $\widehat{\mu_1}(T) = 0.5 < \widehat{\mu_3}(T)=0.6$ but the true means are $\mu_3 = 0.55 < \mu_1 = 0.6$
+
+% \begin{figure}[h!]  % [htbp]
+%     \centering
+%     \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step0.png}
+%     \caption{See }
+%     \label{fig:2:example_of_a_5_arm_bandit_problem__step0}
+% \end{figure}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
+    \caption{Screenshot of the demonstration available online on my website, for a current step of $t=24$.
+See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}.}
+    \label{fig:2:example_of_a_5_arm_bandit_problem}
+\end{figure}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
+    \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.}
+    \label{fig:2:example_of_a_5_arm_bandit_problem__step100}
+\end{figure}
+
+\TODOL{Write this!}
 
@@
 \subsection{Different hypotheses on rewards distributions}
+
+From now on until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
 
@@
 
-Markov models: presented later in Section~\ref{sec:3:markovModels}.
+Markov models: we do not present it in this thesis.
 
@@
 
+Citation: \cite{bouneffouf2019survey}
+
 \subsection{Application for Opportunistic Spectrum Access}
@@
 Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
+
+% ----------------------------------------------------------------------------
+\section{Lower and upper bounds on regret}
+\label{sec:2:lowerUpperBoundsRegret}
+
+\subsection{Measuring performance with the (mean) regret}
+
+Explain what is regret
+
+\subsection{Regret lower bounds: do not detail too much, don't explain the tools behind the results}
+
+- [Lai and Robbins] lower-bound in $\Omega(\log(T))$
+- Worst case lower-bound in $\Omega(\sqrt{T})$
+- Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models)
+
+\subsection{Regret upper bounds, tools and different kinds of bounds}
+
+- useful decomposition, with number of samples (see Claire's PhD)
+- different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
+- tools to show regret upper-bounds, concentration inequalities,
+- short proof of UCB or klUCB? just to start to show the tools
+
+\subsection{Other measures of performances}
+
+- Best Arm Identification?
+- Quantile regret or just histogram of regrets
+- Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
+- Fairness
 
@@
 % ----------------------------------------------------------------------------
-\section{Lower and upper bounds on regret}
-\label{sec:2:lowerUpperBoundsRegret}
-
-\subsection{Measuring performance with the (mean) regret}
-Explain what is regret
-
-\subsection{Regret lower bounds: do not detail too much, don't explain the tools behind the results}
-
-- [Lai and Robbins] lower-bound in $\Omega(\log(T))$
-- Worst case lower-bound in $\Omega(\sqrt{T})$
-- Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models)
-
-\subsection{Regret upper bounds, tools and different kinds of bounds}
-
-- useful decomposition, with number of samples (see Claire's PhD)
-- different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
-- tools to show regret upper-bounds, concentration inequalities,
-- short proof of UCB or klUCB? just to start to show the tools
-
-\subsection{Other measures of performances}
-
-- Quantile regret or just histogram of regrets
-- Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-- Fairness
-
-% ----------------------------------------------------------------------------
 \section{Different approaches on algorithm selection}
@@
 % ----------------------------------------------------------------------------
-\section{Conclusion of Chapter 2}
+\section{Conclusion}
 \label{sec:2:conclusion}
 
-In this chapter, we saw...
+In this chapter, presented
 
-Future works include...
+\TODOL{Write this!}
 
 % ----------------------------------------------------------------------------
-\section{Appendix for Chapter 2}
+\section{Appendix}
 \label{sec:2:appendix}
 
-\subsection{Eficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
+\TODOL{Write this ?}
+
+\subsection{Efficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
+
+\TODOL{On s'en fout en fait de ça, non ? Ma thèse n'a pas étudiée klUCB, et les implémentations en C ne sont pas de moi mais viennent de pymaBandits... Donc inutile de parler de ça !}
 

### 2019-05-15 2-Chapters/2-Chapter/chapter2.tex
Je ne sais juste pas quoi écrire c'est horrible

@@
 After $T=100$ steps, the player obtained a cumulated reward of $56$, by playing mostly arms $4,3,5,1,0$ (in decreasing order of number of plays).
-The empirical means $\widehat{\mu_k}(T)$ correctly identify the best arm (arm $4$), but do not correctly rank the arms as arms $1$ and $3$ obtained means of $\widehat{\mu_1}(T) = 0.5 < \widehat{\mu_3}(T)=0.6$ but the true means are $\mu_3 = 0.55 < \mu_1 = 0.6$
+The empirical means $\widehat{\mu_k}(T)$ correctly identify the best arm (arm $4$), but do not correctly rank the arms as arms $1$ and $3$ obtained means of $\widehat{\mu_1}(T) = 0.5 < \widehat{\mu_3}(T)=0.6$ but the true means are $\mu_3 = 0.55 < \mu_1 = 0.6$.
 
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
+    \includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
     \caption{Screenshot of the demonstration available online on my website, for a current step of $t=24$.
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
+    \includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
     \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.}
@@
 
-\TODOL{Write this!}
-
 % ----------------------------------------------------------------------------
 \subsection{Different hypotheses on rewards distributions}
+
+\TODOL{Write this!}
 

### 2019-05-15 2-Chapters/2-Chapter/chapter2.tex
No Appendix for Chapter 2, and a one-page conclusion.
I'm writing abstracts and intros and conclusions but I fail to write FUCKING TEXT OH SHIT.

@@
 
+\newpage
+
 % ----------------------------------------------------------------------------
@@
 
+\newpage  % FIXME
+
 % ----------------------------------------------------------------------------
@@
 
-In this chapter, presented
+In this chapter, we presented the multi-armed bandit model, focussing on a finite number of arms, and real-valued rewards.
+Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
+By showcasing an interactive demonstration made available online\footnote{~The demo is hosted on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}.}, we presented the notations used in all this thesis.
 
-\TODOL{Write this!}
+We ended this chapter by presenting the first contribution of this manuscript \cite{Besson2018WCNC}, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
+Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on a set of small bandit problems.
+%
+The numerical simulations used our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
+The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.
 
-% ----------------------------------------------------------------------------
-\section{Appendix}
-\label{sec:2:appendix}
+We presented the simplest MAB model studied in this chapter, by focussing on one agent playing a stationary game.
+Both hypotheses are removed or weaken in the rest of this manuscript,
+% first in Chapter~\ref{chapter:4} where we consider many XXX
+first by considering players facing a stationary problem in a decentralized way in Chapter~\ref{chapter:5},
+and then by considering a single player facing a non-stationary or a piece-wise stationary problem in Chapter~\ref{chapter:6}.
+%
+For both directions, we present in the two final chapters natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
+of stationary multi-player and piece-wise stationary bandits.
 
-\TODOL{Write this ?}
+We take another approach in Chapter~\ref{chapter:4}, where the MAB model is generalized to study decentralized learning of a large set of independent players, all having different activation times.
+This extension is significantly harder than the two previously evoked ones, and we were unfortunately unable to obtain any strong theoretical results under these loose hypotheses.
+This model is however more generic and as such it was found suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
 
-\subsection{Efficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
+\paragraph{Possible future works.}
+%
+We focussed in this thesis on finite-arms and one-dimensional bandit problems,
+and thus two possible directions of future works could be to extend our works
+to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
 
-\TODOL{On s'en fout en fait de ça, non ? Ma thèse n'a pas étudiée klUCB, et les implémentations en C ne sont pas de moi mais viennent de pymaBandits... Donc inutile de parler de ça !}
+% % ----------------------------------------------------------------------------
+% \section{Appendix}
+% \label{sec:2:appendix}
 
-I want to detail the maths behind an implementation of klUCB, see https://smpybandits.github.io/docs/Policies.kullback.html and all what is detailed there.
+% \TODOL{Write this ?}
 
-How to write a fast version of klBern and klucbBern (the two most used functions), using C or Numba or Cython, maybe I can include the speed up simulations here?
+% \subsection{Efficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
+
+% \TODOL{On s'en fout en fait de ça, non ? Ma thèse n'a pas étudiée klUCB, et les implémentations en C ne sont pas de moi mais viennent de pymaBandits... Donc inutile de parler de ça !}
+
+% I want to detail the maths behind an implementation of klUCB, see https://smpybandits.github.io/docs/Policies.kullback.html and all what is detailed there.
+
+% How to write a fast version of klBern and klucbBern (the two most used functions), using C or Numba or Cython, maybe I can include the speed up simulations here?
 First piece of code, before really presenting SMPyBandits ?
 
-My work for this part is also published on https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes
-(in Python) and https://github.com/Naereen/KullbackLeibler.jl (in Julia).
+% My work for this part is also published on https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes
+% (in Python) and https://github.com/Naereen/KullbackLeibler.jl (in Julia).

### 2019-05-15 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
So long

@@
 
+\TODOL{Je pense refaire complètement les expériences de ce morceau.
+
+- C'est un peu bête de montrer des problèmes chelous genre Gaussien et "mixed" quand tout le reste de la thèse s'intéresse vraiment aux Bernoulli.
+
+- C'est un peu naze de montrer en log-y ou en log-log, alors que le reste de la thèse s'intéresse uniquement à des plots x-y, ou des histogrammes.
+
+- Pour être plus consistant avec les sections suivantes, je pourrais inclure juste 1/2 courbes de regret, et un tableau de résultats pour d'autres problèmes ?
+}
+
 % - explain the problem studied in the simulation

### 2019-05-15 2-Chapters/2-Chapter/chapter2.tex
So long

@@
 
+\newpage
 \TODOL{L'abstract ici est trop long, mais je l'ai écrit avant de commencer le chapitre, pour avoir une idée de ce que je devais faire ! Je couperai à la fin !}
@@
 In all this thesis, we consider one-dimensional exponential families, and in particular Bernoulli distributions.
-
+%
 We review quickly the possible applications of bandits, that range from recommender systems and information retrieval to healthcare, finance.
@@
 \TODOL{Write this!}
+In the example above, we consider Bernoulli distributions, but other distributions have been studied in the literature.
+From now on until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
+%
+% Everything is implemented and documented at
+% https://smpybandits.github.io/docs/Arms.html
+We also focus on real-valued rewards, meaning that $r_k(t)\in\R$ for all $k$ and $t$.
+An important hypothesis is whether rewards are bounded or not,
+and whether the player knows if they are bounded or not before starting the bandit game.
+Moreover, if rewards are known to be bounded, let say in an interval $[a,b]$, another important hypothesis is whether the player knows the values of $a$ and $b$.
+%
+Intuitively, the bandit game is easier if the player knows the support of the distributions, and we restrict to this case in all the thesis.
+Most of the algorithms proposed in the literature follow this hypothesis as well.
 
-From now on until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
+The mostly used
+infinitely supported distributions are Gaussian, exponential and Poisson,
+while in the literature, the Bernoulli distribution is the most common case of finitely supported distributions.
 
-Everything is implemented and documented at
-https://smpybandits.github.io/docs/Arms.html
-
-Bernoulli, Gaussian, sub-Bernoulli, sub-Gaussian, Exponential, sub-Exponential...
+% Bernoulli, Gaussian, sub-Bernoulli, sub-Gaussian, Exponential, sub-Exponential...
 
@@
 
-Markov models: we do not present it in this thesis.
+\paragraph{About Markov models.}
+%
+% Markov models: we do not present it in this thesis.
+Finally, we note that Markov models, while being implemented in SMPyBandits, are not used in this thesis.
+They were introduced in the 1980s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
+A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and such models come in two flavors: rested or restless.
+For $K$ arms, each Markov chain has a finite number of state $s$, each corresponding to a (constant) reward that the player obtains if she selects this arm while its Markov chain is in state $s$.
+Rested Markov models means that only the state of the selected arm's Markov chain can change, following its Markov transition matrix.
+Restless models remove this hypothesis, making them harder to track and solve.
+%
+Such models were less studied than stationary or adversarial models, but some interesting works focussed on Markov models in the last 10 years.
+For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
+
+\TODOL{On garde ce petit paragraphe, par curiosité ?}
 
@@
 \label{sec:2:applicationsofStationaryMAB}
+
+The blooming success of the research on multi-armed bandits is easily explained by the different spectrum of applications of MAB models to real-world discrete-time decision making problems.
+This research field has been very active since the years 2010s, but it started as early as 1933 with \cite{Thompson33}, and was active since the 1980s \cite{LaiRobbins85,Anantharam87a,Anantharam87b}.
 
@@
 Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
+
+\cite{Jouini09,Jouini10,Jouini12}
+\cite{MoyWSR2014}
+\cite{RobertSDR2014}
+\cite{modiDemo2016,Modi17QoS}
 
@@
 Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
-By showcasing an interactive demonstration made available online\footnote{~The demo is hosted on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}.}, we presented the notations used in all this thesis.
+By showcasing an interactive demonstration made available online,
+% XXX ?
+% (\href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}),
+we presented the notations used in all this thesis.
 
-We ended this chapter by presenting the first contribution of this manuscript \cite{Besson2018WCNC}, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+The first contribution of this manuscript \cite{Besson2018WCNC} concluded this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
 Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on a set of small bandit problems.
+We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
 %

### 2019-05-15 2-Chapters/2-Chapter/chapter2.tex
Just shit

@@
 % https://smpybandits.github.io/docs/Arms.html
-We also focus on real-valued rewards, meaning that $r_k(t)\in\R$ for all $k$ and $t$.
+We also focus only on real-valued rewards, meaning that $r_k(t)\in\R$ for all $k$ and $t$.
 An important hypothesis is whether rewards are bounded or not,
@@
 Most of the algorithms proposed in the literature follow this hypothesis as well.
-
+%
 The mostly used
@@
 while in the literature, the Bernoulli distribution is the most common case of finitely supported distributions.
+Continuous distributions with finite support also included truncated versions of infinitely supported distributions, in particular truncated Gaussian are used for numerical experiments in lots of research articles.
+
+% Just bounded, in $[a,b]$ but without loss of generality in $[0,1]$ etc.
+\paragraph{The normalization trick.}
+If the player knows that reward are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ ($a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = $ that lies in $[0,1]$.
+Note that this ``normalization trick'' is implemented for any policy in SMPyBandits, with the \texttt{lower} and \texttt{amplitude} optional arguments, respectively representing $a$ the lower-bound on rewards and $b-a$ the length of the interval of possible values of rewards.
 
@@
 
-One dimensional exponential family...
+\paragraph{On one-dimensional exponential family.}
+%
+On one-dimensional exponential family...
 
-Just bounded, in $[a,b]$ but without loss of generality in $[0,1]$ etc.
+\paragraph{Other kind of hypothesis: sub-Gaussian and sub-Bernoulli distributions.}
+%
+A lot of research works considers rewards distributions that are not Gaussian but sub-Gaussian, of a known variance, for instance $1/4$.
+For instance, bounded distributions on $[0,1]$ are $1/4$ sub-Gaussian, and this fact is used for instance in \cite{Maillard2018GLR}.
+It means that their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$).
+In Chapter~\ref{chapter:6}, we instead consider sub-Bernoulli distributions, formally introduced in Definition~\ref{def:6:subBernoulliDistributions}.
+%
+Such hypothesis is also proposed for other families of distributions, for instance \cite{KimTewari2019} analyses their Follow-the-Perturbed-Leader algorithm for perturbations following any sub-Weibull distribution (sub-Weibull distributions generalize both sub-Gaussian and sub-Exponential distributions).
 
@@
 The blooming success of the research on multi-armed bandits is easily explained by the different spectrum of applications of MAB models to real-world discrete-time decision making problems.
-This research field has been very active since the years 2010s, but it started as early as 1933 with \cite{Thompson33}, and was active since the 1980s \cite{LaiRobbins85,Anantharam87a,Anantharam87b}.
+This research field has been very active since the years 2010s, but it started as early as 1933 with \cite{Thompson33}, and was active since the 1980s and the seminal works by Lai and Robbins \cite{LaiRobbins85} and by Anantharam and others \cite{Anantharam87a,Anantharam87b}.
+%
+MAB have been applied successfully to various decision making problems, including clinical trials and cognitive radio that we mentioned above.
 
-\subsection{Various possible applications}
+Other popular applications include the following:
+\begin{itemize}
+    \item
+    A/B testing of websites
+    \cite{Kaufmann14,Jamieson17ABTest}
 
-- clinical trial
-- A/B testing of websites
+    \item
     - online content recommandation
-- wireless communication, lots of possibilities, one is frequency/channel selection in the OSA paradigm, detailed after
+    \cite{Li10,Louedec16}
+
+    \item
     - hyper-parameters tuning of deep neural networks etc
+\end{itemize}
 
-Citation: \cite{bouneffouf2019survey}
+% - clinical trial
+% - A/B testing of websites
+% - online content recommandation
+% - wireless communication, lots of possibilities, one is frequency/channel selection in the OSA paradigm, detailed after
+% - hyper-parameters tuning of deep neural networks etc
 
-\subsection{Application for Opportunistic Spectrum Access}
+Other interesting applications are discussed in \cite{bouneffouf2019survey}.
 
-Detail the analogy arm = channel in a licensed spectrum, rewards $r_k(t)$ = 1 if no Primary User or 0 if channel $k$ is busy at time $t$.
-
-Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
-
-\cite{Jouini09,Jouini10,Jouini12}
-\cite{MoyWSR2014}
-\cite{RobertSDR2014}
-\cite{modiDemo2016,Modi17QoS}
+\paragraph{Application for Opportunistic Spectrum Access.}
+% Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
+%
+Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
+In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
+In the model with sensing, the rewards $r_k(t)$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
+A reward of $r_k(t) = 1$ indicates that no Primary User was sensed, while a reward of $r_k(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
+%
+This model was first introduced by Wassim Jouini during his PhD thesis,
+in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
+Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.
+In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
+of using MAB algorithms to optimize channel selection
+on the battery life of a wireless device.
+On the one hand, running a MAB algorithm such as UCB-like algorithms was proved to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
+On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
+Each hardware reconfiguration costs energy for the device, and quickly switching algorithms will lead to a reduction of the battery life.
+The tradeoff between the two aspects is studied empirically in
+\cite{modiDemo2016}.
+%
+Another interesting works is \cite{Modi17QoS}, FIXME
 

### 2019-05-15 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
I launched large scale simulations (for one K, many T + for one T, many K), I'll explore them tomorrow.
Just shit

@@
 
-\begin{definition}[Sub-Bernoulli distributions]
+\begin{definition}[Sub-Bernoulli distributions]\label{def:6:subBernoulliDistributions}
     We call a \emph{sub-Bernoulli distribution} any distribution $\nu$ that satisfies
@@
         \multirow{2}{*}{\GLR{} (Local)}
-            &\UCB{} & XXX & XXX \\  % FIXME do it!
+            &\UCB{} & $191$ & $232$ \\
             & \klUCB{} & $\mathbf{132}$ & $\mathbf{186}$

### 2019-05-16 2-Chapters/2-Chapter/chapter2.tex
Just hard :cry:

@@
 In this chapter, we present the common base of all the models studied in this thesis:
-the single-player and stationary stochastic MAB model.
+the multi-armed bandit (MAB) model,
+restricting to the single-player and stationary stochastic case.
 We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a one-dimensional distribution.
-In all this thesis, we consider one-dimensional exponential families, and in particular Bernoulli distributions.
+In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
 %
-We review quickly the possible applications of bandits, that range from recommender systems and information retrieval to healthcare, finance.
-In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits can be applied with success.
+We review quickly the possible applications of bandits, that range from recommender systems and information retrieval to healthcare and finance.
+In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits has already been applied with success.
 
@@
 %
-We review the main theoretical results from famous previous works, that give a lower-bound on the regret of any algorithm, essentially showing that we cannot find an algorithm achieving a sub-logarithmic regret, in this stochastic MAB model.
+We review the main theoretical results from famous previous works, that give a lower-bound on the regret of any algorithm, essentially showing that we cannot find an algorithm achieving a sub-logarithmic regret, in this stochastic MAB model ($R_T^{\cA} = \Omega(\log(T))$).
 
-We give an overview of the main families of algorithms designed to tackle the stochastic MAB model, and we mainly focus on the \UCB, \klUCB{} and Thompson sampling algorithms, as they are the two we used in the next chapters.
-% Part of the work presented in Chapter~\ref{chapter:4} also used the Thompson sampling algorithm, but no theoretical analyses were done about it.
-When presenting our library SMPyBandits in Chapter~\ref{chapter:3}, designed to perform numerical simulations of MAB problem, we use it to compare the empirical performances of the most famous and most efficient algorithms, on different problems.
+We give an overview of the main families of algorithms designed to tackle the stochastic MAB model, and we mainly focus on the \UCB, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
+When presenting our library SMPyBandits in next Chapter~\ref{chapter:3}, designed to perform numerical simulations of MAB problem, we use it to compare the empirical performances of the most famous and most efficient algorithms, in terms of regret and by evaluating them on different problems.
 %
 In the point-of-view of numerical simulations or real-world applications of bandits on embedded systems, one must also take care of two other measures of performance: time and memory complexities, which can vary greatly between different families of algorithms.
-We also give an empirical comparison of different algorithms for these two measures in Chapter~\ref{chapter:3}.
+We also give an empirical comparison of different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
 
-We also gives regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
-They both attain a logarithmic regret, proving their order optimality (\ie, their regret upper-bound asymptotically matches the lower-bound), moreover \klUCB{} is known to be optimal for a wide range of arms distributions.
+We also give regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
+They both attain a logarithmic regret, proving their order optimality (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
+Moreover, \klUCB{} is known to be optimal for a wide range of arms distributions.
 % A sketch of the proof of both results is given in Appendix.
@@
 %
-For more details and a more pedagogical introduction to multi-armed bandits, we suggest to the interested reader to read the short text-book \cite{Slivkins2019}.
-Another excellent but reasonably short survey is \cite{Bubeck12}, while the more recent book \cite{LattimoreBanditAlgorithmsBook} is the most complete resource about bandit algorithms.
+For more details and a more formal introduction to multi-armed bandits, we suggest to the interested reader to work on a very recent text-book by Slivkins \cite{Slivkins2019}.
+Another excellent but reasonably short survey is the book by Bubeck and Cesa-Bianchi \cite{Bubeck12}, while the more recent book by Lattimore and Szepesv{\'a}ri \cite{LattimoreBanditAlgorithmsBook} is the most complete resource about bandit algorithms.
 Finally, we recommend \cite{bouneffouf2019survey} for a short but good survey on applications of MAB.
@@
 
-Multi-Armed Bandits models were first introduced by Thompson as early as 1933 in \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52}.
+Multi-Armed Bandits models were first introduced by Thompson as early as 1933 in \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
 This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
-The name \emph{multi-armed bandits} comes from the one-armed bandit found in casinos, see Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step0} below.
+Their name comes from the one-armed bandit found in casinos, see Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
 % The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
@@
 
+A MAB is a model of a decision-making game, where a \emph{player} has to sequentially select one action in a (usually finite) set of actions, and only receives a (random) feedback about the selected action, also called a \emph{reward}.
+We consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$.
+When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r_{A(t)}(t)\in\R$, and so on for a finite number of steps $t$, from $t=1$ until $t=T$ for an horizon $T$.
 %
-A MAB is a model of a decision-making game, where a \emph{player} has to sequentially select one action in a (usually finite) set of actions, and only receives a (random) feedback about the action she selected, also called \emph{reward}.
-More formally, we consider $K\in\N, K\geq2$ arms, and the player's decision at time $t$ is denoted $A(t)\in\{1,\dots,K\}$, after what she receives a reward $r_{A(t)}(t)\in\R$, and so on for $t\in\N^*$.
-The goal of the player is to maximize its sum of received rewards, from time $t=1$ to $t=T$ if the game is played at an horizon $T$.
-The difficulty comes from balancing the trade-off between \emph{exploration}, because she has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
-The MAB model is a famous example of a reinforcement learning model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
-We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
+A commonly studied goal for the player is to maximize its sum of received rewards, $\sum_{t=1}^T r_{A(t)}(t)$
+(or its expectancy in stochastic and stationary models).
+
+From a mathematical point-of-view, this quantity $\sum_{t=1}^T r_{A(t)}(t)$ is indeed random, and it depends on two aspects.
+First, the cumulated rewards depend on the randomness from the environment, that is, the unknown and unpredictable values of $r_k(t)$ for all $k$ and $t$.
+Then, this quantity also depends on the decision-making policy followed the player, that is, the choices $A(t)$, that are based on (all) the past observations ($A(1), r_{A(1)}(1), \dots, A(t-1), r_{A(t-1)}(t-1)$), and possibly on an external source of randomness.
+Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, following the uniform distribution on $[0,1]$, and that has to be independent from the rewards $r_k(t)$.
+This source of randomness is used in most MAB algorithm, for instance to break ties in a uniformly random manner.
+%  in the decision making rule.
 
@@
     \centering
-    \resizebox{0.40\textwidth}{!}{
-        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=2]
+    \resizebox{0.50\textwidth}{!}{
+        \begin{tikzpicture}[auto, node distance=5cm, >=latex]
             %
@@
             % We draw an edge between the player and system block to
-            \node [block] (environment) at (2,0) {MAB problem};
+            \node [block] (environment) at (4,0) {MAB problem};
             % Once the nodes are placed, connecting them is easy.
-            \draw [->] (player) to[bend left=90] node[pos=0.5] {Discrete action $k=A(t) \in\{1,\dots,K\}$} (environment);
-            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Real-valued reward $r_k(t) \in \R$} (player);
+            \draw [->] (player) to[bend left=90] node[pos=0.5] {Chooses a discrete action $k=A(t) \in\{1,\dots,K\}$} (environment);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Observes a real-valued reward $r_k(t) \in \R$} (player);
             %
@@
 
-The first application of MAB models was for clinical trials, were an arm represent a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
-The mean of an arm, in this application, represents the mean success rate of a treatment, and the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
-%
-The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication of a device to a gateway (\eg, spreading factor for LoRa, power allocation for NOMA etc).
-%
-After introducing formally the notations used in all this thesis, we review below possible applications of multi-armed bandits.
+The difficulty of this decision-making game comes from balancing the trade-off between \emph{exploration}, because the player has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
+The MAB model is a famous example of a reinforcement learning model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
+We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
+For more details on reinforcement learning, in particular about more generic models, we refer to the famous book \cite{SuttonBarto2018}.
 
@@
 rules out an entire domain of the recent research on MAB, in particular we do not consider linear nor contextual bandits (for more details, see Parts V and VI of \cite{LattimoreBanditAlgorithmsBook}).
-
-We also restrict to stochastic rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in\{1,\dots,\}=[K]$.
-Rewards are stochastic and stationary, meaning that $(r_k(t))_{t\in\N^*}$ is independent and identically distributed (\iid), $r_k(t) \sim \nu_k$ for any $t\geq1$.
+%
+We also restrict to \emph{stochastic} rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in\{1,\dots,\}=[K]$.
+Rewards are stationary, meaning that $(r_k(t))_{t\in\N^*}$ is independent and identically distributed (\iid), $r_k(t) \sim \nu_k$ for any $t\geq1$.
 
@@
 The demo is hosted online, on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}.
-
-The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
-The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r_{A(t)(t)$.
-Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r_{A(t)}(t)\in\{0,1}$, that is sampled \iid{} from $\nu_k$.
-%
-The UI of the demo also shows the current value of $t$ (``total plays'') and $\sum_{s=1}^t r_{A(s)}(s)$ the ``total reward''.
-For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
-%
-The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated probs'' line.
-
-We first illustrate in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} the current state of the game at time $t=24$.
-% Currently at time $t=24$ (out of $T=100$ total time steps),
-At this step, the player has collected a sum of rewards of $14$, by observing $X(t) = [6,2,2,2,2]$ rewards of value $1$ in the $K=5$ different arms.
-Arms were sampled $N(t) = [8,4,4,4,4]$ times, meaning that the value $0$ was seen respectively $[2,2,2,2,2]$ times, and currently arm $1$ appears to be the best one.
-The true means of the arms are $\bm{\mu}=[0.6, 0.2, 0.55, 0.7, 0.5]$, and (much) more samples are needed before the player can accurately identify arm $4$ as the best arm.
-%
-We also illustrate in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} the result of an example of run, when the player was following the \UCB{} algorithm from \cite{Auer02}.
-After $T=100$ steps, the player obtained a cumulated reward of $56$, by playing mostly arms $4,3,5,1,0$ (in decreasing order of number of plays).
-The empirical means $\widehat{\mu_k}(T)$ correctly identify the best arm (arm $4$), but do not correctly rank the arms as arms $1$ and $3$ obtained means of $\widehat{\mu_1}(T) = 0.5 < \widehat{\mu_3}(T)=0.6$ but the true means are $\mu_3 = 0.55 < \mu_1 = 0.6$.
 
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
-    \caption{Screenshot of the demonstration available online on my website, for a current step of $t=24$.
-See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}.}
+    \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
+    \caption{Screenshot of the demonstration available online on my website, for a current step of $t=24$.}
     \label{fig:2:example_of_a_5_arm_bandit_problem}
 \end{figure}
+
+The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} above.
+The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r_{A(t)(t)$.
+Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r_{A(t)}(t)\in\{0,1}$, that is sampled \iid{} from $\nu_k$.
+%
+The UI of the demo also shows the current value of $t$ (``total plays'') and $\sum_{s=1}^t r_{A(s)}(s)$ the ``total reward''.
+For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
+%
+The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated probs'' line.
+
+In the first Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem}, the current state of the game is shown at time $t=24$.
+% Currently at time $t=24$ (out of $T=100$ total time steps),
+At this step, the player has collected a sum of rewards of $14$, by observing $X(t) = [6,2,2,2,2]$ rewards of value $1$ in the $K=5$ different arms.
+Arms were sampled $N(t) = [8,4,4,4,4]$ times, meaning that the value $0$ was seen respectively $[2,2,2,2,2]$ times, and currently arm $1$ appears to be the best one.
+The true means of the arms are $\bm{\mu}=[0.6, 0.2, 0.55, 0.7, 0.5]$, and (much) more samples are needed before the player can accurately identify arm $4$ as the best arm.
+%
+In the second Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} below, we display the result of an example of run, when the player was following the \UCB{} algorithm from \cite{Auer02}.
+After $T=100$ steps, the player obtained a cumulated reward of $56$, by playing mostly arms $4,3,5,1,0$ (in decreasing order of number of plays).
+The empirical means $\widehat{\mu_k}(T)$ correctly identify the best arm (arm $4$), but do not correctly rank the arms as arms $1$ and $3$ obtained means of $\widehat{\mu_1}(T) = 0.5 < \widehat{\mu_3}(T)=0.6$ but the true means are $\mu_3 = 0.55 < \mu_1 = 0.6$.
 
@@
     \includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
-    \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.}
+    \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.
+See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
     \label{fig:2:example_of_a_5_arm_bandit_problem__step100}
 \end{figure}
+
+\paragraph{A bandit policy is a measurable function.}
+
+A bandit algorithm $\cA$, also referred to as a strategy or a policy, and sometimes denoted by $\pi$ or $\rho$, selects an arm $A(t)$ at time $t$ by using the past observations and the past external randomness.
+More formally, an algorithm is thus a sequence of measurable functions $(\cA_t)_{t\geq1}$,
+where $\cA_t$ maps the past observations $\cO_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$
+to an arm $\cA_t(\cO_t) = A(t) \in[K]$,
+if we denote $Y_s = r_{A(s)}(s)$ she $s$-th reward.
+The initial information is reduced to $\cO_1 = (U_0)$, and the first decision is $A(1) = \cA_1(\cO_1)$.
+Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps.
+%
+An algorithm is said to be deterministic if it does not depend on the external randomness $U_0,U_1,\dots$, but in this thesis we only use non-deterministic strategies.
+
+The most common objective for the player is to maximize its expected rewards, that is $\E[\sum_{t=1}^T Y_t] = \E[\sum_{t=1}^T r_{A(t)}(t)]$.
+The expectancy is taken on the randomness of the rewards $(r_k(t))_{k,t}$ as well as the randomness $(U_s)_s$.
+
+\TODOL{A finir, a faire comme dans https://tel.archives-ouvertes.fr/tel-02121614/document page 19}
 
@@
 \paragraph{The normalization trick.}
+%
 If the player knows that reward are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ ($a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = $ that lies in $[0,1]$.
@@
 %
-On one-dimensional exponential family...
+% On one-dimensional exponential family...
+
+In all this thesis we restrict to distributions that have a certain form, and focus on real-valued distributions lying in an exponential families.
+One-dimensional exponential families include Bernoulli and Poisson distributions, as well as Gaussian distributions with a fixed variance.
+Our main interest is Bernoulli distributions, but we prefer to present here the notations on a generic exponential family.
+We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Emilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
+
+Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) have the following form:
+\[ \Pr(x | \lambda) = h(x) \exp(\eta^T T(x) - A(\eta)), \]
+for a parameter vector $\eta$ (the canonical parameter), and given functions $T$ (the sufficient statistic) and $h$.
+$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$, $h$, $T$,
+$A(\eta) = \log \int h(x) \exp(\eta^T T(x)) \lambda(\mathrm{d} x)$.
+%
+The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
+and usually the literature focusses on regular and minimal exponential families (when the natural parameter space is an non-empty open set in $\R$).
+
+An important result on exponential families is that a distribution in such family is entirely characterized by its parameter $\eta$.
+That is why such one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[T(X)]$.
+
+A second important result is a simplified form for the Kullback-Leibler divergence \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
+The KL divergence between two distributions $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ is defined as
+\[ \KL(\Pr(x|\eta_1), \Pr(x|\eta_2)) = \int \Pr(x|\eta_1) \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) \lambda(\mathrm{d}x) = \E_{d_1}[ \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) ]. \]
+%
+So it can be written $\KL( \eta_1,  \eta_2)) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$ and simplified to
+\[ \KL(\Pr(x|\eta_1), \Pr(x|\eta_2)) = \E_{d_1}[ (\eta_1 - \eta_2)^T T(X) ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2)^T \mu_1 - A(\eta_1) + A(\eta_2), \]
+where $\mu_1 = \E_{\eta_1}[T(X)]$ is the mean of distribution $d_1$.
+
+Without diving too much in the details, it is interesting to illustrate these notations for two examples.
+On the first hand, Bernoulli distributions can be seen as an exponential family with
+$T(x) = x$ and $h(x) = 1$,
+and for a Bernoulli distribution of mean $\mu\in[0,1]$,
+the parameter is $\eta = \mu / (1 - \mu)$, giving $A(\eta) = \log(1 + \mathrm{e}^{\eta})$
+(with the limit behavior $\eta=+\infty$ if $\mu=0$).
+%
+On the other hand, Gaussian distributions use
+$T(x) = [x ; x^2]^T$ and $h(x) = 1/\sqrt(2\pi)$,
+and for $\cN(\mu,\sigma^2)$ a Gaussian distribution with mean $\mu\in\R$ and variance $\sigma^2$,
+$\eta = [\mu/\sigma^2 ; -1/\sigma^2]^T$, giving $A(\eta) = -\eta_1^2/4\eta_2 -1/2 \log(-2\eta_2)$ (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
+Thus fixing the variance $\sigma$ gives indeed a one-dimensional exponential family.
+
+In the rest of this thesis, we only consider bounded rewards, and we mostly focus on Bernoulli distributions.
+First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
+% Then in Chapter~\ref{chapter:4}, each static IoT device in the noisy environment is assumed to be following a Bernoulli emission pattern in a fixed channel, FIXME
+%
+For example when we study multi-player bandits in Chapter~\ref{chapter:5} (for the model without sensing) we explain that restricting to the Bernoulli case is interesting as it the harder one, as continuous distributions with a null mass on $r(t)=0$ yield a much simpler problem.
+%
+Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm in the setting on bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use Bernoulli Kullback-Leibler divergences in our analysis.
 
@@
 
-\TODOL{On garde ce petit paragraphe, par curiosité ?}
+\TODOL{On garde ce petit paragraphe sur les modèles de bandits Markovien, par curiosité ?}
 
@@
 MAB have been applied successfully to various decision making problems, including clinical trials and cognitive radio that we mentioned above.
+
+The first application of MAB models was for clinical trials, were an arm represent a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
+The mean of an arm, in this application, represents the mean success rate of a treatment, and the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
+%
+The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication of a device to a gateway (\eg, spreading factor for LoRa, power allocation for NOMA etc).
+%
+% After introducing formally the notations used in all this thesis, we review below possible applications of multi-armed bandits.
 
@@
     \item
-    - hyper-parameters tuning of deep neural networks etc
+    Using bandit algorithms to improve machine learning models or algorithms has been an active research domain for the last ten years or so.
+    As presented below in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, a certain ``master'' bandit algorithm can be used to select on the run the best bandit algorithm from a pool of ``slave'' algorithms.
+    Other possible use cases include hyper-parameter optimization, or feature selection.
+    Hyper-parameters include real-valued parameters, like $\alpha\in[0,\infty)$ for the \UCB{} or $\varepsilon$ for the $\varepsilon$-greedy bandit algorithms, a step size multiplier in a gradient method, or the width of a radial based function (RBF) kernel method.
+    Discrete-valued parameters are also common, like a choice in a fixed set of kernel functions or the depth of neural networks,
+    and higher dimensional or more complex hyper-parameters include the entire architecture of a neural network.
+
+    % FIXME this is copied pasted from \cite{bouneffouf2019survey}
+    It is well known that performance of machine learning algorithms depends critically on identifying a good set of hyper-parameters.
+    While recent approaches use Bayesian optimizationto adaptively select optimal hyper-parameter configurations, they rather focus on speeding up random search
+    through adaptive resource allocation and early-stopping.
+    A different approach was initiated in \cite{LiJamieson2018}, by formulating hyper-parameter optimization as a pure-exploration non-stochastic infinite-armed bandit problem, where a predefined resources, such as iterations, data samples, or features are allocated to randomly sampled configurations.
+This work introduced a novel algorithm, Hyperband, for this framework and analyze its theoretical properties, providing several desirable guarantees.
+    Furthermore, Hyperband was compared with popular Bayesian optimization methods on a suite of hyper-parameter optimization problems; it was observed that Hyperband can provide more than an order-of-magnitude speedup over its competitors on a variety of deep-learning and kernel-based learning problems.
+
 \end{itemize}
@@
 %
-Another interesting works is \cite{Modi17QoS}, FIXME
+Another interesting works is \cite{Modi17QoS}.
+\TODOL{FIXME Finir ce blabla inutile ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
 

### 2019-05-16 2-Chapters/2-Chapter/chapter2.tex
It's hard to write when you know it's so useless.

@@
 We review quickly the possible applications of bandits, that range from recommender systems and information retrieval to healthcare and finance.
-In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits has already been applied with success.
+In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits has already been applied with success in the last ten years.
 
-We then explain the measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and horizon $T$.
+We then explain the main measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and horizon $T$.
 The regret compares the performance of the decision-making player and the oracle who knows in advance the best arm and always selects it.
@@
 We also give an empirical comparison of different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
-
+%
 We also give regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
@@
 %
-Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on \cite{Besson2018WCNC}.
-The notations used in this thesis are the usual notations in all the MAB literature, and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
+Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on the article \cite{Besson2018WCNC}.
 
@@
 Finally, we recommend \cite{bouneffouf2019survey} for a short but good survey on applications of MAB.
+
+The notations used in this thesis are the usual notations in all the MAB literature, and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
+The Nomenclature includes a summary of the notations, see Page~\pageref{chapter:nomenclature}.
 
@@
 Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, following the uniform distribution on $[0,1]$, and that has to be independent from the rewards $r_k(t)$.
-This source of randomness is used in most MAB algorithm, for instance to break ties in a uniformly random manner.
-%  in the decision making rule.
+This source of randomness is used in most MAB algorithms, for instance to break ties uniformly in an index policy.
 
@@
 The MAB model is a famous example of a reinforcement learning model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
-We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
+We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} above.
 For more details on reinforcement learning, in particular about more generic models, we refer to the famous book \cite{SuttonBarto2018}.
@@
 
-\paragraph{Interactive demo for the novice.}
+\paragraph{An interactive demo to discover the MAB problem (for the novice).}
 %
 If you are discovering the concept of bandit model here, I would like to recommend you to go online and play a few times with an interactive demonstration.
-On this demo, you are playing against a MAB problem with $K=5$ arms, when you have $T=100$ decisions to make.
-The demo is hosted online, on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}.
+On this demo, you will be facing a MAB problem with $K=5$ arms, and you have $T=100$ decisions to make.
+The demo is hosted on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}, and illustrated below.
 
@@
 The expectancy is taken on the randomness of the rewards $(r_k(t))_{k,t}$ as well as the randomness $(U_s)_s$.
+%
+Other common objective include best-arm identification (BAI) \cite{audibert2010best}.
 
-\TODOL{A finir, a faire comme dans https://tel.archives-ouvertes.fr/tel-02121614/document page 19}
+% \TODOL{A finir, a faire comme dans https://tel.archives-ouvertes.fr/tel-02121614/document page 19}
 
@@
 
-\TODOL{Write this!}
 In the example above, we consider Bernoulli distributions, but other distributions have been studied in the literature.
@@
 
-\TODOL{On garde ce petit paragraphe sur les modèles de bandits Markovien, par curiosité ?}
+\TODOL{On garde ça sur les modèles de bandits Markovien, par curiosité ?}
 
@@
 Other popular applications include the following:
+%
 \begin{itemize}
     \item
-    A/B testing of websites
-    \cite{Kaufmann14,Jamieson17ABTest}
+    A/B testing of websites is a popular application of the best arm identification problem,
+    where the task is purely exploratory and the player is asked to identify the best of two options (or more), in a finite number of steps \cite{audibert2010best}.
+    The problem can either consider a fixed budget and no freedom on the ending of the game (\ie, an known horizon $T$), or a fixed confidence (\ie, the identified arm must be the true best arm with probability at least $1-\delta$) and a certain freedom on the budget \cite{Garivier16BAI}.
+    The theoretical complexity of this use of bandit for A/B testing was first studied in \cite{Kaufmann14},
+    and later on a more practical point-of-view was proposed in \cite{Jamieson17ABTest}.
 
     \item
-    - online content recommandation
-    \cite{Li10,Louedec16}
+    MAB can also be applied to a broader setting of online content recommandation,
+    with more than two options.
+    The seminal work of \cite{Li10} studies the application of contextual bandit to news article recommendation, as it is used in practice on platforms such as Microsoft's Bing news website,
+    or in applications like Netflix.
+    In such models, the arms correspond to items to recommend (\eg, articles or movies), and the contexts contain features about each user of the system.
+    An interesting work is \cite{Louedec16}, who studies slowly-varying non-stationary models applied to recommender systems.
 
@@
     It is well known that performance of machine learning algorithms depends critically on identifying a good set of hyper-parameters.
-    While recent approaches use Bayesian optimizationto adaptively select optimal hyper-parameter configurations, they rather focus on speeding up random search
-    through adaptive resource allocation and early-stopping.
-    A different approach was initiated in \cite{LiJamieson2018}, by formulating hyper-parameter optimization as a pure-exploration non-stochastic infinite-armed bandit problem, where a predefined resources, such as iterations, data samples, or features are allocated to randomly sampled configurations.
-This work introduced a novel algorithm, Hyperband, for this framework and analyze its theoretical properties, providing several desirable guarantees.
+    While recent approaches use Bayesian optimization to adaptively select optimal hyper-parameter configurations, they rather focus on speeding up random search through adaptive resource allocation and early-stopping.
+    A different approach was initiated in \cite{LiJamieson2018}, by formulating hyper-parameter optimization as a pure-exploration non-stochastic infinite-armed bandit problem, where predefined resources, such as iterations, data samples, or features are allocated to randomly sampled configurations.
+This work introduced a novel algorithm, Hyperband, for this framework and analyzed its theoretical properties, providing several desirable guarantees.
     Furthermore, Hyperband was compared with popular Bayesian optimization methods on a suite of hyper-parameter optimization problems; it was observed that Hyperband can provide more than an order-of-magnitude speedup over its competitors on a variety of deep-learning and kernel-based learning problems.
@@
 
-Other interesting applications are discussed in \cite{bouneffouf2019survey}.
+The three applications detailed above are still active research directions,
+and a curious reader will find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}.
 

### 2019-05-16 2-Chapters/2-Chapter/chapter2.tex
I'm started to be satisfied with the first sections of this chapter

@@
 In the point-of-view of numerical simulations or real-world applications of bandits on embedded systems, one must also take care of two other measures of performance: time and memory complexities, which can vary greatly between different families of algorithms.
-We also give an empirical comparison of different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
+We also empirically compare different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
 %
-We also give regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
+We then give regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
 They both attain a logarithmic regret, proving their order optimality (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
@@
 
-Multi-Armed Bandits models were first introduced by Thompson as early as 1933 in \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
+Multi-Armed Bandits (MAB) models were introduced by Thompson as early as in 1933 \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
 This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
-Their name comes from the one-armed bandit found in casinos, see Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
+Their name refers to one-armed bandits found in casinos, as showed in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
 % The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
@@
 A commonly studied goal for the player is to maximize its sum of received rewards, $\sum_{t=1}^T r_{A(t)}(t)$
-(or its expectancy in stochastic and stationary models).
+(or its mean in stochastic and stationary models).
 
-From a mathematical point-of-view, this quantity $\sum_{t=1}^T r_{A(t)}(t)$ is indeed random, and it depends on two aspects.
-First, the cumulated rewards depend on the randomness from the environment, that is, the unknown and unpredictable values of $r_k(t)$ for all $k$ and $t$.
-Then, this quantity also depends on the decision-making policy followed the player, that is, the choices $A(t)$, that are based on (all) the past observations ($A(1), r_{A(1)}(1), \dots, A(t-1), r_{A(t-1)}(t-1)$), and possibly on an external source of randomness.
-Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, following the uniform distribution on $[0,1]$, and that has to be independent from the rewards $r_k(t)$.
-This source of randomness is used in most MAB algorithms, for instance to break ties uniformly in an index policy.
+The difficulty of this decision-making game comes from balancing the trade-off between \emph{exploration}, because the player has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
+The MAB model is a famous example of a reinforcement learning model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
+We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
+For more details on reinforcement learning, in particular about more generic models, we refer to the famous book \cite{SuttonBarto2018}.
 
@@
 
-The difficulty of this decision-making game comes from balancing the trade-off between \emph{exploration}, because the player has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
-The MAB model is a famous example of a reinforcement learning model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
-We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} above.
-For more details on reinforcement learning, in particular about more generic models, we refer to the famous book \cite{SuttonBarto2018}.
+This quantity $\sum_{t=1}^T r_{A(t)}(t)$ is indeed random, and depends on two aspects.
+First, the rewards depend on the randomness of the environment, \ie, the unknown and unpredictable values $r_k(t)$.
+Then, this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the past observations, \ie, $A(1)$, $r_{A(1)}(1)$, $\dots$, $A(t-1)$, $r_{A(t-1)}(t-1)$, and possibly on an external source of randomness.
+%
+Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, following the uniform distribution on $[0,1]$, and that has to be independent from the rewards $r_k(t)$.
+External randomness is used in most MAB algorithms, for instance to break ties uniformly in an index policy.
 
@@
 %
-An algorithm is said to be deterministic if it does not depend on the external randomness $U_0,U_1,\dots$, but in this thesis we only use non-deterministic strategies.
+An algorithm is said to be deterministic if it does not depend on the external randomness $U_0,U_1,\dots$, but in this thesis we only use non-deterministic algorithms.
 
-The most common objective for the player is to maximize its expected rewards, that is $\E[\sum_{t=1}^T Y_t] = \E[\sum_{t=1}^T r_{A(t)}(t)]$.
-The expectancy is taken on the randomness of the rewards $(r_k(t))_{k,t}$ as well as the randomness $(U_s)_s$.
+The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T Y_t] = \sum_{t=1}^T \E[r_{A(t)}(t)]$.
+The expectancy is taken on the rewards $(r_k(t))_{k,t}$, and the extern random variables $(U_s)_s$.
 %
-Other common objective include best-arm identification (BAI) \cite{audibert2010best}.
-
-% \TODOL{A finir, a faire comme dans https://tel.archives-ouvertes.fr/tel-02121614/document page 19}
+Other common objectives include best-arm identification (BAI) \cite{audibert2010best}.
 
@@
 Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) have the following form:
-\[ \Pr(x | \lambda) = h(x) \exp(\eta^T T(x) - A(\eta)), \]
+\begin{equation}\label{eq:2:exponentialFamily}
+    \Pr(x | \lambda) = h(x) \exp \let( \eta^T T(x) - A(\eta) \right),
+\end{equation}
 for a parameter vector $\eta$ (the canonical parameter), and given functions $T$ (the sufficient statistic) and $h$.
 $A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$, $h$, $T$,
-$A(\eta) = \log \int h(x) \exp(\eta^T T(x)) \lambda(\mathrm{d} x)$.
+$A(\eta) = \log \left( \int h(x) \exp(\eta^T T(x)) \lambda(\mathrm{d} x) \right)$.
 %
@@
 The KL divergence between two distributions $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ is defined as
-\[ \KL(\Pr(x|\eta_1), \Pr(x|\eta_2)) = \int \Pr(x|\eta_1) \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) \lambda(\mathrm{d}x) = \E_{d_1}[ \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) ]. \]
+\begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
+    \KL\left(\Pr(x|\eta_1), \Pr(x|\eta_2)\right) = \int \Pr(x|\eta_1) \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) \lambda(\mathrm{d}x) = \E_{d_1}[ \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) ].
+\end{equation}
 %
-So it can be written $\KL( \eta_1,  \eta_2)) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$ and simplified to
-\[ \KL(\Pr(x|\eta_1), \Pr(x|\eta_2)) = \E_{d_1}[ (\eta_1 - \eta_2)^T T(X) ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2)^T \mu_1 - A(\eta_1) + A(\eta_2), \]
-where $\mu_1 = \E_{\eta_1}[T(X)]$ is the mean of distribution $d_1$.
+So this KL divergence can be written $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, and it can be simplified to an expression using only the two parameters $\eta_1$ and $\eta_2$,
+and $\mu_1 = \E_{\eta_1}[T(X)]$ the mean of distribution $d_1$:
+%
+\begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
+    \KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2)^T T(X) ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2)^T \mu_1 - A(\eta_1) + A(\eta_2),
+\end{equation}
 
-Without diving too much in the details, it is interesting to illustrate these notations for two examples.
+Without diving more in the details of exponential families,
+it is interesting to illustrate this definition and the notations with two important examples:
+%
+\begin{itemize}
+    \item
     On the first hand, Bernoulli distributions can be seen as an exponential family with
@@
     (with the limit behavior $\eta=+\infty$ if $\mu=0$).
-%
+
+    \item
     On the other hand, Gaussian distributions use
@@
     $\eta = [\mu/\sigma^2 ; -1/\sigma^2]^T$, giving $A(\eta) = -\eta_1^2/4\eta_2 -1/2 \log(-2\eta_2)$ (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
-Thus fixing the variance $\sigma$ gives indeed a one-dimensional exponential family.
+    For a generic variance, this form shows that the Gaussian distributions are a two-dimensional exponential families,
+    and fixing the variance $\sigma$ gives indeed a one-dimensional exponential family.
+\end{itemize}
 
-In the rest of this thesis, we only consider bounded rewards, and we mostly focus on Bernoulli distributions.
+In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions.
 First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
@@
 %
-For example when we study multi-player bandits in Chapter~\ref{chapter:5} (for the model without sensing) we explain that restricting to the Bernoulli case is interesting as it the harder one, as continuous distributions with a null mass on $r(t)=0$ yield a much simpler problem.
+For example when we study multi-player bandits in Chapter~\ref{chapter:5}, for the model without sensing, we explain that restricting to the Bernoulli case is interesting and it is actually the hardest one, as continuous distributions with a null mass on $r(t)=0$ yield a much simpler problem.
 %
@@
 
-The first application of MAB models was for clinical trials, were an arm represent a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
+Historically, MAB models were first applied to clinical trials, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
 The mean of an arm, in this application, represents the mean success rate of a treatment, and the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
 %
-The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication of a device to a gateway (\eg, spreading factor for LoRa, power allocation for NOMA etc).
+The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc ).
+In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
 %
@@
 Another interesting works is \cite{Modi17QoS}.
+
 \TODOL{FIXME Finir ce blabla inutile ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
@@
 
+As explained above, the main objective of a player facing a bandit game is to maximize its (expected) cumulated reward.
+In Machine Learning, we usually prefer to aim at minimizing certain quantity, such as error rate in supervised learning, or the distance to the optimum in an optimization problem.
+We introduce in this section the notion of \emph{regret}, and the equivalence for a player between maximizing its sum of rewards and minimizing its regret.
+
 \subsection{Measuring performance with the (mean) regret}
 
-Explain what is regret
+Let us first introduce some notations.
+We consider from now on a stochastic and stationary MAB problem, with $K$ arms of distributions $(\nu_k)_k$, that is, for any time $t$, $r_k(t)\sim \nu_k$.
+We denote $\mu_k$ the mean of the distribution of arm $k$ (it will be referred to as the mean of arm $k$ in the rest of the manuscript), that is $\mu_k = \E[r_k(t)]$.
+%
+% \paragraph{Defining the regret.}
+%
+To define the regret, we first need to distinguish between \emph{optimal} and \emph{sub-optimal} arms.
+
+\begin{defn}[Optimal and sub-optimal arms]\label{def:2:optimalSubOptimalArms}
+    Now consider $\mu^* = \max_k \mu_k$ the mean of the best arm.
+    The best arm can be non unique, and any arm $k$ having $\mu_k = \mu^*$ is said to be \emph{optimal},
+    while arms satisfying $\mu_k < \mu^*$ are called \emph{sub-optimal}.
+\end{defn}
+
+If the goal is to maximize $\E[\sum_{t=1}^T r_{A(t)}(t)]$,
+the optimal strategy for this bandit problem is to always pull an optimal arm, but of course it is unrealistic as the player does not know the true means and cannot know which arm is optimal.
+%
+Comparing the difference between the performance a fixed baseline and that of the player is a common approach in machine learning research,
+and here we can compare with the oracle strategy that obtains an expected reward of $\mu^*$ at each time step.
+%
+For a fixed horizon $T$, if $k^*$ denotes the index of any optimal arm,
+let us introduce the (mean) regret $R_T^{\cA}$ of an algorithm $\cA$ as
+\[ R_T^{\cA} = \E\left[ \sum_{t=1}^T r_{k^*}(t) - r_{A(t)}(t) \right]. \]
+%
+We can rewrite this to obtain the following definition of the regret.
+\begin{align}
+    R_T^{\cA}
+    &= \E\left[ \sum_{t=1}^T r_{k^*}(t) \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right] \\
+    &= T \E\left[ r_{k^*}(1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right] \\
+    &= T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
+\end{align}
+
+\begin{defn}[Regret]\label{def:2:regret}
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized with their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the (mean) regret at horizon $T$ is defined as
+    \begin{equation}
+        R_T^{\cA} = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
+    \end{equation}
+\end{defn}
+
+\TODOL{Est-ce qu'il faut autre chose ?}
+
+\paragraph{A useful decomposition of the regret.}
+%
+As introduced above, let $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ be the number of times arm $k$ was selected until between times $1$ and $t$.
+Remember that the rewards $r_k(t)$ are all \iid{} of mean $\mu_k$.
+Thus one can use the chain rule of expectation, and because the expectation is taken on the randomness of the rewards and of the decisions of the player,
+one can rewrite the expected cumulated rewards like this:
+%
+\begin{align}
+    \E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]
+    &= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right] \nonumber\\
+    &= \sum_{k=1}^K \E [ r_k(1) \underbrace{\sum_{t=1}^T \mathbbm{1}(A(t) = k)}_{= N_k(T)} ] \nonumber\\
+    &= \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].
+\end{align}
+
+If we introduce the \emph{gap} $\Delta_k = \mu^* - \mu_k$ between any arm $k\in[K]$ and an optimal arm,
+then an arm $k$ is sub-optimal if $\Delta_k > 0$, and we can write the following decomposition on the regret,
+
+\begin{lemma}[Regret decomposition]\label{lem:2:RegretDecomposition}
+    The (mean) regret $R_T^{\cA}$ can be written like
+    \begin{equation}
+        R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ].
+    \end{equation}
+\end{lemma}
+
+\begin{proof}\label{proof:2:RegretDecomposition}
+    Thanks to the tower rule, we had
+    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus
+    \[ R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]. \]
+    FIXME more details?
+\end{proof}
+
+\TODOL{Est-ce qu'il faut autre chose ?}
 

### 2019-05-16 2-Chapters/2-Chapter/chapter2.tex
Done for today

@@
     \centering
-    \includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
+    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
     \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.
@@
 We also focus only on real-valued rewards, meaning that $r_k(t)\in\R$ for all $k$ and $t$.
+
 An important hypothesis is whether rewards are bounded or not,
@@
 
-Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) have the following form:
+Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R^d$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) have the following form:
 \begin{equation}\label{eq:2:exponentialFamily}
@@
 
-An important result on exponential families is that a distribution in such family is entirely characterized by its parameter $\eta$.
+-- An important result on exponential families is that a distribution in such family is entirely characterized by its parameter $\eta$.
 That is why such one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[T(X)]$.
 
-A second important result is a simplified form for the Kullback-Leibler divergence \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
+-- A second important result is a simplified form for the Kullback-Leibler divergence \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
 The KL divergence between two distributions $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ is defined as
@@
 A lot of research works considers rewards distributions that are not Gaussian but sub-Gaussian, of a known variance, for instance $1/4$.
-For instance, bounded distributions on $[0,1]$ are $1/4$ sub-Gaussian, and this fact is used for instance in \cite{Maillard2018GLR}.
-It means that their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$).
+For instance, bounded distributions on $[0,1]$ are known to be $1/4$ sub-Gaussian, and this fact is used for instance in \cite{Maillard2018GLR}.
+It means that their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$.
 In Chapter~\ref{chapter:6}, we instead consider sub-Bernoulli distributions, formally introduced in Definition~\ref{def:6:subBernoulliDistributions}.
@@
 
-\paragraph{Application for Opportunistic Spectrum Access.}
+\paragraph{Applications for Opportunistic Spectrum Access.}
 % Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
@@
 
-\TODOL{FIXME Finir ce blabla inutile ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
+\TODOL{FIXME Finir ce blabla rapidement ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
 
 % ----------------------------------------------------------------------------
-\section{Lower and upper bounds on regret}
+\section{Regret: definition, decomposition, and lower bounds}
 \label{sec:2:lowerUpperBoundsRegret}
@@
 \end{lemma}
-
+%
 \begin{proof}\label{proof:2:RegretDecomposition}
@@
     \[ R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]. \]
-    FIXME more details?
+    % FIXME more details?
 \end{proof}
@@
 
+\begin{itemize}
+    \item
     - [Lai and Robbins] lower-bound in $\Omega(\log(T))$
+    \item
     - Worst case lower-bound in $\Omega(\sqrt{T})$
+    \item
     - Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models)
+\end{itemize}
 
@@
 
-- useful decomposition, with number of samples (see Claire's PhD)
+\begin{itemize}
+    \item
     - different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
-- tools to show regret upper-bounds, concentration inequalities,
+    \item
+    - tools to show regret upper-bounds, concentration inequalities?
+    FIXME non j'ai pas la place
+    \item
     - short proof of UCB or klUCB? just to start to show the tools
+    FIXME non j'ai pas la place
+\end{itemize}
 
@@
 - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-- Fairness
+- Fairness ?
+- Switching cost? \cite{modiDemo2016} \cite{Koren17}
 

### 2019-05-16 2-Chapters/2-Chapter/chapter2.tex
Latest version should almost always be online at https://perso.crans.org/besson/articles/.PhD_thesis__Lilian_Besson.pdf

@@
 The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} above.
-The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r_{A(t)(t)$.
-Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r_{A(t)}(t)\in\{0,1}$, that is sampled \iid{} from $\nu_k$.
+The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r_{A(t)}(t)$.
+Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r_{A(t)}(t)\in\{0,1\}$, that is sampled \iid{} from $\nu_k$.
 %
@@
 \begin{equation}\label{eq:2:exponentialFamily}
-    \Pr(x | \lambda) = h(x) \exp \let( \eta^T T(x) - A(\eta) \right),
+    \Pr(x | \lambda) = h(x) \exp \left( \eta^T T(x) - A(\eta) \right),
 \end{equation}
@@
     \begin{equation}
-        R_T^{\cA} = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
+        R_T^{\cA} = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right] = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
     \end{equation}

### 2019-05-20 2-Chapters/2-Chapter/chapter2.tex
Added references about Batch bandit problem (shortly, in Chapter 2, other measures of performance, switching cost).

@@
 \newpage
-\TODOL{L'abstract ici est trop long, mais je l'ai écrit avant de commencer le chapitre, pour avoir une idée de ce que je devais faire ! Je couperai à la fin !}
+% \TODOL{L'abstract ici est trop long, mais je l'ai écrit avant de commencer le chapitre, pour avoir une idée de ce que je devais faire ! Je couperai à la fin !}
 
@@
 %
-We review quickly the possible applications of bandits, that range from recommender systems and information retrieval to healthcare and finance.
+We quickly review the possible applications of bandits, that range from recommender systems and information retrieval to healthcare and finance.
 In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits has already been applied with success in the last ten years.
 
-We then explain the main measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and horizon $T$.
+We then explain the main measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and for horizon $T$.
 The regret compares the performance of the decision-making player and the oracle who knows in advance the best arm and always selects it.
-Other measures of performance, such as best-arm identification, are not used later on in this document, but they are also shortly discussed.
+Other measures of performance, such as best-arm identification, are not used later on in this document.
+% , but they are also shortly discussed.
 %
-We review the main theoretical results from famous previous works, that give a lower-bound on the regret of any algorithm, essentially showing that we cannot find an algorithm achieving a sub-logarithmic regret, in this stochastic MAB model ($R_T^{\cA} = \Omega(\log(T))$).
+We explain the main theoretical result from previous works: a lower-bound on the regret of any algorithm, essentially showing that no algorithm can achieve sub-logarithmic regret on all problems, in this stochastic MAB model (\ie, $R_T^{\cA} = \Omega(\log(T))$).
 
-We give an overview of the main families of algorithms designed to tackle the stochastic MAB model, and we mainly focus on the \UCB, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
-When presenting our library SMPyBandits in next Chapter~\ref{chapter:3}, designed to perform numerical simulations of MAB problem, we use it to compare the empirical performances of the most famous and most efficient algorithms, in terms of regret and by evaluating them on different problems.
+We give an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the \UCB, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
+We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contribution of this thesis.
+We use it to compare the empirical performances of the most famous and most efficient algorithms, in terms of regret and by evaluating them on different problems.
 %
@@
     \centering
-    \resizebox{0.50\textwidth}{!}{
+    \resizebox{0.750\textwidth}{!}{
         \begin{tikzpicture}[auto, node distance=5cm, >=latex]
@@
     \centering
-    \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
+    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
     \caption{Screenshot of the demonstration available online on my website, for a current step of $t=24$.}
@@
     \item
-    On the first hand, Bernoulli distributions can be seen as an exponential family with
+    \textbf{Bernoulli distributions} can be seen as an exponential family with
     $T(x) = x$ and $h(x) = 1$,
@@
     \item
-    On the other hand, Gaussian distributions use
+    \textbf{Gaussian distributions} use
     $T(x) = [x ; x^2]^T$ and $h(x) = 1/\sqrt(2\pi)$,
@@
     \item
-    A/B testing of websites is a popular application of the best arm identification problem,
+    \textbf{A/B testing} of websites is a popular application of the best arm identification problem,
     where the task is purely exploratory and the player is asked to identify the best of two options (or more), in a finite number of steps \cite{audibert2010best}.
@@
     \item
-    MAB can also be applied to a broader setting of online content recommandation,
+    MAB can also be applied to a broader setting of \textbf{online content recommandation},
     with more than two options.
@@
     \item
-    Using bandit algorithms to improve machine learning models or algorithms has been an active research domain for the last ten years or so.
+    Using bandit algorithms for \textbf{improved machine learning} models or algorithms has been an active research domain for the last ten years or so.
     As presented below in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, a certain ``master'' bandit algorithm can be used to select on the run the best bandit algorithm from a pool of ``slave'' algorithms.
@@
 As explained above, the main objective of a player facing a bandit game is to maximize its (expected) cumulated reward.
-In Machine Learning, we usually prefer to aim at minimizing certain quantity, such as error rate in supervised learning, or the distance to the optimum in an optimization problem.
+In Machine Learning, we usually prefer to aim at minimizing certain quantities, such as the error rate in supervised learning, or the distance to the optimum in an optimization problem.
 We introduce in this section the notion of \emph{regret}, and the equivalence for a player between maximizing its sum of rewards and minimizing its regret.
@@
 Let us first introduce some notations.
-We consider from now on a stochastic and stationary MAB problem, with $K$ arms of distributions $(\nu_k)_k$, that is, for any time $t$, $r_k(t)\sim \nu_k$.
-We denote $\mu_k$ the mean of the distribution of arm $k$ (it will be referred to as the mean of arm $k$ in the rest of the manuscript), that is $\mu_k = \E[r_k(t)]$.
+From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $(\nu_k)_k$, that is, for any time $t$, $r_k(t)\sim \nu_k$.
+We denote $\mu_k$ the mean of the distribution of arm $k$ (it will be referred to as the mean of arm $k$ in the rest of the manuscript), that is $\mu_k = \E[r_k(t)] \in \R$.
 %
@@
 \begin{defn}[Optimal and sub-optimal arms]\label{def:2:optimalSubOptimalArms}
-    Now consider $\mu^* = \max_k \mu_k$ the mean of the best arm.
+    Denote $\mu^* = \max_k \mu_k$ the largest mean.
     The best arm can be non unique, and any arm $k$ having $\mu_k = \mu^*$ is said to be \emph{optimal},
@@
 
-If the goal is to maximize $\E[\sum_{t=1}^T r_{A(t)}(t)]$,
+If the goal of the player is to maximize $\E[\sum_{t=1}^T r_{A(t)}(t)]$,
 the optimal strategy for this bandit problem is to always pull an optimal arm, but of course it is unrealistic as the player does not know the true means and cannot know which arm is optimal.
 %
-Comparing the difference between the performance a fixed baseline and that of the player is a common approach in machine learning research,
-and here we can compare with the oracle strategy that obtains an expected reward of $\mu^*$ at each time step.
+Comparing the difference between the performance of a fixed baseline and that of the player is a common approach in machine learning research,
+and here we can compare with the oracle strategy that obtains an (expected) reward of $\mu^*$ at each time step.
 %
@@
 %
-We can rewrite this to obtain the following definition of the regret.
+We can rewrite this expression to obtain the following definition of the regret.
 \begin{align}
@@
 \begin{defn}[Regret]\label{def:2:regret}
-    For an algorithm $\cA$, a bandit problem of $K$ arms characterized with their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the (mean) regret at horizon $T$ is defined as
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the (mean) regret at horizon $T$ is defined as
     \begin{equation}
@@
 %
-As introduced above, let $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ be the number of times arm $k$ was selected until between times $1$ and $t$.
+As introduced above, let $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ be the number of times arm $k$ was selected between times $1$ and $t$.
 Remember that the rewards $r_k(t)$ are all \iid{} of mean $\mu_k$.
-Thus one can use the chain rule of expectation, and because the expectation is taken on the randomness of the rewards and of the decisions of the player,
+Thus one can use the chain rule of expectation, and because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$ (which are measurable wrt to the past observations $\cO_t$),
 one can rewrite the expected cumulated rewards like this:
@@
     \item
-    - [Lai and Robbins] lower-bound in $\Omega(\log(T))$
+    - [Lai and Robbins] lower-bound in $\Omega(\log(T))$ \cite{LaiRobbins85}
     \item
-    - Worst case lower-bound in $\Omega(\sqrt{T})$
+    - Worst case lower-bound in $\Omega(\sqrt{T})$ \cite{Auer02,Auer02NonStochastic,Bubeck12}
     \item
-    - Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models)
+    - Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models) \cite{Auer02NonStochastic}
 \end{itemize}
@@
 
-- Best Arm Identification?
-- Quantile regret or just histogram of regrets
-- Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-- Fairness ?
-- Switching cost? \cite{modiDemo2016} \cite{Koren17}
+In this thesis, we only study analytically the mean regret $R^{\cA}$ (in Chapters \ref{chapter:5,chapter:6}), but we quickly mention other measures of performances introduced in the literature.
+
+% - Quantile regret or just histogram of regrets
+First, when doing numerical experiments about bandits, if one studies the regret as an empirical mean based on a large number of random repetitions (\eg, $N=1000$ repetitions), it is important to not only show the mean value but also the variance of the values taken by the regret on each repetition, to verify that all algorithms perform consistently.
+Indeed, by only visualizing the mean of $1000$ values, it is possible that we miss some ``bad runs'': if $1$ run out of the $1000$ gives linear regret (\ie, $R_T^{\cA} \propto T$) and the $999$ other give logarithmic regret, then the mean will appear logarithmic.
+This is the case of the \Selfish{} algorithm that is defined and explored in Chapter~\ref{chapter:5}.
+By visualizing the entire \emph{distribution} (as an histogram), or the variance of the values of $R_T^{\cA}$, and if the number $N$ is reasonably large, we can verify that the regret appears logarithmic for all runs.
+
+% - Best Arm Identification?
+Other measures of performance that has been studied in the literature include
+the best arm identification (BAI) rate
+and the best arm selection (BAS) rate.
+The BAI rate counts the frequency at which an algorithm $\cA$ correctly identified the best arm(s) at the horizon $T$,
+while the BAS rate counts the total number of time steps during which $\cA$ selected (one of) the best arm(s).
+Both quantities are computed and stored in all numerical simulations using SMPyBandits, but visualizations for both have not been included in this thesis.
+
+% - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
+The switching cost $SC^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $SC^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
+It has recently gained interest in the literature, for instance in \cite{Koren17}.
+% FIXME give the main result proved in this paper?
+In single-player models,
+it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
+But even if $SC^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
+%
+% FIXME we give notations but I don't use them in Chapter5 
+It is more interesting to studying $SC_M^{\cA}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game, as in this case, it is possible that the players follow an efficient centralized or decentralized algorithm (which gives a logarithmic centralized regret) while still suffering from a linear switching cost.
+It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $SC_M^{\cA}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
+% - Switching cost? \cite{modiDemo2016} \cite{Koren17}
+On a more experimental note, it was showed in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-player bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting means that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
+For more details, see \cite{modiDemo2016} for the multi-player case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi}.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
+
+% - Fairness ?
+Finally, another interesting measure of performance is the \emph{fairness} between the $M$ players in a multi-player bandit models, as we explain in Chapter~\ref{chapter:5}.
 

### 2019-05-21 2-Chapters/2-Chapter/chapter2.tex
Starting to quote & rewrite the theorems for the lower bounds.

@@
 
-\subsection{Regret lower bounds: do not detail too much, don't explain the tools behind the results}
+\subsection{Regret lower bounds}
+
+We remind in this section two results about bandit algorithms \emph{cannot} do:
+first a worst case result that states that any algorithm can perform as badly as $\Omega(\sqrt(K T))$ on a certain problem instance (a result from \cite{Auer02NonStochastic}),
+then a problem-dependent lower-bound that states that any algorithm suffers a regret at least $\Omega(\log T)$ on any problem (from \cite{LaiRobbins85}).
+In a certain family of problems, \eg, $K$ Bernoulli distributed arms, the difficulty of a problem is characterized by a measure that depends only on the arms means.
+This measure of difficulty of a problem is hidden in the $\Omega$ notation of the lower-bound.
+For more details, see Chapter~2 in \cite{Slivkins2019}.
+In all this section, we restrict to stationary stochastic bandits with $K\geq2$ arms.
+
+\paragraph{Worst-case lower-bound in $\sqrt{T}$}
+
+The following theorem was proven in \cite{Auer02NonStochastic}.
+
+\begin{theorem}\label{thm:2:worstCaseLowerBound}
+    For 
+\end{theorem}
+
+\paragraph{Problem-dependent lower-bound in $\log T$}
+
+The following two theorems were proven in \cite{LaiRobbins85}.
+
+First, let $\cI$ denote the set of all problem instances, with hypotheses that there are $K \geq 2$ arms whose rewards lie in $[0,1]$.
+
+\begin{theorem}\label{thm:2:firstLogTLowerBound}
+    No algorithm $\cA$ can achieve a regret $R^{\cA}(T) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
+    where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
+\end{theorem}
+
+Then we consider algorithms that uniformly efficient, meaning that we rule out algorithms that can achieve low regret on some problem instance while achieving linear regret on other instances.
+In particular, it is necessary to rule out algorithms that always pick the same arms, as on some problem instance such fixed-arm algorithm can achieve zero regret.
+For instance, the \UCB{} algorithm is uniformly efficient.
+
+\begin{defn}[Uniformly efficient algorithm]\label{def:2:uniformlyEfficientAlgorithm}
+    An algorithm $\cA$ is \emph{uniformly efficient} if it satisfies
+    \begin{equation}
+        R^{\cA}(T) \leq \bigO{C_{I,\alpha} T^{\alpha}}
+    \end{equation}
+    for any value $\alpha>0$ and any problem instance $I\in \cI$,
+    where this notation means that the ``constant'' $C_{I,\alpha}$ can depend on the problem instance $I$ and on $\alpha$, but not on the time horizon $T$.
+\end{defn}
+
+Now we can state the second logarithmic lower-bound, for algorithms in this family:
+
+\begin{theorem}\label{thm:2:secondLogTLowerBound}
+    Let $\cA$ denote a uniformly efficient algorithm,
+    and $I$ an arbitrary problem instance.
+    For this instance,
+    there exists a constant $C_I$ depending only on the problem $I$,
+    and a time $T_0$ such that
+    \begin{equation}
+        \forall T \geq T_0,
+        \;\;\;
+        R^{\cA}(T) \geq C_I \log(T).
+    \end{equation}
+\end{theorem}
+
+% do not detail too much, don't explain the tools behind the result
 

### 2019-05-21 2-Chapters/2-Chapter/chapter2.tex
Done for tonight

@@
 % First chapter begins here
-\chapter{The Single Player and Stationary Multi-Armed Bandits Model}
+\chapter{The Stochastic Multi-Armed Bandits Model}
 \label{chapter:2}
@@
 We present one of the first contribution made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms.
-We illustrate its performance empirically, but no satisfactory theoretical result is given, as the bound we had hoped for have been proved unachievable since then.
+We illustrate its performance empirically, but no satisfactory theoretical result is given, as the bound we had hoped for has been proved unachievable since then.
 
@@
 % ----------------------------------------------------------------------------
-\section{Stationary MAB models}
+\section{The stochastic Multi-Armed Bandit model}
 \label{sec:2:notations}
@@
     \centering
-    \resizebox{0.750\textwidth}{!}{
+    \resizebox{0.350\textwidth}{!}{
         \begin{tikzpicture}[auto, node distance=5cm, >=latex]
@@
 
-This quantity $\sum_{t=1}^T r_{A(t)}(t)$ is indeed random, and depends on two aspects.
-First, the rewards depend on the randomness of the environment, \ie, the unknown and unpredictable values $r_k(t)$.
-Then, this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the past observations, \ie, $A(1)$, $r_{A(1)}(1)$, $\dots$, $A(t-1)$, $r_{A(t-1)}(t-1)$, and possibly on an external source of randomness.
+This quantity $\sum_{t=1}^T r_{A(t)}(t)$ can indeed be random, and usually depends on two aspects.
+First, the rewards may depend on the randomness of the environment, \ie, the unknown and unpredictable values $r_k(t)$.
+Then this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the past observations, \ie, $A(1)$, $r_{A(1)}(1)$, $\dots$, $A(t-1)$, $r_{A(t-1)}(t-1)$, and possibly on an external source of randomness.
 %
-Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, following the uniform distribution on $[0,1]$, and that has to be independent from the rewards $r_k(t)$.
-External randomness is used in most MAB algorithms, for instance to break ties uniformly in an index policy.
+Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, uniformly distributed on $[0,1]$, and that has to be independent from the rewards $r_k(t)$.
+External randomness is used in most MAB algorithms, \eg, to break ties uniformly in an index policy.
 
@@
 %
-Restricting to models with a finite number of arms
-rules out an entire domain of the recent research on MAB, in particular we do not consider linear nor contextual bandits (for more details, see Parts V and VI of \cite{LattimoreBanditAlgorithmsBook}).
+% Restricting to models with a finite number of arms
+% rules out an entire domain of the recent research on MAB, in particular we do not consider linear nor contextual bandits (for more details, see Parts V and VI of \cite{LattimoreBanditAlgorithmsBook}).
 %
@@
 Rewards are stationary, meaning that $(r_k(t))_{t\in\N^*}$ is independent and identically distributed (\iid), $r_k(t) \sim \nu_k$ for any $t\geq1$.
+
+The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T Y_t] = \sum_{t=1}^T \E[r_{A(t)}(t)]$.
+The expectancy is taken on the rewards $(r_k(t))_{k,t}$, and the extern random variables $(U_s)_s$.
+%
+Other common objectives include best-arm identification (BAI) \cite{audibert2010best}.
 
@@
 
-\paragraph{A bandit policy is a measurable function.}
+\paragraph{Decisions can use the past observations.}
 
 A bandit algorithm $\cA$, also referred to as a strategy or a policy, and sometimes denoted by $\pi$ or $\rho$, selects an arm $A(t)$ at time $t$ by using the past observations and the past external randomness.
-More formally, an algorithm is thus a sequence of measurable functions $(\cA_t)_{t\geq1}$,
+More formally, one could state that an algorithm is a sequence of measurable functions $(\cA_t)_{t\geq1}$,
 where $\cA_t$ maps the past observations $\cO_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$
 to an arm $\cA_t(\cO_t) = A(t) \in[K]$,
-if we denote $Y_s = r_{A(s)}(s)$ she $s$-th reward.
+if we denote $Y_s = r_{A(s)}(s)$ the $s$-th reward.
 The initial information is reduced to $\cO_1 = (U_0)$, and the first decision is $A(1) = \cA_1(\cO_1)$.
@@
 
-The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T Y_t] = \sum_{t=1}^T \E[r_{A(t)}(t)]$.
-The expectancy is taken on the rewards $(r_k(t))_{k,t}$, and the extern random variables $(U_s)_s$.
-%
-Other common objectives include best-arm identification (BAI) \cite{audibert2010best}.
-
 % ----------------------------------------------------------------------------
-\subsection{Different hypotheses on rewards distributions}
+\subsection{Hypotheses on rewards distributions}
 
@@
 One-dimensional exponential families include Bernoulli and Poisson distributions, as well as Gaussian distributions with a fixed variance.
-Our main interest is Bernoulli distributions, but we prefer to present here the notations on a generic exponential family.
-We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Emilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
+Our main interest is Bernoulli distributions, but we prefer to present here the notations on a generic exponential family in one dimension.
+% We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Emilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
 
-Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R^d$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) have the following form:
-\begin{equation}\label{eq:2:exponentialFamily}
-    \Pr(x | \lambda) = h(x) \exp \left( \eta^T T(x) - A(\eta) \right),
-\end{equation}
-for a parameter vector $\eta$ (the canonical parameter), and given functions $T$ (the sufficient statistic) and $h$.
-$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$, $h$, $T$,
-$A(\eta) = \log \left( \int h(x) \exp(\eta^T T(x)) \lambda(\mathrm{d} x) \right)$.
+Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) can be written as
+% \begin{equation}\label{eq:2:exponentialFamily}
+$ \Pr(x | \lambda) = h(x) \exp \left( \eta x - A(\eta) \right)$,
+% \end{equation}
+for a parameter vector $\eta$ (the canonical parameter), and a sufficient statistic that we restrict to be $T(x)=x$, and a function $h$.
+$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$ and $h$:
+$A(\eta) = \log \left( \int h(x) \exp(\eta x) \lambda(\mathrm{d} x) \right)$.
 %
@@
 
--- An important result on exponential families is that a distribution in such family is entirely characterized by its parameter $\eta$.
-That is why such one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[T(X)]$.
+-- An important result on exponential families is that \emph{a distribution in such family is entirely characterized by its parameter $\eta$}.
+That is why such one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[X]$.
 
--- A second important result is a simplified form for the Kullback-Leibler divergence \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
+-- A second important result is a simplified form for the \emph{Kullback-Leibler divergence} \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
 The KL divergence between two distributions $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ is defined as
 \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
-    \KL\left(\Pr(x|\eta_1), \Pr(x|\eta_2)\right) = \int \Pr(x|\eta_1) \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) \lambda(\mathrm{d}x) = \E_{d_1}[ \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) ].
+    \KL\left(\Pr(x|\eta_1), \Pr(x|\eta_2)\right) = \int \Pr(x|\eta_1) \log\left(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}\right) \right].
 \end{equation}
@@
 So this KL divergence can be written $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, and it can be simplified to an expression using only the two parameters $\eta_1$ and $\eta_2$,
-and $\mu_1 = \E_{\eta_1}[T(X)]$ the mean of distribution $d_1$:
+and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$:
 %
 \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
-    \KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2)^T T(X) ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2)^T \mu_1 - A(\eta_1) + A(\eta_2),
+    \KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2) X ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2) \mu_1 - A(\eta_1) + A(\eta_2),
 \end{equation}
@@
     \item
-    \textbf{Bernoulli distributions} can be seen as an exponential family with
-    $T(x) = x$ and $h(x) = 1$,
+    \textbf{Bernoulli distributions} can be seen as an exponential family with $h(x) = 1$,
     and for a Bernoulli distribution of mean $\mu\in[0,1]$,
@@
     \item
-    \textbf{Gaussian distributions} use
-    $T(x) = [x ; x^2]^T$ and $h(x) = 1/\sqrt(2\pi)$,
-    and for $\cN(\mu,\sigma^2)$ a Gaussian distribution with mean $\mu\in\R$ and variance $\sigma^2$,
-    $\eta = [\mu/\sigma^2 ; -1/\sigma^2]^T$, giving $A(\eta) = -\eta_1^2/4\eta_2 -1/2 \log(-2\eta_2)$ (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
-    For a generic variance, this form shows that the Gaussian distributions are a two-dimensional exponential families,
-    and fixing the variance $\sigma$ gives indeed a one-dimensional exponential family.
+    \textbf{Gaussian distributions} of a fixed variance are a one-dimensional exponential family,
+    and in this thesis we do not consider Gaussian with an unknown variance.
+    For a variance of $\sigma^2$, the family uses
+    % $T(x) = [x ; x^2]^T$ and
+    $h(x) = 1/\sqrt{2\pi\sigma^2} \exp(-x^2/(2\sigma^2))$,
+    and for a Gaussian distribution with mean $\mu\in\R$ (\ie, $\cN(\mu,\sigma^2)$),
+    the parameter is $\eta = \mu/\sigma^2$, giving $A(\eta) = \mu^2/(2\sigma^2)$
+    (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
+    % % For a generic variance,
+    % This shows that the Gaussian distributions
+    % % are a two-dimensional exponential families,
+    % with a fixed variance $\sigma^2$ form indeed a one-dimensional exponential family.
 \end{itemize}
@@
 They were introduced in the 1980s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
-A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and such models come in two flavors: rested or restless.
+A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and thus they are no longer stochastic nor stationary.
+Such Markov models come in two flavors: rested or restless.
 For $K$ arms, each Markov chain has a finite number of state $s$, each corresponding to a (constant) reward that the player obtains if she selects this arm while its Markov chain is in state $s$.
@@
 
-\TODOL{On garde ça sur les modèles de bandits Markovien, par curiosité ?}
-
 % ----------------------------------------------------------------------------
-\section{Applications of stationary MAB}
-\label{sec:2:applicationsofStationaryMAB}
+\section{Applications of stochastic MAB}
+\label{sec:2:applicationsofStochasticMAB}
 
 The blooming success of the research on multi-armed bandits is easily explained by the different spectrum of applications of MAB models to real-world discrete-time decision making problems.
-This research field has been very active since the years 2010s, but it started as early as 1933 with \cite{Thompson33}, and was active since the 1980s and the seminal works by Lai and Robbins \cite{LaiRobbins85} and by Anantharam and others \cite{Anantharam87a,Anantharam87b}.
+This research field has been very active since the years 2010s, but it started as early as 1933 with \cite{Thompson33}, and was active since the 1980s and the seminal works by Lai and Robbins \cite{LaiRobbins85} and by Anantharam and others \cite{Anantharam87a}.
 %
-MAB have been applied successfully to various decision making problems, including clinical trials and cognitive radio that we mentioned above.
+MAB have been applied successfully to various decision making problems, like the following:
 
-Historically, MAB models were first applied to clinical trials, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
+Historically, MAB models were first applied to \textbf{clinical trials}, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
 The mean of an arm, in this application, represents the mean success rate of a treatment, and the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
-%
-The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc ).
-In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
 %
@@
 
-Other popular applications include the following:
+% Other popular applications include the following.
 %
-\begin{itemize}
-    \item
+% \begin{itemize}
+%     \item
     \textbf{A/B testing} of websites is a popular application of the best arm identification problem,
@@
 
-    \item
+    % \item
     MAB can also be applied to a broader setting of \textbf{online content recommandation},
@@
 
-    \item
+    % \item
     Using bandit algorithms for \textbf{improved machine learning} models or algorithms has been an active research domain for the last ten years or so.
@@
 
-    % FIXME this is copied pasted from \cite{bouneffouf2019survey}
-    It is well known that performance of machine learning algorithms depends critically on identifying a good set of hyper-parameters.
-    While recent approaches use Bayesian optimization to adaptively select optimal hyper-parameter configurations, they rather focus on speeding up random search through adaptive resource allocation and early-stopping.
-    A different approach was initiated in \cite{LiJamieson2018}, by formulating hyper-parameter optimization as a pure-exploration non-stochastic infinite-armed bandit problem, where predefined resources, such as iterations, data samples, or features are allocated to randomly sampled configurations.
+    % % FIXME this is copied pasted from \cite{bouneffouf2019survey}
+    % It is well known that performance of machine learning algorithms depends critically on identifying a good set of hyper-parameters.
+    % While recent approaches use Bayesian optimization to adaptively select optimal hyper-parameter configurations, they rather focus on speeding up random search through adaptive resource allocation and early-stopping.
+    % A different approach was initiated in \cite{LiJamieson2018}, by formulating hyper-parameter optimization as a pure-exploration non-stochastic infinite-armed bandit problem, where predefined resources, such as iterations, data samples, or features are allocated to randomly sampled configurations.
 This work introduced a novel algorithm, Hyperband, for this framework and analyzed its theoretical properties, providing several desirable guarantees.
-    Furthermore, Hyperband was compared with popular Bayesian optimization methods on a suite of hyper-parameter optimization problems; it was observed that Hyperband can provide more than an order-of-magnitude speedup over its competitors on a variety of deep-learning and kernel-based learning problems.
-
-\end{itemize}
+    % Furthermore, Hyperband was compared with popular Bayesian optimization methods on a suite of hyper-parameter optimization problems; it was observed that Hyperband can provide more than an order-of-magnitude speedup over its competitors on a variety of deep-learning and kernel-based learning problems.
+% \end{itemize}
 
@@
 
-The three applications detailed above are still active research directions,
+The different applications detailed above are still active research directions,
 and a curious reader will find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}.
@@
 % Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
-%
+
+The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc ).
+In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
+
 Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
@@
 % ----------------------------------------------------------------------------
-\section{Regret: definition, decomposition, and lower bounds}
+\section{Definition, decomposition and lower-bounds on the regret}
+% : definition, decomposition, and lower bounds
 \label{sec:2:lowerUpperBoundsRegret}
@@
 As explained above, the main objective of a player facing a bandit game is to maximize its (expected) cumulated reward.
-In Machine Learning, we usually prefer to aim at minimizing certain quantities, such as the error rate in supervised learning, or the distance to the optimum in an optimization problem.
+An efficient algorithm should obtain a mean reward converging to the maximum reward.
 We introduce in this section the notion of \emph{regret}, and the equivalence for a player between maximizing its sum of rewards and minimizing its regret.
+In Machine Learning, we usually prefer to aim at minimizing certain quantities, such as the error rate in supervised learning, or the distance to the optimum in an optimization problem,
+but the main interest of studying the regret is to be able to obtain higher-order information about the convergence of the mean reward to the max reward,
+or in other words, to quantify the speed of convergence of a MAB algorithm.
 
@@
 \begin{defn}[Optimal and sub-optimal arms]\label{def:2:optimalSubOptimalArms}
-    Denote $\mu^* = \max_k \mu_k$ the largest mean.
+    Denote $\mu^* = \max_k \mu_k$ the largest mean.\\
     The best arm can be non unique, and any arm $k$ having $\mu_k = \mu^*$ is said to be \emph{optimal},
@@
 %
-We can rewrite this expression to obtain the following definition of the regret.
-\begin{align}
-    R_T^{\cA}
-    &= \E\left[ \sum_{t=1}^T r_{k^*}(t) \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right] \\
-    &= T \E\left[ r_{k^*}(1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right] \\
-    &= T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
-\end{align}
+We can rewrite this expression to obtain the following definition of the regret:
+$R_T^{\cA}$
+$= \E\left[ \sum_{t=1}^T r_{k^*}(t) \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right]$
+$= T \E\left[ r_{k^*}(1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]$
+$= T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]$.
+% This is the definition we give below.
 
@@
 
-\TODOL{Est-ce qu'il faut autre chose ?}
-
 \paragraph{A useful decomposition of the regret.}
 %
-As introduced above, let $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ be the number of times arm $k$ was selected between times $1$ and $t$.
+As introduced above when we described the UI of our online example demo, let $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ be the number of times arm $k$ was selected between times $1$ and $t$.
 Remember that the rewards $r_k(t)$ are all \iid{} of mean $\mu_k$.
 Thus one can use the chain rule of expectation, and because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$ (which are measurable wrt to the past observations $\cO_t$),
-one can rewrite the expected cumulated rewards like this:
-%
-\begin{align}
-    \E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]
-    &= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right] \nonumber\\
-    &= \sum_{k=1}^K \E [ r_k(1) \underbrace{\sum_{t=1}^T \mathbbm{1}(A(t) = k)}_{= N_k(T)} ] \nonumber\\
-    &= \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].
-\end{align}
+one can rewrite the expected cumulated rewards like this,
+$\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]$
+$= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right]$
+$= \sum_{k=1}^K \E [ r_k(1) \sum_{t=1}^T \mathbbm{1}(A(t) = k) ]$
+% \underbrace{XXX}_{= N_k(T)}
+$= \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].$
 
@@
     Thanks to the tower rule, we had
-    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus
-    \[ R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]. \]
-    % FIXME more details?
+    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ].$
 \end{proof}
-
-\TODOL{Est-ce qu'il faut autre chose ?}
 
@@
 
-We remind in this section two results about bandit algorithms \emph{cannot} do:
-first a worst case result that states that any algorithm can perform as badly as $\Omega(\sqrt(K T))$ on a certain problem instance (a result from \cite{Auer02NonStochastic}),
-then a problem-dependent lower-bound that states that any algorithm suffers a regret at least $\Omega(\log T)$ on any problem (from \cite{LaiRobbins85}).
+We include in this section two well-known results about what bandit algorithms \emph{cannot} do.
+First, a \emph{problem-dependent lower-bound} states that any algorithm suffers a regret at least $\Omega(\log T)$ on any problem (from \cite{LaiRobbins85}).
+Then, a \emph{worst-case result} states that any algorithm can perform as badly as $\Omega(\sqrt{K T})$ on a certain problem instance, designed specifically to make it perform badly (a result from \cite{Auer02NonStochastic}),
+%
 In a certain family of problems, \eg, $K$ Bernoulli distributed arms, the difficulty of a problem is characterized by a measure that depends only on the arms means.
 This measure of difficulty of a problem is hidden in the $\Omega$ notation of the lower-bound.
-For more details, see Chapter~2 in \cite{Slivkins2019}.
+
 In all this section, we restrict to stationary stochastic bandits with $K\geq2$ arms.
+We do not give proofs of the following theorems, as they can be found in the historical papers, and simpler proofs are given in recent references, such as \cite{Bubeck12}, \cite{LattimoreBanditAlgorithmsBook}, or Chapter~2 in \cite{Slivkins2019} for instance.
 
-\paragraph{Worst-case lower-bound in $\sqrt{T}$}
-
-The following theorem was proven in \cite{Auer02NonStochastic}.
-
-\begin{theorem}\label{thm:2:worstCaseLowerBound}
-    For 
-\end{theorem}
+Let $\cI$ denote the set of all problem instances, with $K \geq 2$ arms whose rewards lie in $[0,1]$ (and no other hypothesis).
+To specify the dependency on an instance $I$, we denote the regret of $\cA$ on instance $I$ and horizon $T$ by $R_T^{\cA}(I)$.
 
@@
 The following two theorems were proven in \cite{LaiRobbins85}.
+These lower-bounds are of highest interest to design efficient algorithms,
+and a significant part of the research literature on MAB algorithms has focussed on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
+This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically match the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is optimal), or with a larger constant (the algorithm is said order-optimal in such case).
 
-First, let $\cI$ denote the set of all problem instances, with hypotheses that there are $K \geq 2$ arms whose rewards lie in $[0,1]$.
+Moreover, this lower-bound can be directly used to design efficient algorithms, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression in \eqref{eq:2:forSecondLogTLowerBound2} essentially says that any efficient algorithm
+should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times.
+Tracking this quantity is used for instance in the OSSB algorithm proposed in \cite{Combes17}.
+It states that FIXME interpret!
+FIXME explain why this is interesting for the rest of this manuscript!
 
 \begin{theorem}\label{thm:2:firstLogTLowerBound}
-    No algorithm $\cA$ can achieve a regret $R^{\cA}(T) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
+    No algorithm $\cA$ can achieve a regret $R_T^{\cA}(I) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
     where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
@@
     \begin{equation}
-        R^{\cA}(T) \leq \bigO{C_{I,\alpha} T^{\alpha}}
+        R_T^{\cA}(I) \leq \bigO{C_{I,\alpha} T^{\alpha}}
     \end{equation}
@@
     \begin{equation}
-        \forall T \geq T_0,
-        \;\;\;
-        R^{\cA}(T) \geq C_I \log(T).
+        \forall T \geq T_0, \;\;\; R_T^{\cA}(I) \geq C_I \log(T).
     \end{equation}
@@
 
-% do not detail too much, don't explain the tools behind the result
+This third theorem specifies possible values of the constants $C_I$ for the two previous results.
 
+\begin{theorem}\label{thm:2:forSecondLogTLowerBound}
+    Let $\cA$ denote a uniformly efficient algorithm,
+    and $I$ an arbitrary problem instance.
     \begin{itemize}
         \item
-    - [Lai and Robbins] lower-bound in $\Omega(\log(T))$ \cite{LaiRobbins85}
+        The bound from Theorem~\ref{thm:2:secondLogTLowerBound} holds with,
+        \begin{equation}\label{eq:2:forSecondLogTLowerBound}
+            C_I = \mu^* (1 - \mu^*) \sum_{k: \Delta_k > 0} \frac{1}{\Delta_k}.
+        \end{equation}
         \item
-    - Worst case lower-bound in $\Omega(\sqrt{T})$ \cite{Auer02,Auer02NonStochastic,Bubeck12}
-    \item
-    - Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models) \cite{Auer02NonStochastic}
+        For any $\varepsilon>0$, the bound from Theorem~\ref{thm:2:secondLogTLowerBound} holds with,
+        \begin{equation}\label{eq:2:forSecondLogTLowerBound2}
+            C_I = \sum_{k: \Delta_k > 0} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon.
+        \end{equation}
     \end{itemize}
+\end{theorem}
+
+\paragraph{Worst-case lower-bound in $\sqrt{T}$}
+
+For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses the constants $C_I$ no longer bring any information.
+We naively have that $R_T^{\cA} \leq (\max_k \Delta_k) T$, and thus if
+
+For this reason, it is interesting to quote another lower-bound on the regret, that also
+The following theorem was proven in \cite{Auer02NonStochastic}.
+
+It states that FIXME interpret!
+FIXME explain why this is interesting for the rest of this manuscript!
+
+\begin{theorem}\label{thm:2:worstCaseLowerBound}
+    Fix the number of arms $K$, and the algorithm $\cA$.
+    Then for any horizon $T$, there exists a problem instance $I_T$ on which the algorithm suffers a regret $R_T^{\cA}(I_T)$ that verifies
+    \begin{equation}
+        R_T^{\cA}(I_T) \geq \Omega(\sqrt{K T}).
+    \end{equation}
+\end{theorem}
+
+% % do not detail too much, don't explain the tools behind the result
+
+% \begin{itemize}
+%     \item
+%     - [Lai and Robbins] lower-bound in $\Omega(\log(T))$ \cite{LaiRobbins85}
+%     \item
+%     - Worst case lower-bound in $\Omega(\sqrt{T})$ \cite{Auer02,Auer02NonStochastic,Bubeck12}
+%     \item
+%     - Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models) \cite{Auer02NonStochastic}
+% \end{itemize}
 
@@
 
-In this thesis, we only study analytically the mean regret $R^{\cA}$ (in Chapters \ref{chapter:5,chapter:6}), but we quickly mention other measures of performances introduced in the literature.
+In this thesis, we only study analytically the mean regret $R^{\cA}$ (in Chapters \ref{chapter:5,chapter:6}), but we quickly mention other measures of performances used in our work.
 
@@
 
-% - Best Arm Identification?
-Other measures of performance that has been studied in the literature include
-the best arm identification (BAI) rate
-and the best arm selection (BAS) rate.
-The BAI rate counts the frequency at which an algorithm $\cA$ correctly identified the best arm(s) at the horizon $T$,
-while the BAS rate counts the total number of time steps during which $\cA$ selected (one of) the best arm(s).
-Both quantities are computed and stored in all numerical simulations using SMPyBandits, but visualizations for both have not been included in this thesis.
+% DONE on s'en fout !
+% % - Best Arm Identification?
+% Other measures of performance that has been studied in the literature include
+% the best arm identification (BAI) rate
+% and the best arm selection (BAS) rate.
+% The BAI rate counts the frequency at which an algorithm $\cA$ correctly identified the best arm(s) at the horizon $T$,
+% while the BAS rate counts the total number of time steps during which $\cA$ selected (one of) the best arm(s).
+% Both quantities are computed and stored in all numerical simulations using SMPyBandits, but visualizations for both have not been included in this thesis.
 

### 2019-05-22 2-Chapters/3-Chapter/chapter3.tex
So long
Small proofreading of notations of Chapter6

@@
 Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
-We refer to Chapter~\ref{chapter:5} which introduces and details various multi-player bandit models.
+We refer to Chapter~\ref{chapter:5} which introduces the multi-player bandit model.
 
@@
 Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
-Its documentation has seen already about $20000$ visits, the project on GitHub had $76$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
+Its documentation has seen already about $20000$ visits, the project on GitHub had $85$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.

### 2019-05-22 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Small proofreading of notations of Chapter6

@@
 
-\TODOL{Okay ça suffit pour ces travaux ?}
+% \TODOL{Okay ça suffit pour ces travaux ?}
 
@@
 
-Assume that we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$.
+Assume that we collect independent samples $X_1,X_2,\ldots$ all from some distributions supported in $[0,1]$.
 We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
@@
 Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
-Our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm know before hand the values of $a,b$).
+As explained in in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows before hand the values of $a$ and $b$).
 
@@
         &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
-        &= n \left[ \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right].
+        &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
     \end{align*}
@@
         &= \sup\limits_{\mu_0,\mu_1,\tau < n} \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] }, \\
-        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr] \Bigr],  \\
-        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1, 1 < s < n} \Bigl[
+        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr] \Bigr],
+    \end{align*}
+    When the last equation comes from the fact that $\exp$ is increasing.
+    Thus by taking the logarithm of both sides, we obtain
+    \begin{align*}
+        \log \GLR(n)
+        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr],  \\
+        &= \sup\limits_{\mu_0,\mu_1, 1 < s < n} \Bigl[
             s \bigl( \widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0) \bigr) \\
@@
             & \hspace{35pt} - n \bigl( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \bigr)
-        \Bigr] \Bigr].
+        \Bigr].
     \end{align*}
@@
     \begin{align*}
-        \GLR(n)
-        = \exp \Bigl[ \sup\limits_{1 < s < n} \Bigl[
+        \log \GLR(n)
+        = \sup\limits_{1 < s < n} \Bigl[
             s \times \Bigl(
@@
             \Bigr)
-        \Bigr] \Bigr].
+        \Bigr].
     \end{align*}
@@
 \begin{proof}
-    It is actually not used as such in the proofs of the regret upper bound given below, but a similar result is used and proved in the proofs.
-    \TODOL{We have to give the proof, or remove this result?
+    It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proved in the proofs.
+    For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} in page~\pageref{par:6:controllingDelayInOneProof}.
+    % \TODOL{We have to give the proof, or remove this result?
 Je ne vois pas le problème de le laisser là, ça formalise le résultat mais on a pas besoin de celui là, ça permet de bien formaliser ce qu'on entend par contrôler le délai, et dans la preuve plus bas on prouve ce qu'on souhaite.}
@@
 This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the proof of the finite-time regret bound for the variant using ``global restarts''.
-The proof of the other variant using ``local restarts'' follows the skeleton of the unified analysis, and can be found in Appendix~E of \cite{Besson2019GLRT}.
+The proof of the other variant using ``local restarts'' follows the skeleton of the unified analysis, and it can be found in Appendix~E of \cite{Besson2019GLRT}.
 
@@
 
-In this section, we sketch a unified proof for the two Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given below, whose detailed proofs are given in Section~\ref{proof:6:mainRegretBoundGlobal} and in Appendix~E of our paper \cite{Besson2019GLRT} respectively.
+In this section, we sketch a unified proof for the two Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given above.
+%  whose detailed proofs are given in Section~\ref{proof:6:mainRegretBoundGlobal} and in Appendix~E of our paper \cite{Besson2019GLRT} respectively.
 We emphasize that our approach is significantly different from those proposed by \cite{CaoZhenKvetonXie18} for \MUCB{} and by \cite{LiuLeeShroff17} for \CUSUMUCB.
-
-Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
-Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}),
-we observe that we have $\mu_{i^*_t} - \mu_{I_t} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
-and thus we can write a first decomposition of the regret,
+%
+Recall that in this Chapter, the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
+First, we introduce $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}),
+and as we observe that $\mu_{i^*_t} - \mu_{I_t} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
+we can write a first decomposition of the regret,
 \begin{align*}
@@
 \paragraph{Upper bound on term $(b)$: controlling the delay.}
+\label{par:6:controllingDelayInOneProof}
 %
@@
 \left[\left\lfloor \frac{\alpha}{K}\left(\tau^{(k)}-\widehat{\tau}^{(k-1)}\right)\right\rfloor,\left(\tau^{(k)}-\widehat{\tau}^{(k-1)}\right)\right]$.
-Conditionally to $\cC^{(k-1)}$, one obtains furthermore using that $d^{(k-1)} \leq (\tau^{(k)} - \tau^{(k-1)})/2$ -- which follows from Assumption~\ref{ass:6:LongPeriodsGlobal} -- that
+Conditionally to $\cC^{(k-1)}$,
+by using Assumption~\ref{ass:6:LongPeriodsGlobal},
+we can prove that $d^{(k-1)} \leq (\tau^{(k)} - \tau^{(k-1)})/2$,
+and thus we obtain furthermore that
 %
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
-    \caption{Location of the detected change-points for four algorithms on Problem 1.}
+    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
+    \caption{Locations of the detected change-points for four algorithms on Problem 1.}
     \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1}
 \end{figure}
-
-% \TODOL{Conclusion à tirer de cet exemple !}
 
@@
 
-% \TODOL{Conclusion à tirer de cet exemple !}
-
 \begin{figure}[h!]  % [htbp]
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
-    \caption{Location of the detected change-points for four algorithms on Problem 2.}
+    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
+    \caption{Locations of the detected change-points for four algorithms on Problem 2.}
     \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2}
@@
 
-On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there is more changes, we clearly observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much more than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
-Our approach, with the two options of \textbf{Local} or \textbf{Global} restarts, performs very closely to the oracle for problem $4$.
+On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there are more changes, we clearly observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much better than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
+% Our approach, with the two options of \textbf{Local} or \textbf{Global} restarts, performs very closely to the oracle for problem $4$.
+Our approach is the algorithm which performs the closer to the oracle for problem $4$.
 

### 2019-05-22 2-Chapters/Conclusion/conclusion.tex
Meh, a not-too-stupid first quickly-written version of the conclusion

--- 2019-01-17 2-Chapters/Conclusion/conclusion.tex
@@
-%!TEX root = ../thesis.tex
+%!TEX root = ../PhD_thesis__Lilian_Besson
 
 % First chapter begins here
-\chapter*{General Conclusion and Perspectives}
-\addcontentsline{toc}{chapter}{General Conclusion}
-%\minitoc
+\chapter{General Conclusion and Perspectives}
+% \addcontentsline{toc}{chapter}{General Conclusion}
 \label{chapter:conclusion}
+\minitoc
+
 % Write miniTOC just after the title
 \graphicspath{{2-Chapters/6-Chapter/Images/}}
+% ----------------------------------------------------------------------
 
-In this PhD thesis, blabla.
+% \newpage
+\paragraph{Abstract}
 
-In \textbf{Chapter \ref{chapter:4}}, blabla.
+This last chapter concludes the thesis, first by summarizing the previous chapters, and by giving a general conclusion of the thesis.
+%
+We highlight what we consider to be the most important directions of future works, as well as some promising directions that were shortly studied during my PhD but not discussed in the manuscript.
 
-\section*{Perspectives}
-We present in the following the possible areas of future studies
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
+\section{General Conclusion}
 
-\subsubsection*{Blabla}
-\begin{itemize}
-\item Blabla
-\end{itemize}
+\TODOL{Je dois faire quelques phrases résumant la thèse}
+
+This manuscript was organized in two parts, Part~\ref{part:Introduction} and Part~\ref{part:MABIOT}.
+It presented the problems we considered for my PhD research, and we discussed about the main contributions produced.
+
+The first half, Part~\ref{part:Introduction}, starts by motivating our work and situating the context of this research.
+% Chapter 2
+In Chapter~\ref{chapter:2}, we present the multi-armed bandit models, the hypotheses we usually make, and the most important algorithms used to solve MAB problems.
+% Chapter 3
+Then in Chapter~\ref{chapter:3} we discuss in details about our open-source Python library developed to run numerical simulations of MAB problems, SMPyBandits.
+
+The second half of this thesis, Part~\ref{part:MABIOT}, presents our main contributions.
+% Chapter 4
+In Chapter~\ref{chapter:4}, we first propose some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
+% In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithm.
+We also present a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.
+% Chapter 5
+Because the models studied for realistic IoT networks were found to be intractable to analyse from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in a noisy environment,
+in Chapter~\ref{chapter:5}, we study a weaker but still very interesting model.
+We present different variants of the multi-player MAB model, and we propose the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
+% Chapter 6
+Finally, in Chapter~\ref{chapter:6}, we remove the stationary hypothesis on the single-player MAB problem, and we analyze a new actively adaptive algorithm (\GLRklUCB).
+%
+For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
+
+A more precise summary of each chapters is included below.
+
+% % ----------------------------------------------------------------------------
+% \paragraph{Conclusion about \textbf{Chapter \ref{chapter:1}}}
+
+% ----------------------------------------------------------------------------
+\paragraph{Conclusion about \textbf{Chapter \ref{chapter:2}}}
+
+We started to present the multi-armed bandit model, with a finite number of arms and real-valued rewards.
+Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
+The notations used in the manuscript were introduced by showcasing an interactive demonstration made available online\footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
+
+The first contribution of this manuscript \cite{Besson2018WCNC} concluded this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
+Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+
+% ----------------------------------------------------------------------------
+\paragraph{Conclusion about \textbf{Chapter \ref{chapter:3}}}
+
+Then we presented our Python library SMPyBandits \cite{SMPyBandits,SMPyBanditsJMLR}, by detailing
+its purpose and qualities, its organization, and an overview of its usage.
+%
+SMPyBandits allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-player multi-armed bandits.
+SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
+% Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.
+
+We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
+As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
+Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
+One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
+More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.
+All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
+%
+When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
+%
+Our library is used to run numerical simulations on Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
+
+% ----------------------------------------------------------------------------
+\paragraph{Conclusion about \textbf{Chapter \ref{chapter:4}}}
+
+This chapter started the second part of the manuscript, Part~\ref{part:MABIOT}, and presents three main contributions.
+
+First in Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
+with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
+Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
+Similarly, if the number of devices remain constant, increasing the successful transmission rate directly extends the IoT devices battery life, as they suffer less from failed transmissions.
+We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
+%
+This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
+The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each others, and \emph{do not} know the system parameters.
+
+We presented in Section~\ref{sec:4:gnuradio} a proof-of-concept, demonstrated in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
+We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
+Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
+%
+By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.
+
+Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency.
+% , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
+If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
+The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
+We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
+%
+We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
+Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
+Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.
+
+% ----------------------------------------------------------------------------
+\paragraph{Conclusion about \textbf{Chapter \ref{chapter:5}}}
+
+The second chapter of Part~\ref{part:MABIOT} of this manuscript focussed on Multi-Player Multi-Arm Bandits models.
+%
+We present three variants of this model,
+with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
+For the two easiest models --with sensing--, our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret.
+In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
+%
+We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proved to be very active in the last two years.
+For some models, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
+For example, it assumes to know the number of players $M$ before-hand, but we illustrated that previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
+% Further works would be required to adapt the theoretical analysis to these various extensions.
+
+% ----------------------------------------------------------------------------
+\paragraph{Conclusion about \textbf{Chapter \ref{chapter:6}}}
+
+In the last chapter, we studied and presented the piece-wise stationary bandit model.
+%
+We proposed a new algorithm for this problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
+This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but \emph{without any other prior knowledge on the problem}, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
+We also gave numerical evidence of the efficiency of our proposal, which performs usually much better than all the passively adaptive approaches as well as better or comparably to the previous actively adaptive ones.
+
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
+\section{Perspectives}
+
+As discussed at the end of each chapters, our works suggest many interesting directions of future studies.
+
+% ----------------------------------------------------------------------------
+% \paragraph{Perspectives about \textbf{Chapter \ref{chapter:1}}}
+
+% ----------------------------------------------------------------------------
+% \paragraph{Perspectives about }}
+
+As detailed in Chapter~\ref{chapter:2}, we focussed in this thesis on finite-arms and one-dimensional bandit problems,
+and thus two possible directions of future works could be to extend our works
+to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
+The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.
+
+% ----------------------------------------------------------------------------
+% \paragraph{Perspectives about \textbf{Chapter \ref{chapter:3}}}
+
+An interesting future work left on our library SMPyBandits could be to implement new variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
+%
+It would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
+or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+
+% ----------------------------------------------------------------------------
+\paragraph{Perspectives about \textbf{Chapter \ref{chapter:4}}}
+
+% Future works related to this chapter include the following directions.
+
+% \paragraph{Possible extensions of the first model}
+
+The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device.
+Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
+In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
+%
+Moreover, for sake of simplicity we supposed that all devices use the same standard.
+Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
+
+Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
+So additionally to deciding the channel for communication (\ie, \emph{where} to communicate), each device has to also decide \emph{when} to communicate.
+% This is another direction of research, that we will investigate in the future.
+However, this clearly leads to a much larger action space, as there are many time slots in one day (for example), and thus we believe that as soon as the action space becomes too large in this extension, the simple MAB-based learning approach could be no longer appropriate.
+It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
+It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
+
+% We will investigate this behavior in order to understand it better theoretically.
+% We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
+
+% \paragraph{Extensions of the demonstration}
+
+% Possible future extensions of our demonstration include the following points.
+% We could consider more dynamic devices (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more devices with the same card.
+% We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
+% Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card to simulate another gateway).
+
+% \paragraph{Possible extensions for the second model}
+
+% Finally, the utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
+Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
+% %
+% For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
+% or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
+% or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
+% that nevertheless have not been explored for the targeted problem.
+
+In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our demonstration \cite{Besson2018ICT,Besson2019WCNC}.
+%
+A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.
+
+% ----------------------------------------------------------------------------
+\paragraph{Perspectives about \textbf{Chapter \ref{chapter:5}}}
+
+Our study on multi-player bandits suggests several interesting further research directions.
+First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
+So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
+However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
+Similarly to what was done very recently in \cite{wang2019distributed} for communication between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
+
+We also present many extensions of the multi-player bandit model,
+and even if some have already been implemented, a major future work is to implement the most interesting ones in our library SMPyBandits
+(see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
+
+% ----------------------------------------------------------------------------
+\paragraph{Perspectives about \textbf{Chapter \ref{chapter:6}}}
+
+We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected.
+This interesting research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$.
+We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
+We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
+
+We would also like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}.
+
+Another interesting direction of future work is to study non-stationary distributed multi-player bandits, to unify the models from Chapter~\ref{chapter:5} and \ref{chapter:6}.
+A natural extension of the non-stationary model is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
+%
+We could also build on the proof technique used in \cite{WeiSrivastava18Abruptly}, even if we are interested in removing the hypothesis that player $j$ knows its ID $j\in[M]$ before starting to play.
+%
+A promising direction is to directly try to join our contributions from Chapter~\ref{chapter:5} and \ref{chapter:6}, and propose an efficient algorithm using three parts:
+\klUCB{} indexes for arm selection,
+\MCTopM{} for orthogonalization (\ie, dealing with collisions),
+and \GLRklUCB{} for non-stationarity (\ie, dealing with abrupt changes).
+
+% ----------------------------------------------------------------------------
+\paragraph{Other directions of future works}
+
+\TODOL{On doit encore en écrire un peu ?}
+
+% % ----------------------------------------------------------------------------
+% % ----------------------------------------------------------------------------
+% \section{Personal Summary}
+
+% I would like to highlight that from a personal point of view, this thesis was awesome blabla.

### 2019-05-22 2-Chapters/2-Chapter/chapter2.tex
Small proofreading of notations of Chapter6
Small changes on Chap5,6.
- Nice table with multirow in Ch5 like Ch6

@@
 % Just bounded, in $[a,b]$ but without loss of generality in $[0,1]$ etc.
-\paragraph{The normalization trick.}
+\paragraph{The normalization trick.}\label{par:2:normalizationTrick}
 %
@@
 
+Such decomposition of the regret is useful for at least two reasons.
+%
+On the first hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
+Mathematically, both quantities are equal in expectation, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed.
+In our library SMPyBandits, we implement both possibilities, and all values of regret used in this thesis are based on the second estimator, using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives smoother curves and faster convergence to the limit behavior.
+%
+On the second hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T$ and not directly the regret $R_T^{\cA}$.
+For example, we extend this decomposition to the multi-player case in Chapter~\ref{chapter:3}, in Lemma~\ref{lem:5:DecompositionRegret}, and as it also uses the draws $N_k(T)$, it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
+
 \subsection{Regret lower bounds}
@@
 
-Then we consider algorithms that uniformly efficient, meaning that we rule out algorithms that can achieve low regret on some problem instance while achieving linear regret on other instances.
+Then we consider algorithms that uniformly efficient\footnote{This notion is then extended for ``strongly uniformly efficiency'' in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}.}, meaning that we rule out algorithms that can achieve low regret on some problem instance while achieving linear regret on other instances.
 In particular, it is necessary to rule out algorithms that always pick the same arms, as on some problem instance such fixed-arm algorithm can achieve zero regret.
@@
     \begin{equation}
-        R_T^{\cA}(I) \leq \bigO{C_{I,\alpha} T^{\alpha}}
+        R_T^{\cA}(I) = \bigO{C_{I,\alpha} T^{\alpha}}
     \end{equation}

### 2019-05-22 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Small changes on Chap5,6.
- Nice table with multirow in Ch5 like Ch6

@@
 
-% FIXME
-\TODOL{Est-ce qu'on enlève toute mention de la lower-bound et de "order-optimal" du coup ?!}
-% FIXME
+% % FIXME
+% \TODOL{Est-ce qu'on enlève toute mention de la lower-bound et de "order-optimal" du coup ?!}
+% % FIXME
 
@@
     \textbf{Sensing, then collision}: player $j$ observes $Y_{A^j(t),t}$, \emph{then} observes the reward, and thus also $C^j(t)$ only if $Y_{A^j(t),t} = 1$.
-    This common setup, studied for example by \cite{Anandkumar11,Avner15,Rosenski16}, is relevant to model the OSA problem: the device first checks for the presence of primary users in the chosen channel.
-    If this channel is free ($Y_{A^j(t),t}=1$), the transmission is successful ($r^j(t)=1$) if no collision occurs with other smart devices ($\overline{C^j(t)}$).
+    This common setup, studied for example by \cite{Anandkumar11,Rosenski16}, is relevant to model the OSA problem: the device first checks for the presence of primary users in the chosen channel,
+    if this channel is free ($Y_{A^j(t),t}=1$), the transmission is successful ($r^j(t)=1$) if no collision occurs with other smart devices ($\overline{C^j(t)}$).
   \item[\modeltrois]
@@
 % \section{An improved asymptotic regret lower bound (for some algorithms)}
-\section{Decomposing the regret to get an intuition for the design of efficient decentralized algorithms}
+\section[Decomposing the regret to get an intuition about efficient decentralized algorithms]{Decomposing the regret to get an intuition for the design of efficient decentralized algorithms}
 \label{sec:5:lowerbound}
@@
 
-Similarly to what we present above in Chapter~\ref{chapter:2}, Section~\ref{sec:2:lowerUpperBoundsRegret},
+Similarly to what we present above in Section~\ref{sec:2:lowerUpperBoundsRegret},
 we use the Kullback-Leibler divergence to express our lower bound.
@@
 %
-We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
+We first introduce the assumption under which we derive a regret lower bound, that generalizes the assumption of uniform efficiency given in Definition~\ref{def:2:uniformlyEfficientAlgorithm},
+a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
 % Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
@@
 The additional assumption in \eqref{eq:5:SUE} further implies some notion of \emph{fairness}, as it suggests that each of the $M$ players spends on average the same amount of time on each of the $M$ best arms.
-Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\rho^{\gamma} = (\rho^{\gamma(1)},\dots,\rho^{\gamma(M)})$ is independent from the choice of permutation $\gamma \in \Sigma_M$.
+Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\rho^{\sigma} = (\rho^{\sigma(1)},\dots,\rho^{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$.
 In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
@@
 \begin{table}[ht]
-  % \begin{footnotesize}
+  \begin{small}  % XXX
       \centering
-      \begin{tabular}{c|ccc}
-      \textbf{Algorithm} $\;$ \textbackslash $\;$ Number of players & $M=3$ & $M=6$ & $M=9$ \\
+      \begin{tabular}{cc|ccc}
+          \textbf{Algorithm} & \textbf{Index policy} & $M=3$ players & $M=6$ players & $M=9$ players \\
           \hline
-          Centralized multiple-play with \UCB{} & $321 \pm 30$ & $233 \pm 25$ & $0$ \\
-          Centralized multiple-play with \klUCB{} & $94 \pm 17$ & $68 \pm 18$ & $0$ \\
+          \multirow{2}{*}{Centralized multiple-play}
+          & \UCB{} & $321 \pm 30$ & $233 \pm 25$ & $0$ \\
+          & \klUCB{} & $94 \pm 17$ & $68 \pm 18$ & $0$ \\
           \hline
-          \Selfish{} with \UCB{} & $1263 \pm 157$ & $3694 \pm 387$ & $12420 \pm 404$ \\
-          \Selfish{} with \klUCB{} & $243 \pm 31$ & $743 \pm 113$ & $3005 \pm 492$ \\
+          \multirow{2}{*}{\Selfish}
+          & \UCB{} & $1263 \pm 157$ & $3694 \pm 387$ & $12420 \pm 404$ \\
+          & \klUCB{} & $243 \pm 31$ & $743 \pm 113$ & $3005 \pm 492$ \\
           \hline
-          \RhoRand{} with \UCB{} & $1455 \pm 208$ & $4775 \pm 463$ & $11794 \pm 1083$ \\
-          \RhoRand{} with \klUCB{} & $394 \pm 96$ & $2385 \pm 412$ & $7057 \pm 1053$ \\
+          \multirow{2}{*}{\RhoRand}
+          & \UCB{} & $1455 \pm 208$ & $4775 \pm 463$ & $11794 \pm 1083$ \\
+          & \klUCB{} & $394 \pm 96$ & $2385 \pm 412$ & $7057 \pm 1053$ \\
           \hline
-          \RandTopM{} with \UCB{} & $1020 \pm 92$ & $2899 \pm 418$ & $470 \pm 346$ \\
-          \RandTopM{} with \klUCB{} & $258 \pm 43$ & $902 \pm 234$ & $551 \pm 520$ \\
+          \multirow{2}{*}{\RandTopM}
+          & \UCB{} & $1020 \pm 92$ & $2899 \pm 418$ & $470 \pm 346$ \\
+          & \klUCB{} & $258 \pm 43$ & $902 \pm 234$ & $551 \pm 520$ \\
           \hline
-          \MCTopM{} with \UCB{} & $980 \pm 76$ & $1466 \pm 132$ & $43 \pm 13$ \\
-          \MCTopM{} with \klUCB{} & $\mathbf{248 \pm 40}$ & $\mathbf{410 \pm 54}$ & $\mathbf{42 \pm 10}$ \\
+          \multirow{2}{*}{\MCTopM}
+          & \UCB{} & $980 \pm 76$ & $1466 \pm 132$ & $43 \pm 13$ \\
+          & \klUCB{} & $\mathbf{248 \pm 40}$ & $\mathbf{410 \pm 54}$ & $\mathbf{42 \pm 10}$ \\
           \hline
       \end{tabular}
-      \caption{Using \klUCB{} is much more efficient than using \UCB{}, for multi-player bandit (here in a simple problem with $K=9$ arms).}
+      \caption{No matter the orthogonalization policy, using \klUCB{} is much more efficient than using \UCB{}, for multi-player bandit (here in a simple problem with $K=9$ arms).}
       \label{table:5:comparisonUCB_klUCB}
-  % \end{footnotesize}
+  \end{small}  % XXX
   \end{table}

### 2019-05-22 2-Chapters/5-Chapter/chapter5.tex
Small changes on Chap5,6.
- Nice table with multirow in Ch5 like Ch6

@@
 
-For this first model, we present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
+For this first model, we present two algorithms, \RandTopM{} and \MCTopM, combining an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
 Like in previous works, we consider the centralized system regret (multi-player regret).

### 2019-05-22 2-Chapters/Conclusion/conclusion.tex
Small changes on Chap5,6.
- Nice table with multirow in Ch5 like Ch6

@@
 
+\begin{itemize}
+    \item
 The first half, Part~\ref{part:Introduction}, starts by motivating our work and situating the context of this research.
@@
 
+    \item
 The second half of this thesis, Part~\ref{part:MABIOT}, presents our main contributions.
@@
 For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
+\end{itemize}
 
@@
 We proposed a new algorithm for this problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
-This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but \emph{without any other prior knowledge on the problem}, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
+This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
 We also gave numerical evidence of the efficiency of our proposal, which performs usually much better than all the passively adaptive approaches as well as better or comparably to the previous actively adaptive ones.
@@
 However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
-Similarly to what was done very recently in \cite{wang2019distributed} for communication between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
+Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
 
-We also present many extensions of the multi-player bandit model,
+We also presented many extensions of the multi-player bandit model,
 and even if some have already been implemented, a major future work is to implement the most interesting ones in our library SMPyBandits
@@
 
-We would also like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}.
+We would also like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or the rotted bandit model.
 

### 2019-05-22 2-Chapters/5-Chapter/chapter5.tex
So long

@@
 \paragraph{Interpretation of the hypothesis of knowning $M$ for real-world networks.}
+\label{par:5:knowingYourIDinMPBanditsGame}
 %
@@
 \begin{table}[ht]
-% \begin{footnotesize}
+\begin{small}  % WARNING
     \centering
-    \begin{tabular}{c|ccc}
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Number of players} & $M=3$ & $M=6$ & $M=9$ \\
+    \begin{tabular}{cc|ccc}
+    \textbf{Algorithms} & \textbf{Hyp. on $M$} & $M=3$ players & $M=6$ players & $M=9$ players \\
         \hline
-        Centralized multiple-play \klUCB{} & $\mathbf{92 \pm 20}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0}$ \\
-        \Selfish-\klUCB{} & $250 \pm 34$ & $735 \pm 85$ & $3010 \pm 472$ \\
+        Centralized multiple-play
+        & Known $M$ & $\mathbf{92 \pm 20}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0}$ \\
         \hline
-        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 \pm 1035$ \\
-        \RhoRand-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$1422 \pm 1051$} & \textcolor{red}{$9030 \pm 1922$} & \textcolor{red}{$7264 \pm 1009 $} \\
+        \Selfish
+        & Don't need $M$ & $250 \pm 34$ & $735 \pm 85$ & $3010 \pm 472$ \\
         \hline
-        \RandTopM-\klUCB{} & $268 \pm 45$ & $941 \pm 217$ & $437 \pm 367$ \\
-        \RandTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$688 \pm 614$} & \textcolor{red}{$4256 \pm 2701$} & \textcolor{red}{$1155 \pm 544$} \\
+        \multirow{2}{*}{\RhoRand}
+        & Known $M$ & $417 \pm 103$ & $2481 \pm 449$ & $6639 \pm 1035$ \\
+        & \textcolor{red}{Estimate $M$} & \textcolor{red}{$1422 \pm 1051$} & \textcolor{red}{$9030 \pm 1922$} & \textcolor{red}{$7264 \pm 1009 $} \\
         \hline
-        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 \pm 13$ \\
-        \MCTopM-\klUCB{} + \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$618 \pm 29$} \\
+        \multirow{2}{*}{\RandTopM}
+        & Known $M$ & $268 \pm 45$ & $941 \pm 217$ & $437 \pm 367$ \\
+        & \textcolor{red}{Estimate $M$} & \textcolor{red}{$688 \pm 614$} & \textcolor{red}{$4256 \pm 2701$} & \textcolor{red}{$1155 \pm 544$} \\
+        \hline
+        \multirow{2}{*}{\MCTopM}
+        & Known $M$ & $244 \pm 39$ & $401 \pm 55$ & $44 \pm 13$ \\
+        & \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$618 \pm 29$} \\
         \hline
     \end{tabular}
-    \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, comparing algorithms which knows $M$ against algorithms which estimate $M$ on the fly.}
+    \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, comparing algorithms which knows $M$ against algorithms which estimate $M$ on the fly.
+All use \klUCB.}
     \label{table:5:meanRegretSimulationsEstimatingM}
-% \end{footnotesize}
+\end{small}  % WARNING
 \end{table}
@@
 Our proposals are still much more efficient than \RhoRand{} when using the procedure to estimate $M$ described from \cite{Anandkumar11}.
-For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large (on the three algorithms and two cases, about $\times 3.4$), and consistent on the different algorithm.
+For the two non-extreme cases ($M=3,6$), the performance drop when having to estimate $M$ is quite large,
+% (on the three algorithms and two cases, about $\times 3.4$),
+and consistent on the different algorithms.
 The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proved to be $\cO(1)$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
@@
     \end{tabular}
-    \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, for the ``no sensing'' case.}
+    % \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, for the ``no sensing'' case.}
+    \caption{Comparison for different algorithms on the same problem with $M=3,6,9$ players, for the ``no sensing'' case.
+We give the mean regret $\pm$ $1$ std-dev, and more work is needed on our implementation on Improved \MusicalChair.
+The results on \textsc{Sic-MMAB} appear to confirm the numerical experiments in the paper \cite{BoursierPerchet18}.}
     \label{table:5:meanRegretSimulationsNoSensing}
@@
 Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between devices are possible.
-Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
+Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``\emph{graph bandit}'' problem.
 What we are interested here is an extension of the model where players can send some bits to one or all the other players, at every time steps or some time steps.
@@
 % combining Musical Chair with an epoch-based alternative phases of forced exploration and pure exploitation.
-They analyse their ``Game of Thrones'' algorithm and proved that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
+They analyze their ``Game of Thrones'' algorithm and proved that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
 
@@
     \end{tabular}
-    \caption{Comparing \RhoRand{} and \rhoLearn on a simple MP bandit problem with $K=9$ arms.}
+    \caption{Comparing \RhoRand{} and \rhoLearn{} on a simple MP bandit problem with $K=9$ arms.}
     \label{table:5:comparisonRhoRandRhoLearn}
@@
 \indent
-The proposed the RR-SW-UCB\# algorithm, which combines their SW-UCB\# algorithm, previously proposed in \cite{WeiSrivastava18Abruptly}, and a Round-Robin hoping, under the assumption that each user $j\in[M]$ knows its ID $j$ (see above the discussion we give on this unrealistic assumption).
+The proposed the RR-SW-UCB\# algorithm, which combines their SW-UCB\# algorithm, previously proposed in \cite{WeiSrivastava18Abruptly}, and a Round-Robin hoping, under the assumption that each user $j\in[M]$ knows its ID $j$ (we criticize this unrealistic assumption in Section~\ref{par:5:knowingYourIDinMPBanditsGame} above).
 The SW-UCB\# algorithm is discussed more in details in the literature review of Chapter~\ref{chapter:6} below.
@@
 
-\TODOL{Explain that we think the following approach can work very efficiently:
+% \TODOL{Explain that we think the following approach can work very efficiently:
 
-    - Combine \MCTopM{} with \GLRklUCB{} instead of \klUCB,
-    - and incorporate also the detected change-points by B-\GLR{} test in the orthogonalization procedure: after any change-point, the player is not sitting anymore.
-    - Modification: only a change-point on the arm on which player is sitting can make him move?
+%     - Combine \MCTopM{} with \GLRklUCB{} instead of \klUCB,
+%     - and incorporate also the detected change-points by B-\GLR{} test in the orthogonalization procedure: after any change-point, the player is not sitting anymore.
+%     - Modification: only a change-point on the arm on which player is sitting can make him move?
 It's already the case: the \GLR{} test can only detect change-point on an arm that was played, and the \MCTopM{} orthogonalization procedure forces to play the same arm as long as the player is sitting.
-}
+% }
 
@@
 % ----------------------------------------------------------------------------
-\subsection{A note on SMPyBandits}
+\subsection{Some future work on SMPyBandits}
 
@@
 We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future work left on our library.
-For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/X}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
 

### 2019-05-22 2-Chapters/Conclusion/conclusion.tex
So long

@@
 % Chapter 5
-Because the models studied for realistic IoT networks were found to be intractable to analyse from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in a noisy environment,
+Because the models studied for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in a noisy environment,
 in Chapter~\ref{chapter:5}, we study a weaker but still very interesting model.

### 2019-05-22 2-Chapters/1-Chapter/chapter1.tex
Added screenshots of the demo now hosted at https://perso.crans.org/besson/phd/MAB_interactive_demo/
My thesis is so long, and so stupid

@@
 % First chapter begins here
-\chapter{Setting of This Thesis}
+\chapter{Introduction}
 \label{chapter:1}
@@
 
-Abstract for Chapter 1.
-
+\paragraph{Abstract.}
+We start by presenting the problems studied in this thesis,
+then we announce the different directions we considered and the solutions we proposed.
+We give an overview of our main contributions, alongside a list of publications written in the last three years.
 \TODOL{Write this!}
@@
     \centering
-    \resizebox{0.85\textwidth}{!}{
+    \resizebox{0.40\textwidth}{!}{
         \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=2]
@@
 Here we discuss quickly the organization of the thesis.
+The manuscript is organized in two parts and six chapters, and the reading order can be any top-down path between Chapter~\ref{chapter:1} and the Conclusion in Chapter~\ref{chapter:conclusion}, in the graph shown in Figure~\ref{fig:1:organization} below.
+In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2} where we introduce the MAB models, the concepts and the notations used in all this document, and it is mandatory to read it.
+Conversely, Chapter~\ref{chapter:3} presents our simulation library SMPyBandits, and even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use the library for their numerical simulation, reading this chapter is not required to understand the rest of the document.
+%
+Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
+Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where MAB algorithms have been used with success.
+Our two models are interesting and close to reality, but they appeared to be too general to propose a mathematical analysis of the good empirical performance of the considered solutions.
+For this reason, we weaken the models for the rest of the document,
+and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.
 
@@
     \centering
-    \resizebox{0.85\textwidth}{!}{
-    \begin{tikzpicture}[>=latex',line join=bevel,scale=3]
+    % \resizebox{1.00\textwidth}{!}{
+    \begin{tikzpicture}[>=latex',line join=bevel,scale=2.25]
         %
-        \node[align=center] (introduction) at (0,2.5) [rectangle,draw] {Chapter~\ref{chapter:1}\\Introduction};
-        \node[align=center] (chapter2) at (0,2) [rectangle,draw] {Chapter~\ref{chapter:2}\\Multi-Armed Bandit models};
-        \node[align=center] (chapter3) at (4,2) [rectangle,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
-        \node[align=center] (chapter4) at (-2,1) [rectangle,draw] {Chapter~\ref{chapter:4}\\Two MAB models for IoT networks};
-        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapter~\ref{chapter:5}\\Multi-players MAB};
-        \node[align=center] (chapter6) at (2,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary MAB models};
+        \node[align=center] (introduction) at (0,3) [rectangle,draw] {Chapter~\ref{chapter:1}\\Introduction};
+        \node[align=center] (chapter2) at (0,2) [rectangle,draw] {Chapter~\ref{chapter:2}\\Stochastic and Stationary\\Multi-Armed Bandit models};
+        \node[align=center] (chapter3) at (2.5,2) [rectangle,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
+        \node[align=center] (chapter4) at (-2,1) [rectangle,draw] {Chapter~\ref{chapter:4}\\Two MAB models\\for IoT networks};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapter~\ref{chapter:5}\\Multi-players\\MAB models};
+        \node[align=center] (chapter6) at (2,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary\\MAB models};
         \node[align=center] (conclusion) at (0,0) [rectangle,draw] {Chapter~\ref{chapter:conclusion}\\General conclusion};
@@
         \draw [color=black,thick,->] (chapter2) to (chapter5);
+        \draw [color=black,densely dotted,<->]   (chapter4) to (chapter5);
+        \draw [color=black,densely dotted,->]   (chapter3) to (chapter6);
+        \draw [color=black,densely dotted,<->]   (chapter5) to (chapter6);
         \draw [color=black,thick,->] (chapter2) to (chapter6);
@@
     \end{tikzpicture}
-    }
-    \caption{Organization of the thesis.}
+    % }
+    \caption{Organization of the thesis: a reading map.
+Any path containing Chapter~\ref{chapter:1}, Chapter~\ref{chapter:2}, at least one of the three Chapters~\ref{chapter:4}, \ref{chapter:5} and \ref{chapter:6}, and the Conclusion is a self contained way to read this thesis.}
     \label{fig:1:organization}
@@
 \begin{itemize}
+
+\item
+    \emph{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation},\\
+    by Christophe Moy \& \textbf{Lilian Besson}.
+    1st International ISIoT workshop\footnote{\href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}},
+    at \emph{Conference on Distributed Computing in Sensor Systems} (IEEE DCOSS, \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}), Santorini, Greece, May 2019.
+    \cite{MoyBesson2019}
+
+\item
+    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
+    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
+    1st International MOTIoN workshop\footnote{\href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
+    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019.
+    \cite{Bonnefoi2019WCNC}
+
 \item
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019,
+    \emph{Wireless Communication and Networks Conference}\footnote{IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,
     \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann.
-    \emph{Algorithmic Learning Theory} (ALT, \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}), Lanzarote, Spain, April 2018,
+    \emph{Algorithmic Learning Theory}\footnote{ALT, \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}}, Lanzarote, Spain, April 2018,
     \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
@@
     by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}), Barcelone, Spain, April 2018,
+    \emph{Wireless Communication and Networks Conference}\footnote{IEEE WCNC, \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,
     \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
-    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}), Lisboa, Portugal, Septembre $2017$,
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{CROWNCOM, \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,
     \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
@@
 % =============================================================================
-\subsection{Publications in international conferences without proceedings}
+\subsection{Demonstration in international conferences}
 
@@
 \item
-    \emph{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation},\\
-    by Christophe Moy \& \textbf{Lilian Besson}.
-    1st International ISIoT workshop (\href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}),
-    at \emph{Conference on Distributed Computing in Sensor Systems} (IEEE DCOSS, \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}), Santorini, Greece, May 2019.
-    \cite{MoyBesson2019}
-
-\item
-    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
-    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
-    1st International MOTIoN workshop (\href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}),
-    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019.
-    \cite{Bonnefoi2019WCNC}
-
-\item
     \emph{MALIN: ``Multi-Arm bandit Learning for Iot Networks'' with GRC: A TestBed Implementation and Demonstration that Learning Helps},
     by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
-    Demonstration presented at \emph{International Conference on Communication} (ICT, \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}), in Saint-Malo, France in June $2018$.
+    Demonstration presented at \emph{International Conference on Communication}\footnote{ICT, \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}}, in Saint-Malo, France in June $2018$.
     See \href{https://YouTu.be/HospLNQhcMk}{\texttt{YouTu.be/HospLNQhcMk}} for a $6$-minutes presentation video.
@@
 % =============================================================================
+% \subsection{French language conference with proceedings}
 \subsection{Submitted works}
@@
     \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
-    submitted at GRETSI 2019, \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}},
+    submitted at GRETSI 2019\footnote{\href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
     by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
@@
     \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
+    by \textbf{Lilian Besson} \& Emilie Kaufmann, \\
+    February $2019$,
     \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
@@
     About $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
-    with a complete documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}}.
+    and documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}}.
     \cite{SMPyBandits,SMPyBanditsJMLR}
@@
 \paragraph{Tutorial}
-    I gave a tutorial\footnote{Cf. \href{HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}} on the Julia language, at IETR seminar in Vannes in June $2018$,
-    with Pierre Haessig (\texttt{pierreh.eu}).
+    I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
+    with Pierre Haessig (\texttt{pierreh.eu}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.
 
@@
     I also helped for the organization and presented another poster\footnote{Cf.
-\href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Evaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.
+\href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.
 
 \paragraph{Reviews}
-	for \emph{European Workshop on Reinforcement Learning} (Lille) \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}, in October $2018$, I reviewed $5$ papers.
+	for \emph{European Workshop on Reinforcement Learning}\footnote{Cf.
+\href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
 	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

### 2019-05-22 5-Nomenclature/Nomenclature.tex
Starting to change notation \{1,\dots,M\} -> [M]
My thesis is so long, and so stupid

@@
 \nomenclature[A]{wlog}{\textit{with loss of generality}} % TODO not used?
-\nomenclature[A]{i.i.d.}{\textit{identically and independently distributed}}
+\nomenclature[A]{i.i.d.}{\textit{identically and independently distributed} (variables, observations or samples)}
 
@@
 \nomenclature[A]{CNRS}{\textit{Centre National de la Recherche Scientifique}}
-\nomenclature[A]{SCEE}{\textit{Signal Communication Électronique Embarquée} (research team in CentraleSupélec, campus of Rennes)}
+\nomenclature[A]{SCEE}{\textit{Signal, Communication et Électronique Embarquée} (research team in CentraleSupélec, campus of Rennes)}
 
@@
 \nomenclature[A]{PHY}{\textit{PHYsical} (layer)}
-\nomenclature[A]{MAC}{\textit{Medium ACcess} (layer)}
+\nomenclature[A]{MAC}{\textit{Medium Access Control} (layer)}
 \nomenclature[A]{ALOHA}{\textit{ALOHA} (not an acronym)}
@@
 \nomenclature[A]{PU}{\textit{Primary User}}
-\nomenclature[A]{BST}{\textit{Base-Station} (or \textit{gateway})}
+\nomenclature[A]{BTS}{\textit{Base Tranceiver Station} (or \textit{gateway})}
 \nomenclature[A]{QPSK}{\textit{Quadrature Phase-Shift Keying}}
@@
 \nomenclature[L]{$A(t)$, $A^j(t)$}{Decision of algorithm $\cA$ at time $t$, $A(t)\in\{1,\dots,K\}$. Decision for user $j$ in Chapter~\ref{chapter:5}}
-\nomenclature[L]{$Y_{k,t}$}{Random sample from the $k$ arm at time $t$}
+\nomenclature[L]{$Y_{k,t}$}{Random sample from the arm $k$ at time $t$}
 \nomenclature[L]{$\cO_t$}{Vector of observations until time $t$ in Chapter~\ref{chapter:5}}
@@
 \nomenclature[O]{$\cF$, $\cF_t$}{Filtration in a probabilistic model, after $t-1$ prior observations}
-\nomenclature[O]{$I_1,\dots,I_5$}{Used in Chapter~\ref{chapter:5} to denote various events}
+\nomenclature[O]{$[K]$, $[N]$ etc}{For an integer $N\in\N$, $N\geq0$, $[N]$ denotes the set $\{1,\dots,N\}$. If the order is important, it is ordered from $1$ to $N$.}
+% \nomenclature[O]{$I_1,\dots,I_5$}{Used in Chapter~\ref{chapter:5} to denote various events}
 \nomenclature[O]{$\cE$, $\cE_T$}{Used in Chapter~\ref{chapter:6} to denote ``good events'' that happen most of the time}
@@
 
-\nomenclature[O]{$o(\bullet)$, $\cO(\bullet)$, $\Omega(\bullet)$}{Landau notations: $f(x)=o(g(x)$ means that $g(x)\neq0$ and $f(x)/g(x)\to0$ for $x\to\infty$, $f(x)=\cO(g(x))$ means that there exists $x_0,K>0$ such that $f(x)\leq K g(x)$ from $x\geq x_0$, and $f(x)=\Omega(g(x))$ means $g(x)=\cO(f(x))$}
+\nomenclature[O]{$o(\bullet)$, $\cO(\bullet)$, $\Omega(\bullet)$}{Landau notations for positive functions: $f(x)=o(g(x)$ means that $g(x)\neq0$ and $f(x)/g(x)\to0$ for $x\to\infty$, $f(x)=\cO(g(x))$ means that there exists $x_0,K>0$ such that $f(x)\leq K g(x)$ from $x\geq x_0$, and $f(x)=\Omega(g(x))$ means $g(x)=\cO(f(x))$}
 
-\nomenclature[O]{$\overline{E}$}{Complement of an event $E$}
-\nomenclature[O]{$X'$}{Usually denotes a ``wrong value'' of a variable $X$, for instance $M'$ denotes in Section~\ref{par:5:usingWrongValueofM}}
+% \nomenclature[O]{$\overline{E}$}{Complement of an event $E$}
+\nomenclature[O]{$E^c$}{Complement of an event $E$}
+\nomenclature[O]{$X'$}{Usually denotes a ``wrong value'' of a variable $X$, for instance $M'$ denotes in Section~\ref{par:5:usingWrongValueofM}. Also denotes the derivative of a function, \eg, $f'$}
 \nomenclature[O]{$\widehat{X}$}{Usually denotes an ``empirical value'', a mean or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{\mu_k}$ the empirical mean of arm $k$ or $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:towardsRegretUpperBounds}}

### 2019-05-22 2-Chapters/1-Chapter/chapter1.tex
Thanks Christophe for comments on Ch2

@@
 We give an overview of our main contributions, alongside a list of publications written in the last three years.
-\TODOL{Write this!}
 
@@
 
-\TODOL{Write this!}
+We can list the following points to summarize the main contributions of this thesis:
 
-- Literature review of stochastic bandit algorithms (here)
+\begin{itemize}
+    \item
+    We present the concepts and the notations of the multi-armed bandit problem in Chapter~\ref{chapter:2}, from a mathematical point-of-view.
+    But we also follow a didactic approach as we use an online interactive demonstration designed to let anyone play against a small bandit problem from his/her browser, in Section~\ref{par:2:interactiveDemoDiscoverMAB}.
 
-- Aggregation of algorithms for online algorithm selection, state-of-the-art empirical performance with \textbf{Aggregator} (WCNC 2018)
+    \item
+    We give a short literature review of stochastic bandit algorithms in Section~\ref{sec:2:famousMABalgorithms}.
 
-- The most complete open-source simulation library for MAB problems, SMPyBandits (JMLR MLOSS and GitHub)
+    \item
+    We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+    We present an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
+    \cite{Besson2018WCNC}.
+    % with \textbf{Aggregator} (WCNC 2018)
 
-- Different models for MAB learning for IoT networks (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
+    \item
+    We wrote the most complete open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+    We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
+    A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.
 
-- A refined analysis and a new impulse on multi-player bandits (ALT 2018, and 4-8 works inspired by our article since then).
+    \item
+    We propose different models for IoT networks, in Chapter~\ref{chapter:4}, where end-devices with cognitive radio capabilities can implement MAB algorithms on their side, to automatically increase their battery life and allow more devices to use the same network while maintaining a high Quality of Service
+    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC}.
+    % (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
+
+    \item
+    We implemented a proof-of-concept of the aforementioned model \cite{Besson2018ICT}, and we present it in details in Section~\ref{sec:4:gnuradio}.
+We made a video showcasing our demonstration, hosted at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+
+    \item
+    The source code for the two previously mentioned contributions are all published online, along with clear instructions for reproducing our work.
+
+    \item
+    We formalize the multi-player bandit model, and we introduced three variants, in Chapter~\ref{chapter:5}.
+    For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
+    as well as extensive numerical experiments to demonstrate its good performance in comparison with the rest of the literature.
+    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-player bandits, as some recent research works built up on our results.
+
+    \item
+    We also give a detailed literature review of the different extensions of the multi-player MAB model, which we believe was never written before.
+    % were  (ALT 2018, and 4-8 works inspired by our article since then).
 State-of-the-art with our algorithm MCTopM + klUCB for multi-player bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"
 
-- Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
+    \item
+    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
+    Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
+    % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
 
-- Literature review on possible use case of doubling trick, plus a unified and more generic analysis of two families of doubling tricks !
+    \item
+    The last contribution of this thesis is a literature review of the possible use cases of the doubling trick technique for MAB problems,
+    and a unified and more generic analysis of two families of doubling tricks.
+    This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}, but not in the main text of the thesis.
+\end{itemize}
 
@@
 \label{sec:1:organization}
-
-Here we discuss quickly the organization of the thesis.
-The manuscript is organized in two parts and six chapters, and the reading order can be any top-down path between Chapter~\ref{chapter:1} and the Conclusion in Chapter~\ref{chapter:conclusion}, in the graph shown in Figure~\ref{fig:1:organization} below.
-In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2} where we introduce the MAB models, the concepts and the notations used in all this document, and it is mandatory to read it.
-Conversely, Chapter~\ref{chapter:3} presents our simulation library SMPyBandits, and even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use the library for their numerical simulation, reading this chapter is not required to understand the rest of the document.
-%
-Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
-Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where MAB algorithms have been used with success.
-Our two models are interesting and close to reality, but they appeared to be too general to propose a mathematical analysis of the good empirical performance of the considered solutions.
-For this reason, we weaken the models for the rest of the document,
-and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.
 
@@
         %
-        \node[align=center] (introduction) at (0,3) [rectangle,draw] {Chapter~\ref{chapter:1}\\Introduction};
-        \node[align=center] (chapter2) at (0,2) [rectangle,draw] {Chapter~\ref{chapter:2}\\Stochastic and Stationary\\Multi-Armed Bandit models};
-        \node[align=center] (chapter3) at (2.5,2) [rectangle,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
-        \node[align=center] (chapter4) at (-2,1) [rectangle,draw] {Chapter~\ref{chapter:4}\\Two MAB models\\for IoT networks};
+        \node[align=center] (introduction) at (0,3.25) [rectangle,draw] {Chapter~\ref{chapter:1}\\Introduction};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw] {Chapter~\ref{chapter:2}\\Stochastic and Stationary\\Multi-Armed Bandit models};
+        \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapter~\ref{chapter:4}\\Two MAB models\\for IoT networks};
         \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapter~\ref{chapter:5}\\Multi-players\\MAB models};
-        \node[align=center] (chapter6) at (2,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary\\MAB models};
-        \node[align=center] (conclusion) at (0,0) [rectangle,draw] {Chapter~\ref{chapter:conclusion}\\General conclusion};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary\\MAB models};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapter~\ref{chapter:conclusion}\\General conclusion};
         %
@@
         \draw [color=black,densely dotted,<->]   (chapter4) to (chapter5);
-        \draw [color=black,densely dotted,->]   (chapter3) to (chapter6);
+        \draw [color=black,densely dotted,->] -| (chapter3) to (chapter6);
         \draw [color=black,densely dotted,<->]   (chapter5) to (chapter6);
@@
 
+The reading order of the manuscript can be any top-down path between Chapter~\ref{chapter:1} and the Conclusion in Chapter~\ref{chapter:conclusion}, in the graph shown in Figure~\ref{fig:1:organization} above,
+and the thesis is organized in two parts:
+
+In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2} where we introduce the MAB models, the concepts and the notations used in all this document, and it is mandatory to read it.
+Conversely, Chapter~\ref{chapter:3} presents our simulation library SMPyBandits, and even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use the library for their numerical simulation, reading this chapter is not required to understand the rest of the document.
+
+Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
+Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where MAB algorithms have been used with success.
+Our two models are interesting and close to reality, but they appeared to be too general to propose a mathematical analysis of the good empirical performance of the considered solutions.
+For this reason, we weaken the models for the rest of the document,
+and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.
+
 % ----------------------------------------------------------------------------
@@
     by Christophe Moy \& \textbf{Lilian Besson}.
-    1st International ISIoT workshop\footnote{\href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}},
-    at \emph{Conference on Distributed Computing in Sensor Systems} (IEEE DCOSS, \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}), Santorini, Greece, May 2019.
+    1st International ISIoT workshop\footnote{~See \href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}},
+    at \emph{Conference on Distributed Computing in Sensor Systems}\footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}}, Santorini, Greece, May 2019.
     \cite{MoyBesson2019}
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
-    1st International MOTIoN workshop\footnote{\href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
-    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019.
+    1st International MOTIoN workshop\footnote{~MOTIoN 2019, see \href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
+    at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.
     \cite{Bonnefoi2019WCNC}
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,
+    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,
     \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann.
-    \emph{Algorithmic Learning Theory}\footnote{ALT, \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}}, Lanzarote, Spain, April 2018,
+    \emph{Algorithmic Learning Theory}\footnote{~ALT 2018, see \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}}, Lanzarote, Spain, April 2018,
     \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
@@
     by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{IEEE WCNC, \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,
+    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,
     \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
-    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{CROWNCOM, \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{~CROWNCOM, \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,\\
     \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
-    Demonstration presented at \emph{International Conference on Communication}\footnote{ICT, \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}}, in Saint-Malo, France in June $2018$.
+    Demonstration presented at \emph{International Conference on Communication}\footnote{~ICT, \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}}, in Saint-Malo, France in June $2018$.
     See \href{https://YouTu.be/HospLNQhcMk}{\texttt{YouTu.be/HospLNQhcMk}} for a $6$-minutes presentation video.
@@
     \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
-    submitted at GRETSI 2019\footnote{\href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
+    submitted at GRETSI 2019\footnote{~\href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
     by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
@@
 \item
-    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
-    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
+    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},
+    by \textbf{Lilian Besson} \& Emilie Kaufmann,\\
+    September $2018$,
     \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.
@@
     I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
-    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{Cf.
-\href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
-    I also helped for the organization and presented another poster\footnote{Cf.
-\href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.
+    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
+    I also helped for the organization and presented another poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.
 
 \paragraph{Reviews}
-	for \emph{European Workshop on Reinforcement Learning}\footnote{Cf.
-\href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
+	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
 	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

### 2019-05-22 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Thanks Christophe for comments on Ch2

@@
 %
-As expected, the more feedback they get, the faster the underlying algorithms learn, and the better the aggregation algorithm will be.
+As expected, the more feedback they get, the faster the underlying algorithms learn, and the better the aggregation algorithm is.
 This intuition is backed up by theory explained in \cite{Maillard11}.
@@
 First, $a\sim\pi^t$ is sampled first and the arm chosen by $\Alg_a$ is trusted, whereas \Aggr{}
-needs to listen to the $N$ decisions to perform the updates on $\pi^{t+1}$,
-and \ExpQ{} gives back an observation (arm, reward) only to the last trusted algorithm
+needs to listen to the $N$ decisions to perform the updates on $\pi^{t+1}$.
+Then, \ExpQ{} gives back an observation (arm, reward) only to the last trusted algorithm
 whereas \Aggr{} gives it to all algorithms.
@@
 %
-The sequence of non-negative learning rates $(\eta_t)_{t \geq 1}$ used by \ExpQ{} can be arbitrary, it can be constant but should be non-increasing \cite[Theorem 4.2]{Bubeck12}.
+The sequence of non-negative learning rates $(\eta_t)_{t \geq 1}$ used by \ExpQ{} can be arbitrary.
+It can be constant but should be non-increasing \cite[Theorem 4.2]{Bubeck12}.
 % helps to choose a good sequence.
 If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t = \eta := 2 \log(N) / (T K)$.
-However, for real-world communication problems, it is unrealistic to assume a fixed and known time horizon, so we prefer the alternative horizon-free choice%
-\footnote{~It is non-increasing, and is obtained by minimizing the upper-bound on the contextual regret derived in \cite[pp48]{Bubeck12}.}
+However, for real-world communication problems, it can be considered unrealistic\footnote{~For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}.} to assume a fixed and known time horizon, so we prefer the alternative horizon-free choice%
+\footnote{~It is non-increasing, and is obtained by minimizing the upper-bound on the regret derived in \cite[pp48]{Bubeck12}.}
 of learning rates,
@@
 We focus on \emph{i.i.d.} MAB problems, with $K = 9$ channels\footnote{~Similar behaviors are observed for any not-too-large values of $K$, we tried up-to $K = 100$ and the same results were obtained.}.
-For Bernoulli problem, the first one uses $\boldsymbol{\mu}=[0.1,\dots,0.9]$,
-and the second one is divided in three groups:
-2 very bad arms ($\mu = 0.01, 0.02$), 5 average arms ($\mu = 0.3$ to $0.6$) and 3 very good arms ($\mu = 0.78, 0.8, 0.82$).
+For Bernoulli problem, the first one uses $\boldsymbol{\mu}=[0.1,\dots,0.9]$, and is considered as a ``simple'' problem.
+The second one is divided in three groups:
+2 very bad arms ($\mu = 0.01, 0.02$), 5 average arms ($\mu = 0.3$ to $0.6$) and 3 very good arms ($\mu = 0.78, 0.8, 0.82$), and it is considered as a ``harder'' problem.
 The horizon is set to $T = 20000$ (but its value is unknown to all algorithms), and simulations are repeated $1000$ times, to estimate the expected regret.
@@
 using the parameter-free learning rate sequence, $\eta_t := \log{N} / (t K)$.
-The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted, and it is crucial to note that it is only asymptotic and to not be surprised by having regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem).
+The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted, and it is crucial to note that it is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli}).
 %
 Note that for each of the $1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{1\leq k \leq K, 1 \leq t \leq T}$ for every repetition, in order to compare the algorithms on the same realizations%
-\footnote{Similar plots and similar results are obtained if this trick is not used, but it makes more sense to compare them against the same randomization.}
+\footnote{Similar plots and similar results are obtained if this choice is not made, but it makes more sense to compare them against the same randomization.
+Note that this choice is the default in SMPyBandits, but the other possibility is implemented as well, by changing \texttt{cache\_rewards} to \texttt{true} or \texttt{false} in the \texttt{configuration} dictionary.
+See Section~\ref{sub:3:miniConfigurationFileSMPyBandits}.}
 of the MAB problem.
@@
 
-The \Aggr{} does not have satisfying theoretical guarantees in terms of regret $R_T$ yet, unlike many bandit algorithms.
+The \Aggr{} does not have satisfying theoretical guarantees in terms of regret $R_T$, unlike many bandit algorithms.
 %
@@
 between the aggregation algorithm $\Alg_{\mathrm{aggr}}$ and the best aggregated algorithm $\Alg_a$.
-This is in contrast with the (classical) regret, which measure the difference with the best fixed-arm strategy (Eq.~\eqref{eq:5:classicalregret}).
+This is in contrast with the (classical) regret, which measure the difference with the best fixed-arm strategy (Definition~\ref{def:2:regret}).
 Thus, even if the aggregated algorithms have logarithmic (classical) regret, having an adversarial regret scaling as $\sqrt{T}$ does not permit to exhibit a logarithmic (classical) regret for the aggregation algorithm.
@@
 
-Our \Aggr{} algorithm can probably be rewritten as an Online Mirror Descent, as \ExpQ{} and \CORRAL,
-but this does not appear useful as in the case of \CORRAL{}  the analysis cannot bring a logarithmic bound on the regret even by aggregating asymptotically optimal algorithms.
-We will continue investigating regret bounds for \Aggr,
-and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).
+Our \Aggr{} algorithm could be rewritten as an Online Mirror Descent \cite{Hazan2016introduction,Zimmert2018}, as it is the case for both \ExpQ{} and \CORRAL.
+But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
+% We will continue investigating regret bounds for \Aggr,
+% and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).
 
-\paragraph{A note on the simulation code}
-The page \texttt{SMPyBandits.GitHub.io/Aggregation.html} explains how to reproduce the experiments used for this Section.
+\paragraph{Reproducility of the experiments.}
+%
+Experiments presented in this Section are based on our library SMPyBandits, and instructions to reproduce them are given on the page \texttt{SMPyBandits.GitHub.io/Aggregation.html}.
 

### 2019-05-22 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
On se fait jeter dehors !

@@
 
-For any real-world applications of MAB algorithms, several solutions have been explored based on various models.
-However, it is hard to exactly predict which solution could be the best for real-world conditions at every instants.
-As we saw before, a possible approach is to pick one algorithm and simply use it, or to run simulations before-hand, if possible.
-But a third possible approach is to select \emph{on the run} the best algorithm for a specific situation, for which the so-called \emph{expert aggregation algorithms} can be useful.
+As we said, a third possible approach is to select \emph{on the run} the best algorithm for a specific situation, for which the so-called \emph{expert aggregation algorithms} can be useful.
 %
-To the best of our knowledge, aggregation algorithms, such as \ExpQ{} dating back from 2002 (\cite{Auer02}),
-have never been used for stochastic MAB problems,
-and we show that it appears empirically sub-efficient when applied to simple stochastic problems.
-%
-In this section, we present an improved variant of \ExpQ, called \Aggr.
-For synthetic MAB problems,
-with Bernoulli distributions or other distributions for arms,
-simulation results are presented to demonstrate its empirical efficiency.
+To the best of our knowledge, aggregation algorithms, such as \ExpQ{} which dates back from 2002 \cite{Auer02},
+have never been used for stochastic MAB problems.
+% and we show that it appears empirically sub-optimal when applied to simple stochastic problems.
+
+We present an improved variant of \ExpQ, called \Aggr.
+For synthetic MAB problems, with Bernoulli distributions or other distributions for $K$ arms, simulation results are presented to demonstrate its empirical efficiency.
 We combine classical algorithms, such as Thompson sampling, Upper-Confidence Bounds algorithms (\UCB{} and variants), and Bayesian or Kullback-Leibler UCB.
@@
 Our algorithm offers good performance compared to state-of-the-art algorithms
-(\ExpQ{}, \CORRAL{} or \LearnExp{}) \cite{Agarwal16,Singla17}),
+(\ExpQ{}, \CORRAL{} or \LearnExp{}) \cite{Agarwal16,Singla17},
 and appears as a robust approach to select on the run the best algorithm for any stochastic MAB problem, being more realistic to real-world radio settings than any tuning-based approach.
 
-\TODOL{This chapter is basically a raw include from my paper ``Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access'', see https://hal.inria.fr/hal-01705292}
+% \TODOL{This chapter is basically a raw include from my paper ``Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access'', see https://hal.inria.fr/hal-01705292}
 
 % ----------------------------------------------------------------------
-\subsection{Motivation for online algorithm selection}\label{sub:25:introduction}
-
-Many different learning algorithms have been proposed by the machine learning community,
-and most of them depend on several parameters, for instance $\alpha$ for \UCB, the prior for Thompson sampling or BayesUCB,
-the $\kl$ function for \klUCB{} etc.
-Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T = \E_{\boldsymbol{\mu}}[\widetilde{R_T}]$.
-%
-For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms.
-But when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any kind, size and complexity:
-ideally one would like to use an algorithm that can be applied identically against any problem.
-To choose the best algorithm, two approaches can be followed:
-either extensive benchmarks are done beforehand -- if this is possible -- to select the algorithm and its optimal parameters, or an adaptive algorithm is used to learn \emph{on the fly} its parameters.
-We present a simple adaptive solution, that aggregates several learning algorithms in parallel and adaptively chooses which one to trust the most.
-
-\paragraph{Outline}
-This section is organized as follows.
-First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
-Then we present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
-Finally, we present numerical experiments in Section~\ref{sub:25:numExp},
-on Bernoulli and non-Bernoulli MAB problems,
-comparing the regret of several algorithms against different aggregation algorithms.
-Theoretical guarantees are shortly discussed in Section~\ref{sub:25:theory}, and Section~\ref{sub:25:conclusion} concludes.
-
-% ----------------------------------------------------------------------
-\subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
+% \subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
+\paragraph{Aggregating bandit algorithms}\label{sub:25:aggregation}
 
@@
 one arm is chosen with probability $1$ and the others with probability $0$.
-However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $[K]$ at each time step, and draw the next selected arm according to this distribution.
+However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $[K]$ at each time step, and draw the next selected arm according to this distribution
+(\eg, Thompson sampling).
 
@@
 % ----------------------------------------------------------------------
-\subsubsection{The \Aggr{} algorithm}\label{sub:25:Aggr}
+\subsection{Our contribution: the \Aggr{} algorithm}\label{sub:25:Aggr}
 
@@
 Instead
-\footnote{It is the same in the case of constant learning rates, and does not differ much even for decreasing learning rates.}
+\footnote{~It is the same in the case of constant learning rates, and does not differ much even for decreasing learning rates.}
 of updating $\pi^{t}$ multiplicatively as we do for our proposal, \ExpQ{} recomputes it, proportionally to
@@
 
-\TODOL{Je pense refaire complètement les expériences de ce morceau.
+% \TODOL{Je pense refaire complètement les expériences de ce morceau.
 
-- C'est un peu bête de montrer des problèmes chelous genre Gaussien et "mixed" quand tout le reste de la thèse s'intéresse vraiment aux Bernoulli.
+% - C'est un peu bête de montrer des problèmes chelous genre Gaussien et "mixed" quand tout le reste de la thèse s'intéresse vraiment aux Bernoulli.
 
-- C'est un peu naze de montrer en log-y ou en log-log, alors que le reste de la thèse s'intéresse uniquement à des plots x-y, ou des histogrammes.
+% - C'est un peu naze de montrer en log-y ou en log-log, alors que le reste de la thèse s'intéresse uniquement à des plots x-y, ou des histogrammes.
 
-- Pour être plus consistant avec les sections suivantes, je pourrais inclure juste 1/2 courbes de regret, et un tableau de résultats pour d'autres problèmes ?
-}
+% - Pour être plus consistant avec les sections suivantes, je pourrais inclure juste 1/2 courbes de regret, et un tableau de résultats pour d'autres problèmes ?
+% }
 
@@
 Note that this choice is the default in SMPyBandits, but the other possibility is implemented as well, by changing \texttt{cache\_rewards} to \texttt{true} or \texttt{false} in the \texttt{configuration} dictionary.
-See Section~\ref{sub:3:miniConfigurationFileSMPyBandits}.}
+See Section~\ref{sub:3:miniConfigurationFileSMPyBandits} for explanations on this \texttt{configuration.py} file, or the online documentation.}
 of the MAB problem.
@@
 % ----------------------------------------------------------------------
-\subsection{No theoretical guarantees yet}\label{sub:25:theory}
+\paragraph{No theoretical guarantees yet}\label{sub:25:theory}
 
@@
 % ----------------------------------------------------------------------
-\subsection{Conclusion about our \Aggr{} algorithm}\label{sub:25:conclusion}
+\paragraph{Conclusion about our \Aggr{} algorithm}\label{sub:25:conclusion}
 
@@
 %
-Experiments presented in this Section are based on our library SMPyBandits, and instructions to reproduce them are given on the page \texttt{SMPyBandits.GitHub.io/Aggregation.html}.
+Experiments presented in this Section are based on our library SMPyBandits,
+and the page \texttt{SMPyBandits.GitHub.io/Aggregation.html} gives instructions to reproduce them.
 

### 2019-05-22 2-Chapters/2-Chapter/chapter2.tex
My thesis is so long, and so stupid
Thanks Christophe for comments on Ch2
On se fait jeter dehors !

@@
 \minitoc
+% Write miniTOC just after the title
+\graphicspath{{2-Chapters/2-Chapter/Images/}}
 
@@
 %
-Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on the article \cite{Besson2018WCNC}.
+The Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on the article \cite{Besson2018WCNC}.
 
@@
 
-\newpage
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/2-Chapter/Images/}}
+% \newpage
 
@@
     \centering
-    \resizebox{0.350\textwidth}{!}{
+    \resizebox{0.50\textwidth}{!}{
         \begin{tikzpicture}[auto, node distance=5cm, >=latex]
@@
 \paragraph{An interactive demo to discover the MAB problem (for the novice).}
+\label{par:2:interactiveDemoDiscoverMAB}
 %
@@
 
-\paragraph{On one-dimensional exponential family.}
+\paragraph{We restrict to one-dimensional exponential family.}
 %
@@
 -- A second important result is a simplified form for the \emph{Kullback-Leibler divergence} \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
-The KL divergence between two distributions $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ is defined as
+The KL divergence is also called the relative entropy is a measure of how one probability distribution is different from a second, reference probability distribution.
+
+\begin{defn}[Kullback-Leibler divergence]\label{def:2:KLDivergence}
+    The KL divergence between two distributions $d_1$ and $d_2$ is defined as
     \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
-    \KL\left(\Pr(x|\eta_1), \Pr(x|\eta_2)\right) = \int \Pr(x|\eta_1) \log\left(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}\right) \right].
+        \KL\left(d_1, d_2\right) = \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right].
     \end{equation}
-%
-So this KL divergence can be written $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, and it can be simplified to an expression using only the two parameters $\eta_1$ and $\eta_2$,
+\end{defn}
+
+Following this Definition~\ref{def:2:KLDivergence} for two distributions
+$d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$,
+we can write $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$,
 and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$:
@@
     \textbf{Bernoulli distributions} can be seen as an exponential family with $h(x) = 1$,
-    and for a Bernoulli distribution of mean $\mu\in[0,1]$,
+    and for a Bernoulli distribution of mean $\mu\in[0,1]$, denoted $B(\mu)$,
     the parameter is $\eta = \mu / (1 - \mu)$, giving $A(\eta) = \log(1 + \mathrm{e}^{\eta})$
     (with the limit behavior $\eta=+\infty$ if $\mu=0$).
+    %
+    Moreover, the KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
+    $\kl(x,y) = \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$.
 
@@
     $h(x) = 1/\sqrt{2\pi\sigma^2} \exp(-x^2/(2\sigma^2))$,
-    and for a Gaussian distribution with mean $\mu\in\R$ (\ie, $\cN(\mu,\sigma^2)$),
+    and for a Gaussian distribution with mean $\mu\in\R$, denoted $\cN(\mu,\sigma^2)$,
     the parameter is $\eta = \mu/\sigma^2$, giving $A(\eta) = \mu^2/(2\sigma^2)$
@@
     % with a fixed variance $\sigma^2$ form indeed a one-dimensional exponential family.
+    %
+    Moreover, the KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
+    $d(x,y) = \KL_{\cN}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$.
 \end{itemize}
 
-In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions.
+In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the application of choice.
 First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
@@
 %
-Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm in the setting on bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use Bernoulli Kullback-Leibler divergences in our analysis.
+Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm in the setting on bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use the Bernoulli Kullback-Leibler divergence in our analysis.
 
@@
 % ----------------------------------------------------------------------------
-\section{Applications of stochastic MAB}
+\section{Different applications of stochastic MAB}
 \label{sec:2:applicationsofStochasticMAB}
@@
 
-The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc ).
+The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc).
 In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
@@
 
-If the goal of the player is to maximize $\E[\sum_{t=1}^T r_{A(t)}(t)]$,
+If the goal of the player is to maximize $\E\left[\sum_{t=1}^T r_{A(t)}(t)\right]$,
 the optimal strategy for this bandit problem is to always pull an optimal arm, but of course it is unrealistic as the player does not know the true means and cannot know which arm is optimal.
@@
 
+\textbf{Applications:}
+%
 Such decomposition of the regret is useful for at least two reasons.
@@
 %
-On the second hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T$ and not directly the regret $R_T^{\cA}$.
+On the second hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
 For example, we extend this decomposition to the multi-player case in Chapter~\ref{chapter:3}, in Lemma~\ref{lem:5:DecompositionRegret}, and as it also uses the draws $N_k(T)$, it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
@@
 
-In all this section, we restrict to stationary stochastic bandits with $K\geq2$ arms.
+In all this section, we restrict to stationary stochastic problems with $K\geq2$ arms.
 We do not give proofs of the following theorems, as they can be found in the historical papers, and simpler proofs are given in recent references, such as \cite{Bubeck12}, \cite{LattimoreBanditAlgorithmsBook}, or Chapter~2 in \cite{Slivkins2019} for instance.
-
+%
 Let $\cI$ denote the set of all problem instances, with $K \geq 2$ arms whose rewards lie in $[0,1]$ (and no other hypothesis).
@@
 
-\subsection{Summary of classification of the most famous algorithms}
+% \subsection{Summary of classification of the most famous algorithms}
 
-A huge table in one page in portrait mode, giving:
-- name
-- reference, year
-- family, classification
-- time/memory complexity ?
-- regret bound for stochastic regret
+% A huge table in one page in portrait mode, giving:
+% - name
+% - reference, year
+% - family, classification
+% - time/memory complexity ?
+% - regret bound for stochastic regret
 
-of about 20/30 algorithms?
+% of about 20/30 algorithms?
 
@@
 
-\subsection{Naive approach: use UCB or Thompson sampling, and forget about the rest!}
-First solution: be naive, only use UCB everywhere
+For any real-world applications of MAB algorithms,
+several solutions have been explored based on various models.
+Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which model and which algorithms to use.
+It is hard to predict which solution could be the best for real-world conditions at every instants,
+and even by assuming a stationary environment, when one is facing a certain problem but has limited information about it, it is hard to know beforehand which algorithm can be the best solution.
 
-\subsection{Use prior knowledge about the future application to select beforehand the chosen algorithm}
-Second solution, if one has some prior knowledge about the domain or setting for which the learning algorithms will be deployed, she can run synthetic or real-world simulations, compare many algorithms before deployment, and select the best algorithm!
+In this Section, we first present two naive approaches for selecting an algorithm when facing a new problem, then we detail the online approach that uses a ``master'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``master'' algorithm based on \ExpQ.
+This section is based on our article \cite{Besson2018WCNC}.
+
+% ----------------------------------------------------------------------
+\subsection{Motivation for online algorithm selection}\label{sub:25:introduction}
+
+Many different learning algorithms have been proposed by the machine learning community,
+and most of them depend on several parameters, for instance $\alpha$ for \UCB, the prior for Thompson sampling or BayesUCB,
+the $\kl$ function for \klUCB{} etc.
+Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T = \E_{\boldsymbol{\mu}}[\widetilde{R_T}]$.
+%
+For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms.
+But when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any kind, size and complexity:
+ideally one would like to use an algorithm that can be applied identically against any problem.
+
+\paragraph{Naive approaches.}
+%
+On the first hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or because it is simple enough to be used in its application.
+For instance, it can be hard to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and so a practitioner could chose to only consider the \UCB{} algorithm (or other low-cost algorithm).
+%
+On the other hand, if prior knowledge on the application at hand is available, one could implement some benchmarks, and compare a set of algorithms on different problems.
+If a leader appears clearly, it is then possible to choose it for the application.
+
+\paragraph{Small example.}
+%
+For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of a \UCB{} algorithms will perform quite differently in the two problems:
+using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case,
+while using a low $\alpha$, \ie, favoring exploitation, will give low regret in the second case.
+One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of \UCB, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best for the problem at hand.
+
+\paragraph{The online approach.}
+%
+Another possibility is to consider an online approach, which is interested in the case where the computation power or memory storage of the application is not a limitation factor, but where one cannot run benchmarks before deploying the application.
+We consider a fixed set of algorithms, and we use another learning algorithm on top of this set, to learn \emph{on the fly} which one should be trusted more, and eventually, used on its own.
+
+The aggregation approach is especially interesting if we know that the problem the device will face is one of a few known kinds of problems.
+In such cases, if there is $N$ different sorts of problems, and if one has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
+and use the aggregation of $\cA_1,\dots,\cA_N$ when facing the unknown problem.
+
+% To choose the best algorithm, three approaches can be followed:
+% \begin{itemize}
+%     \item
+% \end{itemize}
+% either extensive benchmarks are done beforehand -- if this is possible -- to select the algorithm and its optimal parameters, or an adaptive algorithm is used to learn \emph{on the fly} its parameters.
+
+% We present a simple adaptive solution, that aggregates several learning algorithms in parallel and adaptively chooses which one to trust the most.
+
+% \paragraph{Outline}
+% This section is organized as follows.
+% First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
+% Then we present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
+% Finally, we present numerical experiments in Section~\ref{sub:25:numExp},
+% on Bernoulli and non-Bernoulli MAB problems,
+% comparing the regret of several algorithms against different aggregation algorithms.
+% Theoretical guarantees are shortly discussed in Section~\ref{sub:25:theory}, and Section~\ref{sub:25:conclusion} concludes.
+
+% \subsection{Naive approaches}
+
+% As mentioned, there are two naive approaches
+
+% use UCB or Thompson sampling, and forget about the rest!}
+% First solution: be naive, only use UCB everywhere
+
+% \subsection{Use prior knowledge about the future application to select beforehand the chosen algorithm}
+
+% Second solution, if one has some prior knowledge about the domain or setting for which the learning algorithms will be deployed, she can run synthetic or real-world simulations, compare many algorithms before deployment, and select the best algorithm!
 
@@
 
-The first contribution of this manuscript \cite{Besson2018WCNC} concluded this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, and corresponds to our article \cite{Besson2018WCNC}.
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
@@
 % \TODOL{Write this ?}
-
-% \subsection{Efficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
-
-% \TODOL{On s'en fout en fait de ça, non ? Ma thèse n'a pas étudiée klUCB, et les implémentations en C ne sont pas de moi mais viennent de pymaBandits... Donc inutile de parler de ça !}
-
-% I want to detail the maths behind an implementation of klUCB, see https://smpybandits.github.io/docs/Policies.kullback.html and all what is detailed there.
-
-% How to write a fast version of klBern and klucbBern (the two most used functions), using C or Numba or Cython, maybe I can include the speed up simulations here?
-First piece of code, before really presenting SMPyBandits ?
-
-% My work for this part is also published on https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes
-% (in Python) and https://github.com/Naereen/KullbackLeibler.jl (in Julia).

### 2019-05-22 2-Chapters/2-Chapter/chapter2.tex
Typos in refs and labels

@@
 Tracking this quantity is used for instance in the OSSB algorithm proposed in \cite{Combes17}.
-It states that FIXME interpret!
-FIXME explain why this is interesting for the rest of this manuscript!
+% FIXME explain why this is interesting for the rest of this manuscript!
 
@@
 We naively have that $R_T^{\cA} \leq (\max_k \Delta_k) T$, and thus if 
+\TODOL{Finish this!}
 
@@
 
-It states that FIXME interpret!
-FIXME explain why this is interesting for the rest of this manuscript!
+\TODOL{It states that FIXME interpret! FIXME explain why this is interesting for the rest of this manuscript!}
 
@@
 
-\subsection{Regret upper bounds, tools and different kinds of bounds}
+% \subsection{Regret upper bounds, tools and different kinds of bounds}
 
-\begin{itemize}
-    \item
-    - different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
-    \item
-    - tools to show regret upper-bounds, concentration inequalities?
-    FIXME non j'ai pas la place
-    \item
-    - short proof of UCB or klUCB? just to start to show the tools
-    FIXME non j'ai pas la place
-\end{itemize}
+% \begin{itemize}
+%     \item
+%     - different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
+%     \item
+%     - tools to show regret upper-bounds, concentration inequalities?
+%     FIXME non j'ai pas la place
+%     \item
+%     - short proof of UCB or klUCB? just to start to show the tools
+%     FIXME non j'ai pas la place
+% \end{itemize}
 
@@
 
-In this thesis, we only study analytically the mean regret $R^{\cA}$ (in Chapters \ref{chapter:5,chapter:6}), but we quickly mention other measures of performances used in our work.
+In this thesis, we only study analytically the mean regret $R^{\cA}$ (in Chapters \ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
 
@@
 % - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-The switching cost $SC^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $SC^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
+The \emph{switching cost} $SC^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $SC^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
 It has recently gained interest in the literature, for instance in \cite{Koren17}.
@@
 
+\TODOL{Expliquer UCB}
+\TODOL{Expliquer que les indices UCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
+
+\TODOL{Expliquer klUCB}
+\TODOL{Expliquer que les indices klUCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
+
 \subsubsection{UCB}
@@
 Optimal upper confidence bounds based on Kullback-Leibler divergence and klUCB indexes
+
+\TODOL{Short note on how to implement such policy, using Newton's method ?}
 

### 2019-05-22 2-Chapters/3-Chapter/chapter3.tex
My thesis is so long, and so stupid
Thanks Christophe for comments on Ch2
Typos in refs and labels

@@
 \minitoc
+
+\TODOL{Intégrer les commentaires de Christophe, le plus rapidement possible !}
 
@@
 %
-For instance the following code creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate sthe problem.
+For instance the following code in Listing~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate sthe problem.
 We give another example for truncated Gaussian distributions, and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
@@
     \end{minted}
-    \caption{Small snippet of Bash code to run a simple experiment with SMPyBandits}
-    % \captionof{listing}{Small snippet of Bash code to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
-    \label{lst:3:howToRunBasicLibrary}
+    \caption{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits}
+    % \captionof{listing}{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits.
+\label{lst:3:pythonCodeCreateProblem}}
+    \label{lst:3:pythonCodeCreateProblem}
 \end{listing}
@@
 \subsection{Minimalist example of use of SMPyBandits}
+\label{sub:3:miniExampleSMPyBandits}
 
@@
 \subsection{A minimalist \texttt{main.py} file}
+\label{sub:3:miniMainFileSMPyBandits}
 
@@
 \subsection{A minimalist \texttt{configuration.py} file}
+\label{sub:3:miniConfigurationFileSMPyBandits}
 
@@
 
+% \subsection{Efficient numerical computation of Kullback-Leibler divergences and \klUCB{} indexes}
+
+% \TODOL{On s'en fout en fait de ça, non ? Ma thèse n'a pas étudiée klUCB, et les implémentations en C ne sont pas de moi mais viennent de pymaBandits... Donc inutile de parler de ça !}
+
+% I want to detail the maths behind an implementation of klUCB, see https://smpybandits.github.io/docs/Policies.kullback.html and all what is detailed there.
+
+% How to write a fast version of klBern and klucbBern (the two most used functions), using C or Numba or Cython, maybe I can include the speed up simulations here?
+First piece of code, before really presenting SMPyBandits ?
+
+% My work for this part is also published on https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes
+% (in Python) and https://github.com/Naereen/KullbackLeibler.jl (in Julia).
+
 % --------------------------------------------------------------

### 2019-05-22 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Typos in refs and labels

@@
 Letting $\Delta = |\mu_0 - \mu_1|$, one can essentially establish that for $\tau$ larger than $(1/\Delta^2)\ln(1/\delta)$ (\ie, enough samples before the change), the delay can be of the same magnitude (\ie, enough samples after the change).
-In the bandit analysis to follow, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and \ref{lem:6:GoodEvent}).
+In the bandit analysis to follow, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and the corresponding Lemma not included here, but given in Appendix of \cite{Besson2019GLRT}).
 

### 2019-05-22 2-Chapters/1-Chapter/chapter1.tex
Almost done for refactoring this part of Ch2 from WCNC18

@@
 
-The reading order of the manuscript can be any top-down path between Chapter~\ref{chapter:1} and the Conclusion in Chapter~\ref{chapter:conclusion}, in the graph shown in Figure~\ref{fig:1:organization} above,
-and the thesis is organized in two parts:
+The reading order of the manuscript can be any top-down path between the Introduction in the current Chapter~\ref{chapter:1}, and the Conclusion in the last Chapter~\ref{chapter:conclusion}.
+As the graph in Figure~\ref{fig:1:organization} shows,
+the thesis is organized in two parts:
 

### 2019-05-22 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Almost done for refactoring this part of Ch2 from WCNC18

@@
 
-As we said, a third possible approach is to select \emph{on the run} the best algorithm for a specific situation, for which the so-called \emph{expert aggregation algorithms} can be useful.
+As we said, a third possible approach is to select \emph{on the run} the best algorithm for a specific situation, for which the so called \emph{expert aggregation algorithms} framework can be useful.
 %
 To the best of our knowledge, aggregation algorithms, such as \ExpQ{} which dates back from 2002 \cite{Auer02},
-have never been used for stochastic MAB problems.
+have never been used in pratice for stochastic MAB problems.
 % and we show that it appears empirically sub-optimal when applied to simple stochastic problems.
@@
 We present an improved variant of \ExpQ, called \Aggr.
-For synthetic MAB problems, with Bernoulli distributions or other distributions for $K$ arms, simulation results are presented to demonstrate its empirical efficiency.
+For synthetic MAB problems, with Bernoulli or other distributions for $K$ arms, simulation results are presented to demonstrate its empirical efficiency.
 We combine classical algorithms, such as Thompson sampling, Upper-Confidence Bounds algorithms (\UCB{} and variants), and Bayesian or Kullback-Leibler UCB.
@@
 % \subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
-\paragraph{Aggregating bandit algorithms}\label{sub:25:aggregation}
+\paragraph{Notations}\label{sub:25:aggregation}
 
@@
 			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$}
-			\KwData{Initial uniform distribution, $\pi^{0} = U([N])$}
+			\KwData{Initial uniform distribution, $\pi^{0} = \cU([N])$}
 			\KwResult{$\Alg_{\mathrm{aggr}}=\Aggr\left[\Alg_1, \dots, \Alg_N\right]$}
@@
 				\For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
-					$\Alg_a$ updates its indexes (e.g., \UCB{} indexes)\;
-					It chooses $A^{t+1}_{a} \in [K]$.
+					$\Alg_a$ updates its internal state (e.g., \UCB{} indexes)\;
+					It chooses $A(t+1)_{a} \in [K]$.
 				}
@@
 				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
-				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A^{t+1}_{a} = j\}), \;\forall 1\leq j \leq K$\;
+				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A(t+1)_{a} = j\}), \;\forall 1\leq j \leq K$\;
 				% \tcp*{Sum trusts}
-				Then $\Alg_{\mathrm{aggr}}$ chooses arm $A^{t+1} \sim p^{t+1}$\;
-				% $\P(A^{t+1}_{\mathrm{aggr}} = j) = p^{t+1}_j$
+				Then $\Alg_{\mathrm{aggr}}$ chooses arm $A(t+1) \sim p^{t+1}$\;
+				% $\P(A(t+1)_{\mathrm{aggr}} = j) = p^{t+1}_j$
 				% \tcp*{Sample once}
 				%
-				Give \emph{original} reward $(A^{t+1}, 1 - \ell_{A^{t+1}}(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
-				Compute an \emph{unbiased} estimate of the loss of the trusted algorithms,$$ \ell^{t+1} = \ell_{A^{t+1}}(t+1) / p^{t+1}_{A^{t+1}}\;. $$
+				Give \emph{original} reward $(A(t+1), 1 - \ell_{A(t+1)}(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
+				Compute an \emph{unbiased} estimate of the loss of the trusted algorithms,$$ \ell^{t+1} = \ell_{A(t+1)}(t+1) / p^{t+1}_{A(t+1)}\;. $$
 				% \tcp{Use bandit feedback to update $\pi^{t}$}
-				% \tcp{Finally, use the instantaneous bandit feedback (loss $\ell_{A^{t+1},t}$) to update $\pi^{t}$}
+				% \tcp{Finally, use the instantaneous bandit feedback (loss $\ell_{A(t+1),t}$) to update $\pi^{t}$}
 				% \For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
 				\For{$a = 1, \dots, N$}{
-					% \eIf{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
-					\If{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
+					% \eIf{$\Alg_a$ was trusted, \ie, $A(t+1)_{a} = A(t+1)$}{
+					\If{$\Alg_a$ was trusted, \ie, $A(t+1)_{a} = A(t+1)$}{
 						$ \pi^{t+1}_{a} = \exp(\eta_t \ell^{t+1}) \times \pi^{t}_{a} $
@@
 
-At every time step, after having observed a loss $\ell_{A^{t+1}}(t+1)$ for its chosen action $A^{t+1}$,
+At every time step, after having observed a loss $\ell_{A(t+1)}(t+1)$ for its chosen action $A(t+1)$,
 the algorithm updates the trust probabilities from $\pi^t$ to $\pi^{t+1}$ by
@@
 Only the algorithms $\Alg_a$ who advised the last decision get their trust updated, in order to trust more the ``reliable'' algorithms.
+
 The loss estimate is unbiased in the following sense.
-If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\tilde{\ell}_{a,t+1} = \ell_{A_a^{t+1}}(t+1)$.
-This quantity can only be observed for those algorithms for which $A_{a}^{t+1}=A^{t+1}$.
+If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\tilde{\ell}_{a,t+1} = \ell_{A_A(t+1)}(t+1)$.
+This quantity can only be observed for those algorithms for which $A_{a}^{t+1}=A(t+1)$.
 However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $\tilde{\ell}_{a,t}$.
-More precisely,
-\[\hat{\ell}_{a,t+1} := \frac{\ell_{A_a^{t+1}}(t+1)}{p_{A_a^{t+1}}^{t+1}} \mathbbm{1}(A_a^{t+1} = A^{t+1})\] satisfies $\mathbb{E}[\hat{\ell}_{a,t+1} | \mathcal{H}_t] = \tilde{\ell}_{a,t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$.
-Observe that $\tilde{\ell}_{a,t+1}=\ell^{t+1}$ for all algorithms $a$ such that $A_a^{t+1}=A^{t+1}$, and $\tilde{\ell}_{a,t+1}=0$ otherwise.
+More precisely, if we define the estimate by
+\begin{equation}
+	\hat{\ell}_{a,t+1} = \frac{\ell_{A_A(t+1)}(t+1)}{p_{A_A(t+1)}^{t+1}} \mathbbm{1}(A_A(t+1) = A(t+1))
+\end{equation}
+satisfies $\mathbb{E}[\hat{\ell}_{a,t+1} | \mathcal{H}_t] = \tilde{\ell}_{a,t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$.
+Observe that $\tilde{\ell}_{a,t+1}=\ell^{t+1}$ for all algorithms $a$ such that $A_A(t+1)=A(t+1)$, and $\tilde{\ell}_{a,t+1}=0$ otherwise.
 
 % - Precise some tricky points
-An important feature of \Aggr{} is the feedback provided to each underlying bandit algorithm, upon the observation of arm $A^{t+1}$.
-Rather than updating only the trusted algorithms (that is the algorithms which would have drawn arm $A^{t+1}$) with the observed reward $r_{A^{t+1}}(t+1)=1 - \ell_{A^{t+1}}(t+1)$,  we found that updating each algorithm with the (original) loss observed for arm $A^{t+1}$ improves the performance drastically.
+An important feature of \Aggr{} is the feedback provided to each underlying bandit algorithm, upon the observation of arm $A(t+1)$.
+Rather than updating only the trusted algorithms (that is the algorithms which would have drawn arm $A(t+1)$) with the observed reward $r_{A(t+1)}(t+1)=1 - \ell_{A(t+1)}(t+1)$,  we found that updating each algorithm with the (original) loss observed for arm $A(t+1)$ improves the performance drastically.
 %
@@
 and an alternative would be to
-increase $\pi_{a}$ if $A^{t+1}_a = A^{t+1}$ and to decrease it otherwise.
+increase $\pi_{a}$ if $A(t+1)_a = A(t+1)$ and to decrease it otherwise.
 It would not be so different, as there is a final renormalization step, and empirically this variation has little impact on the performance of \Aggr{}.
@@
 Instead
-\footnote{~It is the same in the case of constant learning rates, and does not differ much even for decreasing learning rates.}
+\footnote{~It is the same for constant learning rates, and does not differ much for decreasing learning rates.}
 of updating $\pi^{t}$ multiplicatively as we do for our proposal, \ExpQ{} recomputes it, proportionally to
@@
 % helps to choose a good sequence.
-If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t = \eta := 2 \log(N) / (T K)$.
+If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t = \eta = 2 \log(N) / (T K)$.
 However, for real-world communication problems, it can be considered unrealistic\footnote{~For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}.} to assume a fixed and known time horizon, so we prefer the alternative horizon-free choice%
@@
 of learning rates,
-$\eta_t := \log(N) / (t K)$ suggested by \cite{Bubeck12}.
+$\eta_t = \log(N) / (t K)$ suggested by \cite{Bubeck12}.
 We compare both approaches empirically, and the second one usually performs better.
@@
 This empirical estimation of the expected regret $R_T$ is plotted below, as a function of $T$, comparing some algorithms $\Alg_1,\dots,\Alg_N$ (for $N=6$), and their aggregation with \Aggr{} (displayed in orange bold),
-using the parameter-free learning rate sequence, $\eta_t := \log{N} / (t K)$.
-The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted, and it is crucial to note that it is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli}).
-%
+using the parameter-free learning rate sequence, $\eta_t = \log{N} / (t K)$.
+
+The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted.
+It corresponds to the second point \eqref{eq:2:forSecondLogTLowerBound2} in Theorem~\ref{thm:2:secondLogTLowerBound}.
+It is crucial to note that the lower-bound is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli}).
+
 Note that for each of the $1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{1\leq k \leq K, 1 \leq t \leq T}$ for every repetition, in order to compare the algorithms on the same realizations%
@@
 We compare our \Aggr{} algorithm,
-as well as other aggregation algorithms, \ExpQ{},
-\CORRAL{} and \LearnExp{} (both with default parameters) \cite{Bubeck12,Agarwal16,Singla17}.
-The aggregated algorithms consist in a naive uniform exploration (to have at least one algorithm with bad performances, \ie{} linear regret, but it is not included in the plots),
-\UCB{} with $\alpha=1/2$, three \klUCB{} with Bernoulli, Gaussian and exponential $\kl$ functions, and BayesUCB and Thompson sampling with uniform prior.
+as well as other aggregation algorithms, \ExpQ{} from \cite{Bubeck12},
+\CORRAL{} from \cite{Agarwal16} and \LearnExp{} from \cite{Singla17}, all with their default parameters.
+%
+The aggregated algorithms consist in a naive uniform exploration (to have at least one algorithm with bad performances, \ie, linear regret, but it is not included in the plots),
+\UCB{} with $\alpha=1/2$, three \klUCB{} algorithms respectively using Bernoulli, Gaussian and exponential $\kl$ functions, and BayesUCB and Thompson sampling with uniform prior.
 
@@
 	\centering
-	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
+	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
 	\caption{On a ``simple'' Bernoulli problem (semilog-$y$ scale).}
@@
 	\centering
-	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
+	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
 	\caption{On a ``harder'' Bernoulli problem, they all have similar performances, except \LearnExp.}
@@
 	\centering
-	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
+	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
 	\caption{On an ``easy'' Gaussian problem, only \Aggr{} shows reasonable performances, thanks to BayesUCB and Thompson sampling.}
@@
 	\centering
-	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
 	\caption{On a harder problem, mixing Bernoulli, Gaussian, Exponential arms, with 3 arms of each types with the \emph{same mean}.}
@@
 	\centering
-	\includegraphics[width=1.00\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
 	\caption{The semilog-$x$ scale clearly shows the logarithmic growth of the regret for the best algorithms and our proposal \Aggr, even in a hard ``mixed'' problem (\emph{cf}.
@@
 \ExpQ{} satisfies a bound on adversarial regret, % $\overline{R_T}$
-$\overline{R_T} \leq 2 \sqrt{T N \log{k}}$,
+$\overline{R_T} \leq 2 \sqrt{T N \log(K)$,
 with the good choice of the learning rate sequence $(\eta_t)_{t \geq 1}$.
@@
 
-This would be a first theoretical guarantee, but not satisfactory as simple algorithms like UCB have regrets scaling as $\log{T}$ \cite{Auer02,Bubeck12}, not $\sqrt{T}$.
+This would be a first theoretical guarantee, but not satisfactory as we saw above that simple algorithms (like \UCB) have regrets scaling as $\log(T)$ \cite{Auer02,Bubeck12}, not $\sqrt{T}$.
 %
@@
 % - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
-We realized experiments on simple MAB problems already used in the community of bandit algorithms for OSA \cite{Jouini10},
-and the simulations results showed that \Aggr{} works as expected, being able to identify on the fly the best algorithm to trust for a specific problem.
+We presented experiments on simple MAB problems coming from previous work on bandits for OSA \cite{Jouini10},
+and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
 Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.

### 2019-05-22 2-Chapters/2-Chapter/chapter2.tex
Almost done for refactoring this part of Ch2 from WCNC18

@@
 % First chapter begins here
-\chapter{The Stochastic Multi-Armed Bandits Model}
+\chapter{Stochastic Multi-Armed Bandits Models}
 \label{chapter:2}
@@
 % ----------------------------------------------------------------------------
-\section{The stochastic Multi-Armed Bandit model}
+\section{The stochastic Multi-Armed Bandit models}
 \label{sec:2:notations}
@@
     Using bandit algorithms for \textbf{improved machine learning} models or algorithms has been an active research domain for the last ten years or so.
-    As presented below in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, a certain ``master'' bandit algorithm can be used to select on the run the best bandit algorithm from a pool of ``slave'' algorithms.
+    As presented below in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, a certain ``leader'' bandit algorithm can be used to select on the run the best bandit algorithm from a pool of ``followers'' algorithms.
     Other possible use cases include hyper-parameter optimization, or feature selection.
@@
 For any real-world applications of MAB algorithms,
-several solutions have been explored based on various models.
-Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which model and which algorithms to use.
+several solutions have been explored based on various models, and for any model, typically there are many algorithms available, as we have seen above.
+%
+Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to use.
 It is hard to predict which solution could be the best for real-world conditions at every instants,
@@
 
-In this Section, we first present two naive approaches for selecting an algorithm when facing a new problem, then we detail the online approach that uses a ``master'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``master'' algorithm based on \ExpQ.
+In this Section, we first present two naive approaches for selecting an algorithm when facing a new problem, and then we detail the online approach that uses a ``leader'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``leader'' algorithm based on \ExpQ.
 This section is based on our article \cite{Besson2018WCNC}.
@@
 the $\kl$ function for \klUCB{} etc.
-Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T = \E_{\boldsymbol{\mu}}[\widetilde{R_T}]$.
+Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
 %
 For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms.
-But when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any kind, size and complexity:
-ideally one would like to use an algorithm that can be applied identically against any problem.
+But when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
+ideally one would like to use an algorithm that can be applied identically against any problem of such family.
 
@@
 %
-On the first hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or because it is simple enough to be used in its application.
-For instance, it can be hard to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and so a practitioner could chose to only consider the \UCB{} algorithm (or other low-cost algorithm).
+On the first hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
+It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the \UCB{} algorithm (or other low-cost algorithms).
 %
@@
 
-\paragraph{Small example.}
+\paragraph{Illustrative example.}
 %
@@
 
-The aggregation approach is especially interesting if we know that the problem the device will face is one of a few known kinds of problems.
-In such cases, if there is $N$ different sorts of problems, and if one has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
+The aggregation approach is especially interesting if we know that the problem the application will face is one of a few known kinds of problems.
+In such cases, if there are $N$ different sorts of problems, and if the practitioner has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
 and use the aggregation of $\cA_1,\dots,\cA_N$ when facing the unknown problem.
@@
 
-The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, and corresponds to our article \cite{Besson2018WCNC}.
+The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+% and corresponds to our article \cite{Besson2018WCNC}.
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.

### 2019-05-22 2-Chapters/1-Chapter/chapter1.tex
Small modifications

@@
 \paragraph{Abstract.}
+%
 We start by presenting the problems studied in this thesis,
@@
 We give an overview of our main contributions, alongside a list of publications written in the last three years.
+%
+The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but use the notations introduced on Chapter~\ref{chapter:2}.
 
@@
     at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.
+    \href{https://HAL.Inria.fr/hal-02049824}{\texttt{HAL.Inria.fr/hal-02049824}}.
     \cite{Bonnefoi2019WCNC}

### 2019-05-22 2-Chapters/2-Chapter/chapter2.tex
Small modifications

@@
 
-The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} above.
+The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
 The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r_{A(t)}(t)$.
@@
 One-dimensional exponential families include Bernoulli and Poisson distributions, as well as Gaussian distributions with a fixed variance.
-Our main interest is Bernoulli distributions, but we prefer to present here the notations on a generic exponential family in one dimension.
+Our main interest is Bernoulli distributions, but we prefer to present the notations of exponential families in one dimension.
 % We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Emilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
@@
     (with the limit behavior $\eta=+\infty$ if $\mu=0$).
-    %
+
     Moreover, the KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
@@
     % with a fixed variance $\sigma^2$ form indeed a one-dimensional exponential family.
-    %
+
     Moreover, the KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
@@
 Such Markov models come in two flavors: rested or restless.
-For $K$ arms, each Markov chain has a finite number of state $s$, each corresponding to a (constant) reward that the player obtains if she selects this arm while its Markov chain is in state $s$.
+For $K$ arms, each Markov chain has a finite number of states $s$, each corresponding to a (constant) reward that the player obtains if she selects this arm while its Markov chain is in state $s$.
 Rested Markov models means that only the state of the selected arm's Markov chain can change, following its Markov transition matrix.
@@
 The different applications detailed above are still active research directions,
-and a curious reader will find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}.
+and a curious reader can find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}.
 
@@
 \begin{defn}[Optimal and sub-optimal arms]\label{def:2:optimalSubOptimalArms}
+    Consider a bandit problem of $K$ arms with distributions characterized by their means $\bm{\mu}=\mu_1,\dots,\mu_k$.
     Denote $\mu^* = \max_k \mu_k$ the largest mean.\\
@@
     \begin{equation}
-        R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ].
+        R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ] = \sum_{k=1,\dots,K : \Delta_k > 0} \Delta_k \E[ N_k(T) ].
     \end{equation}
@@
 \begin{proof}\label{proof:2:RegretDecomposition}
-    Thanks to the tower rule, we had
-    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ].$
+    Thanks to the tower rule, we can do as above and obtain
+    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]$.
+    The sum can then be simplified to only count sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
 \end{proof}
 
-\textbf{Applications:}
+\paragraph{Consequences.}
 %
 Such decomposition of the regret is useful for at least two reasons.
-%
+
 On the first hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
@@
 In our library SMPyBandits, we implement both possibilities, and all values of regret used in this thesis are based on the second estimator, using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives smoother curves and faster convergence to the limit behavior.
-%
+
 On the second hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
@@
 and a significant part of the research literature on MAB algorithms has focussed on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
-This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically match the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is optimal), or with a larger constant (the algorithm is said order-optimal in such case).
+This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically match the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
 
@@
 
-Then we consider algorithms that uniformly efficient\footnote{This notion is then extended for ``strongly uniformly efficiency'' in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}.}, meaning that we rule out algorithms that can achieve low regret on some problem instance while achieving linear regret on other instances.
-In particular, it is necessary to rule out algorithms that always pick the same arms, as on some problem instance such fixed-arm algorithm can achieve zero regret.
-For instance, the \UCB{} algorithm is uniformly efficient.
+We consider uniformly efficient\footnote{This notion is then extended for ``strongly uniformly efficiency'' in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}.} algorithms, to rule out algorithms achieving low regret on some problem instances while achieving linear regret on other instances.
+In particular, it is necessary to rule out algorithms that always pick the same arm, as on some problem instances such fixed-arm algorithms can achieve zero regret.
+This class contains for instance the \UCB{} algorithm, as its regret is proven to be logarithmic on any problem instance.
 
@@
 % ----------------------------------------------------------------------------
-\section{Review of MAB algorithms}
+\section{Review of stochastic MAB algorithms}
 \label{sec:2:famousMABalgorithms}

### 2019-05-22 2-Chapters/3-Chapter/chapter3.tex
Small modifications

@@
         First, BESA is very costly in terms of memory, and $\UCB\dagger$ and \textsc{Tsallis-Inf} are also more costly than others, and both experiences a linear growth of their memory cost.
-        Then, we see PHE, $\mathrm{Exp}3++$, Bayes-UCB and AdBandits having a small but linear growth of memory cost, and it can be explained for each of them (\eg, for Bayes-UCB, computing the $1-1/t$ quantile at time $t$ becomes more and more costly as $t$ grows, and thus as $T$ grows,  ).
+        Then, we see PHE, $\mathrm{Exp}3^{++}$, Bayes-UCB and AdBandits having a small but linear growth of memory cost, and it can be explained for each of them (\eg, for Bayes-UCB, computing the $1-1/t$ quantile at time $t$ becomes more and more costly as $t$ grows, and thus as $T$ grows,  ).
         Finally, the other algorithms (including \UCB, \klUCB{} and Thompson sampling) have a very small footprint and their memory cost do not grow much when $T$ increases.

### 2019-05-23 1-Preamble/4-Acknowledgement/acknowledgement.tex
Suffit pour ce soir…

@@
 Christophe was very supportive on all aspects of the PhD work, and he was also helpful during difficult moments in the last three years.
-Many thanks for being so nice and dynamic\footnote{We had a nice first contact when we met in Spring 2015 in Mahindra Ecole Centrale, in Hyderabad, India, where I taught for one year, and I am honored to have pursued my PhD under his supervision and to have worked with him, and I am also happy to have lived great human moments along with Christophe since 2016.}, and always comprehensive.
+Many thanks for being so nice and dynamic\footnote{~We had a nice first contact when we met in Spring 2015 in Mahindra Ecole Centrale, in Hyderabad, India, where I taught for one year, and I am honored to have pursued my PhD under his supervision and to have worked with him, and I am also happy to have lived great human moments alongside Christophe since 2016.}, and always comprehensive.
 Christophe was helpful for his expertise on wireless communications and engineering, and especially on cognitive radio.
@@
 \indent
-Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years\footnote{Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.}.
+Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years\footnote{~Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{~\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.}.
 Émilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep the highest possible rigour in all parts of our exchanges and productions.
-I was lucky to be able to visit the SequeL team at Inria Lille on regular occasions, and to work in close collaboration with Émilie during these visits\footnote{Merci à Romain et toi pour votre accueil chaleureux, et aussi merci à Florian, Nicolas et Édouard.}.
+I was lucky to be able to visit the SequeL team at Inria Lille on regular occasions, and to work in close collaboration with Émilie during these visits\footnote{~Merci à Romain et toi pour votre accueil chaleureux, et aussi merci à Florian, Nicolas et Édouard.}.
 Intensive white board sessions proved to be fruitful, giving the research work that I am the most proud of \cite{Besson2018ALT}, and rich research papers that join her rigorous expertise in statistical learning and mathematical analysis of bandit algorithms with my passion for numerical experiments, simulations and algorithm designs \cite{Besson2018ALT,Besson2018DoublingTricks,Besson2019GLRT}.
 
-Then I would like to thank the dear members of my PhD jury.
+Then I would like to thank the dear members of my PhD committee.
 I was lucky to be able to discuss with them on regular occasions, and these discussions have been very useful for some of our works.
@@
 I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my studies.
-His solid and serious work kept pushing me further, some of our board games nights were brain teasers comparable with some maths problems solved for this thesis,
-and I was happy to help him for the organization of the \href{http://lettres.sorbonne-universite.fr/Kaamelott}{research colloque}\footnote{Un livre a aussi été publié suite \href{http://lettres.sorbonne-universite.fr/Kaamelott}{au colloque}, \href{https://www.editions-vendemiaire.com/catalogue/a-paraitre/kaamelott-un-livre-d-histoire-florian-besson-et-justine-breton-dir/}{``Kaamelott, un livre d'histoire'', aux éditions Vandémiaire}.} on the \emph{Kaamelott} french TV show in February 2017, it was \href{https://www.fabula.org/actualites/kaamelott-ou-la-relecture-de-l-histoire_75490.php}{the funniest} and \href{https://www.vousnousils.fr/2017/03/22/kaamelott-un-tremplin-vers-lhistoire-et-la-litterature-601428}{most exciting} \href{https://lpcm.hypotheses.org/10718}{research event} I have been to in the last four years!
+His solid and serious work kept pushing me further, some of our board games nights were brain teasers comparable with some maths problems solved for this thesis.
+% and I was happy to help him for the organization of the \href{http://lettres.sorbonne-universite.fr/Kaamelott}{research colloque}\footnote{~Un livre a aussi été publié suite \href{http://lettres.sorbonne-universite.fr/Kaamelott}{au colloque}, \href{https://www.editions-vendemiaire.com/catalogue/a-paraitre/kaamelott-un-livre-d-histoire-florian-besson-et-justine-breton-dir/}{``Kaamelott, un livre d'histoire'', aux éditions Vandémiaire}.} on the \emph{Kaamelott} french TV show in February 2017, it was \href{https://www.fabula.org/actualites/kaamelott-ou-la-relecture-de-l-histoire_75490.php}{the funniest} and \href{https://www.vousnousils.fr/2017/03/22/kaamelott-un-tremplin-vers-lhistoire-et-la-litterature-601428}{most exciting} \href{https://lpcm.hypotheses.org/10718}{research event} I have been to in the last four years!
 %
@@
 
-I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
+I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
 And due to space constraints, I had to chose quite arbitrarily who to include\dots}: Florian, Fabian, Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia from Briançon, Mayotte and Pierre from Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément from Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also from Cachan, Kumudham and Priscilla from India, Thibault, Sébastien, Sélim, Valentin again from Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola from Rennes.
-In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
+% In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
 I loved traveling with you and hope to continue!
@@
 
-Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline and others at CRIStAL in Lille.
+Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline (and others) at CRIStAL in Lille.
 Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially Martine and J-B, your daily smile is priceless!
@@
 and by Ministry of Higher Education and Research, Brittany and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
-I was also lucky to be well received very frequently at Inria Lille Nord Europe, and to be hosted at CentraleSupélec campus of Rennes.
+% I was also lucky to be well received very frequently at Inria Lille Nord Europe, and to be hosted at CentraleSupélec campus of Rennes.
 
@@
 %,
-\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{Que je remercie deux fois plus que ce qu'il me remerciant dans sa propre thèse !
+\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{~Que je remercie deux fois plus que ce qu'il me remerciant dans sa propre thèse !
 Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.

### 2019-05-23 2-Chapters/1-Chapter/chapter1.tex
Suffit pour ce soir…

@@
     1st International MOTIoN workshop\footnote{~MOTIoN 2019, see \href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
-    at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.
+    at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.\\
     \href{https://HAL.Inria.fr/hal-02049824}{\texttt{HAL.Inria.fr/hal-02049824}}.
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
-    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{~CROWNCOM, \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,\\
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{~CROWNCOM 2017, see \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,\\
     \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
-    Demonstration presented at \emph{International Conference on Communication}\footnote{~ICT, \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}}, in Saint-Malo, France in June $2018$.
+    Demonstration presented at \emph{International Conference on Communication}\footnote{~ICT 2018, see \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}}, in Saint-Malo, France in June $2018$.
     See \href{https://YouTu.be/HospLNQhcMk}{\texttt{YouTu.be/HospLNQhcMk}} for a $6$-minutes presentation video.
@@
     \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
-    submitted at GRETSI 2019\footnote{~\href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
+    submitted at GRETSI 2019\footnote{~GRETSI 2019, see \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
     by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
@@
     \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann,\\
-    September $2018$,
+    by \textbf{Lilian Besson} \& Emilie Kaufmann,
+    September $2018$,\\
     \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.

### 2019-05-23 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Small modifications
Suffit pour ce soir…

@@
 % ----------------------------------------------------------------------
-\subsubsection{Explanations of \Aggr{} versus \ExpQ{} }\label{sub:25:Exp4}
+\paragraph{Explanations of \Aggr{} versus \ExpQ{} }\label{sub:25:Exp4}
 
@@
 Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
-It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
-and appears to be more robust than simple non-aggregated algorithms.
+% It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
+% and appears to be more robust than simple non-aggregated algorithms.
 
@@
 
-Our \Aggr{} algorithm could be rewritten as an Online Mirror Descent \cite{Hazan2016introduction,Zimmert2018}, as it is the case for both \ExpQ{} and \CORRAL.
+Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent \cite{Hazan2016introduction,Zimmert2018}, as it is the case for both \ExpQ{} and \CORRAL.
 But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.

### 2019-05-23 2-Chapters/2-Chapter/chapter2.tex
Suffit pour ce soir…

@@
     \centering
-    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
+    \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
     \caption{Screenshot of the demonstration available online on my website, for a current step of $t=24$.}
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
+    \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
     \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.
@@
 %     \item
-    \textbf{A/B testing} of websites is a popular application of the best arm identification problem,
+    \textbf{A/B testing}, for instance for websites, is a popular application of the best arm identification problem,
     where the task is purely exploratory and the player is asked to identify the best of two options (or more), in a finite number of steps \cite{audibert2010best}.
-    The problem can either consider a fixed budget and no freedom on the ending of the game (\ie, an known horizon $T$), or a fixed confidence (\ie, the identified arm must be the true best arm with probability at least $1-\delta$) and a certain freedom on the budget \cite{Garivier16BAI}.
+    The problem can either consider a fixed budget and no freedom on the ending of the game (\ie, an known horizon $T$), or a fixed confidence and a certain freedom on the budget (\ie, the identified arm must be the true best arm with probability at least $1-\delta$) \cite{Garivier16BAI}.
     The theoretical complexity of this use of bandit for A/B testing was first studied in \cite{Kaufmann14},
@@
 
+\begin{itemize}
+    \item Algorithm 1 is
+    an $\varepsilon$-greedy algorithm, as described in \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
+    It has a very small cost in terms of time and memory, but it usually achieves linear regret.
+
+    \item Algorithm 2 is
+    an Explore-then-Commit algorithm from \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
+    Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but only obtain large (linear) regret.
+
+    \item Algorithm 3 is
+    $\mathrm{Exp}3^{++}$ from \cite{Seldin17}, using $\alpha=3$ and $\beta=256$ as advised (\texttt{Exp3PlusPlus}).
+    It is a recent variant of the $\mathrm{Exp3}$ algorithm, using an adaptive tuning of its parameter.
+It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+
+    \item Algorithm 4 is
+    $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
+    It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+
+    \item Algorithm 5 is
+    \klUCB{} from \cite{KLUCBJournal}, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern} (\texttt{klUCB}).
+    It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
+    It is computationally more costly that \UCB, as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
+    In practice, a few steps of a simple Newton method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than an order of magnitude slower than \UCB.
+    % of more than 10 digits
+    \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}
+
+    \item Algorithm 6 is
+    Thompson sampling from \cite{Kaufmann12Thompson}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{Thompson}).
+    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+    It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on C or Cython code, highly optimized).
+
+    \item Algorithm 7 is
+    Bayes-UCB from \cite{Kaufmann12BUCB}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
+    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
+    It is comparable to Thompson sampling in terms of memory complexity, but slower in terms of time complexity as computing a $1-1/t$ quantile is more costly than sampling from a distribution.
+    But in particular for Bernoulli distributed arms and for Beta posteriors and priors, Bayes-UCB is only about twice as slow as Thompson sampling.
+
+    \item Algorithm 8 is
+    AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
+    It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posterior on each arm.
+It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson decision rule (\ie, choose the arm maximizing one draw of random samples from the posteriors), otherwise it uses the Bayes-UCB decision rule (\ie, choose the arm maximizing the $1-1/t$ quantile).
+    It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
+    In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
+
+    \item Algorithm 9 is
+    BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, and it consists in trying to fix a simple observation.
+    For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
+    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
+    It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
+    Analyzing it in the generic case was found to be difficult, and is still an open problem.
+    We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions, \ie, using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
+    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffer from an exponential blow-up when $K$ increases.
+\end{itemize}
+
 % ----------------------------------------------------------------------------

### 2019-05-24 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
SAAAAAD

@@
 and how to update the trusts at each step, and our proposal \Aggr{} differs from \ExpQ{} on these very points.
+%
+The considered aggregation algorithms are special cases of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details see
+the Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
+or this blog post by Jeremy Kun \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-of-the-}}\\
+\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{multiplicative-weights-update-algorithm}},
+which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019).},
+a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
+More details can be found in the survey \cite{Arora2012multiplicative}.
 

### 2019-05-24 2-Chapters/2-Chapter/chapter2.tex
SAAAAAD

@@
 \begin{lemma}[Regret decomposition]\label{lem:2:RegretDecomposition}
-    The (mean) regret $R_T^{\cA}$ can be written like
-    \begin{equation}
+    The (mean) regret $R_T^{\cA}$ can be decomposed as a sum of the number of selections of sub-optimal arms $k$,
+    weighted by their gap, like this:
+    \begin{equation}\label{eq:2:RegretDecomposition}
         R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ] = \sum_{k=1,\dots,K : \Delta_k > 0} \Delta_k \E[ N_k(T) ].
@@
 \begin{proof}\label{proof:2:RegretDecomposition}
-    Thanks to the tower rule, we can do as above and obtain
-    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]$.
+    The tower rule of expectation yields
+    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$ as above, and thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]$.
     The sum can then be simplified to only count sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
@@
 On the first hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
-Mathematically, both quantities are equal in expectation, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed.
-In our library SMPyBandits, we implement both possibilities, and all values of regret used in this thesis are based on the second estimator, using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives smoother curves and faster convergence to the limit behavior.
+Mathematically, both quantities are equal in expectation, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition}.
+In our library SMPyBandits, we implement both estimators, and all values of regret used in this thesis are based on the one using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives smoother curves and faster convergence to the limit behavior.
 
@@
 \begin{theorem}\label{thm:2:firstLogTLowerBound}
-    No algorithm $\cA$ can achieve a regret $R_T^{\cA}(I) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
+    No algorithm $\cA$ can achieve a (mean) regret $R_T^{\cA}(I) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
     where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
@@
 
-We consider uniformly efficient\footnote{This notion is then extended for ``strongly uniformly efficiency'' in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}.} algorithms, to rule out algorithms achieving low regret on some problem instances while achieving linear regret on other instances.
+We consider uniformly efficient\footnote{This notion is then extended for ``strongly uniformly efficiency'' in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}, where we also include a notion of (expected) fairness.} algorithms, to rule out algorithms achieving low regret on some problem instances while achieving linear regret on other instances.
 In particular, it is necessary to rule out algorithms that always pick the same arm, as on some problem instances such fixed-arm algorithms can achieve zero regret.
@@
 
-\begin{defn}[Uniformly efficient algorithm]\label{def:2:uniformlyEfficientAlgorithm}
-    An algorithm $\cA$ is \emph{uniformly efficient} if it satisfies
+\begin{defn}\label{def:2:uniformlyEfficientAlgorithm}
+    An algorithm $\cA$ is \emph{uniformly efficient} if its (mean) regret satisfies
     \begin{equation}
@@
 
+Many algorithms have been proved to achieve logarithmic regret in the stochastic case,
+and in particular it is the case of the algorithms used in this thesis, \UCB{} from \cite{Auer02}, Thompson sampling from \cite{Thompson33} and analyzed in \cite{AgrawalGoyal11,Kaufmann12Thompson}, and \klUCB{} from \cite{Garivier11KL,KLUCBJournal}.
+%
+Such bounds are valid in different settings, and in particular \UCB{} is order-optimal for bounded rewards or one-dimensional exponential families,
+while Thompson sampling is optimal for bounded rewards, and \klUCB{} have been proven to be optimal for both cases.
+
 \paragraph{Worst-case lower-bound in $\sqrt{T}$}
 
-For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses the constants $C_I$ no longer bring any information.
-We naively have that $R_T^{\cA} \leq (\max_k \Delta_k) T$, and thus if 
-\TODOL{Finish this!}
+For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses a constant $C_I$ no longer bring any information.
+Indeed, we can naively bound the regret by $R_T^{\cA} \leq (\max_k \Delta_k) T$, and thus if $(\max_k \Delta_k)$ can be taken so small that $C_I \log(T) \gg (\max_k \Delta_k) T$, and thus a regret upper bound like $R_T^{\cA} \leq C_I \log(T)$ is useless.
 
-For this reason, it is interesting to quote another lower-bound on the regret, that also
+For an example, consider any large horizon $T>10$, and $K=2$ Bernoulli arms.
+Chose any small $\varepsilon$ such that $\varepsilon \ll \sqrt{4 \log(T) / T}$, then if the means are $\mu_1 = 1/2$ and $\mu_2 = 1/2 - \varepsilon$.
+The constant $C_I$ from \eqref{eq:2:forSecondLogTLowerBound} is $C_I = 1 / (4 \varepsilon)$ and $\Delta=\varepsilon$,
+and thus the logarithmic upper-bound is useless in such case, as
+\begin{equation}
+    \varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T.
+\end{equation}
+
+For this reason, the research literature has also looked with interest another family of bounds on regret, that are not problem-dependent but worst-case, also called minimax \cite{Audibert2009minimax,audibert2010minimax}.
+Hence it is interesting to quote another lower-bound on the regret, that can bring useful information in such settings.
 The following theorem was proven in \cite{Auer02NonStochastic}.
-
-\TODOL{It states that FIXME interpret! FIXME explain why this is interesting for the rest of this manuscript!}
 
@@
 \end{theorem}
+
+\TODOL{FIXME explain why this is interesting for the rest of this manuscript!}
+
+Similarly to what is considered for the first lower-bound,
+a natural question is to know if there is an algorithm that achieve a regret upper-bound of the form $R_T^{\cA}(I) \leq \bigO{\sqrt{K T}}$ for any instance, independently on the problem difficulty.
+It is the case for the Exp3 algorithm from \cite{Auer02}, for example.
+Since a few years, some algorithms were shown to achieve both a problem-dependent logarithmic and a minimax upper-bounds,
+like MOSS in \cite{Audibert2009minimax} or recently \KLUCBpp{} in \cite{Menard17}.
 
@@
 %     - Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models) \cite{Auer02NonStochastic}
-% \end{itemize}
-
-% \subsection{Regret upper bounds, tools and different kinds of bounds}
-
-% \begin{itemize}
-%     \item
-%     - different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
-%     \item
-%     - tools to show regret upper-bounds, concentration inequalities?
-%     FIXME non j'ai pas la place
-%     \item
-%     - short proof of UCB or klUCB? just to start to show the tools
-%     FIXME non j'ai pas la place
 % \end{itemize}
@@
 
-Everything is implemented and documented at
-https://smpybandits.github.io/docs/Policies.html
+In this Section, we start by discussing two naive strategies which fail dramatically.
+%
+We then expose two simple strategies, which performs efficiently only if they are tuned using a prior knowledge of the problem difficulty, and are thus unusable for practical applications.
+%
+We continue by presenting a first efficient policy, \UCB, that uses upper confidence bounds (UCB) on the mean estimates, and is known to be order-optimal for bounded rewards.
+Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of \UCB{} but use smaller confidence intervals and thus typically has better theoretical results, and Thompson sampling which takes a Bayesian point-of-view instead of relying on the principle of optimism in the face of uncertainty.
+%
+Finally, we conclude this Section by briefly exposing other families of algorithms.
 
-\subsubsection{Why do we need smart algorithms?}
-Why Follow-the-Leader (EmpiricalMeans https://smpybandits.github.io/docs/Policies.EmpiricalMeans.html) don't work, example.
+\textbf{Implementation:}
+%
+We describe our library SMPyBandits in more details in Chapter~\ref{chapter:3}, but all the described algorithms are implemented and documented in SMPyBandits, alongside with many more algorithms (there were about 65 for single-player stochastic problems at the time of writing this thesis).
+A complete list of the implemented policies can be found on the following web page on the documentation,
+\href{https://smpybandits.github.io/docs/Policies.html}{\texttt{SMPyBandits.GitHub.io/docs/Policies.html}}, and on the \texttt{Policies} module.
 
-\subsection{Simple but efficient algorithms when well tuned: $\varepsilon$-greedy and Explore-then-Commit}
+% ---------------------------------------------------
+\subsubsection{Naive strategies: pure exploitation or pure exploration}
 
-See
-https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
-https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
+Let us first describe two naive strategies, that both fail dramatically.
+We recall the notations introduced above, the sum of rewards $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, and the number of samples $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
+%
+The estimated means, or empirical averages, are $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$).
+
+% Why Follow-the-Leader (EmpiricalMeans https://smpybandits.github.io/docs/Policies.EmpiricalMeans.html) don't work, example.
+On the first hand, consider the ``\emph{Follow-the-Leader}'' strategy which consists in first playing once each arm, then always playing $A(t)\in\argmax \widehat{\mu_k}(t)$.
+The player only exploits the collected information, and this strategy can fail dramatically, \ie, obtain linear regret in some problems.
+Indeed consider $K=2$ Bernoulli arms of means $\mu_1=1/2$ and $\mu_2=\varepsilon$ where $\varepsilon < 1/2$, then with probability $\varepsilon/2$ the player observes a reward of $0$ for arm $1$ then $1$ for arm $2$ on the first rounds, and so she will play arm $2$ for all the $T-2$ rounds, giving a linear regret $R_T \geq \frac{\varepsilon}{2}\left(1/2 - \varepsilon\right) (T-1)$.
+
+On the other hand, consider the uniform strategy which consists in always playing the $K$ uniformly at random.
+The player only explores without using the collected information, and this strategy fails dramatically for any non trivial problem (\ie, linear regret).
+Indeed it obtains a linear regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
+which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all problems expect where $\mu_1=\mu_2=\dots=\mu_K$).
+
+% ---------------------------------------------------
+\subsubsection{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Commit}
+
+As the two examples illustrated, an efficient strategy needs to solve the trade-off between exploration and exploitation.
+The two following solutions both consist in splitting the $T$ time steps into $T_0$ steps of exploration and $T-T_0$ steps of exploitation, either in an alternative way or in a fixed way (``explore then exploit'').
+
+% https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
+On the first hand, the $\varepsilon$-greedy strategy, from \cite{SuttonBarto2018}, consists in alternating exploration and exploitation at a certain ratio.
+Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
+If $\varepsilon$ is constant, then the (mean) regret is still growing linearly as it is lower bounded by $\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k T$.
+But if we consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = 6 K / (d^2 t)$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$,
+then it was shown in \cite{Auer02} that the regret is of the order of $K \log(T) / d + \smallO{T}$.
+
+% https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
+On the other hand, the ``explore-then-commit'' strategy, also introduced in \cite{SuttonBarto2018} and studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
+If $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly.
+
+Both strategies give linear regret if they are tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if lower-bound on the gaps $\Delta_k$ is known.
+%  \ie, they require a prior knowledge on the problem difficulty.
+As such, they are not applicable on an unknown problem and are less interesting from a practical point-of-view.
+
+\newpage
 

### 2019-05-24 2-Chapters/2-Chapter/chapter2.tex
C'est un cauchemar je ne sais rien écrire proprement

@@
 On a more experimental note, it was showed in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-player bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting means that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
-For more details, see \cite{modiDemo2016} for the multi-player case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi}.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
+For more details, see \cite{modiDemo2016} for experiments on the multi-player case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
 
@@
 But if we consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = 6 K / (d^2 t)$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$,
-then it was shown in \cite{Auer02} that the regret is of the order of $K \log(T) / d + \smallO{T}$.
+then it was shown in \cite{Auer02} that the regret is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{\log(T)}$.
 
 % https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
-On the other hand, the ``explore-then-commit'' strategy, also introduced in \cite{SuttonBarto2018} and studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
-If $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly.
-
-Both strategies give linear regret if they are tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if lower-bound on the gaps $\Delta_k$ is known.
+On the other hand, the ``explore-then-exploit'' strategy, also introduced in \cite{SuttonBarto2018} and studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
+If $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, as $R_T \geq (\min_{k: \Delta_k > 0} \Delta_k) K T_0 + p (T - T_0)$, if $p$ denotes the probability that the chosen arm is not optimal.
+As $p$ only depends on the number of collected samples $T_0$ and not $T$, $R_T = \Omega(T)$.
+
+\begin{proof}
+    For the case of two arms,
+    for instance, if the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality from \cite{hoeffding1963probability} to control this probability $p$.
+    %
+    Indeed if $Z_1,\dots,Z_n$ are \iid{} samples from a Bernoulli distribution of parameter $\mu$, of mean $\hat{Z}_n$ we have
+    \begin{align}
+        \forall x < \mu, \bP(\hat{Z}_n) \leq \exp(-2 n (x-\mu)^2),\\
+        \forall x > \mu, \bP(\hat{Z}_n) \geq \exp(-2 n (x-\mu)^2).
+    \end{align}
+    %
+    If $\mu_1 = \mu_2 + \Delta$, then
+    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \widehat{\mu_2}) + \bP(\widehat{\mu_1} < \widehat{\mu_2})$, and both terms can be bounded by using Hoeffding's inequality with $n=T_0/2$ samples,
+    to obtain $p \leq 2 \exp(-T_0 \Delta^2 / 4)$.
+    %
+    For $K$ arms, the same proofs applies and we can obtain
+    $p \leq 2 K \exp(-T_0 d^2 / 4)$ if $d$ is a lower-bound on the positive gaps.
+    %
+    It shows that $p$ is a constant wrt $T$.
+\end{proof}
+
+Using the same analysis, if a lower-bound $d$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret, like for the previous strategy.
+%
+Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) K T_0 + p (T - T_0)$, thus $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$ using Hoeffding's inequality again.
+Optimizing on $T_0$ gives $T_0 = 4/d^2 \log(d^2 T / 4)$
+and $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$.
+
+An extension of the ``exploit-then-exploit'' is to consider not a fixed time $T_0$ but a random time $\tau$ at which exploration stops.
+This time $\tau$ must be a \emph{stopping time} in the sense that it is a measurable random variable, dependent of the past observations, as introduced by \cite{Wald45}.
+The strategy is then referred to as ``explore-then-commit'' (ETC), and the idea is to use a statistical test at every time step $t$, and stop as soon as enough samples were collected to effectively identify the best arm with a certain confidence level $\delta$.
+Choosing $\delta \propto 1/T$ typically lead to a logarithmic upper-bound on the regret.
+
+For both cases, the strategy obtains a linear regret if it is tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if lower-bound on the gaps $\Delta_k$ is known.
 %  \ie, they require a prior knowledge on the problem difficulty.
@@
 
+% \begin{small} % XXX remove if needed
+\begin{figure}[h!]
+	\centering
+    \begin{framed}
+	\begin{algorithm}[H]
+		\begin{small} % XXX remove if needed
+		\For(){$t = 1, \dots, T$}{
+            \uIf{$t \leq K$}{
+                Play arm $A(t) = t$\;
+            }
+			\Else(){
+                For each arm, compute the index $U_k(t)$\;
+                Play arm $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
+            }
+            Observe a reward $r_{A(t)}(t)$\;
+            Update both $X_k(t)$ and $N_k(t)$\;
+		}
+		\caption{A generic index policy.}
+		\label{algo:2:indexPolicy}
+		\end{small} % XXX remove if needed
+	\end{algorithm}
+	\end{framed}
+\end{figure}
+% \end{small} % XXX remove if needed
+
 \subsubsection{UCB}
@@
 
+compute the index $U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$.
+
 \subsubsection{klUCB}
@@
 \TODOL{Short note on how to implement such policy, using Newton's method ?}
+
+compute the index $U_k(t) = XXX$.
 
@@
 
-It's a good opportunity to show off and say that I worked on my own on some non-standard policies.
+It's a good opportunity to show that I worked on my own on some non-standard policies.
 

### 2019-05-24 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
SHIT I'm going home

@@
 					$\Alg_a$ updates its internal state (e.g., \UCB{} indexes)\;
-					It chooses $A(t+1)_{a} \in [K]$.
+					It chooses $A_a(t+1) \in [K]$.
 				}
@@
 				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
-				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A(t+1)_{a} = j\}), \;\forall 1\leq j \leq K$\;
+				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A_a(t+1) = j\}), \;\forall 1\leq j \leq K$\;
 				% \tcp*{Sum trusts}
@@
 				\For{$a = 1, \dots, N$}{
-					% \eIf{$\Alg_a$ was trusted, \ie, $A(t+1)_{a} = A(t+1)$}{
-					\If{$\Alg_a$ was trusted, \ie, $A(t+1)_{a} = A(t+1)$}{
+					% \eIf{$\Alg_a$ was trusted, \ie, $A_a(t+1) = A(t+1)$}{
+					\eIf{$\Alg_a$ was trusted, \ie, $A_a(t+1) = A(t+1)$}{
 						$ \pi^{t+1}_{a} = \exp(\eta_t \ell^{t+1}) \times \pi^{t}_{a} $
-						% \tcp*{More trusted}
-						}%{
+						\tcp*{Trusted more}
+						}{
 						% $ \pi^{t+1}_{a} = \exp(- \eta_t \ell^{t+1}) \times \pi^{t}_{a} $
-						% \tcp*{Less trusted}
-					% }
+						$ \pi^{t+1}_{a} = \pi^{t}_{a} $
+						\tcp*{Do not update the trust now}
 					}
-				Renormalize the new $\pi$: $\pi^{t+1} := \pi^{t+1} / \sum_{a=1}^{N} \pi^{t+1}_{a}$.
+				}
+				Renormalize the new weights $\pi^{t+1}$: $\pi^{t+1} := \pi^{t+1} / \sum_{a=1}^{N} \pi^{t+1}_{a}$.
 				% \tcp*{Project it back to $\Delta_N$}
 			}
-			\caption{Our aggregation algorithm \Aggr.}
+			\caption{Our algorithm \Aggr{} aggregating $\Alg_1,\dots,\Alg_N$.}
 			\label{algo:25:Aggr}
@@
 The loss estimate is unbiased in the following sense.
-If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\tilde{\ell}_{a,t+1} = \ell_{A_A(t+1)}(t+1)$.
+If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\widetilde{\ell_{a,t+1}} = \ell_{A_a(t+1)}(t+1)$.
 This quantity can only be observed for those algorithms for which $A_{a}^{t+1}=A(t+1)$.
-However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $\tilde{\ell}_{a,t}$.
+However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $\widetilde{\ell_{a,t}}$.
 More precisely, if we define the estimate by
 \begin{equation}
-	\hat{\ell}_{a,t+1} = \frac{\ell_{A_A(t+1)}(t+1)}{p_{A_A(t+1)}^{t+1}} \mathbbm{1}(A_A(t+1) = A(t+1))
+	\widehat{\ell_{a,t+1}} = \frac{\ell_{A_a(t+1)}(t+1)}{p_{A_a(t+1)}^{t+1}} \mathbbm{1}(A_a(t+1) = A(t+1))
 \end{equation}
-satisfies $\mathbb{E}[\hat{\ell}_{a,t+1} | \mathcal{H}_t] = \tilde{\ell}_{a,t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$.
-Observe that $\tilde{\ell}_{a,t+1}=\ell^{t+1}$ for all algorithms $a$ such that $A_A(t+1)=A(t+1)$, and $\tilde{\ell}_{a,t+1}=0$ otherwise.
+satisfies $\mathbb{E}[\widehat{\ell_{a,t+1}} | \mathcal{H}_t] = \widetilde{\ell_{a,t+1}}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$.
+Observe that $\widetilde{\ell_{a,t+1}}=\ell^{t+1}$ for all algorithms $a$ such that $A_a(t+1)=A(t+1)$, and $\widetilde{\ell_{a,t+1}}=0$ otherwise.
 
@@
 and an alternative would be to
-increase $\pi_{a}$ if $A(t+1)_a = A(t+1)$ and to decrease it otherwise.
+increase $\pi_{a}$ if $A_a(t+1) = A(t+1)$ and to decrease it otherwise.
 It would not be so different, as there is a final renormalization step, and empirically this variation has little impact on the performance of \Aggr{}.

### 2019-05-24 2-Chapters/2-Chapter/chapter2.tex
SHIT I'm going home

@@
 We give an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the \UCB, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
-We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contribution of this thesis.
-We use it to compare the empirical performances of the most famous and most efficient algorithms, in terms of regret and by evaluating them on different problems.
-%
-In the point-of-view of numerical simulations or real-world applications of bandits on embedded systems, one must also take care of two other measures of performance: time and memory complexities, which can vary greatly between different families of algorithms.
-We also empirically compare different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
+% We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contributions of this thesis.
+% We use it to compare the empirical performances of the most famous and most efficient algorithms, in terms of regret and by evaluating them on different problems.
+% %
+% In the point-of-view of numerical simulations or real-world applications of bandits on embedded systems, one must also take care of two other measures of performance: time and memory complexities, which can vary greatly between different families of algorithms.
+% We also empirically compare different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
 %
@@
 They both attain a logarithmic regret, proving their order optimality (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
-Moreover, \klUCB{} is known to be optimal for a wide range of arms distributions.
+% Moreover, \klUCB{} is known to be optimal for a wide range of arms distributions.
 % A sketch of the proof of both results is given in Appendix.
@@
 Finally, we explore the problem of algorithm selection: as there are many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
-We present one of the first contribution made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms.
-We illustrate its performance empirically, but no satisfactory theoretical result is given, as the bound we had hoped for has been proved unachievable since then.
+We present one of the first contributions made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms,
+and we illustrate its performance empirically.
+% , but no satisfactory theoretical result is given, as the bound we had hoped for has been proved unachievable since then.
 
@@
 %
-As introduced above when we described the UI of our online example demo, let $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ be the number of times arm $k$ was selected between times $1$ and $t$.
-Remember that the rewards $r_k(t)$ are all \iid{} of mean $\mu_k$.
-Thus one can use the chain rule of expectation, and because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$ (which are measurable wrt to the past observations $\cO_t$),
-one can rewrite the expected cumulated rewards like this,
-$\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]$
-$= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right]$
-$= \sum_{k=1}^K \E [ r_k(1) \sum_{t=1}^T \mathbbm{1}(A(t) = k) ]$
-% \underbrace{XXX}_{= N_k(T)}
-$= \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].$
-
-If we introduce the \emph{gap} $\Delta_k = \mu^* - \mu_k$ between any arm $k\in[K]$ and an optimal arm,
-then an arm $k$ is sub-optimal if $\Delta_k > 0$, and we can write the following decomposition on the regret,
+Remember that $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ denotes the number of times arm $k$ was selected between times $1$ and $t$,
+%
+and that the rewards $r_k(t)$ are all \iid{} of mean $\mu_k$.
+The \emph{gap} between any arm $k\in[K]$ and an optimal arm is defined as $\Delta_k = \mu^* - \mu_k$
+(an arm $k$ is sub-optimal if $\Delta_k > 0$),
+and thus we can write the following decomposition on the regret,
 
@@
 \begin{proof}\label{proof:2:RegretDecomposition}
-    The tower rule of expectation yields
-    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$ as above, and thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]$.
+    One can use the chain rule of expectation, and because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$ (which are measurable wrt to the past observations $\cO_t$),
+    one can rewrite the expected cumulated rewards like this,
+    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]$
+    $= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right]$
+    $= \sum_{k=1}^K \E [ r_k(1) \sum_{t=1}^T \mathbbm{1}(A(t) = k) ]$
+    % \underbrace{XXX}_{= N_k(T)}
+    $= \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].$
+    %
+    Thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]$.
     The sum can then be simplified to only count sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
@@
 
-On the first hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
+-- On the first hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
 Mathematically, both quantities are equal in expectation, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition}.
@@
 
-On the second hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
+-- On the other hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
 For example, we extend this decomposition to the multi-player case in Chapter~\ref{chapter:3}, in Lemma~\ref{lem:5:DecompositionRegret}, and as it also uses the draws $N_k(T)$, it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
@@
 %
-Let $\cI$ denote the set of all problem instances, with $K \geq 2$ arms whose rewards lie in $[0,1]$ (and no other hypothesis).
+Let $\cI$ denote the set of all problem instances, with $K \geq 2$ arms. We assume the rewards lie in $[0,1]$ (but no additional hypothesis).
 To specify the dependency on an instance $I$, we denote the regret of $\cA$ on instance $I$ and horizon $T$ by $R_T^{\cA}(I)$.
@@
 For an example, consider any large horizon $T>10$, and $K=2$ Bernoulli arms.
-Chose any small $\varepsilon$ such that $\varepsilon \ll \sqrt{4 \log(T) / T}$, then if the means are $\mu_1 = 1/2$ and $\mu_2 = 1/2 - \varepsilon$.
-The constant $C_I$ from \eqref{eq:2:forSecondLogTLowerBound} is $C_I = 1 / (4 \varepsilon)$ and $\Delta=\varepsilon$,
-and thus the logarithmic upper-bound is useless in such case, as
-\begin{equation}
-    \varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T.
-\end{equation}
+Chose any small $\varepsilon$ such that $\varepsilon \ll \sqrt{4 \log(T) / T}$, then chose let $\mu_1 = 1/2$ and $\mu_2 = 1/2 - \varepsilon$.
+The constant $C_I$ from \eqref{eq:2:forSecondLogTLowerBound} is $C_I = 1 / (4 \varepsilon)$ and $\Delta=\varepsilon$ and it becomes so large that
+the logarithmic upper-bound is useless in such case, as
+\[ \varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T. \]
 
@@
 %
-We then expose two simple strategies, which performs efficiently only if they are tuned using a prior knowledge of the problem difficulty, and are thus unusable for practical applications.
+We then expose two simple strategies, which performs efficiently only if they are tuned using a prior knowledge of the problem difficulty, but are thus unusable for practical applications.
 %
@@
 Let us first describe two naive strategies, that both fail dramatically.
-We recall the notations introduced above, the sum of rewards $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, and the number of samples $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
+We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
 %
@@
 % Why Follow-the-Leader (EmpiricalMeans https://smpybandits.github.io/docs/Policies.EmpiricalMeans.html) don't work, example.
-On the first hand, consider the ``\emph{Follow-the-Leader}'' strategy which consists in first playing once each arm, then always playing $A(t)\in\argmax \widehat{\mu_k}(t)$.
+-- The ``\emph{Follow-the-Leader}'' strategy consists in first playing once each arm, then always playing $A(t)\in\argmax \widehat{\mu_k}(t)$.
 The player only exploits the collected information, and this strategy can fail dramatically, \ie, obtain linear regret in some problems.
-Indeed consider $K=2$ Bernoulli arms of means $\mu_1=1/2$ and $\mu_2=\varepsilon$ where $\varepsilon < 1/2$, then with probability $\varepsilon/2$ the player observes a reward of $0$ for arm $1$ then $1$ for arm $2$ on the first rounds, and so she will play arm $2$ for all the $T-2$ rounds, giving a linear regret $R_T \geq \frac{\varepsilon}{2}\left(1/2 - \varepsilon\right) (T-1)$.
-
-On the other hand, consider the uniform strategy which consists in always playing the $K$ uniformly at random.
+Indeed consider $K=2$ Bernoulli arms of means $\mu_1=1/2$ and $\mu_2=\varepsilon$ where $\varepsilon < 1/2$, then with probability $\varepsilon/2$ the player observes a reward of $0$ for arm $1$ then $1$ for arm $2$ on the first rounds, and so she will play arm $2$ for the $T-2$ remaining rounds, giving a linear (mean) regret $R_T \geq \frac{\varepsilon}{2}\left(\frac{1}{2} - \varepsilon\right) (T-1)$.
+
+-- The uniform strategy always plays the $K$ arms uniformly at random.
 The player only explores without using the collected information, and this strategy fails dramatically for any non trivial problem (\ie, linear regret).
-Indeed it obtains a linear regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
+Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
 which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all problems expect where $\mu_1=\mu_2=\dots=\mu_K$).
@@
 % ---------------------------------------------------
-\subsubsection{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Commit}
-
-As the two examples illustrated, an efficient strategy needs to solve the trade-off between exploration and exploitation.
-The two following solutions both consist in splitting the $T$ time steps into $T_0$ steps of exploration and $T-T_0$ steps of exploitation, either in an alternative way or in a fixed way (``explore then exploit'').
+\subsubsection{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit}
+
+As illustrated by the two previous examples, an efficient strategy needs to solve the trade-off between exploration and exploitation.
+The two following solutions both consist in splitting the $T$ time steps into $T_0$ steps of exploration and $T-T_0$ steps of exploitation.
+%  either in an alternative way or in a fixed way (``explore then exploit'').
 
 % https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
-On the first hand, the $\varepsilon$-greedy strategy, from \cite{SuttonBarto2018}, consists in alternating exploration and exploitation at a certain ratio.
+-- The $\varepsilon$-greedy strategy consists in alternating exploration and exploitation at a certain ratio \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
 Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
 If $\varepsilon$ is constant, then the (mean) regret is still growing linearly as it is lower bounded by $\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k T$.
-But if we consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = 6 K / (d^2 t)$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$,
+In this first case, in average $T_0 = \varepsilon T$ steps are spent on exploration.
+%
+But if we consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 \frac{1}{t}$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$,
 then it was shown in \cite{Auer02} that the regret is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{\log(T)}$.
+In this second case, about $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
 
 % https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
-On the other hand, the ``explore-then-exploit'' strategy, also introduced in \cite{SuttonBarto2018} and studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
-If $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, as $R_T \geq (\min_{k: \Delta_k > 0} \Delta_k) K T_0 + p (T - T_0)$, if $p$ denotes the probability that the chosen arm is not optimal.
-As $p$ only depends on the number of collected samples $T_0$ and not $T$, $R_T = \Omega(T)$.
+-- The ``explore-then-exploit'' strategy, as studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
+We can lower bound the regret of this strategy by
+$R_T \geq K (\min_{k: \Delta_k > 0} \Delta_k) T_0/K + p (T - T_0)$,
+if $p$ denotes the probability that the chosen arm is not optimal (\ie, the arm with highest mean after $T_0/K$ samples of each arms).
+And we can show that $p$ only depends on the number of collected samples $T_0$ and the gaps $\Delta_k$,
+so if $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, \ie, $R_T = \Omega(T)$.
 
 \begin{proof}
-    For the case of two arms,
-    for instance, if the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality from \cite{hoeffding1963probability} to control this probability $p$.
-    %
-    Indeed if $Z_1,\dots,Z_n$ are \iid{} samples from a Bernoulli distribution of parameter $\mu$, of mean $\hat{Z}_n$ we have
-    \begin{align}
-        \forall x < \mu, \bP(\hat{Z}_n) \leq \exp(-2 n (x-\mu)^2),\\
-        \forall x > \mu, \bP(\hat{Z}_n) \geq \exp(-2 n (x-\mu)^2).
-    \end{align}
+    We prove it for the case of two arms.
+    If the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality from \cite{hoeffding1963probability} to control this probability $p$.
+    %
+    Indeed if $Z_1,\dots,Z_n$ are \iid{} samples from a Bernoulli distribution of parameter $\mu$, of mean $\widehat{Z_n}$ we have
+    \begin{equation}
+        \forall x < \mu, \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
+        \;\;\;
+        \forall x > \mu, \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+    \end{equation}
     %
     If $\mu_1 = \mu_2 + \Delta$, then
-    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \widehat{\mu_2}) + \bP(\widehat{\mu_1} < \widehat{\mu_2})$, and both terms can be bounded by using Hoeffding's inequality with $n=T_0/2$ samples,
+    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \mu_1 - \Delta/2) + \bP(\widehat{\mu_2} > \mu_2 + \Delta/2)$, and both terms can be bounded by using Hoeffding's inequality with $n=T_0/2$ samples,
     to obtain $p \leq 2 \exp(-T_0 \Delta^2 / 4)$.
     %
-    For $K$ arms, the same proofs applies and we can obtain
-    $p \leq 2 K \exp(-T_0 d^2 / 4)$ if $d$ is a lower-bound on the positive gaps.
-    %
-    It shows that $p$ is a constant wrt $T$.
+    For $K$ arms, the same proofs applies, and we can obtain
+    $p \leq 2 K \exp(-T_0 d^2 / 4)$, if $d$ is a lower-bound on the positive gaps.
+    %
+    In all cases, it shows that $p$ is a constant wrt $T$.
 \end{proof}
 
-Using the same analysis, if a lower-bound $d$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret, like for the previous strategy.
-%
-Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) K T_0 + p (T - T_0)$, thus $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$ using Hoeffding's inequality again.
+Using the same analysis, if a lower-bound $d$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret.
+%
+Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$ using Hoeffding's inequality again.
 Optimizing on $T_0$ gives $T_0 = 4/d^2 \log(d^2 T / 4)$
-and $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$.
-
-An extension of the ``exploit-then-exploit'' is to consider not a fixed time $T_0$ but a random time $\tau$ at which exploration stops.
+and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$.
+
+An extension of this strategy is to consider not a fixed time $T_0$ but a random time $\tau$ at which exploration stops.
 This time $\tau$ must be a \emph{stopping time} in the sense that it is a measurable random variable, dependent of the past observations, as introduced by \cite{Wald45}.
 The strategy is then referred to as ``explore-then-commit'' (ETC), and the idea is to use a statistical test at every time step $t$, and stop as soon as enough samples were collected to effectively identify the best arm with a certain confidence level $\delta$.
-Choosing $\delta \propto 1/T$ typically lead to a logarithmic upper-bound on the regret.
+Choosing $\delta \propto 1/T$ and using a lower-bound $d$ on the gaps typically lead to an order-optimal algorithm as shown in \cite{GarivierETC2016}.
 
@@
     We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions, \ie, using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
-    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffer from an exponential blow-up when $K$ increases.
+    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffers from an exponential blow-up when $K$ increases.
 \end{itemize}
@@
 	\begin{algorithm}[H]
-		\begin{small} % XXX remove if needed
+		% \begin{small} % XXX remove if needed
 		\For(){$t = 1, \dots, T$}{
@@
 		\label{algo:2:indexPolicy}
-		\end{small} % XXX remove if needed
+		% \end{small} % XXX remove if needed
 	\end{algorithm}

### 2019-05-24 2-Chapters/5-Chapter/chapter5.tex
My thesis is so long, and so stupid
Typos in refs and labels
SHIT I'm going home

@@
 
-\chapter[Multi-Player Multi-Armed Bandits Models]{Multi-Player Multi-Armed Bandits Models and Algorithms}
+% \chapter[Multi-Player Multi-Armed Bandits Models]{Multi-Player Multi-Armed Bandits Models and Algorithms}
+\chapter{Multi-Player Multi-Armed Bandits}
 \label{chapter:5}
 \minitoc
+
+\TODOL{Intégrer les commentaires de Christophe, le plus rapidement possible !}
 
@@
 
-Before concluding this chapter, we give a literature review of some of the many of the three models presented above in Section~\ref{sec:5:model}.
-We wanted to highlight that the community was quite active on research on multi-player bandits in the last $10$ years, but especially active since last spring $2018$.
+Before concluding this chapter, we review many extensions to the three models presented above in Section~\ref{sec:5:model} that were introduced in recent literature.
+%
+We wanted to highlight that the community has been quite active on research on multi-player bandits in the last $10$ years, but especially active since last spring $2018$.
 For instance, our article \cite{Besson2018ALT} was the first to propose the ``no sensing'' case, and it was studied by (at least) two independent group of researchers since its publication in April $2018$ \cite{BoursierPerchet18,LugosiMehrabian18}.
@@
 \subsection{With malicious jammers}
-\label{sub:5:withDifferentMeansAmongPlayers}
+\label{sub:5:withMaliciousJammers}
 

### 2019-05-24 2-Chapters/Conclusion/conclusion.tex
SHIT I'm going home

@@
 
-\begin{itemize}
-    \item
+% \begin{itemize}
+%     \item
 The first half, Part~\ref{part:Introduction}, starts by motivating our work and situating the context of this research.
@@
 
-    \item
+    % \item
 The second half of this thesis, Part~\ref{part:MABIOT}, presents our main contributions.
@@
 For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
-\end{itemize}
+% \end{itemize}
 

### 2019-05-24 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
DONE

@@
 % - explain the problem studied in the simulation
-We focus on \emph{i.i.d.} MAB problems, with $K = 9$ channels\footnote{~Similar behaviors are observed for any not-too-large values of $K$, we tried up-to $K = 100$ and the same results were obtained.}.
+We focus on \emph{i.i.d.} MAB problems, with $K = 9$ channels\footnote{~Similar behaviors are observed for other values, \eg, trying up-to $K = 100$ gave the same results.}.
 For Bernoulli problem, the first one uses $\boldsymbol{\mu}=[0.1,\dots,0.9]$, and is considered as a ``simple'' problem.
@@
 % ----------------------------------------------------------------------
-\paragraph{No theoretical guarantees yet}\label{sub:25:theory}
+\paragraph{No theoretical guarantees}\label{sub:25:theory}
 
@@
 for both Bernoulli and non-Bernoulli problems (\eg, Gaussian, exponential and Poisson distributions).
-Further theoretical developments are left as future work.
+Further theoretical developments were left as future work.
+
+\TODOL{Expliquer qu'en fait on NE PEUT PAS obtenir la borne qu'on espérait...}
 

### 2019-05-24 2-Chapters/2-Chapter/chapter2.tex
DONE

@@
 %
-We describe our library SMPyBandits in more details in Chapter~\ref{chapter:3}, but all the described algorithms are implemented and documented in SMPyBandits, alongside with many more algorithms (there were about 65 for single-player stochastic problems at the time of writing this thesis).
+We describe our library SMPyBandits in more details in Chapter~\ref{chapter:3}, but all the algorithms described in this chapter are implemented in SMPyBandits, in the \texttt{Policies} module, alongside with many more algorithms (there are about 65 for single-player stochastic problems).
 A complete list of the implemented policies can be found on the following web page on the documentation,
-\href{https://smpybandits.github.io/docs/Policies.html}{\texttt{SMPyBandits.GitHub.io/docs/Policies.html}}, and on the \texttt{Policies} module.
+\href{https://smpybandits.github.io/docs/Policies.html}{\texttt{SMPyBandits.GitHub.io/docs/Policies.html}}.
 

### 2019-05-27 2-Chapters/1-Chapter/chapter1.tex
...

@@
 % =============================================================================
-% \subsection{French language conference with proceedings}
-\subsection{Submitted works}
+\subsection{French language conference with proceedings}
+% \subsection{Submitted works}
 
@@
     \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
-    submitted at GRETSI 2019\footnote{~GRETSI 2019, see \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
+    by \textbf{Lilian Besson} \& Emilie Kaufmann,
+    GRETSI 2019\footnote{~GRETSI 2019, see \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
+    August $2019$,
     \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.

### 2019-05-27 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
...

@@
 \ExpQ{} satisfies a bound on adversarial regret, % $\overline{R_T}$
-$\overline{R_T} \leq 2 \sqrt{T N \log(K)$,
+$\overline{R_T} \leq 2 \sqrt{T N \log(K)}$,
 with the good choice of the learning rate sequence $(\eta_t)_{t \geq 1}$.

### 2019-05-27 4-Listofpublications/Listofpublications.tex
Talk to Emilie
SHIT I'm going home
...

@@
 
-\TODOL{J'ai déjà inclus une liste de publication à la fin du Chapitre 1, nécessaire ici aussi ?}
+% \TODOL{J'ai déjà inclus une liste de publication à la fin du Chapitre 1, nécessaire ici aussi ?}
 
@@
 \item
+  \href{https://hal.inria.fr/hal-XXX/document}{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation}, by \href{https://moychris.wordpress.com/}{C. Moy} \& \textbf{L. Besson},
+  \href{https://sites.google.com/view/ISIoT2019/}{ISIoT Workshop} at \href{https://dcoss.org/}{DCOSS}, Santorini, Greece, May 2019.
+  % \href{https://hal.inria.fr/hal-XXX}{HAL-XXX}.
+  \cite{MoyBesson2019}
+\item
   \href{https://perso.crans.org/besson/articles/BMBBM__IEEE_WCNC__2019.pdf}{Upper-Confidence
@@
 \item
+  \href{https://perso.crans.org/besson/articles/BBM__Demo_ICT_2018.pdf}{MALIN:
+  « Multi-Arm bandit Learning for Iot Networks » with GRC: A TestBed
+  Implementation and Demonstration that Learning Helps}, by \textbf{L.
+  Besson}, \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
+  \href{https://moychris.wordpress.com/}{C. Moy}. Demonstration
+  presented at ICT (\emph{International Conference on Communication}),
+  Saint-Malo, France, June 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}, \href{https://hal.inria.fr/hal-02013866}{HAL-02013866}.
+  \cite{Besson2018ICT}
+\end{etaremune}
+
+\section*{International Conferences}
+\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
+\begin{etaremune}
+\item
   \href{https://perso.crans.org/besson/articles/BBM__IEEE_WCNC_2019.pdf}{GNU
@@
   \cite{Besson2019WCNC}
-\item
-  \href{https://perso.crans.org/besson/articles/BBM__Demo_ICT_2018.pdf}{MALIN:
-  « Multi-Arm bandit Learning for Iot Networks » with GRC: A TestBed
-  Implementation and Demonstration that Learning Helps}, by \textbf{L.
-  Besson}, \href{https://remibonnefoi.wordpress.com/}{R. Bonnefoi},
-  \href{https://moychris.wordpress.com/}{C. Moy}. Demonstration
-  presented at ICT (\emph{International Conference on Communication}),
-  Saint-Malo, France, June 2018. Cf. \texttt{YouTu.be/HospLNQhcMk}, \href{https://hal.inria.fr/hal-02013866}{HAL-02013866}.
-  \cite{Besson2018ICT}
 \item
@@
 
-% \section*{National Conferences}
-% \renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
-% \begin{etaremune}
-% 	\item \textbf{L. Besson}, E. Kaufmann, ``XXX'', in \emph{GRETSI}, Lille, France, Aug 2019
-% \end{etaremune}
+\section*{National Conferences}
+\renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
+\begin{etaremune}
+\item
+  \href{https://hal.inria.fr/hal-02006471/document}{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires}, by \textbf{L. Besson} \&
+  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
+  Kaufmann}.
+  GRETSI, Lille, August 2019,
+  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
+  \cite{Besson2019Gretsi}
+\end{etaremune}
 
@@
 Besson}, E.
-Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}.
+Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, July 2017, \href{https://hal.inria.fr/hal-02013847}{HAL-02013847}.
   % \item \textbf{L. Besson}, E. Kaufmann, ``Multi-Players Bandits Revisited'', in \emph{SCEE Seminar}, Rennes, France, March 2017.
@@
 Bonnefoi, \textbf{L.
-Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016, \href{https://hal.inria.fr/hal-02013839}{HAL-02013839}.
+Besson}, ``Multi-Armed Bandit Learning in IoT Networks'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, June 2018, \href{https://hal.inria.fr/hal-02013839}{HAL-02013839}.
 \end{etaremune}
@@
 \begin{etaremune}
-\item
-  \href{https://hal.inria.fr/hal-XXX/document}{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation}, by \href{https://moychris.wordpress.com/}{C.
-Moy} \& \textbf{L.
-Besson}, March 2019.
-  % \href{https://hal.inria.fr/hal-XXX}{HAL-XXX}.
-  \cite{MoyBesson2019}
-\item
-  \href{https://hal.inria.fr/hal-02006471/document}{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires}, by \textbf{L. Besson} \&
-  \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{E.
-  Kaufmann}, March 2019,
-  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
-  \cite{Besson2019Gretsi}
 \item

### 2019-05-28 1-Preamble/4-Acknowledgement/acknowledgement.tex
Working to merge Christophe's feedback of Chapter5

@@
 % TODO financial thanks
-This work and my PhD is supported by
+My PhD was funded by the French Ministry of Higher Education and Research (MENESR),
+thanks to a ``contrat doctoral spécifique normalien'' obtained while I was a ``normalien'' student at \'Ecole Normale Sup\'erieure de Paris-Saclay.
+This work has been supported by
 the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{N ANR-16-CE40-0002}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
@@
 by CNRS under the PEPS project BIO (in 2017),
-by the French Ministry of Higher Education and Research (MENESR),
-by \'Ecole Normale Sup\'erieure de Paris-Saclay.
 by European Union, through the European Regional Development Fund (ERDF),
 and by Ministry of Higher Education and Research, Brittany and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
-% I was also lucky to be well received very frequently at Inria Lille Nord Europe, and to be hosted at CentraleSupélec campus of Rennes.
 
@@
 I'll try hard to do better from now on!
-I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} who gave me energy (\eg, cheese, chocolate, coffee, vegetables etc), and finally my \href{https://perso.crans.org/besson/}{favorite technological tools}: git, GitHub and Bitbucket, \LaTeX, Bash and Python, GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other free and open-source tools.
+I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} which gave me smiles and energy, and finally my \href{https://perso.crans.org/besson/}{favorite technological tools}: git, GitHub and Bitbucket ; \LaTeX, Bash, Julia and Python ; GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools\ldots{}
 
@@
 Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.
+
+\vfill{}
 

### 2019-05-28 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
My thesis is so long, and so stupid
Working to merge Christophe's feedback of Chapter5

@@
 % -----------------------------------------------------------------
+\section{Introduction}
+\label{sec:5:introduction}
+% -----------------------------------------------------------------
+
+For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1} and \ref{chapter:4}.
+%
+In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
+Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
+%
+The decentralized hypothesis means that the central base station does not given any direct order to the devices, they are all in charge of deciding the channel they each want to use at each time step.
+Moreover, the devices are independent and do not directly communicate with each other either.
+We show that dynamic end-devices are able to use limited feedback sent by the gateway to learn to find orthogonal affectations of the group of end-devices to the best radio channels, automatically and without explicit communication with each other.
+In other words, the end-devices are able to learn to cooperate efficiently, in a completely decentralized and autonomous manner.
+
+We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
+%
+For this first model, we present two algorithms, \RandTopM{} and \MCTopM, combining an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
+Like in previous works, we consider the centralized system regret (multi-player regret).
+We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
+We then analyze the \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
+%  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
+We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M \leq K$ accessing $K$ channels.
+
+We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
+Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
+Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, we conjectured that it suffers from linear regret, and it was later confirmed by two works inspired by our article \cite{LugosiMehrabian18,BoursierPerchet18}.
+They both proposed new algorithms, proved to achieve logarithmic regret under the no sensing model.
+We include numerical simulations to compare their proposals, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
+
+Finally, we review other extensions of the models, and give a overview of the current state-of-the-art.
+For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
+
+\paragraph{Outline.}
+%
+This chapter is organized as follows.
+First we motivate the model of multi-player multi-armed bandits in Section~\ref{sec:5:motivation}
+We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
+The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms},
+for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
+The result of our experimental study reported in Section~\ref{sec:5:experiments}.
+Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
+% and finally Section~\ref{sec:5:conclusion} concludes.
+
+% -----------------------------------------------------------------
 \section{Motivations for multi-player MAB models}
-\label{sec:5:introduction}
+\label{sec:5:motivation}
 % -----------------------------------------------------------------
@@
 
-For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1},\ref{chapter:4}.
-%
-In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
 We consider $M \leq K$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
@@
 %
-Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
-Unfortunately, it is very hard to formally analyze the IoT network model we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic learning devices.
+Unfortunately, it is very hard to formally analyze the IoT network models we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic (\ie, learning) devices.
 Even if there are many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step the number of communicating devices is random and unpredictable.
 
-Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider the easier case of at most $M \leq K$ devices, communicating at each time step.
-Each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
+Mainly for these two reasons, for this Chapter~\ref{chapter:5} we prefer to only consider at most $M \leq K$ devices, communicating at each time step.
+Note that in the model of Chapter~\ref{chapter:4}, this hypothesis $M \leq K$ corresponds to choosing a probability of activation of $p=1$, as we assumed that (in average) no more than $K$ devices should be active at the same step.
+%
+In our model in Chapter~\ref{chapter:5}, each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
 %
@@
 
-All these above-mentioned works are motivated by the OSA problem, in which it is assumed that \emph{sensing} occurs, that is each smart device observes the availability of a channel (sample from the arm) \emph{before} trying to transmit and possibly experiment a collision with other smart devices.
+All these above-mentioned works are motivated by the OSA problem, in which it is assumed that \emph{sensing} occurs, that is, each smart device observes the availability of a channel (\ie, a reward from the arm) \emph{before} trying to transmit and possibly experiment a collision with other smart devices.
 However some real radio networks do not use sensing at all, \eg, emerging standards currently or recently developed for \emph{Internet of Things} (IoT) networks, such as LoRaWAN.
@@
 for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
-Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can has a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-player MAB problems).
-We do not propose any other efficient algorithm, but our work presented in April~$2018$ inspired two articles shorlty after \cite{LugosiMehrabian18,BoursierPerchet18}.
+Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can have a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-player MAB problems).
+We do not propose any other efficient algorithm, but our work presented in April~$2018$ has strongly inspired two articles shorlty after \cite{LugosiMehrabian18,BoursierPerchet18}.
 They confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
@@
 
-\paragraph{Outline.}
-%
-The rest of the chapter is organized as follows.
-We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
-The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms},
-for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
-The result of our experimental study reported in Section~\ref{sec:5:experiments}.
-Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
-% and finally Section~\ref{sec:5:conclusion} concludes.
-
 % -----------------------------------------------------------------
@@
 
-\textbf{Only Bernoulli?}
+\textbf{Only Bernoulli.}
 However we mention that our lower bound and all our algorithms (and their analysis) can be easily extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}).
-For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio, as it can model channel availabilities.
+For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio problems, based on channel availability.
 
@@
 \end{equation}
-In words, she receives the reward of the selected arm if she is the only one to select this arm, and a reward zero otherwise.
+In other words, she receives the reward of the selected arm if she is the only one to select this arm, and a reward zero otherwise.
 This provides another reason to focus on the Bernoulli model.
-It is the hardest model, in the sense that receiving a reward zero is not enough to detect collisions.
-For other models, the data streams $(Y_{k,s})_s$ are usually continuously distributed, with no mass at zero.
+It is the hardest model, in the sense that receiving a reward zero is \emph{not enough} to detect collisions.
+For other models, the data streams $(Y_{k,s})_s$ are usually continuously distributed, with no probability mass at the value of zero (\eg, Gaussian).
 Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.
@@
 % -----------------------------------------------------------------
-\subsection{Common analysis for \RandTopM- and \MCTopM-\klUCB{}}\label{sub:5:UpperBoundSelections}
+\subsection{Common analysis for \RandTopM- and \MCTopM}\label{sub:5:UpperBoundSelections}
 
@@
 % -----------------------------------------------------------------
-\subsection{Regret analysis of \MCTopM-\klUCB}\label{sub:5:UpperBoundCollisions}
+\subsection{Regret analysis of \MCTopM{} with \klUCB{} indexes}\label{sub:5:UpperBoundCollisions}
 

### 2019-05-28 2-Chapters/5-Chapter/chapter5.tex
Working to merge Christophe's feedback of Chapter5

@@
 \label{chapter:5}
-\minitoc
 
-\TODOL{Intégrer les commentaires de Christophe, le plus rapidement possible !}
-
-\newpage
 \paragraph{Abstract}
-
+%
 In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
@@
 %
-We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
+We discuss three feedback levels that give variants of the multi-players MAB model previously studied in the literature.
+Using a decomposition of the centralized system regret, we start by explaining the intuition about what an efficient decentralized algorithm should do, and then we propose \RandTopM{} and \MCTopM, two new orthogonalization schemes.
+Combining them with the \klUCB{} index policy gives order-optimal algorithms, that achieve state-of-the-art performance,
+as illustrated by numerical similations.
+We conclude by reviewing different extensions of the multi-player models.
 
-For this first model, we present two algorithms, \RandTopM{} and \MCTopM, combining an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
-Like in previous works, we consider the centralized system regret (multi-player regret).
-We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
-We then analyze the \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
-%  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
-We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M \leq K$ accessing $K$ channels.
-
-We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
-Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
-Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, we conjectured that it suffers from linear regret, and it was later confirmed by two works inspired by our article \cite{LugosiMehrabian18,BoursierPerchet18}.
-They both proposed new algorithms, proved to achieve logarithmic regret under the no sensing model.
-We include numerical simulations to compare their proposals, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
-
-Finally, we review other extensions of the models, and give a overview of the current state-of-the-art.
-For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
+\minitoc
 

### 2019-05-28 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for changes on Ch5
- /!\ it's not shorter than before...

@@
 In the absence of sensing information, we propose the \Selfish{} heuristic and investigate its performance.
-Our study of this algorithm was supported in \cite{Besson2018ALT} by (promising) empirical performance and some first (disappointing) theoretical elements, and it was proved to be inefficient since then.
+Our study of this algorithm was supported in \cite{Besson2018ALT} by (promising) empirical performance and some first (disappointing) theoretical elements, and it was proven to be inefficient since then.
 
@@
 Hence receiving $r^j(t) = 0$ directly gives $\indic(C^j(t)) = 1$.
-
+%
 Note that other models for rewards loss have also been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it.
@@
 
-As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the ``collision information term'' is zero for any algorithm $\rho$.
-However, we wanted to be still able to present the derivations of our lower-bound, mainly because its proof and its form give important insights on how to design an asymptotically optimal algorithm $\rho$.
-% and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
+% As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the ``collision information term'' is zero for any algorithm $\rho$.
+% However, we wanted to be still able to present the derivations of our lower-bound, mainly because its proof and its form give important insights on how to design an asymptotically optimal algorithm $\rho$.
+% % and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
 
@@
 The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
-The theorem stated below is proved in the Appendix of \cite{Besson2018ALT},
+The theorem stated below is proven in the Appendix of \cite{Besson2018ALT},
 % \ref{proof:5:BetterLowerBound},
@@
 
-\paragraph{Price of decentralized learning?}
+\paragraph{Price of decentralized learning.}
 %
@@
 Indeed, for an algorithm to attain the lower bound \eqref{eq:5:ourLowerBound}, the number of selections of each sub-optimal arm should match the lower bound \eqref{eq:5:LBDraws} \emph{and} term \ref{eq:5:term2} and term \ref{eq:5:term3} in the regret decomposition of Lemma~\ref{lem:5:DecompositionRegret} should be negligible compared to  $\log(T)$.
+
 To the best of our knowledge, no algorithm has been shown to experience only $o(\log(T))$ collisions so far,
@@
 %
-Letting $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of sensing information obtained by player $j$ for arm $k$, $\widehat{\mu}_k^j(t) = S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$ for player $j$ and one can define the index
+Letting $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained by player $j$ for arm $k$,
+then $\widehat{\mu}_k^j(t) = S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$ for player $j$, and thus one can define the index
 \begin{equation}\label{eq:5:indexFor_UCB_klUCB}
@@
 
-\paragraph{Two ideas for orthogonalization?}
+\paragraph{Two ideas for orthogonalization.}
 \label{par:5:twoIdeasOrthogonalization}
@@
 that we can then bound by the number of changes of arms.
-%
-An interesting consequence of the proof of this result is that
-it also bounds the number of \emph{switches of arms}, $\sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$,
-and this additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
-Even though minimizing the number of arms switching was not a goal,
-this guarantee is interesting for Cognitive Radio applications,
-where switching arms means re-configuring a radio hardware, an operation that costs energy.
-An algorithm guaranteeing a small number of switches is interesting.
-%
-We note that introducing \emph{switching costs}, like it was done in previous works like \cite{Koren17}, is an interesting future work.
 
@@
 
-  And so we obtain the desired inequality, with explicit constants, that depend only on $\boldsymbol{\mu}$ and $M$.
+  Thus we prove the desired inequality, with explicit constants depending only on $\boldsymbol{\mu}$ and $M$.
   %
@@
 
+\paragraph{Logarithmic ``switching costs'' for \MCTopM.}\label{app:5:NumberSwitches}
+%
+Additionally to bounding the collisions in Lemma~\ref{lem:5:collisionsMCTopM},
+an interesting consequence of the proof of this result is that
+it also bounds the (expected) number of \emph{switches of arms}, $SC^{\cA}(T) = \sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$.
+%
+We controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
+which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
+Thus, the total number of arm switches is also proven to be logarithmic, if all players uses the \MCTopM-\klUCB{} algorithm.
+This additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
+%
+Even though minimizing the number of arms switching was not a goal,
+this guarantee is interesting, in particular for Cognitive Radio applications,
+where switching arms means re-configuring a radio hardware, an operation that costs energy.
+An algorithm guaranteeing a small number of switches is thus interesting,
+and for instance \cite{modiDemo2016} studied the empirical impact of the number of hardware reconfigurations on the battery life of a dynamic end-devices.
+This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
+%
+We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17}, is an interesting future work.
+
 % -----------------------------------------------------------------
@@
 Now that the sub-optimal arms selections and the collisions
-are both proved to be at most logarithmic in Lemmas~\ref{lem:5:SubOptimalSelections} and \ref{lem:5:collisionsMCTopM},
+are both proven to be at most logarithmic in Lemmas~\ref{lem:5:SubOptimalSelections} and \ref{lem:5:collisionsMCTopM},
 it follows from our regret decomposition (Lemma~\ref{lem:5:DecompositionRegret}) together with Lemma~\ref{lem:5:1stUpperBound} that the regret of \MCTopM-\klUCB{} is logarithmic.
@@
 
-\paragraph{Number of switches}\label{app:5:NumberSwitches}
-%
-Note that we controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
-which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
-Thus, the total number of arm switches is also proved to be logarithmic, if all players uses
-the \MCTopM-\klUCB{} algorithm.
+% \paragraph{Number of switches}\label{app:5:NumberSwitches}
+% %
+% Note that we controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
+% which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
+% Thus, the total number of arm switches is also proven to be logarithmic, if all players uses
+% the \MCTopM-\klUCB{} algorithm.
 
@@
 %
-As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proved
+As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proven
 to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
@@
 For example with $M=2$ or $M=3$ players competing for $K=3$ arms,
-with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, the histograms in Figure~\ref{fig:5:selfish_fail1} suggests that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} can be very large.
+with means $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, the histograms in Figure~\ref{fig:5:selfish_fail1} suggest that with a small probability, the regret $R_T$ of \Selfish-\klUCB{} can be very large.
 
@@
 
+In all the numerical simulations, we focus on the model with sensing, \ie, model \modeldeux{} or \modeltrois.
+%
 We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \UCB{} or \klUCB{} indices.
@@
 
-We report in Table~\ref{table:5:comparisonUCB_klUCB} below the results in terms of mean regret, $\pm$ one standard deviation.
+We report in Table~\ref{table:5:comparisonUCB_klUCB} below the results in terms of mean regret, plus or minus one standard deviation.
 The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
@@
 \MusicalChair{} consists of a random exploration phase of length $T_0$ after which the players quickly converge to orthogonal strategies targeting the $M$ best arms.
-With probability $1-\delta$, its regret is proved to be ``constant'' (of order $\log(1/\delta)$).
+With probability $1-\delta$, its regret is proven to be ``constant'' (of order $\log(1/\delta)$).
 The theoretical minimal value for $T_0$ depends on $\delta$, on the horizon $T$ and on a lower bound $\epsilon$ on the gap $\Delta = \mu^*_M - \mu^*_{M+1}$, and the practical tuning is hard too.

### 2019-05-28 2-Chapters/5-Chapter/chapter5.tex
Done for changes on Ch5
- /!\ it's not shorter than before...

@@
 Combining them with the \klUCB{} index policy gives order-optimal algorithms, that achieve state-of-the-art performance,
-as illustrated by numerical similations.
+as illustrated by numerical simulations.
 We conclude by reviewing different extensions of the multi-player models.
@@
 In the model we presented above, we assumed the number of players $M$ to be fixed in time during the learning process.
-Without removing this hypothesis (we study this case below in Section~\ref{sub:5:arrivalDepartures}), it is interesting to remark that we made no hypothesis on whether the players can known this value $M$ or not.
+Without removing this hypothesis (we study this case below in Section~\ref{sub:5:arrivalDepartures}), it is interesting to remark that we made no hypothesis on whether the players can know this value $M$ or not.
 Our proposals, \RandTopM{} and \MCTopM, both assume to know $M$ beforehand, like it was done for their main inspiration, \RhoRand.
 
-\paragraph{Performance of our proposals for a wrong value of $M$?}
+\paragraph{Performance of our proposals for a wrong value of $M$.}
 \label{par:5:usingWrongValueofM}
@@
 Then in such networks, if one wants to apply an algorithm like \MCTopM-\klUCB,
-the $M$ devices need to known their number $M$, and have to receive it from the gateway, as it is the only part of this example of network which knowns $M$.
+the $M$ devices need to know their number $M$, and have to receive it from the gateway, as it is the only part of this example of network which knowns $M$.
 It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in[K]$.
@@
 We consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
-As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require the know $M$.
+As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require to know $M$.
 Then we compare the \RhoRand, \RandTopM{} and \MCTopM{} algorithms, using \klUCB, which know $M$ beforehand, with their extensions implementing the same algorithm as \rhoRandEst, to estimate $M$ on the fly.
@@
 
-As reminded above, the model assumes the number of players $M$ remain fixed during all the bandit game.
+As reminded above, the model assumes that the number of players $M$ remains fixed during all the bandit game.
 However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
@@
 Every time step, another player is either arriving, without any knowledge of the problem (\eg, it is a fresh and new IoT device).
-Then no matter the strategy of these news players, even if they all play \MCTopM-\klUCB{} for instance, if they play a uniformly-efficient strategy there is a non-zero probability that player $1$ suffer from a collision at each time step, thus resulting in linear centralized system regret.
+Then no matter the strategy of these new players, even if they all play \MCTopM-\klUCB{} for instance, if they play a uniformly-efficient strategy there is a non-zero probability that player $1$ suffers from a collision at each time step, thus resulting in linear centralized system regret.
 
@@
 
-We give in Table~\ref{table:5:meanRegretSimulationsNoSensing}, we give the mean regret obtained for these different algorithms, and we observe a drastic difference between the unrealistic centralized algorithm, which achieves a very small regret, the heuristic \Selfish-\klUCB{} which achieve small mean regret (but is small to fail in theory and in some instances), and the two other algorithms.
+We give in Table~\ref{table:5:meanRegretSimulationsNoSensing} the mean regret obtained for these different algorithms, and we observe a drastic difference between the unrealistic centralized algorithm, which achieves a very small regret, the heuristic \Selfish-\klUCB{} which achieve small mean regret (but is small to fail in theory and in some instances), and the two other algorithms.
 We were unable to find any bug in our implementation of Improved \MusicalChair{} from \cite{LugosiMehrabian18} and all the different tuning of $c$ (or $g$) explored gave the same disappointing result (linear regret).
@@
 In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
+This hypothesis is realistic, mostly for energy consumption and spectrum efficiency reasons.
 Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between devices are possible.
@@
 Clearly, the problem is easier by allowing communication, and it is quite immediate to see that the communication capacity between players must be limited otherwise the problem is already known and solved.
-Indeed, imagine that at each time step $t$, all the $M$ players could share an unbounded number of bits with the other players, at no cost, even if this hypothesis is clearly unrealistic for wireless networks.
+Indeed, imagine that at each time step $t$, all the $M$ players could share an unbounded number of bits with the other players, at no cost (even if this hypothesis is clearly unrealistic for wireless networks).
 Then they can share all their observations, and they can all run the same multiple-plays MAB algorithm \cite{Anantharam87a}, like for instance the extension of Thompson sampling for multiple-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
-In this setting, a logarithmic regret is obtained easily, and the regret upper-bound of the two aforementioned algorithms asymptotically achieve the lower-bound from \cite{Anantharam87a}.
+In this setting, a logarithmic regret is easily obtained, and the regret upper-bound of the two aforementioned algorithms asymptotically achieve the lower-bound from \cite{Anantharam87a}.
 
-An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit, at every or only some time steps.
+A more interesting extension is thus to limit the communication between players, either to a small number of bits or just one bit, at every or only some time steps.
 We identified that the state-of-the-art on this direction of research consists in the two very recent works
@@
 This question was first answered in \cite{Bistritz18}, and proposed an algorithm based on alternating three phases, and increasing their lengths after the end of each epoch ($2^p$ for the $p$-th epoch).
-First, players explore in order to estimate the expectations of the arm rewards ; then players use their ``Game of Thrones'' dynamics (inspired by Musical Chair \cite{Rosenski16}) and play the optimal solution most of the time ; finally players play the action they played most of the time in the recent GoT phases.
+First, players explore in order to estimate the expectations of the arm rewards ; then players use their ``Game of Thrones'' dynamics (GoT, inspired by Musical Chair \cite{Rosenski16}) and play the optimal solution most of the time ; finally players play the action they played most of the time in the recent GoT phases.
 % combining Musical Chair with an epoch-based alternative phases of forced exploration and pure exploitation.
-They analyze their ``Game of Thrones'' algorithm and proved that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
+They analyze their GoT algorithm and proved that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
 
 Until the very recent work of \cite{KaufmannAbbas19}, it was unknown if a logarithmic regret was possible.
-They proposed an algorithm that is based on two phases: first, players will learn $M$ and learn orthogonal ranks, using the ``communication trick'' by \cite{BoursierPerchet18}, and then one player is elected as a leader and the others are followers.
+They proposed an algorithm that is based on two phases: first, players will learn $M$ and learn orthogonal ranks, using the ``communication trick'' of~\cite{BoursierPerchet18}, and then one player is elected as a leader and the others are followers.
 This second step, after initialization, is also using epoch of increasing lengths $2^p$.
@@
 Similarly to our demonstration of multi-armed bandit learning in an IoT network that we presented in the previous Chapter~\ref{chapter:4} in Section~\ref{sec:4:gnuradio},
-we list here some related works, dating back from $2016$, who proposed a similar approach.
+we list here some related works, dating back from $2016$, who proposed a similar approach for OSA.
 All the similar works that we are aware of also used USRP boards, and implemented the demonstration using either the Simulink and MATLAB softwares, or the GNU Radio software like we did \cite{Besson2018ICT}.
+%
+Indeed, the demonstrations and the works presented below are coming from our team experiments, and have been extended by the team of Sumit Darak at IIIT Delhi after his postdoctoral fellowship in our team.
 
@@
 But instead of selecting a new rank uniformly at random among $[M]$, they proposed to use a ``second-stage'' learning policy\footnote{This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.}, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
-This algorithm is implemented in SMPyBandits and named \rhoLearn\footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.}, where it can use Bayes-UCB or any of the $65$ or more bandit algorithms available in the library.
+This algorithm is implemented in SMPyBandits and named \rhoLearn,
+% \footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.},
+where it can use Bayes-UCB or any of the $65$ or more bandit algorithms available in the library.
 Even if no theoretical guarantee backs up this idea of second-stage learning with Bayes-UCB, empirical simulations found that it can be efficient.
@@
 % Other works in the same line of research included \cite{modiDemo2016} implemented using the same USRP testbed as for our demonstration in Chapter~\ref{chapter:4} (Section~\ref{sec:4:gnuradio}),
-% and the first work of a research team based in IIT Dehli, \cite{kumar2016two}.
+% and the first work of a research team based in IIIT Dehli, \cite{kumar2016two}.
 % %
@@
 
-When we were working on multi-players bandits in Autumn $2017$ \cite{Besson2018ALT}, we left the study of the piece-wise stationary case as a future work, and shortly after we were excited to see that two independent works tackling this question were published, in December $2018$ and in February $2019$.
-Without diving too much into the details, we review here these two works.
+When we were working on multi-players bandits in Autumn $2017$ \cite{Besson2018ALT}, we left the study of the piece-wise stationary case as a future work, and shortly after we were excited to see that three independent works tackling this question were published, in December $2018$ \cite{WeiSrivastava18Distributed} and in February $2019$ \cite{AlaturLevyKrause19,bande2019adversarial}.
+Without diving too much into the details, we review here these three works.
 Without a more serious analysis, we conjecture that it should not be difficult to merge the regret lower-bound for stationary \emph{multi-players} MAB (Theorem~\ref{thm:5:BetterLowerBound}) and the lower-bound for \emph{piece-wise stationary} single-player MAB from \cite{Garivier11UCBDiscount}.
@@
 In this second work also, we can note the poor experimental section, as the proposed algorithm is only compared against the \MusicalChair{} algorithm, in ``easy'' problems (small number of break-points $\Upsilon_T$).
-Despite being more complicated than other approaches, their algorithm ``\texttt{C \& P}'' should not be too hard to implement ourself, and studying its empirical performances is an interesting future work.
+Despite being more complicated than other approaches, their algorithm ``\texttt{C \& P}'' should not be too hard to implement by ourselves, and studying its empirical performance is an interesting future work.
 
@@
 
-We have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
+The experiments in this chapter use SMPyBandits,
+and we refer to the following page for details on how to reproduce them,
+\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}}
+
+Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
 We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future work left on our library.
@@
 
-\newpage  % FIXME
+% \newpage  % FIXME
 % ----------------------------------------------------------------------------
@@
 
-To sum up, we presented in this Chapter three variants of Multi-Player Multi-Arm Bandits,
-with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
-For the two easiest models --with sensing--, our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret.
+To sum up, we presented in this chapter three variants of Multi-Player Multi-Arm Bandits,
+with different levels of feedback being available to the decentralized players, under which we proposed efficient algorithms.
+The two easiest models are the ones with sensing information (\ie, for OSA), for which our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret.
 In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
@@
 
-This Chapter suggests several interesting further research directions.
+This chapter suggests several interesting further research directions.
 First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
@@
 
-\paragraph{A note on the simulation code.}
-%
-All the experiments used for this chapter were performed using SMPyBandits,
-and we refer to the following page for details on how to reproduce them,
-\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}}
+% \paragraph{A note on the simulation code.}
 

### 2019-05-28 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done for merging Xtof's comments on Ch4

@@
 This second approach is ideal, but not realistic in decentralized IoT networks, and it is just used here as a reference.
-Then, these three reference scenarios allow to evaluate the performance of \UCB{} and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
+Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
 We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.
@@
 % (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
-The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
+The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
+We mean here by ALOHA-based that devices do not use sensing, and we do not consider retransmissions of packets (see Section~\ref{sec:4:retransmissions} below where we remove this last hypothesis).
 %
@@
 In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
-This is realistic in IoT networks such as networks based on LoRa.
+This is realistic in IoT networks such as networks based on LoRaWAN.
 % (of $1\;\mathrm{s}$ in our implementation).
@@
     \emph{Dynamic} (or \emph{smart}) end-devices can use all the available channels, by quickly changing their communication channel at any time slot, following a Machine Learning policy.
-    For that end, they use their history of past communication successes or failures they experienced in each channel, in order to learn about channel availability.
+    For that end, they use their history of past communication successes or failures they experienced in each channel, to learn about channel availability.
     We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
-    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bits IEEE float number for each of the $K$ channels).
+    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bit IEEE 754 float number for each of the $K$ channels).
     Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, to improve their successful communication rate automatically.
@@
 
-We further assume that there are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
-Furthermore, in channel $k \in \llbracket 1; K \rrbracket$ there are $0 \leq S_k \leq S$ static devices (so $S = \sum_{k=1}^{K} S_k$).
+We further assume that there are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices
+with $0 \leq S_k \leq S$ static devices in channel $k \in [K]$ (so $S = \sum_{k=1}^{K} S_k$).
 
@@
 
-We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; K \rrbracket$).
+We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in [K]$).
 This reflects a naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.

### 2019-05-28 2-Chapters/4-Chapter/chapter4.tex
Small proofreading of notations of Chapter6
My thesis is so long, and so stupid
Done for merging Xtof's comments on Ch4

@@
 \minitoc
+% Write miniTOC just after the title
+\graphicspath{{2-Chapters/4-Chapter/Images/}}
 % ----------------------------------------------------------------------
+
+\TODOL{Intégrer les commentaires de Christophe, le plus rapidement possible !}
 
@@
 \newpage
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/4-Chapter/Images/}}
 
@@
 Kaufmann12Thompson}, for the first model.
-For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.
+For the demonstration as well as for the second model, without loss of generality, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.
 
@@
 
-% \newpage  % WARNING ?
+\newpage  % WARNING ?
 % ----------------------------------------------------------------------------

### 2019-05-28 2-Chapters/2-Chapter/chapter2.tex
On the one hand… on the other hand
- NOT on one hand
- NOT on the first hand

@@
 
--- On the first hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
+-- On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
 Mathematically, both quantities are equal in expectation, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition}.
@@
 %
-On the first hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
+On the one hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
 It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the \UCB{} algorithm (or other low-cost algorithms).

### 2019-05-28 2-Chapters/3-Chapter/chapter3.tex
Suffit pour ce soir…
On the one hand… on the other hand
- NOT on one hand
- NOT on the first hand

@@
 (\eg, $\UCB_1$ is \texttt{Policies.UCBalpha}).
+
+\TODOL{Si j'en parle plus dans le chapitre 2, je peux aller très vite sur cette partie !}
 
@@
 \end{itemize}
-Thus, on the first hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
+Thus, on the one hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
 %
@@
 For all the different numerical simulations presented in this thesis, except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4}, the setting is the same: we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=100$ like in Chapter~\ref{chapter:6}).
-On the first hand, because of the fundamentally sequential nature of bandit games, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
+On the one hand, because of the fundamentally sequential nature of bandit games, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
 On the other hand, parallelism can help greatly for the two other axes: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.

### 2019-05-28 2-Chapters/4-Chapter/chapter4.tex
On the one hand… on the other hand
- NOT on one hand
- NOT on the first hand

@@
 
-On the first hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
+On the one hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
 On the other hand, the second model is more interesting if the main target is an improvement of the QoS of the IoT networks rather than a longer battery life.

### 2019-05-28 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
On the one hand… on the other hand
- NOT on one hand
- NOT on the first hand

@@
 %
-On the first hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
+On the one hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
 Even though some alternative without pre-agreement are proposed, they are quite complicated and we prefer to focus on other approaches.

### 2019-05-28 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
On the one hand… on the other hand
- NOT on one hand
- NOT on the first hand

@@
 Drawing conclusions on their behavior from a single simulation is tricky, as this example of behavior is counter intuitive:
-On the first hand, \CUSUM{} uses local restarts, and it should be less efficient than global restarts for this problem, as changes are all global.
+On the one hand, \CUSUM{} uses local restarts, and it should be less efficient than global restarts for this problem, as changes are all global.
 On the other hand, \MklUCB{} uses global restarts, but here it fails to detect any changes.

### 2019-05-28 3-Appendices/2-Appendix/appendix2.tex
On the one hand… on the other hand
- NOT on one hand
- NOT on the first hand

@@
 We give two examples to illustrate where this prior knowledge can be realistic or not.
-On the first hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for instance we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
+On the one hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for instance we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
 On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices: usually a learning step corresponds to an uplink and then downlink message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages.

### 2019-05-28 2-Chapters/2-Chapter/chapter2.tex
SHIT

@@
 the logarithmic upper-bound is useless in such case, as
-\[ \varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T. \]
+$\varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T$.
 
@@
 
-\TODOL{FIXME explain why this is interesting for the rest of this manuscript!}
+% \TODOL{FIXME explain why this is interesting for the rest of this manuscript!}
 
@@
 Optimizing on $T_0$ gives $T_0 = 4/d^2 \log(d^2 T / 4)$
-and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$.
+and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.
+%
+This strategy is also referred to as ``explore-first'', and in the case without prior knowledge on the problem, it can achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ as shown in Section~1 of \cite{Slivkins2019}.
 
@@
 
-    \item Algorithm 6 is
-    Thompson sampling from \cite{Kaufmann12Thompson}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{Thompson}).
-    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
-    It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on C or Cython code, highly optimized).
-
-    \item Algorithm 7 is
-    Bayes-UCB from \cite{Kaufmann12BUCB}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
-    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
-    It is comparable to Thompson sampling in terms of memory complexity, but slower in terms of time complexity as computing a $1-1/t$ quantile is more costly than sampling from a distribution.
-    But in particular for Bernoulli distributed arms and for Beta posteriors and priors, Bayes-UCB is only about twice as slow as Thompson sampling.
-
-    \item Algorithm 8 is
-    AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
-    It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posterior on each arm.
-It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson decision rule (\ie, choose the arm maximizing one draw of random samples from the posteriors), otherwise it uses the Bayes-UCB decision rule (\ie, choose the arm maximizing the $1-1/t$ quantile).
-    It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
-    In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
-
-    \item Algorithm 9 is
-    BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, and it consists in trying to fix a simple observation.
-    For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
-    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
-    It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
-    Analyzing it in the generic case was found to be difficult, and is still an open problem.
-    We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions, \ie, using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
-    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffers from an exponential blow-up when $K$ increases.
 \end{itemize}
@@
 
-\subsubsection{Other policies inspired by UCB}
+\subsubsection{Variants on UCB}
 
@@
 
-% ----------------------------------------------------------------------------
-\subsection{Bayesian policies: Thompson Sampling and others}
-
-Idea behind Thompson sampling and Bayesian philosophy.
-
-The Beta posterior and others (https://smpybandits.github.io/docs/Policies.Posterior.html) and Thompson Sampling.
-
-% ----------------------------------------------------------------------------
-\subsection{Other policies: BESA and others}
-
-It's a good opportunity to show that I worked on my own on some non-standard policies.
-
-- One of the first paper I read in 2016 was defining AdBandits, https://smpybandits.github.io/docs/Policies.AdBandits.html, a mixture between Thompson sampling and BayesUCB (very efficient in practice)
-- I especially love the BESA algorithm by [? and ? and Maillard].
-
-% ----------------------------------------------------------------------------
-\subsection{Policies for adversarial bandits: Exp3 and others}
+\paragraph{Other policies inspired by UCB}
+%
+The ``Successive Elimination'' policy also uses confidence bounds, and use both UCB and Lower Confidence Bounds (LCB).
+After playing each active arms once, it eliminates arm $a$ if it is statistically safe that it is sub-optimal, \ie, if there is another arm $a' \neq a$ such that $\UCB_a(t) < \mathrm{LCB}_{a'}(t)$.
+It also achieves a $\bigO{\log(T)}$ problem-dependent regret upper bound, and a $\bigO{\sqrt{K T \log(T)}}$ problem-independent bound.
+
+% ----------------------------------------------------------------------------
+\subsection{Bayesian policies}
+
+% Idea behind Thompson sampling and Bayesian philosophy.
+
+% The Beta posterior and others (https://smpybandits.github.io/docs/Policies.Posterior.html) and Thompson Sampling.
+
+Thompson sampling is the oldest known MAB algorithms, dating back from the first paper by Thompson in 1933 \cite{Thompson33}.
+It became popular since its asymptotical and finite-time analyses in \cite{AgrawalGoyal11,Kaufmann12Thompson}.
+%
+Thompson sampling uses a Bayesian point of view:
+instead of building confidence intervals on the means of the arms like \UCB{} does,
+it starts with a prior on the $K$ distributions (\eg, a flat uniform prior), and then after every observation $(k, r(t))$ it updates the posterior distribution of the arm $k$ with a new data $r(t)$.
+If $P_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim P_k(t)$, and then playing according to the ``optimism in face of uncertainty'' philosophy if these samples represent the arms (expected) quality: $C(t) \argmax_k s_k$.
+
+For problems with Bernoulli distributions on the $K$ arms,
+the best choice of prior is a Beta distribution, that starts as an initially uniform prior (\ie, $\mathrm{Beta}(1,1)$),
+and maintain a distribution $\mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$ after $X_k(t)$ successes observed on arm $k$ and $N_k(t)-X_k(t)$.
+%
+In this setting as well as for Gaussian distributions,
+Thompson sampling was proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+
+Bayes-UCB from \cite{Kaufmann12BUCB} also uses the Bayesian point of view, updated like for Thompson sampling, but used in a different way.
+Instead of sampling from the posteriors and playing the arms with maximum sample, Bayes-UCB computes the $1-1/t$ quantile of each arm, at time $t$, and plays the arm with the largest quantile.
+It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
+%
+Bayes-UCB is comparable to Thompson sampling in terms of memory complexity, but it can be slower in terms of time complexity as computing a $1-1/t$ quantile is generally more costly than sampling from a distribution.
+% a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
+% But in particular for Bernoulli distributed arms and for Beta posteriors and priors, Bayes-UCB is only about twice as slow as Thompson sampling.
+
+For curiosity, we mention a third Bayesian algorithm, AdBandits introduced in \cite{Truzzi13}, mainly because it has not gained any popularity despite its good empirical performance.
+%  using $\alpha=1$ (\texttt{AdBandits}).
+AdBandits follows the same assumptions as Thompson sampling and Bayes-UCB, and for instance for Bernoulli problems it uses a Beta posterior on each arm.
+It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson sampling decision rule, otherwise it uses the Bayes-UCB decision rule.
+It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given.
+%  and it did not gain popularity since then.
+In practice, AdBandits is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
+
+Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that lies between the principle of optimism in the face of uncertainty and the Bayesian point-of-view.
+%  underlying the approaches based on UCB, might not be necessary at all.
+% With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
+They showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as \UCB{} or Thompson sampling, with uniform perturbations.
+Their RCB algorithm ressembles \UCB, and builds the same confidence intervals $\mathrm{LCB}_k(t), \UCB_k(t)$ on arm $k$ at time $t$.
+But instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
+% https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html
+
+% ----------------------------------------------------------------------------
+\subsection{Other policies}
+
+We briefly mention other families of policies.
+
+% It's a good opportunity to show that I worked on my own on some non-standard policies.
+
+% - One of the first paper I read in 2016 was defining AdBandits, https://smpybandits.github.io/docs/Policies.AdBandits.html, a mixture between Thompson sampling and BayesUCB (very efficient in practice)
+% - I especially love the BESA algorithm by [? and ? and Maillard].
+
+BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, and it consists in trying to fix a simple observation.
+For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
+The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
+It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
+Analyzing it in the generic case was found to be difficult, and it is still an open problem.
+% We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions, \ie, using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
+% By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffers from an exponential blow-up when $K$ increases.
+
+% ----------------------------------------------------------------------------
+\paragraph{Policies for adversarial bandits}
 
@@
 
-Softmax: is it the same, but written differently?
-https://smpybandits.github.io/docs/Policies.Softmax.html
-
-% ----------------------------------------------------------------------------
-\subsection{Hybrid policies: MOSS, klUCB++ and klUCB-Switch and others}
-
-Recent work.
-https://smpybandits.github.io/docs/Policies.MOSS.html
-https://smpybandits.github.io/docs/Policies.IMED.html
-https://smpybandits.github.io/docs/Policies.klUCBPlusPlus.html
-https://smpybandits.github.io/docs/Policies.klUCBSwitch.html
-
-Talk about these?
-https://smpybandits.github.io/docs/Policies.OCUCB.html
-https://smpybandits.github.io/docs/Policies.OSSB.html
-
-With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
-Randomized Confidence Bound (RCB) with uniform perturbations performs as well as UCB!
-https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html
-
-% \subsection{Summary of classification of the most famous algorithms}
-
-% A huge table in one page in portrait mode, giving:
-% - name
-% - reference, year
-% - family, classification
-% - time/memory complexity ?
-% - regret bound for stochastic regret
-
-% of about 20/30 algorithms?
+It is a recent variant of the $\mathrm{Exp3}$ algorithm
+is $\mathrm{Exp}3^{++}$ from \cite{Seldin17},
+and it uses an adaptive tuning of its parameters.
+It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+
+% Softmax: is it the same, but written differently?
+% https://smpybandits.github.io/docs/Policies.Softmax.html
+
+% ----------------------------------------------------------------------------
+\paragraph{Hybrid policies}
+
+As explained before, in the last few years, the MAB research community has been interested in finding algorithms which achieves both a problem-dependent logarithmic and a minimax upper-bounds.
+The first example is MOSS from \cite{Audibert2009minimax}, or recently \KLUCBpp{} in \cite{Menard17}.
+They are both index-based, using XXX
+
+More recently, the \klUCB-switch from \cite{Garivier18} XXX
+
+% Recent work.
+% https://smpybandits.github.io/docs/Policies.MOSS.html
+% https://smpybandits.github.io/docs/Policies.IMED.html
+% https://smpybandits.github.io/docs/Policies.klUCBPlusPlus.html
+% https://smpybandits.github.io/docs/Policies.klUCBswitch.html
+
+% Talk about these?
+% https://smpybandits.github.io/docs/Policies.OCUCB.html
+% https://smpybandits.github.io/docs/Policies.OSSB.html
 

### 2019-05-28 2-Chapters/3-Chapter/chapter3.tex
SHIT

@@
     Thompson sampling from \cite{Kaufmann12Thompson}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{Thompson}).
-    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+    % It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
     It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on C or Cython code, highly optimized).
@@
     Bayes-UCB from \cite{Kaufmann12BUCB}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
-    It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
+    % It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
     It is comparable to Thompson sampling in terms of memory complexity, but slower in terms of time complexity as computing a $1-1/t$ quantile is more costly than sampling from a distribution.
@@
     AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
-    It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posterior on each arm.
+    % It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posterior on each arm.
 It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson decision rule (\ie, choose the arm maximizing one draw of random samples from the posteriors), otherwise it uses the Bayes-UCB decision rule (\ie, choose the arm maximizing the $1-1/t$ quantile).
-    It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
+    % It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
     In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
@@
     \item Algorithm 9 is
-    BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, and it consists in trying to fix a simple observation.
-    For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
-    The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
-    It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
-    Analyzing it in the generic case was found to be difficult, and is still an open problem.
-    We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions, \ie, using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
-    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffer from an exponential blow-up when $K$ increases.
+    BESA (Best Empirical Sampled Average) from \cite{Baransi2014},
+    % , and it consists in trying to fix a simple observation.
+    % For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
+    % The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
+    % It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
+    % Analyzing it in the generic case was found to be difficult, and is still an open problem.
+    which is implemented using a binary tournament, written in a naive but efficient way
+    (\ie, not using recursive functions, by using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
+    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffers from an exponential blow-up when $K$ increases.
 \end{itemize}

### 2019-05-28 2-Chapters/2-Chapter/chapter2.tex
Il me reste à parler bien de Exp3, et de UCB & klUCB
- reste okay !

@@
 
-\textbf{Implementation:}
+\paragraph{Implementation.}
 %
@@
 % ---------------------------------------------------
-\subsubsection{Naive strategies: pure exploitation or pure exploration}
+\paragraph{Naive strategies: pure exploitation or pure exploration}
 
@@
 % ---------------------------------------------------
-\subsubsection{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit}
+\paragraph{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit}
 
@@
 %
-This strategy is also referred to as ``explore-first'', and in the case without prior knowledge on the problem, it can achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ as shown in Section~1 of \cite{Slivkins2019}.
+This strategy is also referred to as ``explore-first'', it can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
 
@@
 
-\newpage
-
-\begin{itemize}
-    \item Algorithm 1 is
-    an $\varepsilon$-greedy algorithm, as described in \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
-    It has a very small cost in terms of time and memory, but it usually achieves linear regret.
-
-    \item Algorithm 2 is
-    an Explore-then-Commit algorithm from \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
-    Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but only obtain large (linear) regret.
-
-    \item Algorithm 3 is
-    $\mathrm{Exp}3^{++}$ from \cite{Seldin17}, using $\alpha=3$ and $\beta=256$ as advised (\texttt{Exp3PlusPlus}).
-    It is a recent variant of the $\mathrm{Exp3}$ algorithm, using an adaptive tuning of its parameter.
-It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
-
-    \item Algorithm 4 is
-    $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
-    It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
-
-    \item Algorithm 5 is
-    \klUCB{} from \cite{KLUCBJournal}, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern} (\texttt{klUCB}).
-    It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
-    It is computationally more costly that \UCB, as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
-    In practice, a few steps of a simple Newton method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than an order of magnitude slower than \UCB.
-    % of more than 10 digits
-    \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}
-
-\end{itemize}
-
-% ----------------------------------------------------------------------------
-\subsection{Index policies: UCB and others}
-
-\TODOL{Expliquer UCB}
-\TODOL{Expliquer que les indices UCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
-
-\TODOL{Expliquer klUCB}
-\TODOL{Expliquer que les indices klUCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
+% ----------------------------------------------------------------------------
+\subsection{Index policies, \UCB, \klUCB{} and others}
+
+A large family of algorithms are index policies, that compute an index $U_k(t)$ on each arm $k$ at time $t$.
+The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $C(t) = \argmax_k U_k(t)$.
+If more than one index maximizes $(U_k(t))_k$, the arm is chosen among the set, usually in a uniform random manner: $C(t) \sim \cU(\argmax_k U_k(t))$.
+%
+The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that include well known and efficient algorithms such as \UCB, \klUCB{} and many others.
 
@@
 		}
-		\caption{A generic index policy.}
+		\caption{A generic index policy, using indexes $U_k(t)$ (\eg, \UCB, \klUCB{} etc).}
 		\label{algo:2:indexPolicy}
@@
 
-\subsubsection{UCB}
+\paragraph{The \UCB{} index policy: using Hoeffding's inequality to build confidence intervals.}
+
+\TODOL{Expliquer UCB}
+\TODOL{Expliquer que les indices UCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
 Optimism in face of uncertainty, for UCB etc.
+$\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
+It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
 
@@
 
-\subsubsection{klUCB}
+The \UCB{} indexes are computed as follows:
+\begin{align}\label{eq:UCB_index}
+    \forall k,\;\;\;
+    U_k^{\UCB}(t) = \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\frac{X_k(t)}{N_k(t)}, q) \leq \frac{c \log(t)}{N_k(t)} \right\}.
+\end{align}
+
+\paragraph{The \klUCB{} index policy: using optimal confidence intervals.}
+
+\TODOL{Expliquer klUCB}
+\TODOL{Expliquer que les indices klUCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
+
 Optimal upper confidence bounds based on Kullback-Leibler divergence and klUCB indexes
 
+If rewards lie in $[a,b]$, then the indexes are computed as follows:
+\begin{align}\label{eq:klUCB_index}
+    \forall k,\;\;\;
+    U_k^{\klUCB}(t) = \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\frac{X_k(t)}{N_k(t)}, q) \leq \frac{c \log(t)}{N_k(t)} \right\}.
+\end{align}
+
 \TODOL{Short note on how to implement such policy, using Newton's method ?}
 
-compute the index $U_k(t) = XXX$.
-
-\subsubsection{Variants on UCB}
-
-- UCB-H and klUCB-H
-- UCB-V
-- UCB-dagger
-- UCBoost
-- UCB+ and klUCB+
-
-\paragraph{Other policies inspired by UCB}
+\paragraph{Variants on \UCB}
+
+Many variants have been proposed in the multi-armed bandits literature.
+For instance, \UCB-H (and \klUCB-H) replaces the $\log(t)$ by a $\log(T)$, and it corresponds to the first algorithm analyzed in \cite{Auer02}.
+Using an estimator for the variance leads to \UCB-V, efficient for instance for problems where the arms are Gaussian distributions with unknown variance.
+Other variants include \UCB$^+$, \UCB-dagger from \cite{Lattimore2018refining}, {\UCB}oost, and many more.
+%
+They are usually comparable to the original \UCB{} algorithm, in terms of time and memory complexity, and achieves improved regret upper bounds in the same setting or extension of the initial setting.
+
+\paragraph{Other policies inspired by \UCB}
 %
@@
 it starts with a prior on the $K$ distributions (\eg, a flat uniform prior), and then after every observation $(k, r(t))$ it updates the posterior distribution of the arm $k$ with a new data $r(t)$.
-If $P_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim P_k(t)$, and then playing according to the ``optimism in face of uncertainty'' philosophy if these samples represent the arms (expected) quality: $C(t) \argmax_k s_k$.
+If $\pi_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim \pi_k(t)$, and then playing according to the ``optimism in face of uncertainty'' philosophy if these samples represent the arms (expected) quality: $C(t) \in \argmax_k s_k$.
 
@@
 and maintain a distribution $\mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$ after $X_k(t)$ successes observed on arm $k$ and $N_k(t)-X_k(t)$.
+We detail the Thompson sampling algorithm in this case, below in Algorithm~\ref{algo:2:ThompsonSampling}, as it corresponds to the case used in Chapter~\ref{chapter:4} in Section~\ref{sub:41:numericalResults}.
 %
@@
 Thompson sampling was proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+
+% \begin{small} % XXX remove if needed
+\begin{figure}[h!]
+	\centering
+    \begin{framed}
+	\begin{algorithm}[H]
+        % \begin{small} % XXX remove if needed
+		\For(){$t = 1, \dots, T$}{
+            For each arm, sample $s_k(t) \sim \pi_k(t-1)$ from the posterior of arm $k$, $\pi_k(t-1)=\mathrm{Beta}(X_k(t-1)+1,N_k(t-1)-X_k(t-1)+1)$\;
+            Play arm $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
+            Observe a reward $r_{A(t)}(t)$\;
+            Update both $X_k(t)$ and $N_k(t)$\;
+		}
+		\caption{Thompson Sampling for Beta prior and posteriors, on Bernoulli rewards.}
+		\label{algo:2:ThompsonSampling}
+		% \end{small} % XXX remove if needed
+	\end{algorithm}
+	\end{framed}
+\end{figure}
+% \end{small} % XXX remove if needed
 
@@
 %  and it did not gain popularity since then.
-In practice, AdBandits is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
-
-Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that lies between the principle of optimism in the face of uncertainty and the Bayesian point-of-view.
+% In practice, AdBandits is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
+
+Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the principle of optimism in the face of uncertainty and the Bayesian point-of-view.
 %  underlying the approaches based on UCB, might not be necessary at all.
 % With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
-They showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as \UCB{} or Thompson sampling, with uniform perturbations.
+The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as \UCB{} or Thompson sampling, with uniform perturbations.
 Their RCB algorithm ressembles \UCB, and builds the same confidence intervals $\mathrm{LCB}_k(t), \UCB_k(t)$ on arm $k$ at time $t$.
@@
 
-BESA (Best Empirical Sampled Average) from \cite{Baransi2014}, and it consists in trying to fix a simple observation.
-For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
-The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
-It was analyzed partially for the two-armed case, and the authors illustrated that it can be very efficient empirically.
+BESA (Best Empirical Sampled Average) was introduced in \cite{Baransi2014}, in order to fix a simple observation:
+For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, their empirical mean).
+The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have respectively $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
+It was analyzed partially, for the two-armed case it was proved to be asymptotically optimal for one-dimensional exponential families, and the authors illustrated that it can be very efficient empirically.
 Analyzing it in the generic case was found to be difficult, and it is still an open problem.
@@
 
+FIXME
 Exp3, Exp3++ etc
@@
 The first example is MOSS from \cite{Audibert2009minimax}, or recently \KLUCBpp{} in \cite{Menard17}.
-They are both index-based, using XXX
-
-More recently, the \klUCB-switch from \cite{Garivier18} XXX
+They are both index-based, using the following indexes:
+\begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp}
+    \forall k,\;\;\;
+    U^{\mathrm{MOSS}}_k(t) &= \frac{X_k(t)}{N_k(t)} + \sqrt{\max\left(0, \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)}\right)}. \\
+    \forall k,\;\;\;
+    U^{\KLUCBpp}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\hat{\mu}_k(t), q) \leq \frac{c g(N_k(t), T, K)}{N_k(t)} \right\}.
+\end{align}
+%
+If we denote $g(t, T, K) = \log^+(y (1 + \log^+(y)^2))$, and $y = \frac{T}{K t}$, where $\log^+(x) = \max(0, \log(x))$.
+
+Both algorithms are index policies, that are not anytime.
+For any $K$, $T$, and problem instances $I$ in a certain family $\cI$ (\eg, with Bernoulli distributions),
+they enjoy regret bounds like the following, for two constants $c_{\cA}$ and $c'_{\cA}$ that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only:
+\begin{equation}\label{eq:2:regretBoundsForHybridPolicies}
+    R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T)).
+\end{equation}
+
+More recently, the \klUCB-switch from \cite{Garivier18} is the first algorithm to be proved to
+enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, that is, matching the $C_I \ln(T)$ lower bound by \cite{LaiRobbins85}.
+It is an index policy that uses a modified version of the indexes of the \klUCB$^+$ algorithm, for arms that are not sampled much, and a modified version of the indexes of the MOSS$^+$ algorithm, for the other arms.
+To detail this, first let $f(T, K) = \lfloor \rfloor$ for $\gamma=1/5$,
+then \klUCB-switch use the index defined as follows:
+%
+\begin{align}\label{eq:2:indexes_for_klUCBswitch}
+    \forall k,\;\;\;
+    U^{\mathrm{MOSS}+}_k(t) &= \frac{X_k(t)}{N_k(t)} + \sqrt{\max\left(0, \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}\right)}. \\
+    \forall k,\;\;\;
+    U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\hat{\mu}_k(t), q) \leq \frac{c \log(T / (K * N_k(t)))}{N_k(t)} \right\} \\
+    \forall k,\;\;\;
+    U^{\klUCB\text{-}\mathrm{switch}}_k(t) &= \begin{cases}
+        U^{KL+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
+        U^{MOSS+}_k(t) & \text{if } N_k(t) > f(T, K).
+    \end{cases}.
+\end{align}
 
@@
 
-\newpage
+% FIXME?
+% \newpage
 

### 2019-05-28 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Almost done for §2.4

@@
 %
-The considered aggregation algorithms are special cases of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details see
-the Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
-or this blog post by Jeremy Kun \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-of-the-}}\\
-\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{multiplicative-weights-update-algorithm}},
-which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019).},
-a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
-More details can be found in the survey \cite{Arora2012multiplicative}.
+The considered aggregation algorithms are special cases of the well-known \emph{multiplicative weights update algorithm}.
 

### 2019-05-28 2-Chapters/2-Chapter/chapter2.tex
Almost done for §2.4

@@
 BESA (Best Empirical Sampled Average) was introduced in \cite{Baransi2014}, in order to fix a simple observation:
-For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, their empirical mean).
+for instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, their empirical mean).
 The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have respectively $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
@@
 % ----------------------------------------------------------------------------
-\paragraph{Policies for adversarial bandits}
-
-FIXME
-Exp3, Exp3++ etc
-https://smpybandits.github.io/docs/Policies.Exp3.html
-https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
-
-It is a recent variant of the $\mathrm{Exp3}$ algorithm
-is $\mathrm{Exp}3^{++}$ from \cite{Seldin17},
-and it uses an adaptive tuning of its parameters.
+\paragraph{Policies for adversarial bandits.}
+
+Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
+The first of them is the Exp3 policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loiration'',
+which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details see
+the Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
+or this blog post by Jeremy Kun \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-of-the-}}\\
+\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{multiplicative-weights-update-algorithm}},
+which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019).
+More details can be found in the survey \cite{Arora2012multiplicative}.},
+a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
+% https://smpybandits.github.io/docs/Policies.Exp3.html
+%
+The Exp3 policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution:
+$\Pr(C(t)=k) = (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
+The weights are initially uniform, \ie, $w_k(0)=\frac{1}{K}$, and then at each time after observing a reward $r_k(t)$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
+The weights are then renormalized, the have a sum $\sum_k w_k(t+1) = 1$.
+This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r_k(t)$, that is, $\widehat{r_k}(t) = \Pr(C(t)=k)$.
+%
+Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a decreasing sequence $(\gamma_t)_t$.
+%
+It was proved in \cite{Auer02NonStochastic} that Exp3 with constant temperature parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
+while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
+For a short proof of this results as well as more details on Exp3, see \cite{Bubeck12}.
+
+% https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
+Finally, it is worth mentioning the Exp3$^{++}$ algorithm, as a recent variant of Exp3, proposed by \cite{Seldin17},
+that uses an adaptive tuning of its parameters.
 It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
-
-% Softmax: is it the same, but written differently?
-% https://smpybandits.github.io/docs/Policies.Softmax.html
 
@@
 For any $K$, $T$, and problem instances $I$ in a certain family $\cI$ (\eg, with Bernoulli distributions),
-they enjoy regret bounds like the following, for two constants $c_{\cA}$ and $c'_{\cA}$ that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only:
-\begin{equation}\label{eq:2:regretBoundsForHybridPolicies}
-    R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T)).
-\end{equation}
+they enjoy regret bounds like the following, for two constants $c_{\cA}$ and $c'_{\cA}$ that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only: $R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T))$.
 
@@
 It is an index policy that uses a modified version of the indexes of the \klUCB$^+$ algorithm, for arms that are not sampled much, and a modified version of the indexes of the MOSS$^+$ algorithm, for the other arms.
-To detail this, first let $f(T, K) = \lfloor \rfloor$ for $\gamma=1/5$,
+To detail this, first let $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,
 then \klUCB-switch use the index defined as follows:
@@
 \begin{align}\label{eq:2:indexes_for_klUCBswitch}
-    \forall k,\;\;\;
     U^{\mathrm{MOSS}+}_k(t) &= \frac{X_k(t)}{N_k(t)} + \sqrt{\max\left(0, \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}\right)}. \\
-    \forall k,\;\;\;
     U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\hat{\mu}_k(t), q) \leq \frac{c \log(T / (K * N_k(t)))}{N_k(t)} \right\} \\
-    \forall k,\;\;\;
     U^{\klUCB\text{-}\mathrm{switch}}_k(t) &= \begin{cases}
@@
 
-% FIXME?
+FIXME?
+\TODOL{Je dois conclure un peu cette section...}
+
 % \newpage

### 2019-05-29 2-Chapters/2-Chapter/chapter2.tex
I need to shower

@@
 
-In this Section, we start by discussing two naive strategies which fail dramatically.
-%
-We then expose two simple strategies, which performs efficiently only if they are tuned using a prior knowledge of the problem difficulty, but are thus unusable for practical applications.
-%
-We continue by presenting a first efficient policy, \UCB, that uses upper confidence bounds (UCB) on the mean estimates, and is known to be order-optimal for bounded rewards.
-Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of \UCB{} but use smaller confidence intervals and thus typically has better theoretical results, and Thompson sampling which takes a Bayesian point-of-view instead of relying on the principle of optimism in the face of uncertainty.
-%
-Finally, we conclude this Section by briefly exposing other families of algorithms.
+This Section starts by discussing two naive strategies which fail dramatically,
+%
+and then we present two simple strategies which performs efficiently, only if they are tuned using a prior knowledge of the problem difficulty, but are thus unusable for practical applications.
+%
+We focus then on index policies, with a first efficient policy, \UCB, that uses upper confidence bounds (UCB) on the means estimates, and is known to be order-optimal for bounded rewards.
+Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of \UCB{} but use smaller confidence intervals and thus it typically obtains better theoretical results, and Thompson sampling which replaces the principle of ``optimism under uncertainty'' by a Bayesian point-of-view.
+%
+Finally, we conclude by briefly exposing other families of algorithms.
 
@@
 % ---------------------------------------------------
+\subsection{Naive or simple strategies}
+\label{sub:2:naiveSimpleStrategies}
+
 \paragraph{Naive strategies: pure exploitation or pure exploration}
@@
 %
-This strategy is also referred to as ``explore-first'', it can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
+This strategy can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
 
@@
 
-\TODOL{Expliquer UCB}
-\TODOL{Expliquer que les indices UCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
-Optimism in face of uncertainty, for UCB etc.
-$\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
-It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
-
-compute the index $U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$.
-
-The \UCB{} indexes are computed as follows:
+Let us consider another approach for adaptive exploration, known as ``optimism under uncertainty'': assume each arm is as good as it can possibly be given the observations so far, and choose the best arm based on these optimistic estimate.
+%
+This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}.
+For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
+the \emph{average reward} $\widehat{\mu_k}(t)=\frac{X_k(t)}{N_k(t)}$
+and a \emph{confidence radius} $\xi_k(t) = \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}\right)}$.
+%
 \begin{align}\label{eq:UCB_index}
-    \forall k,\;\;\;
-    U_k^{\UCB}(t) = \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\frac{X_k(t)}{N_k(t)}, q) \leq \frac{c \log(t)}{N_k(t)} \right\}.
+    U_k^{\UCB}(t) = \underbrace{\frac{X_k(t)}{N_k(t)}}_{\widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}\right)}}_{\xi_k(t)}.
 \end{align}
+
+This selection rule makes sense for the following reason:
+an arm $k$ is chosen at step $t$ because it has a large $U_k^{\UCB}(t) = \UCB_k(t)$, which can happen for two reasons:
+$i)$ because the average reward $\widehat{\mu_k}(t)$ is large, in which case this arm is likely to have a high reward,
+$ii)$ and/or because the confidence radius $\xi_k(t)$ is large, in which case this arm has not been explored much.
+Either reason makes this arm worth choosing.
+%
+Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{}  respectively represent exploitation and exploration, and summing them up is a natural way to trade off the two.
+The parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to chose $\alpha=1/2$ for optimal uniform performance.
+
+To analyze the algorithm $\UCB_1$, we denote again $k^*$ an optimal arm, and $C(t)$ is the arm chosen by the algorithm in round $t$.
+According to the algorithm, $\UCB_k(t) \geq \UCB_{k^*}(t)$.
+%
+Under the clean event,μ(at) +rt(at)≥ ̄μt(at)andUCBt(a∗)≥μ(a∗). Therefore:
+XXX
+It follows that
+XXX
+
+More details on the proofs can be found for instance in Chapter~2 of \cite{Bubeck12}, or Section~1.3 of \cite{Slivkins2019}.
+% \TODOL{Expliquer la fin}
+The $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+
+It has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects.
+The $\UCB_1$ algorithm requires a constant storage capacity for each arm (storing $X_k(t)$ and $N_k(t)$ require two integers or a \texttt{float} and an integer, \ie, $\bigO{1}$), giving a storage capacity of $\bigO{K}$, independently of the horizon $T$.
+It also needs to perform some mathematical operations, but a square-root and a logarithm are cheap, thus it costs a $\bigO{1}$ constant time for each arm and time step, thus a time $\bigO{K T}$ for the $T$ rounds.
+%
+We use it as the base building block for other policies in Chapter~\ref{chapter:4}, for these reasons.
+Moreover, the $\UCB_1$ algorithm has the advantage of being easy to explain and understand, and its decisions can be interpreted clearly: by reading the values of $\widehat{\mu_k}(t)$ and $\UCB_k(t)$, anyone can see why an arm was chosen at anytime.
+This is why it is used in our demonstration presented in Section~\ref{sec:4:gnuradio}.
 
@@
 \begin{align}\label{eq:klUCB_index}
-    \forall k,\;\;\;
-    U_k^{\klUCB}(t) = \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\frac{X_k(t)}{N_k(t)}, q) \leq \frac{c \log(t)}{N_k(t)} \right\}.
+    U_k^{\klUCB}(t) = \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c \log(t)}{N_k(t)} \right\}.
 \end{align}
@@
 
-Many variants have been proposed in the multi-armed bandits literature.
-For instance, \UCB-H (and \klUCB-H) replaces the $\log(t)$ by a $\log(T)$, and it corresponds to the first algorithm analyzed in \cite{Auer02}.
-Using an estimator for the variance leads to \UCB-V, efficient for instance for problems where the arms are Gaussian distributions with unknown variance.
-Other variants include \UCB$^+$, \UCB-dagger from \cite{Lattimore2018refining}, {\UCB}oost, and many more.
-%
-They are usually comparable to the original \UCB{} algorithm, in terms of time and memory complexity, and achieves improved regret upper bounds in the same setting or extension of the initial setting.
+Many variants of the \UCB{} algorithms have been studied in the multi-armed bandits literature.
+For instance, \UCB-H replaces the $\log(t)$ by a $\log(T)$ (as well as \klUCB-H), and it actually corresponds to the first algorithm analyzed in \cite{Auer02}.
+Using an estimator for the variance of the arms' samples gives \UCB-V, which was for instance proven to be efficient against Gaussian distributions with unknown variance.
+Other variants include \UCB$^+$, $\UCB^{\dagger}$ from \cite{Lattimore2018refining}, {\UCB}oost, and many more.
+%
+All these variants are usually comparable to the original \UCB{} algorithm, in terms of time and memory complexity, and they achieve improved regret upper bounds in the same setting or some extensions of the initial setting.
 
@@
 %
-The ``Successive Elimination'' policy also uses confidence bounds, and use both UCB and Lower Confidence Bounds (LCB).
-After playing each active arms once, it eliminates arm $a$ if it is statistically safe that it is sub-optimal, \ie, if there is another arm $a' \neq a$ such that $\UCB_a(t) < \mathrm{LCB}_{a'}(t)$.
+The ``Successive Elimination'' policy also uses confidence bounds, and use both Upper and Lower Confidence Bounds (LCB).
+After playing each active arms once, it eliminates arm $a$ if it is statistically identified as sub-optimal, \ie, if there is another arm $a' \neq a$ such that $\UCB_a(t) < \mathrm{LCB}_{a'}(t)$.
 It also achieves a $\bigO{\log(T)}$ problem-dependent regret upper bound, and a $\bigO{\sqrt{K T \log(T)}}$ problem-independent bound.
@@
 it starts with a prior on the $K$ distributions (\eg, a flat uniform prior), and then after every observation $(k, r(t))$ it updates the posterior distribution of the arm $k$ with a new data $r(t)$.
-If $\pi_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim \pi_k(t)$, and then playing according to the ``optimism in face of uncertainty'' philosophy if these samples represent the arms (expected) quality: $C(t) \in \argmax_k s_k$.
+If $\pi_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim \pi_k(t)$, and then playing according to the ``optimism under uncertainty'' philosophy if these samples represent the arms (expected) quality: $C(t) \in \argmax_k s_k$.
 
@@
 the best choice of prior is a Beta distribution, that starts as an initially uniform prior (\ie, $\mathrm{Beta}(1,1)$),
-and maintain a distribution $\mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$ after $X_k(t)$ successes observed on arm $k$ and $N_k(t)-X_k(t)$.
+and maintain a distribution $\mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$,
+as $X_k(t)$ counts the number of successes observed on arm $k$, and $N_k(t)-X_k(t)$ counts the number of failures.
+%
 We detail the Thompson sampling algorithm in this case, below in Algorithm~\ref{algo:2:ThompsonSampling}, as it corresponds to the case used in Chapter~\ref{chapter:4} in Section~\ref{sub:41:numericalResults}.
@@
 In this setting as well as for Gaussian distributions,
-Thompson sampling was proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+Thompson sampling was proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
 
@@
 It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given.
+We found that in practice it can achieve very good performance, as illustrated in Section~\ref{sec:3:reviewSPAlgorithms}.
 %  and it did not gain popularity since then.
@@
 
-Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the principle of optimism in the face of uncertainty and the Bayesian point-of-view.
+Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the principle of optimism under uncertainty and the Bayesian point-of-view.
 %  underlying the approaches based on UCB, might not be necessary at all.
@@
 
-We briefly mention other families of policies.
+We can briefly mention other families of policies.
 
@@
 which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details see
-the Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
+on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
 or this blog post by Jeremy Kun \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-of-the-}}\\
@@
 For a short proof of this results as well as more details on Exp3, see \cite{Bubeck12}.
-
+%
 % https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
@@
 \begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp}
-    \forall k,\;\;\;
     U^{\mathrm{MOSS}}_k(t) &= \frac{X_k(t)}{N_k(t)} + \sqrt{\max\left(0, \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)}\right)}. \\
-    \forall k,\;\;\;
-    U^{\KLUCBpp}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\hat{\mu}_k(t), q) \leq \frac{c g(N_k(t), T, K)}{N_k(t)} \right\}.
+    U^{\KLUCBpp}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
 \end{align}
@@
     U^{\mathrm{MOSS}+}_k(t) &= \frac{X_k(t)}{N_k(t)} + \sqrt{\max\left(0, \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}\right)}. \\
-    U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \mathrm{kl}(\hat{\mu}_k(t), q) \leq \frac{c \log(T / (K * N_k(t)))}{N_k(t)} \right\} \\
+    U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(T / (K N_k(t)))}{N_k(t)} \right\} \\
     U^{\klUCB\text{-}\mathrm{switch}}_k(t) &= \begin{cases}
-        U^{KL+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
-        U^{MOSS+}_k(t) & \text{if } N_k(t) > f(T, K).
+        U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
+        U^{\mathrm{MOSS}+}_k(t) & \text{if } N_k(t) > f(T, K).
     \end{cases}.

### 2019-05-29 2-Chapters/2-Chapter/chapter2.tex
Oupsie notations, C(t) is NOT used here, use A(t)

@@
 A large family of algorithms are index policies, that compute an index $U_k(t)$ on each arm $k$ at time $t$.
-The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $C(t) = \argmax_k U_k(t)$.
-If more than one index maximizes $(U_k(t))_k$, the arm is chosen among the set, usually in a uniform random manner: $C(t) \sim \cU(\argmax_k U_k(t))$.
+The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
+If more than one index maximizes $(U_k(t))_k$, the arm is chosen among the set, usually in a uniform random manner: $A(t) \sim \cU(\argmax_k U_k(t))$.
 %
@@
 
-To analyze the algorithm $\UCB_1$, we denote again $k^*$ an optimal arm, and $C(t)$ is the arm chosen by the algorithm in round $t$.
+To analyze the algorithm $\UCB_1$, we denote again $k^*$ an optimal arm, and $A(t)$ is the arm chosen by the algorithm in round $t$.
 According to the algorithm, $\UCB_k(t) \geq \UCB_{k^*}(t)$.
@@
 it starts with a prior on the $K$ distributions (\eg, a flat uniform prior), and then after every observation $(k, r(t))$ it updates the posterior distribution of the arm $k$ with a new data $r(t)$.
-If $\pi_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim \pi_k(t)$, and then playing according to the ``optimism under uncertainty'' philosophy if these samples represent the arms (expected) quality: $C(t) \in \argmax_k s_k$.
+If $\pi_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim \pi_k(t)$, and then playing according to the ``optimism under uncertainty'' philosophy if these samples represent the arms (expected) quality: $A(t) \in \argmax_k s_k$.
 
@@
 The Exp3 policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution:
-$\Pr(C(t)=k) = (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
+$\Pr(A(t)=k) = (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
 The weights are initially uniform, \ie, $w_k(0)=\frac{1}{K}$, and then at each time after observing a reward $r_k(t)$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
 The weights are then renormalized, the have a sum $\sum_k w_k(t+1) = 1$.
-This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r_k(t)$, that is, $\widehat{r_k}(t) = \Pr(C(t)=k)$.
+This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r_k(t)$, that is, $\widehat{r_k}(t) = \Pr(A(t)=k)$.
 %
@@
 while using a low $\alpha$, \ie, favoring exploitation, will give low regret in the second case.
-One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of \UCB, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best for the problem at hand.
+One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of \UCB, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
 

### 2019-05-29 2-Chapters/2-Chapter/chapter2.tex
Typos

@@
 \label{chapter:2}
+
+\TODOL{Abstract ici et plus court, et petite introduction !}
+
 \minitoc
@@
 \newpage
+
 % \TODOL{L'abstract ici est trop long, mais je l'ai écrit avant de commencer le chapitre, pour avoir une idée de ce que je devais faire ! Je couperai à la fin !}
@@
     \begin{equation}
-        R_T^{\cA}(I) = \bigO{C_{I,\alpha} T^{\alpha}}
+        R_T^{\cA}(I) = \smallO{C_{I,\alpha} T^{\alpha}}
     \end{equation}
     for any value $\alpha>0$ and any problem instance $I\in \cI$,
-    where this notation means that the ``constant'' $C_{I,\alpha}$ can depend on the problem instance $I$ and on $\alpha$, but not on the time horizon $T$.
+    where this notation means that the ``constant'' $C_{I,\alpha}$ can depend on the problem instance $I$ and on $\alpha$, but \emph{not} on the time horizon $T$.
 \end{defn}
@@
 Since a few years, some algorithms were shown to achieve both a problem-dependent logarithmic and a minimax upper-bounds,
-like MOSS in \cite{Audibert2009minimax} or recently \KLUCBpp{} in \cite{Menard17}.
+like MOSS in \cite{Audibert2009minimax} or recently \KLUCBpp{} in \cite{Menard17},
+and such results are usually referred to as ``best of both worlds''.
 
@@
 %
-Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$ using Hoeffding's inequality again.
+Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus Hoeffding's inequality gives $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$.
 Optimizing on $T_0$ gives $T_0 = 4/d^2 \log(d^2 T / 4)$
@@
 
-For both cases, the strategy obtains a linear regret if it is tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if lower-bound on the gaps $\Delta_k$ is known.
+For both cases, the strategy obtains sub-optimal regret if it is tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if a lower-bound on the gaps $\Delta_k$ is known.
 %  \ie, they require a prior knowledge on the problem difficulty.
-As such, they are not applicable on an unknown problem and are less interesting from a practical point-of-view.
+As such, this weakness make them unapplicable on an unknown problem, and so these first strategies are less interesting from a practical point-of-view.
 

### 2019-05-29 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Typos

@@
 Following a Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
-it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in[K]$.
-Then, the dynamic device waits in this channel $C(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example considered above).
+it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $A(t) = k \in[K]$.
+Then, the dynamic device waits in this channel $A(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example considered above).
 Before sending another message (\ie, at time $t+1$), the dynamic device knows if it received or not this \Ack{} message.
@@
 The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
-$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
+$r_{1 \dots T} := \sum_{t = 1}^T r_{A(t)}(t),$
 which is a usual objective in MAB problems.

### 2019-05-30 2-Chapters/1-Chapter/chapter1.tex
Typos
Added https://hal.inria.fr/hal-02144465/document

@@
 \label{chapter:1}
-\minitoc
 
@@
 The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but use the notations introduced on Chapter~\ref{chapter:2}.
+
+\minitoc
 
@@
     1st International ISIoT workshop\footnote{~See \href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}},
-    at \emph{Conference on Distributed Computing in Sensor Systems}\footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}}, Santorini, Greece, May 2019.
+    at \emph{Conference on Distributed Computing in Sensor Systems}\footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}}, Santorini, Greece, May 2019.\\
+    \href{https://HAL.Inria.fr/hal-02144465}{\texttt{HAL.Inria.fr/hal-02144465}}.
     \cite{MoyBesson2019}

### 2019-05-30 4-Listofpublications/Listofpublications.tex
Added https://hal.inria.fr/hal-02144465/document

@@
 \item
-  \href{https://hal.inria.fr/hal-XXX/document}{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation}, by \href{https://moychris.wordpress.com/}{C. Moy} \& \textbf{L. Besson},
-  \href{https://sites.google.com/view/ISIoT2019/}{ISIoT Workshop} at \href{https://dcoss.org/}{DCOSS}, Santorini, Greece, May 2019.
-  % \href{https://hal.inria.fr/hal-XXX}{HAL-XXX}.
+  \href{https://hal.inria.fr/hal-02144465/document}{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation}, by \href{https://moychris.wordpress.com/}{C.
+Moy} \& \textbf{L.
+Besson},
+  \href{https://sites.google.com/view/ISIoT2019/}{ISIoT Workshop} at \href{https://dcoss.org/}{DCOSS}, Santorini, Greece, May 2019,
+  \href{https://hal.inria.fr/hal-02144465}{HAL-02144465}.
   \cite{MoyBesson2019}

### 2019-06-03 2-Chapters/2-Chapter/chapter2.tex
Proof reading first parts of Chapter2

@@
 We then give regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
-They both attain a logarithmic regret, proving their order optimality (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
+They both attain a logarithmic regret, proving their order optimality.
+% (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
 % Moreover, \klUCB{} is known to be optimal for a wide range of arms distributions.
@@
 
+\paragraph{About notations.}
+%
 The notations used in this thesis are the usual notations in all the MAB literature, and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
@@
 The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T Y_t] = \sum_{t=1}^T \E[r_{A(t)}(t)]$.
-The expectancy is taken on the rewards $(r_k(t))_{k,t}$, and the extern random variables $(U_s)_s$.
-%
-Other common objectives include best-arm identification (BAI) \cite{audibert2010best}.
+The expectancy is taken on the rewards $(r_k(t))_{k,t}$, and the extern random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
+%
+Other commonly studied objectives include best-arm identification (BAI) \cite{audibert2010best}.
 
@@
 %
-If you are discovering the concept of bandit model here, I would like to recommend you to go online and play a few times with an interactive demonstration.
+If you are discovering here the concept of bandits, I would like to recommend you to go online and play a few times with an interactive demonstration.
 On this demo, you will be facing a MAB problem with $K=5$ arms, and you have $T=100$ decisions to make.
-The demo is hosted on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}, and illustrated below.
+The demo is hosted on my website, at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}, and it is illustrated below.
 
@@
 
-A bandit algorithm $\cA$, also referred to as a strategy or a policy, and sometimes denoted by $\pi$ or $\rho$, selects an arm $A(t)$ at time $t$ by using the past observations and the past external randomness.
-More formally, one could state that an algorithm is a sequence of measurable functions $(\cA_t)_{t\geq1}$,
-where $\cA_t$ maps the past observations $\cO_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$
-to an arm $\cA_t(\cO_t) = A(t) \in[K]$,
+A bandit algorithm $\cA$ is also referred to as a strategy or a policy, and sometimes it is denoted by $\pi$ or $\rho$ in the literature.
+The algorithm $\cA$ selects an arm $A(t)$ at time $t$ by using the past observations and the past external randomness.
+More formally, one could state that an algorithm is a sequence of \emph{measurable} functions $(\cA_t)_{t\geq1}$,
+where $\cA_t$ maps the past observations $O_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$
+to an arm $\cA_t(O_t) = A(t) \in[K]$,
 if we denote $Y_s = r_{A(s)}(s)$ the $s$-th reward.
-The initial information is reduced to $\cO_1 = (U_0)$, and the first decision is $A(1) = \cA_1(\cO_1)$.
+The initial information is reduced to $O_1 = (U_0)$, and the first decision is $A(1) = \cA_1(O_1)$.
 Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps.
@@
 % https://smpybandits.github.io/docs/Arms.html
-We also focus only on real-valued rewards, meaning that $r_k(t)\in\R$ for all $k$ and $t$.
+We also focus only on real-valued rewards, meaning that $r_k(t)\in\R$ for all arm $k$ and time $t$.
 
@@
 and whether the player knows if they are bounded or not before starting the bandit game.
-Moreover, if rewards are known to be bounded, let say in an interval $[a,b]$, another important hypothesis is whether the player knows the values of $a$ and $b$.
-%
-Intuitively, the bandit game is easier if the player knows the support of the distributions, and we restrict to this case in all the thesis.
+Moreover, if rewards are known to be bounded, let say in an interval $[a,b]$, another important hypothesis is whether the player knows the values of $a$ and $b$ or not.
+%
+Intuitively, the bandit game is easier if the player knows the support of the distributions, and we restrict to this case in all the thesis, \ie, \emph{$a,b$ are always supposed to be known}.
 Most of the algorithms proposed in the literature follow this hypothesis as well.
@@
 while in the literature, the Bernoulli distribution is the most common case of finitely supported distributions.
-Continuous distributions with finite support also included truncated versions of infinitely supported distributions, in particular truncated Gaussian are used for numerical experiments in lots of research articles.
+Continuous-valued distributions with finite support also included truncated versions of infinitely supported distributions, in particular truncated Gaussian are used for numerical experiments in lots of research articles.
 
@@
 
--- An important result on exponential families is that \emph{a distribution in such family is entirely characterized by its parameter $\eta$}.
-That is why such one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[X]$.
-
--- A second important result is a simplified form for the \emph{Kullback-Leibler divergence} \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
+We have to quote two important results on exponential families:
+
+\begin{enumerate}
+    \item
+    \emph{A distribution in such family is entirely characterized by its parameter $\eta$}.
+    That is why one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[X]$.
+
+    \item
+    A second important result is a simplified form for the \emph{Kullback-Leibler divergence} \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
     The KL divergence is also called the relative entropy is a measure of how one probability distribution is different from a second, reference probability distribution.
+\end{enumerate}
 
@@
     Moreover, the KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
-    $\kl(x,y) = \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$.
+    $\kl(x,y) = \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
+    and it is also called the \emph{relative binary entropy}.
 
@@
     Moreover, the KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
-    $d(x,y) = \KL_{\cN}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$.
+    $d(x,y) = \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$.
 \end{itemize}
 
-In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the application of choice.
+In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the applications of choice.
 First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
-% Then in Chapter~\ref{chapter:4}, each static IoT device in the noisy environment is assumed to be following a Bernoulli emission pattern in a fixed channel, FIXME
-%
-For example when we study multi-player bandits in Chapter~\ref{chapter:5}, for the model without sensing, we explain that restricting to the Bernoulli case is interesting and it is actually the hardest one, as continuous distributions with a null mass on $r(t)=0$ yield a much simpler problem.
-%
-Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm in the setting on bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use the Bernoulli Kullback-Leibler divergence in our analysis.
+In Chapter~\ref{chapter:4}, we only study models with binary rewards.
+%
+Then, when we study multi-player bandits in Chapter~\ref{chapter:5}, for the model without sensing, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $r(t)=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
+%
+Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm in the setting on bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use the Bernoulli Kullback-Leibler divergence $\kl$ in our analysis.
 
@@
 %
-Such hypothesis is also proposed for other families of distributions, for instance \cite{KimTewari2019} analyses their Follow-the-Perturbed-Leader algorithm for perturbations following any sub-Weibull distribution (sub-Weibull distributions generalize both sub-Gaussian and sub-Exponential distributions).
+Such hypothesis is also proposed for other families of distributions, for instance the recent article \cite{KimTewari2019} analyses their Follow-the-Perturbed-Leader algorithm for perturbations following any sub-Weibull distribution (sub-Weibull distributions generalize both sub-Gaussian and sub-Exponential distributions).
 
@@
 %
-Such models were less studied than stationary or adversarial models, but some interesting works focussed on Markov models in the last 10 years.
-For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
+Such models were less studied than stationary or adversarial models, but some interesting works focussed on Markov models in the last 10 years\footnote{~
+For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.}.
 
@@
     The theoretical complexity of this use of bandit for A/B testing was first studied in \cite{Kaufmann14},
-    and later on a more practical point-of-view was proposed in \cite{Jamieson17ABTest}.
+    and a more practical point-of-view was proposed in \cite{Jamieson17ABTest}.
 
@@
     As presented below in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, a certain ``leader'' bandit algorithm can be used to select on the run the best bandit algorithm from a pool of ``followers'' algorithms.
-    Other possible use cases include hyper-parameter optimization, or feature selection.
+    Other possible use cases include hyper-parameter optimization, or features selection.
     Hyper-parameters include real-valued parameters, like $\alpha\in[0,\infty)$ for the \UCB{} or $\varepsilon$ for the $\varepsilon$-greedy bandit algorithms, a step size multiplier in a gradient method, or the width of a radial based function (RBF) kernel method.
     Discrete-valued parameters are also common, like a choice in a fixed set of kernel functions or the depth of neural networks,
-    and higher dimensional or more complex hyper-parameters include the entire architecture of a neural network.
+    and higher dimensional or more complex hyper-parameters can for instance be the entire architecture of a neural network.
 
@@
 As explained above, the main objective of a player facing a bandit game is to maximize its (expected) cumulated reward.
-An efficient algorithm should obtain a mean reward converging to the maximum reward.
+An efficient algorithm should see its mean reward converge to the maximum reward.
 We introduce in this section the notion of \emph{regret}, and the equivalence for a player between maximizing its sum of rewards and minimizing its regret.
@@
 Let us first introduce some notations.
-From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $(\nu_k)_k$, that is, for any time $t$, $r_k(t)\sim \nu_k$.
-We denote $\mu_k$ the mean of the distribution of arm $k$ (it will be referred to as the mean of arm $k$ in the rest of the manuscript), that is $\mu_k = \E[r_k(t)] \in \R$.
+From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that is, for any time $t$, $r_k(t)\sim \nu_k$.
+We denote $\mu_k$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$} in the rest of the manuscript), that is $\mu_k = \E[r_k(t)] \in \R$.
 %
@@
 
-\begin{defn}[Optimal and sub-optimal arms]\label{def:2:optimalSubOptimalArms}
+\begin{defn}\label{def:2:optimalSubOptimalArms}
+    % [Optimal and sub-optimal arms]
     Consider a bandit problem of $K$ arms with distributions characterized by their means $\bm{\mu}=\mu_1,\dots,\mu_k$.
-    Denote $\mu^* = \max_k \mu_k$ the largest mean.\\
+    Denote $\mu^* = \max_k \mu_k$ the largest mean.
     The best arm can be non unique, and any arm $k$ having $\mu_k = \mu^*$ is said to be \emph{optimal},

### 2019-06-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Proof reading first parts of Chapter2

@@
 The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
-Fix a player $j\in[M]$, and we consider the observations $\cO_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
-\[ \cO_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
+Fix a player $j\in[M]$, and we consider the observations $O_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
+\[ O_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
 where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).
@@
 Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
-We denote $\Pr_{\bm{\mu}}^{\cO_t}$ the distribution of the vector $\cO_t$ under the model $\bm{\mu}$ when using algorithm $\rho$.
+We denote $\Pr_{\bm{\mu}}^{O_t}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\rho$.
 Then we introduce the \emph{collision information term} as:
 \begin{equation}\label{eq:5:collisionInformationTerm}
-  \cI_{\bm{\mu},\bm{\lambda}}(\rho,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | \cO_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | \cO_{t-1}}).
+  \cI_{\bm{\mu},\bm{\lambda}}(\rho,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}}).
 \end{equation}

### 2019-06-03 2-Chapters/2-Chapter/chapter2.tex
Added a reference of a paper studying fairness for single-player stochastic problems
- https://arxiv.org/pdf/1905.11260.pdf

@@
 From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that is, for any time $t$, $r_k(t)\sim \nu_k$.
+We focus on distributions fully characterized by their means, that can be Bernoulli or any one-dimensional exponential family.
 We denote $\mu_k$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$} in the rest of the manuscript), that is $\mu_k = \E[r_k(t)] \in \R$.
@@
     % [Optimal and sub-optimal arms]
-    Consider a bandit problem of $K$ arms with distributions characterized by their means $\bm{\mu}=\mu_1,\dots,\mu_k$.
+    Consider a bandit problem of $K$ arms with distributions of means $\bm{\mu}=\mu_1,\dots,\mu_k$.
     Denote $\mu^* = \max_k \mu_k$ the largest mean.
@@
 $= \E\left[ \sum_{t=1}^T r_{k^*}(t) \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right]$
-$= T \E\left[ r_{k^*}(1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]$
+$= T \times \E\left[ r_{k^*}(t=1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]$
 $= T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]$.
@@
 \begin{defn}[Regret]\label{def:2:regret}
-    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the (mean) regret at horizon $T$ is defined as
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the (mean) regret at horizon $T$ is $R_T^{\cA}$ and it is defined as
+    %
     \begin{equation}
@@
 The \emph{gap} between any arm $k\in[K]$ and an optimal arm is defined as $\Delta_k = \mu^* - \mu_k$
-(an arm $k$ is sub-optimal if $\Delta_k > 0$),
-and thus we can write the following decomposition on the regret,
+(an arm $k$ is thus sub-optimal if and only if $\Delta_k > 0$),
+and thus we can write the following decomposition on the regret.
 
@@
     \begin{equation}\label{eq:2:RegretDecomposition}
-        R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ] = \sum_{k=1,\dots,K : \Delta_k > 0} \Delta_k \E[ N_k(T) ].
+        R_T^{\cA} = \sum_{k=1}^K \Delta_k \; \E[ N_k(T) ]
+        = \sum_{\substack{k=1,\dots,K\\\Delta_k > 0}} \Delta_k \; \E[ N_k(T) ].
     \end{equation}
@@
     Thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]$.
-    The sum can then be simplified to only count sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
+    The sum can then be simplified to only account for sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
 \end{proof}
@@
 
--- On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
+\begin{itemize}
+    \item
+On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
 Mathematically, both quantities are equal in expectation, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition}.
@@
 
--- On the other hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
+    \item
+On the other hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
 For example, we extend this decomposition to the multi-player case in Chapter~\ref{chapter:3}, in Lemma~\ref{lem:5:DecompositionRegret}, and as it also uses the draws $N_k(T)$, it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
+\end{itemize}
 
@@
 
-We include in this section two well-known results about what bandit algorithms \emph{cannot} do.
+We include here two well-known results about what bandit algorithms \emph{cannot} do.
 First, a \emph{problem-dependent lower-bound} states that any algorithm suffers a regret at least $\Omega(\log T)$ on any problem (from \cite{LaiRobbins85}).
@@
 %
-In a certain family of problems, \eg, $K$ Bernoulli distributed arms, the difficulty of a problem is characterized by a measure that depends only on the arms means.
-This measure of difficulty of a problem is hidden in the $\Omega$ notation of the lower-bound.
+In certain families of problems, \eg, $K$ Bernoulli distributed arms, the difficulty of a problem is characterized by a constant that depends only on the arms means.
+This measure of difficulty of a problem is hidden in the $\Omega$ notation in these lower-bounds.
 
@@
 
-\paragraph{Problem-dependent lower-bound in $\log T$}
-
+\paragraph{Problem-dependent lower-bound in $\Omega(\log T)$.}
+%
 The following two theorems were proven in \cite{LaiRobbins85}.
@@
 
-Moreover, this lower-bound can be directly used to design efficient algorithms, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression in \eqref{eq:2:forSecondLogTLowerBound2} essentially says that any efficient algorithm
-should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times.
-Tracking this quantity is used for instance in the OSSB algorithm proposed in \cite{Combes17}.
-% FIXME explain why this is interesting for the rest of this manuscript!
+Moreover, it is interesting to note that this lower-bound can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proved to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression in \eqref{eq:2:forSecondLogTLowerBound2} essentially says that any efficient algorithm
+should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times in the total of $T$ time steps.
 
@@
 
-We consider uniformly efficient\footnote{This notion is then extended for ``strongly uniformly efficiency'' in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}, where we also include a notion of (expected) fairness.} algorithms, to rule out algorithms achieving low regret on some problem instances while achieving linear regret on other instances.
+We consider uniformly efficient\footnote{This notion is then extended to ``\emph{strongly uniformly efficient} algorithms'', in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}, where we also include a notion of (expected) fairness.} algorithms, to rule out algorithms achieving low regret on some problem instances while achieving linear regret on other instances.
 In particular, it is necessary to rule out algorithms that always pick the same arm, as on some problem instances such fixed-arm algorithms can achieve zero regret.
-This class contains for instance the \UCB{} algorithm, as its regret is proven to be logarithmic on any problem instance.
 
@@
 
+This family is non-empty as it contains for instance the \UCB{} algorithm, as its regret is proven to be logarithmic on any problem instance.
 Now we can state the second logarithmic lower-bound, for algorithms in this family:
@@
         \begin{equation}\label{eq:2:forSecondLogTLowerBound}
-            C_I = \mu^* (1 - \mu^*) \sum_{k: \Delta_k > 0} \frac{1}{\Delta_k}.
+            C_I = \mu^* (1 - \mu^*) \sum_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{1}{\Delta_k}.
         \end{equation}
         \item
-        For any $\varepsilon>0$, the bound from Theorem~\ref{thm:2:secondLogTLowerBound} holds with,
+        For any $\varepsilon>0$, the bound from Theorem~\ref{thm:2:secondLogTLowerBound} also holds with,
         \begin{equation}\label{eq:2:forSecondLogTLowerBound2}
-            C_I = \sum_{k: \Delta_k > 0} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon.
+            C_I = \sum_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon.
         \end{equation}
@@
 
-\paragraph{Worst-case lower-bound in $\sqrt{T}$}
-
+\paragraph{Worst-case lower-bound in $\Omega(\sqrt{T})$.}
+.
 For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses a constant $C_I$ no longer bring any information.
@@
 
-For this reason, the research literature has also looked with interest another family of bounds on regret, that are not problem-dependent but worst-case, also called minimax \cite{Audibert2009minimax,audibert2010minimax}.
+For this reason, the research literature has also been interested by another family of regret bounds, that are not problem-dependent but worst-case, or also called minimax \cite{Audibert2009minimax,audibert2010minimax} or problem-independent.
 Hence it is interesting to quote another lower-bound on the regret, that can bring useful information in such settings.
-The following theorem was proven in \cite{Auer02NonStochastic}.
+The following theorem is from \cite{Auer02NonStochastic}.
 
@@
 Finally, another interesting measure of performance is the \emph{fairness} between the $M$ players in a multi-player bandit models, as we explain in Chapter~\ref{chapter:5}.
+It was not studied much, and only a very recent article studies this measure of performance under a formal approach, \cite{Patil2019stochastic}.
 

### 2019-06-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Added a reference of a paper studying fairness for single-player stochastic problems
- https://arxiv.org/pdf/1905.11260.pdf

@@
 Note that all our proposed algorithms are permutation invariant and \MCTopM{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.
+
+As we state it here, the constraint on fairness is new, and it was not studied much in previous works.
+In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
+We note that only a very recent article studies this measure of performance under a formal approach, \cite{Patil2019stochastic}, but only for the single-player case and therefore they consider a slightly different notion of fairness.
 

### 2019-06-03 2-Chapters/2-Chapter/chapter2.tex
Done for now

@@
 We then give regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
-They both attain a logarithmic regret, proving their order optimality.
+They are order-optimal as they attain a logarithmic regret.
+% proving their order optimality.
 % (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
@@
 
-\paragraph{Decisions can use the past observations.}
+\paragraph{Decisions can (and should) depend on the past observations.}
 
@@
     The problem can either consider a fixed budget and no freedom on the ending of the game (\ie, an known horizon $T$), or a fixed confidence and a certain freedom on the budget (\ie, the identified arm must be the true best arm with probability at least $1-\delta$) \cite{Garivier16BAI}.
-    The theoretical complexity of this use of bandit for A/B testing was first studied in \cite{Kaufmann14},
+    The theoretical complexity of this problem
+    % use of bandit for A/B testing
+    was first studied in \cite{Kaufmann14},
     and a more practical point-of-view was proposed in \cite{Jamieson17ABTest}.
@@
 \begin{proof}\label{proof:2:RegretDecomposition}
-    One can use the chain rule of expectation, and because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$ (which are measurable wrt to the past observations $\cO_t$),
-    one can rewrite the expected cumulated rewards like this,
-    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]$
-    $= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right]$
-    $= \sum_{k=1}^K \E [ r_k(1) \sum_{t=1}^T \mathbbm{1}(A(t) = k) ]$
+    We use the chain rule of expectation, and because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$ (which are measurable wrt to the past observations $\cO_t$),
+    we can rewrite the expected cumulated rewards:
+    \begin{align*}
+        \E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]
+        &= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right] \\
+        &= \sum_{k=1}^K \E \left[ r_k(1) \sum_{t=1}^T \mathbbm{1}(A(t) = k) \right]
         % \underbrace{XXX}_{= N_k(T)}
-    $= \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].$
+        = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].
+    \end{align*}
+    % FIXME there is a conditionning to do here!
     %
@@
 % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
-On a more experimental note, it was showed in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-player bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting means that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
+On a more experimental note, it was showed in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-player bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
 For more details, see \cite{modiDemo2016} for experiments on the multi-player case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
@@
 Let us first describe two naive strategies, that both fail dramatically.
+They are detailed in Algorithm~\ref{algo:2:naiveStrategies} below.
 We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
@@
 The estimated means, or empirical averages, are $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$).
+
+-- The uniform strategy always plays the $K$ arms uniformly at random.
+The player \textcolor{darkgreen}{\emph{only explores}} without using the collected information, and this strategy fails dramatically for any non trivial problem (\ie, linear regret).
+Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
+which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all problems expect where $\mu_1=\mu_2=\dots=\mu_K$).
 
@@
 -- The ``\emph{Follow-the-Leader}'' strategy consists in first playing once each arm, then always playing $A(t)\in\argmax \widehat{\mu_k}(t)$.
-The player only exploits the collected information, and this strategy can fail dramatically, \ie, obtain linear regret in some problems.
+The player \textcolor{blue}{\emph{only exploits}} the collected information, and this strategy can fail dramatically, \ie, obtain linear regret in some problems.
 Indeed consider $K=2$ Bernoulli arms of means $\mu_1=1/2$ and $\mu_2=\varepsilon$ where $\varepsilon < 1/2$, then with probability $\varepsilon/2$ the player observes a reward of $0$ for arm $1$ then $1$ for arm $2$ on the first rounds, and so she will play arm $2$ for the $T-2$ remaining rounds, giving a linear (mean) regret $R_T \geq \frac{\varepsilon}{2}\left(\frac{1}{2} - \varepsilon\right) (T-1)$.
-
--- The uniform strategy always plays the $K$ arms uniformly at random.
-The player only explores without using the collected information, and this strategy fails dramatically for any non trivial problem (\ie, linear regret).
-Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
-which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all problems expect where $\mu_1=\mu_2=\dots=\mu_K$).
-
-% ---------------------------------------------------
-\paragraph{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit}
-
-As illustrated by the two previous examples, an efficient strategy needs to solve the trade-off between exploration and exploitation.
-The two following solutions both consist in splitting the $T$ time steps into $T_0$ steps of exploration and $T-T_0$ steps of exploitation.
-%  either in an alternative way or in a fixed way (``explore then exploit'').
-
-% https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
--- The $\varepsilon$-greedy strategy consists in alternating exploration and exploitation at a certain ratio \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
-Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
-If $\varepsilon$ is constant, then the (mean) regret is still growing linearly as it is lower bounded by $\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k T$.
-In this first case, in average $T_0 = \varepsilon T$ steps are spent on exploration.
-%
-But if we consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 \frac{1}{t}$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$,
-then it was shown in \cite{Auer02} that the regret is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{\log(T)}$.
-In this second case, about $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
-
-% https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
--- The ``explore-then-exploit'' strategy, as studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
-We can lower bound the regret of this strategy by
-$R_T \geq K (\min_{k: \Delta_k > 0} \Delta_k) T_0/K + p (T - T_0)$,
-if $p$ denotes the probability that the chosen arm is not optimal (\ie, the arm with highest mean after $T_0/K$ samples of each arms).
-And we can show that $p$ only depends on the number of collected samples $T_0$ and the gaps $\Delta_k$,
-so if $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, \ie, $R_T = \Omega(T)$.
-
-\begin{proof}
-    We prove it for the case of two arms.
-    If the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality from \cite{hoeffding1963probability} to control this probability $p$.
-    %
-    Indeed if $Z_1,\dots,Z_n$ are \iid{} samples from a Bernoulli distribution of parameter $\mu$, of mean $\widehat{Z_n}$ we have
-    \begin{equation}
-        \forall x < \mu, \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
-        \;\;\;
-        \forall x > \mu, \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
-    \end{equation}
-    %
-    If $\mu_1 = \mu_2 + \Delta$, then
-    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \mu_1 - \Delta/2) + \bP(\widehat{\mu_2} > \mu_2 + \Delta/2)$, and both terms can be bounded by using Hoeffding's inequality with $n=T_0/2$ samples,
-    to obtain $p \leq 2 \exp(-T_0 \Delta^2 / 4)$.
-    %
-    For $K$ arms, the same proofs applies, and we can obtain
-    $p \leq 2 K \exp(-T_0 d^2 / 4)$, if $d$ is a lower-bound on the positive gaps.
-    %
-    In all cases, it shows that $p$ is a constant wrt $T$.
-\end{proof}
-
-Using the same analysis, if a lower-bound $d$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret.
-%
-Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus Hoeffding's inequality gives $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$.
-Optimizing on $T_0$ gives $T_0 = 4/d^2 \log(d^2 T / 4)$
-and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.
-%
-This strategy can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
-
-An extension of this strategy is to consider not a fixed time $T_0$ but a random time $\tau$ at which exploration stops.
-This time $\tau$ must be a \emph{stopping time} in the sense that it is a measurable random variable, dependent of the past observations, as introduced by \cite{Wald45}.
-The strategy is then referred to as ``explore-then-commit'' (ETC), and the idea is to use a statistical test at every time step $t$, and stop as soon as enough samples were collected to effectively identify the best arm with a certain confidence level $\delta$.
-Choosing $\delta \propto 1/T$ and using a lower-bound $d$ on the gaps typically lead to an order-optimal algorithm as shown in \cite{GarivierETC2016}.
-
-For both cases, the strategy obtains sub-optimal regret if it is tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if a lower-bound on the gaps $\Delta_k$ is known.
-%  \ie, they require a prior knowledge on the problem difficulty.
-As such, this weakness make them unapplicable on an unknown problem, and so these first strategies are less interesting from a practical point-of-view.
-
-% ----------------------------------------------------------------------------
-\subsection{Index policies, \UCB, \klUCB{} and others}
-
-A large family of algorithms are index policies, that compute an index $U_k(t)$ on each arm $k$ at time $t$.
-The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
-If more than one index maximizes $(U_k(t))_k$, the arm is chosen among the set, usually in a uniform random manner: $A(t) \sim \cU(\argmax_k U_k(t))$.
-%
-The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that include well known and efficient algorithms such as \UCB, \klUCB{} and many others.
 
@@
 		% \begin{small} % XXX remove if needed
-		\For(){$t = 1, \dots, T$}{
+		\For(){$t = 1, 2, \dots, T$}{
+            \uIf{\textcolor{darkgreen}{pure exploration}}{
+                Play uniformly at random \textcolor{darkgreen}{$A(t) \in \cU(\{1,\dots,K\})$}
+                \tcp*[f]{Explore}
+            }
+            \uIf{\textcolor{blue}{pure exploitation}}{
+                Play uniformly among the arms of maximal empirical mean: \textcolor{blue}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(t))$}
+                \tcp*[f]{Exploit}\\
+                Observe a reward $r_{A(t)}(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
+            }
+		}
+		\caption{Naive strategies: \textcolor{darkgreen}{pure exploitation} or \textcolor{blue}{pure exploration}.}
+		\label{algo:2:naiveStrategies}
+		% \end{small} % XXX remove if needed
+	\end{algorithm}
+	\end{framed}
+\end{figure}
+% \end{small} % XXX remove if needed
+
+% ---------------------------------------------------
+\paragraph{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit}
+
+As illustrated by the two previous examples, an efficient strategy needs to solve the trade-off between exploration and exploitation.
+The two following solutions both consist in splitting the $T$ time steps into $T_0$ steps of exploration and $T-T_0$ steps of exploitation.
+%  either in an alternative way or in a fixed way (``explore then exploit'').
+They are detailed in Algorithm~\ref{algo:2:naiveStrategies}.
+
+% https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
+-- The \textcolor{deeppurple}{$\varepsilon$-greedy strategy} consists in alternating exploration and exploitation at a certain ratio \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
+Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
+If $\varepsilon$ is constant, then the (mean) regret is still growing linearly as it is lower bounded by $\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k T$.
+In this first case, in average $T_0 = \varepsilon T$ steps are spent on exploration.
+%
+But if we consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$,
+then it was shown in \cite{Auer02} that the regret is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{\log(T)}$.
+In this second case, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
+
+% \begin{small} % XXX remove if needed
+\begin{figure}[h!]
+	\centering
+    \begin{framed}
+	\begin{algorithm}[H]
+		% \begin{small} % XXX remove if needed
+		\For(){$t = 1, 2, \dots, T$}{
+            \uIf{\textcolor{deeppurple}{$\varepsilon$-greedy}}{
+                Sample a value uniformly in $[0,1]$: $U_t\sim\cU([0,1])$\\
+                \uIf{\textcolor{deeppurple}{$U_t < \varepsilon$ (\ie, with probability $\varepsilon$)}}{
+                    Play uniformly at random $A(t) \in \cU(\{1,\dots,K\})$
+                    \tcp*[f]{Explore}
+                }
+                \uElse{
+                    Play uniformly among the arms of maximal empirical mean: $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(t))$
+                    \tcp*[f]{or Exploit}
+                }
+            }
+            \uIf{\textcolor{deepgold}{Explore-then-Exploit}}{
+                \uIf{$t \leq \textcolor{deepgold}{T_0}$}{
+                    Play uniformly at random $A(t) \in \cU(\{1,\dots,K\})$
+                    \tcp*[f]{Explore}
+                }
+                \uElse{
+                    \uIf{$t = \textcolor{deepgold}{T_0}$}{
+                        Pick one arm $A(\textcolor{deepgold}{T_0})=k$ uniformly among the arms of maximal empirical mean: \textcolor{deepgold}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(\textcolor{deepgold}{T_0}))$}
+                        \tcp*[f]{then Exploit}
+                    }
+                    Play the same arm $A(t) = A(\textcolor{deepgold}{T_0})$
+                }
+            }
+            Observe a reward $r_{A(t)}(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
+		}
+		\caption{Simple efficient strategies: \textcolor{deeppurple}{$\varepsilon$-greedy} and \textcolor{deepgold}{Explore-then-Exploit}.}
+		\label{algo:2:naiveStrategies}
+		% \end{small} % XXX remove if needed
+	\end{algorithm}
+	\end{framed}
+\end{figure}
+% \end{small} % XXX remove if needed
+
+% https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
+-- The ``\textcolor{deepgold}{Explore-then-Exploit}'' strategy, as studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
+We can lower bound the regret of this strategy by
+$R_T \geq K (\min_{k: \Delta_k > 0} \Delta_k) T_0/K + p (T - T_0)$,
+if $p$ denotes the probability that the chosen arm is not optimal (\ie, the arm with highest mean after $T_0/K$ samples of each arms).
+And we can show that $p$ only depends on the number of collected samples $T_0$ and the gaps $\Delta_k$,
+so if $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, \ie, $R_T = \Omega(T)$.
+
+\begin{proof}
+    We prove it for the case of two arms.
+    If the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality from \cite{hoeffding1963probability} to control this probability $p$.
+    %
+    Indeed if $Z_1,\dots,Z_n$ are \iid{} samples from a Bernoulli distribution of parameter $\mu$, of mean $\widehat{Z_n} = \frac{1}{n} \sum_{i=1}^n Z_i$, Hoeffding's inequality gives both
+    \begin{equation}
+        \forall x < \mu, \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
+        \;\;\;
+        \forall x > \mu, \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+    \end{equation}
+    %
+    If $\mu_1 = \mu_2 + \Delta$, then
+    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \mu_1 - \Delta/2) + \bP(\widehat{\mu_2} > \mu_2 + \Delta/2)$, and both terms can be bounded by using Hoeffding's inequality with $n=T_0/2$ samples,
+    to obtain $p \leq 2 \exp(-T_0 \Delta^2 / 4)$.
+
+    A simple union bound extends the proof for $K$ arms, giving
+    $p \leq 2 K \exp(-T_0 d^2 / 4)$, if $d$ is a lower-bound on the positive gaps.
+    %
+    In all cases, it shows that $p$ is a constant wrt $T$.
+\end{proof}
+
+Using the same analysis, if a lower-bound $d$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret.
+%
+Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus Hoeffding's inequality gives $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$.
+Optimizing on $T_0$ gives $T_0 = 4/d^2 \log(d^2 T / 4)$
+and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.
+%
+This strategy can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
+
+An extension of this strategy is to consider not a fixed time $T_0$ but a random time $\tau$ at which exploration stops.
+This time $\tau$ must be a \emph{stopping time} in the sense that it is a measurable random variable, dependent of the past observations, as introduced by \cite{Wald45}.
+The strategy is then referred to as ``explore-then-commit'' (ETC), and the idea is to use a statistical test at every time step $t$, and stop as soon as enough samples were collected to effectively identify the best arm with a certain confidence level $\delta$.
+Choosing $\delta \propto 1/T$ and using a lower-bound $d$ on the gaps typically lead to an order-optimal algorithm as shown in \cite{GarivierETC2016}.
+
+For both cases, the strategy obtains sub-optimal regret if it is tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if a lower-bound on the gaps $\Delta_k$ is known.
+%  \ie, they require a prior knowledge on the problem difficulty.
+As such, this weakness make them unapplicable on an unknown problem, and so these first strategies are less interesting from a practical point-of-view.
+
+% ----------------------------------------------------------------------------
+\subsection{Index policies, \UCB, \klUCB{} and others}
+
+A large family of algorithms are index policies, that compute an index $U_k(t)$ on each arm $k$ at time $t$.
+The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
+If more than one index maximizes $(U_k(t))_k$, the arm is chosen among the set, usually in a uniform random manner: $A(t) \sim \cU(\argmax_k U_k(t))$.
+%
+The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that include well known and efficient algorithms such as \UCB, \klUCB{} and many others.
+
+% \begin{small} % XXX remove if needed
+\begin{figure}[h!]
+	\centering
+    \begin{framed}
+	\begin{algorithm}[H]
+		% \begin{small} % XXX remove if needed
+		\For(){$t = 1, 2, \dots, T$}{
             \uIf{$t \leq K$}{
@@
 			\Else(){
-                For each arm, compute the index $U_k(t)$\;
-                Play arm $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
-            }
-            Observe a reward $r_{A(t)}(t)$\;
-            Update both $X_k(t)$ and $N_k(t)$\;
-		}
-		\caption{A generic index policy, using indexes $U_k(t)$ (\eg, \UCB, \klUCB{} etc).}
+                For each arm, compute the \textcolor{blue}{index $U_k(t)$}\;
+                Play arm $A(t)$ uniformly among the arms of maximal index: \textcolor{blue}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k(t))$}\;
+            }
+            Observe a reward $r_{A(t)}(t)$ and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$\;
+		}
+		\caption{A generic \textcolor{blue}{index} policy, using \textcolor{blue}{indexes $U_k(t)$} (\eg, \UCB, \klUCB{} etc).}
 		\label{algo:2:indexPolicy}
@@
 \begin{align}\label{eq:UCB_index}
-    U_k^{\UCB}(t) = \underbrace{\frac{X_k(t)}{N_k(t)}}_{\widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}\right)}}_{\xi_k(t)}.
+    U_k^{\UCB_1}(t) = \UCB_k(t) = \underbrace{\frac{X_k(t)}{N_k(t)}}_{\widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}\right)}}_{\xi_k(t)}.
 \end{align}
 
-This selection rule makes sense for the following reason:
-an arm $k$ is chosen at step $t$ because it has a large $U_k^{\UCB}(t) = \UCB_k(t)$, which can happen for two reasons:
-$i)$ because the average reward $\widehat{\mu_k}(t)$ is large, in which case this arm is likely to have a high reward,
+This selection rule \eqref{eq:UCB_index} makes sense for the following reason:
+an arm $k$ is chosen at step $t$ because it has a large $\UCB_k(t)$, which can happen for two reasons:
+$i)$ because the average reward $\widehat{\mu_k}(t)$ is large, in which case this arm is likely to have a high mean reward $\mu_k$,
 $ii)$ and/or because the confidence radius $\xi_k(t)$ is large, in which case this arm has not been explored much.
 Either reason makes this arm worth choosing.
-%
-Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{}  respectively represent exploitation and exploration, and summing them up is a natural way to trade off the two.
-The parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to chose $\alpha=1/2$ for optimal uniform performance.
+One can also observe that $\xi_k(t)\to\infty$ if $t\to\infty$ but $N_k(t)\ll\log(t)$, thus all arms must be explored at least $\mathcal{O}(\log(T))$ times.
+%
+Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{}  respectively represent exploitation and exploration, and summing them up is a natural way of trading them off.
+The parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to choose $\alpha=1/2$ for optimal uniform performance.
 
@@
 XXX
+\TODOL{Expliquer la fin}
 
 More details on the proofs can be found for instance in Chapter~2 of \cite{Bubeck12}, or Section~1.3 of \cite{Slivkins2019}.
-% \TODOL{Expliquer la fin}
 The $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
@@
 
+The \klUCB{} algorithm XXX
 Optimal upper confidence bounds based on Kullback-Leibler divergence and klUCB indexes
 
-If rewards lie in $[a,b]$, then the indexes are computed as follows:
+% If rewards lie in $[a,b]$, then t
+The indexes are computed as follows, for an exploration function $f(t)$
+that is typically chosen as $f(t)=\log(t)+3\log(\log(t))$ for the theoretical analyses,
+and $f(t)=\log(t)$ in practice.
+The constant $c>0$ is usually chosen as $c=1$.
 \begin{align}\label{eq:klUCB_index}
-    U_k^{\klUCB}(t) = \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c \log(t)}{N_k(t)} \right\}.
+    U_k^{\klUCB}(t) = \sup\limits_{q \in [0, 1]} \left\{ q : \KL\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
 \end{align}
 
-\TODOL{Short note on how to implement such policy, using Newton's method ?}
+\paragraph{A note about the implementation of \klUCB{} algorithmes.}
+%
+% \TODOL{Short note on how to implement such policy, using Newton's method ?}
+Implementing this kind of policy requires to implement two things.
+The first piece is usually easy: one has to implement a way to compute $\KL(x,y)$ the Kullback-Leibler divergence of the considered one-dimensional family,
+for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
+The second piece is less obvious: one has to solve the optimisation problem defining the $\sup$ in \eqref{eq:klUCB_index}.
+Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\KL(x,\bullet)$ function is usually convex (for any $x$), one practical and mathematically correct solution is to use Newton's method.
+The search space for $q$ can be restricted to $[\frac{X_k(t)}{N_k(t)}, 1]$ instead of $[0,1]$, and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
+For instance, $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically Newton's method terminates in between $8$ to $12$ steps.
+
+If we perform numerical simulations with \klUCB{} algorithms, we are interested in having efficient implementations, and ideally \klUCB{} should not be significantly slower than \UCB.
+The bottleneck of the performance is the computations of the $\KL(x,y)$ functions and the optimization problem,
+and we note that our library SMPyBandits implements both parts in both a naive Python implementation (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback.py}{\texttt{kullback.py}}) and in a Python module written in a
+C file\footnote{~We also note that we wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
+    These implementations of $\KL(x,y)$ and \klUCB{} indexes are also published as an independent Python package, hosted on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} and installable from Pypi.
+    A Julia version of the same package is also available at \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}},
+    and installable with \texttt{Pkg.add("KullbackLeibler")}.
+},
+in an optimized way (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/C/kullback_py3.c}{\texttt{kullback\_py3.c}}).
+For the supported families of distributions (including Bernoulli, Gaussian, exponential and Poisson), we wrote both a naive and an optimized version of $\KL(x,y)$ and the $\sup_q \KL(x,q) \leq z$ optimization problem, using Newton's method.
 
@@
         % \begin{small} % XXX remove if needed
-		\For(){$t = 1, \dots, T$}{
+		\For(){$t = 1, 2, \dots, T$}{
             For each arm, sample $s_k(t) \sim \pi_k(t-1)$ from the posterior of arm $k$, $\pi_k(t-1)=\mathrm{Beta}(X_k(t-1)+1,N_k(t-1)-X_k(t-1)+1)$\;
-            Play arm $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
-            Observe a reward $r_{A(t)}(t)$\;
-            Update both $X_k(t)$ and $N_k(t)$\;
+            Play arm $A(t)$ uniformly among the arms of maximal random sample: $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k(t))$\;
+            Observe a reward $r_{A(t)}(t)$ and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$\;
 		}
@@
 Bayes-UCB from \cite{Kaufmann12BUCB} also uses the Bayesian point of view, updated like for Thompson sampling, but used in a different way.
-Instead of sampling from the posteriors and playing the arms with maximum sample, Bayes-UCB computes the $1-1/t$ quantile of each arm, at time $t$, and plays the arm with the largest quantile.
+Instead of sampling from the posteriors and playing the arms with maximum sample, Bayes-UCB computes the $(1-1/t)$ quantile of each arm, at time $t$, and plays the arm with the largest quantile.
 It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
 %
-Bayes-UCB is comparable to Thompson sampling in terms of memory complexity, but it can be slower in terms of time complexity as computing a $1-1/t$ quantile is generally more costly than sampling from a distribution.
+Bayes-UCB is comparable to Thompson sampling in terms of memory complexity, but it can be slower in terms of time complexity as computing a quantile is generally more costly than just sampling from a distribution.
 % a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
@@
 The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as \UCB{} or Thompson sampling, with uniform perturbations.
-Their RCB algorithm ressembles \UCB, and builds the same confidence intervals $\mathrm{LCB}_k(t), \UCB_k(t)$ on arm $k$ at time $t$.
+Their RCB algorithm ressembles \UCB, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$).
 But instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
@@
 
-BESA (Best Empirical Sampled Average) was introduced in \cite{Baransi2014}, in order to fix a simple observation:
+\textbf{BESA} (Best Empirical Sampled Average) was introduced in \cite{Baransi2014}, in order to fix a simple observation:
 for instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, their empirical mean).
@@
 The first example is MOSS from \cite{Audibert2009minimax}, or recently \KLUCBpp{} in \cite{Menard17}.
-They are both index-based, using the following indexes:
+They are both index-based (see Algorithm~\ref{algo:2:indexPolicy}), using the following indexes:
 \begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp}
-    U^{\mathrm{MOSS}}_k(t) &= \frac{X_k(t)}{N_k(t)} + \sqrt{\max\left(0, \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)}\right)}. \\
+    U^{\mathrm{MOSS}}_k(t) &= \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)} } \right). \\
     U^{\KLUCBpp}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
@@
 %
-If we denote $g(t, T, K) = \log^+(y (1 + \log^+(y)^2))$, and $y = \frac{T}{K t}$, where $\log^+(x) = \max(0, \log(x))$.
-
-Both algorithms are index policies, that are not anytime.
+If we denote $g(t, T, K) = \log^+(y(t) (1 + \log^+(y(t))^2))$, and $y(t) = \frac{T}{K t}$, with $\log^+(x) = \max(0, \log(x))$.
+
+Both algorithms are index policies, but they are not anytime.
 For any $K$, $T$, and problem instances $I$ in a certain family $\cI$ (\eg, with Bernoulli distributions),
@@
 \begin{align}\label{eq:2:indexes_for_klUCBswitch}
-    U^{\mathrm{MOSS}+}_k(t) &= \frac{X_k(t)}{N_k(t)} + \sqrt{\max\left(0, \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}\right)}. \\
+    U^{\mathrm{MOSS}+}_k(t) &= \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
     U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(T / (K N_k(t)))}{N_k(t)} \right\} \\
@@
 
-FIXME?
-\TODOL{Je dois conclure un peu cette section...}
+\paragraph{Summary.}
+%
+We gave an overview of the most well known families of algorithms for multi-armed bandits,
+focussing on algorithms designed and analyzed for the single-player stationary and stochastic MAB model.
+Except the naive and simple strategies given as introductory examples, most of the policies presented above are order-optimal for Bernoulli distributed problems or bounded rewards (or one-dimensional exponential families),
+and some of them are known to be optimal in different settings.
+%
+In the rest of this thesis, we mainly use the \UCB{} or Thompson sampling algorithms when simplicity is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when optimality is important and where we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
 

### 2019-06-04 2-Chapters/2-Chapter/chapter2.tex
Better discussion on the two notations of fairness

@@
 
-We give an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the \UCB, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
+We give an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the $\UCB_1$, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
 % We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contributions of this thesis.
@@
 %
-We then give regret upper-bounds of the two main algorithms used in this thesis, \UCB{} and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
+We then give regret upper-bounds of the two main algorithms used in this thesis, $\UCB_1$ and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
 They are order-optimal as they attain a logarithmic regret.
@@
 %
-In the second Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} below, we display the result of an example of run, when the player was following the \UCB{} algorithm from \cite{Auer02}.
+In the second Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} below, we display the result of an example of run, when the player was following the $\UCB_1$ algorithm from \cite{Auer02}.
 After $T=100$ steps, the player obtained a cumulated reward of $56$, by playing mostly arms $4,3,5,1,0$ (in decreasing order of number of plays).
@@
 %
-If the player knows that reward are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ ($a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = $ that lies in $[0,1]$.
-Note that this ``normalization trick'' is implemented for any policy in SMPyBandits, with the \texttt{lower} and \texttt{amplitude} optional arguments, respectively representing $a$ the lower-bound on rewards and $b-a$ the length of the interval of possible values of rewards.
+If the player knows that reward are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ (for $a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = $ that lies in $[0,1]$.
+Note that this ``normalization trick'' is implemented for any policy in our library SMPyBandits, with the \texttt{lower} and \texttt{amplitude} optional arguments, respectively representing $a$ the lower-bound on rewards and $b-a$ the length of the interval of possible values of rewards.
 
@@
 
-In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the applications of choice.
+\textbf{Hypotheses in this thesis.}
+%
+In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the considered applications.
 First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
@@
 %
-Then, when we study multi-player bandits in Chapter~\ref{chapter:5}, for the model without sensing, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $r(t)=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
-%
-Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm in the setting on bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use the Bernoulli Kullback-Leibler divergence $\kl$ in our analysis.
+Then, when we study multi-player bandits in Chapter~\ref{chapter:5}, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $r(t)=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
+%
+Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm for bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use the Bernoulli Kullback-Leibler divergence $\kl$ in our analysis.
 
@@
 
-Historically, MAB models were first applied to \textbf{clinical trials}, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
+\textbf{Clinical trials} have been the first historical application of MAB models, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
 The mean of an arm, in this application, represents the mean success rate of a treatment, and the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
@@
     Other possible use cases include hyper-parameter optimization, or features selection.
-    Hyper-parameters include real-valued parameters, like $\alpha\in[0,\infty)$ for the \UCB{} or $\varepsilon$ for the $\varepsilon$-greedy bandit algorithms, a step size multiplier in a gradient method, or the width of a radial based function (RBF) kernel method.
+    Hyper-parameters include real-valued parameters, like $\alpha\in[0,\infty)$ for the $\UCB_1$ or $\varepsilon$ for the $\varepsilon$-greedy bandit algorithms, a step size multiplier in a gradient method, or the width of a radial based function (RBF) kernel method.
     Discrete-valued parameters are also common, like a choice in a fixed set of kernel functions or the depth of neural networks,
     and higher dimensional or more complex hyper-parameters can for instance be the entire architecture of a neural network.
-
-    % % FIXME this is copied pasted from \cite{bouneffouf2019survey}
-    % It is well known that performance of machine learning algorithms depends critically on identifying a good set of hyper-parameters.
-    % While recent approaches use Bayesian optimization to adaptively select optimal hyper-parameter configurations, they rather focus on speeding up random search through adaptive resource allocation and early-stopping.
-    % A different approach was initiated in \cite{LiJamieson2018}, by formulating hyper-parameter optimization as a pure-exploration non-stochastic infinite-armed bandit problem, where predefined resources, such as iterations, data samples, or features are allocated to randomly sampled configurations.
-This work introduced a novel algorithm, Hyperband, for this framework and analyzed its theoretical properties, providing several desirable guarantees.
-    % Furthermore, Hyperband was compared with popular Bayesian optimization methods on a suite of hyper-parameter optimization problems; it was observed that Hyperband can provide more than an order-of-magnitude speedup over its competitors on a variety of deep-learning and kernel-based learning problems.
-% \end{itemize}
-
-% - clinical trial
-% - A/B testing of websites
-% - online content recommandation
-% - wireless communication, lots of possibilities, one is frequency/channel selection in the OSA paradigm, detailed after
-% - hyper-parameters tuning of deep neural networks etc
 
@@
 Let us first introduce some notations.
-From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that is, for any time $t$, $r_k(t)\sim \nu_k$.
+From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that generates \iid{} rewards $r_k(t)\sim \nu_k$, for any time $t$.
 We focus on distributions fully characterized by their means, that can be Bernoulli or any one-dimensional exponential family.
-We denote $\mu_k$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$} in the rest of the manuscript), that is $\mu_k = \E[r_k(t)] \in \R$.
+We denote $\mu_k = \E[r_k(t)] \in \R$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$}).
 %
@@
 If the goal of the player is to maximize $\E\left[\sum_{t=1}^T r_{A(t)}(t)\right]$,
-the optimal strategy for this bandit problem is to always pull an optimal arm, but of course it is unrealistic as the player does not know the true means and cannot know which arm is optimal.
+the optimal strategy for this bandit problem is to always pull one of the optimal arms (it can be not unique), but it is unrealistic as the player does not know the true means, nor the optimal arms (as long as they are not all optimal, \ie, in non trivial problems).
 %
 Comparing the difference between the performance of a fixed baseline and that of the player is a common approach in machine learning research,
-and here we can compare with the oracle strategy that obtains an (expected) reward of $\mu^*$ at each time step.
+and here we can compare with the oracle strategy that always obtains an (expected) reward of $\mu^*$.
 %
@@
 let us introduce the (mean) regret $R_T^{\cA}$ of an algorithm $\cA$ as
-\[ R_T^{\cA} = \E\left[ \sum_{t=1}^T r_{k^*}(t) - r_{A(t)}(t) \right]. \]
-%
-We can rewrite this expression to obtain the following definition of the regret:
-$R_T^{\cA}$
-$= \E\left[ \sum_{t=1}^T r_{k^*}(t) \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right]$
-$= T \times \E\left[ r_{k^*}(t=1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]$
-$= T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]$.
+$R_T^{\cA} = \E\left[ \sum_{t=1}^T (r_{k^*}(t) - r_{A(t)}(t)) \right]$.
+%
+As the rewards from arm $k^*$ are \iid{} (as for all arms), and by linearity of the expectancy, we can rewrite this expression to obtain the following definition of the regret:
+\begin{align*}
+    R_T^{\cA}
+    & = \E\left[ \sum_{t=1}^T (r_{k^*}(t) - r_{A(t)}(t)) \right]
+    = \E\left[ \sum_{t=1}^T r_{k^*}(t) \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right] \\
+    & = T \times \E\left[ r_{k^*}(t=1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]
+    = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
+\end{align*}
 % This is the definition we give below.
@@
 \begin{defn}[Regret]\label{def:2:regret}
-    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the (mean) regret at horizon $T$ is $R_T^{\cA}$ and it is defined as
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ and it is defined as
     %
     \begin{equation}
-        R_T^{\cA} = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right] = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
+        R_T^{\cA}
+        = \E\left[ \sum_{t=1}^T (r_{k^*}(t) - r_{A(t)}(t)) \right]
+        % = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right]
+        = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
     \end{equation}
@@
     The (mean) regret $R_T^{\cA}$ can be decomposed as a sum of the number of selections of sub-optimal arms $k$,
-    weighted by their gap, like this:
+    weighted by their gaps, like this:
     \begin{equation}\label{eq:2:RegretDecomposition}
@@
 \begin{proof}\label{proof:2:RegretDecomposition}
-    We use the chain rule of expectation, and because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$ (which are measurable wrt to the past observations $\cO_t$),
-    we can rewrite the expected cumulated rewards:
+    We use the chain rule of expectation and a conditioning on $O_t$,
+    because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$, which are measurable wrt to the past observations $O_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$.
+    Thus we can rewrite the expected cumulated rewards as follows:
+
+    % \vspace*{-20pt}
     \begin{align*}
         \E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]
-        &= \E \left[ \sum_{k=1}^K \sum_{t=1}^T r_k(t) \mathbbm{1}(A(t) = k) \right] \\
-        &= \sum_{k=1}^K \E \left[ r_k(1) \sum_{t=1}^T \mathbbm{1}(A(t) = k) \right]
-        % \underbrace{XXX}_{= N_k(T)}
+        &= \E \left[ \sum_{t=1}^T \sum_{k=1}^K r_k(t) \mathbbm{1}(A(t) = k) \right]
+        = \sum_{k=1}^K \sum_{t=1}^T \E \left[ r_k(t) \mathbbm{1}(A(t) = k) \right] \\
+        &= \sum_{k=1}^K \sum_{t=1}^T \E \Bigl[ \E \left[ r_k(t) | O_t \right] \mathbbm{1}(A(t) = k) \Bigr]
+        = \sum_{k=1}^K \sum_{t=1}^T \E \Bigl[ \E \left[ r_k(1) \right] \mathbbm{1}(A(t) = k) \Bigr] \\
+        &= \sum_{k=1}^K \underbrace{\E \left[ r_k(1) \right]}_{\mu_k} \underbrace{\sum_{t=1}^T \E \left[ \mathbbm{1}(A(t) = k) \right]}_{= N_k(T)}
+        % = \sum_{k=1}^K \mu_k \E \left[ \sum_{t=1}^T \mathbbm{1}(A(t) = k) \right]\\
+        %  XXX 
         = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].
     \end{align*}
-    % FIXME there is a conditionning to do here!
-    %
-    Thus $R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]$.
+    %
+    Thus $R_T^{\cA} = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right] = T \mu^* - \sum_{k=1}^K \mu_k \E[ N_k(T) ] = \sum_{k=1}^K (\mu^* - \mu_k) \E[ N_k(T) ]$,
+    as $\sum_{k=1}^K  \E[ N_k(T) ] = T$.
     The sum can then be simplified to only account for sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
@@
 
-\paragraph{Consequences.}
-%
-Such decomposition of the regret is useful for at least two reasons.
+\textbf{Consequences.}
+%
+Such decomposition of the regret is useful for at least two reasons:
 
@@
     \item
-On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if one wants to compute and visualize the regret, she can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
-Mathematically, both quantities are equal in expectation, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition}.
-In our library SMPyBandits, we implement both estimators, and all values of regret used in this thesis are based on the one using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives smoother curves and faster convergence to the limit behavior.
+On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if we want to compute and visualize the regret, we can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
+Both quantities are indeed equal \emph{in expectation}, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition} (thanks to the conditioning on $O_t$ done in the proof).
+In our library SMPyBandits, we implement both estimators, and all values of regret used in this thesis are based on the one using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives faster convergence, and also ``smoother'' plots.
 
     \item
-On the other hand, this decomposition is necessary as theoretical analyses of the regret of MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
-For example, we extend this decomposition to the multi-player case in Chapter~\ref{chapter:3}, in Lemma~\ref{lem:5:DecompositionRegret}, and as it also uses the draws $N_k(T)$, it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
+On the other hand, this decomposition is necessary as theoretical analyses of the regret of (single-player) MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
+Even if we prove that controlling $N_k(T)$ is no longer sufficient to obtain low regret for the multi-player case, we extend this decomposition in Chapter~\ref{chapter:3} (in Lemma~\ref{lem:5:DecompositionRegret}), and it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
 \end{itemize}
@@
 
-Moreover, it is interesting to note that this lower-bound can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proved to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression in \eqref{eq:2:forSecondLogTLowerBound2} essentially says that any efficient algorithm
-should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times in the total of $T$ time steps.
-
 \begin{theorem}\label{thm:2:firstLogTLowerBound}
@@
     An algorithm $\cA$ is \emph{uniformly efficient} if its (mean) regret satisfies
-    \begin{equation}
-        R_T^{\cA}(I) = \smallO{C_{I,\alpha} T^{\alpha}}
-    \end{equation}
+    $R_T^{\cA}(I) = \smallO{C_{I,\alpha} T^{\alpha}}$,
     for any value $\alpha>0$ and any problem instance $I\in \cI$,
@@
 
-This family is non-empty as it contains for instance the \UCB{} algorithm, as its regret is proven to be logarithmic on any problem instance.
+This family is non-empty as it contains for instance the $\UCB_1$ algorithm, as its regret is proven to be logarithmic on any problem instance.
 Now we can state the second logarithmic lower-bound, for algorithms in this family:
@@
         The bound from Theorem~\ref{thm:2:secondLogTLowerBound} holds with,
-        \begin{equation}\label{eq:2:forSecondLogTLowerBound}
-            C_I = \mu^* (1 - \mu^*) \sum_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{1}{\Delta_k}.
-        \end{equation}
+        % \begin{equation}\label{eq:2:forSecondLogTLowerBound}
+        $C_I = \mu^* (1 - \mu^*) \sum\limits_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{1}{\Delta_k}$.
+        % \end{equation}
         \item
         For any $\varepsilon>0$, the bound from Theorem~\ref{thm:2:secondLogTLowerBound} also holds with,
-        \begin{equation}\label{eq:2:forSecondLogTLowerBound2}
-            C_I = \sum_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon.
-        \end{equation}
+        % \begin{equation}\label{eq:2:forSecondLogTLowerBound2}
+        $C_I = \sum\limits_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon$.
+        % \end{equation}
     \end{itemize}
@@
 
+Moreover, it is interesting to note that the second lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proved to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
+should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times in the total of $T$ time steps.
+
 Many algorithms have been proved to achieve logarithmic regret in the stochastic case,
-and in particular it is the case of the algorithms used in this thesis, \UCB{} from \cite{Auer02}, Thompson sampling from \cite{Thompson33} and analyzed in \cite{AgrawalGoyal11,Kaufmann12Thompson}, and \klUCB{} from \cite{Garivier11KL,KLUCBJournal}.
-%
-Such bounds are valid in different settings, and in particular \UCB{} is order-optimal for bounded rewards or one-dimensional exponential families,
+and in particular it is the case of the algorithms used in this thesis, $\UCB_1$ from \cite{Auer02}, Thompson sampling from \cite{Thompson33} and analyzed in \cite{AgrawalGoyal11,Kaufmann12Thompson}, and \klUCB{} from \cite{Garivier11KL,KLUCBJournal}.
+%
+Such bounds are valid in different settings, and in particular $\UCB_1$ is order-optimal for bounded rewards or one-dimensional exponential families,
 while Thompson sampling is optimal for bounded rewards, and \klUCB{} have been proven to be optimal for both cases.
@@
 \paragraph{Worst-case lower-bound in $\Omega(\sqrt{T})$.}
-.
+%
 For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses a constant $C_I$ no longer bring any information.
@@
 Chose any small $\varepsilon$ such that $\varepsilon \ll \sqrt{4 \log(T) / T}$, then chose let $\mu_1 = 1/2$ and $\mu_2 = 1/2 - \varepsilon$.
-The constant $C_I$ from \eqref{eq:2:forSecondLogTLowerBound} is $C_I = 1 / (4 \varepsilon)$ and $\Delta=\varepsilon$ and it becomes so large that
+The constant $C_I$ from the first case of Theorem~\ref{thm:2:forSecondLogTLowerBound} is $C_I = 1 / (4 \varepsilon)$ and $\Delta=\varepsilon$ and it becomes so large that
 the logarithmic upper-bound is useless in such case, as
@@
 \end{theorem}
-
-% \TODOL{FIXME explain why this is interesting for the rest of this manuscript!}
 
@@
 %
-% FIXME we give notations but I don't use them in Chapter5 
 It is more interesting to studying $SC_M^{\cA}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game, as in this case, it is possible that the players follow an efficient centralized or decentralized algorithm (which gives a logarithmic centralized regret) while still suffering from a linear switching cost.
@@
 % - Fairness ?
-Finally, another interesting measure of performance is the \emph{fairness} between the $M$ players in a multi-player bandit models, as we explain in Chapter~\ref{chapter:5}.
-It was not studied much, and only a very recent article studies this measure of performance under a formal approach, \cite{Patil2019stochastic}.
+Finally, another interesting measure of performance is the \emph{fairness}.
+It was not studied much for single-player problems, and usually it means that each arm must be explored at least a given fraction of the total number of rounds $T$ (\ie, ``arm-wise fairness'').
+It was studied in a very recent article \cite{Patil2019stochastic},
+where a different notion of regret is introduced to account for the fairness constraint.
+The authors show that different algorithms, based on $\UCB_1$ or Thompson sampling, can achieve logarithmic regret while respecting the fairness constraints.
+%
+In this thesis, we only consider fairness in the multi-player bandit models,
+in Chapter~\ref{chapter:5}, where fairness refers to a different notion (\ie, ``cooperative fairness'').
+When $M<K$ players learn in a decentralized way, they will essentially converge to an orthogonal affectation of players to arms, so that each player uses a unique arm among the $M$ best arms (most of the times).
+The fairness constraint says that, in average, no user is arbitrarily going to converge more frequently on a better arm than another user, and it is for instance the case for permutation invariant decentralized algorithms.
+We refer to Section~\ref{sub:5:betterLowerBound} for more details.
 
@@
 %
-We focus then on index policies, with a first efficient policy, \UCB, that uses upper confidence bounds (UCB) on the means estimates, and is known to be order-optimal for bounded rewards.
-Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of \UCB{} but use smaller confidence intervals and thus it typically obtains better theoretical results, and Thompson sampling which replaces the principle of ``optimism under uncertainty'' by a Bayesian point-of-view.
+We focus then on index policies, with a first efficient policy, $\UCB_1$, that uses upper confidence bounds (UCB) on the means estimates, and is known to be order-optimal for bounded rewards.
+Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of $\UCB_1$ but use smaller confidence intervals and thus it typically obtains better theoretical results, and Thompson sampling which replaces the principle of ``optimism under uncertainty'' by a Bayesian point-of-view.
 %
@@
 
-For both cases, the strategy obtains sub-optimal regret if it is tuned independently of the problem at hand, but they can be tuned to be efficient (\ie, with logarithmic regret) if a lower-bound on the gaps $\Delta_k$ is known.
+For both cases, the strategy obtains sub-optimal regret if it is tuned independently of the problem at hand, but can be tuned to be efficient (\ie, with logarithmic regret) if a lower-bound on the gaps $\Delta_k$ is known.
 %  \ie, they require a prior knowledge on the problem difficulty.
-As such, this weakness make them unapplicable on an unknown problem, and so these first strategies are less interesting from a practical point-of-view.
+As such, this weakness makes the presented strategies unapplicable on an unknown problem, and so they are less interesting from a practical point-of-view.
 
 % ----------------------------------------------------------------------------
-\subsection{Index policies, \UCB, \klUCB{} and others}
+\subsection{Index policies, $\UCB_1$, \klUCB{} and others}
 
@@
 The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
-If more than one index maximizes $(U_k(t))_k$, the arm is chosen among the set, usually in a uniform random manner: $A(t) \sim \cU(\argmax_k U_k(t))$.
-%
-The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that include well known and efficient algorithms such as \UCB, \klUCB{} and many others.
+If more than one indexes maximize $(U_k(t))_k$, the arm is chosen among the set, usually in a uniformly random manner: $A(t) \sim \cU(\argmax_k U_k(t))$.
+%
+The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that include well known and efficient algorithms such as $\UCB_1$, \klUCB{} and many others.
 
@@
 		}
-		\caption{A generic \textcolor{blue}{index} policy, using \textcolor{blue}{indexes $U_k(t)$} (\eg, \UCB, \klUCB{} etc).}
+		\caption{A generic \textcolor{blue}{index} policy, using \textcolor{blue}{indexes $U_k(t)$} (\eg, $\UCB_1$, \klUCB{} etc).}
 		\label{algo:2:indexPolicy}
@@
 
-\paragraph{The \UCB{} index policy: using Hoeffding's inequality to build confidence intervals.}
+\paragraph{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}
 
@@
 
-\paragraph{A note about the implementation of \klUCB{} algorithmes.}
+\paragraph{A note about the implementation of \klUCB{} algorithms.}
 %
@@
 
-\paragraph{Variants on \UCB}
-
-Many variants of the \UCB{} algorithms have been studied in the multi-armed bandits literature.
+\textbf{Variants on $\UCB_1$.}
+%
+Many variants of the $\UCB_1$ algorithms have been studied in the multi-armed bandits literature.
 For instance, \UCB-H replaces the $\log(t)$ by a $\log(T)$ (as well as \klUCB-H), and it actually corresponds to the first algorithm analyzed in \cite{Auer02}.
@@
 %
-All these variants are usually comparable to the original \UCB{} algorithm, in terms of time and memory complexity, and they achieve improved regret upper bounds in the same setting or some extensions of the initial setting.
-
-\paragraph{Other policies inspired by \UCB}
+All these variants are usually comparable to the original $\UCB_1$ algorithm, in terms of time and memory complexity, and they achieve improved regret upper bounds in the same setting or some extensions of the initial setting.
+
+\textbf{Other policies inspired by \UCB.}
 %
@@
 Thompson sampling uses a Bayesian point of view:
-instead of building confidence intervals on the means of the arms like \UCB{} does,
+instead of building confidence intervals on the means of the arms like $\UCB_1$ does,
 it starts with a prior on the $K$ distributions (\eg, a flat uniform prior), and then after every observation $(k, r(t))$ it updates the posterior distribution of the arm $k$ with a new data $r(t)$.
@@
 Bayes-UCB from \cite{Kaufmann12BUCB} also uses the Bayesian point of view, updated like for Thompson sampling, but used in a different way.
-Instead of sampling from the posteriors and playing the arms with maximum sample, Bayes-UCB computes the $(1-1/t)$ quantile of each arm, at time $t$, and plays the arm with the largest quantile.
+Instead of sampling from the posteriors and playing the arms with maximum sample, Bayes-UCB computes the $(1-1/(t (\log(t))^5)$ quantile of each arm, at time $t$, and plays the arm with the largest quantile.
 It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
@@
 % With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
-The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as \UCB{} or Thompson sampling, with uniform perturbations.
-Their RCB algorithm ressembles \UCB, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$).
+The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as $\UCB_1$ or Thompson sampling, with uniform perturbations.
+Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$).
 But instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
@@
 \textbf{BESA} (Best Empirical Sampled Average) was introduced in \cite{Baransi2014}, in order to fix a simple observation:
-for instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, their empirical mean).
-The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have respectively $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
+if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare their empirical means.
+Based on this idea, the BESA algorithm for $K=2$ arms is quite simple: at each time step, if arms $1$ and $2$ have respectively $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
 It was analyzed partially, for the two-armed case it was proved to be asymptotically optimal for one-dimensional exponential families, and the authors illustrated that it can be very efficient empirically.
@@
 % ----------------------------------------------------------------------------
-\paragraph{Policies for adversarial bandits.}
-
+\textbf{Policies for adversarial bandits.}
+%
 Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
@@
 which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details see
-on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
-or this blog post by Jeremy Kun \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-of-the-}}\\
-\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{multiplicative-weights-update-algorithm}},
+% on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
+this blog post by Jeremy Kun, \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-}}\\
+\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{effectiveness-of-the-multiplicative-weights-update-algorithm}},
 which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019).
@@
 % ----------------------------------------------------------------------------
-\paragraph{Hybrid policies}
-
+\textbf{Hybrid policies.}
+%
 As explained before, in the last few years, the MAB research community has been interested in finding algorithms which achieves both a problem-dependent logarithmic and a minimax upper-bounds.
@@
 %
-In the rest of this thesis, we mainly use the \UCB{} or Thompson sampling algorithms when simplicity is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when optimality is important and where we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
+In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when optimality is important and where we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
 
@@
 Many different learning algorithms have been proposed by the machine learning community,
-and most of them depend on several parameters, for instance $\alpha$ for \UCB, the prior for Thompson sampling or BayesUCB,
+and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
 the $\kl$ function for \klUCB{} etc.
@@
 On the one hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
-It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the \UCB{} algorithm (or other low-cost algorithms).
+It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the $\UCB_1$ algorithm (or other low-cost algorithms).
 %
@@
 %
-For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of a \UCB{} algorithms will perform quite differently in the two problems:
+For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of a $\UCB_1$ algorithms will perform quite differently in the two problems:
 using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case,
 while using a low $\alpha$, \ie, favoring exploitation, will give low regret in the second case.
-One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of \UCB, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
+One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
 

### 2019-06-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Better discussion on the two notations of fairness

--- 2019-06-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
@@
 \end{equation}
+
 The additional assumption in \eqref{eq:5:SUE} further implies some notion of \emph{fairness}, as it suggests that each of the $M$ players spends on average the same amount of time on each of the $M$ best arms.
@@
 Note that all our proposed algorithms are permutation invariant and \MCTopM{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.
+
+As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
+In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
+We note that the notion of fairness we defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
+For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is given and known to the algorithm.
 

### 2019-06-04 2-Chapters/2-Chapter/chapter2.tex
Oups je fais de la merde sur la preuve de UCB

@@
 so if $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, \ie, $R_T = \Omega(T)$.
+%
+To prove a bound on $p$, we use the following lemma, well known as Hoeffding's inequality, from \cite{hoeffding1963probability}.
+
+\begin{lemma}[Hoeffding's inequality]\label{lem:2:HoeffdingInequality}
+    Let $Z_1,\dots,Z_n$ be $n$ \iid{} samples from a Bernoulli distribution of parameter $\mu$, of empirical mean $\widehat{Z_n} = \frac{1}{n} \sum\limits_{i=1}^n Z_i$, Hoeffding's inequality gives both
+    \begin{align}\label{eq:2:HoeffdingInequality}
+        \forall x < \mu, \;\;\; & \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
+        \\
+        \forall x > \mu, \;\;\; & \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+    \end{align}
+\end{lemma}
 
@@
     We prove it for the case of two arms.
-    If the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality from \cite{hoeffding1963probability} to control this probability $p$.
-    %
-    Indeed if $Z_1,\dots,Z_n$ are \iid{} samples from a Bernoulli distribution of parameter $\mu$, of mean $\widehat{Z_n} = \frac{1}{n} \sum_{i=1}^n Z_i$, Hoeffding's inequality gives both
-    \begin{equation}
-        \forall x < \mu, \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
-        \;\;\;
-        \forall x > \mu, \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
-    \end{equation}
+    If the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality to control this probability $p$.
     %
@@
 %
-The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that include well known and efficient algorithms such as $\UCB_1$, \klUCB{} and many others.
+The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that includes well known and efficient algorithms such as $\UCB_1$, \klUCB{} and many others.
 
@@
 
-\paragraph{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}
-
+\textbf{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}
+%
 Let us consider another approach for adaptive exploration, known as ``optimism under uncertainty'': assume each arm is as good as it can possibly be given the observations so far, and choose the best arm based on these optimistic estimate.
@@
 %
-Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{}  respectively represent exploitation and exploration, and summing them up is a natural way of trading them off.
-The parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to choose $\alpha=1/2$ for optimal uniform performance.
-
-To analyze the algorithm $\UCB_1$, we denote again $k^*$ an optimal arm, and $A(t)$ is the arm chosen by the algorithm in round $t$.
-According to the algorithm, $\UCB_k(t) \geq \UCB_{k^*}(t)$.
-%
-Under the clean event,μ(at) +rt(at)≥ ̄μt(at)andUCBt(a∗)≥μ(a∗). Therefore:
-XXX
-It follows that
-XXX
-\TODOL{Expliquer la fin}
-
+Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{}  respectively represent \emph{exploitation} and \emph{exploration}, and summing them up is a natural way of dealing with the exploitation/exploration trade off.
+The constant parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to choose $\alpha=1/2$ for optimal uniform performance (\ie, on all problems).
+
+\begin{theorem}[Regret bound for $\UCB_1$]\label{thm:2:UCBregretBound}
+    For any instance $I$ with $K$ arms with bounded distributions in $[0,1]$,
+    of means $\mu_1,\dots,\mu_k$, and any horizon $T>10$,
+    the $\UCB_1$ algorithm verifies the following asymptotic regret bound
+    \begin{equation}\label{eq:2:UCBregretBound}
+        R_T^{\cA}(I) \leq \sum_{k: \Delta_k>0} \frac{8}{\Delta_k} \log(T) + \smallO{\log(T)}.
+    \end{equation}
+\end{theorem}
+%
+\begin{proof}
+    For simplicity, we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality} as stated above in Lemma~\ref\label{lem:2:HoeffdingInequality}.
+    To analyze the algorithm $\UCB_1$, we restrict to $\alpha=2$ and we proceed as follows:
+
+    $i)$ First, for any arm $k$, Hoeffding's inequality shows that $\Pr(\UCB_k(t) \leq \mu_k) \leq \bigO{1/t^{4}}$ independently of $N_k(t)$,
+    and so $\sum_{t=1}^T \Pr(\UCB_k(t) \leq \mu_k) = \smallO{\log(T)}$: the \UCB{} are almost always above the empirical means (\ie, they are true \emph{upper} confidence bounds).
+
+    $ii)$
+    Then, we denote again $k^*$ an optimal arm, and $A(t)$ is the arm chosen in round $t$.
+    According to the algorithm, the probability that $k$ is chosen is upper-bounded by $\Pr(\UCB_k(t) \leq \UCB_{k^*}(t))$.
+    This can happen if $\UCB_{k^*}(t) \leq \mu_{k^*}$, which from $i)$ is very unlikely, or if $\UCB_k(t) \leq \UCB_{k^*}(t) > \mu_{k^*}$.
+    By writing $\mu_k = \mu_{k^*} - \Delta_k$, and because $\UCB_k(t)$ are using $N_k(t)$ \iid{} samples from arm $k$, of mean $\mu_k$, Hoeffding's inequality also shows that $\Pr(\UCB_k(t) - \mu_k > \Delta_k) \leq 8/(\Delta_k^2 t)$ (the confidence radius $\xi_k(t)$ is actually built exactly to obtain this bound).
+    %
+    And thus\footnote{~We used the partial sum of the harmonic series: $\sum\limits_{t=1}^T \frac{1}{t} \underset{T\to\infty}{\sim} \log(T)$.} $\sum_{t=1}^T \Pr(\UCB_k(t) - \mu_k > \Delta_k) \leq 1/\Delta_k^2 \sum_{t=1}^T 8/t = 1/\Delta_k^2 8\log(T)$.
+
+    To sum up, we obtain the following bound on $N_k(T)$ for any sub-optimal arm $k$:
+    \begin{align*}
+        \E[N_k(T)] = \sum_{t=1}^T \Pr(\ind(A(t)=k)) &\leq \sum_{t=1}^T \Pr(\UCB_k(t) > \UCB_{k^*}(t)) \\
+        &\leq \sum_{t=1}^T \Pr(\UCB_{k^*}(t) < \mu_{k^*}) + \sum_{t=1}^T \Pr(\UCB_{k^*}(t) < \mu_{k^*}) \\
+        &\leq \smallO{\log(T)} + \frac{8}{\Delta_k^2} \log(T).
+    \end{align*}
+    %
+    By using the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, this gives $R_T^{\UCB_1} \leq \sum_{k: \Delta_k>0} \frac{8}{\Delta_k} \log(T) + \smallO{\log(T)}$.
+    We preferred to use the $\smallO{\dots}$ Landau notation, instead of keeping the exact constants, for simplicity and clarity.
     More details on the proofs can be found for instance in Chapter~2 of \cite{Bubeck12}, or Section~1.3 of \cite{Slivkins2019}.
-The $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
-
-It has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects.
+\end{proof}
+
+What this theorem shows is that the $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+The $\UCB_1$ algorithm has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects.
 The $\UCB_1$ algorithm requires a constant storage capacity for each arm (storing $X_k(t)$ and $N_k(t)$ require two integers or a \texttt{float} and an integer, \ie, $\bigO{1}$), giving a storage capacity of $\bigO{K}$, independently of the horizon $T$.

### 2019-06-04 2-Chapters/2-Chapter/chapter2.tex
Almost done for UCB proof, I'll then be VERY quick for klUCB

@@
 Finally, another interesting measure of performance is the \emph{fairness}.
-It was not studied much for single-player problems, and usually it means that each arm must be explored at least a given fraction of the total number of rounds $T$ (\ie, ``arm-wise fairness'').
+It was not studied much for single-player problems, and usually it means that each arm must be explored at least a given fraction of the total horizon $T$ (\ie, ``arm-wise fairness'').
 It was studied in a very recent article \cite{Patil2019stochastic},
@@
 % ----------------------------------------------------------------------------
-\subsection{Index policies, $\UCB_1$, \klUCB{} and others}
-
-A large family of algorithms are index policies, that compute an index $U_k(t)$ on each arm $k$ at time $t$.
+\subsection{Index policies: $\UCB_1$, \klUCB{} and others}
+
+A large family of algorithms are index based, as they compute an index $U_k(t)$ on each arm $k$ at time $t$.
 The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
@@
 
-\textbf{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}
+\textbf{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}\\
 %
@@
 Either reason makes this arm worth choosing.
-One can also observe that $\xi_k(t)\to\infty$ if $t\to\infty$ but $N_k(t)\ll\log(t)$, thus all arms must be explored at least $\mathcal{O}(\log(T))$ times.
+One can also observe that $\xi_k(t)\to\infty$ if $t\to\infty$ while $N_k(t)\ll\log(t)$, which can be used to prove that the $\UCB_1$ algorithm tries all arms at least $\mathcal{O}(\log(T))$ times (in average).
 %
@@
 
+We present the regret bound, which was first stated in \cite{Auer02}, and to explain how Hoeffding's inequality is used we prove it below.
+
 \begin{theorem}[Regret bound for $\UCB_1$]\label{thm:2:UCBregretBound}
+    For simplicity, we restrict to $\alpha=2$.
     For any instance $I$ with $K$ arms with bounded distributions in $[0,1]$,
     of means $\mu_1,\dots,\mu_k$, and any horizon $T>10$,
-    the $\UCB_1$ algorithm verifies the following asymptotic regret bound
+    the $\UCB_1$ algorithm verifies the following finite-time regret bound
     \begin{equation}\label{eq:2:UCBregretBound}
-        R_T^{\cA}(I) \leq \sum_{k: \Delta_k>0} \frac{8}{\Delta_k} \log(T) + \smallO{\log(T)}.
+        R_T^{\cA}(I) \leq \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left(1 + \frac{\pi^2}{3}\right)
+        = \bigO{ \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T)}.
     \end{equation}
@@
 \begin{proof}
-    For simplicity, we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality} as stated above in Lemma~\ref\label{lem:2:HoeffdingInequality}.
-    To analyze the algorithm $\UCB_1$, we restrict to $\alpha=2$ and we proceed as follows:
-
-    $i)$ First, for any arm $k$, Hoeffding's inequality shows that $\Pr(\UCB_k(t) \leq \mu_k) \leq \bigO{1/t^{4}}$ independently of $N_k(t)$,
-    and so $\sum_{t=1}^T \Pr(\UCB_k(t) \leq \mu_k) = \smallO{\log(T)}$: the \UCB{} are almost always above the empirical means (\ie, they are true \emph{upper} confidence bounds).
-
-    $ii)$
-    Then, we denote again $k^*$ an optimal arm, and $A(t)$ is the arm chosen in round $t$.
-    According to the algorithm, the probability that $k$ is chosen is upper-bounded by $\Pr(\UCB_k(t) \leq \UCB_{k^*}(t))$.
-    This can happen if $\UCB_{k^*}(t) \leq \mu_{k^*}$, which from $i)$ is very unlikely, or if $\UCB_k(t) \leq \UCB_{k^*}(t) > \mu_{k^*}$.
-    By writing $\mu_k = \mu_{k^*} - \Delta_k$, and because $\UCB_k(t)$ are using $N_k(t)$ \iid{} samples from arm $k$, of mean $\mu_k$, Hoeffding's inequality also shows that $\Pr(\UCB_k(t) - \mu_k > \Delta_k) \leq 8/(\Delta_k^2 t)$ (the confidence radius $\xi_k(t)$ is actually built exactly to obtain this bound).
-    %
-    And thus\footnote{~We used the partial sum of the harmonic series: $\sum\limits_{t=1}^T \frac{1}{t} \underset{T\to\infty}{\sim} \log(T)$.} $\sum_{t=1}^T \Pr(\UCB_k(t) - \mu_k > \Delta_k) \leq 1/\Delta_k^2 \sum_{t=1}^T 8/t = 1/\Delta_k^2 8\log(T)$.
-
-    To sum up, we obtain the following bound on $N_k(T)$ for any sub-optimal arm $k$:
+    As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref\label{lem:2:HoeffdingInequality}.
+    Thanks to the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.
+
+    The first trick consists in writing the following, which is true for any $n \geq 1$:
+    \begin{equation}\label{eq:2:trickInUCBProof}
+        N_k(T) = 1 + \sum_{t=K+1}^T \indic(A(t) = k)
+        \leq n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n).
+    \end{equation}
+
+    Remember that $\UCB_a(t) = \widehat{\mu_a}(t) + \xi_a(t)$ for any arm $a$.
+    According to the algorithm (see Algorithm~\ref{algo:2:indexPolicy} at line~6),
+    arm $k$ is chosen if $\UCB_k(t) \geq \UCB_{k^*}(t)$.
+    Both $\widehat{\mu_a}(t)$ and $\xi_a(t)$ use $N_k(t)$ samples of arm $a$,
+    but as \eqref{eq:2:trickInUCBProof} above started to isolate $N_k(t)$, we want to add the freedom of considering different number of samples (using $n$), so we introduce the notation
+    $\widehat{\mu_{a,n}}(t)$ and $\xi_{a,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
+    \ie, $\widehat{\mu_{a,n}}(t) = \frac{1}{n} \sum_{s=1}^{t-1} r_k(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
+    %
+    Thus, we can continue from \eqref{eq:2:trickInUCBProof} and write
+    \begin{align}\label{eq:2:decompositionInUCBProof}
+        N_k(T) & \leq n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n) \nonumber \\
+        & \leq n + \sum_{t=n}^T \indic(\widehat{\mu_{k^*,N_{k^*}(t)}}(t) + \xi_{k^*,N_{k^*}(t)}(t) < \widehat{\mu_{k,N_k(t)}}(t) + \xi_{k,N_k(t)}(t), N_k(t) \geq n)  \nonumber \\
+        & \leq n + \sum_{t=n}^T \indic(\min_{0 < n_{k^*} < t} \widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \max_{n < n_k < t} \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t)) \nonumber \\
+        \intertext{So two union bounds on $n_{k^*}$ and $n_k$ give}
+        & \leq n + \sum_{t=n}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n}^t  \indic( \widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t) )
+    \end{align}
+
+    Now, at time $t$, this event $\widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t)$
+    implies at least one of the following three events:
     \begin{align*}
-        \E[N_k(T)] = \sum_{t=1}^T \Pr(\ind(A(t)=k)) &\leq \sum_{t=1}^T \Pr(\UCB_k(t) > \UCB_{k^*}(t)) \\
-        &\leq \sum_{t=1}^T \Pr(\UCB_{k^*}(t) < \mu_{k^*}) + \sum_{t=1}^T \Pr(\UCB_{k^*}(t) < \mu_{k^*}) \\
-        &\leq \smallO{\log(T)} + \frac{8}{\Delta_k^2} \log(T).
+        I_1: & \widehat{\mu_{k^*,n_{k^*}}}(t) \leq \mu_{k^*} - \xi_{k^*,n_{k^*}}(t),
+        I_2: & \widehat{\mu_{k,n_k}}(t) \geq \mu_k + \xi_{k,n_k}(t),
+        I_3: & \mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t).
     \end{align*}
-    %
-    By using the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, this gives $R_T^{\UCB_1} \leq \sum_{k: \Delta_k>0} \frac{8}{\Delta_k} \log(T) + \smallO{\log(T)}$.
-    We preferred to use the $\smallO{\dots}$ Landau notation, instead of keeping the exact constants, for simplicity and clarity.
-    More details on the proofs can be found for instance in Chapter~2 of \cite{Bubeck12}, or Section~1.3 of \cite{Slivkins2019}.
+
+    For the first two events $I_1$ and $I_2$, we carefully fixed the number of samples so they are no longer random (they are respectively to $n_{k^*}$ and $n_k$), we can use Hoeffding's inequality to obtain $\Pr(I_j) \leq \frac{1}{t^{4}}$ for $j=1,2$.
+
+    For the last event $I_3$,
+    observe that $\mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t)$ means $\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0$, where we remind the notation of the gap $\Delta_k = \mu_{k^*} - \mu_k$.
+    Thus if $n_k \geq \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil \geq \lceil \frac{8 \log(t)}{\Delta_k^2} \rceil$,
+    then this inequality actually cannot happen:
+    $\Pr(\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0) = 0$,
+    so as soon as $n_k \geq \frac{8 \log(T)}{\Delta_k^2}$, $\Pr(I_3) = 0$.
+
+    Thus we can take the expectation of $N_k(T)$ from equation~\eqref{eq:2:decompositionInUCBProof}, and continue by setting
+    with $n = n_T = \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil$:
+
+    \begin{align}\label{eq:2:decompositionInUCBProof2}
+        \E[ N_k(T) ]
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr( \widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t) ) \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr( I_1 ) + \Pr( I_2 ) + \Pr( I_3 ) \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \nonumber \\
+        \intertext{We actually have that $\sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \leq 2 \sum_{n_{k^*}=1}^t \sum_{n_{k}=1}^t \frac{1}{t^4} \leq 2 \sum_{n_{k^*}=1}^t \frac{1}{t^3} \leq 2 \frac{1}{t^2}$}
+        & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \sum_{t=1}^T \frac{1}{t^2} \nonumber \\
+        & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \sum_{t=1}^{\infty} \frac{1}{t^2} \nonumber \\
+        \E[ N_k(T) ]
+        & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \frac{\pi^2}{6}.
+    \end{align}
+
+    To sum up, we obtain the following finite-time bound on $N_k(T)$ for any sub-optimal arm $k$:
+    \begin{align*}
+        \E[N_k(T)] \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + \frac{\pi^2}{3}.
+    \end{align*}
+    %
+    By using the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, this gives
+    \begin{align*}
+        R_T^{\UCB_1} &\leq \sum_{k: \Delta_k>0} \Delta_k \E[N_k(T)] \\
+        & \leq \frac{8}{\Delta_k} \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left( 1 + \frac{\pi^2}{3} \right)
+        = \frac{8}{\Delta_k} \log(T) + \smallO{\log(T)}.
+    \end{align*}
+
+    More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs (\eg, for any $\alpha\geq\frac{1}{2}$).
 \end{proof}
 
-What this theorem shows is that the $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
+What this theorem shows is that the $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret (if $\Delta=\min\limits_{k: \Delta_k>0} \Delta$).
 The $\UCB_1$ algorithm has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects.
-The $\UCB_1$ algorithm requires a constant storage capacity for each arm (storing $X_k(t)$ and $N_k(t)$ require two integers or a \texttt{float} and an integer, \ie, $\bigO{1}$), giving a storage capacity of $\bigO{K}$, independently of the horizon $T$.
-It also needs to perform some mathematical operations, but a square-root and a logarithm are cheap, thus it costs a $\bigO{1}$ constant time for each arm and time step, thus a time $\bigO{K T}$ for the $T$ rounds.
+It requires a constant storage capacity for each arm (storing $X_k(t)$ and $N_k(t)$ require two integers or a \texttt{float} and an integer, \ie, $\bigO{1}$), giving a storage capacity of $\bigO{K}$, independently of the horizon $T$.
+It also needs to perform some mathematical operations, but a square-root and a logarithm are cheap, thus it costs a $\bigO{1}$ constant time for each arm and time step, thus a time $\bigO{K T}$ for the $T$ time steps.
 %

### 2019-06-04 2-Chapters/2-Chapter/chapter2.tex
Moved proof of Pinkser's inequality to Chapter 2

@@
 We present the regret bound, which was first stated in \cite{Auer02}, and to explain how Hoeffding's inequality is used we prove it below.
+Note that for the rest of the manuscript, we use the \klUCB{} algorithm (and more subtle concentration inequality) whenever we analyze the regret of a newly introduced algorithm (\ie, in Chapters~\ref{chapter:5} and \ref{chapter:6}),
+but we preferred to start with a simpler proof in the single-player stationary case.
+µ
+What this theorem shows is that the $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret (if $\Delta=\min\limits_{k: \Delta_k>0} \Delta$).
 
@@
     For any instance $I$ with $K$ arms with bounded distributions in $[0,1]$,
-    of means $\mu_1,\dots,\mu_k$, and any horizon $T>10$,
+    of means $\mu_1,\dots,\mu_k$, and any horizon $T>K$,
     the $\UCB_1$ algorithm verifies the following finite-time regret bound
@@
     As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref\label{lem:2:HoeffdingInequality}.
-    Thanks to the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.
-
+    Thanks to the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
+    %
+    \indent
     The first trick consists in writing the following, which is true for any $n \geq 1$:
@@
     but as \eqref{eq:2:trickInUCBProof} above started to isolate $N_k(t)$, we want to add the freedom of considering different number of samples (using $n$), so we introduce the notation
-    $\widehat{\mu_{a,n}}(t)$ and $\xi_{a,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
-    \ie, $\widehat{\mu_{a,n}}(t) = \frac{1}{n} \sum_{s=1}^{t-1} r_k(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
+    $\hat{\mu}_{a,n}(t)$ and $\xi_{a,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
+    \ie, $\hat{\mu}_{a,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r_k(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
     %
@@
         N_k(T) & \leq n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n) \nonumber \\
-        & \leq n + \sum_{t=n}^T \indic(\widehat{\mu_{k^*,N_{k^*}(t)}}(t) + \xi_{k^*,N_{k^*}(t)}(t) < \widehat{\mu_{k,N_k(t)}}(t) + \xi_{k,N_k(t)}(t), N_k(t) \geq n)  \nonumber \\
-        & \leq n + \sum_{t=n}^T \indic(\min_{0 < n_{k^*} < t} \widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \max_{n < n_k < t} \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t)) \nonumber \\
-        \intertext{So two union bounds on $n_{k^*}$ and $n_k$ give}
-        & \leq n + \sum_{t=n}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n}^t  \indic( \widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t) )
+        & \leq n + \sum_{t=n}^T \indic(\hat{\mu}_{k^*,N_{k^*}(t)}(t) + \xi_{k^*,N_{k^*}(t)}(t) < \hat{\mu}_{k,N_k(t)}(t) + \xi_{k,N_k(t)}(t), N_k(t) \geq n)  \nonumber \\
+        & = n + \sum_{t=n}^T \indic\left(\min_{0 < n_{k^*} < t} \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \max_{n < n_k < t} \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right) \nonumber \\
+        \intertext{So two union bounds on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$ give}
+        & \leq n + \sum_{t=n}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n}^t  \indic\left( \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right)
     \end{align}
 
-    Now, at time $t$, this event $\widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t)$
+    Now, at time $t$, this event $\hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t)$
     implies at least one of the following three events:
     \begin{align*}
-        I_1: & \widehat{\mu_{k^*,n_{k^*}}}(t) \leq \mu_{k^*} - \xi_{k^*,n_{k^*}}(t),
-        I_2: & \widehat{\mu_{k,n_k}}(t) \geq \mu_k + \xi_{k,n_k}(t),
-        I_3: & \mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t).
+        \begin{cases}
+        I_1 \;:\; & \hat{\mu}_{k^*,n_{k^*}}(t) \leq \mu_{k^*} - \xi_{k^*,n_{k^*}}(t), \\
+        I_2 \;:\; & \hat{\mu}_{k,n_k}(t) \geq \mu_k + \xi_{k,n_k}(t), \\
+        I_3 \;:\; & \mu_{k^*} < \mu_k + 2 \; \xi_{k,n_k}(t).
+        \end{cases}
     \end{align*}
 
-    For the first two events $I_1$ and $I_2$, we carefully fixed the number of samples so they are no longer random (they are respectively to $n_{k^*}$ and $n_k$), we can use Hoeffding's inequality to obtain $\Pr(I_j) \leq \frac{1}{t^{4}}$ for $j=1,2$.
-
+    \begin{itemize}
+        \item
+        For the first two events $I_1$ and $I_2$, we carefully fixed the number of samples so they are no longer random (they are respectively to $n_{k^*}$ and $n_k$), we can use Hoeffding's inequality to obtain $\Pr(I_j) \leq 1 / t^{4}$ (for $j=1,2$).
+
+        \item
         For the last event $I_3$,
         observe that $\mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t)$ means $\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0$, where we remind the notation of the gap $\Delta_k = \mu_{k^*} - \mu_k$.
-    Thus if $n_k \geq \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil \geq \lceil \frac{8 \log(t)}{\Delta_k^2} \rceil$,
+        Thus if $n_k \geq \left\lceil \frac{8 \log(T)}{\Delta_k^2} \right\rceil \geq \left\lceil \frac{8 \log(t)}{\Delta_k^2} \right\rceil$,
         then this inequality actually cannot happen:
@@
         so as soon as $n_k \geq \frac{8 \log(T)}{\Delta_k^2}$, $\Pr(I_3) = 0$.
-
-    Thus we can take the expectation of $N_k(T)$ from equation~\eqref{eq:2:decompositionInUCBProof}, and continue by setting
-    with $n = n_T = \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil$:
-
+    \end{itemize}
+    %
+    Taking the expectation of $N_k(T)$ from \eqref{eq:2:decompositionInUCBProof},
+    and setting $n = n_T = \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil$ thus gives
+    %
     \begin{align}\label{eq:2:decompositionInUCBProof2}
         \E[ N_k(T) ]
-        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr( \widehat{\mu_{k^*,n_{k^*}}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu_{k,n_k}}(t) + \xi_{k,n_k}(t) ) \nonumber \\
-        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr( I_1 ) + \Pr( I_2 ) + \Pr( I_3 ) \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr( \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) ) \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \underbrace{\Pr(I_1)}_{=1/t^4} + \underbrace{\Pr(I_2)}_{=1/t^4} + \underbrace{\Pr(I_3)}_{=0} \nonumber \\
         & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \nonumber \\
-        \intertext{We actually have that $\sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \leq 2 \sum_{n_{k^*}=1}^t \sum_{n_{k}=1}^t \frac{1}{t^4} \leq 2 \sum_{n_{k^*}=1}^t \frac{1}{t^3} \leq 2 \frac{1}{t^2}$}
-        & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \sum_{t=1}^T \frac{1}{t^2} \nonumber \\
-        & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \sum_{t=1}^{\infty} \frac{1}{t^2} \nonumber \\
-        \E[ N_k(T) ]
-        & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \frac{\pi^2}{6}.
+        \intertext{We actually have that $\sum\limits_{n_{k^*}=1}^t \sum\limits_{n_{k}=n_T}^t  2 \frac{1}{t^4} \leq 2 \sum\limits_{n_{k^*}=1}^t \sum\limits_{n_{k}=1}^t \frac{1}{t^4} \leq 2 \sum\limits_{n_{k^*}=1}^t \frac{1}{t^3} \leq 2 \frac{1}{t^2}$}
+        & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \sum_{t=1}^T \frac{1}{t^2}
+        \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \sum_{t=1}^{\infty} \frac{1}{t^2}
+        \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + \frac{\pi^2}{3}.
     \end{align}
-
-    To sum up, we obtain the following finite-time bound on $N_k(T)$ for any sub-optimal arm $k$:
+    %
+    To sum up, we obtain the following finite-time bound on $N_k(T)$ for any sub-optimal arm $k$
     \begin{align*}
-        \E[N_k(T)] \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + \frac{\pi^2}{3}.
+        \E[N_k(T)] \leq \frac{8 \log(T)}{\Delta_k^2} + (1 + \frac{\pi^2}{3}).
     \end{align*}
@@
     \begin{align*}
-        R_T^{\UCB_1} &\leq \sum_{k: \Delta_k>0} \Delta_k \E[N_k(T)] \\
-        & \leq \frac{8}{\Delta_k} \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left( 1 + \frac{\pi^2}{3} \right)
-        = \frac{8}{\Delta_k} \log(T) + \smallO{\log(T)}.
+        R_T^{\UCB_1} &= \sum_{k: \Delta_k>0} \Delta_k \E[N_k(T)] \\
+        & \leq \sum_{k: \Delta_k>0} \frac{8}{\Delta_k} \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left( 1 + \frac{\pi^2}{3} \right)
+        = \sum_{k: \Delta_k>0} \frac{8}{\Delta_k} \log(T) + \smallO{\log(T)}.
     \end{align*}
@@
 
-What this theorem shows is that the $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret (if $\Delta=\min\limits_{k: \Delta_k>0} \Delta$).
-The $\UCB_1$ algorithm has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects.
-It requires a constant storage capacity for each arm (storing $X_k(t)$ and $N_k(t)$ require two integers or a \texttt{float} and an integer, \ie, $\bigO{1}$), giving a storage capacity of $\bigO{K}$, independently of the horizon $T$.
+Therefore, we see that the $\UCB_1$ algorithm has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO(K\log(T)/\Delta^2)$.
+%
+$\UCB_1$ requires a constant storage capacity for each arm (storing $X_k(t)$ and $N_k(t)$ require two integers or a \texttt{float} and an integer, \ie, $\bigO{1}$), giving a storage capacity of $\bigO{K}$, independently of the horizon $T$.
 It also needs to perform some mathematical operations, but a square-root and a logarithm are cheap, thus it costs a $\bigO{1}$ constant time for each arm and time step, thus a time $\bigO{K T}$ for the $T$ time steps.
 %
-We use it as the base building block for other policies in Chapter~\ref{chapter:4}, for these reasons.
+We use $\UCB_1$ as the base building block for other policies in Chapter~\ref{chapter:4}, for these reasons.
 Moreover, the $\UCB_1$ algorithm has the advantage of being easy to explain and understand, and its decisions can be interpreted clearly: by reading the values of $\widehat{\mu_k}(t)$ and $\UCB_k(t)$, anyone can see why an arm was chosen at anytime.
@@
 
-\paragraph{The \klUCB{} index policy: using optimal confidence intervals.}
-
+\paragraph{The \klUCB{} index policy.}
+%
 \TODOL{Expliquer klUCB}
-\TODOL{Expliquer que les indices klUCB = des UCB sur les moyennes + principle of optimism in face of uncertainty}
-
-The \klUCB{} algorithm XXX
-Optimal upper confidence bounds based on Kullback-Leibler divergence and klUCB indexes
-
-% If rewards lie in $[a,b]$, then t
+
+The \klUCB{} algorithm follows the same spirit as the $\UCB_1$ algorithm presented above,
+it is an index policy (see Algorithm~\ref{algo:2:indexPolicy}),
+which also builds upper confidence bounds on the empirical samples.
+The KL-\UCB{} algorithm was first proposed in \cite{Garivier11KL} for bounded rewards,
+and then the \klUCB{} algorithm was introduced in \cite{KLUCBJournal}.
+In all this thesis, we either restrict to exponential families or to bounded rewards in $[0,1]$ and use the fact that they are sub-Bernoulli, thus we prefer to only present \klUCB.
+
+Instead of using a confidence radius $\xi_(t)$ based on Hoeffding's inequality,
+the \klUCB{} algorithm rather use a specific concentration inequality for the considered exponential family, such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
+It yields improved upper confidence bounds (\ie, with smaller $\xi_(t)$) that are based on Kullback-Leibler divergence.
 The indexes are computed as follows, for an exploration function $f(t)$
@@
 
+For Bernoulli distributions,
+we can easily justify the previous claim that \klUCB{} builds smaller confidence intervals (\ie, more accurate) than $\UCB_1$,
+\ie, $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$,
+because Pinkser's inequality gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix),
+and so XXX
+
 \paragraph{A note about the implementation of \klUCB{} algorithms.}
 %
-% \TODOL{Short note on how to implement such policy, using Newton's method ?}
 Implementing this kind of policy requires to implement two things.
@@
 % % ----------------------------------------------------------------------------
-% \section{Appendix}
-% \label{sec:2:appendix}
+\section{Appendix}
+\label{sec:2:appendix}
+
+% ----------------------------------------------------------------------------
+\subsubsection{Pinsker's inequality}\label{proof:2:Pinsker}
+
+Pinsker's inequality is proved below, using simple real analysis arguments\footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).}.
+%
+This Lemma is again used in Chapter~\ref{chapter:6}.
+
+\begin{lemma}[Pinsker's inequality]\label{lem:2:Pinsker}
+    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as
+    \begin{equation}\label{eq:2:Pinsker1}
+        \kl(p,q) = p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right).
+    \end{equation}
+    Then for any $p,q\in[0,1]$, we have
+    \begin{equation}\label{eq:2:Pinsker2}
+        \kl(p,q) \geq 2 (p-q)^2.
+    \end{equation}
+\end{lemma}
+
+\begin{proof}
+    Fix $p\in[0,1]$, and consider the function $g(x) := \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
+    We first observe that $g(0) = 0$, and we easily prove that $x=0$ is the unique minimizer of $g$ over its domain.
+    By definition, $\kl$ is of class $\cC^1$ on its two variables, on $(0,1)$, so $g$ is of class $\cC^1$ on the interior of its domain, $(-p, 1-p)$.
+    Let us compute its derivative, $g'(x) = (\frac{\partial \; \kl}{\partial 2})(p, p + x) - 4x$.
+    Simple calculations yield
+    \[ \left(\frac{\partial \; \kl}{\partial 2}\right) (p, p + x) = - \frac{p}{p+x} + \frac{1-p}{1-(p+x)} =  = -x \frac{(2(p + x) - 1)^2}{(p + x)(p + x - 1)}. \]
+    % https://www.wolframalpha.com/input/?i=diff(p+*+log(p%2F(p%2Bx))+%2B+(1-p)+*+log((1-p)%2F(1-(p%2Bx)))+-+2*x**2,+x)
+    We can verify that $g'(0) = -1 + 1 - 4 \times 0 = 0$, and it has the sign of $-x$ on its domain, as $p + x - 1 < 0$ because $x < 1 - p$, and $p + x > 0$.
+    So $g'$ is first positive then negative, so $g$ is increasing then decreasing on its domain, with a change in its monotony at $x=0$, and thus it indeed has a unique minimizer.
+    We conclude that $g(x) \geq g(0) = 0$ on $(-p, 1-p)$, and so $\kl(p,q) \geq 2 (p-q)^2$ if $x = q-p$.
+
+    The edge case is the ends of the interval, that is if $x=-p$ or $x=1-p$, that is if $q=0$ or $q=1$.
+    If $p\in(0,1)$, then computing the limit of $\kl(p,q)$ for $q\to0^+$ and $q\to1^-$ is obvious and gives $\kl(p,q) \to +\infty$, while the right-hand side of Pinsker's inequality stays finite (so the inequality is still true at the limit).
+    If $p=0$ or $p=1$, the same computation works, by the convention that $p \log(p) = 0$ if $p=0$.
+    %
+    % Bonus: one can verify our computation in Python with the SymPy module:
+    % \begin{verbatim}
+    %     $ (bash) python
+    %     >>> from sympy import *
+    %     >>> x, p = symbols('x p')
+    %     >>> g = p * log(p/(p+x)) + (1-p) * log((1-p)/(1-(p+x))) - 2*x**2
+    %     >>> print(diff(g, x))
+    %     -p/(p + x) - 4*x + (-p + 1)/(-p - x + 1)
+    % \end{verbatim}
+\end{proof}
 

### 2019-06-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Moved proof of Pinkser's inequality to Chapter 2

@@
 For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to the stopping time \eqref{def:6:GLR} with the choice $d(x,y) = 2(x-y)^2$, when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
-Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$ (see Lemma~\ref{lem:6:Pinsker}), hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
+Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$ (see Lemma~\ref{lem:2:Pinsker}), hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
 %
@@
 Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from $\mu_0$ to $\mu_1$ occurs at time $\tau$.
-We already observed that from Pinsker's inequality (Lemma~\ref{lem:6:Pinsker}), the B-GLRT stops earlier than a Gaussian GLR.
+We already observed that from Pinsker's inequality (Lemma~\ref{lem:2:Pinsker}), the B-GLRT stops earlier than a Gaussian GLR.
 Hence, one can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the B-GLRT.
@@
 
-Using Pinsker's inequality from Lemma~\ref{lem:6:Pinsker} and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = |\Delta_i^k|$), one can write
+Using Pinsker's inequality from Lemma~\ref{lem:2:Pinsker} and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = |\Delta_i^k|$), one can write
 %

### 2019-06-04 2-Chapters/2-Chapter/chapter2.tex
Lots of small change haha

@@
 
-\TODOL{Abstract ici et plus court, et petite introduction !}
+\TODOL{Abstract ici et plus court, et petite introduction ?}
 
@@
 \newpage
-
-% \TODOL{L'abstract ici est trop long, mais je l'ai écrit avant de commencer le chapitre, pour avoir une idée de ce que je devais faire ! Je couperai à la fin !}
 
@@
 %
-The Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on the article \cite{Besson2018WCNC}.
+The first Sections~\ref{sec:2:notations}, \ref{sec:2:applicationsofStochasticMAB}, \ref{sec:2:lowerUpperBoundsRegret} and \ref{sec:2:famousMABalgorithms} of this chapter are based on existing works,
+and Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on our article \cite{Besson2018WCNC}.
 
@@
 We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
-For more details on reinforcement learning, in particular about more generic models, we refer to the famous book \cite{SuttonBarto2018}.
+For more details on reinforcement learning, in particular about more generic models, we refer to the famous book by Sutton and Barto \cite{SuttonBarto2018}.
 
@@
             % Once the nodes are placed, connecting them is easy.
-            \draw [->] (player) to[bend left=90] node[pos=0.5] {Chooses a discrete action $k=A(t) \in\{1,\dots,K\}$} (environment);
-            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Observes a real-valued reward $r_k(t) \in \R$} (player);
+            \draw [->] (player) to[bend left=90] node[pos=0.5] {Chooses a \textbf{discrete} action $k=A(t) \in\{1,\dots,K\}$} (environment);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Observes a \textbf{real-valued} reward $r_k(t) \in \R$} (player);
             %
@@
 
-This quantity $\sum_{t=1}^T r_{A(t)}(t)$ can indeed be random, and usually depends on two aspects.
+This quantity $\sum_{t=1}^T r_{A(t)}(t)$ can indeed be random, and it usually depends on two aspects.
 First, the rewards may depend on the randomness of the environment, \ie, the unknown and unpredictable values $r_k(t)$.
@@
 
-A bandit algorithm $\cA$ is also referred to as a strategy or a policy, and sometimes it is denoted by $\pi$ or $\rho$ in the literature.
-The algorithm $\cA$ selects an arm $A(t)$ at time $t$ by using the past observations and the past external randomness.
-More formally, one could state that an algorithm is a sequence of \emph{measurable} functions $(\cA_t)_{t\geq1}$,
+A \emph{bandit algorithm} $\cA$ is also referred to as a strategy or a policy, and sometimes it is denoted by $\pi$ or $\rho$ in the literature.
+The algorithm $\cA$ selects an arm $A(t)$ at time $t$, possibly by using the past observations and the past external randomness.
+As shown below, being oblivious to the past yields very poor performance (the pure exploration policy is discussed in Section~\ref{sub:2:naiveSimpleStrategies}), so efficient policies indeed depend on the environment feedback.
+%
+More formally, an algorithm can be defined as a sequence of \emph{measurable} functions $(\cA_t)_{t\geq1}$,
 where $\cA_t$ maps the past observations $O_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$
@@
 The initial information is reduced to $O_1 = (U_0)$, and the first decision is $A(1) = \cA_1(O_1)$.
-Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps.
+Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps, as illustrated for example in Algorithm~\ref{algo:2:indexPolicy} below.
 %
@@
 
-In the example above, we consider Bernoulli distributions, but other distributions have been studied in the literature.
-From now on until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
+In the example above, we consider Bernoulli distributions, but other real-valued distributions have been studied in the literature.
+From now on and until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
 %
@@
 while in the literature, the Bernoulli distribution is the most common case of finitely supported distributions.
-Continuous-valued distributions with finite support also included truncated versions of infinitely supported distributions, in particular truncated Gaussian are used for numerical experiments in lots of research articles.
+Continuous-valued distributions with finite support also included truncated versions of infinitely supported distributions, in particular \emph{truncated} Gaussian are used for numerical experiments in lots of research articles.
 
@@
 $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$,
-we can write $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$,
+we can write $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
 and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$:
@@
 
-\textbf{Hypotheses in this thesis.}
+\paragraph{Hypotheses in this thesis.}
 %
@@
 Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.
+
+\TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira...}
 In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
@@
 \begin{lemma}[Regret decomposition]\label{lem:2:RegretDecomposition}
-    The (mean) regret $R_T^{\cA}$ can be decomposed as a sum of the number of selections of sub-optimal arms $k$,
-    weighted by their gaps, like this:
+    The (mean) regret $R_T^{\cA}$ can be decomposed as a sum of the number of selections of sub-optimal arms $k$, weighted by their gaps:
     \begin{equation}\label{eq:2:RegretDecomposition}
@@
     Thus we can rewrite the expected cumulated rewards as follows:
-
+    %
+    % FIXME add a blank line or a negative space if it looks weird
+    %
     % \vspace*{-20pt}
@@
 
-\textbf{Consequences.}
+\textbf{Consequences of Lemma~\ref{lem:2:RegretDecomposition}.}
 %
@@
 
-This family is non-empty as it contains for instance the $\UCB_1$ algorithm, as its regret is proven to be logarithmic on any problem instance.
+This family is non-empty as it contains for instance the $\UCB_1$ algorithm, as its regret is proven to be logarithmic on any problem instance with bounded rewards (in Theorem~\ref{thm:2:UCBregretBound} below).
 Now we can state the second logarithmic lower-bound, for algorithms in this family:
@@
         % \begin{equation}\label{eq:2:forSecondLogTLowerBound}
-        $C_I = \mu^* (1 - \mu^*) \sum\limits_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{1}{\Delta_k}$.
+        $C_I = \mu^* (1 - \mu^*) \sum\limits_{k: \Delta_k > 0} \frac{1}{\Delta_k}$.
         % \end{equation}
@@
         % \begin{equation}\label{eq:2:forSecondLogTLowerBound2}
-        $C_I = \sum\limits_{\substack{k=1,\dots,K\\\Delta_k > 0}} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon$.
+        $C_I = \sum\limits_{k: \Delta_k > 0} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon$.
         % \end{equation}
@@
 
-Moreover, it is interesting to note that the second lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proved to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
+Moreover, it is interesting to note that the second lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
 should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times in the total of $T$ time steps.
@@
 %
-For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses a constant $C_I$ no longer bring any information.
+For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses a constant $C_I$ no longer brings any information.
 Indeed, we can naively bound the regret by $R_T^{\cA} \leq (\max_k \Delta_k) T$, and thus if $(\max_k \Delta_k)$ can be taken so small that $C_I \log(T) \gg (\max_k \Delta_k) T$, and thus a regret upper bound like $R_T^{\cA} \leq C_I \log(T)$ is useless.
-
+%
+\begin{proof}
     For an example, consider any large horizon $T>10$, and $K=2$ Bernoulli arms.
@@
     $\varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T$.
+\end{proof}
 
@@
 
-In this thesis, we only study analytically the mean regret $R^{\cA}$ (in Chapters \ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
+In this thesis, we only study analytically the mean regret $R_T^{\cA}$ (in Chapters \ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
 
@@
 % - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-The \emph{switching cost} $SC^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $SC^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
+The \textbf{switching cost} $SC^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $SC^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
 It has recently gained interest in the literature, for instance in \cite{Koren17}.
@@
 %
-It is more interesting to studying $SC_M^{\cA}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game, as in this case, it is possible that the players follow an efficient centralized or decentralized algorithm (which gives a logarithmic centralized regret) while still suffering from a linear switching cost.
-It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $SC_M^{\cA}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
+It is more interesting to study $SC^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
+Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
+%
+It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $SC^{\cA_1,\dots,\cA_M}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
 % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
@@
 For more details, see \cite{modiDemo2016} for experiments on the multi-player case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
+While such ideas can interest a practitioner, they does not change the asymptotic behavior of $SC^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
 
 % - Fairness ?
-Finally, another interesting measure of performance is the \emph{fairness}.
+Finally, another interesting measure of performance is the \textbf{fairness}.
 It was not studied much for single-player problems, and usually it means that each arm must be explored at least a given fraction of the total horizon $T$ (\ie, ``arm-wise fairness'').
@@
 
-Pinsker's inequality is proved below, using simple real analysis arguments\footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).}.
+Pinsker's inequality is proven below, using simple real analysis arguments\footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).}.
 %

### 2019-06-04 2-Chapters/3-Chapter/chapter3.tex
Typos
Done for now
Lots of small change haha

@@
 
-\TODOL{Intégrer les commentaires de Christophe, le plus rapidement possible !}
+\TODOL{Abstract ici et plus court, et petite introduction !}
+\TODOL{Analyser les expériences les plus récentes et décrire mieux en 3.2, et refaire toutes les courbes pour 3.2 et 3.3}
 
@@
 %
-This chapter is based on the following publication: \cite{SMPyBanditsJMLR}.
+This chapter is based on our article \cite{SMPyBanditsJMLR}.
 The code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is freely publicly released under the MIT open-source License.

### 2019-06-04 2-Chapters/4-Chapter/chapter4.tex
Typos
Lots of small change haha

@@
 
-\TODOL{Intégrer les commentaires de Christophe, le plus rapidement possible !}
+\TODOL{Abstract ici et plus court, et petite introduction !}
 
@@
 
-This chapter is mainly based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
+This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 

### 2019-06-04 2-Chapters/5-Chapter/chapter5.tex
On the one hand… on the other hand
- NOT on one hand
- NOT on the first hand
Lots of small change haha

@@
 
-This chapter is mainly based on the following publication: \cite{Besson2018ALT}.
+This chapter is mainly based on our article \cite{Besson2018ALT}.
 
@@
 
-On the first hand, piece-wise stationary multi-players MAB can be tackled by extending algorithms developed for the single-player case.
+On the one hand, piece-wise stationary multi-players MAB can be tackled by extending algorithms developed for the single-player case.
 In \cite{WeiSrivastava18Distributed}, the authors study exactly the same multi-players bandit model as our model, and they propose two distributed algorithms that can efficiently be used by $M \leq K$ players to achieve sub-linear regret for piece-wise stationary problems, essentially by considering it as a harder case of a stationary problem.

### 2019-06-04 2-Chapters/6-Chapter/chapter6.tex
My thesis is so long, and so stupid
Thanks Christophe for comments on Ch2
SHIT I'm going home
Typos
Lots of small change haha

@@
 % \chapter{Non-Stationary MAB Models and Possible Applications for IoT Networks}
-\chapter{The Piece-wise Stationary MAB Model and the GLR-klUCB Algorithm}
+% \chapter{The Piece-wise Stationary MAB Model and the GLR-klUCB Algorithm}
+\chapter{Piece-Wise Stationary Multi-Armed Bandits}
 \label{chapter:6}
 \minitoc
+
+\TODOL{Abstract ici et plus court, et petite introduction !}
 
@@
 %
-This chapter is based on the following publications: \cite{Besson2019GLRT,Besson2019Gretsi}.
+This chapter is based on our articles \cite{Besson2019GLRT,Besson2019Gretsi}.
 

### 2019-06-04 2-Chapters/2-Chapter/chapter2.tex
Done for tonight

@@
 % - Quantile regret or just histogram of regrets
-First, when doing numerical experiments about bandits, if one studies the regret as an empirical mean based on a large number of random repetitions (\eg, $N=1000$ repetitions), it is important to not only show the mean value but also the variance of the values taken by the regret on each repetition, to verify that all algorithms perform consistently.
+First, when doing numerical experiments about bandits, if one studies the regret as an empirical mean based on a ``large'' number of random repetitions (\eg, $N=1000$ repetitions), it is important to not only show the mean value but also the variance of the values taken by the regret on each repetition, to verify that all algorithms perform consistently.
 Indeed, by only visualizing the mean of $1000$ values, it is possible that we miss some ``bad runs'': if $1$ run out of the $1000$ gives linear regret (\ie, $R_T^{\cA} \propto T$) and the $999$ other give logarithmic regret, then the mean will appear logarithmic.
@@
 % - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-The \textbf{switching cost} $SC^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $SC^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
-It has recently gained interest in the literature, for instance in \cite{Koren17}.
-% FIXME give the main result proved in this paper?
-In single-player models,
-it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
-But even if $SC^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
-%
-It is more interesting to study $SC^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
+The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $\mathrm{SC}^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
+It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptatively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.
+%
+Indeed, in single-player models it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
+But even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
+%
+It is more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
 Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
@@
 Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
-which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all problems expect where $\mu_1=\mu_2=\dots=\mu_K$).
+which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all non trivial problems, rulling out the case where $\mu_1=\mu_2=\dots=\mu_K$).
 
@@
 Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
-If $\varepsilon$ is constant, then the (mean) regret is still growing linearly as it is lower bounded by $\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k T$.
-In this first case, in average $T_0 = \varepsilon T$ steps are spent on exploration.
-%
-But if we consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$,
-then it was shown in \cite{Auer02} that the regret is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{\log(T)}$.
-In this second case, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
+On the one hand, if $\varepsilon$ is constant, then the (mean) regret still grows linearly, as it is lower bounded by $(\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k) T$.
+In average, $T_0 = \varepsilon T$ steps are spent on exploration.
+%
+On the other hand, if a lower-bound $d$ on the positive gaps is known beforehand,
+we can consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
+Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
+Here again, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
 
@@
                 \uIf{$t \leq \textcolor{deepgold}{T_0}$}{
-                    Play uniformly at random $A(t) \in \cU(\{1,\dots,K\})$
+                    Play sequentially $A(t) \in 1 + (t \mod K)$
                     \tcp*[f]{Explore}
@@
 % https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
--- The ``\textcolor{deepgold}{Explore-then-Exploit}'' strategy, as studied for instance in \cite{GarivierETC2016}, consists in first exploring uniformly the $K$ arms for $T_0$ time steps, then only exploiting the best arm.
-We can lower bound the regret of this strategy by
+-- The ``\textcolor{deepgold}{Explore-then-Exploit}'' strategy, as studied for instance in \cite{GarivierETC2016}, first explores uniformly the $K$ arms for $T_0$ time steps, then only exploits the arm identified as the best arm after these first $T_0$ steps
+(\ie, one of the arms with highest empirical means, after $T_0/K$ samples of each arm).
+Usually we restrict to $T_0$ being a multiple of $K$.
+On the one hand, we can lower bound its regret by
 $R_T \geq K (\min_{k: \Delta_k > 0} \Delta_k) T_0/K + p (T - T_0)$,
-if $p$ denotes the probability that the chosen arm is not optimal (\ie, the arm with highest mean after $T_0/K$ samples of each arms).
+if $p$ denotes the probability that the chosen arm is not optimal.
 And we can show that $p$ only depends on the number of collected samples $T_0$ and the gaps $\Delta_k$,
 so if $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, \ie, $R_T = \Omega(T)$.
-%
-To prove a bound on $p$, we use the following lemma, well known as Hoeffding's inequality, from \cite{hoeffding1963probability}.
+
+On the other hand, if a lower-bound $d>0$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret.
+%
+To prove a bound on $p$, we use Hoeffding's inequality from \cite{hoeffding1963probability}, reminded below in Lemma~\ref{lem:2:HoeffdingInequality}.
+%
+We have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus Hoeffding's inequality gives $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$.
+Optimizing on $T_0$ gives $T_0 = K (4/d^2) \log(d^2 T / 4)$
+and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K (4/d^2) (1 + \log(d^2 T / 4)) = \bigO{K \log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.
+%
+We note that this strategy can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
+%
+\begin{proof}
+    Let us prove the inequality $p \leq 2 K \exp(-T_0 d^2 / 4)$.\\
+    \indent
+    %
+    First for the case of two arms:
+    if the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality to control this probability $p$.
+    %
+    If $\mu_1 = \mu_2 + \Delta$, then
+    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \mu_1 - \Delta/2) + \bP(\widehat{\mu_2} > \mu_2 + \Delta/2)$, and both terms can be bounded by using Hoeffding's inequality with a (non-random) number of samples $n=T_0/2$,
+    to obtain $p \leq 2 \exp(-T_0 \Delta^2 / 4)$.\\
+    \indent
+    %
+    Then for $K\geq2$ arms, a simple union bound gives
+    $p \leq 2 K \exp(-T_0 d^2 / 4)$, if $d$ is a lower-bound on the positive gaps $\Delta_k$.
+    %
+    In all cases, we showed that $p$ is controlled by $T_0$.
+\end{proof}
 
@@
     \begin{align}\label{eq:2:HoeffdingInequality}
+        \begin{cases}
             \forall x < \mu, \;\;\; & \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
@@
             \forall x > \mu, \;\;\; & \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+        \end{cases}
     \end{align}
 \end{lemma}
-
-\begin{proof}
-    We prove it for the case of two arms.
-    If the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality to control this probability $p$.
-    %
-    If $\mu_1 = \mu_2 + \Delta$, then
-    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \mu_1 - \Delta/2) + \bP(\widehat{\mu_2} > \mu_2 + \Delta/2)$, and both terms can be bounded by using Hoeffding's inequality with $n=T_0/2$ samples,
-    to obtain $p \leq 2 \exp(-T_0 \Delta^2 / 4)$.
-
-    A simple union bound extends the proof for $K$ arms, giving
-    $p \leq 2 K \exp(-T_0 d^2 / 4)$, if $d$ is a lower-bound on the positive gaps.
-    %
-    In all cases, it shows that $p$ is a constant wrt $T$.
-\end{proof}
-
-Using the same analysis, if a lower-bound $d$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret.
-%
-Like for the lower-bound, we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus Hoeffding's inequality gives $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$.
-Optimizing on $T_0$ gives $T_0 = 4/d^2 \log(d^2 T / 4)$
-and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K 4/d^2 (1 + \log(d^2 T / 4)) = \bigO{\log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.
-%
-This strategy can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
 
@@
 \begin{proof}
+    \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ?}
     As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref\label{lem:2:HoeffdingInequality}.
@@
 %
-\TODOL{Expliquer klUCB}
-
 The \klUCB{} algorithm follows the same spirit as the $\UCB_1$ algorithm presented above,
@@
 As explained before, in the last few years, the MAB research community has been interested in finding algorithms which achieves both a problem-dependent logarithmic and a minimax upper-bounds.
-The first example is MOSS from \cite{Audibert2009minimax}, or recently \KLUCBpp{} in \cite{Menard17}.
+The first example is MOSS from \cite{Audibert2009minimax}, or \KLUCBpp{} in \cite{Menard17}.
 They are both index-based (see Algorithm~\ref{algo:2:indexPolicy}), using the following indexes:

### 2019-06-05 1-Preamble/4-Acknowledgement/acknowledgement.tex
Je viens de faire un premier jet (brouillon)
- de l'abstract FR/EN
- du résumé en français

@@
 I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
-And due to space constraints, I had to chose quite arbitrarily who to include\dots}: Florian, Fabian, Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia from Briançon, Mayotte and Pierre from Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément from Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also from Cachan, Kumudham and Priscilla from India, Thibault, Sébastien, Sélim, Valentin again from Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola from Rennes.
+And due to space constraints, I had to chose quite arbitrarily who to include\dots}: Florian, Fabian, Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham and Priscilla in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
 % In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
@@
 
-Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Edouard in Lille.
+Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
 I want to thank especially the ``coffee friends$^*$'' in Rennes, for the daily interesting discussions, which where sometimes the highlights of our life at the office.
@@
 %
-I am thankful to all the great professors who taught me during my studies at ENS de Cachan, between 2011 and 2016, and for some who pushed me into teaching, and studying statistical learning and computer science, including Frédéric, Alain and Claudine in the maths department, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
+I am thankful to the great professors who taught me while I studied at ENS de Cachan, between 2011 and 2016, and for some who pushed me into teaching, and studying statistical learning and computer science, including Frédéric, Alain, Claudine and Nicolas in the maths department, Hubert, Jean, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
 

### 2019-06-05 1-Preamble/5-Abstract/abstract.tex
Work work work...
Je viens de faire un premier jet (brouillon)
- de l'abstract FR/EN
- du résumé en français

@@
 \begin{abstract}
+% \vspace*{1cm}
+
+% \TODOL{Raccourcir juste un peu ?}
 
 % ----------------------------------------------------------------------------
-Blabla.
 
-Dans cette thèse, blabla.
+% Chapitre 1
+Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
+Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
+%  (Chapitre~\ref{chapter:1}).
+% Chapitre 4
+Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement.
+%  (Chapitre~\ref{chapitre:4}).
+% Chapitre 2
+Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB), mais leur analyse est difficile à réaliser,
+% , Chapitre~\ref{chapitre:2}
+car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+% En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
+% Chapitres 5 et 6
+Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueurs non stationnaires.
+%  (Chapitre~\ref{chapitre:5})
+%  (Chapitre~\ref{chapitre:6})
+% Chapitre 3
+Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres.
+% (Chapitre~\ref{chapter:3})
 
-\newpage
+% \newpage
+% ----------------------------------------------------------------------------
 \begin{center}
@@
 \end{center}
-\vspace{1cm}
+% \vspace*{1cm}
 
-% ----------------------------------------------------------------------------
-Blabla.
-
-In this PhD thesis, blabla.
+% Chapter 1
+In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access large-scale Cognitive Radio networks, in unlicensed bands and without central control.
+More specifically, we focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
+%  (Chapter~\ref{chapter:1})
+% Chapter 4
+We propose different models of IoT networks, and we show empirically on both numerical simulations and real-world validation the possible gain of our methods, that use Reinforcement Learning.
+%  (Chapter~\ref{chapter:4})
+% Chapter 2
+The different network access problems are modeled as Multi-Armed Bandits (MAB), but we found that analyzing the realistic models was intractable,
+% ,  Chapter~\ref{chapter:2}
+because proving the convergence of many end-devices playing a collaborative game without communication no coordination is hard, when end-devices all follow random active pattern.
+% Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.
+% Chapter 5 and 6
+The rest of this manuscript thus studies two restricted models, first multi-player bandits in stationary problems, then non-stationary single-player bandits.
+%  (Chapter~\ref{chapter:5})
+%  (Chapter~\ref{chapter:6})
+% Chapter 3
+We also detail another contribution, the SMPyBandits open-source Python library for numerical simulations of MAB problems, which covers all the studied models and more.
+%  (Chapter~\ref{chapter:3})
 

### 2019-06-05 1-Preamble/5-Abstract/resume_fr.tex
Je suis tellement inefficace c'est triste
Je viens de faire un premier jet (brouillon)
- de l'abstract FR/EN
- du résumé en français

@@
 
-Résumé de la thèse en français.
+% Résumé de la thèse en français.
 
 % ----------------------------------------------------------------------------
-\section*{Partie 1}
-\subsection*{Sous partie 1.1}
-\subsection*{Sous partie 1.2}
-\subsection*{Sous partie 1.3}
 
-% ----------------------------------------------------------------------------
-\section*{Partie 2}
-\subsection*{Sous partie 2.1}
-\subsection*{Sous partie 2.2}
-\subsection*{Sous partie 2.3}
-
-% ----------------------------------------------------------------------------
-\section*{Partie 3}
-\subsection*{Sous partie 3.1}
-\subsection*{Sous partie 3.2}
-\subsection*{Sous partie 3.3}
+% Chapitre 1
+Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
+Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil (Chapitre~\ref{chapter:1}).
+% Chapitre 4
+Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement (Chapitre~\ref{chapter:4}).
+% Chapitre 2
+Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB, Chapitre~\ref{chapter:2}), mais leur analyse est difficile à réaliser,
+% , Chapitre~\ref{chapter:2}
+car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+% En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
+% Chapitres 5 et 6
+Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires (Chapitre~\ref{chapter:5}), puis des bandits simple-joueurs non stationnaires (Chapitre~\ref{chapter:6}).
+% Chapitre 3
+Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres (Chapitre~\ref{chapter:3}).
 
@@
 
-Cette thèse aborde tous les points susmentionnés et propose différentes solutions pour chaque scénario.
-Les contributions importantes sont résumées ci-dessous :
+% Cette thèse aborde tous les points susmentionnés et propose différentes solutions pour chaque scénario.
+Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse :
+
+\TODOL{Finir de traduire}
 \begin{itemize}
-\item Item 1
-\item Item 2
-\item Item 3
-\item Item 4
+    \item
+    Nous présentons les concepts et les notations du problème des bandits à armes multiples dans le chapitre~\ref{chapter:2}, d'un point de vue mathématique.
+    Mais nous suivons également une approche didactique puisque nous utilisons une démonstration interactive en ligne conçue pour permettre à n'importe qui de jouer contre un petit problème de bandit depuis son navigateur, dans Section~\ref{par:2:interactiveDemoDemoDiscoverMAB}.
+
+    \item
+    Nous donnons une brève revue de la littérature sur les algorithmes stochastiques de bandit dans Section~\ref{sec:2:famousMABalgorithms}.
+
+    \item
+    Nous présentons le problème du choix de l'algorithme qu'un praticien devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents MAB disponibles, dans la section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes comme solution en ligne au problème de sélection d'algorithmes, et des simulations numériques pour illustrer qu'il atteint des performances empiriques de pointe.
+    \cite{Besson2018WCNC}.
+    avec \textbf{Aggregateur} (WCNC 2018)
+
+    \item
+    Nous avons écrit la bibliothèque de simulation open-source la plus complète pour les problèmes MAB, appelée SMPyBandits, qui est publiée en ligne sous une licence open-source \cite{SMPyBandits,SMPyBanditsJMLR}.
+    Nous présentons en détail son architecture et ses fonctionnalités dans le chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
+    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences utilisées dans la suite de cette thèse.
+
+    \item
+    Nous proposons différents modèles pour les réseaux IdO, dans le chapitre~\ref{chapter:4}, où les terminaux dotés de capacités radio cognitives peuvent implémenter des algorithmes MAB de leur côté, pour augmenter automatiquement la durée de vie de leur batterie et permettre à davantage de périphériques d'utiliser le même réseau tout en maintenant une haute qualité de service
+    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC}.
+    (CROWNCOM 2017, ICT demo 2018, WCNC 2019 et MOTIoN 2019)
+
+    \item
+    Nous avons implémenté une preuve de concept du modèle susmentionné \cite{Besson2018ICT}, et nous la présentons en détail dans Section~\ref{sec:4:gnuradio}.
+Nous avons réalisé une vidéo présentant notre démonstration, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+
+    \item
+    Le code source des deux contributions susmentionnées est publié en ligne, ainsi que des instructions claires pour la reproduction de nos travaux.
+
+    \item
+    Nous formalisons le modèle du bandit multijoueur, et nous avons introduit trois variantes, dans le chapitre~\ref{chapter:5}.
+    Pour le cas de l'information de détection, nous proposons deux nouveaux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM{} pour montrer qu'il est d'ordre-optimal,
+    ainsi que des expériences numériques approfondies pour démontrer sa bonne performance par rapport au reste de la littérature.
+    Notre travail \cite{Besson2018ALT} a également donné une nouvelle impulsion à la recherche sur les bandits multi-joueurs, car certains travaux de recherche récents se sont construits sur nos résultats.
+
+    \item
+    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle MAB multi-joueurs, qui, selon nous, n'a jamais été écrit auparavant.
+    (ALT 2018, et 4-8 travaux inspirés de notre article depuis lors).
+Etat de l'art avec notre algorithme MCTopM + klUCB pour les bandits multi-joueurs "avec détection", état de l'art empirique avec notre approche simple (mais erronée) "égoïste" en cas de "non détection".
+
+    \item
+    Nous présentons également le modèle MAB stationnaire pièce par pièce, dans le chapitre~\ref{chapter:6}, et une analyse documentaire détaillée de la recherche sur le MAB non stationnaire \cite{Besson2019GLRT,Besson2019Gretsi}.
+    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif pour le problème de station stationnaire pièce par pièce, \GLRklUCB, qui atteint des performances de pointe.
+    Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB
+
+    \item
+    La dernière contribution de cette thèse est une revue de la littérature sur les cas d'utilisation possibles de la technique du doublement pour les problèmes MAB,
+    et une analyse unifiée et plus générique de deux familles de doubles tours.
+    Ceci a conduit à l'article \cite{Besson2018DoublingTricks}, qui est rapidement présenté en Annexe~\ref{app:2:DoublingTricks}, mais pas dans le texte principal de la thèse.
 \end{itemize}
@@
 Ce manuscrit est organisé comme suit.
-Nous rappelons dans le \textbf{Chapitre \ref{chapter:1}} blabla.
 
-Dans le \textbf{Chapitre \ref{chapter:2}}, blabla.
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.95\textwidth}{!}{
+    \begin{tikzpicture}[>=latex',line join=bevel,scale=2.25]
+        %
+        \node[align=center] (introduction) at (0,3.25) [rectangle,draw] {Chapitre~\ref{chapter:1}\\Introduction};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw] {Chapitre~\ref{chapter:2}\\Modèle bandit à plusieurs bras\\stochastique et stationnaire};
+        \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapitre~\ref{chapter:3}\\SMPyBandits : bibliothèque de\\simulation pour MAB};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:4}\\Deux modèles MAB\\pour les réseaux IoT};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapitre~\ref{chapter:5}\\Modèle MAB\\Multi-joueurs};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:6}\\Modèle MAB\\Non-stationnaire};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapitre~\ref{chapter:conclusion}\\Conclusion générale};
+        %
+        \draw [color=black,thick,->] (introduction) to (chapter2);
+        \draw [color=black,thick,<->] (chapter2) to (chapter3);
+        \draw [color=black,thick,->] (chapter2) to (chapter4);
+        \draw [color=black,thick,->] (chapter2) to (chapter5);
+        \draw [color=black,densely dotted,<->]   (chapter4) to (chapter5);
+        \draw [color=black,densely dotted,->] -| (chapter3) to (chapter6);
+        \draw [color=black,densely dotted,<->]   (chapter5) to (chapter6);
+        \draw [color=black,thick,->] (chapter2) to (chapter6);
+        \draw [color=black,thick,->] (chapter4) to (conclusion);
+        \draw [color=black,thick,->] (chapter5) to (conclusion);
+        \draw [color=black,thick,->] (chapter6) to (conclusion);
+        %
+    \end{tikzpicture}
+    }
+    \caption{Organisation de la thèse : une carte de lecture.
+Tout chemin contenant le Chapitre~\ref{chapter:1}, le Chapitre~\ref{chapter:2}, au moins un des trois Chapitres~\ref{chapter:4}, \ref{chapter:5} et \ref{chapter:6}, et la Conclusion est une façon valide de lire cette thèse.}
+    \label{fig:1:organization}
+\end{figure}
 
-Dans le \textbf{Chapitre \ref{chapter:3}}, blabla.
+\TODOL{Finir de traduire}
+L'ordre de lecture du manuscrit peut être n'importe quel chemin descendant entre l'introduction dans le chapitre actuel~\ref{chapter:1}, et la conclusion dans le dernier chapitre~\ref{chapter:conclusion}.
+Comme le montre le graphique de la Figure~\ref{fig:1:organisation},
+la thèse est organisée en deux parties :
 
-Dans le \textbf{Chapitre \ref{chapter:4}}, blabla.
+Dans la partie~\ref{part:Introduction}, nous commençons par le Chapitre~\ref{chapter:2} où nous présentons les modèles MAB, les concepts et les notations utilisés dans tout ce document, et il est nécessaire à la lecture du reste du manuscrit.
+Par contre le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulation SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
 
-Dans le \textbf{Chapitre \ref{chapter:5}}, blabla.
-
-Dans le \textbf{Chapitre \ref{chapter:5}} blabla.
+La deuxième partie~\ref{part:MABIOT} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus presque indépendamment.
+Le Chapitre~\ref{chapter:4} commence par présenter différents modèles de réseaux IdO où les algorithmes MAB ont été utilisés avec succès.
+Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop généraux pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
+Pour cette raison, nous affaiblissons les modèles pour le reste du document,
+et les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
 

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Je vais lancer les simulations avant d'aller au foot

@@
 
+\subsection{A first numerical experiment on the example Bernoulli problem}
+\label{sub:2:shortNumericalExperiments}
+
+We finish this Section~\ref{sec:2:famousMABalgorithms} by illustrating the performance of different algorithms, on the small MAB problem used in the online demonstration shown above (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
+We consider $K=5$ arms following Bernoulli distributions, of means $\mu_1=0.6,\mu_2=0.2,\mu_3=0.55,\mu_4=0.7,\mu_5=0.5$ (unknown to the algorithms), and first an horizon $T=100$ (like in the online demo) then a large value of $T=10000$.
+The best arm is $\mu_4$ and the smallest gap is $\Delta = 0.1$, thus a lower-bound on the gaps is taken as $d=\Delta$.
+% (used to tune the two algorithms from Section~\ref{sub:2:naiveSimpleStrategies}).
+
+The different algorithms being compared are the following, in the order they were presented above:
+%
+first we include the pure-exploration and
+pure-exploitation algorithms,
+then the $\varepsilon$-greedy (with $\varepsilon_t=\varepsilon_0/t$ and $\varepsilon_0=6K/d^2=3000$) and
+Explore-then-Commit (with $T_0=K4/d^2 \log(d^2T/4)=2000\log(T/400)$) algorithms
+from Section~\ref{sub:2:naiveSimpleStrategies}, which both uses a tuning based on a prior knowledge of $\Delta$ (that gives large values of $\varepsilon_0$ and $T_0$ yielding linear regret for small horizon).
+%
+We then include efficient algorithms, that do not need prior knowledge of the $T$ (they are anytime) nor $\Delta$ (their are robust and efficient on unknown problems):
+$\UCB_1$ (with $\alpha=1$),
+\klUCB{} (with Bernoulli $\kl = \KL_{B}$),
+and Thompson sampling (with Beta posteriors and flat uniform priors).
+%
+
+We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the regret of the different algorithms, \emph{for one simulation}.
+Then in Table~\ref{table:2:meanResults} we show the results of the average on $N=1000$ independent runs.
+
+\begin{table}[ht]
+    \begin{small}  % WARNING
+        \centering
+        \begin{tabular}{c|cc|cc}
+        \textbf{Algorithms} $\;$ \textbackslash $\;$ Horizon & Reward & Regret $T=100$ & Reward & Regret $T=10000$ \\
+            \hline
+            Pure exploration
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            Pure exploitation
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \hline
+            $\varepsilon$-greedy
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            Explore-then-Commit
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \hline
+            $\UCB_1$
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \klUCB{}
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            Thompson sampling
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \hline
+        \end{tabular}
+        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{for only one run of the simulation}, for different algorithms.}
+        \label{table:2:oneShotResult}
+    \end{small}  % WARNING
+\end{table}
+
+\TODOL{Lancer les simulations !}
+
+\begin{table}[ht]
+    \begin{small}  % WARNING
+        \centering
+        \begin{tabular}{c|cc|cc}
+        \textbf{Algorithms} $\;$ \textbackslash $\;$ Horizon & Reward & Regret $T=100$ & Reward & Regret $T=10000$ \\
+            \hline
+            Pure exploration
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            Pure exploitation
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \hline
+            $\varepsilon$-greedy
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            Explore-then-Commit
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \hline
+            $\UCB_1$
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \klUCB{}
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            Thompson sampling
+                & $XXX$ & $XXX$
+                & $XXX$ & $XXX$ \\
+            \hline
+        \end{tabular}
+        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{averaged over $N=1000$ independent simulations}, for different algorithms.}
+        \label{table:2:meanResults}
+    \end{small}  % WARNING
+\end{table}
+
+\TODOL{Décrire les résultats !}
+
 \paragraph{Summary.}

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Okay experimenting

@@
         % = \sum_{k=1}^K \mu_k \E \left[ \sum_{t=1}^T \mathbbm{1}(A(t) = k) \right]\\
-        %  XXX 
         = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right].
@@
 
-\textbf{Consequences of Lemma~\ref{lem:2:RegretDecomposition}.}
-%
-Such decomposition of the regret is useful for at least two reasons:
+\textbf{Consequences.}
+%
+Such decomposition of the regret can be useful for at least two reasons:
 
@@
 In our library SMPyBandits, we implement both estimators, and all values of regret used in this thesis are based on the one using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives faster convergence, and also ``smoother'' plots.
+\label{remark:2:moreAccurateCountofRegretForSimulations}
 
@@
 
-\subsection{A first numerical experiment on the example Bernoulli problem}
+% \subsection{A first numerical experiment on the example Bernoulli problem}
+\subsection{Comparison of the main algorithms on the Bernoulli problem of the online demonstration}
 \label{sub:2:shortNumericalExperiments}
@@
 
-We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the regret of the different algorithms, \emph{for one simulation}.
+We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the estimated regret\footnote{~First, the rewards are computed as $\frac{1}{N} \sum_{j=1}^N [ \sum_{t=1}^T r^{(j)}_{A(t)}(t) ]$, where $r^{(j)}_{k}(t)$ means the reward stream of the different runs $j=1,\dots,N$.
+Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N [ \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t) ]$ (as it could be negative), but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N [ \sum_{k=1}^K \mu_k N^{(j)}_{k}(T) ]$.
+We noted in a remark after the proof of Lemma~\ref{lem:2:RegretDecomposition} (in page~\pageref{remark:2:moreAccurateCountofRegretForSimulations}) that this later formula is a more accurate empirical estimate of the regret.}
+of the different algorithms, \emph{for one simulation}.
 Then in Table~\ref{table:2:meanResults} we show the results of the average on $N=1000$ independent runs.
+
+% Final ranking for this environment #0 : (using more accurate estimate of the regret)
+% - Policy 'Pure exploration'     was ranked      7 / 7 for this simulation
+%         (last regret = 20.9,    total regret = 20.9,    total reward = 46,      total weighted selection = 49.1).
+% - Policy 'Pure exploitation'    was ranked      1 / 7 for this simulation
+%         (last regret = 1.15,    total regret = 1.15,    total reward = 65,      total weighted selection = 68.85).
+% - Policy '$\varepsilon$-greedy' was ranked      6 / 7 for this simulation
+%         (last regret = 20.85,   total regret = 20.85,   total reward = 50,      total weighted selection = 49.15).
+% - Policy 'Explore-then-Exploit' was ranked      5 / 7 for this simulation
+%         (last regret = 15,      total regret = 15 ,      total reward = 49,      total weighted selection = 55).
+% - Policy 'UCB($\alpha=1$)'      was ranked      2 / 7 for this simulation
+%         (last regret = 10.3,    total regret = 10.3,    total reward = 63,      total weighted selection = 59.7).
+% - Policy 'Bernoulli kl-UCB'     was ranked      4 / 7 for this simulation
+%         (last regret = 12.4,    total regret = 12.4,    total reward = 56,      total weighted selection = 57.6).
+% - Policy 'Thompson sampling'    was ranked      3 / 7 for this simulation
+%         (last regret = 12.2,    total regret = 12.2,    total reward = 51,      total weighted selection = 57.8).
+
+% Final ranking for this environment #0 : (using more accurate estimate of the regret)
+% - Policy 'Pure exploration'     was ranked      7 / 7 for this simulation
+%         (last regret = 1877.6,  total regret = 1886.4,  total reward = 5171,    total weighted selection = 5113.6).
+% - Policy 'Pure exploitation'    was ranked      1 / 7 for this simulation
+%         (last regret = 0.95,    total regret = 0.95,    total reward = 6997,    total weighted selection = 6999).
+% - Policy '$\varepsilon$-greedy' was ranked      5 / 7 for this simulation
+%         (last regret = 380.44,  total regret = 382.25,  total reward = 6657,    total weighted selection = 6617.8).
+% - Policy 'Explore-then-Exploit' was ranked      6 / 7 for this simulation
+%         (last regret = 1199.8,  total regret = 1199.8,  total reward = 5766,    total weighted selection = 5800.2).
+% - Policy 'UCB($\alpha=1$)'      was ranked      4 / 7 for this simulation
+%         (last regret = 78.45,   total regret = 78.45,   total reward = 6920,    total weighted selection = 6921.5).
+% - Policy 'Bernoulli kl-UCB'     was ranked      3 / 7 for this simulation
+%         (last regret = 70.174,  total regret = 70.25,   total reward = 6922,    total weighted selection = 6929.7).
+% - Policy 'Thompson sampling'    was ranked      2 / 7 for this simulation
+%         (last regret = 44.054,  total regret = 44.2,    total reward = 6949,    total weighted selection = 6955.8).
 
@@
         \begin{tabular}{c|cc|cc}
-        \textbf{Algorithms} $\;$ \textbackslash $\;$ Horizon & Reward & Regret $T=100$ & Reward & Regret $T=10000$ \\
+        \textbf{Algorithms} & $\sum\limits_{t=1}^{T=100} r_{A(t)}(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r_{A(t)}(t)$ & $R_{T=10000}^{\cA}$ \\
             \hline
             Pure exploration
-                & $XXX$ & $XXX$
-                & $XXX$ & $XXX$ \\
-            Pure exploitation
-                & $XXX$ & $XXX$
-                & $XXX$ & $XXX$ \\
+                & $46$ & $20.9$
+                & $5171$ & $1886.4$ \\
+            \textbf{Pure exploitation}
+                & $\mathbf{65}$ & $\mathbf{1.15}$
+                & $\mathbf{6997}$ & $\mathbf{0.95}$ \\
             \hline
             $\varepsilon$-greedy
-                & $XXX$ & $XXX$
-                & $XXX$ & $XXX$ \\
+                & $50$ & $20.9$
+                & $6657$ & $382.2$ \\
             Explore-then-Commit
-                & $XXX$ & $XXX$
-                & $XXX$ & $XXX$ \\
+                & $49$ & $15$
+                & $5766$ & $1199.8$ \\
             \hline
             $\UCB_1$
-                & $XXX$ & $XXX$
-                & $XXX$ & $XXX$ \\
+                & $63$ & $10.3$
+                & $6920$ & $78.4$ \\
             \klUCB{}
-                & $XXX$ & $XXX$
-                & $XXX$ & $XXX$ \\
+                & $56$ & $12.4$
+                & $6922$ & $70.2$ \\
+            \hline
             Thompson sampling
-                & $XXX$ & $XXX$
-                & $XXX$ & $XXX$ \\
+                & $51$ & $12.2$
+                & $6949$ & $44.2$ \\
             \hline
         \end{tabular}
-        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{for only one run of the simulation}, for different algorithms.}
+        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{for only one run of the simulation}, for different algorithms.
+\textcolor{red}{It is very random}, see Table~\ref{table:2:meanResults} for average results on $N=1000$ runs, on both cases the pure exploitation was very lucky and obtains close-to-optimal performance.}
         \label{table:2:oneShotResult}
@@
 
-\TODOL{Lancer les simulations !}
+% Final ranking for this environment #0 : (using more accurate estimate of the regret)
+% - Policy 'Pure exploration'     was ranked      7 / 7 for this simulation
+%         (last regret = 19.04,   total regret = 19.04,   total reward = 50.914,  total weighted selection = 50.96).
+% - Policy 'Pure exploitation'    was ranked      1 / 7 for this simulation
+%         (last regret = 8.3496,  total regret = 8.3496,  total reward = 61.692,  total weighted selection = 61.65).
+% - Policy '$\varepsilon$-greedy' was ranked      6 / 7 for this simulation
+%         (last regret = 18.729,  total regret = 18.729,  total reward = 51.274,  total weighted selection = 51.271).
+% - Policy 'Explore-then-Exploit' was ranked      5 / 7 for this simulation
+%         (last regret = 12.859,  total regret = 12.859,  total reward = 57.193,  total weighted selection = 57.141).
+% - Policy 'UCB($\alpha=1$)'      was ranked      4 / 7 for this simulation
+%         (last regret = 9.8451,  total regret = 9.8451,  total reward = 60.031,  total weighted selection = 60.155).
+% - Policy 'Bernoulli kl-UCB'     was ranked      2 / 7 for this simulation
+%         (last regret = 9.3534,  total regret = 9.3534,  total reward = 60.526,  total weighted selection = 60.647).
+% - Policy 'Thompson sampling'    was ranked      3 / 7 for this simulation
+%         (last regret = 9.6005,  total regret = 9.6005,  total reward = 60.355,  total weighted selection = 60.399).
 
@@
         \begin{tabular}{c|cc|cc}
-        \textbf{Algorithms} $\;$ \textbackslash $\;$ Horizon & Reward & Regret $T=100$ & Reward & Regret $T=10000$ \\
+        \textbf{Algorithms} & $\frac{1}{N} \sum\limits_{j=1}^N [ \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t) ]$ & $\frac{1}{N} \sum\limits_{j=1}^N [ R^{\cA,(j)}_{T=100} ]$ & $\frac{1}{N} \sum\limits_{j=1}^N [ \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t) ]$ & $\frac{1}{N} \sum\limits_{j=1}^N [ R^{\cA,(j)}_{T=10000} ]$ \\
             \hline
             Pure exploration
-                & $XXX$ & $XXX$
+                & $51$ & $19$
                 & $XXX$ & $XXX$ \\
             Pure exploitation
-                & $XXX$ & $XXX$
+                & $62$ & $8.3$
                 & $XXX$ & $XXX$ \\
@@
             $\varepsilon$-greedy
-                & $XXX$ & $XXX$
+                & $51$ & $18.7$
                 & $XXX$ & $XXX$ \\
             Explore-then-Commit
-                & $XXX$ & $XXX$
+                & $57$ & $12.9$
                 & $XXX$ & $XXX$ \\
@@
             $\UCB_1$
-                & $XXX$ & $XXX$
+                & $\mathbf{60}$ & $\mathbf{9.8}$
                 & $XXX$ & $XXX$ \\
             \klUCB{}
-                & $XXX$ & $XXX$
+                & $61$ & $9.3$
                 & $XXX$ & $XXX$ \\
+            \hline
             Thompson sampling
-                & $XXX$ & $XXX$
+                & $\mathbf{60}$ & $\mathbf{9.6}$
                 & $XXX$ & $XXX$ \\
@@
 \end{table}
+
+% FIXME generate at the end of the simulation
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.00\linewidth]{meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000.pdf}
+	\caption{Mean cumulated rewards as function of the current time step $t$ for horizon $T=100$ and $N=1000$ independent runs, for different algorithms }
+	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}
+\end{figure}
+
+% FIXME generate at the end of the simulation
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.00\linewidth]{meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000.pdf}
+	\caption{Mean regret $R_t$ as function of $t$ for $T=100$ and $N=1000$.}
+	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000}
+\end{figure}
+
+% FIXME generate at the end of the simulation
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.00\linewidth]{meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000.pdf}
+	\caption{Mean cumulated rewards as function of $t$ for $T=10000$ and $N=1000$.}
+	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
+\end{figure}
+
+% FIXME generate at the end of the simulation
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.00\linewidth]{meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000.pdf}
+	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
+	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
+\end{figure}
+
+We also include in Figures~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regret as functions of the current time step $t$, for $T=100$ then $T=10000$ for $N=1000$ independent runs.
 

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Waiting for end of simulations

@@
 
+% Final ranking for this environment #0 : (using more accurate estimate of the regret)
+% - Policy 'Pure exploration'     was ranked      7 / 7 for this simulation
+%         (last regret = 1891.1,  total regret = 1900.5,  total reward = 5099,    total weighted selection = 5099.5).
+% - Policy 'Pure exploitation'    was ranked      5 / 7 for this simulation
+%         (last regret = 649.98,  total regret = 653.2,   total reward = 6348.6,  total weighted selection = 6346.8).
+% - Policy '$\varepsilon$-greedy' was ranked      4 / 7 for this simulation
+%         (last regret = 540.54,  total regret = 541.9,   total reward = 6460.7,  total weighted selection = 6458.1).
+% - Policy 'Explore-then-Exploit' was ranked      6 / 7 for this simulation
+%         (last regret = 1207.1,  total regret = 1207.1,  total reward = 5794.9,  total weighted selection = 5792.9).
+% - Policy 'UCB($\alpha=1$)'      was ranked      3 / 7 for this simulation
+%         (last regret = 79.435,  total regret = 79.513,  total reward = 6920.9,  total weighted selection = 6920.5).
+% - Policy 'Bernoulli kl-UCB'     was ranked      2 / 7 for this simulation
+%         (last regret = 72.274,  total regret = 72.368,  total reward = 6927.5,  total weighted selection = 6927.6).
+% - Policy 'Thompson sampling'    was ranked      1 / 7 for this simulation
+%         (last regret = 49.366,  total regret = 49.402,  total reward = 6951.4,  total weighted selection = 6950.6).
+
 \begin{table}[ht]
@@
         \begin{tabular}{c|cc|cc}
-        \textbf{Algorithms} & $\frac{1}{N} \sum\limits_{j=1}^N [ \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t) ]$ & $\frac{1}{N} \sum\limits_{j=1}^N [ R^{\cA,(j)}_{T=100} ]$ & $\frac{1}{N} \sum\limits_{j=1}^N [ \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t) ]$ & $\frac{1}{N} \sum\limits_{j=1}^N [ R^{\cA,(j)}_{T=10000} ]$ \\
+        \textbf{Algorithms} & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=100}$ & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=10000}$ \\
             \hline
@@
                 & $51$ & $19$
-                & $XXX$ & $XXX$ \\
+                & $5099$ & $1900$ \\
             Pure exploitation
                 & $62$ & $8.3$
-                & $XXX$ & $XXX$ \\
+                & $6349$ & $653$ \\
             \hline
@@
                 & $51$ & $18.7$
-                & $XXX$ & $XXX$ \\
+                & $6460$ & $542$ \\
             Explore-then-Commit
                 & $57$ & $12.9$
-                & $XXX$ & $XXX$ \\
+                & $5795$ & $1207$ \\
             \hline
@@
                 & $\mathbf{60}$ & $\mathbf{9.8}$
-                & $XXX$ & $XXX$ \\
+                & $6921$ & $79$ \\
             \klUCB{}
                 & $61$ & $9.3$
-                & $XXX$ & $XXX$ \\
+                & $6927$ & $72$ \\
             \hline
@@
                 & $\mathbf{60}$ & $\mathbf{9.6}$
-                & $XXX$ & $XXX$ \\
+                & $\mathbf{6951}$ & $\mathbf{49}$ \\
             \hline
         \end{tabular}
-        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{averaged over $N=1000$ independent simulations}, for different algorithms.}
+        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{averaged over $N=1000$ independent simulations} ($j=1,\dots,N$), for different algorithms.}
         \label{table:2:meanResults}
@@
 
-% FIXME generate at the end of the simulation
-\begin{figure}[h!]  % [htbp]
-	% \centering
-	\includegraphics[width=1.00\linewidth]{meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000.pdf}
-	\caption{Mean cumulated rewards as function of the current time step $t$ for horizon $T=100$ and $N=1000$ independent runs, for different algorithms }
-	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}
-\end{figure}
-
-% FIXME generate at the end of the simulation
-\begin{figure}[h!]  % [htbp]
-	% \centering
-	\includegraphics[width=1.00\linewidth]{meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000.pdf}
-	\caption{Mean regret $R_t$ as function of $t$ for $T=100$ and $N=1000$.}
-	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000}
-\end{figure}
+% % FIXME generate at the end of the simulation
+% \begin{figure}[h!]  % [htbp]
+% 	% \centering
+% 	\includegraphics[width=1.00\linewidth]{meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000.pdf}
+% 	\caption{Mean cumulated rewards as function of the current time step $t$ for horizon $T=100$ and $N=1000$ independent runs, for different algorithms }
+% 	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}
+% \end{figure}
+
+% % FIXME generate at the end of the simulation
+% \begin{figure}[h!]  % [htbp]
+% 	% \centering
+% 	\includegraphics[width=1.00\linewidth]{meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000.pdf}
+% 	\caption{Mean regret $R_t$ as function of $t$ for $T=100$ and $N=1000$.}
+% 	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000}
+% \end{figure}
 
@@
 
-We also include in Figures~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regret as functions of the current time step $t$, for $T=100$ then $T=10000$ for $N=1000$ independent runs.
+We also include in Figures
+% \ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000},
+\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regret as functions of the current time step $t$, for $T=10000$ for $N=1000$ independent runs.
+The reward of the best arm is also plotted in Figure~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, to illustrate convergence of the average reward $\frac{1}{t} \sum\limits_{s=1}^t r_{A(s)}(s)$, which should converge to $\mu^*$ when $t\to\infty$ (or $t\toT$) for efficient algorithms.
+Conversely, we also add the Lai \& Robbins' logarithmic lower-bound on regret ($C_I \log(t)$ from case 2 of Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plot in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
+%
+We can clearly identify the behavior of the different strategies:
+the three efficient strategies indeed achieve a sub-linear regret, that looks logarithmic as supported by the theory,
+while the naive strategies achieve linear regret.
+The two simple strategies also achieve sub-linear regret but are less efficient:
+both $\varepsilon$-greedy and Explore-then-Commit starts as pure exploration, respectively until $\varepsilon_0/t$ becomes smaller than $1$ and until $t > T_0$, then they almost always correctly identify the best arm and start to catch up with the best policies.
 

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
I'm starting to be happy about Chapter:2

@@
 For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.}.
+A curious reader about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and then refer to the famous book \cite{Norris98}.
 
@@
 The different applications detailed above are still active research directions,
-and a curious reader can find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}.
+and a curious reader can find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}, such as game tree search or network routing in Section~1.2 of \cite{LattimoreBanditAlgorithmsBook}.
 
@@
 while Thompson sampling is optimal for bounded rewards, and \klUCB{} have been proven to be optimal for both cases.
+%
+For more details, we refer to Chapter~16 of \cite{LattimoreBanditAlgorithmsBook}.
 
@@
 The following theorem is from \cite{Auer02NonStochastic}.
+%
+For more details, we refer to Chapter~15 of \cite{LattimoreBanditAlgorithmsBook}.
 
@@
 An extension of this strategy is to consider not a fixed time $T_0$ but a random time $\tau$ at which exploration stops.
-This time $\tau$ must be a \emph{stopping time} in the sense that it is a measurable random variable, dependent of the past observations, as introduced by \cite{Wald45}.
+This time $\tau$ must be a \emph{stopping time}\footnote{See Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}, and we use this notion again in Section~\ref{sec:6:GLRklUCB_Algorithm}.},
+in the sense that it is a measurable random variable, dependent of the past observations.
 The strategy is then referred to as ``explore-then-commit'' (ETC), and the idea is to use a statistical test at every time step $t$, and stop as soon as enough samples were collected to effectively identify the best arm with a certain confidence level $\delta$.
@@
 while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
-For a short proof of this results as well as more details on Exp3, see \cite{Bubeck12}.
-%
-% https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
-Finally, it is worth mentioning the Exp3$^{++}$ algorithm, as a recent variant of Exp3, proposed by \cite{Seldin17},
-that uses an adaptive tuning of its parameters.
-It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+A short proof of this results and more details on Exp3 are given in \cite{Bubeck12}.
+%
+Many other policies have been proposed for adversarial bandits, like Follow-the-Perturbed-Leader or Follow-the-Regularized-Leader for linear adversarial bandits, but we do not study them in this thesis, and we refer to Parts~III and VI of \cite{LattimoreBanditAlgorithmsBook} for more details.
 
@@
 As explained before, in the last few years, the MAB research community has been interested in finding algorithms which achieves both a problem-dependent logarithmic and a minimax upper-bounds.
-The first example is MOSS from \cite{Audibert2009minimax}, or \KLUCBpp{} in \cite{Menard17}.
-They are both index-based (see Algorithm~\ref{algo:2:indexPolicy}), using the following indexes:
+We present here some solutions that are index-based (see Algorithm~\ref{algo:2:indexPolicy}).
+The first example is MOSS from \cite{Audibert2009minimax}, or \KLUCBpp{} in \cite{Menard17},
+and they use the following indexes:
 \begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp}
@@
     U^{\mathrm{MOSS}+}_k(t) &= \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
-    U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(T / (K N_k(t)))}{N_k(t)} \right\} \\
+    U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
     U^{\klUCB\text{-}\mathrm{switch}}_k(t) &= \begin{cases}
@@
 
+% https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
+Finally, it is worth mentioning the Exp3$^{++}$ algorithm, as a recent variant of Exp3, proposed by \cite{Seldin17},
+that uses an adaptive tuning of its parameters.
+It is anytime, and it was also proven to obtain good ``best-of-both-world'' performance.
+% meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+
 % \subsection{A first numerical experiment on the example Bernoulli problem}
@@
 We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the estimated regret\footnote{~First, the rewards are computed as $\frac{1}{N} \sum_{j=1}^N [ \sum_{t=1}^T r^{(j)}_{A(t)}(t) ]$, where $r^{(j)}_{k}(t)$ means the reward stream of the different runs $j=1,\dots,N$.
-Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N [ \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t) ]$ (as it could be negative), but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N [ \sum_{k=1}^K \mu_k N^{(j)}_{k}(T) ]$.
-We noted in a remark after the proof of Lemma~\ref{lem:2:RegretDecomposition} (in page~\pageref{remark:2:moreAccurateCountofRegretForSimulations}) that this later formula is a more accurate empirical estimate of the regret.}
+Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N [ \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t) ]$, as this first formula could give a negative regret and is more noisy, but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N [ \sum_{k=1}^K \mu_k N^{(j)}_{k}(T) ]$, as this later formula gives a more robust estimator, as noted the proof of Lemma~\ref{lem:2:RegretDecomposition} (in page~\pageref{remark:2:moreAccurateCountofRegretForSimulations} above).}
 of the different algorithms, \emph{for one simulation}.
-Then in Table~\ref{table:2:meanResults} we show the results of the average on $N=1000$ independent runs.
+%
+It shows that the performance can be \textcolor{red}{very random}, as on both cases the pure exploitation strategy was very lucky and obtains close-to-optimal performance, even though we proved that it suffers from a linear mean regret.
 
@@
         \end{tabular}
-        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{for only one run of the simulation}, for different algorithms.
-\textcolor{red}{It is very random}, see Table~\ref{table:2:meanResults} for average results on $N=1000$ runs, on both cases the pure exploitation was very lucky and obtains close-to-optimal performance.}
+        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{for only one run of the simulation}, for different algorithms.}
         \label{table:2:oneShotResult}
@@
 \end{table}
+
+Then in Table~\ref{table:2:meanResults} we show the results of the average on $N=1000$ independent runs.
+%
+%
+We also include in Figures
+% \ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000},
+\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regret as functions of the current time step $t$, for $T=10000$ (and $N=1000$).
+The reward of the best arm is also plotted in Figure~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, and it serves as an \emph{upper-bound} to illustrate convergence\footnote{~Conversely, we also add the Lai \& Robbins' logarithmic \emph{lower-bound} on regret ($C_I \log(t)$ from case 2 of Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plot in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
+It should not be surprising that the black line of the lower-bound is \emph{above} the values of the regret as the lower-bound is only asymptotic.}
+of the average cumulated reward (\ie, $\frac{1}{t} \sum\limits_{s=1}^t r_{A(s)}(s)$), which should converge to $\mu^*$ when $t\to\infty$ (or $t\toT$) for efficient algorithms.
+%
+We can clearly identify the behavior of the different strategies:
+the three efficient strategies indeed achieve a sub-linear regret, that looks logarithmic as supported by the theory,
+while the naive strategies achieve linear regret.
+The two simple strategies also achieve sub-linear regret but are less efficient:
+both $\varepsilon$-greedy and Explore-then-Commit starts as pure exploration, respectively until $\varepsilon_0/t$ becomes smaller than $1$ and until $t > T_0$, then they usually identify correctly the best arm and start to catch up with the best policies.
 
@@
 
-We also include in Figures
-% \ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000},
-\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regret as functions of the current time step $t$, for $T=10000$ for $N=1000$ independent runs.
-The reward of the best arm is also plotted in Figure~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, to illustrate convergence of the average reward $\frac{1}{t} \sum\limits_{s=1}^t r_{A(s)}(s)$, which should converge to $\mu^*$ when $t\to\infty$ (or $t\toT$) for efficient algorithms.
-Conversely, we also add the Lai \& Robbins' logarithmic lower-bound on regret ($C_I \log(t)$ from case 2 of Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plot in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
-%
-We can clearly identify the behavior of the different strategies:
-the three efficient strategies indeed achieve a sub-linear regret, that looks logarithmic as supported by the theory,
-while the naive strategies achieve linear regret.
-The two simple strategies also achieve sub-linear regret but are less efficient:
-both $\varepsilon$-greedy and Explore-then-Commit starts as pure exploration, respectively until $\varepsilon_0/t$ becomes smaller than $1$ and until $t > T_0$, then they almost always correctly identify the best arm and start to catch up with the best policies.
-
 \TODOL{Décrire les résultats !}
@@
 and some of them are known to be optimal in different settings.
+%
+To clearly understand the different in their empirical behaviors, we illustrated the performance of two naive and two simple strategies against three efficient ones, on the example of Bernoulli-distributed problem used for the online demonstration from Section~\ref{sec:2:notations} (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
+This also serves as a first example of numerical simulation performed with our library SMPyBandits, before the large-scale experiments of Section~\ref{sec:3:reviewSPAlgorithms} below.
 %
@@
 
-\newpage  % FIXME
+\newpage  % WARNING ?
 

### 2019-06-05 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
I'm starting to be happy about Chapter:2

@@
 %
-A sequential change-point detector is a stopping time $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
+A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
 $\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N^*, \bE[X_i] = \mu_0\right)$.

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Finishing to proof read Ch2

@@
 %
-The notations used in this thesis are the usual notations in all the MAB literature, and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
-The Nomenclature includes a summary of the notations, see Page~\pageref{chapter:nomenclature}.
+We use in this thesis the usual notations from the MAB literature,
+% , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
+and a summary is included in the Nomenclature, in Page~\pageref{chapter:nomenclature}.
 
@@
 A MAB is a model of a decision-making game, where a \emph{player} has to sequentially select one action in a (usually finite) set of actions, and only receives a (random) feedback about the selected action, also called a \emph{reward}.
-We consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$.
-When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r_{A(t)}(t)\in\R$, and so on for a finite number of steps $t$, from $t=1$ until $t=T$ for an horizon $T$.
+\textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
+When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r_{A(t)}(t)\in\R$, and so on for a finite number of steps $t$, from $t=1$ until $t=T$ for an \emph{horizon} $T$.
 %
 A commonly studied goal for the player is to maximize its sum of received rewards, $\sum_{t=1}^T r_{A(t)}(t)$
-(or its mean in stochastic and stationary models).
+(or its \emph{mean} in stochastic and stationary models).
 
 The difficulty of this decision-making game comes from balancing the trade-off between \emph{exploration}, because the player has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
-The MAB model is a famous example of a reinforcement learning model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
+The MAB model is a famous example of a \textbf{reinforcement learning} model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
 We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
@@
 First, the rewards may depend on the randomness of the environment, \ie, the unknown and unpredictable values $r_k(t)$.
-Then this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the past observations, \ie, $A(1)$, $r_{A(1)}(1)$, $\dots$, $A(t-1)$, $r_{A(t-1)}(t-1)$, and possibly on an external source of randomness.
+Then this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the \emph{past} observations, \ie, $A(1)$, $r_{A(1)}(1)$, $\dots$, $A(t-1)$, $r_{A(t-1)}(t-1)$, and possibly on an external source of randomness.
 %
@@
 %
-The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated probs'' line.
+The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated Probs'' line.
 
@@
 %
-In the second Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} below, we display the result of an example of run, when the player was following the $\UCB_1$ algorithm from \cite{Auer02}.
+In the second Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} below, we display the result of an example of run, when the player was following the $\UCB_1$ algorithm from \cite{Auer02} (see Section~\ref{sub:2:IndexPolicies}).
 After $T=100$ steps, the player obtained a cumulated reward of $56$, by playing mostly arms $4,3,5,1,0$ (in decreasing order of number of plays).
 The empirical means $\widehat{\mu_k}(T)$ correctly identify the best arm (arm $4$), but do not correctly rank the arms as arms $1$ and $3$ obtained means of $\widehat{\mu_1}(T) = 0.5 < \widehat{\mu_3}(T)=0.6$ but the true means are $\mu_3 = 0.55 < \mu_1 = 0.6$.
+Other examples of such results, for different algorithms, and $T=100$ or $T=10000$, are given in Section~\ref{sub:2:shortNumericalExperiments}.
 
@@
     \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
-    \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.
-See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
+    \caption{Screenshot of the demonstration, at the end of the game after $T=100$ steps, where the player suffers from an empirical regret of $R^{\UCB_1}_T = \mu^* T - 56 = 0.7 T - 56 = 14$ after following the $\UCB_1$ algorithm.
+        You can try it on \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
     \label{fig:2:example_of_a_5_arm_bandit_problem__step100}
@@
 The algorithm $\cA$ selects an arm $A(t)$ at time $t$, possibly by using the past observations and the past external randomness.
-As shown below, being oblivious to the past yields very poor performance (the pure exploration policy is discussed in Section~\ref{sub:2:naiveSimpleStrategies}), so efficient policies indeed depend on the environment feedback.
+As shown below, being oblivious to the past yields very poor performance (cf.
+the pure exploration policy in Section~\ref{sub:2:naiveSimpleStrategies}), so efficient policies indeed depend on the successive feedbacks.
 %
@@
 The initial information is reduced to $O_1 = (U_0)$, and the first decision is $A(1) = \cA_1(O_1)$.
-Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps, as illustrated for example in Algorithm~\ref{algo:2:indexPolicy} below.
-%
-An algorithm is said to be deterministic if it does not depend on the external randomness $U_0,U_1,\dots$, but in this thesis we only use non-deterministic algorithms.
+Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps (\eg, Algorithm~\ref{algo:2:indexPolicy}).
+% as illustrated for example in Algorithm~\ref{algo:2:indexPolicy} below.
+%
+An algorithm is said to be \emph{deterministic} if it does not depend on the external randomness $U_0,U_1,\dots$, but in this thesis \textbf{we only use non-deterministic algorithms}
+(in particular, index policies need $U_t$ to break ties in the $\argmax$, see Algorithm~\ref{algo:2:indexPolicy} below).
 
@@
 In the example above, we consider Bernoulli distributions, but other real-valued distributions have been studied in the literature.
-From now on and until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
+From now on and until the last chapter of this thesis, \textbf{we only consider stochastic rewards}. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
 %
@@
 % https://smpybandits.github.io/docs/Arms.html
-We also focus only on real-valued rewards, meaning that $r_k(t)\in\R$ for all arm $k$ and time $t$.
+\textbf{We also focus only on real-valued rewards}, meaning that $r_k(t)\in\R$ for all arm $k$ and time $t$.
 
@@
 
-In all this thesis we restrict to distributions that have a certain form, and focus on real-valued distributions lying in an exponential families.
+In all this thesis we restrict to distributions that have a certain form, and \textbf{we focus either on real-valued distributions lying in an exponential families or bounded distributions}.
 One-dimensional exponential families include Bernoulli and Poisson distributions, as well as Gaussian distributions with a fixed variance.
-Our main interest is Bernoulli distributions, but we prefer to present the notations of exponential families in one dimension.
+Our main interest is Bernoulli distributions, but we prefer to present the more general notations of exponential families in one dimension.
 % We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Emilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
@@
 for a parameter vector $\eta$ (the canonical parameter), and a sufficient statistic that we restrict to be $T(x)=x$, and a function $h$.
-$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$ and $h$:
-$A(\eta) = \log \left( \int h(x) \exp(\eta x) \lambda(\mathrm{d} x) \right)$.
+$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$ and $h$,
+as $A(\eta) = \log \left( \int h(x) \exp(\eta x) \lambda(\mathrm{d} x) \right)$.
 %
@@
     \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
-        \KL\left(d_1, d_2\right) = \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right].
+        \KL\left(d_1, d_2\right) = \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}\cup\{\pm\infty\}.
     \end{equation}
@@
 %
-\begin{itemize}
+\begin{itemize}\label{par:2:notationsExponentialFamiliesBernoulliGaussian}
     \item
@@
     and it is also called the \emph{relative binary entropy}.
+    It satisfies $\kl(x,y) \geq d_{1/4}(x,y)$, see Lemma~\ref{lem:2:Pinsker} in Appendix~\ref{proof:2:Pinsker}.
 
@@
     Moreover, the KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
-    $d(x,y) = \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$.
+    $d_{\sigma^2}(x,y) = \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$.
 \end{itemize}
@@
     because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$, which are measurable wrt to the past observations $O_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$.
-    Thus we can rewrite the expected cumulated rewards as follows:
+    Thus we can rewrite the expected cumulated reward as follows:
     %
@@
 %
-\begin{proof}
+\begin{proof}\label{proof:2:worstCaseLowerBound}
     For an example, consider any large horizon $T>10$, and $K=2$ Bernoulli arms.
@@
 %
-\begin{proof}
+\begin{proof}\label{proof:2:tuningExploreThenCommit}
     Let us prove the inequality $p \leq 2 K \exp(-T_0 d^2 / 4)$.\\
@@
 \subsection{Index policies: $\UCB_1$, \klUCB{} and others}
+\label{sub:2:IndexPolicies}
 
@@
 the \emph{average reward} $\widehat{\mu_k}(t)=\frac{X_k(t)}{N_k(t)}$
-and a \emph{confidence radius} $\xi_k(t) = \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}\right)}$.
+and a \emph{confidence radius} $\xi_k(t) = \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
 %
 \begin{align}\label{eq:UCB_index}
-    U_k^{\UCB_1}(t) = \UCB_k(t) = \underbrace{\frac{X_k(t)}{N_k(t)}}_{\widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}\right)}}_{\xi_k(t)}.
+    U_k^{\UCB_1}(t) = \UCB_k(t) = \underbrace{\frac{X_k(t)}{N_k(t)}}_{\widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\xi_k(t)}.
 \end{align}
@@
 %
-\begin{proof}
+\begin{proof}\label{proof:2:UCBregretBound}
     \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ?}
-    As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref\label{lem:2:HoeffdingInequality}.
+    As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
     Thanks to the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
@@
 \ie, $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$,
-because Pinkser's inequality gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix),
-and so XXX
+because Pinkser's inequality gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix).
+% FIXME is this enough?
 
@@
 \subsection{Bayesian policies}
+\label{sub:2:BayesianPolicies}
 
@@
 \subsection{Other policies}
+\label{sub:2:OtherPolicies}
 
@@
 
-We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the estimated regret\footnote{~First, the rewards are computed as $\frac{1}{N} \sum_{j=1}^N [ \sum_{t=1}^T r^{(j)}_{A(t)}(t) ]$, where $r^{(j)}_{k}(t)$ means the reward stream of the different runs $j=1,\dots,N$.
-Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N [ \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t) ]$, as this first formula could give a negative regret and is more noisy, but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N [ \sum_{k=1}^K \mu_k N^{(j)}_{k}(T) ]$, as this later formula gives a more robust estimator, as noted the proof of Lemma~\ref{lem:2:RegretDecomposition} (in page~\pageref{remark:2:moreAccurateCountofRegretForSimulations} above).}
+We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the estimated regrets\footnote{~First, the rewards are computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{A(t)}(t)$, where $r^{(j)}_{k}(t)$ means the reward stream of the different runs $j=1,\dots,N$.
+Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t)$, as this first formula could give a negative regret and is more noisy, but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N \sum_{k=1}^K \mu_k N^{(j)}_{k}(T)$, as this later formula gives a more robust estimator, as noted the proof of Lemma~\ref{lem:2:RegretDecomposition} (in Page~\pageref{remark:2:moreAccurateCountofRegretForSimulations} above).}
 of the different algorithms, \emph{for one simulation}.
@@
 % \ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT100N1000},
-\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regret as functions of the current time step $t$, for $T=10000$ (and $N=1000$).
+\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regrets as functions of the current time step $t$, for $T=10000$ (and $N=1000$).
 The reward of the best arm is also plotted in Figure~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, and it serves as an \emph{upper-bound} to illustrate convergence\footnote{~Conversely, we also add the Lai \& Robbins' logarithmic \emph{lower-bound} on regret ($C_I \log(t)$ from case 2 of Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plot in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
 It should not be surprising that the black line of the lower-bound is \emph{above} the values of the regret as the lower-bound is only asymptotic.}
-of the average cumulated reward (\ie, $\frac{1}{t} \sum\limits_{s=1}^t r_{A(s)}(s)$), which should converge to $\mu^*$ when $t\to\infty$ (or $t\toT$) for efficient algorithms.
+of the average rewards (\ie, $\frac{1}{t} \sum\limits_{s=1}^t r_{A(s)}(s)$), which should converge to $\mu^*$ when $t\to\infty$ for efficient algorithms.
 %
@@
 
-% % FIXME generate at the end of the simulation
 % \begin{figure}[h!]  % [htbp]
@@
 % \end{figure}
-
-% % FIXME generate at the end of the simulation
+%
 % \begin{figure}[h!]  % [htbp]
@@
 
-% FIXME generate at the end of the simulation
 \begin{figure}[h!]  % [htbp]
 	% \centering
-	\includegraphics[width=1.00\linewidth]{meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000.pdf}
-	\caption{Mean cumulated rewards as function of $t$ for $T=10000$ and $N=1000$.}
+    \includegraphics[width=1.20\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
+	\caption{Average of rewards, as function of $t$, for $T=10000$ and $N=1000$.}
 	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
@@
 
-% FIXME generate at the end of the simulation
 \begin{figure}[h!]  % [htbp]
 	% \centering
-	\includegraphics[width=1.00\linewidth]{meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000.pdf}
+	\includegraphics[width=1.20\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
 	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
@@
 \end{figure}
-
-\TODOL{Décrire les résultats !}
 
@@
 
-Pinsker's inequality is proven below, using simple real analysis arguments\footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).}.
-%
-This Lemma is again used in Chapter~\ref{chapter:6}.
+We prove it below using simple real analysis arguments\footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).},
+%
+and for instance we use it again in Chapter~\ref{chapter:6}.
+%
+Moreover, using the notations of the exponential families for Bernoulli and fixed-variance Gaussian distributions (see above Section~\ref{par:2:notationsExponentialFamiliesBernoulliGaussian}), it can be written $\kl(x,y) \geq d_{1/4}(x,y)$.
+%
+As bounded distributions in $[0,1]$ are sub-Bernoulli as well as $1/4$-sub-Gaussian, one consequence of this inequality is that the confidence radius of \klUCB{} is \emph{smaller} than the one in $\UCB_1$, and different concentration inequalities can then be used to prove their finite-time logarithmic regret upper-bounds, and to prove that \klUCB{} is asymptotically optimal while $\UCB_1$ is only order-optimal for bounded rewards (see above in Section~\ref{proof:2:UCBregretBound} or \cite{Auer02} for $\UCB_1$, and \cite{KLUCBJournal} for \klUCB).
 
 \begin{lemma}[Pinsker's inequality]\label{lem:2:Pinsker}
-    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as
-    \begin{equation}\label{eq:2:Pinsker1}
-        \kl(p,q) = p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right).
-    \end{equation}
+    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) = p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
     Then for any $p,q\in[0,1]$, we have
@@
 \begin{proof}
-    Fix $p\in[0,1]$, and consider the function $g(x) := \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
+    Fix $p\in[0,1]$, and consider the function $g(x) = \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
     We first observe that $g(0) = 0$, and we easily prove that $x=0$ is the unique minimizer of $g$ over its domain.
@@
     Simple calculations yield
-    \[ \left(\frac{\partial \; \kl}{\partial 2}\right) (p, p + x) = - \frac{p}{p+x} + \frac{1-p}{1-(p+x)} =  = -x \frac{(2(p + x) - 1)^2}{(p + x)(p + x - 1)}. \]
+    \[ \left(\frac{\partial \; \kl}{\partial 2}\right) (p, p + x) = - \frac{p}{p+x} + \frac{1-p}{1-(p+x)} = -x \frac{(2(p + x) - 1)^2}{(p + x)(p + x - 1)}. \]
     % https://www.wolframalpha.com/input/?i=diff(p+*+log(p%2F(p%2Bx))+%2B+(1-p)+*+log((1-p)%2F(1-(p%2Bx)))+-+2*x**2,+x)

### 2019-06-05 2-Chapters/1-Chapter/chapter1.tex
Je viens de faire un premier jet (brouillon)
- de l'abstract FR/EN
- du résumé en français
Starting to try to add meta data

@@
 % ----------------------------------------------------------------------------
-\section{List of publications}
+\section{List of publications and other experiences}
 \label{sec:1:listPublications}
@@
 
-\paragraph{Seminars}
+\paragraph{Seminars.}
     I gave some presentations in the following events:
@@
 
-\paragraph{Tutorial}
+\paragraph{Tutorial.}
     I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
@@
 
-\paragraph{Training}
+\paragraph{Training.}
     I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
@@
 
-\paragraph{Conferences}
+\paragraph{Conferences.}
     I attended the following conferences:
@@
 
-\paragraph{Seminars}
+\paragraph{Seminars.}
     I attended the following seminars:
@@
 
-\paragraph{Responsabilities}
+\paragraph{Responsabilities.}
     I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
@@
 
-\paragraph{Reviews}
+\paragraph{Reviews.}
 	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
@@
 
-\paragraph{System Administrator}
+\paragraph{System Administrator.}
 	maintaining our workstations running GNU/Linux and Windows
@@
 	from January $2017$ to September $2018$.
+
+% WARNING
+\vfill{}
+
+\paragraph{Copyright notice.}
+%
+This document and the additional resources required to compile it (including \LaTeX{} code, Python snippets, images etc)
+are \href{https://github.com/Naereen/phd-thesis/}{publicly published},
+online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}},
+under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}}.
+
+\TODOL{Open source the repository as soon as I defended my thesis!}
+
+\begin{center}
+    \textbf{Copyright 2016-2019, \copyright ~Lilian~Besson.}
+\end{center}

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Changed \paragraph{XXX} to \paragraph{XXX.}

@@
     \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
-    \caption{Screenshot of the demonstration available online on my website, for a current step of $t=24$.}
+    \caption[Screenshot of the demonstration for a current step of $t=24$.]{Screenshot of the demonstration available online on my website, for a current step of $t=24$.}
     \label{fig:2:example_of_a_5_arm_bandit_problem}
@@
     \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
-    \caption{Screenshot of the demonstration, at the end of the game after $T=100$ steps, where the player suffers from an empirical regret of $R^{\UCB_1}_T = \mu^* T - 56 = 0.7 T - 56 = 14$ after following the $\UCB_1$ algorithm.
+    \caption[Screenshot of the demonstration, at the end of the game after $T=100$ steps]{Screenshot of the demonstration, at the end of the game after $T=100$ steps, where the player suffers from an empirical regret of $R^{\UCB_1}_T = \mu^* T - 56 = 0.7 T - 56 = 14$ after following the $\UCB_1$ algorithm.
         You can try it on \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
@@
 % Just bounded, in $[a,b]$ but without loss of generality in $[0,1]$ etc.
-\paragraph{The normalization trick.}\label{par:2:normalizationTrick}
+\paragraph{The normalization trick.}\label{par:2:normalizationTrick.}
 %
@@
 
-In all this thesis we restrict to distributions that have a certain form, and \textbf{we focus either on real-valued distributions lying in an exponential families or bounded distributions}.
+In all this thesis we restrict to distributions that have a certain form, and we focus either on real-valued distributions lying in an exponential families or bounded distributions.
 One-dimensional exponential families include Bernoulli and Poisson distributions, as well as Gaussian distributions with a fixed variance.
@@
 %
-In the rest of this thesis, \emph{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the considered applications.
+In the rest of this thesis, \textbf{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the considered applications.
 First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
@@
 
-\TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira...}
+\TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira..}
 In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
@@
     & = \E\left[ \sum_{t=1}^T (r_{k^*}(t) - r_{A(t)}(t)) \right]
-    = \E\left[ \sum_{t=1}^T r_{k^*}(t) \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right] \\
+    = \E\left[ \sum_{t=1}^T \underbrace{r_{k^*}(t)}_{\text{\iid{} from $\nu_{k^*}$}} \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right] \\
     & = T \times \E\left[ r_{k^*}(t=1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]
@@
 
-\paragraph{Naive strategies: pure exploitation or pure exploration}
+\paragraph{Naive strategies: pure exploitation or pure exploration.}
 
@@
 % ---------------------------------------------------
-\paragraph{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit}
+\paragraph{Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit.}
 
@@
 
-% \paragraph{Outline}
+% \paragraph{Outline.}
 % This section is organized as follows.

### 2019-06-05 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Commence à intégrer changements d'Emilie, cf. #13
Add Figures for Appendix of Chapter4
Done changing notations \{1,\dots,M\} -> [M] all over the PhD
Changed \paragraph{XXX} to \paragraph{XXX.}

@@
-\subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
-\label{sec:5:proofLagrangian}
+% \subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
+% \label{sec:5:proofLagrangian}
 
-This short subsection contains the missing details of the proof of Proposition \ref{prop:41:Lagrangian}.
+We include here the missing details of the proof of Proposition~\ref{prop:41:Lagrangian}.
 %
@@
 
-\paragraph{Quasi-convexity}
+\paragraph{Quasi-convexity.}
 
@@
     The function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{K}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
-    \hfill{}$\square$
 \end{itemize}
 
-\paragraph{Derivation of the Lagrange multiplier solution}
+\hfill{}$\square$
+
+\paragraph{Derivation of the Lagrange multiplier solution.}
 
 Let us now prove the expression for $D_i^*$ given in \eqref{eq:41:Dilambda}.
+
 \begin{itemize}
@@
     and so the unique solution to the maximization problem \eqref{eq:41:optPb}, thanks to \cite[Theorem 1]{Luenberger68}.
+\end{itemize}
+
 \hfill{}$\square$
-\end{itemize}

### 2019-06-05 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Typos in refs and labels
Changed \paragraph{XXX} to \paragraph{XXX.}

@@
 
-\paragraph{Outline}
+\paragraph{Outline.}
 
@@
 First the system model is introduced in Section~\ref{sub:43:model},
-% A formal description of the MAB learning algorithms is given in Section~\ref{sub:43:motivations},
-and our motivations are exposed in Section~\ref{sub:43:MABalgo}.
+and our motivations are exposed in Section~\ref{sub:43:motivations}.
 The proposed \UCB-based heuristics are presented in Section~\ref{sub:43:heuristics}, while the corresponding numerical results are shown in Section~\ref{sub:43:numExp}.
@@
 % ----------------------------------------------------------------------
-\paragraph{LPWA network}
+\paragraph{LPWA network.}
 
@@
 % ----------------------------------------------------------------------
-\paragraph{Model of IoT devices}
+\paragraph{Model of IoT devices.}
 
@@
 
-\paragraph{Probability of collision at a second transmission slot}
+\paragraph{Probability of collision at a second transmission slot.}
 
@@
 
-\paragraph{Learning is useful for non-congested networks}
+\paragraph{Learning is useful for non-congested networks.}
 
@@
 % % % --- --- --- --- --- --- --- --- --- ---
-% \paragraph{The \UCB{} algorithm}\label{sub:43:algoUCB}
+% \paragraph{The \UCB{} algorithm.}\label{sub:43:algoUCB}
 
@@
 
-% \paragraph{First stage \UCB}\label{sub:43:firstStageUCB}
+% \paragraph{First stage \UCB.}\label{sub:43:firstStageUCB}
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Uniform random retransmission (``Random'')}\label{sub:43:UCBthenRandom}
+\paragraph{Uniform random retransmission (``Random'').}\label{sub:43:UCBthenRandom}
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{\UCB{} for retransmission (``Only \UCB{}'')}\label{sub:43:TwoUCB}
+\paragraph{\UCB{} for retransmission (``Only \UCB{}'').}\label{sub:43:TwoUCB}
 
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'')}\label{sub:43:UCBwithDelay}
+\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'').}\label{sub:43:UCBwithDelay}
 
@@
 
-\paragraph{Note on the simulation code}
+\paragraph{Note on the simulation code.}
 %

### 2019-06-05 1-Preamble/5-Abstract/resume_fr.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Fixed typos in some references

@@
     Nous présentons les concepts et les notations du problème des bandits à armes multiples dans le chapitre~\ref{chapter:2}, d'un point de vue mathématique.
-    Mais nous suivons également une approche didactique puisque nous utilisons une démonstration interactive en ligne conçue pour permettre à n'importe qui de jouer contre un petit problème de bandit depuis son navigateur, dans Section~\ref{par:2:interactiveDemoDemoDiscoverMAB}.
+    Mais nous suivons également une approche didactique puisque nous utilisons une démonstration interactive en ligne conçue pour permettre à n'importe qui de jouer contre un petit problème de bandit depuis son navigateur, dans Section~\ref{par:2:interactiveDemoDiscoverMAB}.
 
@@
     }
-    \caption{Organisation de la thèse : une carte de lecture.
+    \caption[Organisation de la thèse : une carte de lecture]{Organisation de la thèse : une carte de lecture.
 Tout chemin contenant le Chapitre~\ref{chapter:1}, le Chapitre~\ref{chapter:2}, au moins un des trois Chapitres~\ref{chapter:4}, \ref{chapter:5} et \ref{chapter:6}, et la Conclusion est une façon valide de lire cette thèse.}
-    \label{fig:1:organization}
+    \label{fig:1:organization_fr}
 \end{figure}
@@
 L'ordre de lecture du manuscrit peut être n'importe quel chemin descendant entre l'introduction dans le chapitre actuel~\ref{chapter:1}, et la conclusion dans le dernier chapitre~\ref{chapter:conclusion}.
-Comme le montre le graphique de la Figure~\ref{fig:1:organisation},
+Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,
 la thèse est organisée en deux parties :

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Fixed typos in some references

@@
 % Just bounded, in $[a,b]$ but without loss of generality in $[0,1]$ etc.
-\paragraph{The normalization trick.}\label{par:2:normalizationTrick.}
+\paragraph{The normalization trick.}
+\label{par:2:normalizationTrick}
 %
@@
 %  either in an alternative way or in a fixed way (``explore then exploit'').
-They are detailed in Algorithm~\ref{algo:2:naiveStrategies}.
+They are detailed in Algorithm~\ref{algo:2:simpleStrategies}.
 
@@
 		\caption{Simple efficient strategies: \textcolor{deeppurple}{$\varepsilon$-greedy} and \textcolor{deepgold}{Explore-then-Exploit}.}
-		\label{algo:2:naiveStrategies}
+		\label{algo:2:simpleStrategies}
 		% \end{small} % XXX remove if needed
@@
 \begin{proof}\label{proof:2:UCBregretBound}
-    \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ?}
+    \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ? si c'est bien fait, ça peut aider la lecture, par exemple \textcolor{darkgreen}{$k$} et \textcolor{gold}{$k^*$}.}
+    %
     As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.

### 2019-06-05 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Fixed typos in some references

@@
 
-\paragraph{One gateway and many devices}
+\paragraph{One gateway and many devices.}
 %
@@
 
-\paragraph{Slotted protocol}
+\paragraph{Slotted protocol.}
 %
@@
 
-\paragraph{Static vs dynamic devices}
+\paragraph{Static vs dynamic devices.}
 %
@@
 
-\paragraph{Random emission patterns}
+\paragraph{Random emission patterns.}
 %
@@
 
-\paragraph{Some assumptions on the network occupation}
+\paragraph{Some assumptions on the network occupation.}
 %
@@
 
-\paragraph{MAB model}
+\paragraph{MAB model.}
 %
@@
 
-\paragraph{A bandit model for IoT}
+\paragraph{A bandit model for IoT.}
 
@@
 
-\paragraph{Two algorithms}
+\paragraph{Two algorithms.}
 
@@
 
-\paragraph{\emph{Adversarial} bandit algorithms?}
+\paragraph{\emph{Adversarial} bandit algorithms?.}
 
@@
     % FIXME use a shorter caption for the list of figures in Appendix of the thesis: it takes lots of lines to have these (a)...(d) shown in list of figures
-    \caption[Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
+    \caption[XXX Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:from10to100}
@@
 
-\paragraph{First results}
+\paragraph{First results.}
 
@@
 
-\paragraph{Successful transmissions rate as a function of the dynamic devices}
+\paragraph{Successful transmissions rate as a function of the dynamic devices.}
 
@@
 
-\paragraph{A safety check}
+\paragraph{A safety check.}
 
@@
 
-\paragraph{Note on the simulation code}
+\paragraph{Note on the simulation code.}
 

### 2019-06-05 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Fixed a bug of using subfloat and having long captions in the List of Figures.

@@
 % \subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
-\paragraph{Notations}\label{sub:25:aggregation}
+\paragraph{Notations.}\label{sub:25:aggregation}
 
@@
 % ----------------------------------------------------------------------
-\paragraph{Explanations of \Aggr{} versus \ExpQ{} }\label{sub:25:Exp4}
+\paragraph{Explanations of \Aggr{} versus \ExpQ{} .}\label{sub:25:Exp4}
 
@@
 The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted.
-It corresponds to the second point \eqref{eq:2:forSecondLogTLowerBound2} in Theorem~\ref{thm:2:secondLogTLowerBound}.
+It corresponds to the second case in Theorem~\ref{thm:2:secondLogTLowerBound}.
 It is crucial to note that the lower-bound is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli}).
@@
 % ----------------------------------------------------------------------
-\paragraph{No theoretical guarantees}\label{sub:25:theory}
+\paragraph{No theoretical guarantees.}\label{sub:25:theory}
 
@@
 % ----------------------------------------------------------------------
-\paragraph{Conclusion about our \Aggr{} algorithm}\label{sub:25:conclusion}
+\paragraph{Conclusion about our \Aggr{} algorithm.}\label{sub:25:conclusion}
 

### 2019-06-05 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Fixed a bug of using subfloat and having long captions in the List of Figures.

@@
     \centering
-    \subfloat[10\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{10intelligent.eps}
+    \subfloat[][10\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{10intelligent.eps}
     \label{fig:41:10intelligent}}
     %
-    \subfloat[30\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{30intelligent.eps}
+    \subfloat[][30\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{30intelligent.eps}
     \label{fig:41:30intelligent}}
@@
     \vspace*{-10pt}
-    \subfloat[50\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{50intelligent.eps}
+    \subfloat[][50\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{50intelligent.eps}
     \label{fig:41:50intelligent}}
     %
-    \subfloat[100\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{100intelligent.eps}
+    \subfloat[][100\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{100intelligent.eps}
     \label{fig:41:100intelligent}}
-    % FIXME use a shorter caption for the list of figures in Appendix of the thesis: it takes lots of lines to have these (a)...(d) shown in list of figures
-    \caption[XXX Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
+    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:from10to100}
@@
     \centering
-    \subfloat[10\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_10.eps}
+    \subfloat[][10\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_10.eps}
     \label{fig:41:ch2_10}}
     %
-    \subfloat[30\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_30.eps}
+    \subfloat[][30\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_30.eps}
     \label{fig:41:ch2_30}}
@@
     \vspace*{-10pt}
-    \subfloat[50\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_50.eps}
+    \subfloat[][50\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_50.eps}
     \label{fig:41:ch2_50}}
     %
-    \subfloat[100\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_100.eps}
+    \subfloat[][100\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_100.eps}
     \label{fig:41:ch2_100}}

### 2019-06-05 2-Chapters/5-Chapter/chapter5.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Fixed a bug of using subfloat and having long captions in the List of Figures.

@@
 
-\paragraph{Abstract}
+\paragraph{Abstract.}
 %
@@
 
-\paragraph{Publications}
+\paragraph{Publications.}
 
@@
 We give the mean regret $\pm$ $1$ std-dev, and more work is needed on our implementation on Improved \MusicalChair.
-The results on \textsc{Sic-MMAB} appear to confirm the numerical experiments in the paper \cite{BoursierPerchet18}.}
+The results on \textsc{Sic-MMAB} appear to confirm the numerical experiments of \cite{BoursierPerchet18}.}
     \label{table:5:meanRegretSimulationsNoSensing}
@@
     \end{tabular}
-    \caption{Comparing \RhoRand{} and \rhoLearn{} on a simple MP bandit problem with $K=9$ arms.}
+    \caption{Comparing \RhoRand{} and \rhoLearn{} on a simple MP-MAB problem with $K=9$ arms.}
     \label{table:5:comparisonRhoRandRhoLearn}
@@
 
-% \paragraph{Acknowledgments}
+% \paragraph{Acknowledgments.}
 % Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Starting to conclude Ch2

@@
 
-\TODOL{Abstract ici et plus court, et petite introduction ?}
+\paragraph{Abstract.}
+%
+In this chapter, we present the common base of all the models studied in this thesis: the multi-armed bandit (MAB) model, restricting to the single-player and stationary stochastic case.
+We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a one-dimensional distribution.
+%
+We formalize this MAB model, then we quote some of its important applications.
+We define the notion of regret of an algorithm, and quote important results from the literature. A short review of the most important families of MAB algorithms is then given.
+We conclude this chapter by discussing the question of online algorithm selection.
 
@@
 
-\paragraph{Abstract.}
-%
-In this chapter, we present the common base of all the models studied in this thesis:
-the multi-armed bandit (MAB) model,
-restricting to the single-player and stationary stochastic case.
-We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a one-dimensional distribution.
-In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
-%
-We quickly review the possible applications of bandits, that range from recommender systems and information retrieval to healthcare and finance.
+% ----------------------------------------------------------------------------
+\section{Introduction}
+\label{sec:2:introduction}
+
+Multi-Armed Bandits (MAB) models were introduced by Thompson as early as in 1933 \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
+This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
+Their name refers to one-armed bandits found in casinos, as shown in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
+% The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
+% Since then, many efficient algorithms were proposed, XXX
+%
+% In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
+We formally define the model and the main notations used in all this document in Section~\ref{sec:2:notations}.
+%
+Multi-armed bandits have been applied to various fields, from recommender systems and information retrieval to healthcare and finance.
+We discuss some applications in Section~\ref{sec:2:applicationsofStochasticMAB}.
 In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits has already been applied with success in the last ten years.
 
-We then explain the main measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and for horizon $T$.
+% We then explain the main measure of performance used in this thesis, defined as the regret and written $R_T^{\cA}$ for algorithm $\cA$ and for horizon $T$.
 The regret compares the performance of the decision-making player and the oracle who knows in advance the best arm and always selects it.
-Other measures of performance, such as best-arm identification, are not used later on in this document.
-% , but they are also shortly discussed.
-%
-We explain the main theoretical result from previous works: a lower-bound on the regret of any algorithm, essentially showing that no algorithm can achieve sub-logarithmic regret on all problems, in this stochastic MAB model (\ie, $R_T^{\cA} = \Omega(\log(T))$).
-
-We give an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the $\UCB_1$, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
+% Other measures of performance, such as best-arm identification, are not used later on in this document.
+% % , but they are also shortly discussed.
+% %
+% We explain the main theoretical result from previous works: a lower-bound on the regret of any algorithm, essentially showing that no algorithm can achieve sub-logarithmic regret on all problems, in this stochastic MAB model (\ie, $R_T^{\cA} = \Omega(\log(T))$).
+
+A rich research literature has produced many different algorithms tackling the MAB problem, as suggested by the length of the reference book \cite{LattimoreBanditAlgorithmsBook}.
+We give in Section~\ref{sec:2:famousMABalgorithms} an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the $\UCB_1$, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
 % We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contributions of this thesis.
@@
 %
-We then give regret upper-bounds of the two main algorithms used in this thesis, $\UCB_1$ and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
+% We then give regret upper-bounds of the two main algorithms used in this thesis, $\UCB_1$ and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
 They are order-optimal as they attain a logarithmic regret.
@@
 
-Finally, we explore the problem of algorithm selection: as there are many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
+Finally, we explore in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} the problem of algorithm selection: as there are many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
 We present one of the first contributions made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms,
@@
 
-\paragraph{Publication.}
-%
-The first Sections~\ref{sec:2:notations}, \ref{sec:2:applicationsofStochasticMAB}, \ref{sec:2:lowerUpperBoundsRegret} and \ref{sec:2:famousMABalgorithms} of this chapter are based on existing works,
-and Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on our article \cite{Besson2018WCNC}.
+\paragraph{About notations.}
+%
+We use in this thesis the usual notations from the MAB literature,
+% , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
+and a summary is included in the Nomenclature, in Page~\pageref{chapter:nomenclature}.
 
@@
 
-\paragraph{About notations.}
-%
-We use in this thesis the usual notations from the MAB literature,
-% , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
-and a summary is included in the Nomenclature, in Page~\pageref{chapter:nomenclature}.
+\paragraph{Publication.}
+%
+The first Sections~\ref{sec:2:notations}, \ref{sec:2:applicationsofStochasticMAB}, \ref{sec:2:lowerUpperBoundsRegret} and \ref{sec:2:famousMABalgorithms} of this chapter are based on existing works,
+and Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on our article \cite{Besson2018WCNC}.
 
@@
 \label{sec:2:notations}
-
-Multi-Armed Bandits (MAB) models were introduced by Thompson as early as in 1933 \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
-This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
-Their name refers to one-armed bandits found in casinos, as showed in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
-% The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
-% Since then, many efficient algorithms were proposed, XXX
 
@@
 \begin{defn}[Kullback-Leibler divergence]\label{def:2:KLDivergence}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
     The KL divergence between two distributions $d_1$ and $d_2$ is defined as
@@
     \end{equation}
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{defn}
@@
 \begin{defn}\label{def:2:optimalSubOptimalArms}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
     % [Optimal and sub-optimal arms]
@@
     while arms satisfying $\mu_k < \mu^*$ are called \emph{sub-optimal}.
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{defn}
@@
 \begin{defn}[Regret]\label{def:2:regret}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
     For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ and it is defined as
@@
 \begin{lemma}[Regret decomposition]\label{lem:2:RegretDecomposition}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     The (mean) regret $R_T^{\cA}$ can be decomposed as a sum of the number of selections of sub-optimal arms $k$, weighted by their gaps:
@@
     \end{equation}
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
 \begin{theorem}\label{thm:2:firstLogTLowerBound}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
     No algorithm $\cA$ can achieve a (mean) regret $R_T^{\cA}(I) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
     where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
+\end{leftbar}  % XXX leftbar theorembar, comment if needed
 \end{theorem}
@@
 \begin{defn}\label{def:2:uniformlyEfficientAlgorithm}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
     An algorithm $\cA$ is \emph{uniformly efficient} if its (mean) regret satisfies
@@
     where this notation means that the ``constant'' $C_{I,\alpha}$ can depend on the problem instance $I$ and on $\alpha$, but \emph{not} on the time horizon $T$.
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{defn}
@@
 \begin{theorem}\label{thm:2:secondLogTLowerBound}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
     Let $\cA$ denote a uniformly efficient algorithm,
@@
     \end{equation}
+\end{leftbar}  % XXX leftbar theorembar, comment if needed
 \end{theorem}
@@
 \begin{theorem}\label{thm:2:forSecondLogTLowerBound}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
     Let $\cA$ denote a uniformly efficient algorithm,
@@
     \end{itemize}
+\end{leftbar}  % XXX leftbar theorembar, comment if needed
 \end{theorem}
@@
 \begin{theorem}\label{thm:2:worstCaseLowerBound}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
     Fix the number of arms $K$, and the algorithm $\cA$.
@@
 % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
-On a more experimental note, it was showed in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-player bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
+On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-player bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
 For more details, see \cite{modiDemo2016} for experiments on the multi-player case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
@@
 \begin{lemma}[Hoeffding's inequality]\label{lem:2:HoeffdingInequality}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     Let $Z_1,\dots,Z_n$ be $n$ \iid{} samples from a Bernoulli distribution of parameter $\mu$, of empirical mean $\widehat{Z_n} = \frac{1}{n} \sum\limits_{i=1}^n Z_i$, Hoeffding's inequality gives both
@@
     \end{align}
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
 \begin{theorem}[Regret bound for $\UCB_1$]\label{thm:2:UCBregretBound}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
     For simplicity, we restrict to $\alpha=2$.
@@
 \begin{lemma}[Pinsker's inequality]\label{lem:2:Pinsker}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) = p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.

### 2019-06-05 2-Chapters/3-Chapter/chapter3.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Starting to conclude Ch2

@@
 
-\paragraph{Simulation loop}
+\paragraph{Simulation loop.}
 %
@@
 
-\paragraph{Simulation loop for multi-player bandit}
+\paragraph{Simulation loop for multi-player bandit.}
 
@@
 gaussian_problem = MAB({'arm_type': Gaussian, 'params': means})
-gaussian_problem.plotHistogram()  # display the histogram showed below
+gaussian_problem.plotHistogram()  # display the histogram shown below
     \end{minted}
@@
 containing over 65 algorithms (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}).
-It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams showed on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
+It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams shown on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
 For instance, most existing algorithms are index-based, and new ones can be written easily by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
@@
 
-% \paragraph{List of algorithms}
+% \paragraph{List of algorithms.}
 
@@
 
-\paragraph{Two different problems}
+\paragraph{Two different problems.}
 
@@
 
-\paragraph{Other parameters of the experiments}
+\paragraph{Other parameters of the experiments.}
 
@@
 
-% \paragraph{Mean regret}
+% \paragraph{Mean regret.}
 
@@
 
-% \paragraph{Computation time}
+% \paragraph{Computation time.}
 
@@
 
-% \paragraph{Storage cost}
+% \paragraph{Storage cost.}
 
@@
 
-\paragraph{Interpretation}
+\paragraph{Interpretation.}
 
@@
 
-\paragraph{Bash command to run these experiments}
+\paragraph{Bash command to run these experiments.}
 

### 2019-06-05 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Proofreading of all Ch2 and Ch3

@@
 % \subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
-\paragraph{Notations.}\label{sub:25:aggregation}
+\textbf{Notations:}\label{sub:25:aggregation}
 
@@
 % ----------------------------------------------------------------------
-\paragraph{Explanations of \Aggr{} versus \ExpQ{} .}\label{sub:25:Exp4}
+\textbf{Comparison with \ExpQ{}:}\label{sub:25:Exp4}
 
@@
 
-\paragraph{How to choose the learning rates $(\eta_t)$?}
+\textbf{How to choose the learning rates $(\eta_t)$:}
 %
@@
 %
-This empirical estimation of the expected regret $R_T$ is plotted below, as a function of $T$, comparing some algorithms $\Alg_1,\dots,\Alg_N$ (for $N=6$), and their aggregation with \Aggr{} (displayed in orange bold),
+This empirical estimation of the expected regret $R_T$ is plotted below\footnote{~The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted.
+It corresponds to the second case in Theorem~\ref{thm:2:secondLogTLowerBound}.
+It is crucial to note that the lower-bound is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli}).},
+as a function of $T$, comparing some algorithms $\Alg_1,\dots,\Alg_N$ (for $N=6$), and their aggregation with \Aggr{} (displayed in orange bold),
 using the parameter-free learning rate sequence, $\eta_t = \log{N} / (t K)$.
-
-The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted.
-It corresponds to the second case in Theorem~\ref{thm:2:secondLogTLowerBound}.
-It is crucial to note that the lower-bound is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli}).
 
@@
 for both Bernoulli and non-Bernoulli problems (\eg, Gaussian, exponential and Poisson distributions).
-Further theoretical developments were left as future work.
+% Further theoretical developments were left as future work.
 

### 2019-06-05 2-Chapters/2-Chapter/chapter2.tex
Proofreading of all Ch2 and Ch3

@@
     The KL divergence between two distributions $d_1$ and $d_2$ is defined as
-    \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
+    \begin{align}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
         \KL\left(d_1, d_2\right) = \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}\cup\{\pm\infty\}.
-    \end{equation}
+    \end{align}
+\vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
 it is interesting to illustrate this definition and the notations with two important examples:
-%
-\begin{itemize}\label{par:2:notationsExponentialFamiliesBernoulliGaussian}
-    \item
+
+\label{par:2:notationsExponentialFamiliesBernoulliGaussian}
+% \begin{itemize}
+    % \item
+    $\bullet$
     \textbf{Bernoulli distributions} can be seen as an exponential family with $h(x) = 1$,
@@
     (with the limit behavior $\eta=+\infty$ if $\mu=0$).
-
+    %
     Moreover, the KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
@@
 
-    \item
+    $\bullet$
     \textbf{Gaussian distributions} of a fixed variance are a one-dimensional exponential family,
@@
     % with a fixed variance $\sigma^2$ form indeed a one-dimensional exponential family.
-
+    %
     Moreover, the KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
     $d_{\sigma^2}(x,y) = \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$.
-\end{itemize}
+% \end{itemize}
 
@@
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$ and if $\mu^* = \max_k \mu_k$, then the \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ and it is defined as
-    %
-    \begin{equation}
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$, with $\mu^* = \max_k \mu_k$, then its \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ defined as
+    %
+    \begin{align}\label{eq:2:regret}
         R_T^{\cA}
@@
         = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
-    \end{equation}
+    \end{align}
+    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
     The (mean) regret $R_T^{\cA}$ can be decomposed as a sum of the number of selections of sub-optimal arms $k$, weighted by their gaps:
-    \begin{equation}\label{eq:2:RegretDecomposition}
+    \begin{align}\label{eq:2:RegretDecomposition}
         R_T^{\cA} = \sum_{k=1}^K \Delta_k \; \E[ N_k(T) ]
         = \sum_{\substack{k=1,\dots,K\\\Delta_k > 0}} \Delta_k \; \E[ N_k(T) ].
-    \end{equation}
+    \end{align}
+    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    Let $\cA$ denote a uniformly efficient algorithm,
-    and $I$ an arbitrary problem instance.
-    For this instance,
-    there exists a constant $C_I$ depending only on the problem $I$,
+    If $\cA$ is uniformly efficient,
+    then for any arbitrary problem instance $I$,
+    there exists a constant $C_I$ depending only on $I$,
     and a time $T_0$ such that
@@
     \end{equation}
+    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    Fix the number of arms $K$, and the algorithm $\cA$.
-    Then for any horizon $T$, there exists a problem instance $I_T$ on which the algorithm suffers a regret $R_T^{\cA}(I_T)$ that verifies
-    \begin{equation}
-        R_T^{\cA}(I_T) \geq \Omega(\sqrt{K T}).
-    \end{equation}
+    Fix the number of arms $K$, and an algorithm $\cA$.
+    Then for any horizon $T$, there exists a problem instance $I_T$ on which the algorithm suffers at least a regret $R_T^{\cA}(I_T) \geq \Omega(\sqrt{K T})$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
     \end{align}
+    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 \begin{proof}\label{proof:2:UCBregretBound}
-    \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ? si c'est bien fait, ça peut aider la lecture, par exemple \textcolor{darkgreen}{$k$} et \textcolor{gold}{$k^*$}.}
-    %
     As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
@@
     \end{equation}
+
+    \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ? si c'est bien fait, ça peut aider la lecture, par exemple \textcolor{darkgreen}{$k$} et \textcolor{gold}{$k^*$}.}
 
@@
     Thus, we can continue from \eqref{eq:2:trickInUCBProof} and write
+    %
     \begin{align}\label{eq:2:decompositionInUCBProof}
-        N_k(T) & \leq n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n) \nonumber \\
+        N_k(T)
+        & \leq n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n) \nonumber \\
         & \leq n + \sum_{t=n}^T \indic(\hat{\mu}_{k^*,N_{k^*}(t)}(t) + \xi_{k^*,N_{k^*}(t)}(t) < \hat{\mu}_{k,N_k(t)}(t) + \xi_{k,N_k(t)}(t), N_k(t) \geq n)  \nonumber \\
-        & = n + \sum_{t=n}^T \indic\left(\min_{0 < n_{k^*} < t} \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \max_{n < n_k < t} \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right) \nonumber \\
-        \intertext{So two union bounds on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$ give}
+        & = n + \sum_{t=n}^T \indic\left(\min_{0 < n_{k^*} < t} \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \max_{n < n_k < t} \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right)
+    \end{align}
+    So two union bounds on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$ give
+    \begin{align}\label{eq:2:decompositionInUCBProof}
+        N_k(T)
         & \leq n + \sum_{t=n}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n}^t  \indic\left( \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right)
@@
         & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \nonumber \\
-        \intertext{We actually have that $\sum\limits_{n_{k^*}=1}^t \sum\limits_{n_{k}=n_T}^t  2 \frac{1}{t^4} \leq 2 \sum\limits_{n_{k^*}=1}^t \sum\limits_{n_{k}=1}^t \frac{1}{t^4} \leq 2 \sum\limits_{n_{k^*}=1}^t \frac{1}{t^3} \leq 2 \frac{1}{t^2}$}
+        \intertext{We actually have that $\sum\limits_{n_{k^*}=1}^t \sum\limits_{n_{k}=n_T}^t  2 \frac{1}{t^4} \leq 2 \sum\limits_{n_{k^*}=1}^t \frac{1}{t^3} \leq 2 \frac{1}{t^2}$, whose sum for $t=1$ to $\infty$ is $2\frac{\pi^2}{6}$.}
         & \leq \frac{8 \log(T)}{\Delta_k^2} + 1 + 2 \sum_{t=1}^T \frac{1}{t^2}
@@
 
-    More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs (\eg, for any $\alpha\geq\frac{1}{2}$).
+    More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs, in particular for any value of $\alpha\geq\frac{1}{2}$.
 \end{proof}
@@
 %
-Bayes-UCB is comparable to Thompson sampling in terms of memory complexity, but it can be slower in terms of time complexity as computing a quantile is generally more costly than just sampling from a distribution.
+Bayes-UCB is comparable to Thompson sampling in terms of memory complexity, but it can be slower, as computing a quantile is generally more costly than just sampling from a distribution.
 % a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
@@
 	% \centering
-    \includegraphics[width=1.20\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
+    \includegraphics[width=1.15\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
 	\caption{Average of rewards, as function of $t$, for $T=10000$ and $N=1000$.}
@@
 	% \centering
-	\includegraphics[width=1.20\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
+	\includegraphics[width=1.15\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
 	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
@@
 
-\paragraph{Summary.}
+\paragraph{Summary of this review of MAB algorithms.}
 %
@@
 
-\paragraph{Naive approaches.}
+\textbf{Naive approaches:}
 %
@@
 
-\paragraph{Illustrative example.}
+\textbf{Illustrative example:}
 %
@@
 
-\paragraph{The online approach.}
+\textbf{The online approach:}
 %
@@
     \end{equation}
+    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed

### 2019-06-05 2-Chapters/3-Chapter/chapter3.tex
Proofreading of all Ch2 and Ch3

@@
 \label{chapter:3}
+
+\paragraph{Abstract.}
+%
+SMPyBandits is a Python package I have developed since the beginning of my PhD.
+It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
+%  and SMPyBandits is written in Python.
+This library is by far the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
+%
+This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
+Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
+
 \minitoc
 
-\TODOL{Abstract ici et plus court, et petite introduction !}
+\newpage
+
+% \newpage
+% Write miniTOC just after the title
+\graphicspath{{2-Chapters/3-Chapter/Images/}}
+\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
+
+% ----------------------------------------------------------------------------
+\section{Introduction}
+\label{sec:3:Introduction}
+
 \TODOL{Analyser les expériences les plus récentes et décrire mieux en 3.2, et refaire toutes les courbes pour 3.2 et 3.3}
 
-\newpage
-\paragraph{Abstract.}
-%
+% This chapter is intended as a long version of the paper presenting SMPyBandits that I wrote in Summer 2018, see https://hal.inria.fr/hal-01840022
+
+% SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms and models of Multi-Armed Bandits.
+% It aims at being extensive, simple to use and maintain, with a clean and perfectly documented codebase.
+But most of all it allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments while starting them (see below for an example).
+
 SMPyBandits is a Python package I have developed since the beginning of my PhD.
@@
 More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
-It is hosted on GitHub, and uses the most recent technologies, by using two online services of continuous integration to run automated tests and build the documentation.
-
-This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
-Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
-% \TODOL{Do we really have time/space to talk about Markov models?}
-% Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.
-
-% \newpage
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/3-Chapter/Images/}}
-\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
-
-% This chapter is intended as a long version of the paper presenting SMPyBandits that I wrote in Summer 2018, see https://hal.inria.fr/hal-01840022
-
-% SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms and models of Multi-Armed Bandits.
-% It aims at being extensive, simple to use and maintain, with a clean and perfectly documented codebase.
-But most of all it allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments while starting them (see below for an example).
+It is hosted on GitHub, and uses the state-of-the-art development technologies, by using two online services of continuous integration to run automated tests and to build its online documentation.
 
@@
 \texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
-It uses the Sphinx software \cite{sphinx}, and the content is directly written in the Python files as docstring, to allow users to have access to the documentation from within their IDE or the Python console.
-The most interesting component of the library is the many algorithms being implemented: for each of them, the documentation gives at least a reference to a research paper (\eg, \cite{Kaufmann12BUCB} for \texttt{BayesUCB}), as well as a bird-eye view of its behavior.
-For most algorithms, in particular for index policies, the internal variables of the implementation are carefully linked with the notations of each paper, and formulas explaining the way the algorithm selects its arm are usually given.
-Whenever it is appropriate, the documentation also includes warning or information about the empirical performance of costly algorithms.
-The documentation also contains extensive examples of any intermediate numerical functions, that are also used as tests (using \texttt{doctest.testmod} from the standard library).
+It uses the Sphinx software \cite{sphinx}, and the content is directly written in the Python files as docstrings, so users have access to the documentation from within their IDE or the Python console.
+The most interesting component of the library is the many MAB algorithms being implemented: for each of them, the documentation gives a reference to a research paper (\eg, \cite{Kaufmann12BUCB} for \texttt{BayesUCB}), as well as a bird-eye view of its behavior.
+For most algorithms, especially for index policies, the internal variables of the implementation are carefully linked to the notations of each paper, and formulas explaining the way the algorithm selects its arm are usually given.
+Whenever it is needed, we also included warnings or information about the empirical performance of the more costly algorithms.
+
+The documentation also contains extensive examples of all intermediate numerical functions, that are also used as tests (using \texttt{doctest.testmod} from the standard library).
 For example, Figures~\ref{fig:3:twoScreenshotsOfTheDocumentation} and \ref{fig:3:twoMoreScreenshotsOfTheDocumentation} below show four screenshots from the documentation.
@@
 
-\TODOL{Si j'en parle plus dans le chapitre 2, je peux aller très vite sur cette partie !}
-
 \begin{itemize}
     \item Algorithm 1 is
-    an $\varepsilon$-greedy algorithm, as described in \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
+    $\varepsilon$-greedy \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
     It has a very small cost in terms of time and memory, but it usually achieves linear regret.
@@
     \item Algorithm 2 is
-    an Explore-then-Commit algorithm from \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
+    Explore-then-Commit \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
     Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but only obtain large (linear) regret.
@@
     % of more than 10 digits
-    \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}
+    % \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}
 
@@
     (\ie, not using recursive functions, by using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
-    By using such binary tournament, the extension $K>2$ arms is costly in terms of both time and memory, as illustrated below, and suffers from an exponential blow-up when $K$ increases.
+    Despite this technic, the extension to $K>2$ arms is still costly in terms of both time and memory, as illustrated below, and blows-up exponentially  when $K$ increases.
 \end{itemize}
@@
 
-\newpage  % WARNING
+% \newpage  % WARNING
 % ----------------------------------------------------------------------------

### 2019-06-05 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Done for a large proofreading tonight

@@
 \textbf{Notations:}\label{sub:25:aggregation}
-
+%
 We assume to have $N \geq 2$ MAB algorithms, $\Alg_1, \dots, \Alg_N$,
@@
 
+\TODOL{Utiliser des rewards et pas des loss... Aucun autre moment dans la thèse ne parle de loss, c'est un peu louche d'en parler juste là, ça fait vraiment copié/collé de l'article...}
+
 At every time step, after having observed a loss $\ell_{A(t+1)}(t+1)$ for its chosen action $A(t+1)$,
@@
 The loss estimate is unbiased in the following sense.
-If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\widetilde{\ell_{a,t+1}} = \ell_{A_a(t+1)}(t+1)$.
+If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\widetilde{\ell}_{a,t+1} = \ell_{A_a(t+1)}(t+1)$.
 This quantity can only be observed for those algorithms for which $A_{a}^{t+1}=A(t+1)$.
@@
 \begin{equation}
-	\widehat{\ell_{a,t+1}} = \frac{\ell_{A_a(t+1)}(t+1)}{p_{A_a(t+1)}^{t+1}} \mathbbm{1}(A_a(t+1) = A(t+1))
+	\widehat{\ell}_{a,t+1} = \frac{\ell_{A_a(t+1)}(t+1)}{p_{A_a(t+1)}^{t+1}} \mathbbm{1}(A_a(t+1) = A(t+1))
 \end{equation}
-satisfies $\mathbb{E}[\widehat{\ell_{a,t+1}} | \mathcal{H}_t] = \widetilde{\ell_{a,t+1}}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$.
-Observe that $\widetilde{\ell_{a,t+1}}=\ell^{t+1}$ for all algorithms $a$ such that $A_a(t+1)=A(t+1)$, and $\widetilde{\ell_{a,t+1}}=0$ otherwise.
+satisfies $\mathbb{E}[\widehat{\ell}_{a,t+1} | \mathcal{H}_t] = \widetilde{\ell}_{a,t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$.
+Observe that $\widetilde{\ell}_{a,t+1}=\ell^{t+1}$ for all algorithms $a$ such that $A_a(t+1)=A(t+1)$, and $\widetilde{\ell}_{a,t+1}=0$ otherwise.
 
@@
 \textbf{Comparison with \ExpQ{}:}\label{sub:25:Exp4}
-
+%
 % - explain, but don't give pseudo-code ?

### 2019-06-05 2-Chapters/3-Chapter/chapter3.tex
Done for a large proofreading tonight

@@
 
-\paragraph{Abstract.}
+\textbf{Abstract.}
 %
@@
 %  and SMPyBandits is written in Python.
-This library is by far the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
+It is the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
 %
 This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
-Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
+Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, and a study to compare time and memory costs of the main algorithms.
 

### 2019-06-05 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Fixed typos in some references
Done for a large proofreading tonight

@@
 \subsection{The first heuristics: \UCB{} unaware of retransmissions}
-\label{sub:43:heuristics}
+\label{sub:43:UCBnaive}
 % ----------------------------------------------------------------------
@@
 This first heuristic we propose is unaware of retransmission: the same channel is used for retransmissions.
-The \UCB{} algorithm is implemented independently by each device, and we present it in Algorithm~\ref{algo:43:UCB}.
+The \UCB{} algorithm is implemented independently by each device, we denote it ``first-stage'' \UCB, and we present it in Algorithm~\ref{algo:43:UCB}.
+It is the same algorithm as the Algorithm~\ref{algo:2:indexPolicy} in Section~\ref{sub:2:IndexPolicies} (using the same \UCB{} indexes as we defined them in equation \eqref{eq:UCB_index}),
+but we write it again to make clear the difference with first transmission and retransmission of messages.
+%
 Note that a device using this first approach is only able to select a channel for the first transmission (using \UCB), and it uses the same channel for all the corresponding retransmissions of a packet (if retransmissions happen).
@@
 
-\paragraph{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}.}
+\textbf{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}:}
 %

### 2019-06-05 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Done for a large proofreading tonight

@@
 For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
+
+\paragraph{Publications.}
+%
+This chapter is mainly based on our article \cite{Besson2018ALT}.
 
@@
   \end{align*}
+
   Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
+
   \begin{align}
@@
   and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
+
   \begin{align}
@@
 
-\paragraph{Logarithmic Regret for \MCTopM}
+\paragraph{Logarithmic Regret for \MCTopM.}
 %
@@
 
-% \paragraph{Number of switches}\label{app:5:NumberSwitches}
+% \paragraph{Number of switches.}\label{app:5:NumberSwitches}
 % %
@@
 
-\paragraph{Strong uniform efficiency}\label{app:5:JustifyingDefinition5}
+\paragraph{Strong uniform efficiency.}\label{app:5:JustifyingDefinition5}
 %
@@
 Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
-
-We report in Table~\ref{table:5:comparisonUCB_klUCB} below the results in terms of mean regret, plus or minus one standard deviation.
-The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
-Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
-Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved (the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is $\cO(1)$ if there is no arms in \Mworst{} !), and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous (state-of-the-art) approaches using ranks.
-%
-Additional experiments and illustrations are given below.
-We illustrate below the (empirical mean) regret $R_t$ as a function of time and not only results in terms of empirical mean regret $R_T$ at the end of the bandit game.
 
@@
 
+We report in Table~\ref{table:5:comparisonUCB_klUCB} above the results in terms of mean regret, plus or minus one standard deviation.
+The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
+Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
+Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved (the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is $\cO(1)$ if there is no arms in \Mworst{} !), and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous (state-of-the-art) approaches using ranks.
+%
+Additional experiments and illustrations are given below.
+We illustrate below the (empirical mean) regret $R_t$ as a function of time and not only results in terms of empirical mean regret $R_T$ at the end of the bandit game.
+
 The purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,

### 2019-06-05 2-Chapters/5-Chapter/chapter5.tex
Starting to conclude Ch2
Done for a large proofreading tonight

@@
 \minitoc
-
-\vfill{}
-
-\paragraph{Publications.}
-
-This chapter is mainly based on our article \cite{Besson2018ALT}.
 
@@
 In the paper \cite{SawantKumar2018}, the authors study for the first time distributed algorithms that are robust against malicious behavior, also called \emph{jamming attack}.
-They consider both the cases of jammers launching coordinated and uncoordinated attacks, and consider a set of $J$ jammers.
+They consider both the cases of \emph{jammers} launching coordinated and uncoordinated attacks, and consider a set of $J$ jammers.
 In the coordinated attack, the jammers select non-overlapping channels to attack in each time slot and can significantly increase the number of collisions for SUs.

### 2019-06-05 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for proofreading tonight

@@
 If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem introduced above would reduce to a parametric sequential test of
-\[\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0)),\]
+$\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$,
 against the alternative
-\[\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1)).\]
-
+$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
+%
 The Generalized Likelihood Ratio statistic for this test is defined by
@@
 %
-By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
-%
-\begin{equation}\label{eq:6:BernoulliDivergence}
-    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
-\end{equation}
+By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$),
+$\kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$
 %
@@
 %
-where for $k \leq k'$, $\widehat{\mu}_{k:k'} := \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
+where for $k \leq k'$, $\widehat{\mu}_{k:k'} := \frac{1}{k - k' + 1} \sum\limits_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
 
@@
     We can again work with log-likelihoods, and so we have
+
     \begin{align*}
@@
 
-Let us go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}.
+Let us go back to the Bernoulli case with divergence $\kl(x,y)$.
 We first observe that
@@
 
+\TODOL{Fix spacing issue!}
+
 \paragraph{Upper bound on term $(A)$.}
@@
     \centering
-    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
     \caption{Locations of the detected change-points for four algorithms on Problem 1.}
@@
     \centering
-    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
     \caption{Locations of the detected change-points for four algorithms on Problem 2.}
@@
     \centering
-    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
+    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
     \caption{\textbf{Problem $3$}: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
@@
     \centering
-    \includegraphics[width=0.80\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
+    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
     \caption{\textbf{Problem $4$}: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
@@
     \centering
-    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
+    \includegraphics[width=0.95\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
     \caption{\textbf{Pb $5$}: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
@@
     \centering
-    \includegraphics[width=1.20\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
+    \includegraphics[width=1.15\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
     \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 1.}

### 2019-06-06 2-Chapters/Conclusion/conclusion.tex
Typos
Changed \paragraph{XXX} to \paragraph{XXX.}
Done for proofreading tonight
Added leftbar like in my Master thesis
Added https://arxiv.org/abs/1811.11043

@@
 \label{chapter:conclusion}
+
+% \newpage
+\paragraph{Abstract.}
+%
+This last chapter concludes the thesis, first by summarizing the previous chapters, and by giving a general conclusion of the thesis.
+%
+We highlight what we consider to be the most important directions of future works, as well as some promising directions that were shortly studied during my PhD but not discussed in the manuscript.
+
+\vspace*{10pt}
+
 \minitoc
@@
 
-% \newpage
-\paragraph{Abstract}
-
-This last chapter concludes the thesis, first by summarizing the previous chapters, and by giving a general conclusion of the thesis.
-%
-We highlight what we consider to be the most important directions of future works, as well as some promising directions that were shortly studied during my PhD but not discussed in the manuscript.
+\newpage
 
@@
 % % ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter \ref{chapter:1}}}
+% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:1}}.}
 
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter \ref{chapter:2}}}
+\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:2}}.}
 
@@
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
-Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
 We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter \ref{chapter:3}}}
+\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:3}}.}
 
@@
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
-One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
-More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.
-All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
+One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, I we tried to follow actively the state-of-the-art research, since December $2016$ to June $2019$.
+More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary case.
+The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modifications.
 %
@@
 %
-Our library is used to run numerical simulations on Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
+Our library is used to run numerical simulations on Chapters~\ref{chapter:2}~ \ref{chapter:5} and~\ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
 
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter \ref{chapter:4}}}
+\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:4}}.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter \ref{chapter:5}}}
+\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:5}}.}
 
-The second chapter of Part~\ref{part:MABIOT} of this manuscript focussed on Multi-Player Multi-Arm Bandits models.
+The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Player Multi-Arm Bandits models.
 %
@@
 with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
-For the two easiest models --with sensing--, our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret.
+For the two easiest models --with sensing--, our theoretical contribution improves the state-of-the-art upper bounds on the regret.
 In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
@@
 We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proved to be very active in the last two years.
-For some models, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
+For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
 For example, it assumes to know the number of players $M$ before-hand, but we illustrated that previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter \ref{chapter:6}}}
+\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:6}}.}
 
@@
 % ----------------------------------------------------------------------------
-% \paragraph{Perspectives about \textbf{Chapter \ref{chapter:1}}}
+% \paragraph{Perspectives about \textbf{Chapter~\ref{chapter:1}}.}
 
 % ----------------------------------------------------------------------------
-% \paragraph{Perspectives about }}
+\paragraph{Perspectives about \textbf{Part~\ref{part:Introduction}}.}
+
+% ----------------------------------------------------------------------------
+% \paragraph{Perspectives about }.}
 
@@
 % ----------------------------------------------------------------------------
-% \paragraph{Perspectives about \textbf{Chapter \ref{chapter:3}}}
+% \paragraph{Perspectives about \textbf{Chapter~\ref{chapter:3}}.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Perspectives about \textbf{Chapter \ref{chapter:4}}}
+\paragraph{Perspectives about \textbf{Chapter~\ref{chapter:4}}.}
 
@@
 
-% \paragraph{Possible extensions of the first model}
+% \paragraph{Possible extensions of the first model.}
 
@@
 
-% \paragraph{Extensions of the demonstration}
+% \paragraph{Extensions of the demonstration.}
 
@@
 
-% \paragraph{Possible extensions for the second model}
+% \paragraph{Possible extensions for the second model.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Perspectives about \textbf{Chapter \ref{chapter:5}}}
+\paragraph{Perspectives about \textbf{Chapter~\ref{chapter:5}}.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Perspectives about \textbf{Chapter \ref{chapter:6}}}
+\paragraph{Perspectives about \textbf{Chapter~\ref{chapter:6}}.}
 
@@
 
-We would also like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or the rotted bandit model.
+We would also like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or the rotting bandit model \cite{Seznec2018}.
 
-Another interesting direction of future work is to study non-stationary distributed multi-player bandits, to unify the models from Chapter~\ref{chapter:5} and \ref{chapter:6}.
+Another interesting direction of future work is to study non-stationary distributed multi-player bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.
 A natural extension of the non-stationary model is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
@@
 %
-A promising direction is to directly try to join our contributions from Chapter~\ref{chapter:5} and \ref{chapter:6}, and propose an efficient algorithm using three parts:
+A promising direction is to directly try to join our contributions from Chapter~\ref{chapter:5} and~\ref{chapter:6}, and propose an efficient algorithm using three parts:
 \klUCB{} indexes for arm selection,
@@
 % ----------------------------------------------------------------------------
-\paragraph{Other directions of future works}
+\paragraph{Other directions of future works.}
 

### 2019-06-06 4-Listofpublications/Listofpublications.tex
Added https://arxiv.org/abs/1811.11043

@@
 
-\section*{International Conferences}
-\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
+\section*{International Workshops}
+\renewcommand{\labelenumi}{[IW-\arabic{enumi}]}
 \begin{etaremune}
@@
   \href{https://sites.google.com/view/ISIoT2019/}{ISIoT Workshop} at \href{https://dcoss.org/}{DCOSS}, Santorini, Greece, May 2019,
-  \href{https://hal.inria.fr/hal-02144465}{HAL-02144465}.
+  \href{https://hal.inria.fr/hal-02144465}{HAL-02144465},
+  \href{https://arxiv.org/abs/1906.00614}{arxiv:1906.00614}.
   \cite{MoyBesson2019}
@@
   \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
-  Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02049824}{HAL-02049824}.
+  Marrakech, Maroc, April 2019,
+  \href{https://hal.inria.fr/hal-02049824}{HAL-02049824},
+  \href{https://arxiv.org/abs/1902.10615}{arxiv:1902.10615}.
   \cite{Bonnefoi2019WCNC}
+\end{etaremune}
+
+\section*{Demonstrations at International Conferences}
+\renewcommand{\labelenumi}{[DEMO-\arabic{enumi}]}
+\begin{etaremune}
 \item
@@
   \href{http://wcnc2019.ieee-wcnc.org/}{IEEE WCNC},
-  Marrakech, Maroc, April 2019, \href{https://hal.inria.fr/hal-02006825}{HAL-02006825}.
+  Marrakech, Maroc, April 2019,
+  \href{https://hal.inria.fr/hal-02006825}{HAL-02006825},
+  \href{https://arxiv.org/abs/1902.01734}{arxiv:1902.01734}.
   \cite{Besson2019WCNC}
@@
   (\emph{Algorithmic Learning Theory}), Lanzarote, Spain, April 2018,
-  \href{https://hal.inria.fr/hal-01629733}{HAL-01629733}.
+  \href{https://hal.inria.fr/hal-01629733}{HAL-01629733},
+  \href{https://arxiv.org/abs/1711.02317}{arxiv:1711.02317}.
   \cite{Besson2018ALT}
@@
   Lisbonne, Portugal, September 2017,
-  \href{https://hal.inria.fr/hal-01575419}{HAL-01575419}, \textbf{Best Paper Award}.
+  \href{https://hal.inria.fr/hal-01575419}{HAL-01575419},
+  \href{https://arxiv.org/abs/1807.00491}{arxiv:1807.00491},
+  \textbf{Best Paper Award}.
   \cite{Bonnefoi17}
@@
 
-\section*{National Conferences}
+\section*{National Conferences (\emph{in French})}
 \renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
@@
   Kaufmann}, February 2019,
-  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471}.
+  \href{https://hal.inria.fr/hal-02006471}{HAL-02006471},
+  \href{https://arxiv.org/abs/1902.01575}{arxiv:1902.01575}.
   \cite{Besson2019GLRT}
@@
   Kaufmann}, September 2018,
-  \href{https://hal.inria.fr/hal-01736357}{HAL-01736357}.
+  \href{https://hal.inria.fr/hal-01736357}{HAL-01736357},
+  \href{https://arxiv.org/abs/1803.06971}{arxiv:1803.06971}.
   \cite{Besson2018DoublingTricks}

### 2019-06-06 2-Chapters/1-Chapter/chapter1.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Fixed the Nomenclature

@@
     % }
-    \caption{Organization of the thesis: a reading map.
+    \caption[Organization of the thesis: a reading map]{Organization of the thesis: a reading map.
 Any path containing Chapter~\ref{chapter:1}, Chapter~\ref{chapter:2}, at least one of the three Chapters~\ref{chapter:4}, \ref{chapter:5} and \ref{chapter:6}, and the Conclusion is a self contained way to read this thesis.}
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,
+    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,\\
     \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
@@
     by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,
+    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,\\
     \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
@@
     by \textbf{Lilian Besson}, active development since October $2016$,
-    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
+    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.\\
     About $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    September $2018$,\\
+    September $2018$,
     \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.

### 2019-06-06 5-Nomenclature/Nomenclature.tex
Working to merge Christophe's feedback of Chapter5
Fixed the Nomenclature

@@
 \nomenclature[A]{RandTopM}{\textit{Random Hoping on Top-M} (algorithm)}
-\nomenclature[A]{GLR}{\textit{Generalized Likelihood Ratio}}
+\nomenclature[A]{GLR, GLRT}{\textit{Generalized Likelihood Ratio}, \textit{Generalized Likelihood Ratio Test}}
 \nomenclature[A]{MAB}{\textit{Multi-Armed Bandit}}
@@
 \nomenclature[A]{RL}{\textit{Reinforcement Learning}}
-\nomenclature[A]{GLRT}{\textit{Generalized Likelihood Ratio Test}}
 
@@
 \nomenclature[G]{$\nu_k$, $\nu_k(t)$, $\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
-\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$ in Chapter~\ref{chapter:5}, arm $i$ in the $k$-th stationary interval in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$ (in Chapter~\ref{chapter:6}), arm $k$ for user $j$ (in Chapter~\ref{chapter:5}), arm $i$ in the $k$-th stationary interval (in Chapter~\ref{chapter:6})}
 \nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
@@
 In Chapter~\ref{chapter:4} in Algorithm~\ref{algo:43:UCBwithDelay} it denotes a delay.
-In Chapter~\ref{chapter:6}, it can denote different gaps}
+In Chapter~\ref{chapter:6}, it can denote different gaps ($\Delta^{\text{opt}}$ and $\Delta^{\text{change}}$)}
 \nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
@@
 $\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
-\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Usually denotes the location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\tau$}{Usually denotes the length of a sliding-window, \eg, in SW-\UCB{} in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\tau$}{Length of a sliding-window, \eg, in SW-\UCB{} in Chapter~\ref{chapter:6}}
 \nomenclature[G]{$\gamma$}{Usually denotes a discount factor, \eg, in D-\UCB{} or D-TS in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, $\lambda_k = \mu_k + \varepsilon$ in Chapter~\ref{chapter:5} or a drift correction parameter in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, the parameter for the $\varepsilon$-greedy algorithm, or a drift correction parameter for the \CUSUM{} algorithm in Chapter~\ref{chapter:6}}
 \nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\beta$}{Usually denotes a threshold of GLR tests in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\beta(n,\delta)$}{Usually denotes a threshold for the statistical (GLR) tests in Chapter~\ref{chapter:6}}
 
-\nomenclature[G]{$\alpha_0$, $\delta_0$}{Usually denotes a constant scaling of a parameter of an algorithm, for instance $\alpha = \alpha_0 \alpha_T$ is used in Chapter~\ref{chapter:6}, where $\alpha_T$ is the formula suggested by the theoretical analysis and $\alpha_0 > 0$, \eg, $\alpha_0 = 0.01$}
+\nomenclature[G]{$\alpha_0$, $\delta_0$}{Usually denotes a constant scaling of a parameter of an algorithm, for instance $\varepsilon_t = \varepsilon_0 / t$ is used in Chapter~\ref{chapter:2}, or $\alpha = \alpha_0 \alpha_T$ is used in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\pi$}{It usually denotes the number $\pi\simeq 3.14\ldots$, but it also denotes a probability distribution in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} or a permutation in Section~\ref{par:5:twoIdeasOrthogonalization}}
+% \nomenclature[G]{$\rho$, $\rho^k$}{A bandit strategy, denoting the mathematical object which maps past observations to future decisions (mainly in Chapter~\ref{chapter:5})}
 
@@
 % matches Roman Symbols        > L
-\nomenclature[L]{$t$, $s$, $n$, $r$}{Time step, $t\in\{1,\dots,T\}$. Chapter~\ref{chapter:6} also uses $s$, $n$ and $r$, \eg, in the $\sup$ of the stopping times or some technical lemmas}
+\nomenclature[L]{$t$, $s$, $n$, $r$}{Time step, $t\in[T]$. Chapter~\ref{chapter:6} also uses $s$, $n$ and $r$, \eg, in the $\sup$ of the stopping times or some technical lemmas}
 \nomenclature[L]{$T$}{Time horizon, the duration of the bandit game (always $T\geq1$)}
 \nomenclature[L]{$T_0$, $T_1$}{Fixed durations of some algorithms based on different phases, \eg, the Musical Chair algorithm from \cite{Rosenski16}}
-\nomenclature[L]{$\cA$, $\pi$}{A algorithm, a policy, designing the algorithm (\eg, \UCB) rather than the strategy.
-$\pi$ can also denote a permutation in Section~\ref{par:5:twoIdeasOrthogonalization}}
-\nomenclature[L]{$\rho$, $\rho^k$}{A bandit strategy, denoting the mathematical object which maps past observations to future decisions (mainly in Chapter~\ref{chapter:5})}
-\nomenclature[L]{$A(t)$, $A^j(t)$}{Decision of algorithm $\cA$ at time $t$, $A(t)\in\{1,\dots,K\}$. Decision for user $j$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$\cA$}{An algorithm, also referred to as a policy or a strategy.
+$\cA_1,\dots,\cA_N$ denote the $N$ aggregated algorithms in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} and $\cA_1,\dots,\cA_M$ denotes the algorithms of the $M$ players in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$A(t)$, $A^j(t)$}{Decision of algorithm $\cA$ at time $t\in[T]$, $A(t)\in[K]$ (from algorithm $\cA$), decision for user $j\in[M]$ in Chapter~\ref{chapter:5} (from algorithm $\cA^j$)}
 \nomenclature[L]{$Y_{k,t}$}{Random sample from the arm $k$ at time $t$}
@@
 \nomenclature[L]{$K$}{Number of arms for multi-armed bandit games}
-\nomenclature[L]{$k$}{Usually denotes the $k$-th arm, $k\in\{1,\dots,K\}$, mainly used in subscripts}
+% \nomenclature[L]{$[T}$, $[K]$, etc}{The set of integers from $1$ to $T$, $[T]=\{1,\dots,T\} = \{ n \in\N: 1 \leq n \leq T\}$}
+\nomenclature[L]{$k$}{Usually denotes the $k$-th arm, $k\in[K]$, mainly used in subscripts}
 \nomenclature[L]{$M$}{In Chapter~\ref{chapter:4}, denote the maximum number of retransmission of a packet in the ALOHA protocol, and in Chapter~\ref{chapter:5}, denote the number of player for multi-players bandit games}
 \nomenclature[L]{$m$}{In Chapter~\ref{chapter:4}, denote the maximum length of the back-off interval after a collision}
-\nomenclature[L]{$i$, $j$}{Usually denotes the $i$-th or $j$-th player, $i,j\in\{1,\dots,M\}$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$i$, $j$}{Usually denotes the $i$-th or $j$-th player, $i,j\in[M]$ in Chapter~\ref{chapter:5}}
 \nomenclature[L]{$N$}{Usually denotes the number of independent repetitions of the same numerical experiments (\eg, $N=1000$).
@@
 
-\nomenclature[L]{$C_k(t)$, $C^j(t)$}{In Chapters~\ref{chapter:4} and \ref{chapter:5}, collision indicator on arm $k$ or for user $j$ at time $j$}
+\nomenclature[L]{$C_k(t)$, $C^j(t)$}{In Chapters~\ref{chapter:4} and \ref{chapter:5}, collision indicator, on arm $k\in[K]$ or for user $j\in[M]$, at time $j$}
 
-\nomenclature[L]{$C_{\bm{\mu}}$, $D_{\bm{\mu}}$, $G_{M,\bm{\mu}}$}{In Chapter~\ref{chapter:5}, constants depending on the problem parameters only (or also on $M$)}
+\nomenclature[L]{$C_{\bm{\mu}}$, $D_{\bm{\mu}}$, $G_{M,\bm{\mu}}$}{In Chapter~\ref{chapter:5}, constants depending on the problem parameters only and on $M$}
 
@@
 \nomenclature[O]{$\cF$, $\cF_t$}{Filtration in a probabilistic model, after $t-1$ prior observations}
-\nomenclature[O]{$[K]$, $[N]$ etc}{For an integer $N\in\N$, $N\geq0$, $[N]$ denotes the set $\{1,\dots,N\}$. If the order is important, it is ordered from $1$ to $N$.}
+\nomenclature[O]{$[K]$, $T$, $[N]$ etc}{For an integer $N\in\N$, $N\geq0$, $[N]$ denotes the set $\{1,\dots,N\} = \{ n \in\N : 1 \leq n \leq N \}$.
+If the order is important, it is ordered from $1$ to $N$.}
 % \nomenclature[O]{$I_1,\dots,I_5$}{Used in Chapter~\ref{chapter:5} to denote various events}
@@
 
-\nomenclature[O]{$\lfloor \bullet \rfloor$, $\lceil \bullet \rceil$}{Floor and ceiling funtions, $\lfloor x \rfloor = \sup\{n\in\Z, x \leq n\}$ and $\lceil x \rceil = \sup\{n\in\Z, x < n\}$}
+\nomenclature[O]{$\lfloor \bullet \rfloor$, $\lceil \bullet \rceil$}{Floor $\lfloor x \rfloor = \sup\{n\in\Z, n \leq x\}$ and ceil $\lceil x \rceil = \inf\{n\in\Z, x < n\}$ functions}
 
@@
 % matches Subscripts           > I
-\nomenclature[I]{$x_k$}{Usually denotes a variable depending on an arm, for $k\in\{1,\dots,K\}$}
+\nomenclature[I]{$x_k$}{Usually denotes a variable depending on an arm, for $k\in[K]$}
+\nomenclature[I]{$Y_{k,t}$}{Usually denotes a variable depending on an arm, for $k\in[K]$, and on time, for $t\in[T]$}
 
@@
 % matches Superscripts         > E
-\nomenclature[E]{$y^j$}{Usually denotes a variable depending on a player, for $j\in\{1,\dots,M\}$}
-\nomenclature[E]{$y^{(k)}$, $y^{(\ell)}$}{Usually the superscript index $(k)$ (resp.
-$(\ell)$) denotes a variable in the $k$-th stationary interval, under the \emph{global} (resp.
-\emph{local}) changes point-of-view, in Chapter~\ref{chapter:6}}
+\nomenclature[E]{$y^j$}{Usually denotes a variable depending on a player in Chapter~\ref{chapter:5} (for $j\in[M]$), or to distinguish between different independent runs in numerical simulations in Chapter~\ref{chapter:2} (for $j\in[N]$)}
+\nomenclature[E]{$y^{(k)}$, $y^{(\ell)}$}{Usually the superscript index $(k)$ or $(\ell)$ denotes a variable in the $k$-th stationary interval, under the  point-of-views of \emph{global} or \emph{local} changes, in Chapter~\ref{chapter:6}}

### 2019-06-06 2-Chapters/1-Chapter/chapter1.tex
Added leftbar to all defn,cor,thm etc

@@
 are \href{https://github.com/Naereen/phd-thesis/}{publicly published},
-online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}},
-under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}}.
+under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}},
+online at \\
+\href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}},
 

### 2019-06-06 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Added leftbar to all defn,cor,thm etc

--- 2019-06-05 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
@@
 				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
-				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A_a(t+1) = j\}), \;\forall 1\leq j \leq K$\;
+				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(\{A_a(t+1) = j\}), \;\forall 1\leq j \leq K$\;
 				% \tcp*{Sum trusts}
@@
 				%
-				Give \emph{original} reward $(A(t+1), 1 - \ell_{A(t+1)}(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
-				Compute an \emph{unbiased} estimate of the loss of the trusted algorithms,$$ \ell^{t+1} = \ell_{A(t+1)}(t+1) / p^{t+1}_{A(t+1)}\;. $$
+				Give \emph{original} reward $(A(t+1), r(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
+				Compute an \emph{unbiased} estimate of the reward of the trusted algorithms,$$ \widehat{r}(t+1) = r(t+1) / p^{t+1}_{A(t+1)}\;. $$
 				% \tcp{Use bandit feedback to update $\pi^{t}$}
-				% \tcp{Finally, use the instantaneous bandit feedback (loss $\ell_{A(t+1),t}$) to update $\pi^{t}$}
+				% \tcp{Finally, use the instantaneous bandit feedback (reward $\ell_{A(t+1),t}$) to update $\pi^{t}$}
 				% \For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
@@
 					\eIf{$\Alg_a$ was trusted, \ie, $A_a(t+1) = A(t+1)$}{
-						$ \pi^{t+1}_{a} = \exp(\eta_t \ell^{t+1}) \times \pi^{t}_{a} $
+						$ \pi^{t+1}_{a} = \exp(\eta_t r(t+1) \times \pi^{t}_{a} $
 						\tcp*{Trusted more}
 						}{
-						% $ \pi^{t+1}_{a} = \exp(- \eta_t \ell^{t+1}) \times \pi^{t}_{a} $
+						% $ \pi^{t+1}_{a} = \exp(- \eta_t r(t+1) \times \pi^{t}_{a} $
 						$ \pi^{t+1}_{a} = \pi^{t}_{a} $
@@
 
-At every time step, after having observed a loss $\ell_{A(t+1)}(t+1)$ for its chosen action $A(t+1)$,
+\TODOL{Etre sur des notations entre $t$ et $t+1$ pour faire comme le reste du chapitre et de la thèse !}
+
+At every time step, after having observed a reward $r(t+1)=Y_{A(t+1),t+1}$ for its chosen action $A(t+1)$,
 the algorithm updates the trust probabilities from $\pi^t$ to $\pi^{t+1}$ by
-a multiplicative exponential factor (using the learning rate and the \emph{unbiased} loss).
+a multiplicative exponential factor (using the learning rate and the \emph{unbiased} reward).
 %
@@
 
-The loss estimate is unbiased in the following sense.
-If one had access to the rewards $r_k(t+1)$ (or the losses $\ell_k(t+1)$) for all arms $k$, the loss incurred by algorithm $a$ at time $t+1$ would be $\widetilde{\ell_{a,t+1}} = \ell_{A_a(t+1)}(t+1)$.
-This quantity can only be observed for those algorithms for which $A_{a}^{t+1}=A(t+1)$.
-However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $\widetilde{\ell_{a,t}}$.
-More precisely, if we define the estimate by
+The reward estimate is unbiased in the following sense.
+If one had access to the samples $Y_{k,t+1}$ for all arms $k$, the reward incurred by algorithm $a$ at time $t+1$ would be $r^a(t+1) = Y_{A_a(t+1),t+1}$.
+But we are not in the full information setting (see \cite{CesaLugosi06}), but in the bandit setting, so only one of the samples $Y_{k,t+1}$ can be observed, $r(t)=Y_{A(t+1),t+1}$.
+This quantity can only be observed for those algorithms $a$ for which $A_{a}^{t+1}=A(t+1)$.
+However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $r^a(t+1)$. More precisely, if we define the estimate by
 \begin{equation}
-	\widehat{\ell_{a,t+1}} = \frac{\ell_{A_a(t+1)}(t+1)}{p_{A_a(t+1)}^{t+1}} \mathbbm{1}(A_a(t+1) = A(t+1))
+	\widehat{r^a}(t+1) = \frac{Y_{A_a(t+1),t+1}}{p_{A_a(t+1)}^{t+1}} \indic(A_a(t+1) = A(t+1))
 \end{equation}
-satisfies $\mathbb{E}[\widehat{\ell_{a,t+1}} | \mathcal{H}_t] = \widetilde{\ell_{a,t+1}}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\mathcal{H}_{t}$.
-Observe that $\widetilde{\ell_{a,t+1}}=\ell^{t+1}$ for all algorithms $a$ such that $A_a(t+1)=A(t+1)$, and $\widetilde{\ell_{a,t+1}}=0$ otherwise.
+satisfies $\E_{\cO_t}[\widehat{r^a}(t+1) | \cO_t] = Y_{A_a(t+1),t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\cO_t$.
+%
+Observe that $\widehat{r^a}(t+1) = 0$ for all algorithms $a$ such that $A_a(t+1) \neq A(t+1)$.
 
@@
 An important feature of \Aggr{} is the feedback provided to each underlying bandit algorithm, upon the observation of arm $A(t+1)$.
-Rather than updating only the trusted algorithms (that is the algorithms which would have drawn arm $A(t+1)$) with the observed reward $r_{A(t+1)}(t+1)=1 - \ell_{A(t+1)}(t+1)$,  we found that updating each algorithm with the (original) loss observed for arm $A(t+1)$ improves the performance drastically.
+Rather than updating only the trusted algorithms (that is the algorithms which would have drawn arm $A(t+1)$) with the observed reward $r(t+1)$, we found that updating each algorithm with the (original) reward observed for arm $A(t+1)$ improves the performance drastically.
 %
@@
 \textbf{Comparison with \ExpQ{}:}\label{sub:25:Exp4}
-
+%
 % - explain, but don't give pseudo-code ?
-The \ExpQ{} algorithm (see, e.g. \cite[Section 4.2]{Bubeck12})
-is similar to \Aggr{}, presented in Algorithm \ref{algo:25:Aggr},
+The \ExpQ{} algorithm dates from \cite{Auer02NonStochastic}, and for instance it is also well explained in Section~4.2 of \cite{Bubeck12}.
+It is similar to \Aggr{}, presented in Algorithm~\ref{algo:25:Aggr},
 but differs in the two following points.
 %
-First, $a\sim\pi^t$ is sampled first and the arm chosen by $\Alg_a$ is trusted, whereas \Aggr{}
+The first difference is that $a\sim\pi^t$ is sampled first and the arm chosen by $\Alg_a$ is trusted, whereas \Aggr{}
 needs to listen to the $N$ decisions to perform the updates on $\pi^{t+1}$.
@@
 %
-Second, after having computed the loss estimate $\ell$, \ExpQ{} updates the estimated cumulative loss for each algorithm,
-$\widetilde{L}_a(t) = \sum_{s=1}^{t} \ell_{A_a^s}(s) \times \mathbbm{1}(A^{s}_{a} = A^{s}_{\mathrm{aggr}})$.
+The second difference is that after having computed the reward estimate $\widehat{r^a}(t+1)$, \ExpQ{} updates the estimated cumulative reward for each algorithm,
+$\widetilde{R}_a(t) = \sum_{s=1}^{t} \widehat{r^a}(t+1) \times \indic(A^{s}_{a} = A^{s}_{\mathrm{aggr}})$.
 %
@@
 of updating $\pi^{t}$ multiplicatively as we do for our proposal, \ExpQ{} recomputes it, proportionally to
-$\exp(- \eta_t \widetilde{L}_a(t))$.
+$\exp(\eta_t \widetilde{R}_a(t))$.
+%
+We also note that \ExpQ{} uses \emph{losses} instead of rewards, with $\ell(t) = 1 - r(t)$, but this is only a change of vocabulary.
 
@@
 
-Figures~\ref{fig:25:EasyBernoulli} and \ref{fig:25:HarderMixed} are in semilog-$y$ scale, this helps to see that the best algorithms can be an \emph{order of magnitude} more efficient than the worst, and the \Aggr{} performs similarly to the best ones, when the other aggregation algorithms are usually amongst the worst.
+Figures~\ref{fig:25:EasyBernoulli} and~\ref{fig:25:HarderMixed} are in semilog-$y$ scale, this helps to see that the best algorithms can be an \emph{order of magnitude} more efficient than the worst, and the \Aggr{} performs similarly to the best ones, when the other aggregation algorithms are usually amongst the worst.
 Figure~\ref{fig:25:HarderMixed_semilogx} is in semilog-$x$ scale to show that the regret of efficient algorithms are indeed logarithmic.
@@
 
-For Bernoulli problems (Figures~\ref{fig:25:EasyBernoulli} and \ref{fig:25:HardBernoulli}), UCB with $\alpha=1/2$, Thompson sampling, BayesUCB and \klUCB{}$^+$ (with the binary $\kl$ function) all perform similarly, and \Aggr{} is found to be as efficient as all of them.
+For Bernoulli problems (Figures~\ref{fig:25:EasyBernoulli} and~\ref{fig:25:HardBernoulli}), UCB with $\alpha=1/2$, Thompson sampling, BayesUCB and \klUCB{}$^+$ (with the binary $\kl$ function) all perform similarly, and \Aggr{} is found to be as efficient as all of them.
 For Gaussian and exponential arms, rewards are truncated into $[0,1]$, and the variance of Gaussian distributions is fixed to $\sigma^2 = 0.05$ for all arms, and can be known to the algorithms (the \kl{} function is adapted to this one-dimensional exponential family).
@@
 
-For each of the 4 problems considered, the \Aggr{} algorithm with default option (broadcast loss to all players) is the best of all the aggregation algorithms,
+For each of the 4 problems considered, the \Aggr{} algorithm with default option (\ie, broadcast rewards to all the aggregated algorithms) is the best of all the aggregation algorithms,
 and its regret is very close to the best of the aggregated algorithms.
-Especially in difficult problems with mixed or unknown distributions, \Aggr{} showed to be more efficient that \ExpQ{} and orders of magnitude better than the other reference aggregation algorithms \LearnExp{} and \CORRAL{} (see Figures~\ref{fig:25:HarderMixed} and \ref{fig:25:HarderMixed_semilogx}).
+Especially in difficult problems with mixed or unknown distributions, \Aggr{} showed to be more efficient that \ExpQ{} and orders of magnitude better than the other reference aggregation algorithms \LearnExp{} and \CORRAL{} (see Figures~\ref{fig:25:HarderMixed} and~\ref{fig:25:HarderMixed_semilogx}).
 

### 2019-06-06 2-Chapters/2-Chapter/chapter2.tex
Done for a large proofreading tonight
Added leftbar to all defn,cor,thm etc

@@
 % Since then, many efficient algorithms were proposed, XXX
-%
+
 % In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
@@
 
-\paragraph{About notations.}
+\textbf{About notations.}
 %
@@
 
-\paragraph{Main references for the curious reader.}
+\textbf{Main references for the curious reader.}
 %
@@
 \textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
-When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r_{A(t)}(t)\in\R$, and so on for a finite number of steps $t$, from $t=1$ until $t=T$ for an \emph{horizon} $T$.
-%
-A commonly studied goal for the player is to maximize its sum of received rewards, $\sum_{t=1}^T r_{A(t)}(t)$
+When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) = Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$, from $t=1$ until $t=T$ for an \emph{horizon} $T$.
+%
+A commonly studied goal for the player is to maximize its sum of received rewards, $\sum_{t=1}^T r(t)$
 (or its \emph{mean} in stochastic and stationary models).
@@
             \draw [->] (player) to[bend left=90] node[pos=0.5] {Chooses a \textbf{discrete} action $k=A(t) \in\{1,\dots,K\}$} (environment);
-            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Observes a \textbf{real-valued} reward $r_k(t) \in \R$} (player);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Observes a \textbf{real-valued} reward $r(t)=Y_{k,t} \in \R$} (player);
             %
@@
 
-This quantity $\sum_{t=1}^T r_{A(t)}(t)$ can indeed be random, and it usually depends on two aspects.
-First, the rewards may depend on the randomness of the environment, \ie, the unknown and unpredictable values $r_k(t)$.
-Then this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the \emph{past} observations, \ie, $A(1)$, $r_{A(1)}(1)$, $\dots$, $A(t-1)$, $r_{A(t-1)}(t-1)$, and possibly on an external source of randomness.
-%
-Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, uniformly distributed on $[0,1]$, and that has to be independent from the rewards $r_k(t)$.
+This quantity $\sum_{t=1}^T r(t)$ can indeed be random, and it usually depends on two aspects.
+First, the rewards may depend on the randomness of the environment, \ie, the unknown and unpredictable values $r(t)=Y_{k,t}$.
+Then this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the \emph{past} observations, \ie, $A(1)$, $Y_{A(1),1}$, $\dots$, $A(t-1)$, $Y_{A(t-1),t-1}$, and possibly on an external source of randomness.
+%
+Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, uniformly distributed on $[0,1]$, and that has to be independent from the samples $Y_{k,t}$.
 External randomness is used in most MAB algorithms, \eg, to break ties uniformly in an index policy.
@@
 
-In all this thesis, we focus on the model with finite arms, and binary or real-valued stochastic rewards, that is $A(t)\in\{1,\dots,K\}$ for a fixed and known number of arms $K\in\N,K\geq2$, and $r_{A(t)}(t)\in\R$.
+In all this thesis, we focus on the model with finite arms, and binary or real-valued stochastic rewards, that is $A(t)\in\{1,\dots,K\}$ for a fixed and known number of arms $K\in\N,K\geq2$, and $Y_{A(t),t}\in\R$.
 %
@@
 We also restrict to \emph{stochastic} rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in\{1,\dots,\}=[K]$.
-Rewards are stationary, meaning that $(r_k(t))_{t\in\N^*}$ is independent and identically distributed (\iid), $r_k(t) \sim \nu_k$ for any $t\geq1$.
-
-The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T Y_t] = \sum_{t=1}^T \E[r_{A(t)}(t)]$.
-The expectancy is taken on the rewards $(r_k(t))_{k,t}$, and the extern random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
+Rewards are stationary, meaning that $(Y_{k,t})_{t\in\N^*}$ is independent and identically distributed (\iid), $Y_{k,t} \sim \nu_k$ for any $t\geq1$.
+
+The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T r(t)] = \sum_{t=1}^T \E[Y_{A(t),t}]$.
+The expectancy is taken on the rewards $(Y_{k,t})_{k,t}$, and the extern random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
 %
 Other commonly studied objectives include best-arm identification (BAI) \cite{audibert2010best}.
+
+% \TODOL{Je dois unifier les notations, et utiliser $Y_{k,t}$ pour les flux qui sont iid et $r(t)$ pour les récompenses effectivement observées (comme dans les chapitres 5 et 6)}
 
@@
 The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
-The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r_{A(t)}(t)$.
-Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r_{A(t)}(t)\in\{0,1\}$, that is sampled \iid{} from $\nu_k$.
-%
-The UI of the demo also shows the current value of $t$ (``total plays'') and $\sum_{s=1}^t r_{A(s)}(s)$ the ``total reward''.
-For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
+The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r(t)$.
+Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r(t)\in\{0,1\}$, that is sampled \iid{} from $\nu_k$.
+%
+The UI of the demo also shows the current value of $t$ (``total plays'') and $\sum_{s=1}^t r(s)$ the ``total reward''.
+For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) = \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
 %
@@
 More formally, an algorithm can be defined as a sequence of \emph{measurable} functions $(\cA_t)_{t\geq1}$,
-where $\cA_t$ maps the past observations $O_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$
-to an arm $\cA_t(O_t) = A(t) \in[K]$,
-if we denote $Y_s = r_{A(s)}(s)$ the $s$-th reward.
+where $\cA_t$ maps the past observations $O_t = (U_0, Y_{A(1),1}, U_1, \dots, Y_{A(t-1),t-1}, U_{t-1})$
+to an arm $\cA_t(O_t) = A(t) \in[K]$
+(we remind that we denote $r(s) = Y_{A(s),s}$ the $s$-th reward).
 The initial information is reduced to $O_1 = (U_0)$, and the first decision is $A(1) = \cA_1(O_1)$.
@@
 % https://smpybandits.github.io/docs/Arms.html
-\textbf{We also focus only on real-valued rewards}, meaning that $r_k(t)\in\R$ for all arm $k$ and time $t$.
+\textbf{We also focus only on real-valued rewards}, meaning that $Y_{k,t}\in\R$ for all arm $k$ and time $t$.
 
@@
 
-\begin{defn}[Kullback-Leibler divergence]\label{def:2:KLDivergence}
+\begin{definition}[Kullback-Leibler divergence]\label{def:2:KLDivergence}
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
@@
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
-\end{defn}
+\end{definition}
 
@@
 In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-In the model with sensing, the rewards $r_k(t)$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
-A reward of $r_k(t) = 1$ indicates that no Primary User was sensed, while a reward of $r_k(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
+In the model with sensing, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
+A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
 %
@@
 Let us first introduce some notations.
-From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that generates \iid{} rewards $r_k(t)\sim \nu_k$, for any time $t$.
+From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that generates \iid{} samples $Y_{k,t} \sim \nu_k$, for any time $t$.
+% with $r(t)=Y_{A(t),t}$).
 We focus on distributions fully characterized by their means, that can be Bernoulli or any one-dimensional exponential family.
-We denote $\mu_k = \E[r_k(t)] \in \R$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$}).
+We denote $\mu_k = \E[Y_{k,t}] \in \R$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$}).
 %
@@
 
-\begin{defn}\label{def:2:optimalSubOptimalArms}
+\begin{definition}\label{def:2:optimalSubOptimalArms}
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
@@
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
-\end{defn}
-
-If the goal of the player is to maximize $\E\left[\sum_{t=1}^T r_{A(t)}(t)\right]$,
+\end{definition}
+
+If the goal of the player is to maximize $\E\left[\sum_{t=1}^T r(t)\right]$,
 the optimal strategy for this bandit problem is to always pull one of the optimal arms (it can be not unique), but it is unrealistic as the player does not know the true means, nor the optimal arms (as long as they are not all optimal, \ie, in non trivial problems).
@@
 let us introduce the (mean) regret $R_T^{\cA}$ of an algorithm $\cA$ as
-$R_T^{\cA} = \E\left[ \sum_{t=1}^T (r_{k^*}(t) - r_{A(t)}(t)) \right]$.
+$R_T^{\cA} = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]$.
 %
@@
     R_T^{\cA}
-    & = \E\left[ \sum_{t=1}^T (r_{k^*}(t) - r_{A(t)}(t)) \right]
-    = \E\left[ \sum_{t=1}^T \underbrace{r_{k^*}(t)}_{\text{\iid{} from $\nu_{k^*}$}} \right] - \E\left[ \sum_{t=1}^T r_{A(t)}(t) \right] \\
-    & = T \times \E\left[ r_{k^*}(t=1) \right] - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right]
-    = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
+    & = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
+    = \E\left[ \sum_{t=1}^T \underbrace{Y_{k^*,t}}_{\text{\iid{} from $\nu_{k^*}$}} \right] - \E\left[ \sum_{t=1}^T r(t) \right] \\
+    & = T \times \E\left[ Y_{k^*,1} \right] - \sum_{t=1}^T \E\left[ r(t) \right]
+    = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right].
 \end{align*}
@@
 
-\begin{defn}[Regret]\label{def:2:regret}
+\begin{definition}[Regret]\label{def:2:regret}
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
@@
         R_T^{\cA}
-        = \E\left[ \sum_{t=1}^T (r_{k^*}(t) - r_{A(t)}(t)) \right]
+        = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
         % = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right]
-        = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right].
+        = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right].
     \end{align}
@@
 %
-and that the rewards $r_k(t)$ are all \iid{} of mean $\mu_k$.
+and that the samples $Y_{k,t}$ are all \iid{} of mean $\mu_k$.
 The \emph{gap} between any arm $k\in[K]$ and an optimal arm is defined as $\Delta_k = \mu^* - \mu_k$
@@
     We use the chain rule of expectation and a conditioning on $O_t$,
-    because the expectation is taken on the randomness of the (\iid) rewards $(r_k(t))_t$ and on the decisions of the player $(A(t))_t$, which are measurable wrt to the past observations $O_t = (U_0, Y_1, U_1, \dots, Y_{t-1}, U_{t-1})$.
+    because the expectation is taken on the randomness of the (\iid) samples $(Y_{k,t})_t$ and on the decisions of the player $(A(t))_t$, which are measurable wrt to the past observations $O_t = (U_0, Y_{A(1),1}, U_1, \dots, Y_{A(t-1),t-1}, U_{t-1})$.
     Thus we can rewrite the expected cumulated reward as follows:
@@
     \begin{align*}
-        \E \left[ \sum_{t=1}^T r_{A(t)}(t) \right]
-        &= \E \left[ \sum_{t=1}^T \sum_{k=1}^K r_k(t) \mathbbm{1}(A(t) = k) \right]
-        = \sum_{k=1}^K \sum_{t=1}^T \E \left[ r_k(t) \mathbbm{1}(A(t) = k) \right] \\
-        &= \sum_{k=1}^K \sum_{t=1}^T \E \Bigl[ \E \left[ r_k(t) | O_t \right] \mathbbm{1}(A(t) = k) \Bigr]
-        = \sum_{k=1}^K \sum_{t=1}^T \E \Bigl[ \E \left[ r_k(1) \right] \mathbbm{1}(A(t) = k) \Bigr] \\
-        &= \sum_{k=1}^K \underbrace{\E \left[ r_k(1) \right]}_{\mu_k} \underbrace{\sum_{t=1}^T \E \left[ \mathbbm{1}(A(t) = k) \right]}_{= N_k(T)}
+        \E \left[ \sum_{t=1}^T r(t) \right]
+        &= \E \left[ \sum_{t=1}^T \sum_{k=1}^K Y_{k,t} \mathbbm{1}(A(t) = k) \right]
+        = \sum_{k=1}^K \sum_{t=1}^T \E \left[ Y_{k,t} \mathbbm{1}(A(t) = k) \right] \\
+        &= \sum_{k=1}^K \sum_{t=1}^T \E \Bigl[ \E \left[ Y_{k,t} | O_t \right] \mathbbm{1}(A(t) = k) \Bigr]
+        = \sum_{k=1}^K \sum_{t=1}^T \E \Bigl[ \E \left[ Y_{k,1} \right] \mathbbm{1}(A(t) = k) \Bigr] \\
+        &= \sum_{k=1}^K \underbrace{\E \left[ Y_{k,1} \right]}_{\mu_k} \underbrace{\sum_{t=1}^T \E \left[ \mathbbm{1}(A(t) = k) \right]}_{= N_k(T)}
         % = \sum_{k=1}^K \mu_k \E \left[ \sum_{t=1}^T \mathbbm{1}(A(t) = k) \right]\\
@@
     %
-    Thus $R_T^{\cA} = T \mu^* - \sum_{t=1}^T \E\left[ r_{A(t)}(t) \right] = T \mu^* - \sum_{k=1}^K \mu_k \E[ N_k(T) ] = \sum_{k=1}^K (\mu^* - \mu_k) \E[ N_k(T) ]$,
+    Thus $R_T^{\cA} = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right] = T \mu^* - \sum_{k=1}^K \mu_k \E[ N_k(T) ] = \sum_{k=1}^K (\mu^* - \mu_k) \E[ N_k(T) ]$,
     as $\sum_{k=1}^K  \E[ N_k(T) ] = T$.
@@
 
-\begin{defn}\label{def:2:uniformlyEfficientAlgorithm}
+\begin{definition}\label{def:2:uniformlyEfficientAlgorithm}
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
@@
 They are detailed in Algorithm~\ref{algo:2:naiveStrategies} below.
-We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) = \sum_{s=1}^t r_k(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
+We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) = \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
 %
@@
                 \tcp*[f]{Exploit}\\
-                Observe a reward $r_{A(t)}(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
+                Observe a reward $r(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
             }
@@
             }
-            Observe a reward $r_{A(t)}(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
+            Observe a reward $r(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
 		}
@@
             }
-            Observe a reward $r_{A(t)}(t)$ and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$\;
+            Observe a reward $r(t)$ and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$\;
 		}
@@
     $\hat{\mu}_{a,n}(t)$ and $\xi_{a,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
-    \ie, $\hat{\mu}_{a,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r_k(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
+    \ie, $\hat{\mu}_{a,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
     %
@@
     So two union bounds on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$ give
-    \begin{align}\label{eq:2:decompositionInUCBProof}
+    \begin{align}\label{eq:2:decompositionInUCBProof2}
         N_k(T)
@@
     %
-    \begin{align}\label{eq:2:decompositionInUCBProof2}
+    \begin{align}\label{eq:2:decompositionInUCBProof3}
         \E[ N_k(T) ]
@@
             Play arm $A(t)$ uniformly among the arms of maximal random sample: $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k(t))$\;
-            Observe a reward $r_{A(t)}(t)$ and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$\;
+            Observe a reward $r(t)$ and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$\;
 		}
@@
 $\Pr(A(t)=k) = (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
-The weights are initially uniform, \ie, $w_k(0)=\frac{1}{K}$, and then at each time after observing a reward $r_k(t)$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
+The weights are initially uniform, \ie, $w_k(0)=\frac{1}{K}$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
 The weights are then renormalized, the have a sum $\sum_k w_k(t+1) = 1$.
-This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r_k(t)$, that is, $\widehat{r_k}(t) = \Pr(A(t)=k)$.
+This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = \Pr(A(t)=k)$.
 %
@@
         \begin{tabular}{c|cc|cc}
-        \textbf{Algorithms} & $\sum\limits_{t=1}^{T=100} r_{A(t)}(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r_{A(t)}(t)$ & $R_{T=10000}^{\cA}$ \\
+        \textbf{Algorithms} & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
             \hline
@@
 It should not be surprising that the black line of the lower-bound is \emph{above} the values of the regret as the lower-bound is only asymptotic.}
-of the average rewards (\ie, $\frac{1}{t} \sum\limits_{s=1}^t r_{A(s)}(s)$), which should converge to $\mu^*$ when $t\to\infty$ for efficient algorithms.
+of the average rewards (\ie, $\frac{1}{t} \sum\limits_{s=1}^t r(s)$), which should converge to $\mu^*$ when $t\to\infty$ for efficient algorithms.
 %
@@
 
-We prove it below using simple real analysis arguments\footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).},
+We prove it below using simple real analysis arguments,
+% \footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).},
 %
@@
     Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) = p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
-    Then for any $p,q\in[0,1]$, we have
-    \begin{equation}\label{eq:2:Pinsker2}
-        \kl(p,q) \geq 2 (p-q)^2.
-    \end{equation}
-    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+
+    Then for any $p,q\in[0,1]$, $\kl$ verifies $\kl(p,q) \geq 2 (p-q)^2$.
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
     Let us compute its derivative, $g'(x) = (\frac{\partial \; \kl}{\partial 2})(p, p + x) - 4x$.
-    Simple calculations yield
+    Simple calculus yield
     \[ \left(\frac{\partial \; \kl}{\partial 2}\right) (p, p + x) = - \frac{p}{p+x} + \frac{1-p}{1-(p+x)} = -x \frac{(2(p + x) - 1)^2}{(p + x)(p + x - 1)}. \]

### 2019-06-06 2-Chapters/3-Chapter/chapter3.tex
Added leftbar to all defn,cor,thm etc

@@
 a stochastic MAB problem is defined by $K>1$ distributions $\nu_k$ (also called arms),
-used to sample \iid{} rewards $r_k(t) \sim \nu_k, \forall t$.
-An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
-Her goal is to maximize $\sum_{t=1}^T r_{A(t)}(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.
+used to generate the \iid{} samples $Y_{k,t} \sim \nu_k, \forall t$.
+An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r(t) = Y_{A(t),t}$ without knowing the other (hidden) rewards.
+Her goal is to maximize $\sum_{t=1}^T r(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.
 
@@
 Note that the second step is repeated a large number of times (\eg, $N=100$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
-If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms.
+If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(Y_{k,t})_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms.
 
@@
 The
-    sample (random) feedback $Y_k(t)$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^{(j)}(t)$ from the random feedback and the collision model.
+    sample (random) feedback $Y_{k,t}$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^{(j)}(t)$ from the random feedback and the collision model.
 Finally feeds the observation $(A^{(j)}(t), Y_{A^{(j)}(t)}(t), r^{(j)}(t))$ to player $j$, for all the $M$ players,

### 2019-06-06 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Added leftbar to all defn,cor,thm etc

@@
 %
-For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
+For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r(t)=Y_{k,t} \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
 The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
-$r_{1 \dots T} := \sum_{t = 1}^T r_{A(t)}(t),$
+$\sum_{t = 1}^T r(t),$
 which is a usual objective in MAB problems.
@@
 Contrary to many previous work done in the CR field \cite{Jouini10,Jouini12} (\eg, Opportunistic Spectrum Access),
-the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
+the reward $r(t)=Y_{k,t}$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
 In our model, rewards rather come from receiving or not an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
@@
 \begin{proposition}\label{prop:41:Lagrangian}
+\begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
     The \emph{Lagrange multipliers} method \cite{BoydVanderberghe04} can be used to solve the constraint real-valued maximization problem introduced in equation \eqref{eq:41:prob}.
@@
     \end{equation}
+\end{leftbar}  % XXX leftbar propositionbar, comment if needed
 \end{proposition}

### 2019-06-06 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Added leftbar to all defn,cor,thm etc

@@
 The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
-$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r_k(\tau) \mathbbm{1}(A(\tau) = k)$,
-where $r_{k}(t)$ is the reward obtained after transmission in channel $k$ at time $t$.
+$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r(\tau) \mathbbm{1}(A(\tau) = k)$,
+where $r(t)=Y_{k,t}$ is the reward obtained after transmission in channel $k$ at time $t$.
 The upper confidence bound in each channel $k$ is defined as
@@
 				Transmit in channel $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
-				Reward $r_{A(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
+				Reward $r(t) = 1$, if \emph{Ack} is received, else $0$\;
 			}
@@
 Note that, we use the subscript $({}^r)$ to denote the variables
-$\widehat{\mu^r_k}(t)$, $B^r_k(t)$ and $U^r_k(t)$,
+$\widehat{\mu^r}(t)$, $B^r_k(t)$ and $U^r_k(t)$,
 related to the \UCB{} algorithm employed for the retransmission.
@@
 				Transmit in channel $C^r(t) = \arg\max_{1\leq k \leq K} U^r_k(t)$\;
-				Reward $r^r_{C^r(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
+				Reward $r^r(t) = 1$, if \emph{Ack} is received, else $0$\;
 			}

### 2019-06-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Added leftbar to all defn,cor,thm etc

@@
 \begin{definition}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
   A \emph{collision} occurs at time $t$ if at least two players choose the same arm.
@@
   that respectively indicate that a collision occurs at time $t$ for player $j$ and that a collision occurs at time $t$ on arm $k$.
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{definition}
@@
 
-A multi-player MAB strategy is formally defined as a tuple $\rho = (\rho^1,\dots,\rho^M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
-First, each player $j$ should adopt a \emph{sequential} strategy $\rho^j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
+A multi-player MAB strategy is formally defined as a tuple $\cA = (\cA_1,\dots,\cA_M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
+First, each player $j$ should adopt a \emph{sequential} strategy $\cA_j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
 Previous observations for player $j$ at time $t$ always include the previously chosen arms $A^j(s)$ and received rewards $r^j(s)$ for $s<t$, but may also include the \emph{sensing information} $Y_{A^j(t),t}$ or the \emph{collision information} $C^j(t)$.
@@
 \begin{definition}\label{def:5:MbestMworst}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
   We denote by $\mu_1^*$ the best mean, $\mu_2^*$ the second best etc, and
@@
   Note that they are both uniquely defined if $\mu_M^* > \mu_{M+1}^*$.
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{definition}
@@
 Following a natural approach in the bandit literature, we evaluate the performance of a multi-player strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
-The regret of the strategy $\rho$ at horizon $T$ is the difference between the cumulated reward of an oracle strategy, assigning in this case the $M$ players to \Mbest,
-and the cumulated reward of strategy $\rho$:
+The regret of the strategy $\cA$ at horizon $T$ is the difference between the cumulated reward of an oracle strategy, assigning in this case the $M$ players to \Mbest,
+and the cumulated reward of strategy $\cA$:
 
 \begin{definition}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
   The excepted centralized multi-player regret is defined by
   \begin{equation}\label{eq:5:regret}
-    R_T(\boldsymbol{\mu}, M, \rho) := \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
-  \end{equation}
+    R_T(\boldsymbol{\mu}, M, \cA) := \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
+  \end{equation}
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{definition}
@@
 \begin{definition}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
   % [$T_k(T)$ and $\cC_k(T)$]
@@
   \end{equation}
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{definition}
@@
 \begin{lemma}\label{lem:5:DecompositionRegret}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
   For any bandit instance $\boldsymbol{\mu}\in\cP_M$ such that $\mu_M^* > \mu_{M+1}^*$, it holds that
@@
     \begin{align}\label{eq:5:termeundeuxtermetrois}
-      R_T(\boldsymbol{\mu}, M, \rho) &=
+      R_T(\boldsymbol{\mu}, M, \cA) &=
       \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[T_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}} \\
@@
   % \end{small}
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
 \begin{lemma}\label{lem:5:1stLowerBound}
-    For any strategy $\rho$ and $\boldsymbol{\mu}\in\cP_M$, it holds that
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
+    For any strategy $\cA$ and $\boldsymbol{\mu}\in\cP_M$, it holds that
     \begin{align*}
-        R_T(\boldsymbol{\mu}, M, \rho)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)].
+        R_T(\boldsymbol{\mu}, M, \cA)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)].
     \end{align*}
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
   \end{equation*}
-  By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \rho)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
+  By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \cA)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
   \begin{align*}
@@
   And now both terms are non-negative, as $T_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
-  which proves that $R_T(\boldsymbol{\mu}, M, \rho) \geq \ref{eq:5:term1}$, as wanted.
+  which proves that $R_T(\boldsymbol{\mu}, M, \cA) \geq \ref{eq:5:term1}$, as wanted.
 \end{proof}
@@
 \begin{definition}\label{def:5:DecentralizedUniformEfficiency}
-  A strategy $\rho$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$ and for all $\alpha \in (0,1)$,
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
+  An algorithm $\cA$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$ and for all $\alpha \in (0,1)$,
   % \vspace*{-5pt}  % XXX
   \begin{align}
-    & R_T(\boldsymbol{\mu},M,\rho) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
+    & R_T(\boldsymbol{\mu},M,\cA) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
     \text{and} & \;
@@
   \end{align}
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{definition}
@@
 %
-From this assumption $\left(R_T(\boldsymbol{\mu},M,\rho)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
+From this assumption $\left(R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
 ${T}- \E_{\mu}[T_k(T)] {=} o(T^{\alpha})$, and so
@@
 The additional assumption in \eqref{eq:5:SUE} further implies some notion of \emph{fairness}, as it suggests that each of the $M$ players spends on average the same amount of time on each of the $M$ best arms.
-Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\rho^{\sigma} = (\rho^{\sigma(1)},\dots,\rho^{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$.
+Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\cA_{\sigma} = (\cA_{\sigma(1)},\dots,\cA_{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$.
 In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
@@
 
-% As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the ``collision information term'' is zero for any algorithm $\rho$.
-% However, we wanted to be still able to present the derivations of our lower-bound, mainly because its proof and its form give important insights on how to design an asymptotically optimal algorithm $\rho$.
+% As it was discovered in \cite{BoursierPerchet18} and further explained in Appendix~E of \cite{KaufmannAbbas19}, our lower-bound was not true for any algorithm, as we wrongly claimed in \cite{Besson2018ALT} that the ``collision information term'' is zero for any algorithm $\cA$.
+% However, we wanted to be still able to present the derivations of our lower-bound, mainly because its proof and its form give important insights on how to design an asymptotically optimal algorithm $\cA$.
 % % and because it still applies to algorithms of the \RhoRand{} family as well as our proposals \RandTopM{} and \MCTopM.
 
-% FIXME
 The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
-Fix a player $j\in[M]$, and we consider the observations $O_t$ gathered by this player after $t$ rounds of its algorithm $\rho^j$:
+Fix a player $j\in[M]$, and we consider the observations $O_t$ gathered by this player after $t$ rounds of its algorithm $\cA_j$:
 \[ O_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
@@
 Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
-We denote $\Pr_{\bm{\mu}}^{O_t}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\rho$.
+We denote $\Pr_{\bm{\mu}}^{O_t}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\cA$.
 Then we introduce the \emph{collision information term} as:
 \begin{equation}\label{eq:5:collisionInformationTerm}
-  \cI_{\bm{\mu},\bm{\lambda}}(\rho,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}}).
-\end{equation}
-
-For a \emph{decentralized strategy} $\rho$ that has access to the sensing information (\ie, ruling out model \modeltrois), and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T) = \smallO{\log(T)}$,
+  \cI_{\bm{\mu},\bm{\lambda}}(\cA,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}}).
+\end{equation}
+
+For a \emph{decentralized strategy} $\cA$ that has access to the sensing information (\ie, ruling out model \modeltrois), and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$,
 we now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections.
-The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\rho,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
+The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
 The theorem stated below is proven in the Appendix of \cite{Besson2018ALT},
@@
 \begin{theorem}\label{thm:5:BetterLowerBound}
-  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\rho$ and $\boldsymbol{\mu}\in\cP_M$.
-  Furthermore, if $\rho$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$, then
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
+  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\cA$ and $\boldsymbol{\mu}\in\cP_M$.
+  Furthermore, if $\cA$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$, then
   % \vspace*{-10pt}  % XXX
@@
   \begin{equation}\label{eq:5:ourLowerBound}
-    \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \rho)}{\log(T)}
+    \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
     \geq M \times \left( \sum_{k \in \Mworst} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_M^*)} \right) .
   \end{equation}
+\end{leftbar}  % XXX leftbar theorembar, comment if needed
 \end{theorem}
@@
 \begin{equation}\label{eq:5:Zhao10LowerBound}
-  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \rho)}{\log(T)}
+  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
   \geq \sum_{k \in \Mworst} \left( \sum_{j=1}^{M} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_{j}^*)} \right),
@@
 %
-It is worth mentioning that \cite{Zhao10} proved a lower bound under the more general assumption for $\rho$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
+It is worth mentioning that \cite{Zhao10} proved a lower bound under the more general assumption for $\cA$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
 Our result could be extended to this case but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.
@@
 \begin{lemma}\label{lem:5:1stUpperBound}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
   The term \ref{eq:5:term2} in Lemma~\ref{lem:5:DecompositionRegret} is upper bounded as
@@
   % \vspace*{-5pt}  % XXX
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
 \begin{lemma}\label{lem:5:SubOptimalSelections}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
   For any $\boldsymbol{\mu}\in\cP_M$,
@@
   \end{equation}
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
   \begin{lemma}\label{lem:5:Fact1KLUCB}
+  \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
       For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
@@
       \end{equation}
+  \end{leftbar}  % XXX leftbar lemmabar, comment if needed
   \end{lemma}
@@
   \begin{lemma}\label{lem:5:Fact2KLUCB}
+  \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
@@
     \end{small} % XXX
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
 \begin{lemma}\label{lem:5:elementaryLemma_RandTopM_MCTopM}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
   For any $\boldsymbol{\mu}\in\cP_M$,
@@
   \end{equation*}
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
 \begin{lemma}\label{lem:5:collisionsMCTopM}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
   For any $\boldsymbol{\mu}\in\cP_M$,
@@
 \begin{theorem}\label{thm:5:LogarithmicRegret_MCTopMklUCB}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
   If all $M$ players use
@@
   \begin{equation}\label{eq:5:LogarithmicRegret_MCTopMklUCB}
-    R_T(\boldsymbol{\mu}, M, \rho) \leq G_{M,\boldsymbol{\mu}} \log(T) + \smallO{\log T}.
+    R_T(\boldsymbol{\mu}, M, \cA) \leq G_{M,\boldsymbol{\mu}} \log(T) + \smallO{\log T}.
   \end{equation}
   Moreover, the dependency of the constant regarding the number of players is $G_{M,\boldsymbol{\mu}} = \bigO{M^3}$.
+\end{leftbar}  % XXX leftbar theorembar, comment if needed
 \end{theorem}

### 2019-06-06 2-Chapters/6-Chapter/chapter6.tex
Done for a large proofreading tonight
Added leftbar to all defn,cor,thm etc

@@
 \label{chapter:6}
-\minitoc
 
-\TODOL{Abstract ici et plus court, et petite introduction !}
-
-\newpage
 \paragraph{Abstract.}
@@
 to account for possible non-stationarity of the rewards.
-% that is, for problems that could be tackled with MAB algorithms.
-Instead of dealing with changes happening possibly at every time step, like in adversarial models, we rather consider a finite small number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
+% % that is, for problems that could be tackled with MAB algorithms.
+% Instead of dealing with changes happening possibly at every time step, like in adversarial models, we rather consider a finite small number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
+%
+We review existing works on piece-wise stationary MAB models.
+% for which the two main families of proposed solutions are either passively or actively adaptive.
+In practice, it was already observed that actively adaptive strategies usually greatly outperform the passively adaptive ones.
+% The more efficient strategies from the current state-of-the-art are based on a combination of an efficient algorithm for the stationary problem with an efficient online change detection algorithm, like CUSUM, PHT or the Generalized Likelihood Ratio Test (GLRT), and so we pursue on this direction.
+%
+We study the Generalized Likelihood Ratio Test (GLRT) for
+% Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any 
+bounded or sub-Bernoulli distributions.
+Our test enjoys finite-time guarantees
+% its \emph{false alarm probability}, meaning that in a stationary model, the test should not detect any change,
+% and its \emph{detection delay}, meaning that in a model with a change-point, the test should detect within a reasonable delay.
+%
+of its \emph{false alarm probability} and \emph{detection delay}, and can be combined with an efficient bandit policy (\klUCB), to propose an efficient algorithm for piece-wise stationary problems, \GLRklUCB.
+% Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
+% % \MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
+%
+% We give the first unified analysis of a piece-wise stationary bandit algorithm for the two variants, and our analysis
+We analyze its regret, and show that it achieves state-of-the-art non-asymptotic regret bounds.
+% and yield regret bounds with explicit constants.
+% Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$, when the algorithm runs with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
+%
+Finally, we give extensive numerical experiments, to compare our approach against other state-of-the-art solutions.
+% , on different piece-wise stationary bandit problems.
+% % , of increasing difficulty.
+% %
+% We also include additional details about our policy \GLRklUCB, including a sensitivity analysis of its two parameters, a numerical evaluation of its time and memory costs, as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
 
-We start by reviewing the previous work on piece-wise stationary MAB models, for which the two main families of proposed solutions are either passively or actively adaptive.
-In practice, it was already observed that actively adaptive strategies usually greatly outperform the passively adaptive ones.
-The more efficient strategies from the current state-of-the-art are based on a combination of an efficient algorithm for the stationary problem with an efficient online change detection algorithm, like CUSUM, PHT or the Generalized Likelihood Ratio Test (GLRT), and so we pursue on this direction.
-%
-Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any sub-Bernoulli distributions, which include bounded distributions.
-Our test, B-GLRT, enjoys finite-time guarantees in terms of two quantities:
-its \emph{false alarm probability}, meaning that in a stationary model, the test should not detect any change,
-and its \emph{detection delay}, meaning that in a model with a change-point, the test should detect within a reasonable delay.
+\minitoc
 
-Thanks to these two non-asymptotic properties, the B-GLRT test can be combined with an efficient bandit policy, the \klUCB{} index policy, to propose an efficient algorithm for piece-wise stationary bandit problems, \GLRklUCB.
-Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
-% \MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
-%
-We give the first unified analysis of a piece-wise stationary bandit algorithm for the two variants, and our analysis is completely non-asymptotic and yield regret bounds with explicit constants.
-Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$, when the algorithm runs with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
-
-Finally, we give extensive numerical experiments to compare our approach, in its two variants, against the current state-of-the-art policies, on different piece-wise stationary bandit problems, of increasing difficulty.
-%
-We also include additional details about our policy \GLRklUCB, including a sensitivity analysis of its two parameters, a numerical evaluation of its time and memory costs, as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
-
-% \vfill{}
-
-\paragraph{Publications.}
-%
-This chapter is based on our articles \cite{Besson2019GLRT,Besson2019Gretsi}.
+\newpage
 

### 2019-06-06 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Added leftbar to all defn,cor,thm etc

@@
 %
-For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
-Indeed, in cognitive radio new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate; whereas in online recommendation, the popularity of items is also subject to trends.
+For cognitive radio applications, as well as other applications such as recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
+Indeed, in cognitive radio, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate.
+And for instance in online recommendation, the popularity of items is also subject to trends.
 Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.
@@
 
-It could seem that the problem is considered as solved by this simple Exp3.S algorithm, but it was actively studied since the seminal paper \cite{Auer02NonStochastic}.
+It could seem that the problem is considered as solved by the Exp3.S algorithm, but it was actively studied since the seminal paper \cite{Auer02NonStochastic}.
 Two main reasons explain the dynamic research on this problem:
@@
 
-\paragraph{Outline.}
+\textbf{Outline.}
 %
@@
 
+% \vfill{}
+
+\textbf{Publications.}
+%
+This chapter is based on our articles \cite{Besson2019GLRT,Besson2019Gretsi}.
+
 % -----------------------------------------------------------------
@@
 
-A policy $\pi$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
-The performance of $\pi$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\pi$,
+A policy $\cA$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
+Like for stationary problems (see Definition~\ref{def:2:regret}), the performance of $\cA$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\cA$
+We use the same notation as the regret for stationary problem without ambiguity, as in this chapter we only consider the following definition.
 %
 \begin{equation}\label{def:6:nti}
-    R_T^{\pi} = \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
+    R_T^{\cA} = \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
 \end{equation}
@@
 If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint recursively by
-\begin{equation}
-    \tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}.
-\end{equation}
+$\tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}$.
 %
@@
 \begin{definition}[Sub-Bernoulli distributions]\label{def:6:subBernoulliDistributions}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
     We call a \emph{sub-Bernoulli distribution} any distribution $\nu$ that satisfies
     $\ln \bE_{X\sim\nu}\left[e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu=\bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) = \ln(1-\mu + \mu e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$, for any $\mu\in[0,1]$.
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{definition}
@@
 \begin{definition}\label{def:6:GLRDef}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
     The Bernoulli GLRT (B-GLRT) change point detector with threshold function $\beta(n,\delta)$ is the stopping time $\widehat{\tau}_{\delta}$ defined by
@@
     \end{equation}
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
 \end{definition}
@@
 \begin{lemma}\label{lem:6:FalseAlarm}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     Let be $\bP_{\mu_0}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t$.
@@
 \begin{lemma}\label{lem:6:ConcFirst}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     $(X_i)_{i \in \N^*}$ and $(Y_i)_{i \in \N^*}$ two independent \iid{} processes with respective means $\mu$ and $\mu'$ such that
@@
     where $\cT$ is the function defined in \eqref{def:6:function_T}.
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
@@
 \begin{lemma}\label{lem:6:Delay}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
@@
 \begin{proposition}\label{prop:6:EnoughSamples}
+\begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
     For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $i$ (\ie, for a $k\in\{1,\dots,\NCi\}$, one has $\tau_i^{(k)} = \tau_i(t) < s \leq t < \tau_i^{(k+1)}$) it holds that
@@
 
-\begin{assumption}[Enough samples between two global change-points]
-    \label{ass:6:LongPeriodsGlobal}
+\begin{assumption}[Enough samples between two global change-points]\label{ass:6:LongPeriodsGlobal}
+\begin{leftbar}[assumptionbar]  % XXX leftbar assumptionbar, comment if needed
     Define
@@
     $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
+\end{leftbar}  % XXX leftbar assumptionbar, comment if needed
 \end{assumption}
@@
 \begin{theorem}\label{thm:6:mainRegretBoundGlobal}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
     For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
@@
     \end{align}
+\end{leftbar}  % XXX leftbar theorembar, comment if needed
 \end{theorem}
@@
 \begin{corollary}\label{cor:6:Global}
+\begin{leftbar}[corollarybar]  % XXX leftbar corollarybar, comment if needed
     For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriodsGlobal},
@@
     \end{enumerate}
+\end{leftbar}  % XXX leftbar corollarybar, comment if needed
 \end{corollary}
@@
 
-\begin{assumption}[Enough samples between two local change-points]
-    \label{ass:6:LongPeriods}
+\begin{assumption}[Enough samples between two local change-points]\label{ass:6:LongPeriods}
+\begin{leftbar}[assumptionbar]  % XXX leftbar assumptionbar, comment if needed
     Define
@@
     Then \emph{we assume} that for all arm $i$ and all change-point of that arm $\ell \in \{1,\dots,\NCi\}$, $\tau_i^{(\ell)} - \tau_i^{(\ell-1)} \geq 2\max (d_i^{(\ell)},d_i^{(\ell-1)})$.
+\end{leftbar}  % XXX leftbar assumptionbar, comment if needed
 \end{assumption}
@@
 \begin{theorem}\label{thm:6:mainRegretBound}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
     For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on Local Restart satisfies
@@
 \begin{corollary}\label{cor:6:Local}
+\begin{leftbar}[corollarybar]  % XXX leftbar corollarybar, comment if needed
     For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriods},
@@
 \begin{lemma}\label{lem:6:GoodEventGlobal}
+\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
     With $\cE_T(\delta)$ defined as in \eqref{def:6:GoodEvenGlobal}, the ``bad event'' is unlikely:
     $\bP(\cE_T^c(\delta)) \leq \delta(K+1)\Upsilon_T$.
+\end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}

### 2019-06-06 2-Chapters/1-Chapter/chapter1.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment

--- 2019-06-06 2-Chapters/1-Chapter/chapter1.tex
@@
 are \href{https://github.com/Naereen/phd-thesis/}{publicly published},
+under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}},
 online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}},
-under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}}.
 

### 2019-06-06 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment

@@
 \end{equation}
-satisfies $\E_{\cO_t}[\widehat{r^a}(t+1) | \cO_t] = Y_{A_a(t+1),t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $\cO_t$.
+satisfies $\E_{O_t}[\widehat{r^a}(t+1) | O_t] = Y_{A_a(t+1),t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $O_t$.
 %

### 2019-06-06 2-Chapters/2-Chapter/chapter2.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment

@@
 
-% \TODOL{Je dois unifier les notations, et utiliser $Y_{k,t}$ pour les flux qui sont iid et $r(t)$ pour les récompenses effectivement observées (comme dans les chapitres 5 et 6)}
+% DONE Je dois unifier les notations, et utiliser $Y_{k,t}$ pour les flux qui sont iid et $r(t)$ pour les récompenses effectivement observées (comme dans les chapitres 5 et 6)
 
@@
     \end{align}
-\vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+% \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
 \label{par:2:notationsExponentialFamiliesBernoulliGaussian}
-% \begin{itemize}
-    % \item
-    $\bullet$
+\begin{itemize}
+    \item
+    % $\bullet$
     \textbf{Bernoulli distributions} can be seen as an exponential family with $h(x) = 1$,
@@
     (with the limit behavior $\eta=+\infty$ if $\mu=0$).
-    %
-    Moreover, the KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
+
+    The KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
     $\kl(x,y) = \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
@@
 
-    $\bullet$
+    \item
+    % $\bullet$
     \textbf{Gaussian distributions} of a fixed variance are a one-dimensional exponential family,
@@
     and for a Gaussian distribution with mean $\mu\in\R$, denoted $\cN(\mu,\sigma^2)$,
-    the parameter is $\eta = \mu/\sigma^2$, giving $A(\eta) = \mu^2/(2\sigma^2)$
-    (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
+    the parameter is $\eta = \mu/\sigma^2$, giving $A(\eta) = \mu^2/(2\sigma^2)$.
     % % For a generic variance,
@@
     % with a fixed variance $\sigma^2$ form indeed a one-dimensional exponential family.
-    %
-    Moreover, the KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
-    $d_{\sigma^2}(x,y) = \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$.
-% \end{itemize}
+
+    The KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
+    $d_{\sigma^2}(x,y) = \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$
+    (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
+\end{itemize}
 
@@
     \end{align}
-    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+% \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
     \end{align}
-    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    % \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 %
-\begin{proof}\label{proof:2:RegretDecomposition}
+\begin{smallproof}\label{proof:2:RegretDecomposition}
     We use the chain rule of expectation and a conditioning on $O_t$,
@@
     The sum can then be simplified to only account for sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
-\end{proof}
-
-\textbf{Consequences.}
+\end{smallproof}
+
+\paragraph{Consequences.}
 %
@@
     \end{equation}
-    \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    % \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 %
-\begin{proof}\label{proof:2:worstCaseLowerBound}
+\begin{smallproof}\label{proof:2:worstCaseLowerBound}
     For an example, consider any large horizon $T>10$, and $K=2$ Bernoulli arms.
@@
     $\varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T$.
-\end{proof}
+\end{smallproof}
 
@@
 %
-\begin{proof}\label{proof:2:tuningExploreThenCommit}
+\begin{smallproof}\label{proof:2:tuningExploreThenCommit}
     Let us prove the inequality $p \leq 2 K \exp(-T_0 d^2 / 4)$.\\
@@
     In all cases, we showed that $p$ is controlled by $T_0$.
-\end{proof}
+\end{smallproof}
 
@@
 %
-\begin{proof}\label{proof:2:UCBregretBound}
+\begin{smallproof}\label{proof:2:UCBregretBound}
     As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
@@
     More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs, in particular for any value of $\alpha\geq\frac{1}{2}$.
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
     Fix $p\in[0,1]$, and consider the function $g(x) = \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
@@
     % \end{verbatim}
-\end{proof}
-
-% \TODOL{Write this ?}
+\end{smallproof}

### 2019-06-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment

@@
 
-\begin{proof}
+\begin{smallproof}
   Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
@@
   Which is exactly the decomposition we wanted to prove.
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
   Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
@@
   which proves that $R_T(\boldsymbol{\mu}, M, \cA) \geq \ref{eq:5:term1}$, as wanted.
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
   Recall that we want to upper bound
@@
   and Lemma~\ref{lem:5:1stUpperBound} follows.
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
   %
@@
   with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
   We analyze the case when the current arm leaves the set $\TopM$ (Line 4):
@@
   Summing over $k'$ yields the result.
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
   A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small.
@@
   \end{align}
-\end{proof}
+\end{smallproof}
 

### 2019-06-06 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment

@@
 
-\begin{proof}
+\begin{smallproof}
     First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
@@
     and $(n-s) \times \kl(\widehat{\mu}_{s+1:n}, \widehat{\mu}_{1:n})$.
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
 Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
@@
 %
-\end{proof}
+\end{smallproof}
 
@@
 \end{lemma}
-\begin{proof}
+\begin{smallproof}
     This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
-\end{proof}
+\end{smallproof}
 
@@
 %
-\begin{proof}
+\begin{smallproof}
     It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proved in the proofs.
@@
 Je ne vois pas le problème de le laisser là, ça formalise le résultat mais on a pas besoin de celui là, ça permet de bien formaliser ce qu'on entend par contrôler le délai, et dans la preuve plus bas on prouve ce qu'on souhaite.}
-\end{proof}
+\end{smallproof}
 
@@
 
-\begin{proof}
+\begin{smallproof}
     We consider one arm $i\in[K]$, and when the \GLRklUCB{} algorithm is running,
@@
     % Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
-\end{proof}
+\end{smallproof}
 

### 2019-06-06 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Lots of typos changed, I'm reading all the thesis very quickly

@@
 
-Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
-%
 \begin{small}
@@
 					$\Alg_a$ updates its internal state (e.g., \UCB{} indexes)\;
-					It chooses $A_a(t+1) \in [K]$.
+					It chooses $A_a(t) \in [K]$.
 				}
@@
 				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
-				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(\{A_a(t+1) = j\}), \;\forall 1\leq j \leq K$\;
+				Let $p^{t}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(\{A_a(t) = j\}), \;\forall 1\leq j \leq K$\;
 				% \tcp*{Sum trusts}
-				Then $\Alg_{\mathrm{aggr}}$ chooses arm $A(t+1) \sim p^{t+1}$\;
-				% $\P(A(t+1)_{\mathrm{aggr}} = j) = p^{t+1}_j$
+				Then $\Alg_{\mathrm{aggr}}$ chooses arm $A(t) \sim p^{t}$\;
+				% $\P(A(t)_{\mathrm{aggr}} = j) = p^{t}_j$
 				% \tcp*{Sample once}
 				%
-				Give \emph{original} reward $(A(t+1), r(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
-				Compute an \emph{unbiased} estimate of the reward of the trusted algorithms,$$ \widehat{r}(t+1) = r(t+1) / p^{t+1}_{A(t+1)}\;. $$
+				Give \emph{original} reward $(A(t), r(t))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
+				Compute an \emph{unbiased} estimate of the reward of the trusted algorithms,$$ \widehat{r}(t) = r(t) / p^{t}_{A(t)}\;. $$
 				% \tcp{Use bandit feedback to update $\pi^{t}$}
-				% \tcp{Finally, use the instantaneous bandit feedback (reward $\ell_{A(t+1),t}$) to update $\pi^{t}$}
+				% \tcp{Finally, use the instantaneous bandit feedback (reward $\ell_{A(t),t}$) to update $\pi^{t}$}
 				% \For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
 				\For{$a = 1, \dots, N$}{
-					% \eIf{$\Alg_a$ was trusted, \ie, $A_a(t+1) = A(t+1)$}{
-					\eIf{$\Alg_a$ was trusted, \ie, $A_a(t+1) = A(t+1)$}{
-						$ \pi^{t+1}_{a} = \exp(\eta_t r(t+1) \times \pi^{t}_{a} $
+					% \eIf{$\Alg_a$ was trusted, \ie, $A_a(t) = A(t)$}{
+					\eIf{$\Alg_a$ was trusted, \ie, $A_a(t) = A(t)$}{
+						$ \pi^{t+1}_{a} = \exp(\eta_t r(t) \times \pi^{t}_{a} $
 						\tcp*{Trusted more}
 						}{
-						% $ \pi^{t+1}_{a} = \exp(- \eta_t r(t+1) \times \pi^{t}_{a} $
+						% $ \pi^{t+1}_{a} = \exp(- \eta_t r(t) \times \pi^{t}_{a} $
 						$ \pi^{t+1}_{a} = \pi^{t}_{a} $
@@
 
-\TODOL{Etre sur des notations entre $t$ et $t+1$ pour faire comme le reste du chapitre et de la thèse !}
-
-At every time step, after having observed a reward $r(t+1)=Y_{A(t+1),t+1}$ for its chosen action $A(t+1)$,
+Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
+%
+At every time step, after having observed a reward $r(t)=Y_{A(t),t}$ for its chosen action $A(t)$,
 the algorithm updates the trust probabilities from $\pi^t$ to $\pi^{t+1}$ by
@@
 
-The reward estimate is unbiased in the following sense.
-If one had access to the samples $Y_{k,t+1}$ for all arms $k$, the reward incurred by algorithm $a$ at time $t+1$ would be $r^a(t+1) = Y_{A_a(t+1),t+1}$.
-But we are not in the full information setting (see \cite{CesaLugosi06}), but in the bandit setting, so only one of the samples $Y_{k,t+1}$ can be observed, $r(t)=Y_{A(t+1),t+1}$.
-This quantity can only be observed for those algorithms $a$ for which $A_{a}^{t+1}=A(t+1)$.
-However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $r^a(t+1)$. More precisely, if we define the estimate by
+The reward estimate is unbiased in the following sense. If one had access to the samples $Y_{k,t}$ for all arms $k$, the reward incurred by algorithm $a$ at time $t$ would be $r^a(t) = Y_{A_a(t),t}$.
+But we are not in the full information setting (see \cite{CesaLugosi06}), but in the bandit setting, so only one of the samples $Y_{k,t}$ can be observed, $r(t)=Y_{A(t),t}$.
+This quantity can only be observed for those algorithms $a$ for which $A_{a}^{t}=A(t)$.
+However, by dividing by the probability of observing this recommendation, one obtains an unbiased estimate of $r^a(t)$. More precisely, if we define the estimate by
 \begin{equation}
-	\widehat{r^a}(t+1) = \frac{Y_{A_a(t+1),t+1}}{p_{A_a(t+1)}^{t+1}} \indic(A_a(t+1) = A(t+1))
+	\widehat{r^a}(t) = \frac{Y_{A_a(t),t}}{p_{A_a(t)}^{t}} \indic(A_a(t) = A(t))
 \end{equation}
-satisfies $\E_{O_t}[\widehat{r^a}(t+1) | O_t] = Y_{A_a(t+1),t+1}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $O_t$.
+satisfies $\E_{O_t}[\widehat{r^a}(t) | O_t] = Y_{A_a(t),t}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $O_t$.
 %
-Observe that $\widehat{r^a}(t+1) = 0$ for all algorithms $a$ such that $A_a(t+1) \neq A(t+1)$.
+Observe that $\widehat{r^a}(t) = 0$ for all algorithms $a$ such that $A_a(t) \neq A(t)$.
 
 % - Precise some tricky points
-An important feature of \Aggr{} is the feedback provided to each underlying bandit algorithm, upon the observation of arm $A(t+1)$.
-Rather than updating only the trusted algorithms (that is the algorithms which would have drawn arm $A(t+1)$) with the observed reward $r(t+1)$, we found that updating each algorithm with the (original) reward observed for arm $A(t+1)$ improves the performance drastically.
+An important feature of \Aggr{} is the feedback provided to each underlying bandit algorithm, upon the observation of arm $A(t)$.
+Rather than updating only the trusted algorithms (that is the algorithms which would have drawn arm $A(t)$) with the observed reward $r(t)$, we found that updating each algorithm with the (original) reward observed for arm $A(t)$ improves the performance drastically.
 %
@@
 and an alternative would be to
-increase $\pi_{a}$ if $A_a(t+1) = A(t+1)$ and to decrease it otherwise.
+increase $\pi_{a}$ if $A_a(t) = A(t)$ and to decrease it otherwise.
 It would not be so different, as there is a final renormalization step, and empirically this variation has little impact on the performance of \Aggr{}.
@@
 The first difference is that $a\sim\pi^t$ is sampled first and the arm chosen by $\Alg_a$ is trusted, whereas \Aggr{}
-needs to listen to the $N$ decisions to perform the updates on $\pi^{t+1}$.
+needs to listen to the $N$ decisions to perform the updates on $\pi^{t}$.
 Then, \ExpQ{} gives back an observation (arm, reward) only to the last trusted algorithm
@@
 %
-The second difference is that after having computed the reward estimate $\widehat{r^a}(t+1)$, \ExpQ{} updates the estimated cumulative reward for each algorithm,
-$\widetilde{R}_a(t) = \sum_{s=1}^{t} \widehat{r^a}(t+1) \times \indic(A^{s}_{a} = A^{s}_{\mathrm{aggr}})$.
+The second difference is that after having computed the reward estimate $\widehat{r^a}(t)$, \ExpQ{} updates the estimated cumulative reward for each algorithm,
+$\widetilde{R}_a(t) = \sum_{s=1}^{t} \widehat{r^a}(t) \times \indic(A^{s}_{a} = A^{s}_{\mathrm{aggr}})$.
 %
@@
 Note that for each of the $1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{1\leq k \leq K, 1 \leq t \leq T}$ for every repetition, in order to compare the algorithms on the same realizations%
-\footnote{Similar plots and similar results are obtained if this choice is not made, but it makes more sense to compare them against the same randomization.
+\footnote{~Similar plots and similar results are obtained if this choice is not made, but it makes more sense to compare them against the same randomization.
 Note that this choice is the default in SMPyBandits, but the other possibility is implemented as well, by changing \texttt{cache\_rewards} to \texttt{true} or \texttt{false} in the \texttt{configuration} dictionary.
@@
 This is in contrast with the (classical) regret, which measure the difference with the best fixed-arm strategy (Definition~\ref{def:2:regret}).
-Thus, even if the aggregated algorithms have logarithmic (classical) regret, having an adversarial regret scaling as $\sqrt{T}$ does not permit to exhibit a logarithmic (classical) regret for the aggregation algorithm.
+Thus, even if the aggregated algorithms have logarithmic (classical) regret, having an adversarial regret scaling as $\sqrt{T}$ does not permit to obtain a logarithmic (classical) regret for the aggregation algorithm.
 %
@@
 with the good choice of the learning rate sequence $(\eta_t)_{t \geq 1}$.
+
 Our proposed algorithm follows quite closely the architecture of \ExpQ,
 and a similar bound for \Aggr{} is expected to hold.
-% this is left as a future work.
-% For sake of conciseness, we cannot present a proof here.
 %
-
 This would be a first theoretical guarantee, but not satisfactory as we saw above that simple algorithms (like \UCB) have regrets scaling as $\log(T)$ \cite{Auer02,Bubeck12}, not $\sqrt{T}$.
@@
 % Further theoretical developments were left as future work.
-
-\TODOL{Expliquer qu'en fait on NE PEUT PAS obtenir la borne qu'on espérait...}
 
@@
 What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
-
-Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent \cite{Hazan2016introduction,Zimmert2018}, as it is the case for both \ExpQ{} and \CORRAL.
+%
+%
+Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
 But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.

### 2019-06-06 2-Chapters/2-Chapter/chapter2.tex
Lots of typos changed, I'm reading all the thesis very quickly

@@
 
-% First chapter begins here
 \chapter{Stochastic Multi-Armed Bandits Models}
@@
 
-\textbf{Abstract.}
+\graphicspath{{2-Chapters/2-Chapter/Images/}}
+
+\paragraph{Abstract.}
 %
@@
 \minitoc
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/2-Chapter/Images/}}
+% ----------------------------------------------------------------------
 
@@
 %
-It shows that the performance can be \textcolor{red}{very random}, as on both cases the pure exploitation strategy was very lucky and obtains close-to-optimal performance, even though we proved that it suffers from a linear mean regret.
+It shows that the performance can \textcolor{red}{fluctuate a lot}, as on both cases the pure exploitation strategy was very lucky and obtains close-to-optimal performance, even though we proved that it suffers from a linear mean regret.
 
@@
                 & $6949$ & $44.2$ \\
-            \hline
         \end{tabular}
@@
                 & $\mathbf{6951}$ & $\mathbf{49}$ \\
-            \hline
         \end{tabular}
@@
     \includegraphics[width=1.15\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
-	\caption{Average of rewards, as function of $t$, for $T=10000$ and $N=1000$.}
+	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
 	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}

### 2019-06-06 2-Chapters/4-Chapter/chapter4.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Done for a large proofreading tonight
Lots of typos changed, I'm reading all the thesis very quickly

@@
 \label{chapter:4}
+
+\graphicspath{{2-Chapters/4-Chapter/Images/}}
+
+\TODOL{Abstract ici et plus court, et petite introduction !}
+\TODOL{Intégrer les derniers commentaires de Christophe}
+
+\paragraph{Abstract.}
+
+In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithms to learn and improve their access to the network.
+FIXME finish !
+
 \minitoc
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/4-Chapter/Images/}}
 % ----------------------------------------------------------------------
 
-\TODOL{Abstract ici et plus court, et petite introduction !}
+\newpage
 
-\newpage
-\paragraph{Abstract}
+\paragraph{Abstract.}
 
-In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithm to learn and improve their access to the network.
+In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithms to learn and improve their access to the network.
 The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
@@
 % ----------------------------------------------------------------------
-% \paragraph{Summary of the first model}
+% \paragraph{Summary of the first model.}
 \label{sub:41:conclusion}
@@
 % ----------------------------------------------------------------------
-% \paragraph{Conclusions taken from the demonstration}
+% \paragraph{Conclusions taken from the demonstration.}
 \label{sub:42:conclusionFromDemonstration}
@@
 % ----------------------------------------------------------------------
-% \paragraph{Conclusion about this second model}
+% \paragraph{Conclusion about this second model.}
 \label{sub:43:conclusion}
@@
 
-\paragraph{Multi-Player MAB with \emph{collision avoidance}}
+\paragraph{Multi-Player MAB with \emph{collision avoidance}.}
 

### 2019-06-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Lots of typos changed, I'm reading all the thesis very quickly

@@
 
-\paragraph{Publications.}
-%
-This chapter is mainly based on our article \cite{Besson2018ALT}.
-
-\paragraph{Outline.}
+\textbf{Outline.}
 %
@@
 % and finally Section~\ref{sec:5:conclusion} concludes.
+
+\textbf{Publication.}
+%
+This chapter is mainly based on our article \cite{Besson2018ALT}.
 
@@
 
+\TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}
+
 % \vspace*{-5pt}  % XXX remove if problem
@@
               }
-              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for a fixed underlying index policy $g^j$).}
+              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $g^j$).}
           \label{algo:5:RandTopM}
@@
               }
-              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for a fixed underlying index policy $g^j$).}
+              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $g^j$).}
           \label{algo:5:MCTopM}
@@
   We analyze the case when the current arm leaves the set $\TopM$ (Line 4):
+
   \begin{align*}
@@
     \E_{\mu}\left[\sum_{k=1}^K \cC_k(T)\right]
-    \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
+    \leq M^2\left(2 M + 1\right) \left(\sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
   \end{equation}
@@
     &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
-    &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
-    &\leq \left(2 M^3 + M^2\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
+    &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
+    &\leq \left(2 M^3 + M^2\right) \left(\sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
   \end{align}
@@
     \sum_{k=1}^K \E_{\mu}[\cC^k(T)] & = \E_{\mu}[\cC(T)] \nonumber\\
-    & \leq M^2\left(2 M + 1\right) \left(\sum_{a,b=1,\dots,K,\;\mu_a < \mu_b} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
+    & \leq M^2\left(2 M + 1\right) \left(\sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
   \end{align}
@@
 
-\paragraph{Logarithmic ``switching costs'' for \MCTopM.}\label{app:5:NumberSwitches}
+\paragraph{Logarithmic switching cost for \MCTopM.}\label{app:5:NumberSwitches}
 %
@@
 %
-As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proven
+As soon as $R_T = \bigO{\log T}$ for all problems, \MCTopM{} is clearly proven
 to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
@@
           & \klUCB{} & $\mathbf{248 \pm 40}$ & $\mathbf{410 \pm 54}$ & $\mathbf{42 \pm 10}$ \\
-          \hline
+          % \hline
       \end{tabular}

### 2019-06-06 2-Chapters/5-Chapter/chapter5.tex
Added leftbar to all defn,cor,thm etc
Lots of typos changed, I'm reading all the thesis very quickly

@@
 \label{chapter:5}
+
+\graphicspath{{2-Chapters/5-Chapter/Images/}}
+\graphicspath{{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/}}
 
@@
 \minitoc
+% ----------------------------------------------------------------------
 
 \newpage
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/5-Chapter/Images/}}
-\graphicspath{{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/}}
 
@@
 We presented in Section~\ref{sec:5:model} the model \modeltrois, without sensing information.
-Players can only observe $Y_k(t)$, the product of the \iid{} random sensing information from arm $k$ and of the no-collision indicator.
+Players can only observe $Y_{k,t}$, the product of the \iid{} random sensing information from arm $k$ and of the no-collision indicator.
 This model seems harder than the model with sensing information, and even though the proposal \Selfish{} works fine in simulations (in terms of average regret), we proved that it can yield linear regret.
@@
 To illustrate the difficulty of the ``no sensing'' case, we illustrate here the performances of different algorithms designed specifically for this setting.
-As above, we consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
+As above, we consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $N=100$ independent repetitions, and horizon $T=10000$.
 We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
-The centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, but it is interesting to also add it here, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, as it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
+It is also interesting to add the centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, because it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
 %
-For the two other algorithms, we use the parameters advised in all papers: for \textsc{Sic-MMAB}, we used $T_0 = \lceil K \log(T) \rceil$ as advised by the authors.
-For Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis.
+For the two other algorithms, we use the advised tuning of their parameters:
+for \textsc{Sic-MMAB}, we used $T_0 = \lceil K \log(T) \rceil$,
+% as advised by the authors.
+for Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis.
 We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, however we also experimented with larger values of $T$ and found a same linear behavior.}.
@@
         \textsc{Sic-MMAB} & $1705 \pm 340$ & $3915 \pm 300$ & $1713 \pm 52$ \\
-        Improved \MusicalChair{} & $\textcolor{red}{12149 \pm 60}$ & $\textcolor{red}{22341 \pm 77}$ & $\textcolor{red}{27462 \pm 85}$ \\
+        \textcolor{red}{Improved \MusicalChair{}} & $\textcolor{red}{12149 \pm 60}$ & $\textcolor{red}{22341 \pm 77}$ & $\textcolor{red}{27462 \pm 85}$ \\
         \hline
@@
     % \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on the same problem with $M=3,6,9$, for the ``no sensing'' case.}
-    \caption{Comparison for different algorithms on the same problem with $M=3,6,9$ players, for the ``no sensing'' case.
-We give the mean regret $\pm$ $1$ std-dev, and more work is needed on our implementation on Improved \MusicalChair.
+    \caption{Comparison of the mean regret $\pm$ $1$ std-dev, for different algorithms, on the same problem with $M=3,6,9$ players, for the ``no sensing'' case.
+More work is needed on our implementation on Improved \MusicalChair.
 The results on \textsc{Sic-MMAB} appear to confirm the numerical experiments of \cite{BoursierPerchet18}.}
@@
 
-% ----------------------------------------------------------------------------
-\subsection{Some future work on SMPyBandits}
-
-The experiments in this chapter use SMPyBandits,
-and we refer to the following page for details on how to reproduce them,
-\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}}
-
-Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
-We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future work left on our library.
-For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
-
-% \newpage  % FIXME
+\newpage  % FIXME
 % ----------------------------------------------------------------------------
@@
 
+Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
+We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future work left on our library.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
+
+% ----------------------------------------------------------------------------
+% \subsection{Some future work on SMPyBandits}
+\paragraph{Reproducility of the experiments.}
+
+The experiments in this chapter use SMPyBandits,
+and we refer to the following page for details on how to reproduce them,\\
+\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}}
+
 % -----------------------------------------------------------------

### 2019-06-06 2-Chapters/6-Chapter/chapter6.tex
Lots of typos changed, I'm reading all the thesis very quickly

@@
 
-\paragraph{Abstract.}
+\graphicspath{{2-Chapters/6-Chapter/Images/}}
 
-In this last chapter, we study a generalization of the stationary multi-armed bandit model,
-to account for possible non-stationarity of the rewards.
-% % that is, for problems that could be tackled with MAB algorithms.
-% Instead of dealing with changes happening possibly at every time step, like in adversarial models, we rather consider a finite small number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
-%
-We review existing works on piece-wise stationary MAB models.
-% for which the two main families of proposed solutions are either passively or actively adaptive.
+% \paragraph{Abstract.}
+
+% In this last chapter, we study a generalization of the stationary multi-armed bandit model,
+% to account for possible non-stationarity of the rewards.
+% % % that is, for problems that could be tackled with MAB algorithms.
+% % Instead of dealing with changes happening possibly at every time step, like in adversarial models, we rather consider a finite small number $\Upsilon_T$ of \emph{change-points}, and we consider problems that are stationary on any interval between two consecutive change-points.
+% %
+% We review existing works on piece-wise stationary MAB models.
+% % for which the two main families of proposed solutions are either passively or actively adaptive.
 In practice, it was already observed that actively adaptive strategies usually greatly outperform the passively adaptive ones.
-% The more efficient strategies from the current state-of-the-art are based on a combination of an efficient algorithm for the stationary problem with an efficient online change detection algorithm, like CUSUM, PHT or the Generalized Likelihood Ratio Test (GLRT), and so we pursue on this direction.
-%
-We study the Generalized Likelihood Ratio Test (GLRT) for
-% Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any 
-bounded or sub-Bernoulli distributions.
-Our test enjoys finite-time guarantees
-% its \emph{false alarm probability}, meaning that in a stationary model, the test should not detect any change,
-% and its \emph{detection delay}, meaning that in a model with a change-point, the test should detect within a reasonable delay.
-%
-of its \emph{false alarm probability} and \emph{detection delay}, and can be combined with an efficient bandit policy (\klUCB), to propose an efficient algorithm for piece-wise stationary problems, \GLRklUCB.
-% Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
-% % \MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
-%
-% We give the first unified analysis of a piece-wise stationary bandit algorithm for the two variants, and our analysis
-We analyze its regret, and show that it achieves state-of-the-art non-asymptotic regret bounds.
-% and yield regret bounds with explicit constants.
-% Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$, when the algorithm runs with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
-%
-Finally, we give extensive numerical experiments, to compare our approach against other state-of-the-art solutions.
-% , on different piece-wise stationary bandit problems.
-% % , of increasing difficulty.
+% % The more efficient strategies from the current state-of-the-art are based on a combination of an efficient algorithm for the stationary problem with an efficient online change detection algorithm, like CUSUM, PHT or the Generalized Likelihood Ratio Test (GLRT), and so we pursue on this direction.
 % %
-% We also include additional details about our policy \GLRklUCB, including a sensitivity analysis of its two parameters, a numerical evaluation of its time and memory costs, as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
+% We study the Generalized Likelihood Ratio Test (GLRT) for
+% % Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any 
+% bounded or sub-Bernoulli distributions.
+% Our test enjoys finite-time guarantees
+% % its \emph{false alarm probability}, meaning that in a stationary model, the test should not detect any change,
+% % and its \emph{detection delay}, meaning that in a model with a change-point, the test should detect within a reasonable delay.
+% %
+% of its \emph{false alarm probability} and \emph{detection delay}, and can be combined with an efficient bandit policy (\klUCB), to propose an efficient algorithm for piece-wise stationary problems, \GLRklUCB.
+% % Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
+% % % \MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
+% %
+% % We give the first unified analysis of a piece-wise stationary bandit algorithm for the two variants, and our analysis
+% We analyze its regret, and show that it achieves state-of-the-art non-asymptotic regret bounds.
+% % and yield regret bounds with explicit constants.
+% % Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$, when the algorithm runs with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
+% %
+% Finally, we give extensive numerical experiments, to compare our approach against other state-of-the-art solutions.
+% % , on different piece-wise stationary bandit problems.
+% % % , of increasing difficulty.
+% % %
+% % We also include additional details about our policy \GLRklUCB, including a sensitivity analysis of its two parameters, a numerical evaluation of its time and memory costs, as well as details on an easy but efficient way to speed up the B-GLRT test while not loosing too much in terms of regrets.
 
 \minitoc
+% ----------------------------------------------------------------------
 
@@
 
-% \newpage
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/6-Chapter/Images/}}
-
 % This chapter is basically a raw include from my paper ``The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits'', see https://hal.inria.fr/hal-02006471
-
-% \TODOL{Rewrite and reorder most of this chapter, with more subsections, less paragraphs etc, and less copy-paste of our paper \cite{Besson2019GLRT}.}
 

### 2019-06-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
CHAUD

@@
 \begin{equation}
-  A^j(t) := \pi\left(R^j(t), [g^j_\ell(t)]_{\ell=1,\dots,K}\right).
+  A^j(t) := \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right).
 \end{equation}
@@
 \begin{equation}
-  \TopM(t) := \left\{ \pi\left(k, \{g^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
+  \TopM(t) := \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
 \end{equation}
@@
                       }(\tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}){
-                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
+                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                       }
@@
                   Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
-                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
+                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
               }
-              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $g^j$).}
+              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
           \label{algo:5:RandTopM}
@@
                     {
-                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
+                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                       \tcp*[f]{not empty} \\
@@
                   Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
-                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
+                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
               }
-              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $g^j$).}
+              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
           \label{algo:5:MCTopM}
@@
 a new arm is sampled uniformly from a subset of $\TopM(t)$,
-defined with the previous indices $g^j(t-1)$ (Lines~$3$-$5$).
+defined with the previous indices $U^j(t-1)$ (Lines~$3$-$5$).
 %
 The subset enforces a certain inequality on indices,
-$g_{k'}^j(t-1) \leq g^j_{k}(t-1)$ and $g_{k'}^j(t) \geq g^j_{k}(t)$,
+$U_{k'}^j(t-1) \leq U^j_{k}(t-1)$ and $U_{k'}^j(t) \geq U^j_{k}(t)$,
 when switching from $k=A^j(t)$ to $k'=A^j(t+1)$.
@@
   %&= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
-  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in[M]:\; g_{m^*}^j(t) < g_k^j(t) \right).
+  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in[M]:\; U_{m^*}^j(t) < g_k^j(t) \right).
   %
@@
 
-  Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
+  Considering the relative position of the upper-confidence bound $U_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
 
@@
     \E_{\mu}[T_k^j(T)] &\leq
-    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; g_{m_1^*}(t) < \mu_m^* \right) + \notag \\
+    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; U_{m_1^*}(t) < \mu_m^* \right) + \notag \\
     \;\;\;\;
-    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; g_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in [M]: \;  g_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
-    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(g_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
-    %
-    & \leq \sum_{m_1=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( g_{m_1^*}(t) < \mu_{m_1}^* \right)
+    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
+    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
+    %
+    & \leq \sum_{m_1=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( U_{m_1^*}(t) < \mu_{m_1}^* \right)
     + \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
@@
   \begin{align*}
-    & \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
-    & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
-    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
+    & \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': U_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
+    & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right) \\
+    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{:= T_{k'}}
   \end{align*}
@@
   Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
-  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
+  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right)\]
   The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
@@
   \begin{align}
-    \sum_{t=1}^T & \Pr\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
+    \sum_{t=1}^T & \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
     & =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
@@
   If $\mu_k > \mu_{k'}$, we rather use that
-  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \Pr\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
+  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \Pr\left(A^j(t+1) = k', U_{k'}^j(t) \geq \mu_{k}\right)\]
   and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields

### 2019-06-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
I have some small bugs, but all is good

@@
 
-Under observation model \modeltrois{} no sensing information is available and the previous algorithms cannot be used, as the sum of sensing information $S_k^j(t)$ and thus the empirical mean $\widehat{\mu}_k^j(t)$ cannot be computed, hence neither the indices $g_k^j(t)$.
+Under observation model \modeltrois{} no sensing information is available and the previous algorithms cannot be used, as the sum of sensing information $S_k^j(t)$ and thus the empirical mean $\widehat{\mu}_k^j(t)$ cannot be computed, hence neither the indices $U_k^j(t)$.
 However, one can still define a notion of \emph{empirical reward} received from arm $k$ by player $j$, by introducing
@@
 \begin{equation}\label{eq:5:indexTilde}
-  \widetilde{g_k}^j(t) = \begin{cases}
+  \widetilde{U_k}^j(t) = \begin{cases}
       \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
@@
 \begin{equation}
-  A^j(t) = \argmax_{ k \in [K]} \ \widetilde{g_k}^j(t-1).\label{algo:5:Selfish}
+  A^j(t) = \argmax_{ k \in [K]} \ \widetilde{U_k}^j(t-1).\label{algo:5:Selfish}
 \end{equation}
 The name comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
-The reason that this may work precisely comes from the fact that $\widetilde{g_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
+The reason that this may work precisely comes from the fact that $\widetilde{U_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
 but some hybrid index that simultaneously increases when a transmission occurs and decreases when a collision occurs.
@@
 \begin{equation}
-  \widetilde{g_k}^j(t) = g_k^j(t) -
+  \widetilde{U_k}^j(t) = U_k^j(t) -
   \underbrace{\left(\frac{N_k^{j,C}(t)}{N_k^j(t)}\right)}_{\text{fraction of collisions}}
@@
 \begin{equation}\label{eq:5:indexFor_UCB_klUCB}
-  g_k^j(t) := \begin{cases}
+  U_k^j(t) := \begin{cases}
       \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
@@
 where $f(t)$ is some \emph{exploration function}.
-$f(t)$ is usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(g_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
+$f(t)$ is usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
 A classical (single-player) UCB algorithm aims at the arm with largest index.
@@
                       \tcp*[f]{randomly switch}
-                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}){
-                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
+                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB}){
+                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                       }
@@
                     {
-                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
+                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                       \tcp*[f]{not empty} \\
-                      % \tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}
+                      % \tcp*[f]{randomly switch on an arm that had smaller UCB}
                       $s^j(t+1) = \mathrm{False}$
@@
   %
-  %     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
+  %     Recall $U_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
   %     As only user $j$ is considered here, the superscript $j$ is dropped.
@@
   The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
-  \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
+  \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : U_m^j(t) < U_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
   Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
@@
   = \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k \right).\\
-  %&= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
-  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in[M]:\; U_{m^*}^j(t) < g_k^j(t) \right).
+  %&= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\Mbest,\; U_m(t) < U_k(t) \right) \\
+  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in[M]:\; U_{m^*}^j(t) < U_k^j(t) \right).
   %
@@
     \;\;\;\;
-    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
-    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
+    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq U_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
+    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq U_k(t)\right) \notag \\
     %
     & \leq \sum_{m_1=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( U_{m_1^*}(t) < \mu_{m_1}^* \right)
-    + \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+    + \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq U_k(t) \right)
     \label{eq:5:sumOfProbaForTkj}
@@
   \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-      For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
+      For any arm $k$, if $U_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
       \begin{equation}
-        \sum_{t=1}^T \Pr_{\bm{\mu}} \left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
+        \sum_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
       \end{equation}
@@
   \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
+    For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $U_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
     \begin{small} % XXX
       \begin{align*}
-        &\sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
+        &\sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq U_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
         + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.
@@
   \begin{align*}
-    & \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': U_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
+    & \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': U_{k'}^j(t-1) \leq U_k^j(t-1)\}\right) \\
     & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right) \\
@@
   Either she sampled $A^j(t+1)$ uniformly
-  from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
+  from $\TopM(t) \cap \{ m : U_m^j(t) \leq U_k^j(t) \}$
   with transition $(3)$,

### 2019-06-06 2-Chapters/6-Chapter/chapter6.tex
CHAUD
I have some small bugs, but all is good

--- 2019-06-06 2-Chapters/6-Chapter/chapter6.tex
@@
 \label{chapter:6}
+
+\graphicspath{{2-Chapters/6-Chapter/Images/}}
 
@@
 %
-Finally, we give extensive numerical experiments, to compare our approach against other state-of-the-art solutions.
+Finally, we compare our approach against other state-of-the-art solutions on extensive numerical experiments.
 % , on different piece-wise stationary bandit problems.
@@
 
+\TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
+
 \minitoc
+% ----------------------------------------------------------------------
 

### 2019-06-06 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Lots of typos changed, I'm reading all the thesis very quickly
I have some small bugs, but all is good

@@
 Numerical experiments are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.
-
-% \TODOL{Rename GLR, GLRT and GLR Bernoulli to B-GLRT like what we did in our GRETSI 2019 paper \cite{Besson2019Gretsi}.}
-
-% \vfill{}
 
@@
     \begin{enumerate}
-        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \Upsilon_T\sqrt{T\ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2}\sqrt{\Upsilon_T T \ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}$.
+        \item Choosing $\alpha = \sqrt{\ln(T)/T}$, $\delta = 1/\sqrt{T}$ gives
+        \begin{equation}
+            R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \Upsilon_T\sqrt{T\ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)},
+        \end{equation}
+        \item Choosing $\alpha = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ gives
+        \begin{equation}
+            R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2}\sqrt{\Upsilon_T T \ln(T)} + \frac{(K-1)}{\Delta^{\text{opt}}} \Upsilon_T\ln(T)}.
+        \end{equation}
     \end{enumerate}
@@
     \begin{enumerate}
-        \item Choosing $\alpha = \sqrt{\frac{\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{\EffChange\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{\EffChange T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)}$,
-        \item Choosing $\alpha = \sqrt{\frac{K\Upsilon_T\ln(T)}{T}}$, $\delta = \frac{1}{\sqrt{K\Upsilon_T T}}$ gives $R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}$.
+        \item Choosing $\alpha = \sqrt{\ln(T) / T}$, $\delta = 1 / \sqrt{T}$ gives
+        \begin{equation}
+            R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + 1 / \left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)},
+        \end{equation}
+        \item Choosing $\alpha = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ gives
+        \begin{equation}
+            R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)},
+        \end{equation}
+        \item Choosing $\alpha = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ gives
+        \begin{equation}
+            R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)},
+        \end{equation}
+        \item Choosing $\alpha = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ gives
+        \begin{equation}
+            R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}.
+        \end{equation}
     \end{enumerate}
@@
 %
-Note that from Assumption~\ref{ass:6:LongPeriodsGlobal}, as the period between two changes are long enough, if $\cE_T(\delta)$ holds, then for all change $k$, one has $\tau^{(k)} \leq \hat{\tau}^{(k)} \leq \tau^{(k+1)}$.
-Using Assumption~\ref{ass:6:LongPeriods}, one can prove the following.
+From Assumption~\ref{ass:6:LongPeriodsGlobal}, as the period between two changes are long enough, if $\cE_T(\delta)$ holds, then for all change $k$, we have $\tau^{(k)} \leq \hat{\tau}^{(k)} \leq \tau^{(k+1)}$.
+So we can prove the following.
 
@@
     With $\cE_T(\delta)$ defined as in \eqref{def:6:GoodEvenGlobal}, the ``bad event'' is unlikely:
-    $\bP(\cE_T^c(\delta)) \leq \delta(K+1)\Upsilon_T$.
+    \begin{equation}
+        \bP(\cE_T^c(\delta)) \leq \delta(K+1)\Upsilon_T.
+    \end{equation}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 
-\TODOL{Fix spacing issue!}
+% \TODOL{Fix spacing issue!}
 

### 2019-06-07 2-Chapters/2-Chapter/chapter2.tex
CHAUD
I have some small bugs, but all is good
Better font, and lots of small improvements

@@
 We conclude this chapter by discussing the question of online algorithm selection.
+
+% \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
 
@@
 
-\paragraph{Publication.}
+\textbf{Publication.}
 %
@@
 		}
-		\caption{Naive strategies: \textcolor{darkgreen}{pure exploitation} or \textcolor{blue}{pure exploration}.}
+		\caption[Naive strategies: pure exploitation or pure exploration.]{Naive strategies: \textcolor{darkgreen}{pure exploitation} or \textcolor{blue}{pure exploration}.}
 		\label{algo:2:naiveStrategies}
@@
 		}
-		\caption{Simple efficient strategies: \textcolor{deeppurple}{$\varepsilon$-greedy} and \textcolor{deepgold}{Explore-then-Exploit}.}
+		\caption[Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit.]{Simple efficient strategies: \textcolor{deeppurple}{$\varepsilon$-greedy} and \textcolor{deepgold}{Explore-then-Exploit}.}
 		\label{algo:2:simpleStrategies}
@@
 		}
-		\caption{A generic \textcolor{blue}{index} policy, using \textcolor{blue}{indexes $U_k(t)$} (\eg, $\UCB_1$, \klUCB{} etc).}
+		\caption[{A generic index policy, using indexes $U_k(t)$ (\eg, $\UCB_1$, \klUCB{} etc).}]{A generic \textcolor{blue}{index} policy, using \textcolor{blue}{indexes $U_k(t)$} (\eg, $\UCB_1$, \klUCB{} etc).}
 		\label{algo:2:indexPolicy}

### 2019-06-07 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Done changing notations \{1,\dots,M\} -> [M] all over the PhD
Changed \paragraph{XXX} to \paragraph{XXX.}
CHAUD
I have some small bugs, but all is good
Better font, and lots of small improvements

@@
 %       %
-%       \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $\{1,\dots,K\}$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
+%       \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
 %       %
@@
 % \end{eqnarray*}
-% where we denote by $\widehat{S}_t = \{A^j(t), j \in \{1,\dots,M\}\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
+% where we denote by $\widehat{S}_t = \{A^j(t), j \in [M]\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
 % \begin{eqnarray*}
@@
 % \[\forall t > 0, \ \ \cO_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
-% where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $\{1,\dots,M\}$ for new ranks or arms.} useful to select $A^j(t)$.
+% where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $[M]$ for new ranks or arms.} useful to select $A^j(t)$.
 % Formally, one can say that $A^j(t)$ is $\sigma(\cO_{t-1})$ measurable, if $\sigma(\cO_t)$ denotes the sigma-Algebra generated by the observations $\cO_t$, and as $\cF^j(t) \subseteq \sigma(\cO_{t})$, with an equality under observation model \modelun.
@@
 % %
-% %     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
+% %     Recall $U_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
 % %     As only user $j$ is considered here, the superscript $j$ is dropped.
 % %
-% Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
+% Fix $k\in\Mworst$ and a player $j \in [M]$.
 % The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
-% \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
+% \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : U_m^j(t) < U_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
 % Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
@@
 % = \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
-% %&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
-% &= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
+% %&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; U_m(t) < U_k(t) \right) \\
+% &= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in[M]:\; U_{m^*}^j(t) < U_k^j(t) \right).
 % %
-% \intertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
-% \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
-% &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
-% &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
+% \intertext{Considering the relative position of the upper-confidence bound $U_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
+% \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in[M]:\; U_{m^*}(t) \leq U_k(t) , \forall m \in [M]: \;  U_{m^*}(t) \geq \mu_m^* \right)\\
+% &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in[M]: \; U_{m^*}(t) < \mu_m^* \right) \\
+% &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in[M]:\; \mu_{m}^* \leq U_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(U_{m^*}(t) < \mu_m^* \right) \\
 % %
@@
 % \sum_{t=1}^T
-%   \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+%   \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq U_k(t) \right)
 % +
 % \sum_{m=1}^M
-%   \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right),
+%   \sum_{t=1}^T \Pr\left( U_{m^*}(t) < \mu_m^* \right),
 % \end{align*}
-% where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.
+% where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in [M]$.
 
@@
 % \begin{lemma}\label{lem:5:Fact1KLUCB}
-%     For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
+%     For any arm $k$, if $U_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
 %     \begin{equation}
-%       \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
+%       \sum_{t=1}^T \Pr\left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
 %     \end{equation}
@@
 
-% \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
+% \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $U_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
 % \begin{align*}
-%   &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
+%   &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq U_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
 %   + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
@@
 % \begin{align*}
-%   & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
-%   & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
-%   & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
+%   & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': U_{k'}^j(t-1) \leq U_k^j(t-1)\}\right) \\
+%   & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right) \\
+%   & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{:= T_{k'}}
 % \end{align*}
@@
 % Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
-% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
+% \[T_{k'} \leq \sum_{t=1}^T \bP\left(U_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right)\]
 % The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
@@
 % \begin{align}
-%   & \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)
+%   & \sum_{t=1}^T \bP\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right)
 %   =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
@@
 % If $\mu_k > \mu_{k'}$, we rather use that
-% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
+% \[T_{k'} \leq \sum_{t=1}^T \bP\left(U_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', U_{k'}^j(t) \geq \mu_{k}\right)\]
 % and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
@@
 %%%%%%%%%%%%ù PREVIOUS VERSION
-% For any arms $k, k' \in \{1, \dots, K\}$,
+% For any arms $k, k' \in [K]$,
 % let $\phi_{k,k'}^j(t)$ be the event
@@
 %     C^j(t),
-%     A^j(t) = k' \in \TopM(t) \cap \{ m : g_m^j(t-2) \leq g_k^j(t-2) \}
+%     A^j(t) = k' \in \TopM(t) \cap \{ m : U_m^j(t-2) \leq U_k^j(t-2) \}
 %   %
@@
 % This event $\phi_{k,k'}^j(t)$ we focus on implies both
-% $g_{k'}^j(t-1) \leq g_k^j(t-1), A^j(t-1)=k$
+% $U_{k'}^j(t-1) \leq U_k^j(t-1), A^j(t-1)=k$
 % and
-% $g_k^j(t) \leq g_{k'}^j(t), A^j(t)=k'$.
+% $U_k^j(t) \leq U_{k'}^j(t), A^j(t)=k'$.
 % %
@@
 % if $\mu_k \geq \mu_{k'}$,
-% $\phi_{k,k'}^j(t)$ implies $g_k^j(t) \leq g_{k'}^j(t)$,
+% $\phi_{k,k'}^j(t)$ implies $U_k^j(t) \leq U_{k'}^j(t)$,
 % and
 % if $\mu_k < \mu_{k'}$, then
-% $\phi_{k,k'}^j(t)$ implies $g_{k'}^j(t-1) \leq g_k^j(t-1)$,
+% $\phi_{k,k'}^j(t)$ implies $U_{k'}^j(t-1) \leq U_k^j(t-1)$,
 % both with $A^j(t-1)=k, A^j(t)=k'$.
@@
 % Assume $\mu_k \geq \mu_{k'}$.
-% Then, essentially $g_k^j(t) \leq g_{k'}^j(t)$
+% Then, essentially $U_k^j(t) \leq U_{k'}^j(t)$
 % is unlikely (\ie, the sum of probabilities is $\bigO{\log T}$)
-% as the indices $g_m^j(t)$ tends to $\mu_m$ if arm $m$ is sampled enough (\ie, $\E_{\mu}[T_m^j(T)] \to \infty$).
+% as the indices $U_m^j(t)$ tends to $\mu_m$ if arm $m$ is sampled enough (\ie, $\E_{\mu}[T_m^j(T)] \to \infty$).
 % \begin{align}
@@
 %     &\leq
-%     \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t), A^j(t-1)=k, A^j(t)=k'\right)
-%     \intertext{Either $g_{k'}^j(t) > \mu_k$ or $g_{k'}^j(t) \leq \mu_k$, and by introducing the two cases,}
+%     \sum_{t=1}^{T} \Pr\left(U_k^j(t) \leq U_{k'}^j(t), A^j(t-1)=k, A^j(t)=k'\right)
+%     \intertext{Either $U_{k'}^j(t) > \mu_k$ or $U_{k'}^j(t) \leq \mu_k$, and by introducing the two cases,}
 %     &\leq
-%     \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t) \leq \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag \\
-%     &\;\;+ \sum_{t=1}^{T} \Pr\left(g_k^j(t) \leq g_{k'}^j(t), g_{k'}^j(t) > \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag
-%     \intertext{For the first term, $g_k^j(t) \leq \mu_k$ is highly unlikely (\ie, the sum of probabilities is $\smallO{\log T}$) as the indices ${(g_k^j)}_k$ are Upper Confidence Bounds (if arm $k$ is sampled enough).
-For the second term, $g_{k'}^j(t) > \mu_k$ is unlikely if arm $k'$ is sampled enough.
+%     \sum_{t=1}^{T} \Pr\left(U_k^j(t) \leq U_{k'}^j(t) \leq \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag \\
+%     &\;\;+ \sum_{t=1}^{T} \Pr\left(U_k^j(t) \leq U_{k'}^j(t), U_{k'}^j(t) > \mu_k, A^j(t-1)=k, A^j(t)=k'\right) \notag
+%     \intertext{For the first term, $U_k^j(t) \leq \mu_k$ is highly unlikely (\ie, the sum of probabilities is $\smallO{\log T}$) as the indices ${(U_k^j)}_k$ are Upper Confidence Bounds (if arm $k$ is sampled enough).
+For the second term, $U_{k'}^j(t) > \mu_k$ is unlikely if arm $k'$ is sampled enough.
 More formally,}
@@
 %     &\leq
-%     \underbrace{\sum_{t=1}^{T} \Pr\left( g_k^j(t) \leq \mu_k \right)}_{(i)}
-%     + \underbrace{\sum_{t=1}^{T} \Pr\left( g_{k'}^j(t) > \mu_k, A^j(t)=k' \right)}_{(ii)}.
+%     \underbrace{\sum_{t=1}^{T} \Pr\left( U_k^j(t) \leq \mu_k \right)}_{(i)}
+%     + \underbrace{\sum_{t=1}^{T} \Pr\left( U_{k'}^j(t) > \mu_k, A^j(t)=k' \right)}_{(ii)}.
 % \end{align}
@@
 % Exactly the same arguments work for the other case $\mu_k < \mu_{k'}$, by considering
-% whether $g_k^j(t-1) > \mu_{k'}$ (which is also ulikely),
-% or whether $g_k^j(t-1) \leq \mu_k$ which implies $g_{k'}^j(t-1) \leq \mu_{k'}$ (and which is also highly unlikely), and with Lemmas~\ref{lem:5:Fact1KLUCB}, \ref{lem:5:Fact2KLUCB},
+% whether $U_k^j(t-1) > \mu_{k'}$ (which is also ulikely),
+% or whether $U_k^j(t-1) \leq \mu_k$ which implies $U_{k'}^j(t-1) \leq \mu_{k'}$ (and which is also highly unlikely), and with Lemmas~\ref{lem:5:Fact1KLUCB}, \ref{lem:5:Fact2KLUCB},
 % we obtain that
@@
 % Either she sampled $A^j(t+1)$ uniformly
-% from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
+% from $\TopM(t) \cap \{ m : U_m^j(t) \leq U_k^j(t) \}$
 % with transition $(3)$,
@@
 
-% \paragraph{Strong uniform efficiency}\label{app:5:JustifyingDefinition5}
+% \paragraph{Strong uniform efficiency.}\label{app:5:JustifyingDefinition5}
 % %
@@
 Each rectangle represents a configuration, as the matrix $[[\widetilde{S_k}^j(t) / T_k^j(t) ]_j]_k$.
-Absorbing configurations from depth $2$ are case of equality of the two vectors and the \Selfish{} indices $\widetilde{g_k}^j(t)$.
+Absorbing configurations from depth $2$ are case of equality of the two vectors and the \Selfish{} indices $\widetilde{U_k}^j(t)$.
 Transitions are labeled with their probabilities.}
@@
 $\log T$, is asymptotically lower bounded by a constant $\mathrm{LB}(\boldsymbol{\mu}, M)$
-depending on the problem $\boldsymbol{\mu}$ and the number of players $M$, for any $\rho$.
+depending on the problem $\boldsymbol{\mu}$ and the number of players $M$, for any $\cA$.
 \begin{equation}
-  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \rho)}{\log T} \geq \mathrm{LB}(\boldsymbol{\mu}, M).
+  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log T} \geq \mathrm{LB}(\boldsymbol{\mu}, M).
 \end{equation}
@@
 The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} below
-shows the regret $R_T(\boldsymbol{\mu}, M, \rho)$
+shows the regret $R_T(\boldsymbol{\mu}, M, \cA)$
 on the same example problem $\boldsymbol{\mu}$, with $K = 9$ arms and respectively $M = 6$, or $9$ players, for \Selfish-\klUCB.

### 2019-06-07 2-Chapters/1-Chapter/chapter1.tex
Proofread with proselint, writegood, checkwriting

@@
     \item
-    We wrote the most complete open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+    We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
     We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.

### 2019-06-07 2-Chapters/2-Chapter/chapter2.tex
Proofread with proselint, writegood, checkwriting

@@
 The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as $\UCB_1$ or Thompson sampling, with uniform perturbations.
-Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$).
-But instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
+Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$),
+but instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
 % https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html
@@
 %
-For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms.
-But when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
+For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms,
+but when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
 ideally one would like to use an algorithm that can be applied identically against any problem of such family.

### 2019-06-07 2-Chapters/3-Chapter/SMPyBandits_paper.git/longpaper.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Proofread with proselint, writegood, checkwriting

@@
 \hypertarget{arms-and-problems}{%
-\paragraph{Arms and problems}\label{arms-and-problems}}
+\paragraph{Arms and problems.}\label{arms-and-problems}}
 
@@
 {[}@matplotlib{]} for 2D plotting, numpy {[}@numpy{]} for data storing,
-random number generations and and operations on arrays, scipy
+random number generations and operations on arrays, scipy
 {[}@scipy{]} for statistical and special functions, and seaborn

### 2019-06-07 2-Chapters/3-Chapter/chapter3.tex
Lots of typos changed, I'm reading all the thesis very quickly
CHAUD
FAIT CHIER
Proofread with proselint, writegood, checkwriting

@@
 
-\textbf{Abstract.}
+\graphicspath{{2-Chapters/3-Chapter/Images/}}
+\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
+
+\paragraph{Abstract.}
 %
@@
 %  and SMPyBandits is written in Python.
-It is the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
+It is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
 %
@@
 
+\TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
+
 \minitoc
@@
 \newpage
-
-% \newpage
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/3-Chapter/Images/}}
-\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
 
@@
 %  and SMPyBandits is written in Python.
-This library is by far the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
+This library is by far the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
 It aims at being exhaustive, simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
@@
 
-\paragraph{Publications.}
+\textbf{Publications.}
 %
@@
     It is based on the online mirror descent algorithm with a Tsallis entropy regularizer, and it is anytime.
-    It was also proved to to obtain good ``best of both world'' performances.
+    It was also proved to obtain good ``best of both world'' performances.
     Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
@@
         $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1846 \;\text{s}$ \\
-        \hline
+        % \hline
     \end{tabular}

### 2019-06-07 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Proofread with proselint, writegood, checkwriting

@@
 Additionally to bounding the collisions in Lemma~\ref{lem:5:collisionsMCTopM},
-an interesting consequence of the proof of this result is that
+another consequence of our proof of this result is that
 it also bounds the (expected) number of \emph{switches of arms}, $SC^{\cA}(T) = \sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$.
@@
 Even though minimizing the number of arms switching was not a goal,
-this guarantee is interesting, in particular for Cognitive Radio applications,
+this guarantee is appealing, in particular for Cognitive Radio applications,
 where switching arms means re-configuring a radio hardware, an operation that costs energy.
@@
 %
-We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17}, is an interesting future work.
+We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17} for single-player bandits, could also be fruitful.
 

### 2019-06-07 2-Chapters/5-Chapter/chapter5.tex
I have some small bugs, but all is good
Proofread with proselint, writegood, checkwriting

@@
 We conclude by reviewing different extensions of the multi-player models.
+
+\TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
 
@@
 %
-Without giving more details, we let the interested refer to what is is explained for the algorithms presented in \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
+Without giving more details, we let the interested refer to what is explained for the algorithms presented in \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
 %
@@
 In \cite{Rosenski16}, the authors explain that if the model allows arrival or departures of players at \emph{any time step}, then the game is much harder, and sub-linear regret is most likely un-achievable, if we consider the natural extension of the definition of regret as in our model presented in Section~\ref{sec:5:model} above.
-However, this negative results depends on how arrival of players are modeled:
+But this negative results depends on how arrival of players are modeled:
 if an arriving player has no prior memory of the previous observations, \ie, if it model a new device, then if there is no restriction on the number or frequency of arrival or departures, we can most likely prove that sub-linear regret is not achievable in general.
@@
 Every time step, another player is either arriving, without any knowledge of the problem (\eg, it is a fresh and new IoT device).
-Then no matter the strategy of these new players, even if they all play \MCTopM-\klUCB{} for instance, if they play a uniformly-efficient strategy there is a non-zero probability that player $1$ suffers from a collision at each time step, thus resulting in linear centralized system regret.
+Then no matter the strategy of these new players, even if they all play \MCTopM-\klUCB{} for instance, if they play a uniformly efficient strategy there is a non-zero probability that player $1$ suffers from a collision at each time step, thus resulting in linear centralized system regret.
 
@@
 A naive idea, if $L$ is large enough, is simply to restart the underlying algorithm every $L$ time steps, in order to directly benefit from the theoretical guarantees of the ``static setting'' (where $M$ stays constant).
-However, this idea yields a large regret if the length of ``static'' intervals $L$ is too small.
+Unfortunately this idea yields a large regret if the length of ``static'' intervals $L$ is too small.
 Applied to the \MusicalChair{} algorithm, the authors in \cite{Rosenski16} call this basic extension Dynamic \MusicalChair
@@
 Their argument is essentially to prove that for \Selfish-powered players using the simple \UCB-indexes, with a probability $p$ at time $t$ (both independent from $T$), two players might have the same number of pulls and the same observed rewards for each arm.
-In that case, the two players would pull the exact same arms and thus collide for a long time, until they reach a ``tie breaking point'' where they could choose different arms thanks to a random tie breaking rule (\eg, if two values of their \UCB{} indexes are the same, the $\argmax$ is a uniform random choice among the two arms).
+In that case, the two players would pull the same arms and thus collide for a long time, until they reach a ``tie breaking point'' where they could choose different arms thanks to a random tie breaking rule (\eg, if two values of their \UCB{} indexes are the same, the $\argmax$ is a uniform random choice among the two arms).
 They prove that it is unlikely to encounter any of these ``tie breaking points'', in theory if the \UCB{} indexes are real-valued number (that can be rational or irrational).
@@
 for Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis.
-We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, however we also experimented with larger values of $T$ and found a same linear behavior.}.
+We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, and we also experimented with larger values of $T$ and found a similar linear behavior.}.
 
@@
 Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
-As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as showed in Theorem~4.1
-(their notation uses $K$ for $M$ the number of players and $N$ for $K$ the number of arms).
+As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as shown in Theorem~4.1
+(they used $K$ for $M$ the number of players, and $N$ for $K$ the number of arms).
 This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for Exp3 for the single-player adversarial setting in \cite{Auer02NonStochastic}.
@@
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
-However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
-Similarly to what was done very recently in \cite{wang2019distributed} for communication between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
+However, sub-optimal draws turn out to not be the dominant term in the regret, both in our upper bound and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
+Similarly to what was done very recently in \cite{wang2019distributed}, for communications between players who collaborate with each other, it would be interesting to characterize the (minimal) number of collisions needed to achieve logarithmic regret.
 

### 2019-06-07 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Proofread with proselint, writegood, checkwriting

@@
 Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
-As explained in in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows before hand the values of $a$ and $b$).
+As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows before hand the values of $a$ and $b$).
 

### 2019-06-07 2-Chapters/Conclusion/conclusion.tex
Fixed the Nomenclature
I have some small bugs, but all is good
Proofread with proselint, writegood, checkwriting

@@
 \label{chapter:conclusion}
+
+% Write miniTOC just after the title
+\graphicspath{{2-Chapters/6-Chapter/Images/}}
+% ----------------------------------------------------------------------
 
@@
 
-\vspace*{10pt}
+% \vspace*{50pt}
 
 \minitoc
-
-% Write miniTOC just after the title
-\graphicspath{{2-Chapters/6-Chapter/Images/}}
-% ----------------------------------------------------------------------
 
@@
 Because the models studied for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in a noisy environment,
-in Chapter~\ref{chapter:5}, we study a weaker but still very interesting model.
+in Chapter~\ref{chapter:5}, we study a weaker but still interesting model.
 We present different variants of the multi-player MAB model, and we propose the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
@@
 For the two easiest models --with sensing--, our theoretical contribution improves the state-of-the-art upper bounds on the regret.
-In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
+In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
 %
@@
 
-As discussed at the end of each chapters, our works suggest many interesting directions of future studies.
+As discussed at the end of each chapters,
+our works suggest possible directions of future studies.
 
@@
 
-An interesting future work left on our library SMPyBandits could be to implement new variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
+A first future work left on our library SMPyBandits could be to implement new variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
 For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
@@
 It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
-It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
+It could be exciting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
 
@@
 
-Our study on multi-player bandits suggests several interesting further research directions.
+Our study on multi-player bandits suggests several further research directions.
 First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
-However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
-Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
+However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}.
+% Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
 
@@
 We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected.
-This interesting research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$.
+This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$.
 We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.

### 2019-06-07 1-Preamble/5-Abstract/abstract.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment
Typo

@@
 \begin{abstract}
-% \vspace*{1cm}
-
+\vspace*{-30pt}
+%
 % \TODOL{Raccourcir juste un peu ?}
-
 % ----------------------------------------------------------------------------
-
+\begin{small}
 % Chapitre 1
 Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
-Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
+Nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
 %  (Chapitre~\ref{chapter:1}).
@@
 % (Chapitre~\ref{chapter:3})
+\end{small}
+
+% ----------------------------------------------------------------------------
+\hr{}
 
@@
 % \vspace*{1cm}
-
+% \vspace*{-5pt}
+%
+\begin{small}
 % Chapter 1
 In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access large-scale Cognitive Radio networks, in unlicensed bands and without central control.
-More specifically, we focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
+We focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
 %  (Chapter~\ref{chapter:1})
@@
 %  (Chapter~\ref{chapter:3})
+\end{small}
 

### 2019-06-07 1-Preamble/5-Abstract/resume_fr.tex
Typo

--- 2019-01-17 1-Preamble/5-Abstract/resume_fr.tex
@@
 
-Cette thèse aborde tous les points susmentionnés et propose différentes solutions pour chaque scénario.
-Les contributions importantes sont résumées ci-dessous :
+Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
+Nous signalons les conférences internationales auxquelles ont été publiées certains résultats.
+
 \begin{itemize}
-\item Item 1
-\item Item 2
-\item Item 3
-\item Item 4
+    \item
+    Nous présentons les concepts et les notations du modèle des bandits multi-bras, dans le chapitre~\ref{chapter:2} selon un point de vue mathématique formel.
+    Nous suivons aussi une approche didactique, puisque nous utilisons en Section~\ref{par:2:interactiveDemoDiscoverMAB} une démonstration interactive en ligne conçue pour permettre à n'importe qui de jouer contre un petit problème de bandit depuis son navigateur.
+
+    \item
+    Nous donnons une brève revue de la littérature de recherche concernant les algorithmes stochastiques de bandits en Section~\ref{sec:2:famousMABalgorithms}.
+
+    \item
+    Nous présentons le problème du choix de l'algorithme qu'un praticien devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithms MAB disponibles, dans la Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par quelques simulations numériques qu'il atteint de bonnes performances empiriques
+    (WCNC 2018) \cite{Besson2018WCNC}.
+
+    \item
+    Nous avons écrit la plus complète bibliothèque de simulation pour les problèmes MAB, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBandits,SMPyBanditsJMLR}.
+    Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
+    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences utilisées dans la suite de cette thèse.
+
+    \item
+    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les appareils dotés de capacités de radio intelligente peuvent implémenter de leur côté des algorithmes MAB, pour augmenter automatiquement la durée de vie de leur batterie.
+Cela permet à davantage de périphériques d'utiliser le même réseau tout en maintenant une haute qualité de service
+    (CROWNCOM 2017, ICT demo 2018, WCNC 2019 et MOTIoN 2019)
+    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC}.
+
+    \item
+    Nous avons implémenté une preuve de concept du modèle susmentionné \cite{Besson2018ICT}, et nous la présentons en détail en Section~\ref{sec:4:gnuradio}.
+Nous avons aussi réalisé une courte vidéo présentant notre démonstration, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+
+    \item
+    Le code source des deux contributions susmentionnées utilise GNU Octave ou MATLAB et est publié en ligne, ainsi que des instructions claires pour la reproduction de nos travaux.
+
+    \item
+    Nous formalisons le modèle du bandit multi-joueur, pour lequel nous avons introduit trois variantes, dans le Chapitre~\ref{chapter:5}.
+    Pour le cas avec information de détection (``sensing''), nous proposons deux nouveaux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM, qui prouve qu'il est asymptotiquement optimal.
+    Nous présentons aussi des expériences numériques approfondies pour montrer qu'il est bien plus efficace que les autres algorithmes de la littérature précédente.
+    Notre travail \cite{Besson2018ALT} (ALT 2018) a également contribué à une nouvelle impulsion à la recherche sur les bandits multi-joueurs, car certains travaux de recherche récents se sont construits sur nos résultats.
+
+    \item
+    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle MAB multi-joueurs, qui, selon nous, n'a jamais été écrit auparavant.
+    % (ALT 2018, et 4-8 travaux inspirés de notre article depuis lors).
+Etat de l'art avec notre algorithme MCTopM + klUCB pour les bandits multi-joueurs "avec détection", état de l'art empirique avec notre approche simple (mais erronée) "égoïste" en cas de "non détection".
+
+    \item
+    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (GRETSI 2019).
+    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif pour ce problème stationnaire par morceaux, \GLRklUCB, qui atteint des performances comparables à l'état de l'art.
+    % Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB
+
+    \item
+    La dernière contribution de cette thèse est une revue de la littérature sur les cas d'utilisation possibles de la technique du doublement successif de l'horizon (``doubling trick'') pour les problèmes de bandits,
+    ainsi qu'une analyse unifiée et plus générique de deux familles de doublements.
+    Ceci a conduit à l'article \cite{Besson2018DoublingTricks}, qui est rapidement présenté en Annexe~\ref{app:2:DoublingTricks}.
 \end{itemize}
@@
 Ce manuscrit est organisé comme suit.
-Nous rappelons dans le \textbf{Chapitre \ref{chapter:1}} blabla.
 
-Dans le \textbf{Chapitre \ref{chapter:2}}, blabla.
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.95\textwidth}{!}{
+    \begin{tikzpicture}[>=latex',line join=bevel,scale=2.25]
+        %
+        \node[align=center] (introduction) at (0,3.25) [rectangle,draw] {Chapitre~\ref{chapter:1}\\Introduction};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw] {Chapitre~\ref{chapter:2}\\Modèle bandits multi-bras\\stochastique et stationnaire};
+        \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapitre~\ref{chapter:3}\\SMPyBandits : bibliothèque\\de simulations pour MAB};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:4}\\Deux modèles MAB\\pour les réseaux IoT};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapitre~\ref{chapter:5}\\Modèle MAB\\Multi-joueurs};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:6}\\Modèle MAB\\Non-stationnaire};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapitre~\ref{chapter:conclusion}\\Conclusion générale};
+        %
+        \draw [color=black,thick,->] (introduction) to (chapter2);
+        \draw [color=black,thick,<->] (chapter2) to (chapter3);
+        \draw [color=black,thick,->] (chapter2) to (chapter4);
+        \draw [color=black,thick,->] (chapter2) to (chapter5);
+        \draw [color=black,densely dotted,<->]   (chapter4) to (chapter5);
+        \draw [color=black,densely dotted,->] -| (chapter3) to (chapter6);
+        \draw [color=black,densely dotted,<->]   (chapter5) to (chapter6);
+        \draw [color=black,thick,->] (chapter2) to (chapter6);
+        \draw [color=black,thick,->] (chapter4) to (conclusion);
+        \draw [color=black,thick,->] (chapter5) to (conclusion);
+        \draw [color=black,thick,->] (chapter6) to (conclusion);
+        %
+    \end{tikzpicture}
+    }
+    \caption[Organisation de la thèse : une carte de lecture]{Organisation de la thèse.
+Cette thèse peut se lire en suivant n'importe quel chemin contenant le Chapitre~\ref{chapter:1}, le Chapitre~\ref{chapter:2}, au moins un des trois Chapitres~\ref{chapter:4}, \ref{chapter:5} et \ref{chapter:6}, et la Conclusion.}
+    \label{fig:1:organization_fr}
+\end{figure}
 
-Dans le \textbf{Chapitre \ref{chapter:3}}, blabla.
+L'ordre de lecture du manuscrit peut être n'importe quel chemin, descendant entre l'introduction dans le Chapitre~\ref{chapter:1} actuel, et la conclusion dans le dernier Chapitre~\ref{chapter:conclusion}.
+Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,
+la thèse est organisée en deux parties :
 
-Dans le \textbf{Chapitre \ref{chapter:4}}, blabla.
+\begin{itemize}
+    \item
+Dans la Partie~\ref{part:Introduction}, nous commençons par le Chapitre~\ref{chapter:2} où nous présentons les modèles MAB, les concepts et les notations utilisés dans tout ce document.
+Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
+Par contre le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
 
-Dans le \textbf{Chapitre \ref{chapter:5}}, blabla.
-
-Dans le \textbf{Chapitre \ref{chapter:5}} blabla.
+    \item
+La deuxième Partie~\ref{part:MABIOT} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus presque indépendamment.
+Le Chapitre~\ref{chapter:4} commence par présenter différents modèles de réseaux IdO où les algorithmes MAB ont été utilisés avec succès.
+Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop généraux pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
+Pour cette raison, nous affaiblissons les modèles pour le reste du document,
+et les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
+\end{itemize}
 

### 2019-06-07 2-Chapters/1-Chapter/chapter1.tex
Typo

@@
     \item
-    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
+    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi} (GRETSI 2019).
     Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
@@
     \item
-    The last contribution of this thesis is a literature review of the possible use cases of the doubling trick technique for MAB problems,
+    The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
     and a unified and more generic analysis of two families of doubling tricks.
-    This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}, but not in the main text of the thesis.
+    This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
 \end{itemize}
@@
 
+\begin{itemize}
+    \item
 In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2} where we introduce the MAB models, the concepts and the notations used in all this document, and it is mandatory to read it.
@@
 
+    \item
 Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
@@
 and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.
+\end{itemize}
 
@@
 
-\paragraph{Seminars.}
+\textbf{Seminars.}
     I gave some presentations in the following events:
@@
 
-\paragraph{Tutorial.}
+\textbf{Tutorial.}
     I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
@@
 
-\paragraph{Training.}
+\textbf{Training.}
     I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
@@
 
-\paragraph{Conferences.}
+\textbf{Conferences.}
     I attended the following conferences:
@@
 
-\paragraph{Seminars.}
+\textbf{Seminars.}
     I attended the following seminars:
@@
 	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
-	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$,
+	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$.
 
-\paragraph{Responsabilities.}
+\textbf{Responsabilities.}
     I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
@@
 
-\paragraph{Reviews.}
+\textbf{Reviews.}
 	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
@@
 
-\paragraph{System Administrator.}
+\textbf{System Administrator.}
 	maintaining our workstations running GNU/Linux and Windows
@@
 
-\paragraph{Copyright notice.}
+\textbf{Copyright notice.}
 %
@@
 under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}},
-online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}},
+online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}}.
 

### 2019-06-07 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Typo

@@
 				Give \emph{original} reward $(A(t), r(t))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
-				Compute an \emph{unbiased} estimate of the reward of the trusted algorithms,$$ \widehat{r}(t) = r(t) / p^{t}_{A(t)}\;. $$
+				Compute an \emph{unbiased} estimate of the reward, $\widehat{r}(t) = r(t) / p^{t}_{A(t)}$\;
 				% \tcp{Use bandit feedback to update $\pi^{t}$}
@@
 					\eIf{$\Alg_a$ was trusted, \ie, $A_a(t) = A(t)$}{
-						$ \pi^{t+1}_{a} = \exp(\eta_t r(t) \times \pi^{t}_{a} $
+						$ \pi^{t+1}_{a} = \exp(\eta_t r(t)) \times \pi^{t}_{a} $
 						\tcp*{Trusted more}
@@
 \end{equation}
-satisfies $\E_{O_t}[\widehat{r^a}(t) | O_t] = Y_{A_a(t),t}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $O_t$.
+then it satisfies $\E_{O_t}[\widehat{r^a}(t) | O_t] = Y_{A_a(t),t}$, for all $a$, where the expectation is taken conditionally to the history of observations up to round $t$, $O_t$.
 %
@@
 and an alternative would be to
-increase $\pi_{a}$ if $A_a(t) = A(t)$ and to decrease it otherwise.
+increase $\pi_{a}$ if $A_a(t) = A(t)$ and to \emph{decrease it otherwise}.
 It would not be so different, as there is a final renormalization step, and empirically this variation has little impact on the performance of \Aggr{}.

### 2019-06-07 2-Chapters/2-Chapter/chapter2.tex
Typo

@@
 % Since then, many efficient algorithms were proposed, XXX
-
+%
 % In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
@@
 % A sketch of the proof of both results is given in Appendix.
-
+%
 Finally, we explore in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} the problem of algorithm selection: as there are many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
@@
 \textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
-When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) = Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$, from $t=1$ until $t=T$ for an \emph{horizon} $T$.
+When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) = Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
 %
@@
 The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T r(t)] = \sum_{t=1}^T \E[Y_{A(t),t}]$.
-The expectancy is taken on the rewards $(Y_{k,t})_{k,t}$, and the extern random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
+The expectancy $\E[\bullet]$ is taken on the rewards $(Y_{k,t})_{k,t}$, as well as on the external random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
 %
@@
 The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
-The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $100$ steps, $\sum_{t=1}^{100} r(t)$.
-Your decisions have to be made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r(t)\in\{0,1\}$, that is sampled \iid{} from $\nu_k$.
+The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $T=100$ steps, \ie, to maximize $\sum_{t=1}^{100} r(t)$.
+Your decisions are made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r(t)\in\{0,1\}$, that is sampled \iid{} from $\nu_k$.
 %
@@
 %
-In the second Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} below, we display the result of an example of run, when the player was following the $\UCB_1$ algorithm from \cite{Auer02} (see Section~\ref{sub:2:IndexPolicies}).
+In the second Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem__step100} below, we display the result of an example of run, when the player was following the $\UCB_1$ algorithm from \cite{Auer02} (we present it below in Section~\ref{sub:2:IndexPolicies}).
 After $T=100$ steps, the player obtained a cumulated reward of $56$, by playing mostly arms $4,3,5,1,0$ (in decreasing order of number of plays).
@@
 
-A \emph{bandit algorithm} $\cA$ is also referred to as a strategy or a policy, and sometimes it is denoted by $\pi$ or $\rho$ in the literature.
+A \emph{bandit algorithm} $\cA$ is also referred to as a strategy or a policy.
+% and sometimes it is denoted by $\pi$ or $\rho$ in the literature.
 The algorithm $\cA$ selects an arm $A(t)$ at time $t$, possibly by using the past observations and the past external randomness.
@@
 
-\begin{definition}[Kullback-Leibler divergence]\label{def:2:KLDivergence}
+\begin{definition}\label{def:2:KLDivergence}
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-    The KL divergence between two distributions $d_1$ and $d_2$ is defined as
+    The \emph{Kullback-Leibler divergence} between distributions $d_1$ and $d_2$ is defined as
     \begin{align}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
@@
     \end{align}
-% \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+% \vspace*{-5cm}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
 we can write $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
-and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$:
-%
-\begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
-    \KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2) X ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2) \mu_1 - A(\eta_1) + A(\eta_2),
-\end{equation}
-
+and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$,
+%
+% \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
+$\KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2) X ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2) \mu_1 - A(\eta_1) + A(\eta_2)$.
+% \end{equation}
+%
 Without diving more in the details of exponential families,
@@
 
-\begin{itemize}
-    \item
+% \begin{itemize}
+%     \item
 On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if we want to compute and visualize the regret, we can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
@@
 
-    \item
+    % \item
 On the other hand, this decomposition is necessary as theoretical analyses of the regret of (single-player) MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
 Even if we prove that controlling $N_k(T)$ is no longer sufficient to obtain low regret for the multi-player case, we extend this decomposition in Chapter~\ref{chapter:3} (in Lemma~\ref{lem:5:DecompositionRegret}), and it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
-\end{itemize}
+% \end{itemize}
 
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    No algorithm $\cA$ can achieve a (mean) regret $R_T^{\cA}(I) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
+    \emph{No algorithm $\cA$ can achieve} a (mean) regret $R_T^{\cA}(I) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
     where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
@@
 
-This family is non-empty as it contains for instance the $\UCB_1$ algorithm, as its regret is proven to be logarithmic on any problem instance with bounded rewards (in Theorem~\ref{thm:2:UCBregretBound} below).
+This family is non-empty as it contains for instance the $\UCB_1$ algorithm, since its regret is proven to be logarithmic on any problem instance with bounded rewards (in Theorem~\ref{thm:2:UCBregretBound} below).
 Now we can state the second logarithmic lower-bound, for algorithms in this family:
@@
 \begin{theorem}\label{thm:2:forSecondLogTLowerBound}
-\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
+\begin{leftbarnospace}[theorembar]  % XXX leftbar theorembar, comment if needed
     Let $\cA$ denote a uniformly efficient algorithm,
@@
     \end{itemize}
-\end{leftbar}  % XXX leftbar theorembar, comment if needed
+\end{leftbarnospace}  % XXX leftbar theorembar, comment if needed
 \end{theorem}
@@
 The weights are initially uniform, \ie, $w_k(0)=\frac{1}{K}$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
-The weights are then renormalized, the have a sum $\sum_k w_k(t+1) = 1$.
+The weights are then renormalized, to have a sum $\sum_k w_k(t+1) = 1$.
 This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = \Pr(A(t)=k)$.
 %
-Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a decreasing sequence $(\gamma_t)_t$.
-%
-It was proved in \cite{Auer02NonStochastic} that Exp3 with constant temperature parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
+Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a non-increasing sequence $(\gamma_t)_t$.
+%
+It was proved in \cite{Auer02NonStochastic} that Exp3 with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
 while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
@@
 %
-Many other policies have been proposed for adversarial bandits, like Follow-the-Perturbed-Leader or Follow-the-Regularized-Leader for linear adversarial bandits, but we do not study them in this thesis, and we refer to Parts~III and VI of \cite{LattimoreBanditAlgorithmsBook} for more details.
+Many other policies have been proposed for adversarial bandits, like Follow-the-Perturbed-Leader or Follow-the-Regularized-Leader for linear adversarial bandits,
+% but we do not study them in this thesis,
+and we refer to Parts~III and VI of \cite{LattimoreBanditAlgorithmsBook} for more details.
 
@@
 %
-For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of a $\UCB_1$ algorithms will perform quite differently in the two problems:
+For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of $\UCB_1$ will perform quite differently in the two problems:
 using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case,

### 2019-06-08 2-Chapters/3-Chapter/chapter3.tex
Small stuff

@@
 %
+The code for SMPyBandits is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is freely publicly released under the MIT open-source License.
 This chapter is based on our article \cite{SMPyBanditsJMLR}.
-The code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is freely publicly released under the MIT open-source License.
 
@@
 Any simulation library targeting single-player bandit problems must implement at least three components:
-reward distributions, MAB algorithm, and a simulation loop that essentially looks like this:
-\begin{itemize}
-    \item First, initialize the MAB problem and one or more algorithms,
-    \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
+reward distributions, MAB algorithm, and a simulation loop that essentially looks like this.
+% \begin{itemize}
+    % \item
+    First, initialize the MAB problem and one or more algorithms,
+    % \item
+    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
 Ask algorithm $\cA$ her chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm,
-    \item At the end, compute the cumulated reward, the regret, plot visualizations etc.
-\end{itemize}
+    % \item
+    At the end, compute the cumulated reward, the regret, plot visualizations etc.
+% \end{itemize}
 %
@@
 % a collision model, and
-a simulation loop that essentially looks like this:
-\begin{itemize}
-    \item First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
-    \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
+a simulation loop that essentially looks like this.
+% \begin{itemize}
+    % \item
+    First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
+    % \item
+    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
 Ask each player $\cA^{(j)}$ her chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
@@
 Finally feeds the observation $(A^{(j)}(t), Y_{A^{(j)}(t)}(t), r^{(j)}(t))$ to player $j$, for all the $M$ players,
-    \item At the end, compute the accumulated reward, the regret, plot visualizations etc.
-\end{itemize}
+    % \item
+    At the end, compute the accumulated reward, the regret, plot visualizations etc.
+% \end{itemize}
 
@@
 It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams shown on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
-For instance, most existing algorithms are index-based, and new ones can be written easily by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
-An index-based algorithm computes an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$ and simply play $A(t) = \arg\max_k I_k(t)$.
-For instance the code specific to the UCB algorithm \cite{LaiRobbins85,Auer02} is as short as this (and fully documented):
+For instance, most existing algorithms are index policies (see Algorithm~\ref{algo:2:indexPolicy}), and new ones are easily written by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
+They compute an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$, and play $A(t) = \arg\max_k I_k(t)$.
+For instance the code specific to $\UCB_1$ is as short as this (and fully documented):
 
@@
     % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy}
-    \captionof{listing}{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy \label{lst:3:smallIndexPolicy}.}
+    \captionof{listing}{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
     % \label{lst:3:smallIndexPolicy}
@@
 
-% \TODOL{I need to quickly explain that SMPyBandits is using Read the Docs since $2017$ and Travis Ci since $2018$ to: }
-
 Since $2017$ and $2018$, SMPyBandits is using two online continuous integration (CI) service, to automatically build and host the documentation, and to automatically test the library on some numerical simulations.
@@
 \paragraph{Travis CI.}
-Since $2018$, we have been using the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
+Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
 Travis CI allows to automatically run short numerical simulations after every commit.
@@
 Additional results in terms of real measurements of time and memory are given and discussed in the next Section~\ref{sec:3:timeAndMemoryCosts}.
-
+%
 Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on two algorithms depending if simplicity or efficiency is favored:
@@
 
-We consider the nine following algorithms, and seven more are described in Appendix~\ref{sub:3:additionalExperiments}.
+We consider the $9$ following algorithms, and $7$ more are described in Appendix~\ref{sub:3:additionalExperiments}.
 For each of them, we give a bibliographic reference, that corresponds to a recent article studying it and not the first one which introduced it.
 We give the chosen tuning of the parameters, for parametric algorithms.
-Algorithms that are ``not anytime'' use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, they use the correct successive values. \\
-%
-\indent
+Algorithms that are ``not anytime'' use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, they use the correct successive values.
+%
 For reproducibility, we also give in ``\texttt{typewriter font}'' the name of the corresponding class in the \texttt{Policies} module in SMPyBandits
@@
 
-\begin{itemize}
-    \item Algorithm 1 is
+\begin{enumerate}
+    \item
+    %  Algorithm 1 is
     $\varepsilon$-greedy \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
@@
 
-    \item Algorithm 2 is
+    \item
+    %  Algorithm 2 is
     Explore-then-Commit \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
@@
 
-    \item Algorithm 3 is
+    \item
+    %  Algorithm 3 is
     $\mathrm{Exp}3^{++}$ from \cite{Seldin17}, using $\alpha=3$ and $\beta=256$ as advised (\texttt{Exp3PlusPlus}).
@@
 
-    \item Algorithm 4 is
+    \item
+    %  Algorithm 4 is
     $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
@@
 
-    \item Algorithm 5 is
+    \item
+    %  Algorithm 5 is
     \klUCB{} from \cite{KLUCBJournal}, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern} (\texttt{klUCB}).
-    It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
+    It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret.
+    % with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
     It is computationally more costly that \UCB, as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
-    In practice, a few steps of a simple Newton method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than an order of magnitude slower than \UCB.
+    In practice, a few steps of a simple Newton method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than one order of magnitude slower than \UCB.
     % of more than 10 digits
@@
 
-    \item Algorithm 6 is
+    \item
+    %  Algorithm 6 is
     Thompson sampling from \cite{Kaufmann12Thompson}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{Thompson}).
@@
 
-    \item Algorithm 7 is
+    \item
+    %  Algorithm 7 is
     Bayes-UCB from \cite{Kaufmann12BUCB}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
     % It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
-    It is comparable to Thompson sampling in terms of memory complexity, but slower in terms of time complexity as computing a $1-1/t$ quantile is more costly than sampling from a distribution.
+    It is comparable to Thompson sampling in terms of memory complexity, but slower in terms of time complexity as computing a quantile is more costly than sampling from a distribution.
     But in particular for Bernoulli distributed arms and for Beta posteriors and priors, Bayes-UCB is only about twice as slow as Thompson sampling.
 
-    \item Algorithm 8 is
+    \item
+    %  Algorithm 8 is
     AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
@@
 
-    \item Algorithm 9 is
+    \item
+    %  Algorithm 9 is
     BESA (Best Empirical Sampled Average) from \cite{Baransi2014},
@@
     % Analyzing it in the generic case was found to be difficult, and is still an open problem.
-    which is implemented using a binary tournament, written in a naive but efficient way
+    which is implemented with a binary tournament, written in a naive but efficient way
     (\ie, not using recursive functions, by using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
-    Despite this technic, the extension to $K>2$ arms is still costly in terms of both time and memory, as illustrated below, and blows-up exponentially  when $K$ increases.
-\end{itemize}
+    But the extension to $K>2$ arms is still costly in terms of both time and memory, as illustrated below, and blows-up exponentially  when $K$ increases.
+\end{enumerate}
 
 We focussed here on the most well known algorithms, and two others that are efficient but less known (AdBandits and BESA),
-but to illustrate that SMPyBandits implements many other solutions to the MAB problem proposed in very recent works, we explain in Appendix~\ref{sub:3:additionalExperiments} seven other algorithms.
+but to illustrate that SMPyBandits implements many other solutions to the MAB problem proposed in very recent works, we explain in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
 They are more recent (some from $2016$, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or are based on a different point-of-view.
@@
 
-This Appendix section describes seven more algorithms that are also compared with the nine presented above in Section~\ref{sec:3:reviewSPAlgorithms}.
-
-% \TODOL{Si on a la place, j'aimerai bien inclure des résultats pour 7 algorithmes de plus, que voici :}
-
-\begin{itemize}
-    \item Algorithm 10 is
+This Appendix section describes $7$ more algorithms that are also compared with the $9$ presented above in Section~\ref{sec:3:reviewSPAlgorithms}.
+
+\begin{enumerate}
+    \setcounter{enumi}{9}
+    \item
+    % Algorithm 10 is
     $\mathrm{MOSS}$-$\mathrm{Anytime}$ from \cite{Degenne16}, using $\alpha=1$ (\texttt{MOSSAnytime}).
@@
 
-    \item Algorithm 11 is
-    the approximated Finite-Horizon Gittins from \cite{Lattimore16a}, using $\alpha=1$ (\texttt{ApproximatedFHGittins}).
+    \item
+    % Algorithm 11 is
+    Approximated Finite-Horizon Gittins from \cite{Lattimore16a}, using $\alpha=1$ (\texttt{ApproximatedFHGittins}).
     It mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
@@
 
-    \item Algorithm 12 is
+    \item
+    % Algorithm 12 is
     $\UCB\dagger$ from \cite{Lattimore2018refining}, using $\alpha=1$ (\texttt{UCBdagger}).
@@
 
-    \item Algorithm 13 is
-    the anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018},
+    \item
+    % Algorithm 13 is
+    Anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018},
     is a very recent variant of the \klUCB{} algorithm \cite{KLUCBJournal}
@@
 
-    \item Algorithm 14 is
+    \item
+    % Algorithm 14 is
     \textsc{Tsallis-Inf} from \cite{Zimmert2018}, using $\alpha=1/2$ (it seems to be the most efficient of the algorithms described in the article) (\texttt{TsallisInf}).
@@
 
-    \item Algorithm 15 is
+    \item
+    % Algorithm 15 is
     $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019}, using $\alpha=1$ and perturbations uniformly sampled in $[0,1]$.
@@
 
-    \item Algorithm 16 is
+    \item
+    % Algorithm 16 is
     $\mathrm{PHE}$ (Perturbed-History Exploration) from \cite{KvetonSzepesvari2019}, using a perturbation scale of $0.5$ as advised (\texttt{PHE}).
@@
     Interestingly, its time and memory consumption stay constant with respect to the step number $t$ and the horizon $T$, because generating and summing $t$ pseudo-rewards from a Bernoulli distribution can actually be done in $\cO(1)$ time and space, by using efficient sampling methods for sampling from Binomial distributions.
-\end{itemize}
+\end{enumerate}
 
@@
     \centering
-    \begin{tabular}{c|ccc}
+    \begin{tabular}{c|rrr}
     \textbf{Repetitions} $\;$ \textbackslash $\;$ number of jobs \texttt{N\_JOBS} & $1$ & $4$ ($=$ nb cores) & $20$ ($>$ nb cores) \\
@@
 
-\TODOL{Pas sûr que ça vaille la peine d'inclure ce passage, c'est vraiment optionnel !}
-
-% Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).
-\paragraph{Approaches we did not try.}
-%
-The two other approaches we could have consider is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
-
-$(i)$ I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
-Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.
-Since $2016$, we saw a large press coverage as well as an extensive use in research of deep learning libraries that make general-purpose machine learning algorithms train on the GPU of a user's laptop or a cluster of GPU.
-This success is mainly possible because of the heavy parallelism of such training algorithms, and the parallel nature of GPU.
-To the best of the author knowledge, nobody has tried to implement high performance MAB simulations by using the ``parallelism power'' of a GPU (at least, no code for such experiments were made public in $2019$).
-I worked on a GPU, implementing fluid dynamic simulations in an internship in $2012$, and I have since then kept a curiosity on how to use GPU-powered libraries and code.
-I have contributed to and used famous deep-learning libraries, like Theano (\href{http://deeplearning.net/software/theano/}{\texttt{DeepLearning.net/software/theano}}) or Keras (\href{https://keras.io/}{\texttt{keras.io}}), and my limited knowledge on such libraries made me believe that it was not easy to use a GPU for bandit simulations, and most surely it would not have been worth the time.
-I would be very curious to understand how a GPU could be used to implement highly efficient simulations for sequential learning problems, because it seemed hard whenever I thought about it.
-
-$(ii)$ I also did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
-It is partly due to time constraint, as I would have been curious to try, but mainly because we found that it would not have helped us much to use a large scale cluster.
-The main reason is that in the multi-armed bandit and sequential learning literature, most research papers do not even include an experimental section, and for the papers who did take the time to implement and test their proposed algorithms, it is almost done on just a few problems and for short- or medium- duration experiments.
-For instance, the papers we consider to be the best ones regarding their empirical sections are \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, for piece-wise stationary bandits, and they mainly consider reasonable problems of horizon $T=10000$ and no more than $1000$ independent repetitions.
+% \TODOL{Pas sûr que ça vaille la peine d'inclure ce passage, c'est vraiment optionnel !}
+
+% % Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).
+% \paragraph{Approaches we did not try.}
+% %
+% The two other approaches we could have consider is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
+
+% $(i)$ I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
+% Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.
+% Since $2016$, we saw a large press coverage as well as an extensive use in research of deep learning libraries that make general-purpose machine learning algorithms train on the GPU of a user's laptop or a cluster of GPU.
+% This success is mainly possible because of the heavy parallelism of such training algorithms, and the parallel nature of GPU.
+% To the best of the author knowledge, nobody has tried to implement high performance MAB simulations by using the ``parallelism power'' of a GPU (at least, no code for such experiments were made public in $2019$).
+% I worked on a GPU, implementing fluid dynamic simulations in an internship in $2012$, and I have since then kept a curiosity on how to use GPU-powered libraries and code.
+% I have contributed to and used famous deep-learning libraries, like Theano (\href{http://deeplearning.net/software/theano/}{\texttt{DeepLearning.net/software/theano}}) or Keras (\href{https://keras.io/}{\texttt{keras.io}}), and my limited knowledge on such libraries made me believe that it was not easy to use a GPU for bandit simulations, and most surely it would not have been worth the time.
+% I would be very curious to understand how a GPU could be used to implement highly efficient simulations for sequential learning problems, because it seemed hard whenever I thought about it.
+
+% $(ii)$ I also did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
+% It is partly due to time constraint, as I would have been curious to try, but mainly because we found that it would not have helped us much to use a large scale cluster.
+% The main reason is that in the multi-armed bandit and sequential learning literature, most research papers do not even include an experimental section, and for the papers who did take the time to implement and test their proposed algorithms, it is almost done on just a few problems and for short- or medium- duration experiments.
+% For instance, the papers we consider to be the best ones regarding their empirical sections are \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, for piece-wise stationary bandits, and they mainly consider reasonable problems of horizon $T=10000$ and no more than $1000$ independent repetitions.
 Each paper considers one harder problem, of horizon $T=1000000$ and less repetitions.
-%
-In each article written during my thesis, we included extensive numerical simulations, and even the longest ones (for \cite{Besson2019GLRT}) were short enough to run in less than $12$ hours on a $12$-core workstation, so we could run a few large-scale simulations over night.
-For such reasons, we prefer to not try to run simulations on a cluster.
+% %
+% In each article written during my thesis, we included extensive numerical simulations, and even the longest ones (for \cite{Besson2019GLRT}) were short enough to run in less than $12$ hours on a $12$-core workstation, so we could run a few large-scale simulations over night.
+% For such reasons, we prefer to not try to run simulations on a cluster.

### 2019-06-08 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment
Small stuff

@@
 %
-This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a device has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.
-
-Before modeling this goal with a \emph{multi-armed bandit problem}, we present some natural baseline (reference) policies.
+This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a device has to try all channels a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst channels too many times.
+%
+Before presenting \emph{multi-armed bandit algorithms}, we present some natural baseline (reference) policies.
 
@@
 In this case, for one dynamic device, a successful transmission happens if it is the only device to choose channel $k$, at that time slot.
-The probability of successful transmission is computed as follows, because the $S_k$ static devices in each channel $k$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$:
+the $S_k$ static devices in each channel $k$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$,
+so probability of successful transmission is computed as follows
 \begin{equation}
@@
 
-We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of $(S_k)_k$ the repartition of static devices, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
+We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the repartition of static devices $(S_k)_k$, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
 This is not possible in IoT contexts as several completely independent networks may operate in a single place.} is possible.
@@
 
-\begin{proof}
+\begin{smallproof}
 \begin{itemize}
@@
 The strong duality condition is satisfied in this case \cite{BoydVanderberghe04}, so finding the saddle points will be enough to find the maximizers.
-    \item
-    More details are given in Section~\ref{sec:4:proofLagrangian} in the Appendix of this Chapter.
+    % \item
     % \hfill{}$\square$
     \end{itemize}
-\end{proof}
+    More details are given in Section~\ref{sec:4:proofLagrangian} in the Appendix of this Chapter.
+\end{smallproof}
 
-Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (it can be computed numerically in an efficient manner) \cite{Corless96}.
+Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (which can be computed numerically in an efficient manner, \cite{Corless96}).
 Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint
@@
 %
-In our model, every dynamic device implements its own \UCB{} algorithm, \emph{independently}.
+In our model, every dynamic device implements its own algorithm, \emph{independently}.
 For one dynamic device, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
 %
-The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
-It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
+The TS algorithm
+% although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
+is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
 
@@
 % (in their respective case of application).
-Further research on this aspect could lead to interesting future works.
+% Further research on this aspect could lead to interesting future works.
 
@@
 
-\paragraph{A safety check.}
-
+\textbf{A safety check.}
+%
 We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall k, S_k = S/K$), to check that learning approach (here, only \UCB)
@@
 
-\paragraph{Note on the simulation code.}
-
+% \paragraph{Note on the simulation code.}
+\textbf{Reproducibility.}
+%
 The simulation code used for the experiments in Section~\ref{sub:41:numericalResults} is for MATLAB or GNU Octave,
-and is open-sourced under the MIT License, at\\
+and it was written in collaboration with Rémi Bonnefoi, in April and May $2017$.
+Instructions to reproduce our experiments are given, and
+the code is open-sourced under the MIT License, at
 \href{https://Bitbucket.org/scee_ietr/rl_slotted_iot_networks}{\texttt{Bitbucket.org/scee\_ietr/rl\_slotted\_iot\_networks}}.
-It was written in collaboration with Rémi Bonnefoi, in April and May $2017$.

### 2019-06-08 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Small stuff

@@
 \paragraph{Related works.}
-% FIXME?
 % \TODOL{Est-ce que je fais ça ici ou à la fin du chapitre 5 ? Ou peut-être en chapitre 1, ça ferait du contenu au chapitre 1...}
@@
 
-% \paragraph{Outline}
+% \paragraph{Outline.}
 
@@
 
-\paragraph{Physical layer and protocol}
+\paragraph{Physical layer and protocol.}
 
@@
 Thanks to this index, we can have several devices communicating with the same gateway.
+%
+In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
+In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the gateway.
 
@@
 
-In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
-In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the gateway.
-
-\paragraph{Equipment}
+\paragraph{Equipment.}
 % Boards and versions
@@
 
-\paragraph{Implementation}
+\paragraph{Details about our implementation.}
 
@@
 As such, one laptop can run in parallel the control program of any number of boards\footnote{~Even if in practice, maximum efficiency is kept as long as there is not more than one GRC design by CPU core.}.
-
+%
 The GNU Radio software provides the framework and tools to build and run software radio or just general signal-processing applications.
@@
 
-\paragraph{User Interface}
+\paragraph{User Interface.}
 
@@
 
-\paragraph{Availability of data and materials}
+\paragraph{Availability of data and materials.}
 %
@@
 \href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/\\malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
-
+%
 It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.
@@
 
-\paragraph{Video}
+\paragraph{Video.}
 %
-As depicted in Figure~\ref{fig:42:screenshotDemoYouTube},
+As depicted in Figure~\ref{fig:42:screenshotDemoYouTube} above,
 we realized a $6$-minute \textbf{video} to sum-up our demonstration and advertise our work online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
@@
 
-% \paragraph{Special acknowledgment}
+% \paragraph{Special acknowledgment.}
 % %

### 2019-06-08 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
Small stuff

@@
 For the curious reader, we wanted to include here an illustration and a description of the three components of the demonstration we presented above in Section~\ref{sec:4:gnuradio}.
-%
+
 The code corresponding to the following components is written in \texttt{C++} and \texttt{Python} for the hand-written blocks, and using the GUI of the GNU Radio Companion software.
 The flow-graphs are saved as XML files, and the complete code of our demo is fully available online, open-sourced under GPLv3 license, at
-\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit-learning}\\\texttt{-for-iot-networks-with-grc/}}.
+\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit}\\\texttt{-learning-for-iot-networks-with-grc/}}.
 

### 2019-06-08 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Small stuff

@@
 then realistic empirical simulations like in \cite{kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
-,
 and finally applications to multi-hoping networks \cite{Mitton16,Toldov16} or other kinds of networks \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
@@
 
-\paragraph{Outline.}
-
+\textbf{Outline.}
+%
 The rest of the section is organized as follows.
@@
 Our goal here is to obtain a mathematical approximation of the collision probability at the second transmission slot $p_{c1}$, as a function of the first collision probability $p_{c}$.
-
+%
 We make two approximations $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as (they are hypotheses on which the rest of this section is built),
@@
 
-Moreover, from \eqref{eq:43:1} we define the probability $p_{cp}(n)$ that involves the collision of $n$ packets sent by each IoT device (for any $1\leq n \leq N-1$), during the first transmission slot, and is defined by the following equation
+Moreover, from \eqref{eq:43:1} we define the probability $p_{cp}(n)$ that involves the collision of $n$ packets sent by each IoT device (for any $1\leq n \leq N-1$), during the first transmission slot, and is defined by $p_{cp}(n) = {N-1 \choose n} \; x^n \left(1-x\right)^{N-1-n}$.
 %
-\begin{equation*}\label{eq:43:2}
-	p_{cp}(n) = {N-1 \choose n} \; x^n \left(1-x\right)^{N-1-n}.
-\end{equation*}
-
 As explained above, if an IoT device experiences a collision at the first transmission, it proceeds for the retransmission of its packet after a random back-off interval.
@@
 Once again under $\mathcal{H}_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first $N_0 \ll N-1$ terms of the sum in \eqref{eq:43:sumpc2} are predominant.
-We derive,
+We derive $p_{ca} \simeq  1- \frac{1}{p_c}\sum_{n=1}^{N_0}{N-1 \choose n} x^n\left(1-x\right)^{N-1-n} \left( 1-\frac{1}{m}\right)^n$.
 %
-\begin{equation}\label{eq:43:sumpca2}
-	p_{ca} \simeq  1- \frac{1}{p_c}\sum_{n=1}^{N_0}{N-1 \choose n} x^n\left(1-x\right)^{N-1-n} \left( 1-\frac{1}{m}\right)^n.
-\end{equation}
-
 Moreover, for these terms, $n$ is small compared to $N-1$, and so $N-1-n$ can be approximated to $N-1$. Thus it gives,
@@
 
-We interpret this fact in two different situations.
+We interpret this fact in two different situations:
 \begin{itemize}
@@
 
+\begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
 	\textbf{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}:}
@@
 	In all the algorithms and notations presented in this Section, the time steps $t$ do not refer to the global time steps (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
-In other words, each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $M$ times (\eg, $M=10$) until it receives an \Ack.
-
-To be clear, once again, we repeat that the time steps $t$ in the following algorithms denote the number of time the device sent an uplink packet, and these uplink (re)transmissions can happen at different (real-world) times, more or less spread out in time, it does not matter to the learning part of this model and work.
+	Each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $M$ times (\eg, $M=10$) until it receives an \Ack.
+	%
+	In other words, the time steps $t$ in the following algorithms denote the number of time the device sent an uplink packet, and these uplink (re)transmissions can happen at different (real-world) times, more or less spread out in time, it does not matter to the learning part of this model and work.
+\end{leftbar}  % XXX leftbar warningbar, comment if needed
 

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Small stuff

@@
 
+\begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
   % FIXME
-% \begin{framed}
   \textbf{\textcolor{red}{Warning}.}
+  %
   The regret lower bound given in \cite{Besson2018ALT} is not applicable to any algorithm,
@@
   Proving that the proposed algorithms are in this class is left as an open question.
-% \end{framed}
-% FIXME
+\end{leftbar}  % XXX leftbar warningbar, comment if needed
 
@@
 We introduce additional notations in the following definition.
-
-\begin{definition}
+%
+\begin{definition}\label{def:5:nbSelections_nbCollisions}
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
   % [$T_k(T)$ and $\cC_k(T)$]
-  \label{def:5:nbSelections_nbCollisions}
   % \label{def:5:nbSelections}

### 2019-06-08 2-Chapters/5-Chapter/chapter5.tex
Small stuff

@@
 % ----------------------------------------------------------------------------
-\subsection{Inspirations from real-world demonstrations to validate some models}
+\subsection{Inspirations from Cognitive Radio Proof-of-Concepts}
 \label{sub:5:USRPdemos}
@@
 The first \rhoLearn uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
-and we include both the centralized multi-play version of \klUCB{} and \MCTopM-\klUCB{} for comparison.
+and we include both the centralized multiple-play version of \klUCB{} and \MCTopM-\klUCB{} for comparison.
 

### 2019-06-08 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Small stuff

@@
 It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
-The idea behind Memory Bandit is to use an expert aggregation algorithm, like EXP4 from \cite{Auer02}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
+The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} \cite{Besson2018WCNC}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
 At each time step, a new expert is introduced, and experts corresponds to different Thompson sampling algorithms, each using a different history of pulls and rewards.
@@
 Understanding this difference in terms of numerical performance of the two variants is left as future work.
-
+%
 Finally, on the practical side, while we can note that the proposed B-GLRT test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing much in terms of regret, in Appendix~\ref{sub:6:IdeasOptimizations}.
@@
 %-----------------------------------------------------------------------------
-\paragraph{Two options for restarts.}
+\textbf{Two options for restarts.}
 %
@@
 %-----------------------------------------------------------------------------
-\paragraph{Threshold function $\beta$.}
-
+\textbf{Threshold function $\beta$.}
+%
 We present in Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} numerical simulations that compare different choices of threshold functions $\beta$.
@@
 %-----------------------------------------------------------------------------
-\paragraph{Forced exploration.}
-
+\textbf{Forced exploration.}
+%
 The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
@@
     \end{equation}
-    where
-    \begin{equation}
-        {\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}.
-    \end{equation}
+    where ${\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 
-One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ time steps.
-Hence those guarantees are valid on ``easy'' problem instances only, with few changes of a large magnitude, and in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
-%
-However, this does not prevent our algorithms from performing well on more realistic instances, and numerical experiments support this claim,
-as illustrated for instance with problem $3$ presented in the next Section~\ref{sec:6:NumericalExperiments}).
-%
-Note that \MUCB{} \cite{CaoZhenKvetonXie18} is also analyzed for the same type of unrealistic assumptions, while its practical performance is illustrated beyond those.
+One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ rounds.
+Hence those guarantees are valid on ``easy'' problems only, with few changes of large magnitudes, in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
+%
+However, this does not prevent our algorithms from performing well on more realistic instances, as shown by the numerical experiments,
+for instance with problem $3$ presented in the next Section~\ref{sec:6:NumericalExperiments}).
+%
+And \MUCB{} \cite{CaoZhenKvetonXie18} is also analyzed for the same type of unrealistic assumptions, while its practical performance is illustrated beyond those.
 
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    With $\cE_T(\delta)$ defined as in \eqref{def:6:GoodEvenGlobal}, the ``bad event'' is unlikely:
-    \begin{equation}
-        \bP(\cE_T^c(\delta)) \leq \delta(K+1)\Upsilon_T.
-    \end{equation}
+    With $\cE_T(\delta)$ defined as in \eqref{def:6:GoodEvenGlobal}, the ``bad event'' is unlikely
+    % \begin{equation}
+        $\bP(\cE_T^c(\delta)) \leq \delta(K+1)\Upsilon_T$.
+    % \end{equation}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 
-Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1} below displays the locations of the successively detected change-points by four algorithms,
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
+    \caption{Locations of the detected change-points for four algorithms on Problem 1.}
+    \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1}
+\end{figure}
+
+Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1} above displays the locations of the successively detected change-points by four algorithms,
 $1)$ \MklUCB,
@@
 
+\textbf{Problem $\bm{2}$} has only global changes, and important changes happen only at $t=1000$ when arm $2$ sees its mean change from $0.9$ to $0.7$, then at $t=2000$ when it becomes sub-optimal ($\mu_2$ goes from $0.7$ to $0.5$) and arm $0$ becomes optimal, and at $t=3000$ and $t=4000$ when arm $0$ stays optimal but sees its mean change from $0.6$ to $0.7$ and then from $0.7$ to $0.8$.
+
 \begin{figure}[h!]  % [htbp]
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
-    \caption{Locations of the detected change-points for four algorithms on Problem 1.}
-    \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
+    \caption{Locations of the detected change-points for four algorithms on Problem 2.}
+    \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2}
 \end{figure}
 
-\textbf{Problem $\bm{2}$} has only global changes, and important changes happen only at $t=1000$ when arm $2$ sees its mean change from $0.9$ to $0.7$, then at $t=2000$ when it becomes sub-optimal ($\mu_2$ goes from $0.7$ to $0.5$) and arm $0$ becomes optimal, and at $t=3000$ and $t=4000$ when arm $0$ stays optimal but sees its mean change from $0.6$ to $0.7$ and then from $0.7$ to $0.8$.
-
-Like for the first problem, we illustrate the behavior of the same four algorithms in Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2} below.
+Like for the first problem, we illustrate the behavior of the same four algorithms in Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2} above.
 In this other problem, the difference between \MklUCB{} and \CUSUMklUCB{} is clear:
@@
 % \end{itemize}
-
-\begin{figure}[h!]  % [htbp]
-    \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
-    \caption{Locations of the detected change-points for four algorithms on Problem 2.}
-    \label{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2}
-\end{figure}
 
@@
 
+% -----------------------------------------------------------------
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
+    \caption{Histograms of the distributions of regret $R_T$ ($T=5000$) for problem 1.}
+    \label{fig:6:histogramRegretPb1}
+\end{figure}
+
 \newpage

### 2019-06-08 2-Chapters/2-Chapter/chapter2.tex
A la BU le samedi après-midi ouais

@@
 Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to use.
-It is hard to predict which solution could be the best for real-world conditions at every instants,
+It is hard to predict which solution could be the best for real-world conditions at every instant,
 and even by assuming a stationary environment, when one is facing a certain problem but has limited information about it, it is hard to know beforehand which algorithm can be the best solution.

### 2019-06-08 2-Chapters/3-Chapter/chapter3.tex
A la BU le samedi après-midi ouais

@@
 The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-player) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
-After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithms, as well as regret and other visualizations.
+After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
 The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with a low variance.
@@
     % , and it consists in trying to fix a simple observation.
-    % For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arms (\eg, mean).
+    % For instance in \UCB, if arm $1$ has $100$ samples and arm $2$ has $50$ samples, it should make more sense to only use $50$ samples of arm $1$ to compute and compare any statistics for each arm (\eg, mean).
     % The BESA algorithm for $K=2$ arms is then very simple: at each time step, if arms $1$ and $2$ have $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
@@
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
-One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
+One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
 More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.

### 2019-06-08 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
A la BU le samedi après-midi ouais

@@
 We consider the network from the point of view of one dynamic device.
-Following a Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
+Following a Bernoulli emission process, every time a dynamic device has to communicate with the gateway,
 it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $A(t) = k \in[K]$.
@@
 Besides the modified time scale (rewards are no longer collected at every time step), this looks like a usual bandit problem.
-However, it cannot be modeled as a stochastic MAB, as the rewards are (unfortunately) \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning and random activation of each devices).
+However, it cannot be modeled as a stochastic MAB, as the rewards are (unfortunately) \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other dynamic devices, that is not stationary (because of learning and random activation of each device).
 %

### 2019-06-08 2-Chapters/4-Chapter/chapter4.tex
CHAUD
I have some small bugs, but all is good
A la BU le samedi après-midi ouais

@@
 FIXME finish !
+
+\TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
 
@@
 
-\paragraph{Publications.}
+\textbf{Publications.}
 
@@
 This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
-The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each others, and \emph{do not} know the system parameters.
+The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.
 

### 2019-06-08 2-Chapters/5-Chapter/chapter5.tex
A la BU le samedi après-midi ouais

@@
 The solution they proposed combines Multi-Armed Bandit learning with a lightweight signalling-based coordination scheme, and ensures convergence to a stable allocation of resources.
-Their work considers single-user level algorithms for two scenarios: an unknown fixed number of users, and a dynamic number of users, both for different arms means for each users (players).
+Their work considers single-user level algorithms for two scenarios: an unknown fixed number of users, and a dynamic number of users, both for different arms means for each user (players).
 Analytic performance guarantees, proving convergence to stable marriage configurations, are presented for both setups.
@@
 Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
-We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future work left on our library.
+We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future works left on our library.
 For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.

### 2019-06-08 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
A la BU le samedi après-midi ouais

@@
 %
-A small false alarm probability ensures that no detection occurs before it should (\ie, no useless detection occur on stationary segments), and a small detection delay ensures that, if the stationary segments are long enough, then every change-points on every arms will be detected after a short enough amount of samples from that arm.
+A small false alarm probability ensures that no detection occurs before it should (\ie, no useless detection occur on stationary segments), and a small detection delay ensures that, if the stationary segments are long enough, then every change-points on every arm will be detected after a short enough amount of samples from that arm.
 %
@@
 The two Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2} show the final regret $R_T$ obtained for each algorithm.
-Results highlighted in \textbf{bold} show the best non-oracle algorithm for each experiments,
+Results highlighted in \textbf{bold} show the best non-oracle algorithm for each experiment,
 with our proposal being the best non-oracle strategy for problems $1$ and $2$.
@@
 On these problems, \GLRklUCB{} with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts.
-Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instances.
+Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instance.
 The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17},
@@
 We show below on Figure~\ref{fig:6:meanRegretPb1} and additional in Appendix~\ref{app:6:moreFigures} the simulation results for the five problems.
-The results in terms of mean regret $R_T$ are given above in Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2}, but it is also interesting to observe two plots for each experiments.
+The results in terms of mean regret $R_T$ are given above in Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2}, but it is also interesting to observe two plots for each experiment.
 %

### 2019-06-08 2-Chapters/Conclusion/conclusion.tex
A la BU le samedi après-midi ouais

@@
 
-A more precise summary of each chapters is included below.
+A more precise summary of each chapter is included below.
 
@@
 Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
-The notations used in the manuscript were introduced by showcasing an interactive demonstration made available online\footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
+The notations used in the manuscript were introduced by showcasing an interactive demonstration published online\footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
 
@@
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
-Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for the practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
 We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
@@
 
-We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
-As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
+We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
+Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions.
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
-One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, I we tried to follow actively the state-of-the-art research, since December $2016$ to June $2019$.
+One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
+% We indeed followed actively the state-of-the-art research, since December $2016$ to June $2019$.
 More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary case.
-The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modifications.
+The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
 %
@@
 Similarly, if the number of devices remain constant, increasing the successful transmission rate directly extends the IoT devices battery life, as they suffer less from failed transmissions.
-We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
+We show that both \UCB{} and Thompson sampling algorithms have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
 %
 This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
-The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each others, and \emph{do not} know the system parameters.
+The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.
 
@@
 If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
-The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
-We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
+We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
+The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a multi-armed bandit algorithm for this second-stage optimization problem.
 %
@@
 Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
-Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.
+Surprisingly, the main take-away message is that a simple \UCB{} learning approach (that simply retransmits in the same channel), turns out to perform as well as more complicated heuristics.
 
@@
 
-In the last chapter, we studied and presented the piece-wise stationary bandit model.
+In the last chapter, we were interested in relaxing the stationary hypothesis made in Chapter~\ref{chapter:2}.
 %
-We proposed a new algorithm for this problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
+We first describe the break-point detection problem, that we solve by analyzing at finite-time the properties of the Bernoulli GLR change-point detector.
+This sequential break-point detection test can be applied for bounded rewards, and we bound its false alarm probability as well as its detection delay.
+For the piece-wise stationary bandit model,
+we proposed a new efficient algorithm, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
 This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
@@
 
-As discussed at the end of each chapters,
+As discussed at the end of each chapter,
 our works suggest possible directions of future studies.
@@
 As detailed in Chapter~\ref{chapter:2}, we focussed in this thesis on finite-arms and one-dimensional bandit problems,
-and thus two possible directions of future works could be to extend our works
+and thus two possible directions of future research could be to extend our works
 to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
@@
 
-A first future work left on our library SMPyBandits could be to implement new variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
+A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
 For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
 %
-It would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
+A second interesting task could be to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
 or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+%
+A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
+It is already possible to easily reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if one installs or downloads the library, but can also be used on cloud platforms, such as
+Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} for free.},
+or Google Colab\footnote{~See \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}.}.
 
@@
 
-In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our demonstration \cite{Besson2018ICT,Besson2019WCNC}.
+In order to validate our results in a realistic experimental setting and not only with simulations, future works should include a hardware implementation of the analyzed models, in order to complete our demonstration presented in Section~\ref{sec:4:gnuradio} \cite{Besson2018ICT,Besson2019WCNC}.
 %
-A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.
+A hardware demonstrator could be also benefit to study other settings by relaxing some hypotheses, for instance by studying a similar model in non-slotted time.
 
@@
 \MCTopM{} for orthogonalization (\ie, dealing with collisions),
-and \GLRklUCB{} for non-stationarity (\ie, dealing with abrupt changes).
+and the Bernoulli GLR break-point detector for non-stationarity (\ie, dealing with abrupt changes).
+The three parts should be connected, the same way we built \GLRklUCB{} by connecting both components.
+We propose the idea of incorporating the detected change-points by B-\GLR{} test in the orthogonalization procedure: after any change-point is detected, the player is not sitting anymore.
+Analyzing this strategy is left as a future work.
 
-\TODOL{Explain that we think the following approach can work very efficiently:
+% \TODOL{Explain that we think the following approach can work very efficiently:
 
-    - Combine \MCTopM{} with \GLRklUCB{} instead of \klUCB,
-    - and incorporate also the detected change-points by B-\GLR{} test in the orthogonalization procedure: after any change-point, the player is not sitting anymore.
-    - Modification: only a change-point on the arm on which player is sitting can make him move?
+%     - Combine \MCTopM{} with \GLRklUCB{} instead of \klUCB,
+%     - and incorporate also the detected change-points by B-\GLR{} test in the orthogonalization procedure: after any change-point, the player is not sitting anymore.
+%     - Modification: only a change-point on the arm on which player is sitting can make him move?
 It's already the case: the \GLR{} test can only detect change-point on an arm that was played, and the \MCTopM{} orthogonalization procedure forces to play the same arm as long as the player is sitting.
-}
+% }
 
@@
 \paragraph{Other directions of future works.}
+
+Another very interesting direction of research is to propose a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and automatically adapt to the setting at hand.
+For instance, even if \GLRklUCB{} is very efficient for piece-wise stationary problems, its forced exploration makes it sub-optimal for stationary problems.
+Another example is for multi-players bandits, where \MCTopM-\klUCB{} performs sub-optimally on piece-wise stationary problems.
+% if we consider the proposed extension of \MCTopM-\klUCB{} that estimates the number $M$ of players, i
+One approach to obtain a unified ``master'' algorithm could be to use expert aggregation on the different state-of-the-art algorithms presented in this thesis, and as each one is efficient in a different setting, the resulting aggregated algorithm could also be efficient in each of the different settings.
+But as we illustrated in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, this approach does not seem theoretically promising.
 

### 2019-06-08 1-Preamble/5-Abstract/abstract.tex
Starting to read everything to hunt := notations

@@
 % Use `abstract' as an option in the document class to print only the titlepage and the abstract.
+% \vspace*{-30pt}
+%
 \begin{abstract}
@@
 % , Chapitre~\ref{chapitre:2}
-car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication ni aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
 % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.

### 2019-06-08 1-Preamble/5-Abstract/resume_fr.tex
Starting to read everything to hunt := notations

@@
 % Résumé de la thèse en français.
+
+\TODOL{Je ne sais pas quelle longueur ni quel niveau de détail il faut viser ici.
+Est-ce que ces 3/4 pages suffisent ?
+Est-ce qu'il faut donner des maths en français pour CHAQUE CHAPITRE ?
+J'ai un peu la flemme d'écrire 20 pages avec des maths en français juste pour pouvoir citer tel théorème ou tel résultat.}
 
@@
     \item
-    Le code source des deux contributions susmentionnées utilise GNU Octave ou MATLAB et est publié en ligne, ainsi que des instructions claires pour la reproduction de nos travaux.
+    Le code source des deux contributions susmentionnées est publié en ligne, ainsi que des instructions pour la reproduction de nos travaux (avec GNU Octave ou MATLAB).
 
@@
 
+% WARNING
+\vfill{}
+
+\textbf{Note sur le droit intellectuel.}
+%
+Ce document et les ressources additionnelles requises pour le compiler (notamment les fichiers \LaTeX, les morceaux de code Python, les figures etc)
+sont \href{https://github.com/Naereen/phd-thesis/}{publiées publiquement},
+selon les termes de la \href{https://lbesson.mit-license.org/}{\emph{licence MIT}} open-source,
+en ligne sur \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}}.
+
+\TODOL{Open source the repository as soon as I defended my thesis!}
+
+\begin{center}
+    \textbf{Copyright 2016-2019, \copyright ~Lilian~Besson.}
+\end{center}
+
 \end{resume_fr}

### 2019-06-08 2-Chapters/1-Chapter/chapter1.tex
Starting to read everything to hunt := notations

@@
 %
-The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but use the notations introduced on Chapter~\ref{chapter:2}.
+The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but all use the notations introduced on Chapter~\ref{chapter:2}.
 
@@
     for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
-    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in May $2019$.
+    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in June $2019$.
 
@@
     I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
-    with Pierre Haessig (\texttt{pierreh.eu}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.
+    with Pierre Haessig (see \href{https://pierreh.eu/}{\texttt{pierreh.eu}}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.
 

### 2019-06-08 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Starting to read everything to hunt := notations

@@
     \includegraphics[width=0.70\linewidth]{system_model1.eps}
-    \caption{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
+    \caption[In our system model, some dynamic devices (in the IoT network in blue) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in orange).]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
     \label{fig:41:system_model1}
@@
     \includegraphics[scale=0.40]{protocol.eps}
-    \caption{The considered time-frequency slotted protocol.
+    \caption[The considered time-frequency slotted protocol.
+Each frame is composed by a fixed duration uplink slot in which the end-devices transmit their (uplink) packets.
+If a packet is well received, the gateway replies by transmitting an \Ack, after the ack delay.]{The considered time-frequency slotted protocol.
 Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Starting to read everything to hunt := notations

@@
     % \end{subfigure}
-    \caption[Multi-player regret with its three terms]{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
+    \caption[Illustrating the decomposition of the multi-player regret with its three terms]{Illustrating the decomposition of the multi-player regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
     \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}

### 2019-06-08 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Starting to read everything to hunt := notations

@@
     \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
-    \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).
+    \caption{Problem $1$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).
         The means are in $[0,1]$, and there are $C+1=5$ stationary intervals of equal lengths.
@@
     \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
-    \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).
+    \caption{Problem $2$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).
         The means are again in $[0,1]$, and there are also $C+1=5$ stationary intervals of equal lengths.
@@
     \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
-    \caption{\textbf{Problem $3$}: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
+    \caption{Problem $3$: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
     \label{fig:6:Problem_5}
@@
     \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
-    \caption{\textbf{Problem $4$}: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
+    \caption{Problem $4$: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
     % \caption{\textbf{Problem 4}: $K=3$ arms with $T=5000$, $\Upsilon=4$ changes occur on all arms at a time (\ie, $C=12$).}
@@
     \includegraphics[width=0.95\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
-    \caption{\textbf{Pb $5$}: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
+    \caption{Pb $5$: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
     % \caption{\textbf{Problem 5}: $K=5$, $T=100000$, $\Upsilon=81$ changes occurring on some arms at every time ($C=179$).}

### 2019-06-08 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
A la BU le samedi après-midi ouais
Proofread all Ch2 and added x := rhs when needed

@@
 and let $\Alg_{\mathrm{aggr}}$ be an aggregation algorithm,
-which runs the $N$ algorithms in parallel (with the same slotted time), and use them to choose its channels based on a voting from their $N$ decisions.
+which runs the $N$ algorithms in parallel (with the same slotted time), and use them to choose its arms based on a voting from their $N$ decisions.
 %
@@
 \begin{equation}
-	\widehat{r^a}(t) = \frac{Y_{A_a(t),t}}{p_{A_a(t)}^{t}} \indic(A_a(t) = A(t))
+	\widehat{r^a}(t) := \frac{Y_{A_a(t),t}}{p_{A_a(t)}^{t}} \indic(A_a(t) = A(t))
 \end{equation}
@@
 % helps to choose a good sequence.
-If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t = \eta = 2 \log(N) / (T K)$.
+If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t := \eta = 2 \log(N) / (T K)$.
 However, for real-world communication problems, it can be considered unrealistic\footnote{~For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}.} to assume a fixed and known time horizon, so we prefer the alternative horizon-free choice%
@@
 of learning rates,
-$\eta_t = \log(N) / (t K)$ suggested by \cite{Bubeck12}.
+$\eta_t := \log(N) / (t K)$ suggested by \cite{Bubeck12}.
 We compare both approaches empirically, and the second one usually performs better.
@@
 % - explain the problem studied in the simulation
-We focus on \emph{i.i.d.} MAB problems, with $K = 9$ channels\footnote{~Similar behaviors are observed for other values, \eg, trying up-to $K = 100$ gave the same results.}.
+We focus on \emph{i.i.d.} MAB problems, with $K = 9$ arms\footnote{~Similar behaviors are observed for other values, \eg, trying up-to $K = 100$ gave the same results.}.
 For Bernoulli problem, the first one uses $\boldsymbol{\mu}=[0.1,\dots,0.9]$, and is considered as a ``simple'' problem.
@@
 
-Note that for each of the $1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{1\leq k \leq K, 1 \leq t \leq T}$ for every repetition, in order to compare the algorithms on the same realizations%
+Note that for each of the $N=1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{1\leq k \leq K, 1 \leq t \leq T}$ (for every repetition), in order to compare the algorithms on the same realizations%
 \footnote{~Similar plots and similar results are obtained if this choice is not made, but it makes more sense to compare them against the same randomization.
@@
 %
-The aggregated algorithms consist in a naive uniform exploration (to have at least one algorithm with bad performances, \ie, linear regret, but it is not included in the plots),
+The aggregated algorithms consist in:
+a naive uniform exploration (to have at least one algorithm with bad performances, \ie, linear regret, that is \emph{not} included in the plots to reduce clutter),
 \UCB{} with $\alpha=1/2$, three \klUCB{} algorithms respectively using Bernoulli, Gaussian and exponential $\kl$ functions, and BayesUCB and Thompson sampling with uniform prior.
@@
 	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
-	\caption{On a harder problem, mixing Bernoulli, Gaussian, Exponential arms, with 3 arms of each types with the \emph{same mean}.}
+	\caption{On a harder problem, mixing Bernoulli, Gaussian, Exponential arms, with 3 arms of each type with the \emph{same mean}.}
 	\label{fig:25:HarderMixed}
@@
 	\centering
-	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
+	\includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
 	\caption{The semilog-$x$ scale clearly shows the logarithmic growth of the regret for the best algorithms and our proposal \Aggr, even in a hard ``mixed'' problem (\emph{cf}.
@@
 \cite[Theorem 4.2]{Bubeck12} proves that
-\ExpQ{} satisfies a bound on adversarial regret, % $\overline{R_T}$
+\ExpQ{} have a bounded adversarial regret, % $\overline{R_T}$
 $\overline{R_T} \leq 2 \sqrt{T N \log(K)}$,

### 2019-06-08 2-Chapters/2-Chapter/chapter2.tex
Proofread all Ch2 and added x := rhs when needed

@@
 \textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
-When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) = Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
+When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) := Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
 %
@@
 This quantity $\sum_{t=1}^T r(t)$ can indeed be random, and it usually depends on two aspects.
-First, the rewards may depend on the randomness of the environment, \ie, the unknown and unpredictable values $r(t)=Y_{k,t}$.
+First, the rewards may depend on the randomness of the environment, \ie, the unknown and unpredictable values $r(t)=Y_{A(t),t}$.
 Then this quantity also depends on the player's decision-making policy, \ie, the choices $A(t)$, that are based on (all) the \emph{past} observations, \ie, $A(1)$, $Y_{A(1),1}$, $\dots$, $A(t-1)$, $Y_{A(t-1),t-1}$, and possibly on an external source of randomness.
@@
 The UI of the demo also shows the current value of $t$ (``total plays'') and $\sum_{s=1}^t r(s)$ the ``total reward''.
-For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) = \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
-%
-The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated Probs'' line.
+For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) := \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) := \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
+%
+The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) := X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated Probs'' line.
 
@@
 More formally, an algorithm can be defined as a sequence of \emph{measurable} functions $(\cA_t)_{t\geq1}$,
-where $\cA_t$ maps the past observations $O_t = (U_0, Y_{A(1),1}, U_1, \dots, Y_{A(t-1),t-1}, U_{t-1})$
-to an arm $\cA_t(O_t) = A(t) \in[K]$
+where $\cA_t$ maps the past observations $O_t := (U_0, Y_{A(1),1}, U_1, \dots, Y_{A(t-1),t-1}, U_{t-1})$
+to an arm $\cA_t(O_t) := A(t) \in[K]$
 (we remind that we denote $r(s) = Y_{A(s),s}$ the $s$-th reward).
 The initial information is reduced to $O_1 = (U_0)$, and the first decision is $A(1) = \cA_1(O_1)$.
-Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps (\eg, Algorithm~\ref{algo:2:indexPolicy}).
+Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps $t=1,\dots,K$ (\eg, Algorithm~\ref{algo:2:indexPolicy}).
 % as illustrated for example in Algorithm~\ref{algo:2:indexPolicy} below.
@@
 %
-If the player knows that reward are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ (for $a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = $ that lies in $[0,1]$.
+If the player knows that reward are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ (for $a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = \frac{r-a}{b-a}$ that lies in $[0,1]$.
 Note that this ``normalization trick'' is implemented for any policy in our library SMPyBandits, with the \texttt{lower} and \texttt{amplitude} optional arguments, respectively representing $a$ the lower-bound on rewards and $b-a$ the length of the interval of possible values of rewards.
@@
 for a parameter vector $\eta$ (the canonical parameter), and a sufficient statistic that we restrict to be $T(x)=x$, and a function $h$.
-$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$ and $h$,
+The cumulant function $A(\eta)$ is entirely determined by the $\eta$ and $h$,
 as $A(\eta) = \log \left( \int h(x) \exp(\eta x) \lambda(\mathrm{d} x) \right)$.
@@
     \begin{align}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
-        \KL\left(d_1, d_2\right) = \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}\cup\{\pm\infty\}.
+        \KL\left(d_1, d_2\right) := \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}\cup\{\pm\infty\}.
     \end{align}
@@
 $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$,
-we can write $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
+we can use a shorter notation and write $\KL( \eta_1, \eta_2 ) := \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
 and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$,
@@
     The KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
-    $\kl(x,y) = \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
+    $\kl(x,y) := \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
     and it is also called the \emph{relative binary entropy}.
@@
     The KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
-    $d_{\sigma^2}(x,y) = \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$
+    $d_{\sigma^2}(x,y) := \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$
     (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
@@
 %
-Then, when we study multi-player bandits in Chapter~\ref{chapter:5}, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $r(t)=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
+Then, when we study multi-player bandits in Chapter~\ref{chapter:5}, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $Y_{k,t}=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
 %
@@
     Other possible use cases include hyper-parameter optimization, or features selection.
-    Hyper-parameters include real-valued parameters, like $\alpha\in[0,\infty)$ for the $\UCB_1$ or $\varepsilon$ for the $\varepsilon$-greedy bandit algorithms, a step size multiplier in a gradient method, or the width of a radial based function (RBF) kernel method.
+    Hyper-parameters include real-valued parameters, like $\alpha\in[\frac{1}{2},\infty)$ for the $\UCB_1$ or $\varepsilon\in(0,1)$ for the $\varepsilon$-greedy bandit algorithms, a step size multiplier in a gradient method, or the width of a radial based function (RBF) kernel method.
     Discrete-valued parameters are also common, like a choice in a fixed set of kernel functions or the depth of neural networks,
@@
 Let us first introduce some notations.
-From now on, we consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that generates \iid{} samples $Y_{k,t} \sim \nu_k$, for any time $t$.
+We consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that generates \iid{} samples $Y_{k,t} \sim \nu_k$, for any time $t$.
 % with $r(t)=Y_{A(t),t}$).
 We focus on distributions fully characterized by their means, that can be Bernoulli or any one-dimensional exponential family.
-We denote $\mu_k = \E[Y_{k,t}] \in \R$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$}).
+We denote $\mu_k := \E[Y_{k,t}] \in \R$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$}).
 %
@@
     Consider a bandit problem of $K$ arms with distributions of means $\bm{\mu}=\mu_1,\dots,\mu_k$.
-    Denote $\mu^* = \max_k \mu_k$ the largest mean.
+    Denote $\mu^* := \max_k \mu_k$ the largest mean.
     The best arm can be non unique, and any arm $k$ having $\mu_k = \mu^*$ is said to be \emph{optimal},
@@
 let us introduce the (mean) regret $R_T^{\cA}$ of an algorithm $\cA$ as
-$R_T^{\cA} = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]$.
+$R_T^{\cA} := \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]$.
 %
@@
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\bm{\mu} = \mu_1,\dots,\mu_K$, with $\mu^* = \max_k \mu_k$, then its \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ defined as
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\mu_1,\dots,\mu_K$, with $\mu^* := \max_k \mu_k$, then its \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ defined as
     %
@@
         R_T^{\cA}
-        = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
+        := \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
         % = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right]
@@
 
+One first need to observe that $R_T^{\cA} \geq 0$ for any $T$ and $\cA$, and that $R_T^{\cA} \leq \mu^* T$ (if rewards lie in $[0,1]$).
+Thus any algorithm has $R_T^{\cA} = \bigO{T}$, which justifies why we are interested in efficient algorithms that achieve at least a sub-linear regret, \ie, $R_T^{\cA} = \smallO{T}$.
+
 \paragraph{A useful decomposition of the regret.}
 %
-Remember that $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ denotes the number of times arm $k$ was selected between times $1$ and $t$,
+Remember that $N_k(t) := \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ denotes the number of times arm $k$ was selected between times $1$ and $t$,
 %
 and that the samples $Y_{k,t}$ are all \iid{} of mean $\mu_k$.
-The \emph{gap} between any arm $k\in[K]$ and an optimal arm is defined as $\Delta_k = \mu^* - \mu_k$
+The \emph{gap} between any arm $k\in[K]$ and an optimal arm is defined as $\Delta_k := \mu^* - \mu_k$
 (an arm $k$ is thus sub-optimal if and only if $\Delta_k > 0$),
@@
 and a significant part of the research literature on MAB algorithms has focussed on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
-This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically match the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
+This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
 
@@
     there exists a constant $C_I$ depending only on $I$,
-    and a time $T_0$ such that
-    \begin{equation}
-        \forall T \geq T_0, \;\;\; R_T^{\cA}(I) \geq C_I \log(T).
-    \end{equation}
-    % \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    and a time $T_0$ such that $\forall T \geq T_0, \;\;\; R_T^{\cA}(I) \geq C_I \log(T)$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 % - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $\mathrm{SC}^{\cA}(T) = \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
+The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $\mathrm{SC}^{\cA}(T) := \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
 It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptatively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.
@@
 Indeed, in single-player models it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
-But even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
-%
-It is more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
+But even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.\\
+%
+We find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
 Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
@@
 
-\paragraph{Implementation.}
+\textbf{Implementation.}
 %
@@
 They are detailed in Algorithm~\ref{algo:2:naiveStrategies} below.
-We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) = \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) = \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
-%
-The estimated means, or empirical averages, are $\widehat{\mu_k}(t) = X_k(t) / N_k(t)$ (when $N_k(t)>0$).
+We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) := \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) := \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
+%
+The estimated means, or empirical averages, are $\widehat{\mu_k}(t) := X_k(t) / N_k(t)$ (when $N_k(t)>0$).
 
@@
 On the other hand, if a lower-bound $d$ on the positive gaps is known beforehand,
-we can consider $\varepsilon_t$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
+we can consider a sequence $(\varepsilon_t)_{t\in\N^*}$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
 Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
@@
 
-\begin{lemma}[Hoeffding's inequality]\label{lem:2:HoeffdingInequality}
+\begin{lemma}\label{lem:2:HoeffdingInequality}
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    Let $Z_1,\dots,Z_n$ be $n$ \iid{} samples from a Bernoulli distribution of parameter $\mu$, of empirical mean $\widehat{Z_n} = \frac{1}{n} \sum\limits_{i=1}^n Z_i$, Hoeffding's inequality gives both
+    Let $Z_1,\dots,Z_n$ be $n$ \iid{} samples from a Bernoulli distribution of parameter $\mu$ (where $n$ is fixed),
+    of empirical mean $\widehat{Z_n} := \frac{1}{n} \sum\limits_{i=1}^n Z_i$, \emph{Hoeffding's inequality} gives both
     \begin{align}\label{eq:2:HoeffdingInequality}
@@
 For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
-the \emph{average reward} $\widehat{\mu_k}(t)=\frac{X_k(t)}{N_k(t)}$
-and a \emph{confidence radius} $\xi_k(t) = \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
+the \emph{average reward} $\widehat{\mu_k}(t) := \frac{X_k(t)}{N_k(t)}$
+and a \emph{confidence radius} $\xi_k(t) := \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
 %
 \begin{align}\label{eq:UCB_index}
-    U_k^{\UCB_1}(t) = \UCB_k(t) = \underbrace{\frac{X_k(t)}{N_k(t)}}_{\widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\xi_k(t)}.
+    U_k^{\UCB_1}(t) = \UCB_k(t) := \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation} \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration} \xi_k(t)}.
 \end{align}
@@
 but we preferred to start with a simpler proof in the single-player stationary case.
-µ
-What this theorem shows is that the $\UCB_1$ algorithm achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret (if $\Delta=\min\limits_{k: \Delta_k>0} \Delta$).
+%
+This theorem shows that $\UCB_1$ achieves a order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta := \min\limits_{k: \Delta_k>0} \Delta$).
 
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    For simplicity, we restrict to $\alpha=2$.
     For any instance $I$ with $K$ arms with bounded distributions in $[0,1]$,
     of means $\mu_1,\dots,\mu_k$, and any horizon $T>K$,
-    the $\UCB_1$ algorithm verifies the following finite-time regret bound
+    the $\UCB_1$ algorithm with its parameter\footnotemark{} $\alpha=2$ verifies the following finite-time regret bound
     \begin{equation}\label{eq:2:UCBregretBound}
@@
 %
+\footnotetext{~For simplicity we present the theorem and its prove in the restricted case of $\alpha=2$.
+More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs, in particular for any value of $\alpha\geq\frac{1}{2}$.}%
+%
+\vspace*{-10pt}  % FIXME vspace negative
 \begin{smallproof}\label{proof:2:UCBregretBound}
@@
         N_k(T) = 1 + \sum_{t=K+1}^T \indic(A(t) = k)
-        \leq n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n).
+        \leq 1 + n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n).
     \end{equation}
@@
     Remember that $\UCB_a(t) = \widehat{\mu_a}(t) + \xi_a(t)$ for any arm $a$.
-    According to the algorithm (see Algorithm~\ref{algo:2:indexPolicy} at line~6),
-    arm $k$ is chosen if $\UCB_k(t) \geq \UCB_{k^*}(t)$.
-    Both $\widehat{\mu_a}(t)$ and $\xi_a(t)$ use $N_k(t)$ samples of arm $a$,
-    but as \eqref{eq:2:trickInUCBProof} above started to isolate $N_k(t)$, we want to add the freedom of considering different number of samples (using $n$), so we introduce the notation
-    $\hat{\mu}_{a,n}(t)$ and $\xi_{a,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
-    \ie, $\hat{\mu}_{a,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
+    According to the Algorithm~\ref{algo:2:indexPolicy} (see line~6),
+    a suboptimal arm $k$ can be chosen only if $\UCB_k(t) \geq \UCB_{k^*}(t)$.
+    Both $\widehat{\mu_a}(t)$ and $\xi_a(t)$ use $n=N_k(t)$ samples of arm $a$,
+    but as \eqref{eq:2:trickInUCBProof} above started to isolate $N_k(t)$, we add the freedom of considering different number of samples (using $n$), so we introduce the notation
+    $\widehat{\mu}_{a,n}(t)$ and $\xi_{a,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
+    \ie, $\widehat{\mu}_{a,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
     %
@@
         N_k(T)
-        & \leq n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n) \nonumber \\
-        & \leq n + \sum_{t=n}^T \indic(\hat{\mu}_{k^*,N_{k^*}(t)}(t) + \xi_{k^*,N_{k^*}(t)}(t) < \hat{\mu}_{k,N_k(t)}(t) + \xi_{k,N_k(t)}(t), N_k(t) \geq n)  \nonumber \\
-        & = n + \sum_{t=n}^T \indic\left(\min_{0 < n_{k^*} < t} \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \max_{n < n_k < t} \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right)
+        & \leq 1 + n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n) \nonumber \\
+        & \leq 1 + n + \sum_{t=n}^T \indic(\widehat{\mu}_{k^*,N_{k^*}(t)}(t) + \xi_{k^*,N_{k^*}(t)}(t) < \widehat{\mu}_{k,N_k(t)}(t) + \xi_{k,N_k(t)}(t), N_k(t) \geq n)  \nonumber \\
+        & = 1 + n + \sum_{t=n}^T \indic\left(\min_{0 < n_{k^*} \leq t} \widehat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \max_{n < n_k \leq t} \widehat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right)
     \end{align}
-    So two union bounds on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$ give
+    So two union bounds on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$ give two sums
     \begin{align}\label{eq:2:decompositionInUCBProof2}
         N_k(T)
-        & \leq n + \sum_{t=n}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n}^t  \indic\left( \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right)
+        & \leq 1 + n + \sum_{t=n}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n}^t  \indic\left( \widehat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right)
     \end{align}
 
-    Now, at time $t$, this event $\hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t)$
+    Now, at time $t$, this event $\widehat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t)$
     implies at least one of the following three events:
@@
         \begin{cases}
-        I_1 \;:\; & \hat{\mu}_{k^*,n_{k^*}}(t) \leq \mu_{k^*} - \xi_{k^*,n_{k^*}}(t), \\
-        I_2 \;:\; & \hat{\mu}_{k,n_k}(t) \geq \mu_k + \xi_{k,n_k}(t), \\
-        I_3 \;:\; & \mu_{k^*} < \mu_k + 2 \; \xi_{k,n_k}(t).
+        I_1 \;:=\; & ( \widehat{\mu}_{k^*,n_{k^*}}(t) \leq \mu_{k^*} - \xi_{k^*,n_{k^*}}(t) ), \\
+        I_2 \;:=\; & ( \widehat{\mu}_{k,n_k}(t) \geq \mu_k + \xi_{k,n_k}(t) ), \\
+        I_3 \;:=\; & ( \mu_{k^*} < \mu_k + 2 \; \xi_{k,n_k} (t) ).
         \end{cases}
@@
         For the last event $I_3$,
-        observe that $\mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t)$ means $\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0$, where we remind the notation of the gap $\Delta_k = \mu_{k^*} - \mu_k$.
+        observe that $\mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t)$ means $\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0$, where we remind the notation of the gap $\Delta_k := \mu_{k^*} - \mu_k$.
         Thus if $n_k \geq \left\lceil \frac{8 \log(T)}{\Delta_k^2} \right\rceil \geq \left\lceil \frac{8 \log(t)}{\Delta_k^2} \right\rceil$,
@@
     Taking the expectation of $N_k(T)$ from \eqref{eq:2:decompositionInUCBProof},
-    and setting $n = n_T = \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil$ thus gives
+    and setting $n = n_T := \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil$ thus gives
     %
@@
         \E[ N_k(T) ]
-        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr( \hat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \hat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) ) \nonumber \\
-        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \underbrace{\Pr(I_1)}_{=1/t^4} + \underbrace{\Pr(I_2)}_{=1/t^4} + \underbrace{\Pr(I_3)}_{=0} \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr\left( \widehat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right) \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t \bigl[ \underbrace{\Pr(I_1)}_{=1/t^4} + \underbrace{\Pr(I_2)}_{=1/t^4} + \underbrace{\Pr(I_3)}_{=0} \bigr] \nonumber \\
         & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \nonumber \\
@@
     \begin{align*}
-        \E[N_k(T)] \leq \frac{8 \log(T)}{\Delta_k^2} + (1 + \frac{\pi^2}{3}).
+        \E[N_k(T)] \leq \frac{8 \log(T)}{\Delta_k^2} + \left(1 + \frac{\pi^2}{3}\right).
     \end{align*}
@@
     \end{align*}
-
-    More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs, in particular for any value of $\alpha\geq\frac{1}{2}$.
 \end{smallproof}
@@
 The indexes are computed as follows, for an exploration function $f(t)$
-that is typically chosen as $f(t)=\log(t)+3\log(\log(t))$ for the theoretical analyses,
-and $f(t)=\log(t)$ in practice.
-The constant $c>0$ is usually chosen as $c=1$.
+that is typically chosen as $f(t) := \log(t)+3\log(\log(t))$ for the theoretical analyses,
+and $f(t) := \log(t)$ in practice.
+The constant $c>0$ is usually chosen as $c = 1$.
 \begin{align}\label{eq:klUCB_index}
-    U_k^{\klUCB}(t) = \sup\limits_{q \in [0, 1]} \left\{ q : \KL\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
+    U_k^{\klUCB}(t) := \sup\limits_{q \in [0, 1]} \left\{ q : \KL\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
 \end{align}
 
-For Bernoulli distributions,
-we can easily justify the previous claim that \klUCB{} builds smaller confidence intervals (\ie, more accurate) than $\UCB_1$,
-\ie, $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$,
-because Pinkser's inequality gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix).
-% FIXME is this enough?
+For Bernoulli distributions, the previous claim that \klUCB{} builds smaller confidence intervals (\ie, more accurate) than $\UCB_1$, \ie, $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$,
+is justified by Pinkser's inequality gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix).
 
@@
 Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\KL(x,\bullet)$ function is usually convex (for any $x$), one practical and mathematically correct solution is to use Newton's method.
-The search space for $q$ can be restricted to $[\frac{X_k(t)}{N_k(t)}, 1]$ instead of $[0,1]$, and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
+The search space for $q$ can be restricted to $[ X_k(t) / N_k(t), 1 ]$ instead of $[0,1]$, and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
 For instance, $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically Newton's method terminates in between $8$ to $12$ steps.
@@
 If we perform numerical simulations with \klUCB{} algorithms, we are interested in having efficient implementations, and ideally \klUCB{} should not be significantly slower than \UCB.
-The bottleneck of the performance is the computations of the $\KL(x,y)$ functions and the optimization problem,
+The bottleneck of the performance is the computations of the $\KL$ functions and the optimization problem,
 and we note that our library SMPyBandits implements both parts in both a naive Python implementation (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback.py}{\texttt{kullback.py}}) and in a Python module written in a
-C file\footnote{~We also note that we wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
+C file\footnote{~Note that we also wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
     These implementations of $\KL(x,y)$ and \klUCB{} indexes are also published as an independent Python package, hosted on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} and installable from Pypi.
@@
 instead of building confidence intervals on the means of the arms like $\UCB_1$ does,
-it starts with a prior on the $K$ distributions (\eg, a flat uniform prior), and then after every observation $(k, r(t))$ it updates the posterior distribution of the arm $k$ with a new data $r(t)$.
-If $\pi_k(t)$ denotes this posterior distribution, Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim \pi_k(t)$, and then playing according to the ``optimism under uncertainty'' philosophy if these samples represent the arms (expected) quality: $A(t) \in \argmax_k s_k$.
+it starts with a prior $\pi_k(0)$ on the $K$ distributions (\eg, a flat uniform prior), and then after every observation $(k, r(t))$ it updates the posterior distribution of the arm $k$ with a new data $r(t)$, from $\pi_k(t)$ to $\pi_k(t+1)$.
+Thompson sampling makes its decision at time $t$ by first sampling a random sample $s_k \sim \pi_k(t)$, and then playing according to the ``optimism under uncertainty'' philosophy if these samples represent the arms (expected) quality, \ie, $A(t) \in \argmax_k s_k$.
 
 For problems with Bernoulli distributions on the $K$ arms,
-the best choice of prior is a Beta distribution, that starts as an initially uniform prior (\ie, $\mathrm{Beta}(1,1)$),
-and maintain a distribution $\mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$,
+the best choice of prior is a Beta distribution, that starts as an initially uniform prior (\ie, $\forall k, \pi_k(0) := \mathrm{Beta}(1,1)$),
+and maintain a distribution $\pi_k(t) := \mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$,
 as $X_k(t)$ counts the number of successes observed on arm $k$, and $N_k(t)-X_k(t)$ counts the number of failures.
@@
             Play arm $A(t)$ uniformly among the arms of maximal random sample: $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k(t))$\;
-            Observe a reward $r(t)$ and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$\;
+            Observe a reward $r(t)$ and update $X_k(t)$ and $N_k(t)$\;
+            %  and $\widehat{\mu_k}(t)$
 		}
@@
 
-Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the principle of optimism under uncertainty and the Bayesian point-of-view.
+Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the two aforementioned point-of-views.
+%  the principle of optimism under uncertainty and the Bayesian point-of-view.
 %  underlying the approaches based on UCB, might not be necessary at all.
 % With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
-The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as $\UCB_1$ or Thompson sampling, with uniform perturbations.
-Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$),
-but instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
+The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as $\UCB_1$ or Thompson sampling, with uniformly distributed perturbations.
+Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence interval $\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) := \widehat{\mu_k}(t) - \xi_k(t)$),
+but instead of maximizing the UCB, a random sample is taken on each interval $s_k \sim \cU(\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right])$, and the played arm is the one maximizing this sample, \ie, $A(t) \in \argmax_k s_k$.
 % https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html
@@
 %
-The Exp3 policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution:
-$\Pr(A(t)=k) = (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
-The weights are initially uniform, \ie, $w_k(0)=\frac{1}{K}$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
+The Exp3 policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
+$\Pr(A(t)=k) := (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
+The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
 The weights are then renormalized, to have a sum $\sum_k w_k(t+1) = 1$.
-This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = \Pr(A(t)=k)$.
-%
-Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a non-increasing sequence $(\gamma_t)_t$.
+This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = r(t) / \Pr(A(t)=k)$.
+%
+Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a non-increasing sequence $(\gamma_t)_{t\in\N^*}$.
 %
@@
 \begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp}
-    U^{\mathrm{MOSS}}_k(t) &= \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)} } \right). \\
-    U^{\KLUCBpp}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
+    U^{\mathrm{MOSS}}_k(t) & := \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)} } \right). \\
+    U^{\KLUCBpp}_k(t) & := \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
 \end{align}
@@
 \begin{align}\label{eq:2:indexes_for_klUCBswitch}
-    U^{\mathrm{MOSS}+}_k(t) &= \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
-    U^{\KLUCB+}_k(t) &= \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
-    U^{\klUCB\text{-}\mathrm{switch}}_k(t) &= \begin{cases}
+    U^{\mathrm{MOSS}+}_k(t) & := \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
+    U^{\KLUCB+}_k(t) & := \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
+    U^{\klUCB\text{-}\mathrm{switch}}_k(t) & := \begin{cases}
         U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
@@
 % \subsection{A first numerical experiment on the example Bernoulli problem}
-\subsection{Comparison of the main algorithms on the Bernoulli problem of the online demonstration}
+\subsection{Comparison of the main algorithms on the example Bernoulli problem from the online interactive demonstration}
 \label{sub:2:shortNumericalExperiments}
@@
 the $\kl$ function for \klUCB{} etc.
-Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
+Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instances, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
 %
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) = p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
+    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) := p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
 
@@
 \begin{smallproof}
-    Fix $p\in[0,1]$, and consider the function $g(x) = \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
+    Fix $p\in[0,1]$, and consider the function $g(x) := \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
     We first observe that $g(0) = 0$, and we easily prove that $x=0$ is the unique minimizer of $g$ over its domain.
-    By definition, $\kl$ is of class $\cC^1$ on its two variables, on $(0,1)$, so $g$ is of class $\cC^1$ on the interior of its domain, $(-p, 1-p)$.
+    By definition, $\kl$ is of class $\cC^1$ on its two variables, on the open interval $(0,1)$, so $g$ is of class $\cC^1$ on the interior of its domain, $(-p, 1-p)$.
     Let us compute its derivative, $g'(x) = (\frac{\partial \; \kl}{\partial 2})(p, p + x) - 4x$.

### 2019-06-08 2-Chapters/3-Chapter/chapter3.tex
Proofread all Ch2 and added x := rhs when needed

@@
 
-\TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
+% \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
 
@@
 and what kind of information is displayed, saved and plotted after a batch of simulations of bandit problems.
-
+%
+%
 The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-player) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
@@
 
-\paragraph{Simulation loop.}
+\textbf{Simulation loop.}
 %
@@
 
-\paragraph{Simulation loop for multi-player bandit.}
+\textbf{Simulation loop for multi-player bandit.}
 
@@
 For more details, an interested reader can refer to the following Jupyter notebook \cite{jupyter}:
-\begin{small}
-    \href{https://smpybandits.github.io/notebooks/Easily_creating_MAB_problems.html}{\texttt{SMPyBandits.GitHub.io/notebooks/Easily\_creating\_MAB\_problems.html}}
-\end{small}
-
+\href{https://smpybandits.github.io/notebooks/Easily_creating_MAB_problems.html}{\texttt{SMPyBandits.GitHub.io/notebooks/Easily\_creating\_MAB\_problems.html}}.\\
+%
 We have not yet tried to add support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
@@
 
-\subsection{MAB algorithms}
+\subsection{Multi-Armed Bandits algorithms}
 
@@
 
-Launching simulations is easy, for instance this snippet shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
-It takes about $20$ minutes, on a standard $4$-cores $64$ bits GNU/Linux laptop.
-Using environment variables (\texttt{N=1000} etc) in the command line is not required, but it is convenient:
+Simulations are easily executed, \eg, this snippet shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
+It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
+Environment variables (\texttt{N=1000} etc) in the command line are not required, but they are convenient.
 

### 2019-06-08 2-Chapters/3-Chapter/chapter3.tex
Done for proof reading chapter 4

@@
 and summary plots are automatically saved as high-quality PNG, PDF and EPS, using matplotlib \cite{matplotlib} and seaborn \cite{seaborn}.
-Raw data from each simulation is also saved in a HDF5 file using h5py \cite{h5py}, an efficient and compressed binary format, to allow easy post-mortem exploration of simulation results.
+Raw data from each simulation is also saved in an HDF5 file, using h5py \cite{h5py}, an efficient and compressed binary format, to allow easy post-mortem manipulation of any simulation results.
 Making new simulations is very easy, one only needs to write a configuration script (\texttt{configuration.py}), without needing a complete knowledge of the internal code architecture.
@@
 
-Simulations are easily executed, \eg, this snippet shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
+Simulations are easily executed, \eg, Listing~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
 It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
@@
 
-This library is written in Python \cite{python}, for versions 2.7+ or 3.4+, using \texttt{matplotlib} \cite{matplotlib} for 2D plotting, \texttt{numpy} \cite{numpy} for data storing, random number generations and operations on arrays, \texttt{scipy} \cite{scipy} for statistical and special functions, and \texttt{seaborn} \cite{seaborn} for pretty plotting and colorblind-aware colormaps.
+This library is written in Python \cite{python}, for versions 2.7+ or 3.4+, using \texttt{matplotlib} \cite{matplotlib} for 2D plotting, \texttt{numpy} \cite{numpy} for data storing, random number generations and operations on arrays, \texttt{scipy} \cite{scipy} for statistical and special functions, and \texttt{seaborn} \cite{seaborn} for clean plotting and colorblind-aware colormaps.
 
@@
 
-\paragraph{Read the Docs.}
+\textbf{Read the Docs.}
 Since $2017$, we have been using the free web service provided by Read the Docs (\href{https://readthedocs.org/}{\texttt{ReadTheDocs.org}}).
@@
 
-\paragraph{Travis CI.}
+\textbf{Travis CI.}
 Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
@@
 Additional results in terms of real measurements of time and memory are given and discussed in the next Section~\ref{sec:3:timeAndMemoryCosts}.
-%
+
 Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on two algorithms depending if simplicity or efficiency is favored:
@@
     $\mathrm{Exp}3^{++}$ from \cite{Seldin17}, using $\alpha=3$ and $\beta=256$ as advised (\texttt{Exp3PlusPlus}).
-    It is a recent variant of the $\mathrm{Exp3}$ algorithm, using an adaptive tuning of its parameter.
-It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+    It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
 
@@
     which is implemented with a binary tournament, written in a naive but efficient way
-    (\ie, not using recursive functions, by using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
+    (\ie, not using recursive functions, by using the \texttt{BESA} class with the \texttt{''non\_recursive''=True} option).
     But the extension to $K>2$ arms is still costly in terms of both time and memory, as illustrated below, and blows-up exponentially  when $K$ increases.
@@
 \paragraph{Two different problems.}
-
+%
 We consider two different problems, that are described here for a generic value of $K \geq 2$ the number of arms.
@@
 
-\begin{itemize}
-    \item \textbf{Problem 1} considers $K$ arms of equally spaced means, in $[0,1]$.
+% \begin{itemize}
+    % \item
+    \textbf{Problem 1} considers $K$ arms of equally spaced means, in $[0,1]$.
     It uses a minimum mean of $0.05$ and a maximum mean of $0.95$.
@@
 
-    \item \textbf{Problem 2} takes another approach. Instead of simulating $1000$ times the same problem, it generates a new random problem for each independent run.
+    % \item
+    \textbf{Problem 2} takes another approach. Instead of simulating $1000$ times the same problem, it generates a new random problem for each independent run.
     It considers a randomly generated vector of means, each being sampled uniformly at random in $[0,1]$, with a minimum gap of $\Delta^{\min}$, $\bm{\mu} \sim \cE(\Delta^{\min},\mu_{\min},\mu_{\max})$, as well as minimum and maximum values of $\mu_{\min}$ and $\mu_{\max}$.
@@
     % 0.1 if NB_ARMS <= 5 else 1. / (3 * NB_ARMS)
-\end{itemize}
+% \end{itemize}
 
@@
 % We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
-We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} that we shortly present in Appendix~\ref{app:2:DoublingTricks}.
+We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} (see Appendix~\ref{app:2:DoublingTricks}).
 
@@
 \indent
-Since $2017$, I also received by email or by GitHub issues some requests to implement new features or models.
+Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models.
 Most of them were implemented quickly, but some would require more time and are left as possible future works.
@@
     It is anytime, and was prove to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
-    Empirically, it usually outperforms \klUCB{} in terms of regret, and costs about the same memory, but it is slower.
+    Empirically, it usually outperforms (slightly) \klUCB{} in terms of regret, and costs about the same memory, but it is slower.
 
@@
     Like $\mathrm{RCB}$, $\mathrm{PHE}$ was found to be empirically slightly worse than other state-of-the-art algorithms.
-    Interestingly, its time and memory consumption stay constant with respect to the step number $t$ and the horizon $T$, because generating and summing $t$ pseudo-rewards from a Bernoulli distribution can actually be done in $\cO(1)$ time and space, by using efficient sampling methods for sampling from Binomial distributions.
+    Interestingly, its time and memory consumption stay constant with respect to the step number $t$ and the horizon $T$, because generating and summing $t$ pseudo-rewards from a Bernoulli distribution can actually be done in $\cO(1)$ time and space, by using efficient sampling methods for sampling from a Binomial distribution.
 \end{enumerate}

### 2019-06-08 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done for proof reading chapter 4

@@
 
-\paragraph{One gateway and many devices.}
+\textbf{One gateway and many devices.}
 %
@@
 The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
-We mean here by ALOHA-based that devices do not use sensing, and we do not consider retransmissions of packets (see Section~\ref{sec:4:retransmissions} below where we remove this last hypothesis).
+We mean here by ALOHA-based that devices do not use sensing, and we do not consider retransmissions of packets (we remove this hypothesis in Section~\ref{sec:4:retransmissions}).
 %
@@
     We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
-    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bit IEEE 754 float number for each of the $K$ channels).
+    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one float number for each of the $K$ channels).
     Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, to improve their successful communication rate automatically.
@@
 
-\paragraph{MAB model.}
+\paragraph{Link with the MAB model.}
 %
@@
 %
-For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r(t)=Y_{k,t} \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
+For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r(t) := Y_{k,t} \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
 The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
 $\sum_{t = 1}^T r(t),$
-which is a usual objective in MAB problems.
+which is the usual objective in MAB problems (see Section~\ref{sec:2:notations}).
 
@@
 % , in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
-Contrary to many previous work done in the CR field \cite{Jouini10,Jouini12} (\eg, Opportunistic Spectrum Access),
-the reward $r(t)=Y_{k,t}$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
+Contrarily to many previous work done in the CR field (\eg, Opportunistic Spectrum Access) \cite{Jouini10,Jouini12},
+the reward $r(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
 In our model, rewards rather come from receiving or not an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.
@@
 %
-Before presenting \emph{multi-armed bandit algorithms}, we present some natural baseline (reference) policies.
+Before presenting the MAB algorithms used in the experiments, we present some natural baseline (reference) policies.
 
@@
 They are however useful for our numerical simulations, and are used as references, to show that our MAB-based approaches quickly learn to perform almost optimally.
+%
+Their short names are used in the legend on Figures~\ref{fig:41:from10to100}, \ref{fig:41:figure4appendix}), and are given in ``quotes'' in the corresponding paragraphs.
 
-\paragraph{$1^{\text{st}}$ - Naive policy: Random Channel Selection}
-
+\textbf{$1^{\text{st}}$ - Naive policy: Random Channel Selection} ``\textcolor{darkgreen}{Random}''.
+%
 We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in [K]$).
@@
 
-\paragraph{$2^{\text{nd}}$ - (Unachievable) Optimal oracle policy}
-
+\textbf{$2^{\text{nd}}$ - (Unachievable) Optimal oracle policy} ``\textcolor{orange}{Optimal}''.
+%
 We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the repartition of static devices $(S_k)_k$, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
@@
     The \emph{Lagrange multipliers} method \cite{BoydVanderberghe04} can be used to solve the constraint real-valued maximization problem introduced in equation \eqref{eq:41:prob}.
-
+    %
     It gives a closed form expression for the (unique) optimal solution $D_k^*(\lambda)$, depending on the system parameters, and the unknown Lagrange multiplier $\lambda \in \mathbb{R}$.
-
+    %
     \begin{equation}\label{eq:41:Dilambda}
@@
 
-Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} = \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (which can be computed numerically in an efficient manner, \cite{Corless96}).
+Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} := \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (which can be computed numerically in an efficient manner, \cite{Corless96}).
 Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint
@@
 
-\paragraph{$3^{\text{rd}}$ - A greedy approach of the oracle strategy}
-
+\textbf{$3^{\text{rd}}$ - A greedy approach of the oracle strategy} ``\textcolor{deeppurple}{Good sub-optimal}''.
+%
 We propose a \emph{sequential} approximation of the optimal policy:
@@
 
-\paragraph{A bandit model for IoT.}
-
+\textbf{Using a bandit algorithm for IoT devices:}
+%
 Our IoT application is challenging in that there are \emph{multiple} players (the dynamic devices) interacting with the \emph{same} arms (the channels), without any centralized communication (they do not even know the total number of dynamic devices).
@@
 
-\paragraph{Two algorithms.}
-
-The two algorithms, \UCB{} and Thompson sampling (TS), were presented in Section~\ref{sec:2:famousMABalgorithms}.
-The \UCB{} algorithm uses a parameter $\alpha > 0$, set to $\alpha = 1/2$ in our experiments.
+\textbf{Considered algorithms.}
+%
+% In our model, every dynamic device implements its own algorithm, \emph{independently}.
+% For one dynamic device, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
+%
+The two algorithms we consider are \UCB{} (``\textcolor{blue}{UCB}'' in the figures)
+and Thompson sampling (TS) (``\textcolor{red}{Thompson-sampling}''),
+and they are presented in Section~\ref{sec:2:famousMABalgorithms}.
 % originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
 %
-In our model, every dynamic device implements its own algorithm, \emph{independently}.
-For one dynamic device, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.
-%
-The TS algorithm
+The \UCB{} algorithm uses $\alpha = 1/2$ in our experiments.
+The TS algorithm is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
 % although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
-is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.
 
-\paragraph{\emph{Adversarial} bandit algorithms?.}
+\paragraph{\emph{Adversarial} bandit algorithms?}
 
@@
 
-\begin{figure}[!b]
+\begin{figure}[!h]
     \centering
@@
 
-\paragraph{Successful transmissions rate as a function of the dynamic devices.}
-
+\textbf{Successful transmissions rate as a function of the dynamic devices.}
+%
 To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure~\ref{fig:41:perf_learning} the evolution of the gain, in terms of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each IoT device.
@@
 The simulation code used for the experiments in Section~\ref{sub:41:numericalResults} is for MATLAB or GNU Octave,
-and it was written in collaboration with Rémi Bonnefoi, in April and May $2017$.
+and it was written in collaboration with Rémi Bonnefoi, in May $2017$.
 Instructions to reproduce our experiments are given, and

### 2019-06-08 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Done for proof reading chapter 4

@@
 	\item
-	For $0 < \gamma < 1$, the function: $g: x \mapsto x \gamma^x$ is quasi-convex on $[0,\infty)$, \ie, $g(\eta x + (1-\eta)y) \leq \max(g(x), g(y))$ for any $x,y \in [0,\infty)$ and $\eta \in [0,1]$ (definition from \cite{Luenberger68}).
+	For $0 < \gamma < 1$, the function $g(x) = x \gamma^x$ is quasi-convex on $[0,\infty)$, \ie, $g(\eta x + (1-\eta)y) \leq \max(g(x), g(y))$ for any $x,y \in [0,\infty)$ and $\eta \in [0,1]$ (definition from \cite{Luenberger68}).
     Indeed, $g(\eta x + (1-\eta)y) = \eta \left[ x (\gamma^x)^{\eta} \right] \gamma^{((1-\eta)y)} + (1-\eta)\left[ y (\gamma^y)^{1-\eta}\right] \gamma^{\eta x}$, and $\gamma^{((1-\eta)y)} \leq 1$ and $\gamma^{\eta x} \leq 1$.
@@
     \item
-    The function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{K}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
+    The function $f(D_1, \dots, D_{K}) = \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{K}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
+    \hfill{}$\square$
 \end{itemize}
-
-\hfill{}$\square$
 
@@
     and so the unique solution to the maximization problem \eqref{eq:41:optPb}, thanks to \cite[Theorem 1]{Luenberger68}.
+    \hfill{}$\square$
 \end{itemize}
-
-\hfill{}$\square$

### 2019-06-08 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Done for proof reading chapter 4

@@
 
-\paragraph{Related works.}
+\textbf{Related works.}
 % \TODOL{Est-ce que je fais ça ici ou à la fin du chapitre 5 ? Ou peut-être en chapitre 1, ça ferait du contenu au chapitre 1...}
 %
-We evoke quickly here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
+We evoke here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
 Starting from $2010$, the works of Wassim Jouini \cite{Jouini09,Jouini10,Jouini12} were the first ones to propose to use Reinforcement Learning for Cognitive Radio, especially MAB and the \UCB{} algorithm, and from $2014$, proof-of-concepts were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
@@
 
-\paragraph{Physical layer and protocol.}
-
+\textbf{Physical layer and protocol.}
+%
 We implement a simple PHY/MAC layers solution, in order to demonstrate the possibility of improvement of the performance of IoT communications in unlicensed bands.
@@
 % Octoclock
-To ease the synchronization implementation in both time and frequency between the boards representing the dynamic devices and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
+To ease the synchronization in both time and frequency between the boards representing the dynamic devices and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
 % https://www.ettus.com/product/details/OctoClock-G
@@
 The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at
-\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/\\malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
+\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm\\-bandit-learning-for-iot-networks-with-grc/}}.
 %
 It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.
-The demonstration only requires a laptop and free tools,
+The demonstration only requires a laptop and open-source free softwares,
 as the laptop should run a GNU/Linux distribution (like Ubuntu or Debian),
 in addition to USRP platforms from Ettus Research.
+
+\paragraph{Video.}
+%
+As depicted in Figure~\ref{fig:42:screenshotDemoYouTube} below,
+we realized a $6$-minute \textbf{video} to sum-up our demonstration and advertise our work online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
+It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
+% \textcolor{white}{Of course, the video was not seen much, because our work is useless! Yay!}
 
@@
 
-\paragraph{Video.}
-%
-As depicted in Figure~\ref{fig:42:screenshotDemoYouTube} above,
-we realized a $6$-minute \textbf{video} to sum-up our demonstration and advertise our work online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
-The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
-It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
-% \textcolor{white}{Of course, the video was not seen much, because our work is useless! Yay!}
-
 % \paragraph{Special acknowledgment.}

### 2019-06-08 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for proof reading chapter 4

@@
 In this section, we extend the model presented before, to take into account the possibility for retransmissions of a message after a collision.
-This extension is of major importance: most protocols for real-world IoT networks can use retransmissions, it is for instance the case of the LoRa standard.
+This extension is of major importance: most protocols for real-world IoT networks can use retransmissions, it is for instance the case of the LoRaWAN standard.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.
@@
 
-The contributions of this section are summarized as follows:
-\begin{itemize}
-	\item
+The contributions of this section are summarized as follows.
+% \begin{itemize}
+	% \item
 	Firstly, we provide a close form approximation of the radio collision probability after a first retransmission.
 	By doing this, we highlight the need to develop a learning approach for channel selection upon collision.
-
-	\item
+%
+	% \item
 	Secondly, different heuristics are proposed to cope with retransmissions.
-
-	\item Lastly, we conduct simulations in order to compare the performance of the proposed heuristics with a naive uniform random approach, and a \UCB{} strategy (\ie, without any learning for the retransmissions, that is, the same channel is used for retransmission).
-\end{itemize}
+%
+	% \item
+	Lastly, we conduct simulations in order to compare the performance of the proposed heuristics with a naive uniform random approach, and a \UCB{} strategy (\ie, without any learning for the retransmissions, that is, the same channel is used for retransmission).
+% \end{itemize}
 
@@
 these devices will attempt to retransmit their packet up-to $M$ times, with $M \in\mathbb{N}$.
-It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\llbracket 0, m-1 \rrbracket$, where $m \in\mathbb{N}, m>0$ is the length of the back-off interval
+It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\{0, \dots, m-1\}$, where $m \in\mathbb{N}, m>0$ is the length of the back-off interval
 (note the difference between $m$ and $M$).
@@
 Considering one device and one channel,
-we denote $x_t^i$ the probability that it is transmitting a packet for the $(i+1)$-th time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$),
+we denote $x_t^i$ the probability that it is transmitting a packet for the $(i+1)$-th time in a given time slot $t$ (with $i\in \{0, \dots, M-1 \}$),
 and we denote $x_t = \sum_{i=0}^{M-1}x_t^i$ the probability that it transmits a packet (\ie, just the sum on $i$ of $x_t^i$).
@@
 %
-\begin{equation}\label{eq:43:sumpca}
-	p_{ca} = \frac{1}{p_c}\sum_{n=1}^{N-1} p_{bp}(n).
-\end{equation}
-
+% \begin{equation}\label{eq:43:sumpca}
+	$p_{ca} = \frac{1}{p_c}\sum_{n=1}^{N-1} p_{bp}(n)$.
+% \end{equation}
+%
 % \begin{figure}[htp!]  % [htbp]
@@
 % \end{figure}
-
+%
 Therefore, the expression of $p_{ca}$ is
@@
 
-We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca4}, and we rewrite the expression of $p_{ca}$ as
+The binomial theorem expands the sum in \eqref{eq:43:sumpca4}, so we can rewrite the expression of $p_{ca}$ as
 %
@@
 	% \includegraphics[width=0.65\linewidth]{Approximation_m10.eps}
-	\includegraphics[width=0.80\linewidth]{Approximation_m10-eps-converted-to.pdf}
+	\includegraphics[width=0.70\linewidth]{Approximation_m10-eps-converted-to.pdf}
 	\caption[Proposed approximation for the probability of collision at the second transmission]{Proposed approximation for the probability of collision at the second transmission.
@@
 % \begin{equation}\label{eq:43:maxucb}
-% 	A(t) = \arg\max_{1\leq k \leq K} U_k(t).
+% 	A(t) \sim \cU(\argmax_k U_k(t)).
 % \end{equation}
@@
 % 			Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$\;
-% 			Transmit in channel $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
+% 			Transmit in channel $A(t) \sim \cU(\argmax_k U_k(t))$\;
 % 			Reward $r_{A(t)}(t) = 1$, if \emph{Ack} is received, else $0$\;
@@
 
-More formally, for one device, let $N_k(t)$ be the number of times the channel $k$ (for $k\in \llbracket 1, K \rrbracket$) was selected up-to time $t-1$, for $t\geq 0$
-for any $t\in\mathbb{N}$,
-$N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(A(\tau) = k)$.
-The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
-$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r(\tau) \mathbbm{1}(A(\tau) = k)$,
-where $r(t)=Y_{k,t}$ is the reward obtained after transmission in channel $k$ at time $t$.
-The upper confidence bound in each channel $k$ is defined as
-$U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$.
-Finally, the transmission channel at time step $t$
-is the one maximizing this \UCB{} index $U_k(t)$,
-as it is the one expected to be the best one at the current time step $t$,
-$A(t) = \arg\max_{1\leq k \leq K} U_k(t)$.
-
 % \begin{small} % XXX remove if needed
@@
 				Compute $\forall k, U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$\;
-				Transmit in channel $A(t) = \arg\max_{1\leq k \leq K} U_k(t)$\;
+				Transmit in channel $A(t) \sim \cU(\argmax_k U_k(t))$\;
 				Reward $r(t) = 1$, if \emph{Ack} is received, else $0$\;
@@
 % \end{small} % XXX remove if needed
+
+More formally, for one device, let $N_k(t)$ be the number of times the channel $k$ (for $k\in [K]$) was selected up-to time $t-1$, for $t\geq 0$
+for any $t\in\mathbb{N}$,
+$N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(A(\tau) = k)$.
+The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
+$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r(\tau) \mathbbm{1}(A(\tau) = k)$,
+where $r(t)=Y_{k,t}$ is the reward obtained after transmission in channel $k$ at time $t$.
+The upper confidence bound in each channel $k$ is defined as
+$U_k(t) = \widehat{\mu_k}(t) + \sqrt{\alpha \log(t) / N_k(t)}$.
+Finally, the transmission channel at time step $t$
+is the one maximizing this \UCB{} index $U_k(t)$,
+as it is the one expected to be the best one at the current time step $t$,
+$A(t) = \arg\max_k U_k(t)$.
 
@@
 We present below four heuristics for this second stage.
-Their short names are used in the legend on Figures~\ref{fig:43:mainExperiment1}, \ref{fig:43:mainExperiment2}), and are given in ``quotes'' in the corresponding paragraphs.
+Their short names (with their colors) are used in the legend on Figures~\ref{fig:43:mainExperiment1}, \ref{fig:43:mainExperiment2}), and are given in ``quotes'' in the corresponding paragraphs.
 
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Uniform random retransmission (``Random'').}\label{sub:43:UCBthenRandom}
+\paragraph{Uniform random retransmission ``\textcolor{cyan}{(Random)''}.}\label{sub:43:UCBthenRandom}
 
-In this first proposal, the device uses a random channel selection, following a uniform distribution (in $\llbracket 1, K \rrbracket$).
+In this first proposal, the device uses a random channel selection, following a uniform distribution (on $[K]$).
 It is described below in Algorithm~\ref{algo:43:UCBthenRandom}.
@@
 			\uIf{First packet transmission}{
-				Use first-stage \UCB. %as in Algorithm~\ref{algo:43:UCB}.
+				Use first-stage \UCB{} as in Algorithm~\ref{algo:43:UCB}.
 			}
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{\UCB{} for retransmission (``Only \UCB{}'').}\label{sub:43:TwoUCB}
+\paragraph{\UCB{} for retransmission (``\textcolor{blue}{Only \UCB{}}'').}\label{sub:43:TwoUCB}
 
@@
 				Compute $\forall k, U^r_k(t) = \widehat{\mu^r_k}(t) + \sqrt{\alpha \log(t) / N_k^r(t)}$\;
-				Transmit in channel $C^r(t) = \arg\max_{1\leq k \leq K} U^r_k(t)$\;
+				Transmit in channel $C^r(t) \sim \cU(\argmax_k U^r_k(t))$\;
 				Reward $r^r(t) = 1$, if \emph{Ack} is received, else $0$\;
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{$K$ different {\UCB}s for retransmission (``$K$ \UCB{}'')}\label{sub:43:UCBthenKp1}
+\paragraph{$K$ different {\UCB}s for retransmission (``\textcolor{green}{$K$ \UCB}'')}\label{sub:43:UCBthenKp1}
 
@@
 The corresponding algorithm is depicted in Algorithm~\ref{algo:43:UCBthenKp1}.
-Each of these algorithms are denoted using the subscript $({}^{j})$, for $j\in\llbracket 1, K \rrbracket$.
+Each of these algorithms are denoted using the subscript $({}^{j})$, for $j\in[K]$.
 
@@
 				Compute $\forall k, U_k^j(t) = \widehat{\mu_k}^j(t) + \sqrt{\alpha \log(t) / N_k^j(t)}$\;
-				Transmit in channel $A^j(t) = \arg\max_{1\leq k \leq K} U^j_k(t)$\;
+				Transmit in channel $A^j(t) \sim \cU(\argmax_k U^j_k(t))$\;
 				Reward $r^j_{A^j(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'').}\label{sub:43:UCBwithDelay}
+\paragraph{Delayed \UCB{} for retransmission (``\textcolor{red}{Delayed \UCB}'').}\label{sub:43:UCBwithDelay}
 
@@
 				Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + \sqrt{\alpha \log(t) / N_k^d(t)}$\;
-				Transmit in channel $A^d(t) = \arg\max_{1\leq k \leq K} U^d_k(t)$\;
+				Transmit in channel $A^d(t) \sim \cU(\argmax_k U^d_k(t))$\;
 				Reward $r^d_{A^d(t)}(t) = 1$ if \emph{Ack} is received, else $0$\;
@@
 Instead of considering a first-stage \UCB{} followed by a random channel retransmission, we could have considered a random first-stage channel retransmission followed by a second-stage \UCB.
-This was not considered in our study \cite{Bonnefoi2019WCNC}, and we preferred to not add it to the experiments presented below, for three reasons: to avoid clutter in the plots, to win some time as these experiments had to run for quite a long time, but mainly because the first-stage channel selection is most important one as we illustrate by the large difference in terms of performance between the uniform channel selection and the ``only UCB'' heuristic, below in Figures~\ref{fig:43:mainExperiment1} and \ref{fig:43:mainExperiment2}.
+This was not considered in our study \cite{Bonnefoi2019WCNC}, and we preferred to not add it to the experiments presented below, for three reasons: to avoid clutter in the plots, to win some time as these experiments had to run for quite a long time, but mainly because the first-stage channel selection is most important one as we illustrate by the large difference in terms of performance between the uniform channel selection and the ``Only UCB'' heuristic, below in Figures~\ref{fig:43:mainExperiment1} and \ref{fig:43:mainExperiment2}.
 
-\paragraph{Using another bandit policy?}
+\textbf{Using another bandit policy?}
 %
@@
 We could have chose any bandit algorithm for the ``first stage'' component used in this Section, and for simplicity and clarity we focussed on a simple but efficient one (\UCB).
-We believe that the same empirical results, but more importantly, the same conclusions, could be given if we used Bayes-UCB or other bandit algorithm for the base building block in the different heuristics proposed in Section~\ref{sub:43:heuristics}.
-The goal of this Section was not to optimize on the bandit policy (as we presented in Section~\ref{sec:3:reviewSPAlgorithms} some numerical simulations doing precisely this), but rather to compare the different heuristics, for a fixed bandit policy (for which we preferred to use \UCB{} for simplicity).
+We believe that the same empirical results, but more importantly, the same conclusions, could be given if we used Bayes-UCB or other bandit algorithm for the base building block in the different heuristics proposed in this section.
+% ~\ref{sub:43:heuristics}.
+Our goal here was not to optimize on the bandit policy (as we presented in Section~\ref{sec:3:reviewSPAlgorithms} some numerical simulations doing precisely this), but rather to compare the different heuristics, for a fixed bandit policy (for which we preferred to use \UCB{} for simplicity).
 
@@
 
-We simulate our network considering $N$ devices following the contention Markov process described in Section \ref{sub:43:model}, and a LoRa standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
+We simulate our network considering $N$ devices following the contention Markov process described in Section \ref{sub:43:model}, and a LoRaWAN standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
 Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of $1\;\mathrm{s}$.
@@
 
-A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve), which is (still) the current state-of-the-art in IoT networks.
+A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve in black), which is (still) the current state-of-the-art in IoT networks.
 The improvement of the heuristics over the non-learning approach is clear, and for every heuristic that uses a kind of learning mechanism it can observed the successful transmission rate increases rapidly (or equivalently the PLR decreases).
@@
 	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB'']{
-		Comparison between the exposed heuristics for the retransmission: \textcolor{blue}{``Only \UCB''}, \textcolor{cyan}{``Random''}, \textcolor{purple}{``\UCB''}, \textcolor{green}{``$K$ \UCB''}, and \textcolor{red}{``Delayed \UCB''}.
-		The usage of the same learning policy for transmissions and retransmission is named \textcolor{blue}{``Only \UCB{}''},
-		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as \textbf{``no \UCB{}''}.
+		Comparison between the exposed heuristics for the retransmission:`` \textcolor{blue}{Only \UCB}'', ``\textcolor{cyan}{Random}'', ``\textcolor{purple}{\UCB}'', ``\textcolor{green}{$K$ \UCB}'', and ``\textcolor{red}{Delayed \UCB}''.
+		The usage of the same learning policy for transmissions and retransmission is named ``\textcolor{blue}{Only \UCB{}}'',
+		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as ``\textbf{no \UCB{}}''.
 		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
@@
 
-\paragraph{Note on the simulation code.}
+% \paragraph{Note on the simulation code.}
+\paragraph{Reproducility of the experiments.}
 %

### 2019-06-08 2-Chapters/4-Chapter/chapter4.tex
Done for proof reading chapter 4

--- 2019-06-06 2-Chapters/4-Chapter/chapter4.tex
@@
 FIXME finish !
+
+% \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
 
@@
 
-% We give in this section conclusions about the two models presented above, one for long-life IoT devices (without retransmission) and for improved QoS devices (\ie, with retransmissions), and about our demonstration, as well as directions of future works, some of which are currently being studied by researchers of our team.
+% % We give in this section conclusions about the two models presented above, one for long-life IoT devices (without retransmission) and for improved QoS devices (\ie, with retransmissions), and about our demonstration, as well as directions of future works, some of which are currently being studied by researchers of our team.
 
-% \TODOL{Je devrais réécrire toute cette conclusion, si j'ai le temps, pour faire mieux qu'une suite de 3 conclusions et 3 future works, et être plus concis et moins ``short sighted'' comme disait Emilie.}
+% % \subsection*{Conclusions}
 
-% \subsection*{Conclusions}
+% % ----------------------------------------------------------------------
+% % \paragraph{Summary of the first model.}
+% \label{sub:41:conclusion}
 
-% ----------------------------------------------------------------------
-% \paragraph{Summary of the first model.}
-\label{sub:41:conclusion}
+% In Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
+% with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
+% Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
+% Similarly, if the number of devices remain constant, increasing the successful transmission rate directly extends the IoT devices battery life, as they suffer less from failed transmissions.
+% We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
+% %
+% This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
+% The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.
 
-In Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
-with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
-Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
-Similarly, if the number of devices remain constant, increasing the successful transmission rate directly extends the IoT devices battery life, as they suffer less from failed transmissions.
-We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
-%
-This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
-The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each others, and \emph{do not} know the system parameters.
+% % ----------------------------------------------------------------------
+% % \paragraph{Conclusions taken from the demonstration.}
+% \label{sub:42:conclusionFromDemonstration}
 
-% ----------------------------------------------------------------------
-% \paragraph{Conclusions taken from the demonstration.}
-\label{sub:42:conclusionFromDemonstration}
+% We presented in Section~\ref{sec:4:gnuradio} a demonstration, showed in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
+% We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
+% Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
+% %
+% By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.
 
-We presented in Section~\ref{sec:4:gnuradio} a demonstration, showed in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
-We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
-Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
-%
-By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.
+% % ----------------------------------------------------------------------
+% % \paragraph{Conclusion about this second model.}
+% \label{sub:43:conclusion}
 
-% ----------------------------------------------------------------------
-% \paragraph{Conclusion about this second model.}
-\label{sub:43:conclusion}
+% In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency.
+% % , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
+% If the priority is the Quality of Service, for instance with renewable energy capabilities, this second model is more appropriate.
+% The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
+% We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
+% %
+% We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
+% Our simulations show that each of our proposed heuristic outperforms a naive random access scheme.
+% Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.
 
-In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency.
-% , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
-If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
-The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
-We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.
-%
-We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
-Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
-Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.
+% % ----------------------------------------------------------------------
+% \subsection*{Future works}
+% \label{sub:4:futureWorks}
 
-% ----------------------------------------------------------------------
-\subsection*{Future works}
-\label{sub:4:futureWorks}
+% % Future works related to this chapter include the following directions.
 
-% Future works related to this chapter include the following directions.
+% % \paragraph{Possible extensions of the first model.}
 
-% \paragraph{Possible extensions of the first model.}
+% The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device.
+Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
+% In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
+% %
+% Moreover, for sake of simplicity we supposed that all devices use the same standard.
+% Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
 
-The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device.
-Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
-In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
-%
-Moreover, for sake of simplicity we supposed that all devices use the same standard.
-Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
+% Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
+% So additionally to deciding the channel for communication (\ie, \emph{where} to communicate), each device has to also decide \emph{when} to communicate.
+% % This is another direction of research, that we will investigate in the future.
+% However, this clearly leads to a much larger action space, as there are many time slots in one day (for example), and thus we believe that as soon as the action space becomes too large in this extension, the simple MAB-based learning approach could be no longer appropriate.
+% It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
+% It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
 
-Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
-So additionally to deciding the channel for communication (\ie, \emph{where} to communicate), each device has to also decide \emph{when} to communicate.
-% This is another direction of research, that we will investigate in the future.
-However, this clearly leads to a much larger action space, as there are many time slots in one day (for example), and thus we believe that as soon as the action space becomes too large in this extension, the simple MAB-based learning approach could be no longer appropriate.
-It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
-It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
+% % We will investigate this behavior in order to understand it better theoretically.
+% % We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
 
-% We will investigate this behavior in order to understand it better theoretically.
-% We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
+% % \paragraph{Extensions of the demonstration.}
 
-% \paragraph{Extensions of the demonstration.}
+% Possible future extensions of our demonstration include the following points.
+% We could consider more dynamic devices (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more devices with the same card.
+% We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
+% Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card to simulate another gateway).
+% %
+% % Pub pour notre autre article
+% % We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.
 
-Possible future extensions of our demonstration include the following points.
-We could consider more dynamic devices (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more devices with the same card.
-We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
-Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card to simulate another gateway).
-%
-% Pub pour notre autre article
-% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.
+% % \paragraph{Possible extensions for the second model.}
 
-% \paragraph{Possible extensions for the second model.}
+% Finally, the utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
+Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
+% %
+% For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
+% or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
+% or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
+% that nevertheless have not been explored for the targeted problem.
+% Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this Chapter.
 
-Finally, the utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
-Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
-%
-For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
-or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
-or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
-that nevertheless have not been explored for the targeted problem.
-Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this Chapter.
-
-In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our demonstration \cite{Besson2019WCNC}.
-Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes.
-% and Christophe Moy at IETR and University Rennes 1.
-%
-A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.
+% In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our demonstration \cite{Besson2019WCNC}.
+% Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes.
+% % and Christophe Moy at IETR and University Rennes 1.
+% %
+% A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.
 

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
A la BU le samedi après-midi ouais
Almost done for ch5

@@
 
-We consider $M \leq K$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
+We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
 A perfect time and frequency synchronization is assumed,
@@
 
-% Another interesting extension of the studied model is to consider different channels utility for each devices.
-It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
+% Another interesting extension of the studied model is to consider different channels utility for each device.
+It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channel.
 % This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.
@@
 Hence there is small abuse of notation when calling it a number of collisions.
-
-Letting $\cP_M = \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
+%
+Letting $\cP_M := \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
 be the set of bandit instances such that there is a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.
@@
 we use the Kullback-Leibler divergence to express our lower bound.
-Let $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ be KL divergence between the Bernoulli distribution of mean $x$ and that of mean $y$.
+We remind that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
 %
@@
 Having a small regret on every problem instance, \ie, being uniformly efficient,
-is a natural assumption for algorithms,
-that rules out algorithms tuned to perform well on specific instances only.
+is a natural assumption,
+that rules out algorithms tuned to perform well on specific instances only (\eg, fixed-armed algorithms).
 %
@@
 The additional assumption in \eqref{eq:5:SUE} further implies some notion of \emph{fairness}, as it suggests that each of the $M$ players spends on average the same amount of time on each of the $M$ best arms.
-Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\cA_{\sigma} = (\cA_{\sigma(1)},\dots,\cA_{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$.
+Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\cA_{\sigma} = (\cA_{\sigma(1)},\dots,\cA_{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$ of the set $\{1,\dots,M\}$.
 In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
-Note that all our proposed algorithms are permutation invariant and \MCTopM{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.
-
-As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
+Note that the algorithms studied in Section~\ref{sec:5:algorithms} are permutation invariant, and \MCTopM-\klUCB{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.
+
+% As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
 In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
-We note that the notion of fairness we defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
+The notion of fairness defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
 For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is given and known to the algorithm.
@@
 The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
-Fix a player $j\in[M]$, and we consider the observations $O_t$ gathered by this player after $t$ rounds of its algorithm $\cA_j$:
-\[ O_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
+Similarly to what we did in Section~\ref{sec:2:notations},
+consider the observations $O_t^j$ that she gathered after $t$ rounds of its algorithm $\cA_j$ for a fixed player $j\in[M]$,
+defined as $O_t^j := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
 where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).
@@
 Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
-We denote $\Pr_{\bm{\mu}}^{O_t}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\cA$.
+We denote $\Pr_{\bm{\mu}}^{O_t^j}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\cA_j$.
 Then we introduce the \emph{collision information term} as:
 \begin{equation}\label{eq:5:collisionInformationTerm}
-  \cI_{\bm{\mu},\bm{\lambda}}(\cA,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}}).
-\end{equation}
-
-For a \emph{decentralized strategy} $\cA$ that has access to the sensing information (\ie, ruling out model \modeltrois), and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$,
+  \cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}^j}).
+\end{equation}
+
+For a \emph{decentralized strategy} $\cA$ that has access to the sensing information (\ie, ruling out model \modeltrois), and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$,
 we now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections.
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\cA$ and $\boldsymbol{\mu}\in\cP_M$.
-  Furthermore, if $\cA$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$, then
+  Under observation models $\modelun$ and $\modeldeux$, consider a strongly uniformly efficient \emph{decentralized} policy $\cA=(\cA_1,\dots,\cA_M)$ and a problem $\boldsymbol{\mu}\in\cP_M$.
+  Furthermore, if $\cA_j$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$, then
   % \vspace*{-10pt}  % XXX
@@
   % \vspace{-0.3cm}  % XXX
-  \noindent From Lemma~\ref{lem:5:1stLowerBound}, it follows that
+  \noindent And so if $\cA_1,\dots,\cA_M$ all satisfy this hypothesis, from Lemma~\ref{lem:5:1stLowerBound}, it follows that
   % \vspace*{-5pt}  % XXX
@@
 %
-It is worth mentioning that \cite{Zhao10} proved a lower bound under the more general assumption for $\cA$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
-Our result could be extended to this case but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.
+It is worth mentioning that \cite{Zhao10} considered the more general assumption for $\cA$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
+Our result could be extended to this case, but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.
 
@@
 \begin{equation}
-  \widetilde{S_k}^j(t) = \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
+  \widetilde{S_k}^j(t) := \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
   \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) := \widetilde{S_k}^j(t) \;/\; T_k^j(t).
@@
 \begin{equation}\label{eq:5:indexTilde}
-  \widetilde{U_k}^j(t) = \begin{cases}
+  \widetilde{U_k}^j(t) := \begin{cases}
       \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
@@
 
-Given any of these two index policies (\UCB{} or \klUCB), the \Selfish{} algorithm is defined by,
-%
-\begin{equation}
-  A^j(t) = \argmax_{ k \in [K]} \ \widetilde{U_k}^j(t-1).\label{algo:5:Selfish}
-\end{equation}
-The name comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
+Given any of these two index policies (\UCB{} or \klUCB), the \Selfish{} algorithm is then playing like a single-player index policy (see Algorithm~\ref{algo:2:indexPolicy})
+% \begin{equation}\label{algo:5:Selfish}
+$A^j(t) \in \cU(\argmax_{ k \in [K]} \ \widetilde{U_k}^j(t-1))$.
+% \end{equation}
+The name ``selfish'' comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
 The reason that this may work precisely comes from the fact that $\widetilde{U_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
 but some hybrid index that simultaneously increases when a transmission occurs and decreases when a collision occurs.
-
+%
 This behavior is easier to be understood for the case of \Selfish-\UCB{} in which, letting $N_k^{j,C}(t) = \sum_{s=1}^t \indic(C^j(t))$ be the number of collisions on arm $k$, one can show that the hybrid \Selfish{} index induces a penalty proportional to the fraction of collision on this arm and the quality of the arm itself:
@@
 Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or cannot be considered.
+In fact, we used the \Selfish{} heuristic in Chapter~\ref{chapter:4}, in the different models, using \UCB{} or Thompson sampling as the underlying policy.
+%
 Thus we propose next our main contribution, the \MCTopM{} algorithm, proved to be asymptotically optimal for the identification of suboptimal arms (when using \klUCB), attaining order-optimal logarithmic regret, and outperforming all the other algorithms for the ``sensing case''.
@@
 
-In a single-player setting, \emph{index policies} are popular bandit algorithms: at each round one index is computed for each arm, that only depends on the history of plays of this arm and (possibly) some exogenous randomness.
+% In a single-player setting, \emph{index policies} are popular bandit algorithms: at each round one index is computed for each arm, that only depends on the history of plays of this arm and (possibly) some exogenous randomness.
 Then, the arm with highest index is selected.
@@
 
-% concrete examples and how their are used within MPB
-
-\paragraph{Index policies.}
-%
-The approaches we now describe for multi-player bandits can be used in combination with any index policy, but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
+% % concrete examples and how their are used within MPB
+
+% \textbf{Index policies.}
+% %
+The approache we now describe for multi-player bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
 In particular, we focus on two types of indices:
-\UCB{} indices \citep{Auer02}
-and \klUCB{} indices \citep{KLUCBJournal}, that can be defined for each player $j$ in the following way.
-%
-Letting $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained by player $j$ for arm $k$,
-then $\widehat{\mu}_k^j(t) = S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$ for player $j$, and thus one can define the index
+\UCB{} \citep{Auer02}
+and \klUCB{} \citep{KLUCBJournal}.
+For player $j\in[M]$
+%
+denote $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
+then $\widehat{\mu}_k^j(t) := S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
 \begin{equation}\label{eq:5:indexFor_UCB_klUCB}
@@
 \end{equation}
-where $f(t)$ is some \emph{exploration function}.
-$f(t)$ is usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
-A classical (single-player) UCB algorithm aims at the arm with largest index.
+where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
+A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
 However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
@@
 and selects at time $t$,
-\begin{equation}
-  A^j(t) := \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right).
-\end{equation}
+% \begin{equation}
+$A^j(t) := \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right)$.
+% \end{equation}
 If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU([M])$.
@@
 
-\paragraph{The \RandTopM{} algorithm.}
+\textbf{The \RandTopM{} algorithm.}
 %
@@
 When a collision occurs for a player implementing the \RandTopM{} algorithm, that player randomly switches arm within $\TopM$, while \MCTopM{} uses a more sophisticated mechanism, that is reminiscent of ``Musical Chair'' (MC) and inspired by the work of \cite{Rosenski16}: players tend to fix themselves on arms (``chairs'') and ignore future collision when this happens.
+
+\TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}
 
@@
 \begin{figure}[h!]
-  \resizebox{1.00\textwidth}{!}{
-  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.8]
+  \resizebox{1.10\textwidth}{!}{
+  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.0]
       %
@@
       \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
-      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
+      \path [color=blue,thick,->] (notfixed) edge[loop right] node[below,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t)$,\\$A^j(t) \in \TopM(t)$} (1);
       \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
-      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
+      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[below,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
       \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for ch5

@@
   let player $j\in[M]$ use the \RandTopM-, \MCTopM- or \rhoRand-\klUCB{}
-  decentralized policy with exploration function $f(t) = \log(t) + 3 \log\log(t)$.
-  Then for any sub-optimal arm $k \in \Mworst$, there exists two problem-dependent positive constants $C_{\boldsymbol{\mu}}, D_{\boldsymbol{\mu}} > 0$ such that
+  decentralized policy with exploration function $f(t) = \log(t) + 3 \log(\log(t))$.
+  Then for any sub-optimal arm $k \in \Mworst$ there exists problem-dependent constants $C_{\boldsymbol{\mu}}, D_{\boldsymbol{\mu}} > 0$ such that
   % %
@@
       \frac{\log(T)}{\kl(\mu_k,\mu_{M}^*)}
-      + \underbrace{C_{\boldsymbol{\mu}} \sqrt{\log(T)} + D_{\boldsymbol{\mu}}\log\log(T) + 3M + 1}_{= \smallO{\log(T)}}.
+      + \underbrace{C_{\boldsymbol{\mu}} \sqrt{\log(T)} + D_{\boldsymbol{\mu}}\log(\log(T)) + 3M + 1}_{= \smallO{\log(T)}}.
   \end{equation}
@@
   Considering the relative position of the upper-confidence bound $U_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
-
+  \begin{align*}
+    \E_{\mu}[T_k^j(T)] &\leq
+    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; U_{m_1^*}(t) < \mu_m^* \right) + \\
+    \;\;\;\;
+    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq U_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right)
+  \end{align*}
   \begin{align}
-    \E_{\mu}[T_k^j(T)] &\leq
-    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; U_{m_1^*}(t) < \mu_m^* \right) + \notag \\
-    \;\;\;\;
-    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq U_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
-    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq U_k(t)\right) \notag \\
+    \E_{\mu}[T_k^j(T)] &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq U_k(t)\right) \notag \\
     %
@@
   \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-      For any arm $k$, if $U_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
+      For any arm $k$, if $U_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log(\log(t))$,
       \begin{equation}
-        \sum_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
+        \sum_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log(\log(T)).
       \end{equation}
@@
     \E_{\mu}\left[T_k^j(T)\right]
-    & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
-    & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
+    & \leq \frac{\log(T) + 3 \log(\log(T))}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log(\log(T))}  \nonumber\\
+    & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log(\log(T)) + 3M+1,
   \end{align}
@@
   We analyze the case when the current arm leaves the set $\TopM$ (Line 4):
-
+  %
   \begin{align*}
@@
   \[T_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right)\]
-  The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
-To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
+  The first sum is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
+  To control the second sum, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
   Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
   and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
-
+  %
+  \begin{align*}
+    \sum_{t=1}^T & \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right) \\
+    & =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]
+  \end{align*}
   \begin{align}
-    \sum_{t=1}^T & \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
-    & =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
     & \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
@@
   If $\mu_k > \mu_{k'}$, we rather use that
-  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \Pr\left(A^j(t+1) = k', U_{k'}^j(t) \geq \mu_{k}\right)\]
+  $T_{k'} \leq \sum\limits_{t=1}^T \Pr\left(U_{k}^j(t) \leq \mu_{k}\right) + \sum\limits_{t=1}^T \Pr\left(A^j(t+1) = k', U_{k'}^j(t) \geq \mu_{k}\right)$,
   and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
-  \[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
+  $T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T))$.
   Summing over $k'$ yields the result.
@@
   Then, we introduce $\widetilde{C^j}(t)$ as the event that a collision occurs for player $j$ at round $t$ if she is not yet fixed on her arm, that is
-  \begin{equation}
-      \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
-  \end{equation}
+  % \begin{equation}
+  $\widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right)$.
+  % \end{equation}
 
@@
   thus we can use Wald's Lemma \citep{Wald45},
-  to obtain a decomposition with two terms $(\alpha)$ and $(\beta)$:
-  \[
-    \E_{\mu}[N_2^j(t)] \leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}.
-  \]
+  to obtain a decomposition with two terms $(\alpha)$ and $(\beta)$,
+  $\E_{\mu}[N_2^j(t)] \leq \E_{\mu}\left[ n(T) \right] \times \max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]$.
+  % $\E_{\mu}[N_2^j(t)] \leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}$.
 
@@
 %
-Additionally to bounding the collisions in Lemma~\ref{lem:5:collisionsMCTopM},
-another consequence of our proof of this result is that
-it also bounds the (expected) number of \emph{switches of arms}, $SC^{\cA}(T) = \sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$.
+While Lemma~\ref{lem:5:collisionsMCTopM} bounds the number of collisions,
+another consequence of our proof is that we have also bounded
+the (expected) number of \emph{switches of arms}, $SC^{\cA}(T) = \sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$.
 %
@@
 
-\paragraph{Other experiments.}
+\textbf{Other experiments.}
 %
 We now present more results for two bandit instances.
-\begin{itemize}
-  \item
+% \begin{itemize}
+%   \item
   The first instance is a small problem one with $K=3$ arms and means
@@
   It is used to illustrate that in some unlucky runs, \Selfish{} can suffer a linear regret, as illustrated in Figure~\ref{fig:5:selfish_fail1}.
-  \item
+  % \item
   The second instance considers $K=9$ arms with means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
@@
   and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
-\end{itemize}
+% \end{itemize}
 %
@@
 
-\paragraph{Uniformly sampled problems.}
+\textbf{Uniformly sampled problems.}
 %

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Done for ch5

--- 2019-06-07 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
@@
     % \end{subfigure}
-    \caption[Multi-player regret with its three terms]{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
+    \caption[Illustrating the decomposition of the multi-player regret with its three terms]{The three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} of the multi-player regret, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
     \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}

### 2019-06-08 2-Chapters/5-Chapter/chapter5.tex
Almost done for ch5
Done for ch5

@@
 \begin{equation}\label{eq:5:defRegretDynamicSetting}
-    R_T^{\text{dyn}} = \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
+    R_T^{\text{dyn}} := \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
 \end{equation}
@@
 We illustrate this in Table~\ref{table:5:comparisonRhoRandRhoLearn} below, where we consider the same problem as described above in Section~\ref{sub:5:withoutSensing} (Table~\ref{table:5:meanRegretSimulationsNoSensing}).
-The first \rhoLearn uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
+We include different \rhoLearn{} algorithms, that uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
 and we include both the centralized multiple-play version of \klUCB{} and \MCTopM-\klUCB{} for comparison.

### 2019-06-08 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for proofreading Ch6

@@
 
-% Multi-Armed Bandit (MAB) problems form a well-studied class of sequential decision making problems, in which an agent repeatedly chooses an action $I_t \in[K]$ or ``arm'' --in reference to the arm of a one-armed bandit-- among a set of $K$ arms.
-In the most standard version of the stochastic bandit model, each arm $i$ is associated with an \iid{} sequence of rewards $(r_{i,t})$ that follow some distribution of mean $\mu_i$.
-Upon selecting arm $I_t$, the agent receives the reward $r_{I_t,t}$ associated to the chosen arm, and her goal is to adopt a sequential sampling strategy that maximize the expected sum of these rewards.
+% Multi-Armed Bandit (MAB) problems form a well-studied class of sequential decision making problems, in which an agent repeatedly chooses an action $A(t) \in[K]$ or ``arm'' --in reference to the arm of a one-armed bandit-- among a set of $K$ arms.
+In the most standard version of the stochastic bandit model, each arm $k$ is associated with an \iid{} sequence of rewards $(r_{i,t})$ that follow some distribution of mean $\mu_k$.
+Upon selecting arm $A(t)$, the agent receives the reward $r_{A(t),t}$ associated to the chosen arm, and her goal is to adopt a sequential sampling strategy that maximize the expected sum of these rewards.
 This is equivalent to minimizing the \emph{regret}, defined as the difference between the total reward of the oracle strategy always selecting the arm with largest mean, $\mu^*$, and that of our strategy:
 
-% \[ R_T = \bE[\sum_{t=1}^T(\mu^* - \mu_{I_t})]. \]
+% \[ R_T = \bE[\sum_{t=1}^T(\mu^* - \mu_{A(t)})]. \]
 
@@
 %
-In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $i$ at round $t$ has some mean $\mu_i(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $i_t^* = \arg\max_{i} \mu_i(t)$.
+In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $k$ at round $t$ has some mean $\mu_k(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $k_t^* = \arg\max_i \mu_k(t)$.
 
@@
 A \emph{piece-wise stationary bandit model} is characterized by a set of $K$ arms.
-A (random) stream of rewards $(X_{i,t})_{t\in\N^*}$ is associated to each arm $i \in [K]$.
-We assume that the rewards are bounded, and without loss of generality we assume that $X_{i,t} \in [0,1]$.
-We denote by $\mu_{i}(t) :=  \bE[X_{i,t}]$ the mean reward of arm $i$ at round $t$.
-At each round $t$, a decision maker has to select an arm $I_t\in[K]$, based on past observation and receives the corresponding reward $r(t) = X_{I_t,t}$.
-At time $t$, we denote by $i_t^*$ an arm with maximal expected reward, \ie, $\mu_{i_t^*}(t) = \max_i \mu_i(t)$, called an optimal arm (possibly not unique).
+A (random) stream of rewards $(Y_{k,t})_{t\in\N^*}$ is associated to each arm $k \in [K]$. We assume that the rewards are bounded, and without loss of generality we assume that $Y_{k,t} \in [0,1]$.
+We denote by $\mu_k(t) :=  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
+At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) = Y_{A(t),t}$.
+At time $t$, we denote by $k_t^*$ an arm with maximal expected reward, \ie, $\mu_{k_t^*}(t) = \max_k \mu_k(t)$, called an optimal arm (possibly not unique).
 
 A policy $\cA$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
-Like for stationary problems (see Definition~\ref{def:2:regret}), the performance of $\cA$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\cA$
+Like for stationary problems (see Definition~\ref{def:2:regret}), the performance of $\cA$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $k_t^*$ at time $t$, and that of the policy $\cA$
 We use the same notation as the regret for stationary problem without ambiguity, as in this chapter we only consider the following definition.
@@
 \begin{equation}\label{def:6:nti}
-    R_T^{\cA} = \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
-\end{equation}
-
-In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in[K]: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
-If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint recursively by
-$\tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}$.
-%
-Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, that is on any stationary segments, the rewards $(X_{i,t})$ associated to each arm are \iid{} (hence the name of \emph{piece-wise stationary} problems).
+    R_T^{\cA} = \E\left[\sum_{t=1}^T \left(\mu_{k_t^*}(t) - \mu_{A(t)}(t)\right)\right].
+\end{equation}
+
+In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists k\in[K]: \mu_k(t) \neq \mu_k(t+1)\right)$.
+If we denote $\tau^{(0)} = 0$, we define the $i$-th breakpoint recursively by
+$\tau^{(i)} = \inf\{t > \tau^{(i-1)} : \exists k : \mu_k(t) \neq \mu_k(t+1)\}$.
+%
+Hence for $t\in[\tau^{(i)} + 1,\tau^{(i+1)}]$, that is on any stationary segments, the rewards $(Y_{k,t})$ associated to each arm $k$ are \iid{} (hence the name of \emph{piece-wise stationary} problems).
 
@@
 
-We define the \emph{number of change-points} on arm $i$,
-$\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$,
+We define the \emph{number of change-points} on arm $k$,
+$\NCk := \sum_{t=1}^{T-1} \indic\left(\mu_k(t) \neq \mu_k(t+1)\right)$,
 for which it clearly holds that
-$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms.
-Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
+$\NCk \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms.
+Letting $C_T := \sum_{k=1}^K \NCk$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
 We illustrate the two extreme cases in problems $1$ and $2$ presented in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} below.
 
-\paragraph{An interesting interpolation.}
+\textbf{An interesting interpolation.}
 %
@@
 Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the EXP3.S algorithm for the piece-wise stationary problem, as explained in Section~\ref{sec:6:Introduction} above.
-% \TODOL{Explain here that we should have been more careful regarding this sentence?}
 
 \paragraph{Two example of problems.}\label{par:6:benchmark1}
-
+%
 To illustrate the model, we give here two examples of piece-wise stationary problems.
@@
 We consider $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
-The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1} below.
+The arm means $(\mu_k(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1} below.
 % (and Figure~\ref{fig:6:Problem_1} in Appendix),
@@
 Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
-% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
+% \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\alpha/K$ at each time step,
 % and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\alpha \right\rceil$).
@@
 Similarly to what previous approaches did (\eg, \CUSUM{} uses a threshold $\varepsilon$), this phase checks for changes of a certain minimum magnitude, but the innovation in this work is that algorithm does not need any prior knowledge of the minimum magnitude.
-Instead, it uses geometrically decreasing magnitudes $d_i = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only consider a few different values of $d_i$.
+Instead, it uses geometrically decreasing magnitudes $d_k = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only consider a few different values of $d_k$.
 % Our  algorithmAdSwitch(shown  as  Algorithm  1)  proceeds  in  episodes,  where  a  newepisode starts when the algorithm detects a change in one of the arms.
@@
 A different setting where the quantity of interest is not $\Upsilon_T$ but the total variational budget $V_T$, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
-The total variation budget $V_T$ is defined as $\sum_{t=1}^{T-1} \sum_{i=1}^K D(\nu_i(t), \nu_i(t+1))$, for a certain measure of difference $D$, which measures how much the two distributions of any arm $i$ have changed from time $t$ to time $t+1$.
+The total variation budget $V_T$ is defined as $\sum_{t=1}^{T-1} \sum_{k=1}^K D(\nu_k(t), \nu_k(t+1))$, for a certain measure of difference $D$, which measures how much the two distributions of any arm $k$ have changed from time $t$ to time $t+1$.
 Two examples can be the total variation distance $\|\cdot\|_{TV}$ or the Kullback-Leibler divergence.
@@
 %
-\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
+\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_k(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
 %
@@
 In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
-it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{i,t} |\mu_i(t) - \mu_i(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{i,j} |\mu_i(t) - \mu_j(t)|$.
+it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{k,t} |\mu_k(t) - \mu_k(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{k,k'} |\mu_k(t) - \mu_{k'}(t)|$.
 
@@
 
-Assume that we collect independent samples $X_1,X_2,\ldots$ all from some distributions supported in $[0,1]$.
-We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
+Assume that we collect independent samples $X_1,X_2,\dots$ all from some distributions supported in $[0,1]$.
+We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\dots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\dots$ have a different mean $\mu_1 \neq \mu_0$.
 %
@@
 against the alternative
-$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \ldots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \ldots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
+$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
 %
@@
 \begin{equation}\label{eq:6:firstDefGLRT}
-    \GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},
-\end{equation}
-%
-where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the \emph{likelihoods} of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
+    \GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \dots, X_n ; \mu_0)},
+\end{equation}
+%
+where $\ell(X_1, \dots, X_n ; \mu_0)$ and $\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the \emph{likelihoods} of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
 High values of this statistic tend to indicate rejection of $\cH_0$.
@@
     First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
-    We have $\ell(X_1, \ldots, X_n ; \mu_0) = \prod_{i=1}^n \ell(X_i ; \mu_0)$ by independence of the observations $X_i$,
-    and $\ell(X_i ; \mu_0) = \mu_0^{X_i} (1-\mu_0)^{1-X_i}$ for Bernoulli distributions.
+    We have $\ell(X_1, \dots, X_n ; \mu_0) = \prod_{i=1}^n \ell(X_k ; \mu_0)$ by independence of the observations $X_k$,
+    and $\ell(X_k ; \mu_0) = \mu_0^{X_k} (1-\mu_0)^{1-X_k}$ for Bernoulli distributions.
     Therefore, taking the logarithm gives
     \begin{align*}
-        \log\ell(X_1, \ldots, X_n ; \mu_0)
-        &= \sum_{i=1}^n X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) \\
-        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
+        \log\ell(X_1, \dots, X_n ; \mu_0)
+        &= \sum_{i=1}^n X_k \log(\mu_0) + (1-X_k) \log(1-\mu_0) \\
+        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_k \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_k \right) \\
         &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
@@
     \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
-        \sup_{\mu_0} \log\ell(X_1, \ldots, X_n ; \mu_0) = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
+        \sup_{\mu_0} \log\ell(X_1, \dots, X_n ; \mu_0) = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
     \end{equation}
@@
     We can again work with log-likelihoods, and so we have
-
+    %
     \begin{align*}
-        & \log\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau) \\
-        &= \sum_{i=1}^{\tau} X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_i \log(\mu_1) + (1-X_i) \log(1-\mu_1) \\
+        & \log\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau) \\
+        &= \sum_{i=1}^{\tau} X_k \log(\mu_0) + (1-X_k) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_k \log(\mu_1) + (1-X_k) \log(1-\mu_1) \\
         &= s \bigl( \widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0) \bigr) \\
@@
         \GLR(n)
-        % &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)}, \\
-        &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right) \right] }, \\
-        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] }, \\
-        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr] \Bigr],
+        % &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \dots, X_n ; \mu_0)}, \\
+        &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right) \right] }, \\
+        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \frac{\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] }, \\
+        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr] \Bigr],
     \end{align*}
@@
         \log \GLR(n)
-        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr],  \\
+        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr],  \\
         &= \sup\limits_{\mu_0,\mu_1, 1 < s < n} \Bigl[
@@
 $\mu_1$) and then left to be solved for $s$.
-    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_i = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,
+    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_k = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,
     which are both handled similarly.
@@
 \paragraph{False alarm probability.}\label{par:6:falseAlarm}
-
+%
 In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small.
@@
     Let be $\bP_{\mu_0}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t$.
-    % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
+    % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_k \in [0,1]$ for all $i$.
     Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    $(X_i)_{i \in \N^*}$ and $(Y_i)_{i \in \N^*}$ two independent \iid{} processes with respective means $\mu$ and $\mu'$ such that
+    $(X_k)_{i \in \N^*}$ and $(Y_k)_{i \in \N^*}$ two independent \iid{} processes with respective means $\mu$ and $\mu'$ such that
     $\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)}$,
@@
     Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family.
-Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
+Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_k$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_k$, it holds that for every $s,r \in \N^*$,
     \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
@@
 \paragraph{Detection delay.}
-
+%
 Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from $\mu_0$ to $\mu_1$ occurs at time $\tau$.
@@
         % XXX Algorithm
-        \textbf{Initialization: } $\forall i \in [K]$, $\tau_i \leftarrow 0$ and $n_i \leftarrow 0$ \\
-        \For{$t=1,2,\ldots, T$}{
+        \textbf{Initialization: } $\forall k \in [K]$, $\tau_k = 0$ and $n_k = 0$ \\
+        \For{$t=1,2,\dots, T$}{
             \uIf{
@@
             }{
-                $A_t \leftarrow t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor$
+                $A(t) = t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor$
                 \tcp*[f]{forced exploration} \\
@@
             \Else{
-                $A_t \leftarrow \arg\max\limits_{i\in [K]} \mbox{UCB}_i(t)$ as defined in \eqref{def:6:UCB}
+                $A(t) = \arg\max\limits_{k \in [K]} \mbox{UCB}_k(t)$ as defined in \eqref{def:6:UCB}
             }
-            \mbox{Play arm } $A_t$ \mbox{and receive the reward } $X_{A_t,t}$ : $n_{A_t} \leftarrow n_{A_t} + 1; Z_{A_t, n_{A_t}} \leftarrow X_{A_t,t}$ \\
+            \mbox{Play arm } $A(t)$ \mbox{and receive the reward } $Y_{A(t),t}$ : $n_{A(t)} = n_{A(t)} + 1; Z_{A(t), n_{A(t)}} = Y_{A(t),t}$ \\
             \uIf(\tcp*[f]{Change point is detected}){
-                $\GLR_\delta(Z_{A_t,1}, \ldots, Z_{A_t, n_{A_t}})$ = True
+                $\GLR_\delta(Z_{A(t),1}, \dots, Z_{A(t), n_{A(t)}})$ = True
             }
@@
                 \uIf{Global restart}{
-                    $\forall i\in [K], \tau_i \leftarrow t$ and $n_i \leftarrow 0 $.
+                    $\forall k\in [K], \tau_k = t$ and $n_k = 0 $.
                     \tcp*[f]{global restart}
@@
                 \Else{
-                    $\tau_{A_t} \leftarrow t$ and $n_{A_t} \leftarrow 0$
+                    $\tau_{A(t)} = t$ and $n_{A(t)} = 0$
                     \tcp*[f]{local restart}
@@
 \begin{equation}\label{eq:whenDoesGLRklUCBrestart}
-    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
+    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right)\right] \geq \beta(n,\delta),
 \end{equation}
@@
 We define the (\klUCB{} like) index used by our algorithm, by denoting
-$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
-$n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)$
-the number of selections of arm $i$, and
-$\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)$
-their empirical mean (if $n_i(t)\neq0$).
+$\tau_k(t)$ the last restart that happened for arm $k$ before time $t$,
+$n_k(t) = \sum_{s=\tau_k(t)+1}^{t} \indic(A(s)=k)$
+the number of selections of arm $k$, and
+$\widehat{\mu}_k(t) = \frac{1}{n_k(t)} \sum_{s=\tau_k(t)+1}^{t}Y_{k,s} \indic(A(s)=i)$
+their empirical mean (if $n_k(t)\neq0$).
 %
@@
 \begin{equation}\label{def:6:UCB}
-    \UCB_i(t) := \max \Bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \Bigr\}.
+    \UCB_k(t) := \max \Bigl\{ q\in[0,1] : n_k(t) \times \kl\left(\widehat{\mu}_k(t),q\right) \leq f(t - \tau_k(t)) \Bigr\}.
 \end{equation}
@@
 %
-Under local restart, in the general case the times $\tau_i(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate}.
+Under local restart, in the general case the times $\tau_k(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate}.
 %
@@
 $(31)$-$(32)$ in the paper \cite{LiuLeeShroff17}.}:
-$f(t-\tau_i(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{i=1}^K n_i(t)$.
+$f(t-\tau_k(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{k=1}^K n_k(t)$.
 
@@
 Both approaches are parameterized by $\alpha\in(0,1)$.
-More precisely, \CUSUMUCB{} samples arm $i$ with probability $\alpha/K$ at each time step,
+More precisely, \CUSUMUCB{} samples arm $k$ with probability $\alpha/K$ at each time step,
 and \MUCB{} uses a deterministic scheme, to force exploring the $K$ arms: when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\alpha \right\rceil$.
@@
 \begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
-    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $i$ (\ie, for a $k\in\{1,\dots,\NCi\}$, one has $\tau_i^{(k)} = \tau_i(t) < s \leq t < \tau_i^{(k+1)}$) it holds that
+    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $k$ (\ie, for a $i\in[\NCk]$, one has $\tau_k^{(i)} = \tau_k(t) < s \leq t < \tau_k^{(i+1)}$) it holds that
     \begin{equation}
-        n_i(t) - n_i(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor.
+        n_k(t) - n_k(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor.
     \end{equation}
@@
 \begin{smallproof}
-    We consider one arm $i\in[K]$, and when the \GLRklUCB{} algorithm is running,
-    we consider two time steps $s\leq t\in\N^*$, chosen between two restart times for that arm $i$.
-    Lines~$3$-$4$ state that $A_u = u \mod \lceil \frac{K}{\alpha} \rceil$
+    We consider one arm $k\in[K]$, and when the \GLRklUCB{} algorithm is running,
+    we consider two time steps $s\leq t\in\N^*$, chosen between two restart times for that arm $k$.
+    Lines~$3$-$4$ state that $A(u) = u \mod \lceil \frac{K}{\alpha} \rceil$
     if $u \mod \lceil \frac{K}{\alpha} \rceil \in [K]$,
@@
     \begin{align*}
-        n_i(t) - n_i(s)
-        & = \sum_{u=s+1}^{t} \indic(A_u = i) \\
-        & \geq \sum_{u=s+1}^{t} \indic\left(A_u = i, A_u = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
-        & \geq \sum_{u=s+1}^{t} \indic\left(i = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
+        n_k(t) - n_k(s)
+        & = \sum_{u=s+1}^{t} \indic(A(u) = k) \\
+        & \geq \sum_{u=s+1}^{t} \indic\left(A(u) = k, A(u) = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
+        & \geq \sum_{u=s+1}^{t} \indic\left(k = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
         & = \bigl(t - (s+1) + 1\bigr) / \left\lceil \frac{K}{\alpha} \right\rceil
@@
 
-Recall that $\tau^{(k)}$ denotes the position of the $k$-th break-point and let $\mu_i^{(k)}$ be the mean of arm $i$ on the segment between the $k$ and $(k+1)$-th breakpoint:
-$\forall t \in \{ \tau^{(k-1)}+1, \dots, \tau^{(k)} \}, \mu_{i}(t) = \mu_i^{(k)}$.
-We also introduce $k^* = \arg\max_i \mu_i^{(k)} $ and the largest gap at break-point $k$ as $\Delta^{(k)} := \max_{i=1,\dots,K} |\mu_i^{(k)} - \mu_i^{(k-1)}| >0$.
+Recall that $\tau^{(i)}$ denotes the position of the $i$-th break-point and let $\mu_k^{(i)}$ be the mean of arm $k$ on the segment between the $i$ and $(i+1)$-th breakpoint:
+$\forall t \in \{ \tau^{(i-1)}+1, \dots, \tau^{(i)} \}, \mu_k(t) = \mu_k^{(i)}$.
+We also introduce $i^* = \arg\max_k \mu_k^{(i)} $ and the largest gap at break-point $i$ as $\Delta^{(i)} := \max_{k\in[K]} |\mu_k^{(i)} - \mu_k^{(i-1)}| >0$.
 
@@
     \begin{equation}\label{def:6:DelayGlobal}
-        d^{(k)} = d^{(k)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
-    \end{equation}
-    %
-    Then \emph{we assume} that for all change $k \in \{1,\dots,\Upsilon_T\}$,
-    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
+        d^{(i)} = d^{(i)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta^{(i)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
+    \end{equation}
+    %
+    Then \emph{we assume} that for all change $i \in [\Upsilon_T]$,
+    $\tau^{(i)} - \tau^{(i-1)} \geq 2\max (d^{(i)},d^{(i-1)})$.
 \end{leftbar}  % XXX leftbar assumptionbar, comment if needed
@@
 It features the parameters $\alpha$ and $\delta$,
-the KL-divergence terms $\kl(\mu_{i}^{(k)},\mu_{k^*}^{(k)})$ expressing the hardness of the (stationary) MAB problem between two breakpoints,
-and the $\Delta^{(k)}$ terms expressing the hardness of the change-point detection problem.
+the KL-divergence terms $\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})$ expressing the hardness of the (stationary) MAB problem between two breakpoints,
+and the $\Delta^{(i)}$ terms expressing the hardness of the change-point detection problem.
 
@@
     \begin{align}
-        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T \\
-        & + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
+        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T \\
+        & + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
 \nonumber
@@
 and to distinguish notations between local and global changes, we denote $\ell$ instead of $k$ for the indexes of change-points.
-We let $\tau_i^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $i$}: $\tau_i^{(\ell)} = \inf \{ t > \tau_i^{(\ell - 1)} : \mu_i(t) \neq \mu_i(t+1)\}$,
-with the convention $\tau_i^{(0)}=0$, and let $\overline{\mu}_i^{(\ell)}$ be the $\ell$-th value for the mean of arm $i$, such that $\forall t \in [\tau_i^{(\ell-1)}+1, \tau_i^{(\ell)}], \ \ \mu_i(t) = \overline{\mu}_i^{(\ell)}$.
-We also introduce the gap $\Delta_i^{(\ell)} = \overline{\mu}_i^{\ell} - \overline{\mu}_i^{\ell-1} > 0$.
+We let $\tau_k^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $k$}: $\tau_k^{(\ell)} = \inf \{ t > \tau_k^{(\ell - 1)} : \mu_k(t) \neq \mu_k(t+1)\}$,
+with the convention $\tau_k^{(0)}=0$, and let $\overline{\mu}_k^{(\ell)}$ be the $\ell$-th value for the mean of arm $k$, such that $\forall t \in [\tau_k^{(\ell-1)}+1, \tau_k^{(\ell)}], \ \ \mu_k(t) = \overline{\mu}_k^{(\ell)}$.
+We also introduce the gap $\Delta_k^{(\ell)} = \overline{\mu}_k^{\ell} - \overline{\mu}_k^{\ell-1} > 0$.
 
@@
     \begin{equation}\label{def:6:DelayLocal}
-        d_i^{(\ell)} = d_i^{(\ell)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
-    \end{equation}
-    %
-    Then \emph{we assume} that for all arm $i$ and all change-point of that arm $\ell \in \{1,\dots,\NCi\}$, $\tau_i^{(\ell)} - \tau_i^{(\ell-1)} \geq 2\max (d_i^{(\ell)},d_i^{(\ell-1)})$.
+        d_k^{(\ell)} = d_k^{(\ell)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
+    \end{equation}
+    %
+    Then \emph{we assume} that for all arm $k$ and all change-point of that arm $\ell \in [\NCk]$, $\tau_k^{(\ell)} - \tau_k^{(\ell-1)} \geq 2\max (d_k^{(\ell)},d_k^{(\ell-1)})$.
 \end{leftbar}  % XXX leftbar assumptionbar, comment if needed
@@
 It requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
-Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{i=1}^K \NCi \leq K \Upsilon_T$.
+Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{k=1}^K \NCk \leq K \Upsilon_T$.
 
@@
     \begin{equation}
-        R_T \leq 2\sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{4K}{\alpha \left(\Delta_i^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{i=1}^K\sum_{\ell=1}^{\NCi} \frac{\ln(T)}{\kl(\overline{\mu}_i^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
-    \end{equation}
-    where ${\mu}^*_{i,\ell} = \inf \left\{ \mu_{i_t^*}(t) : \mu_{i_t^*}(t) \neq \overline{\mu}_i^{(\ell)}, t \in [\tau_i^{(\ell)}+1, \tau_i^{(\ell+1)}]\right\}$.
+        R_T \leq 2\sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{4K}{\alpha \left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{\ln(T)}{\kl(\overline{\mu}_k^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
+    \end{equation}
+    where ${\mu}^*_{i,\ell} = \inf \left\{ \mu_{k_t^*}(t) : \mu_{k_t^*}(t) \neq \overline{\mu}_k^{(\ell)}, t \in [\tau_k^{(\ell)}+1, \tau_k^{(\ell+1)}]\right\}$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 %
-Recall that in this Chapter, the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$.
+Recall that in this Chapter, the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{k_t^*}(t) - \mu_{A(t)}(t))\right]$.
 First, we introduce $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}),
-and as we observe that $\mu_{i^*_t} - \mu_{I_t} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
+and as we observe that $\mu_{k_t^*} - \mu_{A(t)} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
 we can write a first decomposition of the regret,
 \begin{align*}
- & \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t)) \\
- & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
- & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
+ & \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t)) \\
+ & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{A(t)}(t) \geq \UCB_{k_t^*}(t)\right) \\
+ & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right) + \sum_{k=1}^K\sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_k(t))\indic\left(A(t) = k, \UCB_k(t) \geq \mu_{k_t^*}(t)\right).
 \end{align*}
@@
 \begin{align}\label{eq:6:GeneRegretBound}
-    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
-    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}. \nonumber
+    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right)\right]}_{(A)} \\
+    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(\UCB_{A(t)}(t) \geq \mu_{k_t^*}(t)\right)\right]}_{(B)}. \nonumber
 \end{align}
@@
 %
-To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
+To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(i)}+1, \tau^{(i+1)}]$ for global changes and $[\tau^{(i)}_k+1, \tau^{(i+1)}_k]$ for each arm $k$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
 
@@
 \begin{equation}
-    \cE_T(\alpha,\delta) = \left(\forall i \in \{1, \ldots, K\}, \forall \ell \in \{1, \ldots, \NCi\}, \hat{\tau}^{(\ell)}_i \in \left[\tau_i^{(\ell)} + 1, \tau_i^{(\ell)} + d_i^{(\ell)}\right] \right),
-\end{equation}
-%
-where $\widehat{\tau}_i^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $i$ and $d_i^{(\ell)}=d_i^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
+    \cE_T(\alpha,\delta) = \left(\forall k \in [K], \forall \ell \in [\NCk], \hat{\tau}^{(\ell)}_k \in \left[\tau_k^{(\ell)} + 1, \tau_k^{(\ell)} + d_k^{(\ell)}\right] \right),
+\end{equation}
+%
+where $\widehat{\tau}_k^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $k$ and $d_k^{(\ell)}=d_k^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
 Introducing the event
-\[\cC_i^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_i^{(\ell)} \in \left[\tau_i^{(j)} + 1, \tau_i^{(j)} + d_i^{(j)} \right] \right\},\]
+\[\cC_k^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_k^{(\ell)} \in \left[\tau_k^{(j)} + 1, \tau_k^{(j)} + d_k^{(j)} \right] \right\},\]
 that all the changes up to the $\ell$-th have been detected, a union bound yields this decomposition
@@
 \begin{align}
-    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \leq \tau_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(a)} + \sum\limits_{i=1}^K\sum\limits_{\ell=1}^{\NCi} \underbrace{\bP\left(\widehat{\tau}_i^{(\ell)} \geq \tau_i^{(\ell)} + d_i^{(\ell)} \;|\; \cC_i^{(\ell-1)}\right)}_{(b)}.
+    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{k=1}^K\sum\limits_{\ell=1}^{\NCk} \underbrace{\bP\left(\widehat{\tau}_k^{(\ell)} \leq \tau_k^{(\ell)} \;|\; \cC_k^{(\ell-1)}\right)}_{(a)} + \sum\limits_{k=1}^K\sum\limits_{\ell=1}^{\NCk} \underbrace{\bP\left(\widehat{\tau}_k^{(\ell)} \geq \tau_k^{(\ell)} + d_k^{(\ell)} \;|\; \cC_k^{(\ell-1)}\right)}_{(b)}.
 \end{align}
@@
 Term $(a)$ is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
-Observe that under the bandit algorithm, the change point detector associated to arm $i$ is based on (possibly much) less than $t - \tau_i(t)$ samples from arm $i$, which makes false alarm even less likely to occur.
+Observe that under the bandit algorithm, the change point detector associated to arm $k$ is based on (possibly much) less than $t - \tau_k(t)$ samples from arm $k$, which makes false alarm even less likely to occur.
 Hence, it is easy to show that $(a) \leq \delta$.
@@
 % \begin{align}\label{eq:6:GeneRegretBound}
-%     R_T \leq T \bP\left(\cE_T^c\right) + \alpha T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
-%     & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
+%     R_T \leq T \bP\left(\cE_T^c\right) + \alpha T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right)\right]}_{(A)} \nonumber \\
+%     & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(\UCB_{A(t)}(t) \geq \mu_{k_t^*}(t)\right)\right]}_{(B)}.
 % \end{align}
@@
 We first introduce some notations for the proof:
-let $\widehat{\tau}^{(k)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
-We denote $n_i(t) = \sum_{s=\tau(t)+1}^{t}\indic(A_s = i)$ the number of selections of arm $i$ since the last (global) restart, and $\widehat{\mu}_i(t) = \frac{1}{n_i(t)}\sum_{s=\tau(t)+1}^{t}X_{i,s}\indic(A_s = i)$ their empirical average (if $n_i(t) \neq 0$).
+let $\widehat{\tau}^{(i)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
+We denote $n_k(t) = \sum_{s=\tau(t)+1}^{t}\indic(A(s) = i)$ the number of selections of arm $k$ since the last (global) restart, and $\widehat{\mu}_k(t) = \frac{1}{n_k(t)}\sum_{s=\tau(t)+1}^{t}Y_{k,s}\indic(A(s) = i)$ their empirical average (if $n_k(t) \neq 0$).
 
@@
 Let $d^k$ be defined as in Assumption~\ref{ass:6:LongPeriodsGlobal}, we define
-\begin{equation}\cE_T(\delta) = \left( \forall k \in \{1, \ldots, \Upsilon_T\}, \hat{\tau}^{(k)} \in \left[\tau^{(k)} + 1, \tau^{(k)} + d^{(k)}\right] \right).\label{def:6:GoodEvenGlobal}\end{equation}
+\begin{equation}\cE_T(\delta) = \left( \forall k \in [\Upsilon_T], \hat{\tau}^{(i)} \in \left[\tau^{(i)} + 1, \tau^{(i)} + d^{(i)}\right] \right).\label{def:6:GoodEvenGlobal}\end{equation}
 Under the good event, all the change points are detected within a delay at most $d^(k)$.
 %
-From Assumption~\ref{ass:6:LongPeriodsGlobal}, as the period between two changes are long enough, if $\cE_T(\delta)$ holds, then for all change $k$, we have $\tau^{(k)} \leq \hat{\tau}^{(k)} \leq \tau^{(k+1)}$.
+From Assumption~\ref{ass:6:LongPeriodsGlobal}, as the period between two changes are long enough, if $\cE_T(\delta)$ holds, then for all change $k$, we have $\tau^{(i)} \leq \hat{\tau}^{(i)} \leq \tau^{(i+1)}$.
 So we can prove the following.
@@
 \begin{align*}
-    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{i_t^*}(t) \, \kl\left(\widehat{\mu}_{i_t^*}(t), \mu_{i_t^*}(t)\right) \geq f(t - \tau(t))\right)\right] \\
-     & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}(t))\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(k)} + \indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}^{(k)})\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d^{(k)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(k)})\right)\right],
+    (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{k_t^*}(t) \, \kl\left(\widehat{\mu}_{k_t^*}(t), \mu_{k_t^*}(t)\right) \geq f(t - \tau(t))\right)\right] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}(t))\right)\right] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(i)} + \indic(\cE_T) \sum_{t=\hat{\tau}^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(i)}) \sum_{t=\hat{\tau}^{(i)}}^{\tau^{(i+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right],
 \end{align*}
 %
-where we introduce the event $\cC^{(k)}$ that all the changes up to the $k$-th have been detected:
+where we introduce the event $\cC^{(i)}$ that all the changes up to the $k$-th have been detected:
 %
 \begin{equation}\label{def:6:EventCiGlobal}
-    \cC^{(k)} = \left\{\forall j \leq k, \widehat{\tau}^{(j)} \in \{\tau^{(j)} + 1, \dots, \tau^{(j)} + d^{(j)} \} \right\}.
-\end{equation}
-%
-Clearly, $\cE_T \subseteq \cC^{(k)}$ and $\cC^{(k)}$ is $\cF_{\widehat{\tau}^{(k)}}$-measurable.
-Observe that conditionally to $\cF_{\widehat{\tau}^{(k)}}$, when $\cC^{(k)}$ holds, $\widehat{\mu}_{k^*}(t)$ is the average of samples that have all mean $\mu_{k^*}^{(k)}$.
-%
-Thus, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_{k^*}^{(k)}$, one can write
+    \cC^{(i)} = \left\{\forall j \leq k, \widehat{\tau}^{(j)} \in \{\tau^{(j)} + 1, \dots, \tau^{(j)} + d^{(j)} \} \right\}.
+\end{equation}
+%
+Clearly, $\cE_T \subseteq \cC^{(i)}$ and $\cC^{(i)}$ is $\cF_{\widehat{\tau}^{(i)}}$-measurable.
+Observe that conditionally to $\cF_{\widehat{\tau}^{(i)}}$, when $\cC^{(i)}$ holds, $\widehat{\mu}_{k^*}(t)$ is the average of samples that have all mean $\mu_{k^*}^{(i)}$.
+%
+Thus, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_{k^*}^{(i)}$, one can write
 
 \begin{align*}
-    &\bE\left[\left.\indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}^{(k)})\right) \right| \cF_{\widehat{\tau}^{(k)}}\right]\\
-    & = \indic(\cC^{(k)}) \sum_{t=\hat{\tau}^{(k)}}^{\tau^{(k+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(k)}\right) \geq f(t - \widehat{\tau}^{(k)})\right) \;|\; \cF_{\widehat{\tau}^{(k)}}\right] \\
-    & \leq \indic(\cC^{(k)}) \sum_{t'=1}^{\tau^{(k+1)} - \hat{\tau}^{(k)}} \bP\left(\exists s \leq t' : s \, \kl(\widehat{\mu}_{s},\mu_{k^*}^{(k)}) \geq f(t')\right) \\
+    &\bE\left[\left.\indic(\cC^{(i)}) \sum_{t=\hat{\tau}^{(i)}}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \right| \cF_{\widehat{\tau}^{(i)}}\right]\\
+    & = \indic(\cC^{(i)}) \sum_{t=\hat{\tau}^{(i)}}^{\tau^{(i+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \;|\; \cF_{\widehat{\tau}^{(i)}}\right] \\
+    & \leq \indic(\cC^{(i)}) \sum_{t'=1}^{\tau^{(i+1)} - \hat{\tau}^{(i)}} \bP\left(\exists s \leq t' : s \, \kl(\widehat{\mu}_{s},\mu_{k^*}^{(i)}) \geq f(t')\right) \\
     & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
@@
 \begin{equation}\label{eq:6:TermAFinalGlobal}
-    (A) \leq \sum_{k=1}^{\Upsilon_T}\left[d^{(k)} + \ln(\ln(T))\right].
+    (A) \leq \sum_{k=1}^{\Upsilon_T}\left[d^{(i)} + \ln(\ln(T))\right].
 \end{equation}
@@
 %
-We let $\tilde{\mu}_{i,s}^{(k)}$ denote the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
-Rewriting the sum in $t$ as the sum of consecutive intervals $[\tau^{(k)}+1, \tau^{(k+1)}]$,
+We let $\tilde{\mu}_{k,s}^{(i)}$ denote the empirical mean of the first $s$ observations of arm $k$ made after time $t=\hat{\tau}^{(i)}+1$.
+Rewriting the sum in $t$ as the sum of consecutive intervals $[\tau^{(i)}+1, \tau^{(i+1)}]$,
 %
 \begin{align*}
-    (B) & \leq \bE\Big[\indic(\cE_T)\sum_{k=1}^{\Upsilon_T} \sum_{t=\tau^{(k)}}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{I_t}^{(k)}\right)\indic\left(\UCB_{I_t}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
-     & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\hat{\tau}^{(k)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{I_t}^{(k)}\right)\indic\left(\UCB_{I_t}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\indic\left(I_t=i,\UCB_{i}(t) \geq \mu_{k^*}^{(k)}\right)\Big] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\times \\
-    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(k)}+1}^{\tau^{(k+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(k)}}\indic\left(I_t = i,n_i(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \widehat{\tau}^{(k)})\right)\Big]
+    (B) & \leq \bE\Big[\indic(\cE_T)\sum_{k=1}^{\Upsilon_T} \sum_{t=\tau^{(i)}}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\hat{\tau}^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\indic\left(A(t)=i,\UCB_k(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
+    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
 \end{align*}%
 \begin{align*}
-    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_i(\tau^{(k+1)})}\indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\Big]  \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_i^{(k)}+ \sum_{i=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)\bE\Big[\indic(\cC^{(k)}) \sum_{s=1}^{n_i(\tau^{(k+1)})}\indic\left(s \, \kl(\tilde{\mu}_{i,s}^{(k)}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\Big].
+    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cC^{(i)}) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big].
 \end{align*}
 
-Conditionally to $\cF_{\widehat{\tau}^{(k)}}$, when $\cC^{(k)}$ holds, for $s \in \{1, \ldots, n_i(\tau^{(k+1)})\}$, $\tilde{\mu}_{i,s}^{(k)}$ is the empirical mean from \iid{} observations of mean $\mu_i^{(k)}$.
-Therefore, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_i^{(k)}$, it follows from the law of total expectation that
+Conditionally to $\cF_{\widehat{\tau}^{(i)}}$, when $\cC^{(i)}$ holds, for $s \in [n_k(\tau^{(i+1)})]$, $\tilde{\mu}_{k,s}^{(i)}$ is the empirical mean from \iid{} observations of mean $\mu_k^{(i)}$.
+Therefore, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_k^{(i)}$, it follows from the law of total expectation that
 \begin{align*}
-    (B)& \leq \sum_{k=1}^{\Upsilon_T}d^{(k)}+\sum_{i=1}^K\sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right) \sum_{s=1}^{\tau^{(k+1)}-\tau^{(k)}}\bP\left(s \times \kl(\widehat{\mu}_{s}, \mu_{k^*}^{(k)}) \leq f(\tau^{(k+1)} - \tau^{(k)})\right)\cdot
+    (B)& \leq \sum_{k=1}^{\Upsilon_T}d^{(i)}+\sum_{k=1}^K\sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right) \sum_{s=1}^{\tau^{(i+1)}-\tau^{(i)}}\bP\left(s \times \kl(\widehat{\mu}_{s}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\cdot
 \end{align*}
 %
-If $\mu_{k^*}^{(k)} > \mu_i^{(k)}$ by definition, we can use the analysis of \klUCB{} from \cite{KLUCBJournal} to further upper bound the right-most part, and we obtain
+If $\mu_{k^*}^{(i)} > \mu_k^{(i)}$ by definition, we can use the analysis of \klUCB{} from \cite{KLUCBJournal} to further upper bound the right-most part, and we obtain
 %
 \begin{align}\label{eq:6:TermBFinalGlobal}
-    (B) & \leq \sum_{k=1}^{\Upsilon_T}d^{(k)}+\sum_{i=1}^K\sum_{k=1}^{\Upsilon_T}\indic\left(\mu_i^{(k)} \neq \mu_{k^*}^{(k)}\right)\left[ \frac{\left(\mu_{k^*}^{(k)} - \mu_{i}^{(k)}\right)}{\kl(\mu_i^{(k)},\mu_{k^*}^{(k)})}\log(T) + \bigO{\sqrt{\log(T)}}\right].
+    (B) & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)}+\sum_{k=1}^K\sum_{k=1}^{\Upsilon_T}\indic\left(\mu_k^{(i)} \neq \mu_{k^*}^{(i)}\right)\left[ \frac{\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)}{\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})}\log(T) + \bigO{\sqrt{\log(T)}}\right].
 \end{align}
@@
 Combining the regret decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEventGlobal} and the two upper bounds of $(A)$ in \eqref{eq:6:TermAFinalGlobal} and of $(B)$ in \eqref{eq:6:TermBFinalGlobal},
-\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(k)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_i^{(k)} \neq \mu_{k^*}^{(k)}} \frac{\left( \mu_{k^*}^{(k)}-\mu_i^{(k)}\right)}{\kl\left(\mu_i^{(k)},\mu_{k^*}^{(k)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
+\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
 which concludes the proof.
@@
 
-% We introduce $\widehat{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after $t=\tau^{(k)}+1$ and $\tilde{\mu}_{i,s}^{(k)}$ as the empirical mean of the first $s$ observations of arm $i$ made after time $t=\hat{\tau}^{(k)}+1$.
-
-Recall that $\cC^{(k)}$ defined in \eqref{def:6:EventCiGlobal} is the event that all the breakpoints up to the $k$-th have been correctly detected.
+% We introduce $\widehat{\mu}_{k,s}^{(i)}$ as the empirical mean of the first $s$ observations of arm $k$ made after $t=\tau^{(i)}+1$ and $\tilde{\mu}_{k,s}^{(i)}$ as the empirical mean of the first $s$ observations of arm $k$ made after time $t=\hat{\tau}^{(i)}+1$.
+
+Recall that $\cC^{(i)}$ defined in \eqref{def:6:EventCiGlobal} is the event that all the breakpoints up to the $k$-th have been correctly detected.
 Using a union bound, one can write
@@
     \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T}
-    \bP\left(\left. \widehat{\tau}^{(k)} \notin \{ \tau^{(k)}+1, \dots, \tau^{(k)} + d^{(k)} \} \right|\; \cC^{(k-1)}\right)
+    \bP\left(\left. \widehat{\tau}^{(i)} \notin \{ \tau^{(i)}+1, \dots, \tau^{(i)} + d^{(i)} \} \right|\; \cC^{(i-1)}\right)
 \end{align*}
@@
 \begin{align*}
-    \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \leq \tau^{(k)} \;|\; \cC^{(k-1)}\right)}_{(a)} \;\; + \;\; \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)} \;|\; \cC^{(k-1)}\right)}_{(b)}.
+    \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(i)} \leq \tau^{(i)} \;|\; \cC^{(i-1)}\right)}_{(a)} \;\; + \;\; \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)} \;|\; \cC^{(i-1)}\right)}_{(b)}.
 \end{align*}
@@
 %
-We have $\widehat{\tau}^{(k)} \leq \tau^{(k)}$ and it implies that there exists an arm whose associated change point detector has experienced a false-alarm:
+We have $\widehat{\tau}^{(i)} \leq \tau^{(i)}$ and it implies that there exists an arm whose associated change point detector has experienced a false-alarm:
 %
 \begin{align*}
-    (a) & \leq \bP\left(\exists i, \exists s < t \leq n_i(\tau_i^{(k)}) : s \, \kl\left(\tilde{\mu}_{i,1:s}^{(k-1)},\tilde{\mu}_{i,1:t}^{(k-1)}\right) + (t - s) \, \kl\left(\tilde{\mu}_{i,s+1:t}^{(k-1)},\tilde{\mu}_{i,1:t}^{(k-1)}\right) > \beta(t, \delta) \;|\; \cC^{(k-1)}\right) \\
-    & \leq \sum_{i=1}^K\bP\left(\exists s < t : s \, \kl(\widehat{\mu}_{1:s},\mu_i^{(k-1)}) + (t - s) \, \kl(\widehat{\mu}_{s+1:t}, \mu_{i}^{(k-1)}) > \beta(t, \delta)\right),
+    (a) & \leq \bP\left(\exists i, \exists s < t \leq n_k(\tau_k^{(i)}) : s \, \kl\left(\tilde{\mu}_{i,1:s}^{(i-1)},\tilde{\mu}_{i,1:t}^{(i-1)}\right) + (t - s) \, \kl\left(\tilde{\mu}_{i,s+1:t}^{(i-1)},\tilde{\mu}_{i,1:t}^{(i-1)}\right) > \beta(t, \delta) \;|\; \cC^{(i-1)}\right) \\
+    & \leq \sum_{k=1}^K\bP\left(\exists s < t : s \, \kl(\widehat{\mu}_{1:s},\mu_k^{(i-1)}) + (t - s) \, \kl(\widehat{\mu}_{s+1:t}, \mu_k^{(i-1)}) > \beta(t, \delta)\right),
 \end{align*}
-with $\hat\mu_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_i^{(k-1)}$.
-Indeed, conditionally to $\cC^{(k-1)}$, the $n_i(\tau^{(k)})$ successive observations of arm $i$ arm starting from $\hat \tau^{(k)}$ are \iid{} with mean $\mu_i^{(k-1)}$.
+with $\hat\mu_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_k^{(i-1)}$.
+Indeed, conditionally to $\cC^{(i-1)}$, the $n_k(\tau^{(i)})$ successive observations of arm $k$ arm starting from $\hat \tau^{(i)}$ are \iid{} with mean $\mu_k^{(i-1)}$.
 Using Lemma~\ref{lem:6:ConcFirst}, term $(a)$ is upper bounded by $K\delta$.
@@
 %
-From the definition of $\Delta^{(k)}$, there exists an arm $i$ such that $\Delta^{(k)} = |\mu_i^{(k)} - \mu_i^{(k-1)}|$.
-We shall prove that it is unlikely that the change-point detector associated to $i$ doesn't trigger within the delay $d^{(k)}$.
-
-First, it follows from Proposition~\ref{prop:6:EnoughSamples} that there exists $\overline{t} \in \{\tau^{(k)}, \dots, \tau^{(k)} + d^{(k)} \}$ such that
-$n_i(\overline{t}) - n_i(\tau^{(k)}) = \overline{r}$ where $\overline{r} = \lfloor \frac{\alpha}{K} d^{(k)}\rfloor$
-(as the mapping $t\mapsto n_i(t) - n_i(\tau^{(k)})$ is non-decreasing, is $0$ at $t=\tau^{(k)}$ and its value at $\tau^{(k)}+d^{(k)}$ is larger than $\overline{r}$).
-Using that \[\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)}\right) \subseteq \left(\widehat{\tau}^{(k)} \geq \overline{t}\right)\]
-the event $\left(\widehat{\tau}^{(k)} \geq \tau^{(k)} + d^{(k)}\right)$ further implies that
+From the definition of $\Delta^{(i)}$, there exists an arm $k$ such that $\Delta^{(i)} = |\mu_k^{(i)} - \mu_k^{(i-1)}|$.
+We shall prove that it is unlikely that the change-point detector associated to $i$ doesn't trigger within the delay $d^{(i)}$.
+
+First, it follows from Proposition~\ref{prop:6:EnoughSamples} that there exists $\overline{t} \in \{\tau^{(i)}, \dots, \tau^{(i)} + d^{(i)} \}$ such that
+$n_k(\overline{t}) - n_k(\tau^{(i)}) = \overline{r}$ where $\overline{r} = \lfloor \frac{\alpha}{K} d^{(i)}\rfloor$
+(as the mapping $t\mapsto n_k(t) - n_k(\tau^{(i)})$ is non-decreasing, is $0$ at $t=\tau^{(i)}$ and its value at $\tau^{(i)}+d^{(i)}$ is larger than $\overline{r}$).
+Using that \[\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right) \subseteq \left(\widehat{\tau}^{(i)} \geq \overline{t}\right)\]
+the event $\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right)$ further implies that
 \[
-    n_i(\tau^{(k)}) \, \kl\left(\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)})},\tilde{\mu}^{k-1}_{i,n_i(\overline{t})}\right)
-    + \overline{r} \, \kl\left(\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)}) : n_i(\overline{t})},\tilde{\mu}^{k-1}_{i,n_i(\overline{t})}\right) \leq \beta(n_i(\tau^{(k)}) + \overline{r},\delta),
+    n_k(\tau^{(i)}) \, \kl\left(\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})},\tilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right)
+    + \overline{r} \, \kl\left(\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})},\tilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right) \leq \beta(n_k(\tau^{(i)}) + \overline{r},\delta),
 \]
-where $\tilde{\mu}^{k-1}_{i,s}$ denotes the empirical mean of the $s$ first observation of arm $i$ since the $(k-1)$-th restart $\hat{\tau}^{(k-1)}$ and  $\tilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
-Conditionally to $\cC^{(k-1)}$, $\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)})}$ is the empirical mean of $n_i(\tau^{(k)})$ \iid{} replications of mean $\mu^{k-1}$, whereas $\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)}) : n_i(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_i^{k}$.
-
-Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_i(\tau^{(k)})$ lies in the interval $
-\left[\left\lfloor \frac{\alpha}{K}\left(\tau^{(k)}-\widehat{\tau}^{(k-1)}\right)\right\rfloor,\left(\tau^{(k)}-\widehat{\tau}^{(k-1)}\right)\right]$.
-Conditionally to $\cC^{(k-1)}$,
+where $\tilde{\mu}^{k-1}_{k,s}$ denotes the empirical mean of the $s$ first observation of arm $k$ since the $(k-1)$-th restart $\hat{\tau}^{(i-1)}$ and  $\tilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
+Conditionally to $\cC^{(i-1)}$, $\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})}$ is the empirical mean of $n_k(\tau^{(i)})$ \iid{} replications of mean $\mu^{k-1}$, whereas $\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_k^{k}$.
+
+Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_k(\tau^{(i)})$ lies in the interval $
+\left[\left\lfloor \frac{\alpha}{K}\left(\tau^{(i)}-\widehat{\tau}^{(i-1)}\right)\right\rfloor,\left(\tau^{(i)}-\widehat{\tau}^{(i-1)}\right)\right]$.
+Conditionally to $\cC^{(i-1)}$,
 by using Assumption~\ref{ass:6:LongPeriodsGlobal},
-we can prove that $d^{(k-1)} \leq (\tau^{(k)} - \tau^{(k-1)})/2$,
+we can prove that $d^{(i-1)} \leq (\tau^{(i)} - \tau^{(i-1)})/2$,
 and thus we obtain furthermore that
@@
 \begin{eqnarray*}
-    n_i(\tau^{(k)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(k)}-\tau^{(k-1)} )\right\rfloor, \dots, \tau^{(k)}-\tau^{(k-1)} \right\} := \cI_k.
+    n_k(\tau^{(i)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} := \cI_k.
 \end{eqnarray*}
@@
 Introducing $\widehat{\mu}_{a,s}$ (resp.
-$\widehat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\widehat{\mu}_i^{(k-1)}$ (resp.
-$\widehat{\mu}_i^{(k)}$), such that $\widehat{\mu}_{a,s}$ and $\widehat{\mu}_{b,r}$ are independent, it follows that
+$\widehat{\mu}_{b,s}$) the empirical mean of $s$ \iid{} observations with mean $\widehat{\mu}_k^{(i-1)}$ (resp.
+$\widehat{\mu}_k^{(i)}$), such that $\widehat{\mu}_{a,s}$ and $\widehat{\mu}_{b,r}$ are independent, it follows that
 \[
@@
 \]
-where we have also used that $\tilde{\mu}^{k-1}_{i,n_i(\overline{t})} = \left(n_i(\tau^{(k)})\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)})} + \overline{r}\tilde{\mu}^{k-1}_{i,n_i(\tau^{(k)}) : n_i(\overline{t})} \right) / (n_i(\tau^{(k)}) + \overline{r})$.
-
-Using Pinsker's inequality from Lemma~\ref{lem:2:Pinsker} and introducing the gap $\Delta_i^{(k)} = \mu_i^{(k-1)} - {\mu}_i^{(k)}$ (which is such that $\Delta^{(k)} = |\Delta_i^k|$), one can write
+where we have also used that $\tilde{\mu}^{k-1}_{i,n_k(\overline{t})} = \left(n_k(\tau^{(i)})\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})} + \overline{r}\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})} \right) / (n_k(\tau^{(i)}) + \overline{r})$.
+
+Using Pinsker's inequality from Lemma~\ref{lem:2:Pinsker} and introducing the gap $\Delta_k^{(i)} = \mu_k^{(i-1)} - {\mu}_k^{(i)}$ (which is such that $\Delta^{(i)} = |\Delta_k^k|$), one can write
 %
@@
     (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
-    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,s} - \Delta_i^{(k)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
-    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
+    & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,s} - \Delta_k^{(i)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
+    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_k^{(i)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
 \end{align}
@@
 For the second term, we use the observation
-\[\frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_i^{(k)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}| \geq |\Delta_i^{(k)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
-and, using that $\Delta^{(k)} = |\Delta_i^k|$, one obtains
+\[\frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_k^{(i)}\right)^2  \leq \beta(s+\overline{r},\delta) \ \ \Rightarrow \ \ |\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}| \geq |\Delta_k^{(i)}| - \sqrt{\frac{s+\overline{r}}{2\overline{r}s}\beta(s+\overline{r},\delta)}\]
+and, using that $\Delta^{(i)} = |\Delta_k^k|$, one obtains
 %
 \begin{equation}\label{eq:6:FromHereGlobal}
-    (b) \leq \delta + \bP\left(\exists s \in \cI_k : \Delta^{(k)} \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).
-\end{equation}
-%
-Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau^{(k)}-\tau^{(k-1)})/2\right\rfloor$.
+    (b) \leq \delta + \bP\left(\exists s \in \cI_k : \Delta^{(i)} \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).
+\end{equation}
+%
+Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau^{(i)}-\tau^{(i-1)})/2\right\rfloor$.
 Using that the mappings $s \mapsto (s+\overline{r})/s\overline{r}$ and $s \mapsto \beta(s + \overline{r},\delta)$ are respectively decreasing and increasing in $s$, one has, for all $s\in \cI_k$,
@@
     2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) & \leq &
-    2\frac{s_{\min}+\overline{r}}{s_{\min}\overline{r}}\beta\left(T, \delta\right) \leq \frac{4\beta(T,\delta)}{\left\lfloor \frac{\alpha}{K}d^{(k)}\right\rfloor},
+    2\frac{s_{\min}+\overline{r}}{s_{\min}\overline{r}}\beta\left(T, \delta\right) \leq \frac{4\beta(T,\delta)}{\left\lfloor \frac{\alpha}{K}d^{(i)}\right\rfloor},
 \end{eqnarray*}
 %
-where the last inequality follows from the fact that $\overline{r} \leq s_{\min}$ as $d^{(k)} \leq (\tau^{(k)} - \tau^{(k-1)})/2$ by Assumption~\ref{ass:6:LongPeriodsGlobal}.
-Now the definition of $d^{(k)}$ readily implies that
-\[\left\lfloor \frac{\alpha}{K}d^{(k)}\right\rfloor > \frac{4\beta(T,\delta)}{\left(\Delta^{(k)}\right)^2},\]
+where the last inequality follows from the fact that $\overline{r} \leq s_{\min}$ as $d^{(i)} \leq (\tau^{(i)} - \tau^{(i-1)})/2$ by Assumption~\ref{ass:6:LongPeriodsGlobal}.
+Now the definition of $d^{(i)}$ readily implies that
+\[\left\lfloor \frac{\alpha}{K}d^{(i)}\right\rfloor > \frac{4\beta(T,\delta)}{\left(\Delta^{(i)}\right)^2},\]
 which yields
-\[\forall s \in \cI_k, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta^{(k)}\right)^2.\]
+\[\forall s \in \cI_k, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta^{(i)}\right)^2.\]
 %
@@
 
-% \TODOE{note that in this corollary $\Delta^{\text{change}}$ could be defined as the smallest value of $\Delta^{(k)}$ which can be larger than the smallest change on any arm...
+% \TODOE{note that in this corollary $\Delta^{\text{change}}$ could be defined as the smallest value of $\Delta^{(i)}$ which can be larger than the smallest change on any arm...
 But for the sake of uniformity between the two corollaries, we may prefer to keep it like that}
@@
 
-\textbf{Problem $\bm 3$.} (see Figure~\ref{fig:6:Problem_5}) This problem is harder, with $K=6$, $\Upsilon_T=8$ and $T=20000$.
+\textbf{Problem 3.} (see Figure~\ref{fig:6:Problem_5}) This problem is harder, with $K=6$, $\Upsilon_T=8$ and $T=20000$.
 Most arms change at almost every time steps,
@@
 % - Pb 4 changes are on all or almost arms at a time, but sequences don't have same length
-\paragraph{Problem 4.}
-
+\textbf{Problem 4.}
+%
 Like problem $1$, it uses $K=3$ arms, $\Upsilon=4$ change-points and $T=5000$,
@@
 % - Pb 6 is Yahoo! example from Figure~3 in CUSUM-UCB paper \cite{LiuLeeShroff17}
-\paragraph{Problem 5.}
-
+\textbf{Problem 5.}
+%
 Like problem $3$, this harder problem is also inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!}, but from another competitor paper, see Figure~3 from \cite{LiuLeeShroff17}.

### 2019-06-08 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Done, I changed := to \overset{.}{=}

@@
 				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
-				Let $p^{t}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(\{A_a(t) = j\}), \;\forall 1\leq j \leq K$\;
+				Let $p^{t}_j \eqdef \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(\{A_a(t) = j\}), \;\forall 1\leq j \leq K$\;
 				% \tcp*{Sum trusts}
@@
 				}
-				Renormalize the new weights $\pi^{t+1}$: $\pi^{t+1} := \pi^{t+1} / \sum_{a=1}^{N} \pi^{t+1}_{a}$.
+				Renormalize the new weights $\pi^{t+1}$: $\pi^{t+1} \eqdef \pi^{t+1} / \sum_{a=1}^{N} \pi^{t+1}_{a}$.
 				% \tcp*{Project it back to $\Delta_N$}
@@
 \begin{equation}
-	\widehat{r^a}(t) := \frac{Y_{A_a(t),t}}{p_{A_a(t)}^{t}} \indic(A_a(t) = A(t))
+	\widehat{r^a}(t) \eqdef \frac{Y_{A_a(t),t}}{p_{A_a(t)}^{t}} \indic(A_a(t) = A(t))
 \end{equation}
@@
 % helps to choose a good sequence.
-If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t := \eta = 2 \log(N) / (T K)$.
+If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t \eqdef \eta = 2 \log(N) / (T K)$.
 However, for real-world communication problems, it can be considered unrealistic\footnote{~For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}.} to assume a fixed and known time horizon, so we prefer the alternative horizon-free choice%
@@
 of learning rates,
-$\eta_t := \log(N) / (t K)$ suggested by \cite{Bubeck12}.
+$\eta_t \eqdef \log(N) / (t K)$ suggested by \cite{Bubeck12}.
 We compare both approaches empirically, and the second one usually performs better.

### 2019-06-08 2-Chapters/2-Chapter/chapter2.tex
Done, I changed := to \overset{.}{=}

@@
 \textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
-When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) := Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
+When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) \eqdef Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
 %
@@
 The UI of the demo also shows the current value of $t$ (``total plays'') and $\sum_{s=1}^t r(s)$ the ``total reward''.
-For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) := \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) := \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
-%
-The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) := X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated Probs'' line.
+For each arm, we show the sum of rewards obtained from that arm, \ie, $X_k(t) \eqdef \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, in the ``Rewards'' line, and the number of pulls of that arm, \ie, $N_k(t) \eqdef \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ in the ``Pulls'' line.
+%
+The demo also shows the estimated probability of each arm, that is $\widehat{\mu_k}(t) \eqdef X_k(t) / N_k(t)$ (when $N_k(t)>0$), in the ``Estimated Probs'' line.
 
@@
 More formally, an algorithm can be defined as a sequence of \emph{measurable} functions $(\cA_t)_{t\geq1}$,
-where $\cA_t$ maps the past observations $O_t := (U_0, Y_{A(1),1}, U_1, \dots, Y_{A(t-1),t-1}, U_{t-1})$
-to an arm $\cA_t(O_t) := A(t) \in[K]$
+where $\cA_t$ maps the past observations $O_t \eqdef (U_0, Y_{A(1),1}, U_1, \dots, Y_{A(t-1),t-1}, U_{t-1})$
+to an arm $\cA_t(O_t) \eqdef A(t) \in[K]$
 (we remind that we denote $r(s) = Y_{A(s),s}$ the $s$-th reward).
@@
     \begin{align}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
-        \KL\left(d_1, d_2\right) := \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}\cup\{\pm\infty\}.
+        \KL\left(d_1, d_2\right) \eqdef \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}\cup\{\pm\infty\}.
     \end{align}
@@
 $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$,
-we can use a shorter notation and write $\KL( \eta_1, \eta_2 ) := \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
+we can use a shorter notation and write $\KL( \eta_1, \eta_2 ) \eqdef \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
 and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$,
@@
     The KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
-    $\kl(x,y) := \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
+    $\kl(x,y) \eqdef \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
     and it is also called the \emph{relative binary entropy}.
@@
     The KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
-    $d_{\sigma^2}(x,y) := \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$
+    $d_{\sigma^2}(x,y) \eqdef \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$
     (see Chapter~8 of \cite{JordanCourseStatBerkeley}).
@@
 We focus on distributions fully characterized by their means, that can be Bernoulli or any one-dimensional exponential family.
-We denote $\mu_k := \E[Y_{k,t}] \in \R$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$}).
+We denote $\mu_k \eqdef \E[Y_{k,t}] \in \R$ the mean of the distribution of arm $k$ (it will be referred to as the \emph{mean of arm $k$}).
 %
@@
     Consider a bandit problem of $K$ arms with distributions of means $\bm{\mu}=\mu_1,\dots,\mu_k$.
-    Denote $\mu^* := \max_k \mu_k$ the largest mean.
+    Denote $\mu^* \eqdef \max_k \mu_k$ the largest mean.
     The best arm can be non unique, and any arm $k$ having $\mu_k = \mu^*$ is said to be \emph{optimal},
@@
 let us introduce the (mean) regret $R_T^{\cA}$ of an algorithm $\cA$ as
-$R_T^{\cA} := \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]$.
+$R_T^{\cA} \eqdef \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]$.
 %
@@
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\mu_1,\dots,\mu_K$, with $\mu^* := \max_k \mu_k$, then its \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ defined as
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\mu_1,\dots,\mu_K$, with $\mu^* \eqdef \max_k \mu_k$, then its \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ defined as
     %
@@
         R_T^{\cA}
-        := \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
+        \eqdef \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
         % = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right]
@@
 %
-Remember that $N_k(t) := \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ denotes the number of times arm $k$ was selected between times $1$ and $t$,
+Remember that $N_k(t) \eqdef \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ denotes the number of times arm $k$ was selected between times $1$ and $t$,
 %
 and that the samples $Y_{k,t}$ are all \iid{} of mean $\mu_k$.
-The \emph{gap} between any arm $k\in[K]$ and an optimal arm is defined as $\Delta_k := \mu^* - \mu_k$
+The \emph{gap} between any arm $k\in[K]$ and an optimal arm is defined as $\Delta_k \eqdef \mu^* - \mu_k$
 (an arm $k$ is thus sub-optimal if and only if $\Delta_k > 0$),
@@
 % - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $\mathrm{SC}^{\cA}(T) := \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
+The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $\mathrm{SC}^{\cA}(T) \eqdef \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
 It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptatively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.
@@
 They are detailed in Algorithm~\ref{algo:2:naiveStrategies} below.
-We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) := \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) := \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
-%
-The estimated means, or empirical averages, are $\widehat{\mu_k}(t) := X_k(t) / N_k(t)$ (when $N_k(t)>0$).
+We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) \eqdef \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) \eqdef \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
+%
+The estimated means, or empirical averages, are $\widehat{\mu_k}(t) \eqdef X_k(t) / N_k(t)$ (when $N_k(t)>0$).
 
@@
     Let $Z_1,\dots,Z_n$ be $n$ \iid{} samples from a Bernoulli distribution of parameter $\mu$ (where $n$ is fixed),
-    of empirical mean $\widehat{Z_n} := \frac{1}{n} \sum\limits_{i=1}^n Z_i$, \emph{Hoeffding's inequality} gives both
+    of empirical mean $\widehat{Z_n} \eqdef \frac{1}{n} \sum\limits_{i=1}^n Z_i$, \emph{Hoeffding's inequality} gives both
     \begin{align}\label{eq:2:HoeffdingInequality}
@@
 For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
-the \emph{average reward} $\widehat{\mu_k}(t) := \frac{X_k(t)}{N_k(t)}$
-and a \emph{confidence radius} $\xi_k(t) := \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
+the \emph{average reward} $\widehat{\mu_k}(t) \eqdef \frac{X_k(t)}{N_k(t)}$
+and a \emph{confidence radius} $\xi_k(t) \eqdef \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
 %
 \begin{align}\label{eq:UCB_index}
-    U_k^{\UCB_1}(t) = \UCB_k(t) := \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation} \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration} \xi_k(t)}.
+    U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation} \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration} \xi_k(t)}.
 \end{align}
@@
 %
-This theorem shows that $\UCB_1$ achieves a order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta := \min\limits_{k: \Delta_k>0} \Delta$).
+This theorem shows that $\UCB_1$ achieves a order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta \eqdef \min\limits_{k: \Delta_k>0} \Delta$).
 
@@
         \begin{cases}
-        I_1 \;:=\; & ( \widehat{\mu}_{k^*,n_{k^*}}(t) \leq \mu_{k^*} - \xi_{k^*,n_{k^*}}(t) ), \\
-        I_2 \;:=\; & ( \widehat{\mu}_{k,n_k}(t) \geq \mu_k + \xi_{k,n_k}(t) ), \\
-        I_3 \;:=\; & ( \mu_{k^*} < \mu_k + 2 \; \xi_{k,n_k} (t) ).
+        I_1 \; \eqdef \; & ( \widehat{\mu}_{k^*,n_{k^*}}(t) \leq \mu_{k^*} - \xi_{k^*,n_{k^*}}(t) ), \\
+        I_2 \; \eqdef \; & ( \widehat{\mu}_{k,n_k}(t) \geq \mu_k + \xi_{k,n_k}(t) ), \\
+        I_3 \; \eqdef \; & ( \mu_{k^*} < \mu_k + 2 \; \xi_{k,n_k} (t) ).
         \end{cases}
@@
         For the last event $I_3$,
-        observe that $\mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t)$ means $\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0$, where we remind the notation of the gap $\Delta_k := \mu_{k^*} - \mu_k$.
+        observe that $\mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t)$ means $\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0$, where we remind the notation of the gap $\Delta_k \eqdef \mu_{k^*} - \mu_k$.
         Thus if $n_k \geq \left\lceil \frac{8 \log(T)}{\Delta_k^2} \right\rceil \geq \left\lceil \frac{8 \log(t)}{\Delta_k^2} \right\rceil$,
@@
     Taking the expectation of $N_k(T)$ from \eqref{eq:2:decompositionInUCBProof},
-    and setting $n = n_T := \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil$ thus gives
+    and setting $n = n_T \eqdef \lceil \frac{8 \log(T)}{\Delta_k^2} \rceil$ thus gives
     %
@@
 The indexes are computed as follows, for an exploration function $f(t)$
-that is typically chosen as $f(t) := \log(t)+3\log(\log(t))$ for the theoretical analyses,
-and $f(t) := \log(t)$ in practice.
+that is typically chosen as $f(t) \eqdef \log(t)+3\log(\log(t))$ for the theoretical analyses,
+and $f(t) \eqdef \log(t)$ in practice.
 The constant $c>0$ is usually chosen as $c = 1$.
 \begin{align}\label{eq:klUCB_index}
-    U_k^{\klUCB}(t) := \sup\limits_{q \in [0, 1]} \left\{ q : \KL\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
+    U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \KL\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
 \end{align}
@@
 For problems with Bernoulli distributions on the $K$ arms,
-the best choice of prior is a Beta distribution, that starts as an initially uniform prior (\ie, $\forall k, \pi_k(0) := \mathrm{Beta}(1,1)$),
-and maintain a distribution $\pi_k(t) := \mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$,
+the best choice of prior is a Beta distribution, that starts as an initially uniform prior (\ie, $\forall k, \pi_k(0) \eqdef \mathrm{Beta}(1,1)$),
+and maintain a distribution $\pi_k(t) \eqdef \mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$,
 as $X_k(t)$ counts the number of successes observed on arm $k$, and $N_k(t)-X_k(t)$ counts the number of failures.
@@
 The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as $\UCB_1$ or Thompson sampling, with uniformly distributed perturbations.
-Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence interval $\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) := \widehat{\mu_k}(t) - \xi_k(t)$),
+Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence interval $\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) \eqdef \widehat{\mu_k}(t) - \xi_k(t)$),
 but instead of maximizing the UCB, a random sample is taken on each interval $s_k \sim \cU(\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right])$, and the played arm is the one maximizing this sample, \ie, $A(t) \in \argmax_k s_k$.
@@
 The Exp3 policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
-$\Pr(A(t)=k) := (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
+$\Pr(A(t)=k) \eqdef (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
 The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
@@
 \begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp}
-    U^{\mathrm{MOSS}}_k(t) & := \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)} } \right). \\
-    U^{\KLUCBpp}_k(t) & := \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
+    U^{\mathrm{MOSS}}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)} } \right). \\
+    U^{\KLUCBpp}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
 \end{align}
@@
 \begin{align}\label{eq:2:indexes_for_klUCBswitch}
-    U^{\mathrm{MOSS}+}_k(t) & := \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
-    U^{\KLUCB+}_k(t) & := \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
-    U^{\klUCB\text{-}\mathrm{switch}}_k(t) & := \begin{cases}
+    U^{\mathrm{MOSS}+}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
+    U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
+    U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
         U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) := p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
+    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) \eqdef p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
 
@@
 \begin{smallproof}
-    Fix $p\in[0,1]$, and consider the function $g(x) := \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
+    Fix $p\in[0,1]$, and consider the function $g(x) \eqdef \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
     We first observe that $g(0) = 0$, and we easily prove that $x=0$ is the unique minimizer of $g$ over its domain.

### 2019-06-08 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done, I changed := to \overset{.}{=}

@@
 %
-For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r(t) := Y_{k,t} \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
+For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r(t) \eqdef Y_{k,t} \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
 The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
@@
 
-Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} := \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (which can be computed numerically in an efficient manner, \cite{Corless96}).
+Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} \eqdef \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (which can be computed numerically in an efficient manner, \cite{Corless96}).
 Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done, I changed := to \overset{.}{=}

@@
   \begin{equation}
-    C^j(t) :=  \{ \exists j' \neq j : A^{j'}(t) = A^j(t) \}
-    \ \ \ \text{and} \ \ \ C_k(t) :=  \left\{ \# \{ j : A^j(t) = k\} > 1 \right\},
+    C^j(t) \eqdef  \{ \exists j' \neq j : A^{j'}(t) = A^j(t) \}
+    \ \ \ \text{and} \ \ \ C_k(t) \eqdef  \left\{ \# \{ j : A^j(t) = k\} > 1 \right\},
   \end{equation}
@@
 \begin{equation}
-  r^j(t) := Y_{A^j(t),t} \; \indic(\overline{C^j(t)}).
+  r^j(t) \eqdef Y_{A^j(t),t} \; \indic(\overline{C^j(t)}).
 \end{equation}
@@
   \begin{equation}\label{eq:5:regret}
-    R_T(\boldsymbol{\mu}, M, \cA) := \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
+    R_T(\boldsymbol{\mu}, M, \cA) \eqdef \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
   \end{equation}
@@
   % \label{def:5:nbSelections}
-  Let $T^j_k(T) := \sum_{t=1}^T \indic(A^j(t) = k)$,
-  and denote $T_k(T) := \sum_{j=1}^M T^j_k(T)$ the \emph{number of selections} of arm $k\in[K]$ by any player $j\in[M]$, up to time $T$.
+  Let $T^j_k(T) \eqdef \sum_{t=1}^T \indic(A^j(t) = k)$,
+  and denote $T_k(T) \eqdef \sum_{j=1}^M T^j_k(T)$ the \emph{number of selections} of arm $k\in[K]$ by any player $j\in[M]$, up to time $T$.
   % and $\cC_k(T)$ be the number of collisions on arm $k$ up-to time $T$.
@@
   \begin{equation}
-    \cC_k(T) :=
+    \cC_k(T) \eqdef
     \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(C^j(t)) \indic(A^j(t) = k).
@@
 %
-Letting $\cP_M := \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
+Letting $\cP_M \eqdef \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
 be the set of bandit instances such that there is a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.
@@
 \begin{smallproof}
-  Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
+  Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t) \eqdef \indic(\overline{C^j(t)})$,
   \begin{align*}
@@
       - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
-      % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
+      % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{ \eqdef \overline{\mu}^*} (T M)
     &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
@@
     %
-    \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
+    \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* \eqdef \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
     %
@@
     %
-    \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
+    \intertext{Let $\overline{\Delta_k} \eqdef \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
     %
@@
   %
-  Let $T_k^{!}(T) := \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
+  Let $T_k^{!}(T) \eqdef \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
   where the notation $\exists!$ stands for ``there exists a unique''.
@@
 we use the Kullback-Leibler divergence to express our lower bound.
-We remind that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
+We remind that $\kl(x,y) \eqdef x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
 %
@@
 consider the observations $O_t^j$ that she gathered after $t$ rounds of its algorithm $\cA_j$ for a fixed player $j\in[M]$,
-defined as $O_t^j := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
+defined as $O_t^j \eqdef \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
 where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).
@@
 \begin{equation}\label{eq:5:collisionInformationTerm}
-  \cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}^j}).
+  \cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) \eqdef \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}^j}).
 \end{equation}
@@
 \begin{equation}
-  \widetilde{S_k}^j(t) := \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
-  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) := \widetilde{S_k}^j(t) \;/\; T_k^j(t).
+  \widetilde{S_k}^j(t) \eqdef \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
+  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) \eqdef \widetilde{S_k}^j(t) \;/\; T_k^j(t).
 \end{equation}
@@
 \begin{equation}\label{eq:5:indexTilde}
-  \widetilde{U_k}^j(t) := \begin{cases}
+  \widetilde{U_k}^j(t) \eqdef \begin{cases}
       \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
@@
 %
-denote $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
-then $\widehat{\mu}_k^j(t) := S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
+denote $S_k^j(t) \eqdef \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
+then $\widehat{\mu}_k^j(t) \eqdef S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
 \begin{equation}\label{eq:5:indexFor_UCB_klUCB}
-  U_k^j(t) := \begin{cases}
+  U_k^j(t) \eqdef \begin{cases}
       \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
@@
 % \begin{equation}
-$A^j(t) := \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right)$.
+$A^j(t) \eqdef \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right)$.
 % \end{equation}
@@
 \begin{equation}
-  \TopM(t) := \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
+  \TopM(t) \eqdef \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
 \end{equation}
@@
     & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right) \\
-    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{:= T_{k'}}
+    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{ \eqdef T_{k'}}
   \end{align*}
@@
   To control the second sum, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
-  Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
+  Letting $\kl^+(x, y) \eqdef \kl(x, y) \indic(x \geq y)$,
   and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
@@
   \begin{align*}
-    I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
+    I_1^j(t) & \eqdef \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
     % go sitted
-    I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
+    I_2^j(t) & \eqdef \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
     \ \ \ &\text{and} \ \ \
-    I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
+    I_3(t) & \eqdef \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
     % if sitted
-    I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
+    I_3(t) & \eqdef \left(s^j(t),A^j(t) \in \TopM(t)\right),
     \ \ \ &\text{and} \ \ \
-    I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
+    I_5(t) & \eqdef \left(s^j(t),A^j(t) \notin \TopM(t)\right).
   \end{align*}
@@
   % \begin{equation}
-  $\widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right)$.
+  $\widetilde{C^j}(t) \eqdef \left(C^j(t), \overline{s^j(t)}\right)$.
   % \end{equation}
@@
   Thus, $\sum_{j=1}^M \Pr(C^j(t))$ can be upper bounded by $M \sum_{j=1}^M \Pr(\widetilde{C^j}(t))$ (union bound),
-  and it follows that if $\cC(T) := \sum_{k=1}^K \cC_k(T)$ then
+  and it follows that if $\cC(T) \eqdef \sum_{k=1}^K \cC_k(T)$ then
   \[\E_{\mu}[\cC(T)] \leq M \sum_{j=1}^M \sum_{t=1}^T \Pr(\widetilde{C^j}(t)).\]
@@
   \ie,
-  $n(T) := \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
+  $n(T) \eqdef \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
   (or $0$ if $\theta_1$ does not exist).

### 2019-06-08 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done, I changed := to \overset{.}{=}

@@
 A (random) stream of rewards $(Y_{k,t})_{t\in\N^*}$ is associated to each arm $k \in [K]$. We assume that the rewards are bounded, and without loss of generality we assume that $Y_{k,t} \in [0,1]$.
-We denote by $\mu_k(t) :=  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
+We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
 At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) = Y_{A(t),t}$.
@@
 
-In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists k\in[K]: \mu_k(t) \neq \mu_k(t+1)\right)$.
+In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T \eqdef \sum_{t=1}^{T-1} \indic\left(\exists k\in[K]: \mu_k(t) \neq \mu_k(t+1)\right)$.
 If we denote $\tau^{(0)} = 0$, we define the $i$-th breakpoint recursively by
@@
 We define the \emph{number of change-points} on arm $k$,
-$\NCk := \sum_{t=1}^{T-1} \indic\left(\mu_k(t) \neq \mu_k(t+1)\right)$,
+$\NCk \eqdef \sum_{t=1}^{T-1} \indic\left(\mu_k(t) \neq \mu_k(t+1)\right)$,
 for which it clearly holds that
 $\NCk \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms.
-Letting $C_T := \sum_{k=1}^K \NCk$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
+Letting $C_T \eqdef \sum_{k=1}^K \NCk$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
 We illustrate the two extreme cases in problems $1$ and $2$ presented in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} below.
@@
 \begin{equation}\label{eq:6:firstDefGLRT}
-    \GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \dots, X_n ; \mu_0)},
+    \GLR(n) \eqdef \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \dots, X_n ; \mu_0)},
 \end{equation}
@@
 By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$),
-$\kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$
+$\kl(x,y) \eqdef x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$
 %
@@
 %
-where for $k \leq k'$, $\widehat{\mu}_{k:k'} := \frac{1}{k - k' + 1} \sum\limits_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
+where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
 
@@
     %
-    For a constant $a\in[0,1]$, let $h(x) := a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
+    For a constant $a\in[0,1]$, let $h(x) \eqdef a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
     and we are trying to solve $\sup_{x\in[0,1]} h(x)$.
@@
     \begin{equation}\label{def:6:GLR}
-        \widehat{\tau}_{\delta} := \inf \left\{ n \in \N^* : \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
+        \widehat{\tau}_{\delta} \eqdef \inf \left\{ n \in \N^* : \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
     \end{equation}
@@
 \begin{equation}\label{def:6:function_T}
-    \cT(x) ~:=~ 2 \widetilde{h}\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
+    \cT(x) ~ \eqdef ~ 2 \widetilde{h}\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
 \end{equation}
@@
 \begin{equation}\label{def:6:UCB}
-    \UCB_k(t) := \max \Bigl\{ q\in[0,1] : n_k(t) \times \kl\left(\widehat{\mu}_k(t),q\right) \leq f(t - \tau_k(t)) \Bigr\}.
+    \UCB_k(t) \eqdef \max \Bigl\{ q\in[0,1] : n_k(t) \times \kl\left(\widehat{\mu}_k(t),q\right) \leq f(t - \tau_k(t)) \Bigr\}.
 \end{equation}
@@
 $\forall t \in \{ \tau^{(i-1)}+1, \dots, \tau^{(i)} \}, \mu_k(t) = \mu_k^{(i)}$.
-We also introduce $i^* = \arg\max_k \mu_k^{(i)} $ and the largest gap at break-point $i$ as $\Delta^{(i)} := \max_{k\in[K]} |\mu_k^{(i)} - \mu_k^{(i-1)}| >0$.
+We also introduce $i^* = \arg\max_k \mu_k^{(i)} $ and the largest gap at break-point $i$ as $\Delta^{(i)} \eqdef \max_{k\in[K]} |\mu_k^{(i)} - \mu_k^{(i-1)}| >0$.
 
@@
 \begin{eqnarray*}
-    n_k(\tau^{(i)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} := \cI_k.
+    n_k(\tau^{(i)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} \eqdef \cI_k.
 \end{eqnarray*}

### 2019-06-08 2-Chapters/2-Chapter/chapter2.tex
Lots of small typos.

@@
 enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, that is, matching the $C_I \ln(T)$ lower bound by \cite{LaiRobbins85}.
-It is an index policy that uses a modified version of the indexes of the \klUCB$^+$ algorithm, for arms that are not sampled much, and a modified version of the indexes of the MOSS$^+$ algorithm, for the other arms.
+It is an index policy that uses a modified version of the indexes of the \klUCB$^+$ algorithm (see \cite{Honda2019} for more details), for arms that are not sampled much, and a modified version of the indexes of the MOSS$^+$ algorithm, for the other arms.
 To detail this, first let $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Lots of small typos.

@@
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-  % [$T_k(T)$ and $\cC_k(T)$]
+  % [$N_k(T)$ and $\cC_k(T)$]
   % \label{def:5:nbSelections}
-  Let $T^j_k(T) \eqdef \sum_{t=1}^T \indic(A^j(t) = k)$,
-  and denote $T_k(T) \eqdef \sum_{j=1}^M T^j_k(T)$ the \emph{number of selections} of arm $k\in[K]$ by any player $j\in[M]$, up to time $T$.
+  Let $N_k^j(T) \eqdef \sum_{t=1}^T \indic(A^j(t) = k)$,
+  and denote $N_k(T) \eqdef \sum_{j=1}^M N_k^j(T)$ the \emph{number of selections} of arm $k\in[K]$ by any player $j\in[M]$, up to time $T$.
   % and $\cC_k(T)$ be the number of collisions on arm $k$ up-to time $T$.
@@
       R_T(\boldsymbol{\mu}, M, \cA) &=
-      \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[T_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}} \\
+      \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[N_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}} \\
       & \;\;\;\;\;\;
-      + \;\; \underbrace{\sum_{k \in \Mbest} (\mu_k -  \mu_M^*) (T - \E_{\mu}[T_k(T)])}_{\emph{\mytag{(b)}{eq:5:term2}}}
+      + \;\; \underbrace{\sum_{k \in \Mbest} (\mu_k -  \mu_M^*) (T - \E_{\mu}[N_k(T)])}_{\emph{\mytag{(b)}{eq:5:term2}}}
       + \;\; \underbrace{\sum_{k=1}^{K} \mu_k \E_{\mu}[\cC_k(T)]}_{\emph{\mytag{(c)}{eq:5:term3}}}. \nonumber
@@
 \begin{smallproof}
-  Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t) \eqdef \indic(\overline{C^j(t)})$,
+  Using the definition of regret $R_T(\boldsymbol{\mu}, M, \cA)$ from \eqref{eq:5:regret}, denoted here $R_T$, and this collision indicator $\eta^j(t) \eqdef \indic(\overline{C^j(t)})$,
   \begin{align*}
-    R(T)
+    R_T
     &= \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M Y_{A^j(t),t} \eta^j(t) \right]
@@
 And so}
-    R(T)
+    R_T
     &= \E_{\mu}\left[ \sum_{t=1}^{T}\sum_{j \in \Mbest} \mu_j
@@
     &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
-      - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ T^j_k(T) \right]
+      - \sum_{k=1}^{K} \sum_{j=1}^{M} \mu_k \E_{\mu}\left[ N_k^j(T) \right]
       + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
     %
-    \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* \eqdef \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
-    %
-    &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ T^j_k(T) \right]
+    \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ N_k^j(T)\right]$, and if we denote $\overline{\mu}^* \eqdef \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
+    %
+    &= \sum_{k=1}^{K} \sum_{j=1}^{M} (\overline{\mu}^* - \mu_k) \E_{\mu}\left[ N_k^j(T) \right]
       + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
@@
     %
-    &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
-      + \overline{\Delta_{M^*}} \E_{\mu}\left[ T_{M^*}(T) \right] \\
-      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ T_k(T) \right]
+    &= \sum_{k \in \Mbest\setminus\{M\}} \overline{\Delta_k} \E_{\mu}\left[ N_k(T) \right]
+      + \overline{\Delta_{M^*}} \E_{\mu}\left[ N_{M^*}(T) \right] \\
+      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} \overline{\Delta_k} \E_{\mu}\left[ N_k(T) \right]
       + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
     %
-    \intertext{But for $k = M^*$, $T_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ T_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ T_k(T) \right]$, so by recombining the terms, we obtain,}
-    %
-    &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
+    \intertext{But for $k = M^*$, $N_{M^*}(T) = T M^* - \sum\limits_{k \in \Mbest\setminus\{M\}} \E_{\mu}\left[ N_k(T) \right] - \sum\limits_{k \in \Mworst} \E_{\mu}\left[ N_k(T) \right]$, so by recombining the terms, we obtain,}
+    %
+    &= \sum_{k \in \Mbest\setminus\{M\}} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ N_k(T) \right]
       + \overline{\Delta_{M^*}} T M^* \\
-      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ T_k(T) \right]
+      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\overline{\Delta_k} - \overline{\Delta_{M^*}}) \E_{\mu}\left[ N_k(T) \right]
       + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
@@
   \begin{align*}
-    R(T)
-    &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+    R_T
+    &= \sum_{k \in \Mbest} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ N_k(T) \right]
       + \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) T \\
-      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+      &\;\;\;\;\;\;\;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ N_k(T) \right]
       + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right]
@@
   \begin{align*}
-  R(T)
-    & = \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ T_k(T) \right]\right) \\
-      & \;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ T_k(T) \right]
+  R_T
+    & = \sum_{k \in \Mbest} (\mu_k - \mu_{M^*}) \left(T - \E_{\mu}\left[ N_k(T) \right]\right) \\
+      & \;\; + \sum_{k \in \Mworst} (\mu_{M^*} - \mu_k) \E_{\mu}\left[ N_k(T) \right]
       + \sum_{k=1}^{K} \mu_k \E_{\mu}\left[ \cC_k(T) \right].
@@
 The regret decomposition in Lemma~\ref{lem:5:DecompositionRegret} is valid for both centralized and decentralized algorithms.
-For centralized algorithms, due to the absence of collisions, \ref{eq:5:term3} is obviously zero, and \ref{eq:5:term2} is non-negative, as $T_k(T) \leq T$.
-For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $T_k(T) > T$ (which is counter intuitive with such notations).
+For centralized algorithms, due to the absence of collisions, \ref{eq:5:term3} is obviously zero, and \ref{eq:5:term2} is non-negative, as $N_k(T) \leq T$.
+For decentralized algorithms, \ref{eq:5:term3} may be significantly large, and term \ref{eq:5:term2} may be negative, as many collisions on arm $k$ may lead to $N_k(T) > T$ (which is counter intuitive with such notations).
 However, a careful manipulation of this decomposition shows that the regret is always lower bounded by term \ref{eq:5:term1}.
@@
     \begin{align*}
-        R_T(\boldsymbol{\mu}, M, \cA)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[T_k(T)].
+        R_T(\boldsymbol{\mu}, M, \cA)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[N_k(T)].
     \end{align*}
@@
   Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
-  but it is not obvious for \ref{eq:5:term2} as there is no reason for $T_k(T)$ to be upper bounded by $T$.
-  %
-  Let $T_k^{!}(T) \eqdef \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
+  but it is not obvious for \ref{eq:5:term2} as there is no reason for $N_k(T)$ to be upper bounded by $T$.
+  %
+  Let $N_k^{!}(T) \eqdef \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
   where the notation $\exists!$ stands for ``there exists a unique''.
-  Then $T_k(T) = \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(A^j(t) = k)$ can be decomposed as
+  Then $N_k(T) = \sum_{t=1}^{T} \sum_{j=1}^{M} \indic(A^j(t) = k)$ can be decomposed as
   \begin{equation*}
-    T_k(T) = \sum_{t=1}^{T} \indic(\exists! j, A^j(t) = k) + \sum_{t=1}^{T} \sum_{j=1}^{M} c_{k,t} \indic(A^j(t) = k)
-    = T_k^{!}(T) + C_k(T).
+    N_k(T) = \sum_{t=1}^{T} \indic(\exists! j, A^j(t) = k) + \sum_{t=1}^{T} \sum_{j=1}^{M} c_{k,t} \indic(A^j(t) = k)
+    = N_k^{!}(T) + C_k(T).
   \end{equation*}
@@
     \ref{eq:5:term2} + \ref{eq:5:term3} &=
-    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[N_k^!(T)])
     + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)] \\
@@
     &=
-    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[N_k^!(T)])
     + \sum_{k \in \Mbest} \mu_M^* \E_{\mu}[C_k(T)]
@@
     &=
-    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[T_k^!(T)])
+    \sum_{k \in \Mbest} (\mu_k - \mu_M^*) (T - \E_{\mu}[N_k^!(T)])
     + \sum_{k=1}^{M} \min(\mu_M^*, \mu_k) \E_{\mu}[C_k(T)].
   \end{align*}
-  And now both terms are non-negative, as $T_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
+  And now both terms are non-negative, as $N_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
   which proves that $R_T(\boldsymbol{\mu}, M, \cA) \geq \ref{eq:5:term1}$, as wanted.
@@
     \frac{T}{M}
-    - \E_{\mu}[T^j_k(T)] \mathop{=}\limits_{T \to +\infty} o(T^{\alpha}).
+    - \E_{\mu}[N_k^j(T)] \mathop{=}\limits_{T \to +\infty} o(T^{\alpha}).
     \label{eq:5:SUE}
@@
 From this assumption $\left(R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
-${T}- \E_{\mu}[T_k(T)] {=} o(T^{\alpha})$, and so
+${T}- \E_{\mu}[N_k(T)] {=} o(T^{\alpha})$, and so
 \begin{equation}\label{eq:5:intermediate}
   % \vspace*{-5pt}  % XXX
-  \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[T^j_k(T)]\right) = o(T^{\alpha}).
+  \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[N_k^j(T)]\right) = o(T^{\alpha}).
   % \vspace*{-5pt}  % XXX
@@
 Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\cA_{\sigma} = (\cA_{\sigma(1)},\dots,\cA_{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$ of the set $\{1,\dots,M\}$.
-In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
+In that case, it holds that  $\E_{\mu}[N_k^j(T)]=\E_{\mu}[N_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
 Note that the algorithms studied in Section~\ref{sec:5:algorithms} are permutation invariant, and \MCTopM-\klUCB{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.
@@
   \begin{equation}\label{eq:5:LBDraws}
-    \forall j \in [M], \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
+    \forall j \in [M], \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[N_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
   \end{equation}
@@
 %
-It is worth mentioning that \cite{Zhao10} considered the more general assumption for $\cA$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
+It is worth mentioning that \cite{Zhao10} considered the more general assumption for $\cA$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[N_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
 Our result could be extended to this case, but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.
@@
   \begin{equation}\label{eq:5:1stUpperBound}
-    \ref{eq:5:term2} \leq (\mu_1^* - \mu_M^*) \Bigl(    \sum_{k \in \Mworst} \E_{\mu}[T_k(T)]
+    \ref{eq:5:term2} \leq (\mu_1^* - \mu_M^*) \Bigl(    \sum_{k \in \Mworst} \E_{\mu}[N_k(T)]
     + \sum_{k \in \Mbest} \E_{\mu}[C_{k}(T)]
@@
   Recall that we want to upper bound
-  $ \ref{eq:5:term2} : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[T_k(T)]\right)$.
+  $ \ref{eq:5:term2} : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[N_k(T)]\right)$.
   First, we observe that, for all $k\in \Mbest$,
   \begin{eqnarray*}
-    T - \E_{\mu}[T_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
+    T - \E_{\mu}[N_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
     & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
@@
   \begin{eqnarray*}
-  \ref{eq:5:term2} & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[T_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
+  \ref{eq:5:term2} & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[N_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
   & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
@@
   which yields
-  \[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[T_k(T)\right]\]
+  \[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[N_k(T)\right]\]
   and Lemma~\ref{lem:5:1stUpperBound} follows.
@@
 This result can also be used to recover Proposition~1 from \cite{Anandkumar11}, giving an upper bound on the regret that only depends on
-the \emph{expected number of sub-optimal selections} -- $\E_{\mu}[T_k(T)]$ for $k \in \Mworst$ --
+the \emph{expected number of sub-optimal selections} -- $\E_{\mu}[N_k(T)]$ for $k \in \Mworst$ --
 and the \emph{expected number of colliding players on the optimal arms} -- $\E_{\mu}[\cC_k(T)]$ for $k \in \Mbest$.
-Note that, in term (c) the number of colliding players on the sub-optimal arm $k$ may be upper bounded as $\E_{\mu}[\cC_k(T)] \leq M \E_{\mu}[T_k(T)]$.
+Note that, in term (c) the number of colliding players on the sub-optimal arm $k$ may be upper bounded as $\E_{\mu}[\cC_k(T)] \leq M \E_{\mu}[N_k(T)]$.
 %
@@
   \widetilde{S_k}^j(t) \eqdef \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
-  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) \eqdef \widetilde{S_k}^j(t) \;/\; T_k^j(t).
+  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) \eqdef \widetilde{S_k}^j(t) \;/\; N_k^j(t).
 \end{equation}
@@
   \widetilde{U_k}^j(t) \eqdef \begin{cases}
-      \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
+      \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2N_k^j(t))}
       &\text{for } \UCB, \\
-      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widetilde{\mu_k}^j(t), q) \leq f(t) \right\}
+      \sup\left\{ q \in [0, 1]: N_k^j(t)\times \kl(\widetilde{\mu_k}^j(t), q) \leq f(t) \right\}
       &\text{for } \klUCB.
@@
 denote $S_k^j(t) \eqdef \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
-then $\widehat{\mu}_k^j(t) \eqdef S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
+then $\widehat{\mu}_k^j(t) \eqdef S_k^j(t)/N_k^j(t)$ (if $N_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
 \begin{equation}\label{eq:5:indexFor_UCB_klUCB}
   U_k^j(t) \eqdef \begin{cases}
-      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
+      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2N_k^j(t))}
       &\text{for } \UCB, \\
-      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
+      \sup\left\{ q \in [0, 1]: N_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
       &\text{for } \klUCB,
@@
   \begin{equation}\label{eq:5:SubOptimalSelections}
-      \E_{\mu}[T_k^j(T)] \leq
+      \E_{\mu}[N_k^j(T)] \leq
       \frac{\log(T)}{\kl(\mu_k,\mu_{M}^*)}
@@
   \begin{align*}
-  \E_{\mu}[T_k^j(T)]
+  \E_{\mu}[N_k^j(T)]
   &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
@@
   \begin{align*}
-    \E_{\mu}[T_k^j(T)] &\leq
+    \E_{\mu}[N_k^j(T)] &\leq
     \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; U_{m_1^*}(t) < \mu_m^* \right) + \\
@@
   \begin{align}
-    \E_{\mu}[T_k^j(T)] &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq U_k(t)\right) \notag \\
+    \E_{\mu}[N_k^j(T)] &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq U_k(t)\right) \notag \\
     %
@@
   \begin{align}\label{eq:5:UBprecise}
-    \E_{\mu}\left[T_k^j(T)\right]
+    \E_{\mu}\left[N_k^j(T)\right]
     & \leq \frac{\log(T) + 3 \log(\log(T))}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log(\log(T))}  \nonumber\\
@@
     & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right) \\
-    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{ \eqdef T_{k'}}
+    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{ \eqdef N_{k'}}
   \end{align*}
 
-  Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
-  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right)\]
+  Now, to control $N_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
+  \[N_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right)\]
   The first sum is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}.
@@
   From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
-  \[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]
+  \[N_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]
 
   If $\mu_k > \mu_{k'}$, we rather use that
-  $T_{k'} \leq \sum\limits_{t=1}^T \Pr\left(U_{k}^j(t) \leq \mu_{k}\right) + \sum\limits_{t=1}^T \Pr\left(A^j(t+1) = k', U_{k'}^j(t) \geq \mu_{k}\right)$,
+  $N_{k'} \leq \sum\limits_{t=1}^T \Pr\left(U_{k}^j(t) \leq \mu_{k}\right) + \sum\limits_{t=1}^T \Pr\left(A^j(t+1) = k', U_{k'}^j(t) \geq \mu_{k}\right)$,
   and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
-  $T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T))$.
+  $N_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T))$.
   Summing over $k'$ yields the result.

### 2019-06-08 2-Chapters/5-Chapter/chapter5.tex
Done, I changed := to \overset{.}{=}
Lots of small typos.

@@
 \begin{equation}\label{eq:5:defRegretDynamicSetting}
-    R_T^{\text{dyn}} := \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
+    R_T^{\text{dyn}} \eqdef \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
 \end{equation}
@@
 Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
-However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB, as its regret scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}.\\
+However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (while scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).\\
 \indent

### 2019-06-08 5-Nomenclature/Nomenclature.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment
Lots of small typos.

@@
 \nomenclature[G]{$\gamma$}{Usually denotes a discount factor, \eg, in D-\UCB{} or D-TS in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, the parameter for the $\varepsilon$-greedy algorithm, or a drift correction parameter for the \CUSUM{} algorithm in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, the parameter for the $\varepsilon$-greedy algorithm, or the drift correction parameter for \CUSUM{} in Chapter~\ref{chapter:6}}
 \nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
@@
 \nomenclature[L]{$Y_{k,t}$}{Random sample from the arm $k$ at time $t$}
-\nomenclature[L]{$\cO_t$}{Vector of observations until time $t$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$O_t$}{Vector of observations until time $t$ in Chapter~\ref{chapter:5}}
 \nomenclature[L]{$r(t)$, $r^j(t)$}{Reward obtained at time $t$, for user $j$ at time $t$}
@@
 \nomenclature[L]{$k$}{Usually denotes the $k$-th arm, $k\in[K]$, mainly used in subscripts}
-\nomenclature[L]{$M$}{In Chapter~\ref{chapter:4}, denote the maximum number of retransmission of a packet in the ALOHA protocol, and in Chapter~\ref{chapter:5}, denote the number of player for multi-players bandit games}
-\nomenclature[L]{$m$}{In Chapter~\ref{chapter:4}, denote the maximum length of the back-off interval after a collision}
+\nomenclature[L]{$M$}{Maximum number of retransmission of a packet in the ALOHA protocol in Chapter~\ref{chapter:4}, and number of player for multi-players bandit games in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$m$}{Maximum length of the back-off interval after a collision, in Chapter~\ref{chapter:4}}
 \nomenclature[L]{$i$, $j$}{Usually denotes the $i$-th or $j$-th player, $i,j\in[M]$ in Chapter~\ref{chapter:5}}
@@
 
-\nomenclature[L]{$N_k(t)$ or $T_k(t)$}{Number of samples obtained for arm $k$ at time $t$ ($N_k(t)$ in Chapter~\ref{chapter:4}, $T_k(t)$ in Chapter~\ref{chapter:5})}
+\nomenclature[L]{$N_k(t)$}{Number of samples obtained for arm $k$ at time $t$}
 \nomenclature[L]{$\widehat{\mu}_k(t)$}{Empirical mean of rewards obtained for arm $k$ at time $t$}
@@
 \nomenclature[L]{$p$}{In Chapter~\ref{chapter:4}, probability of transmission for the devices following a Bernoulli random emission pattern (\eg, $p=10^{-5}$)}
-\nomenclature[L]{$S$, $S_k$}{In Chapter~\ref{chapter:4}, total number of \emph{static} devices in the network, resp. in channel $k$}
-\nomenclature[L]{$D$, $D_k$}{In Chapter~\ref{chapter:4}, total number of \emph{dynamic} devices in the network, resp. in channel $k$}
+\nomenclature[L]{$S$, $S_k$}{In Chapter~\ref{chapter:4}, total number of \emph{static} devices in the network or in channel $k$}
+\nomenclature[L]{$D$, $D_k$}{In Chapter~\ref{chapter:4}, total number of \emph{dynamic} devices in the network or in channel $k$}
 
-\nomenclature[L]{$C_k(t)$, $C^j(t)$}{In Chapters~\ref{chapter:4} and \ref{chapter:5}, collision indicator, on arm $k\in[K]$ or for user $j\in[M]$, at time $j$}
+\nomenclature[L]{$C_k(t)$, $C^j(t)$}{In Chapter~\ref{chapter:5}, collision indicator on arm $k\in[K]$ or for user $j\in[M]$, at time $t$}
 
@@
 \nomenclature[I]{$x_k$}{Usually denotes a variable depending on an arm, for $k\in[K]$}
-\nomenclature[I]{$Y_{k,t}$}{Usually denotes a variable depending on an arm, for $k\in[K]$, and on time, for $t\in[T]$}
+\nomenclature[I]{$Y_{k,t}$}{Usually denotes a variable depending on an arm $k\in[K]$ and on time $t\in[T]$}
 

### 2019-06-09 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done, complete proof-reading of Chapter:4

@@
 
-As explained before, the future Internet of Things (IoT) networks will require to support more and more communicating devices.
-In this section, we show that intelligent devices in unlicensed bands can use Multi-Armed Bandit (MAB) learning algorithms to improve spectral resource exploitation.
+As explained before in Chapter~\ref{chapter:1}, the future IoT networks will require to support more and more communicating devices.
+In this section, we show that intelligent devices in unlicensed bands can use MAB learning algorithms to improve spectral resource exploitation.
 %
@@
 The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
-and the network is less and less stationary.
+and the network usage is more and more fluctuatating.
 % and the stochastic hypothesis is more and more violated.
@@
 
-\textbf{Outline.}
-%
-The rest of this section is organized as follows.
+% \textbf{Outline.}
+% %
+% The rest of this section is organized as follows.
 The system model is introduced in Section~\ref{sub:41:systemModel}.
 The reference policies are described in \ref{sub:41:threeReferencePolicies}, and our proposed sequential learning policies are introduced in \ref{sub:41:sequentialPolicies}.
-Numerical results are presented in \ref{sub:41:numericalResults}.
+% Numerical results are presented in \ref{sub:41:numericalResults}.
 
@@
 The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
-We mean here by ALOHA-based that devices do not use sensing, and we do not consider retransmissions of packets (we remove this hypothesis in Section~\ref{sec:4:retransmissions}).
+We mean here by ALOHA-based that \emph{devices do not use sensing}, and \emph{we do not consider retransmissions of packets} (we relax this hypothesis in Section~\ref{sec:4:retransmissions}).
 %
@@
     \includegraphics[width=0.70\linewidth]{system_model1.eps}
-    \caption[In our system model, some dynamic devices (in the IoT network in blue) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in orange).]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
+    \caption[In our system model, some dynamic devices (in the IoT network in blue) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in orange).]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange} left/right).}
     \label{fig:41:system_model1}
@@
 
-These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated co-localized and/or neighboring networks (in \textcolor{orange}{orange} in Figure~\ref{fig:41:system_model1}).
+These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated co-localized and/or neighboring networks (two are shown in \textcolor{orange}{orange} in Figure~\ref{fig:41:system_model1}, in left or right).
 This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels,
@@
 In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
-If two or more devices send an uplink packet in the same slot, the uplink packets collide, and the acknowledgement \Ack{} is not transmitted.
+If two or more devices send an uplink packet in the same slot, the uplink packets collide, and the acknowledgement (\Ack) is not transmitted.
 In this case, we say that there is a \emph{collision} in this time slot.
@@
 
-\begin{figure}[!t]
+\begin{figure}[!h]
     \centering
@@
 
-\paragraph{Static vs dynamic devices.}
+\textbf{Static vs dynamic devices.}
 %
@@
 
-\paragraph{Random emission patterns.}
+\textbf{Random emission patterns:}
 %
@@
 
-\paragraph{Some assumptions on the network occupation.}
+\textbf{Some assumptions on the network occupation:}
 %
@@
 %
-We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for IoT networks.
+We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for realistic IoT networks.\\
 % \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
@@
 Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
-Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to have an average of active devices not larger than the number of channels.
+Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to \emph{have an average of active devices not larger\footnote{~In Chapter~\ref{chapter:5} we simply consider the case of $p=1$ and $D \leq K$ dynamic devices.} than the number of channels}.
 
-\paragraph{Link with the MAB model.}
+\textbf{Link with the MAB model:}
 %
@@
 
-\textbf{$1^{\text{st}}$ - Naive policy: Random Channel Selection} ``\textcolor{darkgreen}{Random}''.
+\paragraph{$1^{\text{st}}$ - Naive policy: Random Channel Selection} (``\textcolor{darkgreen}{Random}'')
 %
@@
 so probability of successful transmission is computed as follows
+%
 \begin{equation}
@@
 \end{align}
+%
 This expression \eqref{eq:41:strategynaive} is constant (in time), and easy to compute numerically, but comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it
@@
 
-\textbf{$2^{\text{nd}}$ - (Unachievable) Optimal oracle policy} ``\textcolor{orange}{Optimal}''.
+\paragraph{$2^{\text{nd}}$ - (Unachievable) Optimal oracle policy} (``\textcolor{orange}{Optimal}'')
 %
@@
 then this probability is computed as before:
+%
 \begin{align}\label{eq:41:prob_col}
@@
     \begin{equation}\label{eq:41:Dilambda}
-        D_k^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_k-1}} \right)-1 \right]\right)^{+} .
+        D_k^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda \e}{(1-p)^{S_k-1}} \right)-1 \right]\right)^{+} .
     \end{equation}
@@
 \end{proposition}
-
+%
 \begin{smallproof}
@@
 Where in the equation~\eqref{eq:41:Dilambda}, $(a)^{+} \eqdef \max(a,0)$, and $\mathcal{W}$ denotes the $\mathcal{W}$-Lambert function which is the reciprocal bijection of $x \mapsto x e^x$ on $\mathbb{R^+} = [0, +\infty)$ (which can be computed numerically in an efficient manner, \cite{Corless96}).
-Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint
-\begin{equation}\label{eq:41:constraintLambda}
-    \sum_{k=1}^{K} D_k^*(\lambda) = D.
-\end{equation}
-
-Equation \eqref{eq:41:constraintLambda} can be solved numerically, with simple one-dimensional root finding algorithms.
+Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint $\sum_{k=1}^{K} D_k^*(\lambda) = D$.
+%
+This single constraint can be solved numerically, with simple one-dimensional root finding algorithms.
 Solving the optimization problem provides the optimal real number value for $D_k^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $k$:
@@
 
-\textbf{$3^{\text{rd}}$ - A greedy approach of the oracle strategy} ``\textcolor{deeppurple}{Good sub-optimal}''.
+\paragraph{$3^{\text{rd}}$ - A greedy approach of the oracle strategy} (``\textcolor{deeppurple}{Good sub-optimal}'')
 %
@@
 
-While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy caused partly by the learning policy used by dynamic devices, but mainly by their random activation processes.
-
+While the stochastic MAB model has been used to describe some aspects of Cognitive Radio systems, it is in principle not suitable for our IoT model, due to the non-stationarity of the channels occupancy, caused partly by the learning policy used by dynamic devices, but mainly by their random activation processes.
+%
 In our model, every dynamic device implements its own learning algorithm, \emph{independently}.
@@
 Our IoT application is challenging in that there are \emph{multiple} players (the dynamic devices) interacting with the \emph{same} arms (the channels), without any centralized communication (they do not even know the total number of dynamic devices).
-
+%
 % Considered alone, each dynamic device
@@
 
-\paragraph{\emph{Adversarial} bandit algorithms?}
-
+\textbf{\emph{Adversarial} bandit algorithms?}
+%
 Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
@@
 therefore we did not report results here.
-% FIXME maybe we should... ?
-%(see Section \ref{sub:41:numericalResults}).
-However, contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary model, as its performance guarantees are true in \emph{any} setting.
+% % FIXME maybe we should... ?
+% %(see Section \ref{sub:41:numericalResults}).
+% However, contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary model, as its performance guarantees are true in \emph{any} setting.
 % But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
@@
 We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions.
+
 Clearly, the problem is less interesting for a homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
@@
 \paragraph{First results.}
-
+%
 Figure \ref{fig:41:from10to100} presents the successful transmission rate as a function of time.

### 2019-06-09 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Done, complete proof-reading of Chapter:4

@@
 
-In this demo, we can generate scenarios with different parameters (number of channels, interfering traffic, etc) in order to evaluate the performance of learning in various settings.
+We can generate scenarios with different parameters (number of channels, interfering traffic, etc) in order to evaluate the performance of learning in various settings.
 Moreover, we compare the performance of learning strategies with that of the random uniform access to channels, which is the current state-of-the-art of commercial LPWAN solutions \cite{Raza17}.
@@
 \paragraph{Details about our implementation.}
-
+%
 We used the GNU Radio Companion software (GRC, version $3.7$ in $2017$),
@@
 We compare in this PoC the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
-For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
+For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (``\textcolor{green}{TS}'', in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
 
@@
 
-In Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} above, we show the results averaged on $10$ repetitions using the same conditions.
+In Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} below, we show the results averaged on $10$ repetitions using the same conditions.
 %

### 2019-06-09 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done, complete proof-reading of Chapter:4

@@
 
-In this section, we extend the model presented before, to take into account the possibility for retransmissions of a message after a collision.
-This extension is of major importance: most protocols for real-world IoT networks can use retransmissions, it is for instance the case of the LoRaWAN standard.
+In this section, we extend the previous model to take into account the possibility for retransmissions of a message after a collision.
+This is of major importance as most protocols for real-world IoT networks can use retransmissions, it is for instance the case of the LoRaWAN standard.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.
@@
 
-\textbf{Outline.}
-%
-The rest of the section is organized as follows.
-First the system model is introduced in Section~\ref{sub:43:model},
-and our motivations are exposed in Section~\ref{sub:43:motivations}.
-The proposed \UCB-based heuristics are presented in Section~\ref{sub:43:heuristics}, while the corresponding numerical results are shown in Section~\ref{sub:43:numExp}.
-Finally, some conclusions are drawn later in Section~\ref{sub:43:conclusion}.
+% \textbf{Outline.}
+% %
+% The rest of the section is organized as follows.
+% First the system model is introduced in Section~\ref{sub:43:model},
+% and our motivations are exposed in Section~\ref{sub:43:motivations}.
+% The proposed \UCB-based heuristics are presented in Section~\ref{sub:43:heuristics}, while the corresponding numerical results are shown in Section~\ref{sub:43:numExp}.
+% Finally, some conclusions are drawn later in Section~\ref{sub:43:conclusion}.
 
@@
 % 	\begin{algorithm}[H]
-% 		% \begin{small} % XXX remove if needed
+		% \begin{small} % XXX remove if needed
 % 		\For(){$t = 1, \dots, T$}{
@@
 % 		\label{algo:43:UCB}
-% 		% \end{small} % XXX remove if needed
+		% \end{small} % XXX remove if needed
 % 	\end{algorithm}
@@
 % ----------------------------------------------------------------------
-\subsection{The first heuristics: \UCB{} unaware of retransmissions}
+\subsection{The first heuristic: \UCB{} unaware of retransmissions}
 \label{sub:43:UCBnaive}
@@
 %
-Note that a device using this first approach is only able to select a channel for the first transmission (using \UCB), and it uses the same channel for all the corresponding retransmissions of a packet (if retransmissions happen).
+Note that a device using this first approach is only able to select a channel for the first transmission (using \UCB, line 3-5), and then it uses the same channel for all the corresponding retransmissions of a packet (if retransmissions happen, line 7-8).
 
@@
 		}
-		\caption{First-stage \UCB{} and retransmission in same channel.}
+		\caption[First-stage \UCB{} and retransmission in same channel.]{First-stage \UCB{} and retransmission in same channel (``\textcolor{blue}{Only \UCB{}}'').}
 		\label{algo:43:UCB}
@@
 
-More formally, for one device, let $N_k(t)$ be the number of times the channel $k$ (for $k\in [K]$) was selected up-to time $t-1$, for $t\geq 0$
+More formally, for one device, let $N_k(t)$ be the number of times the channel $k$ (for $k\in [K]$) was selected up-to time $t-1$, for $t\geq1$
 for any $t\in\mathbb{N}$,
-$N_k(t) = \sum_{\tau=0}^{t-1} \mathbbm{1}(A(\tau) = k)$.
-The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as the mean reward obtained up-to time $t-1$,
-$\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=0}^{t-1} r(\tau) \mathbbm{1}(A(\tau) = k)$,
+$N_k(t) = \sum_{\tau=1}^{t-1} \mathbbm{1}(A(\tau) = k)$.
+The empirical mean estimator $\widehat{\mu_k}(t)$ of channel $k$ is defined as $\widehat{\mu_k}(t) = \frac{1}{N_k(t)} \sum_{\tau=1}^{t-1} r(\tau) \mathbbm{1}(A(\tau) = k)$,
 where $r(t)=Y_{k,t}$ is the reward obtained after transmission in channel $k$ at time $t$.
@@
 Finally, the transmission channel at time step $t$
-is the one maximizing this \UCB{} index $U_k(t)$,
-as it is the one expected to be the best one at the current time step $t$,
-$A(t) = \arg\max_k U_k(t)$.
+is $A(t) \sim \cU(\argmax_k U_k(t))$.
+% the one maximizing this \UCB{} index $U_k(t)$,
+% as it is the one expected to be the best one at the current time step $t$,
 
@@
 	%
-	In all the algorithms and notations presented in this Section, the time steps $t$ do not refer to the global time steps (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
+	In all the algorithms and notations presented in this Section, the time steps $t$ do not refer to the \emph{global} time steps (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
 	Each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $M$ times (\eg, $M=10$) until it receives an \Ack.
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Uniform random retransmission ``\textcolor{cyan}{(Random)''}.}\label{sub:43:UCBthenRandom}
-
+\textbf{Uniform random retransmission ``(\textcolor{cyan}{Random}'').}\label{sub:43:UCBthenRandom}
+%
 In this first proposal, the device uses a random channel selection, following a uniform distribution (on $[K]$).
@@
 		}
-		\caption{Uniform random retransmission.}    % A naive
+		\caption[Heuristic: uniform random retransmission.]{Heuristic: uniform random retransmission ``(\textcolor{cyan}{Random}'').}    % A naive
 		\label{algo:43:UCBthenRandom}
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{\UCB{} for retransmission (``\textcolor{blue}{Only \UCB{}}'').}\label{sub:43:TwoUCB}
-
+\textbf{\UCB{} for retransmission (``\textcolor{purple}{\UCB{}}'').}\label{sub:43:TwoUCB}
+%
 Instead of applying a random channel selection for retransmission,
@@
 		}
-		\caption{\UCB{} for retransmission.}    % A naive
+		\caption[Heuristic: \UCB{} for retransmission.]{Heuristic: \UCB{} for retransmission (``\textcolor{purple}{\UCB{}}'').}    % A naive
 		\label{algo:43:TwoUCB}
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{$K$ different {\UCB}s for retransmission (``\textcolor{green}{$K$ \UCB}'')}\label{sub:43:UCBthenKp1}
-
+\textbf{$K$ different {\UCB}s for retransmission (``\textcolor{green}{$K$ \UCB}'')}\label{sub:43:UCBthenKp1}
+%
 Another heuristic is to not use the same algorithm no matter where the collision occurred, but to use $K$ different second-stage \UCB{} algorithms.
@@
 		}
-		\caption{$K$ different {\UCB}s for retransmission.}
+		\caption[Heuristic: $K$ different {\UCB}s for retransmission.]{Heuristic: $K$ different {\UCB}s for retransmission (``\textcolor{green}{$K$ \UCB}'').}
 		\label{algo:43:UCBthenKp1}
@@
 % --- --- --- --- --- --- --- --- --- ---
-\paragraph{Delayed \UCB{} for retransmission (``\textcolor{red}{Delayed \UCB}'').}\label{sub:43:UCBwithDelay}
-
+\textbf{Delayed \UCB{} for retransmission (``\textcolor{red}{Delayed \UCB}'').}\label{sub:43:UCBwithDelay}
+%
 This last heuristic is a composite of
@@
 		}
-		\caption{Delayed \UCB{} for retransmission.}
+		\caption[Heuristic: delayed \UCB{} for retransmission.]{Heuristic: delayed \UCB{} for retransmission (``\textcolor{red}{Delayed \UCB}'').}
 		\label{algo:43:UCBwithDelay}
@@
 
-\paragraph{Other ideas that we did not explore.}
+\textbf{Other ideas that we did not explore.}
 %

### 2019-06-09 2-Chapters/4-Chapter/chapter4.tex
Almost done for ch5
Done, complete proof-reading of Chapter:4

@@
 
-\TODOL{Abstract ici et plus court, et petite introduction !}
-\TODOL{Intégrer les derniers commentaires de Christophe}
+% \TODOL{Intégrer les derniers commentaires de Christophe}
 
+% FIXME finish !
 \paragraph{Abstract.}
+%
+In this Chapter~\ref{chapter:4}, we propose two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
+that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
+%
+Decentralized RL for IoT lets the devices use acknowledgements sent by their Base Station as a reward, instead of sensing feedback like for OSA.
+We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
+Simulations show that dynamic devices can improve their spectrum efficiency,
+by using MAB algorithms like \UCB.
+% in the case of a non-uniform repartition of the static devices in the channels
+We also developed a proof-of-concept using USRP platforms, for a real-world validation of this approach.
+% Numerical simulations justify the interest of the approach,
+%  of the advantage of using MAB algorithms for the first model.
+%
+% On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
+The first model
+% without retransmission of packets,
+is interesting for its simplicity,
+and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
 
-In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithms to learn and improve their access to the network.
-FIXME finish !
+However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $M=10$).
+We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics significantly improve the network efficiency.
+% Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
+%
+The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
 
@@
 
-\paragraph{Abstract.}
+% \paragraph{Abstract.}
 
-In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithms to learn and improve their access to the network.
-The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
-We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP platforms \cite{USRPDocumentation}.
+% In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithms to learn and improve their access to the network.
+% The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
+% We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP platforms \cite{USRPDocumentation}.
 
-More precisely, we start by motivating the idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
-We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
-Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
-We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every instant.
+% More precisely, we start by motivating the idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
+% We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
+% Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
+% We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every instant.
 If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
-we show in simulations that in the case of a non-uniform repartition of the static devices in the channels, the dynamic devices can drastically improve the network efficiency.
-%
-On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
+% we show in simulations that in the case of a non-uniform repartition of the static devices in the channels, the dynamic devices can drastically improve the network efficiency.
+% %
+% On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
 
-Then we present in Section~\ref{sec:4:retransmissions} a second model, extending our first model, in which devices can retransmit their messages for a fixed number of maximum retransmission, like in the ALOHA protocol.
-In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
-Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
+% Then we present in Section~\ref{sec:4:retransmissions} a second model, extending our first model, in which devices can retransmit their messages for a fixed number of maximum retransmission, like in the ALOHA protocol.
+% In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
+% Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
 
-On the one hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
-On the other hand, the second model is more interesting if the main target is an improvement of the QoS of the IoT networks rather than a longer battery life.
+% On the one hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
+% On the other hand, the second model is more interesting if the main target is an improvement of the QoS of the IoT networks rather than a longer battery life.
 
-\vfill{}
+% \vfill{}
 
-\textbf{Publications.}
-
-This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
-
-\newpage
+% \newpage
 
@@
 Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
-As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary.
-This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
+As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly not stationary''.
+This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms (that is why we relax the hypothesis and only consider up-to $M\leqK$ players in Chapter~\ref{chapter:5}).
 We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
@@
 We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
+
+\textbf{Publications.}
+%
+This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 
@@
 
-\subsection*{Summary of this chapter}
+% \subsection*{Summary of this chapter}
 
@@
 In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
-%
+
 On the one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
@@
 Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-player MAB} model to describe our problem, as proposed by \cite{Zhao10}.
-%
+
 The main difference between the two models is the fact that in this chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).

### 2019-06-09 2-Chapters/1-Chapter/chapter1.tex
Nice work on Ch5

@@
 \subsection{Example: decentralized learning in unlicensed bands with selfish or orthogonalization-based MAB algorithms}
+
+We already explained in Chapter~\ref{chapter:1} that
+unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
+This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.
+
+Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
+As end-devices battery life is a key constraint of IoT networks,
+and as IoT networks are decentralized (\ie, the devices initiate transmissions),
+this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
+%
+In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
+Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
+We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
+Kaufmann12Thompson}, for the first model.
+For the demonstration as well as for the second model, without loss of generality, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.
+
+As detailed in Chapter~\ref{chapter:1},
+MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
+For example,
+proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
+and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
+Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
+Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
+% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
+
+We evoke here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
+Starting from $2010$, the works of Wassim Jouini \cite{Jouini09,Jouini10,Jouini12} were the first ones to propose to use Reinforcement Learning for Cognitive Radio, especially MAB and the \UCB{} algorithm, and from $2014$, proof-of-concepts were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
+In $2015$ and $2016$, Christophe Moy continued to work on this direction, with a PhD student (Navikkumar Modi) and post-doc (Sumit J. Darak) in our team SCEE, for the following publications
+\cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
+Since $2017$, Sumit J.
+Darak and his team at IIIT Delhi have been very active in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+\cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
 

### 2019-06-09 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Nice work on Ch5

@@
   %
-  The regret lower bound given in \cite{Besson2018ALT} is not applicable to any algorithm,
+  The regret lower bound that we gave in \cite{Besson2018ALT} is not applicable to any algorithm,
   as it was discovered by Etienne Boursier and Vianney Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
@@
 %
-Letting $\cP_M \eqdef \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
-be the set of bandit instances such that there is a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.
+By denoting $\cP_M \eqdef \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
+the set of bandit instances with a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.
 
@@
 However, a careful manipulation of this decomposition shows that the regret is always lower bounded by term \ref{eq:5:term1}.
-The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} in Appendix illustrate two cases of $M<K$ and $M=K$ and the different impact of the three terms (\ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3}) on the regret.
+The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} in Appendix illustrate two cases of $M<K$ and $M=K$ and the different impact of the three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3} on the regret.
 
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    For any strategy $\cA$ and $\boldsymbol{\mu}\in\cP_M$, it holds that
+    For any strategy $\cA$ and $\boldsymbol{\mu}\in\cP_M$, the regret is lower-bounded by \ref{eq:5:term1}:
     \begin{align*}
@@
 % This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
-
+%
+\vspace*{-20pt}  % XXX
 \begin{definition}\label{def:5:DecentralizedUniformEfficiency}
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-  An algorithm $\cA$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$ and for all $\alpha \in (0,1)$,
+  An algorithm $\cA$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$,
+  % and for all $\alpha \in (0,1)$,
   % \vspace*{-5pt}  % XXX
   \begin{align}
-    & R_T(\boldsymbol{\mu},M,\cA) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
-    \text{and} & \;
+    \forall a\in(0,1) & R_T(\boldsymbol{\mu},M,\cA) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
+    \text{and} \forall a\in(0,1) & \;
     \forall j \in [M], k \in \Mbest, \;\;
@@
 
-\paragraph{Two ideas for orthogonalization.}
+\textbf{Two ideas for orthogonalization.}
 \label{par:5:twoIdeasOrthogonalization}
@@
 % Now, our algorithms !!
-\paragraph{Our two proposals.}
+\textbf{Our two proposals.}
 We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
@@
 \begin{figure}[h!]
-  \resizebox{1.10\textwidth}{!}{
+  \resizebox{1.05\textwidth}{!}{
   \begin{tikzpicture}[>=latex',line join=bevel,scale=5.0]
@@
       \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
-      \path [color=blue,thick,->] (notfixed) edge[loop right] node[below,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t)$,\\$A^j(t) \in \TopM(t)$} (1);
+      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t)$,\\$A^j(t) \in \TopM(t)$} (1);
       \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
-      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[below,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
+      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
       \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
@@
   \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-      For any arm $k$, if $U_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log(\log(t))$,
-      \begin{equation}
-        \sum_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log(\log(T)).
-      \end{equation}
+      For any arm $k$, if $U_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log(\log(t))$, then
+      % \begin{equation}
+      $\sum\limits_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log(\log(T))$.
+      % \end{equation}
   \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 %
-For the plots showing the regret, our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} is displayed.
+We also include our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} on regret plots.
 
@@
 
-Note that more numerical experiments were conducted for the article \cite{Besson2018ALT}, and in particular the last pages of the paper show more figures.
+Note that more numerical experiments were conducted for the article \cite{Besson2018ALT}, and in particular the last pages of the paper show more figures, in other settings.
 

### 2019-06-09 2-Chapters/5-Chapter/chapter5.tex
Nice work on Ch5

@@
 
-\TODOL{Expliquer pourquoi ce n'est pas très réaliste ? Le paragraphe ci dessous n'est pas très simple à comprendre, je peux faire mieux!}
+% \TODOL{Expliquer pourquoi ce n'est pas très réaliste ? Le paragraphe ci dessous n'est pas très simple à comprendre, je peux faire mieux!}
 
-If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm running on top of a round-Robin phase can achieve order-optimal regret.
-If the players know the time horizon $T$, then fix a confidence level $\delta$.
+If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm (see Algorithm~\ref{algo:2:naiveStrategies}), running on top of a round-Robin phase can achieve order-optimal regret.
+If the players know the time horizon $T$, then they can fix a confidence level $\delta$.
 For the first $T_0$ time steps, the $M$ players will use a simple round-Robin game, using their (unique) ID to stay orthogonal: player $j$ starts at arm $j$, then $j+1$, then cycle in $[K]$.
@@
 The mean regret of such approach is easily bounded by $R_T \leq M T_0 + (1 - \delta) (T - T_0)$, so by using $\delta = 1-1/T$, $R_T = \bigO{T_0}$.
-By calibrating $T_0$ based on $\delta$ (which needs prior knowledge of $T$) and the minimal gap $\Delta$ between \Mbest{} and \Mworst, one can show, using similar arguments as used by \cite{Rosenski16} for the Musical Chair policy, that with large probability a constant regret is obtained.
+By calibrating $T_0$ based on $\delta$ (which needs prior knowledge of $T$, see the proof we gave in Section~\ref{proof:2:tuningExploreThenCommit}) and the minimal gap $\Delta$ between \Mbest{} and \Mworst, one can show, using similar arguments as used by \cite{Rosenski16} for the Musical Chair policy, that with large probability a constant regret is obtained.
 A logarithmic regret can be obtained from the same bound, proving the order-optimality of this simple approach, if we assume that user $j$ knows it is user number $j$.
@@
 \begin{equation}\label{eq:5:defRegretDynamicSetting}
-    R_T^{\text{dyn}} \eqdef \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\Bigl[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \Bigr].
+    R_T^{\text{dyn}} \eqdef \sum_{t=1}^T \sum_{k=1}^{M(t)} \mu^*_{k} - \E\left[ \sum_{t=1}^T \sum_{j=1}^{M(t)} r^j(t) \right].
 \end{equation}
@@
 
-In \cite{Rosenski16}, it is assumed that the players all know a lower-bound $L$ on the length of all the intervals during which arrival or departures of players are allowed.
+The authors of \cite{Rosenski16} also assume that the players all know a lower-bound $L$ on the length of all the intervals during which arrival or departures of players are allowed.
 A naive idea, if $L$ is large enough, is simply to restart the underlying algorithm every $L$ time steps, in order to directly benefit from the theoretical guarantees of the ``static setting'' (where $M$ stays constant).
@@
 Under the same hypothesis, and if the lower-bound $L$ is known before-hand and is large enough, then we could also apply the same idea as from \cite{Rosenski16} to our approach.
-We believe that we could easily prove that a sub-linear regret is achievable, if all players run \MCTopM-\klUCB{} (and if currently active players restart their memory of the past observations every $L$ time steps).
-As the regret guarantees are stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, applying the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to \MCTopM{} would also give a small regret upper-bound.
+We believe that we could easily prove a sub-linear regret bound, if all players run \MCTopM-\klUCB, and if currently active players restart their memory of the past observations every $L$ time steps.
+As the regret guarantee is stronger (\ie, smaller regret upper-bound) for \MCTopM{} than \MusicalChair, we also believe that in this ``slowly varying'' dynamic setting, applying the idea of Dynamic \MusicalChair{} from \cite{Rosenski16} to \MCTopM{} would also give a small regret upper-bound.
 
@@
 This model seems harder than the model with sensing information, and even though the proposal \Selfish{} works fine in simulations (in terms of average regret), we proved that it can yield linear regret.
-In our paper \cite{Besson2018ALT} as well as the previous sections, it was left as a future work to know if an algorithm can achieve sub-linear regret in this harder model without sensing information.
+In our paper \cite{Besson2018ALT} as well as in the previous sections, it was left as a future work to know if an algorithm can achieve sub-linear regret in this harder model without sensing information.
 
@@
 Similarly, Lugosi and Mehrabian studied the same problem \cite{LugosiMehrabian18}.
-In both works, the authors detail algorithms that give in both cases a logarithmic regret if $M \leq K$ players all independently implement the proposed algorithm.
+In both works, the authors detail algorithms that give a logarithmic regret if $M \leq K$ players all independently implement the proposed algorithm.
 
@@
 More work is needed on our implementation on Improved \MusicalChair.
-The results on \textsc{Sic-MMAB} appear to confirm the numerical experiments of \cite{BoursierPerchet18}.}
+The results on \textsc{Sic-MMAB} confirm the numerical experiments of \cite{BoursierPerchet18}.}
     \label{table:5:meanRegretSimulationsNoSensing}
@@
 
-In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
+In the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with their gateway.
 This hypothesis is realistic, mostly for energy consumption and spectrum efficiency reasons.
-Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between devices are possible.
-Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``\emph{graph bandit}'' problem.
-What we are interested here is an extension of the model where players can send some bits to one or all the other players, at every time steps or some time steps.
+Of course, we can relax this hypothesis, and in practice in some families of wireless networks, communication between devices are possible.
+Note that this extension is \emph{not} considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``\emph{graph bandit}'' problem \cite{valko2016bandits}.
+We are interested here by an extension of the model where players can send some bits to one or all the other players, at some or every time steps.
 

### 2019-06-09 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Done, I changed := to \overset{.}{=}
Done proof reading ch5

@@
 
-%     Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t):=\indic(\overline{C^j(t)})$,
+%     Using the definition of regret $R_T$ from \eqref{eq:5:regret}, and this collision indicator $\eta^j(t) \eqdef \indic(\overline{C^j(t)})$,
 %     \begin{align*}
@@
 %         - \sum_{t=1}^{T} \sum_{j=1}^{M} \mu_{A^j(t)} \eta^j(t) \right] \\
-%         % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{:= \overline{\mu}^*} (T M)
+%         % &= \underbrace{\left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)}_{ \eqdef \overline{\mu}^*} (T M)
 %       &= \left( \frac{1}{M} \sum_{j \in \Mbest} \mu_j \right)
@@
 %       %
-%       \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* := \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
+%       \intertext{For the first term, we have $T M = \sum\limits_{k=1}^{K} \sum\limits_{j=1}^{M} \E_{\mu}\left[ T^j_k(T)\right]$, and if we denote $\overline{\mu}^* \eqdef \frac{1}{M} \sum\limits_{j \in \Mbest} \mu_j$ the average mean of the $M$-best arms, then,}
 %       %
@@
 %       %
-%       \intertext{Let $\overline{\Delta_k} := \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
+%       \intertext{Let $\overline{\Delta_k} \eqdef \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
 %       %
@@
 %   %
-%   Let $T_k^{!}(T) := \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
+%   Let $T_k^{!}(T) \eqdef \sum_{t=1}^{T} \indic(\exists! j, A^j(t)=k)$,
 %   where the notation $\exists!$ stands for ``there exists a unique''.
@@
 %   \Pr_{\bm{\mu}}\left(A_T\right)
-%     & \leq 2M\frac{\E_{\bm{\mu}}\left[T^j_k(T)\right]}{T} =: x_T, \\
+%     & \leq 2M\frac{\E_{\bm{\mu}}\left[T^j_k(T)\right]}{T} \reqdef x_T, \\
 %   \Pr_{\bm{\lambda}}\left(\overline{A_T}\right)
-%     & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_{\bm{\lambda}}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} =: \frac{y_T}{T},
+%     & = \Pr_{\bm{\lambda}}\left(\frac{T}{M} - T^j_k(T) > \frac{T}{2M}\right) \leq 2M\frac{\E_{\bm{\lambda}}\left[ \frac{T}{M} - T^j_k(T)\right]}{T} \reqdef \frac{y_T}{T},
 % \end{align*}
@@
 % \[\forall t > 0, \ \ \cO_t = (U_0,Y_1,C_1,U_1,\dots,Y_t,C_t,U_t)\]
-% where $Y_{t} := Y_{A^j(t-1),t}$ denotes the sensing information, $C_t := C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $[M]$ for new ranks or arms.} useful to select $A^j(t)$.
+% where $Y_{t} \eqdef Y_{A^j(t-1),t}$ denotes the sensing information, $C_t \eqdef C^j(t)$ denotes the collision information (not always completely exploited under observation model \modeldeux) and $U_t$ denotes some external source of randomness\footnote{For instance, \MCTopM, \RandTopM{} and \rhoRand{} draws from a uniform variable in $[M]$ for new ranks or arms.} useful to select $A^j(t)$.
 % Formally, one can say that $A^j(t)$ is $\sigma(\cO_{t-1})$ measurable, if $\sigma(\cO_t)$ denotes the sigma-Algebra generated by the observations $\cO_t$, and as $\cF^j(t) \subseteq \sigma(\cO_{t})$, with an equality under observation model \modelun.
@@
 %   & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right) \\
-%   & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{:= T_{k'}}
+%   & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{ \eqdef T_{k'}}
 % \end{align*}
@@
 To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
-% Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
+% Letting $\kl^+(x, y) \eqdef \kl(x, y) \indic(x \geq y)$,
 % and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
@@
 % \begin{equation*}
-%   % \phi_{k}^j(t) :=
+%   % \phi_{k}^j(t) \eqdef
 %   \left(
@@
 % \begin{align*}
-%   I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
+%   I_1^j(t) & \eqdef \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
 %   % go sitted
-%   I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
+%   I_2^j(t) & \eqdef \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
 %   \ \ \ &\text{and} \ \ \
-%   I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
+%   I_3(t) & \eqdef \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
 %   % if sitted
-%   I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
+%   I_3(t) & \eqdef \left(s^j(t),A^j(t) \in \TopM(t)\right),
 %   \ \ \ &\text{and} \ \ \
-%   I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
+%   I_5(t) & \eqdef \left(s^j(t),A^j(t) \notin \TopM(t)\right).
 % \end{align*}
@@
 % \begin{equation}
-%     \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
+%     \widetilde{C^j}(t) \eqdef \left(C^j(t), \overline{s^j(t)}\right).
 % \end{equation}
@@
 % Thus, $\sum_{j=1}^M \Pr(C^j(t))$ can be upper bounded by $M \sum_{j=1}^M \Pr(\widetilde{C^j}(t))$ (union bound),
-% and it follows that if $\cC(T) := \sum_{k=1}^K \cC_k(T)$ then
+% and it follows that if $\cC(T) \eqdef \sum_{k=1}^K \cC_k(T)$ then
 % \[\E_{\mu}[\cC(T)] \leq M \sum_{j=1}^M \sum_{t=1}^T \Pr(\widetilde{C^j}(t)).\]
@@
 % \ie,
-% $n(T) := \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
+% $n(T) \eqdef \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
 % (or $0$ if $\theta_1$ does not exist).
@@
 We proved in Theorem~\ref{thm:5:BetterLowerBound} that the normalized regret, \ie, $R_T$ divided by
-$\log T$, is asymptotically lower bounded by a constant $\mathrm{LB}(\boldsymbol{\mu}, M)$
+$\log T$, is asymptotically lower bounded by a constant $C_{\boldsymbol{\mu}, M}$
 depending on the problem $\boldsymbol{\mu}$ and the number of players $M$, for any $\cA$.
 \begin{equation}
-  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log T} \geq \mathrm{LB}(\boldsymbol{\mu}, M).
+  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log T} \geq C_{\boldsymbol{\mu}, M}.
 \end{equation}
@@
 the number of player, from $1$ player to $9$ players, and on the
-$y$ axis is the value of this constant $\mathrm{LB}(\boldsymbol{\mu}, M)$, from the initial
+$y$ axis is the value of this constant $C_{\boldsymbol{\mu}, M}$, from the initial
 theorem and from our theorem.
@@
 Figure~\ref{fig:5:CompLowerBounds} clearly shows that our improved lower bound is indeed larger than the initial one by \cite{Zhao10},
-and both become uninformative when $M=K$ (\ie, null).
+and both become uninformative when $M=K$ (\ie, it gives $R_T / \log(T) \geq 0$ which is obvious anyway).
 

### 2019-06-09 2-Chapters/6-Chapter/chapter6.tex
Done for proofreading Ch6
Done proof reading ch5

@@
 %
-Finally, we compare our approach against other state-of-the-art solutions on extensive numerical experiments.
+Finally, we compare our approach with other state-of-the-art solutions on large numerical experiments.
 % , on different piece-wise stationary bandit problems.

### 2019-06-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Lots of small typos.
Done proof reading ch5

@@
 % This chapter is structured as follows.
-We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}.
-In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
+We introduce the model in Section~\ref{sec:6:BanditSetting}, and review related works in Section~\ref{sec:6:relatedWork}.
+In Section~\ref{sec:6:ChangePointDetector} we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
 We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on the regret of both variants.
-The unified regret analysis for these two algorithms is given in Section~\ref{sec:6:RegretAnalysis}.
-Numerical experiments are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.
+The unified regret analysis is given in Section~\ref{sec:6:RegretAnalysis}, we prove one result in Section~\ref{sec:6:proofRegret}.
+We discuss numerical experiments in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.
 
@@
 At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) = Y_{A(t),t}$.
-At time $t$, we denote by $k_t^*$ an arm with maximal expected reward, \ie, $\mu_{k_t^*}(t) = \max_k \mu_k(t)$, called an optimal arm (possibly not unique).
+At time $t$, we denote by $k_t^*$ an arm with maximal expected reward, \ie, $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$, called an optimal arm (possibly not unique).
 
 A policy $\cA$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
-Like for stationary problems (see Definition~\ref{def:2:regret}), the performance of $\cA$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $k_t^*$ at time $t$, and that of the policy $\cA$
-We use the same notation as the regret for stationary problem without ambiguity, as in this chapter we only consider the following definition.
+Like for stationary problems (see Definition~\ref{def:2:regret} in Chapter~\ref{chapter:2}), the performance of $\cA$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $k_t^*$ at time $t$, and that of the policy $\cA$
+We use the same notation as the regret for stationary problem without ambiguity, as in this chapter we only consider the following definition,
 %
 \begin{equation}\label{def:6:nti}
-    R_T^{\cA} = \E\left[\sum_{t=1}^T \left(\mu_{k_t^*}(t) - \mu_{A(t)}(t)\right)\right].
+    R_T^{\cA} \eqdef \E\left[\sum_{t=1}^T \left(\mu_{k_t^*}(t) - \mu_{A(t)}(t)\right)\right].
 \end{equation}
@@
 In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T \eqdef \sum_{t=1}^{T-1} \indic\left(\exists k\in[K]: \mu_k(t) \neq \mu_k(t+1)\right)$.
-If we denote $\tau^{(0)} = 0$, we define the $i$-th breakpoint recursively by
-$\tau^{(i)} = \inf\{t > \tau^{(i-1)} : \exists k : \mu_k(t) \neq \mu_k(t+1)\}$.
+If we denote $\tau^{(0)} \eqdef 0$, we define the $i$-th breakpoint recursively by
+$\tau^{(i)} \eqdef \inf\{t > \tau^{(i-1)} : \exists k : \mu_k(t) \neq \mu_k(t+1)\}$.
 %
@@
 Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
-For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fixed channel or subset of channels, and on a particular day in a particular city, if the company deploys a lot of new objects, then one or some channels see their mean occupancy rates abruptly change.
+For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new objects, then one (or some) channel(s) see their mean occupancy rates abruptly change.
 
@@
 %
-The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while some adversarial models can be considered as a special (worst) case, with $\Upsilon_T = T$ (when considering an adaptive or an oblivious adversary).
+The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while some adversarial models can be considered as a special (worst) case, with $\Upsilon_T = T-1$ (when considering an adaptive or an oblivious adversary).
 However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
@@
 Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
-Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
+Active strategies monitor the arms' rewards, by using change detection algorithms \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
 Passive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
@@
         \begin{equation}
-            R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + 1 / \left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)},
+            R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)},
         \end{equation}
@@
 % -----------------------------------------------------------------
-\begin{figure}[h!]  % [htbp]
+\begin{figure}[b!]  % [htbp]
     \centering

### 2019-06-09 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
OMG c'est 00h23 un dimanche soir et je bosse !

@@
 			\KwIn{$N$ bandit algorithms, $\Alg_1, \dots, \Alg_N$, with $N \geq 2$}
-			\KwIn{Number of arms, $K \geq 2$}
-			\KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
+			% \KwIn{Number of arms, $K \geq 2$}
+			% \KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
 			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$}
@@
 a naive uniform exploration (to have at least one algorithm with bad performances, \ie, linear regret, that is \emph{not} included in the plots to reduce clutter),
-\UCB{} with $\alpha=1/2$, three \klUCB{} algorithms respectively using Bernoulli, Gaussian and exponential $\kl$ functions, and BayesUCB and Thompson sampling with uniform prior.
+\UCB{} with $\alpha=1/2$, three \klUCB{} algorithms (resp. with Bernoulli, Gaussian and exponential $\kl$ functions), and BayesUCB and Thompson sampling with uniform prior.
 
@@
 % as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
-
+%
 % - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
@@
 Experiments presented in this Section are based on our library SMPyBandits,
-and the page \texttt{SMPyBandits.GitHub.io/Aggregation.html} gives instructions to reproduce them.
+and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.
 

### 2019-06-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
OMG c'est 00h23 un dimanche soir et je bosse !

@@
 A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
-$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N^*, \bE[X_i] = \mu_0\right)$.
+$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall t\in\N^*, \bE[X_t] = \mu_0\right)$.
 
@@
 Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
-As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows before hand the values of $a$ and $b$).
+As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows beforehand the values of $a$ and $b$).
 
@@
 
-If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem introduced above would reduce to a parametric sequential test of
-$\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$,
+If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem would reduce to a parametric sequential test of
+$\cH_0 : (\exists \mu_0: \forall t\in\N^*, X_t \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$,
 against the alternative
-$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
+$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.\\
 %
@@
 % in Appendix~\ref{app:6:GLR_with_kl}
-that in any one-dimensional exponential family, we have
+that for Bernoulli distributions, we have
 \begin{equation}\label{eq:6:secondDefGLRT}
@@
 %
-where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
+where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{t=k}^{k'} X_t$ denotes the average of the observations $X_t$ collected between the instants $k$ and $k'$.
 
@@
     First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
-    We have $\ell(X_1, \dots, X_n ; \mu_0) = \prod_{i=1}^n \ell(X_k ; \mu_0)$ by independence of the observations $X_k$,
-    and $\ell(X_k ; \mu_0) = \mu_0^{X_k} (1-\mu_0)^{1-X_k}$ for Bernoulli distributions.
+    We have $\ell(X_1, \dots, X_n ; \mu_0) = \prod\limits_{i=1}^n \ell(X_i ; \mu_0)$ by independence of the observations $X_i$,
+    and $\ell(X_i ; \mu_0) = \mu_0^{X_i} (1-\mu_0)^{1-X_i}$ for Bernoulli distributions.
     Therefore, taking the logarithm gives
     \begin{align*}
-        \log\ell(X_1, \dots, X_n ; \mu_0)
-        &= \sum_{i=1}^n X_k \log(\mu_0) + (1-X_k) \log(1-\mu_0) \\
-        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_k \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_k \right) \\
+        \log \left[ \ell(X_1, \dots, X_n ; \mu_0) \right]
+        &= \sum_{i=1}^n X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) \\
+        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
         &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
@@
     If $a=0$ or $a=1$, $h$ is maximum at $x=a$.
-    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative:
+    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative,
     $h'(x) = a/x - (1-a)/(1-x)$, so $h'(x) = 0$ if and only if $x=a$.
@@
     \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
-        \sup_{\mu_0} \log\ell(X_1, \dots, X_n ; \mu_0) = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
+        \sup_{\mu_0} \log \left[ \ell(X_1, \dots, X_n ; \mu_0) \right] = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
     \end{equation}
@@
     \begin{align*}
-        & \log\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau) \\
-        &= \sum_{i=1}^{\tau} X_k \log(\mu_0) + (1-X_k) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_k \log(\mu_1) + (1-X_k) \log(1-\mu_1) \\
+        & \log \left[ \ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau) \right] \\
+        &= \sum_{i=1}^{\tau} X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_i \log(\mu_1) + (1-X_i) \log(1-\mu_1) \\
         &= s \bigl( \widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0) \bigr) \\
@@
 $\mu_1$) and then left to be solved for $s$.
-    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_k = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,
+    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_i = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,
     which are both handled similarly.
@@
     Let be $\bP_{\mu_0}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t$.
-    % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_k \in [0,1]$ for all $i$.
+    % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
     Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
@@
 The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
-
+%
 Let us go back to the Bernoulli case with divergence $\kl(x,y)$.
-We first observe that
+We first have
 \[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\lambda \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
@@
 \begin{align*}
-    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
-    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
+    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists s \in \N^*, n \in \N^*, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
+    & \leq \bP_{\mu_0}\left(\exists s \in \N^*, n \in \N^*, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
 % \end{align*}%FIXME uncomment this if needed
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    $(X_k)_{i \in \N^*}$ and $(Y_k)_{i \in \N^*}$ two independent \iid{} processes with respective means $\mu$ and $\mu'$ such that
-    $\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)}$,
-    where $\phi_\mu(\lambda) = \bE_{X \sim \nu^{\mu}}[e^{\lambda X}]$ is the moment generating function of the distribution $\nu^\mu$, which is the unique distribution in an exponential family that has mean $\mu$.
-
-    Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family.
-Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_k$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_k$, it holds that for every $s,r \in \N^*$,
-    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
+    Consider a one-dimensional exponential family $\cE$,
+    and let $\nu_{\mu}$ be the unique distribution in this family that has mean $\mu$,
+    with moment generating function $\phi_\mu(\lambda) = \bE_{X \sim \nu_{\mu}}[e^{\lambda X}]$.
+    Let $\kl_{\cE}(\mu,\mu') = \KL(\nu_{\mu},\nu_{\mu'})$ be the divergence function associated to $\cE$.
+    Let $(X_i)_{i \in \N^*}$ and $(Y_k)_{i \in \N^*}$ be two independent \iid{} processes, with respective means $\mu$ and $\mu'$, such that
+    $\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)}$.
+    %
+    Denote $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_k$.
+    Then for every $s,r \in \N^*$ we have,
+    \begin{small}
+        \begin{equation}
+            \bP\left(\exists r \in \N^* : s \, \kl_{\cE}\left(\widehat{\mu}_{s},\mu\right) + r \, \kl_{\cE}\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},
+        \end{equation}
+    \end{small}
     where $\cT$ is the function defined in \eqref{def:6:function_T}.
@@
     Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
-    and let $\Delta = |\mu_0 - \mu_1|$.
+    and let $\Delta = |\mu_0 - \mu_1|$.\\
     Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
@@
 
-Also note that, as any test based on scan-statistics, the B-GLRT can be costly to implement as at every time step, it considers all previous time steps as a possible positions for a change-point.
-Thus, in practice the following adaptation may be interesting, based on down-sampling the possible time steps:
+Also note that, as any test based on scan-statistics, the B-GLRT can be costly to implement as at every time step, it considers all previous time steps as a possible position for a change-point.
+Thus, down-sampling the possible time steps is an interesting adaptation in practice:
 %
@@
     For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $k$ (\ie, for a $i\in[\NCk]$, one has $\tau_k^{(i)} = \tau_k(t) < s \leq t < \tau_k^{(i+1)}$) it holds that
-    \begin{equation}
-        n_k(t) - n_k(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor.
-    \end{equation}
+    % \begin{equation}
+    $n_k(t) - n_k(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor$.
+    % \end{equation}
 \end{leftbar}  % XXX leftbar propositionbar, comment if needed
@@
 This section give the finite-time regret bounds for the two variants, and interpretations of the different results.
+Although we only prove one bound, the other proof can be found in \cite{Besson2019GLRT}.
 
@@
 
-Recall that $\tau^{(i)}$ denotes the position of the $i$-th break-point and let $\mu_k^{(i)}$ be the mean of arm $k$ on the segment between the $i$ and $(i+1)$-th breakpoint:
+Recall that $\tau^{(i)}$ denotes the position of the $i$-th break-point and let $\mu_k^{(i)}$ be the mean of arm $k$ on the segment between the $i$- and $(i+1)$-th breakpoint:
 $\forall t \in \{ \tau^{(i-1)}+1, \dots, \tau^{(i)} \}, \mu_k(t) = \mu_k^{(i)}$.
-We also introduce $i^* = \arg\max_k \mu_k^{(i)} $ and the largest gap at break-point $i$ as $\Delta^{(i)} \eqdef \max_{k\in[K]} |\mu_k^{(i)} - \mu_k^{(i-1)}| >0$.
-
-\begin{assumption}[Enough samples between two global change-points]\label{ass:6:LongPeriodsGlobal}
+Let $i^* = \arg\max_k \mu_k^{(i)} $ and the largest gap at break-point $i$ as $\Delta^{(i)} \eqdef \max_{k\in[K]} |\mu_k^{(i)} - \mu_k^{(i-1)}| >0$.
+
+\begin{assumption}\label{ass:6:LongPeriodsGlobal}
 \begin{leftbar}[assumptionbar]  % XXX leftbar assumptionbar, comment if needed
     Define
-    \begin{equation}\label{def:6:DelayGlobal}
-        d^{(i)} = d^{(i)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta^{(i)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
-    \end{equation}
-    %
-    Then \emph{we assume} that for all change $i \in [\Upsilon_T]$,
+    % \begin{equation}\label{def:6:DelayGlobal}
+    $d^{(i)} = d^{(i)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta^{(i)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil$.
+    % \end{equation}
+    %
+    Then \emph{we assume that there are enough samples between two global change-points}.
+    In other words, we assume for all change $i \in [\Upsilon_T]$,
     $\tau^{(i)} - \tau^{(i-1)} \geq 2\max (d^{(i)},d^{(i-1)})$.
@@
         R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T \\
-        & + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
+        & + \sum_{k=1}^{\Upsilon_T}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
 \nonumber
@@
 
-\begin{assumption}[Enough samples between two local change-points]\label{ass:6:LongPeriods}
+\begin{assumption}\label{ass:6:LongPeriods}
 \begin{leftbar}[assumptionbar]  % XXX leftbar assumptionbar, comment if needed
     Define
-    \begin{equation}\label{def:6:DelayLocal}
-        d_k^{(\ell)} = d_k^{(\ell)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil.
-    \end{equation}
-    %
-    Then \emph{we assume} that for all arm $k$ and all change-point of that arm $\ell \in [\NCk]$, $\tau_k^{(\ell)} - \tau_k^{(\ell-1)} \geq 2\max (d_k^{(\ell)},d_k^{(\ell-1)})$.
+    % \begin{equation}\label{def:6:DelayLocal}
+    $d_k^{(\ell)} = d_k^{(\ell)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil$.
+    % \end{equation}
+    %
+    Then \emph{we assume that there are enough samples between two local change-points}.
+    In other words, we assume that for all arm $k$ and all change-point of that arm $\ell \in [\NCk]$, $\tau_k^{(\ell)} - \tau_k^{(\ell-1)} \geq 2\max (d_k^{(\ell)},d_k^{(\ell-1)})$.
 \end{leftbar}  % XXX leftbar assumptionbar, comment if needed
@@
 For example for local changes (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
-\begin{equation}
-    \cE_T(\alpha,\delta) = \left(\forall k \in [K], \forall \ell \in [\NCk], \hat{\tau}^{(\ell)}_k \in \left[\tau_k^{(\ell)} + 1, \tau_k^{(\ell)} + d_k^{(\ell)}\right] \right),
-\end{equation}
+% \begin{equation}
+$\cE_T(\alpha,\delta) = \left(\forall k \in [K], \forall \ell \in [\NCk], \widehat{\tau}^{(\ell)}_k \in \left[\tau_k^{(\ell)} + 1, \tau_k^{(\ell)} + d_k^{(\ell)}\right] \right)$,
+% \end{equation}
 %
@@
 Introducing the event
-\[\cC_k^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_k^{(\ell)} \in \left[\tau_k^{(j)} + 1, \tau_k^{(j)} + d_k^{(j)} \right] \right\},\]
+$\cC_k^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_k^{(\ell)} \in \left[\tau_k^{(j)} + 1, \tau_k^{(j)} + d_k^{(j)} \right] \right\}$,
 that all the changes up to the $\ell$-th have been detected, a union bound yields this decomposition
@@
 
-We first introduce some notations for the proof:
+% We first introduce some notations for the proof:
+The proof uses these notations:
 let $\widehat{\tau}^{(i)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
@@
 Let $d^k$ be defined as in Assumption~\ref{ass:6:LongPeriodsGlobal}, we define
-\begin{equation}\cE_T(\delta) = \left( \forall k \in [\Upsilon_T], \hat{\tau}^{(i)} \in \left[\tau^{(i)} + 1, \tau^{(i)} + d^{(i)}\right] \right).\label{def:6:GoodEvenGlobal}\end{equation}
+\begin{equation}\cE_T(\delta) = \left( \forall k \in [\Upsilon_T], \widehat{\tau}^{(i)} \in \left[\tau^{(i)} + 1, \tau^{(i)} + d^{(i)}\right] \right).\label{def:6:GoodEvenGlobal}\end{equation}
 Under the good event, all the change points are detected within a delay at most $d^(k)$.
 %
-From Assumption~\ref{ass:6:LongPeriodsGlobal}, as the period between two changes are long enough, if $\cE_T(\delta)$ holds, then for all change $k$, we have $\tau^{(i)} \leq \hat{\tau}^{(i)} \leq \tau^{(i+1)}$.
+From Assumption~\ref{ass:6:LongPeriodsGlobal}, as the period between two changes are long enough, if $\cE_T(\delta)$ holds, then for all change $k$, we have $\tau^{(i)} \leq \widehat{\tau}^{(i)} \leq \tau^{(i+1)}$.
 So we can prove the following.
@@
      & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}(t))\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(i)} + \indic(\cE_T) \sum_{t=\hat{\tau}^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(i)}) \sum_{t=\hat{\tau}^{(i)}}^{\tau^{(i+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right],
+    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right] \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(i)}) \sum_{t=wide\hat{\tau}^{(i)}}^{\tau^{(i+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right],
 \end{align*}
@@
 Thus, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_{k^*}^{(i)}$, one can write
-
+%
 \begin{align*}
-    &\bE\left[\left.\indic(\cC^{(i)}) \sum_{t=\hat{\tau}^{(i)}}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \right| \cF_{\widehat{\tau}^{(i)}}\right]\\
-    & = \indic(\cC^{(i)}) \sum_{t=\hat{\tau}^{(i)}}^{\tau^{(i+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \;|\; \cF_{\widehat{\tau}^{(i)}}\right] \\
-    & \leq \indic(\cC^{(i)}) \sum_{t'=1}^{\tau^{(i+1)} - \hat{\tau}^{(i)}} \bP\left(\exists s \leq t' : s \, \kl(\widehat{\mu}_{s},\mu_{k^*}^{(i)}) \geq f(t')\right) \\
+    &\bE\left[\left.\indic(\cC^{(i)}) \sum_{t=wide\hat{\tau}^{(i)}}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \right| \cF_{\widehat{\tau}^{(i)}}\right]\\
+    & = \indic(\cC^{(i)}) \sum_{t=\widehat{\tau}^{(i)}}^{\tau^{(i+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \;|\; \cF_{\widehat{\tau}^{(i)}}\right] \\
+    & \leq \indic(\cC^{(i)}) \sum_{t'=1}^{\tau^{(i+1)} - \widehat{\tau}^{(i)}} \bP\left(\exists s \leq t' : s \, \kl(\widehat{\mu}_{s},\mu_{k^*}^{(i)}) \geq f(t')\right) \\
     & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
@@
 %
-where the last but one inequality relies on the concentration inequality given in Lemma 2 of \cite{KLUCBJournal} and the fact that $f(t) = \ln(t) + 3 \ln(\ln(t))$.
+where the last but one inequality relies on the concentration inequality given in Lemma~2 of \cite{KLUCBJournal} and the choice $f(t) = \ln(t) + 3 \ln(\ln(t))$.
 Finally, using the law of total expectation yields
@@
 %
-We let $\tilde{\mu}_{k,s}^{(i)}$ denote the empirical mean of the first $s$ observations of arm $k$ made after time $t=\hat{\tau}^{(i)}+1$.
-Rewriting the sum in $t$ as the sum of consecutive intervals $[\tau^{(i)}+1, \tau^{(i+1)}]$,
+We let $\widetilde{\mu}_{k,s}^{(i)}$ denote the empirical mean of the first $s$ observations of arm $k$ made after time $t=\widehat{\tau}^{(i)}+1$.
+Rewriting the sum in $t$ as the sum of consecutive intervals $[\tau^{(i)}+1, \tau^{(i+1)}]$, we obtain
 %
@@
     (B) & \leq \bE\Big[\indic(\cE_T)\sum_{k=1}^{\Upsilon_T} \sum_{t=\tau^{(i)}}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
-     & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\hat{\tau}^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+     & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\widehat{\tau}^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
     & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\indic\left(A(t)=i,\UCB_k(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
     & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
-    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\hat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
+    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=wide\hat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
 \end{align*}%
 \begin{align*}
-    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cC^{(i)}) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big].
+    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(wide\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
+    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cC^{(i)}) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big].
 \end{align*}
 
-Conditionally to $\cF_{\widehat{\tau}^{(i)}}$, when $\cC^{(i)}$ holds, for $s \in [n_k(\tau^{(i+1)})]$, $\tilde{\mu}_{k,s}^{(i)}$ is the empirical mean from \iid{} observations of mean $\mu_k^{(i)}$.
+Conditionally to $\cF_{\widehat{\tau}^{(i)}}$, when $\cC^{(i)}$ holds, for $s \in [n_k(\tau^{(i+1)})]$, $\widetilde{\mu}_{k,s}^{(i)}$ is the empirical mean from \iid{} observations of mean $\mu_k^{(i)}$.
 Therefore, introducing $\widehat{\mu}_s$ as a sequence of \iid{} random variables with mean $\mu_k^{(i)}$, it follows from the law of total expectation that
@@
 
-% We introduce $\widehat{\mu}_{k,s}^{(i)}$ as the empirical mean of the first $s$ observations of arm $k$ made after $t=\tau^{(i)}+1$ and $\tilde{\mu}_{k,s}^{(i)}$ as the empirical mean of the first $s$ observations of arm $k$ made after time $t=\hat{\tau}^{(i)}+1$.
+% We introduce $\widehat{\mu}_{k,s}^{(i)}$ as the empirical mean of the first $s$ observations of arm $k$ made after $t=\tau^{(i)}+1$ and $\widetilde{\mu}_{k,s}^{(i)}$ as the empirical mean of the first $s$ observations of arm $k$ made after time $t=\widehat{\tau}^{(i)}+1$.
 
@@
 \begin{align*}
-    (a) & \leq \bP\left(\exists i, \exists s < t \leq n_k(\tau_k^{(i)}) : s \, \kl\left(\tilde{\mu}_{i,1:s}^{(i-1)},\tilde{\mu}_{i,1:t}^{(i-1)}\right) + (t - s) \, \kl\left(\tilde{\mu}_{i,s+1:t}^{(i-1)},\tilde{\mu}_{i,1:t}^{(i-1)}\right) > \beta(t, \delta) \;|\; \cC^{(i-1)}\right) \\
+    (a) & \leq \bP\left(\exists i, \exists s < t \leq n_k(\tau_k^{(i)}) : s \, \kl\left(\widetilde{\mu}_{i,1:s}^{(i-1)},\widetilde{\mu}_{i,1:t}^{(i-1)}\right) + (t - s) \, \kl\left(\widetilde{\mu}_{i,s+1:t}^{(i-1)},\widetilde{\mu}_{i,1:t}^{(i-1)}\right) > \beta(t, \delta) \;|\; \cC^{(i-1)}\right) \\
     & \leq \sum_{k=1}^K\bP\left(\exists s < t : s \, \kl(\widehat{\mu}_{1:s},\mu_k^{(i-1)}) + (t - s) \, \kl(\widehat{\mu}_{s+1:t}, \mu_k^{(i-1)}) > \beta(t, \delta)\right),
 \end{align*}
-with $\hat\mu_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_k^{(i-1)}$.
-Indeed, conditionally to $\cC^{(i-1)}$, the $n_k(\tau^{(i)})$ successive observations of arm $k$ arm starting from $\hat \tau^{(i)}$ are \iid{} with mean $\mu_k^{(i-1)}$.
+with $\widehat{\mu}_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_k^{(i-1)}$.
+Indeed, conditionally to $\cC^{(i-1)}$, the $n_k(\tau^{(i)})$ successive observations of arm $k$ arm starting from $\widehat{\tau}^{(i)}$ are \iid{} with mean $\mu_k^{(i-1)}$.
 Using Lemma~\ref{lem:6:ConcFirst}, term $(a)$ is upper bounded by $K\delta$.
@@
 (as the mapping $t\mapsto n_k(t) - n_k(\tau^{(i)})$ is non-decreasing, is $0$ at $t=\tau^{(i)}$ and its value at $\tau^{(i)}+d^{(i)}$ is larger than $\overline{r}$).
-Using that \[\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right) \subseteq \left(\widehat{\tau}^{(i)} \geq \overline{t}\right)\]
+Using that $\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right) \subseteq \left(\widehat{\tau}^{(i)} \geq \overline{t}\right)$
 the event $\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right)$ further implies that
 \[
-    n_k(\tau^{(i)}) \, \kl\left(\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})},\tilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right)
-    + \overline{r} \, \kl\left(\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})},\tilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right) \leq \beta(n_k(\tau^{(i)}) + \overline{r},\delta),
+    n_k(\tau^{(i)}) \, \kl\left(\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})},\widetilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right)
+    + \overline{r} \, \kl\left(\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})},\widetilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right) \leq \beta(n_k(\tau^{(i)}) + \overline{r},\delta),
 \]
-where $\tilde{\mu}^{k-1}_{k,s}$ denotes the empirical mean of the $s$ first observation of arm $k$ since the $(k-1)$-th restart $\hat{\tau}^{(i-1)}$ and  $\tilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
-Conditionally to $\cC^{(i-1)}$, $\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})}$ is the empirical mean of $n_k(\tau^{(i)})$ \iid{} replications of mean $\mu^{k-1}$, whereas $\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_k^{k}$.
+where $\widetilde{\mu}^{k-1}_{k,s}$ denotes the empirical mean of the $s$ first observation of arm $k$ since the $(k-1)$-th restart $\widehat{\tau}^{(i-1)}$ and  $wide\tilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
+Conditionally to $\cC^{(i-1)}$, $\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})}$ is the empirical mean of $n_k(\tau^{(i)})$ \iid{} replications of mean $\mu^{k-1}$, whereas $\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_k^{k}$.
 
@@
 \end{eqnarray*}
-
+%
 Introducing $\widehat{\mu}_{a,s}$ (resp.
@@
 \]
-where we have also used that $\tilde{\mu}^{k-1}_{i,n_k(\overline{t})} = \left(n_k(\tau^{(i)})\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})} + \overline{r}\tilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})} \right) / (n_k(\tau^{(i)}) + \overline{r})$.
+where we have also used that $\widetilde{\mu}^{k-1}_{i,n_k(\overline{t})} = \left(n_k(\tau^{(i)})\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})} + \overline{r}\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})} \right) / (n_k(\tau^{(i)}) + \overline{r})$.
 
@@
     \centering
-    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
     \caption{Problem $3$: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
@@
     \centering
-    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
     \caption{Problem $4$: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
     \caption{Pb $5$: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
@@
 Thompson sampling and \klUCB{} are efficient,
-and better than Discounted-\klUCB{} which is very inefficient.
+and better than Discounted-\klUCB{} which is inefficient.
 DTS and SW-\klUCB{} can sometimes be more efficient than their stationary counterparts, but perform worse than the Oracle and most actively adaptive algorithms.
-M-\klUCB{} and \CUSUMklUCB{} outperform the previous algorithms, but \GLRklUCB{} is often better.
+%
+For \klUCB{} indexes, M- and \CUSUM- outperform the previous algorithms, but \GLR- is often better.
 %
 On these problems, \GLRklUCB{} with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts.
-Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instance.
-The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17},
-and it usually performs similarly to \MklUCB, however it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.
-
-% \TODOL{I HAVE TO add the following algorithms, just to illustrate more Exp3.S}
+Note that on problem 2, all means change at every breakpoint, hence one could expect the Global variant to be more efficient, yet the experiments show the superiority of Location variant on every instance.
+The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17}.
+Exp3.S usually performs similarly to \MklUCB, but it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.
 
@@
 
-One can note that the best non-oracle strategies are actively adaptive,
+We highlight that the best non-oracle strategies are actively adaptive,
 thus the experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.
@@
 % -----------------------------------------------------------------
-\begin{figure}[b!]  % [htbp]
+\begin{figure}[h!]  % [htbp]
     \centering

### 2019-06-09 2-Chapters/Conclusion/conclusion.tex
OMG c'est 00h23 un dimanche soir et je bosse !

@@
 
-Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency.
+Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted in time and frequency.
 % , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.

### 2019-06-10 2-Chapters/3-Chapter/chapter3.tex
OMG c'est 00h23 un dimanche soir et je bosse !
Proofreading

@@
 
+\TODOL{Analyser les expériences les plus récentes et décrire mieux en 3.2, et refaire toutes les courbes pour 3.2 et 3.3}
+
 \paragraph{Abstract.}
@@
 It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
-%  and SMPyBandits is written in Python.
 It is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
@@
 \label{sec:3:Introduction}
-
-\TODOL{Analyser les expériences les plus récentes et décrire mieux en 3.2, et refaire toutes les courbes pour 3.2 et 3.3}
 
@@
 
-\textbf{Publications.}
+\textbf{References.}
 %
@@
 
-\textbf{Simulation loop.}
+\textbf{Simulation loop for single-player MAB.}
 %
@@
 
-\textbf{Simulation loop for multi-player bandit.}
-
+\textbf{Simulation loop for multi-player MAB.}
+%
 For Cognitive Radio dealing with OSA problems and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
@@
 %
-We have not yet tried to add support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
-%
-However, note that our library does support finite-state real-valued Markov MAB models, where arms do not correspond to distributions but to Markov chains, in the two variants of \emph{rested} or \emph{restless} Markov chains, as introduced by \cite{Anantharam87b}.
+We have not yet added support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
+%
+However, note that our library does support finite-state real-valued Markov MAB models, where arms correspond to Markov chains \cite{Norris98}, in the two \emph{rested}/\emph{restless} variants, as introduced by \cite{Anantharam87b}.
 
@@
 
-Simulations are easily executed, \eg, Listing~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
-It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
-Environment variables (\texttt{N=1000} etc) in the command line are not required, but they are convenient.
-
 % https://tex.stackexchange.com/a/12430/
@@
 \end{small}
+
+Simulations are easily executed, \eg, Listing~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
+It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
+Environment variables (\texttt{N=1000} etc) in the command line are not required, but they are convenient.
 
@@
 
--- In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
+\begin{itemize}
+    \item
+In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
 We designed a variant of the Exp4 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
@@
 
--- In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+    \item
+In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they obtain logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
 
--- In \cite{Besson2018DoublingTricks} (quickly presented in Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
-In sequential learning, an algorithm is said to be ``anytime'' if it does not need to know the horizon $T$ of the experiments.
+    \item
+In \cite{Besson2018DoublingTricks} (quickly presented in Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
+% In sequential learning, an algorithm is said to be ``anytime'' if it does not need to know the horizon $T$ of the experiments.
 A well-known trick for transforming any ``non-anytime'' algorithm to an ``anytime'' variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
@@
 
--- In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
+    \item
+In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
 We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
-We designed the \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.GLR_UCB.html}{GLRklUCB}} policy, and implemented most of the policies from state-of-the-art research on passively or active adaptive policies, both adversarial- or stochastic-based, designed to tackle the piece-wise stationary MAB problems.
+We designed the \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.GLR_UCB.html}{GLRklUCB}} policy, and implemented most of the state-of-the-art passive or active adaptive policies, both adversarial- or stochastic-based, designed to tackle the piece-wise stationary MAB problems.
 We also implemented a large benchmark of different problems.
+\end{itemize}
 
@@
 
-In this section, we use the SMPyBandits library to compare experimentally various state-of-the-art single-player algorithms on some multi-armed bandit problems.
-First, we detail the list of different algorithms that we compare, and the different problems on which they are all tested.
-We give the results of the numerical experiments, in terms of mean regret at the end of the experiment of different horizons $T$.
+In this section, we use the SMPyBandits library to compare experimentally various state-of-the-art (single-player) algorithms on some MAB problems.
+We first detail the list of different algorithms that we compare.
+% , and the different problems on which they are all tested.
+We then give the results of the numerical experiments, in terms of mean regret at the end of the experiment of different horizons $T$.
 Additional results in terms of real measurements of time and memory are given and discussed in the next Section~\ref{sec:3:timeAndMemoryCosts}.
-
-Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on two algorithms depending if simplicity or efficiency is favored:
-\begin{itemize}
-    \item
-    \UCB{} is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
-    \item
-    \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when efficiency is favored.
-\end{itemize}
+% %
+% Finally, we give some observations on the results, and the main take-away message of this section is the following: in all the rest of this thesis, we focus on two algorithms depending if simplicity or efficiency is favored:
+% \begin{itemize}
+%     \item
+%     \UCB{} is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
+%     \item
+%     \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when efficiency is favored.
+% \end{itemize}
 
@@
 For each of them, we give a bibliographic reference, that corresponds to a recent article studying it and not the first one which introduced it.
-We give the chosen tuning of the parameters, for parametric algorithms.
+We give the chosen tuning of its parameters, for parametric algorithms.
 Algorithms that are ``not anytime'' use the exact value of the horizon $T$, and when increasing values of $T$ are studied for the same problem, they use the correct successive values.
@@
     %  Algorithm 1 is
-    $\varepsilon$-greedy \cite{Bubeck12}, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily) (\texttt{EpsilonDecreasing}).
+    $\varepsilon$-greedy \cite{Bubeck12} (\texttt{EpsilonDecreasing}), shown in Algorithm~\ref{algo:2:naiveStrategies} above, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily).
     It has a very small cost in terms of time and memory, but it usually achieves linear regret.
@@
     %  Algorithm 2 is
-    Explore-then-Commit \cite{GarivierETC2016}, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems) (\texttt{ETC\_KnownGap}).
-    Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but only obtain large (linear) regret.
+    Explore-then-Commit \cite{GarivierETC2016} (\texttt{ETC\_KnownGap}), shown in Algorithm~\ref{algo:2:naiveStrategies} above, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems).
+    Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but most of the times it only obtains large (linear) regret.
 
@@
     %  Algorithm 3 is
-    $\mathrm{Exp}3^{++}$ from \cite{Seldin17}, using $\alpha=3$ and $\beta=256$ as advised (\texttt{Exp3PlusPlus}).
-    It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+    $\mathrm{Exp}3^{++}$ from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
+    It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances.
+    % meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
 
@@
     %  Algorithm 4 is
-    $\UCB_1$ from \cite{Auer02}, using $\alpha=1$ (\texttt{UCBalpha}).
+    $\UCB_1$ from \cite{Auer02} (\texttt{UCBalpha}), shown in Algorithm~\ref{algo:2:indexPolicy} above, using $\alpha=1$.
     It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
@@
     %  Algorithm 5 is
-    \klUCB{} from \cite{KLUCBJournal}, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern} (\texttt{klUCB}).
-    It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret.
+    \klUCB{} from \cite{KLUCBJournal} (\texttt{klUCB}), also shown in Algorithm~\ref{algo:2:indexPolicy} above, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern}.
+    % It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret.
     % with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
@@
     %  Algorithm 6 is
-    Thompson sampling from \cite{Kaufmann12Thompson}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{Thompson}).
+    Thompson sampling from \cite{Kaufmann12Thompson} (\texttt{Thompson}), using a Beta prior and posterior, initially uniform (\ie, $\pi_k(0)=\mathrm{Beta}(1,1)$).
     % It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
-    It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on C or Cython code, highly optimized).
+    It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on highly optimized C or Cython code).
 
@@
     %  Algorithm 7 is
-    Bayes-UCB from \cite{Kaufmann12BUCB}, using a Beta prior and posterior, initially uniform (\ie, $\mathrm{Beta}(1,1)$) (\texttt{BayesUCB}).
+    Bayes-UCB from \cite{Kaufmann12BUCB} (\texttt{BayesUCB}), using a Beta prior and posterior, initially uniform, \ie, $\pi_k(0)=\mathrm{Beta}(1,1)$.
     % It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
-    It is comparable to Thompson sampling in terms of memory complexity, but slower in terms of time complexity as computing a quantile is more costly than sampling from a distribution.
+    It is comparable to Thompson sampling in terms of memory complexity, but slower as computing a quantile is more costly than sampling from a distribution.
     But in particular for Bernoulli distributed arms and for Beta posteriors and priors, Bayes-UCB is only about twice as slow as Thompson sampling.
@@
     %  Algorithm 8 is
-    AdBandits from \cite{Truzzi13}, using $\alpha=1$ (\texttt{AdBandits}).
+    AdBandits from \cite{Truzzi13} (\texttt{AdBandits}), using $\alpha=1$.
     % It follows the same assumptions as Thompson sampling and Bayes-UCB, and uses a Beta posterior on each arm.
@@
 
-We focussed here on the most well known algorithms, and two others that are efficient but less known (AdBandits and BESA),
-but to illustrate that SMPyBandits implements many other solutions to the MAB problem proposed in very recent works, we explain in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
+We focussed here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
+To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
 They are more recent (some from $2016$, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or are based on a different point-of-view.
-The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} algorithms in the rest of this thesis, as they stay comparable with the state-of-the-art algorithms but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to PHE or RCB)
-
-\paragraph{Two different problems.}
-%
-We consider two different problems, that are described here for a generic value of $K \geq 2$ the number of arms.
-For simplicity, we focus on Bernoulli distributed arms, even if similar results were observed for other distributions, in particular for truncated or untruncated Gaussian distributions.
-
-% \begin{itemize}
-    % \item
-    \textbf{Problem 1} considers $K$ arms of equally spaced means, in $[0,1]$.
-    It uses a minimum mean of $0.05$ and a maximum mean of $0.95$.
-    For instance, for $K=2$, $\bm{\mu}=[0.05, 0.95]$ and for $K=5$, $\bm{\mu} = [0.05, 0.275, 0.5, 0.725, 0.95]$.
-
-    % \item
-    \textbf{Problem 2} takes another approach. Instead of simulating $1000$ times the same problem, it generates a new random problem for each independent run.
-    It considers a randomly generated vector of means, each being sampled uniformly at random in $[0,1]$, with a minimum gap of $\Delta^{\min}$, $\bm{\mu} \sim \cE(\Delta^{\min},\mu_{\min},\mu_{\max})$, as well as minimum and maximum values of $\mu_{\min}$ and $\mu_{\max}$.
+The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with the state-of-the-art algorithms but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).
+
+\paragraph{Randomly sampled problems.}
+%
+For brevity, we prefer to focus on a single kind of distributions (Bernoulli),
+but similar results were observed for other distributions, in particular for (possibly truncated) Gaussian distributions.
+%
+% % \begin{itemize}
+%     % \item
+%     \textbf{Problem 1} considers $K$ arms of equally spaced means, in $[0,1]$.
+%     It uses a minimum mean of $0.05$ and a maximum mean of $0.95$.
+%     For instance, for $K=2$, $\bm{\mu}=[0.05, 0.95]$ and for $K=5$, $\bm{\mu} = [0.05, 0.275, 0.5, 0.725, 0.95]$.
+
+%     % \item
+% \textbf{Problem 2} takes another approach.
+For a fixed number of arms $K \geq 2$,
+instead of simulating a large number of times the same problem,
+we generate a new random problem for each independent run.
+%
+We consider a randomly generated vector of means, each being sampled uniformly at random in $[0,1]$, with a minimum gap of $\Delta^{\min}$, $\bm{\mu} \sim \cE(\Delta^{\min},\mu_{\min},\mu_{\max})$, as well as minimum and maximum values of $\mu_{\min}$ and $\mu_{\max}$.
 This space $\cE_{\Delta^{\min}}$ is defined as $\{ \bm{\mu} \in [\mu_{\min}, \mu_{\max}]^K : \min_{k\neq k'} |\mu_k - \mu_{k'}| \geq \Delta^{\min} \}$.
@@
 
-\paragraph{Other parameters of the experiments.}
-
+\textbf{Other parameters of the experiments.}
+%
 %    FIXME update if needed?
-We consider increasing values for the horizon, and we focus on $7$ exponentially-spaced values: $T=1000$, $T=2500$, $T=5000$, $T=10000$, $T=20000$, $T=25000$ and $T=50000$.
-We consider as well an increasing number of arms, and we focus again on $7$ values: $K=2$, $K=4$, $K=8$, $K=12$, $K=16$, $K=20$ and $K=32$.
-For all experiments, we run $N=100$ independent simulations.
+On the one hand, we fix $K=8$ and we consider increasing values for the horizon, from $T=1000$ to $T=50000$.
+On the other hand, we fix $T=5000$ and we consider as well an increasing number of arms, from $K=2$ to $K=32$.
+% , and we focus again on $7$ values: $K=2$, $K=4$, $K=8$, $K=12$, $K=16$, $K=20$ and $K=32$.
+For all experiments, we run $N=1000$ independent simulations.
 We show in the snippet of code~\ref{lst:3:BashCodeToLaunchLargeExperiments} in the Appendix below more details about how to run these experiments.
 
-\TODOL{Dans les figures où je montre une seule valeur de $T$ ou de $K$, et $R_T$, $\cT_T$ ou $\cM_T$ en fonction de $K$ ou de $T$, il faudrait avoir plus de valeurs à montrer, et pas une progression exponentielle...
+% \TODOL{Dans les figures où je montre une seule valeur de $T$ ou de $K$, et $R_T$, $\cT_T$ ou $\cM_T$ en fonction de $K$ ou de $T$, il faudrait avoir plus de valeurs à montrer, et pas une progression exponentielle...
 J'ai relancé avec :
-    - $T=10000$ et $K = 6, 10, 14, 18, 22, 24, 26, 28, 32$ pour couvrir $2 \dots 32$ avec un pas de $2$,
-    - $K=32$ et $T = 15000, 30000, 35000, 40000, 45000$ pour couvrir $5000 \dots 50000$ avec un pas de $5000$,
-}
+%     - $T=10000$ et $K = 6, 10, 14, 18, 22, 24, 26, 28, 32$ pour couvrir $2 \dots 32$ avec un pas de $2$,
+%     - $K=32$ et $T = 15000, 30000, 35000, 40000, 45000$ pour couvrir $5000 \dots 50000$ avec un pas de $5000$,
+% }
 
@@
 
-\subsection{Observations and summary of the experiments}
+\subsection{Summary of the experiments}
 
@@
 
-\TODOL{Use UCB and klUCB, it's enough to be efficient!}
-\TODOL{Regret usually scales as $\bigO{K}$ and $\bigO{\log T}$ !}
+We have been able to check empirically that the regret of all the efficient algorithm indeed scales as predicted by the theory, that is
+linearly in the number of arms, and logarithmically in the horizon, \ie, $R_T = \bigO{K \log T}}$.
+%
+The main take-away message is the following: in all the rest of this thesis, we focus on two algorithms depending if simplicity or efficiency is favored:
+\begin{itemize}\tightlist
+    \item
+    \UCB{} is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
+    \item
+    \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when efficiency is favored.
+\end{itemize}
 
@@
 
-% \TODOL{Write text for this Section, write configuration.py file, run short-term experiment, include them, if happy about it, run longer ang larger experiments}
-
 In this section, we report additional experiment results from the simulations described in the previous Section~\ref{sec:3:reviewSPAlgorithms}.
 Instead of studying on the efficiency of the algorithms (\ie, their regret), we report results of real measurements in terms of computational time as well as memory storage.
-%
-The goal of this section is to verify that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
-
+% %
+% The goal of this section is to verify that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
+%
+%
 While the results reported in the previous section should not depend on the implementation of the different algorithms, the results in this section concern real measurements of both time and memory consumptions of the simulation software used for these simulations.
 Hence, the reported results highly depend on many factors, including how the code is written, and where and when it is run.
-
-\begin{itemize}
-    \item Quality and optimization of the code:
-    The same level of optimization is used in all the codebase of SMPyBandits, and most of the time, the code is pure and naive Python and was not optimized.
-    Mathematical functions (\eg, $\sqrt{\bullet}, \log$) and random numbers used are using \texttt{numpy} and \texttt{numpy.random} modules \cite{numpy}, which are based on C and Cython code \cite{cython} and can be considered highly efficient.
-    Computations of Kullback-Leibler and indexes for \klUCB{} algorithms are based on a compiled CPython extension written in C \cite{python} and can also be considered highly efficient.
-    As we can safely affirm that the code of the different algorithms has the same level of optimization, the comparison is fair, and we do not favor any algorithm in their implementation in SMPyBandits.
-
-    \item About the experimental environment:
-    the exact measurements used for the figures displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
-    As long as the different algorithms and simulations are performed on the same environment, the comparison and conclusions we make from them are fair and do make sense.
-
-\end{itemize}
-
-\paragraph{About the measurement protocol.}
-%
-To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.
-\begin{itemize}
-    \item For time, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
-    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in[T]$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
-    This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.
-    \item For memory, we use two different approaches whether the code runs on a UNIX or a Windows system.
-    On UNIX, the \texttt{resource} module allows to measure the memory of the current process, and by counting the difference in memory used by the \texttt{Evaluator} object before and after the ``for'' loop, we can track the memory used by the algorithm to store all its interior variables.
-\end{itemize}
-In both cases, the ``for'' loop takes some time and stores many variables, but the only difference in terms of both time or memory between two algorithms is explained by the difference in the implementation of the algorithms.
-%     def time()
-% time() -> floating point number
-
-% Return the current time in seconds since the Epoch. Fractions of a second may be present if the system clock provides them.
-
-% sys.getsizeof(pickle.dumps(policy))
-% return resource.getrusage(resource.RUSAGE_SELF).ru_maxrss
-
-\paragraph{Comparing measurements.}
-But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
-Relative difference for costs do not depend much on the code and the machine used for the simulation.
-%
-% The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
-The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
-%
-Thus, we can check that efficient algorithms have a running time independent on the horizon $T$, instead of just observing a linear dependency, and we can check that efficient algorithms have a memory cost independent on the horizon.
-All the considered algorithms in this Chapter are efficient in this aspect, but in Chapter~\ref{chapter:6} we study algorithms that essentially have to store, and run computations, on an increasing number of observations (\eg, \CUSUMUCB), so their computation cost at time $t$ increases (linearly) with $t$, and thus they are more costly.
-%
-We can also check that efficient algorithms have a running time and a memory cost linear in the number of arms $K$, while more complex algorithms like BESA suffer from an exponential blow-up of their complexity when $K$ increases.
-
-\paragraph{Normalizing data.}
-%
-The two figures below regret normalized data, in the following way.
-For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
-Simply considering the average of such measurements makes no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
-%
-We know that efficient algorithms are expected to follow these patterns:
-\begin{itemize}
-    \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
-    \item The total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
-    \item The total memory cost $\cM_T$ should scale as $\cO(K)$, independent of the horizon.
-\end{itemize}
-Thus, on the one hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
-%
-On the other hand, when we plot a quantity ($R_T$, $\cT_T$, $\cM_T$) as a function of $T$ (resp. of $K$) we only normalize by $K$ (resp. by $T$).
-
-\subsection{Computational time}
+We take precautions to ensure the fairness of the comparison between the different algorithms, as detailed in Appendix~\ref{sub:3:precautionsTimeMemory}.
+
+% \subsection{Computational time}
 
@@
 
-\subsection{Memory cost}
+% \subsection{Memory cost}
 
@@
 
-\TODOL{Use UCB and klUCB}
-\TODOL{Time complexity usually scales as $\bigO{KT}$}
-\TODOL{Memory complexity usually scales as $\bigO{K}$}
+The main take-away message is the following:
+the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
 
@@
 % --------------------------------------------------------------
-\subsection{Description of seven more recent algorithms for experiments in Section~\ref{sec:3:reviewSPAlgorithms}}
+\subsection{Additional details about numerical experiments in Section~\ref{sec:3:reviewSPAlgorithms}}
 \label{sub:3:additionalExperiments}
 
-This Appendix section describes $7$ more algorithms that are also compared with the $9$ presented above in Section~\ref{sec:3:reviewSPAlgorithms}.
+This Appendix section starts by describing $7$ more algorithms that are also compared with the $9$ presented above in Section~\ref{sec:3:reviewSPAlgorithms}.
 
@@
     % Algorithm 11 is
-    Approximated Finite-Horizon Gittins from \cite{Lattimore16a}, using $\alpha=1$ (\texttt{ApproximatedFHGittins}).
+    Approximated Finite-Horizon Gittins from \cite{Lattimore16a} (\texttt{ApproximatedFHGittins}), using $\alpha=1$.
     It mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
@@
     % Algorithm 12 is
-    $\UCB\dagger$ from \cite{Lattimore2018refining}, using $\alpha=1$ (\texttt{UCBdagger}).
+    $\UCB\dagger$ from the same author \cite{Lattimore2018refining} (\texttt{UCBdagger}), using $\alpha=1$.
     It mimics the \UCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
@@
     % Algorithm 13 is
-    Anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018},
-    is a very recent variant of the \klUCB{} algorithm \cite{KLUCBJournal}
-    (\texttt{klUCBswitchAnytime}).
+    Anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018}
+    (\texttt{klUCBswitchAnytime}),
+    is a very recent variant of the \klUCB{} algorithm \cite{KLUCBJournal}.
     It mimics the \klUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
@@
     % Algorithm 14 is
-    \textsc{Tsallis-Inf} from \cite{Zimmert2018}, using $\alpha=1/2$ (it seems to be the most efficient of the algorithms described in the article) (\texttt{TsallisInf}).
+    \textsc{Tsallis-Inf} from \cite{Zimmert2018} (\texttt{TsallisInf}), using $\alpha=1/2$ (it seems to be the most efficient of the algorithms described in the article).
     It is based on the online mirror descent algorithm with a Tsallis entropy regularizer, and it is anytime.
@@
     % Algorithm 15 is
-    $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019}, using $\alpha=1$ and perturbations uniformly sampled in $[0,1]$.
-We prefer to use the simplest of the algorithms described in the article, and other algorithms in the same article used other families of distributions for the perturbations (\texttt{RCB}).
+    $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019} (\texttt{RCB}), using $\alpha=1$ and perturbations uniformly sampled in $[0,1]$.
+We prefer to use the simplest of the algorithms described in the article, and other algorithms in the same article used other families of distributions for the perturbations.
     It was analyzed and found to be order-optimal for problem-dependent bounds, and empirically we found that it is usually performs slightly worse than \UCB, in terms of regret, time and memory.
@@
     % Algorithm 16 is
-    $\mathrm{PHE}$ (Perturbed-History Exploration) from \cite{KvetonSzepesvari2019}, using a perturbation scale of $0.5$ as advised (\texttt{PHE}).
+    $\mathrm{PHE}$ (Perturbed-History Exploration) from \cite{KvetonSzepesvari2019} (\texttt{PHE}), using a perturbation scale of $0.5$ as advised.
     The algorithm adds $\cO(t)$ \iid{} pseudo-rewards to its history in round $t$, and then pulls the arm with the highest estimated value in its perturbed history.
@@
 
-\paragraph{Interpretation.}
-
-\TODOL{A faire}
+% \paragraph{Interpretation.}
+
+% \TODOL{A faire}
 
 \paragraph{Bash command to run these experiments.}
-
-Finally, we give below in Listing~\ref{lst:3:BashCodeToLaunchLargeExperiments} the Bash command that was used to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
+We give below in Listing~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
 It uses \texttt{for} loops and environment variables from Bash, and not Python, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $N$, $T$ and $K$.
@@
 $ for N in 4 100 1000;        do \
-  for BAYES in False True;        do \
-  for T in 1000 2500 5000 10000 20000 50000; do \
-  for K     in 2 4 8 12 16 20 32; do \
+    for T in $(seq 5000 5000 50000); do \
+      for K in $(seq 2 2 32); do \
         DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
-                         N=$N BAYES=$BAYES T=$T K=$K \
+          N=$N BAYES=True  T=$T K=$K \
           python3 main.py configuration.py \
         && cp logs/main_py3_log.txt \
-          logs/main_py3_log__N${N}_BAYES${BAYES}_T{T}_K{K}.txt \
-  done; done; done; done
+              logs/main_py3_log__N${N}_BAYES_T{T}_K{K}.txt \
+  done; done; done
     \end{minted}
@@
 % --------------------------------------------------------------
-\subsection{Additional remark: about parallelism in SMPyBandits}
+\subsection{Methodology details for measurements of time and memory}
+\label{sub:3:precautionsTimeMemory}
+
+The results reported in Section~\ref{sec:3:timeAndMemoryCosts} highly depend on many factors, including how the code is written, and where and when it is run.
+We detail both below:
+
+\begin{itemize}
+    \item Quality and optimization of the code:
+    The same level of optimization is used in all the codebase of SMPyBandits, and most of the time, the code is pure and naive Python and was not optimized.
+    Mathematical functions (\eg, $\sqrt{\bullet}, \log$) and random numbers used are using \texttt{numpy} and \texttt{numpy.random} modules \cite{numpy}, which are based on C and Cython code \cite{cython} and can be considered highly efficient.
+    Computations of Kullback-Leibler and indexes for \klUCB{} algorithms are based on a compiled CPython extension written in C \cite{python} and can also be considered highly efficient.
+    As we can safely affirm that the code of the different algorithms has the same level of optimization, the comparison is fair, and we do not favor any algorithm in their implementation in SMPyBandits.
+
+    \item About the experimental environment:
+    the exact measurements used for the figures displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
+    As long as the different algorithms and simulations are performed on the same environment, the comparison we make from them are fair and do make sense.
+\end{itemize}
+
+\paragraph{About the measurement protocol.}
+%
+To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.
+\begin{itemize}
+    \item For time, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
+    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in[T]$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
+    This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.
+    \item For memory, we use two different approaches whether the code runs on a UNIX or a Windows system.
+    On UNIX, the \texttt{resource} module allows to measure the memory of the current process, and by counting the difference in memory used by the \texttt{Evaluator} object before and after the ``for'' loop, we can track the memory used by the algorithm to store all its interior variables.
+\end{itemize}
+In both cases, the ``for'' loop takes some time and stores many variables, but the only difference in terms of both time or memory between two algorithms is explained by the difference in the implementation of the algorithms.
+%     def time()
+% time() -> floating point number
+
+% Return the current time in seconds since the Epoch. Fractions of a second may be present if the system clock provides them.
+
+% sys.getsizeof(pickle.dumps(policy))
+% return resource.getrusage(resource.RUSAGE_SELF).ru_maxrss
+
+\paragraph{Comparing measurements.}
+But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
+Relative difference for costs do not depend much on the code and the machine used for the simulation.
+%
+% The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
+The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
+%
+Thus, we can check that efficient algorithms have a running time independent on the horizon $T$, instead of just observing a linear dependency, and we can check that efficient algorithms have a memory cost independent on the horizon.
+All the considered algorithms in this Chapter are efficient in this aspect, but in Chapter~\ref{chapter:6} we study algorithms that essentially have to store, and run computations, on an increasing number of observations (\eg, \CUSUMUCB), so their computation cost at time $t$ increases (linearly) with $t$, and thus they are more costly.
+%
+We can also check that efficient algorithms have a running time and a memory cost linear in the number of arms $K$, while more complex algorithms like BESA suffer from an exponential blow-up of their complexity when $K$ increases.
+
+\paragraph{Normalizing data.}
+%
+The two figures below regret normalized data, in the following way.
+For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
+Simply considering the average of such measurements makes no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
+%
+We know that efficient algorithms are expected to follow these patterns:
+\begin{itemize}\tightlist
+    \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
+    \item The total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
+    \item The total memory cost $\cM_T$ should scale as $\cO(K)$, independent of the horizon.
+\end{itemize}
+Thus, on the one hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
+%
+On the other hand, when we plot a quantity ($R_T$, $\cT_T$, $\cM_T$) as a function of $T$ (resp. of $K$) we only normalize by $K$ (resp. by $T$).
+
+% --------------------------------------------------------------
+\subsection{Final remark: how we used parallelism in SMPyBandits}
 \label{sub:3:parallelSimulations}
@@
 
-\TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ?}
+\TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? C'est vraiment pas crucial.
+    C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
+}
 

### 2019-06-10 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Proofreading

@@
 	\item
-	For $0 < \gamma < 1$, the function $g(x) = x \gamma^x$ is quasi-convex on $[0,\infty)$, \ie, $g(\eta x + (1-\eta)y) \leq \max(g(x), g(y))$ for any $x,y \in [0,\infty)$ and $\eta \in [0,1]$ (definition from \cite{Luenberger68}).
+	For $0 < \gamma < 1$, the function $g(x) \eqdef x \gamma^x$ is quasi-convex on $[0,\infty)$, \ie, $g(\eta x + (1-\eta)y) \leq \max(g(x), g(y))$ for any $x,y \in [0,\infty)$ and $\eta \in [0,1]$ (definition from \cite{Luenberger68}).
     Indeed, $g(\eta x + (1-\eta)y) = \eta \left[ x (\gamma^x)^{\eta} \right] \gamma^{((1-\eta)y)} + (1-\eta)\left[ y (\gamma^y)^{1-\eta}\right] \gamma^{\eta x}$, and $\gamma^{((1-\eta)y)} \leq 1$ and $\gamma^{\eta x} \leq 1$.
@@
     \item
-    The function $f(D_1, \dots, D_{K}) = \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{K}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
+    The function $f(D_1, \dots, D_{K}) \eqdef \sum\limits_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is quasi-convex in each of its coordinates, on $[0,\infty)^{K}$, as a sum of component-wise quasi-convex functions (with $\gamma = (1 - p) \in (0, 1)$, thanks to the first point).
     \hfill{}$\square$
@@
 
-Let us now prove the expression for $D_i^*$ given in \eqref{eq:41:Dilambda}.
+Let us now prove the expression for $D_i^*$ given above in \eqref{eq:41:Dilambda}.
 
@@
 	\item
-	If $\boldsymbol{D} = (D_1,\dots,D_{K})$,
-    the Lagrangian is denoted $\mathcal{L}(\boldsymbol{D}, \lambda) = f(\boldsymbol{D}) + \lambda(D - \sum_{i=1}^{K} D_i)$, and its derivative w.r.t. $D_i$ is
+	If $\boldsymbol{D} \eqdef (D_1,\dots,D_{K})$,
+    the Lagrangian is denoted $\mathcal{L}(\boldsymbol{D}, \lambda) \eqdef f(\boldsymbol{D}) + \lambda(D - \sum\limits_{i=1}^{K} D_i)$, and its derivative w.r.t. $D_i$ is
     $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) = (1-p)^{S_i + D_i - 1} + \log(1 - p) D_i (1-p)^{S_i + D_i - 1} - \lambda$.
@@
     By using the $\mathcal{W}$-Lambert function $\mathcal{W}$ \cite{Corless96}, reciprocal of $y \mapsto \mathrm{e}^{y}y$, we get $x = y - 1 = \mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1$.
-    So the gradient of the Lagrangian is zero iff $D_i^* = \max(0, 1/\log(1-p) x) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1\right]^{+}$, because $D_i^*$ has to be non-negative.
+    So the gradient of the Lagrangian is zero iff $D_i^* = \max(0, x/\log(1-p)) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1\right]^{+}$, because $D_i^*$ has to be non-negative.
     This gives the $i$-th coordinate of the unique saddle point of $f(\boldsymbol{D})$,

### 2019-06-10 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Proofreading

@@
 where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{t=k}^{k'} X_t$ denotes the average of the observations $X_t$ collected between the instants $k$ and $k'$.
+
+\TODOL{Si on veut, on peut déplacer cette preuve en Annexe, voir même l'enlever, elle n'est pas spécialement palpitante et pas requise pour la suite !}
 
@@
         % XXX Input, data and output
-        \KwIn{\emph{Problem parameters}: $T\in\N^*$, $K\in\N^*$\;}
+        % \KwIn{\emph{Problem parameters}: $T\in\N^*$, $K\in\N^*$\;}
         \KwIn{\emph{Parameters}: exploration probability $\alpha \in (0, 1)$, confidence level $\delta>0$\;}
@@
 
-\paragraph{Other forced exploration schemes?}
+\textbf{Other forced exploration schemes?}
 %
@@
 
+\textbf{\textcolor{red}{Warning}:}
+%
 We highlight that \emph{this result is finite-time} and not asymptotic, even if it uses the notation $\bigO{\sqrt{\ln(T)}}$.
@@
     For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriods},
-    with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, the following holds.
+    with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, the following regret bounds hold.
     \begin{enumerate}
-        \item Choosing $\alpha = \sqrt{\ln(T) / T}$, $\delta = 1 / \sqrt{T}$ gives
+        \item Choosing $\alpha = \sqrt{\ln(T) / T}$, $\delta = 1 / \sqrt{T}$ (with no prior knowledge of $\Upsilon_T$ or $\EffChange$) gives
         \begin{equation}
-            R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \EffChange\sqrt{T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)},
-        \end{equation}
-        \item Choosing $\alpha = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ gives
+            R_T = \bigO{\frac{K}{\left(\Delta^{\text{change}}\right)^2} \textcolor{red}{\EffChange}\sqrt{T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2} \ln(T)},
+        \end{equation}
+        \item Choosing $\alpha = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and optimist guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
         \begin{equation}
-            R_T = \bigO{\frac{K^2}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)},
-        \end{equation}
-        \item Choosing $\alpha = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ gives
+            R_T = \bigO{\frac{K^{\textcolor{red}{2}}}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\Upsilon_T T\ln(T)} + \frac{K\Upsilon_T}{\left(\Delta^{\text{opt}}\right)^2}\ln(T)},
+        \end{equation}
+        \item Choosing $\alpha = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ (with prior knowledge of $\EffChange$) gives
         \begin{equation}
@@
         \end{equation}
-        \item Choosing $\alpha = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ gives
+        \item Choosing $\alpha = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and pessimist guess $\EffChange \simeq K \Upsilon_T$) gives
         \begin{equation}
-            R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{1}{\left(\Delta^{\text{opt}}\right)^2} \EffChange\ln(T)}.
+            R_T = \bigO{\frac{K}{\left({\Delta}^{\text{change}}\right)^2}\sqrt{\EffChange T\ln(T)} + \frac{\EffChange}{\left(\Delta^{\text{opt}}\right)^2} \ln(T)}.
         \end{equation}
@@
     & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(i)}) \sum_{t=wide\hat{\tau}^{(i)}}^{\tau^{(i+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right],
+    & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(i)}) \sum_{t=\widehat{\tau}^{(i)}}^{\tau^{(i+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right],
 \end{align*}
@@
 \begin{align*}
-    &\bE\left[\left.\indic(\cC^{(i)}) \sum_{t=wide\hat{\tau}^{(i)}}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \right| \cF_{\widehat{\tau}^{(i)}}\right]\\
+    &\bE\left[\left.\indic(\cC^{(i)}) \sum_{t=\widehat{\tau}^{(i)}}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \right| \cF_{\widehat{\tau}^{(i)}}\right]\\
     & = \indic(\cC^{(i)}) \sum_{t=\widehat{\tau}^{(i)}}^{\tau^{(i+1)}} \bE\left[\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right) \;|\; \cF_{\widehat{\tau}^{(i)}}\right] \\
@@
     & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
-    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=wide\hat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
-\end{align*}%
-\begin{align*}
-    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(wide\tilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
+    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
+% \end{align*}%
+% \begin{align*}
+    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
     & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cC^{(i)}) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big].
@@
 \]
-where $\widetilde{\mu}^{k-1}_{k,s}$ denotes the empirical mean of the $s$ first observation of arm $k$ since the $(k-1)$-th restart $\widehat{\tau}^{(i-1)}$ and  $wide\tilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
+where $\widetilde{\mu}^{k-1}_{k,s}$ denotes the empirical mean of the $s$ first observation of arm $k$ since the $(k-1)$-th restart $\widehat{\tau}^{(i-1)}$ and  $\widetilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
 Conditionally to $\cC^{(i-1)}$, $\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})}$ is the empirical mean of $n_k(\tau^{(i)})$ \iid{} replications of mean $\mu^{k-1}$, whereas $\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_k^{k}$.
@@
 %
-\begin{eqnarray*}
-    n_k(\tau^{(i)})& \in& \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} \eqdef \cI_k.
-\end{eqnarray*}
+\begin{equation*}
+    n_k(\tau^{(i)}) \in \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} \eqdef \cI_k.
+\end{equation*}
 %

### 2019-06-10 2-Chapters/3-Chapter/chapter3.tex
Typeset unit (B, KiB, s, µs) correctly, thanks to siunitx

@@
 We have been able to check empirically that the regret of all the efficient algorithm indeed scales as predicted by the theory, that is
-linearly in the number of arms, and logarithmically in the horizon, \ie, $R_T = \bigO{K \log T}}$.
+linearly in the number of arms, and logarithmically in the horizon, \ie, $R_T = \bigO{K \log T}$.
 %

### 2019-06-10 2-Chapters/4-Chapter/chapter4.tex
Typeset unit (B, KiB, s, µs) correctly, thanks to siunitx

@@
 As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly not stationary''.
-This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms (that is why we relax the hypothesis and only consider up-to $M\leqK$ players in Chapter~\ref{chapter:5}).
+This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms (that is why we relax the hypothesis and only consider up-to $M \leq K$ players in Chapter~\ref{chapter:5}).
 We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.

### 2019-06-11 2-Chapters/2-Chapter/chapter2.tex
OMG c'est 00h23 un dimanche soir et je bosse !
Commentaire Christophe Ch4.1 et 4.2

@@
 
-Instead of using a confidence radius $\xi_(t)$ based on Hoeffding's inequality,
+Instead of using a confidence radius $\xi_k(t)$ based on Hoeffding's inequality,
 the \klUCB{} algorithm rather use a specific concentration inequality for the considered exponential family, such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
-It yields improved upper confidence bounds (\ie, with smaller $\xi_(t)$) that are based on Kullback-Leibler divergence.
+It yields improved upper confidence bounds (\ie, with smaller $\xi_k(t)$) that are based on Kullback-Leibler divergence.
 The indexes are computed as follows, for an exploration function $f(t)$
@@
 \end{smallproof}
+
+% % FIXME
+% \newpage
+% % FIXME

### 2019-06-11 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Commentaire Christophe Ch4.1 et 4.2

@@
 We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
-This second approach is ideal, but not realistic in decentralized IoT networks, and it is just used here as a reference.
+This second approach is ideal, but not realistic the most common situation of decentralized co-located IoT networks, and it is just used here as a reference.
 Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
@@
 Messages can be sent in a fixed number $K\geq2$ of wireless channels,
-that are assumed to be orthogonal, that is, a message in channel $k$ will not cause collision to a message in another channel $j \neq i$.
+that are assumed to be orthogonal, that is, a message in channel $k$ will not cause collision to a message in another channel $j \neq k$.
 % (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
@@
 In order to simulate networks designed for the IoT, we consider a protocol \textbf{with no sensing} and no repetition of uplink messages.
-The gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) who succeeded in sending an uplink packet.
+The gateway is in charge of sending back an acknowledgement, after some fixed-time delay, to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) who succeeded in sending an uplink packet.
 %
@@
     Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one float number for each of the $K$ channels).
-    Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, to improve their successful communication rate automatically.
+    Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, in order to improve their successful communication rate automatically.
 \end{itemize}
@@
 As this problem is only interesting if devices are able to communicate reasonably efficiently with the gateway, we assume that devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
-Note that even unlicensed bands have such limitations, as for instance this is a strict requirement for any device using the $867\;\mathrm{MHz}$ ISM band in Europe (and its regulation also enforces a maximum transmission power, but we do not consider power transmission in this work).
+Note that even unlicensed bands have such limitations, as for instance this is a strict requirement for any device using the \SI{868}{\mega\hertz} ISM band in Europe (and its regulation also enforces a maximum transmission power, but we do not consider power transmission in this work).
 %
@@
 % \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
-Indeed, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$).
+\indente
+On the opposite, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$).
 Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
@@
 %
-We consider the network from the point of view of one dynamic device.
-Following a Bernoulli emission process, every time a dynamic device has to communicate with the gateway,
+All devices follow a Bernoulli emission process.
+Consider the network from the point of view of one dynamic device:
+every time a dynamic device has to communicate with the gateway,
 it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $A(t) = k \in[K]$.
@@
 % are $\mu_1,\dots,\mu_K$, the mean availability of the channels,
-are $S_1, \dots, S_K$ the (stationary) occupancy of the channels by the static devices,
+are $S_1, \dots, S_K$ which represent the (stationary) occupancy of the channels by the static devices,
 and they are unknown to each dynamic device, so to maximize their cumulated rewards, they must learn the distributions of the channels, in order to be able to progressively improve their respective successful communication rate.
@@
 %
-Before presenting the MAB algorithms used in the experiments, we present some natural baseline (reference) policies.
+Before presenting the MAB algorithms used in the experiments, we present some simple baseline (reference) policies.
 
@@
 This reflects a naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.
+Note that even nowadays this is still the solution implemented for real IoT devices deployed in new LoRaWAN networks.
 
@@
 
-If the oracle chooses a fixed configuration of dynamic devices, with $D_k$ devices affected to channel $k$ is fixed (in time, \ie, stationary),
-then this probability is computed as before:
+If the oracle chooses a fixed configuration of dynamic devices, it means that for each dynamic device the oracle affects it to a unique channel for all time steps.
+Then there is a number $D_k$ of devices affected to channel $k$ being fixed in time (\ie, stationary),
+and thus this probability is computed as before:
 %
@@
 For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
-Increasing the number of end-devices decreases the gap between the optimal and random policies: the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of \ref{fig:41:perf_learning}.
+Increasing the number of end-devices decreases the gap between the optimal and random policies:
+as expected intuitively, the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of Figure~\ref{fig:41:perf_learning}.
 
@@
     \includegraphics[scale=0.65]{perf_learning.eps}
-    \caption{Learning with \UCB{} and Thomson Sampling, with many dynamic devices. Both algorithms achieve close to optimal performance after a reasonable learning time.}
+    \caption{Learning with \UCB{} and Thomson Sampling, with many dynamic devices.
+        The learning gain, for each device, decreases with the proportion of dynamic devices in the network.
+        Note that the values ($16\%$ etc) depend on the repartition of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.
+        % Both algorithms achieve close to optimal performance after a reasonable learning time.
+    }
     \label{fig:41:perf_learning}
@@
 
-\textbf{Successful transmissions rate as a function of the dynamic devices.}
+\textbf{Successful transmissions rate as a function of the number of dynamic devices.}
 %
@@
 We can see that when the proportion of end-devices is low (\eg, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
-The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a lowest rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$).
+The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a lowest rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$),
+for the considered values of $S_1,\dots,S_K$.
 
@@
 Figure \ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
-Note that when this proportion increases, the assumptions of a stochastic MAB model are clearly violated, and there is no justification for the efficiency of TS and \UCB{} algorithms.
+Note that when this proportion increases, the assumptions of a stochastic MAB model are clearly violated, and there is no mathematical justification for the efficiency of TS and \UCB{} algorithms.
 Hence, it is surprising to have near optimal performance with stochastic MAB algorithms applied to partly or fully dynamic scenarios.

### 2019-06-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Commentaire Christophe Ch4.1 et 4.2
Cf mail avec Emilie sur la question posée à mon séminaire à PANAMA du 06 juin.

@@
     \end{equation}
+    with the convention that we \emph{enforce} $\left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] = 0$
+    in the case of $\widehat{\mu}_{1:s} = \widehat{\mu}_{s+1:n}$
+    (in particular, if $\widehat{\mu}_{1:n} \in \{0,1\}$, then $\widehat{\mu}_{1:s} = \widehat{\mu}_{s+1:n} = 0$ or $\widehat{\mu}_{1:n}$).
+    % Indeed in such cases, the $\kl(x,y)$
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
     & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
-    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
+    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]\\
 % \end{align*}%

### 2019-06-12 1-Preamble/5-Abstract/abstract.tex
Multi-player -> multi-players

@@
 % Chapter 5 and 6
-The rest of this manuscript thus studies two restricted models, first multi-player bandits in stationary problems, then non-stationary single-player bandits.
+The rest of this manuscript thus studies two restricted models, first multi-players bandits in stationary problems, then non-stationary single-player bandits.
 %  (Chapter~\ref{chapter:5})

### 2019-06-12 2-Chapters/1-Chapter/chapter1.tex
Multi-player -> multi-players

@@
 \subsection{Example: Opportunistic Spectrum Access (OSA) in licensed bands}
+
+\paragraph{Applications for Opportunistic Spectrum Access.}
+% Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
+
+The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc).
+In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
+
+Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
+In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
+In the model with sensing, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
+A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
+%
+This model was first introduced by Wassim Jouini during his PhD thesis,
+in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
+Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.
+
+\TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira..}
+In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
+of using MAB algorithms to optimize channel selection
+on the battery life of a wireless device.
+On the one hand, running a MAB algorithm such as UCB-like algorithms was proved to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
+On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
+Each hardware reconfiguration costs energy for the device, and quickly switching algorithms will lead to a reduction of the battery life.
+The tradeoff between the two aspects is studied empirically in
+\cite{modiDemo2016}.
+%
+Another interesting works is \cite{Modi17QoS}.
+
+\TODOL{FIXME Finir ce blabla rapidement ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
 
@@
     \item
-    We formalize the multi-player bandit model, and we introduced three variants, in Chapter~\ref{chapter:5}.
+    We formalize the multi-players bandit model, and we introduced three variants, in Chapter~\ref{chapter:5}.
     For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
     as well as extensive numerical experiments to demonstrate its good performance in comparison with the rest of the literature.
-    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-player bandits, as some recent research works built up on our results.
+    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works built up on our results.
 
     \item
-    We also give a detailed literature review of the different extensions of the multi-player MAB model, which we believe was never written before.
+    We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
     % were  (ALT 2018, and 4-8 works inspired by our article since then).
-State-of-the-art with our algorithm MCTopM + klUCB for multi-player bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"
+State-of-the-art with our algorithm MCTopM + klUCB for multi-players bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"
 
@@
 
+% FIXME
+\newpage
+
 % ----------------------------------------------------------------------------
-\section{List of publications and other experiences}
+\section{List of publications}
 \label{sec:1:listPublications}
@@
 % =============================================================================
-\subsection{Publications in international conferences with proceedings}
+\paragraph{Publications in international conferences with proceedings}
 
@@
 % =============================================================================
-\subsection{Demonstration in international conferences}
+\textbf{Demonstration in international conferences}
 
@@
 % =============================================================================
-\subsection{French language conference with proceedings}
-% \subsection{Submitted works}
+\textbf{French language conference with proceedings}
+% \textbf{Submitted works}
 
@@
 % =============================================================================
-\subsection{In progress works waiting for a new submission}
+\textbf{In progress works waiting for a new submission}
 
@@
 
-% =============================================================================
-\subsection{Other works}
+% % =============================================================================
+% \textbf{Other works}
 
-\begin{itemize}
-\item
-    \emph{A Note on the Ei Function and a Useful Sum-Inequality},
-    by \textbf{Lilian Besson},
-    February $2018$,
-    \href{https://HAL.Inria.fr/hal-01847480}{\texttt{HAL.Inria.fr/hal-01847480}}.
+% \begin{itemize}
+% \item
+%     \emph{A Note on the Ei Function and a Useful Sum-Inequality},
+%     by \textbf{Lilian Besson},
+%     February $2018$,
+%     \href{https://HAL.Inria.fr/hal-01847480}{\texttt{HAL.Inria.fr/hal-01847480}}.
 
-\end{itemize}
+% \end{itemize}
 
-% =============================================================================
-\subsection{Presentations in seminars and conferences}
+% % =============================================================================
+% \textbf{Presentations in seminars and conferences}
 
-\textbf{Seminars.}
-    I gave some presentations in the following events:
-    SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
-    SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
-    as well as
-    for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
-    for the brown-bag seminar at ENSAI in Bruz in January $2018$,
-    for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
-    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in June $2019$.
+% \textbf{Seminars.}
+%     I gave some presentations in the following events:
+%     SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
+%     SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
+%     as well as
+%     for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
+%     for the brown-bag seminar at ENSAI in Bruz in January $2018$,
+%     for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
+%     for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in June $2019$.
 
-\textbf{Tutorial.}
-    I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
-    with Pierre Haessig (see \href{https://pierreh.eu/}{\texttt{pierreh.eu}}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.
+% \textbf{Tutorial.}
+%     I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
+%     with Pierre Haessig (see \href{https://pierreh.eu/}{\texttt{pierreh.eu}}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.
 
-\textbf{Training.}
-    I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
-    We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
-    and I gave about $12$ training sessions, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.
+% \textbf{Training.}
+%     I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
+%     We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
+%     and I gave about $12$ training sessions, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.
 
-% =============================================================================
-\subsection{Other experiences}
+% % =============================================================================
+% \textbf{Other experiences}
 
-\textbf{Conferences.}
-    I attended the following conferences:
-	\emph{International Conference on Communication} ICC (Paris), May $2017$,
-    \emph{Conference on Learning Theory} COLT (Amsterdam), July $2017$,
-    \emph{Conference on Cognitive Radio Oriented Wireless Networks} CROWNCOM (Lisboa), September $2017$,
-    \emph{Conference on Algorithmic Learning Theory} ALT (Lanzarote), April $2018$,
-    \emph{Wireless Communication and Networking Conference} WCNC (Barcelona), April $2018$,
-	\emph{International Conference on Telecommunication} ICT (Saint-Malo), June $2018$,
-    \emph{Wireless Communication and Networking Conference} WCNC (Marrakech), April $2019$,
-    \emph{Colloque francophone de traitement du signal et des images} GRETSI (Lille), August $2019$.
+% \textbf{Conferences.}
+%     I attended the following conferences:
+% 	\emph{International Conference on Communication} ICC (Paris), May $2017$,
+%     \emph{Conference on Learning Theory} COLT (Amsterdam), July $2017$,
+%     \emph{Conference on Cognitive Radio Oriented Wireless Networks} CROWNCOM (Lisboa), September $2017$,
+%     \emph{Conference on Algorithmic Learning Theory} ALT (Lanzarote), April $2018$,
+%     \emph{Wireless Communication and Networking Conference} WCNC (Barcelona), April $2018$,
+% 	\emph{International Conference on Telecommunication} ICT (Saint-Malo), June $2018$,
+%     \emph{Wireless Communication and Networking Conference} WCNC (Marrakech), April $2019$,
+%     \emph{Colloque francophone de traitement du signal et des images} GRETSI (Lille), August $2019$.
 
-\textbf{Seminars.}
-    I attended the following seminars:
-	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
-	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
-	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$.
+% \textbf{Seminars.}
+%     I attended the following seminars:
+% 	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
+% 	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
+% 	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$.
 
-\textbf{Responsabilities.}
-    I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
-    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
-    I also helped for the organization and presented another poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.
+% \textbf{Responsibilities.}
+%     I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
+%     I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
+%     I also helped for the organization and presented another poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.
 
-\textbf{Reviews.}
-	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
-	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.
+% \textbf{Reviews.}
+% 	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
+% 	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.
 
-\textbf{System Administrator.}
-	maintaining our workstations running GNU/Linux and Windows
-	at SCEE team,
-	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
-	from January $2017$ to September $2018$.
+% \textbf{System Administrator.}
+% 	maintaining our workstations running GNU/Linux and Windows
+% 	at SCEE team,
+% 	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
+% 	from January $2017$ to September $2018$.
 
@@
 
-\TODOL{Open source the repository as soon as I defended my thesis!}
+% FIXME
+% \TODOL{Open source the repository as soon as I defended my thesis!}
 

### 2019-06-12 2-Chapters/2-Chapter/chapter2.tex
Multi-player -> multi-players

@@
 %
-Then, when we study multi-player bandits in Chapter~\ref{chapter:5}, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $Y_{k,t}=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
+Then, when we study multi-players bandits in Chapter~\ref{chapter:5}, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $Y_{k,t}=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
 %
@@
 On the other hand, this decomposition is necessary as theoretical analyses of the regret of (single-player) MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
-Even if we prove that controlling $N_k(T)$ is no longer sufficient to obtain low regret for the multi-player case, we extend this decomposition in Chapter~\ref{chapter:3} (in Lemma~\ref{lem:5:DecompositionRegret}), and it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
+Even if we prove that controlling $N_k(T)$ is no longer sufficient to obtain low regret for the multi-players case, we extend this decomposition in Chapter~\ref{chapter:3} (in Lemma~\ref{lem:5:DecompositionRegret}), and it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
 % \end{itemize}
@@
 
-We consider uniformly efficient\footnote{This notion is then extended to ``\emph{strongly uniformly efficient} algorithms'', in the multi-player case with Definition~\ref{def:5:DecentralizedUniformEfficiency}, where we also include a notion of (expected) fairness.} algorithms, to rule out algorithms achieving low regret on some problem instances while achieving linear regret on other instances.
+We consider uniformly efficient\footnote{This notion is then extended to ``\emph{strongly uniformly efficient} algorithms'', in the multi-players case with Definition~\ref{def:5:DecentralizedUniformEfficiency}, where we also include a notion of (expected) fairness.} algorithms, to rule out algorithms achieving low regret on some problem instances while achieving linear regret on other instances.
 In particular, it is necessary to rule out algorithms that always pick the same arm, as on some problem instances such fixed-arm algorithms can achieve zero regret.
@@
 %
-We find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-player bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
+We find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
 Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
@@
 % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
-On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-player bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
-For more details, see \cite{modiDemo2016} for experiments on the multi-player case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
+On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
+For more details, see \cite{modiDemo2016} for experiments on the multi-players case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
 While such ideas can interest a practitioner, they does not change the asymptotic behavior of $SC^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
@@
 %
-In this thesis, we only consider fairness in the multi-player bandit models,
+In this thesis, we only consider fairness in the multi-players bandit models,
 in Chapter~\ref{chapter:5}, where fairness refers to a different notion (\ie, ``cooperative fairness'').
@@
 			\Else(){
-                For each arm, compute the \textcolor{blue}{index $U_k(t)$}\;
-                Play arm $A(t)$ uniformly among the arms of maximal index: \textcolor{blue}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k(t))$}\;
+                For each arm, compute the \textcolor{blue}{index $U_k^{\cA}(t)$}\;
+                Play arm $A(t)$ uniformly among the arms of maximal index: \textcolor{blue}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$}\;
             }
@@
 		}
-		\caption[{A generic index policy, using indexes $U_k(t)$ (\eg, $\UCB_1$, \klUCB{} etc).}]{A generic \textcolor{blue}{index} policy, using \textcolor{blue}{indexes $U_k(t)$} (\eg, $\UCB_1$, \klUCB{} etc).}
+		\caption[{A generic index policy $\cA$, using indexes $U_k(t)$ (\eg, $\UCB_1$, \klUCB{} etc).}]{A generic \textcolor{blue}{index} policy $\cA$, using \textcolor{blue}{indexes $U_k^{\cA}(t)$} (\eg, $\UCB_1$, \klUCB{} etc).}
 		\label{algo:2:indexPolicy}
@@
     \begin{equation}\label{eq:2:UCBregretBound}
-        R_T^{\cA}(I) \leq \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left(1 + \frac{\pi^2}{3}\right)
+        R_T^{\UCB_1}(I) \leq \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left(1 + \frac{\pi^2}{3}\right)
         = \bigO{ \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T)}.
@@
 For Bernoulli distributions, the previous claim that \klUCB{} builds smaller confidence intervals (\ie, more accurate) than $\UCB_1$, \ie, $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$,
-is justified by Pinkser's inequality gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix).
+is justified by Pinkser's inequality which gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix).
+%
+The \KLUCB{} and \klUCB{} algorithms achieve finite-time regret logarithmic upper-bounds on their regret, and have been proven to be asymptotically optimal, respectively for one-dimensional exponential families and for bounded rewards distributions \cite{Garivier11KL,KLUCBJournal}.
 
@@
 For both directions, we present in the two final chapters natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
-of stationary multi-player and piece-wise stationary bandits.
+of stationary multi-players and piece-wise stationary bandits.
 

### 2019-06-12 2-Chapters/3-Chapter/chapter3.tex
Multi-player -> multi-players

@@
 %
-The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-player) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
+The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-players) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
 After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
@@
 
-\subsection{Single- and multi-player MAB problems}
+\subsection{Single- and multi-players MAB problems}
 
@@
 
-\textbf{Simulation loop for multi-player MAB.}
+\textbf{Simulation loop for multi-players MAB.}
 %
@@
 Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
-We refer to Chapter~\ref{chapter:5} which introduces the multi-player bandit model.
+We refer to Chapter~\ref{chapter:5} which introduces the multi-players bandit model.
 
@@
 
-Any simulation library targeting multi-player bandit problems must implement at least another component:
+Any simulation library targeting multi-players bandit problems must implement at least another component:
 % a collision model, and
@@
 Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
-The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-player model).
+The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
 Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.
@@
     \item
-In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-players bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they obtain logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
@@
 $(i)$
-Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-player multi-armed bandits.
+Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
 SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
@@
 As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
-Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
+Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
 One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
-More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.
+More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
 All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
@@
 %
-An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
+An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
 For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
 For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
-or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
-%
-At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-player bandit model,
+or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+%
+At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-players bandit model,
 and even if some have already been implemented, a major future work is to implement the most interesting ones

### 2019-06-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Multi-player -> multi-players

@@
 % \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
-\indente
+\indent
 On the opposite, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$).

### 2019-06-12 2-Chapters/4-Chapter/chapter4.tex
Commentaire Christophe Ch4.1 et 4.2
Multi-player -> multi-players

@@
 %
-This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
+This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019}.
 
@@
 
-\paragraph{Multi-Player MAB with \emph{collision avoidance}.}
+\paragraph{Multi-Players MAB with \emph{collision avoidance}.}
 
 We are actually able to analyze a simpler model, if we assume to have at most $K$ devices with a transmission probability of $p=1$.
-Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-player MAB} model to describe our problem, as proposed by \cite{Zhao10}.
+Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-players MAB} model to describe our problem, as proposed by \cite{Zhao10}.
 
@@
 and their random activation patter makes the number of active devices a random variable.
-Analyzing multi-player MAB models under this hypothesis is much harder, and is left as a future work.
+Analyzing multi-players MAB models under this hypothesis is much harder, and is left as a future work.
 %

### 2019-06-12 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Multi-player -> multi-players

@@
-% - ``Multi-Player Bandits Revisited'', see https://hal.inria.fr/hal-01629733
-
-% Multi-player Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
-% Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-player MAB algorithms.
+% - ``Multi-Players Bandits Revisited'', see https://hal.inria.fr/hal-01629733
+
+% Multi-players Multi-Armed Bandits (MAB) have been extensively studied in the literature, motivated by applications to Cognitive Radio systems.
+% Driven by such applications as well, we motivate the introduction of several levels of feedback for multi-players MAB algorithms.
 % Existing works assume that \emph{sensing information} is available to the algorithm.
@@
 In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
-Such networks can be modelled using a decentralized multi-player multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
+Such networks can be modelled using a decentralized multi-players multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
 %
@@
 
-We start by reviewing previous works on multi-player MAB models, which all considered the easier case of \emph{sensing feedback}.
+We start by reviewing previous works on multi-players MAB models, which all considered the easier case of \emph{sensing feedback}.
 %
 For this first model, we present two algorithms, \RandTopM{} and \MCTopM, combining an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
-Like in previous works, we consider the centralized system regret (multi-player regret).
+Like in previous works, we consider the centralized system regret (multi-players regret).
 We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
@@
 They both proposed new algorithms, proved to achieve logarithmic regret under the no sensing model.
-We include numerical simulations to compare their proposals, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
+We include numerical simulations to compare their proposals, and we present in details the current state-of-the-art of research on multi-players MAB models without sensing.
 
@@
 This chapter is organized as follows.
-First we motivate the model of multi-player multi-armed bandits in Section~\ref{sec:5:motivation}
-We introduce the multi-player bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
+First we motivate the model of multi-players multi-armed bandits in Section~\ref{sec:5:motivation}
+We introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
 The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms},
@@
 % -----------------------------------------------------------------
-\section{Motivations for multi-player MAB models}
+\section{Motivations for multi-players MAB models}
 \label{sec:5:motivation}
@@
 Due to the communication cost implied by a central controller, a more relevant model is the
-\emph{decentralized multi-player} multi-armed bandit model, introduced by \cite{Zhao10} and further studied shortly after in \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
+\emph{decentralized multi-players} multi-armed bandit model, introduced by \cite{Zhao10} and further studied shortly after in \cite{Anandkumar10,Anandkumar11}, in which players select arms individually and collisions may occur, that yield a loss of reward.
 Further algorithms were proposed in similar models by \cite{Tekin12IEEE} and \cite{Kalathil12} (under the assumption that each arm is a Markov chain),
@@
 We present two algorithms, \RandTopM{} and \MCTopM, based on the combination of an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
-Like in previous works, we consider the centralized system regret (multi-player regret), or simply referred to as regret.
+Like in previous works, we consider the centralized system regret (multi-players regret), or simply referred to as regret.
 We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
@@
 %
-Our article \cite{Besson2018ALT} was the first to study this other model of multi-player bandits for the ``no sensing'' case,
+Our article \cite{Besson2018ALT} was the first to study this other model of multi-players bandits for the ``no sensing'' case,
 % We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
@@
 for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
-Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can have a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-player MAB problems).
+Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can have a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-players MAB problems).
 We do not propose any other efficient algorithm, but our work presented in April~$2018$ has strongly inspired two articles shorlty after \cite{LugosiMehrabian18,BoursierPerchet18}.
 They confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
-We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-player MAB models without sensing.
-
-% In this chapter, we take a step back and present the different feedback levels possible for multi-player MAB algorithms.
+We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-players MAB models without sensing.
+
+% In this chapter, we take a step back and present the different feedback levels possible for multi-players MAB algorithms.
 For each of them, we propose algorithmic solutions supported by both experimental and theoretical guarantees.
@@
 % -----------------------------------------------------------------
-\section{Three feedback levels for the multi-player bandit model}
+\section{Three feedback levels for the multi-players bandit model}
 \label{sec:5:model}
@@
 % explain the interaction protocol and the GOAL
-In the multi-player MAB setting, there are $M \in [K]$ players (or agents),
+In the multi-players MAB setting, there are $M \in [K]$ players (or agents),
 that have to make decisions at some pre-specified time instants.
@@
 
-A multi-player MAB strategy is formally defined as a tuple $\cA = (\cA_1,\dots,\cA_M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
+A multi-players MAB strategy is formally defined as a tuple $\cA = (\cA_1,\dots,\cA_M)$ of arm selection strategies for of each of the $M$ players, and the goal is to propose a strategy that maximizes the total reward of the system, under some constraints.
 First, each player $j$ should adopt a \emph{sequential} strategy $\cA_j$, that decides which arm to select at time $t$ based on \emph{previous observations}.
@@
 
-Following a natural approach in the bandit literature, we evaluate the performance of a multi-player strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
+Following a natural approach in the bandit literature, we evaluate the performance of a multi-players strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
 The regret of the strategy $\cA$ at horizon $T$ is the difference between the cumulated reward of an oracle strategy, assigning in this case the $M$ players to \Mbest,
@@
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-  The excepted centralized multi-player regret is defined by
+  The excepted centralized multi-players regret is defined by
   \begin{equation}\label{eq:5:regret}
@@
 
-With this definition, maximizing the expected sum of global reward of the system is indeed equivalent to minimizing the regret, and we investigate the best possible \emph{regret rate} of a decentralized multi-player algorithm in the next section.
+With this definition, maximizing the expected sum of global reward of the system is indeed equivalent to minimizing the regret, and we investigate the best possible \emph{regret rate} of a decentralized multi-players algorithm in the next section.
 
@@
 % -----------------------------------------------------------------
-\section{New algorithms for multi-player bandits}
+\section{New algorithms for multi-players bandits}
 \label{sec:5:algorithms}
@@
 % %
-The approache we now describe for multi-player bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
+The approache we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
 In particular, we focus on two types of indices:
@@
       \end{tabular}
-      \caption{No matter the orthogonalization policy, using \klUCB{} is much more efficient than using \UCB{}, for multi-player bandit (here in a simple problem with $K=9$ arms).}
+      \caption{No matter the orthogonalization policy, using \klUCB{} is much more efficient than using \UCB{}, for multi-players bandit (here in a simple problem with $K=9$ arms).}
       \label{table:5:comparisonUCB_klUCB}
@@
 % were considered in the analysis of both
-We also compared our algorithms to  %multi-player multi-armed bandits algorithms
+We also compared our algorithms to  %multi-players multi-armed bandits algorithms
 with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}, in the presence of sensing, \ie, observation model \modeldeux, for which they were developed.

### 2019-06-12 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Multi-player -> multi-players

@@
 
-% \begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-player bandit models denoted by $\bm{\mu}$ and $\bm{\lambda}$ respectively, it holds that
+% \begin{lemma}\label{lem:5:CD} Under observation model \modelun{} and \modeldeux, for every event $A$ that is $\cF_{T}^{j}$-measurable, considering two multi-players bandit models denoted by $\bm{\mu}$ and $\bm{\lambda}$ respectively, it holds that
 % \begin{equation}
@@
     % \end{subfigure}
-    \caption[Illustrating the decomposition of the multi-player regret with its three terms]{The three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} of the multi-player regret, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
+    \caption[Illustrating the decomposition of the multi-players regret with its three terms]{The three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} of the multi-players regret, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
     \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}

### 2019-06-12 2-Chapters/5-Chapter/chapter5.tex
Multi-player -> multi-players

@@
 
-% \chapter[Multi-Player Multi-Armed Bandits Models]{Multi-Player Multi-Armed Bandits Models and Algorithms}
-\chapter{Multi-Player Multi-Armed Bandits}
+% \chapter[Multi-Players Multi-Armed Bandits Models]{Multi-Players Multi-Armed Bandits Models and Algorithms}
+\chapter{Multi-Players Multi-Armed Bandits}
 \label{chapter:5}
@@
 as illustrated by numerical simulations.
-We conclude by reviewing different extensions of the multi-player models.
+We conclude by reviewing different extensions of the multi-players models.
 
@@
 %
-We wanted to highlight that the community has been quite active on research on multi-player bandits in the last $10$ years, but especially active since last spring $2018$.
+We wanted to highlight that the community has been quite active on research on multi-players bandits in the last $10$ years, but especially active since last spring $2018$.
 For instance, our article \cite{Besson2018ALT} was the first to propose the ``no sensing'' case, and it was studied by (at least) two independent group of researchers since its publication in April $2018$ \cite{BoursierPerchet18,LugosiMehrabian18}.
@@
 % FIXME écrit pas ça c'est inutile !
-% By following actively the research on multi-player bandit models since the last three years, we believe to be able to give an exhaustive literature review, but it is of course possible that we missed an important research work.
+% By following actively the research on multi-players bandit models since the last three years, we believe to be able to give an exhaustive literature review, but it is of course possible that we missed an important research work.
 In this case, it is our mistake and not an intentional choice as we included all the different extensions of our models that we were aware of.
 
-Many extensions of the simpler model of multi-player bandits have been considered.
+Many extensions of the simpler model of multi-players bandits have been considered.
 The number of players $M$ can be fixed but initially unknown to the players, the sensing information can be absent (like the model \modeltrois{} presented above), there could be communications between players (happening at each time step or only occasionally), or $M$ could evolve over time to model the possibility of arrivals or departures of players (\ie, connexions and disconnections of devices in a wireless networks).
@@
 Finally, we can also be interested by models where $M$ stays fixed, but the environment evolves, for instance with abrupt changes in the means of arms.
-This last model makes a good connexion between this chapter and the next one, and an exciting future work is to further study this model in order to tackle this kind of problems by merging our contributions for stationary multi-player bandits and for non-stationary single-player bandits.
+This last model makes a good connexion between this chapter and the next one, and an exciting future work is to further study this model in order to tackle this kind of problems by merging our contributions for stationary multi-players bandits and for non-stationary single-player bandits.
 
@@
 However, in real-world wireless networks, when players model communicating devices connected to a single gateway, devices can arrive or leave the network at any time.
-The existing previous work on multi-player bandit models are all motivated by possible applications to wireless networks, but most of them assume $M$ to be fixed.
+The existing previous work on multi-players bandit models are all motivated by possible applications to wireless networks, but most of them assume $M$ to be fixed.
 The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
@@
 
-In the multi-player model presented in this chapter, we assume that the arm distributions are the same for all the players.
+In the multi-players model presented in this chapter, we assume that the arm distributions are the same for all the players.
 For cognitive radio applications, where arms model channels and players are radio devices, it can be unrealistic to consider that two players, maybe located at different distances from the gateway or equipped with different hardwares, encounter the same mean quality when accessing the same channel.
 %
-Motivated by this weakness, some researchers studied an interesting extension of the model of interest, considering multi-player MAB models with different arms distributions among players.
+Motivated by this weakness, some researchers studied an interesting extension of the model of interest, considering multi-players MAB models with different arms distributions among players.
 %
@@
 In the coordinated attack, the jammers select non-overlapping channels to attack in each time slot and can significantly increase the number of collisions for SUs.
-They setup the problem in each scenario as a multi-player bandit and develop algorithm,
+They setup the problem in each scenario as a multi-players bandit and develop algorithm,
 and their analysis shows that when the SUs faithfully implement proposed algorithms, the regret is constant with high probability.
@@
 
-To sum up, we presented in this chapter three variants of Multi-Player Multi-Arm Bandits,
+To sum up, we presented in this chapter three variants of Multi-Players Multi-Arm Bandits,
 with different levels of feedback being available to the decentralized players, under which we proposed efficient algorithms.
@@
 This chapter suggests several interesting further research directions.
-First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
+First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
@@
 
-Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-player bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
+Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-players bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
 We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future works left on our library.

### 2019-06-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Multi-player -> multi-players

@@
 
-As mentioned at the end of Chapter~\ref{chapter:5}, another interesting direction of future work is to study non-stationary distributed multi-player bandits.
+As mentioned at the end of Chapter~\ref{chapter:5}, another interesting direction of future work is to study non-stationary distributed multi-players bandits.
 A natural extension of the non-stationary model presented in this Chapter is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
@@
 Their algorithm assume that player $j$ knows its ID $j\in[M]$ as well as $\nu$, and removing these hypotheses is an interesting direction of future work.
-% \TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed} by using \MCTopM{} and \GLRklUCB{} with each other?
+% \TODOL{Evoke here the possibility of study the piece-wise stationary distributed multi-players bandit model introduced in \cite{WeiSrivastava18Distributed} by using \MCTopM{} and \GLRklUCB{} with each other?
 }
@@
 %
-% Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm \cite{Besson2018ALT}, to efficiently tackle the piece-wise stationary distributed multi-player bandit model introduced in \cite{WeiSrivastava18Distributed}.
+% Finally, we hope that our proposal can be combined with our state-of-the-art $\mathrm{MCTopM}$ algorithm \cite{Besson2018ALT}, to efficiently tackle the piece-wise stationary distributed multi-players bandit model introduced in \cite{WeiSrivastava18Distributed}.
 

### 2019-06-12 2-Chapters/Conclusion/conclusion.tex
Multi-player -> multi-players

@@
 in Chapter~\ref{chapter:5}, we study a weaker but still interesting model.
-We present different variants of the multi-player MAB model, and we propose the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
+We present different variants of the multi-players MAB model, and we propose the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
 % Chapter 6
@@
 %
-SMPyBandits allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-player multi-armed bandits.
+SMPyBandits allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
 SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
@@
 Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions.
-Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
+Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
 One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
 % We indeed followed actively the state-of-the-art research, since December $2016$ to June $2019$.
-More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary case.
+More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary case.
 The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
@@
 
-The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Player Multi-Arm Bandits models.
+The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Players Multi-Arm Bandits models.
 %
@@
 
-A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
+A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
 For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
@@
 A second interesting task could be to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
-or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
 %
@@
 
-Our study on multi-player bandits suggests several further research directions.
-First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-player model with sensing information.
+Our study on multi-players bandits suggests several further research directions.
+First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
@@
 
-We also presented many extensions of the multi-player bandit model,
+We also presented many extensions of the multi-players bandit model,
 and even if some have already been implemented, a major future work is to implement the most interesting ones in our library SMPyBandits
@@
 
-Another interesting direction of future work is to study non-stationary distributed multi-player bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.
+Another interesting direction of future work is to study non-stationary distributed multi-players bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.
 A natural extension of the non-stationary model is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.

### 2019-06-12 3-Appendices/2-Appendix/appendix2.tex
Multi-player -> multi-players

@@
 For this later range of applications, it is of highest interest to be able to use low-cost MAB algorithms that do not rely on prior knowledge of the problem for which they will be applied, and especially do not need to know the horizon $T$.
-While we proposed in Chapter~\ref{chapter:5} any-time algorithms solving the presented problem of stationary multi-player bandit, our solution for the problem of non-stationary MAB problem studied in Chapter~\ref{chapter:6} does rely on a prior knowledge of the horizon $T$.
+While we proposed in Chapter~\ref{chapter:5} any-time algorithms solving the presented problem of stationary multi-players bandit, our solution for the problem of non-stationary MAB problem studied in Chapter~\ref{chapter:6} does rely on a prior knowledge of the horizon $T$.
 It is interesting to know what is the best approach to fix this weakness: work more and design algorithms that are inherently any-time, or use a generic technique to avoid depending on a prior knowledge of $T$, and automatically transform our non-anytime approach to make it anytime.

### 2019-06-12 1-Preamble/5-Abstract/abstract.tex
Typo

@@
 % Chapitres 5 et 6
-Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueurs non stationnaires.
+Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueur non stationnaires.
 %  (Chapitre~\ref{chapitre:5})

### 2019-06-12 2-Chapters/1-Chapter/chapter1.tex
Typo

@@
 on the battery life of a wireless device.
-On the one hand, running a MAB algorithm such as UCB-like algorithms was proved to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
+On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
 On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
@@
 \item
-    \emph{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation},\\
+    \emph{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation},
     by Christophe Moy \& \textbf{Lilian Besson}.
     1st International ISIoT workshop\footnote{~See \href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}},
-    at \emph{Conference on Distributed Computing in Sensor Systems}\footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}}, Santorini, Greece, May 2019.\\
+    at \emph{Conference on Distributed Computing in Sensor Systems}\footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}}, Santorini, Greece, May 2019.
     \href{https://HAL.Inria.fr/hal-02144465}{\texttt{HAL.Inria.fr/hal-02144465}}.
@@
 \item
-    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
+    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},
     by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
     1st International MOTIoN workshop\footnote{~MOTIoN 2019, see \href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
-    at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.\\
+    at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.
     \href{https://HAL.Inria.fr/hal-02049824}{\texttt{HAL.Inria.fr/hal-02049824}}.
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,\\
+    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,
     \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
@@
 \item
-    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},\\
+    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},
     by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,\\
+    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,
     \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
-    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{~CROWNCOM 2017, see \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,\\
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{~CROWNCOM 2017, see \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,
     \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
@@
 % =============================================================================
-\textbf{Demonstration in international conferences}
+\paragraph{Demonstration in international conferences}
 
@@
 % =============================================================================
-\textbf{French language conference with proceedings}
+\paragraph{French language conference with proceedings}
 % \textbf{Submitted works}
@@
 % =============================================================================
-\textbf{In progress works waiting for a new submission}
+\paragraph{In progress works waiting for a new submission}
 
@@
     \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann, \\
+    by \textbf{Lilian Besson} \& Emilie Kaufmann,
     February $2019$,
@@
 
-\textbf{Copyright notice.}
+\paragraph{Copyright notice.}
 %

### 2019-06-12 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Typo

@@
 %
-Regret bounds in several different settings are proved for the \CORRAL{} algorithm \cite{Agarwal16}, but no logarithmic upper-bound can be obtained from their technique, even in the simplest setting of stochastic bandits.
+Regret bounds in several different settings are proven for the \CORRAL{} algorithm \cite{Agarwal16}, but no logarithmic upper-bound can be obtained from their technique, even in the simplest setting of stochastic bandits.
 %
@@
 
-\ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proved.
+\ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
 Empirically, \Aggr{} showed to always have a logarithmic

### 2019-06-12 2-Chapters/2-Chapter/chapter2.tex
Typo

@@
 and we illustrate its performance empirically.
-% , but no satisfactory theoretical result is given, as the bound we had hoped for has been proved unachievable since then.
+% , but no satisfactory theoretical result is given, as the bound we had hoped for has been proven unachievable since then.
 
@@
 
-Many algorithms have been proved to achieve logarithmic regret in the stochastic case,
+Many algorithms have been proven to achieve logarithmic regret in the stochastic case,
 and in particular it is the case of the algorithms used in this thesis, $\UCB_1$ from \cite{Auer02}, Thompson sampling from \cite{Thompson33} and analyzed in \cite{AgrawalGoyal11,Kaufmann12Thompson}, and \klUCB{} from \cite{Garivier11KL,KLUCBJournal}.
@@
 Indeed, in single-player models it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
-But even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.\\
-%
+But even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
+
+\TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
 We find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
@@
 %
+\TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
 In this thesis, we only consider fairness in the multi-players bandit models,
@@
 \begin{align}\label{eq:UCB_index}
-    U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation} \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration} \xi_k(t)}.
+    U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation }\; \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration }\; \xi_k(t)}.
 \end{align}
 
-This selection rule \eqref{eq:UCB_index} makes sense for the following reason:
-an arm $k$ is chosen at step $t$ because it has a large $\UCB_k(t)$, which can happen for two reasons:
-$i)$ because the average reward $\widehat{\mu_k}(t)$ is large, in which case this arm is likely to have a high mean reward $\mu_k$,
+This selection rule $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$ makes sense for the following reasons:
+an arm $k$ is chosen at step $t$ because it has a large $\UCB_k(t)$, which can happen for two reasons.
+$i)$ Because the average reward $\widehat{\mu_k}(t)$ is large, in which case this arm is likely to have a high mean reward $\mu_k$,
 $ii)$ and/or because the confidence radius $\xi_k(t)$ is large, in which case this arm has not been explored much.
 Either reason makes this arm worth choosing.
-One can also observe that $\xi_k(t)\to\infty$ if $t\to\infty$ while $N_k(t)\ll\log(t)$, which can be used to prove that the $\UCB_1$ algorithm tries all arms at least $\mathcal{O}(\log(T))$ times (in average).
-%
-Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{}  respectively represent \emph{exploitation} and \emph{exploration}, and summing them up is a natural way of dealing with the exploitation/exploration trade off.
-The constant parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to choose $\alpha=1/2$ for optimal uniform performance (\ie, on all problems).
+One can also observe that $\xi_k(t)\to\infty$ if $t\to\infty$ while $N_k(t)\ll\log(t)$, which can be used to prove that the $\UCB_1$ algorithm tries all arms at least $\Omega(\log(T))$ times (in average).
+%
+Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{} defined in \eqref{eq:UCB_index} respectively represent \emph{exploitation} and \emph{exploration}, and summing them up is a natural way of dealing with the exploitation/exploration trade off.
+The constant parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to choose $\alpha=1/2$ for optimal uniform performance (\ie, on all problems, \cite{Auer02}).
 
@@
 %
-This theorem shows that $\UCB_1$ achieves a order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta \eqdef \min\limits_{k: \Delta_k>0} \Delta$).
+This theorem shows that $\UCB_1$ achieves an order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta \eqdef \min\limits_{k: \Delta_k>0} \Delta$).
 
@@
 \begin{smallproof}\label{proof:2:UCBregretBound}
-    As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
+    As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives the inequality \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
     Thanks to the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
@@
 Instead of sampling from the posteriors and playing the arms with maximum sample, Bayes-UCB computes the $(1-1/(t (\log(t))^5)$ quantile of each arm, at time $t$, and plays the arm with the largest quantile.
-It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
+It was also proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
 %
@@
 Based on this idea, the BESA algorithm for $K=2$ arms is quite simple: at each time step, if arms $1$ and $2$ have respectively $n_1$ and $n_2$ samples, and if $n_1>n_2$ (for instance), first it sub-samples $n_2$ observations from $n_1$, then it computes the mean of these observations, and play the arm maximizing this mean (now that there is the same number of observations for both arms, it makes more sense to compare these quantities).
-It was analyzed partially, for the two-armed case it was proved to be asymptotically optimal for one-dimensional exponential families, and the authors illustrated that it can be very efficient empirically.
+It was analyzed partially, for the two-armed case it was proven to be asymptotically optimal for one-dimensional exponential families, and the authors illustrated that it can be very efficient empirically.
 Analyzing it in the generic case was found to be difficult, and it is still an open problem.
@@
 %
-It was proved in \cite{Auer02NonStochastic} that Exp3 with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
+It was proven in \cite{Auer02NonStochastic} that Exp3 with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
 while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
@@
 
-More recently, the \klUCB-switch from \cite{Garivier18} is the first algorithm to be proved to
+More recently, the \klUCB-switch from \cite{Garivier18} is the first algorithm to be proven to
 enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, that is, matching the $C_I \ln(T)$ lower bound by \cite{LaiRobbins85}.
@@
 \label{sub:2:shortNumericalExperiments}
+
+\TODOL{Si je dois gagner de la place, je peux ne pas inclure cette section...}
 

### 2019-06-12 2-Chapters/3-Chapter/chapter3.tex
Typo

@@
 		Example of a single-player simulation showing the average regret of four algorithms.
-They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proved to match the lower-bound up-to a constant), and each but UCB is known to be asymptotically optimal (\ie, with the constant matching the lower-bound).
+They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proven to match the lower-bound up-to a constant), and each but UCB is known to be asymptotically optimal (\ie, with the constant matching the lower-bound).
 	}
@@
     Thompson sampling from \cite{Kaufmann12Thompson} (\texttt{Thompson}), using a Beta prior and posterior, initially uniform (\ie, $\pi_k(0)=\mathrm{Beta}(1,1)$).
-    % It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+    % It was also proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
     It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on highly optimized C or Cython code).
@@
     Bayes-UCB from \cite{Kaufmann12BUCB} (\texttt{BayesUCB}), using a Beta prior and posterior, initially uniform, \ie, $\pi_k(0)=\mathrm{Beta}(1,1)$.
-    % It was also proved to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
+    % It was also proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
     It is comparable to Thompson sampling in terms of memory complexity, but slower as computing a quantile is more costly than sampling from a distribution.
@@
     It is based on the online mirror descent algorithm with a Tsallis entropy regularizer, and it is anytime.
-    It was also proved to obtain good ``best of both world'' performances.
+    It was also proven to obtain good ``best of both world'' performances.
     Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.

### 2019-06-12 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Typo

@@
 Note that this bound is in $\bigO{M^3}$,
-which significantly improves the $\bigO{M{2M-1 \choose M}}$ proved by \cite{Anandkumar11} for \rhoRand.
-It is worse than the $\bigO{M^2}$ proved by \cite{Rosenski16} for \MusicalChair{}.
+which significantly improves the $\bigO{M{2M-1 \choose M}}$ proven by \cite{Anandkumar11} for \rhoRand.
+It is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}.
 %, due to our trick of focusing on collisions for non-sitted players.

### 2019-06-12 2-Chapters/5-Chapter/chapter5.tex
Typo

@@
 We can start by asking whether the most efficient algorithm, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different than the real number of players $M$.
-Even though we did not include numerical simulations for this case in Section~\ref{sec:5:experiments} above, we did some tests, which confirmed two disappointing results that were also proved analytically.
+Even though we did not include numerical simulations for this case in Section~\ref{sec:5:experiments} above, we did some tests, which confirmed two disappointing results that were also proven analytically.
 First, the three algorithms (\RhoRand, \RandTopM{} and \MCTopM) give linear regret if they are using a value $M'$ strictly smaller than $M$ (and if $\mu^*_{M-1} > \mu^*_M$), as players will converge to play about $T-\bigO{\log(T)}$ times the $M-1$ best arms, leading to at least one collision most of the time, and thus a linear regret.
@@
 and consistent on the different algorithms.
-The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proved to be $\cO(1)$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
+The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proven to be $\cO(1)$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
 In this extreme case, \RhoRand{} and its extension both perform very closely, as expected, and much worse than \MCTopM.
@@
 This behavior is well known for classical bandits, where bounds of the form $\log(T)/\Delta^2$ become worse than a linear regret $T$ if $\Delta$ is very small (\ie, $\Delta\ll \sqrt{\log(T)/T}$).
-For this extension, their proposed algorithm is proved to achieve $\bigO{K^2 M (\log(T))^2 / \mu + K M \min(\sqrt{T \log(T)}, \log(T)/\Delta')}$ regret, if $\mu$ is a lower-bound on $\mu^*_{M}$ and $\Delta' = \max(\Delta, \min\{|\mu^*_M-\mu^*_i| : \mu^*_M > \mu^*_i \})$ (that both have to be known beforehand by the algorithm).
+For this extension, their proposed algorithm is proven to achieve $\bigO{K^2 M (\log(T))^2 / \mu + K M \min(\sqrt{T \log(T)}, \log(T)/\Delta')}$ regret, if $\mu$ is a lower-bound on $\mu^*_{M}$ and $\Delta' = \max(\Delta, \min\{|\mu^*_M-\mu^*_i| : \mu^*_M > \mu^*_i \})$ (that both have to be known beforehand by the algorithm).
 %
@@
 %
-It is also proved in Appendix~A of \cite{BoursierPerchet18} that \Selfish-\UCB{} has a linear regret, in the theoretical case, with a very neat argument from number theory (using Lindemann-Weierstrass theorem).
+It is also proven in Appendix~A of \cite{BoursierPerchet18} that \Selfish-\UCB{} has a linear regret, in the theoretical case, with a very neat argument from number theory (using Lindemann-Weierstrass theorem).
 As we conjectured, there is a gap between the theoretical result and its practical consequence, as the Theorem~4 they gave is only valid for real-valued number, and not for hardware-represented floating point number.
@@
 % combining Musical Chair with an epoch-based alternative phases of forced exploration and pure exploitation.
-They analyze their GoT algorithm and proved that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
+They analyze their GoT algorithm and proven that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
 
@@
 %
-We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proved to be very active in the last two years.
+We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proven to be very active in the last two years.
 For some models, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.

### 2019-06-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Typo

@@
 % Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
-% More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
+% More recently, MAB models have been proven useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
 % Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
@@
 It works by decreasing all the past rewards by a multiplicative factor, when receiving a new reward from an arm, so that the recent rewards weight more in the (discounted) empirical mean used for the computation of the \UCB{} indexes.
-The regret of D-UCB was proved to be upper-bounded by $\cO(\sqrt{\Upsilon_T T }\log(T))$ in \cite{Garivier11UCBDiscount}, with a tuning $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, dependent on $\Upsilon_T$.
+The regret of D-UCB was proven to be upper-bounded by $\cO(\sqrt{\Upsilon_T T }\log(T))$ in \cite{Garivier11UCBDiscount}, with a tuning $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, dependent on $\Upsilon_T$.
 %
@@
 % , so we do not include it in .
-Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proved to be less efficient empirically, or are designed for more specific settings.
+Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proven to be less efficient empirically, or are designed for more specific settings.
 
@@
 The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
-A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proved, even when $\Upsilon_T$ is known.
+A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.
 
@@
 
-When all their parameters are tuned correctly, both approaches are proved to achieve a good regret upper-bound, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
+When all their parameters are tuned correctly, both approaches are proven to achieve a good regret upper-bound, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
 As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
@@
 Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
-A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proved, which is weaker than existing results in our setting.
+A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proven, which is weaker than existing results in our setting.
 
@@
 \begin{smallproof}
-    This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
+    This Lemma~\ref{lem:6:ConcFirst} is proven in Appendix~\ref{app:6:proofConcFirst}.
 \end{smallproof}
@@
 \begin{smallproof}
-    It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proved in the proofs.
+    It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proven in the proofs.
     For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} in page~\pageref{par:6:controllingDelayInOneProof}.
@@
 
-For a threshold $\beta$ chosen as in the Lemma~\ref{lem:6:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lem:6:Delay} is that if the break-point $\tau$ takes place after $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude (with high probability).
+For a threshold $\beta$ chosen as in the Lemma~\ref{lem:6:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lem:6:Delay} is that if the break-point $\tau$ takes place after about $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude (with high probability).
 %
@@
             \uIf{
-                $\alpha>0$ and $t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor \in [K]$
+                $\alpha>0$ and $t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor \in [K]$\;
             }{
-                $A(t) = t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor$
+                $A(t) = t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor$\;
                 \tcp*[f]{forced exploration} \\
@@
             \Else{
-                $A(t) = \arg\max\limits_{k \in [K]} \mbox{UCB}_k(t)$ as defined in \eqref{def:6:UCB}
+                $A(t) \in \cU\left( \arg\max\limits_{k \in [K]} \mbox{UCB}_k(t) \right)$ as defined in \eqref{def:6:UCB}\;
             }
-            \mbox{Play arm } $A(t)$ \mbox{and receive the reward } $Y_{A(t),t}$ : $n_{A(t)} = n_{A(t)} + 1; Z_{A(t), n_{A(t)}} = Y_{A(t),t}$ \\
+            \mbox{Play arm } $A(t)$ \mbox{and receive the reward } $Y_{A(t),t}$ : $n_{A(t)} = n_{A(t)} + 1; Z_{A(t), n_{A(t)}} = Y_{A(t),t}$\;
             \uIf(\tcp*[f]{Change point is detected}){
@@
             {
-                \uIf{Global restart}{
-                    $\forall k\in [K], \tau_k = t$ and $n_k = 0 $.
-                    \tcp*[f]{global restart}
+                \uIf{\textbf{Global} restart}{
+                    $\forall k\in [K], \tau_k = t$ and $n_k = 0$\;
+                    \tcp*[f]{restart all arms}
                     }
                 \Else{
-                    $\tau_{A(t)} = t$ and $n_{A(t)} = 0$
-                    \tcp*[f]{local restart}
+                    $\tau_{A(t)} = t$ and $n_{A(t)} = 0$\;
+                    \tcp*[f]{restart only this arm}
                 }

### 2019-06-12 2-Chapters/Conclusion/conclusion.tex
Typo

@@
 
-\TODOL{Je dois faire quelques phrases résumant la thèse}
+\TODOL{Je dois faire quelques phrases résumant la thèse, mais plus court}
 
@@
 Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
-The notations used in the manuscript were introduced by showcasing an interactive demonstration published online\footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
+The notations used in the manuscript were introduced by showcasing an interactive demonstration\footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
 
@@
 %
-We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proved to be very active in the last two years.
+We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proven to be very active in the last two years.
 For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.

### 2019-06-12 5-Nomenclature/Nomenclature.tex
Typo

@@
 % Institution
-\nomenclature[A]{ANR}{\textit{Agence Nationale de la Recherche} (French National Research Agency)}
-\nomenclature[A]{ENS}{\textit{École Normale Supérieure}}
-\nomenclature[A]{CNRS}{\textit{Centre National de la Recherche Scientifique}}
+% \nomenclature[A]{ANR}{\textit{Agence Nationale de la Recherche} (French National Research Agency)}
+% \nomenclature[A]{ENS}{\textit{École Normale Supérieure}}
+% \nomenclature[A]{CNRS}{\textit{Centre National de la Recherche Scientifique}}
 \nomenclature[A]{SCEE}{\textit{Signal, Communication et Électronique Embarquée} (research team in CentraleSupélec, campus of Rennes)}
@@
 \nomenclature[G]{$\nu_k$, $\nu_k(t)$, $\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
-\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$ (in Chapter~\ref{chapter:6}), arm $k$ for user $j$ (in Chapter~\ref{chapter:5}), arm $i$ in the $k$-th stationary interval (in Chapter~\ref{chapter:6})}
+\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_k^{(i)}$}{Mean of arm $k$, arm $k$ at time $t$ (in Chapter~\ref{chapter:6}), arm $k$ for user $j$ (in Chapter~\ref{chapter:5}), arm $k$ in the $i$-th stationary interval (in Chapter~\ref{chapter:6})}
 \nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
@@
 $\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
-\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\tau_k^j$, $\tau_k^{(i)}$}{Location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
 \nomenclature[G]{$\tau$}{Length of a sliding-window, \eg, in SW-\UCB{} in Chapter~\ref{chapter:6}}
@@
 \nomenclature[L]{$\cH_0$, $\cH_1$, $\cH_2$}{Hypothesis, in Chapters~\ref{chapter:4} and \ref{chapter:6}}
-\nomenclature[L]{$d^{(k)}$, $d^{(\ell)}$}{Lower-bound on number of samples for accurate detection of change-points in Chapter~\ref{chapter:6}}
+\nomenclature[L]{$d^{(i)}$, $d^{(\ell)}$}{Lower-bound on number of samples for accurate detection of change-points in Chapter~\ref{chapter:6}}
 \nomenclature[L]{$\cT$}{In Chapter~\ref{chapter:6}, denote a function in \eqref{def:6:function_T} and a set of time steps in Section~\ref{sub:6:IdeasOptimizations}}
@@
 \nomenclature[E]{$y^j$}{Usually denotes a variable depending on a player in Chapter~\ref{chapter:5} (for $j\in[M]$), or to distinguish between different independent runs in numerical simulations in Chapter~\ref{chapter:2} (for $j\in[N]$)}
-\nomenclature[E]{$y^{(k)}$, $y^{(\ell)}$}{Usually the superscript index $(k)$ or $(\ell)$ denotes a variable in the $k$-th stationary interval, under the  point-of-views of \emph{global} or \emph{local} changes, in Chapter~\ref{chapter:6}}
+\nomenclature[E]{$y^{(i)}$, $y^{(\ell)}$}{Usually the superscript index $(i)$ or $(\ell)$ denotes a variable in the $i$-th stationary interval, under the  point-of-views of \emph{global} or \emph{local} changes, in Chapter~\ref{chapter:6}}

### 2019-06-12 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Commentaire Christophe Ch4.1 et 4.2
Commentaires de Xtof pour le Ch4

@@
 
-In this section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept (PoC) of the model introduced in Section~\ref{sec:4:firstModel}.
+In this section, we present a demonstration showcased at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept (PoC) of the model introduced in Section~\ref{sec:4:firstModel}.
+%
+As far as we know, this is the first demonstration of running learning algorithms on the IoT device side, in real radio conditions, as highlighted it in our paper \cite{MoyBesson2019}.
 
@@
 
-In our demo, we consider a simple wireless network, consisting of one gateway (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
+In our demo, we consider a simple wireless network, that reproduces the model of Section~\ref{sec:4:firstModel}, consisting of one gateway (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
 Some dynamic intelligent devices (end-user or autonomous devices) try to communicate with the gateway, with a low-overhead protocol.
@@
 
-We can generate scenarios with different parameters (number of channels, interfering traffic, etc) in order to evaluate the performance of learning in various settings.
+We can generate scenarios with different parameters (number of channels, interfering traffic load on each channel, etc) in order to evaluate the performance of learning in various settings.
 Moreover, we compare the performance of learning strategies with that of the random uniform access to channels, which is the current state-of-the-art of commercial LPWAN solutions \cite{Raza17}.
@@
 We could have used any physical layer and any ALOHA-based protocol.
-We choose to implement our own physical layer and protocol, for both clarity and conciseness, and because developing a complete IoT network protocol stack is no more a research work and would have fall outside of the scope of this thesis.
+We choose to implement our own physical layer and protocol, for both clarity and conciseness, and because developing a complete IoT network protocol stack is no more my research work and would have fall outside of the scope of this thesis.
 
@@
 Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
-For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a longer sequence if the PoC have to consider more users.
-With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway
+For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a sequence of QPSK symols if the PoC has to consider more users.
+With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway.
 Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
@@
 with version $4$ of their FPGA system and version $5.1$ of the RBX system.
-As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP.
-The gateway, a USRP which emulates the interfering traffic, and at least one dynamic device, each of them having its own USRP and its own learning algorithm.
+As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP:
+the gateway,
+a traffic generator which emulates the interfering traffic (made by surrounding static devices),
+and at least one dynamic device.
+Each dynamic device has its own USRP and its own learning algorithm.
 % Cables and laptop
@@
 The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
-The pictures in Figure~\ref{fig:42:photosSCEETestBed} shows the testbed used for these experiments.
+The pictures in Figure~\ref{fig:42:photosSCEETestBed} show the testbed used for these experiments.
 % Octoclock
@@
 The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.
-Messages of the random traffic are shorter in time by purpose (they could be coming from othe rIoT standard), in order to distinguish them from an intelligent device traffic on the ``waterfall'' visualizations of the traffic.
+Messages of the random traffic, generated by surrounding static devices, are shorter in time by purpose (they could be coming from other IoT standard), in order to distinguish them from an intelligent device traffic on the ``waterfall'' visualizations of the traffic.
 
@@
 At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \emph{(b)}.
-We can see, in this view that the device transmitted its last $9$ packets in the two best channels (channel $\#3$ and $\#4$).
+We can see, in this view that the device transmitted its last $9$ packets in channels $\#3$ and $\#4$.
 Then, at the top right of this interface \emph{(c)}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the gateway (\textcolor{darkblue}{blue}).
 Colors in the ``waterfall'' represent the RF power level received at the device antenna.
-Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (emission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \emph{(c)} in part (2)).
+Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (transmission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \emph{(c)} in part (2)).
 Finally, at the bottom right \emph{(d)}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).
@@
 The last part is the interface of the gateway, where we can see the traffic observed by the gateway \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
+The observed traffic is coherent with \emph{(c)} of part (2), as the elements are very close the one from the others on the testbed.
+Colors may change, as they depend on the exact distance between the different transmitters and receivers.
 \end{enumerate}
@@
 We compare in this PoC the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
+Note that we could run more algorithms, but with no real added-valued in terms of validation of the proposed learning-based approach, which is our goal.
+%
 For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (``\textcolor{green}{TS}'', in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
@@
 %
-Each experiment takes about half a day,
-as we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
+Each experiment has been done on a duration of about half a day,
+due to the IoT sporadic transmission mode that we want to respect (like in our model of Section~\ref{sec:4:firstModel}).
+However, we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
 Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
-Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_i)_i = [15\%, 10\%, 2\%, 1\%]$.
+Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_k)_k = [15\%, 10\%, 2\%, 1\%]$.
+Note that a maximum occupancy rate of $15\%$ could seem not so high, but indeed it is, because for a pure ALOHA access mode, a naive dynamic device only enjoys about $40\%$ success rate, under such occupancy of the channels (see the ``uniform'' plot in Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}).
+%
 The occupancy rate of a channel, which denotes the mean occupancy, is implemented using the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).

### 2019-06-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Commentaires de Xtof pour le Ch4

@@
 other recent directions of research first include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
-then realistic empirical simulations like in \cite{kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
+then realistic empirical PoC like in \cite{MoyWSR2014,RobertSDR2014,modiDemo2016,darak2016bayesian,kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
 and finally applications to multi-hoping networks \cite{Mitton16,Toldov16} or other kinds of networks \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
@@
 
-The aim of this section is to assess the performance of MAB algorithms for channel selection in LPWA networks, while taking into account the impact of retransmissions on the network performance.
+The aim of this section is to assess the performance of MAB algorithms for channel selection in LPWA networks operating in unlicensed bands, while taking into account the impact of retransmissions on the network performance.
 For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
@@
 in case of RF (uplink or downlink) collision that prevent the device to receive the \emph{Ack} from the gateway,
-these devices will attempt to retransmit their packet up-to $M$ times, with $M \in\mathbb{N}$.
-It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\{0, \dots, m-1\}$, where $m \in\mathbb{N}, m>0$ is the length of the back-off interval
-(note the difference between $m$ and $M$).
+these devices will attempt to retransmit their packet up-to $\mathrm{MaxBackOff}$ times\footnote{~We prefer to denote it $\mathrm{MaxBackOff}$ instead of $M$ like we did in our paper \cite{Bonnefoi2019WCNC}, as $M$ is used in the next Chapter~\ref{chapter:5}.},
+with $\mathrm{MaxBackOff} \in\mathbb{N}^*$.
+It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\{0, \dots, m-1\}$, where $m \in\mathbb{N}^*$ is the length of the back-off interval
+(note the difference between $m$ and $\mathrm{MaxBackOff}$).
 
@@
 The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Figure~\ref{fig:43:Markov_model}.
-When a device has a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{M-2} \}$, at each $M$ back-off stage.
+When a device has a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{\mathrm{MaxBackOff}-2} \}$, at each $\mathrm{MaxBackOff}$ back-off stage.
 At each time slot, a transition from an idle state to a transmission state (denoted as \texttt{Trans.}) occurs if a packet transmission is required, while waiting states (denoted as \texttt{Wait}), correspond to a $m$ back-off interval.
@@
 Contrarily to what may appear at first sight,
-the goal is not to minimize the number of retransmissions, but to maximize the probability of successful transmission, considering both the first transmission of a message and the retransmissions of the same message, because the objective of each device is to maximize its battery life by minimizing its number of transmissions.
+\emph{the goal is not to minimize the number of retransmissions}, but to maximize the probability of successful transmission, considering both the first transmission of a message and the retransmissions of the same message.
+Indeed, the objective of each device is to \emph{maximize its battery life by minimizing its \textbf{total} number of transmissions}.
 % , where each channel (also called arms) is viewed as a gambling machine (bandit), and each bandit has a \emph{reward}.
@@
 When such an IoT device experiments a collision, it goes in a back-off state to retransmit the same packet on the same or another channel.
-If all devices remain in the same channel for retransmissions, it could result in a sequence of successive collisions with the same packets' devices that previously collided.
+If all devices remain in the same channel for retransmissions, it is a well-known result that it could result in a sequence of successive collisions with the same packets' devices that previously collided.
 %
@@
 	\item $\mathcal{H}_{2}$:
-	The total number of back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions
+	The total number of back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after {\mathrm{MaxBackOf}}$ successive failed retransmissions
 	(otherwise, its battery life can be threatened if it does not have re-fueling capabilities).
@@
 Considering one device and one channel,
-we denote $x_t^i$ the probability that it is transmitting a packet for the $(i+1)$-th time in a given time slot $t$ (with $i\in \{0, \dots, M-1 \}$),
-and we denote $x_t = \sum_{i=0}^{M-1}x_t^i$ the probability that it transmits a packet (\ie, just the sum on $i$ of $x_t^i$).
+we denote $x_t^i$ the probability that it is transmitting a packet for the $(i+1)$-th time in a given time slot $t$ (with $i\in \{0, \dots, \mathrm{MaxBackOff}-1 \}$),
+and we denote $x_t = \sum_{i=0}^{\mathrm{MaxBackOff}-1}x_t^i$ the probability that it transmits a packet (\ie, just the sum on $i$ of $x_t^i$).
 We consider $N > 0$ active devices following the same policy.
@@
 %{\color{red}Considering one device and a channel,
-%we denote $x_t^i$ the packet transmission probability for the $i+1$ time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$).}
+%we denote $x_t^i$ the packet transmission probability for the $i+1$ time in a given time slot $t$ (with $i\in \llbracket 0, \mathrm{MaxBackOff}-1 \rrbracket$).}
 %
-%{\color{red}Considering one device and a channel, we denote $x_t^i$ the transmission probability for the $i+1$ time, with $i\in \llbracket 0, M-1 \rrbracket$, in a given time slot $t$}
+%{\color{red}Considering one device and a channel, we denote $x_t^i$ the transmission probability for the $i+1$ time, with $i\in \llbracket 0, \mathrm{MaxBackOff}-1 \rrbracket$, in a given time slot $t$}
 
@@
 In order to assess the proposed approximation, we suppose a unique channel where all the devices follow the same contention Markov process.
-We simulate an ALOHA protocol with a maximum number of retransmissions $M=10$, a maximum back-off interval $m=10$, and a transmission probability $p=10^{-3}$.
+We simulate an ALOHA protocol with a maximum number of retransmissions {\mathrm{MaxBackOf}}=10$, a maximum back-off interval $m=10$, and a transmission probability $p=10^{-3}$.
 %
@@
 	In all the algorithms and notations presented in this Section, the time steps $t$ do not refer to the \emph{global} time steps (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
-	Each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $M$ times (\eg, $M=10$) until it receives an \Ack.
+	Each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $\mathrm{MaxBackOff}$ times (\eg, $\mathrm{MaxBackOff}=10$) until it receives an \Ack.
 	%
@@
 			\uIf{First packet transmission}{
-				Use first-stage \UCB{} as in Algorithm~\ref{algo:43:UCB}.
+				Use first-stage \UCB{} as in Algorithm~\ref{algo:43:UCB}\;
 			}
@@
 %
-Similarly, Philippe Mary asked me at the MoTION workshop where I presented \cite{Bonnefoi2019WCNC}, why didn't we rather used Bayes-UCB instead of \UCB{} (from \cite{Kaufmann12BUCB}), as it is known to empirically outperform \UCB{} (as illustrated for instance in \cite{darak2016bayesian,kumar2016two}).
+% Similarly, Philippe Mary asked me at the MoTION workshop where I presented \cite{Bonnefoi2019WCNC}, why didn't we rather used Bayes-UCB instead of \UCB{} (from \cite{Kaufmann12BUCB}), as it is known to empirically outperform \UCB{} (as illustrated for instance in \cite{darak2016bayesian,kumar2016two}).
 We could have chose any bandit algorithm for the ``first stage'' component used in this Section, and for simplicity and clarity we focussed on a simple but efficient one (\UCB).
@@
 In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform repartition of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
-In other words, the channels are occupied respectively $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $M = 5$, and $m=5$.
+In other words, the channels are occupied\footnote{~Note that we consider higher occupancy rates that the ones considered previously in Section~\ref{sec:4:gnuradio}, because the goal of this experimental section is to evaluate our approach that use learning to optimize the way each device retransmits its packets.
+We need to have a large enough probability $p_c$ of retransmission if we want to see a difference between the different learning heuristics, which is why we consider a network more densily occupied than in Section~\ref{sec:4:gnuradio}.}
+respectively $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $\mathrm{MaxBackOff} = 5$, and $m=5$.
 In Figure~\ref{fig:43:mainExperiment1}, we show the successful transmission rate versus the number of slots, for all the proposed heuristics.
@@
 A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve in black), which is (still) the current state-of-the-art in IoT networks.
-The improvement of the heuristics over the non-learning approach is clear, and for every heuristic that uses a kind of learning mechanism it can observed the successful transmission rate increases rapidly (or equivalently the PLR decreases).
+The improvement of the heuristics over the non-learning approach is clear, and for every heuristic that uses a kind of learning mechanism it can be observed that the successful transmission rate increases rapidly (or equivalently the PLR decreases).
 Moreover, all of these approaches show a fast convergence making them suitable for the targeted application.
@@
 		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
+		We considered $N=1000$ static IoT devices, that occupy the $4$ channels in a static way leading to mean occupancy rates of $10\%,30\%,30\%,30\%$.
 	}
@@
 
-We also consider in our experiment the case of an ALOHA protocol using $M=5$, and $m=10$, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
+We also consider in our experiment the case of an ALOHA protocol using $\mathrm{MaxBackOff}=5$, and $m=10$, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
 The corresponding results are depicted in Figure~\ref{fig:43:mainExperiment2}.
@@
 		We observe that the \textcolor{cyan}{random retransmission} achieves poor performance compared to the others.
+		We considered $N=2000$ static IoT devices, that occupy the $4$ channels in a static way leading to mean occupancy rates of $10\%,30\%,30\%,30\%$.
 	}

### 2019-06-12 2-Chapters/Conclusion/conclusion.tex
Commentaires de Xtof pour le Ch4

@@
 Surprisingly, the main take-away message is that a simple \UCB{} learning approach (that simply retransmits in the same channel), turns out to perform as well as more complicated heuristics.
+
+Thus in Chapter~\ref{chapter:4}, we proposed two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
+that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
+%
+Decentralized RL for IoT lets the devices use acknowledgements sent by their Base Station as a reward, instead of sensing feedback like for OSA.
+We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
+The first model (without retransmission of packets),
+is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
+%
+However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $M=10$).
+%
+The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
 

### 2019-06-12 5-Nomenclature/Nomenclature.tex
Commentaires de Xtof pour le Ch4

@@
 \nomenclature[L]{$k$}{Usually denotes the $k$-th arm, $k\in[K]$, mainly used in subscripts}
-\nomenclature[L]{$M$}{Maximum number of retransmission of a packet in the ALOHA protocol in Chapter~\ref{chapter:4}, and number of player for multi-players bandit games in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$\mathrm{MaxBackOff}$}{Maximum number of retransmission of a packet in the ALOHA protocol in Chapter~\ref{chapter:4}}
 \nomenclature[L]{$m$}{Maximum length of the back-off interval after a collision, in Chapter~\ref{chapter:4}}
+\nomenclature[L]{$M$}{Number of player for multi-players bandit games in Chapter~\ref{chapter:5}}
 \nomenclature[L]{$i$, $j$}{Usually denotes the $i$-th or $j$-th player, $i,j\in[M]$ in Chapter~\ref{chapter:5}}
 \nomenclature[L]{$N$}{Usually denotes the number of independent repetitions of the same numerical experiments (\eg, $N=1000$).
-In Chapter~\ref{chapter:4}, denote the number of IoT (dynamic) devices in the studied network}
+In Chapter~\ref{chapter:4}, $N$ denotes the number of IoT (dynamic) devices in the studied network}
 

### 2019-06-12 2-Chapters/2-Chapter/chapter2.tex
Focussed -> focused. Okay... http://www.future-perfect.co.uk/grammar-tip/is-it-focussed-or-focused/

@@
 %
-Such models were less studied than stationary or adversarial models, but some interesting works focussed on Markov models in the last 10 years\footnote{~
+Such models were less studied than stationary or adversarial models, but some interesting works focused on Markov models in the last 10 years\footnote{~
 For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.}.
@@
 These lower-bounds are of highest interest to design efficient algorithms,
-and a significant part of the research literature on MAB algorithms has focussed on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
+and a significant part of the research literature on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
 This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
@@
 %
-We focussed in this thesis on finite-arms and one-dimensional bandit problems,
+We focused in this thesis on finite-arms and one-dimensional bandit problems,
 and thus two possible directions of future works could be to extend our works

### 2019-06-12 2-Chapters/3-Chapter/chapter3.tex
Focussed -> focused. Okay... http://www.future-perfect.co.uk/grammar-tip/is-it-focussed-or-focused/

@@
 
-We focussed here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
+We focused here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
 To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.

### 2019-06-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Focussed -> focused. Okay... http://www.future-perfect.co.uk/grammar-tip/is-it-focussed-or-focused/

@@
 % Similarly, Philippe Mary asked me at the MoTION workshop where I presented \cite{Bonnefoi2019WCNC}, why didn't we rather used Bayes-UCB instead of \UCB{} (from \cite{Kaufmann12BUCB}), as it is known to empirically outperform \UCB{} (as illustrated for instance in \cite{darak2016bayesian,kumar2016two}).
-We could have chose any bandit algorithm for the ``first stage'' component used in this Section, and for simplicity and clarity we focussed on a simple but efficient one (\UCB).
+We could have chose any bandit algorithm for the ``first stage'' component used in this Section, and for simplicity and clarity we focused on a simple but efficient one (\UCB).
 We believe that the same empirical results, but more importantly, the same conclusions, could be given if we used Bayes-UCB or other bandit algorithm for the base building block in the different heuristics proposed in this section.

### 2019-06-12 2-Chapters/4-Chapter/chapter4.tex
Commentaires de Xtof pour le Ch4
Focussed -> focused. Okay... http://www.future-perfect.co.uk/grammar-tip/is-it-focussed-or-focused/

@@
 
-% \TODOL{Intégrer les derniers commentaires de Christophe}
-
-% FIXME finish !
 \paragraph{Abstract.}
@@
 
-However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $M=10$).
+However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
 We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics significantly improve the network efficiency.
@@
 
-To sum-up, we focussed in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices of the IoT network.
+To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices of the IoT network.
 We presented two models of wireless IoT network without sensing, inspired by the ALOHA protocol, and with or without retransmission of uplink packages in case of collisions.

### 2019-06-12 2-Chapters/5-Chapter/chapter5.tex
Focussed -> focused. Okay... http://www.future-perfect.co.uk/grammar-tip/is-it-focussed-or-focused/

@@
 
-Finally, we note that previous works all focussed on the ``sensing'' case, but most likely sub-linear regret can be achieve by decentralized algorithms that leverage the same techniques as introduced by \cite{BoursierPerchet18,LugosiMehrabian18} for the ``no sensing case''.
+Finally, we note that previous works all focused on the ``sensing'' case, but most likely sub-linear regret can be achieve by decentralized algorithms that leverage the same techniques as introduced by \cite{BoursierPerchet18,LugosiMehrabian18} for the ``no sensing case''.
 %
@@
 
-In \cite{darak2016bayesian,Darak16}, the authors study the same model and focussed on the \RhoRand{} policy, which was the state-of-the-art back in $2016$, combined with \UCB{} or \klUCB, as well as an extension using Bayes-UCB from \cite{Kaufmann12BUCB}.
+In \cite{darak2016bayesian,Darak16}, the authors study the same model and focused on the \RhoRand{} policy, which was the state-of-the-art back in $2016$, combined with \UCB{} or \klUCB, as well as an extension using Bayes-UCB from \cite{Kaufmann12BUCB}.
 Their idea is to use the same skeleton as \RhoRand, that is to assign a \emph{rank} to each player, and make players change their rank after any collision.
@@
 Since the beginning of this thesis, all the studied bandit problems were stationary: the rewards coming from choosing arm $k$ are \iid{} and follow the same distribution $\nu_k$.
-The next Chapter~\ref{chapter:6} is focussed on the piece-wise stationary bandit model, a relaxation of this hypothesis.
+The next Chapter~\ref{chapter:6} is focused on the piece-wise stationary bandit model, a relaxation of this hypothesis.
 

### 2019-06-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Focussed -> focused. Okay... http://www.future-perfect.co.uk/grammar-tip/is-it-focussed-or-focused/

@@
 For cognitive radio applications, as well as other applications such as recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
-Indeed, in cognitive radio, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate.
+Indeed, in cognitive radio and IoT networks, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate.
 And for instance in online recommendation, the popularity of items is also subject to trends.
@@
 The second reason is that passively adaptive algorithms function as black-box models: they do not try to detect changes and do not explain why they changed from one arm to another.
-A lot of the recent research has been focussed on actively adaptive algorithms, precisely because they allow to interpret their decisions, by detecting changes on arms with statistical tests.
+A lot of the recent research has been focused on actively adaptive algorithms, precisely because they allow to interpret their decisions, by detecting changes on arms with statistical tests.
 The piece-wise stationary MAB model was then studied by \cite{Kocsis06} and \cite{YuMannor09}.
@@
 Our proposal \GLRklUCB{} is inspired by both the \MUCB{} and \CUSUMUCB{} algorithms.
-Previous works focussed on using \UCB{} \cite{LiuLeeShroff17,CaoZhenKvetonXie18},
+Previous works focused on using \UCB{} \cite{LiuLeeShroff17,CaoZhenKvetonXie18},
 but we propose to use \klUCB{} instead, as it is known to be more efficient for Bernoulli rewards as well as for a more generic case of bounded or one-dimensional exponential families \cite{KLUCBJournal}.

### 2019-06-12 2-Chapters/Conclusion/conclusion.tex
Focussed -> focused. Okay... http://www.future-perfect.co.uk/grammar-tip/is-it-focussed-or-focused/

@@
 
-As detailed in Chapter~\ref{chapter:2}, we focussed in this thesis on finite-arms and one-dimensional bandit problems,
+As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on finite-arms and one-dimensional bandit problems,
 and thus two possible directions of future research could be to extend our works

### 2019-06-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Commentaires Christophe Ch6

@@
 For cognitive radio applications, as well as other applications such as recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
-Indeed, in cognitive radio and IoT networks, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate.
+Indeed, in cognitive radio or IoT networks, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate, making it non stationary.
 And for instance in online recommendation, the popularity of items is also subject to trends.
@@
 %
-In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $k$ at round $t$ has some mean $\mu_k(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $k_t^* = \arg\max_i \mu_k(t)$.
+In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $k$ at round $t$ has some mean $\mu_k(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $k_t^* \eqdef \arg\max_i \mu_k(t)$.
 
@@
 We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
-At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) = Y_{A(t),t}$.
+At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) \eqdef Y_{A(t),t}$.
 At time $t$, we denote by $k_t^*$ an arm with maximal expected reward, \ie, $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$, called an optimal arm (possibly not unique).
@@
 Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
-For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new objects, then one (or some) channel(s) see their mean occupancy rates abruptly change.
+For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company (or network provider) use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new objects, then one (or some) channel(s) see their mean occupancy rates abruptly change.
 
@@
 We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
-To our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
+To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
 Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
@@
 
-Both D-UCB and SW-UCB builds on a stationary policy, but for example EXP3.S from \cite{Auer02NonStochastic}
+Both D-UCB and SW-UCB build on a stationary policy, but for example EXP3.S from \cite{Auer02NonStochastic}
 builds on the EXP3 policy, which is designed for the adversarial case.
@@
 the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
-When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
+When using an efficient Change-point Detector (CD) algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
 %
@@
 %
-The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
+The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but their approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
 %
@@
 %
-Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
-% \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\alpha/K$ at each time step,
-% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\alpha \right\rceil$).
+Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\omega\in(0,1)$.
+% \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\omega/K$ at each time step,
+% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\omega \right\rceil$).
 % Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
@@
 As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
-As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$, but most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
+As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$.
+But most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
 %
 \CUSUMUCB{} achieves a better regret upper bound, of the order of $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but \emph{only} for Bernoulli distributions,
-and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
-%
-\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).
+and when its $4$ parameters ($\omega,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
+%
+\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\omega,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).
 
@@
 It proceeds in episodes, starting a new episode when the algorithm detects a change in one of the arms.
-Each episode is an estimation phase, playing both arms alternatingly until their means can be distinguished, then an exploitation phase.
-With some low probability, an exploration phase is started that checks whether a change has occurred.
-Similarly to what previous approaches did (\eg, \CUSUM{} uses a threshold $\varepsilon$), this phase checks for changes of a certain minimum magnitude, but the innovation in this work is that algorithm does not need any prior knowledge of the minimum magnitude.
-Instead, it uses geometrically decreasing magnitudes $d_k = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only consider a few different values of $d_k$.
+Each episode consists of three phases, estimation, exploitation then exploration.
+First, an estimation phase where both arms are played alternatingly until their means can be distinguished.
+Then follows an exploitation phase, and finally with some low probability, an exploration phase is started that checks whether a change has occurred.
+Similarly to what previous approaches did (\eg, \CUSUM{} uses a threshold $\varepsilon$), this phase checks for changes of a certain minimum magnitude, but the innovation in this work is that this algorithm does not need any prior knowledge of the minimum magnitude.
+Instead, it uses geometrically decreasing magnitudes $d_k = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only considers a few different values of $d_k$.
 % Our  algorithmAdSwitch(shown  as  Algorithm  1)  proceeds  in  episodes,  where  a  newepisode starts when the algorithm detects a change in one of the arms.
@@
 It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
-The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} \cite{Besson2018WCNC}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
-At each time step, a new expert is introduced, and experts corresponds to different Thompson sampling algorithms, each using a different history of pulls and rewards.
+The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} \cite{Besson2018WCNC}, modified to efficiently aggregate a growing number of experts from techniques presented in \cite{Mourtada17}.
+At each time step, a new expert is introduced, and experts correspond to different Thompson sampling algorithms, each using a different history of pulls and rewards.
 Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
@@
 
-This setting is quite different from the setting we are interested by in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
+This setting is quite different from the setting we are considering in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
 hence we do no include any of these algorithms in the experiments presented in Section~\ref{sec:6:NumericalExperiments}.
@@
 In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
-we assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
+we assume that the algorithm knows in advance the number of break-points $\Upsilon_T$.
 Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
@@
 From Table~1 in \cite{Luo18}, one can see that their two algorithms Exp4.S and Ada-ILTCB recover the same regret guarantees as we do, in the non-contextual case:
-that is, a $\cO(\Upsilon_T \sqrt{T}$ regret without knowing the number of changes and $\cO(\sqrt{\Upsilon_T T})$ with this knowledge.
+that is, a $\cO(\Upsilon_T \sqrt{T})$ regret without knowing the number of changes and $\cO(\sqrt{\Upsilon_T T})$ with this knowledge.
 We discuss both these algorithms:
@@
 However, this tuning is only possible in a ``balanced'' instance without stationary sequence longer than $T/\Upsilon_T$.
-In a piece-wise stationary instance with a stationary sequence of length $T/2$, one cannot get better than $\cO(\Upsilon_T \sqrt{T})$ regret, by setting $L=T/2$ (see for instance the problem 4 in our benchmark below in Section~\ref{sec:6:NumericalExperiments}), and as such the regret bound of Ada-ILTCB appear much worse than the announced $\cO(\sqrt{\Upsilon_T T})$ bound.
+In a piece-wise stationary instance with a stationary sequence of length $T/2$, one cannot get better than $\cO(\Upsilon_T \sqrt{T})$ regret, by setting $L=T/2$ (see for instance the problem 4 in our benchmark below in Section~\ref{sec:6:NumericalExperiments}), and as such the regret bound of Ada-ILTCB appears much worse than the announced $\cO(\sqrt{\Upsilon_T T})$ bound.
 %
@@
     \item
-On the one hand, we expect the \emph{local restart} variant of \GLRklUCB{} to outperform the \emph{global restart} variant for problem $1$, as the intuition suggests that it is sub-optimal to reinitialize the memory of the observations of the $K-1$ arms whose distributions did not change after a change-point is detected (as the \emph{global restart} variant do).
+On the one hand, we expect the \emph{local restart} variant of \GLRklUCB{} to outperform the \emph{global restart} variant for problem $1$, as the intuition suggests that it is sub-optimal to reinitialize the memory of the observations of the $K-1$ arms whose distributions did not change after a change-point is detected (as the \emph{global restart} variant does).
     \item
-On the other hand, we expect the \emph{global restart} variant to outperform the \emph{local restart} variant for problem $2$, as detecting a change one any arm is enough to know that all arms changed and to reinitialize the memory of all arms, and thus the intuition suggests that it is sub-optimal to reinitialize only the memory of the observation of the arm on which the change-point was detected (as the \emph{local restart} variant do).
+On the other hand, we expect the \emph{global restart} variant to outperform the \emph{local restart} variant for problem $2$, as detecting a change on any arm is enough to know that all arms changed and to reinitialize the memory of all arms, and thus the intuition suggests that it is sub-optimal to reinitialize only the memory of the observation of the arm on which the change-point was detected (as the \emph{local restart} variant does).
 \end{itemize}
@@
 %
-Surprisingly, numerical experiments show that the \emph{local} variant of \GLRklUCB{} actually outperforms the \emph{glocal} variant for both problem $1$ and $2$, as shown below in Table~\ref{table:6:totalResults1}, and in other problems as shown in Table~\ref{table:6:totalResults2}.
+Surprisingly, numerical experiments show that the \emph{local} variant of \GLRklUCB{} actually outperforms the \emph{global} variant for both problem $1$ and $2$, as shown below in Table~\ref{table:6:totalResults1}, and in other problems as shown in Table~\ref{table:6:totalResults2}.
 %
@@
 
-\paragraph{About \klUCB.}
+\paragraph{About kl-UCB.}
 %
@@
 For a fair comparison, we therefore chose to compare the different change-point detector algorithms combined with \klUCB.
-We also include in Table~\ref{table:6:comparisonUCB_klUCB} a comparison of the performance of different algorithms when using \UCB{} or \klUCB{} indexes, to illustrate that using a more efficient index policy always improves performance, as predicted.
+We also include in Table~\ref{table:6:comparisonUCB_klUCB} a comparison of the performance of different algorithms when using \UCB{} or \klUCB{} indexes, in order to illustrate that using a more efficient index policy always improves performance, as predicted.
 
@@
 %
-A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
+A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t \eqdef \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
 $\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall t\in\N^*, \bE[X_t] = \mu_0\right)$.
@@
     We call a \emph{sub-Bernoulli distribution} any distribution $\nu$ that satisfies
-    $\ln \bE_{X\sim\nu}\left[e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu=\bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) = \ln(1-\mu + \mu e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$, for any $\mu\in[0,1]$.
+    $\ln \bE_{X\sim\nu}\left[e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu \eqdef \bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) \eqdef \ln(1-\mu + \mu e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$, for any $\mu\in[0,1]$.
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
 %
-By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$),
+By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) \eqdef 0$ if $t=0$),
 $\kl(x,y) \eqdef x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$
@@
 %
-We give below two lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.
+We give below two lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there are enough samples before a change.
 
@@
 \end{equation}
-where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$,
-%
-we define $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise, for any $x \ge 0$,
+where for $u \ge 1$ we define $h(u) \eqdef u - \ln(u)$ and its inverse $h^{-1}(u)$,
+%
+we define $\widetilde{h}(x) \eqdef e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\widetilde{h}(x) \eqdef (3/2) (x-\ln(\ln (3/2)))$ otherwise, for any $x \ge 0$,
 and with the value $\zeta(2) = \pi^2 / 6$.
@@
     It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proven in the proofs.
-    For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} in page~\pageref{par:6:controllingDelayInOneProof}.
+    For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} (more specifically, in page~\pageref{par:6:controllingDelayInOneProof}).
     % \TODOL{We have to give the proof, or remove this result?
@@
 We now present the \GLRklUCB{} algorithm, which combines a bandit algorithm with a change-point detector running on each arm.
-It also needs a third ingredient, some forced exploration parameterized by $\alpha\in(0,1)$ to ensure each arm is sampled enough and changes can also be detected on arms currently under-sampled by the bandit algorithm.
+It also needs a third ingredient, some forced exploration parameterized by $\omega\in(0,1)$ to ensure each arm is sampled enough and changes can also be detected on arms currently under-sampled by the bandit algorithm.
 %
@@
         % \KwIn{\emph{Problem parameters}: $T\in\N^*$, $K\in\N^*$\;}
-        \KwIn{\emph{Parameters}: exploration probability $\alpha \in (0, 1)$, confidence level $\delta>0$\;}
-        \KwIn{\emph{Option}: \textbf{Local} or \textbf{Global} restart\;}
+        \KwIn{\emph{Parameters}: exploration rate $\omega \in (0, 1)$, confidence level $\delta>0$}
+        \KwIn{\emph{Option}: \textbf{Local} or \textbf{Global} restart}
         % \KwData{Data}
@@
         % XXX Algorithm
-        \textbf{Initialization: } $\forall k \in [K]$, $\tau_k = 0$ and $n_k = 0$ \\
+        \textbf{initialization: } $\forall k \in [K]$, $\tau_k = 0$ and $n_k = 0$\;
         \For{$t=1,2,\dots, T$}{
-            \uIf{
-                $\alpha>0$ and $t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor \in [K]$\;
+            \uIf(\tcp*[f]{forced exploration}){
+                $t \mod \left\lfloor \frac{K}{\omega}\right\rfloor \in [K]$
             }{
-                $A(t) = t \mod \left\lfloor \frac{K}{\alpha}\right\rfloor$\;
-                \tcp*[f]{forced exploration} \\
+                $A(t) = t \mod \left\lfloor \frac{K}{\omega}\right\rfloor$;
             }
-            \Else{
-                $A(t) \in \cU\left( \arg\max\limits_{k \in [K]} \mbox{UCB}_k(t) \right)$ as defined in \eqref{def:6:UCB}\;
+            \uElse{
+                $A(t) \in \cU\left( \arg\max_{k \in [K]} \mbox{UCB}_k(t) \right)$, with $\mbox{UCB}_k(t)$ defined in \eqref{def:6:UCB}\;
             }
-            \mbox{Play arm } $A(t)$ \mbox{and receive the reward } $Y_{A(t),t}$ : $n_{A(t)} = n_{A(t)} + 1; Z_{A(t), n_{A(t)}} = Y_{A(t),t}$\;
+            \mbox{Play arm } $A(t)$ : $n_{A(t)} = n_{A(t)} + 1$\;
+            \mbox{Observe the reward} $Y_{A(t),t}$ : $Z_{A(t), n_{A(t)}} = Y_{A(t),t}$\;
             \uIf(\tcp*[f]{Change point is detected}){
@@
                 \uIf{\textbf{Global} restart}{
-                    $\forall k\in [K], \tau_k = t$ and $n_k = 0$\;
+                    $\forall k\in [K], \tau_k = t$ and $n_k = 0$;
                     \tcp*[f]{restart all arms}
                     }
-                \Else{
-                    $\tau_{A(t)} = t$ and $n_{A(t)} = 0$\;
+                \uElse{
+                    $\tau_{A(t)} = t$ and $n_{A(t)} = 0$;
                     \tcp*[f]{restart only this arm}
@@
 To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
+\begin{small}
     \begin{equation}\label{eq:whenDoesGLRklUCBrestart}
@@
     \end{equation}
+\end{small}
 %
 with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}.
+We remind in this notation, if $\widehat{Z}_{1:s}=\widehat{Z}_{s+1:n}$, \ie, if all observations $Z_1,\dots,Z_n$ are equal, we set the left-hand side of \eqref{eq:whenDoesGLRklUCBrestart} to $0$ and the B-GLRT cannot detect a change-point.
+%
 We define the (\klUCB{} like) index used by our algorithm, by denoting
@@
 %
-For this algorithm, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
+For this algorithm, we simultaneously investigate two possible behaviors:
+\emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
 %
@@
 $(31)$-$(32)$ in the paper \cite{LiuLeeShroff17}.}:
-$f(t-\tau_k(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{k=1}^K n_k(t)$.
+$f(t-\tau_k(t))$ is replaced by $f(n_t)$ with $n_t \eqdef \sum_{k=1}^K n_k(t)$.
 
@@
 We present in Appendix~\ref{sec:6:exploringDifferentThresholdFunctions} numerical simulations that compare different choices of threshold functions $\beta$.
-We compare the non-explicit function used in Lemma~\ref{lem:6:FalseAlarm} (that make use of a numerical approximation of the function $\cT$),
+We compare the non-explicit function used in Lemma~\ref{lem:6:FalseAlarm} (that makes use of a numerical approximation of the function $\cT$),
 with the simpler value $\beta(n,\delta) = \ln(3n\sqrt{n}/\delta)$, as well as two other choices.
@@
 The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
-Both approaches are parameterized by $\alpha\in(0,1)$.
-More precisely, \CUSUMUCB{} samples arm $k$ with probability $\alpha/K$ at each time step,
-and \MUCB{} uses a deterministic scheme, to force exploring the $K$ arms: when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\alpha \right\rceil$.
-Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
+Both approaches are parameterized by $\omega\in(0,1)$.
+More precisely, \CUSUMUCB{} samples arm $k$ with probability $\omega/K$ at each time step,
+and \MUCB{} uses a deterministic scheme, to force exploring the $K$ arms: when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\omega \right\rceil$.
+Both solutions ensure that all arms are sampled enough on each stationary sequence, so that the CD algorithm has enough \iid{} samples to detect changes.
 %
@@
     % \begin{equation}
-    $n_k(t) - n_k(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor$.
+    $n_k(t) - n_k(s) \geq \left\lfloor \frac{\omega}{K} (t-s) \right\rfloor$.
     % \end{equation}
@@
     we consider two time steps $s\leq t\in\N^*$, chosen between two restart times for that arm $k$.
-    Lines~$3$-$4$ state that $A(u) = u \mod \lceil \frac{K}{\alpha} \rceil$
-    if $u \mod \lceil \frac{K}{\alpha} \rceil \in [K]$,
+    Lines~$3$-$4$ state that $A(u) = u \mod \lceil \frac{K}{\omega} \rceil$
+    if $u \mod \lceil \frac{K}{\omega} \rceil \in [K]$,
     % (see Algorithm~\ref{algo:6:GLRklUCB} for details).
@@
         & = \sum_{u=s+1}^{t} \indic(A(u) = k) \\
-        & \geq \sum_{u=s+1}^{t} \indic\left(A(u) = k, A(u) = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
-        & \geq \sum_{u=s+1}^{t} \indic\left(k = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
-        & = \bigl(t - (s+1) + 1\bigr) / \left\lceil \frac{K}{\alpha} \right\rceil
-        \geq \left\lfloor \frac{\alpha}{K} (t - s) \right\rfloor.
+        & \geq \sum_{u=s+1}^{t} \indic\left(A(u) = k, A(u) = u \mod \left\lceil \frac{K}{\omega} \right\rceil\right) \\
+        & \geq \sum_{u=s+1}^{t} \indic\left(k = u \mod \left\lceil \frac{K}{\omega} \right\rceil\right) \\
+        & = \bigl(t - (s+1) + 1\bigr) / \left\lceil \frac{K}{\omega} \right\rceil
+        \geq \left\lfloor \frac{\omega}{K} (t - s) \right\rfloor.
     \end{align*}
@@
 
-This section give the finite-time regret bounds for the two variants, and interpretations of the different results.
+This section gives the finite-time regret bounds for the two variants, and interpretations of the different results.
 Although we only prove one bound, the other proof can be found in \cite{Besson2019GLRT}.
@@
 % ----------------------------------------------------------------------------
-\subsection{Results for \GLRklUCB{} using Global Changes}
+\subsection{Results for \GLRklUCB{} using \emph{global restarts}}
 
@@
     % \begin{equation}\label{def:6:DelayGlobal}
-    $d^{(i)} = d^{(i)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta^{(i)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil$.
+    $d^{(i)} \eqdef d^{(i)}(\omega,\delta) \eqdef \Bigl\lceil \frac{4K}{\omega\left(\Delta^{(i)}\right)^2}\beta(T,\delta) + \frac{K}{\omega} \Bigr\rceil$.
     % \end{equation}
@@
 Under this assumption, we provide in Theorem~\ref{thm:6:mainRegretBoundGlobal} a finite time problem-dependent regret upper bound.
-It features the parameters $\alpha$ and $\delta$,
+It features the parameters $\omega$ and $\delta$,
 the KL-divergence terms $\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})$ expressing the hardness of the (stationary) MAB problem between two breakpoints,
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on \textbf{Global} Restart satisfies
+    For $\omega$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based on \textbf{Global} Restart satisfies
     \begin{align}
-        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T \\
+        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T \\
         & + \sum_{k=1}^{\Upsilon_T}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
@@
     \begin{enumerate}
-        \item Choosing $\alpha = \sqrt{\ln(T)/T}$, $\delta = 1/\sqrt{T}$ gives
+        \item Choosing $\omega = \sqrt{\ln(T)/T}$, $\delta = 1/\sqrt{T}$ gives
         \begin{equation}
@@
         \end{equation}
-        \item Choosing $\alpha = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ gives
+        \item Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ gives
         \begin{equation}
@@
 % ----------------------------------------------------------------------------
-\subsection{Results for \GLRklUCB{} using Local Changes}
-
-Some new notations are needed to state a regret bound for \GLRklUCB{} using local changes,
-and to distinguish notations between local and global changes, we denote $\ell$ instead of $k$ for the indexes of change-points.
+\subsection{Results for \GLRklUCB{} using Local Restarts}
+
+Some new notations are needed to state a regret bound for \GLRklUCB{} using \emph{local restarts},
+and to distinguish notations between the results for the two variants (local and global restarts), we denote $\ell$ instead of $k$ for the indexes of change-points.
 We let $\tau_k^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $k$}: $\tau_k^{(\ell)} = \inf \{ t > \tau_k^{(\ell - 1)} : \mu_k(t) \neq \mu_k(t+1)\}$,
@@
     % \begin{equation}\label{def:6:DelayLocal}
-    $d_k^{(\ell)} = d_k^{(\ell)}(\alpha,\delta) = \Bigl\lceil \frac{4K}{\alpha\left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \Bigr\rceil$.
+    $d_k^{(\ell)} \eqdef d_k^{(\ell)}(\omega,\delta) \eqdef \Bigl\lceil \frac{4K}{\omega\left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) + \frac{K}{\omega} \Bigr\rceil$.
     % \end{equation}
@@
 
-Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to the best of the authors knowledge,our analysis is the first one to be given for such assumption on the problem difficulty.
+Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to the best of or knowledge, our analysis is the first one to be given for such assumption on the problem difficulty.
 %
 It requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
-Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange = \sum_{k=1}^K \NCk \leq K \Upsilon_T$.
+Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange \eqdef \sum_{k=1}^K \NCk \leq K \Upsilon_T$.
 
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    For $\alpha$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\alpha$ and $\delta$ based on Local Restart satisfies
+    For $\omega$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based on Local Restart satisfies
     %
     \begin{equation}
-        R_T \leq 2\sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{4K}{\alpha \left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) +\alpha T + 2 \delta \EffChange  + \sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{\ln(T)}{\kl(\overline{\mu}_k^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
-    \end{equation}
-    where ${\mu}^*_{i,\ell} = \inf \left\{ \mu_{k_t^*}(t) : \mu_{k_t^*}(t) \neq \overline{\mu}_k^{(\ell)}, t \in [\tau_k^{(\ell)}+1, \tau_k^{(\ell+1)}]\right\}$.
+        R_T \leq 2\sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{4K}{\omega \left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) +\omega T + 2 \delta \EffChange  + \sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{\ln(T)}{\kl(\overline{\mu}_k^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
+    \end{equation}
+    where ${\mu}^*_{i,\ell} \eqdef \inf \left\{ \mu_{k_t^*}(t) : \mu_{k_t^*}(t) \neq \overline{\mu}_k^{(\ell)}, t \in [\tau_k^{(\ell)}+1, \tau_k^{(\ell+1)}]\right\}$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 %
-We can directly obtain different regret upper-bounds for different choices of the two parameters $\alpha$ and $\delta$.
+We can directly obtain different regret upper-bounds for different choices of the two parameters $\omega$ and $\delta$.
 We prefer to state the four cases, to highlight the modularity of the result given by Theorem~\ref{thm:6:mainRegretBound}.
@@
     \begin{enumerate}
-        \item Choosing $\alpha = \sqrt{\ln(T) / T}$, $\delta = 1 / \sqrt{T}$ (with no prior knowledge of $\Upsilon_T$ or $\EffChange$) gives
+        \item Choosing $\omega = \sqrt{\ln(T) / T}$, $\delta = 1 / \sqrt{T}$ (with no prior knowledge of $\Upsilon_T$ or $\EffChange$) gives
         \begin{equation}
@@
         \end{equation}
-        \item Choosing $\alpha = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and optimist guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
+        \item Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and optimist guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
         \begin{equation}
@@
         \end{equation}
-        \item Choosing $\alpha = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ (with prior knowledge of $\EffChange$) gives
+        \item Choosing $\omega = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ (with prior knowledge of $\EffChange$) gives
         \begin{equation}
@@
         \end{equation}
-        \item Choosing $\alpha = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and pessimist guess $\EffChange \simeq K \Upsilon_T$) gives
+        \item Choosing $\omega = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and pessimist guess $\EffChange \simeq K \Upsilon_T$) gives
         \begin{equation}
@@
 
-The proof of \GLRklUCB{} with Local Changes is not included, but it can be found in the Appendix of \cite{Besson2019GLRT}, as it is quite similar to the proof for Global changes given above.
+The proof of \GLRklUCB{} with \emph{local restarts} is not included, but it can be found in the Appendix of \cite{Besson2019GLRT}, as it is quite similar to the proof for \emph{global restarts} given above.
 The main difficulty relies in defining the ``good event'' $\cE_T$, and proving that it happens with high probability (by showing that the complementary event $\cE_T^c$ is highly unlikely).
@@
 Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
-both show that there exists a tuning of $\alpha$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
-%
-For very particular instances such that $\Upsilon_T = \EffChange$, \ie, at each break-point only one arm changes (\eg, problem $1$ from Figure~\ref{fig:6:Problem_1}), the theory advocates the use of local changes.
-%
-Indeed, while the regret guarantees obtained are similar, those obtained for local changes hold for a wider variety of problems as Assumption~\ref{ass:6:LongPeriods} is less stringent than Assumption~\ref{ass:6:LongPeriodsGlobal}.
-%
-Besides those specific instances, our results are essentially worse for local than global changes.
+both show that there exists a tuning of $\omega$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
+%
+For very particular instances such that $\Upsilon_T = \EffChange$, \ie, at each break-point only one arm changes (\eg, problem $1$ from Figure~\ref{fig:6:Problem_1}), the theory advocates the use of local restarts.
+%
+Indeed, while the regret guarantees obtained are similar, those obtained for local restarts hold for a wider variety of problems as Assumption~\ref{ass:6:LongPeriods} is less stringent than Assumption~\ref{ass:6:LongPeriodsGlobal}.
+%
+Besides those specific instances, our results are essentially worse for \emph{local} than \emph{global restarts}.
 However, we only obtain regret upper bounds -- thus providing a theoretical safety net for both variants of our algorithm, and the practical story is different, as discussed in Section~\ref{sec:6:NumericalExperiments}.
@@
 
-One can note that with the tuning of $\alpha$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ rounds.
+One can note that with the tuning of $\omega$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ rounds.
 Hence those guarantees are valid on ``easy'' problems only, with few changes of large magnitudes, in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
@@
 
-This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the proof of the finite-time regret bound for the variant using ``global restarts''.
-The proof of the other variant using ``local restarts'' follows the skeleton of the unified analysis, and it can be found in Appendix~E of \cite{Besson2019GLRT}.
+This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the proof of the finite-time regret bound for the variant using \emph{global restarts}.
+The proof of the other variant using \emph{local restarts} follows the skeleton of the unified analysis, and it can be found in Appendix~E of \cite{Besson2019GLRT}.
 
@@
 %
-Recall that in this Chapter, the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{k_t^*}(t) - \mu_{A(t)}(t))\right]$.
-First, we introduce $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}),
+Recall that in this Chapter, the regret is defined as $R_T \eqdef \bE\left[\sum_{t=1}^T(\mu_{k_t^*}(t) - \mu_{A(t)}(t))\right]$.
+First, we introduce $\cD(T,\omega)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}),
 and as we observe that $\mu_{k_t^*} - \mu_{A(t)} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
@@
     & \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t)) \\
- & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{A(t)}(t) \geq \UCB_{k_t^*}(t)\right) \\
- & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right) + \sum_{k=1}^K\sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_k(t))\indic\left(A(t) = k, \UCB_k(t) \geq \mu_{k_t^*}(t)\right).
+    & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\omega)) + \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(t \notin\cD(T,\omega), \UCB_{A(t)}(t) \geq \UCB_{k_t^*}(t)\right) \\
+    & \hspace{1cm}\leq  \omega T + \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right) + \sum_{k=1}^K\sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_k(t))\indic\left(A(t) = k, \UCB_k(t) \geq \mu_{k_t^*}(t)\right).
 \end{align*}
@@
 \begin{align}\label{eq:6:GeneRegretBound}
-    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right)\right]}_{(A)} \\
+    R_T \leq T \bP\left(\cE_T^c\right) + \omega T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right)\right]}_{(A)} \\
     & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(\UCB_{A(t)}(t) \geq \mu_{k_t^*}(t)\right)\right]}_{(B)}. \nonumber
@@
 %
-To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(i)}+1, \tau^{(i+1)}]$ for global changes and $[\tau^{(i)}_k+1, \tau^{(i+1)}_k]$ for each arm $k$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
+To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(i)}+1, \tau^{(i+1)}]$ for \emph{global restarts} and $[\tau^{(i)}_k+1, \tau^{(i+1)}_k]$ for each arm $k$ for \emph{local restarts}, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
 
 The tricky part of each proof, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$.
-For example for local changes (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
+For example for \emph{local restarts} (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
 % \begin{equation}
-$\cE_T(\alpha,\delta) = \left(\forall k \in [K], \forall \ell \in [\NCk], \widehat{\tau}^{(\ell)}_k \in \left[\tau_k^{(\ell)} + 1, \tau_k^{(\ell)} + d_k^{(\ell)}\right] \right)$,
+$\cE_T(\omega,\delta) \eqdef \left(\forall k \in [K], \forall \ell \in [\NCk], \widehat{\tau}^{(\ell)}_k \in \left[\tau_k^{(\ell)} + 1, \tau_k^{(\ell)} + d_k^{(\ell)}\right] \right)$,
 % \end{equation}
 %
-where $\widehat{\tau}_k^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $k$ and $d_k^{(\ell)}=d_k^{(\ell)}(\alpha,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
+where $\widehat{\tau}_k^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $k$ and $d_k^{(\ell)} = d_k^{(\ell)}(\omega,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
 Introducing the event
-$\cC_k^{(\ell)} = \left\{\forall j \leq \ell, \widehat{\tau}_k^{(\ell)} \in \left[\tau_k^{(j)} + 1, \tau_k^{(j)} + d_k^{(j)} \right] \right\}$,
+$\cC_k^{(\ell)} \eqdef \left\{\forall j \leq \ell, \widehat{\tau}_k^{(\ell)} \in \left[\tau_k^{(j)} + 1, \tau_k^{(j)} + d_k^{(j)} \right] \right\}$,
 that all the changes up to the $\ell$-th have been detected, a union bound yields this decomposition
@@
 \begin{align}
-    \bP(\cE_T(\alpha,\delta)^c) & \leq \sum\limits_{k=1}^K\sum\limits_{\ell=1}^{\NCk} \underbrace{\bP\left(\widehat{\tau}_k^{(\ell)} \leq \tau_k^{(\ell)} \;|\; \cC_k^{(\ell-1)}\right)}_{(a)} + \sum\limits_{k=1}^K\sum\limits_{\ell=1}^{\NCk} \underbrace{\bP\left(\widehat{\tau}_k^{(\ell)} \geq \tau_k^{(\ell)} + d_k^{(\ell)} \;|\; \cC_k^{(\ell-1)}\right)}_{(b)}.
+    \bP(\cE_T(\omega,\delta)^c) & \leq \sum\limits_{k=1}^K\sum\limits_{\ell=1}^{\NCk} \underbrace{\bP\left(\widehat{\tau}_k^{(\ell)} \leq \tau_k^{(\ell)} \;|\; \cC_k^{(\ell-1)}\right)}_{(a)} + \sum\limits_{k=1}^K\sum\limits_{\ell=1}^{\NCk} \underbrace{\bP\left(\widehat{\tau}_k^{(\ell)} \geq \tau_k^{(\ell)} + d_k^{(\ell)} \;|\; \cC_k^{(\ell-1)}\right)}_{(b)}.
 \end{align}
@@
 More precisely, we need to leverage the forced exploration (Proposition~\ref{prop:6:EnoughSamples}) to be sure we have enough samples for detection.
-This explains why delays defined in Assumption~\ref{ass:6:LongPeriods} are scaled by $1/\alpha$.
+This explains why delays defined in Assumption~\ref{ass:6:LongPeriods} are scaled by $1/\omega$.
 Using some elementary calculus and a concentration inequality given in Lemma~\ref{lem:6:Chernoff2arms}, we can finally prove that $(b) \leq \delta$.
@@
 
-\subsection{Proof for \GLRklUCB{} with global changes}
+\subsection{Proof for \GLRklUCB{} with \emph{global restarts}}
 \label{proof:6:mainRegretBoundGlobal}
@@
 % \begin{align}\label{eq:6:GeneRegretBound}
-%     R_T \leq T \bP\left(\cE_T^c\right) + \alpha T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right)\right]}_{(A)} \nonumber \\
+%     R_T \leq T \bP\left(\cE_T^c\right) + \omega T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{k_t^*}(t) \leq \mu_{k_t^*}(t)\right)\right]}_{(A)} \nonumber \\
 %     & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(\UCB_{A(t)}(t) \geq \mu_{k_t^*}(t)\right)\right]}_{(B)}.
@@
 let $\widehat{\tau}^{(i)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
-We denote $n_k(t) = \sum_{s=\tau(t)+1}^{t}\indic(A(s) = i)$ the number of selections of arm $k$ since the last (global) restart, and $\widehat{\mu}_k(t) = \frac{1}{n_k(t)}\sum_{s=\tau(t)+1}^{t}Y_{k,s}\indic(A(s) = i)$ their empirical average (if $n_k(t) \neq 0$).
+We denote $n_k(t) \eqdef \sum_{s=\tau(t)+1}^{t}\indic(A(s) = i)$ the number of selections of arm $k$ since the last (global) restart, and $\widehat{\mu}_k(t) \eqdef \frac{1}{n_k(t)}\sum_{s=\tau(t)+1}^{t}Y_{k,s}\indic(A(s) = i)$ their empirical average (if $n_k(t) \neq 0$).
 
@@
 Let $d^k$ be defined as in Assumption~\ref{ass:6:LongPeriodsGlobal}, we define
-\begin{equation}\cE_T(\delta) = \left( \forall k \in [\Upsilon_T], \widehat{\tau}^{(i)} \in \left[\tau^{(i)} + 1, \tau^{(i)} + d^{(i)}\right] \right).\label{def:6:GoodEvenGlobal}\end{equation}
+\begin{equation}\label{def:6:GoodEvenGlobal}
+    \cE_T(\delta) \eqdef \left( \forall k \in [\Upsilon_T], \widehat{\tau}^{(i)} \in \left[\tau^{(i)} + 1, \tau^{(i)} + d^{(i)}\right] \right).
+\end{equation}
 Under the good event, all the change points are detected within a delay at most $d^(k)$.
@@
 Combining the regret decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEventGlobal} and the two upper bounds of $(A)$ in \eqref{eq:6:TermAFinalGlobal} and of $(B)$ in \eqref{eq:6:TermBFinalGlobal},
-\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\alpha \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\alpha T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
+\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
 which concludes the proof.
@@
 %
-From the definition of $\Delta^{(i)}$, there exists an arm $k$ such that $\Delta^{(i)} = |\mu_k^{(i)} - \mu_k^{(i-1)}|$.
+From the definition of $\Delta^{(i)}$, there exists an arm $k$ such that $\Delta^{(i)} \eqdef |\mu_k^{(i)} - \mu_k^{(i-1)}|$.
 We shall prove that it is unlikely that the change-point detector associated to $i$ doesn't trigger within the delay $d^{(i)}$.
@@
 First, it follows from Proposition~\ref{prop:6:EnoughSamples} that there exists $\overline{t} \in \{\tau^{(i)}, \dots, \tau^{(i)} + d^{(i)} \}$ such that
-$n_k(\overline{t}) - n_k(\tau^{(i)}) = \overline{r}$ where $\overline{r} = \lfloor \frac{\alpha}{K} d^{(i)}\rfloor$
-(as the mapping $t\mapsto n_k(t) - n_k(\tau^{(i)})$ is non-decreasing, is $0$ at $t=\tau^{(i)}$ and its value at $\tau^{(i)}+d^{(i)}$ is larger than $\overline{r}$).
+$n_k(\overline{t}) - n_k(\tau^{(i)}) = \overline{r}$, where $\overline{r} \eqdef \lfloor \frac{\omega}{K} d^{(i)}\rfloor$
+(as the mapping $t \mapsto n_k(t) - n_k(\tau^{(i)})$ is non-decreasing, is $0$ at $t=\tau^{(i)}$ and its value at $\tau^{(i)}+d^{(i)}$ is larger than $\overline{r}$).
 Using that $\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right) \subseteq \left(\widehat{\tau}^{(i)} \geq \overline{t}\right)$
@@
 
-Moreover, due to Proposition~\ref{prop:6:EnoughSamples}, $n_k(\tau^{(i)})$ lies in the interval $
-\left[\left\lfloor \frac{\alpha}{K}\left(\tau^{(i)}-\widehat{\tau}^{(i-1)}\right)\right\rfloor,\left(\tau^{(i)}-\widehat{\tau}^{(i-1)}\right)\right]$.
+Moreover, Proposition~\ref{prop:6:EnoughSamples} shows that $n_k(\tau^{(i)})$ is bounded in a certain interval,
+$n_k(\tau^{(i)}) \in \left[\left\lfloor \frac{\omega}{K}\left(\tau^{(i)}-\widehat{\tau}^{(i-1)}\right)\right\rfloor,\left(\tau^{(i)}-\widehat{\tau}^{(i-1)}\right)\right]$.
 Conditionally to $\cC^{(i-1)}$,
@@
 \begin{equation*}
-    n_k(\tau^{(i)}) \in \left\{ \left\lfloor \frac{\alpha}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} \eqdef \cI_k.
+    n_k(\tau^{(i)}) \in \left\{ \left\lfloor \frac{\omega}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} \eqdef \cI_k.
 \end{equation*}
@@
 
-Using Pinsker's inequality from Lemma~\ref{lem:2:Pinsker} and introducing the gap $\Delta_k^{(i)} = \mu_k^{(i-1)} - {\mu}_k^{(i)}$ (which is such that $\Delta^{(i)} = |\Delta_k^k|$), one can write
+Using Pinsker's inequality from Lemma~\ref{lem:2:Pinsker} and introducing the gap $\Delta_k^{(i)} \eqdef \mu_k^{(i-1)} - {\mu}_k^{(i)}$ (which is such that $\Delta^{(i)} = |\Delta_k^k|$), one can write
 %
@@
 %
-Define $s_{\min} = \left\lfloor \frac{\alpha}{K} (\tau^{(i)}-\tau^{(i-1)})/2\right\rfloor$.
+Define $s_{\min} \eqdef \left\lfloor \frac{\omega}{K} (\tau^{(i)}-\tau^{(i-1)})/2\right\rfloor$.
 Using that the mappings $s \mapsto (s+\overline{r})/s\overline{r}$ and $s \mapsto \beta(s + \overline{r},\delta)$ are respectively decreasing and increasing in $s$, one has, for all $s\in \cI_k$,
@@
     2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) & \leq &
-    2\frac{s_{\min}+\overline{r}}{s_{\min}\overline{r}}\beta\left(T, \delta\right) \leq \frac{4\beta(T,\delta)}{\left\lfloor \frac{\alpha}{K}d^{(i)}\right\rfloor},
+    2\frac{s_{\min}+\overline{r}}{s_{\min}\overline{r}}\beta\left(T, \delta\right) \leq \frac{4\beta(T,\delta)}{\left\lfloor \frac{\omega}{K}d^{(i)}\right\rfloor},
 \end{eqnarray*}
@@
 Now the definition of $d^{(i)}$ readily implies that
-\[\left\lfloor \frac{\alpha}{K}d^{(i)}\right\rfloor > \frac{4\beta(T,\delta)}{\left(\Delta^{(i)}\right)^2},\]
+\[\left\lfloor \frac{\omega}{K}d^{(i)}\right\rfloor > \frac{4\beta(T,\delta)}{\left(\Delta^{(i)}\right)^2},\]
 which yields
@@
 %
-Additionally, we include the Exp3.S algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\alpha=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \mathrm{e}) / ((\mathrm{e}-1)T)})$.
+Additionally, we include the Exp3.S algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \mathrm{e}) / ((\mathrm{e}-1)T)})$.
 
 Our main goal is to compare against \emph{actively adaptive algorithms}.
-We include \CUSUMklUCB{} \cite{LiuLeeShroff17}, tuned with $M=150$ and $\varepsilon=0.1$ for easy problems ($1,2,4$) and $\varepsilon=0.001$ for hard problems ($3,5$),
+We include \CUSUMklUCB, tuned with $M=150$ and $\varepsilon=0.1$ for easy problems ($1,2,4$) and $\varepsilon=0.001$ for hard problems ($3,5$),
 % based on a prior knowledge of the difficulty of the problems,
-and with $h = \log(T/\Upsilon_T)$, $\alpha = \sqrt{\Upsilon_T \log(T/\Upsilon_T) / T}$, as suggested in the paper.
+and with $h = \log(T/\Upsilon_T)$, $\omega = \sqrt{\Upsilon_T \log(T/\Upsilon_T) / T}$, as suggested in the paper \cite{LiuLeeShroff17}.
 % as for the first problems ($1$ to $4$) we consider have large changes in terms of means between change-points.
@@
 %
-Finally, we include \MklUCB{} \cite{CaoZhenKvetonXie18},
+Finally, we include \MklUCB,
 % as a simple adaptation of the \MUCB algorithm,
-tuned with $w=150$, based on a prior knowledge of the problems as the formula using $\delta_{\min}$ given in the paper is too large for small horizons (on all our problem instances), a threshold $b=\sqrt{w \log(2 K T)}$ and $\gamma=\sqrt{\Upsilon_T K (2 b + 3 \sqrt{w}) / (2 T)}$ as suggested by Remark~4 in the paper.
+tuned with $w=150$, based on a prior knowledge of the problems as the formula using $\delta_{\min}$ given in \cite{CaoZhenKvetonXie18} is too large for small horizons (on all our problem instances), a threshold $b=\sqrt{w \log(2 K T)}$ and $\gamma=\sqrt{\Upsilon_T K (2 b + 3 \sqrt{w}) / (2 T)}$ as suggested by Remark~4 in the paper \cite{CaoZhenKvetonXie18}.
 
@@
 using respectively
-$\delta = 1/\sqrt{\Upsilon T}$, $\alpha = \alpha_0 \sqrt{\Upsilon_T \log(T) /T}$
+$\delta = 1/\sqrt{\Upsilon T}$, $\omega = \omega_0 \sqrt{\Upsilon_T \log(T) /T}$
 and
-$\delta = 1/\sqrt{K \Upsilon_T T}$, $\alpha = \alpha_0\sqrt{K \Upsilon_T \log(T) / T}$
+$\delta = 1/\sqrt{K \Upsilon_T T}$, $\omega = \omega_0\sqrt{K \Upsilon_T \log(T) / T}$
 from Corollaries \ref{cor:6:Global} and \ref{cor:6:Local}.
-Both choices of $\alpha$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
-the constant is set to $\alpha_0 = 0.05$.
-We show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\alpha\leq0.1$.
+Both choices of $\omega$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
+the constant is set to $\omega_0 = 0.05$.
+We show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\omega\leq0.1$.
 We do not use a constant $\delta_0$ to change the confidence level to $\delta = \delta_0 \delta_T$, as we show in Appendix~\ref{sec:6:choosingDelta} that $\delta_0=1$ is uniformly better for different problems.
@@
 % \TODOE{Je l'ai déjà dit, mais vraiment on va nous accuser d'avoir optimisé un paramètre chez nous et pas chez les autres.
-Est-ce qu'avec $\alpha_0=1$ on a des résultats comparables ou bien on est complètement à côté ?}
+Est-ce qu'avec $\omega_0=1$ on a des résultats comparables ou bien on est complètement à côté ?}
 %
@@
 \textbf{Problem 3.} (see Figure~\ref{fig:6:Problem_5}) This problem is harder, with $K=6$, $\Upsilon_T=8$ and $T=20000$.
-Most arms change at almost every time steps,
+At every break-points, almost all arms change,
 and means are bounded in $[0.01, 0.07]$.
@@
 for which we compare \CUSUMUCB{} against \CUSUMklUCB, \MUCB{} against \MklUCB, and the oracle policy using \UCB{} against \klUCB.
-We report in the Table~\ref{table:6:comparisonUCB_klUCB} below the results (in terms of mean regret),
-and we observe that using \klUCB{} rather than \UCB{} indices always yield better practical performance.
+We report in Table~\ref{table:6:comparisonUCB_klUCB} below the results (in terms of mean regret),
+and we observe that using \klUCB{} rather than \UCB{} indices always yields better practical performance.
 %
@@
       \end{tabular}
-      \caption{Using \klUCB{} is much more efficient than using \UCB{}, for non-stationary bandit.}
+      \caption{Mean regret $\pm$ $1$ std-dev, on problems $1$ and $2$ with $T=5000$. We conclude that using \klUCB{} is much more efficient than using \UCB{}, for non-stationary bandit.}
       \label{table:6:comparisonUCB_klUCB}
@@
     \begin{tabular}{c|cccccc}
-    \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $1$ & Pb $2$ & Pb $3$ \\
+    \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $1$ & Pb $2$ ($T=5000$) & Pb $3$ ($T=20000$) \\
         \hline

### 2019-06-12 5-Nomenclature/Nomenclature.tex
Commentaires Christophe Ch6

@@
 \nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, the parameter for the $\varepsilon$-greedy algorithm, or the drift correction parameter for \CUSUM{} in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
+\nomenclature[G]{$\alpha$}{Denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}}
+\nomenclature[G]{$\omega$}{Denotes a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
 \nomenclature[G]{$\beta(n,\delta)$}{Usually denotes a threshold for the statistical (GLR) tests in Chapter~\ref{chapter:6}}

### 2019-06-12 2-Chapters/2-Chapter/chapter2.tex
Typos, et commentaires Christophe

@@
 
-In this thesis, we only study analytically the mean regret $R_T^{\cA}$ (in Chapters \ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
+In this thesis, we only study analytically the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
 

### 2019-06-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Typos, et commentaires Christophe

@@
 For simplicity, and for consistency with the rest of this thesis, we only consider a discrete protocol (\ie, slotted).
-As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
+As illustrated in Figure~\ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
 All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
@@
 %
-Figure \ref{fig:41:from10to100} presents the successful transmission rate as a function of time.
+Figure~\ref{fig:41:from10to100} presents the successful transmission rate as a function of time.
 The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
@@
 
-We can see in Figure \ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%.
+We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%.
 When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
-Moreover, we can see in Figures \ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
+Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
 For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
@@
 
-Figure \ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
+Figure~\ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
 Note that when this proportion increases, the assumptions of a stochastic MAB model are clearly violated, and there is no mathematical justification for the efficiency of TS and \UCB{} algorithms.

### 2019-06-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Typos, et commentaires Christophe

@@
 
-We simulate our network considering $N$ devices following the contention Markov process described in Section \ref{sub:43:model}, and a LoRaWAN standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
+We simulate our network considering $N$ devices following the contention Markov process described in Section~\ref{sub:43:model}, and a LoRaWAN standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
 Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of $1\;\mathrm{s}$.

### 2019-06-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Typos, et commentaires Christophe

@@
 It was shown that Exp3.S attains a regret of $\cO(\sqrt{K T S \log(K T)})$, when comparing with an arbitrary sequence of comparators that does not switch for more than $S-1$ times.
-Running their algorithm over the non-stationary problem, and picking their comparators sequence as the best arm in each stationary interval, one can already get $\sqrt(\Upsilon_T T \log(T))$ regret, and the only prior knowledge needed to run the Exp3.S algorithm is $T$ and $\Upsilon_T$.
+Running their algorithm over the non-stationary problem, and picking their comparators sequence as the best arm in each stationary interval, one can already get $\sqrt{\Upsilon_T T \log(T)}$ regret, and the only prior knowledge needed to run the Exp3.S algorithm is $T$ and $\Upsilon_T$.
 Moreover, Exp3.S is efficient both in term of storage and computation time, and simple to implement.
@@
 To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
-Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
-Active strategies monitor the arms' rewards, by using change detection algorithms \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
-Passive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
+Following the vocabulary used in previous works, we make the distinction between \emph{passively} and \emph{actively} adaptive strategies.
+On the one hand, Actively adaptive strategies monitor the arms' rewards, by using change detection algorithms \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
+On the other hand, Passively adaptive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
 % , while active strategies are usually tuned with a threshold for the statistical test.
@@
 where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{t=k}^{k'} X_t$ denotes the average of the observations $X_t$ collected between the instants $k$ and $k'$.
-
-\TODOL{Si on veut, on peut déplacer cette preuve en Annexe, voir même l'enlever, elle n'est pas spécialement palpitante et pas requise pour la suite !}
-
-\begin{smallproof}
-    First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
-    We have $\ell(X_1, \dots, X_n ; \mu_0) = \prod\limits_{i=1}^n \ell(X_i ; \mu_0)$ by independence of the observations $X_i$,
-    and $\ell(X_i ; \mu_0) = \mu_0^{X_i} (1-\mu_0)^{1-X_i}$ for Bernoulli distributions.
-    Therefore, taking the logarithm gives
-    \begin{align*}
-        \log \left[ \ell(X_1, \dots, X_n ; \mu_0) \right]
-        &= \sum_{i=1}^n X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) \\
-        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
-        &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
-    \end{align*}
-    %
-    For a constant $a\in[0,1]$, let $h(x) \eqdef a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
-    and we are trying to solve $\sup_{x\in[0,1]} h(x)$.
-    If $a=0$ or $a=1$, $h$ is maximum at $x=a$.
-    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative,
-    $h'(x) = a/x - (1-a)/(1-x)$, so $h'(x) = 0$ if and only if $x=a$.
-    Thus in all cases, $\sup_{x\in[0,1]} h(x) = h(a)$.
-    %
-    Here, we have $a = \widehat{\mu}_{1:n}$, and thus we solve the $\sup_{\mu_0}$ optimization problem found in the denominator of the $\GLR(n)$ expression.
-    By replacing $\mu_0 = \widehat{\mu}_{1:n}$, we obtained the following (unique) solution,
-    \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
-        \sup_{\mu_0} \log \left[ \ell(X_1, \dots, X_n ; \mu_0) \right] = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
-    \end{equation}
-
-    Now let us consider the nominator of the $\GLR(n)$ expression.
-    We can again work with log-likelihoods, and so we have
-    %
-    \begin{align*}
-        & \log \left[ \ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau) \right] \\
-        &= \sum_{i=1}^{\tau} X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_i \log(\mu_1) + (1-X_i) \log(1-\mu_1) \\
-        &= s \bigl( \widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0) \bigr) \\
-        & \;\;\;\; + (n-s) \bigl( \widehat{\mu}_{s+1:n} \log(\mu_1) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\mu_1) \bigr).
-    \end{align*}
-    Because we solved in \eqref{eq:6:solutionPbOpt_for_1n} the $\sup$ for the denominator,
-    we have
-    \begin{align*}
-        \GLR(n)
-        % &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \dots, X_n ; \mu_0)}, \\
-        &= \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right) \right] }, \\
-        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \frac{\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] }, \\
-        &= \exp \Bigl[ \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr] \Bigr],
-    \end{align*}
-    When the last equation comes from the fact that $\exp$ is increasing.
-    Thus by taking the logarithm of both sides, we obtain
-    \begin{align*}
-        \log \GLR(n)
-        &= \sup\limits_{\mu_0,\mu_1,\tau < n} \Bigl[ \log\Bigl[ \frac{\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{ \exp\left[ n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right) \right] } \Bigr] \Bigr],  \\
-        &= \sup\limits_{\mu_0,\mu_1, 1 < s < n} \Bigl[
-            s \bigl( \widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0) \bigr) \\
-            & \hspace{35pt} + (n-s) \bigl( \widehat{\mu}_{s+1:n} \log(\mu_1) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\mu_1) \bigr) \\
-            & \hspace{35pt} - n \bigl( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \bigr)
-        \Bigr].
-    \end{align*}
-    By linearity and independence, we can separate the joint optimization problem on $\mu_0,\mu_1,s$ in two optimizations problems for $\mu_0,s$ and $\mu_1,s$, that can first be solved explicitly for $\mu_0$ (resp.
-$\mu_1$) and then left to be solved for $s$.
-    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_i = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,
-    which are both handled similarly.
-    For instance for $\mu_0$, we use the same computation as above with the function $h$ to find the optimum:
-    $\widehat{\mu}_{1:s} \log(\mu_0) + (1 - \widehat{\mu}_{1:s}) \log(1-\mu_0)$
-    is optimum for $\mu_0 = \widehat{\mu}_{1:s}$.
-    Similarly, the term for $\mu_1$ gives that $\sup_{\mu_1} \widehat{\mu}_{s+1:n} \log(\mu_1) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\mu_1)$
-    is attained for $\mu_1 = \widehat{\mu}_{s+1:n}$.
-    Finally, by replacing the two expressions of the solutions for $\mu_0$ and $\mu_1$,
-    we obtain
-    \begin{align*}
-        \log \GLR(n)
-        = \sup\limits_{1 < s < n} \Bigl[
-            s \times \Bigl(
-                    & \widehat{\mu}_{1:s} \log(\widehat{\mu}_{1:s}) + (1 - \widehat{\mu}_{1:s}) \log(1-\widehat{\mu}_{1:s}) \\
-                    & - \widehat{\mu}_{1:s} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:s}) \log(1-\widehat{\mu}_{1:n})
-            \Bigr) \\
-            + (n-s) \times \Bigl(
-                & \widehat{\mu}_{s+1:n} \log(\widehat{\mu}_{s+1:n}) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\widehat{\mu}_{s+1:n}) \\
-                & - \widehat{\mu}_{s+1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{s+1:n}) \log(1-\widehat{\mu}_{1:n})
-            \Bigr)
-        \Bigr].
-    \end{align*}
-    We conclude by recognizing the expressions of $s \times \kl(\widehat{\mu}_{1:s}, \widehat{\mu}_{1:n})$
-    and $(n-s) \times \kl(\widehat{\mu}_{s+1:n}, \widehat{\mu}_{1:n})$.
-\end{smallproof}
+This result is proven in Appendix~\ref{app:6:GLR_with_kl}.
+%
 
@@
 
-In this section, we sketch a unified proof for the two Theorems \ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given above.
+In this section, we sketch a unified proof for the two Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given above.
 %  whose detailed proofs are given in Section~\ref{proof:6:mainRegretBoundGlobal} and in Appendix~E of our paper \cite{Besson2019GLRT} respectively.
@@
 $\delta = 1/\sqrt{K \Upsilon_T T}$, $\omega = \omega_0\sqrt{K \Upsilon_T \log(T) / T}$
-from Corollaries \ref{cor:6:Global} and \ref{cor:6:Local}.
+from Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}.
 Both choices of $\omega$ appear to be too large empirically, as they come from minimizing the regret upper-bound rather than the regret itself, thus
@@
 }
-Furthermore, a promising direction is to directly try to join our contributions from Chapter~\ref{chapter:5} and \ref{chapter:6}, and propose an efficient algorithm using three parts:
+Furthermore, a promising direction is to directly try to join our contributions from Chapters~\ref{chapter:5} and \ref{chapter:6}, and propose an efficient algorithm using three parts:
 \klUCB{} indexes for arm selection,

### 2019-06-12 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Fixed typos in notations of Algorithms RandTopM/MCTopM in Ch5

@@
 
-% % FIXME
 % \TODOL{Est-ce qu'on enlève toute mention de la lower-bound et de "order-optimal" du coup ?!}
-% % FIXME
 
@@
 
-\TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}
+% \TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}
 
@@
           % XXX Algorithm
-              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ \\
-              \For{$t = 0, \dots, T - 1$}{
-                  %
-                  \eIf{$A^j(t) \notin \TopM(t)$}{
-                    \eIf(\tcp*[f]{collision}){$C^j(t)$}{
-                      $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
+              Let $A^j(0) \sim \cU([K])$ and $C^j(0)=\mathrm{False}$ \\
+              \For{$t = 1, \dots, T$}{
+                  %
+                  \eIf{$A^j(t-1) \notin \TopM(t)$}{
+                    \eIf(\tcp*[f]{collision at previous step}){$C^j(t-1)$}{
+                      $A^j(t) \sim \cU \left(\TopM(t)\right)$
                       \tcp*[f]{randomly switch}
                       }(\tcp*[f]{randomly switch on an arm that had smaller UCB}){
-                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
+                        $A^j(t) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                       }
                     }{
-                      $A^j(t+1) = A^j(t)$
+                      $A^j(t) = A^j(t-1)$
                       \tcp*[f]{stays on the same arm}
                     }
-                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
+                  Play arm $A^j(t)$, get new observations (sensing and collision), \\
                   Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
@@
           % XXX Algorithm
-              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
-              \For{$t = 0, \dots, T-1$}{
+              Let $A^j(0) \sim \cU([K])$ and $C^j(0)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
+              \For{$t = 1, \dots, T$}{
                    \uIf(\tcp*[f]{transition $(3)$ or $(5)$}){
-                      $A^j(t) \notin \TopM(t)$}
+                      $A^j(t-1) \notin \TopM(t)$}
                     {
-                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
+                      $A^j(t) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                       \tcp*[f]{not empty} \\
                       % \tcp*[f]{randomly switch on an arm that had smaller UCB}
-                      $s^j(t+1) = \mathrm{False}$
+                      $s^j(t) = \mathrm{False}$
                       \tcp*[f]{aim at an arm with a smaller UCB at $t-1$}
@@
                     \uElseIf(\tcp*[f]{collision and not fixed}){
-                        $C^j(t)$ \emph{and} $\overline{s^j(t)}$}
+                        $C^j(t-1)$ \emph{and} $\overline{s^j(t-1)}$}
                       {
-                        $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
+                        $A^j(t) \sim \cU \left(\TopM(t)\right)$
                         \tcp*[f]{transition $(2)$} \\
-                        $s^j(t+1) = \mathrm{False}$
+                        $s^j(t) = \mathrm{False}$
                     }
                     \Else(\tcp*[f]{transition $(1)$ or $(4)$}){
-                      $A^j(t+1) = A^j(t)$ \tcp*[f]{stay on the previous arm} \\
-                      $s^j(t+1) = \mathrm{True}$ \tcp*[f]{become or stay fixed on a ``chair''}
+                      $A^j(t) = A^j(t-1)$ \tcp*[f]{stay on the previous arm} \\
+                      $s^j(t) = \mathrm{True}$ \tcp*[f]{become or stay fixed on a ``chair''}
                     }
-                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
+                  Play arm $A^j(t)$, get new observations (sensing and collision), \\
                   Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
@@
 if player $j$ did not encounter a collision when using arm $k$ at time $t$,
-then she marks her current arm as a ``chair'' ($s^j(t+1)=\mathrm{True}$),
+then she marks her current arm as a ``chair'' ($s^j(t)=\mathrm{True}$),
 and will keep using it even if collisions happen in the future (Lines~$9$-$11$).
@@
 a new arm is sampled uniformly from a subset of $\TopM(t)$,
-defined with the previous indices $U^j(t-1)$ (Lines~$3$-$5$).
+defined with the previous indices $U^j(t)$ (Lines~$3$-$5$).
 %
@@
 $U_{k'}^j(t-1) \leq U^j_{k}(t-1)$ and $U_{k'}^j(t) \geq U^j_{k}(t)$,
-when switching from $k=A^j(t)$ to $k'=A^j(t+1)$.
+when switching from $k=A^j(t-1)$ to $k'=A^j(t)$.
 This helps to control the number of such changes of arm,
@@
 \begin{figure}[h!]
-  \resizebox{1.05\textwidth}{!}{
+  \resizebox{1.03\textwidth}{!}{
   \begin{tikzpicture}[>=latex',line join=bevel,scale=5.0]
@@
       \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
-      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
-      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
+      \node (notfixed) at (1,0) [draw,rectangle,very thick] {Not fixed, $\overline{s^j(t-1)}$};
+      \node (fixed) at (0,0) [draw,rectangle,very thick] {Fixed, $s^j(t-1)$};
       %
       \draw [black,->] (start) -> (notfixed.20);
-      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
-      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t)$,\\$A^j(t) \in \TopM(t)$} (1);
-      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
-      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
-      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
+      \draw [color=cyan,very thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t-1)}, A^j(t-1) \in \TopM(t)$} (fixed);
+      \path [color=blue,very thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t-1)$,\\$A^j(t-1) \in \TopM(t)$} (1);
+      \path [color=red,very thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t-1) \notin \TopM(t)$} (1);
+      \path [color=darkgreen,very thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t-1) \in \TopM(t)$} (fixed);
+      \draw [color=red,very thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t-1) \notin \TopM(t)$} (notfixed);
       %

### 2019-06-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Proof-read Ch5

@@
 	\item $\mathcal{H}_{2}$:
-	The total number of back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after {\mathrm{MaxBackOf}}$ successive failed retransmissions
+	The total number of back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $\mathrm{MaxBackOff}$ successive failed retransmissions
 	(otherwise, its battery life can be threatened if it does not have re-fueling capabilities).
@@
 In order to assess the proposed approximation, we suppose a unique channel where all the devices follow the same contention Markov process.
-We simulate an ALOHA protocol with a maximum number of retransmissions {\mathrm{MaxBackOf}}=10$, a maximum back-off interval $m=10$, and a transmission probability $p=10^{-3}$.
+We simulate an ALOHA protocol with a maximum number of retransmissions $\mathrm{MaxBackOff}=10$, a maximum back-off interval $m=10$, and a transmission probability $p=10^{-3}$.
 %

### 2019-06-12 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Proof-read Ch5

@@
   A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small.
-Its proof is postponed to the end of this section.
+  % Its proof is postponed to the end of this section.
 
@@
 
-  As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
-  Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
+  As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t-1)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
+  Formally, $s^j(0)$ is false, and $s^j(t)$ is defined inductively from $s^j(t-1)$ as
   % following the algorithm of \MCTopM:
   \begin{equation}
-      s^j(t+1) =
-      \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
-      \cap \left( A^j(t) \in \TopM(t) \right).
+      s^j(t) =
+      \left( s^j(t-1) \cup \left( \overline{s^j(t-1)} \cap \overline{C^j(t-1)} \right) \right)
+      \cap \left( A^j(t) \in \TopM(t-1) \right).
   \end{equation}
@@
   For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words.
-At step $t$, if player $j$ is not fixed ($\overline{s^j(t)}$), she can have three behaviors when executing \MCTopM.
+At step $t$, if player $j$ is not fixed ($\overline{s^j(t-1)}$), she can have three behaviors when executing \MCTopM.
   She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
@@
   with two cases.
-  Either she sampled $A^j(t+1)$ uniformly
-  from $\TopM(t) \cap \{ m : U_m^j(t) \leq U_k^j(t) \}$
+  Either she sampled $A^j(t)$ uniformly
+  from $\TopM(t) \cap \{ m : U_m^j(t-1) \leq U_k^j(t-1) \}$
   with transition $(3)$,
-  in case of collision and if $A^j(t+1) \in \TopM(t)$,
-  or she sampled $A^j(t+1)$ uniformly
+  in case of collision and if $A^j(t-1) \in \TopM(t)$,
+  or she sampled $A^j(t)$ uniformly
   from $\TopM(t)$ with transition $(2)$,
-  if $A^j(t+1) \notin \TopM(t)$.
-  In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
+  if $A^j(t-1) \notin \TopM(t)$.
+  In particular, note that if $\overline{C^j(t-1)}$, transition $(3)$ is executed and not $(2)$.
   %
@@
   For player $j$ and round $t$, we now introduce a few events that are useful in the proof.
-First, for every $x=1,2,3,4,5$, we denote $I_x^j(t)$ the event that a transition of type $(x)$ occurs for player $j$ after the first $t$ observations (\ie, between round $t$ and round $t+1$, to decide $A^j(t+1)$).
+First, for every $x=1,2,3,4,5$, we denote $I_x^j(t)$ the event that a transition of type $(x)$ occurs for player $j$ after the first $t$ observations (\ie, in round $t$, to decide $A^j(t)$).
   Formally they are defined by
   \begin{align*}
-    I_1^j(t) & \eqdef \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
+    I_1^j(t) & \eqdef \left(\overline{s^j(t-1)},\overline{C_j(t-1)}, A^j(t-1) \in \TopM(t)\right), \\
     % go sitted
-    I_2^j(t) & \eqdef \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
+    I_2^j(t) & \eqdef \left(\overline{s^j(t-1)},C_j(t-1), A^j(t-1) \in \TopM(t)\right),
     \ \ \ &\text{and} \ \ \
-    I_3(t) & \eqdef \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
+    I_3(t) & \eqdef \left(\overline{s^j(t-1)},A^j(t-1) \notin \TopM(t)\right), \\
     % if sitted
-    I_3(t) & \eqdef \left(s^j(t),A^j(t) \in \TopM(t)\right),
+    I_3(t) & \eqdef \left(s^j(t-1),A^j(t-1) \in \TopM(t)\right),
     \ \ \ &\text{and} \ \ \
-    I_5(t) & \eqdef \left(s^j(t),A^j(t) \notin \TopM(t)\right).
+    I_5(t) & \eqdef \left(s^j(t-1),A^j(t-1) \notin \TopM(t)\right).
   \end{align*}
@@
 
-  A key observation is that $C^j(t)$ implies $\bigcup_{j'=1}^M \widetilde{C^{j'}}(t)$, as a collision necessarily involves at least one player not yet fixed on her arm ($\overline{s^{j'(t)}}$).
-  Otherwise, if they are all fixed, \ie, for all $j$, $s^j(t)$, then by definition of $s^j(t)$, none of the player changed their arm from $t-1$ to $t$, and none experienced any collision at time $t-1$ so by induction there is no collision at time $t$.
+  A key observation is that $C^j(t)$ implies $\bigcup_{j'=1}^M \widetilde{C^{j'}}(t)$, as a collision necessarily involves at least one player $j'\in[K]$ not yet fixed on her arm ($\overline{s^{j'(t-1)}}$).
+  Otherwise, if they are all fixed, \ie, for all $j'$, $s^{j'}(t-1)$, then by definition of $s^j(t-1)$, none of the player changed their arm from $t-1$ to $t$, and none experienced any collision at time $t-1$ so by induction there is no collision at time $t$.
   %
@@
   Let $N_x^j(T)$ be the random variable denoting the number of transitions of type $(x)$.
-  Neglecting the event $\overline{s^j(t)}$ for $x=3$ and $s^j(t)$ for $x=5$, one has
+  Neglecting the event $\overline{s^j(t-1)}$ for $x=3$ and $s^j(t)$ for $x=5$, one has
   \begin{equation}
@@
       = \sum_{t=1}^T \Pr(I_x^j(t))
-      \leq \sum_{t=1}^T \Pr \left(A^j(t) \notin \TopM(t)\right)
-      \leq \sum_{t=1}^T \sum_{k=1}^K \Pr \left(A^j(t)=k, k \notin \TopM(t)\right),
+      \leq \sum_{t=1}^T \Pr \left(A^j(t-1) \notin \TopM(t)\right)
+      \leq \sum_{t=1}^T \sum_{k=1}^K \Pr \left(A^j(t-1)=k, k \notin \TopM(t)\right),
   \end{equation}
@@
   a simple argument can be used.
-  $\bigcup_{t=\theta_i}^{\tau_i-1} I_2^j(t)$
+  The union of events $\bigcup_{t=\theta_i}^{\tau_i-1} I_2^j(t)$
   implies $C^j(t)$ for $\tau_i - \theta_i$ consecutive times.
@@
   the successive collisions (\ie, $C^j(t-1) \cap C^j(t)$)
-  implies that each new arm $A^j(t+1)$ for $t \in \{\theta_i, \tau_i-1\}$ is selected uniformly from
-  $\TopM(t+1)$,
+  implies that each new arm $A^j(t)$ for $t \in \{\theta_i, \tau_i-1\}$ is selected uniformly from
+  $\TopM(t)$,
   a set of size $M$ with at least one available arm.
   %
-  Indeed, as there is $M-1$ other players, at time $t+1$ \emph{at least} one arm in $\TopM(t+1)$ is not selected by any player $k'\neq k$,
+  Indeed, as there is $M-1$ other players, at time $t$ \emph{at least} one arm in $\TopM(t)$ is not selected by any player $k'\neq k$,
   and so player $j$ has \emph{at least} a probability $1/M$ to select
-  a free arm, which implies $\overline{C^j(t+1)}$, and so implies the end of the sequence.
+  a free arm, which implies $\overline{C^j(t)}$, and so implies the end of the sequence.
   %
@@
   $\E_{\mu}\left[ \tau_i - \theta_i \right]$,
-  is bounded by the expected number of failed trial of a repeated Bernoulli experiment, with probability of success larger than $1/M$ (by the uniform choice of $A^j(t+1)$ in a set of size $M$ with at least one available arm).
+  is bounded by the expected number of failed trial of a repeated Bernoulli experiment, with probability of success larger than $1/M$ (by the uniform choice of $A^j(t)$ in a set of size $M$ with at least one available arm).
   We recognize the mean of a geometric random variable, of parameter $\lambda \geq 1/M$, and so $\E_{\mu}\left[ \tau_i - \theta_i \right] = \frac{1}{\lambda} \leq \frac{1}{1/M} = M$.
@@
 another consequence of our proof is that we have also bounded
-the (expected) number of \emph{switches of arms}, $SC^{\cA}(T) = \sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$.
-%
-We controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
+the (expected) number of \emph{switches of arms}, $\mathrm{SC}^{\cA}(T) = \sum_{t=1}^{T-1} \Pr(A^j(t+1) \neq A^j(t))$.
+%
+We controlled the total number of transitions $(2)$, $(3)$ and $(5)$ (see Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM}),
 which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
@@
 % -----------------------------------------------------------------
-\subsection{Discussion on \Selfish} \label{sub:5:SelfishFails}
-
-The analysis of \Selfish{} is harder, but we obtained some understanding of the behavior of this algorithm, that seems to be doing surprisingly well in many contexts, as in our experiments with $K=9$ arms and in extensive experiments not reported in this section.
+\subsection{Discussion on \Selfish}\label{sub:5:SelfishFails}
+
+The analysis of the \Selfish{} algorithm is harder, but we obtained some understanding of the behavior of this algorithm, that seems to be doing surprisingly well in many contexts, as in our experiments with $K=9$ arms and in extensive experiments not reported in this section.
 However, a disappointing result is that we found simple problems, usually with small number of arms, for which the algorithm may fail.
@@
 
-In the Appendix~E of \cite{Besson2018ALT},
+In the Appendix~E of our article \cite{Besson2018ALT},
 we explained when such situations may happen, and we included a conjectured (constant, but small) lower bound on the probability that \Selfish{} experience collision almost at every round.
@@
       \end{tabular}
-      \caption{No matter the orthogonalization policy, using \klUCB{} is much more efficient than using \UCB{}, for multi-players bandit (here in a simple problem with $K=9$ arms).}
+      \caption{No matter the orthogonalization policy and $M$ the number of players, using \klUCB{} is much more efficient than using \UCB{}, for multi-players bandit (here in a simple problem with $K=9$ arms).}
       \label{table:5:comparisonUCB_klUCB}
@@
 Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
-Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved (the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is $\cO(1)$ if there is no arms in \Mworst{} !), and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous (state-of-the-art) approaches using ranks.
+Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved that the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is indeed $\cO(1)$ if there is no arms in \Mworst{}, and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous state-of-the-art approaches using ranks.
 %
@@
   \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem]{Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
-With a perfect knowledge on the horizon and the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly in this simple setting.
+With a perfect knowledge on the horizon and the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly, even in this simple setting.
 The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can bounded with probability $0.5$ and $0.9$ respectively.
 The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.
-The log-$y$ scale is used to easily differentiate between the different algorithms, and highlight that our proposal outperform both \MEGA{} and \MusicalChair{} by three orders of magnitudes!}
+The log-$y$ scale is used to easily differentiate between the different algorithms, and highlight that our proposal (in \textcolor{gold}{yellow $\nabla$}) outperform both \MEGA{} and \MusicalChair{} by three orders of magnitudes!}
   \label{fig:5:MP__K9_M3_T123456_N100__8_algos}
@@
 %
-\begin{figure}[!b]
+\begin{figure}[!t]
   % \centering
@@
 \MusicalChair{} consists of a random exploration phase of length $T_0$ after which the players quickly converge to orthogonal strategies targeting the $M$ best arms.
-With probability $1-\delta$, its regret is proven to be ``constant'' (of order $\log(1/\delta)$).
+With probability at least $1-\delta$, its regret is proven to be ``constant'' (of order $\log(1/\delta)$).
 The theoretical minimal value for $T_0$ depends on $\delta$, on the horizon $T$ and on a lower bound $\epsilon$ on the gap $\Delta = \mu^*_M - \mu^*_{M+1}$, and the practical tuning is hard too.
@@
 
-\textbf{Uniformly sampled problems.}
+\paragraph{Uniformly sampled problems.}
 %

### 2019-06-12 3-Appendices/2-Appendix/appendix2.tex
Reduced size and spacing{0.93} for ALL "list of *" in Appendix

--- 2019-05-28 3-Appendices/2-Appendix/appendix2.tex
@@
 
+\paragraph{Motivations for anytime algorithms.}
+%
 As introduced in Chapter~\ref{chapter:2},
@@
 On the one hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for instance we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
-On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices: usually a learning step corresponds to an uplink and then downlink message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages.
-While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies), many kinds of applications such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.
+On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices (like we present it in Chapter~\ref{chapter:4}): usually a learning step corresponds to an uplink and then downlink message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages.
+While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies), many kinds of application such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.
 
 For this later range of applications, it is of highest interest to be able to use low-cost MAB algorithms that do not rely on prior knowledge of the problem for which they will be applied, and especially do not need to know the horizon $T$.
-While we proposed in Chapter~\ref{chapter:5} any-time algorithms solving the presented problem of stationary multi-player bandit, our solution for the problem of non-stationary MAB problem studied in Chapter~\ref{chapter:6} does rely on a prior knowledge of the horizon $T$.
+In this thesis,
+while we proposed in Chapter~\ref{chapter:5} any-time algorithms solving the presented problem of stationary multi-players bandit, our solution for the problem of non-stationary MAB problem studied in Chapter~\ref{chapter:6} does rely on a prior knowledge of the horizon $T$.
 It is interesting to know what is the best approach to fix this weakness: work more and design algorithms that are inherently any-time, or use a generic technique to avoid depending on a prior knowledge of $T$, and automatically transform our non-anytime approach to make it anytime.
 
-A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06}.
+\paragraph{Solution to ``patch'' a non-anytime algorithm.}
+%
+A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06} and used for instance in \cite{Auer10,AuerGajaneOrtner18}.
 %
@@
 
-In our article \cite{Besson2018DoublingTricks}, we also present numerical experiments in the case of stationary MAB problems, to compare the performance of efficient anytime algorithms, like \klUCB, against efficient non-anytime algorithms, like \KLUCBpp, made anytime with different doubling-trick.
+In our article \cite{Besson2018DoublingTricks}, we also present numerical experiments in the case of stationary MAB problems, to compare the performance of efficient anytime algorithms, like \klUCB, against efficient non-anytime algorithms, like \KLUCBpp{} from \cite{Menard17}, made anytime with different choices of doubling-trick.
 We conclude that for such problems, if $T$ is not known before, it is always more efficient to use policies that were designed to be anytime that to use a doubling trick.
-For example, an applicative paper written in 2018, \cite{li2019useDoublingTrick}, only tested the use of an exponential doubling trick combined with \KLUCBpp, but most surely the \klUCB{} algorithm would have given better empirical performance as well as more robust results.
+For example, an applicative paper written in 2018, \cite{li2019useDoublingTrick}, only tested the use of an exponential doubling trick combined with \KLUCBpp, but most surely the \klUCB{} algorithm would have given better empirical performance as well as more robust results\ldots
 

### 2019-06-12 1-Preamble/5-Abstract/abstract.tex
Changed to a lot of small things for 1st/last page

@@
 \begin{abstract}
-\vspace*{-30pt}
+% \vspace*{-30pt}
+
+% ----------------------------------------------------------------------------
+\begin{center}
+	\textbf{{\Large Résumé}}
+\end{center}
+
 %
@@
 % ----------------------------------------------------------------------------
-\begin{small}
+% \begin{small} % XXX
 % Chapitre 1
@@
 % Chapitre 3
-Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres.
+Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre les modèles étudiés et d'autres.
 % (Chapitre~\ref{chapter:3})
-\end{small}
+% \end{small} % XXX
 
@@
 %
-\begin{small}
+% \begin{small} % XXX
 % Chapter 1
@@
 %  (Chapter~\ref{chapter:3})
-\end{small}
+% \end{small} % XXX
 

### 2019-06-12 2-Chapters/Conclusion/conclusion.tex
Reduced size and spacing{0.93} for ALL "list of *" in Appendix
Changed to a lot of small things for 1st/last page

@@
 % ----------------------------------------------------------------------------
+\newpage
+% FIXME
+
+% ----------------------------------------------------------------------------
 % ----------------------------------------------------------------------------
@@
 A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
-It is already possible to easily reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if one installs or downloads the library, but can also be used on cloud platforms, such as
+It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if one installs or downloads the library, but can also be used on cloud platforms, such as
 Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} for free.},

### 2019-06-12 2-Chapters/1-Chapter/chapter1.tex
\vfill{} before \minitoc at beginning of all Chapters

@@
 The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but all use the notations introduced on Chapter~\ref{chapter:2}.
+
+\vfill{}  % WARNING vfill before minitoc?
 
@@
     by Christophe Moy \& \textbf{Lilian Besson}.
-    1st International ISIoT workshop\footnote{~See \href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}},
-    at \emph{Conference on Distributed Computing in Sensor Systems}\footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}}, Santorini, Greece, May 2019.
-    \href{https://HAL.Inria.fr/hal-02144465}{\texttt{HAL.Inria.fr/hal-02144465}}.
+    1st International ISIoT workshop,
+    % \footnote{~See \href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}},
+    at \emph{Conference on Distributed Computing in Sensor Systems},
+    % \footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}},
+    Santorini, Greece, May 2019.
+    % \href{https://HAL.Inria.fr/hal-02144465}{\texttt{HAL.Inria.fr/hal-02144465}}.
     \cite{MoyBesson2019}
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
-    1st International MOTIoN workshop\footnote{~MOTIoN 2019, see \href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
+    1st International MOTIoN workshop,
+    % \footnote{~MOTIoN 2019, see \href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
     at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.
-    \href{https://HAL.Inria.fr/hal-02049824}{\texttt{HAL.Inria.fr/hal-02049824}}.
+    % \href{https://HAL.Inria.fr/hal-02049824}{\texttt{HAL.Inria.fr/hal-02049824}}.
     \cite{Bonnefoi2019WCNC}
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}}, Marrakech, Morocco, April 2019,
-    \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
+    \emph{Wireless Communication and Networks Conference},
+    % \footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}},
+    Marrakech, Morocco, April 2019,
+    % \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
     \cite{Besson2019WCNC}
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann.
-    \emph{Algorithmic Learning Theory}\footnote{~ALT 2018, see \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}}, Lanzarote, Spain, April 2018,
-    \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
+    \emph{Algorithmic Learning Theory},
+    % \footnote{~ALT 2018, see \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}},
+    Lanzarote, Spain, April 2018,
+    % \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
     \cite{Besson2018ALT}
@@
     by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
-    \emph{Wireless Communication and Networks Conference}\footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}}, Barcelone, Spain, April 2018,
-    \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
+    \emph{Wireless Communication and Networks Conference},
+    % \footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}},
+    Barcelona, Spain, April 2018,
+    % \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
     \cite{Besson2018WCNC}
@@
     by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
-    \emph{Conference on Cognitive Radio Oriented Wireless Networks}\footnote{~CROWNCOM 2017, see \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}}, Lisboa, Portugal, Septembre $2017$,
-    \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
+    \emph{Conference on Cognitive Radio Oriented Wireless Networks},
+    % \footnote{~CROWNCOM 2017, see \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}},
+    Lisboa, Portugal, Septembre $2017$,
+    % \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
     \textbf{Best Paper Award}.
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
-    Demonstration presented at \emph{International Conference on Communication}\footnote{~ICT 2018, see \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}}, in Saint-Malo, France in June $2018$.
+    Demonstration presented at \emph{International Conference on Communication},
+    % \footnote{~ICT 2018, see \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}},
+    Saint-Malo, France in June $2018$.
     See \href{https://YouTu.be/HospLNQhcMk}{\texttt{YouTu.be/HospLNQhcMk}} for a $6$-minutes presentation video.
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    GRETSI 2019\footnote{~GRETSI 2019, see \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
+    GRETSI 2019,
+    % \footnote{~GRETSI 2019, see \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}}},
     August $2019$,
-    \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
+    % \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
     \cite{Besson2019Gretsi}
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    February $2019$,
-    \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
+    February $2019$,\\
+    Preprint at \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
     \cite{Besson2019GLRT}
@@
     by \textbf{Lilian Besson}, active development since October $2016$,
-    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.\\
+    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
     About $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
@@
     September $2018$,
-    \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.
+    Preprint at \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.
     \cite{Besson2018DoublingTricks}

### 2019-06-12 2-Chapters/1-Chapter/chapter1.tex
Done

@@
 
-% FIXME
+% WARNING
 \newpage
@@
 
+We conclude this chapter with a list of publications published during this PhD.
+All the publications are published on HAL (see the \href{https://hal.archives-ouvertes.fr/}{\texttt{HAL.Archives-Ouvertes.fr}} website),
+and the complete list can be found on
+\href{https://cv.archives-ouvertes.fr/lilian-besson/}{\texttt{CV.Archives-Ouvertes.fr/lilian-besson}}.
+
 % =============================================================================
-\paragraph{Publications in international conferences with proceedings}
+\subsection*{Publications in international conferences with proceedings}
 
@@
 % =============================================================================
-\paragraph{Demonstration in international conferences}
+\subsection*{Demonstration in international conferences}
 
@@
 % =============================================================================
-\paragraph{French language conference with proceedings}
+\subsection*{French language conference with proceedings}
 % \textbf{Submitted works}
@@
 % =============================================================================
-\paragraph{In progress works waiting for a new submission}
+\subsection*{In progress works waiting for a new submission}
 

### 2019-06-12 2-Chapters/2-Chapter/chapter2.tex
\vfill{} before \minitoc at beginning of all Chapters
Done

@@
 % \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
+
+\vfill{}  % WARNING vfill before minitoc?
 
@@
 %
-It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $SC^{\cA_1,\dots,\cA_M}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
+It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
 % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
@@
 For more details, see \cite{modiDemo2016} for experiments on the multi-players case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
-While such ideas can interest a practitioner, they does not change the asymptotic behavior of $SC^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
+While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
 
@@
 \begin{align}\label{eq:klUCB_index}
-    U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \KL\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
+    U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
 \end{align}
@@
     U^{\mathrm{MOSS}}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)} } \right). \\
+% \end{align}
+% \begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp2}
     U^{\KLUCBpp}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
@@
 then \klUCB-switch use the index defined as follows:
-%
-\begin{align}\label{eq:2:indexes_for_klUCBswitch}
-    U^{\mathrm{MOSS}+}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
-    U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
+\[
     U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
@@
     \end{cases}.
-\end{align}
+\]
+
+If the two indexes $U^{\KLUCB+}_k(t)$ and $U^{\mathrm{MOSS}+}_k(t)$ are defined by
+
+% \vspace*{-20pt}   % FIXME??
+\begin{align*}
+    U^{\mathrm{MOSS}+}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
+% \end{align*}
+% \begin{align*}
+    U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
+% \end{align*}
+% \begin{align*}
+    % U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
+    %     U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
+    %     U^{\mathrm{MOSS}+}_k(t) & \text{if } N_k(t) > f(T, K).
+    % \end{cases}.
+\end{align*}
 

### 2019-06-13 1-Preamble/5-Abstract/abstract.tex
Small improvements on layout of abstract/minitoc for chapters

@@
 % Chapter 3
-We also detail another contribution, the SMPyBandits open-source Python library for numerical simulations of MAB problems, which covers all the studied models and more.
+We also detail another contribution, SMPyBandits, an open-source Python library for numerical MAB simulations, covering the studied models and more.
 %  (Chapter~\ref{chapter:3})

### 2019-06-13 2-Chapters/3-Chapter/chapter3.tex
\vfill{} before \minitoc at beginning of all Chapters
Small improvements on layout of abstract/minitoc for chapters

@@
 
-\paragraph{Abstract.}
+\abstractStartChapter{}
 %
@@
 
-% \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
-
-\minitoc
-
-\newpage
+\minitocStartChapter{}
 
@@
 %
-For instance the following code in Listing~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate sthe problem.
+For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate sthe problem.
 We give another example for truncated Gaussian distributions, and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
@@
 
-Simulations are easily executed, \eg, Listing~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
+Simulations are easily executed, \eg, Code~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
 It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
@@
 
-Below is presented in Listing~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
+Below is presented in Code~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
 Only the mean reward is printed at the end of the simulation.
@@
 
-We present below in Listing~\ref{lst:3:main} a full example of a minimalist \texttt{main.py} file,
+We present below in Code~\ref{lst:3:main} a full example of a minimalist \texttt{main.py} file,
 which is used to load the configuration, run the experiment and then print and plot various visualizations for the simulation.
@@
 
-We present below in Listing~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
+We present below in Code~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
 which is used to define the experiment.
@@
     (\texttt{klUCBswitchAnytime}),
-    is a very recent variant of the \klUCB{} algorithm \cite{KLUCBJournal}.
-    It mimics the \klUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
+    is a very recent variant of the \KLUCB{} algorithm \cite{KLUCBJournal}.
+    It mimics the \KLUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
     It is anytime, and was prove to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
@@
 \paragraph{Bash command to run these experiments.}
-We give below in Listing~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
+We give below in Code~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
 It uses \texttt{for} loops and environment variables from Bash, and not Python, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $N$, $T$ and $K$.

### 2019-06-13 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Small improvements on layout of abstract/minitoc for chapters

@@
 
-% -----------------------------------------------------------------
-\section{Introduction}
-\label{sec:5:introduction}
-% -----------------------------------------------------------------
-
-For the development of Cognitive Radio, a crucial step is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$), as we explained in Chapters~\ref{chapter:1} and \ref{chapter:4}.
-%
-In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
+% % -----------------------------------------------------------------
+% \section{Introduction}
+% \label{sec:5:introduction}
+% % -----------------------------------------------------------------
+
+% -----------------------------------------------------------------
+\section{Motivations for multi-players MAB models}
+\label{sec:5:motivation}
+% -----------------------------------------------------------------
+
+As we saw in Chapters~\ref{chapter:1} and \ref{chapter:4},
+a crucial step for the development of Cognitive Radio is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
+%
+In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
 Such networks can be modelled using a decentralized multi-players multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
@@
 
-We start by reviewing previous works on multi-players MAB models, which all considered the easier case of \emph{sensing feedback}.
-%
-For this first model, we present two algorithms, \RandTopM{} and \MCTopM, combining an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
-Like in previous works, we consider the centralized system regret (multi-players regret).
-We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
-We then analyze the \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
-%  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
-We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M \leq K$ accessing $K$ channels.
-
-We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
-Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
-Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, we conjectured that it suffers from linear regret, and it was later confirmed by two works inspired by our article \cite{LugosiMehrabian18,BoursierPerchet18}.
-They both proposed new algorithms, proved to achieve logarithmic regret under the no sensing model.
-We include numerical simulations to compare their proposals, and we present in details the current state-of-the-art of research on multi-players MAB models without sensing.
-
-Finally, we review other extensions of the models, and give a overview of the current state-of-the-art.
-For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
-
-\textbf{Outline.}
-%
-This chapter is organized as follows.
-First we motivate the model of multi-players multi-armed bandits in Section~\ref{sec:5:motivation}
-We introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}, and give a new regret lower bound in Section~\ref{sec:5:lowerbound}.
-The \RandTopM, \MCTopM{} and \Selfish{} algorithms are introduced in Section~\ref{sec:5:algorithms},
-for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
-The result of our experimental study reported in Section~\ref{sec:5:experiments}.
-Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
-% and finally Section~\ref{sec:5:conclusion} concludes.
-
-\textbf{Publication.}
-%
-This chapter is mainly based on our article \cite{Besson2018ALT}.
-
-% -----------------------------------------------------------------
-\section{Motivations for multi-players MAB models}
-\label{sec:5:motivation}
-% -----------------------------------------------------------------
-
-% % general intro, UCB
-
-% Several sequential decision making problems under the constraint of partial information
-% have been studied since the 1950s under the name of Multi-Armed Bandit (MAB) problems \citep{Robbins52,LaiRobbins85}.
-% In a stochastic MAB model, an agent is facing $K$ unknown probability distributions, called arms in reference to the arms of a one-armed bandit (or slot machine) in a casino.
-% Each time she selects (or draws) an arm, she receives a reward drawn from the associated distribution.
-% Her goal is to build a sequential selection strategy that maximizes the total reward received.
-% A class of algorithms to solve this problem is based on Upper Confidence Bounds (UCB), first proposed by \cite{LaiRobbins85,Agrawal95} and further popularized by \cite{Auer02}.
-% The field has been very active since then, with several algorithms proposed and analyzed, both theoretically and empirically, even beyond the stochastic assumption on arms, as explained in the survey by \cite{Bubeck12}.
-
-% % appli, from clinical trials to cognitive radios
-
-% The initial motivation to study MAB problems arose from clinical trials (the first MAB model can be traced back to $1933$, by \citeauthor{Thompson33}), in which a doctor sequentially allocates treatments (arms) to patients and observes their efficacy (reward).
-% More recently, applications of MAB have shifted towards sequential content recommendation, \eg{} sequential display of advertising to customers or A/B testing \citep{Li10,Chapelleetal14Ad}.
-% %
-% In the mean time, MAB were found to be relevant to the field of Cognitive Radio (CR, \cite{Mitola99}),
-% and \cite{Jouini09,Jouini10} first proposed to use \UCB{} for the Opportunistic Spectrum Access (OSA) problem,
-% and successfully conducted experiments on real radio networks demonstrating its usefulness.
-% For CR applications, each arm models the quality or availability of a radio channel (a frequency band) in which there is some background traffic (\eg, primary users paying to have a guaranteed access to the channel in the case of OSA).
-% A smart radio device needs to insert itself in the background traffic, by sequentially choosing a channel to access and try to communicate on, seeking to optimize the quality of its global transmissions.
-
-% the need for multiple players in cognitive radio
-
 We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
@@
 %
-In our model in Chapter~\ref{chapter:5}, each device will use a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
+We start by reviewing previous works on multi-players MAB models, which all considered the easier case of \emph{sensing feedback}.
+But in our model in this chapter, each device also uses a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
 %
@@
 Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can have a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-players MAB problems).
-We do not propose any other efficient algorithm, but our work presented in April~$2018$ has strongly inspired two articles shorlty after \cite{LugosiMehrabian18,BoursierPerchet18}.
+We do not propose any other efficient algorithm, but our work presented in April~$2018$ has strongly inspired two articles shortly after \cite{LugosiMehrabian18,BoursierPerchet18}.
 They confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
 We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-players MAB models without sensing.
+
+% In this Chapter~\ref{chapter:5}, we start by reviewing previous works on multi-players MAB models, which all considered the easier case of \emph{sensing feedback}.
+% %
+% For this first model, we present two algorithms, \RandTopM{} and \MCTopM, combining an efficient MAB index policy (we chose \klUCB) and a smart orthogonalization procedure, based on a random hoping procedure called Musical Chair.
+% Like in previous works, we consider the centralized system regret (multi-players regret).
+% We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
+% We then analyze the \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
+% %  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
+% We also present extensive numerical simulations that show that our proposal outperforms all previous solutions and is much more efficient in this easier model of sensing feedback with a fixed and known number of players $M \leq K$ accessing $K$ channels.
+
+% We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
+% Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
+% Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, we conjectured that it suffers from linear regret, and it was later confirmed by two works inspired by our article \cite{LugosiMehrabian18,BoursierPerchet18}.
+% They both proposed new algorithms, proved to achieve logarithmic regret under the no sensing model.
+% We include numerical simulations to compare their proposals, and we present in details the current state-of-the-art of research on multi-players MAB models without sensing.
+
+% Finally, we review other extensions of the models, and give a overview of the current state-of-the-art.
+% For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
 
@@
 % This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.
+
+\textbf{Outline.}
+%
+The rest of this chapter is organized as follows.
+% First we motivate the model of multi-players multi-armed bandits in Section~\ref{sec:5:motivation}
+We introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}.
+We present a useful regret decomposition in Section~\ref{sec:5:lowerbound}, and illustrate it in the Appendix~\ref{sec:5:appendix}.
+The \Selfish, \RandTopM{} and \MCTopM{} algorithms are introduced in Section~\ref{sec:5:algorithms},
+for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
+The result of our experimental study reported in Section~\ref{sec:5:experiments}.
+Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
+For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
+% and finally Section~\ref{sec:5:conclusion} concludes.
+
+\textbf{Publication.}
+%
+This chapter is mainly based on our article \cite{Besson2018ALT}.
 

### 2019-06-13 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Small improvements on layout of abstract/minitoc for chapters

@@
 % -----------------------------------------------------------------
-\section{Motivation for non stationary MAB models}
+\section{Motivations for non stationary MAB models}
 \label{sec:6:Introduction}
@@
 The unified regret analysis is given in Section~\ref{sec:6:RegretAnalysis}, we prove one result in Section~\ref{sec:6:proofRegret}.
-We discuss numerical experiments in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.
+We discuss numerical experiments in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix~\ref{sec:6:appendix}.
 

### 2019-06-13 2-Chapters/Conclusion/conclusion.tex
Small improvements on layout of abstract/minitoc for chapters

@@
 % \newpage
-\paragraph{Abstract.}
+\abstractStartChapter{}
 %
@@
 
-% \vspace*{50pt}
-
-\minitoc
-
-\newpage
+\minitocStartChapter{}
 
@@
 
-% \begin{itemize}
-%     \item
+\begin{itemize}
+    \item
 The first half, Part~\ref{part:Introduction}, starts by motivating our work and situating the context of this research.
@@
 
-    % \item
+    \item
 The second half of this thesis, Part~\ref{part:MABIOT}, presents our main contributions.
@@
 For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
-% \end{itemize}
+\end{itemize}
 
@@
 A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
-It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if one installs or downloads the library, but can also be used on cloud platforms, such as
+It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if you install the library, but can also be used on cloud platforms, such as
 Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} for free.},

### 2019-06-13 2-Chapters/1-Chapter/chapter1.tex
Small improvements on layout of abstract/minitoc for chapters
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1

@@
 
-\paragraph{Abstract.}
+\abstractStartChapter{}%
 %
@@
 %
-The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but all use the notations introduced on Chapter~\ref{chapter:2}.
+The organization of this manuscript is detailed in a flow graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but all use the notations introduced on Chapter~\ref{chapter:2}.
 
-\vfill{}  % WARNING vfill before minitoc?
+\minitocStartChapter{}
 
-\minitoc
-
-\newpage
 % Write miniTOC just after the title
@@
 \begin{itemize}
-    \item
-    We present the concepts and the notations of the multi-armed bandit problem in Chapter~\ref{chapter:2}, from a mathematical point-of-view.
-    But we also follow a didactic approach as we use an online interactive demonstration designed to let anyone play against a small bandit problem from his/her browser, in Section~\ref{par:2:interactiveDemoDiscoverMAB}.
+    % \item
+    % We present the concepts and the notations of the multi-armed bandit problem in Chapter~\ref{chapter:2}, from a mathematical point-of-view.
+    % But we also follow a didactic approach as we use an online interactive demonstration designed to let anyone play against a small bandit problem from his/her browser, in Section~\ref{par:2:interactiveDemoDiscoverMAB}.
+
+    % \item
+    % We give a short literature review of stochastic bandit algorithms in Section~\ref{sec:2:famousMABalgorithms}.
 
     \item
-    We give a short literature review of stochastic bandit algorithms in Section~\ref{sec:2:famousMABalgorithms}.
-
-    \item
-    We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+    % We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
     We present an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
@@
 
-    \item
-    The source code for the two previously mentioned contributions are all published online, along with clear instructions for reproducing our work.
+    % \item
+    % The source code for the two previously mentioned contributions are all published online, along with clear instructions for reproducing our work.
 
@@
     \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
-    About $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
-    and documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}}.
+    It currently consists in about $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
+    and a complete documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} and \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.
     \cite{SMPyBandits,SMPyBanditsJMLR}
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    September $2018$,
+    September $2018$,\\
     Preprint at \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.

### 2019-06-13 2-Chapters/2-Chapter/chapter2.tex
Small improvements on layout of abstract/minitoc for chapters
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1

@@
 
-\paragraph{Abstract.}
+\TODOL{Intégrer commentaires d'Emilie !}
+\TODOL{Découper en deux chapitres, un pour le modèle et un pour Aggregator ? Si je découpe, il faut changer le dessin tikz et d'autres trucs}
+
+\abstractStartChapter{}%
 %
@@
 
-% \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
-
-\vfill{}  % WARNING vfill before minitoc?
-
-\minitoc
+\minitocStartChapter{}
+
 % ----------------------------------------------------------------------
-
-\newpage
 
@@
 % , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
-and a summary is included in the Nomenclature, in Page~\pageref{chapter:nomenclature}.
+and a summary is included in the Nomenclature, included in the Appendix in Part~\ref{part:Conclusion} (Page~\pageref{chapter:nomenclature}).
 

### 2019-06-13 2-Chapters/4-Chapter/chapter4.tex
Typos, et commentaires Christophe
\vfill{} before \minitoc at beginning of all Chapters
Small improvements on layout of abstract/minitoc for chapters
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1

@@
 
-% \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
-
-\minitoc
-% ----------------------------------------------------------------------
-
-\newpage
-
-% \paragraph{Abstract.}
-
-% In this Chapter~\ref{chapter:4}, we present two models of Internet of Things networks, composed of many independent devices that are able to run MAB algorithms to learn and improve their access to the network.
-% The two models are presented mathematically, and numerical simulations justify the interest of the proposed algorithms.
-% We also present a proof-of-concept real-world validation of the advantage of using Multi-Armed Bandit algorithms for the first model, by showing and explaining a demonstration implemented on USRP platforms \cite{USRPDocumentation}.
-
-% More precisely, we start by motivating the idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
-% We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
-% Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
-% We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every instant.
-If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
-% we show in simulations that in the case of a non-uniform repartition of the static devices in the channels, the dynamic devices can drastically improve the network efficiency.
-% %
-% On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
-
-% Then we present in Section~\ref{sec:4:retransmissions} a second model, extending our first model, in which devices can retransmit their messages for a fixed number of maximum retransmission, like in the ALOHA protocol.
-% In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
-% Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
-
-% On the one hand, the first model, without retransmission of packets, is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
-% On the other hand, the second model is more interesting if the main target is an improvement of the QoS of the IoT networks rather than a longer battery life.
-
-% \vfill{}
-
-% \newpage
+\minitocStartChapter{}
 
@@
 
-On the one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
-On the other hand, it is quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
-Sadly, analyzing our model with thousands of independent devices communicating and learning in their own (random) time scale turned out to be of extreme difficulty.
+A satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
+It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
+Sadly, analyzing our model with thousands of independent devices communicating and learning in their own (random) time scales turned out to be of extreme difficulty.
+%
+We can also highlight that the two models we studied are complementary:
+the first model (without retransmission of packets), is interesting for its simplicity, and focused on improving the battery life of IoT devices, at the cost of a lower Quality of Service (QoS),
+%
+while the second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
 
-\paragraph{Multi-Players MAB with \emph{collision avoidance}.}
-
-We are actually able to analyze a simpler model, if we assume to have at most $K$ devices with a transmission probability of $p=1$.
-Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-players MAB} model to describe our problem, as proposed by \cite{Zhao10}.
-
+\textbf{Multi-Players MAB.}
+%
+On the first hand, we are actually able to analyze a simpler model, if we assume to have at most $M \leq K$ devices with a transmission probability of $p=1$.
+Instead of the experiment-driven direction pursued in this chapter, another possibility is to consider a \emph{multi-players MAB} model to describe our problem.
+%
 The main difference between the two models is the fact that in this chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
 If static and dynamic devices that have to transmit at a fixed time are denoted \emph{active devices},
-and their random activation patter makes the number of active devices a random variable.
+then their random activation patter makes the number of active devices an (unpredictable) random variable.
 Analyzing multi-players MAB models under this hypothesis is much harder, and is left as a future work.
 %
-We study this other direction in the next Chapter~\ref{chapter:5}.
+We study this first direction in the next Chapter~\ref{chapter:5}.
 
@@
 
+\textbf{Non-stationary MAB.}
+%
+On the other hand, while it is hard to analyze the models of this chapter due to their non-stationary behaviors, we are also able to analyze a simpler model, if we focus on a single player accessing a network which is assumed to be \emph{piece-wise stationary}
+(MAB problems which are stationary on ``long enough'' intervals, of unknown locations and lengths).
+%
+We also study this second direction in Chapter~\ref{chapter:6}.
+
 \newpage
@@
 
-\subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
+\subsection{Proof of Proposition~\ref{prop:41:Lagrangian}}
 \label{sec:4:proofLagrangian}

### 2019-06-13 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1
Better links to
- https://SMPyBandits.GitHub.io/MultiPlayers.html
- https://SMPyBandits.GitHub.io/NonStationaryBandits.html

@@
 
-\paragraph{A note on the simulation code.}
-%
-All the experiments used for this chapter were performed using SMPyBandits,
+% \paragraph{A note on the simulation code.}
+\paragraph{Reproducility of the experiments.}
+%
+The experiments in this chapter use our library SMPyBandits, as presented in Chapter~\ref{chapter:3},
 and we refer to the following page for details on how to reproduce them,

### 2019-06-13 2-Chapters/1-Chapter/chapter1.tex
Rien de neuf

@@
 \label{chapter:1}
+
+\TODOL{Ecrire les 1.1 et 1.2 en quelques paragraphes}
 
@@
 \TODOL{FIXME Finir ce blabla rapidement ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
+
+We already explained in Chapter~\ref{chapter:1} that
+unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
+This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.
+
+Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
+As end-devices battery life is a key constraint of IoT networks,
+and as IoT networks are decentralized (\ie, the devices initiate transmissions),
+this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
+%
+In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
+Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
+We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
+Kaufmann12Thompson}, for the first model.
+For the demonstration as well as for the second model, without loss of generality, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.
+
+As detailed in Chapter~\ref{chapter:1},
+MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
+For example,
+proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
+and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
+Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
+Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
+% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
+
+We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed but frequency- and time-slotted IoT network.
+Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
+As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly not stationary''.
+This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms (that is why we relax the hypothesis and only consider up-to $M \leq K$ players in Chapter~\ref{chapter:5}).
+We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
+%
+We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
+
+As explained before in Chapter~\ref{chapter:1}, the future IoT networks will require to support more and more communicating devices.
+In this section, we show that intelligent devices in unlicensed bands can use MAB learning algorithms to improve spectral resource exploitation.
+%
+We evaluate the performance of two classical MAB learning algorithms, \UCB{} and Thomson Sampling, to handle the decentralized decision-making of Spectrum Access, applied to IoT networks.
+We mean by \emph{decentralized} that learning is performed on the device side, and is spread on the totality of the devices in a network.
+We also evaluate the learning performance when the number of intelligent end-devices grows.
+%
+We show that using learning algorithms
+firstly extends devices battery autonomy (by reducing the packet-loss-ratio or equivalently, by improving the successful-communication rate),
+and secondly does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
+In the studied settings,
+Stochastic MAB learning
+% provides an up to $16\%$ gain in terms of successful transmission probabilities, and
+has near optimal performance even in non-stationary settings with a majority of intelligent devices.
+
+% \TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}
+
+The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
+and the network usage is more and more fluctuatating.
+% and the stochastic hypothesis is more and more violated.
+To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
+Static devices form an interfering traffic, which could be generated by devices using other standards as well.
+Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel.
+So our first hypothesis is actually not constraining.
+We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
+This second approach is ideal, but not realistic the most common situation of decentralized co-located IoT networks, and it is just used here as a reference.
+Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
+We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.
 

### 2019-06-13 2-Chapters/2-Chapter/chapter2.tex
Commencer à intégrer les commentaires d'Emilie. Chaud

@@
 The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T r(t)] = \sum_{t=1}^T \E[Y_{A(t),t}]$.
-The expectancy $\E[\bullet]$ is taken on the rewards $(Y_{k,t})_{k,t}$, as well as on the external random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
+The expectation $\E[\bullet]$ is taken on the rewards $(Y_{k,t})_{k,t}$, as well as on the external random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
 %
@@
 % ----------------------------------------------------------------------------
-\subsection{Hypotheses on rewards distributions}
+\subsection{Common assumptions on the reward distributions}
 
@@
 
-\paragraph{We restrict to one-dimensional exponential family.}
+% \paragraph{We restrict to one-dimensional exponential family.}
 %
@@
 
-In all this thesis we restrict to distributions that have a certain form, and we focus either on real-valued distributions lying in an exponential families or bounded distributions.
+In the literature, parametric assumptions on the rewards are sometimes considered, typically the assumption that rewards belong to a real-valued distribution lying in an exponential families.
+%  or bounded distributions.
 One-dimensional exponential families include Bernoulli and Poisson distributions, as well as Gaussian distributions with a fixed variance.
@@
     and it is also called the \emph{relative binary entropy}.
-    It satisfies $\kl(x,y) \geq d_{1/4}(x,y)$, see Lemma~\ref{lem:2:Pinsker} in Appendix~\ref{proof:2:Pinsker}.
+    It satisfies $\kl(x,y) \geq d_{1/4}(x,y)$ (Pinsker's inadequality).
 
@@
 %
-As the rewards from arm $k^*$ are \iid{} (as for all arms), and by linearity of the expectancy, we can rewrite this expression to obtain the following definition of the regret:
+As the rewards from arm $k^*$ are \iid{} (as for all arms), and by linearity of the expectation, we can rewrite this expression to obtain the following definition of the regret:
 \begin{align*}
@@
 Note that for the rest of the manuscript, we use the \klUCB{} algorithm (and more subtle concentration inequality) whenever we analyze the regret of a newly introduced algorithm (\ie, in Chapters~\ref{chapter:5} and \ref{chapter:6}),
-but we preferred to start with a simpler proof in the single-player stationary case.
-%
-This theorem shows that $\UCB_1$ achieves an order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta \eqdef \min\limits_{k: \Delta_k>0} \Delta$).
+but we think the simpler proof of $\UCB_1$ is enlightening for the reader.
+% but we preferred to start with a simpler proof in the single-player stationary case.
+%
+This theorem shows that $\UCB_1$ achieves an order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta \eqdef \min\limits_{k: \Delta_k>0} \Delta$), and the result also has the advantage of being non asymptotic.
 
@@
 
-    \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ? si c'est bien fait, ça peut aider la lecture, par exemple \textcolor{darkgreen}{$k$} et \textcolor{gold}{$k^*$}.}
+    % \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ? si c'est bien fait, ça peut aider la lecture, par exemple \textcolor{darkgreen}{$k$} et \textcolor{gold}{$k^*$}.}
 
@@
 
-Therefore, we see that the $\UCB_1$ algorithm has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO(K\log(T)/\Delta^2)$.
+We see that the $\UCB_1$ algorithm has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO((\sum_{k: \Delta_k>0} \frac{1}{\Delta_k})\log(T))$.
 %
@@
 it is an index policy (see Algorithm~\ref{algo:2:indexPolicy}),
-which also builds upper confidence bounds on the empirical samples.
-The KL-\UCB{} algorithm was first proposed in \cite{Garivier11KL} for bounded rewards,
-and then the \klUCB{} algorithm was introduced in \cite{KLUCBJournal}.
+which also builds upper confidence bounds on the unknown mean of each arm.
+The KL-\UCB{} algorithm was first proposed in \cite{Garivier11KL} for one-dimensional exponential families,
+and then the \klUCB{} algorithm was introduced in \cite{KLUCBJournal} for bounded rewards.
 In all this thesis, we either restrict to exponential families or to bounded rewards in $[0,1]$ and use the fact that they are sub-Bernoulli, thus we prefer to only present \klUCB.
@@
 Instead of using a confidence radius $\xi_k(t)$ based on Hoeffding's inequality,
-the \klUCB{} algorithm rather use a specific concentration inequality for the considered exponential family, such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
-It yields improved upper confidence bounds (\ie, with smaller $\xi_k(t)$) that are based on Kullback-Leibler divergence.
+the \klUCB{} algorithm rather use the Chernoff concentration inequality for the considered exponential family
+% such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
+% It yields improved upper confidence bounds (\ie, with smaller $\xi_k(t)$) that are based on Kullback-Leibler divergence.
 The indexes are computed as follows, for an exploration function $f(t)$
-that is typically chosen as $f(t) \eqdef \log(t)+3\log(\log(t))$ for the theoretical analyses,
+that is typically chosen as $f(t) \eqdef \log(t)+c\log(\log(t))$ for the theoretical analyses (with $c \geq 3$),
 and $f(t) \eqdef \log(t)$ in practice.
@@
 \begin{align}\label{eq:klUCB_index}
-    U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{c f(t}{N_k(t)} \right\}.
+    U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{f(t}{N_k(t)} \right\}.
 \end{align}
@@
 For Bernoulli distributions, the previous claim that \klUCB{} builds smaller confidence intervals (\ie, more accurate) than $\UCB_1$, \ie, $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$,
-is justified by Pinkser's inequality which gives $\kl(x,y) \geq 2(p-q)^2$ (see Lemma~\ref{lem:2:Pinsker} proved in appendix).
+is justified by Pinkser's inequality which gives $\kl(x,y) \geq 2(p-q)^2$.
 %
@@
 Implementing this kind of policy requires to implement two things.
-The first piece is usually easy: one has to implement a way to compute $\KL(x,y)$ the Kullback-Leibler divergence of the considered one-dimensional family,
+The first piece is usually easy: one has to implement a way to compute $\kl(x,y)$ the Kullback-Leibler divergence of the considered one-dimensional family,
 for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
 The second piece is less obvious: one has to solve the optimisation problem defining the $\sup$ in \eqref{eq:klUCB_index}.
-Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\KL(x,\bullet)$ function is usually convex (for any $x$), one practical and mathematically correct solution is to use Newton's method.
-The search space for $q$ can be restricted to $[ X_k(t) / N_k(t), 1 ]$ instead of $[0,1]$, and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
-For instance, $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically Newton's method terminates in between $8$ to $12$ steps.
+Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\kl(x,\bullet)$ function is usually convex (for any $x$), two practical and mathematically correct solutions are Newton's method or a bisection search.
+The search space for $q$ can be restricted to $[ X_k(t) / N_k(t), 1 ]$ instead of $[0,1]$,
+and for simplicity and efficiency, in this case a naive bisection search works well.
+% and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
+For instance, $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically the bisection search terminates in between $5$ to $20$ steps.
 
@@
 
-\textbf{Other policies inspired by \UCB.}
-%
-The ``Successive Elimination'' policy also uses confidence bounds, and use both Upper and Lower Confidence Bounds (LCB).
-After playing each active arms once, it eliminates arm $a$ if it is statistically identified as sub-optimal, \ie, if there is another arm $a' \neq a$ such that $\UCB_a(t) < \mathrm{LCB}_{a'}(t)$.
-It also achieves a $\bigO{\log(T)}$ problem-dependent regret upper bound, and a $\bigO{\sqrt{K T \log(T)}}$ problem-independent bound.
+% \textbf{Other policies inspired by \UCB.}
+% %
+% The ``Successive Elimination'' policy also uses confidence bounds, and use both Upper and Lower Confidence Bounds (LCB).
+% After playing each active arms once, it eliminates arm $a$ if it is statistically identified as sub-optimal, \ie, if there is another arm $a' \neq a$ such that $\UCB_a(t) < \mathrm{LCB}_{a'}(t)$.
+% It also achieves a $\bigO{\log(T)}$ problem-dependent regret upper bound, and a $\bigO{\sqrt{K T \log(T)}}$ problem-independent bound.
 
@@
 
-Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the two aforementioned point-of-views.
+Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the two aforementioned point-of-views, frequentist algorithms such as UCB and Bayesian algorithms such as Thompson sampling.
 %  the principle of optimism under uncertainty and the Bayesian point-of-view.
@@
 
-Then in Table~\ref{table:2:meanResults} we show the results of the average on $N=1000$ independent runs.
+Then in Table~\ref{table:2:meanResults} we show the \emph{results of the average on $N=1000$ independent runs}.
 %
@@
 %
-In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when optimality is important and where we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
+In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity and realistic deployment is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
 
@@
 
+% % % ----------------------------------------------------------------------------
+% \section{Appendix}
+% \label{sec:2:appendix}
+
 % % ----------------------------------------------------------------------------
-\section{Appendix}
-\label{sec:2:appendix}
-
-% ----------------------------------------------------------------------------
-\subsubsection{Pinsker's inequality}\label{proof:2:Pinsker}
-
-We prove it below using simple real analysis arguments,
-% \footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).},
-%
-and for instance we use it again in Chapter~\ref{chapter:6}.
-%
-Moreover, using the notations of the exponential families for Bernoulli and fixed-variance Gaussian distributions (see above Section~\ref{par:2:notationsExponentialFamiliesBernoulliGaussian}), it can be written $\kl(x,y) \geq d_{1/4}(x,y)$.
-%
-As bounded distributions in $[0,1]$ are sub-Bernoulli as well as $1/4$-sub-Gaussian, one consequence of this inequality is that the confidence radius of \klUCB{} is \emph{smaller} than the one in $\UCB_1$, and different concentration inequalities can then be used to prove their finite-time logarithmic regret upper-bounds, and to prove that \klUCB{} is asymptotically optimal while $\UCB_1$ is only order-optimal for bounded rewards (see above in Section~\ref{proof:2:UCBregretBound} or \cite{Auer02} for $\UCB_1$, and \cite{KLUCBJournal} for \klUCB).
-
-\begin{lemma}[Pinsker's inequality]\label{lem:2:Pinsker}
-\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) \eqdef p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
-
-    Then for any $p,q\in[0,1]$, $\kl$ verifies $\kl(p,q) \geq 2 (p-q)^2$.
-\end{leftbar}  % XXX leftbar lemmabar, comment if needed
-\end{lemma}
-
-\begin{smallproof}
-    Fix $p\in[0,1]$, and consider the function $g(x) \eqdef \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
-    We first observe that $g(0) = 0$, and we easily prove that $x=0$ is the unique minimizer of $g$ over its domain.
-    By definition, $\kl$ is of class $\cC^1$ on its two variables, on the open interval $(0,1)$, so $g$ is of class $\cC^1$ on the interior of its domain, $(-p, 1-p)$.
-    Let us compute its derivative, $g'(x) = (\frac{\partial \; \kl}{\partial 2})(p, p + x) - 4x$.
-    Simple calculus yield
-    \[ \left(\frac{\partial \; \kl}{\partial 2}\right) (p, p + x) = - \frac{p}{p+x} + \frac{1-p}{1-(p+x)} = -x \frac{(2(p + x) - 1)^2}{(p + x)(p + x - 1)}. \]
-    % https://www.wolframalpha.com/input/?i=diff(p+*+log(p%2F(p%2Bx))+%2B+(1-p)+*+log((1-p)%2F(1-(p%2Bx)))+-+2*x**2,+x)
-    We can verify that $g'(0) = -1 + 1 - 4 \times 0 = 0$, and it has the sign of $-x$ on its domain, as $p + x - 1 < 0$ because $x < 1 - p$, and $p + x > 0$.
-    So $g'$ is first positive then negative, so $g$ is increasing then decreasing on its domain, with a change in its monotony at $x=0$, and thus it indeed has a unique minimizer.
-    We conclude that $g(x) \geq g(0) = 0$ on $(-p, 1-p)$, and so $\kl(p,q) \geq 2 (p-q)^2$ if $x = q-p$.
-
-    The edge case is the ends of the interval, that is if $x=-p$ or $x=1-p$, that is if $q=0$ or $q=1$.
-    If $p\in(0,1)$, then computing the limit of $\kl(p,q)$ for $q\to0^+$ and $q\to1^-$ is obvious and gives $\kl(p,q) \to +\infty$, while the right-hand side of Pinsker's inequality stays finite (so the inequality is still true at the limit).
-    If $p=0$ or $p=1$, the same computation works, by the convention that $p \log(p) = 0$ if $p=0$.
-    %
-    % Bonus: one can verify our computation in Python with the SymPy module:
-    % \begin{verbatim}
-    %     $ (bash) python
-    %     >>> from sympy import *
-    %     >>> x, p = symbols('x p')
-    %     >>> g = p * log(p/(p+x)) + (1-p) * log((1-p)/(1-(p+x))) - 2*x**2
-    %     >>> print(diff(g, x))
-    %     -p/(p + x) - 4*x + (-p + 1)/(-p - x + 1)
-    % \end{verbatim}
-\end{smallproof}
-
-% % FIXME
-% \newpage
-% % FIXME
+% \subsubsection{Pinsker's inequality}\label{proof:2:Pinsker}
+
+% We prove it below using simple real analysis arguments,
+% % \footnote{~The proof is inspired by Lemma~10.2 from \cite{LattimoreBanditAlgorithmsBook} (version $1699$ of January $2019$).},
+% %
+% and for instance we use it again in Chapter~\ref{chapter:6}.
+% %
+% Moreover, using the notations of the exponential families for Bernoulli and fixed-variance Gaussian distributions (see above Section~\ref{par:2:notationsExponentialFamiliesBernoulliGaussian}), it can be written $\kl(x,y) \geq d_{1/4}(x,y)$.
+% %
+% As bounded distributions in $[0,1]$ are sub-Bernoulli as well as $1/4$-sub-Gaussian, one consequence of this inequality is that the confidence radius of \klUCB{} is \emph{smaller} than the one in $\UCB_1$, and different concentration inequalities can then be used to prove their finite-time logarithmic regret upper-bounds, and to prove that \klUCB{} is asymptotically optimal while $\UCB_1$ is only order-optimal for bounded rewards (see above in Section~\ref{proof:2:UCBregretBound} or \cite{Auer02} for $\UCB_1$, and \cite{KLUCBJournal} for \klUCB).
+
+% \begin{lemma}[Pinsker's inequality]\label{lem:2:Pinsker}
+% \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
+%     Let $p,q\in[0,1]$, and remember that the \emph{relative binary entropy} is defined as $\kl(p,q) \eqdef p \ln\left(\frac{p}{q}\right) + (1-p)\ln\left(\frac{1-p}{1-q}\right)$.
+
+%     Then for any $p,q\in[0,1]$, $\kl$ verifies $\kl(p,q) \geq 2 (p-q)^2$.
+% \end{leftbar}  % XXX leftbar lemmabar, comment if needed
+% \end{lemma}
+
+% \begin{smallproof}
+%     Fix $p\in[0,1]$, and consider the function $g(x) \eqdef \kl(p, p+x) - 2 x^2$ ($x$ will be $q-p$), for $x\in(-p, 1-p)$.
+%     We first observe that $g(0) = 0$, and we easily prove that $x=0$ is the unique minimizer of $g$ over its domain.
+%     By definition, $\kl$ is of class $\cC^1$ on its two variables, on the open interval $(0,1)$, so $g$ is of class $\cC^1$ on the interior of its domain, $(-p, 1-p)$.
+%     Let us compute its derivative, $g'(x) = (\frac{\partial \; \kl}{\partial 2})(p, p + x) - 4x$.
+%     Simple calculus yield
+%     \[ \left(\frac{\partial \; \kl}{\partial 2}\right) (p, p + x) = - \frac{p}{p+x} + \frac{1-p}{1-(p+x)} = -x \frac{(2(p + x) - 1)^2}{(p + x)(p + x - 1)}. \]
+%     % https://www.wolframalpha.com/input/?i=diff(p+*+log(p%2F(p%2Bx))+%2B+(1-p)+*+log((1-p)%2F(1-(p%2Bx)))+-+2*x**2,+x)
+%     We can verify that $g'(0) = -1 + 1 - 4 \times 0 = 0$, and it has the sign of $-x$ on its domain, as $p + x - 1 < 0$ because $x < 1 - p$, and $p + x > 0$.
+%     So $g'$ is first positive then negative, so $g$ is increasing then decreasing on its domain, with a change in its monotony at $x=0$, and thus it indeed has a unique minimizer.
+%     We conclude that $g(x) \geq g(0) = 0$ on $(-p, 1-p)$, and so $\kl(p,q) \geq 2 (p-q)^2$ if $x = q-p$.
+
+%     The edge case is the ends of the interval, that is if $x=-p$ or $x=1-p$, that is if $q=0$ or $q=1$.
+%     If $p\in(0,1)$, then computing the limit of $\kl(p,q)$ for $q\to0^+$ and $q\to1^-$ is obvious and gives $\kl(p,q) \to +\infty$, while the right-hand side of Pinsker's inequality stays finite (so the inequality is still true at the limit).
+%     If $p=0$ or $p=1$, the same computation works, by the convention that $p \log(p) = 0$ if $p=0$.
+%     %
+%     % Bonus: one can verify our computation in Python with the SymPy module:
+%     % \begin{verbatim}
+%     %     $ (bash) python
+%     %     >>> from sympy import *
+%     %     >>> x, p = symbols('x p')
+%     %     >>> g = p * log(p/(p+x)) + (1-p) * log((1-p)/(1-(p+x))) - 2*x**2
+%     %     >>> print(diff(g, x))
+%     %     -p/(p + x) - 4*x + (-p + 1)/(-p - x + 1)
+%     % \end{verbatim}
+% \end{smallproof}

### 2019-06-13 2-Chapters/3-Chapter/chapter3.tex
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1
Commencer à intégrer les commentaires d'Emilie. Chaud

@@
 
-\abstractStartChapter{}
+\abstractStartChapter{}%
 %
@@
     It is computationally more costly that \UCB, as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
-    In practice, a few steps of a simple Newton method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than one order of magnitude slower than \UCB.
+    In practice, a few steps of a simple bisection search method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than one order of magnitude slower than \UCB.
     % of more than 10 digits

### 2019-06-13 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Commencer à intégrer les commentaires d'Emilie. Chaud

@@
 For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to the stopping time \eqref{def:6:GLR} with the choice $d(x,y) = 2(x-y)^2$, when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
-Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$ (see Lemma~\ref{lem:2:Pinsker}), hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
+Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
 %
@@
 Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from $\mu_0$ to $\mu_1$ occurs at time $\tau$.
-We already observed that from Pinsker's inequality (Lemma~\ref{lem:2:Pinsker}), the B-GLRT stops earlier than a Gaussian GLR.
+We already observed that from Pinsker's inequality, the B-GLRT stops earlier than a Gaussian GLR.
 Hence, one can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the B-GLRT.
@@
 
-Using Pinsker's inequality from Lemma~\ref{lem:2:Pinsker} and introducing the gap $\Delta_k^{(i)} \eqdef \mu_k^{(i-1)} - {\mu}_k^{(i)}$ (which is such that $\Delta^{(i)} = |\Delta_k^k|$), one can write
+Using Pinsker's inequality, and introducing the gap $\Delta_k^{(i)} \eqdef \mu_k^{(i-1)} - {\mu}_k^{(i)}$ (which is such that $\Delta^{(i)} = |\Delta_k^k|$), one can write
 %

### 2019-06-13 1-Preamble/4-Acknowledgement/acknowledgement.tex
Done for now, I added a line about Hélène ❤

@@
 Finally, I thank sincerely my parents, Christophe and Patricia, for being the best parents since the last 30 years, and for their vital moral support.
-\emph{Merci de nous avoir inculqué une si forte curiosité et d'être si gentils !}
+\emph{Merci infiniement de nous avoir inculqué cette curiosité et d'être si gentils, et merci de votre soutien constant et de votre amour !}
 
@@
 I loved traveling with you and hope to continue!
+\emph{J'ai une pensée particulière pour Hélène, dont l'énergie et la détermination depuis 2018 m'ont poussé de l'avant.}
 

### 2019-06-13 2-Chapters/1-Chapter/chapter1.tex
Last proof reading of Ch6 and conclusion

@@
 The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
-and the network usage is more and more fluctuatating.
+and the network usage is more and more fluctuating.
 % and the stochastic hypothesis is more and more violated.

### 2019-06-13 2-Chapters/2-Chapter/chapter2.tex
Done for now, I added a line about Hélène ❤
Last proof reading of Ch6 and conclusion

@@
     and it is also called the \emph{relative binary entropy}.
-    It satisfies $\kl(x,y) \geq d_{1/4}(x,y)$ (Pinsker's inadequality).
+    It satisfies $\kl(x,y) \geq d_{1/4}(x,y)$ (Pinsker's inequality).
 
@@
 
-\paragraph{Applications for Opportunistic Spectrum Access.}
+\textbf{Applications for Opportunistic Spectrum Access.}
 % Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
-
-The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc).
-In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
-
-Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
-In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-In the model with sensing, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
+%
+As highlighted in Chapter~\ref{chapter:1},
+the focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc).
+% In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
+%
+% Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
+% In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
+In the model of Opportunistic Spectrum Access (OSA) with sensing, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
 A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
@@
 in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
-Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.
-
-\TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira..}
-In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
-of using MAB algorithms to optimize channel selection
-on the battery life of a wireless device.
-On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
-On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
-Each hardware reconfiguration costs energy for the device, and quickly switching algorithms will lead to a reduction of the battery life.
-The tradeoff between the two aspects is studied empirically in
-\cite{modiDemo2016}.
-%
-Another interesting works is \cite{Modi17QoS}.
-
-\TODOL{FIXME Finir ce blabla rapidement ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
+Proof-of-concepts using real-world radio hardware was first proposed in \cite{MoyWSR2014,RobertSDR2014}.
+%
+We study a similar model of using MAB for Cognitive Radio but without the OSA strucutre of Primary and Secondary users, \ie, without sensing information but only a collision indicator, in Chapters~\ref{chapter:4} and \ref{chapter:5}.
 
@@
 The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $\mathrm{SC}^{\cA}(T) \eqdef \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
-It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptatively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.
+It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.
 %
@@
 But even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
-
-\TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
-We find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$).
-Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
-%
-It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
-% - Switching cost? \cite{modiDemo2016} \cite{Koren17}
-On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
+%
+% \TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
+We find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
+% Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
+% %
+% It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
+% % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
+% On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
 For more details, see \cite{modiDemo2016} for experiments on the multi-players case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
@@
 %
-\TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
+% \TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
 In this thesis, we only consider fairness in the multi-players bandit models,
 in Chapter~\ref{chapter:5}, where fairness refers to a different notion (\ie, ``cooperative fairness'').
-When $M<K$ players learn in a decentralized way, they will essentially converge to an orthogonal affectation of players to arms, so that each player uses a unique arm among the $M$ best arms (most of the times).
-The fairness constraint says that, in average, no user is arbitrarily going to converge more frequently on a better arm than another user, and it is for instance the case for permutation invariant decentralized algorithms.
+% When $M<K$ players learn in a decentralized way, they will essentially converge to an orthogonal affectation of players to arms, so that each player uses a unique arm among the $M$ best arms (most of the times).
+% The fairness constraint says that, in average, no user is arbitrarily going to converge more frequently on a better arm than another user, and it is for instance the case for permutation invariant decentralized algorithms.
 We refer to Section~\ref{sub:5:betterLowerBound} for more details.
@@
 %
-\begin{align}\label{eq:UCB_index}
+\begin{align}\label{eq:2:UCB_index}
     U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation }\; \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration }\; \xi_k(t)}.
@@
 %
-Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{} defined in \eqref{eq:UCB_index} respectively represent \emph{exploitation} and \emph{exploration}, and summing them up is a natural way of dealing with the exploitation/exploration trade off.
+Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{} defined in \eqref{eq:2:UCB_index} respectively represent \emph{exploitation} and \emph{exploration}, and summing them up is a natural way of dealing with the exploitation/exploration trade off.
 The constant parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to choose $\alpha=1/2$ for optimal uniform performance (\ie, on all problems, \cite{Auer02}).
@@
 The constant $c>0$ is usually chosen as $c = 1$.
-\begin{align}\label{eq:klUCB_index}
+\begin{align}\label{eq:2:klUCB_index}
     U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{f(t}{N_k(t)} \right\}.
@@
 for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
-The second piece is less obvious: one has to solve the optimisation problem defining the $\sup$ in \eqref{eq:klUCB_index}.
+The second piece is less obvious: one has to solve the optimisation problem defining the $\sup$ in \eqref{eq:2:klUCB_index}.
 Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\kl(x,\bullet)$ function is usually convex (for any $x$), two practical and mathematically correct solutions are Newton's method or a bisection search.
@@
 %  and it did not gain popularity since then.
-% In practice, AdBandits is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
+% In practice, AdBandits is usually (slightly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
 
@@
 Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
-The first of them is the Exp3 policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loiration'',
+The first of them is the Exp3 policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
 which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details see
@@
 If the two indexes $U^{\KLUCB+}_k(t)$ and $U^{\mathrm{MOSS}+}_k(t)$ are defined by
-
+%
 % \vspace*{-20pt}   % FIXME??
@@
 % \begin{align*}
-    U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\} \\
+    U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\}
 % \end{align*}

### 2019-06-13 2-Chapters/3-Chapter/chapter3.tex
Last proof reading of Ch6 and conclusion

@@
 
-SMPyBandits stands for \emph{S}ingle and \emph{M}ulti \emph{P}layer \emph{Bandits} in P\emph{y}thon.
+SMPyBandits stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
 The library does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a pure Python implementation \cite{python}, and uses only standard open-source Python packages (\ie, no hard to install dependencies).
@@
 %
-For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulate sthe problem.
+For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulates the problem.
 We give another example for truncated Gaussian distributions, and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
@@
 
-Simulations are easily executed, \eg, Code~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPUs.
+Simulations are easily executed, \eg, Code~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPU.
 It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
@@
     % It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given, and it did not gain popularity since then.
-    In practice, it is usually (slighly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
+    In practice, it is usually (slightly) more efficient than both Thompson sampling and Bayes-UCB in terms of regret, it is comparable in terms of computation times, but costs more memory.
 

### 2019-06-13 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Last proof reading of Ch6 and conclusion

@@
   The regret lower bound that we gave in \cite{Besson2018ALT} is not applicable to any algorithm,
-  as it was discovered by Etienne Boursier and Vianney Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
+  as it was discovered by Boursier and Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
   The proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step,
   but the results stated in this section are now correct,
-  thanks to the modifications proposed by Emilie Kaufmann and Abbas Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
+  thanks to the modifications proposed by Kaufmann and Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
   The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which \emph{should} contain \RhoRand, \RandTopM{} and \MCTopM, but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
@@
         R_T(\boldsymbol{\mu}, M, \cA)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[N_k(T)].
-    \end{align*}
+    \end{align*}%
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
   Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
-  but it is not obvious for \ref{eq:5:term2} as there is no reason for $N_k(T)$ to be upper bounded by $T$.
+  but it is not obvious for \ref{eq:5:term2} as there is no reason for $N_k(T)$ to be upper bounded by $T$ (it counts the selections of arm $k$ by \emph{all} the players, and for instance if player $1$ is fixed on arm $1$ and player $2$ plays it at least once, then $N_1(T)\geqT+1$).
   %
@@
 This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
+On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
+While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
 %
 We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17} for single-player bandits, could also be fruitful.
+
+% On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps.
+For more details, see \cite{modiDemo2016} for experiments on the multi-players case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
+While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
 

### 2019-06-13 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Last proof reading of Ch6 and conclusion

@@
 
-\textbf{An interesting interpolation.}
+\paragraph{An interesting interpolation.}
 %
@@
 %
-Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\omega\in(0,1)$.
+Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\omega\in(0,1)$ (it is called $\alpha$ in both papers, but to avoid clutter with the $\alpha$ parameter in $\UCB_1$, see \eqref{eq:2:UCB_index}, we rename it $\omega$ in this chapter).
 % \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\omega/K$ at each time step,
@@
 against the alternative
-$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.\\
+$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
 %
@@
     %
+    \begin{small}
         \begin{equation}\label{def:6:GLR}
@@
         \end{equation}
+    \end{small}
     with the convention that we \emph{enforce} $\left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] = 0$
@@
     \begin{equation}\label{def:6:beta}
-        \beta(n,\delta) = 2\cT\left(\frac{\ln(3n\sqrt{n}/\delta)}{2}\right) + 6\ln(1+\ln(n)).
+        \beta(n,\delta) \eqdef 2\cT\left(\frac{\ln(3n\sqrt{n}/\delta)}{2}\right) + 6\ln(1+\ln(n)).
     \end{equation}
@@
 %
-And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
+And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum\limits_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
 %
@@
     with moment generating function $\phi_\mu(\lambda) = \bE_{X \sim \nu_{\mu}}[e^{\lambda X}]$.
-    Let $\kl_{\cE}(\mu,\mu') = \KL(\nu_{\mu},\nu_{\mu'})$ be the divergence function associated to $\cE$.
+    Let $\kl_{\cE}(\mu,\mu') \eqdef \KL(\nu_{\mu},\nu_{\mu'})$ be the divergence function associated to $\cE$.
     Let $(X_i)_{i \in \N^*}$ and $(Y_k)_{i \in \N^*}$ be two independent \iid{} processes, with respective means $\mu$ and $\mu'$, such that
@@
     %
-    Denote $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_k$.
+    Denote $\widehat{\mu}_s \denote \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r \denote \frac{1}{r}\sum\limits_{i=1}^r Y_k$.
     Then for every $s,r \in \N^*$ we have,
@@
 %
-The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
+The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
 Both approaches are parameterized by $\omega\in(0,1)$.
@@
         \end{equation}
-        \item Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and optimist guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
+        \item Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``optimist'' guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
         \begin{equation}
@@
         \end{equation}
-        \item Choosing $\omega = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and pessimist guess $\EffChange \simeq K \Upsilon_T$) gives
+        \item Choosing $\omega = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``pessimist'' guess $\EffChange \simeq K \Upsilon_T$) gives
         \begin{equation}
@@
     & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
-    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]\\
-% \end{align*}%
-% \begin{align*}
+    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
+    % \\
+\end{align*}%
+\begin{align*}
     & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
@@
 which concludes the proof.
+%
+\hfill{}$\square$
 
@@
 Using Lemma~\ref{lem:6:ConcFirst}, term $(a)$ is upper bounded by $K\delta$.
+%
+\hfill{}$\square$
 
@@
 Now the definition of $d^{(i)}$ readily implies that
-\[\left\lfloor \frac{\omega}{K}d^{(i)}\right\rfloor > \frac{4\beta(T,\delta)}{\left(\Delta^{(i)}\right)^2},\]
+$\left\lfloor \frac{\omega}{K}d^{(i)}\right\rfloor > \frac{4\beta(T,\delta)}{\left(\Delta^{(i)}\right)^2}$,
 which yields
@@
 Hence, the probability in the right-hand side of \eqref{eq:6:FromHereGlobal} is zero, which yields $(b) \leq \delta$.
+%
+\hfill{}$\square$
 

### 2019-06-13 2-Chapters/3-Chapter/chapter3.tex
\allowdisplaybreaks{} command from https://tex.stackexchange.com/a/12561/97964

@@
 %
-The code for SMPyBandits is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is freely publicly released under the MIT open-source License.
+The code for SMPyBandits is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source License.
 This chapter is based on our article \cite{SMPyBanditsJMLR}.
@@
     % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy}
-    \captionof{listing}{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
+    \captionof{listing}{Code defining the UCB algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
     % \label{lst:3:smallIndexPolicy}
@@
     \centering
-	\includegraphics[width=0.49\linewidth]{overview_documentation_1.png}
-    \includegraphics[width=0.49\linewidth]{overview_documentation_2.png}
-	\caption{Screenshots from two pages of the documentation.}
+	\includegraphics[width=0.499\linewidth]{overview_documentation_1.png}
+    \includegraphics[width=0.499\linewidth]{overview_documentation_2.png}
+	\caption{Screenshots from two pages of the documentation: homepage, and list of algorithms in the \texttt{Policies} module.}
 	\label{fig:3:twoScreenshotsOfTheDocumentation}
@@
     \centering
-	\includegraphics[width=0.49\linewidth]{overview_documentation_3.png}
-	\includegraphics[width=0.49\linewidth]{overview_documentation_4.png}
-	\caption{Screenshots from two other pages of the documentation.}
+	\includegraphics[width=0.499\linewidth]{overview_documentation_3.png}
+	\includegraphics[width=0.499\linewidth]{overview_documentation_4.png}
+	\caption{Screenshots from two other pages of the documentation: the \texttt{klUCBSwitch} policy, and documentation of the API to add a new policy.}
 	\label{fig:3:twoMoreScreenshotsOfTheDocumentation}
@@
         \end{minted}
-        \caption{Small snippet of Bash code to download and install dependencies of SMPyBandits.}
+        \caption{Example of Bash code to download and install dependencies of SMPyBandits.}
         % \captionof{listing}{Small snippet of Bash code to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
@@
     \end{minted}
-    \caption{Small snippet of Bash code to run a simple experiment with SMPyBandits.}
+    \caption{Example of Bash code to run a simple experiment with SMPyBandits.}
     % \captionof{listing}{Small snippet of Bash code to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
@@
     \end{minted}
-    \caption{Small snippet of Python code to configure the list of algorithms tested on a problem.}
+    \caption{Example of Python code to configure the list of algorithms tested on a problem.}
     % \captionof{listing}{Small snippet of Python code to configure the list of algorithms tested on a problem. \label{lst:3:howToConfigureAlgorithms}.}
@@
     % \caption{Small example of \texttt{main.py} file}
-    \captionof{listing}{Small example of \texttt{main.py} file\label{lst:3:main}.}
+    \captionof{listing}{Example of \texttt{main.py} file\label{lst:3:main}.}
     % \label{lst:3:main}
@@
     \end{minted}
-    \caption{Small snippet of Bash code to run an experiment}
+    \caption{Example of Bash code to run an experiment}
     % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}
@@
     % \caption{Small example of \texttt{configuration.py} file}
-    \captionof{listing}{Small example of \texttt{configuration.py} file\label{lst:3:configuration}.}
+    \captionof{listing}{Example of \texttt{configuration.py} file\label{lst:3:configuration}.}
     % \label{lst:3:configuration}

### 2019-06-13 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
\allowdisplaybreaks{} command from https://tex.stackexchange.com/a/12561/97964

@@
   Note that term \ref{eq:5:term3} is clearly lower bounded by $0$
-  but it is not obvious for \ref{eq:5:term2} as there is no reason for $N_k(T)$ to be upper bounded by $T$ (it counts the selections of arm $k$ by \emph{all} the players, and for instance if player $1$ is fixed on arm $1$ and player $2$ plays it at least once, then $N_1(T)\geqT+1$).
+  but it is not obvious for \ref{eq:5:term2} as there is no reason for $N_k(T)$ to be upper bounded by $T$ (it counts the selections of arm $k$ by \emph{all} the players, and for instance if player $1$ is fixed on arm $1$ and player $2$ plays it at least once, then $N_1(T) \geq T+1$).
   %
@@
   \end{equation*}
-  By focusing on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \cA)$ from Lemma~\ref{lem:5:DecompositionRegret}, we have
+  We focus on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \cA)$ from Lemma~\ref{lem:5:DecompositionRegret},
   \begin{align*}
@@
   And now both terms are non-negative, as $N_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
-  which proves that $R_T(\boldsymbol{\mu}, M, \cA) \geq \ref{eq:5:term1}$, as wanted.
+  which proves that $R_T(\boldsymbol{\mu}, M, \cA) = \ref{eq:5:term1} + \ref{eq:5:term2} + \ref{eq:5:term3} \geq \ref{eq:5:term1}$, as wanted.
 \end{smallproof}
@@
 
+\textcolor{red}{\textbf{Warning:}}
+We want to strengthen the fact that in all the proposed algorithms,
+the players do \emph{not} know their numbers $j\in[M]$,
+and they do not need to know it to achieve low regret.
+%
+We discuss the fact that this would be an unrealistic hypothesis in Section~\ref{par:5:knowingYourIDinMPBanditsGame} below.
+
 % -----------------------------------------------------------------
@@
 In particular, we focus on two types of indices:
-\UCB{} \citep{Auer02}
+$\UCB_1$ \citep{Auer02}
 and \klUCB{} \citep{KLUCBJournal}.
@@
                   %
-                  \eIf{$A^j(t-1) \notin \TopM(t)$}{
-                    \eIf(\tcp*[f]{collision at previous step}){$C^j(t-1)$}{
+                  \uIf{$A^j(t-1) \notin \TopM(t)$}{
+                    \uIf(\tcp*[f]{collision at previous step}){$C^j(t-1)$}{
                       $A^j(t) \sim \cU \left(\TopM(t)\right)$
                       \tcp*[f]{randomly switch}
-                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB}){
+                      }
+                    \uElse(\tcp*[f]{randomly switch on an arm that had smaller UCB}){
                       $A^j(t) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                     }
-                    }{
+                  }
+                  \uElse{
                     $A^j(t) = A^j(t-1)$
@@
 
-\paragraph{The \MCTopM{} algorithm.}
+\textbf{The \MCTopM{} algorithm.}
 %
@@
                     }
-                    \Else(\tcp*[f]{transition $(1)$ or $(4)$}){
+                    \uElse(\tcp*[f]{transition $(1)$ or $(4)$}){
                       $A^j(t) = A^j(t-1)$ \tcp*[f]{stay on the previous arm} \\
@@
   let player $j\in[M]$ use the \RandTopM-, \MCTopM- or \rhoRand-\klUCB{}
-  decentralized policy with exploration function $f(t) = \log(t) + 3 \log(\log(t))$.
+  decentralized policy with exploration function $f(t) \eqdef \log(t) + 3 \log(\log(t))$.
   Then for any sub-optimal arm $k \in \Mworst$ there exists problem-dependent constants $C_{\boldsymbol{\mu}}, D_{\boldsymbol{\mu}} > 0$ such that

### 2019-06-13 2-Chapters/3-Chapter/chapter3.tex
Done for proofreading Ch3

@@
     \centering
-	\includegraphics[width=0.499\linewidth]{overview_documentation_1.png}
-    \includegraphics[width=0.499\linewidth]{overview_documentation_2.png}
-	\caption{Screenshots from two pages of the documentation: homepage, and list of algorithms in the \texttt{Policies} module.}
+	\includegraphics[width=0.495\linewidth]{overview_documentation_1.png}
+    \includegraphics[width=0.495\linewidth]{overview_documentation_2.png}
+	\caption{Screenshots from two pages of the documentation: the homepage (\texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}), and a list of all the algorithms in the \texttt{Policies} module.}
 	\label{fig:3:twoScreenshotsOfTheDocumentation}
@@
     \centering
-	\includegraphics[width=0.499\linewidth]{overview_documentation_3.png}
-	\includegraphics[width=0.499\linewidth]{overview_documentation_4.png}
-	\caption{Screenshots from two other pages of the documentation: the \texttt{klUCBSwitch} policy, and documentation of the API to add a new policy.}
+	\includegraphics[width=0.495\linewidth]{overview_documentation_3.png}
+	\includegraphics[width=0.495\linewidth]{overview_documentation_4.png}
+	\caption{Screenshots from two other pages of the documentation: the \texttt{klUCBSwitch} policy, and details of the code organization, to know how to add a new policy.}
 	\label{fig:3:twoMoreScreenshotsOfTheDocumentation}
@@
 We show how to install SMPyBandits, and an example of how to run a simple experiment.
-This bash snippet\footnote{~See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}.
+This bash code\footnote{~See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}.
 You can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.},
@@
 For all experiments, we run $N=1000$ independent simulations.
-We show in the snippet of code~\ref{lst:3:BashCodeToLaunchLargeExperiments} in the Appendix below more details about how to run these experiments.
+We show in Code~\ref{lst:3:BashCodeToLaunchLargeExperiments} in the Appendix below more details about how to run these experiments.
 
@@
 
-\paragraph{Bash command to run these experiments.}
+\textbf{Bash command to run these experiments.}
 We give below in Code~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
@@
     \end{minted}
-    \caption{Snippet of Bash code to run the large experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.}
+    \caption{Bash code to run the large experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.}
     % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:BashCodeToLaunchLargeExperiments}.}
@@
 
-\begin{itemize}
-    \item Quality and optimization of the code:
+% \begin{itemize}
+    % \item
+    \emph{Quality and optimization of the code:}
     The same level of optimization is used in all the codebase of SMPyBandits, and most of the time, the code is pure and naive Python and was not optimized.
@@
 
-    \item About the experimental environment:
+    % \item
+    \emph{About the experimental environment:}
     the exact measurements used for the figures displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
     As long as the different algorithms and simulations are performed on the same environment, the comparison we make from them are fair and do make sense.
-\end{itemize}
+% \end{itemize}
 
@@
 To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.
-\begin{itemize}
-    \item For time, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
+
+% \begin{itemize}
+%     \item
+    \emph{For time}, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
     In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in[T]$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
     This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.
-    \item For memory, we use two different approaches whether the code runs on a UNIX or a Windows system.
+
+    % \item
+    \emph{For memory}, we use two different approaches whether the code runs on a UNIX or a Windows system.
     On UNIX, the \texttt{resource} module allows to measure the memory of the current process, and by counting the difference in memory used by the \texttt{Evaluator} object before and after the ``for'' loop, we can track the memory used by the algorithm to store all its interior variables.
-\end{itemize}
+% \end{itemize}
+
 In both cases, the ``for'' loop takes some time and stores many variables, but the only difference in terms of both time or memory between two algorithms is explained by the difference in the implementation of the algorithms.
@@
 
-\TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? C'est vraiment pas crucial.
-    C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
+\TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? Ce n'est pas crucial.
+    % C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
 }
 
-Nowadays, parallelism is everywhere in the computational world, and any serious framework for numerical simulations must explore at least one of the three main approaches to (try to) gain performance from parallelism.
-%
-For all the different numerical simulations presented in this thesis, except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4}, the setting is the same: we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=100$ like in Chapter~\ref{chapter:6}).
-On the one hand, because of the fundamentally sequential nature of bandit games, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
-On the other hand, parallelism can help greatly for the two other axes: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
+Nowadays, parallelism is everywhere in the computational world, and any serious framework for numerical simulations must try to gain performance from parallelism.
+In this last section, we quickly review the chosen approach for SMPyBandits (multi-core on one machine).
+
+For all the numerical simulations presented in this thesis,
+% except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4},
+the setting is the same:
+we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=100$ like in Chapter~\ref{chapter:6}).
+%
+Because of the fundamentally sequential nature of bandit games, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
+But parallelism can help greatly for the two other axes: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
 Similarly, if we want to run $100$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $100/4=25$ repetitions on each process also bring a speed-up factor of $4$.
 %
-In this last section, we quickly review the chosen approach for SMPyBandits (multi-core on one machine), and we explain why the two other approaches were less appropriate for our study of multi-armed bandit problems.
-
-\paragraph{What we did implement: \texttt{Joblib} for multi-core simulations.}
+%  and we explain why the two other approaches were less appropriate for our study of multi-armed bandit problems.
+
+\textbf{What we did implement: \texttt{Joblib} for multi-core simulations.}
 %
@@
 As explained in the page \href{https://smpybandits.github.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}, we implemented in SMPyBandits an easy way to run any simulations on $n$ cores of a machine, using the \texttt{Joblib} \cite{joblib} library.
-It is implemented in a completely transparent way, and if someone uses the command-line variable to configure experiments, using one core or all the cores of the machine one changes \texttt{N\_JOBS=1} to \texttt{N\_JOBS=-1}, like in this example.
+It is implemented in a completely transparent way, and if someone uses the command-line variable to configure experiments, switching from using one core to using all the cores of the machine is as simple as changing \texttt{N\_JOBS=1} to \texttt{N\_JOBS=-1}, like in the Code Example~\ref{lst:3:runOneCoreOrMore} below.
 
@@
         \end{minted}
-        \caption{To run such simulation using the maximum number of cores, use \texttt{N\_JOBS=-1} instead.}
+        \caption{To run such simulation using all the CPU cores, use \texttt{N\_JOBS=-1} instead.}
         % \captionof{listing}{Run a simulation using 1 or \texttt{N} cores \label{lst:3:runOneCoreOrMore}.}
@@
 Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
+%
+More details are included in the documentation page
+\href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
 

### 2019-06-13 1-Preamble/5-Abstract/abstract.tex
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1
Quite a weird change of outline, but I like it!

@@
 \begin{center}
-	\textbf{{\Large Résumé}}
+	\textbf{{\LARGE Résumé}}
 \end{center}
@@
 % Chapitres 5 et 6
-Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueur non stationnaires.
+Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits mono-joueur non stationnaires.
 %  (Chapitre~\ref{chapitre:5})
@@
 \begin{center}
-	\textbf{{\Large Abstract}}
+	\textbf{{\LARGE Abstract}}
 \end{center}

### 2019-06-13 1-Preamble/5-Abstract/resume_fr.tex
Quite a weird change of outline, but I like it!

@@
 % Chapitres 5 et 6
-Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires (Chapitre~\ref{chapter:5}), puis des bandits simple-joueurs non stationnaires (Chapitre~\ref{chapter:6}).
+Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires (Chapitre~\ref{chapter:5}), puis des bandits mono-joueurs non stationnaires (Chapitre~\ref{chapter:6}).
 % Chapitre 3

### 2019-06-13 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Quite a weird change of outline, but I like it!

@@
 To the best of our knowledge, aggregation algorithms, such as \ExpQ{} which dates back from 2002 \cite{Auer02},
-have never been used in pratice for stochastic MAB problems.
+have never been used in practice for stochastic MAB problems.
 % and we show that it appears empirically sub-optimal when applied to simple stochastic problems.
@@
 For synthetic MAB problems, with Bernoulli or other distributions for $K$ arms, simulation results are presented to demonstrate its empirical efficiency.
-We combine classical algorithms, such as Thompson sampling, Upper-Confidence Bounds algorithms (\UCB{} and variants), and Bayesian or Kullback-Leibler UCB.
+We combine classical algorithms, such as Thompson sampling, Upper-Confidence Bounds algorithms (\UCB{} and variants), and Bayesian or \klUCB.
 %
@@
 % ----------------------------------------------------------------------
-% \subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
-\textbf{Notations:}\label{sub:25:aggregation}
+\subsection{Aggregating bandit algorithms}\label{sub:25:aggregation}
+% \paragraph{Notations:}\label{sub:25:aggregation}
 %
@@
 % ----------------------------------------------------------------------
-\subsection{Our contribution: the \Aggr{} algorithm}\label{sub:25:Aggr}
+\section{Our contribution: the \Aggr{} algorithm}\label{sub:25:Aggr}
 
-\begin{small}
+Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
+%
+At every time step, after having observed a reward $r(t)=Y_{A(t),t}$ for its chosen action $A(t)$,
+the algorithm updates the trust probabilities from $\pi^t$ to $\pi^{t+1}$ by
+a multiplicative exponential factor (using the learning rate and the \emph{unbiased} reward).
+%
+Only the algorithms $\Alg_a$ who advised the last decision get their trust updated, in order to trust more the ``reliable'' algorithms.
+
+% \begin{small}
 	\begin{figure}[h!]
@@
 			% \KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
-			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$}
+			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$, \eg, $\eta_t=\frac{\log(N)}{t K}$}
 			\KwData{Initial uniform distribution, $\pi^{0} = \cU([N])$}
@@
 	\end{figure}
-\end{small}
+% \end{small}
 
-Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
+\paragraph{Unbiased estimate of the rewards.}
 %
-At every time step, after having observed a reward $r(t)=Y_{A(t),t}$ for its chosen action $A(t)$,
-the algorithm updates the trust probabilities from $\pi^t$ to $\pi^{t+1}$ by
-a multiplicative exponential factor (using the learning rate and the \emph{unbiased} reward).
-%
-Only the algorithms $\Alg_a$ who advised the last decision get their trust updated, in order to trust more the ``reliable'' algorithms.
-
 The reward estimate is unbiased in the following sense. If one had access to the samples $Y_{k,t}$ for all arms $k$, the reward incurred by algorithm $a$ at time $t$ would be $r^a(t) = Y_{A_a(t),t}$.
@@
 % ----------------------------------------------------------------------
-\textbf{Comparison with \ExpQ{}:}\label{sub:25:Exp4}
+\paragraph{Comparison with \ExpQ{}:}\label{sub:25:Exp4}
 %
@@
 
-\textbf{How to choose the learning rates $(\eta_t)$:}
+\paragraph{How to choose the learning rates $(\eta_t)$:}
 %
@@
 % ----------------------------------------------------------------------
-\subsection{Experiments on simulated MAB problems}\label{sub:25:numExp}
+\section{Experiments on simulated MAB problems}\label{sub:25:numExp}
 
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
 	\caption{The semilog-$x$ scale clearly shows the logarithmic growth of the regret for the best algorithms and our proposal \Aggr, even in a hard ``mixed'' problem (\emph{cf}.
@@
 % ----------------------------------------------------------------------
-\paragraph{No theoretical guarantees.}\label{sub:25:theory}
+\subsection{No theoretical guarantees.}\label{sub:25:theory}
 
@@
 
-% ----------------------------------------------------------------------
-\paragraph{Conclusion about our \Aggr{} algorithm.}\label{sub:25:conclusion}
+% % ----------------------------------------------------------------------
+% \paragraph{Conclusion about our \Aggr{} algorithm.}\label{sub:25:conclusion}
 
-We presented the use of aggregation algorithms in the context of Opportunistic Spectrum Access for Cognitive Radio,
-especially for the real-world setting of unknown problem instances,
-when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
-% - \ExpQ and \Aggr works fine
-% Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
-Our proposed \Aggr{} was presented in details,
-and we also highlighted its differences with \ExpQ.
-% as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
-%
-% - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
-We presented experiments on simple MAB problems coming from previous work on bandits for OSA \cite{Jouini10},
-and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
-Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
-% It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
-% and appears to be more robust than simple non-aggregated algorithms.
+% We presented the use of aggregation algorithms in the context of Opportunistic Spectrum Access for Cognitive Radio,
+% especially for the real-world setting of unknown problem instances,
+% when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
+% % - \ExpQ and \Aggr works fine
+% % Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
+% Our proposed \Aggr{} was presented in details,
+% and we also highlighted its differences with \ExpQ.
+% % as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
+% %
+% % - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
+% We presented experiments on simple MAB problems coming from previous work on bandits for OSA \cite{Jouini10},
+% and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
+% Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
+% % It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
+% % and appears to be more robust than simple non-aggregated algorithms.
 
-\ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
-Empirically, \Aggr{} showed to always have a logarithmic
-regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
-It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
-As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.
-%
-What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
-%
-%
-Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
-But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
-% We will continue investigating regret bounds for \Aggr,
-% and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).
+% \ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
+% Empirically, \Aggr{} showed to always have a logarithmic
+% regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
+% It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
+% As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.
+% %
+% What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
+% %
+% %
+% Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
+% But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
+% % We will continue investigating regret bounds for \Aggr,
+% % and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).
 
-\paragraph{Reproducility of the experiments.}
-%
-Experiments presented in this Section are based on our library SMPyBandits,
-and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.
+% \paragraph{Reproducility of the experiments.}
+% %
+% Experiments presented in this Section are based on our library SMPyBandits,
+% and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.
 

### 2019-06-13 2-Chapters/2-Chapter/chapter2.tex
Quite a weird change of outline, but I like it!

@@
 We define the notion of regret of an algorithm, and quote important results from the literature. A short review of the most important families of MAB algorithms is then given.
-We conclude this chapter by discussing the question of online algorithm selection.
+% We conclude this chapter by discussing the question of online algorithm selection.
 
@@
 We give in Section~\ref{sec:2:famousMABalgorithms} an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the $\UCB_1$, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
-% We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contributions of this thesis.
-% We use it to compare the empirical performances of the most famous and most efficient algorithms, in terms of regret and by evaluating them on different problems.
+% % We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contributions of this thesis.
+% % We use it to compare the empirical performances of the most famous and most efficient algorithms, in terms of regret and by evaluating them on different problems.
+% % %
+% % In the point-of-view of numerical simulations or real-world applications of bandits on embedded systems, one must also take care of two other measures of performance: time and memory complexities, which can vary greatly between different families of algorithms.
+% % We also empirically compare different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
 % %
-% In the point-of-view of numerical simulations or real-world applications of bandits on embedded systems, one must also take care of two other measures of performance: time and memory complexities, which can vary greatly between different families of algorithms.
-% We also empirically compare different algorithms for these two measures in Section~\ref{sec:3:timeAndMemoryCosts}.
-%
-% We then give regret upper-bounds of the two main algorithms used in this thesis, $\UCB_1$ and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
+% % We then give regret upper-bounds of the two main algorithms used in this thesis, $\UCB_1$ and \klUCB, to highlight the difference between their finite-time regret upper-bounds.
 They are order-optimal as they attain a logarithmic regret.
-% proving their order optimality.
-% (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
-% Moreover, \klUCB{} is known to be optimal for a wide range of arms distributions.
-% A sketch of the proof of both results is given in Appendix.
-%
-Finally, we explore in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} the problem of algorithm selection: as there are many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
-We present one of the first contributions made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms,
-and we illustrate its performance empirically.
-% , but no satisfactory theoretical result is given, as the bound we had hoped for has been proven unachievable since then.
+% % proving their order optimality.
+% % (\ie, their regret upper-bound asymptotically matches the lower-bound up-to a constant factor).
+% % Moreover, \klUCB{} is known to be optimal for a wide range of arms distributions.
+% % A sketch of the proof of both results is given in Appendix.
+% %
+% Finally, we explore in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} the problem of algorithm selection: as there are many different algorithms, all having different qualities and properties, how one can decide which algorithm to use for a particular application?
+% We present one of the first contributions made during my PhD \cite{Besson2018WCNC}, consisting of an efficient algorithm for aggregating algorithms,
+% and we illustrate its performance empirically.
+% % , but no satisfactory theoretical result is given, as the bound we had hoped for has been proven unachievable since then.
 
@@
 
-\textbf{Publication.}
-%
-The first Sections~\ref{sec:2:notations}, \ref{sec:2:applicationsofStochasticMAB}, \ref{sec:2:lowerUpperBoundsRegret} and \ref{sec:2:famousMABalgorithms} of this chapter are based on existing works,
-and Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on our article \cite{Besson2018WCNC}.
+\textbf{No publication associated with this chapter.}
+%
+% The first Sections~\ref{sec:2:notations}, \ref{sec:2:applicationsofStochasticMAB}, \ref{sec:2:lowerUpperBoundsRegret} and \ref{sec:2:famousMABalgorithms} of this chapter are based on existing works.
+This chapter is based on existing works.
+% and Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on our article \cite{Besson2018WCNC}.
 
@@
 \textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
-When the player decides at time $t$ to play (or pull) the arm $A(t)\in\{1,\dots,K\}=[K]$, she receives a reward $r(t) \eqdef Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
+When the player decides at time $t$ to play (or pull) the arm $A(t)\in[K]$,
+where this notation denotes $[K] \eqdef \{1,\dots,K\}$, she receives a reward $r(t) \eqdef Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
 %
@@
 %
-We also restrict to \emph{stochastic} rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in\{1,\dots,\}=[K]$.
+We also restrict to \emph{stochastic} rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in[K]$.
 Rewards are stationary, meaning that $(Y_{k,t})_{t\in\N^*}$ is independent and identically distributed (\iid), $Y_{k,t} \sim \nu_k$ for any $t\geq1$.
@@
 
-The different applications detailed above are still active research directions,
-and a curious reader can find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}, such as game tree search or network routing in Section~1.2 of \cite{LattimoreBanditAlgorithmsBook}.
-
 \textbf{Applications for Opportunistic Spectrum Access.}
@@
 We study a similar model of using MAB for Cognitive Radio but without the OSA strucutre of Primary and Secondary users, \ie, without sensing information but only a collision indicator, in Chapters~\ref{chapter:4} and \ref{chapter:5}.
+
+The different applications discussed in this Section are still active research directions,
+and a curious reader can find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}, such as game tree search or network routing in Section~1.2 of \cite{LattimoreBanditAlgorithmsBook}.
 
@@
 %
-We describe our library SMPyBandits in more details in Chapter~\ref{chapter:3}, but all the algorithms described in this chapter are implemented in SMPyBandits, in the \texttt{Policies} module, alongside with many more algorithms (there are about 65 for single-player stochastic problems).
+We describe our library SMPyBandits in more details in Chapter~\ref{chapter:3}.
+All the algorithms described in this chapter are implemented in SMPyBandits, in the \texttt{Policies} module, alongside with many more algorithms (there are about 65 for single-player stochastic problems).
 A complete list of the implemented policies can be found on the following web page on the documentation,
@@
 Let us first describe two naive strategies, that both fail dramatically.
-They are detailed in Algorithm~\ref{algo:2:naiveStrategies} below.
+% They are detailed in Algorithm~\ref{algo:2:naiveStrategies} below.
 We recall the notations introduced above in Section~\ref{par:2:interactiveDemoDiscoverMAB}, the sums of rewards are $X_k(t) \eqdef \sum_{s=1}^t r(s) \mathbbm{1}(A(s) = k)$, and the numbers of samples are $N_k(t) \eqdef \sum_{s=1}^t \mathbbm{1}(A(s) = k)$.
@@
 
--- The uniform strategy always plays the $K$ arms uniformly at random.
+-- The uniform strategy always plays the $K$ arms uniformly at random,
+$\forall t\in[T], A(t) \sim \cU([K])$, where $\cU(S)$ denotes the uniform distribution on set $S$, and $\cU([K])$ the uniform distribution on $[K] = \{1,\dots,K\}$ (\ie, $\forall t, \forall k\in[K], \Pr(A(t) = k) = \frac{1}{K}$).
+%
 The player \textcolor{darkgreen}{\emph{only explores}} without using the collected information, and this strategy fails dramatically for any non trivial problem (\ie, linear regret).
@@
 
-% \begin{small} % XXX remove if needed
-\begin{figure}[h!]
-	\centering
-    \begin{framed}
-	\begin{algorithm}[H]
-		% \begin{small} % XXX remove if needed
-		\For(){$t = 1, 2, \dots, T$}{
-            \uIf{\textcolor{darkgreen}{pure exploration}}{
-                Play uniformly at random \textcolor{darkgreen}{$A(t) \in \cU(\{1,\dots,K\})$}
-                \tcp*[f]{Explore}
-            }
-            \uIf{\textcolor{blue}{pure exploitation}}{
-                Play uniformly among the arms of maximal empirical mean: \textcolor{blue}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(t))$}
-                \tcp*[f]{Exploit}\\
-                Observe a reward $r(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
-            }
-		}
-		\caption[Naive strategies: pure exploitation or pure exploration.]{Naive strategies: \textcolor{darkgreen}{pure exploitation} or \textcolor{blue}{pure exploration}.}
-		\label{algo:2:naiveStrategies}
-		% \end{small} % XXX remove if needed
-	\end{algorithm}
-	\end{framed}
-\end{figure}
-% \end{small} % XXX remove if needed
+% % \begin{small} % XXX remove if needed
+% \begin{figure}[h!]
+% 	\centering
+%     \begin{framed}
+% 	\begin{algorithm}[H]
+% 		% \begin{small} % XXX remove if needed
+% 		\For(){$t = 1, 2, \dots, T$}{
+%             \uIf{\textcolor{darkgreen}{pure exploration}}{
+%                 Play uniformly at random \textcolor{darkgreen}{$A(t) \sim \cU(\{1,\dots,K\})$}
+%                 \tcp*[f]{Explore}
+%             }
+%             \uIf{\textcolor{blue}{pure exploitation}}{
+%                 Play uniformly among the arms of maximal empirical mean: \textcolor{blue}{$A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(t))$}
+%                 \tcp*[f]{Exploit}\\
+%                 Observe a reward $r(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
+%             }
+% 		}
+% 		\caption[Naive strategies: pure exploitation or pure exploration.]{Naive strategies: \textcolor{darkgreen}{pure exploitation} or \textcolor{blue}{pure exploration}.}
+% 		\label{algo:2:naiveStrategies}
+% 		% \end{small} % XXX remove if needed
+% 	\end{algorithm}
+% 	\end{framed}
+% \end{figure}
+% % \end{small} % XXX remove if needed
 
@@
                 \uIf{\textcolor{deeppurple}{$U_t < \varepsilon$ (\ie, with probability $\varepsilon$)}}{
-                    Play uniformly at random $A(t) \in \cU(\{1,\dots,K\})$
+                    Play uniformly at random $A(t) \sim \cU(\{1,\dots,K\})$
                     \tcp*[f]{Explore}
@@
                 \uElse{
-                    Play uniformly among the arms of maximal empirical mean: $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(t))$
+                    Play uniformly among the arms of maximal empirical mean: $A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(t))$
                     \tcp*[f]{or Exploit}
@@
                     \uIf{$t = \textcolor{deepgold}{T_0}$}{
-                        Pick one arm $A(\textcolor{deepgold}{T_0})=k$ uniformly among the arms of maximal empirical mean: \textcolor{deepgold}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(\textcolor{deepgold}{T_0}))$}
+                        Pick one arm $A(\textcolor{deepgold}{T_0})=k$ uniformly among the arms of maximal empirical mean: \textcolor{deepgold}{$A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(\textcolor{deepgold}{T_0}))$}
                         \tcp*[f]{then Exploit}
@@
             }
-			\Else(){
+			\uElse(){
                 For each arm, compute the \textcolor{blue}{index $U_k^{\cA}(t)$}\;
-                Play arm $A(t)$ uniformly among the arms of maximal index: \textcolor{blue}{$A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$}\;
+                Play arm $A(t)$ uniformly among the arms of maximal index: \textcolor{blue}{$A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$}\;
             }
@@
 
-This selection rule $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$ makes sense for the following reasons:
+This selection rule $A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$ makes sense for the following reasons:
 an arm $k$ is chosen at step $t$ because it has a large $\UCB_k(t)$, which can happen for two reasons.
@@
             For each arm, sample $s_k(t) \sim \pi_k(t-1)$ from the posterior of arm $k$, $\pi_k(t-1)=\mathrm{Beta}(X_k(t-1)+1,N_k(t-1)-X_k(t-1)+1)$\;
-            Play arm $A(t)$ uniformly among the arms of maximal random sample: $A(t) \in \cU(\arg\max\limits_{1\leq k \leq K} U_k(t))$\;
+            Play arm $A(t)$ uniformly among the arms of maximal random sample: $A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} U_k(t))$\;
             Observe a reward $r(t)$ and update $X_k(t)$ and $N_k(t)$\;
@@
 
-% \newpage
-
-% ----------------------------------------------------------------------------
-\section{Different approaches on algorithm selection}
-\label{sec:2:chooseYourPreferredBanditAlgorithm}
-
-For any real-world applications of MAB algorithms,
-several solutions have been explored based on various models, and for any model, typically there are many algorithms available, as we have seen above.
-%
-Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to use.
-It is hard to predict which solution could be the best for real-world conditions at every instant,
-and even by assuming a stationary environment, when one is facing a certain problem but has limited information about it, it is hard to know beforehand which algorithm can be the best solution.
-
-In this Section, we first present two naive approaches for selecting an algorithm when facing a new problem, and then we detail the online approach that uses a ``leader'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``leader'' algorithm based on \ExpQ.
-This section is based on our article \cite{Besson2018WCNC}.
-
-% ----------------------------------------------------------------------
-\subsection{Motivation for online algorithm selection}\label{sub:25:introduction}
-
-Many different learning algorithms have been proposed by the machine learning community,
-and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
-the $\kl$ function for \klUCB{} etc.
-Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instances, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
-%
-For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms,
-but when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
-ideally one would like to use an algorithm that can be applied identically against any problem of such family.
-
-\textbf{Naive approaches:}
-%
-On the one hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
+% % \newpage
+
+% % ----------------------------------------------------------------------------
+% \section{Different approaches on algorithm selection}
+% \label{sec:2:chooseYourPreferredBanditAlgorithm}
+
+% For any real-world applications of MAB algorithms,
+% several solutions have been explored based on various models, and for any model, typically there are many algorithms available, as we have seen above.
+% %
+% Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to use.
+% It is hard to predict which solution could be the best for real-world conditions at every instant,
+% and even by assuming a stationary environment, when one is facing a certain problem but has limited information about it, it is hard to know beforehand which algorithm can be the best solution.
+
+% In this Section, we first present two naive approaches for selecting an algorithm when facing a new problem, and then we detail the online approach that uses a ``leader'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``leader'' algorithm based on \ExpQ.
+% This section is based on our article \cite{Besson2018WCNC}.
+
+% % ----------------------------------------------------------------------
+% \subsection{Motivation for online algorithm selection}\label{sub:25:introduction}
+
+% Many different learning algorithms have been proposed by the machine learning community,
+% and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
+% the $\kl$ function for \klUCB{} etc.
+% Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instances, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
+% %
+% For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms,
+% but when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
+% ideally one would like to use an algorithm that can be applied identically against any problem of such family.
+
+% \textbf{Naive approaches:}
+% %
+% On the one hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
 It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the $\UCB_1$ algorithm (or other low-cost algorithms).
-%
-On the other hand, if prior knowledge on the application at hand is available, one could implement some benchmarks, and compare a set of algorithms on different problems.
+% %
+% On the other hand, if prior knowledge on the application at hand is available, one could implement some benchmarks, and compare a set of algorithms on different problems.
 If a leader appears clearly, it is then possible to choose it for the application.
 
-\textbf{Illustrative example:}
-%
-For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of $\UCB_1$ will perform quite differently in the two problems:
-using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case,
-while using a low $\alpha$, \ie, favoring exploitation, will give low regret in the second case.
-One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
-
-\textbf{The online approach:}
-%
-Another possibility is to consider an online approach, which is interested in the case where the computation power or memory storage of the application is not a limitation factor, but where one cannot run benchmarks before deploying the application.
-We consider a fixed set of algorithms, and we use another learning algorithm on top of this set, to learn \emph{on the fly} which one should be trusted more, and eventually, used on its own.
-
-The aggregation approach is especially interesting if we know that the problem the application will face is one of a few known kinds of problems.
-In such cases, if there are $N$ different sorts of problems, and if the practitioner has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
-and use the aggregation of $\cA_1,\dots,\cA_N$ when facing the unknown problem.
-
-% To choose the best algorithm, three approaches can be followed:
-% \begin{itemize}
-%     \item
-% \end{itemize}
-% either extensive benchmarks are done beforehand -- if this is possible -- to select the algorithm and its optimal parameters, or an adaptive algorithm is used to learn \emph{on the fly} its parameters.
-
-% We present a simple adaptive solution, that aggregates several learning algorithms in parallel and adaptively chooses which one to trust the most.
-
-% \paragraph{Outline.}
-% This section is organized as follows.
-% First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
-% Then we present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
-% Finally, we present numerical experiments in Section~\ref{sub:25:numExp},
-% on Bernoulli and non-Bernoulli MAB problems,
-% comparing the regret of several algorithms against different aggregation algorithms.
-% Theoretical guarantees are shortly discussed in Section~\ref{sub:25:theory}, and Section~\ref{sub:25:conclusion} concludes.
-
-% \subsection{Naive approaches}
-
-% As mentioned, there are two naive approaches
-
-% use UCB or Thompson sampling, and forget about the rest!}
-% First solution: be naive, only use UCB everywhere
-
-% \subsection{Use prior knowledge about the future application to select beforehand the chosen algorithm}
-
-% Second solution, if one has some prior knowledge about the domain or setting for which the learning algorithms will be deployed, she can run synthetic or real-world simulations, compare many algorithms before deployment, and select the best algorithm!
-
-\subsection{Online algorithm selection with expert aggregation}
-
-\input{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex}
-
-% A last solution is online algorithm selection, inspired from expert aggregation.
-% Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292
+% \textbf{Illustrative example:}
+% %
+% For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of $\UCB_1$ will perform quite differently in the two problems:
+% using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case,
+% while using a low $\alpha$, \ie, favoring exploitation, will give low regret in the second case.
+% One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
+
+% \textbf{The online approach:}
+% %
+% Another possibility is to consider an online approach, which is interested in the case where the computation power or memory storage of the application is not a limitation factor, but where one cannot run benchmarks before deploying the application.
+% We consider a fixed set of algorithms, and we use another learning algorithm on top of this set, to learn \emph{on the fly} which one should be trusted more, and eventually, used on its own.
+
+% The aggregation approach is especially interesting if we know that the problem the application will face is one of a few known kinds of problems.
+% In such cases, if there are $N$ different sorts of problems, and if the practitioner has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
+% and use the aggregation of $\cA_1,\dots,\cA_N$ when facing the unknown problem.
+
+% % To choose the best algorithm, three approaches can be followed:
+% % \begin{itemize}
+% %     \item
+% % \end{itemize}
+% % either extensive benchmarks are done beforehand -- if this is possible -- to select the algorithm and its optimal parameters, or an adaptive algorithm is used to learn \emph{on the fly} its parameters.
+
+% % We present a simple adaptive solution, that aggregates several learning algorithms in parallel and adaptively chooses which one to trust the most.
+
+% % \paragraph{Outline.}
+% % This section is organized as follows.
+% % First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
+% % Then we present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
+% % Finally, we present numerical experiments in Section~\ref{sub:25:numExp},
+% % on Bernoulli and non-Bernoulli MAB problems,
+% % comparing the regret of several algorithms against different aggregation algorithms.
+% % Theoretical guarantees are shortly discussed in Section~\ref{sub:25:theory}, and Section~\ref{sub:25:conclusion} concludes.
+
+% % \subsection{Naive approaches}
+
+% % As mentioned, there are two naive approaches
+
+% % use UCB or Thompson sampling, and forget about the rest!}
+% % First solution: be naive, only use UCB everywhere
+
+% % \subsection{Use prior knowledge about the future application to select beforehand the chosen algorithm}
+
+% % Second solution, if one has some prior knowledge about the domain or setting for which the learning algorithms will be deployed, she can run synthetic or real-world simulations, compare many algorithms before deployment, and select the best algorithm!
+
+% \subsection{Online algorithm selection with expert aggregation}
+
+% \input{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex}
+
+% % A last solution is online algorithm selection, inspired from expert aggregation.
+% % Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292
 
@@
 we presented the notations used in all this thesis.
-
-The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
-% and corresponds to our article \cite{Besson2018WCNC}.
-We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
-Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
-%
-The numerical simulations used our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
-The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.
+%
+% The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+% % and corresponds to our article \cite{Besson2018WCNC}.
+% We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
+% Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+% We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+% %
+We also showed a small numerical simulation, comparing some algorithms on a small Bernoulli-distributed problem, using our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
+% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.
 

### 2019-06-13 2-Chapters/2-Chapter/chapter25.tex
Quite a weird change of outline, but I like it!

%!TEX root = ../PhD_thesis__Lilian_Besson

\chapter{Expert aggregation for online MAB algorithms selection}
\label{chapter:25}

\graphicspath{{2-Chapters/2-Chapter/Images/}}

\abstractStartChapter{}%
%
In this chapter, we discuss the question of online algorithm selection.
%
We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.

\minitocStartChapter{}

% \newpage

% ----------------------------------------------------------------------------
\section{Different approaches on algorithm selection}
\label{sec:25:chooseYourPreferredBanditAlgorithm}

For any real-world applications of MAB algorithms,
several solutions have been explored based on various models, and for any model, typically there are many algorithms available, as we have seen above.
%
Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to use.
It is hard to predict which solution could be the best for real-world conditions at every instant,
and even by assuming a stationary environment, when one is facing a certain problem but has limited information about it, it is hard to know beforehand which algorithm can be the best solution.

In this Section, we first present two naive approaches for selecting an algorithm when facing a new problem, and then we detail the online approach that uses a ``leader'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``leader'' algorithm based on \ExpQ.

\paragraph{Publication.}
%
This chapter is based on our article \cite{Besson2018WCNC}.

\paragraph{Outline.}
This chapter is organized as follows.
First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
Then we present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
Finally, we present numerical experiments in Section~\ref{sub:25:numExp},
on Bernoulli and non-Bernoulli MAB problems,
comparing the regret of several algorithms against different aggregation algorithms.
Theoretical guarantees are shortly discussed in Section~\ref{sub:25:theory}, and Section~\ref{sec:25:conclusion} concludes.

% ----------------------------------------------------------------------
\subsection{Motivation for online algorithm selection}\label{sub:25:introduction}

Many different learning algorithms have been proposed by the machine learning community,
and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
the $\kl$ function for \klUCB{} etc.
Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instances, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
%
For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms,
but when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
ideally one would like to use an algorithm that can be applied identically against any problem of such family.

\paragraph{Naive approaches:}
%
On the one hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application. It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the $\UCB_1$ algorithm (or other low-cost algorithms).
%
On the other hand, if prior knowledge on the application at hand is available, one could implement some benchmarks, and compare a set of algorithms on different problems. If a leader appears clearly, it is then possible to choose it for the application.

\paragraph{Illustrative example:}
%
For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of $\UCB_1$ will perform quite differently in the two problems:
using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case,
while using a low $\alpha$, \ie, favoring exploitation, will give low regret in the second case.
One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.

% \paragraph{The online approach:}
% %
% Another possibility is to consider an online approach, which is interested in the case where the computation power or memory storage of the application is not a limitation factor, but where one cannot run benchmarks before deploying the application.
% We consider a fixed set of algorithms, and we use another learning algorithm on top of this set, to learn \emph{on the fly} which one should be trusted more, and eventually, used on its own.

% The aggregation approach is especially interesting if we know that the problem the application will face is one of a few known kinds of problems.
% In such cases, if there are $N$ different sorts of problems, and if the practitioner has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
% and use the aggregation of $\cA_1,\dots,\cA_N$ when facing the unknown problem.

\subsection{Online algorithm selection with expert aggregation}

Another possibility is to consider an online approach, which is interested in the case where the computation power or memory storage of the application is not a limitation factor, but where one cannot run benchmarks before deploying the application.
We consider a fixed set of algorithms, and we use another learning algorithm on top of this set, to learn \emph{on the fly} which one should be trusted more, and eventually, used on its own.

The aggregation approach is especially interesting if we know that the problem the application will face is one of a few known kinds of problems.
In such cases, if there are $N$ different sorts of problems, and if the practitioner has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
and use the aggregation of $\cA_1,\dots,\cA_N$ when facing the unknown problem.

\input{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex}

% A last solution is online algorithm selection, inspired from expert aggregation.
% Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292

% \newpage  % WARNING ?

% ----------------------------------------------------------------------------
\section{Conclusion}
\label{sec:25:conclusion}

In this chapter, we presented the first contribution of this manuscript,
that corresponds to our article \cite{Besson2018WCNC}.
%
We presented the use of aggregation algorithms,
especially for the real-world setting of unknown problem instances,
when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
% - \ExpQ and \Aggr works fine
% Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
Our proposed \Aggr{} was presented in details,
and we also highlighted its differences with \ExpQ.

% as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
%
% - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
We presented experiments on simple MAB problems, coming from previous work on bandits for OSA \cite{Jouini10},
and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
% It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
% and appears to be more robust than simple non-aggregated algorithms.

\ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
Empirically, \Aggr{} showed to always have a logarithmic
regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.

What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
%
%
Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
% We will continue investigating regret bounds for \Aggr,
% and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).

% The numerical simulations used our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

\paragraph{Reproducility of the experiments.}
%
Experiments presented in this Section are based on our library SMPyBandits,
and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.

### 2019-06-13 2-Chapters/2-Chapter/chapter2.tex
SO MANY CHANGES

@@
 
-\chapter{Stochastic Multi-Armed Bandits Models}
+\chapter{The Stochastic Multi-Armed Bandits Model}
 \label{chapter:2}
@@
 %
-In this chapter, we present the common base of all the models studied in this thesis: the multi-armed bandit (MAB) model, restricting to the single-player and stationary stochastic case.
+In this chapter, we present the common base of all the more sophisticated models studied in this thesis: the stochastic multi-armed bandit (MAB) model, restricting to the single-player and stationary stochastic case.
+
+We motivate this model and discuss its application to and beyond cognitive radio, and we review existing MAB algorithms.
+Some of which are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling.
+
 We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a one-dimensional distribution.
@@
 
-% ----------------------------------------------------------------------------
-\section{Introduction}
-\label{sec:2:introduction}
-
-Multi-Armed Bandits (MAB) models were introduced by Thompson as early as in 1933 \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
-This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
-Their name refers to one-armed bandits found in casinos, as shown in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
-% The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
-% Since then, many efficient algorithms were proposed, XXX
-%
-% In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
-We formally define the model and the main notations used in all this document in Section~\ref{sec:2:notations}.
-%
-Multi-armed bandits have been applied to various fields, from recommender systems and information retrieval to healthcare and finance.
+% % ----------------------------------------------------------------------------
+% \section{Introduction}
+% \label{sec:2:introduction}
+
+% Multi-Armed Bandits (MAB) models were introduced by Thompson as early as in 1933 \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
+% This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
+% Their name refers to one-armed bandits found in casinos, as shown in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
+% % The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
+% % Since then, many efficient algorithms were proposed, XXX
+% %
+% % In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
+% We formally define the model and the main notations used in all this document in Section~\ref{sec:2:notations}.
+% %
+% Multi-armed bandits have been applied to various fields, from recommender systems and information retrieval to healthcare and finance.
 We discuss some applications in Section~\ref{sec:2:applicationsofStochasticMAB}.
-In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits has already been applied with success in the last ten years.
+% In the rest of this thesis, we focus on cognitive radio, an important field of research where bandits has already been applied with success in the last ten years.
 
@@
 
-A rich research literature has produced many different algorithms tackling the MAB problem, as suggested by the length of the reference book \cite{LattimoreBanditAlgorithmsBook}.
-We give in Section~\ref{sec:2:famousMABalgorithms} an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the $\UCB_1$, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
+% A rich research literature has produced many different algorithms tackling the MAB problem, as suggested by the length of the reference book \cite{LattimoreBanditAlgorithmsBook}.
+% We give in Section~\ref{sec:2:famousMABalgorithms} an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focus on the $\UCB_1$, \klUCB{} and Thompson sampling algorithms, as they are used in the rest of the manuscript.
 % % We present SMPyBandits in the next Chapter~\ref{chapter:3}, a Python open-source library, designed to run numerical simulations of MAB problems, one of the main contributions of this thesis.
@@
 
-\textbf{About notations.}
-%
-We use in this thesis the usual notations from the MAB literature,
-% , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
-and a summary is included in the Nomenclature, included in the Appendix in Part~\ref{part:Conclusion} (Page~\pageref{chapter:nomenclature}).
-
-\textbf{Main references for the curious reader.}
-%
-For more details and a more formal introduction to multi-armed bandits, we suggest to the interested reader to work on a very recent text-book by Slivkins \cite{Slivkins2019}.
-Another excellent but reasonably short survey is the book by Bubeck and Cesa-Bianchi \cite{Bubeck12}, while the more recent book by Lattimore and Szepesv{\'a}ri \cite{LattimoreBanditAlgorithmsBook} is the most complete resource about bandit algorithms.
-Finally, we recommend \cite{bouneffouf2019survey} for a short but good survey on applications of MAB.
-
-\textbf{No publication associated with this chapter.}
-%
-% The first Sections~\ref{sec:2:notations}, \ref{sec:2:applicationsofStochasticMAB}, \ref{sec:2:lowerUpperBoundsRegret} and \ref{sec:2:famousMABalgorithms} of this chapter are based on existing works.
-This chapter is based on existing works.
-% and Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on our article \cite{Besson2018WCNC}.
+% \textbf{No publication associated with this chapter.}
+% %
+% % The first Sections~\ref{sec:2:notations}, \ref{sec:2:applicationsofStochasticMAB}, \ref{sec:2:lowerUpperBoundsRegret} and \ref{sec:2:famousMABalgorithms} of this chapter are based on existing works.
+% This chapter is based on existing works.
+% % and Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} in this chapter is based on our article \cite{Besson2018WCNC}.
 
@@
 % ----------------------------------------------------------------------------
-\section{The stochastic Multi-Armed Bandit models}
+\section{The stochastic Multi-Armed Bandit model}
 \label{sec:2:notations}
 
-A MAB is a model of a decision-making game, where a \emph{player} has to sequentially select one action in a (usually finite) set of actions, and only receives a (random) feedback about the selected action, also called a \emph{reward}.
+Multi-Armed Bandits (MAB) models were introduced by Thompson as early as in 1933 \cite{Thompson33}, and later studied from the 1950s by Robbins \cite{Robbins52} and others.
+This family of models was first proposed for clinical trials, and later applied to a wide range of different problems.
+Their name refers to one-armed bandits found in casinos, as shown in Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
+% The seminal paper by Lai and Robbins \cite{LaiRobbins85} was the first to introduce the measure of performance widely used in the literature, the notion of \emph{regret}, and to analyze a lower-bound on the performance of any algorithm against a fixed problem.
+% Since then, many efficient algorithms were proposed, XXX
+%
+% In all this thesis, we only consider one-dimensional exponential families, and in particular Bernoulli distributions.
+We now formally introduce the model.
+%  and the main notations used in all this document in Section~\ref{sec:2:notations}.
+
+A MAB refers to a decision-making game, where a \emph{player} has to sequentially select one action in a usually finite set of actions, and only receives a (random) feedback about the selected action, also called a \emph{reward}.
 \textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
-When the player decides at time $t$ to play (or pull) the arm $A(t)\in[K]$,
-where this notation denotes $[K] \eqdef \{1,\dots,K\}$, she receives a reward $r(t) \eqdef Y_{A(t),t}\in\R$, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
+We denote $[K]$ the set $\{1,\dots,K\}$.
+Each arm $k\in[K]$ is associated with a stream of rewards $(Y_{k,t})_{t\in\N^*}$,
+and when the player decides at time $t$ to play (or pull) the arm $A(t)\in[K]$,
+she receives a reward $r(t) \eqdef Y_{A(t),t}$.
+She keeps playing an arm and observing a reward iteratively, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
 %
@@
 
+\textbf{About notations.}
+%
+We use in this thesis the usual notations from the MAB literature,
+% , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
+and a summary is included in the Nomenclature, included in the Appendix in Part~\ref{part:Conclusion} (Page~\pageref{chapter:nomenclature}).
+
+\textbf{Main references for the curious reader.}
+%
+For more details and a more formal introduction to multi-armed bandits, we suggest to the interested reader to work on a very recent text-book by Slivkins \cite{Slivkins2019}.
+Another excellent but reasonably short survey is the book by Bubeck and Cesa-Bianchi \cite{Bubeck12}, while the more recent book by Lattimore and Szepesv{\'a}ri \cite{LattimoreBanditAlgorithmsBook} is the most complete resource about bandit algorithms.
+Finally, we recommend \cite{bouneffouf2019survey} for a short but good survey on applications of MAB.
+
 % ----------------------------------------------------------------------------
@@
 
-In all this thesis, we focus on the model with finite arms, and binary or real-valued stochastic rewards, that is $A(t)\in\{1,\dots,K\}$ for a fixed and known number of arms $K\in\N,K\geq2$, and $Y_{A(t),t}\in\R$.
+In all this thesis, we focus on the model with finite arms, that is $A(t)\in\{1,\dots,K\}$ for a fixed and known number of arms $K\in\N,K\geq2$.
+%
+We focus on the \textbf{stochastic} MAB, in which the reward stream are drawn from some (unknown) distributions.
 %
@@
 We also restrict to \emph{stochastic} rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in[K]$.
-Rewards are stationary, meaning that $(Y_{k,t})_{t\in\N^*}$ is independent and identically distributed (\iid), $Y_{k,t} \sim \nu_k$ for any $t\geq1$.
+Rewards are stationary, meaning that $(Y_{k,t})_{t\in\N^*}$ is independent and identically distributed (\iid), and $Y_{k,t} \sim \nu_k$ for any $t\geq1$.
+We only consider binary or real-valued rewards, \ie, $Y_{k,t}\in\R$.
 
@@
 
-% \paragraph{We restrict to one-dimensional exponential family.}
+\paragraph{One-dimensional exponential family.}
 %
@@
 % ----------------------------------------------------------------------------
-\section{Different applications of stochastic MAB}
+\section{Applications of stochastic MAB}
 \label{sec:2:applicationsofStochasticMAB}
@@
 We focus then on index policies, with a first efficient policy, $\UCB_1$, that uses upper confidence bounds (UCB) on the means estimates, and is known to be order-optimal for bounded rewards.
-Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of $\UCB_1$ but use smaller confidence intervals and thus it typically obtains better theoretical results, and Thompson sampling which replaces the principle of ``optimism under uncertainty'' by a Bayesian point-of-view.
+Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of $\UCB_1$ but use tighter confidence intervals and thus it typically obtains better theoretical results, and Thompson sampling which replaces the principle of ``optimism under uncertainty'' by a Bayesian point-of-view.
 %
@@
 % ----------------------------------------------------------------------------
-\subsection{Index policies: $\UCB_1$, \klUCB{} and others}
+\subsection{The optimism principle and index policies: $\UCB_1$, \klUCB{} etc}
 \label{sub:2:IndexPolicies}
 
-A large family of algorithms are index based, as they compute an index $U_k(t)$ on each arm $k$ at time $t$.
-The indexes should represent the expected quality of each arm, thus index policies play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
+A large family of algorithms are index based, as they compute an index $U_k(t)$ on each arm $k$ at time $t$,
+used to play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
 If more than one indexes maximize $(U_k(t))_k$, the arm is chosen among the set, usually in a uniformly random manner: $A(t) \sim \cU(\argmax_k U_k(t))$.
@@
 %
-This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}.
+This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}, for bounded rewards.
 For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
@@
     \begin{equation}\label{eq:2:UCBregretBound}
-        R_T^{\UCB_1}(I) \leq \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left(1 + \frac{\pi^2}{3}\right)
-        = \bigO{ \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T)}.
+        R_T^{\UCB_1}(I) \leq \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left(1 + \frac{\pi^2}{3}\right).
+        % = \bigO{ \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T)}.
     \end{equation}
@@
 and $f(t) \eqdef \log(t)$ in practice.
-The constant $c>0$ is usually chosen as $c = 1$.
+% The constant $c>0$ is usually chosen as $c = 1$.
 \begin{align}\label{eq:2:klUCB_index}
@@
 
-For Bernoulli distributions, the previous claim that \klUCB{} builds smaller confidence intervals (\ie, more accurate) than $\UCB_1$, \ie, $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$,
-is justified by Pinkser's inequality which gives $\kl(x,y) \geq 2(p-q)^2$.
+For Bernoulli distributions, \klUCB{} builds tighter confidence intervals than $\UCB_1$, in the sense that $U_k^{\klUCB}(t) < U_k^{\UCB}(t)$
+but with the same probability coverage.
+This is justified by Pinkser's inequality, which gives $\kl(x,y) \geq 2(x-y)^2$.
 %

### 2019-06-13 2-Chapters/2-Chapter/chapter25.tex
SO MANY CHANGES

@@
 %
+We gave in Section~\ref{sec:2:famousMABalgorithms} an overview of the main families of algorithms designed to tackle stochastic MAB problems, and we mainly focused on the $\UCB_1$, \klUCB{} and Thompson sampling algorithms.
+% as they are used in the rest of the manuscript.
+But a rich research literature has produced many different algorithms tackling the MAB problem, as suggested for instance by the length of the reference book \cite{LattimoreBanditAlgorithmsBook}.
+%
 Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to use.
@@
 
-In this Section, we first present two naive approaches for selecting an algorithm when facing a new problem, and then we detail the online approach that uses a ``leader'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``leader'' algorithm based on \ExpQ.
+In this chapter, we first present two naive approaches for selecting an algorithm when facing a new problem, and then we detail the online approach that uses a ``leader'' MAB algorithm running on top of a pool of ``followers'' algorithms, and we present our contribution that is a new ``leader'' algorithm based on \ExpQ.
 

### 2019-06-14 2-Chapters/2-Chapter/chapter25.tex
Starting to add the new figures

@@
 This chapter is organized as follows.
-First, we explain in Section~\ref{sub:25:aggregation} how to combine such algorithms for aggregation.
-Then we present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
+First, we give more motivations for online algorithm selection in the rest of this section, then we explain how to combine such stationary MAB algorithms and aggregate them.
+We present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
 Finally, we present numerical experiments in Section~\ref{sub:25:numExp},
@@
 comparing the regret of several algorithms against different aggregation algorithms.
-Theoretical guarantees are shortly discussed in Section~\ref{sub:25:theory}, and Section~\ref{sec:25:conclusion} concludes.
+Theoretical guarantees are shortly discussed, before we conclude in Section~\ref{sec:25:conclusion}.
 

### 2019-06-14 2-Chapters/3-Chapter/chapter3.tex
Starting to add the new figures

@@
     %  Algorithm 1 is
-    $\varepsilon$-greedy \cite{Bubeck12} (\texttt{EpsilonDecreasing}), shown in Algorithm~\ref{algo:2:naiveStrategies} above, using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily).
+    $\varepsilon$-greedy \cite{Bubeck12} (\texttt{EpsilonDecreasing}),
+    % shown in Algorithm~\ref{algo:2:naiveStrategies} above,
+    using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily).
     It has a very small cost in terms of time and memory, but it usually achieves linear regret.
@@
     %  Algorithm 2 is
-    Explore-then-Commit \cite{GarivierETC2016} (\texttt{ETC\_KnownGap}), shown in Algorithm~\ref{algo:2:naiveStrategies} above, that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems).
+    Explore-then-Commit \cite{GarivierETC2016} (\texttt{ETC\_KnownGap}),
+    % shown in Algorithm~\ref{algo:2:naiveStrategies} above,
+    that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems).
     Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but most of the times it only obtains large (linear) regret.
@@
     %  Algorithm 5 is
-    \klUCB{} from \cite{KLUCBJournal} (\texttt{klUCB}), also shown in Algorithm~\ref{algo:2:indexPolicy} above, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes, implemented as \texttt{kullback.klucbBern}.
+    \klUCB{} from \cite{KLUCBJournal} (\texttt{klUCB}), also shown in Algorithm~\ref{algo:2:indexPolicy} above, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes (coded as \texttt{kullback.klucbBern}).
     % It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret.
@@
 
-% \begin{table}[!t]
-% \begin{footnotesize}  % WARNING
-%     \centering
-%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-%         \hline
-%         $\varepsilon$-greedy &
-%             $R^{1}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{1}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{1}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{1}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{1}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{1}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{1}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{1}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{1}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{1}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{1}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{1}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{1}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{1}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{1}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{1}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{1}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{1}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{1}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{1}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         Explore-then-Commit &
-%             $R^{2}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{2}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{2}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{2}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{2}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{2}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{2}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{2}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{2}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{2}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{2}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{2}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{2}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{2}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{2}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{2}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{2}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{2}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{2}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{2}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         $\mathrm{Exp}3^{++}$ &
-%             $R^3_{T_1,K_1} = 0 \pm 0$
-%                 $R^3_{T_1,K_2} = 0 \pm 0$
-%                 $R^3_{T_1,K_3} = 0 \pm 0$
-%                 $R^3_{T_1,K_4} = 0 \pm 0$
-%                 $R^3_{T_1,K_5} = 0 \pm 0$ &
-%             $R^3_{T_2,K_1} = 0 \pm 0$
-%                 $R^3_{T_2,K_2} = 0 \pm 0$
-%                 $R^3_{T_2,K_3} = 0 \pm 0$
-%                 $R^3_{T_2,K_4} = 0 \pm 0$
-%                 $R^3_{T_2,K_5} = 0 \pm 0$ &
-%             $R^3_{T_3,K_1} = 0 \pm 0$
-%                 $R^3_{T_3,K_2} = 0 \pm 0$
-%                 $R^3_{T_3,K_3} = 0 \pm 0$
-%                 $R^3_{T_3,K_4} = 0 \pm 0$
-%                 $R^3_{T_3,K_5} = 0 \pm 0$ &
-%             $R^3_{T_4,K_1} = 0 \pm 0$
-%                 $R^3_{T_4,K_2} = 0 \pm 0$
-%                 $R^3_{T_4,K_3} = 0 \pm 0$
-%                 $R^3_{T_4,K_4} = 0 \pm 0$
-%                 $R^3_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         $\UCB_1$ &
-%             $R^{4}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{4}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{4}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{4}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{4}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{4}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{4}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{4}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{4}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{4}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{4}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{4}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{4}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{4}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{4}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{4}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{4}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{4}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{4}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{4}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         \klUCB{} &
-%             $R^{5}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{5}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{5}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{5}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{5}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{5}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{5}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{5}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{5}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{5}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{5}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{5}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{5}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{5}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{5}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{5}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{5}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{5}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{5}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{5}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         Thompson sampling &
-%             $R^{6}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{6}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{6}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{6}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{6}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{6}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{6}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{6}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{6}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{6}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{6}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{6}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{6}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{6}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{6}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{6}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{6}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{6}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{6}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{6}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         Thompson sampling &
-%             $R^{7}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{7}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{7}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{7}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{7}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{7}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{7}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{7}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{7}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{7}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{7}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{7}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{7}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{7}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{7}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{7}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{7}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{7}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{7}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{7}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         AdBandits &
-%             $R^{8}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{8}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{8}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{8}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{8}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{8}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{8}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{8}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{8}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{8}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{8}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{8}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{8}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{8}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{8}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{8}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{8}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{8}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{8}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{8}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         BESA &
-%             $R^{9}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{9}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{9}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{9}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{9}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{9}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{9}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{9}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{9}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{9}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{9}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{9}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{9}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{9}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{9}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{9}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{9}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{9}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{9}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{9}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%     \end{tabular}
-%     \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 1} with different values of time $T$ and number of arm $K$.
-%     }
-%     \label{table:3:meanRegret_problem1}
-% \end{footnotesize}  % WARNING
-% \end{table}
-
-% \begin{table}[!t]
-% \begin{footnotesize}  % WARNING
-%     \centering
-%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-%         \hline
-%         $\varepsilon$-greedy &
-%             $R^{'1}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'1}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'1}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'1}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'1}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'1}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'1}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'1}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'1}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'1}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'1}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'1}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'1}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'1}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'1}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'1}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'1}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'1}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'1}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'1}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         Explore-then-Commit &
-%             $R^{'2}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'2}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'2}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'2}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'2}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'2}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'2}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'2}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'2}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'2}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'2}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'2}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'2}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'2}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'2}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'2}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'2}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'2}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'2}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'2}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         $\mathrm{Exp}3^{++}$ &
-%             $R^{'3}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'3}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'3}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'3}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'3}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'3}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'3}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'3}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'3}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'3}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'3}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'3}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'3}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'3}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'3}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'3}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'3}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'3}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'3}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'3}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         $\UCB_1$ &
-%             $R^{'4}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'4}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'4}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'4}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'4}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'4}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'4}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'4}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'4}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'4}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'4}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'4}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'4}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'4}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'4}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'4}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'4}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'4}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'4}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'4}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         \klUCB{} &
-%             $R^{'5}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'5}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'5}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'5}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'5}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'5}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'5}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'5}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'5}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'5}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'5}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'5}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'5}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'5}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'5}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'5}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'5}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'5}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'5}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'5}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         Thompson sampling &
-%             $R^{'6}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'6}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'6}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'6}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'6}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'6}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'6}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'6}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'6}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'6}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'6}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'6}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'6}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'6}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'6}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'6}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'6}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'6}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'6}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'6}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         Bayes-UCB &
-%             $R^{'7}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'7}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'7}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'7}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'7}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'7}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'7}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'7}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'7}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'7}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'7}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'7}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'7}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'7}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'7}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'7}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'7}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'7}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'7}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'7}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         AdBandits &
-%             $R^{'8}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'8}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'8}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'8}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'8}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'8}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'8}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'8}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'8}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'8}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'8}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'8}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'8}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'8}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'8}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'8}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'8}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'8}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'8}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'8}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%         BESA &
-%             $R^{'9}_{T_1,K_1} = 0 \pm 0$
-%                 $R^{'9}_{T_1,K_2} = 0 \pm 0$
-%                 $R^{'9}_{T_1,K_3} = 0 \pm 0$
-%                 $R^{'9}_{T_1,K_4} = 0 \pm 0$
-%                 $R^{'9}_{T_1,K_5} = 0 \pm 0$ &
-%             $R^{'9}_{T_2,K_1} = 0 \pm 0$
-%                 $R^{'9}_{T_2,K_2} = 0 \pm 0$
-%                 $R^{'9}_{T_2,K_3} = 0 \pm 0$
-%                 $R^{'9}_{T_2,K_4} = 0 \pm 0$
-%                 $R^{'9}_{T_2,K_5} = 0 \pm 0$ &
-%             $R^{'9}_{T_3,K_1} = 0 \pm 0$
-%                 $R^{'9}_{T_3,K_2} = 0 \pm 0$
-%                 $R^{'9}_{T_3,K_3} = 0 \pm 0$
-%                 $R^{'9}_{T_3,K_4} = 0 \pm 0$
-%                 $R^{'9}_{T_3,K_5} = 0 \pm 0$ &
-%             $R^{'9}_{T_4,K_1} = 0 \pm 0$
-%                 $R^{'9}_{T_4,K_2} = 0 \pm 0$
-%                 $R^{'9}_{T_4,K_3} = 0 \pm 0$
-%                 $R^{'9}_{T_4,K_4} = 0 \pm 0$
-%                 $R^{'9}_{T_4,K_5} = 0 \pm 0$ \\
-%         \hline
-%     \end{tabular}
-%     \caption{Mean regret $\pm$ $1$ std-dev, for $9$ different algorithms on \textbf{problem 2} with different values of time $T$ and number of arm $K$.
-%     }
-%     \label{table:3:meanRegret_problem2}
-% \end{footnotesize}  % WARNING
-% \end{table}
-
-% Include in appendix of this chapter the results for 7 more algorithms, the most recent and less important ones (MOSSAnytime, ApproximatedFHGittins, UCBdagger, klUCBswitchAnytime, PHE, RCB, Tsallis-Inf)
-
-% FIXME generate from the notebook then copy from WS3
 \begin{figure}[h!]  % [htbp]
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_arms__56pb__7Ks_T50000.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_arms__56pb__7Ks_T50000.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_arms__16pb__T5000.pdf}
 	\caption[Regret vs different values of $K$.]{
         Regret vs different values of $K$ ($R_T / \log(T)$),
-        for $T=50000$ and for $16$ algorithms.
+        for $T=5000$ and for $16$ algorithms.
         The $y$-axis is in log-scale.
@@
 	}
-	\label{fig:3:16_different_algorithms__lognormregret_vs_arms__56pb__7Ks_T50000}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_arms__16pb__T5000}
 \end{figure}
@@
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normregret_vs_horizons__56pb__K32_7Ts.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__normregret_vs_horizons__56pb__K32_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_times__16pb__K8.pdf}
 	\caption[Regret vs different values of $T$.]{
@@
 	}
-	\label{fig:3:16_different_algorithms__normregret_vs_horizons__56pb__K32_7Ts}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_times__16pb__K8}
 \end{figure}
 
-\subsection{Summary of the experiments}
-
-\TODOL{Au lieu d'avoir des longues captions pour les figures, je vais écrire quelques paragraphes donnant les mêmes conclusion et faisant référence aux figures.
-Je ferai ça très vite si on est contents des figures.}
-
-We have been able to check empirically that the regret of all the efficient algorithm indeed scales as predicted by the theory, that is
+\paragraph{Summary of the experiments.}
+%
+In the two Figures~\ref{fig:3:16_different_algorithms__lognormregret_vs_arms__16pb__T5000} and \ref{fig:3:16_different_algorithms__lognormregret_vs_times__16pb__K8}
+included in this section,
+we are able to check empirically that the regret of all the efficient algorithm indeed scales as predicted by the theory, that is
 linearly in the number of arms, and logarithmically in the horizon, \ie, $R_T = \bigO{K \log T}$.
@@
     \item
-    \UCB{} is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
-    \item
-    \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when efficiency is favored.
+    \UCB{} (or Thompson sampling) is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
+    \item
+    \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when efficiency is favored, and when we analyze mathematically the performance of an algorithm that is running on top of a classical stationary MAB policy, like our two contributions, \MCTopM{} and \GLR.
 \end{itemize}
-
-% I want to show that SMPyBandits contains highly documented implementations of all the most important algorithms designed for stationary single-player multi-armed bandit.
-
-% I will include here formulas about the main algorithms (klUCB, MOSS, etc).
-% I will pick some problems (Bernoulli, Gaussian, others), with fixed means or ``Bayesian'' means (\ie, changing at each time) and I will run extensive numerical simulations to compare the main algorithms.
-
-% The take away message is: klUCB beats UCB, Thompson sampling is great too, all the others algorithms are less efficient or comparable, so we only use klUCB for all the rest of this thesis.
-
-% I will include graphs showing the mean regret $R_t$ as a function of time $t$, as well as histograms or boxplots showing the distribution of $R_T$.
 
@@
 \label{sec:3:timeAndMemoryCosts}
-
-\TODOL{Tout ce blabla est un peu trop long, je devrais couper et être plus concis !}
 
@@
 
-% \subsection{Computational time}
-
-% \TODOL{Just one long table for problem 1, not necessary to include other problems}
+\paragraph{Computational time.}
+
+In Figure~\ref{fig:3:16_different_algorithms__lognormregret_vs_logtime__28pb}, we can see FIXME
+
+Moreover, in the two Figures~\ref{fig:3:16_different_algorithms__lognormtime_vs_arms__16pb__T5000} and
+\ref{fig:3:16_different_algorithms__lognormtime_vs_times__16pb__K8},
+we are able to check empirically that the computational time of all the efficient algorithm indeed scales as predicted by the theory, that is
+linearly in the number of arms in the horizon, \ie, $\cT_T = \bigO{KT}$.
 
@@
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormtime__56pb__7Ks_7Ts.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormtime__56pb__7Ks_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_logtime__28pb.pdf}
 	\caption[Normalized mean regret vs normalized running time (in micro-seconds).]{
@@
 	}
-	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormtime__56pb__7Ks_7Ts}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_logtime__28pb}
 \end{figure}
 
-% FIXME generate from the notebook then copy from WS3
 \begin{figure}[h!]  % [htbp]
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormtime_vs_arms__56pb__7Ks_T50000.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormtime_vs_arms__56pb__7Ks_T50000.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormtime_vs_arms__16pb__T5000.pdf}
 	\caption[Normalized running time vs different values of $K$.]{
         Normalized running time vs different values of $K$ ($\cT_T / T$),
-        for $T=50000$ and for $16$ algorithms.
+        for $T=5000$ and for $16$ algorithms.
         $y$-axis is in log-scale.
-        All algorithms except BESA has a linear normalized running time, meaning that for $K$ arms they use a computation time proportional to $K$, as predicted.
+        All algorithms except BESA has a linear normalized running time, meaning that for $K$ arms they use a computation time proportional to $K$, as predicted: $\cT_T = \bigO{KT}$.
 	}
-	\label{fig:3:16_different_algorithms__lognormtime_vs_arms__56pb__7Ks_T50000}
+	\label{fig:3:16_different_algorithms__lognormtime_vs_arms__16pb__T5000}
 \end{figure}
@@
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normtime_vs_horizons__56pb__K32_7Ts.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__normtime_vs_horizons__56pb__K32_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormtime_vs_times__16pb__K8.pdf}
 	\caption[Normalized running time vs different values of $T$.]{
@@
         for $K=32$ and for $16$ algorithms.
-        All algorithms except BESA has a constant normalized running time, meaning that for $T$ rounds they use a computation time proportional to $T$, as predicted.
+        $y$-axis is in log-scale.
+        All algorithms except BESA has a constant normalized running time, meaning that for $T$ rounds they use a computation time proportional to $T$, as predicted: $\cT_T = \bigO{KT}$.
 	}
-	\label{fig:3:16_different_algorithms__normtime_vs_horizons__56pb__K32_7Ts}
+	\label{fig:3:16_different_algorithms__lognormtime_vs_times__16pb__K8}
 \end{figure}
 
-% \begin{table}[!t]
-% \begin{footnotesize}  % WARNING
-%     \centering
-%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-%         \hline
-%         $\varepsilon$-greedy &
-%             $\cT^{1}_{T_1,K_1} = 0$
-%                 $\cT^{1}_{T_1,K_2} = 0$
-%                 $\cT^{1}_{T_1,K_3} = 0$
-%                 $\cT^{1}_{T_1,K_4} = 0$
-%                 $\cT^{1}_{T_1,K_5} = 0$ &
-%             $\cT^{1}_{T_2,K_1} = 0$
-%                 $\cT^{1}_{T_2,K_2} = 0$
-%                 $\cT^{1}_{T_2,K_3} = 0$
-%                 $\cT^{1}_{T_2,K_4} = 0$
-%                 $\cT^{1}_{T_2,K_5} = 0$ &
-%             $\cT^{1}_{T_3,K_1} = 0$
-%                 $\cT^{1}_{T_3,K_2} = 0$
-%                 $\cT^{1}_{T_3,K_3} = 0$
-%                 $\cT^{1}_{T_3,K_4} = 0$
-%                 $\cT^{1}_{T_3,K_5} = 0$ &
-%             $\cT^{1}_{T_4,K_1} = 0$
-%                 $\cT^{1}_{T_4,K_2} = 0$
-%                 $\cT^{1}_{T_4,K_3} = 0$
-%                 $\cT^{1}_{T_4,K_4} = 0$
-%                 $\cT^{1}_{T_4,K_5} = 0$ \\
-%         \hline
-%         Explore-then-Commit &
-%             $\cT^{2}_{T_1,K_1} = 0$
-%                 $\cT^{2}_{T_1,K_2} = 0$
-%                 $\cT^{2}_{T_1,K_3} = 0$
-%                 $\cT^{2}_{T_1,K_4} = 0$
-%                 $\cT^{2}_{T_1,K_5} = 0$ &
-%             $\cT^{2}_{T_2,K_1} = 0$
-%                 $\cT^{2}_{T_2,K_2} = 0$
-%                 $\cT^{2}_{T_2,K_3} = 0$
-%                 $\cT^{2}_{T_2,K_4} = 0$
-%                 $\cT^{2}_{T_2,K_5} = 0$ &
-%             $\cT^{2}_{T_3,K_1} = 0$
-%                 $\cT^{2}_{T_3,K_2} = 0$
-%                 $\cT^{2}_{T_3,K_3} = 0$
-%                 $\cT^{2}_{T_3,K_4} = 0$
-%                 $\cT^{2}_{T_3,K_5} = 0$ &
-%             $\cT^{2}_{T_4,K_1} = 0$
-%                 $\cT^{2}_{T_4,K_2} = 0$
-%                 $\cT^{2}_{T_4,K_3} = 0$
-%                 $\cT^{2}_{T_4,K_4} = 0$
-%                 $\cT^{2}_{T_4,K_5} = 0$ \\
-%         \hline
-%         $\mathrm{Exp}3^{++}$ &
-%             $\cT^3_{T_1,K_1} = 0$
-%                 $\cT^3_{T_1,K_2} = 0$
-%                 $\cT^3_{T_1,K_3} = 0$
-%                 $\cT^3_{T_1,K_4} = 0$
-%                 $\cT^3_{T_1,K_5} = 0$ &
-%             $\cT^3_{T_2,K_1} = 0$
-%                 $\cT^3_{T_2,K_2} = 0$
-%                 $\cT^3_{T_2,K_3} = 0$
-%                 $\cT^3_{T_2,K_4} = 0$
-%                 $\cT^3_{T_2,K_5} = 0$ &
-%             $\cT^3_{T_3,K_1} = 0$
-%                 $\cT^3_{T_3,K_2} = 0$
-%                 $\cT^3_{T_3,K_3} = 0$
-%                 $\cT^3_{T_3,K_4} = 0$
-%                 $\cT^3_{T_3,K_5} = 0$ &
-%             $\cT^3_{T_4,K_1} = 0$
-%                 $\cT^3_{T_4,K_2} = 0$
-%                 $\cT^3_{T_4,K_3} = 0$
-%                 $\cT^3_{T_4,K_4} = 0$
-%                 $\cT^3_{T_4,K_5} = 0$ \\
-%         \hline
-%         $\UCB_1$ &
-%             $\cT^{4}_{T_1,K_1} = 0$
-%                 $\cT^{4}_{T_1,K_2} = 0$
-%                 $\cT^{4}_{T_1,K_3} = 0$
-%                 $\cT^{4}_{T_1,K_4} = 0$
-%                 $\cT^{4}_{T_1,K_5} = 0$ &
-%             $\cT^{4}_{T_2,K_1} = 0$
-%                 $\cT^{4}_{T_2,K_2} = 0$
-%                 $\cT^{4}_{T_2,K_3} = 0$
-%                 $\cT^{4}_{T_2,K_4} = 0$
-%                 $\cT^{4}_{T_2,K_5} = 0$ &
-%             $\cT^{4}_{T_3,K_1} = 0$
-%                 $\cT^{4}_{T_3,K_2} = 0$
-%                 $\cT^{4}_{T_3,K_3} = 0$
-%                 $\cT^{4}_{T_3,K_4} = 0$
-%                 $\cT^{4}_{T_3,K_5} = 0$ &
-%             $\cT^{4}_{T_4,K_1} = 0$
-%                 $\cT^{4}_{T_4,K_2} = 0$
-%                 $\cT^{4}_{T_4,K_3} = 0$
-%                 $\cT^{4}_{T_4,K_4} = 0$
-%                 $\cT^{4}_{T_4,K_5} = 0$ \\
-%         \hline
-%         \klUCB{} &
-%             $\cT^{5}_{T_1,K_1} = 0$
-%                 $\cT^{5}_{T_1,K_2} = 0$
-%                 $\cT^{5}_{T_1,K_3} = 0$
-%                 $\cT^{5}_{T_1,K_4} = 0$
-%                 $\cT^{5}_{T_1,K_5} = 0$ &
-%             $\cT^{5}_{T_2,K_1} = 0$
-%                 $\cT^{5}_{T_2,K_2} = 0$
-%                 $\cT^{5}_{T_2,K_3} = 0$
-%                 $\cT^{5}_{T_2,K_4} = 0$
-%                 $\cT^{5}_{T_2,K_5} = 0$ &
-%             $\cT^{5}_{T_3,K_1} = 0$
-%                 $\cT^{5}_{T_3,K_2} = 0$
-%                 $\cT^{5}_{T_3,K_3} = 0$
-%                 $\cT^{5}_{T_3,K_4} = 0$
-%                 $\cT^{5}_{T_3,K_5} = 0$ &
-%             $\cT^{5}_{T_4,K_1} = 0$
-%                 $\cT^{5}_{T_4,K_2} = 0$
-%                 $\cT^{5}_{T_4,K_3} = 0$
-%                 $\cT^{5}_{T_4,K_4} = 0$
-%                 $\cT^{5}_{T_4,K_5} = 0$ \\
-%         \hline
-%         Thompson sampling &
-%             $\cT^{6}_{T_1,K_1} = 0$
-%                 $\cT^{6}_{T_1,K_2} = 0$
-%                 $\cT^{6}_{T_1,K_3} = 0$
-%                 $\cT^{6}_{T_1,K_4} = 0$
-%                 $\cT^{6}_{T_1,K_5} = 0$ &
-%             $\cT^{6}_{T_2,K_1} = 0$
-%                 $\cT^{6}_{T_2,K_2} = 0$
-%                 $\cT^{6}_{T_2,K_3} = 0$
-%                 $\cT^{6}_{T_2,K_4} = 0$
-%                 $\cT^{6}_{T_2,K_5} = 0$ &
-%             $\cT^{6}_{T_3,K_1} = 0$
-%                 $\cT^{6}_{T_3,K_2} = 0$
-%                 $\cT^{6}_{T_3,K_3} = 0$
-%                 $\cT^{6}_{T_3,K_4} = 0$
-%                 $\cT^{6}_{T_3,K_5} = 0$ &
-%             $\cT^{6}_{T_4,K_1} = 0$
-%                 $\cT^{6}_{T_4,K_2} = 0$
-%                 $\cT^{6}_{T_4,K_3} = 0$
-%                 $\cT^{6}_{T_4,K_4} = 0$
-%                 $\cT^{6}_{T_4,K_5} = 0$ \\
-%         \hline
-%         Thompson sampling &
-%             $\cT^{7}_{T_1,K_1} = 0$
-%                 $\cT^{7}_{T_1,K_2} = 0$
-%                 $\cT^{7}_{T_1,K_3} = 0$
-%                 $\cT^{7}_{T_1,K_4} = 0$
-%                 $\cT^{7}_{T_1,K_5} = 0$ &
-%             $\cT^{7}_{T_2,K_1} = 0$
-%                 $\cT^{7}_{T_2,K_2} = 0$
-%                 $\cT^{7}_{T_2,K_3} = 0$
-%                 $\cT^{7}_{T_2,K_4} = 0$
-%                 $\cT^{7}_{T_2,K_5} = 0$ &
-%             $\cT^{7}_{T_3,K_1} = 0$
-%                 $\cT^{7}_{T_3,K_2} = 0$
-%                 $\cT^{7}_{T_3,K_3} = 0$
-%                 $\cT^{7}_{T_3,K_4} = 0$
-%                 $\cT^{7}_{T_3,K_5} = 0$ &
-%             $\cT^{7}_{T_4,K_1} = 0$
-%                 $\cT^{7}_{T_4,K_2} = 0$
-%                 $\cT^{7}_{T_4,K_3} = 0$
-%                 $\cT^{7}_{T_4,K_4} = 0$
-%                 $\cT^{7}_{T_4,K_5} = 0$ \\
-%         \hline
-%         AdBandits &
-%             $\cT^{8}_{T_1,K_1} = 0$
-%                 $\cT^{8}_{T_1,K_2} = 0$
-%                 $\cT^{8}_{T_1,K_3} = 0$
-%                 $\cT^{8}_{T_1,K_4} = 0$
-%                 $\cT^{8}_{T_1,K_5} = 0$ &
-%             $\cT^{8}_{T_2,K_1} = 0$
-%                 $\cT^{8}_{T_2,K_2} = 0$
-%                 $\cT^{8}_{T_2,K_3} = 0$
-%                 $\cT^{8}_{T_2,K_4} = 0$
-%                 $\cT^{8}_{T_2,K_5} = 0$ &
-%             $\cT^{8}_{T_3,K_1} = 0$
-%                 $\cT^{8}_{T_3,K_2} = 0$
-%                 $\cT^{8}_{T_3,K_3} = 0$
-%                 $\cT^{8}_{T_3,K_4} = 0$
-%                 $\cT^{8}_{T_3,K_5} = 0$ &
-%             $\cT^{8}_{T_4,K_1} = 0$
-%                 $\cT^{8}_{T_4,K_2} = 0$
-%                 $\cT^{8}_{T_4,K_3} = 0$
-%                 $\cT^{8}_{T_4,K_4} = 0$
-%                 $\cT^{8}_{T_4,K_5} = 0$ \\
-%         \hline
-%         BESA &
-%             $\cT^{9}_{T_1,K_1} = 0$
-%                 $\cT^{9}_{T_1,K_2} = 0$
-%                 $\cT^{9}_{T_1,K_3} = 0$
-%                 $\cT^{9}_{T_1,K_4} = 0$
-%                 $\cT^{9}_{T_1,K_5} = 0$ &
-%             $\cT^{9}_{T_2,K_1} = 0$
-%                 $\cT^{9}_{T_2,K_2} = 0$
-%                 $\cT^{9}_{T_2,K_3} = 0$
-%                 $\cT^{9}_{T_2,K_4} = 0$
-%                 $\cT^{9}_{T_2,K_5} = 0$ &
-%             $\cT^{9}_{T_3,K_1} = 0$
-%                 $\cT^{9}_{T_3,K_2} = 0$
-%                 $\cT^{9}_{T_3,K_3} = 0$
-%                 $\cT^{9}_{T_3,K_4} = 0$
-%                 $\cT^{9}_{T_3,K_5} = 0$ &
-%             $\cT^{9}_{T_4,K_1} = 0$
-%                 $\cT^{9}_{T_4,K_2} = 0$
-%                 $\cT^{9}_{T_4,K_3} = 0$
-%                 $\cT^{9}_{T_4,K_4} = 0$
-%                 $\cT^{9}_{T_4,K_5} = 0$ \\
-%         \hline
-%     \end{tabular}
-%     \caption{Normalized running time, rounded and in milli-seconds, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
-%     \label{table:3:time_problem1}
-% \end{footnotesize}  % WARNING
-% \end{table}
-
-% \subsection{Memory cost}
-
-% \TODOL{Just one long table for problem 1, not necessary to include other problems}
+\paragraph{Memory cost.}
+
+In Figure~\ref{fig:3:16_different_algorithms__lognormregret_vs_logmemory__28pb}, we can see FIXME
+
+Moreover, in the two Figures~\ref{fig:3:16_different_algorithms__lognormmemory_vs_arms__16pb__T5000} and
+\ref{fig:3:16_different_algorithms__lognormmemory_vs_times__16pb__K8},
+we are able to check empirically that the memory cost of all the efficient algorithm indeed scales as predicted by the theory, that is
+linearly in the number of arms but independently of the horizon, \ie, $\cM_T = \bigO{K}$.
 
@@
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7Ks_7Ts.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7Ks_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormregret_vs_logmemory__28pb.pdf}
 	\caption[Normalized mean regret vs normalized memory costs (in bytes).]{
@@
 	}
-	\label{fig:3:16_different_algorithms__lognormregret_vs_lognormmemory__56pb__7Ks_7Ts}
+	\label{fig:3:16_different_algorithms__lognormregret_vs_logmemory__28pb}
 \end{figure}
 
-% FIXME generate from the notebook then copy from WS3
 \begin{figure}[h!]  % [htbp]
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normmemory_vs_arms__56pb__7Ks_T50000.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__normmemory_vs_arms__56pb__7Ks_T50000.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormmemory_vs_arms__16pb__T5000.pdf}
 	\caption[Normalized memory cost vs different values of $K$.]{
         Normalized memory cost vs different values of $K$ ($\cM_T / K$),
-        for $T=50000$ and for $16$ algorithms.
-        % $y$-axis is in log-scale.
-        All algorithms (except BESA) has an almost constant memory cost, meaning that for $K$ arms they use a storage proportional to $K$, as predicted.
+        for $T=5000$ and for $16$ algorithms.
+        $y$-axis is in log-scale.
+        All algorithms (except BESA) has an almost constant memory cost, meaning that for $K$ arms they use a storage proportional to $K$, as predicted: $\cM_T = \bigO{K}$.
 	}
-	\label{fig:3:16_different_algorithms__normmemory_vs_arms__56pb__7Ks_T50000}
+	\label{fig:3:16_different_algorithms__lognormmemory_vs_arms__16pb__T5000}
 \end{figure}
@@
 	% \centering
-	\includegraphics[width=1.10\linewidth]{16_different_algorithms__normmemory_vs_horizons__56pb__K32_7Ts.pdf}
+	% \includegraphics[width=1.10\linewidth]{16_different_algorithms__normmemory_vs_horizons__56pb__K32_7Ts.pdf}
+	\includegraphics[width=1.10\linewidth]{16_different_algorithms__lognormmemory_vs_times__16pb__K8.pdf}
 	\caption[Normalized memory cost vs different values of $T$.]{
@@
         for $K=32$ and for $16$ algorithms.
-        All algorithms (except BESA) has a very slowly growing normalized memory cost, meaning that for $T$ rounds they use a computation time almost independent of $T$, as predicted.
+        $y$-axis is in log-scale.
+        All algorithms (except BESA) has a very slowly growing normalized memory cost, meaning that for $T$ rounds they use a computation time almost independent of $T$, as predicted: $\cM_T = \bigO{K}$.
         There is a clear separation between three groups.
@@
 	}
-	\label{fig:3:16_different_algorithms__normmemory_vs_horizons__56pb__K32_7Ts}
+	\label{fig:3:16_different_algorithms__lognormmemory_vs_times__16pb__K8}
 \end{figure}
 
-% \begin{table}[!t]
-% \begin{footnotesize}  % WARNING
-%     \centering
-%     \begin{tabular}{c|*{5}{m{2cm}}} % WARNING size ?
-%     \textbf{Algorithms} $\;$ \textbackslash $\;$ $\mathbf{T=}$
-%         & $T_1 = 1000$ & $T_2 = 5000$ & $T_3 = 10000$ & $T_4 = 50000$ \\
-%         \hline
-%         $\varepsilon$-greedy &
-%             $\cM^{1}_{T_1,K_1} = 0$
-%                 $\cM^{1}_{T_1,K_2} = 0$
-%                 $\cM^{1}_{T_1,K_3} = 0$
-%                 $\cM^{1}_{T_1,K_4} = 0$
-%                 $\cM^{1}_{T_1,K_5} = 0$ &
-%             $\cM^{1}_{T_2,K_1} = 0$
-%                 $\cM^{1}_{T_2,K_2} = 0$
-%                 $\cM^{1}_{T_2,K_3} = 0$
-%                 $\cM^{1}_{T_2,K_4} = 0$
-%                 $\cM^{1}_{T_2,K_5} = 0$ &
-%             $\cM^{1}_{T_3,K_1} = 0$
-%                 $\cM^{1}_{T_3,K_2} = 0$
-%                 $\cM^{1}_{T_3,K_3} = 0$
-%                 $\cM^{1}_{T_3,K_4} = 0$
-%                 $\cM^{1}_{T_3,K_5} = 0$ &
-%             $\cM^{1}_{T_4,K_1} = 0$
-%                 $\cM^{1}_{T_4,K_2} = 0$
-%                 $\cM^{1}_{T_4,K_3} = 0$
-%                 $\cM^{1}_{T_4,K_4} = 0$
-%                 $\cM^{1}_{T_4,K_5} = 0$ \\
-%         \hline
-%         Explore-then-Commit &
-%             $\cM^{2}_{T_1,K_1} = 0$
-%                 $\cM^{2}_{T_1,K_2} = 0$
-%                 $\cM^{2}_{T_1,K_3} = 0$
-%                 $\cM^{2}_{T_1,K_4} = 0$
-%                 $\cM^{2}_{T_1,K_5} = 0$ &
-%             $\cM^{2}_{T_2,K_1} = 0$
-%                 $\cM^{2}_{T_2,K_2} = 0$
-%                 $\cM^{2}_{T_2,K_3} = 0$
-%                 $\cM^{2}_{T_2,K_4} = 0$
-%                 $\cM^{2}_{T_2,K_5} = 0$ &
-%             $\cM^{2}_{T_3,K_1} = 0$
-%                 $\cM^{2}_{T_3,K_2} = 0$
-%                 $\cM^{2}_{T_3,K_3} = 0$
-%                 $\cM^{2}_{T_3,K_4} = 0$
-%                 $\cM^{2}_{T_3,K_5} = 0$ &
-%             $\cM^{2}_{T_4,K_1} = 0$
-%                 $\cM^{2}_{T_4,K_2} = 0$
-%                 $\cM^{2}_{T_4,K_3} = 0$
-%                 $\cM^{2}_{T_4,K_4} = 0$
-%                 $\cM^{2}_{T_4,K_5} = 0$ \\
-%         \hline
-%         $\mathrm{Exp}3^{++}$ &
-%             $\cM^3_{T_1,K_1} = 0$
-%                 $\cM^3_{T_1,K_2} = 0$
-%                 $\cM^3_{T_1,K_3} = 0$
-%                 $\cM^3_{T_1,K_4} = 0$
-%                 $\cM^3_{T_1,K_5} = 0$ &
-%             $\cM^3_{T_2,K_1} = 0$
-%                 $\cM^3_{T_2,K_2} = 0$
-%                 $\cM^3_{T_2,K_3} = 0$
-%                 $\cM^3_{T_2,K_4} = 0$
-%                 $\cM^3_{T_2,K_5} = 0$ &
-%             $\cM^3_{T_3,K_1} = 0$
-%                 $\cM^3_{T_3,K_2} = 0$
-%                 $\cM^3_{T_3,K_3} = 0$
-%                 $\cM^3_{T_3,K_4} = 0$
-%                 $\cM^3_{T_3,K_5} = 0$ &
-%             $\cM^3_{T_4,K_1} = 0$
-%                 $\cM^3_{T_4,K_2} = 0$
-%                 $\cM^3_{T_4,K_3} = 0$
-%                 $\cM^3_{T_4,K_4} = 0$
-%                 $\cM^3_{T_4,K_5} = 0$ \\
-%         \hline
-%         $\UCB_1$ &
-%             $\cM^{4}_{T_1,K_1} = 0$
-%                 $\cM^{4}_{T_1,K_2} = 0$
-%                 $\cM^{4}_{T_1,K_3} = 0$
-%                 $\cM^{4}_{T_1,K_4} = 0$
-%                 $\cM^{4}_{T_1,K_5} = 0$ &
-%             $\cM^{4}_{T_2,K_1} = 0$
-%                 $\cM^{4}_{T_2,K_2} = 0$
-%                 $\cM^{4}_{T_2,K_3} = 0$
-%                 $\cM^{4}_{T_2,K_4} = 0$
-%                 $\cM^{4}_{T_2,K_5} = 0$ &
-%             $\cM^{4}_{T_3,K_1} = 0$
-%                 $\cM^{4}_{T_3,K_2} = 0$
-%                 $\cM^{4}_{T_3,K_3} = 0$
-%                 $\cM^{4}_{T_3,K_4} = 0$
-%                 $\cM^{4}_{T_3,K_5} = 0$ &
-%             $\cM^{4}_{T_4,K_1} = 0$
-%                 $\cM^{4}_{T_4,K_2} = 0$
-%                 $\cM^{4}_{T_4,K_3} = 0$
-%                 $\cM^{4}_{T_4,K_4} = 0$
-%                 $\cM^{4}_{T_4,K_5} = 0$ \\
-%         \hline
-%         \klUCB{} &
-%             $\cM^{5}_{T_1,K_1} = 0$
-%                 $\cM^{5}_{T_1,K_2} = 0$
-%                 $\cM^{5}_{T_1,K_3} = 0$
-%                 $\cM^{5}_{T_1,K_4} = 0$
-%                 $\cM^{5}_{T_1,K_5} = 0$ &
-%             $\cM^{5}_{T_2,K_1} = 0$
-%                 $\cM^{5}_{T_2,K_2} = 0$
-%                 $\cM^{5}_{T_2,K_3} = 0$
-%                 $\cM^{5}_{T_2,K_4} = 0$
-%                 $\cM^{5}_{T_2,K_5} = 0$ &
-%             $\cM^{5}_{T_3,K_1} = 0$
-%                 $\cM^{5}_{T_3,K_2} = 0$
-%                 $\cM^{5}_{T_3,K_3} = 0$
-%                 $\cM^{5}_{T_3,K_4} = 0$
-%                 $\cM^{5}_{T_3,K_5} = 0$ &
-%             $\cM^{5}_{T_4,K_1} = 0$
-%                 $\cM^{5}_{T_4,K_2} = 0$
-%                 $\cM^{5}_{T_4,K_3} = 0$
-%                 $\cM^{5}_{T_4,K_4} = 0$
-%                 $\cM^{5}_{T_4,K_5} = 0$ \\
-%         \hline
-%         Thompson sampling &
-%             $\cM^{6}_{T_1,K_1} = 0$
-%                 $\cM^{6}_{T_1,K_2} = 0$
-%                 $\cM^{6}_{T_1,K_3} = 0$
-%                 $\cM^{6}_{T_1,K_4} = 0$
-%                 $\cM^{6}_{T_1,K_5} = 0$ &
-%             $\cM^{6}_{T_2,K_1} = 0$
-%                 $\cM^{6}_{T_2,K_2} = 0$
-%                 $\cM^{6}_{T_2,K_3} = 0$
-%                 $\cM^{6}_{T_2,K_4} = 0$
-%                 $\cM^{6}_{T_2,K_5} = 0$ &
-%             $\cM^{6}_{T_3,K_1} = 0$
-%                 $\cM^{6}_{T_3,K_2} = 0$
-%                 $\cM^{6}_{T_3,K_3} = 0$
-%                 $\cM^{6}_{T_3,K_4} = 0$
-%                 $\cM^{6}_{T_3,K_5} = 0$ &
-%             $\cM^{6}_{T_4,K_1} = 0$
-%                 $\cM^{6}_{T_4,K_2} = 0$
-%                 $\cM^{6}_{T_4,K_3} = 0$
-%                 $\cM^{6}_{T_4,K_4} = 0$
-%                 $\cM^{6}_{T_4,K_5} = 0$ \\
-%         \hline
-%         Bayes-UCB &
-%             $\cM^{7}_{T_1,K_1} = 0$
-%                 $\cM^{7}_{T_1,K_2} = 0$
-%                 $\cM^{7}_{T_1,K_3} = 0$
-%                 $\cM^{7}_{T_1,K_4} = 0$
-%                 $\cM^{7}_{T_1,K_5} = 0$ &
-%             $\cM^{7}_{T_2,K_1} = 0$
-%                 $\cM^{7}_{T_2,K_2} = 0$
-%                 $\cM^{7}_{T_2,K_3} = 0$
-%                 $\cM^{7}_{T_2,K_4} = 0$
-%                 $\cM^{7}_{T_2,K_5} = 0$ &
-%             $\cM^{7}_{T_3,K_1} = 0$
-%                 $\cM^{7}_{T_3,K_2} = 0$
-%                 $\cM^{7}_{T_3,K_3} = 0$
-%                 $\cM^{7}_{T_3,K_4} = 0$
-%                 $\cM^{7}_{T_3,K_5} = 0$ &
-%             $\cM^{7}_{T_4,K_1} = 0$
-%                 $\cM^{7}_{T_4,K_2} = 0$
-%                 $\cM^{7}_{T_4,K_3} = 0$
-%                 $\cM^{7}_{T_4,K_4} = 0$
-%                 $\cM^{7}_{T_4,K_5} = 0$ \\
-%         \hline
-%         AdBandits &
-%             $\cM^{8}_{T_1,K_1} = 0$
-%                 $\cM^{8}_{T_1,K_2} = 0$
-%                 $\cM^{8}_{T_1,K_3} = 0$
-%                 $\cM^{8}_{T_1,K_4} = 0$
-%                 $\cM^{8}_{T_1,K_5} = 0$ &
-%             $\cM^{8}_{T_2,K_1} = 0$
-%                 $\cM^{8}_{T_2,K_2} = 0$
-%                 $\cM^{8}_{T_2,K_3} = 0$
-%                 $\cM^{8}_{T_2,K_4} = 0$
-%                 $\cM^{8}_{T_2,K_5} = 0$ &
-%             $\cM^{8}_{T_3,K_1} = 0$
-%                 $\cM^{8}_{T_3,K_2} = 0$
-%                 $\cM^{8}_{T_3,K_3} = 0$
-%                 $\cM^{8}_{T_3,K_4} = 0$
-%                 $\cM^{8}_{T_3,K_5} = 0$ &
-%             $\cM^{8}_{T_4,K_1} = 0$
-%                 $\cM^{8}_{T_4,K_2} = 0$
-%                 $\cM^{8}_{T_4,K_3} = 0$
-%                 $\cM^{8}_{T_4,K_4} = 0$
-%                 $\cM^{8}_{T_4,K_5} = 0$ \\
-%         \hline
-%         BESA &
-%             $\cM^{9}_{T_1,K_1} = 0$
-%                 $\cM^{9}_{T_1,K_2} = 0$
-%                 $\cM^{9}_{T_1,K_3} = 0$
-%                 $\cM^{9}_{T_1,K_4} = 0$
-%                 $\cM^{9}_{T_1,K_5} = 0$ &
-%             $\cM^{9}_{T_2,K_1} = 0$
-%                 $\cM^{9}_{T_2,K_2} = 0$
-%                 $\cM^{9}_{T_2,K_3} = 0$
-%                 $\cM^{9}_{T_2,K_4} = 0$
-%                 $\cM^{9}_{T_2,K_5} = 0$ &
-%             $\cM^{9}_{T_3,K_1} = 0$
-%                 $\cM^{9}_{T_3,K_2} = 0$
-%                 $\cM^{9}_{T_3,K_3} = 0$
-%                 $\cM^{9}_{T_3,K_4} = 0$
-%                 $\cM^{9}_{T_3,K_5} = 0$ &
-%             $\cM^{9}_{T_4,K_1} = 0$
-%                 $\cM^{9}_{T_4,K_2} = 0$
-%                 $\cM^{9}_{T_4,K_3} = 0$
-%                 $\cM^{9}_{T_4,K_4} = 0$
-%                 $\cM^{9}_{T_4,K_5} = 0$ \\
-%         \hline
-%     \end{tabular}
-%     \caption{Non-normalized memory cost, rounded and in Bytes, for $9$ different algorithms on problem $1$ with different, values of time $T$ and number of arm $K$.}
-%     \label{table:3:memory_problem1}
-% \end{footnotesize}  % WARNING
-% \end{table}
-
-\subsection{Summary of the experiments}
-
-\TODOL{Au lieu d'avoir des longues captions pour les figures, je vais écrire quelques paragraphes donnant les mêmes conclusion et faisant référence aux figures.
-Je ferai ça très vite si on est contents des figures.}
-
+\paragraph{Summary of the experiments.}
+%
 The main take-away message is the following:
@@
 
-% \newpage  % WARNING
+\newpage  % WARNING
 % ----------------------------------------------------------------------------

### 2019-06-14 2-Chapters/3-Chapter/chapter3.tex
Added figures of Chapter SMPyBandits

@@
 
-\TODOL{Analyser les expériences les plus récentes et décrire mieux en 3.2, et refaire toutes les courbes pour 3.2 et 3.3}
+\TODOL{Si j'ai le temps, je peux améliorer les expériences faites en 3.3 et 3.4. C'est déjà suffisant comme ça, je peux juste faire mieux !}
 
@@
         All algorithms appear to have a regret slowly growing with respect to $K$, as predicted by the regret bounds which are linear with respect to the number of arms, as predicted.
-        \textsc{Tsallis-Inf} is the only state-of-the-art algorithm which appears to be very inefficient for problems with a small number of arms.
-        Both the $\varepsilon$-greedy and Explore-then-Commit algorithms performed poorly, achieving linear regret (not displayed).
+        Bayesian algorithms appear to be the most efficient, and \klUCB{} as well as \UCB{} are also shown to be efficient.
+        % \textsc{Tsallis-Inf} is the only state-of-the-art algorithm which appears to be very inefficient for problems with a small number of arms.
+        Both the $\varepsilon$-greedy and Explore-then-Commit algorithms performed poorly, actually they achieve linear regret.
+        % (not displayed).
 	}
@@
         for $K=32$ and for $16$ algorithms.
-        All efficient algorithms appear to have a logarithmic regret, meaning that their regret bound is indeed logarithmic with respect to the horizon, as predicted.
+        All efficient algorithms appear to have a logarithmic regret with respect to the horizon, as predicted.
         The respective ranking of all the algorithms also appears to remain preserved for different values of $T$, which is also backed-up by theoretical results: if two algorithms $\cA$ and $\cA'$ has a regret close to their regret bounds, of the form $R_T \leq \cO(K \log(T))$, and the bound for $\cA$ use a smaller constant than the bound for $\cA'$, $\cA$ should obtain a smaller regret than $\cA'$ no matter the horizon.
-        Finally, we also observe once more than both the $\varepsilon$-greedy and Explore-then-Commit algorithms performed poorly, achieving linear regret (not displayed after $T=100000$).
+        Bayesian algorithms appear to be the most efficient, and \klUCB{} as well as \UCB{} are also shown to be efficient.
+        Finally, we also observe once more than both the $\varepsilon$-greedy and Explore-then-Commit algorithms performed poorly, achieving linear regret.
 	}
@@
     \item
-    \UCB{} (or Thompson sampling) is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
+    \UCB{} is used in the more applied Chapter~\ref{chapter:4}, when simplicity is favored,
     \item
@@
 
-In Figure~\ref{fig:3:16_different_algorithms__lognormregret_vs_logtime__28pb}, we can see FIXME
+In Figure~\ref{fig:3:16_different_algorithms__lognormregret_vs_logtime__28pb}, we can see that Bayesian algorithms appear to be the most efficient (\ie, in the bottom left corner), and \klUCB{} is very close to their best empirical performances.
+\UCB, algorithms that were proven to perform similarly to \UCB{} (\ie, PHE and RCB), and other index policies inspired by \UCB{} (ApprFHG, $\UCB\dagger$) enjoy very similar performances: a larger regret than Bayesian algorithms and \klUCB, but a shorter running time.
+%
+This highlights a trade-off between optimality in terms of regret and computational efficiency.
 
@@
 
-In Figure~\ref{fig:3:16_different_algorithms__lognormregret_vs_logmemory__28pb}, we can see FIXME
+Like for the computational cost discussed above,
+in Figure~\ref{fig:3:16_different_algorithms__lognormregret_vs_logmemory__28pb}, we can see that Bayesian algorithms like Thompson sampling again appear to be the most efficient (\ie, in the bottom left corner), and \klUCB{} is close to their best empirical performances.
+\UCB{} and other index policies inspired by \UCB{} (ApprFHG, $\UCB\dagger$) enjoy very similar performances: a larger regret than Bayesian algorithms and \klUCB, but a similar memory cost.
+%
+This time, there is no clear trade-off between optimality in terms of regret and memory cost, and based on simply this Figure~\ref{fig:3:16_different_algorithms__lognormregret_vs_logmemory__28pb}, one could advise to use Thompson sampling rather than \klUCB.
 
@@
         $y$-axis is in log-scale.
-        All algorithms (except BESA) has a very slowly growing normalized memory cost, meaning that for $T$ rounds they use a computation time almost independent of $T$, as predicted: $\cM_T = \bigO{K}$.
-        There is a clear separation between three groups.
-        First, BESA is very costly in terms of memory, and $\UCB\dagger$ and \textsc{Tsallis-Inf} are also more costly than others, and both experiences a linear growth of their memory cost.
-        Then, we see PHE, $\mathrm{Exp}3^{++}$, Bayes-UCB and AdBandits having a small but linear growth of memory cost, and it can be explained for each of them (\eg, for Bayes-UCB, computing the $1-1/t$ quantile at time $t$ becomes more and more costly as $t$ grows, and thus as $T$ grows,  ).
-        Finally, the other algorithms (including \UCB, \klUCB{} and Thompson sampling) have a very small footprint and their memory cost do not grow much when $T$ increases.
+        All algorithms have a growing normalized memory cost, meaning that for $T$ rounds they use a computation time almost independent of $T$, as predicted: $\cM_T = \bigO{K}$.
+        % All algorithms (except BESA) have a very slowly growing normalized memory cost, meaning that for $T$ rounds they use a computation time almost independent of $T$, as predicted: $\cM_T = \bigO{K}$.
+        % There is a clear separation between three groups.
+        % First, BESA is very costly in terms of memory, and $\UCB\dagger$ and \textsc{Tsallis-Inf} are also more costly than others, and both experiences a linear growth of their memory cost.
+        % Then, we see PHE, $\mathrm{Exp}3^{++}$, Bayes-UCB and AdBandits having a small but linear growth of memory cost, and it can be explained for each of them (\eg, for Bayes-UCB, computing a quantile at time $t$ becomes more and more costly as $t$ grows, and thus as $T$ grows,  ).
+        % Finally, the other algorithms (including \UCB, \klUCB{} and Thompson sampling) have a very small footprint and their memory cost do not grow much when $T$ increases.
 	}
@@
 
+We also shown a trade-off between optimality in terms of regret, and efficiency in terms of time complexity or memory cost.
+Our position regarding this trade-off is motivated by Occam's razor principle.
+On the first hand, if the algorithm should be implemented on a cognitive radio device that has limited hardware capacity (for instance),
+one can reasonably aim at the simplest yet order-optimal algorithms, and we advise to use \UCB{} and algorithms running on top of this simple index policy, as we do in Chapter~\ref{chapter:4}.
+On the other hand, if one is more interested in the mathematical developments and wants to prove the tightest possible regret upper bounds, aiming at more efficient but more complex algorithms is interesting, and we chose the \klUCB{} algorithm in Chapters~\ref{chapter:5} and \ref{chapter:6}.
+
 % I want to show the real numerical costs in terms of computation time and storage requirement of the various algorithms for single-player bandits.
@@
 
+\newpage  % WARNING
 \newpage  % WARNING
@@
 We give below in Code~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
-It uses \texttt{for} loops and environment variables from Bash, and not Python, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $N$, $T$ and $K$.
+It uses \texttt{for} loops and environment variables from Bash, and not Python, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $T$ and $K$.
 
@@
     \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
-$ for N in 4 100 1000;        do \
-    for T in $(seq 5000 5000 50000); do \
+$ for T in $(seq 5000 5000 50000); do \
       for K in $(seq 2 2 32); do \
       DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
-          N=$N BAYES=True  T=$T K=$K \
+          N=1000  BAYES=True    T=$T         K=$K \
           python3 main.py configuration.py \
       && cp logs/main_py3_log.txt \
-              logs/main_py3_log__N${N}_BAYES_T{T}_K{K}.txt \
-  done; done; done
+              logs/main_py3_log__N1000_BAYES_T{T}_K{K}.txt \
+  done; done
     \end{minted}

### 2019-06-14 2-Chapters/1-Chapter/chapter1.tex
Fixed typos

@@
 \graphicspath{{2-Chapters/1-Chapter/Images/}}
+
+% ----------------------------------------------------------------------------
+\section{Context of this thesis}
+\label{sec:1:problems}
+
+\paragraph{On Cognitive Radio.}
+
+Des thèses précédentes avaient déjà identifié l'apprentissage séquentiel comme étant pertinent pour la radio intelligente.
+On peut citer les travaux de \cite{Jouini09,Jouini10,Jouini12}, qui ont notamment souligné les points suivants :
+
+\begin{itemize}
+    \item implémentation peu couteuse en ressources d'exécution (temps et mémoire), contrairement aux autres approches de décision basées sur l'optimisation, ou d'autres techniques d'IA,
+    \item capacité à apprendre depuis zéro (pas de phase de pré apprentissage durant laquelle le système ne fait rien d'autre qu'apprendre), et forte capacité à s'adapter automatiquement à un nouveau contexte,
+    \item coordination décentralisée possible de plusieurs joueurs, sans qu'il n'y ait de communication autorisée entre eux,
+    \item approche cohérente dans un système décentralisé, si chaque objet sans fil apprend de son côté plutôt que de recevoir des instructions d'une station de base,
+    \item résistent aux erreurs d'apprentissage
+\end{itemize}
+
+Ces techniques ont aussi permis d'identifier les bandits multi-bras comme le bon modèle d'apprentissage séquentiel statistique dans le contexte de la radio intelligente.
+Des algorithmes de MAB ont été utilisés avec succès dans différents contextes :
+
+\begin{itemize}
+    \item le choix de la configuration radio, avec par exemple \cite{KerkoucheAlami18} récemment,
+    \item l'accès dynamique au spectre, et en particulier l'accès opportuniste (OSA) \cite{Jouini09}.
+\end{itemize}
+
+Mais le cas OSA a plusieurs faiblesses :
+
+\begin{itemize}
+    \item il est trop futuriste, car la régulation des fréquences ne permet pas encore (et en est très loin) de l'appliquer en situation réelle,
+    \item en l'absence de régulation coordonnée, chaque cas d'OSA dépendra de la bande considérée (et comment les objets "licencees" s'y comportent), mais les études de l'équipe SCEE ont montrées que les approches MAB sont capables de s'adapter à tous les cas,
+    \item le retour de décision utilisé dans l'OSA repose sur le sensing, qui est le maillon faible de la boucle (pas encore assez fiable pour être utilisé en vrai).
+    \item d'où l'idée d'utiliser les MAB en Cognitive Radio, dans les bandes non licenciées, dès 2011 \cite{Jouini12}. Dans ce cas, le contexte est comme en OSA, mais sans PU, et pas de sensing.
+\end{itemize}
+
+Pour toutes ces raisons, le cadre de l'internet des objets (IoT) apparaît comme idéal car c'est un cadre qui s'affranchit des contraintes de l'OSA.
+
+Par conséquent, cette thèse vise à refermer certaines portes qui étaient encore intéressantes à étudier dans la suite de ce qui a été considéré pour l'OSA
+
+\paragraph{On Multi-Armed Bandit.}
 

### 2019-06-15 2-Chapters/4-Chapter/chapter4.tex
Rien de neuf
Typo

@@
 %
-In this Chapter~\ref{chapter:4}, we propose two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
+In this chapter, we propose two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
 that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
@@
 % \section{Motivations for Selfish MAB Learning for IoT Networks}
-\section{Introduction}
+\section{Introduction and motivations for MAB learning for IoT Networks}
 \label{sec:4:motivations}

### 2019-06-15 2-Chapters/5-Chapter/chapter5.tex
\vfill{} before \minitoc at beginning of all Chapters
Small improvements on layout of abstract/minitoc for chapters
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1
Better links to
- https://SMPyBandits.GitHub.io/MultiPlayers.html
- https://SMPyBandits.GitHub.io/NonStationaryBandits.html
Typo

@@
 
-\paragraph{Abstract.}
+\abstractStartChapter{}%
 %
-In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
+In this chapter, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
 We restrict to the easier case of at most $M \leq K$ devices in a network with $K$ channels, because it was found very hard to analyze the aforementioned IoT network model.
@@
 
-% \TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}
-
-\minitoc
-% ----------------------------------------------------------------------
-
-\newpage
+\minitocStartChapter{}
 
@@
 
-The experiments in this chapter use SMPyBandits,
-and we refer to the following page for details on how to reproduce them,\\
+The experiments in this chapter use our library SMPyBandits, as presented in Chapter~\ref{chapter:3},
+and we refer to the following page for details on how to reproduce them,
 \href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}}

### 2019-06-15 2-Chapters/1-Chapter/chapter1.tex
Tellement difficile de savoir comment avancer

@@
 
-\TODOL{Ecrire les 1.1 et 1.2 en quelques paragraphes}
+\TODOL{Ecrire les 1.1 et 1.2 en quelques pages}
 
@@
 %
-We start by presenting the problems studied in this thesis,
-then we announce the different directions we considered and the solutions we proposed.
+We start by presenting the context then the problems studied in this thesis.
+We further announce the different directions we considered and the solutions we proposed.
 We give an overview of our main contributions, alongside a list of publications written in the last three years.
 %
-The organization of this manuscript is detailed in a flow graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but all use the notations introduced on Chapter~\ref{chapter:2}.
+The organization of this manuscript is detailed in a flow graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but they are all built upon the model and notations introduced on Chapter~\ref{chapter:2}.
 
@@
 \label{sec:1:problems}
+
+At the root of the problems that motivates this thesis is the question of global warming, and the increase of the world population.
+During the last 150 years, mankind has developed a wide range of different communications technology, and since the $1920$s wireless communications between devices have been made possible, and more and more frequent in our lives.
+With the avent of Internet of Things networks, billions of autonomous low-power devices are expected to be deployed world-wide, for a large range of different applications.
+It is now well-known that with the current trend of population and energy increase, any newly deployed technology should be
+cheap and energy efficient,
+as well as adapted to serve a large number of people and devices.
+
+\paragraph{From old TV to the IoT standards.}
+%
+Historically, three families of wireless communication systems have been deployed: first, central broadcasting (\eg, old TV), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things networks (IoT).
+
+% Central broadcasting
+Starting from FM radio music and news, then old black-and-white TV and color TV, the first systems were purely made for centralized broadcasting of information.
+In a large area, one antenna well located emits continuously in a fixed Radio Frequency (RF) band, and many devices can listen to this RF traffic and decode it, for instance to listen to music.
+Different standards have been developed, and some of them are still in use today, and their nature requires a central authority that allocates RF bands to different usages (\eg, TV channels etc).
+For example, in the city of Cesson-Sévigné where the campus of CentraleSupélec is located, a high RF tower emits on many standards, and in particular the \SI{92.3}{\mega\hertz} band is allocated to a French classical music radio (\emph{Radio Classique}).
+Broadcast-based systems are not only used for TV or radio,
+and for instance two other famous and world-wide deployed examples are the Global Positioning System (GPS) and infrared remote controllers.
+
+% Two ways communications
+However, this kind of mono-directional wireless communications were soon found to not be satisfactory for many applications, and thus bi-directional standards were defined from the $1940$s.
+First developed by and for the military, such applications have reached general public usage since the $1970$s.
+The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people lives in the richer countries in the world, with the booming of cellphones since the $2000$s and smartphones since the $2010$s.
+In most systems, there is still an antenna in charge of a large area (or cell), and many devices are able to receive and transmit data to the antenna.
+Nowadays, the most well-known systems include WiFi, and systems deployed for mobile phone communications, with a long list of standards from 1G, 2G, 3G, 4G and now 5G, still in development in $2019$.
+Even if both systems initially differed in usage, as 1G was designed to exchange only voice, and WiFi to link wireless devices at home to an Internet-connected box, they are conceptually very close.
+As the systems are cellular, any device knows the antenna (or base station) it is associated with, and the bi-directional nature of their exchanges makes a centralized control possible.
+In order to optimize the efficiency of the network, the base station is equipped with decision making algorithms that affects the devices in the cell to the available resources (time slots, RF bands, power etc).
+To be able to communicate with the base station without interfering with each other, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA), or more recently to Non-Orthogonal techniques (NOMA).
+Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance of the wireless communications to and from the devices.
+More recently, green radio has emerged as a solution to the tradeoff between maximizing QoS while minimizing the power consumption of base stations and devices.
+
+% The opposite: decentralized broadcasting, and the age of IoT
+Finally, a third kind of systems are of the opposite kind, and can be designed as decentralized broadcasting: an antenna is still in charge of many devices, but such devices can only send up-link packets, and the only down-link data they can receive are short acknowledgements sent from the base-station to indicate success or failure of an up-link packet.
+This family of wireless systems are referred to as Internet of Things,
+and a typical example of application is for sensor networks.
+For the future development of smart grids, smart cities, smart homes, or smart agriculture, sensor networks are promised to be widely deployed.
+Two examples of future applications that are already in deployment, in France or other countries, are connected buildings and connected agriculture.
+For buildings, the main goal is to reduce the cost of heating empty buildings and use temperature sensor networks to get accurate and regular data about the temperature in each room and each floor, and let the centralized heat control optimize its cost and energy consumption.
+For agriculture, one example can be to equipped every cow (in large farms) with a sensors that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and monitors the health of the animals.
+
+\newpage
+
+But de ce chapitre :
+le lecteur va savoir en quelques pages
+
+- les télécoms ? télécoms décentralisées ?
+  => okay
+- l'IoT c'est quoi ?
+- les bandits c'est quoi ?
+- les bandits multi-joueurs, pourquoi ?
+- pourquoi étudier ça ?
+- OK qu'est ce qu'on a proposé et pourquoi ?
+  => plan des contributions, bullet point (1ère version, je pourrai faire mieux après)
+
+- OK comment lire la thèse
+  => le plan Figure~\ref{fig:1:organization}
+
+\TODOL{Je dois faire ça d'ici mercredi matin}
+% Je peux m'inspirer beaucoup de la thèse de Navik...
+% https://tel.archives-ouvertes.fr/tel-01668536/document
+
+commencer par un paragraphe qui explique qu'on étudie les télécom dans le but de les rendre automatiquement plus efficace, plus verte, plus adaptative
+
+commencer par décrire ce sur quoi porte la thèse, en deux phrases qui réduisent le cadre
+
+- technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
+- apprentissage statistique (machine learning) > séquentiel > par renforcement > information partielle (bandits)
+
+Je peux piquer les quelques figures chouettes de Navik (en attendant la version finale ou je referai mes figures)
+
+Il faut expliquer ce que c'est que la radio intelligente, et le cas de l'OSA
+
+je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
+
+Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
+
+expliquer ce que Christophe Moy, Wassim Jouini, Navik Modi etc ont fait depuis 2009
+
+quels nouveaux challenges en 2016 ?
+
+autre direction
+
+- radio intelligente > internet des objets
+
+directions de recherches
+
+- réfléchir aux modèles, quelles hypothèses sont retirées comparé au modèle OSA
+
+- analyser mathématiquement les modèles
+
+- proposer des solutions : algorithmes, dont la convergence est justifiée, ou heuristiques
+
+- des simulations logicielles
+
+- des démonstrations matérielles
+
+- et en bonus on obtient aussi de belles contributions théoriques
+
+\subsection{Reinforcement learning}
+
+\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.40\textwidth}{!}{
+        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=2]
+            %
+            % We start by placing the blocks
+            \node [block] (player) at (0,0) {Player};
+            % We draw an edge between the player and system block to
+            \node [block] (environment) at (2,0) {Environment};
+            % Once the nodes are placed, connecting them is easy.
+            \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Reward} (player);
+            %
+        \end{tikzpicture}
+    }
+\caption{Reinforcement learning cycle.}
+\label{fig:1:ReinforcementLearningCycle}
+\end{figure}
+
+\hline{}
+\newpage
 
@@
 \TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira..}
-In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
+In a second PhD thesis supervised by Christophe Moy \cite{Modi17PhD}, Navikkumar Modi studied the impact
 of using MAB algorithms to optimize channel selection
@@
 Another interesting works is \cite{Modi17QoS}.
-
-\TODOL{FIXME Finir ce blabla rapidement ? Ou alors ce sera déjà fait en intro/chapitre1 ??}
 
@@
 
-% \TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}
-
 The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
@@
 
-We already explained in Chapter~\ref{chapter:1} that
-unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
-This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.
-
-Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
-As end-devices battery life is a key constraint of IoT networks,
-and as IoT networks are decentralized (\ie, the devices initiate transmissions),
-this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
-%
-In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
-Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
-We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
-Kaufmann12Thompson}, for the first model.
-For the demonstration as well as for the second model, without loss of generality, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.
-
-As detailed in Chapter~\ref{chapter:1},
-MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
-For example,
-proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
-and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
-Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
-% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
-
 We evoke here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
@@
 
-\subsection{Reinforcement learning}
+\TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
 
-\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
-\begin{figure}[h!]
-    \centering
-    \resizebox{0.40\textwidth}{!}{
-        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=2]
-            %
-            % We start by placing the blocks
-            \node [block] (player) at (0,0) {Player};
-            % We draw an edge between the player and system block to
-            \node [block] (environment) at (2,0) {Environment};
-            % Once the nodes are placed, connecting them is easy.
-            \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
-            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Reward} (player);
-            %
-        \end{tikzpicture}
-    }
-\caption{Reinforcement learning cycle.}
-\label{fig:1:ReinforcementLearningCycle}
-\end{figure}
+\newpage % WARNING
+% FIXME c'est bon tout ce qu'on a avant ?
 
@@
     \item
-    % We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
-    We present an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
-    \cite{Besson2018WCNC}.
-    % with \textbf{Aggregator} (WCNC 2018)
-
-    \item
     We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
@@
     A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.
+
+    \item
+    % We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+    We present in Chapter~\ref{chapter:25} an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
+    \cite{Besson2018WCNC}.
+    % with \textbf{Aggregator} (WCNC 2018)
 
@@
 
+The reading order of the manuscript can be any top-down path between the Introduction in the current Chapter~\ref{chapter:1}, and the Conclusion in the last Chapter~\ref{chapter:conclusion}.
+As the graph in Figure~\ref{fig:1:organization} shows,
+the thesis is organized in two parts.
+
 \begin{figure}[h!]
     \centering
-    % \resizebox{1.00\textwidth}{!}{
+    \resizebox{0.95\textwidth}{!}{
     \begin{tikzpicture}[>=latex',line join=bevel,scale=2.25]
@@
         \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
+        \node[align=center] (chapter25) at (-2.5,2.25) [rectangle,draw] {Chapter~\ref{chapter:25}\\Online selection\\of the best algorithm};
         \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapter~\ref{chapter:4}\\Two MAB models\\for IoT networks};
-        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapter~\ref{chapter:5}\\Multi-players\\MAB models};
-        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary\\MAB models};
-        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapter~\ref{chapter:conclusion}\\General conclusion};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapter~\ref{chapter:5}\\Multi-players\\Multi-Armed Bandits};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary\\Multi-Armed Bandits};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapter~\ref{chapter:conclusion}\\General Conclusion};
+        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw] {Appendix};
         %
@@
         \draw [color=black,thick,<->] (chapter2) to (chapter3);
+        \draw [color=black,thick,<->] (chapter2) to (chapter25);
         \draw [color=black,thick,->] (chapter2) to (chapter4);
@@
         \draw [color=black,densely dotted,<->]   (chapter4) to (chapter5);
-        \draw [color=black,densely dotted,->] -| (chapter3) to (chapter6);
+        % \draw [color=black,densely dotted,->] -| (chapter3) to (chapter25);
+        % \draw [color=black,densely dotted,->] -| (chapter25) to (chapter6);
         \draw [color=black,densely dotted,<->]   (chapter5) to (chapter6);
@@
         \draw [color=black,thick,->] (chapter6) to (conclusion);
+        \draw [color=black,thick,->] (conclusion) to (appendix);
         %
     \end{tikzpicture}
-    % }
-    \caption[Organization of the thesis: a reading map]{Organization of the thesis: a reading map.
-Any path containing Chapter~\ref{chapter:1}, Chapter~\ref{chapter:2}, at least one of the three Chapters~\ref{chapter:4}, \ref{chapter:5} and \ref{chapter:6}, and the Conclusion is a self contained way to read this thesis.}
+    }
+    \caption[Organization of the thesis: a reading map]{A reading map of the thesis.
+Any top-down path containing Chapter~\ref{chapter:1}, Chapter~\ref{chapter:2}, at least one of the three Chapters~\ref{chapter:4}, \ref{chapter:5} and \ref{chapter:6}, and the Conclusion is a self contained way to read this thesis.}
     \label{fig:1:organization}
@@
 
-The reading order of the manuscript can be any top-down path between the Introduction in the current Chapter~\ref{chapter:1}, and the Conclusion in the last Chapter~\ref{chapter:conclusion}.
-As the graph in Figure~\ref{fig:1:organization} shows,
-the thesis is organized in two parts:
+% \begin{itemize}
+    % \item
+In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2}, required for the rest of the document as we introduce the MAB models, the concepts and the notations used in this thesis.
+Conversely, even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use numerical simulations based on our simulation library SMPyBandits, the Chapter~\ref{chapter:3} where we present it is not required to understand them.
+We conclude this part with Chapter~\ref{chapter:25} which details the first contribution of this thesis, on online algorithms selection.
 
-\begin{itemize}
-    \item
-In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2} where we introduce the MAB models, the concepts and the notations used in all this document, and it is mandatory to read it.
-Conversely, Chapter~\ref{chapter:3} presents our simulation library SMPyBandits, and even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use the library for their numerical simulation, reading this chapter is not required to understand the rest of the document.
-
-    \item
+    % \item
 Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
@@
 and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.
-\end{itemize}
+% \end{itemize}
 
@@
 We conclude this chapter with a list of publications published during this PhD.
-All the publications are published on HAL (see the \href{https://hal.archives-ouvertes.fr/}{\texttt{HAL.Archives-Ouvertes.fr}} website),
-and the complete list can be found on
+All the following works are published entirely and freely, on the HAL platform (see the \href{https://hal.archives-ouvertes.fr/}{\texttt{HAL.Archives-Ouvertes.fr}} website).
+The complete list can be found on
 \href{https://cv.archives-ouvertes.fr/lilian-besson/}{\texttt{CV.Archives-Ouvertes.fr/lilian-besson}}.
@@
 \begin{itemize}
-
 \item
@@
     % \href{https://HAL.Inria.fr/hal-02144465}{\texttt{HAL.Inria.fr/hal-02144465}}.
+    See Section~\ref{sec:4:gnuradio}.
     \cite{MoyBesson2019}
@@
     % \href{https://HAL.Inria.fr/hal-02049824}{\texttt{HAL.Inria.fr/hal-02049824}}.
+    See Section~\ref{sec:4:retransmissions}.
     \cite{Bonnefoi2019WCNC}
@@
     % \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
+    See Section~\ref{sec:4:gnuradio}.
     \cite{Besson2019WCNC}
@@
     % \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
+    See Chapter~\ref{chapter:5}.
     \cite{Besson2018ALT}
@@
     % \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
+    See Chapter~\ref{chapter:25}.
     \cite{Besson2018WCNC}
@@
     \textbf{Best Paper Award}.
+    See Section~\ref{sec:4:firstModel}.
     \cite{Bonnefoi17}
@@
     See \href{https://YouTu.be/HospLNQhcMk}{\texttt{YouTu.be/HospLNQhcMk}} for a $6$-minutes presentation video.
+    See Section~\ref{sec:4:gnuradio}.
     \cite{Besson2018ICT}
@@
     % \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
+    See Chapter~\ref{chapter:6}.
     \cite{Besson2019Gretsi}
@@
     Preprint at \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
+    See Chapter~\ref{chapter:6}.
     \cite{Besson2019GLRT}
@@
     \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
-    by \textbf{Lilian Besson}, active development since October $2016$,
+    by \textbf{Lilian Besson}, active development between October $2016$ and March $2019$,
     \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
@@
     and a complete documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} and \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.
+    See Chapter~\ref{chapter:3}.
     \cite{SMPyBandits,SMPyBanditsJMLR}
@@
     Preprint at \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.
+    See Appendix~\ref{app:2:DoublingTricks}.
     \cite{Besson2018DoublingTricks}

### 2019-06-15 2-Chapters/1-Chapter/chapter1.tex
SO HARD

@@
 and a typical example of application is for sensor networks.
+
 For the future development of smart grids, smart cities, smart homes, or smart agriculture, sensor networks are promised to be widely deployed.
@@
 For agriculture, one example can be to equipped every cow (in large farms) with a sensors that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and monitors the health of the animals.
+
+This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices, it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
+In the present and future IoT networks, many devices of heterogenous nature are using the same antenna for many different applications.
+A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and will use a battery, which duration should be maximized (typically more than 10 years).
+Another common constraint for IoT devices is their low duty cycle, as most applications target one to a few messages sent by day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
+%
+In this thesis, we study Internet of Things networks, and these two constraints of low power consumption (or long battery life) and low duty cycle are essential.
+
+\paragraph{Cognitive Radio.}
+%
 
@@
 
-- les télécoms ? télécoms décentralisées ?
-  => okay
-- l'IoT c'est quoi ?
+- contexte historique technologique => okay
+
+- les télécoms ? télécoms décentralisées ? => okay
+
+- l'IoT c'est quoi ? exemple d'applications ? => okay ?
+
 - les bandits c'est quoi ?
-- les bandits multi-joueurs, pourquoi ?
+
+- les bandits multi-joueurs décentralisés, pourquoi ?
+
 - pourquoi étudier ça ?
-- OK qu'est ce qu'on a proposé et pourquoi ?
-  => plan des contributions, bullet point (1ère version, je pourrai faire mieux après)
 
-- OK comment lire la thèse
-  => le plan Figure~\ref{fig:1:organization}
+- OK qu'est ce qu'on a proposé et pourquoi ? => plan des contributions, bullet point (1ère version, je pourrai faire mieux après)
+
+- OK comment lire la thèse => le plan Figure~\ref{fig:1:organization}
 
@@
 - et en bonus on obtient aussi de belles contributions théoriques
+
+\newpage
 
@@
 % Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
+
+\TODOL{Tous les paragraphes suivants sont copiés collés de plus tard dans la thèse, je peux les réadapter ici, et les supprimer/raccourcir plus loin}
 

### 2019-06-15 2-Chapters/2-Chapter/chapter2.tex
Fixed typos
done

@@
 % \begin{equation}\label{eq:2:exponentialFamily}
-$ \Pr(x | \lambda) = h(x) \exp \left( \eta x - A(\eta) \right)$,
+$ \Pr_{\eta}(x | \lambda) = h(x) \exp \left( \eta x - A(\eta) \right)$,
 % \end{equation}
-for a parameter vector $\eta$ (the canonical parameter), and a sufficient statistic that we restrict to be $T(x)=x$, and a function $h$.
+for a parameter vector $\eta$ (the canonical parameter),
+% and a sufficient statistic that we restrict to be $T(x)=x$,
+and a function $h$.
 The cumulant function $A(\eta)$ is entirely determined by the $\eta$ and $h$,
@@
 
-We have to quote two important results on exponential families:
+We mention two important results on exponential families:
 
@@
     \emph{A distribution in such family is entirely characterized by its parameter $\eta$}.
-    That is why one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[X]$.
+    And the mapping $\eta \mapsto \E_{\eta}[X]$ is one-to-one,
+    if $\E_{\eta}$ is the expectation under the probabilistic model $\bP_{\eta}$.
+    That is why one-dimensional distributions are entirely characterized by their mean, $\mu = \E_{\eta}[X]$.
 
@@
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-    The \emph{Kullback-Leibler divergence} between distributions $d_1$ and $d_2$ is defined as
+    The \emph{Kullback-Leibler divergence} between distributions
+    with densities $d_1$ and $d_2$, wrt to $\lambda$, is defined as
     \begin{align}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
@@
     % $\bullet$
-    \textbf{Gaussian distributions} of a fixed variance are a one-dimensional exponential family,
+    \textbf{Gaussian distributions of a known variance} are a one-dimensional exponential family,
     and in this thesis we do not consider Gaussian with an unknown variance.
@@
 % https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
--- The \textcolor{deeppurple}{$\varepsilon$-greedy strategy} consists in alternating exploration and exploitation at a certain ratio \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
+-- The \textcolor{deeppurple}{$\varepsilon$-greedy strategy} consists in alternating exploration and exploitation at a certain ratio \cite{Auer02}.
+% \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
 Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
@@
 Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
+But this is not satisfactory, as $\min_k \Delta_k$ needs to be known in advance, which is usually not the case for real applications.
 Here again, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
@@
 % https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
--- The ``\textcolor{deepgold}{Explore-then-Exploit}'' strategy, as studied for instance in \cite{GarivierETC2016}, first explores uniformly the $K$ arms for $T_0$ time steps, then only exploits the arm identified as the best arm after these first $T_0$ steps
+-- The ``\textcolor{deepgold}{Explore-then-Exploit}'' strategy, as presented for instance in \cite{Bubeck12}, first explores uniformly the $K$ arms for $T_0$ time steps, then only exploits the arm identified as the best arm after these first $T_0$ steps
 (\ie, one of the arms with highest empirical means, after $T_0/K$ samples of each arm).
 Usually we restrict to $T_0$ being a multiple of $K$.
-On the one hand, we can lower bound its regret by
-$R_T \geq K (\min_{k: \Delta_k > 0} \Delta_k) T_0/K + p (T - T_0)$,
-if $p$ denotes the probability that the chosen arm is not optimal.
-And we can show that $p$ only depends on the number of collected samples $T_0$ and the gaps $\Delta_k$,
-so if $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, \ie, $R_T = \Omega(T)$.
-
-On the other hand, if a lower-bound $d>0$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret.
-%
-To prove a bound on $p$, we use Hoeffding's inequality from \cite{hoeffding1963probability}, reminded below in Lemma~\ref{lem:2:HoeffdingInequality}.
-%
-We have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$, thus Hoeffding's inequality gives $R_T \leq K T_0 + K \exp(-T_0 d^2/4) T$.
-Optimizing on $T_0$ gives $T_0 = K (4/d^2) \log(d^2 T / 4)$
-and proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K (4/d^2) (1 + \log(d^2 T / 4)) = \bigO{K \log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.
-%
-We note that this strategy can also achieve $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$ without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
+% On the one hand, we can lower bound its regret by
+% $R_T \geq K (\min_{k: \Delta_k > 0} \Delta_k) T_0/K + p (T - T_0)$,
+% if $p$ denotes the probability that the chosen arm is not optimal.
+% And we can show that $p$ only depends on the number of collected samples $T_0$ and the gaps $\Delta_k$,
+% so if $T_0$ is fixed independently of $T$ and the problem difficulty, the regret is again growing linearly, \ie, $R_T = \Omega(T)$.
+
+Here again, if a lower-bound $d>0$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret, as stated in Lemma~\ref{lem:2:ExploreThenExploit} below.
+It proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K (4/d^2) (1 + \log(d^2 T / 4)) = \bigO{K \log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.
+%
+%
+We note that this strategy can also achieve a much larger regret bound, $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$, without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
+
+\begin{theorem}\label{thm:2:ExploreThenExploit}
+\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
+    For any instance $I$ with $K$ arms with Bernoulli distributions in $[0,1]$,
+    of means $\mu_1,\dots,\mu_k$, and any horizon $T>K$,
+    and if $d \leq \Delta = \min_k \Delta_k$ is known,
+    let $T_0 = ((2K)/d^2) \log(d^2 T / (2K))$.
+    Then the Explore-then-Exploit algorithm with parameter $T_0$ verifies the following finite-time regret bound
+    \begin{equation}\label{eq:2:ExploreThenExploit}
+        R_T^{\text{Explore-then-Exploit}} \leq \frac{2K}{d^2} \log\left( \frac{d^2 T}{2K} \right) = \bigO{\frac{K \log(T)}{\Delta^2}}.
+    \end{equation}
+\end{leftbar}  % XXX leftbar theorembar, comment if needed
+\end{theorem}
 %
 \begin{smallproof}\label{proof:2:tuningExploreThenCommit}
-    Let us prove the inequality $p \leq 2 K \exp(-T_0 d^2 / 4)$.\\
-    \indent
+    %
+    If $p$ denotes the probability that the chosen arm is not optimal,
+    the algorithm suffers from a linear regret for the first $T_0$ rounds, then with probability $p$ it suffers from a linear regret for the remaining $T-T_0$ rounds, and with probability $1-p$ it suffers no regret.
+    Thus we have $R_T \leq (\max_{k: \Delta_k > 0} \Delta_k) T_0 + p (T - T_0)$.
+    We then prove that $p \leq 2 K \exp(-T_0 d^2 / 4)$.
+
+    We use Hoeffding's inequality from \cite{hoeffding1963probability}, reminded below in Lemma~\ref{lem:2:HoeffdingInequality}.
     %
     First for the case of two arms:
-    if the distributions are Bernoulli (or sub-Gaussian), we can use Hoeffding's inequality to control this probability $p$.
-    %
-    If $\mu_1 = \mu_2 + \Delta$, then
-    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \mu_1 - \Delta/2) + \bP(\widehat{\mu_2} > \mu_2 + \Delta/2)$, and both terms can be bounded by using Hoeffding's inequality with a (non-random) number of samples $n=T_0/2$,
-    to obtain $p \leq 2 \exp(-T_0 \Delta^2 / 4)$.\\
-    \indent
-    %
-    Then for $K\geq2$ arms, a simple union bound gives
-    $p \leq 2 K \exp(-T_0 d^2 / 4)$, if $d$ is a lower-bound on the positive gaps $\Delta_k$.
-    %
-    In all cases, we showed that $p$ is controlled by $T_0$.
+    if $\mu_1 = \mu_2 + \Delta$, then
+    $p = \bP(\widehat{\mu_1} < \widehat{\mu_2}) \leq \bP(\widehat{\mu_1} < \mu_1 - \Delta/2) + \bP(\widehat{\mu_2} > \mu_2 + \Delta/2)$, and both terms can be bounded by using Hoeffding's inequality with a (non-random) number of samples $n \leq T_0/K$,
+    to obtain $p \leq 2 \exp(-T_0 \Delta^2 / (2 K))$.
+    %
+    Then for $K\geq2$ arms, a simple union bound on the sub-optimal arms gives
+    $p \leq 2 (K-1) \exp(-T_0 d^2 / (2 K))$, if $d$ is a lower-bound on the positive gaps $\Delta_k$.
+    %
+    % In all cases, we showed that $p$ is controlled by $T_0$.
+
+    So this bound on $p$ gives $R_T \leq K T_0 + 2 K \exp(-T_0 d^2 / (2 K)) T$ (because $\max_{k: \Delta_k > 0} \Delta_k \leq 1$ and $T - T_0 \leq T$).
+    Optimizing the right hand-side on $T_0$ gives $T_0 = ((2K)/d^2) \log(d^2 T / (2K))$.
+    This gives the announced bound.
 \end{smallproof}
@@
     of empirical mean $\widehat{Z_n} \eqdef \frac{1}{n} \sum\limits_{i=1}^n Z_i$, \emph{Hoeffding's inequality} gives both
+    \vspace*{-10pt}  % FIXME vspace negative to compensate the \end{leftbar}
     \begin{align}\label{eq:2:HoeffdingInequality}
@@
 then \klUCB-switch use the index defined as follows:
-\[
+\begin{align*}
     U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
@@
     \end{cases}.
-\]
+\end{align*}
 

### 2019-06-15 2-Chapters/5-Chapter/chapter5.tex
done

@@
 The first work studying the relaxation of this hypothesis is \cite{Rosenski16}, where this case is called the ``dynamic setting''.
-If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectancy $\E[\bullet]$ is capturing the randomness in the sensing information, as well as in the players' decisions (\ie, collisions):
+If the arrival or departures of players is not random, but determined in advance while still being unknown to any player, the natural notion of regret is the following, where the expectation $\E[\bullet]$ is capturing the randomness in the sensing information, as well as in the players' decisions (\ie, collisions):
 \begin{equation}\label{eq:5:defRegretDynamicSetting}

### 2019-06-16 1-Preamble/4-Acknowledgement/acknowledgement.tex
C'est horrible je sais pas quoi dire ni comment

@@
 I loved traveling with you and hope to continue!
-\emph{J'ai une pensée particulière pour Hélène, dont l'énergie et la détermination depuis 2018 m'ont poussé de l'avant.}
+\emph{J'ai une pensée particulière pour Hélène, merci pour ton énergie et ta détermination depuis 2018.}
 
@@
 \href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{~Que je remercie deux fois plus que ce qu'il me remerciant dans sa propre thèse !
-Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, Rami Othman.
+Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, and Rami Othman.
 

### 2019-06-16 1-Preamble/5-Abstract/resume_fr.tex
C'est horrible je sais pas quoi dire ni comment

@@
 \begin{itemize}
-    \item
-    Nous présentons les concepts et les notations du modèle des bandits multi-bras, dans le chapitre~\ref{chapter:2} selon un point de vue mathématique formel.
-    Nous suivons aussi une approche didactique, puisque nous utilisons en Section~\ref{par:2:interactiveDemoDiscoverMAB} une démonstration interactive en ligne conçue pour permettre à n'importe qui de jouer contre un petit problème de bandit depuis son navigateur.
+    % \item
+    % Nous présentons les concepts et les notations du modèle des bandits multi-bras, dans le chapitre~\ref{chapter:2} selon un point de vue mathématique formel.
+    % Nous suivons aussi une approche didactique, puisque nous utilisons en Section~\ref{par:2:interactiveDemoDiscoverMAB} une démonstration interactive en ligne conçue pour permettre à n'importe qui de jouer contre un petit problème de bandit depuis son navigateur.
 
-    \item
-    Nous donnons une brève revue de la littérature de recherche concernant les algorithmes stochastiques de bandits en Section~\ref{sec:2:famousMABalgorithms}.
-
-    \item
-    Nous présentons le problème du choix de l'algorithme qu'un praticien devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithms MAB disponibles, dans la Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
-    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par quelques simulations numériques qu'il atteint de bonnes performances empiriques
-    (WCNC 2018) \cite{Besson2018WCNC}.
+    % \item
+    % Nous donnons une brève revue de la littérature de recherche concernant les algorithmes stochastiques de bandits en Section~\ref{sec:2:famousMABalgorithms}.
 
@@
     Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences utilisées dans la suite de cette thèse.
+
+    \item
+    Nous présentons le problème du choix de l'algorithme qu'un praticien devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithms MAB disponibles, dans le Chapitre~\ref{chapter:25}.
+    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par quelques simulations numériques qu'il atteint de bonnes performances empiriques
+    (WCNC 2018) \cite{Besson2018WCNC}.
 
@@
         \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapitre~\ref{chapter:3}\\SMPyBandits : bibliothèque\\de simulations pour MAB};
+        \node[align=center] (chapter25) at (-2.5,2.25) [rectangle,draw] {Chapter~\ref{chapter:25}\\Sélection séquentielle\\d'algorithme MAB};
         \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:4}\\Deux modèles MAB\\pour les réseaux IoT};
@@
         \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapitre~\ref{chapter:conclusion}\\Conclusion générale};
+        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw] {Appendices};
         %
@@
         \draw [color=black,densely dotted,<->]   (chapter4) to (chapter5);
-        \draw [color=black,densely dotted,->] -| (chapter3) to (chapter6);
+        % \draw [color=black,densely dotted,->] -| (chapter3) to (chapter6);
+        % \draw [color=black,densely dotted,->] -| (chapter25) to (chapter5);
         \draw [color=black,densely dotted,<->]   (chapter5) to (chapter6);
@@
         \draw [color=black,thick,->] (chapter6) to (conclusion);
+        \draw [color=black,thick,->] (conclusion) to (appendix);
         %
@@
 Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
-Par contre le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
+Par contre le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
+Nous terminons cette première partie par le Chapitre~\ref{chapter:25}, qui n'est pas non plus nécessaire à la compréhension de la suite du manuscrit, et qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
 

### 2019-06-16 2-Chapters/1-Chapter/chapter1.tex
C'est horrible je sais pas quoi dire ni comment

@@
 
+% Where
+This doctoral thesis took place at the IETR laboratory in Rennes (France), in the SCEE team hosted on the Rennes campus of the engineering college CentraleSupélec.
+In Rennes I was supervised by Professor Christophe Moy,
+and I was also co-supervised by Doctor Emilie Kaufmann, whom I visited many times at Inria Lille Nord Europe in Lille (France).
+
+% Main problem
 At the root of the problems that motivates this thesis is the question of global warming, and the increase of the world population.
@@
 Conversely, even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use numerical simulations based on our simulation library SMPyBandits, the Chapter~\ref{chapter:3} where we present it is not required to understand them.
-We conclude this part with Chapter~\ref{chapter:25} which details the first contribution of this thesis, on online algorithms selection.
+We conclude this part with Chapter~\ref{chapter:25}, which is also not mandatory for the rest of this thesis, and which details one of the first contributions of this thesis, a new algorithm for online MAB algorithms selection.
 

### 2019-06-16 2-Chapters/2-Chapter/chapter2.tex
C'est horrible je sais pas quoi dire ni comment

@@
 \TODOL{Intégrer commentaires d'Emilie !}
-\TODOL{Découper en deux chapitres, un pour le modèle et un pour Aggregator ? Si je découpe, il faut changer le dessin tikz et d'autres trucs}
 
@@
 In this chapter, we present the common base of all the more sophisticated models studied in this thesis: the stochastic multi-armed bandit (MAB) model, restricting to the single-player and stationary stochastic case.
-
+%
 We motivate this model and discuss its application to and beyond cognitive radio, and we review existing MAB algorithms.
 Some of which are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling.
-
+%
 We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a one-dimensional distribution.
@@
 % - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
-The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one time step to the next one, \ie, $\mathrm{SC}^{\cA}(T) \eqdef \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
+The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one round to the next one, \ie, $\mathrm{SC}^{\cA}(T) \eqdef \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
 It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.

### 2019-06-16 2-Chapters/Conclusion/conclusion.tex
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1
Last proof reading of Ch6 and conclusion
C'est horrible je sais pas quoi dire ni comment

@@
 % \newpage
-\abstractStartChapter{}
+\abstractStartChapter{}%
 %
-This last chapter concludes the thesis, first by summarizing the previous chapters, and by giving a general conclusion of the thesis.
+This last chapter concludes the document, first by summarizing the previous chapters, and by giving a general conclusion of the thesis.
 %
@@
 This manuscript was organized in two parts, Part~\ref{part:Introduction} and Part~\ref{part:MABIOT}.
-It presented the problems we considered for my PhD research, and we discussed about the main contributions produced.
+It presented the problems we considered for my PhD research, and we discussed about the main contributions produced during the last three years.
 
@@
     \item
-The first half, Part~\ref{part:Introduction}, starts by motivating our work and situating the context of this research.
+The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
 % Chapter 2
-In Chapter~\ref{chapter:2}, we present the multi-armed bandit models, the hypotheses we usually make, and the most important algorithms used to solve MAB problems.
+In Chapter~\ref{chapter:2}, we presented the multi-armed bandit models, the hypotheses we usually make, and the most important algorithms that solve different kinds of MAB problems.
 % Chapter 3
-Then in Chapter~\ref{chapter:3} we discuss in details about our open-source Python library developed to run numerical simulations of MAB problems, SMPyBandits.
+Then in Chapter~\ref{chapter:3} we discussed in details about our open-source Python library, developed to run numerical simulations of MAB problems, SMPyBandits.
+%
+Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an efficient answer to the online algorithm selection problem.
 
     \item
-The second half of this thesis, Part~\ref{part:MABIOT}, presents our main contributions.
+The second half of this thesis, Part~\ref{part:MABIOT}, presented our main contributions.
 % Chapter 4
-In Chapter~\ref{chapter:4}, we first propose some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
-% In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithm.
-We also present a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.
+In Chapter~\ref{chapter:4}, we first proposed some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
+In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithm.
+We also presented a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.
 % Chapter 5
 Because the models studied for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in a noisy environment,
-in Chapter~\ref{chapter:5}, we study a weaker but still interesting model.
-We present different variants of the multi-players MAB model, and we propose the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
+in Chapter~\ref{chapter:5}, we studied a weaker but still interesting model.
+We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
 % Chapter 6
-Finally, in Chapter~\ref{chapter:6}, we remove the stationary hypothesis on the single-player MAB problem, and we analyze a new actively adaptive algorithm (\GLRklUCB).
+Finally, in Chapter~\ref{chapter:6}, we removed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
 %
-For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
+For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations, and on the regret bounds obtained in our theoretical analyses.
 \end{itemize}
@@
 The notations used in the manuscript were introduced by showcasing an interactive demonstration\footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
-
-The first contribution of this manuscript \cite{Besson2018WCNC} concluded this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
-We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
-Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for the practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
 
@@
 Our library is used to run numerical simulations on Chapters~\ref{chapter:2}~ \ref{chapter:5} and~\ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
+
+% ----------------------------------------------------------------------------
+\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:25}}.}
+
+The first contribution of this manuscript \cite{Besson2018WCNC} concluded this part.
+We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
+Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for the practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
 
@@
 % ----------------------------------------------------------------------------
-\newpage
-% FIXME
+% \newpage
 
@@
 One approach to obtain a unified ``master'' algorithm could be to use expert aggregation on the different state-of-the-art algorithms presented in this thesis, and as each one is efficient in a different setting, the resulting aggregated algorithm could also be efficient in each of the different settings.
-But as we illustrated in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, this approach does not seem theoretically promising.
+But as we illustrated in Section~\ref{sub:25:theory}, this approach does not seem theoretically promising.
 

### 2019-06-16 2-Chapters/1-Chapter/chapter1.tex
C'est la merde

@@
 % Where
-This doctoral thesis took place at the IETR laboratory in Rennes (France), in the SCEE team hosted on the Rennes campus of the engineering college CentraleSupélec.
+This doctoral thesis started in October $2016$, and finished in October $2019$.
+My research took place at the IETR laboratory in Rennes (France), in the SCEE team hosted on the Rennes campus of the engineering college CentraleSupélec.
 In Rennes I was supervised by Professor Christophe Moy,
@@
 % Main problem
-At the root of the problems that motivates this thesis is the question of global warming, and the increase of the world population.
+At the root of the problems that motivates this thesis are the questions of global warming, and the increase of the world population.
 During the last 150 years, mankind has developed a wide range of different communications technology, and since the $1920$s wireless communications between devices have been made possible, and more and more frequent in our lives.
-With the avent of Internet of Things networks, billions of autonomous low-power devices are expected to be deployed world-wide, for a large range of different applications.
-It is now well-known that with the current trend of population and energy increase, any newly deployed technology should be
-cheap and energy efficient,
+With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed world-wide, for a large range of different applications.
+It is now well-known that with the current trend of population and with the energy crisis, any newly deployed technology should be cheap and energy efficient,
 as well as adapted to serve a large number of people and devices.
+%
+That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of the deployment of IoT networks, by optimizing the wireless communication thanks to embedded decentralized decision making.
 
@@
 In order to optimize the efficiency of the network, the base station is equipped with decision making algorithms that affects the devices in the cell to the available resources (time slots, RF bands, power etc).
-To be able to communicate with the base station without interfering with each other, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA), or more recently to Non-Orthogonal techniques (NOMA).
+To be able to communicate with the base station without interfering with each others, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA), or more recently to Non-Orthogonal techniques (NOMA).
 Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance of the wireless communications to and from the devices.
@@
 %
+
+\begin{itemize}
+    \item
+    Joseph Mittola \cite{Mitola99}:
+    \emph{``A really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment.}
+
+    \item
+    Simon Haykin \cite{Haykin05}:
+    \emph{``A CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication.}
+\end{itemize}
 
@@
 
-commencer par un paragraphe qui explique qu'on étudie les télécom dans le but de les rendre automatiquement plus efficace, plus verte, plus adaptative
+OK commencer par un paragraphe qui explique qu'on étudie les télécom dans le but de les rendre automatiquement plus efficace, plus verte, plus adaptative
 
-commencer par décrire ce sur quoi porte la thèse, en deux phrases qui réduisent le cadre
+OK commencer par décrire ce sur quoi porte la thèse, en deux phrases qui réduisent le cadre
 
@@
     \item
-    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi} (GRETSI 2019).
+    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
     Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
@@
 
-We conclude this chapter with a list of publications published during this PhD.
+We conclude this chapter with a list of works published during this PhD.
 All the following works are published entirely and freely, on the HAL platform (see the \href{https://hal.archives-ouvertes.fr/}{\texttt{HAL.Archives-Ouvertes.fr}} website).

### 2019-06-16 2-Chapters/2-Chapter/chapter2.tex
C'est la merde

@@
     A second important result is a simplified form for the \emph{Kullback-Leibler divergence} \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
-    The KL divergence is also called the relative entropy is a measure of how one probability distribution is different from a second, reference probability distribution.
+    The KL divergence is also called the \emph{relative entropy},
+    and it is a measure of how one probability distribution is different from a second, reference probability distribution.
 \end{enumerate}
@@
 
-\begin{lemma}[Regret decomposition]\label{lem:2:RegretDecomposition}
-\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
+\begin{proposition}[Regret decomposition]\label{prop:2:RegretDecomposition}
+\begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
     The (mean) regret $R_T^{\cA}$ can be decomposed as a sum of the number of selections of sub-optimal arms $k$, weighted by their gaps:
@@
     % \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
-\end{leftbar}  % XXX leftbar lemmabar, comment if needed
-\end{lemma}
+\end{leftbar}  % XXX leftbar propositionbar, comment if needed
+\end{proposition}
 %
@@
 Both quantities are indeed equal \emph{in expectation}, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition} (thanks to the conditioning on $O_t$ done in the proof).
-In our library SMPyBandits, we implement both estimators, and all values of regret used in this thesis are based on the one using the decomposition of Lemma~\ref{lem:2:RegretDecomposition}, because it gives faster convergence, and also ``smoother'' plots.
+In our library SMPyBandits, we implement both estimators, and all values of regret used in this thesis are based on the one using the decomposition of Proposition~\ref{prop:2:RegretDecomposition}, because it gives faster convergence, and also ``smoother'' plots.
 \label{remark:2:moreAccurateCountofRegretForSimulations}
@@
 
-Moreover, it is interesting to note that the second lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Lemma~\ref{lem:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
+Moreover, it is interesting to note that the second lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Proposition~\ref{prop:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
 should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times in the total of $T$ time steps.
@@
     As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives the inequality \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
-    Thanks to the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
+    Thanks to the regret decomposition from Proposition~\ref{prop:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
     %
@@
     %
-    By using the regret decomposition from Lemma~\ref{lem:2:RegretDecomposition}, this gives
+    By using the regret decomposition from Proposition~\ref{prop:2:RegretDecomposition}, this gives
     \begin{align*}
@@
 We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the estimated regrets\footnote{~First, the rewards are computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{A(t)}(t)$, where $r^{(j)}_{k}(t)$ means the reward stream of the different runs $j=1,\dots,N$.
-Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t)$, as this first formula could give a negative regret and is more noisy, but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N \sum_{k=1}^K \mu_k N^{(j)}_{k}(T)$, as this later formula gives a more robust estimator, as noted the proof of Lemma~\ref{lem:2:RegretDecomposition} (in Page~\pageref{remark:2:moreAccurateCountofRegretForSimulations} above).}
+Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t)$, as this first formula could give a negative regret and is more noisy, but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N \sum_{k=1}^K \mu_k N^{(j)}_{k}(T)$, as this later formula gives a more robust estimator, as noted the proof of Proposition~\ref{prop:2:RegretDecomposition} (in Page~\pageref{remark:2:moreAccurateCountofRegretForSimulations} above).}
 of the different algorithms, \emph{for one simulation}.

### 2019-06-16 1-Preamble/5-Abstract/resume_fr.tex
rude

@@
 Est-ce qu'il faut donner des maths en français pour CHAQUE CHAPITRE ?
-J'ai un peu la flemme d'écrire 20 pages avec des maths en français juste pour pouvoir citer tel théorème ou tel résultat.}
+J'ai peur de ne pas réussir à écrire 20 pages avec des maths en français juste pour pouvoir citer tel théorème ou tel résultat.}
 
@@
 % Chapitre 2
-Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB, Chapitre~\ref{chapter:2}), mais leur analyse est difficile à réaliser,
+Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (MAB, Chapitre~\ref{chapter:2}), mais leur analyse est difficile à réaliser,
 % , Chapitre~\ref{chapter:2}
@@
         %
-        \node[align=center] (introduction) at (0,3.25) [rectangle,draw] {Chapitre~\ref{chapter:1}\\Introduction};
-        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw] {Chapitre~\ref{chapter:2}\\Modèle bandits multi-bras\\stochastique et stationnaire};
-        \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapitre~\ref{chapter:3}\\SMPyBandits : bibliothèque\\de simulations pour MAB};
-        \node[align=center] (chapter25) at (-2.5,2.25) [rectangle,draw] {Chapter~\ref{chapter:25}\\Sélection séquentielle\\d'algorithme MAB};
-        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:4}\\Deux modèles MAB\\pour les réseaux IoT};
-        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapitre~\ref{chapter:5}\\Modèle MAB\\Multi-joueurs};
-        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:6}\\Modèle MAB\\Non-stationnaire};
-        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapitre~\ref{chapter:conclusion}\\Conclusion générale};
-        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw] {Appendices};
+        \node[align=center] (introduction) at (0,3.25) [rectangle,draw,fill=blue!10] {\textbf{Chapitre~\ref{chapter:1}}\\Introduction};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:2}}\\Modèle bandits multi-bras\\stochastique et stationnaire};
+        \node[align=center] (chapter3) at (-2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:3}}\\SMPyBandits : bibliothèque\\de simulations pour MAB};
+        \node[align=center] (chapter25) at (2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:25}}\\Sélection séquentielle\\d'algorithme MAB};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles MAB\\pour les réseaux IoT};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:5}}\\Modèle MAB\\Multi-joueurs};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:6}}\\Modèle MAB\\Non-stationnaire};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!10] {\textbf{Chapitre~\ref{chapter:conclusion}}\\Conclusion générale};
+        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=blue!10] {Appendices};
         %

### 2019-06-16 2-Chapters/1-Chapter/chapter1.tex
rude

@@
 
+% FIXME remove this
+But de ce chapitre :
+le lecteur va savoir en quelques pages
+
+- contexte historique technologique => okay
+
+- les télécoms ? télécoms décentralisées ? => okay
+
+- l'IoT c'est quoi ? exemple d'applications ? => okay ?
+
+- les bandits c'est quoi ?
+
+- les bandits multi-joueurs décentralisés, pourquoi ?
+
+- pourquoi étudier ça ?
+
+- OK qu'est ce qu'on a proposé et pourquoi ? => plan des contributions, bullet point (1ère version, je pourrai faire mieux après)
+
+- OK comment lire la thèse => le plan Figure~\ref{fig:1:organization}
+% FIXME remove this
+
 % ----------------------------------------------------------------------------
@@
 At the root of the problems that motivates this thesis are the questions of global warming, and the increase of the world population.
-During the last 150 years, mankind has developed a wide range of different communications technology, and since the $1920$s wireless communications between devices have been made possible, and more and more frequent in our lives.
+During the last 150 years, mankind has developed a wide range of different communication technologies, and since the $1920$s wireless communications between devices have been made possible, and more and more frequent in our lives.
 With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed world-wide, for a large range of different applications.
@@
 %
-That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of the deployment of IoT networks, by optimizing the wireless communication thanks to embedded decentralized decision making.
+That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing the wireless communication thanks to embedded decentralized decision making.
 
@@
 %
-Historically, three families of wireless communication systems have been deployed: first, central broadcasting (\eg, old TV), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things networks (IoT).
+Historically, three families of wireless communication systems have been deployed: first, centralized broadcasting (\eg, old TV), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things networks (IoT).
 
@@
 
+% OK commencer par un paragraphe qui explique qu'on étudie les télécom dans le but de les rendre automatiquement plus efficace, plus verte, plus adaptative
+
 This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices, it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
@@
 In this thesis, we study Internet of Things networks, and these two constraints of low power consumption (or long battery life) and low duty cycle are essential.
+More precisely, we study the possible interconnections between Cognitive Radio and Machine Learning, applied to IoT networks.
+Let us define and detail both concepts.
 
-\paragraph{Cognitive Radio.}
+% OK commencer par décrire ce sur quoi porte la thèse, en deux phrases qui réduisent le cadre
+
+\paragraph{From IT to Cognitive Radio.}
 %
+We can narrow down the first field of study of this thesis, steps by steps,
+from Information Technologies (IT), to telecommunications, then wireless communications, and finally to Cognitive Radio.
+Cognitive Radio (CR) is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, since the last $20$ years.
+% - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
 
@@
     \item
-    Joseph Mittola \cite{Mitola99}:
-    \emph{``A really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment.}
+    Joseph Mittola in $1999$ said that
+    \emph{``a really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment''} \cite{Mitola99}.
 
     \item
-    Simon Haykin \cite{Haykin05}:
-    \emph{``A CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication.}
+    Then Simon Haykin in $2005$ also said that
+    \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
 \end{itemize}
 
-\newpage
+% FIXME how to explain spectrum scarcity?
+A major problem of current wireless technologies is the issue of spectrum scarcity:
+in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
+% FIXME how to explain PU and SU in a few sentences?
+In such licensed bands, Primary Users (PU) pay to access the network, and they have a priority over 
+But even though the RF bands are allocated, real world measurements often show that some bands are not densely used.
+% FIXME how to explain DSA ?
+The concept of Dynamic Spectrum Access (DSA) was thus proposed in the early $2000$s, 
+% FIXME how to explain OSA ?
+and Opportunistic Spectrum Access.
 
-But de ce chapitre :
-le lecteur va savoir en quelques pages
+\paragraph{From Machine Learning to Reinforcement Learning.}
+%
+Similarly, we can narrow down the second field of study of this thesis, steps by steps.
+Starting from the general concept of Machine Learning (ML), we focused our interest to statistical learning, then sequential learning, then Reinforcement Learning (RL, \cite{SuttonBarto2018}), then RL with limited feedback and finally to multi-armed bandits learning.
+The reference book on RL is \cite{SuttonBarto2018}, so let us quote the definition of RL given by Richard Sutton and Andrew Barto:
+\emph{``Reinforcement learning is learning what to do -— how to map situations to actions -— so
+as to maximize a numerical reward signal. The learner is not told which actions to
+take, but instead must discover which actions yield the most reward by trying them''}.
+% - apprentissage statistique (machine learning) > séquentiel > par renforcement > information partielle (bandits)
 
-- contexte historique technologique => okay
+We illustrate this idea in the following Figure~\ref{fig:1:ReinforcementLearningCycle}, where a player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$,) and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
+The goal of the player is to maximize its reward.
+Many real-world problems can be framed as Reinforcement Learning problems \cite{bouneffouf2019survey}, for instance learning to walk or to drive, learning to play a game, or discovering which treatment is efficient in healing a certain disease.
 
-- les télécoms ? télécoms décentralisées ? => okay
+\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.30\textwidth}{!}{
+        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=1.5]
+            %
+            % We start by placing the blocks
+            \node [block] (player) at (0,0) {Player};
+            % We draw an edge between the player and system block to
+            \node [block] (environment) at (2,0) {Environment};
+            % Once the nodes are placed, connecting them is easy.
+            \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Reward} (player);
+            %
+        \end{tikzpicture}
+    }
+\caption{Reinforcement learning cycle.}
+\label{fig:1:ReinforcementLearningCycle}
+\end{figure}
 
-- l'IoT c'est quoi ? exemple d'applications ? => okay ?
+More precisely, in this thesis we focus on a certain kind of Reinforcement Learning models where the learning does not have access to the entire reaction of the environment after taking its action.
+In other words, we focus on RL with limited feedback, when the player only sees the reward given by its action at every round, and not the reward that would have been given if she chose any other action.
+This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} in details in the next Chapter~\ref{chapter:2}.
+% Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
 
-- les bandits c'est quoi ?
+% Je peux piquer les quelques figures chouettes de Navik (en attendant la version finale ou je referai mes figures)
 
-- les bandits multi-joueurs décentralisés, pourquoi ?
+% Il faut expliquer ce que c'est que la radio intelligente, et le cas de l'OSA
 
-- pourquoi étudier ça ?
+% je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
 
-- OK qu'est ce qu'on a proposé et pourquoi ? => plan des contributions, bullet point (1ère version, je pourrai faire mieux après)
-
-- OK comment lire la thèse => le plan Figure~\ref{fig:1:organization}
-
-\TODOL{Je dois faire ça d'ici mercredi matin}
-% Je peux m'inspirer beaucoup de la thèse de Navik...
-% https://tel.archives-ouvertes.fr/tel-01668536/document
-
-OK commencer par un paragraphe qui explique qu'on étudie les télécom dans le but de les rendre automatiquement plus efficace, plus verte, plus adaptative
-
-OK commencer par décrire ce sur quoi porte la thèse, en deux phrases qui réduisent le cadre
-
-- technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
-- apprentissage statistique (machine learning) > séquentiel > par renforcement > information partielle (bandits)
-
-Je peux piquer les quelques figures chouettes de Navik (en attendant la version finale ou je referai mes figures)
-
-Il faut expliquer ce que c'est que la radio intelligente, et le cas de l'OSA
-
-je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
-
-Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
+\paragraph{Previous work in the SCEE team.}
 
@@
 
+\paragraph{Current challenges.}
+%
 quels nouveaux challenges en 2016 ?
@@
 
+\paragraph{Our approach.}
+%
 directions de recherches
@@
 
-\subsection{Reinforcement learning}
-
-\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
-\begin{figure}[h!]
-    \centering
-    \resizebox{0.40\textwidth}{!}{
-        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=2]
-            %
-            % We start by placing the blocks
-            \node [block] (player) at (0,0) {Player};
-            % We draw an edge between the player and system block to
-            \node [block] (environment) at (2,0) {Environment};
-            % Once the nodes are placed, connecting them is easy.
-            \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
-            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Reward} (player);
-            %
-        \end{tikzpicture}
-    }
-\caption{Reinforcement learning cycle.}
-\label{fig:1:ReinforcementLearningCycle}
-\end{figure}
-
-\hline{}
-\newpage
+\TODOL{Je dois faire ça d'ici mercredi matin}
+% Je peux m'inspirer beaucoup de la thèse de Navik...
+% https://tel.archives-ouvertes.fr/tel-01668536/document
 
@@
         %
-        \node[align=center] (introduction) at (0,3.25) [rectangle,draw] {Chapter~\ref{chapter:1}\\Introduction};
-        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw] {Chapter~\ref{chapter:2}\\Stochastic and Stationary\\Multi-Armed Bandit models};
-        \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapter~\ref{chapter:3}\\SMPyBandits: simulation\\library for MAB};
-        \node[align=center] (chapter25) at (-2.5,2.25) [rectangle,draw] {Chapter~\ref{chapter:25}\\Online selection\\of the best algorithm};
-        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapter~\ref{chapter:4}\\Two MAB models\\for IoT networks};
-        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapter~\ref{chapter:5}\\Multi-players\\Multi-Armed Bandits};
-        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapter~\ref{chapter:6}\\Non-stationary\\Multi-Armed Bandits};
-        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapter~\ref{chapter:conclusion}\\General Conclusion};
-        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw] {Appendix};
+        \node[align=center] (introduction) at (0,3.25) [rectangle,draw,fill=blue!10] {\textbf{Chapter~\ref{chapter:1}}\\Introduction};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapter~\ref{chapter:2}}\\Stochastic and Stationary\\Multi-Armed Bandit models};
+        \node[align=center] (chapter3) at (-2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapter~\ref{chapter:3}}\\SMPyBandits: simulation\\library for MAB};
+        \node[align=center] (chapter25) at (+2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapter~\ref{chapter:25}}\\Online selection\\of the best algorithm};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapter~\ref{chapter:4}}\\Two MAB models\\for IoT networks};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!10] {\textbf{Chapter~\ref{chapter:5}}\\Multi-players\\Multi-Armed Bandits};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapter~\ref{chapter:6}}\\Non-stationary\\Multi-Armed Bandits};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!10] {\textbf{Chapter~\ref{chapter:conclusion}}\\General Conclusion};
+        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=blue!10] {Appendix};
         %
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    February $2019$,\\
+    February $2019$.
+    See Chapter~\ref{chapter:6}.
     Preprint at \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
-    See Chapter~\ref{chapter:6}.
     \cite{Besson2019GLRT}
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    September $2018$,\\
+    September $2018$.
+    See Appendix~\ref{app:2:DoublingTricks}.
     Preprint at \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.
-    See Appendix~\ref{app:2:DoublingTricks}.
     \cite{Besson2018DoublingTricks}

### 2019-06-16 2-Chapters/2-Chapter/chapter2.tex
rude
Proofreading Ch2

@@
 %
-Such models were less studied than stationary or adversarial models, but some interesting works focused on Markov models in the last 10 years\footnote{~
-For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.}.
+Such models were less studied than stationary or adversarial models, but some interesting works focused on Markov models in the last 10 years.
+For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
 A curious reader about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and then refer to the famous book \cite{Norris98}.
@@
     of empirical mean $\widehat{Z_n} \eqdef \frac{1}{n} \sum\limits_{i=1}^n Z_i$, \emph{Hoeffding's inequality} gives both
-    \vspace*{-10pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    % \vspace*{-10pt}  % FIXME vspace negative to compensate the \end{leftbar}
     \begin{align}\label{eq:2:HoeffdingInequality}
@@
 %
-\vspace*{-10pt}  % FIXME vspace negative
 \begin{smallproof}\label{proof:2:UCBregretBound}
-    As we focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives the inequality \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
+    We focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives the inequality \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
     Thanks to the regret decomposition from Proposition~\ref{prop:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
@@
 
-    Remember that $\UCB_a(t) = \widehat{\mu_a}(t) + \xi_a(t)$ for any arm $a$.
+    Remember that $\UCB_k(t) = \widehat{\mu_k}(t) + \xi_k(t)$ for any arm $k$.
     According to the Algorithm~\ref{algo:2:indexPolicy} (see line~6),
     a suboptimal arm $k$ can be chosen only if $\UCB_k(t) \geq \UCB_{k^*}(t)$.
-    Both $\widehat{\mu_a}(t)$ and $\xi_a(t)$ use $n=N_k(t)$ samples of arm $a$,
+    Both $\widehat{\mu_k}(t)$ and $\xi_k(t)$ use $n=N_k(t)$ samples of arm $k$,
     but as \eqref{eq:2:trickInUCBProof} above started to isolate $N_k(t)$, we add the freedom of considering different number of samples (using $n$), so we introduce the notation
-    $\widehat{\mu}_{a,n}(t)$ and $\xi_{a,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
-    \ie, $\widehat{\mu}_{a,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{a,n}(t) = \sqrt{2\log(t) / n}$.
+    $\widehat{\mu}_{k,n}(t)$ and $\xi_{k,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
+    \ie, $\widehat{\mu}_{k,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{k,n}(t) = \sqrt{2\log(t) / n}$.
     %
@@
         & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr\left( \widehat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right) \nonumber \\
-        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t \bigl[ \underbrace{\Pr(I_1)}_{=1/t^4} + \underbrace{\Pr(I_2)}_{=1/t^4} + \underbrace{\Pr(I_3)}_{=0} \bigr] \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t \bigl[ \underbrace{\Pr(I_1)}_{\leq 1/t^4} + \underbrace{\Pr(I_2)}_{\leq 1/t^4} + \underbrace{\Pr(I_3)}_{=0} \bigr] \nonumber \\
         & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \nonumber \\
@@
 
-We see that the $\UCB_1$ algorithm has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO((\sum_{k: \Delta_k>0} \frac{1}{\Delta_k})\log(T))$.
+We see in this Theorem~\ref{thm:2:UCBregretBound} that $\UCB_1$ has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO((\sum_{k: \Delta_k>0} \frac{1}{\Delta_k})\log(T))$.
 %
@@
 Instead of using a confidence radius $\xi_k(t)$ based on Hoeffding's inequality,
-the \klUCB{} algorithm rather use the Chernoff concentration inequality for the considered exponential family
+the \klUCB{} algorithm rather uses the Chernoff concentration inequality for the considered exponential family
 % such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
@@
 
-\paragraph{A note about the implementation of \klUCB{} algorithms.}
+\textbf{A note about the implementation of \klUCB{} algorithms.}
 %
@@
 We present here some solutions that are index-based (see Algorithm~\ref{algo:2:indexPolicy}).
-The first example is MOSS from \cite{Audibert2009minimax}, or \KLUCBpp{} in \cite{Menard17},
-and they use the following indexes:
-\begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp}
+The first example is MOSS from \cite{Audibert2009minimax}, or \KLUCBpp{} in \cite{Menard17}.
+If we denote $g(t, T, K) = \log^+(y(t) (1 + \log^+(y(t))^2))$, and $y(t) = \frac{T}{K t}$, with $\log^+(x) = \max(0, \log(x))$,
+then they use the following indexes:
+\begin{align*}\label{eq:2:indexes_for_MOSS_klUCBpp}
     U^{\mathrm{MOSS}}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{t}{K N_k(t)}\right)}{N_k(t)} } \right). \\
-% \end{align}
-% \begin{align}\label{eq:2:indexes_for_MOSS_klUCBpp2}
-    U^{\KLUCBpp}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
-\end{align}
-%
-If we denote $g(t, T, K) = \log^+(y(t) (1 + \log^+(y(t))^2))$, and $y(t) = \frac{T}{K t}$, with $\log^+(x) = \max(0, \log(x))$.
+% \end{align*}
+% \begin{align*}\label{eq:2:indexes_for_MOSS_klUCBpp2}
+    U^{\KLUCBpp}_k(t) & \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{g(N_k(t), T, K)}{N_k(t)} \right\}.
+\end{align*}
 
@@
 enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, that is, matching the $C_I \ln(T)$ lower bound by \cite{LaiRobbins85}.
-It is an index policy that uses a modified version of the indexes of the \klUCB$^+$ algorithm (see \cite{Honda2019} for more details), for arms that are not sampled much, and a modified version of the indexes of the MOSS$^+$ algorithm, for the other arms.
-To detail this, first let $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,
-then \klUCB-switch use the index defined as follows:
+It is an index policy that uses modified versions of the indexes of \klUCB$^+$ for arms that are not sampled much (see \cite{Honda2019} for more details), and of MOSS$^+$ for the other arms.
+If the two indexes
+% $U^{\KLUCB+}_k(t)$ and $U^{\mathrm{MOSS}+}_k(t)$
+are defined by
+$U^{\mathrm{MOSS}+}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \log\left(T / (K N_k(t))\right) / N_k(t)} \right)$
+and
+$U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \log(T / (K N_k(t))) / N_k(t) \right\}$,
+and if $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,
+then \klUCB-switch uses the index defined as follows:
 \begin{align*}
@@
 \end{align*}
-
-If the two indexes $U^{\KLUCB+}_k(t)$ and $U^{\mathrm{MOSS}+}_k(t)$ are defined by
-%
-% \vspace*{-20pt}   % FIXME??
-\begin{align*}
-    U^{\mathrm{MOSS}+}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
+%
+
+%
+% \begin{align*}
+%     U^{\mathrm{MOSS}+}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \frac{\log\left(\frac{T}{K N_k(t)}\right)}{N_k(t)}} \right). \\
+% % \end{align*}
+% % \begin{align*}
+%     U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\}
+% % \end{align*}
+% % \begin{align*}
+%     % U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
+%     %     U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
+%     %     U^{\mathrm{MOSS}+}_k(t) & \text{if } N_k(t) > f(T, K).
+%     % \end{cases}.
 % \end{align*}
-% \begin{align*}
-    U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [a, b]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{\log(\frac{T}{K N_k(t)})}{N_k(t)} \right\}
-% \end{align*}
-% \begin{align*}
-    % U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
-    %     U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
-    %     U^{\mathrm{MOSS}+}_k(t) & \text{if } N_k(t) > f(T, K).
-    % \end{cases}.
-\end{align*}
 
@@
 
-\TODOL{Si je dois gagner de la place, je peux ne pas inclure cette section...}
+% \TODOL{Si je dois gagner de la place, je peux ne pas inclure cette section...}
 
@@
 It should not be surprising that the black line of the lower-bound is \emph{above} the values of the regret as the lower-bound is only asymptotic.}
-of the average rewards (\ie, $\frac{1}{t} \sum\limits_{s=1}^t r(s)$), which should converge to $\mu^*$ when $t\to\infty$ for efficient algorithms.
+of the average rewards
+(\ie, $\frac{1}{t} (r(1) + \dots + r(t))$),
+% (\ie, $\frac{1}{t} \sum\limits_{s=1}^t r(s)$),
+which should converge to $\mu^*$ when $t\to\infty$ for efficient algorithms.
 %
@@
 	% \centering
-    \includegraphics[width=1.15\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
+    \includegraphics[width=1.14\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
 	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
@@
 	% \centering
-	\includegraphics[width=1.15\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
+	\includegraphics[width=1.14\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
 	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
@@
 
-\newpage  % WARNING ?
+% \newpage  % WARNING ?
 

### 2019-06-17 2-Chapters/1-Chapter/chapter1.tex
Illustrations

@@
 Starting from the general concept of Machine Learning (ML), we focused our interest to statistical learning, then sequential learning, then Reinforcement Learning (RL, \cite{SuttonBarto2018}), then RL with limited feedback and finally to multi-armed bandits learning.
+As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
 The reference book on RL is \cite{SuttonBarto2018}, so let us quote the definition of RL given by Richard Sutton and Andrew Barto:
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.45\textwidth]{Illustration_of_Machine_Learning_and_Reinforcement_Learning.jpg}
+\caption{Main types of Machine Learning, \copyright{} from \href{https://vas3k.com/blog/machine_learning/}{\texttt{VAS3k.com/blog/machine\_learning}}.}
+\label{fig:1:MainTypesOfMachineLearning}
+\end{figure}
+
+% Modified from the https://draw.io/ map from
+% https://github.com/trekhleb/homemade-machine-learning/blob/master/images/machine-learning-map.xml
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.90\textwidth]{machine-learning-map.pdf}
+\caption{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.}
+\label{fig:1:MainTypesOfMachineLearning2}
+\end{figure}
+
 We illustrate this idea in the following Figure~\ref{fig:1:ReinforcementLearningCycle}, where a player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$,) and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
@@
 
-\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
 \begin{figure}[h!]
     \centering
-    \resizebox{0.30\textwidth}{!}{
-        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=1.5]
-            %
-            % We start by placing the blocks
-            \node [block] (player) at (0,0) {Player};
-            % We draw an edge between the player and system block to
-            \node [block] (environment) at (2,0) {Environment};
-            % Once the nodes are placed, connecting them is easy.
-            \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
-            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Reward} (player);
-            %
-        \end{tikzpicture}
-    }
-\caption{Reinforcement learning cycle.}
+    \includegraphics[width=0.20\textwidth]{Illustration_of_Reinforcement_Learning_action-reward_loop.png}
+    \caption{Reinforcement learning cycle, a learner interacts with its environment through actions, and observes a reward, iteratively.
+\copyright{} from \href{https://vas3k.com/blog/machine_learning/}{\texttt{VAS3k.com/blog/machine\_learning}}.}
     \label{fig:1:ReinforcementLearningCycle}
 \end{figure}
+
+% \tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
+% \begin{figure}[h!]
+%     \centering
+%     \resizebox{0.30\textwidth}{!}{
+%         \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=1.5]
+%             %
+%             % We start by placing the blocks
+%             \node [block] (player) at (0,0) {Player};
+%             % We draw an edge between the player and system block to
+%             \node [block] (environment) at (2,0) {Environment};
+%             % Once the nodes are placed, connecting them is easy.
+%             \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
+%             \draw [->] (environment) to[bend left=90] node[pos=0.5] {Reward} (player);
+%             %
+%         \end{tikzpicture}
+%     }
+% \caption{Reinforcement learning cycle.}
+% \label{fig:1:ReinforcementLearningCycle}
+% \end{figure}
 

### 2019-06-17 2-Chapters/2-Chapter/chapter2.tex
Illustrations

@@
 % \TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
-We find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
+Because it is relevant for cognitive radio applications, as a hardware reconfiguration costs energy, and so changing channel from two consecutive time steps has a non trivial energy cost, as highlighted in \cite{modiDemo2016}.
+And thus we find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
+Our proposed algorithm, \MCTopM, achieves order-optimal logarithmic regret as well as a logarithmic number of switches.
 % Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
@@
 
-This Section starts by discussing two naive strategies which fail dramatically,
-%
-and then we present two simple strategies which performs efficiently, only if they are tuned using a prior knowledge of the problem difficulty, but are thus unusable for practical applications.
-%
-We focus then on index policies, with a first efficient policy, $\UCB_1$, that uses upper confidence bounds (UCB) on the means estimates, and is known to be order-optimal for bounded rewards.
-Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of $\UCB_1$ but use tighter confidence intervals and thus it typically obtains better theoretical results, and Thompson sampling which replaces the principle of ``optimism under uncertainty'' by a Bayesian point-of-view.
-%
-Finally, we conclude by briefly exposing other families of algorithms.
+This Section reviews the most important families of stochastic MAB algorithms, from naive and simple strategies, to strategies based on the optimism or the Bayesian principles, to recent ``best-of-both-worlds'' strategies.
+% starts by discussing two naive strategies which fail dramatically,
+% %
+% and then we present two simple strategies which performs efficiently, only if they are tuned using a prior knowledge of the problem difficulty, but are thus unusable for practical applications.
+% %
+% We focus then on index policies, with a first efficient policy, $\UCB_1$, that uses upper confidence bounds (UCB) on the means estimates, and is known to be order-optimal for bounded rewards.
+% Two other well-known and optimal algorithms are exposed: \klUCB{} extends the idea of $\UCB_1$ but use tighter confidence intervals and thus it typically obtains better theoretical results, and Thompson sampling which replaces the principle of ``optimism under uncertainty'' by a Bayesian point-of-view.
+% %
+% Finally, we conclude by briefly exposing other families of algorithms.
 

### 2019-06-17 2-Chapters/1-Chapter/chapter1.tex
Ajout d'une réf

--- 2019-06-16 2-Chapters/1-Chapter/chapter1.tex
@@
 
-% FIXME remove this
-But de ce chapitre :
-le lecteur va savoir en quelques pages
-
-- contexte historique technologique => okay
-
-- les télécoms ? télécoms décentralisées ? => okay
-
-- l'IoT c'est quoi ? exemple d'applications ? => okay ?
-
-- les bandits c'est quoi ?
-
-- les bandits multi-joueurs décentralisés, pourquoi ?
-
-- pourquoi étudier ça ?
-
-- OK qu'est ce qu'on a proposé et pourquoi ? => plan des contributions, bullet point (1ère version, je pourrai faire mieux après)
-
-- OK comment lire la thèse => le plan Figure~\ref{fig:1:organization}
-% FIXME remove this
-
 % ----------------------------------------------------------------------------
@@
 % Where
+\textbf{Where and when?}
+%
 This doctoral thesis started in October $2016$, and finished in October $2019$.
@@
 % Main problem
+\textbf{Main problems.}
+%
 At the root of the problems that motivates this thesis are the questions of global warming, and the increase of the world population.
-During the last 150 years, mankind has developed a wide range of different communication technologies, and since the $1920$s wireless communications between devices have been made possible, and more and more frequent in our lives.
+% https://en.wikipedia.org/wiki/History_of_radio
+During the last 150 years, mankind has developed a wide range of different communication technologies, and since the late $1890$s wireless communications between devices have been made possible, and more and more frequent in our lives.
 With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed world-wide, for a large range of different applications.
-It is now well-known that with the current trend of population and with the energy crisis, any newly deployed technology should be cheap and energy efficient,
+It is now well-known that with the current trend of population increase and with the energy crisis, any newly deployed technology should be cheap and energy efficient,
 as well as adapted to serve a large number of people and devices.
-%
-That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing the wireless communication thanks to embedded decentralized decision making.
+
+That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms (MAB algorithms),
+directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing the wireless communication thanks to embedded decentralized decision making.
 
@@
 Broadcast-based systems are not only used for TV or radio,
-and for instance two other famous and world-wide deployed examples are the Global Positioning System (GPS) and infrared remote controllers.
+and for instance two other famous and world-wide deployed examples are the long-range Global Positioning System (GPS), and short-range infrared remote controllers.
 
 % Two ways communications
-However, this kind of mono-directional wireless communications were soon found to not be satisfactory for many applications, and thus bi-directional standards were defined from the $1940$s.
-First developed by and for the military, such applications have reached general public usage since the $1970$s.
+However, this kind of mono-directional wireless communications were soon found to not be satisfactory for many applications, and thus bi-directional standards were defined from the $1910$s.
+First developed by and for the military, and the navy (\eg, the Titanic sent the first SOS signal over radio in $1912$),
+such applications have reached general public usage since the $1950$s.
 The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people lives in the richer countries in the world, with the booming of cellphones since the $2000$s and smartphones since the $2010$s.
@@
 Another common constraint for IoT devices is their low duty cycle, as most applications target one to a few messages sent by day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
-%
-In this thesis, we study Internet of Things networks, and these two constraints of low power consumption (or long battery life) and low duty cycle are essential.
-More precisely, we study the possible interconnections between Cognitive Radio and Machine Learning, applied to IoT networks.
-Let us define and detail both concepts.
 
@@
 
-\paragraph{From IT to Cognitive Radio.}
-%
-We can narrow down the first field of study of this thesis, steps by steps,
-from Information Technologies (IT), to telecommunications, then wireless communications, and finally to Cognitive Radio.
-Cognitive Radio (CR) is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, since the last $20$ years.
-% - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
+\section{Cognitive Radio and Multi-Armed Bandits}
 
-\begin{itemize}
-    \item
-    Joseph Mittola in $1999$ said that
-    \emph{``a really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment''} \cite{Mitola99}.
+In this thesis, we study Internet of Things networks, and these two constraints of low power consumption (or long battery life) and low duty cycle are essential.
+More precisely, we study the possible interconnections between \textbf{Machine Learning} and \textbf{Cognitive Radio} applied for IoT networks.
+Let us define and detail both concepts.
 
-    \item
-    Then Simon Haykin in $2005$ also said that
-    \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
-\end{itemize}
-
-% FIXME how to explain spectrum scarcity?
-A major problem of current wireless technologies is the issue of spectrum scarcity:
-in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
-% FIXME how to explain PU and SU in a few sentences?
-In such licensed bands, Primary Users (PU) pay to access the network, and they have a priority over 
-But even though the RF bands are allocated, real world measurements often show that some bands are not densely used.
-% FIXME how to explain DSA ?
-The concept of Dynamic Spectrum Access (DSA) was thus proposed in the early $2000$s, 
-% FIXME how to explain OSA ?
-and Opportunistic Spectrum Access.
+% This thesis studies the possible applications of a certain kind of Machine Learning algorithms, Multi-Armed Bandits, to a certain kind of Telecommunications, Cognitive Radio and Internet of Things of Networks.
 
@@
 %
-Similarly, we can narrow down the second field of study of this thesis, steps by steps.
+We can narrow down the first field of study of this thesis, steps by steps.
 Starting from the general concept of Machine Learning (ML), we focused our interest to statistical learning, then sequential learning, then Reinforcement Learning (RL, \cite{SuttonBarto2018}), then RL with limited feedback and finally to multi-armed bandits learning.
+As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
 The reference book on RL is \cite{SuttonBarto2018}, so let us quote the definition of RL given by Richard Sutton and Andrew Barto:
@@
 
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.45\textwidth]{Illustration_of_Machine_Learning_and_Reinforcement_Learning.jpg}
+% \caption{Main types of Machine Learning, \copyright{} from \href{https://vas3k.com/blog/machine_learning/}{\texttt{VAS3k.com/blog/machine\_learning}}.}
+% \label{fig:1:MainTypesOfMachineLearning2}
+% \end{figure}
+
+% Modified from the https://draw.io/ map from
+% https://github.com/trekhleb/homemade-machine-learning/blob/master/images/machine-learning-map.xml
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.90\textwidth]{machine-learning-map.pdf}
+\caption{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.}
+\label{fig:1:MainTypesOfMachineLearning}
+\end{figure}
+
 We illustrate this idea in the following Figure~\ref{fig:1:ReinforcementLearningCycle}, where a player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$,) and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
@@
 Many real-world problems can be framed as Reinforcement Learning problems \cite{bouneffouf2019survey}, for instance learning to walk or to drive, learning to play a game, or discovering which treatment is efficient in healing a certain disease.
+
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.25\textwidth]{Illustration_of_Reinforcement_Learning_action-reward_loop.png}
+%     \caption{Reinforcement learning cycle, a learner interacts with its environment through actions, and observes a reward, iteratively.
+\copyright{} from \href{https://vas3k.com/blog/machine_learning/}{\texttt{VAS3k.com/blog/machine\_learning}}.}
+%     \label{fig:1:ReinforcementLearningCycle}
+% \end{figure}
 
@@
     }
-\caption{Reinforcement learning cycle.}
+\caption{Reinforcement learning cycle: a learner interacts with its environment through actions, and observes a reward, iteratively.}
 \label{fig:1:ReinforcementLearningCycle}
@@
 
-% Je peux piquer les quelques figures chouettes de Navik (en attendant la version finale ou je referai mes figures)
+\paragraph{From IT to Cognitive Radio.}
+%
+Similarly, we can narrow down the second field of study of this thesis, steps by steps:
+from Information Technologies (IT), to telecommunications, then wireless communications, and finally to Cognitive Radio.
+Cognitive Radio (CR) is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
+% - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
+%
+% je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
+\begin{itemize}
+    \item
+    Joseph Mittola in $1999$ said that
+    \emph{``a really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment''} \cite{Mitola99}.
+
+    \item
+    Then Simon Haykin in $2005$ also said that
+    \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
+    He was the first to propose to use CR for dynamic spectrum access.
+
+    \item
+    The \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia page on Cognitive Radio} states
+    \emph{``A cognitive radio (CR) is a radio that can be programmed and configured dynamically to use the best wireless channels in its vicinity to avoid user interference and congestion.
+Such a radio automatically detects available channels in wireless spectrum, then accordingly changes its transmission or reception parameters to allow more concurrent wireless communications in a given spectrum band at one location''}.
+\end{itemize}
+
+\textbf{The spectrum scarcity issue.}
+%
+% FIXME how to explain spectrum scarcity?
+A major problem of current wireless technologies is the issue of spectrum scarcity:
+in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
+%
+% FIXME copy paste from Wikipedia
+% Traditional regulatory structures have been built for an analog model and are not optimized for cognitive radio.
+Regulatory bodies in the world, like the Federal Communications Commission in the United States, as well as different independent measurement campaigns found that most radio frequency spectrum was inefficiently utilized
+meaning that while a band can be allocated to a certain usage, it can be free from any user in certain time and/or place.
+For instance, see \cite{valenta2010survey} for a survey on spectrum utilization in Europe.
+%
+Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are insufficiently utilized.
+Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place.
+Moreover, fixed spectrum allocation prevents rarely used frequencies (those assigned to specific services) from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
+Regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands if they would not cause any interference to licensed users.
+These initiatives have focused cognitive-radio research on \textbf{Dynamic Spectrum Access (DSA)}.
+
+% explain PU and SU in a few sentences
+In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the mobile telephony network.
+The PU have a strict priority over any other non-paying users, referred to as Secondary User (SU).
+So even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU is equipped with an efficient sensing capacity, it can analyze its environment, and use a licensed band if it is free of any PU.
+% explain DSA
+This defines the concept of DSA was thus proposed in the early $2000$s, and we refer to the survey \cite{Zhao07} for more details.
+% FIXME how to explain OSA ?
+We focus on one example, the \textbf{Opportunistic Spectrum Access}.
 
 % Il faut expliquer ce que c'est que la radio intelligente, et le cas de l'OSA
-
-% je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
 
@@
 \paragraph{On Multi-Armed Bandit.}
-
-% ----------------------------------------------------------------------------
-\section{Considered problems}
-\label{sec:1:problems}
-
-\TODOL{Write this!}
-
-\subsection{Spectrum scarcity}
-
-\subsection{Low-consumption IoT devices for the future}
-
-% ----------------------------------------------------------------------------
-\section{Considered solutions}
-\label{sec:1:solutions}
-
-\TODOL{Write this!}
-
-\subsection{Online/sequential machine learning to automatically improve wireless communications on the run}
-
-\subsection{Example: Opportunistic Spectrum Access (OSA) in licensed bands}
 
@@
 
+\newpage % WARNING
+% FIXME c'est bon tout ce qu'on a avant ?
+
+% ----------------------------------------------------------------------------
+\section{Our contributions}
+\label{sec:1:contributions}
+
 \TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
@@
 % ----------------------------------------------------------------------------
-\section{Overview of the contributions}
-\label{sec:1:contributions}
+\section{Summary of the contributions}
+\label{sec:1:summaryOfContributions}
 
@@
 
-    \item
-    We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
-    % were  (ALT 2018, and 4-8 works inspired by our article since then).
+    % \item
+    % We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
+    % % were  (ALT 2018, and 4-8 works inspired by our article since then).
 State-of-the-art with our algorithm MCTopM + klUCB for multi-players bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"

### 2019-06-17 2-Chapters/1-Chapter/chapter1.tex
Fixed a typo in Ch2, and better outline for Ch1

@@
 \TODOL{Ecrire les 1.1 et 1.2 en quelques pages}
+\TODOL{J'enlève l'abstract et la minitoc dès que le chapitre est terminé !}
 
@@
 In order to optimize the efficiency of the network, the base station is equipped with decision making algorithms that affects the devices in the cell to the available resources (time slots, RF bands, power etc).
-To be able to communicate with the base station without interfering with each others, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA), or more recently to Non-Orthogonal techniques (NOMA).
+To be able to communicate with the base station without interfering with each others, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA), or Non-Orthogonal techniques only started to emerge recently (\eg, NOMA).
 Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance of the wireless communications to and from the devices.
@@
 This family of wireless systems are referred to as Internet of Things,
-and a typical example of application is for sensor networks.
-
+and a typical example of application is for sensor networks.\\
+\indent
 For the future development of smart grids, smart cities, smart homes, or smart agriculture, sensor networks are promised to be widely deployed.
@@
 For buildings, the main goal is to reduce the cost of heating empty buildings and use temperature sensor networks to get accurate and regular data about the temperature in each room and each floor, and let the centralized heat control optimize its cost and energy consumption.
-For agriculture, one example can be to equipped every cow (in large farms) with a sensors that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and monitors the health of the animals.
-
+For agriculture, one example can be to equipped every cow (in large farms) with a sensors that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and monitors the health of the animals.\\
+%
 % OK commencer par un paragraphe qui explique qu'on étudie les télécom dans le but de les rendre automatiquement plus efficace, plus verte, plus adaptative
-
+\indent
 This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices, it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
@@
 % OK commencer par décrire ce sur quoi porte la thèse, en deux phrases qui réduisent le cadre
+
+\TODOL{Citer \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull} et dire les trucs suivants}
+To quote a report published by the French government, ``Dynamic Spectrum Management For Innovation And Growth'' \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull}:
+\begin{quote}
+    In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy.
+    Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.
+    %
+    Everyone agrees on the growing need for spectrum.
+    This growing need results from two phenomena.
+    On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
+    On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.
+    Today, there are no more available frequencies in the easily exploitable frequency bands.
+    Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.
+    Growing recourse to spectrum sharing, and in particular dynamic spectrum sharing, constitutes an important spectrum reserve.
+    A particular sharing form, the use of unlicensed bands, open to all and free, has been intensely developed with the growth of Wifi use.
+\end{quote}
 
@@
     \centering
-    \includegraphics[width=0.90\textwidth]{machine-learning-map.pdf}
+    \includegraphics[width=0.95\textwidth]{machine-learning-map.pdf}
 \caption{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.}
@@
 We illustrate this idea in the following Figure~\ref{fig:1:ReinforcementLearningCycle}, where a player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$,) and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
-The goal of the player is to maximize its reward.
-Many real-world problems can be framed as Reinforcement Learning problems \cite{bouneffouf2019survey}, for instance learning to walk or to drive, learning to play a game, or discovering which treatment is efficient in healing a certain disease.
+The goal of the player is to maximize its reward, by trials and errors (\ie,actions and rewards).
+Many real-world problems can be framed as Reinforcement Learning problems \cite{bouneffouf2019survey}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease.
 
@@
 
+% Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
 More precisely, in this thesis we focus on a certain kind of Reinforcement Learning models where the learning does not have access to the entire reaction of the environment after taking its action.
@@
 This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} in details in the next Chapter~\ref{chapter:2}.
-% Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
+Clinical trials and treatment discovery was historically the first application of MAB, in $1933$ with the early work of W. Thompson \cite{Thompson33}.
 
@@
 %
-% FIXME how to explain spectrum scarcity?
 A major problem of current wireless technologies is the issue of spectrum scarcity:
-in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
+in most kinds of frequency bands, as illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
 %
-% FIXME copy paste from Wikipedia
+% FIXME copy paste from Wikipedia, change a little bit!
 % Traditional regulatory structures have been built for an analog model and are not optimized for cognitive radio.
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.65\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
+    \caption{A chart representing the allocation of radio spectrum in the United States of (North) America, in $2016$. \copyright{} Wikimedia.}
+    % https://upload.wikimedia.org/wikipedia/commons/c/c7/United_States_Frequency_Allocations_Chart_2016_-_The_Radio_Spectrum.pdf?uselang=fr
+    \label{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}
+\end{figure}
+
 % explain PU and SU in a few sentences
@@
 This defines the concept of DSA was thus proposed in the early $2000$s, and we refer to the survey \cite{Zhao07} for more details.
+We focus on one example, the \textbf{Opportunistic Spectrum Access} (OSA).
+
+\TODOL{Je dois faire ça d'ici mercredi matin}
+
 % FIXME how to explain OSA ?
-We focus on one example, the \textbf{Opportunistic Spectrum Access}.
+\paragraph{Opportunistic Spectrum Access.}
+%
 
@@
 
-\paragraph{Previous work in the SCEE team.}
+\paragraph{Previous work in the SCEE team on OSA.}
 
@@
 
-\paragraph{Current challenges.}
+\paragraph{New challenges for the IoT networks.}
 %
@@
 
-\paragraph{Our approach.}
-%
-directions de recherches
-
-- réfléchir aux modèles, quelles hypothèses sont retirées comparé au modèle OSA
-
-- analyser mathématiquement les modèles
-
-- proposer des solutions : algorithmes, dont la convergence est justifiée, ou heuristiques
-
-- des simulations logicielles
-
-- des démonstrations matérielles
-
-- et en bonus on obtient aussi de belles contributions théoriques
-
 \newpage
 
-\TODOL{Je dois faire ça d'ici mercredi matin}
 % Je peux m'inspirer beaucoup de la thèse de Navik...
 % https://tel.archives-ouvertes.fr/tel-01668536/document
-
-\paragraph{On Cognitive Radio.}
 
@@
 
-\paragraph{On Multi-Armed Bandit.}
-
-\paragraph{Applications for Opportunistic Spectrum Access.}
 % Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
@@
 
-\TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira..}
 In a second PhD thesis supervised by Christophe Moy \cite{Modi17PhD}, Navikkumar Modi studied the impact
@@
 \cite{modiDemo2016}.
-%
-Another interesting works is \cite{Modi17QoS}.
 
@@
 
-\subsection{Example: decentralized learning in unlicensed bands with selfish or orthogonalization-based MAB algorithms}
-
 We evoke here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
@@
 
+We start by formulating the problem studied in this thesis, and then we develop our approach.
+
+\paragraph{Problematic.}
+%
+\begin{quote}
+    \emph{``Can we adapt the decision making tools already successfully applied to Cognitve Radio for Opportunistic Spectrum Access to the specific needs of CR for the future IoT networks''}
+\end{quote}
+
+We answered partially to this problematic by the following research steps.
+
+\paragraph{Part I.}
+
+FIXME
+We explored the rich literature of multi-armed bandits.
+Many algorithms exist.
+So we have to understand which ones can be applied to CR and IoT.
+
+Our exploration of the large number of MAB algorithms, and of MAB models, developed in the recent literature has produced a Python library of simulation of MAB problems.
+%
+We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
+A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.
+
+Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement in a given IoT object.
+FIXME online algorithm selection and aggregation, with our contribution \Aggr{}.
+
+\paragraph{Part II.}
+%
+
+First model \cite{Bonnefoi17}, CROWNCOM paper, model to detail
+
+Verify this first model empirically: proof of concept \cite{Besson2018ICT}
+
+Second model with decision making also for the retransmissions of packets \cite{Bonnefoi2019WCNC}
+
+Selfish
+
+Understand Selfish in the simpler framework of multi-player MAB
+
+In the OSA setting, we failed
+
+But we proposed new algorithms for this problem, and relaunched the research on this model \cite{Besson2018ALT}
+
+% \paragraph{Our approach.}
+% %
+% directions de recherches
+
+% - réfléchir aux modèles, quelles hypothèses sont retirées comparé au modèle OSA
+
+% - analyser mathématiquement les modèles
+
+% - proposer des solutions : algorithmes, dont la convergence est justifiée, ou heuristiques
+
+% - des simulations logicielles
+
+% - des démonstrations matérielles
+
+% - et en bonus on obtient aussi de belles contributions théoriques
+
+% FIXME c'est bon tout ce qu'on a avant ?
+
 \TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
@@
 \newpage % WARNING
-% FIXME c'est bon tout ce qu'on a avant ?
 

### 2019-06-17 2-Chapters/2-Chapter/chapter2.tex
Fixed a typo in Ch2, and better outline for Ch1

@@
 are defined by
-$U^{\mathrm{MOSS}+}_k(t) & \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \log\left(T / (K N_k(t))\right) / N_k(t)} \right)$
+$U^{\mathrm{MOSS}+}_k(t) \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \log\left(T / (K N_k(t))\right) / N_k(t)} \right)$
 and
-$U^{\KLUCB+}_k(t) & \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \log(T / (K N_k(t))) / N_k(t) \right\}$,
+$U^{\KLUCB+}_k(t) \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \log(T / (K N_k(t))) / N_k(t) \right\}$,
 and if $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,

### 2019-06-17 2-Chapters/1-Chapter/chapter1.tex
J'essaie de terminer cette revue des travaux sur OSA

@@
 
-\paragraph{From Machine Learning to Reinforcement Learning.}
+\paragraph{From Machine Learning to Multi-Armed Bandits.}
 %
 We can narrow down the first field of study of this thesis, steps by steps.
-Starting from the general concept of Machine Learning (ML), we focused our interest to statistical learning, then sequential learning, then Reinforcement Learning (RL, \cite{SuttonBarto2018}), then RL with limited feedback and finally to multi-armed bandits learning.
+Starting from the general concept of Machine Learning (ML), we focused our interest to statistical learning, then sequential learning, then Reinforcement Learning (RL), then RL with limited feedback and finally to Multi-Armed Bandits (MAB) learning.
 As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
@@
 The goal of the player is to maximize its reward, by trials and errors (\ie,actions and rewards).
-Many real-world problems can be framed as Reinforcement Learning problems \cite{bouneffouf2019survey}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease.
+Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial).
 
@@
 Clinical trials and treatment discovery was historically the first application of MAB, in $1933$ with the early work of W. Thompson \cite{Thompson33}.
+%
+Multi-Armed Bandits (MAB) are a simple yet powerful example of the well-known exploration/exploitation dilemma:
+when facing a set of $K$ actions whose effects on the environment are unknown, a learning must balance between
+exploration of the unknown actions, in order to collect more information about them,
+and exploitation of the best action according to its current knowledge.
 
@@
 A major problem of current wireless technologies is the issue of spectrum scarcity:
-in most kinds of frequency bands, as illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
+in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
+As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage) to space research, inter-satellite, and mobile telephony.
 %
-% FIXME copy paste from Wikipedia, change a little bit!
+% FIXME copy paste from Wikipedia, change a little bit?
 % Traditional regulatory structures have been built for an analog model and are not optimized for cognitive radio.
-Regulatory bodies in the world, like the Federal Communications Commission in the United States, as well as different independent measurement campaigns found that most radio frequency spectrum was inefficiently utilized
-meaning that while a band can be allocated to a certain usage, it can be free from any user in certain time and/or place.
-For instance, see \cite{valenta2010survey} for a survey on spectrum utilization in Europe.
-%
+Regulatory bodies in the world, like the Federal Communications Commission in the USA, as well as different independent measurement campaigns found that most radio frequency spectrum was inefficiently utilized,
+meaning that while a band can be allocated to a certain usage, it can be free from any user in certain times and/or places.
+We refer to \cite{valenta2010survey} for a survey on spectrum utilization in Europe.
+
 Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are insufficiently utilized.
@@
 Regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands if they would not cause any interference to licensed users.
-These initiatives have focused cognitive-radio research on \textbf{Dynamic Spectrum Access (DSA)}.
+These initiatives have focused CR research on \textbf{Dynamic Spectrum Access (DSA)}.
 
@@
     \centering
-    \includegraphics[width=0.65\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
-    \caption{A chart representing the allocation of radio spectrum in the United States of (North) America, in $2016$. \copyright{} Wikimedia.}
+    \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
+    \caption{A chart representing the allocation of radio spectrum in the USA in $2016$.
+Only a small share of the spectrum is not yet allocated, the top-left corner in white (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz}).
+\copyright{} Wikimedia.}
     % https://upload.wikimedia.org/wikipedia/commons/c/c7/United_States_Frequency_Allocations_Chart_2016_-_The_Radio_Spectrum.pdf?uselang=fr
@@
 This defines the concept of DSA was thus proposed in the early $2000$s, and we refer to the survey \cite{Zhao07} for more details.
-We focus on one example, the \textbf{Opportunistic Spectrum Access} (OSA).
+We focus on one example, the \textbf{Opportunistic Spectrum Access} (OSA) problem.
+
+\paragraph{Opportunistic Spectrum Access.}
+%
+In OSA, the focus is on spectrum access for one SU accessing a licensed spectrum.
+%
+The following hypothesis are maid:
+the SU is equipped with sensing,
+and a consider a fixed and finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
+For instance, it can be a set of three Wi-Fi channels at different frequency, emitted by the same Wi-Fi station.
+Another hypothesis is that both PU and SU are synchronized in times, by sub-dividing the time in discrete time steps.
+%
+Thus if the SU spends a short time in the beginning of each time slot to do sensing, it can scan for the presence or absence of any PU before transmitting.
+If a SU was able to sense for all the $K$ frequency bands, it could simply transmit in one of the free channels if any, or not transmit if all channels are used at a given time step.
+However, sensing is known to be error-prone and costly, especially for wide-band sensing, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
+This assumption, along with the regulation that PU cannot be disturbed, enforces the SU to transmit in the channel that it sensed.
+% non negligible cost of radio reconfiguration
+
+Focussing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of $K\geq2$ channels), then it performs sensing, and finally it transmits in this channel if it is free.
+The goal of the SU is to minimize its energy consumption (we remind that we focus on green radio) and to maximize its uplink data rate, or equivalently, to maximize its number of successful transmission.
+%
+If the different channels are not uniformly used by the PU, and if we assume a stationary hypothesis on the PU traffic, then the goal of the SU boils down to explore the different channels and exploiting the best ones.
+This frames the Opportunistic Spectrum Access problem as a exploration/exploitation problem, white a finite set of actions (the channels, also called arms),
+in a sequential action-then-feedback cycle (see Figure~\ref{fig:1:ReinforcementLearningCycle}, time steps are $t\in\N^*$),
+under partial information feedback.
+These three hypotheses are the ones that restrict from the general Reinforcement Learning framework to the specific Multi-Armed Bandit case.
 
 \TODOL{Je dois faire ça d'ici mercredi matin}
+\paragraph{MAB for OSA.}
+%
+Previous works of our SCEE team showed that MAB can be used to model the OSA problem:
+orthogonal frequency bands (or channels) are models by arms $k\in\{1,\dots,K\}$,
+and the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$ is modeled by a reward of $r(t) \in \{0,1\}$.
+Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus an uplink message can be sent), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message should be sent.
+%
+This model was first studied by Wassim Jouini during his PhD thesis, ten years ago
+in \cite{Jouini09}, and later studied in both \cite{Jouini10,Jouini12}.
+His works with Christophe Moy were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the \UCB{} algorithm,
+along the early works of \cite{Liu08,Zhao10}.
+%
+Shortly after in $2014$, proof-of-concepts using real-world radio hardware were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
 
-% FIXME how to explain OSA ?
-\paragraph{Opportunistic Spectrum Access.}
-%
-
-% Il faut expliquer ce que c'est que la radio intelligente, et le cas de l'OSA
-
-\paragraph{Previous work in the SCEE team on OSA.}
-
-expliquer ce que Christophe Moy, Wassim Jouini, Navik Modi etc ont fait depuis 2009
-
-\paragraph{New challenges for the IoT networks.}
-%
-quels nouveaux challenges en 2016 ?
-
-autre direction
-
-- radio intelligente > internet des objets
-
-\newpage
-
-% Je peux m'inspirer beaucoup de la thèse de Navik...
-% https://tel.archives-ouvertes.fr/tel-01668536/document
-
-Des thèses précédentes avaient déjà identifié l'apprentissage séquentiel comme étant pertinent pour la radio intelligente.
-On peut citer les travaux de \cite{Jouini09,Jouini10,Jouini12}, qui ont notamment souligné les points suivants :
-
-\begin{itemize}
-    \item implémentation peu couteuse en ressources d'exécution (temps et mémoire), contrairement aux autres approches de décision basées sur l'optimisation, ou d'autres techniques d'IA,
-    \item capacité à apprendre depuis zéro (pas de phase de pré apprentissage durant laquelle le système ne fait rien d'autre qu'apprendre), et forte capacité à s'adapter automatiquement à un nouveau contexte,
-    \item coordination décentralisée possible de plusieurs joueurs, sans qu'il n'y ait de communication autorisée entre eux,
-    \item approche cohérente dans un système décentralisé, si chaque objet sans fil apprend de son côté plutôt que de recevoir des instructions d'une station de base,
-    \item résistent aux erreurs d'apprentissage
-\end{itemize}
-
-Ces techniques ont aussi permis d'identifier les bandits multi-bras comme le bon modèle d'apprentissage séquentiel statistique dans le contexte de la radio intelligente.
-Des algorithmes de MAB ont été utilisés avec succès dans différents contextes :
-
-\begin{itemize}
-    \item le choix de la configuration radio, avec par exemple \cite{KerkoucheAlami18} récemment,
-    \item l'accès dynamique au spectre, et en particulier l'accès opportuniste (OSA) \cite{Jouini09}.
-\end{itemize}
-
-Mais le cas OSA a plusieurs faiblesses :
-
-\begin{itemize}
-    \item il est trop futuriste, car la régulation des fréquences ne permet pas encore (et en est très loin) de l'appliquer en situation réelle,
-    \item en l'absence de régulation coordonnée, chaque cas d'OSA dépendra de la bande considérée (et comment les objets "licencees" s'y comportent), mais les études de l'équipe SCEE ont montrées que les approches MAB sont capables de s'adapter à tous les cas,
-    \item le retour de décision utilisé dans l'OSA repose sur le sensing, qui est le maillon faible de la boucle (pas encore assez fiable pour être utilisé en vrai).
-    \item d'où l'idée d'utiliser les MAB en Cognitive Radio, dans les bandes non licenciées, dès 2011 \cite{Jouini12}. Dans ce cas, le contexte est comme en OSA, mais sans PU, et pas de sensing.
-\end{itemize}
-
-Pour toutes ces raisons, le cadre de l'internet des objets (IoT) apparaît comme idéal car c'est un cadre qui s'affranchit des contraintes de l'OSA.
-
-Par conséquent, cette thèse vise à refermer certaines portes qui étaient encore intéressantes à étudier dans la suite de ce qui a été considéré pour l'OSA
-
-% Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
-
-\TODOL{Tous les paragraphes suivants sont copiés collés de plus tard dans la thèse, je peux les réadapter ici, et les supprimer/raccourcir plus loin}
-
-The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc).
-In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
-
-Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
-In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-In the model with sensing, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
-A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
-%
-This model was first introduced by Wassim Jouini during his PhD thesis,
-in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
-Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.
-
-In a second PhD thesis supervised by Christophe Moy \cite{Modi17PhD}, Navikkumar Modi studied the impact
-of using MAB algorithms to optimize channel selection
-on the battery life of a wireless device.
+In a second PhD thesis supervised by Christophe Moy \cite{Modi17PhD} from $2014$ to $2017$, Navikkumar Modi studied the impact of using MAB algorithms to optimize channel selection on the battery life of a wireless device.
 On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
@@
 \cite{modiDemo2016}.
+%
+In $2017$, Christophe Moy continued to work on this direction, with a post-doctoral student, Sumit J. Darak, leading to the following publications
+\cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
+%
+For example, proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
+%
+Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
+%
+Since $2017$, Sumit J.
+Darak and his team at IIIT Delhi have been very active in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+\cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
+
+\newpage % WARNING
+
+\paragraph{Limitations and specifities of IoT networks.}
+%
+The aforementioned previous works have shown that MAB algorithms can be applied with success for the OSA problem.
+But if we consider CR-ready devices that cannot perform sensing, such as low-cost and low energy consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
+Moreover, in most cases, the IoT networks use unlicensed bands, and as such there are no longer a distinction between PU and SU.
+%
+The specificities of IoT networks can be listed as follows.
+\begin{itemize}
+    \item
+    most IoT networks run in unlicensed bands, meaning that this distinction between PU and SU no longer makes sense,
+    \item
+    most IoT devices are very low-cost devices, with low material and computational capacities,
+    meaning that they are not equipped with any sensing capabilities,
+    \item
+    a single IoT base station will have to handle a very large number of devices,
+    and cannot send coordination orders to them in order to optimize the networks
+    \item
+    IoT networks are decentralized
+    \item
+    very low duty cycles (only a few messages a day).
+    % FIXME more specificities
+\end{itemize}
+
+One could think that in the absence of sensing, Reinforcement Learning is no longer possible, but any real IoT device still receives some information about its environment after some or every transmissions.
+In most IoT, either standards for short-range (ZigBee, Z-Wave, Bluetooth) or long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16},
+an up-link message sent to a base station is usually followed by a down-link message sent back by the base station, to indicate if the up-link message was successfully received and understood.
+%
+By using this feedback, that consist in an \emph{acknowledgement} message received shortly after every trial
+
+\newpage % WARNING
+
+% FIXME c'est bon tout ce qu'on a avant ?
+
+% ----------------------------------------------------------------------------
+\section{Our contributions}
+\label{sec:1:contributions}
+
+We start by formulating the problem studied in this thesis, and then we develop our approach.
+
+\paragraph{Problematic.}
+%
+If we sum-up the studied problems and summarize them in one question, it would be the following:
+\emph{``Can we adapt the decision making tools already successfully applied to Cognitve Radio for Opportunistic Spectrum Access to the specific needs of CR for the future IoT networks ?''}
+%
+We answer partially to this problematic by the following research steps.
+
+\paragraph{Focusing on MAB algorithms.}
+
+FIXME
+We explored the rich literature of multi-armed bandits.
+Many algorithms exist.
+So we have to understand which ones can be applied to CR and IoT.
+
+Our exploration of the large number of MAB algorithms, and of MAB models, developed in the recent literature has produced a Python library of simulation of MAB problems.
+%
+We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
+A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.
+
+Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement in a given IoT object.
+FIXME online algorithm selection and aggregation, with our contribution \Aggr{}.
+
+\paragraph{Focusing on IoT networks.}
+%
+
+First model \cite{Bonnefoi17}, CROWNCOM paper, model to detail
 
@@
 
-As detailed in Chapter~\ref{chapter:1},
-MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
-For example,
-proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
-and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
-Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
-% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
-
-We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed but frequency- and time-slotted IoT network.
-Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
-As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly not stationary''.
-This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms (that is why we relax the hypothesis and only consider up-to $M \leq K$ players in Chapter~\ref{chapter:5}).
-We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
-%
-We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
-
-As explained before in Chapter~\ref{chapter:1}, the future IoT networks will require to support more and more communicating devices.
-In this section, we show that intelligent devices in unlicensed bands can use MAB learning algorithms to improve spectral resource exploitation.
-%
-We evaluate the performance of two classical MAB learning algorithms, \UCB{} and Thomson Sampling, to handle the decentralized decision-making of Spectrum Access, applied to IoT networks.
-We mean by \emph{decentralized} that learning is performed on the device side, and is spread on the totality of the devices in a network.
-We also evaluate the learning performance when the number of intelligent end-devices grows.
-%
-We show that using learning algorithms
-firstly extends devices battery autonomy (by reducing the packet-loss-ratio or equivalently, by improving the successful-communication rate),
-and secondly does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
-In the studied settings,
-Stochastic MAB learning
-% provides an up to $16\%$ gain in terms of successful transmission probabilities, and
-has near optimal performance even in non-stationary settings with a majority of intelligent devices.
-
 The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
@@
 We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.
-
-We evoke here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
-Starting from $2010$, the works of Wassim Jouini \cite{Jouini09,Jouini10,Jouini12} were the first ones to propose to use Reinforcement Learning for Cognitive Radio, especially MAB and the \UCB{} algorithm, and from $2014$, proof-of-concepts were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
-In $2015$ and $2016$, Christophe Moy continued to work on this direction, with a PhD student (Navikkumar Modi) and post-doc (Sumit J. Darak) in our team SCEE, for the following publications
-\cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
-Since $2017$, Sumit J.
-Darak and his team at IIIT Delhi have been very active in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
-\cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
-
-\newpage % WARNING
-% FIXME c'est bon tout ce qu'on a avant ?
-
-% ----------------------------------------------------------------------------
-\section{Our contributions}
-\label{sec:1:contributions}
-
-We start by formulating the problem studied in this thesis, and then we develop our approach.
-
-\paragraph{Problematic.}
-%
-\begin{quote}
-    \emph{``Can we adapt the decision making tools already successfully applied to Cognitve Radio for Opportunistic Spectrum Access to the specific needs of CR for the future IoT networks''}
-\end{quote}
-
-We answered partially to this problematic by the following research steps.
-
-\paragraph{Part I.}
-
-FIXME
-We explored the rich literature of multi-armed bandits.
-Many algorithms exist.
-So we have to understand which ones can be applied to CR and IoT.
-
-Our exploration of the large number of MAB algorithms, and of MAB models, developed in the recent literature has produced a Python library of simulation of MAB problems.
-%
-We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
-We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
-A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.
-
-Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement in a given IoT object.
-FIXME online algorithm selection and aggregation, with our contribution \Aggr{}.
-
-\paragraph{Part II.}
-%
-
-First model \cite{Bonnefoi17}, CROWNCOM paper, model to detail
 
@@
 
-% \paragraph{Our approach.}
+\paragraph{Our approach.}
+%
+Our approach for the aforementioned problems is summarized as follows:
+we propose a couple of new models of decision making for IoT networks, on the device's side, for the spectrum access problem.
+By considering the difference between the hypothesis of previously studied models for the OSA problem and our problem at end for IoT, we identify the constraints and specificities of applying decentralized ML for IoT devices.
+%
+Like for the OSA, we can apply successfully model the decision making problem as a Multi-Armed Bandit problem.
+We showed on numerical simulations as well as on a real-world proof-of-concept that applying low-cost sequential learning algorithms coming from the MAB literature successfully reduce the collision rate of IoT end-devices.
+We proposed new algorithms, some of which are proven to converge efficiently.
+
 % %
@@
 
-We can list the following points to summarize the main contributions of this thesis:
+To summarize the main contributions of this thesis, we can list the following points:
 

### 2019-06-17 1-Preamble/5-Abstract/resume_fr.tex
Probably enough for today? ALmost done for everything in Ch1 but the details of our contributions

@@
 
-    \item
-    Le code source des deux contributions susmentionnées est publié en ligne, ainsi que des instructions pour la reproduction de nos travaux (avec GNU Octave ou MATLAB).
+    % \item
+    % Le code source des deux contributions susmentionnées est publié en ligne, ainsi que des instructions pour la reproduction de nos travaux (avec GNU Octave ou MATLAB).
 
@@
         \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!10] {\textbf{Chapitre~\ref{chapter:conclusion}}\\Conclusion générale};
-        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=blue!10] {Appendices};
+        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=blue!10] {Annexes};
         %
@@
 
-L'ordre de lecture du manuscrit peut être n'importe quel chemin, descendant entre l'introduction dans le Chapitre~\ref{chapter:1} actuel, et la conclusion dans le dernier Chapitre~\ref{chapter:conclusion}.
+L'ordre de lecture du manuscrit peut être n'importe quel chemin, descendant entre l'introduction donnée dans Chapitre~\ref{chapter:1}, et la conclusion générale qui constitue le dernier Chapitre~\ref{chapter:conclusion}.
 Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,

### 2019-06-17 2-Chapters/1-Chapter/chapter1.tex
Probably enough for today? ALmost done for everything in Ch1 but the details of our contributions

@@
 
-\TODOL{Ecrire les 1.1 et 1.2 en quelques pages}
-\TODOL{J'enlève l'abstract et la minitoc dès que le chapitre est terminé !}
+\TODOL{Lundi soir : je dois encore relire le 1.1 pour les fautes d'orthographe (et les éventuels erreurs historiques) !}
+\TODOL{Lundi soir : je dois encore terminer le 1.2 et être plus concis ?}
+\TODOL{Je dois écrire le 1.3 en max 2 pages : j'ai déjà le plan, je dois développer}
 
@@
 The organization of this manuscript is detailed in a flow graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but they are all built upon the model and notations introduced on Chapter~\ref{chapter:2}.
+
+\TODOL{J'enlève l'abstract et la minitoc dès que le chapitre est terminé !}
 
@@
 and I was also co-supervised by Doctor Emilie Kaufmann, whom I visited many times at Inria Lille Nord Europe in Lille (France).
+
+\TODOL{Est-ce que je dois détailler plus ? Quel financement j'ai eu (CDSN), quels autres financements m'ont aidés ?}
 
@@
 Another common constraint for IoT devices is their low duty cycle, as most applications target one to a few messages sent by day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
+%
+Many different standards for IoT networks have been proposed,
+and they consist in a specification for both the \emph{PHY}sical
+and the Medium Access Control (\emph{MAC}) layer.
+To quote some examples of standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications, while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications.
+We refer to the survey \cite{Centenaro16} for more details, and references in \cite{Azari18} or our recent article \cite{MoyBesson2019}.
 
 % OK commencer par décrire ce sur quoi porte la thèse, en deux phrases qui réduisent le cadre
-
-\TODOL{Citer \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull} et dire les trucs suivants}
-To quote a report published by the French government, ``Dynamic Spectrum Management For Innovation And Growth'' \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull}:
-\begin{quote}
-    In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy.
-    Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.
-    %
-    Everyone agrees on the growing need for spectrum.
-    This growing need results from two phenomena.
-    On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
-    On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.
-    Today, there are no more available frequencies in the easily exploitable frequency bands.
-    Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.
-    Growing recourse to spectrum sharing, and in particular dynamic spectrum sharing, constitutes an important spectrum reserve.
-    A particular sharing form, the use of unlicensed bands, open to all and free, has been intensely developed with the growth of Wifi use.
-\end{quote}
 
@@
     \includegraphics[width=0.95\textwidth]{machine-learning-map.pdf}
-\caption{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.}
+\caption{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.
+MAB is RL with partial information, and RL is ML with no supervision or data, just an environment to interact with.}
 \label{fig:1:MainTypesOfMachineLearning}
@@
 
-We illustrate this idea in the following Figure~\ref{fig:1:ReinforcementLearningCycle}, where a player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$,) and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
+We illustrate this idea of a \emph{learning cycle} alternating between actions and feedback,
+in the following Figure~\ref{fig:1:ReinforcementLearningCycle}.
+A player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$), and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
 The goal of the player is to maximize its reward, by trials and errors (\ie,actions and rewards).
@@
 Similarly, we can narrow down the second field of study of this thesis, steps by steps:
-from Information Technologies (IT), to telecommunications, then wireless communications, and finally to Cognitive Radio.
-Cognitive Radio (CR) is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
+from Information Technologies (IT), to telecommunications, to wireless communications, then to Software Defined Radio (SDR),
+and finally to Cognitive Radio (CR).
+%
+The transition from the historical approach of hardware-based radio to SDR architectures and is a gradual process that started in the early $1990$s and has accelerated in the $2000$s.
+% https://en.wikipedia.org/wiki/Software-defined_radio
+A SDR is a radio communication system where components that have been traditionally implemented in hardware (e.g.
+mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are instead implemented by means of software on a personal computer or embedded system, with the exception of a unique pair of receiver/transmiter antennas.
+Even though the SDR paradigm was initiated by the USA defense research, over the past years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
+% CR has been the first answer to solve spectrum scarcity issue which collected a huge attention in early 21st century.
+%
+CR is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
 % - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
@@
     \item
-    The \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia page on Cognitive Radio} states
+    The \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia encyclopedia} states
     \emph{``A cognitive radio (CR) is a radio that can be programmed and configured dynamically to use the best wireless channels in its vicinity to avoid user interference and congestion.
 Such a radio automatically detects available channels in wireless spectrum, then accordingly changes its transmission or reception parameters to allow more concurrent wireless communications in a given spectrum band at one location''}.
+\end{itemize}
+
+\TODOL{Je peux améliorer (ou enlever) ce passage qui site J.Toledano, ou alors le commenter en plus de détails ? J'aime bien son rapport en fait (j'aurai du le lire au début de ma thèse !)}
+% https://www.economie.gouv.fr/files/files/PDF/french-spectrum-mission-executive-summary-2014-06-25.pdf
+% https://www.economie.gouv.fr/files/files/PDF/rapport-gestion-dynamique-spectre-2014-06-30.pdf
+The abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by Joëlle Toledano, commissioned and published by the French government \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull},
+mentions different important problems:
+%
+\begin{itemize}\tightlist
+    \item
+    \emph{``In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy''}
+    \item
+    \emph{``Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.''}
+    %
+    \item
+    \emph{``Everyone agrees on the growing need for spectrum.
+    This growing need results from two phenomena.
+    On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
+    On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.''}
+    \item
+    \emph{``Today, there are no more available frequencies in the easily exploitable frequency bands.
+    Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.''}
+    \item
+    \emph{``Growing recourse to spectrum sharing, and in particular dynamic spectrum sharing, constitutes an important spectrum reserve.
+    A particular sharing form, the use of unlicensed bands, open to all and free, has been intensely developed with the growth of Wi-Fi use.''}
 \end{itemize}
@@
 in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
-As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage) to space research, inter-satellite, and mobile telephony.
+As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication) to space research, inter-satellite, mobile telephony and many other applications.
 %
-% FIXME copy paste from Wikipedia, change a little bit?
 % Traditional regulatory structures have been built for an analog model and are not optimized for cognitive radio.
-Regulatory bodies in the world, like the Federal Communications Commission in the USA, as well as different independent measurement campaigns found that most radio frequency spectrum was inefficiently utilized,
-meaning that while a band can be allocated to a certain usage, it can be free from any user in certain times and/or places.
+Regulatory bodies in the world, like the \href{https://www.fcc.gov/}{Federal Communications Commission} in the USA (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
+or \href{http://www.tdf.fr/}{TDF} in France (see \href{https://www.TDF.fr/}{\texttt{TDF.fr}}, previously known as \emph{``TéléDiffusion de France''}), as well as different independent measurement campaigns, found that most radio frequency spectrum was inefficiently utilized,
+meaning that while a band can be allocated to a certain unique usage, it can be free from any user in certain times and/or places.
 We refer to \cite{valenta2010survey} for a survey on spectrum utilization in Europe.
@@
 Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place.
-Moreover, fixed spectrum allocation prevents rarely used frequencies (those assigned to specific services) from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
-Regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands if they would not cause any interference to licensed users.
+Moreover, the fixed spectrum allocation prevents rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
+Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands if they would not cause any interference to licensed users.
 These initiatives have focused CR research on \textbf{Dynamic Spectrum Access (DSA)}.
@@
 %
-In OSA, the focus is on spectrum access for one SU accessing a licensed spectrum.
+In OSA, the focus is on one SU accessing a licensed spectrum.
 %
@@
 This frames the Opportunistic Spectrum Access problem as a exploration/exploitation problem, white a finite set of actions (the channels, also called arms),
-in a sequential action-then-feedback cycle (see Figure~\ref{fig:1:ReinforcementLearningCycle}, time steps are $t\in\N^*$),
+in a sequential action-then-feedback cycle (time steps are $t\in\N^*$),
 under partial information feedback.
-These three hypotheses are the ones that restrict from the general Reinforcement Learning framework to the specific Multi-Armed Bandit case.
+These three hypotheses are the ones that restrict from the general RL framework to the specific Multi-Armed Bandit case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
 
-\TODOL{Je dois faire ça d'ici mercredi matin}
-\paragraph{MAB for OSA.}
+\paragraph{MAB for OSA, and a short presentation of the history of this research at the SCEE team.}
 %
@@
 %
-This model was first studied by Wassim Jouini during his PhD thesis, ten years ago
-in \cite{Jouini09}, and later studied in both \cite{Jouini10,Jouini12}.
-His works with Christophe Moy were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the \UCB{} algorithm,
-along the early works of \cite{Liu08,Zhao10}.
+This model was first studied by Wassim Jouini during his PhD thesis with Christophe Moy, ten years ago, first with \cite{Jouini09} and later with \cite{Jouini10,Jouini12}.
+His works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the \UCB{} algorithm,
+along the early works of Qing Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
 %
-Shortly after in $2014$, proof-of-concepts using real-world radio hardware were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
+Shortly after in $2014$, proof-of-concepts using real-world radio hardware and Software Defined Radio were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
 
@@
 
-\newpage % WARNING
+\TODOL{Raccourcir toutes les fois où je parle des travaux sur l'OSA après dans la suite de la thèse !}
 
@@
     \item
-    most IoT networks run in unlicensed bands, meaning that this distinction between PU and SU no longer makes sense,
+    most IoT networks run in unlicensed bands (no more distinction between PU and SU),
     \item
@@
     a single IoT base station will have to handle a very large number of devices,
-    and cannot send coordination orders to them in order to optimize the networks
+    and cannot send coordination orders to them in order to optimize the network,
     \item
-    IoT networks are decentralized
-    \item
-    very low duty cycles (only a few messages a day).
-    % FIXME more specificities
+    IoT devices have low duty cycles (only a few messages every minute) to very low duty cycles (only a few messages a day).
+    % FIXME more specificities?
 \end{itemize}
@@
 One could think that in the absence of sensing, Reinforcement Learning is no longer possible, but any real IoT device still receives some information about its environment after some or every transmissions.
-In most IoT, either standards for short-range (ZigBee, Z-Wave, Bluetooth) or long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16},
-an up-link message sent to a base station is usually followed by a down-link message sent back by the base station, to indicate if the up-link message was successfully received and understood.
+In most IoT standards an up-link message sent to a base station is usually followed by a down-link message sent back by the base station, to indicate if the up-link message was successfully received and understood.
 %
-By using this feedback, that consist in an \emph{acknowledgement} message received shortly after every trial
+By using this feedback, that consists in an \emph{acknowledgement} (\Ack) message received shortly after every successful transmission, or in an absence of \Ack{} avery a failed transmission, it is possible that an IoT end-device can also use RL algorithms to optimize its communications.
+%
+To the best of our knowledge, this direction of research has never been studied before the year $2016$ and the beginning of this PhD.
 
-\newpage % WARNING
-
-% FIXME c'est bon tout ce qu'on a avant ?
+% \newpage % WARNING
 
@@
 \paragraph{Focusing on MAB algorithms.}
+% FIXME
+%
+We started by exploring the rich literature of multi-armed bandits,
+as many different algorithms exist, with lots of variants on the simple problem presented above (see \cite{LattimoreBanditAlgorithmsBook,Slivkins2019} for surveys).
+We thus start the first Part~\ref{part:Introduction} of this thesis with Chapter~\ref{chapter:2}, which presents the MAB model and gives a review of the most important algorithms designed to solve stochastic and stationary MAB problems.
+%
+In order to clearly understand which algorithms could be applied to the aforementioned problems of CR and IoT,
+we were not only interested by the usual measure of performance of any MAB algorithm (its regret, see below in Section~\ref{sec:2:lowerUpperBoundsRegret}),
+but also by their empirical performances in terms of computational complexity and storage requirements, both from the point of view of the theoretical analyses and real-world measurements on time and memory footprints of the different algorithms.
 
-FIXME
-We explored the rich literature of multi-armed bandits.
-Many algorithms exist.
-So we have to understand which ones can be applied to CR and IoT.
-
-Our exploration of the large number of MAB algorithms, and of MAB models, developed in the recent literature has produced a Python library of simulation of MAB problems.
+Our exploration of the large number of MAB algorithms and models developed in the recent literature
+has gave us the ambition to write a single piece of software were we could easily implement new models and algorithms, in order to compare the existing ones and empirically explores the performances of newly proposed algorithms.
+To answer this goal, we produced a Python library of simulation of MAB problems.
 %
@@
 Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement in a given IoT object.
-FIXME online algorithm selection and aggregation, with our contribution \Aggr{}.
+To answer this question, we present two approaches.
+The first one is an empirical comparison of the most efficient and well-known existing algorithms in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}, where it appears that simple algorithms such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal} are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
+The second approach is an online algorithm selection, consisting in aggregating a set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
 
@@
 %
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
@@
 Verify this first model empirically: proof of concept \cite{Besson2018ICT}
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
 Second model with decision making also for the retransmissions of packets \cite{Bonnefoi2019WCNC}
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
 Selfish
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
 Understand Selfish in the simpler framework of multi-player MAB
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
 In the OSA setting, we failed
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
 But we proposed new algorithms for this problem, and relaunched the research on this model \cite{Besson2018ALT}
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
@@
 We proposed new algorithms, some of which are proven to converge efficiently.
+
+\TODOL{Je dois faire ça d'ici mercredi matin}
 
@@
 % ----------------------------------------------------------------------------
-\section{Summary of the contributions}
+\paragraph{Summary of the contributions}
 \label{sec:1:summaryOfContributions}

### 2019-06-17 2-Chapters/2-Chapter/chapter2.tex
Probably enough for today? ALmost done for everything in Ch1 but the details of our contributions

@@
 
-\TODOL{Intégrer commentaires d'Emilie !}
+\TODOL{Intégrer les derniers commentaires d'Emilie, entre la fin des détails sur les familles exponentielles (avec les deux exemples), et quand je commence en Section~\ref{sub:2:otherMeasuresPerformance} à parler des autres mesures !}
 
@@
 %
-We study a similar model of using MAB for Cognitive Radio but without the OSA strucutre of Primary and Secondary users, \ie, without sensing information but only a collision indicator, in Chapters~\ref{chapter:4} and \ref{chapter:5}.
+We study a similar model of using MAB for Cognitive Radio but without the OSA structre of Primary and Secondary users, \ie, without sensing information but only a collision indicator, in Chapters~\ref{chapter:4} and \ref{chapter:5}.
 
@@
 % ----------------------------------------------------------------------------
-\section{Definition, decomposition and lower-bounds on the regret}
+% \section{Definition, decomposition and lower-bounds on the regret}
+\section{Measuring the performance of a MAB algorithm}
 % : definition, decomposition, and lower bounds
@@
 \subsection{Other measures of performances}
-
-In this thesis, we only study analytically the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
+\label{sub:2:otherMeasuresPerformance}
+
+The theoretical results obtained in this thesis are all expressed in terms of the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
 

### 2019-06-18 2-Chapters/1-Chapter/chapter1.tex
Ca a vachement bien avancé !
Almost done for this Ch1 ?

@@
 
-\TODOL{Lundi soir : je dois encore relire le 1.1 pour les fautes d'orthographe (et les éventuels erreurs historiques) !}
-\TODOL{Lundi soir : je dois encore terminer le 1.2 et être plus concis ?}
-\TODOL{Je dois écrire le 1.3 en max 2 pages : j'ai déjà le plan, je dois développer}
+\TODOL{Mardi : je dois encore relire le 1.1 pour les fautes d'orthographe (et les éventuels erreurs historiques) !}
+\TODOL{Mardi : je dois encore terminer le 1.2 et être plus concis ?}
+\TODOL{Je dois écrire la fin du 1.3 en max 2 pages : j'ai déjà le plan, je dois développer}
 
@@
 
-% ----------------------------------------------------------------------------
-\section{Context of this thesis}
-\label{sec:1:problems}
-
 % Where
@@
 
-\TODOL{Est-ce que je dois détailler plus ? Quel financement j'ai eu (CDSN), quels autres financements m'ont aidés ?}
+% \TODOL{Est-ce que je dois détailler plus ? Quel financement j'ai eu (CDSN), quels autres financements m'ont aidés ?}
 
-% Main problem
-\textbf{Main problems.}
+% ----------------------------------------------------------------------------
+\section{Context of this thesis}
+\label{sec:1:problems}
+
+% % Main problem
+% \textbf{Main problems.}
 %
@@
 
+\textbf{The spectrum scarcity issue.}
+%
+A major problem of current wireless technologies is the issue of spectrum scarcity:
+in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
+As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication) to space research, inter-satellite, mobile telephony and many other applications.
+%
+% Traditional regulatory structures have been built for an analog model and are not optimized for cognitive radio.
+Regulatory bodies in the world, like the \href{https://www.fcc.gov/}{Federal Communications Commission} in the USA (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
+or \href{http://www.tdf.fr/}{TDF} in France (see \href{https://www.TDF.fr/}{\texttt{TDF.fr}}, previously known as \emph{``TéléDiffusion de France''}), as well as different independent measurement campaigns, found that most radio frequency spectrum was inefficiently utilized,
+meaning that while a band can be allocated to a certain unique usage, it can be free from any user in certain times and/or places.
+We refer to \cite{valenta2010survey} for a survey on spectrum utilization in Europe.
+
+Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are insufficiently utilized.
+Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place.
+Moreover, the fixed spectrum allocation prevents rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
+Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands if they would not cause any interference to licensed users.
+These initiatives have initiated the research on \textbf{Cognitive Radio},
+and especially on one subfield that we detail below, \textbf{Dynamic Spectrum Access (DSA)}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
+    \caption{A chart representing the allocation of radio spectrum in the USA in $2016$.
+Only a small share of the spectrum is not yet allocated, the top-left corner in white (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz}).
+\copyright{} Wikimedia.}
+    % https://upload.wikimedia.org/wikipedia/commons/c/c7/United_States_Frequency_Allocations_Chart_2016_-_The_Radio_Spectrum.pdf?uselang=fr
+    \label{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}
+\end{figure}
+
+\TODOL{Je peux améliorer (ou enlever) ce passage qui site J.Toledano, ou alors le commenter en plus de détails ? J'aime bien son rapport en fait (j'aurai du le lire au début de ma thèse !)}
+% https://www.economie.gouv.fr/files/files/PDF/french-spectrum-mission-executive-summary-2014-06-25.pdf
+% https://www.economie.gouv.fr/files/files/PDF/rapport-gestion-dynamique-spectre-2014-06-30.pdf
+To highlight that the spectrum scarcity is not only an issue in the USA,
+we quote the abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by Joëlle Toledano \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull}, who was commissioned by the French government in $2014$.
+Different important problems are mentioned:
+%
+\begin{itemize}\tightlist
+    \item
+    \emph{``In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy.''}
+    And the usage of wireless communications is nowadays world-wide and omnipresent:
+    \emph{``Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.''}
+    Thus it is of highest interest to set up efficient, green, low-cost and durable new technologies for the future wireless standards, networks and devices.
+    %
+    \item
+    The radio spectrum is not only widely used for all kinds of applications, it is nowadays intensely used for some of them.
+    \emph{``Everyone agrees on the growing need for spectrum.
+    This growing need results from two phenomena.
+    On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
+    On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.''}
+    The first estimation was correct, and the mobile traffic is again \href{https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth}{estimated} to be increased by 10 times between 2019 and 2025.
+    % https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth
+    % https://www.cisco.com/c/en/us/solutions/collateral/service-provider/visual-networking-index-vni/white-paper-c11-738429.html
+    If the second estimation was overestimating the growth of the IoT market, forecasts predict  more than 64 billion IoT devices by 2025.
+    % https://www.businessinsider.com/internet-of-things-report?IR=T
+    \item
+    \emph{``Today, there are no more available frequencies in the easily exploitable frequency bands.
+    Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.''}
+    As illustrated above in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}.
+    \item
+    \emph{``Growing recourse to spectrum sharing, and in particular dynamic spectrum sharing, constitutes an important spectrum reserve.
+    A particular sharing form, the use of unlicensed bands, open to all and free, has been intensely developed with the growth of Wi-Fi use.''}
+    This thesis focusses on spectrum sharing in unlicensed bands, for IoT networks that lie at the opposite direction as the one underlying Wi-Fi networks: the spectrum sharing will be learned automatically by the end-devices, and not centrally decided by the base station.
+\end{itemize}
+
 \section{Cognitive Radio and Multi-Armed Bandits}
@@
 
+\TODOL{Emilie pense que je devrais faire ça mieux :
+    Quant au survey du machine learning (Figure 1.1), je ne pense pas que
+    cela soit nécessaire. Je trouve ça plus pertinent d'introduire le
+    problème de décision séquentielle qui nous intéresse (qui n'est pas
+    juste un sub-field du machine learning d'ailleurs, la communauté
+    statistique s'y intéresse depuis bien plus longtemps !), sur l'exemple
+    de l'OSA.
+}
+
 \paragraph{From Machine Learning to Multi-Armed Bandits.}
@@
 As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
-The reference book on RL is \cite{SuttonBarto2018}, so let us quote the definition of RL given by Richard Sutton and Andrew Barto:
+The reference book on RL is \cite{SuttonBarto2018}, so let us quote the definition of RL given by R. Sutton and A. Barto:
 \emph{``Reinforcement learning is learning what to do -— how to map situations to actions -— so
@@
 The goal of the player is to maximize its reward, by trials and errors (\ie,actions and rewards).
-Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial).
+Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial), etc.
 
@@
 and exploitation of the best action according to its current knowledge.
+%
+The MAB problem have been studied in both the machine learning and the statistics community, since the $1950$s with pioneers like H.
+Robbins \cite{Robbins52}, and more recently it is an active field of research, since the late $1990$s \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
+Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}.
 
@@
 \end{itemize}
-
-\TODOL{Je peux améliorer (ou enlever) ce passage qui site J.Toledano, ou alors le commenter en plus de détails ? J'aime bien son rapport en fait (j'aurai du le lire au début de ma thèse !)}
-% https://www.economie.gouv.fr/files/files/PDF/french-spectrum-mission-executive-summary-2014-06-25.pdf
-% https://www.economie.gouv.fr/files/files/PDF/rapport-gestion-dynamique-spectre-2014-06-30.pdf
-The abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by Joëlle Toledano, commissioned and published by the French government \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull},
-mentions different important problems:
-%
-\begin{itemize}\tightlist
-    \item
-    \emph{``In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy''}
-    \item
-    \emph{``Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.''}
-    %
-    \item
-    \emph{``Everyone agrees on the growing need for spectrum.
-    This growing need results from two phenomena.
-    On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
-    On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.''}
-    \item
-    \emph{``Today, there are no more available frequencies in the easily exploitable frequency bands.
-    Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.''}
-    \item
-    \emph{``Growing recourse to spectrum sharing, and in particular dynamic spectrum sharing, constitutes an important spectrum reserve.
-    A particular sharing form, the use of unlicensed bands, open to all and free, has been intensely developed with the growth of Wi-Fi use.''}
-\end{itemize}
-
-\textbf{The spectrum scarcity issue.}
-%
-A major problem of current wireless technologies is the issue of spectrum scarcity:
-in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
-As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication) to space research, inter-satellite, mobile telephony and many other applications.
-%
-% Traditional regulatory structures have been built for an analog model and are not optimized for cognitive radio.
-Regulatory bodies in the world, like the \href{https://www.fcc.gov/}{Federal Communications Commission} in the USA (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
-or \href{http://www.tdf.fr/}{TDF} in France (see \href{https://www.TDF.fr/}{\texttt{TDF.fr}}, previously known as \emph{``TéléDiffusion de France''}), as well as different independent measurement campaigns, found that most radio frequency spectrum was inefficiently utilized,
-meaning that while a band can be allocated to a certain unique usage, it can be free from any user in certain times and/or places.
-We refer to \cite{valenta2010survey} for a survey on spectrum utilization in Europe.
-
-Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are insufficiently utilized.
-Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place.
-Moreover, the fixed spectrum allocation prevents rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
-Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands if they would not cause any interference to licensed users.
-These initiatives have focused CR research on \textbf{Dynamic Spectrum Access (DSA)}.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
-    \caption{A chart representing the allocation of radio spectrum in the USA in $2016$.
-Only a small share of the spectrum is not yet allocated, the top-left corner in white (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz}).
-\copyright{} Wikimedia.}
-    % https://upload.wikimedia.org/wikipedia/commons/c/c7/United_States_Frequency_Allocations_Chart_2016_-_The_Radio_Spectrum.pdf?uselang=fr
-    \label{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}
-\end{figure}
 
@@
 %
-This model was first studied by Wassim Jouini during his PhD thesis with Christophe Moy, ten years ago, first with \cite{Jouini09} and later with \cite{Jouini10,Jouini12}.
+This model was first studied by W. Jouini during his PhD thesis with C. Moy, ten years ago, first with \cite{Jouini09} and later with \cite{Jouini10,Jouini12}.
+
 His works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the \UCB{} algorithm,
-along the early works of Qing Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
+along the early works of Q. Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
 %
-Shortly after in $2014$, proof-of-concepts using real-world radio hardware and Software Defined Radio were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
-
-In a second PhD thesis supervised by Christophe Moy \cite{Modi17PhD} from $2014$ to $2017$, Navikkumar Modi studied the impact of using MAB algorithms to optimize channel selection on the battery life of a wireless device.
+Shortly after in $2014$, proof-of-concepts using real-world radio hardware and Software Defined Radio were developed by C. Moy and his student C. Robert \cite{RobertSDR2014,MoyWSR2014}.
+In a second PhD thesis supervised by C.
+Moy \cite{Modi17PhD} from $2014$ to $2017$, N.
+Modi studied the impact of using MAB algorithms to optimize channel selection on the battery life of a wireless device.
 On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
@@
 \cite{modiDemo2016}.
-%
-In $2017$, Christophe Moy continued to work on this direction, with a post-doctoral student, Sumit J. Darak, leading to the following publications
+
+In $2017$, C. Moy continued to work on this direction, with a post-doctoral student, S. Darak, leading to the following publications
 \cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
@@
 %
-Since $2017$, Sumit J.
+Since $2017$, S.
 Darak and his team at IIIT Delhi have been very active in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
@@
 
-\TODOL{Raccourcir toutes les fois où je parle des travaux sur l'OSA après dans la suite de la thèse !}
+% \TODOL{Raccourcir toutes les fois où je parle des travaux sur l'OSA après dans la suite de la thèse !}
 
@@
     most IoT devices are very low-cost devices, with low material and computational capacities,
-    meaning that they are not equipped with any sensing capabilities,
+    meaning that they are not equipped with any sensing capabilities.
+    However, they are equipped with a transceiver as well as with a receiver,
+    and with low-complexity storage and computation capacities,
     \item
@@
     IoT devices have low duty cycles (only a few messages every minute) to very low duty cycles (only a few messages a day).
-    % FIXME more specificities?
 \end{itemize}
+\TODOL{FIXME more specificities? Est-ce que je dois mentionner d'autres contraintes ? Il me faudrait aussi une ou deux références genre survey des standards IoT}
 
@@
 If we sum-up the studied problems and summarize them in one question, it would be the following:
-\emph{``Can we adapt the decision making tools already successfully applied to Cognitve Radio for Opportunistic Spectrum Access to the specific needs of CR for the future IoT networks ?''}
+\emph{``Can we adapt the decision making tools already successfully applied to Cognitve Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
 %
@@
 
-\paragraph{Focusing on MAB algorithms.}
-% FIXME
+\subsection{Focusing on MAB algorithms}
+
+\paragraph{Chapter~\ref{chapter:2}.}
 %
@@
 
+\paragraph{Chapter~\ref{chapter:3}.}
+%
 Our exploration of the large number of MAB algorithms and models developed in the recent literature
@@
 
+\paragraph{Chapter~\ref{chapter:25}.}
+%
 Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement in a given IoT object.
@@
 
-\paragraph{Focusing on IoT networks.}
+\subsection{Focusing on IoT networks.}
+
+In the second Part~\ref{part:MABIOT} of this thesis, we start by proposing and studying different models of IoT networks, with simulations and a real-world proof-of-concept, and then we study interesting questions on two mathematical models of MAB that arise from our first contribution.
+
+\paragraph{Chapter~\ref{chapter:4}.}
 %
+In Chapter~\ref{chapter:4}, we study two models of IoT networks.
+The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use the acknowledgement as feedback to learn the best channel for their uplink communications.
+The goal of this first model of applying decentralized RL on the device side is to increase the Quality of Service (QoS) of the application of such IoT network, by reducing the rate of failed transmission.
+The base station will thus receive more data from the devices its serves, if they can successfully learn an efficient spectrum access scheme.
+%
+On the other hand, the second model we study considers packet retransmissions, and while the concept stays similar, the consequence of an efficient decentralized learning will be both an increase battery life for the devices as well as an increase QoS.
+
+We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of up-to thousands of end-devices and a single base station.
+We consider a time-frequency slotted ALOHA-based protocol, by assuming a perfect synchronization between all devices and the base station.
+%
+Our IoT network model contains two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
+Static devices form an interfering traffic, which could be generated by devices using other standards as well.
+Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choices of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) from the base station.
+%
+We first detail three base-lines, a naive approach and two centralized ``oracle'' approaches,
+that are used to evaluate the performance of MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communication rate.
+We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes more and more unpredictable.
+%
+Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
+This model constitutes the first article written during this thesis, \cite{Bonnefoi17}, with R. Bonnefoi, another PhD student of our team SCEE.
+
+Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations is a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
+In Section~\ref{sec:4:firstModel}, we propose the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network \cite{Besson2018ICT}.
+We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements, that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
+% Verify this first model empirically: proof of concept
+
+We then consider extension of our first model, following the idea of packet retransmission that is at the root of the ALOHA protocol \cite{Abramson1970,Roberts75}.
+This second model proposes to also use decision making at the second layer of the MAC protocol, that is to also use RL and MAB algorithms, in order to learn the best channel to use for retransmitting any packet that suffered from a collision at its first transmission.
+We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations \cite{Bonnefoi2019WCNC}.
+This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well as the simplest of the different heuristics we compared is efficient and allow the devices to learn to behave almost optimally.
+
 \TODOL{Je dois faire ça d'ici mercredi matin}
 
-First model \cite{Bonnefoi17}, CROWNCOM paper, model to detail
-
-We already explained in Chapter~\ref{chapter:1} that
-unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
-This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.
-
-Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
-As end-devices battery life is a key constraint of IoT networks,
-and as IoT networks are decentralized (\ie, the devices initiate transmissions),
-this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
+\paragraph{Chapter~\ref{chapter:5}.}
 %
-In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
-Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
-We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
-Kaufmann12Thompson}, for the first model.
-For the demonstration as well as for the second model, without loss of generality, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.
-
-The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
-and the network usage is more and more fluctuating.
-% and the stochastic hypothesis is more and more violated.
-To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
-Static devices form an interfering traffic, which could be generated by devices using other standards as well.
-Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel.
-So our first hypothesis is actually not constraining.
-We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
-This second approach is ideal, but not realistic the most common situation of decentralized co-located IoT networks, and it is just used here as a reference.
-Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
-We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.
-
-Verify this first model empirically: proof of concept \cite{Besson2018ICT}
-\TODOL{Je dois faire ça d'ici mercredi matin}
-
-Second model with decision making also for the retransmissions of packets \cite{Bonnefoi2019WCNC}
-\TODOL{Je dois faire ça d'ici mercredi matin}
-
+In the two models presented before, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm,
+each of them targeting their local objective of maximizing their cumulated reward (in a selfish way),
+in order to optimize the complete system.
+It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the prisoner dilemma.
 Selfish
@@
 
+\TODOL{Emilie me dit de continuer comme ça :
+    En tout cas ce chapitre 1 commence à avoir bonne forme pour l'école
+    doctorale, il n'y aura plus qu'à finaliser la deuxième partie de la
+    section 1.3 (la première est très bien). Tu tiens le bon bout !
+}
+
+\paragraph{Chapter~\ref{chapter:6}.}
+%
+FIXME explain non stationarity !
+
 \paragraph{Our approach.}
@@
 We proposed new algorithms, some of which are proven to converge efficiently.
-
-\TODOL{Je dois faire ça d'ici mercredi matin}
-
-% %
-% directions de recherches
-
-% - réfléchir aux modèles, quelles hypothèses sont retirées comparé au modèle OSA
-
-% - analyser mathématiquement les modèles
-
-% - proposer des solutions : algorithmes, dont la convergence est justifiée, ou heuristiques
-
-% - des simulations logicielles
-
-% - des démonstrations matérielles
-
-% - et en bonus on obtient aussi de belles contributions théoriques
-
-% FIXME c'est bon tout ce qu'on a avant ?
 
@@
     \item
-    We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
-    We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
-    A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.
+    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, published under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+    We present it in details in Chapter~\ref{chapter:3}.
 

### 2019-06-18 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Almost done for this Ch1 ?

@@
 % \TODOL{Est-ce que je fais ça ici ou à la fin du chapitre 5 ? Ou peut-être en chapitre 1, ça ferait du contenu au chapitre 1...}
+% FIXED j'en parle au chapitre 1
 %
-We evoke here some previous works that similarly implemented proof-of-concepts to show that reinforcement learning algorithms can be used within real-world wireless communications.
-Starting from $2010$, the works of Wassim Jouini \cite{Jouini09,Jouini10,Jouini12} were the first ones to propose to use Reinforcement Learning for Cognitive Radio, especially MAB and the \UCB{} algorithm, and from $2014$, proof-of-concepts were developed by Christophe Moy and his student Clément Robert \cite{RobertSDR2014,MoyWSR2014}.
-In $2015$ and $2016$, Christophe Moy continued to work on this direction, with a PhD student (Navikkumar Modi) and post-doc (Sumit J. Darak) in our team SCEE, for the following publications
-\cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
-Since $2017$, Sumit J.
-Darak and his team at IIIT Delhi have been very active in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+This work is new for the IoT context, but previous works have similarly implemented proof-of-concepts to show that Reinforcement Learning (RL) algorithms can be used within real-world wireless communications.
+Starting from $2010$, the works of W.
+Jouini and C.
+Moy \cite{Jouini09,Jouini10,Jouini12} were among the first ones to propose to use RL for Cognitive Radio, especially MAB and the \UCB{} algorithm, and proof-of-concepts were developed with C.
+Robert from $2013$ \cite{RobertSDR2014,MoyWSR2014}.
+Between $2015$ and $2017$, C. Moy, N. Modi and S. Darak (of our team SCEE) continued to work on this direction \cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
+Since $2017$, S.
+Darak and his team at IIIT Delhi (India) have been very active in the research on CR using MAB, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.

### 2019-06-18 2-Chapters/5-Chapter/chapter5.tex
Almost done for this Ch1 ?

@@
 % ----------------------------------------------------------------------------
-\subsection{Inspirations from Cognitive Radio Proof-of-Concepts}
+% \subsection{Inspirations from Cognitive Radio Proof-of-Concepts}
+\subsection{An idea from a Cognitive Radio Proof-of-Concept: also use learning for the rank of \RhoRand}
 \label{sub:5:USRPdemos}
 
-Similarly to our demonstration of multi-armed bandit learning in an IoT network that we presented in the previous Chapter~\ref{chapter:4} in Section~\ref{sec:4:gnuradio},
+Similarly to our demonstration of MAB learning in an IoT network that we presented in the previous Chapter~\ref{chapter:4} in Section~\ref{sec:4:gnuradio},
 we list here some related works, dating back from $2016$, who proposed a similar approach for OSA.
@@
 %
-Indeed, the demonstrations and the works presented below are coming from our team experiments, and have been extended by the team of Sumit Darak at IIIT Delhi after his postdoctoral fellowship in our team.
+Indeed, the demonstrations and the works presented below are coming from our team experiments, and have been extended by the team of S.
+Darak at IIIT Delhi after his postdoctoral fellowship in our SCEE team.
+%
+In \cite{darak2016bayesian,Darak16}, the authors study the same model and focused on the \RhoRand{} policy, which was the state-of-the-art back in $2016$, combined with \UCB{} or \klUCB, as well as an extension using Bayes-UCB from \cite{Kaufmann12BUCB}.
 
-In \cite{darak2016bayesian,Darak16}, the authors study the same model and focused on the \RhoRand{} policy, which was the state-of-the-art back in $2016$, combined with \UCB{} or \klUCB, as well as an extension using Bayes-UCB from \cite{Kaufmann12BUCB}.
-Their idea is to use the same skeleton as \RhoRand, that is to assign a \emph{rank} to each player, and make players change their rank after any collision.
-But instead of selecting a new rank uniformly at random among $[M]$, they proposed to use a ``second-stage'' learning policy\footnote{This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.}, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
-This algorithm is implemented in SMPyBandits and named \rhoLearn,
+Their idea is to use the same skeleton as \RhoRand{} \cite{darak2016bayesian}, that is to assign a \emph{rank} to each player, and make players change their rank after any collision.
+But instead of selecting a new rank uniformly at random among $[M]$, they proposed to use a ``second-stage'' learning policy, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
+This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.
+This algorithm is named \rhoLearn, and it is implemented in SMPyBandits,
 % \footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.},
-where it can use Bayes-UCB or any of the $65$ or more bandit algorithms available in the library.
+where it can use Bayes-UCB, or any of the $65$ or more bandit algorithms available in the library.
+%
 Even if no theoretical guarantee backs up this idea of second-stage learning with Bayes-UCB, empirical simulations found that it can be efficient.

### 2019-06-18 2-Chapters/1-Chapter/chapter1.tex
Lots of details on Part I of the thsis

@@
 Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
-This model constitutes the first article written during this thesis, \cite{Bonnefoi17}, with R. Bonnefoi, another PhD student of our team SCEE.
+This model constitutes the first article written during this thesis with R. Bonnefoi, another PhD student of our team SCEE.
+We presented this first work, which obtained the ``best paper award'', at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$ \cite{Bonnefoi17}.
 
 Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations is a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
-In Section~\ref{sec:4:firstModel}, we propose the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network \cite{Besson2018ICT}.
+In Section~\ref{sec:4:firstModel}, we propose the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network.
 We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements, that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
 % Verify this first model empirically: proof of concept
+This work was again implemented with R. Bonnefoi, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}.
 
-We then consider extension of our first model, following the idea of packet retransmission that is at the root of the ALOHA protocol \cite{Abramson1970,Roberts75}.
+We then consider an extension of our first model, following the idea of packet retransmission that is at the root of the ALOHA protocol \cite{Abramson1970,Roberts75}.
 This second model proposes to also use decision making at the second layer of the MAC protocol, that is to also use RL and MAB algorithms, in order to learn the best channel to use for retransmitting any packet that suffered from a collision at its first transmission.
-We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations \cite{Bonnefoi2019WCNC}.
+We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations.
 This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well as the simplest of the different heuristics we compared is efficient and allow the devices to learn to behave almost optimally.
-
-\TODOL{Je dois faire ça d'ici mercredi matin}
+This third contribution concluded our collaboration with R.
+Bonnefoi, and was presented at the MoTION workshop during the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Bonnefoi2019WCNC}.
 
@@
 %
-In the two models presented before, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm,
-each of them targeting their local objective of maximizing their cumulated reward (in a selfish way),
-in order to optimize the complete system.
+In the two models presented above, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm,
+each of them targeting their local objective of maximizing their cumulated reward (in a selfish way), in order to optimize the complete system.
 It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the prisoner dilemma.
+So it is quite surprising that our numerical simulations as well as the realistic PoC showed that decentralized selfish MAB learning lead to efficient
+
 Selfish
@@
 
-\paragraph{Our approach.}
-%
-Our approach for the aforementioned problems is summarized as follows:
-we propose a couple of new models of decision making for IoT networks, on the device's side, for the spectrum access problem.
-By considering the difference between the hypothesis of previously studied models for the OSA problem and our problem at end for IoT, we identify the constraints and specificities of applying decentralized ML for IoT devices.
-%
-Like for the OSA, we can apply successfully model the decision making problem as a Multi-Armed Bandit problem.
-We showed on numerical simulations as well as on a real-world proof-of-concept that applying low-cost sequential learning algorithms coming from the MAB literature successfully reduce the collision rate of IoT end-devices.
-We proposed new algorithms, some of which are proven to converge efficiently.
+% \paragraph{Our approach.}
+% %
+% Our approach for the aforementioned problems is summarized as follows:
+% we propose a couple of new models of decision making for IoT networks, on the device's side, for the spectrum access problem.
+% By considering the difference between the hypothesis of previously studied models for the OSA problem and our problem at end for IoT, we identify the constraints and specificities of applying decentralized ML for IoT devices.
+% %
+% Like for the OSA, we can apply successfully model the decision making problem as a Multi-Armed Bandit problem.
+% We showed on numerical simulations as well as on a real-world proof-of-concept that applying low-cost sequential learning algorithms coming from the MAB literature successfully reduce the collision rate of IoT end-devices.
+% We proposed new algorithms, some of which are proven to converge efficiently.
 
@@
 % WARNING
-\newpage
+% \newpage
 

### 2019-06-18 2-Chapters/1-Chapter/chapter1.tex
Almost okay for summary/history of Ch4, Ch5, Ch6

@@
 This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices, it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
-In the present and future IoT networks, many devices of heterogenous nature are using the same antenna for many different applications.
+In the present and future IoT networks, many devices of heterogeneous nature are using the same antenna for different applications.
 A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and will use a battery, which duration should be maximized (typically more than 10 years).
@@
 in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
-As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication) to space research, inter-satellite, mobile telephony and many other applications.
+As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication) to space research, inter-satellite, mobile telephony and a large number of other applications.
 %
@@
 The reference book on RL is \cite{SuttonBarto2018}, so let us quote the definition of RL given by R. Sutton and A. Barto:
-\emph{``Reinforcement learning is learning what to do -— how to map situations to actions -— so
+\emph{``Reinforcement learning is learning what to do - how to map situations to actions - so
 as to maximize a numerical reward signal. The learner is not told which actions to
@@
 % https://en.wikipedia.org/wiki/Software-defined_radio
-A SDR is a radio communication system where components that have been traditionally implemented in hardware (e.g.
-mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are instead implemented by means of software on a personal computer or embedded system, with the exception of a unique pair of receiver/transmiter antennas.
+A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are instead implemented by means of software on a personal computer or embedded system, with the exception of a unique pair of receiver/transmitter antennas.
 Even though the SDR paradigm was initiated by the USA defense research, over the past years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
@@
 The specificities of IoT networks can be listed as follows.
-\begin{itemize}
+\begin{itemize}\tightlist
     \item
@@
 The second approach is an online algorithm selection, consisting in aggregating a set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
+This work on aggregating MAB algorithms was motivated for the OSA case of Cognitive Radio, where most previous research works only considered the $\UCB_1$ algorithm without justifying this choice rigorously,
+and was presented at the IEEE WCNC conference in Barcelona, Spain, in April $2018$.
 
@@
 % Verify this first model empirically: proof of concept
-This work was again implemented with R. Bonnefoi, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}.
+This is again the result of a fruitful collaboration with R.
+Bonnefoi, and this work was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
 
@@
 This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well as the simplest of the different heuristics we compared is efficient and allow the devices to learn to behave almost optimally.
-This third contribution concluded our collaboration with R.
-Bonnefoi, and was presented at the MoTION workshop during the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Bonnefoi2019WCNC}.
+This third contribution concluded our collaboration with R. Bonnefoi, and was presented at the $1^{\text{st}}$ MoTION workshop, during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
 
@@
 It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the prisoner dilemma.
-So it is quite surprising that our numerical simulations as well as the realistic PoC showed that decentralized selfish MAB learning lead to efficient
+So it is quite surprising that our numerical simulations as well as the realistic PoC showed that decentralized selfish MAB learning lead to efficient coordination between devices.
 
-Selfish
-\TODOL{Je dois faire ça d'ici mercredi matin}
+We first tried to analyze the performance of this decentralized algorithm, that we refer to as Selfish, in the model of Section~\ref{sec:4:firstModel},
+but due to the random number of active devices at each time step (as soon as the probability $p$ of activation is $p < 1$), we were unable to propose a clean analysis.
+%
+That is why in Chapter~\ref{chapter:5}, we relax the hypothesis of having $M \gg K$ devices in a network with $K$ orthogonal wireless channel, equivalent to having $p < 1$, and we consider the case of devices that each communicate at every step (\ie, $p=1$), and so only up-to $M \leq K$ devices.
+We present the studied model in Section~\ref{sec:5:model}, that comes with different variants depending on the feedback level, and cover both the OSA case (\ie, with sensing information) or the IoT case (\ie, without sensing feedback).
+The goal was to understand the heuristic of Chapter~\ref{chapter:4}, Selfish, in the simpler framework of multi-player MAB, which has been studied before \cite{Zhao10,Anandkumar10,Anandkumar11} for the case of OSA.
 
-Understand Selfish in the simpler framework of multi-player MAB
-\TODOL{Je dois faire ça d'ici mercredi matin}
+On the first hand, in the OSA setting, we failed to obtain positive result for the Selfish policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability.
+This result was later confirmed and generalized by \cite{BoursierPerchet18}.
+%
+On the other hand, we were able to propose new algorithms for this multi-player bandit problem with sensing information, and we analyzed our proposal \MCTopM-\klUCB, to show that it achieves a finite-time logarithmic regret upper bound, improving over previous state-of-the-art results.
+%
+Unfortunately this strong theoretical result heavily depends on the presence of sensing feedback, and as such our results are not (yet) applicable to the IoT model without sensing.
 
-In the OSA setting, we failed
-\TODOL{Je dois faire ça d'ici mercredi matin}
-
-But we proposed new algorithms for this problem, and relaunched the research on this model \cite{Besson2018ALT}
-\TODOL{Je dois faire ça d'ici mercredi matin}
+Our work on multi-player bandit was presented at the ALT conference in Lanzarote, Spain, in April $2019$ \cite{Besson2018ALT},
+%
+%
+and it relaunched the research on this model as a few independent teams started to work on different questions after its publication.
+%
+Since last year, it was proven that the ``no sensing'' case (\ie, IoT) is essentially not harder that the ``sensing'' (\ie, OSA) case, from a theoretical point-of-view, by E.
+Boursier and V.
+Perchet \cite{BoursierPerchet18}.
+Even if their approach is costly empirically, the result is surprising and very interesting.
+We discuss it, along with other recent research works such as \cite{LugosiMehrabian18}, in Section~\ref{sec:5:literatureReviewOtherModels}.
+%
+Moreover, the case of different means of arms for the different players was studied in \cite{KaufmannAbbas19}, and this generalization of our model is very interested for CR applications, as the mean of an arm is the average quality (\eg, availability) of a radio channel, and as such it has no reason to be uniform among end-devices (\ie, players).
+%
+Finally the adversarial case of multi-player bandit has started to be studied, independently by \cite{bande2019adversarial} and \cite{AlaturLevyKrause19}.
 
@@
 
+As explained above, the model of Chapter~\ref{chapter:4} was found intractable to analyze because we consider an IoT network with many end-devices, all following random activation patterns.
+Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, but leads to the model of Chapter~\ref{chapter:5}.
+%
+On the first hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let the devices learn an optimal orthogonal affectation of $M \leq K$ devices in the $K$ channels.
+However, it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid any central control of the devices by the base station, as we explained above.
+%
+On the other hand, FIXME explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
+
 \paragraph{Chapter~\ref{chapter:6}.}
 %
-FIXME explain non stationarity !
+% FIXME explain non stationarity !
+%
+In order to also understand precisely how MAB algorithms behave under non-stationary environment, we studied the literature on adversarial and on non-stationary MAB, for both cases of slowly-varying and abruptly-changing environments.
+We begin our last Chapter~\ref{chapter:6} by reviewing the existing works,
+and we chose to focus on the piece-wise stationary problem.
+Two main families of algorithms have been proposed for this problem,
+that usually consist in combining an efficient policy designed for the stationary MAB problem (\eg, \UCB{} or \klUCB) and a way to adapt to changes in the arms distributions.
+Passively adaptive policies use a window of a fixed or evolving size, or a discount factor to slowly forget about the past observations,
+while actively adaptive policies use a statistic test to detect the change-point (\ie, the time steps between two consecutive stationary intervals).
+As the recent literature showed that the later approach is usually more competitive, and usually obtains better results from both empirical and theoretical aspects, we chose to develop our own actively adaptive algorithm.
+
+Following two recent previous works \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, we combine an efficient index policy (\klUCB) and an efficient change-point detection test, under the assumption of bounded rewards.
+Building on very recent results on the Generalized Likelihood Ratio test (GLRT) for Gaussian and sub-Gaussian variables \cite{Maillard2018GLR}, we instead focus on bounded rewards and Bernoulli distributions.
+Using the fact that bounded variables in $[0,1]$ are not only $1/4$ sub-Gaussian but also sub-Bernoulli, we prove the first finite-time guarantees for the GLRT for bounded variables.
+We first show finite-time bounds on both the false alarm probability of our test, and its detection delay, under mild assumption on the lengths of the stationary sequences.
+Our algorithm, denoted \GLRklUCB, is then presented in two variants, whether change-points are local (\ie, only one arm mean changes at each change-point) or global (\ie, possibly all the arm means change at a time).
+%
+We prove that by combining the asymptotically optimal policy \klUCB, designed for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of our proposed algorithm.
+%
+The performance of \GLRklUCB{} is also illustrated on numerical experiments on synthetic data, where it is shown to outperforms all the passively adaptive policies as well as the previous actively adaptive policies.
+%
+This last chapter is based on our last work, since October $2018$, and it first lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
+Our work also lead to the article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results and submitted soon to a journal (probably the Journal of Machine Learning Research) in fall $2019$.
 

### 2019-06-18 2-Chapters/1-Chapter/chapter1.tex
J'ai hâte de finir ! Pas le temps d'aller voir Xtof désoooolé

@@
 
-\TODOL{Mardi : je dois encore relire le 1.1 pour les fautes d'orthographe (et les éventuels erreurs historiques) !}
-\TODOL{Mardi : je dois encore terminer le 1.2 et être plus concis ?}
-\TODOL{Je dois écrire la fin du 1.3 en max 2 pages : j'ai déjà le plan, je dois développer}
+\TODOL{Mardi soir : je dois encore relire le 1.1 pour les fautes d'orthographe (et les éventuels erreurs historiques) !}
+\TODOL{Mardi soir : je dois encore améliorer le 1.2 et être plus concis ?}
+\TODOL{Mardi soir : je dois relire la fin du 1.3, et terminer le dernier point}
 
@@
 %
-On the other hand, FIXME explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
+On the other hand,
+FIXME explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
+\TODOL{Mardi soir je dois encore expliquer ça !}
 
@@
 As the recent literature showed that the later approach is usually more competitive, and usually obtains better results from both empirical and theoretical aspects, we chose to develop our own actively adaptive algorithm.
+
+\TODOL{Mardi soir je peux mieux expliquer, là je suis long et chiant je trouve}
 
@@
 
-\TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
+% \TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
 
-\newpage % WARNING
+% \newpage % WARNING
 
@@
     \item
-    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
+    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non-stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
     Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
@@
 \end{itemize}
+
+\newpage  % WARNING
 
@@
     \cite{MoyBesson2019}
+    Note that we are already working on an extended version that will be submitted to a journal in July $2019$.
 
@@
     \cite{Besson2019GLRT}
+    Note that we are already working on an extended version that will be submitted to a journal in autumn $2019$.
 

### 2019-06-18 2-Chapters/1-Chapter/chapter1.tex
Typos

@@
 
-\abstractStartChapter{}%
-%
-We start by presenting the context then the problems studied in this thesis.
-We further announce the different directions we considered and the solutions we proposed.
-We give an overview of our main contributions, alongside a list of publications written in the last three years.
-%
-The organization of this manuscript is detailed in a flow graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but they are all built upon the model and notations introduced on Chapter~\ref{chapter:2}.
+% \abstractStartChapter{}%
+% %
+% We start by presenting the context then the problems studied in this thesis.
+% We further announce the different directions we considered and the solutions we proposed.
+% We give an overview of our main contributions, alongside a list of publications written in the last three years.
+% %
+% The organization of this manuscript is detailed in a flow graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but they are all built upon the model and notations introduced on Chapter~\ref{chapter:2}.
 
-\TODOL{J'enlève l'abstract et la minitoc dès que le chapitre est terminé !}
+% \TODOL{J'enlève l'abstract et la minitoc dès que le chapitre est terminé !}
 
-\minitocStartChapter{}
+% \minitocStartChapter{}
 
@@
 
-% \TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
+\TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
 
-% \newpage % WARNING
+\newpage % WARNING
 
@@
 Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
-Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where MAB algorithms have been used with success.
+Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where we show that MAB algorithms can be used with success.
 Our two models are interesting and close to reality, but they appeared to be too general to propose a mathematical analysis of the good empirical performance of the considered solutions.
@@
 \href{https://cv.archives-ouvertes.fr/lilian-besson/}{\texttt{CV.Archives-Ouvertes.fr/lilian-besson}}.
+%
+The following publications are ordered historically, from the most recent to the oldest.
 
@@
     \cite{MoyBesson2019}
-    Note that we are already working on an extended version that will be submitted to a journal in July $2019$.
+
+    Note that we are already working on an extended version that will be submitted to a journal on Machine Learning for Wireless Communications, in July $2019$.
 
@@
     \cite{Besson2019GLRT}
-    Note that we are already working on an extended version that will be submitted to a journal in autumn $2019$.
+
+    Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.
 

### 2019-06-18 2-Chapters/2-Chapter/chapter2.tex
Typos

@@
     Using bandit algorithms for \textbf{improved machine learning} models or algorithms has been an active research domain for the last ten years or so.
-    As presented below in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}, a certain ``leader'' bandit algorithm can be used to select on the run the best bandit algorithm from a pool of ``followers'' algorithms.
+    As presented below in Chapter~\ref{chapter:25}, a certain ``leader'' bandit algorithm can be used to select on the run the best bandit algorithm from a pool of ``followers'' algorithms.
     Other possible use cases include hyper-parameter optimization, or features selection.
@@
 
-Here again, if a lower-bound $d>0$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret, as stated in Lemma~\ref{lem:2:ExploreThenExploit} below.
+Here again, if a lower-bound $d>0$ on the positive gaps is known beforehand, one can find a tuning of $T_0$ that gives a logarithmic regret, as stated in Theorem~\ref{thm:2:ExploreThenExploit} below.
 It proves that the ``explore-then-exploit'' strategy can also obtain an order-optimal regret, $R_T \leq K (4/d^2) (1 + \log(d^2 T / 4)) = \bigO{K \log(T)}$, if it is tuned with a fixed time $T_0$ using prior knowledge on the problem (\ie, $d$) and the horizon $T$.

### 2019-06-18 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
\allowdisplaybreaks{} command from https://tex.stackexchange.com/a/12561/97964
Fixed typos
Typos

@@
 It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
-The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} \cite{Besson2018WCNC}, modified to efficiently aggregate a growing number of experts from techniques presented in \cite{Mourtada17}.
+The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Chapter~\ref{chapter:25} \cite{Besson2018WCNC}, modified to efficiently aggregate a growing number of experts from techniques presented in \cite{Mourtada17}.
 At each time step, a new expert is introduced, and experts correspond to different Thompson sampling algorithms, each using a different history of pulls and rewards.
@@
     %
-    Denote $\widehat{\mu}_s \denote \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r \denote \frac{1}{r}\sum\limits_{i=1}^r Y_k$.
+    Denote $\widehat{\mu}_s \eqdef \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r \eqdef \frac{1}{r}\sum\limits_{i=1}^r Y_k$.
     Then for every $s,r \in \N^*$ we have,
@@
 \[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
-which concludes the proof.
+which concludes the proof.m
 %

### 2019-06-18 5-Nomenclature/Nomenclature.tex
done
Typos

@@
 \nomenclature[G]{$\alpha_0$, $\delta_0$}{Usually denotes a constant scaling of a parameter of an algorithm, for instance $\varepsilon_t = \varepsilon_0 / t$ is used in Chapter~\ref{chapter:2}, or $\alpha = \alpha_0 \alpha_T$ is used in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\pi$}{It usually denotes the number $\pi\simeq 3.14\ldots$, but it also denotes a probability distribution in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} or a permutation in Section~\ref{par:5:twoIdeasOrthogonalization}}
+\nomenclature[G]{$\pi$}{It usually denotes the number $\pi\simeq 3.14\ldots$, but it also denotes a probability distribution in Chapter~\ref{chapter:25} or a permutation in Section~\ref{par:5:twoIdeasOrthogonalization}}
 % \nomenclature[G]{$\rho$, $\rho^k$}{A bandit strategy, denoting the mathematical object which maps past observations to future decisions (mainly in Chapter~\ref{chapter:5})}
@@
 \nomenclature[L]{$\cA$}{An algorithm, also referred to as a policy or a strategy.
-$\cA_1,\dots,\cA_N$ denote the $N$ aggregated algorithms in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} and $\cA_1,\dots,\cA_M$ denotes the algorithms of the $M$ players in Chapter~\ref{chapter:5}}
+$\cA_1,\dots,\cA_N$ denote the $N$ aggregated algorithms in Chapter~\ref{chapter:25} and $\cA_1,\dots,\cA_M$ denotes the algorithms of the $M$ players in Chapter~\ref{chapter:5}}
 \nomenclature[L]{$A(t)$, $A^j(t)$}{Decision of algorithm $\cA$ at time $t\in[T]$, $A(t)\in[K]$ (from algorithm $\cA$), decision for user $j\in[M]$ in Chapter~\ref{chapter:5} (from algorithm $\cA^j$)}
@@
 \nomenclature[O]{$\Pr$}{Probability measure under a probabilistic model}
-\nomenclature[O]{$\E$}{Expectancy under a probabilistic model}
+\nomenclature[O]{$\E$}{Expectation under a probabilistic model}
 \nomenclature[O]{$\mathbbm{1}(E)$}{Indicator function of an event $E$ ($= 1$ if and only if the event $E$ is true)}

### 2019-06-18 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Typo

@@
 The \UCB{} algorithm is implemented independently by each device, we denote it ``first-stage'' \UCB, and we present it in Algorithm~\ref{algo:43:UCB}.
-It is the same algorithm as the Algorithm~\ref{algo:2:indexPolicy} in Section~\ref{sub:2:IndexPolicies} (using the same \UCB{} indexes as we defined them in equation \eqref{eq:UCB_index}),
+It is the same algorithm as the Algorithm~\ref{algo:2:indexPolicy} in Section~\ref{sub:2:IndexPolicies} (using the same \UCB{} indexes as we defined them in equation \eqref{eq:2:UCB_index}),
 but we write it again to make clear the difference with first transmission and retransmission of messages.

### 2019-06-18 2-Chapters/5-Chapter/chapter5.tex
Typo

@@
 
-If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm (see Algorithm~\ref{algo:2:naiveStrategies}), running on top of a round-Robin phase can achieve order-optimal regret.
+If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm (see Section~\ref{sub:2:naiveSimpleStrategies}), running on top of a round-Robin phase can achieve order-optimal regret.
 If the players know the time horizon $T$, then they can fix a confidence level $\delta$.

### 2019-06-18 2-Chapters/1-Chapter/chapter1.tex
CHAUD

@@
 
-\TODOL{Mardi soir : je dois encore relire le 1.1 pour les fautes d'orthographe (et les éventuels erreurs historiques) !}
+\TODOL{Mardi soir : je dois encore relire le 1.1 pour les éventuelles erreurs historiques !}
 \TODOL{Mardi soir : je dois encore améliorer le 1.2 et être plus concis ?}
@@
 Robbins \cite{Robbins52}, and more recently it is an active field of research, since the late $1990$s \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}.
+Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}.
 
@@
 If the different channels are not uniformly used by the PU, and if we assume a stationary hypothesis on the PU traffic, then the goal of the SU boils down to explore the different channels and exploiting the best ones.
-This frames the Opportunistic Spectrum Access problem as a exploration/exploitation problem, white a finite set of actions (the channels, also called arms),
+This frames the Opportunistic Spectrum Access problem as an exploration/exploitation problem, white a finite set of actions (the channels, also called \emph{arms}),
 in a sequential action-then-feedback cycle (time steps are $t\in\N^*$),
@@
 
-His works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the \UCB{} algorithm,
+Their works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the \UCB{} algorithm,
 along the early works of Q. Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
@@
 Shortly after in $2014$, proof-of-concepts using real-world radio hardware and Software Defined Radio were developed by C. Moy and his student C. Robert \cite{RobertSDR2014,MoyWSR2014}.
-In a second PhD thesis supervised by C.
-Moy \cite{Modi17PhD} from $2014$ to $2017$, N.
-Modi studied the impact of using MAB algorithms to optimize channel selection on the battery life of a wireless device.
+In a second PhD thesis \cite{Modi17PhD}, N. Modi studied from $2014$ to $2017$ the impact of using MAB algorithms to optimize channel selection on the battery life of a wireless device.
 On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
@@
 
-In $2017$, C. Moy continued to work on this direction, with a post-doctoral student, S. Darak, leading to the following publications
-\cite{darak2016bayesian,Darak16,modiDemo2016,kumar2016two}.
+In $2017$, C. Moy continued to work on this direction, with a post-doctoral student, S. Darak, leading to publications such as
+\cite{darak2016bayesian,Darak16}.
 %
@@
 
-\TODOL{Je suis content de la suite de l'intro : aperçu des contributions, plans de thèse, liste des publications}
+\TODOL{Je suis content de la SUITE DE L'INTRO : aperçu des contributions, plans de thèse, liste des publications... Ca vous va ?}
 
@@
     % \item
-In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2}, required for the rest of the document as we introduce the MAB models, the concepts and the notations used in this thesis.
+\textcolor{darkred}{First, in Part~\ref{part:Introduction}}, we start by the next Chapter~\ref{chapter:2}, required for the rest of the document as we introduce the MAB models, the concepts and the notations used in this thesis.
 Conversely, even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use numerical simulations based on our simulation library SMPyBandits, the Chapter~\ref{chapter:3} where we present it is not required to understand them.
@@
     % \item
-Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
+\textcolor{darkgreen}{Then, the second Part~\ref{part:MABIOT}} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
 Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where we show that MAB algorithms can be used with success.
@@
 
-    Note that we are already working on an extended version that will be submitted to a journal on Machine Learning for Wireless Communications, in July $2019$.
+    \emph{Note that we are already working on an extended version that will be submitted to a journal on Machine Learning for Wireless Communications, in July $2019$.}
 
@@
 
-    Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.
+    \emph{Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
 

### 2019-06-18 2-Chapters/Conclusion/conclusion.tex
CHAUD

@@
 It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if you install the library, but can also be used on cloud platforms, such as
-Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} for free.},
+Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} from your browser.},
 or Google Colab\footnote{~See \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}.}.

### 2019-06-19 2-Chapters/1-Chapter/chapter1.tex
MIAM

@@
 If we sum-up the studied problems and summarize them in one question, it would be the following:
-\emph{``Can we adapt the decision making tools already successfully applied to Cognitve Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
+\emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
 %
@@
 we were not only interested by the usual measure of performance of any MAB algorithm (its regret, see below in Section~\ref{sec:2:lowerUpperBoundsRegret}),
-but also by their empirical performances in terms of computational complexity and storage requirements, both from the point of view of the theoretical analyses and real-world measurements on time and memory footprints of the different algorithms.
+but also by their empirical performances in terms of computational complexity and storage requirements, both from the point of view of the theoretical analyses and real-world measurements on time and memory footprints.
 
@@
 Our exploration of the large number of MAB algorithms and models developed in the recent literature
-has gave us the ambition to write a single piece of software were we could easily implement new models and algorithms, in order to compare the existing ones and empirically explores the performances of newly proposed algorithms.
+has gave us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explores the performances of newly proposed algorithms.
 To answer this goal, we produced a Python library of simulation of MAB problems.
 %
-We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR} and hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
 We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
@@
 %
-Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement in a given IoT object.
+Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement, in a given IoT object, in order to let this object adapts robustly in any environment.
 To answer this question, we present two approaches.
-The first one is an empirical comparison of the most efficient and well-known existing algorithms in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}, where it appears that simple algorithms such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal} are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
-The second approach is an online algorithm selection, consisting in aggregating a set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
-This work on aggregating MAB algorithms was motivated for the OSA case of Cognitive Radio, where most previous research works only considered the $\UCB_1$ algorithm without justifying this choice rigorously,
+The first one is an empirical comparison of the most efficient and well-known existing algorithms, in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}.
+We confirm that famous and not too sophisticated algorithms, such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
+The second approach is an online algorithm selection, consisting in aggregating a (finite) set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
+This work on aggregating MAB algorithms was motivated for the OSA case of Cognitive Radio, where many previous research works only considered the $\UCB_1$ algorithm without justifying this choice rigorously,
 and was presented at the IEEE WCNC conference in Barcelona, Spain, in April $2018$.
@@
 In Chapter~\ref{chapter:4}, we study two models of IoT networks.
-The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use the acknowledgement as feedback to learn the best channel for their uplink communications.
+The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use the acknowledgement as feedback, in order to learn the best channel for their uplink communications.
 The goal of this first model of applying decentralized RL on the device side is to increase the Quality of Service (QoS) of the application of such IoT network, by reducing the rate of failed transmission.
-The base station will thus receive more data from the devices its serves, if they can successfully learn an efficient spectrum access scheme.
+The base station will thus receive more up-link packets from the devices it serves, if they can successfully learn an efficient spectrum access scheme.
 %
-On the other hand, the second model we study considers packet retransmissions, and while the concept stays similar, the consequence of an efficient decentralized learning will be both an increase battery life for the devices as well as an increase QoS.
+On the other hand, the second model we study considers packet retransmissions, and while the concept stays similar, the application of an efficient decentralized learning results in both an increased battery life for each of the devices, as well as an increase QoS for the complete network.
 
-We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of up-to thousands of end-devices and a single base station.
+We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of thousands (or more) of end-devices and a single base station.
 We consider a time-frequency slotted ALOHA-based protocol, by assuming a perfect synchronization between all devices and the base station.
@@
 We first detail three base-lines, a naive approach and two centralized ``oracle'' approaches,
-that are used to evaluate the performance of MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communication rate.
-We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes more and more unpredictable.
+that are used to evaluate the performance of MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communications rate.
+We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from the other devices becomes more and more fluctuating and unpredictable.
 %
-Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
+Even without changing anything on the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
 This model constitutes the first article written during this thesis with R. Bonnefoi, another PhD student of our team SCEE.
@@
 Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations is a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
-In Section~\ref{sec:4:firstModel}, we propose the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network.
+In Section~\ref{sec:4:firstModel}, we detail what we believe to be the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network.
 We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements, that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
@@
 This is again the result of a fruitful collaboration with R.
-Bonnefoi, and this work was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
+Bonnefoi, and this work was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
 
-We then consider an extension of our first model, following the idea of packet retransmission that is at the root of the ALOHA protocol \cite{Abramson1970,Roberts75}.
+We then consider an extension of our first model, following the idea of packet retransmissions, that is at the root of the famous ALOHA protocol \cite{Abramson1970,Roberts75}.
+The idea is that when a device sends an uplink message but does not receive its \Ack{} before a certain delay, instead of going back to sleep mode and waiting for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, $m \leq 10$ times), after some small random waiting time.
+%
 This second model proposes to also use decision making at the second layer of the MAC protocol, that is to also use RL and MAB algorithms, in order to learn the best channel to use for retransmitting any packet that suffered from a collision at its first transmission.
@@
 %
-In the two models presented above, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm,
-each of them targeting their local objective of maximizing their cumulated reward (in a selfish way), in order to optimize the complete system.
-It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the prisoner dilemma.
-So it is quite surprising that our numerical simulations as well as the realistic PoC showed that decentralized selfish MAB learning lead to efficient coordination between devices.
+In the two models presented above, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm in order to optimize the complete system,
+meaning that each of them target their own local objective, which is to maximize their cumulated reward (in a selfish way).
+It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma} or other popular ``dilemma'' games.
+% , and for more examples we refer to the book \cite{CesaLugosi06}.
+% https://www.ii.uni.wroc.pl/~lukstafi/pmwiki/uploads/AGT/Prediction_Learning_and_Games.pdf
+So it is quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices.
 
-We first tried to analyze the performance of this decentralized algorithm, that we refer to as Selfish, in the model of Section~\ref{sec:4:firstModel},
+We first tried to analyze the performance of this decentralized algorithm, that we refer to as \Selfish, in the model of Section~\ref{sec:4:firstModel},
 but due to the random number of active devices at each time step (as soon as the probability $p$ of activation is $p < 1$), we were unable to propose a clean analysis.
 %
-That is why in Chapter~\ref{chapter:5}, we relax the hypothesis of having $M \gg K$ devices in a network with $K$ orthogonal wireless channel, equivalent to having $p < 1$, and we consider the case of devices that each communicate at every step (\ie, $p=1$), and so only up-to $M \leq K$ devices.
-We present the studied model in Section~\ref{sec:5:model}, that comes with different variants depending on the feedback level, and cover both the OSA case (\ie, with sensing information) or the IoT case (\ie, without sensing feedback).
-The goal was to understand the heuristic of Chapter~\ref{chapter:4}, Selfish, in the simpler framework of multi-player MAB, which has been studied before \cite{Zhao10,Anandkumar10,Anandkumar11} for the case of OSA.
+That is why in Chapter~\ref{chapter:5}, we relax the hypothesis of having $M \gg K$ (or even simply $M > K$) devices in a network with $K$ orthogonal wireless channel, equivalent to having $p < 1$, and we consider the case of devices communicating at every time step (\ie, $p=1$), and so we restrict to only up-to $M \leq K$ devices.
+The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing information) or the IoT case (\ie, without sensing feedback).
+The goal was to understand the heuristic of Chapter~\ref{chapter:4}, \Selfish, in the simpler framework of multi-player MAB, which has been studied before \cite{Zhao10,Anandkumar10,Anandkumar11} for the case of OSA.
+%
+The OSA case covered by our model is actually slightly different as the one considered by \cite{Jouini10} and other previous works,
+as our model requires that an \Ack{} is sent back by the base station if the transmission was successful, even in the case where sensing information is available.
+It differs from the previously studied models of applying MAB for OSA, as they only considered synchronized devices, one SU and PU, and thus if the sensing indicates that one channel is free at one time slot, the up-link message sent by the SU device is sure to be successfully received by the base station (in the ideal model), so there is no risk of collision.
 
-On the first hand, in the OSA setting, we failed to obtain positive result for the Selfish policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability.
+On the first hand, in the OSA setting, we failed to obtain positive result for the Selfish policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
 This result was later confirmed and generalized by \cite{BoursierPerchet18}.
@@
 %
-%
-and it relaunched the research on this model as a few independent teams started to work on different questions after its publication.
+and its publication relaunched the research on this model, as a few independent teams started to work on different questions after our presentation.
 %
@@
 Perchet \cite{BoursierPerchet18}.
-Even if their approach is costly empirically, the result is surprising and very interesting.
-We discuss it, along with other recent research works such as \cite{LugosiMehrabian18}, in Section~\ref{sec:5:literatureReviewOtherModels}.
+Even if their approach is costly empirically, the result is surprising and very interesting: they showed that even under the strong hypothesis of no direct communication between players and no central coordination, the players can use the collision feedback (\ie, receiving or not an \Ack) as an indirect way to communicate bits of information with each other.
+This idea requires that devices coordinate and split the time steps between ``listening'' and ``talking'' phases, and the authors denote it the ``communication trick''.
+As they showed, after a long enough initialization period, the $M$ players have elected a ``master'' player and $M-1$ ``followers'' players (with high probability), and thanks to this communication trick, the problem can boil down to a centralized multiple-play (MP) MAB, which can be solved efficiently with the MP generalization of \klUCB{} \cite{Luedtke16} or Thompson sampling \cite{Komiyama15}.
+
+We discuss more in details this interesting follow-up of our article \cite{BoursierPerchet18}, along with other recent research works such as \cite{LugosiMehrabian18}, in Section~\ref{sec:5:literatureReviewOtherModels}.
 %
-Moreover, the case of different means of arms for the different players was studied in \cite{KaufmannAbbas19}, and this generalization of our model is very interested for CR applications, as the mean of an arm is the average quality (\eg, availability) of a radio channel, and as such it has no reason to be uniform among end-devices (\ie, players).
+Moreover, the case of different means of arms for the different players was studied in \cite{Bistritz18,KaufmannAbbas19}, and this generalization of our model is very interested for CR applications, as the mean of an arm is the average quality (\eg, availability) of a radio channel, and as such it has no reason to be uniform among end-devices (\ie, players).
 %
@@
 
-\TODOL{Emilie me dit de continuer comme ça :
-    En tout cas ce chapitre 1 commence à avoir bonne forme pour l'école
-    doctorale, il n'y aura plus qu'à finaliser la deuxième partie de la
-    section 1.3 (la première est très bien). Tu tiens le bon bout !
-}
-
-As explained above, the model of Chapter~\ref{chapter:4} was found intractable to analyze because we consider an IoT network with many end-devices, all following random activation patterns.
+As explained above, the model of Chapter~\ref{chapter:4} was found intractable to analyze, mainly because we consider an IoT network with many end-devices, all following random activation patterns.
+The difficulty thus resides in the fact that we are trying to analyze MAB algorithms, designed to tackle stationary problems, on a non-stationary problem. %XXX
 Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, but leads to the model of Chapter~\ref{chapter:5}.
@@
 %
-On the other hand,
-FIXME explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
-\TODOL{Mardi soir je dois encore expliquer ça !}
+On the other hand, another way to relax the non-stationary hypothesis is to take the point-of-view of one device, like in the OSA case mentioned above, where the focus is on one SU surrounded by many PU having stationary behaviors.
+If we focus on one IoT device, we can assume that the environment (\ie, the surrounding devices) is stationary on some time intervals, which leads to the last contribution and chapter of this thesis.
+% DONE explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
 
 \paragraph{Chapter~\ref{chapter:6}.}
-%
-% FIXME explain non stationarity !
 %
@@
 We prove that by combining the asymptotically optimal policy \klUCB, designed for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of our proposed algorithm.
+We make the same kind XXX
 %
@@
 Our work also lead to the article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results and submitted soon to a journal (probably the Journal of Machine Learning Research) in fall $2019$.
+
+\paragraph{Conclusion.}
+%
+We conclude this thesis in Chapter~\ref{chapter:conclusion}, to give some perspectives and directions about possible future works.
+Most of the contributions that we present in this thesis call for more developments, on both directions of providing more mathematical analysis and implementing new proof-of-concepts using real wireless hardware and realistic Internet of Things networks.
+%
+The Appendix of this manuscript starts with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks} we discuss about our work on doubling-trick
 
@@
 % We proposed new algorithms, some of which are proven to converge efficiently.
-
-\TODOL{Je suis content de la SUITE DE L'INTRO : aperçu des contributions, plans de thèse, liste des publications... Ca vous va ?}
 

### 2019-06-19 2-Chapters/4-Chapter/chapter4.tex
MIAM

@@
 In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
-Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
+We highlight that even without changing anything on the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
 We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,

### 2019-06-19 1-Preamble/5-Abstract/resume_fr.tex
New references. I need to finish Ch1 and Ch2

@@
 J'ai peur de ne pas réussir à écrire 20 pages avec des maths en français juste pour pouvoir citer tel théorème ou tel résultat.}
+
+\TODOL{En fait, maintenant que le ch1 est écrit, je peux traduire (et couper un peu) ce chapitre 1, et ça ira !}
 
@@
     \item
-Dans la Partie~\ref{part:Introduction}, nous commençons par le Chapitre~\ref{chapter:2} où nous présentons les modèles MAB, les concepts et les notations utilisés dans tout ce document.
+\textcolor{darkred}{Dans la Partie~\ref{part:Introduction}}, nous commençons par le Chapitre~\ref{chapter:2} où nous présentons les modèles MAB, les concepts et les notations utilisés dans tout ce document.
 Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
@@
     \item
-La deuxième Partie~\ref{part:MABIOT} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus presque indépendamment.
+\textcolor{darkgreen}{La deuxième Partie~\ref{part:MABIOT}} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus presque indépendamment.
 Le Chapitre~\ref{chapter:4} commence par présenter différents modèles de réseaux IdO où les algorithmes MAB ont été utilisés avec succès.

### 2019-06-19 2-Chapters/1-Chapter/chapter1.tex
New references. I need to finish Ch1 and Ch2

@@
 \TODOL{Mardi soir : je dois encore relire le 1.1 pour les éventuelles erreurs historiques !}
-\TODOL{Mardi soir : je dois encore améliorer le 1.2 et être plus concis ?}
-\TODOL{Mardi soir : je dois relire la fin du 1.3, et terminer le dernier point}
+\TODOL{Mardi soir : je dois encore améliorer le 1.2 et 1.3 ?}
 
@@
 During the last 150 years, mankind has developed a wide range of different communication technologies, and since the late $1890$s wireless communications between devices have been made possible, and more and more frequent in our lives.
-With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed world-wide, for a large range of different applications.
+With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, for a large range of different applications.
 It is now well-known that with the current trend of population increase and with the energy crisis, any newly deployed technology should be cheap and energy efficient,
@@
 Broadcast-based systems are not only used for TV or radio,
-and for instance two other famous and world-wide deployed examples are the long-range Global Positioning System (GPS), and short-range infrared remote controllers.
+and for instance two other famous and worldwide deployed examples are the long-range global navigation satellite systems (such as USA's GPS or European Union's Galileo), and short-range infrared remote controllers.
+% https://en.wikipedia.org/wiki/Satellite_navigation
 
@@
 meaning that while a band can be allocated to a certain unique usage, it can be free from any user in certain times and/or places.
-We refer to \cite{valenta2010survey} for a survey on spectrum utilization in Europe.
+We refer to \cite{patil2011survey} for a survey on worldwide spectrum utilization, and to \cite{valenta2010survey} for the situation in Europe.
 
@@
     \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
-    \caption{A chart representing the allocation of radio spectrum in the USA in $2016$.
+    \caption[A chart representing the allocation of radio spectrum in the USA in $2016$]{A chart representing the allocation of radio spectrum in the USA in $2016$.
 Only a small share of the spectrum is not yet allocated, the top-left corner in white (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz}).
@@
 To highlight that the spectrum scarcity is not only an issue in the USA,
-we quote the abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by Joëlle Toledano \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull}, who was commissioned by the French government in $2014$.
+we quote the abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by J.
+Toledano \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull}, who was commissioned by the French government in $2014$.
 Different important problems are mentioned:
 %
-\begin{itemize}\tightlist
+\begin{itemize}%\tightlist
     \item
     \emph{``In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy.''}
-    And the usage of wireless communications is nowadays world-wide and omnipresent:
+    And the usage of wireless communications is nowadays worldwide and omnipresent:
     \emph{``Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.''}
@@
     Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.''}
-    As illustrated above in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}.
+    As illustrated above in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}, and discussed in \cite{patil2011survey,valenta2010survey}.
     \item
@@
 In this thesis, we study Internet of Things networks, and these two constraints of low power consumption (or long battery life) and low duty cycle are essential.
-More precisely, we study the possible interconnections between \textbf{Machine Learning} and \textbf{Cognitive Radio} applied for IoT networks.
+More precisely, we study the possible interconnections between \textbf{Cognitive Radio} and \textbf{Machine Learning} applied for IoT networks.
 Let us define and detail both concepts.
@@
 
-\TODOL{Emilie pense que je devrais faire ça mieux :
-    Quant au survey du machine learning (Figure 1.1), je ne pense pas que
-    cela soit nécessaire. Je trouve ça plus pertinent d'introduire le
-    problème de décision séquentielle qui nous intéresse (qui n'est pas
-    juste un sub-field du machine learning d'ailleurs, la communauté
-    statistique s'y intéresse depuis bien plus longtemps !), sur l'exemple
-    de l'OSA.
-}
+% \TODOL{Emilie pense que je devrais faire ça mieux :
+%     Quant au survey du machine learning (Figure 1.1), je ne pense pas que
+%     cela soit nécessaire. Je trouve ça plus pertinent d'introduire le
+%     problème de décision séquentielle qui nous intéresse (qui n'est pas
+%     juste un sub-field du machine learning d'ailleurs, la communauté
+%     statistique s'y intéresse depuis bien plus longtemps !), sur l'exemple
+%     de l'OSA.
+% }
 
@@
     \includegraphics[width=0.95\textwidth]{machine-learning-map.pdf}
-\caption{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.
+\caption[Classification of the main types of Machine Learning]{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.
 MAB is RL with partial information, and RL is ML with no supervision or data, just an environment to interact with.}
@@
 A player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$), and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
-The goal of the player is to maximize its reward, by trials and errors (\ie,actions and rewards).
+The goal of the player is to maximize its reward, by trials and errors (\ie, actions and rewards).
 Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial), etc.
@@
 % Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
-More precisely, in this thesis we focus on a certain kind of Reinforcement Learning models where the learning does not have access to the entire reaction of the environment after taking its action.
+More precisely, in this thesis we focus on a certain kind of Reinforcement Learning models where the learner does not have access to the entire reaction of the environment after taking its action.
 In other words, we focus on RL with limited feedback, when the player only sees the reward given by its action at every round, and not the reward that would have been given if she chose any other action.
 This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} in details in the next Chapter~\ref{chapter:2}.
-Clinical trials and treatment discovery was historically the first application of MAB, in $1933$ with the early work of W. Thompson \cite{Thompson33}.
+Clinical trials and treatment discovery were historically the first applications of MAB since $1930$s, with the early work of W. Thompson \cite{Thompson33}.
 %
@@
 Robbins \cite{Robbins52}, and more recently it is an active field of research, since the late $1990$s \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}.
+Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}, and the wide range of applications of MAB in the recent years.
 
@@
 The first one is an empirical comparison of the most efficient and well-known existing algorithms, in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}.
-We confirm that famous and not too sophisticated algorithms, such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
+We confirm that well-known and not too sophisticated algorithms, such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
 The second approach is an online algorithm selection, consisting in aggregating a (finite) set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
-This work on aggregating MAB algorithms was motivated for the OSA case of Cognitive Radio, where many previous research works only considered the $\UCB_1$ algorithm without justifying this choice rigorously,
-and was presented at the IEEE WCNC conference in Barcelona, Spain, in April $2018$.
+This work on aggregating MAB algorithms was motivated for the OSA case of Cognitive Radio, where many previous research works only considered the $\UCB_1$ algorithm without really justifying this choice.
+This contribution was presented at the IEEE WCNC conference in Barcelona, Spain, in April $2018$.
 
@@
 %
-On the other hand, the second model we study considers packet retransmissions, and while the concept stays similar, the application of an efficient decentralized learning results in both an increased battery life for each of the devices, as well as an increase QoS for the complete network.
+The second model we study then considers packet retransmissions, and while the concept stays similar, the application of an efficient decentralized learning results in both an increased battery life for each of the devices, as well as an increase QoS for the complete network.
 
@@
 %
-Even without changing anything on the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
-This model constitutes the first article written during this thesis with R. Bonnefoi, another PhD student of our team SCEE.
-We presented this first work, which obtained the ``best paper award'', at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$ \cite{Bonnefoi17}.
+Even without changing anything on the level of the considered IoT standard, our proposal is just an add-on capability, that can be set-up on each device, on a unit-per-unit basis.
+This model constitutes the first article written during this thesis, and initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
+We presented this first work at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, and obtained the ``best paper award'' \cite{Bonnefoi17}.
 
@@
 On the other hand, we were able to propose new algorithms for this multi-player bandit problem with sensing information, and we analyzed our proposal \MCTopM-\klUCB, to show that it achieves a finite-time logarithmic regret upper bound, improving over previous state-of-the-art results.
+Our algorithm also achieves a logarithmic number of collisions and arm switches, and allows a fixed group of $M$ devices to efficiently learn to use the $M$ best channels orthogonally for almost all their up-link communications.
 %
@@
 Following two recent previous works \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, we combine an efficient index policy (\klUCB) and an efficient change-point detection test, under the assumption of bounded rewards.
-Building on very recent results on the Generalized Likelihood Ratio test (GLRT) for Gaussian and sub-Gaussian variables \cite{Maillard2018GLR}, we instead focus on bounded rewards and Bernoulli distributions.
+We build on very recent results on the Generalized Likelihood Ratio test (GLRT) for Gaussian and sub-Gaussian variables \cite{Maillard2018GLR}, and we instead focus on bounded rewards and Bernoulli distributions.
 Using the fact that bounded variables in $[0,1]$ are not only $1/4$ sub-Gaussian but also sub-Bernoulli, we prove the first finite-time guarantees for the GLRT for bounded variables.
@@
 We prove that by combining the asymptotically optimal policy \klUCB, designed for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of our proposed algorithm.
-We make the same kind XXX
+Like our competitors, we make the some hypothesis on the length of the stationary intervals, and the best regret upper-bound is obtained when the algorithm knows before-hand the horizon and the number of change-points, but our algorithm does not need to know any other knowledge of the problem difficulty to be tuned optimally.
 %
@@
 %
-The Appendix of this manuscript starts with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks} we discuss about our work on doubling-trick
+This manuscript ends with the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks} we discuss about our work on doubling-trick, in order to remove the need for an algorithm to know the horizon of the bandit game before starting to play it.
+The rest of the appendix contains a list of abbreviations and notations, then lists of figures, algorithms, code samples and tables.
+The list of the bibliographical references used in this thesis are given at the end.
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Summary of the contributions}
+\subsection{Summary of the contributions}
 \label{sec:1:summaryOfContributions}

### 2019-06-19 2-Chapters/2-Chapter/chapter2.tex
New references. I need to finish Ch1 and Ch2

@@
 
-More recently, the \klUCB-switch from \cite{Garivier18} is the first algorithm to be proven to
+More recently, the \klUCB-switch from \cite{GarivierHadiji2018} is the first algorithm to be proven to
 enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, that is, matching the $C_I \ln(T)$ lower bound by \cite{LaiRobbins85}.

### 2019-06-19 2-Chapters/3-Chapter/chapter3.tex
New references. I need to finish Ch1 and Ch2

@@
 For example, Figures~\ref{fig:3:twoScreenshotsOfTheDocumentation} and \ref{fig:3:twoMoreScreenshotsOfTheDocumentation} below show four screenshots from the documentation.
-We show the main page, the list of algorithms in the \texttt{Policies} module, an example of the documentation of one algorithm (for the \klUCB-Switch algorithm from \cite{Garivier18}), and a page detailing the API and organization of the library as well as instructions to follow if someone wants to implement a new model or a new algorithm.
+We show the main page, the list of algorithms in the \texttt{Policies} module, an example of the documentation of one algorithm (for the \klUCB-Switch algorithm from \cite{GarivierHadiji2018}), and a page detailing the API and organization of the library as well as instructions to follow if someone wants to implement a new model or a new algorithm.
 

### 2019-06-19 2-Chapters/Conclusion/conclusion.tex
New references. I need to finish Ch1 and Ch2

--- 2019-06-16 2-Chapters/Conclusion/conclusion.tex
@@
 It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if you install the library, but can also be used on cloud platforms, such as
-Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} for free.},
+Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} from your browser.}
 or Google Colab\footnote{~See \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}.}.

### 2019-06-19 2-Chapters/1-Chapter/chapter1.tex
Beautify bibtex

@@
 %
-The specificities of IoT networks can be listed as follows.
+The specificities of IoT networks can be listed as follows,
+and we refer to \cite{Centenaro16} for more details.
 \begin{itemize}\tightlist
@@
 \end{itemize}
-\TODOL{FIXME more specificities? Est-ce que je dois mentionner d'autres contraintes ? Il me faudrait aussi une ou deux références genre survey des standards IoT}
+\TODOL{FIXME more specificities? Est-ce que je dois mentionner d'autres contraintes }
 
@@
 
-\subsection{Focusing on MAB algorithms}
+% \subsection{Focusing on MAB algorithms}
+\subsection{Part~\ref{part:Introduction}}
 
@@
 
-\subsection{Focusing on IoT networks.}
+% \subsection{Focusing on IoT networks.}
+\subsection{Part~\ref{part:MABIOT}}
 
@@
 As explained above, the model of Chapter~\ref{chapter:4} was found intractable to analyze, mainly because we consider an IoT network with many end-devices, all following random activation patterns.
-The difficulty thus resides in the fact that we are trying to analyze MAB algorithms, designed to tackle stationary problems, on a non-stationary problem. %XXX
-Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, but leads to the model of Chapter~\ref{chapter:5}.
+The difficulty does not reside in the fact that we are trying to analyze MAB algorithms, designed to tackle stationary problems, on a non-stationary problem,
+but rather that we are analyzing algorithms which are all playing in different (random) subsets of the global synchronized time.
+Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, but leads to the model of Chapter~\ref{chapter:5}.\\
 %
+\indent
 On the first hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let the devices learn an optimal orthogonal affectation of $M \leq K$ devices in the $K$ channels.
@@
 On the other hand, another way to relax the non-stationary hypothesis is to take the point-of-view of one device, like in the OSA case mentioned above, where the focus is on one SU surrounded by many PU having stationary behaviors.
-If we focus on one IoT device, we can assume that the environment (\ie, the surrounding devices) is stationary on some time intervals, which leads to the last contribution and chapter of this thesis.
+If we focus on one IoT device, we could assume that the environment (\ie, the surrounding devices) is stationary on some time intervals, which leads to the last contribution and chapter of this thesis.
 % DONE explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
@@
 We begin our last Chapter~\ref{chapter:6} by reviewing the existing works,
-and we chose to focus on the piece-wise stationary problem.
-Two main families of algorithms have been proposed for this problem,
+and we chose to focus on the piece-wise stationary problem,
+meaning that the underlying bandit problem is stationary on some intervals, separated by change-points located at unknown times.
+Assuming the environment to be piece-wise stationary indeed makes sense for applications to wireless networks, where a change-point can for instance correspond to the arrival of a new group of end-devices in the network.
+Examples of such situation can be a new company arrives on the market, or a farmer installs sensors on a herd of cows, etc.
+%
+Two main families of algorithms have been proposed for the piece-wise stationary problem,
 that usually consist in combining an efficient policy designed for the stationary MAB problem (\eg, \UCB{} or \klUCB) and a way to adapt to changes in the arms distributions.
-Passively adaptive policies use a window of a fixed or evolving size, or a discount factor to slowly forget about the past observations,
-while actively adaptive policies use a statistic test to detect the change-point (\ie, the time steps between two consecutive stationary intervals).
+Passively adaptive policies use a window of a fixed or evolving size \cite{Garivier11UCBDiscount}, or a discount factor, in order to forget about the past observations \cite{Kocsis06,Gupta11thompson},
+while actively adaptive policies use a statistical test to detect the change-points \cite{MellorShapiro13,Allesiardo15}.
+%
 As the recent literature showed that the later approach is usually more competitive, and usually obtains better results from both empirical and theoretical aspects, we chose to develop our own actively adaptive algorithm.
-
-\TODOL{Mardi soir je peux mieux expliquer, là je suis long et chiant je trouve}
 
@@
 This manuscript ends with the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks} we discuss about our work on doubling-trick, in order to remove the need for an algorithm to know the horizon of the bandit game before starting to play it.
-The rest of the appendix contains a list of abbreviations and notations, then lists of figures, algorithms, code samples and tables.
-The list of the bibliographical references used in this thesis are given at the end.
+The rest of the appendix contains a list of abbreviations and notations, then lists of figures, algorithms, code samples and tables,
+and finally the list of the bibliographical references.
 
@@
 
-\newpage % WARNING
+% \newpage % WARNING
 
@@
 
-\newpage  % WARNING
+% \newpage  % WARNING
 
@@
 The reading order of the manuscript can be any top-down path between the Introduction in the current Chapter~\ref{chapter:1}, and the Conclusion in the last Chapter~\ref{chapter:conclusion}.
-As the graph in Figure~\ref{fig:1:organization} shows,
-the thesis is organized in two parts.
+As the graph in Figure~\ref{fig:1:organization} shows, the thesis is organized in two parts.
 

### 2019-06-19 2-Chapters/3-Chapter/chapter3.tex
Relaunching these simulations

@@
 Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
-Its documentation has seen already about $20000$ visits, the project on GitHub had $85$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
+Its documentation has seen already about $20000$ visits, the project on GitHub had $88$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
@@
 We give below in Code~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
-It uses \texttt{for} loops and environment variables from Bash, and not Python, but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $T$ and $K$.
+It uses \texttt{for} loops and environment variables from Bash, and not Python, mainly for concision,
+but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $T$ and $K$.
 
@@
 $ for T in $(seq 5000 5000 50000); do \
-      for K in $(seq 2 2 32); do \
     DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
-          N=1000  BAYES=True    T=$T         K=$K \
+        N=1000  BAYES=True    T=$T         K=8 \
         python3 main.py configuration.py \
     && cp logs/main_py3_log.txt \
-              logs/main_py3_log__N1000_BAYES_T{T}_K{K}.txt \
-  done; done
+          logs/main_py3_log__N1000_BAYES_T{T}_K8.txt \
+  done
+$ for K in $(seq 2 2 32); do \
+    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
+        N=1000  BAYES=True    T=5000       K=$K \
+        python3 main.py configuration.py \
+    && cp logs/main_py3_log.txt \
+          logs/main_py3_log__N1000_BAYES_T5000_K{K}.txt \
+done
     \end{minted}
-    \caption{Bash code to run the large experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.}
+    \caption{Bash code to run the large-scale experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}, for $K=8$ and $T\in\{5000,\dots,50000\}$ and $T=5000$ and $K\in\{2,\dots,32\}$.}
     % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:BashCodeToLaunchLargeExperiments}.}
@@
 % --------------------------------------------------------------
-\subsection{Final remark: how we used parallelism in SMPyBandits}
+\subsection{Using multi-cores parallelism to speed-up simulations in SMPyBandits}
 \label{sub:3:parallelSimulations}
@@
 the setting is the same:
-we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=100$ like in Chapter~\ref{chapter:6}).
+we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=1000$ like in Chapter~\ref{chapter:6}).
 %
@@
 But parallelism can help greatly for the two other axes: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
-Similarly, if we want to run $100$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $100/4=25$ repetitions on each process also bring a speed-up factor of $4$.
+Similarly, if we want to run $1000$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $1000/4=25$ repetitions on each process also bring a speed-up factor close to $4$.
 %
@@
 
-\textbf{What we did implement: \texttt{Joblib} for multi-core simulations.}
+\textbf{\texttt{Joblib} for multi-core simulations.}
 %
@@
 As explained in the page \href{https://smpybandits.github.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}, we implemented in SMPyBandits an easy way to run any simulations on $n$ cores of a machine, using the \texttt{Joblib} \cite{joblib} library.
+
 It is implemented in a completely transparent way, and if someone uses the command-line variable to configure experiments, switching from using one core to using all the cores of the machine is as simple as changing \texttt{N\_JOBS=1} to \texttt{N\_JOBS=-1}, like in the Code Example~\ref{lst:3:runOneCoreOrMore} below.
+Similarly, changing from within the Python code is simple, by changing the \texttt{configuration["n\_jobs"]} value.
 
@@
         \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
-$ BAYES=False ARM_TYPE=Bernoulli N=100 T=10000 K=9 N_JOBS=1 \
-  python3 main.py configuration.py
+$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=1 \
+  python3 main.py configuration.py  # on only one core!
+
+$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=-1 \
+  python3 main.py configuration.py  # parallel on all the cores!
+
+$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=100 \
+  python3 main.py configuration.py  # parallel with too many jobs!
         \end{minted}
-        \caption{To run such simulation using all the CPU cores, use \texttt{N\_JOBS=-1} instead.}
+        \caption{Running a simulation using either one CPU core with \texttt{N\_JOBS=1}, or all of them with \texttt{N\_JOBS=-1}, or more jobs than CPU cores with \texttt{N\_JOBS=20} (if there are $4$ cores).
+The later option is not advised, and \texttt{N\_JOBS=-1} gives the optimal speed-up factor.}
         % \captionof{listing}{Run a simulation using 1 or \texttt{N} cores \label{lst:3:runOneCoreOrMore}.}
@@
 
+\textbf{Effect of the number of jobs.}
+%
 As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
@@
 We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $20$ jobs.
-We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=1000$, and different number of repetitions and number of jobs.
-It clearly illustrates that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
+%
+We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=10000$, and different number of repetitions and number of jobs.
+These three experiments correspond to the three lines of Code Example~\ref{lst:3:runOneCoreOrMore} shown above.
+%
+The results of Table~\ref{table:3:speedUpTimeParallelComputations} and of Figure~\ref{fig:3:analyze_speedup_as_nb_of_jobs}
+clearly illustrate that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
 Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
@@
     \begin{tabular}{c|rrr}
-    \textbf{Repetitions} $\;$ \textbackslash $\;$ number of jobs \texttt{N\_JOBS} & $1$ & $4$ ($=$ nb cores) & $20$ ($>$ nb cores) \\
+    \textbf{Repetitions} & \texttt{N\_JOBS}$=1$ & \textcolor{darkgreen}{\texttt{N\_JOBS}$=4$ ($= n_{\text{cores}}$)} & \texttt{N\_JOBS}=$20$ ($> n_{\text{cores}}$) \\
         \hline
-        $1$ repetition    & $15 \;\text{s}$ & $26 \;\text{s}$ & $43 \;\text{s}$ \\
-        $10$ repetitions  & $87 \;\text{s}$ & $51 \;\text{s}$ & $76 \;\text{s}$ \\
-        $100$ repetitions & $749 \;\text{s}$ & $272 \;\text{s}$ & $308 \;\text{s}$ \\
-        $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1846 \;\text{s}$ \\
+        $N=1$ repetition    & \SI{15}{\second}
+        & \SI{26}{\second} ($0.57 \%$)  % 15/26
+        & \SI{43}{\second} ($0.34 \%$)  % 15/43
+        \\
+        $N=10$ repetitions  & \SI{87}{\second}
+        & \SI{51}{\second} ($1.70 \%$)  % 87/51
+        & \SI{76}{\second} ($1.14 \%$)  % 87/76
+        \\
+        $N=100$ repetitions & \SI{749}{\second}
+        & \SI{272}{\second} ($2.75 \%$)  % 749/272
+        & \SI{308}{\second} ($2.43 \%$)  % 749/308
+        \\
+        $N=500$ repetitions & \SI{2944}{\second}
+        & \SI{1530}{\second} ($1.92 \%$)  % 2944/1530
+        & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
+        \\
         % \hline
     \end{tabular}
-    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.}
+    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, on a machine with $n_{\text{cores}}=12$ CPU cores, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.
+The numbers in parentheses are the absolute gain in comparison with the non-parallel simulation (first column), in percentages.}
     \label{table:3:speedUpTimeParallelComputations}
@@
 \end{table}
+
+\begin{figure}[h!]  % [htbp]
+	\centering
+	% \includegraphics[width=0.75\linewidth]{logs/analyze_speedup_as_nb_of_jobs.pdf}
+	\includegraphics[width=0.75\linewidth]{logs/analyze_speedup_as_nb_of_jobs_normalize.pdf}
+	\caption{Absolute gain of the running time, for the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine.}
+	\label{fig:3:analyze_speedup_as_nb_of_jobs}
+\end{figure}
+
+% for N in 1 10 100 1000; do for NJ in 1 12 100; do echo "With N=$N repetitions and N_JOBS=$NJ jobs..."; time NOPLOTS=True SAVEALL=False BAYES=False ARM_TYPE=Bernoulli N=$N T=1000 K=9 N_JOBS=$NJ python3 main.py configuration.py 2>/dev/null | grep "real "; done; done && FreeSMS.py "DONE"
+
+% $N=1$ repetition    & \SI{15}{\second}
+% & \SI{26}{\second} ($0.57 \%$)  % 15/26
+% & \SI{43}{\second} ($0.34 \%$)  % 15/43
+% \\
+% $N=10$ repetitions  & \SI{87}{\second}
+% & \SI{51}{\second} ($1.70 \%$)  % 87/51
+% & \SI{76}{\second} ($1.14 \%$)  % 87/76
+% \\
+% $N=100$ repetitions & \SI{749}{\second}
+% & \SI{272}{\second} ($2.75 \%$)  % 749/272
+% & \SI{308}{\second} ($2.43 \%$)  % 749/308
+% \\
+% $N=500$ repetitions & \SI{2944}{\second}
+% & \SI{1530}{\second} ($1.92 \%$)  % 2944/1530
+% & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
 

### 2019-06-19 2-Chapters/3-Chapter/chapter3.tex
Plots for end of Chapter 3

@@
 
-\graphicspath{{2-Chapters/3-Chapter/Images/}}
-\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
+\graphicspath{{2-Chapters/3-Chapter/Images/}{2-Chapters/3-Chapter/logs/}{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
 
@@
 
-\textbf{References.}
+\paragraph{References.}
 %
@@
 
+% FIXME?
+\vfill{}
+% FIXME?
+
 \begin{figure}[h!]  % [htbp]
@@
 We show how to install SMPyBandits, and an example of how to run a simple experiment.
-This bash code\footnote{~See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}.
-You can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.},
-and install the requirements for Python 3 (once):
+See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.
+SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}, you can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.
+This bash code shows how to clone the code, and install the requirements for Python 3 (once):
 
@@
 
+Simulations are easily executed, \eg, Code~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPU.
+It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
+Environment variables (\texttt{N=1000} etc) in the command line are not required, but they are convenient.
+
 % https://tex.stackexchange.com/a/12430/
@@
 \end{small}
-
-Simulations are easily executed, \eg, Code~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPU.
-It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
-Environment variables (\texttt{N=1000} etc) in the command line are not required, but they are convenient.
 
@@
 $ for T in $(seq 5000 5000 50000); do \
-    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
-        N=1000  BAYES=True    T=$T         K=8 \
-        python3 main.py configuration.py \
-    && cp logs/main_py3_log.txt \
-          logs/main_py3_log__N1000_BAYES_T{T}_K8.txt \
+    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 N=1000 BAYES=True \
+    T=$T   K=8  python3 main.py configuration.py \
+    && cp logs/main_py3_log.txt logs/main_py3_log__N1000_BAYES_T{T}_K8.txt \
   done
 $ for K in $(seq 2 2 32); do \
-    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 \
-        N=1000  BAYES=True    T=5000       K=$K \
-        python3 main.py configuration.py \
-    && cp logs/main_py3_log.txt \
-          logs/main_py3_log__N1000_BAYES_T5000_K{K}.txt \
+    DEBUG=False NOPLOTS=False SAVEALL=True N_JOBS=-1 N=1000 BAYES=True \
+    T=5000 K=$K python3 main.py configuration.py \
+    && cp logs/main_py3_log.txt logs/main_py3_log__N1000_BAYES_T5000_K{K}.txt \
   done
@@
 
-\TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? Ce n'est pas crucial.
-    % C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
-}
+% \TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? Ce n'est pas crucial.
+%     % C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
+% }
 
@@
 $ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=1 \
-  python3 main.py configuration.py  # on only one core!
+  python3 main.py configuration.py  # on only one core
 
 $ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=-1 \
-  python3 main.py configuration.py  # parallel on all the cores!
+  python3 main.py configuration.py  # parallel on all the cores
 
 $ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=100 \
-  python3 main.py configuration.py  # parallel with too many jobs!
+  python3 main.py configuration.py  # parallel with too many jobs
         \end{minted}
@@
 Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits.
-We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $20$ jobs.
+We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $100$ jobs.
 %
@@
 %
-The results of Table~\ref{table:3:speedUpTimeParallelComputations} and of Figure~\ref{fig:3:analyze_speedup_as_nb_of_jobs}
-clearly illustrate that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
+The results of Table~\ref{table:3:speedUpTimeParallelComputations} clearly illustrate that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
 Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
 %
+Similar conclusions can be drawn from Figure~\ref{fig:3:analyze_speedup_as_nb_of_jobs}, which reports the results of the same experiment on a machine with $n_{\text{cores}}=12$ CPU cores.
 More details are included in the documentation page
@@
         $N=100$ repetitions & \SI{749}{\second}
-        & \SI{272}{\second} ($2.75 \%$)  % 749/272
+        & \SI{272}{\second} (\textcolor{darkgreen}{$2.75 \%$})  % 749/272
         & \SI{308}{\second} ($2.43 \%$)  % 749/308
@@
         $N=500$ repetitions & \SI{2944}{\second}
-        & \SI{1530}{\second} ($1.92 \%$)  % 2944/1530
+        & \SI{1530}{\second} (\textcolor{darkgreen}{$1.92 \%$})  % 2944/1530
         & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
@@
     \end{tabular}
-    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, on a machine with $n_{\text{cores}}=12$ CPU cores, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.
-The numbers in parentheses are the absolute gain in comparison with the non-parallel simulation (first column), in percentages.}
+    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, on a machine with $n_{\text{cores}}=4$ CPU cores, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.
+The percentages are the relative gains in comparison with the non-parallel simulation.}
     \label{table:3:speedUpTimeParallelComputations}
@@
 	\centering
-	% \includegraphics[width=0.75\linewidth]{logs/analyze_speedup_as_nb_of_jobs.pdf}
-	\includegraphics[width=0.75\linewidth]{logs/analyze_speedup_as_nb_of_jobs_normalize.pdf}
-	\caption{Absolute gain of the running time, for the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine.}
+	% \includegraphics[width=0.85\linewidth]{analyze_speedup_as_nb_of_jobs.pdf}
+	\includegraphics[width=0.95\linewidth]{analyze_speedup_as_nb_of_jobs_normalize.pdf}
+	\caption{Relative gain on the running time in comparison to using one job, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine running the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}.}
 	\label{fig:3:analyze_speedup_as_nb_of_jobs}

### 2019-06-19 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Starting to add the new figures
Little bit more math detail in Ch25

@@
 have never been used in practice for stochastic MAB problems.
+For instance, if we look at the applications of MAB for Opportunistic Spectrum Access, like it was proposed in \cite{Jouini09,Jouini10,Jouini12},
+online selection of MAB algorithms was never discussed,
+and usually the considered algorithm is simply $\UCB_1$, without discussion regarding this choice.
 % and we show that it appears empirically sub-optimal when applied to simple stochastic problems.
 
-We present an improved variant of \ExpQ, called \Aggr.
-For synthetic MAB problems, with Bernoulli or other distributions for $K$ arms, simulation results are presented to demonstrate its empirical efficiency.
-We combine classical algorithms, such as Thompson sampling, Upper-Confidence Bounds algorithms (\UCB{} and variants), and Bayesian or \klUCB.
-%
-Our algorithm offers good performance compared to state-of-the-art algorithms
-(\ExpQ{}, \CORRAL{} or \LearnExp{}) \cite{Agarwal16,Singla17},
-and appears as a robust approach to select on the run the best algorithm for any stochastic MAB problem, being more realistic to real-world radio settings than any tuning-based approach.
+% We present an improved variant of \ExpQ, called \Aggr.
+% For synthetic MAB problems, with Bernoulli or other distributions for $K$ arms, simulation results are presented to demonstrate its empirical efficiency.
+% We combine classical algorithms, such as Thompson sampling, Upper-Confidence Bounds algorithms (\UCB{} and variants), and Bayesian or \klUCB.
+% %
+% Our algorithm offers good performance compared to state-of-the-art algorithms
+% (\ExpQ{}, \CORRAL{} or \LearnExp{}) \cite{Agarwal16,Singla17},
+% and appears as a robust approach to select on the run the best algorithm for any stochastic MAB problem, being more realistic to real-world radio settings than any tuning-based approach.
 
@@
 %
-To simplify the presentation, we only aggregate bandit algorithms that give deterministic recommendations:
-one arm is chosen with probability $1$ and the others with probability $0$.
-However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $[K]$ at each time step, and draw the next selected arm according to this distribution
-(\eg, Thompson sampling).
+To simplify the presentation, we restrict in Algorithm~\ref{algo:25:Aggr} to bandit algorithms that give deterministic recommendations: one arm is chosen with probability $1$ and the others with probability $0$.
+%
+However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $q(t) = (q_k(t))_{k\in[K]}$ over the $K$ arms at each time step, and draw the next selected arm according to this distribution.
+For instance, Thompson sampling uses its posterior distribution, or \UCB{} can use a distribution with a mass of $1/n$ if there are $n\geq2$ arms of maximal index, in case of tie in the $\argmax$.
 
-The aggregation algorithm maintains a probability distribution $\pi^{t}$ on the $N$ algorithms $\Alg_a$, starting from a uniform distribution:
+The aggregation algorithm maintains a probability distribution $\pi^{t}$ on the $N$ algorithms $\Alg_a$ (\ie, $\pi^{t}\in\Delta_N$ the simplex of dimension $N$), starting from a uniform distribution:
 $\pi^t_a$ is the probability of trusting the decision made by algorithm $\Alg_a$ at time $t$.
-$\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it decides whom to trust by sampling $a \in [N]$ from $\pi^t$, then follows $\Alg_a$'s decision.
+$\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it first decides whom to trust by sampling $a \in [N]$ from $\pi^t$, then follows $\Alg_a$'s decision: $a \sim \pi^t, A_{\mathrm{aggr}}(t) = A(t) = A_a(t)$.
+We prove below that it is equivalent to first let all the algorithms decide their arm, \ie, $A_a(t)$, and then to compute
+$\forall k\in[K], p^{t}_k \eqdef \sum_{a = 1}^{N} \pi^{t}_a \times \indic(\{A_a(t) = k\})$, the weighted probability that one of the $N$ aggregated algorithms chose arm $k$, and finally to sample $A(t)\sim\pi^t$.
+
+\begin{proposition}\label{prop:25:equivalenceOfSamplingSchemes}
+\begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
+	The two following sampling schemes for the aggregated algorithm $\Alg_{\mathrm{aggr}} \left[\Alg_1, \dots, \Alg_N\right]$ are equivalent,
+	\ie, they give the same distribution on the action chosen by $\Alg_{\mathrm{aggr}}$, $A_{\mathrm{aggr}}(t) = A(t)$.
+	%
+	\begin{enumerate}
+		\item Sample $a(t)\sim \pi^t$ first, then trusts $\Alg_{a(t)}$'s decision: $A_{\mathrm{aggr}}(t) = A(t) = A_{a(t)}(t)$,
+		\item Let $A_a(t)$ be the arm chosen by algorithm $\Alg_a$ (sampled from $q_a^t$), for each $a\in[N]$, and compute $\forall k\in[K], p^{t}_k \eqdef \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(A_a(t) = k)$.
+Then, play $A_{\mathrm{aggr}}(t) = A(t) \sim p^t$.
+	\end{enumerate}
+	% \begin{align}\label{eq:25:equivalenceOfSamplingSchemes}
+	% 	 \\
+	% \end{align}
+\end{leftbar}  % XXX leftbar propositionbar, comment if needed
+\end{proposition}
+\begin{proof}
+	For any fixed time step $t\in[T]$,
+	we denote respectively $P(t)$ and $P'(t)$ the distribution of action of the aggregated algorithm $\Alg_{\mathrm{aggr}}$ at time $t$, respectively under the first and second sampling scheme.
+	Let $q_a^t\in\Delta_K$ be the distribution of the chosen action by algorithm $\Alg_a$  (for $a\in[N]$), that depends on the past observations (as well as some external randomness, see Section~\ref{par:2:externalrandomness}).
+	% \begin{enumerate}
+		% \item
+		The first sampling scheme gives
+		$P_k(t) = \bP(A_{\mathrm{aggr}}(t) = k) = \bP(\cupdot_{a=1}^N A_{\mathrm{aggr}}(t) = A_a(t) = k, a(t) = a) = \sum_{a=1}^N \bP(A_a(t) = k) \bP(a(t) = a) = \sum_{a=1}^N q_{a,k}(t) \pi_a^t$,
+		by independence of $a(t)\sim \pi^t$ and $A_i(t) \sim q_i(t)$ (for all $i$).
+		% \item
+		The second sampling scheme gives
+		$P'_k(t) = \bP(A_{\mathrm{aggr}}(t) = k) = \E[p_k^t]$ by definition,
+		and $\E[p_k^t] = \sum_{a = 1}^{N} \pi^{t}_a \E[\indic(A_a(t) = k)] = \sum_{a = 1}^{N} \pi^{t}_a \bP(A_a(t) = k) = \sum_{a=1}^N q_{a,k}(t) \pi_a^t$.
+	% \end{enumerate}
+	Thus we showed that $P_k(t) = P'_k(t)$, so the two sampling schemes are equivalent, as claimed.
+\end{proof}
+
 The main questions are then to know what observations (\ie, arms and rewards) should be given as feedback to which algorithms,
@@
 			% \KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
-			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$, \eg, $\eta_t=\frac{\log(N)}{t K}$}
+			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$, \eg, $\eta_t= \log(N) / (t K)$}
 			\KwData{Initial uniform distribution, $\pi^{0} = \cU([N])$}
@@
 				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
-				Let $p^{t}_j \eqdef \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(\{A_a(t) = j\}), \;\forall 1\leq j \leq K$\;
+				Let $p^{t}_k \eqdef \sum\limits_{a = 1}^{N} \pi^{t}_a \times \indic(A_a(t) = k), \;\forall k\in[K]$\;
 				% \tcp*{Sum trusts}
@@
 					% \eIf{$\Alg_a$ was trusted, \ie, $A_a(t) = A(t)$}{
-					\eIf{$\Alg_a$ was trusted, \ie, $A_a(t) = A(t)$}{
-						$ \pi^{t+1}_{a} = \exp(\eta_t r(t)) \times \pi^{t}_{a} $
+					\uIf{$\Alg_a$ was trusted, \ie, $A_a(t) = A(t)$}{
+						$ \pi^{t+1}_{a} = \exp(\eta_t \widehat{r}(t)) \times \pi^{t}_{a} $
 						\tcp*{Trusted more}
-					}{
+					}
+					\uElse{
 						% $ \pi^{t+1}_{a} = \exp(- \eta_t r(t) \times \pi^{t}_{a} $
@@
 needs to listen to the $N$ decisions to perform the updates on $\pi^{t}$.
-Then, \ExpQ{} gives back an observation (arm, reward) only to the last trusted algorithm
+Then, \ExpQ{} gives back an observation (\ie, a pair arm, reward) only to the last trusted algorithm
 whereas \Aggr{} gives it to all algorithms.
@@
 %
-Instead
-\footnote{~It is the same for constant learning rates, and does not differ much for decreasing learning rates.}
-of updating $\pi^{t}$ multiplicatively as we do for our proposal, \ExpQ{} recomputes it, proportionally to
+Instead of updating $\pi^{t}$ multiplicatively as we do for our proposal, \ExpQ{} recomputes it, proportionally to
 $\exp(\eta_t \widetilde{R}_a(t))$.
-%
+Note that there is no difference for the case of constant learning rates, and it does not differ much for decreasing learning rates as well.
+
 We also note that \ExpQ{} uses \emph{losses} instead of rewards, with $\ell(t) = 1 - r(t)$, but this is only a change of vocabulary.
@@
 If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t \eqdef \eta = 2 \log(N) / (T K)$.
-However, for real-world communication problems, it can be considered unrealistic\footnote{~For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}.} to assume a fixed and known time horizon, so we prefer the alternative horizon-free choice%
-\footnote{~It is non-increasing, and is obtained by minimizing the upper-bound on the regret derived in \cite[pp48]{Bubeck12}.}
-of learning rates,
+However, for real-world communication problems, it can be considered unrealistic to assume a fixed and known time horizon.
+For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}, that we shortly present in Appendix~\ref{app:2:DoublingTricks}.
+%
+Thus we prefer the alternative horizon-free choice of learning rates,
 $\eta_t \eqdef \log(N) / (t K)$ suggested by \cite{Bubeck12}.
-We compare both approaches empirically, and the second one usually performs better.
-We also stick to this choice of $(\eta_t)_{t \geq 1}$ for \Aggr.
+It is non-increasing, and is obtained by minimizing the upper-bound on the regret derived in \cite[pp48]{Bubeck12}.
+%
+We compared both approaches empirically, and the second one usually performs better.
+We also stick to this sequence $(\eta_t)_{t \geq 1}$ of decreasing learning rates for \Aggr.
 
@@
 as a function of $T$, comparing some algorithms $\Alg_1,\dots,\Alg_N$ (for $N=6$), and their aggregation with \Aggr{} (displayed in orange bold),
-using the parameter-free learning rate sequence, $\eta_t = \log{N} / (t K)$.
+using the parameter-free learning rate sequence, $\eta_t \eqdef \log{N} / (t K)$.
 
@@
 	\centering
-	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
+	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
 	\caption{On a ``simple'' Bernoulli problem (semilog-$y$ scale).}
@@
 
-\begin{figure}[b!]  % [htbp]
+\begin{figure}[h!]  % [htbp]
 	\centering
-	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
-	\caption{On a ``harder'' Bernoulli problem, they all have similar performances, except \LearnExp.}
+	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
+	\caption{On a ``harder'' Bernoulli problem, they all have similar performances, except \LearnExp, and our proposal \Aggr{} outperforms its competitors.}
 	\label{fig:25:HardBernoulli}
@@
 
-\begin{figure}[b!]  % [htbp]
+\begin{figure}[h!]  % [htbp]
 	\centering
-	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
+	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
 	\caption{On an ``easy'' Gaussian problem, only \Aggr{} shows reasonable performances, thanks to BayesUCB and Thompson sampling.}
@@
 	\centering
-	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
 	\caption{On a harder problem, mixing Bernoulli, Gaussian, Exponential arms, with 3 arms of each type with the \emph{same mean}.}
@@
 
-\begin{figure}[b!]  % [htbp]
+\begin{figure}[h!]  % [htbp]
 	\centering
-	\includegraphics[width=1.10\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
 	\caption{The semilog-$x$ scale clearly shows the logarithmic growth of the regret for the best algorithms and our proposal \Aggr, even in a hard ``mixed'' problem (\emph{cf}.
@@
 %
-And Figure~\ref{fig:25:HarderMixed} shows a considerably harder ``mixed'' problem, when the distributions are no longer in the same one-dimensional exponential family and so the Lai \& Robbins' lower-bound no longer holds (even if there still exists a lower-bound).
+And Figure~\ref{fig:25:HarderMixed} shows a considerably harder ``mixed'' problem, when the distributions are no longer in the same one-dimensional exponential family and so the Lai \& Robbins' lower-bound no longer holds.
 

### 2019-06-19 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Rien de neuf
Correcting some stuff in Ch4

@@
 and secondly does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
-In the studied settings,
-Stochastic MAB learning
+In the studied settings, stochastic MAB learning
 % provides an up to $16\%$ gain in terms of successful transmission probabilities, and
@@
 Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
-We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.
+We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic, more and more fluctuating and unpredictable.
 
@@
 %
-By considering a small number of wireless channels, and a unique PHY layer configuration\footnote{~But MAB learning to also optimize on the PHY layer parameters have started to be explored, after our work \cite{Bonnefoi17}, see for instance \cite{KerkoucheAlami18}.} (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
+By considering a small number of orthogonal wireless channels, and a unique PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
 the device can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use, that is, the most vacant one.
 Indeed, in this model, the \emph{quality} of the channels (\ie, the ressources) are identified by their vacancy, or one minus their occupancy rates.
+
+We note that MAB learning has also started to be applied to also optimize the PHY layer parameters, see for instance \cite{KerkoucheAlami18} following our article \cite{Bonnefoi17},
+but in this thesis we chose to restrict to the spectrum access problem and thus we only MAB models where the arms correspond to orthogonal frequency bands.
 
@@
 Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
-Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to \emph{have an average of active devices not larger\footnote{~In Chapter~\ref{chapter:5} we simply consider the case of $p=1$ and $D \leq K$ dynamic devices.} than the number of channels}.
+Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to \emph{have an average of active devices not larger\footnote{~In Chapter~\ref{chapter:5} we simply consider the case of $p=1$ and $M = D \leq K$ dynamic devices, referred to as \emph{players}.} than the number of channels}.
 
@@
 \paragraph{$1^{\text{st}}$ - Naive policy: Random Channel Selection} (``\textcolor{darkgreen}{Random}'')
-%
+
 We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in [K]$).
@@
 so probability of successful transmission is computed as follows
-%
+
 \begin{equation}
@@
 \paragraph{$2^{\text{nd}}$ - (Unachievable) Optimal oracle policy} (``\textcolor{orange}{Optimal}'')
-%
+
 We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the repartition of static devices $(S_k)_k$, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
@@
 \paragraph{$3^{\text{rd}}$ - A greedy approach of the oracle strategy} (``\textcolor{deeppurple}{Good sub-optimal}'')
-%
+
 We propose a \emph{sequential} approximation of the optimal policy:

### 2019-06-19 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Correcting some stuff in Ch4

@@
 %
-As far as we know, this is the first demonstration of running learning algorithms on the IoT device side, in real radio conditions, as highlighted it in our paper \cite{MoyBesson2019}.
+As far as we know, this is the first demonstration of running learning algorithms on the end-device side for Internet of Things networks, in real radio conditions, as highlighted it in our paper \cite{MoyBesson2019}.
 

### 2019-06-19 2-Chapters/4-Chapter/chapter4.tex
Correcting some stuff in Ch4

@@
 
-We already explained in Chapter~\ref{chapter:1} that
+After the first chapters that presented the model of MAB, we go back to the initial problems studied in this thesis, and thus we focus on Internet of Things (IoT) networks.
+%
+As explained in the introduction in Chapter~\ref{chapter:1},
 unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
@@
 As end-devices battery life is a key constraint of IoT networks,
-and as IoT networks are decentralized (\ie, the devices initiate transmissions),
+and as IoT networks are decentralized, because the devices initiate transmissions,
 this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
@@
 
-As detailed in Chapter~\ref{chapter:1},
-MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
-For example,
-proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
-and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
-Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
-Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
-Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
-% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
+% As detailed in Chapter~\ref{chapter:1},
+% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
+% For example,
+% proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
+% and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
+% Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
+% Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
+% Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
+% % However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
 
@@
 
+\newpage  % WARNING ?
 % ----------------------------------------------------------------------------

### 2019-06-19 2-Chapters/2-Chapter/chapter2.tex
details
Finishing to fix Emilie's comment of Ch2

@@
 
+\label{par:2:externalrandomness}
 This quantity $\sum_{t=1}^T r(t)$ can indeed be random, and it usually depends on two aspects.
@@
 In the rest of this thesis, \textbf{we only consider bounded rewards in the mathematical developments}, and we mostly focus on Bernoulli distributions, because they are usually the most relevant choice for the considered applications.
-First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
+% First, we restrict for simplicity to Bernoulli distributions for the simulations shown in Chapter~\ref{chapter:3}, even if our library does implement all the distributions commonly found in the literature (including unbounded one-dimensional exponential families and even Markov models).
 In Chapter~\ref{chapter:4}, we only study models with binary rewards.
 %
-Then, when we study multi-players bandits in Chapter~\ref{chapter:5}, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $Y_{k,t}=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense).
+Then, when we study multi-players bandits in Chapter~\ref{chapter:5}, we explain that restricting to the Bernoulli case is interesting and not restrictive, as it is actually the hardest case (since continuous distributions has a null mass on $Y_{k,t}=0$, they yield a much simpler problem as the sensing/no sensing distinction no longer makes sense), but this work can be extended to any one-dimensional exponential families.
 %
@@
 
-\paragraph{Other kind of hypothesis: sub-Gaussian and sub-Bernoulli distributions.}
-%
-A lot of research works considers rewards distributions that are not Gaussian but sub-Gaussian, of a known variance, for instance $1/4$.
+\paragraph{Sub-Gaussian and sub-Bernoulli distributions.}
+%
+A lot of research works considers rewards distributions that are not Gaussian but sub-Gaussian,
+% of a known variance, for instance $1/4$.
+meaning distributions whose moment generating function is dominated by that of a Gaussian distribution with the same mean (and a known variance).
 For instance, bounded distributions on $[0,1]$ are known to be $1/4$ sub-Gaussian, and this fact is used for instance in \cite{Maillard2018GLR}.
-It means that their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$.
 In Chapter~\ref{chapter:6}, we instead consider sub-Bernoulli distributions, formally introduced in Definition~\ref{def:6:subBernoulliDistributions}.
@@
 % Markov models: we do not present it in this thesis.
-Finally, we note that Markov models, while being implemented in SMPyBandits, are not used in this thesis.
-They were introduced in the 1980s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
+Finally, we note that Markov bandit models have also been considered for Cognitive Radio applications \cite{Liu08}.
+They are not studied in this thesis, even though we did implemented them in SMPyBandits,
+They were introduced in the $1980$s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
 A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and thus they are no longer stochastic nor stationary.
@@
 %
-Such models were less studied than stationary or adversarial models, but some interesting works focused on Markov models in the last 10 years.
-For instance, \cite{Melian15} proposed a cognitive radio model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
-A curious reader about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and then refer to the famous book \cite{Norris98}.
+Such models were less studied than stationary or adversarial models, but some interesting works applied Markov models or CR applications in the last 10 years
+For instance, \cite{Melian15} proposed a CR model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
+A curious reader about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and refer to \cite{Norris98}.
 
@@
 \textbf{Clinical trials} have been the first historical application of MAB models, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
-The mean of an arm, in this application, represents the mean success rate of a treatment, and the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
+The mean of an arm, in this application, represents the mean success rate of a treatment.
+Following the ``best arm identification'' model, the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
+And following our model of interest, maximizing the rewards corresponds to maximize the number of patients being successfully treated.
 %
@@
 %     \item
-    \textbf{A/B testing}, for instance for websites, is a popular application of the best arm identification problem,
-    where the task is purely exploratory and the player is asked to identify the best of two options (or more), in a finite number of steps \cite{audibert2010best}.
-    The problem can either consider a fixed budget and no freedom on the ending of the game (\ie, an known horizon $T$), or a fixed confidence and a certain freedom on the budget (\ie, the identified arm must be the true best arm with probability at least $1-\delta$) \cite{Garivier16BAI}.
-    The theoretical complexity of this problem
-    % use of bandit for A/B testing
-    was first studied in \cite{Kaufmann14},
-    and a more practical point-of-view was proposed in \cite{Jamieson17ABTest}.
+    % \textbf{A/B testing}, for instance for websites, is a popular application of the best arm identification problem,
+    % where the task is purely exploratory and the player is asked to identify the best of two options (or more), in a finite number of steps \cite{audibert2010best}.
+    % The problem can either consider a fixed budget and no freedom on the ending of the game (\ie, an known horizon $T$), or a fixed confidence and a certain freedom on the budget (\ie, the identified arm must be the true best arm with probability at least $1-\delta$) \cite{Garivier16BAI}.
+    % The theoretical complexity of this problem
+    % % use of bandit for A/B testing
+    % was first studied in \cite{Kaufmann14},
+    % and a more practical point-of-view was proposed in \cite{Jamieson17ABTest}.
 
@@
     Other possible use cases include hyper-parameter optimization, or features selection.
-    Hyper-parameters include real-valued parameters, like $\alpha\in[\frac{1}{2},\infty)$ for the $\UCB_1$ or $\varepsilon\in(0,1)$ for the $\varepsilon$-greedy bandit algorithms, a step size multiplier in a gradient method, or the width of a radial based function (RBF) kernel method.
+    Hyper-parameters include real-valued parameters,
+    % like $\alpha\in[\frac{1}{2},\infty)$ for the $\UCB_1$ or $\varepsilon\in(0,1)$ for the $\varepsilon$-greedy bandit algorithms,
+    like a step size multiplier $\gamma$ in a gradient descent,
+    the width $\rho$ of a Radial Based Function (RBF) kernel,
+    the margin $C$ in a Support Vector Machines (SVM), etc.
     Discrete-valued parameters are also common, like a choice in a fixed set of kernel functions or the depth of neural networks,
@@
 
-\textbf{Applications for Opportunistic Spectrum Access.}
+\textbf{Applications for Cognitive Radio.}
 % Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
@@
 % In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-In the model of Opportunistic Spectrum Access (OSA) with sensing, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
+In the model of Opportunistic Spectrum Access (OSA) with sensing \cite{Jouini09,Jouini10}, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
 A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
-%
-This model was first introduced by Wassim Jouini during his PhD thesis,
-in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
-Proof-of-concepts using real-world radio hardware was first proposed in \cite{MoyWSR2014,RobertSDR2014}.
+% %
+% This model was first introduced by W. Jouini during his PhD thesis,
+% in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
+% Proof-of-concepts using real-world radio hardware were first proposed in \cite{MoyWSR2014,RobertSDR2014}.
 %
@@
 As explained above, the main objective of a player facing a bandit game is to maximize its (expected) cumulated reward.
-An efficient algorithm should see its mean reward converge to the maximum reward.
-We introduce in this section the notion of \emph{regret}, and the equivalence for a player between maximizing its sum of rewards and minimizing its regret.
-In Machine Learning, we usually prefer to aim at minimizing certain quantities, such as the error rate in supervised learning, or the distance to the optimum in an optimization problem,
-but the main interest of studying the regret is to be able to obtain higher-order information about the convergence of the mean reward to the max reward,
-or in other words, to quantify the speed of convergence of a MAB algorithm.
+%
+In particular, an efficient algorithm should see its mean reward converge to the maximum reward.
+In the bandit literature, the performance of an algorithm is often measured in terms of regret, a performance measure that we now introduce.
+% We introduce in this section the notion of \emph{regret}, and the equivalence for a player between maximizing its sum of rewards and minimizing its regret.
+% In Machine Learning, we usually prefer to aim at minimizing certain quantities, such as the error rate in supervised learning, or the distance to the optimum in an optimization problem,
+% but the main interest of studying the regret is to be able to obtain higher-order information about the convergence of the mean reward to the max reward,
+% or in other words, to quantify the speed of convergence of a MAB algorithm.
 
@@
 %
-As the rewards from arm $k^*$ are \iid{} (as for all arms), and by linearity of the expectation, we can rewrite this expression to obtain the following definition of the regret:
-\begin{align*}
+As the rewards from arm $k^*$ are \iid{} (as for all arms), and by linearity of the expectation, we can rewrite this expression to obtain the following definition of the regret.
+% \begin{align*}
+%     R_T^{\cA}
+%     & = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
+%     = \E\left[ \sum_{t=1}^T \underbrace{Y_{k^*,t}}_{\text{\iid{} from $\nu_{k^*}$}} \right] - \E\left[ \sum_{t=1}^T r(t) \right] \\
+%     & = T \times \E\left[ Y_{k^*,1} \right] - \sum_{t=1}^T \E\left[ r(t) \right]
+%     = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right].
+% \end{align*}
+% % This is the definition we give below.
+
+\begin{definition}[Regret]\label{def:2:regret}
+\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
+    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\mu_1,\dots,\mu_K$, with $\mu^* \eqdef \max_k \mu_k$, then its \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ defined as
+    %
+    \begin{align}\label{eq:2:regret}
         R_T^{\cA}
@@
         = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right].
-\end{align*}
-% This is the definition we give below.
-
-\begin{definition}[Regret]\label{def:2:regret}
-\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-    For an algorithm $\cA$, a bandit problem of $K$ arms characterized by their means $\mu_1,\dots,\mu_K$, with $\mu^* \eqdef \max_k \mu_k$, then its \emph{(mean) regret} at horizon $T$ is $R_T^{\cA}$ defined as
-    %
-    \begin{align}\label{eq:2:regret}
-        R_T^{\cA}
-        \eqdef \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
-        % = T \mu^* - \sum_{t=1}^T \E\left[ Y(t) \right]
-        = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right].
     \end{align}
@@
 
-One first need to observe that $R_T^{\cA} \geq 0$ for any $T$ and $\cA$, and that $R_T^{\cA} \leq \mu^* T$ (if rewards lie in $[0,1]$).
+One first need to observe that $R_T^{\cA} \geq 0$ for any $T$ and $\cA$, and that $R_T^{\cA} \leq \mu^* T$, so $R_T^{\cA} \leqT$ if the rewards lie in $[0,1]$.
 Thus any algorithm has $R_T^{\cA} = \bigO{T}$, which justifies why we are interested in efficient algorithms that achieve at least a sub-linear regret, \ie, $R_T^{\cA} = \smallO{T}$.
@@
 %
-Remember that $N_k(t) \eqdef \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ denotes the number of times arm $k$ was selected between times $1$ and $t$,
+Recall that $N_k(t) \eqdef \sum_{s=1}^t \mathbbm{1}(A(s) = k)$ denotes the number of times arm $k$ was selected between times $1$ and $t$,
 %

### 2019-06-19 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Finishing to fix Emilie's comment of Ch2

@@
 
-\textbf{Static vs dynamic devices.}
+\paragraph{Static vs dynamic devices.}
 %
@@
 % , in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
-Contrarily to many previous work done in the CR field (\eg, Opportunistic Spectrum Access) \cite{Jouini10,Jouini12},
+Contrarily to many previous work done in the CR field and for Opportunistic Spectrum Access \cite{Jouini10,Jouini12},
 the reward $r(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.

### 2019-06-19 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Finishing to fix Emilie's comment of Ch2

@@
 We presented in Section~\ref{sec:4:firstModel} the impact of non-stationarity on the network performance using MAB algorithms is studied.
-Low-cost algorithms following two well-known approaches, such as the Upper-Confidence Bound (\UCB{}) \cite{Auer02,Auer02}, and the Thompson Sampling (TS) algorithms \cite{Thompson33} have reported encouraging results.
+Low-cost algorithms following two well-known approaches, such as the Upper-Confidence Bound (\UCB{}), and the Thompson Sampling (TS) algorithms have reported encouraging results.
 When considering the application of MAB algorithms for slotted wireless protocols in a decentralized manner,
-other recent directions of research first include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
-then realistic empirical PoC like in \cite{MoyWSR2014,RobertSDR2014,modiDemo2016,darak2016bayesian,kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
+other recent directions of research include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
+or realistic empirical PoC like in \cite{MoyWSR2014,RobertSDR2014,modiDemo2016,darak2016bayesian,kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
 and finally applications to multi-hoping networks \cite{Mitton16,Toldov16} or other kinds of networks \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
@@
 in case of RF (uplink or downlink) collision that prevent the device to receive the \emph{Ack} from the gateway,
-these devices will attempt to retransmit their packet up-to $\mathrm{MaxBackOff}$ times\footnote{~We prefer to denote it $\mathrm{MaxBackOff}$ instead of $M$ like we did in our paper \cite{Bonnefoi2019WCNC}, as $M$ is used in the next Chapter~\ref{chapter:5}.},
+these devices will attempt to retransmit their packet up-to $\mathrm{MaxBackOff}$ times\footnote{~We denote it $\mathrm{MaxBackOff}$ instead of $M$ like we did in our paper \cite{Bonnefoi2019WCNC}, as $M$ is used in next Chapter~\ref{chapter:5}.},
 with $\mathrm{MaxBackOff} \in\mathbb{N}^*$.

### 2019-06-19 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
New references. I need to finish Ch1 and Ch2
Finishing to fix Emilie's comment of Ch2

@@
 Among those, the \klUCB{} algorithm \cite{KLUCBJournal} has been shown to be asymptotically optimal for Bernoulli distributions (in that it exactly matches the lower bound given by \cite{LaiRobbins85}) and can also be employed when the rewards are assumed to be bounded in $[0,1]$.
-% \emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{Garivier18}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.
+% \emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{GarivierHadiji2018}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.
 
@@
 %
-\begin{equation}\label{def:6:nti}
+\begin{equation}\label{def:6:regret}
     R_T^{\cA} \eqdef \E\left[\sum_{t=1}^T \left(\mu_{k_t^*}(t) - \mu_{A(t)}(t)\right)\right].
@@
 
-In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T \eqdef \sum_{t=1}^{T-1} \indic\left(\exists k\in[K]: \mu_k(t) \neq \mu_k(t+1)\right)$.
+In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) \emph{number of breakpoints} $\Upsilon_T$, defined by
+\begin{equation}\label{def:6:breakpoints}
+    \Upsilon_T \eqdef \sum_{t=1}^{T-1} \indic\left(\exists k\in[K]: \mu_k(t) \neq \mu_k(t+1)\right).
+\end{equation}
 If we denote $\tau^{(0)} \eqdef 0$, we define the $i$-th breakpoint recursively by
@@
 %
-Hence for $t\in[\tau^{(i)} + 1,\tau^{(i+1)}]$, that is on any stationary segments, the rewards $(Y_{k,t})$ associated to each arm $k$ are \iid{} (hence the name of \emph{piece-wise stationary} problems).
+Hence for $t\in[\tau^{(i)} + 1,\tau^{(i+1)}]$, that is on any stationary segments, the rewards $(Y_{k,t})$ associated to each arm $k$ are \iid{} (hence the name \emph{piece-wise stationary}).
 
@@
 
-We define the \emph{number of change-points} on arm $k$,
+We define $\NCk$ the \emph{number of change-points} on arm $k$ by
 $\NCk \eqdef \sum_{t=1}^{T-1} \indic\left(\mu_k(t) \neq \mu_k(t+1)\right)$,
@@
 The following theorem provides an upper bound on the expected regret ofour algorithm
-They prove it achieves a regret bound of $\cO(\log(T) \sqrt{\Upsilon_T T})$, without a prior knowledge of $\Upsilon_T$.
+They prove that their algorithm achieves a regret bound of $\cO(\log(T) \sqrt{\Upsilon_T T})$, without a prior knowledge of $\Upsilon_T$.
 However, this work as two drawbacks:
@@
 The Generalized Likelihood Ratio statistic for this test is defined by
+
 \begin{equation}\label{eq:6:firstDefGLRT}
@@
     with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, the following regret bounds hold.
+
     \begin{enumerate}
@@
         \end{equation}
+
         \item Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``optimist'' guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
@@
         \end{equation}
+
         \item Choosing $\omega = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ (with prior knowledge of $\EffChange$) gives
@@
         \end{equation}
+
         \item Choosing $\omega = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``pessimist'' guess $\EffChange \simeq K \Upsilon_T$) gives

### 2019-06-19 2-Chapters/2-Chapter/chapter2.tex
Fixed according to Emilie's remarks

@@
 and thus we can write the following decomposition on the regret.
+%
+\emph{Its proof is simple and it is not a contribution, but we include it for readers that are unfamiliar with conditioning arguments}.
 
@@
 First, a \emph{problem-dependent lower-bound} states that any algorithm suffers a regret at least $\Omega(\log T)$ on any problem (from \cite{LaiRobbins85}).
-Then, a \emph{worst-case result} states that any algorithm can perform as badly as $\Omega(\sqrt{K T})$ on a certain problem instance, designed specifically to make it perform badly (a result from \cite{Auer02NonStochastic}),
+Then, a \emph{worst-case result} states that for any algorithm, there exists a certain problem instance such that the algorithm can perform as badly as $\Omega(\sqrt{K T})$ (a result from \cite{Auer02NonStochastic}),
 %
@@
 %
-Let $\cI$ denote the set of all problem instances, with $K \geq 2$ arms. We assume the rewards lie in $[0,1]$ (but no additional hypothesis).
-To specify the dependency on an instance $I$, we denote the regret of $\cA$ on instance $I$ and horizon $T$ by $R_T^{\cA}(I)$.
+Let $\cI$ denote the set of all problem instances, with $K \geq 2$ arms.
+We assume the rewards lie in $[0,1]$,
+to present the lower-bounds, we prefer to focus on Bernoulli distributions, for simplicity.
+
+To specify the dependency on an instance $I\in\cI$, we denote the regret of $\cA$ on instance $I$ and horizon $T$ by $R_T^{\cA}(I)$.
+We also index the Landau notations $o(\bullet)$ and $\bigO{\bullet}$ by $I$,
+for instance $R_T^{\cA}(I) = \smallO_I{\log(T)}$ means $R_T^{\cA}(I) = \smallO{c_I \log(T)}$
+where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
 
@@
 and a significant part of the research literature on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
-This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
+This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is asymptotically \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
 
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    \emph{No algorithm $\cA$ can achieve} a (mean) regret $R_T^{\cA}(I) = \smallO{c_I \log(T)}$ for all problem instances $I \in \cI$,
-    where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
+    \emph{No algorithm $\cA$ can achieve} a (mean) regret $R_T^{\cA}(I) = \smallO_I{\log(T)}$ for all Bernoulli problem instances $I \in \cI$.
+    \cite[Theorem~2.12]{Slivkins2019}
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
     An algorithm $\cA$ is \emph{uniformly efficient} if its (mean) regret satisfies
-    $R_T^{\cA}(I) = \smallO{C_{I,\alpha} T^{\alpha}}$,
-    for any value $\alpha>0$ and any problem instance $I\in \cI$,
-    where this notation means that the ``constant'' $C_{I,\alpha}$ can depend on the problem instance $I$ and on $\alpha$, but \emph{not} on the time horizon $T$.
+    $R_T^{\cA}(I) = \smallO_I{T^{\alpha}}$,
+    for any value $\alpha>0$ and any problem instance $I\in \cI$.
+    % where this notation means that the ``constant'' $C_{I,\alpha}$ can depend on the problem instance $I$ and on $\alpha$, but \emph{not} on the time horizon $T$.
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
 
-This family is non-empty as it contains for instance the $\UCB_1$ algorithm, since its regret is proven to be logarithmic on any problem instance with bounded rewards (in Theorem~\ref{thm:2:UCBregretBound} below).
-Now we can state the second logarithmic lower-bound, for algorithms in this family:
+This family is non-empty as it contains for instance the $\UCB_1$ algorithm, since its regret is proven to be logarithmic on any Bernoulli problem instance (in Theorem~\ref{thm:2:UCBregretBound} below).
+Now we can state the second logarithmic lower-bound, for algorithms in this family.
 
@@
     If $\cA$ is uniformly efficient,
-    then for any arbitrary problem instance $I$,
-    there exists a constant $C_I$ depending only on $I$,
-    and a time $T_0$ such that $\forall T \geq T_0, \;\;\; R_T^{\cA}(I) \geq C_I \log(T)$.
+    then for any arbitrary Bernoulli problem instance $I$,
+    the regret is asymptotically lower-bounded by $C_I \log(T)$
+    \cite[Theorem~2.13]{Slivkins2019},
+    or in other words
+    \[ \liminf_{T\to\infty} \frac{R_T^{\cA}(I)}{\log(T)} \geq C_I. \]
+    % there exists a constant $C_I$ depending only on $I$,
+    % and a time $T_0$ such that $\forall T \geq T_0, \;\;\; R_T^{\cA}(I) \geq C_I \log(T)$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
     Let $\cA$ denote a uniformly efficient algorithm,
-    and $I$ an arbitrary problem instance.
+    and $I$ an arbitrary Bernoulli problem instance.
+    \cite[Theorem~2.12]{Slivkins2019}
     \begin{itemize}
         \item
-        The bound from Theorem~\ref{thm:2:secondLogTLowerBound} holds with,
+        The bound from Theorem~\ref{thm:2:secondLogTLowerBound} holds with
         % \begin{equation}\label{eq:2:forSecondLogTLowerBound}
@@
         \item
-        For any $\varepsilon>0$, the bound from Theorem~\ref{thm:2:secondLogTLowerBound} also holds with,
+        For any $\varepsilon>0$, the bound from Theorem~\ref{thm:2:secondLogTLowerBound} also holds with
         % \begin{equation}\label{eq:2:forSecondLogTLowerBound2}
-        $C_I = \sum\limits_{k: \Delta_k > 0} \frac{\Delta_k}{\KL(\mu_k, \mu^*)} - \varepsilon$.
+        $C_I = \sum\limits_{k: \Delta_k > 0} \frac{\Delta_k}{\kl(\mu_k, \mu^*)} - \varepsilon$.
         % \end{equation}
@@
 Moreover, it is interesting to note that the second lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Proposition~\ref{prop:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
-should sample each sub-optimal arm $k$ about $\log(T)/\KL(\mu_k,\mu^*)$ times in the total of $T$ time steps.
+should sample each sub-optimal arm $k$ about $\log(T)/\kl(\mu_k,\mu^*)$ times in the total of $T$ time steps.
 
@@
 For a fixed horizon $T$, it is interesting to note that one can find instances $I$ that are so ``hard'' that a logarithmic regret (lower or upper) bound that uses a constant $C_I$ no longer brings any information.
-Indeed, we can naively bound the regret by $R_T^{\cA} \leq (\max_k \Delta_k) T$, and thus if $(\max_k \Delta_k)$ can be taken so small that $C_I \log(T) \gg (\max_k \Delta_k) T$, and thus a regret upper bound like $R_T^{\cA} \leq C_I \log(T)$ is useless.
-%
-\begin{smallproof}\label{proof:2:worstCaseLowerBound}
-    For an example, consider any large horizon $T>10$, and $K=2$ Bernoulli arms.
-    Chose any small $\varepsilon$ such that $\varepsilon \ll \sqrt{4 \log(T) / T}$, then chose let $\mu_1 = 1/2$ and $\mu_2 = 1/2 - \varepsilon$.
-    The constant $C_I$ from the first case of Theorem~\ref{thm:2:forSecondLogTLowerBound} is $C_I = 1 / (4 \varepsilon)$ and $\Delta=\varepsilon$ and it becomes so large that
-    the logarithmic upper-bound is useless in such case, as
-    $\varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T$.
-\end{smallproof}
+Indeed, we can naively bound the regret by $R_T^{\cA} \leq (\max_k \Delta_k) T$, and thus if $(\max_k \Delta_k)$ can be taken so small that $C_I \log(T) \gg (\max_k \Delta_k) T$, a regret upper bound like $R_T^{\cA} \leq C_I \log(T)$ is useless.
+%
+% \begin{smallproof}\label{proof:2:worstCaseLowerBound}
+%     For an example, consider any large horizon $T>10$, and $K=2$ Bernoulli arms.
+%     Chose any small $\varepsilon$ such that $\varepsilon \ll \sqrt{4 \log(T) / T}$, then chose let $\mu_1 = 1/2$ and $\mu_2 = 1/2 - \varepsilon$.
+%     The constant $C_I$ from the first case of Theorem~\ref{thm:2:forSecondLogTLowerBound} is $C_I = 1 / (4 \varepsilon)$ and $\Delta=\varepsilon$ and it becomes so large that
+%     the logarithmic upper-bound is useless in such case, as
+%     $\varepsilon \ll \sqrt{\frac{4 \log(T)}{T}} \Longleftrightarrow \frac{1}{4\varepsilon} \log(T) \gg \Delta T$.
+% \end{smallproof}
 

### 2019-06-19 2-Chapters/2-Chapter/chapter2.tex
Long proof reading!

@@
     as $\sum_{k=1}^K  \E[ N_k(T) ] = T$.
-    The sum can then be simplified to only account for sub-optimal arms, \ie, arms $k$ such that $\Delta_k > 0$.
+    The sum is then simplified to only count sub-optimal arms (\ie, $\Delta_k > 0$).
 \end{smallproof}
@@
 To specify the dependency on an instance $I\in\cI$, we denote the regret of $\cA$ on instance $I$ and horizon $T$ by $R_T^{\cA}(I)$.
-We also index the Landau notations $o(\bullet)$ and $\bigO{\bullet}$ by $I$,
-for instance $R_T^{\cA}(I) = \smallO_I{\log(T)}$ means $R_T^{\cA}(I) = \smallO{c_I \log(T)}$
-where this notation means that the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on the means and on $K$) but not on the time horizon $T$.
+We also index the Landau notations $o(\dots)$ and $\bigO{\dots}$ by $I$,
+for instance $R_T^{\cA}(I) = o_I(\log(T))$ means $R_T^{\cA}(I) = \smallO{c_I \log(T)}$
+where the ``constant'' $c_I$ can depend on the problem instance $I$ (\eg, on its means and on $K$) but \emph{not} on the time horizon $T$.
 
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    \emph{No algorithm $\cA$ can achieve} a (mean) regret $R_T^{\cA}(I) = \smallO_I{\log(T)}$ for all Bernoulli problem instances $I \in \cI$.
-    \cite[Theorem~2.12]{Slivkins2019}
+    \emph{No algorithm $\cA$ can achieve} a (mean) regret $R_T^{\cA}(I) = o_I(\log(T))$ for all Bernoulli problem instances $I \in \cI$.
+    \hfill{} \cite[Theorem~2.12]{Slivkins2019}
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
     then for any arbitrary Bernoulli problem instance $I$,
-    the regret is asymptotically lower-bounded by $C_I \log(T)$
-    \cite[Theorem~2.13]{Slivkins2019},
-    or in other words
+    its regret is asymptotically lower-bounded by $C_I \log(T)$,
+    or in other words,
+    \hfill{} \cite[Theorem~2.13]{Slivkins2019},
     \[ \liminf_{T\to\infty} \frac{R_T^{\cA}(I)}{\log(T)} \geq C_I. \]
@@
 
-This third theorem specifies possible values of the constants $C_I$ for the two previous results.
+We state below Theorem~2.12 from \cite{Slivkins2019}, to specify a possible value for the constant $C_I$.
 
@@
 \begin{leftbarnospace}[theorembar]  % XXX leftbar theorembar, comment if needed
-    Let $\cA$ denote a uniformly efficient algorithm,
+    If $\cA$ is uniformly efficient,
     and $I$ an arbitrary Bernoulli problem instance.
-    \cite[Theorem~2.12]{Slivkins2019}
     \begin{itemize}
@@
         % \begin{equation}\label{eq:2:forSecondLogTLowerBound}
-        $C_I = \mu^* (1 - \mu^*) \sum\limits_{k: \Delta_k > 0} \frac{1}{\Delta_k}$.
+            % $C_I = \mu^* (1 - \mu^*) \sum\limits_{k: \Delta_k > 0} \frac{1}{\Delta_k}$.
+            $C_I = \sum\limits_{k: \Delta_k > 0} \frac{\Delta_k}{\kl(\mu_k, \mu^*)}$.
         % \end{equation}
         \item
-        For any $\varepsilon>0$, the bound from Theorem~\ref{thm:2:secondLogTLowerBound} also holds with
+        For any $\varepsilon>0$, the bound also holds at finite time with
         % \begin{equation}\label{eq:2:forSecondLogTLowerBound2}
-        $C_I = \sum\limits_{k: \Delta_k > 0} \frac{\Delta_k}{\kl(\mu_k, \mu^*)} - \varepsilon$.
+            $C'_I = C_I - \varepsilon$,
         % \end{equation}
+        meaning that there exists a constant $C_I$ depending only on $I$,
+        and a time $T_0$ such that $\forall T \geq T_0, \;\;\; R_T^{\cA}(I) \geq C'_I \log(T)$.
     \end{itemize}
@@
 
-For this reason, the research literature has also been interested by another family of regret bounds, that are not problem-dependent but worst-case, or also called minimax \cite{Audibert2009minimax,audibert2010minimax} or problem-independent.
-Hence it is interesting to quote another lower-bound on the regret, that can bring useful information in such settings.
-The following theorem is from \cite{Auer02NonStochastic}.
+For this reason, another family of regret bounds is relevant, that are not problem-dependent but worst-case, or also called minimax \cite{Audibert2009minimax,audibert2010minimax} or problem-independent.
+We give an example of such result below.
+% Hence it is interesting to quote another lower-bound on the regret, that can bring useful information in such settings.
+The following theorem is from \cite{Auer02NonStochastic}, and \cite[Theorem~2.1]{Slivkins2019}.
 %
@@
     Fix the number of arms $K$, and an algorithm $\cA$.
-    Then for any horizon $T$, there exists a problem instance $I_T$ on which the algorithm suffers at least a regret $R_T^{\cA}(I_T) \geq \Omega(\sqrt{K T})$.
+    Then for any horizon $T$, there exists a Bernoulli problem instance $I_T$ on which the algorithm suffers at least a regret $R_T^{\cA}(I_T) \geq \Omega(\sqrt{K T})$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 
-The theoretical results obtained in this thesis are all expressed in terms of the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.
+The theoretical results obtained in the rest of this thesis are all expressed in terms of the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances that we consider in our works and in this manuscript.
 
@@
 
-\paragraph{Naive strategies: pure exploitation or pure exploration.}
+\paragraph{Pure exploitation or pure exploration.}
 
@@
 A large family of algorithms are index based, as they compute an index $U_k(t)$ on each arm $k$ at time $t$,
-used to play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
-If more than one indexes maximize $(U_k(t))_k$, the arm is chosen among the set, usually in a uniformly random manner: $A(t) \sim \cU(\argmax_k U_k(t))$.
+and they play the arm that maximizes their index, \ie, $A(t) = \argmax_k U_k(t)$.
+If more than one indexes maximize $(U_k(t))_k$, the arm is chosen from the set, usually in a uniformly random manner: $A(t) \sim \cU(\argmax_k U_k(t))$ (using the external randomness, as explained above).
 %
@@
     These implementations of $\KL(x,y)$ and \klUCB{} indexes are also published as an independent Python package, hosted on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} and installable from Pypi.
-    A Julia version of the same package is also available at \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}},
-    and installable with \texttt{Pkg.add("KullbackLeibler")}.
+    A Julia version of the same package is also available at \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}}.
+    % and installable with \texttt{Pkg.add("KullbackLeibler")}.
 },
@@
 
-We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the estimated regrets\footnote{~First, the rewards are computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{A(t)}(t)$, where $r^{(j)}_{k}(t)$ means the reward stream of the different runs $j=1,\dots,N$.
-Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t)$, as this first formula could give a negative regret and is more noisy, but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N \sum_{k=1}^K \mu_k N^{(j)}_{k}(T)$, as this later formula gives a more robust estimator, as noted the proof of Proposition~\ref{prop:2:RegretDecomposition} (in Page~\pageref{remark:2:moreAccurateCountofRegretForSimulations} above).}
-of the different algorithms, \emph{for one simulation}.
+We first give in Table~\ref{table:2:oneShotResult} the cumulated rewards as well as the estimated regrets of the different algorithms, \emph{for one simulation}.
+First, the rewards are computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{A(t)}(t)$, where $r^{(j)}_{k}(t)$ means the reward stream of the different runs $j=1,\dots,N$.
+Then the estimated regret is \emph{not} computed as $\frac{1}{N} \sum_{j=1}^N \sum_{t=1}^T r^{(j)}_{k^*}(t) - r^{(j)}_{A(t)}(t)$, as this first formula could give a negative regret and is more noisy, but as $\mu^* T - \frac{1}{N} \sum_{j=1}^N \sum_{k=1}^K \mu_k N^{(j)}_{k}(T)$, as this later formula gives a more robust estimator, as noted the proof of Proposition~\ref{prop:2:RegretDecomposition} (in Page~\pageref{remark:2:moreAccurateCountofRegretForSimulations} above).
 %
@@
 
-\begin{table}[ht]
+\begin{table}[!h]
     \begin{small}  % WARNING
@@
 \ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, \ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000} below the mean cumulated rewards and mean regrets as functions of the current time step $t$, for $T=10000$ (and $N=1000$).
-The reward of the best arm is also plotted in Figure~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, and it serves as an \emph{upper-bound} to illustrate convergence\footnote{~Conversely, we also add the Lai \& Robbins' logarithmic \emph{lower-bound} on regret ($C_I \log(t)$ from case 2 of Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plot in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
-It should not be surprising that the black line of the lower-bound is \emph{above} the values of the regret as the lower-bound is only asymptotic.}
+The reward of the best arm is also plotted in Figure~\ref{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}, and it serves as an \emph{upper-bound} to illustrate convergence
 of the average rewards
@@
 which should converge to $\mu^*$ when $t\to\infty$ for efficient algorithms.
+Conversely, we also add the Lai \& Robbins' logarithmic \emph{lower-bound} on regret ($C_I \log(t)$ from Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plot in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
+It should not be surprising that the black line of the lower-bound is \emph{above} the values of the regret as the lower-bound is only asymptotic.
 %
@@
 
-% \newpage  % WARNING ?
+\newpage  % WARNING ?
 

### 2019-06-19 2-Chapters/2-Chapter/chapter25.tex
Little bit more math detail in Ch25
Long proof reading!

@@
 
-\paragraph{Outline.}
+\textbf{Outline.}
+%
 This chapter is organized as follows.
-First, we give more motivations for online algorithm selection in the rest of this section, then we explain how to combine such stationary MAB algorithms and aggregate them.
+First, we give more motivations in the rest of this section, then we explain how to combine such stationary MAB algorithms and aggregate them.
 We present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},

### 2019-06-19 2-Chapters/3-Chapter/chapter3.tex
Long proof reading!

@@
 and even if some have already been implemented, a major future work is to implement the most interesting ones
-(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).\\
-%
-\indent
+(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
+
 Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models.
@@
 %
-We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.\\
-%
-\indent
+We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.
+
 As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
@@
 
-\textbf{Bash command to run these experiments.}
-We give below in Code~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments}.
+\paragraph{Bash command to run these experiments.}
+%
+We give below in Code~\ref{lst:3:BashCodeToLaunchLargeExperiments} a Bash command to run the experiments presented in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sub:3:additionalExperiments} above.
 It uses \texttt{for} loops and environment variables from Bash, and not Python, mainly for concision,
-but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for the different values of $T$ and $K$.
+but we could also have done the same by writing a Python script that calls the \texttt{main.py} script for these values of $T$ and $K$.
 
@@
 	% \includegraphics[width=0.85\linewidth]{analyze_speedup_as_nb_of_jobs.pdf}
-	\includegraphics[width=0.95\linewidth]{analyze_speedup_as_nb_of_jobs_normalize.pdf}
+	\includegraphics[width=1.10\linewidth]{analyze_speedup_as_nb_of_jobs_normalize.pdf}
 	\caption{Relative gain on the running time in comparison to using one job, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine running the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}.}

### 2019-06-19 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
C'est horrible je sais pas quoi dire ni comment
Finishing to fix Emilie's comment of Ch2
Long proof reading!

@@
 %
-\vspace*{-20pt}  % XXX
+% \vspace*{-20pt}  % XXX
 \begin{definition}\label{def:5:DecentralizedUniformEfficiency}
@@
 %   \item
-  The first instance is a small problem one with $K=3$ arms and means
+  The first instance is a small problem with $K=3$ arms and means
   $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for the case of $M=2$ players.
@@
   % \begin{subfigure}[!h]{0.49\textwidth}
-      \includegraphics[width=1.00\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_5016720151160452442.pdf}
+      \includegraphics[width=1.10\textwidth]{MP__K3_M2_T5000_N1000__4_algos/all_HistogramsRegret____env1-1_5016720151160452442.pdf}
   % \end{subfigure}

### 2019-06-19 2-Chapters/5-Chapter/chapter5.tex
details
Long proof reading!

@@
 
-\graphicspath{{2-Chapters/5-Chapter/Images/}}
-\graphicspath{{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/}}
+\graphicspath{{2-Chapters/5-Chapter/Images/}{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/}}
 
@@
 % \subsection{Inspirations from Cognitive Radio Proof-of-Concepts}
-\subsection{An idea from a Cognitive Radio Proof-of-Concept: also use learning for the rank of \RhoRand}
+% \subsection{An idea from a Cognitive Radio PoC: also use learning for the rank of \RhoRand}
+\subsection{Can MAB learning also be used to learn the rank for \RhoRand{} ?}
 \label{sub:5:USRPdemos}
@@
 
-\newpage  % FIXME
+% \newpage  % FIXME
+
 % ----------------------------------------------------------------------------

### 2019-06-19 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Long proof reading!

@@
 \begin{corollary}\label{cor:6:Global}
-\begin{leftbar}[corollarybar]  % XXX leftbar corollarybar, comment if needed
+% \begin{leftbar}[corollarybar]  % XXX leftbar corollarybar, comment if needed
     For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriodsGlobal},
@@
     \end{enumerate}
-\end{leftbar}  % XXX leftbar corollarybar, comment if needed
+% \end{leftbar}  % XXX leftbar corollarybar, comment if needed
 \end{corollary}
@@
 we highlight that \emph{this result is finite-time}.
+The proof of \GLRklUCB{} with \emph{local restarts} is not included, but it can be found in the Appendix of \cite{Besson2019GLRT}, as it is quite similar to the proof for \emph{global restarts} given above.
+The main difficulty relies in defining the ``good event'' $\cE_T$, and proving that it happens with high probability (by showing that the complementary event $\cE_T^c$ is highly unlikely).
 %
@@
 \begin{corollary}\label{cor:6:Local}
-\begin{leftbar}[corollarybar]  % XXX leftbar corollarybar, comment if needed
+% \begin{leftbar}[corollarybar]  % XXX leftbar corollarybar, comment if needed
     For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriods},
@@
     \end{enumerate}
-\end{leftbar}  % XXX leftbar corollarybar, comment if needed
+% \end{leftbar}  % XXX leftbar corollarybar, comment if needed
 \end{corollary}
-
-The proof of \GLRklUCB{} with \emph{local restarts} is not included, but it can be found in the Appendix of \cite{Besson2019GLRT}, as it is quite similar to the proof for \emph{global restarts} given above.
-The main difficulty relies in defining the ``good event'' $\cE_T$, and proving that it happens with high probability (by showing that the complementary event $\cE_T^c$ is highly unlikely).
 

### 2019-06-19 2-Chapters/1-Chapter/chapter1.tex
details
Printing to proof read abstract/intro/conclusion

@@
 
-\TODOL{Mardi soir : je dois encore relire le 1.1 pour les éventuelles erreurs historiques !}
-\TODOL{Mardi soir : je dois encore améliorer le 1.2 et 1.3 ?}
+\TODOL{Mardi soir : je dois encore tout relire et améliorer un peu !}
 
@@
 % \subsection{Focusing on MAB algorithms}
-\subsection{Part~\ref{part:Introduction}}
+\subsection{Exploring the literature of MAB algorithms, Part~\ref{part:Introduction}}
 
@@
 % \subsection{Focusing on IoT networks.}
-\subsection{Part~\ref{part:MABIOT}}
+\subsection{Our model of IoT network and decentralized MAB, Part~\ref{part:MABIOT}}
 
@@
 
-% \newpage % WARNING
+\newpage % WARNING
 
@@
 % WARNING
-% \newpage
+\newpage
 

### 2019-06-19 2-Chapters/2-Chapter/chapter2.tex
Printing to proof read abstract/intro/conclusion

@@
 \graphicspath{{2-Chapters/2-Chapter/Images/}}
-
-\TODOL{Intégrer les derniers commentaires d'Emilie, entre la fin des détails sur les familles exponentielles (avec les deux exemples), et quand je commence en Section~\ref{sub:2:otherMeasuresPerformance} à parler des autres mesures !}
 
@@
     as $\sum_{k=1}^K  \E[ N_k(T) ] = T$.
-    The sum is then simplified to only count sub-optimal arms (\ie, $\Delta_k > 0$).
+    The sum is then simplified to only count sub-optimal arms.
 \end{smallproof}

### 2019-06-20 1-Preamble/5-Abstract/abstract.tex
Probably enough for today? ALmost done for everything in Ch1 but the details of our contributions
Huge proof reading of Ch1 & Conclusion

@@
 % Chapitre 2
-Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB), mais leur analyse est difficile à réaliser,
+Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (MAB), mais leur analyse est difficile à réaliser,
 % , Chapitre~\ref{chapitre:2}
-car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication ni aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+car il est délicat de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication ni aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
 % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
@@
 % Chapter 3
-We also detail another contribution, SMPyBandits, an open-source Python library for numerical MAB simulations, covering the studied models and more.
+We also detail another contribution, SMPyBandits, an open-source Python library for numerical MAB simulations, covering all the studied models and more.
 %  (Chapter~\ref{chapter:3})

### 2019-06-20 2-Chapters/1-Chapter/chapter1.tex
Huge proof reading of Ch1 & Conclusion

@@
 %
-This doctoral thesis started in October $2016$, and finished in October $2019$.
+This manuscript concludes my doctoral thesis, which started in October $2016$ and finished in October $2019$.
 My research took place at the IETR laboratory in Rennes (France), in the SCEE team hosted on the Rennes campus of the engineering college CentraleSupélec.
@@
 %
-At the root of the problems that motivates this thesis are the questions of global warming, and the increase of the world population.
+At the root of the problems that motivate this thesis are the two questions of global warming, and the increase of the world population.
 % https://en.wikipedia.org/wiki/History_of_radio
-During the last 150 years, mankind has developed a wide range of different communication technologies, and since the late $1890$s wireless communications between devices have been made possible, and more and more frequent in our lives.
+During the last 150 years, mankind has developed a wide range of different communication technologies, and since the late $1890$s wireless communications between manufactured devices have been made possible, and more and more frequent in our lives.
 With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, for a large range of different applications.
@@
 
-That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms (MAB algorithms),
+That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms (Multi-Armed Bandit algorithms),
 directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing the wireless communication thanks to embedded decentralized decision making.
@@
 %
-Historically, three families of wireless communication systems have been deployed: first, centralized broadcasting (\eg, old TV), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things networks (IoT).
+Historically, three families of wireless communication systems have been deployed: first, centralized broadcasting (\eg, old TV), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things (IoT) networks.
 
@@
 % The opposite: decentralized broadcasting, and the age of IoT
-Finally, a third kind of systems are of the opposite kind, and can be designed as decentralized broadcasting: an antenna is still in charge of many devices, but such devices can only send up-link packets, and the only down-link data they can receive are short acknowledgements sent from the base-station to indicate success or failure of an up-link packet.
+Finally, a third kind of systems are of the opposite kind, and can be designed as decentralized broadcasting: an antenna is still in charge of many devices, but such devices can only send up-link packets, and the only down-link data they can receive are short acknowledgements sent from the base-station to indicate success or failure of every up-link packet.
 This family of wireless systems are referred to as Internet of Things,
@@
 For buildings, the main goal is to reduce the cost of heating empty buildings and use temperature sensor networks to get accurate and regular data about the temperature in each room and each floor, and let the centralized heat control optimize its cost and energy consumption.
-For agriculture, one example can be to equipped every cow (in large farms) with a sensors that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and monitors the health of the animals.\\
+For agriculture, one example can be to equipped every cow (in large farms) with a sensor that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and to monitor the health of the animals.\\
 %
@@
 In the present and future IoT networks, many devices of heterogeneous nature are using the same antenna for different applications.
-A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and will use a battery, which duration should be maximized (typically more than 10 years).
-Another common constraint for IoT devices is their low duty cycle, as most applications target one to a few messages sent by day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
+A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and run on a battery, which duration should be maximized (typically more than 10 years).
+Another common constraint for IoT devices is their low duty cycle, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
 %
@@
 and they consist in a specification for both the \emph{PHY}sical
-and the Medium Access Control (\emph{MAC}) layer.
+and the \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}) layer.
 To quote some examples of standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications, while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications.
@@
 and especially on one subfield that we detail below, \textbf{Dynamic Spectrum Access (DSA)}.
-
+%
+%
 \begin{figure}[h!]
@@
 \end{figure}
-
+%
+%
 \TODOL{Je peux améliorer (ou enlever) ce passage qui site J.Toledano, ou alors le commenter en plus de détails ? J'aime bien son rapport en fait (j'aurai du le lire au début de ma thèse !)}
@@
 %
-In OSA, the focus is on one SU accessing a licensed spectrum.
+In OSA, the focus is on one SU accessing a licensed spectrum, occupied by PU that have a strict priority over the SU, which has to follow a ``listen-before-talk'' access scheme.
 %
@@
 If the different channels are not uniformly used by the PU, and if we assume a stationary hypothesis on the PU traffic, then the goal of the SU boils down to explore the different channels and exploiting the best ones.
-This frames the Opportunistic Spectrum Access problem as an exploration/exploitation problem, white a finite set of actions (the channels, also called \emph{arms}),
+This frames the Opportunistic Spectrum Access problem as an exploration/exploitation problem, with a finite set of actions (the channels, also called \emph{arms}),
 in a sequential action-then-feedback cycle (time steps are $t\in\N^*$),
@@
 Since $2017$, S.
-Darak and his team at IIIT Delhi have been very active in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+Darak and his team at IIIT Delhi have actively worked in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
@@
 The aforementioned previous works have shown that MAB algorithms can be applied with success for the OSA problem.
-But if we consider CR-ready devices that cannot perform sensing, such as low-cost and low energy consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
+But if we consider CR-ready devices that cannot perform sensing, such as low-cost and low energy-consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
 Moreover, in most cases, the IoT networks use unlicensed bands, and as such there are no longer a distinction between PU and SU.
@@
 %
-If we sum-up the studied problems and summarize them in one question, it would be the following:
+If we sum-up the studied problems and summarize them in one question, it could be the following:
 \emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
@@
 %
-In order to clearly understand which algorithms could be applied to the aforementioned problems of CR and IoT,
+In order to clearly understand which algorithms could be adapted to the aforementioned constraints of CR for IoT networks,
 we were not only interested by the usual measure of performance of any MAB algorithm (its regret, see below in Section~\ref{sec:2:lowerUpperBoundsRegret}),
@@
 Our exploration of the large number of MAB algorithms and models developed in the recent literature
-has gave us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explores the performances of newly proposed algorithms.
+has given us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explores the performances of newly proposed algorithms.
 To answer this goal, we produced a Python library of simulation of MAB problems.
@@
 To answer this question, we present two approaches.
-The first one is an empirical comparison of the most efficient and well-known existing algorithms, in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}.
-We confirm that well-known and not too sophisticated algorithms, such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
+The first one is an empirical comparison of the most efficient and well known existing algorithms, in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}.
+We confirm that widely used and not too sophisticated algorithms, such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
 The second approach is an online algorithm selection, consisting in aggregating a (finite) set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
@@
 %
-In Chapter~\ref{chapter:4}, we study two models of IoT networks.
-The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use the acknowledgement as feedback, in order to learn the best channel for their uplink communications.
+We study two models of IoT networks in Chapter~\ref{chapter:4}.
+The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use the acknowledgement as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the less occupied by the surrounding traffic).
 The goal of this first model of applying decentralized RL on the device side is to increase the Quality of Service (QoS) of the application of such IoT network, by reducing the rate of failed transmission.
@@
 %
-The second model we study then considers packet retransmissions, and while the concept stays similar, the application of an efficient decentralized learning results in both an increased battery life for each of the devices, as well as an increase QoS for the complete network.
+The second model we propose then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increase QoS for the entire network.
 
@@
 Static devices form an interfering traffic, which could be generated by devices using other standards as well.
-Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choices of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) from the base station.
+Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choices of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) received from the base station.
 %
 We first detail three base-lines, a naive approach and two centralized ``oracle'' approaches,
-that are used to evaluate the performance of MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communications rate.
+that are used to evaluate the performance of the considered MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communications rate.
 We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from the other devices becomes more and more fluctuating and unpredictable.
@@
 
-Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations is a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
+Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations was a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
 In Section~\ref{sec:4:firstModel}, we detail what we believe to be the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network.
@@
 We then consider an extension of our first model, following the idea of packet retransmissions, that is at the root of the famous ALOHA protocol \cite{Abramson1970,Roberts75}.
-The idea is that when a device sends an uplink message but does not receive its \Ack{} before a certain delay, instead of going back to sleep mode and waiting for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, $m \leq 10$ times), after some small random waiting time.
+The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode and waiting for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, $m \leq 10$ times), after some small random waiting times.
 %
-This second model proposes to also use decision making at the second layer of the MAC protocol, that is to also use RL and MAB algorithms, in order to learn the best channel to use for retransmitting any packet that suffered from a collision at its first transmission.
+We propose to also use decision making, that is to also use RL and MAB algorithms, for this other model, at the second layer of the MAC protocol, in order to learn the best channels to use for retransmitting the packets that suffered from a collision at their first transmissions.
 We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations.
-This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well as the simplest of the different heuristics we compared is efficient and allow the devices to learn to behave almost optimally.
+This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well that the simplest of the different heuristics we compared is efficient, and allows the devices to learn to behave almost optimally.
 This third contribution concluded our collaboration with R. Bonnefoi, and was presented at the $1^{\text{st}}$ MoTION workshop, during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
@@
 %
-In the two models presented above, the core concept is to let all the dynamic devices of a IoT network run a fully decentralized learning algorithm in order to optimize the complete system,
-meaning that each of them target their own local objective, which is to maximize their cumulated reward (in a selfish way).
+In the two models presented above, the core concept is to let every dynamic device of a IoT network runs a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
+It means that each of them target its own local objective, which is to maximize its cumulated reward (in a selfish way).
 It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma} or other popular ``dilemma'' games.
@@
 % https://www.ii.uni.wroc.pl/~lukstafi/pmwiki/uploads/AGT/Prediction_Learning_and_Games.pdf
-So it is quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices.
+So it was quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices, despite the fact that these IoT objects cannot communicate directly with each other and receive only a collision feedback from the base station they are associated to.
 
 We first tried to analyze the performance of this decentralized algorithm, that we refer to as \Selfish, in the model of Section~\ref{sec:4:firstModel},
-but due to the random number of active devices at each time step (as soon as the probability $p$ of activation is $p < 1$), we were unable to propose a clean analysis.
+but due to the random number of active devices at each time step (\ie, as soon as the probability $p$ of activation is $p < 1$), we were unable to develop a clean analysis.
 %
@@
 The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing information) or the IoT case (\ie, without sensing feedback).
-The goal was to understand the heuristic of Chapter~\ref{chapter:4}, \Selfish, in the simpler framework of multi-player MAB, which has been studied before \cite{Zhao10,Anandkumar10,Anandkumar11} for the case of OSA.
+The goal was to understand the heuristic of Chapter~\ref{chapter:4}, \Selfish, in the simpler framework of multi-player MAB, which has been studied before for the case of OSA, as studied for instance by \cite{Zhao10,Anandkumar10,Anandkumar11}.
 %
@@
 On the first hand, in the OSA setting, we failed to obtain positive result for the Selfish policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
-This result was later confirmed and generalized by \cite{BoursierPerchet18}.
+This result was later confirmed and analyzed by \cite{BoursierPerchet18}.
 %
@@
 %
-Unfortunately this strong theoretical result heavily depends on the presence of sensing feedback, and as such our results are not (yet) applicable to the IoT model without sensing.
+This strong theoretical result heavily depends on the presence of sensing feedback, and as such our results are not (yet) applicable to the IoT model without sensing.
 
@@
 %
-and its publication relaunched the research on this model, as a few independent teams started to work on different questions after our presentation.
+and it relaunched the research on this model, as a few independent teams started to work on different questions shortly after its publication.
 %
@@
 \indent
-On the first hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let the devices learn an optimal orthogonal affectation of $M \leq K$ devices in the $K$ channels.
-However, it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid any central control of the devices by the base station, as we explained above.
+On the first hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let all groups of $M \leq K$ devices learn an optimal orthogonal affectation in the $K$ channels (\eg, groups can be the set of devices that all emit at the same time).
+%
+However, even if we continue to assume a synchronized time in all the rest of the thesis,
+it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid any central control of the devices by the base station, as we explained above.
 %
 On the other hand, another way to relax the non-stationary hypothesis is to take the point-of-view of one device, like in the OSA case mentioned above, where the focus is on one SU surrounded by many PU having stationary behaviors.
-If we focus on one IoT device, we could assume that the environment (\ie, the surrounding devices) is stationary on some time intervals, which leads to the last contribution and chapter of this thesis.
+If we focus on one IoT device, its environment (\ie, the surrounding devices) is non-stationary, meaning that its average properties can fluctuate with time, but we could assume a certain structure on this non-stationarity.
+For instance, enforcing stationarity on some time intervals leads to the last contribution and chapter of this thesis.
 % DONE explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
@@
 Assuming the environment to be piece-wise stationary indeed makes sense for applications to wireless networks, where a change-point can for instance correspond to the arrival of a new group of end-devices in the network.
-Examples of such situation can be a new company arrives on the market, or a farmer installs sensors on a herd of cows, etc.
+Examples of such situation can be a new company arriving on the market in one city (\eg, the Linky counters being set-up nowadays in France), or the neighbor farmer installing sensors on his own herd of cows, etc.
 %
 Two main families of algorithms have been proposed for the piece-wise stationary problem,
-that usually consist in combining an efficient policy designed for the stationary MAB problem (\eg, \UCB{} or \klUCB) and a way to adapt to changes in the arms distributions.
+that usually consist in combining an efficient policy designed for the stationary MAB problem (\eg, Thompson Sampling or \klUCB) and a way to adapt to changes in the arms distributions.
 Passively adaptive policies use a window of a fixed or evolving size \cite{Garivier11UCBDiscount}, or a discount factor, in order to forget about the past observations \cite{Kocsis06,Gupta11thompson},
@@
 We build on very recent results on the Generalized Likelihood Ratio test (GLRT) for Gaussian and sub-Gaussian variables \cite{Maillard2018GLR}, and we instead focus on bounded rewards and Bernoulli distributions.
+Bounded rewards are indeed usually more appropriate for Cognitive Radio applications.
 Using the fact that bounded variables in $[0,1]$ are not only $1/4$ sub-Gaussian but also sub-Bernoulli, we prove the first finite-time guarantees for the GLRT for bounded variables.
@@
 This last chapter is based on our last work, since October $2018$, and it first lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
-Our work also lead to the article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results and submitted soon to a journal (probably the Journal of Machine Learning Research) in fall $2019$.
+Our work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results and submitted soon to a journal (probably the Journal of Machine Learning Research) in fall $2019$.
 
@@
 We conclude this thesis in Chapter~\ref{chapter:conclusion}, to give some perspectives and directions about possible future works.
-Most of the contributions that we present in this thesis call for more developments, on both directions of providing more mathematical analysis and implementing new proof-of-concepts using real wireless hardware and realistic Internet of Things networks.
-%
+Most of the contributions that we present in this thesis call for more developments, on both directions of providing more mathematical analysis and implementing new proof-of-concepts, using real wireless hardware and realistic Internet of Things networks.
+
 This manuscript ends with the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks} we discuss about our work on doubling-trick, in order to remove the need for an algorithm to know the horizon of the bandit game before starting to play it.
@@
     \item
-    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, published under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
     We present it in details in Chapter~\ref{chapter:3}.
@@
     We propose different models for IoT networks, in Chapter~\ref{chapter:4}, where end-devices with cognitive radio capabilities can implement MAB algorithms on their side, to automatically increase their battery life and allow more devices to use the same network while maintaining a high Quality of Service
-    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC}.
+    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019}.
     % (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
@@
     \item
-    We formalize the multi-players bandit model, and we introduced three variants, in Chapter~\ref{chapter:5}.
+    We present the multi-players bandit model, for which we introduce three variants, in Chapter~\ref{chapter:5}.
     For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
-    as well as extensive numerical experiments to demonstrate its good performance in comparison with the rest of the literature.
-    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works built up on our results.
+    as well as extensive numerical experiments to demonstrate its excellent performance in comparison with the rest of the literature.
+    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works heavily built up on our results.
 
@@
     % \item
-\textcolor{darkred}{First, in Part~\ref{part:Introduction}}, we start by the next Chapter~\ref{chapter:2}, required for the rest of the document as we introduce the MAB models, the concepts and the notations used in this thesis.
+\textcolor{darkred}{First, in Part~\ref{part:Introduction}} (second line), we start by the next Chapter~\ref{chapter:2}, required for the rest of the document, as we introduce the MAB models and the notations used in this thesis.
 Conversely, even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use numerical simulations based on our simulation library SMPyBandits, the Chapter~\ref{chapter:3} where we present it is not required to understand them.
@@
     % \item
-\textcolor{darkgreen}{Then, the second Part~\ref{part:MABIOT}} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
+\textcolor{darkgreen}{Then, the second Part~\ref{part:MABIOT}} contain three chapters (third line), that are included in both the logical and chronological orders, but can be read almost independently.
 Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where we show that MAB algorithms can be used with success.
@@
 
-    \emph{Note that we are already working on an extended version that will be submitted to a journal on Machine Learning for Wireless Communications, in July $2019$.}
+    \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning for Wireless Communications, in July $2019$.}
 
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    February $2019$.
+    February $2019$.\\
     See Chapter~\ref{chapter:6}.
@@
 
-    \emph{Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
+    \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
 
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    September $2018$.
+    September $2018$.\\
     See Appendix~\ref{app:2:DoublingTricks}.

### 2019-06-20 2-Chapters/2-Chapter/chapter2.tex
Huge proof reading of Ch1 & Conclusion

@@
 
-One first need to observe that $R_T^{\cA} \geq 0$ for any $T$ and $\cA$, and that $R_T^{\cA} \leq \mu^* T$, so $R_T^{\cA} \leqT$ if the rewards lie in $[0,1]$.
+One first need to observe that $R_T^{\cA} \geq 0$ for any $T$ and $\cA$, and that $R_T^{\cA} \leq \mu^* T$, so $R_T^{\cA} \leq T$ if the rewards lie in $[0,1]$.
 Thus any algorithm has $R_T^{\cA} = \bigO{T}$, which justifies why we are interested in efficient algorithms that achieve at least a sub-linear regret, \ie, $R_T^{\cA} = \smallO{T}$.
@@
     An algorithm $\cA$ is \emph{uniformly efficient} if its (mean) regret satisfies
-    $R_T^{\cA}(I) = \smallO_I{T^{\alpha}}$,
+    $R_T^{\cA}(I) = o_I(T^{\alpha})$,
     for any value $\alpha>0$ and any problem instance $I\in \cI$.

### 2019-06-20 2-Chapters/Conclusion/conclusion.tex
Huge proof reading of Ch1 & Conclusion

@@
 
-% \newpage
-\abstractStartChapter{}%
-%
-This last chapter concludes the document, first by summarizing the previous chapters, and by giving a general conclusion of the thesis.
-%
-We highlight what we consider to be the most important directions of future works, as well as some promising directions that were shortly studied during my PhD but not discussed in the manuscript.
+% % \newpage
+% \abstractStartChapter{}%
+% %
+% This last chapter concludes the document, first by summarizing the previous chapters, and by giving a general conclusion of the thesis.
+% %
+% We highlight what we consider to be the most important directions of future works, as well as some promising directions that were shortly studied during my PhD but not discussed in the manuscript.
 
-\minitocStartChapter{}
+% \minitocStartChapter{}
 

### 2019-06-20 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done, I should regenerate the cover/back pages in MS Word...

@@
 
-\textbf{Successful transmissions rate as a function of the number of dynamic devices.}
+\textbf{Successful transmission rate as a function of the number of dynamic devices.}
 %
-To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure~\ref{fig:41:perf_learning} the evolution of the gain, in terms of successful transmissions rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each IoT device.
+To better assess the evolution of the optimal policy compared to the random one, we have displayed on Figure~\ref{fig:41:perf_learning} the evolution of the gain, in terms of successful transmission rate, provided by the optimal oracle and the two learning policies, after $10^6$ time slots, \ie, about $1000$ transmissions for each IoT device.
 We can see that when the proportion of end-devices is low (\eg, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.

### 2019-06-20 2-Chapters/Conclusion/conclusion.tex
Done, I should regenerate the cover/back pages in MS Word...

@@
 % Chapter 3
-Then in Chapter~\ref{chapter:3} we discussed in details about our open-source Python library, developed to run numerical simulations of MAB problems, SMPyBandits.
+Then in Chapter~\ref{chapter:3} we discussed in details about our open-source Python library, SMPyBandits, developed to run numerical simulations of MAB problems.
 %
@@
 In Chapter~\ref{chapter:4}, we first proposed some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
-In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithm.
+In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithms.
 We also presented a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.
 % Chapter 5
-Because the models studied for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in a noisy environment,
+Because the models studied for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
 in Chapter~\ref{chapter:5}, we studied a weaker but still interesting model.
@@
 %
-For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations, and on the regret bounds obtained in our theoretical analyses.
+For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
 \end{itemize}
 
-A more precise summary of each chapter is included below.
+A more precise summary of each chapter is given below.
 
@@
 We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
-Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions.
+Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions (as well as rested or restless Markov models).
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
@@
 \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:25}}.}
-
-The first contribution of this manuscript \cite{Besson2018WCNC} concluded this part.
+%
+This first part ends by presenting one of our first contributions \cite{Besson2018WCNC}.
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
-Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for the practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for a practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
 We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
@@
 
-First in Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
-with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
+First in Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of applying MAB learning algorithms in IoT networks,
+with a focus on the convergence of algorithms, in terms of successful transmission rates.
+We restrict to the case of a protocol slotted in both time and frequency, and we assume a certain clock pre-agreement and time synchronisation between devices and the base station.
+We studied the behavior of the network when the proportion of intelligent dynamic devices changes.
 Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
-Similarly, if the number of devices remain constant, increasing the successful transmission rate directly extends the IoT devices battery life, as they suffer less from failed transmissions.
+Similarly, if the number of devices remain constant, increasing the successful transmission rate directly improves the network QoS.
+%  as the base station receives more accurately the data sent by the end-devices since less transmissions failed.
 We show that both \UCB{} and Thompson sampling algorithms have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
@@
 %
-By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.
+By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using onlt the acknowledgement (\Ack) feedback sent from the base station.
 
-Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted in time and frequency.
+Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency and with perfect synchronisation.
 % , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
@@
 Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
-Surprisingly, the main take-away message is that a simple \UCB{} learning approach (that simply retransmits in the same channel), turns out to perform as well as more complicated heuristics.
+Surprisingly, the main take-away message is that a simple \UCB{} learning approach (that simply retransmits in the same channel), turns out to perform almost as well as more complicated heuristics.
 
@@
 The first model (without retransmission of packets),
-is interesting for its simplicity, and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
+is interesting for its simplicity, and because considering no retransmission can improve the battery life of the IoT devices, at the cost of a lower Quality of Service (QoS).
 %
-However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $M=10$).
+However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $10$ times).
 %
 The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
+
+\TODOL{Je dois clarifier les conclusions des deux modèles, est-ce que l'un est vraiment mieux pour améliorer la QoS mais coûte en batterie, et l'autre améliore la batterie mais coûte en QoS ?
+Pas très clair en ayant la tête dans le guidon là !...}
 
@@
 The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device.
-Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
+Also, other emission patterns could be considered, instead of a Bernoulli process for each user.
 In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.

### 2019-06-20 1-Preamble/4-Acknowledgement/acknowledgement.tex
Huge proofreading!

@@
 
-Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
-I want to thank especially the ``coffee friends$^*$'' in Rennes, for the daily interesting discussions, which where sometimes the highlights of our life at the office.
+Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Alessandro, Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
+I want to thank especially the ``coffee friends$^*$'' in Rennes, for the daily interesting discussions, the best moments of our life at the office.
 %

### 2019-06-20 1-Preamble/5-Abstract/abstract.tex
Huge proofreading!

@@
 % ,  Chapter~\ref{chapter:2}
-because proving the convergence of many end-devices playing a collaborative game without communication no coordination is hard, when end-devices all follow random active pattern.
+because proving the convergence of many end-devices playing a collaborative game without communication nor coordination is hard, when end-devices all follow random active pattern.
 % Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.

### 2019-06-20 2-Chapters/2-Chapter/chapter2.tex
Huge proofreading!

@@
 We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
-For more details on reinforcement learning, in particular about more generic models, we refer to the famous book by Sutton and Barto \cite{SuttonBarto2018}.
+For more details on reinforcement learning, in particular about more generic models, we refer to the famous book \cite{SuttonBarto2018}.
 
@@
 and a function $h$.
-The cumulant function $A(\eta)$ is entirely determined by the $\eta$ and $h$,
+The cumulant function $A(\eta)$ is entirely determined by $\eta$ and $h$,
 as $A(\eta) = \log \left( \int h(x) \exp(\eta x) \lambda(\mathrm{d} x) \right)$.
@@
 to present the lower-bounds, we prefer to focus on Bernoulli distributions, for simplicity.
-
+%
 To specify the dependency on an instance $I\in\cI$, we denote the regret of $\cA$ on instance $I$ and horizon $T$ by $R_T^{\cA}(I)$.
@@
         \begin{tabular}{c|cc|cc}
-        \textbf{Algorithms} & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
+        \textbf{Algorithms}
+        & rewards
+        & regret for $T=100$
+        & rewards
+        & regret for $T=10000$ \\
+        % & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
+        % & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
             \hline
@@
         \begin{tabular}{c|cc|cc}
-        \textbf{Algorithms} & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=100}$ & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=10000}$ \\
+        \textbf{Algorithms}
+        & rewards
+        & regret for $T=100$
+        & rewards
+        & regret for $T=10000$ \\
+        % & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=100}$ & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=10000}$ \\
             \hline

### 2019-06-20 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
DONE

@@
 
-The binomial theorem expands the sum in \eqref{eq:43:sumpca4}, so we can rewrite the expression of $p_{ca}$ as
+The binomial theorem expands the sum in \eqref{eq:43:sumpca4}, so we can rewrite the expression of $p_{ca}$
 %
@@
 the smaller the $p_c$, the larger is the gap between $p_c$ and $p_{c1}$.
-
+%
 We interpret this fact in two different situations:
@@
 			}
-			\Else(\tcp*[f]{Retransmit in same channel}){
+			\uElse(\tcp*[f]{Retransmit in same channel}){
 				j $\leftarrow$ last channel selected by first-stage \UCB\;
@@
 			}
-			\Else(\tcp*[f]{Random retransmission}){
+			\uElse(\tcp*[f]{Random retransmission}){
 				Transmit in channel $A(t) \sim \mathcal{U}(1,\ldots,K)$\;
@@
 			}
-			\Else(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^r$}){
+			\uElse(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^r$}){
 				Compute $\forall k, U^r_k(t) = \widehat{\mu^r_k}(t) + \sqrt{\alpha \log(t) / N_k^r(t)}$\;
@@
 			}
-			\Else(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^j$}){ % Retr. after trying channel $j$
+			\uElse(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^j$}){ % Retr. after trying channel $j$
 				j $\leftarrow$ last channel selected by first-stage \UCB\;
@@
 			}
-			\Else(\tcp*[f]{Packet retransmission with delayed $\mathrm{UCB}^d$}){
+			\uElse(\tcp*[f]{Packet retransmission with delayed $\mathrm{UCB}^d$}){
 				Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + \sqrt{\alpha \log(t) / N_k^d(t)}$\;

### 2019-06-21 1-Preamble/5-Abstract/resume_fr.tex
Should be nice, for print version i have these colored pages before chapters

@@
         %
-        \node[align=center] (introduction) at (0,3.25) [rectangle,draw,fill=blue!10] {\textbf{Chapitre~\ref{chapter:1}}\\Introduction};
-        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:2}}\\Modèle bandits multi-bras\\stochastique et stationnaire};
+        \node[align=center] (introduction) at (0,3.25) [rectangle,draw,fill=blue!15] {\textbf{Chapitre~\ref{chapter:1}}\\Introduction};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!15] {\textbf{Chapitre~\ref{chapter:2}}\\Modèle bandits multi-bras\\stochastique et stationnaire};
         \node[align=center] (chapter3) at (-2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:3}}\\SMPyBandits : bibliothèque\\de simulations pour MAB};
-        \node[align=center] (chapter25) at (2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:25}}\\Sélection séquentielle\\d'algorithme MAB};
+        \node[align=center] (chapter25) at (2.5,2.25) [rectangle,draw,fill=red!20] {\textbf{Chapitre~\ref{chapter:25}}\\Sélection séquentielle\\d'algorithme MAB};
         \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles MAB\\pour les réseaux IoT};
-        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:5}}\\Modèle MAB\\Multi-joueurs};
-        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:6}}\\Modèle MAB\\Non-stationnaire};
-        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!10] {\textbf{Chapitre~\ref{chapter:conclusion}}\\Conclusion générale};
-        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=blue!10] {Annexes};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!15] {\textbf{Chapitre~\ref{chapter:5}}\\Modèle MAB\\Multi-joueurs};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!20] {\textbf{Chapitre~\ref{chapter:6}}\\Modèle MAB\\Non-stationnaire};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!20] {\textbf{Chapitre~\ref{chapter:conclusion}}\\Conclusion générale};
+        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=yellow!10] {Annexes};
         %

### 2019-06-21 2-Chapters/1-Chapter/chapter1.tex
Should be nice, for print version i have these colored pages before chapters

@@
         %
-        \node[align=center] (introduction) at (0,3.25) [rectangle,draw,fill=blue!10] {\textbf{Chapter~\ref{chapter:1}}\\Introduction};
-        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapter~\ref{chapter:2}}\\Stochastic and Stationary\\Multi-Armed Bandit models};
+        \node[align=center] (introduction) at (0,3.25) [rectangle,draw,fill=blue!15] {\textbf{Chapter~\ref{chapter:1}}\\Introduction};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!15] {\textbf{Chapter~\ref{chapter:2}}\\Stochastic and Stationary\\Multi-Armed Bandit models};
         \node[align=center] (chapter3) at (-2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapter~\ref{chapter:3}}\\SMPyBandits: simulation\\library for MAB};
-        \node[align=center] (chapter25) at (+2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapter~\ref{chapter:25}}\\Online selection\\of the best algorithm};
+        \node[align=center] (chapter25) at (+2.5,2.25) [rectangle,draw,fill=red!20] {\textbf{Chapter~\ref{chapter:25}}\\Online selection\\of the best algorithm};
         \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapter~\ref{chapter:4}}\\Two MAB models\\for IoT networks};
-        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!10] {\textbf{Chapter~\ref{chapter:5}}\\Multi-players\\Multi-Armed Bandits};
-        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapter~\ref{chapter:6}}\\Non-stationary\\Multi-Armed Bandits};
-        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!10] {\textbf{Chapter~\ref{chapter:conclusion}}\\General Conclusion};
-        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=blue!10] {Appendix};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!15] {\textbf{Chapter~\ref{chapter:5}}\\Multi-players\\Multi-Armed Bandits};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!20] {\textbf{Chapter~\ref{chapter:6}}\\Non-stationary\\Multi-Armed Bandits};
+        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!20] {\textbf{Chapter~\ref{chapter:conclusion}}\\General Conclusion};
+        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=yellow!10] {Appendix};
         %

### 2019-06-21 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
DONE
Should be nice, for print version i have these colored pages before chapters

@@
     \includegraphics[width=0.70\linewidth]{system_model1.eps}
-    \caption[In our system model, some dynamic devices (in the IoT network in blue) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in orange).]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange} left/right).}
+    \caption[In our system model, some dynamic devices (in the IoT network) transmit packets to a gateway and suffer from the interference generated by neighboring networks.]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange} left/right).}
     \label{fig:41:system_model1}
@@
 
-\textbf{A safety check.}
+\paragraph{A safety check.}
 %
@@
 % \paragraph{Note on the simulation code.}
-\textbf{Reproducibility.}
+\paragraph{Reproducibility.}
 %

### 2019-06-21 2-Chapters/1-Chapter/chapter1.tex
Toujours pas sur de comment faire les transitions entre chapitres

@@
 \end{figure}
-%
-%
+
 \TODOL{Je peux améliorer (ou enlever) ce passage qui site J.Toledano, ou alors le commenter en plus de détails ? J'aime bien son rapport en fait (j'aurai du le lire au début de ma thèse !)}
@@
     The \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia encyclopedia} states
-    \emph{``A cognitive radio (CR) is a radio that can be programmed and configured dynamically to use the best wireless channels in its vicinity to avoid user interference and congestion.
+    \emph{``a CR is a radio that can be programmed and configured dynamically to use the best wireless channels in its vicinity to avoid user interference and congestion.
 Such a radio automatically detects available channels in wireless spectrum, then accordingly changes its transmission or reception parameters to allow more concurrent wireless communications in a given spectrum band at one location''}.
@@
 % explain PU and SU in a few sentences
-In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the mobile telephony network.
-The PU have a strict priority over any other non-paying users, referred to as Secondary User (SU).
-So even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU is equipped with an efficient sensing capacity, it can analyze its environment, and use a licensed band if it is free of any PU.
+In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the network.
+As exposed in the definition above, the CR philosophy suggests that if the network is not used by any PU in a certain location and time, then non-paying users, referred to as Secondary User (SU), could use it.
+Regulations state that PU have a strict priority over any other SU,
+but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU is equipped with an efficient sensing capacity, it can analyze its environment, and use a licensed band if it is free of any PU.
 % explain DSA
-This defines the concept of DSA was thus proposed in the early $2000$s, and we refer to the survey \cite{Zhao07} for more details.
+This defines the concept of DSA, proposed in the early $2000$s, and we refer to the surveys \cite{akyildiz2006next,Zhao07,garhwal2012survey} for more details.
 We focus on one example, the \textbf{Opportunistic Spectrum Access} (OSA) problem.
@@
 If a SU was able to sense for all the $K$ frequency bands, it could simply transmit in one of the free channels if any, or not transmit if all channels are used at a given time step.
-However, sensing is known to be error-prone and costly, especially for wide-band sensing, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
+However, sensing is known to be error-prone and costly, especially for wide-band sensing (for more details, see the surveys \cite{yucek2009survey,subhedar2011spectrum}), thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
 This assumption, along with the regulation that PU cannot be disturbed, enforces the SU to transmit in the channel that it sensed.
@@
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
+%
+For more details on the entire research on Cognitive Radio, we refer to the surveys \cite{garhwal2012survey,marinho2012cognitive}.
 
@@
 The goal of this first model of applying decentralized RL on the device side is to increase the Quality of Service (QoS) of the application of such IoT network, by reducing the rate of failed transmission.
-The base station will thus receive more up-link packets from the devices it serves, if they can successfully learn an efficient spectrum access scheme.
+The base station will thus receive more up-link packets from the devices being served, if they can successfully learn an efficient spectrum access scheme.
+It implies that the network can accept more end-devices while maintaining the same QoS.
 %
-The second model we propose then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increase QoS for the entire network.
+Our second model then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increase QoS for the entire network.
 
@@
 %
-In the two models presented above, the core concept is to let every dynamic device of a IoT network runs a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
+In both models presented above, the core concept is to let every dynamic device of a IoT network runs a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
 It means that each of them target its own local objective, which is to maximize its cumulated reward (in a selfish way).
-It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma} or other popular ``dilemma'' games.
+It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma} or other popular ``dilemmas''.
 % , and for more examples we refer to the book \cite{CesaLugosi06}.
@@
 We conclude this thesis in Chapter~\ref{chapter:conclusion}, to give some perspectives and directions about possible future works.
-Most of the contributions that we present in this thesis call for more developments, on both directions of providing more mathematical analysis and implementing new proof-of-concepts, using real wireless hardware and realistic Internet of Things networks.
+Most of the contributions that we present in this thesis call for more developments, on both directions of providing more mathematical analysis and implementing new proof-of-concepts, using real wireless hardware and realistic IoT networks.
 

### 2019-06-21 2-Chapters/2-Chapter/chapter2.tex
Toujours pas sur de comment faire les transitions entre chapitres

@@
         % \end{equation}
-        meaning that there exists a constant $C_I$ depending only on $I$,
+        meaning that there exists a constant $C'_I$ depending only on $I$,
         and a time $T_0$ such that $\forall T \geq T_0, \;\;\; R_T^{\cA}(I) \geq C'_I \log(T)$.
@@
 Similarly to what is considered for the first lower-bound,
-a natural question is to know if there is an algorithm that achieve a regret upper-bound of the form $R_T^{\cA}(I) \leq \bigO{\sqrt{K T}}$ for any instance, independently on the problem difficulty.
-It is the case for the Exp3 algorithm from \cite{Auer02}, for example.
-Since a few years, some algorithms were shown to achieve both a problem-dependent logarithmic and a minimax upper-bounds,
-like MOSS in \cite{Audibert2009minimax} or recently \KLUCBpp{} in \cite{Menard17},
+a natural question is to know if there is an algorithm achieving a regret upper-bound of the form $R_T^{\cA}(I) \leq \bigO{\sqrt{K T}}$ for any instance, independently on the problem difficulty.
+The question has been answered since the early $2000$s, as it is for instance the case for the Exp3 algorithm from \cite{Auer02}.
+In the last decade, some algorithms were shown to achieve both a problem-dependent logarithmic and a minimax upper-bounds,
+like MOSS from \cite{Audibert2009minimax} or recently \KLUCBpp{} from \cite{Menard17},
 and such results are usually referred to as ``best of both worlds''.
@@
 They are detailed in Algorithm~\ref{algo:2:simpleStrategies}.
-
-% https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
--- The \textcolor{deeppurple}{$\varepsilon$-greedy strategy} consists in alternating exploration and exploitation at a certain ratio \cite{Auer02}.
-% \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
-Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
-On the one hand, if $\varepsilon$ is constant, then the (mean) regret still grows linearly, as it is lower bounded by $(\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k) T$.
-In average, $T_0 = \varepsilon T$ steps are spent on exploration.
-%
-On the other hand, if a lower-bound $d$ on the positive gaps is known beforehand,
-we can consider a sequence $(\varepsilon_t)_{t\in\N^*}$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
-Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
-But this is not satisfactory, as $\min_k \Delta_k$ needs to be known in advance, which is usually not the case for real applications.
-Here again, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
 
@@
 
+% https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
+-- The \textcolor{deeppurple}{$\varepsilon$-greedy strategy} consists in alternating exploration and exploitation at a certain ratio \cite{Auer02}.
+% \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
+Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
+On the one hand, if $\varepsilon$ is constant, then the (mean) regret still grows linearly, as it is lower bounded by $(\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k) T$.
+In average, $T_0 = \varepsilon T$ steps are spent on exploration.
+%
+On the other hand, if a lower-bound $d$ on the positive gaps is known beforehand,
+we can consider a sequence $(\varepsilon_t)_{t\in\N^*}$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
+Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
+But this is not satisfactory, as $\min_k \Delta_k$ needs to be known in advance, which is usually not the case for real applications.
+Here again, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
+
 % https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
@@
 The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that includes well known and efficient algorithms such as $\UCB_1$, \klUCB{} and many others.
+
+\textbf{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}\\
+%
+Let us consider another approach for adaptive exploration, known as ``optimism under uncertainty'': assume each arm is as good as it can possibly be given the observations so far, and choose the best arm based on these optimistic estimate.
+%
+This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}, for bounded rewards.
+For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
+the \emph{average reward} $\widehat{\mu_k}(t) \eqdef \frac{X_k(t)}{N_k(t)}$
+and a \emph{confidence radius} $\xi_k(t) \eqdef \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
+%
+\begin{align}\label{eq:2:UCB_index}
+    U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation }\; \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration }\; \xi_k(t)}.
+\end{align}
 
@@
 
-\textbf{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}\\
-%
-Let us consider another approach for adaptive exploration, known as ``optimism under uncertainty'': assume each arm is as good as it can possibly be given the observations so far, and choose the best arm based on these optimistic estimate.
-%
-This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}, for bounded rewards.
-For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
-the \emph{average reward} $\widehat{\mu_k}(t) \eqdef \frac{X_k(t)}{N_k(t)}$
-and a \emph{confidence radius} $\xi_k(t) \eqdef \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
-%
-\begin{align}\label{eq:2:UCB_index}
-    U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation }\; \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration }\; \xi_k(t)}.
-\end{align}
-
 This selection rule $A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$ makes sense for the following reasons:
@@
 
-\textbf{A note about the implementation of \klUCB{} algorithms.}
+\paragraph{A note about the implementation of \klUCB{} algorithms.}
 %
 Implementing this kind of policy requires to implement two things.
-The first piece is usually easy: one has to implement a way to compute $\kl(x,y)$ the Kullback-Leibler divergence of the considered one-dimensional family,
+The first one is usually easy: one has to implement $\kl(x,y)$ the Kullback-Leibler divergence of the considered one-dimensional family,
 for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
-The second piece is less obvious: one has to solve the optimisation problem defining the $\sup$ in \eqref{eq:2:klUCB_index}.
+The second one is less obvious: one has to solve the optimisation problem defining the $\sup$ in \eqref{eq:2:klUCB_index}.
 Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\kl(x,\bullet)$ function is usually convex (for any $x$), two practical and mathematically correct solutions are Newton's method or a bisection search.
@@
 For instance, \UCB-H replaces the $\log(t)$ by a $\log(T)$ (as well as \klUCB-H), and it actually corresponds to the first algorithm analyzed in \cite{Auer02}.
-Using an estimator for the variance of the arms' samples gives \UCB-V, which was for instance proven to be efficient against Gaussian distributions with unknown variance.
+Using an estimator for the variance of the arms' samples gives \UCB-V, which was proven to be efficient against Gaussian distributions with unknown variance.
 Other variants include \UCB$^+$, $\UCB^{\dagger}$ from \cite{Lattimore2018refining}, {\UCB}oost, and many more.
@@
 %  using $\alpha=1$ (\texttt{AdBandits}).
-AdBandits follows the same assumptions as Thompson sampling and Bayes-UCB, and for instance for Bernoulli problems it uses a Beta posterior on each arm.
-It uses a random mixture between the two algorithms, and with a certain probability it uses the Thompson sampling decision rule, otherwise it uses the Bayes-UCB decision rule.
+AdBandits follows the same assumptions as Thompson sampling, and for instance for Bernoulli problems it uses a Beta posterior on each arm.
+It uses a random mixture between the two algorithms, and with a certain probability it follows Thompson sampling decision rule, otherwise it samples the arm with highest empirical mean (\ie, forced exploitation).
 It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given.

### 2019-06-21 2-Chapters/2-Chapter/chapter2.tex
Proofread Ch2 and CH3

@@
 %
-If you are discovering here the concept of bandits, I would like to recommend you to go online and play a few times with an interactive demonstration.
+If you are discovering here the concept of bandits, we would like to recommend you to go online and play a few times with an interactive demonstration.
 On this demo, you will be facing a MAB problem with $K=5$ arms, and you have $T=100$ decisions to make.
@@
 \end{smallproof}
-
+%
 \begin{lemma}\label{lem:2:HoeffdingInequality}
@@
 The Algorithm~\ref{algo:2:indexPolicy} below details a generic index policy, that includes well known and efficient algorithms such as $\UCB_1$, \klUCB{} and many others.
-
-\textbf{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}\\
-%
-Let us consider another approach for adaptive exploration, known as ``optimism under uncertainty'': assume each arm is as good as it can possibly be given the observations so far, and choose the best arm based on these optimistic estimate.
-%
-This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}, for bounded rewards.
-For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
-the \emph{average reward} $\widehat{\mu_k}(t) \eqdef \frac{X_k(t)}{N_k(t)}$
-and a \emph{confidence radius} $\xi_k(t) \eqdef \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
-%
-\begin{align}\label{eq:2:UCB_index}
-    U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation }\; \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration }\; \xi_k(t)}.
-\end{align}
 
@@
 
+\paragraph{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}\\
+%
+Let us consider another approach for adaptive exploration, known as ``optimism under uncertainty'': assume each arm is as good as it can possibly be given the observations so far, and choose the best arm based on these optimistic estimate.
+%
+This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}, for bounded rewards.
+
+For a parameter $\alpha$, the \UCB{} indexes are computed as follows, as a sum of
+the \emph{average reward} $\widehat{\mu_k}(t) \eqdef \frac{X_k(t)}{N_k(t)}$
+and a \emph{confidence radius} $\xi_k(t) \eqdef \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
+%
+\begin{align}\label{eq:2:UCB_index}
+    U_k^{\UCB_1}(t) = \UCB_k(t) \eqdef \underbrace{\frac{X_k(t)}{N_k(t)}}_{\text{Exploitation }\; \widehat{\mu_k}(t)} + \underbrace{\sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}}_{\text{Exploration }\; \xi_k(t)}.
+\end{align}
+
 This selection rule $A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} U_k^{\cA}(t))$ makes sense for the following reasons:
@@
         & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  \Pr\left( \widehat{\mu}_{k^*,n_{k^*}}(t) + \xi_{k^*,n_{k^*}}(t) < \widehat{\mu}_{k,n_k}(t) + \xi_{k,n_k}(t) \right) \nonumber \\
-        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t \bigl[ \underbrace{\Pr(I_1)}_{\leq 1/t^4} + \underbrace{\Pr(I_2)}_{\leq 1/t^4} + \underbrace{\Pr(I_3)}_{=0} \bigr] \nonumber \\
+        & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t \bigl( \underbrace{\Pr(I_1)}_{\leq 1/t^4} + \underbrace{\Pr(I_2)}_{\leq 1/t^4} + \underbrace{\Pr(I_3)}_{=0} \bigr) \nonumber \\
         & \leq 1 + \frac{8 \log(T)}{\Delta_k^2} + \sum_{t=n_T}^T \sum_{n_{k^*}=1}^t \sum_{n_{k}=n_T}^t  2 \frac{1}{t^4} \nonumber \\
@@
 
-We see in this Theorem~\ref{thm:2:UCBregretBound} that $\UCB_1$ has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime, and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO((\sum_{k: \Delta_k>0} \frac{1}{\Delta_k})\log(T))$.
+We see in this Theorem~\ref{thm:2:UCBregretBound} that $\UCB_1$ has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO((\sum_{k: \Delta_k>0} \frac{1}{\Delta_k})\log(T))$.
 %
@@
 It also needs to perform some mathematical operations, but a square-root and a logarithm are cheap, thus it costs a $\bigO{1}$ constant time for each arm and time step, thus a time $\bigO{K T}$ for the $T$ time steps.
-%
+
 We use $\UCB_1$ as the base building block for other policies in Chapter~\ref{chapter:4}, for these reasons.
@@
 % It yields improved upper confidence bounds (\ie, with smaller $\xi_k(t)$) that are based on Kullback-Leibler divergence.
-The indexes are computed as follows, for an exploration function $f(t)$
+We remind that we consider bounded rewards in $[0,1]$.
+Consider an exploration function $f(t)$,
 that is typically chosen as $f(t) \eqdef \log(t)+c\log(\log(t))$ for the theoretical analyses (with $c \geq 3$),
-and $f(t) \eqdef \log(t)$ in practice.
+and $f(t) \eqdef \log(t)$ in practice,
+then the \klUCB{} indexes are defined as follows
 % The constant $c>0$ is usually chosen as $c = 1$.
@@
 
-\textbf{Variants on $\UCB_1$.}
+\paragraph{Variants on $\UCB_1$.}
 %
@@
 		}
-		\caption{Thompson Sampling for Beta prior and posteriors, on Bernoulli rewards.}
+		\caption{Thompson Sampling for Bernoulli rewards, with a Beta prior on each arm, and also Beta posteriors.}
 		\label{algo:2:ThompsonSampling}
@@
 
-For curiosity, we mention a third Bayesian algorithm, AdBandits introduced in \cite{Truzzi13}, mainly because it has not gained any popularity despite its good empirical performance.
+We mention a third Bayesian algorithm for curiosity, AdBandits introduced in \cite{Truzzi13}, because it has not gained any popularity despite its good empirical performance.
 %  using $\alpha=1$ (\texttt{AdBandits}).
 AdBandits follows the same assumptions as Thompson sampling, and for instance for Bernoulli problems it uses a Beta posterior on each arm.
-It uses a random mixture between the two algorithms, and with a certain probability it follows Thompson sampling decision rule, otherwise it samples the arm with highest empirical mean (\ie, forced exploitation).
-It was proposed in this article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given.
-We found that in practice it can achieve very good performance, as illustrated in Section~\ref{sec:3:reviewSPAlgorithms}.
+It uses a random mixture of two algorithms: with a certain probability it follows Thompson sampling decision rule, otherwise it samples the arm with highest empirical mean (\ie, forced exploitation).
+It was proposed in the small article \cite{Truzzi13} and illustrated on empirical simulations, but no theoretical analysis was given.
+We illustrate in Section~\ref{sec:3:reviewSPAlgorithms} that it can achieve good performance in practice.
 %  and it did not gain popularity since then.

### 2019-06-21 2-Chapters/3-Chapter/chapter3.tex
Proofread Ch2 and CH3

@@
 
-SMPyBandits is a Python package I have developed since the beginning of my PhD.
-It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
-%  and SMPyBandits is written in Python.
+Since December $2016$ I worked regularly on the SMPyBandits library, which is a Python package  designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
+%
 This library is by far the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
@@
 %
-The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-players) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
+The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-players) defined by the number of arms $K$ and their distributions $(\nu_k)_{k\in[K]}$, for some time steps that range from $t=1$ to the horizon $T$.
 After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
@@
 For the classical single-player stochastic MAB model, as defined in Chapter~\ref{chapter:2},
-a stochastic MAB problem is defined by $K>1$ distributions $\nu_k$ (also called arms),
+a stochastic MAB problem is defined by $K>1$ distributions $(\nu_k)_{k\in[K]}$ (also called arms),
 used to generate the \iid{} samples $Y_{k,t} \sim \nu_k, \forall t$.
-An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r(t) = Y_{A(t),t}$ without knowing the other (hidden) rewards.
-Her goal is to maximize $\sum_{t=1}^T r(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.
+%
+An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r(t) = Y_{A(t),t}$ without knowing the other (hidden) rewards (in opposition to the full-information setting where $(Y_{1,t},\dots,Y_{K,t})$ is observed)).
+%
+Her goal is to maximize $\sum_{t=1}^T r(t)$, which require to trade off between exploring the unknown $K$ arms, exploiting the arm which is currently estimated to be the best one.
 
@@
 %
-It includes
+In particular, it includes
 \texttt{rhoRand}
@@
 % (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}})
-from \cite{Besson2018ALT}.
-For comparison, realistic (\eg, \UCB{} for multiple play) or full-knowledge centralized algorithms are also implemented.
+from \cite{Besson2018ALT}, and more recent solutions.
+For comparison, realistic (\eg, \UCB{} for multiple play) or full-knowledge \emph{centralized} algorithms are also implemented.
 
@@
     % \item
-    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
-Ask each player $\cA^{(j)}$ her chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
-The
-    sample (random) feedback $Y_{k,t}$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^{(j)}(t)$ from the random feedback and the collision model.
-Finally feeds the observation $(A^{(j)}(t), Y_{A^{(j)}(t)}(t), r^{(j)}(t))$ to player $j$, for all the $M$ players,
+    For $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
+Ask each player $\cA^j$ her chosen arm $A^j(t)$, then query the collision model to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
+    Then, sample (random) feedback $Y_{k,t}$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^j(t)$ from the random feedback using the collision model.
+    Finally, feed the observation $(A^j(t), Y_{A^j(t),t}, r^j(t))$ to player $j$, for all the $M$ players,
+    %
+    The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $r^j(t)=0$) if she is not the only one to have chosen an arm $k=A^j(t)$, otherwise $r^j(t)=Y_{A^j(t),t}$ if she is the only one playing this arm.
     % \item
@@
 For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulates the problem.
-We give another example for truncated Gaussian distributions, and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
+We give another example for \emph{truncated} Gaussian distributions (on $[0,1]$), and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
 
@@
 For more details, an interested reader can refer to the following Jupyter notebook \cite{jupyter}:
-\href{https://smpybandits.github.io/notebooks/Easily_creating_MAB_problems.html}{\texttt{SMPyBandits.GitHub.io/notebooks/Easily\_creating\_MAB\_problems.html}}.\\
-%
+\href{https://smpybandits.github.io/notebooks/Easily_creating_MAB_problems.html}{\texttt{SMPyBandits.GitHub.io/notebooks/Easily\_creating\_MAB\_problems.html}}.
+
 We have not yet added support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
@@
 %
-These CI services are free for open-source works, and are both triggered whenever a modification of the codebase is sent to the hosting platform (\ie, GitHub).
+These CI services are free for open-source works, and are both triggered whenever a modification of the codebase is sent to the hosting platform GitHub.
 
@@
 I have also been building the documentation manually on a weekly basis, to also host it on \href{https://smpybandits.github.io/}{\texttt{SMPyBandits.GitHub.io}} thanks to GitHub pages.
-In the last two years, the documentation saw about 8500 unique visits, while the GitHub project have been visited about 3500 unique visits.
+In the last two years, the documentation saw about 11000 unique visits, while the GitHub project have been visited about 4500 unique visits.
 
@@
 Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
-The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
+The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
 Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.
@@
 SMPyBandits has been used for the following research articles since $2017$, and it is used for the previous and the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
+%
+We give a link to a page on the documentation, that gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.
 
@@
     \item
-In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}.
+In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms.
 We designed a variant of the Exp4 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
+The documentation page is on \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}}
 % Aggregating experts is a well-studied idea in sequential learning and in machine learning in general.
@@
     \item
-In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-players bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}.
+In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-players bandit algorithms.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they obtain logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
-
-    \item
-In \cite{Besson2018DoublingTricks} (quickly presented in Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}.
+Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}}
+
+    \item
+In \cite{Besson2018DoublingTricks} (see Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes.
+\texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}}
 % In sequential learning, an algorithm is said to be ``anytime'' if it does not need to know the horizon $T$ of the experiments.
@@
 In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
-We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
+We illustrate and compare different algorithms designed for this family of non-stationary problems.
 We designed the \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.GLR_UCB.html}{GLRklUCB}} policy, and implemented most of the state-of-the-art passive or active adaptive policies, both adversarial- or stochastic-based, designed to tackle the piece-wise stationary MAB problems.
-We also implemented a large benchmark of different problems.
+We also implemented a large benchmark of different piece-wise stationary problems.
+See \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}}
 \end{itemize}
@@
     using $\varepsilon_t = \varepsilon_0 / t$, and $\varepsilon_0 = 0.1$ (chosen arbitrarily).
-    It has a very small cost in terms of time and memory, but it usually achieves linear regret.
+    It has a very small cost in terms of time and memory, but it usually achieves linear regret in practice, despite of the theoretical regret bound.
 
@@
 To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
-They are more recent (some from $2016$, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or are based on a different point-of-view.
+They are more recent (\eg, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or are based on a different point-of-view.
 The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with the state-of-the-art algorithms but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).
@@
 
-\textbf{Other parameters of the experiments.}
+\paragraph{Other parameters of the experiments.}
 %
@@
 
-We give in the figures below the results of the experiments.
+We give in Figures~\ref{fig:3:16_different_algorithms__lognormregret_vs_arms__16pb__T5000} and \ref{fig:3:16_different_algorithms__lognormregret_vs_times__16pb__K8} below the results of the experiments.
 % For the two different problems and different values of $T$ and $K$, we give the mean $\pm 1$ standard-variation of the empirical regret $R_T$ obtained by the different algorithms at the end of the experiments.
@@
         The $y$-axis is in log-scale.
-        All algorithms appear to have a regret slowly growing with respect to $K$, as predicted by the regret bounds which are linear with respect to the number of arms, as predicted.
-        Bayesian algorithms appear to be the most efficient, and \klUCB{} as well as \UCB{} are also shown to be efficient.
+        All algorithms appear to have a regret slowly growing with respect to $K$, as predicted by the regret bounds which are linear with respect to the number of arms.
+        Bayesian algorithms appear to be the most efficient, and \klUCB{} as well as \UCB{} are also seen to be efficient.
         % \textsc{Tsallis-Inf} is the only state-of-the-art algorithm which appears to be very inefficient for problems with a small number of arms.
@@
         Bayesian algorithms appear to be the most efficient, and \klUCB{} as well as \UCB{} are also shown to be efficient.
-        Finally, we also observe once more than both the $\varepsilon$-greedy and Explore-then-Commit algorithms performed poorly, achieving linear regret.
+        Finally, we also observe once more that both $\varepsilon$-greedy and Explore-then-Commit performed poorly, achieving linear regret.
 	}
@@
 %
-In the two Figures~\ref{fig:3:16_different_algorithms__lognormregret_vs_arms__16pb__T5000} and \ref{fig:3:16_different_algorithms__lognormregret_vs_times__16pb__K8}
-included in this section,
-we are able to check empirically that the regret of all the efficient algorithm indeed scales as predicted by the theory, that is
+% In the two Figures~\ref{fig:3:16_different_algorithms__lognormregret_vs_arms__16pb__T5000} and \ref{fig:3:16_different_algorithms__lognormregret_vs_times__16pb__K8}
+% included in this section,
+We are able to check empirically that the regret of all the efficient algorithm indeed scales as predicted by the theory, that is
 linearly in the number of arms, and logarithmically in the horizon, \ie, $R_T = \bigO{K \log T}$.
@@
 The main take-away message is the following: in all the rest of this thesis, we focus on two algorithms depending if simplicity or efficiency is favored:
-\begin{itemize}\tightlist
+\begin{itemize}%\tightlist
     \item
@@
     \item
-    \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when efficiency is favored, and when we analyze mathematically the performance of an algorithm that is running on top of a classical stationary MAB policy, like our two contributions, \MCTopM{} and \GLR.
+    \klUCB{} is used, in the two more theoretical Chapters~\ref{chapter:5} and \ref{chapter:6}, when we analyze mathematically the performance of an algorithm that is running on top of a classical stationary MAB policy, like our two contributions, \MCTopM{} and \GLR.
 \end{itemize}
+
+% FIXME
+\newpage
 
@@
 
-\paragraph{Computational time.}
+\subsection{Computational time.}
 
@@
 %
-This highlights a trade-off between optimality in terms of regret and computational efficiency.
+This observation highlights an interesting trade-off between having a small regret, and being fast and computationally efficient.
+%
+We also observe that BESA has a low regret but is much slower than all the other algorithms, because its complexity is exponentially growing wrt $K$ the number of arms.
 
@@
 
-\paragraph{Memory cost.}
+\subsection{Memory cost.}
 
@@
 
-\paragraph{Summary of the experiments.}
-%
-The main take-away message is the following:
-the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.
+% FIXME
+% \newpage
+
+\paragraph{Conclusion.}
+%
+% The main take-away message is the following:
+The simplest (but most efficient) algorithms have
+a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independently of $t$, and thus have a total time complexity bounded by $\cT_T = \bigO{KT}$,
+as well as a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\cM_T = \bigO{K}$.
 

### 2019-06-21 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Proof reading Ch4

@@
 % ----------------------------------------------------------------------
-\paragraph{Comparison with \ExpQ{}:}\label{sub:25:Exp4}
+\paragraph{Comparison with \ExpQ}.\label{sub:25:Exp4}
 %
@@
 We also note that \ExpQ{} uses \emph{losses} instead of rewards, with $\ell(t) = 1 - r(t)$, but this is only a change of vocabulary.
+A learner can equivalently play to maximize its cumulated rewards or to minimize its cumulated losses.
+Most works on using expert advice and game theory usually prefer to think about losses, as for instance the book \cite{CesaLugosi06}.
 
-\paragraph{How to choose the learning rates $(\eta_t)$:}
+\paragraph{How to choose $\eta_t$.}
 %
@@
 % helps to choose a good sequence.
-If the horizon $T$ is known (and fixed), the best choice is given by $\eta_t \eqdef \eta = 2 \log(N) / (T K)$.
+If the horizon $T$ is known (and fixed), they advise to use $\eta_t \eqdef \eta = 2 \log(N) / (T K)$.
 However, for real-world communication problems, it can be considered unrealistic to assume a fixed and known time horizon.
@@
 % - explain the problem studied in the simulation
-We focus on \emph{i.i.d.} MAB problems, with $K = 9$ arms\footnote{~Similar behaviors are observed for other values, \eg, trying up-to $K = 100$ gave the same results.}.
+We focus on \emph{i.i.d.} MAB problems, with $K = 9$ arms.
+Similar behaviors are observed for other values, \eg, trying up-to $K = 100$ gave the same results.
 For Bernoulli problem, the first one uses $\boldsymbol{\mu}=[0.1,\dots,0.9]$, and is considered as a ``simple'' problem.
@@
 %
-This empirical estimation of the expected regret $R_T$ is plotted below\footnote{~The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted.
-It corresponds to the second case in Theorem~\ref{thm:2:secondLogTLowerBound}.
-It is crucial to note that the lower-bound is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli}).},
+This empirical estimation of the expected regret $R_T$ is plotted below,
 as a function of $T$, comparing some algorithms $\Alg_1,\dots,\Alg_N$ (for $N=6$), and their aggregation with \Aggr{} (displayed in orange bold),
@@
 
-Note that for each of the $N=1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{1\leq k \leq K, 1 \leq t \leq T}$ (for every repetition), in order to compare the algorithms on the same realizations%
-\footnote{~Similar plots and similar results are obtained if this choice is not made, but it makes more sense to compare them against the same randomization.
+\textbf{Lower-bound.}
+%
+The Lai \& Robbins' logarithmic lower-bound \cite{LaiRobbins85} is also plotted.
+It corresponds to the second case in Theorem~\ref{thm:2:secondLogTLowerBound}.
+It is crucial to note that the lower-bound is only asymptotic, and as such one should not be surprised to see regret curves smaller than the lower-bound (\eg, for the easier Bernoulli problem in Figure~\ref{fig:25:EasyBernoulli})
+
+\textbf{Comparing the algorithms against the same realizations.}
+%
+Note that for each of the $N=1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{k\in[K], t\in[T]}$, for each of the $N$ repetitions, in order to compare the algorithms on the same realizations of the MAB problem.
+%
+Similar plots and similar results are obtained if this choice is not made, but it makes more sense to compare them against the same randomization.
 Note that this choice is the default in SMPyBandits, but the other possibility is implemented as well, by changing \texttt{cache\_rewards} to \texttt{true} or \texttt{false} in the \texttt{configuration} dictionary.
-See Section~\ref{sub:3:miniConfigurationFileSMPyBandits} for explanations on this \texttt{configuration.py} file, or the online documentation.}
-of the MAB problem.
+See Section~\ref{sub:3:miniConfigurationFileSMPyBandits} for explanations on this \texttt{configuration.py} file, or the online documentation.
 
@@
 	\centering
-	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
 	\caption{On a ``simple'' Bernoulli problem (semilog-$y$ scale).}
@@
 	\centering
-	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
 	\caption{On a ``harder'' Bernoulli problem, they all have similar performances, except \LearnExp, and our proposal \Aggr{} outperforms its competitors.}
@@
 	\centering
-	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
 	\caption{On an ``easy'' Gaussian problem, only \Aggr{} shows reasonable performances, thanks to BayesUCB and Thompson sampling.}
@@
 	\centering
-	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
 	\caption{On a harder problem, mixing Bernoulli, Gaussian, Exponential arms, with 3 arms of each type with the \emph{same mean}.}
@@
 	\centering
-	\includegraphics[width=1.07\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
 	\caption{The semilog-$x$ scale clearly shows the logarithmic growth of the regret for the best algorithms and our proposal \Aggr, even in a hard ``mixed'' problem (\emph{cf}.

### 2019-06-21 2-Chapters/2-Chapter/chapter25.tex
Proof reading Ch4

@@
 
-\paragraph{Naive approaches:}
+\paragraph{Naive approaches.}
 %
@@
 
-\paragraph{Illustrative example:}
+\paragraph{Illustrative example.}
 %
 For instance, if you know that the considered problem can either have $K$ arms with very close means, or one optimal arm far away from the other, two versions of $\UCB_1$ will perform quite differently in the two problems:
-using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case,
-while using a low $\alpha$, \ie, favoring exploitation, will give low regret in the second case.
-One approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let another learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
+using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case and high regret in the second case,
+while using a low $\alpha$, \ie, favoring exploitation, will obtain opposite performance.
+A first approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let a ``master'' learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
 

### 2019-06-21 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Proof reading Ch4

@@
 
+\begin{figure}[!h]
+    \centering
+    \includegraphics[scale=0.40]{protocol.eps}
+    \caption[The considered time-frequency slotted protocol.
+Each frame is composed by a fixed duration uplink slot in which the end-devices transmit their (uplink) packets.
+If a packet is well received, the gateway replies by transmitting an \Ack, after the ack delay.]{The considered time-frequency ALOHA-based communication protocol \cite{Abramson1970,Roberts75}.
+Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
+If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
+    \label{fig:41:protocol}
+\end{figure}
+
 \paragraph{Slotted protocol.}
 %
-For simplicity, and for consistency with the rest of this thesis, we only consider a discrete protocol (\ie, slotted).
+For simplicity, and as in all this thesis, we only consider a discrete protocol (\ie, slotted).
 As illustrated in Figure~\ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
 All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
-In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication, with no sensing.
+In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication protocol \cite{Abramson1970,Roberts75}, with no sensing.
 Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the gateway.
@@
 This way, no collision can occur on the downlink messages, easing the analysis of collisions.
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[scale=0.40]{protocol.eps}
-    \caption[The considered time-frequency slotted protocol.
-Each frame is composed by a fixed duration uplink slot in which the end-devices transmit their (uplink) packets.
-If a packet is well received, the gateway replies by transmitting an \Ack, after the ack delay.]{The considered time-frequency slotted protocol.
-Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
-If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
-    \label{fig:41:protocol}
-\end{figure}
 
@@
 
-\textbf{Random emission patterns:}
+\textbf{Random emission patterns.}
 %
 We suppose that all devices follow the same emission pattern, being fixed in time, and we choose to model it as a simple Bernoulli process:
-all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability\footnote{~In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $K / (S + D)$ for all devices to communicate successfully (in average).}.
+all devices have the same probability to send a packet in any (discrete) temporal slot, and we denote $p \in (0, 1)$ this probability.
 The parameter $p$ essentially controls the frequency of communication for each device, once the time scale is fixed, and $1/p$ is proportional to the \emph{duty cycle}.
@@
 For instance, for IoT devices sending messages with a duration of $1$ second and possibly at every second, then on a daily basis we have $p = 1 / (12 \times 60 \times 60) \simeq 1.5 \times 10^{-5}$.
+In the experiments below, $p$ is about $10^{-3}$, because in a crowded network $p$ should be smaller than $K / (S + D)$ for all devices to communicate successfully (in average).
 
-\textbf{Some assumptions on the network occupation:}
+\textbf{Some assumptions on the network occupation.}
 %
@@
 
-\textbf{Link with the MAB model:}
+\textbf{Link with the MAB model.}
 %
@@
 the $S_k$ static devices in each channel $k$ are assumed to be independent, and static and dynamic devices are assumed to \emph{not} transmit at each time $t$ with a fixed probability $1-p$,
-so probability of successful transmission is computed as follows
+so probability of successful transmission is computed as follows.
 
@@
 % $\Pr(\text{no other dynamic device sent in}\;i)$,
-is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device}\;k\;\text{did not sent in}\;i) = \prod_{k=1}^{D-1} \Pr(\text{device}\;k\;\text{did not sent in}\;i)$.
-And $\Pr(\text{device}\;k\;\text{sent in}\;i) = p \times 1 / K$, by uniform choice on channels and the Bernoulli emission hypothesis.
+is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device } k \text{ did not sent in } i) = \prod_{k=1}^{D-1} \Pr(\text{device } k \text{ did not sent in } i)$.
+And $\Pr(\text{device } k \text{ sent in } i) = p \times 1 / K$, by uniform choice on channels and the Bernoulli emission hypothesis.
 So $q = \prod_{k=1}^{D-1} (1 - p/K) = (1-p/K)^{D-1}$.
@@
 %
-This expression \eqref{eq:41:strategynaive} is constant (in time), and easy to compute numerically, but comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it
-(maybe after a long enough initial learning period).
+This probability \eqref{eq:41:strategynaive} is constant wrt time, and easy to compute numerically.
+Comparing the successful transmission rate of any policy against this naive policy is important, as any efficient learning algorithm should outperform it,
+maybe after a long enough initial learning period.
 
@@
 
-Consequently, an optimal allocation vector $(D_1,\dots,D_{K})$ is a solution of the following real-valued constraint optimization problem:
+Consequently, an optimal allocation vector $(D_1,\dots,D_{K}) \in \R^K$
+is a solution of the following real-valued constraint optimization problem:
 %
@@
     \text{such that}\;\; & \sum_{k=1}^{K} D_k = D, \label{eq:41:eqCstr}\\
-    & D_k \geq 0 \qquad \forall k\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
+    & D_k \geq 0 \qquad \forall k\in [K]. \label{eq:41:ineqCstr}
     % \end{cases}
@@
 Moreover, condition \eqref{eq:41:eqCstr} implies that the Lagrange multiplier $\lambda$ is the solution of the constraint $\sum_{k=1}^{K} D_k^*(\lambda) = D$.
-%
 This single constraint can be solved numerically, with simple one-dimensional root finding algorithms.
-Solving the optimization problem provides the optimal real number value for $D_k^*$, which has to be rounded\footnote{~Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels.} to find the optimal number of devices for channel $k$:
+
+Note that solving the optimization problem provides the optimal \emph{real values} $(D_k^*)_k$, which have to be rounded to find the optimal \emph{integer} number of devices for channel $k$.
+Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
 %
@@
 
-\textbf{Using a bandit algorithm for IoT devices:}
+\textbf{Using a bandit algorithm for IoT devices.}
 %
@@
 
-Clearly, the problem is less interesting for a homogeneous repartition, as all channels appear the same for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
+Clearly, the problem is less interesting for a homogeneous repartition, as all channels are equivalent for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
 
@@
 
-\textbf{Successful transmission rate as a function of the number of dynamic devices.}
+\paragraph{Successful transmission rate as a function of the number of dynamic devices.}
 %

### 2019-06-21 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Finishing to fix Emilie's comment of Ch2
Proof reading Ch4

@@
 
-This PoC implements an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) devices, embedding the proposed solution,
+This PoC implements an IoT network the following way: one base station, one or several intelligent (\emph{i.e.}, learning) devices, embedding the proposed solution,
 and a traffic generator that emulates radio interference from many other devices.
-Intelligent devices communicate with the gateway with a wireless ALOHA-based protocol with acknowledgements, which does not require any specific overhead for the learning.
+Intelligent devices communicate with the base station with a wireless ALOHA-based protocol with acknowledgements, which does not require any specific overhead for the learning.
 %
@@
 
-\textbf{Related works.}
-% \TODOL{Est-ce que je fais ça ici ou à la fin du chapitre 5 ? Ou peut-être en chapitre 1, ça ferait du contenu au chapitre 1...}
-% FIXED j'en parle au chapitre 1
+\paragraph{Related works.}
 %
@@
 
-In our demo, we consider a simple wireless network, that reproduces the model of Section~\ref{sec:4:firstModel}, consisting of one gateway (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
-Some dynamic intelligent devices (end-user or autonomous devices) try to communicate with the gateway, with a low-overhead protocol.
+In our demo, we consider a simple wireless network, that reproduces the model of Section~\ref{sec:4:firstModel}, consisting of one base station (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
+Some dynamic intelligent devices (end-user or autonomous devices) try to communicate with the base station, with a low-overhead protocol.
 This communication can be done in different channels which are also shared by devices using other networks.
-Once the gateway receives a packet transmitted by a dynamic device in one channel (\ie, if no collision occurred), it transmits back to it an acknowledgement in the same channel, after a fixed-time delay, as it is done in the LoRaWAN standard.
+Once the base station receives a packet transmitted by a dynamic device in one channel (\ie, if no collision occurred), it transmits back to it an acknowledgement in the same channel, after a fixed-time delay, as it is done in the LoRaWAN standard.
 This \emph{Ack} allows the device to learn about the channel quality (\ie, mean availability) and thus, to use learning algorithms for the purpose of best channel selection.
@@
 For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a sequence of QPSK symols if the PoC has to consider more users.
-With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway.
-Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
-Thanks to this index, we can have several devices communicating with the same gateway.
+With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the base station.
+Once the base station receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
+Thanks to this index, we can have several devices communicating with the same base station.
 %
 In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
-In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the gateway.
+In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the base station.
 
@@
 As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP:
-the gateway,
+the base station,
 a traffic generator which emulates the interfering traffic (made by surrounding static devices),
@@
 % Octoclock
-To ease the synchronization in both time and frequency between the boards representing the dynamic devices and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
+To ease the synchronization in both time and frequency between the boards representing the dynamic devices and the base station, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
 % https://www.ettus.com/product/details/OctoClock-G
@@
 
-\begin{enumerate}[leftmargin=6mm]
-    \item[(1)]
+\begin{enumerate}[label=\textcolor{darkred}{\protect\boldcircled{\arabic*}},leftmargin=6mm]
+    \item %[(1)]
+    % \item[\textcolor{darkred}{\circled{1}}]
 The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.
@@
 
-    \item[(2)]
+    \item %[(2)]
+    % \item[\textcolor{darkred}{\circled{2}}]
 The second part is the interface of the intelligent device which is made of four parts.
-At the top left, we observe the constellation of the transmitted packet \emph{(a)}.
-At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \emph{(b)}.
+At the top left, we observe the constellation of the transmitted packet \circled{a}.
+At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \circled{b}.
 We can see, in this view that the device transmitted its last $9$ packets in channels $\#3$ and $\#4$.
-Then, at the top right of this interface \emph{(c)}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the gateway (\textcolor{darkblue}{blue}).
+Then, at the top right of this interface \circled{c}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the base station (\textcolor{darkblue}{blue}).
 Colors in the ``waterfall'' represent the RF power level received at the device antenna.
-Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (transmission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \emph{(c)} in part (2)).
-Finally, at the bottom right \emph{(d)}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).
+Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (transmission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \circled{c} in part \textcolor{darkred}{\boldcircled{2}}).
+Finally, at the bottom right \circled{d}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).
 
-    \item[(3)]
-The last part is the interface of the gateway, where we can see the traffic observed by the gateway \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
+    \item %[(3)]
+    % \item[\textcolor{darkred}{\circled{3}}]
+The last part is the interface of the base station, where we can see the traffic observed by the base station \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
 The observed traffic is coherent with \emph{(c)} of part (2), as the elements are very close the one from the others on the testbed.
@@
 The occupancy rate of a channel, which denotes the mean occupancy, is implemented using the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).
+
 % They did not overlap each other, as

### 2019-06-21 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
Proofreading
Proof reading Ch4

@@
 \texttt{Generator} is the only hand-written block, which is configured with a list of active channels, a list of occupancy rate, and constants about the PHY layer (preamble length and data length).
-It uses one USRP only for emitting data (``Sink'' mode).
+It uses one USRP equipped with one antenna, as it only emits data (by using the ``UHD: USRP'' block in``Sink'' mode).
 
@@
 
-\subsubsection*{IoT gateway}
+\subsubsection*{IoT base station}
 
-The \textbf{IoT gateway} is presented in Figure~\ref{fig:4app:USRP_RX_BTS__v1__simple_grc} below.
+The \textbf{IoT base station} is presented in Figure~\ref{fig:4app:USRP_RX_BTS__v1__simple_grc} below.
 %
@@
 \end{itemize}
-All blocks need to know constants about the PHY layer (preamble length and data length). The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).
+All blocks need to know the constants about the PHY layer (preamble length and data length).
+The base station uses one USRP board equipped with two antennas,
+to emit and receive data (by using ``Sink'' and ``Source'' modes).
 
@@
     \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_RX_BTS__v1__simple_grc.png}
-    \caption{The \textbf{IoT gateway} flow-graph.}
+    \caption{The \textbf{IoT base station} flow-graph.}
     \label{fig:4app:USRP_RX_BTS__v1__simple_grc}
@@
 Figure~\ref{fig:4app:USRP_TX_SU__v1__simple_grc} below shows the \textbf{IoT dynamic device} flow-graph.
-The hand-written blocks are the
+The hand-written blocks are
 \begin{itemize}
@@
 Most blocks need to know constants about the PHY layer (preamble length and data length).
-Any of the IoT dynamic devices use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes)
+Any of the IoT dynamic devices also uses one USRP board equipped with two antennas,
+to emit and receive data (``Sink'' and ``Source'' modes).
+
+We note the difference between the base station and the end-devices.
+The base station uses its Rx antenna to scan the entire range of the $K$ channels, at all time steps ; while dynamic devices use their Rx antennas only to listen for the acknowledgement sent by the base station if it received and decoded its uplink packet.
+After each transmission, the dynamic device listens to one channel during a certain time interval, in the channel used for uplink transmission.
 

### 2019-06-21 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Proof reading Ch4

@@
 In this section, we extend the previous model to take into account the possibility for retransmissions of a message after a collision.
-This is of major importance as most protocols for real-world IoT networks can use retransmissions, it is for instance the case of the LoRaWAN standard.
+This is of major importance as most protocols for real-world IoT networks can use retransmissions, it is for instance the case of the LoRaWAN standard \cite{Raza17}.
 As before, we propose and evaluate different learning strategies based on MAB algorithms.

### 2019-06-21 2-Chapters/4-Chapter/chapter4.tex
Proof reading Ch4

@@
 
-A satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
+We conclude that this learning-based approach is efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rates.
 It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
-Sadly, analyzing our model with thousands of independent devices communicating and learning in their own (random) time scales turned out to be of extreme difficulty.
+Unfortunately, analyzing our model with thousands of independent devices, all communicating and learning in their own (random) time scales, turned out to be of extreme difficulty.
 %
 We can also highlight that the two models we studied are complementary:
-the first model (without retransmission of packets), is interesting for its simplicity, and focused on improving the battery life of IoT devices, at the cost of a lower Quality of Service (QoS),
+the first model (without retransmission of packets), is interesting for its simplicity, and focused on improving the battery life of IoT devices,
+% , at the cost of a lower Quality of Service (QoS),
 %
-while the second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
+while the second model is more interesting if the main target is an improvement of the Quality of Service (QoS).
+% , rather than a longer battery life.
 
-\textbf{Multi-Players MAB.}
+\paragraph{Multi-Players MAB.}
 %
@@
 
-\textbf{Non-stationary MAB.}
+\paragraph{Non-stationary MAB.}
 %

### 2019-06-21 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done proofreading quickly Ch5

@@
 %
-The decentralized hypothesis means that the central base station does not given any direct order to the devices, they are all in charge of deciding the channel they each want to use at each time step.
+The decentralized hypothesis means that the central base station (or gateway) does not given any direct order to the devices, they are all in charge of deciding the channel they each want to use at each time step.
 Moreover, the devices are independent and do not directly communicate with each other either.
@@
 
-\textbf{Outline.}
+\paragraph{Outline.}
 %
@@
 
-\textbf{Publication.}
+\paragraph{Publication.}
 %
@@
     \forall a\in(0,1) & R_T(\boldsymbol{\mu},M,\cA) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
-    \text{and} \forall a\in(0,1) & \;
+    \text{and } \; \forall a\in(0,1) & \;
     \forall j \in [M], k \in \Mbest, \;\;
@@
 
-Having a small regret on every problem instance, \ie, being uniformly efficient,
+Having a small regret on every problem instance (\ie, being uniformly efficient)
 is a natural assumption,
-that rules out algorithms tuned to perform well on specific instances only (\eg, fixed-armed algorithms).
-%
-From this assumption $\left(R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
+that rules out policies tuned to perform well only on specific instances (\eg, fixed-armed policies).
+%
+From this assumption $R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
 ${T}- \E_{\mu}[N_k(T)] {=} o(T^{\alpha})$, and so
@@
 % As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
+\textbf{About fairness.}
+%
 In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
 The notion of fairness defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
-For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is given and known to the algorithm.
+For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is known by the algorithm.
 
@@
 
-The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
+\textbf{Collision information.}
+%
+The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} comes from Appendix~E of \cite{KaufmannAbbas19}.
 Similarly to what we did in Section~\ref{sec:2:notations},
@@
 defined as $O_t^j \eqdef \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
-where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).
+where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$.
+For instance, an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index the decision is an $\argmax$, and ties are usually broken by a uniform random selection among the arms with maximum index.
 
@@
 This result can also be used to recover Proposition~1 from \cite{Anandkumar11}, giving an upper bound on the regret that only depends on
-the \emph{expected number of sub-optimal selections} -- $\E_{\mu}[N_k(T)]$ for $k \in \Mworst$ --
-and the \emph{expected number of colliding players on the optimal arms} -- $\E_{\mu}[\cC_k(T)]$ for $k \in \Mbest$.
+the \emph{expected number of sub-optimal selections}, $\E_{\mu}[N_k(T)]$ for $k \in \Mworst$,
+and the \emph{expected number of colliding players on the optimal arms}, $\E_{\mu}[\cC_k(T)]$ for $k \in \Mbest$.
 Note that, in term (c) the number of colliding players on the sub-optimal arm $k$ may be upper bounded as $\E_{\mu}[\cC_k(T)] \leq M \E_{\mu}[N_k(T)]$.
@@
 In the next section, we present an algorithm that has a logarithmic regret,
+by controlling its sub-optimal selections,
 while ensuring that the number of sub-optimal selections is matching the lower bound of Theorem~\ref{thm:5:BetterLowerBound}.
@@
 
-\textcolor{red}{\textbf{Warning:}}
+\begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
+  \textcolor{red}{\textbf{Warning.}}
   We want to strengthen the fact that in all the proposed algorithms,
@@
   %
-We discuss the fact that this would be an unrealistic hypothesis in Section~\ref{par:5:knowingYourIDinMPBanditsGame} below.
+  This would be an unrealistic hypothesis, as explained in Section~\ref{par:5:knowingYourIDinMPBanditsGame} below.
+\end{leftbar}  % XXX leftbar warningbar, comment if needed
 
@@
 % %
-The approache we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
+The approaches we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
 In particular, we focus on two types of indices:
@@
 
-\textbf{Two ideas for orthogonalization.}
+\paragraph{Two ideas for orthogonalization.}
 \label{par:5:twoIdeasOrthogonalization}
@@
 %
-More formally, letting $\pi(k,\mathbf{g})$ be the index of the $k$-th largest entry in a vector $\mathbf{g}$,
+More formally, letting $\pi(k,\mathbf{U})$ be the index of the $k$-th largest entry in a vector $\mathbf{U}$,
 % (for $k\in[K]$),
@@
 % \begin{equation}
-$A^j(t) \eqdef \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right)$.
+$A^j(t) \eqdef \pi\left(R^j(t), [U^j_\ell(t)]_{\ell\in[K]}\right)$.
 % \end{equation}
@@
 % Now, our algorithms !!
-\textbf{Our two proposals.}
+\paragraph{Our two proposals.}
 We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
 \begin{equation}
-  \TopM(t) \eqdef \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
-\end{equation}
-
-\textbf{The \RandTopM{} algorithm.}
+  \TopM(t) \eqdef \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell\in[K]}\right), k=1,\dots,M\right\}.
+\end{equation}
+
+\paragraph{The \RandTopM{} algorithm.}
 %
 We precisely state our first proposal below in Algorithm~\ref{algo:5:RandTopM}.
+%
+\RandTopM{} is essentially a refinement over \RhoRand, to not use the indirection of ranks, and a simpler version of \MCTopM.
+The difference with \MCTopM{} is that the later introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
+%
+In \RandTopM{}, player $j$ is always considered ``not fixed'',
+and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.
 
@@
 
-\RandTopM{} is essentially a refinement over \RhoRand, to not use the indirection of ranks, and a simpler version of \MCTopM.
-The difference with \MCTopM{} is that the later introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
-%
-In \RandTopM{}, player $j$ is always considered ``not fixed'',
-and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.
-
-\textbf{The \MCTopM{} algorithm.}
+\paragraph{The \MCTopM{} algorithm.}
 %
@@
   As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t-1)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
-  Formally, $s^j(0)$ is false, and $s^j(t)$ is defined inductively from $s^j(t-1)$ as
+  Thus $s^j(0)$ is false, and $s^j(t)$ is defined inductively from $s^j(t-1)$ as
   % following the algorithm of \MCTopM:
@@
   She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
-  or she stays in state $\overline{s^j(t)}$,
-  with two cases.
+  or she stays in state $\overline{s^j(t)}$
+  in two cases.
   Either she sampled $A^j(t)$ uniformly
@@
   if $A^j(t-1) \notin \TopM(t)$.
-  In particular, note that if $\overline{C^j(t-1)}$, transition $(3)$ is executed and not $(2)$.
+  In particular, note that $(3)$ is executed and not $(2)$ if $\overline{C^j(t-1)}$.
   %
@@
   (possibly larger than $T$),
-  of sequences during which $I_2(s)$ is true for all $s=\theta_i,\dots,\tau_i-1$ but not before and after,
+  of sequences during which $I_2(s)$ is true for all $s \in [\theta_i,\dots,\tau_i-1]$, but not before and not after,
   that is
@@
   to obtain a decomposition with two terms $(\alpha)$ and $(\beta)$,
-  $\E_{\mu}[N_2^j(t)] \leq \E_{\mu}\left[ n(T) \right] \times \max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]$.
+  $\E_{\mu}[N_2^j(t)] \leq \E_{\mu}\left[ n(T) \right] \times \max\limits_{i \geq 1} \E_{\mu}\left[ \tau_i - \theta_i \right]$.
   % $\E_{\mu}[N_2^j(t)] \leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}$.
@@
 
-  We can be more precise about the constants, all the previous arguments can be used successively:
-  \begin{align}
+  We can be more precise about the constants, by using the previous arguments successively.
+  \begin{align*}
     \E_{\mu}[\cC(T)]
@@
     = M \left(\sum_{j=1}^M \E_{\mu}[N_2^j(T)] + \E_{\mu}[N_3^j(T)]\right) \\
-    &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \notag
-  \end{align}%
-  \begin{align}
-    &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \notag \\
-    &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \left(\sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T} \notag  \\
-    &\leq \left(2 M^3 + M^2\right) \left(\sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
-  \end{align}
+    &\leq M^2 \left(\E_{\mu}[n(T)] \E_{\mu}[\theta_i - \tau_i] \right) + M^2 \E_{\mu}[N_3^1(T)] \\
+    &\leq M^2 (1+\E_{\mu}[N_3^1(T)] + \E_{\mu}[N_5^1(T)]) M + M^2 \E_{\mu}[N_3^1(T)] \\
+    &\leq 2 M^3 \E_{\mu}[N_3^1(T)] + \smallO{\log T} + M^2 \sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{\log(T)}{\kl(\mu_a,\mu_b)} + \smallO{\log T}  \\
+    &\leq \left(2 M^3 + M^2\right) \sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{\log(T)}{\kl(\mu_a,\mu_b)} + \smallO{\log T}.
+  \end{align*}
 
@@
   %
-  \begin{align}
-    \sum_{k=1}^K \E_{\mu}[\cC^k(T)] & = \E_{\mu}[\cC(T)] \nonumber\\
-    & \leq M^2\left(2 M + 1\right) \left(\sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{1}{\kl(\mu_a,\mu_b)}\right) \log(T) + \smallO{\log T}.
-  \end{align}
+  \begin{equation}
+    \sum_{k=1}^K \E_{\mu}[\cC^k(T)] = \E_{\mu}[\cC(T)]
+    \leq M^2\left(2 M + 1\right) \sum_{\substack{a,b=1,\dots,K\\\mu_a < \mu_b}} \frac{\log(T)}{\kl(\mu_a,\mu_b)} + \smallO{\log T}.
+  \end{equation}
 \end{smallproof}
@@
       \begin{tabular}{cc|ccc}
-          \textbf{Algorithm} & \textbf{Index policy} & $M=3$ players & $M=6$ players & $M=9$ players \\
+          \textbf{Algorithm} & \textbf{Index policy} & $M=3$ players & $M=6$ & $M=9$ \\
           \hline
@@
       \end{tabular}
-      \caption{No matter the orthogonalization policy and $M$ the number of players, using \klUCB{} is much more efficient than using \UCB{}, for multi-players bandit (here in a simple problem with $K=9$ arms).}
+      \caption{For all orthogonalization policies and different numbers of players, using \klUCB{} is much more efficient than using \UCB{}, for multi-players bandit (here in a simple problem with $K=9$ arms).}
       \label{table:5:comparisonUCB_klUCB}
@@
 
-\textbf{Other experiments.}
+\paragraph{Other experiments.}
 %

### 2019-06-21 2-Chapters/5-Chapter/chapter5.tex
DONE
Done proofreading quickly Ch5

@@
         & \textcolor{red}{Estimate $M$} & \textcolor{red}{$563 \pm 546$} & \textcolor{red}{$1560 \pm 1293$} & \textcolor{red}{$618 \pm 29$} \\
-        \hline
+        % \hline
     \end{tabular}
@@
 
-Inspired by \cite{Besson2018ALT},
-Boursier and Perchet studied this question in the summer $2018$ following its publication \cite{BoursierPerchet18}.
-Their answer is that the model without sensing information is essentially not harder than the model with sensing information.
+Inspired by the publication of our article \cite{Besson2018ALT},
+V. Boursier and E. Perchet studied this question in the following summer $2018$ \cite{BoursierPerchet18}.
+Their main contribution is an answer to the aforementioned open problem,
+as they showed that the model without sensing information is essentially not harder than the model with sensing information.
 %
-Similarly, Lugosi and Mehrabian studied the same problem \cite{LugosiMehrabian18}.
-In both works, the authors detail algorithms that give a logarithmic regret if $M \leq K$ players all independently implement the proposed algorithm.
+Similarly, G. Lugosi and A. Mehrabian studied the same problem \cite{LugosiMehrabian18}.
+In both works, the authors detail algorithms that give a logarithmic system regret, if $M \leq K$ players all independently implement the proposed solution.
 
@@
         \textcolor{red}{Improved \MusicalChair{}} & $\textcolor{red}{12149 \pm 60}$ & $\textcolor{red}{22341 \pm 77}$ & $\textcolor{red}{27462 \pm 85}$ \\
-        \hline
+        % \hline
     \end{tabular}
@@
 Another line of research is to consider models that are closer to realistic wireless communication networks, as it is done in \cite{Avner16,AvnerMannor18}.
-Explaining in details their model and proposed solutions would be quite lengthy, and we rather prefer to let the interested reader refer to the later work \cite{AvnerMannor18}.
-We sum-up its contributions quickly.
+Explaining in details their model and proposed solutions is out of the scope of this thesis,
+but we sum-up the contributions of the latest \cite{AvnerMannor18}.
+%
 They address several aspects of the challenge of communication networks shared by many users simultaneously: learning unknown stochastic network characteristics, sharing resources with other users while keeping coordination overhead to a minimum.
 The solution they proposed combines Multi-Armed Bandit learning with a lightweight signalling-based coordination scheme, and ensures convergence to a stable allocation of resources.
-Their work considers single-user level algorithms for two scenarios: an unknown fixed number of users, and a dynamic number of users, both for different arms means for each user (players).
+Their work considers single-user level algorithms for two scenarios: an unknown fixed number of users, and a dynamic number of users, both for different arms means for each user.
 Analytic performance guarantees, proving convergence to stable marriage configurations, are presented for both setups.
-The algorithms are based on a system-wide perspective, rather than focusing on single user welfare.
+Similarly to our work, the algorithms are based on a system-wide perspective, rather than focusing on single user welfare.
+%
 
-% \TODOL{A conclusion from this sub-section presenting this model?}
+\TODOL{A conclusion from this sub-section presenting this model?}
 
@@
 %
-Starting in 2012 by \cite{Kalathil12} where arms are Markov chain, this model was studied more actively recently, \cite{DarakHanawal18,Bistritz18,KaufmannAbbas19,Tibrewal2019}.
+Starting in 2012 by \cite{Kalathil12} where arms are Markov chain, this model was studied more actively recently, in two articles published in $2018$ \cite{DarakHanawal18,Bistritz18} and two more in $2019$ \cite{KaufmannAbbas19,Tibrewal2019}.
 
@@
 % combining Musical Chair with an epoch-based alternative phases of forced exploration and pure exploitation.
-They analyze their GoT algorithm and proven that it achieves a regret of $\cO((\log(T))^{2+\kappa})$ for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
+They proved that their GoT algorithm achieves a regret of $\cO((\log(T))^{2+\kappa})$, for any positive constant $\kappa$, as small as possible, if $\kappa$ is known by the algorithm.
 
@@
         \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 \pm 13$ \\
-        \hline
+        % \hline
     \end{tabular}
@@
 
-When we were working on multi-players bandits in Autumn $2017$ \cite{Besson2018ALT}, we left the study of the piece-wise stationary case as a future work, and shortly after we were excited to see that three independent works tackling this question were published, in December $2018$ \cite{WeiSrivastava18Distributed} and in February $2019$ \cite{AlaturLevyKrause19,bande2019adversarial}.
+When we were working on multi-players bandits, in Autumn $2017$ and $2018$ \cite{Besson2018ALT}, we left the study of the piece-wise stationary case as a future work, and shortly after we were excited to see that three independent works tackling this question were published, in December $2018$ \cite{WeiSrivastava18Distributed} and in February $2019$ \cite{AlaturLevyKrause19,bande2019adversarial}.
 Without diving too much into the details, we review here these three works.
@@
 
+\paragraph{Extending the single-player case.}
+%
 On the one hand, piece-wise stationary multi-players MAB can be tackled by extending algorithms developed for the single-player case.
 In \cite{WeiSrivastava18Distributed}, the authors study exactly the same multi-players bandit model as our model, and they propose two distributed algorithms that can efficiently be used by $M \leq K$ players to achieve sub-linear regret for piece-wise stationary problems, essentially by considering it as a harder case of a stationary problem.
-\\
-\indent
+
 The proposed the RR-SW-UCB\# algorithm, which combines their SW-UCB\# algorithm, previously proposed in \cite{WeiSrivastava18Abruptly}, and a Round-Robin hoping, under the assumption that each user $j\in[M]$ knows its ID $j$ (we criticize this unrealistic assumption in Section~\ref{par:5:knowingYourIDinMPBanditsGame} above).
@@
 Even if their analysis is not explicit regarding the constant, in the case where $\Upsilon_T = \bigO{T^{\gamma}}$ for a known $\gamma$, their regret upper-bound actually matches the conjectured lower-bound, up to a logarithmic factor $\log(T)$.
-\\
-%
-\indent
+
 And finally, even though numerical simulations in their papers \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} are interesting and confirm the regret upper-bounds, they do not compare with any other algorithm, and it is left as an interesting future work to study this direction in more details.
@@
 
+\paragraph{In the adversarial setting.}
+%
 On the other hand, another possibility is to extend ideas developed for the adversarial setting.
@@
 The recent work \cite{bande2019adversarial} also proposes a decentralized algorithm that can achieve sub-linear regret ($\cO(T^{3/4})$) under an adversarial multi-players bandit model.
-\\
-\indent
+
 In \cite{AlaturLevyKrause19}, the authors build on the \MusicalChair{} algorithm from \cite{Rosenski16}, to let the $M$ players converge to a ranking in an efficient and decentralized way (cf.
@@
 Then, to address the no-communication constraint, we enforce the players to keep the same decisions (arms) within long periods of time (blocks).
-This gives them the chance to coordinate between themselves via a simple protocol that uses collisions as a primitive, yet effective manner of communication.\\
-%
-\indent
+This gives them the chance to coordinate between themselves via a simple protocol that uses collisions as a primitive, yet effective manner of communication.
+
 Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
@@
 Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
-However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (while scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).\\
-\indent
+However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (while scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).
+
 Moreover, at first sight, this idea of ``meta'' arms implies an exponential blow up in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
-the authors provide in Section~4.1 and Lemma~4.1 an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} for a good review).
+the authors provide in Section~4.1 and Lemma~4.1 \cite{AlaturLevyKrause19} an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} for a good review).
 %
@@
 
-% \newpage  % FIXME
+\newpage  % FIXME
 

### 2019-06-21 2-Chapters/1-Chapter/chapter1.tex
Almost done for Ch6

@@
 This family of wireless systems are referred to as Internet of Things,
-and a typical example of application is for sensor networks.\\
-\indent
+and a typical example of application is for sensor networks.
+
 For the future development of smart grids, smart cities, smart homes, or smart agriculture, sensor networks are promised to be widely deployed.
@@
 For buildings, the main goal is to reduce the cost of heating empty buildings and use temperature sensor networks to get accurate and regular data about the temperature in each room and each floor, and let the centralized heat control optimize its cost and energy consumption.
-For agriculture, one example can be to equipped every cow (in large farms) with a sensor that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and to monitor the health of the animals.\\
+For agriculture, one example can be to equipped every cow (in large farms) with a sensor that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and to monitor the health of the animals.
 %
 % OK commencer par un paragraphe qui explique qu'on étudie les télécom dans le but de les rendre automatiquement plus efficace, plus verte, plus adaptative
-\indent
+
 This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices, it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
@@
 but rather that we are analyzing algorithms which are all playing in different (random) subsets of the global synchronized time.
-Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, but leads to the model of Chapter~\ref{chapter:5}.\\
-%
-\indent
+Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, but leads to the model of Chapter~\ref{chapter:5}.
+
 On the first hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let all groups of $M \leq K$ devices learn an optimal orthogonal affectation in the $K$ channels (\eg, groups can be the set of devices that all emit at the same time).
@@
 it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid any central control of the devices by the base station, as we explained above.
-%
+
 On the other hand, another way to relax the non-stationary hypothesis is to take the point-of-view of one device, like in the OSA case mentioned above, where the focus is on one SU surrounded by many PU having stationary behaviors.

### 2019-06-21 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done proofreading quickly Ch5
Almost done for Ch6

@@
 %
-We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for realistic IoT networks.\\
+We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for realistic IoT networks.
 % \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
-\indent
+
 On the opposite, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$).
@@
     \item
-    In a realistic scenario, we can assume that $D_k\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall k\in\llbracket 1;K \rrbracket$.
+    In a realistic scenario, we can assume that $D_k\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall k\in[K]$.
 For such values for $D_k$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{k=1}^{K} D_k (1 - p)^{S_k + D_k -1}$ is concave as the sum of concave functions

### 2019-06-21 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
DONE
Almost done for Ch6

@@
 
-\textbf{Publications.}
+\paragraph{Publications.}
 %
@@
     \centering
-    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
     \caption{Problem $1$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).
@@
     \centering
-    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
     \caption{Problem $2$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).
@@
     \begin{enumerate}
-        \item Choosing $\omega = \sqrt{\ln(T)/T}$, $\delta = 1/\sqrt{T}$ gives
+        \item
+        Choosing $\omega = \sqrt{\ln(T)/T}$, $\delta = 1/\sqrt{T}$ (with no prior knowledge of $\Upsilon_T$) gives
         \begin{equation}
@@
         \end{equation}
-        \item Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ gives
+        \item
+        Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$) gives
         \begin{equation}
@@
     \begin{enumerate}
-        \item Choosing $\omega = \sqrt{\ln(T) / T}$, $\delta = 1 / \sqrt{T}$ (with no prior knowledge of $\Upsilon_T$ or $\EffChange$) gives
+        \item
+        Choosing $\omega = \sqrt{\ln(T) / T}$, $\delta = 1 / \sqrt{T}$ (with no prior knowledge of $\Upsilon_T$ or $\EffChange$) gives
         \begin{equation}
@@
 
-        \item Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``optimist'' guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
+        \item
+        Choosing $\omega = \sqrt{\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``optimist'' guess $\Upsilon_T \simeq \EffChange \ll K \Upsilon_T$) gives
         \begin{equation}
@@
 
-        \item Choosing $\omega = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ (with prior knowledge of $\EffChange$) gives
+        \item
+        Choosing $\omega = \sqrt{\EffChange\ln(T) / T}$, $\delta = 1 / \sqrt{\EffChange T}$ (with prior knowledge of $\EffChange$) gives
         \begin{equation}
@@
 
-        \item Choosing $\omega = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``pessimist'' guess $\EffChange \simeq K \Upsilon_T$) gives
+        \item
+        Choosing $\omega = \sqrt{K\Upsilon_T\ln(T) / T}$, $\delta = 1 / \sqrt{K\Upsilon_T T}$ (with prior knowledge of $\Upsilon_T$ and ``pessimist'' guess $\EffChange \simeq K \Upsilon_T$) gives
         \begin{equation}
@@
 \begin{align*}
-    & \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t)) \\
+    & R_T = \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t)) \leq \ldots \\
     & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\omega)) + \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(t \notin\cD(T,\omega), \UCB_{A(t)}(t) \geq \UCB_{k_t^*}(t)\right) \\
@@
 \[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
-which concludes the proof.m
+which concludes the proof.
 %
@@
 
+% -----------------------------------------------------------------
+\begin{figure}[h!]  % [htbp]
+    \centering
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
+    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem $2$.}
+    \label{fig:6:meanRegretPb2}
+\end{figure}
+
 On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there are more changes, we clearly observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much better than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).

### 2019-06-21 2-Chapters/Conclusion/conclusion.tex
I'll print myself a baby version!

@@
 
-\begin{itemize}
-    \item
+% \begin{itemize}
+    % \item
 The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
@@
 
-    \item
+    % \item
 The second half of this thesis, Part~\ref{part:MABIOT}, presented our main contributions.

### 2019-06-24 1-Preamble/4-Acknowledgement/acknowledgement.tex
Done for thanks

@@
 
-% \TODOL{FIXME à finir, en septembre !}
 % See https://perso.crans.org/besson/phd/thanks/
+\vspace*{-20pt}
 
-I would like to acknowledge and thank the following people, and also acknowledge the various fundings who supported our works during the last three years and made this thesis possible.
+I would like to acknowledge and thank the following people, and also acknowledge the various foundings who supported our works during the last three years and made this thesis possible.
 
-My first thoughts go to my two PhD advisors.\\
+My PhD was founded by the French Ministry of Higher Education and Research (MENESR),
+thanks to a ``contrat doctoral spécifique normalien'' obtained while I was a ``normalien'' student at \'Ecole Normale Sup\'erieure de Paris-Saclay.
+This work has also been supported by
+the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{N ANR-16-CE40-0002}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
+by CNRS under the PEPS project BIO,
+by European Union, through the European Regional Development Fund (ERDF),
+and by Ministry of Higher Education and Research,
+R\'egion Bretagne and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
+
+My thoughts then go to my two PhD advisors:\\
 %
@@
 Firstly, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in IETR lab and in CentraleSupélec campus of Rennes, a very nice and warm environment for research and day-to-day life.
-Christophe was very supportive on all aspects of the PhD work, and he was also helpful during difficult moments in the last three years.
-Many thanks for being so nice and dynamic\footnote{~We had a nice first contact when we met in Spring 2015 in Mahindra Ecole Centrale, in Hyderabad, India, where I taught for one year, and I am honored to have pursued my PhD under his supervision and to have worked with him, and I am also happy to have lived great human moments alongside Christophe since 2016.}, and always comprehensive.
+Christophe was very supportive on all aspects of the PhD work, and he was also helpful during difficult moments in the last three years. Many thanks for being so nice and dynamic,
+% \footnote{~We had a nice first contact when we met in Spring 2015 in Mahindra Ecole Centrale, in Hyderabad, India, where I taught for one year, and I am honored to have pursued my PhD under his supervision and to have worked with him, and I am also happy to have lived great human moments alongside Christophe since 2016.}
+and always comprehensive.
 Christophe was helpful for his expertise on wireless communications and engineering, and especially on cognitive radio.
@@
 \indent
-Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years\footnote{~Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{~\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.}.
+Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years.
+% \footnote{~Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{~\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.}.
 Émilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep the highest possible rigour in all parts of our exchanges and productions.
-I was lucky to be able to visit the SequeL team at Inria Lille on regular occasions, and to work in close collaboration with Émilie during these visits\footnote{~Merci à Romain et toi pour votre accueil chaleureux, et aussi merci à Florian, Nicolas et Édouard.}.
+Visiting the SequeL team at Inria Lille on regular occasions was a rich experience, and I enjoyed working in close collaboration with Émilie during these visits.
+% \footnote{~Merci à Romain et toi pour votre accueil chaleureux, et aussi merci à Florian, Nicolas et Édouard.}.
 Intensive white board sessions proved to be fruitful, giving the research work that I am the most proud of \cite{Besson2018ALT}, and rich research papers that join her rigorous expertise in statistical learning and mathematical analysis of bandit algorithms with my passion for numerical experiments, simulations and algorithm designs \cite{Besson2018ALT,Besson2018DoublingTricks,Besson2019GLRT}.
@@
 Then I would like to thank the dear members of my PhD committee.
-I was lucky to be able to discuss with them on regular occasions, and these discussions have been very useful for some of our works.
+It was interesting to be able to discuss with them on regular occasions, and these discussions have been very useful for some of our works.
 Many thanks to \href{https://sites.google.com/site/vianneyperchet/}{Vianney Perchet} and \href{http://researchers.lille.inria.fr/~mitton/}{Nathalie Mitton} for accepting to review my thesis during the summer 2019, and for their very valuable comments.
@@
 I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my studies.
-His solid and serious work kept pushing me further, some of our board games nights were brain teasers comparable with some maths problems solved for this thesis.
+His solid and serious work kept pushing me further, and I am immensely proud of him.
+Some of our board game nights were brain teasers comparable with some maths problems solved for this thesis.
 % and I was happy to help him for the organization of the \href{http://lettres.sorbonne-universite.fr/Kaamelott}{research colloque}\footnote{~Un livre a aussi été publié suite \href{http://lettres.sorbonne-universite.fr/Kaamelott}{au colloque}, \href{https://www.editions-vendemiaire.com/catalogue/a-paraitre/kaamelott-un-livre-d-histoire-florian-besson-et-justine-breton-dir/}{``Kaamelott, un livre d'histoire'', aux éditions Vandémiaire}.} on the \emph{Kaamelott} french TV show in February 2017, it was \href{https://www.fabula.org/actualites/kaamelott-ou-la-relecture-de-l-histoire_75490.php}{the funniest} and \href{https://www.vousnousils.fr/2017/03/22/kaamelott-un-tremplin-vers-lhistoire-et-la-litterature-601428}{most exciting} \href{https://lpcm.hypotheses.org/10718}{research event} I have been to in the last four years!
 %
-I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life.
-Very warm thanks then go to Marine, to my grand-parents, Jeannine, Jean-Marc and Jacqueline, my uncle Jacques and aunt Bernadette, and my cousins.
+I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life. \emph{Ton courage m'inspire quotidiennement, merci beaucoup !}
+Very warm thanks then go to Marine and my grand-parents, Jeannine \& Jean-Marc and Jacqueline, and to my uncle Jacques and aunt Bernadette.
 Finally, I thank sincerely my parents, Christophe and Patricia, for being the best parents since the last 30 years, and for their vital moral support.
-\emph{Merci infiniement de nous avoir inculqué cette curiosité et d'être si gentils, et merci de votre soutien constant et de votre amour !}
-
+\emph{Merci infiniment de nous avoir offert cette curiosité, merci d'être si gentils, et merci de votre soutien constant et de votre amour !}
+\\
+\indent
 I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
-And due to space constraints, I had to chose quite arbitrarily who to include\dots}: Florian, Fabian, Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham and Priscilla in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
+And due to space constraints, I had to chose quite arbitrarily who to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham and Priscilla in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
 % In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
@@
 
-Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline (and others) at CRIStAL in Lille.
-Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially Martine and J-B, your daily smile is priceless!
-I warmly thank David P.
-for his trust, and François, David P.
-and David C.
-at ENS de Rennes, and Romaric at ENSAI : I loved every single of the $3 \times 64$ hours I spent teaching with you in the last three years.
-I am grateful to Amélie at Lycée Chateaubriand and Gilbert at Lycée Joliot-Curie for useful exchanges.
-%
-I am thankful to the great professors who taught me while I studied at ENS de Cachan, between 2011 and 2016, and for some who pushed me into teaching, and studying statistical learning and computer science, including Frédéric, Alain, Claudine and Nicolas in the maths department, Hubert, Jean, Sylvain, Paul, Serge in the computer science department, and thanks to Carine, Delphine and the late Nicolas for their support.
-
-% TODO financial thanks
-My PhD was funded by the French Ministry of Higher Education and Research (MENESR),
-thanks to a ``contrat doctoral spécifique normalien'' obtained while I was a ``normalien'' student at \'Ecole Normale Sup\'erieure de Paris-Saclay.
-This work has been supported by
-the French National Research Agency (ANR), under the projects BADASS, SOGREEN and EPHYL (grants \emph{N ANR-16-CE40-0002}, \emph{N ANR-14-CE28-0025-02} and \emph{N ANR-16-CE25-0002-03}),
-by R\'egion Bretagne, France,
-by CNRS under the PEPS project BIO (in 2017),
-by European Union, through the European Regional Development Fund (ERDF),
-and by Ministry of Higher Education and Research, Brittany and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
-
-I would also like to thank some unusual things.
-Some people thank God in their thesis, I feel that I prefer to thank the Earth, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
-I'll try hard to do better from now on!
-I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} which gave me smiles and energy, and finally my \href{https://perso.crans.org/besson/}{favorite technological tools}: git, GitHub and Bitbucket ; \LaTeX, Bash, Julia and Python ; GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools\ldots{}
-
 In the last few years, some friends thanked me in their own PhD thesis, and it is a pleasure to do the same, and cross-link to their own PhD thesis.
@@
 Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, and Rami Othman.
+
+Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile, Maud, Frédéric, Gabriel (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline (and others) at CRIStAL in Lille.
+Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially Martine and J-B, your daily smile was priceless!
+
+During my PhD I was lucky to teach in Mathematics and Computer Science.
+I warmly thank David Pichardie for his trust, and also François Schwarzentruber, Nathalie Bertrand and David Cachera at ENS de Rennes, and Romaric Gaudel at ENSAI.
+Thanks to them, I loved all of the $3 \times 64$ hours spent teaching, since October $2016$.
+% I am trully honored to join \emph{the Département d'Informatique} at ENS de Rennes in next September $2019$.
+Warm thanks to my amazing students, I am proud of our interesting exchanges, and I feel lucky to be able to keep working alongside some of you.
+I am grateful to Amélie Stainer and Gilbert Gabillard at Lycée Chateaubriand and Joliot-Curie for useful exchanges.
+\\
+\indent
+I am thankful to amazing professors who shared their passion, first of mathematics and then of computer science:
+first, Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Marseille.
+Then at ENS de Cachan ($2011$-$16$), I especially think about Frédéric P., Alain T., Claudine P., Florian de V.
+and Nicolas V.
+in Maths, to Hubert C., Jean G., Sylvain S., Paul G., Serge H.
+in Computer Science, and to Carine S.-P., Delphine L.
+and the late Nicolas P.
+for their support.
+I also thank Didier C. and Vijaysekhar C. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$).
+
+I would also like to thank some unusual things.
+I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} which gave me smiles and energy, and finally my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git (and GitHub and Bitbucket) ; \LaTeX, Bash, Julia and Python ; GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
+Some people thank God in their thesis, I feel that I prefer to thank the Earth, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
+I'll try hard to do better from now on!
 

### 2019-06-25 2-Chapters/1-Chapter/chapter1.tex
Update and fix typos

@@
     \includegraphics[width=0.95\textwidth]{machine-learning-map.pdf}
-\caption[Classification of the main types of Machine Learning]{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.
+\caption[Classification of the main types of Machine Learning.]{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.
 MAB is RL with partial information, and RL is ML with no supervision or data, just an environment to interact with.}
@@
     }
-    \caption[Organization of the thesis: a reading map]{A reading map of the thesis.
+    \caption[Organization of the thesis: a reading map.]{A reading map of the thesis.
 Any top-down path containing Chapter~\ref{chapter:1}, Chapter~\ref{chapter:2}, at least one of the three Chapters~\ref{chapter:4}, \ref{chapter:5} and \ref{chapter:6}, and the Conclusion is a self contained way to read this thesis.}

### 2019-06-25 2-Chapters/2-Chapter/chapter2.tex
I'll print myself a baby version!
Update and fix typos

@@
     \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
-    \caption[Screenshot of the demonstration, at the end of the game after $T=100$ steps]{Screenshot of the demonstration, at the end of the game after $T=100$ steps, where the player suffers from an empirical regret of $R^{\UCB_1}_T = \mu^* T - 56 = 0.7 T - 56 = 14$ after following the $\UCB_1$ algorithm.
+    \caption[Screenshot of the demonstration, at the end of the game after $T=100$ steps.]{Screenshot of the demonstration, at the end of the game after $T=100$ steps, where the player suffers from an empirical regret of $R^{\UCB_1}_T = \mu^* T - 56 = 0.7 T - 56 = 14$ after following the $\UCB_1$ algorithm.
         You can try it on \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
@@
 
-\paragraph{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}\\
+\paragraph{The $\UCB_1$ index policy: using Hoeffding's inequality to build confidence intervals.}
 %
@@
 % 	\includegraphics[width=1.00\linewidth]{meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000.pdf}
-% 	\caption{Mean cumulated rewards as function of the current time step $t$ for horizon $T=100$ and $N=1000$ independent runs, for different algorithms }
+% 	\caption{Mean cumulated rewards as function of the current time step $t$ for horizon $T=100$ and $N=1000$ independent runs, for different algorithms .}
 % 	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT100N1000}

### 2019-06-25 2-Chapters/3-Chapter/chapter3.tex
Update and fix typos

@@
     \end{minted}
-    \caption{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits}
+    \caption{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits.}
     % \captionof{listing}{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits.
@@
     \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_a_IndexPolicy_UCB.py}
-    % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy}
+    % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy.}
     \captionof{listing}{Code defining the UCB algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
@@
 	\includegraphics[width=1.05\linewidth]{3.pdf}
-	\caption[Example of a single-player simulation showing the average regret of $4$ algorithms]{
+	\caption[Example of a single-player simulation showing the average regret of $4$ algorithms.]{
 		Example of a single-player simulation showing the average regret of four algorithms.
@@
     \end{minted}
-    \caption{Example of Bash code to run an experiment}
+    \caption{Example of Bash code to run an experiment.}
     % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}

### 2019-06-25 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Update and fix typos

@@
 	\includegraphics[width=0.70\linewidth]{Approximation_m10-eps-converted-to.pdf}
-	\caption[Proposed approximation for the probability of collision at the second transmission]{Proposed approximation for the probability of collision at the second transmission.
+	\caption[Proposed approximation for the probability of collision at the second transmission.]{Proposed approximation for the probability of collision at the second transmission.
 It is more precise for smaller values of $N$.}
@@
 	\includegraphics[width=0.80\linewidth]{ResultsUCB.eps}
-	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB'']{
+	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{
 		Comparison between the exposed heuristics for the retransmission:`` \textcolor{blue}{Only \UCB}'', ``\textcolor{cyan}{Random}'', ``\textcolor{purple}{\UCB}'', ``\textcolor{green}{$K$ \UCB}'', and ``\textcolor{red}{Delayed \UCB}''.
@@
 	\includegraphics[width=0.80\linewidth]{ResultsUCB2.eps}
-	\caption[First comparison between the exposed heuristics for the retransmission: Only \UCB, Random, \UCB, $K$ \UCB, and Delayed \UCB]{
+	\caption[Second comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random'', ``\UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{
 		Second scenario: learning helps a lot (a gain of $30\%$ in terms of collision probability), and learning to retransmit smartly is needed.

### 2019-06-25 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Update and fix typos

@@
               }
-              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
+              \caption[The \RandTopM{} decentralized learning policy.]{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
           \label{algo:5:RandTopM}
@@
               }
-              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
+              \caption[The \MCTopM{} decentralized learning policy.]{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
           \label{algo:5:MCTopM}
@@
   % \end{subfigure}
-  \caption[Failure case of \Selfish]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
+  \caption[Failure case of \Selfish.]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
 Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$).
@@
   % \end{subfigure}
-  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+  \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem.]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
 \RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand.
@@
   % \end{subfigure}
-  \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
+  \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem.]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
 In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand.
@@
   % \end{subfigure}
-  \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem]{Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
+  \caption[Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on a fixed problem.]{Regret for $M=3$ players for $K=9$ arms, horizon $T=123456$, for $100$ repetitions on problem $\mu=[0.1,\dots,0.9]$.
 With a perfect knowledge on the horizon and the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly, even in this simple setting.
@@
   % \end{subfigure}
-  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
 \rhoRand{} (top \textcolor{blue}{blue} curve) is outperformed by the other algorithms (and the gain increases with $M$).
@@
   % \end{subfigure}
-  \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+  \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
 \rhoRand{} (top \textcolor{blue}{blue}) is outperformed by the other algorithms (and the gain increases when $M$ increases), which all perform similarly in such configurations.
@@
   % \end{subfigure}
-  \caption[Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled]{Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
+  \caption[Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
 This extreme case $M=K$ shows the drastic difference of behavior between \RandTopM{} and \MCTopM, having constant regret, and \rhoRand{} and \Selfish, having large regret.}

### 2019-06-25 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Typo
Update and fix typos

@@
 % which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
-% Thus, the total number of arm switches is also proved to be logarithmic, if all players uses
+% Thus, the total number of arm switches is also proven to be logarithmic, if all players uses
 % the \MCTopM-\klUCB{} algorithm.
@@
 % %
-% As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proved
+% As soon as $R_T = \bigO{\log T}$ for all problem, \MCTopM{} is clearly proven
 % to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
@@
   \includegraphics[width=0.95\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
-  \caption[Comparison of our lower bound against the one from \cite{Zhao10}]{Comparison of our lower bound against the one from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
+  \caption[Comparison of our lower bound against the one from \cite{Zhao10}.]{Comparison of our lower bound against the one from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
   \label{fig:5:CompLowerBounds}
@@
     % \end{subfigure}
-    \caption[Illustrating the decomposition of the multi-players regret with its three terms]{The three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} of the multi-players regret, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
+    \caption[Illustrating the decomposition of the multi-players regret with its three terms.]{The three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} of the multi-players regret, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
     \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}

### 2019-06-25 1-Preamble/4-Acknowledgement/acknowledgement.tex
Lots of typos

@@
 \indent
-Firstly, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in IETR lab and in CentraleSupélec campus of Rennes, a very nice and warm environment for research and day-to-day life.
+Firstly, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in the IETR lab and CentraleSupélec campus of Rennes, very nice and warm environments for research and day-to-day life.
 Christophe was very supportive on all aspects of the PhD work, and he was also helpful during difficult moments in the last three years. Many thanks for being so nice and dynamic,
@@
 and to \href{http://rcombes.supelec.free.fr/}{Richard Combes} as well for being part of my jury.
-Some of their articles were strong inspirations for our own works, and I wish to also thank their co-authors, and in particular the Masters or PhD students, and useful discussions we had whenever we met, especially thanks to \href{https://www.cvernade.com/}{Claire Vernade}, \href{https://www.linkedin.com/in/reda-alami-067304109}{Réda Alami}, \href{https://www.researchgate.net/profile/Etienne_Boursier}{Étienne Boursier}, \href{https://www.researchgate.net/profile/Frederic_Loge_Munerel}{Frédéric Loge-Munerel} and \href{https://pratikgajane.wordpress.com/}{Pratik Gajane}, and others I could not meet (\eg, Viktor Toldov or \href{https://fr.linkedin.com/in/robin-allesiardo-phd-39221a92}{Robin Allesiardo}).
+Some of their research works were strong inspirations for our own works, and I wish to also thank their co-authors, and in particular the Masters or PhD students, and useful discussions we had whenever we met: thanks to \href{https://www.cvernade.com/}{Claire Vernade}, \href{https://www.linkedin.com/in/reda-alami-067304109}{Réda Alami}, \href{https://www.researchgate.net/profile/Etienne_Boursier}{Étienne Boursier}, \href{https://www.researchgate.net/profile/Frederic_Loge_Munerel}{Frédéric Loge-Munerel} and \href{https://pratikgajane.wordpress.com/}{Pratik Gajane}, and others I could not meet (\eg, Viktor Toldov or \href{https://fr.linkedin.com/in/robin-allesiardo-phd-39221a92}{Robin Allesiardo}).
 
@@
 His solid and serious work kept pushing me further, and I am immensely proud of him.
-Some of our board game nights were brain teasers comparable with some maths problems solved for this thesis.
+Some of our board game nights were brain teasers comparable with some maths problems solved for this thesis!
 % and I was happy to help him for the organization of the \href{http://lettres.sorbonne-universite.fr/Kaamelott}{research colloque}\footnote{~Un livre a aussi été publié suite \href{http://lettres.sorbonne-universite.fr/Kaamelott}{au colloque}, \href{https://www.editions-vendemiaire.com/catalogue/a-paraitre/kaamelott-un-livre-d-histoire-florian-besson-et-justine-breton-dir/}{``Kaamelott, un livre d'histoire'', aux éditions Vandémiaire}.} on the \emph{Kaamelott} french TV show in February 2017, it was \href{https://www.fabula.org/actualites/kaamelott-ou-la-relecture-de-l-histoire_75490.php}{the funniest} and \href{https://www.vousnousils.fr/2017/03/22/kaamelott-un-tremplin-vers-lhistoire-et-la-litterature-601428}{most exciting} \href{https://lpcm.hypotheses.org/10718}{research event} I have been to in the last four years!
@@
 I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
-And due to space constraints, I had to chose quite arbitrarily who to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham and Priscilla in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
+And due to space constraints, I had to choose quite arbitrarily whom to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham and Priscilla in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
 % In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
 I loved traveling with you and hope to continue!
-\emph{J'ai une pensée particulière pour Hélène, merci pour ton énergie et ta détermination depuis 2018.}
+\emph{Une pensée particulière pour Hélène, merci pour ton énergie et ta détermination depuis 2018.}
 
-Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali, Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali, Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Alessandro, Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
+Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali C., Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali Z., Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Alessandro, Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
 I want to thank especially the ``coffee friends$^*$'' in Rennes, for the daily interesting discussions, the best moments of our life at the office.
@@
 I would also like to thank some unusual things.
-I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} which gave me smiles and energy, and finally my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git (and GitHub and Bitbucket) ; \LaTeX, Bash, Julia and Python ; GNU/Linux and XUbuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
-Some people thank God in their thesis, I feel that I prefer to thank the Earth, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
+I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and frequent rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} which gave me smiles and energy, and finally my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git (and GitHub and Bitbucket) ; \LaTeX, Bash, Julia and Python ; GNU/Linux and (X)Ubuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
+Some people thank God in their thesis, I feel that I prefer to thank \emph{the Earth}, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
 I'll try hard to do better from now on!

### 2019-06-25 1-Preamble/5-Abstract/resume_fr.tex
Update and fix typos
Lots of typos

@@
 Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
-Nous signalons les conférences internationales auxquelles ont été publiées certains résultats.
+Nous signalons les conférences nationales et internationales auxquelles ont été publié certains résultats.
 
@@
     Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
-    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences utilisées dans la suite de cette thèse.
+    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences et simulations numériques utilisées dans cette thèse.
 
@@
     Nous présentons le problème du choix de l'algorithme qu'un praticien devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithms MAB disponibles, dans le Chapitre~\ref{chapter:25}.
-    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par quelques simulations numériques qu'il atteint de bonnes performances empiriques
-    (WCNC 2018) \cite{Besson2018WCNC}.
+    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par des simulations numériques qu'il atteint de bonnes performances empiriques
+    (publié à WCNC 2018) \cite{Besson2018WCNC}.
 
@@
 Cela permet à davantage de périphériques d'utiliser le même réseau tout en maintenant une haute qualité de service
-    (CROWNCOM 2017, ICT demo 2018, WCNC 2019 et MOTIoN 2019)
-    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC}.
+    (publiés à CROWNCOM 2017, ICT demo 2018, WCNC 2019 et MOTIoN 2019)
+    \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
 
     \item
-    Nous avons implémenté une preuve de concept du modèle susmentionné \cite{Besson2018ICT}, et nous la présentons en détail en Section~\ref{sec:4:gnuradio}.
+    Nous avons implémenté une démonstration réaliste du modèle susmentionné \cite{Besson2018ICT,Besson2019WCNC}, et nous la présentons en détail en Section~\ref{sec:4:gnuradio}.
 Nous avons aussi réalisé une courte vidéo présentant notre démonstration, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
@@
     \item
-    Nous formalisons le modèle du bandit multi-joueur, pour lequel nous avons introduit trois variantes, dans le Chapitre~\ref{chapter:5}.
-    Pour le cas avec information de détection (``sensing''), nous proposons deux nouveaux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM, qui prouve qu'il est asymptotiquement optimal.
+    Nous formalisons le modèle du bandit multi-joueur, pour lequel nous avons introduit deux variantes, dans le Chapitre~\ref{chapter:5}.
+    Pour le cas avec information de détection (``sensing''), nous proposons deux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM, qui prouve qu'il est asymptotiquement optimal.
     Nous présentons aussi des expériences numériques approfondies pour montrer qu'il est bien plus efficace que les autres algorithmes de la littérature précédente.
-    Notre travail \cite{Besson2018ALT} (ALT 2018) a également contribué à une nouvelle impulsion à la recherche sur les bandits multi-joueurs, car certains travaux de recherche récents se sont construits sur nos résultats.
+    Notre travail \cite{Besson2018ALT}, publié à ALT 2018, a également contribué à une nouvelle impulsion à la recherche sur les bandits multi-joueurs, car certains travaux de recherche récents se sont construits sur nos résultats.
 
@@
     Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (GRETSI 2019).
-    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif pour ce problème stationnaire par morceaux, \GLRklUCB, qui atteint des performances comparables à l'état de l'art.
+    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif pour ce problème stationnaire par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles .
     % Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB
@@
     \item
-    La dernière contribution de cette thèse est une revue de la littérature sur les cas d'utilisation possibles de la technique du doublement successif de l'horizon (``doubling trick'') pour les problèmes de bandits,
+    La dernière contribution de cette thèse est notre article \cite{Besson2018DoublingTricks}, résumé en Annexe~\ref{app:2:DoublingTricks}.
+    Nous donnes une revue de la littérature sur l'utilisation de la technique de doublements successifs de l'horizon (``doubling trick'') pour les problèmes de bandits,
     ainsi qu'une analyse unifiée et plus générique de deux familles de doublements.
-    Ceci a conduit à l'article \cite{Besson2018DoublingTricks}, qui est rapidement présenté en Annexe~\ref{app:2:DoublingTricks}.
 \end{itemize}
@@
     }
-    \caption[Organisation de la thèse : une carte de lecture]{Organisation de la thèse.
+    \caption[Organisation de la thèse : une carte de lecture.]{Organisation de la thèse.
 Cette thèse peut se lire en suivant n'importe quel chemin contenant le Chapitre~\ref{chapter:1}, le Chapitre~\ref{chapter:2}, au moins un des trois Chapitres~\ref{chapter:4}, \ref{chapter:5} et \ref{chapter:6}, et la Conclusion.}
@@
 
-L'ordre de lecture du manuscrit peut être n'importe quel chemin, descendant entre l'introduction donnée dans Chapitre~\ref{chapter:1}, et la conclusion générale qui constitue le dernier Chapitre~\ref{chapter:conclusion}.
+L'ordre de lecture du manuscrit peut être n'importe quel chemin, entre l'introduction donnée dans Chapitre~\ref{chapter:1}, et la conclusion générale qui constitue le dernier Chapitre~\ref{chapter:conclusion}.
 Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,
@@
 Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
-Par contre le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
-Nous terminons cette première partie par le Chapitre~\ref{chapter:25}, qui n'est pas non plus nécessaire à la compréhension de la suite du manuscrit, et qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
+Par contre, le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
+Nous terminons cette première partie par le Chapitre~\ref{chapter:25}, qui n'est pas non plus nécessaire à la compréhension du manuscrit, et qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
 
@@
 \textcolor{darkgreen}{La deuxième Partie~\ref{part:MABIOT}} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus presque indépendamment.
-Le Chapitre~\ref{chapter:4} commence par présenter différents modèles de réseaux IdO où les algorithmes MAB ont été utilisés avec succès.
-Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop généraux pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
+Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux IdO, dans lesquels les algorithmes MAB ont été utilisés avec succès.
+Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop complexes pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
 Pour cette raison, nous affaiblissons les modèles pour le reste du document,
-et les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
+et les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
 \end{itemize}

### 2019-06-25 2-Chapters/1-Chapter/chapter1.tex
Lots of typos

@@
 % Where
-\textbf{Where and when?}
+% \textbf{Where and when?}
 %
@@
 
-\TODOL{Je peux améliorer (ou enlever) ce passage qui site J.Toledano, ou alors le commenter en plus de détails ? J'aime bien son rapport en fait (j'aurai du le lire au début de ma thèse !)}
-% https://www.economie.gouv.fr/files/files/PDF/french-spectrum-mission-executive-summary-2014-06-25.pdf
-% https://www.economie.gouv.fr/files/files/PDF/rapport-gestion-dynamique-spectre-2014-06-30.pdf
-To highlight that the spectrum scarcity is not only an issue in the USA,
-we quote the abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by J.
+% \TODOL{Je peux améliorer (ou enlever) ce passage qui site J.Toledano, ou alors le commenter en plus de détails ? J'aime bien son rapport en fait (j'aurai du le lire au début de ma thèse !)}
+% % https://www.economie.gouv.fr/files/files/PDF/french-spectrum-mission-executive-summary-2014-06-25.pdf
+% % https://www.economie.gouv.fr/files/files/PDF/rapport-gestion-dynamique-spectre-2014-06-30.pdf
+% To highlight that the spectrum scarcity is not only an issue in the USA,
+% we quote the abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by J.
 Toledano \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull}, who was commissioned by the French government in $2014$.
-Different important problems are mentioned:
-%
-\begin{itemize}%\tightlist
-    \item
-    \emph{``In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy.''}
-    And the usage of wireless communications is nowadays worldwide and omnipresent:
-    \emph{``Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.''}
-    Thus it is of highest interest to set up efficient, green, low-cost and durable new technologies for the future wireless standards, networks and devices.
-    %
-    \item
-    The radio spectrum is not only widely used for all kinds of applications, it is nowadays intensely used for some of them.
-    \emph{``Everyone agrees on the growing need for spectrum.
-    This growing need results from two phenomena.
-    On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
-    On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.''}
-    The first estimation was correct, and the mobile traffic is again \href{https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth}{estimated} to be increased by 10 times between 2019 and 2025.
-    % https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth
-    % https://www.cisco.com/c/en/us/solutions/collateral/service-provider/visual-networking-index-vni/white-paper-c11-738429.html
-    If the second estimation was overestimating the growth of the IoT market, forecasts predict  more than 64 billion IoT devices by 2025.
-    % https://www.businessinsider.com/internet-of-things-report?IR=T
-    \item
-    \emph{``Today, there are no more available frequencies in the easily exploitable frequency bands.
-    Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.''}
-    As illustrated above in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}, and discussed in \cite{patil2011survey,valenta2010survey}.
-    \item
-    \emph{``Growing recourse to spectrum sharing, and in particular dynamic spectrum sharing, constitutes an important spectrum reserve.
-    A particular sharing form, the use of unlicensed bands, open to all and free, has been intensely developed with the growth of Wi-Fi use.''}
-    This thesis focusses on spectrum sharing in unlicensed bands, for IoT networks that lie at the opposite direction as the one underlying Wi-Fi networks: the spectrum sharing will be learned automatically by the end-devices, and not centrally decided by the base station.
-\end{itemize}
+% Different important problems are mentioned:
+% %
+% \begin{itemize}%\tightlist
+%     \item
+%     \emph{``In France, digital economy represents $5\%$ of GDP and affects $80\%$ of the French economy.''}
+%     And the usage of wireless communications is nowadays worldwide and omnipresent:
+%     \emph{``Radio frequencies are essential to many sectors: communications, broadcasting, transportation, satellite networks, energy networks and smart grids, public or private security, defense, etc.''}
+%     Thus it is of highest interest to set up efficient, green, low-cost and durable new technologies for the future wireless standards, networks and devices.
+%     %
+%     \item
+%     The radio spectrum is not only widely used for all kinds of applications, it is nowadays intensely used for some of them.
+%     \emph{``Everyone agrees on the growing need for spectrum.
+%     This growing need results from two phenomena.
+%     On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
+%     On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.''}
+%     The first estimation was correct, and the mobile traffic is again \href{https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth}{estimated} to be increased by 10 times between 2019 and 2025.
+%     % https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth
+%     % https://www.cisco.com/c/en/us/solutions/collateral/service-provider/visual-networking-index-vni/white-paper-c11-738429.html
+%     If the second estimation was overestimating the growth of the IoT market, forecasts predict  more than 64 billion IoT devices by 2025.
+%     % https://www.businessinsider.com/internet-of-things-report?IR=T
+%     \item
+%     \emph{``Today, there are no more available frequencies in the easily exploitable frequency bands.
+%     Moreover, it is getting more and more difficult to resort to classical methods of frequency bands liberation.''}
+%     As illustrated above in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum}, and discussed in \cite{patil2011survey,valenta2010survey}.
+%     \item
+%     \emph{``Growing recourse to spectrum sharing, and in particular dynamic spectrum sharing, constitutes an important spectrum reserve.
+%     A particular sharing form, the use of unlicensed bands, open to all and free, has been intensely developed with the growth of Wi-Fi use.''}
+%     This thesis focusses on spectrum sharing in unlicensed bands, for IoT networks that lie at the opposite direction as the one underlying Wi-Fi networks: the spectrum sharing will be learned automatically by the end-devices, and not centrally decided by the base station.
+% \end{itemize}
 

### 2019-06-25 2-Chapters/Conclusion/conclusion.tex
Lots of typos

@@
 This manuscript was organized in two parts, Part~\ref{part:Introduction} and Part~\ref{part:MABIOT}.
-It presented the problems we considered for my PhD research, and we discussed about the main contributions produced during the last three years.
+%
+The main problem we studied was \emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
+%
+This document detailed the historical, scientific and technical contexts of the problems we considered for our research, and presented the main contributions produced during the last three years.
 
@@
 %
-Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an efficient answer to the online algorithm selection problem.
+Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an empirically efficient answer to the online algorithm selection problem.
 
@@
 % Chapter 5
-Because the models studied for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
+Because the models proposed for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
 in Chapter~\ref{chapter:5}, we studied a weaker but still interesting model.
@@
 % Chapter 6
-Finally, in Chapter~\ref{chapter:6}, we removed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
+Finally, in Chapter~\ref{chapter:6}, we relaxed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
 %
@@
 %
-SMPyBandits allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
-SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
+Our library allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
+SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license (MIT License), and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
 % Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.
@@
 We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
-Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions (as well as rested or restless Markov models).
+Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions (and rested/restless Markov models).
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
-One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
+One of the main qualities of our library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
 % We indeed followed actively the state-of-the-art research, since December $2016$ to June $2019$.
@@
 %
-Our library is used to run numerical simulations on Chapters~\ref{chapter:2}~ \ref{chapter:5} and~\ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
+Our library is used to run numerical simulations on Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
 
@@
 Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
-Similarly, if the number of devices remain constant, increasing the successful transmission rate directly improves the network QoS.
+Another way of viewing this result is the opposite point-of-view: if the number of devices remains constant, increasing the successful transmission rate directly improves the network QoS.
 %  as the base station receives more accurately the data sent by the end-devices since less transmissions failed.
-We show that both \UCB{} and Thompson sampling algorithms have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which follow a random activation process.
+We show that both \UCB{} and Thompson sampling algorithms have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which all follow a random activation process.
 %
-This is both surprising and encouraging, it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
+This was both surprising and encouraging, as it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
 The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.
@@
 %
-By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using onlt the acknowledgement (\Ack) feedback sent from the base station.
+By using such low-cost algorithms, we confirmed empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using only the acknowledgement (\Ack) feedback sent from the base station.
 
-Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency and with perfect synchronisation.
+Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of Low-Power Wide-Area (LPWA) networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency, and assuming a perfect synchronisation.
 % , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
 If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
-We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
+We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for the first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
 The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a multi-armed bandit algorithm for this second-stage optimization problem.
@@
 Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
-Surprisingly, the main take-away message is that a simple \UCB{} learning approach (that simply retransmits in the same channel), turns out to perform almost as well as more complicated heuristics.
+Surprisingly, the simulations showed that a simple \UCB{} learning approach, that simply retransmits in the same channel, turns out to perform almost as well as more complicated heuristics, that can change the channel used for retransmissions.
 
 Thus in Chapter~\ref{chapter:4}, we proposed two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
-that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
+that can all use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
 %
-Decentralized RL for IoT lets the devices use acknowledgements sent by their Base Station as a reward, instead of sensing feedback like for OSA.
+Decentralized RL for IoT lets the devices use the acknowledgement sent by their Base Station as a reward, instead of requiring some sensing capacity, like in the OSA case.
 We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
 The first model (without retransmission of packets),
-is interesting for its simplicity, and because considering no retransmission can improve the battery life of the IoT devices, at the cost of a lower Quality of Service (QoS).
+is interesting for its simplicity, and because considering no retransmission can improve the battery life of the IoT devices.
+% , at the cost of a lower Quality of Service (QoS).
 %
@@
 with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
-For the two easiest models --with sensing--, our theoretical contribution improves the state-of-the-art upper bounds on the regret.
+For the two easiest models --with sensing, our theoretical contribution improves the state-of-the-art upper bounds on the regret.
 In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
+\Selfish{} corresponds to the heuristic used in the first model studied in Chapter~\ref{chapter:4}.
 %
@@
 For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
-For example, it assumes to know the number of players $M$ before-hand, but we illustrated that previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
+For example, it assumes to know the number of players $M$ before-hand, but we illustrated that a previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
 % Further works would be required to adapt the theoretical analysis to these various extensions.
@@
 %
-We first describe the break-point detection problem, that we solve by analyzing at finite-time the properties of the Bernoulli GLR change-point detector.
+We first describe the break-point (or change-point) detection problem, that we solve by analyzing at finite-time the properties of the Bernoulli GLR change-point detector.
 This sequential break-point detection test can be applied for bounded rewards, and we bound its false alarm probability as well as its detection delay.
@@
 
-As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on finite-arms and one-dimensional bandit problems,
-and thus two possible directions of future research could be to extend our works
+As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on MAB problems with a finite number of arms, and binary, bounded or real-valued rewards.
+Thus two possible directions of future research could be to extend our works
 to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
@@
 %
-A second interesting task could be to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
+A second interesting task could be to extend the library for problems with non-finite arms, such as linear or contextual bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
 or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
@@
 %
-Moreover, for sake of simplicity we supposed that all devices use the same standard.
+Moreover, for the sake of simplicity we supposed that all devices use the same standard.
 Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
@@
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
-However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}.
+However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}, like a recent work \cite{wang2019distributed} studied the minimal amount of communication needed to achieve logarithmic regret, in a similar model that authorizes direct communications between players.
 % Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
@@
 
-We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected.
-This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$.
+We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
+This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of the number of break-points $\Upsilon_T$.
 We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
-We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
+We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of more efficient algorithms, as suggested by \cite{Maillard2018GLR}.
 

### 2019-06-25 1-Preamble/5-Abstract/resume_fr.tex
Lots of typos on Ch1 etc

@@
     \item
-    Nous avons écrit la plus complète bibliothèque de simulation pour les problèmes MAB, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBandits,SMPyBanditsJMLR}.
+    Nous avons écrit la plus complète bibliothèque de simulation pour les problèmes MAB, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBanditsJMLR,SMPyBandits}.
     Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.

### 2019-06-25 2-Chapters/1-Chapter/chapter1.tex
Lots of typos on Ch1 etc

@@
 % https://en.wikipedia.org/wiki/History_of_radio
-During the last 150 years, mankind has developed a wide range of different communication technologies, and since the late $1890$s wireless communications between manufactured devices have been made possible, and more and more frequent in our lives.
+During the last 150 years, mankind has developed a wide range of different communication technologies, and since the late $1890$s wireless communications between man-made manufactured devices have been made possible, and more and more frequent in our lives.
 With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, for a large range of different applications.
-It is now well-known that with the current trend of population increase and with the energy crisis, any newly deployed technology should be cheap and energy efficient,
+It is now a world-wide consensus that with the current trend of population increase and with the energy crisis, any newly deployed technology should be both cheap and energy efficient,
 as well as adapted to serve a large number of people and devices.
@@
 That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms (Multi-Armed Bandit algorithms),
-directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing the wireless communication thanks to embedded decentralized decision making.
+directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing their wireless communications and learning to be self-organized, thanks to embedded decentralized decision making.
 
@@
 Different standards have been developed, and some of them are still in use today, and their nature requires a central authority that allocates RF bands to different usages (\eg, TV channels etc).
-For example, in the city of Cesson-Sévigné where the campus of CentraleSupélec is located, a high RF tower emits on many standards, and in particular the \SI{92.3}{\mega\hertz} band is allocated to a French classical music radio (\emph{Radio Classique}).
+For example, in the city of Cesson-Sévigné where my campus of CentraleSupélec is located, a high RF tower emits on many standards, and in particular the \SI{92.3}{\mega\hertz} band is allocated to a French classical music radio (\emph{Radio Classique}).
 Broadcast-based systems are not only used for TV or radio,
@@
 However, this kind of mono-directional wireless communications were soon found to not be satisfactory for many applications, and thus bi-directional standards were defined from the $1910$s.
-First developed by and for the military, and the navy (\eg, the Titanic sent the first SOS signal over radio in $1912$),
+First developed by and for the military and the navy (\eg, the Titanic sent the first SOS signal over radio in $1912$),
 such applications have reached general public usage since the $1950$s.
-The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people lives in the richer countries in the world, with the booming of cellphones since the $2000$s and smartphones since the $2010$s.
-In most systems, there is still an antenna in charge of a large area (or cell), and many devices are able to receive and transmit data to the antenna.
+The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people's lives in the richer countries in the world, with the booming of cellphones since the $2000$s and smartphones since the $2010$s.
+In most systems, there is still a base station (with one or more antenna) in charge of a large area (or cell), and many devices are able to receive and transmit data to the base station.
 Nowadays, the most well-known systems include WiFi, and systems deployed for mobile phone communications, with a long list of standards from 1G, 2G, 3G, 4G and now 5G, still in development in $2019$.
@@
 Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance of the wireless communications to and from the devices.
-More recently, green radio has emerged as a solution to the tradeoff between maximizing QoS while minimizing the power consumption of base stations and devices.
+More recently, green radio has emerged as a solution to the tradeoff between maximizing QoS of the system, while minimizing the power consumption of base stations and devices.
 
@@
 Finally, a third kind of systems are of the opposite kind, and can be designed as decentralized broadcasting: an antenna is still in charge of many devices, but such devices can only send up-link packets, and the only down-link data they can receive are short acknowledgements sent from the base-station to indicate success or failure of every up-link packet.
-This family of wireless systems are referred to as Internet of Things,
+This family of wireless systems are referred to as Internet of Things (IoT),
 and a typical example of application is for sensor networks.
@@
 Two examples of future applications that are already in deployment, in France or other countries, are connected buildings and connected agriculture.
-For buildings, the main goal is to reduce the cost of heating empty buildings and use temperature sensor networks to get accurate and regular data about the temperature in each room and each floor, and let the centralized heat control optimize its cost and energy consumption.
+For buildings, the main goal is to reduce the cost of heating empty buildings (at night etc) and use temperature sensor networks to get accurate and regular data about the temperature in all rooms and floors, and let the centralized heat control optimize its cost and energy consumption.
 For agriculture, one example can be to equipped every cow (in large farms) with a sensor that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and to monitor the health of the animals.
@@
 This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices, it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
-In the present and future IoT networks, many devices of heterogeneous nature are using the same antenna for different applications.
-A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and run on a battery, which duration should be maximized (typically more than 10 years).
-Another common constraint for IoT devices is their low duty cycle, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
+In the present and future IoT networks, many devices of heterogeneous natures are using the same antenna for different applications.
+A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and run on a tiny battery, which lifespan should be maximized, as typically commercial IoT companies sell a lifespan more than 10 years \cite{Centenaro16}.
+Another common constraint for IoT devices is their low duty cycles, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
 %
@@
 
-\textbf{The spectrum scarcity issue.}
+\textbf{Spectrum Scarcity.}
 %
@@
 Moreover, the fixed spectrum allocation prevents rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
-Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands if they would not cause any interference to licensed users.
+Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands, if they would not cause any interference to the (paying) licensed users.
 These initiatives have initiated the research on \textbf{Cognitive Radio},
-and especially on one subfield that we detail below, \textbf{Dynamic Spectrum Access (DSA)}.
+and especially on a field that we detail below, \textbf{Dynamic Spectrum Access (DSA)}.
 %
@@
 % To highlight that the spectrum scarcity is not only an issue in the USA,
-% we quote the abstract of the ``Dynamic Spectrum Management For Innovation And Growth'' report by J.
+% we quote the abstract of the ``Dynamic Spectrum Management For Innovation and Growth'' report by J.
 Toledano \cite{Toledano2014EnglishSummary,Toledano2014FrenchFull}, who was commissioned by the French government in $2014$.
@@
 %     This growing need results from two phenomena.
-%     On the one hand, mobile traffic should be multiplied by 13 to 25 fold between 2011 and 2017.
-%     On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to 50 billion in 2020, according to estimations.''}
-%     The first estimation was correct, and the mobile traffic is again \href{https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth}{estimated} to be increased by 10 times between 2019 and 2025.
+%     On the one hand, mobile traffic should be multiplied by $13$ to $25$ fold between $2011$ and $2017$.
+%     On the other hand, the development of new innovative services such as the Internet of Things and its multiple applications (smart cities, e-health...), could lead to a growing number of connected devices, up to $50$ billion in $2020$, according to estimations.''}
+%     The first estimation was correct, and the mobile traffic is again \href{https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth}{estimated} to be increased by $10$ times between $2019$ and $2025$.
 %     % https://www.ericsson.com/en/mobility-report/future-mobile-data-usage-and-traffic-growth
@@
 In other words, we focus on RL with limited feedback, when the player only sees the reward given by its action at every round, and not the reward that would have been given if she chose any other action.
-This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} in details in the next Chapter~\ref{chapter:2}.
+This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} (MAB) in details in the next Chapter~\ref{chapter:2}.
 Clinical trials and treatment discovery were historically the first applications of MAB since $1930$s, with the early work of W. Thompson \cite{Thompson33}.
 %
-Multi-Armed Bandits (MAB) are a simple yet powerful example of the well-known exploration/exploitation dilemma:
-when facing a set of $K$ actions whose effects on the environment are unknown, a learning must balance between
+MAB are a simple yet powerful example of the well-known exploration/exploitation dilemma:
+when facing a set of $K$ actions whose effects on the environment are unknown, a learner must balance between
 exploration of the unknown actions, in order to collect more information about them,
-and exploitation of the best action according to its current knowledge.
+and exploitation of the best action, according to its current knowledge.
 %
@@
 Similarly, we can narrow down the second field of study of this thesis, steps by steps:
-from Information Technologies (IT), to telecommunications, to wireless communications, then to Software Defined Radio (SDR),
+from Information Technologies (IT), to Telecommunications, to Wireless Communications, then to Software Defined Radio (SDR),
 and finally to Cognitive Radio (CR).
@@
 % https://en.wikipedia.org/wiki/Software-defined_radio
-A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are instead implemented by means of software on a personal computer or embedded system, with the exception of a unique pair of receiver/transmitter antennas.
+A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are all instead implemented by means of software on a personal computer or an embedded system, with the exception of a unique pair of receiver/transmitter antennas.
 Even though the SDR paradigm was initiated by the USA defense research, over the past years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
@@
     \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
-    He was the first to propose to use CR for dynamic spectrum access.
+    He was the first to propose to use CR for Dynamic Spectrum Access.
 
@@
 If a SU was able to sense for all the $K$ frequency bands, it could simply transmit in one of the free channels if any, or not transmit if all channels are used at a given time step.
-However, sensing is known to be error-prone and costly, especially for wide-band sensing (for more details, see the surveys \cite{yucek2009survey,subhedar2011spectrum}), thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
-This assumption, along with the regulation that PU cannot be disturbed, enforces the SU to transmit in the channel that it sensed.
+However, sensing is known to be error-prone and costly, especially for wide-band sensing, as detailed in the surveys \cite{yucek2009survey,subhedar2011spectrum}, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
+This assumption, along with the regulation that PU cannot be disturbed, enforces the SU to transmit in the channel that it sensed, in each time step, if and only if it was sensed free from any PU.
 % non negligible cost of radio reconfiguration
 
-Focussing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of $K\geq2$ channels), then it performs sensing, and finally it transmits in this channel if it is free.
+Focussing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of $K\geq2$ channels), then it performs sensing, and finally it transmits in this channel if it was sensed free.
 The goal of the SU is to minimize its energy consumption (we remind that we focus on green radio) and to maximize its uplink data rate, or equivalently, to maximize its number of successful transmission.
@@
 Shortly after in $2014$, proof-of-concepts using real-world radio hardware and Software Defined Radio were developed by C. Moy and his student C. Robert \cite{RobertSDR2014,MoyWSR2014}.
-In a second PhD thesis \cite{Modi17PhD}, N. Modi studied from $2014$ to $2017$ the impact of using MAB algorithms to optimize channel selection on the battery life of a wireless device.
+In a second PhD thesis \cite{Modi17PhD}, N. Modi studied from $2014$ to $2017$ the impact of using MAB algorithms to optimize channel selections, on the battery life of a wireless device.
 On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
-On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
-Each hardware reconfiguration costs energy for the device, and quickly switching algorithms will lead to a reduction of the battery life.
+On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfigurations for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
+Each hardware reconfiguration costs energy for the device, and quickly-switching algorithms lead to a reduction of the battery life.
 The tradeoff between the two aspects is studied empirically in
@@
 %
-For example, proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
+For example, proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA.
 %
-Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
+Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning are appropriate, and solve efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
 %
@@
 %
-For more details on the entire research on Cognitive Radio, we refer to the surveys \cite{garhwal2012survey,marinho2012cognitive}.
+For more details on the state of research on Cognitive Radio, we refer to the surveys \cite{garhwal2012survey,marinho2012cognitive}.
 
@@
     \item
-    IoT devices have low duty cycles (only a few messages every minute) to very low duty cycles (only a few messages a day).
+    IoT devices have low duty cycles (only a few messages every minute) to very low duty cycles (only a few messages every day).
 \end{itemize}
@@
 %
-By using this feedback, that consists in an \emph{acknowledgement} (\Ack) message received shortly after every successful transmission, or in an absence of \Ack{} avery a failed transmission, it is possible that an IoT end-device can also use RL algorithms to optimize its communications.
+By using this feedback, that consists in an \emph{acknowledgement} (\Ack) message received shortly after every successful transmission, or in an absence of \Ack{} after a failed transmission, it is possible that an IoT end-device can also use RL algorithms to optimize its communications.
 %
-To the best of our knowledge, this direction of research has never been studied before the year $2016$ and the beginning of this PhD.
+% To the best of our knowledge, this direction of research has never been studied before the year $2016$ and the beginning of this PhD.
 
@@
 
-\paragraph{Problematic.}
+\textbf{Problematic.}
 %
@@
 we were not only interested by the usual measure of performance of any MAB algorithm (its regret, see below in Section~\ref{sec:2:lowerUpperBoundsRegret}),
-but also by their empirical performances in terms of computational complexity and storage requirements, both from the point of view of the theoretical analyses and real-world measurements on time and memory footprints.
+but also by their empirical performances in terms of computational complexity and storage requirements, from the point of view of both theoretical analyses and real-world measurements on time and memory footprints.
 
@@
 %
-We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR} and hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
+We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBanditsJMLR,SMPyBandits} and hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
 We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
@@
 We study two models of IoT networks in Chapter~\ref{chapter:4}.
-The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use the acknowledgement as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the less occupied by the surrounding traffic).
+The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use acknowledgements as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the less occupied by the surrounding traffic).
 The goal of this first model of applying decentralized RL on the device side is to increase the Quality of Service (QoS) of the application of such IoT network, by reducing the rate of failed transmission.
@@
 This model constitutes the first article written during this thesis, and initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
-We presented this first work at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, and obtained the ``best paper award'' \cite{Bonnefoi17}.
+We presented this first work at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, and it obtained the ``best paper award'' \cite{Bonnefoi17}.
 
@@
 In Section~\ref{sec:4:firstModel}, we detail what we believe to be the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network.
-We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements, that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
+We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements.
+% , that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
 % Verify this first model empirically: proof of concept
@@
 We then consider an extension of our first model, following the idea of packet retransmissions, that is at the root of the famous ALOHA protocol \cite{Abramson1970,Roberts75}.
-The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode and waiting for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, $m \leq 10$ times), after some small random waiting times.
+The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode in order to wait for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, $m \leq 10$ times), after some small random waiting times.
 %
@@
 %
-In both models presented above, the core concept is to let every dynamic device of a IoT network runs a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
+In both models presented above, the core concept is to let every dynamic device of an IoT network runs a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
 It means that each of them target its own local objective, which is to maximize its cumulated reward (in a selfish way).
@@
 That is why in Chapter~\ref{chapter:5}, we relax the hypothesis of having $M \gg K$ (or even simply $M > K$) devices in a network with $K$ orthogonal wireless channel, equivalent to having $p < 1$, and we consider the case of devices communicating at every time step (\ie, $p=1$), and so we restrict to only up-to $M \leq K$ devices.
-The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing information) or the IoT case (\ie, without sensing feedback).
+The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing) or the IoT case (\ie, without sensing).
 The goal was to understand the heuristic of Chapter~\ref{chapter:4}, \Selfish, in the simpler framework of multi-player MAB, which has been studied before for the case of OSA, as studied for instance by \cite{Zhao10,Anandkumar10,Anandkumar11}.
@@
 Perchet \cite{BoursierPerchet18}.
-Even if their approach is costly empirically, the result is surprising and very interesting: they showed that even under the strong hypothesis of no direct communication between players and no central coordination, the players can use the collision feedback (\ie, receiving or not an \Ack) as an indirect way to communicate bits of information with each other.
+Even if their approach is costly empirically, the result is surprising and very interesting.
+They showed that even under the strong hypothesis of no direct communication between players and no central coordination (in the same model as the one of our paper), the players can actually use the collision feedback (\ie, receiving or not an \Ack) as an indirect way to communicate bits of information with each other.
 This idea requires that devices coordinate and split the time steps between ``listening'' and ``talking'' phases, and the authors denote it the ``communication trick''.
@@
 On the other hand, another way to relax the non-stationary hypothesis is to take the point-of-view of one device, like in the OSA case mentioned above, where the focus is on one SU surrounded by many PU having stationary behaviors.
-If we focus on one IoT device, its environment (\ie, the surrounding devices) is non-stationary, meaning that its average properties can fluctuate with time, but we could assume a certain structure on this non-stationarity.
-For instance, enforcing stationarity on some time intervals leads to the last contribution and chapter of this thesis.
+If we focus on one IoT device, its environment (\ie, the surrounding devices) is non-stationary, meaning that its average properties can fluctuate with time, but realistic environments usually yield a certain structure on this non-stationarity.
+For instance, enforcing stationarity on some time intervals leads to the last contribution and the last chapter of this thesis.
 % DONE explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
@@
 Following two recent previous works \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, we combine an efficient index policy (\klUCB) and an efficient change-point detection test, under the assumption of bounded rewards.
-We build on very recent results on the Generalized Likelihood Ratio test (GLRT) for Gaussian and sub-Gaussian variables \cite{Maillard2018GLR}, and we instead focus on bounded rewards and Bernoulli distributions.
+We build on very recent results on the Generalized Likelihood Ratio test (GLRT) for Gaussian and sub-Gaussian variables \cite{Maillard2018GLR}, but we instead focus on bounded rewards and Bernoulli distributions.
 Bounded rewards are indeed usually more appropriate for Cognitive Radio applications.
@@
 We prove that by combining the asymptotically optimal policy \klUCB, designed for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of our proposed algorithm.
-Like our competitors, we make the some hypothesis on the length of the stationary intervals, and the best regret upper-bound is obtained when the algorithm knows before-hand the horizon and the number of change-points, but our algorithm does not need to know any other knowledge of the problem difficulty to be tuned optimally.
+We consider the same hypothesis as our competitors, by assuming that the length of the stationary intervals to be ``long enough'' for the change-points to be detectable.
+The best regret upper-bound is obtained when the algorithm knows before-hand the horizon and the number of change-points, but our algorithm does not need to have any other knowledge of the problem difficulty to be tuned optimally.
 %
-The performance of \GLRklUCB{} is also illustrated on numerical experiments on synthetic data, where it is shown to outperforms all the passively adaptive policies as well as the previous actively adaptive policies.
+The performance of \GLRklUCB{} is also illustrated on numerical experiments on synthetic data, where it is shown to outperform all the passively adaptive policies as well as the previous actively adaptive policies.
 %
-This last chapter is based on our last work, since October $2018$, and it first lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
-Our work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results and submitted soon to a journal (probably the Journal of Machine Learning Research) in fall $2019$.
+This last chapter is based on our last work, and it first lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
+Our work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results, in order to submit it soon to a journal (probably the Journal of Machine Learning Research) before December $2019$.
 
@@
 
-This manuscript ends with the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks} we discuss about our work on doubling-trick, in order to remove the need for an algorithm to know the horizon of the bandit game before starting to play it.
+This manuscript ends with the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks}.
+We discuss about our work on doubling-trick, in order to remove the need for an algorithm to know the horizon of the bandit game before starting to play it.
 The rest of the appendix contains a list of abbreviations and notations, then lists of figures, algorithms, code samples and tables,
@@
     \item
-    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
+    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence (MIT) \cite{SMPyBanditsJMLR,SMPyBandits}.
     We present it in details in Chapter~\ref{chapter:3}.
@@
     See Chapter~\ref{chapter:3}.
-    \cite{SMPyBandits,SMPyBanditsJMLR}
+    \cite{SMPyBanditsJMLR,SMPyBandits}
 

### 2019-06-25 2-Chapters/Conclusion/conclusion.tex
Lots of typos on Ch1 etc

@@
 
-Then we presented our Python library SMPyBandits \cite{SMPyBandits,SMPyBanditsJMLR}, by detailing
+Then we presented our Python library SMPyBandits \cite{SMPyBanditsJMLR,SMPyBandits}, by detailing
 its purpose and qualities, its organization, and an overview of its usage.
@@
 The three parts should be connected, the same way we built \GLRklUCB{} by connecting both components.
-We propose the idea of incorporating the detected change-points by B-\GLR{} test in the orthogonalization procedure: after any change-point is detected, the player is not sitting anymore.
+We propose the idea of incorporating the detected change-points by B-\GLR{} test in the orthogonalization procedure \MCTopM: after any change-point is detected, the player is not sitting anymore.
 Analyzing this strategy is left as a future work.

### 2019-06-25 3-Appendices/2-Appendix/appendix2.tex
Lots of typos on Ch1 etc

@@
 We give two examples to illustrate where this prior knowledge can be realistic or not.
-On the one hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for instance we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
+On the one hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for this model we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
 On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices (like we present it in Chapter~\ref{chapter:4}): usually a learning step corresponds to an uplink and then downlink message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages.
-While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies), many kinds of application such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.
+While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies \cite{Centenaro16}), many kinds of application such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.
 
@@
 while we proposed in Chapter~\ref{chapter:5} any-time algorithms solving the presented problem of stationary multi-players bandit, our solution for the problem of non-stationary MAB problem studied in Chapter~\ref{chapter:6} does rely on a prior knowledge of the horizon $T$.
-It is interesting to know what is the best approach to fix this weakness: work more and design algorithms that are inherently any-time, or use a generic technique to avoid depending on a prior knowledge of $T$, and automatically transform our non-anytime approach to make it anytime.
+It is an interesting direction of research to know what is the best approach to fix this weakness: work more and design algorithms that are inherently any-time, or use a generic technique to avoid depending on a prior knowledge of $T$, and automatically transform our non-anytime approach to make it anytime.
 
@@
 to generalize previously known results that certain doubling tricks can be used to conserve certain regret bounds.
-In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log T)$.
-We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log T)$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
+In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log(T))$.
+We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log(T))$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
 Interestingly, we prove that they conserve bounds of the mixed form $R_T = \cO(T^{\gamma} (\log(T))^{\delta})$, for $0<\gamma<1$ and $\delta>0$, and so an exponential doubling trick could be used to obtain an anytime version of the algorithm we propose for non-stationary bandits, \GLRklUCB{} in Chapter~\ref{chapter:6}.
@@
 
-To conclude this work, we would need to complete the study of the doubling tricks, and instead of focussing on two families (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bound.
+To conclude this work, we would need to complete the study of the doubling tricks, and instead of focussing on two families (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bounds.
 Such doubling is given either by the function mapping the current time $t$ to the current estimate of the horizon $T(t)$, or the sequence $(T_i)_{i\in\N^*}$ of successive estimates of the horizons.

### 2019-06-25 2-Chapters/1-Chapter/chapter1.tex
Thanks to Emilie's comments

@@
 %
-At the root of the problems that motivate this thesis are the two questions of global warming, and the increase of the world population.
+At the root of the problems that motivate this thesis are the questions of global warming and the increase of the world population.
 % https://en.wikipedia.org/wiki/History_of_radio
@@
 Two examples of future applications that are already in deployment, in France or other countries, are connected buildings and connected agriculture.
-For buildings, the main goal is to reduce the cost of heating empty buildings (at night etc) and use temperature sensor networks to get accurate and regular data about the temperature in all rooms and floors, and let the centralized heat control optimize its cost and energy consumption.
+For buildings, the main goal is to reduce the cost of heating empty buildings and use temperature sensor networks to get accurate and regular data about the temperature in all rooms and floors, and let the centralized heat control optimize its cost and energy consumption.
 For agriculture, one example can be to equipped every cow (in large farms) with a sensor that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and to monitor the health of the animals.
@@
 % Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
-More precisely, in this thesis we focus on a certain kind of Reinforcement Learning models where the learner does not have access to the entire reaction of the environment after taking its action.
+More precisely, in this thesis we focus on Reinforcement Learning models, in particular the learner does not have access to the entire reaction of the environment after taking its action.
 In other words, we focus on RL with limited feedback, when the player only sees the reward given by its action at every round, and not the reward that would have been given if she chose any other action.
@@
 We start by formulating the problem studied in this thesis, and then we develop our approach.
-
-\textbf{Problematic.}
+%
+% \textbf{Problematic.}
 %
@@
 % \subsection{Focusing on MAB algorithms}
-\subsection{Exploring the literature of MAB algorithms, Part~\ref{part:Introduction}}
+% \subsection{Exploring the literature of MAB algorithms}
+\subsection{Exploring the Jungle of MAB algorithms}
 
@@
 We started by exploring the rich literature of multi-armed bandits,
-as many different algorithms exist, with lots of variants on the simple problem presented above (see \cite{LattimoreBanditAlgorithmsBook,Slivkins2019} for surveys).
+as many different algorithms exist, with lots of variants on the simple problem presented above.
+% (see \cite{LattimoreBanditAlgorithmsBook,Slivkins2019} for surveys).
 We thus start the first Part~\ref{part:Introduction} of this thesis with Chapter~\ref{chapter:2}, which presents the MAB model and gives a review of the most important algorithms designed to solve stochastic and stationary MAB problems.
@@
 has given us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explores the performances of newly proposed algorithms.
-To answer this goal, we produced a Python library of simulation of MAB problems.
+To tackle this goal, we produced a Python library of simulation of MAB problems.
 %
@@
 % \subsection{Focusing on IoT networks.}
-\subsection{Our model of IoT network and decentralized MAB, Part~\ref{part:MABIOT}}
+\subsection{Our model of IoT network and decentralized MAB}
 
@@
 %
-In both models presented above, the core concept is to let every dynamic device of an IoT network runs a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
+In both models presented above, the core concept is to let every dynamic device of an IoT network run a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
 It means that each of them target its own local objective, which is to maximize its cumulated reward (in a selfish way).
-It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure: for instance, one can think of the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma} or other popular ``dilemmas''.
+It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure, one can think of popular ``dilemmas'', such as the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma}.
 % , and for more examples we refer to the book \cite{CesaLugosi06}.
 % https://www.ii.uni.wroc.pl/~lukstafi/pmwiki/uploads/AGT/Prediction_Learning_and_Games.pdf
-So it was quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices, despite the fact that these IoT objects cannot communicate directly with each other and receive only a collision feedback from the base station they are associated to.
+So it was quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices, despite the fact that these IoT objects cannot communicate directly with each other, and receive only a collision feedback from the base station they are associated to.
 
@@
 %
-and it relaunched the research on this model, as a few independent teams started to work on different questions shortly after its publication.
+and since then, many work have studied extension of our framework or alternative algorithms, as a few independent teams started to work on different questions shortly after its publication.
 %
@@
 This idea requires that devices coordinate and split the time steps between ``listening'' and ``talking'' phases, and the authors denote it the ``communication trick''.
-As they showed, after a long enough initialization period, the $M$ players have elected a ``master'' player and $M-1$ ``followers'' players (with high probability), and thanks to this communication trick, the problem can boil down to a centralized multiple-play (MP) MAB, which can be solved efficiently with the MP generalization of \klUCB{} \cite{Luedtke16} or Thompson sampling \cite{Komiyama15}.
+As they showed, the players can start by a ``musical chair'' phase, to reach an orthogonal allocation, and then they can alternate exploration, communication and exploitation phases until the end of the game at horizon $T$.
+The communication trick is used by each player to send and receive observations about all arms from the other players, using one forced collision to send one bit at a time.
+They use about $T-\bigO{\log(T)}$ rounds in exploitation of the best arms, and only $\bigO{\log(T)}$ rounds in exploration and communication, yielding a logarithmic regret.
+% As they showed, after a long enough initialization period, the $M$ players have elected a ``master'' player and $M-1$ ``followers'' players (with high probability),
+Another direction, as followed by \cite{Bistritz18}, is to use this communication trick and reduce the problem to a centralized multiple-play (MP) MAB, which can be solved efficiently with the MP generalization of \klUCB{} \cite{Luedtke16} or Thompson sampling \cite{Komiyama15}.
 
@@
 
-\paragraph{Conclusion.}
+\paragraph{Appendix.}
 %
-We conclude this thesis in Chapter~\ref{chapter:conclusion}, to give some perspectives and directions about possible future works.
-Most of the contributions that we present in this thesis call for more developments, on both directions of providing more mathematical analysis and implementing new proof-of-concepts, using real wireless hardware and realistic IoT networks.
-
+% We conclude this thesis in Chapter~\ref{chapter:conclusion}, to give some perspectives and directions about possible future works.
+% Most of the contributions that we present in this thesis call for more developments, on both directions of providing more mathematical analysis and implementing new proof-of-concepts, using real wireless hardware and realistic IoT networks.
+%
 This manuscript ends with the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks}.
@@
 
-\newpage % WARNING
+% \newpage % WARNING
 
-% ----------------------------------------------------------------------------
-\subsection{Summary of the contributions}
-\label{sec:1:summaryOfContributions}
+% % ----------------------------------------------------------------------------
+% \subsection{Summary of the contributions}
+% \label{sec:1:summaryOfContributions}
 
-To summarize the main contributions of this thesis, we can list the following points:
+% To summarize the main contributions of this thesis, we can list the following points:
 
-\begin{itemize}
-    % \item
-    % We present the concepts and the notations of the multi-armed bandit problem in Chapter~\ref{chapter:2}, from a mathematical point-of-view.
-    % But we also follow a didactic approach as we use an online interactive demonstration designed to let anyone play against a small bandit problem from his/her browser, in Section~\ref{par:2:interactiveDemoDiscoverMAB}.
+% \begin{itemize}
+%     % \item
+%     % We present the concepts and the notations of the multi-armed bandit problem in Chapter~\ref{chapter:2}, from a mathematical point-of-view.
+%     % But we also follow a didactic approach as we use an online interactive demonstration designed to let anyone play against a small bandit problem from his/her browser, in Section~\ref{par:2:interactiveDemoDiscoverMAB}.
+
+%     % \item
+%     % We give a short literature review of stochastic bandit algorithms in Section~\ref{sec:2:famousMABalgorithms}.
 
 %     \item
-    % We give a short literature review of stochastic bandit algorithms in Section~\ref{sec:2:famousMABalgorithms}.
+%     We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence (MIT) \cite{SMPyBanditsJMLR,SMPyBandits}.
+%     We present it in details in Chapter~\ref{chapter:3}.
 
-    \item
-    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence (MIT) \cite{SMPyBanditsJMLR,SMPyBandits}.
-    We present it in details in Chapter~\ref{chapter:3}.
+%     \item
+%     % We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+%     We present in Chapter~\ref{chapter:25} an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
+%     \cite{Besson2018WCNC}.
+%     % with \textbf{Aggregator} (WCNC 2018)
 
-    \item
-    % We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
-    We present in Chapter~\ref{chapter:25} an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
-    \cite{Besson2018WCNC}.
-    % with \textbf{Aggregator} (WCNC 2018)
+%     \item
+%     We propose different models for IoT networks, in Chapter~\ref{chapter:4}, where end-devices with cognitive radio capabilities can implement MAB algorithms on their side, to automatically increase their battery life and allow more devices to use the same network while maintaining a high Quality of Service
+%     \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019}.
+%     % (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
 
-    \item
-    We propose different models for IoT networks, in Chapter~\ref{chapter:4}, where end-devices with cognitive radio capabilities can implement MAB algorithms on their side, to automatically increase their battery life and allow more devices to use the same network while maintaining a high Quality of Service
-    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019}.
-    % (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
-
-    \item
-    We implemented a proof-of-concept of the aforementioned model \cite{Besson2018ICT}, and we present it in details in Section~\ref{sec:4:gnuradio}.
+%     \item
+%     We implemented a proof-of-concept of the aforementioned model \cite{Besson2018ICT}, and we present it in details in Section~\ref{sec:4:gnuradio}.
 We made a video showcasing our demonstration, hosted at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
 
-    % \item
-    % The source code for the two previously mentioned contributions are all published online, along with clear instructions for reproducing our work.
-
-    \item
-    We present the multi-players bandit model, for which we introduce three variants, in Chapter~\ref{chapter:5}.
-    For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
-    as well as extensive numerical experiments to demonstrate its excellent performance in comparison with the rest of the literature.
-    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works heavily built up on our results.
+%     % \item
+%     % The source code for the two previously mentioned contributions are all published online, along with clear instructions for reproducing our work.
 
 %     \item
-    % We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
-    % % were  (ALT 2018, and 4-8 works inspired by our article since then).
+%     We present the multi-players bandit model, for which we introduce three variants, in Chapter~\ref{chapter:5}.
+%     For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
+%     as well as extensive numerical experiments to demonstrate its excellent performance in comparison with the rest of the literature.
+%     Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works heavily built up on our results.
+
+%     % \item
+%     % We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
+%     % % were  (ALT 2018, and 4-8 works inspired by our article since then).
 State-of-the-art with our algorithm MCTopM + klUCB for multi-players bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"
 
-    \item
-    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non-stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
-    Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
-    % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
+%     \item
+%     We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non-stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
+%     Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
+%     % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
 
-    \item
-    The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
-    and a unified and more generic analysis of two families of doubling tricks.
-    This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
-\end{itemize}
+%     \item
+%     The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
+%     and a unified and more generic analysis of two families of doubling tricks.
+%     This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
+% \end{itemize}
 

### 2019-06-26 2-Chapters/1-Chapter/chapter1.tex
Lots of things to do, oh my god.

@@
 \label{chapter:1}
-
-\TODOL{Mardi soir : je dois encore tout relire et améliorer un peu !}
 
@@
 
-% \TODOL{Est-ce que je dois détailler plus ? Quel financement j'ai eu (CDSN), quels autres financements m'ont aidés ?}
-
 % ----------------------------------------------------------------------------
@@
 %
-At the root of the problems that motivate this thesis are the questions of global warming and the increase of the world population.
+The root of the problems motivating this thesis are the questions of global warming and the increase of the world population.
 % https://en.wikipedia.org/wiki/History_of_radio
-During the last 150 years, mankind has developed a wide range of different communication technologies, and since the late $1890$s wireless communications between man-made manufactured devices have been made possible, and more and more frequent in our lives.
+During the last $150$ years, mankind has developed many different communication technologies, and since the late $1890$s, wireless communications between man-made manufactured devices have been made possible, and more and more frequent in our lives.
 With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, for a large range of different applications.
@@
 as well as adapted to serve a large number of people and devices.
+%
+Such IoT technologies should be able to adapt automatically to different environments and application contexts, and be as efficient as possible.
+Instead of devoting a lot of time, energy and money on Research and Development to set up efficient radio access schemes, for all possible cases, a cheap and promising direction is to use Machine Learning to (try to) attain the same performance gain automatically.
 
 That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms (Multi-Armed Bandit algorithms),
-directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing their wireless communications and learning to be self-organized, thanks to embedded decentralized decision making.
+directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing their wireless communications and by learning to be self-organized, thanks to embedded decentralized decision making.
 
@@
 In the present and future IoT networks, many devices of heterogeneous natures are using the same antenna for different applications.
-A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and run on a tiny battery, which lifespan should be maximized, as typically commercial IoT companies sell a lifespan more than 10 years \cite{Centenaro16}.
+A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and run on a tiny battery, which lifespan should be maximized, as typically commercial IoT companies sell a lifespan more than $10$ years \cite{Centenaro16}.
 Another common constraint for IoT devices is their low duty cycles, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
@@
 
-Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are insufficiently utilized.
-Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place.
-Moreover, the fixed spectrum allocation prevents rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
-Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands, if they would not cause any interference to the (paying) licensed users.
-These initiatives have initiated the research on \textbf{Cognitive Radio},
-and especially on a field that we detail below, \textbf{Dynamic Spectrum Access (DSA)}.
-%
-%
 \begin{figure}[h!]
@@
 \end{figure}
+
+Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are insufficiently utilized.
+Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place.
+Moreover, the fixed spectrum allocation prevents rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
+Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands, if they would not cause any interference to the (paying) licensed users.
+These initiatives have initiated the research on \textbf{Cognitive Radio},
+and especially on a field that we detail below, \textbf{Dynamic Spectrum Access (DSA)}.
 
@@
 
-\paragraph{From Machine Learning to Multi-Armed Bandits.}
+\paragraph{From Statistics or Machine Learning to Multi-Armed Bandits.}
 %
 We can narrow down the first field of study of this thesis, steps by steps.
-Starting from the general concept of Machine Learning (ML), we focused our interest to statistical learning, then sequential learning, then Reinforcement Learning (RL), then RL with limited feedback and finally to Multi-Armed Bandits (MAB) learning.
+% Stats > sequential learning > bandits
+A first point-of-view is the historical one, as Multi-Armed Bandits (MAB) learning emerged as an interesting problem in the statistics community and among researchers interested by sequential learning, as shown by the earlier works of \cite{Thompson33,Robbins52,LaiRobbins85}.
+% ML > RL > MAB
+We can also expose another point-of-view, which starts from the general concept of Machine Learning (ML), and focus on Reinforcement Learning (RL), then on MAB learning.
 As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
+%
 The reference book on RL is \cite{SuttonBarto2018}, so let us quote the definition of RL given by R. Sutton and A. Barto:
@@
 
-We illustrate this idea of a \emph{learning cycle} alternating between actions and feedback,
+We illustrate this idea of a \emph{learning cycle} alternating between action and feedback,
 in the following Figure~\ref{fig:1:ReinforcementLearningCycle}.
@@
 % Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
-More precisely, in this thesis we focus on Reinforcement Learning models, in particular the learner does not have access to the entire reaction of the environment after taking its action.
-In other words, we focus on RL with limited feedback, when the player only sees the reward given by its action at every round, and not the reward that would have been given if she chose any other action.
+As this thesis focus on Reinforcement Learning models,
+it is important to highlight that in such decision making models, the learner does not have access to the entire reaction of the environment after taking its action.
+%
+In other words, the player only sees the reward given by its action at every round, and not the reward that would have been given if she chose any other action.
 This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} (MAB) in details in the next Chapter~\ref{chapter:2}.
@@
 % \subsection{Focusing on IoT networks.}
-\subsection{Our model of IoT network and decentralized MAB}
+\subsection{Our models of IoT networks and decentralized MAB}
 
@@
 
-On the first hand, in the OSA setting, we failed to obtain positive result for the Selfish policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
+On the one hand, in the OSA setting, we failed to obtain positive result for the Selfish policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
 This result was later confirmed and analyzed by \cite{BoursierPerchet18}.
@@
 
-On the first hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let all groups of $M \leq K$ devices learn an optimal orthogonal affectation in the $K$ channels (\eg, groups can be the set of devices that all emit at the same time).
+On the one hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let all groups of $M \leq K$ devices learn an optimal orthogonal affectation in the $K$ channels (\eg, groups can be the set of devices that all emit at the same time).
 %
@@
 \label{sec:1:listPublications}
+
+\TODOL{Guy Carrault me dit : Sinon les publications du doctorant sont indiquées en début de document, je ne me rappelle plus si elles ne doivent pas figurées à la fin}
 
@@
     % \footnote{~IEEE DCOSS 2019, see \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}},
-    Santorini, Greece, May 2019.
+    Santorini, Greece, May $2019$.
     % \href{https://HAL.Inria.fr/hal-02144465}{\texttt{HAL.Inria.fr/hal-02144465}}.
@@
     % \footnote{~MOTIoN 2019, see \href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}},
-    at \emph{IEEE WCNC}, Marrakech, Morocco, April 2019.
+    at \emph{IEEE WCNC}, Marrakech, Morocco, April $2019$.
     % \href{https://HAL.Inria.fr/hal-02049824}{\texttt{HAL.Inria.fr/hal-02049824}}.
@@
     % \footnote{~IEEE WCNC 2019, see \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}},
-    Marrakech, Morocco, April 2019,
+    Marrakech, Morocco, April $2019$,
     % \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
@@
     % \footnote{~ALT 2018, see \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}},
-    Lanzarote, Spain, April 2018,
+    Lanzarote, Spain, April $2018$,
     % \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
@@
     % \footnote{~IEEE WCNC 2018, see \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}},
-    Barcelona, Spain, April 2018,
+    Barcelona, Spain, April $2018$,
     % \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
@@
 \begin{center}
-    \textbf{Copyright 2016-2019, \copyright ~Lilian~Besson.}
+    \textbf{Copyright $2016$-$2019$, \copyright ~Lilian~Besson.}
 \end{center}

### 2019-06-26 2-Chapters/Conclusion/conclusion.tex
Lots of things to do, oh my god.

@@
 %
-This document detailed the historical, scientific and technical contexts of the problems we considered for our research, and presented the main contributions produced during the last three years.
+We detailed the historical, scientific and technical contexts of the problems we considered for our research, and presented the main contributions produced during the last three years.
 
@@
 Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
-The notations used in the manuscript were introduced by showcasing an interactive demonstration\footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
+The notations used in the manuscript were introduced by showcasing an interactive demonstration.
+% \footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
 
@@
 If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
-We presented and evaluated several heuristics that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for the first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
-The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a multi-armed bandit algorithm for this second-stage optimization problem.
+We presented and evaluated heuristics that learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for the first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
+The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a MAB algorithm for this second-stage optimization problem.
 %
@@
 For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
-For example, it assumes to know the number of players $M$ before-hand, but we illustrated that a previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
+For example, it assumes to know the number of players $M$ before-hand, but we illustrated that a previously introduced technique to estimate $M$ can also be applied successfully to our proposal.
+%  and give satisfactory empirical performances.
 % Further works would be required to adapt the theoretical analysis to these various extensions.

### 2019-06-27 2-Chapters/1-Chapter/chapter1.tex
Changed order of two parts. And send for first printed version!

@@
 
+\paragraph{From IT to Cognitive Radio.}
+%
+Similarly, we can narrow down the second field of study of this thesis, steps by steps:
+from Information Technologies (IT), to Telecommunications, to Wireless Communications, then to Software Defined Radio (SDR),
+and finally to Cognitive Radio (CR).
+%
+The transition from the historical approach of hardware-based radio to SDR architectures and is a gradual process that started in the early $1990$s and has accelerated in the $2000$s.
+% https://en.wikipedia.org/wiki/Software-defined_radio
+A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are all instead implemented by means of software on a personal computer or an embedded system, with the exception of a unique pair of receiver/transmitter antennas.
+Even though the SDR paradigm was initiated by the USA defense research, over the past years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
+% CR has been the first answer to solve spectrum scarcity issue which collected a huge attention in early 21st century.
+%
+CR is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
+% - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
+%
+% je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
+\begin{itemize}
+    \item
+    Joseph Mittola in $1999$ said that
+    \emph{``a really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment''} \cite{Mitola99}.
+
+    \item
+    Then Simon Haykin in $2005$ also said that
+    \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
+    He was the first to propose to use CR for Dynamic Spectrum Access.
+
+    \item
+    The \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia encyclopedia} states
+    \emph{``a CR is a radio that can be programmed and configured dynamically to use the best wireless channels in its vicinity to avoid user interference and congestion.
+Such a radio automatically detects available channels in wireless spectrum, then accordingly changes its transmission or reception parameters to allow more concurrent wireless communications in a given spectrum band at one location''}.
+\end{itemize}
+
+% explain PU and SU in a few sentences
+In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the network.
+As exposed in the definition above, the CR philosophy suggests that if the network is not used by any PU in a certain location and time, then non-paying users, referred to as Secondary User (SU), could use it.
+Regulations state that PU have a strict priority over any other SU,
+but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU is equipped with an efficient sensing capacity, it can analyze its environment, and use a licensed band if it is free of any PU.
+% explain DSA
+This defines the concept of DSA, proposed in the early $2000$s, and we refer to the surveys \cite{akyildiz2006next,Zhao07,garhwal2012survey} for more details.
+After a short paragraph on Reinforcement Learning, we then focus on one example, the \textbf{Opportunistic Spectrum Access} (OSA) problem.
+
 \paragraph{From Statistics or Machine Learning to Multi-Armed Bandits.}
@@
 We can also expose another point-of-view, which starts from the general concept of Machine Learning (ML), and focus on Reinforcement Learning (RL), then on MAB learning.
-As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
+% As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
 %
@@
 
-% Modified from the https://draw.io/ map from
-% https://github.com/trekhleb/homemade-machine-learning/blob/master/images/machine-learning-map.xml
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.95\textwidth]{machine-learning-map.pdf}
-    \caption[Classification of the main types of Machine Learning.]{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.
+% % Modified from the https://draw.io/ map from
+% % https://github.com/trekhleb/homemade-machine-learning/blob/master/images/machine-learning-map.xml
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.95\textwidth]{machine-learning-map.pdf}
+%     \caption[Classification of the main types of Machine Learning.]{Classification of the main types of Machine Learning: MAB is a subset of Reinforcement Learning and RL is a subset of ML.
 MAB is RL with partial information, and RL is ML with no supervision or data, just an environment to interact with.}
-    \label{fig:1:MainTypesOfMachineLearning}
-\end{figure}
+%     \label{fig:1:MainTypesOfMachineLearning}
+% \end{figure}
 
@@
 Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}, and the wide range of applications of MAB in the recent years.
-
-\paragraph{From IT to Cognitive Radio.}
-%
-Similarly, we can narrow down the second field of study of this thesis, steps by steps:
-from Information Technologies (IT), to Telecommunications, to Wireless Communications, then to Software Defined Radio (SDR),
-and finally to Cognitive Radio (CR).
-%
-The transition from the historical approach of hardware-based radio to SDR architectures and is a gradual process that started in the early $1990$s and has accelerated in the $2000$s.
-% https://en.wikipedia.org/wiki/Software-defined_radio
-A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are all instead implemented by means of software on a personal computer or an embedded system, with the exception of a unique pair of receiver/transmitter antennas.
-Even though the SDR paradigm was initiated by the USA defense research, over the past years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
-% CR has been the first answer to solve spectrum scarcity issue which collected a huge attention in early 21st century.
-%
-CR is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
-% - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
-%
-% je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
-\begin{itemize}
-    \item
-    Joseph Mittola in $1999$ said that
-    \emph{``a really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment''} \cite{Mitola99}.
-
-    \item
-    Then Simon Haykin in $2005$ also said that
-    \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
-    He was the first to propose to use CR for Dynamic Spectrum Access.
-
-    \item
-    The \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia encyclopedia} states
-    \emph{``a CR is a radio that can be programmed and configured dynamically to use the best wireless channels in its vicinity to avoid user interference and congestion.
-Such a radio automatically detects available channels in wireless spectrum, then accordingly changes its transmission or reception parameters to allow more concurrent wireless communications in a given spectrum band at one location''}.
-\end{itemize}
-
-% explain PU and SU in a few sentences
-In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the network.
-As exposed in the definition above, the CR philosophy suggests that if the network is not used by any PU in a certain location and time, then non-paying users, referred to as Secondary User (SU), could use it.
-Regulations state that PU have a strict priority over any other SU,
-but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU is equipped with an efficient sensing capacity, it can analyze its environment, and use a licensed band if it is free of any PU.
-% explain DSA
-This defines the concept of DSA, proposed in the early $2000$s, and we refer to the surveys \cite{akyildiz2006next,Zhao07,garhwal2012survey} for more details.
-We focus on one example, the \textbf{Opportunistic Spectrum Access} (OSA) problem.
 

### 2019-07-02 1-Preamble/4-Acknowledgement/acknowledgement.tex
Proofreading intro

@@
 
-I would like to acknowledge and thank the following people, and also acknowledge the various foundings who supported our works during the last three years and made this thesis possible.
+I would like to acknowledge and thank the following people, and also the various foundings which supported our works during the last three years and made this thesis possible.
 
 My PhD was founded by the French Ministry of Higher Education and Research (MENESR),
-thanks to a ``contrat doctoral spécifique normalien'' obtained while I was a ``normalien'' student at \'Ecole Normale Sup\'erieure de Paris-Saclay.
+thanks to a ``contrat doctoral spécifique normalien'' obtained while I was a ``normalien'' student at \'Ecole Normale Sup\'erieure de Cachan (now ENS de Paris-Saclay).
 This work has also been supported by
@@
 by European Union, through the European Regional Development Fund (ERDF),
-and by Ministry of Higher Education and Research,
-R\'egion Bretagne and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
+% and by the Ministry of Higher Education and Research,
+and by R\'egion Bretagne and Rennes Métropole, through the CPER Project \emph{SOPHIE / STIC \& Ondes}.
 
@@
 \indent
-Firstly, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in the IETR lab and CentraleSupélec campus of Rennes, very nice and warm environments for research and day-to-day life.
-Christophe was very supportive on all aspects of the PhD work, and he was also helpful during difficult moments in the last three years. Many thanks for being so nice and dynamic,
+First, thanks to \href{https://MoyChristophe.Wordpress.com/}{Christophe Moy} who welcomed me in the SCEE team, in the IETR lab and in CentraleSupélec campus of Rennes, a nice and warm environment for research and day-to-day life.
+Christophe was very supportive on every aspect of my work as a PhD student, and he was also helpful during difficult moments in the last three years. Many thanks for being so nice and dynamic,
 % \footnote{~We had a nice first contact when we met in Spring 2015 in Mahindra Ecole Centrale, in Hyderabad, India, where I taught for one year, and I am honored to have pursued my PhD under his supervision and to have worked with him, and I am also happy to have lived great human moments alongside Christophe since 2016.}
-and always comprehensive.
+and always comprehensive, especially regarding my teaching activities.
 Christophe was helpful for his expertise on wireless communications and engineering, and especially on cognitive radio.
-Regardless of what we did together, whether it was debugging a difficult GNU Radio block \cite{Besson2018ICT,Besson2019WCNC}, writing a paper by night while Christophe was in a plane coming back from India \cite{MoyBesson2019}, chatting at a barbecue or playing ``palets'' together along the ocean in Vannes, I was lucky to collaborate with and be around Christophe.\\
+Regardless of what we did together, whether it was debugging a GNU Radio block \cite{Besson2018ICT,Besson2019WCNC}, writing a paper by night while Christophe was in a plane coming back from India \cite{MoyBesson2019}, chatting at a barbecue or playing ``palets bretons'' together along the ocean in Vannes, I was lucky to collaborate with and be around Christophe.\\
 %
 \indent
-Secondly, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervised my thesis, and did an amazing job in the last four years.
+Then, thanks to \href{http://chercheurs.lille.inria.fr/ekaufman/research.html}{Émilie Kaufmann}, who accepted to co-supervise my thesis, and who did an amazing job in the last four years.
 % \footnote{~Since our first discussions, in Autumn 2015 when she was giving practical sessions for the \href{http://researchers.lille.inria.fr/~lazaric/Webpage/MVA-RL_Course15.html}{Reinforcement Learning course} of \href{http://researchers.lille.inria.fr/~lazaric/}{Alessandro Lazaric} at MVA\footnote{~\href{http://math.ens-paris-saclay.fr/version-francaise/formations/master-mva/}{Master Mathématiques Vision Apprentissage} at \href{https://www.ens-paris-saclay.fr/}{ENS de Paris-Saclay}, previously named \href{https://www.ens-cachan.fr}{ENS de Cachan}.}, to the last repetitions of my PhD presentation in October 2019.}.
-Émilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep the highest possible rigour in all parts of our exchanges and productions.
+Émilie was highly motivated by our collaboration, and she always pushed me further, to explore the mathematical aspects of my PhD work, to keep faith in the research world, and to keep a strong rigour in every part of our exchanges and productions.
 Visiting the SequeL team at Inria Lille on regular occasions was a rich experience, and I enjoyed working in close collaboration with Émilie during these visits.
 % \footnote{~Merci à Romain et toi pour votre accueil chaleureux, et aussi merci à Florian, Nicolas et Édouard.}.
-Intensive white board sessions proved to be fruitful, giving the research work that I am the most proud of \cite{Besson2018ALT}, and rich research papers that join her rigorous expertise in statistical learning and mathematical analysis of bandit algorithms with my passion for numerical experiments, simulations and algorithm designs \cite{Besson2018ALT,Besson2018DoublingTricks,Besson2019GLRT}.
+Our intensive white board sessions proved to be fruitful, giving the research work that I am the most proud of \cite{Besson2018ALT}, and rich research papers that join her rigorous expertise in statistical learning and mathematical analysis of bandit algorithms with my passion for numerical experiments, simulations and algorithm designs \cite{Besson2018ALT,Besson2018DoublingTricks,Besson2019GLRT}.
 
@@
 I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my studies.
-His solid and serious work kept pushing me further, and I am immensely proud of him.
+His rigour, dedication and seriousness kept pushing me further, and I am immensely proud of him.
 Some of our board game nights were brain teasers comparable with some maths problems solved for this thesis!
@@
 %
-I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life. \emph{Ton courage m'inspire quotidiennement, merci beaucoup !}
-Very warm thanks then go to Marine and my grand-parents, Jeannine \& Jean-Marc and Jacqueline, and to my uncle Jacques and aunt Bernadette.
-Finally, I thank sincerely my parents, Christophe and Patricia, for being the best parents since the last 30 years, and for their vital moral support.
+I also thank my younger brother F., for being an inspiration for other (highly important) directions of my life. \emph{Ton courage et ta force m'inspirent quotidiennement, merci beaucoup !}
+Very warm thanks then go to M. and my grand-parents, J. \& J.-M. and J., and to my uncle J. and aunt B.!
+Finally, I thank sincerely my parents, C. and P., for being the best parents since the last $26$ years, and for their (vital) moral support.
 \emph{Merci infiniment de nous avoir offert cette curiosité, merci d'être si gentils, et merci de votre soutien constant et de votre amour !}
@@
 \indent
-I would like to thank my closest friends, sorted in chronological order of first encounter in the tall grass\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
-And due to space constraints, I had to choose quite arbitrarily whom to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham and Priscilla in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
+I would then like to thank my closest friends, sorted in chronological order of our first encounter in the \href{https://bulbapedia.bulbagarden.net/wiki/Tall_grass}{tall grass}\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
+And due to space constraints, I had to choose quite arbitrarily whom to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
 % In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
@@
 
-Thanks to my wonderful colleagues during these three years: Rémi$^*$, Pascal$^*$, Yves, Jacques, Ali C., Haïfa, Marwa, Navik, Quentin$^*$, Vincent$^*$, Rami$^*$, Muhammad, Cristo$^*$, Adrien$^*$, Esteban$^*$, Ali Z., Éloïse$^*$, Nabil$^*$, Bastien$^*$, Corentin$^*$, Georgios, Morgane$^*$ in Rennes ; and Alessandro, Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
-I want to thank especially the ``coffee friends$^*$'' in Rennes, for the daily interesting discussions, the best moments of our life at the office.
+Thanks to my wonderful colleagues during these three years: Rémi\textsuperscript{\tiny{\Coffeecup}}, Pascal\textsuperscript{\tiny{\Coffeecup}}, Yves, Amor, Carlos Faouzi, Jacques P., Jacques W., Ali C., Haïfa, Marwa, Navik, Quentin\textsuperscript{\tiny{\Coffeecup}}, Vincent\textsuperscript{\tiny{\Coffeecup}}, Rami\textsuperscript{\tiny{\Coffeecup}}, Muhammad, Cristo\textsuperscript{\tiny{\Coffeecup}}, Adrien\textsuperscript{\tiny{\Coffeecup}}, Esteban\textsuperscript{\tiny{\Coffeecup}}, Ali Z., Éloïse\textsuperscript{\tiny{\Coffeecup}}, Nabil\textsuperscript{\tiny{\Coffeecup}}, Bastien\textsuperscript{\tiny{\Coffeecup}}, Corentin\textsuperscript{\tiny{\Coffeecup}}, Georgios, Morgane\textsuperscript{\tiny{\Coffeecup}} in Rennes ; and Alessandro, Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
+I want to thank especially the ``coffee friends\textsuperscript{\tiny{\Coffeecup}}'' in Rennes, for the daily interesting discussions.
+%  the best moments of our life at the office.
 %
@@
 %,
-\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{thesis}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{thesis}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{~Que je remercie deux fois plus que ce qu'il me remerciant dans sa propre thèse !
-Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{sis}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{thesis}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{thesis}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{thesis}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{thesis}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{thesis}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, and Rami Othman.
+\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis link $\nearrow$}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{$\nearrow$}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{$\nearrow$}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{~Que je remercie deux fois plus que ce qu'il me remerciant dans sa propre thèse !
+Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{$\nearrow$}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{$\nearrow$}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{$\nearrow$}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{$\nearrow$}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{$\nearrow$}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{$\nearrow$}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, and Rami Othman.
 
 Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile, Maud, Frédéric, Gabriel (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline (and others) at CRIStAL in Lille.
-Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially Martine and J-B, your daily smile was priceless!
+Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially to Martine and J-B, your daily smile was priceless!
 
-During my PhD I was lucky to teach in Mathematics and Computer Science.
-I warmly thank David Pichardie for his trust, and also François Schwarzentruber, Nathalie Bertrand and David Cachera at ENS de Rennes, and Romaric Gaudel at ENSAI.
-Thanks to them, I loved all of the $3 \times 64$ hours spent teaching, since October $2016$.
+During my PhD, I was lucky to teach regularly in Mathematics and Computer Science.
+I warmly thank David Pichardie for his trust and support, and also François Schwarzentruber, Nathalie Bertrand and David Cachera at ENS de Rennes, and Romaric Gaudel and his colleagues at ENSAI.
+Thanks to them, I loved all of the $3 \times 64$ hours I spent teaching, since October $2016$.
 % I am trully honored to join \emph{the Département d'Informatique} at ENS de Rennes in next September $2019$.
-Warm thanks to my amazing students, I am proud of our interesting exchanges, and I feel lucky to be able to keep working alongside some of you.
+Warm thanks to my amazing students, I am proud of you and of our interesting exchanges.
+%  and I feel lucky to be able to keep working alongside some of you.
 I am grateful to Amélie Stainer and Gilbert Gabillard at Lycée Chateaubriand and Joliot-Curie for useful exchanges.
@@
 \indent
-I am thankful to amazing professors who shared their passion, first of mathematics and then of computer science:
-first, Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Marseille.
+I am thankful to all the amazing professors who shared their passion for mathematics and computer science with me.
+A partial list includes Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers, Marseille.
 Then at ENS de Cachan ($2011$-$16$), I especially think about Frédéric P., Alain T., Claudine P., Florian de V.
 and Nicolas V.
-in Maths, to Hubert C., Jean G., Sylvain S., Paul G., Serge H.
+in Maths, to Hubert C.-L., Jean G.-L., Sylvain S., Paul G., Serge H.
 in Computer Science, and to Carine S.-P., Delphine L.
@@
 for their support.
-I also thank Didier C. and Vijaysekhar C. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$).
+I also thank Didier C., Vijaysekhar C. and Jai P. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$).
 
 I would also like to thank some unusual things.
-I also thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday despite the too many accidents and frequent rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} which gave me smiles and energy, and finally my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git (and GitHub and Bitbucket) ; \LaTeX, Bash, Julia and Python ; GNU/Linux and (X)Ubuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
-Some people thank God in their thesis, I feel that I prefer to thank \emph{the Earth}, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
-I'll try hard to do better from now on!
+I thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday, despite the too many accidents and frequent rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} that gave me smiles and energy, and my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git (and GitHub and Bitbucket), \LaTeX, Bash, Julia and Python, GNU/Linux and (X)Ubuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
+Some people thank (a/some) God in their thesis, I feel that I prefer to thank \emph{the Earth}, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
+I'am trying hard to do better from now on!
 
@@
 %
-\emph{Deseo terminar este agradecimiento con un pensamiento especial para Lola, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con sus risas y su amor, durante el año pasado.}\\
+\small{
+    \emph{Deseo terminar este agradecimiento con un pensamiento especial para mi gatita \emph{L.}, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con sus risas y su amor, durante el año pasado.}\\
     \indent
     \textcolor{white}{$\heartsuit$ Te quiero, Lola, ¿quieres vivir conmigo?}
+}
 

### 2019-07-02 1-Preamble/5-Abstract/abstract.tex
Proofreading intro

--- 2019-06-20 1-Preamble/5-Abstract/abstract.tex
@@
 % Chapitre 1
-Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
-Nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
+Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de type radio intelligente, dans des bandes non licenciées et sans supervision centrale.
+Nous considérons des réseaux actuels ou futures de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
 %  (Chapitre~\ref{chapter:1}).
 % Chapitre 4
-Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement.
+Nous proposons deux modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui se reposent sur l'apprentissage par renforcement.
 %  (Chapitre~\ref{chapitre:4}).
@@
 % , Chapitre~\ref{chapitre:2}
-car il est délicat de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication ni aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+car il est délicat de prouver la convergence d'un grand nombre d'appareils jouant à un jeu collaboratif sans communication ni aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
 % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
@@
 % Chapitre 3
-Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre les modèles étudiés et d'autres.
+Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits, qui permet des simulations numériques de problèmes MAB, qui couvre les modèles étudiés et d'autres.
 % (Chapitre~\ref{chapter:3})
@@
 % Chapter 1
-In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access large-scale Cognitive Radio networks, in unlicensed bands and without central control.
-We focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
+In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access Cognitive Radio networks, in unlicensed bands and without central control.
+We focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, in order to let them automatically improve the efficiency of their wireless communications.
 %  (Chapter~\ref{chapter:1})
@@
 % ,  Chapter~\ref{chapter:2}
-because proving the convergence of many end-devices playing a collaborative game without communication no coordination is hard, when end-devices all follow random active pattern.
+because proving the convergence of many IoT devices playing a collaborative game, without communication nor coordination is hard, when they all follow random active pattern.
 % Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.
@@
 % Chapter 3
-We also detail another contribution, SMPyBandits, an open-source Python library for numerical MAB simulations, covering all the studied models and more.
+We also detail another contribution, SMPyBandits, our open-source Python library for numerical MAB simulations, that covers all the studied models and more.
 %  (Chapter~\ref{chapter:3})

### 2019-07-02 1-Preamble/5-Abstract/resume_fr.tex
Proofreading intro

@@
 Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
-Nous signalons les conférences nationales et internationales auxquelles ont été publié certains résultats.
+Nous signalons le chapitre correspondant, ainsi que les conférences nationales ou internationales auxquelles ont été publié certains résultats.
 
@@
     \item
-    Nous présentons le problème du choix de l'algorithme qu'un praticien devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithms MAB disponibles, dans le Chapitre~\ref{chapter:25}.
-    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par des simulations numériques qu'il atteint de bonnes performances empiriques
+    Nous présentons le problème du choix de l'algorithme qu'un-e ingénieur-e devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithmes MAB disponibles, dans le Chapitre~\ref{chapter:25}.
+    Nous présentons un algorithme, appelé \Aggr, pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par des simulations numériques qu'il peut atteindre de bonnes performances empiriques
     (publié à WCNC 2018) \cite{Besson2018WCNC}.
@@
 Cela permet à davantage de périphériques d'utiliser le même réseau tout en maintenant une haute qualité de service
-    (publiés à CROWNCOM 2017, ICT demo 2018, WCNC 2019 et MOTIoN 2019)
+    (publiés à CROWNCOM 2017, démonstration ICT 2018, WCNC 2019 et MOTIoN 2019)
     \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
@@
     \item
-    Nous avons implémenté une démonstration réaliste du modèle susmentionné \cite{Besson2018ICT,Besson2019WCNC}, et nous la présentons en détail en Section~\ref{sec:4:gnuradio}.
+    Nous avons implémenté une démonstration réaliste du modèle susmentionné \cite{Besson2018ICT,Besson2019WCNC}, présentée à ICT 2018, et nous la détaillons en Section~\ref{sec:4:gnuradio}.
 Nous avons aussi réalisé une courte vidéo présentant notre démonstration, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
@@
     Nous formalisons le modèle du bandit multi-joueur, pour lequel nous avons introduit deux variantes, dans le Chapitre~\ref{chapter:5}.
-    Pour le cas avec information de détection (``sensing''), nous proposons deux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM, qui prouve qu'il est asymptotiquement optimal.
-    Nous présentons aussi des expériences numériques approfondies pour montrer qu'il est bien plus efficace que les autres algorithmes de la littérature précédente.
+    Pour le cas avec information de détection (``sensing''), nous proposons deux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM, qui prouve qu'il est asymptotiquement très efficace.
+    Nous présentons aussi des expériences numériques approfondies pour montrer qu'il est bien plus efficace que les autres algorithmes de la littérature.
     Notre travail \cite{Besson2018ALT}, publié à ALT 2018, a également contribué à une nouvelle impulsion à la recherche sur les bandits multi-joueurs, car certains travaux de recherche récents se sont construits sur nos résultats.
@@
     \item
-    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle MAB multi-joueurs, qui, selon nous, n'a jamais été écrit auparavant.
+    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle MAB multi-joueurs, qui a été particulièrement active ces deux dernières années.
+    % qui, selon nous, n'a jamais été écrit auparavant.
     % (ALT 2018, et 4-8 travaux inspirés de notre article depuis lors).
@@
     \item
-    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (GRETSI 2019).
-    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif pour ce problème stationnaire par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles .
+    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (publié à GRETSI 2019).
+    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles .
     % Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB
@@
     \caption[Organisation de la thèse : une carte de lecture.]{Organisation de la thèse.
-Cette thèse peut se lire en suivant n'importe quel chemin contenant le Chapitre~\ref{chapter:1}, le Chapitre~\ref{chapter:2}, au moins un des trois Chapitres~\ref{chapter:4}, \ref{chapter:5} et \ref{chapter:6}, et la Conclusion.}
+Cette thèse peut se lire en suivant n'importe quel chemin contenant le Chapitre~\ref{chapter:1}, le Chapitre~\ref{chapter:2}, au moins un des trois Chapitres~\ref{chapter:4}, \ref{chapter:5} ou \ref{chapter:6}, et la Conclusion.}
     \label{fig:1:organization_fr}
@@
     \item
-\textcolor{darkred}{Dans la Partie~\ref{part:Introduction}}, nous commençons par le Chapitre~\ref{chapter:2} où nous présentons les modèles MAB, les concepts et les notations utilisés dans tout ce document.
+\textcolor{darkred}{Dans la Partie~\ref{part:Introduction}}, nous commençons par le Chapitre~\ref{chapter:2} qui présente les modèles de bandits multi-bras (MAB), les concepts et les notations utilisés dans tout ce document.
 Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
-Par contre, le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
-Nous terminons cette première partie par le Chapitre~\ref{chapter:25}, qui n'est pas non plus nécessaire à la compréhension du manuscrit, et qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
+Le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
+Nous terminons cette première partie par le Chapitre~\ref{chapter:25}, qui n'est pas non plus nécessaire à la compréhension globale du manuscrit, mais qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
 
     \item
-\textcolor{darkgreen}{La deuxième Partie~\ref{part:MABIOT}} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus presque indépendamment.
+\textcolor{darkgreen}{La deuxième Partie~\ref{part:MABIOT}} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus quasiment indépendamment.
 Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux IdO, dans lesquels les algorithmes MAB ont été utilisés avec succès.
@@
 %
-Ce document et les ressources additionnelles requises pour le compiler (notamment les fichiers \LaTeX, les morceaux de code Python, les figures etc)
-sont \href{https://github.com/Naereen/phd-thesis/}{publiées publiquement},
+Ce document et les ressources additionnelles requises pour le générer (notamment les fichiers \LaTeX, les morceaux de code Python, les figures etc)
+sont \href{https://github.com/Naereen/phd-thesis/}{distribuées publiquement},
 selon les termes de la \href{https://lbesson.mit-license.org/}{\emph{licence MIT}} open-source,

### 2019-07-02 2-Chapters/1-Chapter/chapter1.tex
Proofreading intro

@@
 %
-This manuscript concludes my doctoral thesis, which started in October $2016$ and finished in October $2019$.
-My research took place at the IETR laboratory in Rennes (France), in the SCEE team hosted on the Rennes campus of the engineering college CentraleSupélec.
-In Rennes I was supervised by Professor Christophe Moy,
+This manuscript concludes my doctoral thesis, which started in October $2016$ and finished in November $2019$.
+My research took place at the IETR laboratory in Rennes (France), in the SCEE team hosted on the Rennes campus of the engineering school CentraleSupélec.
+I was supervised by Professor Christophe Moy, in Rennes,
 and I was also co-supervised by Doctor Emilie Kaufmann, whom I visited many times at Inria Lille Nord Europe in Lille (France).
@@
 With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, for a large range of different applications.
-It is now a world-wide consensus that with the current trend of population increase and with the energy crisis, any newly deployed technology should be both cheap and energy efficient,
+It is now a world-wide consensus that with the current trend of population increase and with the ongoing energy crisis, any newly deployed technology should be both cheap and energy efficient,
 as well as adapted to serve a large number of people and devices.
@@
 Such IoT technologies should be able to adapt automatically to different environments and application contexts, and be as efficient as possible.
-Instead of devoting a lot of time, energy and money on Research and Development to set up efficient radio access schemes, for all possible cases, a cheap and promising direction is to use Machine Learning to (try to) attain the same performance gain automatically.
+Instead of devoting a lot of time, energy and funding on Research and Development to design many different efficient radio access schemes, for all possible cases, a cheap and promising direction is to use Machine Learning to (try to) attain the same performance gain automatically.
 
-That is why we are interested in this thesis about the possible applications of embedding a certain kind of Machine Learning algorithms (Multi-Armed Bandit algorithms),
-directly into the future IoT devices, in order to improve their battery life and reduce the energy cost of IoT networks, by optimizing their wireless communications and by learning to be self-organized, thanks to embedded decentralized decision making.
+That is why we are interested in this thesis about
+improving the battery life of IoT end-devices and reduce the energy cost of IoT networks.
+We propose to attain these two goals jointly, by embedding low-cost decentralized decision making on directly into the future IoT devices.
+%
+Our focus in this PhD thesis is thus on the possible applications of embedding a certain kind of Machine Learning algorithms (Multi-Armed Bandit algorithms), in order to let the IoT devices optimize their wireless communications and learn to be self-organized automatically and without central control nor coordination.
 
@@
 %
-Historically, three families of wireless communication systems have been deployed: first, centralized broadcasting (\eg, old TV), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things (IoT) networks.
+Historically, three main families of wireless communication systems have been deployed: first, centralized broadcasting (\eg, old TV or radio), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things (IoT) networks (\eg, sensor networks).
 
@@
 Starting from FM radio music and news, then old black-and-white TV and color TV, the first systems were purely made for centralized broadcasting of information.
-In a large area, one antenna well located emits continuously in a fixed Radio Frequency (RF) band, and many devices can listen to this RF traffic and decode it, for instance to listen to music.
-Different standards have been developed, and some of them are still in use today, and their nature requires a central authority that allocates RF bands to different usages (\eg, TV channels etc).
+In a large area, one well located antenna emits continuously in a fixed Radio Frequency (RF) band, and many devices can listen to this RF traffic and decode it, for instance to listen to some music.
+Different standards have been developed, and some of them are still in use today, and their nature requires a central authority that allocates RF bands to different usages (\eg, TV or FM radio channels etc).
 For example, in the city of Cesson-Sévigné where my campus of CentraleSupélec is located, a high RF tower emits on many standards, and in particular the \SI{92.3}{\mega\hertz} band is allocated to a French classical music radio (\emph{Radio Classique}).
 Broadcast-based systems are not only used for TV or radio,
-and for instance two other famous and worldwide deployed examples are the long-range global navigation satellite systems (such as USA's GPS or European Union's Galileo), and short-range infrared remote controllers.
+and for instance two other famous and worldwide deployed examples of mono-directional radio systems are the long-range global navigation satellite systems (such as USA's GPS or European Union's Galileo), and short-range infrared remote controllers.
 % https://en.wikipedia.org/wiki/Satellite_navigation
@@
 However, this kind of mono-directional wireless communications were soon found to not be satisfactory for many applications, and thus bi-directional standards were defined from the $1910$s.
-First developed by and for the military and the navy (\eg, the Titanic sent the first SOS signal over radio in $1912$),
+First developed by and for the military and the navy (\eg, the Titanic sent the first ``SOS'' signal over radio in $1912$),
 such applications have reached general public usage since the $1950$s.
-The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people's lives in the richer countries in the world, with the booming of cellphones since the $2000$s and smartphones since the $2010$s.
+The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people's lives in the richer countries in the world, with the booming of personal cellphones since the $2000$s and smartphones since the $2010$s.
 In most systems, there is still a base station (with one or more antenna) in charge of a large area (or cell), and many devices are able to receive and transmit data to the base station.
@@
 Even if both systems initially differed in usage, as 1G was designed to exchange only voice, and WiFi to link wireless devices at home to an Internet-connected box, they are conceptually very close.
-As the systems are cellular, any device knows the antenna (or base station) it is associated with, and the bi-directional nature of their exchanges makes a centralized control possible.
-In order to optimize the efficiency of the network, the base station is equipped with decision making algorithms that affects the devices in the cell to the available resources (time slots, RF bands, power etc).
-To be able to communicate with the base station without interfering with each others, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA), or Non-Orthogonal techniques only started to emerge recently (\eg, NOMA).
-Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance of the wireless communications to and from the devices.
-More recently, green radio has emerged as a solution to the tradeoff between maximizing QoS of the system, while minimizing the power consumption of base stations and devices.
+As the systems are cellular, any device knows the base station it is associated with, and the bi-directional nature of their exchanges makes a centralized control possible.
+In order to optimize the efficiency of the network, the base station is equipped with decision making algorithms that affects the devices in the cell to the available resources (time slots, RF bands, transmission power etc).
+To be able to communicate with the base station without interfering with each others, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA).
+%  and Non-Orthogonal techniques only started to emerge recently (\eg, NOMA).
+Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance, of the wireless communications to and from the devices.
+More recently, ``green radio'' has emerged as a solution to the tradeoff between maximizing the QoS of the system, and minimizing the power consumption of the base station and the devices it serves.
 
 % The opposite: decentralized broadcasting, and the age of IoT
-Finally, a third kind of systems are of the opposite kind, and can be designed as decentralized broadcasting: an antenna is still in charge of many devices, but such devices can only send up-link packets, and the only down-link data they can receive are short acknowledgements sent from the base-station to indicate success or failure of every up-link packet.
+Finally, a third kind of systems are of the opposite kind, and could be designed as decentralized broadcasting: one antenna is still in charge of many devices, but such devices can only send up-link packets, and the only down-link data they can receive are short acknowledgements, sent from the base-station to indicate success or failure of every up-link packet.
 This family of wireless systems are referred to as Internet of Things (IoT),
-and a typical example of application is for sensor networks.
+and a typical example of application of such IoT networks is for sensor networks.
 
-For the future development of smart grids, smart cities, smart homes, or smart agriculture, sensor networks are promised to be widely deployed.
-Two examples of future applications that are already in deployment, in France or other countries, are connected buildings and connected agriculture.
-For buildings, the main goal is to reduce the cost of heating empty buildings and use temperature sensor networks to get accurate and regular data about the temperature in all rooms and floors, and let the centralized heat control optimize its cost and energy consumption.
-For agriculture, one example can be to equipped every cow (in large farms) with a sensor that emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking and to monitor the health of the animals.
+For the future development of ``smart grids'', ``smart cities'', ``smart homes'', or ``smart agriculture'', sensor networks are promised to be widely deployed.
+Two examples of future applications that are already in deployment, in France or other countries, are ``connected buildings'' and ``connected agriculture''.
+For buildings, the main goal is to reduce the cost of heating empty buildings and use sensor networks to get accurate and regular data about the temperature in all rooms and floors, and let the centralized heat control optimize its cost and energy consumption.
+For agriculture, one example can be to equip every cow (in large farms) with a sensor that regularly emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking, to monitor the health of the animals etc.
 %
@@
 
-This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices, it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
+This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices (in downlink), it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
+Indeed, a central control requires signaling packets that have been identified as too heavy for these kinds of systems.
 In the present and future IoT networks, many devices of heterogeneous natures are using the same antenna for different applications.
-A common hypothesis is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and run on a tiny battery, which lifespan should be maximized, as typically commercial IoT companies sell a lifespan more than $10$ years \cite{Centenaro16}.
-Another common constraint for IoT devices is their low duty cycles, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems such as 4G/5G and WiFi.
+A common problem is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and will run on a tiny battery, which lifespan should be maximized.
+For instance, most commercial IoT companies nowadays sell a lifespan more than $10$ years, like SIGFOX \cite{Centenaro16}.
+Another common constraint for IoT devices is their low duty cycles, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems (such as 4G/5G and WiFi).
 %
-Many different standards for IoT networks have been proposed,
+Many different standards for IoT networks have been proposed in the recent years,
 and they consist in a specification for both the \emph{PHY}sical
@@
 A major problem of current wireless technologies is the issue of spectrum scarcity:
-in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, limiting the possibility of adding new usage.
-As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication) to space research, inter-satellite, mobile telephony and a large number of other applications.
+in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, and this limits the possibility of adding any new usage.
+As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication, \eg, in the Titanic), to space research, inter-satellite, mobile telephony and many other applications.
 %
@@
 Regulatory bodies in the world, like the \href{https://www.fcc.gov/}{Federal Communications Commission} in the USA (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
-or \href{http://www.tdf.fr/}{TDF} in France (see \href{https://www.TDF.fr/}{\texttt{TDF.fr}}, previously known as \emph{``TéléDiffusion de France''}), as well as different independent measurement campaigns, found that most radio frequency spectrum was inefficiently utilized,
+or \href{http://www.tdf.fr/}{TDF} in France (see \href{https://www.TDF.fr/}{\texttt{TDF.fr}}), as well as different independent measurement campaigns, found that most radio frequency spectrum are inefficiently utilized,
 meaning that while a band can be allocated to a certain unique usage, it can be free from any user in certain times and/or places.
-We refer to \cite{patil2011survey} for a survey on worldwide spectrum utilization, and to \cite{valenta2010survey} for the situation in Europe.
+We refer to \cite{patil2011survey} for a survey on the worldwide spectrum utilization, and to \cite{valenta2010survey} for the situation in Europe.
 
@@
     \centering
-    \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
+    % \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
+    \includegraphics[width=0.99\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
     \caption[A chart representing the allocation of radio spectrum in the USA in $2016$]{A chart representing the allocation of radio spectrum in the USA in $2016$.
-Only a small share of the spectrum is not yet allocated, the top-left corner in white (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz}).
-\copyright{} Wikimedia.}
+Only a small share of the spectrum is not yet allocated, the top-left corner in white and indicated by \textcolor{red}{\textbf{Free} in red} (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz}).
+\copyright{} United States Department of Commerce, \href{https://www.ntia.doc.gov/files/ntia/publications/january_2016_spectrum_wall_chart.pdf}{\texttt{www.ntia.doc.gov/files/ntia/publications/january\_2016\_spectrum\_wall\_chart.pdf}}.}
+    % https://commons.wikimedia.org/wiki/File:United_States_Frequency_Allocations_Chart_2016_-_The_Radio_Spectrum.pdf
     % https://upload.wikimedia.org/wikipedia/commons/c/c7/United_States_Frequency_Allocations_Chart_2016_-_The_Radio_Spectrum.pdf?uselang=fr
@@
 
-In this thesis, we study Internet of Things networks, and these two constraints of low power consumption (or long battery life) and low duty cycle are essential.
+In this thesis, we study Internet of Things networks, and the two constraints of low power consumption (or long battery life) and low duty cycle are essential.
 More precisely, we study the possible interconnections between \textbf{Cognitive Radio} and \textbf{Machine Learning} applied for IoT networks.
@@
 %
-Similarly, we can narrow down the second field of study of this thesis, steps by steps:
+We can narrow down the first field of study of this thesis, steps by steps:
 from Information Technologies (IT), to Telecommunications, to Wireless Communications, then to Software Defined Radio (SDR),
@@
 %
-The transition from the historical approach of hardware-based radio to SDR architectures and is a gradual process that started in the early $1990$s and has accelerated in the $2000$s.
+The transition from the historical approach of hardware-based radio to SDR architectures is a gradual process, that started in the early $1990$s and has accelerated in the $2000$s.
 % https://en.wikipedia.org/wiki/Software-defined_radio
 A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are all instead implemented by means of software on a personal computer or an embedded system, with the exception of a unique pair of receiver/transmitter antennas.
-Even though the SDR paradigm was initiated by the USA defense research, over the past years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
+Even though the SDR paradigm was initiated by the USA defense research, over the past $20$ years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
 % CR has been the first answer to solve spectrum scarcity issue which collected a huge attention in early 21st century.
 %
-CR is not a standard technology, and as such it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
+As CR is not a standard technology, it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
 % - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
@@
     \item
-    Joseph Mittola in $1999$ said that
+    Joseph Mittola in $1999$ proposed that
     \emph{``a really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment''} \cite{Mitola99}.

### 2019-07-02 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Proofreading intro

@@
 % -----------------------------------------------------------------
-\section{Theoretical elements, and regret upper bound for \MCTopM}
+\section{Finite-time upper-bound on the regret of \MCTopM}
 \label{sec:5:upperbounds}
@@
 % -----------------------------------------------------------------
-\section{Numerical simulations in presence of sensing}
+\section{Experimental results for multi-player bandits, with sensing information}
 \label{sec:5:experiments}
@@
 
-In all the numerical simulations, we focus on the model with sensing, \ie, model \modeldeux{} or \modeltrois.
+In all the numerical simulations, we focus on the model with sensing, \ie, model \modelun{} or \modeldeux.
 %

### 2019-07-02 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Update and fix typos
Proofreading intro

@@
         }
-        \caption{\GLRklUCB, with \textbf{Local} or \textbf{Global} restarts}
+        \caption{\GLRklUCB, with \textbf{Local} or \textbf{Global} restarts.}
         \label{algo:6:GLRklUCB}
@@
 
-\section{Experimental results for piece-wise stationary bandit problems}
+\section{Experimental results for piece-wise stationary bandits}
 \label{sec:6:NumericalExperiments}
@@
     \end{tabular}
-    \caption{Mean regret $\pm$ $1$ std-dev, on problems $1$, $2$ (with $T=5000$) and $3$ ($T=20000$).
-    }
+    \caption{Mean regret $\pm$ $1$ std-dev, on problems $1$, $2$ (with $T=5000$) and $3$ ($T=20000$).}
     \label{table:6:totalResults1}

### 2019-07-02 2-Chapters/1-Chapter/chapter1.tex
Done for Chapters 2 and 3

@@
     % \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
-    \includegraphics[width=0.99\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
+    \includegraphics[width=0.95\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
     \caption[A chart representing the allocation of radio spectrum in the USA in $2016$]{A chart representing the allocation of radio spectrum in the USA in $2016$.
@@
 %
-We can narrow down the first field of study of this thesis, steps by steps.
+Similarly, we can narrow down the second field of study of this thesis, steps by steps.
 % Stats > sequential learning > bandits
@@
 
-We illustrate this idea of a \emph{learning cycle} alternating between action and feedback,
-in the following Figure~\ref{fig:1:ReinforcementLearningCycle}.
-A player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$), and then observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
-The goal of the player is to maximize its reward, by trials and errors (\ie, actions and rewards).
+We illustrate below the idea of a \emph{learning cycle}, alternating between action and feedback,
+in Figure~\ref{fig:1:ReinforcementLearningCycle}.
+A player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$), and then by observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
+The goal of the player is to maximize its rewards, by trials and errors (\ie, actions and rewards).
 Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial), etc.
@@
 This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} (MAB) in details in the next Chapter~\ref{chapter:2}.
-Clinical trials and treatment discovery were historically the first applications of MAB since $1930$s, with the early work of W. Thompson \cite{Thompson33}.
+Clinical trials and treatment discovery were historically the first application of MAB since the $1930$s, with the early work of W. Thompson \cite{Thompson33}.
 %
-MAB are a simple yet powerful example of the well-known exploration/exploitation dilemma:
+MAB are a simple yet powerful example of the well-known \emph{exploration/exploitation dilemma}:
 when facing a set of $K$ actions whose effects on the environment are unknown, a learner must balance between
-exploration of the unknown actions, in order to collect more information about them,
-and exploitation of the best action, according to its current knowledge.
+exploring the unknown actions, in order to collect more information about them,
+and exploiting the best action, according to its current knowledge.
 %
@@
 Robbins \cite{Robbins52}, and more recently it is an active field of research, since the late $1990$s \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}, and the wide range of applications of MAB in the recent years.
+Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}, and the wide range of applications of MAB in the recent years \cite{bouneffouf2019survey}.
 
@@
 %
-The following hypothesis are maid:
-the SU is equipped with sensing,
-and a consider a fixed and finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
+The following hypotheses are maid:
+the SU is equipped with a sensing capability,
+and we consider a fixed and finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
 For instance, it can be a set of three Wi-Fi channels at different frequency, emitted by the same Wi-Fi station.
-Another hypothesis is that both PU and SU are synchronized in times, by sub-dividing the time in discrete time steps.
+Another hypothesis is that both PU and SU are synchronized in times, by sub-dividing the time in discrete time steps (or rounds).
 %
-Thus if the SU spends a short time in the beginning of each time slot to do sensing, it can scan for the presence or absence of any PU before transmitting.
-If a SU was able to sense for all the $K$ frequency bands, it could simply transmit in one of the free channels if any, or not transmit if all channels are used at a given time step.
+Thus if the SU spends a short time in the beginning of each time slot to perform sensing, it can scan for the presence or absence of any PU before transmitting.
+If the SU was able to sense for all the $K$ frequency bands, it could simply transmit in one of the free channels if any, or not transmit if all channels are used at a given time step.
 However, sensing is known to be error-prone and costly, especially for wide-band sensing, as detailed in the surveys \cite{yucek2009survey,subhedar2011spectrum}, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
@@
 Focussing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of $K\geq2$ channels), then it performs sensing, and finally it transmits in this channel if it was sensed free.
-The goal of the SU is to minimize its energy consumption (we remind that we focus on green radio) and to maximize its uplink data rate, or equivalently, to maximize its number of successful transmission.
+The goal of the SU is to minimize its energy consumption (we remind that we focus on ``green'' cognitive radio) and to maximize its uplink data rate, or equivalently, to maximize its number of successful transmission.
 %
-If the different channels are not uniformly used by the PU, and if we assume a stationary hypothesis on the PU traffic, then the goal of the SU boils down to explore the different channels and exploiting the best ones.
+If the different channels are not uniformly used by the PU, and if we assume a stationary hypothesis on the PU traffic, then the goal of the SU boils down to exploring the different channels and exploiting the best ones.
 This frames the Opportunistic Spectrum Access problem as an exploration/exploitation problem, with a finite set of actions (the channels, also called \emph{arms}),
@@
 under partial information feedback.
-These three hypotheses are the ones that restrict from the general RL framework to the specific Multi-Armed Bandit case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
+These three hypotheses are the ones that restrict from the general ML framework to the specific Multi-Armed Bandit case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
 
@@
 %
-This model was first studied by W. Jouini during his PhD thesis with C. Moy, ten years ago, first with \cite{Jouini09} and later with \cite{Jouini10,Jouini12}.
+This model was first studied by W. Jouini during his PhD thesis with C. Moy \cite{Jouini12PhD}, ten years ago, first with \cite{Jouini09} and later with \cite{Jouini10,Jouini12}.
 
-Their works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the \UCB{} algorithm,
+Their works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the $\UCB_1$ algorithm,
 along the early works of Q. Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
@@
 Since $2017$, S.
-Darak and his team at IIIT Delhi have actively worked in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+Darak, and his team at IIIT Delhi in India, have actively worked in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
@@
 %
-The aforementioned previous works have shown that MAB algorithms can be applied with success for the OSA problem.
+The aforementioned literature have essentially shown that MAB algorithms can be applied with success for the OSA problem.
 But if we consider CR-ready devices that cannot perform sensing, such as low-cost and low energy-consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
@@
 The specificities of IoT networks can be listed as follows,
-and we refer to \cite{Centenaro16} for more details.
+and we refer to the survey \cite{Centenaro16} for more details.
 \begin{itemize}\tightlist
@@
 %
-By using this feedback, that consists in an \emph{acknowledgement} (\Ack) message received shortly after every successful transmission, or in an absence of \Ack{} after a failed transmission, it is possible that an IoT end-device can also use RL algorithms to optimize its communications.
+By using this feedback, that consists in an \emph{acknowledgement} (\Ack) message received shortly after every successful transmission, or in an absence of \Ack{} after a failed transmission, it is possible that an IoT end-device can also use a well-designed RL algorithm in order to optimize its communications.
 %
@@
 %
-Because there are so many different MAB algorithms, we are also interested by the question of how practitioner can chose the one she will implement, in a given IoT object, in order to let this object adapts robustly in any environment.
+Because there are so many different MAB algorithms, we are also interested by the question of how a practitioner can chose the one she will implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
 To answer this question, we present two approaches.
@@
 
-In the second Part~\ref{part:MABIOT} of this thesis, we start by proposing and studying different models of IoT networks, with simulations and a real-world proof-of-concept, and then we study interesting questions on two mathematical models of MAB that arise from our first contribution.
+In the second Part~\ref{part:MABIOT} of this thesis, we start by proposing and studying different models of IoT networks, with simulations and a real-world proof-of-concept, and then we study interesting questions on two mathematical models of MAB, arising from simplification of our first model.
 
@@
 The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use acknowledgements as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the less occupied by the surrounding traffic).
-The goal of this first model of applying decentralized RL on the device side is to increase the Quality of Service (QoS) of the application of such IoT network, by reducing the rate of failed transmission.
+The goal of this first model is to increase the Quality of Service (QoS) of the application of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmission.
 The base station will thus receive more up-link packets from the devices being served, if they can successfully learn an efficient spectrum access scheme.
@@
 %
-Our second model then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increase QoS for the entire network.
+Our second model then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increased QoS for the entire network.
 
 We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of thousands (or more) of end-devices and a single base station.
-We consider a time-frequency slotted ALOHA-based protocol, by assuming a perfect synchronization between all devices and the base station.
+We consider an ALOHA-based protocol, slotted in both time and frequency, by assuming a simple pre-agreed synchronization between all devices and the base station.
+This synchronization is the only aspect of the network that requires central control from the base station, and it only requires one downlink packet when a new device joins the network operated by this base station \cite{Centenaro16}.
 %
@@
 Static devices form an interfering traffic, which could be generated by devices using other standards as well.
-Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choices of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) received from the base station.
+Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choice of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) received from the base station.
 %
@@
 %
-Even without changing anything on the level of the considered IoT standard, our proposal is just an add-on capability, that can be set-up on each device, on a unit-per-unit basis.
+Even without changing anything on the level of the considered IoT standard (neither the PHY nor the MAC layers), our proposal is just an add-on capability, that can be set-up on each device, on a unit-per-unit basis.
 This model constitutes the first article written during this thesis, and initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
-We presented this first work at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, and it obtained the ``best paper award'' \cite{Bonnefoi17}.
+We presented this first work at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, where it obtained the ``best paper award'' \cite{Bonnefoi17}.
 
 Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations was a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
-In Section~\ref{sec:4:firstModel}, we detail what we believe to be the first PoC of using decentralized MAB learning in a simplified ``IoT'' like wireless network.
-We were able to implement a network with a base station which receives incoming messages from different end-devices and replies with acknowledgements.
+In Section~\ref{sec:4:firstModel}, we detail what we believe to be the first PoC of using decentralized MAB learning in a simplified IoT-like wireless network.
+We implemented a network with a base station, which receives incoming messages from different end-devices and replies with acknowledgements.
 % , that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
 % Verify this first model empirically: proof of concept
-This is again the result of a fruitful collaboration with R.
+This is a second result of a fruitful collaboration with R.
 Bonnefoi, and this work was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
@@
 We then consider an extension of our first model, following the idea of packet retransmissions, that is at the root of the famous ALOHA protocol \cite{Abramson1970,Roberts75}.
-The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode in order to wait for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, $m \leq 10$ times), after some small random waiting times.
+The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode in order to wait for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, up-to $m \leq 10$ times), after some small random waiting times.
 %
@@
 This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well that the simplest of the different heuristics we compared is efficient, and allows the devices to learn to behave almost optimally.
-This third contribution concluded our collaboration with R. Bonnefoi, and was presented at the $1^{\text{st}}$ MoTION workshop, during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
+This third contribution concluded our collaboration with R. Bonnefoi, and was presented at the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
 
@@
 %
-In both models presented above, the core concept is to let every dynamic device of an IoT network run a fully decentralized learning algorithm, each running on its own (random) activation times, in order to optimize the complete system.
+In both models presented above, the core concept is to let every dynamic device of an IoT network run a learning algorithm in a fully decentralized way, in order to optimize the complete system.
+It is important to note that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own (random) activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
 It means that each of them target its own local objective, which is to maximize its cumulated reward (in a selfish way).
@@
 % https://www.ii.uni.wroc.pl/~lukstafi/pmwiki/uploads/AGT/Prediction_Learning_and_Games.pdf
-So it was quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices, despite the fact that these IoT objects cannot communicate directly with each other, and receive only a collision feedback from the base station they are associated to.
+So it was quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices in all the considered scenarios, despite the fact that these IoT objects cannot communicate directly with each other, and receive only a collision feedback from the base station they are associated to.
 
@@
 %
-That is why in Chapter~\ref{chapter:5}, we relax the hypothesis of having $M \gg K$ (or even simply $M > K$) devices in a network with $K$ orthogonal wireless channel, equivalent to having $p < 1$, and we consider the case of devices communicating at every time step (\ie, $p=1$), and so we restrict to only up-to $M \leq K$ devices.
+That is why in Chapter~\ref{chapter:5}, we relax the hypothesis of having $M \gg K$ (or even simply $M > K$) devices in a network with $K$ orthogonal wireless channel, that is equivalent to having $p < 1$, and we consider the case of IoT devices communicating at every time step (\ie, $p=1$), and so we restrict to only up-to $M \leq K$ devices.
 The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing) or the IoT case (\ie, without sensing).
@@
 as our model requires that an \Ack{} is sent back by the base station if the transmission was successful, even in the case where sensing information is available.
-It differs from the previously studied models of applying MAB for OSA, as they only considered synchronized devices, one SU and PU, and thus if the sensing indicates that one channel is free at one time slot, the up-link message sent by the SU device is sure to be successfully received by the base station (in the ideal model), so there is no risk of collision.
+It differs from the previously studied models of applying MAB for OSA, as they only considered synchronized devices, one SU and PU, and thus if the sensing indicates that one channel is free at one time slot, the up-link message sent by the SU device is sure to be successfully received by the base station (in the ideal model), so there is no risk of collision, and thus no need for an \Ack.
 
@@
 %
-On the other hand, we were able to propose new algorithms for this multi-player bandit problem with sensing information, and we analyzed our proposal \MCTopM-\klUCB, to show that it achieves a finite-time logarithmic regret upper bound, improving over previous state-of-the-art results.
+On the other hand, we were able to propose new algorithms for this multi-player bandit problem with sensing information, and we analyzed our proposal \MCTopM-\klUCB, to show that it achieves a finite-time logarithmic regret upper bound, improving significantly over previous state-of-the-art results.
 Our algorithm also achieves a logarithmic number of collisions and arm switches, and allows a fixed group of $M$ devices to efficiently learn to use the $M$ best channels orthogonally for almost all their up-link communications.
@@
 %
-and since then, many work have studied extension of our framework or alternative algorithms, as a few independent teams started to work on different questions shortly after its publication.
+and since then, some works have studied extensions of our framework or alternative algorithms, as a few independent teams started to work on different questions shortly after its publication.
 %
@@
 They showed that even under the strong hypothesis of no direct communication between players and no central coordination (in the same model as the one of our paper), the players can actually use the collision feedback (\ie, receiving or not an \Ack) as an indirect way to communicate bits of information with each other.
-This idea requires that devices coordinate and split the time steps between ``listening'' and ``talking'' phases, and the authors denote it the ``communication trick''.
+This idea requires that devices coordinate at least a little bit, by an initial pre-agreement (which is simply the fact that the IoT devices run the same algorithms).
 As they showed, the players can start by a ``musical chair'' phase, to reach an orthogonal allocation, and then they can alternate exploration, communication and exploitation phases until the end of the game at horizon $T$.
+Because they all follow the same algorithm, the devices are able to split the time steps between ``listening'' and ``talking'' phases, and the authors denote it the ``communication trick''.
 The communication trick is used by each player to send and receive observations about all arms from the other players, using one forced collision to send one bit at a time.
@@
 The difficulty does not reside in the fact that we are trying to analyze MAB algorithms, designed to tackle stationary problems, on a non-stationary problem,
-but rather that we are analyzing algorithms which are all playing in different (random) subsets of the global synchronized time.
-Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, but leads to the model of Chapter~\ref{chapter:5}.
+but rather that we are analyzing algorithms which are all playing in different (random) subsets of the global synchronized time steps.
+Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, even though this lead to the model of Chapter~\ref{chapter:5}.
 
-On the one hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let all groups of $M \leq K$ devices learn an optimal orthogonal affectation in the $K$ channels (\eg, groups can be the set of devices that all emit at the same time).
+On the one hand, if the activation pattern of all devices could be fixed, for instance based on a centralized affectation of the devices to different time slots, then the model of multi-player bandit from Chapter~\ref{chapter:5} can be used to let every group of $M \leq K$ devices learn an optimal orthogonal affectation in the $K$ channels (\eg, groups can be the set of devices that all emit at the same time).
 %
 However, even if we continue to assume a synchronized time in all the rest of the thesis,
-it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid any central control of the devices by the base station, as we explained above.
+it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid signaling and any central control of the devices by the base station, as we explained above.
 
@@
 %
-In order to also understand precisely how MAB algorithms behave under non-stationary environment, we studied the literature on adversarial and on non-stationary MAB, for both cases of slowly-varying and abruptly-changing environments.
+In order to also understand precisely how MAB algorithms behave under non-stationary environment, we studied the literature on adversarial and on non-stationary MAB, for the two cases of slowly-varying and abruptly-changing environments.
 We begin our last Chapter~\ref{chapter:6} by reviewing the existing works,
@@
 %
-As the recent literature showed that the later approach is usually more competitive, and usually obtains better results from both empirical and theoretical aspects, we chose to develop our own actively adaptive algorithm.
+As the recent literature showed that the later approach is usually more competitive, by obtaining better results from both empirical and theoretical aspects, we chose to develop our own actively adaptive algorithm.
 
@@
 We build on very recent results on the Generalized Likelihood Ratio test (GLRT) for Gaussian and sub-Gaussian variables \cite{Maillard2018GLR}, but we instead focus on bounded rewards and Bernoulli distributions.
-Bounded rewards are indeed usually more appropriate for Cognitive Radio applications.
+Bounded rewards are indeed usually more appropriate for Cognitive Radio applications, as shown above.
 Using the fact that bounded variables in $[0,1]$ are not only $1/4$ sub-Gaussian but also sub-Bernoulli, we prove the first finite-time guarantees for the GLRT for bounded variables.
-We first show finite-time bounds on both the false alarm probability of our test, and its detection delay, under mild assumption on the lengths of the stationary sequences.
+We first show finite-time bounds on both the false alarm probability of our test, and its detection delay, under mild assumptions on the lengths of the stationary sequences.
 Our algorithm, denoted \GLRklUCB, is then presented in two variants, whether change-points are local (\ie, only one arm mean changes at each change-point) or global (\ie, possibly all the arm means change at a time).
 %
-We prove that by combining the asymptotically optimal policy \klUCB, designed for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of our proposed algorithm.
+We prove that by combining the policy \klUCB, asymptotically optimal for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of our proposed algorithm.
 We consider the same hypothesis as our competitors, by assuming that the length of the stationary intervals to be ``long enough'' for the change-points to be detectable.
@@
 %
-This last chapter is based on our last work, and it first lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
+This last chapter is based on our last work, which lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
 Our work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results, in order to submit it soon to a journal (probably the Journal of Machine Learning Research) before December $2019$.
@@
 We conclude this chapter with a list of works published during this PhD.
-All the following works are published entirely and freely, on the HAL platform (see the \href{https://hal.archives-ouvertes.fr/}{\texttt{HAL.Archives-Ouvertes.fr}} website).
+All the following works are published entirely and freely, on the HAL platform (on \href{https://hal.archives-ouvertes.fr/}{\texttt{HAL.Archives-Ouvertes.fr}}).
 The complete list can be found on
@@
 
-    \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning for Wireless Communications, in July $2019$.}
+    % \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning for Wireless Communications, in July $2019$.}
 
@@
     \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},
-    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
+    by Rémi Bonnefoi, \textbf{Lilian Besson}, Julio Manco-Vasquez \& Christophe Moy.
     1st International MOTIoN workshop,
@@
     \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings},
-    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
+    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& Jacques Palicot.
     \emph{Conference on Cognitive Radio Oriented Wireless Networks},
@@
     \cite{Besson2019Gretsi}
+
+\end{itemize}
+
+% =============================================================================
+\subsection*{Submitted works}
+
+\begin{itemize}
+
+\item
+    \emph{Decentralized Spectrum Learning for Radio Collision Mitigation in Ultra-Dense IoT Networks: LoRaWAN Case Study and Measurements},
+    by Christophe Moy, \textbf{Lilian Besson}, Guillaume Delbarre \& Laurent Toutain,
+    July $2019$.
+    See Chapter~\ref{chapter:4}.
+    Submitted for a special volume of \href{https://annalsoftelecommunications.wp.imt.fr/}{the Annals of Telecommunications} journal, on ``Machine Learning for Intelligent Wireless Communications and Networking''.
+    % Preprint at \href{https://HAL.Inria.fr/hal-XXX}{\texttt{HAL.Inria.fr/hal-XXX}}.
+    \cite{MoyBesson2019Annales}
+
+    % \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
 

### 2019-07-02 2-Chapters/2-Chapter/chapter2.tex
Done for Chapters 2 and 3

@@
 %
-In this chapter, we present the common base of all the more sophisticated models studied in this thesis: the stochastic multi-armed bandit (MAB) model, restricting to the single-player and stationary stochastic case.
-%
-We motivate this model and discuss its application to and beyond cognitive radio, and we review existing MAB algorithms.
-Some of which are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling.
-%
-We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a one-dimensional distribution.
+In this chapter, we present the common base of all the more sophisticated mathematical models studied in this thesis: the stochastic multi-armed bandit (MAB) model, by restricting to the single-player and stationary stochastic case.
+%
+We motivate this model and discuss its application to and beyond cognitive radio.
+We review existing MAB algorithms,
+some of which are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling.
+%
+We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a binary or real-valued distribution.
 %
 We formalize this MAB model, then we quote some of its important applications.
-We define the notion of regret of an algorithm, and quote important results from the literature. A short review of the most important families of MAB algorithms is then given.
+We define the notion of regret of an algorithm, and present important results from the literature. A short review of the most important families of MAB algorithms is then given.
 % We conclude this chapter by discussing the question of online algorithm selection.
@@
 
-A MAB refers to a decision-making game, where a \emph{player} has to sequentially select one action in a usually finite set of actions, and only receives a (random) feedback about the selected action, also called a \emph{reward}.
+A MAB refers to a decision-making game, where a \emph{player} has to sequentially select one action in a usually finite set of actions (or arms), and only receives a (random) feedback about the selected action, also called a \emph{reward}.
 \textbf{We always consider $K\in\N, K\geq2$ arms, and discrete times $t\in\N^*$}.
@@
 and when the player decides at time $t$ to play (or pull) the arm $A(t)\in[K]$,
-she receives a reward $r(t) \eqdef Y_{A(t),t}$.
+she receives a reward $r(t) \eqdef Y_{A(t),t}$ from this stream.
 She keeps playing an arm and observing a reward iteratively, and so on for a finite number of steps $t$ (or rounds), from $t=1$ until $t=T$ for an \emph{horizon} $T$.
@@
 
-The difficulty of this decision-making game comes from balancing the trade-off between \emph{exploration}, because the player has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm.
-The MAB model is a famous example of a \textbf{reinforcement learning} model, where the decision-making process has to adapt to an unknown environment using noisy observations and a discrete time, alternating between decisions and observations.
+The difficulty of this decision-making game comes from balancing the trade-off between \emph{exploration}, because the player has to observe as much as possible all the arms to get more knowledge about the distributions of their rewards, and \emph{exploitation}, because she also has to select as much as possible the best arm in insight.
+The MAB model is a famous example of a \textbf{reinforcement learning} model, where the decision-making process has to adapt to an unknown environment using noisy observations, alternating between decisions and observations.
 We illustrate this cycle in Figure~\ref{fig:2:ReinforcementLearningCycleMABmodel} below.
@@
 Without loss of generality, it can be given by a sequence $U_0,U_1,\dots$ of \iid{} random values, uniformly distributed on $[0,1]$, and that has to be independent from the samples $Y_{k,t}$.
-External randomness is used in most MAB algorithms, \eg, to break ties uniformly in an index policy.
+External randomness is used in most MAB algorithms, \eg, to uniformly break ties in an index policy.
 
@@
 % , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
-and a summary is included in the Nomenclature, included in the Appendix in Part~\ref{part:Conclusion} (Page~\pageref{chapter:nomenclature}).
+and a summary is included in the Nomenclature, in the Appendix in Part~\ref{part:Conclusion} (Page~\pageref{chapter:nomenclature}).
 
@@
 %
-We focus on the \textbf{stochastic} MAB, in which the reward stream are drawn from some (unknown) distributions.
+We focus on the \textbf{stochastic} MAB, in which the reward stream are drawn from some (unknown) distributions, before the beginning of the game.
 %
@@
 %
-We also restrict to \emph{stochastic} rewards, that is we associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in[K]$.
-Rewards are stationary, meaning that $(Y_{k,t})_{t\in\N^*}$ is independent and identically distributed (\iid), and $Y_{k,t} \sim \nu_k$ for any $t\geq1$.
-We only consider binary or real-valued rewards, \ie, $Y_{k,t}\in\R$.
+% We also restrict to \emph{stochastic} rewards, that is
+We associate a real-valued distribution to each arm, denoted $\nu_k$ for arm $k\in[K]$.
+We assume that rewards are stationary, meaning that $(Y_{k,t})_{t\in\N^*}$ is independent and identically distributed (\iid), and $Y_{k,t} \sim \nu_k$ for any $t\geq1$.
+We only consider binary rewards, \ie, $Y_{k,t}\in\{0,1\}$, or real-valued rewards, \ie, $Y_{k,t}\in\R$, and in all this thesis rewards are one-dimensional.
 
 The most common objective for the player is to maximize its expected rewards $\E[\sum_{t=1}^T r(t)] = \sum_{t=1}^T \E[Y_{A(t),t}]$.
-The expectation $\E[\bullet]$ is taken on the rewards $(Y_{k,t})_{k,t}$, as well as on the external random variables $(U_s)_s$ (\ie, on the algorithm's decisions).
-%
-Other commonly studied objectives include best-arm identification (BAI) \cite{audibert2010best}.
+The expectation $\E[\bullet]$ is taken on the rewards $(Y_{k,t})_{k,t}$, as well as on the external random variables $(U_s)_s$, \ie, on the algorithm's decisions.
+%
+Other commonly studied objectives include best-arm identification (BAI) \cite{audibert2010best}, risk- or cost-aware reward maximization etc.
 
@@
 The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
-The arms follow Bernoulli distributions, \ie, $\nu_k = \mathcal{B}(\mu_k)$, of unknown means $\mu_k\in[0,1]$, and your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $T=100$ steps, \ie, to maximize $\sum_{t=1}^{100} r(t)$.
-Your decisions are made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{gold}{yellow}), \ie, the binary value $r(t)\in\{0,1\}$, that is sampled \iid{} from $\nu_k$.
+The arms follow Bernoulli distributions, \ie, $\nu_k = \cB(\mu_k)$, of unknown means $\mu_k\in[0,1]$ (this means that $\forall k\in[K], \forall t\in[T], \Pr(Y_{k,t} = 1) = \mu_k$ and $\Pr(Y_{k,t} = 0) = 1 - \mu_k$).
+Your goal in this interactive demonstration is to obtain the highest possible cumulated reward in $T=100$ steps, \ie, to maximize $\sum_{t=1}^{100} r(t)$.
+Your decisions are made sequentially: at time $t$, you pick one of the arms, $A(t) \in\{1,2,3,4,5\}=[5]$, then the demo shows the random reward obtained from this (virtual) casino machine (in \textcolor{deepgold}{yellow}), \ie, the binary value $r(t)\in\{0,1\}$, that is sampled \iid{} from $\nu_k$.
 %
@@
 while in the literature, the Bernoulli distribution is the most common case of finitely supported distributions.
-Continuous-valued distributions with finite support also included truncated versions of infinitely supported distributions, in particular \emph{truncated} Gaussian are used for numerical experiments in lots of research articles.
+Continuous-valued distributions with finite support also included truncated versions of infinitely supported distributions, in particular \emph{truncated} Gaussian are used for numerical experiments in some research articles.
 
@@
 %
-If the player knows that reward are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ (for $a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = \frac{r-a}{b-a}$ that lies in $[0,1]$.
+If the player knows that rewards are bounded in an interval $[a,b]$, and if she knows $a$ and $b$ (for $a<b$), then with no loss of generality we can restrict to the interval $[0,1]$, as if $r\in[a,b]$, the player can instead consider the normalized reward $r' = \frac{r-a}{b-a}$ that lies in $[0,1]$.
 Note that this ``normalization trick'' is implemented for any policy in our library SMPyBandits, with the \texttt{lower} and \texttt{amplitude} optional arguments, respectively representing $a$ the lower-bound on rewards and $b-a$ the length of the interval of possible values of rewards.
@@
 
-In the literature, parametric assumptions on the rewards are sometimes considered, typically the assumption that rewards belong to a real-valued distribution lying in an exponential families.
+In the literature, parametric assumptions on the rewards are sometimes considered, typically the assumption is that rewards belong to a real-valued distribution lying in an exponential families.
 %  or bounded distributions.
 One-dimensional exponential families include Bernoulli and Poisson distributions, as well as Gaussian distributions with a fixed variance.
-Our main interest is Bernoulli distributions, but we prefer to present the more general notations of exponential families in one dimension.
+In most cases, our main interest is Bernoulli distributions, but we prefer to present the more general notations of exponential families in one dimension.
 % We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Emilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
@@
 % \end{equation}
-for a parameter vector $\eta$ (the canonical parameter),
+for a parameter $\eta$ (the canonical parameter),
 % and a sufficient statistic that we restrict to be $T(x)=x$,
@@
 The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
-and usually the literature focusses on regular and minimal exponential families (when the natural parameter space is an non-empty open set in $\R$).
+and usually the literature focusses on regular and minimal exponential families (\ie, when the natural parameter space is an non-empty open set in $\R$).
 
@@
     \begin{align}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
-        \KL\left(d_1, d_2\right) \eqdef \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}\cup\{\pm\infty\}.
+        \KL\left(d_1, d_2\right) \eqdef \int d_1 \log\left(\frac{d_2}{d_1}\right) \lambda(\mathrm{d}x) = \E_{d_1} \left[ \log\left(\frac{d_2}{d_1}\right) \right] \in \mathbb{R}_{+}\cup\{+\infty\}.
     \end{align}
@@
     the parameter is $\eta = \mu / (1 - \mu)$, giving $A(\eta) = \log(1 + \mathrm{e}^{\eta})$
-    (with the limit behavior $\eta=+\infty$ if $\mu=0$).
-
-    The KL divergence between $B(x)$ and $B(y)$, of parameters $\eta$, $\eta'$ is given by
-    $\kl(x,y) \eqdef \KL_{B}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
+    (with the limit behavior $\eta=+\infty$ if $\mu=1$).
+
+    The KL divergence between $\cB(x)$ and $\cB(y)$, of parameters $\eta$, $\eta'$ is given by
+    $\kl(x,y) \eqdef \KL_{\cB}(\eta,\eta') = x \log(x/y) + (1-x) \log((1-x)/(1-y))$,
     and it is also called the \emph{relative binary entropy}.
@@
 
-    The KL divergence between $\cN(x,\sigma^2)$ and $\cN(x,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
+    The KL divergence between $\cN(x,\sigma^2)$ and $\cN(y,\sigma^2)$, of parameters $\eta$, $\eta'$ is given by
     $d_{\sigma^2}(x,y) \eqdef \KL_{\cN,\sigma^2}(\eta,\eta') = (x-y)^2 / (2\sigma^2)$
@@
 %
-Such hypothesis is also proposed for other families of distributions, for instance the recent article \cite{KimTewari2019} analyses their Follow-the-Perturbed-Leader algorithm for perturbations following any sub-Weibull distribution (sub-Weibull distributions generalize both sub-Gaussian and sub-Exponential distributions).
+Such hypothesis is also proposed for other families of distributions, for instance the recent article \cite{KimTewari2019} analyses their Follow-the-Perturbed-Leader algorithm for perturbations following any sub-Weibull distribution, that generalize both sub-Gaussian and sub-Exponential distributions.
 
@@
 Finally, we note that Markov bandit models have also been considered for Cognitive Radio applications \cite{Liu08}.
-They are not studied in this thesis, even though we did implemented them in SMPyBandits,
+They are not studied in this thesis, even though we did implemented them in SMPyBandits.
 They were introduced in the $1980$s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
-A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and thus they are no longer stochastic nor stationary.
-Such Markov models come in two flavors: rested or restless.
+A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and thus they are no longer stochastic.
+A Markov model exists in two flavors: rested or restless.
 For $K$ arms, each Markov chain has a finite number of states $s$, each corresponding to a (constant) reward that the player obtains if she selects this arm while its Markov chain is in state $s$.
@@
 %
-Such models were less studied than stationary or adversarial models, but some interesting works applied Markov models or CR applications in the last 10 years
+Such models were less studied than stationary or adversarial models, but some interesting works applied Markov models to Cognitive Radio in the last 10 years
 For instance, \cite{Melian15} proposed a CR model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
-A curious reader about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and refer to \cite{Norris98}.
+A reader curious about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and then refer to \cite{Norris98}.
 
@@
 
-The blooming success of the research on multi-armed bandits is easily explained by the different spectrum of applications of MAB models to real-world discrete-time decision making problems.
+The blooming success of the research on multi-armed bandits is easily explained by the different possibilities of applying MAB models to real-world discrete-time decision making problems.
 This research field has been very active since the years 2010s, but it started as early as 1933 with \cite{Thompson33}, and was active since the 1980s and the seminal works by Lai and Robbins \cite{LaiRobbins85} and by Anantharam and others \cite{Anantharam87a}.
@@
 Following the ``best arm identification'' model, the goal of a doctor in a clinical trial is to identify the best treatment, \ie, the arm with highest mean, in a number of trials as short as possible.
-And following our model of interest, maximizing the rewards corresponds to maximize the number of patients being successfully treated.
+And following our model of interest, maximizing the rewards corresponds to maximizing the number of patients being successfully treated.
 %
@@
     MAB can also be applied to a broader setting of \textbf{online content recommandation},
-    with more than two options.
-    The seminal work of \cite{Li10} studies the application of contextual bandit to news article recommendation, as it is used in practice on platforms such as Microsoft's Bing news website,
-    or in applications like Netflix.
+    with two options (A/B testing), or more than two options.
+    The seminal work of \cite{Li10} studies the application of contextual bandit to news articles recommendation, as it is used in practice on platforms such as Microsoft's Bing news website,
+    or in video streaming platform like YouTube or Netflix.
     In such models, the arms correspond to items to recommend (\eg, articles or movies), and the contexts contain features about each user of the system.
@@
     % like $\alpha\in[\frac{1}{2},\infty)$ for the $\UCB_1$ or $\varepsilon\in(0,1)$ for the $\varepsilon$-greedy bandit algorithms,
-    like a step size multiplier $\gamma$ in a gradient descent,
+    for instance for supervised machine learning we can think of
+    the step size multiplier $\gamma$ in a gradient descent,
     the width $\rho$ of a Radial Based Function (RBF) kernel,
@@
 As highlighted in Chapter~\ref{chapter:1},
-the focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc).
+the focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, the spreading factor for LoRa \cite{KerkoucheAlami18}, the power allocation for NOMA \cite{TianWangWangSong2019}, etc).
 % In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.
@@
 % In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-In the model of Opportunistic Spectrum Access (OSA) with sensing \cite{Jouini09,Jouini10}, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
+In the model of Opportunistic Spectrum Access (OSA) with sensing \cite{Jouini09,Jouini10}, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$, at the beginning of its $t$-th transmission slot.
 A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
@@
 
-The different applications discussed in this Section are still active research directions,
+The different applications discussed in this section are still active research directions,
 and a curious reader can find other interesting applications of MAB models and algorithms in \cite{bouneffouf2019survey}, such as game tree search or network routing in Section~1.2 of \cite{LattimoreBanditAlgorithmsBook}.
@@
 %
-In particular, an efficient algorithm should see its mean reward converge to the maximum reward.
+In particular, an efficient algorithm should see its mean reward converge to the maximum reward, if the horizon grows.
 In the bandit literature, the performance of an algorithm is often measured in terms of regret, a performance measure that we now introduce.
@@
 
-\subsection{Measuring performance with the (mean) regret}
+\subsection{Introducing the (mean) regret}
 
 Let us first introduce some notations.
-We consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that generates \iid{} samples $Y_{k,t} \sim \nu_k$, for any time $t$.
+We consider a stochastic and stationary MAB problem, with $K$ arms of distributions $\nu_1,\dots,\nu_K=(\nu_k)_k$, that generate \iid{} samples $Y_{k,t} \sim \nu_k$, for any time $t$.
 % with $r(t)=Y_{A(t),t}$).
@@
 
-One first need to observe that $R_T^{\cA} \geq 0$ for any $T$ and $\cA$, and that $R_T^{\cA} \leq \mu^* T$, so $R_T^{\cA} \leq T$ if the rewards lie in $[0,1]$.
+One first need to observe that $R_T^{\cA} \geq 0$ for any $T$ and $\cA$, and that $R_T^{\cA} \leq \mu^* T$, in particular $R_T^{\cA} \leq T$ if the rewards lie in $[0,1]$.
 Thus any algorithm has $R_T^{\cA} = \bigO{T}$, which justifies why we are interested in efficient algorithms that achieve at least a sub-linear regret, \ie, $R_T^{\cA} = \smallO{T}$.
@@
 %
-\emph{Its proof is simple and it is not a contribution, but we include it for readers that are unfamiliar with conditioning arguments}.
+\emph{Its proof is simple and it is not a contribution, but we include it below for readers that are unfamiliar with conditioning arguments}.
 
@@
 %     \item
-On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if we want to compute and visualize the regret, we can either use the definition with the rewards, or the decomposition and simply sum the gaps $\Delta_k$ with the number of sub-optimal draws.
-Both quantities are indeed equal \emph{in expectation}, but with only a finite number of trajectories and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition} (thanks to the conditioning on $O_t$ done in the proof).
+On the one hand, from a numerical simulation point-of-view, when we run a finite number of repetitions of the same stochastic experiment, if we want to compute and visualize the regret, we can either use the definition with the rewards, or the decomposition and simply sum the product of the gaps $\Delta_k$ with the number of sub-optimal draws.
+Both quantities are indeed equal \emph{in expectation}, but with only a finite number of experiments and observations, the first estimate is more noisy than the second one, since the randomness on the rewards is (partially) removed in the decomposition \eqref{eq:2:RegretDecomposition} (thanks to the conditioning on $O_t$ done in the proof).
 In our library SMPyBandits, we implement both estimators, and all values of regret used in this thesis are based on the one using the decomposition of Proposition~\ref{prop:2:RegretDecomposition}, because it gives faster convergence, and also ``smoother'' plots.
@@
 We include here two well-known results about what bandit algorithms \emph{cannot} do.
-First, a \emph{problem-dependent lower-bound} states that any algorithm suffers a regret at least $\Omega(\log T)$ on any problem (from \cite{LaiRobbins85}).
-Then, a \emph{worst-case result} states that for any algorithm, there exists a certain problem instance such that the algorithm can perform as badly as $\Omega(\sqrt{K T})$ (a result from \cite{Auer02NonStochastic}),
+First, a \emph{problem-dependent lower-bound} states that any algorithm suffers a regret at least $\Omega(\log T)$ on any parametric problem \cite{LaiRobbins85}.
+Then, a \emph{worst-case result} states that for any algorithm, there exists a certain problem instance such that the algorithm can perform as badly as $\Omega(\sqrt{K T})$  \cite{Auer02NonStochastic}.
 %
@@
 Let $\cI$ denote the set of all problem instances, with $K \geq 2$ arms.
-We assume the rewards lie in $[0,1]$,
+We assume that rewards lie in $[0,1]$,
 to present the lower-bounds, we prefer to focus on Bernoulli distributions, for simplicity.
@@
 These lower-bounds are of highest interest to design efficient algorithms,
-and a significant part of the research literature on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
-This means that an upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is asymptotically \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
+and a significant part of the research on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
+This means that a finite-time or an asymptotic upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is asymptotically \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
 
@@
 
-This family is non-empty as it contains for instance the $\UCB_1$ algorithm, since its regret is proven to be logarithmic on any Bernoulli problem instance (in Theorem~\ref{thm:2:UCBregretBound} below).
+This family is non-empty, as it contains for instance the $\UCB_1$ algorithm, since its regret is proven to be logarithmic on any Bernoulli problem instance (in Theorem~\ref{thm:2:UCBregretBound} below).
 Now we can state the second logarithmic lower-bound, for algorithms in this family.
@@
     or in other words,
+    $\liminf_{T\to\infty} \frac{R_T^{\cA}(I)}{\log(T)} \geq C_I$.
+    %
     \hfill{} \cite[Theorem~2.13]{Slivkins2019},
-    \[ \liminf_{T\to\infty} \frac{R_T^{\cA}(I)}{\log(T)} \geq C_I. \]
     % there exists a constant $C_I$ depending only on $I$,
@@
 
-We state below Theorem~2.12 from \cite{Slivkins2019}, to specify a possible value for the constant $C_I$.
+Theorem~2.12 from \cite{Slivkins2019} specifies a possible value for the constant $C_I$.
 
@@
 
-Moreover, it is interesting to note that the second lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Proposition~\ref{prop:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
+Moreover, it is interesting to note that the first lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Proposition~\ref{prop:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
 should sample each sub-optimal arm $k$ about $\log(T)/\kl(\mu_k,\mu^*)$ times in the total of $T$ time steps.
-
+%
 Many algorithms have been proven to achieve logarithmic regret in the stochastic case,
@@
 %
-For more details, we refer to Chapter~16 of \cite{LattimoreBanditAlgorithmsBook}.
+Chapter~16 of \cite{LattimoreBanditAlgorithmsBook} contains more details.
 
@@
 %
-For more details, we refer to Chapter~15 of \cite{LattimoreBanditAlgorithmsBook}.
+We also refer to Chapter~15 of \cite{LattimoreBanditAlgorithmsBook}.
 
@@
     Fix the number of arms $K$, and an algorithm $\cA$.
-    Then for any horizon $T$, there exists a Bernoulli problem instance $I_T$ on which the algorithm suffers at least a regret $R_T^{\cA}(I_T) \geq \Omega(\sqrt{K T})$.
+    Then for any horizon $T$, there exists a Bernoulli problem instance $I_T$ on which $\cA$ suffers at least a regret $R_T^{\cA}(I_T) \geq \Omega(\sqrt{K T})$.
 \end{leftbar}  % XXX leftbar theorembar, comment if needed
@@
 
-The theoretical results obtained in the rest of this thesis are all expressed in terms of the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances that we consider in our works and in this manuscript.
+In the rest of this thesis, the theoretical performance analysis of an algorithm that we obtained are all expressed in terms of the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances that we consider in our works and in this manuscript.
 
@@
 This is the case of the \Selfish{} algorithm that is defined and explored in Chapter~\ref{chapter:5}.
-By visualizing the entire \emph{distribution} (as an histogram), or the variance of the values of $R_T^{\cA}$, and if the number $N$ is reasonably large, we can verify that the regret appears logarithmic for all runs.
+On the contrary, for efficient algorithms,
+we can visualize the entire \emph{distribution} as an histogram, or the variance of the values of $R_T^{\cA}$, and if the number $N$ is reasonably large, we can verify that the regret appears logarithmic for all runs.
 
@@
 %
-% \TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
-Because it is relevant for cognitive radio applications, as a hardware reconfiguration costs energy, and so changing channel from two consecutive time steps has a non trivial energy cost, as highlighted in \cite{modiDemo2016}.
-And thus we find more interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
+It is relevant for cognitive radio applications, as a hardware reconfiguration costs energy, and so changing channel from two consecutive time steps has a non trivial energy cost, as highlighted in \cite{modiDemo2016}.
+We also find interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
 Our proposed algorithm, \MCTopM, achieves order-optimal logarithmic regret as well as a logarithmic number of switches.
@@
 %
-% \TODOL{Enlever ça, juste une phrase qui dit "on le fait plus tard".}
 In this thesis, we only consider fairness in the multi-players bandit models,
-in Chapter~\ref{chapter:5}, where fairness refers to a different notion (\ie, ``cooperative fairness'').
+in Chapter~\ref{chapter:5}, where fairness refers to a different notion (\ie, ``cooperative fairness''),
 % When $M<K$ players learn in a decentralized way, they will essentially converge to an orthogonal affectation of players to arms, so that each player uses a unique arm among the $M$ best arms (most of the times).
 % The fairness constraint says that, in average, no user is arbitrarily going to converge more frequently on a better arm than another user, and it is for instance the case for permutation invariant decentralized algorithms.
-We refer to Section~\ref{sub:5:betterLowerBound} for more details.
+in Section~\ref{sub:5:betterLowerBound}.
 
@@
 -- The uniform strategy always plays the $K$ arms uniformly at random,
-$\forall t\in[T], A(t) \sim \cU([K])$, where $\cU(S)$ denotes the uniform distribution on set $S$, and $\cU([K])$ the uniform distribution on $[K] = \{1,\dots,K\}$ (\ie, $\forall t, \forall k\in[K], \Pr(A(t) = k) = \frac{1}{K}$).
+$\forall t\in[T], A(t) \sim \cU([K])$, where $\cU(S)$ denotes the uniform distribution on a set $S$, and $\cU([K])$ the uniform distribution on $[K] = \{1,\dots,K\}$ (\ie, $\forall t, \forall k\in[K], \Pr(A(t) = k) = \frac{1}{K}$).
 %
@@
 Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
-which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all non trivial problems, rulling out the case where $\mu_1=\mu_2=\dots=\mu_K$).
+which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all non trivial problems, ruling out the case where $\mu_1=\mu_2=\dots=\mu_K$).
 
@@
 %
-%
-We note that this strategy can also achieve a much larger regret bound, $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$, without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
+% We note that this strategy can also achieve a much larger regret bound, $R_T = \bigO{T^{2/3} (K \log(T))^{1/3}}$, without prior knowledge on the problem, as shown in Section~1 of \cite{Slivkins2019}.
 
@@
 
-    So this bound on $p$ gives $R_T \leq K T_0 + 2 K \exp(-T_0 d^2 / (2 K)) T$ (because $\max_{k: \Delta_k > 0} \Delta_k \leq 1$ and $T - T_0 \leq T$).
+    So this bound on $p$ gives $R_T \leq K T_0 + 2 K \exp(-T_0 d^2 / (2 K)) T$ (as $\max_{k: \Delta_k > 0} \Delta_k \leq 1$ and $T - T_0 \leq T$).
     Optimizing the right hand-side on $T_0$ gives $T_0 = ((2K)/d^2) \log(d^2 T / (2K))$.
-    This gives the announced bound.
+    % This gives the announced bound.
 \end{smallproof}
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    Let $Z_1,\dots,Z_n$ be $n$ \iid{} samples from a Bernoulli distribution of parameter $\mu$ (where $n$ is fixed),
-    of empirical mean $\widehat{Z_n} \eqdef \frac{1}{n} \sum\limits_{i=1}^n Z_i$, \emph{Hoeffding's inequality} gives both
-    % \vspace*{-10pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    Let $Z_1,\dots,Z_n$ be $n$ \iid{} samples from a Bernoulli distribution of parameter $\mu\in[0,1]$ (where $n$ is fixed),
+    of empirical mean $\widehat{Z_n} \eqdef \frac{1}{n} \sum_{i=1}^n Z_i$, \emph{Hoeffding's inequality} gives
+    \vspace*{-5pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    % \begin{small}
     \begin{align}\label{eq:2:HoeffdingInequality}
-        \begin{cases}
-            \forall x < \mu, \;\;\; & \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
-            \\
-            \forall x > \mu, \;\;\; & \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
-        \end{cases}
+        % \begin{cases}
+            \forall x < \mu, \;\;\; \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
+            \;\;\;
+            \forall x > \mu, \;\;\; \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+        % \end{cases}
+        % \begin{cases}
+        %     \forall x < \mu, \;\;\; & \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
+        %     \\
+        %     \forall x > \mu, \;\;\; & \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+        % \end{cases}
     \end{align}
-    % \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    % \end{small}
+    \vspace*{-5pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 %
-This intuition leads to the $\UCB_1$ algorithm, initially introduced in \cite{Auer02}, for bounded rewards.
+This intuition took roots in \cite{LaiRobbins85}, and later lead to the $\UCB_1$ algorithm, analyzed in \cite{Auer02} for bounded rewards.
 
@@
 
-    % \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ? si c'est bien fait, ça peut aider la lecture, par exemple \textcolor{darkgreen}{$k$} et \textcolor{gold}{$k^*$}.}
+    % \TODOL{J'aimerai bien rajouter des couleurs sur $k$ et $k^*$ partout plus bas ?
+si c'est bien fait, ça peut aider la lecture, par exemple \textcolor{darkgreen}{$k$} et \textcolor{deepgold}{$k^*$}.}
 
@@
     but as \eqref{eq:2:trickInUCBProof} above started to isolate $N_k(t)$, we add the freedom of considering different number of samples (using $n$), so we introduce the notation
-    $\widehat{\mu}_{k,n}(t)$ and $\xi_{k,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \leq N_k(t)$,
+    $\widehat{\mu}_{k,n}(t)$ and $\xi_{k,n}(t)$ where $N_k(t)$ is replaced with $n$ if $n \geq N_k(t)$,
     \ie, $\widehat{\mu}_{k,n(t)} = \frac{1}{n} \sum_{s=1}^{t-1} r(s) \indic(A(s)=k, N_k(s) \leq n)$ and $\xi_{k,n}(t) = \sqrt{2\log(t) / n}$.
@@
     \end{align}
-    So two union bounds on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$ give two sums
+    So two union bounds, on $n_{k^*}$ for the $\min$ and on $n_k$ for the $\max$, give two sums
     \begin{align}\label{eq:2:decompositionInUCBProof2}
@@
         observe that $\mu_{k^*} < \mu_k + 2 \xi_{k,n_k}(t)$ means $\Delta_k - 2 \sqrt{2\log(t) / n_k} < 0$, where we remind the notation of the gap $\Delta_k \eqdef \mu_{k^*} - \mu_k$.
-        Thus if $n_k \geq \left\lceil \frac{8 \log(T)}{\Delta_k^2} \right\rceil \geq \left\lceil \frac{8 \log(t)}{\Delta_k^2} \right\rceil$,
+        Thus if $n_k \geq \left\lceil \frac{8 \log(T)}{\Delta_k^2} \right\rceil \geq \left\lceil \frac{8 \log(t)}{\Delta_k^2} \right\rceil$ (as $T \geq t$),
         then this inequality actually cannot happen:
@@
 and then the \klUCB{} algorithm was introduced in \cite{KLUCBJournal} for bounded rewards.
-In all this thesis, we either restrict to exponential families or to bounded rewards in $[0,1]$ and use the fact that they are sub-Bernoulli, thus we prefer to only present \klUCB.
+In all this thesis, we either restrict to exponential families or to bounded rewards in $[0,1]$, for which we use the fact that they are sub-Bernoulli, thus we prefer to only present \klUCB.
 
 Instead of using a confidence radius $\xi_k(t)$ based on Hoeffding's inequality,
-the \klUCB{} algorithm rather uses the Chernoff concentration inequality for the considered exponential family
+the \klUCB{} algorithm rather uses the Chernoff's concentration inequality for the considered exponential family \cite{chernoff1952measure,chernoff1981note}.
 % such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
 % It yields improved upper confidence bounds (\ie, with smaller $\xi_k(t)$) that are based on Kullback-Leibler divergence.
-We remind that we consider bounded rewards in $[0,1]$.
+We remind that we consider rewards bounded in $[0,1]$.
 Consider an exploration function $f(t)$,
@@
 and $f(t) \eqdef \log(t)$ in practice,
-then the \klUCB{} indexes are defined as follows
+then the \klUCB{} indexes are defined by
 % The constant $c>0$ is usually chosen as $c = 1$.
@@
 for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
-The second one is less obvious: one has to solve the optimisation problem defining the $\sup$ in \eqref{eq:2:klUCB_index}.
+The second one is less obvious: one has to solve the optimisation problem in \eqref{eq:2:klUCB_index}.
 Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\kl(x,\bullet)$ function is usually convex (for any $x$), two practical and mathematically correct solutions are Newton's method or a bisection search.
@@
 % and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
-For instance, $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically the bisection search terminates in between $5$ to $20$ steps.
+For instance, a precision level of $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically the bisection search terminates in between $5$ to $20$ steps.
 
@@
 The bottleneck of the performance is the computations of the $\KL$ functions and the optimization problem,
-and we note that our library SMPyBandits implements both parts in both a naive Python implementation (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback.py}{\texttt{kullback.py}}) and in a Python module written in a
+and we note that our library SMPyBandits implements the two parts in both a naive Python implementation (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback.py}{\texttt{kullback.py}}) and in a Python module written in a
 C file\footnote{~Note that we also wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
@@
 in an optimized way (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/C/kullback_py3.c}{\texttt{kullback\_py3.c}}).
-For the supported families of distributions (including Bernoulli, Gaussian, exponential and Poisson), we wrote both a naive and an optimized version of $\KL(x,y)$ and the $\sup_q \KL(x,q) \leq z$ optimization problem, using Newton's method.
+For the supported families of distributions (including Bernoulli, Gaussian, exponential and Poisson), we wrote both a naive Python and an optimized CPython version of their Kullback-Leibler divergence (\ie, $\KL(x,y)$) and a bisection search for the optimization problem for the \klUCB{} index (\ie, $\sup_q \KL(x,q) \leq z$).
 
@@
 the best choice of prior is a Beta distribution, that starts as an initially uniform prior (\ie, $\forall k, \pi_k(0) \eqdef \mathrm{Beta}(1,1)$),
-and maintain a distribution $\pi_k(t) \eqdef \mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$,
+and maintains a distribution $\pi_k(t) \eqdef \mathrm{Beta}(X_k(t)+1,N_k(t)-X_k(t)+1)$,
 as $X_k(t)$ counts the number of successes observed on arm $k$, and $N_k(t)-X_k(t)$ counts the number of failures.
@@
 %
-In this setting as well as for Gaussian distributions,
+In this setting, as well as any one-dimensional exponential families,
+% for Gaussian distributions,
 Thompson sampling was proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
@@
 		}
-		\caption{Thompson Sampling for Bernoulli rewards, with a Beta prior on each arm, and also Beta posteriors.}
+		\caption{Thompson Sampling for Bernoulli rewards, with Beta prior/posteriors.}
 		\label{algo:2:ThompsonSampling}
@@
 The first of them is the Exp3 policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
-which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details see
+which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details, see
 % on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
-this blog post by Jeremy Kun, \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-}}\\
-\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{effectiveness-of-the-multiplicative-weights-update-algorithm}},
-which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019).
-More details can be found in the survey \cite{Arora2012multiplicative}.},
+this article by J.
+Kun, \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-}}\\
+\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{of-the-multiplicative-weights-update-algorithm}}.
+The survey \cite{Arora2012multiplicative} gives more details.},
+% which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019)
 a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
@@
 For any $K$, $T$, and problem instances $I$ in a certain family $\cI$ (\eg, with Bernoulli distributions),
-they enjoy regret bounds like the following, for two constants $c_{\cA}$ and $c'_{\cA}$ that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only: $R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T))$.
+they enjoy regret bounds like the following, for two constants $c_{\cA} \geq 1$ and $c'_{\cA} \geq 1$, that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only: $R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T))$.
 
@@
 % https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
-Finally, it is worth mentioning the Exp3$^{++}$ algorithm, as a recent variant of Exp3, proposed by \cite{Seldin17},
-that uses an adaptive tuning of its parameters.
+Finally, it is worth mentioning the Exp3$^{++}$ algorithm, as a recent variant of Exp3 that uses an adaptive tuning of its parameters, proposed by \cite{Seldin17}.
 It is anytime, and it was also proven to obtain good ``best-of-both-world'' performance.
@@
 We finish this Section~\ref{sec:2:famousMABalgorithms} by illustrating the performance of different algorithms, on the small MAB problem used in the online demonstration shown above (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
-We consider $K=5$ arms following Bernoulli distributions, of means $\mu_1=0.6,\mu_2=0.2,\mu_3=0.55,\mu_4=0.7,\mu_5=0.5$ (unknown to the algorithms), and first an horizon $T=100$ (like in the online demo) then a large value of $T=10000$.
+We consider $K=5$ arms following Bernoulli distributions, of means $\mu_1=0.6,\mu_2=0.2,\mu_3=0.55,\mu_4=0.7,\mu_5=0.5$ (unknown to the algorithms), and first an horizon $T=100$ (like in the online demo) then a larger value of $T=10000$.
 The best arm is $\mu_4$ and the smallest gap is $\Delta = 0.1$, thus a lower-bound on the gaps is taken as $d=\Delta$.
@@
         \textbf{Algorithms}
-        & rewards
-        & regret for $T=100$
-        & rewards
-        & regret for $T=10000$ \\
+        & Rewards
+        & Regret for $T=100$
+        & Rewards
+        & Regret for $T=10000$ \\
         % & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
@@
         \textbf{Algorithms}
-        & rewards
-        & regret for $T=100$
-        & rewards
-        & regret for $T=10000$ \\
+        & Rewards
+        & Regret for $T=100$
+        & Rewards
+        & Regret for $T=10000$ \\
         % & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=100}$ & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=10000}$ \\
@@
 
+\paragraph{Summary of this review of MAB algorithms.}
+%
+We gave an overview of the most well known families of algorithms for multi-armed bandits,
+focussing on algorithms designed and analyzed for the single-player stationary and stochastic MAB model.
+Except the naive and simple strategies given as introductory examples, most of the policies presented above are order-optimal for Bernoulli distributed problems or bounded rewards (or one-dimensional exponential families),
+and some of them are known to be optimal in different settings.
+%
+To clearly understand the difference in their empirical behaviors, we illustrated the performance of two naive and two simple strategies against three efficient ones, on the example of Bernoulli-distributed problem used for the online demonstration from Section~\ref{sec:2:notations} (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
+This also serves as a first example of numerical simulation performed with our library SMPyBandits, before the large-scale experiments of Section~\ref{sec:3:reviewSPAlgorithms} below.
+%
+In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity and realistic deployment is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
+
 \begin{figure}[h!]  % [htbp]
@@
 
-\paragraph{Summary of this review of MAB algorithms.}
-%
-We gave an overview of the most well known families of algorithms for multi-armed bandits,
-focussing on algorithms designed and analyzed for the single-player stationary and stochastic MAB model.
-Except the naive and simple strategies given as introductory examples, most of the policies presented above are order-optimal for Bernoulli distributed problems or bounded rewards (or one-dimensional exponential families),
-and some of them are known to be optimal in different settings.
-%
-To clearly understand the different in their empirical behaviors, we illustrated the performance of two naive and two simple strategies against three efficient ones, on the example of Bernoulli-distributed problem used for the online demonstration from Section~\ref{sec:2:notations} (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
-This also serves as a first example of numerical simulation performed with our library SMPyBandits, before the large-scale experiments of Section~\ref{sec:3:reviewSPAlgorithms} below.
-%
-In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity and realistic deployment is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
-
-% % \newpage
+\newpage
 
@@
 % %
-We also showed a small numerical simulation, comparing some algorithms on a small Bernoulli-distributed problem, using our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
+We then use the small Bernoulli-distributed problem behind this demonstration
+to present a small numerical simulation, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
 % The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

### 2019-07-02 2-Chapters/3-Chapter/chapter3.tex
Lots of things to do, oh my god.
Done for Chapters 2 and 3

@@
 
-\chapter{SMPyBandits: a state-of-the-art Python library to simulate MAB problems}
+% \chapter{SMPyBandits: a state-of-the-art Python library to simulate MAB problems}
+\chapter{SMPyBandits: our Python library to simulate MAB problems}
 \label{chapter:3}
@@
 It allows fast prototyping of simulations, with an easy configuration system and command-line options to customize experiments.
-Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
+Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of useful visualizations.
 %
@@
 SMPyBandits stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
-The library does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a pure Python implementation \cite{python}, and uses only standard open-source Python packages (\ie, no hard to install dependencies).
+The library does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a Python implementation \cite{python}, and uses only standard open-source Python packages (\ie, no hard to install dependencies).
 Some critical parts are also available as a \texttt{C} Python extension, and the just-in-time Numba compiler \cite{numba} is used whenever it is possible, so we can note that we optimized the time efficiency of what could be (easily) optimized.
@@
 We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
-then we discuss in Section~\ref{sec:3:timeAndMemoryCosts} about the time and memory costs of MAB algorithms.
+then we discuss in Section~\ref{sec:3:timeAndMemoryCosts} about the time and memory costs the same algorithms.
 The take away messages are two-fold:
 from a practical point-of-view, it is usually advised to use the simplest algorithm and we advise to use \UCB{} (as we do in the next Chapter~\ref{chapter:4}),
-and that for theoretical works where optimality of the MAB algorithm matters, we advise to use \klUCB{} (as we do in the next Chapters~\ref{chapter:5} and \ref{chapter:6}).
+and that for theoretical works where optimality of the MAB algorithm is analyzed, we advise to use \klUCB{} (as we do in the next Chapters~\ref{chapter:5} and \ref{chapter:6}).
 % To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
@@
 After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
-The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with a low variance.
+The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with low variances.
 
@@
 For the classical single-player stochastic MAB model, as defined in Chapter~\ref{chapter:2},
-a stochastic MAB problem is defined by $K>1$ distributions $(\nu_k)_{k\in[K]}$ (also called arms),
+a stochastic MAB problem is defined by $K \geq 2$ distributions $(\nu_k)_{k\in[K]}$ (also called arms),
 used to generate the \iid{} samples $Y_{k,t} \sim \nu_k, \forall t$.
 %
-An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r(t) = Y_{A(t),t}$ without knowing the other (hidden) rewards (in opposition to the full-information setting where $(Y_{1,t},\dots,Y_{K,t})$ is observed)).
+An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r(t) = Y_{A(t),t}$ without knowing the other (hidden) rewards (in opposition to the full-information setting where $(Y_{1,t},\dots,Y_{K,t})$ is observed).
 %
@@
 
-\textbf{Simulation loop for single-player MAB.}
+\paragraph{Simulation loop for single-player MAB.}
 %
 Any simulation library targeting single-player bandit problems must implement at least three components:
-reward distributions, MAB algorithm, and a simulation loop that essentially looks like this.
+reward distributions, MAB algorithms, and a simulation loop that essentially looks like this.
 % \begin{itemize}
@@
     Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
-Ask algorithm $\cA$ her chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm,
+Ask algorithm $\cA$ her chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm.
     % \item
-    At the end, compute the cumulated reward, the regret, plot visualizations etc.
+    At the end, compute the cumulated reward, the regret, then plot visualizations etc.
 % \end{itemize}
 %
-Note that the second step is repeated a large number of times (\eg, $N=100$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
-If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(Y_{k,t})_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms.
-
-\textbf{Simulation loop for multi-players MAB.}
-%
-For Cognitive Radio dealing with OSA problems and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
+Note that the second step should be repeated a large number of times (\eg, $N=100$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
+If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(Y_{k,t})_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms (with the option \texttt{configuration["cache\_reward"] = True}).
+
+\paragraph{Simulation loop for multi-players MAB.}
+%
+For Cognitive Radio dealing with OSA problems and other applications, an extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
 Whenever two or more players select the same arm at the same time (\eg, for OSA, the same frequency channel), they all suffer from a radio collision.
@@
 
-SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.html}{PoliciesMultiPlayers}}).
+SMPyBandits implements all the collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.html}{PoliciesMultiPlayers}}).
 %
@@
     % \item
-    For $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
+    For $t=1$ to $t=T$ (known before hand), repeat the following block (for each cohort).
 Ask each player $\cA^j$ her chosen arm $A^j(t)$, then query the collision model to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
@@
     % \item
-    At the end, compute the accumulated reward, the regret, plot visualizations etc.
+    At the end, compute the accumulated reward, the regret, then plot visualizations etc.
 % \end{itemize}
@@
 %
-We focus on one dimensional distributions, implemented in the \texttt{\href{https://smpybandits.github.io/docs/Arms.html}{Arms}} module.
+We focus on one dimensional distributions (\ie, $r(t)\in\R$), implemented in the \texttt{\href{https://smpybandits.github.io/docs/Arms.html}{Arms}} module.
 The library supports discrete distributions: \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Poisson}}), and a generic \emph{discrete} distribution (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{DiscreteArm}}),
@@
 %
-For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulates the problem.
-We give another example for \emph{truncated} Gaussian distributions (on $[0,1]$), and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.
+For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following Bernoulli distributions, with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulates the problem.
+We give another example for \emph{truncated} Gaussian distributions on $[0,1]$, and a visualization of a histogram of $10000$ rewards, below in Figure~\ref{fig:3:exampleOfRewards}.
 
@@
 
-We have not yet added support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
+We have not yet added support for higher dimensional distributions of rewards, such as linear bandits, but it would be an interesting extension of SMPyBandits.
 %
@@
 For instance, most existing algorithms are index policies (see Algorithm~\ref{algo:2:indexPolicy}), and new ones are easily written by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
-They compute an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$, and play $A(t) = \arg\max_k I_k(t)$.
-For instance the code specific to $\UCB_1$ is as short as this (and fully documented):
+% They compute an index $U_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$, and play $A(t) \in \arg\max_k U_k(t)$.
+For instance the code specific to $\UCB_1$ (with $\alpha=2$) is as written and fully documented in a file as short as this:
 
@@
     % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy.}
-    \captionof{listing}{Code defining the UCB algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
+    \captionof{listing}{Code defining the $\UCB_1$ algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
     % \label{lst:3:smallIndexPolicy}
@@
 
-\subsection{Documentation}
+\subsection{Documentation of the library}
 
@@
 \texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
-It uses the Sphinx software \cite{sphinx}, and the content is directly written in the Python files as docstrings, so users have access to the documentation from within their IDE or the Python console.
-The most interesting component of the library is the many MAB algorithms being implemented: for each of them, the documentation gives a reference to a research paper (\eg, \cite{Kaufmann12BUCB} for \texttt{BayesUCB}), as well as a bird-eye view of its behavior.
+It uses the Sphinx software \cite{sphinx}, and the content is directly written in the Python files as docstrings (in \texttt{"""\dots"""}, see in the example of Code~\ref{lst:3:smallIndexPolicy} given above), so users have access to the documentation from within their IDE or the Python console.
+The most interesting component of the library is the many MAB algorithms being implemented: for each of them, the documentation gives a reference to a research paper (\eg, \cite{Auer02} for \texttt{UCB} in Code~\ref{lst:3:smallIndexPolicy}), as well as a bird-eye view of its behavior.
 For most algorithms, especially for index policies, the internal variables of the implementation are carefully linked to the notations of each paper, and formulas explaining the way the algorithm selects its arm are usually given.
@@
 
-\subsection{Dependencies}
+\subsection{Dependencies on other Python libraries}
 \label{sub:3:dependencies}
@@
 
-Optional dependencies include \texttt{joblib} \cite{joblib} for parallel simulations, \texttt{numba} \cite{numba} for automatic speed-up on small functions using a Just-in-Time compiler, \texttt{sphinx} \cite{sphinx} for generating the documentation, \texttt{virtualenv} \cite{virtualenv} for launching simulations in isolated environments, and \texttt{jupyter} \cite{jupyter} used with \texttt{ipython} \cite{ipython} to experiment with the code.
+Optional dependencies include \texttt{joblib} \cite{joblib} for parallel simulations, \texttt{numba} \cite{numba} for automatic speed-up on small functions (using a Just-in-Time compiler), \texttt{sphinx} \cite{sphinx} for generating the documentation, \texttt{virtualenv} \cite{virtualenv} for launching simulations in isolated environments, and \texttt{jupyter} \cite{jupyter} used with \texttt{ipython} \cite{ipython} to experiment with the code.
 
@@
 
-Since $2017$ and $2018$, SMPyBandits is using two online continuous integration (CI) service, to automatically build and host the documentation, and to automatically test the library on some numerical simulations.
+Since $2017$ and $2018$, SMPyBandits is using two online continuous integration (CI) services, to automatically build and host its documentation, and to automatically test the library on some numerical simulations.
 %
@@
 Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
-Travis CI allows to automatically run short numerical simulations after every commit.
-This allows to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
+It allows to automatically run short numerical simulations after every commit,
+in order to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
 The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
@@
 %
-Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
+Since its first use, the service did about $190$ builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
 The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
@@
 To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
-They are more recent (\eg, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or are based on a different point-of-view.
-The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with the state-of-the-art algorithms but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).
+They are more recent (\eg, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or based on a different point-of-view.
+The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with or more efficient than other recent algorithms, but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).
 
@@
 %
-We consider a randomly generated vector of means, each being sampled uniformly at random in $[0,1]$, with a minimum gap of $\Delta^{\min}$, $\bm{\mu} \sim \cE(\Delta^{\min},\mu_{\min},\mu_{\max})$, as well as minimum and maximum values of $\mu_{\min}$ and $\mu_{\max}$.
+We consider a randomly generated vector of means, each being sampled uniformly at random in $[0,1]$, with a minimum gap of $\Delta^{\min}$, as well as minimum and maximum values of $\mu_{\min}$ and $\mu_{\max}$, that is denoted by $\bm{\mu} \sim \cE(\Delta^{\min},\mu_{\min},\mu_{\max})$.
 This space $\cE_{\Delta^{\min}}$ is defined as $\{ \bm{\mu} \in [\mu_{\min}, \mu_{\max}]^K : \min_{k\neq k'} |\mu_k - \mu_{k'}| \geq \Delta^{\min} \}$.
@@
 We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
-As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
+As far as now, SMPyBandits is restricted to the finite-arm case, but it supports a wide range of arms distributions.
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
-One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
+One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research from December $2016$ to June $2019$.
 More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
@@
 Most of them were implemented quickly, but some would require more time and are left as possible future works.
-Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).
+Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits of \cite{Seznec2018} (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).
 
 Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
-Its documentation has seen already about $20000$ visits, the project on GitHub had $88$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
+Its documentation has seen already about $20000$ visits, the project on GitHub had $88$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
 %
-We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.
+We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to our work.
 
 As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
-I would also love to be able in the next to teach a graduate or under-graduate course on multi-armed bandits, and use my library as a support to illustrate the course, as well as practical sessions.
+I would also love to be able to teach a graduate or under-graduate course on multi-armed bandits in the next years, and use my library as a support to illustrate the course, as well as practical sessions.
 

### 2019-07-02 2-Chapters/3-Chapter/src/example_of_a_IndexPolicy_UCB.py

@@ -1,22 +1,19 @@
-from numpy import sqrt, log
-from .IndexPolicy import IndexPolicy
+from numpy import sqrt, log          # mathematical functions
+from IndexPolicy import IndexPolicy  # base class
 
 class UCB(IndexPolicy):
-  """ The UCB policy for bounded bandits.
-  Reference: [Lai & Robbins, 1985]. """
+    """ The UCB policy for bounded bandits.
+    Reference: [Lai & Robbins, 1985], [Auer et al. 2020]. """
 
-  def computeIndex(self, arm):
-    r""" Compute the current index, at time t and
-    after :math:`N_k(t)` pulls of arm k:
+    def computeIndex(self, arm):
+        r""" Compute the current index :math:`U_k(t)`, at time t (:attr:`self.t`)
+        and after :math:`N_k(t)` pulls of arm k (``arm``):
 
-    .. math::
-
-        I_k(t) = \frac{X_k(t)}{N_k(t)}
-        + \sqrt{\frac{2 \log(t)}{N_k(t)}}.
-    """
-    if self.pulls[arm] < 1:  # forced exploration
-      return float('+inf')   # in the first steps
-    else:                    # or compute UCB index
-      estimated_mean = (self.rewards[arm] / self.pulls[arm])
-      exploration_bias = sqrt((2 * log(self.t)) / self.pulls[arm])
-      return estimated_mean + exploration_bias
+        .. math:: U_k(t) = \frac{X_k(t)}{N_k(t)} + \sqrt{\frac{2 \log(t)}{N_k(t)}}.
+        """
+        if self.pulls[arm] < 1:    # forced exploration
+            return float('+inf')     # in the first steps
+        else:                      # or compute UCB index
+            estimated_mean = self.rewards[arm] / self.pulls[arm]
+            exploration_bias = sqrt((2 * log(self.t)) / self.pulls[arm])
+            return estimated_mean + exploration_bias

### 2019-07-02 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Done for tonight

@@
 
-As we said, a third possible approach is to select \emph{on the run} the best algorithm for a specific situation, for which the so called \emph{expert aggregation algorithms} framework can be useful.
+The aggregation approach is especially interesting if we know that the problem the application will face is one of a few known kinds of problems.
+In such cases, if there are $N$ different sorts of problems, and if the practitioner has prior knowledge on it, she can use the naive approach by selecting algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$.
+Then she can use the aggregation of $\cA_1,\dots,\cA_N$ when facing an unknown problem.
+%
+% As we said, a third possible approach is to select \emph{on the run} the best algorithm for a specific situation, for which the so called \emph{expert aggregation algorithms} framework can be useful.
 %
@@
 online selection of MAB algorithms was never discussed,
-and usually the considered algorithm is simply $\UCB_1$, without discussion regarding this choice.
+and usually the considered algorithm is simply $\UCB_1$, with no discussion regarding this choice.
 % and we show that it appears empirically sub-optimal when applied to simple stochastic problems.
@@
 However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $q(t) = (q_k(t))_{k\in[K]}$ over the $K$ arms at each time step, and draw the next selected arm according to this distribution.
-For instance, Thompson sampling uses its posterior distribution, or \UCB{} can use a distribution with a mass of $1/n$ if there are $n\geq2$ arms of maximal index, in case of tie in the $\argmax$.
+For instance, Thompson sampling uses its posterior distribution, or \UCB{} can use a distribution with a mass of $1/n$ on the $n \geq 1$ arms of maximal index, and a mass of $0$ on the other arms.
+% in case of tie in the $\argmax$.
 
-The aggregation algorithm maintains a probability distribution $\pi^{t}$ on the $N$ algorithms $\Alg_a$ (\ie, $\pi^{t}\in\Delta_N$ the simplex of dimension $N$), starting from a uniform distribution:
-$\pi^t_a$ is the probability of trusting the decision made by algorithm $\Alg_a$ at time $t$.
+The aggregation algorithm maintains a probability distribution $\pi^{t}$ on the $N$ algorithms $\Alg_a$,
+% (\ie, $\pi^{t}\in\Delta_N$ the simplex of dimension $N$)
+starting from a uniform distribution.
+The probability of trusting the decision made by algorithm $\Alg_a$ at time $t$ is thus $\pi^t_a$.
 $\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it first decides whom to trust by sampling $a \in [N]$ from $\pi^t$, then follows $\Alg_a$'s decision: $a \sim \pi^t, A_{\mathrm{aggr}}(t) = A(t) = A_a(t)$.
@@
 The main questions are then to know what observations (\ie, arms and rewards) should be given as feedback to which algorithms,
-and how to update the trusts at each step, and our proposal \Aggr{} differs from \ExpQ{} on these very points.
+and how to update the trusts at each step, and our proposal \Aggr{} differs from \ExpQ{} on these very two points.
 %
@@
 					$\Alg_a$ updates its internal state (e.g., \UCB{} indexes)\;
-					It chooses $A_a(t) \in [K]$.
+					It chooses $A_a(t) \in [K]$\;
 				}
@@
 \begin{equation}
-	\widehat{r^a}(t) \eqdef \frac{Y_{A_a(t),t}}{p_{A_a(t)}^{t}} \indic(A_a(t) = A(t))
+	\widehat{r^a}(t) \eqdef \frac{Y_{A_a(t),t}}{p_{A_a(t)}^{t}} \indic(A_a(t) = A(t)),
 \end{equation}
@@
 increase $\pi_{a}$ if $A_a(t) = A(t)$ and to \emph{decrease it otherwise}.
-It would not be so different, as there is a final renormalization step, and empirically this variation has little impact on the performance of \Aggr{}.
+It would not be so different, as there is a final renormalization step, and empirically this variation has little or no impact on the performance of \Aggr{}.
 
 % ----------------------------------------------------------------------
-\paragraph{Comparison with \ExpQ}.\label{sub:25:Exp4}
+\paragraph{Comparison with \ExpQ.}\label{sub:25:Exp4}
 %
@@
 %
-Thus we prefer the alternative horizon-free choice of learning rates,
+Thus we prefer an alternative horizon-free choice of learning rates,
 $\eta_t \eqdef \log(N) / (t K)$ suggested by \cite{Bubeck12}.
-It is non-increasing, and is obtained by minimizing the upper-bound on the regret derived in \cite[pp48]{Bubeck12}.
+It is non-increasing, and it is obtained by minimizing the upper-bound on the regret derived in \cite[pp48]{Bubeck12}.
 %
@@
 as a function of $T$, comparing some algorithms $\Alg_1,\dots,\Alg_N$ (for $N=6$), and their aggregation with \Aggr{} (displayed in orange bold),
-using the parameter-free learning rate sequence, $\eta_t \eqdef \log{N} / (t K)$.
+using the parameter-free learning rate sequence, $\eta_t \eqdef \log(N) / (t K)$.
 
@@
 %
-Note that for each of the $N=1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(r_k(t))_{k\in[K], t\in[T]}$, for each of the $N$ repetitions, in order to compare the algorithms on the same realizations of the MAB problem.
+Note that for each of the $N=1000$ simulations, we choose to generate all the rewards beforehand, \ie, one full matrix $(Y_{k,t})_{k\in[K], t\in[T]}$, for each of the $N$ repetitions, in order to compare the algorithms on the same realizations of the MAB problem.
 %
@@
 
-For each of the 4 problems considered, the \Aggr{} algorithm with default option (\ie, broadcast rewards to all the aggregated algorithms) is the best of all the aggregation algorithms,
-and its regret is very close to the best of the aggregated algorithms.
-Especially in difficult problems with mixed or unknown distributions, \Aggr{} showed to be more efficient that \ExpQ{} and orders of magnitude better than the other reference aggregation algorithms \LearnExp{} and \CORRAL{} (see Figures~\ref{fig:25:HarderMixed} and~\ref{fig:25:HarderMixed_semilogx}).
+For each of the 4 problems considered, our \Aggr{} algorithm  is the best of all the aggregation algorithms,
+and its regret is close to the best of the aggregated algorithms.
+Especially in difficult problems with mixed distributions, \Aggr{} showed to be more efficient that \ExpQ{} and orders of magnitude better than the other reference aggregation algorithms \LearnExp{} and \CORRAL{} (see Figures~\ref{fig:25:HarderMixed} and~\ref{fig:25:HarderMixed_semilogx}).
 

### 2019-07-02 2-Chapters/2-Chapter/chapter25.tex
Done for tonight

@@
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
-Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, another solution could be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+We propose an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
 
@@
 %
-Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to use.
+Thus, when a practitioner is facing a problem where MAB algorithms could be used, it is not an easy task to decide which algorithm to implement.
 It is hard to predict which solution could be the best for real-world conditions at every instant,
@@
 using a large $\alpha$, \ie, favoring exploration, will give low regret in the first case and high regret in the second case,
-while using a low $\alpha$, \ie, favoring exploitation, will obtain opposite performance.
-A first approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let a ``master'' learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
+while using a low $\alpha$, \ie, favoring exploitation, will obtain opposite performances.
+A first approach can be to use an intermediate value, as $\alpha=1/2$ suggested by theory, but another approach could be to consider an aggregated vote of different versions of $\UCB_1$, each running with a different value of $\alpha$ (\eg, in a logarithmic grid), and let a ``leader'' learning algorithm decide which value of $\alpha$ is the best \emph{for the problem at hand}.
 
@@
 
-Another possibility is to consider an online approach, which is interested in the case where the computation power or memory storage of the application is not a limitation factor, but where one cannot run benchmarks before deploying the application.
+The online approach is interesting in the case where the computation power or memory storage of the application is not a limitation factor, but where one cannot run benchmarks before deploying the application.
 We consider a fixed set of algorithms, and we use another learning algorithm on top of this set, to learn \emph{on the fly} which one should be trusted more, and eventually, used on its own.
-
-The aggregation approach is especially interesting if we know that the problem the application will face is one of a few known kinds of problems.
-In such cases, if there are $N$ different sorts of problems, and if the practitioner has prior knowledge on it, one can use the naive approach to select an algorithm $\cA_i$ which should perform well on problem $i$, for $i\in[N]$,
-and use the aggregation of $\cA_1,\dots,\cA_N$ when facing the unknown problem.
 
@@
 Empirically, \Aggr{} showed to always have a logarithmic
-regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
+regret if it aggregates at least one algorithm with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
 It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
-As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.
+As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this is an over-optimistic goal.
 
@@
 Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
-But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
+But this direction does not appear very useful for our objective at hand, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
 % We will continue investigating regret bounds for \Aggr,

### 2019-07-02 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done for tonight

@@
 %
-We show that using learning algorithms
-firstly extends devices battery autonomy (by reducing the packet-loss-ratio or equivalently, by improving the successful-communication rate),
-and secondly does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
-In the studied settings, stochastic MAB learning
-% provides an up to $16\%$ gain in terms of successful transmission probabilities, and
-has near optimal performance even in non-stationary settings with a majority of intelligent devices.
+% We show that using learning algorithms
+% firstly extends devices battery autonomy (by reducing the packet-loss-ratio or equivalently, by improving the successful-communication rate),
+% and secondly does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
+% In the studied settings, stochastic MAB learning
+% % provides an up to $16\%$ gain in terms of successful transmission probabilities, and
+% has near optimal performance even in non-stationary settings with a majority of intelligent devices.
 
@@
 So our first hypothesis is actually not constraining.
-We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
-This second approach is ideal, but not realistic the most common situation of decentralized co-located IoT networks, and it is just used here as a reference.
+We first evaluate the probability of collision if the dynamic devices randomly select their channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
+This second approach is ideal, but not realistic for the most common situation of decentralized co-located IoT networks, and it is just used here as a reference.
 Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
@@
 This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels,
-as each IoT protocol or IoT networks can choose to use a different subset of channels.
+as each IoT protocol or IoT network can choose to use a different subset of channels.
 %
 In order to simulate networks designed for the IoT, we consider a protocol \textbf{with no sensing} and no repetition of uplink messages.
-The gateway is in charge of sending back an acknowledgement, after some fixed-time delay, to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) who succeeded in sending an uplink packet.
+The gateway is in charge of sending back an acknowledgement, after some fixed-time delay, to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) which succeeded in sending an uplink packet.
 %
@@
     \emph{Dynamic} (or \emph{smart}) end-devices can use all the available channels, by quickly changing their communication channel at any time slot, following a Machine Learning policy.
-    For that end, they use their history of past communication successes or failures they experienced in each channel, to learn about channel availability.
+    For that end, they use their history of successes or failures of their past communication they experienced in each channel, to learn about channel availability.
     We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
-    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one float number for each of the $K$ channels).
+    Note that of course we assume a limited storage capacity, so no device stores the full history of communications successes and failures, but they only store average rates (which can be stored using two integers for each of the $K$ channels).
     Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, in order to improve their successful communication rate automatically.
@@
     \centering
-    \includegraphics[scale=0.65]{perf_learning.eps}
+    \includegraphics[scale=0.80]{perf_learning.eps}
     \caption{Learning with \UCB{} and Thomson Sampling, with many dynamic devices.
@@
 
+\newpage % XXX
+
 % \paragraph{Note on the simulation code.}

### 2019-07-02 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Done for tonight

@@
     % \item[\textcolor{darkred}{\circled{3}}]
-The last part is the interface of the base station, where we can see the traffic observed by the base station \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
-The observed traffic is coherent with \emph{(c)} of part (2), as the elements are very close the one from the others on the testbed.
+The last part is the interface of the base station, where we can see the traffic observed by the base station \circled{a} and the channels in which the last acknowledgements have been sent \circled{b}.
+The observed traffic is coherent with \circled{c} of part \textcolor{darkred}{\boldcircled{2}}, as the elements are very close the one from the others on the testbed.
 Colors may change, as they depend on the exact distance between the different transmitters and receivers.
+See more details in \cite{MoyBesson2019}.
 \end{enumerate}

### 2019-07-02 2-Chapters/4-Chapter/chapter4.tex
Done for tonight

@@
 % On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
-The first model
-% without retransmission of packets,
-is interesting for its simplicity,
-and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
-
+% The first model
+% % without retransmission of packets,
+% is interesting for its simplicity,
+% and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
+%
 However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
@@
 %
-The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
+% The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
 
@@
 Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
-As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly not stationary''.
-This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms (that is why we relax the hypothesis and only consider up-to $M \leq K$ players in Chapter~\ref{chapter:5}).
+As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``strongly not stationary''.
+This specific context makes it very hard to give mathematical proofs of convergence and of efficiency of bandit algorithms (that is why we relax the hypothesis and only consider up-to $M \leq K$ players in Chapter~\ref{chapter:5}).
 We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
 %
-We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.
+We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packets if the \emph{Ack} was not received.
 
@@
 %
-This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019}.
+This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019,MoyBesson2019Annales}.
 

### 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Done for §5.2 and §5.3

@@
 Since $2017$, S.
-Darak and his team at IIIT Delhi (India) have been very active in the research on CR using MAB, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+Darak and his team at IIIT Delhi (India) have been quite active in the research on CR using MAB, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
@@
     \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
-Similar gains of performance were obtained in many different scenarios.}
+Similar gains of performance were obtained in other scenarios.}
     \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
@@
 % They did not overlap each other, as
-When facing the same stationary background traffic, we see that the learning devices are both very quickly more efficient than the naive uniform device.
+When facing the same stationary background traffic, we see that the learning devices are both quickly more efficient than the naive uniform device.
 We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
 %
-In stationary environments, both the TS and \UCB{} algorithms are very efficient and converge quickly, resulting in a very strong decrease in collisions and failed communication slots.
+In stationary environments, both the TS and \UCB{} algorithms are efficient and converge quickly, resulting in a strong decrease in collisions and failed communication slots.
 \UCB{} is faster to learn but eventually TS gives a (slightly) better average performance.

### 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for §5.2 and §5.3

@@
 The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Figure~\ref{fig:43:Markov_model}.
-When a device has a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{\mathrm{MaxBackOff}-2} \}$, at each $\mathrm{MaxBackOff}$ back-off stage.
+When a device has a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{c\mathrm{MaxBackOff}-2} \}$, at each $\mathrm{MaxBackOff}$ back-off stage.
 At each time slot, a transition from an idle state to a transmission state (denoted as \texttt{Trans.}) occurs if a packet transmission is required, while waiting states (denoted as \texttt{Wait}), correspond to a $m$ back-off interval.
@@
 	% \includegraphics[width=0.70\linewidth]{Markov_model-eps-converted-to.pdf}
-	\caption{The Markov model of the behavior of all devices paired to the considered IoT network using the ALOHA protocol.}
+	\caption{The Markov model of the behavior of all devices paired to the considered IoT network using the ALOHA protocol. Here, $\mathrm{MaxBackOff}$ is written $M$, to have smaller labels.}
 	\label{fig:43:Markov_model}

### 2019-07-03 1-Preamble/4-Acknowledgement/acknowledgement.tex
Done for thanks and Ch4

@@
 I would then like to thank my closest friends, sorted in chronological order of our first encounter in the \href{https://bulbapedia.bulbagarden.net/wiki/Tall_grass}{tall grass}\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
-And due to space constraints, I had to choose quite arbitrarily whom to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
+And due to space constraints, I had to choose quite arbitrarily whom to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Audrey, Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
 % In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
@@
 
-Thanks to my wonderful colleagues during these three years: Rémi\textsuperscript{\tiny{\Coffeecup}}, Pascal\textsuperscript{\tiny{\Coffeecup}}, Yves, Amor, Carlos Faouzi, Jacques P., Jacques W., Ali C., Haïfa, Marwa, Navik, Quentin\textsuperscript{\tiny{\Coffeecup}}, Vincent\textsuperscript{\tiny{\Coffeecup}}, Rami\textsuperscript{\tiny{\Coffeecup}}, Muhammad, Cristo\textsuperscript{\tiny{\Coffeecup}}, Adrien\textsuperscript{\tiny{\Coffeecup}}, Esteban\textsuperscript{\tiny{\Coffeecup}}, Ali Z., Éloïse\textsuperscript{\tiny{\Coffeecup}}, Nabil\textsuperscript{\tiny{\Coffeecup}}, Bastien\textsuperscript{\tiny{\Coffeecup}}, Corentin\textsuperscript{\tiny{\Coffeecup}}, Georgios, Morgane\textsuperscript{\tiny{\Coffeecup}} in Rennes ; and Alessandro, Ronan, Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Julien, Mathieu, Matteo, Mahsa, Nicolas, Édouard in Lille.
-I want to thank especially the ``coffee friends\textsuperscript{\tiny{\Coffeecup}}'' in Rennes, for the daily interesting discussions.
+Thanks to my wonderful colleagues during these three years: Rémi\textsuperscript{\tiny{\Coffeecup}}, Pascal\textsuperscript{\tiny{\Coffeecup}}, Yves, Amor, Carlos Faouzi, Jacques P., Jacques W., Ali C., Haïfa, Marwa, Navik, Quentin\textsuperscript{\tiny{\Coffeecup}}, Vincent\textsuperscript{\tiny{\Coffeecup}}, Rami\textsuperscript{\tiny{\Coffeecup}}, Pierre H., Muhammad, Cristo\textsuperscript{\tiny{\Coffeecup}}, Adrien\textsuperscript{\tiny{\Coffeecup}}, Esteban\textsuperscript{\tiny{\Coffeecup}}, Ali Z., Éloïse\textsuperscript{\tiny{\Coffeecup}}, Nabil\textsuperscript{\tiny{\Coffeecup}}, Bastien\textsuperscript{\tiny{\Coffeecup}}, Corentin\textsuperscript{\tiny{\Coffeecup}}, Julio, Georgios, Morgane\textsuperscript{\tiny{\Coffeecup}} in Rennes ; and Alessandro, Ronan, Julien P., Odalric-Ambrym, Philippe, Olivier, Florian, Xuedong, Guillaume, Benjamin, Julien S., Mathieu, Matteo, Mahsa, Nicolas, Sadegh, Édouard, Yannis and Omar in Lille.
+I want to especially thank the ``coffee friends\textsuperscript{\tiny{\Coffeecup}}'' in Rennes, for our daily interesting discussions.
 %  the best moments of our life at the office.
 %
-Thanks to everybody who played \emph{futsal} on Wednesday lunch, I still have to progress immensely before I could pretend to be useful, but it's been a lot of fun to play with you guys!
+Thanks to everybody who played \emph{``futsal''} on Wednesday lunch,
+% I still have to progress immensely,
+% before I could pretend to be useful
+but it's been a lot of fun to play with you guys!
 
@@
 %,
-\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis link $\nearrow$}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{$\nearrow$}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{$\nearrow$}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{~Que je remercie deux fois plus que ce qu'il me remerciant dans sa propre thèse !
+\href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis link $\nearrow$}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{$\nearrow$}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{$\nearrow$}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{~Que je remercie deux fois plus que ce qu'il me remerciait dans sa propre thèse !
 Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{$\nearrow$}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{$\nearrow$}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{$\nearrow$}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{$\nearrow$}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{$\nearrow$}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{$\nearrow$}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, and Rami Othman.
 
-Thanks to some people for their administrative support regarding research, travelling or teaching: Karine, Jeannine, Grégory, Anne, Cécile, Maud, Frédéric, Gabriel (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline (and others) at CRIStAL in Lille.
-Thanks to the nice people at Sodexo in the CentraleSupélec campus of Rennes, especially to Martine and J-B, your daily smile was priceless!
+For their administrative support regarding research, travelling or teaching,
+thanks to Karine, Jeannine, Grégory, Anne, Cécile, Maud, Frédéric, Gabriel (and others) at CentraleSupélec campus of Rennes, Amélie at Inria Lille and Maryline (and others) at CRIStAL in Lille.
+Thanks to the nice people at Sodexo in our campus, especially to Martine and J.-B.!
+%  your daily smile was priceless!
 
@@
 I warmly thank David Pichardie for his trust and support, and also François Schwarzentruber, Nathalie Bertrand and David Cachera at ENS de Rennes, and Romaric Gaudel and his colleagues at ENSAI.
-Thanks to them, I loved all of the $3 \times 64$ hours I spent teaching, since October $2016$.
+Thanks to them, I loved to teach these $3 \times 64$ hours.
+% since October $2016$.
 % I am trully honored to join \emph{the Département d'Informatique} at ENS de Rennes in next September $2019$.
-Warm thanks to my amazing students, I am proud of you and of our interesting exchanges.
+Thanks to my amazing students, I am proud of you and of our interesting exchanges.
 %  and I feel lucky to be able to keep working alongside some of you.
-I am grateful to Amélie Stainer and Gilbert Gabillard at Lycée Chateaubriand and Joliot-Curie for useful exchanges.
+I am thankful to Amélie Stainer and Gilbert Gabillard at Lycée Chateaubriand and Joliot-Curie for useful discussions.
 \\
 \indent
-I am thankful to all the amazing professors who shared their passion for mathematics and computer science with me.
+I am grateful to all the amazing professors who shared their passion for mathematics and computer science with me.
 A partial list includes Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers, Marseille.
@@
 in Computer Science, and to Carine S.-P., Delphine L.
-and the late Nicolas P.
-for their support.
+and the late Nicolas P.\textsuperscript{$\dagger$} for their support.
 I also thank Didier C., Vijaysekhar C. and Jai P. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$).
@@
 
+\emph{Je suis désolé si je vous ai oublié, merci de corriger} : Je remercie sincèrement ....................... pour .............
+\emph{I am sorry I forgot you, please fix this mistake}: I warmly thank .................... for ..........
+
+\vfill{}
+
 % I would like to conclude these acknowledgements by a special thought toward Lola, who was of great support on a daily basis, who marvelled me by her enthusiasm and dynamism, and who comforted me with her laughs and love, in the last year.
@@
 %
-\small{
-    \emph{Deseo terminar este agradecimiento con un pensamiento especial para mi gatita \emph{L.}, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con sus risas y su amor, durante el año pasado.}\\
-    \indent
+% \small{
+    \emph{Deseo terminar este agradecimiento con un pensamiento especial para mi gatita \emph{L.}, quien me apoyó diariamente, me sorprendió con su entusiasmo y me consoló con sus risas y su amor, durante el año pasado.
+¡Muchísimas gracias, gatita!}
     \textcolor{white}{$\heartsuit$ Te quiero, Lola, ¿quieres vivir conmigo?}
-}
+% }
 

### 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for thanks and Ch4

@@
 \graphicspath{{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/}}
+
+% In the context of Cognitive Radio \cite{Mitola99,Haykin05},
+% Multi-Arm Bandit (MAB) algorithms \cite{Auer02,Auer02,Bubeck12} have been recently proposed as a potential solution for channel access in LPWA networks \cite{Bonnefoi18,Azari18,Bonnefoi17}.
+We presented in Section~\ref{sec:4:firstModel} low-cost algorithms following well-known approaches, such as Upper-Confidence Bound (\UCB), and we have reported encouraging results.
+When considering the application of MAB algorithms for slotted wireless protocols in a decentralized manner,
+other recent directions of research include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
+or realistic empirical PoC like in \cite{RobertSDR2014,modiDemo2016,darak2016bayesian,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
+and finally applications to multi-hoping networks \cite{Mitton16,Toldov16} or other kinds of networks \cite{Azari18,Wilhelmi19collaborative,Wilhelmi19potential}.
+%
+None of the mentioned works discuss the impact of retransmissions on the performance of MAB learning algorithms.
 
@@
 
-The presented strategies allow IoT devices to improve their access to the network and their autonomy, while taking into account the impact of encountered radio collisions.
-For that end, several heuristics employing \UCB{} algorithms are examined, to explore the information provided by the number of retransmissions.
-In this section, our results show that approaches based on \UCB{} obtain a significant improvement in terms of successful transmission probabilities, compared to a naive approach which does not learn.
-Furthermore, it also reveals that a pure \UCB{} channel access is as efficient as more sophisticated learning strategies.
+% The presented strategies allow IoT devices to improve their access to the network and their autonomy, while taking into account the impact of encountered radio collisions.
+% For that end, several heuristics employing \UCB{} algorithms are examined, to explore the information provided by the number of retransmissions.
+% In this section, our results show that approaches based on \UCB{} obtain a significant improvement in terms of successful transmission probabilities, compared to a naive approach which does not learn.
+% Furthermore, it also reveals that a pure \UCB{} channel access is as efficient as more sophisticated learning strategies.
 
@@
 
-% In the context of Cognitive Radio \cite{Mitola99,Haykin05},
-% Multi-Arm Bandit (MAB) algorithms \cite{Auer02,Auer02,Bubeck12} have been recently proposed as a potential solution for channel access in LPWA networks \cite{Bonnefoi18,Azari18,Bonnefoi17}.
-We presented in Section~\ref{sec:4:firstModel} the impact of non-stationarity on the network performance using MAB algorithms is studied.
-Low-cost algorithms following two well-known approaches, such as the Upper-Confidence Bound (\UCB{}), and the Thompson Sampling (TS) algorithms have reported encouraging results.
-When considering the application of MAB algorithms for slotted wireless protocols in a decentralized manner,
-other recent directions of research include theoretical analysis, like what we present in the next Chapter~\ref{chapter:5},
-or realistic empirical PoC like in \cite{MoyWSR2014,RobertSDR2014,modiDemo2016,darak2016bayesian,kumar2016two,kumar2017channel} or the previous Section~\ref{sec:4:gnuradio},
-and finally applications to multi-hoping networks \cite{Mitton16,Toldov16} or other kinds of networks \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
-%
-None of the mentioned works discuss in detail the impact of retransmissions on the performance of MAB learning algorithms as we do now.
-
-The aim of this section is to assess the performance of MAB algorithms for channel selection in LPWA networks operating in unlicensed bands, while taking into account the impact of retransmissions on the network performance.
+We want to assess the performance of MAB algorithms for channel selection in LPWA networks operating in unlicensed bands, while taking into account the impact of retransmissions on the network performance.
 For this reason, several decision making strategies are applied after a first retransmission (\ie, when a collision occurs).
 The proposed approach employs contextual information provided by the number of retransmissions, and is again implemented independently by each device, so that no coordination among them is needed.
-Moreover, our \UCB{}-based heuristics show low complexity making them suitable for being embedded in LPWA devices.
+Moreover, our \UCB{}-based heuristics again show low complexity, which make them suitable for being embedded in LPWA devices, like in the previous sections.
 
-The contributions of this section are summarized as follows.
+The contributions of this section can be thus summarized as follows.
 % \begin{itemize}
@@
 
-Like before in this chapter, we consider an LPWA network composed of a gateway and a large number of end-devices that regularly send short data packets, where $K$ channels ($K>1$) are available for the transmission of their packets.
+Like in the previous sections of this chapter, we consider an LPWA network \cite{Raza17}, composed of a gateway and a large number of end-devices that regularly send short data packets, where $K$ channels ($K>1$) are available for the transmission of their packets.
 %
@@
 The aforementioned contention process can be described by a Markov chain model \cite{Norris98} similar to the one presented in \cite{Yang12}, as it is depicted in Figure~\ref{fig:43:Markov_model}.
-When a device has a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $\{p_{c}, p_{c1}, \dots, p_{c\mathrm{MaxBackOff}-2} \}$, at each $\mathrm{MaxBackOff}$ back-off stage.
+When a device has a packet to transmit, it goes from an idle state to a transmission state, while considering retransmissions due to different collision probabilities, $p_{c}, p_{c1}, \dots, p_{c\mathrm{MaxBackOff}-2}$, at each $\mathrm{MaxBackOff}$ back-off stage.
 At each time slot, a transition from an idle state to a transmission state (denoted as \texttt{Trans.}) occurs if a packet transmission is required, while waiting states (denoted as \texttt{Wait}), correspond to a $m$ back-off interval.
@@
 	% \includegraphics[width=0.70\linewidth]{Markov_model-eps-converted-to.pdf}
-	\caption{The Markov model of the behavior of all devices paired to the considered IoT network using the ALOHA protocol. Here, $\mathrm{MaxBackOff}$ is written $M$, to have smaller labels.}
+	\caption{The Markov model of the behavior of all devices paired to the considered IoT network using the ALOHA protocol.
+\emph{Note}: $\mathrm{MaxBackOff}$ is written $M$, to have smaller and simpler labels.}
 	\label{fig:43:Markov_model}
@@
 %
-In Figure~\ref{fig:43:Approximation_m10}, we show the collision probabilities for different number of devices $N$ (from $N=50$ up-to $N=400$), for both $p_{c}$ and $p_{c1}$.
+In Figure~\ref{fig:43:Approximation_m10}, we show the collision probabilities for different number of devices $N$ (from $N=50$ to $N=400$), for both $p_{c}$ and $p_{c1}$.
 %
-From this simulations, we can verify that our approximation is very precise for lower values $p_{c1} \leq 30 \%$ (\ie, red and orange curves are quite close).
+We can verify that our approximation is very precise for lower values $p_{c1} \leq 30 \%$ (\ie, red and orange curves are quite close).
 Moreover, a significant gap between $p_{c1}$ and $p_c$,
@@
 	% \includegraphics[width=0.65\linewidth]{Approximation_m10.eps}
-	\includegraphics[width=0.70\linewidth]{Approximation_m10-eps-converted-to.pdf}
+	\includegraphics[width=0.60\linewidth]{Approximation_m10-eps-converted-to.pdf}
 	\caption[Proposed approximation for the probability of collision at the second transmission.]{Proposed approximation for the probability of collision at the second transmission.
@@
 \begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
-	\textbf{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}:}
+	% \begin{small}  % XXX WARNING
+		\textbf{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}.}
 		%
-	In all the algorithms and notations presented in this Section, the time steps $t$ do not refer to the \emph{global} time steps (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
-	Each device wakes up whenever it has to send some data, following its random emission process (here, we propose a Bernoulli process of small probability, \eg, $p=10^{-3}$), and when it woke up, it sends its packet, and will send again for at most $\mathrm{MaxBackOff}$ times (\eg, $\mathrm{MaxBackOff}=10$) until it receives an \Ack.
+		In the algorithms presented in this chapter, the time step $t$ does not refer to the \emph{global} time step (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
+		Each device wakes up whenever it has to send some data, following its own emission process (we restrict to a Bernoulli process of small probability, \eg, $p=10^{-3}$), and whenever it wakes up, it sends its packet, and will send again for at most $\mathrm{MaxBackOff}$ times (\eg, $\mathrm{MaxBackOff}=10$) until it receives an \Ack.
 		%
-	In other words, the time steps $t$ in the following algorithms denote the number of time the device sent an uplink packet, and these uplink (re)transmissions can happen at different (real-world) times, more or less spread out in time, it does not matter to the learning part of this model and work.
+		In other words, the time steps $t$ in the following algorithms denote the number of uplink packets sent by the device, and these uplink transmissions or retransmissions can happen at different (global or real-world) times.
+		The local index $t$ is independent on the global time (\ie, the time of the gateway), and IoT device does not need to care about this difference.
+		From the device point-of-view, the learning occurs at successive times (whenever the device wakes up), regardless of the global time.
+	% \end{small}  % XXX WARNING
 \end{leftbar}  % XXX leftbar warningbar, comment if needed
@@
 % --- --- --- --- --- --- --- --- --- ---
-\textbf{Uniform random retransmission ``(\textcolor{cyan}{Random}'').}\label{sub:43:UCBthenRandom}
+\paragraph{Uniform random retransmission ``(\textcolor{cyan}{Random}'').}\label{sub:43:UCBthenRandom}
 %
@@
 % --- --- --- --- --- --- --- --- --- ---
-\textbf{\UCB{} for retransmission (``\textcolor{purple}{\UCB{}}'').}\label{sub:43:TwoUCB}
+\paragraph{\UCB{} for retransmission (``\textcolor{purple}{\UCB{}}'').}\label{sub:43:TwoUCB}
 %
@@
 Note that, we use the subscript $({}^r)$ to denote the variables
-$\widehat{\mu^r}(t)$, $B^r_k(t)$ and $U^r_k(t)$,
+$\widehat{\mu}^r(t)$, $B^r_k(t)$ and $U^r_k(t)$,
 related to the \UCB{} algorithm employed for the retransmission.
@@
 			}
-			\uElse(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^r$}){
-				Compute $\forall k, U^r_k(t) = \widehat{\mu^r_k}(t) + \sqrt{\alpha \log(t) / N_k^r(t)}$\;
+			\uElse(\tcp*[f]{Packet retransmission with the other $\mathrm{UCB}^r$}){
+				Compute $\forall k, U^r_k(t) = \widehat{\mu_k}^r(t) + \sqrt{\alpha \log(t) / N_k^r(t)}$\;
 				Transmit in channel $C^r(t) \sim \cU(\argmax_k U^r_k(t))$\;
@@
 % --- --- --- --- --- --- --- --- --- ---
-\textbf{$K$ different {\UCB}s for retransmission (``\textcolor{green}{$K$ \UCB}'')}\label{sub:43:UCBthenKp1}
+\paragraph{$K$ different {\UCB}s for retransmission (``\textcolor{green}{$K$ \UCB}'')}\label{sub:43:UCBthenKp1}
 %
@@
 			}
-			\uElse(\tcp*[f]{Packet retransmission with $\mathrm{UCB}^j$}){ % Retr. after trying channel $j$
+			\uElse(\tcp*[f]{Packet retransmission with one of the $K$ $\mathrm{UCB}^j$}){ % Retr. after trying channel $j$
 				j $\leftarrow$ last channel selected by first-stage \UCB\;
@@
 % --- --- --- --- --- --- --- --- --- ---
-\textbf{Delayed \UCB{} for retransmission (``\textcolor{red}{Delayed \UCB}'').}\label{sub:43:UCBwithDelay}
+\paragraph{Delayed \UCB{} for retransmission (``\textcolor{red}{Delayed \UCB}'').}\label{sub:43:UCBwithDelay}
 %
@@
 			\uElse(\tcp*[f]{Packet retransmission with delayed $\mathrm{UCB}^d$}){
-				Compute $\forall k, U^d_k(t) = \widehat{\mu^d_k}(t) + \sqrt{\alpha \log(t) / N_k^d(t)}$\;
+				Compute $\forall k, U^d_k(t) = \widehat{\mu_k}^d(t) + \sqrt{\alpha \log(t) / N_k^d(t)}$\;
 				Transmit in channel $A^d(t) \sim \cU(\argmax_k U^d_k(t))$\;
@@
 
-\textbf{Other ideas that we did not explore.}
+\paragraph{Other ideas that we did not explore.}
 %
@@
 
-\textbf{Using another bandit policy?}
+\paragraph{Using another bandit policy?}
 %
@@
 We simulate our network considering $N$ devices following the contention Markov process described in Section~\ref{sub:43:model}, and a LoRaWAN standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
-Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of $1\;\mathrm{s}$.
+Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of \SI{1}{\second}.
+% $1\;\mathrm{s}$.
 %
@@
 In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform repartition of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
-In other words, the channels are occupied\footnote{~Note that we consider higher occupancy rates that the ones considered previously in Section~\ref{sec:4:gnuradio}, because the goal of this experimental section is to evaluate our approach that use learning to optimize the way each device retransmits its packets.
-We need to have a large enough probability $p_c$ of retransmission if we want to see a difference between the different learning heuristics, which is why we consider a network more densily occupied than in Section~\ref{sec:4:gnuradio}.}
+In other words, the channels are occupied\footnote{~Note that we consider higher occupancy rates that the ones considered previously in Section~\ref{sec:4:gnuradio}, because the goal of this experimental section is to evaluate our approach that uses learning to optimize the way each device retransmits its packets.
+We need to have a large enough probability $p_c$ of retransmission if we want to see a difference between the different learning heuristics, which is why we consider a network more densely occupied than the one considered in Section~\ref{sec:4:gnuradio}.}
 respectively $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $\mathrm{MaxBackOff} = 5$, and $m=5$.
@@
 
-A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``\textbf{no \UCB{}}'' curve in black), which is (still) the current state-of-the-art in IoT networks.
+A first result is that all the heuristics clearly outperform the non-learning approach that simply uses random channel selection for both transmissions and retransmissions (\ie, the ``no \UCB{}'' curve in black), which is (still) the current state-of-the-art in IoT networks.
 The improvement of the heuristics over the non-learning approach is clear, and for every heuristic that uses a kind of learning mechanism it can be observed that the successful transmission rate increases rapidly (or equivalently the PLR decreases).
@@
 		The usage of the same learning policy for transmissions and retransmission is named ``\textcolor{blue}{Only \UCB{}}'',
-		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as ``\textbf{no \UCB{}}''.
+		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as ``no \UCB{}''.
 		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.

### 2019-07-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for thanks and Ch4

@@
 As we saw in Chapters~\ref{chapter:1} and \ref{chapter:4},
-a crucial step for the development of Cognitive Radio is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
+a crucial step for the development of Cognitive Radio is to insert \emph{multiple} smart devices (at least $M \geq 2$), in the \emph{same} environment, or background traffic.
 %
@@
 A perfect time and frequency synchronization is assumed,
-and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
+and some \iid{} background traffic is assumed to be non-uniformly distributed in the $K$ channels.
 As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
@@
 %
-With the presence of a central controller that can assign the devices to separate channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
+With the presence of a central controller that can assign the devices to different channels, this amounts to choosing at each time step \emph{several} arms of a MAB in order to maximize the global rewards, and can thus be viewed as an application of the multiple-play bandit, introduced by \cite{Anantharam87a} and recently studied by \cite{Komiyama15}.
 They essentially proved that existing algorithms can be easily extended to the multiple-play case, with provable guarantees on their regret.
@@
 However some real radio networks do not use sensing at all, \eg, emerging standards currently or recently developed for \emph{Internet of Things} (IoT) networks, such as LoRaWAN.
-Thus, to take into account these new applications, algorithms with additional constraints on the available feedback have to be proposed within the multiple-player MAB model.
+Thus, to take into account these new applications, algorithms with additional constraints on the available feedback have to be proposed, within the multiple-player MAB model.
 Especially, the typical approach that combines a (single-player) bandit algorithm based on the sensing information --to learn the quality of the channels while targeting the best ones-- with a low-complexity decentralized collision avoidance protocol, is no longer possible.
@@
 We do not propose any other efficient algorithm, but our work presented in April~$2018$ has strongly inspired two articles shortly after \cite{LugosiMehrabian18,BoursierPerchet18}.
-They confirmed our finding that \Selfish{} can have a linear regret, and they both proposed new algorithms.
+They confirmed our findings that \Selfish{} can have a linear regret, and they both proposed new algorithms.
 We include some numerical simulations to compare some of them, and we present in details the current state-of-the-art of research on multi-players MAB models without sensing.
@@
 % First we motivate the model of multi-players multi-armed bandits in Section~\ref{sec:5:motivation}
-We introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}.
+We first introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}.
 We present a useful regret decomposition in Section~\ref{sec:5:lowerbound}, and illustrate it in the Appendix~\ref{sec:5:appendix}.
@@
 for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
-The result of our experimental study reported in Section~\ref{sec:5:experiments}.
+We report the results of our experimental study in Section~\ref{sec:5:experiments}.
 Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
@@
 we consider a $K$-armed Bernoulli bandit model, % (for $K\geq2$),
+of horizon $T \geq 2$,
 in which arm $k$ is a Bernoulli distribution with mean $\mu_k\in[0,1]$.
-We denote $(Y_{k,t})_{t\in\N}$ the \iid{} (binary) \emph{reward stream} for arm $k$, that satisfies $\Pr(Y_{k,t}=1) = \mu_k$ and that is independent from the other rewards streams.
+We denote $(Y_{k,t})_{t\in[T]}$ the \iid{} (binary) \emph{reward stream} for arm $k$, that satisfies $\Pr(Y_{k,t}=1) = \mu_k$ and that is independent from the other rewards streams.
 
@@
 However we mention that our lower bound and all our algorithms (and their analysis) can be easily extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}).
-For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio problems, based on channel availability.
+For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio problems, by considering channel availability.
 
@@
   \end{equation}
-  that respectively indicate that a collision occurs at time $t$ for player $j$ and that a collision occurs at time $t$ on arm $k$.
+  that respectively indicate that a collision occurs at time $t$ for player $j$ ($C^j(t)$, $j$ in superscript), and that a collision occurs at time $t$ on arm $k$ ($C_k(t)$, $k$ in subscript).
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
 %
-Note that other models for rewards loss have also been proposed in the literature, for instance the reward is randomly allocated to one of the players selecting it.
-To stay consistent and for simplicity, with the model presented in the previous Chapter~\ref{chapter:4},
-we preferred to focus on full reward occlusion in this chapter.
+Note that other models for rewards loss have also been proposed in the literature, for instance the reward could be randomly allocated to one of the players selecting it.
+To stay consistent with the model presented in the previous Chapter~\ref{chapter:4},
+and for simplicity, we preferred to focus on full reward occlusion in this chapter.
 
@@
 
-Following a natural approach in the bandit literature, we evaluate the performance of a multi-players strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
+Following a natural approach in the bandit literature, like in Section~\ref{sec:2:lowerUpperBoundsRegret},
+we evaluate the performance of a multi-players strategy using the \emph{expected regret} (later simply referred to as regret), that measures the performance gap with respect to the best possible strategy.
 The regret of the strategy $\cA$ at horizon $T$ is the difference between the cumulated reward of an oracle strategy, assigning in this case the $M$ players to \Mbest,
-and the cumulated reward of strategy $\cA$:
+and its cumulated reward:
 
@@
   \begin{equation}\label{eq:5:regret}
-    R_T(\boldsymbol{\mu}, M, \cA) \eqdef \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
+    R_T^{\cA}(\boldsymbol{\mu}, M) \eqdef \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M r^j(t)\right].
   \end{equation}
@@
     \begin{align}\label{eq:5:termeundeuxtermetrois}
-      R_T(\boldsymbol{\mu}, M, \cA) &=
+      R_T^{\cA}(\boldsymbol{\mu}, M) &=
       \underbrace{\sum_{k \in \Mworst} (\mu_M^* -  \mu_k) \E_{\mu}[N_k(T)]}_{\emph{\mytag{(a)}{eq:5:term1}}} \\
@@
   % \end{small}
+  \vspace*{-5pt}  % XXX
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 \begin{smallproof}
-  Using the definition of regret $R_T(\boldsymbol{\mu}, M, \cA)$ from \eqref{eq:5:regret}, denoted here $R_T$, and this collision indicator $\eta^j(t) \eqdef \indic(\overline{C^j(t)})$,
+  Using the definition of regret $R_T^{\cA}(\boldsymbol{\mu}, M)$ from \eqref{eq:5:regret}, denoted here $R_T$, and this collision indicator $\eta^j(t) \eqdef \indic(\overline{C^j(t)})$,
   \begin{align*}
@@
      = \left(\sum_{k=1}^{M}\mu_k^*\right)T - \E_{\mu}\left[\sum_{t=1}^T\sum_{j=1}^M \mu_{A^j(t)} \eta^j(t)\right]
-    \intertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence from $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$).
+    \intertext{The last equality comes from the linearity of expectations, and the fact that $\E_{\mu}[Y_{k,t}] = \mu_k$ (for all $t$, from the \iid{} hypothesis), and the independence with $A^j(t)$, $\eta^j(t)$ and $Y_{k,t}$ (observed \emph{after} playing $A^j(t)$).
 So $\E_{\mu}[Y_{A^j(t),t} \eta^j(t)] = \sum_{k} \E_{\mu}[\mu_k \indic(A^j(t),t) \eta^j(t)] = \E_{\mu}[\mu_{A^j(t)} \eta^j(t)]$.
@@
     %
-    \intertext{Let $\overline{\Delta_k} \eqdef \overline{\mu}^* - \mu_k$ be the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), then by splitting $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$, we get}
+    \intertext{If $\overline{\Delta_k} \eqdef \overline{\mu}^* - \mu_k$ is the gap between the mean of the arm $k$ and the $M$-best average mean, and if $M^*$ denotes the index of the worst of the $M$-best arms (\ie, $M^* = \arg\min_{k\in\Mbest}(\mu_k)$), we can split $[K]$ into three disjoint sets $\Mbest \cupdot \Mworst = (\Mbest\setminus\{M^*\}) \cupdot \{M^*\} \cupdot \Mworst$}
     %
@@
     \begin{align*}
-        R_T(\boldsymbol{\mu}, M, \cA)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[N_k(T)].
+        R_T^{\cA}(\boldsymbol{\mu}, M)    \geq \sum\limits_{k \in \Mworst} (\mu_M^*- \mu_k) \E_{\mu}[N_k(T)].
     \end{align*}%
@@
   \end{equation*}
-  We focus on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T(\boldsymbol{\mu}, M, \cA)$ from Lemma~\ref{lem:5:DecompositionRegret},
+  We focus on the two terms $\ref{eq:5:term2} + \ref{eq:5:term3}$ from the decomposition of $R_T^{\cA}(\boldsymbol{\mu}, M)$ from Lemma~\ref{lem:5:DecompositionRegret},
   \begin{align*}
@@
   And now both terms are non-negative, as $N_k^!(T) \leq T$, $\min(\mu_M^*, \mu_k)\geq 0$, and $C_k(T) \geq 0$, so $\ref{eq:5:term2} + \ref{eq:5:term3} \geq 0$
-  which proves that $R_T(\boldsymbol{\mu}, M, \cA) = \ref{eq:5:term1} + \ref{eq:5:term2} + \ref{eq:5:term3} \geq \ref{eq:5:term1}$, as wanted.
+  which proves that $R_T^{\cA}(\boldsymbol{\mu}, M) = \ref{eq:5:term1} + \ref{eq:5:term2} + \ref{eq:5:term3} \geq \ref{eq:5:term1}$, as wanted.
 \end{smallproof}
@@
   \begin{align}
-    \forall a\in(0,1) & R_T(\boldsymbol{\mu},M,\cA) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
-    \text{and } \; \forall a\in(0,1) & \;
+    \forall a\in(0,1), & R_T^{\cA}(\boldsymbol{\mu},M) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
+    \text{and } \; \forall a\in(0,1), & \;
     \forall j \in [M], k \in \Mbest, \;\;
@@
 %
-From this assumption $R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
+From this assumption $R_T^{\cA}(\boldsymbol{\mu},M)=o(T^\alpha)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
 ${T}- \E_{\mu}[N_k(T)] {=} o(T^{\alpha})$, and so
@@
 The notion of fairness defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
-For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is known by the algorithm.
+For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of bandit game, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K] \in [0,1]^K$ is known by the algorithm.
 
@@
 Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
-We denote $\Pr_{\bm{\mu}}^{O_t^j}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\cA_j$.
+We denote $\Pr_{\bm{\mu}}^{O_t^j}$ the distribution of the vector $O_t^j$ under the model $\bm{\mu}$ when using algorithm $\cA_j$.
 Then we introduce the \emph{collision information term} as:
 \begin{equation}\label{eq:5:collisionInformationTerm}
-  \cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) \eqdef \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}^j}).
+  \cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) \eqdef \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}^j}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}^j}).
 \end{equation}
@@
   \begin{equation}\label{eq:5:ourLowerBound}
-    \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
+    \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T^{\cA}(\boldsymbol{\mu}, M)}{\log(T)}
     \geq M \times \left( \sum_{k \in \Mworst} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_M^*)} \right) .
@@
 \begin{equation}\label{eq:5:Zhao10LowerBound}
-  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
+  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T^{\cA}(\boldsymbol{\mu}, M)}{\log(T)}
   \geq \sum_{k \in \Mworst} \left( \sum_{j=1}^{M} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_{j}^*)} \right),
@@
   \begin{equation}\label{eq:5:LogarithmicRegret_MCTopMklUCB}
-    R_T(\boldsymbol{\mu}, M, \cA) \leq G_{M,\boldsymbol{\mu}} \log(T) + \smallO{\log T}.
+    R_T^{\cA}(\boldsymbol{\mu}, M) \leq G_{M,\boldsymbol{\mu}} \log(T) + \smallO{\log T}.
   \end{equation}

### 2019-07-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Done for thanks and Ch4

@@
 \begin{equation}
-  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log T} \geq C_{\boldsymbol{\mu}, M}.
+  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T^{\cA}(\boldsymbol{\mu}, M)}{\log T} \geq C_{\boldsymbol{\mu}, M}.
 \end{equation}
@@
 The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} below
-shows the regret $R_T(\boldsymbol{\mu}, M, \cA)$
+shows the regret $R_T^{\cA}(\boldsymbol{\mu}, M)$
 on the same example problem $\boldsymbol{\mu}$, with $K = 9$ arms and respectively $M = 6$, or $9$ players, for \Selfish-\klUCB.

### 2019-07-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for Chap5

@@
   \end{equation}
-\end{leftbar}  % XXX leftbar defnbar, comment if needed
-\end{definition}
-
   Note that when $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one.
@@
   Hence there is small abuse of notation when calling it a number of collisions.
-%
-By denoting $\cP_M \eqdef \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
-the set of bandit instances with a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.
+\end{leftbar}  % XXX leftbar defnbar, comment if needed
+\end{definition}
+
+%
+% By denoting $\cP_M \eqdef \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
+% the set of bandit instances with a strict gap between the $M$ best arms and the other arms,
+We now provide a regret decomposition for any bandit instances with a strict gap between the $M$ best arms and the other arms (\ie, $\mu_M^* > \mu_{M+1}^*$).
 
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-  For any bandit instance $\boldsymbol{\mu}\in\cP_M$ such that $\mu_M^* > \mu_{M+1}^*$, it holds that
+  For any bandit instance $\boldsymbol{\mu}\in\cP_M  \eqdef \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$, it holds that
   % \vspace*{-5pt}  % XXX
@@
     \end{align}
+    % \vspace*{-5pt}  % XXX
+    In this decomposition, term \ref{eq:5:term1} counts the lost rewards due to \emph{sub-optimal arms} selections ($k \in \Mworst$), term \ref{eq:5:term2} counts the number of times the \emph{best arms} were \emph{not} selected ($k \in \Mbest$), and term \ref{eq:5:term3} counts the weighted number of collisions, on \emph{all arms}.
   % \end{small}
-  \vspace*{-5pt}  % XXX
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
 \end{lemma}
-
-In this decomposition, term \ref{eq:5:term1} counts the lost rewards due to \emph{sub-optimal arms} selections ($k \in \Mworst$), term \ref{eq:5:term2} counts the number of times the \emph{best arms} were \emph{not} selected ($k \in \Mbest$), and term \ref{eq:5:term3} counts the weighted number of collisions, on \emph{all arms}.
 
@@
   \begin{align}
-    \forall a\in(0,1), & R_T^{\cA}(\boldsymbol{\mu},M) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
+    \forall a\in(0,1), & R_T^{\cA}(\boldsymbol{\mu},M) \mathop{=}\limits_{T \to +\infty} o(T^\alpha), \\
     \text{and } \; \forall a\in(0,1), & \;
@@
 
-\textbf{Collision information.}
+\paragraph{Collision information.}
 %
@@
 Similarly to what we did in Section~\ref{sec:2:notations},
-consider the observations $O_t^j$ that she gathered after $t$ rounds of its algorithm $\cA_j$ for a fixed player $j\in[M]$,
+consider the observations $O_t^j$ that player $j$ gathered after $t$ rounds of its algorithm $\cA_j$ for a fixed player $j\in[M]$,
 defined as $O_t^j \eqdef \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
@@
 at least the arm replacing the $k\in\TopM(t-1)$ in $\TopM(t)$.
-Collisions are dealt with only for non-fixed player $j$,
+Collisions are dealt with only for non-fixed players $j$,
 and when the previous arm is still in $\TopM(t)$.
@@
 %
-The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM, and they are illustrated in Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below.
+The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM{} (in the next section), and they are illustrated in Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below.
 
@@
   \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
-  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
+  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t)$ using information of previous steps.}
   \label{fig:5:StateMachineAlgorithm_MCTopM}
@@
 for our two proposals \RandTopM{} and \MCTopM{} as well as for \rhoRand{}, when they are combined with \klUCB{} indices,
-and Section~\ref{sub:5:UpperBoundCollisions} proves that the number of collisions are logarithmic, and hence the regret of \MCTopM{} is also logarithmic.
+and then we restrict to the most efficient algorithm \MCTopM, by proving in Section~\ref{sub:5:UpperBoundCollisions} that the number of collisions are logarithmic, and hence the regret of \MCTopM{} is also logarithmic.
 %
@@
   The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
-  \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : U_m^j(t) < U_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
-  Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
-  If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
-  In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.
+  \begin{equation}\label{eq:5:KeyInclusion}
+    \left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : U_m^j(t) < U_k^j(t) \right).
+  \end{equation}
+
+  For each algorithm, the arm selected at time $t$ always belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
+  Selecting the sub-optimal arm $k$ at time $t$ implies that $k \in \TopM(t)$, and that one of the arms in \Mbest{} must be excluded from $\TopM(t)$, because there are $M$ arms in both \Mbest{} and $\TopM(t)$.
+  In particular, arm $k$ must have a larger index than this particular arm $m$.
 
@@
   Considering the relative position of the upper-confidence bound $U_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition
-  \begin{align*}
+  \begin{align}
     \E_{\mu}[N_k^j(T)] &\leq
-    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; U_{m_1^*}(t) < \mu_m^* \right) + \\
+    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; U_{m_1^*}(t) < \mu_m^* \right) + \nonumber \\
     \;\;\;\;
-    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq U_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right)
-  \end{align*}
-  \begin{align}
+    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq U_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \nonumber \\
+  % \end{align}
+  % \begin{align}
     \E_{\mu}[N_k^j(T)] &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq U_k(t)\right) \notag \\
@@
       % \begin{equation}
-      $\sum\limits_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log(\log(T))$.
+      $\sum\limits_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4 \mathrm{e} \log(\log(T))$.
       % \end{equation}
@@
     & \leq \frac{\log(T) + 3 \log(\log(T))}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log(\log(T))}  \nonumber\\
-    & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log(\log(T)) + 3M+1,
+    & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4M \mathrm{e} \log(\log(T)) + 3M+1,
   \end{align}
@@
 Note that this bound is in $\bigO{M^3}$,
-which significantly improves the $\bigO{M{2M-1 \choose M}}$ proven by \cite{Anandkumar11} for \rhoRand.
-It is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}.
-%, due to our trick of focusing on collisions for non-sitted players.
+which significantly improves over the $\bigO{M{2M-1 \choose M}}$ bound, proven by \cite{Anandkumar11} for \rhoRand.
+Our bound is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}. %, due to our trick of focusing on collisions for non-sitted players.
 However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity.
-It is indeed not very satisfying, from an applicative point of view, to require a prior knowledge
-of $\mu^*_{M}-\mu^*_{M+1}$ if one wants to run the \MusicalChair{} algorithm.
+The \MusicalChair{} algorithm is indeed not very satisfying from an applicative point of view, as it requires a prior knowledge of $\mu^*_{M}-\mu^*_{M+1}$.
 
@@
 This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
-On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret.
+On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches, in order to reduce by a certain multiplicative factor its switching cost while multiplying its regret by the same factor.
 While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
@@
 Now that the sub-optimal arms selections and the collisions
-are both proven to be at most logarithmic in Lemmas~\ref{lem:5:SubOptimalSelections} and \ref{lem:5:collisionsMCTopM},
+are both proven to be at most logarithmic in Lemmas~\ref{lem:5:SubOptimalSelections} and \ref{lem:5:collisionsMCTopM} above,
 it follows from our regret decomposition (Lemma~\ref{lem:5:DecompositionRegret}) together with Lemma~\ref{lem:5:1stUpperBound} that the regret of \MCTopM-\klUCB{} is logarithmic.
-More precisely, one obtains a finite-time problem-depend upper bound on the regret of this algorithm.
+More precisely, we obtain a finite-time problem-depend upper bound on the regret of this algorithm.
 
@@
 %
-As soon as $R_T = \bigO{\log T}$ for all problems, \MCTopM{} is clearly proven
-to be uniformly efficient, as $\log T$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
+As soon as $R_T = \bigO{\log(T)}$ for all problems, \MCTopM{} is clearly proven
+to be uniformly efficient, as $\log(T)$ is $\smallO{T^{\alpha}}$ for any $\alpha\in(0,1)$.
 %
@@
 % -----------------------------------------------------------------
-\section{Experimental results for multi-player bandits, with sensing information}
+\section{Experimental results for multi-player bandits with sensing}
 \label{sec:5:experiments}
@@
 Our proposed algorithms, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
+
 We also include a (unrealistic) centralized multiple-play \klUCB{} algorithm,
@@
 essentially to check that the \emph{``price of decentralized learning''} is not too large.
+Centralized multiple-play index policies are very similar to single-player index policies (see Algorithm~\ref{algo:2:indexPolicy}), they work by computing an index $U_k'(t)$ for each arm $k\in[K]$, and then sort the arm by decreasing index, and affect the $M$ players to the $M$ arms with highest index.
+They are based on a central supervision, as a single algorithm is in charge of deciding the arms used by the $M$ players, and as such they are both unrealistic to set-up in practice, and falling outside of our focus that is on decentralized algorithms.
 
@@
 with \MEGA{} \citep{Avner15} and \MusicalChair{} \citep{Rosenski16}, in the presence of sensing, \ie, observation model \modeldeux, for which they were developed.
-Yet these two algorithms were found hard to use efficiently in practice and we show in
+Yet these two algorithms were found to both be very hard to use efficiently in practice, and we show in
 % that were also proposed

### 2019-07-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Done for Chap5

@@
 $\log T$, is asymptotically lower bounded by a constant $C_{\boldsymbol{\mu}, M}$
-depending on the problem $\boldsymbol{\mu}$ and the number of players $M$, for any $\cA$.
+depending on the problem $\boldsymbol{\mu}$ and the number of players $M$, for any algorithm $\cA$ of a certain class.
+%
+We remind that if  $\cA=(\cA_1,\dots,\cA_M)$ is a strongly uniformly efficient \emph{decentralized} policy,
+and if $\cA_j$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$, then we obtained the following asymptotical result:
+%
 \begin{equation}
@@
 %
-For an example problem with $K = 9$ arms, we display below on the $x$ axis is
+For an example problem with $K = 9$ arms, we display below on the $x$ axis
 the number of player, from $1$ player to $9$ players, and on the
-$y$ axis is the value of this constant $C_{\boldsymbol{\mu}, M}$, from the initial
+$y$ axis the value of this constant $C_{\boldsymbol{\mu}, M}$, from the initial
 theorem and from our theorem.

### 2019-07-03 2-Chapters/5-Chapter/chapter5.tex
Proofreading intro
Done for Chap5

@@
 % ----------------------------------------------------------------------------
-\section{Literature review of many extensions of the models}
+\section{Literature review of extensions of the multi-player MAB model}
 \label{sec:5:literatureReviewOtherModels}
@@
 % ----------------------------------------------------------------------------
-\subsection{With communication or coordination between players}
+\subsection{With direct communication or coordination between players}
 \label{sub:5:withCommunicationOrCoordination}
@@
 %
-In the easier case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO(M K \log(K T) + K M^3 \log(M K T) / \Delta) + K M^2 (\log(M \log(K T) / \Delta))^2$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
+In the easier case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO\left( M K \log(K T) + K M^3 \log(M K T) / \Delta + K M^2 (\log(M \log(K T) / \Delta))^2 \right)$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
 In the generic case, their algorithm M-ETC with Elimination achieves a regret of $\cO((\log(T))^{1+\kappa})$ for any positive constant $\kappa$.
@@
 This chapter suggests several interesting further research directions.
-First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
+First, one could want to investigate the notion of \emph{optimal algorithm}, in the decentralized multi-players model with sensing information.
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
 However, sub-optimal draws turn out to not be the dominant term in the regret, both in our upper bound and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
-Similarly to what was done very recently in \cite{wang2019distributed}, for communications between players who collaborate with each other, it would be interesting to characterize the (minimal) number of collisions needed to achieve logarithmic regret.
+Similarly to what was done very recently in \cite{wang2019distributed}, for communications between players who collaborate with each other, it would be interesting to characterize the (minimal) number of collisions needed to achieve logarithmic regret in this decentralized model.
 

### 2019-07-03 2-Chapters/6-Chapter/chapter6.tex
\vfill{} before \minitoc at beginning of all Chapters
Small improvements on layout of abstract/minitoc for chapters
Lots of small changes. I'll start fixing Chapter 2 this weekend, today my goal is Ch1
Typo
Done for Chap5

@@
 
-\paragraph{Abstract.}
-
+\abstractStartChapter{}%
+%
 In this last chapter, we study a generalization of the stationary multi-armed bandit model,
@@
 %
-of its \emph{false alarm probability} and \emph{detection delay}, and can be combined with an efficient bandit policy (\klUCB), to propose an efficient algorithm for piece-wise stationary problems, \GLRklUCB.
+on its \emph{false alarm probability} and \emph{detection delay}, and can be combined with an efficient bandit policy (\klUCB), to propose an efficient algorithm for piece-wise stationary problems, \GLRklUCB.
 % Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
@@
 %
-Finally, we compare our approach with other state-of-the-art solutions on large numerical experiments.
+Finally, we compare our approach with other state-of-the-art solutions on different numerical experiments.
 % , on different piece-wise stationary bandit problems.

### 2019-07-03 2-Chapters/Conclusion/conclusion.tex
Done for Chap5

@@
 In Chapter~\ref{chapter:4}, we first proposed some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
-In two different models, with or without retransmissions, we showed that end-devices can learn independently to increase their successful transmission rates in the network, by using low-cost MAB algorithms.
+In two different models, with or without retransmissions, we showed that end-devices can independently learn to increase their successful transmission rates in the network, by using low-cost MAB algorithms.
 We also presented a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.

### 2019-07-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Typo
Lots of tiny typos on appendix

@@
               }
-              \caption[The \RandTopM{} decentralized learning policy.]{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
+              \caption{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
           \label{algo:5:RandTopM}
@@
               }
-              \caption[The \MCTopM{} decentralized learning policy.]{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
+              \caption{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
           \label{algo:5:MCTopM}
@@
 \rhoRand{} (top \textcolor{blue}{blue} curve) is outperformed by the other algorithms (and the gain increases with $M$).
-\MCTopM{} (bottom \textcolor{gold}{yellow}) outperforms all the other algorithms is most cases.}
+\MCTopM{} (bottom \textcolor{gold}{yellow}) outperforms all the other algorithms in most cases.}
   % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
@@
   % ~
+  \vspace{15pt}%  FIXME?
   % \begin{subfigure}[!h]{1.00\textwidth}
@@
   % ~
+  \vspace{20pt}%  FIXME?
   % \begin{subfigure}[!h]{1.00\textwidth}

### 2019-07-03 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Lots of tiny typos on appendix

@@
         }
-        \caption{\GLRklUCB, with \textbf{Local} or \textbf{Global} restarts.}
+        \caption{The \GLRklUCB{} algorithm, with \textbf{Local} or \textbf{Global} restarts.}
         \label{algo:6:GLRklUCB}
@@
     \includegraphics[width=1.15\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
-    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem 1.}
+    \caption{Mean regret as a function of time, $R_t$ for horizon $T = 5000$, for problem 1.}
     \label{fig:6:meanRegretPb1}
@@
     \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
-    \caption{Mean regret as a function of time, $R_t$ ($1 \leq t \leq T = 5000$) for problem $2$.}
+    \caption{Mean regret as a function of time, $R_t$ for horizon $T = 5000$, for problem $2$.}
     \label{fig:6:meanRegretPb2}

### 2019-07-03 2-Chapters/Conclusion/conclusion.tex
Lots of tiny typos on appendix

@@
 %
-For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret bounds obtained in our theoretical analyses.
+For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in our theoretical analyses.
 % \end{itemize}
@@
 
-We started to present the multi-armed bandit model, with a finite number of arms and real-valued rewards.
+We started by presenting the multi-armed bandit model, with a finite number of arms and real-valued rewards.
 Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
@@
 %
-Our library is used to run numerical simulations on Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
+Our library is used to run numerical simulations in Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
 
@@
 %  as the base station receives more accurately the data sent by the end-devices since less transmissions failed.
-We show that both \UCB{} and Thompson sampling algorithms have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which all follow a random activation process.
+We show that \UCB{} and Thompson sampling algorithms both have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which all follow a random activation process.
 %
@@
 We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
-Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
+Results obtained in practice were discussed, to highlight the interest of using learning algorithms for cognitive radio online optimization problems, and especially multi-armed bandit learning algorithms.
 %
@@
 %
-We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
-Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
-Surprisingly, the simulations showed that a simple \UCB{} learning approach, that simply retransmits in the same channel, turns out to perform almost as well as more complicated heuristics, that can change the channel used for retransmissions.
+We showed that incorporating learning for the first transmissions can help to achieve optimal performance, with significant gains in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
+Our simulations showed that the proposed heuristics outperform a naive random access scheme.
+Surprisingly, we showed that a simple \UCB{} learning approach, that simply retransmits in the same channel, turns out to perform almost as well as more complicated heuristics, that can change the channel used for retransmissions.
 
@@
 with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
-For the two easiest models --with sensing, our theoretical contribution improves the state-of-the-art upper bounds on the regret.
+For the two easiest models (\ie, with sensing), our theoretical contribution improves the state-of-the-art upper bounds on the regret.
 In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
@@
 %
-We first describe the break-point (or change-point) detection problem, that we solve by analyzing at finite-time the properties of the Bernoulli GLR change-point detector.
+We first describe the break-point (or change-point) detection problem, that we solve by the Bernoulli GLR change-point detector, and by analyzing its properties at finite-time.
 This sequential break-point detection test can be applied for bounded rewards, and we bound its false alarm probability as well as its detection delay.
@@
 A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
-It is already possible to reproduce some of the experiments presented in this thesis, by using the Jupyter notebooks \cite{jupyter}, hosted on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}, which can be used locally if you install the library, but can also be used on cloud platforms, such as
-Binder\footnote{~You can try on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} from your browser.}
-or Google Colab\footnote{~See \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}.}.
+It is already possible to reproduce some of the experiments presented in this thesis,
+by following the instructions given in the documentation (see Section~\ref{sub:3:listResearchWorksUsingSMPyBandits}),
+by using the Jupyter notebooks \cite{jupyter} made available
+\footnote{~These notebooks are hosted on both the GitHub repository of SMPyBandits and on my website on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}.
+They can be used locally if you install the library, but can also be used on cloud platforms, such as Binder, on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} or Google Colab, on \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}, from your browser.}.
 
@@
 
-The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device.
+The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if the $S$ static and $D$ dynamic devices have different transmission patterns, and less easily with one probability per device.
 Also, other emission patterns could be considered, instead of a Bernoulli process for each user.
@@
 It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
-It could be exciting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
+Thus, it could be exciting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
 

### 2019-07-03 5-Nomenclature/Nomenclature.tex
Lots of tiny typos on appendix

@@
 \nomenclature[L]{$T$}{Time horizon, the duration of the bandit game (always $T\geq1$)}
-\nomenclature[L]{$T_0$, $T_1$}{Fixed durations of some algorithms based on different phases, \eg, the Musical Chair algorithm from \cite{Rosenski16}}
+\nomenclature[L]{$T_0$, $T_1$}{Fixed durations of some algorithms based on different phases, \eg, for Explore-then-Exploit or Musical Chair from \cite{Rosenski16}}
 \nomenclature[L]{$\cA$}{An algorithm, also referred to as a policy or a strategy.
@@
 \nomenclature[L]{$Y_{k,t}$}{Random sample from the arm $k$ at time $t$}
-\nomenclature[L]{$O_t$}{Vector of observations until time $t$ in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$O_t$, $O_t^j$}{Vector of observations until time $t$ in Chapters~\ref{chapter:2} and \ref{chapter:5} (for player $j$)}
 \nomenclature[L]{$r(t)$, $r^j(t)$}{Reward obtained at time $t$, for user $j$ at time $t$}
@@
 
-\nomenclature[L]{$R_T$, $R_T^{\cA}$}{Regret (of an algorithm $\cA$) for horizon $T$}
+\nomenclature[L]{$R_T$, $R_T^{\cA}$, $R_T^{\cA}(I)$}{Regret of an algorithm $\cA$ for horizon $T$ (on instance $I$)}
 
@@
 \nomenclature[L]{$\widehat{\mu}_k(t)$}{Empirical mean of rewards obtained for arm $k$ at time $t$}
-\nomenclature[L]{$I_k(t)$}{Index of arm $k$ at time $t$ for an index policy}
+\nomenclature[L]{$U_k(t)$}{Index of arm $k$ at time $t$ for an index policy}
 \nomenclature[L]{$\UCB_k(t)$}{Upper-Confidence Bound (\UCB) of arm $k$ at time $t$ for an index policy}
-\nomenclature[L]{$g_k^j(t)$, $g^j(t)$}{Index of (arm $k$) at time $t$ of user $j$ for an index policy in Chapter~\ref{chapter:5}}
+\nomenclature[L]{$U_k^j(t)$, $U^j(t)$}{Index of (arm $k$) at time $t$ of user $j$ for an index policy in Chapter~\ref{chapter:5}}
 
@@
 \nomenclature[O]{$\cW$}{Lambert $\cW$ function, the first branch of the inverse of $x\mapsto x \exp(x)$, cf. \cite{Corless96}}
-\nomenclature[O]{$f$, $g$, $h$}{Real-valued functions, \eg, exploration function used for \klUCB{} indexes, $f(t) = \log(t) + 3 \log(\log(t))$}
+\nomenclature[O]{$f$, $g$, $h$}{Real-valued functions, \eg, the exploration function used for \klUCB{} indexes, $f(t) = \log(t) + 3 \log(\log(t))$}
 
@@
 \nomenclature[O]{$X'$}{Usually denotes a ``wrong value'' of a variable $X$, for instance $M'$ denotes in Section~\ref{par:5:usingWrongValueofM}. Also denotes the derivative of a function, \eg, $f'$}
-\nomenclature[O]{$\widehat{X}$}{Usually denotes an ``empirical value'', a mean or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{\mu_k}$ the empirical mean of arm $k$ or $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:towardsRegretUpperBounds}}
-\nomenclature[O]{$\widetilde{X}$}{Usually denotes another ``empirical value'' or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{S}_t$ the set of selected arms in Section~\ref{sub:5:Selfish}.
-Also denote a surrogate for a function without a closed form, \eg, $\widetilde{T}$ in Section~\ref{sec:6:exploringDifferentThresholdFunctions}}
+\nomenclature[O]{$\widehat{X}$}{Usually denotes an ``empirical value'', a mean or an estimate of a quantity $X$ that depends on time, \eg, $\widehat{\mu_k}$ the empirical mean of arm $k$}
+\nomenclature[O]{$\widetilde{X}$}{Usually denotes another ``empirical value'' or an estimate of a quantity $X$ that depends on time, \eg, $\widetilde{S}_t$ the set of selected arms in Section~\ref{sub:5:Selfish}.
+Also denotes a surrogate for a function without a closed form, \eg, $\widetilde{T}$ in Section~\ref{sec:6:exploringDifferentThresholdFunctions}}
 

### 2019-07-03 2-Chapters/1-Chapter/chapter1.tex
Last pass on all the manuscript, I want to improve description using legends

@@
 For example, in the city of Cesson-Sévigné where my campus of CentraleSupélec is located, a high RF tower emits on many standards, and in particular the \SI{92.3}{\mega\hertz} band is allocated to a French classical music radio (\emph{Radio Classique}).
-Broadcast-based systems are not only used for TV or radio,
-and for instance two other famous and worldwide deployed examples of mono-directional radio systems are the long-range global navigation satellite systems (such as USA's GPS or European Union's Galileo), and short-range infrared remote controllers.
+Broadcast-based systems are not only used for TV or radio:
+two other famous and worldwide deployed examples of mono-directional radio systems are the long-range global navigation satellite systems (such as USA's GPS or European Union's Galileo), and short-range infrared remote controllers.
 % https://en.wikipedia.org/wiki/Satellite_navigation
@@
 We study two models of IoT networks in Chapter~\ref{chapter:4}.
-The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use acknowledgements as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the less occupied by the surrounding traffic).
+The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use acknowledgements as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the channel less occupied by the surrounding traffic).
 The goal of this first model is to increase the Quality of Service (QoS) of the application of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmission.

### 2019-07-03 2-Chapters/2-Chapter/chapter2.tex
Last pass on all the manuscript, I want to improve description using legends

@@
     \includegraphics[width=0.85\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem.png}
-    \caption[Screenshot of the demonstration for a current step of $t=24$.]{Screenshot of the demonstration available online on my website, for a current step of $t=24$.}
+    \caption[Screenshot of the demonstration for a current step of $t=24$.]{Screenshot of the demonstration available on my website, for a current step of $t=24$.}
     \label{fig:2:example_of_a_5_arm_bandit_problem}
@@
     \caption[Screenshot of the demonstration, at the end of the game after $T=100$ steps.]{Screenshot of the demonstration, at the end of the game after $T=100$ steps, where the player suffers from an empirical regret of $R^{\UCB_1}_T = \mu^* T - 56 = 0.7 T - 56 = 14$ after following the $\UCB_1$ algorithm.
-        You can try it on \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
+        You can try it online at \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
     \label{fig:2:example_of_a_5_arm_bandit_problem__step100}
@@
 Following this Definition~\ref{def:2:KLDivergence} for two distributions
-$d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$,
+$d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ in the same exponential family,
 we can use a shorter notation and write $\KL( \eta_1, \eta_2 ) \eqdef \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
@@
 and we note that our library SMPyBandits implements the two parts in both a naive Python implementation (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback.py}{\texttt{kullback.py}}) and in a Python module written in a
-C file\footnote{~Note that we also wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
-    These implementations of $\KL(x,y)$ and \klUCB{} indexes are also published as an independent Python package, hosted on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} and installable from Pypi.
-    A Julia version of the same package is also available at \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}}.
+C file\footnote{~We also wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
+    These implementations are also published an independent Python and Julia packages, on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} (also on Pypi),
+    and on \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}}.
     % and installable with \texttt{Pkg.add("KullbackLeibler")}.

### 2019-07-03 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Last pass on all the manuscript, I want to improve description using legends

@@
 This second approach is ideal, but not realistic for the most common situation of decentralized co-located IoT networks, and it is just used here as a reference.
-Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
+Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communications rate, as it reflects the network efficiency.
 We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic, more and more fluctuating and unpredictable.
@@
     Note that of course we assume a limited storage capacity, so no device stores the full history of communications successes and failures, but they only store average rates (which can be stored using two integers for each of the $K$ channels).
-    Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, in order to improve their successful communication rate automatically.
+    Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, in order to improve their successful communications rate automatically.
 \end{itemize}
@@
 On the opposite, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$).
-Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
+Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communications rate) will be so close to zero that using learning algorithms cannot improve much.
 Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to \emph{have an average of active devices not larger\footnote{~In Chapter~\ref{chapter:5} we simply consider the case of $p=1$ and $M = D \leq K$ dynamic devices, referred to as \emph{players}.} than the number of channels}.
@@
 are $S_1, \dots, S_K$ which represent the (stationary) occupancy of the channels by the static devices,
-and they are unknown to each dynamic device, so to maximize their cumulated rewards, they must learn the distributions of the channels, in order to be able to progressively improve their respective successful communication rate.
+and they are unknown to each dynamic device, so to maximize their cumulated rewards, they must learn the distributions of the channels, in order to be able to progressively improve their respective successful communications rate.
 %

### 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Last pass on all the manuscript, I want to improve description using legends

@@
 %
-This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent devices at least do not reduce their successful communication rate in comparison to the naive devices.
+This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent devices at least do not reduce their successful communications rate in comparison to the naive devices.
 This also shows that in case of non-uniform stationary traffic, MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent devices.
@@
 %
-For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (``\textcolor{green}{TS}'', in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
+For one dynamic device, three algorithms are compared by their mean successful communications rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (``\textcolor{green}{TS}'', in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).
 
@@
     \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
-    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
+    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
 Similar gains of performance were obtained in other scenarios.}
@@
 When facing the same stationary background traffic, we see that the learning devices are both quickly more efficient than the naive uniform device.
-We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
+We obtain an improvement in terms of successful communications rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
 %

### 2019-07-03 2-Chapters/2-Chapter/chapter2.tex
Done for 2day

@@
 C file\footnote{~We also wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
-    These implementations are also published an independent Python and Julia packages, on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} (also on Pypi),
-    and on \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}}.
+    These implementations are also published an independent Python package, on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} (also on Pypi),
+    as well as a Julia package, on \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}}.
     % and installable with \texttt{Pkg.add("KullbackLeibler")}.
@@
 
-\paragraph{Variants on $\UCB_1$.}
+\textbf{Variants on $\UCB_1$.}
 %
@@
 \begin{table}[!h]
-    \begin{small}  % WARNING
+    % \begin{small}  % WARNING
         \centering
@@
         \label{table:2:oneShotResult}
-    \end{small}  % WARNING
+    % \end{small}  % WARNING
 \end{table}
@@
 \begin{table}[ht]
-    \begin{small}  % WARNING
+    % \begin{small}  % WARNING
         \centering
@@
         \label{table:2:meanResults}
-    \end{small}  % WARNING
+    % \end{small}  % WARNING
 \end{table}
@@
 	% \centering
-    \includegraphics[width=1.14\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
+    \includegraphics[width=1.23\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
 	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
@@
 	% \centering
-	\includegraphics[width=1.14\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
+	\includegraphics[width=1.17\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
 	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
@@
 %
-For both directions, we present in the two final chapters natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
+For both directions, we present in the two final chapters of this thesis natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
 of stationary multi-players and piece-wise stationary bandits.
@@
 %
-We focused in this thesis on finite-arms and one-dimensional bandit problems,
-and thus two possible directions of future works could be to extend our works
-to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
+We focus in this thesis on finite-arms and one-dimensional bandit problems,
+and therefore, two possible directions of future works could be to extend our works
+to MAB models with either multidimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
 

### 2019-07-03 2-Chapters/3-Chapter/chapter3.tex
Lots of tiny typos on appendix
Done for 2day

@@
 	\includegraphics[width=0.75\linewidth]{exampleOfRewards.pdf}
-	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$, $0.5$ and $0.9$.}
+	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
 	\label{fig:3:exampleOfRewards}
@@
 configuration["policies"] = [
-  {"archtype": klUCB, "params": {"klucb": klucbBern}},
-  {"archtype": klUCBPlusPlus,
-   "params": {"horizon": T, "klucb": klucbBern}},
-  {"archtype": UCBalpha, "params": {"alpha": 1}},
-  {"archtype": Thompson, "params": {"posterior": Beta}}
+    { "archtype": klUCB, "params": { "klucb": klucbBern } },
+    { "archtype": klUCBPlusPlus, "params": {"horizon": T, "klucb": klucbBern}},
+    { "archtype": UCBalpha, "params": { "alpha": 1 } },
+    { "archtype": Thompson, "params": { "posterior": Beta } }
 ]
@@
 	\caption[Example of a single-player simulation showing the average regret of $4$ algorithms.]{
-		Example of a single-player simulation showing the average regret of four algorithms.
+		Example of a single-player simulation showing the average regret of four algorithms (\textcolor{red}{\UCB}, \textcolor{blue}{\KLUCBpp}, \textcolor{darkgreen}{\klUCB} and \textcolor{gold}{Thompson sampling}).
 They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proven to match the lower-bound up-to a constant), and each but UCB is known to be asymptotically optimal (\ie, with the constant matching the lower-bound).
@@
 	\caption{Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}.
-For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).}
+For instance, \textcolor{gold}{Thomson sampling} is very efficient in average (\textcolor{gold}{in yellow}), and \textcolor{red}{\UCB} shows a larger variance (\textcolor{red}{in red}).}
 	\label{fig:3:firstPlot_hist}
@@
 \subsection{List of research works using SMPyBandits}
+\label{sub:3:listResearchWorksUsingSMPyBandits}
 
@@
 To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
-They are more recent (\eg, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or based on a different point-of-view.
+They are more recent (\eg, $3$ from $2018$ and $2$ from $2019$) and are usually more complex, or based on a different point-of-view.
 The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with or more efficient than other recent algorithms, but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).
@@
 
-\subsection{Computational time.}
+\subsection{Computational time}
 
@@
 
-\subsection{Memory cost.}
+\subsection{Memory cost}
 

### 2019-07-03 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done for 2day

@@
     \includegraphics[width=0.70\linewidth]{system_model1.eps}
-    \caption[In our system model, some dynamic devices (in the IoT network) transmit packets to a gateway and suffer from the interference generated by neighboring networks.]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange} left/right).}
+    \caption[In our system model, some dynamic devices (in the IoT network) transmit packets to a gateway and suffer from the interference generated by neighboring networks.]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange left/right}).}
     \label{fig:41:system_model1}
@@
     \label{fig:41:100intelligent}}
-    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
+    \caption[Performance of two MAB algorithms, \UCB{} and Thompson Sampling in red, compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms, \textcolor{blue}{\UCB{} in blue} and \textcolor{red}{Thompson Sampling in red}, compared to extreme reference policies without learning or \textcolor{gold}{oracle knowledge}, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:from10to100}
@@
 
-We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%.
+We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red $+$}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue} $*$), when the number of end-devices is below 50\%.
 When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
@@
     \includegraphics[scale=0.80]{perf_learning.eps}
-    \caption{Learning with \UCB{} and Thomson Sampling, with many dynamic devices.
+    \caption[Learning with \UCB{} and Thomson Sampling, with many dynamic devices]{Learning with \textcolor{red}{\UCB} and \textcolor{gold}{Thomson Sampling}, with many dynamic devices.
 The learning gain, for each device, decreases with the proportion of dynamic devices in the network.
-        Note that the values ($16\%$ etc) depend on the repartition of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.
-        % Both algorithms achieve close to optimal performance after a reasonable learning time.
-    }
+Note that the values ($16\%$ etc) depend on the repartition of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.}
     \label{fig:41:perf_learning}
@@
     \label{fig:41:ch2_100}}
-    \caption{Performance of the UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
+    \caption[Performance of the \UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.]{Performance of the \textcolor{blue}{\UCB} bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:figure4appendix}

### 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Done for 2day

@@
     \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
-    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
+    \caption{Less than $400$ communication slots (\ie, less than $100$ trials in each channel) are sufficient for the two learning devices (\textcolor{red}{\UCB} and \textcolor{green}{Thompson Sampling}) to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the \textcolor{cyan}{non-learning (uniform)} device, which stays around $40\%$ of success.
 Similar gains of performance were obtained in other scenarios.}

### 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for 2day

@@
 %
-We can verify that our approximation is very precise for lower values $p_{c1} \leq 30 \%$ (\ie, red and orange curves are quite close).
+We can verify that our approximation is very precise for lower values $p_{c1} \leq 30 \%$ (\ie, \textcolor{red}{red} and \textcolor{gold}{orange} curves are quite close).
 Moreover, a significant gap between $p_{c1}$ and $p_c$,
@@
 	\includegraphics[width=0.60\linewidth]{Approximation_m10-eps-converted-to.pdf}
-	\caption[Proposed approximation for the probability of collision at the second transmission.]{Proposed approximation for the probability of collision at the second transmission.
-It is more precise for smaller values of $N$.}
+	\caption[Our approximation for the probability of collision at the second transmission.]{\textcolor{red}{Our approximation (in red)} of \textcolor{gold}{$p_{c1}$ (in orange)}, probability of collision at the second transmission, is more precise for smaller $N$.
+We also include \textcolor{blue}{$p_{c}$ (in blue)}.}
 	\label{fig:43:Approximation_m10}

### 2019-07-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for 2day

@@
   \caption[Failure case of \Selfish.]{Regret for $M=2$ players, $K=3$ arms, horizon $T=5000$, $1000$ repetitions and $\boldsymbol{\mu} = [0.1, 0.5, 0.9]$.
-Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green} curve for \Selfish{} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$).
+Axis $x$ is for regret (different scale for each part), and the \textcolor{darkgreen}{green curve for \Selfish} shows a small probability of having a linear regret ($17$ cases of $R_T \geq T$, out of $1000$).
 The regret for the three other algorithms is very small for this problem, always smaller than $100$ here.}
@@
   \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem.]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-\RandTopM{} (\textcolor{gold}{yellow} curve) outperforms \Selfish{} (\textcolor{darkgreen}{green}), both clearly outperform \rhoRand.
+\textcolor{gold}{\RandTopM{} (in yellow)} outperforms \textcolor{darkgreen}{\Selfish{} (in green)}, both clearly outperform \rhoRand.
 The regret of \MCTopM{} is logarithmic, empirically with the same slope as the lower bound.
@@
   \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem.]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-In different settings, \RandTopM{} (\textcolor{gold}{yellow} curve) and \Selfish{} (\textcolor{darkgreen}{green}) can outperform each other, and always outperform \rhoRand.
+In different settings, \textcolor{gold}{\RandTopM{} (in yellow)} and \textcolor{darkgreen}{\Selfish{} (in green)} can outperform each other, and always outperform \rhoRand.
 \MCTopM{} is always among the best algorithms, and for $M$ not too small, its regret seems logarithmic with a constant matching the lower bound.}
@@
   % \begin{subfigure}[!h]{0.49\textwidth}
-    \includegraphics[width=1.10\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
+    \includegraphics[width=0.95\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
   % \end{subfigure}
   \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-\rhoRand{} (top \textcolor{blue}{blue} curve) is outperformed by the other algorithms (and the gain increases with $M$).
-\MCTopM{} (bottom \textcolor{gold}{yellow}) outperforms all the other algorithms in most cases.}
+\textcolor{blue}{\rhoRand{} (top blue)} is outperformed by the other algorithms (and the gain increases with $M$).
+\textcolor{gold}{\MCTopM{} (bottom yellow)} outperforms all the other algorithms in most cases.}
   % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
@@
   \caption[Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=2$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-\rhoRand{} (top \textcolor{blue}{blue}) is outperformed by the other algorithms (and the gain increases when $M$ increases), which all perform similarly in such configurations.
+\textcolor{blue}{\rhoRand{} (top blue)} is outperformed by the other algorithms (and the gain increases when $M$ increases), which all perform similarly in such configurations.
 Note that the (small) tail of the histograms come from complicated problems $\boldsymbol{\mu}$ and not failure cases.}
@@
   \caption[Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-This extreme case $M=K$ shows the drastic difference of behavior between \RandTopM{} and \MCTopM, having constant regret, and \rhoRand{} and \Selfish, having large regret.}
+This extreme case $M=K$ shows the drastic difference of behavior between \textcolor{red}{\RandTopM{} (red)} and \textcolor{gold}{\MCTopM{} (yellow)}, with constant regret, and \textcolor{blue}{\rhoRand{} (blue)} and \textcolor{darkgreen}{\Selfish{} (green)}, with large regret.}
   \label{fig:5:MP__K9_M9_T5000_N500__4_algos__all_HistogramsRegret}

### 2019-07-03 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Done for 2day

@@
 We remind that if  $\cA=(\cA_1,\dots,\cA_M)$ is a strongly uniformly efficient \emph{decentralized} policy,
-and if $\cA_j$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$, then we obtained the following asymptotical result:
-%
-\begin{equation}
-  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T^{\cA}(\boldsymbol{\mu}, M)}{\log T} \geq C_{\boldsymbol{\mu}, M}.
-\end{equation}
+and if $\cA_j$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$, then we obtained the following asymptotical result
+$\mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T^{\cA}(\boldsymbol{\mu}, M)}{\log T} \geq C_{\boldsymbol{\mu}, M}$.
 %
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
-  \caption[Comparison of our lower bound against the one from \cite{Zhao10}.]{Comparison of our lower bound against the one from \cite{Zhao10}, on a simple problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$, as a function of the number of players $M$.}
+  \includegraphics[width=0.90\textwidth]{Comparison_of_our_lowerbound_and_Anandkumar_2010__9_arms.pdf}
+  \caption[Comparison of our lower bound against the one from \cite{Zhao10}.]{Comparison of \textcolor{red}{our lower bound (in red)} against \textcolor{blue}{the one from \cite{Zhao10} (in blue)}, on a small problem with $9$ Bernoulli arms, of means $\boldsymbol{\mu} = [0.1, \dots, 0.9]$, as a function of the number of players $M$.}
   \label{fig:5:CompLowerBounds}
@@
 on the same example problem $\boldsymbol{\mu}$, with $K = 9$ arms and respectively $M = 6$, or $9$ players, for \Selfish-\klUCB.
+We considered $1000$ independent runs and plot the average regret.
 %
@@
     % \end{subfigure}
-    \caption[Illustrating the decomposition of the multi-players regret with its three terms.]{The three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3} of the multi-players regret, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
+    \caption[Illustrating the decomposition of the multi-players regret with its three terms.]{The three terms \ref{eq:5:term1} (\textcolor{red}{in red}), \ref{eq:5:term2} (\textcolor{cyan}{in cyan}), \ref{eq:5:term3} (\textcolor{green}{in green}) of the multi-players regret, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} (\textbf{in black}), for \Selfish-\klUCB.
+Two cases for $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$.}
     \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}

### 2019-07-03 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for 2day

@@
 $4)$ \GLRklUCB{} with \emph{global restart}.
-In this ``easy'' problem, all algorithms correctly detect the important changes, except \CUSUM{} which detected a change on arm $1$ (in \textcolor{red}{red}) twice after the change-point located at $t=3000$.
-The three other algorithms have a very small delay, for instance only $9$ samples from arm $0$ (in \textcolor{blue}{blue}) after its change at time $t=2000$ are enough for the \GLR{} test to detect a change.
+In this ``easy'' problem, all algorithms correctly detect the important changes, except \CUSUM{} which detected a change on arm $1$ (\textcolor{red}{in red}) twice after the change-point located at $t=3000$.
+The three other algorithms have a very small delay, for instance only $9$ samples from arm $0$ (\textcolor{blue}{in blue}) after its change at time $t=2000$ are enough for the \GLR{} test to detect a change.
 Very small delays are possible only after having collected a lot of samples of the arm which changed, and for instance the same algorithm obtains a delay of $32$ samples from \textcolor{red}{arm $1$} for the next change at $t=3000$, because this arm was sampled less.
@@
 The two \GLRklUCB{} variants correctly detect the important changes, and we observe that their delays to detect a change can vary, mainly due to the randomness (we remind that we illustrate only one repetition of the simulation here).
-For instance, after the change on arm $1$ (in \textcolor{red}{red}), the \emph{local restart} detects it at time $t=2376$ and the \emph{global restart} detects it at time $t=2701$.
+For instance, after the change on arm $1$ (\textcolor{red}{in red}), the \emph{local restart} detects it at time $t=2376$ and the \emph{global restart} detects it at time $t=2701$.
 
@@
 %
-Like problem $3$, this harder problem is also inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!}, but from another competitor paper, see Figure~3 from \cite{LiuLeeShroff17}.
+Like problem $3$, this harder problem is inspired from synthetic data obtained from a real-world database of clicks from \emph{Yahoo!}, but from another competitor paper, see Figure~3 from \cite{LiuLeeShroff17}.
 %

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
Shorter appendix in Ch6 and Ch7

@@
 
-With this numerical framework, simulations can run on a single CPU or a multi-core machine using joblib \cite{joblib} (see Appendix~\ref{sub:3:parallelSimulations}),
+With this numerical framework, simulations can run on a single CPU or a multi-core machine using joblib \cite{joblib},
+% (see Appendix~\ref{sub:3:parallelSimulations}),
 and summary plots are automatically saved as high-quality PNG, PDF and EPS, using matplotlib \cite{matplotlib} and seaborn \cite{seaborn}.
@@
 
-\paragraph{Bash command to run these experiments.}
+\paragraph{How to run such experiments.}
 %
@@
 We know that efficient algorithms are expected to follow these patterns:
-\begin{itemize}\tightlist
+\begin{itemize}%\tightlist
     \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
@@
 
-% --------------------------------------------------------------
-\subsection{Using multi-cores parallelism to speed-up simulations in SMPyBandits}
-\label{sub:3:parallelSimulations}
-% Using multi-core to speed-up simulations
-
-% \TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? Ce n'est pas crucial.
-%     % C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
-% }
-
-Nowadays, parallelism is everywhere in the computational world, and any serious framework for numerical simulations must try to gain performance from parallelism.
-In this last section, we quickly review the chosen approach for SMPyBandits (multi-core on one machine).
-
-For all the numerical simulations presented in this thesis,
-% except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4},
-the setting is the same:
-we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=1000$ like in Chapter~\ref{chapter:6}).
-%
-Because of the fundamentally sequential nature of bandit games, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
-But parallelism can help greatly for the two other axes: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
-Similarly, if we want to run $1000$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $1000/4=25$ repetitions on each process also bring a speed-up factor close to $4$.
-%
-%  and we explain why the two other approaches were less appropriate for our study of multi-armed bandit problems.
-
-\textbf{\texttt{Joblib} for multi-core simulations.}
-%
-The first approach is to use multiple cores of the same machines, and because it is both the simplest and the less financially as well as ecologically costly, this is the approach implemented in SMPyBandits.
-The machines I had access to during my thesis, either my own laptop or a workstation hosted the SCEE team in CentraleSupélec campus, were equipped with $i5$ or $i7$ Intel CPU with $4$ or $12$ cores.
-%
-As explained in the page \href{https://smpybandits.github.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}, we implemented in SMPyBandits an easy way to run any simulations on $n$ cores of a machine, using the \texttt{Joblib} \cite{joblib} library.
-
-It is implemented in a completely transparent way, and if someone uses the command-line variable to configure experiments, switching from using one core to using all the cores of the machine is as simple as changing \texttt{N\_JOBS=1} to \texttt{N\_JOBS=-1}, like in the Code Example~\ref{lst:3:runOneCoreOrMore} below.
-Similarly, changing from within the Python code is simple, by changing the \texttt{configuration["n\_jobs"]} value.
-
-% https://tex.stackexchange.com/a/12430/
-\begin{small}
-    \begin{listing}[h!]
-        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
-$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=1 \
-  python3 main.py configuration.py  # on only one core
-
-$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=-1 \
-  python3 main.py configuration.py  # parallel on all the cores
-
-$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=100 \
-  python3 main.py configuration.py  # parallel with too many jobs
-        \end{minted}
-        \caption{Running a simulation using either one CPU core with \texttt{N\_JOBS=1}, or all of them with \texttt{N\_JOBS=-1}, or more jobs than CPU cores with \texttt{N\_JOBS=20} (if there are $4$ cores).
+% % --------------------------------------------------------------
+% \subsection{Using multi-cores parallelism to speed-up simulations in SMPyBandits}
+% \label{sub:3:parallelSimulations}
+% % Using multi-core to speed-up simulations
+
+% % \TODOL{Si on a la place, j'aimerai bien inclure ces pages suivantes ? Ce n'est pas crucial.
+% %     % C'est en ligne sur \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
+% % }
+
+% Nowadays, parallelism is everywhere in the computational world, and any serious framework for numerical simulations must try to gain performance from parallelism.
+% In this last section, we quickly review the chosen approach for SMPyBandits (multi-core on one machine).
+
+% For all the numerical simulations presented in this thesis,
+% % except the ones based on MATLAB or GNU Radio in Chapter~\ref{chapter:4},
+% the setting is the same:
+% we consider a small set of $p$ different problems, of time horizon $T$ that we want to simulate for $N$ independent runs (\eg, $p=6$, $T=10000$ and $N=1000$ like in Chapter~\ref{chapter:6}).
+% %
+% Because of the fundamentally sequential nature of bandit games, each repetition of the simulation must be sequential regarding the time steps $t=1,\dots,T$, and so no parallelism can be done to speed up this axis.
+% But parallelism can help greatly for the two other axes: if we have a way to run in parallel $4$ processes, and we have $p=4$ problems to simulate, then running a process for each problem directly brings a speed-up factor of $4$.
+% Similarly, if we want to run $1000$ repetitions of the same (random) problem, and we can run $4$ processes in parallel, then running $1000/4=25$ repetitions on each process also bring a speed-up factor close to $4$.
+% %
+% %  and we explain why the two other approaches were less appropriate for our study of multi-armed bandit problems.
+
+% \textbf{\texttt{Joblib} for multi-core simulations.}
+% %
+% The first approach is to use multiple cores of the same machines, and because it is both the simplest and the less financially as well as ecologically costly, this is the approach implemented in SMPyBandits.
+% The machines I had access to during my thesis, either my own laptop or a workstation hosted the SCEE team in CentraleSupélec campus, were equipped with $i5$ or $i7$ Intel CPU with $4$ or $12$ cores.
+% %
+% As explained in the page \href{https://smpybandits.github.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}, we implemented in SMPyBandits an easy way to run any simulations on $n$ cores of a machine, using the \texttt{Joblib} \cite{joblib} library.
+
+% It is implemented in a completely transparent way, and if someone uses the command-line variable to configure experiments, switching from using one core to using all the cores of the machine is as simple as changing \texttt{N\_JOBS=1} to \texttt{N\_JOBS=-1}, like in the Code Example~\ref{lst:3:runOneCoreOrMore} below.
+% Similarly, changing from within the Python code is simple, by changing the \texttt{configuration["n\_jobs"]} value.
+
+% % https://tex.stackexchange.com/a/12430/
+% \begin{small}
+%     \begin{listing}[h!]
+%         \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
+% $ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=1 \
+%   python3 main.py configuration.py  # on only one core
+
+% $ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=-1 \
+%   python3 main.py configuration.py  # parallel on all the cores
+
+% $ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=100 \
+%   python3 main.py configuration.py  # parallel with too many jobs
+%         \end{minted}
+%         \caption{Running a simulation using either one CPU core with \texttt{N\_JOBS=1}, or all of them with \texttt{N\_JOBS=-1}, or more jobs than CPU cores with \texttt{N\_JOBS=20} (if there are $4$ cores).
 The later option is not advised, and \texttt{N\_JOBS=-1} gives the optimal speed-up factor.}
-        % \captionof{listing}{Run a simulation using 1 or \texttt{N} cores \label{lst:3:runOneCoreOrMore}.}
-        \label{lst:3:runOneCoreOrMore}
-    \end{listing}
-\end{small}
-
-\textbf{Effect of the number of jobs.}
-%
-As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
-%  dividing the running time by \texttt{N\_JOBS}.
-But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is sub-optimal, as the CPU will essentially spend all its time (and memory) managing the different jobs, and not actually doing the simulations.
-Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits.
+%         % \captionof{listing}{Run a simulation using 1 or \texttt{N} cores \label{lst:3:runOneCoreOrMore}.}
+%         \label{lst:3:runOneCoreOrMore}
+%     \end{listing}
+% \end{small}
+
+% \textbf{Effect of the number of jobs.}
+% %
+% As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
+% %  dividing the running time by \texttt{N\_JOBS}.
+% But jobs are implemented as threads, so the speed-up cannot be more than the number of cores, and using for instance $20$ jobs on $4$-cores for the $20$ repetitions is sub-optimal, as the CPU will essentially spend all its time (and memory) managing the different jobs, and not actually doing the simulations.
+% Using the above example, we illustrate the effect of using multi-jobs and multi-cores on the time efficiency of simulations using SMPyBandits.
 We consider three values of \texttt{N\_JOBS}, $1$ to use only one core and one job, $4$ to use all the $4$ cores of my $i5$ Intel CPU, and $100$ jobs.
-%
-We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=10000$, and different number of repetitions and number of jobs.
-These three experiments correspond to the three lines of Code Example~\ref{lst:3:runOneCoreOrMore} shown above.
-%
-The results of Table~\ref{table:3:speedUpTimeParallelComputations} clearly illustrate that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
-Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
-%
-Similar conclusions can be drawn from Figure~\ref{fig:3:analyze_speedup_as_nb_of_jobs}, which reports the results of the same experiment on a machine with $n_{\text{cores}}=12$ CPU cores.
-More details are included in the documentation page
-\href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
-
-% for N in 1 10 100; do for NJOBS in 1 4 100; do DEBUG=True SAVEALL=False NOPLOTS=True N=$N T=10000 N_JOBS=$NJOBS make single; read; done
-
-\begin{table}[ht]
-% \begin{footnotesize}
-    \centering
-    \begin{tabular}{c|rrr}
-    \textbf{Repetitions} & \texttt{N\_JOBS}$=1$ & \textcolor{darkgreen}{\texttt{N\_JOBS}$=4$ ($= n_{\text{cores}}$)} & \texttt{N\_JOBS}=$20$ ($> n_{\text{cores}}$) \\
-        \hline
-        $N=1$ repetition    & \SI{15}{\second}
-        & \SI{26}{\second} ($0.57 \%$)  % 15/26
-        & \SI{43}{\second} ($0.34 \%$)  % 15/43
-        \\
-        $N=10$ repetitions  & \SI{87}{\second}
-        & \SI{51}{\second} ($1.70 \%$)  % 87/51
-        & \SI{76}{\second} ($1.14 \%$)  % 87/76
-        \\
-        $N=100$ repetitions & \SI{749}{\second}
-        & \SI{272}{\second} (\textcolor{darkgreen}{$2.75 \%$})  % 749/272
-        & \SI{308}{\second} ($2.43 \%$)  % 749/308
-        \\
-        $N=500$ repetitions & \SI{2944}{\second}
-        & \SI{1530}{\second} (\textcolor{darkgreen}{$1.92 \%$})  % 2944/1530
-        & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
-        \\
-        % \hline
-    \end{tabular}
-    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, on a machine with $n_{\text{cores}}=4$ CPU cores, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.
-The percentages are the relative gains in comparison with the non-parallel simulation.}
-    \label{table:3:speedUpTimeParallelComputations}
-% \end{footnotesize}
-\end{table}
-
-\begin{figure}[h!]  % [htbp]
-	\centering
-	% \includegraphics[width=0.85\linewidth]{analyze_speedup_as_nb_of_jobs.pdf}
-	\includegraphics[width=1.10\linewidth]{analyze_speedup_as_nb_of_jobs_normalize.pdf}
-	\caption{Relative gain on the running time in comparison to using one job, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine running the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}.}
-	\label{fig:3:analyze_speedup_as_nb_of_jobs}
-\end{figure}
-
-% for N in 1 10 100 1000; do for NJ in 1 12 100; do echo "With N=$N repetitions and N_JOBS=$NJ jobs..."; time NOPLOTS=True SAVEALL=False BAYES=False ARM_TYPE=Bernoulli N=$N T=1000 K=9 N_JOBS=$NJ python3 main.py configuration.py 2>/dev/null | grep "real "; done; done && FreeSMS.py "DONE"
-
+% %
+% We give in Table~\ref{table:3:speedUpTimeParallelComputations} an example of running time of an experiment with $T=10000$, and different number of repetitions and number of jobs.
+% These three experiments correspond to the three lines of Code Example~\ref{lst:3:runOneCoreOrMore} shown above.
+% %
+% The results of Table~\ref{table:3:speedUpTimeParallelComputations} clearly illustrate that using more jobs than the number of CPU is sub-optimal, and that as soon as the number of repetitions is large enough, using one job by available CPU core (\ie, here $4$ jobs) gives a significant speed-up time.
+% Due to the cost of orchestrating the different jobs, and memory exchanges at the end of each repetition, the parallel version is \emph{not} $4$ times faster, but empirically we always found it to be $2$ to $3.5$ times faster.
+% %
+% Similar conclusions can be drawn from Figure~\ref{fig:3:analyze_speedup_as_nb_of_jobs}, which reports the results of the same experiment on a machine with $n_{\text{cores}}=12$ CPU cores.
+% More details are included in the documentation page
+% \href{https://smpybandits.github.io/About_parallel_computations.html}{\texttt{SMPyBandits.github.io/About\_parallel\_computations.html}}
+
+% % for N in 1 10 100; do for NJOBS in 1 4 100; do DEBUG=True SAVEALL=False NOPLOTS=True N=$N T=10000 N_JOBS=$NJOBS make single; read; done
+
+% \begin{table}[ht]
+% % \begin{footnotesize}
+%     \centering
+%     \begin{tabular}{c|rrr}
+%     \textbf{Repetitions} & \texttt{N\_JOBS}$=1$ & \textcolor{darkgreen}{\texttt{N\_JOBS}$=4$ ($= n_{\text{cores}}$)} & \texttt{N\_JOBS}=$20$ ($> n_{\text{cores}}$) \\
+%         \hline
 %         $N=1$ repetition    & \SI{15}{\second}
@@
 %         $N=100$ repetitions & \SI{749}{\second}
-% & \SI{272}{\second} ($2.75 \%$)  % 749/272
+%         & \SI{272}{\second} (\textcolor{darkgreen}{$2.75 \%$})  % 749/272
 %         & \SI{308}{\second} ($2.43 \%$)  % 749/308
@@
 %         $N=500$ repetitions & \SI{2944}{\second}
-% & \SI{1530}{\second} ($1.92 \%$)  % 2944/1530
+%         & \SI{1530}{\second} (\textcolor{darkgreen}{$1.92 \%$})  % 2944/1530
 %         & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
-
-% \TODOL{Pas sûr que ça vaille la peine d'inclure ce passage, c'est vraiment optionnel !}
-
-% % Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).
-% \paragraph{Approaches we did not try.}
-% %
-% The two other approaches we could have consider is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
-
-% $(i)$ I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
-% Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.
-% Since $2016$, we saw a large press coverage as well as an extensive use in research of deep learning libraries that make general-purpose machine learning algorithms train on the GPU of a user's laptop or a cluster of GPU.
-% This success is mainly possible because of the heavy parallelism of such training algorithms, and the parallel nature of GPU.
-% To the best of the author knowledge, nobody has tried to implement high performance MAB simulations by using the ``parallelism power'' of a GPU (at least, no code for such experiments were made public in $2019$).
-% I worked on a GPU, implementing fluid dynamic simulations in an internship in $2012$, and I have since then kept a curiosity on how to use GPU-powered libraries and code.
-% I have contributed to and used famous deep-learning libraries, like Theano (\href{http://deeplearning.net/software/theano/}{\texttt{DeepLearning.net/software/theano}}) or Keras (\href{https://keras.io/}{\texttt{keras.io}}), and my limited knowledge on such libraries made me believe that it was not easy to use a GPU for bandit simulations, and most surely it would not have been worth the time.
-% I would be very curious to understand how a GPU could be used to implement highly efficient simulations for sequential learning problems, because it seemed hard whenever I thought about it.
-
-% $(ii)$ I also did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
-% It is partly due to time constraint, as I would have been curious to try, but mainly because we found that it would not have helped us much to use a large scale cluster.
-% The main reason is that in the multi-armed bandit and sequential learning literature, most research papers do not even include an experimental section, and for the papers who did take the time to implement and test their proposed algorithms, it is almost done on just a few problems and for short- or medium- duration experiments.
-% For instance, the papers we consider to be the best ones regarding their empirical sections are \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, for piece-wise stationary bandits, and they mainly consider reasonable problems of horizon $T=10000$ and no more than $1000$ independent repetitions.
+%         \\
+%         % \hline
+%     \end{tabular}
+%     \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, on a machine with $n_{\text{cores}}=4$ CPU cores, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.
+The percentages are the relative gains in comparison with the non-parallel simulation.}
+%     \label{table:3:speedUpTimeParallelComputations}
+% % \end{footnotesize}
+% \end{table}
+
+% \begin{figure}[h!]  % [htbp]
+% 	\centering
+% 	% \includegraphics[width=0.85\linewidth]{analyze_speedup_as_nb_of_jobs.pdf}
+% 	\includegraphics[width=1.10\linewidth]{analyze_speedup_as_nb_of_jobs_normalize.pdf}
+% 	\caption{Relative gain on the running time in comparison to using one job, as a function of the number of jobs \texttt{N\_JOBS}, for a $n_{\text{cores}}=12$-core machine running the simulation of Code Example~\ref{lst:3:runOneCoreOrMore}.}
+% 	\label{fig:3:analyze_speedup_as_nb_of_jobs}
+% \end{figure}
+
+% % for N in 1 10 100 1000; do for NJ in 1 12 100; do echo "With N=$N repetitions and N_JOBS=$NJ jobs..."; time NOPLOTS=True SAVEALL=False BAYES=False ARM_TYPE=Bernoulli N=$N T=1000 K=9 N_JOBS=$NJ python3 main.py configuration.py 2>/dev/null | grep "real "; done; done && FreeSMS.py "DONE"
+
+% % $N=1$ repetition    & \SI{15}{\second}
+% % & \SI{26}{\second} ($0.57 \%$)  % 15/26
+% % & \SI{43}{\second} ($0.34 \%$)  % 15/43
+% % \\
+% % $N=10$ repetitions  & \SI{87}{\second}
+% % & \SI{51}{\second} ($1.70 \%$)  % 87/51
+% % & \SI{76}{\second} ($1.14 \%$)  % 87/76
+% % \\
+% % $N=100$ repetitions & \SI{749}{\second}
+% % & \SI{272}{\second} ($2.75 \%$)  % 749/272
+% % & \SI{308}{\second} ($2.43 \%$)  % 749/308
+% % \\
+% % $N=500$ repetitions & \SI{2944}{\second}
+% % & \SI{1530}{\second} ($1.92 \%$)  % 2944/1530
+% % & \SI{1846}{\second} ($1.59 \%$)  % 2944/1846
+
+% % \TODOL{Pas sûr que ça vaille la peine d'inclure ce passage, c'est vraiment optionnel !}
+
+% % % Explain why I didn't try more on parallelisation (eg on Grid5000 or other large scale computer).
+% % \paragraph{Approaches we did not try.}
+% % %
+% % The two other approaches we could have consider is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
+
+% % $(i)$ I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
+% % Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.
+% % Since $2016$, we saw a large press coverage as well as an extensive use in research of deep learning libraries that make general-purpose machine learning algorithms train on the GPU of a user's laptop or a cluster of GPU.
+% % This success is mainly possible because of the heavy parallelism of such training algorithms, and the parallel nature of GPU.
+% % To the best of the author knowledge, nobody has tried to implement high performance MAB simulations by using the ``parallelism power'' of a GPU (at least, no code for such experiments were made public in $2019$).
+% % I worked on a GPU, implementing fluid dynamic simulations in an internship in $2012$, and I have since then kept a curiosity on how to use GPU-powered libraries and code.
+% % I have contributed to and used famous deep-learning libraries, like Theano (\href{http://deeplearning.net/software/theano/}{\texttt{DeepLearning.net/software/theano}}) or Keras (\href{https://keras.io/}{\texttt{keras.io}}), and my limited knowledge on such libraries made me believe that it was not easy to use a GPU for bandit simulations, and most surely it would not have been worth the time.
+% % I would be very curious to understand how a GPU could be used to implement highly efficient simulations for sequential learning problems, because it seemed hard whenever I thought about it.
+
+% % $(ii)$ I also did not try to use any large scale computer cluster, even if I was aware of the possibility offered by the Grid 5000 project, for instance.
+% % It is partly due to time constraint, as I would have been curious to try, but mainly because we found that it would not have helped us much to use a large scale cluster.
+% % The main reason is that in the multi-armed bandit and sequential learning literature, most research papers do not even include an experimental section, and for the papers who did take the time to implement and test their proposed algorithms, it is almost done on just a few problems and for short- or medium- duration experiments.
+% % For instance, the papers we consider to be the best ones regarding their empirical sections are \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, for piece-wise stationary bandits, and they mainly consider reasonable problems of horizon $T=10000$ and no more than $1000$ independent repetitions.
 Each paper considers one harder problem, of horizon $T=1000000$ and less repetitions.
-% %
-% In each article written during my thesis, we included extensive numerical simulations, and even the longest ones (for \cite{Besson2019GLRT}) were short enough to run in less than $12$ hours on a $12$-core workstation, so we could run a few large-scale simulations over night.
-% For such reasons, we prefer to not try to run simulations on a cluster.
+% % %
+% % In each article written during my thesis, we included extensive numerical simulations, and even the longest ones (for \cite{Besson2019GLRT}) were short enough to run in less than $12$ hours on a $12$-core workstation, so we could run a few large-scale simulations over night.
+% % For such reasons, we prefer to not try to run simulations on a cluster.

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Shorter appendix in Ch6 and Ch7

@@
 
-This section gives the finite-time regret bounds for the two variants, and interpretations of the different results.
-Although we only prove one bound, the other proof can be found in \cite{Besson2019GLRT}.
+This section gives the finite-time regret bounds for the two variants, and interpretations of our results.
+We include only one proof, as the other is given in \cite{Besson2019GLRT}.
 
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    For $\omega$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based on \textbf{Global} Restart satisfies
+    For $\omega$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriodsGlobal} is satisfied, the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based on \textbf{Global} Restart satisfies the following \emph{finite-time} regret bound
     \begin{align}
         R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T \\
-        & + \sum_{k=1}^{\Upsilon_T}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}}.
+        & + \sum_{k=1}^{\Upsilon_T}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \textcolor{red}{\bigO{\sqrt{\ln(T)}}}.
 \nonumber
@@
 %
-We highlight that \emph{this result is finite-time} and not asymptotic, even if it uses the notation $\bigO{\sqrt{\ln(T)}}$.
-This last term does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a constant times $\sqrt{\ln(T)}$ from a certain time $T_0$.
+We highlight that \emph{this result is finite-time} and not asymptotic, but if it uses the notation $\textcolor{red}{\bigO{\sqrt{\ln(T)}}}$ for simplicity.
+This last term $\bigO{\dots}$ does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a certain $c_0 \times \sqrt{\ln(T)}$, from a certain time $T_0$ (\ie, $\forall T \geq T_0$).
+It is also the case of the next Theorem~\ref{thm:6:mainRegretBound}.
+
+% \vspace*{-20pt}  % FIXME
 
@@
 \begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
-    For $\omega$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based on Local Restart satisfies
+    For $\omega$ and $\delta$ for which Assumption~\ref{ass:6:LongPeriods} is satisfied, the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based on Local Restart satisfies the following \emph{finite-time} regret bound
     %
     \begin{equation}
-        R_T \leq 2\sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{4K}{\omega \left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) +\omega T + 2 \delta \EffChange  + \sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{\ln(T)}{\kl(\overline{\mu}_k^{(\ell)},{\mu}^*_{i,\ell})} +\bigO{\sqrt{\ln(T)}},
+        R_T \leq 2\sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{4K}{\omega \left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) +\omega T + 2 \delta \EffChange  + \sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{\ln(T)}{\kl(\overline{\mu}_k^{(\ell)},{\mu}^*_{i,\ell})} + \textcolor{red}{\bigO{\sqrt{\ln(T)}}},
     \end{equation}
@@
 Like for the first Theorem~\ref{thm:6:mainRegretBoundGlobal},
-we highlight that \emph{this result is finite-time}.
+we highlight that \textcolor{red}{\emph{this result is finite-time}}.
 The proof of \GLRklUCB{} with \emph{local restarts} is not included, but it can be found in the Appendix of \cite{Besson2019GLRT}, as it is quite similar to the proof for \emph{global restarts} given above.
@@
     \end{enumerate}
+    \vspace*{-20pt}  % FIXME
 % \end{leftbar}  % XXX leftbar corollarybar, comment if needed
@@
 the constant is set to $\omega_0 = 0.05$.
-We show in Appendix~\ref{sec:6:choosingAlpha0} a certain robustness, with similar regret as soon as $\omega\leq0.1$.
-We do not use a constant $\delta_0$ to change the confidence level to $\delta = \delta_0 \delta_T$, as we show in Appendix~\ref{sec:6:choosingDelta} that $\delta_0=1$ is uniformly better for different problems.
+% We show in Appendix~\ref{sec:6:choosingAlpha0}
+We show in Appendix~G of \cite{Besson2019GLRT} a sensitivity analysis on $\omega$,
+which concludes to a robustness regarding this parameter, with similar regret as soon as $\omega\leq0.1$.
+We do not use a constant $\delta_0$ to change the confidence level to $\delta = \delta_0 \delta_T$, as we
+verified empirically
+% show in Appendix~\ref{sec:6:choosingDelta}
+that $\delta_0=1$ is uniformly better for different problems.
 %
@@
 and \CUSUM{} also uses the first trick with $\Delta n = 10$ (see Appendix~\ref{sub:6:IdeasOptimizations} for more details).
+
+\paragraph{Time and memory costs of \GLRklUCB.}
+%
+Finally, we mention that we discuss in details in Appendix~F of \cite{Besson2019GLRT} about the time and memory complexities of \GLRklUCB, and the other algorithms considered for the numerical experiments.
+%
+To sum up, at every time step the CPD algorithm needs a time $\bigO{n_i}=\bigO{d_{\max}}$, and at the end, the time complexity of \CUSUMklUCB{} as well as \GLRklUCB{} is $\bigO{K T d_{\max}}$, which can be up-to $\bigO{K T^2}$, much more costly than $\bigO{K T}$ for \klUCB{} for instance.
+%
+Our proposal \GLRklUCB{} requires a storage of the order of $\bigO{K d_{\max}}$ and a running time of the order of $\bigO{K T d_{\max}}$.
+We validated the two bounds experimentally, with results presented in Tables~2 and 3 in \cite{Besson2019GLRT}.
 

### 2019-07-04 2-Chapters/1-Chapter/chapter1.tex
First changes from Emilie's LONG comments

@@
 %
-This manuscript ends with the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks}.
-We discuss about our work on doubling-trick, in order to remove the need for an algorithm to know the horizon of the bandit game before starting to play it.
-The rest of the appendix contains a list of abbreviations and notations, then lists of figures, algorithms, code samples and tables,
+This manuscript ends with
+% the appendixes, first with a short overview of another contribution, in Appendix~\ref{app:2:DoublingTricks}.
+% We discuss about our work on doubling-trick, in order to remove the need for an algorithm to know the horizon of the bandit game before starting to play it.
+% The rest of the appendix
+a list of abbreviations and notations, then lists of figures, algorithms, code samples and tables,
 and finally the list of the bibliographical references.
@@
     September $2018$.\\
-    See Appendix~\ref{app:2:DoublingTricks}.
+    % See Appendix~\ref{app:2:DoublingTricks}.
     Preprint at \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.

### 2019-07-04 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Lots of tiny typos on appendix
Done for 2day
First changes from Emilie's LONG comments

@@
 			}
-			\caption{Our algorithm \Aggr{} aggregating $\Alg_1,\dots,\Alg_N$.}
+			\caption{The \Aggr{} algorithm, aggregating $N$ MAB algorithms $\Alg_1,\dots,\Alg_N$.}
 			\label{algo:25:Aggr}
@@
 However, for real-world communication problems, it can be considered unrealistic to assume a fixed and known time horizon.
-For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}, that we shortly present in Appendix~\ref{app:2:DoublingTricks}.
+For more discussion on this hypothesis, we refer to Section~1 of our article \cite{Besson2018DoublingTricks}.
+% that we shortly present in Appendix~\ref{app:2:DoublingTricks}.
 %
@@
 	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
-	\caption{On a ``simple'' Bernoulli problem (semilog-$y$ scale).}
+	\caption{On a ``simple'' Bernoulli problem (semilog-$y$ scale). \Aggr{} is in \textbf{bold \textcolor{red}{red}}.}
 	\label{fig:25:EasyBernoulli}
@@
 Especially in difficult problems with mixed distributions, \Aggr{} showed to be more efficient that \ExpQ{} and orders of magnitude better than the other reference aggregation algorithms \LearnExp{} and \CORRAL{} (see Figures~\ref{fig:25:HarderMixed} and~\ref{fig:25:HarderMixed_semilogx}).
+
+\paragraph{Reproducibility.}
+%
+The experiments in this section are based on our library SMPyBandits,
+and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.
 
@@
 
-% \paragraph{Reproducility of the experiments.}
+% \paragraph{Reproducibility.}
 % %
-% Experiments presented in this Section are based on our library SMPyBandits,
+% The experiments in this section are based on our library SMPyBandits,
 % and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
First changes from Emilie's LONG comments

@@
 The documentation also contains extensive examples of all intermediate numerical functions, that are also used as tests (using \texttt{doctest.testmod} from the standard library).
-For example, Figures~\ref{fig:3:twoScreenshotsOfTheDocumentation} and \ref{fig:3:twoMoreScreenshotsOfTheDocumentation} below show four screenshots from the documentation.
-We show the main page, the list of algorithms in the \texttt{Policies} module, an example of the documentation of one algorithm (for the \klUCB-Switch algorithm from \cite{GarivierHadiji2018}), and a page detailing the API and organization of the library as well as instructions to follow if someone wants to implement a new model or a new algorithm.
-
-% FIXME?
-\vfill{}
-% FIXME?
+% and \ref{fig:3:twoMoreScreenshotsOfTheDocumentation}
+For example, Figure~\ref{fig:3:twoScreenshotsOfTheDocumentation} below show two screenshots from the documentation.
+We show the main page, and the list of algorithms in the \texttt{Policies} module.
+% an example of the documentation of one algorithm (for the \klUCB-Switch algorithm from \cite{GarivierHadiji2018}), and a page detailing the API and organization of the library as well as instructions to follow if someone wants to implement a new model or a new algorithm.
 
@@
 
-\begin{figure}[h!]  % [htbp]
-    \centering
-	\includegraphics[width=0.495\linewidth]{overview_documentation_3.png}
-	\includegraphics[width=0.495\linewidth]{overview_documentation_4.png}
-	\caption{Screenshots from two other pages of the documentation: the \texttt{klUCBSwitch} policy, and details of the code organization, to know how to add a new policy.}
-	\label{fig:3:twoMoreScreenshotsOfTheDocumentation}
-\end{figure}
+% \begin{figure}[h!]  % [htbp]
+%     \centering
+% 	\includegraphics[width=0.495\linewidth]{overview_documentation_3.png}
+% 	\includegraphics[width=0.495\linewidth]{overview_documentation_4.png}
+% 	\caption{Screenshots from two other pages of the documentation: the \texttt{klUCBSwitch} policy, and details of the code organization, to know how to add a new policy.}
+% 	\label{fig:3:twoMoreScreenshotsOfTheDocumentation}
+% \end{figure}
 
@@
     \item
-In \cite{Besson2018DoublingTricks} (see Appendix~\ref{app:2:DoublingTricks}), we used SMPyBandits to illustrate and compare different "doubling trick" schemes.
+In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes.
 \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}}
+% (see Appendix~\ref{app:2:DoublingTricks})
 % In sequential learning, an algorithm is said to be ``anytime'' if it does not need to know the horizon $T$ of the experiments.
@@
 Our position regarding this trade-off is motivated by Occam's razor principle.
-On the first hand, if the algorithm should be implemented on a cognitive radio device that has limited hardware capacity (for instance),
+On the one hand, if the algorithm should be implemented on a cognitive radio device that has limited hardware capacity (for instance),
 one can reasonably aim at the simplest yet order-optimal algorithms, and we advise to use \UCB{} and algorithms running on top of this simple index policy, as we do in Chapter~\ref{chapter:4}.
@@
 % We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
-We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} (see Appendix~\ref{app:2:DoublingTricks}).
+We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks}.
+%  (see Appendix~\ref{app:2:DoublingTricks}).
 

### 2019-07-04 2-Chapters/4-Chapter/chapter4.tex
First changes from Emilie's LONG comments

@@
 %
-On the first hand, we are actually able to analyze a simpler model, if we assume to have at most $M \leq K$ devices with a transmission probability of $p=1$.
+On the one hand, we are actually able to analyze a simpler model, if we assume to have at most $M \leq K$ devices with a transmission probability of $p=1$.
 Instead of the experiment-driven direction pursued in this chapter, another possibility is to consider a \emph{multi-players MAB} model to describe our problem.

### 2019-07-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Shorter appendix in Ch6 and Ch7
First changes from Emilie's LONG comments

@@
 However, a careful manipulation of this decomposition shows that the regret is always lower bounded by term \ref{eq:5:term1}.
-The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} in Appendix illustrate two cases of $M<K$ and $M=K$ and the different impact of the three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3} on the regret.
+% The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} in Appendix illustrate two cases of $M<K$ and $M=K$ and the different impact of the three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3} on the regret.
+The Figure~6 in Appendix~F of \cite{Besson2018ALT} illustrates two cases of $M<K$ and $M=K$, and the different impact of the three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3} on the regret.
 
@@
 \end{equation}
-as for every $k \in \Mworst$ and $j \in [M]$, $\kl(\mu_k, \mu_j^*) \geq \kl(\mu_k, \mu_M^*)$
-(see Figure~\ref{fig:5:CompLowerBounds} in Appendix~\ref{app:5:illustrationLowerBound}).
+as for every $k \in \Mworst$ and $j \in [M]$, $\kl(\mu_k, \mu_j^*) \geq \kl(\mu_k, \mu_M^*)$.
+% (see Figure~\ref{fig:5:CompLowerBounds} in Appendix~\ref{app:5:illustrationLowerBound}).
 %
@@
 % \end{framed}
+
+% ----------------------------------------------------------------------------
+% \subsection{Some future work on SMPyBandits}
+\paragraph{Reproducibility.}
+
+The experiments in this chapter use our library SMPyBandits,
+and the page \href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}} gives instructions to reproduce them.

### 2019-07-04 2-Chapters/5-Chapter/chapter5.tex
First changes from Emilie's LONG comments

@@
 For the same example, the ``doubling trick'' technique can be used to obtain a regret upper bound for an algorithm unaware of the value of $T$, within a constant multiplicative factor of the upper bound given for the algorithm aware of $T$ (Section~4.1).
-Because the regret bound is of the form $\bigO{\sqrt{T} \log(T)}$, a simple ``doubling trick'' of increasing horizons of lengths $T_i = 2^i$ works well, as proposed in \cite{CesaLugosi06} and as studied in depth in our article \cite{Besson2018DoublingTricks}, and quickly presented in Appendix~\ref{app:2:DoublingTricks}.
+Because the regret bound is of the form $\bigO{\sqrt{T} \log(T)}$, a simple ``doubling trick'' of increasing horizons of lengths $T_i = 2^i$ works well, as proposed in \cite{CesaLugosi06} and as studied in depth in our article \cite{Besson2018DoublingTricks}.
+% and quickly presented in Appendix~\ref{app:2:DoublingTricks}.
 %
@@
 
-% ----------------------------------------------------------------------------
-% \subsection{Some future work on SMPyBandits}
-\paragraph{Reproducility of the experiments.}
-
-The experiments in this chapter use our library SMPyBandits, as presented in Chapter~\ref{chapter:3},
-and we refer to the following page for details on how to reproduce them,
-\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{\texttt{SMPyBandits.GitHub.io/MultiPlayers.html}}
-
 % -----------------------------------------------------------------
@@
 
-\newpage
-% ----------------------------------------------------------------------------
-\section{Appendix}
-\label{sec:5:appendix}
+% \newpage
+% % ----------------------------------------------------------------------------
+% \section{Appendix}
+% \label{sec:5:appendix}
 
-% We include here some missing proofs and additional simulation results for this Chapter.
+% % We include here some missing proofs and additional simulation results for this Chapter.
 
-% ----------------------------------------------------------------------------
-\input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex}
+% % ----------------------------------------------------------------------------
+% \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex}

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
First changes from Emilie's LONG comments

@@
 
+% \paragraph{A note on the simulation code.}
+\paragraph{Reproducibility.}
+%
+The experiments in this chapter use our library SMPyBandits,
+and the page \href{https://SMPyBandits.GitHub.io/NonStationaryBandits.html}{\texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html}} gives instructions to reproduce them.
+
 \newpage
@@
 % Thanks to Odalric-Ambrym Maillard at Inria Lille for useful discussions, and thanks to Christophe Moy at University Rennes 1.
-
-% \paragraph{A note on the simulation code.}
-\paragraph{Reproducility of the experiments.}
-%
-The experiments in this chapter use our library SMPyBandits, as presented in Chapter~\ref{chapter:3},
-and we refer to the following page for details on how to reproduce them,
-\href{https://SMPyBandits.GitHub.io/NonStationaryBandits.html}{\texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html}}

### 2019-07-04 2-Chapters/2-Chapter/chapter25.tex
First changes from Emilie's LONG comments
Cleanup empty pages between chapters etc. Less than 300 pages YAY

@@
 
-% A last solution is online algorithm selection, inspired from expert aggregation.
-% Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292
+% % A last solution is online algorithm selection, inspired from expert aggregation.
+% % Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292
 
-\newpage  % WARNING ?
+% \newpage  % WARNING ?
 
-% ----------------------------------------------------------------------------
-\section{Conclusion}
-\label{sec:25:conclusion}
+% % ----------------------------------------------------------------------------
+% \section{Conclusion}
+% \label{sec:25:conclusion}
 
-In this chapter, we presented the first contribution of this manuscript,
-that corresponds to our article \cite{Besson2018WCNC}.
-%
-We presented the use of aggregation algorithms,
-especially for the real-world setting of unknown problem instances,
-when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
-% - \ExpQ and \Aggr works fine
-% Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
-Our proposed \Aggr{} was presented in details,
-and we also highlighted its differences with \ExpQ.
+% In this chapter, we presented the first contribution of this manuscript,
+% that corresponds to our article \cite{Besson2018WCNC}.
+% %
+% We presented the use of aggregation algorithms,
+% especially for the real-world setting of unknown problem instances,
+% when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
+% % - \ExpQ and \Aggr works fine
+% % Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
+% Our proposed \Aggr{} was presented in details,
+% and we also highlighted its differences with \ExpQ.
 
-% as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
-%
-% - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
-We presented experiments on simple MAB problems, coming from previous work on bandits for OSA \cite{Jouini10},
-and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
-Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
-% It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
-% and appears to be more robust than simple non-aggregated algorithms.
+% % as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
+% %
+% % - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
+% We presented experiments on simple MAB problems, coming from previous work on bandits for OSA \cite{Jouini10},
+% and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
+% Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
+% % It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
+% % and appears to be more robust than simple non-aggregated algorithms.
 
-\ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
-Empirically, \Aggr{} showed to always have a logarithmic
-regret if it aggregates at least one algorithm with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
-It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
-As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this is an over-optimistic goal.
+% \ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
+% Empirically, \Aggr{} showed to always have a logarithmic
+% regret if it aggregates at least one algorithm with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
+% It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
+% As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this is an over-optimistic goal.
 
-What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
-%
-%
-Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
-But this direction does not appear very useful for our objective at hand, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
-% We will continue investigating regret bounds for \Aggr,
-% and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).
+% What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
+% %
+% %
+% Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
+% But this direction does not appear very useful for our objective at hand, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
+% % We will continue investigating regret bounds for \Aggr,
+% % and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).
 
-% The numerical simulations used our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
-% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.
-
-\paragraph{Reproducility of the experiments.}
-%
-Experiments presented in this Section are based on our library SMPyBandits,
-and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.
+% % The numerical simulations used our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
+% % The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
Cleanup empty pages between chapters etc. Less than 300 pages YAY

@@
 
-We focused here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
+We focus on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
 To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
@@
 %
-An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
-For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
-For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
-or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
-%
-At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-players bandit model,
-and even if some have already been implemented, a major future work is to implement the most interesting ones
-(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
-
-Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models.
+% An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
+% For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
+% For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
+% or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+% %
+% At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-players bandit model,
+% and even if some have already been implemented, a major future work is to implement the most interesting ones
+% (tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
+
+% Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models.
 Most of them were implemented quickly, but some would require more time and are left as possible future works.
-Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits of \cite{Seznec2018} (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).
-
-Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
+% Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits of \cite{Seznec2018} (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).
+
+The most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
 Its documentation has seen already about $20000$ visits, the project on GitHub had $88$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used SMPyBandits.
@@
 We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to our work.
-
+%
 As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
@@
 
-We start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
-We conclude by a discussion on parallel multi-core CPU computations to speed-up SMPyBandits.
-
-\subsection{Minimalist example of use of SMPyBandits}
-\label{sub:3:miniExampleSMPyBandits}
-
-Below is presented in Code~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
-Only the mean reward is printed at the end of the simulation.
-%
-With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
-\begin{verbatim}
-The UCB algorithm obtains here a mean reward = 0.886
-\end{verbatim}
-% \vspace*{-10pt} % WARNING
-%
-% https://tex.stackexchange.com/a/12430/
-\begin{small}
+% We start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
+We give in this appendix more details regarding the numerical experiments presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above,
+for the other algorithms used in our benchmark,
+and details on the methodology used to measure the time and memory.
+% We conclude by a discussion on parallel multi-core CPU computations to speed-up SMPyBandits.
+
+% \subsection{Minimalist example of use of SMPyBandits}
+% \label{sub:3:miniExampleSMPyBandits}
+
+% Below is presented in Code~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
+% Only the mean reward is printed at the end of the simulation.
+% %
+% With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
+% \begin{verbatim}
+% The UCB algorithm obtains here a mean reward = 0.886
+% \end{verbatim}
+% % \vspace*{-10pt} % WARNING
+% %
+% % https://tex.stackexchange.com/a/12430/
+% \begin{small}
+% % \begin{listing}[h!]
+%     \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py}
+%     % \caption{Small example of \texttt{main.py} file}
+%     \captionof{listing}{Example of use of SMPyBandits\label{lst:3:exampleOfUse}.}
+%     % \label{lst:3:exampleOfUse}
+% % \end{listing}
+% \end{small}
+
+% \subsection{A minimalist \texttt{main.py} file}
+% \label{sub:3:miniMainFileSMPyBandits}
+
+% We present below in Code~\ref{lst:3:main} a full example of a minimalist \texttt{main.py} file,
+% which is used to load the configuration, run the experiment and then print and plot various visualizations for the simulation.
+
+% % https://tex.stackexchange.com/a/12430/
+% \begin{small}
+% % \begin{listing}[h!]
+%     \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
+%     % \caption{Small example of \texttt{main.py} file}
+%     \captionof{listing}{Example of \texttt{main.py} file\label{lst:3:main}.}
+%     % \label{lst:3:main}
+% % \end{listing}
+% \end{small}
+
+% \subsection{A minimalist \texttt{configuration.py} file}
+% \label{sub:3:miniConfigurationFileSMPyBandits}
+
+% We present below in Code~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
+% which is used to define the experiment.
+% It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
+% It uses environment variables, so a user can configure some parameters of the simulation when launching them from a terminal:
+
+% % \begin{verbatim}
+% %     (bash) $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
+% % \end{verbatim}
+
+% % https://tex.stackexchange.com/a/12430/
 % \begin{listing}[h!]
-    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py}
-    % \caption{Small example of \texttt{main.py} file}
-    \captionof{listing}{Example of use of SMPyBandits\label{lst:3:exampleOfUse}.}
-    % \label{lst:3:exampleOfUse}
+%     \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
+% $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
+%     \end{minted}
+%     \caption{Example of Bash code to run an experiment.}
+%     % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}
+%     \label{lst:3:howToRunExperiment2}
 % \end{listing}
-\end{small}
-
-\subsection{A minimalist \texttt{main.py} file}
-\label{sub:3:miniMainFileSMPyBandits}
-
-We present below in Code~\ref{lst:3:main} a full example of a minimalist \texttt{main.py} file,
-which is used to load the configuration, run the experiment and then print and plot various visualizations for the simulation.
-
-% https://tex.stackexchange.com/a/12430/
-\begin{small}
-% \begin{listing}[h!]
-    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
-    % \caption{Small example of \texttt{main.py} file}
-    \captionof{listing}{Example of \texttt{main.py} file\label{lst:3:main}.}
-    % \label{lst:3:main}
-% \end{listing}
-\end{small}
-
-\subsection{A minimalist \texttt{configuration.py} file}
-\label{sub:3:miniConfigurationFileSMPyBandits}
-
-We present below in Code~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
-which is used to define the experiment.
-It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
-It uses environment variables, so a user can configure some parameters of the simulation when launching them from a terminal:
-
-% \begin{verbatim}
-%     (bash) $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
-% \end{verbatim}
-
-% https://tex.stackexchange.com/a/12430/
-\begin{listing}[h!]
-    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
-$ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
-    \end{minted}
-    \caption{Example of Bash code to run an experiment.}
-    % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}
-    \label{lst:3:howToRunExperiment2}
-\end{listing}
-
-% https://tex.stackexchange.com/a/12430/
-\begin{small}
-% \begin{listing}[h!]
-    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
-    % \caption{Small example of \texttt{configuration.py} file}
-    \captionof{listing}{Example of \texttt{configuration.py} file\label{lst:3:configuration}.}
-    % \label{lst:3:configuration}
-% \end{listing}
-\end{small}
+
+% % https://tex.stackexchange.com/a/12430/
+% \begin{small}
+% % \begin{listing}[h!]
+%     \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
+%     % \caption{Small example of \texttt{configuration.py} file}
+%     \captionof{listing}{Example of \texttt{configuration.py} file\label{lst:3:configuration}.}
+%     % \label{lst:3:configuration}
+% % \end{listing}
+% \end{small}
 
@@
     % Algorithm 10 is
-    $\mathrm{MOSS}$-$\mathrm{Anytime}$ from \cite{Degenne16}, using $\alpha=1$ (\texttt{MOSSAnytime}).
-    It is a recent variant of the $\mathrm{MOSS}$ algorithm \cite{BubeckSlivkins12}, using an adaptive tuning of its parameter.
-It is anytime, and obtains good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+    $\mathrm{MOSS}$-$\mathrm{Anytime}$ from \cite{Degenne16}, using $\alpha=1$ (\texttt{MOSSAnytime}),
+    is an anytime variant of $\mathrm{MOSS}$ \cite{BubeckSlivkins12}, using an adaptive tuning of its parameter.
+It obtains good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
     Empirically, it performs usually similarly to \UCB, and is slightly more costly in terms of both time and memory.
@@
     % Algorithm 11 is
-    Approximated Finite-Horizon Gittins from \cite{Lattimore16a} (\texttt{ApproximatedFHGittins}), using $\alpha=1$.
-    It mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
+    Approximated Finite-Horizon Gittins from \cite{Lattimore16a} (\texttt{ApproximatedFHGittins}), using $\alpha=1$,
+    mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
     It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
@@
     % Algorithm 12 is
-    $\UCB\dagger$ from the same author \cite{Lattimore2018refining} (\texttt{UCBdagger}), using $\alpha=1$.
-    It mimics the \UCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
+    $\UCB\dagger$ from the same author \cite{Lattimore2018refining} (\texttt{UCBdagger}), using $\alpha=1$,
+    mimics the \UCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
     It also achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
@@
     (\texttt{klUCBswitchAnytime}),
-    is a very recent variant of the \KLUCB{} algorithm \cite{KLUCBJournal}.
+    is a very recent variant of \KLUCB{} \cite{KLUCBJournal}.
     It mimics the \KLUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
@@
     $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019} (\texttt{RCB}), using $\alpha=1$ and perturbations uniformly sampled in $[0,1]$.
-We prefer to use the simplest of the algorithms described in the article, and other algorithms in the same article used other families of distributions for the perturbations.
+We prefer to use the simplest of the algorithms described in the article.
+    %  and other algorithms in the same article used other families of distributions for the perturbations.
     It was analyzed and found to be order-optimal for problem-dependent bounds, and empirically we found that it is usually performs slightly worse than \UCB, in terms of regret, time and memory.
@@
 
-\paragraph{About the measurement protocol.}
+\textbf{About the measurement protocol.}
 %
@@
 
-\paragraph{Comparing measurements.}
+\textbf{Comparing measurements.}
+%
 But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
@@
 
-\paragraph{Normalizing data.}
+\textbf{Normalizing data.}
 %

### 2019-07-04 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
First changes from Emilie's LONG comments
Cleanup empty pages between chapters etc. Less than 300 pages YAY

@@
 % \paragraph{Note on the simulation code.}
-\paragraph{Reproducility of the experiments.}
+\paragraph{Reproducibility.}
 %
-The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
-It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in July and August $2018$.
+The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License.
+It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in the summer $2018$,
+and is published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.

### 2019-07-04 2-Chapters/4-Chapter/chapter4.tex
Cleanup empty pages between chapters etc. Less than 300 pages YAY

@@
 % ----------------------------------------------------------------------------
-\section{Conclusion}
+\section{Towards theoretical guarantees}
+% \section{Conclusion}
 \label{sec:4:conclusion}
@@
 We presented two models of wireless IoT network without sensing, inspired by the ALOHA protocol, and with or without retransmission of uplink packages in case of collisions.
-In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
+% In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
+%
+We conclude that this learning-based approach is efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rates.
 
-We conclude that this learning-based approach is efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rates.
 It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
 Unfortunately, analyzing our model with thousands of independent devices, all communicating and learning in their own (random) time scales, turned out to be of extreme difficulty.
-%
-We can also highlight that the two models we studied are complementary:
-the first model (without retransmission of packets), is interesting for its simplicity, and focused on improving the battery life of IoT devices,
-% , at the cost of a lower Quality of Service (QoS),
-%
-while the second model is more interesting if the main target is an improvement of the Quality of Service (QoS).
-% , rather than a longer battery life.
+That is why we focus on two different directions of theoretical analysis in the next chapters of this thesis.
+% %
+% We can also highlight that the two models we studied are complementary:
+% the first model (without retransmission of packets), is interesting for its simplicity, and focused on improving the battery life of IoT devices,
+% % , at the cost of a lower Quality of Service (QoS),
+% %
+% while the second model is more interesting if the main target is an improvement of the Quality of Service (QoS).
+% % , rather than a longer battery life.
 

### 2019-07-04 2-Chapters/5-Chapter/chapter5.tex
Cleanup empty pages between chapters etc. Less than 300 pages YAY

@@
 This hypothesis is realistic, mostly for energy consumption and spectrum efficiency reasons.
-Of course, we can relax this hypothesis, and in practice in some families of wireless networks, communication between devices are possible.
+Of course, we can relax it, and in practice in some families of wireless networks, communication between devices are possible.
 Note that this extension is \emph{not} considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``\emph{graph bandit}'' problem \cite{valko2016bandits}.
@@
 
-\newpage  % FIXME
+% \newpage  % FIXME
 
@@
 
-This chapter suggests several interesting further research directions.
-First, one could want to investigate the notion of \emph{optimal algorithm}, in the decentralized multi-players model with sensing information.
-So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
-However, sub-optimal draws turn out to not be the dominant term in the regret, both in our upper bound and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
-Similarly to what was done very recently in \cite{wang2019distributed}, for communications between players who collaborate with each other, it would be interesting to characterize the (minimal) number of collisions needed to achieve logarithmic regret in this decentralized model.
+% This chapter suggests several interesting further research directions.
+% First, one could want to investigate the notion of \emph{optimal algorithm}, in the decentralized multi-players model with sensing information.
+% So far we provided a state-of-the-art upper bound on the expected number of sub-optimal arms selections.
+% % , which suggests some form of (asymptotic) optimality.
+% However, sub-optimal draws turn out to not be the dominant term in the regret, both in our upper bound and in practice, thus an interesting future work is to identify some notion of \emph{minimal number of collisions}.
+% Similarly to what was done very recently in \cite{wang2019distributed}, for communications between players who collaborate with each other, it would be interesting to characterize the (minimal) number of collisions needed to achieve logarithmic regret in this decentralized model.
 
@@
 
-Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-players bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
-We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future works left on our library.
-For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
+% Additionally, we have implemented in SMPyBandits \cite{SMPyBandits} the following extensions: evaluating $M$ on the fly (Section~\ref{sub:5:unknownNumberOfPlayers}), and piece-wise stationary multi-players bandits (Section~\ref{sub:5:towardsNonStationaryModels}).
+% We have not yet implemented the other extensions, but we are interested to do it, and it is one of the major future works left on our library.
+% For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}, for tickets number \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}.
 

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Cleanup empty pages between chapters etc. Less than 300 pages YAY

@@
     \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
     \caption{Histograms of the distributions of regret $R_T$ ($T=5000$) for problem 1.}
@@
 %
-The experiments in this chapter use our library SMPyBandits,
-and the page \href{https://SMPyBandits.GitHub.io/NonStationaryBandits.html}{\texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html}} gives instructions to reproduce them.
+The experiments in this chapter use SMPyBandits,
+and the instructions to reproduce them are given on \href{https://SMPyBandits.GitHub.io/NonStationaryBandits.html}{\texttt{SMPyBandits.GitHub.io/NonStationaryBandits.html}}.
 
@@
 
-We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected.
+% We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected.
 This interesting research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$.
 We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
-We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
-Finally, we would like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,WeiSrivastava18Abruptly}.
-
-As mentioned at the end of Chapter~\ref{chapter:5}, another interesting direction of future work is to study non-stationary distributed multi-players bandits.
+% We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of algorithms, as suggested by \cite{Maillard2018GLR}.
+% Finally, we would like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,WeiSrivastava18Abruptly}.
+
+As mentioned at the end of Chapter~\ref{chapter:5}, an interesting direction of future work is to study non-stationary distributed multi-players bandits.
 A natural extension of the non-stationary model presented in this Chapter is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.

### 2019-07-04 2-Chapters/2-Chapter/chapter2.tex
Shorter conclusion for Ch2 and Ch3

@@
 	% \centering
-    \includegraphics[width=1.23\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
+    \includegraphics[width=1.20\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
 	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
@@
 	% \centering
-	\includegraphics[width=1.17\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
+	\includegraphics[width=1.12\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
 	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
@@
 %
-For both directions, we present in the two final chapters of this thesis natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
+For both directions, we present natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
 of stationary multi-players and piece-wise stationary bandits.

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
Shorter conclusion for Ch2 and Ch3

@@
 
-In this chapter, we presented our Python library SMPyBandits, by detailing three main points:
+We chose to present our Python library SMPyBandits in this chapter, before the rest of this thesis, instead of including this chapter in the Appendix of the thesis, because all the numerical experiments on multi-player bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on our library.
+%
+We detailed three main points of SMPyBandits:
 $(i)$ its purpose and qualities,

### 2019-07-04 2-Chapters/2-Chapter/chapter2.tex
A few lines shorter for list of algorithms in 3.3.1

@@
 %  either in an alternative way or in a fixed way (``explore then exploit'').
-They are detailed in Algorithm~\ref{algo:2:simpleStrategies}.
+% They are detailed in
+% % The first is detailed in
+% % Algorithm~\ref{algo:2:simpleStrategies}.
+% Algorithms~\ref{algo:2:simpleStrategies_EpsilonGreedy} and \ref{algo:2:simpleStrategies_ExploreThenExploit} below.
+% % Algorithm~\ref{algo:2:simpleStrategies_EpsilonGreedy}.
+% % and \ref{algo:2:simpleStrategies_ExploreThenExploit}.
 
@@
 		\For(){$t = 1, 2, \dots, T$}{
-            \uIf{\textcolor{deeppurple}{$\varepsilon$-greedy}}{
             Sample a value uniformly in $[0,1]$: $U_t\sim\cU([0,1])$\\
@@
             }
+            Observe a reward $r(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
 		}
-            \uIf{\textcolor{deepgold}{Explore-then-Exploit}}{
+		\caption[A simple efficient strategy, the $\varepsilon$-greedy algorithm.]{A simple efficient strategy, the \textcolor{deeppurple}{$\varepsilon$-greedy} algorithm.}
+		\label{algo:2:simpleStrategies_EpsilonGreedy}
+		% \end{small} % XXX remove if needed
+	\end{algorithm}
+	\end{framed}
+\end{figure}
+% \end{small} % XXX remove if needed
+
+% % \begin{small} % XXX remove if needed
+% \begin{figure}[h!]
+% 	\centering
+%     \begin{framed}
+% 	\begin{algorithm}[H]
+% 		% \begin{small} % XXX remove if needed
+% 		\For(){$t = 1, 2, \dots, T$}{
+%             \uIf{\textcolor{deeppurple}{$\varepsilon$-greedy}}{
+%                 Sample a value uniformly in $[0,1]$: $U_t\sim\cU([0,1])$\\
+%                 \uIf{\textcolor{deeppurple}{$U_t < \varepsilon$ (\ie, with probability $\varepsilon$)}}{
+%                     Play uniformly at random $A(t) \sim \cU(\{1,\dots,K\})$
+%                     \tcp*[f]{Explore}
+%                 }
+%                 \uElse{
+%                     Play uniformly among the arms of maximal empirical mean: $A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(t))$
+%                     \tcp*[f]{or Exploit}
+%                 }
+%             }
+%             \uIf{\textcolor{deepgold}{Explore-then-Exploit}}{
+%                 \uIf{$t \leq \textcolor{deepgold}{T_0}$}{
+%                     Play sequentially $A(t) \in 1 + (t \mod K)$
+%                     \tcp*[f]{Explore}
+%                 }
+%                 \uElse{
+%                     \uIf{$t = \textcolor{deepgold}{T_0}$}{
+%                         Pick one arm $A(\textcolor{deepgold}{T_0})=k$ uniformly among the arms of maximal empirical mean: \textcolor{deepgold}{$A(t) \sim \cU(\arg\max\limits_{1\leq k \leq K} \widehat{\mu_k}(\textcolor{deepgold}{T_0}))$}
+%                         \tcp*[f]{then Exploit}
+%                     }
+%                     Play the same arm $A(t) = A(\textcolor{deepgold}{T_0})$
+%                 }
+%             }
+%             Observe a reward $r(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
+% 		}
+% 		\caption[Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit.]{Simple efficient strategies: \textcolor{deeppurple}{$\varepsilon$-greedy} and \textcolor{deepgold}{Explore-then-Exploit}.}
+% 		\label{algo:2:simpleStrategies}
+% 		% \end{small} % XXX remove if needed
+% 	\end{algorithm}
+% 	\end{framed}
+% \end{figure}
+% % \end{small} % XXX remove if needed
+
+% https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
+-- The \textcolor{deeppurple}{$\varepsilon$-greedy strategy} (in Algorithm~\ref{algo:2:simpleStrategies_EpsilonGreedy})
+consists in alternating exploration and exploitation at a certain ratio \cite{Auer02}.
+% \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
+Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
+On the one hand, if $\varepsilon$ is constant, then the (mean) regret still grows linearly, as it is lower bounded by $(\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k) T$.
+In average, $T_0 = \varepsilon T$ steps are spent on exploration.
+%
+On the other hand, if a lower-bound $d$ on the positive gaps is known beforehand,
+we can consider a sequence $(\varepsilon_t)_{t\in\N^*}$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
+Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
+But this is not satisfactory, as $\min_k \Delta_k$ needs to be known in advance, which is usually not the case for real applications.
+Here again, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
+
+% \begin{small} % XXX remove if needed
+\begin{figure}[h!]
+	\centering
+    \begin{framed}
+	\begin{algorithm}[H]
+		% \begin{small} % XXX remove if needed
+		\For(){$t = 1, 2, \dots, T$}{
             \uIf{$t \leq \textcolor{deepgold}{T_0}$}{
@@
             }
-            }
             Observe a reward $r(t)$, and update $X_k(t)$, $N_k(t)$ and $\widehat{\mu_k}(t)$
 		}
-		\caption[Simple efficient strategies: $\varepsilon$-greedy and Explore-then-Exploit.]{Simple efficient strategies: \textcolor{deeppurple}{$\varepsilon$-greedy} and \textcolor{deepgold}{Explore-then-Exploit}.}
-		\label{algo:2:simpleStrategies}
+		\caption[A simple efficient strategy, the Explore-then-Exploit algorithm.]{A simple efficient strategy, the \textcolor{deepgold}{Explore-then-Exploit} algorithm.}
+		\label{algo:2:simpleStrategies_ExploreThenExploit}
 		% \end{small} % XXX remove if needed
@@
 
-% https://smpybandits.github.io/docs/Policies.EpsilonGreedy.html
--- The \textcolor{deeppurple}{$\varepsilon$-greedy strategy} consists in alternating exploration and exploitation at a certain ratio \cite{Auer02}.
-% \cite{SuttonBarto2018,Bubeck12,LattimoreBanditAlgorithmsBook}.
-Fix $0<\varepsilon<1$, then at each round, with probability $\varepsilon$ the player selects an arm uniformly at random (exploration) and with probability $1-\varepsilon$ the arm with highest empirical mean is selected (exploitation).
-On the one hand, if $\varepsilon$ is constant, then the (mean) regret still grows linearly, as it is lower bounded by $(\varepsilon \frac{1}{K} \sum_{k=1}^K \Delta_k) T$.
-In average, $T_0 = \varepsilon T$ steps are spent on exploration.
-%
-On the other hand, if a lower-bound $d$ on the positive gaps is known beforehand,
-we can consider a sequence $(\varepsilon_t)_{t\in\N^*}$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
-Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
-But this is not satisfactory, as $\min_k \Delta_k$ needs to be known in advance, which is usually not the case for real applications.
-Here again, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.
-
 % https://smpybandits.github.io/docs/Policies.ExploreThenCommit.html
--- The ``\textcolor{deepgold}{Explore-then-Exploit}'' strategy, as presented for instance in \cite{Bubeck12}, first explores uniformly the $K$ arms for $T_0$ time steps, then only exploits the arm identified as the best arm after these first $T_0$ steps
+-- The ``\textcolor{deepgold}{Explore-then-Exploit}'' strategy (in Algorithm~\ref{algo:2:simpleStrategies_ExploreThenExploit}), presented for instance in \cite{Bubeck12}, first explores uniformly the $K$ arms for $T_0$ time steps, then only exploits the arm identified as the best arm after these first $T_0$ steps
 (\ie, one of the arms with highest empirical means, after $T_0/K$ samples of each arm).
@@
 and a \emph{confidence radius} $\xi_k(t) \eqdef \sqrt{\alpha \frac{\log\left(t\right)}{N_k(t)}}$.
-%
+
 \begin{align}\label{eq:2:UCB_index}
@@
 %
-In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity and realistic deployment is favored (\ie, Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when we give mathematical developments (\ie, Chapters~\ref{chapter:5} and \ref{chapter:6}).
+In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity and realistic deployment is favored (in Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when we give mathematical developments (in Chapters~\ref{chapter:5} and \ref{chapter:6}).
 
@@
 
-\newpage
+% % FIXME
+% \newpage
 
@@
 % %
-We then use the small Bernoulli-distributed problem behind this demonstration
+We then used the small Bernoulli-distributed problem behind this demonstration
 to present a small numerical simulation, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
@@
 
-We presented the simplest MAB model studied in this chapter, by focussing on one agent playing a stationary game.
-Both hypotheses are removed or weaken in the rest of this manuscript,
+% We take another approach in Chapter~\ref{chapter:4}, where the
+We generalize this MAB model in Chapter~\ref{chapter:4}, by studying decentralized learning of a large set of independent players, all having different activation times.
+This extension is significantly harder than the two previously evoked ones, and we were unfortunately unable to obtain any strong theoretical results under these loose hypotheses.
+This model is however more generic and as such it was found suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
+
+We take a more formal approach in the rest of the thesis,
+% We presented the simplest MAB model studied in this chapter, by focussing on one agent playing a stationary game.
+% Both hypotheses are removed or weaken in the rest of this manuscript,
 % first in Chapter~\ref{chapter:4} where we consider many XXX
@@
 
-We take another approach in Chapter~\ref{chapter:4}, where the MAB model is generalized to study decentralized learning of a large set of independent players, all having different activation times.
-This extension is significantly harder than the two previously evoked ones, and we were unfortunately unable to obtain any strong theoretical results under these loose hypotheses.
-This model is however more generic and as such it was found suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
-
-\paragraph{Possible future works.}
-%
-We focus in this thesis on finite-arms and one-dimensional bandit problems,
-and therefore, two possible directions of future works could be to extend our works
-to MAB models with either multidimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
+% \paragraph{Possible future works.}
+% %
+% We focus in this thesis on finite-arms and one-dimensional bandit problems,
+% and therefore, two possible directions of future works could be to extend our works
+% to MAB models with either multidimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
 

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
A few lines shorter for list of algorithms in 3.3.1

--- 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
@@
 from SMPyBandits.Arms import *
-arm1 = Bernoulli(0.1)
-arm2 = Bernoulli(0.5)
-arm3 = Bernoulli(0.9)
-
+from SMPyBandits.Environment import MAB
+
+arm1 = Bernoulli(0.1) ; arm2 = Bernoulli(0.5) ; arm3 = Bernoulli(0.9)
+bernoulli_problem = MAB([arm1, arm2, arm3])
+
+# but also
 means = [0.1, 0.5, 0.9]
-from SMPyBandits.Environment import MAB
-bernoulli_problem = MAB([arm1, arm2, arm3])
-# but also
 bernoulli_problem = MAB({'arm_type': Bernoulli, 'params': means})
@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{exampleOfRewards.pdf}
+	\includegraphics[width=0.82\linewidth]{exampleOfRewards.pdf}
 	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
@@
 
-\textbf{Read the Docs.}
+\paragraph{Read the Docs.}
 Since $2017$, we have been using the free web service provided by Read the Docs (\href{https://readthedocs.org/}{\texttt{ReadTheDocs.org}}).
@@
 
-\textbf{Travis CI.}
+\paragraph{Travis CI.}
 Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
@@
     %  Algorithm 2 is
-    Explore-then-Commit \cite{GarivierETC2016} (\texttt{ETC\_KnownGap}),
+    Explore-then-Commit (\texttt{ETC\_KnownGap}),
     % shown in Algorithm~\ref{algo:2:naiveStrategies} above,
@@
     $\mathrm{Exp}3^{++}$ from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
-    It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances.
+    % It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances.
     % meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
@@
     % with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
-    It is computationally more costly that \UCB, as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
-    In practice, a few steps of a simple bisection search method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than one order of magnitude slower than \UCB.
+    It is computationally more costly that \UCB,
+    % as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
+    % In practice, a few steps of a simple bisection search method are enough to obtain a good numerical precision,
+    even if empirically we found that even in the worst cases \klUCB{} is no more than one order of magnitude slower than \UCB.
     % of more than 10 digits
@@
 
-In this chapter, we presented our Python library SMPyBandits, by detailing three main points:
+We chose to present here our Python library SMPyBandits in this chapter rather than in the Appendix, because all the numerical experiments on single-player multi-arm bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on our library.
+%
+We detailed three main points of SMPyBandits:
 $(i)$ its purpose and qualities,
@@
 %
-When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
+When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate, mean cumulated reward, real time or memory costs.
 
@@
 
-\textbf{About the measurement protocol.}
+\paragraph{About the measurement protocol.}
 %
@@
 
-\textbf{Comparing measurements.}
+\paragraph{Comparing measurements.}
 %
@@
 
-\textbf{Normalizing data.}
+\paragraph{Normalizing data.}
 %

### 2019-07-04 2-Chapters/Conclusion/conclusion.tex
A few lines shorter for list of algorithms in 3.3.1

@@
 Thus two possible directions of future research could be to extend our works
-to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
+to MAB models with either multidimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
 The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.

### 2019-07-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
First few pages of Ch7

@@
 %
-As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like $\mathrm{EXP3}$ \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
-%
-However, we found out empirically that \Selfish{} is doing surprisingly well when using \UCB-like indexes, greatly outperforming \Selfish{} based on $\mathrm{EXP3}$,
+As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like $\mathrm{Exp3}$ \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
+%
+However, we found out empirically that \Selfish{} is doing surprisingly well when using \UCB-like indexes, greatly outperforming \Selfish{} based on $\mathrm{Exp3}$,
 like what we found in Chapter~\ref{chapter:4} in harder settings.

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
First few pages of Ch7

@@
 %
-For cognitive radio applications, as well as other applications such as recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
+
+As highlighted in Chapters~\ref{chapter:1} and \ref{chapter:4} for cognitive radio applications, as well as other applications such as recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
 Indeed, in cognitive radio or IoT networks, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate, making it non stationary.
@@
 For adversarial environments, the pseudo-regret, which compares the accumulated reward of a given strategy with that of the best fixed-arm policy, is often studied.
-The pseudo regret of the EXP3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}.
+The pseudo regret of the Exp3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}.
 However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round.
 For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced in Section~8 of the seminal paper \cite{Auer02NonStochastic}.
-They propose a simple extension of the EXP3 policy, referred to as Exp3.S, that uses exponential weights like EXP3 tweaked with an adaptive forced exploration probability to passively adapt to changes.
+They propose a simple extension of the Exp3 policy, referred to as Exp3.S, that uses exponential weights like Exp3 tweaked with an adaptive forced exploration probability to passively adapt to changes.
 It was shown that Exp3.S attains a regret of $\cO(\sqrt{K T S \log(K T)})$, when comparing with an arbitrary sequence of comparators that does not switch for more than $S-1$ times.
 Running their algorithm over the non-stationary problem, and picking their comparators sequence as the best arm in each stationary interval, one can already get $\sqrt{\Upsilon_T T \log(T)}$ regret, and the only prior knowledge needed to run the Exp3.S algorithm is $T$ and $\Upsilon_T$.
-Moreover, Exp3.S is efficient both in term of storage and computation time, and simple to implement.
+Moreover, Exp3.S is efficient in terms of both storage and computation time, and is simple to implement.
 
@@
 Two main reasons explain the dynamic research on this problem:
-first, it is well known that despite their qualities, exponential-weights algorithms like Exp3 can usually be greatly outperformed by UCB-based algorithms, for stochastic problems, and so it is expected that Exp3.S can be outperformed by other algorithms, thus any practical application might benefit from algorithms more efficient than Exp3.S.
+first, it is well known that despite their qualities, exponential-weights algorithms like Exp3 can usually be greatly outperformed by UCB-based algorithms, for stochastic problems, and so it is expected that Exp3.S can be outperformed by other algorithms, thus any practical application might benefit from algorithms that are more efficient than Exp3.S.
 %
@@
 A lot of the recent research has been focused on actively adaptive algorithms, precisely because they allow to interpret their decisions, by detecting changes on arms with statistical tests.
-The piece-wise stationary MAB model was then studied by \cite{Kocsis06} and \cite{YuMannor09}.
+The piece-wise stationary MAB model has then been studied by \cite{Kocsis06} and \cite{YuMannor09}.
 %
@@
 % \TODOL{I need to write a paragraph about using the ``adversarial setting'' for piece-wise stationary.
-It is very important to state that EXP3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with NO prior knowledge of $T$ and $\Upsilon_T$.}
+It is very important to state that Exp3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with NO prior knowledge of $T$ and $\Upsilon_T$.}
 
@@
 In Section~\ref{sec:6:ChangePointDetector} we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
-We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on the regret of both variants.
+We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on their regret.
 The unified regret analysis is given in Section~\ref{sec:6:RegretAnalysis}, we prove one result in Section~\ref{sec:6:proofRegret}.
@@
 
-A \emph{piece-wise stationary bandit model} is characterized by a set of $K$ arms.
-A (random) stream of rewards $(Y_{k,t})_{t\in\N^*}$ is associated to each arm $k \in [K]$. We assume that the rewards are bounded, and without loss of generality we assume that $Y_{k,t} \in [0,1]$.
+A \emph{piece-wise stationary bandit model} generalizes the model of Chapter~\ref{chapter:2}.
+It is characterized by a set of $K$ arms, and an horizon $T$.
+A (random) stream of rewards $(Y_{k,t})_{t\in[T]}$ is associated to each arm $k \in [K]$.
+We assume that the rewards are bounded, and wlog we assume that $Y_{k,t} \in [0,1]$.
 We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
-At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) \eqdef Y_{A(t),t}$.
-At time $t$, we denote by $k_t^*$ an arm with maximal expected reward, \ie, $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$, called an optimal arm (possibly not unique).
+At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on her past observations, and receives the corresponding reward $r(t) \eqdef Y_{A(t),t}$.
+At time $t$, we denote by $k_t^*$ an optimal arm, \ie, an arm with maximal expected reward $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$ (possibly not unique).
 
 A policy $\cA$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
-Like for stationary problems (see Definition~\ref{def:2:regret} in Chapter~\ref{chapter:2}), the performance of $\cA$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $k_t^*$ at time $t$, and that of the policy $\cA$
+Like for stationary problems (see Definition~\ref{def:2:regret} in Chapter~\ref{chapter:2}), the performance of $\cA$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $k_t^*$ at time $t$ (that can change at each round), and that of the policy $\cA$.
 We use the same notation as the regret for stationary problem without ambiguity, as in this chapter we only consider the following definition,
@@
 
-In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) \emph{number of breakpoints} $\Upsilon_T$, defined by
+In the piece-wise \iid{} model, we furthermore assume that a relatively small \emph{number of breakpoints} $\Upsilon_T$ (wrt the horizon $T$, \ie, $\Upsilon_T = \smallO{T}$), which is defined by
 \begin{equation}\label{def:6:breakpoints}
@@
 \end{equation}
-If we denote $\tau^{(0)} \eqdef 0$, we define the $i$-th breakpoint recursively by
-$\tau^{(i)} \eqdef \inf\{t > \tau^{(i-1)} : \exists k : \mu_k(t) \neq \mu_k(t+1)\}$.
+If we denote $\tau^{(0)} \eqdef 0$, we define the $i$-th breakpoint by
+$\tau^{(i)} \eqdef \inf\{t > \tau^{(i-1)} : \exists k : \mu_k(t) \neq \mu_k(t+1)\}$ for $i\in[\Upsilon_T]$.
 %
@@
 Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
-Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
-For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company (or network provider) use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new objects, then one (or some) channel(s) see their mean occupancy rates abruptly change.
+Depending on the application, many scenario can be meaningful: changes can occur on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
+For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the devices of one company (or network provider) use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new devices, then one (or some) channel(s) see their mean occupancy rates abruptly change.
 
@@
 $\NCk \eqdef \sum_{t=1}^{T-1} \indic\left(\mu_k(t) \neq \mu_k(t+1)\right)$,
-for which it clearly holds that
+which is clearly bounded by
+% for which it clearly holds that
 $\NCk \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms.
-Letting $C_T \eqdef \sum_{k=1}^K \NCk$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
+If $C_T \eqdef \sum_{k=1}^K \NCk$ is the total number of change-points on the arms, we have $C_T \in \{ \Upsilon_T, \Upsilon_T + 1, \dots, K\Upsilon_T \}$.
 We illustrate the two extreme cases in problems $1$ and $2$ presented in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} below.
@@
 %
-The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while some adversarial models can be considered as a special (worst) case, with $\Upsilon_T = T-1$ (when considering an adaptive or an oblivious adversary).
-However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
-Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the EXP3.S algorithm for the piece-wise stationary problem, as explained in Section~\ref{sec:6:Introduction} above.
+The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$ (\ie, one stationary segment), while some adversarial models can be considered as a special (worst) case, with $\Upsilon_T = T-1$ (when considering an adaptive or an oblivious adversary).
+However, typical analyzes of algorithms designed for the piece-wise stationary model assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
+Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the Exp3.S algorithm for the piece-wise stationary problem, as explained in Section~\ref{sec:6:Introduction} above.
 
@@
 We consider $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
-The arm means $(\mu_k(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1} below.
+The arm means $(\mu_k(t))_{k\in[K],t\in[T]}$ are shown in Figure~\ref{fig:6:Problem_1} below.
 % (and Figure~\ref{fig:6:Problem_1} in Appendix),
@@
 This problem is close to Problem 1, with a minimum optimality gap of $0.1$ (at any time, the smallest difference between two means is at least $0.1$),
-and shown in Figure~\ref{fig:6:Problem_2}.
+and illustrated in Figure~\ref{fig:6:Problem_2}.
 However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$ for arms $0$ and $1$, and of $0.2$ for arm $2$ (between two breakpoints, the mean change of $+0.1$ for arm $0$ and $-0.1$ for arm $1$ and $-0.2$ for arm $2$).
@@
 We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
-To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
+To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or Exp3, with a strategy to account for changes in arms' distributions.
 Following the vocabulary used in previous works, we make the distinction between \emph{passively} and \emph{actively} adaptive strategies.
-On the one hand, Actively adaptive strategies monitor the arms' rewards, by using change detection algorithms \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
-On the other hand, Passively adaptive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
+On the one hand, actively adaptive strategies monitor the arms' rewards, by using change detection algorithms \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
+On the other hand, passively adaptive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
 % , while active strategies are usually tuned with a threshold for the statistical test.
@@
 %
-Their common idea is to adapt a classical policy into forgetting old rewards, .
-If the forgetting behavior is done efficiently, then the policy can efficiently focus mostly on the most recent rewards, and passively adapt to changes when they happen.
-%
-The Discounted UCB (D-UCB) algorithm was first introduced in \cite{Kocsis06}, and it is an adaptation of the \UCB{} algorithm, with a discount factor $\gamma\in(0,1)$.
-It works by decreasing all the past rewards by a multiplicative factor, when receiving a new reward from an arm, so that the recent rewards weight more in the (discounted) empirical mean used for the computation of the \UCB{} indexes.
+Their common idea is to adapt a classical policy into forgetting old rewards.
+If the forgetting behavior is done efficiently, then the policy can efficiently focus mostly on the most recent rewards, and passively adapt to changes whenever they happen.
+%
+The Discounted UCB (D-UCB) algorithm is an adaptation of the \UCB{} algorithm, first introduced in \cite{Kocsis06}.
+It works by decreasing all the past rewards by a discount factor $\gamma\in(0,1)$, when receiving a new reward from an arm, so that the recent rewards weight more in the (discounted) empirical means, that are used for the computation of its \UCB{} indexes.
 The regret of D-UCB was proven to be upper-bounded by $\cO(\sqrt{\Upsilon_T T }\log(T))$ in \cite{Garivier11UCBDiscount}, with a tuning $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, dependent on $\Upsilon_T$.
 %
-The Sliding-Window UCB (SW-UCB) uses a sliding window of a fixed size $\tau$, to store only the most $\tau$ recent rewards for each arm,
-and it was proposed by \cite{Garivier11UCBDiscount}.
+The Sliding-Window UCB (SW-UCB), proposed by \cite{Garivier11UCBDiscount}, uses a sliding window of a fixed size $\tau$, to store only the most $\tau$ recent rewards for each arm.
 They prove that tuning its window-size to $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$, gives a bound on the regret of SW-UCB of the form $\cO(\sqrt{\Upsilon_T T \log(T)})$.
 
-Both D-UCB and SW-UCB build on a stationary policy, but for example EXP3.S from \cite{Auer02NonStochastic}
-builds on the EXP3 policy, which is designed for the adversarial case.
-EXP3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
+Both D-UCB and SW-UCB build on a stationary policy, but for example Exp3.S from \cite{Auer02NonStochastic}
+builds on the Exp3 policy, which is designed for the adversarial case.
+Exp3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
 It constitues a good baseline for our numerical experiments in Section~\ref{sec:6:NumericalExperiments},
-even if we did find that EXP3.S performs worse that the most of the other approaches based on extensions of stationary bandit algorithms (\eg, SW-UCB).
+even if we did find that Exp3.S performs worse that the most of the other approaches based on extending stationary bandit algorithms (\eg, SW-UCB).
 % , so we do not include it in .
@@
 The idea of using a simple discount factor, as for D-UCB, was recently adapted to a Bayesian policy, with the Discounted Thompson sampling (DTS) algorithm presented by \cite{RajKalyani17}.
-Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust in the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in the paper.
+Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust on the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in the paper.
 The DTS algorithm can perform well in practice, for instance with $\gamma=0.75$ on Problems $1$ and $2$ (see Section~\ref{sec:6:NumericalExperiments}).
-However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
+% However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
 %
@@
 %
-We distinguish two families,
-the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
+% We distinguish two families,
+A first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
 When using an efficient Change-point Detector (CD) algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
 %
-The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
+The Adapt-EVE algorithm from \cite{Hartland06} used a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
 %
@@
 %
-The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
+The Exp3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with Exp3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
 A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.
@@
 %
-The RExp3 algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm.
+The RExp3 algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the Exp3 algorithm.
 Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).

### 2019-07-04 2-Chapters/2-Chapter/chapter2.tex
Using uniform notations for Exp3, Exp3.S, Exp3.R

@@
 a natural question is to know if there is an algorithm achieving a regret upper-bound of the form $R_T^{\cA}(I) \leq \bigO{\sqrt{K T}}$ for any instance, independently on the problem difficulty.
-The question has been answered since the early $2000$s, as it is for instance the case for the Exp3 algorithm from \cite{Auer02}.
+The question has been answered since the early $2000$s, as it is for instance the case for the \ExpThree{} algorithm from \cite{Auer02}.
 In the last decade, some algorithms were shown to achieve both a problem-dependent logarithmic and a minimax upper-bounds,
@@
 Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
-The first of them is the Exp3 policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
+The first of them is the \ExpThree{} policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
 which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details, see
@@
 %
-The Exp3 policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
+The \ExpThree{} policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
 $\Pr(A(t)=k) \eqdef (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
-The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
+The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, \ExpThree{} updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$.
 The weights are then renormalized, to have a sum $\sum_k w_k(t+1) = 1$.
@@
 %
-It was proven in \cite{Auer02NonStochastic} that Exp3 with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
+It was proven in \cite{Auer02NonStochastic} that \ExpThree{} with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
 while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
-A short proof of this results and more details on Exp3 are given in \cite{Bubeck12}.
+A short proof of this results and more details on \ExpThree{} are given in \cite{Bubeck12}.
 %
@@
 % https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
-Finally, it is worth mentioning the Exp3$^{++}$ algorithm, as a recent variant of Exp3 that uses an adaptive tuning of its parameters, proposed by \cite{Seldin17}.
+Finally, it is worth mentioning the \Exppp{} algorithm, as a recent variant of \ExpThree{} that uses an adaptive tuning of its parameters, proposed by \cite{Seldin17}.
 It is anytime, and it was also proven to obtain good ``best-of-both-world'' performance.

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
First few pages of Ch7
Using uniform notations for Exp3, Exp3.S, Exp3.R

@@
 the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
-For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{klUCB}} in \texttt{Policies} module) against the non-anytime variant $\klUCB^{++}$ algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Posterior.Beta}})).
+For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{klUCB}} in \texttt{Policies} module) against the non-anytime variant \KLUCBpp{} algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Posterior.Beta}})).
 
@@
     %  Algorithm 3 is
-    $\mathrm{Exp}3^{++}$ from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
-    % It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances.
+    \Exppp{} from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
+    % It is a recent anytime variant of the \ExpThree{} algorithm, which was proven to obtain good ``best of both world'' performances.
     % meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
@@
 
-% FIXME
-\newpage
+% % FIXME
+% \newpage
 
@@
 
-% FIXME
-% \newpage
-
 \paragraph{Conclusion.}
@@
 as well as a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\cM_T = \bigO{K}$.
+
+% FIXME
+% \vspace*{30pt}
+\clearpage
+% \newpage
 
@@
 
-\newpage  % WARNING
-\newpage  % WARNING
+% FIXME?
+% \newpage  % WARNING
+% \newpage  % WARNING
 % ----------------------------------------------------------------------------

### 2019-07-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Using uniform notations for Exp3, Exp3.S, Exp3.R

@@
 %
-As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like $\mathrm{Exp3}$ \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
-%
-However, we found out empirically that \Selfish{} is doing surprisingly well when using \UCB-like indexes, greatly outperforming \Selfish{} based on $\mathrm{Exp3}$,
+As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like \ExpThree{} \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
+%
+However, we found out empirically that \Selfish{} is doing surprisingly well when using \UCB-like indexes, greatly outperforming \Selfish{} based on \ExpThree,
 like what we found in Chapter~\ref{chapter:4} in harder settings.

### 2019-07-04 2-Chapters/5-Chapter/chapter5.tex
Using uniform notations for Exp3, Exp3.S, Exp3.R

@@
 We illustrate this in Table~\ref{table:5:comparisonRhoRandRhoLearn} below, where we consider the same problem as described above in Section~\ref{sub:5:withoutSensing} (Table~\ref{table:5:meanRegretSimulationsNoSensing}).
-We include different \rhoLearn{} algorithms, that uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
+We include different \rhoLearn{} algorithms, that uses \klUCB{}, Bayes-UCB or \ExpThree{} for the second-stage learning,
 and we include both the centralized multiple-play version of \klUCB{} and \MCTopM-\klUCB{} for comparison.
@@
         \rhoLearn-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 \pm 331$ \\
-        \rhoLearn-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5134 \pm 976$ \\
+        \rhoLearn-\klUCB{} + \ExpThree{} & $529 \pm 175$ & $1659 \pm 331$ & $5134 \pm 976$ \\
         \hline
@@
 
-Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
+Their proposal is using the \ExpThree{} algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
 As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as shown in Theorem~4.1
 (they used $K$ for $M$ the number of players, and $N$ for $K$ the number of arms).
-This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for Exp3 for the single-player adversarial setting in \cite{Auer02NonStochastic}.
+This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for \ExpThree{} for the single-player adversarial setting in \cite{Auer02NonStochastic}.
 Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Using uniform notations for Exp3, Exp3.S, Exp3.R

@@
 For adversarial environments, the pseudo-regret, which compares the accumulated reward of a given strategy with that of the best fixed-arm policy, is often studied.
-The pseudo regret of the Exp3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}.
+The pseudo regret of the \ExpThree{} algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}.
 However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round.
 For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced in Section~8 of the seminal paper \cite{Auer02NonStochastic}.
-They propose a simple extension of the Exp3 policy, referred to as Exp3.S, that uses exponential weights like Exp3 tweaked with an adaptive forced exploration probability to passively adapt to changes.
-It was shown that Exp3.S attains a regret of $\cO(\sqrt{K T S \log(K T)})$, when comparing with an arbitrary sequence of comparators that does not switch for more than $S-1$ times.
-Running their algorithm over the non-stationary problem, and picking their comparators sequence as the best arm in each stationary interval, one can already get $\sqrt{\Upsilon_T T \log(T)}$ regret, and the only prior knowledge needed to run the Exp3.S algorithm is $T$ and $\Upsilon_T$.
-Moreover, Exp3.S is efficient in terms of both storage and computation time, and is simple to implement.
-
-It could seem that the problem is considered as solved by the Exp3.S algorithm, but it was actively studied since the seminal paper \cite{Auer02NonStochastic}.
+They propose a simple extension of the \ExpThree{} policy, referred to as \ExpThreeS, that uses exponential weights like \ExpThree{} tweaked with an adaptive forced exploration probability to passively adapt to changes.
+It was shown that \ExpThreeS{} attains a regret of $\cO(\sqrt{K T S \log(K T)})$, when comparing with an arbitrary sequence of comparators that does not switch for more than $S-1$ times.
+Running their algorithm over the non-stationary problem, and picking their comparators sequence as the best arm in each stationary interval, one can already get $\sqrt{\Upsilon_T T \log(T)}$ regret, and the only prior knowledge needed to run the \ExpThreeS{} algorithm is $T$ and $\Upsilon_T$.
+Moreover, \ExpThreeS{} is efficient in terms of both storage and computation time, and is simple to implement.
+
+It could seem that the problem is considered as solved by the \ExpThreeS{} algorithm, but it was actively studied since the seminal paper \cite{Auer02NonStochastic}.
 Two main reasons explain the dynamic research on this problem:
-first, it is well known that despite their qualities, exponential-weights algorithms like Exp3 can usually be greatly outperformed by UCB-based algorithms, for stochastic problems, and so it is expected that Exp3.S can be outperformed by other algorithms, thus any practical application might benefit from algorithms that are more efficient than Exp3.S.
+first, it is well known that despite their qualities, exponential-weights algorithms like \ExpThree{} can usually be greatly outperformed by UCB-based algorithms, for stochastic problems, and so it is expected that \ExpThreeS{} can be outperformed by other algorithms, thus any practical application might benefit from algorithms that are more efficient than \ExpThreeS.
 %
@@
 % \TODOL{I need to write a paragraph about using the ``adversarial setting'' for piece-wise stationary.
-It is very important to state that Exp3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with NO prior knowledge of $T$ and $\Upsilon_T$.}
+It is very important to state that \ExpThreeS{} from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with NO prior knowledge of $T$ and $\Upsilon_T$.}
 
@@
 However, typical analyzes of algorithms designed for the piece-wise stationary model assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
-Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the Exp3.S algorithm for the piece-wise stationary problem, as explained in Section~\ref{sec:6:Introduction} above.
+Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the \ExpThreeS{} algorithm for the piece-wise stationary problem, as explained in Section~\ref{sec:6:Introduction} above.
 
@@
 We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
-To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or Exp3, with a strategy to account for changes in arms' distributions.
+To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or \ExpThree, with a strategy to account for changes in arms' distributions.
 Following the vocabulary used in previous works, we make the distinction between \emph{passively} and \emph{actively} adaptive strategies.
@@
 
-Both D-UCB and SW-UCB build on a stationary policy, but for example Exp3.S from \cite{Auer02NonStochastic}
-builds on the Exp3 policy, which is designed for the adversarial case.
-Exp3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
+Both D-UCB and SW-UCB build on a stationary policy, but for example \ExpThreeS{} from \cite{Auer02NonStochastic}
+builds on the \ExpThree{} policy, which is designed for the adversarial case.
+\ExpThreeS{} actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
 It constitues a good baseline for our numerical experiments in Section~\ref{sec:6:NumericalExperiments},
-even if we did find that Exp3.S performs worse that the most of the other approaches based on extending stationary bandit algorithms (\eg, SW-UCB).
+even if we did find that \ExpThreeS{} performs worse that the most of the other approaches based on extending stationary bandit algorithms (\eg, SW-UCB).
 % , so we do not include it in .
@@
 %
-The Exp3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with Exp3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
+The \ExpThreeR{} algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with \ExpThree, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
 A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.
@@
 
-A different setting where the quantity of interest is not $\Upsilon_T$ but the total variational budget $V_T$, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
+A different setting where the quantity of interest is not $\Upsilon_T$ but the total variational budget $V_T$, was introduced by \cite{Besbes14stochastic}, for which they proposed the R\ExpThree{} algorithm.
 The total variation budget $V_T$ is defined as $\sum_{t=1}^{T-1} \sum_{k=1}^K D(\nu_k(t), \nu_k(t+1))$, for a certain measure of difference $D$, which measures how much the two distributions of any arm $k$ have changed from time $t$ to time $t+1$.
@@
 %
-The RExp3 algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the Exp3 algorithm.
+The R\ExpThree{} algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the \ExpThree{} algorithm.
 Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
@@
 %
-first, Exp4.S actually reduces to Exp3.S in the non-contextual case, and Exp3.S is discussed above.
+first, Exp4.S actually reduces to \ExpThreeS{} in the non-contextual case, and \ExpThreeS{} is discussed above.
 % As already acknowledged in the answer to  Reviewer 1,  we were indeed unaware of the very good theoretical properties of this algorithm in the piece-wise non-stationary setting.
-We will revise our state-of-the-art section and include the Exp3.S algorithm as a baseline in our experiments.
+We will revise our state-of-the-art section and include the \ExpThreeS{} algorithm as a baseline in our experiments.
 However, preliminary experiments reveal that this algorithm is much less efficient that GLR-klUCB.
@@
 %
-Additionally, we include the Exp3.S algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \mathrm{e}) / ((\mathrm{e}-1)T)})$.
+Additionally, we include the \ExpThreeS{} algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \mathrm{e}) / ((\mathrm{e}-1)T)})$.
 
@@
 Note that on problem 2, all means change at every breakpoint, hence one could expect the Global variant to be more efficient, yet the experiments show the superiority of Location variant on every instance.
-The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17}.
-Exp3.S usually performs similarly to \MklUCB, but it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.
+The \ExpThreeS{} algorithm was found to outperform other algorithms based on \ExpThree{}, including recent variants like \Exppp{} \cite{Seldin17} or \ExpThreeR{} from \cite{Allesiardo17}.
+\ExpThreeS{} usually performs similarly to \MklUCB, but it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.
 
@@
         \hline
-        Exp3.S & $352 \pm 51$ & $310 \pm 62$ & $665 \pm 93$ \\
+        \ExpThreeS{} & $352 \pm 51$ & $310 \pm 62$ & $665 \pm 93$ \\
         \hline
@@
         \hline
-        Exp3.S & $551 \pm 63$ & $860 \pm 109$ & $723 \pm 121$ \\
+        \ExpThreeS{} & $551 \pm 63$ & $860 \pm 109$ & $723 \pm 121$ \\
         \hline
@@
 %
-First, we show the mean regret as a function of time (\ie, $R_t$ for $1 \leq t \leq T$), for $9$ of the considered algorithms (as they are outperformed by the others, Exp3.S and Discounted-\klUCB{} are not included to avoid clutter).
+First, we show the mean regret as a function of time (\ie, $R_t$ for $1 \leq t \leq T$), for $9$ of the considered algorithms (as they are outperformed by the others, \ExpThreeS{} and Discounted-\klUCB{} are not included to avoid clutter).
 %

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
Lots of changes on Exp3, Exp4 etc notations

@@
 In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms.
-We designed a variant of the Exp4 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
+We designed a variant of the \ExpFour{} algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
 The documentation page is on \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}}
@@
     % Algorithm 13 is
-    Anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018}
+    Anytime variant of \KLUCBswitch{} from \cite{GarivierHadiji2018}
     (\texttt{klUCBswitchAnytime}),
@@
 Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
-    \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as KL-UCB-Switch), and costs an-order-of-magnitude more memory.
+    \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as \KLUCBswitch), and costs an-order-of-magnitude more memory.
 
@@
 %         \hline
-%         KL-UCB-Switch &
+%         \KLUCBswitch{} &
 %             $R^{13}_{T_1,K_1} = 0 \pm 0$
@@
 %         \hline
-%         KL-UCB-Switch &
+%         \KLUCBswitch{} &
 %             $\cT^{13}_{T_1,K_1} = 0$
@@
 %         \hline
-%         KL-UCB-Switch &
+%         \KLUCBswitch{} &
 %             $\cM^{13}_{T_1,K_1} = 0$

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Lots of changes on Exp3, Exp4 etc notations

@@
 % In this chapter we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
-% We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
+% We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB, matching existing results (when $\Upsilon_T$ is known).
 % Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.
@@
 the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
-or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
-%
-The \CUSUM{} test is rather complicated and parametric: it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm.
-A change is detected when either random walks cross a threshold $h$.
+and a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
+%
+The \CUSUM{} test is rather complicated and it is parametric: it uses the first $M$ samples (for a fixed $M\in\N^*$) from one arm to compute an average $\widehat{u_0}$, and then builds two random walks, using the remaining observations for this arm.
+A change is detected when one of the random walks crosses a threshold $h$.
 It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
 The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
-and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments.
+and we found in our experiments that \PHT-\UCB{} performs very similarly to \CUSUMUCB.
 Even if it has the advantage of not requiring a parameter $M$,
@@
 %
-In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half.
-It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
-%
-Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\omega\in(0,1)$ (it is called $\alpha$ in both papers, but to avoid clutter with the $\alpha$ parameter in $\UCB_1$, see \eqref{eq:2:UCB_index}, we rename it $\omega$ in this chapter).
+In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w\in 2\N^*$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half.
+It is numerically much simpler, and has the advantage of using only a bounded memory (of order $\bigO{K w}$ for $K$ arms).
+%
+Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\omega\in(0,1)$ (note that it is called $\alpha$ in both papers, but we rename it $\omega$ in this chapter to avoid clutter with $\alpha$ of $\UCB_1$ in \eqref{eq:2:UCB_index}).
 % \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\omega/K$ at each time step,
@@
 
-When all their parameters are tuned correctly, both approaches are proven to achieve a good regret upper-bound, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
-As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
+When all their parameters are tuned correctly, both approaches are proven to achieve good regret upper-bounds, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
+As both results are valid under (slightly) different assumptions, we consider the two results to be the current state-of-the-art.
 As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$.
-But most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
+But most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) or on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
 %
 \CUSUMUCB{} achieves a better regret upper bound, of the order of $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but \emph{only} for Bernoulli distributions,
-and when its $4$ parameters ($\omega,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
-%
-\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\omega,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).
+and when its $4$ parameters ($\omega, \varepsilon, M, h$) are tuned based on problem-dependent knowledge.
+%
+\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\omega, b, w, \gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).
 
@@
 So it requires the tuning of two parameters, $w$ and $h$.
-On the other hand, \MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
-Compared to \CUSUMUCB{}, note that \MUCB{} is numerically much simpler as it only uses a bounded memory, of order $\bigO{K w}$ for $K$ arms.
+On the other hand, \MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change \emph{on one} of the arms.
+Compared to \CUSUMUCB, note that \MUCB{} is numerically much simpler as it only uses a sub-linear memory (wrt $T$), of order $\bigO{K w}$ for $K$ arms (as $w$ grows with $T$).
 
 Another interesting recent work is \cite{AuerGajaneOrtner18}, where the authors propose an efficient algorithm, AdSwitch, for the two-armed case ($K=2$).
-It proceeds in episodes, starting a new episode when the algorithm detects a change in one of the arms.
+It proceeds in episodes, starting a new episode whenever the algorithm detects a change in one of the arms.
 Each episode consists of three phases, estimation, exploitation then exploration.
@@
 %
-Another of research on actively adaptive algorithms uses a Bayesian point of view.
+Another line of research on actively adaptive algorithms uses a Bayesian point of view.
 A Bayesian CD algorithm is combined with Thompson Sampling in \cite{MellorShapiro13},
@@
 It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
-The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Chapter~\ref{chapter:25} \cite{Besson2018WCNC}, modified to efficiently aggregate a growing number of experts from techniques presented in \cite{Mourtada17}.
-At each time step, a new expert is introduced, and experts correspond to different Thompson sampling algorithms, each using a different history of pulls and rewards.
+The idea behind Memory Bandit is to use an expert aggregation algorithm, like \ExpFour{} from \cite{Auer02} or our \Aggr{} algorithm from Chapter~\ref{chapter:25} \cite{Besson2018WCNC}, modified to efficiently aggregate a growing number of experts, using techniques presented in \cite{Mourtada17}.
+%
+A new expert is introduced at each time step, and experts correspond to different Thompson sampling algorithms, each using a different history of pulls and rewards.
 Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
 %
-Note that the regret bound for these methods are still open-problems,
-and they are computationally more costly,
-therefore we choose to not include them in our experiments.
+Note that obtaining regret bounds for these methods is still an open-problem,
+and as they are computationally more costly, we chose to not include them in our experiments.
 
@@
 
-A different setting where the quantity of interest is not $\Upsilon_T$ but the total variational budget $V_T$, was introduced by \cite{Besbes14stochastic}, for which they proposed the R\ExpThree{} algorithm.
+A different setting where the quantity of interest is not $\Upsilon_T$ but the total variational budget $V_T$, was introduced by \cite{Besbes14stochastic}, for which they proposed the \RExpThree{} algorithm.
 The total variation budget $V_T$ is defined as $\sum_{t=1}^{T-1} \sum_{k=1}^K D(\nu_k(t), \nu_k(t+1))$, for a certain measure of difference $D$, which measures how much the two distributions of any arm $k$ have changed from time $t$ to time $t+1$.
@@
 %
-The R\ExpThree{} algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the \ExpThree{} algorithm.
-Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
-A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proven, which is weaker than existing results in our setting.
-
-This setting is quite different from the setting we are considering in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
+The \RExpThree{} algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the \ExpThree{} algorithm.
+% Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$
+As the total variational budget satisfies $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$, with $\Delta^{\text{change}}$ the minimum magnitude of a change-point,
+their regret bound of order of $\cO(V_T^{1/3} T^{2/3})$ is actually weaker than existing results in our setting.
+
+The slowly-varying model is quite different from the setting we are considering in this chapter, and even if the most recent works on this model are very interesting \cite{WeiSrivastava18Abruptly}, we preferred to focus on the piece-wise stationary (also called abruptly changing) model,
 hence we do no include any of these algorithms in the experiments presented in Section~\ref{sec:6:NumericalExperiments}.
@@
 where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
-and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
+and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), in order to tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
 %
@@
 For contextual bandits, two very recent works are \cite{Luo18} and \cite{ChenLeeLuoWei2019}.
-While both works target a more general setting, their algorithms are also applicable to non-contextual bandits, \ie, classical bandits like the model in this chapter.
-From Table~1 in \cite{Luo18}, one can see that their two algorithms Exp4.S and Ada-ILTCB recover the same regret guarantees as we do, in the non-contextual case:
+While both works target a more general setting, their algorithms are also applicable to non-contextual bandits, \ie, classical bandits like the model studied in this chapter.
+From Table~1 in \cite{Luo18}, one can see that their two algorithms \ExpFourS{} and Ada-ILTCB recover the same regret guarantees as we do, in the non-contextual case:
 that is, a $\cO(\Upsilon_T \sqrt{T})$ regret without knowing the number of changes and $\cO(\sqrt{\Upsilon_T T})$ with this knowledge.
@@
 %
-first, Exp4.S actually reduces to \ExpThreeS{} in the non-contextual case, and \ExpThreeS{} is discussed above.
+first, in the non-contextual case, \ExpFourS{} actually reduces to \ExpThreeS, which we discussed above.
 % As already acknowledged in the answer to  Reviewer 1,  we were indeed unaware of the very good theoretical properties of this algorithm in the piece-wise non-stationary setting.
@@
 %
-\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_k(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
+The two algorithms \CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_k(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
 %
@@
 In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
-it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{k,t} |\mu_k(t) - \mu_k(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{k,k'} |\mu_k(t) - \mu_{k'}(t)|$.
+it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\Delta^{\text{change}} = \max_{k,t} |\mu_k(t) - \mu_k(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\Delta^{\text{opt}} = \max_{k,k'} |\mu_k(t) - \mu_{k'}(t)|$.
 
@@
 Problem $1$ has only \emph{local changes}, meaning that at any change-point, only one arm sees its distribution change, while problem $2$ has only \emph{global changes}, meaning that all arms see their distribution change at each change-point.
-They illustrate both the extreme cases of having $C_T = \Upsilon_T$ (for problem $1$) and $C_T = K \Upsilon_T$ (for problem $2$).
+They illustrate the two extreme cases of having $C_T = \Upsilon_T$ (for problem $1$) and $C_T = K \Upsilon_T$ (for problem $2$).
 \begin{itemize}
@@
 %
-Of course, on a given problem, without additional prior knowledge on the difficulty of the problem at hand, the algorithm cannot know which situation is more likely to happen, only local changes ($C_T = \Upsilon_T$) or only global changes ($C_T = K \Upsilon_T$) or any intermediate setting between the two extreme cases.
+Of course, in a given environment, without additional prior knowledge on the difficulty of the problem at hand, the algorithm cannot know which situation is more likely to happen, between having only local changes ($C_T = \Upsilon_T$) or only global changes ($C_T = K \Upsilon_T$), or any intermediate setting between the two extreme cases.
 Thus we do not believe to be able to design a policy that could be uniformly better than the two variants of \GLRklUCB.
@@
 %
-Finally, on the practical side, while we can note that the proposed B-GLRT test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing much in terms of regret, in Appendix~\ref{sub:6:IdeasOptimizations}.
+Finally, on the practical side, we can note that while the proposed B-GLRT test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing much in terms of regret, in Appendix~\ref{sub:6:IdeasOptimizations}.
 
@@
 %
-Our proposal \GLRklUCB{} is inspired by both the \MUCB{} and \CUSUMUCB{} algorithms.
-Previous works focused on using \UCB{} \cite{LiuLeeShroff17,CaoZhenKvetonXie18},
+Our proposal \GLRklUCB{} is inspired by both the \MUCB{} and \CUSUMUCB{} algorithms \cite{LiuLeeShroff17,CaoZhenKvetonXie18}.
+Previous works focused on using \UCB,
 but we propose to use \klUCB{} instead, as it is known to be more efficient for Bernoulli rewards as well as for a more generic case of bounded or one-dimensional exponential families \cite{KLUCBJournal}.
-Theoretical results for these works were only given for \UCB, but they both suggest that extending the results to \klUCB{} (or other efficient stationary policies) should not be difficult.
+Theoretical results for \cite{LiuLeeShroff17} and \cite{CaoZhenKvetonXie18} were only given for \UCB, but they suggested that extending the results to \klUCB{} (or other efficient stationary policies) should not be difficult.
 For a fair comparison, we therefore chose to compare the different change-point detector algorithms combined with \klUCB.
@@
 
-Sequential change-point detection has been extensively studied in the statistical community.
+Sequential change-point detection has been extensively studied in the statistical community, from the $1930$s with pioneer works like \cite{Wilks1938} and later on with seminal works like \cite{barnard1959control,siegmund1995using}.
 We refer to the book \cite{Basseville93} for a survey.
@@
 Generalized Likelihood Ratio tests date back to the seminal work of \cite{Wilks1938}, and were for instance studied for change-point detection by \cite{barnard1959control,siegmund1995using}.
-Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, can be used for this problem.
+Exploiting the fact that bounded distribution are $(1/4)$-sub Gaussian (\ie, their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$), the (Gaussian) GLRT, recently studied in depth by \cite{Maillard2018GLR}, and can be used for this problem.
 We propose instead to exploit the fact that bounded distributions are also dominated by Bernoulli distributions.
@@
 
-The regret bounds we obtain for the two variants of our proposal \GLRklUCB{},
+The regret bounds we obtain for the two variants of our proposal \GLRklUCB,
 Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
@@
       \end{tabular}
-      \caption{Mean regret $\pm$ $1$ std-dev, on problems $1$ and $2$ with $T=5000$. We conclude that using \klUCB{} is much more efficient than using \UCB{}, for non-stationary bandit.}
+      \caption{Mean regret $\pm$ $1$ std-dev, on problems $1$ and $2$ with $T=5000$. We conclude that using \klUCB{} is much more efficient than using \UCB, for non-stationary bandit.}
       \label{table:6:comparisonUCB_klUCB}
@@
 Note that on problem 2, all means change at every breakpoint, hence one could expect the Global variant to be more efficient, yet the experiments show the superiority of Location variant on every instance.
-The \ExpThreeS{} algorithm was found to outperform other algorithms based on \ExpThree{}, including recent variants like \Exppp{} \cite{Seldin17} or \ExpThreeR{} from \cite{Allesiardo17}.
+The \ExpThreeS{} algorithm was found to outperform other algorithms based on \ExpThree, including recent variants like \Exppp{} \cite{Seldin17} or \ExpThreeR{} from \cite{Allesiardo17}.
 \ExpThreeS{} usually performs similarly to \MklUCB, but it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.

### 2019-07-04 2-Chapters/2-Chapter/chapter2.tex
Lots of changes on Exp3, Exp4 etc notations
Still changing the macros

@@
 Following this Definition~\ref{def:2:KLDivergence} for two distributions
-$d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ in the same exponential family,
+$d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ in the same exponential family $\cE$,
 we can use a shorter notation and write $\KL( \eta_1, \eta_2 ) \eqdef \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
@@
 % \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
-$\KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2) X ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2) \mu_1 - A(\eta_1) + A(\eta_2)$.
+$\KL_{\cE}\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2) X ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2) \mu_1 - A(\eta_1) + A(\eta_2)$.
 % \end{equation}
@@
     and for a Bernoulli distribution of mean $\mu\in[0,1]$, denoted $B(\mu)$,
-    the parameter is $\eta = \mu / (1 - \mu)$, giving $A(\eta) = \log(1 + \mathrm{e}^{\eta})$
+    the parameter is $\eta = \mu / (1 - \mu)$, giving $A(\eta) = \log(1 + \e^{\eta})$
     (with the limit behavior $\eta=+\infty$ if $\mu=1$).
@@
 $\UCB_1$ (with $\alpha=1$),
-\klUCB{} (with Bernoulli $\kl = \KL_{B}$),
+\klUCB{} (with Bernoulli $\kl = \KL_{\cB}$),
 and Thompson sampling (with Beta posteriors and flat uniform priors).
@@
 % Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-% We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+% We proposed an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
 % %

### 2019-07-04 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Typos, et commentaires Christophe
Still changing the macros

@@
-% \subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
+% \subsection{Proof of Proposition~\ref{prop:41:Lagrangian}}
 % \label{sec:5:proofLagrangian}
@@
     So the gradient is zero $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) |_{D_i=D_i^*} = 0$
-    iff $D_i^*$ satisfies $(1-p)^{D_i^*}( 1 + \log(1-p) D_i^*) = \lambda / (1-p)^{S_i - 1}$. Let $x = \log(1-p) D_i^*$ this is equivalent to $\mathrm{e}^{x}(1 + x) = \lambda / (1-p)^{S_i - 1}$
-    and with $y = 1 + x$, we get $\mathrm{e}^{y}y = \lambda\mathrm{e} / (1-p)^{S_i - 1}$.
+    iff $D_i^*$ satisfies $(1-p)^{D_i^*}( 1 + \log(1-p) D_i^*) = \lambda / (1-p)^{S_i - 1}$. Let $x = \log(1-p) D_i^*$ this is equivalent to $\e^{x}(1 + x) = \lambda / (1-p)^{S_i - 1}$
+    and with $y = 1 + x$, we get $\e^{y}y = \lambda\e / (1-p)^{S_i - 1}$.
 
 	\item
-    By using the $\mathcal{W}$-Lambert function $\mathcal{W}$ \cite{Corless96}, reciprocal of $y \mapsto \mathrm{e}^{y}y$, we get $x = y - 1 = \mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1$.
-    So the gradient of the Lagrangian is zero iff $D_i^* = \max(0, x/\log(1-p)) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1\right]^{+}$, because $D_i^*$ has to be non-negative.
+    By using the $\mathcal{W}$-Lambert function $\mathcal{W}$ \cite{Corless96}, reciprocal of $y \mapsto \e^{y}y$, we get $x = y - 1 = \mathcal{W}(\lambda\e / (1-p)^{S_i - 1}) - 1$.
+    So the gradient of the Lagrangian is zero iff $D_i^* = \max(0, x/\log(1-p)) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\e / (1-p)^{S_i - 1}) - 1\right]^{+}$, because $D_i^*$ has to be non-negative.
     This gives the $i$-th coordinate of the unique saddle point of $f(\boldsymbol{D})$,

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Still changing the macros

@@
     We call a \emph{sub-Bernoulli distribution} any distribution $\nu$ that satisfies
-    $\ln \bE_{X\sim\nu}\left[e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu \eqdef \bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) \eqdef \ln(1-\mu + \mu e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$, for any $\mu\in[0,1]$.
+    $\ln \bE_{X\sim\nu}\left[\e^{\lambda X}\right] \leq \phi_{\mu}(\lambda)$ with $\mu \eqdef \bE_{X\sim \nu}[X]$ and $\phi_{\mu}(\lambda) \eqdef \ln(1-\mu + \mu \e^\lambda)$ is the log moment generating function of a Bernoulli distribution with mean $\mu$, for any $\mu\in[0,1]$.
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
 Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
-As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows beforehand the values of $a$ and $b$).
+As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$.
+We assume that both the decision maker and the algorithm know beforehand the values of $a$ and $b$.
 
@@
 
-If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem would reduce to a parametric sequential test of
+If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem would reduce to a simple parametric sequential test of
 $\cH_0 : (\exists \mu_0: \forall t\in\N^*, X_t \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$,
@@
 %
-By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) \eqdef 0$ if $t=0$),
-$\kl(x,y) \eqdef x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$
+By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$,
+defined by $\kl(x,y) \eqdef x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$ for $x,y\in[0,1]$ (with the usual convention that $t \log(t) \eqdef 0$ if $t=0$).
 %
 Indeed, we show below
-% in Appendix~\ref{app:6:GLR_with_kl}
+in Appendix~\ref{app:6:GLR_with_kl}
 that for Bernoulli distributions, we have
 \begin{equation}\label{eq:6:secondDefGLRT}
-    \log \GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right].
-\end{equation}
-%
-where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{t=k}^{k'} X_t$ denotes the average of the observations $X_t$ collected between the instants $k$ and $k'$.
-This result is proven in Appendix~\ref{app:6:GLR_with_kl}.
-%
-
+    \log \GLR(n) = \sup_{s \in [n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right].
+\end{equation}
+%
+where for $s \leq s'$, $\widehat{\mu}_{s:s'} \eqdef \frac{1}{s - s' + 1} \sum_{t=s}^{s'} X_t$ denotes the average of the observations $X_s,\dots,X_{s'}$.
+% This result is proven in Appendix~\ref{app:6:GLR_with_kl}.
+%
+%
 This motivates the following definition of the Bernoulli GLRT change point detector.
@@
         \begin{equation}\label{def:6:GLR}
-            \widehat{\tau}_{\delta} \eqdef \inf \left\{ n \in \N^* : \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
+            \widehat{\tau}_{\delta} \eqdef \inf \left\{ n \in \N^* : \sup_{s \in [n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\}.
         \end{equation}
@@
     with the convention that we \emph{enforce} $\left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] = 0$
-    in the case of $\widehat{\mu}_{1:s} = \widehat{\mu}_{s+1:n}$
-    (in particular, if $\widehat{\mu}_{1:n} \in \{0,1\}$, then $\widehat{\mu}_{1:s} = \widehat{\mu}_{s+1:n} = 0$ or $\widehat{\mu}_{1:n}$).
+    whenever we have $\widehat{\mu}_{1:s} = \widehat{\mu}_{s+1:n}$,
+    in particular if $\widehat{\mu}_{1:n} \in \{0,1\}$, then $\widehat{\mu}_{1:s} = \widehat{\mu}_{s+1:n} = \widehat{\mu}_{1:n} \in \{0,1\}$.
     % Indeed in such cases, the $\kl(x,y)$
@@
 
-Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR}, but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence $d(x,y)$ between two elements in that exponential family that have mean $x$ and $y$.
-For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to the stopping time \eqref{def:6:GLR} with the choice $d(x,y) = 2(x-y)^2$, when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
-Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
-%
-In the next section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
+Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for any one-parameter exponential family $\cE$, for which the GLR test is defined as in~\eqref{def:6:GLR}, but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence $\kl_{\cE}(x,y)$ between two elements in that exponential family that have mean $x$ and $y$ (see Definition~\ref{def:2:KLDivergence}).
+For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to the stopping time \eqref{def:6:GLR} with the quadratic divergence $d(x,y) = 2(x-y)^2$, when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
+Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR.
+% , based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
+%
+In the next section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which in particular holds for any distribution supported in $[0,1]$.
 
@@
 
-When used for a bandit problem, the two main properties of a sequential change-point detection test are its false alarm probability and its detection delay.
+When used for a bandit problem, the two main properties of a sequential change-point detection test are its \emph{false alarm probability} and its \emph{detection delay}.
 %
@@
 %
-We give below two lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there are enough samples before a change.
+We give below two lemmas:
+Lemma~\ref{lem:6:FalseAlarm} bounds the false alarm probability for a certain choice of threshold function,
+and Lemma~\ref{lem:6:Delay} bounds the detection delay if there are enough samples before the change-point.
 
@@
 %
-In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small.
-To define $\beta$, we need to introduce the function $\cT$, defined for $x>0$ by
+In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is as small as we want.
+To define $\beta$, we first introduce the function $\cT$, defined for $x>0$ by
 %
@@
 %
-we define $\widetilde{h}(x) \eqdef e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\widetilde{h}(x) \eqdef (3/2) (x-\ln(\ln (3/2)))$ otherwise, for any $x \ge 0$,
+we define $\widetilde{h}(x) \eqdef \e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\widetilde{h}(x) \eqdef (3/2) (x-\ln(\ln (3/2)))$ otherwise, for any $x \ge 0$,
 and with the value $\zeta(2) = \pi^2 / 6$.
@@
 %
-The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96}, which is the inverse of $x \mapsto x \exp(x)$, as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
+The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96}, which is the inverse of $x \mapsto x \e^x$, as $h^{-1}(x) = - \mathcal{W}(- \e^{-x})$.
 
@@
 $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
+% %
+% The B-GLRT test is defined in Definition~\ref{def:6:GLRDef}.
 
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    Let be $\bP_{\mu_0}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t$.
+    If $\bP_{\mu_0}$ is a probabilistic model under which $X_t \in [0,1]$,
+    and $\forall t, \E_{\bP_{\mu_0}}[X_t] = \mu_0$,
+    % and has an average of $\mu_0$ for all $t$.
     % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
-    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
-    $\bP_{\mu_0}(\widehat{\tau}_\delta < \infty) \leq \delta$, with the threshold function
+    then the B-GLRT test $\widehat{\tau}_\delta$ (from Definition~\ref{def:6:GLRDef}) satisfies
+    % given in Definition~\ref{def:6:GLRDef}
+    $\bP_{\mu_0}(\widehat{\tau}_\delta < \infty) \leq \delta$, with the threshold
     %
@@
 \begin{equation}
-    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
-\end{equation}
-where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family.
-The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
+    \bE[\e^{\lambda X} ] \leq \e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
+\end{equation}
+where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family $\cE$.
+The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence $\kl_{\cE}$ in that exponential family.
+% (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
 %
@@
 where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
-As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$, if we denote $F(x)=\ln(1+\ln(x))$,
+As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$, if we denote $F(x) \eqdef \ln(1+\ln(x))$,
 %
@@
 %
-And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum\limits_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
+And so we have
+$\bP_{\mu_0}\left(T_{\delta} < \infty\right)
+\leq \sum\limits_{s =1}^{\infty} \exp(- \frac{\ln(3s^{3/2}/\delta)}{2} )
+= \sum\limits_{s =1}^{\infty} \frac{\delta}{3s^{3/2}}
+\leq \delta$.
 %
@@
     with moment generating function $\phi_\mu(\lambda) = \bE_{X \sim \nu_{\mu}}[e^{\lambda X}]$.
-    Let $\kl_{\cE}(\mu,\mu') \eqdef \KL(\nu_{\mu},\nu_{\mu'})$ be the divergence function associated to $\cE$.
+    Let $\kl_{\cE}(\mu,\mu') \eqdef \KL(\nu_{\mu},\nu_{\mu'})$ be the KL divergence associated to $\cE$.
     Let $(X_i)_{i \in \N^*}$ and $(Y_k)_{i \in \N^*}$ be two independent \iid{} processes, with respective means $\mu$ and $\mu'$, such that
-    $\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)}$.
+    $\bE[e^{\lambda X_1}] \leq \e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq \e^{\phi_{\mu'}(\lambda)}$.
     %
@@
         \begin{equation}
-            \bP\left(\exists r \in \N^* : s \, \kl_{\cE}\left(\widehat{\mu}_{s},\mu\right) + r \, \kl_{\cE}\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},
+            \bP\left(\exists r \in \N^* : s \, \kl_{\cE}\left(\widehat{\mu}_{s},\mu\right) + r \, \kl_{\cE}\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq \e^{-x},
         \end{equation}
@@
 %
-Additionally, we include the \ExpThreeS{} algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \mathrm{e}) / ((\mathrm{e}-1)T)})$.
+Additionally, we include the \ExpThreeS{} algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \e) / ((\e-1)T)})$.
 

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Almost done for proofreading Ch7, only typos on \mu_i... to check

@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    If $\bP_{\mu_0}$ is a probabilistic model under which $X_t \in [0,1]$,
-    and $\forall t, \E_{\bP_{\mu_0}}[X_t] = \mu_0$,
+    For a probabilistic model $P = \bP_{\mu_0}$ under which $X_t \in [0,1]$,
+    and $\forall t, \E_P[X_t] = \mu_0$,
     % and has an average of $\mu_0$ for all $t$.
     % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
-    then the B-GLRT test $\widehat{\tau}_\delta$ (from Definition~\ref{def:6:GLRDef}) satisfies
+    the B-GLRT test $\widehat{\tau}_\delta$ (from Definition~\ref{def:6:GLRDef}) satisfies
     % given in Definition~\ref{def:6:GLRDef}
@@
 %
-Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from $\mu_0$ to $\mu_1$ occurs at time $\tau$.
+Another key feature of a change-point detector is its \emph{detection delay} under a model in which a change from mean $\mu_0$ to mean $\mu_1$ occurs at time $\tau$.
 We already observed that from Pinsker's inequality, the B-GLRT stops earlier than a Gaussian GLR.
-Hence, one can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the B-GLRT.
+Hence, we can leverage some techniques from \cite{Maillard2018GLR} to upper bound the detection delay of the B-GLRT.
 Letting $\Delta = |\mu_0 - \mu_1|$, one can essentially establish that for $\tau$ larger than $(1/\Delta^2)\ln(1/\delta)$ (\ie, enough samples before the change), the delay can be of the same magnitude (\ie, enough samples after the change).
-In the bandit analysis to follow, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and the corresponding Lemma not included here, but given in Appendix of \cite{Besson2019GLRT}).
+In the bandit analysis in Section~\ref{sec:6:RegretAnalysis}, the detection delay will be crucially used to control the probability of the good event (in Lemma~\ref{lem:6:GoodEventGlobal} and the corresponding Lemma not included here, but given in Appendix of \cite{Besson2019GLRT}).
 
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
-    and let $\Delta = |\mu_0 - \mu_1|$.\\
-    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
+    For a probabilistic model $P' = \bP_{\mu_0,\mu_1,\tau}$ under which $X_t \in [0,1]$,
+    % and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
+    and $\forall t \leq \tau, \E_{P'}[X_t] = \mu_0$,
+    $\forall t > \tau, \E_{P'}[X_t] = \mu_1$,
+    % with $\Delta = |\mu_0 - \mu_1|$,
+    the B-GLRT test satisfies
     \begin{equation}
@@
     \end{equation}
+    where $\Delta = |\mu_0 - \mu_1|$ denotes the gap of the change-point.
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 
-For a threshold $\beta$ chosen as in the Lemma~\ref{lem:6:FalseAlarm}, we can show that a consequence of the Lemma~\ref{lem:6:Delay} is that if the break-point $\tau$ takes place after about $\Delta^{-2} \ln(1/\delta)$ samples, the detection time may be of the same magnitude (with high probability).
-%
-In other words, if the B-GLRT test observes enough samples before a change-point,
-its delay $\widehat{\tau}_\delta$ is in the order of $O(\Delta^{-2} \ln(1/\delta))$, with high probability.
+For a threshold $\beta$ chosen as in the Lemma~\ref{lem:6:FalseAlarm}, a consequence of the Lemma~\ref{lem:6:Delay} is that if the break-point $\tau$ happens after about $\cO(\Delta^{-2} \ln(1/\delta))$ samples, the detection time will be of the same magnitude (with high probability).
+%
+In other words, if the B-GLRT test gathers enough samples before a change-point,
+its delay $\widehat{\tau}_\delta$ is of order $\cO(\Delta^{-2} \ln(1/\delta))$, with high probability.
 
@@
 $\beta(n,\delta) = \left(1 + \frac{1}{n}\right)\ln\left(\frac{3n\sqrt{n}}{\delta}\right)$,
-can be used to prove an upper bound of $\delta$ for the false alarm probability of the GLR, with quadratic divergence $\kl(x,y)=2(x-y)^2$.
+can be used to prove an upper bound of $\delta$ for the false alarm probability of the GLR, with quadratic divergence $d(x,y)=2(x-y)^2$.
 %
@@
 
-Also note that, as any test based on scan-statistics, the B-GLRT can be costly to implement as at every time step, it considers all previous time steps as a possible position for a change-point.
-Thus, down-sampling the possible time steps is an interesting adaptation in practice:
+Also note that, as any test based on scan-statistics, the B-GLRT can be costly to implement as at every time step $t$, it considers all previous time steps as a possible position for a change-point (\ie, $s\in[n-1]$ if there are $n$ samples).
+Thus, it can be interesting in practice to down-sample the possible values of both $n$ and $s$, that is to use a large stopping time, defined by
 %
@@
 However, the detection delay may be slightly increased.
-In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups
-has little impact on the regret of the bandit strategy introduced below.
+In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical tweaks
+has little impact on the regret of the bandit strategy introduced below, while speeding up its computation time.
 
@@
 
-We start by giving the pseudo-code of our algorithm, by explaining every part, then we give its finite-time regret upper bounds, for the two variants of using local or global restarts.
+We start by giving the pseudo-code of our algorithm, and by explaining every part, then we give its finite-time regret upper bounds, for the two variants of using local or global restarts.
 The analysis in both cases is using a unified proof technique, that we present before giving more details about the proof of half of the results (the other proof can be found in \cite{Besson2019GLRT}).
@@
 
-We now present the \GLRklUCB{} algorithm, which combines a bandit algorithm with a change-point detector running on each arm.
-It also needs a third ingredient, some forced exploration parameterized by $\omega\in(0,1)$ to ensure each arm is sampled enough and changes can also be detected on arms currently under-sampled by the bandit algorithm.
-%
-\GLRklUCB{} combines the \klUCB{} algorithm as it is given and analyzed by \cite{KLUCBJournal}, known to be optimal for Bernoulli bandits, with the B-GLRT change-point detector introduced in Section~\ref{sec:6:ChangePointDetector}.
-%
-This algorithm, formally stated as Algorithm~\ref{algo:6:GLRklUCB}, can be used in any bandit model with bounded rewards, and we expect it to be very efficient for Bernoulli distributions, which are relevant for practical applications.
+We now present our \GLRklUCB{} algorithm, which combines a bandit algorithm with a change-point detector running on each arm.
+It also needs a third ingredient, some forced exploration parameterized by $\omega\in(0,1)$ to ensure that each arm is sampled enough and in order to also be able to detect changes on arms currently under-sampled by the bandit algorithm.
+%
+\GLRklUCB{} combines the \klUCB{} algorithm, as it is given and analyzed by \cite{KLUCBJournal}, with the B-GLRT change-point detector introduced in Section~\ref{sec:6:ChangePointDetector}.
+%
+This algorithm, formally stated as Algorithm~\ref{algo:6:GLRklUCB}, can be used in any bandit model with bounded rewards, and we expect it to be very efficient at least for Bernoulli distributions, which are the most relevant for our applications of interest (see Chapter~\ref{chapter:1} and \ref{chapter:4}).
 
@@
 The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms.
-A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
+A restart happens when the B-GLRT change-point detector detects a change on the arm $k$ that has been played (line $9$).
 %
@@
     \begin{equation}\label{eq:whenDoesGLRklUCBrestart}
-        \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right)\right] \geq \beta(n,\delta),
+        \sup_{s \in [n-1]} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right)\right] \geq \beta(n,\delta),
     \end{equation}
@@
 %
-with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}.
-We remind in this notation, if $\widehat{Z}_{1:s}=\widehat{Z}_{s+1:n}$, \ie, if all observations $Z_1,\dots,Z_n$ are equal, we set the left-hand side of \eqref{eq:whenDoesGLRklUCBrestart} to $0$ and the B-GLRT cannot detect a change-point.
+with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as we recommend in practice (see Section~\ref{sub:6:PracticalConsiderations}).
+We remind that in this notation, if $\widehat{Z}_{1:s}=\widehat{Z}_{s+1:n}$ (\ie, if all observations $Z_1,\dots,Z_n$ are equal), then we set the left-hand side of \eqref{eq:whenDoesGLRklUCBrestart} to $0$, and the B-GLRT cannot detect a change-point.
 %
@@
 $\tau_k(t)$ the last restart that happened for arm $k$ before time $t$,
-$n_k(t) = \sum_{s=\tau_k(t)+1}^{t} \indic(A(s)=k)$
+$n_k(t) \eqdef \sum_{s=\tau_k(t)+1}^{t} \indic(A(s)=k)$
 the number of selections of arm $k$, and
-$\widehat{\mu}_k(t) = \frac{1}{n_k(t)} \sum_{s=\tau_k(t)+1}^{t}Y_{k,s} \indic(A(s)=i)$
+$\widehat{\mu}_k(t) \eqdef \frac{1}{n_k(t)} \sum_{s=\tau_k(t)+1}^{t}Y_{k,s} \indic(A(s)=i)$
 their empirical mean (if $n_k(t)\neq0$).
 %
-With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
+With the exploration function $f(t) \eqdef \ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
 the index is defined using the binary relative entropy $\kl$ as
@@
 %-----------------------------------------------------------------------------
-\textbf{Two options for restarts.}
+\paragraph{Two options for restarts.}
 %
@@
 %
-Under local restart, in the general case the times $\tau_k(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate}.
-%
-One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative is currently not correctly supported by theory, as we found mistakes in the analysis of \CUSUMUCB: the main problem resides in the use of Hoeffding's inequality with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
+Under local restart, in the general case the times $\tau_k(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate} $f(t - \tau_k(t))$.
+%
+It is important to highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} indexes are defined differently\footnote{~This alternative is currently not correctly supported by theory, as we found mistakes in the analysis of \CUSUMUCB: the main problem resides in the use of Hoeffding's inequality with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq.
 $(31)$-$(32)$ in the paper \cite{LiuLeeShroff17}.}:
@@
 %-----------------------------------------------------------------------------
-\textbf{Threshold function $\beta$.}
+\paragraph{Threshold function $\beta$.}
 %
@@
 %-----------------------------------------------------------------------------
-\textbf{Forced exploration.}
+\paragraph{Forced exploration.}
 %
@@
 
-\textbf{Other forced exploration schemes?}
+\paragraph{Other forced exploration schemes?}
 %

### 2019-07-04 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Almost done

@@
 %
-In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $k$ at round $t$ has some mean $\mu_k(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $k_t^* \eqdef \arg\max_i \mu_k(t)$.
+In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $k$ at round $t$ has some mean $\mu_k(t)$, that is constant on intervals between two \emph{break-points}, and the regret is measured with respect to the \emph{current} best arm $k_t^* \eqdef \arg\max_i \mu_k(t)$.
 
@@
 We introduce the model in Section~\ref{sec:6:BanditSetting}, and review related works in Section~\ref{sec:6:relatedWork}.
-In Section~\ref{sec:6:ChangePointDetector} we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
+In Section~\ref{sec:6:ChangePointDetector} we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a change-point Detector (CPD) algorithm.
 We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on their regret.
@@
 
-In the piece-wise \iid{} model, we furthermore assume that a relatively small \emph{number of breakpoints} $\Upsilon_T$ (wrt the horizon $T$, \ie, $\Upsilon_T = \smallO{T}$), which is defined by
-\begin{equation}\label{def:6:breakpoints}
+In the piece-wise \iid{} model, we furthermore assume that a relatively small \emph{number of break-points} $\Upsilon_T$ (wrt the horizon $T$, \ie, $\Upsilon_T = \smallO{T}$), which is defined by
+\begin{equation}\label{def:6:break-points}
     \Upsilon_T \eqdef \sum_{t=1}^{T-1} \indic\left(\exists k\in[K]: \mu_k(t) \neq \mu_k(t+1)\right).
 \end{equation}
-If we denote $\tau^{(0)} \eqdef 0$, we define the $i$-th breakpoint by
+If we denote $\tau^{(0)} \eqdef 0$, we define the $i$-th break-point by
 $\tau^{(i)} \eqdef \inf\{t > \tau^{(i-1)} : \exists k : \mu_k(t) \neq \mu_k(t+1)\}$ for $i\in[\Upsilon_T]$.
@@
 
-Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
-Depending on the application, many scenario can be meaningful: changes can occur on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
+Note than when a break-point occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
+Depending on the application, many scenario can be meaningful: changes can occur on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each break-point.
 For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the devices of one company (or network provider) use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new devices, then one (or some) channel(s) see their mean occupancy rates abruptly change.
@@
 and illustrated in Figure~\ref{fig:6:Problem_2}.
-However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$ for arms $0$ and $1$, and of $0.2$ for arm $2$ (between two breakpoints, the mean change of $+0.1$ for arm $0$ and $-0.1$ for arm $1$ and $-0.2$ for arm $2$).
+However, all arms change at every break-point (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$ for arms $0$ and $1$, and of $0.2$ for arm $2$ (between two break-points, the mean change of $+0.1$ for arm $0$ and $-0.1$ for arm $1$ and $-0.2$ for arm $2$).
 The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$ markers}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$ markers}).
@@
         The means are again in $[0,1]$, and there are also $C+1=5$ stationary intervals of equal lengths.
-        % Some changes do not modify the optimal arm, but the mean of each arm does change at all change points.
+        % Some changes do not modify the optimal arm, but the mean of each arm does change at all change-points.
     }
@@
 A first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
-When using an efficient Change-point Detector (CD) algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
+When using an efficient change-point Detector (CD) algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
 %
@@
 As both results are valid under (slightly) different assumptions, we consider the two results to be the current state-of-the-art.
-As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$.
+As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of break-points $\Upsilon_T$.
 But most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) or on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
@@
 %
-The two algorithms \CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_k(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
+The two algorithms \CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_k(t))$-- mostly saying that the break-points are sufficiently far away from each other.
 %
 However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
-Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
+Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of break-points  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
 %
@@
 % -----------------------------------------------------------------
-\section{The Bernoulli GLRT Change-Point Detector}
+\section{The Bernoulli GLRT change-point Detector}
 \label{sec:6:ChangePointDetector}
@@
 %
-This motivates the following definition of the Bernoulli GLRT change point detector.
+This motivates the following definition of the Bernoulli GLRT change-point detector.
 % Note that the $\exp$ appearing in the \eqref{eq:6:secondDefGLRT} is absorbed as a $\log$ in the threshold function $\beta$.
@@
 \begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
-    The Bernoulli GLRT (B-GLRT) change point detector with threshold function $\beta(n,\delta)$ is the stopping time $\widehat{\tau}_{\delta}$ defined by
+    The Bernoulli GLRT (B-GLRT) change-point detector with threshold function $\beta(n,\delta)$ is the stopping time $\widehat{\tau}_{\delta}$ defined by
     %
@@
             \mbox{Observe the reward} $Y_{A(t),t}$ : $Z_{A(t), n_{A(t)}} = Y_{A(t),t}$\;
-            \uIf(\tcp*[f]{Change point is detected}){
+            \uIf(\tcp*[f]{change-point is detected}){
                 $\GLR_\delta(Z_{A(t),1}, \dots, Z_{A(t), n_{A(t)}})$ = True
@@
 %
-The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
+Our proposal \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
 Both approaches are parameterized by $\omega\in(0,1)$.
@@
 \begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
-    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $k$ (\ie, for a $i\in[\NCk]$, one has $\tau_k^{(i)} = \tau_k(t) < s \leq t < \tau_k^{(i+1)}$) it holds that
+    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $k$, \ie, for a $i\in[\NCk]$, one has $\tau_k^{(i)} = \tau_k(t) < s \leq t < \tau_k^{(i+1)}$, it holds that
     % \begin{equation}
@@
 We present in Appendix~\ref{sec:6:exploringDifferentForcedExplorationMechanisms} numerical simulations that compare three different options of forced exploration schemes.
-We compare the uniformly random exploration used by \CUSUMUCB, against the deterministic scheme used in Algorithm~\ref{algo:6:GLRklUCB}, and against a more complicated scheme based on \emph{tracking}.
+We compare the uniformly random exploration used by \CUSUMUCB, against the deterministic scheme used in Algorithm~\ref{algo:6:GLRklUCB}, and against a more complicated scheme based on \emph{tracking} and inspired by \cite{Garivier16BAI}.
 To sum-up these simulations, the deterministic scheme gives an efficient change-point detection algorithm, and an efficient \GLRklUCB{} policy.
@@
 
-Recall that $\tau^{(i)}$ denotes the position of the $i$-th break-point and let $\mu_k^{(i)}$ be the mean of arm $k$ on the segment between the $i$- and $(i+1)$-th breakpoint:
+Recall that $\tau^{(i)}$ denotes the position of the $i$-th break-point and let $\mu_k^{(i)}$ be the mean of arm $k$ on the segment between the $i$- and $(i+1)$-th break-point:
 $\forall t \in \{ \tau^{(i-1)}+1, \dots, \tau^{(i)} \}, \mu_k(t) = \mu_k^{(i)}$.
@@
 
-Assumption~\ref{ass:6:LongPeriodsGlobal} is easy to interpret and standard in non-stationary bandits.
-It requires that the distance between two consecutive breakpoints is large enough: how large depends on the magnitude of the largest change that happen at those two breakpoints.
+Assumption~\ref{ass:6:LongPeriodsGlobal} is easy to interpret,
+and it is actually a standard assumption in non-stationary bandits.
+It requires that the distance between two consecutive break-points is large enough:
+how large depends on the magnitude of the largest change that happen at those two break-points.
 Under this assumption, we provide in Theorem~\ref{thm:6:mainRegretBoundGlobal} a finite time problem-dependent regret upper bound.
-It features the parameters $\omega$ and $\delta$,
-the KL-divergence terms $\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})$ expressing the hardness of the (stationary) MAB problem between two breakpoints,
-and the $\Delta^{(i)}$ terms expressing the hardness of the change-point detection problem.
+It features the algorithm's parameters $\omega$ and $\delta$,
+the KL-divergence terms $\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})$ expressing the hardness of the (stationary) MAB problem between two break-points,
+and the $\Delta^{(i)}$ terms expressing the hardness of the change-point detection problem in the $i$-th segment.
 
@@
     \begin{align}
-        R_T \;\; \leq \;\; & 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T \\
-        & + \sum_{k=1}^{\Upsilon_T}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \textcolor{red}{\bigO{\sqrt{\ln(T)}}}.
+        R_T \;\; \leq \;\; & 2\sum_{i=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T \\
+        & + \sum_{k=1}^{K}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \textcolor{red}{\bigO{\sqrt{\ln(T)}}}.
 \nonumber
@@
     For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriodsGlobal},
-    let $\Delta^{\text{opt}}$ denote the smallest value of a sub-optimality gap on one of the stationary segments, and $\Delta^{\text{change}}$ be the smallest magnitude of any change point on any arm.
+    let $\Delta^{\text{opt}}$ denote the smallest value of a sub-optimality gap on one of the stationary segments, and $\Delta^{\text{change}}$ be the smallest magnitude of any change-point on any arm,
+    then the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based Global Restarts satisfies
     \begin{enumerate}
@@
 Some new notations are needed to state a regret bound for \GLRklUCB{} using \emph{local restarts},
-and to distinguish notations between the results for the two variants (local and global restarts), we denote $\ell$ instead of $k$ for the indexes of change-points.
-We let $\tau_k^{(\ell)}$ denote the position of the $\ell$-th change point \emph{for arm $k$}: $\tau_k^{(\ell)} = \inf \{ t > \tau_k^{(\ell - 1)} : \mu_k(t) \neq \mu_k(t+1)\}$,
+and to distinguish notations between the results for the two variants (local and global restarts), we denote $\ell$ instead of $i$ for the indexes of change-points.
+We let $\tau_k^{(\ell)}$ denote the position of the $\ell$-th change-point \emph{for arm $k$}: $\tau_k^{(\ell)} = \inf \{ t > \tau_k^{(\ell - 1)} : \mu_k(t) \neq \mu_k(t+1)\}$,
 with the convention $\tau_k^{(0)}=0$, and let $\overline{\mu}_k^{(\ell)}$ be the $\ell$-th value for the mean of arm $k$, such that $\forall t \in [\tau_k^{(\ell-1)}+1, \tau_k^{(\ell)}], \ \ \mu_k(t) = \overline{\mu}_k^{(\ell)}$.
@@
 It requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
-Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of breakpoints $\Upsilon_T$ is replaced with the \emph{total} number of change points $\EffChange \eqdef \sum_{k=1}^K \NCk \leq K \Upsilon_T$.
+Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of break-points $\Upsilon_T$ is replaced with the \emph{total} number of change-points $\EffChange \eqdef \sum_{k=1}^K \NCk \leq K \Upsilon_T$.
 
@@
 We can directly obtain different regret upper-bounds for different choices of the two parameters $\omega$ and $\delta$.
-We prefer to state the four cases, to highlight the modularity of the result given by Theorem~\ref{thm:6:mainRegretBound}.
+We prefer to state the four cases, as this highlights the modularity of our analysis given by Theorem~\ref{thm:6:mainRegretBound}.
 
@@
     For ``easy'' problems satisfying the corresponding Assumption~\ref{ass:6:LongPeriods},
-    with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, the following regret bounds hold.
+    with $\Delta^{\text{opt}}$ and ${\Delta}^{\text{change}}$ defined as in Corollary~\ref{cor:6:Global}, then the regret of \GLRklUCB{} with parameters $\omega$ and $\delta$ based Local Restarts satisfies
 
@@
 The regret bounds we obtain for the two variants of our proposal \GLRklUCB,
-Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
+in Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
 both show that there exists a tuning of $\omega$ and $\delta$ as a function and $T$ \emph{and the number of changes} such that the regret is of order $\cO_h(K\sqrt{\Upsilon_T T \ln(T)})$ and $\cO_h(K\sqrt{\EffChange T \ln(T)})$ respectively, where the $\cO_h$ notations ignore the gap terms.
@@
 %
-Besides those specific instances, our results are essentially worse for \emph{local} than \emph{global restarts}.
+Besides those specific instances, our results are essentially worse for \emph{local} than for \emph{global restarts}.
 However, we only obtain regret upper bounds -- thus providing a theoretical safety net for both variants of our algorithm, and the practical story is different, as discussed in Section~\ref{sec:6:NumericalExperiments}.
-We find that \GLRklUCB{} performs better with local restarts, uniformly on all problems.
-
-One can note that with the tuning of $\omega$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive breakpoints (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ rounds.
-Hence those guarantees are valid on ``easy'' problems only, with few changes of large magnitudes, in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
+Indeed, we find that \GLRklUCB{} performs better with local restarts, uniformly on all problems.
+
+One can note that with the tuning of $\omega$ and $\delta$ prescribed by Corollaries~\ref{cor:6:Global} and \ref{cor:6:Local}, the regret bounds of \GLRklUCB{} hold for problem instances for which two consecutive break-points (or change-points on an arm) are separated by more than (about) $\sqrt{T\log(T)}/(\Delta^{\text{change}})^2$ rounds.
+Hence those guarantees are valid on ``easy'' problems only, with ``few'' changes of ``large'' magnitudes, in particular they do not hold for the harder problems $3$ or $5$ presented in Section~\ref{sec:6:NumericalExperiments}.
 %
 However, this does not prevent our algorithms from performing well on more realistic instances, as shown by the numerical experiments,
-for instance with problem $3$ presented in the next Section~\ref{sec:6:NumericalExperiments}).
+for instance with problem $3$ presented in the next Section~\ref{sec:6:NumericalExperiments}.
 %
@@
 
-This Section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the proof of the finite-time regret bound for the variant using \emph{global restarts}.
+This section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the proof of the finite-time regret bound for the variant using \emph{global restarts}.
 The proof of the other variant using \emph{local restarts} follows the skeleton of the unified analysis, and it can be found in Appendix~E of \cite{Besson2019GLRT}.
@@
 
-In this section, we sketch a unified proof for the two Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given above.
+% In this section, we sketch a unified proof for the two Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} given above.
 %  whose detailed proofs are given in Section~\ref{proof:6:mainRegretBoundGlobal} and in Appendix~E of our paper \cite{Besson2019GLRT} respectively.
@@
 and as we observe that $\mu_{k_t^*} - \mu_{A(t)} \leq 1$ because rewards are assumed to be bounded in $[0,1]$,
-we can write a first decomposition of the regret,
+we can write a first decomposition of the regret, first without the expectation defining the mean regret:
 \begin{align*}
-    & R_T = \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t)) \leq \ldots \\
+    & \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t)) \\
     & \hspace{1cm} \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\omega)) + \sum_{t=1}^T (\mu_{k_t^*}(t) - \mu_{A(t)}(t))\indic\left(t \notin\cD(T,\omega), \UCB_{A(t)}(t) \geq \UCB_{k_t^*}(t)\right) \\
@@
 %
-To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(i)}+1, \tau^{(i+1)}]$ for \emph{global restarts} and $[\tau^{(i)}_k+1, \tau^{(i+1)}_k]$ for each arm $k$ for \emph{local restarts}, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.
-
-The tricky part of each proof, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$.
+To control $(A)$ and $(B)$, we split the sum over consecutive segments, $[\tau^{(i)}+1, \tau^{(i+1)}]$ for \emph{global restarts} and $[\tau^{(i)}_k+1, \tau^{(i+1)}_k]$ for each arm $k$ for \emph{local restarts}, and use elements from the analysis of \klUCB{} from \cite{KLUCBJournal}.
+
+The tricky part of both proofs, which crucially exploits Assumption~\ref{ass:6:LongPeriodsGlobal} or \ref{ass:6:LongPeriods}, is actually to obtain an upper bound on $\bP(\cE_T^c)$.
 For example for \emph{local restarts} (Theorem~\ref{thm:6:mainRegretBound}), the good event is defined as
@@
 %
-where $\widehat{\tau}_k^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $k$ and $d_k^{(\ell)} = d_k^{(\ell)}(\omega,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
+where $\widehat{\tau}_k^{(\ell)}$ is defined as the $\ell$-th change detected by the algorithm on arm $k$, and $d_k^{(\ell)} = d_k^{(\ell)}(\omega,\delta)$ is defined in Assumption~\ref{ass:6:LongPeriods}.
 Introducing the event
@@
 \item
-Term $(a)$ is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector run in isolation.
-Observe that under the bandit algorithm, the change point detector associated to arm $k$ is based on (possibly much) less than $t - \tau_k(t)$ samples from arm $k$, which makes false alarm even less likely to occur.
+Term $(a)$ is related to the control of probability of false alarm, which is given by Lemma~\ref{lem:6:FalseAlarm} for a change-point detector running in isolation.
+Observe that under the bandit algorithm, the change-point detector associated to arm $k$ is based on (possibly much) less than $t - \tau_k(t)$ samples from arm $k$, which makes false alarm even less likely to occur.
 Hence, it is easy to show that $(a) \leq \delta$.
@@
 Term $(b)$ is related to the control of the detection delay, which is more tricky to obtain under the \GLRklUCB{} adaptive sampling scheme,
-when compared to a result like Theorem~6 in \cite{Maillard2018GLR} for the change-point detector run in isolation.
-%
-More precisely, we need to leverage the forced exploration (Proposition~\ref{prop:6:EnoughSamples}) to be sure we have enough samples for detection.
+when compared to a result like Lemma~\ref{lem:6:Delay}, or Theorem~6 in \cite{Maillard2018GLR} for the change-point detector running in isolation.
+%
+More precisely, we need to leverage the forced exploration to be sure we have enough samples for detection, thanks to Proposition~\ref{prop:6:EnoughSamples}.
 This explains why delays defined in Assumption~\ref{ass:6:LongPeriods} are scaled by $1/\omega$.
-Using some elementary calculus and a concentration inequality given in Lemma~\ref{lem:6:Chernoff2arms}, we can finally prove that $(b) \leq \delta$.
+Using some elementary calculus and a concentration inequality given in Lemma~\ref{lem:6:Chernoff2arms}, we can finally also prove that $(b) \leq \delta$.
 \end{itemize}
 
-Finally, the ``bad event'' is unlikely, $\bP(\cE_T^c) \leq 2\EffChange \delta$.
-By putting together the three pieces, that are a bound on $(A)$, on $(B)$ and on $\bP(\cE_T^c)$, we obtain the desired finite-time upper-bound on the regret of our proposal \GLRklUCB.
+Finally by controlling terms $(a)$ and $(b)$, the ``bad event'' is unlikely: $\bP(\cE_T^c) \leq 2\EffChange \delta$.
+By putting together the three pieces, that are a bound on $(A)$, $(B)$ and $\bP(\cE_T^c)$, we obtain the desired finite-time upper-bound on the regret of our proposal \GLRklUCB.
 
@@
 \end{equation}
-Under the good event, all the change points are detected within a delay at most $d^(k)$.
+Under the good event, all the change-points are detected within a delay at most $d^{(i)}$.
 %
@@
 \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
-    With $\cE_T(\delta)$ defined as in \eqref{def:6:GoodEvenGlobal}, the ``bad event'' is unlikely
+    % With $\cE_T(\delta)$ defined as in \eqref{def:6:GoodEvenGlobal}, the ``bad event'' is unlikely
+    The ``bad event''  $\cE_T(\delta)$ defined in \eqref{def:6:GoodEvenGlobal} is unlikely:
     % \begin{equation}
@@
     (A) & \leq \bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(n_{k_t^*}(t) \, \kl\left(\widehat{\mu}_{k_t^*}(t), \mu_{k_t^*}(t)\right) \geq f(t - \tau(t))\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}(t))\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}\bE\left[ d^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)} + \sum_{k=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(i)}) \sum_{t=\widehat{\tau}^{(i)}}^{\tau^{(i+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right],
+    & \leq \sum_{i=1}^{\Upsilon_T}\bE\left[\indic(\cE_T) \sum_{t=\tau^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}(t))\right)\right] \\
+    & \leq \sum_{i=1}^{\Upsilon_T}\bE\left[ d^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}} \indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}^{(i)}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right] \\
+    & \leq \sum_{i=1}^{\Upsilon_T}d^{(i)} + \sum_{i=1}^{\Upsilon_T}\bE\left[\indic(\cC^{(i)}) \sum_{t=\widehat{\tau}^{(i)}}^{\tau^{(i+1)}}\indic\left(n_{k^*}(t) \, \kl\left(\widehat{\mu}_{k^*}(t), \mu_{k^*}\right) \geq f(t - \widehat{\tau}^{(i)})\right)\right],
 \end{align*}
@@
     & \leq \indic(\cC^{(i)}) \sum_{t'=1}^{\tau^{(i+1)} - \widehat{\tau}^{(i)}} \bP\left(\exists s \leq t' : s \, \kl(\widehat{\mu}_{s},\mu_{k^*}^{(i)}) \geq f(t')\right) \\
-    & \leq \sum_{t=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
+    & \leq \sum_{2=1}^T \frac{1}{t\ln(t)} \leq \ln(\ln(T)),
 \end{align*}
@@
 where the last but one inequality relies on the concentration inequality given in Lemma~2 of \cite{KLUCBJournal} and the choice $f(t) = \ln(t) + 3 \ln(\ln(t))$.
+The last inequality comes from a sum-integral comparison, as $\int_{1}^T \frac{1}{t \ln(t)} = \ln(\ln(T))$.
 Finally, using the law of total expectation yields
 \begin{equation}\label{eq:6:TermAFinalGlobal}
-    (A) \leq \sum_{k=1}^{\Upsilon_T}\left[d^{(i)} + \ln(\ln(T))\right].
+    (A) \leq \sum_{i=1}^{\Upsilon_T}\left[d^{(i)} + \ln(\ln(T))\right].
 \end{equation}
@@
 \begin{align*}
-    (B) & \leq \bE\Big[\indic(\cE_T)\sum_{k=1}^{\Upsilon_T} \sum_{t=\tau^{(i)}}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}\bE\Big[\indic(\cE_T)\widehat{\tau}^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\indic\left(A(t)=i,\UCB_k(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
+    (B) & \leq \bE\Big[\indic(\cE_T)\sum_{i=1}^{\Upsilon_T} \sum_{t=\tau^{(i)}}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+    & \leq \sum_{i=1}^{\Upsilon_T} \bE\Big[\indic(\cE_T)\widehat{\tau}^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+    & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\indic\left(A(t)=i,\UCB_k(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+    & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \sum_{i=1}^{\Upsilon_T} \left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
     &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
@@
 \begin{align*}
-    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
-    & \leq \sum_{k=1}^{\Upsilon_T}d_k^{(i)}+ \sum_{k=1}^K \sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cC^{(i)}) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big].
+    & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \sum_{i=1}^{\Upsilon_T} \left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
+    & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \sum_{i=1}^{\Upsilon_T} \left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cC^{(i)}) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big].
 \end{align*}
@@
 \begin{align*}
-    (B)& \leq \sum_{k=1}^{\Upsilon_T}d^{(i)}+\sum_{k=1}^K\sum_{k=1}^{\Upsilon_T}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right) \sum_{s=1}^{\tau^{(i+1)}-\tau^{(i)}}\bP\left(s \times \kl(\widehat{\mu}_{s}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\cdot
+    (B)& \leq \sum_{i=1}^{\Upsilon_T} d^{(i)}+\sum_{k=1}^K\sum_{i=1}^{\Upsilon_T} \left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right) \sum_{s=1}^{\tau^{(i+1)}-\tau^{(i)}}\bP\left(s \times \kl(\widehat{\mu}_{s}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\cdot
 \end{align*}
@@
 \begin{align}\label{eq:6:TermBFinalGlobal}
-    (B) & \leq \sum_{k=1}^{\Upsilon_T}d^{(i)}+\sum_{k=1}^K\sum_{k=1}^{\Upsilon_T}\indic\left(\mu_k^{(i)} \neq \mu_{k^*}^{(i)}\right)\left[ \frac{\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)}{\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})}\log(T) + \bigO{\sqrt{\log(T)}}\right].
+    (B) & \leq \sum_{i=1}^{\Upsilon_T} d^{(i)}+\sum_{k=1}^K\sum_ik=1}^{\Upsilon_T} \indic\left(\mu_k^{(i)} \neq \mu_{k^*}^{(i)}\right)\left[ \frac{\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)}{\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})}\log(T) + \bigO{\sqrt{\log(T)}}\right].
 \end{align}
@@
 Combining the regret decomposition~\eqref{eq:6:GeneRegretBound} with Lemma~\ref{lem:6:GoodEventGlobal} and the two upper bounds of $(A)$ in \eqref{eq:6:TermAFinalGlobal} and of $(B)$ in \eqref{eq:6:TermBFinalGlobal},
-\[R_T \leq 2\sum_{k=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{\Upsilon_T}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
+\[R_T \leq 2\sum_{i=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T  + \sum_{k=1}^{K}\sum_{i : \mu_k^{(i)} \neq \mu_{k^*}^{(i)}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \bigO{\sqrt{\ln(T)}},\]
 which concludes the proof.
@@
 
-Recall that $\cC^{(i)}$ defined in \eqref{def:6:EventCiGlobal} is the event that all the breakpoints up to the $k$-th have been correctly detected.
+Recall that $\cC^{(i)}$ defined in \eqref{def:6:EventCiGlobal} is the event that all the break-points up to the $k$-th have been correctly detected.
 Using a union bound, one can write
@@
 \begin{align*}
-    \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T}
+    \bP(\cE_T^c) & \leq \sum\limits_{i=1}^{\Upsilon_T}
     \bP\left(\left. \widehat{\tau}^{(i)} \notin \{ \tau^{(i)}+1, \dots, \tau^{(i)} + d^{(i)} \} \right|\; \cC^{(i-1)}\right)
@@
 \begin{align*}
-    \bP(\cE_T^c) & \leq \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(i)} \leq \tau^{(i)} \;|\; \cC^{(i-1)}\right)}_{(a)} \;\; + \;\; \sum\limits_{k=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)} \;|\; \cC^{(i-1)}\right)}_{(b)}.
+    \bP(\cE_T^c) & \leq \sum\limits_{i=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(i)} \leq \tau^{(i)} \;|\; \cC^{(i-1)}\right)}_{(a)} \;\; + \;\; \sum\limits_{i=1}^{\Upsilon_T} \underbrace{\bP\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)} \;|\; \cC^{(i-1)}\right)}_{(b)}.
 \end{align*}
 %
-The final result follows by proving that $(a) \leq K \delta$ and $(b)\leq \delta$, as detailed below.
+The final result $\bP(\cE_T^c) \leq (K+1)\delta$ follows by proving that $(a) \leq K \delta$ and $(b)\leq \delta$, as detailed below.
 
@@
 %
-We have $\widehat{\tau}^{(i)} \leq \tau^{(i)}$ and it implies that there exists an arm whose associated change point detector has experienced a false-alarm:
+We have $\widehat{\tau}^{(i)} \leq \tau^{(i)}$, which implies that there exists an arm whose associated change-point detector has experienced a false-alarm, \ie,
 %
 \begin{align*}
-    (a) & \leq \bP\left(\exists i, \exists s < t \leq n_k(\tau_k^{(i)}) : s \, \kl\left(\widetilde{\mu}_{i,1:s}^{(i-1)},\widetilde{\mu}_{i,1:t}^{(i-1)}\right) + (t - s) \, \kl\left(\widetilde{\mu}_{i,s+1:t}^{(i-1)},\widetilde{\mu}_{i,1:t}^{(i-1)}\right) > \beta(t, \delta) \;|\; \cC^{(i-1)}\right) \\
+    (a) & \leq \bP\left(\exists k\in[K], \exists s < t \leq n_k(\tau_k^{(i)}) : s \, \kl\left(\widetilde{\mu}_{k,1:s}^{(i-1)},\widetilde{\mu}_{k,1:t}^{(i-1)}\right) + (t - s) \, \kl\left(\widetilde{\mu}_{k,s+1:t}^{(i-1)},\widetilde{\mu}_{k,1:t}^{(i-1)}\right) > \beta(t, \delta) \;|\; \cC^{(i-1)}\right) \\
     & \leq \sum_{k=1}^K\bP\left(\exists s < t : s \, \kl(\widehat{\mu}_{1:s},\mu_k^{(i-1)}) + (t - s) \, \kl(\widehat{\mu}_{s+1:t}, \mu_k^{(i-1)}) > \beta(t, \delta)\right),
 \end{align*}
-with $\widehat{\mu}_{s:s'} = \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_k^{(i-1)}$.
+where teh last inequality comes from a union bound on $k\in[K]$,
+and with $\widehat{\mu}_{s:s'} \eqdef \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_k^{(i-1)}$.
 Indeed, conditionally to $\cC^{(i-1)}$, the $n_k(\tau^{(i)})$ successive observations of arm $k$ arm starting from $\widehat{\tau}^{(i)}$ are \iid{} with mean $\mu_k^{(i-1)}$.
@@
 From the definition of $\Delta^{(i)}$, there exists an arm $k$ such that $\Delta^{(i)} \eqdef |\mu_k^{(i)} - \mu_k^{(i-1)}|$.
-We shall prove that it is unlikely that the change-point detector associated to $i$ doesn't trigger within the delay $d^{(i)}$.
+We shall prove that it is unlikely that the change-point detector associated to $i$ does not trigger within the delay $d^{(i)}$.
 
@@
 We include in this study two algorithms designed for the classical MAB, \klUCB{} \cite{Garivier11KL}, and Thompson sampling \cite{AgrawalGoyal11,Kaufmann12Thompson},
-as well as an ``oracle'' version of \klUCB, that we call Oracle-Restart. This algorithm knows the exact locations of the breakpoints, and restarts \klUCB{} at those locations (without any delay).
+as well as an ``oracle'' version of \klUCB, that we call Oracle-Restart. This algorithm knows the exact locations of the break-points, and restarts \klUCB{} at those locations (without any delay).
 
@@
 For a fair comparison, all algorithms that use  \UCB{} as a sub-routine were adapted to use \klUCB{} instead, which yields better performance\footnote{\cite{LiuLeeShroff17,CaoZhenKvetonXie18} both mention that extending their analysis to the use of \klUCB{} should not be too difficult.}.
-Moreover, all the algorithms are tuned as described in the corresponding paper, using in particular the knowledge of the number of breakpoints $\Upsilon_T$ and the horizon $T$.
+Moreover, all the algorithms are tuned as described in the corresponding paper, using in particular the knowledge of the number of break-points $\Upsilon_T$ and the horizon $T$.
 We first include three \emph{passively adaptive algorithms}:
@@
 It is much harder, with $\Upsilon=81$ change-points on $K=5$ arms for a longer horizon of $T=100000$.
-Some arms change at almost every time steps, for a total number of breakpoints $C=179$, but the optimal arm is almost always the same one (\textcolor{red}{arm $0$, with $\bullet$}).
+Some arms change at almost every time steps, for a total number of break-points $C=179$, but the optimal arm is almost always the same one (\textcolor{red}{arm $0$, with $\bullet$}).
 It is a good benchmark to see if the actively adaptive policies do not detect \emph{too many changes}, as the Oracle-Restart policy suffers higher regret in comparison to \klUCB.
@@
 On these problems, \GLRklUCB{} with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts.
-Note that on problem 2, all means change at every breakpoint, hence one could expect the Global variant to be more efficient, yet the experiments show the superiority of Location variant on every instance.
+Note that on problem 2, all means change at every break-point, hence one could expect the Global variant to be more efficient, yet the experiments show the superiority of Location variant on every instance.
 The \ExpThreeS{} algorithm was found to outperform other algorithms based on \ExpThree, including recent variants like \Exppp{} \cite{Seldin17} or \ExpThreeR{} from \cite{Allesiardo17}.
@@
 Problem $4$ use $K=3$ arms, and a first long stationary sequence.
-Problem $5$ use $K=5$, $T=100000$ and is much harder with $\Upsilon=82$ breakpoints and $C=179$ changes.}
+Problem $5$ use $K=5$, $T=100000$ and is much harder with $\Upsilon=82$ break-points and $C=179$ changes.}
     \label{table:6:totalResults2}
@@
 We highlight that the best non-oracle strategies are actively adaptive,
-thus the experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.
+thus the experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change-point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.
 

### 2019-07-05 2-Chapters/3-Chapter/chapter3.tex
Few typos

@@
 	\includegraphics[width=0.82\linewidth]{exampleOfRewards.pdf}
-	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
+	\caption[Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$, $0.5$ and $0.9$.]{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
 	\label{fig:3:exampleOfRewards}
@@
 	\includegraphics[width=1.05\linewidth]{3_hist.pdf}
-	\caption{Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}.
+	\caption[Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}.
+For instance, Thomson sampling is very efficient in average, and \UCB shows a larger variance.]{Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}.
 For instance, \textcolor{gold}{Thomson sampling} is very efficient in average (\textcolor{gold}{in yellow}), and \textcolor{red}{\UCB} shows a larger variance (\textcolor{red}{in red}).}

### 2019-07-05 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Few typos

--- 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
@@
     \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
-    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
+    \caption[Less than $400$ communication slots (\ie, less than $100$ trials in each channel) are sufficient for the two learning devices (\UCB{} and Thompson Sampling) to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the \textcolor{cyan}{non-learning (uniform)} device, which stays around $40\%$ of success.
+Similar gains of performance were obtained in other scenarios]{Less than $400$ communication slots (\ie, less than $100$ trials in each channel) are sufficient for the two learning devices (\textcolor{red}{\UCB} and \textcolor{green}{Thompson Sampling}) to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the \textcolor{cyan}{non-learning (uniform)} device, which stays around $40\%$ of success.
 Similar gains of performance were obtained in other scenarios.}

### 2019-07-05 2-Chapters/Conclusion/conclusion.tex
Lots of changes on Exp3, Exp4 etc notations
Few typos

@@
 
-A more precise summary of each chapter is given below.
+% A more precise summary of each chapter is given below.
 
@@
 Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for a practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+We proposed an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
 
@@
 \footnote{~These notebooks are hosted on both the GitHub repository of SMPyBandits and on my website on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}.
-They can be used locally if you install the library, but can also be used on cloud platforms, such as Binder, on \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} or Google Colab, on \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}, from your browser.}.
+They can be used locally if you install the library, but can also be used on cloud platforms, like Binder \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} or Google Colab \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}}.
 
@@
 
-Another very interesting direction of research is to propose a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and automatically adapt to the setting at hand.
+Another interesting question is to look for a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and automatically adapt to the setting at hand.
 For instance, even if \GLRklUCB{} is very efficient for piece-wise stationary problems, its forced exploration makes it sub-optimal for stationary problems.

### 2019-07-05 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Typos, and slowly updating references (old arXiv -> conf/journal version)

@@
     \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
-    \caption[Less than $400$ communication slots (\ie, less than $100$ trials in each channel) are sufficient for the two learning devices (\UCB{} and Thompson Sampling) to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the \textcolor{cyan}{non-learning (uniform)} device, which stays around $40\%$ of success.
+    \caption[Less than $400$ communication slots (\ie, less than $100$ trials in each channel) are sufficient for the two learning devices (\UCB{} and Thompson Sampling) to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success.
 Similar gains of performance were obtained in other scenarios]{Less than $400$ communication slots (\ie, less than $100$ trials in each channel) are sufficient for the two learning devices (\textcolor{red}{\UCB} and \textcolor{green}{Thompson Sampling}) to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the \textcolor{cyan}{non-learning (uniform)} device, which stays around $40\%$ of success.

### 2019-07-05 2-Chapters/2-Chapter/chapter25.tex
Lots of changes on Exp3, Exp4 etc notations
Lots of improvements on the arXiv papers

@@
 Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, another solution could be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We propose an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+We propose an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
 
@@
 comparing the regret of several algorithms against different aggregation algorithms.
-Theoretical guarantees are shortly discussed, before we conclude in Section~\ref{sec:25:conclusion}.
+Theoretical guarantees are shortly discussed in Section~\ref{sec:25:theory}.
+% , before we conclude in Section~\ref{sec:25:conclusion}.
 

### 2019-07-05 2-Chapters/3-Chapter/chapter3.tex
Lots of improvements on the arXiv papers

@@
 	\includegraphics[width=0.82\linewidth]{exampleOfRewards.pdf}
-	\caption[Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$, $0.5$ and $0.9$.]{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
+	\caption[Histogram of $10000$ \iid{} rewards obtained from three arms with a truncated Gaussian distribution, of respective means $0.1$, $0.5$ and $0.9$.]{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
 	\label{fig:3:exampleOfRewards}

### 2019-07-05 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Typos, and slowly updating references (old arXiv -> conf/journal version)
Okay relire toutes ces notations compliquées _k ^i

@@
     \begin{equation}\label{eq:whenDoesGLRklUCBrestart}
-        \sup_{s \in [n-1]} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right)\right] \geq \beta(n,\delta),
+        \sup_{s \in [n-1]} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^n Z_k , \frac{1}{n}\sum_{i=1}^n Z_k\right)\right] \geq \beta(n,\delta),
     \end{equation}
@@
 the number of selections of arm $k$, and
-$\widehat{\mu}_k(t) \eqdef \frac{1}{n_k(t)} \sum_{s=\tau_k(t)+1}^{t}Y_{k,s} \indic(A(s)=i)$
+$\widehat{\mu}_k(t) \eqdef \frac{1}{n_k(t)} \sum_{s=\tau_k(t)+1}^{t}Y_{k,s} \indic(A(s)=k)$
 their empirical mean (if $n_k(t)\neq0$).
@@
     % Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
+    \vspace*{-30pt}% FIXME
 \end{smallproof}
@@
 with the convention $\tau_k^{(0)}=0$, and let $\overline{\mu}_k^{(\ell)}$ be the $\ell$-th value for the mean of arm $k$, such that $\forall t \in [\tau_k^{(\ell-1)}+1, \tau_k^{(\ell)}], \ \ \mu_k(t) = \overline{\mu}_k^{(\ell)}$.
-We also introduce the gap $\Delta_k^{(\ell)} = \overline{\mu}_k^{\ell} - \overline{\mu}_k^{\ell-1} > 0$.
+We also introduce the gap $\Delta_k^{(\ell)} = \overline{\mu}_k^{(\ell)} - \overline{\mu}_k^{(\ell-1)} > 0$.
 
@@
 
-Assumption~\ref{ass:6:LongPeriods} is again easy to interpret, like Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard in non-stationary bandits, and to the best of or knowledge, our analysis is the first one to be given for such assumption on the problem difficulty.
+Assumption~\ref{ass:6:LongPeriods} is easy to interpret, as Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard the literature, and to the best of or knowledge, our analysis is the first one to consider such assumption on the problem difficulty.
 %
 It requires that any two consecutive change-points \emph{on a given arm} are sufficiently spaced (relatively to the magnitude of those two change-points).
-Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of break-points $\Upsilon_T$ is replaced with the \emph{total} number of change-points $\EffChange \eqdef \sum_{k=1}^K \NCk \leq K \Upsilon_T$.
+Under that assumption, Theorem~\ref{thm:6:mainRegretBound} provides a regret upper bound that scales with similar quantities as that of Theorem~\ref{thm:6:mainRegretBoundGlobal}, except that the number of break-points $\Upsilon_T$ is replaced with the \emph{total} number of change-points $\EffChange \eqdef \sum_{k=1}^K \NCk$, which verifies $\Upsilon_T \leq \EffChange \leq K \Upsilon_T$.
 
@@
     \end{enumerate}
-    \vspace*{-20pt}  % FIXME
+    % \vspace*{-10pt}  % FIXME
 % \end{leftbar}  % XXX leftbar corollarybar, comment if needed
@@
 The proof uses these notations:
-let $\widehat{\tau}^{(i)}$ be the $k$-th change detected by the algorithm, leading to the $k$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
-We denote $n_k(t) \eqdef \sum_{s=\tau(t)+1}^{t}\indic(A(s) = i)$ the number of selections of arm $k$ since the last (global) restart, and $\widehat{\mu}_k(t) \eqdef \frac{1}{n_k(t)}\sum_{s=\tau(t)+1}^{t}Y_{k,s}\indic(A(s) = i)$ their empirical average (if $n_k(t) \neq 0$).
+let $\widehat{\tau}^{(i)}$ be the $i$-th change detected by the algorithm, leading to the $i$-th (full) restart and let $\widehat{\tau}(t)$ be the last time before $t$ that the algorithm restarted.
+We denote $n_k(t) \eqdef \sum_{s=\tau(t)+1}^{t}\indic(A(s) = k)$ the number of selections of arm $k$ since the last (global) restart, and $\widehat{\mu}_k(t) \eqdef \frac{1}{n_k(t)}\sum_{s=\tau(t)+1}^{t}Y_{k,s}\indic(A(s) = k)$ their empirical average (if $n_k(t) \neq 0$).
 
 As explained before, our analysis relies on the general regret decomposition~\eqref{eq:6:GeneRegretBound}, with the following appropriate good event.
-Let $d^k$ be defined as in Assumption~\ref{ass:6:LongPeriodsGlobal}, we define
+Let $d^(i)$ be defined as in Assumption~\ref{ass:6:LongPeriodsGlobal}, we define
 \begin{equation}\label{def:6:GoodEvenGlobal}
-    \cE_T(\delta) \eqdef \left( \forall k \in [\Upsilon_T], \widehat{\tau}^{(i)} \in \left[\tau^{(i)} + 1, \tau^{(i)} + d^{(i)}\right] \right).
+    \cE_T(\delta) \eqdef \left( \forall i \in [\Upsilon_T], \widehat{\tau}^{(i)} \in \left[\tau^{(i)} + 1, \tau^{(i)} + d^{(i)}\right] \right).
 \end{equation}
@@
 %
-where we introduce the event $\cC^{(i)}$ that all the changes up to the $k$-th have been detected:
+where we introduce the event $\cC^{(i)}$ that all the changes up to the $i$-th have been detected:
 %
@@
 %
-where the last but one inequality relies on the concentration inequality given in Lemma~2 of \cite{KLUCBJournal} and the choice $f(t) = \ln(t) + 3 \ln(\ln(t))$.
-The last inequality comes from a sum-integral comparison, as $\int_{1}^T \frac{1}{t \ln(t)} = \ln(\ln(T))$.
-Finally, using the law of total expectation yields
+where the last but one inequality relies on the concentration inequality given in Lemma~2 of \cite{KLUCBJournal}, for the choice $f(t) = \ln(t) + 3 \ln(\ln(t))$.
+The last inequality comes from a sum-integral comparison, as $\int\limits_{1}^T \frac{1}{t \ln(t)} \mathrm{d}t = \ln(\ln(T))$.
+Finally, the law of total expectation gives
 \begin{equation}\label{eq:6:TermAFinalGlobal}
@@
     & \leq \sum_{i=1}^{\Upsilon_T} \bE\Big[\indic(\cE_T)\widehat{\tau}^{(i)} + \indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_{A(t)}^{(i)}\right)\indic\left(\UCB_{A(t)}(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
-    & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\indic\left(A(t)=i,\UCB_k(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
+    & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\indic\left(A(t)=k,\UCB_k(t) \geq \mu_{k^*}^{(i)}\right)\Big] \\
     & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \sum_{i=1}^{\Upsilon_T} \left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\times \\
-    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = i,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
-    % \\
-\end{align*}%
-\begin{align*}
+    &\hspace{1cm}\bE\Big[\indic(\cE_T) \sum_{t=\widehat{\tau}^{(i)}+1}^{\tau^{(i+1)}}\sum_{s=1}^{t-\widehat{\tau}^{(i)}}\indic\left(A(t) = k,n_k(t) = s\right) \indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \widehat{\tau}^{(i)})\right)\Big]
+    \\
+% \end{align*}%
+% \begin{align*}
     & \leq \sum_{i=1}^{\Upsilon_T} d_k^{(i)}+ \sum_{k=1}^K \sum_{i=1}^{\Upsilon_T} \left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)\bE\Big[\indic(\cE_T) \sum_{s=1}^{n_k(\tau^{(i+1)})}\indic\left(s \, \kl(\widetilde{\mu}_{k,s}^{(i)}, \mu_{k^*}^{(i)}) \leq f(\tau^{(i+1)} - \tau^{(i)})\right)\Big]  \\
@@
 \begin{align}\label{eq:6:TermBFinalGlobal}
-    (B) & \leq \sum_{i=1}^{\Upsilon_T} d^{(i)}+\sum_{k=1}^K\sum_ik=1}^{\Upsilon_T} \indic\left(\mu_k^{(i)} \neq \mu_{k^*}^{(i)}\right)\left[ \frac{\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)}{\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})}\log(T) + \bigO{\sqrt{\log(T)}}\right].
+    (B) & \leq \sum_{i=1}^{\Upsilon_T} d^{(i)}+\sum_{k=1}^K\sum_{i=1}^{\Upsilon_T} \indic\left(\mu_k^{(i)} \neq \mu_{k^*}^{(i)}\right)\left[ \frac{\left(\mu_{k^*}^{(i)} - \mu_k^{(i)}\right)}{\kl(\mu_k^{(i)},\mu_{k^*}^{(i)})}\log(T) + \bigO{\sqrt{\log(T)}}\right].
 \end{align}
@@
 \end{align*}
-And thus we have
+And thus we can split between a term corresponding to a false alarm and a term corresponding to a bounded detection delay,
 \begin{align*}
@@
 %
-The final result $\bP(\cE_T^c) \leq (K+1)\delta$ follows by proving that $(a) \leq K \delta$ and $(b)\leq \delta$, as detailed below.
+The final result follows by proving that $(a) \leq K \delta$ and $(b)\leq \delta$, as detailed below.
 
@@
 \begin{align*}
-    (a) & \leq \bP\left(\exists k\in[K], \exists s < t \leq n_k(\tau_k^{(i)}) : s \, \kl\left(\widetilde{\mu}_{k,1:s}^{(i-1)},\widetilde{\mu}_{k,1:t}^{(i-1)}\right) + (t - s) \, \kl\left(\widetilde{\mu}_{k,s+1:t}^{(i-1)},\widetilde{\mu}_{k,1:t}^{(i-1)}\right) > \beta(t, \delta) \;|\; \cC^{(i-1)}\right) \\
+    (a) & \leq \bP\bigl(\exists k\in[K], \exists s < t \leq n_k(\tau_k^{(i)})
+        : s \, \kl\left(\widehat{\mu}_{k,1:s}^{(i-1)},\widehat{\mu}_{k,1:t}^{(i-1)}\right) \\
+    & \hspace*{150pt} + (t - s) \, \kl\left(\widehat{\mu}_{k,s+1:t}^{(i-1)},\widehat{\mu}_{k,1:t}^{(i-1)}\right) > \beta(t, \delta) \;|\; \cC^{(i-1)}\bigr) \\
     & \leq \sum_{k=1}^K\bP\left(\exists s < t : s \, \kl(\widehat{\mu}_{1:s},\mu_k^{(i-1)}) + (t - s) \, \kl(\widehat{\mu}_{s+1:t}, \mu_k^{(i-1)}) > \beta(t, \delta)\right),
 \end{align*}
-where teh last inequality comes from a union bound on $k\in[K]$,
+where the last inequality simply comes from a union bound on $k\in[K]$,
 and with $\widehat{\mu}_{s:s'} \eqdef \sum_{r=s}^{s'} Z_{i,r}$ where $Z_{i,r}$ is an \iid{} sequence with mean $\mu_k^{(i-1)}$.
@@
 From the definition of $\Delta^{(i)}$, there exists an arm $k$ such that $\Delta^{(i)} \eqdef |\mu_k^{(i)} - \mu_k^{(i-1)}|$.
-We shall prove that it is unlikely that the change-point detector associated to $i$ does not trigger within the delay $d^{(i)}$.
+We shall prove that it is unlikely that the change-point detector associated to this arm $k$ does not trigger within the delay $d^{(i)}$.
 
@@
 (as the mapping $t \mapsto n_k(t) - n_k(\tau^{(i)})$ is non-decreasing, is $0$ at $t=\tau^{(i)}$ and its value at $\tau^{(i)}+d^{(i)}$ is larger than $\overline{r}$).
-Using that $\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right) \subseteq \left(\widehat{\tau}^{(i)} \geq \overline{t}\right)$
+Using the inclusion $\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right) \subseteq \left(\widehat{\tau}^{(i)} \geq \overline{t}\right)$,
 the event $\left(\widehat{\tau}^{(i)} \geq \tau^{(i)} + d^{(i)}\right)$ further implies that
 \[
-    n_k(\tau^{(i)}) \, \kl\left(\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})},\widetilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right)
-    + \overline{r} \, \kl\left(\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})},\widetilde{\mu}^{k-1}_{i,n_k(\overline{t})}\right) \leq \beta(n_k(\tau^{(i)}) + \overline{r},\delta),
+    n_k(\tau^{(i)}) \, \kl\left(\widetilde{\mu}^{(i-1)}_{i,n_k(\tau^{(i)})},\widetilde{\mu}^{(i-1)}_{i,n_k(\overline{t})}\right)
+    + \overline{r} \, \kl\left(\widetilde{\mu}^{(i-1)}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})},\widetilde{\mu}^{(i-1)}_{i,n_k(\overline{t})}\right) \leq \beta(n_k(\tau^{(i)}) + \overline{r},\delta),
 \]
-where $\widetilde{\mu}^{k-1}_{k,s}$ denotes the empirical mean of the $s$ first observation of arm $k$ since the $(k-1)$-th restart $\widehat{\tau}^{(i-1)}$ and  $\widetilde{\mu}^{k-1}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
-Conditionally to $\cC^{(i-1)}$, $\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})}$ is the empirical mean of $n_k(\tau^{(i)})$ \iid{} replications of mean $\mu^{k-1}$, whereas $\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_k^{k}$.
+where $\widetilde{\mu}^{(i-1)}_{k,s}$ denotes the empirical mean of the $s$ first observation of arm $k$ since the $(i-1)$-th restart $\widehat{\tau}^{(i-1)}$ and  $\widetilde{\mu}^{(i-1)}_{i,s:s'}$ the empirical mean that includes observation number $s$ to number $s'$.
+Conditionally to $\cC^{(i-1)}$, $\widetilde{\mu}^{(i-1)}_{i,n_k(\tau^{(i)})}$ is the empirical mean of $n_k(\tau^{(i)})$ \iid{} replications of mean $\mu^{(i-1)}$, whereas $\widetilde{\mu}^{(i-1)}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})}$ is the empirical mean of $\overline{r}$ \iid{} replications of mean $\mu_k^{(i)}$.
 
@@
 \begin{equation*}
-    n_k(\tau^{(i)}) \in \left\{ \left\lfloor \frac{\omega}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} \eqdef \cI_k.
+    n_k(\tau^{(i)}) \in \left\{ \left\lfloor \frac{\omega}{2K} (\tau^{(i)}-\tau^{(i-1)} )\right\rfloor, \dots, \tau^{(i)}-\tau^{(i-1)} \right\} \eqdef \cI_{i}.
 \end{equation*}
@@
 \[
-    (b)  \leq \bP\left(\exists s \in \cI_k : s \, \kl\left(\widehat{\mu}_{a,s},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\widehat{\mu}_{b,\overline{r}},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
+    (b)  \leq \bP\left(\exists s \in \cI_{i} : s \, \kl\left(\widehat{\mu}_{a,s},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}} \right) +  \overline{r} \, \kl\left(\widehat{\mu}_{b,\overline{r}},\frac{s\widehat{\mu}_{a,s}+\overline{r}\widehat{\mu}_{b,\overline{r}}}{s+\overline{r}}  \right)  \leq \beta(s+\overline{r},\delta)\right),
 \]
-where we have also used that $\widetilde{\mu}^{k-1}_{i,n_k(\overline{t})} = \left(n_k(\tau^{(i)})\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)})} + \overline{r}\widetilde{\mu}^{k-1}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})} \right) / (n_k(\tau^{(i)}) + \overline{r})$.
-
-Using Pinsker's inequality, and introducing the gap $\Delta_k^{(i)} \eqdef \mu_k^{(i-1)} - {\mu}_k^{(i)}$ (which is such that $\Delta^{(i)} = |\Delta_k^k|$), one can write
+where we used the following property on means of $x=y+z$ observations $Z_i$:
+$\widetile{\mu}_{1:x} = \widetile{\mu}_{1:y+z} = \frac{1}{y+z} \sum_{i=1}^{y+z} Z_i = \frac{1}{y+z} \left( y \widetile{\mu}_{1:y} + z \widetile{\mu}_{y+1:z} \right)$,
+with $x = n_k(\overline{t}) = y + z = n_k(\tau^{(i)}) + \overline{r}$,
+in order to compute
+$\widetilde{\mu}^{(i-1)}_{i,n_k(\overline{t})} = \left(n_k(\tau^{(i)})\widetilde{\mu}^{(i-1)}_{i,n_k(\tau^{(i)})} + \overline{r}\widetilde{\mu}^{(i-1)}_{i,n_k(\tau^{(i)}) : n_k(\overline{t})} \right) / (n_k(\tau^{(i)}) + \overline{r})$.
+
+Using Pinsker's inequality, and introducing the gap $\Delta_k^{(i)} \eqdef \mu_k^{(i-1)} - {\mu}_k^{(i)}$ (which is such that $\Delta^{(i)} = |\Delta_k^{(i)}|$), one can write
 %
 \begin{align}
-    (b) & \leq \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
+    (b) & \leq \bP\left(\exists s \in \cI_{i} : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} \right)^2  \leq \beta(s+\overline{r},\delta)\right)\nonumber\\
     & \leq \bP\left(\exists s \in \N : \frac{2sr}{s+r}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,s} - \Delta_k^{(i)}\right)^2  \geq \beta(s+r,\delta)\right) \nonumber\\
-    & + \bP\left(\exists s \in \cI_k : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_k^{(i)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
+    & + \bP\left(\exists s \in \cI_{i} : \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}} - \Delta_k^{(i)}\right)^2  \leq \beta(s+\overline{r},\delta), \frac{2s\overline{r}}{s+\overline{r}}\left(\widehat{\mu}_{a,s} - \widehat{\mu}_{b,\overline{r}}\right)^2  \leq \beta(s+\overline{r},\delta)\right) \nonumber
 \end{align}
@@
 \begin{equation}\label{eq:6:FromHereGlobal}
-    (b) \leq \delta + \bP\left(\exists s \in \cI_k : \Delta^{(i)} \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).
+    (b) \leq \delta + \bP\left(\exists s \in \cI_{i} : \Delta^{(i)} \leq 2\sqrt{\frac{s+\overline{r}}{2s\overline{r}}\beta(s+\overline{r},\delta)}\right).
 \end{equation}
@@
 Define $s_{\min} \eqdef \left\lfloor \frac{\omega}{K} (\tau^{(i)}-\tau^{(i-1)})/2\right\rfloor$.
-Using that the mappings $s \mapsto (s+\overline{r})/s\overline{r}$ and $s \mapsto \beta(s + \overline{r},\delta)$ are respectively decreasing and increasing in $s$, one has, for all $s\in \cI_k$,
+Using the fact that the two mappings $s \mapsto (s+\overline{r})/s\overline{r}$ and $s \mapsto \beta(s + \overline{r},\delta)$ are respectively decreasing and increasing in $s$, one has, for all $s\in \cI_{i}$,
 \begin{eqnarray*}
@@
 which yields
-\[\forall s \in \cI_k, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta^{(i)}\right)^2.\]
+\[\forall s \in \cI_{i}, \ \ 2\frac{s+\overline{r}}{s\overline{r}}\beta\left(s+\overline{r}, \delta\right) \leq \left(\Delta^{(i)}\right)^2.\]
 %
@@
 
-In this section we report results of numerical simulations performed on synthetic data to compare the performance of \GLRklUCB{} against other state-of-the-art approaches, on some piece-wise stationary bandit problems.
+In this section we report results of numerical simulations performed on synthetic data, in order to compare the performance of \GLRklUCB{} against other state-of-the-art approaches, on some piece-wise stationary bandit problems.
 %
@@
 We present regret tables and regret plots,
-for which the regret was estimated using $1000$ independent runs.
+for which the mean regret was estimated using $1000$ independent runs.
 
@@
 
-We include in this study two algorithms designed for the classical MAB, \klUCB{} \cite{Garivier11KL}, and Thompson sampling \cite{AgrawalGoyal11,Kaufmann12Thompson},
-as well as an ``oracle'' version of \klUCB, that we call Oracle-Restart. This algorithm knows the exact locations of the break-points, and restarts \klUCB{} at those locations (without any delay).
-
-Then, we compare our algorithms to several competitor designed for a piece-wise stationary model.
+We include in this study two algorithms designed for the classical stationary MAB, \klUCB{} \cite{Garivier11KL}, and Thompson sampling \cite{AgrawalGoyal11,Kaufmann12Thompson},
+and an ``oracle'' version of \klUCB, that we call Oracle-Restart. This algorithm knows the exact locations of the break-points, and restarts \klUCB{} at those locations (without any delay).
+
+Then, we compare our algorithms to several competitors designed for a piece-wise stationary model.
 For a fair comparison, all algorithms that use  \UCB{} as a sub-routine were adapted to use \klUCB{} instead, which yields better performance\footnote{\cite{LiuLeeShroff17,CaoZhenKvetonXie18} both mention that extending their analysis to the use of \klUCB{} should not be too difficult.}.
@@
 We first include three \emph{passively adaptive algorithms}:
-Discounted \klUCB{} (D-\klUCB, \cite{Kocsis06}), with discount factor $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$; Sliding-Window \klUCB{} (SW-\klUCB,  \cite{Garivier11UCBDiscount}) using window-size $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$ and Discounted Thompson sampling (DTS, \cite{RajKalyani17}) with discount factor $\gamma = 0.95$.
+Discounted \klUCB{} (D-\klUCB, \cite{Kocsis06}), with discount factor $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, Sliding-Window \klUCB{} (SW-\klUCB,  \cite{Garivier11UCBDiscount}) using window-size $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$ and Discounted Thompson sampling (DTS, \cite{RajKalyani17}) with discount factor $\gamma = 0.95$.
 For this last algorithm, the discount factor $\gamma=0.75$ suggested by the authors was performing significantly worse on the problem instances we tried.
-More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).
+More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (problems $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (problems $3,5$).
 %
@@
 %
-Additionally, we include the \ExpThreeS{} algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \e) / ((\e-1)T)})$.
+Additionally, we include the \ExpThreeS{} algorithm from \cite{Auer02NonStochastic}, setting its parameters as in their Corollary~8.3: $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \e) / ((\e-1)T)})$, based on a prior knowledge of $T$ and $\Upsilon_T$.
 

### 2019-07-05 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Almost done for Ch7

@@
 $\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
-%
+
 The Generalized Likelihood Ratio statistic for this test is defined by
@@
     % Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
-    \vspace*{-30pt}% FIXME
+    % \vspace*{-30pt}% FIXME
 \end{smallproof}
@@
 we can write a first decomposition of the regret, first without the expectation defining the mean regret:
+\begin{small}
 \begin{align*}
@@
 \end{align*}
+\end{small}
 %
@@
 %
-Additionally, we include the \ExpThreeS{} algorithm from \cite{Auer02NonStochastic}, setting its parameters as in their Corollary~8.3: $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \e) / ((\e-1)T)})$, based on a prior knowledge of $T$ and $\Upsilon_T$.
+Additionally, we include the \ExpThreeS{} algorithm from \cite{Auer02NonStochastic}, setting its parameters as in the Corollary~8.3 of the paper: $\omega=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \e) / ((\e-1)T)})$, based on a prior knowledge of $T$ and $\Upsilon_T$.
 
@@
 by showing the times at which they detect a change and reinitialize their memory of observations of one or all the arms.
-We also display the number of observations (since the last restart) of the arm on which the change was detected.
-As shown below, in the experiments summary, the three change-point detection algorithm give similar performances in terms of regret, but they have different behaviors.
+% We also display the number of observations (since the last restart) of the arm on which the change was detected.
+As shown below, in the experiments summary, the three algorithms based on change-point detection obtain sublinear regret, but they have different behaviors.
 The test in \MUCB{} usually detects less changes, while the test in \CUSUMUCB{} usually detects more changes and even unnecessary changes.
 %
-On the examples presented below as well as large scale examples, all the tests appear to have no false alarm, and always a small detection delay for changes that are ``easy enough'' to be efficiently detected.
-The four algorithms are presented in the order of decreasing performance (\ie, the first one is empirically the most efficient, that is, it obtains a smaller regret in average).
+On the examples presented below as well as large scale examples, all the tests appear to have no (or a small number of) false alarm, and always a small detection delay for changes that are ``easy enough'' to be efficiently detected.
+% The four algorithms are presented in the order of decreasing performance (\ie, the first one is empirically the most efficient, that is, it obtains a smaller regret in average).
 
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
+    \includegraphics[width=0.96\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1.pdf}
     \caption{Locations of the detected change-points for four algorithms on Problem 1.}
@@
 
-Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1} above displays the locations of the successively detected change-points by four algorithms,
+Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1} displays the locations of the successively detected change-points by four algorithms,
 $1)$ \MklUCB,
@@
 Very small delays are possible only after having collected a lot of samples of the arm which changed, and for instance the same algorithm obtains a delay of $32$ samples from \textcolor{red}{arm $1$} for the next change at $t=3000$, because this arm was sampled less.
-On this problem, while \MklUCB{} detects the same changes as \GLRklUCB, with larger but comparable delays, it is seen to obtain a larger regret than the two \GLR{} variants, because the tuning of its forced exploration probability makes it explore suboptimal arms more often.
+On this problem, while \MklUCB{} detects the same changes as \GLRklUCB, with larger but comparable delays, it is seen to obtain a larger regret than the two \GLR{} variants, simply because the tuning of its forced exploration probability (\ie, $\omega$) makes it explore suboptimal arms more often.
 
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
+    \includegraphics[width=0.96\linewidth]{2-Chapters/6-Chapter/Images/Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2.pdf}
     \caption{Locations of the detected change-points for four algorithms on Problem 2.}
@@
 
-Like for the first problem, we illustrate the behavior of the same four algorithms in Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2} above.
+Like for the first problem, we illustrate the behavior of the same four algorithms in Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb2}.
 In this other problem, the difference between \MklUCB{} and \CUSUMklUCB{} is clear:
@@
 while the second one detects the changes that happen on the currently optimal arm and some other changes.
-Drawing conclusions on their behavior from a single simulation is tricky, as this example of behavior is counter intuitive:
+Drawing conclusions on their behavior from a single simulation is meaningless, as this example of behavior is counter intuitive:
 On the one hand, \CUSUM{} uses local restarts, and it should be less efficient than global restarts for this problem, as changes are all global.
-On the other hand, \MklUCB{} uses global restarts, but here it fails to detect any changes.
+On the other hand, \MklUCB{} uses global restarts, but here it fails to detect any change.
 %
@@
 But it is still interesting to compare the final regret in one run, as given above, with the mean regret for $1000$ independent runs, as given below in Table~\ref{table:6:totalResults1}.
-The values are not the same, but the ranking is: our proposal outperforms both \CUSUMklUCB{} and \MklUCB, and the test based on \CUSUM{} seems more efficient than the test based of \MklUCB{} in the problems at hand.
+The values change, but the ranking stays the same: our proposal outperforms both \CUSUMklUCB{} and \MklUCB, and the test based on \CUSUM{} seems more efficient than the test based of \MklUCB{} in the problems at hand.
 
@@
 
-We continue here the presentation of the other problems of our benchmark, see Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} above for the two problems $1$ and $2$.
+We continue here the presentation of the other problems of our benchmark, after Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} above for the two problems $1$ and $2$.
 
@@
 It is interesting to say that this problem is inspired from Figure~3 of \cite{CaoZhenKvetonXie18}, where the synthetic data was obtained from manipulations on a real-world database of clicks from \emph{Yahoo!}.
+%
+The details fall outside of the scope of this section, and are well explained in the paper \cite{CaoZhenKvetonXie18}.
 
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
     \caption{Problem $3$: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
@@
 but the stationary sequences between successive change-points no longer have the same length, as illustrated in Figure~\ref{fig:6:Problem_4}.
-Classical (stationary) algorithms such as \klUCB{} can be ``tricked'' by large enough stationary sequences, as they learn with a large confidence the optimal arm, and then fail to adapt to a new optimal arm after a change-point.
-We observe below in Table~\ref{table:6:totalResults2} that they can suffer higher regret when the change-points are more spaced out, as this problem starts with a longer stationary sequence of length $T/2$.
+Classical (stationary) algorithms such as \klUCB{} can be ``tricked'' by large enough stationary sequences, as they start by learning with a large confidence the optimal arm, and then fail to adapt to a new optimal arm after a change-point.
+We observe below in Table~\ref{table:6:totalResults2} that they can suffer higher regret when the change-points are more spaced out, like for instance for this problem starting with a longer stationary sequence of length $T/2$.
 
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
     \caption{Problem $4$: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
@@
 %
-It is much harder, with $\Upsilon=81$ change-points on $K=5$ arms for a longer horizon of $T=100000$.
+It is a much harder piece-wise stationary problem, with $\Upsilon=81$ change-points on $K=5$ arms for a longer horizon of $T=100000$.
 Some arms change at almost every time steps, for a total number of break-points $C=179$, but the optimal arm is almost always the same one (\textcolor{red}{arm $0$, with $\bullet$}).
@@
 as shown in Figure~\ref{fig:6:Problem_6}.
+% In practice, this problem is very hard and and
 
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
     \caption{Pb $5$: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
@@
 Similarly to what is presented in Chapter~\ref{chapter:5} in Table~\ref{table:5:comparisonUCB_klUCB}, we start by some experiments that justify the focus on the \klUCB{} index policy.
-The purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices for piece-wise stationary problems.
-%
-Consider two experiments on \textbf{problems 1} and \textbf{2}, with a horizon of $T=20000$ and $1000$ independent repetitions,
-for which we compare \CUSUMUCB{} against \CUSUMklUCB, \MUCB{} against \MklUCB, and the oracle policy using \UCB{} against \klUCB.
+Indeed, the purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices for piece-wise stationary problems.
+%
+We run two experiments on \textbf{problems 1} and \textbf{2}, with a horizon of $T=20000$ and $1000$ independent repetitions,
+for which we compare \UCB{} against \klUCB, for \CUSUM-, M-, \GLR{} and the oracle policy.
 We report in Table~\ref{table:6:comparisonUCB_klUCB} below the results (in terms of mean regret),
 and we observe that using \klUCB{} rather than \UCB{} indices always yields better practical performance.
-%
-Thus it is fair to compare our proposal against the modified versions of the two algorithms \CUSUM{} and $\mathrm{M}$- that use the \klUCB{} index policy instead of \UCB, as they perform uniformly better with \klUCB{} than with \UCB{} (the same tendency was observed on all other problems).
+
+Thus it is fair to compare our proposal against the modified versions of the two algorithms \CUSUM- and $\mathrm{M}$- that use the \klUCB{} index policy instead of \UCB, as they perform uniformly better with \klUCB{} than with \UCB{} (the same tendency was observed on all other problems).
 We also note that, without surprise, the oracle policy also performs (much) better with \klUCB{} than with \UCB.

### 2019-07-05 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Almost almost done for CH7

@@
 % -----------------------------------------------------------------
-\paragraph{Results.}
+\paragraph{Results on the entire benchmark.}
 \label{sub:6:NumericalResults}
@@
 We show below on Figure~\ref{fig:6:meanRegretPb1} and additional in Appendix~\ref{app:6:moreFigures} the simulation results for the five problems.
-The results in terms of mean regret $R_T$ are given above in Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2}, but it is also interesting to observe two plots for each experiment.
-%
-First, we show the mean regret as a function of time (\ie, $R_t$ for $1 \leq t \leq T$), for $9$ of the considered algorithms (as they are outperformed by the others, \ExpThreeS{} and Discounted-\klUCB{} are not included to avoid clutter).
-%
-Efficient stationary algorithms, like TS and \klUCB, typically suffer a linear regret after a change on the optimal arm changes if they had ``too'' many samples before the change-points (\eg, on Figure~\ref{fig:6:meanRegretPb1} and even more on Figures~\ref{fig:6:meanRegretPb4} and \ref{fig:6:meanRegretPb5}).
+The results of Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2} focus on the value of $R_T$ at the end of each bandit game, but it is also interesting to observe two plots for each experiment.
+%
+First, we show the mean regret as a function of time (\ie, $R_t$ for $t \in [T]$), for $9$ of the considered algorithms (as they are outperformed by the others, \ExpThreeS{} and Discounted-\klUCB{} are not included in order to avoid clutter).
+%
+Efficient stationary algorithms, like TS and \klUCB, typically suffer a linear regret after a change on the optimal arm, if they had ``too'' many samples before the change-points (\eg, on Figure~\ref{fig:6:meanRegretPb1} and even more on Figures~\ref{fig:6:meanRegretPb4} and \ref{fig:6:meanRegretPb5}).
 This illustrates the conjecture that classical algorithms can suffer linear regret even on simple piece-wise stationary problems.
@@
 
-On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there are more changes, we clearly observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much better than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
+On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there are more changes, we observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much better than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
 % Our approach, with the two options of \textbf{Local} or \textbf{Global} restarts, performs very closely to the oracle for problem $4$.
@@
 
-We also show the empirical distribution of the regret $R_T$, on Figure~\ref{fig:6:histogramRegretPb1}.
-It shows that all algorithms have a rather small variance on their regret, except Thompson Sampling which has a large tail due to its large mean regret on this (easy) non-stationary problems.
+We also show the empirical distribution of the regret $R_T$, on Figure~\ref{fig:6:histogramRegretPb1} below.
+It shows that algorithms efficient in terms of regret also have a small variance on their regret.
+% , except Thompson Sampling which has a large tail due to its large mean regret on this (easy) non-stationary problems.
 

### 2019-07-05 1-Preamble/5-Abstract/abstract.tex
Longer intro, I need to cut in §1.1 and §1.2
- ❤ Je t'aime chaton ❤

@@
 Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de type radio intelligente, dans des bandes non licenciées et sans supervision centrale.
-Nous considérons des réseaux actuels ou futures de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
+Nous considérons notamment des réseaux actuels ou futures de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
 %  (Chapitre~\ref{chapter:1}).
@@
 % Chapitre 2
-Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (MAB), mais leur analyse est difficile à réaliser,
-% , Chapitre~\ref{chapitre:2}
-car il est délicat de prouver la convergence d'un grand nombre d'appareils jouant à un jeu collaboratif sans communication ni aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (MAB), mais l'analyse de la convergence d'un grand nombre d'appareils jouant à un jeu collaboratif sans communication ni aucune coordination reste délicate,
+lorsque les appareils suivent tous un modèle d'activation aléatoire.
 % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.

### 2019-07-05 1-Preamble/5-Abstract/resume_fr.tex
First changes from Emilie's LONG comments
Longer intro, I need to cut in §1.1 and §1.2
- ❤ Je t'aime chaton ❤

@@
 % Chapitre 1
-Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
-Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil (Chapitre~\ref{chapter:1}).
+Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio sans fil de manière intelligente, dans des bandes non licenciées et sans supervision centrale.
+Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IdO), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil (Chapitre~\ref{chapter:1}).
 % Chapitre 4
-Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement (Chapitre~\ref{chapter:4}).
+Nous proposons différents modèles de réseaux IdO, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement (Chapitre~\ref{chapter:4}).
 % Chapitre 2
-Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (MAB, Chapitre~\ref{chapter:2}), mais leur analyse est difficile à réaliser,
+Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (MAB, Chapitre~\ref{chapter:2}), mais leur analyse est complexe,
 % , Chapitre~\ref{chapter:2}
-car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+notamment si l'on veut prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
 % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
 % Chapitres 5 et 6
-Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires (Chapitre~\ref{chapter:5}), puis des bandits mono-joueurs non stationnaires (Chapitre~\ref{chapter:6}).
+Le reste de ce manuscrit étudie donc deux modèles restreints, afin de pouvoir en faire une analyse théorique exhaustive.
+Nous considérons d'abord des bandits multi-joueurs dans des problèmes stationnaires (Chapitre~\ref{chapter:5}), puis des bandits mono-joueurs non stationnaires (Chapitre~\ref{chapter:6}).
 % Chapitre 3
 Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres (Chapitre~\ref{chapter:3}).
+
+% ----------------------------------------------------------------------------
+\section*{Contexte}
+
+FIXME 0.5 à 2 pages de contexte, historique, technique, humain, etc.
 
@@
 Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
-Nous signalons le chapitre correspondant, ainsi que les conférences nationales ou internationales auxquelles ont été publié certains résultats.
+Elles se situent à trois niveaux : des modèles MAB pour plusieurs contextes radio liés à la radio intelligente ; des preuves théoriques associées aux problèmes d'apprentissages correspond ; et enfin des implémentations d'algorithmes à but de simulations, de partage avec la communauté, et de validation par des démonstrations radio réalistes.
+
+Nous signalons le chapitre, ainsi que les conférences nationales ou internationales dans lesquelles ont été publiés les résultats correspondants.
 
@@
     Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
-    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences et simulations numériques utilisées dans cette thèse.
+    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences et simulations numériques utilisées tout au long de thèse.
 
     \item
-    Nous présentons le problème du choix de l'algorithme qu'un-e ingénieur-e devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithmes MAB disponibles, dans le Chapitre~\ref{chapter:25}.
+    Nous présentons le problème du choix de l'algorithme qu'un-e ingénieur-e devrait utiliser, ou de la sélection d'algorithme parmi la riche collection des différents algorithmes de bandits dans la litérature, dans le Chapitre~\ref{chapter:25}.
     Nous présentons un algorithme, appelé \Aggr, pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par des simulations numériques qu'il peut atteindre de bonnes performances empiriques
@@
     \item
-    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les appareils dotés de capacités de radio intelligente peuvent implémenter de leur côté des algorithmes MAB, pour augmenter automatiquement la durée de vie de leur batterie.
-Cela permet à davantage de périphériques d'utiliser le même réseau tout en maintenant une haute qualité de service
+    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les appareils dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes MAB, pour augmenter automatiquement la durée de vie de leur batterie.
+    Cela permet également à davantage d'objets d'utiliser le même réseau tout en maintenant un taux élevé d'accès au spectre sans souffrir de collisions radio
     (publiés à CROWNCOM 2017, démonstration ICT 2018, WCNC 2019 et MOTIoN 2019)
@@
     \item
-    Nous avons implémenté une démonstration réaliste du modèle susmentionné \cite{Besson2018ICT,Besson2019WCNC}, présentée à ICT 2018, et nous la détaillons en Section~\ref{sec:4:gnuradio}.
+    Nous avons réalisé une démonstration réaliste du modèle proposé ci-dessus \cite{Besson2018ICT,Besson2019WCNC}, présentée à ICT 2018, et nous la détaillons en Section~\ref{sec:4:gnuradio}.
 Nous avons aussi réalisé une courte vidéo présentant notre démonstration, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
@@
     \item
-    La dernière contribution de cette thèse est notre article \cite{Besson2018DoublingTricks}, résumé en Annexe~\ref{app:2:DoublingTricks}.
-    Nous donnes une revue de la littérature sur l'utilisation de la technique de doublements successifs de l'horizon (``doubling trick'') pour les problèmes de bandits,
+    La dernière contribution de cette thèse est notre article \cite{Besson2018DoublingTricks}.
+    % , résumé en Annexe~\ref{app:2:DoublingTricks}.
+    Cet article donne une revue de la littérature sur l'utilisation de la technique de doublements successifs de l'horizon (``doubling trick'') pour les problèmes de bandits,
     ainsi qu'une analyse unifiée et plus générique de deux familles de doublements.
@@
 
-Ce manuscrit est organisé comme suit.
+% Ce manuscrit est organisé comme suit.
+%
+L'ordre de lecture du manuscrit peut suivre n'importe quel chemin, entre l'introduction donnée dans le Chapitre~\ref{chapter:1}, et la conclusion générale qui constitue le dernier Chapitre~\ref{chapter:conclusion}.
+Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,
+la thèse est organisée en deux parties, correspondant aux deux lignes intermédiaires de la figure suivante.
 
@@
         \node[align=center] (chapter25) at (2.5,2.25) [rectangle,draw,fill=red!20] {\textbf{Chapitre~\ref{chapter:25}}\\Sélection séquentielle\\d'algorithme MAB};
-        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles MAB\\pour les réseaux IoT};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles MAB\\pour les réseaux IdO};
         \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!15] {\textbf{Chapitre~\ref{chapter:5}}\\Modèle MAB\\Multi-joueurs};
@@
 
-L'ordre de lecture du manuscrit peut être n'importe quel chemin, entre l'introduction donnée dans Chapitre~\ref{chapter:1}, et la conclusion générale qui constitue le dernier Chapitre~\ref{chapter:conclusion}.
-Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,
-la thèse est organisée en deux parties :
-
 \begin{itemize}
@@
 Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
-Le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
-Nous terminons cette première partie par le Chapitre~\ref{chapter:25}, qui n'est pas non plus nécessaire à la compréhension globale du manuscrit, mais qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
+Le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, qui est utilisées par les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} pour leurs simulations numériques.
+%  lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
+Nous terminons cette première partie par le Chapitre~\ref{chapter:25},
+% qui n'est pas non plus nécessaire à la compréhension globale du manuscrit, mais
+qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
 
@@
 Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop complexes pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
-Pour cette raison, nous affaiblissons les modèles pour le reste du document,
-et les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
+Pour cette raison, nous simplifions les modèles précédents dans la suite du document,
+afin d'établir des preuves mathématiques garantissant la convergence, ainsi que les gains en performance apportés par ces algorithmes de bandit.
+Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
+Ces deux études ont chacune donné des résultats théoriques à la pointe de la recherche sur les modèles de bandits multi-joueurs et de bandits stationnaires par morceaux, que nous avons aussi validé par des expériences numériques.
+
 \end{itemize}

### 2019-07-05 2-Chapters/1-Chapter/chapter1.tex
Longer intro, I need to cut in §1.1 and §1.2
- ❤ Je t'aime chaton ❤

@@
 % https://en.wikipedia.org/wiki/History_of_radio
-During the last $150$ years, mankind has developed many different communication technologies, and since the late $1890$s, wireless communications between man-made manufactured devices have been made possible, and more and more frequent in our lives.
+In the last $150$ years, mankind has developed many different communication technologies, and since the late $1890$s, wireless communications between manufactured devices have been made possible, and more and more frequent in our lives.
 With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, for a large range of different applications.
@@
 Such IoT technologies should be able to adapt automatically to different environments and application contexts, and be as efficient as possible.
-Instead of devoting a lot of time, energy and funding on Research and Development to design many different efficient radio access schemes, for all possible cases, a cheap and promising direction is to use Machine Learning to (try to) attain the same performance gain automatically.
+%
+In addition to the usual Research and Development effort to design many different efficient radio access schemes, for all the possible cases,
+now the time has come to combine it with cheap and promising Machine Learning in order to (try to) attain the level of performance gain necessary for the IoT promises to become reality.
 
@@
 Our exploration of the large number of MAB algorithms and models developed in the recent literature
-has given us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explores the performances of newly proposed algorithms.
+has given us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explore the performances of newly proposed algorithms.
 To tackle this goal, we produced a Python library of simulation of MAB problems.
@@
 We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
-A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.
+A full documentation is available online, as well as exhaustive instructions to reproduce the simulations discussed in the rest of this thesis.
 
@@
 %
-Because there are so many different MAB algorithms, we are also interested by the question of how a practitioner can chose the one she will implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
+Because there are so many different MAB algorithms, we are also interested by the question of how a practitioner can chose the algorithm to implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
 To answer this question, we present two approaches.
@@
 We study two models of IoT networks in Chapter~\ref{chapter:4}.
-The first model considers IoT devices that simply have data to send to a base station, at a regular frequency (but random times), and that use acknowledgements as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the channel less occupied by the surrounding traffic).
+The first model considers IoT devices that simply have data to regularly send to a base station (at unpredictable or random times).
+Such devices use acknowledgements as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the channel less occupied by the surrounding traffic).
 The goal of this first model is to increase the Quality of Service (QoS) of the application of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmission.
@@
 %
-Our second model then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increased QoS for the entire network.
-
-We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of thousands (or more) of end-devices and a single base station.
-We consider an ALOHA-based protocol, slotted in both time and frequency, by assuming a simple pre-agreed synchronization between all devices and the base station.
+Our second model then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increased QoS for the entire network, as less retransmissions reduce the local spectrum load.
+
+% We present in Section~\ref{sec:4:firstModel} a first model of an IoT network, that is made of thousands (or more) of end-devices and a single base station.
+% We consider an ALOHA-based protocol, slotted in both time and frequency, by assuming a simple pre-agreed synchronization between all devices and the base station.
 This synchronization is the only aspect of the network that requires central control from the base station, and it only requires one downlink packet when a new device joins the network operated by this base station \cite{Centenaro16}.
-%
-Our IoT network model contains two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
-Static devices form an interfering traffic, which could be generated by devices using other standards as well.
-Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choice of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) received from the base station.
-%
-We first detail three base-lines, a naive approach and two centralized ``oracle'' approaches,
-that are used to evaluate the performance of the considered MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communications rate.
-We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from the other devices becomes more and more fluctuating and unpredictable.
-%
-Even without changing anything on the level of the considered IoT standard (neither the PHY nor the MAC layers), our proposal is just an add-on capability, that can be set-up on each device, on a unit-per-unit basis.
-This model constitutes the first article written during this thesis, and initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
-We presented this first work at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, where it obtained the ``best paper award'' \cite{Bonnefoi17}.
-
-Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations was a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
-In Section~\ref{sec:4:firstModel}, we detail what we believe to be the first PoC of using decentralized MAB learning in a simplified IoT-like wireless network.
-We implemented a network with a base station, which receives incoming messages from different end-devices and replies with acknowledgements.
-% , that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
-% Verify this first model empirically: proof of concept
-This is a second result of a fruitful collaboration with R.
-Bonnefoi, and this work was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
-
-We then consider an extension of our first model, following the idea of packet retransmissions, that is at the root of the famous ALOHA protocol \cite{Abramson1970,Roberts75}.
-The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode in order to wait for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, up-to $m \leq 10$ times), after some small random waiting times.
-%
-We propose to also use decision making, that is to also use RL and MAB algorithms, for this other model, at the second layer of the MAC protocol, in order to learn the best channels to use for retransmitting the packets that suffered from a collision at their first transmissions.
-We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations.
-This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well that the simplest of the different heuristics we compared is efficient, and allows the devices to learn to behave almost optimally.
-This third contribution concluded our collaboration with R. Bonnefoi, and was presented at the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
+% %
+% Our IoT network model contains two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions.
+% Static devices form an interfering traffic, which could be generated by devices using other standards as well.
+% Dynamic devices can use RL, and especially MAB algorithms, to optimize their sequential choice of channels, in order to maximize the number of successful up-link transmissions, thanks to the success/failure feedback that is the presence/absence of acknowledgement (\Ack) received from the base station.
+% %
+% We first detail three base-lines, a naive approach and two centralized ``oracle'' approaches,
+% that are used to evaluate the performance of the considered MAB algorithms (\UCB{} and TS), in terms of the centralized system-wide successful communications rate.
+% We show that our approach is interesting, as these low-cost MAB algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from the other devices becomes more and more fluctuating and unpredictable.
+% %
+% Even without changing anything on the level of the considered IoT standard (neither the PHY nor the MAC layers), our proposal is just an add-on capability, that can be set-up on each device, on a unit-per-unit basis.
+The first model constitutes the first article written during this thesis, and initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
+We presented it at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, where it obtained the ``best paper award'' \cite{Bonnefoi17}.
+%
+% Verifying the potential gain of performance brought by these MAB algorithms in numerical simulations was a first step, but it is satisfactory to also verify our results on a real-world proof-of-concept (PoC).
+% In Section~\ref{sec:4:firstModel}, we detail what we believe to be the first PoC of using decentralized MAB learning in a simplified IoT-like wireless network.
+% We implemented a network with a base station, which receives incoming messages from different end-devices and replies with acknowledgements.
+% % , that are used as feedback by the devices and let them learn an efficient spectrum access scheme.
+% % Verify this first model empirically: proof of concept
+The proof-of-concept (PoC) continued our fruitful collaboration with R.
+Bonnefoi, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
+%
+% We then consider an extension of our first model, following the idea of packet retransmissions, that is at the root of the famous ALOHA protocol \cite{Abramson1970,Roberts75}.
+% The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode in order to wait for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, up-to $m \leq 10$ times), after some small random waiting times.
+% %
+% We propose to also use decision making, that is to also use RL and MAB algorithms, for this other model, at the second layer of the MAC protocol, in order to learn the best channels to use for retransmitting the packets that suffered from a collision at their first transmissions.
+% We present the model in Section~\ref{sec:4:retransmissions}, along with mathematical justifications of the potential of using this second-stage MAB learning for retransmissions, and numerical illustrations.
+% This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well that the simplest of the different heuristics we compared is efficient, and allows the devices to learn to behave almost optimally.
+Our second model, with retransmissions, corresponds to our last collaboration with R.
+Bonnefoi, which was presented at the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
 
@@
 
-Our work on multi-player bandit was presented at the ALT conference in Lanzarote, Spain, in April $2019$ \cite{Besson2018ALT},
-%
-and since then, some works have studied extensions of our framework or alternative algorithms, as a few independent teams started to work on different questions shortly after its publication.
-%
-Since last year, it was proven that the ``no sensing'' case (\ie, IoT) is essentially not harder that the ``sensing'' (\ie, OSA) case, from a theoretical point-of-view, by E.
+% Our work on multi-player bandit was presented at the ALT conference in Lanzarote, Spain, in April $2019$ \cite{Besson2018ALT},
+% %
+% and since then, some works have studied extensions of our framework or alternative algorithms, as a few independent teams started to work on different questions shortly after its publication.
+% %
+% Since last year, it was proven that the ``no sensing'' case (\ie, IoT) is essentially not harder that the ``sensing'' (\ie, OSA) case, from a theoretical point-of-view, by E.
 Boursier and V.
 Perchet \cite{BoursierPerchet18}.
-Even if their approach is costly empirically, the result is surprising and very interesting.
-They showed that even under the strong hypothesis of no direct communication between players and no central coordination (in the same model as the one of our paper), the players can actually use the collision feedback (\ie, receiving or not an \Ack) as an indirect way to communicate bits of information with each other.
-This idea requires that devices coordinate at least a little bit, by an initial pre-agreement (which is simply the fact that the IoT devices run the same algorithms).
-As they showed, the players can start by a ``musical chair'' phase, to reach an orthogonal allocation, and then they can alternate exploration, communication and exploitation phases until the end of the game at horizon $T$.
-Because they all follow the same algorithm, the devices are able to split the time steps between ``listening'' and ``talking'' phases, and the authors denote it the ``communication trick''.
-The communication trick is used by each player to send and receive observations about all arms from the other players, using one forced collision to send one bit at a time.
-They use about $T-\bigO{\log(T)}$ rounds in exploitation of the best arms, and only $\bigO{\log(T)}$ rounds in exploration and communication, yielding a logarithmic regret.
-% As they showed, after a long enough initialization period, the $M$ players have elected a ``master'' player and $M-1$ ``followers'' players (with high probability),
-Another direction, as followed by \cite{Bistritz18}, is to use this communication trick and reduce the problem to a centralized multiple-play (MP) MAB, which can be solved efficiently with the MP generalization of \klUCB{} \cite{Luedtke16} or Thompson sampling \cite{Komiyama15}.
-
-We discuss more in details this interesting follow-up of our article \cite{BoursierPerchet18}, along with other recent research works such as \cite{LugosiMehrabian18}, in Section~\ref{sec:5:literatureReviewOtherModels}.
-%
-Moreover, the case of different means of arms for the different players was studied in \cite{Bistritz18,KaufmannAbbas19}, and this generalization of our model is very interested for CR applications, as the mean of an arm is the average quality (\eg, availability) of a radio channel, and as such it has no reason to be uniform among end-devices (\ie, players).
-%
-Finally the adversarial case of multi-player bandit has started to be studied, independently by \cite{bande2019adversarial} and \cite{AlaturLevyKrause19}.
+% Even if their approach is costly empirically, the result is surprising and very interesting.
+% They showed that even under the strong hypothesis of no direct communication between players and no central coordination (in the same model as the one of our paper), the players can actually use the collision feedback (\ie, receiving or not an \Ack) as an indirect way to communicate bits of information with each other.
+% This idea requires that devices coordinate at least a little bit, by an initial pre-agreement (which is simply the fact that the IoT devices run the same algorithms).
+% As they showed, the players can start by a ``musical chair'' phase, to reach an orthogonal allocation, and then they can alternate exploration, communication and exploitation phases until the end of the game at horizon $T$.
+% Because they all follow the same algorithm, the devices are able to split the time steps between ``listening'' and ``talking'' phases, and the authors denote it the ``communication trick''.
+% The communication trick is used by each player to send and receive observations about all arms from the other players, using one forced collision to send one bit at a time.
+% They use about $T-\bigO{\log(T)}$ rounds in exploitation of the best arms, and only $\bigO{\log(T)}$ rounds in exploration and communication, yielding a logarithmic regret.
+% % As they showed, after a long enough initialization period, the $M$ players have elected a ``master'' player and $M-1$ ``followers'' players (with high probability),
+% Another direction, as followed by \cite{Bistritz18}, is to use this communication trick and reduce the problem to a centralized multiple-play (MP) MAB, which can be solved efficiently with the MP generalization of \klUCB{} \cite{Luedtke16} or Thompson sampling \cite{Komiyama15}.
+
+% We discuss more in details this interesting follow-up of our article \cite{BoursierPerchet18}, along with other recent research works such as \cite{LugosiMehrabian18}, in Section~\ref{sec:5:literatureReviewOtherModels}.
+% %
+% Moreover, the case of different means of arms for the different players was studied in \cite{Bistritz18,KaufmannAbbas19}, and this generalization of our model is very interested for CR applications, as the mean of an arm is the average quality (\eg, availability) of a radio channel, and as such it has no reason to be uniform among end-devices (\ie, players).
+% %
+% Finally the adversarial case of multi-player bandit has started to be studied, independently by \cite{bande2019adversarial} and \cite{AlaturLevyKrause19}.
 
@@
 
-% % ----------------------------------------------------------------------------
-% \subsection{Summary of the contributions}
-% \label{sec:1:summaryOfContributions}
-
-% To summarize the main contributions of this thesis, we can list the following points:
-
-% \begin{itemize}
-%     % \item
-%     % We present the concepts and the notations of the multi-armed bandit problem in Chapter~\ref{chapter:2}, from a mathematical point-of-view.
-%     % But we also follow a didactic approach as we use an online interactive demonstration designed to let anyone play against a small bandit problem from his/her browser, in Section~\ref{par:2:interactiveDemoDiscoverMAB}.
-
-%     % \item
-%     % We give a short literature review of stochastic bandit algorithms in Section~\ref{sec:2:famousMABalgorithms}.
-
+% ----------------------------------------------------------------------------
+\subsection{Summary of the contributions}
+\label{sec:1:summaryOfContributions}
+
+We can list the following points to summarize the main contributions of this thesis.
+
+\begin{itemize}
     % \item
-%     We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence (MIT) \cite{SMPyBanditsJMLR,SMPyBandits}.
-%     We present it in details in Chapter~\ref{chapter:3}.
+    % We present the concepts and the notations of the multi-armed bandit problem in Chapter~\ref{chapter:2}, from a mathematical point-of-view.
+    % But we also follow a didactic approach as we use an online interactive demonstration designed to let anyone play against a small bandit problem from his/her browser, in Section~\ref{par:2:interactiveDemoDiscoverMAB}.
 
     % \item
-%     % We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
-%     We present in Chapter~\ref{chapter:25} an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
-%     \cite{Besson2018WCNC}.
-%     % with \textbf{Aggregator} (WCNC 2018)
+    % We give a short literature review of stochastic bandit algorithms in Section~\ref{sec:2:famousMABalgorithms}.
+
+    \item
+    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence (MIT) \cite{SMPyBanditsJMLR,SMPyBandits}.
+    We present it in details in Chapter~\ref{chapter:3}.
+
+    \item
+    % We present the problem of choosing which algorithm a practitioner should use, or algorithm selection from the rich collection of different MAB available, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
+    We present in Chapter~\ref{chapter:25} an algorithm called \Aggr{} for aggregation of algorithms as an online solution to the algorithm selection problem, and numerical simulations to illustrate that it achieves state-of-the-art empirical performances
+    \cite{Besson2018WCNC}.
+    % with \textbf{Aggregator} (WCNC 2018)
+
+    \item
+    We propose different models for IoT networks, in Chapter~\ref{chapter:4}, where end-devices with cognitive radio capabilities can implement MAB algorithms on their side, to automatically increase their battery life and allow more devices to use the same network while maintaining a high Quality of Service
+    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019,MoyBesson2019Annales}.
+    % (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
+
+    \item
+    We implemented a proof-of-concept of the aforementioned model \cite{Besson2018ICT}, and we present it in Section~\ref{sec:4:gnuradio},
+    % We made a video showcasing our demonstration, hosted at
+    and in a video, hosted at
+    \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
 
     % \item
-%     We propose different models for IoT networks, in Chapter~\ref{chapter:4}, where end-devices with cognitive radio capabilities can implement MAB algorithms on their side, to automatically increase their battery life and allow more devices to use the same network while maintaining a high Quality of Service
-%     \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019}.
-%     % (CROWNCOM 2017, ICT demo 2018, WCNC 2019 and MOTIoN 2019)
+    % The source code for the two previously mentioned contributions are all published online, along with clear instructions for reproducing our work.
+
+    \item
+    We present three variants of the multi-players bandit model in Chapter~\ref{chapter:5}.
+    For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
+    as well as extensive numerical experiments to demonstrate its excellent performance in comparison with the rest of the literature.
+    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works heavily built up on our results.
 
     % \item
-%     We implemented a proof-of-concept of the aforementioned model \cite{Besson2018ICT}, and we present it in details in Section~\ref{sec:4:gnuradio}.
-We made a video showcasing our demonstration, hosted at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
-
-%     % \item
-%     % The source code for the two previously mentioned contributions are all published online, along with clear instructions for reproducing our work.
-
-%     \item
-%     We present the multi-players bandit model, for which we introduce three variants, in Chapter~\ref{chapter:5}.
-%     For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
-%     as well as extensive numerical experiments to demonstrate its excellent performance in comparison with the rest of the literature.
-%     Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works heavily built up on our results.
-
-%     % \item
-%     % We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
-%     % % were  (ALT 2018, and 4-8 works inspired by our article since then).
+    % We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
+    % % were  (ALT 2018, and 4-8 works inspired by our article since then).
 State-of-the-art with our algorithm MCTopM + klUCB for multi-players bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"
 
-%     \item
-%     We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non-stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
-%     Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
-%     % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
-
-%     \item
-%     The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
-%     and a unified and more generic analysis of two families of doubling tricks.
-%     This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
-% \end{itemize}
+    \item
+    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non-stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
+    Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
+    % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
+
+    \item
+    The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
+    and a unified and more generic analysis of two families of doubling tricks.
+    This lead to the article \cite{Besson2018DoublingTricks}.
+    % which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
+\end{itemize}
 
@@
 The reading order of the manuscript can be any top-down path between the Introduction in the current Chapter~\ref{chapter:1}, and the Conclusion in the last Chapter~\ref{chapter:conclusion}.
-As the graph in Figure~\ref{fig:1:organization} shows, the thesis is organized in two parts.
+The thesis is organized in two parts, corresponding to the two intermediate lines of the following Figure~\ref{fig:1:organization}.
 
@@
     \centering
-    \resizebox{0.95\textwidth}{!}{
+    \resizebox{1.00\textwidth}{!}{
     \begin{tikzpicture}[>=latex',line join=bevel,scale=2.25]
@@
 Conversely, even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use numerical simulations based on our simulation library SMPyBandits, the Chapter~\ref{chapter:3} where we present it is not required to understand them.
-We conclude this part with Chapter~\ref{chapter:25}, which is also not mandatory for the rest of this thesis, and which details one of the first contributions of this thesis, a new algorithm for online MAB algorithms selection.
+We conclude this part with Chapter~\ref{chapter:25}, where we detail one of the first contributions of this thesis, a new algorithm for online MAB algorithms selection.
 

### 2019-07-05 2-Chapters/1-Chapter/chapter1.tex
Added a reference

@@
 %
-Historically, three main families of wireless communication systems have been deployed: first, centralized broadcasting (\eg, old TV or radio), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized broadcasting for the Internet of Things (IoT) networks (\eg, sensor networks).
-
-% Central broadcasting
-Starting from FM radio music and news, then old black-and-white TV and color TV, the first systems were purely made for centralized broadcasting of information.
-In a large area, one well located antenna emits continuously in a fixed Radio Frequency (RF) band, and many devices can listen to this RF traffic and decode it, for instance to listen to some music.
-Different standards have been developed, and some of them are still in use today, and their nature requires a central authority that allocates RF bands to different usages (\eg, TV or FM radio channels etc).
-For example, in the city of Cesson-Sévigné where my campus of CentraleSupélec is located, a high RF tower emits on many standards, and in particular the \SI{92.3}{\mega\hertz} band is allocated to a French classical music radio (\emph{Radio Classique}).
-Broadcast-based systems are not only used for TV or radio:
-two other famous and worldwide deployed examples of mono-directional radio systems are the long-range global navigation satellite systems (such as USA's GPS or European Union's Galileo), and short-range infrared remote controllers.
-% https://en.wikipedia.org/wiki/Satellite_navigation
-
-% Two ways communications
-However, this kind of mono-directional wireless communications were soon found to not be satisfactory for many applications, and thus bi-directional standards were defined from the $1910$s.
-First developed by and for the military and the navy (\eg, the Titanic sent the first ``SOS'' signal over radio in $1912$),
-such applications have reached general public usage since the $1950$s.
-The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people's lives in the richer countries in the world, with the booming of personal cellphones since the $2000$s and smartphones since the $2010$s.
-In most systems, there is still a base station (with one or more antenna) in charge of a large area (or cell), and many devices are able to receive and transmit data to the base station.
-Nowadays, the most well-known systems include WiFi, and systems deployed for mobile phone communications, with a long list of standards from 1G, 2G, 3G, 4G and now 5G, still in development in $2019$.
-Even if both systems initially differed in usage, as 1G was designed to exchange only voice, and WiFi to link wireless devices at home to an Internet-connected box, they are conceptually very close.
-As the systems are cellular, any device knows the base station it is associated with, and the bi-directional nature of their exchanges makes a centralized control possible.
-In order to optimize the efficiency of the network, the base station is equipped with decision making algorithms that affects the devices in the cell to the available resources (time slots, RF bands, transmission power etc).
-To be able to communicate with the base station without interfering with each others, such affectation should be orthogonal, and led to systems based on Time Division Multiple Access (TDMA) and Frequency Division Multiple Access (FDMA).
-%  and Non-Orthogonal techniques only started to emerge recently (\eg, NOMA).
-Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance, of the wireless communications to and from the devices.
-More recently, ``green radio'' has emerged as a solution to the tradeoff between maximizing the QoS of the system, and minimizing the power consumption of the base station and the devices it serves.
-
+Historically, three main families of wireless communication systems have been deployed in massive commercial networks: first, centralized broadcasting (\eg, radio or TV broadcasting), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized data harvesting for the Internet of Things (IoT) networks (\eg, sensor networks).
+%
+% % Central broadcasting
+% Starting from FM radio music and news, then old black-and-white TV and color TV, the first systems were purely made for centralized broadcasting of information.
+In a large area, one well located antenna emits continuously in a fixed Radio Frequency (RF) band, and many devices can listen to this RF signal and demodulate it, for instance to listen to some music.
+% Different standards have been developed, and some of them are still in use today, and their nature requires a central authority that allocates RF bands to different usages (\eg, TV or FM radio channels etc).
+% For example, in the city of Cesson-Sévigné where the campus of CentraleSupélec is located, a high RF tower emits on many standards, and in particular the \SI{92.3}{\mega\hertz} band is allocated to a French classical music radio (\emph{Radio Classique}).
+% Broadcast-based systems are not only used for TV or radio:
+% another famous and worldwide deployed example of mono-directional radio systems are the long-range global navigation satellite systems (such as USA's GPS or European Union's Galileo).
+% %  and short-range infrared remote controllers.
+% % https://en.wikipedia.org/wiki/Satellite_navigation
+%
+% % Two ways communications
+% However, this kind of mono-directional wireless communications were soon found to not be satisfactory for many applications, and thus bi-directional standards were defined from the $1910$s.
+% First developed by and for the military and the navy (\eg, the Titanic sent the first ``SOS'' signal over radio in $1912$),
+% such applications have reached general public usage since the $1950$s.
+% The greatest achievement of this long-running R\&D domain was the avent of mobile telephony in the $1990$s, which has continued to be more and more present in people's lives in all over the world, with the booming of personal cellphones since the $2000$s and smartphones since the $2010$s.
+% In most systems, there is still a base station (with one or more antenna) in charge of a large area (or cell), and many devices are able to receive and transmit data to the base station.
+% Nowadays, the most well-known systems include WiFi, and systems deployed for mobile phone communications, with a long list of standards from 2G, 3G, 4G and now 5G, still in development in $2019$.
+% Even if both systems initially differed in usage, as 2G was designed to exchange only voice, and WiFi to link wireless devices at home to an Internet-connected box, they are conceptually very close.
+% As the systems are cellular, any device knows the base station it is associated with, and the bi-directional nature of their exchanges makes a centralized control possible.
+% In order to optimize the efficiency of the network, the base station is equipped with decision making algorithms that affects the devices in the cell to the available resources (time slots, RF bands, transmission power etc).
+% To be able to communicate with the base station without interfering with each others, such affectation should be orthogonal, and led to systems based on Time (TDMA), Frequency (FDMA) and Code Division Multiple Access (CDMA).
+% %  and Non-Orthogonal techniques only started to emerge recently (\eg, NOMA).
+% Traditionally, the base station aimed only at maximizing the Quality of Service (QoS), that englobes the data rate, latency, availability and other measures of performance, of the wireless communications to and from the devices.
+% More recently, ``green radio'' has emerged as a solution to the tradeoff between maximizing the QoS of the system, and minimizing the power consumption of the base station and the devices it serves.
+%
+%
 % The opposite: decentralized broadcasting, and the age of IoT
-Finally, a third kind of systems are of the opposite kind, and could be designed as decentralized broadcasting: one antenna is still in charge of many devices, but such devices can only send up-link packets, and the only down-link data they can receive are short acknowledgements, sent from the base-station to indicate success or failure of every up-link packet.
+% Finally, a third kind of systems are of the opposite kind,
+This third kind of systems can be designed as decentralized:
+even if a central base station is
+% one antenna is still
+still in charge of many devices,
+the devices initiate the sending of up-link packets, and the only down-link data they can receive are short acknowledgements, sent from the base-station to indicate success or failure of every up-link packet.
 This family of wireless systems are referred to as Internet of Things (IoT),
-and a typical example of application of such IoT networks is for sensor networks.
+and a typical example of application of such IoT networks is for sensor wireless networks.
 
@@
 
-This third kind of wireless systems are characterized by their decentralized nature: due to the really limited information that the base station can send to the devices (in downlink), it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
+This third kind of wireless systems are characterized by their decentralized nature,
+where communications are initialized and regulated by the devices, not a centralized control systems
+% due to the really limited information that the base station can send to the devices (in downlink), it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
 Indeed, a central control requires signaling packets that have been identified as too heavy for these kinds of systems.
@@
 in most kinds of frequency bands, the entire RF spectrum is now allocated and free bands no longer exist, and this limits the possibility of adding any new usage.
-As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below, only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz}, while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz}, is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication, \eg, in the Titanic), to space research, inter-satellite, mobile telephony and many other applications.
+As illustrated in Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} below,
+% only a very small portion of the RF spectrum in the USA is not yet allocated, from $0$ to \SI{9}{\kilo\hertz},
+% while the rest of the displayed portion of the spectrum, from \SI{9}{\kilo\hertz} to \SI{300}{\giga\hertz},
+almost the entire spectrum
+is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication, \eg, in the Titanic), to space research, inter-satellite, mobile telephony and many other applications.
 %
 % Traditional regulatory structures have been built for an analog model and are not optimized for cognitive radio.
-Regulatory bodies in the world, like the \href{https://www.fcc.gov/}{Federal Communications Commission} in the USA (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
-or \href{http://www.tdf.fr/}{TDF} in France (see \href{https://www.TDF.fr/}{\texttt{TDF.fr}}), as well as different independent measurement campaigns, found that most radio frequency spectrum are inefficiently utilized,
-meaning that while a band can be allocated to a certain unique usage, it can be free from any user in certain times and/or places.
+Regulatory bodies in the world, like the
+\href{https://www.itu.int/en/Pages/default.aspx}{International Telecommunication Union} (see \href{https://www.itu.int/}{\texttt{www.ITU.int}}),
+the \href{https://www.fcc.gov/}{Federal Communications Commission} in the USA (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
+or the \href{https://cept.org}{European Conference of Postal and Telecommunications Administrations} in Europe (see \href{https://www.CEPT.org/}{\texttt{CEPT.org}}),
+% or \href{http://www.tdf.fr/}{TDF} in France (see \href{https://www.TDF.fr/}{\texttt{TDF.fr}}),
+as well as different independent measurement campaigns, found that most radio frequency spectrum are inefficiently utilized.
+This means that while a band can be allocated to a certain unique usage, it can be free from any user in certain times and/or places.
 We refer to \cite{patil2011survey} for a survey on the worldwide spectrum utilization, and to \cite{valenta2010survey} for the situation in Europe.
@@
     \caption[A chart representing the allocation of radio spectrum in the USA in $2016$]{A chart representing the allocation of radio spectrum in the USA in $2016$.
-Only a small share of the spectrum is not yet allocated, the top-left corner in white and indicated by \textcolor{red}{\textbf{Free} in red} (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz}).
 \copyright{} United States Department of Commerce, \href{https://www.ntia.doc.gov/files/ntia/publications/january_2016_spectrum_wall_chart.pdf}{\texttt{www.ntia.doc.gov/files/ntia/publications/january\_2016\_spectrum\_wall\_chart.pdf}}.}
+    % Only a small share of the spectrum is not yet allocated, the top-left corner in white and indicated by \textcolor{red}{\textbf{Free} in red} (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz})
     % https://commons.wikimedia.org/wiki/File:United_States_Frequency_Allocations_Chart_2016_-_The_Radio_Spectrum.pdf
@@
 
-Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are insufficiently utilized.
-Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place.
-Moreover, the fixed spectrum allocation prevents rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
-Therefore, in the last $15$ years, regulatory bodies in the world have been considering whether to allow unlicensed users in licensed bands, if they would not cause any interference to the (paying) licensed users.
-These initiatives have initiated the research on \textbf{Cognitive Radio},
-and especially on a field that we detail below, \textbf{Dynamic Spectrum Access (DSA)}.
+Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are less utilized.
+Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place, as shown by \cite{Lopez2009spectral}.
+Moreover, the fixed spectrum allocation prevents
+the introduction of new services, especially for low-cost or small markets equipments.
+% rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
+Therefore, in the last $15$ years, thanks to an active lobbying by the cognitive radio community,
+regulatory bodies in the world have been considering whether to allow
+a new wireless communication paradigm:
+enable unlicensed users in licensed bands, if they would not cause any interference to the (paying) licensed users.
+These initiatives are considered by the \textbf{Cognitive Radio},
+field that we detail below, and especially for \textbf{Dynamic Spectrum Access (DSA)}.
 
@@
 
-In this thesis, we study Internet of Things networks, and the two constraints of low power consumption (or long battery life) and low duty cycle are essential.
+In this thesis, we study long-range Internet of Things networks, characterized by mainly three essential constraints:
+low power consumption (or long battery life),
+long-range
+and low duty cycle.
+%
 More precisely, we study the possible interconnections between \textbf{Cognitive Radio} and \textbf{Machine Learning} applied for IoT networks.
@@
 We can narrow down the first field of study of this thesis, steps by steps:
-from Information Technologies (IT), to Telecommunications, to Wireless Communications, then to Software Defined Radio (SDR),
-and finally to Cognitive Radio (CR).
+from Information Technologies (IT), to Telecommunications, to Wireless Communications, then to Cognitive Radio (CR),
+and finally to Software Defined Radio (SDR).
 %
@@
 % https://en.wikipedia.org/wiki/Software-defined_radio
-A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, amplifiers, modulators/demodulators, detectors, etc) are all instead implemented by means of software on a personal computer or an embedded system, with the exception of a unique pair of receiver/transmitter antennas.
-Even though the SDR paradigm was initiated by the USA defense research, over the past $20$ years the industry has started to be interested by SDR, and CR have got significant attention from the academia and industry as well.
+A SDR is a radio communication system where components that have been traditionally implemented in hardware (\eg, mixers, filters, modulators/demodulators, detectors, etc) are more and more implemented by means of software on a processor.
+%  with the exception of a unique pair of receiver/transmitter antennas.
+Even though the SDR paradigm was initiated by the USA defense research in the $1980$s, over the past $20$ years the civil industry has started to be interested by SDR, and CR has got significant attention from the academia and industry as well.
 % CR has been the first answer to solve spectrum scarcity issue which collected a huge attention in early 21st century.
 %
-As CR is not a standard technology, it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, in the last $20$ years.
+As CR is not a standard technology, it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, that just appeared $20$ years ago.
 % - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
@@
     \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
-    He was the first to propose to use CR for Dynamic Spectrum Access.
+    % He was the first to propose to use CR for Dynamic Spectrum Access.
 
@@
     \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
-    by \textbf{Lilian Besson}, active development between October $2016$ and March $2019$,
+    by \textbf{Lilian Besson}, active development since October $2016$,
     \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
-    It currently consists in about $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
-    and a complete documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} and \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.
+    It currently consists in about $45000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
+    and a complete documentation accessible on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} or \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.
     See Chapter~\ref{chapter:3}.

### 2019-07-05 2-Chapters/1-Chapter/chapter1.tex
Small typos, detected with proselint **/*.tex

@@
 even if a central base station is
-% one antenna is still
 still in charge of many devices,
@@
 These initiatives are considered by the \textbf{Cognitive Radio},
-field that we detail below, and especially for \textbf{Dynamic Spectrum Access (DSA)}.
+field that we detail below, and especially for \textbf{Dynamic Spectrum Access (DSA)},
+for which we refer to the surveys \cite{akyildiz2006next,garhwal2012survey} for more details.
 
@@
 % explain PU and SU in a few sentences
+One possible way of considering spectrum flexibility is the following.
+%
 In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the network.
-As exposed in the definition above, the CR philosophy suggests that if the network is not used by any PU in a certain location and time, then non-paying users, referred to as Secondary User (SU), could use it.
-Regulations state that PU have a strict priority over any other SU,
-but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU is equipped with an efficient sensing capacity, it can analyze its environment, and use a licensed band if it is free of any PU.
+As exposed in the definition above of the CR philosophy, we could imagine that if the network is not used by any PU in a certain location and time, then non licensed users, referred to as Secondary User (SU), could use it, by paying some rent to the PU.
+%
+Regulation states that PU have a strict priority,
+but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU was equipped with an efficient sensing capacity, it could analyze its environment, and use a licensed band if it is free of any PU.
 % explain DSA
-This defines the concept of DSA, proposed in the early $2000$s, and we refer to the surveys \cite{akyildiz2006next,Zhao07,garhwal2012survey} for more details.
-After a short paragraph on Reinforcement Learning, we then focus on one example, the \textbf{Opportunistic Spectrum Access} (OSA) problem.
+This defines the concept of \textbf{Opportunistic Spectrum Access} (OSA), for which we refer to the survey \cite{Zhao07} for more details.
+% After a short paragraph on Reinforcement Learning, we then focus on the OSA problem.
 
@@
 %
-Similarly, we can narrow down the second field of study of this thesis, steps by steps.
+% Similarly, we can narrow down the second field of study of this thesis, steps by steps.
 % Stats > sequential learning > bandits
-A first point-of-view is the historical one, as Multi-Armed Bandits (MAB) learning emerged as an interesting problem in the statistics community and among researchers interested by sequential learning, as shown by the earlier works of \cite{Thompson33,Robbins52,LaiRobbins85}.
+% A first point-of-view is the historical one, as
+We are interested by Multi-Armed Bandits (MAB) learning,
+which emerged as an interesting problem in the statistics community and among researchers interested by sequential learning, as shown by the earlier works of \cite{Thompson33,Robbins52,LaiRobbins85}.
+It is also included in the more general concept of Reinforcement Learning (RL), itself one of the field of Machine Learning (ML).
 % ML > RL > MAB
-We can also expose another point-of-view, which starts from the general concept of Machine Learning (ML), and focus on Reinforcement Learning (RL), then on MAB learning.
+% We can also expose another point-of-view, which starts from the general concept of Machine Learning (ML), and focus on , then on MAB learning.
 % As illustrated in Figure~\ref{fig:1:MainTypesOfMachineLearning}, there are other types of ML, that we do not focus on in this thesis.
@@
 %
-In OSA, the focus is on one SU accessing a licensed spectrum, occupied by PU that have a strict priority over the SU, which has to follow a ``listen-before-talk'' access scheme.
+The two communities of OSA and MAB have started to interact, and the resulting works have received a great interest from both communities, since the late $2000$s and early $2010$s, with pioneers works like \cite{Liu08,Zhao10,Jouini09,Jouini10}.
+The focus is on one SU accessing a licensed spectrum, occupied by PU that have a strict priority over the SU, which has to follow a ``listen-before-talk'' access scheme.
 %
@@
 the SU is equipped with a sensing capability,
-and we consider a fixed and finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
+and considers a fixed and finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
 For instance, it can be a set of three Wi-Fi channels at different frequency, emitted by the same Wi-Fi station.
-Another hypothesis is that both PU and SU are synchronized in times, by sub-dividing the time in discrete time steps (or rounds).
-%
-Thus if the SU spends a short time in the beginning of each time slot to perform sensing, it can scan for the presence or absence of any PU before transmitting.
+Another hypothesis is that both PU and SU are synchronized in time, by sub-dividing the time in discrete time steps (or rounds).
+%
+Thus if the SU spends a short time in the beginning of each time slot to perform sensing, it can scan for the presence or absence of any PU before transmitting, during the rest of the time slot without colliding with the PU.
 If the SU was able to sense for all the $K$ frequency bands, it could simply transmit in one of the free channels if any, or not transmit if all channels are used at a given time step.
-However, sensing is known to be error-prone and costly, especially for wide-band sensing, as detailed in the surveys \cite{yucek2009survey,subhedar2011spectrum}, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
-This assumption, along with the regulation that PU cannot be disturbed, enforces the SU to transmit in the channel that it sensed, in each time step, if and only if it was sensed free from any PU.
+However, sensing is known to be costly, in terms of energy consumption, especially for wide-band sensing, as detailed in the surveys \cite{yucek2009survey,subhedar2011spectrum}, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
+This assumption, along with the hypothesis that PU cannot be disturbed (without which no regulation will never be accepted for OSA), enforces the SU to transmit in the channel that it sensed, in each time step, if and only if it was sensed free from any PU.
 % non negligible cost of radio reconfiguration
 
-Focussing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of $K\geq2$ channels), then it performs sensing, and finally it transmits in this channel if it was sensed free.
+Focussing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of channels, $[K]=\{1,\dots,K\}$), then it performs sensing, and finally it transmits in this channel if it was sensed free.
 The goal of the SU is to minimize its energy consumption (we remind that we focus on ``green'' cognitive radio) and to maximize its uplink data rate, or equivalently, to maximize its number of successful transmission.
@@
 in a sequential action-then-feedback cycle (time steps are $t\in\N^*$),
-under partial information feedback.
+under partial information feedback (\ie, about one channel over $K$ at each time step).
 These three hypotheses are the ones that restrict from the general ML framework to the specific Multi-Armed Bandit case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
@@
 and the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$ is modeled by a reward of $r(t) \in \{0,1\}$.
-Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus an uplink message can be sent), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message should be sent.
+Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus a SU message can be sent), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no SU message should be sent, until the end of that slot.
 %
@@
 Their works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the $\UCB_1$ algorithm,
-along the early works of Q. Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
-%
-Shortly after in $2014$, proof-of-concepts using real-world radio hardware and Software Defined Radio were developed by C. Moy and his student C. Robert \cite{RobertSDR2014,MoyWSR2014}.
+along with the early works of Q. Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
+%
+Shortly after in $2014$, proof-of-concepts using real-world radio hardware and Software Defined Radio were developed by C.
+Moy and his student C.
+Robert \cite{RobertSDR2014,MoyWSR2014}, using USRP boards and the MATLAB/Simulink software.
 In a second PhD thesis \cite{Modi17PhD}, N. Modi studied from $2014$ to $2017$ the impact of using MAB algorithms to optimize channel selections, on the battery life of a wireless device.
-On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
+On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and could bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
 On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfigurations for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
@@
 %
-Since $2017$, S.
-Darak, and his team at IIIT Delhi in India, have actively worked in the research on cognitive radio using multi-armed bandits, and some of their recent works are illustrated with real-world demo using USRP and the MATLAB/Simulink system
+Since $2017$, S. Darak, and his team at IIIT Delhi in India, have actively worked in the research on cognitive radio using multi-armed bandits.
+Inheriting from his work with the SCEE team,
+some of their recent works are also illustrated with real-world demo using USRP and the MATLAB/Simulink system
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
@@
 The aforementioned literature have essentially shown that MAB algorithms can be applied with success for the OSA problem.
-But if we consider CR-ready devices that cannot perform sensing, such as low-cost and low energy-consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
+However, if we consider CR-ready devices that cannot perform sensing, such as low-cost and low energy-consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
 Moreover, in most cases, the IoT networks use unlicensed bands, and as such there are no longer a distinction between PU and SU.
+In other words, there is no longer any priority between users in IoT networks.
+%  (no more PU, all are SU).
 %
@@
     \item
-    most IoT networks run in unlicensed bands (no more distinction between PU and SU),
-    \item
-    most IoT devices are very low-cost devices, with low material and computational capacities,
-    meaning that they are not equipped with any sensing capabilities.
-    However, they are equipped with a transceiver as well as with a receiver,
-    and with low-complexity storage and computation capacities,
+    we consider here IoT networks that run in unlicensed bands (no more distinction between PU and SU),
+    \item
+    most IoT devices are very low-cost devices, possibly running on low-quality hardware, and having limited software and power capabilities (\ie, tiny batteries).
+    It means that they even if they are not equipped with particular sensing capabilities,
+    they are equipped with a transmitter as well as with a receiver (Tx and Rx, \ie, a transceiver).
+    and they have low-complexity storage and computation capabilities (\ie, small embedded processors),
     \item
@@
     \item
-    IoT devices have low duty cycles (only a few messages every minute) to very low duty cycles (only a few messages every day).
+    IoT devices have low duty cycles (only a few messages every minute) to very low duty cycles (only a few messages every month).
 \end{itemize}
-\TODOL{FIXME more specificities? Est-ce que je dois mentionner d'autres contraintes }
+
+A central coordination of the devices with their base station is impossible as it would require to communicate continuously with the base station, which violates the two last constraints.
+%
+Moreover, due to both their limited hardware (Rx/Tx), limited computational power and limited battery life, the IoT devices are able to use their received antenna to sense one channel occasionally (typically, for a few time slots after every single up-link message), but they cannot perform spectrum sensing like it is considers for OSA and CR (\ie, at the beginning of a time slot, in a ``listen-before-talk'' scheme).
 
 One could think that in the absence of sensing, Reinforcement Learning is no longer possible, but any real IoT device still receives some information about its environment after some or every transmissions.
-In most IoT standards an up-link message sent to a base station is usually followed by a down-link message sent back by the base station, to indicate if the up-link message was successfully received and understood.
+In most IoT standards an up-link message sent to a base station can be followed by a down-link message sent back by the base station, to indicate if the up-link message was successfully received and understood.
 %

### 2019-07-05 2-Chapters/2-Chapter/chapter2.tex
Lots of improvements on the arXiv papers
Small typos, detected with proselint **/*.tex

@@
 % , and we follow closely the formalism of \cite{Kaufmann12PhD}, as well as \cite{Slivkins2019,LattimoreBanditAlgorithmsBook,Bubeck12}.
-and a summary is included in the Nomenclature, in the Appendix in Part~\ref{part:Conclusion} (Page~\pageref{chapter:nomenclature}).
+and a summary is included in the Nomenclature, in the Appendix (Page~\pageref{chapter:nomenclature}).
 
@@
 Indeed, in single-player models it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
-But even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
+Even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
 %
@@
 we can consider a sequence $(\varepsilon_t)_{t\in\N^*}$ decreasing with time $t$, for instance $\varepsilon_t = \varepsilon_0 / t$ with $\varepsilon_0 = 6 K / d^2$, for a constant $0 < d < \min_{k: \Delta_k > 0} \Delta_k$.
-Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$.
-But this is not satisfactory, as $\min_k \Delta_k$ needs to be known in advance, which is usually not the case for real applications.
+Then it was shown in \cite{Auer02} that the regret of $\varepsilon$-greedy is of the order of $K \log(T) / d + \smallO{T}$, which leads to an order-optimal regret of $\bigO{K \log(T)}$,
+but this is not satisfactory, as $\min_k \Delta_k$ needs to be known in advance, which is usually not the case for real applications.
 Here again, an average of $T_0 \leq \varepsilon_0 \log(T)$ steps are spent on exploration.

### 2019-07-05 2-Chapters/3-Chapter/chapter3.tex
Small typos, detected with proselint **/*.tex

@@
     It was also proven to obtain good ``best of both world'' performances.
-    Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
+    Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{~Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
 Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.

### 2019-07-05 2-Chapters/5-Chapter/chapter5.tex
Small typos, detected with proselint **/*.tex

@@
 for Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis.
-We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, and we also experimented with larger values of $T$ and found a similar linear behavior.}.
+We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{~For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, and we also experimented with larger values of $T$ and found a similar linear behavior.}.
 

### 2019-07-05 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Small typos, detected with proselint **/*.tex
Typo on one label/reference.

@@
 Instead, it uses geometrically decreasing magnitudes $d_k = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only considers a few different values of $d_k$.
-% Our  algorithmAdSwitch(shown  as  Algorithm  1)  proceeds  in  episodes,  where  a  newepisode starts when the algorithm detects a change in one of the arms.
-Each episode startswith an estimation phase (Step 2), in which both arms are chosen alternatingly until theirmeans can be distinguished (Step 3).
-After the estimation phase,  the algorithm exploitsthe empirical best arm (Step 5).
-However,  in order to detect changes,  at each step withsome probability (Step 4) an exploration phase is started that checks whether a change hasoccurred (comparing the resulting empirical means to that obtained from the estimationphase).
-During such exploration phases the algorithm checks for changes of certain magni-tudes (denoted bydi).
-Note that our algorithm does not require knowledge of the numberof changesL.
-The following theorem provides an upper bound on the expected regret ofour algorithm
+% Our algorithmAdSwitch(shown as Algorithm 1) proceeds in episodes, where a new episode starts when the algorithm detects a change in one of the arms.
+Each episode starts with an estimation phase (Step 2), in which both arms are chosen alternatingly until their means can be distinguished (Step 3).
+After the estimation phase, the algorithm exploits the empirical best arm (Step 5).
+However, in order to detect changes, at each step with some probability (Step 4) an exploration phase is started that checks whether a change has occurred (comparing the resulting empirical means to that obtained from the estimation phase).
+During such exploration phases the algorithm checks for changes of certain magnitudes (denoted by di).
+Note that our algorithm does not require knowledge of the number of changesL.
+The following theorem provides an upper bound on the expected regret of our algorithm
 They prove that their algorithm achieves a regret bound of $\cO(\log(T) \sqrt{\Upsilon_T T})$, without a prior knowledge of $\Upsilon_T$.
@@
 where we used the following property on means of $x=y+z$ observations $Z_i$:
-$\widetile{\mu}_{1:x} = \widetile{\mu}_{1:y+z} = \frac{1}{y+z} \sum_{i=1}^{y+z} Z_i = \frac{1}{y+z} \left( y \widetile{\mu}_{1:y} + z \widetile{\mu}_{y+1:z} \right)$,
+$\widetilde{\mu}_{1:x} = \widetilde{\mu}_{1:y+z} = \frac{1}{y+z} \sum_{i=1}^{y+z} Z_i = \frac{1}{y+z} \left( y \widetilde{\mu}_{1:y} + z \widetilde{\mu}_{y+1:z} \right)$,
 with $x = n_k(\overline{t}) = y + z = n_k(\tau^{(i)}) + \overline{r}$,

### 2019-07-05 2-Chapters/Conclusion/conclusion.tex
Small typos, detected with proselint **/*.tex
Typo on one label/reference.

@@
 We started by presenting the multi-armed bandit model, with a finite number of arms and real-valued rewards.
-Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
+Our focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
 The notations used in the manuscript were introduced by showcasing an interactive demonstration.
@@
 One approach to obtain a unified ``master'' algorithm could be to use expert aggregation on the different state-of-the-art algorithms presented in this thesis, and as each one is efficient in a different setting, the resulting aggregated algorithm could also be efficient in each of the different settings.
-But as we illustrated in Section~\ref{sub:25:theory}, this approach does not seem theoretically promising.
+But as we illustrated in Section~\ref{sec:25:theory}, this approach does not seem theoretically promising.
 

### 2019-07-05 1-Preamble/5-Abstract/resume_fr.tex
Small modifs

@@
 
-\TODOL{Je ne sais pas quelle longueur ni quel niveau de détail il faut viser ici.
-Est-ce que ces 3/4 pages suffisent ?
-Est-ce qu'il faut donner des maths en français pour CHAQUE CHAPITRE ?
-J'ai peur de ne pas réussir à écrire 20 pages avec des maths en français juste pour pouvoir citer tel théorème ou tel résultat.}
+% ----------------------------------------------------------------------------
 
-\TODOL{En fait, maintenant que le ch1 est écrit, je peux traduire (et couper un peu) ce chapitre 1, et ça ira !}
-
-% ----------------------------------------------------------------------------
+Ce manuscrit conclut ma thèse de doctorat, qui a débuté en octobre $2016$ et s'est terminée en novembre $2019$.
+Mes recherches se sont déroulées au laboratoire de l'IETR à Rennes (France), dans l'équipe SCEE hébergée sur le campus de Rennes de l'école d'ingénieurs CentraleSupélec.
+J'étais supervisé par le professeur Christophe Moy, à Rennes,
+et j'étais également co-encadré par le Docteur Emilie Kaufmann, que j'ai visité plusieurs fois à l'Inria Lille Nord Europe à Lille (France).
 
@@
 FIXME 0.5 à 2 pages de contexte, historique, technique, humain, etc.
+
+% Problème principal
+%
+La racine des problèmes qui motivent cette thèse sont les questions du réchauffement climatique et de l'augmentation de la population mondiale.
+% https://en.wikipedia.org/wiki/History_of_radio
+Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, les communications sans fil entre les appareils fabriqués ont été rendues possibles, et de plus en plus fréquentes dans nos vies.
+Avec l'avènement des réseaux Internet des objets (IdO), des milliards de dispositifs autonomes basse consommation devraient être déployés dans le monde entier, pour un large éventail d'applications différentes.
+Il existe aujourd'hui un consensus mondial sur le fait qu'avec la tendance actuelle à l'accroissement démographique et la crise énergétique actuelle, toute nouvelle technologie déployée doit être à la fois bon marché et efficace sur le plan énergétique,
+ainsi qu'adaptés pour desservir un grand nombre de personnes et d'appareils.
+%
+Ces technologies de l'IdO devraient pouvoir s'adapter automatiquement à différents environnements et contextes d'application, et être aussi efficaces que possible.
+%
+En plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, pour tous les cas possibles,
+le moment est venu de le combiner avec un apprentissage machine bon marché et prometteur pour (essayer d') atteindre le niveau de gain de performance nécessaire pour que les promesses de l'IdO deviennent réalité.
+
+C'est pourquoi nous nous intéressons à cette thèse sur
+l'amélioration de la durée de vie des batteries des équipements terminaux IoT et la réduction du coût énergétique des réseaux IoT.
+Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant la prise de décision décentralisée à faible coût directement dans les futurs dispositifs IdO.
+%
+Notre thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (Multi-Armed Bandit algorithms), afin de permettre aux appareils IoT d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
+
+\paragraph{Des anciens téléviseurs aux normes de l'IdO.}
+%
+Historiquement, trois grandes familles de systèmes de communication sans fil ont été déployées dans les grands réseaux commerciaux : d'abord, la radiodiffusion centralisée (\eg, radio ou télédiffusion), puis les systèmes bidirectionnels centralisés (\eg, 4G ou WiFi), et aujourd'hui la collecte décentralisée de données pour les réseaux Internet des objets (\eg, réseaux capteurs).
+%
+Ce troisième type de systèmes peut être conçu comme décentralisé :
+même si une station de base centrale est
+toujours en charge de nombreux appareils,
+les appareils déclenchent l'envoi de paquets de liaison montante, et les seules données de liaison descendante qu'ils peuvent recevoir sont de courts accusés de réception, envoyés par la station de base pour indiquer le succès ou l'échec de chaque paquet de liaison montante.
+Cette famille de systèmes sans fil est appelée Internet des objets (IoT),
+et un exemple typique d'application de tels réseaux IoT est celui des réseaux sans fil de capteurs.
+
+Pour le développement futur des ``réseaux intelligents'', des ``villes intelligentes'', des ``maisons intelligentes'' ou de l'''agriculture intelligente'', les réseaux de capteurs devraient être largement déployés.
+Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les ``bâtiments connectés'' et l'''agriculture connectée''.
+Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides et d'utiliser des réseaux de capteurs pour obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, et de permettre au contrôle centralisé de la chaleur d'optimiser son coût et sa consommation énergétique.
+Pour l'agriculture, un exemple peut être d'équiper chaque vache (dans les grandes exploitations) d'un capteur qui émet régulièrement des informations biologiques, telles que la température corporelle ou le niveau de stress, afin d'optimiser le temps de traite, de surveiller la santé des animaux, etc.
+
+Ce troisième type de systèmes sans fil se caractérise par leur nature décentralisée,
+où les communications sont initialisées et régulées par les appareils, et non par un système de contrôle centralisé.
+%
+En effet, une commande centrale nécessite la signalisation de paquets qui ont été identifiés comme trop lourds pour ce type de systèmes.
+Dans les réseaux IdO actuels et futurs, de nombreux dispositifs de nature hétérogène utilisent la même antenne pour des applications différentes.
+Un problème commun est que ces appareils IoT ont une forte contrainte sur leur consommation d'énergie, car la plupart d'entre eux seront déployés sans accès direct à l'alimentation et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
+Par exemple, la plupart des entreprises commerciales d'IdO vendent aujourd'hui une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
+Une autre contrainte commune aux dispositifs IoT est leur faible rapport cyclique, car la plupart des applications ciblent un ou quelques messages à envoyer chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
+%
+De nombreuses normes différentes pour les réseaux IdO ont été proposées ces dernières années,
+et ils consistent en une spécification à la fois pour le \emph{PHY}sical
+et le calque \emph{M}edium \emph{A}access \emph{C}ontrol (\emph{MAC}).
+Pour citer quelques exemples de normes, ZigBee, Z-Wave ou Bluetooth visent les communications à courte portée, tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à longue portée.
+Nous nous référons à l'enquête \cite{Centenaro16} pour plus de détails, et des références dans \cite{Azari18} ou notre récent article \cite{MoyBesson2019}.
+
+\paragraph{Pénurie du spectre radio.}
+%
+Un problème majeur des technologies sans fil actuelles est la question de la pénurie du spectre radio :
+dans la plupart des bandes de fréquences, la totalité du spectre RF est désormais attribuée et les bandes libres n'existent plus, ce qui limite la possibilité d'ajouter tout nouvel usage.
+Tel qu'illustré dans la Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} ci-dessous,
+presque tout le spectre
+est affecté à divers usages, qui vont de la radionavigation maritime (historiquement le premier usage des radiocommunications, \eg, dans le Titanic), à la recherche spatiale, l'inter-satellite, la téléphonie mobile et de nombreuses autres applications.
+%
+Les organismes de réglementation dans le monde, comme le
+\href{https://www.itu.int/en/Pages/default.aspx}{International Telecommunication Union} (voir \href{https://www.itu.int/}{\texttt{www.ITU.int}}),
+la \href{https://www.fcc.gov/}{Federal Communications Commission} aux États-Unis (voir \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
+ou la \href{https://cept.org}{ Conférence européenne des administrations des postes et des télécommunications} en Europe (voir \href{https://www.CEPT.org/}{\texttt{CEPT.org}}),
+ainsi que différentes campagnes de mesure indépendantes, ont montré que la plupart des fréquences du spectre des fréquences radioélectriques sont utilisées de manière inefficace.
+Cela signifie qu'une bande peut être attribuée à un certain usage unique, mais qu'elle peut être libre de tout utilisateur à certains moments et/ou endroits.
+Nous nous référons à \cite{patil2011survey} pour une étude sur l'utilisation mondiale du spectre, et à \cite{valenta2010survey} pour la situation en Europe.
+
+Les bandes des réseaux cellulaires sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radioamateurs) sont moins utilisées.
+Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend fortement du moment et du lieu, comme le montre \cite{Lopez2009spectral}.
+En outre, l'attribution fixe du spectre empêche
+l'introduction de nouveaux services, en particulier pour les équipements à bas prix ou pour les petits marchés.
+%
+Donc, au cours des $15$ dernières années, grâce à un lobbying actif de la communauté de la radio intelligente,
+les organismes de réglementation dans le monde se sont demandé s'il fallait permettre à l
+un nouveau paradigme de communication sans fil :
+permettre aux utilisateurs non titulaires d'une licence d'utiliser des bandes sous licence s'ils ne causent pas d'interférence aux utilisateurs (payants) titulaires d'une licence.
+Ces initiatives sont examinées par le \textbf{Radio Cognitive},
+que nous détaillons ci-dessous, et en particulier pour \textbf{Dynamic Spectrum Access (DSA)},
+pour lesquelles nous renvoyons aux enquêtes \cite{akyildiz2006next,garhwal2012survey} pour plus de détails.
+
+\section*{Radio Intelligente et Bandits Multi-Bras}
+
+Dans cette thèse, nous étudions les réseaux Internet des Objets à longue portée, caractérisés principalement par trois contraintes essentielles :
+faible consommation d'énergie (ou longue durée de vie des piles),
+à long terme
+et un cycle d'utilisation réduit.
+%
+Plus précisément, nous étudions les interconnexions possibles entre \textbf{Radio Intelligente} et \textbf{l'apprentissage statistique} appliquées aux réseaux IdO.
+Définissons et détaillons les deux concepts.
+
+\paragraph{Des TIC à la Radio Intelligente (CR).}
+%
+Nous pouvons affiner le premier champ d'étude de cette thèse, étape par étape :
+des technologies de l'information et communication (TIC), aux télécommunications, aux communications sans fil, puis à la radio intelligente (RC),
+et enfin à la radio logicielle (Software Defined Radio, SDR).
+%
+La transition de l'approche historique de la radio matérielle aux architectures SDR est un processus graduel, qui a commencé au début des années 1990 et s'est accéléré dans les années 2000.
+% https://en.wikipedia.org/wiki/Software-defined_radio
+Un SDR est un système de radiocommunication où les composants qui ont été traditionnellement implémentés dans le matériel (\eg, mélangeurs, filtres, modulateurs/démodulateurs, détecteurs, etc) sont de plus en plus implémentés au moyen de logiciels sur un processeur.
+%
+Même si le paradigme de la RRL a été initié par la recherche américaine en matière de défense dans les années 1980, au cours des 20 dernières années, l'industrie civile a commencé à s'intéresser à la RRL, et la RRL a également suscité beaucoup d'intérêt de la part du milieu universitaire et de l'industrie.
+%
+Comme la RC n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RC, qui vient de paraître il y a $20$ il y a quelques années.
+%
+\begin{itemize}
+    \item
+    J. Mittola, en $1999$, a proposé que
+    \emph{``une radio vraiment intelligente qui serait autonome, sensible aux RF et à l'utilisateur, et qui inclurait la technologie logicielle et les capacités d'apprentissage machine ainsi qu'une grande connaissance de l'environnement radio haute-fidélité''} \cite{Mitola99}.
+
+    \item
+    Puis S. Haykin en $2005$ a aussi dit que
+    \emph{``une radio intelligente (CR) est un système de communication sans fil intelligent qui est capable de connaître son environnement, d'apprendre et d'adapter ses paramètres de fonctionnement (puissance d'émission et fréquence porteuse) à la volée dans le but de fournir une communication fiable à tout moment, en tout lieu et efficace sur le plan spectral''} \cite{Haykin05}.
+
+    \item
+    La \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia encyclopedia} dit ceci
+    \emph{``une CR est une radio qui peut être programmée et configurée dynamiquement pour utiliser les meilleurs canaux sans fil à proximité afin d'éviter les interférences et la congestion des utilisateurs.
+Une telle radio détecte automatiquement les canaux disponibles dans le spectre sans fil, puis modifie en conséquence ses paramètres de transmission ou de réception pour permettre plus de communications sans fil simultanées dans une bande de spectre donnée à un endroit donné''}.
+\end{itemize}
+
+L'une des façons possibles d'envisager la flexibilité du spectre est la suivante.
+%
+Dans les bandes sous licence, il y a des utilisateurs primaires (PU) qui paient pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
+Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est pas utilisé par une UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (UD), pourraient l'utiliser en payant un loyer à l'UP.
+%
+La réglementation stipule que le PU a une priorité stricte,
+mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si une UD était équipée d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si elle est exempte de toute PU.
+Ceci définit le concept de \textbf{Opportunistic Spectrum Access} (OSA), pour lequel nous nous référons à l'enquête \cite{Zhao07} pour plus de détails.
+
+\paragraph{De la statistique ou de l'apprentissage machine aux bandits multi-bras.}
+%
+Nous sommes intéressés par l'apprentissage des Bandits Multi-Bras (MAB),
+qui est apparu comme un problème intéressant dans la communauté statistique et parmi les chercheurs intéressés par l'apprentissage séquentiel, comme le montrent les travaux antérieurs de \cite{Thompson33,Robbins52,LaiRobbins85}.
+Il est également inclus dans le concept plus général d'apprentissage du renforcement (RL), lui-même un des domaines de l'apprentissage machine (ML).
+ML > RL > MAB
+%
+Le livre de référence sur RL est \cite{SuttonBarto2018}, citons donc la définition de RL donnée par R. Sutton et A. Barto :
+\emph{``l'apprentissage par renforcement, c'est apprendre quoi faire - comment faire le lien entre les situations et les actions -
+pour maximiser un signal de récompense numérique. L'apprenant n'est pas informé des mesures à prendre pour
+mais doit plutôt découvrir quelles actions sont les plus gratifiantes en les essayant''}.
+% apprentissage statistique > séquentiel > par renforcement > information partielle (bandits)
+
+Nous illustrons ci-dessous l'idée d'un cycle d'apprentissage, alternant entre l'action et le feedback,
+dans la Figure~\ref{fig:1:RenforcementLearningCycle}.
+Un joueur (ou un apprenant) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
+Le but du joueur est de maximiser ses récompenses, par des essais et des erreurs (@@ie, actions et récompenses).
+De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage de renforcement, comme l'illustrent l'enquête \cite{bouneffouf2019survey} et la section~\ref{sec:2:applicationsdeStochasticMAB}, par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie (essai clinique), etc.
+
+% Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
+Comme cette thèse se concentre sur les modèles d'apprentissage de renforcement,
+il est important de souligner que dans de tels modèles de prise de décision, l'apprenant n'a pas accès à l'ensemble de la réaction de l'environnement après avoir pris ses mesures.
+%
+En d'autres termes, le joueur ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une autre action.
+Ce genre de rétroaction limitée s'appelle \emph{information sur les bandits}, et nous discutons de l'histoire et du concept des \emph{Multi-Armed Bandits}.
+(MAB) en détail dans le prochain chapitre~\ref{chapitre:2}.
+Les essais cliniques et la découverte de traitements ont été historiquement la première application du MAB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33}.
+%
+Le MAB est un exemple simple mais puissant du dilemme bien connu de l'exploration/exploitation :
+lorsqu'il est confronté à un ensemble d'actions de $K$ dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
+explorer les actions inconnues, afin de recueillir plus d'informations à leur sujet,
+et en exploitant la meilleure action, selon ses connaissances actuelles.
+%
+Le problème du MAB a été étudié à la fois dans la communauté de l'apprentissage machine et de la statistique, depuis les $1950$ avec des pionniers comme H.
+Robbins \cite{Robbins52}, et plus récemment c'est un domaine de recherche actif, depuis les $1990$ \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
+La recherche sur le MAB a produit une vaste littérature dans le domaine des $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet de grand intérêt, comme l'illustrent les enquêtes et livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications du MAB dans les récentes années \cite{bouneffouf2019survey}.
+
+\paragraph{Accès Opportuniste au Spectre (OSA).}
+%
+Les deux communautés de l'OSA et du MAB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des $2000$ et le début des $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
+L'accent est mis sur une UD accédant à un spectre sous licence, occupé par une UD, qui a une priorité stricte sur l'UD, qui doit suivre un schéma d'accès " écoute avant conversation ".
+%
+Les hypothèses suivantes sont bonnes :
+l'UD est équipée d'une capacité de détection,
+et considère un ensemble fixe et fini de canaux orthogonaux, c'est-à-dire différentes bandes de fréquences dans un spectre sous licence.
+Par exemple, il peut s'agir d'un ensemble de trois canaux Wi-Fi à fréquence différente, émis par la même station Wi-Fi.
+Une autre hypothèse est que PU et SU sont synchronisés dans le temps, en subdivisant le temps en pas de temps discrets (ou ronds).
+%
+Ainsi, si l'unité de station de base passe peu de temps au début de chaque plage horaire pour effectuer la détection, elle peut rechercher la présence ou l'absence d'une UP avant de transmettre, pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
+Si l'UD était capable de détecter toutes les bandes de fréquences $K$, elle pouvait simplement émettre dans l'un des canaux libres s'il y en avait, ou ne pas émettre si tous les canaux sont utilisés à un pas de temps donné.
+Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les enquêtes \cite{yucek2009survey,subhedar2011spectrum}, de sorte que la plupart des travaux sur l'OSA limitent la capacité de détection de l'UD à une seule chaîne à la fois.
+Cette hypothèse, ainsi que l'hypothèse selon laquelle la PU ne peut pas être perturbée (sans laquelle aucune régulation ne sera jamais acceptée pour l'OSA), impose à l'UU de transmettre dans le canal qu'elle a détecté, à chaque pas de temps, si et seulement si elle a été détectée sans PU quelconque.
+coût non négligeable de la reconfiguration radio
+
+En se concentrant sur un SU dans un réseau OSA, il doit décider séquentiellement d'un canal à détecter (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il effectue la détection, et enfin il transmet dans ce canal si il a été détecté libre.
+Le but de l'UD est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou l'équivalent, pour maximiser son nombre de transmissions réussies.
+%
+Si les différents canaux ne sont pas uniformément utilisés par la PU, et si l'on suppose une hypothèse stationnaire sur le trafic de la PU, alors l'objectif de la SU se résume à explorer les différents canaux et à exploiter les meilleurs.
+Le problème de l'accès opportuniste au spectre est donc un problème d'exploration/exploitation, avec un ensemble fini d'actions (les canaux, également appelés \emph{arms}),
+dans un cycle séquentiel d'action puis de rétroaction (les pas de temps sont $t\in\N^*$),
+sous la rétroaction partielle d'information (\ie, environ un canal de plus de $K$ à chaque pas de temps).
+Ces trois hypothèses sont celles qui limitent le cadre général du blanchiment d'argent au cas spécifique du bandit multibras (voir Figure~\ref{fig:1:ReinforcementLearningCycle}).
+
+\paragraph{MAB pour l'OSA, et une brève présentation de l'historique de cette recherche dans l'équipe SCEE.}
+%
+Les travaux antérieurs de notre équipe SCEE ont montré que le MAB peut être utilisé pour modéliser le problème de l'AOS :
+les bandes de fréquences orthogonales (ou canaux) sont des modèles par bras $k\in\{1,\dots,K\}$,
+et le feedback obtenu par le dispositif équipé de CR après avoir détecté le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
+En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message SU peut être envoyé), alors qu'une récompense de $r(t)=0$ indique que le canal $k$ est occupé au moment $t$ et qu'aucun message SU ne doit être envoyé, jusqu'à la fin de cet intervalle.
+%
+Ce modèle a d'abord été étudié par W. Jouini lors de sa thèse de doctorat avec C. Moy \cite{Jouini12PhD}, il y a dix ans, d'abord avec \cite{Jouini09} et ensuite avec \cite{Jouini10,Jouini12}.
+
+Leurs travaux ont été parmi les premiers à proposer l'utilisation de Reinforcement Learning for Cognitive Radio et le problème OSA, notamment le modèle MAB et l'algorithme $\UCB_1$,
+ainsi que les premiers travaux de Q. Zhao et de son équipe, par exemple avec les travaux \cite{Liu08,Zhao10}.
+%
+Peu de temps après, en $2014$, C.
+Moy et son étudiant C.
+Robert \cite{RobertSDR2014,MoyWSR2014,MoyWSR2014} ont développé des preuves de concept utilisant le matériel radio et la radio logicielle du monde réel en utilisant des cartes USRP et le logiciel MATLAB/Simulink.
+Dans une deuxième thèse de doctorat \cite{Modi17PhD}, N.
+Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes MAB pour optimiser la sélection des canaux, sur la durée de vie des piles d'un appareil sans fil.
+D'une part, l'utilisation d'un algorithme MAB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du dispositif.
+D'autre part, les algorithmes MAB classiques ont tendance à changer souvent de bras, surtout au début du processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour le dispositif sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par le dispositif.
+Chaque reconfiguration matérielle coûte de l'énergie à l'appareil et les algorithmes de commutation rapide entraînent une réduction de l'autonomie de la batterie.
+Le compromis entre les deux aspects est étudié de façon empirique dans les études suivantes
+\cite{modiDemo2016}.
+
+En $2017$, C. Moy a continué à travailler dans cette direction, avec un étudiant postdoctoral, S. Darak, qui a mené à des publications telles que
+\cite{darak2016bayesian,Darak16}.
+%
+Par exemple, des preuves de concepts comme \cite{kumar2016two} ont prouvé la capacité de telles approches sur des signaux radio réels pour l'OSA.
+%
+Certaines analyses des mesures radio réelles effectuées pour les canaux ionosphériques HF ont également prouvé que les solutions basées sur l'apprentissage MAB sont appropriées et résolvent efficacement ce type de problèmes de prise de décision sur les signaux sans fil du monde réel \cite{Melian15}.
+%
+Depuis $2017$, S. Darak et son équipe à l'IIIT Delhi en Inde, ont travaillé activement dans la recherche sur la radio intelligente à l'aide de bandits à armes multiples.
+Héritage de son travail au sein de l'équipe SCEE,
+certains de leurs travaux récents sont également illustrés par des démos du monde réel utilisant USRP et le système MATLAB/Simulink
+\cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
+%
+Pour plus de détails sur l'état de la recherche sur la radio intelligente, nous renvoyons aux enquêtes \cite{garhwal2012survey,marinho2012cognitive}.
+
+\paragraph{Limitations et spécificités des réseaux IdO.}
+%
+La littérature susmentionnée a essentiellement montré que les algorithmes MAB peuvent être appliqués avec succès au problème de l'AOS.
+Toutefois, si l'on considère les dispositifs prêts à l'emploi qui ne peuvent pas effectuer la détection, tels que les dispositifs finaux à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle MAB qui utilise la détection pour détecter le PU dans le cas de l'OSA ne peut plus être appliqué.
+En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre PU et SU.
+En d'autres termes, il n'y a plus de priorité entre les utilisateurs des réseaux IdO.
+(plus de PU, tous sont SU).
+%
+Les spécificités des réseaux IdO peuvent être énumérées comme suit,
+et nous nous référons à l'enquête \cite{Centenaro16} pour plus de détails.
+%
+\begin{itemize}\tightlist
+    \item
+    nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre PU et SU),
+    \item
+    la plupart des dispositifs IoT sont des dispositifs à très faible coût, fonctionnant éventuellement sur du matériel de mauvaise qualité et ayant des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
+    Cela signifie que même s'ils ne sont pas équipés de capacités de détection particulières,
+    ils sont équipés d'un émetteur ainsi que d'un récepteur (Tx et Rx, \ie, un émetteur-récepteur).
+    et ils ont des capacités de stockage et de calcul peu complexes (petits processeurs embarqués),
+    \item
+    une seule station de base IoT devra gérer un très grand nombre d'appareils,
+    et ne peut leur envoyer des ordres de coordination afin d'optimiser le réseau,
+    \item
+    Les dispositifs IoT ont des cycles de fonctionnement faibles (seulement quelques messages par minute) à très faibles cycles de fonctionnement (seulement quelques messages par mois).
+\end{itemize}
+
+Une coordination centrale des appareils avec leur station de base est impossible car il faudrait communiquer en permanence avec la station de base, ce qui viole les deux dernières contraintes.
+%
+De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de la batterie, les dispositifs IoT sont capables d'utiliser leur antenne reçue pour détecter un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), mais ils ne peuvent pas effectuer de détection de spectre comme cela est envisagé pour OSA et CR (\ie, au début de chaque intervalle, dans un schéma " écoute avant conversation ").
+
+On pourrait penser qu'en l'absence de détection, l'apprentissage du renforcement n'est plus possible, mais tout appareil IoT réel reçoit encore des informations sur son environnement après certaines ou toutes les transmissions.
+Dans la plupart des normes IoT, un message de liaison montante envoyé à une station de base peut être suivi d'un message de liaison descendante renvoyé par la station de base pour indiquer si le message de liaison montante a été bien reçu et compris.
+%
+En utilisant cette rétroaction, ça consiste en un \emph{acknowledgement} (\Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IoT puisse également utiliser un algorithme RL bien conçu afin d'optimiser ses communications.
+
+% ----------------------------------------------------------------------------
+\section*{Nos contributions}
+\label{sec:1:contributions}
+
+Nous commençons par formuler le problème étudié dans cette thèse, puis nous développons notre approche.
+%
+Si nous résumons les problèmes étudiés et les résumons en une seule question, ce pourrait être la suivante :
+\Insister sur le fait que l'on peut adapter les outils de prise de décision déjà appliqués avec succès à la radio intelligente pour l'accès opportuniste au spectre aux besoins spécifiques de la CR pour les (futurs) réseaux de l'Internet des objets ?
+%
+Nous répondons partiellement à cette problématique par les étapes de recherche suivantes.
+
+\subsection*{Explorer la jungle des algorithmes MAB}
+
+\paragraph{Chapter~\ref{chapter:2}.}
+%
+Nous avons commencé par explorer la riche littérature des bandits aux armes multiples,
+car il existe de nombreux algorithmes différents, avec de nombreuses variantes sur le problème simple présenté ci-dessus.
+Nous commençons donc la première partie~\ref{part:Introduction} de cette thèse par le chapitre~\ref{chapitre:2}, qui présente le modèle MAB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires du MAB.
+%
+Afin de bien comprendre quels algorithmes pourraient être adaptés aux contraintes susmentionnées de la RC pour les réseaux IdO,
+nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme MAB (son regret, voir ci-dessous dans Section~\ref{sec:2:lowerUpperBoundsRegret}),
+mais aussi par leurs performances empiriques en termes de complexité de calcul et d'exigences de stockage, tant du point de vue des analyses théoriques que des mesures en temps réel et des empreintes mémoire.
+
+\paragraph{Chapter~\ref{chapter:3}.}
+%
+Notre exploration du grand nombre d'algorithmes et de modèles MAB développés dans la littérature récente.
+nous a donné l'ambition d'écrire un seul logiciel permettant à chacun d'implémenter facilement de nouveaux modèles et algorithmes, afin de comparer les modèles existants et d'explorer empiriquement les performances des nouveaux algorithmes proposés.
+Pour atteindre cet objectif, nous avons produit une bibliothèque Python de simulation des problèmes MAB.
+%
+Nous avons écrit la bibliothèque de simulation open-source la plus exhaustive pour les problèmes MAB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
+Nous présentons en détail son architecture et ses fonctionnalités dans le chapitre~\ref{chapitre:3}, ainsi que différents exemples de son utilisation.
+Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les simulations discutées dans la suite de cette thèse.
+
+\paragraph{Chapter~\ref{chapter:25}.}
+%
+Étant donné le grand nombre d'algorithmes MAB différents, nous nous intéressons aussi à la question de savoir comment un praticien peut choisir l'algorithme à mettre en œuvre, dans un objet IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
+Pour répondre à cette question, nous présentons deux approches.
+La première est une comparaison empirique des algorithmes existants les plus efficaces et les mieux connus, dans Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
+Nous confirmons que les algorithmes largement utilisés et pas trop sophistiqués, tels que \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
+La seconde approche est une sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons dans Chapitre~\ref{chapitre:25}.
+Ce travail sur l'agrégation des algorithmes MAB a été motivé pour le cas OSA de la radio cognitive, où de nombreux travaux de recherche antérieurs ne prenaient en compte que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
+Cette contribution a été présentée à la conférence WCNC de l'IEEE à Barcelone, Espagne, en avril 2010.
+
+% FIXME VRAIMENT BEAUCOUP raccourcir la suite !
+
+\subsection*{Nos modèles de réseaux d'IdO et de MAB décentralisé}
+
+Dans la deuxième partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux d'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du MAB, issues de la simplification de notre premier modèle.
+
+\paragraph{Chapter~\ref{chapter:4}.}
+%
+Nous étudions deux modèles de réseaux IdO dans le chapitre~\ref{chapitre:4}.
+Le premier modèle considère les dispositifs IdO qui ont simplement des données à envoyer régulièrement à une station de base (à des moments imprévisibles ou aléatoires).
+De tels dispositifs utilisent les acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
+L'objectif de ce premier modèle est d'améliorer la qualité de service (QoS) de l'application de ce réseau IoT, en appliquant une RL décentralisée côté appareil afin de réduire le taux d'échec de transmission.
+La station de base recevra donc plus de paquets de liaison montante des appareils desservis, si elle peut apprendre à utiliser un système efficace d'accès au spectre.
+Cela implique que le réseau peut accepter plus d'équipements finaux tout en maintenant la même QoS.
+%
+Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet d'augmenter la durée de vie des piles de chacun des appareils, ainsi que la QoS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
+
+Le premier modèle constitue le premier article écrit au cours de cette thèse, et a initié une collaboration avec R. Bonnefoi, un autre doctorant de notre équipe SCEE.
+Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre 2010, où il a obtenu le "prix du meilleur papier" \cite{Bonnefoi17}.
+%
+La preuve de concept (PoC) a continué notre collaboration fructueuse avec R.
+Bonnefoi, et a été démontrée pendant trois jours à la conférence ICT de l'IEEE à Saint-Malo, France, en juin $2018$ \cite{Besson2018ICT}, et ensuite présentée à la conférence IEEE WCNC à Marrakech, Maroc, en avril $2019$ \cite{Besson2019WCNC}.
+%
+Notre deuxième modèle, avec retransmissions, correspond à notre dernière collaboration avec R.
+Bonnefoi, qui a été présentée à l'atelier $1^{\text{st}}$ MoTION, également pendant la conférence IEEE WCNC $2019$ \cite{Bonnefoi2019WCNC}.
+
+\paragraph{Chapter~\ref{chapter:5}.}
+%
+Dans les deux modèles présentés ci-dessus, le concept de base est de laisser chaque dispositif dynamique d'un réseau IdO exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
+Il est important de noter que chaque dispositif ne communique, et donc n'apprend, qu'à quelques tours et non à toutes les étapes, en suivant son propre processus d'activation (aléatoire) (nous nous limitons à un processus d'activation de Bernoulli purement aléatoire de probabilité $p$).
+Cela signifie que chacun d'eux vise son propre objectif local, qui est de maximiser sa récompense cumulée (d'une manière égoïste).
+Il est bien connu dans la littérature sur la théorie des jeux que le jeu égoïste peut être désastreux pour la mesure centralisée de la performance, on peut penser aux "dilemmes" populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
+Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage MAB décentralisé et égoïste conduisait à une coordination efficace entre les dispositifs dans tous les scénarios considérés, malgré le fait que ces objets IoT ne peuvent communiquer directement entre eux, et reçoivent seulement un retour de collision depuis la station de base à laquelle ils sont associés.
+
+Nous avons d'abord essayé d'analyser la performance de cet algorithme décentralisé, que nous appelons \Selfish, dans le modèle de Section~\ref{sec:4:firstModel},
+mais en raison du nombre aléatoire de dispositifs actifs à chaque pas de temps (@@ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
+%
+C'est pourquoi dans le chapitre~\ref{chapitre:5}, nous relâchons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$, qui est équivalent à avoir $p < 1$, et nous considérons le cas des dispositifs IoT communiquant à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ dispositifs à jour.
+Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas OSA (\ie, avec détection) ou le cas IoT (\ie, sans détection).
+L'objectif était de comprendre l'heuristique de Chapter~\ref{chapter:4}, \Selfish, dans le cadre plus simple du multi-player MAB, qui a été étudié auparavant pour le cas de l'OSA, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
+%
+Le cas OSA couvert par notre modèle est en fait légèrement différent de celui considéré par \cite{Jouini10} et d'autres travaux antérieurs,
+car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
+Il diffère des modèles précédemment étudiés d'application du MAB pour l'AOS, puisqu'ils ne prennent en compte que les dispositifs synchronisés, une unité de base et une unité centrale de traitement, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'unité de base sera certainement reçu avec succès par la base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
+
+D'une part, dans le cadre de la LVMO, nous n'avons pas réussi à obtenir un résultat positif pour la politique égoïste, car nous avons prouvé que dans certains contextes limités (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre de regret moyen linéaire.
+Ce résultat a ensuite été confirmé et analysé par \cite{BoursierPerchet18}.
+%
+D'autre part, nous avons été en mesure de proposer de nouveaux algorithmes pour ce problème de bandit multi-joueurs avec des informations de détection, et nous avons analysé notre proposition \MCTopM-\klUCB, pour montrer qu'elle atteint une limite supérieure de regret logarithmique à temps fini, améliorant considérablement les résultats précédents.
+Notre algorithme atteint également un nombre logarithmique de collisions et de commutateurs de bras, et permet à un groupe fixe de dispositifs $M$ d'apprendre efficacement à utiliser les meilleurs canaux $M$ orthogonalement pour presque toutes leurs communications montantes.
+%
+Ce résultat théorique solide dépend fortement de la présence d'une rétroaction de détection et, par conséquent, nos résultats ne sont pas (encore) applicables au modèle IdO sans détection.
+
+Comme expliqué plus haut, le modèle de Chapitre~\ref{chapitre:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux équipements finaux, tous suivant des modèles d'activation aléatoires.
+La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes MAB, conçus pour résoudre des problèmes stationnaires, sur un problème non stationnaire,
+mais plutôt que nous analysons des algorithmes qui jouent tous dans différents sous-ensembles (aléatoires) des pas de temps synchronisés globaux.
+Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et appliquer au maximum $M \leq K$ aux dispositifs n'est pas vraiment réaliste pour les réseaux IdO, même si cela conduit au modèle du Chapitre~\ref{chapitre:5}.
+
+D'une part, si le modèle d'activation de tous les périphériques peut être corrigé, par exemple en se basant sur une affectation centralisée des périphériques à différents créneaux horaires, alors le modèle de bandit multi-joueurs du Chapitre~\ref{chapitre:5} peut être utilisé pour permettre à chaque groupe de périphériques $M \leq K$ d'apprendre une affectation orthogonale optimale dans les canaux $K$ (\eg, groupes peuvent être le jeu des périphériques que tous émettent au même instant).
+%
+Cependant, même si nous continuons à supposer un temps synchronisé dans tout le reste de la thèse,
+il est difficile de soutenir que cette hypothèse d'un calendrier centralisé pour les équipements finaux peut être réaliste, car nous étudions le cas de l'apprentissage décentralisé pour l'AOS et l'IdO précisément afin d'éviter la signalisation et tout contrôle central des équipements par la station de base, comme nous l'avons expliqué ci-dessus.
+
+D'autre part, une autre façon de relâcher l'hypothèse non stationnaire est de prendre le point de vue d'un dispositif, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur une unité de soins entourée de plusieurs PU ayant des comportements fixes.
+Si nous nous concentrons sur un dispositif IoT, son environnement (\ie, les dispositifs environnants) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réalistes donnent généralement une certaine structure sur cette non-stationnalité.
+Par exemple, l'application de la stationnarité à certains intervalles de temps conduit à la dernière contribution et au dernier chapitre de cette thèse.
+
+\paragraph{Chapter~\ref{chapter:6}.}
+%
+Afin de comprendre également avec précision comment les algorithmes MAB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur le MAB accusatoire et sur le MAB non stationnaire, pour les deux cas d'environnements variant lentement et changeant brusquement.
+Nous commençons notre dernier Chapitre~\ref{chapitre:6} en passant en revue les travaux existants,
+et nous avons choisi de nous concentrer sur le problème des pièces stationnaires,
+ce qui signifie que le problème de bandit sous-jacent est stationnaire à certains intervalles, séparés par des points de changement situés à des moments inconnus.
+Supposer que l'environnement est stationnaire au niveau de la pièce est en effet logique pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'équipements terminaux dans le réseau.
+Des exemples d'une telle situation peuvent être une nouvelle entreprise arrivant sur le marché dans une ville (par exemple, les comptoirs Linky qui sont installés aujourd'hui en France), ou l'agriculteur voisin installant des capteurs sur son propre troupeau de vaches, etc.
+%
+Deux grandes familles d'algorithmes ont été proposées pour le problème stationnaire par pièce,
+qui consistent généralement à combiner une politique efficace conçue pour le problème stationnaire du MAB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des armes.
+Les politiques passivement adaptatives utilisent une fenêtre de taille fixe ou évolutive \cite{Garivier11UCBDiscount}, ou un facteur d'actualisation, pour oublier les observations passées \cite{Kocsis06,Gupta11thompson},
+tandis que les politiques activement adaptatives utilisent un test statistique pour détecter les points de changement \cite{MellorShapiro13,Allesiardo15}.
+%
+Comme la littérature récente a montré que cette dernière approche est généralement plus compétitive, en obtenant de meilleurs résultats tant sur le plan empirique que théorique, nous avons choisi de développer notre propre algorithme activement adaptatif.
+
+Suite à deux travaux précédents récents \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, nous combinons une politique d'index efficace (\klUCB) et un test efficace de détection des points de changement, dans l'hypothèse de récompenses limitées.
+Nous nous appuyons sur des résultats très récents du test du rapport de vraisemblance généralisé (TRGL) pour les variables gaussiennes et sous-gaussiennes \cite{Maillard2018GLR}, mais nous nous concentrons plutôt sur les récompenses limitées et les distributions de Bernoulli.
+Les récompenses limitées sont en effet généralement plus appropriées pour les applications de radio cognitive, comme indiqué ci-dessus.
+En utilisant le fait que les variables bornées en $[0,1]$ ne sont pas seulement sous-Gaussiennes à $1/4$ mais aussi sous-Bernoulli, nous prouvons les premières garanties à temps fini pour le GLRT pour les variables bornées.
+Nous montrons d'abord des limites temporelles finies à la fois sur la probabilité de fausse alarme de notre test, et sur son délai de détection, sous de légères hypothèses sur la longueur des séquences stationnaires.
+Notre algorithme, noté \GLRklUCB, est alors présenté en deux variantes, que les points de changement soient locaux (\ie, une seule moyenne de bras change à chaque point de changement) ou globaux (\ie, peut-être tous les moyens de bras changent à la fois).
+%
+Nous prouvons qu'en combinant la politique \klUCB, asymptotiquement optimale pour les problèmes stationnaires, et notre nouvelle analyse du GLRT pour les récompenses limitées, nous obtenons des garanties de pointe sur le regret de notre algorithme proposé.
+Nous considérons la même hypothèse que nos concurrents, en supposant que la longueur des intervalles stationnaires est " assez longue " pour que les points de changement soient détectables.
+La limite supérieure du meilleur regret est obtenue lorsque l'algorithme connaît à l'avance l'horizon et le nombre de points de changement, mais notre algorithme n'a pas besoin d'avoir d'autres connaissances sur la difficulté du problème pour être accordé de manière optimale.
+%
+La performance de \GLRklUCB{} est également illustrée par des expériences numériques sur des données synthétiques, où il est démontré qu'elle surpasse toutes les politiques passivement adaptatives ainsi que les précédentes politiques activement adaptatives.
+%
+Ce dernier chapitre est basé sur notre dernier travail, qui a mené à une publication à la conférence du GRETSI à Lille, France, en août $2019$ \cite{Besson2019Gretsi}.
+Notre travail a aussi mené à l'article de la version longue \cite{Besson2019GLRT}, qui sera partiellement réécrit et complété avec des résultats plus récents, afin de le soumettre bientôt à une revue (probablement le Journal of Machine Learning Research) avant décembre $2019$.
+
+\paragraph{Annexe.}
+%
+Ce manuscrit se termine par
+une liste d'abréviations et de notations, puis des listes de figures, d'algorithmes, d'échantillons de code et de tableaux,
+et enfin la liste des références bibliographiques.
+
+% FIXME VRAIMENT BEAUCOUP raccourcir ce qui précéde !
 
@@
     % (ALT 2018, et 4-8 travaux inspirés de notre article depuis lors).
-Etat de l'art avec notre algorithme MCTopM + klUCB pour les bandits multi-joueurs "avec détection", état de l'art empirique avec notre approche simple (mais erronée) "égoïste" en cas de "non détection".
+Etat de l'art avec notre algorithme MCTopM + klUCB pour les bandits multi-joueurs ``avec détection'', état de l'art empirique avec notre approche simple (mais erronée) ``égoïste'' en cas de ``non détection''.
 
@@
         \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!20] {\textbf{Chapitre~\ref{chapter:conclusion}}\\Conclusion générale};
-        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=yellow!10] {Annexes};
+        % \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=yellow!10] {Annexes};
         %
@@
         \draw [color=black,thick,->] (chapter6) to (conclusion);
-        \draw [color=black,thick,->] (conclusion) to (appendix);
+        % \draw [color=black,thick,->] (conclusion) to (appendix);
         %

### 2019-07-05 2-Chapters/1-Chapter/chapter1.tex
Small modifs

@@
 
-    \item
-    The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
-    and a unified and more generic analysis of two families of doubling tricks.
-    This lead to the article \cite{Besson2018DoublingTricks}.
-    % which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
+    % \item
+    % The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
+    % and a unified and more generic analysis of two families of doubling tricks.
+    % This lead to the article \cite{Besson2018DoublingTricks}.
+    % % which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
 \end{itemize}
@@
         \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!20] {\textbf{Chapter~\ref{chapter:conclusion}}\\General Conclusion};
-        \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=yellow!10] {Appendix};
+        % \node[align=center] (appendix) at (2.5,-0.25) [rectangle,draw,fill=yellow!10] {Appendix};
         %

### 2019-07-05 2-Chapters/2-Chapter/chapter2.tex
Small modifs

@@
     % \item
-    MAB can also be applied to a broader setting of \textbf{online content recommandation},
-    with two options (A/B testing), or more than two options.
-    The seminal work of \cite{Li10} studies the application of contextual bandit to news articles recommendation, as it is used in practice on platforms such as Microsoft's Bing news website,
+    MAB can also be applied to a broader setting of \textbf{online content recommandation}.
+    % with two options (A/B testing), or more than two options.
+    The seminal work of \cite{Li10} studies the application of contextual bandit to online advertisement
+    or news articles recommendation, as it is used in practice on platforms such as Microsoft's Bing news website,
     or in video streaming platform like YouTube or Netflix.
@@
     \begin{align}\label{eq:2:HoeffdingInequality}
-        % \begin{cases}
-            \forall x < \mu, \;\;\; \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
-            \;\;\;
-            \forall x > \mu, \;\;\; \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
-        % \end{cases}
-        % \begin{cases}
-        %     \forall x < \mu, \;\;\; & \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
-        %     \\
-        %     \forall x > \mu, \;\;\; & \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
-        % \end{cases}
+        % % \begin{cases}
+        %     \forall x < \mu, \;\;\; \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
+        %     \;\;\;
+        %     \forall x > \mu, \;\;\; \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+        % % \end{cases}
+        \begin{cases}
+            \forall x < \mu, \;\;\; & \bP(\widehat{Z_n} < x) \leq \exp(-2 n (x-\mu)^2),
+            \\
+            \forall x > \mu, \;\;\; & \bP(\widehat{Z_n} > x) \leq \exp(-2 n (x-\mu)^2).
+        \end{cases}
     \end{align}
@@
 The \KLUCB{} and \klUCB{} algorithms achieve finite-time regret logarithmic upper-bounds on their regret, and have been proven to be asymptotically optimal, respectively for one-dimensional exponential families and for bounded rewards distributions \cite{Garivier11KL,KLUCBJournal}.
+% FIXME
 
@@
 % \subsection{A first numerical experiment on the example Bernoulli problem}
-\subsection{Comparison of the main algorithms on the example Bernoulli problem from the online interactive demonstration}
+% \subsection{Comparison of the main algorithms on the example Bernoulli problem from the online interactive demonstration}
+\subsection{Comparison of the main algorithms on an example Bernoulli problem}
 \label{sub:2:shortNumericalExperiments}

### 2019-07-06 2-Chapters/1-Chapter/chapter1.tex
J'utilise moins "our ..."

@@
 
-We start by formulating the problem studied in this thesis, and then we develop our approach.
+We start by formulating the problem studied in this thesis, and then we develop the organization of the contributions.
 %
@@
 % \subsection{Focusing on IoT networks.}
-\subsection{Our models of IoT networks and decentralized MAB}
+\subsection{Two models of IoT networks and decentralized MAB}
 
@@
 % % Verify this first model empirically: proof of concept
-The proof-of-concept (PoC) continued our fruitful collaboration with R.
+The proof-of-concept (PoC) continued the fruitful collaboration with R.
 Bonnefoi, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
@@
 % This shows that using learning in a decentralized way on the devices' side is again a powerful idea under this model with retransmissions, and we show as well that the simplest of the different heuristics we compared is efficient, and allows the devices to learn to behave almost optimally.
-Our second model, with retransmissions, corresponds to our last collaboration with R.
-Bonnefoi, which was presented at the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
+The second model with retransmissions concluded our collaboration with R.
+Bonnefoi, and resulted to a paper presented at the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
 
@@
 % https://www.ii.uni.wroc.pl/~lukstafi/pmwiki/uploads/AGT/Prediction_Learning_and_Games.pdf
-So it was quite surprising that our numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices in all the considered scenarios, despite the fact that these IoT objects cannot communicate directly with each other, and receive only a collision feedback from the base station they are associated to.
+So it was quite surprising that the numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices in all the considered scenarios, despite the fact that these IoT objects cannot communicate directly with each other, and receive only a collision feedback from the base station they are associated to.
 
@@
 %
-The OSA case covered by our model is actually slightly different as the one considered by \cite{Jouini10} and other previous works,
+The OSA case covered by the studied model is actually slightly different as the one considered by \cite{Jouini10} and other previous works,
 as our model requires that an \Ack{} is sent back by the base station if the transmission was successful, even in the case where sensing information is available.
@@
 In order to also understand precisely how MAB algorithms behave under non-stationary environment, we studied the literature on adversarial and on non-stationary MAB, for the two cases of slowly-varying and abruptly-changing environments.
-We begin our last Chapter~\ref{chapter:6} by reviewing the existing works,
+We begin the last Chapter~\ref{chapter:6} by reviewing the existing works,
 and we chose to focus on the piece-wise stationary problem,
@@
 Using the fact that bounded variables in $[0,1]$ are not only $1/4$ sub-Gaussian but also sub-Bernoulli, we prove the first finite-time guarantees for the GLRT for bounded variables.
-We first show finite-time bounds on both the false alarm probability of our test, and its detection delay, under mild assumptions on the lengths of the stationary sequences.
+We first show finite-time bounds on both the false alarm probability of the test, and its detection delay, under mild assumptions on the lengths of the stationary sequences.
 Our algorithm, denoted \GLRklUCB, is then presented in two variants, whether change-points are local (\ie, only one arm mean changes at each change-point) or global (\ie, possibly all the arm means change at a time).
 %
-We prove that by combining the policy \klUCB, asymptotically optimal for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of our proposed algorithm.
+We prove that by combining the policy \klUCB, asymptotically optimal for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of the proposed algorithm.
 We consider the same hypothesis as our competitors, by assuming that the length of the stationary intervals to be ``long enough'' for the change-points to be detectable.
@@
 This last chapter is based on our last work, which lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
-Our work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results, in order to submit it soon to a journal (probably the Journal of Machine Learning Research) before December $2019$.
+This work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results, in order to submit it soon to a journal (probably the Journal of Machine Learning Research) before December $2019$.
 
@@
     as well as extensive numerical experiments to demonstrate its excellent performance in comparison with the rest of the literature.
-    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works heavily built up on our results.
+    Some recent research works heavily built up on our results, and the research on multi-players bandits has been quite active since its publication \cite{Besson2018ALT}.
 
@@
 % =============================================================================
-\subsection*{Demonstration in international conferences}
+\subsection*{Demonstrations in international conferences}
 
@@
 % =============================================================================
-\subsection*{French language conference with proceedings}
+\subsection*{French language conferences with proceedings}
 % \textbf{Submitted works}

### 2019-07-06 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Cleanup empty pages between chapters etc. Less than 300 pages YAY
Lots of improvements on the arXiv papers
Small modifs
J'utilise moins "our ..."

@@
 %
-To the best of our knowledge, aggregation algorithms, such as \ExpQ{} which dates back from 2002 \cite{Auer02},
-have never been used in practice for stochastic MAB problems.
+To the best of the authors' knowledge, aggregation algorithms, such as \ExpQ{} which dates back from 2002 \cite{Auer02},
+have actually never been used in practice for stochastic MAB problems.
+%
 For instance, if we look at the applications of MAB for Opportunistic Spectrum Access, like it was proposed in \cite{Jouini09,Jouini10,Jouini12},
@@
 % ----------------------------------------------------------------------
-\section{Our contribution: the \Aggr{} algorithm}\label{sub:25:Aggr}
+\section{The \Aggr{} algorithm}\label{sub:25:Aggr}
 
-Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
+The \Aggr{} algorithm is detailed in Algorithm~\ref{algo:25:Aggr}.
 %
@@
 Note that this choice is the default in SMPyBandits, but the other possibility is implemented as well, by changing \texttt{cache\_rewards} to \texttt{true} or \texttt{false} in the \texttt{configuration} dictionary.
-See Section~\ref{sub:3:miniConfigurationFileSMPyBandits} for explanations on this \texttt{configuration.py} file, or the online documentation.
+We refer to the online documentation for explanations on all these details.
+% See Section~\ref{sub:3:miniConfigurationFileSMPyBandits} for explanations on this \texttt{configuration.py} file, or the online documentation.
 
-We compare our \Aggr{} algorithm,
+We compare the \Aggr{} algorithm,
 as well as other aggregation algorithms, \ExpQ{} from \cite{Bubeck12},
@@
 
-For each of the 4 problems considered, our \Aggr{} algorithm  is the best of all the aggregation algorithms,
+For each of the 4 problems considered, the \Aggr{} algorithm  is the best of all the aggregation algorithms,
 and its regret is close to the best of the aggregated algorithms.
@@
 % ----------------------------------------------------------------------
-\subsection{No theoretical guarantees.}\label{sub:25:theory}
+% \subsection{No theoretical guarantees.}\label{sec:25:theory}
+\section{Towards theoretical guarantees}\label{sec:25:theory}
 
@@
 
-Our proposed algorithm follows quite closely the architecture of \ExpQ,
+The proposed algorithm follows quite closely the architecture of \ExpQ,
 and a similar bound for \Aggr{} is expected to hold.

### 2019-07-06 2-Chapters/2-Chapter/chapter2.tex
J'utilise moins "our ..."

@@
 %
-Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm for bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and we use the Bernoulli Kullback-Leibler divergence $\kl$ in our analysis.
+Finally, in Chapter~\ref{chapter:6} we analyze our proposed algorithm for bounded distributions, without restricting to Bernoulli distributions, but we use the fact that bounded distributions on $[0,1]$ are sub-Bernoulli, and our analysis use the Bernoulli Kullback-Leibler divergence $\kl$.
 
@@
 Moreover, the $\UCB_1$ algorithm has the advantage of being easy to explain and understand, and its decisions can be interpreted clearly: by reading the values of $\widehat{\mu_k}(t)$ and $\UCB_k(t)$, anyone can see why an arm was chosen at anytime.
-This is why it is used in our demonstration presented in Section~\ref{sec:4:gnuradio}.
+This is why it is used in the proof-of-concept demonstration presented in Section~\ref{sec:4:gnuradio}.
 
@@
 %
-For both directions, we present natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
-of stationary multi-players and piece-wise stationary bandits.
+For both directions, we present natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results,
+for the two problems of stationary multi-players and piece-wise stationary bandits.
 

### 2019-07-06 2-Chapters/2-Chapter/chapter25.tex
J'utilise moins "our ..."

@@
 First, we give more motivations in the rest of this section, then we explain how to combine such stationary MAB algorithms and aggregate them.
-We present our proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
+We present the proposed algorithm, called \Aggr, in Section~\ref{sub:25:Aggr},
 Finally, we present numerical experiments in Section~\ref{sub:25:numExp},

### 2019-07-06 2-Chapters/3-Chapter/chapter3.tex
Small modifs
J'utilise moins "our ..."

@@
 
-% \chapter{SMPyBandits: a state-of-the-art Python library to simulate MAB problems}
-\chapter{SMPyBandits: our Python library to simulate MAB problems}
+\chapter{SMPyBandits: an exhaustive Python library to simulate MAB problems}
+% \chapter{SMPyBandits: our Python library to simulate MAB problems}
 \label{chapter:3}
@@
 The code for SMPyBandits is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source License.
-This chapter is based on our article \cite{SMPyBanditsJMLR}.
+This chapter is based on the article \cite{SMPyBanditsJMLR}.
 
@@
 
-We chose to present here our Python library SMPyBandits in this chapter rather than in the Appendix, because all the numerical experiments on single-player multi-arm bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on our library.
-%
-We detailed three main points of SMPyBandits:
+We chose to present here the Python library SMPyBandits in this chapter, rather than in the Appendix, because all the numerical experiments on single-player multi-arm bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on SMPyBandits.
+%
+We detailed three main points of our library:
 $(i)$ its purpose and qualities,
@@
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
-One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research from December $2016$ to June $2019$.
+One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we followed the active research from December $2016$ to June $2019$.
 More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
@@
 %
-We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to our work.
+We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to this work.
 %

### 2019-07-06 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
J'utilise moins "our ..."

@@
 Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel.
-So our first hypothesis is actually not constraining.
+So this first hypothesis is actually not constraining.
 We first evaluate the probability of collision if the dynamic devices randomly select their channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
@@
 
-We note that MAB learning has also started to be applied to also optimize the PHY layer parameters, see for instance \cite{KerkoucheAlami18} following our article \cite{Bonnefoi17},
+We note that MAB learning has also started to be applied to also optimize the PHY layer parameters, see for instance \cite{KerkoucheAlami18} which followed our article \cite{Bonnefoi17},
 but in this thesis we chose to restrict to the spectrum access problem and thus we only MAB models where the arms correspond to orthogonal frequency bands.
@@
 %
-Our IoT application is challenging in that there are \emph{multiple} players (the dynamic devices) interacting with the \emph{same} arms (the channels), without any centralized communication (they do not even know the total number of dynamic devices).
+The IoT application is challenging in that there are \emph{multiple} players (the dynamic devices) interacting with the \emph{same} arms (the channels), without any centralized communication (they do not even know the total number of dynamic devices).
 %

### 2019-07-06 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
J'utilise moins "our ..."

@@
 
-In this section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
+In this section, we present the implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
 We first describe the simplified physical layer of this demo, then we present our GNU Radio implementation.
@@
 
-Regarding the physical layer, we consider a QPSK constellation (\emph{Quadrature Phase-Shift Keying}). Moreover, we use simplified packets composed of two parts.
+Regarding the physical layer, we consider a QPSK constellation (\emph{Quadrature Phase-Shift Keying}, \eg, see references in \cite{Bodinier17}).
+Moreover, we use simplified packets composed of two parts.
 The first part is the \emph{preamble} which is used for the purpose of synchronization (phase correction).
@@
 with version $4$ of their FPGA system and version $5.1$ of the RBX system.
-As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP:
+As illustrated in Figure~\ref{fig:42:our_demo}, the implementation is composed of at least $3$ USRP:
 the base station,
@@
 
-We have designed a user interface in order to visualize the results obtained  with our experimental demonstration. This user interface is shown in Figure~\ref{fig:42:UI}.
+We have designed a user interface in order to visualize the results obtained  with the experimental demonstration. This user interface is shown in Figure~\ref{fig:42:UI}.
 We can see that it is made of three parts, one for each USRP, as highlighted in \textcolor{darkred}{circled red numbers}.
@@
 Each experiment has been done on a duration of about half a day,
-due to the IoT sporadic transmission mode that we want to respect (like in our model of Section~\ref{sec:4:firstModel}).
+due to the IoT sporadic transmission mode that we want to respect (like in the model of Section~\ref{sec:4:firstModel}).
 However, we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
@@
 Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
-However, our experimental testbed can not run hundreds of intelligent devices.
+However, our experimental testbed cannot run hundreds of intelligent devices.
 Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).

### 2019-07-06 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
J'utilise moins "our ..."

@@
 % It motivates our present work for which we consider the retransmissions in the analysis of MAB algorithms.
-It motivates our present work for which we consider the number of retransmissions, carried out by each device.
+It motivates this section for which we consider the number of retransmissions, carried out by each device.
 % to address several MAB algorithms.
@@
 Thus, it seems interesting to consider in the decision making policy the possibility for a device to retransmit in a different channel.
-One of our motivations to develop new MAB algorithms for our problem is this option of using a different communication channel between the first transmission and the next retransmissions.
+One of our motivations to develop new MAB algorithms for the considered problem is this option of using a different communication channel between the first transmission and the next retransmissions.
 
@@
 
-We assume to be in the steady state \cite{Norris98}, in our Markov chain model depicted in Figure~\ref{fig:43:Markov_model}, and thus the probabilities no longer depend on the slot number $t$ (\ie, $\forall t, x_t=x$).
+We assume to be in the steady state \cite{Norris98}, in the Markov chain model depicted in Figure~\ref{fig:43:Markov_model}, and thus the probabilities no longer depend on the slot number $t$ (\ie, $\forall t, x_t=x$).
 Therefore, the probability that this device has a collision at the first transmission is $p_c$, and has the following expression
@@
 %
-We can verify that our approximation is very precise for lower values $p_{c1} \leq 30 \%$ (\ie, \textcolor{red}{red} and \textcolor{gold}{orange} curves are quite close).
+We can verify that the obtained approximation is very precise for lower values $p_{c1} \leq 30 \%$ (\ie, \textcolor{red}{red} and \textcolor{gold}{orange} curves are quite close).
 Moreover, a significant gap between $p_{c1}$ and $p_c$,
@@
 	\item
-	On the other hand, if $p_c$ is small enough, \ie, in a network not yet too congested, then our derivation shows that $p_{c1} > p_c$, meaning that the possible gain of retransmitting in a different channel that the one used for the first transmission can be large, in terms of collision probability (\eg, up-to $10\%$ in this experimental setting).
+	On the other hand, if $p_c$ is small enough, \ie, in a network not yet too congested, then the derivation above shows that $p_{c1} > p_c$, meaning that the possible gain of retransmitting in a different channel that the one used for the first transmission can be large, in terms of collision probability (\eg, up-to $10\%$ in this experimental setting).
 	In other words, when learning can be useful (small $p_c$), learning to retransmit in a different channel can have a large impact on the global collision rate,
@@
 Although, this approach increases the complexity and storage requirements (now, of order $\mathcal{O}(K^2)$).
-For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
+For the LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
 For instance, for $K=4$ channels,
@@
 
-We also consider in our experiment the case of an ALOHA protocol using $\mathrm{MaxBackOff}=5$, and $m=10$, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
+We also consider in these experiment the case of an ALOHA protocol using $\mathrm{MaxBackOff}=5$, and $m=10$, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
 The corresponding results are depicted in Figure~\ref{fig:43:mainExperiment2}.
 In this case the successful transmission rate is degraded compared with achieved results in Figure~\ref{fig:43:mainExperiment1}.
-We can explain this by the fact that we are considering in our network more devices, which increases the collision probability.
+We can explain this by the fact that we are considering in this network an increased number of devices, which increases the collision probability.
 It is important to highlight, that the ``Random'' retransmission heuristic shows a poor performance in comparison to the other heuristics, and it can be attributed to the fact that the number of retransmission is increased, and consequently a

### 2019-07-06 2-Chapters/4-Chapter/chapter4.tex
Small modifs
J'utilise moins "our ..."

@@
 % ----------------------------------------------------------------------------
-\section{Test-bed implementation of our model for real-world validation}
+% \section{Test-bed implementation of the first model for real-world validation}
+\section{Proof-of-concept of our model for real-world validation}
 \label{sec:4:gnuradio}
@@
 It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
-Unfortunately, analyzing our model with thousands of independent devices, all communicating and learning in their own (random) time scales, turned out to be of extreme difficulty.
-That is why we focus on two different directions of theoretical analysis in the next chapters of this thesis.
+Unfortunately, analyzing the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales, turned out to be of extreme difficulty.
+That is why we focus on two different simplifications of this model in the next chapters, for which we are able to provide clean theoretical analyses.
 % %

### 2019-07-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Still changing the macros
Lots of improvements on the arXiv papers
J'utilise moins "our ..."

@@
 We start by reviewing previous works on multi-players MAB models, which all considered the easier case of \emph{sensing feedback}.
-But in our model in this chapter, each device also uses a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
+But in the model studied in this chapter, each device also uses a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
 %
@@
 Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can have a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-players MAB problems).
-We do not propose any other efficient algorithm, but our work presented in April~$2018$ has strongly inspired two articles shortly after \cite{LugosiMehrabian18,BoursierPerchet18}.
+We do not propose any other efficient algorithm, but our work presented in April~$2018$ has strongly inspired two articles published shortly after \cite{LugosiMehrabian18,BoursierPerchet18}.
 They confirmed our findings that \Selfish{} can have a linear regret, and they both proposed new algorithms.
@@
 We first introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}.
-We present a useful regret decomposition in Section~\ref{sec:5:lowerbound}, and illustrate it in the Appendix~\ref{sec:5:appendix}.
+We define the regret, and then present a useful decomposition in Section~\ref{sec:5:lowerbound}.
+% and illustrate it in the Appendix~\ref{sec:5:appendix}.
 The \Selfish, \RandTopM{} and \MCTopM{} algorithms are introduced in Section~\ref{sec:5:algorithms},
-for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
-We report the results of our experimental study in Section~\ref{sec:5:experiments}.
+for which we present a theoretical analysis in Section~\ref{sec:5:upperbounds}.
+We report the results of an experimental study in Section~\ref{sec:5:experiments}.
 Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
@@
     However, the sensing and collision information are censored.
-    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications, but as explained above we did not present theoretical results for this third model (yet).
+    Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications.
+As explained above, this third model was found to be too general to analyze mathematically, and we did not present theoretical results of convergence in Chapter~\ref{chapter:4}.
     %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
@@
   The regret lower bound that we gave in \cite{Besson2018ALT} is not applicable to any algorithm,
-  as it was discovered by Boursier and Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
+  as it was discovered by V. Boursier and E. Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
   The proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step,
   but the results stated in this section are now correct,
-  thanks to the modifications proposed by Kaufmann and Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
+  thanks to the modifications proposed by E. Kaufmann and A. Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
   The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which \emph{should} contain \RhoRand, \RandTopM{} and \MCTopM, but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
@@
 Similarly to what we present above in Section~\ref{sec:2:lowerUpperBoundsRegret},
-we use the Kullback-Leibler divergence to express our lower bound.
+we use the Kullback-Leibler divergence to express this lower bound.
 We remind that $\kl(x,y) \eqdef x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
@@
 This intuitively suggests the number of players $M$ as the (multiplicative) \emph{``price of decentralized learning''}.
-However, to establish our regret bound, we lower bounded the number of collisions by zero, which may be too optimistic.
+However, to establish this regret bound, we lower bounded the number of collisions by zero, which may be too optimistic.
 %
@@
 
-To the best of our knowledge, no algorithm has been shown to experience only $o(\log(T))$ collisions so far,
+To the best of the authors' knowledge, no algorithm has been shown to experience only $o(\log(T))$ collisions so far,
 for every $M \in \{2,\dots,K\}$ and $\boldsymbol{\mu} \in \cP_M$.
@@
 % %
-The approaches we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained.
+The approaches we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to \UCB{} algorithms, for which strong theoretical guarantees can be obtained.
 In particular, we focus on two types of indices:
@@
 %
-We precisely state our first proposal below in Algorithm~\ref{algo:5:RandTopM}.
+We precisely state the first proposal below in Algorithm~\ref{algo:5:RandTopM}.
 %
@@
 %
-Our second proposal \MCTopM{} is stated below as Algorithm~\ref{algo:5:MCTopM},
+The second proposal \MCTopM{} is stated below as Algorithm~\ref{algo:5:MCTopM},
 it is a slightly more complex extension of the \RandTopM{} algorithm.
@@
       % \begin{equation}
-      $\sum\limits_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4 \mathrm{e} \log(\log(T))$.
+      $\sum\limits_{t=1}^T \Pr_{\bm{\mu}} \left(U_k^{j}(t) < \mu_k\right) \leq 3 + 4 \e \log(\log(T))$.
       % \end{equation}
@@
     & \leq \frac{\log(T) + 3 \log(\log(T))}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log(\log(T))}  \nonumber\\
-    & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4M \mathrm{e} \log(\log(T)) + 3M+1,
+    & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4M \e \log(\log(T)) + 3M+1,
   \end{align}
@@
 
-For our proposal \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
+For the proposal \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
 First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
@@
 which significantly improves over the $\bigO{M{2M-1 \choose M}}$ bound, proven by \cite{Anandkumar11} for \rhoRand.
-Our bound is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}. %, due to our trick of focusing on collisions for non-sitted players.
+The obtained bound is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}. %, due to our trick of focusing on collisions for non-sitted players.
 However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity.
@@
 The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
-Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
+Second, we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
 Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved that the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is indeed $\cO(1)$ if there is no arms in \Mworst{}, and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous state-of-the-art approaches using ranks.
@@
 %
-We also include our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} on regret plots.
+We also include the \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} on regret plots.
 

### 2019-07-06 2-Chapters/5-Chapter/chapter5.tex
J'utilise moins "our ..."

@@
 
-Some works studied the same model as our model \modeldeux{} ``with sensing'', under the hypothesis that players do not know in the value of $M$.
+Some works studied the same model as model \modeldeux{} ``with sensing'', under the hypothesis that players do not know in the value of $M$.
 As illustrated above, if we consider algorithms building on the same ideas as \RhoRand{} or \MCTopM, it seems mandatory to first build an estimate of the value of $M$ then run the initial algorithm that assumed a perfect knowledge of $M$.
@@
 
-In \cite{Rosenski16}, the authors explain that if the model allows arrival or departures of players at \emph{any time step}, then the game is much harder, and sub-linear regret is most likely un-achievable, if we consider the natural extension of the definition of regret as in our model presented in Section~\ref{sec:5:model} above.
+In \cite{Rosenski16}, the authors explain that if the model allows arrival or departures of players at \emph{any time step}, then the game is much harder, and sub-linear regret is most likely un-achievable, if we consider the natural extension of the definition of regret as in the model presented in Section~\ref{sec:5:model} above.
 But this negative results depends on how arrival of players are modeled:
@@
 We give in Table~\ref{table:5:meanRegretSimulationsNoSensing} the mean regret obtained for these different algorithms, and we observe a drastic difference between the unrealistic centralized algorithm, which achieves a very small regret, the heuristic \Selfish-\klUCB{} which achieve small mean regret (but is small to fail in theory and in some instances), and the two other algorithms.
-We were unable to find any bug in our implementation of Improved \MusicalChair{} from \cite{LugosiMehrabian18} and all the different tuning of $c$ (or $g$) explored gave the same disappointing result (linear regret).
+We were unable to find any bug in our implementation of Improved \MusicalChair{} from \cite{LugosiMehrabian18} and all the different tuning of $c$ (or $g$) explored gave the same disappointing result (\ie, linear regret).
 The \textsc{Sic-MMAB} algorithm performs better than the \Selfish{} heuristic for the extreme case of $M=K$, but its large regret in this case shows that the orthogonalization protocol developed by \cite{BoursierPerchet18} is not fast to converge (for illustration, for the same problem for $M=K=9$ for the ``sensing'' case, \MCTopM-\klUCB{} achieves a mean regret of about $40$, two orders of magnitude smaller!).
@@
 
-Similarly to our demonstration of MAB learning in an IoT network that we presented in the previous Chapter~\ref{chapter:4} in Section~\ref{sec:4:gnuradio},
-we list here some related works, dating back from $2016$, who proposed a similar approach for OSA.
-All the similar works that we are aware of also used USRP boards, and implemented the demonstration using either the Simulink and MATLAB softwares, or the GNU Radio software like we did \cite{Besson2018ICT}.
+Similarly to the proof-of-concept of applying MAB learning in an IoT network that we presented in Section~\ref{sec:4:gnuradio},
+% we list here some related works, dating back from $2016$, which proposed a similar approach for OSA.
+% All the similar works that we are aware of also used USRP boards, and implemented the demonstration using either MATLAB/Simulink or GNU Radio software like we did \cite{Besson2018ICT}.
 %
-Indeed, the demonstrations and the works presented below are coming from our team experiments, and have been extended by the team of S.
+% Indeed, the demonstrations and the works presented below are coming from our team experiments, and have been extended by the team of S.
 Darak at IIIT Delhi after his postdoctoral fellowship in our SCEE team.
 %
-In \cite{darak2016bayesian,Darak16}, the authors study the same model and focused on the \RhoRand{} policy, which was the state-of-the-art back in $2016$, combined with \UCB{} or \klUCB, as well as an extension using Bayes-UCB from \cite{Kaufmann12BUCB}.
-
-Their idea is to use the same skeleton as \RhoRand{} \cite{darak2016bayesian}, that is to assign a \emph{rank} to each player, and make players change their rank after any collision.
+in \cite{darak2016bayesian,Darak16}, the authors study the same model and focused on the \RhoRand{} policy, for the OSA case.
+It was the state-of-the-art back in $2016$, and the authors combine it with \UCB{} or \klUCB, as well as an extension using Bayes-UCB from \cite{Kaufmann12BUCB}.
+%
+Their idea is to use the same skeleton as \RhoRand, that is to assign a \emph{rank} to each player, and make players change their rank after any collision.
 But instead of selecting a new rank uniformly at random among $[M]$, they proposed to use a ``second-stage'' learning policy, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.

### 2019-07-06 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
J'utilise moins "our ..."

@@
 We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
-To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or \ExpThree, with a strategy to account for changes in arms' distributions.
+To the best of the authors' knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or \ExpThree, with a strategy to account for changes in arms' distributions.
 Following the vocabulary used in previous works, we make the distinction between \emph{passively} and \emph{actively} adaptive strategies.
@@
 \ExpThreeS{} actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
-It constitues a good baseline for our numerical experiments in Section~\ref{sec:6:NumericalExperiments},
+It constitues a good baseline for the numerical experiments in Section~\ref{sec:6:NumericalExperiments},
 even if we did find that \ExpThreeS{} performs worse that the most of the other approaches based on extending stationary bandit algorithms (\eg, SW-UCB).
@@
 The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
-and we found in our experiments that \PHT-\UCB{} performs very similarly to \CUSUMUCB.
+and we found in numerical experiments that \PHT-\UCB{} performs very similarly to \CUSUMUCB.
 Even if it has the advantage of not requiring a parameter $M$,
@@
 %
-Then, the Ada-ILTCB algorithm is however a less appealing candidate for our setting, for the following reason.
+Then, the Ada-ILTCB algorithm is however a less appealing candidate for the setting of this chapter, for the following reason.
 This algorithm requires as input a parameter $L$ which needs to be an upper bound on the largest stationary sequence in the problem, in order for its regret to scale as $\Upsilon_T \sqrt{L}$ (neglecting $K$ and $\log$ factors).
@@
 In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
-Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
+Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of the proposed algorithm.
 % This study is supported by both theoretical and empirical results presented in this chapter.
@@
 
-Moreover, we can use the two problems $1$ and $2$ to illustrate the expected empirical behavior of our proposal \GLRklUCB.
+Moreover, we can use the two problems $1$ and $2$ to illustrate the expected empirical behavior of \GLRklUCB.
 Problem $1$ has only \emph{local changes}, meaning that at any change-point, only one arm sees its distribution change, while problem $2$ has only \emph{global changes}, meaning that all arms see their distribution change at each change-point.
@@
 Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
-As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$.
+As explained in Section~\ref{par:2:normalizationTrick}, this work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$.
 We assume that both the decision maker and the algorithm know beforehand the values of $a$ and $b$.
@@
 
-We now present our \GLRklUCB{} algorithm, which combines a bandit algorithm with a change-point detector running on each arm.
+We now present the \GLRklUCB{} algorithm, which combines a bandit algorithm with a change-point detector running on each arm.
 It also needs a third ingredient, some forced exploration parameterized by $\omega\in(0,1)$ to ensure that each arm is sampled enough and in order to also be able to detect changes on arms currently under-sampled by the bandit algorithm.
@@
 %
-Our proposal \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
+\GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs a uniform random exploration.
 Both approaches are parameterized by $\omega\in(0,1)$.
@@
 
-This section gives the finite-time regret bounds for the two variants, and interpretations of our results.
+This section gives the finite-time regret bounds for the two variants, and interpretations of both results.
 We include only one proof, as the other is given in \cite{Besson2019GLRT}.
@@
 
-Assumption~\ref{ass:6:LongPeriods} is easy to interpret, as Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard the literature, and to the best of or knowledge, our analysis is the first one to consider such assumption on the problem difficulty.
+Assumption~\ref{ass:6:LongPeriods} is easy to interpret, as Assumption~\ref{ass:6:LongPeriodsGlobal}, but it is non standard the literature, and to the best of the authors' knowledge, this analysis is the first one to consider such assumption on the problem difficulty.
 %
@@
 
-The regret bounds we obtain for the two variants of our proposal \GLRklUCB,
+The regret bounds we obtain for the two variants of \GLRklUCB,
 in Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} respectively,
@@
 
-This section starts by giving a unified analysis of regret of the two variants of our algorithm, then we give the proof of the finite-time regret bound for the variant using \emph{global restarts}.
+This section starts by giving a unified analysis of regret of the two variants of \GLRklUCB, then we give the proof of the finite-time regret bound for the variant using \emph{global restarts}.
 The proof of the other variant using \emph{local restarts} follows the skeleton of the unified analysis, and it can be found in Appendix~E of \cite{Besson2019GLRT}.
@@
 Finally by controlling terms $(a)$ and $(b)$, the ``bad event'' is unlikely: $\bP(\cE_T^c) \leq 2\EffChange \delta$.
-By putting together the three pieces, that are a bound on $(A)$, $(B)$ and $\bP(\cE_T^c)$, we obtain the desired finite-time upper-bound on the regret of our proposal \GLRklUCB.
+By putting together the three pieces, that are a bound on $(A)$, $(B)$ and $\bP(\cE_T^c)$, we obtain the desired finite-time upper-bound on the regret of \GLRklUCB.
 
@@
 
-Our main goal is to compare against \emph{actively adaptive algorithms}.
+The main goal here is to compare against \emph{actively adaptive algorithms}.
 We include \CUSUMklUCB, tuned with $M=150$ and $\varepsilon=0.1$ for easy problems ($1,2,4$) and $\varepsilon=0.001$ for hard problems ($3,5$),
@@
 % as a simple adaptation of the \MUCB algorithm,
-tuned with $w=150$, based on a prior knowledge of the problems as the formula using $\delta_{\min}$ given in \cite{CaoZhenKvetonXie18} is too large for small horizons (on all our problem instances), a threshold $b=\sqrt{w \log(2 K T)}$ and $\gamma=\sqrt{\Upsilon_T K (2 b + 3 \sqrt{w}) / (2 T)}$ as suggested by Remark~4 in the paper \cite{CaoZhenKvetonXie18}.
+tuned with $w=150$, based on a prior knowledge of the problems as the formula using $\delta_{\min}$ given in \cite{CaoZhenKvetonXie18} is too large for small horizons (on all the problem instances), a threshold $b=\sqrt{w \log(2 K T)}$ and $\gamma=\sqrt{\Upsilon_T K (2 b + 3 \sqrt{w}) / (2 T)}$ as suggested by Remark~4 in the paper \cite{CaoZhenKvetonXie18}.
 
@@
 
-In these two examples, we compare the two variants of our proposal with the two state-of-the-art policies \CUSUMklUCB{} and \MklUCB,
+In these two examples, we compare the two variants of \GLRklUCB, with the two state-of-the-art policies \CUSUMklUCB{} and \MklUCB,
 that are all based on combining \klUCB{} with a change-point detection algorithm.
@@
 But it is still interesting to compare the final regret in one run, as given above, with the mean regret for $1000$ independent runs, as given below in Table~\ref{table:6:totalResults1}.
-The values change, but the ranking stays the same: our proposal outperforms both \CUSUMklUCB{} and \MklUCB, and the test based on \CUSUM{} seems more efficient than the test based of \MklUCB{} in the problems at hand.
+The values change, but the ranking stays the same: \GLRklUCB{} outperforms both \CUSUMklUCB{} and \MklUCB, and the test based on \CUSUM{} seems more efficient than the test based of \MklUCB{} in the problems at hand.
 
@@
 Results highlighted in \textbf{bold} show the best non-oracle algorithm for each experiment,
-with our proposal being the best non-oracle strategy for problems $1$ and $2$.
+with \GLRklUCB{} being the best non-oracle strategy for problems $1$ and $2$.
 Thompson sampling and \klUCB{} are efficient,
@@
 
-On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there are more changes, we observe that our approach \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much better than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
+On simple problems, like problem $1$, all the algorithms being designed for piece-wise stationary environments perform similarly, but as soon as the gaps are smaller or there are more changes, we observe that \GLRklUCB{} can outperform the two other actively adaptive algorithms \CUSUMklUCB{} and \MklUCB{} (\eg, on Figure~\ref{fig:6:meanRegretPb2}), and performs much better than passively adaptive algorithms DTS and SW-\klUCB{}  (\eg, on Figure~\ref{fig:6:meanRegretPb3}).
 % Our approach, with the two options of \textbf{Local} or \textbf{Global} restarts, performs very closely to the oracle for problem $4$.

### 2019-07-06 2-Chapters/Conclusion/conclusion.tex
J'utilise moins "our ..."

@@
     % \item
-The second half of this thesis, Part~\ref{part:MABIOT}, presented our main contributions.
+The second half of this thesis, Part~\ref{part:MABIOT}, presented the main contributions.
 % Chapter 4
@@
 In two different models, with or without retransmissions, we showed that end-devices can independently learn to increase their successful transmission rates in the network, by using low-cost MAB algorithms.
-We also presented a proof-of-concept that implements such network on real radio hardware and validates our model and proposed algorithms.
+We also presented a proof-of-concept that implements such network on real radio hardware and validates the proposed model and algorithms.
 % Chapter 5
@@
 %
-For the last two chapters, our proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in our theoretical analyses.
+For the last two chapters, the proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in our theoretical analyses.
 % \end{itemize}
@@
 %
-Our library allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
+This library allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
 SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license (MIT License), and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
@@
 
-Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of our model of Low-Power Wide-Area (LPWA) networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency, and assuming a perfect synchronisation.
+Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of the previous model of Low-Power Wide-Area (LPWA) networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency, and assuming a perfect synchronisation.
 % , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
@@
 Thus two possible directions of future research could be to extend our works
-to MAB models with either multidimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
+to MAB models with either multi-dimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
 The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.
@@
 
-Our study on multi-players bandits suggests several further research directions.
+The study on multi-players bandits suggests several further research directions.
 First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
@@
 We also presented many extensions of the multi-players bandit model,
-and even if some have already been implemented, a major future work is to implement the most interesting ones in our library SMPyBandits
+and even if some have already been implemented, a major future work is to implement the most interesting ones in SMPyBandits
 (see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).

### 2019-07-06 1-Preamble/5-Abstract/abstract.tex
Not sure if I say enough about the lower-bounds...

@@
 % Chapitre 1
+\noindent
 Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de type radio intelligente, dans des bandes non licenciées et sans supervision centrale.

### 2019-07-06 1-Preamble/5-Abstract/resume_fr.tex
Not sure if I say enough about the lower-bounds...

@@
 
-    \item
-    La dernière contribution de cette thèse est notre article \cite{Besson2018DoublingTricks}.
-    % , résumé en Annexe~\ref{app:2:DoublingTricks}.
-    Cet article donne une revue de la littérature sur l'utilisation de la technique de doublements successifs de l'horizon (``doubling trick'') pour les problèmes de bandits,
-    ainsi qu'une analyse unifiée et plus générique de deux familles de doublements.
+    % \item
+    % La dernière contribution de cette thèse est notre article \cite{Besson2018DoublingTricks}.
+    % % , résumé en Annexe~\ref{app:2:DoublingTricks}.
+    % Cet article donne une revue de la littérature sur l'utilisation de la technique de doublements successifs de l'horizon (``doubling trick'') pour les problèmes de bandits,
+    % ainsi qu'une analyse unifiée et plus générique de deux familles de doublements.
 \end{itemize}

### 2019-07-06 2-Chapters/1-Chapter/chapter1.tex
Not sure if I say enough about the lower-bounds...

@@
 %
-We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBanditsJMLR,SMPyBandits} and hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
+To the best of our knowledge, we wrote the most exhaustive open-source library for bandits.
+Our library SMPyBandits is published online, under an open-source licence \cite{SMPyBanditsJMLR,SMPyBandits}, and hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
 We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
@@
     \item
-    We developed SMPyBandits, the most exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence (MIT) \cite{SMPyBanditsJMLR,SMPyBandits}.
+    We developed SMPyBandits, an exhaustive open-source simulation library for MAB problems, that we published on-line under an open-source licence (MIT) \cite{SMPyBanditsJMLR,SMPyBandits}.
     We present it in details in Chapter~\ref{chapter:3}.

### 2019-07-06 2-Chapters/2-Chapter/chapter2.tex
Not sure if I say enough about the lower-bounds...

@@
     % with two options (A/B testing), or more than two options.
-    The seminal work of \cite{Li10} studies the application of contextual bandit to online advertisement
-    or news articles recommendation, as it is used in practice on platforms such as Microsoft's Bing news website,
-    or in video streaming platform like YouTube or Netflix.
-    In such models, the arms correspond to items to recommend (\eg, articles or movies), and the contexts contain features about each user of the system.
+    The seminal work of \cite{Li10} studies the application of contextual bandit to news articles recommendation, as it is used in practice on platforms such as Microsoft's Bing news website.
+    Other famous examples include online advertisement or in other kinds of content recommendation, such as for video streaming platforms like YouTube or Netflix.
+    In such models, the arms correspond to items to recommend (\eg, ads, articles or movies), and the contexts contain features about each user of the system.
     An interesting work is \cite{Louedec16}, who studies slowly-varying non-stationary models applied to recommender systems.
@@
 \paragraph{Problem-dependent lower-bound in $\Omega(\log T)$.}
+\label{par:2:def_optimal_orderoptimal}
 %
@@
 These lower-bounds are of highest interest to design efficient algorithms,
-and a significant part of the research on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically.
-This means that a finite-time or an asymptotic upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound, with the same constant in the big-$\cO$ notation (in which case we say that the algorithm is asymptotically \emph{optimal}), or with a larger constant (the algorithm is said to be \emph{order-optimal} in such case).
+and a significant part of the research on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically \cite{LaiRobbins85}.
+%
+This means that a finite-time or an asymptotic upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound.
+When the regret upper-bound matches the lower-bound with the same constant as in the big-$\cO$ notation, we say that the algorithm is asymptotically \emph{optimal},
+otherwise the algorithm is said to be \emph{order-optimal} if it matches the lower-bound with with a larger constant.
 
@@
 Theorem~2.12 from \cite{Slivkins2019} specifies a possible value for the constant $C_I$.
+Note that the lower-bound of Lai and Robbins, proved in \cite{LaiRobbins85}, is more general and applies to any one-dimensional exponential family $\cE$, by replacing the Bernoulli Kullback-Leibler divergence $\kl$ with the KL divergence of this family, $\kl_{\cE}$.
 
@@
 
-Moreover, it is interesting to note that the first lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Proposition~\ref{prop:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
+It is interesting to note that the first lower-bound of Theorem~\ref{thm:2:forSecondLogTLowerBound} can be directly used to design efficient algorithms\footnote{~Tracking this quantity, by using empirical estimates of the means, is used for instance in the OSSB algorithm proposed in \cite{Combes17}, which is proven to attain the lower-bound for a wider range of problems (for problems called structured stochastic bandits).}, as thanks to the regret decomposition given in Proposition~\ref{prop:2:RegretDecomposition} above, the expression of $C_I$ essentially says that any efficient algorithm
 should sample each sub-optimal arm $k$ about $\log(T)/\kl(\mu_k,\mu^*)$ times in the total of $T$ time steps.
@@
 Many algorithms have been proven to achieve logarithmic regret in the stochastic case,
-and in particular it is the case of the algorithms used in this thesis, $\UCB_1$ from \cite{Auer02}, Thompson sampling from \cite{Thompson33} and analyzed in \cite{AgrawalGoyal11,Kaufmann12Thompson}, and \klUCB{} from \cite{Garivier11KL,KLUCBJournal}.
-%
-Such bounds are valid in different settings, and in particular $\UCB_1$ is order-optimal for bounded rewards or one-dimensional exponential families,
-while Thompson sampling is optimal for bounded rewards, and \klUCB{} have been proven to be optimal for both cases.
+and in particular it is the case of the algorithms used in the rest of thesis, $\UCB_1$ from \cite{Auer02}, Thompson sampling from \cite{Thompson33} and analyzed in \cite{AgrawalGoyal11,Kaufmann12Thompson}, and \klUCB{} from \cite{Garivier11KL,KLUCBJournal}.
+%
+Such bounds are valid in different settings:
+$\UCB_1$ is order-optimal for bounded rewards or one-dimensional exponential families (\ie, it matches the asymptotic lower-bound of \cite{LaiRobbins85}),
+while Thompson sampling is optimal for binary rewards, and \klUCB{} and \KLUCB{} have been proven to be optimal respectively for both cases.
 %
@@
 The strategy is then referred to as ``explore-then-commit'' (ETC), and the idea is to use a statistical test at every time step $t$, and stop as soon as enough samples were collected to effectively identify the best arm with a certain confidence level $\delta$.
-Choosing $\delta \propto 1/T$ and using a lower-bound $d$ on the gaps typically lead to an order-optimal algorithm as shown in \cite{GarivierETC2016}.
+Choosing $\delta \propto 1/T$ and using a lower-bound $d$ on the gaps typically lead to an order-optimal algorithm, as shown in \cite{GarivierETC2016}.
 
@@
 %
-This theorem shows that $\UCB_1$ achieves an order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta \eqdef \min\limits_{k: \Delta_k>0} \Delta$), and the result also has the advantage of being non asymptotic.
+% This theorem shows that $\UCB_1$ achieves an order-optimal problem-dependent regret, bounded by $\cO(K\log(T)/\Delta^2)$ (if $\Delta \eqdef \min\limits_{k: \Delta_k>0} \Delta$), and the result also has the advantage of being non asymptotic.
+This has the advantage of being non asymptotic.
 
@@
 
-We see in this Theorem~\ref{thm:2:UCBregretBound} that $\UCB_1$ has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime and computationally simple, for both memory and time aspects, and it achieves an order-optimal problem-dependent regret upper bound in $\cO((\sum_{k: \Delta_k>0} \frac{1}{\Delta_k})\log(T))$.
+We see in this Theorem~\ref{thm:2:UCBregretBound} that $\UCB_1$ has the advantage of not relying on any prior knowledge of the problem difficulty, of being anytime and computationally simple, for both memory and time aspects.
+Moreover, $\UCB_1$ achieves an order-optimal problem-dependent regret upper bound in $\cO((\sum_{k: \Delta_k>0} \frac{1}{\Delta_k})\log(T))$, meaning that it matches the asymptotic regret lower-bound of Lai and Robbins up-to a constant factor \cite{LaiRobbins85} (see the definition above in page \pageref{par:2:def_optimal_orderoptimal}).
 %
@@
 %
-The \KLUCB{} and \klUCB{} algorithms achieve finite-time regret logarithmic upper-bounds on their regret, and have been proven to be asymptotically optimal, respectively for one-dimensional exponential families and for bounded rewards distributions \cite{Garivier11KL,KLUCBJournal}.
-% FIXME
+The \KLUCB{} and \klUCB{} algorithms achieve finite-time regret logarithmic upper-bounds on their regret, and have been proven to be asymptotically optimal, respectively for one-dimensional exponential families and for binary rewards \cite{Garivier11KL,KLUCBJournal}.
+In particular, we highlight that \klUCB{} achieves a finite-time regret upper-bound, that matches the asymptotical lower-bound of Lai and Robbins \cite{LaiRobbins85} for Bernoulli distributed problems, which is our main focus of interest in both Chapters~\ref{chapter:5} and \ref{chapter:6}.
 
@@
 %
-In this setting, as well as any one-dimensional exponential families,
+In this setting, as well as for one-dimensional exponential families,
 % for Gaussian distributions,
-Thompson sampling was proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
+Thompson sampling was proven to achieve finite-time and asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
 
@@
 Instead of sampling from the posteriors and playing the arms with maximum sample, Bayes-UCB computes the $(1-1/(t (\log(t))^5)$ quantile of each arm, at time $t$, and plays the arm with the largest quantile.
-It was also proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12BUCB}.
+It was also proven to achieve an asymptotical optimal problem-dependent bound for binary (Bernoulli) bandits \cite{Kaufmann12BUCB}.
 %
@@
 It was analyzed partially, for the two-armed case it was proven to be asymptotically optimal for one-dimensional exponential families, and the authors illustrated that it can be very efficient empirically.
-Analyzing it in the generic case was found to be difficult, and it is still an open problem.
+However, analyzing it in the generic case was found to be difficult, and it is still an open problem.
 % We implemented it using a binary tournament implement in a naive but efficient way (not using recursive functions, \ie, using the \texttt{BESA} class with \texttt{''non\_recursive''=True}).
@@
 More recently, the \klUCB-switch from \cite{GarivierHadiji2018} is the first algorithm to be proven to
-enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, that is, matching the $C_I \ln(T)$ lower bound by \cite{LaiRobbins85}.
+enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, by matching the $C_I \ln(T)$ asymptotic lower-bound by \cite{LaiRobbins85}.
 It is an index policy that uses modified versions of the indexes of \klUCB$^+$ for arms that are not sampled much (see \cite{Honda2019} for more details), and of MOSS$^+$ for the other arms.
@@
 focussing on algorithms designed and analyzed for the single-player stationary and stochastic MAB model.
-Except the naive and simple strategies given as introductory examples, most of the policies presented above are order-optimal for Bernoulli distributed problems or bounded rewards (or one-dimensional exponential families),
+Except the naive and simple strategies given as introductory examples, most of the policies presented above are order-optimal for at least one of the model of Bernoulli distributed problems, bounded rewards or one-dimensional exponential families,
 and some of them are known to be optimal in different settings.

### 2019-07-06 2-Chapters/3-Chapter/chapter3.tex
Not sure if I say enough about the lower-bounds...

@@
 It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
-It is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
+To the best of the authors' knowledge, it is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
 %
@@
 %
-This library is by far the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
-It aims at being exhaustive, simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
+This library is by far the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models, according to the authors' knowledge.
+It aims at being simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
 It allows fast prototyping of simulations, with an easy configuration system and command-line options to customize experiments.

### 2019-07-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Improved all mentions to the (wrong) lower-bound of Ch6

@@
 Like in previous works, we consider the centralized system regret (multi-players regret), or simply referred to as regret.
-We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class, including previous solutions such as \rhoRand{} and our two solutions.
-We then analyze the \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic and order-optimal, improving over the previous state-of-the-art.
+We start by showing an improved asymptotical regret lower-bound for any algorithm of a certain class.
+% , including previous solutions such as \rhoRand{} and our two solutions.
+We then analyze the \MCTopM{} algorithm and we show that its regret upper-bound is logarithmic, improving over the previous state-of-the-art.
 %  with a regret bounded by $\bigO{\log(T)}$ for a game at horizon $T$.
@@
 
-\textbf{Only Bernoulli.}
-However we mention that our lower bound and all our algorithms (and their analysis) can be easily extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}).
-For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio problems, by considering channel availability.
+\textbf{Generalization to other distributions.}
+However we mention that our lower bound and all our algorithms (and their analysis) could be extended to one-dimensional exponential families (just like for the \klUCB{} algorithm of \cite{KLUCBJournal}).
+For simplicity, we focus on the Bernoulli case, that is also the most relevant for Cognitive Radio problems, by considering channels' availabilities as a source of binary rewards.
 
@@
     Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications.
-As explained above, this third model was found to be too general to analyze mathematically, and we did not present theoretical results of convergence in Chapter~\ref{chapter:4}.
+As explained above, this third model was found to be too general to analyze mathematically, and we did not present theoretical results of convergence of the algorithms considered in Chapter~\ref{chapter:4}.
     %\footnote{\cite{Bonnefoi17} considers a different setting, more suited for large-scale IoT networks, and a future work will be to extend our work to this setting with more than $K$ players but small probability of activations.}.
@@
 
-  Note that they are both uniquely defined if $\mu_M^* > \mu_{M+1}^*$.
+  Note that they are both uniquely defined whenever $\mu_M^* > \mu_{M+1}^*$.
 \end{leftbar}  % XXX leftbar defnbar, comment if needed
@@
   The regret lower bound that we gave in \cite{Besson2018ALT} is not applicable to any algorithm,
-  as it was discovered by V. Boursier and E. Perchet, as they explain in Section~2.4 in \cite{BoursierPerchet18}.
-  The proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step,
-  but the results stated in this section are now correct,
-  thanks to the modifications proposed by E. Kaufmann and A. Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
+  as it was discovered by V. Boursier and E. Perchet.
+  As they explain in Section~2.4 in \cite{BoursierPerchet18},
+  the proof we gave in the Appendix of our paper \cite{Besson2018ALT} was wrong in just one step.
+  The results stated in this section are now correct,
+  thanks to the modifications that we added here, that is, to consider this information term $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T)$.
+  % thanks to the modifications proposed by E. Kaufmann and A. Mehrabian in Appendix~E of \cite{KaufmannAbbas19}.
   The cost for this modification is to restrict the lower-bound to a smaller class of algorithms, which \emph{should} contain \RhoRand, \RandTopM{} and \MCTopM, but not the two algorithms proposed SIC-MMAB from \cite{BoursierPerchet18} and Multiplayer Explore-then-Commit (M-ETC) from \cite{KaufmannAbbas19}.
@@
 Similarly to what we present above in Section~\ref{sec:2:lowerUpperBoundsRegret},
-we use the Kullback-Leibler divergence to express this lower bound.
-We remind that $\kl(x,y) \eqdef x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
+we use the Kullback-Leibler divergence $\kl$ to express this lower bound.
+% We remind that $\kl(x,y) \eqdef x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
 %
@@
 we now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections.
-The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
+The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T)$ essentially says that ``the collisions do not bring too much information on the arm means''.
+% as stated in \cite{KaufmannAbbas19}.
 The theorem stated below is proven in the Appendix of \cite{Besson2018ALT},
@@
 
-Observe that the regret lower bound \eqref{eq:5:ourLowerBound} is tighter than the state-of-the-art lower bound in this setup
-given by \cite{Zhao10}, that states that
+Observe that the regret lower bound \eqref{eq:5:ourLowerBound} is tighter than the state-of-the-art lower bound in this setup, given by \cite{Zhao10}, that states that
 \begin{equation}\label{eq:5:Zhao10LowerBound}
@@
 Our result could be extended to this case, but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.
+%
+We also highlight that the proof of the bound of \cite{Zhao10} contained the same mistake as our proof in \cite{Besson2018ALT}, but by using the same hypothesis and modification in the proof, it is applicable to the same class of algorithms as our Theorem~\ref{thm:5:BetterLowerBound} (\ie, algorithms $\cA$ such that $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$).
 
@@
 
-To the best of the authors' knowledge, no algorithm has been shown to experience only $o(\log(T))$ collisions so far,
+To the best of the authors' knowledge, no algorithm has been shown to experience only $o(\log(T))$ collisions in expectation so far,
 for every $M \in \{2,\dots,K\}$ and $\boldsymbol{\mu} \in \cP_M$.
-%
-Since our article \cite{Besson2018ALT}, we kept as a future work the following question.
-A lower bound on the minimal number of collisions experienced by any strongly uniformly efficient decentralized algorithm would thus be a nice complement to our Theorem~\ref{thm:5:BetterLowerBound}.
+The best candidate would be the \textsc{Sic-MMAB} algorithm \cite{BoursierPerchet18}, whose number of collisions is proven to be $\bigO{(\log(T))^2} = o(\log(T))$ in high probability (Lemma~8), but not $o(\log(T))$ in expectation.
+%
+Since our article \cite{Besson2018ALT}, we kept as a future work the question
+of finding a lower bound on the minimal number of collisions experienced by any strongly uniformly efficient decentralized algorithm.
+Such result would thus be a nice complement to our Theorem~\ref{thm:5:BetterLowerBound}.
 
@@
 Regardless of whether sensing is or not possible, we start by presenting formally in Section~\ref{sub:5:Selfish} the \Selfish{} heuristic, as it was used by all the IoT devices in the first model presented in Chapter~\ref{chapter:4}.
-It does not use an orthogonalization strategy as the collisions are directly accounted for in the \UCB-like indices that are used by each device to select its channel.
-\Selfish{} can also be used under observation model \modeltrois{} --\emph{without sensing}--, and without the knowledge of $M$.
+It does not use an orthogonalization strategy as the collisions are directly accounted for in the \UCB-like indices that are used by each device to select its channel (\ie, by each player to select its arm, in the bandit vocabulary).
+\Selfish{} can also be used under observation model \modeltrois{} --\emph{without sensing}, and without the knowledge of $M$.
 It was conjectured in \cite{Besson2018ALT} that \Selfish{} can suffer linear regret, and later confirmed in \cite{LugosiMehrabian18,BoursierPerchet18}.
@@
 \begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
-  \textcolor{red}{\textbf{Warning.}}
+  \textcolor{red}{\textbf{Remark.}}
   We want to strengthen the fact that in all the proposed algorithms,
@@
 %
-This is a sanity check: the lower-bound of
-Theorem~\ref{thm:5:BetterLowerBound} indeed applies to our algorithm \MCTopM,
-and finally this highlights that it is order-optimal for the regret, in the sense that it matches the lower-bound up-to a multiplicative constant,
-and optimal for the term \ref{eq:5:term1}.
+% This is a sanity check: the lower-bound of
+% Theorem~\ref{thm:5:BetterLowerBound} indeed applies to our algorithm \MCTopM,
+% and finally this highlights that it is order-optimal for the regret, in the sense that it matches the lower-bound up-to a multiplicative constant,
+% and optimal for the term \ref{eq:5:term1}.
 

### 2019-07-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for improving Ch6

@@
   First, we observe that, for all $k\in \Mbest$,
-  \begin{eqnarray*}
-    T - \E_{\mu}[N_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
-    & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
-  \end{eqnarray*}
+  % \begin{eqnarray*}
+    $T - \E_{\mu}[N_k(T)] \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right]$,
+  % \end{eqnarray*}
   where we denote by $\widehat{S}_t = \{A^j(t), j \in [M]\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
@@
 \end{equation}
-where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
+where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory (\eg, $f(t)=\log(t)+3\log(\log(t))$), which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
 A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
@@
 for our two proposals \RandTopM{} and \MCTopM{} as well as for \rhoRand{}, when they are combined with \klUCB{} indices,
-and then we restrict to the most efficient algorithm \MCTopM, by proving in Section~\ref{sub:5:UpperBoundCollisions} that the number of collisions are logarithmic, and hence the regret of \MCTopM{} is also logarithmic.
-%
-Finally, Section~\ref{sub:5:SelfishFails} shortly discusses a disappointing result regarding \Selfish.
+and then we restrict to the most efficient algorithm \MCTopM, by proving in Section~\ref{sub:5:UpperBoundCollisions} that it achieves a logarithmic number of collisions and regret.
+%
+Finally, Section~\ref{sub:5:SelfishFails} shortly discusses a negative result regarding \Selfish.
 %  with more insights provided in Appendix~\ref{app:5:SelfishFails}.
@@
     For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $U_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
-    \begin{small} % XXX
-      \begin{align*}
-        &\sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq U_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
-        + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.
-      \end{align*}
-    \end{small} % XXX
+    % \begin{small} % XXX
+      \begin{align}
+        \sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq U_k^{j}(t) \right)
+        & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
+        + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} \nonumber\\
+        & + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.
+      \end{align}
+    % \end{small} % XXX
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
   it holds that
-  \begin{equation*}
-    \sum_{t=1}^T
-    \Pr\left(A^j(t)=k, k\notin \TopM(t)\right)
-    = \left(\sum_{k', \mu_{k'} < \mu_k}\frac{1}{\kl(\mu_k,\mu_{k'})} + \sum_{k', \mu_{k'} > \mu_k}\frac{1}{\kl(\mu_{k'},\mu_{k})}\right) \log(T) + o(\log(T)).
-  \end{equation*}
+  \begin{align}
+    \sum_{t=1}^T \Pr\left(A^j(t)=k, k\notin \TopM(t)\right) =
+    & \left( \sum_{k', \mu_{k'} < \mu_k}\frac{1}{\kl(\mu_k,\mu_{k'})} \right) \log(T) \nonumber\\
+    & \left( \sum_{k', \mu_{k'} > \mu_k}\frac{1}{\kl(\mu_{k'},\mu_{k})}\right ) \log(T)
+    + o(\log(T)).
+  \end{align}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 
-For the proposal \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
+For \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
 First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
@@
 The obtained bound is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}. %, due to our trick of focusing on collisions for non-sitted players.
-However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity.
-The \MusicalChair{} algorithm is indeed not very satisfying from an applicative point of view, as it requires a prior knowledge of $\mu^*_{M}-\mu^*_{M+1}$.
+However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity,
+at it is indeed not very satisfying from an applicative point of view to require a prior knowledge of $\mu^*_{M}-\mu^*_{M+1}$.
 
@@
 We controlled the total number of transitions $(2)$, $(3)$ and $(5)$ (see Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM}),
-which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
-Thus, the total number of arm switches is also proven to be logarithmic, if all players uses the \MCTopM-\klUCB{} algorithm.
-This additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
+which are the only transitions when a player can change its arm.
+%  switch from arm $k$ to arm $k'\neq k$.
+Thus, the total number of arm switches is also logarithmic, if the $M$ players use \MCTopM-\klUCB.
+This additional guarantee was never clearly stated for previous works, like \rhoRand.
 %
@@
   % \begin{subfigure}[!h]{0.49\textwidth}
-    \includegraphics[width=0.95\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
+    \includegraphics[width=1.05\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
   % \end{subfigure}

### 2019-07-06 1-Preamble/5-Abstract/resume_fr.tex
Done for tonight

@@
 % https://en.wikipedia.org/wiki/History_of_radio
-Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, les communications sans fil entre les appareils fabriqués ont été rendues possibles, et de plus en plus fréquentes dans nos vies.
-Avec l'avènement des réseaux Internet des objets (IdO), des milliards de dispositifs autonomes basse consommation devraient être déployés dans le monde entier, pour un large éventail d'applications différentes.
+Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, la télé-communication sans fil entre des appareils fabriqués a été rendue possible, et de plus en plus fréquentes dans nos vies.
+Avec l'avènement des réseaux Internet des Objets (IdO), des milliards de dispositifs autonomes à basse consommation devraient être déployés dans le monde entier, pour un large éventail d'applications différentes.
 Il existe aujourd'hui un consensus mondial sur le fait qu'avec la tendance actuelle à l'accroissement démographique et la crise énergétique actuelle, toute nouvelle technologie déployée doit être à la fois bon marché et efficace sur le plan énergétique,
-ainsi qu'adaptés pour desservir un grand nombre de personnes et d'appareils.
+ainsi qu'adaptée pour desservir un grand nombre de personnes et d'appareils.
 %
-Ces technologies de l'IdO devraient pouvoir s'adapter automatiquement à différents environnements et contextes d'application, et être aussi efficaces que possible.
+Ces technologies de l'IdO devraient pouvoir s'adapter automatiquement à différents environnements et différents contextes d'application, et être aussi efficaces que possible.
 %
@@
 C'est pourquoi nous nous intéressons à cette thèse sur
-l'amélioration de la durée de vie des batteries des équipements terminaux IoT et la réduction du coût énergétique des réseaux IoT.
+l'amélioration de la durée de vie des batteries des équipements terminaux IdO et la réduction du coût énergétique des réseaux IdO.
 Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant la prise de décision décentralisée à faible coût directement dans les futurs dispositifs IdO.
 %
-Notre thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (Multi-Armed Bandit algorithms), afin de permettre aux appareils IoT d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
+Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (Multi-Armed Bandit algorithms), afin de permettre aux appareils IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
 
@@
 Ce troisième type de systèmes peut être conçu comme décentralisé :
-même si une station de base centrale est
-toujours en charge de nombreux appareils,
+même si une station de base centrale est toujours en charge de nombreux appareils,
 les appareils déclenchent l'envoi de paquets de liaison montante, et les seules données de liaison descendante qu'ils peuvent recevoir sont de courts accusés de réception, envoyés par la station de base pour indiquer le succès ou l'échec de chaque paquet de liaison montante.
-Cette famille de systèmes sans fil est appelée Internet des objets (IoT),
-et un exemple typique d'application de tels réseaux IoT est celui des réseaux sans fil de capteurs.
+Cette famille de systèmes sans fil est appelée Internet des Objets (IdO), ou \emph{Internet of Things} (IoT),
+et un exemple typique d'application de tels réseaux IdO est celui des réseaux sans fil de capteurs.
 
-Pour le développement futur des ``réseaux intelligents'', des ``villes intelligentes'', des ``maisons intelligentes'' ou de l'''agriculture intelligente'', les réseaux de capteurs devraient être largement déployés.
-Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les ``bâtiments connectés'' et l'''agriculture connectée''.
-Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides et d'utiliser des réseaux de capteurs pour obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, et de permettre au contrôle centralisé de la chaleur d'optimiser son coût et sa consommation énergétique.
-Pour l'agriculture, un exemple peut être d'équiper chaque vache (dans les grandes exploitations) d'un capteur qui émet régulièrement des informations biologiques, telles que la température corporelle ou le niveau de stress, afin d'optimiser le temps de traite, de surveiller la santé des animaux, etc.
+Pour le développement futur des ``réseaux intelligents'', des ``villes intelligentes'', des ``maisons intelligentes'' ou de l'``agriculture intelligente'', les réseaux de capteurs devraient être largement déployés.
+Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les ``bâtiments connectés'' et l'``agriculture connectée''.
+Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides et d'utiliser des réseaux de capteurs afin d'obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, dans le but de permettre au contrôle centralisé du chauffage d'optimiser son coût et sa consommation énergétique.
+Pour l'agriculture, un exemple peut être d'équiper chaque tête de bétail (dans les grandes fermes) d'un capteur qui émet régulièrement des informations biologiques, telles que la température corporelle ou le niveau de stress, afin d'optimiser l'heure de traite des vaches, de surveiller la santé des animaux, etc.
 
@@
 Dans les réseaux IdO actuels et futurs, de nombreux dispositifs de nature hétérogène utilisent la même antenne pour des applications différentes.
-Un problème commun est que ces appareils IoT ont une forte contrainte sur leur consommation d'énergie, car la plupart d'entre eux seront déployés sans accès direct à l'alimentation et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
-Par exemple, la plupart des entreprises commerciales d'IdO vendent aujourd'hui une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
-Une autre contrainte commune aux dispositifs IoT est leur faible rapport cyclique, car la plupart des applications ciblent un ou quelques messages à envoyer chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
+Un problème commun est que ces appareils IdO ont une forte contrainte sur leur consommation d'énergie, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
+Par exemple, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des appareils ayant une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
+Une autre contrainte commune aux dispositifs IdO est leur faible besoin en communication, car la plupart des applications ciblent des besoins d'un à quelques messages à envoyer chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
 %
@@
 
-\paragraph{Pénurie du spectre radio.}
+\paragraph{Épuisement du spectre radio.}
 %
-Un problème majeur des technologies sans fil actuelles est la question de la pénurie du spectre radio :
+Un problème majeur des technologies sans fil actuelles est la question de l'épuisement du spectre radio :
 dans la plupart des bandes de fréquences, la totalité du spectre RF est désormais attribuée et les bandes libres n'existent plus, ce qui limite la possibilité d'ajouter tout nouvel usage.
-Tel qu'illustré dans la Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} ci-dessous,
+Tel qu'illustré en Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} ci-dessous,
 presque tout le spectre
-est affecté à divers usages, qui vont de la radionavigation maritime (historiquement le premier usage des radiocommunications, \eg, dans le Titanic), à la recherche spatiale, l'inter-satellite, la téléphonie mobile et de nombreuses autres applications.
+est affecté à divers usages, qui vont de la radionavigation maritime (historiquement le premier usage des radiocommunications, \eg, dans le Titanic), à la recherche spatiale, les communications inter satellitaire, la téléphonie mobile et de nombreuses autres applications.
 %
@@
 
-Les bandes des réseaux cellulaires sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radioamateurs) sont moins utilisées.
-Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend fortement du moment et du lieu, comme le montre \cite{Lopez2009spectral}.
+Les bandes des réseaux cellulaires sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radio amateures) sont moins utilisées.
+Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend fortement du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
 En outre, l'attribution fixe du spectre empêche
@@
 Donc, au cours des $15$ dernières années, grâce à un lobbying actif de la communauté de la radio intelligente,
-les organismes de réglementation dans le monde se sont demandé s'il fallait permettre à l
-un nouveau paradigme de communication sans fil :
+les organismes de réglementation dans le monde se sont demandé s'il fallait permettre un nouveau paradigme de communication sans fil :
 permettre aux utilisateurs non titulaires d'une licence d'utiliser des bandes sous licence s'ils ne causent pas d'interférence aux utilisateurs (payants) titulaires d'une licence.
@@
 
-\paragraph{Des TIC à la Radio Intelligente (CR).}
+\paragraph{Des TIC à la Radio Intelligente (RI).}
 %
 Nous pouvons affiner le premier champ d'étude de cette thèse, étape par étape :
-des technologies de l'information et communication (TIC), aux télécommunications, aux communications sans fil, puis à la radio intelligente (RC),
+des technologies de l'information et communication (TIC), aux télécommunications, aux communications sans fil, puis à la radio intelligente (en anglais, \emph{cognitive radio} CR),
 et enfin à la radio logicielle (Software Defined Radio, SDR).
@@
 % https://en.wikipedia.org/wiki/Software-defined_radio
-Un SDR est un système de radiocommunication où les composants qui ont été traditionnellement implémentés dans le matériel (\eg, mélangeurs, filtres, modulateurs/démodulateurs, détecteurs, etc) sont de plus en plus implémentés au moyen de logiciels sur un processeur.
+Un SDR est un système de radiocommunication où les composants qui ont été traditionnellement implémentés par du matériel dédiés (\eg, mélangeurs, filtres, modulateurs/démodulateurs, détecteurs, etc) sont de plus en plus implémentés au moyen de logiciels sur un processeur.
 %
-Même si le paradigme de la RRL a été initié par la recherche américaine en matière de défense dans les années 1980, au cours des 20 dernières années, l'industrie civile a commencé à s'intéresser à la RRL, et la RRL a également suscité beaucoup d'intérêt de la part du milieu universitaire et de l'industrie.
+Même si le paradigme de la RI a été initié par la recherche américaine en matière de défense dans les années $1980$, au cours des $20$ dernières années, l'industrie civile a commencé à s'intéresser à la RI, et la RI a également suscité beaucoup d'intérêt de la part du milieu universitaire et de l'industrie.
 %
-Comme la RC n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RC, qui vient de paraître il y a $20$ il y a quelques années.
+Comme la RI n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RI, qui vient de paraître il y a $20$ il y a quelques années.
 %
@@
     Puis S. Haykin en $2005$ a aussi dit que
-    \emph{``une radio intelligente (CR) est un système de communication sans fil intelligent qui est capable de connaître son environnement, d'apprendre et d'adapter ses paramètres de fonctionnement (puissance d'émission et fréquence porteuse) à la volée dans le but de fournir une communication fiable à tout moment, en tout lieu et efficace sur le plan spectral''} \cite{Haykin05}.
+    \emph{``une radio intelligente (RI) est un système de communication sans fil intelligent qui est capable de connaître son environnement, d'apprendre et d'adapter ses paramètres de fonctionnement (puissance d'émission et fréquence porteuse) à la volée dans le but de fournir une communication fiable à tout moment, en tout lieu et efficace sur le plan spectral''} \cite{Haykin05}.
 
@@
     La \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia encyclopedia} dit ceci
-    \emph{``une CR est une radio qui peut être programmée et configurée dynamiquement pour utiliser les meilleurs canaux sans fil à proximité afin d'éviter les interférences et la congestion des utilisateurs.
+    \emph{``une RI est une radio qui peut être programmée et configurée dynamiquement pour utiliser les meilleurs canaux sans fil à proximité afin d'éviter les interférences et la congestion des utilisateurs.
 Une telle radio détecte automatiquement les canaux disponibles dans le spectre sans fil, puis modifie en conséquence ses paramètres de transmission ou de réception pour permettre plus de communications sans fil simultanées dans une bande de spectre donnée à un endroit donné''}.
@@
 %
-Dans les bandes sous licence, il y a des utilisateurs primaires (PU) qui paient pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
-Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est pas utilisé par une UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (UD), pourraient l'utiliser en payant un loyer à l'UP.
+Dans les bandes sous licence, il y a des utilisateurs primaires (UP) qui paient pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
+Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est pas utilisé par une UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser en payant un loyer à l'UP.
 %
-La réglementation stipule que le PU a une priorité stricte,
-mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si une UD était équipée d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si elle est exempte de toute PU.
-Ceci définit le concept de \textbf{Opportunistic Spectrum Access} (OSA), pour lequel nous nous référons à l'enquête \cite{Zhao07} pour plus de détails.
+La réglementation stipule que le UP a une priorité stricte,
+mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si une US était équipée d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si elle est exempte de toute UP.
+Ceci définit le concept d'\textbf{Accès Opportuniste au Spectre} (AOS, \emph{Opportunistic Spectrum Access}, OSA), pour lequel nous nous référons à l'article \cite{Zhao07} pour plus de détails.
 
@@
 Il est également inclus dans le concept plus général d'apprentissage du renforcement (RL), lui-même un des domaines de l'apprentissage machine (ML).
-ML > RL > MAB
+% ML > RL > MAB
 %
-Le livre de référence sur RL est \cite{SuttonBarto2018}, citons donc la définition de RL donnée par R. Sutton et A. Barto :
+Le livre de référence sur RL est \cite{SuttonBarto2018}, citons donc la définition qu'en donnent R. Sutton et A. Barto :
 \emph{``l'apprentissage par renforcement, c'est apprendre quoi faire - comment faire le lien entre les situations et les actions -
@@
 
-Nous illustrons ci-dessous l'idée d'un cycle d'apprentissage, alternant entre l'action et le feedback,
-dans la Figure~\ref{fig:1:RenforcementLearningCycle}.
+Nous illustrons ci-dessous l'idée d'un cycle d'apprentissage, alternant entre l'action et la réaction,
+dans la Figure~\ref{fig:1:ReinforcementLearningCycle}.
 Un joueur (ou un apprenant) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
-Le but du joueur est de maximiser ses récompenses, par des essais et des erreurs (@@ie, actions et récompenses).
-De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage de renforcement, comme l'illustrent l'enquête \cite{bouneffouf2019survey} et la section~\ref{sec:2:applicationsdeStochasticMAB}, par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie (essai clinique), etc.
+Le but du joueur est de maximiser ses récompenses, par des essais et des erreurs (\ie, actions et récompenses).
+De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage de renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie (essai clinique), etc.
 
@@
 Ce genre de rétroaction limitée s'appelle \emph{information sur les bandits}, et nous discutons de l'histoire et du concept des \emph{Multi-Armed Bandits}.
-(MAB) en détail dans le prochain chapitre~\ref{chapitre:2}.
+(MAB) en détail dans le prochain Chapitre~\ref{chapter:2}.
 Les essais cliniques et la découverte de traitements ont été historiquement la première application du MAB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33}.
@@
 
-\paragraph{Accès Opportuniste au Spectre (OSA).}
+\paragraph{Accès Opportuniste au Spectre (AOS).}
 %
-Les deux communautés de l'OSA et du MAB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des $2000$ et le début des $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
-L'accent est mis sur une UD accédant à un spectre sous licence, occupé par une UD, qui a une priorité stricte sur l'UD, qui doit suivre un schéma d'accès " écoute avant conversation ".
+Les deux communautés de l'AOS et du MAB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des $2000$ et le début des $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
+L'accent est mis sur une US accédant à un spectre sous licence, occupé par une US, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès " écoute avant conversation ".
 %
 Les hypothèses suivantes sont bonnes :
-l'UD est équipée d'une capacité de détection,
+l'US est équipée d'une capacité de détection,
 et considère un ensemble fixe et fini de canaux orthogonaux, c'est-à-dire différentes bandes de fréquences dans un spectre sous licence.
 Par exemple, il peut s'agir d'un ensemble de trois canaux Wi-Fi à fréquence différente, émis par la même station Wi-Fi.
-Une autre hypothèse est que PU et SU sont synchronisés dans le temps, en subdivisant le temps en pas de temps discrets (ou ronds).
+Une autre hypothèse est que UP et SU sont synchronisés dans le temps, en subdivisant le temps en pas de temps discrets (ou ronds).
 %
 Ainsi, si l'unité de station de base passe peu de temps au début de chaque plage horaire pour effectuer la détection, elle peut rechercher la présence ou l'absence d'une UP avant de transmettre, pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
-Si l'UD était capable de détecter toutes les bandes de fréquences $K$, elle pouvait simplement émettre dans l'un des canaux libres s'il y en avait, ou ne pas émettre si tous les canaux sont utilisés à un pas de temps donné.
-Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les enquêtes \cite{yucek2009survey,subhedar2011spectrum}, de sorte que la plupart des travaux sur l'OSA limitent la capacité de détection de l'UD à une seule chaîne à la fois.
-Cette hypothèse, ainsi que l'hypothèse selon laquelle la PU ne peut pas être perturbée (sans laquelle aucune régulation ne sera jamais acceptée pour l'OSA), impose à l'UU de transmettre dans le canal qu'elle a détecté, à chaque pas de temps, si et seulement si elle a été détectée sans PU quelconque.
+Si l'US était capable de détecter toutes les bandes de fréquences $K$, elle pouvait simplement émettre dans l'un des canaux libres s'il y en avait, ou ne pas émettre si tous les canaux sont utilisés à un pas de temps donné.
+Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les enquêtes \cite{yucek2009survey,subhedar2011spectrum}, de sorte que la plupart des travaux sur l'AOS limitent la capacité de détection de l'US à une seule chaîne à la fois.
+Cette hypothèse, ainsi que l'hypothèse selon laquelle la UP ne peut pas être perturbée (sans laquelle aucune régulation ne sera jamais acceptée pour l'AOS), impose à l'UU de transmettre dans le canal qu'elle a détecté, à chaque pas de temps, si et seulement si elle a été détectée sans UP quelconque.
 coût non négligeable de la reconfiguration radio
 
-En se concentrant sur un SU dans un réseau OSA, il doit décider séquentiellement d'un canal à détecter (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il effectue la détection, et enfin il transmet dans ce canal si il a été détecté libre.
-Le but de l'UD est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou l'équivalent, pour maximiser son nombre de transmissions réussies.
+En se concentrant sur un SU dans un réseau AOS, il doit décider séquentiellement d'un canal à détecter (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il effectue la détection, et enfin il transmet dans ce canal si il a été détecté libre.
+Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou l'équivalent, pour maximiser son nombre de transmissions réussies.
 %
-Si les différents canaux ne sont pas uniformément utilisés par la PU, et si l'on suppose une hypothèse stationnaire sur le trafic de la PU, alors l'objectif de la SU se résume à explorer les différents canaux et à exploiter les meilleurs.
+Si les différents canaux ne sont pas uniformément utilisés par la UP, et si l'on suppose une hypothèse stationnaire sur le trafic de la UP, alors l'objectif de la SU se résume à explorer les différents canaux et à exploiter les meilleurs.
 Le problème de l'accès opportuniste au spectre est donc un problème d'exploration/exploitation, avec un ensemble fini d'actions (les canaux, également appelés \emph{arms}),
@@
 
-\paragraph{MAB pour l'OSA, et une brève présentation de l'historique de cette recherche dans l'équipe SCEE.}
+\paragraph{MAB pour l'AOS, et un bref historique de ces recherches par l'équipe SCEE.}
 %
@@
 les bandes de fréquences orthogonales (ou canaux) sont des modèles par bras $k\in\{1,\dots,K\}$,
-et le feedback obtenu par le dispositif équipé de CR après avoir détecté le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
+et la récompense obtenue par le dispositif équipé de CR après avoir détecté le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
 En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message SU peut être envoyé), alors qu'une récompense de $r(t)=0$ indique que le canal $k$ est occupé au moment $t$ et qu'aucun message SU ne doit être envoyé, jusqu'à la fin de cet intervalle.
@@
 
-Leurs travaux ont été parmi les premiers à proposer l'utilisation de Reinforcement Learning for Cognitive Radio et le problème OSA, notamment le modèle MAB et l'algorithme $\UCB_1$,
+Leurs travaux ont été parmi les premiers à proposer l'utilisation de Reinforcement Learning for Cognitive Radio et le problème AOS, notamment le modèle MAB et l'algorithme $\UCB_1$,
 ainsi que les premiers travaux de Q. Zhao et de son équipe, par exemple avec les travaux \cite{Liu08,Zhao10}.
@@
 %
-Par exemple, des preuves de concepts comme \cite{kumar2016two} ont prouvé la capacité de telles approches sur des signaux radio réels pour l'OSA.
+Par exemple, des preuves de concepts comme \cite{kumar2016two} ont prouvé la capacité de telles approches sur des signaux radio réels pour l'AOS.
 %
@@
 La littérature susmentionnée a essentiellement montré que les algorithmes MAB peuvent être appliqués avec succès au problème de l'AOS.
-Toutefois, si l'on considère les dispositifs prêts à l'emploi qui ne peuvent pas effectuer la détection, tels que les dispositifs finaux à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle MAB qui utilise la détection pour détecter le PU dans le cas de l'OSA ne peut plus être appliqué.
-En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre PU et SU.
+Toutefois, si l'on considère les dispositifs prêts à l'emploi qui ne peuvent pas effectuer la détection, tels que les dispositifs finaux à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle MAB qui utilise la détection pour détecter le UP dans le cas de l'AOS ne peut plus être appliqué.
+En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et SU.
 En d'autres termes, il n'y a plus de priorité entre les utilisateurs des réseaux IdO.
-(plus de PU, tous sont SU).
+(plus de UP, tous sont SU).
 %
@@
     \item
-    nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre PU et SU),
+    nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre UP et SU),
     \item
-    la plupart des dispositifs IoT sont des dispositifs à très faible coût, fonctionnant éventuellement sur du matériel de mauvaise qualité et ayant des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
+    la plupart des dispositifs IdO sont des dispositifs à très faible coût, fonctionnant éventuellement sur du matériel de mauvaise qualité et ayant des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
     Cela signifie que même s'ils ne sont pas équipés de capacités de détection particulières,
@@
     \item
-    une seule station de base IoT devra gérer un très grand nombre d'appareils,
+    une seule station de base IdO devra gérer un très grand nombre d'appareils,
     et ne peut leur envoyer des ordres de coordination afin d'optimiser le réseau,
     \item
-    Les dispositifs IoT ont des cycles de fonctionnement faibles (seulement quelques messages par minute) à très faibles cycles de fonctionnement (seulement quelques messages par mois).
+    les dispositifs IdO ont des cycles de fonctionnement faibles (seulement quelques messages par minute) à très faibles cycles de fonctionnement (seulement quelques messages par mois).
 \end{itemize}
@@
 %
-De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de la batterie, les dispositifs IoT sont capables d'utiliser leur antenne reçue pour détecter un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), mais ils ne peuvent pas effectuer de détection de spectre comme cela est envisagé pour OSA et CR (\ie, au début de chaque intervalle, dans un schéma " écoute avant conversation ").
+De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de la batterie, les dispositifs IdO sont capables d'utiliser leur antenne reçue pour détecter un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), mais ils ne peuvent pas effectuer de détection de spectre comme cela est envisagé pour AOS et RI (\ie, au début de chaque intervalle, dans un schéma " écoute avant conversation ").
 
-On pourrait penser qu'en l'absence de détection, l'apprentissage du renforcement n'est plus possible, mais tout appareil IoT réel reçoit encore des informations sur son environnement après certaines ou toutes les transmissions.
-Dans la plupart des normes IoT, un message de liaison montante envoyé à une station de base peut être suivi d'un message de liaison descendante renvoyé par la station de base pour indiquer si le message de liaison montante a été bien reçu et compris.
+On pourrait penser qu'en l'absence de détection, l'apprentissage du renforcement n'est plus possible, mais tout appareil IdO réel reçoit encore des informations sur son environnement après certaines ou toutes les transmissions.
+Dans la plupart des normes IdO, un message de liaison montante envoyé à une station de base peut être suivi d'un message de liaison descendante renvoyé par la station de base pour indiquer si le message de liaison montante a été bien reçu et compris.
 %
-En utilisant cette rétroaction, ça consiste en un \emph{acknowledgement} (\Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IoT puisse également utiliser un algorithme RL bien conçu afin d'optimiser ses communications.
+En utilisant cette rétroaction, ça consiste en un \emph{acknowledgement} (\Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IdO puisse également utiliser un algorithme RL bien conçu afin d'optimiser ses communications.
 
@@
 Si nous résumons les problèmes étudiés et les résumons en une seule question, ce pourrait être la suivante :
-\Insister sur le fait que l'on peut adapter les outils de prise de décision déjà appliqués avec succès à la radio intelligente pour l'accès opportuniste au spectre aux besoins spécifiques de la CR pour les (futurs) réseaux de l'Internet des objets ?
+\Insister sur le fait que l'on peut adapter les outils de prise de décision déjà appliqués avec succès à la radio intelligente pour l'accès opportuniste au spectre aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des objets ?
 %
@@
 
-\paragraph{Chapter~\ref{chapter:2}.}
+\paragraph{Chapitre~\ref{chapter:2}.}
 %
@@
 car il existe de nombreux algorithmes différents, avec de nombreuses variantes sur le problème simple présenté ci-dessus.
-Nous commençons donc la première partie~\ref{part:Introduction} de cette thèse par le chapitre~\ref{chapitre:2}, qui présente le modèle MAB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires du MAB.
+Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle MAB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires du MAB.
 %
-Afin de bien comprendre quels algorithmes pourraient être adaptés aux contraintes susmentionnées de la RC pour les réseaux IdO,
-nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme MAB (son regret, voir ci-dessous dans Section~\ref{sec:2:lowerUpperBoundsRegret}),
+Afin de bien comprendre quels algorithmes pourraient être adaptés aux contraintes susmentionnées de la RI pour les réseaux IdO,
+nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme MAB (son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
 mais aussi par leurs performances empiriques en termes de complexité de calcul et d'exigences de stockage, tant du point de vue des analyses théoriques que des mesures en temps réel et des empreintes mémoire.
 
-\paragraph{Chapter~\ref{chapter:3}.}
+\paragraph{Chapitre~\ref{chapter:3}.}
 %
@@
 Nous avons écrit la bibliothèque de simulation open-source la plus exhaustive pour les problèmes MAB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
-Nous présentons en détail son architecture et ses fonctionnalités dans le chapitre~\ref{chapitre:3}, ainsi que différents exemples de son utilisation.
-Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les simulations discutées dans la suite de cette thèse.
+Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
+Une documentation complète est disponible en ligne, ainsi que des instructions pour reproduire les simulations discutées dans la suite de cette thèse.
 
-\paragraph{Chapter~\ref{chapter:25}.}
+\paragraph{Chapitre~\ref{chapter:25}.}
 %
@@
 La première est une comparaison empirique des algorithmes existants les plus efficaces et les mieux connus, dans Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
-Nous confirmons que les algorithmes largement utilisés et pas trop sophistiqués, tels que \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
-La seconde approche est une sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons dans Chapitre~\ref{chapitre:25}.
-Ce travail sur l'agrégation des algorithmes MAB a été motivé pour le cas OSA de la radio cognitive, où de nombreux travaux de recherche antérieurs ne prenaient en compte que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
-Cette contribution a été présentée à la conférence WCNC de l'IEEE à Barcelone, Espagne, en avril 2010.
+Nous confirmons que les algorithmes largement utilisés, tels que \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
+La seconde approche est une sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons dans Chapitre~\ref{chapter:25}.
+Ce travail sur l'agrégation des algorithmes MAB a été motivé pour le cas AOS de la radio cognitive, où de nombreux travaux de recherche antérieurs ne prenaient en compte que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
+Cette contribution a été présentée à la conférence IEEE WCNC à Barcelone, Espagne, en avril 2018.
 
@@
 
-Dans la deuxième partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux d'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du MAB, issues de la simplification de notre premier modèle.
+Dans la deuxième Partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux d'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du MAB, issues de la simplification de notre premier modèle.
 
-\paragraph{Chapter~\ref{chapter:4}.}
+\paragraph{Chapitre~\ref{chapter:4}.}
 %
-Nous étudions deux modèles de réseaux IdO dans le chapitre~\ref{chapitre:4}.
+Nous étudions deux modèles de réseaux IdO dans le Chapitre~\ref{chapter:4}.
 Le premier modèle considère les dispositifs IdO qui ont simplement des données à envoyer régulièrement à une station de base (à des moments imprévisibles ou aléatoires).
 De tels dispositifs utilisent les acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
-L'objectif de ce premier modèle est d'améliorer la qualité de service (QoS) de l'application de ce réseau IoT, en appliquant une RL décentralisée côté appareil afin de réduire le taux d'échec de transmission.
+L'objectif de ce premier modèle est d'améliorer la qualité de service (QoS) de l'application de ce réseau IdO, en appliquant une RL décentralisée côté appareil afin de réduire le taux d'échec de transmission.
 La station de base recevra donc plus de paquets de liaison montante des appareils desservis, si elle peut apprendre à utiliser un système efficace d'accès au spectre.
@@
 %
-La preuve de concept (PoC) a continué notre collaboration fructueuse avec R.
-Bonnefoi, et a été démontrée pendant trois jours à la conférence ICT de l'IEEE à Saint-Malo, France, en juin $2018$ \cite{Besson2018ICT}, et ensuite présentée à la conférence IEEE WCNC à Marrakech, Maroc, en avril $2019$ \cite{Besson2019WCNC}.
+La preuve de concept (PoC) a continué notre collaboration fructueuse, et a été démontrée pendant trois jours à la conférence IEEE ICT à Saint-Malo, France, en juin $2018$ \cite{Besson2018ICT}, et ensuite présentée à la conférence IEEE WCNC à Marrakech, Maroc, en avril $2019$ \cite{Besson2019WCNC}.
 %
 Notre deuxième modèle, avec retransmissions, correspond à notre dernière collaboration avec R.
-Bonnefoi, qui a été présentée à l'atelier $1^{\text{st}}$ MoTION, également pendant la conférence IEEE WCNC $2019$ \cite{Bonnefoi2019WCNC}.
+Bonnefoi, qui a été présentée à l'atelier MoTION, également pendant la conférence IEEE WCNC $2019$ \cite{Bonnefoi2019WCNC}.
 
-\paragraph{Chapter~\ref{chapter:5}.}
+\paragraph{Chapitre~\ref{chapter:5}.}
 %
-Dans les deux modèles présentés ci-dessus, le concept de base est de laisser chaque dispositif dynamique d'un réseau IdO exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
+Dans les deux modèles présentés ci-dessus, le concept est de laisser chaque dispositif dynamique d'un réseau IdO exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
 Il est important de noter que chaque dispositif ne communique, et donc n'apprend, qu'à quelques tours et non à toutes les étapes, en suivant son propre processus d'activation (aléatoire) (nous nous limitons à un processus d'activation de Bernoulli purement aléatoire de probabilité $p$).
@@
 Il est bien connu dans la littérature sur la théorie des jeux que le jeu égoïste peut être désastreux pour la mesure centralisée de la performance, on peut penser aux "dilemmes" populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
-Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage MAB décentralisé et égoïste conduisait à une coordination efficace entre les dispositifs dans tous les scénarios considérés, malgré le fait que ces objets IoT ne peuvent communiquer directement entre eux, et reçoivent seulement un retour de collision depuis la station de base à laquelle ils sont associés.
+Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage MAB décentralisé et égoïste conduisait à une coordination efficace entre les dispositifs dans tous les scénarios considérés, malgré le fait que ces objets IdO ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
 
 Nous avons d'abord essayé d'analyser la performance de cet algorithme décentralisé, que nous appelons \Selfish, dans le modèle de Section~\ref{sec:4:firstModel},
-mais en raison du nombre aléatoire de dispositifs actifs à chaque pas de temps (@@ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
+mais en raison du nombre aléatoire de dispositifs actifs à chaque pas de temps (\ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
 %
-C'est pourquoi dans le chapitre~\ref{chapitre:5}, nous relâchons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$, qui est équivalent à avoir $p < 1$, et nous considérons le cas des dispositifs IoT communiquant à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ dispositifs à jour.
-Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas OSA (\ie, avec détection) ou le cas IoT (\ie, sans détection).
-L'objectif était de comprendre l'heuristique de Chapter~\ref{chapter:4}, \Selfish, dans le cadre plus simple du multi-player MAB, qui a été étudié auparavant pour le cas de l'OSA, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
+C'est pourquoi dans le Chapitre~\ref{chapter:5}, nous relâchons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$, qui est équivalent à avoir $p < 1$, et nous considérons le cas des dispositifs IdO communiquant à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ dispositifs à jour.
+Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas AOS (\ie, avec détection) ou le cas IdO (\ie, sans détection).
+L'objectif était de comprendre l'heuristique du Chapitre~\ref{chapter:4}, \Selfish, dans le cadre plus simple du multi-player MAB, qui a été étudié auparavant pour le cas de l'AOS, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
 %
-Le cas OSA couvert par notre modèle est en fait légèrement différent de celui considéré par \cite{Jouini10} et d'autres travaux antérieurs,
+Le cas AOS couvert par notre modèle est en fait légèrement différent de celui considéré par \cite{Jouini10} et d'autres travaux antérieurs,
 car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
@@
 
-Comme expliqué plus haut, le modèle de Chapitre~\ref{chapitre:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux équipements finaux, tous suivant des modèles d'activation aléatoires.
+Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux équipements finaux, tous suivant des modèles d'activation aléatoires.
 La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes MAB, conçus pour résoudre des problèmes stationnaires, sur un problème non stationnaire,
 mais plutôt que nous analysons des algorithmes qui jouent tous dans différents sous-ensembles (aléatoires) des pas de temps synchronisés globaux.
-Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et appliquer au maximum $M \leq K$ aux dispositifs n'est pas vraiment réaliste pour les réseaux IdO, même si cela conduit au modèle du Chapitre~\ref{chapitre:5}.
+Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et appliquer au maximum $M \leq K$ aux dispositifs n'est pas vraiment réaliste pour les réseaux IdO, même si cela conduit au modèle du Chapitre~\ref{chapter:5}.
 
-D'une part, si le modèle d'activation de tous les périphériques peut être corrigé, par exemple en se basant sur une affectation centralisée des périphériques à différents créneaux horaires, alors le modèle de bandit multi-joueurs du Chapitre~\ref{chapitre:5} peut être utilisé pour permettre à chaque groupe de périphériques $M \leq K$ d'apprendre une affectation orthogonale optimale dans les canaux $K$ (\eg, groupes peuvent être le jeu des périphériques que tous émettent au même instant).
+D'une part, si le modèle d'activation de tous les périphériques peut être corrigé, par exemple en se basant sur une affectation centralisée des périphériques à différents créneaux horaires, alors le modèle de bandit multi-joueurs du Chapitre~\ref{chapter:5} peut être utilisé pour permettre à chaque groupe de périphériques $M \leq K$ d'apprendre une affectation orthogonale optimale dans les canaux $K$ (\eg, groupes peuvent être le jeu des périphériques que tous émettent au même instant).
 %
@@
 
-D'autre part, une autre façon de relâcher l'hypothèse non stationnaire est de prendre le point de vue d'un dispositif, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur une unité de soins entourée de plusieurs PU ayant des comportements fixes.
-Si nous nous concentrons sur un dispositif IoT, son environnement (\ie, les dispositifs environnants) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réalistes donnent généralement une certaine structure sur cette non-stationnalité.
+D'autre part, une autre façon de relâcher l'hypothèse non stationnaire est de prendre le point de vue d'un dispositif, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur une unité de soins entourée de plusieurs UP ayant des comportements fixes.
+Si nous nous concentrons sur un dispositif IdO, son environnement (\ie, les dispositifs environnants) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réalistes donnent généralement une certaine structure sur cette non-stationnarité.
 Par exemple, l'application de la stationnarité à certains intervalles de temps conduit à la dernière contribution et au dernier chapitre de cette thèse.
 
-\paragraph{Chapter~\ref{chapter:6}.}
+\paragraph{Chapitre~\ref{chapter:6}.}
 %
 Afin de comprendre également avec précision comment les algorithmes MAB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur le MAB accusatoire et sur le MAB non stationnaire, pour les deux cas d'environnements variant lentement et changeant brusquement.
-Nous commençons notre dernier Chapitre~\ref{chapitre:6} en passant en revue les travaux existants,
+Nous commençons notre dernier Chapitre~\ref{chapter:6} en passant en revue les travaux existants,
 et nous avons choisi de nous concentrer sur le problème des pièces stationnaires,
@@
     \item
-    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (publié à GRETSI 2019).
+    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le Chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (publié à GRETSI 2019).
     Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles .
@@
 afin d'établir des preuves mathématiques garantissant la convergence, ainsi que les gains en performance apportés par ces algorithmes de bandit.
-Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
+Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du Chapitre~\ref{chapter:2} et les modèles de réseaux IdO du Chapitre~\ref{chapter:4}.
 Ces deux études ont chacune donné des résultats théoriques à la pointe de la recherche sur les modèles de bandits multi-joueurs et de bandits stationnaires par morceaux, que nous avons aussi validé par des expériences numériques.

### 2019-07-06 2-Chapters/1-Chapter/chapter1.tex
Done for tonight

@@
 % % Verify this first model empirically: proof of concept
-The proof-of-concept (PoC) continued the fruitful collaboration with R.
-Bonnefoi, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
+The proof-of-concept (PoC) continued the fruitful collaboration, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
 %

### 2019-07-07 1-Preamble/5-Abstract/resume_fr.tex
Almost done for this French translation of Ch1, useless as I'll remove it tonight, but still...

@@
 % Chapitre 1
-Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio sans fil de manière intelligente, dans des bandes non licenciées et sans supervision centrale.
-Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IdO), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil (Chapitre~\ref{chapter:1}).
+Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les objets reconfigurables qui peuvent accéder à des réseaux de radio sans fil de manière intelligente, dans des bandes non licenciées et sans supervision centrale.
+Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IdO), avec l'objectif d'augmenter la durée de vie de la batterie des objets, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil (Chapitre~\ref{chapter:1}).
 % Chapitre 4
@@
 % , Chapitre~\ref{chapter:2}
-notamment si l'on veut prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
+notamment si l'on veut prouver la convergence de nombreux objets jouant à un jeu collaboratif sans aucune coordination, lorsque les objets suivent tous un modèle d'activation aléatoire.
 % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
@@
 Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, la télé-communication sans fil entre des appareils fabriqués a été rendue possible, et de plus en plus fréquentes dans nos vies.
-Avec l'avènement des réseaux Internet des Objets (IdO), des milliards de dispositifs autonomes à basse consommation devraient être déployés dans le monde entier, pour un large éventail d'applications différentes.
+Avec l'avènement des réseaux Internet des Objets (IdO), des milliards de $M$ objets basse consommation devraient être déployés dans le monde entier, pour un large éventail d'applications différentes.
 Il existe aujourd'hui un consensus mondial sur le fait qu'avec la tendance actuelle à l'accroissement démographique et la crise énergétique actuelle, toute nouvelle technologie déployée doit être à la fois bon marché et efficace sur le plan énergétique,
@@
 C'est pourquoi nous nous intéressons à cette thèse sur
-l'amélioration de la durée de vie des batteries des équipements terminaux IdO et la réduction du coût énergétique des réseaux IdO.
-Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant la prise de décision décentralisée à faible coût directement dans les futurs dispositifs IdO.
+l'amélioration de la durée de vie des batteries des objets IdO et la réduction du coût énergétique des réseaux IdO.
+Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant la prise de décision décentralisée à faible coût directement dans les futurs $M$ objets
 %
-Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (Multi-Armed Bandit algorithms), afin de permettre aux appareils IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
+Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (Multi-Armed Bandit algorithms), afin de permettre aux objets IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
 
@@
 %
-Ce troisième type de systèmes peut être conçu comme décentralisé :
-même si une station de base centrale est toujours en charge de nombreux appareils,
-les appareils déclenchent l'envoi de paquets de liaison montante, et les seules données de liaison descendante qu'ils peuvent recevoir sont de courts accusés de réception, envoyés par la station de base pour indiquer le succès ou l'échec de chaque paquet de liaison montante.
-Cette famille de systèmes sans fil est appelée Internet des Objets (IdO), ou \emph{Internet of Things} (IoT),
+Ce troisième type de systèmes peut être désigné comme décentralisé :
+même si une station de base centrale est toujours en charge de nombreux objets (appareils),
+ce sont les objets qui déclenchent l'envoi de paquets radio, et les seules données descendantes qu'ils peuvent recevoir sont de courts accusés de réception (\emph{acknowledgements}, \Ack), envoyés par la station de base pour indiquer le succès ou l'échec de chaque paquet envoyé.
+Cette famille de systèmes sans fil est appelée Internet des Objets (IdO, ou \emph{Internet of Things}, IoT),
 et un exemple typique d'application de tels réseaux IdO est celui des réseaux sans fil de capteurs.
@@
 
-Ce troisième type de systèmes sans fil se caractérise par leur nature décentralisée,
-où les communications sont initialisées et régulées par les appareils, et non par un système de contrôle centralisé.
+Ce troisième type de systèmes sans fil est caractérisé par leur nature décentralisée,
+où les communications sont initialisées et régulées par les objets, et non par un système de contrôle centralisé.
 %
-En effet, une commande centrale nécessite la signalisation de paquets qui ont été identifiés comme trop lourds pour ce type de systèmes.
-Dans les réseaux IdO actuels et futurs, de nombreux dispositifs de nature hétérogène utilisent la même antenne pour des applications différentes.
-Un problème commun est que ces appareils IdO ont une forte contrainte sur leur consommation d'énergie, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
-Par exemple, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des appareils ayant une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
-Une autre contrainte commune aux dispositifs IdO est leur faible besoin en communication, car la plupart des applications ciblent des besoins d'un à quelques messages à envoyer chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
+En effet, un contrôle centrale nécessite des paquets de signalisation qui ont été identifiés comme trop lourds pour ce type de systèmes.
+Dans les réseaux IdO actuels et futurs, de nombreux $M$ objets hétérogènes utilisent la même antenne pour des applications différentes.
+Un problème commun est que ces objets IdO ont une forte contrainte sur leur consommation d'énergie, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
+Ainsi, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des objets ayant une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
+Une autre contrainte commune aux objets est leur faible besoin en communication, car la plupart des applications n'auront besoin que d'un à quelques messages à envoyer chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
 %
 De nombreuses normes différentes pour les réseaux IdO ont été proposées ces dernières années,
-et ils consistent en une spécification à la fois pour le \emph{PHY}sical
-et le calque \emph{M}edium \emph{A}access \emph{C}ontrol (\emph{MAC}).
-Pour citer quelques exemples de normes, ZigBee, Z-Wave ou Bluetooth visent les communications à courte portée, tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à longue portée.
-Nous nous référons à l'enquête \cite{Centenaro16} pour plus de détails, et des références dans \cite{Azari18} ou notre récent article \cite{MoyBesson2019}.
+et ils consistent en une spécification à la fois pour la couche \emph{PHY}sique
+et la couche \emph{M}edium \emph{A}access \emph{C}ontrol (\emph{MAC}).
+Pour citer quelques exemples de normes, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée, tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
+Nous nous référons à l'article \cite{Centenaro16} pour plus de détails, et les références dans \cite{Azari18} ou notre récent article \cite{MoyBesson2019}.
 
@@
 presque tout le spectre
-est affecté à divers usages, qui vont de la radionavigation maritime (historiquement le premier usage des radiocommunications, \eg, dans le Titanic), à la recherche spatiale, les communications inter satellitaire, la téléphonie mobile et de nombreuses autres applications.
+est affecté à divers usages, qui vont de la radio-navigation maritime (historiquement le premier usage des radio-communications, \eg, dans le Titanic), à la recherche spatiale, les communications inter satellitaire, la téléphonie mobile et de nombreuses autres applications.
 %
@@
 ou la \href{https://cept.org}{ Conférence européenne des administrations des postes et des télécommunications} en Europe (voir \href{https://www.CEPT.org/}{\texttt{CEPT.org}}),
-ainsi que différentes campagnes de mesure indépendantes, ont montré que la plupart des fréquences du spectre des fréquences radioélectriques sont utilisées de manière inefficace.
+ainsi que différentes campagnes de mesure indépendantes, ont montré que la plupart des fréquences du spectre des fréquences radio-électriques sont utilisées de manière inefficace.
 Cela signifie qu'une bande peut être attribuée à un certain usage unique, mais qu'elle peut être libre de tout utilisateur à certains moments et/ou endroits.
@@
 En outre, l'attribution fixe du spectre empêche
-l'introduction de nouveaux services, en particulier pour les équipements à bas prix ou pour les petits marchés.
+l'introduction de nouveaux services, en particulier pour les objets à bas prix ou pour les marchés de niche.
 %
-Donc, au cours des $15$ dernières années, grâce à un lobbying actif de la communauté de la radio intelligente,
+C'est ainsi qu'au cours des $15$ dernières années, grâce à un lobbying actif de la communauté de la radio intelligente,
 les organismes de réglementation dans le monde se sont demandé s'il fallait permettre un nouveau paradigme de communication sans fil :
-permettre aux utilisateurs non titulaires d'une licence d'utiliser des bandes sous licence s'ils ne causent pas d'interférence aux utilisateurs (payants) titulaires d'une licence.
-Ces initiatives sont examinées par le \textbf{Radio Cognitive},
-que nous détaillons ci-dessous, et en particulier pour \textbf{Dynamic Spectrum Access (DSA)},
-pour lesquelles nous renvoyons aux enquêtes \cite{akyildiz2006next,garhwal2012survey} pour plus de détails.
+permettre aux utilisateurs non titulaires d'une licence d'utiliser des bandes sous licence s'ils ne causent pas d'interférence aux utilisateurs payants, \ie, titulaires d'une licence.
+Ces initiatives sont examinées par la \textbf{Radio Intelligente} (RI, ou \emph{Cognitive Radio}, CR),
+que nous détaillons ci-dessous, et en particulier pour l'\textbf{Accès Dynamique au Spectre} (ADS, ou \emph{Dynamic Spectrum Access}, DSA),
+pour lesquelles nous renvoyons aux articles \cite{akyildiz2006next,garhwal2012survey} pour plus de détails.
 
@@
 Dans cette thèse, nous étudions les réseaux Internet des Objets à longue portée, caractérisés principalement par trois contraintes essentielles :
-faible consommation d'énergie (ou longue durée de vie des piles),
-à long terme
-et un cycle d'utilisation réduit.
+faible consommation d'énergie (ou longue durée de vie des batteries),
+longue portée,
+et un faible cycle d'utilisation (\ie, faible à très faible nombre de messages par jour).
 %
@@
 Nous pouvons affiner le premier champ d'étude de cette thèse, étape par étape :
-des technologies de l'information et communication (TIC), aux télécommunications, aux communications sans fil, puis à la radio intelligente (en anglais, \emph{cognitive radio} CR),
-et enfin à la radio logicielle (Software Defined Radio, SDR).
+des technologies de l'information et communication (TIC), aux télécommunications, aux communications sans fil, puis à la Radio Intelligente,
+et enfin à la Radio Logicielle (\emph{Software Defined Radio}, SDR).
 %
-La transition de l'approche historique de la radio matérielle aux architectures SDR est un processus graduel, qui a commencé au début des années 1990 et s'est accéléré dans les années 2000.
+La transition de l'approche historique de la radio matérielle aux architectures SDR est un processus graduel, qui a commencé au début des années $1990$ et s'est accéléré dans les années $2000$.
 % https://en.wikipedia.org/wiki/Software-defined_radio
@@
     \item
-    La \href{https://en.wikipedia.org/wiki/Cognitive_radio}{Wikipedia encyclopedia} dit ceci
+    L'\href{https://en.wikipedia.org/wiki/Cognitive_radio}{encyclopédie Wikipedia} dit ceci
     \emph{``une RI est une radio qui peut être programmée et configurée dynamiquement pour utiliser les meilleurs canaux sans fil à proximité afin d'éviter les interférences et la congestion des utilisateurs.
-Une telle radio détecte automatiquement les canaux disponibles dans le spectre sans fil, puis modifie en conséquence ses paramètres de transmission ou de réception pour permettre plus de communications sans fil simultanées dans une bande de spectre donnée à un endroit donné''}.
+Une telle radio détecte automatiquement les canaux disponibles dans le spectre radio, puis modifie en conséquence ses paramètres de transmission ou de réception pour permettre plus de communications sans fil simultanées dans une bande de spectre donnée à un endroit donné''}.
 \end{itemize}
@@
 %
-Dans les bandes sous licence, il y a des utilisateurs primaires (UP) qui paient pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
-Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est pas utilisé par une UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser en payant un loyer à l'UP.
+Dans les bandes sous licences, il y a des utilisateurs primaires (UP) qui paient pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
+Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est pas utilisé par un UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser aussi, éventuellement en payant un abonnement.
 %
-La réglementation stipule que le UP a une priorité stricte,
-mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si une US était équipée d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si elle est exempte de toute UP.
+La réglementation stipule que les UP ont une priorité stricte,
+mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de toute UP.
 Ceci définit le concept d'\textbf{Accès Opportuniste au Spectre} (AOS, \emph{Opportunistic Spectrum Access}, OSA), pour lequel nous nous référons à l'article \cite{Zhao07} pour plus de détails.
 
-\paragraph{De la statistique ou de l'apprentissage machine aux bandits multi-bras.}
+\paragraph{Des statistiques ou de l'apprentissage machine aux bandits multi-bras.}
 %
-Nous sommes intéressés par l'apprentissage des Bandits Multi-Bras (MAB),
+Nous sommes intéressés par l'apprentissage séquentiel et en particulier par les Bandits Multi-Bras (\emph{Multi-Armed Bandits}, MAB),
 qui est apparu comme un problème intéressant dans la communauté statistique et parmi les chercheurs intéressés par l'apprentissage séquentiel, comme le montrent les travaux antérieurs de \cite{Thompson33,Robbins52,LaiRobbins85}.
-Il est également inclus dans le concept plus général d'apprentissage du renforcement (RL), lui-même un des domaines de l'apprentissage machine (ML).
+Il est également inclus dans le concept plus général d'apprentissage par renforcement (\emph{Reinforcement Learning}, RL), lui-même un des domaines de l'apprentissage machine (\emph{Machine Learning}, ML).
 % ML > RL > MAB
 %
-Le livre de référence sur RL est \cite{SuttonBarto2018}, citons donc la définition qu'en donnent R. Sutton et A. Barto :
+Le livre de référence sur le RL est \cite{SuttonBarto2018}, citons donc la définition qu'en donnent R. Sutton et A. Barto :
 \emph{``l'apprentissage par renforcement, c'est apprendre quoi faire - comment faire le lien entre les situations et les actions -
@@
 Nous illustrons ci-dessous l'idée d'un cycle d'apprentissage, alternant entre l'action et la réaction,
-dans la Figure~\ref{fig:1:ReinforcementLearningCycle}.
-Un joueur (ou un apprenant) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
-Le but du joueur est de maximiser ses récompenses, par des essais et des erreurs (\ie, actions et récompenses).
-De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage de renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie (essai clinique), etc.
+en Figure~\ref{fig:1:ReinforcementLearningCycle}.
+Une joueuse (ou un apprenant) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
+Le but de la joueuse est de maximiser ses récompenses, par des essais et des erreurs (\ie, actions et récompenses).
+De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage de renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, comme par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie, etc.
 
@@
 %
-En d'autres termes, le joueur ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une autre action.
+En d'autres termes, la joueuse ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une autre action.
 Ce genre de rétroaction limitée s'appelle \emph{information sur les bandits}, et nous discutons de l'histoire et du concept des \emph{Multi-Armed Bandits}.
@@
 %
-Le MAB est un exemple simple mais puissant du dilemme bien connu de l'exploration/exploitation :
-lorsqu'il est confronté à un ensemble d'actions de $K$ dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
+Le MAB est un exemple simple mais puissant du dilemme bien connu de entre \emph{exploration et exploitation} :
+lorsqu'elle est confronté à un ensemble d'actions de $K$ dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
 explorer les actions inconnues, afin de recueillir plus d'informations à leur sujet,
@@
 Robbins \cite{Robbins52}, et plus récemment c'est un domaine de recherche actif, depuis les $1990$ \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-La recherche sur le MAB a produit une vaste littérature dans le domaine des $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet de grand intérêt, comme l'illustrent les enquêtes et livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications du MAB dans les récentes années \cite{bouneffouf2019survey}.
+La recherche sur le MAB a produit une vaste littérature dans le domaine des $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet actif, comme l'illustrent les livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications des MAB dans dernières années \cite{bouneffouf2019survey}.
 
@@
 Les deux communautés de l'AOS et du MAB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des $2000$ et le début des $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
-L'accent est mis sur une US accédant à un spectre sous licence, occupé par une US, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès " écoute avant conversation ".
+L'accent est mis sur un US accédant à un spectre sous licence, occupé par un US, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès d'``écoute avant transmission''.
 %
-Les hypothèses suivantes sont bonnes :
-l'US est équipée d'une capacité de détection,
+Les hypothèses sont les suivantes :
+l'US est équipé d'une capacité de détection,
 et considère un ensemble fixe et fini de canaux orthogonaux, c'est-à-dire différentes bandes de fréquences dans un spectre sous licence.
 Par exemple, il peut s'agir d'un ensemble de trois canaux Wi-Fi à fréquence différente, émis par la même station Wi-Fi.
-Une autre hypothèse est que UP et SU sont synchronisés dans le temps, en subdivisant le temps en pas de temps discrets (ou ronds).
+Une autre hypothèse est que UP et US sont synchronisés dans le temps, en subdivisant le temps en pas de temps discrets.
 %
-Ainsi, si l'unité de station de base passe peu de temps au début de chaque plage horaire pour effectuer la détection, elle peut rechercher la présence ou l'absence d'une UP avant de transmettre, pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
-Si l'US était capable de détecter toutes les bandes de fréquences $K$, elle pouvait simplement émettre dans l'un des canaux libres s'il y en avait, ou ne pas émettre si tous les canaux sont utilisés à un pas de temps donné.
-Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les enquêtes \cite{yucek2009survey,subhedar2011spectrum}, de sorte que la plupart des travaux sur l'AOS limitent la capacité de détection de l'US à une seule chaîne à la fois.
-Cette hypothèse, ainsi que l'hypothèse selon laquelle la UP ne peut pas être perturbée (sans laquelle aucune régulation ne sera jamais acceptée pour l'AOS), impose à l'UU de transmettre dans le canal qu'elle a détecté, à chaque pas de temps, si et seulement si elle a été détectée sans UP quelconque.
-coût non négligeable de la reconfiguration radio
+Ainsi, si l'US passe peu de temps au début de chaque plage horaire pour effectuer la détection, il peut rechercher la présence ou l'absence d'un UP avant de transmettre, pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
+Si l'US était capable d'analyser tous les $K$ canaux, il pourrait simplement émettre dans l'un des canaux libres s'il y en avait, ou ne pas émettre si tous les canaux sont utilisés à un pas de temps donné.
+%
+Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les articles \cite{yucek2009survey,subhedar2011spectrum}, de sorte que la plupart des travaux sur l'AOS limitent la capacité de détection de l'US à un seul canal à la fois.
+Cette hypothèse, ainsi que l'hypothèse selon laquelle les UP ne peuvent pas être perturbés (sans laquelle aucune régulation ne sera jamais acceptée pour l'AOS), impose à l'US de transmettre dans le canal qu'il a détecté, à chaque pas de temps, si et seulement si il a été détectée sans aucun UP.
+% coût non négligeable de la reconfiguration radio
 
-En se concentrant sur un SU dans un réseau AOS, il doit décider séquentiellement d'un canal à détecter (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il effectue la détection, et enfin il transmet dans ce canal si il a été détecté libre.
+En se concentrant sur un US dans un réseau AOS, il doit décider séquentiellement d'un canal à détecter (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il effectue la détection, et enfin il transmet dans ce canal si il a été détecté libre.
 Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou l'équivalent, pour maximiser son nombre de transmissions réussies.
 %
-Si les différents canaux ne sont pas uniformément utilisés par la UP, et si l'on suppose une hypothèse stationnaire sur le trafic de la UP, alors l'objectif de la SU se résume à explorer les différents canaux et à exploiter les meilleurs.
+Si les différents canaux ne sont pas uniformément utilisés par les UP, et si l'on suppose une hypothèse stationnaire sur le trafic de les UP, alors l'objectif de l'US se résume à explorer les différents canaux et à exploiter les meilleurs.
 Le problème de l'accès opportuniste au spectre est donc un problème d'exploration/exploitation, avec un ensemble fini d'actions (les canaux, également appelés \emph{arms}),
@@
 les bandes de fréquences orthogonales (ou canaux) sont des modèles par bras $k\in\{1,\dots,K\}$,
-et la récompense obtenue par le dispositif équipé de CR après avoir détecté le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
-En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message SU peut être envoyé), alors qu'une récompense de $r(t)=0$ indique que le canal $k$ est occupé au moment $t$ et qu'aucun message SU ne doit être envoyé, jusqu'à la fin de cet intervalle.
+et la récompense obtenue par le objet équipé de CR après avoir détecté le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
+En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors qu'une récompense de $r(t)=0$ indique que le canal $k$ est occupé au moment $t$ et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
 %
@@
 Dans une deuxième thèse de doctorat \cite{Modi17PhD}, N.
-Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes MAB pour optimiser la sélection des canaux, sur la durée de vie des piles d'un appareil sans fil.
-D'une part, l'utilisation d'un algorithme MAB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du dispositif.
-D'autre part, les algorithmes MAB classiques ont tendance à changer souvent de bras, surtout au début du processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour le dispositif sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par le dispositif.
-Chaque reconfiguration matérielle coûte de l'énergie à l'appareil et les algorithmes de commutation rapide entraînent une réduction de l'autonomie de la batterie.
+Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes MAB pour optimiser la sélection des canaux, sur la durée de vie des piles d'un objet sans fil.
+D'une part, l'utilisation d'un algorithme MAB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du objet.
+D'autre part, les algorithmes MAB classiques ont tendance à changer souvent de bras, surtout au début du processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour le objet sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par le objet.
+Chaque reconfiguration matérielle coûte de l'énergie à l'objet et les algorithmes de commutation rapide entraînent une réduction de l'autonomie de la batterie.
 Le compromis entre les deux aspects est étudié de façon empirique dans les études suivantes
@@
 %
-Depuis $2017$, S. Darak et son équipe à l'IIIT Delhi en Inde, ont travaillé activement dans la recherche sur la radio intelligente à l'aide de bandits à armes multiples.
+Depuis $2017$, S. Darak et son équipe à l'IIIT Delhi en Inde, ont travaillé activement dans la recherche sur la radio intelligente à l'aide de bandits multi-bras.
 Héritage de son travail au sein de l'équipe SCEE,
@@
 La littérature susmentionnée a essentiellement montré que les algorithmes MAB peuvent être appliqués avec succès au problème de l'AOS.
-Toutefois, si l'on considère les dispositifs prêts à l'emploi qui ne peuvent pas effectuer la détection, tels que les dispositifs finaux à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle MAB qui utilise la détection pour détecter le UP dans le cas de l'AOS ne peut plus être appliqué.
-En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et SU.
+Toutefois, si l'on considère les $M$ objets l'emploi qui ne peuvent pas effectuer de détection active, tels que les $M$ objets faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle MAB qui utilise la détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
+En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et US.
 En d'autres termes, il n'y a plus de priorité entre les utilisateurs des réseaux IdO.
-(plus de UP, tous sont SU).
+(plus de UP, tous sont US).
 %
@@
     \item
-    nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre UP et SU),
+    nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre UP et US),
     \item
-    la plupart des dispositifs IdO sont des dispositifs à très faible coût, fonctionnant éventuellement sur du matériel de mauvaise qualité et ayant des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
+    la plupart des $M$ objets des $M$ objets faible coût, fonctionnant éventuellement sur du matériel de mauvaise qualité et ayant des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
     Cela signifie que même s'ils ne sont pas équipés de capacités de détection particulières,
@@
     \item
-    une seule station de base IdO devra gérer un très grand nombre d'appareils,
+    une seule station de base IdO devra gérer un très grand nombre d'objets,
     et ne peut leur envoyer des ordres de coordination afin d'optimiser le réseau,
     \item
-    les dispositifs IdO ont des cycles de fonctionnement faibles (seulement quelques messages par minute) à très faibles cycles de fonctionnement (seulement quelques messages par mois).
+    les $M$ objets ont des cycles de fonctionnement faibles à très faibles (seulement quelques messages par minute ou par mois).
 \end{itemize}
 
-Une coordination centrale des appareils avec leur station de base est impossible car il faudrait communiquer en permanence avec la station de base, ce qui viole les deux dernières contraintes.
+Une coordination centrale des objets avec leur station de base est impossible car il faudrait communiquer en permanence avec la station de base, ce qui viole les deux dernières contraintes.
 %
-De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de la batterie, les dispositifs IdO sont capables d'utiliser leur antenne reçue pour détecter un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), mais ils ne peuvent pas effectuer de détection de spectre comme cela est envisagé pour AOS et RI (\ie, au début de chaque intervalle, dans un schéma " écoute avant conversation ").
+De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de la batterie, les $M$ objets sont capables d'utiliser leur antenne reçue pour écouter dans un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), mais ils ne peuvent pas effectuer de détection de spectre à chaque instant, comme cela est envisagé pour l'AOS (\ie, au début de chaque intervalle, dans un schéma ``écoute avant transmission'').
 
-On pourrait penser qu'en l'absence de détection, l'apprentissage du renforcement n'est plus possible, mais tout appareil IdO réel reçoit encore des informations sur son environnement après certaines ou toutes les transmissions.
+On pourrait penser qu'en l'absence de détection, l'apprentissage du renforcement n'est plus possible, mais tout objet IdO réel reçoit encore des informations sur son environnement après certaines ou toutes les transmissions.
 Dans la plupart des normes IdO, un message de liaison montante envoyé à une station de base peut être suivi d'un message de liaison descendante renvoyé par la station de base pour indiquer si le message de liaison montante a été bien reçu et compris.
@@
 \section*{Nos contributions}
-\label{sec:1:contributions}
 
@@
 Si nous résumons les problèmes étudiés et les résumons en une seule question, ce pourrait être la suivante :
-\Insister sur le fait que l'on peut adapter les outils de prise de décision déjà appliqués avec succès à la radio intelligente pour l'accès opportuniste au spectre aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des objets ?
+\emph{peut-on adapter les outils de prise de décision déjà appliqués avec succès à la radio intelligente pour l'accès opportuniste au spectre aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des objets ?}
 %
@@
 %
-Nous avons commencé par explorer la riche littérature des bandits aux armes multiples,
-car il existe de nombreux algorithmes différents, avec de nombreuses variantes sur le problème simple présenté ci-dessus.
+Nous avons commencé par explorer la riche littérature des bandits multi-bras,
+car il existe de nombreux algorithmes différents, avec de nombreuses variantes du problème simple présenté ci-dessus.
 Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle MAB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires du MAB.
@@
 nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme MAB (son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
-mais aussi par leurs performances empiriques en termes de complexité de calcul et d'exigences de stockage, tant du point de vue des analyses théoriques que des mesures en temps réel et des empreintes mémoire.
+mais aussi par leurs performances empiriques en termes de complexité de calcul et de stockage, tant du point de vue des analyses théoriques que des mesures réelle.
 
@@
 Nous étudions deux modèles de réseaux IdO dans le Chapitre~\ref{chapter:4}.
-Le premier modèle considère les dispositifs IdO qui ont simplement des données à envoyer régulièrement à une station de base (à des moments imprévisibles ou aléatoires).
-De tels dispositifs utilisent les acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
-L'objectif de ce premier modèle est d'améliorer la qualité de service (QoS) de l'application de ce réseau IdO, en appliquant une RL décentralisée côté appareil afin de réduire le taux d'échec de transmission.
-La station de base recevra donc plus de paquets de liaison montante des appareils desservis, si elle peut apprendre à utiliser un système efficace d'accès au spectre.
-Cela implique que le réseau peut accepter plus d'équipements finaux tout en maintenant la même QoS.
+Le premier modèle considère les $M$ objets ont simplement des données à envoyer régulièrement à une station de base (à des moments imprévisibles ou aléatoires).
+De tels $M$ objets acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
+L'objectif de ce premier modèle est d'améliorer la qualité de service (QoS) de l'application de ce réseau IdO, en appliquant une RL décentralisée côté objet afin de réduire le taux d'échec de transmission.
+La station de base recevra donc plus de paquets de liaison montante des objets desservis, si elle peut apprendre à utiliser un système efficace d'accès au spectre.
+Cela implique que le réseau peut accepter plus d'objets tout en maintenant la même QoS.
 %
-Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet d'augmenter la durée de vie des piles de chacun des appareils, ainsi que la QoS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
+Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet d'augmenter la durée de vie des piles de chacun des objets, ainsi que la QoS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
 
@@
 %
-Dans les deux modèles présentés ci-dessus, le concept est de laisser chaque dispositif dynamique d'un réseau IdO exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
-Il est important de noter que chaque dispositif ne communique, et donc n'apprend, qu'à quelques tours et non à toutes les étapes, en suivant son propre processus d'activation (aléatoire) (nous nous limitons à un processus d'activation de Bernoulli purement aléatoire de probabilité $p$).
+Dans les deux modèles présentés ci-dessus, le concept est de laisser chaque objet dynamique d'un réseau IdO exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
+Il est important de noter que chaque objet ne communique, et donc n'apprend, qu'à quelques tours et non à toutes les étapes, en suivant son propre processus d'activation (aléatoire) (nous nous limitons à un processus d'activation de Bernoulli purement aléatoire de probabilité $p$).
 Cela signifie que chacun d'eux vise son propre objectif local, qui est de maximiser sa récompense cumulée (d'une manière égoïste).
 Il est bien connu dans la littérature sur la théorie des jeux que le jeu égoïste peut être désastreux pour la mesure centralisée de la performance, on peut penser aux "dilemmes" populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
-Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage MAB décentralisé et égoïste conduisait à une coordination efficace entre les dispositifs dans tous les scénarios considérés, malgré le fait que ces objets IdO ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
+Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage MAB décentralisé et égoïste conduisait à une coordination efficace entre les $M$ objets les scénarios considérés, malgré le fait que ces objets IdO ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
 
 Nous avons d'abord essayé d'analyser la performance de cet algorithme décentralisé, que nous appelons \Selfish, dans le modèle de Section~\ref{sec:4:firstModel},
-mais en raison du nombre aléatoire de dispositifs actifs à chaque pas de temps (\ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
+mais en raison du nombre aléatoire de $M$ objets chaque pas de temps (\ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
 %
-C'est pourquoi dans le Chapitre~\ref{chapter:5}, nous relâchons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$, qui est équivalent à avoir $p < 1$, et nous considérons le cas des dispositifs IdO communiquant à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ dispositifs à jour.
+C'est pourquoi dans le Chapitre~\ref{chapter:5}, nous relâchons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$, qui est équivalent à avoir $p < 1$, et nous considérons le cas des $M$ objets à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ $M$ objets.
 Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas AOS (\ie, avec détection) ou le cas IdO (\ie, sans détection).
@@
 car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
-Il diffère des modèles précédemment étudiés d'application du MAB pour l'AOS, puisqu'ils ne prennent en compte que les dispositifs synchronisés, une unité de base et une unité centrale de traitement, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'unité de base sera certainement reçu avec succès par la base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
+Il diffère des modèles précédemment étudiés d'application du MAB pour l'AOS, puisqu'ils ne prennent en compte que les $M$ objets une unité de base et une unité centrale de traitement, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'unité de base sera certainement reçu avec succès par la base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
 
-D'une part, dans le cadre de la LVMO, nous n'avons pas réussi à obtenir un résultat positif pour la politique égoïste, car nous avons prouvé que dans certains contextes limités (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre de regret moyen linéaire.
+D'une part, dans le cadre de l'AOS, nous n'avons pas réussi à obtenir un résultat positif pour la politique égoïste, car nous avons prouvé que dans certains contextes limités (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre de regret moyen linéaire.
 Ce résultat a ensuite été confirmé et analysé par \cite{BoursierPerchet18}.
@@
 D'autre part, nous avons été en mesure de proposer de nouveaux algorithmes pour ce problème de bandit multi-joueurs avec des informations de détection, et nous avons analysé notre proposition \MCTopM-\klUCB, pour montrer qu'elle atteint une limite supérieure de regret logarithmique à temps fini, améliorant considérablement les résultats précédents.
-Notre algorithme atteint également un nombre logarithmique de collisions et de commutateurs de bras, et permet à un groupe fixe de dispositifs $M$ d'apprendre efficacement à utiliser les meilleurs canaux $M$ orthogonalement pour presque toutes leurs communications montantes.
+Notre algorithme atteint également un nombre logarithmique de collisions et de commutateurs de bras, et permet à un groupe fixe de $M$ objets d'apprendre efficacement à utiliser les $M$ meilleurs canaux orthogonalement pour presque toutes leurs communications montantes.
 %
@@
 
-Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux équipements finaux, tous suivant des modèles d'activation aléatoires.
+Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux objets, tous suivant des modèles d'activation aléatoires.
 La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes MAB, conçus pour résoudre des problèmes stationnaires, sur un problème non stationnaire,
 mais plutôt que nous analysons des algorithmes qui jouent tous dans différents sous-ensembles (aléatoires) des pas de temps synchronisés globaux.
-Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et appliquer au maximum $M \leq K$ aux dispositifs n'est pas vraiment réaliste pour les réseaux IdO, même si cela conduit au modèle du Chapitre~\ref{chapter:5}.
+Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et imposer un maximum de $M \leq K$ objets n'est pas vraiment réaliste pour les réseaux IdO, même si cela conduit au modèle du Chapitre~\ref{chapter:5}.
 
@@
 Cependant, même si nous continuons à supposer un temps synchronisé dans tout le reste de la thèse,
-il est difficile de soutenir que cette hypothèse d'un calendrier centralisé pour les équipements finaux peut être réaliste, car nous étudions le cas de l'apprentissage décentralisé pour l'AOS et l'IdO précisément afin d'éviter la signalisation et tout contrôle central des équipements par la station de base, comme nous l'avons expliqué ci-dessus.
+il est difficile de soutenir que cette hypothèse d'un calendrier centralisé pour les objets peut être réaliste, car nous étudions le cas de l'apprentissage décentralisé pour l'AOS et l'IdO précisément afin d'éviter la signalisation et tout contrôle central des objets par la station de base, comme nous l'avons expliqué ci-dessus.
 
-D'autre part, une autre façon de relâcher l'hypothèse non stationnaire est de prendre le point de vue d'un dispositif, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur une unité de soins entourée de plusieurs UP ayant des comportements fixes.
-Si nous nous concentrons sur un dispositif IdO, son environnement (\ie, les dispositifs environnants) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réalistes donnent généralement une certaine structure sur cette non-stationnarité.
+D'autre part, une autre façon de relâcher l'hypothèse non stationnaire est de prendre le point de vue d'un objet, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur une unité de soins entourée de plusieurs UP ayant des comportements fixes.
+Si nous nous concentrons sur un objet IdO, son environnement (\ie, les $M$ objets) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réalistes donnent généralement une certaine structure sur cette non-stationnarité.
 Par exemple, l'application de la stationnarité à certains intervalles de temps conduit à la dernière contribution et au dernier chapitre de cette thèse.
@@
 ce qui signifie que le problème de bandit sous-jacent est stationnaire à certains intervalles, séparés par des points de changement situés à des moments inconnus.
-Supposer que l'environnement est stationnaire au niveau de la pièce est en effet logique pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'équipements terminaux dans le réseau.
+Supposer que l'environnement est stationnaire au niveau de la pièce est en effet logique pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'objets dans le réseau.
 Des exemples d'une telle situation peuvent être une nouvelle entreprise arrivant sur le marché dans une ville (par exemple, les comptoirs Linky qui sont installés aujourd'hui en France), ou l'agriculteur voisin installant des capteurs sur son propre troupeau de vaches, etc.
@@
 Deux grandes familles d'algorithmes ont été proposées pour le problème stationnaire par pièce,
-qui consistent généralement à combiner une politique efficace conçue pour le problème stationnaire du MAB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des armes.
+qui consistent généralement à combiner une politique efficace conçue pour le problème stationnaire du MAB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des bras.
 Les politiques passivement adaptatives utilisent une fenêtre de taille fixe ou évolutive \cite{Garivier11UCBDiscount}, ou un facteur d'actualisation, pour oublier les observations passées \cite{Kocsis06,Gupta11thompson},
@@
     \item
-    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les appareils dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes MAB, pour augmenter automatiquement la durée de vie de leur batterie.
+    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les objets dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes MAB, pour augmenter automatiquement la durée de vie de leur batterie.
     Cela permet également à davantage d'objets d'utiliser le même réseau tout en maintenant un taux élevé d'accès au spectre sans souffrir de collisions radio

### 2019-07-07 2-Chapters/1-Chapter/chapter1.tex
Almost done for this French translation of Ch1, useless as I'll remove it tonight, but still...

@@
 A common problem is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and will run on a tiny battery, which lifespan should be maximized.
-For instance, most commercial IoT companies nowadays sell a lifespan more than $10$ years, like SIGFOX \cite{Centenaro16}.
+Most commercial IoT companies nowadays indeed sell a lifespan larger than $10$ years, like SIGFOX \cite{Centenaro16}.
 Another common constraint for IoT devices is their low duty cycles, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems (such as 4G/5G and WiFi).
@@
 and the \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}) layer.
-To quote some examples of standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications, while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications.
+To quote some examples of standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications, while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications (up-to \SI{50}{\kilo\meter}).
 We refer to the survey \cite{Centenaro16} for more details, and references in \cite{Azari18} or our recent article \cite{MoyBesson2019}.
@@
 In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the network.
-As exposed in the definition above of the CR philosophy, we could imagine that if the network is not used by any PU in a certain location and time, then non licensed users, referred to as Secondary User (SU), could use it, by paying some rent to the PU.
+As exposed in the definition above of the CR philosophy, we could imagine that if the network is not used by any PU in a certain location and time, then non licensed users, referred to as Secondary User (SU), could use it, possibly by paying some rent to the network operator.
 %
 Regulation states that PU have a strict priority,
-but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU was equipped with an efficient sensing capacity, it could analyze its environment, and use a licensed band if it is free of any PU.
+but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU was equipped with an efficient sensing capacity, it could analyze its environment, and use a licensed band if and only if it is free of any PU.
 % explain DSA
@@
 %
-MAB are a simple yet powerful example of the well-known \emph{exploration/exploitation dilemma}:
+MAB are a simple yet powerful example of the well-known \emph{exploration vs exploitation dilemma}:
 when facing a set of $K$ actions whose effects on the environment are unknown, a learner must balance between

### 2019-07-07 1-Preamble/5-Abstract/resume_fr.tex
Dernière grosse relecture du résumé en français, avant la compression

@@
 % Chapitre 2
-Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (MAB, Chapitre~\ref{chapter:2}), mais leur analyse est complexe,
+Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (BMB, ou \emph{Multi-Armed Bandits}, MAB, en Chapitre~\ref{chapter:2}), mais leur analyse est complexe,
 % , Chapitre~\ref{chapter:2}
@@
 % Chapitre 3
-Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres (Chapitre~\ref{chapter:3}).
+Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes BMB, qui couvre tous les modèles étudiés et d'autres (Chapitre~\ref{chapter:3}).
 
@@
 %
-Nous sommes intéressés par l'apprentissage séquentiel et en particulier par les Bandits Multi-Bras (\emph{Multi-Armed Bandits}, MAB),
+Nous sommes intéressés par l'apprentissage séquentiel et en particulier par les Bandits Multi-Bras (BMB),
 qui est apparu comme un problème intéressant dans la communauté statistique et parmi les chercheurs intéressés par l'apprentissage séquentiel, comme le montrent les travaux antérieurs de \cite{Thompson33,Robbins52,LaiRobbins85}.
@@
 En d'autres termes, la joueuse ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une autre action.
-Ce genre de rétroaction limitée s'appelle \emph{information sur les bandits}, et nous discutons de l'histoire et du concept des \emph{Multi-Armed Bandits}.
-(MAB) en détail dans le prochain Chapitre~\ref{chapter:2}.
-Les essais cliniques et la découverte de traitements ont été historiquement la première application du MAB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33}.
+Ce genre de rétroaction limitée s'appelle l'information de type bandit (\emph{bandit feedback}), et nous discutons de l'histoire et du concept des \emph{bandits multi-bras} (BMB, ou MAB) en détail dans le prochain Chapitre~\ref{chapter:2}.
+Les essais cliniques et la découverte de traitements ont été historiquement la première application des BMB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33}.
 %
-Le MAB est un exemple simple mais puissant du dilemme bien connu de entre \emph{exploration et exploitation} :
+Le BMB est un exemple simple mais puissant du dilemme bien connu de entre \emph{exploration et exploitation} :
 lorsqu'elle est confronté à un ensemble d'actions de $K$ dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
@@
 %
-Le problème du MAB a été étudié à la fois dans la communauté de l'apprentissage machine et de la statistique, depuis les $1950$ avec des pionniers comme H.
+Le problème du BMB a été étudié à la fois dans la communauté de l'apprentissage machine et de la statistique, depuis les $1950$ avec des pionniers comme H.
 Robbins \cite{Robbins52}, et plus récemment c'est un domaine de recherche actif, depuis les $1990$ \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-La recherche sur le MAB a produit une vaste littérature dans le domaine des $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet actif, comme l'illustrent les livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications des MAB dans dernières années \cite{bouneffouf2019survey}.
+La recherche sur le BMB a produit une vaste littérature dans le domaine des $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet actif, comme l'illustrent les livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications des BMB dans dernières années \cite{bouneffouf2019survey}.
 
@@
 %
-Les deux communautés de l'AOS et du MAB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des $2000$ et le début des $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
+Les deux communautés de l'AOS et du BMB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des $2000$ et le début des $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
 L'accent est mis sur un US accédant à un spectre sous licence, occupé par un US, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès d'``écoute avant transmission''.
@@
 
-\paragraph{MAB pour l'AOS, et un bref historique de ces recherches par l'équipe SCEE.}
+\paragraph{BMB pour l'AOS, et un bref historique de ces recherches par l'équipe SCEE.}
 %
-Les travaux antérieurs de notre équipe SCEE ont montré que le MAB peut être utilisé pour modéliser le problème de l'AOS :
+Les travaux antérieurs de notre équipe SCEE ont montré que le BMB peut être utilisé pour modéliser le problème de l'AOS :
 les bandes de fréquences orthogonales (ou canaux) sont des modèles par bras $k\in\{1,\dots,K\}$,
@@
 
-Leurs travaux ont été parmi les premiers à proposer l'utilisation de Reinforcement Learning for Cognitive Radio et le problème AOS, notamment le modèle MAB et l'algorithme $\UCB_1$,
+Leurs travaux ont été parmi les premiers à proposer l'utilisation de Reinforcement Learning for Cognitive Radio et le problème AOS, notamment le modèle BMB et l'algorithme $\UCB_1$,
 ainsi que les premiers travaux de Q. Zhao et de son équipe, par exemple avec les travaux \cite{Liu08,Zhao10}.
@@
 Dans une deuxième thèse de doctorat \cite{Modi17PhD}, N.
-Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes MAB pour optimiser la sélection des canaux, sur la durée de vie des piles d'un objet sans fil.
-D'une part, l'utilisation d'un algorithme MAB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du objet.
-D'autre part, les algorithmes MAB classiques ont tendance à changer souvent de bras, surtout au début du processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour le objet sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par le objet.
+Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes BMB pour optimiser la sélection des canaux, sur la durée de vie des piles d'un objet sans fil.
+D'une part, l'utilisation d'un algorithme BMB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du objet.
+D'autre part, les algorithmes BMB classiques ont tendance à changer souvent de bras, surtout au début du processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour le objet sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par le objet.
 Chaque reconfiguration matérielle coûte de l'énergie à l'objet et les algorithmes de commutation rapide entraînent une réduction de l'autonomie de la batterie.
@@
 %
-Certaines analyses des mesures radio réelles effectuées pour les canaux ionosphériques HF ont également prouvé que les solutions basées sur l'apprentissage MAB sont appropriées et résolvent efficacement ce type de problèmes de prise de décision sur les signaux sans fil du monde réel \cite{Melian15}.
+Certaines analyses des mesures radio réelles effectuées pour les canaux ionosphériques HF ont également prouvé que les solutions basées sur l'apprentissage BMB sont appropriées et résolvent efficacement ce type de problèmes de prise de décision sur les signaux sans fil du monde réel \cite{Melian15}.
 %
@@
 %
-La littérature susmentionnée a essentiellement montré que les algorithmes MAB peuvent être appliqués avec succès au problème de l'AOS.
-Toutefois, si l'on considère les $M$ objets l'emploi qui ne peuvent pas effectuer de détection active, tels que les $M$ objets faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle MAB qui utilise la détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
+La littérature susmentionnée a essentiellement montré que les algorithmes BMB peuvent être appliqués avec succès au problème de l'AOS.
+Toutefois, si l'on considère les $M$ objets l'emploi qui ne peuvent pas effectuer de détection active, tels que les $M$ objets faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle BMB qui utilise la détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
 En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et US.
@@
 
-\subsection*{Explorer la jungle des algorithmes MAB}
+\subsection*{Explorer la jungle des algorithmes BMB}
 
@@
 car il existe de nombreux algorithmes différents, avec de nombreuses variantes du problème simple présenté ci-dessus.
-Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle MAB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires du MAB.
+Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle BMB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires du BMB.
 %
 Afin de bien comprendre quels algorithmes pourraient être adaptés aux contraintes susmentionnées de la RI pour les réseaux IdO,
-nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme MAB (son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
+nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme BMB (son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
 mais aussi par leurs performances empiriques en termes de complexité de calcul et de stockage, tant du point de vue des analyses théoriques que des mesures réelle.
@@
 %
-Notre exploration du grand nombre d'algorithmes et de modèles MAB développés dans la littérature récente.
+Notre exploration du grand nombre d'algorithmes et de modèles BMB développés dans la littérature récente.
 nous a donné l'ambition d'écrire un seul logiciel permettant à chacun d'implémenter facilement de nouveaux modèles et algorithmes, afin de comparer les modèles existants et d'explorer empiriquement les performances des nouveaux algorithmes proposés.
-Pour atteindre cet objectif, nous avons produit une bibliothèque Python de simulation des problèmes MAB.
+Pour atteindre cet objectif, nous avons produit une bibliothèque Python de simulation des problèmes BMB.
 %
-Nous avons écrit la bibliothèque de simulation open-source la plus exhaustive pour les problèmes MAB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
+Nous avons écrit la bibliothèque de simulation open-source la plus exhaustive pour les problèmes BMB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
 Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
@@
 %
-Étant donné le grand nombre d'algorithmes MAB différents, nous nous intéressons aussi à la question de savoir comment un praticien peut choisir l'algorithme à mettre en œuvre, dans un objet IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
+Étant donné le grand nombre d'algorithmes BMB différents, nous nous intéressons aussi à la question de savoir comment un praticien peut choisir l'algorithme à mettre en œuvre, dans un objet IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
 Pour répondre à cette question, nous présentons deux approches.
@@
 La seconde approche est une sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons dans Chapitre~\ref{chapter:25}.
-Ce travail sur l'agrégation des algorithmes MAB a été motivé pour le cas AOS de la radio cognitive, où de nombreux travaux de recherche antérieurs ne prenaient en compte que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
+Ce travail sur l'agrégation des algorithmes BMB a été motivé pour le cas AOS de la radio cognitive, où de nombreux travaux de recherche antérieurs ne prenaient en compte que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
 Cette contribution a été présentée à la conférence IEEE WCNC à Barcelone, Espagne, en avril 2018.
@@
 
-\subsection*{Nos modèles de réseaux d'IdO et de MAB décentralisé}
+\subsection*{Nos modèles de réseaux d'IdO et de BMB décentralisé}
 
-Dans la deuxième Partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux d'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du MAB, issues de la simplification de notre premier modèle.
+Dans la deuxième Partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux d'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du BMB, issues de la simplification de notre premier modèle.
 
@@
 Il est bien connu dans la littérature sur la théorie des jeux que le jeu égoïste peut être désastreux pour la mesure centralisée de la performance, on peut penser aux "dilemmes" populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
-Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage MAB décentralisé et égoïste conduisait à une coordination efficace entre les $M$ objets les scénarios considérés, malgré le fait que ces objets IdO ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
+Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage BMB décentralisé et égoïste conduisait à une coordination efficace entre les $M$ objets les scénarios considérés, malgré le fait que ces objets IdO ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
 
@@
 Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas AOS (\ie, avec détection) ou le cas IdO (\ie, sans détection).
-L'objectif était de comprendre l'heuristique du Chapitre~\ref{chapter:4}, \Selfish, dans le cadre plus simple du multi-player MAB, qui a été étudié auparavant pour le cas de l'AOS, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
+L'objectif était de comprendre l'heuristique du Chapitre~\ref{chapter:4}, \Selfish, dans le cadre plus simple du BMB multi-joueur, qui a été étudié auparavant pour le cas de l'AOS, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
 %
@@
 car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
-Il diffère des modèles précédemment étudiés d'application du MAB pour l'AOS, puisqu'ils ne prennent en compte que les $M$ objets une unité de base et une unité centrale de traitement, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'unité de base sera certainement reçu avec succès par la base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
+Il diffère des modèles précédemment étudiés d'application du BMB pour l'AOS, puisqu'ils ne prennent en compte que les $M$ objets une unité de base et une unité centrale de traitement, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'unité de base sera certainement reçu avec succès par la base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
 
@@
 Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux objets, tous suivant des modèles d'activation aléatoires.
-La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes MAB, conçus pour résoudre des problèmes stationnaires, sur un problème non stationnaire,
+La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes BMB, conçus pour résoudre des problèmes stationnaires, sur un problème non stationnaire,
 mais plutôt que nous analysons des algorithmes qui jouent tous dans différents sous-ensembles (aléatoires) des pas de temps synchronisés globaux.
@@
 %
-Afin de comprendre également avec précision comment les algorithmes MAB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur le MAB accusatoire et sur le MAB non stationnaire, pour les deux cas d'environnements variant lentement et changeant brusquement.
+Afin de comprendre également avec précision comment les algorithmes BMB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur les BMB adverse ou non stationnaire, pour les deux cas d'environnements variant lentement et changeant brusquement.
 Nous commençons notre dernier Chapitre~\ref{chapter:6} en passant en revue les travaux existants,
@@
 Deux grandes familles d'algorithmes ont été proposées pour le problème stationnaire par pièce,
-qui consistent généralement à combiner une politique efficace conçue pour le problème stationnaire du MAB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des bras.
+qui consistent généralement à combiner une politique efficace conçue pour le problème stationnaire du BMB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des bras.
 Les politiques passivement adaptatives utilisent une fenêtre de taille fixe ou évolutive \cite{Garivier11UCBDiscount}, ou un facteur d'actualisation, pour oublier les observations passées \cite{Kocsis06,Gupta11thompson},
@@
 Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
-Elles se situent à trois niveaux : des modèles MAB pour plusieurs contextes radio liés à la radio intelligente ; des preuves théoriques associées aux problèmes d'apprentissages correspond ; et enfin des implémentations d'algorithmes à but de simulations, de partage avec la communauté, et de validation par des démonstrations radio réalistes.
+Elles se situent à trois niveaux : des modèles BMB pour plusieurs contextes radio liés à la radio intelligente ; des preuves théoriques associées aux problèmes d'apprentissages correspond ; et enfin des implémentations d'algorithmes à but de simulations, de partage avec la communauté, et de validation par des démonstrations radio réalistes.
 
@@
     \item
-    Nous avons écrit la plus complète bibliothèque de simulation pour les problèmes MAB, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBanditsJMLR,SMPyBandits}.
+    Nous avons écrit la plus complète bibliothèque de simulation pour les problèmes BMB, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBanditsJMLR,SMPyBandits}.
     Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
@@
     \item
-    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les objets dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes MAB, pour augmenter automatiquement la durée de vie de leur batterie.
+    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les objets dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes BMB, pour augmenter automatiquement la durée de vie de leur batterie.
     Cela permet également à davantage d'objets d'utiliser le même réseau tout en maintenant un taux élevé d'accès au spectre sans souffrir de collisions radio
@@
     \item
-    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle MAB multi-joueurs, qui a été particulièrement active ces deux dernières années.
+    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle BMB multi-joueurs, qui a été particulièrement active ces deux dernières années.
     % qui, selon nous, n'a jamais été écrit auparavant.
@@
         \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!15] {\textbf{Chapitre~\ref{chapter:2}}\\Modèle bandits multi-bras\\stochastique et stationnaire};
-        \node[align=center] (chapter3) at (-2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:3}}\\SMPyBandits : bibliothèque\\de simulations pour MAB};
-        \node[align=center] (chapter25) at (2.5,2.25) [rectangle,draw,fill=red!20] {\textbf{Chapitre~\ref{chapter:25}}\\Sélection séquentielle\\d'algorithme MAB};
-        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles MAB\\pour les réseaux IdO};
-        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!15] {\textbf{Chapitre~\ref{chapter:5}}\\Modèle MAB\\Multi-joueurs};
-        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!20] {\textbf{Chapitre~\ref{chapter:6}}\\Modèle MAB\\Non-stationnaire};
+        \node[align=center] (chapter3) at (-2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapitre~\ref{chapter:3}}\\SMPyBandits : bibliothèque\\de simulations pour BMB};
+        \node[align=center] (chapter25) at (2.5,2.25) [rectangle,draw,fill=red!20] {\textbf{Chapitre~\ref{chapter:25}}\\Sélection séquentielle\\d'algorithme BMB};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles BMB\\pour les réseaux IdO};
+        \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!15] {\textbf{Chapitre~\ref{chapter:5}}\\Modèle BMB\\Multi-joueurs};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!20] {\textbf{Chapitre~\ref{chapter:6}}\\Modèle BMB\\Non-stationnaire};
         \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!20] {\textbf{Chapitre~\ref{chapter:conclusion}}\\Conclusion générale};
@@
     \item
-\textcolor{darkred}{Dans la Partie~\ref{part:Introduction}}, nous commençons par le Chapitre~\ref{chapter:2} qui présente les modèles de bandits multi-bras (MAB), les concepts et les notations utilisés dans tout ce document.
+\textcolor{darkred}{Dans la Partie~\ref{part:Introduction}}, nous commençons par le Chapitre~\ref{chapter:2} qui présente les modèles de bandits multi-bras (BMB), les concepts et les notations utilisés dans tout ce document.
 Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
@@
 % qui n'est pas non plus nécessaire à la compréhension globale du manuscrit, mais
-qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes MAB.
+qui détaille la première contribution : un nouvel algorithme pour la sélection séquentielle d'algorithmes BMB.
 
@@
 \textcolor{darkgreen}{La deuxième Partie~\ref{part:MABIOT}} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus quasiment indépendamment.
-Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux IdO, dans lesquels les algorithmes MAB ont été utilisés avec succès.
+Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux IdO, dans lesquels les algorithmes BMB ont été utilisés avec succès.
 Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop complexes pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
@@
 afin d'établir des preuves mathématiques garantissant la convergence, ainsi que les gains en performance apportés par ces algorithmes de bandit.
-Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du Chapitre~\ref{chapter:2} et les modèles de réseaux IdO du Chapitre~\ref{chapter:4}.
+Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle BMB stationnaire à un joueur du Chapitre~\ref{chapter:2} et les modèles de réseaux IdO du Chapitre~\ref{chapter:4}.
 Ces deux études ont chacune donné des résultats théoriques à la pointe de la recherche sur les modèles de bandits multi-joueurs et de bandits stationnaires par morceaux, que nous avons aussi validé par des expériences numériques.

### 2019-07-07 1-Preamble/5-Abstract/resume_fr.tex
Je relis tout le ch1 traduit en français, ça aide à trouver des typos sur la version anglaise

@@
 J'étais supervisé par le professeur Christophe Moy, à Rennes,
-et j'étais également co-encadré par le Docteur Emilie Kaufmann, que j'ai visité plusieurs fois à l'Inria Lille Nord Europe à Lille (France).
+et j'étais également co-encadré par le docteur Emilie Kaufmann, que j'ai visité de nombreuses fois à Inria Lille Nord Europe à Lille (France).
 
-% Chapitre 1
-Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les objets reconfigurables qui peuvent accéder à des réseaux de radio sans fil de manière intelligente, dans des bandes non licenciées et sans supervision centrale.
-Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IdO), avec l'objectif d'augmenter la durée de vie de la batterie des objets, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil (Chapitre~\ref{chapter:1}).
-% Chapitre 4
-Nous proposons différents modèles de réseaux IdO, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement (Chapitre~\ref{chapter:4}).
-% Chapitre 2
-Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (BMB, ou \emph{Multi-Armed Bandits}, MAB, en Chapitre~\ref{chapter:2}), mais leur analyse est complexe,
-% , Chapitre~\ref{chapter:2}
-notamment si l'on veut prouver la convergence de nombreux objets jouant à un jeu collaboratif sans aucune coordination, lorsque les objets suivent tous un modèle d'activation aléatoire.
-% En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
-% Chapitres 5 et 6
-Le reste de ce manuscrit étudie donc deux modèles restreints, afin de pouvoir en faire une analyse théorique exhaustive.
-Nous considérons d'abord des bandits multi-joueurs dans des problèmes stationnaires (Chapitre~\ref{chapter:5}), puis des bandits mono-joueurs non stationnaires (Chapitre~\ref{chapter:6}).
-% Chapitre 3
-Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes BMB, qui couvre tous les modèles étudiés et d'autres (Chapitre~\ref{chapter:3}).
+% % Chapitre 1
+% Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les objets reconfigurables qui peuvent accéder à des réseaux de radio sans fil de manière intelligente, dans des bandes non licenciées et sans supervision centrale.
+% Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IdO), avec l'objectif d'augmenter la durée de vie de la batterie des objets, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil (Chapitre~\ref{chapter:1}).
+% % Chapitre 4
+% Nous proposons différents modèles de réseaux IdO, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement (Chapitre~\ref{chapter:4}).
+% % Chapitre 2
+% Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Bras (BMB, ou \emph{Multi-Armed Bandits}, MAB, en Chapitre~\ref{chapter:2}), mais leur analyse est complexe,
+% % , Chapitre~\ref{chapter:2}
+% notamment si l'on veut prouver la convergence de nombreux objets jouant à un jeu collaboratif sans aucune coordination, lorsque les objets suivent tous un modèle d'activation aléatoire.
+% % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
+% % Chapitres 5 et 6
+% Le reste de ce manuscrit étudie donc deux modèles restreints, afin de pouvoir en faire une analyse théorique exhaustive.
+% Nous considérons d'abord des bandits multi-joueurs dans des problèmes stationnaires (Chapitre~\ref{chapter:5}), puis des bandits mono-joueurs non stationnaires (Chapitre~\ref{chapter:6}).
+% % Chapitre 3
+% Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes BMB, qui couvre tous les modèles étudiés et d'autres (Chapitre~\ref{chapter:3}).
 
 % ----------------------------------------------------------------------------
-\section*{Contexte}
+\section*{Contexte de cette thèse}
 
@@
 % https://en.wikipedia.org/wiki/History_of_radio
-Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, la télé-communication sans fil entre des appareils fabriqués a été rendue possible, et de plus en plus fréquentes dans nos vies.
-Avec l'avènement des réseaux Internet des Objets (IdO), des milliards de $M$ objets basse consommation devraient être déployés dans le monde entier, pour un large éventail d'applications différentes.
+Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, la télécommunication sans fil entre des appareils fabriqués par l'homme a été rendue possible, et de plus en plus fréquentes dans nos vies.
+Avec l'avènement des réseaux de l'Internet des Objets (IdO), des milliards d'objets à basse consommation devraient être déployés dans le monde entier, permettant un large éventail d'applications différentes.
 Il existe aujourd'hui un consensus mondial sur le fait qu'avec la tendance actuelle à l'accroissement démographique et la crise énergétique actuelle, toute nouvelle technologie déployée doit être à la fois bon marché et efficace sur le plan énergétique,
@@
 %
-En plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, pour tous les cas possibles,
+En plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, couvrant tous les cas possibles,
 le moment est venu de le combiner avec un apprentissage machine bon marché et prometteur pour (essayer d') atteindre le niveau de gain de performance nécessaire pour que les promesses de l'IdO deviennent réalité.
 
-C'est pourquoi nous nous intéressons à cette thèse sur
+C'est pourquoi nous nous intéressons dans cette thèse sur
 l'amélioration de la durée de vie des batteries des objets IdO et la réduction du coût énergétique des réseaux IdO.
-Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant la prise de décision décentralisée à faible coût directement dans les futurs $M$ objets
+Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant une prise de décision décentralisée à faible coût, directement dans les futurs objets IdO.
 %
-Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (Multi-Armed Bandit algorithms), afin de permettre aux objets IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
+Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (algorithmes de types bandits multi-bras), afin de permettre aux objets IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
 
@@
 Ce troisième type de systèmes peut être désigné comme décentralisé :
-même si une station de base centrale est toujours en charge de nombreux objets (appareils),
+même si une station de base centrale est toujours en charge de nombreux objets (ou appareils),
 ce sont les objets qui déclenchent l'envoi de paquets radio, et les seules données descendantes qu'ils peuvent recevoir sont de courts accusés de réception (\emph{acknowledgements}, \Ack), envoyés par la station de base pour indiquer le succès ou l'échec de chaque paquet envoyé.
@@
 
-Pour le développement futur des ``réseaux intelligents'', des ``villes intelligentes'', des ``maisons intelligentes'' ou de l'``agriculture intelligente'', les réseaux de capteurs devraient être largement déployés.
+Pour le développement futur des ``réseaux intelligents'' (\emph{smart grid}), des ``villes intelligentes'' (\emph{smart cities}), des ``maisons intelligentes'' (\emph{smart homes}) ou de l'``agriculture intelligente'' (\emph{smart agriculture}), des réseaux de capteurs devraient être largement déployés.
 Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les ``bâtiments connectés'' et l'``agriculture connectée''.
 Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides et d'utiliser des réseaux de capteurs afin d'obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, dans le but de permettre au contrôle centralisé du chauffage d'optimiser son coût et sa consommation énergétique.
-Pour l'agriculture, un exemple peut être d'équiper chaque tête de bétail (dans les grandes fermes) d'un capteur qui émet régulièrement des informations biologiques, telles que la température corporelle ou le niveau de stress, afin d'optimiser l'heure de traite des vaches, de surveiller la santé des animaux, etc.
+Pour l'agriculture, un exemple peut être d'équiper chaque tête de bétail (dans les grandes fermes) de capteurs qui émettent régulièrement des informations biologiques, telles que la température corporelle ou le niveau de stress, afin d'optimiser l'heure de traite des vaches, de surveiller la santé des animaux, etc.
 
@@
 En effet, un contrôle centrale nécessite des paquets de signalisation qui ont été identifiés comme trop lourds pour ce type de systèmes.
-Dans les réseaux IdO actuels et futurs, de nombreux $M$ objets hétérogènes utilisent la même antenne pour des applications différentes.
-Un problème commun est que ces objets IdO ont une forte contrainte sur leur consommation d'énergie, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
+Dans les réseaux IdO actuels et futurs, de nombreux objets hétérogènes utilisent la même station de base (aussi appelée point d'accès, ou \emph{base station}, \emph{access point} ou \emph{gateway}) pour des applications différentes.
+Un problème commun est la forte contrainte en terme de consommation d'énergie de ces objets IdO, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
 Ainsi, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des objets ayant une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
-Une autre contrainte commune aux objets est leur faible besoin en communication, car la plupart des applications n'auront besoin que d'un à quelques messages à envoyer chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
+Une autre contrainte commune aux objets IdO est leur faible besoin en communication, car la plupart des applications n'auront besoin d'envoyer qu'un ou quelques messages chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
 %
@@
 et la couche \emph{M}edium \emph{A}access \emph{C}ontrol (\emph{MAC}).
-Pour citer quelques exemples de normes, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée, tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
+Pour citer quelques exemples de normes pour l'IdO, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée (jusqu'à \SI{2}{\meter}), tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
 Nous nous référons à l'article \cite{Centenaro16} pour plus de détails, et les références dans \cite{Azari18} ou notre récent article \cite{MoyBesson2019}.
@@
 %
-Les organismes de réglementation dans le monde, comme le
-\href{https://www.itu.int/en/Pages/default.aspx}{International Telecommunication Union} (voir \href{https://www.itu.int/}{\texttt{www.ITU.int}}),
-la \href{https://www.fcc.gov/}{Federal Communications Commission} aux États-Unis (voir \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
-ou la \href{https://cept.org}{ Conférence européenne des administrations des postes et des télécommunications} en Europe (voir \href{https://www.CEPT.org/}{\texttt{CEPT.org}}),
+Les organismes de réglementation dans le monde, comme l'
+\href{https://www.itu.int/en/Pages/default.aspx}{Union Internationale des Télécommunications} (\emph{ITU}, voir \href{https://www.itu.int/}{\texttt{www.ITU.int}}),
+la \href{https://www.fcc.gov/}{Commission Fédérale des Communications} aux États-Unis d'Amérique du Nord (\emph{FCC}, voir \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
+ou la \href{https://cept.org}{Conférence Européenne des Administrations des Postes et des Télécommunications} en Europe (\emph{CEPT}, voir \href{https://www.CEPT.org/}{\texttt{CEPT.org}}),
 ainsi que différentes campagnes de mesure indépendantes, ont montré que la plupart des fréquences du spectre des fréquences radio-électriques sont utilisées de manière inefficace.
@@
 Les bandes des réseaux cellulaires sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radio amateures) sont moins utilisées.
-Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend fortement du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
+Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend fortement à la fois du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
 En outre, l'attribution fixe du spectre empêche
@@
 Plus précisément, nous étudions les interconnexions possibles entre \textbf{Radio Intelligente} et \textbf{l'apprentissage statistique} appliquées aux réseaux IdO.
-Définissons et détaillons les deux concepts.
+Définissons et détaillons les deux concepts séparément.
 
@@
 %
-Comme la RI n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RI, qui vient de paraître il y a $20$ il y a quelques années.
+Comme la RI n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RI, qui vient de paraître il y a $20$ ans.
 %
-\begin{itemize}
+\begin{itemize}\tightlist
     \item
@@
     Puis S. Haykin en $2005$ a aussi dit que
-    \emph{``une radio intelligente (RI) est un système de communication sans fil intelligent qui est capable de connaître son environnement, d'apprendre et d'adapter ses paramètres de fonctionnement (puissance d'émission et fréquence porteuse) à la volée dans le but de fournir une communication fiable à tout moment, en tout lieu et efficace sur le plan spectral''} \cite{Haykin05}.
+    \emph{``une radio intelligente (RI) est un système de communication sans fil intelligent qui est capable de connaître son environnement, d'apprendre et d'adapter ses paramètres de fonctionnement (\eg, puissance d'émission et fréquence porteuse) à la volée, dans le but de fournir une communication fiable à tout moment, en tout lieu et efficace sur le plan spectral''} \cite{Haykin05}.
 
@@
 Dans les bandes sous licences, il y a des utilisateurs primaires (UP) qui paient pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
-Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est pas utilisé par un UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser aussi, éventuellement en payant un abonnement.
+Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est utilisé par aucun UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser aussi, éventuellement en payant un abonnement auprès de l'opérateur du réseau.
 %
 La réglementation stipule que les UP ont une priorité stricte,
-mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de toute UP.
+mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de tout UP.
 Ceci définit le concept d'\textbf{Accès Opportuniste au Spectre} (AOS, \emph{Opportunistic Spectrum Access}, OSA), pour lequel nous nous référons à l'article \cite{Zhao07} pour plus de détails.
@@
 qui est apparu comme un problème intéressant dans la communauté statistique et parmi les chercheurs intéressés par l'apprentissage séquentiel, comme le montrent les travaux antérieurs de \cite{Thompson33,Robbins52,LaiRobbins85}.
-Il est également inclus dans le concept plus général d'apprentissage par renforcement (\emph{Reinforcement Learning}, RL), lui-même un des domaines de l'apprentissage machine (\emph{Machine Learning}, ML).
+Les BMB sont également inclus dans le concept plus général d'apprentissage par renforcement (\emph{Reinforcement Learning}, RL), lui-même un des domaines de l'apprentissage machine (\emph{Machine Learning}, ML).
 % ML > RL > MAB
@@
 
-Nous illustrons ci-dessous l'idée d'un cycle d'apprentissage, alternant entre l'action et la réaction,
+Nous illustrons ci-dessous l'idée d'un cycle d'apprentissage, alternant entre action et réaction,
 en Figure~\ref{fig:1:ReinforcementLearningCycle}.
-Une joueuse (ou un apprenant) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
+Une joueuse (ou un apprenant) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de cette action, fournie par l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
 Le but de la joueuse est de maximiser ses récompenses, par des essais et des erreurs (\ie, actions et récompenses).
-De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage de renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, comme par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie, etc.
+De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage par renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie, etc.
 
 % Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
-Comme cette thèse se concentre sur les modèles d'apprentissage de renforcement,
-il est important de souligner que dans de tels modèles de prise de décision, l'apprenant n'a pas accès à l'ensemble de la réaction de l'environnement après avoir pris ses mesures.
+Comme cette thèse se concentre sur les modèles d'apprentissage par renforcement (RL),
+il est important de souligner que dans de tels modèles de prise de décision, l'apprenant n'a pas accès à l'ensemble des réactions possibles de l'environnement après avoir choisi son action.
 %
 En d'autres termes, la joueuse ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une autre action.
-Ce genre de rétroaction limitée s'appelle l'information de type bandit (\emph{bandit feedback}), et nous discutons de l'histoire et du concept des \emph{bandits multi-bras} (BMB, ou MAB) en détail dans le prochain Chapitre~\ref{chapter:2}.
+Ce genre de rétroaction limitée s'appelle l'information de type bandit (\emph{bandit feedback}), et nous discutons de l'histoire et du concept des \emph{bandits multi-bras} (BMB) en détail dans le prochain Chapitre~\ref{chapter:2}.
 Les essais cliniques et la découverte de traitements ont été historiquement la première application des BMB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33}.
 %
-Le BMB est un exemple simple mais puissant du dilemme bien connu de entre \emph{exploration et exploitation} :
-lorsqu'elle est confronté à un ensemble d'actions de $K$ dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
+Le BMB est un exemple simple mais puissant du dilemme bien connu entre \emph{exploration et exploitation} :
+lorsqu'elle est confronté à un ensemble de $K$ actions dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
 explorer les actions inconnues, afin de recueillir plus d'informations à leur sujet,
@@
 %
-Le problème du BMB a été étudié à la fois dans la communauté de l'apprentissage machine et de la statistique, depuis les $1950$ avec des pionniers comme H.
-Robbins \cite{Robbins52}, et plus récemment c'est un domaine de recherche actif, depuis les $1990$ \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-La recherche sur le BMB a produit une vaste littérature dans le domaine des $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet actif, comme l'illustrent les livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications des BMB dans dernières années \cite{bouneffouf2019survey}.
+Le problème du BMB a été étudié à la fois dans la communauté de l'apprentissage machine et de la statistique, depuis les années $1950$ avec des pionniers comme H.
+Robbins \cite{Robbins52}, et plus récemment c'est un domaine de recherche actif, depuis les années $1990$ \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
+La recherche sur le BMB a produit une vaste littérature des années $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet actif, comme l'illustrent les livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications des BMB les dernières années \cite{bouneffouf2019survey}.
 
@@
 %
-Les deux communautés de l'AOS et du BMB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des $2000$ et le début des $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
-L'accent est mis sur un US accédant à un spectre sous licence, occupé par un US, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès d'``écoute avant transmission''.
+Les deux communautés de l'AOS et du BMB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des années $2000$ et le début des années $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
+L'accent est mis sur un US accédant à un spectre sous licence, occupé par un UP, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès d'``écoute avant transmission''.
 %
@@
 l'US est équipé d'une capacité de détection,
-et considère un ensemble fixe et fini de canaux orthogonaux, c'est-à-dire différentes bandes de fréquences dans un spectre sous licence.
-Par exemple, il peut s'agir d'un ensemble de trois canaux Wi-Fi à fréquence différente, émis par la même station Wi-Fi.
+et considère un ensemble fixe et fini de $K$ canaux orthogonaux, \ie, différentes bandes de fréquences dans un spectre sous licence.
+Par exemple, il peut s'agir d'un ensemble de $K=3$ canaux Wi-Fi à fréquence différente, émis par la même station Wi-Fi.
 Une autre hypothèse est que UP et US sont synchronisés dans le temps, en subdivisant le temps en pas de temps discrets.
 %
-Ainsi, si l'US passe peu de temps au début de chaque plage horaire pour effectuer la détection, il peut rechercher la présence ou l'absence d'un UP avant de transmettre, pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
+Ainsi, si l'US passe peu de temps au début de chaque plage horaire pour essayer de détecter un UP, il peut rechercher la présence ou l'absence d'un UP, avant de transmettre pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
 Si l'US était capable d'analyser tous les $K$ canaux, il pourrait simplement émettre dans l'un des canaux libres s'il y en avait, ou ne pas émettre si tous les canaux sont utilisés à un pas de temps donné.
 %
-Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les articles \cite{yucek2009survey,subhedar2011spectrum}, de sorte que la plupart des travaux sur l'AOS limitent la capacité de détection de l'US à un seul canal à la fois.
-Cette hypothèse, ainsi que l'hypothèse selon laquelle les UP ne peuvent pas être perturbés (sans laquelle aucune régulation ne sera jamais acceptée pour l'AOS), impose à l'US de transmettre dans le canal qu'il a détecté, à chaque pas de temps, si et seulement si il a été détectée sans aucun UP.
+Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les articles \cite{yucek2009survey,subhedar2011spectrum} (\emph{spectrum sensing} en anglais), de sorte que la plupart des travaux sur l'AOS limitent la capacité de détection de l'US à un seul canal à la fois.
+Cette hypothèse, ainsi que l'hypothèse selon laquelle les UP ne peuvent pas être perturbés (sans laquelle aucune régulation ne sera jamais acceptée pour l'AOS), impose à l'US de transmettre dans le canal qu'il a analysé, à chaque pas de temps, si et seulement si il a été détectée sans aucun UP.
 % coût non négligeable de la reconfiguration radio
 
-En se concentrant sur un US dans un réseau AOS, il doit décider séquentiellement d'un canal à détecter (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il effectue la détection, et enfin il transmet dans ce canal si il a été détecté libre.
+En se concentrant sur un US dans un réseau AOS, il doit décider séquentiellement d'un canal à analyser (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il écoute ce canal et effectue une détection de PU, et enfin il transmet dans ce canal si il a été détecté libre.
 Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou l'équivalent, pour maximiser son nombre de transmissions réussies.
 %
-Si les différents canaux ne sont pas uniformément utilisés par les UP, et si l'on suppose une hypothèse stationnaire sur le trafic de les UP, alors l'objectif de l'US se résume à explorer les différents canaux et à exploiter les meilleurs.
-Le problème de l'accès opportuniste au spectre est donc un problème d'exploration/exploitation, avec un ensemble fini d'actions (les canaux, également appelés \emph{arms}),
-dans un cycle séquentiel d'action puis de rétroaction (les pas de temps sont $t\in\N^*$),
-sous la rétroaction partielle d'information (\ie, environ un canal de plus de $K$ à chaque pas de temps).
-Ces trois hypothèses sont celles qui limitent le cadre général du blanchiment d'argent au cas spécifique du bandit multibras (voir Figure~\ref{fig:1:ReinforcementLearningCycle}).
+Si les différents canaux ne sont pas uniformément utilisés par les UP, et si l'on suppose une hypothèse stationnaire sur le trafic des UP, alors l'objectif de l'US se résume à explorer les différents canaux et à exploiter les meilleurs.
+Le problème de l'accès opportuniste au spectre est donc un problème d'exploration/exploitation, avec un ensemble fini d'actions (les canaux, également appelés \emph{bras}),
+dans un cycle séquentiel d'action-réponse (les pas de temps sont $t\in\N^*$),
+sous l'information de rétroaction partielle (\ie, l'US reçoit l'information concernant un seul canal parmi $K$).
+Ces trois hypothèses sont celles qui limitent le cadre général de l'apprentissage machine au cas spécifique des bandit multi-bras (voir Figure~\ref{fig:1:ReinforcementLearningCycle}).
 
@@
 Les travaux antérieurs de notre équipe SCEE ont montré que le BMB peut être utilisé pour modéliser le problème de l'AOS :
-les bandes de fréquences orthogonales (ou canaux) sont des modèles par bras $k\in\{1,\dots,K\}$,
-et la récompense obtenue par le objet équipé de CR après avoir détecté le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
-En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors qu'une récompense de $r(t)=0$ indique que le canal $k$ est occupé au moment $t$ et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
+les bandes de fréquences orthogonales (ou canaux) sont modélisées par des bras $k\in\{1,\dots,K\}$,
+et la récompense obtenue par l'objet équipé de RI après avoir analysé le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
+En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors qu'une récompense de $r(t)=0$ indique que le canal $k$ est occupé durant l'interval de temps $t$, et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
 %
-Ce modèle a d'abord été étudié par W. Jouini lors de sa thèse de doctorat avec C. Moy \cite{Jouini12PhD}, il y a dix ans, d'abord avec \cite{Jouini09} et ensuite avec \cite{Jouini10,Jouini12}.
+Ce modèle a d'abord été étudié par W. Jouini lors de sa thèse de doctorat avec C. Moy \cite{Jouini12PhD}, il y a dix ans, d'abord dans l'article \cite{Jouini09} puis avec \cite{Jouini10,Jouini12}.
 
-Leurs travaux ont été parmi les premiers à proposer l'utilisation de Reinforcement Learning for Cognitive Radio et le problème AOS, notamment le modèle BMB et l'algorithme $\UCB_1$,
-ainsi que les premiers travaux de Q. Zhao et de son équipe, par exemple avec les travaux \cite{Liu08,Zhao10}.
+Leurs travaux ont été parmi les premiers à proposer l'utilisation de l'apprentissage par renforcement pour la radio intelligente et le problème AOS, notamment le modèle BMB et l'algorithme $\UCB_1$,
+ainsi que les premiers travaux de Q. Zhao et de son équipe, par exemple \cite{Liu08,Zhao10}.
 %
@@
 Moy et son étudiant C.
-Robert \cite{RobertSDR2014,MoyWSR2014,MoyWSR2014} ont développé des preuves de concept utilisant le matériel radio et la radio logicielle du monde réel en utilisant des cartes USRP et le logiciel MATLAB/Simulink.
+Robert \cite{RobertSDR2014,MoyWSR2014,MoyWSR2014} ont développé des preuves de concept utilisant du matériel radio réaliste et la radio logicielle, avec des cartes USRP et le logiciel MATLAB/Simulink.
 Dans une deuxième thèse de doctorat \cite{Modi17PhD}, N.
-Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes BMB pour optimiser la sélection des canaux, sur la durée de vie des piles d'un objet sans fil.
+Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes BMB pour optimiser la sélection des canaux, sur la durée de vie des batteries d'un objet sans fil.
 D'une part, l'utilisation d'un algorithme BMB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du objet.
-D'autre part, les algorithmes BMB classiques ont tendance à changer souvent de bras, surtout au début du processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour le objet sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par le objet.
-Chaque reconfiguration matérielle coûte de l'énergie à l'objet et les algorithmes de commutation rapide entraînent une réduction de l'autonomie de la batterie.
+D'autre part, les algorithmes BMB classiques ont tendance à changer souvent de bras, surtout au début de leur processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour l'objet sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par l'objet.
+Chaque reconfiguration matérielle coûte de l'énergie à l'objet, et les algorithmes à commutation rapide entraînent donc une réduction de l'autonomie de la batterie.
 Le compromis entre les deux aspects est étudié de façon empirique dans les études suivantes
@@
 
-En $2017$, C. Moy a continué à travailler dans cette direction, avec un étudiant postdoctoral, S. Darak, qui a mené à des publications telles que
+En $2017$, C. Moy a continué à travailler dans cette direction, avec un étudiant post-doctoral, S. Darak, qui a mené à des publications telles que
 \cite{darak2016bayesian,Darak16}.
@@
 Depuis $2017$, S. Darak et son équipe à l'IIIT Delhi en Inde, ont travaillé activement dans la recherche sur la radio intelligente à l'aide de bandits multi-bras.
-Héritage de son travail au sein de l'équipe SCEE,
-certains de leurs travaux récents sont également illustrés par des démos du monde réel utilisant USRP et le système MATLAB/Simulink
+Par héritage de son travail au sein de l'équipe SCEE,
+certains de leurs travaux récents sont également illustrés par des démos réalistes utilisant USRP et le système MATLAB/Simulink
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
@@
 La littérature susmentionnée a essentiellement montré que les algorithmes BMB peuvent être appliqués avec succès au problème de l'AOS.
-Toutefois, si l'on considère les $M$ objets l'emploi qui ne peuvent pas effectuer de détection active, tels que les $M$ objets faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des objets, le modèle BMB qui utilise la détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
+Toutefois, si l'on considère des objets de RI qui ne peuvent pas effectuer de détection active, tels que des objets à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des Objets, le modèle BMB qui utilise la détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
 En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et US.
@@
 %
-Les spécificités des réseaux IdO peuvent être énumérées comme suit,
-et nous nous référons à l'enquête \cite{Centenaro16} pour plus de détails.
+Les autres spécificités des réseaux IdO peuvent être énumérées comme suit,
+et nous nous référons à l'article \cite{Centenaro16} pour plus de détails.
 %
 \begin{itemize}\tightlist
+    % \item
+    % nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre UP et US),
     \item
-    nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre UP et US),
-    \item
-    la plupart des $M$ objets des $M$ objets faible coût, fonctionnant éventuellement sur du matériel de mauvaise qualité et ayant des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
+    la plupart des objets sont à faible coût, fonctionnent sur du matériel de mauvaise qualité et ont des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
     Cela signifie que même s'ils ne sont pas équipés de capacités de détection particulières,
-    ils sont équipés d'un émetteur ainsi que d'un récepteur (Tx et Rx, \ie, un émetteur-récepteur).
+    ils sont équipés d'un émetteur ainsi que d'un récepteur (Tx et Rx, \ie, un émetteur-récepteur),
     et ils ont des capacités de stockage et de calcul peu complexes (petits processeurs embarqués),
@@
     \item
-    les $M$ objets ont des cycles de fonctionnement faibles à très faibles (seulement quelques messages par minute ou par mois).
+    les objets ont des cycles de fonctionnement faibles à très faibles (seulement quelques messages par minute ou par mois).
 \end{itemize}
 
-Une coordination centrale des objets avec leur station de base est impossible car il faudrait communiquer en permanence avec la station de base, ce qui viole les deux dernières contraintes.
+Une coordination centrale des objets avec leur station de base est donc impossible car il faudrait qu'ils communiquent en permanence avec la station de base, ce qui viole les deux dernières contraintes.
 %
@@
 
-On pourrait penser qu'en l'absence de détection, l'apprentissage du renforcement n'est plus possible, mais tout objet IdO réel reçoit encore des informations sur son environnement après certaines ou toutes les transmissions.
-Dans la plupart des normes IdO, un message de liaison montante envoyé à une station de base peut être suivi d'un message de liaison descendante renvoyé par la station de base pour indiquer si le message de liaison montante a été bien reçu et compris.
+On pourrait penser qu'en l'absence de détection, l'apprentissage par renforcement n'est plus possible, mais tout objet IdO réel reçoit encore des informations sur son environnement, après certaines ou toutes ses transmissions.
+Dans la plupart des normes IdO, un message (montant) envoyé à une station de base peut être suivi d'un message (descendant) renvoyé par la station de base pour indiquer si le message a été bien reçu et compris.
 %
-En utilisant cette rétroaction, ça consiste en un \emph{acknowledgement} (\Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IdO puisse également utiliser un algorithme RL bien conçu afin d'optimiser ses communications.
+En utilisant cette rétroaction, qui consiste en un \emph{retour} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IdO puisse également utiliser un algorithme d'apprentissage par renforcement bien conçu, afin d'optimiser ses communications.
 
@@
 %
-Si nous résumons les problèmes étudiés et les résumons en une seule question, ce pourrait être la suivante :
+Si nous résumons les problèmes étudiés en une seule question, ce pourrait être la suivante :
 \emph{peut-on adapter les outils de prise de décision déjà appliqués avec succès à la radio intelligente pour l'accès opportuniste au spectre aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des objets ?}
@@
 Nous avons commencé par explorer la riche littérature des bandits multi-bras,
-car il existe de nombreux algorithmes différents, avec de nombreuses variantes du problème simple présenté ci-dessus.
-Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle BMB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires du BMB.
+car il existe de nombreux algorithmes différents, avec de nombreuses variantes du simple problème présenté ci-dessus.
+Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle BMB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires de BMB.
 %
@@
 %
-Notre exploration du grand nombre d'algorithmes et de modèles BMB développés dans la littérature récente.
+Notre exploration du grand nombre d'algorithmes et de modèles BMB développés dans la littérature récente
 nous a donné l'ambition d'écrire un seul logiciel permettant à chacun d'implémenter facilement de nouveaux modèles et algorithmes, afin de comparer les modèles existants et d'explorer empiriquement les performances des nouveaux algorithmes proposés.
-Pour atteindre cet objectif, nous avons produit une bibliothèque Python de simulation des problèmes BMB.
+Pour atteindre cet objectif, nous avons développé une bibliothèque Python de simulation des problèmes BMB.
 %
-Nous avons écrit la bibliothèque de simulation open-source la plus exhaustive pour les problèmes BMB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
-Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
+D'après ce que les auteurs ont identifiés,
+nous avons écrit la bibliothèque libre de simulation la plus exhaustive pour les problèmes BMB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
+Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, et différents exemples de son utilisation.
 Une documentation complète est disponible en ligne, ainsi que des instructions pour reproduire les simulations discutées dans la suite de cette thèse.
@@
 %
-Étant donné le grand nombre d'algorithmes BMB différents, nous nous intéressons aussi à la question de savoir comment un praticien peut choisir l'algorithme à mettre en œuvre, dans un objet IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
+Étant donné le grand nombre d'algorithmes BMB disponibles, nous nous intéressons aussi à la question de savoir comment une ingénieure peut choisir l'algorithme à mettre en œuvre, dans un objet IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
 Pour répondre à cette question, nous présentons deux approches.
-La première est une comparaison empirique des algorithmes existants les plus efficaces et les mieux connus, dans Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
-Nous confirmons que les algorithmes largement utilisés, tels que \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
-La seconde approche est une sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons dans Chapitre~\ref{chapter:25}.
-Ce travail sur l'agrégation des algorithmes BMB a été motivé pour le cas AOS de la radio cognitive, où de nombreux travaux de recherche antérieurs ne prenaient en compte que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
-Cette contribution a été présentée à la conférence IEEE WCNC à Barcelone, Espagne, en avril 2018.
-
-% FIXME VRAIMENT BEAUCOUP raccourcir la suite !
+La première est une comparaison empirique d'algorithmes existants, se focalisant sur les plus efficaces et les mieux connus, dans Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
+Nous confirmons que les algorithmes largement utilisés, tels que \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} et \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
+La seconde approche est une sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons en Chapitre~\ref{chapter:25}.
+Ce travail sur l'agrégation des algorithmes BMB a été motivé pour le cas AOS de la radio cognitive, où de nombreux travaux antérieurs ne considéraient que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
+Cette contribution a été présentée à la conférence IEEE WCNC à Barcelone, en Espagne, en avril 2018.
 
@@
 %
-Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet d'augmenter la durée de vie des piles de chacun des objets, ainsi que la QoS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
+Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet d'augmenter la durée de vie des batteries de chacun des objets, ainsi que la QoS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
 

### 2019-07-07 2-Chapters/1-Chapter/chapter1.tex
Je relis tout le ch1 traduit en français, ça aide à trouver des typos sur la version anglaise

@@
 In the last $150$ years, mankind has developed many different communication technologies, and since the late $1890$s, wireless communications between manufactured devices have been made possible, and more and more frequent in our lives.
-With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, for a large range of different applications.
+With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, allowing a large range of different applications.
 It is now a world-wide consensus that with the current trend of population increase and with the ongoing energy crisis, any newly deployed technology should be both cheap and energy efficient,
@@
 %
-In addition to the usual Research and Development effort to design many different efficient radio access schemes, for all the possible cases,
+In addition to the usual Research and Development effort to design many different efficient radio access schemes, covering all the possible cases,
 now the time has come to combine it with cheap and promising Machine Learning in order to (try to) attain the level of performance gain necessary for the IoT promises to become reality.
@@
 For buildings, the main goal is to reduce the cost of heating empty buildings and use sensor networks to get accurate and regular data about the temperature in all rooms and floors, and let the centralized heat control optimize its cost and energy consumption.
-For agriculture, one example can be to equip every cow (in large farms) with a sensor that regularly emits biological information, such as body temperature or stress level etc, in order to optimize the time of milking, to monitor the health of the animals etc.
+For agriculture, one example can be to equip every cow (in large farms) with sensors that regularly emit biological information, such as body temperature or stress level etc, in order to optimize the time of milking, to monitor the health of the animals etc.
 %
@@
 In the present and future IoT networks, many devices of heterogeneous natures are using the same antenna for different applications.
-A common problem is that such IoT devices have a strong constraint on their power consumption, as most of them will be deployed without a direct power access and will run on a tiny battery, which lifespan should be maximized.
+A common problem is the strong constraint that such such IoT devices have on their power consumption, as most of them will be deployed without a direct power access and will run on a tiny battery, which lifespan should be maximized.
 Most commercial IoT companies nowadays indeed sell a lifespan larger than $10$ years, like SIGFOX \cite{Centenaro16}.
@@
 and the \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}) layer.
-To quote some examples of standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications, while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications (up-to \SI{50}{\kilo\meter}).
-We refer to the survey \cite{Centenaro16} for more details, and references in \cite{Azari18} or our recent article \cite{MoyBesson2019}.
+To quote some examples of IdO standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications (up-to \SI{2}{\meter}), while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications (up-to \SI{50}{\kilo\meter}).
+We refer to the survey \cite{Centenaro16} for more details, and \cite{Azari18} or our recent article \cite{MoyBesson2019}, and references therein.
 
@@
 \href{https://www.itu.int/en/Pages/default.aspx}{International Telecommunication Union} (see \href{https://www.itu.int/}{\texttt{www.ITU.int}}),
-the \href{https://www.fcc.gov/}{Federal Communications Commission} in the USA (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
+the \href{https://www.fcc.gov/}{Federal Communications Commission} in the United States of North America (see \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
 or the \href{https://cept.org}{European Conference of Postal and Telecommunications Administrations} in Europe (see \href{https://www.CEPT.org/}{\texttt{CEPT.org}}),
@@
     \includegraphics[width=0.95\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
-    \caption[A chart representing the allocation of radio spectrum in the USA in $2016$]{A chart representing the allocation of radio spectrum in the USA in $2016$.
-\copyright{} United States Department of Commerce, \href{https://www.ntia.doc.gov/files/ntia/publications/january_2016_spectrum_wall_chart.pdf}{\texttt{www.ntia.doc.gov/files/ntia/publications/january\_2016\_spectrum\_wall\_chart.pdf}}.}
+    \caption[A chart representing the allocation of radio spectrum in the United States of North America in $2016$]{A chart representing the allocation of radio spectrum in the United States of North America in $2016$.
+\copyright{} United States Department of Commerce, published online at \href{https://www.ntia.doc.gov/files/ntia/publications/january_2016_spectrum_wall_chart.pdf}{\texttt{www.ntia.doc.gov/files/ntia/publications/january\_2016\_spectrum\_wall\_chart.pdf}}.}
     % Only a small share of the spectrum is not yet allocated, the top-left corner in white and indicated by \textcolor{red}{\textbf{Free} in red} (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz})
@@
 Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are less utilized.
-Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on time and place, as shown by \cite{Lopez2009spectral}.
+Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on both time and place, as shown by \cite{Lopez2009spectral} for instance.
 Moreover, the fixed spectrum allocation prevents
@@
 %  with the exception of a unique pair of receiver/transmitter antennas.
-Even though the SDR paradigm was initiated by the USA defense research in the $1980$s, over the past $20$ years the civil industry has started to be interested by SDR, and CR has got significant attention from the academia and industry as well.
+Even though the SDR paradigm was initiated by the United States of North America defense research in the $1980$s, over the past $20$ years the civil industry has started to be interested by SDR, and CR has got significant attention from the academia and industry as well.
 % CR has been the first answer to solve spectrum scarcity issue which collected a huge attention in early 21st century.
@@
 % je peux faire comme Navik, donner des définitions de la CR, citant d'autres travaux
-\begin{itemize}
+\begin{itemize}\tightlist
     \item
@@
 In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the network.
-As exposed in the definition above of the CR philosophy, we could imagine that if the network is not used by any PU in a certain location and time, then non licensed users, referred to as Secondary User (SU), could use it, possibly by paying some rent to the network operator.
+As exposed in the definition above of the CR philosophy, we could imagine that if the network is unused by any PU in a certain location and time, then non licensed users, referred to as Secondary User (SU), could use it, possibly by paying some rent to the network operator.
 %
@@
 in Figure~\ref{fig:1:ReinforcementLearningCycle}.
-A player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$), and then by observing a reward $r(t)$, which is a certain measure of success from the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
+A player (or learner) interacts with its environment by taking an action $A(t)$ (\eg, a choice in a finite set, $A(t)\in\{1,\dots,K\}$, or a vector $A(t)\in\R^d$), and then by observing a reward $r(t)$, which is a certain measure of success of this action, produced by the environment (\eg, $r(t)\in\{0,1\}$ for binary failure/success, or $r(t)\in\R$).
 The goal of the player is to maximize its rewards, by trials and errors (\ie, actions and rewards).
@@
 the SU is equipped with a sensing capability,
-and considers a fixed and finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-For instance, it can be a set of three Wi-Fi channels at different frequency, emitted by the same Wi-Fi station.
+and considers a fixed and finite set of $K$ orthogonal channels, \ie, different frequency bands in a licensed spectrum.
+For instance, it can be a set of $K=3$ Wi-Fi channels at different frequency, emitted by the same Wi-Fi station.
 Another hypothesis is that both PU and SU are synchronized in time, by sub-dividing the time in discrete time steps (or rounds).
@@
 If the different channels are not uniformly used by the PU, and if we assume a stationary hypothesis on the PU traffic, then the goal of the SU boils down to exploring the different channels and exploiting the best ones.
-This frames the Opportunistic Spectrum Access problem as an exploration/exploitation problem, with a finite set of actions (the channels, also called \emph{arms}),
+This frames the Opportunistic Spectrum Access problem as an exploration/exploitation problem,
+with a finite set of actions (the channels, also called \emph{arms}),
 in a sequential action-then-feedback cycle (time steps are $t\in\N^*$),
-under partial information feedback (\ie, about one channel over $K$ at each time step).
+under partial information feedback (\ie, the player receives feedback about only one channel over $K$).
 These three hypotheses are the ones that restrict from the general ML framework to the specific Multi-Armed Bandit case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
 
-\paragraph{MAB for OSA, and a short presentation of the history of this research at the SCEE team.}
+\paragraph{MAB for OSA, and a short history of these researchs at the SCEE team.}
 %
 Previous works of our SCEE team showed that MAB can be used to model the OSA problem:
-orthogonal frequency bands (or channels) are models by arms $k\in\{1,\dots,K\}$,
+orthogonal frequency bands (or channels) are modeled by arms $k\in\{1,\dots,K\}$,
 and the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$ is modeled by a reward of $r(t) \in \{0,1\}$.
-Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus a SU message can be sent), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no SU message should be sent, until the end of that slot.
-%
-This model was first studied by W. Jouini during his PhD thesis with C. Moy \cite{Jouini12PhD}, ten years ago, first with \cite{Jouini09} and later with \cite{Jouini10,Jouini12}.
+Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus a SU message can be sent), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at during the time slot $t$, and no SU message should be sent, until the end of that slot.
+%
+This model was first studied by W. Jouini during his PhD thesis with C. Moy \cite{Jouini12PhD}, ten years ago, first in the article \cite{Jouini09} and later in \cite{Jouini10,Jouini12}.
 
 Their works were among the first ones to propose to use Reinforcement Learning for Cognitive Radio and the OSA problem, especially the MAB model and the $\UCB_1$ algorithm,
-along with the early works of Q. Zhao and her team, for instance with the works \cite{Liu08,Zhao10}.
+along with the early works of Q. Zhao and her team, for instance with \cite{Liu08,Zhao10}.
 %
@@
 On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and could bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
-On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfigurations for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
-Each hardware reconfiguration costs energy for the device, and quickly-switching algorithms lead to a reduction of the battery life.
+On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of their learning process, and this induces a lot of dynamical hardware reconfigurations for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
+Each hardware reconfiguration costs energy for the device, and so quickly-switching algorithms lead to a reduction of the battery life.
 The tradeoff between the two aspects is studied empirically in
@@
 Inheriting from his work with the SCEE team,
-some of their recent works are also illustrated with real-world demo using USRP and the MATLAB/Simulink system
+some of their recent works are also illustrated with realistic demos using USRP and the MATLAB/Simulink system
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
@@
 %
-The specificities of IoT networks can be listed as follows,
+The other specificities of IoT networks can be listed as follows,
 and we refer to the survey \cite{Centenaro16} for more details.
 \begin{itemize}\tightlist
-    \item
-    we consider here IoT networks that run in unlicensed bands (no more distinction between PU and SU),
-    \item
-    most IoT devices are very low-cost devices, possibly running on low-quality hardware, and having limited software and power capabilities (\ie, tiny batteries).
-    It means that they even if they are not equipped with particular sensing capabilities,
-    they are equipped with a transmitter as well as with a receiver (Tx and Rx, \ie, a transceiver).
+    % \item
+    % we consider here IoT networks that run in unlicensed bands (no more distinction between PU and SU),
+    \item
+    most IoT devices are very low-cost devices, run on low-quality hardware, and have limited software and power capabilities (\ie, tiny batteries).
+    It means that even if they are not equipped with particular sensing capabilities,
+    they are equipped with a transmitter as well as with a receiver (Tx and Rx, \ie, a transceiver),
     and they have low-complexity storage and computation capabilities (\ie, small embedded processors),
@@
     \item
-    IoT devices have low duty cycles (only a few messages every minute) to very low duty cycles (only a few messages every month).
+    IoT devices have low to very low duty cycles (few messages every minute to only a few every month).
 \end{itemize}
 
-A central coordination of the devices with their base station is impossible as it would require to communicate continuously with the base station, which violates the two last constraints.
+A central coordination of the devices with their base station is thus impossible as it would require them to communicate continuously with the base station, which violates the two last constraints.
 %
@@
 
-One could think that in the absence of sensing, Reinforcement Learning is no longer possible, but any real IoT device still receives some information about its environment after some or every transmissions.
+One could think that in the absence of sensing, Reinforcement Learning is no longer possible, but any real IoT device still receives some information about its environment after some or every of its transmissions.
 In most IoT standards an up-link message sent to a base station can be followed by a down-link message sent back by the base station, to indicate if the up-link message was successfully received and understood.
 %
-By using this feedback, that consists in an \emph{acknowledgement} (\Ack) message received shortly after every successful transmission, or in an absence of \Ack{} after a failed transmission, it is possible that an IoT end-device can also use a well-designed RL algorithm in order to optimize its communications.
+By using this feedback, which consists in an \emph{acknowledgement} (\Ack) message received shortly after every successful transmission, or in an absence of \Ack{} after a failed transmission, it is possible that an IoT end-device can also use a well-designed RL algorithm, in order to optimize its communications.
 %
@@
 we were not only interested by the usual measure of performance of any MAB algorithm (its regret, see below in Section~\ref{sec:2:lowerUpperBoundsRegret}),
-but also by their empirical performances in terms of computational complexity and storage requirements, from the point of view of both theoretical analyses and real-world measurements on time and memory footprints.
+but also by their empirical performances in terms of computational complexity and storage requirements, from the point of view of both theoretical analyses and real-world measurements.
+%  on time and memory footprints.
 
@@
 has given us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explore the performances of newly proposed algorithms.
-To tackle this goal, we produced a Python library of simulation of MAB problems.
+To tackle this goal, we developed a Python library of simulation of MAB problems.
 %
@@
 %
-Because there are so many different MAB algorithms, we are also interested by the question of how a practitioner can chose the algorithm to implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
+Because there are so many MAB algorithms available, we are also interested by the question of how a practitioner can chose the algorithm to implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
 To answer this question, we present two approaches.

### 2019-07-07 1-Preamble/4-Acknowledgement/acknowledgement.tex
Citations de Kaamelott pour le début de la thèse… ??

@@
 %
-I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my studies.
+I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my life and in my studies.
 His rigour, dedication and seriousness kept pushing me further, and I am immensely proud of him.
@@
 %
-I also thank my younger brother F., for being an inspiration for other (highly important) directions of my life. \emph{Ton courage et ta force m'inspirent quotidiennement, merci beaucoup !}
+I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life. \emph{Ton courage et ta force m'inspirent quotidiennement, merci beaucoup !}
 Very warm thanks then go to M. and my grand-parents, J. \& J.-M. and J., and to my uncle J. and aunt B.!

### 2019-07-07 1-Preamble/5-Abstract/resume_fr.tex
Citations de Kaamelott pour le début de la thèse… ??

@@
 % % Chapitre 4
-% Nous proposons différents modèles de réseaux IdO, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement (Chapitre~\ref{chapter:4}).
+% Nous proposons différents modèles de réseaux de l'IdO, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement (Chapitre~\ref{chapter:4}).
 % % Chapitre 2
@@
 
-FIXME 0.5 à 2 pages de contexte, historique, technique, humain, etc.
+% FIXME 0.5 à 2 pages de contexte, historique, technique, humain, etc.
 
@@
 %
-La racine des problèmes qui motivent cette thèse sont les questions du réchauffement climatique et de l'augmentation de la population mondiale.
+Les problèmes qui motivent cette thèse sont les questions du réchauffement climatique et de l'augmentation de la population mondiale.
 % https://en.wikipedia.org/wiki/History_of_radio
@@
 C'est pourquoi nous nous intéressons dans cette thèse sur
-l'amélioration de la durée de vie des batteries des objets IdO et la réduction du coût énergétique des réseaux IdO.
-Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant une prise de décision décentralisée à faible coût, directement dans les futurs objets IdO.
+l'amélioration de la durée de vie des batteries des objets de l'IdO et la réduction du coût énergétique des réseaux de l'IdO.
+Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant une prise de décision décentralisée à faible coût, directement dans les futurs objets de l'IdO.
 %
-Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (algorithmes de types bandits multi-bras), afin de permettre aux objets IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
+Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (algorithmes de types bandits multi-bras), afin de permettre aux objets de l'IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
 
@@
 Cette famille de systèmes sans fil est appelée Internet des Objets (IdO, ou \emph{Internet of Things}, IoT),
-et un exemple typique d'application de tels réseaux IdO est celui des réseaux sans fil de capteurs.
+et un exemple typique d'application de tels réseaux de l'IdO est celui des réseaux sans fil de capteurs.
 
@@
 En effet, un contrôle centrale nécessite des paquets de signalisation qui ont été identifiés comme trop lourds pour ce type de systèmes.
-Dans les réseaux IdO actuels et futurs, de nombreux objets hétérogènes utilisent la même station de base (aussi appelée point d'accès, ou \emph{base station}, \emph{access point} ou \emph{gateway}) pour des applications différentes.
-Un problème commun est la forte contrainte en terme de consommation d'énergie de ces objets IdO, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
+Dans les réseaux de l'IdO actuels et futurs, de nombreux objets hétérogènes utilisent la même station de base (aussi appelée point d'accès, ou \emph{base station}, \emph{access point} ou \emph{gateway}) pour des applications différentes.
+Un problème commun est la forte contrainte en terme de consommation d'énergie de ces objets de l'IdO, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
 Ainsi, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des objets ayant une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
-Une autre contrainte commune aux objets IdO est leur faible besoin en communication, car la plupart des applications n'auront besoin d'envoyer qu'un ou quelques messages chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
+Une autre contrainte commune aux objets de l'IdO est leur faible besoin en communication, car la plupart des applications n'auront besoin d'envoyer qu'un ou quelques messages chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
 %
-De nombreuses normes différentes pour les réseaux IdO ont été proposées ces dernières années,
+De nombreuses normes différentes pour les réseaux de l'IdO ont été proposées ces dernières années,
 et ils consistent en une spécification à la fois pour la couche \emph{PHY}sique
@@
 %
-Plus précisément, nous étudions les interconnexions possibles entre \textbf{Radio Intelligente} et \textbf{l'apprentissage statistique} appliquées aux réseaux IdO.
+Plus précisément, nous étudions les interconnexions possibles entre \textbf{Radio Intelligente} et \textbf{l'apprentissage statistique} appliquées aux réseaux de l'IdO.
 Définissons et détaillons les deux concepts séparément.
@@
 
-\paragraph{Limitations et spécificités des réseaux IdO.}
+\paragraph{Limitations et spécificités des réseaux de l'IdO.}
 %
@@
 Toutefois, si l'on considère des objets de RI qui ne peuvent pas effectuer de détection active, tels que des objets à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des Objets, le modèle BMB qui utilise la détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
-En outre, dans la plupart des cas, les réseaux IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et US.
-En d'autres termes, il n'y a plus de priorité entre les utilisateurs des réseaux IdO.
+En outre, dans la plupart des cas, les réseaux de l'IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et US.
+En d'autres termes, il n'y a plus de priorité entre les utilisateurs des réseaux de l'IdO.
 (plus de UP, tous sont US).
 %
-Les autres spécificités des réseaux IdO peuvent être énumérées comme suit,
+Les autres spécificités des réseaux de l'IdO peuvent être énumérées comme suit,
 et nous nous référons à l'article \cite{Centenaro16} pour plus de détails.
@@
     % \item
-    % nous considérons ici les réseaux IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre UP et US),
+    % nous considérons ici les réseaux de l'IdO qui fonctionnent dans des bandes sans licence (plus de distinction entre UP et US),
     \item
@@
 
-On pourrait penser qu'en l'absence de détection, l'apprentissage par renforcement n'est plus possible, mais tout objet IdO réel reçoit encore des informations sur son environnement, après certaines ou toutes ses transmissions.
-Dans la plupart des normes IdO, un message (montant) envoyé à une station de base peut être suivi d'un message (descendant) renvoyé par la station de base pour indiquer si le message a été bien reçu et compris.
+On pourrait penser qu'en l'absence de détection, l'apprentissage par renforcement n'est plus possible, mais tout objet de l'IdO réel reçoit encore des informations sur son environnement, après certaines ou toutes ses transmissions.
+Dans la plupart des normes IdO, un message (sur la voie montante) envoyé à une station de base peut être suivi d'un message (sur la voie descendante) renvoyé par la station de base pour indiquer si le message a été bien reçu et compris.
 %
-En utilisant cette rétroaction, qui consiste en un \emph{retour} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IdO puisse également utiliser un algorithme d'apprentissage par renforcement bien conçu, afin d'optimiser ses communications.
+En utilisant cette rétroaction, qui consiste en un \emph{acquittement} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IdO puisse également utiliser un algorithme d'apprentissage par renforcement bien conçu, afin d'optimiser ses communications.
 
@@
 
-\paragraph{Chapitre~\ref{chapter:2}.}
+% \paragraph{Chapitre~\ref{chapter:2}.}
 %
@@
 %
-Afin de bien comprendre quels algorithmes pourraient être adaptés aux contraintes susmentionnées de la RI pour les réseaux IdO,
+Afin de bien comprendre quels algorithmes pourraient être adaptés aux contraintes susmentionnées de la RI pour les réseaux de l'IdO,
 nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme BMB (son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
@@
 
-\paragraph{Chapitre~\ref{chapter:3}.}
+% \paragraph{Chapitre~\ref{chapter:3}.}
 %
@@
 
-\paragraph{Chapitre~\ref{chapter:25}.}
+% \paragraph{Chapitre~\ref{chapter:25}.}
 %
-Étant donné le grand nombre d'algorithmes BMB disponibles, nous nous intéressons aussi à la question de savoir comment une ingénieure peut choisir l'algorithme à mettre en œuvre, dans un objet IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
+Étant donné le grand nombre d'algorithmes BMB disponibles, nous nous intéressons aussi à la question de savoir comment une ingénieure peut choisir l'algorithme à mettre en œuvre, dans un objet de l'IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
 Pour répondre à cette question, nous présentons deux approches.
@@
 
-\subsection*{Nos modèles de réseaux d'IdO et de BMB décentralisé}
+\subsection*{Nos modèles de réseaux de l'IdO et de BMB décentralisé}
 
-Dans la deuxième Partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux d'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du BMB, issues de la simplification de notre premier modèle.
+Dans la deuxième Partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux de l'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du BMB, issues de la simplification de notre premier modèle.
 
-\paragraph{Chapitre~\ref{chapter:4}.}
+% \paragraph{Chapitre~\ref{chapter:4}.}
 %
-Nous étudions deux modèles de réseaux IdO dans le Chapitre~\ref{chapter:4}.
-Le premier modèle considère les $M$ objets ont simplement des données à envoyer régulièrement à une station de base (à des moments imprévisibles ou aléatoires).
-De tels $M$ objets acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
-L'objectif de ce premier modèle est d'améliorer la qualité de service (QoS) de l'application de ce réseau IdO, en appliquant une RL décentralisée côté objet afin de réduire le taux d'échec de transmission.
-La station de base recevra donc plus de paquets de liaison montante des objets desservis, si elle peut apprendre à utiliser un système efficace d'accès au spectre.
-Cela implique que le réseau peut accepter plus d'objets tout en maintenant la même QoS.
+Nous étudions deux modèles de réseaux de l'IdO dans le Chapitre~\ref{chapter:4}.
+Le premier modèle considère que les objets ont simplement des données à envoyer régulièrement à une station de base (à des moments imprévisibles ou aléatoires).
+De tels objets utilisent les acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
+L'objectif de ce premier modèle est d'améliorer la qualité de service (QdS, ou \emph{Quality of Service} QoS) de l'application de ce réseau de l'IdO, en appliquant une RL décentralisée côté objet afin de réduire le taux d'échec de transmission.
+La station de base recevra donc plus de paquets de liaison montante des objets desservis, si ils peuvent apprendre par eux-même un accès efficace au spectre.
+Cela implique que le réseau peut accepter plus d'objets tout en maintenant la même QdS.
 %
-Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet d'augmenter la durée de vie des batteries de chacun des objets, ainsi que la QoS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
+Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet dans ce cas d'augmenter la durée de vie des batteries de chacun des objets, ainsi que la QdS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
 
-Le premier modèle constitue le premier article écrit au cours de cette thèse, et a initié une collaboration avec R. Bonnefoi, un autre doctorant de notre équipe SCEE.
-Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre 2010, où il a obtenu le "prix du meilleur papier" \cite{Bonnefoi17}.
+Le premier modèle est issu du premier article écrit au cours de cette thèse, qui a initié une collaboration avec R. Bonnefoi, un autre doctorant de notre équipe SCEE.
+Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre 2010, où il a obtenu le ``prix du meilleur papier'' \cite{Bonnefoi17}.
 %
@@
 
-\paragraph{Chapitre~\ref{chapter:5}.}
+% \paragraph{Chapitre~\ref{chapter:5}.}
 %
-Dans les deux modèles présentés ci-dessus, le concept est de laisser chaque objet dynamique d'un réseau IdO exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
-Il est important de noter que chaque objet ne communique, et donc n'apprend, qu'à quelques tours et non à toutes les étapes, en suivant son propre processus d'activation (aléatoire) (nous nous limitons à un processus d'activation de Bernoulli purement aléatoire de probabilité $p$).
-Cela signifie que chacun d'eux vise son propre objectif local, qui est de maximiser sa récompense cumulée (d'une manière égoïste).
-Il est bien connu dans la littérature sur la théorie des jeux que le jeu égoïste peut être désastreux pour la mesure centralisée de la performance, on peut penser aux "dilemmes" populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
-Il était donc assez surprenant que nos simulations numériques, ainsi que le PoC réaliste, aient montré que l'apprentissage BMB décentralisé et égoïste conduisait à une coordination efficace entre les $M$ objets les scénarios considérés, malgré le fait que ces objets IdO ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
+Dans les deux modèles de réseaux de l'IdO présentés ci-dessus, le concept est de laisser chaque objet dynamique d'un tel réseau exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
+Il est important de noter que chaque objet ne communique, et donc n'apprend, qu'à quelques tours et non à toutes les étapes, en suivant son propre processus d'activation aléatoire (nous nous limitons à un processus d'activation de Bernoulli purement aléatoire de probabilité $p$).
+Cela signifie que chacun d'eux vise son propre objectif local, qui est de maximiser sa récompense cumulée, d'une manière égoïste et complètement agnostique des autres objets ayant le même objectif.
+Il est bien connu dans la littérature sur la théorie des jeux que jouer égoïstement peut être désastreux pour la mesure centralisée de la performance, on peut penser aux ``dilemmes'' populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
+Il était donc assez surprenant que nos simulations numériques, ainsi que la preuve de concept réaliste, aient montré que l'apprentissage BMB décentralisé et égoïste conduisait à une coordination efficace entre les objets, dans tous les scénarios considérés, malgré le fait que ces objets ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
 
-Nous avons d'abord essayé d'analyser la performance de cet algorithme décentralisé, que nous appelons \Selfish, dans le modèle de Section~\ref{sec:4:firstModel},
-mais en raison du nombre aléatoire de $M$ objets chaque pas de temps (\ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
+C'est pourquoi nous avons d'abord essayé d'analyser la performance de cet algorithme décentralisé, que nous appelons \Selfish, dans le modèle de la Section~\ref{sec:4:firstModel},
+mais en raison du nombre aléatoire d'objets actifs à chaque pas de temps (\ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
 %
-C'est pourquoi dans le Chapitre~\ref{chapter:5}, nous relâchons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$, qui est équivalent à avoir $p < 1$, et nous considérons le cas des $M$ objets à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ $M$ objets.
-Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas AOS (\ie, avec détection) ou le cas IdO (\ie, sans détection).
+C'est la raison pour laquelle nous affaiblissons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$ dans le Chapitre~\ref{chapter:5}, qui est équivalent à avoir $p < 1$, et nous considérons le cas des objets actifs à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ objets.
+Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas AOS (\ie, avec détection des UP) ou le cas IdO (\ie, sans détection).
 L'objectif était de comprendre l'heuristique du Chapitre~\ref{chapter:4}, \Selfish, dans le cadre plus simple du BMB multi-joueur, qui a été étudié auparavant pour le cas de l'AOS, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
 %
-Le cas AOS couvert par notre modèle est en fait légèrement différent de celui considéré par \cite{Jouini10} et d'autres travaux antérieurs,
+Le cas AOS couvert par notre modèle peut sembler légèrement différent de celui considéré par \cite{Jouini10} et d'autres travaux antérieurs,
 car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
-Il diffère des modèles précédemment étudiés d'application du BMB pour l'AOS, puisqu'ils ne prennent en compte que les $M$ objets une unité de base et une unité centrale de traitement, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'unité de base sera certainement reçu avec succès par la base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
+Il est en fait équivalent aux modèles précédemment étudiés d'application du BMB pour l'AOS, puisqu'ils ne prennent en compte que des objets synchronisés, un US et des UP, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'US sera certainement reçu avec succès par la station de base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
 
-D'une part, dans le cadre de l'AOS, nous n'avons pas réussi à obtenir un résultat positif pour la politique égoïste, car nous avons prouvé que dans certains contextes limités (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre de regret moyen linéaire.
-Ce résultat a ensuite été confirmé et analysé par \cite{BoursierPerchet18}.
+D'une part, dans le cadre de l'AOS, nous n'avons pas réussi à obtenir un résultat positif pour l'algorithme \Selfish, car nous avons prouvé que dans certains contextes limités (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre d'un regret moyen linéaire.
+Ce résultat a ensuite été confirmé et analysé par \cite{BoursierPerchet18} (Annexe~F).
 %
-D'autre part, nous avons été en mesure de proposer de nouveaux algorithmes pour ce problème de bandit multi-joueurs avec des informations de détection, et nous avons analysé notre proposition \MCTopM-\klUCB, pour montrer qu'elle atteint une limite supérieure de regret logarithmique à temps fini, améliorant considérablement les résultats précédents.
-Notre algorithme atteint également un nombre logarithmique de collisions et de commutateurs de bras, et permet à un groupe fixe de $M$ objets d'apprendre efficacement à utiliser les $M$ meilleurs canaux orthogonalement pour presque toutes leurs communications montantes.
+D'autre part, nous avons été en mesure de proposer de nouveaux algorithmes pour ce problème de bandit multi-joueurs (avec informations de détection), et nous avons analysé notre proposition \MCTopM-\klUCB, pour montrer qu'elle atteint une limite supérieure de regret logarithmique à temps fini, améliorant considérablement les résultats précédents.
+Notre algorithme atteint également un nombre logarithmique de collisions et de changements de bras, et permet à un groupe fixe de $M$ objets d'apprendre efficacement à utiliser les $M$ meilleurs canaux orthogonalement pour presque toutes leurs communications montantes.
 %
-Ce résultat théorique solide dépend fortement de la présence d'une rétroaction de détection et, par conséquent, nos résultats ne sont pas (encore) applicables au modèle IdO sans détection.
+Ce fort résultat théorique dépend fortement de la présence d'une information de détection et, par conséquent, nos résultats ne sont pas (encore) applicables au modèle IdO sans détection.
 
@@
 La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes BMB, conçus pour résoudre des problèmes stationnaires, sur un problème non stationnaire,
-mais plutôt que nous analysons des algorithmes qui jouent tous dans différents sous-ensembles (aléatoires) des pas de temps synchronisés globaux.
-Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et imposer un maximum de $M \leq K$ objets n'est pas vraiment réaliste pour les réseaux IdO, même si cela conduit au modèle du Chapitre~\ref{chapter:5}.
+mais plutôt que nous analysons des algorithmes qui jouent tous dans différents d'instants d'activations (aléatoires), sous-ensembles des pas de temps synchronisés globaux.
+Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et imposer un maximum de $M \leq K$ objets n'est en fait pas vraiment réaliste pour les réseaux de l'IdO, même si cela conduit au modèle du Chapitre~\ref{chapter:5}.
 
-D'une part, si le modèle d'activation de tous les périphériques peut être corrigé, par exemple en se basant sur une affectation centralisée des périphériques à différents créneaux horaires, alors le modèle de bandit multi-joueurs du Chapitre~\ref{chapter:5} peut être utilisé pour permettre à chaque groupe de périphériques $M \leq K$ d'apprendre une affectation orthogonale optimale dans les canaux $K$ (\eg, groupes peuvent être le jeu des périphériques que tous émettent au même instant).
+D'une part, si le modèle d'activation de tous les périphériques peut être corrigé, par exemple en se basant sur une affectation centralisée des périphériques à différents créneaux horaires, alors le modèle de bandit multi-joueurs du Chapitre~\ref{chapter:5} peut être utilisé pour permettre à chaque groupe de $M \leq K$ périphériques d'apprendre une affectation orthogonale optimale dans les $K$ canaux (\eg, les groupes peuvent être le jeu des périphériques que tous émettent au même instant).
 %
@@
 
-D'autre part, une autre façon de relâcher l'hypothèse non stationnaire est de prendre le point de vue d'un objet, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur une unité de soins entourée de plusieurs UP ayant des comportements fixes.
-Si nous nous concentrons sur un objet IdO, son environnement (\ie, les $M$ objets) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réalistes donnent généralement une certaine structure sur cette non-stationnarité.
-Par exemple, l'application de la stationnarité à certains intervalles de temps conduit à la dernière contribution et au dernier chapitre de cette thèse.
+D'autre part, une autre façon d'affaiblir l'hypothèse non stationnaire est de prendre le point de vue d'un seul objet, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur un US entouré de plusieurs UP ayant des comportements fixes.
+Si nous nous concentrons sur un objet de l'IdO, son environnement (\ie, les autres objets) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réels présentent généralement des non-stationnarité avec une certaine structure.
+Par exemple, l'hypothèses de stationnarité par intervalles de temps conduit à la dernière contribution et au dernier chapitre de cette thèse.
 
-\paragraph{Chapitre~\ref{chapter:6}.}
+% \paragraph{Chapitre~\ref{chapter:6}.}
 %
-Afin de comprendre également avec précision comment les algorithmes BMB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur les BMB adverse ou non stationnaire, pour les deux cas d'environnements variant lentement et changeant brusquement.
+Afin de bien comprendre également comment les algorithmes BMB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur les BMB adverses ou non stationnaires, pour les deux cas d'environnements variant lentement et changeant brusquement.
 Nous commençons notre dernier Chapitre~\ref{chapter:6} en passant en revue les travaux existants,
-et nous avons choisi de nous concentrer sur le problème des pièces stationnaires,
-ce qui signifie que le problème de bandit sous-jacent est stationnaire à certains intervalles, séparés par des points de changement situés à des moments inconnus.
-Supposer que l'environnement est stationnaire au niveau de la pièce est en effet logique pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'objets dans le réseau.
+et nous avons choisi de nous concentrer sur les problèmes de bandits stationnaires par morceaux,
+ce qui signifie que le problème de bandit sous-jacent est stationnaire sur certains intervalles, séparés par des points de changement situés à des moments inconnus.
+Supposer que l'environnement est stationnaire par morceaux est en effet logique pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'objets dans le réseau.
 Des exemples d'une telle situation peuvent être une nouvelle entreprise arrivant sur le marché dans une ville (par exemple, les comptoirs Linky qui sont installés aujourd'hui en France), ou l'agriculteur voisin installant des capteurs sur son propre troupeau de vaches, etc.
 %
-Deux grandes familles d'algorithmes ont été proposées pour le problème stationnaire par pièce,
-qui consistent généralement à combiner une politique efficace conçue pour le problème stationnaire du BMB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des bras.
-Les politiques passivement adaptatives utilisent une fenêtre de taille fixe ou évolutive \cite{Garivier11UCBDiscount}, ou un facteur d'actualisation, pour oublier les observations passées \cite{Kocsis06,Gupta11thompson},
-tandis que les politiques activement adaptatives utilisent un test statistique pour détecter les points de changement \cite{MellorShapiro13,Allesiardo15}.
+Deux grandes familles d'algorithmes ont été proposées pour le problème stationnaire par morceaux,
+qui consistent généralement à combiner un algorithme efficace conçue pour le problème stationnaire du BMB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des bras.
+Les algorithmes passivement adaptatives utilisent une fenêtre de taille fixe ou évolutive \cite{Garivier11UCBDiscount}, ou un facteur d'actualisation, pour oublier les observations passées \cite{Kocsis06,Gupta11thompson},
+tandis que les algorithmes activement adaptatifs utilisent un test statistique pour détecter les points de changement \cite{MellorShapiro13,Allesiardo15}.
 %
@@
 %
-Nous prouvons qu'en combinant la politique \klUCB, asymptotiquement optimale pour les problèmes stationnaires, et notre nouvelle analyse du GLRT pour les récompenses limitées, nous obtenons des garanties de pointe sur le regret de notre algorithme proposé.
+Nous prouvons qu'en combinant l'algorithme \klUCB, asymptotiquement optimale pour les problèmes stationnaires, et notre nouvelle analyse du GLRT pour les récompenses limitées, nous obtenons des garanties de pointe sur le regret de notre algorithme proposé.
 Nous considérons la même hypothèse que nos concurrents, en supposant que la longueur des intervalles stationnaires est " assez longue " pour que les points de changement soient détectables.
@@
 %
-La performance de \GLRklUCB{} est également illustrée par des expériences numériques sur des données synthétiques, où il est démontré qu'elle surpasse toutes les politiques passivement adaptatives ainsi que les précédentes politiques activement adaptatives.
+La performance de \GLRklUCB{} est également illustrée par des expériences numériques sur des données synthétiques, où il est démontré qu'elle surpasse toutes les algorithms passivement adaptatifs ainsi que les précédentes algorithms activement adaptatifs.
 %
-Ce dernier chapitre est basé sur notre dernier travail, qui a mené à une publication à la conférence du GRETSI à Lille, France, en août $2019$ \cite{Besson2019Gretsi}.
+Ce dernier chapitre est basé sur notre dernier travail, publié à la conférence GRETSI à Lille, en août $2019$ \cite{Besson2019Gretsi}.
 Notre travail a aussi mené à l'article de la version longue \cite{Besson2019GLRT}, qui sera partiellement réécrit et complété avec des résultats plus récents, afin de le soumettre bientôt à une revue (probablement le Journal of Machine Learning Research) avant décembre $2019$.
 
-\paragraph{Annexe.}
+% \paragraph{Annexe.}
 %
@@
     \item
-    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les objets dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes BMB, pour augmenter automatiquement la durée de vie de leur batterie.
+    Nous proposons différents modèles pour les réseaux de l'IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les objets dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes BMB, pour augmenter automatiquement la durée de vie de leur batterie.
     Cela permet également à davantage d'objets d'utiliser le même réseau tout en maintenant un taux élevé d'accès au spectre sans souffrir de collisions radio
@@
         \node[align=center] (chapter25) at (2.5,2.25) [rectangle,draw,fill=red!20] {\textbf{Chapitre~\ref{chapter:25}}\\Sélection séquentielle\\d'algorithme BMB};
-        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles BMB\\pour les réseaux IdO};
+        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw,fill=green!10] {\textbf{Chapitre~\ref{chapter:4}}\\Deux modèles BMB\\pour les réseaux de l'IdO};
         \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!15] {\textbf{Chapitre~\ref{chapter:5}}\\Modèle BMB\\Multi-joueurs};
@@
 \textcolor{darkgreen}{La deuxième Partie~\ref{part:MABIOT}} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus quasiment indépendamment.
-Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux IdO, dans lesquels les algorithmes BMB ont été utilisés avec succès.
+Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux de l'IdO, dans lesquels les algorithmes BMB ont été utilisés avec succès.
 Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop complexes pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
@@
 afin d'établir des preuves mathématiques garantissant la convergence, ainsi que les gains en performance apportés par ces algorithmes de bandit.
-Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle BMB stationnaire à un joueur du Chapitre~\ref{chapter:2} et les modèles de réseaux IdO du Chapitre~\ref{chapter:4}.
+Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle BMB stationnaire à un joueur du Chapitre~\ref{chapter:2} et les modèles de réseaux de l'IdO du Chapitre~\ref{chapter:4}.
 Ces deux études ont chacune donné des résultats théoriques à la pointe de la recherche sur les modèles de bandits multi-joueurs et de bandits stationnaires par morceaux, que nous avons aussi validé par des expériences numériques.

### 2019-07-07 2-Chapters/1-Chapter/chapter1.tex
Citations de Kaamelott pour le début de la thèse… ??

@@
 
-\paragraph{Chapter~\ref{chapter:2}.}
+% \paragraph{Chapter~\ref{chapter:2}.}
 %
@@
 
-\paragraph{Chapter~\ref{chapter:3}.}
+% \paragraph{Chapter~\ref{chapter:3}.}
 %
@@
 
-\paragraph{Chapter~\ref{chapter:25}.}
+% \paragraph{Chapter~\ref{chapter:25}.}
 %
@@
 
-\paragraph{Chapter~\ref{chapter:4}.}
+% \paragraph{Chapter~\ref{chapter:4}.}
 %
@@
 The goal of this first model is to increase the Quality of Service (QoS) of the application of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmission.
-The base station will thus receive more up-link packets from the devices being served, if they can successfully learn an efficient spectrum access scheme.
+The base station will thus receive more up-link packets from the devices being served, if they can successfully learn by themselves an efficient spectrum access scheme.
 It implies that the network can accept more end-devices while maintaining the same QoS.
@@
 % Even without changing anything on the level of the considered IoT standard (neither the PHY nor the MAC layers), our proposal is just an add-on capability, that can be set-up on each device, on a unit-per-unit basis.
-The first model constitutes the first article written during this thesis, and initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
+The first model comes from the first article written during this thesis, which initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
 We presented it at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, where it obtained the ``best paper award'' \cite{Bonnefoi17}.
@@
 
-\paragraph{Chapter~\ref{chapter:5}.}
-%
-In both models presented above, the core concept is to let every dynamic device of an IoT network run a learning algorithm in a fully decentralized way, in order to optimize the complete system.
-It is important to note that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own (random) activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
-It means that each of them target its own local objective, which is to maximize its cumulated reward (in a selfish way).
+% \paragraph{Chapter~\ref{chapter:5}.}
+%
+In both models of IoT networks presented above, the core concept is to let every dynamic device of such network run a learning algorithm in a fully decentralized way, in order to optimize the complete system.
+It is important to note that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own random activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
+It means that each of them target its own local objective, which is to maximize its cumulated reward, in a selfish way and by being agnostic to the other devices following the same objective.
 It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure, one can think of popular ``dilemmas'', such as the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma}.
@@
 
-We first tried to analyze the performance of this decentralized algorithm, that we refer to as \Selfish, in the model of Section~\ref{sec:4:firstModel},
+That is why we first tried to analyze the performance of this decentralized algorithm, that we refer to as \Selfish, in the model of Section~\ref{sec:4:firstModel},
 but due to the random number of active devices at each time step (\ie, as soon as the probability $p$ of activation is $p < 1$), we were unable to develop a clean analysis.
 %
-That is why in Chapter~\ref{chapter:5}, we relax the hypothesis of having $M \gg K$ (or even simply $M > K$) devices in a network with $K$ orthogonal wireless channel, that is equivalent to having $p < 1$, and we consider the case of IoT devices communicating at every time step (\ie, $p=1$), and so we restrict to only up-to $M \leq K$ devices.
-The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing) or the IoT case (\ie, without sensing).
+That is the reason why we relax the hypothesis of having $M \gg K$ (or even simply $M > K$) devices in a network with $K$ orthogonal wireless channel in Chapter~\ref{chapter:5}, that is equivalent to having $p < 1$, and we consider the case of IoT devices communicating at every time step (\ie, $p=1$), and so we restrict to only up-to $M \leq K$ devices.
+The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing of PU) or the IoT case (\ie, without sensing).
 The goal was to understand the heuristic of Chapter~\ref{chapter:4}, \Selfish, in the simpler framework of multi-player MAB, which has been studied before for the case of OSA, as studied for instance by \cite{Zhao10,Anandkumar10,Anandkumar11}.
 %
-The OSA case covered by the studied model is actually slightly different as the one considered by \cite{Jouini10} and other previous works,
+The OSA case covered by the studied model can seem to slightly differ with the one considered by \cite{Jouini10} and other previous works,
 as our model requires that an \Ack{} is sent back by the base station if the transmission was successful, even in the case where sensing information is available.
-It differs from the previously studied models of applying MAB for OSA, as they only considered synchronized devices, one SU and PU, and thus if the sensing indicates that one channel is free at one time slot, the up-link message sent by the SU device is sure to be successfully received by the base station (in the ideal model), so there is no risk of collision, and thus no need for an \Ack.
-
-On the one hand, in the OSA setting, we failed to obtain positive result for the Selfish policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
-This result was later confirmed and analyzed by \cite{BoursierPerchet18}.
+It is actually equivalent to the previously studied models of applying MAB for OSA, as they only considered synchronized devices, one SU and PU, and thus if the sensing indicates that one channel is free at one time slot, the up-link message sent by the SU device is sure to be successfully received by the base station (in the ideal model), so there is no risk of collision, and thus no need for an \Ack.
+
+On the one hand, in the OSA setting, we failed to obtain positive result for the \Selfish{} policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
+This result was later confirmed and analyzed by \cite{BoursierPerchet18} (Appendix~F).
 %
@@
 The difficulty does not reside in the fact that we are trying to analyze MAB algorithms, designed to tackle stationary problems, on a non-stationary problem,
-but rather that we are analyzing algorithms which are all playing in different (random) subsets of the global synchronized time steps.
-Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is not really realistic for IoT networks, even though this lead to the model of Chapter~\ref{chapter:5}.
+but rather that we are analyzing algorithms which are all playing in different (random) activation times, subsets of the global synchronized time steps.
+Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is in fact not really realistic for IoT networks, even though this lead to the interesting model of Chapter~\ref{chapter:5}.
 
@@
 
-On the other hand, another way to relax the non-stationary hypothesis is to take the point-of-view of one device, like in the OSA case mentioned above, where the focus is on one SU surrounded by many PU having stationary behaviors.
-If we focus on one IoT device, its environment (\ie, the surrounding devices) is non-stationary, meaning that its average properties can fluctuate with time, but realistic environments usually yield a certain structure on this non-stationarity.
+On the other hand, another way to relax the non-stationary hypothesis is to take the point-of-view of a single device, like in the OSA case mentioned above, where the focus is on one SU surrounded by many PU having stationary behaviors.
+If we focus on one IoT device, its environment (\ie, the surrounding devices) is non-stationary, meaning that its average properties can fluctuate with time, but real environments usually present non-stationarities with certain structures.
 For instance, enforcing stationarity on some time intervals leads to the last contribution and the last chapter of this thesis.
@@
 
-\paragraph{Chapter~\ref{chapter:6}.}
+% \paragraph{Chapter~\ref{chapter:6}.}
 %
@@
 %
-This last chapter is based on our last work, which lead to a publication at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
+This last chapter is based on our last work, published at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
 This work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results, in order to submit it soon to a journal (probably the Journal of Machine Learning Research) before December $2019$.
 
-\paragraph{Appendix.}
+% \paragraph{Appendix.}
 %

### 2019-07-07 2-Chapters/Conclusion/conclusion.tex
Citations de Kaamelott pour le début de la thèse… ??

@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:2}}.}
+% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:2}}.}
 
-We started by presenting the multi-armed bandit model, with a finite number of arms and real-valued rewards.
+We started in Chapter~\ref{chapter:2} by presenting the multi-armed bandit model, with a finite number of arms and real-valued rewards.
 Our focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:3}}.}
+% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:3}}.}
 
-Then we presented our Python library SMPyBandits \cite{SMPyBanditsJMLR,SMPyBandits}, by detailing
+Then we presented in Chapter~\ref{chapter:3} our Python library SMPyBandits \cite{SMPyBanditsJMLR,SMPyBandits}, by detailing
 its purpose and qualities, its organization, and an overview of its usage.
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:25}}.}
+% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:25}}.}
 %
-This first part ends by presenting one of our first contributions \cite{Besson2018WCNC}.
+This first part ends by presenting one of our first contributions \cite{Besson2018WCNC}, in Chapter~\ref{chapter:25}.
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:4}}.}
+% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:4}}.}
 
-This chapter started the second part of the manuscript, Part~\ref{part:MABIOT}, and presents three main contributions.
-
+Chapter~\ref{chapter:4} started the second part of the manuscript, Part~\ref{part:MABIOT}, and presents three main contributions.
+%
 First in Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of applying MAB learning algorithms in IoT networks,
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:5}}.}
+% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:5}}.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Conclusion about \textbf{Chapter~\ref{chapter:6}}.}
+% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:6}}.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Perspectives about \textbf{Part~\ref{part:Introduction}}.}
+% \paragraph{Perspectives about \textbf{Part~\ref{part:Introduction}}.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Perspectives about \textbf{Chapter~\ref{chapter:4}}.}
+% \paragraph{Perspectives about \textbf{Chapter~\ref{chapter:4}}.}
 
@@
 % ----------------------------------------------------------------------------
-\paragraph{Perspectives about \textbf{Chapter~\ref{chapter:5}}.}
+% \paragraph{Perspectives about \textbf{Chapter~\ref{chapter:5}}.}
 
-The study on multi-players bandits suggests several further research directions.
+The study on multi-players bandits in Chapter~\ref{chapter:5} suggests several further research directions.
 First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
@@
 % ----------------------------------------------------------------------------
-\paragraph{Perspectives about \textbf{Chapter~\ref{chapter:6}}.}
+% \paragraph{Perspectives about \textbf{Chapter~\ref{chapter:6}}.}
 
-We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
+We believe that our new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
 This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of the number of break-points $\Upsilon_T$.

### 2019-07-07 1-Preamble/5-Abstract/resume_fr.tex
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 Ce dernier chapitre est basé sur notre dernier travail, publié à la conférence GRETSI à Lille, en août $2019$ \cite{Besson2019Gretsi}.
-Notre travail a aussi mené à l'article de la version longue \cite{Besson2019GLRT}, qui sera partiellement réécrit et complété avec des résultats plus récents, afin de le soumettre bientôt à une revue (probablement le Journal of Machine Learning Research) avant décembre $2019$.
+Notre travail a aussi mené à une version longue \cite{Besson2019GLRT}, qui sera complétée avec des résultats plus récents, afin de le soumettre bientôt à une revue (probablement le Journal of Machine Learning Research) avant décembre $2019$.
 
@@
     \item
-    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle BMB multi-joueurs, qui a été particulièrement active ces deux dernières années.
+    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle BMB multi-joueurs, particulièrement active ces deux dernières années.
     % qui, selon nous, n'a jamais été écrit auparavant.
@@
 
+\hr{}
+
 \textbf{Note sur le droit intellectuel.}

### 2019-07-07 2-Chapters/1-Chapter/chapter1.tex
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 The goal of the player is to maximize its rewards, by trials and errors (\ie, actions and rewards).
-Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or to drive, learning to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial), etc.
+Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or drive, to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial), etc.
 
@@
 However, sensing is known to be costly, in terms of energy consumption, especially for wide-band sensing, as detailed in the surveys \cite{yucek2009survey,subhedar2011spectrum}, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
-This assumption, along with the hypothesis that PU cannot be disturbed (without which no regulation will never be accepted for OSA), enforces the SU to transmit in the channel that it sensed, in each time step, if and only if it was sensed free from any PU.
+This assumption, along with the hypothesis that PU cannot be disturbed (without which no regulation will never be accepted for OSA), enforces the SU to transmit in the channel that it sensed, if and only if it was sensed free from any PU.
 % non negligible cost of radio reconfiguration
@@
 under partial information feedback (\ie, the player receives feedback about only one channel over $K$).
-These three hypotheses are the ones that restrict from the general ML framework to the specific Multi-Armed Bandit case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
+These three hypotheses are the ones that restrict from the general ML framework to the specific MAB case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
 
@@
 %
-We answer partially to this problematic by the following research steps.
+We answer partially to this problematic by the following steps.
 
@@
 %
-In both models of IoT networks presented above, the core concept is to let every dynamic device of such network run a learning algorithm in a fully decentralized way, in order to optimize the complete system.
+In both models of IoT networks presented above, the core concept is to let every dynamic device run a learning algorithm in a fully decentralized way, in order to optimize the complete system.
 It is important to note that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own random activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
 It means that each of them target its own local objective, which is to maximize its cumulated reward, in a selfish way and by being agnostic to the other devices following the same objective.
-It is well known in the game theory literature that playing selfishly can be disastrous for the centralized performance measure, one can think of popular ``dilemmas'', such as the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma}.
+It is well known in game theory that playing selfishly can be disastrous for the centralized performance measure, one can think of popular ``dilemmas'', such as the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma}.
 % , and for more examples we refer to the book \cite{CesaLugosi06}.
@@
     by \textbf{Lilian Besson} \& Emilie Kaufmann,
-    September $2018$.\\
+    September $2018$.
     % See Appendix~\ref{app:2:DoublingTricks}.
@@
 
+\hr{}
+
 \paragraph{Copyright notice.}

### 2019-07-07 2-Chapters/2-Chapter/chapter2.tex
Improved all mentions to the (wrong) lower-bound of Ch6
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 %
+As explained in Chapter~\ref{chapter:1}, we focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated to a binary or real-valued distribution.
 In this chapter, we present the common base of all the more sophisticated mathematical models studied in this thesis: the stochastic multi-armed bandit (MAB) model, by restricting to the single-player and stationary stochastic case.
@@
 We motivate this model and discuss its application to and beyond cognitive radio.
-We review existing MAB algorithms,
-some of which are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling.
-%
-We focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated with a binary or real-valued distribution.
 %
 We formalize this MAB model, then we quote some of its important applications.
-We define the notion of regret of an algorithm, and present important results from the literature. A short review of the most important families of MAB algorithms is then given.
+We define the notion of regret of an algorithm, and present important results from the literature.
+A short review of the most important families of MAB algorithms is then given,
+and we highlight some algorithms that are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling.
+%
 % We conclude this chapter by discussing the question of online algorithm selection.
@@
 
-\subsection{Regret lower bounds}
+\subsection{Regret lower-bounds}
 
@@
 
+\begin{figure}[h!]  % [htbp]
+	% \centering
+	\includegraphics[width=1.12\linewidth]{SP__K5_T10000_N1000__3_algos/main____env1-1_776571591054077529.pdf}
+	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$, for the 3 most efficient algorithms: $\UCB_1$, \klUCB{} and Thompson Sampling.}
+	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000__FIXME}
+\end{figure}
+
 % % FIXME

### 2019-07-07 2-Chapters/3-Chapter/chapter3.tex
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
-Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, and a study to compare time and memory costs of the main algorithms.
+Then we use it to perform a numerical comparisons of the main state-of-the-art single-player bandit algorithms, and a study to compare time and memory costs of some of the most well-known and most widely used MAB algorithms.
 
@@
 
-$(i)$
+\begin{itemize}
+    \item[$(i)$]
 Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
@@
 
-$(ii)$
+    \item[$(ii)$]
 We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
@@
 
-$(iii)$
+    \item[$(iii)$]
 Finally, we presented how to use SMPyBandits, if one wants to run some pre-designed simulations or design new simulations.
@@
 We give below in Appendix~\ref{sec:3:appendix} a small but complete example of the two files that has to be written for a new simulation: a \texttt{main} file which essentially defines the desired visualizations after the simulation, and a \texttt{configuration} file which defines the problems to simulate and their parameters, as well as the algorithms that will be compared.
+\end{itemize}
 

### 2019-07-07 2-Chapters/4-Chapter/chapter4.tex
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 %
-In this chapter, we propose two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
-that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
+After detailing in Part~\ref{part:Introduction} the MAB model, we now come back to our main question of interest.
+In this chapter, we focus on wireless networks following the hypotheses of common to the present and future Internet of Things (IoT) networks exposed in Chapter~\ref{chapter:1}.
+Our goal is to show that the IoT devices can automatically learn to increase their battery life as well as their successful transmission rates, without changing anything on the IoT standard.
+%
+We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
 %
@@
 However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
-We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics significantly improve the network efficiency.
+We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics significantly also improve the network efficiency.
 % Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.

### 2019-07-07 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 % \section{An improved asymptotic regret lower bound (for some algorithms)}
-\section[Decomposing the regret to get an intuition about efficient decentralized algorithms]{Decomposing the regret to get an intuition for the design of efficient decentralized algorithms}
+\section[Regret decomposition and intuition about efficient decentralized algorithms]{Decomposing the multi-player regret to get an intuition for the design of efficient decentralized algorithms}
 \label{sec:5:lowerbound}

### 2019-07-07 2-Chapters/5-Chapter/chapter5.tex
Not sure if I say enough about the lower-bounds...
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 %
-In this chapter, we are interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
-We restrict to the easier case of at most $M \leq K$ devices in a network with $K$ channels, because it was found very hard to analyze the aforementioned IoT network model.
+In this chapter, we are interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
+We restrict to the easier case of at most $M \leq K$ devices in a network with $K$ channels, that each have data to transmit at every instant, that is they have an activation probability of $p=1$, because it was found very hard to analyze the previously studied IoT network models.
 %
@@
 and their analysis shows that when the SUs faithfully implement proposed algorithms, the regret is constant with high probability.
-They validate their claims through exhaustive synthetic experiments and also through a realistic USRP.
+They validate their claims through extensive synthetic experiments and also through a realistic USRP.
 Their synthetic experiments consider different cases, for different values of $K$ the number of channels, $M$ the number of players and $J$ the number of jammers.

### 2019-07-07 2-Chapters/6-Chapter/chapter6.tex
De meilleures transitions, j'ai retravaillé chaque abstract

@@
 %
-In this last chapter, we study a generalization of the stationary multi-armed bandit model,
+In this last chapter, we are also interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
+Instead of keeping the stationary hypothesis but considering more players, as we did in Chapter~\ref{chapter:5}, we are now interested in another direction of formal analysis of the intractable models of IoT networks of Chapter~\ref{chapter:4}.
+%
+We study a generalization of the single-player stationary multi-armed bandit model,
 to account for possible non-stationarity of the rewards.

### 2019-07-08 2-Chapters/2-Chapter/chapter2.tex
Okay for figures 2.5

@@
 %
+% \begin{small}
 \begin{smallproof}\label{proof:2:RegretDecomposition}
@@
 \end{smallproof}
+% \end{small}
 
@@
 %
+% \begin{small}
 \begin{smallproof}\label{proof:2:tuningExploreThenCommit}
@@
 \end{smallproof}
+% \end{small}
 %
@@
 %
+% \begin{small}
 \begin{smallproof}\label{proof:2:UCBregretBound}
@@
 The \KLUCB{} and \klUCB{} algorithms achieve finite-time regret logarithmic upper-bounds on their regret, and have been proven to be asymptotically optimal, respectively for one-dimensional exponential families and for binary rewards \cite{Garivier11KL,KLUCBJournal}.
-In particular, we highlight that \klUCB{} achieves a finite-time regret upper-bound, that matches the asymptotical lower-bound of Lai and Robbins \cite{LaiRobbins85} for Bernoulli distributed problems, which is our main focus of interest in both Chapters~\ref{chapter:5} and \ref{chapter:6}.
+In particular, we highlight that \klUCB{} achieves a finite-time regret upper-bound, that matches the asymptotical lower-bound of Lai and Robbins \cite{LaiRobbins85} for Bernoulli distributed problems, which is our main focus in both Chapters~\ref{chapter:5} and \ref{chapter:6}.
 
@@
 then they use the following indexes:
+\vspace*{-5pt} % FIXME
 \begin{align*}\label{eq:2:indexes_for_MOSS_klUCBpp}
@@
 are defined by
-$U^{\mathrm{MOSS}+}_k(t) \eqdef \frac{X_k(t)}{N_k(t)} + \max\left(0, \sqrt{ \log\left(T / (K N_k(t))\right) / N_k(t)} \right)$
+$U^{\mathrm{MOSS}+}_k(t) \eqdef \frac{X_k(t)}{N_k(t)} + \max (0, \sqrt{ \log\left(\frac{T}{K N_k(t)}\right) / N_k(t)} )$
 and
-$U^{\KLUCB+}_k(t) \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \log(T / (K N_k(t))) / N_k(t) \right\}$,
+$U^{\KLUCB+}_k(t) \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \log(\frac{T}{K N_k(t)}) / N_k(t) \right\}$,
 and if $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,
@@
 \begin{table}[!h]
-    % \begin{small}  % WARNING
+    \begin{small}  % WARNING
         \centering
@@
         \label{table:2:oneShotResult}
-    % \end{small}  % WARNING
+    \end{small}  % WARNING
 \end{table}
@@
 which should converge to $\mu^*$ when $t\to\infty$ for efficient algorithms.
-Conversely, we also add the Lai \& Robbins' logarithmic \emph{lower-bound} on regret ($C_I \log(t)$ from Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plot in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
+Conversely, we also add the Lai \& Robbins' logarithmic \emph{lower-bound} on regret ($C_I \log(t)$ from Theorem~\ref{thm:2:forSecondLogTLowerBound}) on the regret plots in Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}.
 It should not be surprising that the black line of the lower-bound is \emph{above} the values of the regret as the lower-bound is only asymptotic.
@@
 We can clearly identify the behavior of the different strategies:
-the three efficient strategies indeed achieve a sub-linear regret, that looks logarithmic as supported by the theory,
+the three efficient strategies indeed achieve a sub-linear regret (see Figure~\ref{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000_3_algos}), that looks logarithmic, as supported by the theory,
 while the naive strategies achieve linear regret.
 The two simple strategies also achieve sub-linear regret but are less efficient:
-both $\varepsilon$-greedy and Explore-then-Commit starts as pure exploration, respectively until $\varepsilon_0/t$ becomes smaller than $1$ and until $t > T_0$, then they usually identify correctly the best arm and start to catch up with the best policies.
+both $\varepsilon$-greedy and E-t-C starts as pure exploration, respectively until $\varepsilon_0/t$ becomes smaller than $1$ and until $t > T_0$, then they usually identify correctly the best arm and start to catch up with the best policies.
+
+\begin{figure}[h!]  % [htbp]
+	% \centering
+    % \includegraphics[width=1.05\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
+    \includegraphics[width=1.10\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_3198350886837790844.pdf}
+	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
+	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
+\end{figure}
 
@@
 \begin{table}[ht]
-    % \begin{small}  % WARNING
+    \begin{small}  % WARNING
         \centering
@@
         \label{table:2:meanResults}
-    % \end{small}  % WARNING
+    \end{small}  % WARNING
 \end{table}
+
+\begin{figure}[h!]  % [htbp]
+    \centering
+    % [][$7$ algorithms, including unefficient ones]
+    \subfloat{
+        % \includegraphics[width=0.84\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
+        \includegraphics[width=0.92\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_3198350886837790844.pdf}
+        \label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000_7_algos}
+    }
+    \newline
+    \vspace*{-20pt}%FIXME
+    % [][$3$ algorithms, only the most efficient]
+    \subfloat{
+        \includegraphics[width=0.92\linewidth]{SP__K5_T10000_N1000__3_algos/main____env1-1_776571591054077529.pdf}
+        \label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000_3_algos}
+    }
+	% \caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
+	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$. The 3 most efficient algorithms: $\UCB_1$, \klUCB{} and Thompson Sampling achieve logarithmic regret.}
+	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
+\end{figure}
 
@@
 
-\paragraph{Summary of this review of MAB algorithms.}
-%
-We gave an overview of the most well known families of algorithms for multi-armed bandits,
+\subsection{Summary of this review of MAB algorithms}
+%
+We gave an overview of the most well known families of MAB algorithms,
 focussing on algorithms designed and analyzed for the single-player stationary and stochastic MAB model.
@@
 To clearly understand the difference in their empirical behaviors, we illustrated the performance of two naive and two simple strategies against three efficient ones, on the example of Bernoulli-distributed problem used for the online demonstration from Section~\ref{sec:2:notations} (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
-This also serves as a first example of numerical simulation performed with our library SMPyBandits, before the large-scale experiments of Section~\ref{sec:3:reviewSPAlgorithms} below.
-%
-In the rest of this thesis, we mainly use the $\UCB_1$ or Thompson sampling algorithms when simplicity and realistic deployment is favored (in Chapter~\ref{chapter:4}), and the \klUCB{} algorithm when we give mathematical developments (in Chapters~\ref{chapter:5} and \ref{chapter:6}).
-
-\begin{figure}[h!]  % [htbp]
-	% \centering
-    \includegraphics[width=1.20\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
-	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
-	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
-\end{figure}
-
-\begin{figure}[h!]  % [htbp]
-	% \centering
-	\includegraphics[width=1.12\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
-	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
-	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
-\end{figure}
-
-\begin{figure}[h!]  % [htbp]
-	% \centering
-	\includegraphics[width=1.12\linewidth]{SP__K5_T10000_N1000__3_algos/main____env1-1_776571591054077529.pdf}
-	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$, for the 3 most efficient algorithms: $\UCB_1$, \klUCB{} and Thompson Sampling.}
-	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000__FIXME}
-\end{figure}
+This also serves as a first example of simulation performed with our library SMPyBandits, before the large-scale experiments of Section~\ref{sec:3:reviewSPAlgorithms} below.
+%
+In the rest of this thesis, we mainly use $\UCB_1$ or Thompson sampling when simplicity and realistic deployment is favored (in Chapter~\ref{chapter:4}), and \klUCB{} when we give mathematical developments (in Chapters~\ref{chapter:5} and \ref{chapter:6}).
+
+% \begin{figure}[h!]  % [htbp]
+% 	% \centering
+% 	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$, for the 3 most efficient algorithms: $\UCB_1$, \klUCB{} and Thompson Sampling.}
+% 	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000_onlyLogarithic}
+% \end{figure}
 

### 2019-07-08 1-Preamble/4-Acknowledgement/acknowledgement.tex
Je vais au bureau, une dernière relecture de résumé / intro fr / intro anglais / conclusion et ça ira

@@
 \emph{Merci infiniment de nous avoir offert cette curiosité, merci d'être si gentils, et merci de votre soutien constant et de votre amour !}
-\\
-\indent
-I would then like to thank my closest friends, sorted in chronological order of our first encounter in the \href{https://bulbapedia.bulbagarden.net/wiki/Tall_grass}{tall grass}\footnote{~Sorry if I forgot you, it's much harder to keep track of wild encounters in the tall grass in real wife, without a Pokédex!
-And due to space constraints, I had to choose quite arbitrarily whom to include\dots}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Audrey, Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
+% \\
+% \indent
+
+I would then like to thank my closest friends, sorted in chronological order of our first encounter in the \href{https://bulbapedia.bulbagarden.net/wiki/Tall_grass}{tall grass}\footnote{~Sorry if I forgot you, it's harder to keep track of wild encounters in the tall grass without a Pokédex!}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Audrey, Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
+%  And due to space constraints, I had to choose quite arbitrarily whom to include\dots
 % In the last four years, thank you so much my dear friends, for great moments and an even greater moral support, for the holidays in India, Switzerland, Hungary/Austria, Strasbourg, Greece, Saint-Malo, Normandy, Scotland, Marseille, Madrid, Brittany etc.
 I loved traveling with you and hope to continue!
-\emph{Une pensée particulière pour Hélène, merci pour ton énergie et ta détermination depuis 2018.}
+\emph{J'ai une pensée particulière pour Hélène, merci pour ton énergie et ta détermination, notamment en 2018-19.}
 
@@
 \indent
-I am grateful to all the amazing professors who shared their passion for mathematics and computer science with me.
+I am grateful to amazing professors who passed on their passion for mathematics and computer science to me.
 A partial list includes Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers, Marseille.
@@
 in Computer Science, and to Carine S.-P., Delphine L.
-and the late Nicolas P.\textsuperscript{$\dagger$} for their support.
-I also thank Didier C., Vijaysekhar C. and Jai P. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$).
+and Nicolas P.\textsuperscript{$\dagger$} for their support.
+Thanks to Didier C. and Vijaysekhar C. in Hyderabad ($2014$-$15$), Michael U. and Julien F. in Lausanne ($2016$) for our collaborations.
 

### 2019-07-08 1-Preamble/4-Acknowledgement/acknowledgement.tex
J'ai voulu ajouter Malone, Nadine et Bruno ❤

@@
 %
-I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life. \emph{Ton courage et ta force m'inspirent quotidiennement, merci beaucoup !}
-Very warm thanks then go to M. and my grand-parents, J. \& J.-M. and J., and to my uncle J. and aunt B.!
-Finally, I thank sincerely my parents, C. and P., for being the best parents since the last $26$ years, and for their (vital) moral support.
+I also thank my younger brother Fabian, for being an inspiration for other (highly important) directions of my life.
+\emph{Tes voyages, ton courage et ta force m'inspirent quotidiennement, merci beaucoup !}
+I sincerely thank my parents C. \& P., for being the best parents since the last $26$ years, and for their (vital) moral support.
 \emph{Merci infiniment de nous avoir offert cette curiosité, merci d'être si gentils, et merci de votre soutien constant et de votre amour !}
+Very warm thanks then go to M., to my cousin L. and my grand-parents, J. \& J.-M. and J., to my uncle J. and aunt B.!
+\emph{Merci enfin à M., N. \& B. pour votre accueil à La Gouesnière !}
 % \\
@@
 
-I would then like to thank my closest friends, sorted in chronological order of our first encounter in the \href{https://bulbapedia.bulbagarden.net/wiki/Tall_grass}{tall grass}\footnote{~Sorry if I forgot you, it's harder to keep track of wild encounters in the tall grass without a Pokédex!}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Audrey, Rémi, Marine, Adrien, Bastien, Claire, Corentin and Lola in Rennes.
+I would then like to thank my closest friends, sorted in chronological order of our first encounter in the \href{https://bulbapedia.bulbagarden.net/wiki/Tall_grass}{tall grass}\footnote{~Sorry if I forgot you, it's harder to keep track of wild encounters in the tall grass without a Pokédex!}: Guillaume, Florian, Marian, Hélène, Lucie, Lætitia and Julia in Briançon, Mayotte and Pierre in Marseille, Laurent, Simon, Tristan, Ludovic, Jessica, Romain, Benjamin, Angèle, Clément in Cachan, Jill-Jênn, Alain, Claire, Édouard, Loïc and Damien also in Cachan, Kumudham and Priscilla in India, Thibault, Sébastien, Sélim, Valentin again in Cachan, and Audrey, Rémi, Marine, Adrien, Bastien, Claire  and Lola in Rennes.
 %  And due to space constraints, I had to choose quite arbitrarily whom to include\dots
@@
 I loved traveling with you and hope to continue!
-\emph{J'ai une pensée particulière pour Hélène, merci pour ton énergie et ta détermination, notamment en 2018-19.}
+\emph{J'ai une pensée particulière pour Hélène, merci pour ta foi, ton énergie et ta détermination, notamment en 2018-19.}
 
@@
 %
-Thanks to everybody who played \emph{``futsal''} on Wednesday lunch,
+Thanks to the \emph{``futsal''} players on Wednesday lunch,
 % I still have to progress immensely,
 % before I could pretend to be useful
-but it's been a lot of fun to play with you guys!
+I haven't been useful, but it's been a lot of fun to play with you guys!
 
@@
 I am grateful to amazing professors who passed on their passion for mathematics and computer science to me.
-A partial list includes Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers, Marseille.
+A partial list includes Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers in Marseille.
 Then at ENS de Cachan ($2011$-$16$), I especially think about Frédéric P., Alain T., Claudine P., Florian de V.
@@
 \emph{Je suis désolé si je vous ai oublié, merci de corriger} : Je remercie sincèrement ....................... pour .............
-\emph{I am sorry I forgot you, please fix this mistake}: I warmly thank .................... for ..........
+\emph{I am sorry I forgot you, please fix this mistake}: I warmly thank ..................... for ...............
 

### 2019-07-08 1-Preamble/5-Abstract/resume_fr.tex
Okay j'ai relu l'intro en français

@@
 Ce manuscrit conclut ma thèse de doctorat, qui a débuté en octobre $2016$ et s'est terminée en novembre $2019$.
-Mes recherches se sont déroulées au laboratoire de l'IETR à Rennes (France), dans l'équipe SCEE hébergée sur le campus de Rennes de l'école d'ingénieurs CentraleSupélec.
+Mes recherches se sont déroulées au laboratoire IETR à Rennes (France), dans l'équipe SCEE hébergée sur le campus de Rennes de l'école d'ingénieurs CentraleSupélec.
 J'étais supervisé par le professeur Christophe Moy, à Rennes,
@@
 %
-Les problèmes qui motivent cette thèse sont les questions du réchauffement climatique et de l'augmentation de la population mondiale.
+Les problèmes sous-jacents qui motivent cette thèse sont les questions du réchauffement climatique et de l'augmentation de la population mondiale.
 % https://en.wikipedia.org/wiki/History_of_radio
-Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, la télécommunication sans fil entre des appareils fabriqués par l'homme a été rendue possible, et de plus en plus fréquentes dans nos vies.
-Avec l'avènement des réseaux de l'Internet des Objets (IdO), des milliards d'objets à basse consommation devraient être déployés dans le monde entier, permettant un large éventail d'applications différentes.
-Il existe aujourd'hui un consensus mondial sur le fait qu'avec la tendance actuelle à l'accroissement démographique et la crise énergétique actuelle, toute nouvelle technologie déployée doit être à la fois bon marché et efficace sur le plan énergétique,
+Au cours des $150$ dernières années, l'humanité a développé de nombreuses technologies de communication différentes, et depuis la fin des années $1890$, les télécommunications sans fil entre des appareils fabriqués par l'homme ont été rendues possibles, et de plus en plus fréquentes dans nos vies.
+Avec l'avènement des réseaux de l'Internet des Objets (IdO), des milliards d'objets à basse consommation devraient être déployés dans le monde entier, permettant un large éventail de nouvelles applications.
+Il existe aujourd'hui un consensus mondial sur le fait qu'avec la tendance actuelle à l'accroissement démographique et la crise énergétique, toute nouvelle technologie déployée doit être à la fois bon marché et efficace sur le plan énergétique,
 ainsi qu'adaptée pour desservir un grand nombre de personnes et d'appareils.
@@
 %
-En plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, couvrant tous les cas possibles,
+C'est pourquoi en plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, couvrant tous les cas possibles,
 le moment est venu de le combiner avec un apprentissage machine bon marché et prometteur pour (essayer d') atteindre le niveau de gain de performance nécessaire pour que les promesses de l'IdO deviennent réalité.
 
-C'est pourquoi nous nous intéressons dans cette thèse sur
+C'est pourquoi nous nous intéressons dans cette thèse à
 l'amélioration de la durée de vie des batteries des objets de l'IdO et la réduction du coût énergétique des réseaux de l'IdO.
@@
 %
-Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (algorithmes de types bandits multi-bras), afin de permettre aux objets de l'IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
+Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (des algorithmes de types bandits multi-bras), afin de permettre aux objets de l'IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
 
-\paragraph{Des anciens téléviseurs aux normes de l'IdO.}
+\paragraph{Des anciens téléviseurs aux standards de l'IdO.}
 %
-Historiquement, trois grandes familles de systèmes de communication sans fil ont été déployées dans les grands réseaux commerciaux : d'abord, la radiodiffusion centralisée (\eg, radio ou télédiffusion), puis les systèmes bidirectionnels centralisés (\eg, 4G ou WiFi), et aujourd'hui la collecte décentralisée de données pour les réseaux Internet des objets (\eg, réseaux capteurs).
+Historiquement, trois grandes familles de systèmes de communication sans fil ont été déployées dans les grands réseaux commerciaux : d'abord, la radiodiffusion centralisée (\eg, radio ou télédiffusion), puis les systèmes bidirectionnels centralisés (\eg, 4G ou WiFi), et aujourd'hui la collecte décentralisée de données pour les réseaux de l'Internet des objets (\eg, réseaux capteurs).
 %
@@
 Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les ``bâtiments connectés'' et l'``agriculture connectée''.
-Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides et d'utiliser des réseaux de capteurs afin d'obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, dans le but de permettre au contrôle centralisé du chauffage d'optimiser son coût et sa consommation énergétique.
+Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides en utilisant des réseaux de capteurs afin d'obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, pour permettre au contrôle centralisé du chauffage d'optimiser son coût et sa consommation énergétique.
 Pour l'agriculture, un exemple peut être d'équiper chaque tête de bétail (dans les grandes fermes) de capteurs qui émettent régulièrement des informations biologiques, telles que la température corporelle ou le niveau de stress, afin d'optimiser l'heure de traite des vaches, de surveiller la santé des animaux, etc.
 
-Ce troisième type de systèmes sans fil est caractérisé par leur nature décentralisée,
+Ce troisième type de systèmes sans fil sont caractérisés par leur nature décentralisée,
 où les communications sont initialisées et régulées par les objets, et non par un système de contrôle centralisé.
 %
-En effet, un contrôle centrale nécessite des paquets de signalisation qui ont été identifiés comme trop lourds pour ce type de systèmes.
+En effet, un contrôle centrale nécessite des paquets de signalisation très réguliers, qui ont été identifiés comme trop lourds pour ce type de systèmes.
 Dans les réseaux de l'IdO actuels et futurs, de nombreux objets hétérogènes utilisent la même station de base (aussi appelée point d'accès, ou \emph{base station}, \emph{access point} ou \emph{gateway}) pour des applications différentes.
-Un problème commun est la forte contrainte en terme de consommation d'énergie de ces objets de l'IdO, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
-Ainsi, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des objets ayant une durée de vie de plus de $10$ ans, comme SIGFOX \cite{Centenaro16}.
+Un problème commun est la forte contrainte en terme de consommation énergétique de ces objets, car la plupart d'entre eux seront déployés sans alimentation directe et fonctionneront sur une batterie minuscule, dont la durée de vie devrait être maximisée.
+Ainsi, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des objets ayant une durée de vie supérieure à $10$ ans, comme SIGFOX \cite{Centenaro16}.
 Une autre contrainte commune aux objets de l'IdO est leur faible besoin en communication, car la plupart des applications n'auront besoin d'envoyer qu'un ou quelques messages chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
@@
 De nombreuses normes différentes pour les réseaux de l'IdO ont été proposées ces dernières années,
-et ils consistent en une spécification à la fois pour la couche \emph{PHY}sique
+et elles consistent en une spécification à la fois pour la couche \emph{PHY}sique
 et la couche \emph{M}edium \emph{A}access \emph{C}ontrol (\emph{MAC}).
-Pour citer quelques exemples de normes pour l'IdO, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée (jusqu'à \SI{2}{\meter}), tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
-Nous nous référons à l'article \cite{Centenaro16} pour plus de détails, et les références dans \cite{Azari18} ou notre récent article \cite{MoyBesson2019}.
+Pour citer quelques exemples de normes pour des réseaux de type IdO, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée (jusqu'à \SI{2}{\meter}), tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
+Nous nous référons à l'article \cite{Centenaro16} pour plus de détails, et les références dans \cite{Azari18} ou notre dernier travail \cite{MoyBesson2019,MoyBesson2019Annales}.
 
@@
 
-Les bandes des réseaux cellulaires sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radio amateures) sont moins utilisées.
+Les bandes des réseaux cellulaires (2G/3G/4G/5G) sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radio amateures) sont moins utilisées.
 Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend fortement à la fois du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
@@
 les organismes de réglementation dans le monde se sont demandé s'il fallait permettre un nouveau paradigme de communication sans fil :
-permettre aux utilisateurs non titulaires d'une licence d'utiliser des bandes sous licence s'ils ne causent pas d'interférence aux utilisateurs payants, \ie, titulaires d'une licence.
+permettre aux utilisateurs non titulaires d'une licence d'utiliser des bandes sous licence s'ils ne causent pas d'interférence aux utilisateurs payants, titulaires d'une licence.
 Ces initiatives sont examinées par la \textbf{Radio Intelligente} (RI, ou \emph{Cognitive Radio}, CR),
@@
 des technologies de l'information et communication (TIC), aux télécommunications, aux communications sans fil, puis à la Radio Intelligente,
-et enfin à la Radio Logicielle (\emph{Software Defined Radio}, SDR).
+et enfin à la Radio Logicielle (RL, \emph{Software Defined Radio}, SDR).
 %
-La transition de l'approche historique de la radio matérielle aux architectures SDR est un processus graduel, qui a commencé au début des années $1990$ et s'est accéléré dans les années $2000$.
+La transition de l'approche historique de la radio matérielle aux architectures RL est un processus graduel, qui a commencé au début des années $1990$ et s'est accéléré dans les années $2000$.
 % https://en.wikipedia.org/wiki/Software-defined_radio
-Un SDR est un système de radiocommunication où les composants qui ont été traditionnellement implémentés par du matériel dédiés (\eg, mélangeurs, filtres, modulateurs/démodulateurs, détecteurs, etc) sont de plus en plus implémentés au moyen de logiciels sur un processeur.
+Une RL est un système de radiocommunication où les composants qui ont été traditionnellement implémentés par du matériel dédiés (\eg, mélangeurs, filtres, modulateurs/démodulateurs, détecteurs, etc) sont de plus en plus implémentés au moyen de logiciels sur un processeur.
 %
-Même si le paradigme de la RI a été initié par la recherche américaine en matière de défense dans les années $1980$, au cours des $20$ dernières années, l'industrie civile a commencé à s'intéresser à la RI, et la RI a également suscité beaucoup d'intérêt de la part du milieu universitaire et de l'industrie.
+Même si le paradigme de la RI a été initié par des recherches de l'armée des États-Unis d'Amérique du Nord dans les années $1980$, l'industrie civile a commencé à s'intéresser à la RI au cours des $20$ dernières années, et la RI a également suscité beaucoup d'intérêt de la part du milieu universitaire.
 %
-Comme la RI n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RI, qui vient de paraître il y a $20$ ans.
+Comme la RI n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RI, dont le premier vient de paraître il y a $20$ ans.
 %
@@
 %
-Dans les bandes sous licences, il y a des utilisateurs primaires (UP) qui paient pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
+Dans les bandes sous licences, il y a des utilisateurs primaires (UP) qui paient un abonnement pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
 Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est utilisé par aucun UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser aussi, éventuellement en payant un abonnement auprès de l'opérateur du réseau.
@@
 Nous sommes intéressés par l'apprentissage séquentiel et en particulier par les Bandits Multi-Bras (BMB),
-qui est apparu comme un problème intéressant dans la communauté statistique et parmi les chercheurs intéressés par l'apprentissage séquentiel, comme le montrent les travaux antérieurs de \cite{Thompson33,Robbins52,LaiRobbins85}.
-Les BMB sont également inclus dans le concept plus général d'apprentissage par renforcement (\emph{Reinforcement Learning}, RL), lui-même un des domaines de l'apprentissage machine (\emph{Machine Learning}, ML).
+qui sont apparus comme des problèmes intéressants au sein de la communauté des statistiques et de l'apprentissage séquentiel, comme le montrent les travaux pionniers de \cite{Thompson33,Robbins52,LaiRobbins85}.
+Les BMB sont également inclus dans le concept plus général d'apprentissage par renforcement (ApR, ou \emph{Reinforcement Learning}, RL), lui-même un des domaines de l'apprentissage machine (AM, ou \emph{Machine Learning}, ML).
 % ML > RL > MAB
 %
-Le livre de référence sur le RL est \cite{SuttonBarto2018}, citons donc la définition qu'en donnent R. Sutton et A. Barto :
+Le livre de référence sur l'ApR est \cite{SuttonBarto2018}, citons donc la définition qu'en donnent R. Sutton et A. Barto :
 \emph{``l'apprentissage par renforcement, c'est apprendre quoi faire - comment faire le lien entre les situations et les actions -
@@
 en Figure~\ref{fig:1:ReinforcementLearningCycle}.
-Une joueuse (ou un apprenant) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de cette action, fournie par l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
+Un-e joueur-se (ou un-e apprenant-e) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de cette action, fournie par l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
 Le but de la joueuse est de maximiser ses récompenses, par des essais et des erreurs (\ie, actions et récompenses).
-De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage par renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, par exemple apprendre à marcher ou à conduire, apprendre à jouer à un jeu vidéo, ou découvrir quel traitement est efficace pour guérir une certaine maladie, etc.
+De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage par renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, par exemple apprendre à marcher, à conduire, à jouer à un jeu vidéo, découvrir quel traitement est efficace pour guérir une certaine maladie, etc.
 
 % Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
-Comme cette thèse se concentre sur les modèles d'apprentissage par renforcement (RL),
+Comme cette thèse se concentre sur les modèles d'apprentissage par renforcement,
 il est important de souligner que dans de tels modèles de prise de décision, l'apprenant n'a pas accès à l'ensemble des réactions possibles de l'environnement après avoir choisi son action.
 %
-En d'autres termes, la joueuse ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une autre action.
-Ce genre de rétroaction limitée s'appelle l'information de type bandit (\emph{bandit feedback}), et nous discutons de l'histoire et du concept des \emph{bandits multi-bras} (BMB) en détail dans le prochain Chapitre~\ref{chapter:2}.
-Les essais cliniques et la découverte de traitements ont été historiquement la première application des BMB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33}.
+En d'autres termes, la joueuse ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une des autres actions.
+Ce genre de rétroaction limitée s'appelle l'information de type bandit (\emph{bandit feedback}), et nous discutons de l'histoire et du concept des \emph{bandits} (BMB) en détail dans le prochain Chapitre~\ref{chapter:2}.
+Les essais cliniques et la découverte de traitements ont été historiquement la première application des BMB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33},
 %
-Le BMB est un exemple simple mais puissant du dilemme bien connu entre \emph{exploration et exploitation} :
-lorsqu'elle est confronté à un ensemble de $K$ actions dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
+car les BMB sont un exemple simple mais puissant du dilemme bien connu entre \emph{exploration et exploitation}.
+Lorsqu'il est confronté à un ensemble de $K$ actions dont les effets sur l'environnement sont inconnus, l'apprenant doit trouver un équilibre entre
 explorer les actions inconnues, afin de recueillir plus d'informations à leur sujet,
@@
 %
-Le problème du BMB a été étudié à la fois dans la communauté de l'apprentissage machine et de la statistique, depuis les années $1950$ avec des pionniers comme H.
-Robbins \cite{Robbins52}, et plus récemment c'est un domaine de recherche actif, depuis les années $1990$ \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-La recherche sur le BMB a produit une vaste littérature des années $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet actif, comme l'illustrent les livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et la large gamme des applications des BMB les dernières années \cite{bouneffouf2019survey}.
+Les problèmes de bandits ont été étudiés à la fois dans la communauté de l'apprentissage machine et de la statistique, depuis les années $1950$ avec des pionniers comme H.
+Robbins \cite{Robbins52}, et c'est un domaine de recherche actif depuis les années $1990$ \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
+La recherche sur les bandits a produit une vaste littérature durant les années $2000$ \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} et continue d'être un sujet actif, comme l'illustrent les livres \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019} et les nombreuses applications des bandits les dernières années \cite{bouneffouf2019survey}.
 
@@
 %
-Les deux communautés de l'AOS et du BMB ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des années $2000$ et le début des années $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
+Les deux communautés de l'AOS et des bandits ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des années $2000$ et le début des années $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
 L'accent est mis sur un US accédant à un spectre sous licence, occupé par un UP, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès d'``écoute avant transmission''.
@@
 et considère un ensemble fixe et fini de $K$ canaux orthogonaux, \ie, différentes bandes de fréquences dans un spectre sous licence.
-Par exemple, il peut s'agir d'un ensemble de $K=3$ canaux Wi-Fi à fréquence différente, émis par la même station Wi-Fi.
+Par exemple, il peut s'agir d'un ensemble de $K=3$ canaux Wi-Fi à différentes fréquences, émis par la même station Wi-Fi.
 Une autre hypothèse est que UP et US sont synchronisés dans le temps, en subdivisant le temps en pas de temps discrets.
 %
-Ainsi, si l'US passe peu de temps au début de chaque plage horaire pour essayer de détecter un UP, il peut rechercher la présence ou l'absence d'un UP, avant de transmettre pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
+Ainsi, si l'US passe un peu de temps au début de chaque plage horaire pour essayer de détecter un UP, il peut rechercher la présence ou l'absence d'un UP, avant de transmettre pendant le reste de la plage horaire, sans entrer en collision avec l'UP.
 Si l'US était capable d'analyser tous les $K$ canaux, il pourrait simplement émettre dans l'un des canaux libres s'il y en avait, ou ne pas émettre si tous les canaux sont utilisés à un pas de temps donné.
 %
-Cependant, on sait que la détection est coûteuse en termes de consommation d'énergie, en particulier pour la détection à large bande, comme le montrent les articles \cite{yucek2009survey,subhedar2011spectrum} (\emph{spectrum sensing} en anglais), de sorte que la plupart des travaux sur l'AOS limitent la capacité de détection de l'US à un seul canal à la fois.
-Cette hypothèse, ainsi que l'hypothèse selon laquelle les UP ne peuvent pas être perturbés (sans laquelle aucune régulation ne sera jamais acceptée pour l'AOS), impose à l'US de transmettre dans le canal qu'il a analysé, à chaque pas de temps, si et seulement si il a été détectée sans aucun UP.
+Cependant, on sait que la \emph{détection spectrale} est coûteuse en énergie, en particulier pour la détection à large bande, comme le montrent les articles \cite{yucek2009survey,subhedar2011spectrum} (\emph{spectrum sensing}).
+C'est pourquoi la plupart des travaux sur l'AOS limitent la capacité de détection de l'US à un seul canal à la fois.
+Cette hypothèse, ainsi que l'hypothèse selon laquelle les UP ne peuvent pas être perturbés (sans laquelle aucune réglementation ne sera jamais acceptée pour l'AOS), impose à l'US de transmettre dans le canal qu'il a analysé, si et seulement s'il a été détectée sans aucun UP, et ce à chaque pas de temps.
 % coût non négligeable de la reconfiguration radio
 
-En se concentrant sur un US dans un réseau AOS, il doit décider séquentiellement d'un canal à analyser (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il écoute ce canal et effectue une détection de PU, et enfin il transmet dans ce canal si il a été détecté libre.
-Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou l'équivalent, pour maximiser son nombre de transmissions réussies.
+En se concentrant sur un US dans un réseau AOS, il doit décider séquentiellement d'un canal à analyser (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il écoute ce canal et effectue une détection de PU, et enfin il transmet dans ce canal s'il a été détecté libre.
+Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou de façon équivalente, de maximiser son nombre de transmissions réussies.
 %
-Si les différents canaux ne sont pas uniformément utilisés par les UP, et si l'on suppose une hypothèse stationnaire sur le trafic des UP, alors l'objectif de l'US se résume à explorer les différents canaux et à exploiter les meilleurs.
-Le problème de l'accès opportuniste au spectre est donc un problème d'exploration/exploitation, avec un ensemble fini d'actions (les canaux, également appelés \emph{bras}),
+Si les différents canaux ne sont pas uniformément occupés par les UP, et si l'on suppose une certaine stationnarité sur le trafic des UP, alors l'objectif de l'US se résume à explorer les différents canaux et à exploiter les meilleurs.
+Le problème de l'AOS est donc un problème d'exploration/exploitation, avec un ensemble fini d'actions (les canaux, également appelés \emph{bras}),
 dans un cycle séquentiel d'action-réponse (les pas de temps sont $t\in\N^*$),
 sous l'information de rétroaction partielle (\ie, l'US reçoit l'information concernant un seul canal parmi $K$).
-Ces trois hypothèses sont celles qui limitent le cadre général de l'apprentissage machine au cas spécifique des bandit multi-bras (voir Figure~\ref{fig:1:ReinforcementLearningCycle}).
+Ces trois hypothèses sont celles qui limitent le cadre général de l'apprentissage séquentiel au cas spécifique des bandit multi-bras (voir Figure~\ref{fig:1:ReinforcementLearningCycle}).
 
@@
 %
-Les travaux antérieurs de notre équipe SCEE ont montré que le BMB peut être utilisé pour modéliser le problème de l'AOS :
+Les travaux antérieurs de notre équipe SCEE ont montré que les BMB peuvent être utilisés pour modéliser le problème de l'AOS :
 les bandes de fréquences orthogonales (ou canaux) sont modélisées par des bras $k\in\{1,\dots,K\}$,
-et la récompense obtenue par l'objet équipé de RI après avoir analysé le canal $k$ au temps $t$ est modélisé par une récompense de $r(t) \in \in \{0,1\}$.
-En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors qu'une récompense de $r(t)=0$ indique que le canal $k$ est occupé durant l'interval de temps $t$, et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
+et la récompense obtenue par l'objet équipé de RI après avoir analysé le canal $k$ au temps $t$ est modélisée par une récompense $r(t) \in \{0,1\}$.
+En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors que $r(t)=0$ indique que le canal $k$ est occupé durant l'interval de temps $t$, et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
 %
@@
 Leurs travaux ont été parmi les premiers à proposer l'utilisation de l'apprentissage par renforcement pour la radio intelligente et le problème AOS, notamment le modèle BMB et l'algorithme $\UCB_1$,
-ainsi que les premiers travaux de Q. Zhao et de son équipe, par exemple \cite{Liu08,Zhao10}.
+en parallèle des premiers travaux de Q. Zhao et de son équipe, par exemple \cite{Liu08,Zhao10}.
 %
@@
 Robert \cite{RobertSDR2014,MoyWSR2014,MoyWSR2014} ont développé des preuves de concept utilisant du matériel radio réaliste et la radio logicielle, avec des cartes USRP et le logiciel MATLAB/Simulink.
-Dans une deuxième thèse de doctorat \cite{Modi17PhD}, N.
-Modi a étudié de $2014$ à $2017$ l'impact de l'utilisation des algorithmes BMB pour optimiser la sélection des canaux, sur la durée de vie des batteries d'un objet sans fil.
+Dans une seconde thèse, N.
+Modi a étudié de $2014$ à $2017$ l'impact sur la durée de vie des batteries d'un objet sans fil de l'utilisation des algorithmes BMB pour optimiser la sélection des canaux, \cite{Modi17PhD}.
 D'une part, l'utilisation d'un algorithme BMB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du objet.
@@
 %
-Certaines analyses des mesures radio réelles effectuées pour les canaux ionosphériques HF ont également prouvé que les solutions basées sur l'apprentissage BMB sont appropriées et résolvent efficacement ce type de problèmes de prise de décision sur les signaux sans fil du monde réel \cite{Melian15}.
+Certaines analyses de réelles mesures radio effectuées pour les canaux ionosphériques HF ont également prouvé que les solutions basées sur l'apprentissage BMB sont appropriées et résolvent efficacement ce type de problèmes de prise de décision sur les signaux sans fil du monde réel \cite{Melian15}.
 %
@@
 Par héritage de son travail au sein de l'équipe SCEE,
-certains de leurs travaux récents sont également illustrés par des démos réalistes utilisant USRP et le système MATLAB/Simulink
+certains de leurs travaux récents sont également illustrés par des démonstrations réalistes utilisant USRP et le système MATLAB/Simulink
 \cite{KumarYadav2018,SawantKumar2018,JoshiKumar2018}.
@@
 La littérature susmentionnée a essentiellement montré que les algorithmes BMB peuvent être appliqués avec succès au problème de l'AOS.
-Toutefois, si l'on considère des objets de RI qui ne peuvent pas effectuer de détection active, tels que des objets à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux Internet des Objets, le modèle BMB qui utilise la détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
-En outre, dans la plupart des cas, les réseaux de l'IdO utilisent des bandes sans licence et il n'y a donc plus de distinction entre UP et US.
+Toutefois, si l'on considère des objets de RI qui ne peuvent pas effectuer de détection spectrale active, tels que des objets à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux de l'IdO, le modèle BMB qui utilise cette détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
+En outre, dans la plupart des cas, les réseaux de l'IdO utilisent des bandes non licenciées et il n'y a donc plus de distinction entre UP et US.
 En d'autres termes, il n'y a plus de priorité entre les utilisateurs des réseaux de l'IdO.
-(plus de UP, tous sont US).
+% (plus de UP, tous sont US).
 %
@@
     \item
-    les objets ont des cycles de fonctionnement faibles à très faibles (seulement quelques messages par minute ou par mois).
+    les objets ont des cycles de fonctionnement faibles à très faibles (seulement quelques messages par minute ou par mois), et sont en charge d'initier leurs communications en voie montante (\emph{up-link}).
 \end{itemize}
@@
 %
-De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de la batterie, les $M$ objets sont capables d'utiliser leur antenne reçue pour écouter dans un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), mais ils ne peuvent pas effectuer de détection de spectre à chaque instant, comme cela est envisagé pour l'AOS (\ie, au début de chaque intervalle, dans un schéma ``écoute avant transmission'').
+De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de leur batterie, bien que les objets soient capables d'utiliser leur antenne de réception pour écouter dans un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), ils ne peuvent pas effectuer de détection spectrale à chaque instant, comme cela est envisagé pour l'AOS (\ie, au début de chaque intervalle, dans un schéma de type ``écoute avant de transmettre'').
 
-On pourrait penser qu'en l'absence de détection, l'apprentissage par renforcement n'est plus possible, mais tout objet de l'IdO réel reçoit encore des informations sur son environnement, après certaines ou toutes ses transmissions.
-Dans la plupart des normes IdO, un message (sur la voie montante) envoyé à une station de base peut être suivi d'un message (sur la voie descendante) renvoyé par la station de base pour indiquer si le message a été bien reçu et compris.
+On pourrait penser qu'en l'absence de détection spectrale, l'apprentissage par renforcement n'est plus applicable, mais n'importe quel objet d'un vrai réseau de l'IdO reçoit encore des informations sur son environnement, après certaines ou toutes ses transmissions.
+Dans la plupart des normes IdO, un message (sur la voie montante) envoyé à une station de base doit être suivi d'un message (sur la voie descendante) renvoyé par la station de base pour indiquer si le message a été bien reçu, décodé et compris.
 %
-En utilisant cette rétroaction, qui consiste en un \emph{acquittement} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence de \Ack{} après une transmission échouée, il est possible qu'un terminal IdO puisse également utiliser un algorithme d'apprentissage par renforcement bien conçu, afin d'optimiser ses communications.
+En utilisant cette rétroaction, qui consiste en un \emph{acquittement} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence ' \Ack{} après une transmission échouée, il est possible qu'un objet de l'IdO puisse également optimiser ses communications, grâce à un algorithme d'apprentissage par renforcement bien conçu.
 
@@
 Si nous résumons les problèmes étudiés en une seule question, ce pourrait être la suivante :
-\emph{peut-on adapter les outils de prise de décision déjà appliqués avec succès à la radio intelligente pour l'accès opportuniste au spectre aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des objets ?}
+\emph{peut-on adapter les outils de prise de décision déjà appliqués avec succès à la Radio Intelligente (RI) pour l'Accès Opportuniste au Spectre (AOS) aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des Objets (IdO) ?}
 %
@@
 Nous avons commencé par explorer la riche littérature des bandits multi-bras,
-car il existe de nombreux algorithmes différents, avec de nombreuses variantes du simple problème présenté ci-dessus.
+car il existe de nombreux algorithmes différents, avec de nombreuses variantes du problème présenté ci-dessus.
 Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle BMB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires de BMB.
@@
 Afin de bien comprendre quels algorithmes pourraient être adaptés aux contraintes susmentionnées de la RI pour les réseaux de l'IdO,
-nous n'étions pas seulement intéressés par la mesure habituelle des performances d'un algorithme BMB (son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
+nous ne nous sommes pas seulement intéressés par la mesure habituelle des performances d'un algorithme BMB (\ie, son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
 mais aussi par leurs performances empiriques en termes de complexité de calcul et de stockage, tant du point de vue des analyses théoriques que des mesures réelle.
@@
 Notre exploration du grand nombre d'algorithmes et de modèles BMB développés dans la littérature récente
-nous a donné l'ambition d'écrire un seul logiciel permettant à chacun d'implémenter facilement de nouveaux modèles et algorithmes, afin de comparer les modèles existants et d'explorer empiriquement les performances des nouveaux algorithmes proposés.
-Pour atteindre cet objectif, nous avons développé une bibliothèque Python de simulation des problèmes BMB.
+nous a donné l'ambition d'écrire un seul logiciel permettant à n'importe quel-le chercheur-euse d'implémenter facilement de nouveaux modèles et algorithmes, afin de comparer les modèles existants et d'explorer empiriquement les performances de nouveaux algorithmes.
+Pour atteindre cet objectif, nous avons développé une bibliothèque de simulation des problèmes BMB, dans le langage populaire Python.
 %
@@
 nous avons écrit la bibliothèque libre de simulation la plus exhaustive pour les problèmes BMB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
-Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, et différents exemples de son utilisation.
-Une documentation complète est disponible en ligne, ainsi que des instructions pour reproduire les simulations discutées dans la suite de cette thèse.
+Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}.
+%  et différents exemples de son utilisation.
+Une documentation complète est disponible en ligne, ainsi que des instructions pour reproduire les simulations présentées dans la suite de cette thèse.
 
@@
 %
-Étant donné le grand nombre d'algorithmes BMB disponibles, nous nous intéressons aussi à la question de savoir comment une ingénieure peut choisir l'algorithme à mettre en œuvre, dans un objet de l'IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à tout environnement.
+Étant donné le grand nombre d'algorithmes BMB disponibles, nous nous intéressons aussi à la question de savoir comment un-e ingénieur-e peut choisir l'algorithme à mettre en œuvre, dans un objet de l'IdO donné, afin de doter cet objet de la capacité de s'adapter de manière robuste à son futur environnement.
 Pour répondre à cette question, nous présentons deux approches.
-La première est une comparaison empirique d'algorithmes existants, se focalisant sur les plus efficaces et les mieux connus, dans Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
+La première est une comparaison empirique d'algorithmes existants, se focalisant sur les plus efficaces et les mieux connus, en Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
 Nous confirmons que les algorithmes largement utilisés, tels que \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} et \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
-La seconde approche est une sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons en Chapitre~\ref{chapter:25}.
+La seconde approche est la sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons en Chapitre~\ref{chapter:25}.
 Ce travail sur l'agrégation des algorithmes BMB a été motivé pour le cas AOS de la radio cognitive, où de nombreux travaux antérieurs ne considéraient que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
@@
 
-Dans la deuxième Partie~\ref{part:MABIOT} de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux de l'IdO, avec des simulations et une validation de concept réelle, puis nous étudions des questions intéressantes sur deux modèles mathématiques du BMB, issues de la simplification de notre premier modèle.
+Dans la deuxième partie de cette thèse, nous commençons par proposer et étudier différents modèles de réseaux de l'IdO, avec des simulations et un prototype réel (\ie, une preuve de concept), puis nous étudions des questions intéressantes sur deux modèles mathématiques de bandits, issues de la simplification de notre premier modèle.
 
@@
 De tels objets utilisent les acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
-L'objectif de ce premier modèle est d'améliorer la qualité de service (QdS, ou \emph{Quality of Service} QoS) de l'application de ce réseau de l'IdO, en appliquant une RL décentralisée côté objet afin de réduire le taux d'échec de transmission.
-La station de base recevra donc plus de paquets de liaison montante des objets desservis, si ils peuvent apprendre par eux-même un accès efficace au spectre.
-Cela implique que le réseau peut accepter plus d'objets tout en maintenant la même QdS.
+L'objectif de ce premier modèle est d'améliorer la qualité de service (QdS, ou \emph{Quality of Service} QoS) de l'application de ce réseau de l'IdO, en appliquant un ApR décentralisé côté objet afin de réduire le taux d'échec de transmission.
+La station de base recevra donc plus de paquets de liaison montante des objets desservis, s'ils peuvent apprendre par eux-même un accès efficace au spectre.
+Cela implique aussi que le réseau peut accepter plus d'objets tout en maintenant la même QdS.
 %
-Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux concepts soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet dans ce cas d'augmenter la durée de vie des batteries de chacun des objets, ainsi que la QdS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
+Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux modèles soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet dans ce cas d'augmenter la durée de vie des batteries de chacun des objets, ainsi que la QdS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
 
-Le premier modèle est issu du premier article écrit au cours de cette thèse, qui a initié une collaboration avec R. Bonnefoi, un autre doctorant de notre équipe SCEE.
+Le premier modèle sans retransmission est issu du premier article écrit au cours de cette thèse, qui a initié une collaboration avec R. Bonnefoi, un autre doctorant de notre équipe SCEE.
 Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre 2010, où il a obtenu le ``prix du meilleur papier'' \cite{Bonnefoi17}.
 %
-La preuve de concept (PoC) a continué notre collaboration fructueuse, et a été démontrée pendant trois jours à la conférence IEEE ICT à Saint-Malo, France, en juin $2018$ \cite{Besson2018ICT}, et ensuite présentée à la conférence IEEE WCNC à Marrakech, Maroc, en avril $2019$ \cite{Besson2019WCNC}.
+La preuve de concept (PdC) a continué notre collaboration fructueuse, et a été montrée pendant trois jours à la conférence IEEE ICT à Saint-Malo, France, en juin $2018$ \cite{Besson2018ICT}, et aussi à la conférence IEEE WCNC à Marrakech, Maroc, en avril $2019$ \cite{Besson2019WCNC}.
 %
@@
 %
-Dans les deux modèles de réseaux de l'IdO présentés ci-dessus, le concept est de laisser chaque objet dynamique d'un tel réseau exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
-Il est important de noter que chaque objet ne communique, et donc n'apprend, qu'à quelques tours et non à toutes les étapes, en suivant son propre processus d'activation aléatoire (nous nous limitons à un processus d'activation de Bernoulli purement aléatoire de probabilité $p$).
-Cela signifie que chacun d'eux vise son propre objectif local, qui est de maximiser sa récompense cumulée, d'une manière égoïste et complètement agnostique des autres objets ayant le même objectif.
-Il est bien connu dans la littérature sur la théorie des jeux que jouer égoïstement peut être désastreux pour la mesure centralisée de la performance, on peut penser aux ``dilemmes'' populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
+Dans les deux modèles de réseaux de l'IdO présentés ci-dessus, l'idée maîtresse est de laisser chaque objet dynamique d'un tel réseau exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
+Il est important de noter que chaque objet ne communique, et donc n'apprend, qu'à quelques unes et non à toutes les étapes, en suivant son propre processus d'activation aléatoire (nous nous limitons à un processus d'activation purement aléatoire, Bernouli de probabilité $p$).
+Cela signifie que chaque objet vise son propre objectif local, qui est de maximiser sa récompense cumulée, d'une manière égoïste et complètement agnostique des autres objets ayant le même objectif.
+Il est bien connu en théorie des jeux que jouer égoïstement peut être désastreux pour la mesure centralisée de la performance, on peut penser aux ``dilemmes'' populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
 Il était donc assez surprenant que nos simulations numériques, ainsi que la preuve de concept réaliste, aient montré que l'apprentissage BMB décentralisé et égoïste conduisait à une coordination efficace entre les objets, dans tous les scénarios considérés, malgré le fait que ces objets ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
 
-C'est pourquoi nous avons d'abord essayé d'analyser la performance de cet algorithme décentralisé, que nous appelons \Selfish, dans le modèle de la Section~\ref{sec:4:firstModel},
+Nous avons donc d'abord essayé d'analyser la performance de cet algorithme décentralisé, que nous appelons \Selfish, dans le modèle de la Section~\ref{sec:4:firstModel},
 mais en raison du nombre aléatoire d'objets actifs à chaque pas de temps (\ie, dès que la probabilité $p$ d'activation est $p < 1$), nous n'avons pas pu développer une analyse propre.
 %
-C'est la raison pour laquelle nous affaiblissons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$ dans le Chapitre~\ref{chapter:5}, qui est équivalent à avoir $p < 1$, et nous considérons le cas des objets actifs à chaque étape de temps (\ie, $p=1$), et nous limitons donc à $M \leq K$ objets.
+C'est la raison pour laquelle nous affaiblissons l'hypothèse d'avoir $M \gg K$ (ou même simplement $M > K$) dans un réseau avec un canal sans fil orthogonal $K$ dans le Chapitre~\ref{chapter:5}, qui est équivalent à avoir $p < 1$, et nous considérons le cas d'objets actifs à chaque étape de temps (\ie, $p=1$), et nous nous limitons donc à $M \leq K$ objets.
 Le modèle que nous avons étudié comporte différentes variantes, selon le niveau de rétroaction, et couvre à la fois le cas AOS (\ie, avec détection des UP) ou le cas IdO (\ie, sans détection).
-L'objectif était de comprendre l'heuristique du Chapitre~\ref{chapter:4}, \Selfish, dans le cadre plus simple du BMB multi-joueur, qui a été étudié auparavant pour le cas de l'AOS, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
+L'objectif était de comprendre l'heuristique du Chapitre~\ref{chapter:4}, \Selfish, dans le cadre plus simple des BMB multi-joueurs, qui a été étudié auparavant pour le cas de l'AOS, comme étudié par exemple par \cite{Zhao10,Anandkumar10,Anandkumar11}.
 %
@@
 car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
-Il est en fait équivalent aux modèles précédemment étudiés d'application du BMB pour l'AOS, puisqu'ils ne prennent en compte que des objets synchronisés, un US et des UP, et donc si la détection indique qu'un canal est libre à un intervalle de temps, le message de liaison montante envoyé par l'US sera certainement reçu avec succès par la station de base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin de \Ack.
+Il est en fait équivalent aux modèles précédemment étudiés d'applications des BMB pour l'AOS, puisqu'ils ne prennent en compte que des objets synchronisés, un US et des UP, et donc si la détection indique qu'un canal est libre pour un intervalle de temps, le message envoyé par l'US sera certainement reçu avec succès par la station de base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin d'\Ack.
 
-D'une part, dans le cadre de l'AOS, nous n'avons pas réussi à obtenir un résultat positif pour l'algorithme \Selfish, car nous avons prouvé que dans certains contextes limités (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre d'un regret moyen linéaire.
-Ce résultat a ensuite été confirmé et analysé par \cite{BoursierPerchet18} (Annexe~F).
+D'une part, dans le cadre de l'AOS, nous n'avons pas réussi à obtenir un résultat positif pour l'algorithme \Selfish, car nous avons prouvé que dans certains petits problèmes (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre d'un regret moyen linéaire.
+Ce résultat a ensuite été confirmé et analysé par d'autres chercheurs dans \cite{BoursierPerchet18} (Annexe~F).
 %
-D'autre part, nous avons été en mesure de proposer de nouveaux algorithmes pour ce problème de bandit multi-joueurs (avec informations de détection), et nous avons analysé notre proposition \MCTopM-\klUCB, pour montrer qu'elle atteint une limite supérieure de regret logarithmique à temps fini, améliorant considérablement les résultats précédents.
+D'autre part, nous avons été en mesure de proposer de nouveaux algorithmes pour ce problème de bandit multi-joueurs (avec informations de détection), et nous avons analysé notre proposition, appelée \MCTopM, pour montrer qu'elle atteint une borne supérieure de regret logarithmique à temps fini, qui améliore considérablement les résultats précédents.
 Notre algorithme atteint également un nombre logarithmique de collisions et de changements de bras, et permet à un groupe fixe de $M$ objets d'apprendre efficacement à utiliser les $M$ meilleurs canaux orthogonalement pour presque toutes leurs communications montantes.
 %
-Ce fort résultat théorique dépend fortement de la présence d'une information de détection et, par conséquent, nos résultats ne sont pas (encore) applicables au modèle IdO sans détection.
+Ce fort résultat théorique dépend fortement de la présence d'une information de détection spectrale et, par conséquent, nos résultats ne sont pas (encore) applicables au modèle IdO sans détection.
 
-Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux objets, tous suivant des modèles d'activation aléatoires.
-La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes BMB, conçus pour résoudre des problèmes stationnaires, sur un problème non stationnaire,
-mais plutôt que nous analysons des algorithmes qui jouent tous dans différents d'instants d'activations (aléatoires), sous-ensembles des pas de temps synchronisés globaux.
+Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux objets, chacun suivant des modèles d'activation aléatoires.
+La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes BMB, conçus pour résoudre des problèmes stationnaires, appliqués à un problème non stationnaire,
+mais plutôt que nous considérons des algorithmes qui jouent tous à différents instants d'activations (aléatoires), ce qui donne des sous-ensembles différents et imprédictibles des pas de temps synchronisés globaux.
 Généraliser à différentes probabilités d'activation conduirait à un modèle encore plus difficile à analyser, et imposer un maximum de $M \leq K$ objets n'est en fait pas vraiment réaliste pour les réseaux de l'IdO, même si cela conduit au modèle du Chapitre~\ref{chapter:5}.
@@
 Cependant, même si nous continuons à supposer un temps synchronisé dans tout le reste de la thèse,
-il est difficile de soutenir que cette hypothèse d'un calendrier centralisé pour les objets peut être réaliste, car nous étudions le cas de l'apprentissage décentralisé pour l'AOS et l'IdO précisément afin d'éviter la signalisation et tout contrôle central des objets par la station de base, comme nous l'avons expliqué ci-dessus.
+il est difficile de soutenir que cette hypothèse d'un calendrier centralisé pour les objets peut être réaliste, car nous étudions le cas de l'apprentissage décentralisé pour l'AOS et l'IdO précisément afin d'éviter un surcoût dû à une signalisation régulière et un contrôle central des objets par la station de base, comme nous l'avons expliqué précedemment.
 
-D'autre part, une autre façon d'affaiblir l'hypothèse non stationnaire est de prendre le point de vue d'un seul objet, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur un US entouré de plusieurs UP ayant des comportements fixes.
+D'autre part, une autre façon d'affaiblir l'hypothèse de non stationnarité est de prendre le point de vue d'un seul objet, comme dans le cas de l'AOS mentionné ci-dessus, où l'accent est mis sur un US entouré de plusieurs UP ayant des comportements stochastiques et stationnaires.
 Si nous nous concentrons sur un objet de l'IdO, son environnement (\ie, les autres objets) est non stationnaire, ce qui signifie que ses propriétés moyennes peuvent fluctuer dans le temps, mais les environnements réels présentent généralement des non-stationnarité avec une certaine structure.
-Par exemple, l'hypothèses de stationnarité par intervalles de temps conduit à la dernière contribution et au dernier chapitre de cette thèse.
+C'est pourquoi nous considérons l'hypothèse de stationnarité par intervalles de temps, dans la dernière contribution et le dernier chapitre de cette thèse.
 
@@
 %
-Afin de bien comprendre également comment les algorithmes BMB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur les BMB adverses ou non stationnaires, pour les deux cas d'environnements variant lentement et changeant brusquement.
+Afin de bien comprendre également comment les algorithmes BMB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur les BMB adverses ou non stationnaires, pour les deux cas d'environnements variant lentement ou changeant brusquement.
 Nous commençons notre dernier Chapitre~\ref{chapter:6} en passant en revue les travaux existants,
@@
 ce qui signifie que le problème de bandit sous-jacent est stationnaire sur certains intervalles, séparés par des points de changement situés à des moments inconnus.
-Supposer que l'environnement est stationnaire par morceaux est en effet logique pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'objets dans le réseau.
-Des exemples d'une telle situation peuvent être une nouvelle entreprise arrivant sur le marché dans une ville (par exemple, les comptoirs Linky qui sont installés aujourd'hui en France), ou l'agriculteur voisin installant des capteurs sur son propre troupeau de vaches, etc.
+Supposer que l'environnement est stationnaire par morceaux est en effet cohéret pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'objets dans le réseau.
+Des exemples d'une telle situation peuvent être une nouvelle entreprise arrivant sur le marché dans une ville (par exemple, les compteurs Linky qui sont installés aujourd'hui en France), ou l'agriculteur voisin installant des capteurs sur son propre troupeau de vaches, etc.
 %
-Deux grandes familles d'algorithmes ont été proposées pour le problème stationnaire par morceaux,
-qui consistent généralement à combiner un algorithme efficace conçue pour le problème stationnaire du BMB (\eg, Thompson Sampling ou \klUCB) et un moyen de s'adapter aux changements dans la distribution des bras.
-Les algorithmes passivement adaptatives utilisent une fenêtre de taille fixe ou évolutive \cite{Garivier11UCBDiscount}, ou un facteur d'actualisation, pour oublier les observations passées \cite{Kocsis06,Gupta11thompson},
+Deux grandes familles d'algorithmes ont été proposées pour les problèmes stationnaires par morceaux,
+qui consistent généralement à combiner un algorithme efficace conçu pour le problème des bandits stationnaires (\eg, Thompson Sampling ou \klUCB), avec un moyen de s'adapter aux changements dans les distributions des bras.
+Les algorithmes passivement adaptatives utilisent une fenêtre de taille fixe ou évolutive \cite{Garivier11UCBDiscount}, ou un facteur d'oubli, pour oublier les observations passées \cite{Kocsis06,Gupta11thompson},
 tandis que les algorithmes activement adaptatifs utilisent un test statistique pour détecter les points de changement \cite{MellorShapiro13,Allesiardo15}.
@@
 
-Suite à deux travaux précédents récents \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, nous combinons une politique d'index efficace (\klUCB) et un test efficace de détection des points de changement, dans l'hypothèse de récompenses limitées.
-Nous nous appuyons sur des résultats très récents du test du rapport de vraisemblance généralisé (TRGL) pour les variables gaussiennes et sous-gaussiennes \cite{Maillard2018GLR}, mais nous nous concentrons plutôt sur les récompenses limitées et les distributions de Bernoulli.
-Les récompenses limitées sont en effet généralement plus appropriées pour les applications de radio cognitive, comme indiqué ci-dessus.
-En utilisant le fait que les variables bornées en $[0,1]$ ne sont pas seulement sous-Gaussiennes à $1/4$ mais aussi sous-Bernoulli, nous prouvons les premières garanties à temps fini pour le GLRT pour les variables bornées.
-Nous montrons d'abord des limites temporelles finies à la fois sur la probabilité de fausse alarme de notre test, et sur son délai de détection, sous de légères hypothèses sur la longueur des séquences stationnaires.
-Notre algorithme, noté \GLRklUCB, est alors présenté en deux variantes, que les points de changement soient locaux (\ie, une seule moyenne de bras change à chaque point de changement) ou globaux (\ie, peut-être tous les moyens de bras changent à la fois).
+Suite à deux travaux précédents récents \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, nous combinons une stratégie d'indices efficace (\klUCB) et un test efficace de détection des points de changement, dans l'hypothèse de récompenses bornées.
+Nous nous appuyons sur des résultats très récents du Test du Rapport de Vraisemblance Généralisé (\emph{Generalized Likelihood Ratio Test}, GLRT) pour les variables gaussiennes et sous-gaussiennes \cite{Maillard2018GLR}, mais nous nous concentrons plutôt sur les récompenses bornées et les distributions de Bernoulli.
+Les récompenses bornées sont en effet généralement plus appropriées pour les applications de RI, comme nous l'avons discuté ci-dessus.
+En utilisant le fait que les variables bornées dans $[0,1]$ ne sont pas seulement $1/4$ sous-Gaussiennes mais aussi sous-Bernoulli, nous prouvons les premières garanties à temps fini pour le GLRT pour des variables bornées.
+Nous montrons d'abord des bornes à temps finis à la fois sur la probabilité de fausse alarme de notre test, et sur son délai de détection, sous des hypothèses raisonnables sur la longueur des séquences stationnaires.
+Notre algorithme est alors présenté en deux variantes, selon que les points de changement soient locaux (\ie, une seule moyenne de bras change à chaque point de changement) ou globaux (\ie, peut-être que toutes les moyennes des bras changent à la fois).
 %
-Nous prouvons qu'en combinant l'algorithme \klUCB, asymptotiquement optimale pour les problèmes stationnaires, et notre nouvelle analyse du GLRT pour les récompenses limitées, nous obtenons des garanties de pointe sur le regret de notre algorithme proposé.
-Nous considérons la même hypothèse que nos concurrents, en supposant que la longueur des intervalles stationnaires est " assez longue " pour que les points de changement soient détectables.
-La limite supérieure du meilleur regret est obtenue lorsque l'algorithme connaît à l'avance l'horizon et le nombre de points de changement, mais notre algorithme n'a pas besoin d'avoir d'autres connaissances sur la difficulté du problème pour être accordé de manière optimale.
+Nous prouvons qu'en combinant l'algorithme \klUCB, asymptotiquement optimale pour les problèmes stationnaires, et notre nouvelle analyse du GLRT pour les récompenses bornées, nous obtenons des garanties de pointe sur le regret de notre algorithme proposé, noté \GLRklUCB.
+Nous considérons la même hypothèse que nos concurrents, en supposant que la longueur des intervalles stationnaires est ``assez longue'' pour que les points de changement soient détectables.
+La meilleure borne supérieure de regret est obtenue lorsque l'algorithme connaît à l'avance l'horizon et le nombre de points de changement, mais notre algorithme n'a pas besoin d'avoir d'autres connaissances sur la difficulté du problème pour être paramétré de manière optimale.
 %
-La performance de \GLRklUCB{} est également illustrée par des expériences numériques sur des données synthétiques, où il est démontré qu'elle surpasse toutes les algorithms passivement adaptatifs ainsi que les précédentes algorithms activement adaptatifs.
+La performance de \GLRklUCB{} est également illustrée par des expériences numériques sur des données synthétiques, où il est démontré qu'il surpasse tous les algorithmes passivement adaptatifs ainsi que les précédents algorithmes activement adaptatifs.
 %
@@
 Ce manuscrit se termine par
-une liste d'abréviations et de notations, puis des listes de figures, d'algorithmes, d'échantillons de code et de tableaux,
-et enfin la liste des références bibliographiques.
+une liste d'abréviations et de notations, puis de figures, d'algorithmes, de morceaux de code et de tableaux,
+et enfin des références bibliographiques.
 
@@
 Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
-Elles se situent à trois niveaux : des modèles BMB pour plusieurs contextes radio liés à la radio intelligente ; des preuves théoriques associées aux problèmes d'apprentissages correspond ; et enfin des implémentations d'algorithmes à but de simulations, de partage avec la communauté, et de validation par des démonstrations radio réalistes.
+Elles se situent à trois niveaux : des modèles BMB pour plusieurs contextes radio liés à la radio intelligente et l'IdO ; des preuves théoriques associées aux problèmes d'apprentissages correspondants ; et enfin des implémentations d'algorithmes à but de simulations, de partage avec la communauté, et de validation par des démonstrations radio réalistes.
 
@@
     \item
-    Nous avons écrit la plus complète bibliothèque de simulation pour les problèmes BMB, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBanditsJMLR,SMPyBandits}.
+    Nous avons écrit la bibliothèque de simulation pour les problèmes BMB la plus complète, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBanditsJMLR,SMPyBandits}.
     Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
-    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences et simulations numériques utilisées tout au long de thèse.
+    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les simulations numériques utilisées tout au long de ce manuscrit de thèse.
 
@@
     Nous présentons le problème du choix de l'algorithme qu'un-e ingénieur-e devrait utiliser, ou de la sélection d'algorithme parmi la riche collection des différents algorithmes de bandits dans la litérature, dans le Chapitre~\ref{chapter:25}.
-    Nous présentons un algorithme, appelé \Aggr, pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par des simulations numériques qu'il peut atteindre de bonnes performances empiriques
-    (publié à WCNC 2018) \cite{Besson2018WCNC}.
+    Nous présentons un algorithme d'agrégation d'algorithmes, appelé \Aggr, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par des simulations numériques qu'il peut atteindre de bonnes performances empiriques
+    (\cite{Besson2018WCNC}, publié à WCNC 2018).
 
     \item
-    Nous proposons différents modèles pour les réseaux de l'IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les objets dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes BMB, pour augmenter automatiquement la durée de vie de leur batterie.
+    Nous proposons différents modèles pour les réseaux de l'Internet des Objets (IdO), dans le Chapitre~\ref{chapter:4}, où les objets dotés de capacités de radio intelligente peuvent mettre en œuvre de leur côté des algorithmes BMB, pour augmenter automatiquement la durée de vie de leur batterie.
     Cela permet également à davantage d'objets d'utiliser le même réseau tout en maintenant un taux élevé d'accès au spectre sans souffrir de collisions radio
-    (publiés à CROWNCOM 2017, démonstration ICT 2018, WCNC 2019 et MOTIoN 2019)
-    \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.
+    (publiés à CROWNCOM 2017 \cite{Bonnefoi17}, démonstration ICT 2018 \cite{Besson2018ICT}, WCNC 2019 \cite{Besson2019WCNC} et MOTIoN 2019 \cite{Bonnefoi2019WCNC}).
 
@@
     Nous avons réalisé une démonstration réaliste du modèle proposé ci-dessus \cite{Besson2018ICT,Besson2019WCNC}, présentée à ICT 2018, et nous la détaillons en Section~\ref{sec:4:gnuradio}.
-Nous avons aussi réalisé une courte vidéo présentant notre démonstration, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+Nous avons aussi réalisé une vidéo de présentation de $6$ minutes, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
 
@@
     Nous formalisons le modèle du bandit multi-joueur, pour lequel nous avons introduit deux variantes, dans le Chapitre~\ref{chapter:5}.
-    Pour le cas avec information de détection (``sensing''), nous proposons deux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM, qui prouve qu'il est asymptotiquement très efficace.
+    Pour le cas avec information de détection spectrale (``spectral sensing''), nous proposons deux algorithmes, et nous analysons notre algorithme \MCTopM, pour prouver qu'il est très efficace, à temps fini et asymptotiquement.
     Nous présentons aussi des expériences numériques approfondies pour montrer qu'il est bien plus efficace que les autres algorithmes de la littérature.
-    Notre travail \cite{Besson2018ALT}, publié à ALT 2018, a également contribué à une nouvelle impulsion à la recherche sur les bandits multi-joueurs, car certains travaux de recherche récents se sont construits sur nos résultats.
+    Notre travail \cite{Besson2018ALT}, publié à ALT 2018, a également contribué à un nouvel élan à la recherche sur les bandits multi-joueurs, certains travaux de recherche récents s'étant construits sur nos résultats.
 
     \item
-    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle BMB multi-joueurs, particulièrement active ces deux dernières années.
+    Nous donnons une revue détaillée de la littérature sur les différentes extensions du modèle BMB multi-joueurs, particulièrement active ces deux dernières années.
     % qui, selon nous, n'a jamais été écrit auparavant.
@@
     \item
-    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le Chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (publié à GRETSI 2019).
-    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles .
+    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le Chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle (\cite{Besson2019Gretsi} publié à GRETSI 2019, \cite{Besson2019GLRT}).
+    Nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles car notre approche n'a pas besoin de connaîre de bornes sur la difficulté du problème à part le nombre de points de ruptures.
     % Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB
@@
 L'ordre de lecture du manuscrit peut suivre n'importe quel chemin, entre l'introduction donnée dans le Chapitre~\ref{chapter:1}, et la conclusion générale qui constitue le dernier Chapitre~\ref{chapter:conclusion}.
-Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,
+Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessous,
 la thèse est organisée en deux parties, correspondant aux deux lignes intermédiaires de la figure suivante.
@@
 Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
-Le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, qui est utilisées par les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} pour leurs simulations numériques.
+Le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, qui est utilisée par les Chapitres~\ref{chapter:2}, \ref{chapter:25}, \ref{chapter:5} et \ref{chapter:6} pour leurs simulations numériques.
 %  lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.
@@
 \textcolor{darkgreen}{La deuxième Partie~\ref{part:MABIOT}} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus quasiment indépendamment.
-Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux de l'IdO, dans lesquels les algorithmes BMB ont été utilisés avec succès.
+Le Chapitre~\ref{chapter:4} commence par proposer et étudier différents modèles de réseaux de l'IdO, pour lesquels nous montrons que les algorithmes BMB peuvent être utilisés avec succès.
 Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop complexes pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
@@
 Les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient chacun un modèle intermédiaire, situé entre le modèle BMB stationnaire à un joueur du Chapitre~\ref{chapter:2} et les modèles de réseaux de l'IdO du Chapitre~\ref{chapter:4}.
-Ces deux études ont chacune donné des résultats théoriques à la pointe de la recherche sur les modèles de bandits multi-joueurs et de bandits stationnaires par morceaux, que nous avons aussi validé par des expériences numériques.
+Ces deux études ont chacune donné des résultats théoriques à la pointe de la recherche, sur les deux modèles de bandits multi-joueurs et de bandits stationnaires par morceaux, que nous avons aussi validé par des expériences numériques.
 

### 2019-07-08 2-Chapters/1-Chapter/chapter1.tex
Okay j'ai relu l'intro en français

@@
 In the last $150$ years, mankind has developed many different communication technologies, and since the late $1890$s, wireless communications between manufactured devices have been made possible, and more and more frequent in our lives.
-With the avent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, allowing a large range of different applications.
+With the advent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, allowing a large range of different applications.
 It is now a world-wide consensus that with the current trend of population increase and with the ongoing energy crisis, any newly deployed technology should be both cheap and energy efficient,
@@
 and the \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}) layer.
-To quote some examples of IdO standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications (up-to \SI{2}{\meter}), while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications (up-to \SI{50}{\kilo\meter}).
+To quote some examples of IoT standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications (up-to \SI{2}{\meter}), while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications (up-to \SI{50}{\kilo\meter}).
 We refer to the survey \cite{Centenaro16} for more details, and \cite{Azari18} or our recent article \cite{MoyBesson2019}, and references therein.
@@
 almost the entire spectrum
-is allocated to various usages, that goes from maritime radionavigation (historically the first usage of radio telecommunication, \eg, in the Titanic), to space research, inter-satellite, mobile telephony and many other applications.
+is allocated to various usages, that goes from maritime radio-navigation (historically the first usage of radio telecommunication, \eg, in the Titanic), to space research, inter-satellite, mobile telephony and many other applications.
 %
@@
 Moreover, the fixed spectrum allocation prevents
-the introduction of new services, especially for low-cost or small markets equipments.
+the introduction of new services, especially for low-cost or small markets equipment.
 % rarely used frequencies, like those assigned to specific services, from being used, even when any unlicensed users would not cause noticeable interference to the assigned service.
@@
 The goal of the player is to maximize its rewards, by trials and errors (\ie, actions and rewards).
-Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk or drive, to play a board or computer game, or discovering which treatment is efficient in healing a certain disease (clinical trial), etc.
+Many real-world problems can be framed as Reinforcement Learning problems, as illustrated by the survey \cite{bouneffouf2019survey} and Section~\ref{sec:2:applicationsofStochasticMAB}, for instance learning to walk, to drive, to play a board or computer game, discovering which treatment is efficient in healing a certain disease (clinical trial), etc.
 
@@
 
-Focussing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of channels, $[K]=\{1,\dots,K\}$), then it performs sensing, and finally it transmits in this channel if it was sensed free.
-The goal of the SU is to minimize its energy consumption (we remind that we focus on ``green'' cognitive radio) and to maximize its uplink data rate, or equivalently, to maximize its number of successful transmission.
+Focusing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of channels, $[K]=\{1,\dots,K\}$), then it performs sensing, and finally it transmits in this channel if it was sensed free.
+The goal of the SU is to minimize its energy consumption (we remind that we focus on ``green'' cognitive radio) and to maximize its up-link data rate, or equivalently, to maximize its number of successful transmission.
 %
@@
 
-\paragraph{MAB for OSA, and a short history of these researchs at the SCEE team.}
+\paragraph{MAB for OSA, and a short history of these researches at the SCEE team.}
 %
@@
 Robert \cite{RobertSDR2014,MoyWSR2014}, using USRP boards and the MATLAB/Simulink software.
-In a second PhD thesis \cite{Modi17PhD}, N. Modi studied from $2014$ to $2017$ the impact of using MAB algorithms to optimize channel selections, on the battery life of a wireless device.
+In a second PhD thesis, N. Modi studied from $2014$ to $2017$ the impact on the battery life of a wireless device of using MAB algorithms to optimize channel selections \cite{Modi17PhD}.
 On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and could bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
@@
 Each hardware reconfiguration costs energy for the device, and so quickly-switching algorithms lead to a reduction of the battery life.
-The tradeoff between the two aspects is studied empirically in
+The trade-off between the two aspects is studied empirically in
 \cite{modiDemo2016}.
@@
     \item
-    IoT devices have low to very low duty cycles (few messages every minute to only a few every month).
+    IoT devices have low to very low duty cycles (few messages every minute to only a few every month), and are in charge of initiating their up-link communications.
 \end{itemize}
@@
 We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
-A full documentation is available online, as well as exhaustive instructions to reproduce the simulations discussed in the rest of this thesis.
+A full documentation is available online, as well as exhaustive instructions to reproduce the simulations presented in the rest of this thesis.
 
@@
 The first model considers IoT devices that simply have data to regularly send to a base station (at unpredictable or random times).
-Such devices use acknowledgements as feedback, in order to optimize their uplink communications, by accessing more the best channels (\ie, the channel less occupied by the surrounding traffic).
+Such devices use acknowledgements as feedback, in order to optimize their up-link communications, by accessing more the best channels (\ie, the channel less occupied by the surrounding traffic).
 The goal of this first model is to increase the Quality of Service (QoS) of the application of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmission.
@@
 On the one hand, in the OSA setting, we failed to obtain positive result for the \Selfish{} policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
-This result was later confirmed and analyzed by \cite{BoursierPerchet18} (Appendix~F).
+This result was later confirmed and analyzed by other researchers in \cite{BoursierPerchet18} (Appendix~F).
 %
@@
     We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non-stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
-    Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
+    We propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance, while requiring no prior knowledge on the problem difficulty other than the number of break-points.
     % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB
@@
     % \footnote{~CROWNCOM 2017, see \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}},
-    Lisboa, Portugal, Septembre $2017$,
+    Lisboa, Portugal, September $2017$,
     % \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},

### 2019-07-08 2-Chapters/2-Chapter/chapter2.tex
Okay j'ai relu l'intro en français

@@
 %
-Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{} defined in \eqref{eq:2:UCB_index} respectively represent \emph{exploitation} and \emph{exploration}, and summing them up is a natural way of dealing with the exploitation/exploration trade off.
+Furthermore, these two terms $\widehat{\mu_k}(t)$ and $\xi_k(t)$ in the \UCB{} defined in \eqref{eq:2:UCB_index} respectively represent \emph{exploitation} and \emph{exploration}, and summing them up is a natural way of dealing with the exploitation/exploration trade-off.
 The constant parameter $\alpha$ in $\xi_k(t)$ also controls this trade-off, and the theoretical analysis suggests to restrict to $\alpha\geq1/2$, and to choose $\alpha=1/2$ for optimal uniform performance (\ie, on all problems, \cite{Auer02}).

### 2019-07-08 2-Chapters/3-Chapter/chapter3.tex
Okay j'ai relu l'intro en français

@@
 %
-Her goal is to maximize $\sum_{t=1}^T r(t)$, which require to trade off between exploring the unknown $K$ arms, exploiting the arm which is currently estimated to be the best one.
+Her goal is to maximize $\sum_{t=1}^T r(t)$, which require to trade-off between exploring the unknown $K$ arms, exploiting the arm which is currently estimated to be the best one.
 

### 2019-07-08 1-Preamble/5-Abstract/resume_fr.tex
LOTS of works…

@@
 
-Pour le développement futur des ``réseaux intelligents'' (\emph{smart grid}), des ``villes intelligentes'' (\emph{smart cities}), des ``maisons intelligentes'' (\emph{smart homes}) ou de l'``agriculture intelligente'' (\emph{smart agriculture}), des réseaux de capteurs devraient être largement déployés.
-Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les ``bâtiments connectés'' et l'``agriculture connectée''.
+Pour le développement futur des \guillemotleft{} réseaux intelligents \guillemotright{} (\emph{smart grid}), des \guillemotleft{} villes intelligentes \guillemotright{} (\emph{smart cities}), des \guillemotleft{} maisons intelligentes \guillemotright{} (\emph{smart homes}) ou de l'\guillemotleft{} agriculture intelligente \guillemotright{} (\emph{smart agriculture}), des réseaux de capteurs devraient être largement déployés.
+Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les \guillemotleft{} bâtiments connectés \guillemotright{} et l'\guillemotleft{} agriculture connectée \guillemotright{}.
 Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides en utilisant des réseaux de capteurs afin d'obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, pour permettre au contrôle centralisé du chauffage d'optimiser son coût et sa consommation énergétique.
@@
     J. Mittola, en $1999$, a proposé que
-    \emph{``une radio vraiment intelligente qui serait autonome, sensible aux RF et à l'utilisateur, et qui inclurait la technologie logicielle et les capacités d'apprentissage machine ainsi qu'une grande connaissance de l'environnement radio haute-fidélité''} \cite{Mitola99}.
+    \guillemotleft{} \emph{une radio vraiment intelligente qui serait autonome, sensible aux RF et à l'utilisateur, et qui inclurait la technologie logicielle et les capacités d'apprentissage machine ainsi qu'une grande connaissance de l'environnement radio haute-fidélité} \guillemotright{} \cite{Mitola99}.
 
@@
     Puis S. Haykin en $2005$ a aussi dit que
-    \emph{``une radio intelligente (RI) est un système de communication sans fil intelligent qui est capable de connaître son environnement, d'apprendre et d'adapter ses paramètres de fonctionnement (\eg, puissance d'émission et fréquence porteuse) à la volée, dans le but de fournir une communication fiable à tout moment, en tout lieu et efficace sur le plan spectral''} \cite{Haykin05}.
+    \guillemotleft{} \emph{une radio intelligente (RI) est un système de communication sans fil intelligent qui est capable de connaître son environnement, d'apprendre et d'adapter ses paramètres de fonctionnement (\eg, puissance d'émission et fréquence porteuse) à la volée, dans le but de fournir une communication fiable à tout moment, en tout lieu et efficace sur le plan spectral} \guillemotright{} \cite{Haykin05}.
 
@@
     L'\href{https://en.wikipedia.org/wiki/Cognitive_radio}{encyclopédie Wikipedia} dit ceci
-    \emph{``une RI est une radio qui peut être programmée et configurée dynamiquement pour utiliser les meilleurs canaux sans fil à proximité afin d'éviter les interférences et la congestion des utilisateurs.
-Une telle radio détecte automatiquement les canaux disponibles dans le spectre radio, puis modifie en conséquence ses paramètres de transmission ou de réception pour permettre plus de communications sans fil simultanées dans une bande de spectre donnée à un endroit donné''}.
+    \guillemotleft{} \emph{une RI est une radio qui peut être programmée et configurée dynamiquement pour utiliser les meilleurs canaux sans fil à proximité afin d'éviter les interférences et la congestion des utilisateurs.
+Une telle radio détecte automatiquement les canaux disponibles dans le spectre radio, puis modifie en conséquence ses paramètres de transmission ou de réception pour permettre plus de communications sans fil simultanées dans une bande de spectre donnée à un endroit donné} \guillemotright{}.
 \end{itemize}
@@
 La réglementation stipule que les UP ont une priorité stricte,
-mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de tout UP.
+mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection spectrale efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de tout UP.
 Ceci définit le concept d'\textbf{Accès Opportuniste au Spectre} (AOS, \emph{Opportunistic Spectrum Access}, OSA), pour lequel nous nous référons à l'article \cite{Zhao07} pour plus de détails.
@@
 Le livre de référence sur l'ApR est \cite{SuttonBarto2018}, citons donc la définition qu'en donnent R. Sutton et A. Barto :
-\emph{``l'apprentissage par renforcement, c'est apprendre quoi faire - comment faire le lien entre les situations et les actions -
+\guillemotleft{} \emph{l'apprentissage par renforcement, c'est apprendre quoi faire - comment faire le lien entre les situations et les actions -
 pour maximiser un signal de récompense numérique. L'apprenant n'est pas informé des mesures à prendre pour
-mais doit plutôt découvrir quelles actions sont les plus gratifiantes en les essayant''}.
+mais doit plutôt découvrir quelles actions sont les plus gratifiantes en les essayant} \guillemotright.
 % apprentissage statistique > séquentiel > par renforcement > information partielle (bandits)
@@
 Les deux communautés de l'AOS et des bandits ont commencé à interagir, et les travaux qui en ont résulté ont suscité un grand intérêt de la part des deux communautés, depuis la fin des années $2000$ et le début des années $2010$, avec des travaux pionniers comme \cite{Liu08,Zhao10,Jouini09,Jouini10}.
-L'accent est mis sur un US accédant à un spectre sous licence, occupé par un UP, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès d'``écoute avant transmission''.
+L'accent est mis sur un US accédant à un spectre sous licence, occupé par un UP, qui a une priorité stricte sur l'US, qui doit suivre un schéma d'accès de type \guillemotleft{} écoute avant de transmettre \guillemotright{}.
 %
@@
 En se concentrant sur un US dans un réseau AOS, il doit décider séquentiellement d'un canal à analyser (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il écoute ce canal et effectue une détection de PU, et enfin il transmet dans ce canal s'il a été détecté libre.
-Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente ``écologique'', \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou de façon équivalente, de maximiser son nombre de transmissions réussies.
+Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente \guillemotleft{} écologique \guillemotright{}, \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou de façon équivalente, de maximiser son nombre de transmissions réussies.
 %
@@
 %
-\begin{itemize}\tightlist
+\begin{itemize}%\tightlist
     % \item
@@
 %
-De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de leur batterie, bien que les objets soient capables d'utiliser leur antenne de réception pour écouter dans un canal occasionnellement (généralement, pendant quelques intervalles de temps après chaque message de liaison montante), ils ne peuvent pas effectuer de détection spectrale à chaque instant, comme cela est envisagé pour l'AOS (\ie, au début de chaque intervalle, dans un schéma de type ``écoute avant de transmettre'').
+De plus, en raison de leur matériel limité (Rx/Tx), de leur puissance de calcul limitée et de la durée de vie limitée de leur batterie, bien que les objets soient capables d'utiliser leur antenne de réception pour écouter dans un canal occasionnellement (pendant quelques intervalles de temps après chaque message envoyé), ils ne peuvent pas effectuer de détection spectrale à chaque instant, comme cela est envisagé pour l'AOS (\ie, au début de chaque intervalle, dans un schéma de type \guillemotleft{} écoute avant de transmettre \guillemotright{}).
 
@@
 Le premier modèle sans retransmission est issu du premier article écrit au cours de cette thèse, qui a initié une collaboration avec R. Bonnefoi, un autre doctorant de notre équipe SCEE.
-Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre 2010, où il a obtenu le ``prix du meilleur papier'' \cite{Bonnefoi17}.
+Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre 2010, où il a obtenu le \guillemotleft{} prix du meilleur papier \guillemotright{} \cite{Bonnefoi17}.
 %
@@
 Cela signifie que chaque objet vise son propre objectif local, qui est de maximiser sa récompense cumulée, d'une manière égoïste et complètement agnostique des autres objets ayant le même objectif.
-Il est bien connu en théorie des jeux que jouer égoïstement peut être désastreux pour la mesure centralisée de la performance, on peut penser aux ``dilemmes'' populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
+Il est bien connu en théorie des jeux que jouer égoïstement peut être désastreux pour la mesure centralisée de la performance, on peut penser aux \guillemotleft{} dilemmes \guillemotright{} populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
 Il était donc assez surprenant que nos simulations numériques, ainsi que la preuve de concept réaliste, aient montré que l'apprentissage BMB décentralisé et égoïste conduisait à une coordination efficace entre les objets, dans tous les scénarios considérés, malgré le fait que ces objets ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
@@
 Nous prouvons qu'en combinant l'algorithme \klUCB, asymptotiquement optimale pour les problèmes stationnaires, et notre nouvelle analyse du GLRT pour les récompenses bornées, nous obtenons des garanties de pointe sur le regret de notre algorithme proposé, noté \GLRklUCB.
-Nous considérons la même hypothèse que nos concurrents, en supposant que la longueur des intervalles stationnaires est ``assez longue'' pour que les points de changement soient détectables.
+Nous considérons la même hypothèse que nos concurrents, en supposant que la longueur des intervalles stationnaires est \guillemotleft{} assez longue \guillemotright{} pour que les points de changement soient détectables.
 La meilleure borne supérieure de regret est obtenue lorsque l'algorithme connaît à l'avance l'horizon et le nombre de points de changement, mais notre algorithme n'a pas besoin d'avoir d'autres connaissances sur la difficulté du problème pour être paramétré de manière optimale.
@@
     Nous formalisons le modèle du bandit multi-joueur, pour lequel nous avons introduit deux variantes, dans le Chapitre~\ref{chapter:5}.
-    Pour le cas avec information de détection spectrale (``spectral sensing''), nous proposons deux algorithmes, et nous analysons notre algorithme \MCTopM, pour prouver qu'il est très efficace, à temps fini et asymptotiquement.
+    Pour le cas avec information de détection spectrale (\emph{spectral sensing}), nous proposons deux algorithmes, et nous analysons notre algorithme \MCTopM, pour prouver qu'il est très efficace, à temps fini et asymptotiquement.
     Nous présentons aussi des expériences numériques approfondies pour montrer qu'il est bien plus efficace que les autres algorithmes de la littérature.

### 2019-07-08 2-Chapters/1-Chapter/chapter1.tex
LOTS of works…

@@
 This manuscript concludes my doctoral thesis, which started in October $2016$ and finished in November $2019$.
-My research took place at the IETR laboratory in Rennes (France), in the SCEE team hosted on the Rennes campus of the engineering school CentraleSupélec.
+My research took place at the IETR laboratory in Rennes (France), in the SCEE team, hosted on the Rennes campus of the engineering school CentraleSupélec.
 I was supervised by Professor Christophe Moy, in Rennes,
@@
 That is why we are interested in this thesis about
-improving the battery life of IoT end-devices and reduce the energy cost of IoT networks.
+improving the battery life of IoT end-devices and reducing the energy cost of IoT networks.
 We propose to attain these two goals jointly, by embedding low-cost decentralized decision making on directly into the future IoT devices.
@@
 This third kind of systems can be designed as decentralized:
-even if a central base station is
-still in charge of many devices,
+because even if a central base station is still in charge of many devices,
 the devices initiate the sending of up-link packets, and the only down-link data they can receive are short acknowledgements, sent from the base-station to indicate success or failure of every up-link packet.
@@
 This third kind of wireless systems are characterized by their decentralized nature,
-where communications are initialized and regulated by the devices, not a centralized control systems
+where communications are initialized and regulated by the devices, not by a centralized control systems.
 % due to the really limited information that the base station can send to the devices (in downlink), it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
@@
 Most commercial IoT companies nowadays indeed sell a lifespan larger than $10$ years, like SIGFOX \cite{Centenaro16}.
-Another common constraint for IoT devices is their low duty cycles, as most applications target one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems (such as 4G/5G and WiFi).
+Another common constraint for IoT devices is their low duty cycles, as most applications target a need for one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems (such as 4G/5G and WiFi).
 %
@@
 and they consist in a specification for both the \emph{PHY}sical
-and the \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}) layer.
-To quote some examples of IoT standards, ZigBee, Z-Wave or Bluetooth are targeting for short-range communications (up-to \SI{2}{\meter}), while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications (up-to \SI{50}{\kilo\meter}).
-We refer to the survey \cite{Centenaro16} for more details, and \cite{Azari18} or our recent article \cite{MoyBesson2019}, and references therein.
+and the \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}) layers.
+To quote some examples of IoT standards, ZigBee, Z-Wave or Bluetooth are targeting short-range communications (up-to \SI{2}{\meter}), while LoRaWAN, SIGFOX, Ingenu or Weightless are designed for long-range communications (up-to \SI{50}{\kilo\meter}).
+We refer to the survey \cite{Centenaro16} and references in our recent articles \cite{MoyBesson2019,MoyBesson2019Annales} for more details.
 
@@
 
-\textbf{Spectrum Scarcity.}
+\paragraph{Spectrum Scarcity.}
 %
@@
     % \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
-    \includegraphics[width=0.95\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
+    \includegraphics[width=0.92\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
     \caption[A chart representing the allocation of radio spectrum in the United States of North America in $2016$]{A chart representing the allocation of radio spectrum in the United States of North America in $2016$.
-\copyright{} United States Department of Commerce, published online at \href{https://www.ntia.doc.gov/files/ntia/publications/january_2016_spectrum_wall_chart.pdf}{\texttt{www.ntia.doc.gov/files/ntia/publications/january\_2016\_spectrum\_wall\_chart.pdf}}.}
+\copyright{} United States of North America, Department of Commerce, published online at \href{https://www.ntia.doc.gov/files/ntia/publications/january_2016_spectrum_wall_chart.pdf}{\texttt{www.ntia.doc.gov/files/ntia/publications/january\_2016\_spectrum\_wall\_chart.pdf}}.}
     % Only a small share of the spectrum is not yet allocated, the top-left corner in white and indicated by \textcolor{red}{\textbf{Free} in red} (it is in logarithmic scale from \SI{0}{\kilo\hertz} to \SI{300}{\giga\hertz})
@@
 
-Cellular network bands are overloaded in most parts of the world, but other frequency bands (such as military, amateur radio and paging frequencies) are less utilized.
+Cellular network bands are overloaded in most parts of the world (2G/3G/4G/5G), but other frequency bands (such as military, amateur radio and paging frequencies) are less utilized.
 Independent studies performed in some countries confirmed that observation, and concluded that spectrum usage highly depends on both time and place, as shown by \cite{Lopez2009spectral} for instance.
@@
 Therefore, in the last $15$ years, thanks to an active lobbying by the cognitive radio community,
-regulatory bodies in the world have been considering whether to allow
+regulatory bodies all over the world have been considering whether to allow
 a new wireless communication paradigm:
 enable unlicensed users in licensed bands, if they would not cause any interference to the (paying) licensed users.
-These initiatives are considered by the \textbf{Cognitive Radio},
-field that we detail below, and especially for \textbf{Dynamic Spectrum Access (DSA)},
+These initiatives are considered by the \textbf{Cognitive Radio} field that we detail below, and especially for \textbf{Dynamic Spectrum Access (DSA)},
 for which we refer to the surveys \cite{akyildiz2006next,garhwal2012survey} for more details.
@@
 
-In this thesis, we study long-range Internet of Things networks, characterized by mainly three essential constraints:
+In this thesis, we study Internet of Things networks characterized by mainly three essential constraints:
 low power consumption (or long battery life),
-long-range
+long-range communications,
 and low duty cycle.
@@
 
-\paragraph{From IT to Cognitive Radio.}
+\textbf{From IT to Cognitive Radio.}
 %
@@
 %  with the exception of a unique pair of receiver/transmitter antennas.
-Even though the SDR paradigm was initiated by the United States of North America defense research in the $1980$s, over the past $20$ years the civil industry has started to be interested by SDR, and CR has got significant attention from the academia and industry as well.
+Even though the SDR paradigm was initiated by the United States of North America defense research in the $1980$s, over the past $20$ years the civil industry has started to be interested by SDR, and CR has got significant attention from the academia as well.
 % CR has been the first answer to solve spectrum scarcity issue which collected a huge attention in early 21st century.
 %
-As CR is not a standard technology, it does not have a single definition, so let us start by quoting the definition of two researchers whose works were the roots of the development of CR, that just appeared $20$ years ago.
+As CR is not a standard technology, it does not have a single definition, so let us start by quoting the definition of two researchers whose works constitutes the roots of the development of CR, the first one having appeared just $20$ years ago.
 % - technologies de l'information > télécommunications > communications sans fil > radio intelligente > DSA > OSA
@@
 %
-In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use the network.
-As exposed in the definition above of the CR philosophy, we could imagine that if the network is unused by any PU in a certain location and time, then non licensed users, referred to as Secondary User (SU), could use it, possibly by paying some rent to the network operator.
+In licensed bands, there are Primary Users (PU) paying to access the network, for instance anybody has to pay to have a mobile phone number and use cellular networks.
+As exposed in the definition above of the CR philosophy, we could imagine that if the network is unused by any PU in a certain location and time, then non licensed users, referred to as Secondary User (SU), could use it, possibly by also paying some rent to the network operator.
 %
 Regulation states that PU have a strict priority,
-but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU was equipped with an efficient sensing capacity, it could analyze its environment, and use a licensed band if and only if it is free of any PU.
+but even though the RF bands are allocated, real world measurements often show that some bands are not densely used, and thus if a SU was equipped with an efficient \emph{spectrum sensing} capacity, it could analyze its environment, and use a licensed band if and only if it is free of any PU.
 % explain DSA
@@
 
-\paragraph{From Statistics or Machine Learning to Multi-Armed Bandits.}
+\textbf{From Statistics or Machine Learning to Multi-Armed Bandits.}
 %
@@
 We are interested by Multi-Armed Bandits (MAB) learning,
-which emerged as an interesting problem in the statistics community and among researchers interested by sequential learning, as shown by the earlier works of \cite{Thompson33,Robbins52,LaiRobbins85}.
+which emerged as an interesting problem in the statistics community and among researchers interested by sequential learning, as shown by the pioneer works of \cite{Thompson33,Robbins52,LaiRobbins85}.
 It is also included in the more general concept of Reinforcement Learning (RL), itself one of the field of Machine Learning (ML).
@@
 %
-In other words, the player only sees the reward given by its action at every round, and not the reward that would have been given if she chose any other action.
-This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of \emph{Multi-Armed Bandits} (MAB) in details in the next Chapter~\ref{chapter:2}.
-Clinical trials and treatment discovery were historically the first application of MAB since the $1930$s, with the early work of W. Thompson \cite{Thompson33}.
-%
-MAB are a simple yet powerful example of the well-known \emph{exploration vs exploitation dilemma}:
-when facing a set of $K$ actions whose effects on the environment are unknown, a learner must balance between
+In other words, the player only sees the reward generated by playing its action at every round, and not the reward that would have been given if she chose any other action.
+This kind of limited feedback is called \emph{bandit information}, and we discuss the history and the concept of MAB in details in the next Chapter~\ref{chapter:2}.
+Clinical trials and treatment discovery were historically the first application of MAB since the $1930$s, with the early work of W. Thompson \cite{Thompson33},
+%
+as MAB are indeed a simple yet powerful example of the well-known \emph{exploration vs exploitation dilemma}.
+When facing a set of $K$ actions whose effects on the environment are unknown, a learner must balance between
 exploring the unknown actions, in order to collect more information about them,
@@
 Robbins \cite{Robbins52}, and more recently it is an active field of research, since the late $1990$s \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
-Research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}, and the wide range of applications of MAB in the recent years \cite{bouneffouf2019survey}.
-
-\paragraph{Opportunistic Spectrum Access.}
+The research on MAB has produced a vast literature in the $2000$s \cite{Auer02,Auer02NonStochastic,Audibert2009minimax} and continues to be a topic of high interest, as illustrated by the surveys and books \cite{Bubeck12,LattimoreBanditAlgorithmsBook,Slivkins2019}, and the wide range of applications of MAB in the recent years \cite{bouneffouf2019survey}.
+
+\textbf{Opportunistic Spectrum Access.}
 %
@@
 The following hypotheses are maid:
-the SU is equipped with a sensing capability,
+the SU is equipped with a spectrum sensing capability,
 and considers a fixed and finite set of $K$ orthogonal channels, \ie, different frequency bands in a licensed spectrum.
@@
 %
-Thus if the SU spends a short time in the beginning of each time slot to perform sensing, it can scan for the presence or absence of any PU before transmitting, during the rest of the time slot without colliding with the PU.
+Thus if the SU spends a short time in the beginning of each time slot to perform sensing, it can scan for the presence or absence of any PU before transmitting, without colliding with the PU during the rest of the time slot.
 If the SU was able to sense for all the $K$ frequency bands, it could simply transmit in one of the free channels if any, or not transmit if all channels are used at a given time step.
 However, sensing is known to be costly, in terms of energy consumption, especially for wide-band sensing, as detailed in the surveys \cite{yucek2009survey,subhedar2011spectrum}, thus most works on OSA limit the sensing capacity of SU to sensing only one channel at a time.
-This assumption, along with the hypothesis that PU cannot be disturbed (without which no regulation will never be accepted for OSA), enforces the SU to transmit in the channel that it sensed, if and only if it was sensed free from any PU.
+This assumption, along with the hypothesis that PU cannot be disturbed (without which no regulation will ever be accepted for OSA), enforces the SU to transmit in the channel that it sensed, if and only if it was sensed free from any PU.
 % non negligible cost of radio reconfiguration
 
-Focusing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of channels, $[K]=\{1,\dots,K\}$), then it performs sensing, and finally it transmits in this channel if it was sensed free.
+Focusing on one SU in an OSA network, it has to sequentially decide a channel to sense (in the set of channels, $[K]=\{1,\dots,K\}$), then it performs spectrum sensing, and finally it transmits in this channel if it was sensed free.
 The goal of the SU is to minimize its energy consumption (we remind that we focus on ``green'' cognitive radio) and to maximize its up-link data rate, or equivalently, to maximize its number of successful transmission.
@@
 in a sequential action-then-feedback cycle (time steps are $t\in\N^*$),
-under partial information feedback (\ie, the player receives feedback about only one channel over $K$).
-These three hypotheses are the ones that restrict from the general ML framework to the specific MAB case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
-
-\paragraph{MAB for OSA, and a short history of these researches at the SCEE team.}
+under partial information feedback (\ie, the player receives sensing feedback about only one channel over $K$).
+These three hypotheses are the ones that restrict from the general sequential learning framework to the specific MAB case (see Figure~\ref{fig:1:ReinforcementLearningCycle}).
+
+\textbf{MAB for OSA, and a short history of these researches at the SCEE team.}
 %
@@
 and the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$ is modeled by a reward of $r(t) \in \{0,1\}$.
-Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus a SU message can be sent), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at during the time slot $t$, and no SU message should be sent, until the end of that slot.
-%
-This model was first studied by W. Jouini during his PhD thesis with C. Moy \cite{Jouini12PhD}, ten years ago, first in the article \cite{Jouini09} and later in \cite{Jouini10,Jouini12}.
+Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus the SU can send a message), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at during the time slot $t$, and no message should be sent, until the end of that slot.
+%
+This model was first studied by W. Jouini during his PhD thesis with C. Moy \cite{Jouini12PhD}, about ten years ago, first in the article \cite{Jouini09} and later in \cite{Jouini10,Jouini12}.
 
@@
 Moy and his student C.
-Robert \cite{RobertSDR2014,MoyWSR2014}, using USRP boards and the MATLAB/Simulink software.
+Robert, using USRP boards and the MATLAB/Simulink software \cite{RobertSDR2014,MoyWSR2014}.
 In a second PhD thesis, N. Modi studied from $2014$ to $2017$ the impact on the battery life of a wireless device of using MAB algorithms to optimize channel selections \cite{Modi17PhD}.
@@
 
-\paragraph{Limitations and specifities of IoT networks.}
-%
-The aforementioned literature have essentially shown that MAB algorithms can be applied with success for the OSA problem.
-However, if we consider CR-ready devices that cannot perform sensing, such as low-cost and low energy-consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
-Moreover, in most cases, the IoT networks use unlicensed bands, and as such there are no longer a distinction between PU and SU.
+\textbf{Limitations and specifities of IoT networks.}
+%
+The aforementioned literature have essentially shown that MAB algorithms can be applied successfully on the SU side for the OSA problem.
+However, if we consider CR-ready devices that cannot perform spectrum sensing, such as low-cost and low energy-consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
+Moreover, in most cases, the IoT networks use unlicensed bands, and as such there is no longer a distinction between PU and SU.
 In other words, there is no longer any priority between users in IoT networks.
@@
     most IoT devices are very low-cost devices, run on low-quality hardware, and have limited software and power capabilities (\ie, tiny batteries).
-    It means that even if they are not equipped with particular sensing capabilities,
+    It means that even if they are not equipped with particular spectrum sensing capabilities,
     they are equipped with a transmitter as well as with a receiver (Tx and Rx, \ie, a transceiver),

### 2019-07-08 1-Preamble/5-Abstract/resume_fr.tex
Done for proof reading intro FR/EN

@@
 Modi a étudié de $2014$ à $2017$ l'impact sur la durée de vie des batteries d'un objet sans fil de l'utilisation des algorithmes BMB pour optimiser la sélection des canaux, \cite{Modi17PhD}.
-D'une part, l'utilisation d'un algorithme BMB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du objet.
-D'autre part, les algorithmes BMB classiques ont tendance à changer souvent de bras, surtout au début de leur processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour l'objet sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par l'objet.
-Chaque reconfiguration matérielle coûte de l'énergie à l'objet, et les algorithmes à commutation rapide entraînent donc une réduction de l'autonomie de la batterie.
-Le compromis entre les deux aspects est étudié de façon empirique dans les études suivantes
-\cite{modiDemo2016}.
+% D'une part, l'utilisation d'un algorithme BMB tel que les algorithmes de type UCB s'est avérée utile et pourrait apporter des améliorations significatives en termes de taux de transmission, augmentant directement la durée de vie de la batterie du objet.
+% D'autre part, les algorithmes BMB classiques ont tendance à changer souvent de bras, surtout au début de leur processus d'apprentissage, ce qui induit beaucoup de reconfigurations matérielles dynamiques pour l'objet sans fil, car la sélection d'un canal différent nécessite un changement dans le matériel radio utilisé par l'objet.
+% Chaque reconfiguration matérielle coûte de l'énergie à l'objet, et les algorithmes à commutation rapide entraînent donc une réduction de l'autonomie de la batterie.
+Le compromis entre des reconfigurations fréquentes, qui coûtent en énergie, mais qui permettent d'apprendre rapidement est étudié de façon empirique dans \cite{modiDemo2016}.
 

### 2019-07-08 2-Chapters/1-Chapter/chapter1.tex
Done for proof reading intro FR/EN

@@
 
-\textbf{From IT to Cognitive Radio.}
+\paragraph{From IT to Cognitive Radio.}
 %
@@
 
-\textbf{From Statistics or Machine Learning to Multi-Armed Bandits.}
+\paragraph{From Statistics or Machine Learning to Multi-Armed Bandits.}
 %
@@
 
-\textbf{Opportunistic Spectrum Access.}
+\paragraph{Opportunistic Spectrum Access.}
 %
@@
 
-\textbf{MAB for OSA, and a short history of these researches at the SCEE team.}
+\paragraph{MAB for OSA, and a short history of these researches at the SCEE team.}
 %
@@
 In a second PhD thesis, N. Modi studied from $2014$ to $2017$ the impact on the battery life of a wireless device of using MAB algorithms to optimize channel selections \cite{Modi17PhD}.
-On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and could bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
-On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of their learning process, and this induces a lot of dynamical hardware reconfigurations for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
-Each hardware reconfiguration costs energy for the device, and so quickly-switching algorithms lead to a reduction of the battery life.
-The trade-off between the two aspects is studied empirically in
-\cite{modiDemo2016}.
+% On the one hand, running a MAB algorithm such as UCB-like algorithms was proven to be useful and could bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
+% On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of their learning process, and this induces a lot of dynamical hardware reconfigurations for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
+% Each hardware reconfiguration costs energy for the device, and so quickly-switching algorithms lead to a reduction of the battery life.
+The trade-off between rapid software reconfigurations that cost energy but allow to learn quickly is studied empirically in \cite{modiDemo2016}.
 
@@
 
-\textbf{Limitations and specifities of IoT networks.}
+\paragraph{Limitations and specifities of IoT networks.}
 %
@@
 % (see \cite{LattimoreBanditAlgorithmsBook,Slivkins2019} for surveys).
-We thus start the first Part~\ref{part:Introduction} of this thesis with Chapter~\ref{chapter:2}, which presents the MAB model and gives a review of the most important algorithms designed to solve stochastic and stationary MAB problems.
+We thus start the first part of this thesis with Chapter~\ref{chapter:2}, which presents the MAB model and reviews the most important algorithms designed to solve stochastic and stationary MAB problems.
 %
@@
 we were not only interested by the usual measure of performance of any MAB algorithm (its regret, see below in Section~\ref{sec:2:lowerUpperBoundsRegret}),
-but also by their empirical performances in terms of computational complexity and storage requirements, from the point of view of both theoretical analyses and real-world measurements.
+but also by their empirical performances in terms of computational complexity and storage requirements, from the points of view of both theoretical analyses and real-world measurements.
 %  on time and memory footprints.
@@
 Our exploration of the large number of MAB algorithms and models developed in the recent literature
-has given us the ambition to write a single piece of software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explore the performances of newly proposed algorithms.
-To tackle this goal, we developed a Python library of simulation of MAB problems.
+has given us the ambition to write a single software allowing anyone to easily implement new models and algorithms, in order to compare the existing ones and empirically explore the performances of newly proposed algorithms.
+To tackle this goal, we developed a library of simulation of MAB problems, in the Python language.
 %
@@
 %
-Because there are so many MAB algorithms available, we are also interested by the question of how a practitioner can chose the algorithm to implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
+Because there are so many MAB algorithms available, we are also interested by the question of how an engineer can chose the algorithm to implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
 To answer this question, we present two approaches.
@@
 We confirm that widely used and not too sophisticated algorithms, such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
-The second approach is an online algorithm selection, consisting in aggregating a (finite) set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
-This work on aggregating MAB algorithms was motivated for the OSA case of Cognitive Radio, where many previous research works only considered the $\UCB_1$ algorithm without really justifying this choice.
+The second approach is online algorithm selection, consisting in aggregating a (finite) set of algorithms and automatically discovering which one performs the most efficiently, against a given problem, with our contribution \Aggr{} that we detail in Chapter~\ref{chapter:25}.
+This work on aggregating MAB algorithms was motivated for the OSA case of CR, where many previous research works only considered the $\UCB_1$ algorithm without really justifying their choice.
 This contribution was presented at the IEEE WCNC conference in Barcelona, Spain, in April $2018$.
@@
 
-In the second Part~\ref{part:MABIOT} of this thesis, we start by proposing and studying different models of IoT networks, with simulations and a real-world proof-of-concept, and then we study interesting questions on two mathematical models of MAB, arising from simplification of our first model.
+In the second part of this thesis, we start by proposing and studying different models of IoT networks, with simulations and a real-world proof-of-concept, and then we study interesting questions on two mathematical models of MAB, arising from simplification of our first model.
 
@@
 Such devices use acknowledgements as feedback, in order to optimize their up-link communications, by accessing more the best channels (\ie, the channel less occupied by the surrounding traffic).
-The goal of this first model is to increase the Quality of Service (QoS) of the application of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmission.
+The goal of this first model is to increase the Quality of Service (QoS) of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmissions.
 The base station will thus receive more up-link packets from the devices being served, if they can successfully learn by themselves an efficient spectrum access scheme.
-It implies that the network can accept more end-devices while maintaining the same QoS.
-%
-Our second model then considers packet retransmissions, and while the two concepts are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increased QoS for the entire network, as less retransmissions reduce the local spectrum load.
+It implies that the network can serve more end-devices while maintaining the same QoS.
+%
+Our second model then considers packet retransmissions, and while the two models are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increased QoS for the entire network, as less retransmissions reduce the local spectrum load.
 
@@
 % Even without changing anything on the level of the considered IoT standard (neither the PHY nor the MAC layers), our proposal is just an add-on capability, that can be set-up on each device, on a unit-per-unit basis.
-The first model comes from the first article written during this thesis, which initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
+The model without retransmission comes from the first article written during this thesis, which initiated a collaboration with R. Bonnefoi, another PhD student of our team SCEE.
 We presented it at the EAI CROWNCOM conference in Lisboa, Portugal in September $2017$, where it obtained the ``best paper award'' \cite{Bonnefoi17}.
@@
 % % Verify this first model empirically: proof of concept
-The proof-of-concept (PoC) continued the fruitful collaboration, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later presented at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
+The proof-of-concept (PoC) continued the fruitful collaboration, and was demonstrated during three days at the IEEE ICT conference in Saint-Malo, France, in June $2018$ \cite{Besson2018ICT}, and later at the IEEE WCNC conference in Marrakech, Morocco, in April $2019$ \cite{Besson2019WCNC}.
 %
@@
 %
-In both models of IoT networks presented above, the core concept is to let every dynamic device run a learning algorithm in a fully decentralized way, in order to optimize the complete system.
+In both models of IoT networks presented above, the core idea is to let every dynamic device run a learning algorithm in a fully decentralized way, in order to optimize the entire system.
 It is important to note that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own random activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
-It means that each of them target its own local objective, which is to maximize its cumulated reward, in a selfish way and by being agnostic to the other devices following the same objective.
+It means that each of them target its own local objective, which is to maximize its cumulated reward, in a selfish way and by being agnostic about the other devices following the same objective.
 It is well known in game theory that playing selfishly can be disastrous for the centralized performance measure, one can think of popular ``dilemmas'', such as the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma}.
@@
 The model we studied comes with different variants, depending on the feedback level, and covers both the OSA case (\ie, with sensing of PU) or the IoT case (\ie, without sensing).
-The goal was to understand the heuristic of Chapter~\ref{chapter:4}, \Selfish, in the simpler framework of multi-player MAB, which has been studied before for the case of OSA, as studied for instance by \cite{Zhao10,Anandkumar10,Anandkumar11}.
-%
-The OSA case covered by the studied model can seem to slightly differ with the one considered by \cite{Jouini10} and other previous works,
+The goal was to understand the heuristic of Chapter~\ref{chapter:4}, \Selfish, in the simpler framework of multi-player MAB, which has been studied before for the case of OSA, for instance by \cite{Zhao10,Anandkumar10,Anandkumar11}.
+%
+The OSA case covered by the studied model can seem to slightly differ with the one exposed above, from \cite{Jouini10} and other previous works,
 as our model requires that an \Ack{} is sent back by the base station if the transmission was successful, even in the case where sensing information is available.
-It is actually equivalent to the previously studied models of applying MAB for OSA, as they only considered synchronized devices, one SU and PU, and thus if the sensing indicates that one channel is free at one time slot, the up-link message sent by the SU device is sure to be successfully received by the base station (in the ideal model), so there is no risk of collision, and thus no need for an \Ack.
-
-On the one hand, in the OSA setting, we failed to obtain positive result for the \Selfish{} policy, as we proved that in some limited settings (\eg, $K=3$ and $M=2$), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
+It is actually equivalent to the previously studied models of applying MAB for OSA, as they only considered synchronized devices, only one SU and PU, and thus if the sensing indicates that one channel is free at one time slot, the up-link message sent by the SU device is sure to be successfully received by the base station (in the ideal model), so there is no risk of collision, and thus no need for an \Ack.
+
+On the one hand, in the OSA setting, we failed to obtain positive result for the \Selfish{} policy, as we proved that in some limited settings (\eg, $K=3$ channels and $M=2$ devices), \Selfish-\UCB{} can suffer from linear regret with a small probability, and thus suffers from linear mean regret.
 This result was later confirmed and analyzed by other researchers in \cite{BoursierPerchet18} (Appendix~F).
 %
-On the other hand, we were able to propose new algorithms for this multi-player bandit problem with sensing information, and we analyzed our proposal \MCTopM-\klUCB, to show that it achieves a finite-time logarithmic regret upper bound, improving significantly over previous state-of-the-art results.
+On the other hand, we were able to propose new algorithms for this multi-player bandit problem with sensing information, and we analyzed our proposal \MCTopM{} to show that it achieves a finite-time logarithmic regret upper bound, and improves significantly over previous state-of-the-art results.
 Our algorithm also achieves a logarithmic number of collisions and arm switches, and allows a fixed group of $M$ devices to efficiently learn to use the $M$ best channels orthogonally for almost all their up-link communications.
@@
 As explained above, the model of Chapter~\ref{chapter:4} was found intractable to analyze, mainly because we consider an IoT network with many end-devices, all following random activation patterns.
-The difficulty does not reside in the fact that we are trying to analyze MAB algorithms, designed to tackle stationary problems, on a non-stationary problem,
-but rather that we are analyzing algorithms which are all playing in different (random) activation times, subsets of the global synchronized time steps.
+The difficulty does not reside in the fact that we are trying to analyze MAB algorithms, designed to tackle stationary problems, applied on a non-stationary problem,
+but rather that we consider algorithms which are all playing in different (random) activation times, subsets of the global synchronized time steps.
 Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is in fact not really realistic for IoT networks, even though this lead to the interesting model of Chapter~\ref{chapter:5}.
@@
 However, even if we continue to assume a synchronized time in all the rest of the thesis,
-it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid signaling and any central control of the devices by the base station, as we explained above.
+it is hard to argue that this hypothesis of a centralized time schedule for the end-devices can be realistic, because we study the case of decentralized learning for OSA and IoT precisely in order to avoid signaling and any central control of the devices by the base station, as we explained before.
 
@@
 If we focus on one IoT device, its environment (\ie, the surrounding devices) is non-stationary, meaning that its average properties can fluctuate with time, but real environments usually present non-stationarities with certain structures.
-For instance, enforcing stationarity on some time intervals leads to the last contribution and the last chapter of this thesis.
+%
+It is thus interesting and realistic to enforce stationarity on some time intervals, and this leads to the last contribution and the last chapter of this thesis.
 % DONE explain why is it interesting to study piece-wise stationary problems from the point-of-view of one player...
@@
 We begin the last Chapter~\ref{chapter:6} by reviewing the existing works,
-and we chose to focus on the piece-wise stationary problem,
-meaning that the underlying bandit problem is stationary on some intervals, separated by change-points located at unknown times.
-Assuming the environment to be piece-wise stationary indeed makes sense for applications to wireless networks, where a change-point can for instance correspond to the arrival of a new group of end-devices in the network.
+and we chose to focus on the piece-wise stationary problem (\ie, abruptly-changing),
+meaning that the underlying bandit problem is stationary on consecutive intervals, separated by change-points located at unknown times.
+Assuming the environment to be piece-wise stationary indeed makes sense for applications to wireless networks, where a change-point can for instance corresponds to the arrival of a new group of end-devices in the network.
 Examples of such situation can be a new company arriving on the market in one city (\eg, the Linky counters being set-up nowadays in France), or the neighbor farmer installing sensors on his own herd of cows, etc.
@@
 We first show finite-time bounds on both the false alarm probability of the test, and its detection delay, under mild assumptions on the lengths of the stationary sequences.
-Our algorithm, denoted \GLRklUCB, is then presented in two variants, whether change-points are local (\ie, only one arm mean changes at each change-point) or global (\ie, possibly all the arm means change at a time).
-%
-We prove that by combining the policy \klUCB, asymptotically optimal for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of the proposed algorithm.
+Our algorithm is then presented in two variants, whether change-points are local (\ie, only one arm mean changes at each change-point) or global (\ie, possibly all the arm means change at a time).
+%
+We prove that by combining the policy \klUCB, efficient for stationary problems, and our new analysis of the GLRT for bounded rewards, we obtain state-of-the-art guarantees on the regret of the proposed algorithm, denoted \GLRklUCB.
 We consider the same hypothesis as our competitors, by assuming that the length of the stationary intervals to be ``long enough'' for the change-points to be detectable.
@@
 % The rest of the appendix
-a list of abbreviations and notations, then lists of figures, algorithms, code samples and tables,
+a list of abbreviations and notations, then with lists of figures, algorithms, code samples and tables,
 and finally the list of the bibliographical references.
@@
     % We made a video showcasing our demonstration, hosted at
-    and in a video, hosted at
+    and in a $6$-minute video, hosted at
     \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
@@
     as well as extensive numerical experiments to demonstrate its excellent performance in comparison with the rest of the literature.
-    Some recent research works heavily built up on our results, and the research on multi-players bandits has been quite active since its publication \cite{Besson2018ALT}.
+    Some recent research works built up on our results, and the research on multi-players bandits has been quite active since its publication \cite{Besson2018ALT}.
 
@@
         \node[align=center] (introduction) at (0,3.25) [rectangle,draw,fill=blue!15] {\textbf{Chapter~\ref{chapter:1}}\\Introduction};
-        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!15] {\textbf{Chapter~\ref{chapter:2}}\\Stochastic and Stationary\\Multi-Armed Bandit models};
+        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw,fill=red!15] {\textbf{Chapter~\ref{chapter:2}}\\The Stochastic\\Multi-Armed Bandit models};
         \node[align=center] (chapter3) at (-2.5,2.25) [rectangle,draw,fill=red!10] {\textbf{Chapter~\ref{chapter:3}}\\SMPyBandits: simulation\\library for MAB};
@@
         \node[align=center] (chapter5) at (0,1) [rectangle,draw,fill=green!15] {\textbf{Chapter~\ref{chapter:5}}\\Multi-players\\Multi-Armed Bandits};
-        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!20] {\textbf{Chapter~\ref{chapter:6}}\\Non-stationary\\Multi-Armed Bandits};
+        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw,fill=green!20] {\textbf{Chapter~\ref{chapter:6}}\\Piece-Wise Stationary\\Multi-Armed Bandits};
         \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw,fill=blue!20] {\textbf{Chapter~\ref{chapter:conclusion}}\\General Conclusion};
@@
 \textcolor{darkred}{First, in Part~\ref{part:Introduction}} (second line), we start by the next Chapter~\ref{chapter:2}, required for the rest of the document, as we introduce the MAB models and the notations used in this thesis.
-Conversely, even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use numerical simulations based on our simulation library SMPyBandits, the Chapter~\ref{chapter:3} where we present it is not required to understand them.
+Conversely, even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use numerical simulations based on our simulation library SMPyBandits, the Chapter~\ref{chapter:3} where we present SMPyBandits is not required to understand them.
 We conclude this part with Chapter~\ref{chapter:25}, where we detail one of the first contributions of this thesis, a new algorithm for online MAB algorithms selection.
@@
 For this reason, we weaken the models for the rest of the document,
-and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.
+and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the intractable IoT networks models from Chapter~\ref{chapter:4}.
 % \end{itemize}
@@
 
-We conclude this chapter with a list of works published during this PhD.
+We conclude this chapter with a list of works published during my PhD.
 All the following works are published entirely and freely, on the HAL platform (on \href{https://hal.archives-ouvertes.fr/}{\texttt{HAL.Archives-Ouvertes.fr}}).

### 2019-07-08 2-Chapters/2-Chapter/chapter2.tex
Proof read abstract & conclusion of Ch2, Ch3, Ch4, Ch5

@@
 %
-As explained in Chapter~\ref{chapter:1}, we focus on decision-making models with a finite number of resources, called arms, and on stochastic models, where an arm is associated to a binary or real-valued distribution.
+As explained in Chapter~\ref{chapter:1}, we focus in this thesis on Reinforcement Learning (RL), and especially on decision-making models with a finite number of resources, called arms, and under a stochastic hypothesis: an arm being associated to a binary or real-valued distribution.
 In this chapter, we present the common base of all the more sophisticated mathematical models studied in this thesis: the stochastic multi-armed bandit (MAB) model, by restricting to the single-player and stationary stochastic case.
 %
-We motivate this model and discuss its application to and beyond cognitive radio.
-%
-We formalize this MAB model, then we quote some of its important applications.
-We define the notion of regret of an algorithm, and present important results from the literature.
+We motivate this model and discuss its application to and beyond Cognitive Radio (CR).
+%
+% We formalize this MAB model, then we quote some of its important applications.
+We define the notion of regret of an algorithm, and present important results of regret lower-bound from the literature, which quantify what no algorithm can achieve.
 A short review of the most important families of MAB algorithms is then given,
-and we highlight some algorithms that are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling.
+and we highlight some algorithms that are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling, which all achieve logarithmic regret upper-bounds for our main target of interest, that are binary distributed problems.
 %
@@
 We gave an overview of the most well known families of MAB algorithms,
-focussing on algorithms designed and analyzed for the single-player stationary and stochastic MAB model.
-Except the naive and simple strategies given as introductory examples, most of the policies presented above are order-optimal for at least one of the model of Bernoulli distributed problems, bounded rewards or one-dimensional exponential families,
-and some of them are known to be optimal in different settings.
-%
-To clearly understand the difference in their empirical behaviors, we illustrated the performance of two naive and two simple strategies against three efficient ones, on the example of Bernoulli-distributed problem used for the online demonstration from Section~\ref{sec:2:notations} (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
-This also serves as a first example of simulation performed with our library SMPyBandits, before the large-scale experiments of Section~\ref{sec:3:reviewSPAlgorithms} below.
-%
-In the rest of this thesis, we mainly use $\UCB_1$ or Thompson sampling when simplicity and realistic deployment is favored (in Chapter~\ref{chapter:4}), and \klUCB{} when we give mathematical developments (in Chapters~\ref{chapter:5} and \ref{chapter:6}).
+designed and analyzed for the single-player case of the stationary and stochastic bandit model.
+Except the naive and simplest strategies given as introductory examples, most of the policies presented above are order-optimal for at least one of the models of binary or bounded rewards, or one-dimensional exponential families,
+and some of them are known to be asymptotically optimal in different settings.
+%
+% To clearly understand the difference in their empirical behaviors,
+We illustrated the performance of two naive, two simple and three efficient strategies, on the example of Bernoulli-distributed problem used for the online demonstration from Section~\ref{sec:2:notations}.
+%  (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
+This also serves as a first example of simulation performed with our library SMPyBandits, before its presentation in the next chapter.
+% the large-scale experiments of Section~\ref{sec:3:reviewSPAlgorithms} below.
+%
+In the rest of this thesis, we mainly use $\UCB_1$ or Thompson sampling when simplicity and realistic deployment is favored in Chapter~\ref{chapter:4}, and \klUCB{} when we focus on mathematical developments in Chapters~\ref{chapter:5} and \ref{chapter:6}.
 
@@
 
-In this chapter, we presented the multi-armed bandit model, focussing on a finite number of arms, and real-valued rewards.
-Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
-By showcasing an interactive demonstration made available online,
+In this chapter, we presented the multi-armed bandit model, by restricting to a finite number of arms and real-valued rewards.
+Our main focus is on binary or bounded rewards, or one-dimensional exponential families of distributions, and on stochastic and stationary problems.
+By showcasing a small interactive demonstration made available online,
 % XXX ?
@@
 % %
-We then used the small Bernoulli-distributed problem behind this demonstration
+We then used the small Bernoulli-distributed problem underlying this demonstration
 to present a small numerical simulation, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
@@
 % We take another approach in Chapter~\ref{chapter:4}, where the
-We generalize this MAB model in Chapter~\ref{chapter:4}, by studying decentralized learning of a large set of independent players, all having different activation times.
-This extension is significantly harder than the two previously evoked ones, and we were unfortunately unable to obtain any strong theoretical results under these loose hypotheses.
-This model is however more generic and as such it was found suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
-
+We generalize this MAB model in Chapter~\ref{chapter:4}, by studying decentralized learning of a large set of independent players, all having different random activation times.
+This extension is significantly too complicated to analyze, and we were unfortunately unable to obtain any strong theoretical results under such loose hypotheses.
+This model is however realistic, and as such it was found suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
+%
 We take a more formal approach in the rest of the thesis,
@@
 % first in Chapter~\ref{chapter:4} where we consider many XXX
-first by considering players facing a stationary problem in a decentralized way in Chapter~\ref{chapter:5},
-and then by considering a single player facing a non-stationary or a piece-wise stationary problem in Chapter~\ref{chapter:6}.
+first by considering $M \leq K$ devices facing a stationary environment, that is an IoT network of $K$ orthogonal frequency channels, in a decentralized way in Chapter~\ref{chapter:5},
+and then by considering a single device facing a non-stationary (piece-wise stationary) problem in Chapter~\ref{chapter:6}.
 %
 For both directions, we present natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results,
-for the two problems of stationary multi-players and piece-wise stationary bandits.
+for the two problems of stationary multi-players and single-player piece-wise stationary bandits.
 

### 2019-07-08 2-Chapters/2-Chapter/chapter25.tex
De meilleures transitions, j'ai retravaillé chaque abstract
Proof read abstract & conclusion of Ch2, Ch3, Ch4, Ch5

@@
 %
+The review of MAB algorithms given in the previous Chapter~\ref{chapter:2} showed that there are a large collection of different algorithms designed for different kinds of bandit problems.
 In this chapter, we discuss the question of online algorithm selection.
@@
 Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, another solution could be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We propose an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+We propose an extension of the \ExpFour{} algorithm for this problem, that we called \Aggr, and illustrate its performance on some bandit problems.
 

### 2019-07-08 2-Chapters/3-Chapter/chapter3.tex
Proof read abstract & conclusion of Ch2, Ch3, Ch4, Ch5

@@
 %
-SMPyBandits is a Python package I have developed since the beginning of my PhD.
+\SMPyBandits{} is a library I have developed in the Python language since the beginning of my PhD.
 It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
+Its name stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
 To the best of the authors' knowledge, it is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
 %
-This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
-Then we use it to perform a numerical comparisons of the main state-of-the-art single-player bandit algorithms, and a study to compare time and memory costs of some of the most well-known and most widely used MAB algorithms.
+This chapter details the organization of the library and what it implements in terms of arm distributions, models, algorithms and visualizations.
+Then we use it to perform a numerical comparison of the main state-of-the-art single-player bandit algorithms, and a study to compare time and memory costs of some of the most well-known and most widely used MAB algorithms.
 
@@
 
-% This chapter is intended as a long version of the paper presenting SMPyBandits that I wrote in Summer 2018, see https://hal.inria.fr/hal-01840022
-
-% SMPyBandits is the most complete open-source implementation of state-of-the-art algorithms and models of Multi-Armed Bandits.
+% This chapter is intended as a long version of the paper presenting \SMPyBandits{} that I wrote in Summer 2018, see https://hal.inria.fr/hal-01840022
+
+% \SMPyBandits{} is the most complete open-source implementation of state-of-the-art algorithms and models of Multi-Armed Bandits.
 % It aims at being extensive, simple to use and maintain, with a clean and perfectly documented codebase.
@@
 
-Since December $2016$ I worked regularly on the SMPyBandits library, which is a Python package  designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
+Since December $2016$ I worked regularly on the \SMPyBandits{} library, which is a Python package  designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
 %
@@
 
-SMPyBandits stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
+\SMPyBandits{} stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
 The library does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a Python implementation \cite{python}, and uses only standard open-source Python packages (\ie, no hard to install dependencies).
@@
 and that for theoretical works where optimality of the MAB algorithm is analyzed, we advise to use \klUCB{} (as we do in the next Chapters~\ref{chapter:5} and \ref{chapter:6}).
-% To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
+% To also illustrate how one can easily extend the \SMPyBandits{} library to add not only add new distributions or new algorithms.
 % but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
@@
 %
-The code for SMPyBandits is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source License.
+The code for \SMPyBandits{} is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source License.
 This chapter is based on the article \cite{SMPyBanditsJMLR}.
@@
 
-% I want to show what is SMPyBandits, what problems does it solve, what did I implement, how to use it.
-
-We start by explaining how SMPyBandits simulates MAB problems, by detailing its components:
+% I want to show what is \SMPyBandits, what problems does it solve, what did I implement, how to use it.
+
+We start by explaining how \SMPyBandits{} simulates MAB problems, by detailing its components:
 MAB problems (single- or multi-players),
@@
 
-SMPyBandits implements all the collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.html}{PoliciesMultiPlayers}}).
+\SMPyBandits{} implements all the collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.html}{PoliciesMultiPlayers}}).
 %
@@
     \end{minted}
-    \caption{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits.}
+    \caption{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with \SMPyBandits.}
     % \captionof{listing}{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits.
@@
 
-We have not yet added support for higher dimensional distributions of rewards, such as linear bandits, but it would be an interesting extension of SMPyBandits.
+We have not yet added support for higher dimensional distributions of rewards, such as linear bandits, but it would be an interesting extension of \SMPyBandits.
 %
@@
 
-SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
+\SMPyBandits{} is a complete open-source implementation of single-player (classical) bandit algorithms,
 containing over 65 algorithms (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}).
@@
 
-We show how to install SMPyBandits, and an example of how to run a simple experiment.
+We show how to install \SMPyBandits, and an example of how to run a simple experiment.
 See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.
-SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}, you can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.
+\SMPyBandits{} is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}, you can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.
 This bash code shows how to clone the code, and install the requirements for Python 3 (once):
@@
         \end{minted}
-        \caption{Example of Bash code to download and install dependencies of SMPyBandits.}
-        % \captionof{listing}{Small snippet of Bash code to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
+        \caption{Example of Bash code to download and install dependencies of \SMPyBandits.}
+        % \captionof{listing}{Small snippet of Bash code to download and install dependencies of \SMPyBandits{} \label{lst:3:howToInstallLibrary}.}
         \label{lst:3:howToInstallLibrary}
@@
     \end{minted}
-    \caption{Example of Bash code to run a simple experiment with SMPyBandits.}
-    % \captionof{listing}{Small snippet of Bash code to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
+    \caption{Example of Bash code to run a simple experiment with \SMPyBandits.}
+    % \captionof{listing}{Small snippet of Bash code to run a simple experiment with \SMPyBandits{} \label{lst:3:howToRunBasicLibrary}.}
     \label{lst:3:howToRunBasicLibrary}
@@
 %
-When SMPyBandits is installed using \texttt{pip install SMPyBandits}, the dependencies are of course automatically installed if not already present.
+When \SMPyBandits{} is installed using \texttt{pip install SMPyBandits}, the dependencies are of course automatically installed if not already present.
 
@@
 
-Since $2017$ and $2018$, SMPyBandits is using two online continuous integration (CI) services, to automatically build and host its documentation, and to automatically test the library on some numerical simulations.
+Since $2017$ and $2018$, \SMPyBandits{} is using two online continuous integration (CI) services, to automatically build and host its documentation, and to automatically test the library on some numerical simulations.
 %
@@
 Since its first use, the service did about 180 builds, and about $10\%$ of them were useful to detect newly introduced issues or bugs in the code.
-The build is configured using the \texttt{.readthedocs.yml} file in SMPyBandits main folder, and it builds the documentation in about two minutes.
+The build is configured using the \texttt{.readthedocs.yml} file in \SMPyBandits{} main folder, and it builds the documentation in about two minutes.
 I have also been building the documentation manually on a weekly basis, to also host it on \href{https://smpybandits.github.io/}{\texttt{SMPyBandits.GitHub.io}} thanks to GitHub pages.
@@
 in order to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
-The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
+The tests covers all the main models and almost all the algorithms implemented in \SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
 For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
@@
 Since its first use, the service did about $190$ builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
-The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
+The build is configured using the \texttt{.travis.yml} file in \SMPyBandits{} main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by \SMPyBandits{} (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
 Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.
 
-\subsection{List of research works using SMPyBandits}
+\subsection{List of research works using \SMPyBandits}
 \label{sub:3:listResearchWorksUsingSMPyBandits}
 
-SMPyBandits has been used for the following research articles since $2017$, and it is used for the previous and the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
+\SMPyBandits{} has been used for the following research articles since $2017$, and it is used for the previous and the next chapters of this thesis (except in Chapter~\ref{chapter:4}).
 %
@@
     \item
-In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used SMPyBandits to illustrate and compare different aggregation algorithms.
+In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2} above, we used \SMPyBandits{} to illustrate and compare different aggregation algorithms.
 We designed a variant of the \ExpFour{} algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
@@
     \item
-In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-players bandit algorithms.
+In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used \SMPyBandits{} for all the simulations for multi-players bandit algorithms.
 We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they obtain logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).
@@
     \item
-In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes.
+In \cite{Besson2018DoublingTricks}, we used \SMPyBandits{} to illustrate and compare different "doubling trick" schemes.
 \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}}
@@
     \item
-In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case).
+In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6} below, we used \SMPyBandits{} for piece-wise stationary MAB models (only for the single-player case).
 We illustrate and compare different algorithms designed for this family of non-stationary problems.
@@
 
-In this section, we use the SMPyBandits library to compare experimentally various state-of-the-art (single-player) algorithms on some MAB problems.
+In this section, we use the \SMPyBandits{} library to compare experimentally various state-of-the-art (single-player) algorithms on some MAB problems.
 We first detail the list of different algorithms that we compare.
@@
 %
-For reproducibility, we also give in ``\texttt{typewriter font}'' the name of the corresponding class in the \texttt{Policies} module in SMPyBandits
-(\eg, $\UCB_1$ is \texttt{Policies.UCBalpha}).
+For reproducibility, we also give in ``\texttt{typewriter font}'' the name of the corresponding class in the \texttt{Policies} module in \SMPyBandits{} (\eg, $\UCB_1$ is \texttt{Policies.UCBalpha}).
 
@@
     % It was also proven to achieve asymptotical optimal problem-dependent bounds \cite{Kaufmann12Thompson,AgrawalGoyal11}.
-    It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for SMPyBandits, as we use \texttt{numpy.random} module which relies on highly optimized C or Cython code).
+    It is efficient in terms of storage, and even if $K$ random samples must be sampled from the arms posteriors at each time step $t$, Thompson sampling is not much slower than \UCB, if the random number generator used for the simulations is efficient (it is the case for \SMPyBandits, as we use \texttt{numpy.random} module which relies on highly optimized C or Cython code).
 
@@
 We focus on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
-To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
+To highlight that \SMPyBandits{} implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
 They are more recent (\eg, $3$ from $2018$ and $2$ from $2019$) and are usually more complex, or based on a different point-of-view.
@@
 one can reasonably aim at the simplest yet order-optimal algorithms, and we advise to use \UCB{} and algorithms running on top of this simple index policy, as we do in Chapter~\ref{chapter:4}.
-On the other hand, if one is more interested in the mathematical developments and wants to prove the tightest possible regret upper bounds, aiming at more efficient but more complex algorithms is interesting, and we chose the \klUCB{} algorithm in Chapters~\ref{chapter:5} and \ref{chapter:6}.
+On the other hand, if one is more interested in the mathematical developments and wants to prove the tightest possible regret upper bounds, aiming at more efficient but more complex algorithms is interesting, and we chose the \klUCB{} algorithm as the base bandit policy for our works in Chapters~\ref{chapter:5} and \ref{chapter:6}.
 
@@
 
-% SMPyBandits also implemented sparse MAB models (with $s<K$ arms with a mean $\mu_i>0$ and $K-s$ arms with a mean $\mu_i \leq 0$), I could also present this model, the various algorithms designed to tackle it, and some numerical experiments.
+% \SMPyBandits{} also implemented sparse MAB models (with $s<K$ arms with a mean $\mu_i>0$ and $K-s$ arms with a mean $\mu_i \leq 0$), I could also present this model, the various algorithms designed to tackle it, and some numerical experiments.
 
@@
 
-We chose to present here the Python library SMPyBandits in this chapter, rather than in the Appendix, because all the numerical experiments on single-player multi-arm bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on SMPyBandits.
+We chose to present here the Python library \SMPyBandits{} in this chapter, rather than in the Appendix, because all the numerical experiments on single-player multi-arm bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on \SMPyBandits.
 %
@@
     \item[$(i)$]
-Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
-SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
+The purpose of \SMPyBandits{} is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
+\SMPyBandits{} is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
 Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.
@@
     \item[$(ii)$]
-We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
-As far as now, SMPyBandits is restricted to the finite-arm case, but it supports a wide range of arms distributions.
+We detailed how \SMPyBandits{} is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
+As far as now, \SMPyBandits{} is restricted to the finite-arm case, but it supports a wide range of arms distributions.
 Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
 One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we followed the active research from December $2016$ to June $2019$.
-More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
+More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
 All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
 %
-When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate, mean cumulated reward, real time or memory costs.
+When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate, mean cumulated reward, as well as real time or memory costs.
 
     \item[$(iii)$]
-Finally, we presented how to use SMPyBandits, if one wants to run some pre-designed simulations or design new simulations.
+Finally, we presented how to use \SMPyBandits, if one wants to run some pre-designed simulations or design new simulations.
 Running a simulation is very easy, and different examples showed that the main parameters such as the time horizon $T$ or the number of repetitions can be configured directly from the command line when running the Python script, or by modifying the code.
 Designing a new simulation requires to have a basic knowledge of Python, but not to dive into the implementation of the library.
-We give below in Appendix~\ref{sec:3:appendix} a small but complete example of the two files that has to be written for a new simulation: a \texttt{main} file which essentially defines the desired visualizations after the simulation, and a \texttt{configuration} file which defines the problems to simulate and their parameters, as well as the algorithms that will be compared.
+% We give below in Appendix~\ref{sec:3:appendix} a small but complete example of the two files that has to be written for a new simulation: a \texttt{main} file which essentially defines the desired visualizations after the simulation, and a \texttt{configuration} file which defines the problems to simulate and their parameters, as well as the algorithms that will be compared.
 \end{itemize}
@@
 % \paragraph{Remark.}
-We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of SMPyBandits, and as such the library, its documentation and this chapter  are considered an important contribution of this thesis.
-%
-Our library is used for the numerical simulations in all this thesis, except Chapter~\ref{chapter:4}.
+We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of \SMPyBandits, and as such the library, its documentation and this chapter  are considered an important contribution of this thesis.
+%
+Our library is used for the numerical simulations in the rest of this document, except Chapter~\ref{chapter:4}.
 % We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
@@
 %
-% An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
+% An interesting future work left on our library \SMPyBandits{} could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
 % For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
@@
 
-The most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
-Its documentation has seen already about $20000$ visits, the project on GitHub had $88$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used SMPyBandits.
+The most exciting thing that could happen to our \SMPyBandits{} library would be to see it gaining popularity!
+Its documentation has seen already about \href{https://github.com/SMPyBandits/SMPyBandits/graphs/traffic}{$20000$ visits}, \href{https://github.com/SMPyBandits/SMPyBandits/stargazers}{the project on GitHub had $90$ stars} in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used \SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
 %
-We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to this work.
-%
-As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
-I would also love to be able to teach a graduate or under-graduate course on multi-armed bandits in the next years, and use my library as a support to illustrate the course, as well as practical sessions.
+We have submitted a summary paper presenting \SMPyBandits{} \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to this work.
+%
+As a personal note, I would like to continue working on \SMPyBandits, and implement the most interesting requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
+I would also love to be able to teach a graduate or under-graduate course on reinforcement learning and multi-armed bandits in the next years, and use my library as a support to illustrate such course, as well as practical sessions.
 
@@
 
-% We start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
+% We start by giving a small but complete example of use of \SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
 We give in this appendix more details regarding the numerical experiments presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above,
@@
     Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{~Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
-Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
+Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in \SMPyBandits{} will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
     \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as \KLUCBswitch), and costs an-order-of-magnitude more memory.
@@
 % How to write a fast version of klBern and klucbBern (the two most used functions), using C or Numba or Cython, maybe I can include the speed up simulations here?
-First piece of code, before really presenting SMPyBandits ?
+First piece of code, before really presenting \SMPyBandits{} ?
 
@@
     \emph{Quality and optimization of the code:}
-    The same level of optimization is used in all the codebase of SMPyBandits, and most of the time, the code is pure and naive Python and was not optimized.
+    The same level of optimization is used in all the codebase of \SMPyBandits, and most of the time, the code is pure and naive Python and was not optimized.
     Mathematical functions (\eg, $\sqrt{\bullet}, \log$) and random numbers used are using \texttt{numpy} and \texttt{numpy.random} modules \cite{numpy}, which are based on C and Cython code \cite{cython} and can be considered highly efficient.
     Computations of Kullback-Leibler and indexes for \klUCB{} algorithms are based on a compiled CPython extension written in C \cite{python} and can also be considered highly efficient.
-    As we can safely affirm that the code of the different algorithms has the same level of optimization, the comparison is fair, and we do not favor any algorithm in their implementation in SMPyBandits.
+    As we can safely affirm that the code of the different algorithms has the same level of optimization, the comparison is fair, and we do not favor any algorithm in their implementation in \SMPyBandits.
 
@@
     \emph{About the experimental environment:}
-    the exact measurements used for the figures displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
+    the exact measurements used for the figures displayed below highly depend on the machine used for the simulation, the version of the language and its libraries (see above in Section~\ref{sub:3:dependencies} for details about \SMPyBandits), as well as the number of CPU cores being used (here, we used $12$ cores in a $12$-core desktop) etc.
     As long as the different algorithms and simulations are performed on the same environment, the comparison we make from them are fair and do make sense.
@@
     \emph{For time}, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
-    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in[T]$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
+    In the \texttt{Evaluator} object of \SMPyBandits, before launching the ``for'' loop on $t\in[T]$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
     This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.
@@
 % Nowadays, parallelism is everywhere in the computational world, and any serious framework for numerical simulations must try to gain performance from parallelism.
-% In this last section, we quickly review the chosen approach for SMPyBandits (multi-core on one machine).
+% In this last section, we quickly review the chosen approach for \SMPyBandits{} (multi-core on one machine).
 
@@
 % %
-% As explained in the page \href{https://smpybandits.github.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}, we implemented in SMPyBandits an easy way to run any simulations on $n$ cores of a machine, using the \texttt{Joblib} \cite{joblib} library.
+% As explained in the page \href{https://smpybandits.github.io/How_to_run_the_code.html}{\texttt{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}}, we implemented in \SMPyBandits{} an easy way to run any simulations on $n$ cores of a machine, using the \texttt{Joblib} \cite{joblib} library.
 
@@
 
-% % $(i)$ I did not try to add in SMPyBandits the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
+% % $(i)$ I did not try to add in \SMPyBandits{} the possibility to run simulations using a GPU, or any general purpose computation libraries offering a GPU-backend.
 % % Initially designed for graphical simulations and mainly for video-games applications, the use of GPU for scientific computations have been gaining attention for numerical simulation in the research world since the last $15$ years, and NVidia CUDA for GPGPU (General Purpose GPU) started to become popular in $2011$.

### 2019-07-08 2-Chapters/4-Chapter/chapter4.tex
Proof read abstract & conclusion of Ch2, Ch3, Ch4, Ch5

--- 2019-07-06 2-Chapters/4-Chapter/chapter4.tex
@@
 %
-In this chapter, we propose two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
-that can use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
+After detailing in Part~\ref{part:Introduction} the MAB model, we now come back to our main question of interest.
+In this chapter, we focus on wireless networks following the hypotheses common to present and future Internet of Things (IoT) networks exposed in Chapter~\ref{chapter:1}.
+Our goal is to show that the IoT devices can automatically learn to increase their battery life and their successful transmission rates, without changing anything on the IoT standard side.
 %
-Decentralized RL for IoT lets the devices use acknowledgements sent by their Base Station as a reward, instead of sensing feedback like for OSA.
-We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
-Simulations show that dynamic devices can improve their spectrum efficiency,
+We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost Reinforcement Learning (RL) algorithms to learn to improve their spectrum access.
+%
+Decentralized RL for IoT lets the devices use acknowledgements sent back by their Base Station as a reward, instead of sensing feedback like for OSA.
+We consider many independent ``dynamic'' devices, each communicating with a small probability at every instant.
+Simulations show that dynamic devices can greatly improve their spectrum efficiency,
 by using MAB algorithms like \UCB.
@@
 %
-However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
-We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics significantly improve the network efficiency.
+However, our second model considers that in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
+We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics also significantly improve the network efficiency.
 % Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
@@
 
-To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices of the IoT network.
-We presented two models of wireless IoT network without sensing, inspired by the ALOHA protocol, and with or without retransmission of uplink packages in case of collisions.
+To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
+We presented two models of wireless IoT networks, without relying on the feedback provided by spectrum sensing, and inspired by the ALOHA protocol.
+We proposed two versions, with or without retransmissions of uplink packages in case of collisions.
 % In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
@@
 It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
-Unfortunately, analyzing the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales, turned out to be of extreme difficulty.
-That is why we focus on two different simplifications of this model in the next chapters, for which we are able to provide clean theoretical analyses.
+Unfortunately, it turned out to be of extreme difficulty to analyze the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
+That is why we focus on two different simplifications of this model in the next chapters, for which we are able to provide a clean theoretical analysis.
 % %
@@
 %
-The main difference between the two models is the fact that in this chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
+The main difference between the two models is the fact that in Chapter~\ref{chapter:4}, $M \gg K$ devices transmit their messages at every time step, by following a random activation process (with a fixed transmission probability $p < 1$).
 If static and dynamic devices that have to transmit at a fixed time are denoted \emph{active devices},
@@
 %
-We study this first direction in the next Chapter~\ref{chapter:5}.
+We study the case of $M \leq K$ devices learning independently to play a $K$-armed bandit in the next Chapter~\ref{chapter:5}.
 
@@
 %
-On the other hand, while it is hard to analyze the models of this chapter due to their non-stationary behaviors, we are also able to analyze a simpler model, if we focus on a single player accessing a network which is assumed to be \emph{piece-wise stationary}
-(MAB problems which are stationary on ``long enough'' intervals, of unknown locations and lengths).
+On the other hand, while it is hard to analyze the models of this chapter because of the unpredictable behaviors of the IoT devices' activation patterns and the evolving number of active devices, we are also able to analyze a simpler model, if we focus on a single player accessing a network which is assumed to be \emph{piece-wise stationary},
+that are bandit problems which are stationary on ``long enough'' intervals, of unknown locations and lengths.
 %
-We also study this second direction in Chapter~\ref{chapter:6}.
+We then study this second direction in Chapter~\ref{chapter:6}.
 

### 2019-07-08 2-Chapters/3-Chapter/chapter3.tex
Je fini Ch7 et conclusion demain, je suis épuisé

@@
 I have also been building the documentation manually on a weekly basis, to also host it on \href{https://smpybandits.github.io/}{\texttt{SMPyBandits.GitHub.io}} thanks to GitHub pages.
-In the last two years, the documentation saw about 11000 unique visits, while the GitHub project have been visited about 4500 unique visits.
+In the last two years, the documentation saw about 20000 unique visits, while \href{https://github.com/SMPyBandits/SMPyBandits/graphs/traffic}{the GitHub project have been visited about 4500 unique visits}.
 
@@
 The most exciting thing that could happen to our \SMPyBandits{} library would be to see it gaining popularity!
-Its documentation has seen already about \href{https://github.com/SMPyBandits/SMPyBandits/graphs/traffic}{$20000$ visits}, \href{https://github.com/SMPyBandits/SMPyBandits/stargazers}{the project on GitHub had $90$ stars} in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used \SMPyBandits.
+Its documentation has seen already about $20000$ visits, \href{https://github.com/SMPyBandits/SMPyBandits/stargazers}{the project on GitHub had $90$ stars} in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used \SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.

### 2019-07-08 2-Chapters/5-Chapter/chapter5.tex
Je fini Ch7 et conclusion demain, je suis épuisé

@@
 %
-In this chapter, we are interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
-We restrict to the easier case of at most $M \leq K$ devices in a network with $K$ channels, that each have data to transmit at every instant, that is they have an activation probability of $p=1$, because it was found very hard to analyze the previously studied IoT network models.
+In this chapter, we are interested in a more formal approach to the decentralized learning problem presented of many end-devices accessing the same IoT wireless network
+in the previous Chapter~\ref{chapter:4}.
+We restrict to the easier case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
+Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$, because it was found very hard to analyze the previously studied IoT network models, under the hypothesis of random activation patterns of Chapter~\ref{chapter:4}.
 %
-We discuss three feedback levels that give variants of the multi-players MAB model previously studied in the literature.
-Using a decomposition of the centralized system regret, we start by explaining the intuition about what an efficient decentralized algorithm should do, and then we propose \RandTopM{} and \MCTopM, two new orthogonalization schemes.
-Combining them with the \klUCB{} index policy gives order-optimal algorithms, that achieve state-of-the-art performance,
-as illustrated by numerical simulations.
+We discuss three feedback levels that give variants of the multi-players MAB model previously studied in the literature, and we define the centralized regret as a measure of performance of multi-player bandits algorithms.
+%
+Using a decomposition of the centralized system regret, we start by explaining the intuition about the expected behavior of any efficient decentralized algorithm, and then we propose \RandTopM{} and \MCTopM, two new orthogonalization schemes.
+Combining them with the \klUCB{} index policy gives efficient algorithms, that achieve state-of-the-art performance in terms of their regret,
+which is also backed up by numerical simulations.
 We conclude by reviewing different extensions of the multi-players models.
@@
 Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
-However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (while scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).
+However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (which scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).
 
 Moreover, at first sight, this idea of ``meta'' arms implies an exponential blow up in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
-the authors provide in Section~4.1 and Lemma~4.1 \cite{AlaturLevyKrause19} an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} for a good review).
+the authors provide in Section~4.1 and Lemma~4.1 \cite{AlaturLevyKrause19} an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} and references therein for a good introduction).
 %
@@
 The two easiest models are the ones with sensing information (\ie, for OSA), for which our theoretical contribution improves both the state-of-the-art upper and lower bounds on the regret.
-In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
-%
-We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proven to be very active in the last two years.
+In the absence of sensing, we also provide some motivation for the practical use of the interesting \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking into account the collision information, like in Chapter~\ref{chapter:4}.
+
+We also reviewed various variants of this model and we discussed the related literature, which has proven to be very active in the last two years.
 For some models, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
-For example, it assumes to know the number of players $M$ before-hand, but we illustrated that previously introduced technique to estimate $M$ can also be applied to our proposal and give satisfactory empirical performances.
-Further works would be required to adapt the theoretical analysis to these various extensions.
+For example, it assumes to know the number of players $M$ before-hand, but we illustrated that a previously introduced technique to estimate $M$ on the run can also be applied to our proposal and give satisfactory empirical performances.
+Further works would be required to adapt the theoretical analysis to these various extensions, such as the ``dynamic'' case which allows arrivals or departures of devices, and which is especially interesting for Cognitive Radio applications.
 

### 2019-07-09 2-Chapters/2-Chapter/chapter2.tex
J'ai la conclusion qui dépasse sur une page… zut

@@
 
-\chapter{The Stochastic Multi-Armed Bandits Model}
+% \chapter{The Stochastic Multi-Armed Bandits Model}
+\chapter{Stochastic Multi-Armed Bandits}
 \label{chapter:2}
@@
 % We conclude this chapter by discussing the question of online algorithm selection.
+
+\begin{small}
+    \begin{flushright}
+        \emph{-- C'est cool ça, des bandits haha !} Le Barbare,
+        \emph{Le Donjon de Naheulbeuk}\\
+        Saison III, \href{https://soundcloud.com/penofchaos/naheulbeuk-saison-3-episode-5?in=penofchaos/sets/naheulbeuk-saison-3#t=2:05}{Épisode 37 : ``La stratélique''}.
+        écrit par John Lang (Pen of Chaos).
+    \end{flushright}
+\end{small}
 

### 2019-07-09 2-Chapters/5-Chapter/chapter5.tex
J'ai la conclusion qui dépasse sur une page… zut

@@
 
-And finally, even though numerical simulations in their papers \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} are interesting and confirm the regret upper-bounds, they do not compare with any other algorithm, and it is left as an interesting future work to study this direction in more details.
+And finally, even though numerical simulations in their papers \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} are interesting and confirm the regret upper-bounds, they do not compare with any other algorithm, and thus a future work can be to study this direction in more details.
 Sadly, a major drawback of their work is that assume that player $j\in[M]$ knows its index $j$, and as we explained above this small hypothesis is actually quite strong, as it allows players to be already orthogonal, and it reduces greatly the difficulty of the decentralized bandit problem.

### 2019-07-09 2-Chapters/6-Chapter/chapter6.tex
J'ai la conclusion qui dépasse sur une page… zut

--- 2019-07-03 2-Chapters/6-Chapter/chapter6.tex
@@
 % \chapter{The Piece-wise Stationary MAB Model and the GLR-klUCB Algorithm}
-\chapter{Piece-Wise Stationary Multi-Armed Bandits}
+% \chapter{Piece-Wise Stationary Multi-Armed Bandits}
+\chapter{Piece-Wise Stationary Bandits}
 \label{chapter:6}
@@
 %
-In this last chapter, we study a generalization of the stationary multi-armed bandit model,
+In this last chapter, we are also interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
+Instead of keeping the stationary hypothesis but considering $2 \leq M \leq K$ devices accessing a wireless networks with $K$ orthogonal frequency channels, as we did in Chapter~\ref{chapter:5}, we are now interested in another direction of formal analysis of the intractable models of IoT networks of Chapter~\ref{chapter:4}.
+%
+We study a generalization of the single-player stationary bandit model,
 to account for possible non-stationarity of the rewards.
@@
 %
-We review existing works on piece-wise stationary MAB models.
+We review existing works on piece-wise stationary MAB models,
 % for which the two main families of proposed solutions are either passively or actively adaptive.
@@
 %
-We study the Generalized Likelihood Ratio Test (GLRT) for
+then we study the Generalized Likelihood Ratio Test (GLRT) for
 % Using finite-time results of a recent paper studying the GLRT for sub-Gaussian distributions, we then extend the results to the GLRT for any
 bounded or sub-Bernoulli distributions.
-Our test enjoys finite-time guarantees
+Our test enjoys finite-time guarantees,
 % its \emph{false alarm probability}, meaning that in a stationary model, the test should not detect any change,
@@
 % We give the first unified analysis of a piece-wise stationary bandit algorithm for the two variants, and our analysis
-We analyze its regret, and show that it achieves state-of-the-art non-asymptotic regret bounds.
+We analyze its regret, and show that it achieves state-of-the-art finite-time regret bounds.
 % and yield regret bounds with explicit constants.
@@
 %
-Finally, we compare our approach with other state-of-the-art solutions on different numerical experiments.
+Finally, we showcase numerical experiments on which our approach outperform other state-of-the-art solutions.
 % , on different piece-wise stationary bandit problems.

### 2019-07-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
De meilleures transitions, j'ai retravaillé chaque abstract
J'ai la conclusion qui dépasse sur une page… zut

@@
 %-----------------------------------------------------------------------------
-\section{The \GLRklUCB{} algorithm}
+% \section{The \GLRklUCB{} algorithm}
+\section{A new algorithm for piece-wise stationary bandits}
 \label{sec:6:GLRklUCB_Algorithm}
@@
 
-In this chapter, we studied and presented the piece-wise stationary bandit model, and reviewed existing work.
+In this chapter, we studied and presented the piece-wise stationary bandit model, starting by reviewing existing works.
 %
 We proposed a new algorithm for this problem, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
-This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but \emph{without any other prior knowledge on the problem}, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
+This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but \emph{without any other prior knowledge on the problem}, unlike its best two competitors, \CUSUMUCB{} and \MUCB, that require to know a lower bound on the smallest magnitude of a change.
 We also gave numerical evidence of the efficiency of our proposal.
@@
 
-As mentioned at the end of Chapter~\ref{chapter:5}, an interesting direction of future work is to study non-stationary distributed multi-players bandits.
-A natural extension of the non-stationary model presented in this Chapter is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
+This chapter was an interesting direction of analysis of the intractable model from Chapter~\ref{chapter:4}, of independent end-devices running embedded MAB learning algorithms to improve their spectrum access in an IoT network.
+Instead of dealing with the multi-players aspect, as we did in the previous Chapter~\ref{chapter:5}, we dealt in Chapter~\ref{chapter:6} about the non-stationary aspect.
+Different kinds of non-stationary bandit models have been considered in the literature, including adversarial, slowly-varying or abruptly-changing bandits, and as found the later to be the most appropriate to model non-stationarity of cognitive radio networks, we focus on abruptly-changing or piece-wise stationary bandit problems.
+%
+Our proposed algorithm achieves state-of-the-art results for this model, by requiring a weaker prior knowledge on the problem difficulty when compared with its best competitors.
+An important future work is to investigate whether actively adaptive approaches can attain an order-optimal $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$ the number of break-point.
+%
+Another possible future work is to propose an algorithm that could blindly adapt to both the slowly-varying and the abruptly-changing models, as well as being efficient against stationary models.
+
+As suggested by the ``dynamic case'' of multi-players bandits presented in Chapter~\ref{chapter:5}, another promising direction of future work is to study non-stationary distributed multi-players bandits.
+A natural extension of the model presented in this chapter is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
 The authors build on their recent work \cite{WeiSrivastava18Abruptly} and show that a regret bounded by $\cO(\sqrt{T^{\frac{1+\nu}{2}} \log(T)})$ can still be achieved by $M$ players, in the number of breaking points is $\Upsilon_T = \cO(T^{\nu})$.

### 2019-07-09 2-Chapters/Conclusion/conclusion.tex
Je vais au bureau, une dernière relecture de résumé / intro fr / intro anglais / conclusion et ça ira
J'ai la conclusion qui dépasse sur une page… zut

@@
 
-This manuscript was organized in two parts, Part~\ref{part:Introduction} and Part~\ref{part:MABIOT}.
+This manuscript was organized in two parts.
 %
@@
 %
-We detailed the historical, scientific and technical contexts of the problems we considered for our research, and presented the main contributions produced during the last three years.
+We started by detailing the historical, scientific and technical contexts of the problems we considered for our research, and presented the main contributions produced during the last three years.
 
@@
 % Chapter 2
-In Chapter~\ref{chapter:2}, we presented the multi-armed bandit models, the hypotheses we usually make, and the most important algorithms that solve different kinds of MAB problems.
+In Chapter~\ref{chapter:2}, we presented the multi-armed bandit (MAB) models, the hypotheses we usually make on such models, and the most important algorithms that solve different kinds of MAB problems.
 % Chapter 3
-Then in Chapter~\ref{chapter:3} we discussed in details about our open-source Python library, SMPyBandits, developed to run numerical simulations of MAB problems.
+Then in Chapter~\ref{chapter:3} we showcased our open-source Python library SMPyBandits in details, developed to run numerical simulations of MAB problems.
 %
@@
 % Chapter 4
-In Chapter~\ref{chapter:4}, we first proposed some models for realistic IoT wireless networks, in discrete times and with no central coordination between end-devices.
-In two different models, with or without retransmissions, we showed that end-devices can independently learn to increase their successful transmission rates in the network, by using low-cost MAB algorithms.
+In Chapter~\ref{chapter:4}, we first proposed models for realistic IoT wireless networks, in a discrete time and with no central coordination between end-devices.
+In two different models, with or without retransmissions, we showed that end-devices can independently learn favor the less occupied radio channels, in order to increase their successful transmission rates in the network, by embedding low-cost decentralized MAB algorithms.
 We also presented a proof-of-concept that implements such network on real radio hardware and validates the proposed model and algorithms.
 % Chapter 5
-Because the models proposed for realistic IoT networks were found to be intractable to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
-in Chapter~\ref{chapter:5}, we studied a weaker but still interesting model.
-We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision information (\MCTopM-\klUCB).
+Because the models proposed for realistic IoT networks were found to be too general to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
+we started to study a weaker but still interesting model in Chapter~\ref{chapter:5}.
+We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision and spectrum sensing information (\MCTopM).
 % Chapter 6
@@
 %
-For the last two chapters, the proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in our theoretical analyses.
+For the last two chapters, the proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in their theoretical analyses.
 % \end{itemize}
 
+\paragraph{Summary of each chapter.}
 % A more precise summary of each chapter is given below.
@@
 This library allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
-SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license (MIT License), and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
-% Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.
-
+SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license (MIT License), and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).\\
+% Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.\\
+\indent
 We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
@@
 %
-Our library is used to run numerical simulations in Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
+Our library is used to run numerical simulations in Chapters~\ref{chapter:2}, \ref{chapter:3}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
 
@@
 %
-This first part ends by presenting one of our first contributions \cite{Besson2018WCNC}, in Chapter~\ref{chapter:25}.
+This first part ends with Chapter~\ref{chapter:25} by presenting one of our first contributions \cite{Besson2018WCNC}.
 We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
 Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for a practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We proposed an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
+We proposed an extension of the \ExpFour{} algorithm for this problem, that we named \Aggr, and illustrate its performance on some bandit problems.
 
@@
 %
-First in Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of applying MAB learning algorithms in IoT networks,
-with a focus on the convergence of algorithms, in terms of successful transmission rates.
+First in Section~\ref{sec:4:firstModel}, we proposed a model of IoT networks,
+and an evaluation of the performance of applying low-cost and low-complexity decentralized MAB learning algorithms in such, directly on the end-devices sides.
+Our focus is the convergence of such algorithms, and the gain they can provide for the IoT devices, in terms of successful transmission rates.
 We restrict to the case of a protocol slotted in both time and frequency, and we assume a certain clock pre-agreement and time synchronisation between devices and the base station.
 We studied the behavior of the network when the proportion of intelligent dynamic devices changes.
-Concretely, increasing this probability allows to insert more devices in the same network, while maintaining a good Quality of Service.
+Concretely, increasing their successful transmission rates allows to insert more devices in the same network, while maintaining a good Quality of Service.
 Another way of viewing this result is the opposite point-of-view: if the number of devices remains constant, increasing the successful transmission rate directly improves the network QoS.
@@
 %
-This was both surprising and encouraging, as it shows that applying bandit algorithms tailored for a stochastic model is still useful in broader settings.
+This was both surprising and encouraging, as it shows that applying bandit algorithms tailored for a stochastic model is still useful in our broader setting.
 The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.
-
+\\
+%
+\indent
 We presented in Section~\ref{sec:4:gnuradio} a proof-of-concept, demonstrated in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
 We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
-Results obtained in practice were discussed, to highlight the interest of using learning algorithms for cognitive radio online optimization problems, and especially multi-armed bandit learning algorithms.
+We discussed results obtained in practice, to highlight the interest of using learning algorithms for cognitive radio online optimization problems, and especially multi-armed bandit learning algorithms for IoT networks.
 %
-By using such low-cost algorithms, we confirmed empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using only the acknowledgement (\Ack) feedback sent from the base station.
-
+By using such low-cost algorithms, we confirmed empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using only the acknowledgement (\Ack) feedback as a collision binary indicator sent from the base station.
+\\
+%
+\indent
 Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of the previous model of Low-Power Wide-Area (LPWA) networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency, and assuming a perfect synchronisation.
@@
 %
-We showed that incorporating learning for the first transmissions can help to achieve optimal performance, with significant gains in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
-Our simulations showed that the proposed heuristics outperform a naive random access scheme.
+Like for the first model, we showed that incorporating learning for the first transmissions allows to achieve optimal performance, with significant gains in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
+Our simulations showed that the proposed heuristics greatly outperform a naive random access scheme.
 Surprisingly, we showed that a simple \UCB{} learning approach, that simply retransmits in the same channel, turns out to perform almost as well as more complicated heuristics, that can change the channel used for retransmissions.
-
+\\
+%
+\indent
 Thus in Chapter~\ref{chapter:4}, we proposed two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
-that can all use low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
+that can all use embedded low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
 %
-Decentralized RL for IoT lets the devices use the acknowledgement sent by their Base Station as a reward, instead of requiring some sensing capacity, like in the OSA case.
+Decentralized RL for IoT lets the devices use the acknowledgement sent by their Base Station as a reward, instead of requiring some spectrum sensing capacity, like in the OSA case.
 We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
@@
 %
-However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission (\eg, $10$ times).
+However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
 %
 The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
-
+%
 \TODOL{Je dois clarifier les conclusions des deux modèles, est-ce que l'un est vraiment mieux pour améliorer la QoS mais coûte en batterie, et l'autre améliore la batterie mais coûte en QoS ?
 Pas très clair en ayant la tête dans le guidon là !...}
+%
+%
+To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
+We presented two models of wireless IoT networks, without relying on the feedback provided by spectrum sensing, and inspired by the ALOHA protocol.
+% In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
+%
+It was also quite surprising to be able to conclude that this learning-based approach can be as efficient as we showed,
+% , as it allows the IoT devices to automatically and independently increase their successful transmission rates.
+because it is based on applying stochastic MAB algorithms in a non-stationary environment.
+Unfortunately, it turned out to be of extreme difficulty to analyze the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
+That is why we focus on two different simplifications of this model in the next two chapters, for which we are able to provide a clean theoretical analysis,
+either by restricting to $M \leq K$ devices with a transmission probability of $p=1$, or by restricting to a single player accessing a network which is assumed to be \emph{piece-wise stationary}.
 
@@
 
-The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Players Multi-Arm Bandits models.
+The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Players Multi-Arm Bandits models (Chapter~\ref{chapter:5}).
 %
@@
 with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
-For the two easiest models (\ie, with sensing), our theoretical contribution improves the state-of-the-art upper bounds on the regret.
+For the two easiest models (\ie, with sensing), our theoretical contribution \MCTopM-\klUCB{} improves the state-of-the-art regret upper-bounds.
 In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
@@
 
-In the last chapter, we were interested in relaxing the stationary hypothesis made in Chapter~\ref{chapter:2}.
+In the last Chapter~\ref{chapter:6}, we were interested in relaxing the stationary hypothesis made on the bandit model presented in Chapter~\ref{chapter:2}.
 %
@@
 we proposed a new efficient algorithm, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
-This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike \CUSUMUCB{} and \MUCB{} that require to know a lower bound on the smallest magnitude of a change.
+This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike its two competitors \CUSUMUCB{} and \MUCB, that require to know a lower bound on the smallest magnitude of a change.
 We also gave numerical evidence of the efficiency of our proposal, which performs usually much better than all the passively adaptive approaches as well as better or comparably to the previous actively adaptive ones.
@@
 to MAB models with either multi-dimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
-The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.
+The hypotheses we made were motivated by the target application, but these other models could also be used for cognitive radio applications.
 
@@
 by following the instructions given in the documentation (see Section~\ref{sub:3:listResearchWorksUsingSMPyBandits}),
-by using the Jupyter notebooks \cite{jupyter} made available
+by using the Jupyter notebooks \cite{jupyter} made available%
 \footnote{~These notebooks are hosted on both the GitHub repository of SMPyBandits and on my website on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}.
@@
 
-The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if the $S$ static and $D$ dynamic devices have different transmission patterns, and less easily with one probability per device.
+The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if the $S$ static and $D$ dynamic devices have different transmission patterns, and less easily with a different probability per device.
 Also, other emission patterns could be considered, instead of a Bernoulli process for each user.
-In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
+In this whole Chapter~\ref{chapter:4}, we prefer to consider that all the devices have the same activation probability, to keep the notations and the model as simple as possible.
 %
@@
 
-Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
-So additionally to deciding the channel for communication (\ie, \emph{where} to communicate), each device has to also decide \emph{when} to communicate.
+Another extension could be to consider not a Bernoulli activation process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
+In this case, additionally to deciding the channel for communication (\ie, \emph{where} to communicate), each device could thus also have to decide \emph{when} to communicate.
 % This is another direction of research, that we will investigate in the future.
@@
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
-However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}, like a recent work \cite{wang2019distributed} studied the minimal amount of communication needed to achieve logarithmic regret, in a similar model that authorizes direct communications between players.
+However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}, similarly to what a recent work \cite{wang2019distributed} studied for the minimal amount of communication needed to achieve logarithmic regret, in a similar model that authorizes direct communications between players.
 % Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
 
-We also presented many extensions of the multi-players bandit model,
+We also presented many extensions of the multi-players bandit model in Section~\ref{sec:5:literatureReviewOtherModels},
 and even if some have already been implemented, a major future work is to implement the most interesting ones in SMPyBandits
-(see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
+(see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}), as well as extending our theoretical analyses for such extensions of the models considered in this thesis.
 
@@
 
-We believe that our new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
-This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of the number of break-points $\Upsilon_T$.
+We believe that the new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
+This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of the number of such meaningful changes, instead of the number of break-points $\Upsilon_T$.
 We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
@@
 
-We would also like to study in the future possible extension of our approach to the slowly varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or the rotting bandit model \cite{Seznec2018}.
+We would also like to study in the future possible extension of our approach to the slowly-varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or also to the rotting bandit model \cite{Seznec2018}, which is interesting for many applications even if it does not seem appropriate for the cognitive radio setting.
 
@@
 %
-We could also build on the proof technique used in \cite{WeiSrivastava18Abruptly}, even if we are interested in removing the hypothesis that player $j$ knows its ID $j\in[M]$ before starting to play.
+We could also build on the proof technique used in this paper, even if we are interested in removing the hypothesis that player $j$ knows its ID $j\in[M]$ before starting to play.
 %
@@
 and the Bernoulli GLR break-point detector for non-stationarity (\ie, dealing with abrupt changes).
-The three parts should be connected, the same way we built \GLRklUCB{} by connecting both components.
+The three parts should be connected, the same way we built \MCTopM-\klUCB{} and \GLRklUCB, by connecting two of the three components together.
 We propose the idea of incorporating the detected change-points by B-\GLR{} test in the orthogonalization procedure \MCTopM: after any change-point is detected, the player is not sitting anymore.
-Analyzing this strategy is left as a future work.
+%
+Even if preliminary numerical simulations showed promising results,
+analyzing this strategy that combine \MCTopM, \GLR{} and \klUCB{} is left as a challenging future work.
 
@@
 
-Another interesting question is to look for a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and automatically adapt to the setting at hand.
+Another question of highest interest is to look for a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and automatically adapt to the setting at hand.
 For instance, even if \GLRklUCB{} is very efficient for piece-wise stationary problems, its forced exploration makes it sub-optimal for stationary problems.
@@
 But as we illustrated in Section~\ref{sec:25:theory}, this approach does not seem theoretically promising.
-
-\TODOL{On doit encore en écrire un peu ?}
+Thus a preferred approach could be to design a ``unified'' algorithm which adapts automatically to the kind of problem it is facing.
 
@@
 % I would like to highlight that from a personal point of view, this thesis was awesome blabla.
+
+\hr{}
+
+\vfill{}
+
+% https://fr.wikipedia.org/wiki/Liste_des_%C3%A9pisodes_de_Kaamelott
+
+\begin{small}
+\begin{quote}
+    \emph{-- Les rêves, ça se compare pas\ldots}\\
+    Alexandre Astier, \emph{Kaamelott}, Livre VI, Épisode 9, ``Dies Irae''.
+    %  écrit par Alexandre Astier.
+\end{quote}
+\end{small}

### 2019-07-09 2-Chapters/Conclusion/conclusion.tex
Better, the conclusion now fits in the same number of page: I removed the « short summary of two paragraphs for Part I and II »

@@
 
-% \begin{itemize}
-    % \item
-The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
-% Chapter 2
-In Chapter~\ref{chapter:2}, we presented the multi-armed bandit (MAB) models, the hypotheses we usually make on such models, and the most important algorithms that solve different kinds of MAB problems.
-% Chapter 3
-Then in Chapter~\ref{chapter:3} we showcased our open-source Python library SMPyBandits in details, developed to run numerical simulations of MAB problems.
-%
-Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an empirically efficient answer to the online algorithm selection problem.
+% % \begin{itemize}
+%     % \item
+% The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
+% % Chapter 2
+% In Chapter~\ref{chapter:2}, we presented the multi-armed bandit (MAB) models, the hypotheses we usually make on such models, and the most important algorithms that solve different kinds of MAB problems.
+% % Chapter 3
+% Then in Chapter~\ref{chapter:3} we showcased our open-source Python library SMPyBandits in details, developed to run numerical simulations of MAB problems.
+% %
+% Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an empirically efficient answer to the online algorithm selection problem.
 
-    % \item
-The second half of this thesis, Part~\ref{part:MABIOT}, presented the main contributions.
-% Chapter 4
-In Chapter~\ref{chapter:4}, we first proposed models for realistic IoT wireless networks, in a discrete time and with no central coordination between end-devices.
-In two different models, with or without retransmissions, we showed that end-devices can independently learn favor the less occupied radio channels, in order to increase their successful transmission rates in the network, by embedding low-cost decentralized MAB algorithms.
-We also presented a proof-of-concept that implements such network on real radio hardware and validates the proposed model and algorithms.
-% Chapter 5
-Because the models proposed for realistic IoT networks were found to be too general to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
-we started to study a weaker but still interesting model in Chapter~\ref{chapter:5}.
-We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision and spectrum sensing information (\MCTopM).
-% Chapter 6
-Finally, in Chapter~\ref{chapter:6}, we relaxed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
-%
-For the last two chapters, the proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in their theoretical analyses.
-% \end{itemize}
+%     % \item
+% The second half of this thesis, Part~\ref{part:MABIOT}, presented the main contributions.
+% % Chapter 4
+% In Chapter~\ref{chapter:4}, we first proposed models for realistic IoT wireless networks, in a discrete time and with no central coordination between end-devices.
+% In two different models, with or without retransmissions, we showed that end-devices can independently learn favor the less occupied radio channels, in order to increase their successful transmission rates in the network, by embedding low-cost decentralized MAB algorithms.
+% We also presented a proof-of-concept that implements such network on real radio hardware and validates the proposed model and algorithms.
+% % Chapter 5
+% Because the models proposed for realistic IoT networks were found to be too general to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
+% we started to study a weaker but still interesting model in Chapter~\ref{chapter:5}.
+% We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision and spectrum sensing information (\MCTopM).
+% % Chapter 6
+% Finally, in Chapter~\ref{chapter:6}, we relaxed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
+% %
+% For the last two chapters, the proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in their theoretical analyses.
+% % \end{itemize}
 
-\paragraph{Summary of each chapter.}
+% \paragraph{Summary of each chapter.}
 % A more precise summary of each chapter is given below.
@@
 
-As discussed at the end of each chapter,
-our works suggest possible directions of future studies.
+% As discussed at the end of each chapter,
+The works presented in each chapter suggest possible directions of future studies.
 
@@
 Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
-
+\\
+\indent
 Another extension could be to consider not a Bernoulli activation process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
@@
 Thus, it could be exciting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo18} or structured MAB \cite{Combes17} models and algorithms for this extension.
-
+%
 % We will investigate this behavior in order to understand it better theoretically.
 % We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.
-
+%
 % \paragraph{Extensions of the demonstration.}
-
+%
 % Possible future extensions of our demonstration include the following points.
@@
 % Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card to simulate another gateway).
-
+%
+%
 % \paragraph{Possible extensions for the second model.}
-
+%
 % Finally, the utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works.
@@
 % that nevertheless have not been explored for the targeted problem.
-
+\\
+\indent
 In order to validate our results in a realistic experimental setting and not only with simulations, future works should include a hardware implementation of the analyzed models, in order to complete our demonstration presented in Section~\ref{sec:4:gnuradio} \cite{Besson2018ICT,Besson2019WCNC}.
@@
 % Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
-
+\\
+\indent
 We also presented many extensions of the multi-players bandit model in Section~\ref{sec:5:literatureReviewOtherModels},
@@
 We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of more efficient algorithms, as suggested by \cite{Maillard2018GLR}.
+\\
+\indent
+We would also like to study in the future possible extension of our approach to the slowly-varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or also to the rotting bandit model \cite{Seznec2018}, which is interesting for many applications even if it does not seem appropriate for the cognitive radio setting.
 
-We would also like to study in the future possible extension of our approach to the slowly-varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or also to the rotting bandit model \cite{Seznec2018}, which is interesting for many applications even if it does not seem appropriate for the cognitive radio setting.
+% ----------------------------------------------------------------------------
+\subsection*{Other directions of future works}
 
@@
 % }
-
-% ----------------------------------------------------------------------------
-\paragraph{Other directions of future works.}
 

### 2019-07-09 2-Chapters/4-Chapter/chapter4.tex
Sent to my advisors, I'm printing it in 2 DAYS

@@
 As explained in the introduction in Chapter~\ref{chapter:1},
-unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
+unlicensed bands are more and more used and considered for mobile and LAN (Local Area Network) and for Internet of Things communication standards.
+%  (WiFi, LTE-U), (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
 This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.
 
-Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
+Efficient Medium Access Control (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
 As end-devices battery life is a key constraint of IoT networks,
@@
 We highlight that even without changing anything on the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
-We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
+We consider the \UCB \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
 Kaufmann12Thompson}, for the first model.

### 2019-07-09 1-Preamble/4-Acknowledgement/acknowledgement.tex
Corrections fautes orthographes

@@
 
-I would like to acknowledge and thank the following people, and also the various foundings which supported our works during the last three years and made this thesis possible.
+I would like to acknowledge and thank the following people, and also the various fundings which supported our works during the last three years and made this thesis possible.
 
@@
 Some people thank (a/some) God in their thesis, I feel that I prefer to thank \emph{the Earth}, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.
-I'am trying hard to do better from now on!
+I am trying hard to do better from now on!
 

### 2019-07-09 1-Preamble/5-Abstract/resume_fr.tex
Corrections fautes orthographes

@@
     \item
-    L'\href{https://en.wikipedia.org/wiki/Cognitive_radio}{encyclopédie Wikipedia} dit ceci
+    L'\href{https://en.wikipedia.org/wiki/Cognitive_radio}{encyclopédie Wikipédia} dit ceci
     \guillemotleft{} \emph{une RI est une radio qui peut être programmée et configurée dynamiquement pour utiliser les meilleurs canaux sans fil à proximité afin d'éviter les interférences et la congestion des utilisateurs.
@@
 et la récompense obtenue par l'objet équipé de RI après avoir analysé le canal $k$ au temps $t$ est modélisée par une récompense $r(t) \in \{0,1\}$.
-En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors que $r(t)=0$ indique que le canal $k$ est occupé durant l'interval de temps $t$, et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
+En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors que $r(t)=0$ indique que le canal $k$ est occupé durant l'intervalle de temps $t$, et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
 %
@@
 L'objectif de ce premier modèle est d'améliorer la qualité de service (QdS, ou \emph{Quality of Service} QoS) de l'application de ce réseau de l'IdO, en appliquant un ApR décentralisé côté objet afin de réduire le taux d'échec de transmission.
-La station de base recevra donc plus de paquets de liaison montante des objets desservis, s'ils peuvent apprendre par eux-même un accès efficace au spectre.
+La station de base recevra donc plus de paquets de liaison montante des objets desservis, s'ils peuvent apprendre par eux-mêmes un accès efficace au spectre.
 Cela implique aussi que le réseau peut accepter plus d'objets tout en maintenant la même QdS.
@@
 Dans les deux modèles de réseaux de l'IdO présentés ci-dessus, l'idée maîtresse est de laisser chaque objet dynamique d'un tel réseau exécuter un algorithme d'apprentissage de manière totalement décentralisée, afin d'optimiser le système complet.
-Il est important de noter que chaque objet ne communique, et donc n'apprend, qu'à quelques unes et non à toutes les étapes, en suivant son propre processus d'activation aléatoire (nous nous limitons à un processus d'activation purement aléatoire, Bernouli de probabilité $p$).
+Il est important de noter que chaque objet ne communique, et donc n'apprend, qu'à quelques unes et non à toutes les étapes, en suivant son propre processus d'activation aléatoire (nous nous limitons à un processus d'activation purement aléatoire, Bernoulli de probabilité $p$).
 Cela signifie que chaque objet vise son propre objectif local, qui est de maximiser sa récompense cumulée, d'une manière égoïste et complètement agnostique des autres objets ayant le même objectif.
@@
 Cependant, même si nous continuons à supposer un temps synchronisé dans tout le reste de la thèse,
-il est difficile de soutenir que cette hypothèse d'un calendrier centralisé pour les objets peut être réaliste, car nous étudions le cas de l'apprentissage décentralisé pour l'AOS et l'IdO précisément afin d'éviter un surcoût dû à une signalisation régulière et un contrôle central des objets par la station de base, comme nous l'avons expliqué précedemment.
+il est difficile de soutenir que cette hypothèse d'un calendrier centralisé pour les objets peut être réaliste, car nous étudions le cas de l'apprentissage décentralisé pour l'AOS et l'IdO précisément afin d'éviter un surcoût dû à une signalisation régulière et un contrôle central des objets par la station de base, comme nous l'avons expliqué précédemment.
 
@@
 ce qui signifie que le problème de bandit sous-jacent est stationnaire sur certains intervalles, séparés par des points de changement situés à des moments inconnus.
-Supposer que l'environnement est stationnaire par morceaux est en effet cohéret pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'objets dans le réseau.
+Supposer que l'environnement est stationnaire par morceaux est en effet cohérent pour les applications aux réseaux sans fil, où un point de changement peut par exemple correspondre à l'arrivée d'un nouveau groupe d'objets dans le réseau.
 Des exemples d'une telle situation peuvent être une nouvelle entreprise arrivant sur le marché dans une ville (par exemple, les compteurs Linky qui sont installés aujourd'hui en France), ou l'agriculteur voisin installant des capteurs sur son propre troupeau de vaches, etc.
@@
 Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
-Elles se situent à trois niveaux : des modèles BMB pour plusieurs contextes radio liés à la radio intelligente et l'IdO ; des preuves théoriques associées aux problèmes d'apprentissages correspondants ; et enfin des implémentations d'algorithmes à but de simulations, de partage avec la communauté, et de validation par des démonstrations radio réalistes.
+Elles se situent à trois niveaux : des modèles BMB pour plusieurs contextes radio liés à la radio intelligente et l'IdO ; des preuves théoriques associées aux problèmes d'apprentissages correspondants ; et enfin des implantations d'algorithmes à but de simulations, de partage avec la communauté, et de validation par des démonstrations radio réalistes.
 
@@
     \item
-    Nous présentons le problème du choix de l'algorithme qu'un-e ingénieur-e devrait utiliser, ou de la sélection d'algorithme parmi la riche collection des différents algorithmes de bandits dans la litérature, dans le Chapitre~\ref{chapter:25}.
+    Nous présentons le problème du choix de l'algorithme qu'un-e ingénieur-e devrait utiliser, ou de la sélection d'algorithme parmi la riche collection des différents algorithmes de bandits dans la littérature, dans le Chapitre~\ref{chapter:25}.
     Nous présentons un algorithme d'agrégation d'algorithmes, appelé \Aggr, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par des simulations numériques qu'il peut atteindre de bonnes performances empiriques
@@
     Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le Chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle (\cite{Besson2019Gretsi} publié à GRETSI 2019, \cite{Besson2019GLRT}).
-    Nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles car notre approche n'a pas besoin de connaîre de bornes sur la difficulté du problème à part le nombre de points de ruptures.
+    Nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles car notre approche n'a pas besoin de connaître de bornes sur la difficulté du problème à part le nombre de points de ruptures.
     % Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB

### 2019-07-09 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Sent to my advisors, I'm printing it in 2 DAYS
Corrections fautes orthographes

@@
 
-\textbf{Lower-bound.}
+\paragraph{Lower-bound.}
 %
@@
 
-\textbf{Comparing the algorithms against the same realizations.}
+\paragraph{Comparing the algorithms against the same realizations.}
 %
@@
 a naive uniform exploration (to have at least one algorithm with bad performances, \ie, linear regret, that is \emph{not} included in the plots to reduce clutter),
-\UCB{} with $\alpha=1/2$, three \klUCB{} algorithms (resp. with Bernoulli, Gaussian and exponential $\kl$ functions), and BayesUCB and Thompson sampling with uniform prior.
+\UCB{} with $\alpha=1/2$, three \klUCB{} algorithms (resp. with Bernoulli, Gaussian and exponential $\kl$ functions), and Bayes-UCB and Thompson sampling with uniform prior.
 
@@
 	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
-	\caption{On an ``easy'' Gaussian problem, only \Aggr{} shows reasonable performances, thanks to BayesUCB and Thompson sampling.}
+	\caption{On an ``easy'' Gaussian problem, only \Aggr{} shows reasonable performances, thanks to Bayes-UCB and Thompson sampling.}
 	\label{fig:25:EasyGaussian}
@@
 
-For Bernoulli problems (Figures~\ref{fig:25:EasyBernoulli} and~\ref{fig:25:HardBernoulli}), UCB with $\alpha=1/2$, Thompson sampling, BayesUCB and \klUCB{}$^+$ (with the binary $\kl$ function) all perform similarly, and \Aggr{} is found to be as efficient as all of them.
+For Bernoulli problems (Figures~\ref{fig:25:EasyBernoulli} and~\ref{fig:25:HardBernoulli}), UCB with $\alpha=1/2$, Thompson sampling, Bayes-UCB and \klUCB{}$^+$ (with the binary $\kl$ function) all perform similarly, and \Aggr{} is found to be as efficient as all of them.
 For Gaussian and exponential arms, rewards are truncated into $[0,1]$, and the variance of Gaussian distributions is fixed to $\sigma^2 = 0.05$ for all arms, and can be known to the algorithms (the \kl{} function is adapted to this one-dimensional exponential family).
@@
 % Empirically, \Aggr{} showed to always have a logarithmic
-% regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
+% regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, Bayes-UCB etc).
 % It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.

### 2019-07-09 2-Chapters/2-Chapter/chapter2.tex
Sent to my advisors, I'm printing it in 2 DAYS
Corrections fautes orthographes

@@
 
-\textbf{About notations.}
+\paragraph{About notations.}
 %
@@
 
-\textbf{Main references for the curious reader.}
+\paragraph{Main references for the curious reader.}
 %
@@
 
-The webpage looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
+The web-page looks like Figure~\ref{fig:2:example_of_a_5_arm_bandit_problem} below.
 The arms follow Bernoulli distributions, \ie, $\nu_k = \cB(\mu_k)$, of unknown means $\mu_k\in[0,1]$ (this means that $\forall k\in[K], \forall t\in[T], \Pr(Y_{k,t} = 1) = \mu_k$ and $\Pr(Y_{k,t} = 0) = 1 - \mu_k$).
@@
 The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
-and usually the literature focusses on regular and minimal exponential families (\ie, when the natural parameter space is an non-empty open set in $\R$).
+and usually the literature focuses on regular and minimal exponential families (\ie, when the natural parameter space is an non-empty open set in $\R$).
 
@@
     % \item
-    MAB can also be applied to a broader setting of \textbf{online content recommandation}.
+    MAB can also be applied to a broader setting of \textbf{online content recommendation}.
     % with two options (A/B testing), or more than two options.
@@
 In the model of Opportunistic Spectrum Access (OSA) with sensing \cite{Jouini09,Jouini10}, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$, at the beginning of its $t$-th transmission slot.
-A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
+A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no up-link message can be sent.
 % %
@@
 %
-We study a similar model of using MAB for Cognitive Radio but without the OSA structre of Primary and Secondary users, \ie, without sensing information but only a collision indicator, in Chapters~\ref{chapter:4} and \ref{chapter:5}.
+We study a similar model of using MAB for Cognitive Radio but without the OSA structure of Primary and Secondary users, \ie, without sensing information but only a collision indicator, in Chapters~\ref{chapter:4} and \ref{chapter:5}.
 
@@
     % \item
-On the other hand, this decomposition is necessary as theoretical analyses of the regret of (single-player) MAB algorithms are usually based on controlling the suboptimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
+On the other hand, this decomposition is necessary as theoretical analyses of the regret of (single-player) MAB algorithms are usually based on controlling the sub-optimal draws $N_k(T)$ and not directly the regret $R_T^{\cA}$.
 Even if we prove that controlling $N_k(T)$ is no longer sufficient to obtain low regret for the multi-players case, we extend this decomposition in Chapter~\ref{chapter:3} (in Lemma~\ref{lem:5:DecompositionRegret}), and it is the first quantity we prove to be bounded by $\bigO{\log(T)}$ when we prove the regret upper-bound of our proposal \MCTopM-\klUCB.
@@
 %  \ie, they require a prior knowledge on the problem difficulty.
-As such, this weakness makes the presented strategies unapplicable on an unknown problem, and so they are less interesting from a practical point-of-view.
+As such, this weakness makes the presented strategies inapplicable on an unknown problem, and so they are less interesting from a practical point-of-view.
 
@@
     According to the Algorithm~\ref{algo:2:indexPolicy} (see line~6),
-    a suboptimal arm $k$ can be chosen only if $\UCB_k(t) \geq \UCB_{k^*}(t)$.
+    a sub-optimal arm $k$ can be chosen only if $\UCB_k(t) \geq \UCB_{k^*}(t)$.
     Both $\widehat{\mu_k}(t)$ and $\xi_k(t)$ use $n=N_k(t)$ samples of arm $k$,
@@
 
-% - One of the first paper I read in 2016 was defining AdBandits, https://smpybandits.github.io/docs/Policies.AdBandits.html, a mixture between Thompson sampling and BayesUCB (very efficient in practice)
+% - One of the first paper I read in 2016 was defining AdBandits, https://smpybandits.github.io/docs/Policies.AdBandits.html, a mixture between Thompson sampling and Bayes-UCB (very efficient in practice)
 % - I especially love the BESA algorithm by [? and ? and Maillard].
@@
 % Many different learning algorithms have been proposed by the machine learning community,
-% and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
+% and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or Bayes-UCB,
 % the $\kl$ function for \klUCB{} etc.

### 2019-07-09 2-Chapters/2-Chapter/chapter25.tex
J'ai la conclusion qui dépasse sur une page… zut
Sent to my advisors, I'm printing it in 2 DAYS
Corrections fautes orthographes

@@
 We propose an extension of the \ExpFour{} algorithm for this problem, that we called \Aggr, and illustrate its performance on some bandit problems.
+
+\begin{small}
+    \begin{flushright}
+        % https://fr.wikiquote.org/wiki/Kaamelott/%C3%89lias_de_Kelliwic%E2%80%99h
+        \emph{-- Y'a toujours au moins deux solutions à un problème.}\\
+        Bruno Fontaine, \emph{Kaamelott}, Livre III, Épisode 67 ``La Potion de Fécondité II''.
+    \end{flushright}
+\end{small}
 
@@
 Many different learning algorithms have been proposed by the machine learning community,
-and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
+and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or Bayes-UCB,
 the $\kl$ function for \klUCB{} etc.

### 2019-07-09 2-Chapters/3-Chapter/chapter3.tex
J'ai la conclusion qui dépasse sur une page… zut
Corrections fautes orthographes

@@
 Then we use it to perform a numerical comparison of the main state-of-the-art single-player bandit algorithms, and a study to compare time and memory costs of some of the most well-known and most widely used MAB algorithms.
+
+\begin{small}
+    \begin{flushright}
+        \emph{\href{https://zen-of-python.info/beautiful-is-better-than-ugly.html#1}{1. Beautiful is better than ugly.}}\\
+        \emph{\href{https://zen-of-python.info/simple-is-better-than-complex.html#3}{3. Simple is better than complex.}}\\
+        \href{https://zen-of-python.info/}{The Zen of Python},
+        by Tim Peters (in \href{http://legacy.python.org/dev/peps/pep-0020/}{PEP20}).
+    \end{flushright}
+\end{small}
 
@@
     Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{~Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
-Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in \SMPyBandits{} will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
+Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connection between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in \SMPyBandits{} will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
     \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as \KLUCBswitch), and costs an-order-of-magnitude more memory.

### 2019-07-09 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Corrections fautes orthographes

@@
 %
-We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of devices all sends uplink packets to a (unique) network gateway.
+We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of devices all sends up-link packets to a (unique) network gateway.
 Messages can be sent in a fixed number $K\geq2$ of wireless channels,
@@
 % (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
-The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
+The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit up-link packets of fixed duration whenever they want.
 We mean here by ALOHA-based that \emph{devices do not use sensing}, and \emph{we do not consider retransmissions of packets} (we relax this hypothesis in Section~\ref{sec:4:retransmissions}).
@@
 The devices can transmit their packets in one of the $K$ channels.
-In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
+In the case where the gateway --of the corresponding IoT network-- receives an up-link message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
 This is realistic in IoT networks such as networks based on LoRaWAN.
@@
 In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the device by letting only one gateway sends back an acknowledgement.
-The natural choice is for the network to send this acknowledgement by the gateway which received the uplink message.
+The natural choice is for the network to send this acknowledgement by the gateway which received the up-link message.
 For simplicity but without loss of generality, we consider a network with only one gateway,
-but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming uplink messages from the IoT devices being paired in the network.
+but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming up-link messages from the IoT devices being paired in the network.
 %
@@
 %
-In order to simulate networks designed for the IoT, we consider a protocol \textbf{with no sensing} and no repetition of uplink messages.
-The gateway is in charge of sending back an acknowledgement, after some fixed-time delay, to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) which succeeded in sending an uplink packet.
+In order to simulate networks designed for the IoT, we consider a protocol \textbf{with no sensing} and no repetition of up-link messages.
+The gateway is in charge of sending back an acknowledgement, after some fixed-time delay, to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) which succeeded in sending an up-link packet.
 %
@@
 the device can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use, that is, the most vacant one.
-Indeed, in this model, the \emph{quality} of the channels (\ie, the ressources) are identified by their vacancy, or one minus their occupancy rates.
+Indeed, in this model, the \emph{quality} of the channels (\ie, the resources) are identified by their vacancy, or one minus their occupancy rates.
 
@@
     \caption[The considered time-frequency slotted protocol.
-Each frame is composed by a fixed duration uplink slot in which the end-devices transmit their (uplink) packets.
+Each frame is composed by a fixed duration up-link slot in which the end-devices transmit their (up-link) packets.
 If a packet is well received, the gateway replies by transmitting an \Ack, after the ack delay.]{The considered time-frequency ALOHA-based communication protocol \cite{Abramson1970,Roberts75}.
-Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}.
+Each frame is composed by a fixed duration \textcolor{blue}{up-link slot} in which the end-devices transmit their \textcolor{blue}{(up-link) packets}.
 If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
@@
 In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication protocol \cite{Abramson1970,Roberts75}, with no sensing.
-Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the gateway.
+Each time slot is divided in two parts: first for up-link communications in which data packets are sent by end-devices to the gateway.
 In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
-If two or more devices send an uplink packet in the same slot, the uplink packets collide, and the acknowledgement (\Ack) is not transmitted.
+If two or more devices send an up-link packet in the same slot, the up-link packets collide, and the acknowledgement (\Ack) is not transmitted.
 In this case, we say that there is a \emph{collision} in this time slot.
-In other words, if the gateway cannot successfully decode the incoming (uplink) messages because they were corrupted by collisions, it does not send any \Ack{} back.
-This way, no collision can occur on the downlink messages, easing the analysis of collisions.
+In other words, if the gateway cannot successfully decode the incoming (up-link) messages because they were corrupted by collisions, it does not send any \Ack{} back.
+This way, no collision can occur on the down-link messages, easing the analysis of collisions.
 
@@
 When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
-Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
+Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and sub-optimal oracle policies.
 For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.

### 2019-07-09 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Corrections fautes orthographes

@@
 %
-We use a TestBed designed in 2017 by the SCEE team at the Rennes campus of CentraleSupélec \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop running the GNU Radio software \cite{GNURadioDocumentation},
+We use a test-bed designed in 2017 by the SCEE team at the Rennes campus of CentraleSupélec \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop running the GNU Radio software \cite{GNURadioDocumentation},
 and where the intelligence of each device corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.
@@
 With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the base station.
-Once the base station receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
+Once the base station receives an up-link packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the up-link packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
 Thanks to this index, we can have several devices communicating with the same base station.
@@
     \includegraphics[width=0.70\linewidth]{SCEE_TestBed2.jpg}
-    \caption{Two pictures showing the SCEE testbed \cite[Appendix~3]{Bodinier17}, taken in $2018$.}
+    \caption{Two pictures showing the SCEE test-bed \cite[Appendix~3]{Bodinier17}, taken in $2018$.}
     \label{fig:42:photosSCEETestBed}
@@
 The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
-The pictures in Figure~\ref{fig:42:photosSCEETestBed} show the testbed used for these experiments.
+The pictures in Figure~\ref{fig:42:photosSCEETestBed} show the test-bed used for these experiments.
 % Octoclock
@@
 We can see, in this view that the device transmitted its last $9$ packets in channels $\#3$ and $\#4$.
-Then, at the top right of this interface \circled{c}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the base station (\textcolor{darkblue}{blue}).
+Then, at the top right of this interface \circled{c}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the up-link packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the base station (\textcolor{darkblue}{blue}).
 Colors in the ``waterfall'' represent the RF power level received at the device antenna.
@@
 The last part is the interface of the base station, where we can see the traffic observed by the base station \circled{a} and the channels in which the last acknowledgements have been sent \circled{b}.
-The observed traffic is coherent with \circled{c} of part \textcolor{darkred}{\boldcircled{2}}, as the elements are very close the one from the others on the testbed.
+The observed traffic is coherent with \circled{c} of part \textcolor{darkred}{\boldcircled{2}}, as the elements are very close the one from the others on the test-bed.
 Colors may change, as they depend on the exact distance between the different transmitters and receivers.
@@
 Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
-However, our experimental testbed cannot run hundreds of intelligent devices.
+However, our experimental test-bed cannot run hundreds of intelligent devices.
 Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
@@
 The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
-It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
+It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE test-bed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
 % \textcolor{white}{Of course, the video was not seen much, because our work is useless! Yay!}

### 2019-07-09 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
J'utilise moins "our ..."
Corrections fautes orthographes

@@
 The code corresponding to the following components is written in \texttt{C++} and \texttt{Python} for the hand-written blocks, and using the GUI of the GNU Radio Companion software.
-The flow-graphs are saved as XML files, and the complete code of our demo is fully available online, open-sourced under GPLv3 license, at
+The flow-graphs are saved as XML files, and the complete code of our proof-of-concept is fully available online, open-sourced under GPLv3 license, at
 \href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit}\\\texttt{-learning-for-iot-networks-with-grc/}}.
@@
 We note the difference between the base station and the end-devices.
-The base station uses its Rx antenna to scan the entire range of the $K$ channels, at all time steps ; while dynamic devices use their Rx antennas only to listen for the acknowledgement sent by the base station if it received and decoded its uplink packet.
-After each transmission, the dynamic device listens to one channel during a certain time interval, in the channel used for uplink transmission.
+The base station uses its Rx antenna to scan the entire range of the $K$ channels, at all time steps ; while dynamic devices use their Rx antennas only to listen for the acknowledgement sent by the base station if it received and decoded its up-link packet.
+After each transmission, the dynamic device listens to one channel during a certain time interval, in the channel used for up-link transmission.
 

### 2019-07-09 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Corrections fautes orthographes

@@
 The novelty compare to the previous model is that
-in case of RF (uplink or downlink) collision that prevent the device to receive the \emph{Ack} from the gateway,
+in case of RF (up-link or down-link) collision that prevent the device to receive the \emph{Ack} from the gateway,
 these devices will attempt to retransmit their packet up-to $\mathrm{MaxBackOff}$ times\footnote{~We denote it $\mathrm{MaxBackOff}$ instead of $M$ like we did in our paper \cite{Bonnefoi2019WCNC}, as $M$ is used in next Chapter~\ref{chapter:5}.},
@@
 		%
-		In the algorithms presented in this chapter, the time step $t$ does not refer to the \emph{global} time step (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
+		In the algorithms presented in this chapter, the time step $t$ does not refer to the \emph{global} time step (as seen from the gateway), but rather the current number of up-link transmissions (or retransmissions) that was carried out by the device.
 		Each device wakes up whenever it has to send some data, following its own emission process (we restrict to a Bernoulli process of small probability, \eg, $p=10^{-3}$), and whenever it wakes up, it sends its packet, and will send again for at most $\mathrm{MaxBackOff}$ times (\eg, $\mathrm{MaxBackOff}=10$) until it receives an \Ack.
 		%
-		In other words, the time steps $t$ in the following algorithms denote the number of uplink packets sent by the device, and these uplink transmissions or retransmissions can happen at different (global or real-world) times.
+		In other words, the time steps $t$ in the following algorithms denote the number of up-link packets sent by the device, and these up-link transmissions or retransmissions can happen at different (global or real-world) times.
 		The local index $t$ is independent on the global time (\ie, the time of the gateway), and IoT device does not need to care about this difference.

### 2019-07-09 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Corrections fautes orthographes

@@
 and some \iid{} background traffic is assumed to be non-uniformly distributed in the $K$ channels.
-As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
+As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent up-link packet can be received by the gateway.
 %
@@
 We start by reviewing previous works on multi-players MAB models, which all considered the easier case of \emph{sensing feedback}.
-But in the model studied in this chapter, each device also uses a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each uplink message (\ie, at each time step).
+But in the model studied in this chapter, each device also uses a Multi-Armed Bandit algorithm, to maximize its number of successful communications, by using the received acknowledgement \Ack{} as a (random) binary reward after each up-link message (\ie, at each time step).
 %
@@
 In the point of view of the wireless protocol, each time frame is separated as before in a \emph{sensing} phase (during which the device senses for the background traffic),
-an \emph{uplink} phase (during which the device sends a packet to the gateway if it sensed the chosen channel to be free),
-and a \emph{downlink} phase (during which it waits for an \Ack{} from the gateway).
+an \emph{up-link} phase (during which the device sends a packet to the gateway if it sensed the chosen channel to be free),
+and a \emph{down-link} phase (during which it waits for an \Ack{} from the gateway).
 In this first model, the binary reward is $1$ only if the channel was sensed to be free of background traffic \textbf{and} if \Ack{} was received, and the device has access to both information.
@@
 % and as we do in Appendix~\ref{proof:5:BetterLowerBound},
-then introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
+then introduce an alternative model parameterized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
 Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
@@
 %
-Thus we propose next our main contribution, the \MCTopM{} algorithm, proved to be asymptotically optimal for the identification of suboptimal arms (when using \klUCB), attaining order-optimal logarithmic regret, and outperforming all the other algorithms for the ``sensing case''.
+Thus we propose next our main contribution, the \MCTopM{} algorithm, proved to be asymptotically optimal for the identification of sub-optimal arms (when using \klUCB), attaining order-optimal logarithmic regret, and outperforming all the other algorithms for the ``sensing case''.
 
@@
 and for instance \cite{modiDemo2016} studied the empirical impact of the number of hardware reconfigurations on the battery life of a dynamic end-devices.
-This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
+This work highlighted a trade-off between minimizing regret and minimizing the number of switches of arms.
 On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches, in order to reduce by a certain multiplicative factor its switching cost while multiplying its regret by the same factor.
@@
 We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \UCB{} or \klUCB{} indices.
-The analysis given in Section~\ref{sec:5:upperbounds} was focussing on \klUCB, because it is well known that it is more efficient both in practice and in theory.
+The analysis given in Section~\ref{sec:5:upperbounds} was Focusing on \klUCB, because it is well known that it is more efficient both in practice and in theory.
 To illustrate this, we first show below the result of some experiments comparing different algorithms that use either the \UCB{} or the \klUCB{} indexes.
 %
-Our proposed algorithms, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
+Our proposed algorithms, \MCTopM{}, \RandTopM{} and \Selfish{} are bench-marked against the state-of-the-art \RhoRand{} algorithm.
 

### 2019-07-09 2-Chapters/5-Chapter/chapter5.tex
Corrections fautes orthographes

@@
 Many extensions of the simpler model of multi-players bandits have been considered.
-The number of players $M$ can be fixed but initially unknown to the players, the sensing information can be absent (like the model \modeltrois{} presented above), there could be communications between players (happening at each time step or only occasionally), or $M$ could evolve over time to model the possibility of arrivals or departures of players (\ie, connexions and disconnections of devices in a wireless networks).
+The number of players $M$ can be fixed but initially unknown to the players, the sensing information can be absent (like the model \modeltrois{} presented above), there could be communications between players (happening at each time step or only occasionally), or $M$ could evolve over time to model the possibility of arrivals or departures of players (\ie, connection's and disconnections of devices in a wireless networks).
 Also inspired by real-world wireless networks, the arm means may vary among players, for instance to model players located at different distance of the gateway, and we can also consider networks with some ``jammers'' whose purpose is not to communicate to the gateway in a collaborative way, like the $M$ devices, but to interfere with the communications of the $M$ devices.
@@
 Finally, we can also be interested by models where $M$ stays fixed, but the environment evolves, for instance with abrupt changes in the means of arms.
-This last model makes a good connexion between this chapter and the next one, and an exciting future work is to further study this model in order to tackle this kind of problems by merging our contributions for stationary multi-players bandits and for non-stationary single-player bandits.
+This last model makes a good connection between this chapter and the next one, and an exciting future work is to further study this model in order to tackle this kind of problems by merging our contributions for stationary multi-players bandits and for non-stationary single-player bandits.
 
@@
 Second, if the $M$ players falsely use a value $M' > M$, then there is also a certain (fixed) probability to achieve a linear regret.
-Indeed, imagine one player ($M=1$) running \MCTopM{} with the false knowledge of $M'=2$, and if it learned to accurately identify the set of the $2$ best arms, then the \MCTopM{} orthogonalization scheme can make it play the worst of the two arms, and as $M=1$ this player will never encounter any collision, thus playing this suboptimal arm for about $T-\bigO{\log(T)}$ times, also leading to linear regret.
+Indeed, imagine one player ($M=1$) running \MCTopM{} with the false knowledge of $M'=2$, and if it learned to accurately identify the set of the $2$ best arms, then the \MCTopM{} orthogonalization scheme can make it play the worst of the two arms, and as $M=1$ this player will never encounter any collision, thus playing this sub-optimal arm for about $T-\bigO{\log(T)}$ times, also leading to linear regret.
 
-\paragraph{Interpretation of the hypothesis of knowning $M$ for real-world networks.}
+\paragraph{Interpretation of the hypothesis of knowing $M$ for real-world networks.}
 \label{par:5:knowingYourIDinMPBanditsGame}
@@
 Then in such networks, if one wants to apply an algorithm like \MCTopM-\klUCB,
-the $M$ devices need to know their number $M$, and have to receive it from the gateway, as it is the only part of this example of network which knowns $M$.
+the $M$ devices need to know their number $M$, and have to receive it from the gateway, as it is the only part of this example of network which known $M$.
 It seems unrealistic to ask the gateway to send the fixed value $M$ to each device, and not a unique ID to each device, for instance $\mathrm{id}^j\in[K]$.
@@
 By calibrating $T_0$ based on $\delta$ (which needs prior knowledge of $T$, see the proof we gave in Section~\ref{proof:2:tuningExploreThenCommit}) and the minimal gap $\Delta$ between \Mbest{} and \Mworst, one can show, using similar arguments as used by \cite{Rosenski16} for the Musical Chair policy, that with large probability a constant regret is obtained.
-A logarithmic regret can be obtained from the same bound, proving the order-optimality of this simple approach, if we assume that user $j$ knows it is user number $j$.
+A logarithmic regret can be obtained from the same bound, proving the order-optimist of this simple approach, if we assume that user $j$ knows it is user number $j$.
 %
@@
 
-\paragraph{Additionnal numerical simulations.}
+\paragraph{Additional numerical simulations.}
 
@@
 %
-The first work is \cite{NaparstekCohen17}, which proposes to use a decentralized learning based on deep learning, in the ``no sensing'' case, but in a model where the users have to learn not only the channels to use for their uplink messages (only from the feedback through the acknowledgements, like in Section~\ref{sec:4:firstModel}), but where they have to learn the whole ``spectrum access actions''.
+The first work is \cite{NaparstekCohen17}, which proposes to use a decentralized learning based on deep learning, in the ``no sensing'' case, but in a model where the users have to learn not only the channels to use for their up-link messages (only from the feedback through the acknowledgements, like in Section~\ref{sec:4:firstModel}), but where they have to learn the whole ``spectrum access actions''.
 % Deep Multi-User Reinforcement Learning for Dynamic Spectrum Access in Multichannel Wireless Networks
@@
 The mathematics behind their model are actually quite close to the model, but they explain it in terms of OFDMA and Quality-of-Service (QoS).
-Their algorithm is essentially based on a pre-agreement of the $M$ players, that will deterministically run an alternance of exploration phase, auction phase and exploitation phase, of (exponentially) increasing durations.
-By diving into the details of the modulation and giving explicitly the form of the uplink messages sent by the devices, the authors are able to set-up two different uplink packets, to efficiently perform the auction phase (see Figure~2).
+Their algorithm is essentially based on a pre-agreement of the $M$ players, that will deterministically run an alternation of exploration phase, auction phase and exploitation phase, of (exponentially) increasing durations.
+By diving into the details of the modulation and giving explicitly the form of the up-link messages sent by the devices, the authors are able to set-up two different up-link packets, to efficiently perform the auction phase (see Figure~2).
 They prove a regret upper-bound of the order $R_T = \bigO{\log(T)}$, with no special care regarding the constants, but we can also note that their algorithm require a prior knowledge of $\Delta_{\min}$ a problem-dependent constant (Theorem~3).
@@
 In all this chapter and the previous research literature, another common hypothesis is that all the $M$ players are pursuing the same goal, and are all behaving nicely by following the same algorithm.
-% Efficient utilization of licensed spectrum in the cognitive radio network is challenging due to lack of coordination among the Secondary Users (SUs).
-Distributed algorithms proposed in the literature aim to maximize the network throughput by ensuring orthogonal channel allocation for the SUs.
-However, these algorithms work under the assumption that all the SUs faithfully follow the algorithms which may not always hold due to the decentralized nature of the network.
+% Efficient utilization of licensed spectrum in the cognitive radio network is challenging due to lack of coordination among the Secondary Users (SU).
+Distributed algorithms proposed in the literature aim to maximize the network throughput by ensuring orthogonal channel allocation for the SU.
+However, these algorithms work under the assumption that all the SU faithfully follow the algorithms which may not always hold due to the decentralized nature of the network.
 In the paper \cite{SawantKumar2018}, the authors study for the first time distributed algorithms that are robust against malicious behavior, also called \emph{jamming attack}.
 They consider both the cases of \emph{jammers} launching coordinated and uncoordinated attacks, and consider a set of $J$ jammers.
-In the coordinated attack, the jammers select non-overlapping channels to attack in each time slot and can significantly increase the number of collisions for SUs.
+In the coordinated attack, the jammers select non-overlapping channels to attack in each time slot and can significantly increase the number of collisions for SU.
 They setup the problem in each scenario as a multi-players bandit and develop algorithm,
-and their analysis shows that when the SUs faithfully implement proposed algorithms, the regret is constant with high probability.
+and their analysis shows that when the SU faithfully implement proposed algorithms, the regret is constant with high probability.
 They validate their claims through extensive synthetic experiments and also through a realistic USRP.

### 2019-07-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Corrections fautes orthographes

@@
 \ExpThreeS{} actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
-It constitues a good baseline for the numerical experiments in Section~\ref{sec:6:NumericalExperiments},
+It constitutes a good baseline for the numerical experiments in Section~\ref{sec:6:NumericalExperiments},
 even if we did find that \ExpThreeS{} performs worse that the most of the other approaches based on extending stationary bandit algorithms (\eg, SW-UCB).
@@
 %
-The \ExpThreeR{} algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with \ExpThree, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
+The \ExpThreeR{} algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with \ExpThree, and the history of all arms are reset as soon as a sub-optimal arm is detecting to become optimal.
 A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.
@@
 %
-Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\omega\in(0,1)$ (note that it is called $\alpha$ in both papers, but we rename it $\omega$ in this chapter to avoid clutter with $\alpha$ of $\UCB_1$ in \eqref{eq:2:UCB_index}).
+Both approaches also introduced a mechanism to force a uniform exploration of each arm, parameterized by $\omega\in(0,1)$ (note that it is called $\alpha$ in both papers, but we rename it $\omega$ in this chapter to avoid clutter with $\alpha$ of $\UCB_1$ in \eqref{eq:2:UCB_index}).
 % \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\omega/K$ at each time step,
@@
 
-\paragraph{Positionning of our results.}
+\paragraph{Positioning of our results.}
 %
@@
 % We also display the number of observations (since the last restart) of the arm on which the change was detected.
-As shown below, in the experiments summary, the three algorithms based on change-point detection obtain sublinear regret, but they have different behaviors.
+As shown below, in the experiments summary, the three algorithms based on change-point detection obtain sub-linear regret, but they have different behaviors.
 The test in \MUCB{} usually detects less changes, while the test in \CUSUMUCB{} usually detects more changes and even unnecessary changes.
@@
 Very small delays are possible only after having collected a lot of samples of the arm which changed, and for instance the same algorithm obtains a delay of $32$ samples from \textcolor{red}{arm $1$} for the next change at $t=3000$, because this arm was sampled less.
-On this problem, while \MklUCB{} detects the same changes as \GLRklUCB, with larger but comparable delays, it is seen to obtain a larger regret than the two \GLR{} variants, simply because the tuning of its forced exploration probability (\ie, $\omega$) makes it explore suboptimal arms more often.
+On this problem, while \MklUCB{} detects the same changes as \GLRklUCB, with larger but comparable delays, it is seen to obtain a larger regret than the two \GLR{} variants, simply because the tuning of its forced exploration probability (\ie, $\omega$) makes it explore sub-optimal arms more often.
 
@@
 Finally, in the case of hard problems, like problems $3$ and $5$, that have a lot of changes but where the optimal arm barely changes, we verify in Figure~\ref{fig:6:meanRegretPb5} that \klUCB{} and TS can outperform the oracle policy.
-Indeed the oracle policy is suboptimal as it restarts as soon as one arm change but is unaware of the meaningful changes, and stationary policies which quickly identify the best arm will play it most of the times, achieving a smaller regret.
+Indeed the oracle policy is sub-optimal as it restarts as soon as one arm change but is unaware of the meaningful changes, and stationary policies which quickly identify the best arm will play it most of the times, achieving a smaller regret.
 We note that, sadly, all actively adaptive policies fail to outperform stationary policies on such hard problems, because they do not observe enough rewards from each arm between two restarts (\ie, the Assumptions~\ref{ass:6:LongPeriodsGlobal} and \ref{ass:6:LongPeriods} for the two Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} are not satisfied).

### 2019-07-09 3-Appendices/2-Appendix/appendix2.tex
Corrections fautes orthographes

@@
 On the one hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for this model we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
-On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices (like we present it in Chapter~\ref{chapter:4}): usually a learning step corresponds to an uplink and then downlink message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages.
+On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices (like we present it in Chapter~\ref{chapter:4}): usually a learning step corresponds to an up-link and then down-link message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages.
 While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies \cite{Centenaro16}), many kinds of application such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.
@@
 Interestingly, we prove that they conserve bounds of the mixed form $R_T = \cO(T^{\gamma} (\log(T))^{\delta})$, for $0<\gamma<1$ and $\delta>0$, and so an exponential doubling trick could be used to obtain an anytime version of the algorithm we propose for non-stationary bandits, \GLRklUCB{} in Chapter~\ref{chapter:6}.
-However, our study was only focussing on stationary bandit, and it is left as a future work to explore the harder case of piece-wise stationary problems.
+However, our study was only focusing on stationary bandit, and it is left as a future work to explore the harder case of piece-wise stationary problems.
 
@@
 
-To conclude this work, we would need to complete the study of the doubling tricks, and instead of focussing on two families (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bounds.
+To conclude this work, we would need to complete the study of the doubling tricks, and instead of focusing on two families (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bounds.
 Such doubling is given either by the function mapping the current time $t$ to the current estimate of the horizon $T(t)$, or the sequence $(T_i)_{i\in\N^*}$ of successive estimates of the horizons.

### 2019-07-10 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Merci Christophe de toutes ces remarques !

@@
 % \subsection{No theoretical guarantees.}\label{sec:25:theory}
-\section{Towards theoretical guarantees}\label{sec:25:theory}
+\section{Conclusion -- Towards theoretical guarantees}\label{sec:25:theory}
 
@@
 %
-However, \Aggr{} always seems to have a (finite-horizon) logarithmic regret in all the experiments we performed,
+However, \Aggr{} seems to have a (finite-horizon) logarithmic regret in all the experiments we performed,
 for both Bernoulli and non-Bernoulli problems (\eg, Gaussian, exponential and Poisson distributions).
@@
 % regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, Bayes-UCB etc).
-% It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
+% It usually succeeds to be close to the best of the aggregated algorithms, both in terms of regret and best arm pull frequency.
 % As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.

### 2019-07-10 2-Chapters/2-Chapter/chapter2.tex
Merci Christophe de toutes ces remarques !

@@
 As explained in Chapter~\ref{chapter:1}, we focus in this thesis on Reinforcement Learning (RL), and especially on decision-making models with a finite number of resources, called arms, and under a stochastic hypothesis: an arm being associated to a binary or real-valued distribution.
-In this chapter, we present the common base of all the more sophisticated mathematical models studied in this thesis: the stochastic multi-armed bandit (MAB) model, by restricting to the single-player and stationary stochastic case.
+In this chapter, we present the common base of the more sophisticated mathematical models studied in this thesis: the stochastic multi-armed bandit (MAB) model, by restricting to the single-player and stationary stochastic case.
 %
@@
 For any $K$, $T$, and problem instances $I$ in a certain family $\cI$ (\eg, with Bernoulli distributions),
-they enjoy regret bounds like the following, for two constants $c_{\cA} \geq 1$ and $c'_{\cA} \geq 1$, that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only: $R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T))$.
+they satisfy regret bounds like the following, for two constants $c_{\cA} \geq 1$ and $c'_{\cA} \geq 1$, that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only: $R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T))$.
 
 More recently, the \klUCB-switch from \cite{GarivierHadiji2018} is the first algorithm to be proven to
-enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, by matching the $C_I \ln(T)$ asymptotic lower-bound by \cite{LaiRobbins85}.
+obtain simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, by matching the $C_I \ln(T)$ asymptotic lower-bound by \cite{LaiRobbins85}.
 It is an index policy that uses modified versions of the indexes of \klUCB$^+$ for arms that are not sampled much (see \cite{Honda2019} for more details), and of MOSS$^+$ for the other arms.
@@
 In this chapter, we presented the multi-armed bandit model, by restricting to a finite number of arms and real-valued rewards.
-Our main focus is on binary or bounded rewards, or one-dimensional exponential families of distributions, and on stochastic and stationary problems.
+Our main focus is binary or bounded rewards, or one-dimensional exponential families of distributions, and stochastic and stationary problems.
 By showcasing a small interactive demonstration made available online,
@@
 % %
-We then used the small Bernoulli-distributed problem underlying this demonstration
-to present a small numerical simulation, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
+We then used the Bernoulli-distributed problem underlying this demonstration
+to present numerical simulations, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
 % The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.
@@
 We generalize this MAB model in Chapter~\ref{chapter:4}, by studying decentralized learning of a large set of independent players, all having different random activation times.
-This extension is significantly too complicated to analyze, and we were unfortunately unable to obtain any strong theoretical results under such loose hypotheses.
-This model is however realistic, and as such it was found suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
-%
+This extension is significantly harder than the base model presented in this chapter, and it is too complicated to analyze.
+% and we were unfortunately unable to obtain any strong theoretical results under such loose hypotheses.
+This model is more generic and realistic, and as such more suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
+%
+John is very unique.
 We take a more formal approach in the rest of the thesis,

### 2019-07-10 2-Chapters/3-Chapter/chapter3.tex
Merci Christophe de toutes ces remarques !

@@
 Then we use it to perform a numerical comparison of the main state-of-the-art single-player bandit algorithms, and a study to compare time and memory costs of some of the most well-known and most widely used MAB algorithms.
+%
+Note that almost all the numerical simulations used in the rest of the thesis are based on SMPyBandits.
 
@@
     \begin{flushright}
-        \emph{\href{https://zen-of-python.info/beautiful-is-better-than-ugly.html#1}{1. Beautiful is better than ugly.}}\\
-        \emph{\href{https://zen-of-python.info/simple-is-better-than-complex.html#3}{3. Simple is better than complex.}}\\
+        \emph{\href{https://zen-of-python.info/beautiful-is-better-than-ugly.html\#1}{1. Beautiful is better than ugly.}}\\
+        \emph{\href{https://zen-of-python.info/simple-is-better-than-complex.html\#3}{3. Simple is better than complex.}}\\
         \href{https://zen-of-python.info/}{The Zen of Python},

### 2019-07-10 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Merci Christophe de toutes ces remarques !

@@
 Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_k)_k = [15\%, 10\%, 2\%, 1\%]$.
-Note that a maximum occupancy rate of $15\%$ could seem not so high, but indeed it is, because for a pure ALOHA access mode, a naive dynamic device only enjoys about $40\%$ success rate, under such occupancy of the channels (see the ``uniform'' plot in Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}).
+Note that a maximum occupancy rate of $15\%$ could seem not so high, but indeed it is, because for a pure ALOHA access mode, a naive dynamic device only obtains about $40\%$ success rate, under such occupancy of the channels (see the ``uniform'' plot in Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}).
 %

### 2019-07-10 2-Chapters/4-Chapter/chapter4.tex
Corrections fautes orthographes
Merci Christophe de toutes ces remarques !

@@
 %
-We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost Reinforcement Learning (RL) algorithms to learn to improve their spectrum access.
+We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost Reinforcement Learning (RL) algorithms in order to learn how to improve their spectrum access.
 %
@@
 %
-However, our second model considers that in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
-We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics also significantly improve the network efficiency.
+In a second step, we consider a second model where a dynamic device has to (try to) retransmit a message, in case of a failed first transmission, up-to a fixed number of retransmissions.
+We compare heuristics based on \UCB, and simulations confirm that the non-naive heuristics also significantly improve the network efficiency.
 % Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
@@
 % ----------------------------------------------------------------------------
-\section{Towards theoretical guarantees}
+\section{Conclusion -- Towards theoretical guarantees}
 % \section{Conclusion}
@@
 
-To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
+We focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
 We presented two models of wireless IoT networks, without relying on the feedback provided by spectrum sensing, and inspired by the ALOHA protocol.
-We proposed two versions, with or without retransmissions of uplink packages in case of collisions.
+We proposed two versions, with or without retransmissions of up-link packages in case of collisions.
 % In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
@@
 It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
-Unfortunately, it turned out to be of extreme difficulty to analyze the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
-That is why we focus on two different simplifications of this model in the next chapters, for which we are able to provide a clean theoretical analysis.
+Unfortunately, it turned out to be of extreme difficulty to analyze analytically the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
+That is why we focus on two different simplifications of this model in the next chapters, for which we are able to provide a rigorous theoretical analysis.
 % %

### 2019-07-10 2-Chapters/5-Chapter/chapter5.tex
Merci Christophe de toutes ces remarques !

@@
 %
+% https://english.stackexchange.com/questions/35472/approach-to-or-approach-for
+% approach to ? approach for ? approach of ?
 In this chapter, we are interested in a more formal approach to the decentralized learning problem presented of many end-devices accessing the same IoT wireless network
@@
 We restrict to the easier case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
-Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$, because it was found very hard to analyze the previously studied IoT network models, under the hypothesis of random activation patterns of Chapter~\ref{chapter:4}.
+Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$,
+%  because it was found very hard to analyze the previously studied IoT network models,
+under the hypothesis of random activation patterns of Chapter~\ref{chapter:4}.
 %

### 2019-07-10 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Merci Christophe de toutes ces remarques !

@@
 % This chapter is structured as follows.
-We introduce the model in Section~\ref{sec:6:BanditSetting}, and review related works in Section~\ref{sec:6:relatedWork}.
-In Section~\ref{sec:6:ChangePointDetector} we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a change-point Detector (CPD) algorithm.
+We introduce the model in Section~\ref{sec:6:BanditSetting}, and we review related works in Section~\ref{sec:6:relatedWork}.
+In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a change-point Detector (CPD) algorithm.
 We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on their regret.
@@
 %
-for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant enjoys the same false-alarm control.
+for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant obtains the same false-alarm control.
 However, the detection delay may be slightly increased.
@@
 % We believe that our new proof technique could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous work), that would require only a few ``meaningful'' changes to be detected.
-This interesting research direction is left for future work,  but the hope is that the regret would be expressed in term of this number of meaningful changes instead of $\Upsilon_T$.
+This interesting research direction is left for future work,  but the hope is that the regret would be expressed in terms of this number of meaningful changes instead of $\Upsilon_T$.
 We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.

### 2019-07-10 2-Chapters/Conclusion/conclusion.tex
Merci Christophe de toutes ces remarques !

@@
 We believe that the new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
-This promising research direction is left for future work,  but the hope is that the regret would be expressed in term of the number of such meaningful changes, instead of the number of break-points $\Upsilon_T$.
+This promising research direction is left for future work,  but the hope is that the regret would be expressed in terms of the number of such meaningful changes, instead of the number of break-points $\Upsilon_T$.
 We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.

### 2019-07-10 2-Chapters/1-Chapter/chapter1.tex
Corrections fautes orthographes
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 where communications are initialized and regulated by the devices, not by a centralized control systems.
-% due to the really limited information that the base station can send to the devices (in downlink), it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
+% due to the really limited information that the base station can send to the devices (in down-link), it is no longer possible to optimize the efficiency of the network from the centralized point-of-view of the base station.
 Indeed, a central control requires signaling packets that have been identified as too heavy for these kinds of systems.
 In the present and future IoT networks, many devices of heterogeneous natures are using the same antenna for different applications.
-A common problem is the strong constraint that such such IoT devices have on their power consumption, as most of them will be deployed without a direct power access and will run on a tiny battery, which lifespan should be maximized.
+A common problem is the strong constraint that such IoT devices have on their power consumption, as most of them will be deployed without a direct power access and will run on a tiny battery, which lifespan should be maximized.
 Most commercial IoT companies nowadays indeed sell a lifespan larger than $10$ years, like SIGFOX \cite{Centenaro16}.
@@
     \item
-    Joseph Mittola in $1999$ proposed that
+    J. Mittola in $1999$ proposed that
     \emph{``a really smart radio that would be self-, RF- and user-aware, and that would include software technology and machine learning capabilities along with a lot of high-fidelity knowledge of the radio environment''} \cite{Mitola99}.
@@
     \item
-    Then Simon Haykin in $2005$ also said that
+    Then S. Haykin in $2005$ also said that
     \emph{``a CR is an intelligent wireless communication system that is capable of being aware of its surroundings, learning, and adapting its operating parameters (\eg, transmit power and carrier frequency) on the fly with an objective of providing reliable anytime, anywhere, and spectrally efficient communication''} \cite{Haykin05}.
@@
 %
-Because there are so many MAB algorithms available, we are also interested by the question of how an engineer can chose the algorithm to implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
+Because there are so many MAB algorithms available, we are also interested by the question of how an engineer can choose the algorithm to implement, in a given IoT object, in order to equip this object with the capacity to adapt robustly to any environment.
 To answer this question, we present two approaches.
@@
 This work on aggregating MAB algorithms was motivated for the OSA case of CR, where many previous research works only considered the $\UCB_1$ algorithm without really justifying their choice.
-This contribution was presented at the IEEE WCNC conference in Barcelona, Spain, in April $2018$.
+This contribution was presented in the IEEE WCNC conference in Barcelona, Spain, in April $2018$.
 
@@
 %
-Our second model then considers packet retransmissions, and while the two models are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, as well as an increased QoS for the entire network, as less retransmissions reduce the local spectrum load.
+Our second model then considers packet retransmissions, and while the two models are similar, the application of an efficient decentralized learning algorithm now results in both an increased battery life for each of the devices, and an increased QoS for the entire network, as less retransmissions reduce the local spectrum load.
 
@@
 % We consider an ALOHA-based protocol, slotted in both time and frequency, by assuming a simple pre-agreed synchronization between all devices and the base station.
-This synchronization is the only aspect of the network that requires central control from the base station, and it only requires one downlink packet when a new device joins the network operated by this base station \cite{Centenaro16}.
+This synchronization is the only aspect of the network that requires central control from the base station, and it only requires one down-link packet when a new device joins the network operated by this base station \cite{Centenaro16}.
 % %
@@
 % We then consider an extension of our first model, following the idea of packet retransmissions, that is at the root of the famous ALOHA protocol \cite{Abramson1970,Roberts75}.
-% The idea is that when a device sends an uplink message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode in order to wait for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, up-to $m \leq 10$ times), after some small random waiting times.
+% The idea is that when a device sends an up-link message but does not receive its \Ack{} (before a certain delay), instead of going back to sleep mode in order to wait for its next activation (\eg, following the Bernoulli random activation pattern, or waiting for the application to need to send some data), the device will try again to send the same packet, a few times (\eg, up-to $m \leq 10$ times), after some small random waiting times.
 % %
@@
 The second model with retransmissions concluded our collaboration with R.
-Bonnefoi, and resulted to a paper presented at the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
+Bonnefoi, and resulted to a paper presented in the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
 
@@
 In both models of IoT networks presented above, the core idea is to let every dynamic device run a learning algorithm in a fully decentralized way, in order to optimize the entire system.
-It is important to note that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own random activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
-It means that each of them target its own local objective, which is to maximize its cumulated reward, in a selfish way and by being agnostic about the other devices following the same objective.
+We highlight that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own random activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
+This means that each of them target its own local objective, which is to maximize its cumulated reward, in a selfish way and by being agnostic about the other devices following the same objective.
 It is well known in game theory that playing selfishly can be disastrous for the centralized performance measure, one can think of popular ``dilemmas'', such as the \href{https://en.wikipedia.org/wiki/Prisoner%27s_dilemma}{prisoner dilemma}.
@@
 
-% Our work on multi-player bandit was presented at the ALT conference in Lanzarote, Spain, in April $2019$ \cite{Besson2018ALT},
+% Our work on multi-player bandit was presented in the ALT conference in Lanzarote, Spain, in April $2019$ \cite{Besson2018ALT},
 % %
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
-    Demonstration presented at \emph{International Conference on Communication},
+    Demonstration presented in \emph{International Conference on Communication},
     % \footnote{~ICT 2018, see \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}},

### 2019-07-10 2-Chapters/2-Chapter/chapter2.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
-and usually the literature focuses on regular and minimal exponential families (\ie, when the natural parameter space is an non-empty open set in $\R$).
+and usually the literature focuses on regular and minimal exponential families (\ie, when the natural parameter space is a non-empty open set in $\R$).
 
@@
 Finally, we note that Markov bandit models have also been considered for Cognitive Radio applications \cite{Liu08}.
-They are not studied in this thesis, even though we did implemented them in SMPyBandits.
+They are not studied in this thesis, even though we did implement them in SMPyBandits.
 They were introduced in the $1980$s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
@@
 If the goal of the player is to maximize $\E\left[\sum_{t=1}^T r(t)\right]$,
-the optimal strategy for this bandit problem is to always pull one of the optimal arms (it can be not unique), but it is unrealistic as the player does not know the true means, nor the optimal arms (as long as they are not all optimal, \ie, in non trivial problems).
+the optimal strategy for this bandit problem is to always pull one of the optimal arms (it can be not unique), but it is unrealistic as the player does not know the true means, nor the optimal arms (as long as they are not all optimal, \ie, in nontrivial problems).
 %
@@
 When the regret upper-bound matches the lower-bound with the same constant as in the big-$\cO$ notation, we say that the algorithm is asymptotically \emph{optimal},
-otherwise the algorithm is said to be \emph{order-optimal} if it matches the lower-bound with with a larger constant.
+otherwise the algorithm is said to be \emph{order-optimal} if it matches the lower-bound with a larger constant.
 
@@
 On the contrary, for efficient algorithms,
-we can visualize the entire \emph{distribution} as an histogram, or the variance of the values of $R_T^{\cA}$, and if the number $N$ is reasonably large, we can verify that the regret appears logarithmic for all runs.
+we can visualize the entire \emph{distribution} as a histogram, or the variance of the values of $R_T^{\cA}$, and if the number $N$ is reasonably large, we can verify that the regret appears logarithmic for all runs.
 
@@
 %
-It is relevant for cognitive radio applications, as a hardware reconfiguration costs energy, and so changing channel from two consecutive time steps has a non trivial energy cost, as highlighted in \cite{modiDemo2016}.
+It is relevant for cognitive radio applications, as a hardware reconfiguration costs energy, and so changing channel from two consecutive time steps has a nontrivial energy cost, as highlighted in \cite{modiDemo2016}.
 We also find interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
@@
 %
-The player \textcolor{darkgreen}{\emph{only explores}} without using the collected information, and this strategy fails dramatically for any non trivial problem (\ie, linear regret).
+The player \textcolor{darkgreen}{\emph{only explores}} without using the collected information, and this strategy fails dramatically for any nontrivial problem (\ie, linear regret).
 Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
-which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all non trivial problems, ruling out the case where $\mu_1=\mu_2=\dots=\mu_K$).
+which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all nontrivial problems, ruling out the case where $\mu_1=\mu_2=\dots=\mu_K$).
 
@@
 %
-\footnotetext{~For simplicity we present the theorem and its prove in the restricted case of $\alpha=2$.
+\footnotetext{~For simplicity we present the theorem and its proof in the restricted case of $\alpha=2$.
 More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs, in particular for any value of $\alpha\geq\frac{1}{2}$.}%
@@
 Instead of using a confidence radius $\xi_k(t)$ based on Hoeffding's inequality,
-the \klUCB{} algorithm rather uses the Chernoff's concentration inequality for the considered exponential family \cite{chernoff1952measure,chernoff1981note}.
+the \klUCB{} algorithm rather uses Chernoff's concentration inequality for the considered exponential family \cite{chernoff1952measure,chernoff1981note}.
 % such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
@@
 The survey \cite{Arora2012multiplicative} gives more details.},
-% which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019)
+% which comes with an interactive online demonstration in order to better understand the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019)
 a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
@@
 We finish this Section~\ref{sec:2:famousMABalgorithms} by illustrating the performance of different algorithms, on the small MAB problem used in the online demonstration shown above (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
-We consider $K=5$ arms following Bernoulli distributions, of means $\mu_1=0.6,\mu_2=0.2,\mu_3=0.55,\mu_4=0.7,\mu_5=0.5$ (unknown to the algorithms), and first an horizon $T=100$ (like in the online demo) then a larger value of $T=10000$.
+We consider $K=5$ arms following Bernoulli distributions, of means $\mu_1=0.6,\mu_2=0.2,\mu_3=0.55,\mu_4=0.7,\mu_5=0.5$ (unknown to the algorithms), and first a horizon $T=100$ (like in the online demo) then a larger value of $T=10000$.
 The best arm is $\mu_4$ and the smallest gap is $\Delta = 0.1$, thus a lower-bound on the gaps is taken as $d=\Delta$.
@@
 %
-We then include efficient algorithms, that do not need prior knowledge of the $T$ (they are anytime) nor $\Delta$ (their are robust and efficient on unknown problems):
+We then include efficient algorithms, that do not need prior knowledge of the $T$ (they are anytime) nor $\Delta$ (they are robust and efficient on unknown problems):
 $\UCB_1$ (with $\alpha=1$),
@@
 % On the one hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
-It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the $\UCB_1$ algorithm (or other low-cost algorithms).
+It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could choose to only consider the $\UCB_1$ algorithm (or other low-cost algorithms).
 % %
@@
 %
-John is very unique.
 We take a more formal approach in the rest of the thesis,

### 2019-07-10 2-Chapters/3-Chapter/SMPyBandits_paper.git/longpaper.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 repetitions of the default non-Bayesian Bernoulli-distributed problem,
-for \(K=9\) arms, an horizon of \(T=10000\) and on \(4\) CPUs \footnote{It
+for \(K=9\) arms, a horizon of \(T=10000\) and on \(4\) CPUs \footnote{It
   takes about \(20\) to \(40\) minutes for each simulation, on a
@@
   for transforming any non-anytime algorithm to an anytime variant is
-  the ``Doubling Trick'': start with an horizon \(T_0\in\mathbb{N}\),
+  the ``Doubling Trick'': start with a horizon \(T_0\in\mathbb{N}\),
   and when \(t > T_i\), use \(T_{i+1} = 2 T_i\). We studied two generic

### 2019-07-10 2-Chapters/3-Chapter/SMPyBandits_paper.git/paper.tex
Done changing notations \{1,\dots,M\} -> [M] all over the PhD
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 and \emph{i.i.d.} rewards $r_k(t) \sim \nu_k, \forall t$.
-An agent choose arm $A(t)\in\{1,\dots,K\}$ at time $t$ and
+An agent choose arm $A(t)\in[K]$ at time $t$ and
 observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
@@
 start \(N=1000\) repetitions of a simple non-Bayesian
-Bernoulli-distributed problem, for \(K=9\) arms, an horizon of
+Bernoulli-distributed problem, for \(K=9\) arms, a horizon of
 \(T=10000\) and on \(4\) CPUs. Such simulation takes about \(20\) minutes, on a standard \(4\)-cores \(64\) bits GNU/Linux
@@
   algorithm to an anytime variant is the ``Doubling Trick'': start with
-  an horizon \(T_0\in\mathbb{N}^*\), and when \(t > T_i\), use
+  a horizon \(T_0\in\mathbb{N}^*\), and when \(t > T_i\), use
   \(T_{i+1} = 2 T_i\). We studied two generic sequences of growing

### 2019-07-10 2-Chapters/3-Chapter/chapter3.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 %
-The code for \SMPyBandits{} is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source License.
+The code for \SMPyBandits{} is hosted at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}}, and all the library is freely publicly released under the MIT open-source license.
 This chapter is based on the article \cite{SMPyBanditsJMLR}.
@@
     % \item
-    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm).
+    Then, for $t=1$ to $t=T$ (known beforehand), repeat the following block (for each algorithm).
 Ask algorithm $\cA$ her chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm.
@@
     % \item
-    For $t=1$ to $t=T$ (known before hand), repeat the following block (for each cohort).
+    For $t=1$ to $t=T$ (known beforehand), repeat the following block (for each cohort).
 Ask each player $\cA^j$ her chosen arm $A^j(t)$, then query the collision model to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
@@
     \end{minted}
-    \caption{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with \SMPyBandits.}
-    % \captionof{listing}{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, a plot a histogram of rewards, with SMPyBandits.
+    \caption{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, and to plot a histogram of rewards, with \SMPyBandits.}
+    % \captionof{listing}{Example of Python code to create Bernoulli and Gaussian arms, a MAB problem with $K=3$ arms, and to plot a histogram of rewards, with SMPyBandits.
 \label{lst:3:pythonCodeCreateProblem}}
@@
 
-Simulations are easily executed, \eg, Code~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, an horizon of $T=10000$ and on $4$ CPU.
+Simulations are easily executed, \eg, Code~\ref{lst:3:howToRunBasicLibrary} shows how to start $N=1000$ repetitions of a simple non-Bayesian Bernoulli-distributed problem, for $K=9$ arms, a horizon of $T=10000$ and on $4$ CPU.
 It takes about $20$ min, on a $4$-core $64$-bit GNU/Linux laptop.
@@
 in order to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments.
-The tests covers all the main models and almost all the algorithms implemented in \SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
+The tests cover all the main models and almost all the algorithms implemented in \SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
 For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
@@
 % In sequential learning, an algorithm is said to be ``anytime'' if it does not need to know the horizon $T$ of the experiments.
-A well-known trick for transforming any ``non-anytime'' algorithm to an ``anytime'' variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
+A well-known trick for transforming any ``non-anytime'' algorithm to an ``anytime'' variant is the ``Doubling Trick'': start with a horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$.
 We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results.
@@
     It mimics the \KLUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
-    It is anytime, and was prove to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
+    It is anytime, and was proven to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
     Empirically, it usually outperforms (slightly) \klUCB{} in terms of regret, and costs about the same memory, but it is slower.
@@
     It was also proven to obtain good ``best of both world'' performances.
-    Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{~Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
+    Empirically, we were unable to find any problem where it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{~Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations.
 Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connection between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in \SMPyBandits{} will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
@@
 % %
-% As long as the number of jobs (\texttt{N\_JOBS} here) is less then or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
+% As long as the number of jobs (\texttt{N\_JOBS} here) is less than or equal to the number of physical cores in the CPU of the computer, the final speed-up in terms of total computation runtime is almost optimal.
 % %  dividing the running time by \texttt{N\_JOBS}.
@@
 % % %
-% % The two other approaches we could have consider is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
+% % The two other approaches we could have considered is parallel computations running on not multiple cores but multiple machines, in a computer cluster, or parallel computations running in a Graphical Processing Unit (GPU).
 

### 2019-07-10 2-Chapters/5-Chapter/chapter5.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 
-We consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
+We consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and a horizon of $T=10000$.
 As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require to know $M$.

### 2019-07-10 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 A \emph{piece-wise stationary bandit model} generalizes the model of Chapter~\ref{chapter:2}.
-It is characterized by a set of $K$ arms, and an horizon $T$.
+It is characterized by a set of $K$ arms, and a horizon $T$.
 A (random) stream of rewards $(Y_{k,t})_{t\in[T]}$ is associated to each arm $k \in [K]$.
@@
 In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
-it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\Delta^{\text{change}} = \max_{k,t} |\mu_k(t) - \mu_k(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\Delta^{\text{opt}} = \max_{k,k'} |\mu_k(t) - \mu_{k'}(t)|$.
+it only needs to know beforehand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\Delta^{\text{change}} = \max_{k,t} |\mu_k(t) - \mu_k(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\Delta^{\text{opt}} = \max_{k,k'} |\mu_k(t) - \mu_{k'}(t)|$.
 

### 2019-07-10 5-Nomenclature/Nomenclature.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

@@
 \nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
-\nomenclature[G]{$\delta$}{Usually denotes a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}.
+\nomenclature[G]{$\delta$}{Usually denotes a lower-bound on the gap $\Delta$, known beforehandby an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}.
 Also denote a confidence level of an algorithm in Section~\ref{sec:6:ChangePointDetector}}

### 2019-07-10 2-Chapters/2-Chapter/chapter25.tex
Merci Christophe de toutes ces remarques !
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
 On the one hand, a practitioner can decide to pick one algorithm, maybe because it seems efficient on other problems, or maybe because it is simple enough to be used in its application.
-It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could chose to only consider the $\UCB_1$ algorithm (or other low-cost algorithms).
+It might be unrealistic to implement complicated algorithms on limited hardware such as embedded chips in a very low-cost IoT end-device, and for instance a practitioner could choose to only consider the $\UCB_1$ algorithm (or other low-cost algorithms).
 %
@@
 % regret if it aggregates at least one algorithm with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
-% It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
+% It usually succeeds to be close to the best of the aggregated algorithms, both in terms of regret and best arm pull frequency.
 % As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this is an over-optimistic goal.

### 2019-07-10 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
 The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
-$\sum_{t = 1}^T r(t),$
-which is the usual objective in MAB problems (see Section~\ref{sec:2:notations}).
+$\sum_{t = 1}^T r(t)$, which is the usual objective in MAB problems (see Section~\ref{sec:2:notations}).
 
@@
 % $\Pr(\text{no other dynamic device sent in}\;i)$,
-is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device } k \text{ did not sent in } i) = \prod_{k=1}^{D-1} \Pr(\text{device } k \text{ did not sent in } i)$.
+is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device } k \text{ did not send in } i) = \prod_{k=1}^{D-1} \Pr(\text{device } k \text{ did not send in } i)$.
 And $\Pr(\text{device } k \text{ sent in } i) = p \times 1 / K$, by uniform choice on channels and the Bernoulli emission hypothesis.
@@
 We can see that when the proportion of end-devices is low (\eg, $1\%$ of devices are dynamic), the optimal policy provides an improvement of $16\%$ compared to random channel selection.
-The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a lowest rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$),
+The TS algorithm always provides near-optimal performance, but the \UCB{} algorithm has a slower rate of convergence and performs consequently worse after $1000$ transmissions, for instance it only provides a gain of $12\%$ for the same proportion of dynamic devices ($1\%$),
 for the considered values of $S_1,\dots,S_K$.

### 2019-07-10 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
 The idea behind this delay is to allow the first stage \UCB{} to start learning the best channel, before starting the second stage \UCB{} (see details in Algorithm~\ref{algo:43:UCBwithDelay}).
-The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to simulate the problem before-hand to find a good value of some internal parameter.
+The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: a heuristic should work against any problem, without the need to simulate the problem before-hand to find a good value of some internal parameter.
 As such, we only consider a delay of $\Delta=100$ in our experiments, and we did not try to optimize it.}.
@@
 % Similarly, Philippe Mary asked me at the MoTION workshop where I presented \cite{Bonnefoi2019WCNC}, why didn't we rather used Bayes-UCB instead of \UCB{} (from \cite{Kaufmann12BUCB}), as it is known to empirically outperform \UCB{} (as illustrated for instance in \cite{darak2016bayesian,kumar2016two}).
-We could have chose any bandit algorithm for the ``first stage'' component used in this Section, and for simplicity and clarity we focused on a simple but efficient one (\UCB).
+We could have chosen any bandit algorithm for the ``first stage'' component used in this Section, and for simplicity and clarity we focused on a simple but efficient one (\UCB).
 We believe that the same empirical results, but more importantly, the same conclusions, could be given if we used Bayes-UCB or other bandit algorithm for the base building block in the different heuristics proposed in this section.
@@
 
-We also consider in these experiment the case of an ALOHA protocol using $\mathrm{MaxBackOff}=5$, and $m=10$, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
+We also consider in these experiments the case of an ALOHA protocol using $\mathrm{MaxBackOff}=5$, and $m=10$, a statistic distribution of the devices about $40\%, 30\%, 20\%, 10\%$ for the four channels, and $N=2000$ IoT devices.
 The corresponding results are depicted in Figure~\ref{fig:43:mainExperiment2}.

### 2019-07-10 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
 
-We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
+We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in an acknowledgment-based wireless protocol slotted in time.
 A perfect time and frequency synchronization is assumed,
@@
 The goal for every player is to select most of the time one of the $M$ best arms, without colliding too often with other players.
-A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all the arms to estimate their means, while trying to focus on the best arms to gain as much rewards as possible.
+A first difficulty relies in the well-known trade-off between \emph{exploration} and \emph{exploitation}: players need to explore all the arms to estimate their means, while trying to focus on the best arms to gain as many rewards as possible.
 The decentralized setting considers wireless protocol with no direct or explicit exchange of information between players, and assumes that the players only know $K$ and $M$.
@@
 % Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model,
-for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
+for which we only proposed a heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
 Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, it is a mistake to only consider mean regret, as we found on numerical simulations as well as formal derivation on a simple example of $K=M=2$ that the \Selfish{} heuristic can have a linear regret with a low probability (and so asymptotically it has linear expected regret and fails to solve ``no sensing'' multi-players MAB problems).
@@
 % We are also interested in the harder model where devices do not have access to the two feedback information (sensing and \Ack) but only have access to the \Ack.
-% Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed an heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
+% Our article \cite{Besson2018ALT} was the first to propose this ``no sensing'' model, for which we only proposed a heuristic, the naive \Selfish{} strategy as it was already used in Chapter~\ref{chapter:4}.
 % Even if empirical simulations showed that \Selfish-\klUCB{} performs very well, we conjectured that it suffers from linear regret, and it was later confirmed by two works inspired by our article \cite{LugosiMehrabian18,BoursierPerchet18}.
@@
 It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channel.
-% This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.
+% This extension is not studied per se but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.
 
@@
 
-% % concrete examples and how their are used within MPB
+% % concrete examples and how they are used within MPB
 
@@
 then she marks her current arm as a ``chair'' ($s^j(t)=\mathrm{True}$),
-and will keep using it even if collisions happen in the future (Lines~$9$-$11$).
+and will keep using it even if collisions happen in the future (lines~$9$-$11$).
 %
@@
 a new arm is sampled uniformly from a subset of $\TopM(t)$,
-defined with the previous indices $U^j(t)$ (Lines~$3$-$5$).
+defined with the previous indices $U^j(t)$ (lines~$3$-$5$).
 %
@@
 %
-In this case, a new arm is sampled uniformly from $\TopM(t)$ (Lines~$6$-$8$).
+In this case, a new arm is sampled uniformly from $\TopM(t)$ (lines~$6$-$8$).
 %
@@
 the number of switches of arm that are due to the current arm leaving $\TopM(t)$,
-for both \RandTopM{} and \MCTopM{}. It essentially proves that Lines~$3$-$5$ in Algorithm~\ref{algo:5:MCTopM} (when a new arm is sampled from the non-empty subset of $\TopM(t)$)
+for both \RandTopM{} and \MCTopM{}. It essentially proves that lines~$3$-$5$ in Algorithm~\ref{algo:5:MCTopM} (when a new arm is sampled from the non-empty subset of $\TopM(t)$)
 happen a logarithmic number of times. The proof of this result is given below.
@@
 We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \UCB{} or \klUCB{} indices.
-The analysis given in Section~\ref{sec:5:upperbounds} was Focusing on \klUCB, because it is well known that it is more efficient both in practice and in theory.
+The analysis given in Section~\ref{sec:5:upperbounds} was focusing on \klUCB, because it is well known that it is more efficient both in practice and in theory.
 To illustrate this, we first show below the result of some experiments comparing different algorithms that use either the \UCB{} or the \klUCB{} indexes.
@@
 With a perfect knowledge on the horizon and the gap ($\Delta=0.1$ here) and by using the parameters suggested from their respective articles, \MEGA{} and \MusicalChair{} perform badly, even in this simple setting.
-The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can bounded with probability $0.5$ and $0.9$ respectively.
+The first two \MusicalChair{} instances use the optimal $T_0$ value from \cite{Rosenski16}, with $\varepsilon$ taken slightly smaller than the gap $\Delta$ ($\varepsilon=0.99 \Delta$), and respectively with $\delta=0.5$ and $\delta=0.1$, for which the regret can be bounded with probability $0.5$ and $0.9$ respectively.
 The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.

### 2019-07-10 2-Chapters/5-Chapter/chapter5.tex
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
 %
-We can start by asking whether the most efficient algorithm, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different than the real number of players $M$.
+We can start by asking whether the most efficient algorithm, \MCTopM-\klUCB, is still efficient if it uses a value $M'$ different from the real number of players $M$.
 Even though we did not include numerical simulations for this case in Section~\ref{sec:5:experiments} above, we did some tests, which confirmed two disappointing results that were also proven analytically.
@@
 We consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and a horizon of $T=10000$.
-As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require to know $M$.
+As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as a heuristic which does not require to know $M$.
 Then we compare the \RhoRand, \RandTopM{} and \MCTopM{} algorithms, using \klUCB, which know $M$ beforehand, with their extensions implementing the same algorithm as \rhoRandEst, to estimate $M$ on the fly.
@@
 This trick essentially use the fact that the players share a synchronized time, thus they can use a collision as a way to directly exchange one bit of information between two players.
-This process is slow and not efficient, as all but two players must do nothing when player $i$ is sending a bit to player $j$, but it does build a communication protocol in the model without explicit communication between players.
-This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
-We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.
+Such process is slow and not efficient, as all but two players must do nothing when player $i$ is sending a bit to player $j$, but it does build a communication protocol in the model without explicit communication between players.
+We note that this ``communication trick'' is used for instance in \cite{KaufmannAbbas19},
+and we let the interested reader refer to these two recent works \cite{BoursierPerchet18,KaufmannAbbas19}.
+%  as they are both solid, and well explained.
 
@@
 As above, we consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $N=100$ independent repetitions, and horizon $T=10000$.
-We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
+We include the \Selfish-\klUCB{} algorithm as a heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
 It is also interesting to add the centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, because it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
@@
 
-Moreover, at first sight, this idea of ``meta'' arms implies an exponential blow up in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
+Moreover, at first sight, this idea of ``meta'' arms implies an exponential blowup in terms of computational and storage cost, as there is ${M \choose K}$ such ``meta arms'', but
 the authors provide in Section~4.1 and Lemma~4.1 \cite{AlaturLevyKrause19} an interesting discussion regarding the time efficiency of their proposed algorithm, and they show it can stay polynomial in $M$ and $K$ by leveraging techniques from the Determinental Point Processes (DPP, see \cite{GaBaVa18} and references therein for a good introduction).

### 2019-07-10 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
 Note than when a break-point occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
-Depending on the application, many scenario can be meaningful: changes can occur on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each break-point.
+Depending on the application, many scenarios can be meaningful: changes can occur on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each break-point.
 For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the devices of one company (or network provider) use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new devices, then one (or some) channel(s) see their mean occupancy rates abruptly change.
@@
 Then, we compare our algorithms to several competitors designed for a piece-wise stationary model.
-For a fair comparison, all algorithms that use  \UCB{} as a sub-routine were adapted to use \klUCB{} instead, which yields better performance\footnote{\cite{LiuLeeShroff17,CaoZhenKvetonXie18} both mention that extending their analysis to the use of \klUCB{} should not be too difficult.}.
+For a fair comparison, all algorithms that use  \UCB{} as a sub-routine were adapted to use \klUCB{} instead, which yields better performance\footnote{~\cite{LiuLeeShroff17,CaoZhenKvetonXie18} both mention that extending their analysis to the use of \klUCB{} should not be too difficult.}.
 Moreover, all the algorithms are tuned as described in the corresponding paper, using in particular the knowledge of the number of break-points $\Upsilon_T$ and the horizon $T$.

### 2019-07-10 3-Appendices/2-Appendix/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.git/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.tex
Using \begin{framed}...\end{framed} for all algorithms.
New references. I need to finish Ch1 and Ch2
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
         Update the literature review:
-        Talk about the \KLUCBswitch{} algorithm from \cite{Garivier18}.
+        Talk about the \KLUCBswitch{} algorithm from \cite{GarivierHadiji2018}.
     \end{itemize}
@@
 Complementary elements of proofs are deferred to the appendix.
-% We give missing proofs and additional results in Appendix~\ref{sec:missingproofs}, an heuristic $\DTnr$ improving $\DTr$ in App.~\ref{sec:DTnr} along with more experimental results in App.~\ref{sec:otherExperiments}, and other technical tools are given in App.~\ref{sec:otherproofs}.
+% We give missing proofs and additional results in Appendix~\ref{sec:missingproofs}, a heuristic $\DTnr$ improving $\DTr$ in App.~\ref{sec:DTnr} along with more experimental results in App.~\ref{sec:otherExperiments}, and other technical tools are given in App.~\ref{sec:otherproofs}.
 
@@
 % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
+\begin{framed}
 \begin{algorithm}[H]
@@
 \end{algorithm}
+\end{framed}
 % \vspace*{-5pt}  % XXX remove if problem
@@
 We first provide some details on these two algorithms, and then illustrate the behavior of Doubling Tricks applied to these algorithms with different doubling sequences.
-% \cite{Garivier18}
+% \cite{GarivierHadiji2018}
 
@@
 
-Another version of the Doubling Trick with ``no restart'', denoted $\DTnr$, is presented in Appendix~\ref{sec:DTnr}, but it is only an heuristic and cannot be applied to any algorithm $\cA$.
+Another version of the Doubling Trick with ``no restart'', denoted $\DTnr$, is presented in Appendix~\ref{sec:DTnr}, but it is only a heuristic and cannot be applied to any algorithm $\cA$.
 Algorithm~\ref{algo:DTnr} can be applied to \KLUCBpp{} or \AFHG{} for instance, as they use $T$ just as a numerical parameter (see Eqs.~\ref{eq:index_AFHG} and \ref{eq:index_KLUCBpp}), but its first limitation is that it cannot be applied to \DMEDp{} \citep{Honda10} or \Exppp{} \citep{Seldin17}, or any algorithms based on arm eliminations, for example.
@@
 As mentioned previously, the \DTnr{} algorithm (Algorithm~\ref{algo:DTnr})
-is only an heuristic so far, as no theoretical guarantee was proved for it.
+is only a heuristic so far, as no theoretical guarantee was proved for it.
 %

### 2019-07-10 5-Nomenclature/Nomenclature.tex
Last pass with http://emccabedj.ucc.asn.au/checktex.html

@@
 \nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
-\nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda},$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
+\nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda}$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
 \nomenclature[G]{$\Delta$}{Usually denotes the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}.

### 2019-07-10 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Using https://sylvainhalle.github.io/textidote/

@@
 
-This PoC implements an IoT network the following way: one base station, one or several intelligent (\emph{i.e.}, learning) devices, embedding the proposed solution,
+This PoC implements an IoT network the following way: one base station, one or several intelligent (\ie, learning) devices, embedding the proposed solution,
 and a traffic generator that emulates radio interference from many other devices.

### 2019-07-10 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Using https://sylvainhalle.github.io/textidote/

@@
   Recall that we want to upper bound
-  $ \ref{eq:5:term2} : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[N_k(T)]\right)$.
+  $ \ref{eq:5:term2} \eqdef \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[N_k(T)]\right)$.
   First, we observe that, for all $k\in \Mbest$,

### 2019-07-10 2-Chapters/5-Chapter/chapter5.tex
Using https://sylvainhalle.github.io/textidote/

@@
 This second approach works empirically very fine, but it requires a fine tuning of the stopping time $T_0$.
-Rosenski et al. studies the performance of their algorithm with high-probability bounds \cite{Rosenski16}, and the tuning of $T_0$ they propose depends on prior knowledge on the problem.
+J.
+Rosenski and O.
+Shamir and L.
+Szlak study the performance of their algorithm with high-probability bounds \cite{Rosenski16}, and the tuning of $T_0$ they propose depends on prior knowledge on the problem.
 %

### 2019-07-10 2-Chapters/1-Chapter/chapter1.tex
A couple of changes, thanks to https://github.com/get-alex/alex#command-line

@@
 % https://en.wikipedia.org/wiki/History_of_radio
-In the last $150$ years, mankind has developed many different communication technologies, and since the late $1890$s, wireless communications between manufactured devices have been made possible, and more and more frequent in our lives.
-With the advent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, allowing a large range of different applications.
+In the last $150$ years, humankind has developed different communication technologies, and since the late $1890$s, wireless communications between manufactured devices have been made possible, and more and more frequent in our lives.
+With the advent of Internet of Things networks (IoT), billions of autonomous low-power devices are expected to be deployed worldwide, allowing a wide range of different applications.
 It is now a world-wide consensus that with the current trend of population increase and with the ongoing energy crisis, any newly deployed technology should be both cheap and energy efficient,
@@
 %
-In addition to the usual Research and Development effort to design many different efficient radio access schemes, covering all the possible cases,
+In addition to the usual Research and Development effort to design different efficient radio access schemes, covering all the possible cases,
 now the time has come to combine it with cheap and promising Machine Learning in order to (try to) attain the level of performance gain necessary for the IoT promises to become reality.

### 2019-07-10 2-Chapters/2-Chapter/chapter2.tex
A couple of changes, thanks to https://github.com/get-alex/alex#command-line

@@
 
-\textbf{Clinical trials} have been the first historical application of MAB models, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the drug did not heal the disease, and a reward of $1$ indicates a success.
+\textbf{Clinical trials} have been the first historical application of MAB models, where an arm represents a treatment, and the distribution associated with such treatment can be a Bernoulli distribution: a reward of $0$ means the treatment did not heal the disease, and a reward of $1$ indicates a success.
 The mean of an arm, in this application, represents the mean success rate of a treatment.

### 2019-07-10 2-Chapters/3-Chapter/chapter3.tex
Presque terminé cette fichue conclusion

@@
 
-We chose to present here the Python library \SMPyBandits{} in this chapter, rather than in the Appendix, because all the numerical experiments on single-player multi-arm bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on \SMPyBandits.
+We chose to present here the Python library \SMPyBandits{} in this chapter, rather than in the Appendix, because all the numerical experiments on single-player multi-armed bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on \SMPyBandits.
 %

### 2019-07-10 2-Chapters/5-Chapter/chapter5.tex
Presque terminé cette fichue conclusion

@@
 
-To sum up, we presented in this chapter three variants of Multi-Players Multi-Arm Bandits,
+To sum up, we presented in this chapter three variants of Multi-Players Multi-Armed Bandits,
 with different levels of feedback being available to the decentralized players, under which we proposed efficient algorithms.

### 2019-07-10 2-Chapters/Conclusion/conclusion.tex
Presque terminé cette fichue conclusion

@@
 % ----------------------------------------------------------------------------
-\section{Perspectives}
+% \section{Perspectives}
+\subsection*{Future works}
 
 % As discussed at the end of each chapter,
-The works presented in each chapter suggest possible directions of future studies.
+% All the models, simulations,
+What we presented above in each chapter suggest the following directions of future studies.
 
@@
 
-As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on MAB problems with a finite number of arms, and binary, bounded or real-valued rewards.
-Thus two possible directions of future research could be to extend our works
-to MAB models with either multi-dimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
-The hypotheses we made were motivated by the target application, but these other models could also be used for cognitive radio applications.
+% As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on MAB problems with a finite number of arms, and binary, bounded or real-valued rewards.
+% Thus two possible directions of future research could be to extend our works
+% to MAB models with either multi-dimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
+% The hypotheses we made were motivated by the target application, but these other models could also be used for cognitive radio applications.
 
@@
 Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
-\\
-\indent
+% \\
+% \indent
 Another extension could be to consider not a Bernoulli activation process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
@@
 % that nevertheless have not been explored for the targeted problem.
-\\
-\indent
+% \\
+% \indent
+
 In order to validate our results in a realistic experimental setting and not only with simulations, future works should include a hardware implementation of the analyzed models, in order to complete our demonstration presented in Section~\ref{sec:4:gnuradio} \cite{Besson2018ICT,Besson2019WCNC}.
@@
 We also presented many extensions of the multi-players bandit model in Section~\ref{sec:5:literatureReviewOtherModels},
-and even if some have already been implemented, a major future work is to implement the most interesting ones in SMPyBandits
-(see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}), as well as extending our theoretical analyses for such extensions of the models considered in this thesis.
+and even if some have already been implemented,
+an important future work is to implement the most interesting ones in SMPyBandits
+(see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
+From the point-of-view of the theoretical analysis, we are especially interesting about extending our proposed algorithm \MCTopM{} to the case of an unknown number of player, and then to the most interesting extension for the application to wireless networks, the ``dynamic case'' which allows for arrivals and departures of players in the multi-player bandit game.
+%
+% as well as extending our theoretical analyses for such extensions of the models considered in this thesis.
 
@@
 
-We believe that the new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
+Regarding non-stationary bandits, we believe that the new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
 This promising research direction is left for future work,  but the hope is that the regret would be expressed in terms of the number of such meaningful changes, instead of the number of break-points $\Upsilon_T$.
-We shall also investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
+We shall moreover investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
 We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of more efficient algorithms, as suggested by \cite{Maillard2018GLR}.
-\\
-\indent
-We would also like to study in the future possible extension of our approach to the slowly-varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or also to the rotting bandit model \cite{Seznec2018}, which is interesting for many applications even if it does not seem appropriate for the cognitive radio setting.
+% \\
+% \indent
+Another very interesting future work is to study possible extensions of our approach, especially to the slowly-varying model, as studied in \cite{Besbes14stochastic,Louedec16,WeiSrivastava18Abruptly}, or to other models such as the rotting bandit model \cite{Seznec2018}, which is interesting for many applications even if it does not seem directly usable for the cognitive radio setting.
 
 % ----------------------------------------------------------------------------
-\subsection*{Other directions of future works}
-
+% \subsection*{Other directions of future works}
+\paragraph{Unifying multi-player and non-stationary bandits.}
+%
+%
 Another interesting direction of future work is to study non-stationary distributed multi-players bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.

### 2019-07-10 2-Chapters/Conclusion/conclusion.tex
Too long. A short break, and I proofread and compress

@@
 % ----------------------------------------------------------------------------
-\section{General Conclusion}
+% \section{General Conclusion}
+\subsection*{General Conclusion}
 
@@
 
-This manuscript was organized in two parts.
+We started this manuscript by detailing the historical, scientific and technical contexts of our research during this PhD, which studied how decentralized low-cost machine learning technique can be used to automatically improve a certain aspect of wireless telecommunications.
 %
-The main problem we studied was \emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
+We explained the problems that motivate and justify our works,
+and we presented the main contributions produced during the last three years.
 %
-We started by detailing the historical, scientific and technical contexts of the problems we considered for our research, and presented the main contributions produced during the last three years.
+The main question we studied was \emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
+%
+This PhD manuscript is organized in two parts.
+It first focuses on giving a strong theoretical background on the reinforcement learning model considered in this thesis, that are the multi-armed bandits models.
+The manuscript then focuses on our problems of solving the spectrum scarcity issue in unlicensed bands, in the context of the future Internet of Things networks,
+and extending to the specificities of such IoT networks the ideas underlying previously studied applications of machine learning to solve the spectrum scarcity issue in licensed networks.
 
-% % \begin{itemize}
-%     % \item
-% The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
-% % Chapter 2
-% In Chapter~\ref{chapter:2}, we presented the multi-armed bandit (MAB) models, the hypotheses we usually make on such models, and the most important algorithms that solve different kinds of MAB problems.
-% % Chapter 3
-% Then in Chapter~\ref{chapter:3} we showcased our open-source Python library SMPyBandits in details, developed to run numerical simulations of MAB problems.
-% %
-% Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an empirically efficient answer to the online algorithm selection problem.
+Our main contribution to answer the problematic is a model of Internet of Things networks, based on an ALOHA-like protocol slotted in time and frequency, and that considers one IoT base station, serving a large number of devices following the IoT constraints (low-cost, low duty cycle, long range etc).
+In such IoT network, we focused on many dynamically reconfigurable IoT end-devices, that are able to run low-cost and low-complexity decision making algorithms, embedded on each device using their limited computational and storage capacity (limited hardware)
+We proposed two models, whether the considered IoT standard asks the devices to try to retransmit a few times their up-link packet in case of a failed transmission (\ie, a collision with other devices),
+or to drop their packet after a failed transmission (\ie, without retransmission).
+Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order for the numerous IoT devices to improve on their own their spectrum access scheme, by learning which of the $K$ frequency channels are the less occupied by the environmental trafic.
 
-%     % \item
-% The second half of this thesis, Part~\ref{part:MABIOT}, presented the main contributions.
-% % Chapter 4
-% In Chapter~\ref{chapter:4}, we first proposed models for realistic IoT wireless networks, in a discrete time and with no central coordination between end-devices.
-% In two different models, with or without retransmissions, we showed that end-devices can independently learn favor the less occupied radio channels, in order to increase their successful transmission rates in the network, by embedding low-cost decentralized MAB algorithms.
-% We also presented a proof-of-concept that implements such network on real radio hardware and validates the proposed model and algorithms.
-% % Chapter 5
-% Because the models proposed for realistic IoT networks were found to be too general to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
-% we started to study a weaker but still interesting model in Chapter~\ref{chapter:5}.
-% We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision and spectrum sensing information (\MCTopM).
-% % Chapter 6
-% Finally, in Chapter~\ref{chapter:6}, we relaxed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
-% %
-% For the last two chapters, the proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in their theoretical analyses.
-% % \end{itemize}
+The approach we thus advocate is to use low-complexity MAB algorithms such as \UCB, or variants tuned to be robust in slowly-evolving non-stationary scenario (\eg, Discounted-\UCB{} or actively adaptive approach such as \GLR-\klUCB).
+For a given company manufacturing IoT end-devices, it is very simple and cheap in terms of both hardware and software engineering workloads to implement such MAB algorithm on the low-cost embedded processors, and equip all the produced IoT devices with this capacity of using online machine-learning for their spectrum access.
+The embedded decision-making algorithm will automatically improve the channel selection of each device, in a fully decentralized way, and will increase the total number of successful up-link transmissions, thus it will increase the Quality of Service of the considered IoT application, as well as allowing more devices to be served by the same IoT gateway.
+%
+The advocated approach also has the advantage of not requiring the change anything from the IoT standard side, as the only modification happens on the device side, by letting it actively decide the channels it uses for up-link transmissions, instead of relying on a naive uniform channel access.
+%
+We also note that this approach can easily be used to let the IoT devices optimize by themselves and on the fly other parameters of their wireless communications, as the multi-armed bandit framework is not restricted to be applied for the frequency channel selections but can be applied to any exploration-exploitation problem with a finite set of possible options.
+The solution we propose can thus be easily extended to use the same kinds of MAB algorithms to select dynamically parameters such as the coding rate or spreading factor of a LoRa network (\eg, \cite{KerkoucheAlami18}), the emission power in a NOMA standard, or any other discrete-valued parameters that have an impact on the successful transmission rate and can be optimized on the device's side by using reinforcement learning.
 
-% \paragraph{Summary of each chapter.}
-% A more precise summary of each chapter is given below.
+Moreover, from a theoretical aspect, even though it was found hard to propose a rigorous analysis of the model of IoT networks, because of the large number of end-devices having different random and unpredictable activation patterns, our contributions also include an analysis of two restricted models.
+%
+On the one hand, we considered a model where the devices have data to transmit at each time step, relaxing the hypothesis of a random activation process (with a small probability, as IoT devices have low duty cycles), and thus by restricting to the case of at most $M \leq K$ devices in a wireless standard with $K$ orthogonal frequency channels.
+This first model is similar to the multi-player multi-armed bandits studied in the last $10$ years, and for stationary and stochastic environments we were able to propose a state-of-the-art algorithm, \MCTopM.
+Our proposed algorithm performs very well in practice and obtains interesting guarantees on its regret, as well as on the number of radio collisions and on radio reconfigurations that the devices will suffer if they all implement our solution.
+%
+On the other hand, we studied a model that focuses on only one device but relaxes the hypothesis of stationarity for the surrounding radio trafic, and considers that is can be stationary in consecutive intervals of time, of unknown durations.
+This second model is similar to the piece-wise stationary (or abruptly-changing) multi-armed bandits studied since the $2000$s, and we solved it by proposing a new actively adaptive bandit policy, for which we obtained state-of-the-art results, in terms of its regret but also its false alarm probability and delay of detection.
+
+Finally, from an applicative point-of-view, we also presented a realistic implementation of a proof-of-concept of our first model of IoT network, that was used to validate empirically the proposed approach.
+We showed that the IoT devices can effectively use a multi-armed bandit algorithm to automatically learn to favor certain frequency channels over others, in order to optimize their spectrum access, to reduce their number of failed transmissions.
+%
+Our proposed approach is to use decentralized reinforcement learning algorithms, directly embedded on the IoT end-devices, in order to improve their spectrum access and channel selection schemes.
+We confirmed from both a numerical simulations and an validation on real radio trafic and hardware that our proposal is an excellent candidate to start to solve the spectrum scarcity issue for unlicensed bands, that is cheap and easy to deploy, and requires .
+
+% % % \begin{itemize}
+% %     % \item
+% % The first half, Part~\ref{part:Introduction}, started by motivating our work and situating the context of this research.
+% % % Chapter 2
+% % In Chapter~\ref{chapter:2}, we presented the multi-armed bandit (MAB) models, the hypotheses we usually make on such models, and the most important algorithms that solve different kinds of MAB problems.
+% % % Chapter 3
+% % Then in Chapter~\ref{chapter:3} we showcased our open-source Python library SMPyBandits in details, developed to run numerical simulations of MAB problems.
+% % %
+% % Finally, we presented in Chapter~\ref{chapter:25} a first mathematical contribution, our algorithm \Aggr, as an empirically efficient answer to the online algorithm selection problem.
+
+% %     % \item
+% % The second half of this thesis, Part~\ref{part:MABIOT}, presented the main contributions.
+% % % Chapter 4
+% % In Chapter~\ref{chapter:4}, we first proposed models for realistic IoT wireless networks, in a discrete time and with no central coordination between end-devices.
+% % In two different models, with or without retransmissions, we showed that end-devices can independently learn favor the less occupied radio channels, in order to increase their successful transmission rates in the network, by embedding low-cost decentralized MAB algorithms.
+% % We also presented a proof-of-concept that implements such network on real radio hardware and validates the proposed model and algorithms.
+% % % Chapter 5
+% % Because the models proposed for realistic IoT networks were found to be too general to analyze from a theoretical point of view, mainly due to the large number of algorithms interacting at random times in an unpredictable environment,
+% % we started to study a weaker but still interesting model in Chapter~\ref{chapter:5}.
+% % We presented different variants of the multi-players MAB model, and we proposed the state-of-the-art algorithm for the variant with collision and spectrum sensing information (\MCTopM).
+% % % Chapter 6
+% % Finally, in Chapter~\ref{chapter:6}, we relaxed the stationary hypothesis on the single-player MAB problem, and we analyzed a new actively adaptive algorithm, for which we also proved state-of-the-art results (\GLRklUCB).
+% % %
+% % For the last two chapters, the proposed algorithms attain state-of-the-art performances both on numerical simulations and on the regret upper bounds obtained in their theoretical analyses.
+% % % \end{itemize}
+
+% % \paragraph{Summary of each chapter.}
+% % A more precise summary of each chapter is given below.
+
+% % % ----------------------------------------------------------------------------
+% % \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:1}}.}
 
 % % ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:1}}.}
+% % \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:2}}.}
 
-% ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:2}}.}
+% We started in Chapter~\ref{chapter:2} by presenting the multi-armed bandit model, with a finite number of arms and real-valued rewards.
+% Our focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
+% The notations used in the manuscript were introduced by showcasing an interactive demonstration.
+% % \footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
 
-We started in Chapter~\ref{chapter:2} by presenting the multi-armed bandit model, with a finite number of arms and real-valued rewards.
-Our focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
-The notations used in the manuscript were introduced by showcasing an interactive demonstration.
-% \footnote{See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}.
+% % ----------------------------------------------------------------------------
+% % \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:3}}.}
 
-% ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:3}}.}
+% Then we presented in Chapter~\ref{chapter:3} our Python library SMPyBandits \cite{SMPyBanditsJMLR,SMPyBandits}, by detailing
+% its purpose and qualities, its organization, and an overview of its usage.
+% %
+% This library allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
+% SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license (MIT License), and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).\\
+% % Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.\\
+% \indent
+% We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
+% Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions (and rested/restless Markov models).
+% Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
+% One of the main qualities of our library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
+% % We indeed followed actively the state-of-the-art research, since December $2016$ to June $2019$.
+% More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary case.
+% The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
+% %
+% When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
+% %
+% Our library is used to run numerical simulations in Chapters~\ref{chapter:2}, \ref{chapter:3}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
 
-Then we presented in Chapter~\ref{chapter:3} our Python library SMPyBandits \cite{SMPyBanditsJMLR,SMPyBandits}, by detailing
-its purpose and qualities, its organization, and an overview of its usage.
-%
-This library allows any researcher to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
-SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license (MIT License), and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).\\
-% Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.\\
-\indent
-We detailed how SMPyBandits is implementing arms, problems, algorithms, and how the library uses these components to implement a simulation loop, with various visualizations being performed after a simulation.
-Even if SMPyBandits is still restricted to the finite-arm case, it supports a wide range of arm distributions (and rested/restless Markov models).
-Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
-One of the main qualities of our library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature.
-% We indeed followed actively the state-of-the-art research, since December $2016$ to June $2019$.
-More than $65$ algorithms (or variants) are implemented for the single-player case, $5$ for the experts aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary case.
-The entire codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
-%
-When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
-%
-Our library is used to run numerical simulations in Chapters~\ref{chapter:2}, \ref{chapter:3}, \ref{chapter:5} and \ref{chapter:6}, and in other contributions such as our article \cite{Besson2018DoublingTricks}.
+% % ----------------------------------------------------------------------------
+% % \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:25}}.}
+% %
+% This first part ends with Chapter~\ref{chapter:25} by presenting one of our first contributions \cite{Besson2018WCNC}.
+% We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
+% Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for a practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
+% We proposed an extension of the \ExpFour{} algorithm for this problem, that we named \Aggr, and illustrate its performance on some bandit problems.
 
-% ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:25}}.}
-%
-This first part ends with Chapter~\ref{chapter:25} by presenting one of our first contributions \cite{Besson2018WCNC}.
-We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
-Instead of always choosing her favorite algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, a third solution for a practitioner can be to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
-We proposed an extension of the \ExpFour{} algorithm for this problem, that we named \Aggr, and illustrate its performance on some bandit problems.
+% % ----------------------------------------------------------------------------
+% % \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:4}}.}
 
-% ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:4}}.}
+% Chapter~\ref{chapter:4} started the second part of the manuscript, Part~\ref{part:MABIOT}, and presents three main contributions.
+% %
+% First in Section~\ref{sec:4:firstModel}, we proposed a model of IoT networks,
+% and an evaluation of the performance of applying low-cost and low-complexity decentralized MAB learning algorithms in such, directly on the end-devices sides.
+% Our focus is the convergence of such algorithms, and the gain they can provide for the IoT devices, in terms of successful transmission rates.
+% We restrict to the case of a protocol slotted in both time and frequency, and we assume a certain clock pre-agreement and time synchronisation between devices and the base station.
+% We studied the behavior of the network when the proportion of intelligent dynamic devices changes.
+% Concretely, increasing their successful transmission rates allows to insert more devices in the same network, while maintaining a good Quality of Service.
+% Another way of viewing this result is the opposite point-of-view: if the number of devices remains constant, increasing the successful transmission rate directly improves the network QoS.
+% %  as the base station receives more accurately the data sent by the end-devices since less transmissions failed.
+% We show that \UCB{} and Thompson sampling algorithms both have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which all follow a random activation process.
+% %
+% This was both surprising and encouraging, as it shows that applying bandit algorithms tailored for a stochastic model is still useful in our broader setting.
+% The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.
+% \\
+% %
+% \indent
+% We presented in Section~\ref{sec:4:gnuradio} a proof-of-concept, demonstrated in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
+% We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
+% We discussed results obtained in practice, to highlight the interest of using learning algorithms for cognitive radio online optimization problems, and especially multi-armed bandit learning algorithms for IoT networks.
+% %
+% By using such low-cost algorithms, we confirmed empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using only the acknowledgement (\Ack) feedback as a collision binary indicator sent from the base station.
+% \\
+% %
+% \indent
+% Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of the previous model of Low-Power Wide-Area (LPWA) networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency, and assuming a perfect synchronisation.
+% % , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
+% If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
+% We presented and evaluated heuristics that learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for the first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
+% The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a MAB algorithm for this second-stage optimization problem.
+% %
+% Like for the first model, we showed that incorporating learning for the first transmissions allows to achieve optimal performance, with significant gains in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
+% Our simulations showed that the proposed heuristics greatly outperform a naive random access scheme.
+% Surprisingly, we showed that a simple \UCB{} learning approach, that simply retransmits in the same channel, turns out to perform almost as well as more complicated heuristics, that can change the channel used for retransmissions.
+% \\
+% %
+% \indent
+% Thus in Chapter~\ref{chapter:4}, we proposed two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
+% that can all use embedded low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
+% %
+% Decentralized RL for IoT lets the devices use the acknowledgement sent by their Base Station as a reward, instead of requiring some spectrum sensing capacity, like in the OSA case.
+% We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
+% The first model (without retransmission of packets),
+% is interesting for its simplicity, and because considering no retransmission can improve the battery life of the IoT devices.
+% % , at the cost of a lower Quality of Service (QoS).
+% %
+% However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
+% %
+% The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
+% %
+% \TODOL{Je dois clarifier les conclusions des deux modèles, est-ce que l'un est vraiment mieux pour améliorer la QoS mais coûte en batterie, et l'autre améliore la batterie mais coûte en QoS ?
+Pas très clair en ayant la tête dans le guidon là !...}
+% %
+% %
+% To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
+% We presented two models of wireless IoT networks, without relying on the feedback provided by spectrum sensing, and inspired by the ALOHA protocol.
+% % In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
+% %
+% It was also quite surprising to be able to conclude that this learning-based approach can be as efficient as we showed,
+% % , as it allows the IoT devices to automatically and independently increase their successful transmission rates.
+% because it is based on applying stochastic MAB algorithms in a non-stationary environment.
+% Unfortunately, it turned out to be of extreme difficulty to analyze the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
+% That is why we focus on two different simplifications of this model in the next two chapters, for which we are able to provide a clean theoretical analysis,
+% either by restricting to $M \leq K$ devices with a transmission probability of $p=1$, or by restricting to a single player accessing a network which is assumed to be \emph{piece-wise stationary}.
 
-Chapter~\ref{chapter:4} started the second part of the manuscript, Part~\ref{part:MABIOT}, and presents three main contributions.
-%
-First in Section~\ref{sec:4:firstModel}, we proposed a model of IoT networks,
-and an evaluation of the performance of applying low-cost and low-complexity decentralized MAB learning algorithms in such, directly on the end-devices sides.
-Our focus is the convergence of such algorithms, and the gain they can provide for the IoT devices, in terms of successful transmission rates.
-We restrict to the case of a protocol slotted in both time and frequency, and we assume a certain clock pre-agreement and time synchronisation between devices and the base station.
-We studied the behavior of the network when the proportion of intelligent dynamic devices changes.
-Concretely, increasing their successful transmission rates allows to insert more devices in the same network, while maintaining a good Quality of Service.
-Another way of viewing this result is the opposite point-of-view: if the number of devices remains constant, increasing the successful transmission rate directly improves the network QoS.
-%  as the base station receives more accurately the data sent by the end-devices since less transmissions failed.
-We show that \UCB{} and Thompson sampling algorithms both have near-optimal performance, even when their underlying \iid{} assumption (see Chapter~\ref{chapter:2}) is violated by the presence of many ``intelligent'' end-devices which all follow a random activation process.
-%
-This was both surprising and encouraging, as it shows that applying bandit algorithms tailored for a stochastic model is still useful in our broader setting.
-The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{cannot} communicate with each other, and \emph{do not} know the system parameters.
-\\
-%
-\indent
-We presented in Section~\ref{sec:4:gnuradio} a proof-of-concept, demonstrated in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
-We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
-We discussed results obtained in practice, to highlight the interest of using learning algorithms for cognitive radio online optimization problems, and especially multi-armed bandit learning algorithms for IoT networks.
-%
-By using such low-cost algorithms, we confirmed empirically that a dynamically re-configurable device can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using only the acknowledgement (\Ack) feedback as a collision binary indicator sent from the base station.
-\\
-%
-\indent
-Finally in Section~\ref{sec:4:retransmissions}, we presented an extension of the previous model of Low-Power Wide-Area (LPWA) networks based on an ALOHA protocol, again under the hypothesis of a protocol slotted in time and frequency, and assuming a perfect synchronisation.
-% , in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
-If the priority is the Quality of Service (QoS), for instance with renewable energy capabilities, this second model is more appropriate.
-We presented and evaluated heuristics that learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for the first transmission (first-stage), and different proposals based on \UCB{} for the retransmissions upon collisions (second-stage).
-The main novelty of this model is to also learn the most efficient way to retransmit packets upon radio collision, by also using a MAB algorithm for this second-stage optimization problem.
-%
-Like for the first model, we showed that incorporating learning for the first transmissions allows to achieve optimal performance, with significant gains in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
-Our simulations showed that the proposed heuristics greatly outperform a naive random access scheme.
-Surprisingly, we showed that a simple \UCB{} learning approach, that simply retransmits in the same channel, turns out to perform almost as well as more complicated heuristics, that can change the channel used for retransmissions.
-\\
-%
-\indent
-Thus in Chapter~\ref{chapter:4}, we proposed two models of Internet of Things (IoT) networks, composed of many independent IoT end-devices,
-that can all use embedded low-cost RL (Reinforcement Learning) algorithms to learn to improve their spectrum access.
-%
-Decentralized RL for IoT lets the devices use the acknowledgement sent by their Base Station as a reward, instead of requiring some spectrum sensing capacity, like in the OSA case.
-We consider many independent ``dynamic'' devices, communicating with a small probability at every instant.
-The first model (without retransmission of packets),
-is interesting for its simplicity, and because considering no retransmission can improve the battery life of the IoT devices.
-% , at the cost of a lower Quality of Service (QoS).
-%
-However, in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
-%
-The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.
-%
-\TODOL{Je dois clarifier les conclusions des deux modèles, est-ce que l'un est vraiment mieux pour améliorer la QoS mais coûte en batterie, et l'autre améliore la batterie mais coûte en QoS ?
-Pas très clair en ayant la tête dans le guidon là !...}
-%
-%
-To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
-We presented two models of wireless IoT networks, without relying on the feedback provided by spectrum sensing, and inspired by the ALOHA protocol.
-% In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
-%
-It was also quite surprising to be able to conclude that this learning-based approach can be as efficient as we showed,
-% , as it allows the IoT devices to automatically and independently increase their successful transmission rates.
-because it is based on applying stochastic MAB algorithms in a non-stationary environment.
-Unfortunately, it turned out to be of extreme difficulty to analyze the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
-That is why we focus on two different simplifications of this model in the next two chapters, for which we are able to provide a clean theoretical analysis,
-either by restricting to $M \leq K$ devices with a transmission probability of $p=1$, or by restricting to a single player accessing a network which is assumed to be \emph{piece-wise stationary}.
+% % ----------------------------------------------------------------------------
+% % \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:5}}.}
 
-% ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:5}}.}
+% The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Players Multi-Arm Bandits models (Chapter~\ref{chapter:5}).
+% %
+% We present three variants of this model,
+% with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
+% For the two easiest models (\ie, with sensing), our theoretical contribution \MCTopM-\klUCB{} improves the state-of-the-art regret upper-bounds.
+% In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
+% \Selfish{} corresponds to the heuristic used in the first model studied in Chapter~\ref{chapter:4}.
+% %
+% We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proven to be very active in the last two years.
+% For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
+% For example, it assumes to know the number of players $M$ before-hand, but we illustrated that a previously introduced technique to estimate $M$ can also be applied successfully to our proposal.
+% %  and give satisfactory empirical performances.
+% % Further works would be required to adapt the theoretical analysis to these various extensions.
 
-The second chapter of Part~\ref{part:MABIOT} of this manuscript studied Multi-Players Multi-Arm Bandits models (Chapter~\ref{chapter:5}).
-%
-We present three variants of this model,
-with different level of feedback being available to the decentralized players, under which we proposed efficient algorithms.
-For the two easiest models (\ie, with sensing), our theoretical contribution \MCTopM-\klUCB{} improves the state-of-the-art regret upper-bounds.
-In the absence of sensing, we also provide some motivation for the practical use of the \Selfish{} heuristic, a simple index policy based on hybrid indices that are directly taking  into account the collision information.
-\Selfish{} corresponds to the heuristic used in the first model studied in Chapter~\ref{chapter:4}.
-%
-We also reviewed various variants of this model, and for some interesting variants we discussed the related literature, which has proven to be very active in the last two years.
-For some of these variants, we explained why our approach does not work efficiently without modifications, but we detailed and illustrated how to adapt \MCTopM{} to other settings.
-For example, it assumes to know the number of players $M$ before-hand, but we illustrated that a previously introduced technique to estimate $M$ can also be applied successfully to our proposal.
-%  and give satisfactory empirical performances.
-% Further works would be required to adapt the theoretical analysis to these various extensions.
+% % ----------------------------------------------------------------------------
+% % \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:6}}.}
 
-% ----------------------------------------------------------------------------
-% \paragraph{Conclusion about \textbf{Chapter~\ref{chapter:6}}.}
-
-In the last Chapter~\ref{chapter:6}, we were interested in relaxing the stationary hypothesis made on the bandit model presented in Chapter~\ref{chapter:2}.
-%
-We first describe the break-point (or change-point) detection problem, that we solve by the Bernoulli GLR change-point detector, and by analyzing its properties at finite-time.
-This sequential break-point detection test can be applied for bounded rewards, and we bound its false alarm probability as well as its detection delay.
-For the piece-wise stationary bandit model,
-we proposed a new efficient algorithm, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
-This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike its two competitors \CUSUMUCB{} and \MUCB, that require to know a lower bound on the smallest magnitude of a change.
-We also gave numerical evidence of the efficiency of our proposal, which performs usually much better than all the passively adaptive approaches as well as better or comparably to the previous actively adaptive ones.
+% In the last Chapter~\ref{chapter:6}, we were interested in relaxing the stationary hypothesis made on the bandit model presented in Chapter~\ref{chapter:2}.
+% %
+% We first describe the break-point (or change-point) detection problem, that we solve by the Bernoulli GLR change-point detector, and by analyzing its properties at finite-time.
+% This sequential break-point detection test can be applied for bounded rewards, and we bound its false alarm probability as well as its detection delay.
+% For the piece-wise stationary bandit model,
+% we proposed a new efficient algorithm, \GLRklUCB, which combines the \klUCB{} algorithm with the Bernoulli GLR change-point detector.
+% This actively adaptive method attains state-of-the-art regret upper-bounds when tuned with a prior knowledge of the number of changes $\Upsilon_T$, but without any other prior knowledge on the problem, unlike its two competitors \CUSUMUCB{} and \MUCB, that require to know a lower bound on the smallest magnitude of a change.
+% We also gave numerical evidence of the efficiency of our proposal, which performs usually much better than all the passively adaptive approaches as well as better or comparably to the previous actively adaptive ones.
 
@@
 % \subsection*{Other directions of future works}
-\paragraph{Unifying multi-player and non-stationary bandits.}
+\subsection*{Unifying multi-player and non-stationary bandits.}
 %

### 2019-07-10 2-Chapters/Conclusion/conclusion.tex
Last proof reading

@@
 or to drop their packet after a failed transmission (\ie, without retransmission).
-Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order for the numerous IoT devices to improve on their own their spectrum access scheme, by learning which of the $K$ frequency channels are the less occupied by the environmental trafic.
+Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order for the numerous IoT devices to improve on their own their spectrum access scheme, by learning which of the $K$ frequency channels are the less occupied by the environmental traffic.
 
@@
 %
-On the other hand, we studied a model that focuses on only one device but relaxes the hypothesis of stationarity for the surrounding radio trafic, and considers that is can be stationary in consecutive intervals of time, of unknown durations.
+On the other hand, we studied a model that focuses on only one device but relaxes the hypothesis of stationarity for the surrounding radio traffic, and considers that is can be stationary in consecutive intervals of time, of unknown durations.
 This second model is similar to the piece-wise stationary (or abruptly-changing) multi-armed bandits studied since the $2000$s, and we solved it by proposing a new actively adaptive bandit policy, for which we obtained state-of-the-art results, in terms of its regret but also its false alarm probability and delay of detection.
 
-Finally, from an applicative point-of-view, we also presented a realistic implementation of a proof-of-concept of our first model of IoT network, that was used to validate empirically the proposed approach.
+Finally, from an application point-of-view, we also presented a realistic implementation of a proof-of-concept of our first model of IoT network, that was used to validate empirically the proposed approach.
 We showed that the IoT devices can effectively use a multi-armed bandit algorithm to automatically learn to favor certain frequency channels over others, in order to optimize their spectrum access, to reduce their number of failed transmissions.
@@
 Our proposed approach is to use decentralized reinforcement learning algorithms, directly embedded on the IoT end-devices, in order to improve their spectrum access and channel selection schemes.
-We confirmed from both a numerical simulations and an validation on real radio trafic and hardware that our proposal is an excellent candidate to start to solve the spectrum scarcity issue for unlicensed bands, that is cheap and easy to deploy, and requires .
+We confirmed from both a numerical simulations and an validation on real radio traffic and hardware that our proposal is an excellent candidate to start to solve the spectrum scarcity issue for unlicensed bands, that is cheap and easy to deploy, and requires .
 
@@
 One approach to obtain a unified ``master'' algorithm could be to use expert aggregation on the different state-of-the-art algorithms presented in this thesis, and as each one is efficient in a different setting, the resulting aggregated algorithm could also be efficient in each of the different settings.
-But as we illustrated in Section~\ref{sec:25:theory}, this approach does not seem theoretically promising.
-Thus a preferred approach could be to design a ``unified'' algorithm which adapts automatically to the kind of problem it is facing.
+As this approach does not seem theoretically promising, as we illustrated in Section~\ref{sec:25:theory},
+a preferred approach could be to design a ``unified'' algorithm which adapts automatically to the kind of problem it is facing.
 

### 2019-07-10 2-Chapters/Conclusion/conclusion.tex
Trop long de quelques lignes, ZUT

@@
 %
-We explained the problems that motivate and justify our works,
-and we presented the main contributions produced during the last three years.
+We explained the problems that motivate and justify our works.
+% and we presented the main contributions produced during the last three years.
 %
@@
 This PhD manuscript is organized in two parts.
-It first focuses on giving a strong theoretical background on the reinforcement learning model considered in this thesis, that are the multi-armed bandits models.
-The manuscript then focuses on our problems of solving the spectrum scarcity issue in unlicensed bands, in the context of the future Internet of Things networks,
-and extending to the specificities of such IoT networks the ideas underlying previously studied applications of machine learning to solve the spectrum scarcity issue in licensed networks.
+It first focused on giving a strong theoretical background on the reinforcement learning model considered in this thesis, that is the multi-armed bandit model.
+The manuscript then focused on our problems of solving the spectrum scarcity issue in unlicensed bands, in the context of the future Internet of Things networks,
+and on extending to the specificities of such IoT networks the underlying ideas of previously studied applications of machine learning to solve the spectrum scarcity issue in licensed networks.
 
-Our main contribution to answer the problematic is a model of Internet of Things networks, based on an ALOHA-like protocol slotted in time and frequency, and that considers one IoT base station, serving a large number of devices following the IoT constraints (low-cost, low duty cycle, long range etc).
-In such IoT network, we focused on many dynamically reconfigurable IoT end-devices, that are able to run low-cost and low-complexity decision making algorithms, embedded on each device using their limited computational and storage capacity (limited hardware)
-We proposed two models, whether the considered IoT standard asks the devices to try to retransmit a few times their up-link packet in case of a failed transmission (\ie, a collision with other devices),
-or to drop their packet after a failed transmission (\ie, without retransmission).
-Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order for the numerous IoT devices to improve on their own their spectrum access scheme, by learning which of the $K$ frequency channels are the less occupied by the environmental traffic.
+Our main contribution to answer our problematic is a model of Internet of Things networks, based on an ALOHA-like protocol slotted in time and frequency, and that considers one IoT base station serving a large number of devices following the IoT constraints (low-cost, low duty cycle, long range etc).
+In such IoT network, we focused on many dynamically reconfigurable IoT end-devices, that are able to run low-cost and low-complexity decision making algorithms, embedded on each device using their limited computational and storage capacities.
+%  (limited hardware)
+We proposed two models, whether the considered IoT standard enforces the devices to try to retransmit a few times their up-link packet in case of a failed transmission (\ie, a collision with other devices),
+or to drop their packet after a failed transmission (\ie, with or without retransmission).
+Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order for the numerous IoT devices to improve on their own their spectrum access scheme, by learning which of the $K$ frequency channels are the less occupied by the environmental traffic (in average).
 
-The approach we thus advocate is to use low-complexity MAB algorithms such as \UCB, or variants tuned to be robust in slowly-evolving non-stationary scenario (\eg, Discounted-\UCB{} or actively adaptive approach such as \GLR-\klUCB).
-For a given company manufacturing IoT end-devices, it is very simple and cheap in terms of both hardware and software engineering workloads to implement such MAB algorithm on the low-cost embedded processors, and equip all the produced IoT devices with this capacity of using online machine-learning for their spectrum access.
-The embedded decision-making algorithm will automatically improve the channel selection of each device, in a fully decentralized way, and will increase the total number of successful up-link transmissions, thus it will increase the Quality of Service of the considered IoT application, as well as allowing more devices to be served by the same IoT gateway.
+The approach we thus advocate is to use low-complexity MAB algorithms such as \klUCB{} or Thompson sampling, or variants tuned to be robust in slowly-evolving non-stationary scenario (\eg, Sliding-Window \UCB, Discounted-Thompson, or actively adaptive algorithm such as the one of Chapter~\ref{chapter:6}, \GLR-\klUCB).
 %
-The advocated approach also has the advantage of not requiring the change anything from the IoT standard side, as the only modification happens on the device side, by letting it actively decide the channels it uses for up-link transmissions, instead of relying on a naive uniform channel access.
+For a company that manufactures IoT end-devices, it is very simple and cheap in terms of both hardware and software costs to implement such MAB algorithms on the low-cost embedded processors, and equip all the produced IoT devices with this capacity of using online machine-learning for their spectrum access.
+The embedded decision-making algorithm automatically improves the channel selection of each device, in a fully decentralized way, and increases the total number of successful up-link transmissions, and thus this increases the Quality of Service of the considered IoT application and allow more devices to be served by the same IoT gateway.
 %
-We also note that this approach can easily be used to let the IoT devices optimize by themselves and on the fly other parameters of their wireless communications, as the multi-armed bandit framework is not restricted to be applied for the frequency channel selections but can be applied to any exploration-exploitation problem with a finite set of possible options.
-The solution we propose can thus be easily extended to use the same kinds of MAB algorithms to select dynamically parameters such as the coding rate or spreading factor of a LoRa network (\eg, \cite{KerkoucheAlami18}), the emission power in a NOMA standard, or any other discrete-valued parameters that have an impact on the successful transmission rate and can be optimized on the device's side by using reinforcement learning.
+Another consequence is a reduced energy consumption and an increased battery life for the IoT end-devices.
+% , and thus the solution studied in this thesis can also be used to reduce the global energy consumption of the IoT networks.
+% solve both the spectrum scarcity issue and the energy
 
-Moreover, from a theoretical aspect, even though it was found hard to propose a rigorous analysis of the model of IoT networks, because of the large number of end-devices having different random and unpredictable activation patterns, our contributions also include an analysis of two restricted models.
+The advocated approach also has the advantage of not requiring to change anything on the IoT standard, as the only modification happens on the device side, by letting it actively decide the channels it uses for up-link transmissions, instead of relying on a naive uniform channel access.
 %
-On the one hand, we considered a model where the devices have data to transmit at each time step, relaxing the hypothesis of a random activation process (with a small probability, as IoT devices have low duty cycles), and thus by restricting to the case of at most $M \leq K$ devices in a wireless standard with $K$ orthogonal frequency channels.
-This first model is similar to the multi-player multi-armed bandits studied in the last $10$ years, and for stationary and stochastic environments we were able to propose a state-of-the-art algorithm, \MCTopM.
+We also note that this approach can easily be used to allow the IoT devices to optimize by themselves other parameters of their wireless communications on the fly, as the multi-armed bandit framework is not restricted to be applied to the selection of frequency channel but can be applied to any exploration-exploitation problem with a finite set of possible options.
+The solution we propose can thus be easily extended to also select dynamically other parameters such as the coding rate or spreading factor of a LoRa network (\eg, \cite{KerkoucheAlami18}), the emission power in a NOMA standard, or any other discrete-valued parameters that have an impact on the successful transmissions rates and can be optimized on the device's side by using reinforcement learning.
+
+Moreover, from a theoretical aspect, even though it was found hard to propose a rigorous analysis of our model of IoT networks, because of the large number of devices having different random and unpredictable activation patterns, our contributions also include an analysis of two restricted models.
+%
+On the one hand, we considered a model where the devices have data to transmit at each time step, relaxing the hypothesis of a random activation process (with a small probability, as IoT devices have low duty cycles), and thus by restricting to the case of at most $M \leq K$ devices using a wireless standard with $K$ orthogonal frequency channels.
+This first model is similar to the multi-player bandit model studied in the last $10$ years, and for stationary and stochastic environments we were able to propose a state-of-the-art algorithm, \MCTopM.
 Our proposed algorithm performs very well in practice and obtains interesting guarantees on its regret, as well as on the number of radio collisions and on radio reconfigurations that the devices will suffer if they all implement our solution.
 %
-On the other hand, we studied a model that focuses on only one device but relaxes the hypothesis of stationarity for the surrounding radio traffic, and considers that is can be stationary in consecutive intervals of time, of unknown durations.
-This second model is similar to the piece-wise stationary (or abruptly-changing) multi-armed bandits studied since the $2000$s, and we solved it by proposing a new actively adaptive bandit policy, for which we obtained state-of-the-art results, in terms of its regret but also its false alarm probability and delay of detection.
+On the other hand, we studied a model that focuses on only one device but relaxes the hypothesis of stationarity on the surrounding radio traffic, and considers that is can be stationary in consecutive intervals of time, of unknown durations.
+This second model is similar to the piece-wise stationary (or abruptly-changing) bandit model studied since the $2000$s, and we solved it by proposing a new actively adaptive bandit policy, for which we obtained state-of-the-art results, in terms of its regret and its false alarm probability and delay of detection.
 
@@
 Our proposed approach is to use decentralized reinforcement learning algorithms, directly embedded on the IoT end-devices, in order to improve their spectrum access and channel selection schemes.
-We confirmed from both a numerical simulations and an validation on real radio traffic and hardware that our proposal is an excellent candidate to start to solve the spectrum scarcity issue for unlicensed bands, that is cheap and easy to deploy, and requires .
+We confirmed from both numerical simulations and a validation on real radio traffic and real hardware that our proposal is an excellent candidate to start to solve the spectrum scarcity issue for unlicensed bands, that is cheap and easy to deploy.
 
@@
 
-% As discussed at the end of each chapter,
-% All the models, simulations,
-What we presented above in each chapter suggest the following directions of future studies.
+% % As discussed at the end of each chapter,
+% % All the models, simulations,
+% What we presented above in each chapter suggest the following directions of future studies.
 
@@
 %
-A second interesting task could be to extend the library for problems with non-finite arms, such as linear or contextual bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
-or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+A second interesting task could be
+% to extend the library for problems with non-finite arms, such as linear or contextual bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
+to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
 %
@@
 Also, other emission patterns could be considered, instead of a Bernoulli process for each user.
-In this whole Chapter~\ref{chapter:4}, we prefer to consider that all the devices have the same activation probability, to keep the notations and the model as simple as possible.
+In this entire Chapter~\ref{chapter:4}, we prefer to consider that all the devices have the same activation probability, to keep the notations and the model as simple as possible.
 %
 Moreover, for the sake of simplicity we supposed that all devices use the same standard.
-Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.
+Future works could consider more realistic interference scenarios and IoT networks, with non-slotted time, more than one base station etc.
 % \\
@@
 We shall moreover investigate whether actively adaptive approaches can attain a $\cO(\sqrt{\Upsilon_T T})$ regret upper-bound without the knowledge of $\Upsilon_T$.
-We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of more efficient algorithms, as suggested by \cite{Maillard2018GLR}.
+We also believe that combining change-point \emph{localization} with an efficient change-point detection algorithm, such as \GLRklUCB, could lead to an interesting class of more efficient algorithms.
+% as suggested by \cite{Maillard2018GLR}.
 % \\
@@
 %
-Another interesting direction of future work is to study non-stationary distributed multi-players bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.
+Finally, another interesting direction of future work is to study non-stationary distributed multi-players bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.
 A natural extension of the non-stationary model is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
@@
 and the Bernoulli GLR break-point detector for non-stationarity (\ie, dealing with abrupt changes).
-The three parts should be connected, the same way we built \MCTopM-\klUCB{} and \GLRklUCB, by connecting two of the three components together.
-We propose the idea of incorporating the detected change-points by B-\GLR{} test in the orthogonalization procedure \MCTopM: after any change-point is detected, the player is not sitting anymore.
+The three parts should be inter-connected, the same way we built \MCTopM-\klUCB{} and \GLRklUCB, by connecting two of the three components together.
+We propose a simple idea to incorporate the detected change-points by B-\GLR{} test in the orthogonalization procedure \MCTopM: whenever a change-point is detected, the player is no longer ``fixed'' on its chosen arm.
 %
@@
 
-Another question of highest interest is to look for a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and automatically adapt to the setting at hand.
+Another question of highest interest is to look for a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and maybe others, and automatically adapt to the setting at hand.
 For instance, even if \GLRklUCB{} is very efficient for piece-wise stationary problems, its forced exploration makes it sub-optimal for stationary problems.
@@
 % if we consider the proposed extension of \MCTopM-\klUCB{} that estimates the number $M$ of players, i
-One approach to obtain a unified ``master'' algorithm could be to use expert aggregation on the different state-of-the-art algorithms presented in this thesis, and as each one is efficient in a different setting, the resulting aggregated algorithm could also be efficient in each of the different settings.
-As this approach does not seem theoretically promising, as we illustrated in Section~\ref{sec:25:theory},
+One approach to obtain a unified ``master'' algorithm could be to use expert aggregation on the different state-of-the-art algorithms presented in this thesis, and as each one was proven to be efficient in one setting, the resulting aggregated algorithm would also be efficient in each of the different settings.
+As this approach does not seem theoretically promising, as illustrated in Section~\ref{sec:25:theory},
 a preferred approach could be to design a ``unified'' algorithm which adapts automatically to the kind of problem it is facing.
@@
 
+\begin{center}
+    Merci d'avoir lu cette thèse.
+    Thanks for reading this document.
+\end{center}
+
+\hr{}
+
+\vfill{}
+
+% https://fr.wikipedia.org/wiki/Liste_des_%C3%A9pisodes_de_Kaamelott
+
 \begin{small}

### 2019-07-10 2-Chapters/Conclusion/conclusion.tex
DONE

@@
 % \section{General Conclusion}
-\subsection*{General Conclusion}
+% \subsection*{General Conclusion}
 
@@
 
-We started this manuscript by detailing the historical, scientific and technical contexts of our research during this PhD, which studied how decentralized low-cost machine learning technique can be used to automatically improve a certain aspect of wireless telecommunications.
+We started this manuscript by detailing the historical, scientific and technical contexts of our research during this PhD.
+% which studied how decentralized low-cost machine learning technique can be used to automatically improve a certain aspect of wireless telecommunications.
 %
@@
 %
-This PhD manuscript is organized in two parts.
-It first focused on giving a strong theoretical background on the reinforcement learning model considered in this thesis, that is the multi-armed bandit model.
-The manuscript then focused on our problems of solving the spectrum scarcity issue in unlicensed bands, in the context of the future Internet of Things networks,
+% This PhD manuscript
+% is organized in two parts.
+% It
+We first gave a strong theoretical background on the reinforcement learning model considered in this thesis, that is the multi-armed bandit model.
+We then focused on our problems of solving the spectrum scarcity issue in unlicensed bands, in the context of the future Internet of Things networks,
 and on extending to the specificities of such IoT networks the underlying ideas of previously studied applications of machine learning to solve the spectrum scarcity issue in licensed networks.
@@
 
-\hr{}
+% \hr{}
 
-\vfill{}
+% \vfill{}
 

### 2019-07-10 2-Chapters/Conclusion/conclusion.tex
BOUM

@@
 
-In order to validate our results in a realistic experimental setting and not only with simulations, future works should include a hardware implementation of the analyzed models, in order to complete our demonstration presented in Section~\ref{sec:4:gnuradio} \cite{Besson2018ICT,Besson2019WCNC}.
-%
-A hardware demonstrator could be also benefit to study other settings by relaxing some hypotheses, for instance by studying a similar model in non-slotted time.
+% In order to validate our results in a realistic experimental setting and not only with simulations, future works should include a hardware implementation of the analyzed models, in order to complete our demonstration presented in Section~\ref{sec:4:gnuradio} \cite{Besson2018ICT,Besson2019WCNC}.
+% %
+% A hardware demonstrator could be also benefit to study other settings by relaxing some hypotheses, for instance by studying a similar model in non-slotted time.
 
@@
 
-% https://fr.wikipedia.org/wiki/Liste_des_%C3%A9pisodes_de_Kaamelott
-
 \begin{center}
-    Merci d'avoir lu cette thèse.
-    Thanks for reading this document.
+    Je vous remercie d'avoir lu cette thèse. Merci de me signaler toute erreur par courriel :
+    \\
+    Thanks for reading this document. Please notify me about any mistake by e-mail:\\
+    \href{https://github.com/Naereen/phd-thesis/issues/new}{\texttt{Lilian.Besson}\texttt{@}\texttt{Crans.org}}
 \end{center}
@@
 
-% \vfill{}
+\vfill{}
 

### 2019-07-11 1-Preamble/5-Abstract/abstract.tex
Okay, fixed bug of transparent last page

@@
 Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de type radio intelligente, dans des bandes non licenciées et sans supervision centrale.
-Nous considérons notamment des réseaux actuels ou futures de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
+Nous considérons notamment des réseaux actuels ou futurs de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
 %  (Chapitre~\ref{chapter:1}).

### 2019-07-11 1-Preamble/5-Abstract/resume_fr.tex
Okay, fixed bug of transparent last page

@@
 % % , Chapitre~\ref{chapter:2}
-% notamment si l'on veut prouver la convergence de nombreux objets jouant à un jeu collaboratif sans aucune coordination, lorsque les objets suivent tous un modèle d'activation aléatoire.
+% notamment si l'on veut prouver la convergence de nombreux objets jouant à un jeu collaboratif sans aucune coordination, lorsque les objets suivent tous un profil d'activation aléatoire.
 % % En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
@@
 %
-C'est pourquoi en plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, couvrant tous les cas possibles,
+C'est pourquoi, en plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, couvrant tous les cas possibles,
 le moment est venu de le combiner avec un apprentissage machine bon marché et prometteur pour (essayer d') atteindre le niveau de gain de performance nécessaire pour que les promesses de l'IdO deviennent réalité.
 
-C'est pourquoi nous nous intéressons dans cette thèse à
+C'est ainsi que nous avons décidé de nous intéresser dans cette thèse à
 l'amélioration de la durée de vie des batteries des objets de l'IdO et la réduction du coût énergétique des réseaux de l'IdO.
@@
 et elles consistent en une spécification à la fois pour la couche \emph{PHY}sique
-et la couche \emph{M}edium \emph{A}access \emph{C}ontrol (\emph{MAC}).
+et la couche \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}).
 Pour citer quelques exemples de normes pour des réseaux de type IdO, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée (jusqu'à \SI{2}{\meter}), tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
@@
 Nous illustrons ci-dessous l'idée d'un cycle d'apprentissage, alternant entre action et réaction,
-en Figure~\ref{fig:1:ReinforcementLearningCycle}.
+en Figure~\ref{fig:1:ReinforcementLearningCycle_enFr}.
 Un-e joueur-se (ou un-e apprenant-e) interagit avec son environnement en prenant une action $A(t)$ (\eg, un choix dans un ensemble fini, $A(t)\in\{1,\dots,K\}$, ou un vecteur $A(t)\in\R^d$), et ensuite en observant une récompense $r(t)$, qui est une certaine mesure du succès de cette action, fournie par l'environnement (\eg, $r(t)\in\{0,1\}$ pour échec/succès binaire, ou $r(t)\in\R$).
@@
 De nombreux problèmes du monde réel peuvent être présentés comme des problèmes d'apprentissage par renforcement, comme l'illustrent l'article \cite{bouneffouf2019survey} et la Section~\ref{sec:2:applicationsofStochasticMAB}, par exemple apprendre à marcher, à conduire, à jouer à un jeu vidéo, découvrir quel traitement est efficace pour guérir une certaine maladie, etc.
+
+\tikzstyle{block} = [align=center, draw, fill=gray!25, rectangle, minimum height=3em, minimum width=6em]
+\begin{figure}[h!]
+    \centering
+    \resizebox{0.30\textwidth}{!}{
+        \begin{tikzpicture}[auto,node distance=5cm,>=latex,scale=1.5]
+            %
+            % We start by placing the blocks
+            \node [block] (player) at (0,0) {Joueuse};
+            % We draw an edge between the player and system block to
+            \node [block] (environment) at (2,0) {Environnement};
+            % Once the nodes are placed, connecting them is easy.
+            \draw [->] (player) to[bend left=90] node[pos=0.5] {Action} (environment);
+            \draw [->] (environment) to[bend left=90] node[pos=0.5] {Récompense} (player);
+            %
+        \end{tikzpicture}
+    }
+\caption{Cycle de l'apprentissage par renforcement : une joueuse interagit avec son environnement par des actions, et observe une récompense, de façon itérative.}
+\label{fig:1:ReinforcementLearningCycle_enFr}
+\end{figure}
 
@@
 %
-En utilisant cette rétroaction, qui consiste en un \emph{acquittement} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence ' \Ack{} après une transmission échouée, il est possible qu'un objet de l'IdO puisse également optimiser ses communications, grâce à un algorithme d'apprentissage par renforcement bien conçu.
+En utilisant cette rétroaction, qui consiste en un \emph{acquittement} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence ' \Ack{} après une transmission échouée, il est possible qu'un objet de l'IdO puisse également optimiser ses communications, grâce à un algorithme d'apprentissage par renforcement correctement conçu.
 
@@
 
-Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé insoluble à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux objets, chacun suivant des modèles d'activation aléatoires.
+Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé trop complexe à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux objets, chacun suivant des profils d'activation aléatoires.
 La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes BMB, conçus pour résoudre des problèmes stationnaires, appliqués à un problème non stationnaire,
@@
 
-D'une part, si le modèle d'activation de tous les périphériques peut être corrigé, par exemple en se basant sur une affectation centralisée des périphériques à différents créneaux horaires, alors le modèle de bandit multi-joueurs du Chapitre~\ref{chapter:5} peut être utilisé pour permettre à chaque groupe de $M \leq K$ périphériques d'apprendre une affectation orthogonale optimale dans les $K$ canaux (\eg, les groupes peuvent être le jeu des périphériques que tous émettent au même instant).
+D'une part, si le profil d'activation de tous les périphériques peut être corrigé, par exemple en se basant sur une affectation centralisée des périphériques à différents créneaux horaires, alors le modèle de bandit multi-joueurs du Chapitre~\ref{chapter:5} peut être utilisé pour permettre à chaque groupe de $M \leq K$ périphériques d'apprendre une affectation orthogonale optimale dans les $K$ canaux (\eg, les groupes peuvent être le jeu des périphériques que tous émettent au même instant).
 %
@@
 %
-Afin de bien comprendre également comment les algorithmes BMB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur les BMB adverses ou non stationnaires, pour les deux cas d'environnements variant lentement ou changeant brusquement.
+Afin de bien comprendre également comment les algorithmes BMB se comportent dans un environnement non stationnaire, nous avons étudié la littérature sur les BMB adverses ou non stationnaires, principalement pour les deux cas d'environnements variant lentement ou changeant brusquement.
 Nous commençons notre dernier Chapitre~\ref{chapter:6} en passant en revue les travaux existants,
@@
 Suite à deux travaux précédents récents \cite{LiuLeeShroff17,CaoZhenKvetonXie18}, nous combinons une stratégie d'indices efficace (\klUCB) et un test efficace de détection des points de changement, dans l'hypothèse de récompenses bornées.
-Nous nous appuyons sur des résultats très récents du Test du Rapport de Vraisemblance Généralisé (\emph{Generalized Likelihood Ratio Test}, GLRT) pour les variables gaussiennes et sous-gaussiennes \cite{Maillard2018GLR}, mais nous nous concentrons plutôt sur les récompenses bornées et les distributions de Bernoulli.
+Nous nous appuyons sur des résultats très récents à propos du Test du Rapport de Vraisemblance Généralisé (\emph{Generalized Likelihood Ratio Test}, GLRT) pour les variables gaussiennes et sous-gaussiennes \cite{Maillard2018GLR}, mais nous nous concentrons plutôt sur les récompenses bornées et les distributions de Bernoulli.
 Les récompenses bornées sont en effet généralement plus appropriées pour les applications de RI, comme nous l'avons discuté ci-dessus.

### 2019-07-11 2-Chapters/2-Chapter/chapter2.tex
Okay, fixed bug of transparent last page

@@
         \emph{-- C'est cool ça, des bandits haha !} Le Barbare,
-        \emph{Le Donjon de Naheulbeuk}\\
+        \href{http://www.penofchaos.com/warham/donjon.htm}{\emph{Le Donjon de Naheulbeuk}}\\
         Saison III, \href{https://soundcloud.com/penofchaos/naheulbeuk-saison-3-episode-5?in=penofchaos/sets/naheulbeuk-saison-3#t=2:05}{Épisode 37 : ``La stratélique''}.

### 2019-07-11 2-Chapters/Conclusion/conclusion.tex
Okay, fixed bug of transparent last page

@@
 On the one hand, we considered a model where the devices have data to transmit at each time step, relaxing the hypothesis of a random activation process (with a small probability, as IoT devices have low duty cycles), and thus by restricting to the case of at most $M \leq K$ devices using a wireless standard with $K$ orthogonal frequency channels.
-This first model is similar to the multi-player bandit model studied in the last $10$ years, and for stationary and stochastic environments we were able to propose a state-of-the-art algorithm, \MCTopM.
-Our proposed algorithm performs very well in practice and obtains interesting guarantees on its regret, as well as on the number of radio collisions and on radio reconfigurations that the devices will suffer if they all implement our solution.
+This first model is similar to the multi-player bandit model studied in the last $10$ years, and for stationary and stochastic environments we were able to propose an efficient algorithm, \MCTopM.
+Our proposed algorithm performs very well in practice and we obtained interesting guarantees on its regret, as well as on the number of radio collisions and on radio reconfigurations that the devices will suffer if they all implement our solution.
 %
@@
 % % All the models, simulations,
-% What we presented above in each chapter suggest the following directions of future studies.
+The works presented in each chapter suggest possible directions of future studies.
 
@@
 % The hypotheses we made were motivated by the target application, but these other models could also be used for cognitive radio applications.
-
-% ----------------------------------------------------------------------------
-% \paragraph{Perspectives about \textbf{Chapter~\ref{chapter:3}}.}
-
-A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
-For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
-%
-A second interesting task could be
-% to extend the library for problems with non-finite arms, such as linear or contextual bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
-to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
-%
-A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
-It is already possible to reproduce some of the experiments presented in this thesis,
-by following the instructions given in the documentation (see Section~\ref{sub:3:listResearchWorksUsingSMPyBandits}),
-by using the Jupyter notebooks \cite{jupyter} made available%
-\footnote{~These notebooks are hosted on both the GitHub repository of SMPyBandits and on my website on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}.
-They can be used locally if you install the library, but can also be used on cloud platforms, like Binder \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} or Google Colab \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}}.
 
@@
 % ----------------------------------------------------------------------------
+% \paragraph{Perspectives about \textbf{Chapter~\ref{chapter:3}}.}
+
+A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
+%
+A second interesting task could be
+% to extend the library for problems with non-finite arms, such as linear or contextual bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
+to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
+%
+A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
+It is already possible to reproduce some of the experiments presented in this thesis,
+by following the instructions given in the documentation (see Section~\ref{sub:3:listResearchWorksUsingSMPyBandits}),
+by using the Jupyter notebooks \cite{jupyter} made available%
+\footnote{~These notebooks are hosted on both the GitHub repository of SMPyBandits and on my website on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}.
+They can be used locally if you install the library, but can also be used on cloud platforms, like Binder \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} or Google Colab \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}}.
+
+% ----------------------------------------------------------------------------
 % \subsection*{Other directions of future works}
@@
 %
-Finally, another interesting direction of future work is to study non-stationary distributed multi-players bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.
+A natural next step after this thesis is to study non-stationary distributed multi-players bandits, to unify the models from Chapter~\ref{chapter:5} and~\ref{chapter:6}.
 A natural extension of the non-stationary model is to consider non-communicating players cooperating in a decentralized way to play the same bandit game, as it was proposed recently in \cite{WeiSrivastava18Distributed}.
@@
 %
-Even if preliminary numerical simulations showed promising results,
-analyzing this strategy that combine \MCTopM, \GLR{} and \klUCB{} is left as a challenging future work.
+Even if we already obtained promising results on preliminary numerical simulations,
+analyzing such a strategy that combines \MCTopM, \GLR{} and \klUCB{} is left as a challenging future work.
 
@@
 
-Another question of highest interest is to look for a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and maybe others, and automatically adapt to the setting at hand.
+% Another question of highest interest is to look for 
+Finally, I would be very interested in looking at a unique ``unified'' algorithm that can be used in all the different settings studied in this thesis, and maybe others, and automatically adapt to the setting at hand.
 For instance, even if \GLRklUCB{} is very efficient for piece-wise stationary problems, its forced exploration makes it sub-optimal for stationary problems.
@@
 
+\vfill{}
+
 \hr{}
 
-\vfill{}
+\begin{small}
+\begin{center}
+    Je vous remercie d'avoir lu cette thèse. Merci de me signaler toute erreur par courriel ou sur GitHub :
+    \\
+    Thanks for reading this document. Please notify me about any mistake by e-mail or on GitHub:\\
+    $\hookrightarrow$ \href{https://github.com/Naereen/phd-thesis/issues/new}{\texttt{GitHub.com/Naereen/phd-thesis/issues/new}}\\
+    $\hookrightarrow$ \href{mailto:Lilian.Besson_at_Crans.org}{\texttt{Lilian.Besson}\texttt{@}$\{$\texttt{Crans.org},\texttt{Inria.fr},\texttt{live.fr}$\}$}
+\end{center}
+\end{small}
 
-\begin{center}
-    Je vous remercie d'avoir lu cette thèse. Merci de me signaler toute erreur par courriel :
-    \\
-    Thanks for reading this document. Please notify me about any mistake by e-mail:\\
-    \href{https://github.com/Naereen/phd-thesis/issues/new}{\texttt{Lilian.Besson}\texttt{@}\texttt{Crans.org}}
-\end{center}
+\hr{}
 
-% \hr{}
-
-\vfill{}
+% \vfill{}
 
@@
 \begin{quote}
-    \emph{-- Les rêves, ça se compare pas\ldots}\\
+    \emph{-- Les rêves, ça se compare pas\ldots}
     Alexandre Astier, \emph{Kaamelott}, Livre VI, Épisode 9, ``Dies Irae''.

### 2019-07-11 1-Preamble/5-Abstract/resume_fr.tex
Done for now

@@
 
-Ce troisième type de systèmes sans fil sont caractérisés par leur nature décentralisée,
+Ce troisième type de systèmes sans fil est caractérisé par leur nature décentralisée,
 où les communications sont initialisées et régulées par les objets, et non par un système de contrôle centralisé.
@@
 Les bandes des réseaux cellulaires (2G/3G/4G/5G) sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radio amateures) sont moins utilisées.
-Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend fortement à la fois du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
+Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend à la fois du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
 En outre, l'attribution fixe du spectre empêche
@@
 %
-Même si le paradigme de la RI a été initié par des recherches de l'armée des États-Unis d'Amérique du Nord dans les années $1980$, l'industrie civile a commencé à s'intéresser à la RI au cours des $20$ dernières années, et la RI a également suscité beaucoup d'intérêt de la part du milieu universitaire.
+Même si le paradigme de la RI a été initié par des recherches de l'armée des États-Unis d'Amérique du Nord dans les années $1980$, l'industrie civile a commencé à s'intéresser à la RI au cours des vingt dernières années, et la RI a également suscité beaucoup d'intérêt de la part du milieu universitaire.
 %
-Comme la RI n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RI, dont le premier vient de paraître il y a $20$ ans.
+Comme la RI n'est pas une technologie standard, elle n'a pas de définition unique, commençons donc par citer la définition de deux chercheurs dont les travaux ont été à l'origine du développement de la RI, dont le premier vient de paraître il y a vingt ans.
 %
@@
 %
-D'après ce que les auteurs ont identifiés,
+D'après ce que les auteurs ont identifié,
 nous avons écrit la bibliothèque libre de simulation la plus exhaustive pour les problèmes BMB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
@@
     Nous avons réalisé une démonstration réaliste du modèle proposé ci-dessus \cite{Besson2018ICT,Besson2019WCNC}, présentée à ICT 2018, et nous la détaillons en Section~\ref{sec:4:gnuradio}.
-Nous avons aussi réalisé une vidéo de présentation de $6$ minutes, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
+Nous avons aussi réalisé une vidéo de présentation de six minutes, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
 

### 2019-07-11 2-Chapters/2-Chapter/chapter2.tex
Done for now

--- 2019-07-10 2-Chapters/2-Chapter/chapter2.tex
@@
     \begin{flushright}
-        \emph{-- C'est cool ça, des bandits haha !} Le Barbare,
-        \emph{Le Donjon de Naheulbeuk}\\
+        \emph{-- C'est cool ça, des \emph{bandits} haha !}\\
+        Le Barbare,
+        John Lang, \href{http://www.penofchaos.com/warham/donjon.htm}{\emph{Le Donjon de Naheulbeuk}},\\
         Saison III, \href{https://soundcloud.com/penofchaos/naheulbeuk-saison-3-episode-5?in=penofchaos/sets/naheulbeuk-saison-3#t=2:05}{Épisode 37 : ``La stratélique''}.
-        écrit par John Lang (Pen of Chaos).
     \end{flushright}
@@
         % \includegraphics[width=0.84\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
-        \includegraphics[width=0.92\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_3198350886837790844.pdf}
+        \includegraphics[width=0.93\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_3198350886837790844.pdf}
         \label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000_7_algos}
@@
     \subfloat{
-        \includegraphics[width=0.92\linewidth]{SP__K5_T10000_N1000__3_algos/main____env1-1_776571591054077529.pdf}
+        \includegraphics[width=0.93\linewidth]{SP__K5_T10000_N1000__3_algos/main____env1-1_776571591054077529.pdf}
         \label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000_3_algos}

### 2019-07-11 2-Chapters/2-Chapter/chapter25.tex
Done for now

@@
         \emph{-- Y'a toujours au moins deux solutions à un problème.}\\
-        Bruno Fontaine, \emph{Kaamelott}, Livre III, Épisode 67 ``La Potion de Fécondité II''.
+        Élias de Kelliwic'h, interprété par Bruno Fontaine,\\
+        \emph{Kaamelott}, Livre III, Épisode 67 ``La Potion de Fécondité II''.
     \end{flushright}

### 2019-07-11 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done for now

@@
     \centering
-    \includegraphics[scale=0.80]{perf_learning.eps}
+    \includegraphics[scale=0.90]{perf_learning.eps}
     \caption[Learning with \UCB{} and Thomson Sampling, with many dynamic devices]{Learning with \textcolor{red}{\UCB} and \textcolor{gold}{Thomson Sampling}, with many dynamic devices.
@@
     \centering
-    \subfloat[][10\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_10.eps}
+    \subfloat[][10\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_10.eps}
     \label{fig:41:ch2_10}}
     %
-    \subfloat[][30\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_30.eps}
+    \subfloat[][30\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_30.eps}
     \label{fig:41:ch2_30}}
@@
     \vspace*{-10pt}
-    \subfloat[][50\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_50.eps}
+    \subfloat[][50\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_50.eps}
     \label{fig:41:ch2_50}}
     %
-    \subfloat[][100\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_100.eps}
+    \subfloat[][100\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_100.eps}
     \label{fig:41:ch2_100}}

### 2019-07-11 2-Chapters/Conclusion/conclusion.tex
Done for now

@@
     $\hookrightarrow$ \href{https://github.com/Naereen/phd-thesis/issues/new}{\texttt{GitHub.com/Naereen/phd-thesis/issues/new}}\\
-    $\hookrightarrow$ \href{mailto:Lilian.Besson_at_Crans.org}{\texttt{Lilian.Besson}\texttt{@}$\{$\texttt{Crans.org},\texttt{Inria.fr},\texttt{live.fr}$\}$}
+    $\hookrightarrow$ \href{mailto:Lilian.Besson_at_Crans.org}{\texttt{Lilian.Besson}\texttt{@}$\{$\texttt{Crans.org}, \texttt{Inria.fr}, \texttt{live.fr}$\}$}
 \end{center}

### 2019-07-11 1-Preamble/5-Abstract/abstract.tex
Done

@@
 % ,  Chapter~\ref{chapter:2}
-because proving the convergence of many IoT devices playing a collaborative game, without communication nor coordination is hard, when they all follow random active pattern.
+because proving the convergence of many IoT devices playing a collaborative game, without communication nor coordination is hard, when they all follow random activation patterns.
 % Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.

### 2019-07-11 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Trying to kill this last empty page...

@@
     In a realistic scenario, we can assume that $D_k\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall k\in[K]$.
-For such values for $D_k$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{k=1}^{K} D_k (1 - p)^{S_k + D_k -1}$ is concave as the sum of concave functions
+For such values for $D_k$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{k=1}^{K} D_k (1 - p)^{S_k + D_k -1}$ is concave as the sum of concave functions%
     \footnote{~It is worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
@@
 
-\textbf{Using a bandit algorithm for IoT devices.}
+\paragraph{Using a bandit algorithm for IoT devices.}
 %
@@
 
-\textbf{Considered algorithms.}
+\paragraph{Considered algorithms.}
 %
@@
 
-\textbf{\emph{Adversarial} bandit algorithms?}
+\paragraph{\emph{Adversarial} bandit algorithms?}
 %
@@
     \label{fig:41:from10to100}
-    \vspace*{-10pt}
+    % \vspace*{-10pt}
 \end{figure}
@@
 
-\newpage % XXX
+% \newpage % XXX
 

### 2019-07-11 1-Preamble/5-Abstract/resume_fr.tex
Almost done proofreading résumé fr

@@
 Ce manuscrit conclut ma thèse de doctorat, qui a débuté en octobre $2016$ et s'est terminée en novembre $2019$.
-Mes recherches se sont déroulées au laboratoire IETR à Rennes (France), dans l'équipe SCEE hébergée sur le campus de Rennes de l'école d'ingénieurs CentraleSupélec.
+Mes recherches se sont déroulées au laboratoire IETR à Rennes en France, dans l'équipe SCEE hébergée sur le campus de Rennes de l'école d'ingénieurs CentraleSupélec.
 J'étais supervisé par le professeur Christophe Moy, à Rennes,
-et j'étais également co-encadré par le docteur Emilie Kaufmann, que j'ai visité de nombreuses fois à Inria Lille Nord Europe à Lille (France).
+et j'étais également co-encadré par le docteur Émilie Kaufmann, à qui j'ai souvent rendu visite à Inria Lille Nord Europe.
 
@@
 %
-Ces technologies de l'IdO devraient pouvoir s'adapter automatiquement à différents environnements et différents contextes d'application, et être aussi efficaces que possible.
+Ces technologies des nouveaux réseaux sans fil de l'IdO devraient pouvoir s'adapter automatiquement à différents environnements et différents contextes d'application, et être aussi efficaces que possible.
 %
 C'est pourquoi, en plus de l'effort habituel de recherche et développement pour concevoir de nombreux systèmes d'accès radio efficaces, couvrant tous les cas possibles,
-le moment est venu de le combiner avec un apprentissage machine bon marché et prometteur pour (essayer d') atteindre le niveau de gain de performance nécessaire pour que les promesses de l'IdO deviennent réalité.
+le moment est venu de le combiner avec un apprentissage machine peu coûteux, dans le but d'atteindre le niveau de gain de performance nécessaire pour que les promesses de l'IdO deviennent réalité.
 
@@
 Nous proposons d'atteindre ces deux objectifs conjointement, en intégrant une prise de décision décentralisée à faible coût, directement dans les futurs objets de l'IdO.
-%
-Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (des algorithmes de types bandits multi-bras), afin de permettre aux objets de l'IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
+
+Ma thèse de doctorat porte donc sur les applications possibles de l'intégration d'un certain type d'algorithmes d'apprentissage machine (des algorithmes de type bandits multi-bras), afin de permettre aux objets de l'IdO d'optimiser leurs communications sans fil et d'apprendre à s'organiser automatiquement et sans contrôle central ni coordination.
 
@@
 %
-Historiquement, trois grandes familles de systèmes de communication sans fil ont été déployées dans les grands réseaux commerciaux : d'abord, la radiodiffusion centralisée (\eg, radio ou télédiffusion), puis les systèmes bidirectionnels centralisés (\eg, 4G ou WiFi), et aujourd'hui la collecte décentralisée de données pour les réseaux de l'Internet des objets (\eg, réseaux capteurs).
+Historiquement, trois grandes familles de systèmes de communication sans fil ont été déployées dans les grands réseaux commerciaux : d'abord, la radiodiffusion centralisée (\eg, radio ou télédiffusion), puis les systèmes bidirectionnels centralisés (\eg, 4G ou Wi-Fi), et aujourd'hui la collecte décentralisée de données pour les réseaux de l'Internet des objets (\eg, réseaux capteurs).
 %
@@
 même si une station de base centrale est toujours en charge de nombreux objets (ou appareils),
-ce sont les objets qui déclenchent l'envoi de paquets radio, et les seules données descendantes qu'ils peuvent recevoir sont de courts accusés de réception (\emph{acknowledgements}, \Ack), envoyés par la station de base pour indiquer le succès ou l'échec de chaque paquet envoyé.
+ce sont les objets qui déclenchent l'envoi de paquets radio, et les seules données descendantes qu'ils peuvent recevoir sont de courts accusés de réception (ou acquittements, en anglais \emph{acknowledgements}, noté \Ack), envoyés par la station de base pour indiquer le succès ou l'échec de chaque paquet envoyé.
 Cette famille de systèmes sans fil est appelée Internet des Objets (IdO, ou \emph{Internet of Things}, IoT),
@@
 
-Pour le développement futur des \guillemotleft{} réseaux intelligents \guillemotright{} (\emph{smart grid}), des \guillemotleft{} villes intelligentes \guillemotright{} (\emph{smart cities}), des \guillemotleft{} maisons intelligentes \guillemotright{} (\emph{smart homes}) ou de l'\guillemotleft{} agriculture intelligente \guillemotright{} (\emph{smart agriculture}), des réseaux de capteurs devraient être largement déployés.
-Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les \guillemotleft{} bâtiments connectés \guillemotright{} et l'\guillemotleft{} agriculture connectée \guillemotright{}.
-Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides en utilisant des réseaux de capteurs afin d'obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, pour permettre au contrôle centralisé du chauffage d'optimiser son coût et sa consommation énergétique.
+Pour le développement futur des \guillemotleft{} réseaux intelligents \guillemotright{} (\emph{smart grid}), des \guillemotleft{} villes intelligentes \guillemotright{} (\emph{smart cities}), des \guillemotleft{} maisons intelligentes \guillemotright{} (\emph{smart homes}) ou de \guillemotleft{} l'agriculture intelligente \guillemotright{} (\emph{smart agriculture}), des réseaux de capteurs devraient être largement déployés.
+Deux exemples d'applications futures qui sont déjà en cours de déploiement, en France ou à l'étranger, sont les \guillemotleft{} bâtiments connectés \guillemotright{} et \guillemotleft{} l'agriculture connectée \guillemotright{}.
+Pour les bâtiments, l'objectif principal est de réduire le coût de chauffage des bâtiments vides en utilisant des réseaux de capteurs afin d'obtenir des données précises et régulières sur la température dans toutes les pièces et tous les étages, pour permettre au contrôle centralisé du chauffage de minimiser son coût et sa consommation énergétique.
 Pour l'agriculture, un exemple peut être d'équiper chaque tête de bétail (dans les grandes fermes) de capteurs qui émettent régulièrement des informations biologiques, telles que la température corporelle ou le niveau de stress, afin d'optimiser l'heure de traite des vaches, de surveiller la santé des animaux, etc.
 
-Ce troisième type de systèmes sans fil est caractérisé par leur nature décentralisée,
-où les communications sont initialisées et régulées par les objets, et non par un système de contrôle centralisé.
+Ce troisième type de systèmes sans fil est caractérisé par sa nature décentralisée,
+puisque les communications sont initialisées et régulées par les objets, et non par un système de contrôle centralisé.
 %
-En effet, un contrôle centrale nécessite des paquets de signalisation très réguliers, qui ont été identifiés comme trop lourds pour ce type de systèmes.
+En effet, un contrôle central nécessite des paquets de signalisation très réguliers, qui ont été identifiés comme trop lourds pour ce type de systèmes.
 Dans les réseaux de l'IdO actuels et futurs, de nombreux objets hétérogènes utilisent la même station de base (aussi appelée point d'accès, ou \emph{base station}, \emph{access point} ou \emph{gateway}) pour des applications différentes.
@@
 Ainsi, la plupart des entreprises promettant des solutions d'IdO vendent aujourd'hui des objets ayant une durée de vie supérieure à $10$ ans, comme SIGFOX \cite{Centenaro16}.
-Une autre contrainte commune aux objets de l'IdO est leur faible besoin en communication, car la plupart des applications n'auront besoin d'envoyer qu'un ou quelques messages chaque jour, en opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et WiFi).
+Une autre contrainte commune aux objets de l'IdO est leur faible besoin en communication, car la plupart des applications n'auront besoin d'envoyer qu'un ou quelques messages chaque jour, en nette opposition avec le débit de données élevé recherché pour les systèmes centralisés (tels que 4G/5G et Wi-Fi).
 %
@@
 et la couche \emph{M}edium \emph{A}ccess \emph{C}ontrol (\emph{MAC}).
-Pour citer quelques exemples de normes pour des réseaux de type IdO, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée (jusqu'à \SI{2}{\meter}), tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
-Nous nous référons à l'article \cite{Centenaro16} pour plus de détails, et les références dans \cite{Azari18} ou notre dernier travail \cite{MoyBesson2019,MoyBesson2019Annales}.
+Pour citer quelques exemples de normes pouvant être utilisées pour des réseaux de type IdO, ZigBee, Z-Wave ou Bluetooth visent des communications à courte portée (jusqu'à \SI{2}{\meter}), tandis que LoRaWAN, SIGFOX, Ingenu ou Weightless sont destinés aux communications à plus longue portée (jusqu'à \SI{50}{\kilo\meter}).
+Nous référons à l'article \cite{Centenaro16} pour plus de détails, et les références dans \cite{Azari18} ou notre dernier travail \cite{MoyBesson2019Annales}.
 
@@
 Un problème majeur des technologies sans fil actuelles est la question de l'épuisement du spectre radio :
-dans la plupart des bandes de fréquences, la totalité du spectre RF est désormais attribuée et les bandes libres n'existent plus, ce qui limite la possibilité d'ajouter tout nouvel usage.
-Tel qu'illustré en Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} ci-dessous,
+dans la plupart des bandes de fréquences, la totalité du spectre de radio fréquences (RF) est désormais attribuée et les bandes libres n'existent plus, ce qui limite la possibilité d'ajouter de nouveaux usages.
+Comme le montre la Figure~\ref{fig:1:United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum} ci-dessous,
 presque tout le spectre
@@
 Les bandes des réseaux cellulaires (2G/3G/4G/5G) sont surchargées dans la plupart des régions du monde, mais d'autres bandes de fréquences (comme les fréquences militaires ou de radio amateures) sont moins utilisées.
-Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre dépend à la fois du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
+Des études indépendantes réalisées dans certains pays ont confirmé cette observation et conclu que l'utilisation du spectre peut fortement dépendre à la fois du moment et du lieu, comme le montre \cite{Lopez2009spectral} par exemple.
 En outre, l'attribution fixe du spectre empêche
@@
 %
-C'est ainsi qu'au cours des $15$ dernières années, grâce à un lobbying actif de la communauté de la radio intelligente,
+C'est ainsi qu'au cours des quinze dernières années, grâce à un lobbying actif de la communauté de la radio intelligente,
 les organismes de réglementation dans le monde se sont demandé s'il fallait permettre un nouveau paradigme de communication sans fil :
@@
 Une telle radio détecte automatiquement les canaux disponibles dans le spectre radio, puis modifie en conséquence ses paramètres de transmission ou de réception pour permettre plus de communications sans fil simultanées dans une bande de spectre donnée à un endroit donné} \guillemotright{}.
+    (\href{https://en.wikipedia.org/wiki/Cognitive_radio}{\texttt{en.Wikipedia.org/wiki/Cognitive\_radio}})
 \end{itemize}
@@
 La réglementation stipule que les UP ont une priorité stricte,
-mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection spectrale efficace, elle pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de tout UP.
+mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection spectrale efficace, il pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de tout UP.
 Ceci définit le concept d'\textbf{Accès Opportuniste au Spectre} (AOS, \emph{Opportunistic Spectrum Access}, OSA), pour lequel nous nous référons à l'article \cite{Zhao07} pour plus de détails.
@@
 Le livre de référence sur l'ApR est \cite{SuttonBarto2018}, citons donc la définition qu'en donnent R. Sutton et A. Barto :
-\guillemotleft{} \emph{l'apprentissage par renforcement, c'est apprendre quoi faire - comment faire le lien entre les situations et les actions -
-pour maximiser un signal de récompense numérique. L'apprenant n'est pas informé des mesures à prendre pour
-mais doit plutôt découvrir quelles actions sont les plus gratifiantes en les essayant} \guillemotright.
+\guillemotleft{} \emph{l'apprentissage par renforcement, c'est apprendre quoi faire -- comment faire le lien entre les situations et les actions -- pour maximiser un signal de récompense numérique.
+L'apprenant n'est pas informé des mesures à prendre mais doit plutôt découvrir quelles actions sont les plus gratifiantes en les essayant} \guillemotright.
 % apprentissage statistique > séquentiel > par renforcement > information partielle (bandits)
@@
     }
-\caption{Cycle de l'apprentissage par renforcement : une joueuse interagit avec son environnement par des actions, et observe une récompense, de façon itérative.}
+\caption{Cycle de l'apprentissage par renforcement : un-e joueur-se interagit avec son environnement par des actions, et observe une récompense, de façon itérative.}
 \label{fig:1:ReinforcementLearningCycle_enFr}
@@
 En d'autres termes, la joueuse ne voit que la récompense donnée par son action à chaque tour, et non la récompense qui aurait été donnée si elle avait choisi une des autres actions.
-Ce genre de rétroaction limitée s'appelle l'information de type bandit (\emph{bandit feedback}), et nous discutons de l'histoire et du concept des \emph{bandits} (BMB) en détail dans le prochain Chapitre~\ref{chapter:2}.
-Les essais cliniques et la découverte de traitements ont été historiquement la première application des BMB depuis les $1930$, avec les premiers travaux de W. Thompson \cite{Thompson33},
+Ce genre de rétroaction limitée s'appelle l'information de type bandit (\emph{bandit feedback}), et nous discutons de l'histoire et du concept des \emph{bandits} (BMB) en détails dans le prochain Chapitre~\ref{chapter:2}.
+Les essais cliniques et l'identification du meilleur traitement ont été historiquement les premières applications des BMB depuis les années $1930$, avec les premiers travaux de W.
+Thompson \cite{Thompson33},
 %
@@
 
-En se concentrant sur un US dans un réseau AOS, il doit décider séquentiellement d'un canal à analyser (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il écoute ce canal et effectue une détection de PU, et enfin il transmet dans ce canal s'il a été détecté libre.
+En se concentrant sur un US s'insérant dans un réseau permettant l'AOS, il doit décider séquentiellement d'un canal à analyser (dans l'ensemble des canaux, $[K]=\{1,\dots,K\}$), puis il écoute ce canal et effectue une détection de PU, et enfin il transmet dans ce canal s'il a été détecté libre.
 Le but de l'US est de minimiser sa consommation d'énergie (nous rappelons que nous nous concentrons sur la radio intelligente \guillemotleft{} écologique \guillemotright{}, \emph{green radio}) et de maximiser son débit de données sur la liaison montante, ou de façon équivalente, de maximiser son nombre de transmissions réussies.
@@
 sous l'information de rétroaction partielle (\ie, l'US reçoit l'information concernant un seul canal parmi $K$).
-Ces trois hypothèses sont celles qui limitent le cadre général de l'apprentissage séquentiel au cas spécifique des bandit multi-bras (voir Figure~\ref{fig:1:ReinforcementLearningCycle}).
+Ces trois hypothèses sont celles qui limitent le cadre général de l'apprentissage séquentiel au cas spécifique des bandit multi-bras (voir Figure~\ref{fig:1:ReinforcementLearningCycle_enFr}).
 
@@
 %
-Les travaux antérieurs de notre équipe SCEE ont montré que les BMB peuvent être utilisés pour modéliser le problème de l'AOS :
+Les travaux antérieurs de notre équipe SCEE ont montré que les BMB peuvent être utilisés pour modéliser le problème de prise de décision de l'AOS :
 les bandes de fréquences orthogonales (ou canaux) sont modélisées par des bras $k\in\{1,\dots,K\}$,
-et la récompense obtenue par l'objet équipé de RI après avoir analysé le canal $k$ au temps $t$ est modélisée par une récompense $r(t) \in \{0,1\}$.
+et la récompense obtenue par l'objet après avoir analysé le canal $k$ au temps $t$ est modélisée par une récompense $r(t) \in \{0,1\}$.
 En effet, $r(t) = 1$ indique qu'aucun utilisateur primaire n'a été détecté (et donc qu'un message d'US peut être envoyé), alors que $r(t)=0$ indique que le canal $k$ est occupé durant l'intervalle de temps $t$, et qu'aucun message d'US ne doit être envoyé, jusqu'à la fin de cet intervalle.
@@
 La littérature susmentionnée a essentiellement montré que les algorithmes BMB peuvent être appliqués avec succès au problème de l'AOS.
-Toutefois, si l'on considère des objets de RI qui ne peuvent pas effectuer de détection spectrale active, tels que des objets à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux de l'IdO, le modèle BMB qui utilise cette détection active pour vérifier l'absence d'UP dans le cas de l'AOS ne peut plus être appliqué.
+Toutefois, si l'on considère des objets de RI qui ne peuvent pas effectuer de détection spectrale active, tels que des objets à faible coût et à faible consommation d'énergie conçus pour les futurs réseaux de l'IdO, le modèle BMB qui utilise cette détection active pour vérifier l'absence d'UP dans le cadre de l'AOS ne peut plus être appliqué.
 En outre, dans la plupart des cas, les réseaux de l'IdO utilisent des bandes non licenciées et il n'y a donc plus de distinction entre UP et US.
@@
 %
-\begin{itemize}%\tightlist
+\begin{itemize}\tightlist
     % \item
@@
     \item
-    la plupart des objets sont à faible coût, fonctionnent sur du matériel de mauvaise qualité et ont des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
+    La plupart des objets sont à faible coût, fonctionnent sur du matériel de mauvaise qualité et ont des capacités logicielles et d'alimentation limitées (c'est-à-dire de minuscules batteries).
     Cela signifie que même s'ils ne sont pas équipés de capacités de détection particulières,
@@
 %
-En utilisant cette rétroaction, qui consiste en un \emph{acquittement} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence ' \Ack{} après une transmission échouée, il est possible qu'un objet de l'IdO puisse également optimiser ses communications, grâce à un algorithme d'apprentissage par renforcement correctement conçu.
+En utilisant cette rétroaction, qui consiste en un \emph{acquittement} (ou \emph{acknowledgement}, \Ack) reçu peu après chaque transmission réussie, ou en l'absence d'\Ack{} après une transmission échouée, il est possible qu'un objet de l'IdO puisse également optimiser ses communications, grâce à un algorithme d'apprentissage par renforcement correctement conçu.
 
@@
 Si nous résumons les problèmes étudiés en une seule question, ce pourrait être la suivante :
-\emph{peut-on adapter les outils de prise de décision déjà appliqués avec succès à la Radio Intelligente (RI) pour l'Accès Opportuniste au Spectre (AOS) aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des Objets (IdO) ?}
+\guillemotleft{} \emph{peut-on adapter les outils de prise de décision déjà appliqués avec succès à la Radio Intelligente (RI) pour l'Accès Opportuniste au Spectre (AOS) aux besoins spécifiques de la RI pour les (futurs) réseaux de l'Internet des Objets (IdO) ?} \guillemotright.
 %
@@
 %
-Ce dernier chapitre est basé sur notre dernier travail, publié à la conférence GRETSI à Lille, en août $2019$ \cite{Besson2019Gretsi}.
+Ce dernier chapitre est basé sur un de nos derniers travaux, publié à la conférence GRETSI à Lille, en août $2019$ \cite{Besson2019Gretsi}.
 Notre travail a aussi mené à une version longue \cite{Besson2019GLRT}, qui sera complétée avec des résultats plus récents, afin de le soumettre bientôt à une revue (probablement le Journal of Machine Learning Research) avant décembre $2019$.

### 2019-07-11 2-Chapters/1-Chapter/chapter1.tex
Okay, fixed bug of transparent last page
Almost done proofreading résumé fr

@@
 I was supervised by Professor Christophe Moy, in Rennes,
-and I was also co-supervised by Doctor Emilie Kaufmann, whom I visited many times at Inria Lille Nord Europe in Lille (France).
+and I was also co-supervised by Doctor Émilie Kaufmann, whom I visited many times at Inria Lille Nord Europe in Lille (France).
 
@@
 %
-Historically, three main families of wireless communication systems have been deployed in massive commercial networks: first, centralized broadcasting (\eg, radio or TV broadcasting), then centralized bi-directional systems (\eg, 4G or WiFi), and nowadays decentralized data harvesting for the Internet of Things (IoT) networks (\eg, sensor networks).
+Historically, three main families of wireless communication systems have been deployed in massive commercial networks: first, centralized broadcasting (\eg, radio or TV broadcasting), then centralized bi-directional systems (\eg, 4G or Wi-Fi), and nowadays decentralized data harvesting for the Internet of Things (IoT) networks (\eg, sensor networks).
 %
@@
 % In most systems, there is still a base station (with one or more antenna) in charge of a large area (or cell), and many devices are able to receive and transmit data to the base station.
-% Nowadays, the most well-known systems include WiFi, and systems deployed for mobile phone communications, with a long list of standards from 2G, 3G, 4G and now 5G, still in development in $2019$.
-% Even if both systems initially differed in usage, as 2G was designed to exchange only voice, and WiFi to link wireless devices at home to an Internet-connected box, they are conceptually very close.
+% Nowadays, the most well-known systems include Wi-Fi, and systems deployed for mobile phone communications, with a long list of standards from 2G, 3G, 4G and now 5G, still in development in $2019$.
+% Even if both systems initially differed in usage, as 2G was designed to exchange only voice, and Wi-Fi to link wireless devices at home to an Internet-connected box, they are conceptually very close.
 % As the systems are cellular, any device knows the base station it is associated with, and the bi-directional nature of their exchanges makes a centralized control possible.
@@
 Most commercial IoT companies nowadays indeed sell a lifespan larger than $10$ years, like SIGFOX \cite{Centenaro16}.
-Another common constraint for IoT devices is their low duty cycles, as most applications target a need for one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems (such as 4G/5G and WiFi).
+Another common constraint for IoT devices is their low duty cycles, as most applications target a need for one or a few messages to send every day, in strike opposition to the high data-rate pursued for centralized systems (such as 4G/5G and Wi-Fi).
 %
@@
     % \includegraphics[width=0.65\textwidth,angle=90]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum.pdf}
-    \includegraphics[width=0.92\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
+    \includegraphics[width=0.99\textwidth]{United_States_Frequency_Allocations_Chart_2016_The_Radio_Spectrum_3.pdf}
     \caption[A chart representing the allocation of radio spectrum in the United States of North America in $2016$]{A chart representing the allocation of radio spectrum in the United States of North America in $2016$.
@@
 
-% \TODOL{Emilie pense que je devrais faire ça mieux :
+% \TODOL{Émilie pense que je devrais faire ça mieux :
 %     Quant au survey du machine learning (Figure 1.1), je ne pense pas que
@@
 Such a radio automatically detects available channels in wireless spectrum, then accordingly changes its transmission or reception parameters to allow more concurrent wireless communications in a given spectrum band at one location''}.
+    (\href{https://en.wikipedia.org/wiki/Cognitive_radio}{\texttt{en.Wikipedia.org/wiki/Cognitive\_radio}})
 \end{itemize}
@@
     \emph{Multi-Player Bandits Revisited},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann.
+    by \textbf{Lilian Besson} \& Émilie Kaufmann.
     \emph{Algorithmic Learning Theory},
@@
     \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},
-    by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
+    by \textbf{Lilian Besson}, Émilie Kaufmann \& Christophe Moy.
     \emph{Wireless Communication and Networks Conference},
@@
     \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings},
-    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& Jacques Palicot.
+    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Émilie Kaufmann \& Jacques Palicot.
     \emph{Conference on Cognitive Radio Oriented Wireless Networks},
@@
     \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
-    by \textbf{Lilian Besson} \& Emilie Kaufmann,
+    by \textbf{Lilian Besson} \& Émilie Kaufmann,
     GRETSI 2019,
@@
     \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann,
+    by \textbf{Lilian Besson} \& Émilie Kaufmann,
     February $2019$.\\
@@
     \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},
-    by \textbf{Lilian Besson} \& Emilie Kaufmann,
+    by \textbf{Lilian Besson} \& Émilie Kaufmann,
     September $2018$.

### 2019-07-11 1-Preamble/5-Abstract/resume_fr.tex
Merci Christophe

@@
 car il existe de nombreux algorithmes différents, avec de nombreuses variantes du problème présenté ci-dessus.
-Nous commençons donc la première Partie~\ref{part:Introduction} de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle BMB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires de BMB.
+Nous commençons donc la première partie de cette thèse par le Chapitre~\ref{chapter:2}, qui présente le modèle BMB et passe en revue les algorithmes les plus importants conçus pour résoudre les problèmes stochastiques et stationnaires de bandits.
 %
@@
 nous ne nous sommes pas seulement intéressés par la mesure habituelle des performances d'un algorithme BMB (\ie, son regret, voir ci-dessous en Section~\ref{sec:2:lowerUpperBoundsRegret}),
-mais aussi par leurs performances empiriques en termes de complexité de calcul et de stockage, tant du point de vue des analyses théoriques que des mesures réelle.
+mais aussi par leurs performances empiriques en termes de complexité de calcul et de stockage, tant du point de vue des analyses théoriques que des mesures réelles.
 
@@
 Notre exploration du grand nombre d'algorithmes et de modèles BMB développés dans la littérature récente
-nous a donné l'ambition d'écrire un seul logiciel permettant à n'importe quel-le chercheur-euse d'implémenter facilement de nouveaux modèles et algorithmes, afin de comparer les modèles existants et d'explorer empiriquement les performances de nouveaux algorithmes.
+nous a donné l'ambition d'écrire un seul logiciel permettant à n'importe quel-le chercheur-euse d'implanter facilement de nouveaux modèles et algorithmes, afin de comparer les modèles existants et d'explorer empiriquement les performances de nouveaux algorithmes.
 Pour atteindre cet objectif, nous avons développé une bibliothèque de simulation des problèmes BMB, dans le langage populaire Python.
 %
-D'après ce que les auteurs ont identifié,
-nous avons écrit la bibliothèque libre de simulation la plus exhaustive pour les problèmes BMB, appelée SMPyBandits, qui est publiée en ligne sous licence open-source \cite{SMPyBanditsJMLR,SMPyBandits} et hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
+À notre connaissance, nous avons écrit la bibliothèque libre de simulation la plus exhaustive pour les problèmes BMB, appelée SMPyBandits \cite{SMPyBanditsJMLR,SMPyBandits}, qui est publiée en ligne sous licence libre (open-source) et qui est hébergée sur \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}}.
 Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}.
@@
 La première est une comparaison empirique d'algorithmes existants, se focalisant sur les plus efficaces et les mieux connus, en Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
-Nous confirmons que les algorithmes largement utilisés, tels que \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} et \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
-La seconde approche est la sélection d'algorithmes en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons en Chapitre~\ref{chapter:25}.
-Ce travail sur l'agrégation des algorithmes BMB a été motivé pour le cas AOS de la radio cognitive, où de nombreux travaux antérieurs ne considéraient que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
+Nous confirmons que les algorithmes les plus utilisés, tels que \UCB{} \cite{Auer02}, l'échantillonage de Thompson (Thompson sampling) \cite{Thompson33} et \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
+La seconde approche est la sélection d'algorithme en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons en Chapitre~\ref{chapter:25}.
+Ce travail sur l'agrégation des algorithmes BMB était motivé par le cadre AOS, pour lequel les travaux antérieurs ne considéraient que l'algorithme $\UCB_1$ sans vraiment justifier ce choix.
 Cette contribution a été présentée à la conférence IEEE WCNC à Barcelone, en Espagne, en avril 2018.
@@
 Nous étudions deux modèles de réseaux de l'IdO dans le Chapitre~\ref{chapter:4}.
-Le premier modèle considère que les objets ont simplement des données à envoyer régulièrement à une station de base (à des moments imprévisibles ou aléatoires).
-De tels objets utilisent les acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux (c'est-à-dire le canal le moins occupé par le trafic qui les entoure).
-L'objectif de ce premier modèle est d'améliorer la qualité de service (QdS, ou \emph{Quality of Service} QoS) de l'application de ce réseau de l'IdO, en appliquant un ApR décentralisé côté objet afin de réduire le taux d'échec de transmission.
+Le premier modèle considère que les objets ont simplement des données à envoyer régulièrement à une station de base, à des moments imprévisibles ou aléatoires.
+De tels objets utilisent les acquittements comme un retour d'information, afin d'optimiser leurs communications sur la liaison montante, en accédant aux meilleurs canaux, c'est-à-dire le canal le moins occupé par le trafic ambiant.
+L'objectif de ce premier modèle est d'améliorer la qualité de service (QdS, ou \emph{Quality of Service}, QoS) de l'application de ce réseau de l'IdO, en appliquant un ApR décentralisé (côté objet), afin de réduire le taux d'échec de transmission.
 La station de base recevra donc plus de paquets de liaison montante des objets desservis, s'ils peuvent apprendre par eux-mêmes un accès efficace au spectre.
@@
 %
-Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux modèles soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet dans ce cas d'augmenter la durée de vie des batteries de chacun des objets, ainsi que la QdS de l'ensemble du réseau, puisque moins de retransmissions réduisent la charge locale du spectre.
+Notre deuxième modèle considère ensuite les retransmissions de paquets, et bien que les deux modèles soient similaires, l'application d'un algorithme d'apprentissage décentralisé efficace permet dans ce cas d'augmenter la durée de vie des batteries de chacun des objets, ainsi que la QdS de l'ensemble du réseau, puisque moins de retransmissions réduisent aussi la charge locale du spectre.
 
 Le premier modèle sans retransmission est issu du premier article écrit au cours de cette thèse, qui a initié une collaboration avec R. Bonnefoi, un autre doctorant de notre équipe SCEE.
-Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre 2010, où il a obtenu le \guillemotleft{} prix du meilleur papier \guillemotright{} \cite{Bonnefoi17}.
+Nous l'avons présenté à la conférence EAI CROWNCOM à Lisbonne, au Portugal, en septembre $2017$, où il a obtenu le \guillemotleft{} prix du meilleur papier \guillemotright{} \cite{Bonnefoi17}.
 %
@@
 car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
-Il est en fait équivalent aux modèles précédemment étudiés d'applications des BMB pour l'AOS, puisqu'ils ne prennent en compte que des objets synchronisés, un US et des UP, et donc si la détection indique qu'un canal est libre pour un intervalle de temps, le message envoyé par l'US sera certainement reçu avec succès par la station de base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc pas besoin d'\Ack.
+Il est en fait équivalent aux modèles précédemment étudiés d'applications des BMB pour l'AOS, puisqu'ils ne prennent en compte que des objets synchronisés, un US et des UP, et donc si la détection indique qu'un canal est libre pour un intervalle de temps, le message envoyé par l'US sera certainement reçu avec succès par la station de base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc il n'a pas besoin d'\Ack.
 
-D'une part, dans le cadre de l'AOS, nous n'avons pas réussi à obtenir un résultat positif pour l'algorithme \Selfish, car nous avons prouvé que dans certains petits problèmes (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre d'un regret moyen linéaire.
+D'une part, dans le cadre de l'AOS, nous n'avons pas réussi à obtenir de résultat positif pour l'algorithme \Selfish, car nous avons prouvé que dans certains petits problèmes (\eg, $K=3$ et $M=2$), \Selfish-\UCB{} peut présenter un regret linéaire avec une faible probabilité, et donc souffre d'un regret moyen linéaire.
 Ce résultat a ensuite été confirmé et analysé par d'autres chercheurs dans \cite{BoursierPerchet18} (Annexe~F).
@@
 Comme expliqué plus haut, le modèle de Chapitre~\ref{chapter:4} s'est révélé trop complexe à analyser, principalement parce que nous considérons un réseau d'IdO avec de nombreux objets, chacun suivant des profils d'activation aléatoires.
-La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes BMB, conçus pour résoudre des problèmes stationnaires, appliqués à un problème non stationnaire,
+La difficulté ne réside pas dans le fait que nous essayons d'analyser des algorithmes de bandits, conçus pour résoudre des problèmes stationnaires, appliqués à un problème non stationnaire,
 mais plutôt que nous considérons des algorithmes qui jouent tous à différents instants d'activations (aléatoires), ce qui donne des sous-ensembles différents et imprédictibles des pas de temps synchronisés globaux.

### 2019-07-11 2-Chapters/1-Chapter/chapter1.tex
Merci Christophe

@@
 If we sum-up the studied problems and summarize them in one question, it could be the following:
-\emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
+\emph{``Can we adapt the decision making tools, already successfully applied to Cognitive Radio for Opportunistic Spectrum Access, to the specific needs of CR for the (future) Internet of Things networks?''}
 %
@@
 We study two models of IoT networks in Chapter~\ref{chapter:4}.
-The first model considers IoT devices that simply have data to regularly send to a base station (at unpredictable or random times).
-Such devices use acknowledgements as feedback, in order to optimize their up-link communications, by accessing more the best channels (\ie, the channel less occupied by the surrounding traffic).
+The first model considers IoT devices that simply have data to regularly send to a base station, at unpredictable or random times.
+Such devices use acknowledgements as feedback, in order to optimize their up-link communications, by accessing more the best channels, \ie, the channel less occupied by the surrounding traffic.
 The goal of this first model is to increase the Quality of Service (QoS) of such IoT network, by applying decentralized RL on the device side in order to reduce the rate of failed transmissions.

### 2019-07-11 2-Chapters/Conclusion/conclusion.tex
Trying to kill this last empty page...
Merci Christophe

@@
 
-\TODOL{Je dois faire quelques phrases résumant la thèse, mais plus court}
-
 We started this manuscript by detailing the historical, scientific and technical contexts of our research during this PhD.
@@
 %
-The main question we studied was \emph{``Can we adapt the decision making tools already successfully applied to Cognitive Radio for Opportunistic Spectrum Access to the specific needs of CR for the (future) Internet of Things networks?''}
+The main question we studied was \emph{``Can we adapt the decision making tools, already successfully applied to Cognitive Radio for Opportunistic Spectrum Access, to the specific needs of CR for the (future) Internet of Things networks?''}
 %
@@
 We first gave a strong theoretical background on the reinforcement learning model considered in this thesis, that is the multi-armed bandit model.
-We then focused on our problems of solving the spectrum scarcity issue in unlicensed bands, in the context of the future Internet of Things networks,
-and on extending to the specificities of such IoT networks the underlying ideas of previously studied applications of machine learning to solve the spectrum scarcity issue in licensed networks.
+We then focused on our problems of solving the spectrum scarcity issue in unlicensed bands, in the context of the future Internet of Things networks.
+We worked on extending to the specificities of such IoT networks the ideas underlying the previously studied applications of machine learning to solve the spectrum scarcity issue in licensed networks.
 
@@
 or to drop their packet after a failed transmission (\ie, with or without retransmission).
-Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order for the numerous IoT devices to improve on their own their spectrum access scheme, by learning which of the $K$ frequency channels are the less occupied by the environmental traffic (in average).
+Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order that the numerous IoT devices improve their spectrum access scheme, on their own, by learning which of the $K$ frequency channels are the less occupied by the environmental traffic (in average).
 
@@
 For a company that manufactures IoT end-devices, it is very simple and cheap in terms of both hardware and software costs to implement such MAB algorithms on the low-cost embedded processors, and equip all the produced IoT devices with this capacity of using online machine-learning for their spectrum access.
-The embedded decision-making algorithm automatically improves the channel selection of each device, in a fully decentralized way, and increases the total number of successful up-link transmissions, and thus this increases the Quality of Service of the considered IoT application and allow more devices to be served by the same IoT gateway.
+The embedded decision-making algorithm automatically improves the channel selection of each device, in a fully decentralized way.
+It also increases the total number of successful up-link transmissions, and thus this increases the Quality of Service of the considered IoT application.
+Finally, such algorithm also allows more devices to be served by the same IoT gateway.
 %
@@
 
-Moreover, from a theoretical aspect, even though it was found hard to propose a rigorous analysis of our model of IoT networks, because of the large number of devices having different random and unpredictable activation patterns, our contributions also include an analysis of two restricted models.
+Moreover, from a theoretical point-of-view, even though it was found hard to propose a rigorous analysis of our model of IoT networks, because of the large number of devices having different random and unpredictable activation patterns, our contributions also include an analysis of two restricted models.
 %
@@
 Finally, from an application point-of-view, we also presented a realistic implementation of a proof-of-concept of our first model of IoT network, that was used to validate empirically the proposed approach.
-We showed that the IoT devices can effectively use a multi-armed bandit algorithm to automatically learn to favor certain frequency channels over others, in order to optimize their spectrum access, to reduce their number of failed transmissions.
+We showed that the IoT devices can effectively use a multi-armed bandit algorithm to automatically learn to favor certain frequency channels over others, and finally to optimize their spectrum access, to reduce their number of failed transmissions.
 %
@@
 % ----------------------------------------------------------------------------
-% \section{Perspectives}
-\subsection*{Future works}
+\subsection*{Perspectives}
+% \subsection*{Future works}
 
@@
 
+\paragraph{For our models of IoT networks.}
+%
 The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if the $S$ static and $D$ dynamic devices have different transmission patterns, and less easily with a different probability per device.
@@
 
+\paragraph{For multi-player bandits.}
+%
 The study on multi-players bandits in Chapter~\ref{chapter:5} suggests several further research directions.
-First, one could want to investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
+First, one could investigate the notion of \emph{optimal algorithms} in the decentralized multi-players model with sensing information.
 So far we provided the first matching upper and lower bound on the expected number of sub-optimal arms selections, which suggests some form of (asymptotic) optimality.
-However, sub-optimal draws turn out not be the dominant terms in the regret, both in our upper bounds and in practice, thus a promising future work is to identify some notion of \emph{minimal number of collisions}, similarly to what a recent work \cite{wang2019distributed} studied for the minimal amount of communication needed to achieve logarithmic regret, in a similar model that authorizes direct communications between players.
+However, sub-optimal draws turn out to not be the dominant terms in the regret, both in our upper bounds and in practice.
+Thus a promising future work is to identify some notion of \emph{minimal number of collisions}.
+This could be similar to what a recent work \cite{wang2019distributed} studied for the minimal amount of communication needed to achieve logarithmic regret, in a similar model that authorizes direct communications between players.
 % Similarly to what was done very recently in \cite{wang2019distributed} for the communications between players who collaborate with each other, it would be interesting to characterize the number of collisions needed to achieve logarithmic regret.
@@
 (see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
-From the point-of-view of the theoretical analysis, we are especially interesting about extending our proposed algorithm \MCTopM{} to the case of an unknown number of player, and then to the most interesting extension for the application to wireless networks, the ``dynamic case'' which allows for arrivals and departures of players in the multi-player bandit game.
+From the point-of-view of the theoretical analysis, we are especially interesting about first extending our proposed algorithm \MCTopM{} to the case of an unknown number of players.
+Then we could also address the most interesting extension for the application to wireless networks, \ie, the ``dynamic case'' which allows for arrivals and departures of players in the multi-player bandit game.
 %
@@
 
-Regarding non-stationary bandits, we believe that the new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
+\paragraph{Regarding non-stationary bandits.}
+%
+We believe that the new proof technique of Chapter~\ref{chapter:6} could be used to analyze \GLRklUCB{} under less stringent assumptions than the one made in this chapter (and in previous works), that would require only a few ``meaningful'' changes to be detected.
 This promising research direction is left for future work,  but the hope is that the regret would be expressed in terms of the number of such meaningful changes, instead of the number of break-points $\Upsilon_T$.
@@
 
+\paragraph{For SMPyBandits.}
+%
 A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
@@
 % \subsection*{Other directions of future works}
-\subsection*{Unifying multi-player and non-stationary bandits.}
+\paragraph{Unifying multi-player and non-stationary bandits.}
 %
@@
 
-\vfill{}
+% \vfill{}
 
-\hr{}
+% \hr{}
 

### 2019-07-11 2-Chapters/3-Chapter/chapter3.tex
Done for now
Done for end of conclusion

@@
 %
-\SMPyBandits{} is a library I have developed in the Python language since the beginning of my PhD.
+\SMPyBandits{} is a library that has been developed in the Python language since the beginning of this PhD.
 It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
@@
 %
-We detailed three main points of our library:
-$(i)$ its purpose and qualities,
-$(ii)$ its organization,
-and $(iii)$ an overview of its usage:
+We detailed the three following main points of our library:
+% $(i)$ its purpose and qualities,
+% $(ii)$ its organization,
+% and $(iii)$ an overview of its usage:
 %
@@
 
-\paragraph{Future works.}
+\subsection*{Perspectives}
 %
@@
 
-The most exciting thing that could happen to our \SMPyBandits{} library would be to see it gaining popularity!
-Its documentation has seen already about $20000$ visits, \href{https://github.com/SMPyBandits/SMPyBandits/stargazers}{the project on GitHub had $90$ stars} in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used \SMPyBandits.
+An interesting task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
+It is already possible to reproduce some of the experiments presented in this thesis,
+by following the instructions given in the documentation (see Section~\ref{sub:3:listResearchWorksUsingSMPyBandits}),
+by using the Jupyter notebooks \cite{jupyter} made available%
+\footnote{~These notebooks are hosted on both the GitHub repository of SMPyBandits and on my website on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}.
+They can be used locally if you install the library, but can also be used on cloud platforms, like Binder \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} or Google Colab \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}}.
+
+Finally, the most exciting thing that could happen to our \SMPyBandits{} library would be to see it gaining popularity!
+Its documentation has seen already about $20000$ visits, \href{https://github.com/SMPyBandits/SMPyBandits/stargazers}{the project on GitHub had $90$ stars} in June $2019$, and based on the features requested and the emails received about it, we counted between ten to twenty researchers in other labs who use or used \SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
@@
 As a personal note, I would like to continue working on \SMPyBandits, and implement the most interesting requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
-I would also love to be able to teach a graduate or under-graduate course on reinforcement learning and multi-armed bandits in the next years, and use my library as a support to illustrate such course, as well as practical sessions.
+I would also love to be able to teach an under-graduate course on reinforcement learning and bandits in the future, and to use my library as a support to illustrate a course and its practical sessions.
 

### 2019-07-11 2-Chapters/Conclusion/conclusion.tex
Done for end of conclusion

@@
 % \section{General Conclusion}
-% \subsection*{General Conclusion}
+\subsection*{Conclusion on our contributions}
 
@@
 
-\paragraph{For SMPyBandits.}
+\paragraph{For our library SMPyBandits.}
 %
-A few tasks are left on our library SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
+A few tasks are left on SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
 For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
@@
 to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
-%
-A third task that I would have liked to complete is to interface the library with a web-based interactive demonstration, in order to allow anyone to launch simulations without any knowledge of programming.
-It is already possible to reproduce some of the experiments presented in this thesis,
-by following the instructions given in the documentation (see Section~\ref{sub:3:listResearchWorksUsingSMPyBandits}),
-by using the Jupyter notebooks \cite{jupyter} made available%
-\footnote{~These notebooks are hosted on both the GitHub repository of SMPyBandits and on my website on \href{https://perso.crans.org/besson/PhD/notebooks/}{\texttt{perso.crans.org/besson/PhD/notebooks/}}.
-They can be used locally if you install the library, but can also be used on cloud platforms, like Binder \href{https://mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}{\texttt{mybinder.org/v2/gh/SMPyBandits/SMPyBandits/master}} or Google Colab \href{https://colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}{\texttt{colab.research.google.com/github/SMPyBandits/SMPyBandits/tree/master/notebooks/}}}.
 
@@
 
-% \vfill{}
+\vfill{}
 
-% \hr{}
+\hr{}
 

### 2019-07-11 1-Preamble/5-Abstract/resume_fr.tex
Petit changements sur les légendes

@@
 J'étais supervisé par le professeur Christophe Moy, à Rennes,
-et j'étais également co-encadré par le docteur Émilie Kaufmann, à qui j'ai souvent rendu visite à Inria Lille Nord Europe.
+et j'étais également co-encadré par la docteure Émilie Kaufmann, à qui j'ai souvent rendu visite à Inria Lille Nord Europe.
 
@@
 des technologies de l'information et communication (TIC), aux télécommunications, aux communications sans fil, puis à la Radio Intelligente,
-et enfin à la Radio Logicielle (RL, \emph{Software Defined Radio}, SDR).
+et enfin à la Radio Logicielle (RL, ou \emph{Software Defined Radio}, SDR).
 %
@@
 mais même si les bandes RF sont attribuées, les mesures dans le monde réel montrent souvent que certaines bandes ne sont pas utilisées de manière intensive, et donc si un US était équipé d'une capacité de détection spectrale efficace, il pourrait analyser son environnement, et utiliser une bande sous licence si et seulement si elle est exempte de tout UP.
-Ceci définit le concept d'\textbf{Accès Opportuniste au Spectre} (AOS, \emph{Opportunistic Spectrum Access}, OSA), pour lequel nous nous référons à l'article \cite{Zhao07} pour plus de détails.
+Ceci définit le concept d'\textbf{Accès Opportuniste au Spectre} (AOS, ou \emph{Opportunistic Spectrum Access}, OSA), pour lequel nous nous référons à l'article \cite{Zhao07} pour plus de détails.
 
@@
 %
-Notre deuxième modèle, avec retransmissions, correspond à notre dernière collaboration avec R.
+Notre deuxième modèle, avec retransmissions, correspond à la dernière collaboration avec R.
 Bonnefoi, qui a été présentée à l'atelier MoTION, également pendant la conférence IEEE WCNC $2019$ \cite{Bonnefoi2019WCNC}.

### 2019-07-11 2-Chapters/3-Chapter/chapter3.tex
Petit changements sur les légendes

@@
 
-Since December $2016$ I worked regularly on the \SMPyBandits{} library, which is a Python package  designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
+I started the development of the \SMPyBandits{} library in December $2016$, and I worked on it actively in $2017$ and $2018$.
+It is a Python package, designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
 %
@@
 	\caption[Example of a single-player simulation showing the average regret of $4$ algorithms.]{
-		Example of a single-player simulation showing the average regret of four algorithms (\textcolor{red}{\UCB}, \textcolor{blue}{\KLUCBpp}, \textcolor{darkgreen}{\klUCB} and \textcolor{gold}{Thompson sampling}).
+		Example of a single-player simulation showing the average regret of four algorithms (\textcolor{red}{\UCB}, \textcolor{blue}{\KLUCBpp}, \textcolor{darkgreen}{\klUCB} and \textcolor{lightgreen}{Thompson sampling}).
 They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proven to match the lower-bound up-to a constant), and each but UCB is known to be asymptotically optimal (\ie, with the constant matching the lower-bound).
@@
 For instance, Thomson sampling is very efficient in average, and \UCB shows a larger variance.]{Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}.
-For instance, \textcolor{gold}{Thomson sampling} is very efficient in average (\textcolor{gold}{in yellow}), and \textcolor{red}{\UCB} shows a larger variance (\textcolor{red}{in red}).}
+For instance, \textcolor{lightgreen}{Thomson sampling} is very efficient in average (\textcolor{lightgreen}{in green}), and \textcolor{red}{\UCB} shows a larger variance (\textcolor{red}{in red}).}
 	\label{fig:3:firstPlot_hist}

### 2019-07-11 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Petit changements sur les légendes

@@
               \For{$t = 1, \dots, T$}{
-                   \uIf(\tcp*[f]{transition $(3)$ or $(5)$}){
+                   \uIf(\tcp*[f]{\textcolor{red}{transition $(3)$ or $(5)$}}){
                       $A^j(t-1) \notin \TopM(t)$}
@@
                         $A^j(t) \sim \cU \left(\TopM(t)\right)$
-                        \tcp*[f]{transition $(2)$} \\
+                        \tcp*[f]{\textcolor{blue}{transition $(2)$}} \\
                         $s^j(t) = \mathrm{False}$
                     }
-                    \uElse(\tcp*[f]{transition $(1)$ or $(4)$}){
+                    \uElse(\tcp*[f]{transition \textcolor{cyan}{$(1)$} or \textcolor{darkgreen}{$(4)$}}){
                       $A^j(t) = A^j(t-1)$ \tcp*[f]{stay on the previous arm} \\
@@
   \caption[Regret for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on a fixed problem.]{Regret (in log-log scale), for $M=6$ players for $K=9$ arms, horizon $T=5000$, for $1000$ repetitions on problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-\textcolor{gold}{\RandTopM{} (in yellow)} outperforms \textcolor{darkgreen}{\Selfish{} (in green)}, both clearly outperform \rhoRand.
+\textcolor{lightgreen}{\RandTopM{} (in light green)} outperforms \textcolor{darkgreen}{\Selfish{} (in green)}, both clearly outperform \rhoRand.
 The regret of \MCTopM{} is logarithmic, empirically with the same slope as the lower bound.
@@
   \caption[Regret for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for a fixed problem.]{Regret (in log-log scale), for $M=2$ and $9$ players for $K=9$ arms, horizon $T=5000$, for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$] for problem $\boldsymbol{\mu}=[0.1,\dots,0.9]$.
-In different settings, \textcolor{gold}{\RandTopM{} (in yellow)} and \textcolor{darkgreen}{\Selfish{} (in green)} can outperform each other, and always outperform \rhoRand.
+In different settings, \textcolor{lightgreen}{\RandTopM{} (in light green)} and \textcolor{darkgreen}{\Selfish{} (in green)} can outperform each other, and always outperform \rhoRand.
 \MCTopM{} is always among the best algorithms, and for $M$ not too small, its regret seems logarithmic with a constant matching the lower bound.}
@@
 The third instance uses the optimal $T_0$ corresponding to $\delta=1/T$, that is guaranteed to have an expected regret of order $\log(T)$.
-The log-$y$ scale is used to easily differentiate between the different algorithms, and highlight that our proposal (in \textcolor{gold}{yellow $\nabla$}) outperform both \MEGA{} and \MusicalChair{} by three orders of magnitudes!}
+The log-$y$ scale is used to easily differentiate between the different algorithms, and highlight that our proposal (in \textcolor{lightgreen}{light green $\nabla$}) outperform both \MEGA{} and \MusicalChair{} by three orders of magnitudes!}
   \label{fig:5:MP__K9_M3_T123456_N100__8_algos}
@@
 \textcolor{blue}{\rhoRand{} (top blue)} is outperformed by the other algorithms (and the gain increases with $M$).
-\textcolor{gold}{\MCTopM{} (bottom yellow)} outperforms all the other algorithms in most cases.}
+\textcolor{lightgreen}{\MCTopM{} (bottom green)} outperforms all the other algorithms in most cases.}
   % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
@@
   \caption[Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=K=9$, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$.
-This extreme case $M=K$ shows the drastic difference of behavior between \textcolor{red}{\RandTopM{} (red)} and \textcolor{gold}{\MCTopM{} (yellow)}, with constant regret, and \textcolor{blue}{\rhoRand{} (blue)} and \textcolor{darkgreen}{\Selfish{} (green)}, with large regret.}
+This extreme case $M=K$ shows the drastic difference of behavior between \textcolor{red}{\RandTopM{} (red)} and \textcolor{lightgreen}{\MCTopM{} (light green)}, with constant regret, and \textcolor{blue}{\rhoRand{} (blue)} and \textcolor{darkgreen}{\Selfish{} (green)}, with large regret.}
   \label{fig:5:MP__K9_M9_T5000_N500__4_algos__all_HistogramsRegret}

### 2019-07-11 2-Chapters/5-Chapter/chapter5.tex
Petit changements sur les légendes

@@
 The first idea comes from \cite{Anandkumar11}, and the intuition behind it is quite simple, even if the mathematical derivations are not.
-All players will build an estimate $\hat{M}^j(t)$ of the number of player, that start by $\hat{M}^j(0)=1$. As soon as one collision is observed, a player knows that $M\geq2$, so $\hat{M}^j(t+1) = 2$.
-Then, based on probabilistic computations on the expected number of collisions if there were $m$ players, all following the same strategy (\eg, \RhoRand{} in the case of ), and because the formula is simple to compute for different $m$, the authors proposed a statistical test of the hypothesis $M \leq \hat{M}^j(t)$ against $M > \hat{M}^j(t)$ which is expressed as a simple comparison of the current number of collisions (since last update of $\hat{M}^j(t)$) against a threshold.
-If a player observed ``too many'' collisions (that are unlikely to be caused by only $\hat{M}^j(t) - 1$ other players), then she increases her current estimate $\hat{M}^j(t+1) = \hat{M}^j(t) + 1$.
+All players will build an estimate $\widehat{M}^j(t)$ of the number of player, that start by $\widehat{M}^j(0)=1$.
+As soon as one collision is observed, a player knows that $M\geq2$, so $\widehat{M}^j(t+1) = 2$.
+Then, based on probabilistic computations on the expected number of collisions if there were $m$ players, all following the same strategy (\eg, \RhoRand{} in the case of ), and because the formula is simple to compute for different $m$, the authors proposed a statistical test of the hypothesis $M \leq \widehat{M}^j(t)$ against $M > \widehat{M}^j(t)$ which is expressed as a simple comparison of the current number of collisions (since last update of $\widehat{M}^j(t)$) against a threshold.
+If a player observed ``too many'' collisions (that are unlikely to be caused by only $\widehat{M}^j(t) - 1$ other players), then she increases her current estimate $\widehat{M}^j(t+1) = \widehat{M}^j(t) + 1$.
 

### 2019-07-11 2-Chapters/6-Chapter/chapter6.tex
Merci Christophe de toutes ces remarques !
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion
Petit changements sur les légendes

@@
 %
-In this last chapter, we are also interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
+In this last chapter, we are also interested in a more formal approach to the decentralized learning problem presented in Chapter~\ref{chapter:4}.
 Instead of keeping the stationary hypothesis but considering $2 \leq M \leq K$ devices accessing a wireless networks with $K$ orthogonal frequency channels, as we did in Chapter~\ref{chapter:5}, we are now interested in another direction of formal analysis of the intractable models of IoT networks of Chapter~\ref{chapter:4}.
@@
 bounded or sub-Bernoulli distributions.
-Our test enjoys finite-time guarantees,
+We are able to prove finite-time guarantees for our test,
 % its \emph{false alarm probability}, meaning that in a stationary model, the test should not detect any change,
@@
 on its \emph{false alarm probability} and \emph{detection delay}, and can be combined with an efficient bandit policy (\klUCB), to propose an efficient algorithm for piece-wise stationary problems, \GLRklUCB.
-% Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known before hand).
+% Our policy takes ideas from the two most recent and most efficient policies, \CUSUMUCB{} and \MUCB, that enjoys good regret bounds of the order of $R_T = \bigO{\sqrt{\Upsilon_T T \log(T)}}$, if $T$ is the horizon and $\Upsilon_T$ the number of change-points (known beforehand).
 % % \MUCB{} restarts the underlying \UCB{} bandit policy on all arms when a change is detected, while \CUSUMUCB{} restarts the observations of only the arm on which a change is detected.
@@
 % and yield regret bounds with explicit constants.
-% Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$, when the algorithm runs with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know before hand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
+% Moreover, our regret bounds are the first ones to be of the order $\bigO{\sqrt{\Upsilon_T T \log(T)}}$, when the algorithm runs with no prior knowledge of the problem other than the horizon $T$ and the number of change-points $\Upsilon_T$ (both \MUCB{} and \CUSUMUCB{} require to know beforehand a certain measure $\Delta^{\text{change}}$ of the difficulty of the problem).
 %

### 2019-07-11 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Petit changements sur les légendes

@@
         R_T \;\; \leq \;\; & 2\sum_{i=1}^{\Upsilon_T} \frac{4K}{\omega \left(\Delta^{(i)}\right)^2}\beta(T,\delta) +\omega T + \delta (K+1)\Upsilon_T \\
-        & + \sum_{k=1}^{K}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \textcolor{red}{\bigO{\sqrt{\ln(T)}}}.
+        & + \sum_{k=1}^{K}\sum_{\substack{i = 1,\dots,Y_T\\\mu_k^{(i)} \neq \mu_{k^*}^{(i)}}} \frac{\left( \mu_{k^*}^{(i)}-\mu_k^{(i)}\right)}{\kl\left(\mu_k^{(i)},\mu_{k^*}^{(i)}\right)}\ln(T) + \textcolor{red}{\underline{\cO}\left(\sqrt{\ln(T)}\right)}.
 \nonumber
@@
 %
-We highlight that \emph{this result is finite-time} and not asymptotic, but if it uses the notation $\textcolor{red}{\bigO{\sqrt{\ln(T)}}}$ for simplicity.
-This last term $\bigO{\dots}$ does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a certain $c_0 \times \sqrt{\ln(T)}$, from a certain time $T_0$ (\ie, $\forall T \geq T_0$).
+We highlight that \emph{this result is finite-time} and not asymptotic, but if it uses the notation $\textcolor{red}{\underline{\cO}\left(\sqrt{\ln(T)}\right)}$ for simplicity.
+This last term $\underline{\cO}\left(\dots)$ does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a certain $c_0 \times \sqrt{\ln(T)}$, from a certain time $T_0$ (\ie, $\forall T \geq T_0$).
 It is also the case of the next Theorem~\ref{thm:6:mainRegretBound}.
@@
     \begin{equation}
-        R_T \leq 2\sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{4K}{\omega \left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) +\omega T + 2 \delta \EffChange  + \sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{\ln(T)}{\kl(\overline{\mu}_k^{(\ell)},{\mu}^*_{i,\ell})} + \textcolor{red}{\bigO{\sqrt{\ln(T)}}},
+        R_T \leq 2\sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{4K}{\omega \left(\Delta_k^{(\ell)}\right)^2}\beta(T,\delta) +\omega T + 2 \delta \EffChange  + \sum_{k=1}^K\sum_{\ell=1}^{\NCk} \frac{\ln(T)}{\kl(\overline{\mu}_k^{(\ell)},{\mu}^*_{i,\ell})} + \textcolor{red}{\underline{\cO}\left(\sqrt{\ln(T)}\right)},
     \end{equation}
@@
 where the last but one inequality relies on the concentration inequality given in Lemma~2 of \cite{KLUCBJournal}, for the choice $f(t) = \ln(t) + 3 \ln(\ln(t))$.
-The last inequality comes from a sum-integral comparison, as $\int\limits_{1}^T \frac{1}{t \ln(t)} \mathrm{d}t = \ln(\ln(T))$.
+The last inequality comes from a sum-integral comparison, as $\int_{1}^T \frac{1}{t \ln(t)} \mathrm{d}t = \ln(\ln(T))$.
 Finally, the law of total expectation gives

### 2019-07-11 2-Chapters/3-Chapter/src/example_of_a_IndexPolicy_UCB.py

@@ -11,8 +11,8 @@ class UCB(IndexPolicy):
 
         .. math:: U_k(t) = \frac{X_k(t)}{N_k(t)} + \sqrt{\frac{2 \log(t)}{N_k(t)}}.
         """
-        if self.pulls[arm] < 1:    # forced exploration
-            return float('+inf')     # in the first steps
+        if self.pulls[arm] < 1:    # forced exploration in the first steps
+            return float('+inf')
         else:                      # or compute UCB index
             estimated_mean = self.rewards[arm] / self.pulls[arm]
             exploration_bias = sqrt((2 * log(self.t)) / self.pulls[arm])

### 2019-07-11 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Typo

@@
 We highlight that \emph{this result is finite-time} and not asymptotic, but if it uses the notation $\textcolor{red}{\underline{\cO}\left(\sqrt{\ln(T)}\right)}$ for simplicity.
-This last term $\underline{\cO}\left(\dots)$ does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a certain $c_0 \times \sqrt{\ln(T)}$, from a certain time $T_0$ (\ie, $\forall T \geq T_0$).
+This last term $\underline{\cO}\left(\dots\right)$ does not mean the whole inequality is only valid for $T\to\infty$, but it rather means that at finite time, the inequality is valid with the last term being a function $g(T)$, which is upper-bounded by a certain $c_0 \times \sqrt{\ln(T)}$, from a certain time $T_0$ (\ie, $\forall T \geq T_0$).
 It is also the case of the next Theorem~\ref{thm:6:mainRegretBound}.

### 2019-07-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
OH LA BOULETTE : en anglais, « repartition » n'existe pas. On dit « distribution » ou « allocation »
- https://www.wordreference.com/enfr/repartition
- https://www.deepl.com/translator#en/fr/repartition

@@
 
-We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the repartition of static devices $(S_k)_k$, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
+We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the distribution of static devices $(S_k)_k$, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead.
 This is not possible in IoT contexts as several completely independent networks may operate in a single place.} is possible.
-We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.
+We want to find the stationary distribution of devices into channels that maximizes the probability of having a successful transmission.
 
@@
 Note that solving the optimization problem provides the optimal \emph{real values} $(D_k^*)_k$, which have to be rounded to find the optimal \emph{integer} number of devices for channel $k$.
-Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
+Any rounding choice will give about the same distribution, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
 %
@@
 The RF band is divided in $K = 10$ channels.
-Each static device only uses one channel, and their uneven repartition in the $10$ channels is chosen as $(S_1,\cdots, S_{K}) = S \times (0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
+Each static device only uses one channel, and their uneven distribution in the $10$ channels is chosen as $(S_1,\cdots, S_{K}) = S \times (0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases.
 The dynamic devices have access to all the channels, and use learning algorithms.
@@
 We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
-We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions.
+We tried similar experiments with other values for $K$ and this distribution vector, and results were similar for non-homogeneous repartitions.
 
-Clearly, the problem is less interesting for a homogeneous repartition, as all channels are equivalent for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
+Clearly, the problem is less interesting for a homogeneous distribution, as all channels are equivalent for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.
 
@@
 The learning gain, for each device, decreases with the proportion of dynamic devices in the network.
-Note that the values ($16\%$ etc) depend on the repartition of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.}
+Note that the values ($16\%$ etc) depend on the distribution of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.}
     \label{fig:41:perf_learning}
@@
     \label{fig:41:ch2_100}}
-    \caption[Performance of the \UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.]{Performance of the \textcolor{blue}{\UCB} bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
+    \caption[Performance of the \UCB bandit algorithm for the special case of uniform distribution of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.]{Performance of the \textcolor{blue}{\UCB} bandit algorithm for the special case of uniform distribution of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
     \label{fig:41:figure4appendix}
@@
 %
-We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall k, S_k = S/K$), to check that learning approach (here, only \UCB)
+We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform distribution of static devices (\ie, $\forall k, S_k = S/K$), to check that learning approach (here, only \UCB)
 also gives interesting gain of performance, and achieve a close-to optimal successful transmission rate.

### 2019-07-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for now
OH LA BOULETTE : en anglais, « repartition » n'existe pas. On dit « distribution » ou « allocation »
- https://www.wordreference.com/enfr/repartition
- https://www.deepl.com/translator#en/fr/repartition

@@
 
-In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform repartition of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
+In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform distribution of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
 In other words, the channels are occupied\footnote{~Note that we consider higher occupancy rates that the ones considered previously in Section~\ref{sec:4:gnuradio}, because the goal of this experimental section is to evaluate our approach that uses learning to optimize the way each device retransmits its packets.
@@
 	\centering
-	\includegraphics[width=0.80\linewidth]{ResultsUCB.eps}
+	\includegraphics[width=0.90\linewidth]{ResultsUCB.eps}
 	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{
@@
 	\centering
-	\includegraphics[width=0.80\linewidth]{ResultsUCB2.eps}
+	\includegraphics[width=0.90\linewidth]{ResultsUCB2.eps}
 	\caption[Second comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random'', ``\UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{

### 2019-08-30 2-Chapters/Conclusion/conclusion.tex
Petit changements sur les légendes
Typo

@@
 (see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
-From the point-of-view of the theoretical analysis, we are especially interesting about first extending our proposed algorithm \MCTopM{} to the case of an unknown number of players.
+From the point-of-view of the theoretical analysis, we are especially interested about first extending our proposed algorithm \MCTopM{} to the case of an unknown number of players.
 Then we could also address the most interesting extension for the application to wireless networks, \ie, the ``dynamic case'' which allows for arrivals and departures of players in the multi-player bandit game.
@@
 \begin{quote}
-    \emph{-- Les rêves, ça se compare pas\ldots}
+    \emph{-- Les rêves, ça se compare pas\ldots}\\
     Alexandre Astier, \emph{Kaamelott}, Livre VI, Épisode 9, ``Dies Irae''.

### 2019-09-09 1-Preamble/4-Acknowledgement/acknowledgement.tex
Une typo pendant une relecture occasionnelle.

@@
 %
-I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been of a constant support during all my life and in my studies.
+I first want to thank my \href{https://paris-sorbonne.academia.edu/FBesson}{older brother Florian}, who has always been my main model and who has been a constant support during all my life and in my studies.
 His rigour, dedication and seriousness kept pushing me further, and I am immensely proud of him.

### 2019-09-12 2-Chapters/2-Chapter/agregation.tex
Removed raw inclusion of https://hal.inria.fr/hal-01705292

--- 2019-07-10 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
@@
 	\centering
-	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env1-4_932221613383548446.pdf}
-	\caption{On a ``simple'' Bernoulli problem (semilog-$y$ scale). \Aggr{} is in \textbf{bold \textcolor{red}{red}}.}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/Images/plots-aggregation/main_semilogy____env1-4_932221613383548446.pdf}
+	\caption[On a ``simple'' Bernoulli problem (semilog-$y$ scale). \Aggr{} is in bold red.]{On a ``simple'' Bernoulli problem (semilog-$y$ scale). \Aggr{} is in \textbf{bold \textcolor{red}{red}}.}
 	\label{fig:25:EasyBernoulli}
@@
 	\centering
-	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env2-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/Images/plots-aggregation/main____env2-4_932221613383548446.pdf}
 	\caption{On a ``harder'' Bernoulli problem, they all have similar performances, except \LearnExp, and our proposal \Aggr{} outperforms its competitors.}
@@
 	\centering
-	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main____env3-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/Images/plots-aggregation/main____env3-4_932221613383548446.pdf}
 	\caption{On an ``easy'' Gaussian problem, only \Aggr{} shows reasonable performances, thanks to Bayes-UCB and Thompson sampling.}
@@
 	\centering
-	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogy____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/Images/plots-aggregation/main_semilogy____env4-4_932221613383548446.pdf}
 	\caption{On a harder problem, mixing Bernoulli, Gaussian, Exponential arms, with 3 arms of each type with the \emph{same mean}.}
@@
 	\centering
-	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/plots/main_semilogx____env4-4_932221613383548446.pdf}
+	\includegraphics[width=1.09\linewidth]{2-Chapters/2-Chapter/Images/plots-aggregation/main_semilogx____env4-4_932221613383548446.pdf}
 	\caption{The semilog-$x$ scale clearly shows the logarithmic growth of the regret for the best algorithms and our proposal \Aggr, even in a hard ``mixed'' problem (\emph{cf}.

### 2019-09-12 2-Chapters/4-Chapter/CrownCom_17/IoT_slotted_appendix.tex
Removed raw inclusion of https://hal.inria.fr/hal-02006825 https://hal.inria.fr/hal-02049824v2 https://hal.archives-ouvertes.fr/hal-01575419v2

--- 2019-07-04 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
@@
 %
-Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($K, p$), the repartition of static devices ($S_1,\dots,S_{K}$) and the Lagrange multiplier $\lambda$.
+Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($K, p$), the distribution of static devices ($S_1,\dots,S_{K}$) and the Lagrange multiplier $\lambda$.
 

### 2019-09-12 2-Chapters/4-Chapter/chapter4.tex
Almost done proofreading résumé fr
OH LA BOULETTE : en anglais, « repartition » n'existe pas. On dit « distribution » ou « allocation »
- https://www.wordreference.com/enfr/repartition
- https://www.deepl.com/translator#en/fr/repartition
Removed raw inclusion of https://hal.inria.fr/hal-02006825 https://hal.inria.fr/hal-02049824v2 https://hal.archives-ouvertes.fr/hal-01575419v2

@@
 by using MAB algorithms like \UCB.
-% in the case of a non-uniform repartition of the static devices in the channels
+% in the case of a non-uniform distribution of the static devices in the channels
 We also developed a proof-of-concept using USRP platforms, for a real-world validation of this approach.
@@
 unlicensed bands are more and more used and considered for mobile and LAN (Local Area Network) and for Internet of Things communication standards.
-%  (WiFi, LTE-U), (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
+%  (Wi-Fi, LTE-U), (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
 This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.
@@
 % Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
-% Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
+% Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) Wi-Fi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
 % % However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.
@@
 
-\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex}
+\input{2-Chapters/4-Chapter/CrownCom_17/IoT_slotted.tex}
 
@@
 
-\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex}
+\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT/IEEE_WCNC_2019__DemoICT.tex}
 
@@
 
-\input{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex}
+\input{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM/IEEE_WCNC__2019__Paper__BMBM.tex}
 
@@
 
-\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex}
+\input{2-Chapters/4-Chapter/CrownCom_17/IoT_slotted_appendix.tex}
 
@@
 
-\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex}
+\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT/IEEE_WCNC_2019__DemoICT_appendix.tex}

### 2019-09-12 2-Chapters/5-Chapter/ALT_2018__MPBandits/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__leafs__onlyfig.gv.tex
Removed raw inclusion of https://hal.inria.fr/hal-01629733v2

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (302.0bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (0) at (247.1bp,128.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (19) at (134.1bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (68) at (361.1bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \draw [black,->] (0) ..controls (159.85bp,113.55bp) and (145.0bp,105.2bp)  .. (135.1bp,92.0bp) .. controls (130.31bp,85.627bp) and (128.83bp,77.397bp)  .. (19);
  \draw (237.1bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1} - 1)^{2}/2$};
  \draw [black,->] (0) ..controls (315.21bp,109.06bp) and (328.69bp,101.89bp)  .. (339.1bp,92.0bp) .. controls (345.72bp,85.708bp) and (350.47bp,77.073bp)  .. (68);
  \draw (451.1bp,84.5bp) node {$\mu_{1}^{2}(\mu_{2} - 1)^{2}/2$};
%
\end{tikzpicture}

### 2019-09-12 2-Chapters/5-Chapter/ALT_2018__MPBandits/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__absorbing__onlyfig.gv.tex
Removed raw inclusion of https://hal.inria.fr/hal-01629733v2

--- 2019-03-25 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/Tree_exploration_K2_M2_depth3__Selfish_UCB__absorbing__onlyfig.gv.tex
@@
-\begin{tikzpicture}[>=latex,line join=bevel,scale=0.5]
+\begin{tikzpicture}[>=latex,line join=bevel,]
 %%
-  \node (155) at (451.19bp,41.0bp) [draw=red,rectangle,thick] {[[1/2,0/1], [1/2,0/1]]};
-  \node (10) at (635.19bp,215.0bp) [draw,rectangle] {[[0/0,1/1], [0/1,0/0]]};
-  \node (3) at (267.19bp,215.0bp) [draw,rectangle] {[[0/0,0/1], [1/1,0/0]]};
-  \node (6) at (451.19bp,215.0bp) [draw,rectangle] {[[1/1,0/0], [0/0,0/1]]};
-  \node (14) at (83.193bp,128.0bp) [draw=red,rectangle] {[[0/1,1/1], [0/1,1/1]]};
-  \node (1) at (83.193bp,215.0bp) [draw,rectangle] {[[0/1,0/0], [0/0,1/1]]};
-  \node (0) at (359.19bp,302.0bp) [draw=darkgreen,rectangle] {[[0/0,0/0], [0/0,0/0]]};
-  \node (22) at (267.19bp,128.0bp) [draw=red,rectangle] {[[1/1,0/1], [1/1,0/1]]};
-  \node (234) at (635.19bp,41.0bp) [draw=red,rectangle,thick] {[[0/1,1/2], [0/1,1/2]]};
-  \node (65) at (83.193bp,41.0bp) [draw=red,rectangle,thick] {[[0/1,1/2], [0/1,1/2]]};
-  \node (110) at (267.19bp,41.0bp) [draw=red,rectangle,thick] {[[1/2,0/1], [1/2,0/1]]};
-  \node (43) at (635.19bp,128.0bp) [draw=red,rectangle] {[[0/1,1/1], [0/1,1/1]]};
-  \node (30) at (451.19bp,128.0bp) [draw=red,rectangle] {[[1/1,0/1], [1/1,0/1]]};
+\begin{scope}
+  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
+  \pgfsetstrokecolor{strokecol}
+  \draw (383.6bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
+\end{scope}
+  \node (155) at (451.19bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
+  \node (10) at (635.19bp,215.0bp) [draw,ellipse] {[[0/0,1/1], [0/1,0/0]]};
+  \node (3) at (267.19bp,215.0bp) [draw,ellipse] {[[0/0,0/1], [1/1,0/0]]};
+  \node (6) at (451.19bp,215.0bp) [draw,ellipse] {[[1/1,0/0], [0/0,0/1]]};
+  \node (14) at (83.193bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
+  \node (1) at (83.193bp,215.0bp) [draw,ellipse] {[[0/1,0/0], [0/0,1/1]]};
+  \node (0) at (359.19bp,302.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
+  \node (22) at (267.19bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
+  \node (234) at (635.19bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
+  \node (65) at (83.193bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
+  \node (110) at (267.19bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
+  \node (43) at (635.19bp,128.0bp) [draw=red,ellipse] {[[0/1,1/1], [0/1,1/1]]};
+  \node (30) at (451.19bp,128.0bp) [draw=red,ellipse] {[[1/1,0/1], [1/1,0/1]]};
   \draw [red,->] (43) ..controls (635.19bp,98.163bp) and (635.19bp,82.548bp)  .. (234);

### 2019-09-12 2-Chapters/5-Chapter/ALT_2018__MPBandits/figures/Tree_exploration_K=2_M=2_depth=3__Selfish_KLUCB__leafs__onlyfig.gv.tex
Removed raw inclusion of https://hal.inria.fr/hal-01629733v2

\begin{tikzpicture}[>=latex,line join=bevel,]
%%
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (13917.0bp,11.5bp) node {Tree exploration for K=2 arms and M=2 players using Selfish KLUCB, for depth=3 : 68 leafs, 2 absorbing};
\end{scope}
  \node (17) at (6572.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,1/2]]};
  \node (39) at (16256.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,0/1]]};
  \node (30) at (12781.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,2/2]]};
  \node (13) at (4517.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,1/1]]};
  \node (56) at (24250.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/2,0/1]]};
  \node (7) at (2786.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/2,1/1]]};
  \node (47) at (19388.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/2,0/1]]};
  \node (25) at (11345.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,1/1]]};
  \node (55) at (23905.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [0/2,0/1]]};
  \node (60) at (25419.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,0/1]]};
  \node (36) at (15327.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,1/2]]};
  \node (40) at (16604.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,2/2]]};
  \node (5) at (1344.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,0/1]]};
  \node (65) at (26747.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,2/2]]};
  \node (43) at (18325.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,1/1]]};
  \node (2) at (351.19bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [0/2,0/1]]};
  \node (62) at (25894.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/1,2/2]]};
  \node (10) at (3643.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,1/2]]};
  \node (3) at (642.19bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,0/1]]};
  \node (67) at (27303.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,2/2]]};
  \node (19) at (7257.2bp,41.0bp) [draw=red,ellipse] {[[0/1,1/2], [0/1,1/2]]};
  \node (22) at (8945.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/1,2/2]]};
  \node (8) at (3027.2bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [1/2,1/1]]};
  \node (1) at (83.193bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,1/1]]};
  \node (20) at (7484.2bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,1/2]]};
  \node (59) at (25068.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,1/1]]};
  \node (35) at (14990.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,2/2]]};
  \node (0) at (14855.0bp,128.0bp) [draw=green,ellipse] {[[0/0,0/0], [0/0,0/0]]};
  \node (29) at (12437.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,1/2]]};
  \node (16) at (6230.2bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [0/1,0/2]]};
  \node (26) at (11613.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,0/1]]};
  \node (51) at (21910.0bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [1/1,2/2]]};
  \node (49) at (21443.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,0/1]]};
  \node (34) at (14721.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/1,0/2]]};
  \node (44) at (18608.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,0/1]]};
  \node (21) at (7832.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,1/2]]};
  \node (64) at (26506.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,1/2]]};
  \node (9) at (3303.2bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,0/2]]};
  \node (24) at (10390.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,0/2]]};
  \node (52) at (22254.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/2,0/1]]};
  \node (18) at (6909.2bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [2/2,0/1]]};
  \node (31) at (13057.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/2,0/1]]};
  \node (63) at (26162.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [1/2,0/1]]};
  \node (46) at (19048.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/2,0/1]]};
  \node (11) at (3983.2bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [2/2,1/1]]};
  \node (12) at (4251.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [1/2,0/1]]};
  \node (66) at (27015.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [0/1,0/2]]};
  \node (38) at (15916.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/2,1/1]]};
  \node (54) at (23637.0bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,2/2]]};
  \node (33) at (14438.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [1/1,1/2]]};
  \node (57) at (24516.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [0/1,0/2]]};
  \node (32) at (14170.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,1/2]]};
  \node (15) at (5954.2bp,41.0bp) [draw,ellipse] {[[2/2,0/1], [1/1,1/2]]};
  \node (42) at (17988.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/2,1/1]]};
  \node (27) at (11901.0bp,41.0bp) [draw,ellipse] {[[1/1,1/2], [1/1,1/2]]};
  \node (28) at (12100.0bp,41.0bp) [draw,ellipse] {[[0/1,2/2], [2/2,0/1]]};
  \node (41) at (16895.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/2,0/1]]};
  \node (6) at (2444.2bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [2/2,1/1]]};
  \node (68) at (27654.0bp,41.0bp) [draw=red,ellipse] {[[1/2,0/1], [1/2,0/1]]};
  \node (23) at (9290.2bp,41.0bp) [draw,ellipse] {[[1/2,0/1], [0/1,0/2]]};
  \node (58) at (24799.0bp,41.0bp) [draw,ellipse] {[[0/1,0/2], [0/1,1/2]]};
  \node (4) at (993.19bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [0/1,1/2]]};
  \node (53) at (22537.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [0/2,0/1]]};
  \node (37) at (15675.0bp,41.0bp) [draw,ellipse] {[[0/1,1/2], [1/1,1/2]]};
  \node (61) at (25695.0bp,41.0bp) [draw,ellipse] {[[2/2,1/1], [1/1,2/2]]};
  \node (14) at (4861.2bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,0/2]]};
  \node (48) at (20343.0bp,41.0bp) [draw,ellipse] {[[0/2,0/1], [0/1,1/2]]};
  \node (50) at (21711.0bp,41.0bp) [draw,ellipse] {[[1/2,1/1], [1/2,1/1]]};
  \node (45) at (18849.0bp,41.0bp) [draw,ellipse] {[[1/1,2/2], [2/2,1/1]]};
  \draw [black,->] (0) ..controls (15310.0bp,123.9bp) and (16807.0bp,112.39bp)  .. (16851.0bp,92.0bp) .. controls (16863.0bp,86.743bp) and (16872.0bp,76.987bp)  .. (41);
  \draw (17406.0bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (13590.0bp,127.68bp) and (212.52bp,133.93bp)  .. (124.19bp,92.0bp) .. controls (113.32bp,86.84bp) and (104.37bp,77.236bp)  .. (1);
  \draw (246.69bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (14395.0bp,124.58bp) and (12861.0bp,115.02bp)  .. (12817.0bp,92.0bp) .. controls (12807.0bp,86.74bp) and (12799.0bp,77.252bp)  .. (30);
  \draw (12944.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{3}(\mu_{1} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (13931.0bp,125.57bp) and (6996.8bp,114.16bp)  .. (6950.2bp,92.0bp) .. controls (6939.3bp,86.829bp) and (6930.4bp,77.223bp)  .. (18);
  \draw (7112.7bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (15861.0bp,126.07bp) and (24152.0bp,117.69bp)  .. (24207.0bp,92.0bp) .. controls (24218.0bp,86.86bp) and (24228.0bp,77.259bp)  .. (56);
  \draw (24348.0bp,84.5bp) node {$-\mu_{1}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (14086.0bp,126.55bp) and (9395.4bp,122.62bp)  .. (9331.2bp,92.0bp) .. controls (9320.3bp,86.819bp) and (9311.4bp,77.211bp)  .. (23);
  \draw (9869.7bp,84.5bp) node {$\mu_{1}(1-\mu_{2})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (14322.0bp,126.17bp) and (12202.0bp,121.22bp)  .. (12141.0bp,92.0bp) .. controls (12130.0bp,86.783bp) and (12121.0bp,77.169bp)  .. (28);
  \draw (12298.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (0) ..controls (13739.0bp,126.85bp) and (3412.8bp,124.59bp)  .. (3344.2bp,92.0bp) .. controls (3333.3bp,86.837bp) and (3324.4bp,77.232bp)  .. (9);
  \draw (3502.7bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13876.0bp,125.96bp) and (6047.6bp,116.91bp)  .. (5995.2bp,92.0bp) .. controls (5984.3bp,86.832bp) and (5975.4bp,77.226bp)  .. (15);
  \draw (6121.7bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}(\mu_{2} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (16003.0bp,126.94bp) and (26899.0bp,125.49bp)  .. (26972.0bp,92.0bp) .. controls (26983.0bp,86.865bp) and (26993.0bp,77.265bp)  .. (66);
  \draw (27124.0bp,84.5bp) node {$\mu_{1}^{2}(1-\mu_{1})(1-\mu_{2})^{3}/4$};
  \draw [black,->] (0) ..controls (15693.0bp,127.24bp) and (21323.0bp,127.58bp)  .. (21400.0bp,92.0bp) .. controls (21411.0bp,86.851bp) and (21421.0bp,77.249bp)  .. (49);
  \draw (21542.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15547.0bp,125.46bp) and (19293.0bp,116.1bp)  .. (19345.0bp,92.0bp) .. controls (19356.0bp,86.838bp) and (19366.0bp,77.233bp)  .. (47);
  \draw (19830.0bp,84.5bp) node {$(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (13695.0bp,127.09bp) and (2559.0bp,127.07bp)  .. (2485.2bp,92.0bp) .. controls (2474.3bp,86.838bp) and (2465.4bp,77.233bp)  .. (6);
  \draw (2644.7bp,84.5bp) node {$-\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13968.0bp,127.9bp) and (7611.3bp,133.0bp)  .. (7525.2bp,92.0bp) .. controls (7514.3bp,86.827bp) and (7505.4bp,77.221bp)  .. (20);
  \draw (7687.7bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (14428.0bp,123.92bp) and (13125.0bp,113.08bp)  .. (13088.0bp,92.0bp) .. controls (13079.0bp,86.747bp) and (13072.0bp,77.521bp)  .. (31);
  \draw (13633.0bp,84.5bp) node {$\mu_{1}\mu_{2}(9\mu_{1}^{2}\mu_{2}^{2} - 18\mu_{1}^{2}\mu_{2} + 9\mu_{1}^{2} - 18\mu_{1}\mu_{2}^{2} + 36\mu_{1}\mu_{2} - 18\mu_{1} + 9\mu_{2}^{2} - 18\mu_{2} + 10)/8$};
  \draw [black,->] (0) ..controls (13615.0bp,127.55bp) and (768.1bp,132.31bp)  .. (683.19bp,92.0bp) .. controls (672.32bp,86.84bp) and (663.37bp,77.235bp)  .. (3);
  \draw (847.19bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15978.0bp,126.79bp) and (26394.0bp,124.05bp)  .. (26463.0bp,92.0bp) .. controls (26474.0bp,86.864bp) and (26484.0bp,77.264bp)  .. (64);
  \draw (26591.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (14167.0bp,125.48bp) and (10482.0bp,116.33bp)  .. (10431.0bp,92.0bp) .. controls (10420.0bp,86.81bp) and (10411.0bp,77.201bp)  .. (24);
  \draw (10897.0bp,84.5bp) node {$(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (15425.0bp,126.88bp) and (17875.0bp,124.64bp)  .. (17945.0bp,92.0bp) .. controls (17956.0bp,86.818bp) and (17966.0bp,77.21bp)  .. (42);
  \draw (18121.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13955.0bp,128.05bp) and (7387.1bp,134.31bp)  .. (7298.2bp,92.0bp) .. controls (7287.3bp,86.828bp) and (7278.4bp,77.222bp)  .. (19);
  \draw (7400.2bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1} - 1)^{2}/2$};
  \draw [black,->] (0) ..controls (15710.0bp,127.42bp) and (21588.0bp,129.08bp)  .. (21668.0bp,92.0bp) .. controls (21679.0bp,86.853bp) and (21689.0bp,77.25bp)  .. (50);
  \draw (21775.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}/8$};
  \draw [black,->] (0) ..controls (13912.0bp,125.72bp) and (6661.8bp,115.13bp)  .. (6613.2bp,92.0bp) .. controls (6602.3bp,86.83bp) and (6593.4bp,77.224bp)  .. (17);
  \draw (6770.2bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15988.0bp,126.86bp) and (26633.0bp,124.73bp)  .. (26704.0bp,92.0bp) .. controls (26715.0bp,86.865bp) and (26725.0bp,77.264bp)  .. (65);
  \draw (26846.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (13599.0bp,127.62bp) and (478.87bp,133.15bp)  .. (392.19bp,92.0bp) .. controls (381.32bp,86.84bp) and (372.37bp,77.235bp)  .. (2);
  \draw (526.19bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1}  1)^{3}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15743.0bp,127.79bp) and (22124.0bp,132.13bp)  .. (22211.0bp,92.0bp) .. controls (22222.0bp,86.855bp) and (22232.0bp,77.253bp)  .. (52);
  \draw (22360.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13990.0bp,127.67bp) and (7955.1bp,131.0bp)  .. (7873.2bp,92.0bp) .. controls (7862.3bp,86.826bp) and (7853.4bp,77.219bp)  .. (21);
  \draw (8418.2bp,84.5bp) node {$\mu_{1}\mu_{2}(9\mu_{1}^{2}\mu_{2}^{2} - 18\mu_{1}^{2}\mu_{2} + 9\mu_{1}^{2} - 18\mu_{1}\mu_{2}^{2} + 36\mu_{1}\mu_{2} - 18\mu_{1} + 9\mu_{2}^{2} - 18\mu_{2} + 10)/8$};
  \draw [black,->] (0) ..controls (15827.0bp,125.84bp) and (23542.0bp,115.96bp)  .. (23594.0bp,92.0bp) .. controls (23605.0bp,86.859bp) and (23615.0bp,77.257bp)  .. (54);
  \draw (23736.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (13630.0bp,127.46bp) and (1116.9bp,131.28bp)  .. (1034.2bp,92.0bp) .. controls (1023.3bp,86.839bp) and (1014.4bp,77.235bp)  .. (4);
  \draw (1198.2bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13893.0bp,125.85bp) and (6321.9bp,116.12bp)  .. (6271.2bp,92.0bp) .. controls (6260.3bp,86.831bp) and (6251.4bp,77.225bp)  .. (16);
  \draw (6430.7bp,84.5bp) node {$-\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13772.0bp,126.64bp) and (4088.6bp,122.62bp)  .. (4024.2bp,92.0bp) .. controls (4013.3bp,86.836bp) and (4004.4bp,77.23bp)  .. (11);
  \draw (4146.7bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15057.0bp,120.35bp) and (15267.0bp,110.45bp)  .. (15297.0bp,92.0bp) .. controls (15306.0bp,86.539bp) and (15313.0bp,77.444bp)  .. (36);
  \draw (15475.0bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (15760.0bp,127.97bp) and (22404.0bp,133.72bp)  .. (22494.0bp,92.0bp) .. controls (22505.0bp,86.856bp) and (22515.0bp,77.254bp)  .. (53);
  \draw (23052.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (13800.0bp,126.47bp) and (4619.4bp,121.07bp)  .. (4558.2bp,92.0bp) .. controls (4547.3bp,86.835bp) and (4538.4bp,77.229bp)  .. (13);
  \draw (4718.7bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15961.0bp,126.69bp) and (26052.0bp,123.08bp)  .. (26119.0bp,92.0bp) .. controls (26130.0bp,86.864bp) and (26140.0bp,77.263bp)  .. (63);
  \draw (26299.0bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (16015.0bp,127.01bp) and (27186.0bp,126.3bp)  .. (27260.0bp,92.0bp) .. controls (27271.0bp,86.865bp) and (27281.0bp,77.265bp)  .. (67);
  \draw (27443.0bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (13645.0bp,127.38bp) and (1465.8bp,130.27bp)  .. (1385.2bp,92.0bp) .. controls (1374.3bp,86.839bp) and (1365.4bp,77.234bp)  .. (5);
  \draw (1923.7bp,84.5bp) node {$\mu_{1}(1-\mu_{2})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (14663.0bp,120.94bp) and (14477.0bp,111.7bp)  .. (14453.0bp,92.0bp) .. controls (14446.0bp,86.326bp) and (14442.0bp,77.543bp)  .. (33);
  \draw (14583.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (14303.0bp,126.64bp) and (12008.0bp,123.47bp)  .. (11942.0bp,92.0bp) .. controls (11931.0bp,86.788bp) and (11922.0bp,77.175bp)  .. (27);
  \draw (12030.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}/8$};
  \draw [black,->] (0) ..controls (13817.0bp,126.35bp) and (4961.2bp,120.07bp)  .. (4902.2bp,92.0bp) .. controls (4891.3bp,86.834bp) and (4882.4bp,77.229bp)  .. (14);
  \draw (5437.2bp,84.5bp) node {$(1-\mu_{1})(1-\mu_{2})(\mu_{1}^{2}\mu_{2}^{2} - 2\mu_{1}^{2}\mu_{2} + \mu_{1}^{2} - 2\mu_{1}\mu_{2}^{2} + 4\mu_{1}\mu_{2} - 2\mu_{1} + \mu_{2}^{2} - 2\mu_{2} + 2)/8$};
  \draw [black,->] (0) ..controls (15504.0bp,124.85bp) and (18761.0bp,113.1bp)  .. (18806.0bp,92.0bp) .. controls (18817.0bp,86.832bp) and (18827.0bp,77.226bp)  .. (45);
  \draw (18913.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{3}/8$};
  \draw [black,->] (0) ..controls (15225.0bp,125.7bp) and (16158.0bp,120.08bp)  .. (16215.0bp,92.0bp) .. controls (16226.0bp,86.714bp) and (16235.0bp,77.089bp)  .. (39);
  \draw (16395.0bp,84.5bp) node {$-3\mu_{1}^{2}\mu_{2}(1-\mu_{1})(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (16034.0bp,127.11bp) and (27535.0bp,127.3bp)  .. (27611.0bp,92.0bp) .. controls (27622.0bp,86.866bp) and (27632.0bp,77.266bp)  .. (68);
  \draw (27732.0bp,84.5bp) node {$\mu_{1}^{2}(\mu_{2} - 1)^{2}/2$};
  \draw [black,->] (0) ..controls (15947.0bp,126.61bp) and (25785.0bp,122.32bp)  .. (25851.0bp,92.0bp) .. controls (25862.0bp,86.863bp) and (25872.0bp,77.263bp)  .. (62);
  \draw (25993.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/8$};
  \draw [black,->] (0) ..controls (15924.0bp,126.46bp) and (25313.0bp,120.98bp)  .. (25376.0bp,92.0bp) .. controls (25387.0bp,86.863bp) and (25397.0bp,77.262bp)  .. (60);
  \draw (25522.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}(\mu_{2} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (15877.0bp,126.16bp) and (24416.0bp,118.43bp)  .. (24473.0bp,92.0bp) .. controls (24484.0bp,86.861bp) and (24494.0bp,77.26bp)  .. (57);
  \draw (24622.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (14276.0bp,127.28bp) and (11727.0bp,126.73bp)  .. (11654.0bp,92.0bp) .. controls (11643.0bp,86.794bp) and (11634.0bp,77.182bp)  .. (26);
  \draw (11787.0bp,84.5bp) node {$\mu_{1}^{2}(1-\mu_{1})(1-\mu_{2})^{3}/4$};
  \draw [black,->] (0) ..controls (14252.0bp,127.82bp) and (11465.0bp,129.76bp)  .. (11386.0bp,92.0bp) .. controls (11375.0bp,86.798bp) and (11366.0bp,77.187bp)  .. (25);
  \draw (11509.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{3}(1-\mu_{1})/8$};
  \draw [black,->] (0) ..controls (15271.0bp,127.62bp) and (16487.0bp,126.88bp)  .. (16561.0bp,92.0bp) .. controls (16572.0bp,86.767bp) and (16582.0bp,77.151bp)  .. (40);
  \draw (16713.0bp,84.5bp) node {$\mu_{2}^{2}(\mu_{1}  1)^{3}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15905.0bp,126.35bp) and (24964.0bp,119.99bp)  .. (25025.0bp,92.0bp) .. controls (25036.0bp,86.862bp) and (25046.0bp,77.261bp)  .. (59);
  \draw (25208.0bp,84.5bp) node {$3\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15128.0bp,127.97bp) and (15581.0bp,125.56bp)  .. (15641.0bp,92.0bp) .. controls (15651.0bp,86.628bp) and (15658.0bp,77.252bp)  .. (37);
  \draw (15765.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(1-\mu_{1})/4$};
  \draw [black,->] (0) ..controls (15843.0bp,125.94bp) and (23809.0bp,116.71bp)  .. (23862.0bp,92.0bp) .. controls (23873.0bp,86.86bp) and (23883.0bp,77.258bp)  .. (55);
  \draw (24042.0bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (14357.0bp,125.42bp) and (12529.0bp,118.08bp)  .. (12476.0bp,92.0bp) .. controls (12466.0bp,86.799bp) and (12457.0bp,77.321bp)  .. (29);
  \draw (12637.0bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15520.0bp,125.08bp) and (18957.0bp,114.21bp)  .. (19005.0bp,92.0bp) .. controls (19016.0bp,86.834bp) and (19026.0bp,77.229bp)  .. (46);
  \draw (19183.0bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (13725.0bp,126.93bp) and (3138.5bp,125.39bp)  .. (3068.2bp,92.0bp) .. controls (3057.3bp,86.837bp) and (3048.4bp,77.232bp)  .. (8);
  \draw (3194.7bp,84.5bp) node {$\mu_{1}\mu_{2}^{3}(\mu_{1} - 1)^{2}/4$};
  \draw [black,->] (0) ..controls (15456.0bp,127.57bp) and (18204.0bp,128.37bp)  .. (18282.0bp,92.0bp) .. controls (18293.0bp,86.824bp) and (18303.0bp,77.217bp)  .. (43);
  \draw (18431.0bp,84.5bp) node {$\mu_{1}\mu_{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13754.0bp,126.74bp) and (3750.7bp,123.6bp)  .. (3684.2bp,92.0bp) .. controls (3673.3bp,86.836bp) and (3664.4bp,77.231bp)  .. (10);
  \draw (3842.7bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (14760.0bp,111.37bp) and (14731.0bp,103.07bp)  .. (14722.0bp,92.0bp) .. controls (14717.0bp,85.643bp) and (14716.0bp,77.29bp)  .. (34);
  \draw (14845.0bp,84.5bp) node {$\mu_{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15937.0bp,126.55bp) and (25588.0bp,121.76bp)  .. (25652.0bp,92.0bp) .. controls (25663.0bp,86.863bp) and (25673.0bp,77.262bp)  .. (61);
  \draw (25759.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{3}/8$};
  \draw [black,->] (0) ..controls (14605.0bp,127.17bp) and (14241.0bp,123.3bp)  .. (14194.0bp,92.0bp) .. controls (14186.0bp,86.644bp) and (14181.0bp,77.814bp)  .. (32);
  \draw (14317.0bp,84.5bp) node {$\mu_{1}^{3}\mu_{2}^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15618.0bp,126.37bp) and (20237.0bp,121.43bp)  .. (20300.0bp,92.0bp) .. controls (20311.0bp,86.845bp) and (20321.0bp,77.242bp)  .. (48);
  \draw (20858.0bp,84.5bp) node {$\mu_{2}(1-\mu_{1})(3\mu_{1}^{2}\mu_{2}^{2} - 6\mu_{1}^{2}\mu_{2} + 3\mu_{1}^{2} - 6\mu_{1}\mu_{2}^{2} + 12\mu_{1}\mu_{2} - 6\mu_{1} + 3\mu_{2}^{2} - 6\mu_{2} + 4)/8$};
  \draw [black,->] (0) ..controls (13784.0bp,126.56bp) and (4355.0bp,121.84bp)  .. (4292.2bp,92.0bp) .. controls (4281.3bp,86.835bp) and (4272.4bp,77.23bp)  .. (12);
  \draw (4413.7bp,84.5bp) node {$-\mu_{1}(1-\mu_{1})(1-\mu_{2})^{2}/4$};
  \draw [black,->] (0) ..controls (14938.0bp,111.65bp) and (14955.0bp,103.82bp)  .. (14968.0bp,92.0bp) .. controls (14975.0bp,85.829bp) and (14980.0bp,77.081bp)  .. (35);
  \draw (15136.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/2$};
  \draw [black,->] (0) ..controls (15722.0bp,127.56bp) and (21785.0bp,130.2bp)  .. (21867.0bp,92.0bp) .. controls (21878.0bp,86.853bp) and (21888.0bp,77.251bp)  .. (51);
  \draw (22047.0bp,84.5bp) node {$3\mu_{1}^{2}\mu_{2}^{2}(1-\mu_{1})(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (15480.0bp,128.1bp) and (18480.0bp,131.51bp)  .. (18565.0bp,92.0bp) .. controls (18576.0bp,86.829bp) and (18586.0bp,77.222bp)  .. (44);
  \draw (18693.0bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (15174.0bp,123.27bp) and (15837.0bp,113.45bp)  .. (15878.0bp,92.0bp) .. controls (15888.0bp,86.683bp) and (15897.0bp,77.186bp)  .. (38);
  \draw (16053.0bp,84.5bp) node {$\mu_{1}\mu_{2}(\mu_{1}  1)^{2}(1-\mu_{2})^{2}/8$};
  \draw [black,->] (0) ..controls (15891.0bp,126.26bp) and (24697.0bp,119.23bp)  .. (24756.0bp,92.0bp) .. controls (24767.0bp,86.862bp) and (24777.0bp,77.26bp)  .. (58);
  \draw (24898.0bp,84.5bp) node {$\mu_{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/4$};
  \draw [black,->] (0) ..controls (14063.0bp,126.84bp) and (9054.6bp,124.6bp)  .. (8986.2bp,92.0bp) .. controls (8975.3bp,86.821bp) and (8966.4bp,77.213bp)  .. (22);
  \draw (9147.2bp,84.5bp) node {$\mu_{1}\mu_{2}^{2}(\mu_{1} - 1)^{2}(1-\mu_{2})/8$};
  \draw [black,->] (0) ..controls (13713.0bp,126.99bp) and (2898.9bp,126.08bp)  .. (2827.2bp,92.0bp) .. controls (2816.3bp,86.837bp) and (2807.4bp,77.232bp)  .. (7);
  \draw (2936.2bp,84.5bp) node {$\mu_{1}^{2}\mu_{2}(1-\mu_{2})/4$};
%
\end{tikzpicture}

### 2019-09-12 2-Chapters/5-Chapter/chapter5.tex
Typo: Etienne Boursier and Vianney Perchet → E. Boursier and V. Perchet, not V. Boursier and E. Perchet
Removed raw inclusion of https://hal.inria.fr/hal-01629733v2

@@
 
-\graphicspath{{2-Chapters/5-Chapter/Images/}{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/figures/}}
+\graphicspath{{2-Chapters/5-Chapter/Images/}{2-Chapters/5-Chapter/ALT_2018__MPBandits/figures/}}
 
@@
 % ----------------------------------------------------------------------------
-\input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex}
+\input{2-Chapters/5-Chapter/ALT_2018__MPBandits/mpBandits.tex}
 
@@
 Inspired by the publication of our article \cite{Besson2018ALT},
-V. Boursier and E. Perchet studied this question in the following summer $2018$ \cite{BoursierPerchet18}.
+E. Boursier and V. Perchet studied this question in the following summer $2018$ \cite{BoursierPerchet18}.
 Their main contribution is an answer to the aforementioned open problem,
@@
 % % ----------------------------------------------------------------------------
-% \input{2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex}
+% \input{2-Chapters/5-Chapter/ALT_2018__MPBandits/mpBandits_appendix.tex}

### 2019-09-12 2-Chapters/6-Chapter/nonstatbandits/NonStatB.tex
Removed raw inclusion of https://hal.inria.fr/hal-02006471

--- 2019-07-11 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
@@
     \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/6_Problems/Problem_1.pdf}
     \caption{Problem $1$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).
@@
     \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/6_Problems/Problem_2.pdf}
     \caption{Problem $2$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).
@@
     \centering
-    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/6_Problems/Problem_5.pdf}
     \caption{Problem $3$: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
@@
     \centering
-    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/6_Problems/Problem_4.pdf}
     \caption{Problem $4$: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
@@
     \centering
-    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
+    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/6_Problems/Problem_6.pdf}
     \caption{Pb $5$: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
@@
     \centering
-    \includegraphics[width=1.15\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
+    \includegraphics[width=1.15\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/SP__K3_T5000_N1000__9_algos_pb1/main____env1-1_1316297932259102962.pdf}
     \caption{Mean regret as a function of time, $R_t$ for horizon $T = 5000$, for problem 1.}
@@
     \centering
-    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
+    \includegraphics[width=1.09\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/SP__K3_T5000_N1000__9_algos_pb2/main____env1-1_7882655236216606305.pdf}
     \caption{Mean regret as a function of time, $R_t$ for horizon $T = 5000$, for problem $2$.}
@@
     \centering
-    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
+    \includegraphics[width=1.00\linewidth]{2-Chapters/6-Chapter/nonstatbandits/figures/SP__K3_T5000_N1000__9_algos_pb1/main_HistogramsRegret_shareX____env1-1_1316297932259102962.pdf}
     \caption{Histograms of the distributions of regret $R_T$ ($T=5000$) for problem 1.}

### 2019-09-22 1-Preamble/4-Acknowledgement/acknowledgement.tex
❤
Update link to one PhD thesis that was thanking me.

@@
 \href{http://jill-jenn.net/}{Jill-Jênn Vie} (\href{https://github.com/jilljenn/phd}{thesis link $\nearrow$}), \href{https://navikkumarmodi.wordpress.com/}{Navikkumar Modi} (\href{https://tel.archives-ouvertes.fr/tel-01668536/document}{$\nearrow$}), \href{https://quentinbodinier.wordpress.com/}{Quentin Bodinier} (\href{https://tel.archives-ouvertes.fr/tel-01731022/document}{$\nearrow$}), \href{https://paris-sorbonne.academia.edu/FBesson}{Florian Besson}\footnote{~Que je remercie deux fois plus que ce qu'il me remerciait dans sa propre thèse !
-Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{the}\href{https://www.theses.fr/s175725}{$\nearrow$}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{$\nearrow$}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{$\nearrow$}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{$\nearrow$}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{$\nearrow$}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/documents/recherche/these/Main.pdf}{$\nearrow$}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, and Rami Othman.
+Pour répondre à la question, deux ans plus tard, tu remerciais Tom et moi un nombre non nul $x\in\N\cup\{+\infty\}$ satisfaisant $x=2x$, soit $x=+\infty$, et je t'en remercie $2x$ fois !}, \href{https://sites.google.com/view/guerand}{Jessica Guérand} (\href{https://www.theses.fr/s175725}{$\nearrow$}), \href{https://sites.google.com/view/ducasse/}{Romain Ducasse} (\href{https://drive.google.com/open?id=1u2oxRRimcO0jQfuYwSVwgKfcHU5DdoPK}{$\nearrow$}), \href{https://members.loria.fr/SAbelard/}{Simon Abelard} (\href{https://members.loria.fr/SAbelard/theseabelard.pdf}{$\nearrow$}), \href{http://www.cmap.polytechnique.fr/~sacchelli/}{Ludovic Sacchelli} (\href{https://tel.archives-ouvertes.fr/tel-01893068/document}{$\nearrow$}), \href{http://pages.saclay.inria.fr/claire.brecheteau/}{Claire Brécheteau} (\href{https://hal.archives-ouvertes.fr/tel-01897787/document}{$\nearrow$}), \href{http://www.i2m.univ-amu.fr/perso/damien.allonsius/}{Damien Allonsius} (\href{https://hal.archives-ouvertes.fr/tel-01887023/document}{$\nearrow$}), \href{https://remibonnefoi.wordpress.com/}{Rémi Bonnefoi}, and Rami Othman.
+% Romain : https://www.theses.fr/s175711
+% TODO: Rami Othman
 
@@
 ¡Muchísimas gracias, gatita!}
-    \textcolor{white}{$\heartsuit$ Te quiero, Lola, ¿quieres vivir conmigo?}
+    \textcolor{white}{$\heartsuit$ Te quiero, Lola, ¿Quieres compartir tu vida conmigo?}
 % }

### 2019-11-15 1-Preamble/4-Acknowledgement/acknowledgement.tex
Mise à jour de la dédicace et des remerciements. Quelle déception et que de tristesse…
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

@@
 \indent
-I am grateful to amazing professors who passed on their passion for mathematics and computer science to me.
-A partial list includes Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers in Marseille.
+I am grateful to amazing professors who passed on their passion for mathematics and computer science to me,
+% A partial list 
+including Hubert P. in Briançon ($2006$-$09$), Yassine D. ($2009$-$10$) and Agnès B. ($2010$-$11$) in Lycée Thiers in Marseille.
 Then at ENS de Cachan ($2011$-$16$), I especially think about Frédéric P., Alain T., Claudine P., Florian de V.
@@
 I would also like to thank some unusual things.
-I thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday, despite the too many accidents and frequent rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} that gave me smiles and energy, and my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git (and GitHub and Bitbucket), \LaTeX, Bash, Julia and Python, GNU/Linux and (X)Ubuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
+I thank my \href{https://perso.crans.org/besson/zero-dechet/}{bike} which I enjoyed riding everyday, despite the too many accidents and frequent rain, my \href{https://perso.crans.org/besson/cuisine/}{enthusiastic daily cooking} that gave me smiles and energy, and my \href{https://perso.crans.org/besson/}{favorite open-soure technological tools}: git, GitHub and Bitbucket, \LaTeX, Bash, Julia and Python, GNU/Linux and (X)Ubuntu, Firefox, Konsole and Visual Studio Code, and many other great free and open-source tools.
 Some people thank (a/some) God in their thesis, I feel that I prefer to thank \emph{the Earth}, and apologize for the too many times I took a plane for a conference abroad, and in general apologize for the environmental impact of this research work.

### 2019-11-15 2-Chapters/1-Chapter/chapter1.tex
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

@@
     See Chapter~\ref{chapter:4}.
+    % Preprint at \href{https://HAL.Inria.fr/hal-XXX}{\texttt{HAL.Inria.fr/hal-XXX}}.
+    \cite{MoyBesson2019Annales}\\
     Submitted for a special volume of \href{https://annalsoftelecommunications.wp.imt.fr/}{the Annals of Telecommunications} journal, on ``Machine Learning for Intelligent Wireless Communications and Networking''.
-    % Preprint at \href{https://HAL.Inria.fr/hal-XXX}{\texttt{HAL.Inria.fr/hal-XXX}}.
-    \cite{MoyBesson2019Annales}
-
-    % \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
-
-\end{itemize}
-
-% =============================================================================
-\subsection*{In progress works waiting for a new submission}
-
-\begin{itemize}
+
+\item
+    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
+    by \textbf{Lilian Besson} and others, active development since October $2016$,
+    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
+    It currently consists in about $45000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
+    and a complete documentation accessible on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} or \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.\\
+    See Chapter~\ref{chapter:3}.
+    \cite{SMPyBanditsJMLR,SMPyBandits}\\
+    Submitted for a special track of the \href{http://jmlr.org/mloss}{Journal of Machine Learning Research}, for Machine Learning Open-Source Software (MLOSS), in October $2019$.
 
@@
     Preprint at \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
-    \cite{Besson2019GLRT}
-
-    \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
-
-\item
-    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
-    by \textbf{Lilian Besson}, active development since October $2016$,
-    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
-    It currently consists in about $45000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
-    and a complete documentation accessible on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}} or \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}.
-    See Chapter~\ref{chapter:3}.
-    \cite{SMPyBanditsJMLR,SMPyBandits}
+    \cite{Besson2019GLRT}\\
+    An updated version, written with Odalric-Ambrym Maillard, was submitted to \href{http://www.aistats.org}{AISTATS 2020} in October $2020$.
+
+    % \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
+
+    % \emph{$\hookrightarrow$ Note that we are already working on an extended version that will be submitted to a journal on Machine Learning and Statistical Learning, in autumn $2019$.}
+
+\end{itemize}
+
+% =============================================================================
+\subsection*{In progress works waiting for a new submission}
+
+\begin{itemize}
 

### 2019-11-15 2-Chapters/5-Chapter/chapter5.tex
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

@@
 % (on the three algorithms and two cases, about $\times 3.4$),
-and consistent on the different algorithms.
+and consistent on the different algorithms.\newline
 The extreme case of $M=K$ players is of highest interest, as \MCTopM{} achieves a very small regret (proven to be $\cO(1)$ as it is just the expected time of the orthogonalization process), while the same algorithm with unknown $M$ suffers from a much larger regret.
@@
 
+\vfill{}
+
+\hr{}
+\begin{small}
+    \begin{quote}
+        % https://fr.wikiquote.org/wiki/Kaamelott/
+        \emph{-- Victoriae mundis et mundis lacrima.\\
+            Bon, ça ne veut absolument rien dire, mais je trouve que c’est assez dans le ton.}\\
+        Le Roi Loth, interprété par François Rollin,\\
+        \emph{Kaamelott}, Livre IV, ``Le désordre et la nuit''.
+
+        \vspace*{10pt}
+
+        \emph{-- C'est pas moi qu'explique mal, c'est les autres qui sont cons !}\\
+        Perceval, interprété par Franck Pitiot,\\
+        \emph{Kaamelott}, Livre IV, ``Perceval Fait Raitournelle''.
+    \end{quote}
+\end{small}
+
 % -----------------------------------------------------------------

### 2019-11-15 2-Chapters/6-Chapter/chapter6.tex
Removed raw inclusion of https://hal.inria.fr/hal-02006471
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist

@@
 
-\input{2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex}
+\input{2-Chapters/6-Chapter/nonstatbandits/NonStatB.tex}
 
@@
 
-\input{2-Chapters/6-Chapter/nonstatbandits.git/NonStatB_appendix.tex}
+\input{2-Chapters/6-Chapter/nonstatbandits/NonStatB_appendix.tex}
+
+\vfill{}
+
+\hr{}
+\begin{small}
+    \begin{quote}
+        % https://fr.wikiquote.org/wiki/Kaamelott/
+        \emph{-- Par exemple, vous prenez aujourd’hui.
+Vous comptez sept jours.
+Ça vous emmène dans une semaine.
+Et bien on sera exactement le même jour qu’aujourd’hui\dots{} À une vache près, hein… C’est pas une science exacte.}\\
+        Karadoc, interprété par Jean-Christophe Hembert,\\
+        \emph{Kaamelott}, Livre II, ``Sept cent quarante-quatre''.
+
+        % \emph{-- Les chiffres, c'est pas une science exacte figurez-vous !}\\
+        % Karadoc, interprété par Jean-Christophe Hembert,\\
+        % \emph{Kaamelott}, Livre V, ``Le dernier recours''.
+    \end{quote}
+\end{small}

### 2019-11-16 1-Preamble/5-Abstract/resume_fr.tex
Typos

@@
 %
-Historiquement, trois grandes familles de systèmes de communication sans fil ont été déployées dans les grands réseaux commerciaux : d'abord, la radiodiffusion centralisée (\eg, radio ou télédiffusion), puis les systèmes bidirectionnels centralisés (\eg, 4G ou Wi-Fi), et aujourd'hui la collecte décentralisée de données pour les réseaux de l'Internet des objets (\eg, réseaux capteurs).
+Historiquement, trois grandes familles de systèmes de communication sans fil ont été déployées dans les grands réseaux commerciaux : d'abord, la radiodiffusion centralisée (\eg, radio ou télédiffusion), puis les systèmes bidirectionnels centralisés (\eg, 4G ou Wi-Fi), et aujourd'hui la collecte décentralisée de données pour les réseaux de l'Internet des Objets (\eg, réseaux capteurs).
 %
@@
 %
-Les organismes de réglementation dans le monde, comme l'
-\href{https://www.itu.int/en/Pages/default.aspx}{Union Internationale des Télécommunications} (\emph{ITU}, voir \href{https://www.itu.int/}{\texttt{www.ITU.int}}),
+Les organismes de réglementation dans le monde, comme l'\href{https://www.itu.int/en/Pages/default.aspx}{Union Internationale des Télécommunications} (\emph{ITU}, voir \href{https://www.itu.int/}{\texttt{www.ITU.int}}),
 la \href{https://www.fcc.gov/}{Commission Fédérale des Communications} aux États-Unis d'Amérique du Nord (\emph{FCC}, voir \href{https://www.fcc.gov/}{\texttt{FCC.gov}})
@@
 longue portée,
-et un faible cycle d'utilisation (\ie, faible à très faible nombre de messages par jour).
+et un faible cycle d'utilisation (\ie, faible à très faible nombres de messages par jour).
 %
@@
 Dans les bandes sous licences, il y a des utilisateurs primaires (UP) qui paient un abonnement pour accéder au réseau, par exemple n'importe qui doit payer pour avoir un numéro de téléphone mobile et utiliser le réseau.
-Comme nous l'avons vu dans la définition ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est utilisé par aucun UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser aussi, éventuellement en payant un abonnement auprès de l'opérateur du réseau.
+Comme nous l'avons vu dans les définitions ci-dessus de la philosophie de la CR, nous pouvons imaginer que si le réseau n'est utilisé par aucun UP à un certain endroit et à une certaine heure, les utilisateurs non licenciés, appelés utilisateurs secondaires (US), pourraient l'utiliser aussi, éventuellement en payant un abonnement auprès de l'opérateur du réseau.
 %
@@
 La première est une comparaison empirique d'algorithmes existants, se focalisant sur les plus efficaces et les mieux connus, en Section~\ref{sec:3:reviewSPAlgorithms} et \ref{sec:3:timeAndMemoryCosts}.
-Nous confirmons que les algorithmes les plus utilisés, tels que \UCB{} \cite{Auer02}, l'échantillonage de Thompson (Thompson sampling) \cite{Thompson33} et \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
+Nous confirmons que les algorithmes les plus utilisés, tels que \UCB{} \cite{Auer02}, l'échantillonage de Thompson (\emph{Thompson sampling}) \cite{Thompson33} et \klUCB{} \cite{KLUCBJournal}, sont les plus efficaces en termes de regret, et offrent un bon équilibre entre regret et complexité (temps et mémoire).
 La seconde approche est la sélection d'algorithme en ligne, consistant à agréger un ensemble (fini) d'algorithmes et à découvrir automatiquement lequel est le plus efficace, contre un problème donné, avec notre contribution \Aggr{} que nous détaillons en Chapitre~\ref{chapter:25}.
@@
 Cela signifie que chaque objet vise son propre objectif local, qui est de maximiser sa récompense cumulée, d'une manière égoïste et complètement agnostique des autres objets ayant le même objectif.
-Il est bien connu en théorie des jeux que jouer égoïstement peut être désastreux pour la mesure centralisée de la performance, on peut penser aux \guillemotleft{} dilemmes \guillemotright{} populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
+Il est bien connu en théorie des jeux que jouer égoïstement peut être désastreux pour la mesure de performance centralisée, on peut penser aux \guillemotleft{} dilemmes \guillemotright{} populaires, tels que le \href{https://fr.wikipedia.org/wiki/Dilemme_du_prisonnier}{dilemme du prisonnier}.
 Il était donc assez surprenant que nos simulations numériques, ainsi que la preuve de concept réaliste, aient montré que l'apprentissage BMB décentralisé et égoïste conduisait à une coordination efficace entre les objets, dans tous les scénarios considérés, malgré le fait que ces objets ne peuvent communiquer directement entre eux, et reçoivent seulement un indicateur de collision depuis la station de base à laquelle ils sont associés.
@@
 Le cas AOS couvert par notre modèle peut sembler légèrement différent de celui considéré par \cite{Jouini10} et d'autres travaux antérieurs,
-car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est disponible.
+car notre modèle exige qu'un \Ack{} soit renvoyé par la station de base si la transmission a réussi, même dans le cas où l'information de détection est indisponible.
 Il est en fait équivalent aux modèles précédemment étudiés d'applications des BMB pour l'AOS, puisqu'ils ne prennent en compte que des objets synchronisés, un US et des UP, et donc si la détection indique qu'un canal est libre pour un intervalle de temps, le message envoyé par l'US sera certainement reçu avec succès par la station de base (dans le modèle idéal), donc il n'y aura aucun risque de collision, donc il n'a pas besoin d'\Ack.
@@
     Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le Chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle (\cite{Besson2019Gretsi} publié à GRETSI 2019, \cite{Besson2019GLRT}).
-    Nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles car notre approche n'a pas besoin de connaître de bornes sur la difficulté du problème à part le nombre de points de ruptures.
+    Nous proposons un nouvel algorithme activement adaptatif, pour ces problèmes stationnaires par morceaux, \GLRklUCB, qui atteint des performances empiriques et des garanties théoriques comparables à l'état de l'art, tout en utilisant des hypothèses plus faibles car notre approche n'a pas besoin de connaître de borne sur la difficulté du problème à part le nombre de points de ruptures.
     % Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB

### 2019-11-16 2-Chapters/Conclusion/conclusion.tex
Small changes, 3 more citations
- Comme dirait Christopher Walk dans Django
- I'm sorry, I couldn't resist
Typos

--- 2019-07-11 2-Chapters/Conclusion/conclusion.tex
@@
 or to drop their packet after a failed transmission (\ie, with or without retransmission).
-Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order that the numerous IoT devices improve their spectrum access scheme, on their own, by learning which of the $K$ frequency channels are the less occupied by the environmental traffic (in average).
+Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order that the numerous IoT devices improve their spectrum access scheme, on their own, by learning which of the $K$ frequency channels are the less occupied ones by the environmental traffic (in average).
 
@@
 (see tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).
-From the point-of-view of the theoretical analysis, we are especially interesting about first extending our proposed algorithm \MCTopM{} to the case of an unknown number of players.
+From the point-of-view of the theoretical analysis, we are especially interested by first extending our proposed algorithm \MCTopM{} to the case of an unknown number of players.
 Then we could also address the most interesting extension for the application to wireless networks, \ie, the ``dynamic case'' which allows for arrivals and departures of players in the multi-player bandit game.
@@
 A few tasks are left on SMPyBandits, a first one could be to implement new variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
-For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
+For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits/SMPyBandits/issues/}}.
 %
@@
 \begin{quote}
-    \emph{-- Les rêves, ça se compare pas\ldots}
-    Alexandre Astier, \emph{Kaamelott}, Livre VI, Épisode 9, ``Dies Irae''.
+    \emph{-- Les rêves, ça se compare pas\ldots}\\
+    Le Roi Arthur, interprété par Alexandre Astier,\\
+    \emph{Kaamelott}, Livre VI, Épisode 9, ``Dies Irae''.
     %  écrit par Alexandre Astier.

### 2019-11-27 2-Chapters/1-Chapter/chapter1.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 
-This third kind of wireless systems are characterized by their decentralized nature,
+This third kind of wireless systems is characterized by its decentralized nature,
 where communications are initialized and regulated by the devices, not by a centralized control systems.
@@
 % Il faut expliquer ce que c'est que les bandits, rapidement, et leur application à l'OSA
-As this thesis focus on Reinforcement Learning models,
+As this thesis focuses on Reinforcement Learning models,
 it is important to highlight that in such decision making models, the learner does not have access to the entire reaction of the environment after taking its action.
@@
 %
-The MAB problem have been studied in both the machine learning and the statistics community, since the $1950$s with pioneers like H.
+The MAB problem has been studied in both the machine learning and the statistics communities, since the $1950$s with pioneers like H.
 Robbins \cite{Robbins52}, and more recently it is an active field of research, since the late $1990$s \cite{Anantharam87a,Anantharam87b,auer1995gambling,Agrawal95}.
@@
 %
-The following hypotheses are maid:
+The following hypotheses are made:
 the SU is equipped with a spectrum sensing capability,
 and considers a fixed and finite set of $K$ orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-For instance, it can be a set of $K=3$ Wi-Fi channels at different frequency, emitted by the same Wi-Fi station.
+For instance, it can be a set of $K=3$ Wi-Fi channels at different frequencies, emitted by the same Wi-Fi station.
 Another hypothesis is that both PU and SU are synchronized in time, by sub-dividing the time in discrete time steps (or rounds).
@@
 and the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$ is modeled by a reward of $r(t) \in \{0,1\}$.
-Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus the SU can send a message), while a reward of $r(t)=0$ indicates that the channel $k$ is busy at during the time slot $t$, and no message should be sent, until the end of that slot.
+Indeed, $r(t) = 1$ indicates that no Primary User was sensed (and thus the SU can send a message), while a reward of $r(t)=0$ indicates that the channel $k$ is busy during the time slot $t$, and no message should be sent, until the end of that slot.
 %
@@
 %
-The aforementioned literature have essentially shown that MAB algorithms can be applied successfully on the SU side for the OSA problem.
+The aforementioned literature has essentially shown that MAB algorithms can be applied successfully on the SU side for the OSA problem.
 However, if we consider CR-ready devices that cannot perform spectrum sensing, such as low-cost and low energy-consumption end-devices designed for the future Internet of Things networks, the MAB model that uses sensing to detect PU in the OSA case can no longer be applied.
@@
 %
-Moreover, due to both their limited hardware (Rx/Tx), limited computational power and limited battery life, the IoT devices are able to use their received antenna to sense one channel occasionally (typically, for a few time slots after every single up-link message), but they cannot perform spectrum sensing like it is considers for OSA and CR (\ie, at the beginning of a time slot, in a ``listen-before-talk'' scheme).
+Moreover, due to both their limited hardware (Rx/Tx), limited computational power and limited battery life, the IoT devices are able to use their received antenna to sense one channel occasionally (typically, for a few time slots after every single up-link message), but they cannot perform spectrum sensing like it is considered for OSA and CR (\ie, at the beginning of a time slot, in a ``listen-before-talk'' scheme).
 
@@
 To answer this question, we present two approaches.
-The first one is an empirical comparison of the most efficient and well known existing algorithms, in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}.
+The first one is an empirical comparison of the most efficient and well known existing algorithms, in Sections~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts}.
 We confirm that widely used and not too sophisticated algorithms, such as \UCB{} \cite{Auer02}, Thompson sampling \cite{Thompson33} and \klUCB{} \cite{KLUCBJournal}, are the most efficient in terms of regret, and offer a good balance between regret and (time and memory) complexity.
@@
 The second model with retransmissions concluded our collaboration with R.
-Bonnefoi, and resulted to a paper presented in the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
+Bonnefoi, and resulted in a paper presented in the $1^{\text{st}}$ MoTION workshop, also during the IEEE WCNC $2019$ conference \cite{Bonnefoi2019WCNC}.
 
@@
 In both models of IoT networks presented above, the core idea is to let every dynamic device run a learning algorithm in a fully decentralized way, in order to optimize the entire system.
-We highlight that each devices communicates, and thus learns, only at some rounds and not at all the time steps, by following its own random activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
+We highlight that each device communicates, and thus learns, only at some rounds and not at all the time steps, by following its own random activation process (we restrict to a purely random Bernoulli activation process of probability $p$).
 This means that each of them target its own local objective, which is to maximize its cumulated reward, in a selfish way and by being agnostic about the other devices following the same objective.
@@
 % https://www.ii.uni.wroc.pl/~lukstafi/pmwiki/uploads/AGT/Prediction_Learning_and_Games.pdf
-So it was quite surprising that the numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning lead to efficient coordination between devices in all the considered scenarios, despite the fact that these IoT objects cannot communicate directly with each other, and receive only a collision feedback from the base station they are associated to.
+So it was quite surprising that the numerical simulations, as well as the realistic PoC, showed that decentralized selfish MAB learning leads to efficient coordination between devices in all the considered scenarios, despite the fact that these IoT objects cannot communicate directly with each other, and receive only a collision feedback from the base station they are associated to.
 
@@
 but rather that we consider algorithms which are all playing in different (random) activation times, subsets of the global synchronized time steps.
-Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is in fact not really realistic for IoT networks, even though this lead to the interesting model of Chapter~\ref{chapter:5}.
+Generalizing to different activation probabilities would lead to a model even harder to analyze, and enforcing at most $M \leq K$ devices is in fact not really realistic for IoT networks, even though this leads to the interesting model of Chapter~\ref{chapter:5}.
 
@@
 Assuming the environment to be piece-wise stationary indeed makes sense for applications to wireless networks, where a change-point can for instance corresponds to the arrival of a new group of end-devices in the network.
-Examples of such situation can be a new company arriving on the market in one city (\eg, the Linky counters being set-up nowadays in France), or the neighbor farmer installing sensors on his own herd of cows, etc.
+Examples of such a situation can be a new company arriving on the market in one city (\eg, the Linky counters being set-up nowadays in France), or the neighbor farmer installing sensors on his own herd of cows, etc.
 %
 Two main families of algorithms have been proposed for the piece-wise stationary problem,
-that usually consist in combining an efficient policy designed for the stationary MAB problem (\eg, Thompson Sampling or \klUCB) and a way to adapt to changes in the arms distributions.
+that usually consists in combining an efficient policy designed for the stationary MAB problem (\eg, Thompson Sampling or \klUCB) and a way to adapt to changes in the arms distributions.
 Passively adaptive policies use a window of a fixed or evolving size \cite{Garivier11UCBDiscount}, or a discount factor, in order to forget about the past observations \cite{Kocsis06,Gupta11thompson},
@@
 This last chapter is based on our last work, published at the GRETSI conference in Lille, France, in August $2019$ \cite{Besson2019Gretsi}.
-This work also lead to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results, in order to submit it soon to a journal (probably the Journal of Machine Learning Research) before December $2019$.
+This work also leads to the long version article \cite{Besson2019GLRT}, that will be partially rewritten and completed with more recent results, in order to submit it soon to a journal (probably the Journal of Machine Learning Research) in $2020$.
 
@@
     % \item
-\textcolor{darkgreen}{Then, the second Part~\ref{part:MABIOT}} contain three chapters (third line), that are included in both the logical and chronological orders, but can be read almost independently.
+\textcolor{darkgreen}{Then, the second Part~\ref{part:MABIOT}} contains three chapters (third line), that are included in both the logical and chronological orders, but can be read almost independently.
 Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where we show that MAB algorithms can be used with success.
@@
 For this reason, we weaken the models for the rest of the document,
-and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the intractable IoT networks models from Chapter~\ref{chapter:4}.
+and both Chapters~\ref{chapter:5} and \ref{chapter:6} study an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the intractable IoT networks models from Chapter~\ref{chapter:4}.
 % \end{itemize}
@@
     by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
-    Demonstration presented in \emph{International Conference on Communication},
+    Demonstration presented in \emph{International Conference on Telecommunications},
     % \footnote{~ICT 2018, see \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}},
@@
     \cite{Besson2019GLRT}\\
-    An updated version, written with Odalric-Ambrym Maillard, was submitted to \href{http://www.aistats.org}{AISTATS 2020} in October $2020$.
+    An updated version is in writing, with Julien Seznec and Odalric-Ambrym Maillard.
 

### 2019-11-27 2-Chapters/2-Chapter/agregation.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 \cite[Theorem 4.2]{Bubeck12} proves that
-\ExpQ{} have a bounded adversarial regret, % $\overline{R_T}$
+\ExpQ{} has a bounded adversarial regret, % $\overline{R_T}$
 $\overline{R_T} \leq 2 \sqrt{T N \log(K)}$,

### 2019-11-27 2-Chapters/2-Chapter/chapter2.tex
Trying to kill this last empty page...
Almost done proofreading résumé fr
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 %
-For more details and a more formal introduction to multi-armed bandits, we suggest to the interested reader to work on a very recent text-book by Slivkins \cite{Slivkins2019}.
+For more details and a more formal introduction to multi-armed bandits, we suggest the interested reader to work on a very recent text-book by Slivkins \cite{Slivkins2019}.
 Another excellent but reasonably short survey is the book by Bubeck and Cesa-Bianchi \cite{Bubeck12}, while the more recent book by Lattimore and Szepesv{\'a}ri \cite{LattimoreBanditAlgorithmsBook} is the most complete resource about bandit algorithms.
@@
 %
-We focus on the \textbf{stochastic} MAB, in which the reward stream are drawn from some (unknown) distributions, before the beginning of the game.
+We focus on the \textbf{stochastic} MAB, in which the reward stream is drawn from some (unknown) distributions, before the beginning of the game.
 %
@@
 The initial information is reduced to $O_1 = (U_0)$, and the first decision is $A(1) = \cA_1(O_1)$.
-Usually, most algorithms starts by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps $t=1,\dots,K$ (\eg, Algorithm~\ref{algo:2:indexPolicy}).
+Usually, most algorithms start by selecting $A(1)=1,\dots,A(K)=K$ (or a permutation of the $K$ arms) in the $K$ first steps $t=1,\dots,K$ (\eg, Algorithm~\ref{algo:2:indexPolicy}).
 % as illustrated for example in Algorithm~\ref{algo:2:indexPolicy} below.
@@
 
-In the literature, parametric assumptions on the rewards are sometimes considered, typically the assumption is that rewards belong to a real-valued distribution lying in an exponential families.
+In the literature, parametric assumptions on the rewards are sometimes considered, typically the assumption is that rewards belong to a real-valued distribution lying in an exponential family.
 %  or bounded distributions.
@@
 In most cases, our main interest is Bernoulli distributions, but we prefer to present the more general notations of exponential families in one dimension.
-% We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Emilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
+% We follow the notations from the course on statistical learning (Stat 260) taught in 2010 by Michael Jordan at the University of Berkeley \cite{JordanCourseStatBerkeley}, that are also the notations used for instance in Émilie Kaufmann's PhD in 2014 \cite{Kaufmann12PhD}.
 
@@
     \item
-    \emph{A distribution in such family is entirely characterized by its parameter $\eta$}.
+    \emph{A distribution in such a family is entirely characterized by its parameter $\eta$}.
     And the mapping $\eta \mapsto \E_{\eta}[X]$ is one-to-one,
@@
 % In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
-In the model of Opportunistic Spectrum Access (OSA) with sensing \cite{Jouini09,Jouini10}, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$, at the beginning of its $t$-th transmission slot.
+In the model of Opportunistic Spectrum Access (OSA) with sensing \cite{Jouini09,Jouini10}, the samples $Y_{k,t}$ represent the feedback obtained by the CR-equipped device after sensing the channel $k$, at the beginning of its $t$-th transmission slot.
 A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no up-link message can be sent.
@@
 %
-Such decomposition of the regret can be useful for at least two reasons:
+Such a decomposition of the regret can be useful for at least two reasons:
 
@@
 These lower-bounds are of highest interest to design efficient algorithms,
-and a significant part of the research on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically \cite{LaiRobbins85}.
+and a significant part of the research on MAB algorithms has focused on finding algorithms that match the Lai and Robbins' lower-bound asymptotically \cite{LaiRobbins85}.
 %
@@
 Theorem~2.12 from \cite{Slivkins2019} specifies a possible value for the constant $C_I$.
-Note that the lower-bound of Lai and Robbins, proved in \cite{LaiRobbins85}, is more general and applies to any one-dimensional exponential family $\cE$, by replacing the Bernoulli Kullback-Leibler divergence $\kl$ with the KL divergence of this family, $\kl_{\cE}$.
+Note that the lower-bound of Lai and Robbins, proven in \cite{LaiRobbins85}, is more general and applies to any one-dimensional exponential family $\cE$, by replacing the Bernoulli Kullback-Leibler divergence $\kl$ with the KL divergence of this family, $\kl_{\cE}$.
 
@@
 For this reason, another family of regret bounds is relevant, that are not problem-dependent but worst-case, or also called minimax \cite{Audibert2009minimax,audibert2010minimax} or problem-independent.
-We give an example of such result below.
+We give an example of such a result below.
 % Hence it is interesting to quote another lower-bound on the regret, that can bring useful information in such settings.
@@
 Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
-which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all nontrivial problems, ruling out the case where $\mu_1=\mu_2=\dots=\mu_K$).
+which gives $R_T \propto T$ for any problem with at least one sub-optimal arm (\ie, all nontrivial problems, ruling out the case where $\mu_1=\mu_2=\dots=\mu_K$).
 
@@
     % \end{small}
-    \vspace*{-5pt}  % FIXME vspace negative to compensate the \end{leftbar}
+    % \vspace*{-5pt}  % FIXME vspace negative to compensate the \end{leftbar}
 \end{leftbar}  % XXX leftbar lemmabar, comment if needed
@@
 for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
-The second one is less obvious: one has to solve the optimisation problem in \eqref{eq:2:klUCB_index}.
+The second one is less obvious: one has to solve the optimization problem in \eqref{eq:2:klUCB_index}.
 Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\kl(x,\bullet)$ function is usually convex (for any $x$), two practical and mathematically correct solutions are Newton's method or a bisection search.
@@
 Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence interval $\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) \eqdef \widehat{\mu_k}(t) - \xi_k(t)$),
-but instead of maximizing the UCB, a random sample is taken on each interval $s_k \sim \cU(\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right])$, and the played arm is the one maximizing this sample, \ie, $A(t) \in \argmax_k s_k$.
+but instead of maximizing the UCB, a random sample is simply taken on each interval $s_k \sim \cU(\left[ \mathrm{LCB}_k(t), \UCB_k(t) \right])$, and the played arm is the one maximizing this sample, \ie, $A(t) \in \argmax_k s_k$.
 % https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html

### 2019-11-27 2-Chapters/2-Chapter/chapter25.tex
Removed raw inclusion of https://hal.inria.fr/hal-01705292
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 %
-The review of MAB algorithms given in the previous Chapter~\ref{chapter:2} showed that there are a large collection of different algorithms designed for different kinds of bandit problems.
+The review of MAB algorithms given in the previous Chapter~\ref{chapter:2} showed that there is a large collection of different algorithms designed for different kinds of bandit problems.
 In this chapter, we discuss the question of online algorithm selection.
@@
 
-\input{2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex}
+\input{2-Chapters/2-Chapter/agregation.tex}
 

### 2019-11-27 2-Chapters/3-Chapter/chapter3.tex
Removed raw inclusion of https://hal.inria.fr/hal-01840022
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 
-\graphicspath{{2-Chapters/3-Chapter/Images/}{2-Chapters/3-Chapter/logs/}{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}
+\graphicspath{{2-Chapters/3-Chapter/Images/}{2-Chapters/3-Chapter/logs/}{2-Chapters/3-Chapter/Images/plots-smpybandits-paper/}}
 
@@
 Finally, the most exciting thing that could happen to our \SMPyBandits{} library would be to see it gaining popularity!
-Its documentation has seen already about $20000$ visits, \href{https://github.com/SMPyBandits/SMPyBandits/stargazers}{the project on GitHub had $90$ stars} in June $2019$, and based on the features requested and the emails received about it, we counted between ten to twenty researchers in other labs who use or used \SMPyBandits.
+Its documentation has seen already about $25000$ visits, \href{https://github.com/SMPyBandits/SMPyBandits/stargazers}{the project on GitHub had $130$ stars} in November $2019$, and based on the features requested and the emails received about it, we counted between ten to twenty researchers in other labs who use or used \SMPyBandits.
 While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.

### 2019-11-27 2-Chapters/4-Chapter/CrownCom_17/IoT_slotted.tex
Removed raw inclusion of https://hal.inria.fr/hal-02006825 https://hal.inria.fr/hal-02049824v2 https://hal.archives-ouvertes.fr/hal-01575419v2
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

--- 2019-07-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
@@
 
-\graphicspath{{2-Chapters/4-Chapter/CrownCom_17.git/}}
+\graphicspath{{2-Chapters/4-Chapter/CrownCom_17/}}
 
@@
 We note that MAB learning has also started to be applied to also optimize the PHY layer parameters, see for instance \cite{KerkoucheAlami18} which followed our article \cite{Bonnefoi17},
-but in this thesis we chose to restrict to the spectrum access problem and thus we only MAB models where the arms correspond to orthogonal frequency bands.
+but in this thesis we chose to restrict to the spectrum access problem and thus we only consider MAB models where the arms correspond to orthogonal frequency bands.
 
@@
 % , in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
-Contrarily to many previous work done in the CR field and for Opportunistic Spectrum Access \cite{Jouini10,Jouini12},
+Contrarily to many previous works done in the CR field and for Opportunistic Spectrum Access \cite{Jouini10,Jouini12},
 the reward $r(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model.
@@
 All dynamic devices follow the same policy in this case, so the probability of transmitting at that time in channel $k$ for any dynamic device is $p / K$, and there are $D-1$ other dynamic devices.
-As they are independent, the probability that no other dynamic device sent in $i$
+As they are independent, the probability that no other dynamic device sent in $k$
 % $\Pr(\text{no other dynamic device sent in}\;i)$,
-is $q = \Pr(\bigcap_{k=1}^{D-1} \text{device } k \text{ did not send in } i) = \prod_{k=1}^{D-1} \Pr(\text{device } k \text{ did not send in } i)$.
-And $\Pr(\text{device } k \text{ sent in } i) = p \times 1 / K$, by uniform choice on channels and the Bernoulli emission hypothesis.
-So $q = \prod_{k=1}^{D-1} (1 - p/K) = (1-p/K)^{D-1}$.
+is $q = \Pr(\bigcap_{j=1}^{D-1} \text{device } j \text{ did not send in } k) = \prod_{j=1}^{D-1} \Pr(\text{device } j \text{ did not send in } k)$.
+And $\Pr(\text{device } j \text{ sent in } k) = p \times 1 / K$, by uniform choice on channels and the Bernoulli emission hypothesis.
+So $q = \prod_{j=1}^{D-1} (1 - p/K) = (1-p/K)^{D-1}$.
 Thus we can conclude,
@@
 In our model, every dynamic device implements its own learning algorithm, \emph{independently}.
-For one device, the time $t$ refers to the number of time it accessed the network (following its Bernoulli transmission process, \ie, its duty cycle), \emph{not} the total number of time slots from the beginning, as rewards are only obtained after a transmission, and IoT devices only transmit sporadically, due to low transmission duty cycles.
+For one device, the time $t$ refers to the number of times it accessed the network (following its Bernoulli transmission process, \ie, its duty cycle), \emph{not} the total number of time slots from the beginning, as rewards are only obtained after a transmission, and IoT devices only transmit sporadically, due to low transmission duty cycles.
 
@@
 Figure~\ref{fig:41:from10to100} presents the successful transmission rate as a function of time.
-The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
+The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which is outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
 The results are displayed when $10\%$, $30\%$, $50\%$ and $100\%$ of the traffic is generated by dynamic devices.

### 2019-11-27 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM/IEEE_WCNC__2019__Paper__BMBM.tex
Removed raw inclusion of https://hal.inria.fr/hal-02006825 https://hal.inria.fr/hal-02049824v2 https://hal.archives-ouvertes.fr/hal-01575419v2
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

--- 2019-07-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
@@
 
-\graphicspath{{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/}}
+\graphicspath{{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM/}}
 
@@
 We make the hypothesis that the transmission is successful if the channel is available, otherwise it fails upon radio collision.
-The novelty compare to the previous model is that
-in case of RF (up-link or down-link) collision that prevent the device to receive the \emph{Ack} from the gateway,
+The novelty compared to the previous model is that
+in case of RF (up-link or down-link) collision that prevents the devices from receiving the \emph{Ack} from the gateway,
 these devices will attempt to retransmit their packet up-to $\mathrm{MaxBackOff}$ times\footnote{~We denote it $\mathrm{MaxBackOff}$ instead of $M$ like we did in our paper \cite{Bonnefoi2019WCNC}, as $M$ is used in next Chapter~\ref{chapter:5}.},
@@
 
-It is worth to highlight that, if we write \eqref{eq:43:decomppc1} as $p_{c1} = p_c + p_{ca} \left(1-p_c\right)$,
+It is worth highlighting that, if we write \eqref{eq:43:decomppc1} as $p_{c1} = p_c + p_{ca} \left(1-p_c\right)$,
 then it is obvious that $p_{c1}$ is always larger than $p_c$ (as $p_{ca} \left(1-p_c\right) > 0$).

### 2019-11-27 2-Chapters/5-Chapter/ALT_2018__MPBandits/mpBandits.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

--- 2019-07-11 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
@@
     with a device from the background traffic $(Y_{A^j(t),t}=0)$,
-    or to a collision with one of the others players ($C^j(t)$).
-    However, the sensing and collision information are censored.
+    or to a collision with one of the other players ($C^j(t)$).
+    However, the sensing and collision information is censored.
     Recently, our work presented in Chapter~\ref{chapter:4} \cite{Bonnefoi17} presented the first (bandit-based) algorithmic solutions under this (harder) feedback model, in a slightly different setup, more suited to large scale IoT applications.
@@
   The regret lower bound that we gave in \cite{Besson2018ALT} is not applicable to any algorithm,
-  as it was discovered by V. Boursier and E. Perchet.
+  as it was discovered by E. Boursier and V. Perchet.
   As they explain in Section~2.4 in \cite{BoursierPerchet18},
@@
 % The Figure~\ref{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6} in Appendix illustrate two cases of $M<K$ and $M=K$ and the different impact of the three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3} on the regret.
-The Figure~6 in Appendix~F of \cite{Besson2018ALT} illustrates two cases of $M<K$ and $M=K$, and the different impact of the three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3} on the regret.
+Figure~6 in Appendix~F of \cite{Besson2018ALT} illustrates two cases of $M<K$ and $M=K$, and the different impact of the three terms \ref{eq:5:term1}, \ref{eq:5:term2} and \ref{eq:5:term3} on the regret.
 
@@
 Note that $\widetilde{\mu_k}^j(t)$ is no longer meant to be an unbiased estimate of $\mu_k$ as it also takes into account the collision information, that is present in the reward.
-Based on this empirical reward, one can similarly defined modified indices as
+Based on this empirical reward, one can similarly defined modified indices as (where $f(t)=\mathcal{O}(\log(t))$ is an exploration function)
 %
@@
 where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory (\eg, $f(t)=\log(t)+3\log(\log(t))$), which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
-A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
-However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
+A classical (single-player) UCB algorithm aims at the arm with the largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
+However, if each of the $M$ players selects the arm with the largest UCB, all the players will end up colliding most of the time on the best arm.
 To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.
@@
 \RandTopM{} is essentially a refinement over \RhoRand, to not use the indirection of ranks, and a simpler version of \MCTopM.
-The difference with \MCTopM{} is that the later introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
+The difference with \MCTopM{} is that the latter introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
 %
@@
       \end{tabular}
-      \caption{For all orthogonalization policies and different numbers of players, using \klUCB{} is much more efficient than using \UCB{}, for multi-players bandit (here in a simple problem with $K=9$ arms).}
+      \caption{Regret for all orthogonalization policies and different numbers of players.
+Using \klUCB{} is much more efficient than using \UCB{}, for multi-players bandit (here in a simple problem with $K=9$ arms).}
       \label{table:5:comparisonUCB_klUCB}

### 2019-11-27 2-Chapters/5-Chapter/chapter5.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 In this chapter, we are interested in a more formal approach to the decentralized learning problem presented of many end-devices accessing the same IoT wireless network
-in the previous Chapter~\ref{chapter:4}.
-We restrict to the easier case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
+than in the previous Chapter~\ref{chapter:4}.
+We restrict to the easiest case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
 Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$,
@@
 The first idea comes from \cite{Anandkumar11}, and the intuition behind it is quite simple, even if the mathematical derivations are not.
-All players will build an estimate $\widehat{M}^j(t)$ of the number of player, that start by $\widehat{M}^j(0)=1$.
+All players will build an estimate $\widehat{M}^j(t)$ of the number of players, that start by $\widehat{M}^j(0)=1$.
 As soon as one collision is observed, a player knows that $M\geq2$, so $\widehat{M}^j(t+1) = 2$.
@@
 
-Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
+Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of players.
 We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for \RandTopM{} and \MCTopM.
@@
 %
-In the easier case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO\left( M K \log(K T) + K M^3 \log(M K T) / \Delta + K M^2 (\log(M \log(K T) / \Delta))^2 \right)$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
+In the easiest case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO\left( M K \log(K T) + K M^3 \log(M K T) / \Delta + K M^2 (\log(M \log(K T) / \Delta))^2 \right)$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
 In the generic case, their algorithm M-ETC with Elimination achieves a regret of $\cO((\log(T))^{1+\kappa})$ for any positive constant $\kappa$.

### 2019-11-27 2-Chapters/6-Chapter/nonstatbandits/NonStatB.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 A (random) stream of rewards $(Y_{k,t})_{t\in[T]}$ is associated to each arm $k \in [K]$.
-We assume that the rewards are bounded, and wlog we assume that $Y_{k,t} \in [0,1]$.
+We assume that the rewards are bounded, and without loss of generality we assume that $Y_{k,t} \in [0,1]$.
 We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
@@
 
-\paragraph{Two example of problems.}\label{par:6:benchmark1}
-%
-To illustrate the model, we give here two examples of piece-wise stationary problems.
+\paragraph{Two examples of problems.}\label{par:6:benchmark1}
+%
+To illustrate the model, we give here two examples of piece-wise stationary problems
 % problems of our benchmark,
@@
 \textbf{Problem 3.} (see Figure~\ref{fig:6:Problem_5}) This problem is harder, with $K=6$, $\Upsilon_T=8$ and $T=20000$.
-At every break-points, almost all arms change,
+At every break-point, almost all arms change,
 and means are bounded in $[0.01, 0.07]$.

### 2019-11-27 2-Chapters/Conclusion/conclusion.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

@@
 Our main contribution to answer our problematic is a model of Internet of Things networks, based on an ALOHA-like protocol slotted in time and frequency, and that considers one IoT base station serving a large number of devices following the IoT constraints (low-cost, low duty cycle, long range etc).
-In such IoT network, we focused on many dynamically reconfigurable IoT end-devices, that are able to run low-cost and low-complexity decision making algorithms, embedded on each device using their limited computational and storage capacities.
+In such a IoT network, we focused on many dynamically reconfigurable IoT end-devices, that are able to run low-cost and low-complexity decision making algorithms, embedded on each device using their limited computational and storage capacities.
 %  (limited hardware)

### 2019-12-06 1-Preamble/5-Abstract/resume_fr.tex
Ajout licence CC BY-NC-SA https://creativecommons.org/licenses/by-nc-sa/4.0/

@@
 %
-Ce document et les ressources additionnelles requises pour le générer (notamment les fichiers \LaTeX, les morceaux de code Python, les figures etc)
-sont \href{https://github.com/Naereen/phd-thesis/}{distribuées publiquement},
-selon les termes de la \href{https://lbesson.mit-license.org/}{\emph{licence MIT}} open-source,
+Ce document est \href{https://github.com/Naereen/phd-thesis/}{publié} selon les termes de la
+\href{https://creativecommons.org/licenses/by-nc-sa/4.0/}{licence Creative Commons CC BY-NC-SA 4.0} (attribution - pas d'utilisation commerciale - partage dans les mêmes conditions),
+et les ressources additionnelles requises pour le générer (notamment les fichiers \LaTeX, les morceaux de code Python, les figures etc)
+sont \href{https://github.com/Naereen/phd-thesis/}{distribuées},
+sous la \href{https://lbesson.mit-license.org/}{\emph{licence MIT}} open-source,
 en ligne sur \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}}.
 
-\TODOL{Open source the repository as soon as I defended my thesis!}
-
 \begin{center}
+    % \includegraphics[height=1em]{1-Preamble/1-Front/CC-BY-NC-SA-4.png}
     \textbf{Copyright 2016-2019, \copyright ~Lilian~Besson.}

### 2019-12-06 2-Chapters/1-Chapter/chapter1.tex
Ajout licence CC BY-NC-SA https://creativecommons.org/licenses/by-nc-sa/4.0/

@@
 %
-This document and the additional resources required to compile it (including \LaTeX{} code, Python snippets, images etc)
-are \href{https://github.com/Naereen/phd-thesis/}{publicly published},
-under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}},
+This document is publicly published under the terms of the
+\href{https://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative Commons CC BY-NC-SA 4.0 license} (attribution - non commercial - share alike),
+and the additional resources required to compile it (including \LaTeX{} code, Python snippets, images etc)
+are \href{https://github.com/Naereen/phd-thesis/}{published},
+under open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}},
 online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}}.
 
-% FIXME
-% \TODOL{Open source the repository as soon as I defended my thesis!}
-
 \begin{center}
+    % \includegraphics[height=1em]{1-Preamble/1-Front/CC-BY-NC-SA-4.png}
     \textbf{Copyright $2016$-$2019$, \copyright ~Lilian~Besson.}
